\NormalFont\ShortTitle{1 Chronicles}
{\MT The First Book of the Chronicles

\par }\ChapOne{1}{\PP \VerseOne{1}Adam,
Sheth,
Enosh,
\VS{2}Kenan,
Mahalaleel,
Jered,
\VS{3}Henoch,
Methuselah,
Lamech,
\VS{4}Noah,
Shem,
Ham, and
Japheth.
\par }{\PP \VS{5}¶ The
sons of
Japheth;
Gomer, and
Magog, and
Madai, and
Javan, and
Tubal, and
Meshech, and
Tiras.
\VS{6}And the
sons of
Gomer;
Ashchenaz, and
Riphath, and
Togarmah.\FTNT{*}{{\FR 1.6 }
Riphath: or, Diphath as it is in some copies}
\VS{7}And the
sons of
Javan;
Elishah, and
Tarshish,
Kittim, and
Dodanim.\FTNT{†}{{\FR 1.7 }
Dodanim: or, Rodanim, according to some copies}
\par }{\PP \VS{8}¶ The
sons of
Ham;
Cush, and
Mizraim,
Put, and
Canaan.
\VS{9}And the
sons of
Cush;
Seba, and
Havilah, and
Sabta, and
Raamah, and
Sabtecha. And the
sons of
Raamah;
Sheba, and
Dedan.
\VS{10}And
Cush
begat
Nimrod: he
began to be
mighty upon the
earth.
\VS{11}And
Mizraim
begat
Ludim, and
Anamim, and
Lehabim, and
Naphtuhim,
\VS{12}And
Pathrusim, and
Casluhim, (of whom
came the
Philistines,) and
Caphthorim.
\VS{13}And
Canaan
begat
Zidon his
firstborn, and
Heth,
\VS{14}The
Jebusite also, and the
Amorite, and the
Girgashite,
\VS{15}And the
Hivite, and the
Arkite, and the
Sinite,
\VS{16}And the
Arvadite, and the
Zemarite, and the
Hamathite.
\par }{\PP \VS{17}¶ The
sons of
Shem;
Elam, and
Asshur, and
Arphaxad, and
Lud, and
Aram, and
Uz, and
Hul, and
Gether, and
Meshech.\FTNT{‡}{{\FR 1.17 }
Meshech: or, Mash}
\VS{18}And
Arphaxad
begat
Shelah, and
Shelah
begat
Eber.
\VS{19}And unto
Eber were
born
two
sons: the
name of the
one
{\ADD{was}}
Peleg; because in his
days the
earth was
divided: and his
brother’s
name
{\ADD{was}}
Joktan.\FTNT{§}{{\FR 1.19 }
Peleg: that is, division}
\VS{20}And
Joktan
begat
Almodad, and
Sheleph, and
Hazarmaveth, and
Jerah,
\VS{21}Hadoram also, and
Uzal, and
Diklah,
\VS{22}And
Ebal, and
Abimael, and
Sheba,
\VS{23}And
Ophir, and
Havilah, and
Jobab. All these
{\ADD{were}} the
sons of
Joktan.
\par }{\PP \VS{24}¶
Shem,
Arphaxad,
Shelah,
\VS{25}Eber,
Peleg,
Reu,
\VS{26}Serug,
Nahor,
Terah,
\VS{27}Abram; the same
{\ADD{is}}
Abraham.
\par }{\PP \VS{28}The
sons of
Abraham;
Isaac, and
Ishmael.
\par }{\PP \VS{29}¶ These
{\ADD{are}} their
generations: The
firstborn of
Ishmael,
Nebaioth; then
Kedar, and
Adbeel, and
Mibsam,
\VS{30}Mishma, and
Dumah,
Massa,
Hadad, and
Tema,\FTNT{**}{{\FR 1.30 }
Hadad: also called, Hadar}
\VS{31}Jetur,
Naphish, and
Kedemah. These are the
sons of
Ishmael.
\par }{\PP \VS{32}¶ Now the
sons of
Keturah,
Abraham’s
concubine: she
bare
Zimran, and
Jokshan, and
Medan, and
Midian, and
Ishbak, and
Shuah. And the
sons of
Jokshan;
Sheba, and
Dedan.
\VS{33}And the
sons of
Midian;
Ephah, and
Epher, and
Henoch, and
Abida, and
Eldaah. All these
{\ADD{are}} the
sons of
Keturah.
\VS{34}And
Abraham
begat
Isaac. The
sons of
Isaac;
Esau and
Israel.
\par }{\PP \VS{35}¶ The
sons of
Esau;
Eliphaz,
Reuel, and
Jeush, and
Jaalam, and
Korah.
\VS{36}The
sons of
Eliphaz;
Teman, and
Omar,
Zephi, and
Gatam,
Kenaz, and
Timna, and
Amalek.\FTNT{††}{{\FR 1.36 }
Zephi: or, Zepho}
\VS{37}The
sons of
Reuel;
Nahath,
Zerah,
Shammah, and
Mizzah.
\VS{38}And the
sons of
Seir;
Lotan, and
Shobal, and
Zibeon, and
Anah, and
Dishon, and
Ezer, and
Dishan.
\VS{39}And the
sons of
Lotan;
Hori, and
Homam: and
Timna
{\ADD{was}}
Lotan’s
sister.\FTNT{‡‡}{{\FR 1.39 }
Homam: or, Hemam}
\VS{40}The
sons of
Shobal;
Alian, and
Manahath, and
Ebal,
Shephi, and
Onam. And the
sons of
Zibeon;
Aiah, and
Anah.\FTNT{§§}{{\FR 1.40 }
Alian: also called, Alvan}\FTNT{***}{{\FR 1.40 }
Shephi: also called, Shepho}
\VS{41}The
sons of
Anah;
Dishon. And the
sons of
Dishon;
Amram, and
Eshban, and
Ithran, and
Cheran.\FTNT{†††}{{\FR 1.41 }
Amram: or, Hemdan}
\VS{42}The
sons of
Ezer;
Bilhan, and
Zavan,
{\ADD{and}}
Jakan. The
sons of
Dishan;
Uz, and
Aran.\FTNT{‡‡‡}{{\FR 1.42 }
Jakan: or, Akan}
\par }{\PP \VS{43}¶ Now these
{\ADD{are}} the
kings that
reigned in the
land of
Edom
before
{\ADD{any}}
king
reigned over the
children of
Israel;
Bela the
son of
Beor: and the
name of his
city
{\ADD{was}}
Dinhabah.
\VS{44}And when
Bela was
dead,
Jobab the
son of
Zerah of
Bozrah
reigned in his stead.
\VS{45}And when
Jobab was
dead,
Husham of the
land of the
Temanites
reigned in his stead.
\VS{46}And when
Husham was
dead,
Hadad the
son of
Bedad, which
smote
Midian in the
field of
Moab,
reigned in his stead: and the
name of his
city
{\ADD{was}}
Avith.
\VS{47}And when
Hadad was
dead,
Samlah of
Masrekah
reigned in his stead.
\VS{48}And when
Samlah was
dead,
Shaul of
Rehoboth by the
river
reigned in his stead.
\VS{49}And when
Shaul was
dead,
Baal-hanan the
son of
Achbor
reigned in his stead.
\VS{50}And when
Baal-hanan was
dead,
Hadad
reigned in his stead: and the
name of his
city
{\ADD{was}}
Pai; and his
wife’s
name
{\ADD{was}}
Mehetabel, the
daughter of
Matred, the
daughter of
Mezahab.\FTNT{§§§}{{\FR 1.50 }
Hadad: or, Hadar}\FTNT{*}{{\FR 1.50 }
Pai: or, Pau}
\par }{\PP \VS{51}¶
Hadad
died also. And the
dukes of
Edom were;
duke
Timnah,
duke
Aliah,
duke
Jetheth,
\VS{52}Duke
Aholibamah,
duke
Elah,
duke
Pinon,
\VS{53}Duke
Kenaz,
duke
Teman,
duke
Mibzar,
\VS{54}Duke
Magdiel,
duke
Iram. These
{\ADD{are}} the
dukes of
Edom.

\par }\Chap{2}{\PP \VerseOne{1}These
{\ADD{are}} the
sons of
Israel;
Reuben,
Simeon,
Levi, and
Judah,
Issachar, and
Zebulun,\FTNT{*}{{\FR 2.1 }
Israel: or, Jacob}
\VS{2}Dan,
Joseph, and
Benjamin,
Naphtali,
Gad, and
Asher.
\par }{\PP \VS{3}¶ The
sons of
Judah;
Er, and
Onan, and
Shelah:
{\ADD{which}}
three were
born unto him of the
daughter of
Shua the
Canaanitess. And
Er, the
firstborn of
Judah, was
evil in the
sight of the
{\ND{
LORD}}; and he
slew him.
\VS{4}And
Tamar his daughter in
law
bare him
Pharez and
Zerah. All the
sons of
Judah
{\ADD{were}}
five.
\VS{5}The
sons of
Pharez;
Hezron, and
Hamul.
\VS{6}And the
sons of
Zerah;
Zimri, and
Ethan, and
Heman, and
Calcol, and
Dara:
five of them in all.\FTNT{†}{{\FR 2.6 }
Zimri: or, Zabdi}\FTNT{‡}{{\FR 2.6 }
Dara: or, Darda}
\VS{7}And the
sons of
Carmi;
Achar, the
troubler of
Israel, who
transgressed in the thing
accursed.\FTNT{§}{{\FR 2.7 }
Achar: or, Achan}
\VS{8}And the
sons of
Ethan;
Azariah.
\VS{9}The
sons also of
Hezron, that were
born unto him;
Jerahmeel, and
Ram, and
Chelubai.\FTNT{**}{{\FR 2.9 }
Ram: Gr. Aram}\FTNT{††}{{\FR 2.9 }
Chelubai: or, Caleb}
\VS{10}And
Ram
begat
Amminadab; and
Amminadab
begat
Nahshon,
prince of the
children of
Judah;
\VS{11}And
Nahshon
begat
Salma, and
Salma
begat
Boaz,\FTNT{‡‡}{{\FR 2.11 }
Salma: also called, Salmon}
\VS{12}And
Boaz
begat
Obed, and
Obed
begat
Jesse,
\par }{\PP \VS{13}¶ And
Jesse
begat his
firstborn
Eliab, and
Abinadab the
second, and
Shimma the
third,\FTNT{§§}{{\FR 2.13 }
Shimma: or, Shammah}
\VS{14}Nethaneel the
fourth,
Raddai the
fifth,
\VS{15}Ozem the
sixth,
David the
seventh:
\VS{16}Whose
sisters
{\ADD{were}}
Zeruiah, and
Abigail. And the
sons of
Zeruiah;
Abishai, and
Joab, and
Asahel,
three.
\VS{17}And
Abigail
bare
Amasa: and the
father of
Amasa
{\ADD{was}}
Jether the
Ishmeelite.\FTNT{***}{{\FR 2.17 }
Jether…: also called, Ithra an Israelite}
\par }{\PP \VS{18}¶ And
Caleb the
son of
Hezron
begat
{\ADD{children}}
of
Azubah
{\ADD{his}}
wife, and of
Jerioth: her
sons
{\ADD{are}} these;
Jesher, and
Shobab, and
Ardon.
\VS{19}And when
Azubah was
dead,
Caleb
took unto him
Ephrath, which
bare him
Hur.
\VS{20}And
Hur
begat
Uri, and
Uri
begat
Bezaleel.
\par }{\PP \VS{21}¶ And
afterward
Hezron went
in to the
daughter of
Machir the
father of
Gilead, whom he
married when he
{\ADD{was}}
threescore
years
old; and she
bare him
Segub.\FTNT{†††}{{\FR 2.21 }
married: Heb. took}
\VS{22}And
Segub
begat
Jair, who had
three and
twenty
cities in the
land of
Gilead.
\VS{23}And he
took
Geshur, and
Aram, with the
towns of
Jair, from them, with
Kenath, and the
towns thereof,
{\ADD{even}}
threescore
cities. All these
{\ADD{belonged to}} the
sons of
Machir the
father of
Gilead.
\VS{24}And
after that
Hezron was
dead in
Caleb-ephratah, then
Abiah
Hezron’s
wife
bare him
Ashur the
father of
Tekoa.
\par }{\PP \VS{25}¶ And the
sons of
Jerahmeel the
firstborn of
Hezron were,
Ram the
firstborn, and
Bunah, and
Oren, and
Ozem,
{\ADD{and}}
Ahijah.
\VS{26}Jerahmeel had also
another
wife, whose
name
{\ADD{was}}
Atarah; she
{\ADD{was}} the
mother of
Onam.
\VS{27}And the
sons of
Ram the
firstborn of
Jerahmeel were,
Maaz, and
Jamin, and
Eker.
\VS{28}And the
sons of
Onam were,
Shammai, and
Jada. And the
sons of
Shammai;
Nadab, and
Abishur.
\VS{29}And the
name of the
wife of
Abishur
{\ADD{was}}
Abihail, and she
bare him
Ahban, and
Molid.
\VS{30}And the
sons of
Nadab;
Seled, and
Appaim: but
Seled
died without
children.
\VS{31}And the
sons of
Appaim;
Ishi. And the
sons of
Ishi;
Sheshan. And the
children of
Sheshan;
Ahlai.
\VS{32}And the
sons of
Jada the
brother of
Shammai;
Jether, and
Jonathan: and
Jether
died without
children.
\VS{33}And the
sons of
Jonathan;
Peleth, and
Zaza. These were the
sons of
Jerahmeel.
\par }{\PP \VS{34}¶ Now
Sheshan had no
sons, but
daughters. And
Sheshan had a
servant, an
Egyptian, whose
name
{\ADD{was}}
Jarha.
\VS{35}And
Sheshan
gave his
daughter to
Jarha his
servant to
wife; and she
bare him
Attai.
\VS{36}And
Attai
begat
Nathan, and
Nathan
begat
Zabad,
\VS{37}And
Zabad
begat
Ephlal, and
Ephlal
begat
Obed,
\VS{38}And
Obed
begat
Jehu, and
Jehu
begat
Azariah,
\VS{39}And
Azariah
begat
Helez, and
Helez
begat
Eleasah,
\VS{40}And
Eleasah
begat
Sisamai, and
Sisamai
begat
Shallum,
\VS{41}And
Shallum
begat
Jekamiah, and
Jekamiah
begat
Elishama.
\par }{\PP \VS{42}¶ Now the
sons of
Caleb the
brother of
Jerahmeel
{\ADD{were}},
Mesha his
firstborn, which was the
father of
Ziph; and the
sons of
Mareshah the
father of
Hebron.
\VS{43}And the
sons of
Hebron;
Korah, and
Tappuah, and
Rekem, and
Shema.
\VS{44}And
Shema
begat
Raham, the
father of
Jorkoam: and
Rekem
begat
Shammai.
\VS{45}And the
son of
Shammai
{\ADD{was}}
Maon: and
Maon
{\ADD{was}} the
father of
Beth-zur.
\VS{46}And
Ephah,
Caleb’s
concubine,
bare
Haran, and
Moza, and
Gazez: and
Haran
begat
Gazez.
\VS{47}And the
sons of
Jahdai;
Regem, and
Jotham, and
Geshan, and
Pelet, and
Ephah, and
Shaaph.
\VS{48}Maachah,
Caleb’s
concubine,
bare
Sheber, and
Tirhanah.
\VS{49}She
bare also
Shaaph the
father of
Madmannah,
Sheva the
father of
Machbenah, and the
father of
Gibea: and the
daughter of
Caleb
{\ADD{was}}
Achsah.
\par }{\PP \VS{50}¶ These were the
sons of
Caleb the
son of
Hur, the
firstborn of
Ephratah;
Shobal the
father of
Kirjath-jearim,\FTNT{‡‡‡}{{\FR 2.50 }
Ephratah: also called, Ephreth}
\VS{51}Salma the
father of
Beth-lehem,
Hareph the
father of
Beth-gader.
\VS{52}And
Shobal the
father of
Kirjath-jearim had
sons;
Haroeh,
{\ADD{and}}
half of the
Manahethites.\FTNT{§§§}{{\FR 2.52 }
Haroeh: or, Reaiah}\FTNT{*}{{\FR 2.52 }
half…: or, half of the Menuchites, or, Hatsi-ham-menuchoth}
\VS{53}And the
families of
Kirjath-jearim; the
Ithrites, and the
Puhites, and the
Shumathites, and the
Mishraites; of them
came the
Zareathites, and the
Eshtaulites.
\VS{54}The
sons of
Salma;
Beth-lehem, and the
Netophathites,
Ataroth, the house of
Joab, and
half of the
Manahethites, the
Zorites.\FTNT{†}{{\FR 2.54 }
Ataroth…: or, Atarites, or, crowns of the house of Joab}
\VS{55}And the
families of the
scribes which
dwelt at
Jabez; the
Tirathites, the
Shimeathites,
{\ADD{and}}
Suchathites.
These
{\ADD{are}} the
Kenites that
came of
Hemath, the
father of the
house of
Rechab.

\par }\Chap{3}{\PP \VerseOne{1}Now these were the
sons of
David, which were
born unto him in
Hebron; the
firstborn
Amnon, of
Ahinoam the
Jezreelitess; the
second
Daniel, of
Abigail the
Carmelitess:\FTNT{*}{{\FR 3.1 }
Daniel: or, Chileab}
\VS{2}The
third,
Absalom the
son of
Maachah the
daughter of
Talmai
king of
Geshur: the
fourth,
Adonijah the
son of
Haggith:
\VS{3}The
fifth,
Shephatiah of
Abital: the
sixth,
Ithream by
Eglah his
wife.
\VS{4}{\ADD{These}}
six were
born unto him in
Hebron; and there he
reigned
seven
years and
six
months: and in
Jerusalem he
reigned
thirty and
three
years.
\VS{5}And these were
born unto him in
Jerusalem;
Shimea, and
Shobab, and
Nathan, and
Solomon,
four, of
Bath-shua the
daughter of
Ammiel:\FTNT{†}{{\FR 3.5 }
Shimea: or, Shammua}\FTNT{‡}{{\FR 3.5 }
Bath-shua: or, Bath-sheba}\FTNT{§}{{\FR 3.5 }
Ammiel: or, Eliam}
\VS{6}Ibhar also, and
Elishama, and
Eliphelet,\FTNT{**}{{\FR 3.6 }
Elishama: also called, Elishua}
\VS{7}And
Nogah, and
Nepheg, and
Japhia,
\VS{8}And
Elishama, and
Eliada, and
Eliphelet,
nine.\FTNT{††}{{\FR 3.8 }
Eliada: or, Beeliada}
\VS{9}{\ADD{These were}} all the
sons of
David, beside the
sons of the
concubines, and
Tamar their
sister.
\par }{\PP \VS{10}¶ And
Solomon’s
son
{\ADD{was}}
Rehoboam,
Abia his
son,
Asa his
son,
Jehoshaphat his
son,\FTNT{‡‡}{{\FR 3.10 }
Abia: or, Abijam}
\VS{11}Joram his
son,
Ahaziah his
son,
Joash his
son,\FTNT{§§}{{\FR 3.11 }
Ahaziah: or, Azariah}
\VS{12}Amaziah his
son,
Azariah his
son,
Jotham his
son,\FTNT{***}{{\FR 3.12 }
Azariah: or, Uzziah}
\VS{13}Ahaz his
son,
Hezekiah his
son,
Manasseh his
son,
\VS{14}Amon his
son,
Josiah his
son.
\VS{15}And the
sons of
Josiah
{\ADD{were}}, the
firstborn
Johanan, the
second
Jehoiakim, the
third
Zedekiah, the
fourth
Shallum.\FTNT{†††}{{\FR 3.15 }
Johanan: or, Jehoahaz}\FTNT{‡‡‡}{{\FR 3.15 }
Jehoiakim: or, Eliakim}\FTNT{§§§}{{\FR 3.15 }
Zedekiah: or, Mattaniah}
\VS{16}And the
sons of
Jehoiakim:
Jeconiah his
son,
Zedekiah his
son.\FTNT{*}{{\FR 3.16 }
Jeconiah: also called, Jehoiachin or Coniah}
\par }{\PP \VS{17}¶ And the
sons of
Jeconiah;
Assir,
Salathiel his
son,\FTNT{†}{{\FR 3.17 }
Salathiel: Heb. Shealtiel}
\VS{18}Malchiram also, and
Pedaiah, and
Shenazar,
Jecamiah,
Hoshama, and
Nedabiah.
\VS{19}And the
sons of
Pedaiah
{\ADD{were}},
Zerubbabel, and
Shimei: and the
sons of
Zerubbabel;
Meshullam, and
Hananiah, and
Shelomith their
sister:
\VS{20}And
Hashubah, and
Ohel, and
Berechiah, and
Hasadiah,
Jushab-hesed,
five.
\VS{21}And the
sons of
Hananiah;
Pelatiah, and
Jesaiah: the
sons of
Rephaiah, the
sons of
Arnan, the
sons of
Obadiah, the
sons of
Shechaniah.
\VS{22}And the
sons of
Shechaniah;
Shemaiah: and the
sons of
Shemaiah;
Hattush, and
Igeal, and
Bariah, and
Neariah, and
Shaphat,
six.
\VS{23}And the
sons of
Neariah;
Elioenai, and
Hezekiah, and
Azrikam,
three.\FTNT{‡}{{\FR 3.23 }
Hezekiah: Heb. Hiskijah}
\VS{24}And the
sons of
Elioenai
{\ADD{were}},
Hodaiah, and
Eliashib, and
Pelaiah, and
Akkub, and
Johanan, and
Dalaiah, and
Anani,
seven.

\par }\Chap{4}{\PP \VerseOne{1}The
sons of
Judah;
Pharez,
Hezron, and
Carmi, and
Hur, and
Shobal.\FTNT{*}{{\FR 4.1 }
Carmi: also called, Chelubai or Caleb}
\VS{2}And
Reaiah the
son of
Shobal
begat
Jahath; and
Jahath
begat
Ahumai, and
Lahad. These
{\ADD{are}} the
families of the
Zorathites.\FTNT{†}{{\FR 4.2 }
Reaiah: or, Haroeh}
\VS{3}And these
{\ADD{were of}} the
father of
Etam;
Jezreel, and
Ishma, and
Idbash: and the
name of their
sister
{\ADD{was}}
Hazelelponi:
\VS{4}And
Penuel the
father of
Gedor, and
Ezer the
father of
Hushah. These
{\ADD{are}} the
sons of
Hur, the
firstborn of
Ephratah, the
father of
Beth-lehem.
\par }{\PP \VS{5}¶ And
Ashur the
father of
Tekoa had
two
wives,
Helah and
Naarah.
\VS{6}And
Naarah
bare him
Ahuzam, and
Hepher, and
Temeni, and
Haahashtari. These
{\ADD{were}} the
sons of
Naarah.
\VS{7}And the
sons of
Helah
{\ADD{were}},
Zereth, and
Jezoar, and
Ethnan.
\VS{8}And
Coz
begat
Anub, and
Zobebah, and the
families of
Aharhel the
son of
Harum.
\par }{\PP \VS{9}¶ And
Jabez was more
honourable than his
brethren: and his
mother
called his
name
Jabez,
saying, Because I
bare him with
sorrow.\FTNT{‡}{{\FR 4.9 }
Jabez: that is, Sorrowful}
\VS{10}And
Jabez
called on the
God of
Israel,
saying, Oh
that thou wouldest
bless me
indeed, and
enlarge my
coast, and that thine
hand might be with me, and that thou wouldest
keep
{\ADD{me}} from
evil, that it may not
grieve me! And
God
granted him that which he
requested.\FTNT{§}{{\FR 4.10 }
Oh…: Heb. If thou wilt, etc}\FTNT{**}{{\FR 4.10 }
keep…: Heb. do me}
\par }{\PP \VS{11}¶ And
Chelub the
brother of
Shuah
begat
Mehir, which
{\ADD{was}} the
father of
Eshton.
\VS{12}And
Eshton
begat
Beth-rapha, and
Paseah, and
Tehinnah the
father of
Ir-nahash. These
{\ADD{are}} the
men of
Rechah.\FTNT{††}{{\FR 4.12 }
Ir-nahash: or, the city of Nahash}
\VS{13}And the
sons of
Kenaz;
Othniel, and
Seraiah: and the
sons of
Othniel;
Hathath.\FTNT{‡‡}{{\FR 4.13 }
Hathath…: or, Hathath, and Meonothai, who begat, etc}
\VS{14}And
Meonothai
begat
Ophrah: and
Seraiah
begat
Joab, the
father of the
valley of
Charashim; for they were
craftsmen.\FTNT{§§}{{\FR 4.14 }
valley: or, inhabitants of the valley}\FTNT{***}{{\FR 4.14 }
Charashim: that is, craftsmen}\FTNT{†††}{{\FR 4.14 }
Hathath…: or, Hathath, and Meonothai, who begat, etc}
\VS{15}And the
sons of
Caleb the
son of
Jephunneh;
Iru,
Elah, and
Naam: and the
sons of
Elah, even
Kenaz.\FTNT{‡‡‡}{{\FR 4.15 }
even Kenaz: or, Uknaz}
\VS{16}And the
sons of
Jehaleleel;
Ziph, and
Ziphah,
Tiria, and
Asareel.
\VS{17}And the
sons of
Ezra
{\ADD{were}},
Jether, and
Mered, and
Epher, and
Jalon: and she
bare
Miriam, and
Shammai, and
Ishbah the
father of
Eshtemoa.
\VS{18}And his
wife
Jehudijah
bare
Jered the
father of
Gedor, and
Heber the
father of
Socho, and
Jekuthiel the
father of
Zanoah. And these
{\ADD{are}} the
sons of
Bithiah the
daughter of
Pharaoh, which
Mered
took.\FTNT{§§§}{{\FR 4.18 }
Jehudijah: or, the Jewess}
\VS{19}And the
sons of
{\ADD{his}}
wife
Hodiah the
sister of
Naham, the
father of
Keilah the
Garmite, and
Eshtemoa the
Maachathite.\FTNT{*}{{\FR 4.19 }
Hodiah: or, Jehudijah, mentioned before}
\VS{20}And the
sons of
Shimon
{\ADD{were}},
Amnon, and
Rinnah,
Ben-hanan, and
Tilon. And the
sons of
Ishi
{\ADD{were}},
Zoheth, and
Ben-zoheth.
\par }{\PP \VS{21}¶ The
sons of
Shelah the
son of
Judah
{\ADD{were}},
Er the
father of
Lecah, and
Laadah the
father of
Mareshah, and the
families of the
house of them that
wrought fine
linen, of the
house of
Ashbea,
\VS{22}And
Jokim, and the
men of
Chozeba, and
Joash, and
Saraph, who had the
dominion in
Moab, and
Jashubi-lehem. And
{\ADD{these are}}
ancient
things.
\VS{23}These
{\ADD{were}} the
potters, and those that
dwelt among
plants and
hedges: there they
dwelt with the
king for his
work.
\par }{\PP \VS{24}¶ The
sons of
Simeon
{\ADD{were}},
Nemuel, and
Jamin,
Jarib,
Zerah,
{\ADD{and}}
Shaul:\FTNT{†}{{\FR 4.24 }
Nemuel: or, Jemuel}\FTNT{‡}{{\FR 4.24 }
Jarib, Zerah: or, Jachin Zohar}
\VS{25}Shallum his
son,
Mibsam his
son,
Mishma his
son.
\VS{26}And the
sons of
Mishma;
Hamuel his
son,
Zacchur his
son,
Shimei his
son.
\VS{27}And
Shimei had
sixteen
sons and
six
daughters; but his
brethren had not
many
children, neither did all their
family
multiply, like to the
children of
Judah.\FTNT{§}{{\FR 4.27 }
like…: Heb. unto}
\VS{28}And they
dwelt at
Beer-sheba, and
Moladah, and
Hazar-shual,
\VS{29}And at
Bilhah, and at
Ezem, and at
Tolad,\FTNT{**}{{\FR 4.29 }
Bilhah: or, Balah}\FTNT{††}{{\FR 4.29 }
Tolad: or, Eltolad}
\VS{30}And at
Bethuel, and at
Hormah, and at
Ziklag,
\VS{31}And at
Beth-marcaboth, and
Hazar-susim, and at
Beth-birei, and at
Shaaraim. These
{\ADD{were}} their
cities unto the
reign of
David.\FTNT{‡‡}{{\FR 4.31 }
Hazar-susim: or, Hazar-susah}
\VS{32}And their
villages
{\ADD{were}},
Etam, and
Ain,
Rimmon, and
Tochen, and
Ashan,
five
cities:\FTNT{§§}{{\FR 4.32 }
Etam: or, Ether}
\VS{33}And all their
villages that
{\ADD{were}} round
about the same
cities, unto
Baal. These
{\ADD{were}} their
habitations, and their
genealogy.\FTNT{***}{{\FR 4.33 }
Baal: or, Baalath-beer}\FTNT{†††}{{\FR 4.33 }
their genealogy: or, as they divided themselves by nations among them}
\VS{34}And
Meshobab, and
Jamlech, and
Joshah the
son of
Amaziah,
\VS{35}And
Joel, and
Jehu the
son of
Josibiah, the
son of
Seraiah, the
son of
Asiel,
\VS{36}And
Elioenai, and
Jaakobah, and
Jeshohaiah, and
Asaiah, and
Adiel, and
Jesimiel, and
Benaiah,
\VS{37}And
Ziza the
son of
Shiphi, the
son of
Allon, the
son of
Jedaiah, the
son of
Shimri, the
son of
Shemaiah;
\VS{38}These
mentioned by
{\ADD{their}}
names
{\ADD{were}}
princes in their
families: and the
house of their
fathers
increased
greatly.\FTNT{‡‡‡}{{\FR 4.38 }
mentioned: Heb. coming}
\par }{\PP \VS{39}¶ And they
went to the
entrance of
Gedor,
{\ADD{even}} unto the east
side of the
valley, to
seek
pasture for their
flocks.
\VS{40}And they
found
fat
pasture and
good, and the
land
{\ADD{was}}
wide, and
quiet, and
peaceable; for
{\ADD{they}} of
Ham had
dwelt there of
old.
\VS{41}And these
written by
name
came in the
days of
Hezekiah
king of
Judah, and
smote their
tents, and the
habitations that were
found there, and
destroyed them utterly unto this
day, and
dwelt in their rooms: because
{\ADD{there was}}
pasture there for their
flocks.
\VS{42}And
{\ADD{some}} of them,
{\ADD{even}} of the
sons of
Simeon,
five
hundred
men,
went to
mount
Seir, having for their
captains
Pelatiah, and
Neariah, and
Rephaiah, and
Uzziel, the
sons of
Ishi.
\VS{43}And they
smote the
rest of the
Amalekites that were
escaped, and
dwelt there unto this
day.

\par }\Chap{5}{\PP \VerseOne{1}Now the
sons of
Reuben the
firstborn of
Israel, (for he
{\ADD{was}} the
firstborn; but, forasmuch as he
defiled his
father’s
bed, his
birthright was
given unto the
sons of
Joseph the
son of
Israel: and the genealogy is not to be
reckoned after the
birthright.
\VS{2}For
Judah
prevailed above his
brethren, and of him
{\ADD{came}} the chief
ruler; but the
birthright
{\ADD{was}}
Joseph’s:)\FTNT{*}{{\FR 5.2 }
chief…: or, prince}
\VS{3}The
sons,
{\ADD{I say}}, of
Reuben the
firstborn of
Israel
{\ADD{were}},
Hanoch, and
Pallu,
Hezron, and
Carmi.
\VS{4}The
sons of
Joel;
Shemaiah his
son,
Gog his
son,
Shimei his
son,
\VS{5}Micah his
son,
Reaia his
son,
Baal his
son,
\VS{6}Beerah his
son, whom
Tilgath-pilneser
king of
Assyria carried
away
{\ADD{captive:}} he
{\ADD{was}}
prince of the
Reubenites.\FTNT{†}{{\FR 5.6 }
Tilgath-pilneser: also called, Tiglath-pileser}
\VS{7}And his
brethren by their
families, when the
genealogy of their
generations was
reckoned,
{\ADD{were}} the
chief,
Jeiel, and
Zechariah,
\VS{8}And
Bela the
son of
Azaz, the
son of
Shema, the
son of
Joel, who
dwelt in
Aroer, even unto
Nebo and
Baal-meon:\FTNT{‡}{{\FR 5.8 }
Shema: or, Shemaiah}
\VS{9}And
eastward he
inhabited unto the
entering in of the
wilderness from the
river
Euphrates: because their
cattle were
multiplied in the
land of
Gilead.
\VS{10}And in the
days of
Saul they
made
war with the
Hagarites, who
fell by their
hand: and they
dwelt in their
tents
throughout all the
east
{\ADD{land}} of
Gilead.\FTNT{§}{{\FR 5.10 }
throughout…: Heb. upon all the face of the east}
\par }{\PP \VS{11}¶ And the
children of
Gad
dwelt over against them, in the
land of
Bashan unto
Salchah:
\VS{12}Joel the
chief, and
Shapham the
next, and
Jaanai, and
Shaphat in
Bashan.
\VS{13}And their
brethren of the
house of their
fathers
{\ADD{were}},
Michael, and
Meshullam, and
Sheba, and
Jorai, and
Jachan, and
Zia, and
Heber,
seven.
\VS{14}These
{\ADD{are}} the
children of
Abihail the
son of
Huri, the
son of
Jaroah, the
son of
Gilead, the
son of
Michael, the
son of
Jeshishai, the
son of
Jahdo, the
son of
Buz;
\VS{15}Ahi the
son of
Abdiel, the
son of
Guni,
chief of the
house of their
fathers.
\VS{16}And they
dwelt in
Gilead in
Bashan, and in her
towns, and in all the
suburbs of
Sharon, upon their
borders.\FTNT{**}{{\FR 5.16 }
their…: Heb. their goings forth}
\VS{17}All these were reckoned by
genealogies in the
days of
Jotham
king of
Judah, and in the
days of
Jeroboam
king of
Israel.
\par }{\PP \VS{18}¶ The
sons of
Reuben, and the
Gadites, and
half the
tribe of
Manasseh, of valiant
men,
men able to
bear
buckler and
sword, and to
shoot with
bow, and
skilful in
war,
{\ADD{were}}
four and
forty
thousand
seven
hundred and
threescore, that went
out to the
war.\FTNT{††}{{\FR 5.18 }
valiant…: Heb. sons of valour}
\VS{19}And they
made
war with the
Hagarites, with
Jetur, and
Nephish, and
Nodab.
\VS{20}And they were
helped against them, and the
Hagarites were
delivered into their
hand, and all that
{\ADD{were}} with them: for they
cried to
God in the
battle, and he was
intreated of them; because they put their
trust in him.
\VS{21}And they took
away their
cattle; of their
camels
fifty
thousand, and of
sheep two
hundred and
fifty
thousand, and of
asses two
thousand, and of
men an
hundred
thousand.\FTNT{‡‡}{{\FR 5.21 }
took…: Heb. led captive}\FTNT{§§}{{\FR 5.21 }
men: Heb. souls of men}
\VS{22}For there
fell down
many
slain, because the
war
{\ADD{was}} of
God. And they
dwelt in their steads until the
captivity.
\par }{\PP \VS{23}¶ And the
children of the
half
tribe of
Manasseh
dwelt in the
land: they
increased from
Bashan unto
Baal-hermon and
Senir, and unto
mount
Hermon.
\VS{24}And these
{\ADD{were}} the
heads of the
house of their
fathers, even
Epher, and
Ishi, and
Eliel, and
Azriel, and
Jeremiah, and
Hodaviah, and
Jahdiel,
mighty
men of
valour,
famous
men,
{\ADD{and}}
heads of the
house of their
fathers.\FTNT{***}{{\FR 5.24 }
famous…: Heb. men of names}
\par }{\PP \VS{25}¶ And they
transgressed against the
God of their
fathers, and went a
whoring
after the
gods of the
people of the
land, whom
God
destroyed
before them.
\VS{26}And the
God of
Israel stirred
up the
spirit of
Pul
king of
Assyria, and the
spirit of
Tilgath-pilneser
king of
Assyria, and he carried them
away, even the
Reubenites, and the
Gadites, and the
half
tribe of
Manasseh, and
brought them unto
Halah, and
Habor, and
Hara, and to the
river
Gozan, unto this
day.

\par }\Chap{6}{\PP \VerseOne{1}The
sons of
Levi;
Gershon,
Kohath, and
Merari.\FTNT{*}{{\FR 6.1 }
Gershon: or, Gershom}
\VS{2}And the
sons of
Kohath;
Amram,
Izhar, and
Hebron, and
Uzziel.
\VS{3}And the
children of
Amram;
Aaron, and
Moses, and
Miriam. The
sons also of
Aaron;
Nadab, and
Abihu,
Eleazar, and
Ithamar.
\par }{\PP \VS{4}¶
Eleazar
begat
Phinehas,
Phinehas
begat
Abishua,
\VS{5}And
Abishua
begat
Bukki, and
Bukki
begat
Uzzi,
\VS{6}And
Uzzi
begat
Zerahiah, and
Zerahiah
begat
Meraioth,
\VS{7}Meraioth
begat
Amariah, and
Amariah
begat
Ahitub,
\VS{8}And
Ahitub
begat
Zadok, and
Zadok
begat
Ahimaaz,
\VS{9}And
Ahimaaz
begat
Azariah, and
Azariah
begat
Johanan,
\VS{10}And
Johanan
begat
Azariah, (he
{\ADD{it is}} that executed the
priest’s office in the
temple that
Solomon
built in
Jerusalem:)\FTNT{†}{{\FR 6.10 }
in the temple: Heb. in the house}
\VS{11}And
Azariah
begat
Amariah, and
Amariah
begat
Ahitub,
\VS{12}And
Ahitub
begat
Zadok, and
Zadok
begat
Shallum,\FTNT{‡}{{\FR 6.12 }
Shallum: or, Meshullam}
\VS{13}And
Shallum
begat
Hilkiah, and
Hilkiah
begat
Azariah,
\VS{14}And
Azariah
begat
Seraiah, and
Seraiah
begat
Jehozadak,
\VS{15}And
Jehozadak
went
{\ADD{into captivity}}, when the
{\ND{
LORD}} carried
away
Judah and
Jerusalem by the
hand of
Nebuchadnezzar.
\par }{\PP \VS{16}¶ The
sons of
Levi;
Gershom,
Kohath, and
Merari.\FTNT{§}{{\FR 6.16 }
Gershom: or, Gershon}
\VS{17}And these
{\ADD{be}} the
names of the
sons of
Gershom;
Libni, and
Shimei.
\VS{18}And the
sons of
Kohath
{\ADD{were}},
Amram, and
Izhar, and
Hebron, and
Uzziel.
\VS{19}The
sons of
Merari;
Mahli, and
Mushi. And these
{\ADD{are}} the
families of the
Levites according to their
fathers.
\VS{20}Of
Gershom;
Libni his
son,
Jahath his
son,
Zimmah his
son,
\VS{21}Joah his
son,
Iddo his
son,
Zerah his
son,
Jeaterai his
son.\FTNT{**}{{\FR 6.21 }
Joah: or, Ethan}\FTNT{††}{{\FR 6.21 }
Iddo: or, Adaiah}\FTNT{‡‡}{{\FR 6.21 }
Jeaterai: also called, Ethni}
\VS{22}The
sons of
Kohath;
Amminadab his
son,
Korah his
son,
Assir his
son,\FTNT{§§}{{\FR 6.22 }
Amminadab: or, Izhar}
\VS{23}Elkanah his
son, and
Ebiasaph his
son, and
Assir his
son,
\VS{24}Tahath his
son,
Uriel his
son,
Uzziah his
son, and
Shaul his
son.
\VS{25}And the
sons of
Elkanah;
Amasai, and
Ahimoth.
\VS{26}{\ADD{As for}}
Elkanah: the
sons of
Elkanah;
Zophai his
son, and
Nahath his
son,\FTNT{***}{{\FR 6.26 }
Zophai: or, Zuph}
\VS{27}Eliab his
son,
Jeroham his
son,
Elkanah his
son.
\VS{28}And the
sons of
Samuel; the
firstborn
Vashni, and
Abiah.\FTNT{†††}{{\FR 6.28 }
Vashni: called also Joel}
\VS{29}The
sons of
Merari;
Mahli,
Libni his
son,
Shimei his
son,
Uzza his
son,
\VS{30}Shimea his
son,
Haggiah his
son,
Asaiah his
son.
\par }{\PP \VS{31}And these
{\ADD{are they}} whom
David
set over the
service of
song in the
house of the
{\ND{
LORD}}, after that the
ark had
rest.
\VS{32}And they
ministered
before the
dwelling place of the
tabernacle of the
congregation with
singing, until
Solomon had
built the
house of the
{\ND{
LORD}} in
Jerusalem: and
{\ADD{then}} they
waited on their
office according to their
order.
\VS{33}And these
{\ADD{are}} they that
waited with their
children. Of the
sons of the
Kohathites:
Heman a
singer, the
son of
Joel, the
son of
Shemuel,\FTNT{‡‡‡}{{\FR 6.33 }
waited: Heb. stood}
\VS{34}The
son of
Elkanah, the
son of
Jeroham, the
son of
Eliel, the
son of
Toah,
\VS{35}The
son of
Zuph, the
son of
Elkanah, the
son of
Mahath, the
son of
Amasai,
\VS{36}The
son of
Elkanah, the
son of
Joel, the
son of
Azariah, the
son of
Zephaniah,
\VS{37}The
son of
Tahath, the
son of
Assir, the
son of
Ebiasaph, the
son of
Korah,
\VS{38}The
son of
Izhar, the
son of
Kohath, the
son of
Levi, the
son of
Israel.
\VS{39}And his
brother
Asaph, who
stood on his right
hand,
{\ADD{even}}
Asaph the
son of
Berachiah, the
son of
Shimea,
\VS{40}The
son of
Michael, the
son of
Baaseiah, the
son of
Malchiah,
\VS{41}The
son of
Ethni, the
son of
Zerah, the
son of
Adaiah,
\VS{42}The
son of
Ethan, the
son of
Zimmah, the
son of
Shimei,
\VS{43}The
son of
Jahath, the
son of
Gershom, the
son of
Levi.
\VS{44}And their
brethren the
sons of
Merari
{\ADD{stood}} on the left
hand:
Ethan the
son of
Kishi, the
son of
Abdi, the
son of
Malluch,\FTNT{§§§}{{\FR 6.44 }
Kishi: or, Kushaiah}
\VS{45}The
son of
Hashabiah, the
son of
Amaziah, the
son of
Hilkiah,
\VS{46}The
son of
Amzi, the
son of
Bani, the
son of
Shamer,
\VS{47}The
son of
Mahli, the
son of
Mushi, the
son of
Merari, the
son of
Levi.
\VS{48}Their
brethren also the
Levites
{\ADD{were}}
appointed unto all manner of
service of the
tabernacle of the
house of
God.
\par }{\PP \VS{49}¶ But
Aaron and his
sons
offered upon the
altar of the burnt
offering, and on the
altar of
incense,
{\ADD{and were appointed}} for all the
work of the
{\ADD{place}}
most
holy, and to make an
atonement for
Israel, according to all that
Moses the
servant of
God had
commanded.
\VS{50}And these
{\ADD{are}} the
sons of
Aaron;
Eleazar his
son,
Phinehas his
son,
Abishua his
son,
\VS{51}Bukki his
son,
Uzzi his
son,
Zerahiah his
son,
\VS{52}Meraioth his
son,
Amariah his
son,
Ahitub his
son,
\VS{53}Zadok his
son,
Ahimaaz his
son.
\par }{\PP \VS{54}¶ Now these
{\ADD{are}} their dwelling
places throughout their
castles in their
coasts, of the
sons of
Aaron, of the
families of the
Kohathites: for theirs was the
lot.
\VS{55}And they
gave them
Hebron in the
land of
Judah, and the
suburbs thereof round
about it.
\VS{56}But the
fields of the
city, and the
villages thereof, they
gave to
Caleb the
son of
Jephunneh.
\VS{57}And to the
sons of
Aaron they
gave the
cities of Judah,
{\ADD{namely}},
Hebron,
{\ADD{the city}} of
refuge, and
Libnah with her
suburbs, and
Jattir, and
Eshtemoa, with their
suburbs,
\VS{58}And
Hilen with her
suburbs,
Debir with her
suburbs,\FTNT{*}{{\FR 6.58 }
Hilen: or, Holon}
\VS{59}And
Ashan with her
suburbs, and
Beth-shemesh with her
suburbs:\FTNT{†}{{\FR 6.59 }
Ashan: or, Ain}
\VS{60}And out of the
tribe of
Benjamin;
Geba with her
suburbs, and
Alemeth with her
suburbs, and
Anathoth with her
suburbs. All their
cities throughout their
families
{\ADD{were}}
thirteen
cities.\FTNT{‡}{{\FR 6.60 }
Alemeth: or, Almon}
\VS{61}And unto the
sons of
Kohath,
{\ADD{which were}}
left of the
family of that
tribe,
{\ADD{were cities given}} out of the
half
tribe,
{\ADD{namely, out of}} the
half
{\ADD{tribe}} of
Manasseh, by
lot,
ten
cities.
\VS{62}And to the
sons of
Gershom throughout their
families out of the
tribe of
Issachar, and out of the
tribe of
Asher, and out of the
tribe of
Naphtali, and out of the
tribe of
Manasseh in
Bashan,
thirteen
cities.
\VS{63}Unto the
sons of
Merari
{\ADD{were given}} by
lot, throughout their
families, out of the
tribe of
Reuben, and out of the
tribe of
Gad, and out of the
tribe of
Zebulun,
twelve
cities.
\VS{64}And the
children of
Israel
gave to the
Levites
{\ADD{these}}
cities with their
suburbs.
\VS{65}And they
gave by
lot out of the
tribe of the
children of
Judah, and out of the
tribe of the
children of
Simeon, and out of the
tribe of the
children of
Benjamin, these
cities, which are
called by
{\ADD{their}}
names.
\VS{66}And
{\ADD{the residue}} of the
families of the
sons of
Kohath had
cities of their
coasts out of the
tribe of
Ephraim.
\VS{67}And they
gave unto them,
{\ADD{of}} the
cities of
refuge,
Shechem in
mount
Ephraim with her
suburbs;
{\ADD{they gave}} also
Gezer with her
suburbs,
\VS{68}And
Jokmeam with her
suburbs, and
Beth-horon with her
suburbs,
\VS{69}And
Aijalon with her
suburbs, and
Gath-rimmon with her
suburbs:
\VS{70}And out of the
half
tribe of
Manasseh;
Aner with her
suburbs, and
Bileam with her
suburbs, for the
family of the
remnant of the
sons of
Kohath.
\VS{71}Unto the
sons of
Gershom
{\ADD{were given}} out of the
family of the
half
tribe of
Manasseh,
Golan in
Bashan with her
suburbs, and
Ashtaroth with her
suburbs:
\VS{72}And out of the
tribe of
Issachar;
Kedesh with her
suburbs,
Daberath with her
suburbs,
\VS{73}And
Ramoth with her
suburbs, and
Anem with her
suburbs:
\VS{74}And out of the
tribe of
Asher;
Mashal with her
suburbs, and
Abdon with her
suburbs,
\VS{75}And
Hukok with her
suburbs, and
Rehob with her
suburbs:
\VS{76}And out of the
tribe of
Naphtali;
Kedesh in
Galilee with her
suburbs, and
Hammon with her
suburbs, and
Kirjathaim with her
suburbs.
\VS{77}Unto the
rest of the
children of
Merari
{\ADD{were given}} out of the
tribe of
Zebulun,
Rimmon with her
suburbs,
Tabor with her
suburbs:
\VS{78}And on the other
side
Jordan by
Jericho, on the east
side of
Jordan,
{\ADD{were given them}} out of the
tribe of
Reuben,
Bezer in the
wilderness with her
suburbs, and
Jahzah with her
suburbs,
\VS{79}Kedemoth also with her
suburbs, and
Mephaath with her
suburbs:
\VS{80}And out of the
tribe of
Gad;
Ramoth in
Gilead with her
suburbs, and
Mahanaim with her
suburbs,
\VS{81}And
Heshbon with her
suburbs, and
Jazer with her
suburbs.

\par }\Chap{7}{\PP \VerseOne{1}Now the
sons of
Issachar
{\ADD{were}},
Tola, and
Puah,
Jashub, and
Shimron,
four.
\VS{2}And the
sons of
Tola;
Uzzi, and
Rephaiah, and
Jeriel, and
Jahmai, and
Jibsam, and
Shemuel,
heads of their
father’s
house,
{\ADD{to wit}}, of
Tola:
{\ADD{they were}} valiant
men of
might in their
generations; whose
number
{\ADD{was}} in the
days of
David
two and
twenty
thousand and
six
hundred.
\VS{3}And the
sons of
Uzzi;
Izrahiah: and the
sons of
Izrahiah;
Michael, and
Obadiah, and
Joel,
Ishiah,
five: all of them chief
men.
\VS{4}And with them, by their
generations, after the
house of their
fathers,
{\ADD{were}}
bands of
soldiers for
war,
six and
thirty
thousand
{\ADD{men:}} for they had
many
wives and
sons.
\VS{5}And their
brethren among all the
families of
Issachar
{\ADD{were}}
valiant men of
might, reckoned in all by their
genealogies
fourscore and
seven
thousand.
\par }{\PP \VS{6}¶
{\ADD{The sons}} of
Benjamin;
Bela, and
Becher, and
Jediael,
three.
\VS{7}And the
sons of
Bela;
Ezbon, and
Uzzi, and
Uzziel, and
Jerimoth, and
Iri,
five;
heads of the
house of
{\ADD{their}}
fathers,
mighty men of
valour; and were reckoned by their
genealogies
twenty and
two
thousand and
thirty and
four.
\VS{8}And the
sons of
Becher;
Zemira, and
Joash, and
Eliezer, and
Elioenai, and
Omri, and
Jerimoth, and
Abiah, and
Anathoth, and
Alameth. All these
{\ADD{are}} the
sons of
Becher.
\VS{9}And the
number of them, after their genealogy by their
generations,
heads of the
house of their
fathers, mighty
men of
valour,
{\ADD{was}}
twenty
thousand and two
hundred.
\VS{10}The
sons also of
Jediael;
Bilhan: and the
sons of
Bilhan;
Jeush, and
Benjamin, and
Ehud, and
Chenaanah, and
Zethan, and
Tharshish, and
Ahishahar.
\VS{11}All these the
sons of
Jediael, by the
heads of their
fathers,
mighty men of
valour,
{\ADD{were}}
seventeen
thousand and two
hundred
{\ADD{soldiers}}, fit to go
out for
war
{\ADD{and}}
battle.
\VS{12}Shuppim also, and
Huppim, the
children of
Ir,
{\ADD{and}}
Hushim, the
sons of
Aher.\FTNT{*}{{\FR 7.12 }
Ir: or, Iri}\FTNT{†}{{\FR 7.12 }
Aher: or, Ahiram}
\par }{\PP \VS{13}¶ The
sons of
Naphtali;
Jahziel, and
Guni, and
Jezer, and
Shallum, the
sons of
Bilhah.
\par }{\PP \VS{14}¶ The
sons of
Manasseh;
Ashriel, whom she
bare: ({\ADD{but}} his
concubine the
Aramitess
bare
Machir the
father of
Gilead:
\VS{15}And
Machir
took to
wife
{\ADD{the sister}} of
Huppim and
Shuppim, whose
sister’s
name
{\ADD{was}}
Maachah;) and the
name of the
second
{\ADD{was}}
Zelophehad: and
Zelophehad had
daughters.
\VS{16}And
Maachah the
wife of
Machir
bare a
son, and she
called his
name
Peresh; and the
name of his
brother
{\ADD{was}}
Sheresh; and his
sons
{\ADD{were}}
Ulam and
Rakem.
\VS{17}And the
sons of
Ulam;
Bedan. These
{\ADD{were}} the
sons of
Gilead, the
son of
Machir, the
son of
Manasseh.
\VS{18}And his
sister
Hammoleketh
bare
Ishod, and
Abiezer, and
Mahalah.
\VS{19}And the
sons of
Shemida were,
Ahian, and
Shechem, and
Likhi, and
Aniam.
\par }{\PP \VS{20}¶ And the
sons of
Ephraim;
Shuthelah, and
Bered his
son, and
Tahath his
son, and
Eladah his
son, and
Tahath his
son,
\par }{\PP \VS{21}¶ And
Zabad his
son, and
Shuthelah his
son, and
Ezer, and
Elead, whom the
men of
Gath
{\ADD{that were}}
born in
{\ADD{that}}
land
slew, because they came
down to take
away their
cattle.
\VS{22}And
Ephraim their
father
mourned
many
days, and his
brethren
came to
comfort him.
\par }{\PP \VS{23}¶ And when he went
in to his
wife, she
conceived, and
bare a
son, and he
called his
name
Beriah, because it went
evil with his
house.
\VS{24}(And his
daughter
{\ADD{was}}
Sherah, who
built
Beth-horon the
nether, and the
upper, and
Uzzen-sherah.)
\VS{25}And
Rephah
{\ADD{was}} his
son, also
Resheph, and
Telah his
son, and
Tahan his
son,
\VS{26}Laadan his
son,
Ammihud his
son,
Elishama his
son,
\VS{27}Non his
son,
Jehoshua his
son.\FTNT{‡}{{\FR 7.27 }
Non: or, Nun}\FTNT{§}{{\FR 7.27 }
Jehoshua: or, Joshua}
\par }{\PP \VS{28}¶ And their
possessions and
habitations
{\ADD{were}},
Beth-el and the
towns thereof, and
eastward
Naaran, and
westward
Gezer, with the
towns thereof;
Shechem also and the
towns thereof, unto
Gaza and the
towns thereof:\FTNT{**}{{\FR 7.28 }
towns: Heb. daughters}\FTNT{††}{{\FR 7.28 }
unto Gaza: or, Adassa}
\VS{29}And by the
borders of the
children of
Manasseh,
Beth-shean and her
towns,
Taanach and her
towns,
Megiddo and her
towns,
Dor and her
towns. In these
dwelt the
children of
Joseph the
son of
Israel.\FTNT{‡‡}{{\FR 7.29 }
towns: Heb. daughters}
\par }{\PP \VS{30}¶ The
sons of
Asher;
Imnah, and
Isuah, and
Ishuai, and
Beriah, and
Serah their
sister.
\VS{31}And the
sons of
Beriah;
Heber, and
Malchiel, who
{\ADD{is}} the
father of
Birzavith.
\VS{32}And
Heber
begat
Japhlet, and
Shomer, and
Hotham, and
Shua their
sister.
\VS{33}And the
sons of
Japhlet;
Pasach, and
Bimhal, and
Ashvath. These
{\ADD{are}} the
children of
Japhlet.
\VS{34}And the
sons of
Shamer;
Ahi, and
Rohgah,
Jehubbah, and
Aram.
\VS{35}And the
sons of his
brother
Helem;
Zophah, and
Imna, and
Shelesh, and
Amal.
\VS{36}The
sons of
Zophah;
Suah, and
Harnepher, and
Shual, and
Beri, and
Imrah,
\VS{37}Bezer, and
Hod, and
Shamma, and
Shilshah, and
Ithran, and
Beera.
\VS{38}And the
sons of
Jether;
Jephunneh, and
Pispah, and
Ara.
\VS{39}And the
sons of
Ulla;
Arah, and
Haniel, and
Rezia.
\VS{40}All these
{\ADD{were}} the
children of
Asher,
heads of
{\ADD{their}}
father’s
house,
choice
{\ADD{and}} mighty
men of
valour,
chief of the
princes. And the
number throughout the
genealogy of them that were apt to the
war
{\ADD{and}} to
battle
{\ADD{was}}
twenty and
six
thousand
men.

\par }\Chap{8}{\PP \VerseOne{1}Now
Benjamin
begat
Bela his
firstborn,
Ashbel the
second, and
Aharah the
third,
\VS{2}Nohah the
fourth, and
Rapha the
fifth.
\VS{3}And the
sons of
Bela were,
Addar, and
Gera, and
Abihud,\FTNT{*}{{\FR 8.3 }
Addar: or, Ard}
\VS{4}And
Abishua, and
Naaman, and
Ahoah,
\VS{5}And
Gera, and
Shephuphan, and
Huram.\FTNT{†}{{\FR 8.5 }
Shephuphan: or, Shupham}
\VS{6}And these
{\ADD{are}} the
sons of
Ehud: these are the
heads of the
fathers of the
inhabitants of
Geba, and they
removed them to
Manahath:
\VS{7}And
Naaman, and
Ahiah, and
Gera, he
removed them, and
begat
Uzza, and
Ahihud.
\VS{8}And
Shaharaim
begat
{\ADD{children}} in the
country of
Moab,
after he had sent them
away;
Hushim and
Baara
{\ADD{were}} his
wives.
\VS{9}And he
begat of
Hodesh his
wife,
Jobab, and
Zibia, and
Mesha, and
Malcham,
\VS{10}And
Jeuz, and
Shachia, and
Mirma. These
{\ADD{were}} his
sons,
heads of the
fathers.
\VS{11}And of
Hushim he
begat
Abitub, and
Elpaal.
\VS{12}The
sons of
Elpaal;
Eber, and
Misham, and
Shamed, who
built
Ono, and
Lod, with the
towns thereof:
\VS{13}Beriah also, and
Shema,
who
{\ADD{were}}
heads of the
fathers of the
inhabitants of
Aijalon, who drove
away the
inhabitants of
Gath:
\VS{14}And
Ahio,
Shashak, and
Jeremoth,
\VS{15}And
Zebadiah, and
Arad, and
Ader,
\VS{16}And
Michael, and
Ispah, and
Joha, the
sons of
Beriah;
\VS{17}And
Zebadiah, and
Meshullam, and
Hezeki, and
Heber,
\VS{18}Ishmerai also, and
Jezliah, and
Jobab, the
sons of
Elpaal;
\VS{19}And
Jakim, and
Zichri, and
Zabdi,
\VS{20}And
Elienai, and
Zilthai, and
Eliel,
\VS{21}And
Adaiah, and
Beraiah, and
Shimrath, the
sons of
Shimhi;\FTNT{‡}{{\FR 8.21 }
Shimhi: or, Shema}
\VS{22}And
Ishpan, and
Heber, and
Eliel,
\VS{23}And
Abdon, and
Zichri, and
Hanan,
\VS{24}And
Hananiah, and
Elam, and
Antothijah,
\VS{25}And
Iphedeiah, and
Penuel, the
sons of
Shashak;
\VS{26}And
Shamsherai, and
Shehariah, and
Athaliah,
\VS{27}And
Jaresiah, and
Eliah, and
Zichri, the
sons of
Jeroham.
\VS{28}These
{\ADD{were}}
heads of the
fathers, by their
generations,
chief
{\ADD{men}}. These
dwelt in
Jerusalem.
\VS{29}And at
Gibeon
dwelt the
father of
Gibeon; whose
wife’s
name
{\ADD{was}}
Maachah:\FTNT{§}{{\FR 8.29 }
father…: also called Jehiel}
\VS{30}And his
firstborn
son
Abdon, and
Zur, and
Kish, and
Baal, and
Nadab,
\VS{31}And
Gedor, and
Ahio, and
Zacher.\FTNT{**}{{\FR 8.31 }
Zacher: or, Zechariah}
\VS{32}And
Mikloth
begat
Shimeah. And these
also
dwelt with their
brethren in
Jerusalem, over against
them.\FTNT{††}{{\FR 8.32 }
Shimeah: or, Shimeam}
\par }{\PP \VS{33}¶ And
Ner
begat
Kish, and
Kish
begat
Saul, and
Saul
begat
Jonathan, and
Malchi-shua, and
Abinadab, and
Esh-baal.\FTNT{‡‡}{{\FR 8.33 }
Abinadab: also called, Ishui}\FTNT{§§}{{\FR 8.33 }
Esh-baal: or, Ish-bosheth}
\VS{34}And the
son of
Jonathan
{\ADD{was}}
Merib-baal; and
Merib-baal
begat
Micah.\FTNT{***}{{\FR 8.34 }
Merib-baal: or, Mephibosheth}
\VS{35}And the
sons of
Micah
{\ADD{were}},
Pithon, and
Melech, and
Tarea, and
Ahaz.\FTNT{†††}{{\FR 8.35 }
Tarea: or, Tahrea}
\VS{36}And
Ahaz
begat
Jehoadah; and
Jehoadah
begat
Alemeth, and
Azmaveth, and
Zimri; and
Zimri
begat
Moza,\FTNT{‡‡‡}{{\FR 8.36 }
Jehoadah: also called, Jarah}
\VS{37}And
Moza
begat
Binea:
Rapha
{\ADD{was}} his
son,
Eleasah his
son,
Azel his
son:\FTNT{§§§}{{\FR 8.37 }
Rapha: also called, Rephaiah}
\VS{38}And
Azel had
six
sons, whose
names
{\ADD{are}} these,
Azrikam,
Bocheru, and
Ishmael, and
Sheariah, and
Obadiah, and
Hanan. All these
{\ADD{were}} the
sons of
Azel.
\VS{39}And the
sons of
Eshek his
brother
{\ADD{were}},
Ulam his
firstborn,
Jehush the
second, and
Eliphelet the
third.
\VS{40}And the
sons of
Ulam were
mighty
men of
valour,
archers, and had
many
sons, and
sons’
sons, an
hundred and
fifty. All these
{\ADD{are}} of the
sons of
Benjamin.

\par }\Chap{9}{\PP \VerseOne{1}So all
Israel were reckoned by
genealogies; and, behold, they
{\ADD{were}}
written in the
book of the
kings of
Israel and
Judah,
{\ADD{who}} were carried
away to
Babylon for their
transgression.
\par }{\PP \VS{2}¶ Now the
first
inhabitants that
{\ADD{dwelt}} in their
possessions in their
cities
{\ADD{were}}, the
Israelites, the
priests,
Levites, and the
Nethinims.
\VS{3}And in
Jerusalem
dwelt of the
children of
Judah, and of the
children of
Benjamin, and of the
children of
Ephraim, and
Manasseh;
\VS{4}Uthai the
son of
Ammihud, the
son of
Omri, the
son of
Imri, the
son of
Bani, of the
children of
Pharez the
son of
Judah.
\VS{5}And of the
Shilonites;
Asaiah the
firstborn, and his
sons.
\VS{6}And of the
sons of
Zerah;
Jeuel, and their
brethren,
six
hundred and
ninety.
\VS{7}And of the
sons of
Benjamin;
Sallu the
son of
Meshullam, the
son of
Hodaviah, the
son of
Hasenuah,
\VS{8}And
Ibneiah the
son of
Jeroham, and
Elah the
son of
Uzzi, the
son of
Michri, and
Meshullam the
son of
Shephathiah, the
son of
Reuel, the
son of
Ibnijah;
\VS{9}And their
brethren, according to their
generations,
nine
hundred and
fifty and
six. All these
men
{\ADD{were}}
chief of the
fathers in the
house of their
fathers.
\par }{\PP \VS{10}¶ And of the
priests;
Jedaiah, and
Jehoiarib, and
Jachin,
\VS{11}And
Azariah the
son of
Hilkiah, the
son of
Meshullam, the
son of
Zadok, the
son of
Meraioth, the
son of
Ahitub, the
ruler of the
house of
God;\FTNT{*}{{\FR 9.11 }
Azariah: also called, Seraiah}
\VS{12}And
Adaiah the
son of
Jeroham, the
son of
Pashur, the
son of
Malchijah, and
Maasiai the
son of
Adiel, the
son of
Jahzerah, the
son of
Meshullam, the
son of
Meshillemith, the
son of
Immer;
\VS{13}And their
brethren,
heads of the
house of their
fathers, a
thousand and
seven
hundred and
threescore; very
able
men for the
work of the
service of the
house of
God.\FTNT{†}{{\FR 9.13 }
very…: Heb. mighty men of valour}
\par }{\PP \VS{14}And of the
Levites;
Shemaiah the
son of
Hasshub, the
son of
Azrikam, the
son of
Hashabiah, of the
sons of
Merari;
\VS{15}And
Bakbakkar,
Heresh, and
Galal, and
Mattaniah the
son of
Micah, the
son of
Zichri, the
son of
Asaph;
\VS{16}And
Obadiah the
son of
Shemaiah, the
son of
Galal, the
son of
Jeduthun, and
Berechiah the
son of
Asa, the
son of
Elkanah, that
dwelt in the
villages of the
Netophathites.
\VS{17}And the
porters
{\ADD{were}},
Shallum, and
Akkub, and
Talmon, and
Ahiman, and their
brethren:
Shallum
{\ADD{was}} the
chief;
\VS{18}Who hitherto
{\ADD{waited}} in the
king’s
gate
eastward: they
{\ADD{were}}
porters in the
companies of the
children of
Levi.
\VS{19}And
Shallum the
son of
Kore, the
son of
Ebiasaph, the
son of
Korah, and his
brethren, of the
house of his
father, the
Korahites,
{\ADD{were}} over the
work of the
service,
keepers of the
gates of the
tabernacle: and their
fathers,
{\ADD{being}} over the
host of the
{\ND{
LORD}},
{\ADD{were}}
keepers of the
entry.\FTNT{‡}{{\FR 9.19 }
gates: Heb. thresholds}
\VS{20}And
Phinehas the
son of
Eleazar was the
ruler over them in time
past,
{\ADD{and}} the
{\ND{
LORD}}
{\ADD{was}} with him.
\VS{21}{\ADD{And}}
Zechariah the
son of
Meshelemiah
{\ADD{was}}
porter of the
door of the
tabernacle of the
congregation.
\VS{22}All these
{\ADD{which were}}
chosen to be
porters in the
gates
{\ADD{were}} two
hundred and
twelve. These were reckoned by their
genealogy in their
villages,
whom
David and
Samuel the
seer did
ordain in their set
office.\FTNT{§}{{\FR 9.22 }
did…: Heb. founded}\FTNT{**}{{\FR 9.22 }
set…: or, trust}
\VS{23}So they and their
children
{\ADD{had}} the oversight of the
gates of the
house of the
{\ND{
LORD}},
{\ADD{namely}}, the
house of the
tabernacle, by
wards.
\VS{24}In
four
quarters were the
porters, toward the
east,
west,
north, and
south.
\VS{25}And their
brethren,
{\ADD{which were}} in their
villages,
{\ADD{were}} to
come after
seven
days from
time to
time with them.
\VS{26}For these
Levites, the
four
chief
porters, were in
{\ADD{their}} set
office, and were over the
chambers and
treasuries of the
house of
God.\FTNT{††}{{\FR 9.26 }
set…: or, trust}\FTNT{‡‡}{{\FR 9.26 }
chambers: or, storehouses}
\par }{\PP \VS{27}¶ And they
lodged round
about the
house of
God, because the
charge
{\ADD{was}} upon them, and the
opening thereof every
morning
{\ADD{pertained}} to them.
\VS{28}And
{\ADD{certain}} of them had the charge
of the
ministering
vessels, that they should bring them
in and
out by
tale.\FTNT{§§}{{\FR 9.28 }
bring…: Heb. bring them in by tale, and carry them out by tale}
\VS{29}{\ADD{Some}} of them also
{\ADD{were}}
appointed to oversee the
vessels, and all the
instruments of the
sanctuary, and the fine
flour, and the
wine, and the
oil, and the
frankincense, and the
spices.\FTNT{***}{{\FR 9.29 }
instruments: or, vessels}
\VS{30}And
{\ADD{some}} of the
sons of the
priests
made the
ointment of the
spices.
\VS{31}And
Mattithiah,
{\ADD{one}} of the
Levites, who
{\ADD{was}} the
firstborn of
Shallum the
Korahite, had the set
office over the things that were
made in the
pans.\FTNT{†††}{{\FR 9.31 }
set…: or, trust}\FTNT{‡‡‡}{{\FR 9.31 }
in…: or, on flat plates, or, slices}
\VS{32}And
{\ADD{other}} of their
brethren, of the
sons of the
Kohathites,
{\ADD{were}} over the
shewbread, to
prepare
{\ADD{it}} every
sabbath.\FTNT{§§§}{{\FR 9.32 }
shewbread: Heb. bread of ordering}
\VS{33}And these
{\ADD{are}} the
singers,
chief of the
fathers of the
Levites,
{\ADD{who remaining}} in the
chambers
{\ADD{were}}
free: for they were employed in
{\ADD{that}}
work
day and
night.\FTNT{*}{{\FR 9.33 }
they…: Heb. upon them}
\VS{34}These
chief
fathers of the
Levites
{\ADD{were}}
chief throughout their
generations; these
dwelt at
Jerusalem.
\par }{\PP \VS{35}¶ And in
Gibeon
dwelt the
father of
Gibeon,
Jehiel, whose
wife’s
name
{\ADD{was}}
Maachah:
\VS{36}And his
firstborn
son
Abdon, then
Zur, and
Kish, and
Baal, and
Ner, and
Nadab,
\VS{37}And
Gedor, and
Ahio, and
Zechariah, and
Mikloth.
\VS{38}And
Mikloth
begat
Shimeam. And they also
dwelt with their
brethren at
Jerusalem, over against their
brethren.
\VS{39}And
Ner
begat
Kish; and
Kish
begat
Saul; and
Saul
begat
Jonathan, and
Malchi-shua, and
Abinadab, and
Esh-baal.
\VS{40}And the
son of
Jonathan
{\ADD{was}}
Merib-baal: and
Merib-baal
begat
Micah.
\VS{41}And the
sons of
Micah
{\ADD{were}},
Pithon, and
Melech, and
Tahrea,
{\ADD{and Ahaz}}.
\VS{42}And
Ahaz
begat
Jarah; and
Jarah
begat
Alemeth, and
Azmaveth, and
Zimri; and
Zimri
begat
Moza;
\VS{43}And
Moza
begat
Binea; and
Rephaiah his
son,
Eleasah his
son,
Azel his
son.
\VS{44}And
Azel had
six
sons, whose
names
{\ADD{are}} these,
Azrikam,
Bocheru, and
Ishmael, and
Sheariah, and
Obadiah, and
Hanan: these
{\ADD{were}} the
sons of
Azel.

\par }\Chap{10}{\PP \VerseOne{1}Now the
Philistines
fought against
Israel; and the
men of
Israel
fled from
before the
Philistines, and fell
down
slain in
mount
Gilboa.\FTNT{*}{{\FR 10.1 }
slain: or, wounded}
\VS{2}And the
Philistines followed
hard
after
Saul, and
after his
sons; and the
Philistines
slew
Jonathan, and
Abinadab, and
Malchi-shua, the
sons of
Saul.\FTNT{†}{{\FR 10.2 }
Abinadab: also called, Ishui}
\VS{3}And the
battle went
sore against
Saul, and the
archers
hit him, and he was
wounded of the
archers.\FTNT{‡}{{\FR 10.3 }
and the archers: Heb. and the shooters with bows}\FTNT{§}{{\FR 10.3 }
hit: Heb. found}
\VS{4}Then
said
Saul to his
armourbearer,
Draw thy
sword, and
thrust me through therewith; lest these
uncircumcised
come and
abuse me. But his
armourbearer
would not; for he was
sore
afraid. So
Saul
took a
sword, and
fell upon it.\FTNT{**}{{\FR 10.4 }
abuse me: or, mock me}
\VS{5}And when his
armourbearer
saw that
Saul was
dead, he
fell likewise on the
sword, and
died.
\VS{6}So
Saul
died, and his
three
sons, and all his
house
died
together.
\VS{7}And when all the
men of
Israel that
{\ADD{were}} in the
valley
saw that they
fled, and that
Saul and his
sons were
dead, then they
forsook their
cities, and
fled: and the
Philistines
came and
dwelt in them.
\par }{\PP \VS{8}¶ And it came to pass on the
morrow, when the
Philistines
came to
strip the
slain, that they
found
Saul and his
sons
fallen in
mount
Gilboa.
\VS{9}And when they had
stripped him, they
took his
head, and his
armour, and
sent into the
land of the
Philistines round
about, to carry
tidings unto their
idols, and to the
people.
\VS{10}And they
put his
armour in the
house of their
gods, and
fastened his
head in the
temple of
Dagon.
\par }{\PP \VS{11}¶ And when all
Jabesh-gilead
heard all that the
Philistines had
done to
Saul,
\VS{12}They
arose, all the
valiant
men, and took
away the
body of
Saul, and the
bodies of his
sons, and
brought them to
Jabesh, and
buried their
bones under the
oak in
Jabesh, and
fasted
seven
days.
\par }{\PP \VS{13}¶ So
Saul
died for his
transgression which he
committed against the
{\ND{
LORD}},
{\ADD{even}} against the
word of the
{\ND{
LORD}}, which he
kept not, and also for
asking
{\ADD{counsel}} of
{\ADD{one that had}} a familiar
spirit, to
enquire
{\ADD{of it;}}\FTNT{††}{{\FR 10.13 }
committed: Heb. transgressed}
\VS{14}And
enquired not of the
{\ND{
LORD}}: therefore he
slew him, and
turned the
kingdom unto
David the
son of
Jesse.\FTNT{‡‡}{{\FR 10.14 }
Jesse: Heb. Isai}

\par }\Chap{11}{\PP \VerseOne{1}Then all
Israel
gathered themselves to
David unto
Hebron,
saying, Behold, we
{\ADD{are}} thy
bone and thy
flesh.
\VS{2}And moreover in
time
past, even when
Saul was
king, thou
{\ADD{wast}} he that leddest
out and broughtest
in
Israel: and the
{\ND{
LORD}} thy
God
said unto thee, Thou shalt
feed my
people
Israel, and thou shalt be
ruler over my
people
Israel.\FTNT{*}{{\FR 11.2 }
in time…: Heb. both yesterday and the third day}\FTNT{†}{{\FR 11.2 }
feed: or, rule}
\VS{3}Therefore
came all the
elders of
Israel to the
king to
Hebron; and
David
made a
covenant with them in
Hebron
before the
{\ND{
LORD}}; and they
anointed
David
king over
Israel, according to the
word of the
{\ND{
LORD}}
by
Samuel.\FTNT{‡}{{\FR 11.3 }
by: Heb. by the hand of}
\par }{\PP \VS{4}¶ And
David and all
Israel
went to
Jerusalem, which
{\ADD{is}}
Jebus; where the
Jebusites
{\ADD{were}}, the
inhabitants of the
land.
\VS{5}And the
inhabitants of
Jebus
said to
David, Thou shalt not
come hither. Nevertheless
David
took the
castle of
Zion, which
{\ADD{is}} the
city of
David.
\VS{6}And
David
said, Whosoever
smiteth the
Jebusites
first shall be
chief and
captain. So
Joab the
son of
Zeruiah
went
first
up, and was
chief.\FTNT{§}{{\FR 11.6 }
chief: Heb. head}
\VS{7}And
David
dwelt in the
castle; therefore they
called it the
city of
David.\FTNT{**}{{\FR 11.7 }
it: that is, Zion}
\VS{8}And he
built the
city round
about, even from
Millo round
about: and
Joab
repaired the
rest of the
city.\FTNT{††}{{\FR 11.8 }
repaired: Heb. revived}
\VS{9}So
David
waxed
greater and
greater: for the
{\ND{
LORD}} of
hosts
{\ADD{was}} with him.\FTNT{‡‡}{{\FR 11.9 }
waxed…: Heb. went in going and increasing}
\par }{\PP \VS{10}¶ These also
{\ADD{are}} the
chief of the mighty
men whom
David had, who
strengthened themselves with him in his
kingdom,
{\ADD{and}} with all
Israel, to make him
king, according to the
word of the
{\ND{
LORD}} concerning
Israel.\FTNT{§§}{{\FR 11.10 }
strengthened…: or, held strongly with him}
\VS{11}And this
{\ADD{is}} the
number of the mighty
men whom
David had;
Jashobeam, an
Hachmonite, the
chief of the
captains: he lifted
up his
spear against
three
hundred
slain
{\ADD{by him}} at
one
time.\FTNT{***}{{\FR 11.11 }
an Hachmonite: or, son of Hachmoni}
\VS{12}And
after him
{\ADD{was}}
Eleazar the
son of
Dodo, the
Ahohite, who
{\ADD{was one}} of the
three
mighties.
\VS{13}He was with
David at
Pas-dammim, and there the
Philistines were gathered
together to
battle, where was a
parcel of
ground
full of
barley; and the
people
fled from
before the
Philistines.\FTNT{†††}{{\FR 11.13 }
Pas-dammim: also called, Ephes-dammim}
\VS{14}And they
set themselves in the
midst of
{\ADD{that}}
parcel, and
delivered it, and
slew the
Philistines; and the
{\ND{
LORD}}
saved
{\ADD{them}} by a
great
deliverance.\FTNT{‡‡‡}{{\FR 11.14 }
set…: or, stood}\FTNT{§§§}{{\FR 11.14 }
deliverance: or, salvation}
\par }{\PP \VS{15}¶ Now
three of the
thirty
captains went
down to the
rock to
David, into the
cave of
Adullam; and the
host of the
Philistines
encamped in the
valley of
Rephaim.\FTNT{*}{{\FR 11.15 }
three…: or, three captains over the thirty}
\VS{16}And
David
{\ADD{was}} then in the
hold, and the
Philistines’
garrison
{\ADD{was}} then at
Beth-lehem.
\VS{17}And
David
longed, and
said, Oh that one would give me
drink of the
water of the
well of
Beth-lehem, that
{\ADD{is}} at the
gate!
\VS{18}And the
three brake
through the
host of the
Philistines, and
drew
water out of the
well of
Beth-lehem, that
{\ADD{was}} by the
gate, and
took
{\ADD{it}}, and
brought
{\ADD{it}} to
David: but
David
would not
drink
{\ADD{of}} it, but poured it
out to the
{\ND{
LORD}},
\VS{19}And
said, My
God
forbid it me, that I should
do this
thing: shall I
drink the
blood of these
men that have put their lives in
jeopardy? for with
{\ADD{the jeopardy of}} their
lives they
brought it. Therefore he
would not
drink it. These things
did these
three
mightiest.\FTNT{†}{{\FR 11.19 }
that have…: Heb. with their lives?}
\par }{\PP \VS{20}¶ And
Abishai the
brother of
Joab, he was
chief of the
three: for lifting
up his
spear against
three
hundred, he
slew
{\ADD{them}}, and had a
name among the
three.
\VS{21}Of the
three, he was more
honourable than the
two; for he was their
captain: howbeit he
attained not to the
{\ADD{first}}
three.
\VS{22}Benaiah the
son of
Jehoiada, the
son of a
valiant
man of
Kabzeel, who had done
many
acts; he
slew
two lionlike
men of
Moab: also he went
down and
slew a
lion
in a
pit in a
snowy
day.\FTNT{‡}{{\FR 11.22 }
who had…: Heb. great of deeds}
\VS{23}And he
slew an
Egyptian, a
man of
{\ADD{great}}
stature,
five
cubits high; and in the
Egyptian’s
hand
{\ADD{was}} a
spear like a
weaver’s
beam; and he went
down to him with a
staff, and
plucked the
spear out of the
Egyptian’s
hand, and
slew him with his own
spear.\FTNT{§}{{\FR 11.23 }
great…: Heb. measure}
\VS{24}These
{\ADD{things}}
did
Benaiah the
son of
Jehoiada, and had the
name among the
three
mighties.
\VS{25}Behold, he was
honourable among the
thirty, but
attained not to the
{\ADD{first}}
three: and
David
set him over his
guard.
\par }{\PP \VS{26}¶ Also the valiant
men of the
armies
{\ADD{were}},
Asahel the
brother of
Joab,
Elhanan the
son of
Dodo of
Beth-lehem,
\VS{27}Shammoth the
Harorite,
Helez the
Pelonite,\FTNT{**}{{\FR 11.27 }
Shammoth: or, Shammah}\FTNT{††}{{\FR 11.27 }
Harorite: or, Harodite}\FTNT{‡‡}{{\FR 11.27 }
Pelonite: or, Paltite}
\VS{28}Ira the
son of
Ikkesh the
Tekoite,
Abi-ezer the
Antothite,
\VS{29}Sibbecai the
Hushathite,
Ilai the
Ahohite,\FTNT{§§}{{\FR 11.29 }
Sibbecai: or, Mebunnai}\FTNT{***}{{\FR 11.29 }
Ilai: or, Zalmon}
\VS{30}Maharai the
Netophathite,
Heled the
son of
Baanah the
Netophathite,\FTNT{†††}{{\FR 11.30 }
Heled: or, Heleb}
\VS{31}Ithai the
son of
Ribai of
Gibeah,
{\ADD{that pertained}} to the
children of
Benjamin,
Benaiah the
Pirathonite,
\VS{32}Hurai of the
brooks of
Gaash,
Abiel the
Arbathite,\FTNT{‡‡‡}{{\FR 11.32 }
Hurai: or, Hiddai}\FTNT{§§§}{{\FR 11.32 }
Abiel: or, Abi-albon}
\VS{33}Azmaveth the
Baharumite,
Eliahba the
Shaalbonite,
\VS{34}The
sons of
Hashem the
Gizonite,
Jonathan the
son of
Shage the
Hararite,\FTNT{*}{{\FR 11.34 }
Hashem: or, Jashen}
\VS{35}Ahiam the
son of
Sacar the
Hararite,
Eliphal the
son of
Ur,\FTNT{†}{{\FR 11.35 }
Sacar: or, Sharar}\FTNT{‡}{{\FR 11.35 }
Eliphal: or, Eliphelet}\FTNT{§}{{\FR 11.35 }
Ur: or, Ahasbai}
\VS{36}Hepher the
Mecherathite,
Ahijah the
Pelonite,
\VS{37}Hezro the
Carmelite,
Naarai the
son of
Ezbai,\FTNT{**}{{\FR 11.37 }
Hezro: or Hezrai}\FTNT{††}{{\FR 11.37 }
Naarai: or Paarai the Arbite}
\VS{38}Joel the
brother of
Nathan,
Mibhar the
son of
Haggeri,\FTNT{‡‡}{{\FR 11.38 }
the son…: or, the Haggerite}
\VS{39}Zelek the
Ammonite,
Naharai the
Berothite, the
armourbearer of
Joab the
son of
Zeruiah,
\VS{40}Ira the
Ithrite,
Gareb the
Ithrite,
\VS{41}Uriah the
Hittite,
Zabad the
son of
Ahlai,
\VS{42}Adina the
son of
Shiza the
Reubenite, a
captain of the
Reubenites, and
thirty with him,
\VS{43}Hanan the
son of
Maachah, and
Joshaphat the
Mithnite,
\VS{44}Uzzia the
Ashterathite,
Shama and
Jehiel the
sons of
Hothan the
Aroerite,
\VS{45}Jediael the
son of
Shimri, and
Joha his
brother, the
Tizite,\FTNT{§§}{{\FR 11.45 }
son…: or, Shimrite}
\VS{46}Eliel the
Mahavite, and
Jeribai, and
Joshaviah, the
sons of
Elnaam, and
Ithmah the
Moabite,
\VS{47}Eliel, and
Obed, and
Jasiel the
Mesobaite.

\par }\Chap{12}{\PP \VerseOne{1}Now these
{\ADD{are}} they that
came to
David to
Ziklag, while he yet kept himself
close
because of
Saul the
son of
Kish: and they
{\ADD{were}} among the mighty
men,
helpers of the
war.\FTNT{*}{{\FR 12.1 }
while…: Heb. being yet shut up}
\VS{2}{\ADD{They were}}
armed with
bows, and could use both the right
hand and the
left in
{\ADD{hurling}}
stones and
{\ADD{shooting}}
arrows out of a
bow,
{\ADD{even}} of
Saul’s
brethren of
Benjamin.
\VS{3}The
chief
{\ADD{was}}
Ahiezer, then
Joash, the
sons of
Shemaah the
Gibeathite; and
Jeziel, and
Pelet, the
sons of
Azmaveth; and
Berachah, and
Jehu the
Antothite,\FTNT{†}{{\FR 12.3 }
Shemaah: or, Hasmaah}
\VS{4}And
Ismaiah the
Gibeonite, a mighty
man among the
thirty, and over the
thirty; and
Jeremiah, and
Jahaziel, and
Johanan, and
Josabad the
Gederathite,
\VS{5}Eluzai, and
Jerimoth, and
Bealiah, and
Shemariah, and
Shephatiah the
Haruphite,
\VS{6}Elkanah, and
Jesiah, and
Azareel, and
Joezer, and
Jashobeam, the
Korhites,
\VS{7}And
Joelah, and
Zebadiah, the
sons of
Jeroham of
Gedor.
\VS{8}And of the
Gadites there
separated themselves unto
David into the
hold to the
wilderness
men of
might,
{\ADD{and}}
men of
war
{\ADD{fit}} for the
battle, that could
handle
shield and
buckler, whose
faces
{\ADD{were like}} the
faces of
lions, and
{\ADD{were}} as
swift as the
roes upon the
mountains;\FTNT{‡}{{\FR 12.8 }
of war: Heb. of the host}\FTNT{§}{{\FR 12.8 }
as swift…: Heb. as the roes upon the mountains to make haste}
\VS{9}Ezer the
first,
Obadiah the
second,
Eliab the
third,
\VS{10}Mishmannah the
fourth,
Jeremiah the
fifth,
\VS{11}Attai the
sixth,
Eliel the
seventh,
\VS{12}Johanan the
eighth,
Elzabad the
ninth,
\VS{13}Jeremiah the
tenth,
Machbanai the
eleventh.
\VS{14}These
{\ADD{were}} of the
sons of
Gad,
captains of the
host:
one of the
least
{\ADD{was}} over an
hundred, and the
greatest over a
thousand.\FTNT{**}{{\FR 12.14 }
one…: or, one that was least could resist an hundred, and the greatest a thousand}
\VS{15}These
{\ADD{are}} they that went
over
Jordan in the
first
month, when it had
overflown all his
banks; and they put to
flight all
{\ADD{them}} of the
valleys,
{\ADD{both}} toward the
east, and toward the
west.\FTNT{††}{{\FR 12.15 }
overflown: Heb. filled over}
\VS{16}And there
came of the
children of
Benjamin and
Judah to the
hold unto
David.
\VS{17}And
David went
out to
meet them, and
answered and
said unto them, If ye be
come
peaceably unto me to
help me, mine
heart shall be
knit unto you: but if
{\ADD{ye be come}} to
betray me to mine
enemies, seeing
{\ADD{there is}}
no
wrong in mine
hands, the
God of our
fathers
look
{\ADD{thereon}}, and
rebuke
{\ADD{it}}.\FTNT{‡‡}{{\FR 12.17 }
to meet…: Heb. before them}\FTNT{§§}{{\FR 12.17 }
be knit: Heb. be one}\FTNT{***}{{\FR 12.17 }
wrong: or, violence}
\VS{18}Then the
spirit
came upon
Amasai,
{\ADD{who was}}
chief of the
captains,
{\ADD{and he said}}, Thine
{\ADD{are we}},
David, and on thy side, thou
son of
Jesse:
peace,
peace
{\ADD{be}} unto thee, and
peace
{\ADD{be}} to thine
helpers; for thy
God
helpeth thee. Then
David
received them, and
made them
captains of the
band.\FTNT{†††}{{\FR 12.18 }
came…: Heb. clothed}
\VS{19}And there
fell
{\ADD{some}} of
Manasseh to
David, when he
came with the
Philistines against
Saul to
battle: but they
helped them not: for the
lords of the
Philistines upon
advisement
sent him away,
saying, He will
fall to his
master
Saul to
{\ADD{the jeopardy of}} our
heads.\FTNT{‡‡‡}{{\FR 12.19 }
to the…: Heb. on our heads}
\VS{20}As he
went to
Ziklag, there
fell to him of
Manasseh,
Adnah, and
Jozabad, and
Jediael, and
Michael, and
Jozabad, and
Elihu, and
Zilthai,
captains of the
thousands that
{\ADD{were}} of
Manasseh.
\VS{21}And they
helped
David against the
band
{\ADD{of the rovers:}} for they
{\ADD{were}} all mighty
men of
valour, and were
captains in the
host.\FTNT{§§§}{{\FR 12.21 }
against…: or, with a band}
\VS{22}For at
{\ADD{that}}
time
day by
day there
came to
David to
help him, until
{\ADD{it was}} a
great
host, like the
host of
God.
\par }{\PP \VS{23}¶ And these
{\ADD{are}} the
numbers of the
bands
{\ADD{that were}} ready
armed to the
war,
{\ADD{and}}
came to
David to
Hebron, to
turn the
kingdom of
Saul to him, according to the
word of the
{\ND{
LORD}}.\FTNT{*}{{\FR 12.23 }
bands: or, captains, or, men: Heb. heads}
\VS{24}The
children of
Judah that
bare
shield and
spear
{\ADD{were}}
six
thousand and
eight
hundred, ready
armed to the
war.\FTNT{†}{{\FR 12.24 }
armed: or, prepared}
\VS{25}Of the
children of
Simeon, mighty
men of
valour for the
war,
seven
thousand and one
hundred.
\VS{26}Of the
children of
Levi
four
thousand and
six
hundred.
\VS{27}And
Jehoiada
{\ADD{was}} the
leader of the
Aaronites, and with him
{\ADD{were}}
three
thousand and
seven
hundred;
\VS{28}And
Zadok, a young
man
mighty of
valour, and of his
father’s
house
twenty and
two
captains.
\VS{29}And of the
children of
Benjamin, the
kindred of
Saul,
three
thousand: for hitherto the
greatest part of them had
kept the
ward of the
house of
Saul.\FTNT{‡}{{\FR 12.29 }
kindred: Heb. brethren}\FTNT{§}{{\FR 12.29 }
the greatest…: Heb. a multitude of them}
\VS{30}And of the
children of
Ephraim
twenty
thousand and
eight
hundred,
mighty
men of
valour,
famous throughout the
house of their
fathers.\FTNT{**}{{\FR 12.30 }
famous: Heb. men of names}
\VS{31}And of the
half
tribe of
Manasseh
eighteen
thousand, which were
expressed by
name, to
come and make
David
king.
\VS{32}And of the
children of
Issachar,
{\ADD{which were men}} that had
understanding of the
times, to
know what
Israel ought to
do; the
heads of them
{\ADD{were}} two
hundred; and all their
brethren
{\ADD{were}} at their
commandment.
\VS{33}Of
Zebulun, such as went
forth to
battle,
expert in
war, with all
instruments of
war,
fifty
thousand, which could keep
rank:
{\ADD{they were}} not of
double
heart.\FTNT{††}{{\FR 12.33 }
expert…: or, rangers of battle, or, ranged in battle}\FTNT{‡‡}{{\FR 12.33 }
keep…: or, set the battle in array}\FTNT{§§}{{\FR 12.33 }
not…: Heb. without a heart and a heart}
\VS{34}And of
Naphtali a
thousand
captains, and with them with
shield and
spear
thirty and
seven
thousand.
\VS{35}And of the
Danites
expert in
war
twenty and
eight
thousand and
six
hundred.
\VS{36}And of
Asher, such as went
forth to
battle,
expert in
war,
forty
thousand.\FTNT{***}{{\FR 12.36 }
expert: or, keeping their rank}
\VS{37}And on the other
side of
Jordan, of the
Reubenites, and the
Gadites, and of the
half
tribe of
Manasseh, with all manner of
instruments of
war for the
battle, an
hundred and
twenty
thousand.
\VS{38}All these
men of
war, that could
keep
rank,
came with a
perfect
heart to
Hebron, to make
David
king over all
Israel: and all the
rest also of
Israel
{\ADD{were}} of
one
heart to make
David
king.
\VS{39}And there they were with
David
three
days,
eating and
drinking: for their
brethren had
prepared for them.
\VS{40}Moreover they that were
nigh them,
{\ADD{even}} unto
Issachar and
Zebulun and
Naphtali,
brought
bread on
asses, and on
camels, and on
mules, and on
oxen,
{\ADD{and}}
meat,
meal,
cakes of figs, and bunches of
raisins, and
wine, and
oil, and
oxen, and
sheep
abundantly: for
{\ADD{there was}}
joy in
Israel.\FTNT{†††}{{\FR 12.40 }
meat…: or, victual of meal}

\par }\Chap{13}{\PP \VerseOne{1}And
David
consulted with the
captains of
thousands and
hundreds,
{\ADD{and}} with every
leader.
\VS{2}And
David
said unto all the
congregation of
Israel, If
{\ADD{it seem}}
good unto you, and
{\ADD{that it be}} of the
{\ND{
LORD}} our
God, let us
send
abroad unto our
brethren every where,
{\ADD{that are}}
left in all the
land of
Israel, and with them
{\ADD{also}} to the
priests and
Levites
{\ADD{which are}} in their
cities
{\ADD{and}}
suburbs, that they may
gather themselves unto us:\FTNT{*}{{\FR 13.2 }
send…: Heb. break forth and send}\FTNT{†}{{\FR 13.2 }
in their…: Heb. in the cities of their suburbs}
\VS{3}And let us bring
again the
ark of our
God to us: for we
enquired not at it in the
days of
Saul.\FTNT{‡}{{\FR 13.3 }
bring…: Heb. bring about}
\VS{4}And all the
congregation
said that they would do
so: for the
thing was
right in the
eyes of all the
people.
\VS{5}So
David
gathered all
Israel
together, from
Shihor of
Egypt even unto the
entering of
Hemath, to
bring the
ark of
God from
Kirjath-jearim.
\VS{6}And
David went
up, and all
Israel, to
Baalah,
{\ADD{that is}}, to
Kirjath-jearim, which
{\ADD{belonged}} to
Judah, to bring
up thence the
ark of
God the
{\ND{
LORD}}, that
dwelleth
{\ADD{between}} the
cherubims, whose
name is
called
{\ADD{on it}}.
\VS{7}And they
carried the
ark of
God in a
new
cart out of the
house of
Abinadab: and
Uzza and
Ahio
drave the
cart.\FTNT{§}{{\FR 13.7 }
carried…: Heb. made the ark to ride}
\VS{8}And
David and all
Israel
played
before
God with all
{\ADD{their}}
might, and with
singing, and with
harps, and with
psalteries, and with
timbrels, and with
cymbals, and with
trumpets.\FTNT{**}{{\FR 13.8 }
singing: Heb. songs}
\par }{\PP \VS{9}¶ And when they
came unto the
threshingfloor of
Chidon,
Uzza put
forth his
hand to
hold the
ark; for the
oxen
stumbled.\FTNT{††}{{\FR 13.9 }
Chidon: also called Nachon}\FTNT{‡‡}{{\FR 13.9 }
stumbled: or, shook it}
\VS{10}And the
anger of the
{\ND{
LORD}} was
kindled against
Uzza, and he
smote him, because he
put his
hand to the
ark: and there he
died
before
God.
\VS{11}And
David was
displeased, because the
{\ND{
LORD}} had
made a
breach upon
Uzza: wherefore that
place is
called
Perez-uzza to this
day.\FTNT{§§}{{\FR 13.11 }
Perez-uzza: that is, The breach of Uzza}
\VS{12}And
David was
afraid of
God that
day,
saying,
How shall I
bring the
ark of
God
{\ADD{home}} to me?
\VS{13}So
David
brought not the
ark
{\ADD{home}} to himself to the
city of
David, but carried it
aside into the
house of
Obed-edom the
Gittite.\FTNT{***}{{\FR 13.13 }
brought: Heb. removed}
\VS{14}And the
ark of
God
remained with the
family of
Obed-edom in his
house
three
months. And the
{\ND{
LORD}}
blessed the
house of
Obed-edom, and all that he had.

\par }\Chap{14}{\PP \VerseOne{1}Now
Hiram
king of
Tyre
sent
messengers to
David, and
timber of
cedars, with
masons and
carpenters, to
build him an
house.
\VS{2}And
David
perceived that the
{\ND{
LORD}} had
confirmed him
king over
Israel, for his
kingdom was lifted
up on
high, because of his
people
Israel.
\par }{\PP \VS{3}¶ And
David
took more
wives at
Jerusalem: and
David
begat more
sons and
daughters.\FTNT{*}{{\FR 14.3 }
more: Heb. yet}
\VS{4}Now these
{\ADD{are}} the
names of
{\ADD{his}}
children which he had in
Jerusalem;
Shammua, and
Shobab,
Nathan, and
Solomon,
\VS{5}And
Ibhar, and
Elishua, and
Elpalet,
\VS{6}And
Nogah, and
Nepheg, and
Japhia,
\VS{7}And
Elishama, and
Beeliada, and
Eliphalet.\FTNT{†}{{\FR 14.7 }
Beeliada: also called, Eliada}
\par }{\PP \VS{8}¶ And when the
Philistines
heard that
David was
anointed
king over all
Israel, all the
Philistines went
up to
seek
David. And
David
heard
{\ADD{of it}}, and went
out
against them.
\VS{9}And the
Philistines
came and
spread themselves in the
valley of
Rephaim.
\VS{10}And
David
enquired of
God,
saying, Shall I go
up against the
Philistines? and wilt thou
deliver them into mine
hand? And the
{\ND{
LORD}}
said unto him, Go
up; for I will
deliver them into thine
hand.
\VS{11}So they came
up to
Baal-perazim; and
David
smote them there. Then
David
said,
God hath broken
in upon mine
enemies by mine
hand like the breaking
forth of
waters: therefore they
called the
name of that
place
Baal-perazim.\FTNT{‡}{{\FR 14.11 }
Baal-perazim: that is, A place of breaches}
\VS{12}And when they had
left their
gods there,
David gave a
commandment, and they were
burned with
fire.
\VS{13}And the
Philistines yet
again spread themselves
abroad in the
valley.
\VS{14}Therefore
David
enquired again of
God; and
God
said unto him, Go not
up
after them; turn
away from them, and
come upon them over
against the mulberry
trees.
\VS{15}And it shall be, when thou shalt
hear a
sound of
going in the
tops of the mulberry
trees,
{\ADD{that}} then thou shalt go
out to
battle: for
God is gone
forth
before thee to
smite the
host of the
Philistines.
\VS{16}David therefore
did as
God
commanded him: and they
smote the
host of the
Philistines from
Gibeon even to
Gazer.\FTNT{§}{{\FR 14.16 }
Gibeon: also called, Geba}
\VS{17}And the
fame of
David went
out into all
lands; and the
{\ND{
LORD}}
brought the
fear of him upon all
nations.

\par }\Chap{15}{\PP \VerseOne{1}And
{\ADD{David}}
made him
houses in the
city of
David, and
prepared a
place for the
ark of
God, and
pitched for it a
tent.
\VS{2}Then
David
said, None ought to
carry the
ark of
God but the
Levites: for them hath the
{\ND{
LORD}}
chosen to
carry the
ark of
God, and to
minister unto him
for
ever.\FTNT{*}{{\FR 15.2 }
None…: Heb. It is not to carry the ark of God, but for the Levites}
\VS{3}And
David
gathered all
Israel
together to
Jerusalem, to bring
up the
ark of the
{\ND{
LORD}} unto his
place, which he had
prepared for it.
\VS{4}And
David
assembled the
children of
Aaron, and the
Levites:
\VS{5}Of the
sons of
Kohath;
Uriel the
chief, and his
brethren an
hundred and
twenty:\FTNT{†}{{\FR 15.5 }
brethren: or, kinsmen}
\VS{6}Of the
sons of
Merari;
Asaiah the
chief, and his
brethren two
hundred and
twenty:
\VS{7}Of the
sons of
Gershom;
Joel the
chief, and his
brethren an
hundred and
thirty:
\VS{8}Of the
sons of
Elizaphan;
Shemaiah the
chief, and his
brethren two
hundred:
\VS{9}Of the
sons of
Hebron;
Eliel the
chief, and his
brethren
fourscore:
\VS{10}Of the
sons of
Uzziel;
Amminadab the
chief, and his
brethren an
hundred and
twelve.
\VS{11}And
David
called for
Zadok and
Abiathar the
priests, and for the
Levites, for
Uriel,
Asaiah, and
Joel,
Shemaiah, and
Eliel, and
Amminadab,
\VS{12}And
said unto them, Ye
{\ADD{are}} the
chief of the
fathers of the
Levites:
sanctify yourselves,
{\ADD{both}} ye and your
brethren, that ye may bring
up the
ark of the
{\ND{
LORD}}
God of
Israel unto
{\ADD{the place that}} I have
prepared for it.
\VS{13}For because ye
{\ADD{did it}} not at the
first, the
{\ND{
LORD}} our
God made a
breach upon us, for that we
sought him not after the due
order.
\VS{14}So the
priests and the
Levites
sanctified themselves to bring
up the
ark of the
{\ND{
LORD}}
God of
Israel.
\VS{15}And the
children of the
Levites
bare the
ark of
God upon their
shoulders with the
staves thereon, as
Moses
commanded according to the
word of the
{\ND{
LORD}}.
\VS{16}And
David
spake to the
chief of the
Levites to
appoint their
brethren
{\ADD{to be}} the
singers with
instruments of
musick,
psalteries and
harps and
cymbals,
sounding, by lifting
up the
voice with
joy.
\VS{17}So the
Levites
appointed
Heman the
son of
Joel; and of his
brethren,
Asaph the
son of
Berechiah; and of the
sons of
Merari their
brethren,
Ethan the
son of
Kushaiah;
\VS{18}And with them their
brethren of the
second
{\ADD{degree}},
Zechariah,
Ben, and
Jaaziel, and
Shemiramoth, and
Jehiel, and
Unni,
Eliab, and
Benaiah, and
Maaseiah, and
Mattithiah, and
Elipheleh, and
Mikneiah, and
Obed-edom, and
Jeiel, the
porters.
\VS{19}So the
singers,
Heman,
Asaph, and
Ethan,
{\ADD{were appointed}} to
sound with
cymbals of
brass;
\VS{20}And
Zechariah, and
Aziel, and
Shemiramoth, and
Jehiel, and
Unni, and
Eliab, and
Maaseiah, and
Benaiah, with
psalteries on
Alamoth;
\VS{21}And
Mattithiah, and
Elipheleh, and
Mikneiah, and
Obed-edom, and
Jeiel, and
Azaziah, with
harps on the
Sheminith to
excel.\FTNT{‡}{{\FR 15.21 }
on the…: or, on the eighth to oversee}
\VS{22}And
Chenaniah,
chief of the
Levites,
{\ADD{was}} for
song: he
instructed about the
song, because he
{\ADD{was}}
skilful.\FTNT{§}{{\FR 15.22 }
was for…: or, was for the carriage: he instructed about the carriage}\FTNT{**}{{\FR 15.22 }
song: Heb. lifting up}
\VS{23}And
Berechiah and
Elkanah
{\ADD{were}}
doorkeepers for the
ark.
\VS{24}And
Shebaniah, and
Jehoshaphat, and
Nethaneel, and
Amasai, and
Zechariah, and
Benaiah, and
Eliezer, the
priests, did
blow with the
trumpets
before the
ark of
God: and
Obed-edom and
Jehiah
{\ADD{were}}
doorkeepers for the
ark.
\par }{\PP \VS{25}¶ So
David, and the
elders of
Israel, and the
captains over
thousands,
went to bring
up the
ark of the
covenant of the
{\ND{
LORD}} out of the
house of
Obed-edom with
joy.
\VS{26}And it came to pass, when
God
helped the
Levites that
bare the
ark of the
covenant of the
{\ND{
LORD}}, that they
offered
seven
bullocks and
seven
rams.
\VS{27}And
David
{\ADD{was}}
clothed with a
robe of fine
linen, and all the
Levites that
bare the
ark, and the
singers, and
Chenaniah the
master of the
song with the
singers:
David also
{\ADD{had}} upon him an
ephod of
linen.\FTNT{††}{{\FR 15.27 }
song: or, carriage}
\VS{28}Thus all
Israel brought
up the
ark of the
covenant of the
{\ND{
LORD}} with
shouting, and with
sound of the
cornet, and with
trumpets, and with
cymbals, making a
noise with
psalteries and
harps.
\par }{\PP \VS{29}¶ And it came to pass,
{\ADD{as}} the
ark of the
covenant of the
{\ND{
LORD}}
came to the
city of
David, that
Michal the
daughter of
Saul looking
out at a
window
saw
king
David
dancing and
playing: and she
despised him in her
heart.

\par }\Chap{16}{\PP \VerseOne{1}So they
brought the
ark of
God, and
set it in the
midst of the
tent that
David had
pitched for it: and they
offered burnt
sacrifices and peace
offerings
before
God.
\VS{2}And when
David had made an
end of
offering the burnt
offerings and the peace
offerings, he
blessed the
people in the
name of the
{\ND{
LORD}}.
\VS{3}And he
dealt to every
one of
Israel, both
man and
woman, to every
one a
loaf of
bread, and a good piece of
flesh, and a
flagon
{\ADD{of wine}}.
\par }{\PP \VS{4}¶ And he
appointed
{\ADD{certain}} of the
Levites to
minister
before the
ark of the
{\ND{
LORD}}, and to
record, and to
thank and
praise the
{\ND{
LORD}}
God of
Israel:
\VS{5}Asaph the
chief, and
next to him
Zechariah,
Jeiel, and
Shemiramoth, and
Jehiel, and
Mattithiah, and
Eliab, and
Benaiah, and
Obed-edom: and
Jeiel with
psalteries and with
harps; but
Asaph made a
sound with
cymbals;\FTNT{*}{{\FR 16.5 }
with psalteries…: Heb. with instruments of psalteries and harps}
\VS{6}Benaiah also and
Jahaziel the
priests with
trumpets
continually
before the
ark of the
covenant of
God.
\par }{\PP \VS{7}¶ Then on that
day
David
delivered
first
{\ADD{this psalm}} to
thank the
{\ND{
LORD}} into the
hand of
Asaph and his
brethren.
\VS{8}Give
thanks unto the
{\ND{
LORD}},
call upon his
name, make
known his
deeds among the
people.
\VS{9}Sing unto him, sing
psalms unto him,
talk ye of all his
wondrous works.
\VS{10}Glory ye in his
holy
name: let the
heart of them
rejoice that
seek the
{\ND{
LORD}}.
\VS{11}Seek the
{\ND{
LORD}} and his
strength,
seek his
face
continually.
\VS{12}Remember his
marvellous works that he hath
done, his
wonders, and the
judgments of his
mouth;
\VS{13}O ye
seed of
Israel his
servant, ye
children of
Jacob, his chosen
ones.
\VS{14}He
{\ADD{is}} the
{\ND{
LORD}} our
God; his
judgments
{\ADD{are}} in all the
earth.
\VS{15}Be ye
mindful
always of his
covenant; the
word
{\ADD{which}} he
commanded to a
thousand
generations;
\VS{16}{\ADD{Even of the covenant}} which he
made with
Abraham, and of his
oath unto
Isaac;
\VS{17}And hath
confirmed the same to
Jacob for a
law,
{\ADD{and}} to
Israel
{\ADD{for}} an
everlasting
covenant,
\VS{18}Saying, Unto thee will I
give the
land of
Canaan, the
lot of your
inheritance;\FTNT{†}{{\FR 16.18 }
the lot: Heb. the cord}
\VS{19}When ye were but
few, even a
few, and
strangers in it.\FTNT{‡}{{\FR 16.19 }
few, even: Heb. men of number, etc}
\VS{20}And
{\ADD{when}} they
went from
nation to
nation, and from
{\ADD{one}}
kingdom to
another
people;
\VS{21}He
suffered no
man to do them
wrong: yea, he
reproved
kings for their sakes,
\VS{22}{\ADD{Saying}},
Touch not mine
anointed, and do my
prophets no
harm.
\VS{23}Sing unto the
{\ND{
LORD}}, all the
earth; shew
forth from
day to
day his
salvation.
\VS{24}Declare his
glory among the
heathen; his marvellous
works among all
nations.
\VS{25}For
great
{\ADD{is}} the
{\ND{
LORD}}, and
greatly to be
praised: he also
{\ADD{is}} to be
feared above all
gods.
\VS{26}For all the
gods of the
people
{\ADD{are}}
idols: but the
{\ND{
LORD}}
made the
heavens.
\VS{27}Glory and
honour
{\ADD{are}} in his
presence;
strength and
gladness
{\ADD{are}} in his
place.
\VS{28}Give unto the
{\ND{
LORD}}, ye
kindreds of the
people,
give unto the
{\ND{
LORD}}
glory and
strength.
\VS{29}Give unto the
{\ND{
LORD}} the
glory
{\ADD{due}} unto his
name:
bring an
offering, and
come
before him:
worship the
{\ND{
LORD}} in the
beauty of
holiness.
\VS{30}Fear
before him, all the
earth: the
world also shall be
stable, that it be not
moved.
\VS{31}Let the
heavens be
glad, and let the
earth
rejoice: and let
{\ADD{men}}
say among the
nations, The
{\ND{
LORD}}
reigneth.
\VS{32}Let the
sea
roar, and the
fulness thereof: let the
fields
rejoice, and all that
{\ADD{is}} therein.
\VS{33}Then shall the
trees of the
wood sing
out at the
presence of the
{\ND{
LORD}}, because he
cometh to
judge the
earth.
\VS{34}O give
thanks unto the
{\ND{
LORD}}; for
{\ADD{he is}}
good; for his
mercy
{\ADD{endureth}} for
ever.
\VS{35}And
say ye,
Save us, O
God of our
salvation, and gather us
together, and
deliver us from the
heathen, that we may give
thanks to thy
holy
name,
{\ADD{and}}
glory in thy
praise.
\VS{36}Blessed
{\ADD{be}} the
{\ND{
LORD}}
God of
Israel for
ever and
ever. And all the
people
said,
Amen, and
praised the
{\ND{
LORD}}.
\par }{\PP \VS{37}¶ So he
left there
before the
ark of the
covenant of the
{\ND{
LORD}}
Asaph and his
brethren, to
minister
before the
ark
continually, as every
day’s
work
required:
\VS{38}And
Obed-edom with their
brethren,
threescore and
eight;
Obed-edom also the
son of
Jeduthun and
Hosah
{\ADD{to be}}
porters:
\VS{39}And
Zadok the
priest, and his
brethren the
priests,
before the
tabernacle of the
{\ND{
LORD}} in the high
place that
{\ADD{was}} at
Gibeon,
\VS{40}To
offer burnt
offerings unto the
{\ND{
LORD}} upon the
altar of the burnt
offering
continually
morning and
evening, and
{\ADD{to do}} according to all that is
written in the
law of the
{\ND{
LORD}}, which he
commanded
Israel;\FTNT{§}{{\FR 16.40 }
morning…: Heb. in the morning, and in the evening}
\VS{41}And with them
Heman and
Jeduthun, and the
rest that were
chosen, who were
expressed by
name, to give
thanks to the
{\ND{
LORD}}, because his
mercy
{\ADD{endureth}} for
ever;
\VS{42}And with them
Heman and
Jeduthun with
trumpets and
cymbals for those that should make a
sound, and with
musical
instruments of
God. And the
sons of
Jeduthun
{\ADD{were}}
porters.\FTNT{**}{{\FR 16.42 }
porters: Heb. for the gate}
\VS{43}And all the
people
departed every
man to his
house: and
David
returned to
bless his
house.

\par }\Chap{17}{\PP \VerseOne{1}Now it came to pass, as
David
sat in his
house, that
David
said to
Nathan the
prophet, Lo, I
dwell in an
house of
cedars, but the
ark of the
covenant of the
{\ND{
LORD}}
{\ADD{remaineth}} under
curtains.
\VS{2}Then
Nathan
said unto
David,
Do all that
{\ADD{is}} in thine
heart; for
God
{\ADD{is}} with thee.
\par }{\PP \VS{3}¶ And it came to pass the same
night, that the
word of
God came to
Nathan,
saying,
\VS{4}Go and
tell
David my
servant, Thus
saith the
{\ND{
LORD}}, Thou shalt not
build me an
house to dwell
in:
\VS{5}For I have not dwelt
in an
house since the
day that I brought
up
Israel unto this
day; but have gone from
tent to
tent, and from
{\ADD{one}}
tabernacle
{\ADD{to another}}.\FTNT{*}{{\FR 17.5 }
have gone: Heb. have been}
\VS{6}Wheresoever I have
walked with all
Israel,
spake I a
word to
any of the
judges of
Israel, whom I
commanded to
feed my
people,
saying, Why have ye not
built me an
house of
cedars?
\VS{7}Now therefore thus shalt thou
say unto my
servant
David, Thus
saith the
{\ND{
LORD}} of
hosts, I
took thee from the
sheepcote,
{\ADD{even}} from
following the
sheep, that thou shouldest be
ruler over my
people
Israel:\FTNT{†}{{\FR 17.7 }
from following: Heb. from after}
\VS{8}And I have been with thee whithersoever thou hast
walked, and have cut
off all thine
enemies from
before thee, and have
made thee a
name like the
name of the great
men that
{\ADD{are}} in the
earth.
\VS{9}Also I will
ordain a
place for my
people
Israel, and will
plant them, and they shall
dwell in their place, and shall be
moved no more; neither shall the
children of
wickedness
waste them any
more, as at the
beginning,
\VS{10}And since the
time that I
commanded
judges
{\ADD{to be}} over my
people
Israel. Moreover I will
subdue all thine
enemies. Furthermore I
tell thee that the
{\ND{
LORD}} will
build thee an
house.
\par }{\PP \VS{11}¶ And it shall come to pass, when thy
days be
expired that thou must
go
{\ADD{to be}} with thy
fathers, that I will raise
up thy
seed
after thee, which shall be of thy
sons; and I will
establish his
kingdom.
\VS{12}He shall
build me an
house, and I will
stablish his
throne
for
ever.
\VS{13}I will be his
father, and he shall be my
son: and I will not
take my
mercy
away from him, as I
took
{\ADD{it}} from
{\ADD{him}} that was
before thee:
\VS{14}But I will
settle him in mine
house and in my
kingdom for
ever: and his
throne shall be
established
for
evermore.
\VS{15}According to all these
words, and according to all this
vision, so did
Nathan
speak unto
David.
\par }{\PP \VS{16}¶ And
David the
king
came and
sat
before the
{\ND{
LORD}}, and
said, Who
{\ADD{am}} I, O
{\ND{
LORD}}
God, and what
{\ADD{is}} mine
house, that thou hast
brought me
hitherto?
\VS{17}And
{\ADD{yet}} this was a small
thing in thine
eyes, O
God; for thou hast
{\ADD{also}}
spoken of thy
servant’s
house for a great while to
come, and hast
regarded me according to the
estate of a
man of high
degree, O
{\ND{
LORD}}
God.
\VS{18}What
can
David
{\ADD{speak}}
more to thee for the
honour of thy
servant? for thou
knowest thy
servant.
\VS{19}O
{\ND{
LORD}}, for thy
servant’s sake, and according to thine own
heart, hast thou
done all this
greatness, in making
known all
{\ADD{these}} great
things.\FTNT{‡}{{\FR 17.19 }
great…: Heb. greatnesses}
\VS{20}O
{\ND{
LORD}},
{\ADD{there is}} none like thee, neither
{\ADD{is there any}}
God
beside thee, according to all that we have
heard with our
ears.
\VS{21}And what
one
nation in the
earth
{\ADD{is}} like thy
people
Israel, whom
God
went to
redeem
{\ADD{to be}} his own
people, to
make thee a
name of
greatness and
terribleness, by driving
out
nations from
before thy
people, whom thou hast
redeemed out of
Egypt?
\VS{22}For thy
people
Israel didst thou
make thine own
people
for
ever; and thou,
{\ND{
LORD}}, becamest their
God.
\VS{23}Therefore now,
{\ND{
LORD}}, let the
thing that thou hast
spoken concerning thy
servant and concerning his
house be
established
for
ever, and
do as thou hast
said.
\VS{24}Let it even be
established, that thy
name may be
magnified
for
ever,
saying, The
{\ND{
LORD}} of
hosts
{\ADD{is}} the
God of
Israel,
{\ADD{even}} a
God to
Israel: and
{\ADD{let}} the
house of
David thy
servant
{\ADD{be}}
established
before thee.
\VS{25}For thou, O my
God, hast
told thy
servant that thou wilt
build him an
house: therefore thy
servant hath
found
{\ADD{in his heart}} to
pray
before thee.\FTNT{§}{{\FR 17.25 }
hast…: Heb. hast revealed the ear of thy servant}
\VS{26}And now,
{\ND{
LORD}}, thou art
God, and hast
promised this
goodness unto thy
servant:
\VS{27}Now therefore let it
please thee to
bless the
house of thy
servant, that it may be
before thee for
ever: for thou
blessest, O
{\ND{
LORD}}, and
{\ADD{it shall be}}
blessed for
ever.\FTNT{**}{{\FR 17.27 }
let…: or, it hath pleased thee}

\par }\Chap{18}{\PP \VerseOne{1}Now after
this it came to pass, that
David
smote the
Philistines, and
subdued them, and
took
Gath and her
towns out of the
hand of the
Philistines.
\VS{2}And he
smote
Moab; and the
Moabites became
David’s
servants,
{\ADD{and}}
brought
gifts.
\par }{\PP \VS{3}¶ And
David
smote
Hadarezer
king of
Zobah unto
Hamath, as he
went to
stablish his
dominion by the
river
Euphrates.\FTNT{*}{{\FR 18.3 }
Hadarezer: or, Hadadezer}
\VS{4}And
David
took from him a
thousand
chariots, and
seven
thousand
horsemen, and
twenty
thousand
footmen:
David also
houghed all the
chariot
{\ADD{horses}}, but
reserved of them an
hundred
chariots.\FTNT{†}{{\FR 18.4 }
seven…: or, seven hundred}
\VS{5}And when the
Syrians of
Damascus
came to
help
Hadarezer
king of
Zobah,
David
slew of the
Syrians
two and
twenty
thousand
men.\FTNT{‡}{{\FR 18.5 }
Damascus: Heb. Darmesek}
\VS{6}Then
David
put
{\ADD{garrisons}} in
Syria-damascus; and the
Syrians became
David’s
servants,
{\ADD{and}}
brought
gifts. Thus the
{\ND{
LORD}}
preserved
David whithersoever he
went.
\VS{7}And
David
took the
shields of
gold that were on the
servants of
Hadarezer, and
brought them to
Jerusalem.
\VS{8}Likewise from
Tibhath, and from
Chun,
cities of
Hadarezer,
brought
David
very
much
brass, wherewith
Solomon
made the
brasen
sea, and the
pillars, and the
vessels of
brass.\FTNT{§}{{\FR 18.8 }
Tibhath…: called in the book of Samuel Betah, and Berothai}
\par }{\PP \VS{9}¶ Now when
Tou
king of
Hamath
heard how
David had
smitten all the
host of
Hadarezer
king of
Zobah;\FTNT{**}{{\FR 18.9 }
Tou: also called, Toi}
\VS{10}He
sent
Hadoram his
son to
king
David, to
enquire of his
welfare, and to
congratulate him, because he had
fought against
Hadarezer, and
smitten him; (for
Hadarezer had
war with
Tou;) and
{\ADD{with him}} all manner of
vessels of
gold and
silver and
brass.\FTNT{††}{{\FR 18.10 }
Hadoram: also called, Joram}\FTNT{‡‡}{{\FR 18.10 }
to enquire…: or, to salute}\FTNT{§§}{{\FR 18.10 }
to congratulate: Heb. to bless}\FTNT{***}{{\FR 18.10 }
had war: Heb. was the man of wars}
\par }{\PP \VS{11}¶ Them also
king
David
dedicated unto the
{\ND{
LORD}}, with the
silver and the
gold that he
brought from all
{\ADD{these}}
nations; from
Edom, and from
Moab, and from the
children of
Ammon, and from the
Philistines, and from
Amalek.
\VS{12}Moreover
Abishai the
son of
Zeruiah
slew of the
Edomites in the
valley of
salt
eighteen
thousand.\FTNT{†††}{{\FR 18.12 }
Abishai: Heb. Abshai}
\par }{\PP \VS{13}¶ And he
put
garrisons in
Edom; and all the
Edomites became
David’s
servants. Thus the
{\ND{
LORD}}
preserved
David whithersoever he
went.
\par }{\PP \VS{14}¶ So
David
reigned over all
Israel, and
executed
judgment and
justice among all his
people.
\VS{15}And
Joab the
son of
Zeruiah
{\ADD{was}} over the
host; and
Jehoshaphat the
son of
Ahilud,
recorder.\FTNT{‡‡‡}{{\FR 18.15 }
recorder: or, remembrancer}
\VS{16}And
Zadok the
son of
Ahitub, and
Abimelech the
son of
Abiathar,
{\ADD{were}} the
priests; and
Shavsha was
scribe;\FTNT{§§§}{{\FR 18.16 }
Abimelech: also called, Ahimelech}\FTNT{*}{{\FR 18.16 }
Shavsha: also called Seraiah or Shisha}
\VS{17}And
Benaiah the
son of
Jehoiada
{\ADD{was}} over the
Cherethites and the
Pelethites; and the
sons of
David
{\ADD{were}}
chief
about the
king.\FTNT{†}{{\FR 18.17 }
about…: Heb. at the hand of the king}

\par }\Chap{19}{\PP \VerseOne{1}Now it came to pass after
this, that
Nahash the
king of the
children of
Ammon
died, and his
son
reigned in his stead.
\VS{2}And
David
said, I will
shew
kindness unto
Hanun the
son of
Nahash, because his
father
shewed
kindness to me. And
David
sent
messengers to
comfort him concerning his
father. So the
servants of
David
came into the
land of the
children of
Ammon to
Hanun, to
comfort him.
\VS{3}But the
princes of the
children of
Ammon
said to
Hanun,
Thinkest thou that
David doth
honour thy
father, that he hath
sent
comforters unto thee? are not his
servants
come unto thee for to
search, and to
overthrow, and to spy
out the
land?\FTNT{*}{{\FR 19.3 }
Thinkest…: Heb. In thine eyes doth David, etc}
\VS{4}Wherefore
Hanun
took
David’s
servants, and
shaved them, and cut
off their
garments in the
midst hard by their
buttocks, and sent them
away.
\VS{5}Then there
went
{\ADD{certain}}, and
told
David how the
men were served. And he
sent to
meet them: for the
men were
greatly
ashamed. And the
king
said,
Tarry at
Jericho until your
beards be
grown, and
{\ADD{then}}
return.
\par }{\PP \VS{6}¶ And when the
children of
Ammon
saw that they had made themselves
odious to
David,
Hanun and the
children of
Ammon
sent a
thousand
talents of
silver to
hire them
chariots and
horsemen out of
Mesopotamia, and out of
Syria-maachah, and out of
Zobah.\FTNT{†}{{\FR 19.6 }
odious: Heb. to stink}
\VS{7}So they
hired
thirty and
two
thousand
chariots, and the
king of
Maachah and his
people; who
came and
pitched
before
Medeba. And the
children of
Ammon gathered themselves
together from their
cities, and
came to
battle.
\VS{8}And when
David
heard
{\ADD{of it}}, he
sent
Joab, and all the
host of the mighty
men.
\VS{9}And the
children of
Ammon came
out, and put the
battle in
array before the
gate of the
city: and the
kings that were
come
{\ADD{were}} by themselves in the
field.
\VS{10}Now when
Joab
saw that the
battle was set against him
before and
behind, he chose
out of all the
choice of
Israel, and put
{\ADD{them}} in
array
against the
Syrians.\FTNT{‡}{{\FR 19.10 }
the battle…: Heb. the face of the battle was}\FTNT{§}{{\FR 19.10 }
choice: or, young men}
\VS{11}And the
rest of the
people he
delivered unto the
hand of
Abishai his
brother, and they set
{\ADD{themselves}} in
array
against the
children of
Ammon.\FTNT{**}{{\FR 19.11 }
Abishai: Heb. Abshai}
\VS{12}And he
said, If the
Syrians be too
strong for me, then thou shalt
help me: but if the
children of
Ammon be too
strong for thee, then I will
help thee.
\VS{13}Be of good
courage, and let us behave ourselves
valiantly for our
people, and for the
cities of our
God: and let the
{\ND{
LORD}}
do
{\ADD{that which is}}
good in his
sight.
\VS{14}So
Joab and the
people that
{\ADD{were}} with him drew
nigh
before the
Syrians unto the
battle; and they
fled
before him.
\VS{15}And when the
children of
Ammon
saw that the
Syrians were
fled, they likewise
fled
before
Abishai his
brother, and
entered into the
city. Then
Joab
came to
Jerusalem.
\par }{\PP \VS{16}¶ And when the
Syrians
saw that they were put to the
worse
before
Israel, they
sent
messengers, and drew
forth the
Syrians that
{\ADD{were}}
beyond the
river: and
Shophach the
captain of the
host of
Hadarezer
{\ADD{went}}
before them.\FTNT{††}{{\FR 19.16 }
river: that is, Euphrates}\FTNT{‡‡}{{\FR 19.16 }
Shophach: also called, Shobach}
\VS{17}And it was
told
David; and he
gathered all
Israel, and passed
over
Jordan, and
came upon them, and set
{\ADD{the battle}} in
array against them. So when
David had put the
battle in
array
against the
Syrians, they
fought with him.
\VS{18}But the
Syrians
fled
before
Israel; and
David
slew of the
Syrians
seven
thousand
{\ADD{men which fought in}}
chariots, and
forty
thousand
footmen, and
killed
Shophach the
captain of the
host.
\VS{19}And when the
servants of
Hadarezer
saw that they were put to the
worse
before
Israel, they made
peace with
David, and became his
servants: neither
would the
Syrians
help the
children of
Ammon any more.

\par }\Chap{20}{\PP \VerseOne{1}And it came to pass, that
after the
year was
expired, at the
time that
kings go
out
{\ADD{to battle}},
Joab led
forth the
power of the
army, and
wasted the
country of the
children of
Ammon, and
came and
besieged
Rabbah. But
David
tarried at
Jerusalem. And
Joab
smote
Rabbah, and
destroyed it.\FTNT{*}{{\FR 20.1 }
after…: Heb. at the return of the year}
\VS{2}And
David
took the
crown of their
king from off his
head, and
found it to
weigh a
talent of
gold, and
{\ADD{there were}}
precious
stones in it; and it was set upon
David’s
head: and he
brought also
exceeding
much
spoil out of the
city.\FTNT{†}{{\FR 20.2 }
to weigh: Heb. the weight of}
\VS{3}And he brought
out the
people that
{\ADD{were}} in it, and
cut
{\ADD{them}} with
saws, and with
harrows of
iron, and with
axes. Even so
dealt
David with all the
cities of the
children of
Ammon. And
David and all the
people
returned to
Jerusalem.
\par }{\PP \VS{4}¶ And it came to pass after
this, that there
arose
war at
Gezer with the
Philistines; at which
time
Sibbechai the
Hushathite
slew
Sippai,
{\ADD{that was}} of the
children of the
giant: and they were
subdued.\FTNT{‡}{{\FR 20.4 }
arose: or, continued: Heb. stood}\FTNT{§}{{\FR 20.4 }
Gezer: also called, Gob}\FTNT{**}{{\FR 20.4 }
Sippai: also called, Saph}\FTNT{††}{{\FR 20.4 }
the giant: or, Rapha}
\VS{5}And there was
war again with the
Philistines; and
Elhanan the
son of
Jair
slew
Lahmi the
brother of
Goliath the
Gittite, whose
spear
staff
{\ADD{was}} like a
weaver’s
beam.\FTNT{‡‡}{{\FR 20.5 }
Jair: also called, Jaare-oregim}
\VS{6}And yet again there was
war at
Gath, where was a
man of
{\ADD{great}}
stature, whose
fingers and
toes
{\ADD{were}}
four and
twenty,
six
{\ADD{on each hand}}, and
six
{\ADD{on each foot:}} and he also was the
son of the
giant.\FTNT{§§}{{\FR 20.6 }
great…: Heb. measure}\FTNT{***}{{\FR 20.6 }
the son…: Heb. born to the giant, or, Rapha}
\VS{7}But when he
defied
Israel,
Jonathan the
son of
Shimea
David’s
brother
slew him.\FTNT{†††}{{\FR 20.7 }
defied: or, reproached}\FTNT{‡‡‡}{{\FR 20.7 }
Shimea: also called Shammah}
\VS{8}These were
born unto the
giant in
Gath; and they
fell by the
hand of
David, and by the
hand of his
servants.

\par }\Chap{21}{\PP \VerseOne{1}And
Satan stood
up against
Israel, and
provoked
David to
number
Israel.
\VS{2}And
David
said to
Joab and to the
rulers of the
people,
Go,
number
Israel from
Beer-sheba even to
Dan; and
bring the
number of them to me, that I may
know
{\ADD{it}}.
\VS{3}And
Joab
answered, The
{\ND{
LORD}} make his
people an
hundred
times so many
more as
they
{\ADD{be:}} but, my
lord the
king,
{\ADD{are}} they not all my
lord’s
servants? why then doth my
lord
require this thing? why will he be a cause of
trespass to
Israel?
\VS{4}Nevertheless the
king’s
word
prevailed against
Joab. Wherefore
Joab
departed, and
went throughout all
Israel, and
came to
Jerusalem.
\par }{\PP \VS{5}¶ And
Joab
gave the
sum of the
number of the
people unto
David. And all
{\ADD{they of}}
Israel were a
thousand
thousand and an
hundred
thousand
men that
drew
sword: and
Judah
{\ADD{was}}
four
hundred threescore and
ten
thousand
men that
drew
sword.
\VS{6}But
Levi and
Benjamin
counted he not
among them: for the
king’s
word was
abominable to
Joab.
\par }{\PP \VS{7}And
God was
displeased with this
thing; therefore he
smote
Israel.\FTNT{*}{{\FR 21.7 }
And…: Heb. And it was evil in the eyes of the
{\ND{LORD}} concerning this thing}
\VS{8}And
David
said unto
God, I have
sinned
greatly, because I have
done this
thing: but now, I beseech thee, do
away the
iniquity of thy
servant; for I have done
very
foolishly.
\par }{\PP \VS{9}¶ And the
{\ND{
LORD}}
spake unto
Gad,
David’s
seer,
saying,
\VS{10}Go and
tell
David,
saying, Thus
saith the
{\ND{
LORD}}, I
offer thee
three
{\ADD{things:}}
choose thee
one of
them, that I may
do
{\ADD{it}} unto thee.\FTNT{†}{{\FR 21.10 }
offer: Heb. stretch out}
\VS{11}So
Gad
came to
David, and
said unto him, Thus
saith the
{\ND{
LORD}},
Choose thee\FTNT{‡}{{\FR 21.11 }
Choose…: Heb. Take to thee}
\VS{12}Either
three
years’
famine; or
three
months to be
destroyed
before thy
foes, while that the
sword of thine
enemies
overtaketh
{\ADD{thee;}}
or else
three
days the
sword of the
{\ND{
LORD}}, even the
pestilence, in the
land, and the
angel of the
{\ND{
LORD}}
destroying throughout all the
coasts of
Israel. Now therefore
advise thyself what
word I shall bring
again to him that
sent me.
\VS{13}And
David
said unto
Gad, I am in a
great
strait: let me
fall now into the
hand of the
{\ND{
LORD}}; for
very
great
{\ADD{are}} his
mercies: but let me not
fall into the
hand of
man.\FTNT{§}{{\FR 21.13 }
very great: or, very many}
\par }{\PP \VS{14}¶ So the
{\ND{
LORD}}
sent
pestilence upon
Israel: and there
fell of
Israel
seventy
thousand
men.
\VS{15}And
God
sent an
angel unto
Jerusalem to
destroy it: and as he was
destroying, the
{\ND{
LORD}}
beheld, and he
repented him of the
evil, and
said to the
angel that
destroyed, It is
enough,
stay now thine
hand. And the
angel of the
{\ND{
LORD}}
stood by the
threshingfloor of
Ornan the
Jebusite.\FTNT{**}{{\FR 21.15 }
Ornan: also called, Araunah}
\VS{16}And
David lifted
up his
eyes, and
saw the
angel of the
{\ND{
LORD}}
stand between the
earth and the
heaven, having a
drawn
sword in his
hand stretched
out over
Jerusalem. Then
David and the
elders
{\ADD{of Israel, who were}}
clothed in
sackcloth,
fell upon their
faces.
\VS{17}And
David
said unto
God,
{\ADD{Is it}} not I
{\ADD{that}}
commanded the
people to be
numbered? even I it is that have
sinned and done
evil
indeed; but
{\ADD{as for}} these
sheep, what have they
done? let thine
hand, I pray thee, O
{\ND{
LORD}} my
God, be on me, and on my
father’s
house; but not on thy
people, that they should be
plagued.
\par }{\PP \VS{18}¶ Then the
angel of the
{\ND{
LORD}}
commanded
Gad to
say to
David, that
David should go
up, and set
up an
altar unto the
{\ND{
LORD}} in the
threshingfloor of
Ornan the
Jebusite.
\VS{19}And
David went
up at the
saying of
Gad, which he
spake in the
name of the
{\ND{
LORD}}.
\VS{20}And
Ornan turned
back, and
saw the
angel; and his
four
sons with him
hid themselves. Now
Ornan was
threshing
wheat.\FTNT{††}{{\FR 21.20 }
And Ornan…: or, When Ornan turned back and saw the angel, then he and his four sons with him hid themselves}
\VS{21}And as
David
came to
Ornan,
Ornan
looked and
saw
David, and went
out of the
threshingfloor, and
bowed himself to
David with
{\ADD{his}}
face to the
ground.
\VS{22}Then
David
said to
Ornan,
Grant me the
place of
{\ADD{this}}
threshingfloor, that I may
build an
altar therein unto the
{\ND{
LORD}}: thou shalt
grant it me for the
full
price: that the
plague may be
stayed from the
people.\FTNT{‡‡}{{\FR 21.22 }
Grant: Heb. Give}
\VS{23}And
Ornan
said unto
David,
Take
{\ADD{it}} to thee, and let my
lord the
king
do
{\ADD{that which is}}
good in his
eyes:
lo, I
give
{\ADD{thee}} the
oxen
{\ADD{also}} for burnt
offerings, and the threshing
instruments for
wood, and the
wheat for the meat
offering; I
give it all.
\VS{24}And
king
David
said to
Ornan, Nay; but I will
verily
buy it for the
full
price: for I will not
take
{\ADD{that}} which
{\ADD{is}} thine for the
{\ND{
LORD}}, nor
offer burnt
offerings without
cost.
\VS{25}So
David
gave to
Ornan for the
place
six
hundred
shekels of
gold by
weight.
\VS{26}And
David
built there an
altar unto the
{\ND{
LORD}}, and
offered burnt
offerings and peace
offerings, and
called upon the
{\ND{
LORD}}; and he
answered him from
heaven by
fire upon the
altar of burnt
offering.
\VS{27}And the
{\ND{
LORD}}
commanded the
angel; and he put
up his
sword
again into the
sheath thereof.
\par }{\PP \VS{28}¶ At that
time when
David
saw that the
{\ND{
LORD}} had
answered him in the
threshingfloor of
Ornan the
Jebusite, then he
sacrificed there.
\VS{29}For the
tabernacle of the
{\ND{
LORD}}, which
Moses
made in the
wilderness, and the
altar of the burnt
offering,
{\ADD{were}} at that
season in the high
place at
Gibeon.
\VS{30}But
David
could not
go
before it to
enquire of
God: for he was
afraid
because of the
sword of the
angel of the
{\ND{
LORD}}.

\par }\Chap{22}{\PP \VerseOne{1}Then
David
said, This
{\ADD{is}} the
house of the
{\ND{
LORD}}
God, and this
{\ADD{is}} the
altar of the burnt
offering for
Israel.
\VS{2}And
David
commanded to gather
together the
strangers that
{\ADD{were}} in the
land of
Israel; and he
set
masons to
hew
wrought
stones to
build the
house of
God.
\VS{3}And
David
prepared
iron in
abundance for the
nails for the
doors of the
gates, and for the
joinings; and
brass in
abundance without
weight;
\VS{4}Also
cedar
trees in
abundance: for the
Zidonians and they of
Tyre
brought
much
cedar
wood to
David.
\VS{5}And
David
said,
Solomon my
son
{\ADD{is}}
young and
tender, and the
house
{\ADD{that is}} to be
builded for the
{\ND{
LORD}}
{\ADD{must be}}
exceeding
magnifical, of
fame and of
glory throughout all
countries: I will
{\ADD{therefore}} now make
preparation for it. So
David
prepared
abundantly
before his
death.
\par }{\PP \VS{6}¶ Then he
called for
Solomon his
son, and
charged him to
build an
house for the
{\ND{
LORD}}
God of
Israel.
\VS{7}And
David
said to
Solomon, My
son, as for me, it was in my
mind to
build an
house unto the
name of the
{\ND{
LORD}} my
God:
\VS{8}But the
word of the
{\ND{
LORD}} came to me,
saying, Thou hast
shed
blood
abundantly, and hast
made
great
wars: thou shalt not
build an
house unto my
name, because thou hast
shed
much
blood upon the
earth in my
sight.
\VS{9}Behold, a
son shall be
born to thee, who shall be a
man of
rest; and I will give him
rest from all his
enemies round
about: for his
name shall be
Solomon, and I will
give
peace and
quietness unto
Israel in his
days.\FTNT{*}{{\FR 22.9 }
Solomon: that is, Peaceable}
\VS{10}He shall
build an
house for my
name; and he shall be my
son, and I
{\ADD{will be}} his
father; and I will
establish the
throne of his
kingdom over
Israel
for
ever.
\VS{11}Now, my
son, the
{\ND{
LORD}} be with thee; and
prosper thou, and
build the
house of the
{\ND{
LORD}} thy
God, as he hath
said of thee.
\VS{12}Only the
{\ND{
LORD}}
give thee
wisdom and
understanding, and give thee
charge concerning
Israel, that thou mayest
keep the
law of the
{\ND{
LORD}} thy
God.
\VS{13}Then shalt thou
prosper, if thou takest
heed to
fulfil the
statutes and
judgments which the
{\ND{
LORD}}
charged
Moses with concerning
Israel: be
strong, and of good
courage;
dread not, nor be
dismayed.
\VS{14}Now, behold, in my
trouble I have
prepared for the
house of the
{\ND{
LORD}} an
hundred
thousand
talents of
gold, and a
thousand
thousand
talents of
silver; and of
brass and
iron without
weight; for it is in
abundance:
timber also and
stone have I
prepared; and thou mayest
add thereto.\FTNT{†}{{\FR 22.14 }
trouble: or, poverty}
\VS{15}Moreover
{\ADD{there are}}
workmen with thee in
abundance,
hewers and
workers of
stone and
timber, and all manner of cunning
men for every manner of
work.\FTNT{‡}{{\FR 22.15 }
workers…: that is, masons and carpenters}
\VS{16}Of the
gold, the
silver, and the
brass, and the
iron,
{\ADD{there is}} no
number.
Arise
{\ADD{therefore}}, and be
doing, and the
{\ND{
LORD}} be with thee.
\par }{\PP \VS{17}¶
David also
commanded all the
princes of
Israel to
help
Solomon his
son,
{\ADD{saying}},
\VS{18}{\ADD{Is}} not the
{\ND{
LORD}} your
God with you? and hath he
{\ADD{not}} given you
rest on every
side? for he hath
given the
inhabitants of the
land into mine
hand; and the
land is
subdued
before the
{\ND{
LORD}}, and
before his
people.
\VS{19}Now
set your
heart and your
soul to
seek the
{\ND{
LORD}} your
God;
arise therefore, and
build ye the
sanctuary of the
{\ND{
LORD}}
God, to
bring the
ark of the
covenant of the
{\ND{
LORD}}, and the
holy
vessels of
God, into the
house that is to be
built to the
name of the
{\ND{
LORD}}.

\par }\Chap{23}{\PP \VerseOne{1}So when
David was
old and
full of
days, he made
Solomon his
son
king over
Israel.
\par }{\PP \VS{2}¶ And he gathered
together all the
princes of
Israel, with the
priests and the
Levites.
\VS{3}Now the
Levites were
numbered from the
age of
thirty
years and
upward: and their
number by their
polls, man by
man, was
thirty and
eight
thousand.
\VS{4}Of
which,
twenty and
four
thousand
{\ADD{were}} to set
forward the
work of the
house of the
{\ND{
LORD}}; and
six
thousand
{\ADD{were}}
officers and
judges:\FTNT{*}{{\FR 23.4 }
set…: or, oversee}
\VS{5}Moreover
four
thousand
{\ADD{were}}
porters; and
four
thousand
praised the
{\ND{
LORD}} with the
instruments which I
made,
{\ADD{said David}}, to
praise
{\ADD{therewith}}.
\VS{6}And
David
divided them into
courses among the
sons of
Levi,
{\ADD{namely}},
Gershon,
Kohath, and
Merari.\FTNT{†}{{\FR 23.6 }
courses: Heb. divisions}
\par }{\PP \VS{7}¶ Of the
Gershonites
{\ADD{were}},
Laadan, and
Shimei.\FTNT{‡}{{\FR 23.7 }
Laadan: or, Libni}
\VS{8}The
sons of
Laadan; the
chief
{\ADD{was}}
Jehiel, and
Zetham, and
Joel,
three.
\VS{9}The
sons of
Shimei;
Shelomith, and
Haziel, and
Haran,
three. These
{\ADD{were}} the
chief of the
fathers of
Laadan.
\VS{10}And the
sons of
Shimei
{\ADD{were}},
Jahath,
Zina, and
Jeush, and
Beriah. These
four
{\ADD{were}} the
sons of
Shimei.\FTNT{§}{{\FR 23.10 }
Zina: or, Zizah}
\VS{11}And
Jahath was the
chief, and
Zizah the
second: but
Jeush and
Beriah had not
many
sons; therefore they were in
one
reckoning, according to
{\ADD{their}}
father’s
house.\FTNT{**}{{\FR 23.11 }
had…: Heb. did not multiply sons}
\par }{\PP \VS{12}¶ The
sons of
Kohath;
Amram,
Izhar,
Hebron, and
Uzziel,
four.
\VS{13}The
sons of
Amram;
Aaron and
Moses: and
Aaron was
separated, that he should
sanctify the
most holy
things, he and his
sons
for
ever, to burn
incense
before the
{\ND{
LORD}}, to
minister unto him, and to
bless in his
name
for
ever.
\VS{14}Now
{\ADD{concerning}}
Moses the
man of
God, his
sons were
named of the
tribe of
Levi.
\VS{15}The
sons of
Moses
{\ADD{were}},
Gershom, and
Eliezer.
\VS{16}Of the
sons of
Gershom,
Shebuel
{\ADD{was}} the
chief.\FTNT{††}{{\FR 23.16 }
Shebuel: also called, Shubael}
\VS{17}And the
sons of
Eliezer
{\ADD{were}},
Rehabiah the
chief. And
Eliezer had none
other
sons; but the
sons of
Rehabiah were
very
many.\FTNT{‡‡}{{\FR 23.17 }
the chief: or, the first}\FTNT{§§}{{\FR 23.17 }
very many: Heb. highly multiplied}
\VS{18}Of the
sons of
Izhar;
Shelomith the
chief.\FTNT{***}{{\FR 23.18 }
Shelomith: also called, Shelomoth}
\VS{19}Of the
sons of
Hebron;
Jeriah the
first,
Amariah the
second,
Jahaziel the
third, and
Jekameam the
fourth.
\VS{20}Of the
sons of
Uzziel;
Michah the
first, and
Jesiah the
second.
\par }{\PP \VS{21}¶ The
sons of
Merari;
Mahli, and
Mushi. The
sons of
Mahli;
Eleazar, and
Kish.
\VS{22}And
Eleazar
died, and had no
sons, but
daughters: and their
brethren the
sons of
Kish
took them.\FTNT{†††}{{\FR 23.22 }
brethren: or, kinsmen}
\VS{23}The
sons of
Mushi;
Mahli, and
Eder, and
Jeremoth,
three.
\par }{\PP \VS{24}¶ These
{\ADD{were}} the
sons of
Levi after the
house of their
fathers;
{\ADD{even}} the
chief of the
fathers, as they were
counted by
number of
names by their
polls, that
did the
work for the
service of the
house of the
{\ND{
LORD}}, from the
age of
twenty
years and
upward.
\VS{25}For
David
said, The
{\ND{
LORD}}
God of
Israel hath given
rest unto his
people, that they may
dwell in
Jerusalem for
ever:\FTNT{‡‡‡}{{\FR 23.25 }
that…: or, and he dwelleth in Jerusalem, etc}
\VS{26}And also unto the
Levites; they shall no
{\ADD{more}}
carry the
tabernacle, nor any
vessels of it for the
service thereof.
\VS{27}For by the
last
words of
David the
Levites
{\ADD{were}}
numbered from
twenty
years
old and
above:\FTNT{§§§}{{\FR 23.27 }
numbered: Heb. number}
\VS{28}Because their
office
{\ADD{was}} to
wait on the
sons of
Aaron for the
service of the
house of the
{\ND{
LORD}}, in the
courts, and in the
chambers, and in the
purifying of all holy
things, and the
work of the
service of the
house of
God;\FTNT{*}{{\FR 23.28 }
their…: Heb. their station was at the hand of the sons of Aaron}
\VS{29}Both for the
shewbread, and for the fine
flour for meat
offering, and for the
unleavened
cakes, and for
{\ADD{that which is baked in}} the
pan, and for that which is
fried, and for all manner of
measure and
size;\FTNT{†}{{\FR 23.29 }
pan: or, flat plate}
\VS{30}And to
stand every
morning to
thank and
praise the
{\ND{
LORD}}, and likewise at
even;
\VS{31}And to
offer all burnt
sacrifices unto the
{\ND{
LORD}} in the
sabbaths, in the new
moons, and on the set
feasts, by
number, according to the
order commanded unto them,
continually
before the
{\ND{
LORD}}:
\VS{32}And that they should
keep the
charge of the
tabernacle of the
congregation, and the
charge of the
holy
{\ADD{place}}, and the
charge of the
sons of
Aaron their
brethren, in the
service of the
house of the
{\ND{
LORD}}.

\par }\Chap{24}{\PP \VerseOne{1}Now
{\ADD{these are}} the
divisions of the
sons of
Aaron. The
sons of
Aaron;
Nadab, and
Abihu,
Eleazar, and
Ithamar.
\VS{2}But
Nadab and
Abihu
died
before their
father, and had no
children: therefore
Eleazar and
Ithamar executed the
priest’s office.
\VS{3}And
David
distributed them, both
Zadok of the
sons of
Eleazar, and
Ahimelech of the
sons of
Ithamar, according to their
offices in their
service.
\VS{4}And there were
more
chief
men
found of the
sons of
Eleazar than of the
sons of
Ithamar; and
{\ADD{thus}} were they
divided. Among the
sons of
Eleazar
{\ADD{there were}}
sixteen chief
men of the
house of
{\ADD{their}}
fathers, and
eight among the
sons of
Ithamar according to the
house of their
fathers.
\VS{5}Thus were they
divided by
lot, one
sort with
another; for the
governors of the
sanctuary, and
governors
{\ADD{of the house}} of
God, were of the
sons of
Eleazar, and of the
sons of
Ithamar.
\VS{6}And
Shemaiah the
son of
Nethaneel the
scribe,
{\ADD{one}} of the
Levites,
wrote them
before the
king, and the
princes, and
Zadok the
priest, and
Ahimelech the
son of
Abiathar, and
{\ADD{before}} the
chief of the
fathers of the
priests and
Levites:
one
principal
household being
taken for
Eleazar, and
{\ADD{one}}
taken for
Ithamar.\FTNT{*}{{\FR 24.6 }
principal…: Heb. house of the father}
\VS{7}Now the
first
lot came
forth to
Jehoiarib, the
second to
Jedaiah,
\VS{8}The
third to
Harim, the
fourth to
Seorim,
\VS{9}The
fifth to
Malchijah, the
sixth to
Mijamin,
\VS{10}The
seventh to
Hakkoz, the
eighth to
Abijah,
\VS{11}The
ninth to
Jeshua, the
tenth to
Shecaniah,
\VS{12}The
eleventh to
Eliashib, the
twelfth to
Jakim,
\VS{13}The
thirteenth to
Huppah, the
fourteenth to
Jeshebeab,
\VS{14}The
fifteenth to
Bilgah, the
sixteenth to
Immer,
\VS{15}The
seventeenth to
Hezir, the
eighteenth to
Aphses,
\VS{16}The
nineteenth to
Pethahiah, the
twentieth to
Jehezekel,
\VS{17}The
one and
twentieth to
Jachin, the
two and
twentieth to
Gamul,
\VS{18}The
three and
twentieth to
Delaiah, the
four and
twentieth to
Maaziah.
\VS{19}These
{\ADD{were}} the
orderings of them in their
service to
come into the
house of the
{\ND{
LORD}}, according to their
manner,
under
Aaron their
father, as the
{\ND{
LORD}}
God of
Israel had
commanded him.
\par }{\PP \VS{20}¶ And the
rest of the
sons of
Levi
{\ADD{were these:}} Of the
sons of
Amram;
Shubael: of the
sons of
Shubael;
Jehdeiah.\FTNT{†}{{\FR 24.20 }
Shubael: also called, Shebuel}
\VS{21}Concerning
Rehabiah: of the
sons of
Rehabiah, the
first
{\ADD{was}}
Isshiah.
\VS{22}Of the
Izharites;
Shelomoth: of the
sons of
Shelomoth;
Jahath.\FTNT{‡}{{\FR 24.22 }
Shelomoth: also called, Shelomith}
\VS{23}And the
sons
{\ADD{of Hebron;}}
Jeriah
{\ADD{the first}},
Amariah the
second,
Jahaziel the
third,
Jekameam the
fourth.
\VS{24}{\ADD{Of}} the
sons of
Uzziel;
Michah: of the
sons of
Michah;
Shamir.
\VS{25}The
brother of
Michah
{\ADD{was}}
Isshiah: of the
sons of
Isshiah;
Zechariah.
\VS{26}The
sons of
Merari
{\ADD{were}}
Mahli and
Mushi: the
sons of
Jaaziah;
Beno.
\par }{\PP \VS{27}¶ The
sons of
Merari by
Jaaziah;
Beno, and
Shoham, and
Zaccur, and
Ibri.
\VS{28}Of
Mahli
{\ADD{came}}
Eleazar, who had no
sons.
\VS{29}Concerning
Kish: the
son of
Kish
{\ADD{was}}
Jerahmeel.
\VS{30}The
sons also of
Mushi;
Mahli, and
Eder, and
Jerimoth. These
{\ADD{were}} the
sons of the
Levites after the
house of their
fathers.
\VS{31}These likewise
cast
lots over
against their
brethren the
sons of
Aaron in the
presence of
David the
king, and
Zadok, and
Ahimelech, and the
chief of the
fathers of the
priests and
Levites, even the
principal
fathers over
against their
younger
brethren.

\par }\Chap{25}{\PP \VerseOne{1}Moreover
David and the
captains of the
host
separated to the
service of the
sons of
Asaph, and of
Heman, and of
Jeduthun, who should
prophesy with
harps, with
psalteries, and with
cymbals: and the
number of the
workmen according to their
service was:
\VS{2}Of the
sons of
Asaph;
Zaccur, and
Joseph, and
Nethaniah, and
Asarelah, the
sons of
Asaph under the
hands of
Asaph, which
prophesied according to the
order of the
king.\FTNT{*}{{\FR 25.2 }
Asarelah: otherwise called Jesharelah}\FTNT{†}{{\FR 25.2 }
according…: Heb. by the hands of the king}
\VS{3}Of
Jeduthun: the
sons of
Jeduthun;
Gedaliah, and
Zeri, and
Jeshaiah,
Hashabiah, and
Mattithiah,
six, under the
hands of their
father
Jeduthun, who
prophesied with a
harp, to give
thanks and to
praise the
{\ND{
LORD}}.\FTNT{‡}{{\FR 25.3 }
Zeri: or, Izri}
\VS{4}Of
Heman: the
sons of
Heman;
Bukkiah,
Mattaniah,
Uzziel,
Shebuel, and
Jerimoth,
Hananiah,
Hanani,
Eliathah,
Giddalti, and
Romamti-ezer,
Joshbekashah,
Mallothi,
Hothir,
{\ADD{and}}
Mahazioth:\FTNT{§}{{\FR 25.4 }
Uzziel: also called, Azareel}\FTNT{**}{{\FR 25.4 }
Shebuel: also called, Shubael}
\VS{5}All these
{\ADD{were}} the
sons of
Heman the
king’s
seer in the
words of
God, to lift
up the
horn. And
God
gave to
Heman
fourteen
sons and
three
daughters.\FTNT{††}{{\FR 25.5 }
words: or, matters}
\VS{6}All these
{\ADD{were}} under the
hands of their
father for
song
{\ADD{in}} the
house of the
{\ND{
LORD}}, with
cymbals,
psalteries, and
harps, for the
service of the
house of
God, according to the
king’s
order to
Asaph,
Jeduthun, and
Heman.\FTNT{‡‡}{{\FR 25.6 }
according…: Heb. by the hands of the king}
\VS{7}So the
number of them, with their
brethren that were
instructed in the
songs of the
{\ND{
LORD}},
{\ADD{even}} all that were
cunning, was two
hundred
fourscore and
eight.
\par }{\PP \VS{8}¶ And they
cast
lots,
ward
against
{\ADD{ward}}, as well the
small as the
great, the
teacher as the
scholar.
\VS{9}Now the
first
lot came
forth for
Asaph to
Joseph: the
second to
Gedaliah, who with his
brethren and
sons
{\ADD{were}}
twelve:
\VS{10}The
third to
Zaccur,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{11}The
fourth to
Izri,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{12}The
fifth to
Nethaniah,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{13}The
sixth to
Bukkiah,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{14}The
seventh to
Jesharelah,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{15}The
eighth to
Jeshaiah,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{16}The
ninth to
Mattaniah,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{17}The
tenth to
Shimei,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{18}The
eleventh to
Azareel,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{19}The
twelfth to
Hashabiah,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{20}The
thirteenth to
Shubael,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{21}The
fourteenth to
Mattithiah,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{22}The
fifteenth to
Jeremoth,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{23}The
sixteenth to
Hananiah,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{24}The
seventeenth to
Joshbekashah,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{25}The
eighteenth to
Hanani,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{26}The
nineteenth to
Mallothi,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{27}The
twentieth to
Eliathah,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{28}The
one and
twentieth to
Hothir,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{29}The
two and
twentieth to
Giddalti,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{30}The
three and
twentieth to
Mahazioth,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve:
\VS{31}The
four and
twentieth to
Romamti-ezer,
{\ADD{he}}, his
sons, and his
brethren,
{\ADD{were}}
twelve.

\par }\Chap{26}{\PP \VerseOne{1}Concerning the
divisions of the
porters: Of the
Korhites
{\ADD{was}}
Meshelemiah the
son of
Kore, of the
sons of
Asaph.\FTNT{*}{{\FR 26.1 }
Meshelemiah: also called, Shelemiah}\FTNT{†}{{\FR 26.1 }
Asaph: also called, Ebiasaph}
\VS{2}And the
sons of
Meshelemiah
{\ADD{were}},
Zechariah the
firstborn,
Jediael the
second,
Zebadiah the
third,
Jathniel the
fourth,
\VS{3}Elam the
fifth,
Jehohanan the
sixth,
Elioenai the
seventh.
\VS{4}Moreover the
sons of
Obed-edom
{\ADD{were}},
Shemaiah the
firstborn,
Jehozabad the
second,
Joah the
third, and
Sacar the
fourth, and
Nethaneel the
fifth,
\VS{5}Ammiel the
sixth,
Issachar the
seventh,
Peulthai the
eighth: for
God
blessed him.\FTNT{‡}{{\FR 26.5 }
him: that is, Obed-edom}
\VS{6}Also unto
Shemaiah his
son were
sons
born, that
ruled throughout the
house of their
father: for they
{\ADD{were}} mighty
men of
valour.
\VS{7}The
sons of
Shemaiah;
Othni, and
Rephael, and
Obed,
Elzabad, whose
brethren
{\ADD{were}}
strong
men,
Elihu, and
Semachiah.
\VS{8}All these of the
sons of
Obed-edom: they and their
sons and their
brethren,
able
men for
strength for the
service,
{\ADD{were}}
threescore and
two of
Obed-edom.
\VS{9}And
Meshelemiah had
sons and
brethren,
strong
men,
eighteen.
\VS{10}Also
Hosah, of the
children of
Merari, had
sons;
Simri the
chief, (for
{\ADD{though}} he was not the
firstborn, yet his
father
made him the
chief;)
\VS{11}Hilkiah the
second,
Tebaliah the
third,
Zechariah the
fourth: all the
sons and
brethren of
Hosah
{\ADD{were}}
thirteen.
\VS{12}Among these
{\ADD{were}} the
divisions of the
porters,
{\ADD{even}} among the
chief
men,
{\ADD{having}}
wards one
against
another, to
minister in the
house of the
{\ND{
LORD}}.
\par }{\PP \VS{13}¶ And they
cast
lots, as well the
small as the
great, according to the
house of their
fathers, for every
gate.\FTNT{§}{{\FR 26.13 }
as well…: or, as well for the small as for the great}
\VS{14}And the
lot
eastward
fell to
Shelemiah. Then for
Zechariah his
son, a
wise
counsellor, they
cast
lots; and his
lot came
out
northward.\FTNT{**}{{\FR 26.14 }
Shelemiah: also called Meshelemiah}
\VS{15}To
Obed-edom
southward; and to his
sons the
house of
Asuppim.\FTNT{††}{{\FR 26.15 }
Asuppim: Heb. gatherings}
\VS{16}To
Shuppim and
Hosah
{\ADD{the lot came forth}}
westward, with the
gate
Shallecheth, by the
causeway of the going
up,
ward
against
ward.
\VS{17}Eastward
{\ADD{were}}
six
Levites,
northward
four a
day,
southward
four a
day, and toward
Asuppim
two
{\ADD{and}}
two.
\VS{18}At
Parbar
westward,
four at the
causeway,
{\ADD{and}}
two at
Parbar.
\VS{19}These
{\ADD{are}} the
divisions of the
porters among the
sons of
Kore, and among the
sons of
Merari.
\par }{\PP \VS{20}¶ And of the
Levites,
Ahijah
{\ADD{was}} over the
treasures of the
house of
God, and over the
treasures of the dedicated
things.\FTNT{‡‡}{{\FR 26.20 }
dedicated…: Heb. holy things}
\VS{21}{\ADD{As concerning}} the
sons of
Laadan; the
sons of the
Gershonite
Laadan,
chief
fathers,
{\ADD{even}} of
Laadan the
Gershonite,
{\ADD{were}}
Jehieli.\FTNT{§§}{{\FR 26.21 }
Laadan: also called, Libni}\FTNT{***}{{\FR 26.21 }
Jehieli: also called, Jehiel}
\VS{22}The
sons of
Jehieli;
Zetham, and
Joel his
brother,
{\ADD{which were}} over the
treasures of the
house of the
{\ND{
LORD}}.
\VS{23}Of the
Amramites,
{\ADD{and}} the
Izharites, the
Hebronites,
{\ADD{and}} the
Uzzielites:
\VS{24}And
Shebuel the
son of
Gershom, the
son of
Moses,
{\ADD{was}}
ruler of the
treasures.
\VS{25}And his
brethren by
Eliezer;
Rehabiah his
son, and
Jeshaiah his
son, and
Joram his
son, and
Zichri his
son, and
Shelomith his
son.
\VS{26}Which
Shelomith and his
brethren
{\ADD{were}} over all the
treasures of the dedicated
things, which
David the
king, and the
chief
fathers, the
captains over
thousands and
hundreds, and the
captains of the
host, had
dedicated.
\VS{27}Out of the
spoils won in
battles did they
dedicate to
maintain the
house of the
{\ND{
LORD}}.\FTNT{†††}{{\FR 26.27 }
spoils…: Heb. battles and spoils}
\VS{28}And all that
Samuel the
seer, and
Saul the
son of
Kish, and
Abner the
son of
Ner, and
Joab the
son of
Zeruiah, had
dedicated;
{\ADD{and}} whosoever had
dedicated
{\ADD{any thing, it was}} under the
hand of
Shelomith, and of his
brethren.
\par }{\PP \VS{29}¶ Of the
Izharites,
Chenaniah and his
sons
{\ADD{were}} for the
outward
business over
Israel, for
officers and
judges.
\VS{30}{\ADD{And}} of the
Hebronites,
Hashabiah and his
brethren,
men of
valour, a
thousand and
seven
hundred,
{\ADD{were}}
officers among them of
Israel on this
side
Jordan
westward in all the
business of the
{\ND{
LORD}}, and in the
service of the
king.\FTNT{‡‡‡}{{\FR 26.30 }
officers…: Heb. over the charge}
\VS{31}Among the
Hebronites
{\ADD{was}}
Jerijah the
chief,
{\ADD{even}} among the
Hebronites, according to the
generations of his
fathers. In the
fortieth
year of the
reign of
David they were
sought for, and there were
found among them mighty
men of
valour at
Jazer of
Gilead.
\VS{32}And his
brethren,
men of
valour,
{\ADD{were}} two
thousand and
seven
hundred
chief
fathers, whom
king
David made
rulers over the
Reubenites, the
Gadites, and the
half
tribe of
Manasseh, for every
matter pertaining to
God, and
affairs of the
king.\FTNT{§§§}{{\FR 26.32 }
affairs: Heb. thing}

\par }\Chap{27}{\PP \VerseOne{1}Now the
children of
Israel after their
number,
{\ADD{to wit}}, the
chief
fathers and
captains of
thousands and
hundreds, and their
officers that
served the
king in any
matter of the
courses, which came
in and went
out
month by
month throughout all the
months of the
year, of
every
course
{\ADD{were}}
twenty and
four
thousand.
\VS{2}Over the
first
course for the
first
month
{\ADD{was}}
Jashobeam the
son of
Zabdiel: and in his
course
{\ADD{were}}
twenty and
four
thousand.
\VS{3}Of the
children of
Perez
{\ADD{was}} the
chief of all the
captains of the
host for the
first
month.
\VS{4}And over the
course of the
second
month
{\ADD{was}}
Dodai an
Ahohite, and of his
course
{\ADD{was}}
Mikloth also the
ruler: in his
course likewise
{\ADD{were}}
twenty and
four
thousand.\FTNT{*}{{\FR 27.4 }
Dodai: also called, Dodo}
\VS{5}The
third
captain of the
host for the
third
month
{\ADD{was}}
Benaiah the
son of
Jehoiada, a
chief
priest: and in his
course
{\ADD{were}}
twenty and
four
thousand.\FTNT{†}{{\FR 27.5 }
chief…: or, principal officer}
\VS{6}This
{\ADD{is that}}
Benaiah,
{\ADD{who was}}
mighty
{\ADD{among}} the
thirty, and above the
thirty: and in his
course
{\ADD{was}}
Ammizabad his
son.
\VS{7}The
fourth
{\ADD{captain}} for the
fourth
month
{\ADD{was}}
Asahel the
brother of
Joab, and
Zebadiah his
son
after him: and in his
course
{\ADD{were}}
twenty and
four
thousand.
\VS{8}The
fifth
captain for the
fifth
month
{\ADD{was}}
Shamhuth the
Izrahite: and in his
course
{\ADD{were}}
twenty and
four
thousand.
\VS{9}The
sixth
{\ADD{captain}} for the
sixth
month
{\ADD{was}}
Ira the
son of
Ikkesh the
Tekoite: and in his
course
{\ADD{were}}
twenty and
four
thousand.
\VS{10}The
seventh
{\ADD{captain}} for the
seventh
month
{\ADD{was}}
Helez the
Pelonite, of the
children of
Ephraim: and in his
course
{\ADD{were}}
twenty and
four
thousand.
\VS{11}The
eighth
{\ADD{captain}} for the
eighth
month
{\ADD{was}}
Sibbecai the
Hushathite, of the
Zarhites: and in his
course
{\ADD{were}}
twenty and
four
thousand.
\VS{12}The
ninth
{\ADD{captain}} for the
ninth
month
{\ADD{was}}
Abi-ezer the
Anetothite, of the
Benjamites: and in his
course
{\ADD{were}}
twenty and
four
thousand.
\VS{13}The
tenth
{\ADD{captain}} for the
tenth
month
{\ADD{was}}
Maharai the
Netophathite, of the
Zarhites: and in his
course
{\ADD{were}}
twenty and
four
thousand.
\VS{14}The
eleventh
{\ADD{captain}} for the
eleventh
month
{\ADD{was}}
Benaiah the
Pirathonite, of the
children of
Ephraim: and in his
course
{\ADD{were}}
twenty and
four
thousand.
\VS{15}The
twelfth
{\ADD{captain}} for the
twelfth
month
{\ADD{was}}
Heldai the
Netophathite, of
Othniel: and in his
course
{\ADD{were}}
twenty and
four
thousand.\FTNT{‡}{{\FR 27.15 }
Heldai: also called, Heled}
\par }{\PP \VS{16}¶ Furthermore over the
tribes of
Israel: the
ruler of the
Reubenites
{\ADD{was}}
Eliezer the
son of
Zichri: of the
Simeonites,
Shephatiah the
son of
Maachah:
\VS{17}Of the
Levites,
Hashabiah the
son of
Kemuel: of the
Aaronites,
Zadok:
\VS{18}Of
Judah,
Elihu,
{\ADD{one}} of the
brethren of
David: of
Issachar,
Omri the
son of
Michael:\FTNT{§}{{\FR 27.18 }
Elihu: also called, Eliab}
\VS{19}Of
Zebulun,
Ishmaiah the
son of
Obadiah: of
Naphtali,
Jerimoth the
son of
Azriel:
\VS{20}Of the
children of
Ephraim,
Hoshea the
son of
Azaziah: of the
half
tribe of
Manasseh,
Joel the
son of
Pedaiah:
\VS{21}Of the
half
{\ADD{tribe}} of
Manasseh in
Gilead,
Iddo the
son of
Zechariah: of
Benjamin,
Jaasiel the
son of
Abner:
\VS{22}Of
Dan,
Azareel the
son of
Jeroham. These
{\ADD{were}} the
princes of the
tribes of
Israel.
\par }{\PP \VS{23}¶ But
David
took not the
number of them from
twenty
years
old and
under: because the
{\ND{
LORD}} had
said he would
increase
Israel like to the
stars of the
heavens.
\VS{24}Joab the
son of
Zeruiah
began to
number, but he
finished not, because there fell
wrath for
it against
Israel; neither was the
number
put in the
account of the
chronicles of
king
David.\FTNT{**}{{\FR 27.24 }
was: Heb. ascended}
\par }{\PP \VS{25}¶ And over the
king’s
treasures
{\ADD{was}}
Azmaveth the
son of
Adiel: and over the
storehouses in the
fields, in the
cities, and in the
villages, and in the
castles,
{\ADD{was}}
Jehonathan the
son of
Uzziah:
\VS{26}And over them that
did the
work of the
field for
tillage of the
ground
{\ADD{was}}
Ezri the
son of
Chelub:
\VS{27}And over the
vineyards
{\ADD{was}}
Shimei the
Ramathite: over the increase of the
vineyards for the
wine
cellars
{\ADD{was}}
Zabdi the
Shiphmite:\FTNT{††}{{\FR 27.27 }
over the increase…: Heb. over that which was of the vineyards}
\VS{28}And over the olive
trees and the sycomore
trees that
{\ADD{were}} in the low
plains
{\ADD{was}}
Baal-hanan the
Gederite: and over the
cellars of
oil
{\ADD{was}}
Joash:
\VS{29}And over the
herds that
fed in
Sharon
{\ADD{was}}
Shitrai the
Sharonite: and over the
herds
{\ADD{that were}} in the
valleys
{\ADD{was}}
Shaphat the
son of
Adlai:
\VS{30}Over the
camels also
{\ADD{was}}
Obil the
Ishmaelite: and over the
asses
{\ADD{was}}
Jehdeiah the
Meronothite:
\VS{31}And over the
flocks
{\ADD{was}}
Jaziz the
Hagerite. All these
{\ADD{were}} the
rulers of the
substance which
{\ADD{was}}
king
David’s.
\VS{32}Also
Jonathan
David’s
uncle was a
counsellor, a
wise
man, and a
scribe: and
Jehiel the
son of
Hachmoni
{\ADD{was}} with the
king’s
sons:\FTNT{‡‡}{{\FR 27.32 }
scribe: or, secretary}\FTNT{§§}{{\FR 27.32 }
son…: or, Hachmonite}
\VS{33}And
Ahithophel
{\ADD{was}} the
king’s
counsellor: and
Hushai the
Archite
{\ADD{was}} the
king’s
companion:
\VS{34}And
after
Ahithophel
{\ADD{was}}
Jehoiada the
son of
Benaiah, and
Abiathar: and the
general of the
king’s
army
{\ADD{was}}
Joab.

\par }\Chap{28}{\PP \VerseOne{1}And
David
assembled all the
princes of
Israel, the
princes of the
tribes, and the
captains of the
companies that
ministered to the
king by
course, and the
captains over the
thousands, and
captains over the
hundreds, and the
stewards over all the
substance and
possession of the
king, and of his
sons, with the
officers, and with the mighty
men, and with all the valiant
men, unto
Jerusalem.\FTNT{*}{{\FR 28.1 }
possession: or, cattle}\FTNT{†}{{\FR 28.1 }
and of…: or, and his sons}\FTNT{‡}{{\FR 28.1 }
officers: or, eunuchs}
\VS{2}Then
David the
king stood
up upon his
feet, and
said,
Hear me, my
brethren, and my
people:
{\ADD{As for me}}, I
{\ADD{had}} in mine
heart to
build an
house of
rest for the
ark of the
covenant of the
{\ND{
LORD}}, and for the
footstool of our
God, and had made
ready for the
building:
\VS{3}But
God
said unto me, Thou shalt not
build an
house for my
name, because thou
{\ADD{hast been}} a
man of
war, and hast
shed
blood.\FTNT{§}{{\FR 28.3 }
blood: Heb. bloods}
\VS{4}Howbeit the
{\ND{
LORD}}
God of
Israel
chose me before all the
house of my
father to be
king over
Israel for
ever: for he hath
chosen
Judah
{\ADD{to be}} the
ruler; and of the
house of
Judah, the
house of my
father; and among the
sons of my
father he
liked me to make
{\ADD{me}}
king over all
Israel:
\VS{5}And of all my
sons, (for the
{\ND{
LORD}} hath
given me
many
sons,) he hath
chosen
Solomon my
son to
sit upon the
throne of the
kingdom of the
{\ND{
LORD}} over
Israel.
\VS{6}And he
said unto me,
Solomon thy
son, he shall
build my
house and my
courts: for I have
chosen him
{\ADD{to be}} my
son, and I will be his
father.
\VS{7}Moreover I will
establish his
kingdom for
ever, if he be
constant to
do my
commandments and my
judgments, as at this
day.\FTNT{**}{{\FR 28.7 }
constant: Heb. strong}
\VS{8}Now therefore in the
sight of all
Israel the
congregation of the
{\ND{
LORD}}, and in the
audience of our
God,
keep and
seek for all the
commandments of the
{\ND{
LORD}} your
God: that ye may
possess this
good
land, and leave
{\ADD{it}} for an
inheritance for your
children
after you
for
ever.
\par }{\PP \VS{9}¶ And thou,
Solomon my
son,
know thou the
God of thy
father, and
serve him with a
perfect
heart and with a
willing
mind: for the
{\ND{
LORD}}
searcheth all
hearts, and
understandeth all the
imaginations of the
thoughts: if thou
seek him, he will be
found of thee; but if thou
forsake him, he will cast thee
off for
ever.
\VS{10}Take
heed now; for the
{\ND{
LORD}} hath
chosen thee to
build an
house for the
sanctuary: be
strong, and
do
{\ADD{it}}.
\par }{\PP \VS{11}¶ Then
David
gave to
Solomon his
son the
pattern of the
porch, and of the
houses thereof, and of the
treasuries thereof, and of the upper
chambers thereof, and of the
inner
parlours thereof, and of the
place of the mercy
seat,
\VS{12}And the
pattern of all that he had by the
spirit, of the
courts of the
house of the
{\ND{
LORD}}, and of all the
chambers round
about, of the
treasuries of the
house of
God, and of the
treasuries of the dedicated
things:\FTNT{††}{{\FR 28.12 }
of all that…: Heb. of all that was with him}
\VS{13}Also for the
courses of the
priests and the
Levites, and for all the
work of the
service of the
house of the
{\ND{
LORD}}, and for all the
vessels of
service in the
house of the
{\ND{
LORD}}.
\VS{14}{\ADD{He gave}} of
gold by
weight for
{\ADD{things}} of
gold, for all
instruments of all manner of
service;
{\ADD{silver also}} for all
instruments of
silver by
weight, for all
instruments of every kind of
service:
\VS{15}Even the
weight for the
candlesticks of
gold, and for their
lamps of
gold, by
weight for every
candlestick, and for the
lamps thereof: and for the
candlesticks of
silver by
weight,
{\ADD{both}} for the
candlestick, and
{\ADD{also}} for the
lamps thereof, according to the
use of every
candlestick.
\VS{16}And by
weight
{\ADD{he gave}}
gold for the
tables of
shewbread, for every
table; and
{\ADD{likewise}}
silver for the
tables of
silver:
\VS{17}Also
pure
gold for the
fleshhooks, and the
bowls, and the
cups: and for the
golden
basons
{\ADD{he gave gold}} by
weight for every
bason; and
{\ADD{likewise silver}} by
weight for every
bason of
silver:
\VS{18}And for the
altar of
incense
refined
gold by
weight; and
gold for the
pattern of the
chariot of the
cherubims, that spread
out
{\ADD{their wings}}, and
covered the
ark of the
covenant of the
{\ND{
LORD}}.
\VS{19}All
{\ADD{this, said David}}, the
{\ND{
LORD}} made me
understand in
writing by
{\ADD{his}}
hand upon me,
{\ADD{even}} all the
works of this
pattern.
\VS{20}And
David
said to
Solomon his
son, Be
strong and of good
courage, and
do
{\ADD{it:}}
fear not, nor be
dismayed: for the
{\ND{
LORD}}
God,
{\ADD{even}} my
God,
{\ADD{will be}} with thee; he will not
fail thee, nor
forsake thee, until thou hast
finished all the
work for the
service of the
house of the
{\ND{
LORD}}.
\VS{21}And, behold, the
courses of the
priests and the
Levites,
{\ADD{even they shall be with thee}} for all the
service of the
house of
God: and
{\ADD{there shall be}} with thee for all manner of
workmanship every
willing skilful
man, for any manner of
service: also the
princes and all the
people
{\ADD{will be}} wholly at thy
commandment.

\par }\Chap{29}{\PP \VerseOne{1}Furthermore
David the
king
said unto all the
congregation,
Solomon my
son, whom
alone
God hath
chosen,
{\ADD{is yet}}
young and
tender, and the
work
{\ADD{is}}
great: for the
palace
{\ADD{is}} not for
man, but for the
{\ND{
LORD}}
God.
\VS{2}Now I have
prepared with all my
might for the
house of my
God the
gold for
{\ADD{things to be made}} of
gold, and the
silver for
{\ADD{things}} of
silver, and the
brass for
{\ADD{things}} of
brass, the
iron for
{\ADD{things}} of
iron, and
wood for
{\ADD{things}} of
wood;
onyx
stones, and
{\ADD{stones}} to be
set,
glistering
stones, and of divers
colours, and all manner of
precious
stones, and
marble
stones in
abundance.
\VS{3}Moreover, because I have set my
affection to the
house of my
God, I
have of mine own proper
good, of
gold and
silver,
{\ADD{which}} I have
given to the
house of my
God, over and
above all that I have
prepared for the
holy
house,
\VS{4}{\ADD{Even}}
three
thousand
talents of
gold, of the
gold of
Ophir, and
seven
thousand
talents of
refined
silver, to
overlay the
walls of the
houses
{\ADD{withal:}}
\VS{5}The
gold for
{\ADD{things}} of
gold, and the
silver for
{\ADD{things}} of
silver, and for all manner of
work
{\ADD{to be made}} by the
hands of
artificers. And who
{\ADD{then}} is
willing to
consecrate his
service this
day unto the
{\ND{
LORD}}?\FTNT{*}{{\FR 29.5 }
consecrate his service: Heb. fill his hand}
\par }{\PP \VS{6}¶ Then the
chief of the
fathers and
princes of the
tribes of
Israel, and the
captains of
thousands and of
hundreds, with the
rulers of the
king’s
work, offered
willingly,
\VS{7}And
gave for the
service of the
house of
God of
gold
five
thousand
talents and ten
thousand
drams, and of
silver
ten
thousand
talents, and of
brass
eighteen
thousand
talents, and one
hundred
thousand
talents of
iron.
\VS{8}And they with whom
{\ADD{precious}}
stones were
found
gave
{\ADD{them}} to the
treasure of the
house of the
{\ND{
LORD}}, by the
hand of
Jehiel the
Gershonite.
\VS{9}Then the
people
rejoiced, for that they offered
willingly, because with
perfect
heart they offered
willingly to the
{\ND{
LORD}}: and
David the
king also
rejoiced with
great
joy.
\par }{\PP \VS{10}¶ Wherefore
David
blessed the
{\ND{
LORD}}
before all the
congregation: and
David
said,
Blessed
{\ADD{be}} thou,
{\ND{
LORD}}
God of
Israel our
father,
for
ever and
ever.
\VS{11}Thine, O
{\ND{
LORD}},
{\ADD{is}} the
greatness, and the
power, and the
glory, and the
victory, and the
majesty: for all
{\ADD{that is}} in the
heaven and in the
earth
{\ADD{is thine;}} thine
{\ADD{is}} the
kingdom, O
{\ND{
LORD}}, and thou art
exalted as
head above all.
\VS{12}Both
riches and
honour
{\ADD{come}} of
thee, and thou
reignest over all; and in thine
hand
{\ADD{is}}
power and
might; and in thine
hand
{\ADD{it is}} to make
great, and to give
strength unto all.
\VS{13}Now therefore, our
God, we
thank thee, and
praise thy
glorious
name.
\VS{14}But who
{\ADD{am}} I, and what
{\ADD{is}} my
people, that we should
be
able to offer so
willingly after this sort? for all things
{\ADD{come}} of thee, and of thine
own have we
given thee.\FTNT{†}{{\FR 29.14 }
be able: Heb. retain, or, obtain strength}\FTNT{‡}{{\FR 29.14 }
of thine…: Heb. of thine hand}
\VS{15}For we
{\ADD{are}}
strangers
before thee, and
sojourners, as
{\ADD{were}} all our
fathers: our
days on the
earth
{\ADD{are}} as a
shadow, and
{\ADD{there is}} none
abiding.\FTNT{§}{{\FR 29.15 }
abiding: Heb. expectation}
\VS{16}O
{\ND{
LORD}} our
God, all this
store that we have
prepared to
build thee an
house for thine
holy
name
{\ADD{cometh}} of thine
hand, and
{\ADD{is}} all thine own.
\VS{17}I
know also, my
God, that thou
triest the
heart, and hast
pleasure in
uprightness. As for me, in the
uprightness of mine
heart I have willingly
offered all these things: and now have I
seen with
joy thy
people, which are
present here, to offer
willingly unto thee.\FTNT{**}{{\FR 29.17 }
present: Heb. found}
\VS{18}O
{\ND{
LORD}}
God of
Abraham,
Isaac, and of
Israel, our
fathers,
keep this for
ever in the
imagination of the
thoughts of the
heart of thy
people, and
prepare their
heart unto thee:\FTNT{††}{{\FR 29.18 }
prepare: or, stablish}
\VS{19}And
give unto
Solomon my
son a
perfect
heart, to
keep thy
commandments, thy
testimonies, and thy
statutes, and to
do all
{\ADD{these things}}, and to
build the
palace,
{\ADD{for}} the which I have made
provision.
\par }{\PP \VS{20}¶ And
David
said to all the
congregation, Now
bless the
{\ND{
LORD}} your
God. And all the
congregation
blessed the
{\ND{
LORD}}
God of their
fathers, and bowed down their
heads, and
worshipped the
{\ND{
LORD}}, and the
king.
\VS{21}And they
sacrificed
sacrifices unto the
{\ND{
LORD}}, and
offered burnt
offerings unto the
{\ND{
LORD}}, on the
morrow after that
day,
{\ADD{even}} a
thousand
bullocks, a
thousand
rams,
{\ADD{and}} a
thousand
lambs, with their drink
offerings, and
sacrifices in
abundance for all
Israel:
\VS{22}And did
eat and
drink
before the
{\ND{
LORD}} on that
day with
great
gladness. And they made
Solomon the
son of
David
king the second
time, and
anointed
{\ADD{him}} unto the
{\ND{
LORD}}
{\ADD{to be}} the chief
governor, and
Zadok
{\ADD{to be}}
priest.
\par }{\PP \VS{23}Then
Solomon sat
on the
throne of the
{\ND{
LORD}} as
king instead of
David his
father, and
prospered; and all
Israel
obeyed him.
\VS{24}And all the
princes, and the mighty
men, and all the
sons likewise of
king
David, submitted
themselves unto
Solomon the
king.\FTNT{‡‡}{{\FR 29.24 }
submitted…: Heb. gave the hand under Solomon}
\VS{25}And the
{\ND{
LORD}}
magnified
Solomon
exceedingly in the
sight of all
Israel, and
bestowed upon him
{\ADD{such}}
royal
majesty as had not been on any
king
before him in
Israel.
\par }{\PP \VS{26}¶ Thus
David the
son of
Jesse
reigned over all
Israel.
\VS{27}And the
time that he
reigned over
Israel
{\ADD{was}}
forty
years;
seven
years
reigned he in
Hebron, and
thirty and
three
{\ADD{years}}
reigned he in
Jerusalem.
\VS{28}And he
died in a
good old
age,
full of
days,
riches, and
honour: and
Solomon his
son
reigned in his stead.
\VS{29}Now the
acts of
David the
king,
first and
last, behold, they
{\ADD{are}}
written in the
book of
Samuel the
seer, and in the
book of
Nathan the
prophet, and in the
book of
Gad the
seer,\FTNT{§§}{{\FR 29.29 }
book: or, history: Heb. words}
\VS{30}With all his
reign and his
might, and the
times that went
over him, and over
Israel, and over all the
kingdoms of the
countries.
\par }