\NormalFont\ShortTitle{1 Peter}
{\MT THE FIRST EPISTLE GENERAL OF PETER

\par }\ChapOne{1}{\PP \VerseOne{1}Peter, an
apostle of
Jesus
Christ, to the
strangers scattered
throughout
Pontus,
Galatia,
Cappadocia,
Asia,
and
Bithynia,
\VS{2}Elect according
to the
foreknowledge of
God the
Father,
through
sanctification of the
Spirit,
unto
obedience
and
sprinkling of the
blood of
Jesus
Christ:
Grace unto
you,
and
peace, be
multiplied.
\par }{\PP \VS{3}Blessed
{\ADD{be}} the
God
and
Father
of
our
Lord
Jesus
Christ,
which according
to
his
abundant
mercy hath
begotten
us
again
unto a
lively
hope
by the
resurrection of
Jesus
Christ
from the
dead,
\VS{4}To an
inheritance
incorruptible,
and
undefiled,
and that fadeth not
away,
reserved
in
heaven
for
you,
\VS{5}Who are
kept
by the
power of
God
through
faith
unto
salvation
ready to be
revealed
in the
last
time.
\VS{6}Wherein ye greatly
rejoice, though
now for a
season,
if need
be, ye
are in
heaviness
through
manifold
temptations:
\VS{7}That the
trial of
your
faith, being
much more precious
than of
gold
that
perisheth,
though it be
tried
with
fire, might be
found
unto
praise
and
honour
and
glory
at the
appearing of
Jesus
Christ:
\VS{8}Whom
having
not
seen, ye
love;
in
whom, though
now ye
see
{\ADD{him}}
not,
yet
believing, ye
rejoice with
joy
unspeakable
and full of
glory:
\VS{9}Receiving the
end of
your
faith,
{\ADD{even}} the
salvation
of
{\ADD{your}}
souls.
\VS{10}Of
which
salvation the
prophets have
enquired
and searched
diligently, who
prophesied
of the
grace
{\ADD{that should come}}
unto
you:
\VS{11}Searching
what,
or what manner
of
time the
Spirit of
Christ
which was
in
them did
signify, when it
testified beforehand the
sufferings
of
Christ,
and the
glory
that should
follow.
\VS{12}Unto
whom it was
revealed,
that
not unto
themselves,
but unto
us they did
minister the
things,
which
are
now
reported unto
you
by
them that have preached the
gospel unto
you
with the
Holy
Ghost sent
down
from
heaven; which
things the
angels
desire to
look
into.
\VS{13}Wherefore gird
up the
loins of
your
mind, be
sober, and
hope to the
end
for the
grace that is to be
brought unto
you
at the
revelation of
Jesus
Christ;
\VS{14}As
obedient
children,
not fashioning yourselves according
to the
former
lusts
in
your
ignorance:
\VS{15}But
as he which hath
called
you is
holy,
so
be
ye
holy
in
all manner of
conversation;
\VS{16}Because it is
written, Be
ye
holy;
for
I
am
holy.
\VS{17}And
if ye call
on the
Father,
who without respect of
persons
judgeth according
to every
man’s
work,
pass the
time of
your
sojourning
{\ADD{here}}
in
fear:
\VS{18}Forasmuch as ye
know
that ye
were
not
redeemed with corruptible
things,
{\ADD{as}}
silver
and
gold,
from
your
vain
conversation
{\ADD{received}} by tradition from your
fathers;
\VS{19}But with the
precious
blood of
Christ,
as of a
lamb without
blemish
and without
spot:
\VS{20}Who
verily was
foreordained
before the
foundation of the
world,
but was
manifest
in these
last
times
for
you,
\VS{21}Who
by
him do
believe
in
God,
that
raised
him
up
from the
dead,
and
gave
him
glory;
that
your
faith
and
hope might
be
in
God.
\VS{22}Seeing ye have
purified
your
souls
in
obeying the
truth
through the
Spirit
unto
unfeigned love of the
brethren,
{\ADD{see that ye}}
love one
another
with a
pure
heart
fervently:
\VS{23}Being born
again,
not
of
corruptible
seed,
but of
incorruptible,
by the
word of
God, which
liveth
and
abideth
for
ever.
\VS{24}For
all
flesh
{\ADD{is}}
as
grass,
and
all the
glory of
man
as the
flower of
grass. The
grass
withereth,
and the
flower
thereof falleth
away:
\VS{25}But the
word of the
Lord
endureth
for
ever.
And
this
is the
word
which by the gospel is
preached
unto
you.

\par }\Chap{2}{\PP \VerseOne{1}Wherefore laying
aside
all
malice,
and
all
guile,
and
hypocrisies,
and
envies,
and
all evil
speakings,
\VS{2}As
newborn
babes,
desire the
sincere
milk of the
word,
that ye may
grow
thereby:
\VS{3}If so
be ye have
tasted
that the
Lord
{\ADD{is}}
gracious.
\VS{4}To
whom
coming,
{\ADD{as unto}} a
living
stone,
disallowed
indeed
of
men,
but
chosen
of
God,
{\ADD{and}}
precious,
\VS{5}Ye
also,
as
lively
stones, are built
up a
spiritual
house, an
holy
priesthood, to offer
up
spiritual
sacrifices,
acceptable to
God
by
Jesus
Christ.
\VS{6}Wherefore
also it is
contained
in the
scripture,
Behold, I
lay
in
Sion a chief
corner
stone,
elect,
precious:
and he that
believeth
on
him
shall
not be
confounded.
\VS{7}Unto
you
therefore
which
believe
{\ADD{he is}}
precious:
but unto them which be
disobedient, the
stone
which the
builders
disallowed, the
same is
made the
head of the
corner,
\VS{8}And a
stone of
stumbling,
and a
rock of
offence,
{\ADD{even to them}}
which stumble
at the
word, being
disobedient:
whereunto
also they were
appointed.
\VS{9}But
ye
{\ADD{are}} a
chosen
generation, a
royal
priesthood, an
holy
nation, a
peculiar
people;
that ye should shew
forth the
praises of him who hath
called
you out
of
darkness
into
his
marvellous
light:
\VS{10}Which in time
past
{\ADD{were}}
not a
people,
but
{\ADD{are}}
now the
people of
God:
which
had
not obtained
mercy,
but
now have obtained
mercy.
\par }{\PP \VS{11}Dearly
beloved, I
beseech
{\ADD{you}}
as
strangers
and
pilgrims,
abstain from
fleshly
lusts,
which
war
against the
soul;
\VS{12}Having
your
conversation
honest
among the
Gentiles:
that,
whereas they speak
against
you
as
evildoers, they
may
by
{\ADD{your}}
good
works, which they shall
behold,
glorify
God
in the
day of
visitation.
\par }{\PP \VS{13}Submit
yourselves to
every
ordinance of
man
for the
Lord’s sake: whether it
be to the
king,
as
supreme;
\VS{14}Or unto
governors,
as unto them that are
sent
by
him
for the
punishment of
evildoers,
and for the
praise of them that do
well.
\VS{15}For
so
is the
will of
God, that with well
doing ye may put to
silence the
ignorance of
foolish
men:
\VS{16}As
free,
and
not
using
{\ADD{your}}
liberty
for a
cloke of
maliciousness,
but
as the
servants of
God.
\VS{17}Honour
all
{\ADD{men}}.
Love the
brotherhood.
Fear
God.
Honour the
king.
\VS{18}Servants,
{\ADD{be}}
subject
to
{\ADD{your}}
masters
with
all
fear;
not
only to the
good
and
gentle,
but
also to the
froward.
\VS{19}For
this
{\ADD{is}}
thankworthy,
if a
man
for
conscience toward
God
endure
grief,
suffering
wrongfully.
\VS{20}For
what
glory
{\ADD{is it}},
if,
when ye be
buffeted for your
faults, ye shall take it
patiently?
but
if, when ye do
well,
and
suffer
{\ADD{for it}}, ye take it
patiently,
this
{\ADD{is}}
acceptable
with
God.
\VS{21}For
even
hereunto were ye
called:
because
Christ
also
suffered
for
us,
leaving
us an
example,
that ye should
follow
his
steps:
\VS{22}Who
did
no
sin,
neither
was
guile
found
in
his
mouth:
\VS{23}Who, when he was
reviled,
reviled
not
again; when he
suffered, he
threatened
not;
but
committed
{\ADD{himself}} to him that
judgeth
righteously:
\VS{24}Who his own
self
bare
our
sins
in his
own
body
on the
tree,
that
we, being
dead to
sins, should
live unto
righteousness: by
whose
stripes ye were
healed.
\VS{25}For ye
were
as
sheep going
astray;
but
are
now
returned
unto the
Shepherd
and
Bishop of
your
souls.

\par }\Chap{3}{\PP \VerseOne{1}Likewise, ye
wives,
{\ADD{be}} in
subjection to your
own
husbands;
that,
if
any obey
not the
word, they
also
may
without the
word be
won
by the conversation of the
wives;
\VS{2}While they
behold
your
chaste
conversation
{\ADD{coupled}}
with
fear.
\VS{3}Whose
adorning let
it
not
be that
outward
{\ADD{adorning}} of
plaiting the
hair,
and of
wearing of
gold,
or of putting
on of
apparel;
\VS{4}But
{\ADD{let it be}} the
hidden
man of the
heart,
in that which is not
corruptible,
{\ADD{even the ornament}} of a
meek
and
quiet
spirit,
which
is in the
sight of
God of great
price.
\VS{5}For after this
manner in the old
time the
holy
women
also, who
trusted
in
God,
adorned
themselves, being in
subjection unto their
own
husbands:
\VS{6}Even
as
Sara
obeyed
Abraham,
calling
him
lord:
whose
daughters ye
are, as long as ye do
well,
and
are
not
afraid with
any
amazement.
\VS{7}Likewise, ye
husbands, dwell
with
{\ADD{them}} according
to
knowledge,
giving
honour unto the
wife,
as unto the
weaker
vessel,
and
as being heirs
together of the
grace of
life;
that
your
prayers
be
not
hindered.
\VS{8}Finally,
{\ADD{be ye}}
all of one
mind, having compassion one of
another, love as
brethren,
{\ADD{be}}
pitiful,
{\ADD{be}}
courteous:
\VS{9}Not
rendering
evil
for
evil,
or
railing
for
railing:
but
contrariwise
blessing;
knowing
that ye
are
thereunto
called,
that ye should
inherit a
blessing.
\VS{10}For he that
will
love
life,
and
see
good
days, let him
refrain
his
tongue
from
evil,
and
his
lips that they
speak
no
guile:
\VS{11}Let him
eschew
evil,
and
do
good; let him
seek
peace,
and
ensue
it.
\VS{12}For the
eyes of the
Lord
{\ADD{are}}
over the
righteous,
and
his
ears
{\ADD{are open}}
unto
their
prayers:
but the
face of the
Lord
{\ADD{is}}
against them that
do
evil.
\par }{\PP \VS{13}And
who
{\ADD{is}} he that will
harm
you,
if ye
be
followers of that which is
good?
\VS{14}But
and
if ye
suffer
for
righteousness’ sake,
happy
{\ADD{are ye}}:
and
be
not
afraid of
their
terror,
neither be
troubled;
\VS{15}But
sanctify the
Lord
God
in
your
hearts:
and
{\ADD{be}}
ready
always
to
{\ADD{give}} an
answer to every
man that
asketh
you a
reason
of the
hope
that is
in
you
with
meekness
and
fear:
\VS{16}Having a
good
conscience;
that,
whereas they speak
evil of
you,
as of
evildoers, they may be
ashamed that falsely
accuse
your
good
conversation
in
Christ.
\VS{17}For
{\ADD{it is}}
better,
if the
will of
God be
so, that ye
suffer for well
doing,
than for evil
doing.
\VS{18}For
Christ
also
hath
once
suffered
for
sins, the
just
for the
unjust,
that he might
bring
us to
God, being put to
death in the
flesh,
but
quickened by the
Spirit:
\VS{19}By
which
also he
went and
preached unto the
spirits
in
prison;
\VS{20}Which
sometime were
disobedient,
when
once the
longsuffering of
God
waited
in the
days of
Noah, while the
ark was a
preparing,
wherein
few,
that
is,
eight
souls were
saved
by
water.
\VS{21}The like
figure
whereunto
{\ADD{even}}
baptism
doth
also
now
save
us (not the putting
away of the
filth of the
flesh,
but the
answer of a
good
conscience
toward
God,)
by the
resurrection of
Jesus
Christ:
\VS{22}Who is
gone
into
heaven, and
is
on the right
hand of
God;
angels
and
authorities
and
powers being made
subject unto
him.

\par }\Chap{4}{\PP \VerseOne{1}Forasmuch then
as
Christ hath
suffered
for
us in the
flesh,
arm
yourselves
likewise with the
same
mind:
for he that hath
suffered
in the
flesh hath
ceased from
sin;
\VS{2}That
he no
longer should
live the
rest
of
{\ADD{his}}
time
in the
flesh to the
lusts of
men,
but to the
will of
God.
\VS{3}For the
time
past
of
{\ADD{our}}
life may
suffice
us to have
wrought the
will of the
Gentiles, when we
walked
in
lasciviousness,
lusts, excess of
wine,
revellings,
banquetings,
and
abominable
idolatries:
\VS{4}Wherein they think it
strange that
ye
run
not
with
{\ADD{them}}
to the
same
excess of
riot, speaking evil
of
{\ADD{you}}:
\VS{5}Who shall
give
account to him that
is
ready to
judge the
quick
and the
dead.
\VS{6}For
for this
cause was the gospel
preached
also to them that are
dead,
that they might be
judged according
to
men in the
flesh,
but
live according
to
God in the
spirit.
\par }{\PP \VS{7}But the
end of all
things is at
hand: be
ye
therefore
sober,
and
watch
unto
prayer.
\VS{8}And
above all
things
have
fervent
charity
among
yourselves:
for
charity shall
cover the
multitude of
sins.
\VS{9}Use
hospitality
one
to
another
without
grudging.
\VS{10}As every
man hath
received the
gift,
{\ADD{even so}}
minister the
same
one
to
another,
as
good
stewards of the
manifold
grace of
God.
\VS{11}If any
man
speak,
{\ADD{let him speak}}
as the
oracles of
God;
if any
man
minister,
{\ADD{let him do it}}
as
of the
ability
which
God
giveth:
that
God
in all
things may be
glorified
through
Jesus
Christ, to
whom
be
praise
and
dominion
for
ever and
ever.
Amen.
\par }{\PP \VS{12}Beloved, think
it
not
strange
concerning the fiery
trial which
is
to
try
you, as
though some strange
thing
happened unto
you:
\VS{13}But
rejoice, inasmuch
as ye are
partakers of
Christ’s
sufferings;
that,
when
his
glory shall be
revealed, ye may be
glad
also with exceeding
joy.
\VS{14}If ye be
reproached
for the
name of
Christ,
happy
{\ADD{are ye}};
for the
spirit of
glory
and of
God
resteth
upon
you:
on
their
part he is evil spoken
of,
but
on
your
part he is
glorified.
\VS{15}But
let
none of
you
suffer
as a
murderer,
or
{\ADD{as}} a
thief,
or
{\ADD{as}} an
evildoer,
or
as a busybody in other
men’s matters.
\VS{16}Yet
if
{\ADD{any man suffer}}
as a
Christian, let
him
not be
ashamed;
but let him
glorify
God
on
this
behalf.
\VS{17}For the
time
{\ADD{is come}} that
judgment must
begin
at the
house of
God:
and
if
{\ADD{it}}
first
{\ADD{begin}}
at
us,
what shall the
end
{\ADD{be}} of them that obey
not the
gospel of
God?
\VS{18}And
if the
righteous
scarcely be
saved,
where
shall the
ungodly
and the
sinner
appear?
\VS{19}Wherefore let them that
suffer according
to the
will of
God commit the
keeping of
their
souls
{\ADD{to him}}
in well
doing,
as unto a
faithful
Creator.

\par }\Chap{5}{\PP \VerseOne{1}The
elders
which are
among
you I
exhort,
who am also an
elder,
and a
witness of the
sufferings of
Christ, and
also a
partaker of the
glory that
shall be
revealed:
\VS{2}Feed the
flock of
God which is
among
you, taking the
oversight
{\ADD{thereof}},
not by
constraint,
but
willingly;
not for filthy
lucre,
but of a ready
mind;
\VS{3}Neither
as being lords
over
{\ADD{God’s}}
heritage,
but
being
ensamples to the
flock.
\VS{4}And
when the chief
Shepherd shall
appear, ye shall
receive a
crown of
glory that fadeth not
away.
\VS{5}Likewise, ye
younger,
submit
yourselves unto the
elder.
Yea,
all
{\ADD{of you}} be subject one to
another, and be clothed
with
humility:
for
God
resisteth the
proud,
and
giveth
grace to the
humble.
\par }{\PP \VS{6}Humble
yourselves
therefore
under the
mighty
hand of
God,
that he may
exalt
you
in due
time:
\VS{7}Casting
all
your
care
upon
him;
for
he
careth
for
you.
\VS{8}Be
sober, be
vigilant;
because
your
adversary the
devil,
as a
roaring
lion, walketh
about,
seeking
whom he may
devour:
\VS{9}Whom
resist
stedfast in the
faith, knowing
that the
same
afflictions are
accomplished in
your
brethren that are
in the
world.
\VS{10}But the
God of
all
grace,
who hath
called
us
unto
his
eternal
glory
by
Christ
Jesus, after that ye have
suffered a
while,
make
you
perfect,
stablish,
strengthen,
settle
{\ADD{you}}.
\VS{11}To
him
{\ADD{be}}
glory
and
dominion
for
ever and
ever.
Amen.
\par }{\PP \VS{12}By
Silvanus, a
faithful
brother unto
you,
as I
suppose, I have
written
briefly,
exhorting,
and
testifying that
this
is the
true
grace of
God
wherein ye
stand.
\VS{13}The
{\ADD{church that is}}
at
Babylon, elected together
with
{\ADD{you}},
saluteth
you;
and
{\ADD{so doth}}
Marcus
my
son.
\VS{14}Greet
ye one
another
with a
kiss of
charity.
Peace
{\ADD{be}} with
you
all
that are
in
Christ
Jesus.
Amen.
\par }