\NormalFont\ShortTitle{1 Samuel}
{\MT The First Book of Samuel Otherwise Called the First Book of the Kings

\par }\ChapOne{1}{\PP \VerseOne{1}Now there was a
certain
man of
Ramathaim-zophim, of
mount
Ephraim, and his
name
{\ADD{was}}
Elkanah, the
son of
Jeroham, the
son of
Elihu, the
son of
Tohu, the
son of
Zuph, an
Ephrathite:
\VS{2}And he had
two
wives; the
name of the
one
{\ADD{was}}
Hannah, and the
name of the
other
Peninnah: and
Peninnah had
children, but
Hannah had no
children.
\VS{3}And
this
man went
up out of his
city
yearly to
worship and to
sacrifice unto the
{\ND{
LORD}} of
hosts in
Shiloh. And the
two
sons of
Eli,
Hophni and
Phinehas, the
priests of the
{\ND{
LORD}},
{\ADD{were}} there.\FTNT{*}{{\FR 1.3 }
yearly: Heb. from year to year}
\par }{\PP \VS{4}¶ And when the
time was that
Elkanah
offered, he
gave to
Peninnah his
wife, and to all her
sons and her
daughters,
portions:
\VS{5}But unto
Hannah he
gave
a
worthy
portion; for he
loved
Hannah: but the
{\ND{
LORD}} had shut
up her
womb.\FTNT{†}{{\FR 1.5 }
worthy: or, double}
\VS{6}And her
adversary also
provoked her
sore, for to make her
fret, because the
{\ND{
LORD}} had
shut
up her
womb.\FTNT{‡}{{\FR 1.6 }
provoked: Heb. angered}
\VS{7}And
{\ADD{as}} he
did so
year by
year,
when she went
up to the
house of the
{\ND{
LORD}}, so she
provoked her; therefore she
wept, and did not
eat.\FTNT{§}{{\FR 1.7 }
when…: or, from the time that she, etc: Heb. from her going up}
\VS{8}Then
said
Elkanah her
husband to her,
Hannah, why
weepest thou? and why
eatest thou not? and why is thy
heart
grieved?
{\ADD{am}} not I
better to thee than
ten
sons?
\par }{\PP \VS{9}¶ So
Hannah rose
up
after they had
eaten in
Shiloh, and
after they had
drunk. Now
Eli the
priest
sat upon a
seat by a
post of the
temple of the
{\ND{
LORD}}.
\VS{10}And she
{\ADD{was}} in
bitterness of
soul, and
prayed unto the
{\ND{
LORD}}, and
wept
sore.\FTNT{**}{{\FR 1.10 }
in…: Heb. bitter of soul}
\VS{11}And she
vowed a
vow, and
said, O
{\ND{
LORD}} of
hosts, if thou wilt
indeed
look on the
affliction of thine
handmaid, and
remember me, and not
forget thine
handmaid, but wilt
give unto thine
handmaid a
man
child, then I will
give him unto the
{\ND{
LORD}} all the
days of his
life, and there shall no
razor
come upon his
head.\FTNT{††}{{\FR 1.11 }
a man…: Heb. seed of men}
\VS{12}And it came to pass, as she
continued
praying
before the
{\ND{
LORD}}, that
Eli
marked her
mouth.\FTNT{‡‡}{{\FR 1.12 }
continued…: Heb. multiplied to pray}
\VS{13}Now
Hannah, she
spake in her
heart; only her
lips
moved, but her
voice was not
heard: therefore
Eli
thought she had been
drunken.
\VS{14}And
Eli
said unto her, How long wilt thou be
drunken? put
away thy
wine from thee.
\VS{15}And
Hannah
answered and
said, No, my
lord, I
{\ADD{am}} a
woman of a
sorrowful
spirit: I have
drunk neither
wine nor strong
drink, but have poured
out my
soul
before the
{\ND{
LORD}}.\FTNT{§§}{{\FR 1.15 }
of a sorrowful…: Heb. hard of spirit}
\VS{16}Count not thine
handmaid
for a
daughter of
Belial: for out of the
abundance of my
complaint and
grief have I
spoken hitherto.\FTNT{***}{{\FR 1.16 }
complaint: or, meditation}
\VS{17}Then
Eli
answered and
said,
Go in
peace: and the
God of
Israel
grant
{\ADD{thee}} thy
petition that thou hast
asked of him.
\VS{18}And she
said, Let thine
handmaid
find
grace in thy
sight. So the
woman
went her
way, and did
eat, and her
countenance was no more
{\ADD{sad}}.
\par }{\PP \VS{19}¶ And they rose
up in the
morning
early, and
worshipped
before the
{\ND{
LORD}}, and
returned, and
came to their
house to
Ramah: and
Elkanah
knew
Hannah his
wife; and the
{\ND{
LORD}}
remembered her.
\VS{20}Wherefore it came to pass, when the
time was
come about after
Hannah had
conceived, that she
bare a
son, and
called his
name
Samuel,
{\ADD{saying}}, Because I have
asked him of the
{\ND{
LORD}}.\FTNT{†††}{{\FR 1.20 }
when…: Heb. in revolution of days}\FTNT{‡‡‡}{{\FR 1.20 }
Samuel: that is, Asked of God}
\VS{21}And the
man
Elkanah, and all his
house, went
up to
offer unto the
{\ND{
LORD}} the
yearly
sacrifice, and his
vow.
\VS{22}But
Hannah went not
up; for she
said unto her
husband,
{\ADD{I will not go up}} until the
child be
weaned, and
{\ADD{then}} I will
bring him, that he may
appear
before the
{\ND{
LORD}}, and there
abide
for
ever.
\VS{23}And
Elkanah her
husband
said unto her,
Do what
seemeth thee
good;
tarry until thou have
weaned him; only the
{\ND{
LORD}}
establish his
word. So the
woman
abode, and gave her
son
suck until she
weaned him.
\par }{\PP \VS{24}¶ And when she had
weaned him, she took him
up with her, with
three
bullocks, and
one
ephah of
flour, and a
bottle of
wine, and
brought him unto the
house of the
{\ND{
LORD}} in
Shiloh: and the
child
{\ADD{was}}
young.
\VS{25}And they
slew a
bullock, and
brought the
child to
Eli.
\VS{26}And she
said,
Oh my
lord,
{\ADD{as}} thy
soul
liveth, my
lord, I
{\ADD{am}} the
woman that
stood by thee here,
praying unto the
{\ND{
LORD}}.
\VS{27}For this
child I
prayed; and the
{\ND{
LORD}} hath
given me my
petition which I
asked of him:
\VS{28}Therefore also I have
lent him to the
{\ND{
LORD}}; as long as he
liveth he shall be
lent to the
{\ND{
LORD}}. And he
worshipped the
{\ND{
LORD}} there.\FTNT{§§§}{{\FR 1.28 }
lent him: or, returned him, whom I have obtained by petition}\FTNT{*}{{\FR 1.28 }
he shall…: or, he whom I have obtained by petition shall be returned}

\par }\Chap{2}{\PP \VerseOne{1}And
Hannah
prayed, and
said, My
heart
rejoiceth in the
{\ND{
LORD}}, mine
horn is
exalted in the
{\ND{
LORD}}: my
mouth is
enlarged over mine
enemies; because I
rejoice in thy
salvation.
\VS{2}{\ADD{There is}} none
holy as the
{\ND{
LORD}}: for
{\ADD{there is}} none
beside thee: neither
{\ADD{is there}} any
rock like our
God.
\VS{3}Talk no
more so
exceeding
proudly; let
{\ADD{not}}
arrogancy
come out of your
mouth: for the
{\ND{
LORD}}
{\ADD{is}} a
God of
knowledge, and by him
actions are
weighed.\FTNT{*}{{\FR 2.3 }
arrogancy: Heb. hard}
\VS{4}The
bows of the mighty
men
{\ADD{are}}
broken, and they that
stumbled are
girded with
strength.
\VS{5}{\ADD{They that were}}
full have hired
out themselves for
bread; and
{\ADD{they that were}}
hungry
ceased: so that the
barren hath
born
seven; and she that hath
many
children is waxed
feeble.
\VS{6}The
{\ND{
LORD}}
killeth, and maketh
alive: he bringeth
down to the
grave, and bringeth
up.
\VS{7}The
{\ND{
LORD}} maketh
poor, and maketh
rich: he bringeth
low,
and lifteth
up.
\VS{8}He raiseth
up the
poor out of the
dust,
{\ADD{and}} lifteth
up the
beggar from the
dunghill, to
set
{\ADD{them}} among
princes, and to make them
inherit the
throne of
glory: for the
pillars of the
earth
{\ADD{are}} the
{\ND{
LORD’s}}, and he hath
set the
world upon them.
\VS{9}He will
keep the
feet of his
saints, and the
wicked shall be
silent in
darkness; for by
strength shall no
man
prevail.
\VS{10}The
adversaries of the
{\ND{
LORD}} shall be broken to
pieces; out of
heaven shall he
thunder upon them: the
{\ND{
LORD}} shall
judge the
ends of the
earth; and he shall
give
strength unto his
king, and
exalt the
horn of his
anointed.
\par }{\PP \VS{11}And
Elkanah
went to
Ramah to his
house. And the
child
did
minister unto the
{\ND{
LORD}}
before
Eli the
priest.
\par }{\PP \VS{12}¶ Now the
sons of
Eli
{\ADD{were}}
sons of
Belial; they
knew not the
{\ND{
LORD}}.
\VS{13}And the
priests’
custom with the
people
{\ADD{was, that}}, when any
man
offered
sacrifice, the
priest’s
servant
came, while the
flesh was in
seething, with a
fleshhook of
three
teeth in his
hand;
\VS{14}And he
struck
{\ADD{it}} into the
pan, or
kettle, or
caldron, or
pot; all that the
fleshhook brought
up the
priest
took for himself. So they
did in
Shiloh unto all the
Israelites that
came thither.
\VS{15}Also before they
burnt the
fat, the
priest’s
servant
came, and
said to the
man that
sacrificed,
Give
flesh to
roast for the
priest; for he will not
have
sodden
flesh of thee, but
raw.
\VS{16}And
{\ADD{if}} any
man
said unto him, Let them not
fail to
burn the
fat
presently, and
{\ADD{then}}
take
{\ADD{as much}} as thy
soul
desireth; then he would
answer him,
{\ADD{Nay;}} but thou shalt
give
{\ADD{it me}} now: and if not, I will
take
{\ADD{it}} by
force.\FTNT{†}{{\FR 2.16 }
presently: Heb. as on the day}
\VS{17}Wherefore the
sin of the young
men was
very
great
before the
{\ND{
LORD}}: for
men
abhorred the
offering of the
{\ND{
LORD}}.
\par }{\PP \VS{18}¶ But
Samuel
ministered
before the
{\ND{
LORD}},
{\ADD{being}} a
child,
girded with a
linen
ephod.
\VS{19}Moreover his
mother
made him a
little
coat, and
brought
{\ADD{it}} to him from
year to
year, when she came
up with her
husband to
offer the
yearly
sacrifice.
\par }{\PP \VS{20}¶ And
Eli
blessed
Elkanah and his
wife, and
said, The
{\ND{
LORD}}
give thee
seed of this
woman for the
loan which is
lent to the
{\ND{
LORD}}. And they
went unto their own
home.\FTNT{‡}{{\FR 2.20 }
loan…: or, petition which she asked, etc}
\VS{21}And the
{\ND{
LORD}}
visited
Hannah, so that she
conceived, and
bare
three
sons and
two
daughters. And the
child
Samuel
grew
before the
{\ND{
LORD}}.
\par }{\PP \VS{22}¶ Now
Eli was
very
old, and
heard all that his
sons
did unto all
Israel; and how they
lay with the
women that
assembled
{\ADD{at}} the
door of the
tabernacle of the
congregation.\FTNT{§}{{\FR 2.22 }
assembled: Heb. assembled by troops}
\VS{23}And he
said unto them, Why
do ye such
things? for I
hear of your
evil
dealings by all this
people.\FTNT{**}{{\FR 2.23 }
of your…: or, evil words of you}
\VS{24}Nay, my
sons; for
{\ADD{it is}} no
good
report that I
hear: ye make the
{\ND{
LORD’s}}
people to
transgress.\FTNT{††}{{\FR 2.24 }
transgress: or, cry out}
\VS{25}If one
man
sin against
another, the
judge shall
judge him: but if a
man
sin against the
{\ND{
LORD}}, who shall
intreat for him? Notwithstanding they
hearkened not unto the
voice of their
father, because the
{\ND{
LORD}}
would
slay them.
\VS{26}And the
child
Samuel
grew
on, and was in
favour both with the
{\ND{
LORD}}, and also with
men.
\par }{\PP \VS{27}¶ And there
came a
man of
God unto
Eli, and
said unto him, Thus
saith the
{\ND{
LORD}}, Did I
plainly
appear unto the
house of thy
father, when they were in
Egypt in
Pharaoh’s
house?
\VS{28}And did I
choose him out of all the
tribes of
Israel
{\ADD{to be}} my
priest, to
offer upon mine
altar, to
burn
incense, to
wear an
ephod
before me? and did I
give unto the
house of thy
father all the offerings made by
fire of the
children of
Israel?
\VS{29}Wherefore
kick ye at my
sacrifice and at mine
offering, which I have
commanded
{\ADD{in my}}
habitation; and
honourest thy
sons above me, to make yourselves
fat with the
chiefest of all the
offerings of
Israel my
people?
\VS{30}Wherefore the
{\ND{
LORD}}
God of
Israel
saith, I
said
indeed
{\ADD{that}} thy
house, and the
house of thy
father, should
walk
before me
for
ever: but now the
{\ND{
LORD}}
saith, Be it far from
me; for them that
honour me I will
honour, and they that
despise me shall be lightly
esteemed.
\VS{31}Behold, the
days
come, that I will cut
off thine
arm, and the
arm of thy
father’s
house, that there shall not be an old
man in thine
house.
\VS{32}And thou shalt
see an
enemy
{\ADD{in my}}
habitation, in all
{\ADD{the wealth}} which
{\ADD{God}} shall
give
Israel: and there shall not be an old
man in thine
house for
ever.\FTNT{‡‡}{{\FR 2.32 }
an enemy…: or, the affliction of the tabernacle, for all the wealth which God would have given Israel}
\VS{33}And the
man of thine,
{\ADD{whom}} I shall not cut
off from mine
altar,
{\ADD{shall be}} to
consume thine
eyes, and to
grieve thine
heart: and all the
increase of thine
house shall
die in the flower of their
age.\FTNT{§§}{{\FR 2.33 }
in the flower…: Heb. men}
\VS{34}And this
{\ADD{shall be}} a
sign unto thee, that shall
come upon thy
two
sons, on
Hophni and
Phinehas; in
one
day they shall
die
both of them.
\VS{35}And I will raise me
up a
faithful
priest,
{\ADD{that}} shall
do according to
{\ADD{that}} which
{\ADD{is}} in mine
heart and in my
mind: and I will
build him a
sure
house; and he shall
walk
before mine
anointed for
ever.
\VS{36}And it shall come to pass,
{\ADD{that}} every one that is
left in thine
house shall
come
{\ADD{and}}
crouch to him for a
piece of
silver and a
morsel of
bread, and shall
say,
Put me, I pray thee, into
one of the
priests’ offices, that I may
eat a
piece of
bread.\FTNT{***}{{\FR 2.36 }
Put: Heb. Join}\FTNT{†††}{{\FR 2.36 }
one of…: or, somewhat about the priesthood}

\par }\Chap{3}{\PP \VerseOne{1}And the
child
Samuel
ministered unto the
{\ND{
LORD}}
before
Eli. And the
word of the
{\ND{
LORD}} was
precious in those
days;
{\ADD{there was}} no
open
vision.
\VS{2}And it came to pass at that
time, when
Eli
{\ADD{was}} laid
down in his
place, and his
eyes
began to wax
dim,
{\ADD{that}} he
could not
see;
\VS{3}And ere the
lamp of
God went
out in the
temple of the
{\ND{
LORD}}, where the
ark of
God
{\ADD{was}}, and
Samuel was laid
down
{\ADD{to sleep;}}
\VS{4}That the
{\ND{
LORD}}
called
Samuel: and he
answered, Here
{\ADD{am}} I.
\VS{5}And he
ran unto
Eli, and
said, Here
{\ADD{am}}
I; for thou
calledst me. And he
said, I
called not; lie
down
again. And he
went and lay
down.
\VS{6}And the
{\ND{
LORD}}
called yet
again,
Samuel. And
Samuel
arose and
went to
Eli, and
said, Here
{\ADD{am}} I; for thou didst
call me. And he
answered, I
called not, my
son; lie
down
again.
\VS{7}Now
Samuel did not
yet
know the
{\ND{
LORD}}, neither was the
word of the
{\ND{
LORD}} yet
revealed unto him.\FTNT{*}{{\FR 3.7 }
Now…: or, Thus did Samuel before he knew the
{\ND{LORD}}, and before the word of the
{\ND{LORD}} was revealed unto him}
\VS{8}And the
{\ND{
LORD}}
called
Samuel
again the third
time. And he
arose and
went to
Eli, and
said, Here
{\ADD{am}} I; for thou didst
call me. And
Eli
perceived that the
{\ND{
LORD}} had
called the
child.
\VS{9}Therefore
Eli
said unto
Samuel,
Go, lie
down: and it shall be, if he
call thee, that thou shalt
say,
Speak,
{\ND{
LORD}}; for thy
servant
heareth. So
Samuel
went and lay
down in his
place.
\VS{10}And the
{\ND{
LORD}}
came, and
stood, and
called as at other
times,
Samuel,
Samuel. Then
Samuel
answered,
Speak; for thy
servant
heareth.
\par }{\PP \VS{11}¶ And the
{\ND{
LORD}}
said to
Samuel, Behold, I will
do a
thing in
Israel, at which
both the
ears of every one that
heareth it shall
tingle.
\VS{12}In that
day I will
perform against
Eli all
{\ADD{things}} which I have
spoken
concerning his
house: when I
begin, I will also make an
end.\FTNT{†}{{\FR 3.12 }
when…: Heb. beginning and ending}
\VS{13}For I have
told him that I will
judge his
house
for
ever for the
iniquity which he
knoweth; because his
sons made themselves
vile, and he
restrained them not.\FTNT{‡}{{\FR 3.13 }
For I…: or, And I will tell him}\FTNT{§}{{\FR 3.13 }
vile: or, accursed}\FTNT{**}{{\FR 3.13 }
restrained…: Heb. frowned not upon them}
\VS{14}And therefore I have
sworn unto the
house of
Eli, that the
iniquity of
Eli’s
house shall not be
purged with
sacrifice nor
offering
for
ever.
\par }{\PP \VS{15}¶ And
Samuel
lay until the
morning, and
opened the
doors of the
house of the
{\ND{
LORD}}. And
Samuel
feared to
shew
Eli the
vision.
\VS{16}Then
Eli
called
Samuel, and
said,
Samuel, my
son. And he
answered, Here
{\ADD{am}} I.
\VS{17}And he
said, What
{\ADD{is}} the
thing that
{\ADD{the
{\ND{Lord}}}} hath
said unto thee? I pray thee
hide
{\ADD{it}} not from me:
God do
so to thee, and more
also, if thou
hide
{\ADD{any}}
thing from me of all the
things that he
said unto thee.\FTNT{††}{{\FR 3.17 }
more also: Heb. so add}\FTNT{‡‡}{{\FR 3.17 }
thing: or, word}
\VS{18}And
Samuel
told him every
whit, and
hid nothing from him. And he
said, It
{\ADD{is}} the
{\ND{
LORD}}: let him
do what
seemeth him
good.\FTNT{§§}{{\FR 3.18 }
every…: Heb. all the things, or, words}
\par }{\PP \VS{19}¶ And
Samuel
grew, and the
{\ND{
LORD}} was with him, and did let none of his
words
fall to the
ground.
\VS{20}And all
Israel from
Dan even to
Beer-sheba
knew that
Samuel
{\ADD{was}}
established
{\ADD{to be}} a
prophet of the
{\ND{
LORD}}.\FTNT{***}{{\FR 3.20 }
established: or, faithful}
\VS{21}And the
{\ND{
LORD}}
appeared
again in
Shiloh: for the
{\ND{
LORD}}
revealed himself to
Samuel in
Shiloh by the
word of the
{\ND{
LORD}}.

\par }\Chap{4}{\PP \VerseOne{1}And the
word of
Samuel came to all
Israel. Now
Israel went
out
against the
Philistines to
battle, and
pitched beside
Eben-ezer: and the
Philistines
pitched in
Aphek.\FTNT{*}{{\FR 4.1 }
came: or, came to pass: Heb. was}
\VS{2}And the
Philistines put themselves in
array
against
Israel: and when they
joined
battle,
Israel was
smitten
before the
Philistines: and they
slew of the
army in the
field about
four
thousand
men.\FTNT{†}{{\FR 4.2 }
they joined…: Heb. the battle was spread}\FTNT{‡}{{\FR 4.2 }
army: Heb. array}
\par }{\PP \VS{3}¶ And when the
people were
come into the
camp, the
elders of
Israel
said, Wherefore hath the
{\ND{
LORD}}
smitten us to
day
before the
Philistines? Let us
fetch the
ark of the
covenant of the
{\ND{
LORD}} out of
Shiloh unto us, that, when it
cometh
among us, it may
save us out of the
hand of our
enemies.\FTNT{§}{{\FR 4.3 }
fetch: Heb. take unto us}
\VS{4}So the
people
sent to
Shiloh, that they might
bring from thence the
ark of the
covenant of the
{\ND{
LORD}} of
hosts, which
dwelleth
{\ADD{between}} the
cherubims: and the
two
sons of
Eli,
Hophni and
Phinehas,
{\ADD{were}} there with the
ark of the
covenant of
God.
\VS{5}And when the
ark of the
covenant of the
{\ND{
LORD}}
came into the
camp, all
Israel
shouted with a
great
shout, so that the
earth rang
again.
\VS{6}And when the
Philistines
heard the
noise of the
shout, they
said, What
{\ADD{meaneth}} the
noise of this
great
shout in the
camp of the
Hebrews? And they
understood that the
ark of the
{\ND{
LORD}} was
come into the
camp.
\VS{7}And the
Philistines were
afraid, for they
said,
God is
come into the
camp. And they
said,
Woe unto us! for there hath not been such a thing
heretofore.\FTNT{**}{{\FR 4.7 }
heretofore: Heb. yesterday, or, the third day}
\VS{8}Woe unto us! who shall
deliver us out of the
hand of these
mighty
Gods? these
{\ADD{are}} the
Gods that
smote the
Egyptians with all the
plagues in the
wilderness.
\VS{9}Be
strong, and
quit yourselves like
men, O ye
Philistines, that ye be not
servants unto the
Hebrews, as they have
been to you:
quit yourselves like
men, and
fight.\FTNT{††}{{\FR 4.9 }
quit…: Heb. be men}
\par }{\PP \VS{10}¶ And the
Philistines
fought, and
Israel was
smitten, and they
fled every
man into his
tent: and there was a
very
great
slaughter; for there
fell of
Israel
thirty
thousand
footmen.
\VS{11}And the
ark of
God was
taken; and the
two
sons of
Eli,
Hophni and
Phinehas, were
slain.\FTNT{‡‡}{{\FR 4.11 }
were slain: Heb. died}
\par }{\PP \VS{12}¶ And there
ran a
man of
Benjamin out of the
army, and
came to
Shiloh the same
day with his
clothes
rent, and with
earth upon his
head.
\VS{13}And when he
came, lo,
Eli
sat upon a
seat by the
wayside
watching: for his
heart
trembled for the
ark of
God. And when the
man
came into the
city, and
told
{\ADD{it}}, all the
city cried
out.
\VS{14}And when
Eli
heard the
noise of the
crying, he
said, What
{\ADD{meaneth}} the
noise of this
tumult? And the
man
came in
hastily, and
told
Eli.
\VS{15}Now
Eli was
ninety and
eight
years
old; and his
eyes were
dim, that he
could not
see.\FTNT{§§}{{\FR 4.15 }
were dim: Heb. stood}
\VS{16}And the
man
said unto
Eli, I
{\ADD{am}} he that
came out of the
army, and I
fled to
day out of the
army. And he
said, What is there
done, my
son?\FTNT{***}{{\FR 4.16 }
is…: Heb. is the thing}
\VS{17}And the
messenger
answered and
said,
Israel is
fled
before the
Philistines, and there hath been also a
great
slaughter among the
people, and thy
two
sons also,
Hophni and
Phinehas, are
dead, and the
ark of
God is
taken.
\VS{18}And it came to pass, when he made
mention of the
ark of
God, that he
fell from off the
seat
backward
by the
side of the
gate, and his
neck
brake, and he
died: for he was an
old
man, and
heavy. And he had
judged
Israel
forty
years.
\par }{\PP \VS{19}¶ And his daughter in
law,
Phinehas’
wife, was with
child,
{\ADD{near}} to be
delivered: and when she
heard the
tidings
that the
ark of
God was
taken, and that her father in
law and her
husband were
dead, she
bowed herself and
travailed; for her
pains
came upon her.\FTNT{†††}{{\FR 4.19 }
be delivered: or, cry out}\FTNT{‡‡‡}{{\FR 4.19 }
came: Heb. were turned}
\VS{20}And about the
time of her
death the women that
stood by her
said unto her,
Fear not; for thou hast
born a
son. But she
answered not, neither did she
regard
{\ADD{it}}.\FTNT{§§§}{{\FR 4.20 }
neither…: Heb. set not her heart}
\VS{21}And she
named the
child
I-chabod,
saying, The
glory is
departed from
Israel:
because the
ark of
God was
taken, and because of her father in
law and her
husband.\FTNT{*}{{\FR 4.21 }
I-chabod: that is, Where is the glory? or, There is no glory}
\VS{22}And she
said, The
glory is
departed from
Israel: for the
ark of
God is
taken.

\par }\Chap{5}{\PP \VerseOne{1}And the
Philistines
took the
ark of
God, and
brought it from
Eben-ezer unto
Ashdod.
\VS{2}When the
Philistines
took the
ark of
God, they
brought it into the
house of
Dagon, and
set it
by
Dagon.
\par }{\PP \VS{3}¶ And when they of
Ashdod arose
early on the
morrow, behold,
Dagon
{\ADD{was}}
fallen upon his
face to the
earth
before the
ark of the
{\ND{
LORD}}. And they
took
Dagon, and
set him in his
place
again.
\VS{4}And when they arose
early on the
morrow
morning, behold,
Dagon
{\ADD{was}}
fallen upon his
face to the
ground
before the
ark of the
{\ND{
LORD}}; and the
head of
Dagon and
both the
palms of his
hands
{\ADD{were}} cut
off upon the
threshold; only
{\ADD{the stump of}}
Dagon was
left to him.\FTNT{*}{{\FR 5.4 }
the stump…: or, the fishy part}
\VS{5}Therefore neither the
priests of
Dagon, nor any that
come into
Dagon’s
house,
tread on the
threshold of
Dagon in
Ashdod unto this
day.
\par }{\PP \VS{6}But the
hand of the
{\ND{
LORD}} was
heavy upon them of
Ashdod, and he
destroyed them, and
smote them with
emerods,
{\ADD{even}}
Ashdod and the
coasts thereof.
\VS{7}And when the
men of
Ashdod
saw that
{\ADD{it was}} so, they
said, The
ark of the
God of
Israel shall not
abide with us: for his
hand is
sore upon us, and upon
Dagon our
god.
\VS{8}They
sent therefore and
gathered all the
lords of the
Philistines unto them, and
said, What shall we
do with the
ark of the
God of
Israel? And they
answered, Let the
ark of the
God of
Israel be carried
about unto
Gath. And they
carried the
ark of the
God of
Israel
about
{\ADD{thither}}.
\VS{9}And it was
{\ADD{so}}, that,
after they had carried it
about, the
hand of the
{\ND{
LORD}} was against the
city with a
very
great
destruction: and he
smote the
men of the
city, both
small and
great, and they had
emerods in their secret
parts.
\par }{\PP \VS{10}¶ Therefore they
sent the
ark of
God to
Ekron. And it came to pass, as the
ark of
God
came to
Ekron, that the
Ekronites cried
out,
saying, They have brought
about the
ark of the
God of
Israel to us, to
slay us and our
people.\FTNT{†}{{\FR 5.10 }
us, to…: Heb. me to slay me and my}
\VS{11}So they
sent and gathered
together all the
lords of the
Philistines, and
said, Send
away the
ark of the
God of
Israel, and let it go
again to his own
place, that it
slay us not, and our
people: for there was a
deadly
destruction throughout all the
city; the
hand of
God was
very
heavy there.\FTNT{‡}{{\FR 5.11 }
us not…: Heb. me not, and my}
\VS{12}And the
men that
died not were
smitten with the
emerods: and the
cry of the
city went
up to
heaven.

\par }\Chap{6}{\PP \VerseOne{1}And the
ark of the
{\ND{
LORD}} was in the
country of the
Philistines
seven
months.
\VS{2}And the
Philistines
called for the
priests and the
diviners,
saying, What shall we
do to the
ark of the
{\ND{
LORD}}?
tell us wherewith we shall
send it to his
place.
\VS{3}And they
said, If ye send
away the
ark of the
God of
Israel,
send it not
empty; but in any
wise
return him a trespass
offering: then ye shall be
healed, and it shall be
known to you why his
hand is not
removed from you.
\VS{4}Then
said they, What
{\ADD{shall be}} the trespass
offering which we shall
return to him? They
answered,
Five
golden
emerods, and
five
golden
mice,
{\ADD{according to}} the
number of the
lords of the
Philistines: for
one
plague
{\ADD{was}} on you all, and on your
lords.\FTNT{*}{{\FR 6.4 }
you: Heb. them}
\VS{5}Wherefore ye shall
make
images of your
emerods, and
images of your
mice that
mar the
land; and ye shall
give
glory unto the
God of
Israel: peradventure he will
lighten his
hand from off you, and from off your
gods, and from off your
land.
\VS{6}Wherefore then do ye
harden your
hearts, as the
Egyptians and
Pharaoh
hardened their
hearts? when he had wrought
wonderfully among them, did they not let the people
go, and they
departed?\FTNT{†}{{\FR 6.6 }
wonderfully: or, reproachfully}\FTNT{‡}{{\FR 6.6 }
the people: Heb. them}
\VS{7}Now therefore
make
a
new
cart, and
take
two
milch
kine, on which there hath
come no
yoke, and
tie the
kine to the
cart, and
bring their
calves
home from
them:
\VS{8}And
take the
ark of the
{\ND{
LORD}}, and
lay it upon the
cart; and
put the
jewels of
gold, which ye
return him
{\ADD{for}} a trespass
offering, in a
coffer by the
side thereof; and send it
away, that it may
go.
\VS{9}And
see, if it goeth
up by the
way of his own
coast to
Beth-shemesh,
{\ADD{then}} he hath
done us this
great
evil: but if not, then we shall
know that
{\ADD{it is}} not his
hand
{\ADD{that}}
smote us; it
{\ADD{was}} a
chance
{\ADD{that}} happened to
us.\FTNT{§}{{\FR 6.9 }
he: or, it}
\par }{\PP \VS{10}¶ And the
men did
so; and
took
two
milch
kine, and
tied them to the
cart, and shut
up their
calves at
home:
\VS{11}And they
laid the
ark of the
{\ND{
LORD}} upon the
cart, and the
coffer with the
mice of
gold and the
images of their
emerods.
\VS{12}And the
kine took the
straight
way to the
way of
Beth-shemesh,
{\ADD{and}} went
along the
highway,
lowing as they
went, and turned not
aside
{\ADD{to}} the right
hand or
{\ADD{to}} the
left; and the
lords of the
Philistines
went
after them unto the
border of
Beth-shemesh.
\VS{13}And
{\ADD{they of}}
Beth-shemesh
{\ADD{were}}
reaping their
wheat
harvest in the
valley: and they lifted
up their
eyes, and
saw the
ark, and
rejoiced to
see
{\ADD{it}}.
\VS{14}And the
cart
came into the
field of
Joshua, a
Beth-shemite, and
stood there, where
{\ADD{there was}} a
great
stone: and they
clave the
wood of the
cart, and
offered the
kine a burnt
offering unto the
{\ND{
LORD}}.
\VS{15}And the
Levites took
down the
ark of the
{\ND{
LORD}}, and the
coffer that
{\ADD{was}} with it, wherein the
jewels of
gold
{\ADD{were}}, and
put
{\ADD{them}} on the
great
stone: and the
men of
Beth-shemesh
offered burnt
offerings and
sacrificed
sacrifices the same
day unto the
{\ND{
LORD}}.
\VS{16}And when the
five
lords of the
Philistines had
seen
{\ADD{it}}, they
returned to
Ekron the same
day.
\VS{17}And these
{\ADD{are}} the
golden
emerods which the
Philistines
returned
{\ADD{for}} a trespass
offering unto the
{\ND{
LORD}}; for
Ashdod
one, for
Gaza
one, for
Askelon
one, for
Gath
one, for
Ekron
one;
\VS{18}And the
golden
mice,
{\ADD{according to}} the
number of all the
cities of the
Philistines
{\ADD{belonging}} to the
five
lords,
{\ADD{both}} of
fenced
cities, and of
country
villages, even unto the
great
{\ADD{stone of}}
Abel, whereon they set
down the
ark of the
{\ND{
LORD}}:
{\ADD{which stone remaineth}} unto this
day in the
field of
Joshua, the
Beth-shemite.\FTNT{**}{{\FR 6.18 }
great…: or, great stone}
\par }{\PP \VS{19}¶ And he
smote the
men of
Beth-shemesh, because they had
looked into the
ark of the
{\ND{
LORD}}, even he
smote of the
people
fifty
thousand and threescore and
ten
men: and the
people
lamented, because the
{\ND{
LORD}} had
smitten
{\ADD{many}} of the
people with a
great
slaughter.
\VS{20}And the
men of
Beth-shemesh
said, Who is
able to
stand
before this
holy
{\ND{
LORD}}
God? and to whom shall he go
up from us?
\par }{\PP \VS{21}¶ And they
sent
messengers to the
inhabitants of
Kirjath-jearim,
saying, The
Philistines have brought
again the
ark of the
{\ND{
LORD}}; come ye
down,
{\ADD{and}} fetch it
up to you.

\par }\Chap{7}{\PP \VerseOne{1}And the
men of
Kirjath-jearim
came, and fetched
up the
ark of the
{\ND{
LORD}}, and
brought it into the
house of
Abinadab in the
hill, and
sanctified
Eleazar his
son to
keep the
ark of the
{\ND{
LORD}}.
\VS{2}And it came to pass,
while the
ark
abode in
Kirjath-jearim, that the
time was
long; for it was
twenty
years: and all the
house of
Israel
lamented
after the
{\ND{
LORD}}.
\par }{\PP \VS{3}¶ And
Samuel
spake unto all the
house of
Israel,
saying, If ye do
return unto the
{\ND{
LORD}} with all your
hearts,
{\ADD{then}} put
away the
strange
gods and
Ashtaroth from
among you, and
prepare your
hearts unto the
{\ND{
LORD}}, and
serve him only: and he will
deliver you out of the
hand of the
Philistines.
\VS{4}Then the
children of
Israel did put
away
Baalim and
Ashtaroth, and
served the
{\ND{
LORD}} only.
\VS{5}And
Samuel
said,
Gather all
Israel to
Mizpeh, and I will
pray for you unto the
{\ND{
LORD}}.
\VS{6}And they gathered
together to
Mizpeh, and
drew
water, and poured
{\ADD{it}}
out
before the
{\ND{
LORD}}, and
fasted on that
day, and
said there, We have
sinned against the
{\ND{
LORD}}. And
Samuel
judged the
children of
Israel in
Mizpeh.
\par }{\PP \VS{7}And when the
Philistines
heard that the
children of
Israel were gathered
together to
Mizpeh, the
lords of the
Philistines went
up against
Israel. And when the
children of
Israel
heard
{\ADD{it}}, they were
afraid
of the
Philistines.
\VS{8}And the
children of
Israel
said to
Samuel,
Cease not to
cry unto the
{\ND{
LORD}} our
God for us, that he will
save us out of the
hand of the
Philistines.\FTNT{*}{{\FR 7.8 }
Cease…: Heb. Be not silent from us from crying}
\par }{\PP \VS{9}¶ And
Samuel
took
a
sucking
lamb, and
offered
{\ADD{it for}} a burnt
offering
wholly unto the
{\ND{
LORD}}: and
Samuel
cried unto the
{\ND{
LORD}} for
Israel; and the
{\ND{
LORD}}
heard him.\FTNT{†}{{\FR 7.9 }
heard: or, answered}
\VS{10}And as
Samuel was offering
up the burnt
offering, the
Philistines drew
near to
battle against
Israel: but the
{\ND{
LORD}}
thundered with a
great
thunder on that
day upon the
Philistines, and
discomfited them; and they were
smitten
before
Israel.
\VS{11}And the
men of
Israel went
out of
Mizpeh, and
pursued the
Philistines, and
smote them, until
{\ADD{they came}} under
Beth-car.
\VS{12}Then
Samuel
took
a
stone, and
set
{\ADD{it}} between
Mizpeh and
Shen, and
called the
name of it
Eben-ezer,
saying, Hitherto hath the
{\ND{
LORD}}
helped us.\FTNT{‡}{{\FR 7.12 }
Eben-ezer: that is, The stone of help}
\par }{\PP \VS{13}¶ So the
Philistines were
subdued, and they
came no
more into the
coast of
Israel: and the
hand of the
{\ND{
LORD}} was against the
Philistines all the
days of
Samuel.
\VS{14}And the
cities which the
Philistines had
taken from
Israel were
restored to
Israel, from
Ekron even unto
Gath; and the
coasts thereof did
Israel
deliver out of the
hands of the
Philistines. And there was
peace between
Israel and the
Amorites.
\VS{15}And
Samuel
judged
Israel all the
days of his
life.
\VS{16}And he
went
from
year to
year in
circuit to
Beth-el, and
Gilgal, and
Mizpeh, and
judged
Israel in all those
places.\FTNT{§}{{\FR 7.16 }
in circuit: Heb. and he circuited}
\VS{17}And his
return
{\ADD{was}} to
Ramah; for there
{\ADD{was}} his
house; and there he
judged
Israel; and there he
built an
altar unto the
{\ND{
LORD}}.

\par }\Chap{8}{\PP \VerseOne{1}And it came to pass, when
Samuel was
old, that he
made his
sons
judges over
Israel.
\VS{2}Now the
name of his
firstborn was
Joel; and the
name of his
second,
Abiah:
{\ADD{they were}}
judges in
Beer-sheba.
\VS{3}And his
sons
walked not in his
ways, but turned
aside
after
lucre, and
took
bribes, and
perverted
judgment.
\par }{\PP \VS{4}Then all the
elders of
Israel gathered themselves
together, and
came to
Samuel unto
Ramah,
\VS{5}And
said unto him, Behold, thou art
old, and thy
sons
walk not in thy
ways: now
make us a
king to
judge us like all the
nations.
\par }{\PP \VS{6}¶ But the
thing
displeased
Samuel, when they
said,
Give us a
king to
judge us. And
Samuel
prayed unto the
{\ND{
LORD}}.\FTNT{*}{{\FR 8.6 }
displeased: Heb. was evil in the eyes of}
\VS{7}And the
{\ND{
LORD}}
said unto
Samuel,
Hearken unto the
voice of the
people in all that they
say unto thee: for they have not
rejected thee, but they have
rejected me, that I should not
reign over them.
\VS{8}According to all the
works which they have
done since the
day that I brought them
up out of
Egypt even unto this
day, wherewith they have
forsaken me, and
served
other
gods, so
do they also unto thee.
\VS{9}Now therefore
hearken unto their
voice:
howbeit yet
protest
solemnly unto them, and
shew them the
manner of the
king that shall
reign over them.\FTNT{†}{{\FR 8.9 }
hearken…: or, obey}\FTNT{‡}{{\FR 8.9 }
howbeit…: or, notwithstanding when thou hast solemnly protested against them then thou shalt}
\par }{\PP \VS{10}¶ And
Samuel
told all the
words of the
{\ND{
LORD}} unto the
people that
asked of him a
king.
\VS{11}And he
said, This will be the
manner of the
king that shall
reign over you: He will
take your
sons, and
appoint
{\ADD{them}} for himself, for his
chariots, and
{\ADD{to be}} his
horsemen; and
{\ADD{some}} shall
run
before his
chariots.
\VS{12}And he will
appoint him
captains over
thousands, and
captains over
fifties; and
{\ADD{will set them}} to
ear his
ground, and to
reap his
harvest, and to
make his
instruments of
war, and
instruments of his
chariots.
\VS{13}And he will
take your
daughters
{\ADD{to be}}
confectionaries, and
{\ADD{to be}}
cooks, and
{\ADD{to be}}
bakers.
\VS{14}And he will
take your
fields, and your
vineyards, and your
oliveyards,
{\ADD{even}} the
best
{\ADD{of them}}, and
give
{\ADD{them}} to his
servants.
\VS{15}And he will take the
tenth of your
seed, and of your
vineyards, and
give to his
officers, and to his
servants.\FTNT{§}{{\FR 8.15 }
officers: Heb. eunuchs}
\VS{16}And he will
take your
menservants, and your
maidservants, and your
goodliest young
men, and your
asses, and
put
{\ADD{them}} to his
work.
\VS{17}He will take the
tenth of your
sheep: and ye shall be his
servants.
\VS{18}And ye shall cry
out in that
day
because of your
king which ye shall have
chosen you; and the
{\ND{
LORD}} will not
hear you in that
day.
\par }{\PP \VS{19}¶ Nevertheless the
people
refused to
obey the
voice of
Samuel; and they
said, Nay; but we will have a
king over us;
\VS{20}That we also may be like all the
nations; and that our
king may
judge us, and go
out
before us, and
fight our
battles.
\VS{21}And
Samuel
heard all the
words of the
people, and he
rehearsed them in the
ears of the
{\ND{
LORD}}.
\VS{22}And the
{\ND{
LORD}}
said to
Samuel,
Hearken unto their
voice, and
make them a
king. And
Samuel
said unto the
men of
Israel,
Go ye every
man unto his
city.

\par }\Chap{9}{\PP \VerseOne{1}Now there was a
man of
Benjamin, whose
name
{\ADD{was}}
Kish, the
son of
Abiel, the
son of
Zeror, the
son of
Bechorath, the
son of
Aphiah, a
Benjamite, a
mighty
man of
power.\FTNT{*}{{\FR 9.1 }
a Benjamite: or, the son of a man of Jemini}\FTNT{†}{{\FR 9.1 }
power: or, substance}
\VS{2}And he had a
son, whose
name
{\ADD{was}}
Saul, a choice young
man, and a
goodly: and
{\ADD{there was}} not among the
children of
Israel a
goodlier
person than he: from his
shoulders and
upward
{\ADD{he was}}
higher than any of the
people.
\par }{\PP \VS{3}And the
asses of
Kish
Saul’s
father were
lost. And
Kish
said to
Saul his
son,
Take now
one of the
servants with thee, and
arise,
go
seek the
asses.
\VS{4}And he passed
through
mount
Ephraim, and passed
through the
land of
Shalisha, but they
found
{\ADD{them}} not: then they passed
through the
land of
Shalim, and
{\ADD{there they were}} not: and he passed
through the
land of the
Benjamites, but they
found
{\ADD{them}} not.
\VS{5}{\ADD{And}} when they were
come to the
land of
Zuph,
Saul
said to his
servant that
{\ADD{was}} with him,
Come, and let us
return; lest my
father
leave
{\ADD{caring}} for the
asses, and take
thought for us.
\VS{6}And he
said unto him, Behold now,
{\ADD{there is}} in this
city a
man of
God, and
{\ADD{he is}} an
honourable
man; all that he
saith
cometh surely to
pass: now let us
go thither; peradventure he can
shew us our
way that we should
go.
\VS{7}Then
said
Saul to his
servant, But, behold,
{\ADD{if}} we
go, what shall we
bring the
man? for the
bread is
spent in our
vessels, and
{\ADD{there is}} not a
present to
bring to the
man of
God: what have we?\FTNT{‡}{{\FR 9.7 }
is spent…: Heb. is gone out of, etc}\FTNT{§}{{\FR 9.7 }
have we: Heb. is with us?}
\VS{8}And the
servant
answered
Saul
again, and
said, Behold, I
have here at
hand the fourth
part of a
shekel of
silver:
{\ADD{that}} will I
give to the
man of
God, to
tell us our
way.\FTNT{**}{{\FR 9.8 }
I have…: Heb. there is found in my hand}
\VS{9}(Beforetime in
Israel, when a
man
went to
enquire of
God, thus he
spake,
Come, and let us
go to the
seer: for
{\ADD{he that is}}
now
{\ADD{called}} a
Prophet was
beforetime
called a
Seer.)
\VS{10}Then
said
Saul to his
servant,
Well
said;
come, let us
go. So they
went unto the
city where the
man of
God
{\ADD{was}}.\FTNT{††}{{\FR 9.10 }
Well said: Heb. Thy word is good}
\par }{\PP \VS{11}¶
{\ADD{And}} as they went
up the
hill to the
city, they
found young
maidens going
out to
draw
water, and
said unto them,
Is the
seer here?\FTNT{‡‡}{{\FR 9.11 }
the hill…: Heb. in the ascent of the city}
\VS{12}And they
answered them, and
said, He
is; behold,
{\ADD{he is}}
before you: make
haste now, for he
came to
day to the
city; for
{\ADD{there is}} a
sacrifice of the
people to
day in the high
place:\FTNT{§§}{{\FR 9.12 }
sacrifice: or, feast}
\VS{13}As soon as ye be
come into the
city, ye shall
straightway
find him, before he go
up to the high
place to
eat: for the
people will not
eat until he
come, because he doth
bless the
sacrifice;
{\ADD{and}}
afterwards they
eat that be
bidden. Now therefore get you
up; for about this
time ye shall
find him.\FTNT{***}{{\FR 9.13 }
this time: Heb. to day}
\VS{14}And they went
up
into the
city:
{\ADD{and}} when they were
come into the
city, behold,
Samuel came
out
against them, for to go
up to the high
place.
\par }{\PP \VS{15}¶ Now the
{\ND{
LORD}} had
told
Samuel in his
ear
a
day
before
Saul
came,
saying,\FTNT{†††}{{\FR 9.15 }
told…: Heb. revealed the ear of Samuel}
\VS{16}To
morrow about this
time I will
send thee a
man out of the
land of
Benjamin, and thou shalt
anoint him
{\ADD{to be}}
captain over my
people
Israel, that he may
save my
people out of the
hand of the
Philistines: for I have
looked upon my
people, because their
cry is
come unto me.
\VS{17}And when
Samuel
saw
Saul, the
{\ND{
LORD}}
said unto him, Behold the
man whom I
spake to thee of! this same shall
reign over my
people.\FTNT{‡‡‡}{{\FR 9.17 }
reign over: Heb. restrain in}
\par }{\PP \VS{18}Then
Saul drew
near to
Samuel
in the
gate, and
said,
Tell me, I pray thee,
where the
seer’s
house
{\ADD{is}}.
\VS{19}And
Samuel
answered
Saul, and
said, I
{\ADD{am}} the
seer: go
up
before me unto the high
place; for ye shall
eat with me to
day, and to
morrow I will let thee
go, and will
tell thee all that
{\ADD{is}} in thine
heart.
\VS{20}And as for thine
asses that were
lost
three
days
ago,
set not thy
mind on them; for they are
found. And on whom
{\ADD{is}} all the
desire of
Israel?
{\ADD{Is it}} not on thee, and on all thy
father’s
house?\FTNT{§§§}{{\FR 9.20 }
three…: Heb. to day three days}
\VS{21}And
Saul
answered and
said,
{\ADD{Am}} not I a
Benjamite, of the
smallest of the
tribes of
Israel? and my
family the
least of all the
families of the
tribe of
Benjamin? wherefore then
speakest thou
so to me?\FTNT{*}{{\FR 9.21 }
so…: Heb. according to this word}
\VS{22}And
Samuel
took
Saul and his
servant, and
brought them into the
parlour, and made them
sit in the
chiefest
place among them that were
bidden, which
{\ADD{were}} about
thirty
persons.
\VS{23}And
Samuel
said unto the
cook,
Bring the
portion which I
gave thee, of which I
said unto thee,
Set it by thee.
\VS{24}And the
cook took
up the
shoulder, and
{\ADD{that}} which
{\ADD{was}} upon it, and
set
{\ADD{it}}
before
Saul. And
{\ADD{Samuel}}
said, Behold that which is
left!
set
{\ADD{it}}
before thee,
{\ADD{and}}
eat: for unto this
time hath it been
kept for thee since I
said, I have
invited the
people. So
Saul did
eat with
Samuel that
day.\FTNT{†}{{\FR 9.24 }
left: or, reserved}
\par }{\PP \VS{25}¶ And when they were come
down from the high
place into the
city,
{\ADD{Samuel}}
communed with
Saul upon the top of the
house.
\VS{26}And they arose
early: and it came to pass about the
spring of the
day, that
Samuel
called
Saul to the top of the
house,
saying,
Up, that I may send thee
away. And
Saul
arose, and they went
out
both of them, he and
Samuel,
abroad.
\VS{27}{\ADD{And}} as they were going
down to the
end of the
city,
Samuel
said to
Saul,
Bid the
servant pass
on
before us, (and he passed
on,) but
stand thou still a
while, that I may
shew thee the
word of
God.\FTNT{‡}{{\FR 9.27 }
a while: Heb. to day}

\par }\Chap{10}{\PP \VerseOne{1}Then
Samuel
took a
vial of
oil, and
poured
{\ADD{it}} upon his
head, and
kissed him, and
said,
{\ADD{Is it}} not because the
{\ND{
LORD}} hath
anointed thee
{\ADD{to be}}
captain over his
inheritance?
\VS{2}When thou art
departed from
me to
day, then thou shalt
find
two
men by
Rachel’s
sepulchre in the
border of
Benjamin at
Zelzah; and they will
say unto thee, The
asses which thou
wentest to
seek are
found: and, lo, thy
father hath
left the
care of the
asses, and
sorroweth for you,
saying, What shall I
do for my
son?\FTNT{*}{{\FR 10.2 }
care: Heb. business}
\VS{3}Then shalt thou go
on
forward from thence, and thou shalt
come to the
plain of
Tabor, and there shall
meet thee
three
men going
up to
God to
Beth-el,
one
carrying
three
kids, and
another
carrying
three
loaves of
bread, and
another
carrying a
bottle of
wine:
\VS{4}And they
will
salute thee, and
give thee
two
{\ADD{loaves}} of
bread; which thou shalt
receive of their
hands.\FTNT{†}{{\FR 10.4 }
salute…: Heb. ask thee of peace}
\VS{5}After that thou shalt
come to the
hill of
God, where
{\ADD{is}} the
garrison of the
Philistines: and it shall come to pass, when thou art come
thither to the
city, that thou shalt
meet a
company of
prophets coming
down from the high
place with a
psaltery, and a
tabret, and a
pipe, and a
harp,
before them; and they shall
prophesy:
\VS{6}And the
Spirit of the
{\ND{
LORD}} will
come upon thee, and thou shalt
prophesy with them, and shalt be
turned into
another
man.
\VS{7}And let it be, when these
signs are
come unto thee,
{\ADD{that}} thou
do as
occasion
serve thee; for
God
{\ADD{is}} with thee.\FTNT{‡}{{\FR 10.7 }
And…: Heb. And it shall come to pass, that when these signs, etc}\FTNT{§}{{\FR 10.7 }
that…: Heb. do for thee as thine hand shall find}
\VS{8}And thou shalt go
down
before me to
Gilgal; and, behold, I will come
down unto thee, to
offer burnt
offerings,
{\ADD{and}} to
sacrifice
sacrifices of peace
offerings:
seven
days shalt thou
tarry, till I
come to thee, and
shew thee what thou shalt
do.
\par }{\PP \VS{9}¶ And it was
{\ADD{so}}, that when he had
turned his
back to
go from
Samuel,
God
gave him
another
heart: and all those
signs
came to pass that
day.\FTNT{**}{{\FR 10.9 }
back: Heb. shoulder}\FTNT{††}{{\FR 10.9 }
gave: Heb. turned}
\VS{10}And when they
came thither to the
hill, behold, a
company of
prophets
met him; and the
Spirit of
God
came upon him, and he
prophesied
among them.
\VS{11}And it came to pass, when all that
knew him
beforetime
saw that, behold, he
prophesied among the
prophets, then the
people
said
one to
another, What
{\ADD{is}} this
{\ADD{that}} is come unto the
son of
Kish?
{\ADD{Is}}
Saul also among the
prophets?\FTNT{‡‡}{{\FR 10.11 }
one…: Heb. a man to his neighbour}
\VS{12}And
one of the same place
answered and
said, But who
{\ADD{is}} their
father? Therefore it became a
proverb,
{\ADD{Is}}
Saul also among the
prophets?\FTNT{§§}{{\FR 10.12 }
of…: Heb. from thence}
\VS{13}And when he had made an
end of
prophesying, he
came to the high
place.
\par }{\PP \VS{14}¶ And
Saul’s
uncle
said unto him and to his
servant,
Whither
went ye? And he
said, To
seek the
asses: and when we
saw that
{\ADD{they were}} no
where, we
came to
Samuel.
\VS{15}And
Saul’s
uncle
said,
Tell me, I pray thee, what
Samuel
said unto you.
\VS{16}And
Saul
said unto his
uncle, He
told us
plainly that the
asses were
found. But of the
matter of the
kingdom, whereof
Samuel
spake, he
told him not.
\par }{\PP \VS{17}¶ And
Samuel
called the
people together unto the
{\ND{
LORD}} to
Mizpeh;
\VS{18}And
said unto the
children of
Israel, Thus
saith the
{\ND{
LORD}}
God of
Israel, I brought
up
Israel out of
Egypt, and
delivered you out of the
hand of the
Egyptians, and out of the
hand of all
kingdoms,
{\ADD{and}} of them that
oppressed you:
\VS{19}And ye have this
day
rejected your
God, who himself
saved you out of all your
adversities and your
tribulations; and ye have
said unto him,
{\ADD{Nay}}, but
set a
king over us. Now therefore
present yourselves
before the
{\ND{
LORD}} by your
tribes, and by your
thousands.
\VS{20}And when
Samuel had caused all the
tribes of
Israel to come
near, the
tribe of
Benjamin was
taken.
\VS{21}When he had caused the
tribe of
Benjamin to come
near by their
families, the
family of
Matri was
taken, and
Saul the
son of
Kish was
taken: and when they
sought him, he could not be
found.
\VS{22}Therefore they
enquired of the
{\ND{
LORD}} further, if the
man should yet
come
thither. And the
{\ND{
LORD}}
answered, Behold, he hath
hid himself among the
stuff.
\VS{23}And they
ran and
fetched him thence: and when he
stood
among the
people, he was
higher than any of the
people from his
shoulders and
upward.
\VS{24}And
Samuel
said to all the
people,
See ye him whom the
{\ND{
LORD}} hath
chosen, that
{\ADD{there is}} none like him among all the
people? And all the
people
shouted, and
said, God
save the
king.\FTNT{***}{{\FR 10.24 }
God…: Heb. Let the king live}
\VS{25}Then
Samuel
told the
people the
manner of the
kingdom, and
wrote
{\ADD{it}} in a
book, and laid
{\ADD{it}}
up
before the
{\ND{
LORD}}. And
Samuel
sent all the
people
away, every
man to his
house.
\par }{\PP \VS{26}¶ And
Saul also
went
home to
Gibeah; and there
went with him a band of
men, whose
hearts
God had
touched.
\VS{27}But the
children of
Belial
said, How shall this man
save us? And they
despised him, and
brought him no
presents. But he held his
peace.\FTNT{†††}{{\FR 10.27 }
held…: or, was as though he had been deaf}

\par }\Chap{11}{\PP \VerseOne{1}Then
Nahash the
Ammonite came
up, and
encamped against
Jabesh-gilead: and all the
men of
Jabesh
said unto
Nahash,
Make a
covenant with us, and we will
serve thee.
\VS{2}And
Nahash the
Ammonite
answered them, On
this
{\ADD{condition}} will I
make
{\ADD{a covenant}} with you, that I may thrust
out all your
right
eyes, and
lay it
{\ADD{for}} a
reproach upon all
Israel.
\VS{3}And the
elders of
Jabesh
said unto him, Give us
seven
days’
respite, that we may
send
messengers unto all the
coasts of
Israel: and then, if
{\ADD{there be}} no man to
save us, we will come
out to thee.\FTNT{*}{{\FR 11.3 }
Give…: Heb. Forbear us}
\par }{\PP \VS{4}¶ Then
came the
messengers to
Gibeah of
Saul, and
told the
tidings in the
ears of the
people: and all the
people lifted
up their
voices, and
wept.
\par }{\PP \VS{5}And, behold,
Saul
came
after the
herd out of the
field; and
Saul
said, What
{\ADD{aileth}} the
people that they
weep? And they
told him the
tidings of the
men of
Jabesh.
\VS{6}And the
Spirit of
God
came upon
Saul when he
heard those
tidings, and his
anger was
kindled
greatly.
\VS{7}And he
took a
yoke of
oxen, and hewed them in
pieces, and
sent
{\ADD{them}} throughout all the
coasts of
Israel by the
hands of
messengers,
saying, Whosoever cometh not
forth
after
Saul and
after
Samuel, so shall it be
done unto his
oxen. And the
fear of the
{\ND{
LORD}}
fell on the
people, and they came
out with
one
consent.\FTNT{†}{{\FR 11.7 }
with…: Heb. as one man}
\VS{8}And when he
numbered them in
Bezek, the
children of
Israel were
three
hundred
thousand, and the
men of
Judah
thirty
thousand.
\VS{9}And they
said unto the
messengers that
came, Thus shall ye
say unto the
men of
Jabesh-gilead, To
morrow, by
{\ADD{that time}} the
sun be
hot, ye shall have
help. And the
messengers
came and
shewed
{\ADD{it}} to the
men of
Jabesh; and they were
glad.\FTNT{‡}{{\FR 11.9 }
help: or, deliverance}
\VS{10}Therefore the
men of
Jabesh
said, To
morrow we will come
out unto you, and ye shall
do with us all that
seemeth
good unto you.
\VS{11}And it was
{\ADD{so}} on the
morrow, that
Saul
put the
people in
three
companies; and they
came into the
midst of the
host in the
morning
watch, and
slew the
Ammonites until the
heat of the
day: and it came to pass, that they which
remained were
scattered, so that
two of them were not
left
together.
\par }{\PP \VS{12}¶ And the
people
said unto
Samuel, Who
{\ADD{is}} he that
said, Shall
Saul
reign over us?
bring the
men, that we may put them to
death.
\VS{13}And
Saul
said, There shall not a
man be put to
death this
day: for to
day the
{\ND{
LORD}} hath
wrought
salvation in
Israel.
\VS{14}Then
said
Samuel to the
people,
Come, and let us
go to
Gilgal, and
renew the
kingdom there.
\VS{15}And all the
people
went to
Gilgal; and there they made
Saul
king
before the
{\ND{
LORD}} in
Gilgal; and there they
sacrificed
sacrifices of peace
offerings
before the
{\ND{
LORD}}; and there
Saul and all the
men of
Israel
rejoiced
greatly.

\par }\Chap{12}{\PP \VerseOne{1}And
Samuel
said unto all
Israel, Behold, I have
hearkened unto your
voice in all that ye
said unto me, and have
made a
king over you.
\VS{2}And now, behold, the
king
walketh
before you: and I am
old and
grayheaded; and, behold, my
sons
{\ADD{are}} with you: and I have
walked
before you from my
childhood unto this
day.
\VS{3}Behold, here I
{\ADD{am:}}
witness against me before the
{\ND{
LORD}}, and before his
anointed: whose
ox have I
taken? or whose
ass have I
taken? or whom have I
defrauded? whom have I
oppressed? or of whose
hand have I received
{\ADD{any}}
bribe to
blind mine
eyes therewith? and I will
restore it you.\FTNT{*}{{\FR 12.3 }
bribe: Heb. ransom}\FTNT{†}{{\FR 12.3 }
to blind…: or, that I should hide mine eyes at him}
\VS{4}And they
said, Thou hast not
defrauded us, nor
oppressed us, neither hast thou
taken
ought of any
man’s
hand.
\VS{5}And he
said unto them, The
{\ND{
LORD}}
{\ADD{is}}
witness against you, and his
anointed
{\ADD{is}}
witness this
day, that ye have not
found
ought in my
hand. And they
answered,
{\ADD{He is}}
witness.
\par }{\PP \VS{6}¶ And
Samuel
said unto the
people,
{\ADD{It is}} the
{\ND{
LORD}} that
advanced
Moses and
Aaron, and that
brought your
fathers
up out of the
land of
Egypt.\FTNT{‡}{{\FR 12.6 }
advanced: or, made}
\VS{7}Now therefore stand
still, that I may
reason with you
before the
{\ND{
LORD}} of all the righteous
acts of the
{\ND{
LORD}}, which he
did to you and to your
fathers.\FTNT{§}{{\FR 12.7 }
righteous…: Heb. righteousnesses, or, benefits}\FTNT{**}{{\FR 12.7 }
to: Heb. with}
\VS{8}When
Jacob was
come into
Egypt, and your
fathers
cried unto the
{\ND{
LORD}}, then the
{\ND{
LORD}}
sent
Moses and
Aaron, which brought
forth your
fathers out of
Egypt, and made them
dwell in this
place.
\VS{9}And when they
forgat the
{\ND{
LORD}} their
God, he
sold them into the
hand of
Sisera,
captain of the
host of
Hazor, and into the
hand of the
Philistines, and into the
hand of the
king of
Moab, and they
fought against them.
\VS{10}And they
cried unto the
{\ND{
LORD}}, and
said, We have
sinned, because we have
forsaken the
{\ND{
LORD}}, and have
served
Baalim and
Ashtaroth: but now
deliver us out of the
hand of our
enemies, and we will
serve thee.
\VS{11}And the
{\ND{
LORD}}
sent
Jerubbaal, and
Bedan, and
Jephthah, and
Samuel, and
delivered you out of the
hand of your
enemies on every
side, and ye
dwelled
safe.
\VS{12}And when ye
saw that
Nahash the
king of the
children of
Ammon
came against you, ye
said unto me, Nay; but a
king shall
reign over us: when the
{\ND{
LORD}} your
God
{\ADD{was}} your
king.
\VS{13}Now therefore behold the
king whom ye have
chosen,
{\ADD{and}} whom ye have
desired! and, behold, the
{\ND{
LORD}} hath
set a
king over you.
\VS{14}If ye will
fear the
{\ND{
LORD}}, and
serve him, and
obey his
voice, and not
rebel against the
commandment of the
{\ND{
LORD}}, then shall both ye and also the
king that
reigneth over you continue
following the
{\ND{
LORD}} your
God:\FTNT{††}{{\FR 12.14 }
commandment: Heb. mouth}\FTNT{‡‡}{{\FR 12.14 }
continue…: Heb. be after}
\VS{15}But if ye will not
obey the
voice of the
{\ND{
LORD}}, but
rebel against the
commandment of the
{\ND{
LORD}}, then shall the
hand of the
{\ND{
LORD}} be against you, as
{\ADD{it was}} against your
fathers.
\par }{\PP \VS{16}¶ Now
therefore
stand and
see this
great
thing, which the
{\ND{
LORD}} will
do before your
eyes.
\VS{17}{\ADD{Is it}} not
wheat
harvest to
day? I will
call unto the
{\ND{
LORD}}, and he shall
send
thunder and
rain; that ye may
perceive and
see that your
wickedness
{\ADD{is}}
great, which ye have
done in the
sight of the
{\ND{
LORD}}, in
asking you a
king.
\VS{18}So
Samuel
called unto the
{\ND{
LORD}}; and the
{\ND{
LORD}}
sent
thunder and
rain that
day: and all the
people
greatly
feared the
{\ND{
LORD}} and
Samuel.
\VS{19}And all the
people
said unto
Samuel,
Pray for thy
servants unto the
{\ND{
LORD}} thy
God, that we
die not: for we have
added unto all our
sins
{\ADD{this}}
evil, to
ask us a
king.
\par }{\PP \VS{20}¶ And
Samuel
said unto the
people,
Fear not: ye have
done all this
wickedness: yet turn not
aside from
following the
{\ND{
LORD}}, but
serve the
{\ND{
LORD}} with all your
heart;
\VS{21}And turn ye not
aside: for
{\ADD{then should ye go}}
after
vain
{\ADD{things}}, which cannot
profit nor
deliver; for they
{\ADD{are}}
vain.
\VS{22}For the
{\ND{
LORD}} will not
forsake his
people for his
great
name’s sake: because it hath
pleased the
{\ND{
LORD}} to
make you his
people.
\VS{23}Moreover as for
me, God
forbid that I should
sin against the
{\ND{
LORD}} in
ceasing to
pray for you: but I will
teach you the
good and the
right
way:\FTNT{§§}{{\FR 12.23 }
in: Heb. from}
\VS{24}Only
fear the
{\ND{
LORD}}, and
serve him in
truth with all your
heart: for
consider how great
{\ADD{things}} he hath
done for you.\FTNT{***}{{\FR 12.24 }
how…: or, what a great thing}
\VS{25}But if ye shall
still do
wickedly, ye shall be
consumed, both ye and your
king.

\par }\Chap{13}{\PP \VerseOne{1}Saul
reigned one
year; and when he had
reigned
two
years over
Israel,\FTNT{*}{{\FR 13.1 }
reigned one…: Heb. the son of one year in his reigning}
\VS{2}Saul
chose him
three
thousand
{\ADD{men}} of
Israel;
{\ADD{whereof}} two
thousand were with
Saul in
Michmash and in
mount
Beth-el, and a
thousand were with
Jonathan in
Gibeah of
Benjamin: and the
rest of the
people he
sent every
man to his
tent.
\VS{3}And
Jonathan
smote the
garrison of the
Philistines that
{\ADD{was}} in
Geba, and the
Philistines
heard
{\ADD{of it}}. And
Saul
blew the
trumpet throughout all the
land,
saying, Let the
Hebrews
hear.\FTNT{†}{{\FR 13.3 }
Geba: or, the hill}
\VS{4}And all
Israel
heard
say
{\ADD{that}}
Saul had
smitten a
garrison of the
Philistines, and
{\ADD{that}}
Israel also was had in
abomination with the
Philistines. And the
people were called
together
after
Saul to
Gilgal.\FTNT{‡}{{\FR 13.4 }
was…: Heb. did stink}
\par }{\PP \VS{5}¶ And the
Philistines gathered themselves
together to
fight with
Israel,
thirty
thousand
chariots, and
six
thousand
horsemen, and
people as the
sand which
{\ADD{is}} on the
sea
shore in
multitude: and they came
up, and
pitched in
Michmash,
eastward from
Beth-aven.
\VS{6}When the
men of
Israel
saw that they were in a
strait, (for the
people were
distressed,) then the
people did
hide themselves in
caves, and in
thickets, and in
rocks, and in high
places, and in
pits.
\VS{7}And
{\ADD{some of}} the
Hebrews went
over
Jordan to the
land of
Gad and
Gilead. As for
Saul, he
{\ADD{was}} yet in
Gilgal, and all the
people
followed him
trembling.\FTNT{§}{{\FR 13.7 }
followed…: Heb. trembled after him}
\par }{\PP \VS{8}¶ And he
tarried
seven
days, according to the set
time that
Samuel
{\ADD{had appointed:}} but
Samuel
came not to
Gilgal; and the
people were
scattered from him.
\VS{9}And
Saul
said, Bring
hither a burnt
offering to me, and peace
offerings. And he
offered the burnt
offering.
\VS{10}And it came to pass, that as soon as he had made an
end of
offering the burnt
offering, behold,
Samuel
came; and
Saul went
out to
meet him, that he might
salute him.\FTNT{**}{{\FR 13.10 }
salute: Heb. bless}
\par }{\PP \VS{11}¶ And
Samuel
said, What hast thou
done? And
Saul
said, Because I
saw that the
people were
scattered from me, and
{\ADD{that}} thou
camest not within the
days
appointed, and
{\ADD{that}} the
Philistines gathered themselves
together at
Michmash;
\VS{12}Therefore
said I, The
Philistines will come
down now upon me to
Gilgal, and I have not made
supplication
unto the
{\ND{
LORD}}: I
forced myself therefore, and
offered a burnt
offering.\FTNT{††}{{\FR 13.12 }
made…: Heb. intreated the face}
\VS{13}And
Samuel
said to
Saul, Thou hast done
foolishly: thou hast not
kept the
commandment of the
{\ND{
LORD}} thy
God, which he
commanded thee: for now would the
{\ND{
LORD}} have
established thy
kingdom upon
Israel
for
ever.
\VS{14}But now thy
kingdom shall not
continue: the
{\ND{
LORD}} hath
sought him a
man after his own
heart, and the
{\ND{
LORD}} hath
commanded him
{\ADD{to be}}
captain over his
people, because thou hast not
kept
{\ADD{that}} which the
{\ND{
LORD}}
commanded thee.
\par }{\PP \VS{15}And
Samuel
arose, and gat him
up from
Gilgal unto
Gibeah of
Benjamin. And
Saul
numbered the
people
{\ADD{that were}}
present with him, about
six
hundred
men.\FTNT{‡‡}{{\FR 13.15 }
present: Heb. found}
\VS{16}And
Saul, and
Jonathan his
son, and the
people
{\ADD{that were}}
present with them,
abode in
Gibeah of
Benjamin: but the
Philistines
encamped in
Michmash.\FTNT{§§}{{\FR 13.16 }
present: Heb. found}\FTNT{***}{{\FR 13.16 }
Gibeah: Heb. Geba}
\par }{\PP \VS{17}¶ And the
spoilers came
out of the
camp of the
Philistines in
three
companies:
one
company
turned unto the
way
{\ADD{that leadeth to}}
Ophrah, unto the
land of
Shual:
\VS{18}And
another
company
turned the
way
{\ADD{to}}
Beth-horon: and
another
company
turned
{\ADD{to}} the
way of the
border that
looketh to the
valley of
Zeboim toward the
wilderness.
\par }{\PP \VS{19}¶ Now there was no
smith
found throughout all the
land of
Israel: for the
Philistines
said, Lest the
Hebrews
make
{\ADD{them}}
swords or
spears:
\VS{20}But all the
Israelites went
down to the
Philistines, to
sharpen every
man his
share, and his
coulter, and his
axe, and his
mattock.
\VS{21}Yet they had a
file for the
mattocks, and for the
coulters, and for the
forks, and for the
axes, and to
sharpen the
goads.\FTNT{†††}{{\FR 13.21 }
a file: Heb. a file with mouths, or Heb. a pim, a third of a shekel}\FTNT{‡‡‡}{{\FR 13.21 }
sharpen: Heb. set}
\VS{22}So it came to pass in the
day of
battle, that there was neither
sword nor
spear
found in the
hand of any of the
people that
{\ADD{were}} with
Saul and
Jonathan: but with
Saul and with
Jonathan his
son was there
found.
\VS{23}And the
garrison of the
Philistines went
out to the
passage of
Michmash.\FTNT{§§§}{{\FR 13.23 }
garrison: or, standing camp}

\par }\Chap{14}{\PP \VerseOne{1}Now it came to pass upon a
day, that
Jonathan the
son of
Saul
said unto the young
man that
bare his
armour,
Come, and let us go
over to the
Philistines’
garrison, that
{\ADD{is}} on the other
side. But he
told not his
father.\FTNT{*}{{\FR 14.1 }
it came…: or, there was a day}
\VS{2}And
Saul
tarried in the
uttermost part of
Gibeah under a pomegranate
tree which
{\ADD{is}} in
Migron: and the
people that
{\ADD{were}} with him
{\ADD{were}} about
six
hundred
men;
\VS{3}And
Ahiah, the
son of
Ahitub,
I-chabod’s
brother, the
son of
Phinehas, the
son of
Eli, the
{\ND{
LORD’s}}
priest in
Shiloh,
wearing an
ephod. And the
people
knew not that
Jonathan was
gone.\FTNT{†}{{\FR 14.3 }
Ahiah: called Ahimelech}
\par }{\PP \VS{4}¶ And between the
passages, by which
Jonathan
sought to go
over unto the
Philistines’
garrison,
{\ADD{there was}} a
sharp
rock on the one
side, and a
sharp
rock on the other
side: and the
name of the
one
{\ADD{was}}
Bozez, and the
name of the
other
Seneh.
\VS{5}The
forefront of the
one
{\ADD{was}}
situate
northward over
against
Michmash, and the
other
southward over
against
Gibeah.\FTNT{‡}{{\FR 14.5 }
forefront: Heb. tooth}
\VS{6}And
Jonathan
said to the young
man that
bare his
armour,
Come, and let us go
over unto the
garrison of these
uncircumcised: it may be that the
{\ND{
LORD}} will
work for us: for
{\ADD{there is}} no
restraint to the
{\ND{
LORD}} to
save by
many or by
few.
\VS{7}And his
armourbearer
said unto him,
Do all that
{\ADD{is}} in thine
heart:
turn thee; behold, I
{\ADD{am}} with thee according to thy
heart.
\VS{8}Then
said
Jonathan, Behold, we will pass
over unto
{\ADD{these}}
men, and we will
discover ourselves unto them.
\VS{9}If they
say thus unto us,
Tarry until we
come to you; then we will stand
still in our place, and will not go
up unto them.\FTNT{§}{{\FR 14.9 }
Tarry: Heb. Be still}
\VS{10}But if they
say thus, Come
up unto us; then we will go
up: for the
{\ND{
LORD}} hath
delivered them into our
hand: and this
{\ADD{shall be}} a
sign unto us.
\VS{11}And
both of them
discovered themselves unto the
garrison of the
Philistines: and the
Philistines
said, Behold, the
Hebrews come
forth out of the
holes where they had
hid themselves.
\VS{12}And the
men of the
garrison
answered
Jonathan and his
armourbearer, and
said, Come
up to us, and we will
shew you a
thing. And
Jonathan
said unto his
armourbearer, Come
up
after me: for the
{\ND{
LORD}} hath
delivered them into the
hand of
Israel.
\VS{13}And
Jonathan climbed
up upon his
hands and upon his
feet, and his
armourbearer
after him: and they
fell
before
Jonathan; and his
armourbearer
slew
after him.
\VS{14}And that
first
slaughter, which
Jonathan and his
armourbearer
made, was about
twenty
men, within as it were an
half
acre of
land,
{\ADD{which}} a
yoke
{\ADD{of oxen might plow}}.\FTNT{**}{{\FR 14.14 }
an…: or, half a furrow of an acre of land}
\VS{15}And there was
trembling in the
host, in the
field, and among all the
people: the
garrison, and the
spoilers, they also
trembled, and the
earth
quaked: so it was a very
great
trembling.\FTNT{††}{{\FR 14.15 }
a very…: Heb. a trembling of God}
\par }{\PP \VS{16}And the
watchmen of
Saul in
Gibeah of
Benjamin
looked; and, behold, the
multitude melted
away, and they
went on beating
down
{\ADD{one another}}.
\VS{17}Then
said
Saul unto the
people that
{\ADD{were}} with him,
Number now, and
see who is
gone from us. And when they had
numbered, behold,
Jonathan and his
armourbearer
{\ADD{were}} not
{\ADD{there}}.
\VS{18}And
Saul
said unto
Ahiah, Bring
hither the
ark of
God. For the
ark of
God was at that
time with the
children of
Israel.
\par }{\PP \VS{19}¶ And it came to pass, while
Saul
talked unto the
priest, that the
noise that
{\ADD{was}} in the
host of the
Philistines
went
on and
increased: and
Saul
said unto the
priest,
Withdraw thine
hand.\FTNT{‡‡}{{\FR 14.19 }
noise: or, tumult}
\VS{20}And
Saul and all the
people that
{\ADD{were}} with him
assembled themselves, and they
came to the
battle: and, behold, every
man’s
sword was against his
fellow,
{\ADD{and there was}} a
very
great
discomfiture.\FTNT{§§}{{\FR 14.20 }
assembled…: Heb. were cried together}
\VS{21}Moreover the
Hebrews
{\ADD{that}} were with the
Philistines
before that
time, which went
up with them into the
camp
{\ADD{from the country}} round
about, even they also
{\ADD{turned}} to be with the
Israelites that
{\ADD{were}} with
Saul and
Jonathan.
\VS{22}Likewise all the
men of
Israel which had
hid themselves in
mount
Ephraim,
{\ADD{when}} they
heard that the
Philistines
fled, even they also followed
hard
after them in the
battle.
\VS{23}So the
{\ND{
LORD}}
saved
Israel that
day: and the
battle passed
over unto
Beth-aven.
\par }{\PP \VS{24}¶ And the
men of
Israel were
distressed that
day: for
Saul had
adjured the
people,
saying,
Cursed
{\ADD{be}} the
man that
eateth
{\ADD{any}}
food until
evening, that I may be
avenged on mine
enemies. So none of the
people
tasted
{\ADD{any}}
food.
\VS{25}And all
{\ADD{they of}} the
land
came to a
wood; and there was
honey
upon the
ground.
\VS{26}And when the
people were
come into the
wood, behold, the
honey
dropped; but no man
put his
hand to his
mouth: for the
people
feared the
oath.
\VS{27}But
Jonathan
heard not when his
father
charged the
people with the
oath: wherefore he put
forth the
end of the
rod that
{\ADD{was}} in his
hand, and
dipped it in an
honeycomb, and
put his
hand to his
mouth; and his
eyes were
enlightened.
\VS{28}Then
answered
one of the
people, and
said, Thy
father
straitly
charged the
people with an
oath,
saying,
Cursed
{\ADD{be}} the
man that
eateth
{\ADD{any}}
food this
day. And the
people were
faint.\FTNT{***}{{\FR 14.28 }
faint: or, weary}
\VS{29}Then
said
Jonathan, My
father hath
troubled the
land:
see, I pray you, how mine
eyes have been
enlightened, because I
tasted a
little of this
honey.
\VS{30}How much
more, if
haply the
people had
eaten
freely to
day of the
spoil of their
enemies which they
found? for had there not been now a much
greater
slaughter among the
Philistines?
\VS{31}And they
smote the
Philistines that
day from
Michmash to
Aijalon: and the
people were
very
faint.
\VS{32}And the
people
flew upon the
spoil, and
took
sheep, and
oxen, and
calves, and
slew
{\ADD{them}} on the
ground: and the
people did
eat
{\ADD{them}} with the
blood.
\par }{\PP \VS{33}¶ Then they
told
Saul,
saying, Behold, the
people
sin against the
{\ND{
LORD}}, in that they
eat with the
blood. And he
said, Ye have
transgressed:
roll a
great
stone unto me this
day.\FTNT{†††}{{\FR 14.33 }
transgressed: or, dealt treacherously}
\VS{34}And
Saul
said,
Disperse yourselves among the
people, and
say unto them, Bring me
hither every
man his
ox, and every
man his
sheep, and
slay
{\ADD{them}} here, and
eat; and
sin not against the
{\ND{
LORD}} in
eating with the
blood. And all the
people
brought every
man his
ox with
him that
night, and
slew
{\ADD{them}} there.\FTNT{‡‡‡}{{\FR 14.34 }
with him: Heb. in his hand}
\VS{35}And
Saul
built an
altar unto the
{\ND{
LORD}}: the same was the
first
altar that he
built unto the
{\ND{
LORD}}.\FTNT{§§§}{{\FR 14.35 }
the same…: Heb. that altar he began to build unto the
{\ND{LORD}}}
\par }{\PP \VS{36}¶ And
Saul
said, Let us go
down
after the
Philistines by
night, and
spoil them until the
morning
light, and let us not
leave a
man of them. And they
said,
Do whatsoever
seemeth
good unto thee. Then
said the
priest, Let us draw
near
hither unto
God.
\VS{37}And
Saul
asked counsel of
God, Shall I go
down
after the
Philistines? wilt thou
deliver them into the
hand of
Israel? But he
answered him not that
day.
\VS{38}And
Saul
said, Draw ye
near
hither, all the
chief of the
people: and
know and
see
wherein this
sin hath been this
day.\FTNT{*}{{\FR 14.38 }
chief: Heb. corners}
\VS{39}For,
{\ADD{as}} the
{\ND{
LORD}}
liveth, which
saveth
Israel, though it
be in
Jonathan my
son, he shall
surely
die. But
{\ADD{there was}} not a man among all the
people
{\ADD{that}}
answered him.
\VS{40}Then
said he unto all
Israel, Be ye on
one
side, and I and
Jonathan my
son will be on the
other
side. And the
people
said unto
Saul,
Do what
seemeth
good unto thee.
\VS{41}Therefore
Saul
said unto the
{\ND{
LORD}}
God of
Israel,
Give a
perfect
{\ADD{lot}}. And
Saul and
Jonathan were
taken: but the
people
escaped.\FTNT{†}{{\FR 14.41 }
Give…: or, Shew the innocent}\FTNT{‡}{{\FR 14.41 }
escaped: Heb. went forth}
\VS{42}And
Saul
said,
Cast
{\ADD{lots}} between me and
Jonathan my
son. And
Jonathan was
taken.
\VS{43}Then
Saul
said to
Jonathan,
Tell me what thou hast
done. And
Jonathan
told him, and
said, I did
but
taste a
little
honey with the
end of the
rod that
{\ADD{was}} in mine
hand,
{\ADD{and}},
lo, I must
die.
\VS{44}And
Saul
answered,
God do
so and more
also: for thou shalt
surely
die,
Jonathan.
\VS{45}And the
people
said unto
Saul, Shall
Jonathan
die, who hath
wrought this
great
salvation in
Israel? God
forbid:
{\ADD{as}} the
{\ND{
LORD}}
liveth, there shall not one
hair of his
head
fall to the
ground; for he hath
wrought with
God this
day. So the
people
rescued
Jonathan, that he
died not.
\VS{46}Then
Saul went
up from
following the
Philistines: and the
Philistines
went to their own
place.
\par }{\PP \VS{47}¶ So
Saul
took the
kingdom over
Israel, and
fought against all his
enemies on every
side, against
Moab, and against the
children of
Ammon, and against
Edom, and against the
kings of
Zobah, and against the
Philistines: and whithersoever he
turned himself, he
vexed
{\ADD{them}}.
\VS{48}And he
gathered an
host, and
smote the
Amalekites, and
delivered
Israel out of the
hands of them that
spoiled them.\FTNT{§}{{\FR 14.48 }
gathered…: or, wrought mightily}
\VS{49}Now the
sons of
Saul were
Jonathan, and
Ishui, and
Melchi-shua: and the
names of his
two
daughters
{\ADD{were these;}} the
name of the
firstborn
Merab, and the
name of the
younger
Michal:
\VS{50}And the
name of
Saul’s
wife
{\ADD{was}}
Ahinoam, the
daughter of
Ahimaaz: and the
name of the
captain of his
host
{\ADD{was}}
Abner, the
son of
Ner,
Saul’s
uncle.\FTNT{**}{{\FR 14.50 }
Abner: Heb. Abiner}
\VS{51}And
Kish
{\ADD{was}} the
father of
Saul; and
Ner the
father of
Abner
{\ADD{was}} the
son of
Abiel.
\VS{52}And there was
sore
war against the
Philistines all the
days of
Saul: and when
Saul
saw any
strong
man, or any
valiant
man, he
took him unto him.

\par }\Chap{15}{\PP \VerseOne{1}Samuel also
said unto
Saul, The
{\ND{
LORD}}
sent me to
anoint thee
{\ADD{to be}}
king over his
people, over
Israel: now therefore
hearken thou unto the
voice of the
words of the
{\ND{
LORD}}.
\VS{2}Thus
saith the
{\ND{
LORD}} of
hosts, I
remember
{\ADD{that}} which
Amalek
did to
Israel, how he
laid
{\ADD{wait}} for him in the
way, when he came
up from
Egypt.
\VS{3}Now
go and
smite
Amalek, and utterly
destroy all that they have, and
spare them not; but
slay both
man and
woman,
infant and
suckling,
ox and
sheep,
camel and
ass.
\VS{4}And
Saul
gathered the
people
together, and
numbered them in
Telaim, two
hundred
thousand
footmen, and
ten
thousand
men of
Judah.
\VS{5}And
Saul
came to a
city of
Amalek, and laid
wait in the
valley.\FTNT{*}{{\FR 15.5 }
laid…: or, fought}
\par }{\PP \VS{6}¶ And
Saul
said unto the
Kenites,
Go,
depart, get you
down from
among the
Amalekites, lest I
destroy you with them: for ye
shewed
kindness to all the
children of
Israel, when they came
up out of
Egypt. So the
Kenites
departed from
among the
Amalekites.
\VS{7}And
Saul
smote the
Amalekites from
Havilah
{\ADD{until}} thou
comest to
Shur, that
{\ADD{is}} over
against
Egypt.
\VS{8}And he
took
Agag the
king of the
Amalekites
alive, and utterly
destroyed all the
people with the
edge of the
sword.
\VS{9}But
Saul and the
people
spared
Agag, and the
best of the
sheep, and of the
oxen, and of the
fatlings, and the
lambs, and all
{\ADD{that was}}
good, and
would not utterly
destroy them: but every
thing
{\ADD{that was}}
vile and
refuse, that they destroyed
utterly.\FTNT{†}{{\FR 15.9 }
fatlings: or, second sort}
\par }{\PP \VS{10}¶ Then came the
word of the
{\ND{
LORD}} unto
Samuel,
saying,
\VS{11}It
repenteth me that I have set
up
Saul
{\ADD{to be}}
king: for he is turned
back from
following me, and hath not
performed my
commandments. And it
grieved
Samuel; and he
cried unto the
{\ND{
LORD}} all
night.
\VS{12}And when
Samuel rose
early to
meet
Saul in the
morning, it was
told
Samuel,
saying,
Saul
came to
Carmel, and, behold, he set him
up a
place, and is gone
about, and passed
on, and gone
down to
Gilgal.
\VS{13}And
Samuel
came to
Saul: and
Saul
said unto him,
Blessed
{\ADD{be}} thou of the
{\ND{
LORD}}: I have
performed the
commandment of the
{\ND{
LORD}}.
\VS{14}And
Samuel
said, What
{\ADD{meaneth}} then this
bleating of the
sheep in mine
ears, and the
lowing of the
oxen which I
hear?
\VS{15}And
Saul
said, They have
brought them from the
Amalekites: for the
people
spared the
best of the
sheep and of the
oxen, to
sacrifice unto the
{\ND{
LORD}} thy
God; and the
rest we have utterly
destroyed.
\VS{16}Then
Samuel
said unto
Saul,
Stay, and I will
tell thee what the
{\ND{
LORD}} hath
said to me this
night. And he
said unto him, Say
on.
\VS{17}And
Samuel
said, When thou
{\ADD{wast}}
little in thine own
sight,
{\ADD{wast}} thou not
{\ADD{made}} the
head of the
tribes of
Israel, and the
{\ND{
LORD}}
anointed thee
king over
Israel?
\VS{18}And the
{\ND{
LORD}}
sent thee on a
journey, and
said,
Go and utterly
destroy the
sinners the
Amalekites, and
fight against them until they be
consumed.\FTNT{‡}{{\FR 15.18 }
they…: Heb. they consume them}
\VS{19}Wherefore then didst thou not
obey the
voice of the
{\ND{
LORD}}, but didst
fly upon the
spoil, and
didst
evil in the
sight of the
{\ND{
LORD}}?
\VS{20}And
Saul
said unto
Samuel, Yea, I have
obeyed the
voice of the
{\ND{
LORD}}, and have
gone the
way which the
{\ND{
LORD}}
sent me, and have
brought
Agag the
king of
Amalek, and have utterly
destroyed the
Amalekites.
\VS{21}But the
people
took of the
spoil,
sheep and
oxen, the
chief of the things which should have been utterly
destroyed, to
sacrifice unto the
{\ND{
LORD}} thy
God in
Gilgal.
\VS{22}And
Samuel
said, Hath the
{\ND{
LORD}}
{\ADD{as great}}
delight in burnt
offerings and
sacrifices, as in
obeying the
voice of the
{\ND{
LORD}}? Behold, to
obey
{\ADD{is}}
better than
sacrifice,
{\ADD{and}} to
hearken than the
fat of
rams.
\VS{23}For
rebellion
{\ADD{is as}} the
sin of
witchcraft, and
stubbornness
{\ADD{is as}}
iniquity and
idolatry. Because thou hast
rejected the
word of the
{\ND{
LORD}}, he hath also
rejected thee from
{\ADD{being}}
king.\FTNT{§}{{\FR 15.23 }
witchcraft: Heb. divination}
\par }{\PP \VS{24}¶ And
Saul
said unto
Samuel, I have
sinned: for I have
transgressed the
commandment of the
{\ND{
LORD}}, and thy
words: because I
feared the
people, and
obeyed their
voice.
\VS{25}Now therefore, I pray thee,
pardon my
sin, and turn
again with me, that I may
worship the
{\ND{
LORD}}.
\VS{26}And
Samuel
said unto
Saul, I will not
return with thee: for thou hast
rejected the
word of the
{\ND{
LORD}}, and the
{\ND{
LORD}} hath
rejected thee from being
king over
Israel.
\VS{27}And as
Samuel turned
about to go
away, he laid
hold upon the
skirt of his
mantle, and it
rent.
\VS{28}And
Samuel
said unto him, The
{\ND{
LORD}} hath
rent the
kingdom of
Israel from thee this
day, and hath
given it to a
neighbour of thine,
{\ADD{that is}}
better than thou.
\VS{29}And also the
Strength of
Israel will not
lie nor
repent: for he
{\ADD{is}} not a
man, that he should
repent.\FTNT{**}{{\FR 15.29 }
Strength: or, Eternity, or, Victory}
\VS{30}Then he
said, I have
sinned:
{\ADD{yet}}
honour me now, I pray thee, before the
elders of my
people, and before
Israel, and turn
again with me, that I may
worship the
{\ND{
LORD}} thy
God.
\VS{31}So
Samuel turned
again
after
Saul; and
Saul
worshipped the
{\ND{
LORD}}.
\par }{\PP \VS{32}¶ Then
said
Samuel, Bring ye
hither to me
Agag the
king of the
Amalekites. And
Agag
came unto him
delicately. And
Agag
said,
Surely the
bitterness of
death is
past.
\VS{33}And
Samuel
said, As thy
sword hath made
women
childless, so shall thy
mother be
childless among
women. And
Samuel
hewed
Agag in
pieces
before the
{\ND{
LORD}} in
Gilgal.
\par }{\PP \VS{34}¶ Then
Samuel
went to
Ramah; and
Saul went
up to his
house to
Gibeah of
Saul.
\VS{35}And
Samuel came no
more to
see
Saul until the
day of his
death:
nevertheless
Samuel
mourned for
Saul: and the
{\ND{
LORD}}
repented that he had made
Saul
king over
Israel.

\par }\Chap{16}{\PP \VerseOne{1}And the
{\ND{
LORD}}
said unto
Samuel, How long wilt thou
mourn for
Saul, seeing I have
rejected him from
reigning over
Israel?
fill thine
horn with
oil, and
go, I will
send thee to
Jesse the
Beth-lehemite: for I have
provided me a
king among his
sons.
\VS{2}And
Samuel
said, How can I
go? if
Saul
hear
{\ADD{it}}, he will
kill me. And the
{\ND{
LORD}}
said,
Take an
heifer with
thee, and
say, I am
come to
sacrifice to the
{\ND{
LORD}}.\FTNT{*}{{\FR 16.2 }
with thee: Heb. in thine hand}
\VS{3}And
call
Jesse to the
sacrifice, and I will
shew thee what thou shalt
do: and thou shalt
anoint unto me
{\ADD{him}} whom I
name unto thee.
\VS{4}And
Samuel
did that which the
{\ND{
LORD}}
spake, and
came to
Beth-lehem. And the
elders of the
town
trembled at his
coming, and
said,
Comest thou
peaceably?\FTNT{†}{{\FR 16.4 }
coming: Heb. meeting}
\VS{5}And he
said,
Peaceably: I am
come to
sacrifice unto the
{\ND{
LORD}}:
sanctify yourselves, and
come with me to the
sacrifice. And he
sanctified
Jesse and his
sons, and
called them to the
sacrifice.
\par }{\PP \VS{6}¶ And it came to pass, when they were
come, that he
looked on
Eliab, and
said, Surely the
{\ND{
LORD’s}}
anointed
{\ADD{is}} before him.\FTNT{‡}{{\FR 16.6 }
Eliab: called Elihu}
\VS{7}But the
{\ND{
LORD}}
said unto
Samuel,
Look not on his
countenance, or on the
height of his
stature; because I have
refused him: for
{\ADD{the
{\ND{Lord}} seeth}} not as
man
seeth; for
man
looketh on the outward
appearance, but the
{\ND{
LORD}}
looketh on the
heart.\FTNT{§}{{\FR 16.7 }
outward…: Heb. eyes}
\VS{8}Then
Jesse
called
Abinadab, and made him
pass
before
Samuel. And he
said, Neither hath the
{\ND{
LORD}}
chosen this.
\VS{9}Then
Jesse made
Shammah to pass
by. And he
said, Neither hath the
{\ND{
LORD}}
chosen this.\FTNT{**}{{\FR 16.9 }
Shammah: Shimeah, also called, Shimma}
\VS{10}Again,
Jesse made
seven of his
sons to
pass
before
Samuel. And
Samuel
said unto
Jesse, The
{\ND{
LORD}} hath not
chosen these.
\VS{11}And
Samuel
said unto
Jesse, Are here
all
{\ADD{thy}}
children? And he
said, There
remaineth yet the
youngest, and, behold, he
keepeth the
sheep. And
Samuel
said unto
Jesse,
Send and
fetch him: for we will not sit
down till he
come
hither.\FTNT{††}{{\FR 16.11 }
down: Heb. round}
\VS{12}And he
sent, and brought him
in. Now he
{\ADD{was}}
ruddy,
{\ADD{and}}
withal of a
beautiful
countenance, and
goodly to look
to. And the
{\ND{
LORD}}
said,
Arise,
anoint him: for this
{\ADD{is}} he.\FTNT{‡‡}{{\FR 16.12 }
of a…: Heb. fair of eyes}
\VS{13}Then
Samuel
took the
horn of
oil, and
anointed him in the
midst of his
brethren: and the
Spirit of the
{\ND{
LORD}}
came upon
David from that
day
forward. So
Samuel rose
up, and
went to
Ramah.
\par }{\PP \VS{14}¶ But the
Spirit of the
{\ND{
LORD}}
departed from
Saul, and an
evil
spirit from the
{\ND{
LORD}}
troubled him.\FTNT{§§}{{\FR 16.14 }
troubled: or, terrified}
\VS{15}And
Saul’s
servants
said unto him, Behold now, an
evil
spirit from
God
troubleth thee.
\VS{16}Let our
lord now
command thy
servants,
{\ADD{which are}}
before thee, to seek
out a
man,
{\ADD{who is}} a
cunning
player on an
harp: and it shall come to pass, when the
evil
spirit from
God is upon thee, that he shall
play with his
hand, and thou shalt be
well.
\VS{17}And
Saul
said unto his
servants,
Provide me now a
man that can
play
well, and
bring
{\ADD{him}} to me.
\VS{18}Then
answered
one of the
servants, and
said, Behold, I have
seen a
son of
Jesse the
Beth-lehemite,
{\ADD{that is}}
cunning in
playing, and a
mighty valiant
man, and a
man of
war, and
prudent in
matters, and a
comely
person, and the
{\ND{
LORD}}
{\ADD{is}} with him.\FTNT{***}{{\FR 16.18 }
matters: or, speech}
\par }{\PP \VS{19}¶ Wherefore
Saul
sent
messengers unto
Jesse, and
said,
Send me
David thy
son, which
{\ADD{is}} with the
sheep.
\VS{20}And
Jesse
took an
ass
{\ADD{laden}} with
bread, and a
bottle of
wine, and
a
kid, and
sent
{\ADD{them}}
by
David his
son unto
Saul.
\VS{21}And
David
came to
Saul, and
stood
before him: and he
loved him
greatly; and he became his
armourbearer.
\VS{22}And
Saul
sent to
Jesse,
saying, Let
David, I pray thee,
stand
before me; for he hath
found
favour in my
sight.
\VS{23}And it came to pass, when the
{\ADD{evil}}
spirit from
God was upon
Saul, that
David
took an
harp, and
played with his
hand: so
Saul was
refreshed, and was
well, and the
evil
spirit
departed from him.

\par }\Chap{17}{\PP \VerseOne{1}Now the
Philistines gathered
together their
armies to
battle, and were gathered
together at
Shochoh, which
{\ADD{belongeth}} to
Judah, and
pitched between
Shochoh and
Azekah, in
Ephes-dammim.\FTNT{*}{{\FR 17.1 }
Ephes-dammim: or, the coast of Dammim, called Pas-dammim}
\VS{2}And
Saul and the
men of
Israel were gathered
together, and
pitched by the
valley of
Elah, and set the
battle in
array
against the
Philistines.\FTNT{†}{{\FR 17.2 }
set…: Heb. ranged the battle}
\VS{3}And the
Philistines
stood on a
mountain on the one side, and
Israel
stood on a
mountain on the other side: and
{\ADD{there was}} a
valley between them.
\par }{\PP \VS{4}¶ And there went
out a
champion out of the
camp of the
Philistines,
named
Goliath, of
Gath, whose
height
{\ADD{was}}
six
cubits and a
span.
\VS{5}And
{\ADD{he had}} an
helmet of
brass upon his
head, and he
{\ADD{was}}
armed with a
coat of
mail; and the
weight of the
coat
{\ADD{was}}
five
thousand
shekels of
brass.\FTNT{‡}{{\FR 17.5 }
armed: Heb. clothed}
\VS{6}And
{\ADD{he had}}
greaves of
brass upon his
legs, and a
target of
brass between his
shoulders.\FTNT{§}{{\FR 17.6 }
target: or, gorget}
\VS{7}And the
staff of his
spear
{\ADD{was}} like a
weaver’s
beam; and his
spear’s
head
{\ADD{weighed}}
six
hundred
shekels of
iron: and one
bearing a
shield
went
before him.
\VS{8}And he
stood and
cried unto the
armies of
Israel, and
said unto them, Why are ye come
out to set
{\ADD{your}}
battle in
array?
{\ADD{am}} not I a
Philistine, and ye
servants to
Saul?
choose you a
man for you, and let him come
down to me.
\VS{9}If he be
able to
fight with me, and to
kill me, then will we be your
servants: but if I
prevail against him, and
kill him, then shall ye be our
servants, and
serve us.
\VS{10}And the
Philistine
said, I
defy the
armies of
Israel this
day;
give me a
man, that we may
fight
together.
\VS{11}When
Saul and all
Israel
heard those
words of the
Philistine, they were
dismayed, and
greatly
afraid.
\par }{\PP \VS{12}¶ Now
David
{\ADD{was}} the
son of that
Ephrathite of
Beth-lehem-judah, whose
name
{\ADD{was}}
Jesse; and he had
eight
sons: and the
man
went among
men
{\ADD{for}} an old
man in the
days of
Saul.
\VS{13}And the
three
eldest
sons of
Jesse
went
{\ADD{and}}
followed
Saul to the
battle: and the
names of his
three
sons that
went to the
battle
{\ADD{were}}
Eliab the
firstborn, and
next unto him
Abinadab, and the
third
Shammah.
\VS{14}And
David
{\ADD{was}} the
youngest: and the
three
eldest
followed
Saul.
\VS{15}But
David
went and
returned from
Saul to
feed his
father’s
sheep at
Beth-lehem.
\VS{16}And the
Philistine drew
near
morning and
evening, and
presented himself
forty
days.
\VS{17}And
Jesse
said unto
David his
son,
Take now for thy
brethren an
ephah of
this
parched
{\ADD{corn}}, and these
ten
loaves, and
run to the
camp to thy
brethren;
\VS{18}And
carry these
ten
cheeses unto the
captain of
{\ADD{their}}
thousand, and
look how thy
brethren
fare, and
take their
pledge.\FTNT{**}{{\FR 17.18 }
cheeses: Heb. cheeses of milk}\FTNT{††}{{\FR 17.18 }
of…: Heb. of a thousand}
\VS{19}Now
Saul, and they, and all the
men of
Israel,
{\ADD{were}} in the
valley of
Elah,
fighting with the
Philistines.
\par }{\PP \VS{20}¶ And
David rose up
early in the
morning, and
left the
sheep with a
keeper, and
took, and
went, as
Jesse had
commanded him; and he
came to the
trench, as the
host was going
forth to the
fight, and
shouted for the
battle.\FTNT{‡‡}{{\FR 17.20 }
trench: or, place of the carriage}\FTNT{§§}{{\FR 17.20 }
fight: or, battle array, or, place of fight}
\VS{21}For
Israel and the
Philistines had put the battle in
array,
army
against
army.
\VS{22}And
David
left his
carriage in the
hand of the
keeper of the
carriage, and
ran into the
army, and
came and
saluted his
brethren.\FTNT{***}{{\FR 17.22 }
his carriage: Heb. the vessels from upon him}\FTNT{†††}{{\FR 17.22 }
saluted…: Heb. asked his brethren of peace}
\VS{23}And as he
talked with them, behold, there came
up the
champion, the
Philistine of
Gath,
Goliath by
name, out of the
armies of the
Philistines, and
spake according to the
same
words: and
David
heard
{\ADD{them}}.
\VS{24}And all the
men of
Israel, when they
saw the
man,
fled from
him, and were
sore
afraid.\FTNT{‡‡‡}{{\FR 17.24 }
from…: Heb. from his face}
\VS{25}And the
men of
Israel
said, Have ye
seen this
man that is come
up? surely to
defy
Israel is he come
up: and it shall be,
{\ADD{that}} the
man who
killeth him, the
king will
enrich him with
great
riches, and will
give him his
daughter, and
make his
father’s
house
free in
Israel.
\VS{26}And
David
spake to the
men that
stood by him,
saying, What shall be
done to the
man that
killeth
this
Philistine, and taketh
away the
reproach from
Israel? for who
{\ADD{is}} this
uncircumcised
Philistine, that he should
defy the
armies of the
living
God?
\VS{27}And the
people
answered him after this
manner,
saying, So shall it be
done to the
man that
killeth him.
\par }{\PP \VS{28}¶ And
Eliab his
eldest
brother
heard when he
spake unto the
men; and
Eliab’s
anger was
kindled against
David, and he
said, Why camest thou
down hither? and with whom hast thou
left
those
few
sheep in the
wilderness? I
know thy
pride, and the
naughtiness of thine
heart; for thou art come
down that thou mightest
see the
battle.
\VS{29}And
David
said, What have I now
done?
{\ADD{Is there}} not a
cause?
\par }{\PP \VS{30}¶ And he
turned from
him
toward
another, and
spake after the same
manner: and the
people
answered him
again after the
former
manner.\FTNT{§§§}{{\FR 17.30 }
manner: Heb. word}
\par }{\PP \VS{31}And when the
words were
heard which
David
spake, they
rehearsed
{\ADD{them}}
before
Saul: and he
sent for him.\FTNT{*}{{\FR 17.31 }
sent…: Heb. took him}
\par }{\PP \VS{32}¶ And
David
said to
Saul, Let no
man’s
heart
fail because of him; thy
servant will
go and
fight with this
Philistine.
\VS{33}And
Saul
said to
David, Thou art not
able to
go against this
Philistine to
fight with him: for thou
{\ADD{art but}} a
youth, and he a
man of
war from his
youth.
\VS{34}And
David
said unto
Saul, Thy
servant
kept his
father’s
sheep, and there
came a
lion, and a
bear, and
took a
lamb out of the
flock:\FTNT{†}{{\FR 17.34 }
lamb: or, kid}
\VS{35}And I went
out
after him, and
smote him, and
delivered
{\ADD{it}} out of his
mouth: and when he
arose against me, I
caught
{\ADD{him}} by his
beard, and
smote him, and
slew him.
\VS{36}Thy
servant
slew both the
lion and the
bear: and this
uncircumcised
Philistine shall be as
one of them, seeing he hath
defied the
armies of the
living
God.
\VS{37}David
said moreover, The
{\ND{
LORD}} that
delivered me out of the
paw of the
lion, and out of the
paw of the
bear, he will
deliver me out of the
hand of this
Philistine. And
Saul
said unto
David,
Go, and the
{\ND{
LORD}} be with thee.
\par }{\PP \VS{38}¶ And
Saul
armed
David with his
armour, and he
put an
helmet of
brass upon his
head; also he
armed him with a coat of
mail.\FTNT{‡}{{\FR 17.38 }
armed David…: Heb. clothed David with his clothes}
\VS{39}And
David
girded his
sword upon his
armour, and he
assayed to
go; for he had not
proved
{\ADD{it}}. And
David
said unto
Saul, I
cannot
go with these; for I have not
proved
{\ADD{them}}. And
David
put them off him.
\par }{\PP \VS{40}And he
took his
staff in his
hand, and
chose him
five
smooth
stones out of the
brook, and
put them in a
shepherd’s
bag which he had, even in a
scrip; and his
sling
{\ADD{was}} in his
hand: and he drew
near to the
Philistine.\FTNT{§}{{\FR 17.40 }
brook: or, valley}\FTNT{**}{{\FR 17.40 }
bag: Heb. vessel}
\VS{41}And the
Philistine
came
on and drew
near unto
David; and the
man that
bare the
shield
{\ADD{went}}
before him.
\VS{42}And when the
Philistine looked
about, and
saw
David, he
disdained him: for he was
{\ADD{but}} a
youth, and
ruddy, and of a
fair
countenance.
\VS{43}And the
Philistine
said unto
David,
{\ADD{Am}} I a
dog, that thou
comest to me with
staves? And the
Philistine
cursed
David by his
gods.
\VS{44}And the
Philistine
said to
David,
Come to me, and I will
give thy
flesh unto the
fowls of the
air, and to the
beasts of the
field.
\VS{45}Then
said
David to the
Philistine, Thou
comest to me with a
sword, and with a
spear, and with a
shield: but I
come to thee in the
name of the
{\ND{
LORD}} of
hosts, the
God of the
armies of
Israel, whom thou hast
defied.
\VS{46}This
day will the
{\ND{
LORD}}
deliver thee into mine
hand; and I will
smite thee, and
take thine
head from thee; and I will
give the
carcases of the
host of the
Philistines this
day unto the
fowls of the
air, and to the wild
beasts of the
earth; that all the
earth may
know that there
is a
God in
Israel.\FTNT{††}{{\FR 17.46 }
deliver…: Heb. shut thee up}
\VS{47}And all this
assembly shall
know that the
{\ND{
LORD}}
saveth not with
sword and
spear: for the
battle
{\ADD{is}} the
{\ND{
LORD’s}}, and he will
give you into our
hands.
\par }{\PP \VS{48}And it came to pass, when the
Philistine
arose, and
came and drew
nigh to
meet
David, that
David
hasted, and
ran toward the
army to
meet the
Philistine.
\VS{49}And
David
put his
hand in his
bag, and
took thence a
stone, and
slang
{\ADD{it}}, and
smote the
Philistine in his
forehead, that the
stone
sunk into his
forehead; and he
fell upon his
face to the
earth.
\VS{50}So
David
prevailed
over the
Philistine with a
sling and with a
stone, and
smote the
Philistine, and
slew him; but
{\ADD{there was}} no
sword in the
hand of
David.
\VS{51}Therefore
David
ran, and
stood upon the
Philistine, and
took his
sword, and
drew it out of the
sheath thereof, and
slew him, and cut
off his
head therewith. And when the
Philistines
saw their
champion was
dead, they
fled.
\VS{52}And the
men of
Israel and of
Judah
arose, and
shouted, and
pursued the
Philistines, until thou
come to the
valley, and to the
gates of
Ekron. And the
wounded of the
Philistines fell
down by the
way to
Shaaraim, even unto
Gath, and unto
Ekron.
\VS{53}And the
children of
Israel
returned from
chasing
after the
Philistines, and they
spoiled their
tents.
\VS{54}And
David
took the
head of the
Philistine, and
brought it to
Jerusalem; but he
put his
armour in his
tent.
\par }{\PP \VS{55}¶ And when
Saul
saw
David go
forth
against the
Philistine, he
said unto
Abner, the
captain of the
host,
Abner, whose
son
{\ADD{is}} this
youth? And
Abner
said,
{\ADD{As}} thy
soul
liveth, O
king, I
cannot
tell.
\VS{56}And the
king
said,
Enquire thou whose
son the
stripling
{\ADD{is}}.
\VS{57}And as
David
returned from the
slaughter of the
Philistine,
Abner
took him, and
brought him
before
Saul with the
head of the
Philistine in his
hand.
\VS{58}And
Saul
said to him, Whose
son
{\ADD{art}} thou,
{\ADD{thou}} young
man? And
David
answered, I
{\ADD{am}} the
son of thy
servant
Jesse the
Beth-lehemite.

\par }\Chap{18}{\PP \VerseOne{1}And it came to pass, when he had made an
end of
speaking unto
Saul, that the
soul of
Jonathan was
knit with the
soul of
David, and
Jonathan
loved him as his own
soul.
\VS{2}And
Saul
took him that
day, and would
let him go no more
home to his
father’s
house.
\VS{3}Then
Jonathan and
David
made a
covenant, because he
loved him as his own
soul.
\VS{4}And
Jonathan
stripped himself of the
robe that
{\ADD{was}} upon him, and
gave it to
David, and his
garments, even to his
sword, and to his
bow, and to his
girdle.
\par }{\PP \VS{5}¶ And
David went
out whithersoever
Saul
sent him,
{\ADD{and}} behaved himself
wisely: and
Saul
set him over the
men of
war, and he was
accepted in the
sight of all the
people, and also in the
sight of
Saul’s
servants.\FTNT{*}{{\FR 18.5 }
behaved…: or, prospered}
\par }{\PP \VS{6}And it came to pass as they
came, when
David was
returned from the
slaughter of the
Philistine, that the
women came
out of all
cities of
Israel,
singing and
dancing, to
meet
king
Saul, with
tabrets, with
joy, and with instruments of
musick.\FTNT{†}{{\FR 18.6 }
Philistine: or, Philistines}\FTNT{‡}{{\FR 18.6 }
instruments…: Heb. three stringed instruments}
\VS{7}And the
women
answered
{\ADD{one another}} as they
played, and
said,
Saul hath
slain his
thousands, and
David his ten
thousands.
\VS{8}And
Saul was
very
wroth, and the
saying
displeased him; and he
said, They have
ascribed unto
David ten
thousands, and to me they have
ascribed
{\ADD{but}}
thousands: and
{\ADD{what}} can he have more but the
kingdom?\FTNT{§}{{\FR 18.8 }
displeased him: Heb. was evil in his eyes}
\VS{9}And
Saul
eyed
David from that
day and
forward.
\par }{\PP \VS{10}¶ And it came to pass on the
morrow, that the
evil
spirit from
God
came upon
Saul, and he
prophesied in the
midst of the
house: and
David
played with his
hand, as at other
times: and
{\ADD{there was}} a
javelin in
Saul’s
hand.
\VS{11}And
Saul
cast the
javelin; for he
said, I will
smite
David even to the
wall
{\ADD{with it}}. And
David avoided
out of his
presence
twice.
\par }{\PP \VS{12}¶ And
Saul was
afraid
of
David, because the
{\ND{
LORD}} was with him, and was
departed from
Saul.
\VS{13}Therefore
Saul
removed him from him, and
made him his
captain over a
thousand; and he went
out and came
in
before the
people.
\VS{14}And
David behaved himself
wisely in all his
ways; and the
{\ND{
LORD}}
{\ADD{was}} with him.\FTNT{**}{{\FR 18.14 }
behaved…: or, prospered}
\VS{15}Wherefore when
Saul
saw that he
behaved himself
very
wisely, he was
afraid
of him.
\VS{16}But all
Israel and
Judah
loved
David, because he went
out and came
in
before them.
\par }{\PP \VS{17}¶ And
Saul
said to
David, Behold my
elder
daughter
Merab, her will I
give thee to
wife: only be thou
valiant for me, and
fight the
{\ND{
LORD’s}}
battles. For
Saul
said, Let not mine
hand be upon him, but let the
hand of the
Philistines be upon him.\FTNT{††}{{\FR 18.17 }
valiant: Heb. a son of valour}
\VS{18}And
David
said unto
Saul, Who
{\ADD{am}} I? and what
{\ADD{is}} my
life,
{\ADD{or}} my
father’s
family in
Israel, that I should be son in
law to the
king?
\VS{19}But it came to pass at the
time when
Merab
Saul’s
daughter should have been
given to
David, that she was
given unto
Adriel the
Meholathite to
wife.
\VS{20}And
Michal
Saul’s
daughter
loved
David: and they
told
Saul, and the
thing
pleased him.\FTNT{‡‡}{{\FR 18.20 }
pleased him: Heb. was right in his eyes}
\VS{21}And
Saul
said, I will
give him her, that she may be a
snare to him, and that the
hand of the
Philistines may be against him. Wherefore
Saul
said to
David, Thou shalt this
day be my son in
law in
{\ADD{the one of}} the
twain.
\par }{\PP \VS{22}¶ And
Saul
commanded his
servants,
{\ADD{saying}},
Commune with
David
secretly, and
say, Behold, the
king hath
delight in thee, and all his
servants
love thee: now therefore be the
king’s son in
law.
\VS{23}And
Saul’s
servants
spake those
words in the
ears of
David. And
David
said,
Seemeth it to you
{\ADD{a}}
light
{\ADD{thing}} to be a
king’s son in
law, seeing that I
{\ADD{am}} a
poor
man, and lightly
esteemed?
\VS{24}And the
servants of
Saul
told him,
saying, On this
manner
spake
David.\FTNT{§§}{{\FR 18.24 }
On…: Heb. According to these words}
\VS{25}And
Saul
said, Thus shall ye
say to
David, The
king
desireth not any
dowry, but an
hundred
foreskins of the
Philistines, to be
avenged of the
king’s
enemies. But
Saul
thought to make
David
fall by the
hand of the
Philistines.
\VS{26}And when his
servants
told
David these
words, it
pleased
David well to be the
king’s son in
law: and the
days were not
expired.\FTNT{***}{{\FR 18.26 }
expired: Heb. fulfilled}
\VS{27}Wherefore
David
arose and
went, he and his
men, and
slew of the
Philistines two
hundred
men; and
David
brought their
foreskins, and they gave them in full
tale to the
king, that he might be the
king’s son in
law. And
Saul
gave him
Michal his
daughter to
wife.
\par }{\PP \VS{28}¶ And
Saul
saw and
knew that the
{\ND{
LORD}}
{\ADD{was}} with
David, and
{\ADD{that}}
Michal
Saul’s
daughter
loved him.
\VS{29}And
Saul was yet the
more
afraid
of
David; and
Saul became
David’s
enemy
continually.
\VS{30}Then the
princes of the
Philistines went
forth: and it came to pass,
after they went
forth,
{\ADD{that}}
David behaved himself more
wisely than all the
servants of
Saul; so that his
name was
much set
by.\FTNT{†††}{{\FR 18.30 }
set by: Heb. precious}

\par }\Chap{19}{\PP \VerseOne{1}And
Saul
spake to
Jonathan his
son, and to all his
servants, that they should
kill
David.
\VS{2}But
Jonathan
Saul’s
son
delighted
much in
David: and
Jonathan
told
David,
saying,
Saul my
father
seeketh to
kill thee: now therefore, I pray thee, take
heed to thyself until the
morning, and
abide in a
secret
{\ADD{place}}, and
hide thyself:
\VS{3}And I will go
out and
stand
beside my
father in the
field where thou
{\ADD{art}}, and I will
commune with my
father of thee; and
what I
see, that I will
tell thee.
\par }{\PP \VS{4}¶ And
Jonathan
spake
good of
David unto
Saul his
father, and
said unto him, Let not the
king
sin against his
servant, against
David; because he hath not
sinned against thee, and because his
works
{\ADD{have been}} to
thee-ward very
good:
\VS{5}For he did
put his
life in his
hand, and
slew the
Philistine, and the
{\ND{
LORD}}
wrought a
great
salvation for all
Israel: thou
sawest
{\ADD{it}}, and didst
rejoice: wherefore then wilt thou
sin against
innocent
blood, to
slay
David without a
cause?
\VS{6}And
Saul
hearkened unto the
voice of
Jonathan: and
Saul
sware,
{\ADD{As}} the
{\ND{
LORD}}
liveth, he shall
not be
slain.
\VS{7}And
Jonathan
called
David, and
Jonathan
shewed him all those
things. And
Jonathan
brought
David to
Saul, and he was in his
presence, as in times
past.\FTNT{*}{{\FR 19.7 }
in times…: Heb. yesterday, third day}
\par }{\PP \VS{8}¶ And there was
war
again: and
David went
out, and
fought with the
Philistines, and
slew them with a
great
slaughter; and they
fled
from him.\FTNT{†}{{\FR 19.8 }
him: Heb. his face}
\VS{9}And the
evil
spirit from the
{\ND{
LORD}} was upon
Saul, as he
sat in his
house with his
javelin in his
hand: and
David
played with
{\ADD{his}}
hand.
\VS{10}And
Saul
sought to
smite
David even to the
wall with the
javelin; but he slipped
away out of
Saul’s
presence, and he
smote the
javelin into the
wall: and
David
fled, and
escaped that
night.
\par }{\PP \VS{11}Saul also
sent
messengers unto
David’s
house, to
watch him, and to
slay him in the
morning: and
Michal
David’s
wife
told him,
saying, If thou
save not thy
life to
night, to
morrow thou shalt be
slain.
\par }{\PP \VS{12}¶ So
Michal
let
David
down through a
window: and he
went, and
fled, and
escaped.
\VS{13}And
Michal
took an
image, and
laid
{\ADD{it}} in the
bed, and
put a
pillow of
goats’
{\ADD{hair}} for his
bolster, and
covered
{\ADD{it}} with a
cloth.\FTNT{‡}{{\FR 19.13 }
image: Heb. teraphim}
\VS{14}And when
Saul
sent
messengers to
take
David, she
said, He
{\ADD{is}}
sick.
\VS{15}And
Saul
sent the
messengers
{\ADD{again}} to
see
David,
saying, Bring him
up to me in the
bed, that I may
slay him.
\VS{16}And when the
messengers were come
in, behold,
{\ADD{there was}} an
image in the
bed, with a
pillow of
goats’
{\ADD{hair}} for his
bolster.
\VS{17}And
Saul
said unto
Michal, Why hast thou deceived me
so, and sent
away mine
enemy, that he is
escaped? And
Michal
answered
Saul, He
said unto me, Let me
go; why should I
kill thee?
\par }{\PP \VS{18}¶ So
David
fled, and
escaped, and
came to
Samuel to
Ramah, and
told him all that
Saul had
done to him. And he and
Samuel
went and
dwelt in
Naioth.
\VS{19}And it was
told
Saul,
saying, Behold,
David
{\ADD{is}} at
Naioth in
Ramah.
\VS{20}And
Saul
sent
messengers to
take
David: and when they
saw the
company of the
prophets
prophesying, and
Samuel
standing
{\ADD{as}}
appointed over them, the
Spirit of
God was upon the
messengers of
Saul, and they also
prophesied.
\VS{21}And when it was
told
Saul, he
sent
other
messengers, and they
prophesied likewise. And
Saul
sent
messengers
again the
third time, and they
prophesied also.
\VS{22}Then
went he also to
Ramah, and
came to a
great
well that
{\ADD{is}} in
Sechu: and he
asked and
said,
Where
{\ADD{are}}
Samuel and
David? And
{\ADD{one}}
said, Behold,
{\ADD{they be}} at
Naioth in
Ramah.
\VS{23}And he
went thither to
Naioth in
Ramah: and the
Spirit of
God was upon him also, and he
went
on, and
prophesied, until he
came to
Naioth in
Ramah.
\VS{24}And he stripped
off his
clothes
also, and
prophesied
before
Samuel in like manner, and lay
down
naked all that
day and all that
night. Wherefore they
say,
{\ADD{Is}}
Saul also among the
prophets?\FTNT{§}{{\FR 19.24 }
lay: Heb. fell}

\par }\Chap{20}{\PP \VerseOne{1}And
David
fled from
Naioth in
Ramah, and
came and
said
before
Jonathan, What have I
done? what
{\ADD{is}} mine
iniquity? and what
{\ADD{is}} my
sin
before thy
father, that he
seeketh my
life?
\VS{2}And he
said unto him, God
forbid; thou shalt not
die: behold, my
father will
do
nothing either
great or
small, but that he will
shew it me: and why should my
father
hide this
thing from me? it
{\ADD{is}} not
{\ADD{so}}.\FTNT{*}{{\FR 20.2 }
shew…: Heb. uncover mine ear}
\VS{3}And
David
sware moreover, and
said, Thy
father
certainly
knoweth that I have
found
grace in thine
eyes; and he
saith, Let not
Jonathan
know this, lest he be
grieved: but
truly
{\ADD{as}} the
{\ND{
LORD}}
liveth, and
{\ADD{as}} thy
soul
liveth,
{\ADD{there is}} but a
step between me and
death.
\VS{4}Then
said
Jonathan unto
David, Whatsoever thy
soul
desireth, I will even
do
{\ADD{it}} for thee.\FTNT{†}{{\FR 20.4 }
Whatsoever…: or, Say what is thy mind and I will do, etc}\FTNT{‡}{{\FR 20.4 }
desireth: Heb. speaketh, or, thinketh}
\VS{5}And
David
said unto
Jonathan, Behold, to
morrow
{\ADD{is}} the new
moon, and I should not
fail to
sit with the
king at
meat: but let me
go, that I may
hide myself in the
field unto the
third
{\ADD{day}} at
even.
\VS{6}If thy
father at
all
miss me, then
say,
David
earnestly
asked
{\ADD{leave}} of me that he might
run to
Beth-lehem his
city: for
{\ADD{there is}} a
yearly
sacrifice there for all the
family.\FTNT{§}{{\FR 20.6 }
sacrifice: or, feast}
\VS{7}If he
say thus,
{\ADD{It is}}
well; thy
servant shall have
peace: but if he be
very
wroth,
{\ADD{then}} be
sure that
evil is
determined by him.
\VS{8}Therefore thou shalt
deal
kindly with thy
servant; for thou hast
brought thy
servant into a
covenant of the
{\ND{
LORD}} with thee: notwithstanding, if there
be in me
iniquity,
slay me thyself; for why shouldest thou
bring me to thy
father?
\par }{\PP \VS{9}And
Jonathan
said, Far be it from
thee: for if I
knew
certainly that
evil were
determined by my
father to
come upon thee, then would not I
tell it thee?
\VS{10}Then
said
David to
Jonathan, Who shall
tell me? or what
{\ADD{if}} thy
father
answer thee
roughly?
\par }{\PP \VS{11}¶ And
Jonathan
said unto
David,
Come, and let us go
out into the
field. And they went
out
both of them into the
field.
\VS{12}And
Jonathan
said unto
David, O
{\ND{
LORD}}
God of
Israel, when I have
sounded my
father about to
morrow any
time,
{\ADD{or}} the
third
{\ADD{day}}, and, behold,
{\ADD{if there be}}
good toward
David, and I then
send not unto thee, and
shew it thee;\FTNT{**}{{\FR 20.12 }
sounded: Heb. searched}\FTNT{††}{{\FR 20.12 }
shew…: Heb. uncover thine ear}
\VS{13}The
{\ND{
LORD}}
do
so and much
more to
Jonathan: but if it
please my
father
{\ADD{to do}} thee
evil, then I will
shew it thee, and send thee
away, that thou mayest
go in
peace: and the
{\ND{
LORD}} be with thee, as he hath been with my
father.\FTNT{‡‡}{{\FR 20.13 }
shew…: Heb. uncover thine ear}
\VS{14}And thou shalt not only while
yet I
live
shew me the
kindness of the
{\ND{
LORD}}, that I
die not:
\VS{15}But
{\ADD{also}} thou shalt not cut
off thy
kindness from my
house
for
ever: no, not when the
{\ND{
LORD}} hath cut
off the
enemies of
David every
one from the
face of the
earth.
\VS{16}So
Jonathan
made
{\ADD{a covenant}} with the
house of
David,
{\ADD{saying}}, Let the
{\ND{
LORD}} even
require
{\ADD{it}} at the
hand of
David’s
enemies.\FTNT{§§}{{\FR 20.16 }
made: Heb. cut}
\VS{17}And
Jonathan caused
David to
swear
again, because he
loved him: for he
loved him as he
loved his own
soul.\FTNT{***}{{\FR 20.17 }
because…: or, by his love toward him}
\VS{18}Then
Jonathan
said to
David, To
morrow
{\ADD{is}} the new
moon: and thou shalt be
missed, because thy
seat will be
empty.\FTNT{†††}{{\FR 20.18 }
empty: Heb. missed}
\VS{19}And
{\ADD{when}} thou hast stayed three
days,
{\ADD{then}} thou shalt go
down
quickly, and
come to the
place where thou didst
hide thyself
when the
business was
{\ADD{in hand}}, and shalt
remain
by the
stone
Ezel.\FTNT{‡‡‡}{{\FR 20.19 }
quickly: or, diligently: Heb. greatly}\FTNT{§§§}{{\FR 20.19 }
when the…: Heb. in the day of the business}\FTNT{*}{{\FR 20.19 }
Ezel: or, that sheweth the way}
\VS{20}And I will
shoot
three
arrows on the
side
{\ADD{thereof}}, as though I
shot at a
mark.
\VS{21}And, behold, I will
send a
lad,
{\ADD{saying}},
Go, find
out the
arrows. If I
expressly
say unto the
lad, Behold, the
arrows
{\ADD{are}} on this side of thee,
take them; then
come thou: for
{\ADD{there is}}
peace to thee, and no
hurt;
{\ADD{as}} the
{\ND{
LORD}}
liveth.\FTNT{†}{{\FR 20.21 }
no hurt: Heb. not any thing}
\VS{22}But if I
say thus unto the young
man, Behold, the
arrows
{\ADD{are}}
beyond thee; go thy
way: for the
{\ND{
LORD}} hath sent thee
away.
\VS{23}And
{\ADD{as touching}} the
matter which thou and I have
spoken of, behold, the
{\ND{
LORD}}
{\ADD{be}} between thee and me
for
ever.
\par }{\PP \VS{24}¶ So
David
hid himself in the
field: and when the new
moon was come, the
king sat him
down to
eat
meat.
\VS{25}And the
king
sat upon his
seat, as at other
times,
{\ADD{even}} upon a
seat by the
wall: and
Jonathan
arose, and
Abner
sat by
Saul’s
side, and
David’s
place was
empty.
\VS{26}Nevertheless
Saul
spake not any
thing that
day: for he
thought, Something hath
befallen him, he
{\ADD{is}}
not
clean; surely he
{\ADD{is}} not
clean.
\VS{27}And it came to pass on the
morrow,
{\ADD{which was}} the
second
{\ADD{day}} of the
month, that
David’s
place was
empty: and
Saul
said unto
Jonathan his
son, Wherefore
cometh not the
son of
Jesse to
meat, neither
yesterday, nor to
day?
\VS{28}And
Jonathan
answered
Saul,
David
earnestly
asked
{\ADD{leave}} of
me
{\ADD{to go}} to
Beth-lehem:
\VS{29}And he
said, Let me
go, I pray thee; for our
family hath a
sacrifice in the
city; and my
brother, he hath
commanded me
{\ADD{to be there:}} and now, if I have
found
favour in thine
eyes, let me get
away, I pray thee, and
see my
brethren. Therefore he
cometh not unto the
king’s
table.
\VS{30}Then
Saul’s
anger was
kindled against
Jonathan, and he
said unto him, Thou
son of the
perverse
rebellious
{\ADD{woman}}, do not I
know that thou hast
chosen the
son of
Jesse to thine own
confusion, and unto the
confusion of thy
mother’s
nakedness?\FTNT{‡}{{\FR 20.30 }
Thou…: or, Thou perverse rebel: Heb. Son of perverse rebellion}
\VS{31}For as long
as the
son of
Jesse
liveth upon the
ground, thou shalt not be
established, nor thy
kingdom. Wherefore now
send and
fetch him unto me, for he shall surely
die.\FTNT{§}{{\FR 20.31 }
shall…: Heb. is the son of death}
\VS{32}And
Jonathan
answered
Saul his
father, and
said unto him, Wherefore shall he be
slain? what hath he
done?
\VS{33}And
Saul
cast a
javelin at him to
smite him: whereby
Jonathan
knew that it was
determined of his
father to
slay
David.
\VS{34}So
Jonathan
arose from the
table in
fierce
anger, and did
eat no
meat the
second
day of the
month: for he was
grieved for
David, because his
father had done him
shame.
\par }{\PP \VS{35}¶ And it came to pass in the
morning, that
Jonathan went
out into the
field at the time
appointed with
David, and a
little
lad with him.
\VS{36}And he
said unto his
lad,
Run, find
out now the
arrows which I
shoot.
{\ADD{And}} as the
lad
ran, he
shot an
arrow
beyond him.\FTNT{**}{{\FR 20.36 }
beyond…: Heb. to pass over him}
\VS{37}And when the
lad was
come to the
place of the
arrow which
Jonathan had
shot,
Jonathan
cried
after the
lad, and
said,
{\ADD{Is}} not the
arrow
beyond thee?
\VS{38}And
Jonathan
cried
after the
lad, Make
speed,
haste,
stay not. And
Jonathan’s
lad gathered
up the
arrows, and
came to his
master.
\VS{39}But the
lad
knew not any
thing: only
Jonathan and
David
knew the
matter.
\VS{40}And
Jonathan
gave his
artillery unto his
lad, and
said unto him,
Go,
carry
{\ADD{them}} to the
city.\FTNT{††}{{\FR 20.40 }
artillery: Heb. instruments}\FTNT{‡‡}{{\FR 20.40 }
his lad: Heb. the lad that was his}
\par }{\PP \VS{41}¶
{\ADD{And}} as soon as the
lad was
gone,
David
arose out of
{\ADD{a place}}
toward the
south, and
fell on his
face to the
ground, and
bowed himself
three
times: and they
kissed
one
another, and
wept
one with
another, until
David
exceeded.
\VS{42}And
Jonathan
said to
David,
Go in
peace, forasmuch as we have
sworn
both of us in the
name of the
{\ND{
LORD}},
saying, The
{\ND{
LORD}} be between me and thee, and between my
seed and thy
seed
for
ever. And he
arose and
departed: and
Jonathan
went into the
city.\FTNT{§§}{{\FR 20.42 }
forasmuch…: or, the
{\ND{LORD}} be witness of that which etc}

\par }\Chap{21}{\PP \VerseOne{1}Then
came
David to
Nob to
Ahimelech the
priest: and
Ahimelech was
afraid at the
meeting of
David, and
said unto him, Why
{\ADD{art}} thou alone, and no
man with thee?\FTNT{*}{{\FR 21.1 }
Ahimelech: also called Ahiah}
\VS{2}And
David
said unto
Ahimelech the
priest, The
king hath
commanded me a
business, and hath
said unto me, Let no
man
know any
thing of the
business whereabout I
send thee, and what I have
commanded thee: and I have
appointed
{\ADD{my}}
servants to
such and
such a
place.
\VS{3}Now therefore what
is under thine
hand?
give
{\ADD{me}}
five
{\ADD{loaves of}}
bread in mine
hand, or what there is
present.\FTNT{†}{{\FR 21.3 }
present: Heb. found}
\VS{4}And the
priest
answered
David, and
said,
{\ADD{There is}} no
common
bread under mine
hand, but there
is
hallowed
bread; if the young
men have
kept themselves at
least from
women.
\VS{5}And
David
answered the
priest, and
said unto him, Of a
truth
women
{\ADD{have been}}
kept from us about these
three
days, since I came
out, and the
vessels of the young
men are
holy, and
{\ADD{the bread is}} in a
manner
common, yea, though it were
sanctified this
day in the
vessel.\FTNT{‡}{{\FR 21.5 }
yea…: or, especially when this day there is other sanctified in the vessel}
\VS{6}So the
priest
gave him
hallowed
{\ADD{bread:}} for there was no
bread there but the
shewbread, that was
taken from
before the
{\ND{
LORD}}, to
put
hot
bread in the
day when it was taken
away.
\VS{7}Now a certain
man of the
servants of
Saul
{\ADD{was}} there that
day,
detained
before the
{\ND{
LORD}}; and his
name
{\ADD{was}}
Doeg, an
Edomite, the
chiefest of the
herdmen that
{\ADD{belonged}} to
Saul.
\par }{\PP \VS{8}¶ And
David
said unto
Ahimelech, And is
there
not here under thine
hand
spear or
sword? for I have neither
brought my
sword nor my
weapons with
me, because the
king’s
business
required
haste.
\VS{9}And the
priest
said, The
sword of
Goliath the
Philistine, whom thou
slewest in the
valley of
Elah, behold, it
{\ADD{is here}}
wrapped in a
cloth
behind the
ephod: if thou wilt
take that,
take
{\ADD{it:}} for
{\ADD{there is}} no
other
save that
here. And
David
said,
{\ADD{There is}} none like that;
give it me.
\par }{\PP \VS{10}¶ And
David
arose, and
fled that
day for
fear of
Saul, and
went to
Achish the
king of
Gath.
\VS{11}And the
servants of
Achish
said unto him,
{\ADD{Is}} not this
David the
king of the
land? did they not
sing one to another of
him in
dances,
saying,
Saul hath
slain his
thousands, and
David his ten
thousands?
\VS{12}And
David laid
up these
words in his
heart, and was
sore
afraid
of
Achish the
king of
Gath.
\VS{13}And he
changed his
behaviour
before them, and feigned himself
mad in their
hands, and
scrabbled on the
doors of the
gate, and let his
spittle fall
down upon his
beard.\FTNT{§}{{\FR 21.13 }
scrabbled: or, made marks}
\VS{14}Then
said
Achish unto his
servants, Lo, ye
see the
man is
mad: wherefore
{\ADD{then}} have ye
brought him to me?\FTNT{**}{{\FR 21.14 }
is mad: or, playeth the mad man}
\VS{15}Have I
need of mad
men, that ye have
brought this
{\ADD{fellow}} to play the mad
man in my presence? shall this
{\ADD{fellow}}
come into my
house?

\par }\Chap{22}{\PP \VerseOne{1}David therefore
departed thence, and
escaped to the
cave
Adullam: and when his
brethren and all his
father’s
house
heard
{\ADD{it}}, they went
down thither to him.
\VS{2}And every
one
{\ADD{that was}} in
distress, and every
one that
{\ADD{was}} in
debt, and every
one
{\ADD{that was}}
discontented,
gathered themselves unto him; and he became a
captain over them: and there were with him about
four
hundred
men.\FTNT{*}{{\FR 22.2 }
was in debt: Heb. had a creditor}\FTNT{†}{{\FR 22.2 }
discontented: Heb. bitter of soul}
\par }{\PP \VS{3}¶ And
David
went thence to
Mizpeh of
Moab: and he
said unto the
king of
Moab, Let my
father and my
mother, I pray thee, come
forth,
{\ADD{and be}} with you, till I
know what
God will
do for me.
\VS{4}And he
brought them
before the
king of
Moab: and they
dwelt with him all the
while that
David was in the
hold.
\par }{\PP \VS{5}¶ And the
prophet
Gad
said unto
David,
Abide not in the
hold;
depart, and
get thee into the
land of
Judah. Then
David
departed, and
came into the
forest of
Hareth.
\par }{\PP \VS{6}¶ When
Saul
heard that
David was
discovered, and the
men that
{\ADD{were}} with him, (now
Saul
abode in
Gibeah under a
tree in
Ramah, having his
spear in his
hand, and all his
servants
{\ADD{were}}
standing about him;)\FTNT{‡}{{\FR 22.6 }
tree…: or, grove in a high place}
\VS{7}Then
Saul
said unto his
servants that
stood about him, Hear
now, ye
Benjamites; will the
son of
Jesse
give every one of you
fields and
vineyards,
{\ADD{and}}
make you all
captains of
thousands, and
captains of
hundreds;
\VS{8}That all of you have
conspired against me, and
{\ADD{there is}} none that
sheweth me that my
son hath made a
league with the
son of
Jesse, and
{\ADD{there is}} none of you that is
sorry for me, or
sheweth unto me that my
son hath stirred
up my
servant against me, to lie in
wait, as at this
day?\FTNT{§}{{\FR 22.8 }
sheweth…: Heb. uncovereth mine ear}
\par }{\PP \VS{9}¶ Then
answered
Doeg the
Edomite, which was
set over the
servants of
Saul, and
said, I
saw the
son of
Jesse
coming to
Nob, to
Ahimelech the
son of
Ahitub.
\VS{10}And he
enquired of the
{\ND{
LORD}} for him, and
gave him
victuals, and
gave him the
sword of
Goliath the
Philistine.
\VS{11}Then the
king
sent to
call
Ahimelech the
priest, the
son of
Ahitub, and all his
father’s
house, the
priests that
{\ADD{were}} in
Nob: and they
came all of them to the
king.
\VS{12}And
Saul
said,
Hear now, thou
son of
Ahitub. And he
answered, Here I
{\ADD{am}}, my
lord.\FTNT{**}{{\FR 22.12 }
Here…: Heb. Behold me}
\VS{13}And
Saul
said unto him, Why have ye
conspired against me, thou and the
son of
Jesse, in that thou hast
given him
bread, and a
sword, and hast
enquired of
God for him, that he should
rise against me, to lie in
wait, as at this
day?
\VS{14}Then
Ahimelech
answered the
king, and
said, And who
{\ADD{is so}}
faithful among all thy
servants as
David, which is the
king’s son in
law, and
goeth at thy
bidding, and is
honourable in thine
house?
\VS{15}Did I
then
begin to
enquire of
God for him? be it far from
me: let not the
king
impute
{\ADD{any}}
thing unto his
servant,
{\ADD{nor}} to all the
house of my
father: for thy
servant
knew
nothing of all this,
less or
more.\FTNT{††}{{\FR 22.15 }
less…: Heb. little or great}
\VS{16}And the
king
said, Thou shalt
surely
die,
Ahimelech, thou, and all thy
father’s
house.
\par }{\PP \VS{17}¶ And the
king
said unto the
footmen that
stood about him,
Turn, and
slay the
priests of the
{\ND{
LORD}}; because their
hand also
{\ADD{is}} with
David, and because they
knew when he
fled, and did not
shew it to me. But the
servants of the
king
would not put
forth their
hand to
fall upon the
priests of the
{\ND{
LORD}}.\FTNT{‡‡}{{\FR 22.17 }
footmen: or, guard: Heb. runners}
\VS{18}And the
king
said to
Doeg,
Turn thou, and
fall upon the
priests. And
Doeg the
Edomite
turned, and he
fell upon the
priests, and
slew on that
day
fourscore and
five
persons that did
wear a
linen
ephod.
\VS{19}And
Nob, the
city of the
priests,
smote he with the
edge of the
sword, both
men and
women,
children and
sucklings, and
oxen, and
asses, and
sheep, with the
edge of the
sword.
\par }{\PP \VS{20}¶ And
one of the
sons of
Ahimelech the
son of
Ahitub,
named
Abiathar,
escaped, and
fled
after
David.
\VS{21}And
Abiathar
shewed
David that
Saul had
slain the
{\ND{
LORD’s}}
priests.
\VS{22}And
David
said unto
Abiathar, I
knew
{\ADD{it}} that
day, when
Doeg the
Edomite
{\ADD{was}} there, that he would
surely
tell
Saul: I have
occasioned
{\ADD{the death}} of all the
persons of thy
father’s
house.
\VS{23}Abide thou with me,
fear not: for he that
seeketh my
life
seeketh thy
life: but with me thou
{\ADD{shalt be}} in
safeguard.

\par }\Chap{23}{\PP \VerseOne{1}Then they
told
David,
saying, Behold, the
Philistines
fight against
Keilah, and they
rob the
threshingfloors.
\VS{2}Therefore
David
enquired of the
{\ND{
LORD}},
saying, Shall I
go and
smite these
Philistines? And the
{\ND{
LORD}}
said unto
David,
Go, and
smite the
Philistines, and
save
Keilah.
\VS{3}And
David’s
men
said unto him, Behold, we be
afraid here in
Judah: how much more then if we
come to
Keilah against the
armies of the
Philistines?
\VS{4}Then
David
enquired of the
{\ND{
LORD}} yet
again. And the
{\ND{
LORD}}
answered him and
said,
Arise, go
down to
Keilah; for I will
deliver the
Philistines into thine
hand.
\VS{5}So
David and his
men
went to
Keilah, and
fought with the
Philistines, and brought
away their
cattle, and
smote them with a
great
slaughter. So
David
saved the
inhabitants of
Keilah.
\VS{6}And it came to pass, when
Abiathar the
son of
Ahimelech
fled to
David to
Keilah,
{\ADD{that}} he came
down
{\ADD{with}} an
ephod in his
hand.
\par }{\PP \VS{7}¶ And it was
told
Saul that
David was
come to
Keilah. And
Saul
said,
God hath
delivered him into mine
hand; for he is shut
in, by
entering into a
town that hath
gates and
bars.
\VS{8}And
Saul
called all the
people
together to
war, to go
down to
Keilah, to
besiege
David and his
men.
\par }{\PP \VS{9}¶ And
David
knew that
Saul secretly
practised
mischief against him; and he
said to
Abiathar the
priest, Bring
hither the
ephod.
\VS{10}Then
said
David, O
{\ND{
LORD}}
God of
Israel, thy
servant hath
certainly
heard that
Saul
seeketh to
come to
Keilah, to
destroy the
city for my sake.
\VS{11}Will the
men of
Keilah deliver me
up into his
hand? will
Saul come
down, as thy
servant hath
heard? O
{\ND{
LORD}}
God of
Israel, I beseech thee,
tell thy
servant. And the
{\ND{
LORD}}
said, He will come
down.
\VS{12}Then
said
David, Will the
men of
Keilah
deliver me and my
men into the
hand of
Saul? And the
{\ND{
LORD}}
said, They will deliver
{\ADD{thee}}
up.\FTNT{*}{{\FR 23.12 }
deliver: Heb. shut up}
\par }{\PP \VS{13}¶ Then
David and his
men,
{\ADD{which were}} about
six
hundred,
arose and
departed out of
Keilah, and
went
whithersoever they could
go. And it was
told
Saul that
David was
escaped from
Keilah; and he
forbare to go
forth.
\par }{\PP \VS{14}And
David
abode in the
wilderness in strong
holds, and
remained in a
mountain in the
wilderness of
Ziph. And
Saul
sought him every
day, but
God
delivered him not into his
hand.
\VS{15}And
David
saw that
Saul was come
out to
seek his
life: and
David
{\ADD{was}} in the
wilderness of
Ziph in a
wood.
\par }{\PP \VS{16}¶ And
Jonathan
Saul’s
son
arose, and
went to
David into the
wood, and
strengthened his
hand in
God.
\VS{17}And he
said unto him,
Fear not: for the
hand of
Saul my
father shall not
find thee; and thou shalt be
king over
Israel, and I shall be
next unto thee; and that also
Saul my
father
knoweth.
\VS{18}And they
two
made a
covenant
before the
{\ND{
LORD}}: and
David
abode in the
wood, and
Jonathan
went to his
house.
\par }{\PP \VS{19}¶ Then came
up the
Ziphites to
Saul to
Gibeah,
saying, Doth not
David
hide himself with us in strong
holds in the
wood, in the
hill of
Hachilah, which
{\ADD{is}} on the
south of
Jeshimon?\FTNT{†}{{\FR 23.19 }
on…: Heb. on the right hand}\FTNT{‡}{{\FR 23.19 }
Jeshimon: or, the wilderness}
\VS{20}Now therefore, O
king, come
down according to all the
desire of thy
soul to come
down; and our part
{\ADD{shall be}} to
deliver him into the
king’s
hand.
\VS{21}And
Saul
said,
Blessed
{\ADD{be}} ye of the
{\ND{
LORD}}; for ye have
compassion on me.
\VS{22}Go, I pray you,
prepare yet, and
know and
see his
place where his
haunt is,
{\ADD{and}} who hath
seen him there: for it is
told me
{\ADD{that}} he dealeth
very
subtilly.\FTNT{§}{{\FR 23.22 }
haunt…: Heb. foot shall be}
\VS{23}See therefore, and take
knowledge of all the lurking
places where he
hideth himself, and come ye
again to me with the
certainty, and I will
go with you: and it shall come to pass, if he
be in the
land, that I will
search him out throughout all the
thousands of
Judah.
\VS{24}And they
arose, and
went to
Ziph
before
Saul: but
David and his
men
{\ADD{were}} in the
wilderness of
Maon, in the
plain on the
south of
Jeshimon.
\VS{25}Saul also and his
men
went to
seek
{\ADD{him}}. And they
told
David: wherefore he came
down into a
rock, and
abode in the
wilderness of
Maon. And when
Saul
heard
{\ADD{that}}, he
pursued
after
David in the
wilderness of
Maon.\FTNT{**}{{\FR 23.25 }
into…: or, from the rock}
\VS{26}And
Saul
went on this
side of the
mountain, and
David and his
men on that
side of the
mountain: and
David made
haste to get
away for
fear of
Saul; for
Saul and his
men
compassed
David and his
men round
about to
take them.
\par }{\PP \VS{27}¶ But there
came a
messenger unto
Saul,
saying,
Haste thee, and
come; for the
Philistines have
invaded the
land.\FTNT{††}{{\FR 23.27 }
invaded…: Heb. spread themselves upon, etc}
\VS{28}Wherefore
Saul
returned from
pursuing
after
David, and
went
against the
Philistines: therefore they
called that
place
Sela-hammahlekoth.\FTNT{‡‡}{{\FR 23.28 }
Sela-hammahlekoth: that is, The rock of divisions}
\par }{\PP \VS{29}¶ And
David went
up from thence, and
dwelt in strong
holds at
En-gedi.

\par }\Chap{24}{\PP \VerseOne{1}And it came to pass, when
Saul was
returned from
following the
Philistines, that it was
told him,
saying, Behold,
David
{\ADD{is}} in the
wilderness of
En-gedi.\FTNT{*}{{\FR 24.1 }
following: Heb. after}
\VS{2}Then
Saul
took
three
thousand
chosen
men out of all
Israel, and
went to
seek
David and his
men
upon the
rocks of the wild
goats.
\VS{3}And he
came to the
sheepcotes by the
way, where
{\ADD{was}} a
cave; and
Saul went
in to
cover his
feet: and
David and his
men
remained in the
sides of the
cave.
\VS{4}And the
men of
David
said unto him, Behold the
day of which the
{\ND{
LORD}}
said unto thee, Behold, I will
deliver thine
enemy into thine
hand, that thou mayest
do to him as it shall seem
good unto
thee. Then
David
arose, and cut
off the
skirt of
Saul’s
robe
privily.\FTNT{†}{{\FR 24.4 }
Saul’s…: Heb. the robe which was Saul’s}
\VS{5}And it came to pass
afterward, that
David’s
heart
smote him, because he had cut
off
Saul’s
skirt.
\VS{6}And he
said unto his
men, The
{\ND{
LORD}}
forbid that I should
do this
thing unto my
master, the
{\ND{
LORD’s}}
anointed, to stretch
forth mine
hand against him, seeing he
{\ADD{is}} the
anointed of the
{\ND{
LORD}}.
\VS{7}So
David
stayed his
servants with these
words, and
suffered them not to
rise against
Saul. But
Saul rose
up out of the
cave, and
went on
{\ADD{his}}
way.\FTNT{‡}{{\FR 24.7 }
stayed: Heb. cut off}
\VS{8}David also
arose afterward, and went
out of the
cave, and
cried
after
Saul,
saying, My
lord the
king. And when
Saul
looked
behind him,
David
stooped with his
face to the
earth, and
bowed himself.
\par }{\PP \VS{9}¶ And
David
said to
Saul, Wherefore
hearest thou
men’s
words,
saying, Behold,
David
seeketh thy
hurt?
\VS{10}Behold, this
day thine
eyes have
seen how that the
{\ND{
LORD}} had
delivered thee to
day into mine
hand in the
cave: and
{\ADD{some}}
bade
{\ADD{me}}
kill thee: but
{\ADD{mine eye}}
spared thee; and I
said, I will not put
forth mine
hand against my
lord; for he
{\ADD{is}} the
{\ND{
LORD’s}}
anointed.
\VS{11}Moreover, my
father,
see, yea,
see the
skirt of thy
robe in my
hand: for in that I cut
off the
skirt of thy
robe, and
killed thee not,
know thou and
see that
{\ADD{there is}} neither
evil nor
transgression in mine
hand, and I have not
sinned against thee; yet thou
huntest my
soul to
take it.
\VS{12}The
{\ND{
LORD}}
judge between me and thee, and the
{\ND{
LORD}}
avenge me of thee: but mine
hand shall not be upon thee.
\VS{13}As
saith the
proverb of the
ancients,
Wickedness
proceedeth from the
wicked: but mine
hand shall not be upon thee.
\VS{14}After whom is the
king of
Israel come
out?
after whom dost thou
pursue?
after a
dead
dog,
after
a
flea.
\VS{15}The
{\ND{
LORD}} therefore be
judge, and
judge between me and thee, and
see, and
plead my
cause, and
deliver me out of thine
hand.\FTNT{§}{{\FR 24.15 }
deliver: Heb. judge}
\par }{\PP \VS{16}¶ And it came to pass, when
David had made an
end of
speaking these
words unto
Saul, that
Saul
said,
{\ADD{Is}} this thy
voice, my
son
David? And
Saul lifted
up his
voice, and
wept.
\VS{17}And he
said to
David, Thou
{\ADD{art}} more
righteous than I: for thou hast
rewarded me
good, whereas I have
rewarded thee
evil.
\VS{18}And thou hast
shewed this
day how that thou hast
dealt
well with me: forasmuch as when the
{\ND{
LORD}} had
delivered me into thine
hand, thou
killedst me not.\FTNT{**}{{\FR 24.18 }
delivered: Heb. shut up}
\VS{19}For if a
man
find his
enemy, will he let him
go
well
away? wherefore the
{\ND{
LORD}}
reward thee
good for that thou hast
done unto me this
day.
\VS{20}And now, behold, I know
well that thou shalt
surely be
king, and that the
kingdom of
Israel shall be
established in thine
hand.
\VS{21}Swear now therefore unto me by the
{\ND{
LORD}}, that thou wilt
not cut
off my
seed
after me, and that thou wilt not
destroy my
name out of my
father’s
house.
\VS{22}And
David
sware unto
Saul. And
Saul
went
home; but
David and his
men gat them
up unto the
hold.

\par }\Chap{25}{\PP \VerseOne{1}And
Samuel
died; and all the
Israelites were gathered
together, and
lamented him, and
buried him in his
house at
Ramah. And
David
arose, and went
down to the
wilderness of
Paran.
\par }{\PP \VS{2}And
{\ADD{there was}} a
man in
Maon, whose
possessions
{\ADD{were}} in
Carmel; and the
man
{\ADD{was}}
very
great, and he had
three
thousand
sheep, and a
thousand
goats: and he was
shearing his
sheep in
Carmel.\FTNT{*}{{\FR 25.2 }
possessions: or, business}
\VS{3}Now the
name of the
man
{\ADD{was}}
Nabal; and the
name of his
wife
Abigail: and
{\ADD{she was}} a
woman of
good
understanding, and of a
beautiful
countenance: but the
man
{\ADD{was}}
churlish and
evil in his
doings; and he
{\ADD{was}} of the house of
Caleb.
\par }{\PP \VS{4}¶ And
David
heard in the
wilderness that
Nabal did
shear his
sheep.
\VS{5}And
David sent
out
ten young
men, and
David
said unto the young
men, Get you
up to
Carmel, and
go to
Nabal, and
greet him in my
name:\FTNT{†}{{\FR 25.5 }
greet…: Heb. ask him in my name of peace}
\VS{6}And thus shall ye
say to him that
liveth
{\ADD{in prosperity}},
Peace
{\ADD{be}} both to thee, and
peace
{\ADD{be}} to thine
house, and
peace
{\ADD{be}} unto all that thou hast.
\VS{7}And now I have
heard that thou hast
shearers: now thy
shepherds which were with us, we
hurt them not, neither was there
ought
missing unto them, all the
while they were in
Carmel.\FTNT{‡}{{\FR 25.7 }
hurt: Heb. shamed}
\VS{8}Ask thy young
men, and they will
shew thee. Wherefore let the young
men
find
favour in thine
eyes: for we
come in a
good
day:
give, I pray thee, whatsoever
cometh to thine
hand unto thy
servants, and to thy
son
David.
\VS{9}And when
David’s young
men
came, they
spake to
Nabal according to all those
words in the
name of
David, and
ceased.\FTNT{§}{{\FR 25.9 }
ceased: Heb. rested}
\par }{\PP \VS{10}¶ And
Nabal
answered
David’s
servants, and
said, Who
{\ADD{is}}
David? and who
{\ADD{is}} the
son of
Jesse? there be
many
servants now a
days that break
away every
man
from his
master.
\VS{11}Shall I then
take my
bread, and my
water, and my
flesh that I have
killed for my
shearers, and
give
{\ADD{it}} unto
men, whom I
know not whence they
{\ADD{be}}?\FTNT{**}{{\FR 25.11 }
flesh: Heb. slaughter}
\par }{\PP \VS{12}So
David’s young
men
turned their
way, and went
again, and
came and
told him all those
sayings.
\VS{13}And
David
said unto his
men, Gird ye
on every
man his
sword. And they girded
on every
man his
sword; and
David also girded
on his
sword: and there went
up
after
David about
four
hundred
men; and two
hundred
abode by the
stuff.
\par }{\PP \VS{14}¶ But
one of the young
men
told
Abigail,
Nabal’s
wife,
saying, Behold,
David
sent
messengers out of the
wilderness to
salute our
master; and he
railed on them.\FTNT{††}{{\FR 25.14 }
railed…: Heb. flew upon them}
\VS{15}But the
men
{\ADD{were}}
very
good unto us, and we were not
hurt, neither
missed we any
thing, as long
as we were
conversant with them, when we were in the
fields:\FTNT{‡‡}{{\FR 25.15 }
hurt: Heb. shamed}
\VS{16}They were a
wall unto us both by
night and
day, all the
while we were with them
keeping the
sheep.
\VS{17}Now therefore
know and
consider what thou wilt
do; for
evil is
determined against our
master, and against all his
household: for he
{\ADD{is such}} a
son of
Belial, that
{\ADD{a man}} cannot
speak to him.
\par }{\PP \VS{18}¶ Then
Abigail made
haste, and
took two
hundred
loaves, and
two
bottles of
wine, and
five
sheep ready
dressed, and
five
measures of
parched
{\ADD{corn}}, and an
hundred clusters of
raisins, and two
hundred
cakes of figs, and
laid
{\ADD{them}} on
asses.\FTNT{§§}{{\FR 25.18 }
clusters: or, lumps}
\VS{19}And she
said unto her
servants, Go
on
before me; behold, I
come
after you. But she
told not her
husband
Nabal.
\VS{20}And it was
{\ADD{so, as}} she
rode on the
ass, that she came
down by the
covert of the
hill, and, behold,
David and his
men came
down
against her; and she
met them.
\VS{21}Now
David had
said, Surely in
vain have I
kept all that
this
{\ADD{fellow}} hath in the
wilderness, so that
nothing was
missed of all that
{\ADD{pertained}} unto him: and he hath
requited me
evil for
good.
\VS{22}So and
more also
do
God unto the
enemies of
David, if I
leave of all that
{\ADD{pertain}} to him
by the
morning
light any that
pisseth against the
wall.
\VS{23}And when
Abigail
saw
David, she
hasted, and lighted
off the
ass, and
fell
before
David on her
face, and
bowed herself to the
ground,
\VS{24}And
fell at his
feet, and
said, Upon me, my
lord,
{\ADD{upon}} me
{\ADD{let this}}
iniquity
{\ADD{be:}} and let thine
handmaid, I pray thee,
speak in thine
audience, and
hear the
words of thine
handmaid.\FTNT{***}{{\FR 25.24 }
audience: Heb. ears}
\VS{25}Let not my
lord, I pray thee,
regard this
man of
Belial,
{\ADD{even}}
Nabal: for as his
name
{\ADD{is}}, so
{\ADD{is}} he;
Nabal
{\ADD{is}} his
name, and
folly
{\ADD{is}} with him: but I thine
handmaid
saw not the young
men of my
lord, whom thou didst
send.\FTNT{†††}{{\FR 25.25 }
regard: Heb. lay it to his heart}\FTNT{‡‡‡}{{\FR 25.25 }
Nabal: that is, Fool}
\VS{26}Now therefore, my
lord,
{\ADD{as}} the
{\ND{
LORD}}
liveth, and
{\ADD{as}} thy
soul
liveth, seeing the
{\ND{
LORD}} hath
withholden thee from
coming to
{\ADD{shed}}
blood, and from
avenging thyself with thine own
hand, now let thine
enemies, and they that
seek
evil to my
lord, be as
Nabal.\FTNT{§§§}{{\FR 25.26 }
avenging…: Heb. saving thyself}
\VS{27}And now this
blessing which thine
handmaid hath
brought unto my
lord, let it even be
given unto the young
men that
follow my
lord.\FTNT{*}{{\FR 25.27 }
blessing: or, present}\FTNT{†}{{\FR 25.27 }
follow…: Heb. walk at the feet of, etc}
\VS{28}I pray thee,
forgive the
trespass of thine
handmaid: for the
{\ND{
LORD}} will
certainly
make my
lord a
sure
house; because my
lord
fighteth the
battles of the
{\ND{
LORD}}, and
evil hath not been
found in thee
{\ADD{all}} thy
days.
\VS{29}Yet a
man is
risen to
pursue thee, and to
seek thy
soul: but the
soul of my
lord shall be
bound in the
bundle of
life with the
{\ND{
LORD}} thy
God; and the
souls of thine
enemies, them shall he sling
out,
{\ADD{as out}} of the
middle of a
sling.\FTNT{‡}{{\FR 25.29 }
as out…: Heb. in the midst of the bought of a sling}
\VS{30}And it shall come to pass, when the
{\ND{
LORD}} shall have
done to my
lord according to all the
good that he hath
spoken concerning thee, and shall have
appointed thee
ruler over
Israel;
\VS{31}That this shall be no
grief unto thee, nor
offence of
heart unto my
lord, either that thou hast
shed
blood
causeless, or that my
lord hath
avenged himself: but when the
{\ND{
LORD}} shall have dealt
well with my
lord, then
remember thine
handmaid.\FTNT{§}{{\FR 25.31 }
no grief: Heb. no staggering, or, stumbling}
\par }{\PP \VS{32}¶ And
David
said to
Abigail,
Blessed
{\ADD{be}} the
{\ND{
LORD}}
God of
Israel, which
sent thee this
day to
meet me:
\VS{33}And
blessed
{\ADD{be}} thy
advice, and
blessed
{\ADD{be}} thou, which hast
kept me this
day from
coming to
{\ADD{shed}}
blood, and from
avenging myself with mine own
hand.
\VS{34}For in very
deed,
{\ADD{as}} the
{\ND{
LORD}}
God of
Israel
liveth, which hath kept me
back from
hurting thee,
except thou hadst
hasted and
come to
meet me, surely there had not been
left unto
Nabal by the
morning
light any that
pisseth against the
wall.
\VS{35}So
David
received of her
hand
{\ADD{that}} which she had
brought him, and
said unto her, Go
up in
peace to thine
house;
see, I have
hearkened to thy
voice, and have
accepted thy
person.
\par }{\PP \VS{36}¶ And
Abigail
came to
Nabal; and, behold, he held a
feast in his
house, like the
feast of a
king; and
Nabal’s
heart
{\ADD{was}}
merry within him, for he
{\ADD{was}}
very
drunken: wherefore she
told him
nothing,
less or
more, until the
morning
light.
\VS{37}But it came to pass in the
morning, when the
wine was gone
out of
Nabal, and his
wife had
told him these
things, that his
heart
died
within him, and he became
{\ADD{as}} a
stone.
\VS{38}And it came to pass about
ten
days
{\ADD{after}}, that the
{\ND{
LORD}}
smote
Nabal, that he
died.
\par }{\PP \VS{39}¶ And when
David
heard that
Nabal was
dead, he
said,
Blessed
{\ADD{be}} the
{\ND{
LORD}}, that hath
pleaded the
cause of my
reproach from the
hand of
Nabal, and hath
kept his
servant from
evil: for the
{\ND{
LORD}} hath
returned the
wickedness of
Nabal upon his own
head. And
David
sent and
communed with
Abigail, to
take her to him to
wife.
\VS{40}And when the
servants of
David were
come to
Abigail to
Carmel, they
spake unto her,
saying,
David
sent us unto thee, to
take thee to him to
wife.
\VS{41}And she
arose, and
bowed herself on
{\ADD{her}}
face to the
earth, and
said, Behold,
{\ADD{let}} thine
handmaid
{\ADD{be}} a
servant to
wash the
feet of the
servants of my
lord.
\VS{42}And
Abigail
hasted, and
arose, and
rode upon an
ass, with
five
damsels of hers that
went
after her; and she
went
after the
messengers of
David, and became his
wife.\FTNT{**}{{\FR 25.42 }
after her: Heb. at her feet}
\VS{43}David also
took
Ahinoam of
Jezreel; and they were also
both of them his
wives.
\par }{\PP \VS{44}¶ But
Saul had
given
Michal his
daughter,
David’s
wife, to
Phalti the
son of
Laish, which
{\ADD{was}} of
Gallim.\FTNT{††}{{\FR 25.44 }
Phalti: also called, Phaltiel}

\par }\Chap{26}{\PP \VerseOne{1}And the
Ziphites
came unto
Saul to
Gibeah,
saying, Doth not
David
hide himself in the
hill of
Hachilah,
{\ADD{which is}}
before
Jeshimon?
\VS{2}Then
Saul
arose, and went
down to the
wilderness of
Ziph, having
three
thousand
chosen
men of
Israel with him, to
seek
David in the
wilderness of
Ziph.
\VS{3}And
Saul
pitched in the
hill of
Hachilah, which
{\ADD{is}}
before
Jeshimon, by the
way. But
David
abode in the
wilderness, and he
saw that
Saul
came
after him into the
wilderness.
\VS{4}David therefore sent
out
spies, and
understood that
Saul was
come in very
deed.
\par }{\PP \VS{5}¶ And
David
arose, and
came to the
place where
Saul had
pitched: and
David
beheld the
place where
Saul
lay, and
Abner the
son of
Ner, the
captain of his
host: and
Saul
lay in the
trench, and the
people
pitched round
about him.\FTNT{*}{{\FR 26.5 }
trench: or, midst of his carriages}
\par }{\PP \VS{6}Then
answered
David and
said to
Ahimelech the
Hittite, and to
Abishai the
son of
Zeruiah,
brother to
Joab,
saying, Who will go
down with me to
Saul to the
camp? And
Abishai
said, I will go
down with thee.
\VS{7}So
David and
Abishai
came to the
people by
night: and, behold,
Saul
lay
sleeping within the
trench, and his
spear
stuck in the
ground at his
bolster: but
Abner and the
people
lay round
about him.
\VS{8}Then
said
Abishai to
David,
God hath
delivered thine
enemy into thine
hand this
day: now therefore let me
smite him, I pray thee, with the
spear even to the
earth at
once, and I will not
{\ADD{smite}} him the second
time.\FTNT{†}{{\FR 26.8 }
delivered: Heb. shut up}
\VS{9}And
David
said to
Abishai,
Destroy him not: for who can stretch
forth his
hand against the
{\ND{
LORD’s}}
anointed, and be
guiltless?
\VS{10}David
said furthermore,
{\ADD{As}} the
{\ND{
LORD}}
liveth, the
{\ND{
LORD}} shall
smite him; or his
day shall
come to
die; or he shall
descend into
battle, and
perish.
\VS{11}The
{\ND{
LORD}}
forbid that I should stretch
forth mine
hand against the
{\ND{
LORD’s}}
anointed: but, I pray thee,
take thou now the
spear that
{\ADD{is}} at his
bolster, and the
cruse of
water, and let us
go.
\VS{12}So
David
took the
spear and the
cruse of
water from
Saul’s
bolster; and they gat them
away, and no man
saw
{\ADD{it}}, nor
knew
{\ADD{it}}, neither
awaked: for they
{\ADD{were}} all
asleep; because a deep
sleep from the
{\ND{
LORD}} was
fallen upon them.
\par }{\PP \VS{13}¶ Then
David went
over to the other
side, and
stood on the
top of an
hill afar
off; a
great
space
{\ADD{being}} between them:
\VS{14}And
David
cried to the
people, and to
Abner the
son of
Ner,
saying,
Answerest thou not,
Abner? Then
Abner
answered and
said, Who
{\ADD{art}} thou
{\ADD{that}}
criest to the
king?
\VS{15}And
David
said to
Abner,
{\ADD{Art}} not thou a
{\ADD{valiant}}
man? and who
{\ADD{is}} like to thee in
Israel? wherefore then hast thou not
kept thy
lord the
king? for there
came
one of the
people in to
destroy the
king thy
lord.
\VS{16}This
thing
{\ADD{is}} not
good that thou hast
done.
{\ADD{As}} the
{\ND{
LORD}}
liveth, ye
{\ADD{are}}
worthy to
die, because ye have not
kept your
master, the
{\ND{
LORD’s}}
anointed. And now
see where the
king’s
spear
{\ADD{is}}, and the
cruse of
water that
{\ADD{was}} at his
bolster.\FTNT{‡}{{\FR 26.16 }
worthy to die: Heb. the sons of death}
\VS{17}And
Saul
knew
David’s
voice, and
said,
{\ADD{Is}} this thy
voice, my
son
David? And
David
said,
{\ADD{It is}} my
voice, my
lord, O
king.
\VS{18}And he
said, Wherefore doth my
lord thus
pursue
after his
servant? for what have I
done? or what
evil
{\ADD{is}} in mine
hand?
\VS{19}Now therefore, I pray thee, let my
lord the
king
hear the
words of his
servant. If the
{\ND{
LORD}} have stirred thee
up against me, let him
accept an
offering: but if
{\ADD{they be}} the
children of
men,
cursed
{\ADD{be}} they
before the
{\ND{
LORD}}; for they have
driven me out this
day from
abiding in the
inheritance of the
{\ND{
LORD}},
saying,
Go,
serve
other
gods.\FTNT{§}{{\FR 26.19 }
accept: Heb. smell}\FTNT{**}{{\FR 26.19 }
abiding: Heb. cleaving}
\VS{20}Now therefore, let not my
blood
fall to the
earth before the
face of the
{\ND{
LORD}}: for the
king of
Israel is come
out to
seek
a
flea, as when one doth
hunt a
partridge in the
mountains.
\par }{\PP \VS{21}¶ Then
said
Saul, I have
sinned:
return, my
son
David: for I will no more do thee
harm, because my
soul was
precious in thine
eyes this
day: behold, I have played the
fool, and have
erred
exceedingly.
\VS{22}And
David
answered and
said, Behold the
king’s
spear! and let
one of the young
men come
over and
fetch it.
\VS{23}The
{\ND{
LORD}}
render to every
man his
righteousness and his
faithfulness: for the
{\ND{
LORD}}
delivered thee into
{\ADD{my}}
hand to
day, but I
would not stretch
forth mine
hand against the
{\ND{
LORD’s}}
anointed.
\VS{24}And, behold, as thy
life was much set
by this
day in mine
eyes, so let my
life be much set
by in the
eyes of the
{\ND{
LORD}}, and let him
deliver me out of all
tribulation.
\VS{25}Then
Saul
said to
David,
Blessed
{\ADD{be}} thou, my
son
David: thou shalt both
do
great
{\ADD{things}}, and also shalt
still
prevail. So
David
went on his
way, and
Saul
returned to his
place.

\par }\Chap{27}{\PP \VerseOne{1}And
David
said in his
heart, I shall now
perish
one
day by the
hand of
Saul:
{\ADD{there is}} nothing
better for me than
that I should
speedily
escape into the
land of the
Philistines; and
Saul shall
despair of me, to
seek me any more in any
coast of
Israel: so shall I
escape out of his
hand.\FTNT{*}{{\FR 27.1 }
perish: Heb. be consumed}
\VS{2}And
David
arose, and he passed
over with the
six
hundred
men that
{\ADD{were}} with him unto
Achish, the
son of
Maoch,
king of
Gath.
\VS{3}And
David
dwelt with
Achish at
Gath, he and his
men, every
man with his
household,
{\ADD{even}}
David with his
two
wives,
Ahinoam the
Jezreelitess, and
Abigail the
Carmelitess,
Nabal’s
wife.
\VS{4}And it was
told
Saul that
David was
fled to
Gath: and he
sought no more
again for him.
\par }{\PP \VS{5}¶ And
David
said unto
Achish, If I have now
found
grace in thine
eyes, let them
give me a
place in
some
town in the
country, that I may
dwell there: for why should thy
servant
dwell in the
royal
city with thee?
\VS{6}Then
Achish
gave him
Ziklag that
day: wherefore
Ziklag pertaineth unto the
kings of
Judah unto this
day.
\VS{7}And the
time that
David
dwelt in the
country of the
Philistines was a full
year and
four
months.\FTNT{†}{{\FR 27.7 }
the time: Heb. the number of days}\FTNT{‡}{{\FR 27.7 }
a full year: Heb. a year of days}
\par }{\PP \VS{8}¶ And
David and his
men went
up, and
invaded the
Geshurites, and the
Gezrites, and the
Amalekites: for
those
{\ADD{nations were}} of
old the
inhabitants of the
land, as thou
goest to
Shur, even unto the
land of
Egypt.\FTNT{§}{{\FR 27.8 }
Gezrites: or, Gerzites}
\VS{9}And
David
smote the
land, and left neither
man nor
woman
alive, and took
away the
sheep, and the
oxen, and the
asses, and the
camels, and the
apparel, and
returned, and
came to
Achish.
\VS{10}And
Achish
said,
Whither have ye made a
road to
day? And
David
said, Against the
south of
Judah, and against the
south of the
Jerahmeelites, and against the
south of the
Kenites.\FTNT{**}{{\FR 27.10 }
Whither…: or, Did you not make a road, etc}
\VS{11}And
David
saved neither
man nor
woman
alive, to
bring
{\ADD{tidings}} to
Gath,
saying, Lest they should
tell on us,
saying, So
did
David, and so
{\ADD{will be}} his
manner all the
while he
dwelleth in the
country of the
Philistines.
\VS{12}And
Achish
believed
David,
saying, He hath made his
people
Israel
utterly to
abhor him; therefore he shall be my
servant for
ever.\FTNT{††}{{\FR 27.12 }
utterly…: Heb. to stink}

\par }\Chap{28}{\PP \VerseOne{1}And it came to pass in those
days, that the
Philistines
gathered their
armies
together for
warfare, to
fight with
Israel. And
Achish
said unto
David,
Know thou
assuredly, that thou shalt go
out with me to
battle, thou and thy
men.
\VS{2}And
David
said to
Achish,
Surely thou shalt
know what thy
servant can
do. And
Achish
said to
David, Therefore will I
make thee
keeper of mine
head for
ever.
\par }{\PP \VS{3}¶ Now
Samuel was
dead, and all
Israel had
lamented him, and
buried him in
Ramah, even in his own
city. And
Saul had put
away those that had familiar
spirits, and the
wizards, out of the
land.
\VS{4}And the
Philistines gathered themselves
together, and
came and
pitched in
Shunem: and
Saul
gathered all
Israel
together, and they
pitched in
Gilboa.
\VS{5}And when
Saul
saw the
host of the
Philistines, he was
afraid, and his
heart
greatly
trembled.
\VS{6}And when
Saul
enquired of the
{\ND{
LORD}}, the
{\ND{
LORD}}
answered him not, neither by
dreams, nor by
Urim, nor by
prophets.
\par }{\PP \VS{7}¶ Then
said
Saul unto his
servants,
Seek me a
woman that
hath a familiar
spirit, that I may
go to her, and
enquire of her. And his
servants
said to him, Behold,
{\ADD{there is}} a
woman that
hath a familiar
spirit at
Endor.
\VS{8}And
Saul
disguised himself, and put
on
other
raiment, and he
went, and
two
men with him, and they
came to the
woman by
night: and he
said, I pray thee,
divine unto me by the familiar
spirit, and bring me
{\ADD{him}}
up, whom I shall
name unto thee.
\VS{9}And the
woman
said unto him, Behold, thou
knowest what
Saul hath
done, how he hath cut
off those that have familiar
spirits, and the
wizards, out of the
land: wherefore then layest thou a
snare for my
life, to cause me to
die?
\VS{10}And
Saul
sware to her by the
{\ND{
LORD}},
saying,
{\ADD{As}} the
{\ND{
LORD}}
liveth, there
shall no
punishment
happen to thee for this
thing.
\VS{11}Then
said the
woman, Whom shall I bring
up unto thee? And he
said, Bring me
up
Samuel.
\VS{12}And when the
woman
saw
Samuel, she
cried with a
loud
voice: and the
woman
spake to
Saul,
saying, Why hast thou
deceived me? for thou
{\ADD{art}}
Saul.
\VS{13}And the
king
said unto her, Be not
afraid: for what
sawest thou? And the
woman
said unto
Saul, I
saw
gods
ascending out of the
earth.
\VS{14}And he
said unto her, What
form
{\ADD{is}} he of? And she
said, An
old
man cometh
up; and he
{\ADD{is}}
covered with a
mantle. And
Saul
perceived that it
{\ADD{was}}
Samuel, and he
stooped with
{\ADD{his}}
face to the
ground, and
bowed himself.\FTNT{*}{{\FR 28.14 }
What…: Heb. What is his form?}
\par }{\PP \VS{15}¶ And
Samuel
said to
Saul, Why hast thou
disquieted me, to bring me
up? And
Saul
answered, I am
sore
distressed; for the
Philistines make
war against me, and
God is
departed from me, and
answereth me no more, neither
by
prophets, nor by
dreams: therefore I have
called thee, that thou mayest make
known unto me what I shall
do.\FTNT{†}{{\FR 28.15 }
by prophets: Heb. by the hand of prophets}
\VS{16}Then
said
Samuel, Wherefore then dost thou
ask of me, seeing the
{\ND{
LORD}} is
departed from thee, and is become thine
enemy?
\VS{17}And the
{\ND{
LORD}} hath
done to him, as he
spake by
me: for the
{\ND{
LORD}} hath
rent the
kingdom out of thine
hand, and
given it to thy
neighbour,
{\ADD{even}} to
David:\FTNT{‡}{{\FR 28.17 }
to him: or, for himself}\FTNT{§}{{\FR 28.17 }
me: Heb. mine hand}
\VS{18}Because thou
obeyedst not the
voice of the
{\ND{
LORD}}, nor
executedst his
fierce
wrath upon
Amalek, therefore hath the
{\ND{
LORD}}
done this
thing unto thee this
day.
\VS{19}Moreover the
{\ND{
LORD}} will also
deliver
Israel with thee into the
hand of the
Philistines: and to
morrow
{\ADD{shalt}} thou and thy
sons
{\ADD{be}} with me: the
{\ND{
LORD}} also shall
deliver the
host of
Israel into the
hand of the
Philistines.
\par }{\PP \VS{20}Then
Saul
fell
straightway
all
along on the
earth, and was
sore
afraid, because of the
words of
Samuel: and there was no
strength in him; for he had
eaten no
bread all the
day, nor all the
night.\FTNT{**}{{\FR 28.20 }
fell…: Heb. made haste, and fell with the fulness of his stature}
\par }{\PP \VS{21}¶ And the
woman
came unto
Saul, and
saw that he was
sore
troubled, and
said unto him, Behold, thine
handmaid hath
obeyed thy
voice, and I have
put my
life in my
hand, and have
hearkened unto thy
words which thou
spakest unto me.
\VS{22}Now therefore, I pray thee,
hearken thou also unto the
voice of thine
handmaid, and let me
set a
morsel of
bread
before thee; and
eat, that thou mayest have
strength, when thou
goest on thy
way.
\VS{23}But he
refused, and
said, I will not
eat. But his
servants, together with the
woman,
compelled him; and he
hearkened unto their
voice. So he
arose from the
earth, and
sat upon the
bed.
\VS{24}And the
woman had a
fat
calf in the
house; and she
hasted, and
killed it, and
took
flour, and
kneaded
{\ADD{it}}, and did
bake unleavened
bread thereof:
\VS{25}And she
brought
{\ADD{it}}
before
Saul, and
before his
servants; and they did
eat. Then they rose
up, and went
away that
night.

\par }\Chap{29}{\PP \VerseOne{1}Now the
Philistines gathered
together all their
armies to
Aphek: and the
Israelites
pitched by a
fountain which
{\ADD{is}} in
Jezreel.
\VS{2}And the
lords of the
Philistines passed
on by
hundreds, and by
thousands: but
David and his
men passed
on in the
rereward with
Achish.
\VS{3}Then
said the
princes of the
Philistines, What
{\ADD{do}} these
Hebrews
{\ADD{here}}? And
Achish
said unto the
princes of the
Philistines,
{\ADD{Is}} not this
David, the
servant of
Saul the
king of
Israel, which hath been with me these
days, or these
years, and I have
found no
fault in him
since he
fell
{\ADD{unto me}} unto this
day?
\VS{4}And the
princes of the
Philistines were
wroth with him; and the
princes of the
Philistines
said unto him,
Make this
fellow
return, that he may go
again to his
place which thou hast
appointed him, and let him not go
down with us to
battle, lest in the
battle he be an
adversary to us: for wherewith should he
reconcile himself unto his
master?
{\ADD{should it}} not
{\ADD{be}} with the
heads of these
men?
\VS{5}{\ADD{Is}} not this
David, of whom they
sang one to another in
dances,
saying,
Saul
slew his
thousands, and
David his ten
thousands?
\par }{\PP \VS{6}¶ Then
Achish
called
David, and
said unto him, Surely,
{\ADD{as}} the
{\ND{
LORD}}
liveth, thou hast been
upright, and thy going
out and thy coming
in with me in the
host
{\ADD{is}}
good in my
sight: for I have not
found
evil in thee since the
day of thy
coming unto me unto this
day: nevertheless the
lords
favour thee
not.\FTNT{*}{{\FR 29.6 }
the lords…: Heb. thou art not good in the eyes of the lords}
\VS{7}Wherefore now
return, and
go in
peace, that thou
displease not the
lords of the
Philistines.\FTNT{†}{{\FR 29.7 }
displease…: Heb. do not evil in the eyes of the lords}
\par }{\PP \VS{8}¶ And
David
said unto
Achish, But what have I
done? and what hast thou
found in thy
servant so long
as I have been with
thee unto this
day, that I may not
go
fight against the
enemies of my
lord the
king?\FTNT{‡}{{\FR 29.8 }
with…: Heb. before thee}
\VS{9}And
Achish
answered and
said to
David, I
know that thou
{\ADD{art}}
good in my
sight, as an
angel of
God: notwithstanding the
princes of the
Philistines have
said, He shall not go
up with us to the
battle.
\VS{10}Wherefore now rise up
early in the
morning with thy
master’s
servants that are
come with thee: and as soon as ye be up
early in the
morning, and have
light,
depart.
\VS{11}So
David and his
men rose up
early to
depart in the
morning, to
return into the
land of the
Philistines. And the
Philistines went
up to
Jezreel.

\par }\Chap{30}{\PP \VerseOne{1}And it came to pass, when
David and his
men were
come to
Ziklag on the
third
day, that the
Amalekites had
invaded the
south, and
Ziklag, and
smitten
Ziklag, and
burned it with
fire;
\VS{2}And had taken the
women
captives, that
{\ADD{were}} therein: they
slew not
any, either
great or
small, but carried
{\ADD{them}}
away, and
went on their
way.
\par }{\PP \VS{3}¶ So
David and his
men
came to the
city, and, behold,
{\ADD{it was}}
burned with
fire; and their
wives, and their
sons, and their
daughters, were taken
captives.
\VS{4}Then
David and the
people that
{\ADD{were}} with him lifted
up their
voice and
wept, until they had no more
power to
weep.
\VS{5}And
David’s
two
wives were taken
captives,
Ahinoam the
Jezreelitess, and
Abigail the
wife of
Nabal the
Carmelite.
\VS{6}And
David was
greatly
distressed; for the
people
spake of
stoning him, because the
soul of all the
people was
grieved, every
man for his
sons and for his
daughters: but
David
encouraged himself in the
{\ND{
LORD}} his
God.\FTNT{*}{{\FR 30.6 }
grieved: Heb. bitter}
\par }{\PP \VS{7}And
David
said to
Abiathar the
priest,
Ahimelech’s
son, I pray thee, bring me
hither the
ephod. And
Abiathar
brought thither the
ephod to
David.
\VS{8}And
David
enquired at the
{\ND{
LORD}},
saying, Shall I
pursue
after this
troop? shall I
overtake them? And he
answered him,
Pursue: for thou shalt
surely
overtake
{\ADD{them}}, and without
fail
recover
{\ADD{all}}.
\VS{9}So
David
went, he and the
six
hundred
men that
{\ADD{were}} with him, and
came to the
brook
Besor, where those that were left
behind
stayed.
\VS{10}But
David
pursued, he and
four
hundred
men: for two
hundred abode
behind, which were so
faint that they could not go
over the
brook
Besor.
\par }{\PP \VS{11}¶ And they
found an
Egyptian in the
field, and
brought him to
David, and
gave him
bread, and he did
eat; and they made him
drink
water;
\VS{12}And they
gave him a
piece of a
cake of figs, and
two clusters of
raisins: and when he had
eaten, his
spirit came
again to him: for he had
eaten no
bread, nor
drunk
{\ADD{any}}
water,
three
days and
three
nights.
\VS{13}And
David
said unto him, To whom
{\ADD{belongest}} thou? and whence
{\ADD{art}} thou? And he
said, I
{\ADD{am}} a young
man of
Egypt,
servant to an
Amalekite; and my
master
left me, because
three
days agone I fell
sick.
\VS{14}We made an
invasion
{\ADD{upon}} the
south of the
Cherethites, and upon
{\ADD{the coast}} which
{\ADD{belongeth}} to
Judah, and upon the
south of
Caleb; and we
burned
Ziklag with
fire.
\VS{15}And
David
said to him, Canst thou bring me
down to this
company? And he
said,
Swear unto me by
God, that thou wilt neither
kill me,
nor
deliver me into the
hands of my
master, and I will bring thee
down to this
company.
\par }{\PP \VS{16}¶ And when he had brought him
down, behold,
{\ADD{they were}} spread
abroad upon
all the
earth,
eating and
drinking, and
dancing, because of all the
great
spoil that they had
taken out of the
land of the
Philistines, and out of the
land of
Judah.
\VS{17}And
David
smote them from the
twilight even unto the
evening of the next
day: and there
escaped not a
man of them, save
four
hundred
young
men, which
rode upon
camels, and
fled.\FTNT{†}{{\FR 30.17 }
the next…: Heb. their morrow}
\VS{18}And
David
recovered all that the
Amalekites had carried
away: and
David
rescued his
two
wives.
\VS{19}And there was nothing
lacking to them,
neither
small nor
great, neither
sons nor
daughters, neither
spoil, nor any
{\ADD{thing}} that they had
taken to them:
David
recovered all.
\VS{20}And
David
took all the
flocks and the
herds,
{\ADD{which}} they
drave
before
those
{\ADD{other}}
cattle, and
said, This
{\ADD{is}}
David’s
spoil.
\par }{\PP \VS{21}¶ And
David
came to the two
hundred
men, which were so
faint that they could not
follow
David, whom they had made also to
abide at the
brook
Besor: and they went
forth to
meet
David, and to
meet the
people that
{\ADD{were}} with him: and when
David came
near to the
people, he
saluted them.\FTNT{‡}{{\FR 30.21 }
saluted…: or, asked them how they did}
\VS{22}Then
answered all the
wicked
men and
{\ADD{men}} of
Belial, of
those that
went with
David, and
said, Because they
went not with us, we will not
give them
{\ADD{ought}} of the
spoil that we have
recovered, save to every
man his
wife and his
children, that they may lead
{\ADD{them}}
away, and
depart.\FTNT{§}{{\FR 30.22 }
those: Heb. men}
\VS{23}Then
said
David, Ye shall not do
so, my
brethren, with that which the
{\ND{
LORD}} hath
given us, who hath
preserved us, and
delivered the
company that
came against us into our
hand.
\VS{24}For who will
hearken unto you in this
matter? but as his
part
{\ADD{is}} that goeth
down to the
battle, so
{\ADD{shall}} his
part
{\ADD{be}} that
tarrieth by the
stuff: they shall
part
alike.
\VS{25}And it was
{\ADD{so}} from that
day
forward, that he
made it a
statute and an
ordinance for
Israel unto this
day.\FTNT{**}{{\FR 30.25 }
forward: Heb. and forward}
\par }{\PP \VS{26}¶ And when
David
came to
Ziklag, he
sent of the
spoil unto the
elders of
Judah,
{\ADD{even}} to his
friends,
saying, Behold a
present for you of the
spoil of the
enemies of the
{\ND{
LORD}};\FTNT{††}{{\FR 30.26 }
present: Heb. blessing}
\VS{27}To
{\ADD{
them}} which
{\ADD{were}} in
Beth-el, and to
{\ADD{them}} which
{\ADD{were}} in
south
Ramoth, and to
{\ADD{them}} which
{\ADD{were}} in
Jattir,
\VS{28}And to
{\ADD{them}} which
{\ADD{were}} in
Aroer, and to
{\ADD{them}} which
{\ADD{were}} in
Siphmoth, and to
{\ADD{them}} which
{\ADD{were}} in
Eshtemoa,
\VS{29}And to
{\ADD{them}} which
{\ADD{were}} in
Rachal, and to
{\ADD{them}} which
{\ADD{were}} in the
cities of the
Jerahmeelites, and to
{\ADD{them}} which
{\ADD{were}} in the
cities of the
Kenites,
\VS{30}And to
{\ADD{them}} which
{\ADD{were}} in
Hormah, and to
{\ADD{them}} which
{\ADD{were}} in
Chor-ashan, and to
{\ADD{them}} which
{\ADD{were}} in
Athach,
\VS{31}And to
{\ADD{them}} which
{\ADD{were}} in
Hebron, and to all the
places where
David himself and his
men were wont to
haunt.

\par }\Chap{31}{\PP \VerseOne{1}Now the
Philistines
fought against
Israel: and the
men of
Israel
fled from
before the
Philistines, and fell
down
slain in
mount
Gilboa.\FTNT{*}{{\FR 31.1 }
slain: or, wounded}
\VS{2}And the
Philistines followed
hard upon
Saul and upon his
sons; and the
Philistines
slew
Jonathan, and
Abinadab, and
Malchi-shua,
Saul’s
sons.
\VS{3}And the
battle went
sore against
Saul, and the
archers
hit him; and he was
sore
wounded of the
archers.\FTNT{†}{{\FR 31.3 }
and the archers: Heb. and the shooters, men with bows}\FTNT{‡}{{\FR 31.3 }
hit…: Heb. found him}
\VS{4}Then
said
Saul unto his
armourbearer,
Draw thy
sword, and thrust me
through therewith; lest these
uncircumcised
come and thrust me
through, and
abuse me. But his
armourbearer
would not; for he was
sore
afraid. Therefore
Saul
took a
sword, and
fell upon it.\FTNT{§}{{\FR 31.4 }
abuse…: or, mock me}
\VS{5}And when his
armourbearer
saw that
Saul was
dead, he
fell likewise upon his
sword, and
died with him.
\VS{6}So
Saul
died, and his
three
sons, and his
armourbearer, and all his
men, that same
day
together.
\par }{\PP \VS{7}¶ And when the
men of
Israel that
{\ADD{were}} on the other
side of the
valley, and
{\ADD{they}} that
{\ADD{were}} on the other
side
Jordan,
saw that the
men of
Israel
fled, and that
Saul and his
sons were
dead, they
forsook the
cities, and
fled; and the
Philistines
came and
dwelt in them.
\par }{\PP \VS{8}And it came to pass on the
morrow, when the
Philistines
came to
strip the
slain, that they
found
Saul and his
three
sons
fallen in
mount
Gilboa.
\VS{9}And they cut
off his
head, and stripped
off his
armour, and
sent into the
land of the
Philistines round
about, to
publish
{\ADD{it in}} the
house of their
idols, and among the
people.
\VS{10}And they
put his
armour in the
house of
Ashtaroth: and they
fastened his
body to the
wall of
Beth-shan.
\par }{\PP \VS{11}¶ And when the
inhabitants of
Jabesh-gilead
heard of that which the
Philistines had
done to
Saul;\FTNT{**}{{\FR 31.11 }
of that: or, concerning him that}
\VS{12}All the
valiant
men
arose, and
went all
night, and
took the
body of
Saul and the
bodies of his
sons from the
wall of
Beth-shan, and
came to
Jabesh, and
burnt them there.
\VS{13}And they
took their
bones, and
buried
{\ADD{them}} under a
tree at
Jabesh, and
fasted
seven
days.
\par }