\NormalFont\ShortTitle{1 Timothy}
{\MT THE FIRST EPISTLE OF PAUL THE APOSTLE TO TIMOTHY

\par }\ChapOne{1}{\PP \VerseOne{1}Paul, an
apostle of
Jesus
Christ
by the
commandment of
God
our
Saviour,
and
Lord
Jesus
Christ,
{\ADD{which is}}
our
hope;
\VS{2}Unto
Timothy,
{\ADD{my}}
own
son
in the
faith:
Grace,
mercy,
{\ADD{and}}
peace,
from
God
our
Father
and
Jesus
Christ
our
Lord.
\par }{\PP \VS{3}As I
besought
thee to abide
still
at
Ephesus, when I
went
into
Macedonia,
that thou mightest
charge
some that they
teach
no other
doctrine,
\VS{4}Neither give heed
to
fables
and
endless
genealogies,
which
minister
questions,
rather
than
godly
edifying which
is
in
faith:
{\ADD{so do}}.
\VS{5}Now the
end of the
commandment
is
charity out
of a
pure
heart,
and
{\ADD{of}} a
good
conscience,
and
{\ADD{of}}
faith
unfeigned:
\VS{6}From
which
some having
swerved have turned
aside
unto vain
jangling;
\VS{7}Desiring to
be teachers of the
law;
understanding
neither
what they
say,
nor
whereof they
affirm.
\VS{8}But we
know
that the
law
{\ADD{is}}
good,
if a
man
use
it
lawfully;
\VS{9}Knowing
this,
that the
law
is
not made
for a righteous
man,
but for the
lawless
and
disobedient, for the
ungodly
and for
sinners, for
unholy
and
profane, for murderers of
fathers
and murderers of
mothers, for
manslayers,
\VS{10}For
whoremongers, for them that defile themselves with
mankind, for
menstealers, for
liars, for perjured
persons,
and
if there
be
any other
thing that is contrary
to
sound
doctrine;
\VS{11}According
to the
glorious
gospel of the
blessed
God,
which was committed
to
my
trust.
\VS{12}And
I
thank
Christ
Jesus
our
Lord, who hath
enabled
me, for
that he
counted
me
faithful, putting
me
into the
ministry;
\VS{13}Who
was
before a
blasphemer,
and a
persecutor,
and
injurious:
but I obtained
mercy,
because I
did
{\ADD{it}}
ignorantly
in
unbelief.
\VS{14}And the
grace of
our
Lord was exceeding
abundant
with
faith
and
love which
is
in
Christ
Jesus.
\VS{15}This
{\ADD{is}} a
faithful
saying,
and
worthy of
all
acceptation,
that
Christ
Jesus
came
into the
world to
save
sinners; of
whom
I
am
chief.
\VS{16}Howbeit
for this
cause I obtained
mercy,
that
in
me
first
Jesus
Christ might shew
forth
all
longsuffering,
for a
pattern to them which should
hereafter
believe
on
him
to
life
everlasting.
\VS{17}Now unto the
King
eternal,
immortal,
invisible, the
only
wise
God,
{\ADD{be}}
honour
and
glory
for
ever and
ever.
Amen.
\VS{18}This
charge I
commit unto
thee,
son
Timothy, according
to the
prophecies which went
before
on
thee,
that
thou
by
them mightest
war a
good
warfare;
\VS{19}Holding
faith,
and a
good
conscience;
which
some having put
away
concerning
faith have made
shipwreck:
\VS{20}Of
whom
is
Hymenaeus
and
Alexander;
whom I have
delivered unto
Satan,
that they may
learn
not to
blaspheme.

\par }\Chap{2}{\PP \VerseOne{1}I
exhort
therefore,
that,
first of
all,
supplications,
prayers,
intercessions,
{\ADD{and}} giving of
thanks, be
made
for
all
men;
\VS{2}For
kings,
and
{\ADD{for}}
all that
are
in
authority;
that we may
lead a
quiet
and
peaceable
life
in
all
godliness
and
honesty.
\VS{3}For
this
{\ADD{is}}
good
and
acceptable in the sight
of
God
our
Saviour;
\VS{4}Who will
have
all
men to be
saved,
and to
come
unto the
knowledge of the
truth.
\VS{5}For
{\ADD{there is}}
one
God,
and
one
mediator between
God
and
men, the
man
Christ
Jesus;
\VS{6}Who
gave
himself a
ransom
for
all, to be
testified in
due
time.
\VS{7}Where
unto
I am
ordained a
preacher,
and an
apostle, (I
speak the
truth
in
Christ,
{\ADD{and}}
lie
not;) a
teacher of the
Gentiles
in
faith
and
verity.
\VS{8}I
will
therefore
that
men
pray
every
where, lifting
up
holy
hands,
without
wrath
and
doubting.
\VS{9}In like
manner
also,
that
women
adorn
themselves
in
modest
apparel,
with
shamefacedness
and
sobriety;
not
with broided
hair,
or
gold,
or
pearls,
or
costly
array;
\VS{10}But (which
becometh
women
professing
godliness)
with
good
works.
\VS{11}Let the
woman
learn
in
silence
with
all
subjection.
\VS{12}But I
suffer
not a
woman to
teach,
nor to usurp authority
over the
man,
but to
be
in
silence.
\VS{13}For
Adam
was
first
formed,
then
Eve.
\VS{14}And
Adam
was
not
deceived,
but the
woman being
deceived
was
in the
transgression.
\VS{15}Notwithstanding she shall be
saved
in
childbearing,
if they
continue
in
faith
and
charity
and
holiness
with
sobriety.

\par }\Chap{3}{\PP \VerseOne{1}This
{\ADD{is}} a
true
saying,
If a
man
desire the office of a
bishop, he
desireth a
good
work.
\VS{2}A
bishop
then
must
be
blameless, the
husband of
one
wife,
vigilant,
sober, of good
behaviour, given to
hospitality, apt to
teach;
\VS{3}Not given to
wine,
no
striker,
not greedy of filthy
lucre;
but
patient, not a
brawler, not
covetous;
\VS{4}One that
ruleth
well his
own
house,
having his
children
in
subjection
with
all
gravity;
\VS{5}(For
if a
man
know
not
how to
rule his
own
house,
how shall he take
care of the
church of
God?)
\VS{6}Not a
novice,
lest being lifted up with
pride he
fall
into the
condemnation of the
devil.
\VS{7}Moreover
he
must
have a
good
report
of them which are
without;
lest he
fall
into
reproach
and the
snare of the
devil.
\VS{8}Likewise
{\ADD{must}} the
deacons
{\ADD{be}}
grave,
not
doubletongued,
not given
to
much
wine,
not greedy of filthy
lucre;
\VS{9}Holding the
mystery of the
faith
in a
pure
conscience.
\VS{10}And
let
these
also
first be
proved;
then let them use the office of a
deacon,
being
{\ADD{found}}
blameless.
\VS{11}Even
so
{\ADD{must their}}
wives
{\ADD{be}}
grave,
not
slanderers,
sober,
faithful
in all
things.
\VS{12}Let the
deacons
be the
husbands of
one
wife,
ruling their
children
and their
own
houses
well.
\VS{13}For they that have used the office of a
deacon
well
purchase to
themselves a
good
degree,
and
great
boldness
in the
faith which
is
in
Christ
Jesus.
\VS{14}These
things write
I unto
thee,
hoping to
come
unto
thee
shortly:
\VS{15}But
if I tarry
long,
that thou mayest
know
how thou
oughtest to behave
thyself
in the
house of
God,
which
is the
church of the
living
God, the
pillar
and
ground of the
truth.
\VS{16}And without
controversy
great
is the
mystery of
godliness:
God was
manifest
in the
flesh,
justified
in the
Spirit,
seen of
angels,
preached
unto the
Gentiles, believed
on
in the
world, received
up
into
glory.

\par }\Chap{4}{\PP \VerseOne{1}Now the
Spirit
speaketh
expressly,
that
in the
latter
times
some shall depart
from the
faith, giving heed
to
seducing
spirits,
and
doctrines of
devils;
\VS{2}Speaking
lies
in
hypocrisy;
having
their
conscience seared with a hot
iron;
\VS{3}Forbidding to
marry,
{\ADD{and commanding}} to abstain
from
meats,
which
God hath
created
to be
received
with
thanksgiving of them which
believe
and
know the
truth.
\VS{4}For
every
creature of
God
{\ADD{is}}
good,
and
nothing to be
refused, if it be
received
with
thanksgiving:
\VS{5}For it is
sanctified
by the
word of
God
and
prayer.
\VS{6}If thou
put the
brethren in remembrance
of these
things, thou shalt
be a
good
minister of
Jesus
Christ, nourished up
in the
words of
faith
and of
good
doctrine,
whereunto thou hast
attained.
\VS{7}But
refuse
profane
and old
wives’
fables,
and
exercise
thyself
{\ADD{rather}}
unto
godliness.
\VS{8}For
bodily
exercise
profiteth
little:
but
godliness
is
profitable
unto all
things,
having
promise of the
life that now
is,
and of that which is to
come.
\VS{9}This
{\ADD{is}} a
faithful
saying
and
worthy of
all
acceptation.
\VS{10}For
therefore
we
both
labour
and suffer
reproach,
because we
trust
in the
living
God,
who
is the
Saviour of
all
men,
specially of those that
believe.
\par }{\PP \VS{11}These
things
command
and
teach.
\VS{12}Let no
man
despise
thy
youth;
but be
thou an
example of the
believers,
in
word,
in
conversation,
in
charity,
in
spirit,
in
faith,
in
purity.
\VS{13}Till I
come, give
attendance to
reading, to
exhortation, to
doctrine.
\VS{14}Neglect
not the
gift that is
in
thee,
which was
given
thee
by
prophecy,
with the laying
on of the
hands of the
presbytery.
\VS{15}Meditate
upon these
things; give thyself
wholly
to
them;
that
thy
profiting
may
appear
to
all.
\VS{16}Take heed
unto
thyself,
and unto the
doctrine; continue
in
them:
for in
doing
this thou
shalt
both
save
thyself,
and them that
hear
thee.

\par }\Chap{5}{\PP \VerseOne{1}Rebuke
not an
elder,
but
intreat
{\ADD{him}}
as a
father;
{\ADD{and}} the younger
men
as
brethren;
\VS{2}The elder
women
as
mothers; the
younger
as
sisters,
with
all
purity.
\VS{3}Honour
widows that
are
widows
indeed.
\VS{4}But
if
any
widow
have
children
or
nephews, let them
learn
first to shew
piety at
home,
and
to
requite their
parents:
for
that
is
good
and
acceptable
before
God.
\VS{5}Now she
that
is a
widow
indeed,
and
desolate,
trusteth
in
God,
and continueth
in
supplications
and
prayers
night
and
day.
\VS{6}But she
that liveth in
pleasure is
dead while she
liveth.
\VS{7}And these
things give in
charge,
that they may
be
blameless.
\VS{8}But
if
any
provide
not
for his
own,
and
specially for those of his own
house, he hath
denied the
faith,
and
is
worse than an
infidel.
\VS{9}Let
not a
widow be taken into the
number
under
threescore years
old, having
been the
wife of
one
man,
\VS{10}Well reported
of
for
good
works;
if she have brought up
children,
if she have lodged
strangers,
if she have
washed the
saints’
feet,
if she have
relieved the
afflicted,
if she have diligently
followed
every
good
work.
\VS{11}But the
younger
widows
refuse:
for
when they have begun to wax wanton
against
Christ, they
will
marry;
\VS{12}Having
damnation,
because they have cast
off
their
first
faith.
\VS{13}And
withal they
learn
{\ADD{to be}}
idle, wandering
about from house to
house;
and
not
only
idle,
but
tattlers
also
and
busybodies,
speaking things
which they
ought
not.
\VS{14}I
will
therefore
that the younger
women
marry, bear
children, guide the
house,
give
none
occasion to the
adversary to speak
reproachfully.
\VS{15}For
some
are
already turned
aside
after
Satan.
\VS{16}If
any
man
or
woman that
believeth
have
widows, let them
relieve
them,
and
let
not the
church be
charged;
that it may
relieve them
that
are
widows
indeed.
\VS{17}Let the
elders that
rule
well be counted
worthy of
double
honour,
especially they who
labour
in the
word
and
doctrine.
\VS{18}For the
scripture
saith, Thou
shalt
not
muzzle the
ox that treadeth out the
corn.
And, The
labourer
{\ADD{is}}
worthy of
his
reward.
\VS{19}Against an
elder
receive
not an
accusation,
but
before
two
or
three
witnesses.
\VS{20}Them that
sin
rebuke
before
all,
that
others
also
may
fear.
\VS{21}I
charge
{\ADD{thee}}
before
God,
and the
Lord
Jesus
Christ,
and the
elect
angels,
that thou
observe these
things
without preferring one before
another,
doing
nothing
by
partiality.
\VS{22}Lay
hands
suddenly
on no
man,
neither be partaker
of other
men’s
sins:
keep
thyself
pure.
\VS{23}Drink no
longer
water,
but
use a
little
wine
for
thy
stomach’s
sake
and
thine
often
infirmities.
\VS{24}Some
men’s
sins
are open
beforehand, going
before
to
judgment;
and
some
{\ADD{men}} they follow
after.
\VS{25}Likewise
also the
good
works
{\ADD{of some}}
are manifest
beforehand;
and they that
are
otherwise
can
not be
hid.

\par }\Chap{6}{\PP \VerseOne{1}Let as
many
servants
as
are
under the
yoke
count their
own
masters
worthy of
all
honour,
that the
name of
God
and
{\ADD{his}}
doctrine
be
not
blasphemed.
\VS{2}And they that
have
believing
masters, let
them
not
despise
{\ADD{them}},
because they
are
brethren;
but
rather
do
{\ADD{them}}
service,
because they
are
faithful
and
beloved,
partakers of the
benefit. These
things
teach
and
exhort.
\par }{\PP \VS{3}If any
man teach
otherwise,
and
consent
not to
wholesome
words,
{\ADD{even}}
the words of
our
Lord
Jesus
Christ,
and to the
doctrine which is according
to
godliness;
\VS{4}He is
proud,
knowing
nothing,
but
doting
about
questions
and strifes of
words,
whereof
cometh
envy,
strife,
railings,
evil
surmisings,
\VS{5}Perverse
disputings of
men of
corrupt
minds,
and
destitute of the
truth, supposing
that
gain
is
godliness:
from
such withdraw
thyself.
\VS{6}But
godliness
with
contentment
is
great
gain.
\VS{7}For we
brought
nothing
into
{\ADD{this}}
world,
{\ADD{and it is}}
certain we
can
carry
nothing
out.
\VS{8}And
having
food
and
raiment let us
be
therewith
content.
\VS{9}But they that
will be
rich
fall
into
temptation
and a
snare,
and
{\ADD{into}}
many
foolish
and
hurtful
lusts,
which
drown
men
in
destruction
and
perdition.
\VS{10}For the love of
money
is the
root of
all
evil:
which
while
some coveted
after, they have
erred
from the
faith,
and
pierced
themselves
through with
many
sorrows.
\VS{11}But
thou,
O
man of
God,
flee these
things;
and follow
after
righteousness,
godliness,
faith,
love,
patience,
meekness.
\VS{12}Fight the
good
fight of
faith, lay hold
on
eternal
life,
where
unto thou
art
also
called,
and hast
professed a
good
profession
before
many
witnesses.
\VS{13}I
give
thee
charge in the sight
of
God, who
quickeneth all
things,
and
{\ADD{before}}
Christ
Jesus,
who
before
Pontius
Pilate
witnessed a
good
confession;
\VS{14}That
thou
keep
{\ADD{this}}
commandment without
spot,
unrebukeable,
until the
appearing of
our
Lord
Jesus
Christ:
\VS{15}Which in
his
times he shall
shew,
{\ADD{who is}} the
blessed
and
only
Potentate, the
King of
kings,
and
Lord of
lords;
\VS{16}Who
only
hath
immortality, dwelling
in the
light which no man can approach
unto;
whom
no
man hath
seen,
nor
can
see: to
whom
{\ADD{be}}
honour
and
power
everlasting.
Amen.
\VS{17}Charge them that are
rich
in
this
world, that they
be
not
highminded,
nor
trust
in
uncertain
riches,
but
in the
living
God, who
giveth
us
richly all
things
to enjoy;
\VS{18}That they do
good, that they be
rich
in
good
works, ready to
distribute, willing to
communicate;
\VS{19}Laying up in
store for
themselves a
good
foundation
against the time to
come,
that they may lay hold
on
eternal
life.
\VS{20}O
Timothy,
keep that which is committed to thy
trust,
avoiding
profane
{\ADD{and}} vain
babblings,
and
oppositions of
science falsely so
called:
\VS{21}Which
some
professing have
erred
concerning the
faith.
Grace
{\ADD{be}}
with
thee.
Amen.
\par }{\SH The first to Timothy was written from Laodicea, which is the chiefest city of Phrygia Pacatiana.
\par }