\NormalFont\ShortTitle{2 Chronicles}
{\MT The Second Book of the Chronicles

\par }\ChapOne{1}{\PP \VerseOne{1}And
Solomon the
son of
David was
strengthened in his
kingdom, and the
{\ND{
LORD}} his
God
{\ADD{was}} with him, and
magnified him
exceedingly.
\VS{2}Then
Solomon
spake unto all
Israel, to the
captains of
thousands and of
hundreds, and to the
judges, and to every
governor in all
Israel, the
chief of the
fathers.
\VS{3}So
Solomon, and all the
congregation with him,
went to the high
place that
{\ADD{was}} at
Gibeon; for there was the
tabernacle of the
congregation of
God, which
Moses the
servant of the
{\ND{
LORD}} had
made in the
wilderness.
\VS{4}But the
ark of
God had
David brought
up from
Kirjath-jearim to
{\ADD{the place which}}
David had
prepared for it: for he had
pitched a
tent for it at
Jerusalem.
\VS{5}Moreover the
brasen
altar, that
Bezaleel the
son of
Uri, the
son of
Hur, had
made, he
put
before the
tabernacle of the
{\ND{
LORD}}: and
Solomon and the
congregation
sought unto it.\FTNT{*}{{\FR 1.5 }
he put: or, was there}
\VS{6}And
Solomon went
up thither to the
brasen
altar
before the
{\ND{
LORD}}, which
{\ADD{was}} at the
tabernacle of the
congregation, and
offered a
thousand burnt
offerings upon it.
\par }{\PP \VS{7}¶ In that
night did
God
appear unto
Solomon, and
said unto him,
Ask what I shall
give thee.
\VS{8}And
Solomon
said unto
God, Thou hast
shewed
great
mercy unto
David my
father, and hast made me to
reign in his stead.
\VS{9}Now, O
{\ND{
LORD}}
God, let thy
promise unto
David my
father be
established: for thou hast made me
king over a
people like the
dust of the
earth in
multitude.\FTNT{†}{{\FR 1.9 }
like…: Heb. much as the dust of the earth}
\VS{10}Give me now
wisdom and
knowledge, that I may go
out and come
in
before this
people: for who can
judge this thy
people,
{\ADD{that is so}}
great?
\VS{11}And
God
said to
Solomon, Because this was in thine
heart, and thou hast not
asked
riches,
wealth, or
honour, nor the
life of thine
enemies, neither yet hast
asked
long
life; but hast
asked
wisdom and
knowledge for thyself, that thou mayest
judge my
people, over whom I have made thee
king:
\VS{12}Wisdom and
knowledge
{\ADD{is}}
granted unto thee; and I will
give thee
riches, and
wealth, and
honour, such as none of the
kings have had that
{\ADD{have been}}
before thee, neither shall there any
after thee have the like.
\par }{\PP \VS{13}¶ Then
Solomon
came
{\ADD{from his journey}} to the high
place that
{\ADD{was}} at
Gibeon to
Jerusalem, from
before the
tabernacle of the
congregation, and
reigned over
Israel.
\VS{14}And
Solomon
gathered
chariots and
horsemen: and he had a
thousand and
four
hundred
chariots, and
twelve
thousand
horsemen, which he
placed in the
chariot
cities, and with the
king at
Jerusalem.
\VS{15}And the
king
made
silver and
gold at
Jerusalem
{\ADD{as plenteous}} as
stones, and cedar
trees
made he as the sycomore
trees that
{\ADD{are}} in the
vale for
abundance.\FTNT{‡}{{\FR 1.15 }
made: Heb. gave}
\VS{16}And
Solomon had
horses
brought out of
Egypt, and linen
yarn: the
king’s
merchants
received the linen
yarn at a
price.\FTNT{§}{{\FR 1.16 }
Solomon…: Heb. the going forth of the horses which was Solomon’s}
\VS{17}And they fetched
up, and brought
forth out of
Egypt a
chariot for
six
hundred
{\ADD{shekels}} of
silver, and an
horse for an
hundred and
fifty: and so brought they
out
{\ADD{horses}} for all the
kings of the
Hittites, and for the
kings of
Syria, by their
means.\FTNT{**}{{\FR 1.17 }
means: Heb. hand}

\par }\Chap{2}{\PP \VerseOne{1}And
Solomon
determined to
build an
house for the
name of the
{\ND{
LORD}}, and an
house for his
kingdom.
\VS{2}And
Solomon told
out threescore and
ten
thousand
men to bear
burdens, and
fourscore
thousand to
hew in the
mountain, and
three
thousand and
six
hundred to
oversee them.
\par }{\PP \VS{3}¶ And
Solomon
sent to
Huram the
king of
Tyre,
saying, As thou didst
deal with
David my
father, and didst
send him
cedars to
build him an
house to
dwell therein,
{\ADD{even so deal with me}}.\FTNT{*}{{\FR 2.3 }
Huram: or, Hiram}
\VS{4}Behold, I
build an
house to the
name of the
{\ND{
LORD}} my
God, to
dedicate
{\ADD{it}} to him,
{\ADD{and}} to
burn
before him
sweet
incense, and for the
continual
shewbread, and for the burnt
offerings
morning and
evening, on the
sabbaths, and on the new
moons, and on the solemn
feasts of the
{\ND{
LORD}} our
God. This
{\ADD{is an ordinance}} for
ever to
Israel.\FTNT{†}{{\FR 2.4 }
sweet…: Heb. incense of spices}
\VS{5}And the
house which I
build
{\ADD{is}}
great: for
great
{\ADD{is}} our
God above all
gods.
\VS{6}But who is
able to
build him an
house, seeing the
heaven and
heaven of
heavens cannot
contain him? who
{\ADD{am}} I then, that I should
build him an
house, save
only to burn
sacrifice
before him?\FTNT{‡}{{\FR 2.6 }
is able: Heb. hath retained, or, obtained strength}
\VS{7}Send me now therefore a
man
cunning to
work in
gold, and in
silver, and in
brass, and in
iron, and in
purple, and
crimson, and
blue, and that can
skill to
grave with the cunning
men that
{\ADD{are}} with me in
Judah and in
Jerusalem, whom
David my
father did
provide.\FTNT{§}{{\FR 2.7 }
to grave: Heb. to grave gravings}
\VS{8}Send me also
cedar
trees, fir
trees, and algum
trees, out of
Lebanon: for I
know that thy
servants can
skill to
cut
timber in
Lebanon; and, behold, my
servants
{\ADD{shall be}} with thy
servants,\FTNT{**}{{\FR 2.8 }
algum: also called, almuggim}
\VS{9}Even to
prepare me
timber in
abundance: for the
house which I am about to
build
{\ADD{shall be}}
wonderful
great.\FTNT{††}{{\FR 2.9 }
wonderful…: Heb. great and wonderful}
\VS{10}And, behold, I will
give to thy
servants, the
hewers that
cut
timber,
twenty
thousand
measures of
beaten
wheat, and
twenty
thousand
measures of
barley, and
twenty
thousand
baths of
wine, and
twenty
thousand
baths of
oil.
\par }{\PP \VS{11}¶ Then
Huram the
king of
Tyre
answered in
writing, which he
sent to
Solomon, Because the
{\ND{
LORD}} hath
loved his
people, he hath
made thee
king over them.
\VS{12}Huram
said moreover,
Blessed
{\ADD{be}} the
{\ND{
LORD}}
God of
Israel, that
made
heaven and
earth, who hath
given to
David the
king a
wise
son,
endued with
prudence and
understanding, that might
build an
house for the
{\ND{
LORD}}, and an
house for his
kingdom.\FTNT{‡‡}{{\FR 2.12 }
endued…: Heb. knowing prudence and understanding}
\VS{13}And now I have
sent a
cunning
man,
endued with
understanding, of
Huram my
father’s,
\VS{14}The
son of a
woman of the
daughters of
Dan, and his
father
{\ADD{was}} a
man of
Tyre,
skilful to
work in
gold, and in
silver, in
brass, in
iron, in
stone, and in
timber, in
purple, in
blue, and in fine
linen, and in
crimson; also to
grave any manner of
graving, and to find
out every
device which shall be
put to him, with thy cunning
men, and with the cunning
men of my
lord
David thy
father.
\VS{15}Now therefore the
wheat, and the
barley, the
oil, and the
wine, which my
lord hath
spoken of, let him
send unto his
servants:
\VS{16}And we will
cut
wood out of
Lebanon, as much as thou shalt
need: and we will
bring it to thee in
floats by
sea to
Joppa; and thou shalt carry it
up to
Jerusalem.\FTNT{§§}{{\FR 2.16 }
as much…: Heb according to all thy need}\FTNT{***}{{\FR 2.16 }
Joppa: Heb. Japho}
\par }{\PP \VS{17}¶ And
Solomon
numbered all the
strangers that
{\ADD{were}} in the
land of
Israel,
after the
numbering wherewith
David his
father had
numbered them; and they were
found an
hundred and
fifty
thousand and
three
thousand and
six
hundred.\FTNT{†††}{{\FR 2.17 }
the strangers: Heb. the men the strangers}
\VS{18}And he
set threescore and
ten
thousand of them
{\ADD{to be}} bearers of
burdens, and
fourscore
thousand
{\ADD{to be}}
hewers in the
mountain, and
three
thousand and
six
hundred
overseers to
set the
people a
work.

\par }\Chap{3}{\PP \VerseOne{1}Then
Solomon
began to
build the
house of the
{\ND{
LORD}} at
Jerusalem in
mount
Moriah, where
{\ADD{the
{\ND{Lord}}}}
appeared unto
David his
father, in the
place that
David had
prepared in the
threshingfloor of
Ornan the
Jebusite.\FTNT{*}{{\FR 3.1 }
where…: or, which was seen of David his father}\FTNT{†}{{\FR 3.1 }
Ornan: also called, Araunah}
\VS{2}And he
began to
build in the
second
{\ADD{day}} of the
second
month, in the
fourth
year of his
reign.
\par }{\PP \VS{3}¶ Now these
{\ADD{are the things wherein}}
Solomon was
instructed for the
building of the
house of
God. The
length by
cubits after the
first
measure
{\ADD{was}}
threescore
cubits, and the
breadth
twenty
cubits.\FTNT{‡}{{\FR 3.3 }
instructed: Heb. founded}
\VS{4}And the
porch that
{\ADD{was}} in the
front
{\ADD{of the house}}, the
length
{\ADD{of it was}} according to the
breadth of the
house,
twenty
cubits, and the
height
{\ADD{was}} an
hundred and
twenty: and he
overlaid it
within with
pure
gold.
\VS{5}And the
greater
house he
cieled with
fir
tree, which he
overlaid with
fine
gold, and
set thereon palm
trees and
chains.
\VS{6}And he
garnished the
house with
precious
stones for
beauty: and the
gold
{\ADD{was}}
gold of
Parvaim.\FTNT{§}{{\FR 3.6 }
garnished: Heb. covered}
\VS{7}He
overlaid also the
house, the
beams, the
posts, and the
walls thereof, and the
doors thereof, with
gold; and
graved
cherubims on the
walls.
\VS{8}And he
made the
most
holy
house, the
length whereof
{\ADD{was}} according
to the
breadth of the
house,
twenty
cubits, and the
breadth thereof
twenty
cubits: and he
overlaid it with
fine
gold,
{\ADD{amounting}} to
six
hundred
talents.
\VS{9}And the
weight of the
nails
{\ADD{was}}
fifty
shekels of
gold. And he
overlaid the upper
chambers with
gold.
\par }{\PP \VS{10}And in the
most
holy
house he
made
two
cherubims of
image
work, and
overlaid them with
gold.\FTNT{**}{{\FR 3.10 }
image…: or, (as some think) of moveable work}
\par }{\PP \VS{11}¶ And the
wings of the
cherubims
{\ADD{were}}
twenty
cubits
long:
one
wing
{\ADD{of the one cherub was}}
five
cubits,
reaching to the
wall of the
house: and the
other
wing
{\ADD{was likewise}}
five
cubits,
reaching to the
wing of the
other
cherub.
\VS{12}And
{\ADD{one}}
wing of the
other
cherub
{\ADD{was}}
five
cubits,
reaching to the
wall of the
house: and the
other
wing
{\ADD{was}}
five
cubits
{\ADD{also}},
joining to the
wing of the
other
cherub.
\VS{13}The
wings of these
cherubims spread themselves
forth
twenty
cubits: and they
stood on their
feet, and their
faces
{\ADD{were}}
inward.\FTNT{††}{{\FR 3.13 }
inward: or, toward the house}
\par }{\PP \VS{14}¶ And he
made the
vail
{\ADD{of}}
blue, and
purple, and
crimson, and fine
linen, and
wrought
cherubims thereon.\FTNT{‡‡}{{\FR 3.14 }
wrought: Heb. caused to ascend}
\VS{15}Also he
made
before the
house
two
pillars of
thirty and
five
cubits
high, and the
chapiter that
{\ADD{was}} on the
top of each of them
{\ADD{was}}
five
cubits.\FTNT{§§}{{\FR 3.15 }
high: Heb. long}
\VS{16}And he
made
chains,
{\ADD{as}} in the
oracle, and
put
{\ADD{them}} on the
heads of the
pillars; and
made an
hundred
pomegranates, and
put
{\ADD{them}} on the
chains.
\VS{17}And he reared
up the
pillars
before the
temple,
one on the right
hand, and the
other on the
left; and
called the
name of that on the right
hand
Jachin, and the
name of that on the
left
Boaz.\FTNT{***}{{\FR 3.17 }
Jachin: that is, He shall establish}\FTNT{†††}{{\FR 3.17 }
Boaz: that is, In it is strength}

\par }\Chap{4}{\PP \VerseOne{1}Moreover he
made an
altar of
brass,
twenty
cubits the
length thereof, and
twenty
cubits the
breadth thereof, and
ten
cubits the
height thereof.
\par }{\PP \VS{2}¶ Also he
made a
molten
sea of
ten
cubits from
brim to
brim,
round in
compass, and
five
cubits the
height thereof; and a
line of
thirty
cubits did
compass it round
about.\FTNT{*}{{\FR 4.2 }
from…: Heb. from his brim to his brim}
\VS{3}And under it
{\ADD{was}} the
similitude of
oxen, which did
compass it round
about:
ten in a
cubit,
compassing the
sea round
about.
Two
rows of
oxen
{\ADD{were}}
cast, when it was
cast.
\VS{4}It
stood upon
twelve
oxen,
three
looking toward the
north, and
three
looking toward the
west, and
three
looking toward the
south, and
three
looking toward the
east: and the
sea
{\ADD{was set}}
above upon them, and all their hinder
parts
{\ADD{were}}
inward.
\VS{5}And the
thickness of it
{\ADD{was}} an
handbreadth, and the
brim of it like the
work of the
brim of a
cup, with
flowers of
lilies;
{\ADD{and}} it
received and
held
three
thousand
baths.\FTNT{†}{{\FR 4.5 }
with flowers…: or, like a lilyflower}
\par }{\PP \VS{6}¶ He
made also
ten
lavers, and
put
five on the right
hand, and
five on the
left, to
wash in them: such things as they
offered for the burnt
offering they
washed in them; but the
sea
{\ADD{was}} for the
priests to wash
in.\FTNT{‡}{{\FR 4.6 }
such…: Heb. the work of burnt offering}
\VS{7}And he
made
ten
candlesticks of
gold according to their
form, and
set
{\ADD{them}} in the
temple,
five on the right
hand, and
five on the
left.
\VS{8}He
made also
ten
tables, and
placed
{\ADD{them}} in the
temple,
five on the right
side, and
five on the
left. And he
made an
hundred
basons of
gold.\FTNT{§}{{\FR 4.8 }
basons: or, bowls}
\par }{\PP \VS{9}¶ Furthermore he
made the
court of the
priests, and the
great
court, and
doors for the
court, and
overlaid the
doors of them with
brass.
\VS{10}And he
set the
sea on the
right
side of the east
end, over
against the
south.
\par }{\PP \VS{11}And
Huram
made the
pots, and the
shovels, and the
basons. And
Huram
finished the
work that he was to
make for
king
Solomon for the
house of
God;\FTNT{**}{{\FR 4.11 }
basons: or, bowls}\FTNT{††}{{\FR 4.11 }
finished: Heb. finished to make}
\VS{12}{\ADD{To wit}}, the
two
pillars, and the
pommels, and the
chapiters
{\ADD{which were}} on the
top of the two
pillars, and the
two
wreaths to
cover the
two
pommels of the
chapiters which
{\ADD{were}} on the
top of the
pillars;
\VS{13}And
four
hundred
pomegranates on the
two
wreaths;
two
rows of
pomegranates on
each
wreath, to
cover the
two
pommels of the
chapiters which
{\ADD{were}}
upon the
pillars.\FTNT{‡‡}{{\FR 4.13 }
upon: Heb. upon the face of}
\VS{14}He
made also
bases, and
lavers
made he upon the
bases;\FTNT{§§}{{\FR 4.14 }
lavers: or, caldrons}
\VS{15}One
sea, and
twelve
oxen under it.
\VS{16}The
pots also, and the
shovels, and the
fleshhooks, and all their
instruments, did
Huram his
father
make to
king
Solomon for the
house of the
{\ND{
LORD}} of
bright
brass.\FTNT{***}{{\FR 4.16 }
bright: Heb. made bright, or, scoured}
\VS{17}In the
plain of
Jordan did the
king
cast them, in the
clay
ground between
Succoth and
Zeredathah.\FTNT{†††}{{\FR 4.17 }
clay…: Heb. thicknesses of the ground}
\VS{18}Thus
Solomon
made all these
vessels in
great
abundance: for the
weight of the
brass could not be found
out.
\par }{\PP \VS{19}¶ And
Solomon
made all the
vessels that
{\ADD{were for}} the
house of
God, the
golden
altar also, and the
tables whereon the
shewbread
{\ADD{was set;}}
\VS{20}Moreover the
candlesticks with their
lamps, that they should
burn after the
manner
before the
oracle, of
pure
gold;
\VS{21}And the
flowers, and the
lamps, and the
tongs,
{\ADD{made he of}}
gold,
{\ADD{and}} that
perfect
gold;\FTNT{‡‡‡}{{\FR 4.21 }
perfect…: Heb. perfections of gold}
\VS{22}And the
snuffers, and the
basons, and the
spoons, and the
censers,
{\ADD{of}}
pure
gold: and the
entry of the
house, the
inner
doors thereof for the
most
holy
{\ADD{place}}, and the
doors of the
house of the
temple,
{\ADD{were of}}
gold.\FTNT{§§§}{{\FR 4.22 }
basons: or, bowls}

\par }\Chap{5}{\PP \VerseOne{1}Thus all the
work that
Solomon
made for the
house of the
{\ND{
LORD}} was
finished: and
Solomon
brought in
{\ADD{all}} the
things that
David his
father had
dedicated; and the
silver, and the
gold, and all the
instruments,
put he among the
treasures of the
house of
God.
\par }{\PP \VS{2}¶ Then
Solomon
assembled the
elders of
Israel, and all the
heads of the
tribes, the
chief of the
fathers of the
children of
Israel, unto
Jerusalem, to bring
up the
ark of the
covenant of the
{\ND{
LORD}} out of the
city of
David, which
{\ADD{is}}
Zion.
\VS{3}Wherefore all the
men of
Israel
assembled themselves unto the
king in the
feast which
{\ADD{was}} in the
seventh
month.
\VS{4}And all the
elders of
Israel
came; and the
Levites took
up the
ark.
\VS{5}And they brought
up the
ark, and the
tabernacle of the
congregation, and all the
holy
vessels that
{\ADD{were}} in the
tabernacle, these did the
priests
{\ADD{and}} the
Levites bring
up.
\VS{6}Also
king
Solomon, and all the
congregation of
Israel that were
assembled unto him
before the
ark,
sacrificed
sheep and
oxen, which could not be
told nor
numbered for
multitude.
\VS{7}And the
priests brought
in the
ark of the
covenant of the
{\ND{
LORD}} unto his
place, to the
oracle of the
house, into the
most
holy
{\ADD{place, even}} under the
wings of the
cherubims:
\VS{8}For the
cherubims spread
forth
{\ADD{their}}
wings over the
place of the
ark, and the
cherubims
covered the
ark and the
staves thereof
above.
\VS{9}And they drew
out the
staves
{\ADD{of the ark}}, that the
ends of the
staves were
seen from the
ark
before the
oracle; but they were not
seen
without. And there it is unto this
day.\FTNT{*}{{\FR 5.9 }
there…: or, they are there}
\VS{10}{\ADD{There was}} nothing in the
ark save the
two
tables which
Moses
put
{\ADD{therein}} at
Horeb, when the
{\ND{
LORD}}
made
{\ADD{a covenant}} with the
children of
Israel, when they came
out of
Egypt.\FTNT{†}{{\FR 5.10 }
when the
{\ND{LORD}}: or, where, etc}
\par }{\PP \VS{11}¶ And it came to pass, when the
priests were come
out of the
holy
{\ADD{place:}} (for all the
priests
{\ADD{that were}}
present were
sanctified,
{\ADD{and}} did not
{\ADD{then}}
wait by
course:\FTNT{‡}{{\FR 5.11 }
present: Heb. found}
\VS{12}Also the
Levites
{\ADD{which were}} the
singers, all of them of
Asaph, of
Heman, of
Jeduthun, with their
sons and their
brethren,
{\ADD{being}}
arrayed in white
linen, having
cymbals and
psalteries and
harps,
stood at the east
end of the
altar, and with them an
hundred and
twenty
priests
sounding with
trumpets:)
\VS{13}It came even to pass, as the
trumpeters and
singers
{\ADD{were}} as
one, to make
one
sound to be
heard in
praising and
thanking the
{\ND{
LORD}}; and when they lifted
up
{\ADD{their}}
voice with the
trumpets and
cymbals and
instruments of
musick, and
praised the
{\ND{
LORD}},
{\ADD{saying}}, For
{\ADD{he is}}
good; for his
mercy
{\ADD{endureth}} for
ever: that
{\ADD{then}} the
house was
filled with a
cloud,
{\ADD{even}} the
house of the
{\ND{
LORD}};
\VS{14}So that the
priests
could not
stand to
minister by
reason of the
cloud: for the
glory of the
{\ND{
LORD}} had
filled the
house of
God.

\par }\Chap{6}{\PP \VerseOne{1}Then
said
Solomon, The
{\ND{
LORD}} hath
said that he would
dwell in the thick
darkness.
\VS{2}But I have
built an
house of
habitation for thee, and a
place for thy
dwelling for
ever.
\VS{3}And the
king
turned his
face, and
blessed the whole
congregation of
Israel: and all the
congregation of
Israel
stood.
\VS{4}And he
said,
Blessed
{\ADD{be}} the
{\ND{
LORD}}
God of
Israel, who hath with his
hands
fulfilled
{\ADD{that}} which he
spake with his
mouth to my
father
David,
saying,
\VS{5}Since the
day that I brought
forth my
people out of the
land of
Egypt I
chose no
city among all the
tribes of
Israel to
build an
house in, that my
name might be there; neither
chose I any
man to be a
ruler over my
people
Israel:
\VS{6}But I have
chosen
Jerusalem, that my
name might be there; and have
chosen
David to be over my
people
Israel.
\VS{7}Now it was in the
heart of
David my
father to
build an
house for the
name of the
{\ND{
LORD}}
God of
Israel.
\VS{8}But the
{\ND{
LORD}}
said to
David my
father, Forasmuch as it was in thine
heart to
build an
house for my
name, thou didst
well in that it was in thine
heart:
\VS{9}Notwithstanding thou shalt not
build the
house; but thy
son which shall come
forth out of thy
loins, he shall
build the
house for my
name.
\VS{10}The
{\ND{
LORD}} therefore hath
performed his
word that he hath
spoken: for I am risen
up in the room of
David my
father, and am
set on the
throne of
Israel, as the
{\ND{
LORD}}
promised, and have
built the
house for the
name of the
{\ND{
LORD}}
God of
Israel.
\VS{11}And in
it have I
put the
ark, wherein
{\ADD{is}} the
covenant of the
{\ND{
LORD}}, that he
made with the
children of
Israel.
\par }{\PP \VS{12}¶ And he
stood
before the
altar of the
{\ND{
LORD}} in the presence of all the
congregation of
Israel, and spread
forth his
hands:
\VS{13}For
Solomon had
made a
brasen
scaffold, of
five
cubits
long, and
five
cubits
broad, and
three
cubits
high, and had
set it in the
midst of the
court: and upon it he
stood, and kneeled
down upon his
knees before all the
congregation of
Israel, and spread
forth his
hands toward
heaven,\FTNT{*}{{\FR 6.13 }
long…: Heb. the length thereof, etc}
\VS{14}And
said, O
{\ND{
LORD}}
God of
Israel,
{\ADD{there is}} no
God like thee in the
heaven, nor in the
earth; which
keepest
covenant, and
{\ADD{shewest}}
mercy unto thy
servants, that
walk
before thee with all their
hearts:
\VS{15}Thou which hast
kept with thy
servant
David my
father that which thou hast
promised him; and
spakest with thy
mouth, and hast
fulfilled
{\ADD{it}} with thine
hand, as
{\ADD{it is}} this
day.
\VS{16}Now therefore, O
{\ND{
LORD}}
God of
Israel,
keep with thy
servant
David my
father that which thou hast
promised him,
saying, There shall not
fail thee a
man in my
sight to
sit upon the
throne of
Israel; yet
so that thy
children take
heed to their
way to
walk in my
law, as thou hast
walked
before me.\FTNT{†}{{\FR 6.16 }
There…: Heb. There shall not a man be cut off}
\VS{17}Now then, O
{\ND{
LORD}}
God of
Israel, let thy
word be
verified, which thou hast
spoken unto thy
servant
David.
\VS{18}But will
God in very
deed
dwell with
men on the
earth? behold,
heaven and the
heaven of
heavens cannot
contain thee; how much less this
house which I have
built!
\VS{19}Have
respect therefore to the
prayer of thy
servant, and to his
supplication, O
{\ND{
LORD}} my
God, to
hearken unto the
cry and the
prayer which thy
servant
prayeth
before thee:
\VS{20}That thine
eyes may be
open upon this
house
day and
night, upon the
place whereof thou hast
said that thou wouldest
put thy
name there; to
hearken unto the
prayer which thy
servant
prayeth toward this
place.\FTNT{‡}{{\FR 6.20 }
toward…: or, in this place}
\VS{21}Hearken therefore unto the
supplications of thy
servant, and of thy
people
Israel, which they shall
make toward this
place:
hear thou from thy
dwelling
place,
{\ADD{even}} from
heaven; and when thou
hearest,
forgive.\FTNT{§}{{\FR 6.21 }
make: Heb. pray}\FTNT{**}{{\FR 6.21 }
toward…: or, in this place}
\par }{\PP \VS{22}¶ If a
man
sin against his
neighbour, and an
oath be
laid upon him to make him
swear, and the
oath
come
before thine
altar in this
house;\FTNT{††}{{\FR 6.22 }
and an oath…: Heb. and he require an oath of him}
\VS{23}Then
hear thou from
heaven, and
do, and
judge thy
servants, by
requiting the
wicked, by
recompensing his
way upon his own
head; and by
justifying the
righteous, by
giving him according to his
righteousness.
\par }{\PP \VS{24}¶ And if thy
people
Israel be put to the
worse
before the
enemy, because they have
sinned against thee; and shall
return and
confess thy
name, and
pray and make
supplication
before thee in this
house;\FTNT{‡‡}{{\FR 6.24 }
be put…: or, be smitten}\FTNT{§§}{{\FR 6.24 }
in: or, toward}
\VS{25}Then
hear thou from the
heavens, and
forgive the
sin of thy
people
Israel, and bring them
again unto the
land which thou
gavest to them and to their
fathers.
\par }{\PP \VS{26}¶ When the
heaven is shut
up, and there is no
rain, because they have
sinned against thee;
{\ADD{yet}} if they
pray toward this
place, and
confess thy
name, and
turn from their
sin, when thou dost
afflict them;\FTNT{***}{{\FR 6.26 }
toward…: or, in this place}
\VS{27}Then
hear thou from
heaven, and
forgive the
sin of thy
servants, and of thy
people
Israel, when thou hast
taught them the
good
way, wherein they should
walk; and
send
rain upon thy
land, which thou hast
given unto thy
people for an
inheritance.
\par }{\PP \VS{28}¶ If there be
dearth in the
land, if there be
pestilence, if there be
blasting, or
mildew,
locusts, or
caterpillers; if their
enemies
besiege them in the
cities of their
land; whatsoever
sore or whatsoever
sickness
{\ADD{there be:}}\FTNT{†††}{{\FR 6.28 }
in the cities…: Heb. in the land of their gates}
\VS{29}{\ADD{Then}} what
prayer
{\ADD{or}} what
supplication soever shall be made of any
man, or of all thy
people
Israel, when every
one shall
know his own
sore and his own
grief, and shall spread
forth his
hands in this
house:\FTNT{‡‡‡}{{\FR 6.29 }
in: or, toward}\FTNT{§§§}{{\FR 6.29 }
toward…: or, in this place}
\VS{30}Then
hear thou from
heaven thy
dwelling
place, and
forgive, and
render unto every
man according unto all his
ways, whose
heart thou
knowest; (for thou only
knowest the
hearts of the
children of
men:)
\VS{31}That they may
fear thee, to
walk in thy
ways, so long
as they
live in the
land which thou
gavest unto our
fathers.\FTNT{*}{{\FR 6.31 }
so long…: Heb. all the days which}\FTNT{†}{{\FR 6.31 }
in the land: Heb. upon the face of the land}
\par }{\PP \VS{32}¶ Moreover concerning the
stranger, which is not of thy
people
Israel, but is
come from a
far
country for thy
great
name’s sake, and thy
mighty
hand, and thy stretched
out
arm; if they
come and
pray in this
house;\FTNT{‡}{{\FR 6.32 }
toward…: or, in this place}
\VS{33}Then
hear thou from the
heavens,
{\ADD{even}} from thy
dwelling
place, and
do according to all that the
stranger
calleth to thee for; that all
people of the
earth may
know thy
name, and
fear thee, as
{\ADD{doth}} thy
people
Israel, and may
know that this
house which I have
built is
called by thy
name.\FTNT{§}{{\FR 6.33 }
this…: Heb. thy name is called upon this house}
\VS{34}If thy
people go
out to
war against their
enemies by the
way that thou shalt
send them, and they
pray unto thee
toward this
city which thou hast
chosen, and the
house which I have
built for thy
name;
\VS{35}Then
hear thou from the
heavens their
prayer and their
supplication, and
maintain their
cause.\FTNT{**}{{\FR 6.35 }
cause: or, right}
\VS{36}If they
sin against thee, (for
{\ADD{there is}} no
man which
sinneth not,) and thou be
angry with them, and
deliver them over
before
{\ADD{their}}
enemies, and they carry them
away
captives unto a
land far
off or
near;\FTNT{††}{{\FR 6.36 }
they carry them away captives: Heb. they that take them captives carry them away}
\VS{37}Yet
{\ADD{if}} they
bethink themselves in the
land whither they are carried
captive, and
turn and
pray unto thee in the
land of their
captivity,
saying, We have
sinned, we have done
amiss, and have dealt
wickedly;\FTNT{‡‡}{{\FR 6.37 }
bethink…: Heb. bring back to their heart}
\VS{38}If they
return to thee with all their
heart and with all their
soul in the
land of their
captivity, whither they have carried them
captives, and
pray
toward their
land, which thou
gavest unto their
fathers, and
{\ADD{toward}} the
city which thou hast
chosen, and toward the
house which I have
built for thy
name:
\VS{39}Then
hear thou from the
heavens,
{\ADD{even}} from thy
dwelling
place, their
prayer and their
supplications, and
maintain their
cause, and
forgive thy
people which have
sinned against thee.\FTNT{§§}{{\FR 6.39 }
cause: or, right}
\VS{40}Now, my
God, let, I beseech thee, thine
eyes be
open, and
{\ADD{let}} thine
ears
{\ADD{be}}
attent unto the
prayer
{\ADD{that is made}} in this
place.\FTNT{***}{{\FR 6.40 }
unto…: Heb. to the prayer of this place}
\VS{41}Now therefore
arise, O
{\ND{
LORD}}
God, into thy
resting place, thou, and the
ark of thy
strength: let thy
priests, O
{\ND{
LORD}}
God, be
clothed with
salvation, and let thy
saints
rejoice in
goodness.
\VS{42}O
{\ND{
LORD}}
God, turn not
away the
face of thine
anointed:
remember the
mercies of
David thy
servant.

\par }\Chap{7}{\PP \VerseOne{1}Now when
Solomon had made an
end of
praying, the
fire came
down from
heaven, and
consumed the burnt
offering and the
sacrifices; and the
glory of the
{\ND{
LORD}}
filled the
house.
\VS{2}And the
priests
could not
enter into the
house of the
{\ND{
LORD}}, because the
glory of the
{\ND{
LORD}} had
filled the
{\ND{
LORD’s}}
house.
\VS{3}And when all the
children of
Israel
saw how the
fire came
down, and the
glory of the
{\ND{
LORD}} upon the
house, they
bowed themselves with their
faces to the
ground upon the
pavement, and
worshipped, and
praised the
{\ND{
LORD}},
{\ADD{saying}}, For
{\ADD{he is}}
good; for his
mercy
{\ADD{endureth}} for
ever.
\par }{\PP \VS{4}¶ Then the
king and all the
people
offered
sacrifices
before the
{\ND{
LORD}}.
\VS{5}And
king
Solomon
offered a
sacrifice of
twenty and
two
thousand
oxen, and an
hundred and
twenty
thousand
sheep: so the
king and all the
people
dedicated the
house of
God.
\VS{6}And the
priests
waited on their
offices: the
Levites also with
instruments of
musick of the
{\ND{
LORD}}, which
David the
king had
made to
praise the
{\ND{
LORD}}, because his
mercy
{\ADD{endureth}} for
ever, when
David
praised by their
ministry; and the
priests sounded
trumpets before them, and all
Israel
stood.\FTNT{*}{{\FR 7.6 }
by their…: Heb. by their hand}
\VS{7}Moreover
Solomon
hallowed the
middle of the
court that
{\ADD{was}}
before the
house of the
{\ND{
LORD}}: for there he
offered burnt
offerings, and the
fat of the peace
offerings, because the
brasen
altar which
Solomon had
made was not
able to
receive the burnt
offerings, and the meat
offerings, and the
fat.
\par }{\PP \VS{8}¶ Also at the same
time
Solomon
kept the
feast
seven
days, and all
Israel with him, a
very
great
congregation, from the entering
in of
Hamath unto the
river of
Egypt.
\VS{9}And in the
eighth
day they
made a solemn
assembly: for they
kept the
dedication of the
altar
seven
days, and the
feast
seven
days.\FTNT{†}{{\FR 7.9 }
a solemn…: Heb. a restraint}
\VS{10}And on the
three and
twentieth
day of the
seventh
month he
sent the
people
away into their
tents,
glad and
merry in
heart for the
goodness that the
{\ND{
LORD}} had
shewed unto
David, and to
Solomon, and to
Israel his
people.
\VS{11}Thus
Solomon
finished the
house of the
{\ND{
LORD}}, and the
king’s
house: and all that
came into
Solomon’s
heart to
make in the
house of the
{\ND{
LORD}}, and in his own
house, he prosperously
effected.
\par }{\PP \VS{12}¶ And the
{\ND{
LORD}}
appeared to
Solomon by
night, and
said unto him, I have
heard thy
prayer, and have
chosen this
place to myself for an
house of
sacrifice.
\VS{13}If I shut
up
heaven that there be no
rain, or if I
command the
locusts to
devour the
land, or if I
send
pestilence among my
people;
\VS{14}If my
people, which are
called by my
name, shall
humble themselves, and
pray, and
seek my
face, and
turn from their
wicked
ways; then will I
hear from
heaven, and will
forgive their
sin, and will
heal their
land.\FTNT{‡}{{\FR 7.14 }
which…: Heb. upon whom my name is called}
\VS{15}Now mine
eyes shall be
open, and mine
ears
attent unto the
prayer
{\ADD{that is made}} in this
place.\FTNT{§}{{\FR 7.15 }
unto…: Heb. to the prayer of this place}
\VS{16}For now have I
chosen and
sanctified this
house, that my
name may be there
for
ever: and mine
eyes and mine
heart shall be there
perpetually.
\VS{17}And as for thee, if thou wilt
walk
before me, as
David thy
father
walked, and
do according to all that I have
commanded thee, and shalt
observe my
statutes and my
judgments;
\VS{18}Then will I
stablish the
throne of thy
kingdom, according as I have
covenanted with
David thy
father,
saying, There shall not
fail thee a
man
{\ADD{to be}}
ruler in
Israel.\FTNT{**}{{\FR 7.18 }
There…: Heb. There shall not be cut off to thee}
\VS{19}But if ye turn
away, and
forsake my
statutes and my
commandments, which I have
set
before you, and shall
go and
serve
other
gods, and
worship them;
\VS{20}Then will I pluck them up by the
roots out of my
land which I have
given them; and this
house, which I have
sanctified for my
name, will I cast
out of my
sight, and will
make it
{\ADD{to be}} a
proverb and a
byword among all
nations.
\VS{21}And this
house, which is
high, shall be an
astonishment to every one that
passeth by it; so that he shall
say, Why hath the
{\ND{
LORD}}
done thus unto this
land, and unto this
house?
\VS{22}And it shall be
answered, Because they
forsook the
{\ND{
LORD}}
God of their
fathers, which brought them
forth out of the
land of
Egypt, and laid
hold on
other
gods, and
worshipped them, and
served them: therefore hath he
brought all this
evil upon them.

\par }\Chap{8}{\PP \VerseOne{1}And it came to pass at the
end of
twenty
years, wherein
Solomon had
built the
house of the
{\ND{
LORD}}, and his own
house,
\VS{2}That the
cities which
Huram had
restored to
Solomon,
Solomon
built them, and caused the
children of
Israel to
dwell there.
\VS{3}And
Solomon
went to
Hamath-zobah, and
prevailed against it.
\VS{4}And he
built
Tadmor in the
wilderness, and all the
store
cities, which he
built in
Hamath.
\VS{5}Also he
built
Beth-horon the
upper, and
Beth-horon the
nether,
fenced
cities, with
walls,
gates, and
bars;
\VS{6}And
Baalath, and all the
store
cities that
Solomon had, and all the
chariot
cities, and the
cities of the
horsemen, and all that
Solomon
desired to
build in
Jerusalem, and in
Lebanon, and throughout all the
land of his
dominion.\FTNT{*}{{\FR 8.6 }
all that…: Heb. all the desire of Solomon which he desired to build}
\par }{\PP \VS{7}¶
{\ADD{As for}} all the
people
{\ADD{that were}}
left of the
Hittites, and the
Amorites, and the
Perizzites, and the
Hivites, and the
Jebusites, which
{\ADD{were}} not of
Israel,
\VS{8}{\ADD{But}} of their
children, who were
left
after them in the
land, whom the
children of
Israel
consumed not, them did
Solomon make to
pay
tribute until this
day.
\VS{9}But of the
children of
Israel did
Solomon
make no
servants for his
work; but they
{\ADD{were}}
men of
war, and
chief of his
captains, and
captains of his
chariots and
horsemen.
\VS{10}And these
{\ADD{were}} the
chief of
king
Solomon’s
officers,
{\ADD{even}} two
hundred and
fifty, that bare
rule over the
people.
\par }{\PP \VS{11}¶ And
Solomon brought
up the
daughter of
Pharaoh out of the
city of
David unto the
house that he had
built for her: for he
said, My
wife shall not
dwell in the
house of
David
king of
Israel, because
{\ADD{the places are}}
holy, whereunto the
ark of the
{\ND{
LORD}} hath
come.\FTNT{†}{{\FR 8.11 }
holy: Heb. holiness}
\par }{\PP \VS{12}¶ Then
Solomon
offered burnt
offerings unto the
{\ND{
LORD}} on the
altar of the
{\ND{
LORD}}, which he had
built
before the
porch,
\VS{13}Even after a certain
rate every
day,
offering according to the
commandment of
Moses, on the
sabbaths, and on the new
moons, and on the solemn
feasts,
three
times in the
year,
{\ADD{even}} in the
feast of unleavened
bread, and in the
feast of
weeks, and in the
feast of
tabernacles.
\par }{\PP \VS{14}¶ And he
appointed, according to the
order of
David his
father, the
courses of the
priests to their
service, and the
Levites to their
charges, to
praise and
minister before the
priests, as the
duty of every
day
required: the
porters also by their
courses at every
gate: for so had
David the
man of
God
commanded.\FTNT{‡}{{\FR 8.14 }
so had David…: Heb. so was the commandment of David the man of God}
\VS{15}And they
departed not from the
commandment of the
king unto the
priests and
Levites concerning any
matter, or concerning the
treasures.
\VS{16}Now all the
work of
Solomon was
prepared unto the
day of the
foundation of the
house of the
{\ND{
LORD}}, and until it was
finished.
{\ADD{So}} the
house of the
{\ND{
LORD}} was
perfected.
\par }{\PP \VS{17}¶ Then
went
Solomon to
Ezion-geber, and to
Eloth, at the
sea
side in the
land of
Edom.\FTNT{§}{{\FR 8.17 }
Eloth: also called, Elath}
\VS{18}And
Huram
sent him by the
hands of his
servants
ships, and
servants that had
knowledge of the
sea; and they
went with the
servants of
Solomon to
Ophir, and
took thence
four
hundred and
fifty
talents of
gold, and
brought
{\ADD{them}} to
king
Solomon.

\par }\Chap{9}{\PP \VerseOne{1}And when the
queen of
Sheba
heard of the
fame of
Solomon, she
came to
prove
Solomon with hard
questions at
Jerusalem, with a
very
great
company, and
camels that
bare
spices, and
gold in
abundance, and
precious
stones: and when she was
come to
Solomon, she
communed with him of all that was in her
heart.
\VS{2}And
Solomon
told her all her
questions: and there was nothing
hid from
Solomon which he
told her not.
\VS{3}And when the
queen of
Sheba had
seen the
wisdom of
Solomon, and the
house that he had
built,
\VS{4}And the
meat of his
table, and the
sitting of his
servants, and the
attendance of his
ministers, and their
apparel; his
cupbearers also, and their
apparel; and his
ascent by which he went
up into the
house of the
{\ND{
LORD}}; there was no more
spirit in her.\FTNT{*}{{\FR 9.4 }
cupbearers: or, butlers}
\VS{5}And she
said to the
king,
{\ADD{It was}} a
true
report which I
heard in mine own
land of thine
acts, and of thy
wisdom:\FTNT{†}{{\FR 9.5 }
report: Heb. word}\FTNT{‡}{{\FR 9.5 }
acts: or, sayings}
\VS{6}Howbeit I
believed not their
words, until I
came, and mine
eyes had
seen
{\ADD{it:}} and, behold, the one
half of the
greatness of thy
wisdom was not
told me:
{\ADD{for}} thou
exceedest the
fame that I
heard.
\VS{7}Happy
{\ADD{are}} thy
men, and
happy
{\ADD{are}} these thy
servants, which
stand
continually
before thee, and
hear thy
wisdom.
\VS{8}Blessed be the
{\ND{
LORD}} thy
God, which
delighted in thee to
set thee on his
throne,
{\ADD{to be}}
king for the
{\ND{
LORD}} thy
God: because thy
God
loved
Israel, to
establish them for
ever, therefore
made he thee
king over them, to
do
judgment and
justice.
\VS{9}And she
gave the
king an
hundred and
twenty
talents of
gold, and of
spices
great
abundance, and
precious
stones: neither was there any such
spice as the
queen of
Sheba
gave
king
Solomon.
\VS{10}And the
servants also of
Huram, and the
servants of
Solomon, which
brought
gold from
Ophir,
brought
algum
trees and
precious
stones.\FTNT{§}{{\FR 9.10 }
algum…: also called, almug trees}
\VS{11}And the
king
made
{\ADD{of}} the
algum
trees
terraces to the
house of the
{\ND{
LORD}}, and to the
king’s
palace, and
harps and
psalteries for
singers: and there were none
such
seen
before in the
land of
Judah.\FTNT{**}{{\FR 9.11 }
terraces: or, stairs: Heb. highways}
\VS{12}And
king
Solomon
gave to the
queen of
Sheba all her
desire, whatsoever she
asked, beside
{\ADD{that}} which she had
brought unto the
king. So she
turned, and went
away to her own
land, she and her
servants.
\par }{\PP \VS{13}¶ Now the
weight of
gold that
came to
Solomon in
one
year was
six
hundred and
threescore and
six
talents of
gold;
\VS{14}Beside
{\ADD{that which}}
chapmen and
merchants
brought. And all the
kings of
Arabia and
governors of the
country
brought
gold and
silver to
Solomon.\FTNT{††}{{\FR 9.14 }
governors: or, captains}
\par }{\PP \VS{15}¶ And
king
Solomon
made two
hundred
targets
{\ADD{of}}
beaten
gold:
six
hundred
{\ADD{shekels}} of
beaten
gold
went to
one
target.
\VS{16}And
three
hundred
shields
{\ADD{made he of}}
beaten
gold:
three
hundred
{\ADD{shekels}} of
gold
went to
one
shield. And the
king
put them in the
house of the
forest of
Lebanon.
\VS{17}Moreover the
king
made a
great
throne of
ivory, and
overlaid it with
pure
gold.
\VS{18}And
{\ADD{there were}}
six
steps to the
throne, with a
footstool of
gold,
{\ADD{which were}}
fastened to the
throne, and
stays on each side of the
sitting
place, and
two
lions
standing
by the
stays:\FTNT{‡‡}{{\FR 9.18 }
stays: Heb. hands}
\VS{19}And
twelve
lions
stood there on the one side and on the other upon the
six
steps. There was not the like
made in any
kingdom.
\par }{\PP \VS{20}¶ And all the
drinking
vessels of
king
Solomon
{\ADD{were of}}
gold, and all the
vessels of the
house of the
forest of
Lebanon
{\ADD{were of}}
pure
gold: none
{\ADD{were of}}
silver; it was
{\ADD{not}}
any thing
accounted of in the
days of
Solomon.\FTNT{§§}{{\FR 9.20 }
pure: Heb. shut up}\FTNT{***}{{\FR 9.20 }
none were of silver: or, there was no silver in them}
\VS{21}For the
king’s
ships
went to
Tarshish with the
servants of
Huram: every
three
years
once
came the
ships of
Tarshish
bringing
gold, and
silver,
ivory, and
apes, and
peacocks.\FTNT{†††}{{\FR 9.21 }
ivory: or, elephants’ teeth}
\VS{22}And
king
Solomon
passed all the
kings of the
earth in
riches and
wisdom.
\par }{\PP \VS{23}¶ And all the
kings of the
earth
sought the
presence of
Solomon, to
hear his
wisdom, that
God had
put in his
heart.
\VS{24}And they
brought every
man his
present,
vessels of
silver, and
vessels of
gold, and
raiment,
harness, and
spices,
horses, and
mules, a
rate
year by
year.
\par }{\PP \VS{25}¶ And
Solomon had
four
thousand
stalls for
horses and
chariots, and
twelve
thousand
horsemen; whom he
bestowed in the
chariot
cities, and with the
king at
Jerusalem.
\par }{\PP \VS{26}¶ And he
reigned over all the
kings from the
river even unto the
land of the
Philistines, and to the
border of
Egypt.\FTNT{‡‡‡}{{\FR 9.26 }
river: that is, Euphrates}
\VS{27}And the
king
made
silver in
Jerusalem as
stones, and cedar
trees
made he as the sycomore
trees that
{\ADD{are}} in the low
plains in
abundance.\FTNT{§§§}{{\FR 9.27 }
made silver: Heb. gave silver}
\VS{28}And they
brought unto
Solomon
horses out of
Egypt, and out of all
lands.
\par }{\PP \VS{29}¶ Now the
rest of the
acts of
Solomon,
first and
last,
{\ADD{are}} they not
written in the
book of
Nathan the
prophet, and in the
prophecy of
Ahijah the
Shilonite, and in the
visions of
Iddo the
seer against
Jeroboam the
son of
Nebat?\FTNT{*}{{\FR 9.29 }
book: Heb. words}
\VS{30}And
Solomon
reigned in
Jerusalem over all
Israel
forty
years.
\VS{31}And
Solomon
slept with his
fathers, and he was
buried in the
city of
David his
father: and
Rehoboam his
son
reigned in his stead.

\par }\Chap{10}{\PP \VerseOne{1}And
Rehoboam
went to
Shechem: for to
Shechem were all
Israel
come to make him
king.
\VS{2}And it came to pass, when
Jeroboam the
son of
Nebat, who
{\ADD{was}} in
Egypt, whither he had
fled from the
presence of
Solomon the
king,
heard
{\ADD{it}}, that
Jeroboam
returned out of
Egypt.
\VS{3}And they
sent and
called him. So
Jeroboam and all
Israel
came and
spake to
Rehoboam,
saying,
\VS{4}Thy
father made our
yoke
grievous: now therefore
ease thou somewhat the
grievous
servitude of thy
father, and his
heavy
yoke that he
put upon us, and we will
serve thee.
\VS{5}And he
said unto them, Come
again unto me after
three
days. And the
people
departed.
\par }{\PP \VS{6}¶ And
king
Rehoboam took
counsel with the
old men that had
stood
before
Solomon his
father while he yet
lived,
saying,
What counsel
give ye
{\ADD{me}} to
return
answer to this
people?
\VS{7}And they
spake unto him,
saying, If thou be
kind to this
people, and
please them, and
speak
good
words to them, they will be thy
servants for
ever.
\VS{8}But he
forsook the
counsel which the old
men
gave him, and took
counsel with the young
men that were brought
up with him, that
stood
before him.
\VS{9}And he
said unto them, What
advice give ye that we may
return
answer to this
people, which have
spoken to me,
saying,
Ease somewhat the
yoke that thy
father did
put upon us?
\VS{10}And the young
men that were brought
up with him
spake unto him,
saying, Thus shalt thou
answer the
people that
spake unto thee,
saying, Thy
father
made our
yoke
heavy, but make thou
{\ADD{it}} somewhat
lighter for us; thus shalt thou
say unto them, My
little
{\ADD{finger}} shall be
thicker than my
father’s
loins.
\VS{11}For
whereas my
father
put a
heavy
yoke upon you, I will put
more to your
yoke: my
father
chastised you with
whips, but I
{\ADD{will chastise you}} with
scorpions.\FTNT{*}{{\FR 10.11 }
my father put: Heb. my father laded}
\par }{\PP \VS{12}So
Jeroboam and all the
people
came to
Rehoboam on the
third
day, as the
king
bade,
saying, Come
again to me on the
third
day.
\VS{13}And the
king
answered them
roughly; and
king
Rehoboam
forsook the
counsel of the old
men,
\VS{14}And
answered them after the
advice of the young
men,
saying, My
father made your
yoke
heavy, but I will
add thereto: my
father
chastised you with
whips, but I
{\ADD{will chastise you}} with
scorpions.
\VS{15}So the
king
hearkened not unto the
people: for the
cause was of
God, that the
{\ND{
LORD}} might
perform his
word, which he
spake by the
hand of
Ahijah the
Shilonite to
Jeroboam the
son of
Nebat.
\par }{\PP \VS{16}¶ And when all
Israel
{\ADD{
saw}} that the
king would not
hearken unto them, the
people
answered the
king,
saying, What
portion have we in
David? and
{\ADD{we have}} none
inheritance in the
son of
Jesse: every
man to your
tents, O
Israel:
{\ADD{and}} now,
David,
see to thine own
house. So all
Israel
went to their
tents.
\VS{17}But
{\ADD{as for}} the
children of
Israel that
dwelt in the
cities of
Judah,
Rehoboam
reigned over them.
\VS{18}Then
king
Rehoboam
sent
Hadoram that
{\ADD{was}} over the
tribute; and the
children of
Israel
stoned him with
stones, that he
died. But
king
Rehoboam made
speed to get him
up to
{\ADD{his}}
chariot, to
flee to
Jerusalem.\FTNT{†}{{\FR 10.18 }
made speed: Heb. strengthened himself}
\VS{19}And
Israel
rebelled against the
house of
David unto this
day.

\par }\Chap{11}{\PP \VerseOne{1}And when
Rehoboam was
come to
Jerusalem, he
gathered of the
house of
Judah and
Benjamin an
hundred and
fourscore
thousand
chosen
{\ADD{men}}, which were
warriors, to
fight against
Israel, that he might
bring the
kingdom
again to
Rehoboam.
\VS{2}But the
word of the
{\ND{
LORD}} came to
Shemaiah the
man of
God,
saying,
\VS{3}Speak unto
Rehoboam the
son of
Solomon,
king of
Judah, and to all
Israel in
Judah and
Benjamin,
saying,
\VS{4}Thus
saith the
{\ND{
LORD}}, Ye shall not go
up, nor
fight against your
brethren:
return every
man to his
house: for this
thing is
done of me. And they
obeyed the
words of the
{\ND{
LORD}}, and
returned from
going against
Jeroboam.
\par }{\PP \VS{5}¶ And
Rehoboam
dwelt in
Jerusalem, and
built
cities for
defence in
Judah.
\VS{6}He
built even
Beth-lehem, and
Etam, and
Tekoa,
\VS{7}And
Beth-zur, and
Shoco, and
Adullam,
\VS{8}And
Gath, and
Mareshah, and
Ziph,
\VS{9}And
Adoraim, and
Lachish, and
Azekah,
\VS{10}And
Zorah, and
Aijalon, and
Hebron, which
{\ADD{are}} in
Judah and in
Benjamin
fenced
cities.
\VS{11}And he
fortified the strong
holds, and
put
captains in them, and
store of
victual, and of
oil and
wine.
\VS{12}And in every several
city
{\ADD{he put}}
shields and
spears, and made them
exceeding
strong, having
Judah and
Benjamin on his side.
\par }{\PP \VS{13}¶ And the
priests and the
Levites that
{\ADD{were}} in all
Israel
resorted to him out of all their
coasts.\FTNT{*}{{\FR 11.13 }
resorted…: Heb. presented themselves to him}
\VS{14}For the
Levites
left their
suburbs and their
possession, and
came to
Judah and
Jerusalem: for
Jeroboam and his
sons had cast them
off from executing the
priest’s office unto the
{\ND{
LORD}}:
\VS{15}And he
ordained him
priests for the high
places, and for the
devils, and for the
calves which he had
made.
\VS{16}And
after them out of all the
tribes of
Israel such as
set their
hearts to
seek the
{\ND{
LORD}}
God of
Israel
came to
Jerusalem, to
sacrifice unto the
{\ND{
LORD}}
God of their
fathers.
\VS{17}So they
strengthened the
kingdom of
Judah, and made
Rehoboam the
son of
Solomon
strong,
three
years: for
three
years they
walked in the
way of
David and
Solomon.
\par }{\PP \VS{18}¶ And
Rehoboam
took him
Mahalath the
daughter of
Jerimoth the
son of
David to
wife,
{\ADD{and}}
Abihail the
daughter of
Eliab the
son of
Jesse;
\VS{19}Which
bare him
children;
Jeush, and
Shamariah, and
Zaham.
\VS{20}And
after her he
took
Maachah the
daughter of
Absalom; which
bare him
Abijah, and
Attai, and
Ziza, and
Shelomith.\FTNT{†}{{\FR 11.20 }
Maachah: she is also called Michaiah the daughter of Uriel}
\VS{21}And
Rehoboam
loved
Maachah the
daughter of
Absalom above all his
wives and his
concubines: (for he
took
eighteen
wives, and
threescore
concubines; and
begat
twenty and
eight
sons, and
threescore
daughters.)
\VS{22}And
Rehoboam
made
Abijah the
son of
Maachah the
chief,
{\ADD{to be}}
ruler among his
brethren: for
{\ADD{he thought}} to make him
king.
\VS{23}And he dealt
wisely, and
dispersed of all his
children throughout all the
countries of
Judah and
Benjamin, unto every
fenced
city: and he
gave them
victual in
abundance. And he
desired
many
wives.\FTNT{‡}{{\FR 11.23 }
many wives: Heb. a multitude of wives}

\par }\Chap{12}{\PP \VerseOne{1}And it came to pass, when
Rehoboam had
established the
kingdom, and had
strengthened himself, he
forsook the
law of the
{\ND{
LORD}}, and all
Israel with him.
\VS{2}And it came to pass,
{\ADD{that}} in the
fifth
year of
king
Rehoboam
Shishak
king of
Egypt came
up against
Jerusalem, because they had
transgressed against the
{\ND{
LORD}},
\VS{3}With twelve
hundred
chariots, and
threescore
thousand
horsemen: and the
people
{\ADD{were}} without
number that
came with him out of
Egypt; the
Lubims, the
Sukkiims, and the
Ethiopians.
\VS{4}And he
took the
fenced
cities which
{\ADD{pertained}} to
Judah, and
came to
Jerusalem.
\par }{\PP \VS{5}¶ Then
came
Shemaiah the
prophet to
Rehoboam, and
{\ADD{to}} the
princes of
Judah, that were gathered
together to
Jerusalem
because of
Shishak, and
said unto them, Thus
saith the
{\ND{
LORD}}, Ye have
forsaken me, and therefore have I
also
left you in the
hand of
Shishak.
\VS{6}Whereupon the
princes of
Israel and the
king
humbled themselves; and they
said, The
{\ND{
LORD}}
{\ADD{is}}
righteous.
\VS{7}And when the
{\ND{
LORD}}
saw that they
humbled themselves, the
word of the
{\ND{
LORD}} came to
Shemaiah,
saying, They have
humbled themselves;
{\ADD{therefore}} I will not
destroy them, but I will
grant them
some
deliverance; and my
wrath shall not be poured
out upon
Jerusalem by the
hand of
Shishak.\FTNT{*}{{\FR 12.7 }
some: or, a little while}
\VS{8}Nevertheless they shall be his
servants; that they may
know my
service, and the
service of the
kingdoms of the
countries.
\VS{9}So
Shishak
king of
Egypt came
up against
Jerusalem, and took
away the
treasures of the
house of the
{\ND{
LORD}}, and the
treasures of the
king’s
house; he
took all: he carried
away also the
shields of
gold which
Solomon had
made.
\VS{10}Instead of which
king
Rehoboam
made
shields of
brass, and
committed
{\ADD{them}} to the
hands of the
chief of the
guard, that
kept the
entrance of the
king’s
house.
\VS{11}And
when the
king
entered into the
house of the
{\ND{
LORD}}, the
guard
came and
fetched them, and brought them
again into the
guard
chamber.
\VS{12}And when he
humbled himself, the
wrath of the
{\ND{
LORD}}
turned from him, that he would not
destroy
{\ADD{him}}
altogether: and also in
Judah
things went
well.\FTNT{†}{{\FR 12.12 }
and also…: or, and yet in Judah there were good things}
\par }{\PP \VS{13}¶ So
king
Rehoboam
strengthened himself in
Jerusalem, and
reigned: for
Rehoboam
{\ADD{was}}
one and
forty
years
old when he began to
reign, and he
reigned
seventeen
years in
Jerusalem, the
city which the
{\ND{
LORD}} had
chosen out of all the
tribes of
Israel, to
put his
name there. And his
mother’s
name
{\ADD{was}}
Naamah an
Ammonitess.
\VS{14}And he
did
evil, because he
prepared not his
heart to
seek the
{\ND{
LORD}}.\FTNT{‡}{{\FR 12.14 }
prepared: or, fixed}
\VS{15}Now the
acts of
Rehoboam,
first and
last,
{\ADD{are}} they not
written in the
book of
Shemaiah the
prophet, and of
Iddo the
seer concerning
genealogies? And
{\ADD{there were}}
wars between
Rehoboam and
Jeroboam
continually.\FTNT{§}{{\FR 12.15 }
book: Heb. words}
\VS{16}And
Rehoboam
slept with his
fathers, and was
buried in the
city of
David: and
Abijah his
son
reigned in his stead.\FTNT{**}{{\FR 12.16 }
Abijah: also called, Abijam}

\par }\Chap{13}{\PP \VerseOne{1}Now in the
eighteenth
year of
king
Jeroboam began
Abijah to
reign over
Judah.
\VS{2}He
reigned
three
years in
Jerusalem. His
mother’s
name also
{\ADD{was}}
Michaiah the
daughter of
Uriel of
Gibeah. And there was
war between
Abijah and
Jeroboam.
\VS{3}And
Abijah
set the
battle in
array with an
army of valiant
men of
war,
{\ADD{even}}
four
hundred
thousand
chosen
men:
Jeroboam also set the
battle in
array against him with
eight
hundred
thousand
chosen
men,
{\ADD{being}} mighty
men of
valour.\FTNT{*}{{\FR 13.3 }
Abijah set: Heb. Abijah bound together}
\par }{\PP \VS{4}¶ And
Abijah stood
up upon
mount
Zemaraim, which
{\ADD{is}} in
mount
Ephraim, and
said,
Hear me, thou
Jeroboam, and all
Israel;
\VS{5}Ought ye not to
know that the
{\ND{
LORD}}
God of
Israel
gave the
kingdom over
Israel to
David for
ever,
{\ADD{even}} to him and to his
sons by a
covenant of
salt?
\VS{6}Yet
Jeroboam the
son of
Nebat, the
servant of
Solomon the
son of
David, is risen
up, and hath
rebelled against his
lord.
\VS{7}And there are
gathered unto him
vain
men, the
children of
Belial, and have
strengthened themselves against
Rehoboam the
son of
Solomon, when
Rehoboam was
young and
tenderhearted, and could not
withstand
them.
\VS{8}And now ye
think to
withstand the
kingdom of the
{\ND{
LORD}} in the
hand of the
sons of
David; and ye
{\ADD{be}} a
great
multitude, and
{\ADD{there are}} with you
golden
calves, which
Jeroboam
made you for
gods.
\VS{9}Have ye not cast
out the
priests of the
{\ND{
LORD}}, the
sons of
Aaron, and the
Levites, and have
made you
priests after the manner of the
nations of
{\ADD{other}}
lands? so that whosoever
cometh to
consecrate himself with a
young
bullock and
seven
rams,
{\ADD{the same}} may be a
priest of
{\ADD{them that are}}
no
gods.\FTNT{†}{{\FR 13.9 }
to consecrate…: Heb. to fill his hand}
\VS{10}But as for us, the
{\ND{
LORD}}
{\ADD{is}} our
God, and we have not
forsaken him; and the
priests, which
minister unto the
{\ND{
LORD}},
{\ADD{are}} the
sons of
Aaron, and the
Levites
{\ADD{wait}} upon
{\ADD{their}}
business:
\VS{11}And they
burn unto the
{\ND{
LORD}} every
morning and every
evening burnt
sacrifices and
sweet
incense: the
shewbread also
{\ADD{set they in order}} upon the
pure
table; and the
candlestick of
gold with the
lamps thereof, to
burn every
evening: for we
keep the
charge of the
{\ND{
LORD}} our
God; but ye have
forsaken him.
\VS{12}And, behold,
God himself
{\ADD{is}} with us for
{\ADD{our}}
captain, and his
priests with
sounding
trumpets to cry
alarm against you. O
children of
Israel,
fight ye not against the
{\ND{
LORD}}
God of your
fathers; for ye shall not
prosper.
\par }{\PP \VS{13}¶ But
Jeroboam caused an
ambushment to
come
about
behind them: so they were
before
Judah, and the
ambushment
{\ADD{was}}
behind them.
\VS{14}And when
Judah looked
back, behold, the
battle
{\ADD{was}}
before and
behind: and they
cried unto the
{\ND{
LORD}}, and the
priests
sounded with the
trumpets.
\VS{15}Then the
men of
Judah gave a
shout: and as the
men of
Judah
shouted, it came to pass, that
God
smote
Jeroboam and all
Israel
before
Abijah and
Judah.
\VS{16}And the
children of
Israel
fled
before
Judah: and
God
delivered them into their
hand.
\VS{17}And
Abijah and his
people
slew them with a
great
slaughter: so there fell
down
slain of
Israel
five
hundred
thousand
chosen
men.
\VS{18}Thus the
children of
Israel were
brought under at that
time, and the
children of
Judah
prevailed, because they
relied upon the
{\ND{
LORD}}
God of their
fathers.
\VS{19}And
Abijah
pursued
after
Jeroboam, and
took
cities from him,
Beth-el with the
towns thereof, and
Jeshanah with the
towns thereof, and
Ephrain with the
towns thereof.
\VS{20}Neither did
Jeroboam
recover
strength again in the
days of
Abijah: and the
{\ND{
LORD}}
struck him, and he
died.
\par }{\PP \VS{21}¶ But
Abijah waxed
mighty, and
married
fourteen
wives, and
begat
twenty and
two
sons, and
sixteen
daughters.
\VS{22}And the
rest of the
acts of
Abijah, and his
ways, and his
sayings,
{\ADD{are}}
written in the
story of the
prophet
Iddo.

\par }\Chap{14}{\PP \VerseOne{1}So
Abijah
slept with his
fathers, and they
buried him in the
city of
David: and
Asa his
son
reigned in his stead. In his
days the
land was
quiet
ten
years.
\VS{2}And
Asa
did
{\ADD{that which was}}
good and
right in the
eyes of the
{\ND{
LORD}} his
God:
\VS{3}For he took
away the
altars of the
strange
{\ADD{gods}}, and the high
places, and brake
down the
images, and cut
down the
groves:\FTNT{*}{{\FR 14.3 }
images: Heb. statues}
\VS{4}And
commanded
Judah to
seek the
{\ND{
LORD}}
God of their
fathers, and to
do the
law and the
commandment.
\VS{5}Also he took
away out of all the
cities of
Judah the high
places and the
images: and the
kingdom was
quiet
before him.\FTNT{†}{{\FR 14.5 }
images: Heb. sun images}
\par }{\PP \VS{6}¶ And he
built
fenced
cities in
Judah: for the
land had
rest, and he had no
war in those
years; because the
{\ND{
LORD}} had given him
rest.
\VS{7}Therefore he
said unto
Judah, Let us
build these
cities, and make
about
{\ADD{them}}
walls, and
towers,
gates, and
bars,
{\ADD{while}} the
land
{\ADD{is}} yet
before us; because we have
sought the
{\ND{
LORD}} our
God, we have
sought
{\ADD{him}}, and he hath given us
rest on every
side. So they
built and
prospered.
\VS{8}And
Asa had an
army
{\ADD{of men}} that
bare
targets and
spears, out of
Judah
three
hundred
thousand; and out of
Benjamin, that
bare
shields and
drew
bows, two
hundred and
fourscore
thousand: all these
{\ADD{were}} mighty
men of
valour.
\par }{\PP \VS{9}¶ And there came
out against them
Zerah the
Ethiopian with an
host of a
thousand
thousand, and
three
hundred
chariots; and
came unto
Mareshah.
\VS{10}Then
Asa went
out
against him, and they set the
battle in
array in the
valley of
Zephathah at
Mareshah.
\VS{11}And
Asa
cried unto the
{\ND{
LORD}} his
God, and
said,
{\ND{
LORD}},
{\ADD{it is}} nothing with thee to
help,
whether with
many, or with them that have no
power:
help us, O
{\ND{
LORD}} our
God; for we
rest on thee, and in thy
name we
go against this
multitude. O
{\ND{
LORD}}, thou
{\ADD{art}} our
God; let not
man
prevail against thee.\FTNT{‡}{{\FR 14.11 }
man: or, mortal man}
\VS{12}So the
{\ND{
LORD}}
smote the
Ethiopians
before
Asa, and
before
Judah; and the
Ethiopians
fled.
\VS{13}And
Asa and the
people that
{\ADD{were}} with him
pursued them unto
Gerar: and the
Ethiopians were
overthrown, that they could not
recover themselves; for they were
destroyed
before the
{\ND{
LORD}}, and
before his
host; and they carried
away
very
much
spoil.\FTNT{§}{{\FR 14.13 }
destroyed: Heb. broken}
\VS{14}And they
smote all the
cities round
about
Gerar; for the
fear of the
{\ND{
LORD}} came upon them: and they
spoiled all the
cities; for there was exceeding
much
spoil in them.
\VS{15}They
smote also the
tents of
cattle, and carried
away
sheep and
camels in
abundance, and
returned to
Jerusalem.

\par }\Chap{15}{\PP \VerseOne{1}And the
Spirit of
God came upon
Azariah the
son of
Oded:
\VS{2}And he went
out to
meet
Asa, and
said unto him,
Hear ye me,
Asa, and all
Judah and
Benjamin; The
{\ND{
LORD}}
{\ADD{is}} with you, while ye be with him; and if ye
seek him, he will be
found of you; but if ye
forsake him, he will
forsake you.\FTNT{*}{{\FR 15.2 }
to meet…: Heb. before Asa}
\VS{3}Now for a
long
season
Israel
{\ADD{hath been}}
without the
true
God, and without a
teaching
priest, and without
law.
\VS{4}But when they in their
trouble did
turn unto the
{\ND{
LORD}}
God of
Israel, and
sought him, he was
found of them.
\VS{5}And in those
times
{\ADD{there was}} no
peace to him that went
out, nor to him that came
in, but
great
vexations
{\ADD{were}} upon all the
inhabitants of the
countries.
\VS{6}And
nation was
destroyed of
nation, and
city of
city: for
God did
vex them with all
adversity.\FTNT{†}{{\FR 15.6 }
destroyed: Heb. beaten in pieces}
\VS{7}Be ye
strong therefore, and let not your
hands be
weak: for your
work shall
be
rewarded.
\par }{\PP \VS{8}And when
Asa
heard these
words, and the
prophecy of
Oded the
prophet, he took
courage, and put
away the abominable
idols out of all the
land of
Judah and
Benjamin, and out of the
cities which he had
taken from
mount
Ephraim, and
renewed the
altar of the
{\ND{
LORD}}, that
{\ADD{was}}
before the
porch of the
{\ND{
LORD}}.\FTNT{‡}{{\FR 15.8 }
abominable…: Heb. abominations}
\VS{9}And he
gathered all
Judah and
Benjamin, and the
strangers with them out of
Ephraim and
Manasseh, and out of
Simeon: for they
fell to him out of
Israel in
abundance, when they
saw that the
{\ND{
LORD}} his
God
{\ADD{was}} with him.
\VS{10}So they gathered themselves
together at
Jerusalem in the
third
month, in the
fifteenth
year of the
reign of
Asa.
\VS{11}And they
offered unto the
{\ND{
LORD}} the same
time, of the
spoil
{\ADD{which}} they had
brought,
seven
hundred
oxen and
seven
thousand
sheep.\FTNT{§}{{\FR 15.11 }
the same…: Heb. in that day}
\VS{12}And they
entered into a
covenant to
seek the
{\ND{
LORD}}
God of their
fathers with all their
heart and with all their
soul;
\VS{13}That whosoever would not
seek the
{\ND{
LORD}}
God of
Israel should be put to
death,
whether
small or
great, whether
man or
woman.
\VS{14}And they
sware unto the
{\ND{
LORD}} with a
loud
voice, and with
shouting, and with
trumpets, and with
cornets.
\VS{15}And all
Judah
rejoiced at the
oath: for they had
sworn with all their
heart, and
sought him with their whole
desire; and he was
found of them: and the
{\ND{
LORD}} gave them
rest round
about.
\par }{\PP \VS{16}¶ And also
{\ADD{concerning}}
Maachah the
mother of
Asa the
king, he
removed her from
{\ADD{being}}
queen, because she had
made an
idol in a
grove: and
Asa cut
down her
idol, and
stamped
{\ADD{it}}, and
burnt
{\ADD{it}} at the
brook
Kidron.\FTNT{**}{{\FR 15.16 }
mother: that is, grandmother}\FTNT{††}{{\FR 15.16 }
idol: Heb. horror}
\VS{17}But the high
places were not taken
away out of
Israel: nevertheless the
heart of
Asa was
perfect all his
days.
\par }{\PP \VS{18}¶ And he
brought into the
house of
God the things that his
father had
dedicated, and that he himself had
dedicated,
silver, and
gold, and
vessels.
\VS{19}And there was no
{\ADD{more}}
war unto the
five and
thirtieth
year of the
reign of
Asa.

\par }\Chap{16}{\PP \VerseOne{1}In the
six and
thirtieth
year of the
reign of
Asa
Baasha
king of
Israel came
up against
Judah, and
built
Ramah, to the intent
that he might
let none go
out or come
in to
Asa
king of
Judah.
\VS{2}Then
Asa brought
out
silver and
gold out of the
treasures of the
house of the
{\ND{
LORD}} and of the
king’s
house, and
sent to
Ben-hadad
king of
Syria, that
dwelt at
Damascus,
saying,\FTNT{*}{{\FR 16.2 }
Damascus: Heb. Darmesek}
\VS{3}{\ADD{There is}} a
league between me and thee, as
{\ADD{there was}} between my
father and thy
father: behold, I have
sent thee
silver and
gold;
go,
break thy
league with
Baasha
king of
Israel, that he may
depart from me.
\VS{4}And
Ben-hadad
hearkened unto
king
Asa, and
sent the
captains of his
armies against the
cities of
Israel; and they
smote
Ijon, and
Dan, and
Abel-maim, and all the
store
cities of
Naphtali.\FTNT{†}{{\FR 16.4 }
his: Heb. which were his}
\VS{5}And it came to pass, when
Baasha
heard
{\ADD{it}}, that he left
off
building of
Ramah, and let his
work
cease.
\VS{6}Then
Asa the
king
took all
Judah; and they carried
away the
stones of
Ramah, and the
timber thereof, wherewith
Baasha was
building; and he
built therewith
Geba and
Mizpah.
\par }{\PP \VS{7}¶ And at that
time
Hanani the
seer
came to
Asa
king of
Judah, and
said unto him, Because thou hast
relied on the
king of
Syria, and not
relied on the
{\ND{
LORD}} thy
God, therefore is the
host of the
king of
Syria
escaped out of thine
hand.
\VS{8}Were not the
Ethiopians and the
Lubims a
huge
host, with
very
many
chariots and
horsemen? yet, because thou didst
rely on the
{\ND{
LORD}}, he
delivered them into thine
hand.\FTNT{‡}{{\FR 16.8 }
a huge: Heb. in abundance}
\VS{9}For the
eyes of the
{\ND{
LORD}} run to and
fro throughout the whole
earth, to shew himself
strong in the behalf of
{\ADD{them}} whose
heart
{\ADD{is}}
perfect toward him. Herein thou hast done
foolishly: therefore from
henceforth thou shalt
have
wars.\FTNT{§}{{\FR 16.9 }
to shew…: or, strongly to hold with them, etc}
\VS{10}Then
Asa was
wroth with the
seer, and
put him in a
prison
house; for
{\ADD{he was}} in a
rage with him because of this
{\ADD{thing}}. And
Asa
oppressed
{\ADD{some}} of the
people the same
time.\FTNT{**}{{\FR 16.10 }
oppressed: Heb. crushed}
\par }{\PP \VS{11}¶ And, behold, the
acts of
Asa,
first and
last, lo, they
{\ADD{are}}
written in the
book of the
kings of
Judah and
Israel.
\VS{12}And
Asa in the
thirty and
ninth
year of his
reign was
diseased in his
feet, until his
disease
{\ADD{was}}
exceeding
{\ADD{great:}} yet in his
disease he
sought not to the
{\ND{
LORD}}, but to the
physicians.
\par }{\PP \VS{13}¶ And
Asa
slept with his
fathers, and
died in the
one and
fortieth
year of his
reign.
\VS{14}And they
buried him in his own
sepulchres, which he had
made for himself in the
city of
David, and
laid him in the
bed which was
filled with sweet
odours and divers
kinds
{\ADD{of spices}}
prepared by the
apothecaries’
art: and they
made a
very
great
burning for him.\FTNT{††}{{\FR 16.14 }
had made: Heb. had digged}

\par }\Chap{17}{\PP \VerseOne{1}And
Jehoshaphat his
son
reigned in his stead, and
strengthened himself against
Israel.
\VS{2}And he
placed
forces in all the
fenced
cities of
Judah, and
set
garrisons in the
land of
Judah, and in the
cities of
Ephraim, which
Asa his
father had
taken.
\VS{3}And the
{\ND{
LORD}} was with
Jehoshaphat, because he
walked in the
first
ways of his
father
David, and
sought not unto
Baalim;\FTNT{*}{{\FR 17.3 }
of…: or, of his father, and of David}
\VS{4}But
sought to the
{\ADD{
{\ND{Lord}}}}
God of his
father, and
walked in his
commandments, and not after the
doings of
Israel.
\VS{5}Therefore the
{\ND{
LORD}}
stablished the
kingdom in his
hand; and all
Judah
brought to
Jehoshaphat
presents; and he had
riches and
honour in
abundance.\FTNT{†}{{\FR 17.5 }
brought: Heb. gave}
\VS{6}And his
heart was lifted
up in the
ways of the
{\ND{
LORD}}: moreover he took
away the high
places and
groves out of
Judah.\FTNT{‡}{{\FR 17.6 }
was…: that is, was encouraged}
\par }{\PP \VS{7}¶ Also in the
third
year of his
reign he
sent to his
princes,
{\ADD{even}} to
Ben-hail, and to
Obadiah, and to
Zechariah, and to
Nethaneel, and to
Michaiah, to
teach in the
cities of
Judah.
\VS{8}And with them
{\ADD{he sent}}
Levites,
{\ADD{even}}
Shemaiah, and
Nethaniah, and
Zebadiah, and
Asahel, and
Shemiramoth, and
Jehonathan, and
Adonijah, and
Tobijah, and
Tob-adonijah,
Levites; and with them
Elishama and
Jehoram,
priests.
\VS{9}And they
taught in
Judah, and
{\ADD{had}} the
book of the
law of the
{\ND{
LORD}} with them, and went
about throughout all the
cities of
Judah, and
taught the
people.
\par }{\PP \VS{10}¶ And the
fear of the
{\ND{
LORD}} fell upon all the
kingdoms of the
lands that
{\ADD{were}} round
about
Judah, so that they made no
war against
Jehoshaphat.\FTNT{§}{{\FR 17.10 }
fell: Heb. was}
\VS{11}Also
{\ADD{some}} of the
Philistines
brought
Jehoshaphat
presents, and
tribute
silver; and the
Arabians
brought him
flocks,
seven
thousand and
seven
hundred
rams, and
seven
thousand and
seven
hundred he
goats.
\par }{\PP \VS{12}¶ And
Jehoshaphat
waxed
great
exceedingly; and he
built in
Judah
castles, and
cities of
store.\FTNT{**}{{\FR 17.12 }
castles: or, palaces}
\VS{13}And he had
much
business in the
cities of
Judah: and the
men of
war, mighty
men of
valour,
{\ADD{were}} in
Jerusalem.
\VS{14}And these
{\ADD{are}} the
numbers of them according to the
house of their
fathers: Of
Judah, the
captains of
thousands;
Adnah the
chief, and with him mighty
men of
valour
three
hundred
thousand.
\VS{15}And
next to him
{\ADD{was}}
Jehohanan the
captain, and with him two
hundred and
fourscore
thousand.\FTNT{††}{{\FR 17.15 }
next…: Heb. at his hand}
\VS{16}And
next him
{\ADD{was}}
Amasiah the
son of
Zichri, who willingly
offered himself unto the
{\ND{
LORD}}; and with him two
hundred
thousand mighty
men of
valour.
\VS{17}And of
Benjamin;
Eliada a mighty
man of
valour, and with him armed
men with
bow and
shield two
hundred
thousand.
\VS{18}And
next him
{\ADD{was}}
Jehozabad, and with him an
hundred and
fourscore
thousand ready
prepared for the
war.
\VS{19}These
waited on the
king, beside
{\ADD{those}} whom the
king
put in the
fenced
cities throughout all
Judah.

\par }\Chap{18}{\PP \VerseOne{1}Now
Jehoshaphat had
riches and
honour in
abundance, and joined
affinity with
Ahab.
\VS{2}And
after
{\ADD{certain}}
years he went
down to
Ahab to
Samaria. And
Ahab
killed
sheep and
oxen for him in
abundance, and for the
people that
{\ADD{he had}} with him, and
persuaded him to go
up
{\ADD{with him}} to
Ramoth-gilead.\FTNT{*}{{\FR 18.2 }
after…: Heb. at the end of years}
\VS{3}And
Ahab
king of
Israel
said unto
Jehoshaphat
king of
Judah, Wilt thou
go with me to
Ramoth-gilead? And he
answered him, I
{\ADD{am}} as thou
{\ADD{art}}, and my
people as thy
people; and
{\ADD{we will be}} with thee in the
war.
\par }{\PP \VS{4}¶ And
Jehoshaphat
said unto the
king of
Israel,
Enquire, I pray thee, at the
word of the
{\ND{
LORD}} to
day.
\VS{5}Therefore the
king of
Israel gathered
together of
prophets
four
hundred
men, and
said unto them, Shall we
go to
Ramoth-gilead to
battle, or shall I
forbear? And they
said, Go
up; for
God will
deliver
{\ADD{it}} into the
king’s
hand.
\VS{6}But
Jehoshaphat
said,
{\ADD{Is there}} not here a
prophet of the
{\ND{
LORD}} besides, that we might
enquire of him?\FTNT{†}{{\FR 18.6 }
besides: Heb. yet, or, more}
\VS{7}And the
king of
Israel
said unto
Jehoshaphat,
{\ADD{There is}} yet
one
man, by whom we may
enquire of the
{\ND{
LORD}}: but I
hate him; for he never
prophesied
good unto me, but
always
evil: the same
{\ADD{is}}
Micaiah the
son of
Imla. And
Jehoshaphat
said, Let not the
king
say so.
\VS{8}And the
king of
Israel
called for
one
{\ADD{of his}}
officers, and
said, Fetch
quickly
Micaiah the
son of
Imla.\FTNT{‡}{{\FR 18.8 }
officers: or, eunuchs}\FTNT{§}{{\FR 18.8 }
Fetch…: Heb. Hasten}
\VS{9}And the
king of
Israel and
Jehoshaphat
king of
Judah
sat
either of them on his
throne,
clothed in
{\ADD{their}}
robes, and they
sat in a void
place at the entering
in of the
gate of
Samaria; and all the
prophets
prophesied
before them.\FTNT{**}{{\FR 18.9 }
void…: or, floor}
\VS{10}And
Zedekiah the
son of
Chenaanah had
made him
horns of
iron, and
said, Thus
saith the
{\ND{
LORD}}, With these thou shalt
push
Syria until they be
consumed.\FTNT{††}{{\FR 18.10 }
they…: Heb. thou consume them}
\VS{11}And all the
prophets
prophesied so,
saying, Go
up to
Ramoth-gilead, and
prosper: for the
{\ND{
LORD}} shall
deliver
{\ADD{it}} into the
hand of the
king.
\VS{12}And the
messenger that
went to
call
Micaiah
spake to him,
saying, Behold, the
words of the
prophets
{\ADD{declare}}
good to the
king with
one
assent; let thy
word therefore, I pray thee, be like
one of theirs, and
speak thou
good.\FTNT{‡‡}{{\FR 18.12 }
with…: Heb. with one mouth}
\VS{13}And
Micaiah
said,
{\ADD{As}} the
{\ND{
LORD}}
liveth, even what my
God
saith, that will I
speak.
\VS{14}And when he was
come to the
king, the
king
said unto him,
Micaiah, shall we
go to
Ramoth-gilead to
battle, or shall I
forbear? And he
said, Go ye
up, and
prosper, and they shall be
delivered into your
hand.
\VS{15}And the
king
said to him, How many
times shall I
adjure thee that thou
say nothing but the
truth to me in the
name of the
{\ND{
LORD}}?
\VS{16}Then he
said, I did
see all
Israel
scattered upon the
mountains, as
sheep that have no
shepherd: and the
{\ND{
LORD}}
said, These have no
master; let them
return
{\ADD{therefore}} every
man to his
house in
peace.
\VS{17}And the
king of
Israel
said to
Jehoshaphat, Did I not
tell thee
{\ADD{that}} he would not
prophesy
good unto me, but
evil?\FTNT{§§}{{\FR 18.17 }
but evil: or, but for evil}
\VS{18}Again he
said, Therefore
hear the
word of the
{\ND{
LORD}}; I
saw the
{\ND{
LORD}}
sitting upon his
throne, and all the
host of
heaven
standing on his right
hand and
{\ADD{on}} his
left.
\VS{19}And the
{\ND{
LORD}}
said, Who shall
entice
Ahab
king of
Israel, that he may go
up and
fall at
Ramoth-gilead? And one
spake
saying after this
manner, and another
saying after that
manner.
\VS{20}Then there came
out a
spirit, and
stood
before the
{\ND{
LORD}}, and
said, I will
entice him. And the
{\ND{
LORD}}
said unto him, Wherewith?
\VS{21}And he
said, I will go
out, and be a
lying
spirit in the
mouth of all his
prophets. And
{\ADD{the
{\ND{Lord}}}}
said, Thou shalt
entice
{\ADD{him}}, and thou shalt also
prevail: go
out, and
do
{\ADD{even}} so.
\VS{22}Now therefore, behold, the
{\ND{
LORD}} hath
put a
lying
spirit in the
mouth of these thy
prophets, and the
{\ND{
LORD}} hath
spoken
evil against thee.
\VS{23}Then
Zedekiah the
son of
Chenaanah came
near, and
smote
Micaiah upon the
cheek, and
said,
Which
way
went the
Spirit of the
{\ND{
LORD}} from me to
speak unto thee?
\VS{24}And
Micaiah
said, Behold, thou shalt
see on that
day when thou shalt
go into an
inner
chamber to
hide thyself.\FTNT{***}{{\FR 18.24 }
into…: or, from chamber to chamber: Heb. a chamber in a chamber}
\VS{25}Then the
king of
Israel
said,
Take ye
Micaiah, and carry him
back to
Amon the
governor of the
city, and to
Joash the
king’s
son;
\VS{26}And
say, Thus
saith the
king,
Put this
{\ADD{fellow}} in the
prison, and
feed him with
bread of
affliction and with
water of
affliction, until I
return in
peace.
\VS{27}And
Micaiah
said, If thou
certainly
return in
peace,
{\ADD{then}} hath not the
{\ND{
LORD}}
spoken by me. And he
said,
Hearken, all ye
people.
\par }{\PP \VS{28}So the
king of
Israel and
Jehoshaphat the
king of
Judah went
up to
Ramoth-gilead.
\VS{29}And the
king of
Israel
said unto
Jehoshaphat, I will
disguise myself, and will
go to the
battle; but
put thou on thy
robes. So the
king of
Israel
disguised himself; and they
went to the
battle.
\VS{30}Now the
king of
Syria had
commanded the
captains of the
chariots that
{\ADD{were}} with him,
saying,
Fight ye not with
small or
great, save only with the
king of
Israel.
\VS{31}And it came to pass, when the
captains of the
chariots
saw
Jehoshaphat, that they
said, It
{\ADD{is}} the
king of
Israel. Therefore they compassed
about him to
fight: but
Jehoshaphat cried
out, and the
{\ND{
LORD}}
helped him; and
God
moved them
{\ADD{to depart}} from him.
\VS{32}For it came to pass, that, when the
captains of the
chariots
perceived that it was not the
king of
Israel, they turned back
again from
pursuing him.\FTNT{†††}{{\FR 18.32 }
from…: Heb. from after him}
\VS{33}And a
{\ADD{certain}} man
drew a
bow at a
venture, and
smote the
king of
Israel between the
joints of the
harness: therefore he
said to his
chariot
man,
Turn thine
hand, that thou mayest carry me
out of the
host; for I am
wounded.\FTNT{‡‡‡}{{\FR 18.33 }
at a venture: Heb. in his simplicity}\FTNT{§§§}{{\FR 18.33 }
between…: Heb. between the joints and between the breastplate}\FTNT{*}{{\FR 18.33 }
wounded: Heb. made sick}
\VS{34}And the
battle
increased that
day: howbeit the
king of
Israel stayed
{\ADD{himself}}
up in
{\ADD{his}}
chariot
against the
Syrians until the
even: and about the
time of the
sun going
down he
died.

\par }\Chap{19}{\PP \VerseOne{1}And
Jehoshaphat the
king of
Judah
returned to his
house in
peace to
Jerusalem.
\VS{2}And
Jehu the
son of
Hanani the
seer went
out to
meet him, and
said to
king
Jehoshaphat, Shouldest thou
help the
ungodly, and
love them that
hate the
{\ND{
LORD}}?
therefore
{\ADD{is}}
wrath upon thee from
before the
{\ND{
LORD}}.
\VS{3}Nevertheless there are
good
things
found in thee, in that thou hast taken
away the
groves out of the
land, and hast
prepared thine
heart to
seek
God.
\VS{4}And
Jehoshaphat
dwelt at
Jerusalem: and he went
out
again through the
people from
Beer-sheba to
mount
Ephraim, and brought them
back unto the
{\ND{
LORD}}
God of their
fathers.\FTNT{*}{{\FR 19.4 }
he went…: Heb. he returned and went out}
\par }{\PP \VS{5}¶ And he
set
judges in the
land throughout all the
fenced
cities of
Judah,
city by
city,
\VS{6}And
said to the
judges, Take
heed what ye
do: for ye
judge not for
man, but for the
{\ND{
LORD}}, who
{\ADD{is}} with you in the
judgment.\FTNT{†}{{\FR 19.6 }
in…: Heb. in the matter of judgment}
\VS{7}Wherefore now let the
fear of the
{\ND{
LORD}} be upon you; take
heed and
do
{\ADD{it:}} for
{\ADD{there is}} no
iniquity with the
{\ND{
LORD}} our
God, nor
respect of
persons, nor
taking of
gifts.
\par }{\PP \VS{8}¶ Moreover in
Jerusalem did
Jehoshaphat
set of the
Levites, and
{\ADD{of}} the
priests, and of the
chief of the
fathers of
Israel, for the
judgment of the
{\ND{
LORD}}, and for
controversies, when they
returned to
Jerusalem.
\VS{9}And he
charged them,
saying, Thus shall ye
do in the
fear of the
{\ND{
LORD}},
faithfully, and with a
perfect
heart.
\VS{10}And what
cause soever shall
come to you of your
brethren that
dwell in their
cities, between
blood and
blood, between
law and
commandment,
statutes and
judgments, ye shall even
warn them that they
trespass not against the
{\ND{
LORD}}, and
{\ADD{so}}
wrath come upon you, and upon your
brethren:
this
do, and ye shall not
trespass.
\VS{11}And, behold,
Amariah the
chief
priest
{\ADD{is}} over you in all
matters of the
{\ND{
LORD}}; and
Zebadiah the
son of
Ishmael, the
ruler of the
house of
Judah, for all the
king’s
matters: also the
Levites
{\ADD{shall be}}
officers
before you.
Deal
courageously, and the
{\ND{
LORD}} shall be with the
good.\FTNT{‡}{{\FR 19.11 }
Deal…: Heb. Take courage and do}

\par }\Chap{20}{\PP \VerseOne{1}It came to pass after
this also,
{\ADD{that}} the
children of
Moab, and the
children of
Ammon, and with them
{\ADD{other}} beside the
Ammonites,
came against
Jehoshaphat to
battle.
\VS{2}Then there
came some that
told
Jehoshaphat,
saying, There
cometh a
great
multitude against thee from
beyond the
sea on this side
Syria; and, behold, they
{\ADD{be}} in
Hazazon-tamar, which
{\ADD{is}}
En-gedi.
\VS{3}And
Jehoshaphat
feared, and
set
himself to
seek the
{\ND{
LORD}}, and
proclaimed a
fast
throughout all
Judah.\FTNT{*}{{\FR 20.3 }
himself: Heb. his face}
\VS{4}And
Judah gathered themselves
together, to
ask
{\ADD{help}} of the
{\ND{
LORD}}: even out of all the
cities of
Judah they
came to
seek the
{\ND{
LORD}}.
\par }{\PP \VS{5}¶ And
Jehoshaphat
stood in the
congregation of
Judah and
Jerusalem, in the
house of the
{\ND{
LORD}},
before the
new
court,
\VS{6}And
said, O
{\ND{
LORD}}
God of our
fathers,
{\ADD{art}} not thou
God in
heaven? and
rulest
{\ADD{not}} thou over all the
kingdoms of the
heathen? and in thine
hand
{\ADD{is there not}}
power and
might, so that none is able to
withstand thee?
\VS{7}{\ADD{Art}} not thou our
God,
{\ADD{who}} didst drive
out the
inhabitants of this
land
before thy
people
Israel, and
gavest it to the
seed of
Abraham thy
friend for
ever?\FTNT{†}{{\FR 20.7 }
who: Heb. thou}
\VS{8}And they
dwelt therein, and have
built thee a
sanctuary therein for thy
name,
saying,
\VS{9}If,
{\ADD{when}}
evil
cometh upon us,
{\ADD{as}} the
sword,
judgment, or
pestilence, or
famine, we
stand
before this
house, and in thy
presence, (for thy
name
{\ADD{is}} in this
house,) and
cry unto thee in our
affliction, then thou wilt
hear and
help.
\VS{10}And now, behold, the
children of
Ammon and
Moab and
mount
Seir, whom thou wouldest not
let
Israel
invade, when they came
out of the
land of
Egypt, but they
turned from them, and
destroyed them not;
\VS{11}Behold,
{\ADD{I say, how}} they
reward us, to
come to cast us
out of thy
possession, which thou hast given us to
inherit.
\VS{12}O our
God, wilt thou not
judge them? for we have no
might
against this
great
company that
cometh against us; neither
know
we what to
do: but our
eyes
{\ADD{are}} upon thee.
\VS{13}And all
Judah
stood
before the
{\ND{
LORD}},
with their little
ones, their
wives, and their
children.
\par }{\PP \VS{14}¶ Then upon
Jahaziel the
son of
Zechariah, the
son of
Benaiah, the
son of
Jeiel, the
son of
Mattaniah, a
Levite of the
sons of
Asaph, came the
Spirit of the
{\ND{
LORD}} in the
midst of the
congregation;
\VS{15}And he
said,
Hearken ye, all
Judah, and ye
inhabitants of
Jerusalem, and thou
king
Jehoshaphat, Thus
saith the
{\ND{
LORD}} unto you, Be not
afraid nor
dismayed by
reason of this
great
multitude; for the
battle
{\ADD{is}} not yours, but
God’s.
\VS{16}To
morrow go ye
down against them: behold, they come
up by the
cliff of
Ziz; and ye shall
find them at the
end of the
brook,
before the
wilderness of
Jeruel.\FTNT{‡}{{\FR 20.16 }
cliff: Heb. ascent}\FTNT{§}{{\FR 20.16 }
brook: or, valley}
\VS{17}Ye shall not
{\ADD{need}} to
fight in
this
{\ADD{battle:}}
set yourselves,
stand ye
{\ADD{still}}, and
see the
salvation of the
{\ND{
LORD}} with you, O
Judah and
Jerusalem:
fear not, nor be
dismayed; to
morrow go
out
against them: for the
{\ND{
LORD}}
{\ADD{will be}} with you.
\VS{18}And
Jehoshaphat
bowed his head with
{\ADD{his}}
face to the
ground: and all
Judah and the
inhabitants of
Jerusalem
fell
before the
{\ND{
LORD}},
worshipping the
{\ND{
LORD}}.
\VS{19}And the
Levites, of the
children of the
Kohathites, and of the
children of the
Korhites, stood
up to
praise the
{\ND{
LORD}}
God of
Israel with a
loud
voice on
high.
\par }{\PP \VS{20}¶ And they rose
early in the
morning, and went
forth into the
wilderness of
Tekoa: and as they went
forth,
Jehoshaphat
stood and
said,
Hear me, O
Judah, and ye
inhabitants of
Jerusalem;
Believe in the
{\ND{
LORD}} your
God, so shall ye be
established;
believe his
prophets, so shall ye
prosper.
\VS{21}And when he had
consulted with the
people, he
appointed
singers unto the
{\ND{
LORD}}, and that should
praise the
beauty of
holiness, as they went
out
before the
army, and to
say,
Praise the
{\ND{
LORD}}; for his
mercy
{\ADD{endureth}} for
ever.\FTNT{**}{{\FR 20.21 }
that…: Heb. praisers}
\par }{\PP \VS{22}¶ And
when they
began to
sing and to
praise, the
{\ND{
LORD}}
set
ambushments against the
children of
Ammon,
Moab, and
mount
Seir, which were
come against
Judah; and they were
smitten.\FTNT{††}{{\FR 20.22 }
And when…: Heb. And in the time that they, etc}\FTNT{‡‡}{{\FR 20.22 }
to sing…: Heb. in singing and praise}\FTNT{§§}{{\FR 20.22 }
they were…: or, they smote one another}
\VS{23}For the
children of
Ammon and
Moab stood
up against the
inhabitants of
mount
Seir, utterly to
slay and
destroy
{\ADD{them:}} and when they had made an
end of the
inhabitants of
Seir, every
one
helped to
destroy
another.\FTNT{***}{{\FR 20.23 }
to destroy: Heb. for the destruction}
\VS{24}And when
Judah
came toward the watch
tower in the
wilderness, they
looked unto the
multitude, and, behold, they
{\ADD{were}} dead
bodies
fallen to the
earth, and none
escaped.\FTNT{†††}{{\FR 20.24 }
none…: Heb. there was not an escaping}
\VS{25}And when
Jehoshaphat and his
people
came to take
away the
spoil of them, they
found among them in
abundance both
riches with the dead
bodies, and
precious
jewels, which they stripped
off for themselves, more
than they could carry
away: and they were
three
days in
gathering of the
spoil, it was so
much.
\par }{\PP \VS{26}¶ And on the
fourth
day they
assembled themselves in the
valley of
Berachah; for there they
blessed the
{\ND{
LORD}}: therefore the
name of the same
place was
called, The
valley of
Berachah, unto this
day.\FTNT{‡‡‡}{{\FR 20.26 }
Berachah: that is, blessing}
\VS{27}Then they
returned, every
man of
Judah and
Jerusalem, and
Jehoshaphat in the
forefront of them, to go
again to
Jerusalem with
joy; for the
{\ND{
LORD}} had made them to
rejoice over their
enemies.\FTNT{§§§}{{\FR 20.27 }
forefront: Heb. head}
\VS{28}And they
came to
Jerusalem with
psalteries and
harps and
trumpets unto the
house of the
{\ND{
LORD}}.
\VS{29}And the
fear of
God was on all the
kingdoms of
{\ADD{those}}
countries, when they had
heard that the
{\ND{
LORD}}
fought against the
enemies of
Israel.
\VS{30}So the
realm of
Jehoshaphat was
quiet: for his
God gave him
rest round
about.
\par }{\PP \VS{31}¶ And
Jehoshaphat
reigned over
Judah:
{\ADD{he was}}
thirty and
five
years
old when he began to
reign, and he
reigned
twenty and
five
years in
Jerusalem. And his
mother’s
name
{\ADD{was}}
Azubah the
daughter of
Shilhi.
\VS{32}And he
walked in the
way of
Asa his
father, and
departed not from it,
doing
{\ADD{that which was}}
right in the
sight of the
{\ND{
LORD}}.
\VS{33}Howbeit the high
places were not taken
away: for as yet the
people had not
prepared their
hearts unto the
God of their
fathers.
\VS{34}Now the
rest of the
acts of
Jehoshaphat,
first and
last, behold, they
{\ADD{are}}
written in the
book of
Jehu the
son of
Hanani, who
{\ADD{is}}
mentioned in the
book of the
kings of
Israel.\FTNT{*}{{\FR 20.34 }
book of Jehu: Heb. words, etc}\FTNT{†}{{\FR 20.34 }
is mentioned: Heb. was made to ascend}
\par }{\PP \VS{35}¶ And
after this did
Jehoshaphat
king of
Judah
join himself with
Ahaziah
king of
Israel, who
did very
wickedly:
\VS{36}And he
joined himself with him to
make
ships to
go to
Tarshish: and they
made the
ships in
Ezion-geber.
\VS{37}Then
Eliezer the
son of
Dodavah of
Mareshah
prophesied against
Jehoshaphat,
saying, Because thou hast
joined thyself with
Ahaziah, the
{\ND{
LORD}} hath
broken thy
works. And the
ships were
broken, that they were not
able to
go to
Tarshish.

\par }\Chap{21}{\PP \VerseOne{1}Now
Jehoshaphat
slept with his
fathers, and was
buried with his
fathers in the
city of
David. And
Jehoram his
son
reigned in his stead.
\VS{2}And he had
brethren the
sons of
Jehoshaphat,
Azariah, and
Jehiel, and
Zechariah, and
Azariah, and
Michael, and
Shephatiah: all these
{\ADD{were}} the
sons of
Jehoshaphat
king of
Israel.
\VS{3}And their
father
gave them
great
gifts of
silver, and of
gold, and of precious
things, with
fenced
cities in
Judah: but the
kingdom
gave he to
Jehoram; because he
{\ADD{was}} the
firstborn.
\VS{4}Now when
Jehoram was risen
up to the
kingdom of his
father, he
strengthened himself, and
slew all his
brethren with the
sword, and
{\ADD{divers}} also of the
princes of
Israel.
\par }{\PP \VS{5}¶
Jehoram
{\ADD{was}}
thirty and
two
years
old when he began to
reign, and he
reigned
eight
years in
Jerusalem.
\VS{6}And he
walked in the
way of the
kings of
Israel, like as
did the
house of
Ahab: for he had the
daughter of
Ahab to
wife: and he
wrought
{\ADD{that which was}}
evil in the
eyes of the
{\ND{
LORD}}.
\VS{7}Howbeit the
{\ND{
LORD}}
would not
destroy the
house of
David, because of the
covenant that he had
made with
David, and as he
promised to
give a
light to him and to his
sons for
ever.\FTNT{*}{{\FR 21.7 }
light: Heb. lamp, or, candle}
\par }{\PP \VS{8}¶ In his
days the
Edomites
revolted from under the
dominion of
Judah, and
made themselves a
king.\FTNT{†}{{\FR 21.8 }
dominion: Heb. hand}
\VS{9}Then
Jehoram went
forth with his
princes, and all his
chariots with him: and he rose
up by
night, and
smote the
Edomites which compassed him
in, and the
captains of the
chariots.
\VS{10}So the
Edomites
revolted from under the
hand of
Judah unto this
day. The same
time
{\ADD{also}} did
Libnah
revolt from under his
hand; because he had
forsaken the
{\ND{
LORD}}
God of his
fathers.
\VS{11}Moreover he
made high
places in the
mountains of
Judah, and caused the
inhabitants of
Jerusalem to commit
fornication, and
compelled
Judah
{\ADD{thereto}}.
\par }{\PP \VS{12}¶ And there
came a
writing to him from
Elijah the
prophet,
saying, Thus
saith the
{\ND{
LORD}}
God of
David thy
father, Because thou hast not
walked in the
ways of
Jehoshaphat thy
father, nor in the
ways of
Asa
king of
Judah,
\VS{13}But hast
walked in the
way of the
kings of
Israel, and hast made
Judah and the
inhabitants of
Jerusalem to go a
whoring, like to the
whoredoms of the
house of
Ahab, and also hast
slain thy
brethren of thy
father’s
house,
{\ADD{which were}}
better than thyself:
\VS{14}Behold, with a
great
plague will the
{\ND{
LORD}}
smite thy
people, and thy
children, and thy
wives, and all thy
goods:\FTNT{‡}{{\FR 21.14 }
a great…: Heb. a great stroke}
\VS{15}And thou
{\ADD{shalt have}}
great
sickness by
disease of thy
bowels, until thy
bowels fall
out by reason of the
sickness
day by
day.
\par }{\PP \VS{16}¶ Moreover the
{\ND{
LORD}} stirred
up against
Jehoram the
spirit of the
Philistines, and of the
Arabians, that
{\ADD{were}}
near the
Ethiopians:
\VS{17}And they came
up into
Judah, and
brake into it, and carried
away all the
substance that was
found in the
king’s
house, and his
sons also, and his
wives; so that there was never a
son
left him, save
Jehoahaz, the
youngest of his
sons.\FTNT{§}{{\FR 21.17 }
carried…: Heb. carried captive}\FTNT{**}{{\FR 21.17 }
Jehoahaz: also called, Ahaziah, or, Azariah}
\par }{\PP \VS{18}¶ And
after all this the
{\ND{
LORD}}
smote him in his
bowels with an
incurable
disease.
\VS{19}And it came to pass, that in process of
time,
after the
end of
two
years, his
bowels fell
out by
reason of his
sickness: so he
died of
sore
diseases. And his
people
made no
burning for him, like the
burning of his
fathers.
\VS{20}Thirty and
two years
old was he when he began to
reign, and he
reigned in
Jerusalem
eight
years, and
departed without being
desired. Howbeit they
buried him in the
city of
David, but not in the
sepulchres of the
kings.\FTNT{††}{{\FR 21.20 }
without…: Heb. without desire}

\par }\Chap{22}{\PP \VerseOne{1}And the
inhabitants of
Jerusalem
made
Ahaziah his
youngest
son
king in his stead: for the band of
men that
came with the
Arabians to the
camp had
slain all the
eldest. So
Ahaziah the
son of
Jehoram
king of
Judah
reigned.
\VS{2}Forty and
two
years
old
{\ADD{was}}
Ahaziah when he began to
reign, and he
reigned
one
year in
Jerusalem. His
mother’s
name also
{\ADD{was}}
Athaliah the
daughter of
Omri.
\VS{3}He also
walked in the
ways of the
house of
Ahab: for his
mother was his
counsellor to do
wickedly.
\VS{4}Wherefore he
did
evil in the
sight of the
{\ND{
LORD}} like the
house of
Ahab: for they were his
counsellors
after the
death of his
father to his
destruction.
\par }{\PP \VS{5}¶ He
walked also after their
counsel, and
went with
Jehoram the
son of
Ahab
king of
Israel to
war against
Hazael
king of
Syria at
Ramoth-gilead: and the
Syrians
smote
Joram.
\VS{6}And he
returned to be
healed in
Jezreel because of the
wounds which were
given him at
Ramah, when he
fought with
Hazael
king of
Syria. And
Azariah the
son of
Jehoram
king of
Judah went
down to
see
Jehoram the
son of
Ahab at
Jezreel, because he was
sick.\FTNT{*}{{\FR 22.6 }
which…: Heb. wherewith they wounded him}\FTNT{†}{{\FR 22.6 }
Azariah: also called, Ahaziah, or, Jehoahaz}
\VS{7}And the
destruction of
Ahaziah was of
God by
coming to
Joram: for when he was
come, he went
out with
Jehoram against
Jehu the
son of
Nimshi, whom the
{\ND{
LORD}} had
anointed to cut
off the
house of
Ahab.\FTNT{‡}{{\FR 22.7 }
destruction: Heb. treading down}
\VS{8}And it came to pass, that, when
Jehu was executing
judgment upon the
house of
Ahab, and
found the
princes of
Judah, and the
sons of the
brethren of
Ahaziah, that
ministered to
Ahaziah, he
slew them.
\VS{9}And he
sought
Ahaziah: and they
caught him, (for he was
hid in
Samaria,) and
brought him to
Jehu: and when they had
slain him, they
buried him: Because,
said they, he
{\ADD{is}} the
son of
Jehoshaphat, who
sought the
{\ND{
LORD}} with all his
heart. So the
house of
Ahaziah had no
power to
keep still the
kingdom.
\par }{\PP \VS{10}¶ But when
Athaliah the
mother of
Ahaziah
saw that her
son was
dead, she
arose and
destroyed all the
seed
royal of the
house of
Judah.
\VS{11}But
Jehoshabeath, the
daughter of the
king,
took
Joash the
son of
Ahaziah, and
stole him from
among the
king’s
sons that were
slain, and
put him and his
nurse in a
bedchamber. So
Jehoshabeath, the
daughter of
king
Jehoram, the
wife of
Jehoiada the
priest, (for she was the
sister of
Ahaziah,)
hid him
from
Athaliah, so that she
slew him not.\FTNT{§}{{\FR 22.11 }
Jehoshabeath: also called, Jehosheba}
\VS{12}And he was with them
hid in the
house of
God
six
years: and
Athaliah
reigned over the
land.

\par }\Chap{23}{\PP \VerseOne{1}And in the
seventh
year
Jehoiada
strengthened himself, and
took the
captains of
hundreds,
Azariah the
son of
Jeroham, and
Ishmael the
son of
Jehohanan, and
Azariah the
son of
Obed, and
Maaseiah the
son of
Adaiah, and
Elishaphat the
son of
Zichri, into
covenant with him.
\VS{2}And they went
about in
Judah, and
gathered the
Levites out of all the
cities of
Judah, and the
chief of the
fathers of
Israel, and they
came to
Jerusalem.
\VS{3}And all the
congregation
made a
covenant with the
king in the
house of
God. And he
said unto them, Behold, the
king’s
son shall
reign, as the
{\ND{
LORD}} hath
said of the
sons of
David.
\VS{4}This
{\ADD{is}} the
thing that ye shall
do; A third
part of you
entering on the
sabbath, of the
priests and of the
Levites,
{\ADD{shall be}}
porters of the
doors;\FTNT{*}{{\FR 23.4 }
doors: Heb. thresholds}
\VS{5}And a third
part
{\ADD{shall be}} at the
king’s
house; and a third
part at the
gate of the
foundation: and all the
people
{\ADD{shall be}} in the
courts of the
house of the
{\ND{
LORD}}.
\VS{6}But let none
come into the
house of the
{\ND{
LORD}}, save the
priests, and they that
minister of the
Levites; they shall go
in, for they
{\ADD{are}}
holy: but all the
people shall
keep the
watch of the
{\ND{
LORD}}.
\VS{7}And the
Levites shall
compass the
king round
about, every
man with his
weapons in his
hand; and whosoever
{\ADD{else}}
cometh into the
house, he shall be put to
death: but be ye with the
king when he cometh
in, and when he goeth
out.
\VS{8}So the
Levites and all
Judah
did according to all things that
Jehoiada the
priest had
commanded, and
took every
man his
men that were to come
in on the
sabbath, with them that were to
go
{\ADD{out}} on the
sabbath: for
Jehoiada the
priest
dismissed not the
courses.
\VS{9}Moreover
Jehoiada the
priest
delivered to the
captains of
hundreds
spears, and
bucklers, and
shields, that
{\ADD{had been}}
king
David’s, which
{\ADD{were}} in the
house of
God.
\VS{10}And he
set all the
people, every
man having his
weapon in his
hand, from the
right
side of the
temple to the
left
side of the
temple, along by the
altar and the
temple, by the
king round
about.\FTNT{†}{{\FR 23.10 }
side: Heb. shoulder}\FTNT{‡}{{\FR 23.10 }
temple: Heb. house}
\VS{11}Then they brought
out the
king’s
son, and
put upon him the
crown, and
{\ADD{gave him}} the
testimony, and made him
king. And
Jehoiada and his
sons
anointed him, and
said, God
save the
king.\FTNT{§}{{\FR 23.11 }
God…: Heb. Let the king live}
\par }{\PP \VS{12}¶ Now when
Athaliah
heard the
noise of the
people
running and
praising the
king, she
came to the
people into the
house of the
{\ND{
LORD}}:
\VS{13}And she
looked, and, behold, the
king
stood at his
pillar at the entering
in, and the
princes and the
trumpets by the
king: and all the
people of the
land
rejoiced, and
sounded with
trumpets, also the
singers with
instruments of
musick, and such as
taught to sing
praise. Then
Athaliah
rent her
clothes, and
said,
Treason,
Treason.\FTNT{**}{{\FR 23.13 }
Treason: Heb. Conspiracy}
\VS{14}Then
Jehoiada the
priest brought
out the
captains of
hundreds that were set
over the
host, and
said unto them, Have her
forth of the
ranges: and whoso
followeth her, let him be
slain with the
sword. For the
priest
said,
Slay her not in the
house of the
{\ND{
LORD}}.
\VS{15}So they
laid
hands on her; and when she was
come to the
entering of the
horse
gate by the
king’s
house, they
slew her there.
\par }{\PP \VS{16}¶ And
Jehoiada
made a
covenant between him, and between all the
people, and between the
king, that they should be the
{\ND{
LORD’s}}
people.
\VS{17}Then all the
people
went to the
house of
Baal, and brake it
down, and
brake his
altars and his
images in
pieces, and
slew
Mattan the
priest of
Baal
before the
altars.
\VS{18}Also
Jehoiada
appointed the
offices of the
house of the
{\ND{
LORD}} by the
hand of the
priests the
Levites, whom
David had
distributed in the
house of the
{\ND{
LORD}}, to
offer the burnt
offerings of the
{\ND{
LORD}}, as
{\ADD{it is}}
written in the
law of
Moses, with
rejoicing and with
singing,
{\ADD{as it was ordained}}
by
David.\FTNT{††}{{\FR 23.18 }
as it was…: Heb. by the hands of David}
\VS{19}And he
set the
porters at the
gates of the
house of the
{\ND{
LORD}}, that none
{\ADD{which was}}
unclean in any
thing should enter
in.
\VS{20}And he
took the
captains of
hundreds, and the
nobles, and the
governors of the
people, and all the
people of the
land, and brought
down the
king from the
house of the
{\ND{
LORD}}: and they
came
through the
high
gate into the
king’s
house, and
set the
king upon the
throne of the
kingdom.
\VS{21}And all the
people of the
land
rejoiced: and the
city was
quiet, after that they had
slain
Athaliah with the
sword.

\par }\Chap{24}{\PP \VerseOne{1}Joash
{\ADD{was}}
seven
years
old when he began to
reign, and he
reigned
forty
years in
Jerusalem. His
mother’s
name also
{\ADD{was}}
Zibiah of
Beer-sheba.
\VS{2}And
Joash
did
{\ADD{that which was}}
right in the
sight of the
{\ND{
LORD}} all the
days of
Jehoiada the
priest.
\VS{3}And
Jehoiada
took for him
two
wives; and he
begat
sons and
daughters.
\par }{\PP \VS{4}¶ And it came to pass
after this,
{\ADD{that}}
Joash was
minded to
repair the
house of the
{\ND{
LORD}}.\FTNT{*}{{\FR 24.4 }
to repair: Heb. to renew}
\VS{5}And he gathered
together the
priests and the
Levites, and
said to them, Go
out unto the
cities of
Judah, and
gather of all
Israel
money to
repair the
house of your
God
from
year to
year, and see that ye
hasten the
matter. Howbeit the
Levites
hastened
{\ADD{it}} not.
\VS{6}And the
king
called for
Jehoiada the
chief, and
said unto him, Why hast thou not
required of the
Levites to bring
in out of
Judah and out of
Jerusalem the
collection,
{\ADD{according to the commandment}} of
Moses the
servant of the
{\ND{
LORD}}, and of the
congregation of
Israel, for the
tabernacle of
witness?
\VS{7}For the
sons of
Athaliah, that
wicked woman, had broken
up the
house of
God; and also all the dedicated
things of the
house of the
{\ND{
LORD}} did they
bestow upon
Baalim.
\VS{8}And at the
king’s
commandment they
made
a
chest, and
set it
without at the
gate of the
house of the
{\ND{
LORD}}.
\VS{9}And they
made a
proclamation through
Judah and
Jerusalem, to bring
in to the
{\ND{
LORD}} the
collection
{\ADD{that}}
Moses the
servant of
God
{\ADD{laid}} upon
Israel in the
wilderness.\FTNT{†}{{\FR 24.9 }
a proclamation: Heb. a voice}
\VS{10}And all the
princes and all the
people
rejoiced, and brought
in, and
cast into the
chest, until they had made an
end.
\VS{11}Now it came to pass, that at what
time the
chest was
brought unto the
king’s
office by the
hand of the
Levites, and when they
saw that
{\ADD{there was}}
much
money, the
king’s
scribe and the
high
priest’s
officer
came and
emptied the
chest, and
took it, and
carried it to his
place
again. Thus they
did
day by
day, and
gathered
money in
abundance.
\VS{12}And the
king and
Jehoiada
gave it to such as
did the
work of the
service of the
house of the
{\ND{
LORD}}, and
hired
masons and
carpenters to
repair the
house of the
{\ND{
LORD}}, and also such as
wrought
iron and
brass to
mend the
house of the
{\ND{
LORD}}.
\VS{13}So the
workmen
wrought, and the
work
was
perfected by
them, and they
set the
house of
God in his
state, and
strengthened it.\FTNT{‡}{{\FR 24.13 }
the work…: Heb. the healing went up upon the work}
\VS{14}And when they had
finished
{\ADD{it}}, they
brought the
rest of the
money
before the
king and
Jehoiada, whereof were
made
vessels for the
house of the
{\ND{
LORD}},
{\ADD{even}}
vessels to
minister, and to
offer
{\ADD{withal}}, and
spoons, and
vessels of
gold and
silver. And they
offered burnt
offerings in the
house of the
{\ND{
LORD}}
continually all the
days of
Jehoiada.\FTNT{§}{{\FR 24.14 }
to offer…: or, pestils}
\par }{\PP \VS{15}¶ But
Jehoiada waxed
old, and was
full of
days when he
died; an
hundred and
thirty
years
old
{\ADD{was he}} when he
died.
\VS{16}And they
buried him in the
city of
David among the
kings, because he had
done
good in
Israel, both toward
God, and toward his
house.
\VS{17}Now
after the
death of
Jehoiada
came the
princes of
Judah, and made
obeisance to the
king. Then the
king
hearkened unto them.
\VS{18}And they
left the
house of the
{\ND{
LORD}}
God of their
fathers, and
served
groves and
idols: and
wrath came upon
Judah and
Jerusalem for this their
trespass.
\VS{19}Yet he
sent
prophets to them, to bring them
again unto the
{\ND{
LORD}}; and they
testified against them: but they would not give
ear.
\VS{20}And the
Spirit of
God
came upon
Zechariah the
son of
Jehoiada the
priest, which
stood above the
people, and
said unto them, Thus
saith
God, Why
transgress ye the
commandments of the
{\ND{
LORD}}, that ye cannot
prosper? because ye have
forsaken the
{\ND{
LORD}}, he hath also
forsaken you.\FTNT{**}{{\FR 24.20 }
came…: Heb. clothed}
\VS{21}And they
conspired against him, and
stoned him with
stones at the
commandment of the
king in the
court of the
house of the
{\ND{
LORD}}.
\VS{22}Thus
Joash the
king
remembered not the
kindness which
Jehoiada his
father had
done to him, but
slew his
son. And when he
died, he
said, The
{\ND{
LORD}}
look upon
{\ADD{it}}, and
require
{\ADD{it}}.
\par }{\PP \VS{23}¶ And it came to pass at the
end of the
year,
{\ADD{that}} the
host of
Syria came
up against him: and they
came to
Judah and
Jerusalem, and
destroyed all the
princes of the
people from among the
people, and
sent all the
spoil of them unto the
king of
Damascus.\FTNT{††}{{\FR 24.23 }
at the…: Heb. in the revolution of the year}\FTNT{‡‡}{{\FR 24.23 }
Damascus: Heb. Darmesek}
\VS{24}For the
army of the
Syrians
came with a
small company of
men, and the
{\ND{
LORD}}
delivered a
very
great
host into their
hand, because they had
forsaken the
{\ND{
LORD}}
God of their
fathers. So they
executed
judgment against
Joash.
\VS{25}And when they were
departed from him, (for they
left him in
great
diseases,) his own
servants
conspired against him for the
blood of the
sons of
Jehoiada the
priest, and
slew him on his
bed, and he
died: and they
buried him in the
city of
David, but they
buried him not in the
sepulchres of the
kings.
\VS{26}And these are they that
conspired against him;
Zabad the
son of
Shimeath an
Ammonitess, and
Jehozabad the
son of
Shimrith a
Moabitess.\FTNT{§§}{{\FR 24.26 }
Zabad: or, Jozacher}\FTNT{***}{{\FR 24.26 }
Shimrith: or, Shomer}
\par }{\PP \VS{27}¶ Now
{\ADD{concerning}} his
sons, and the
greatness of the
burdens
{\ADD{laid}} upon him, and the
repairing of the
house of
God, behold, they
{\ADD{are}}
written in the
story of the
book of the
kings. And
Amaziah his
son
reigned in his stead.\FTNT{†††}{{\FR 24.27 }
repairing: Heb. founding}\FTNT{‡‡‡}{{\FR 24.27 }
story: or, commentary}

\par }\Chap{25}{\PP \VerseOne{1}Amaziah
{\ADD{was}}
twenty and
five
years
old
{\ADD{when}} he began to
reign, and he
reigned
twenty and
nine
years in
Jerusalem. And his
mother’s
name
{\ADD{was}}
Jehoaddan of
Jerusalem.
\VS{2}And he
did
{\ADD{that which was}}
right in the
sight of the
{\ND{
LORD}}, but not with a
perfect
heart.
\par }{\PP \VS{3}¶ Now it came to pass, when the
kingdom was
established to him, that he
slew his
servants that had
killed the
king his
father.\FTNT{*}{{\FR 25.3 }
established…: Heb. confirmed upon him}
\VS{4}But he
slew not their
children, but
{\ADD{did}} as
{\ADD{it is}}
written in the
law in the
book of
Moses, where the
{\ND{
LORD}}
commanded,
saying, The
fathers shall not
die for the
children, neither shall the
children
die for the
fathers, but every
man shall
die for his own
sin.
\par }{\PP \VS{5}¶ Moreover
Amaziah
gathered
Judah
together, and
made them
captains over
thousands, and
captains over
hundreds, according to the
houses of
{\ADD{their}}
fathers, throughout all
Judah and
Benjamin: and he
numbered them from
twenty
years
old and
above, and
found them
three
hundred
thousand
choice
{\ADD{men, able}} to go
forth to
war, that could
handle
spear and
shield.
\VS{6}He
hired also an
hundred
thousand
mighty men of
valour out of
Israel for an
hundred
talents of
silver.
\VS{7}But there
came a
man of
God to him,
saying, O
king, let not the
army of
Israel
go with thee; for the
{\ND{
LORD}}
{\ADD{is}} not with
Israel,
{\ADD{to wit, with}} all the
children of
Ephraim.
\VS{8}But if thou wilt
go,
do
{\ADD{it}}, be
strong for the
battle:
God shall make thee
fall
before the
enemy: for
God
hath
power to
help, and to cast
down.
\VS{9}And
Amaziah
said to the
man of
God, But what shall we
do for the
hundred
talents which I have
given to the
army of
Israel? And the
man of
God
answered, The
{\ND{
LORD}}
is able to
give thee much
more than this.\FTNT{†}{{\FR 25.9 }
army: Heb. band}
\VS{10}Then
Amaziah
separated them,
{\ADD{to wit}}, the
army that was
come to him out of
Ephraim, to
go
home
again: wherefore their
anger was
greatly
kindled against
Judah, and they
returned
home in
great
anger.\FTNT{‡}{{\FR 25.10 }
home again: Heb. to their place}\FTNT{§}{{\FR 25.10 }
in great…: Heb. in heat of anger}
\par }{\PP \VS{11}¶ And
Amaziah
strengthened himself, and led
forth his
people, and
went to the
valley of
salt, and
smote of the
children of
Seir
ten
thousand.
\VS{12}And
{\ADD{other}}
ten
thousand
{\ADD{left}}
alive did the
children of
Judah carry away
captive, and
brought them unto the
top of the
rock, and cast them
down from the
top of the
rock, that they all were broken in
pieces.
\par }{\PP \VS{13}¶ But the
soldiers of the
army which
Amaziah sent
back, that they should not
go with him to
battle,
fell upon the
cities of
Judah, from
Samaria even unto
Beth-horon, and
smote
three
thousand of them, and
took
much
spoil.\FTNT{**}{{\FR 25.13 }
the soldiers…: Heb. the sons of the band}
\par }{\PP \VS{14}¶ Now it came to pass,
after that
Amaziah was
come from the
slaughter of the
Edomites, that he
brought the
gods of the
children of
Seir, and set them
up
{\ADD{to be}} his
gods, and bowed
down himself
before them, and burned
incense unto them.
\VS{15}Wherefore the
anger of the
{\ND{
LORD}} was
kindled against
Amaziah, and he
sent unto him a
prophet, which
said unto him, Why hast thou
sought after the
gods of the
people, which could not
deliver their own
people out of thine
hand?
\VS{16}And it came to pass, as he
talked with him, that
{\ADD{the king}}
said unto him, Art thou
made of the
king’s
counsel?
forbear; why shouldest thou be
smitten? Then the
prophet
forbare, and
said, I
know that
God hath
determined to
destroy thee, because thou hast
done this, and hast not
hearkened unto my
counsel.\FTNT{††}{{\FR 25.16 }
determined: Heb. counselled}
\par }{\PP \VS{17}¶ Then
Amaziah
king of
Judah took
advice, and
sent to
Joash, the
son of
Jehoahaz, the
son of
Jehu,
king of
Israel,
saying,
Come, let us see one
another in the
face.
\VS{18}And
Joash
king of
Israel
sent to
Amaziah
king of
Judah,
saying, The
thistle that
{\ADD{was}} in
Lebanon
sent to the
cedar that
{\ADD{was}} in
Lebanon,
saying,
Give thy
daughter to my
son to
wife: and there passed
by a
wild
beast that
{\ADD{was}} in
Lebanon, and trode
down the
thistle.\FTNT{‡‡}{{\FR 25.18 }
thistle: or, furze bush, or, thorn}\FTNT{§§}{{\FR 25.18 }
a wild…: Heb. a beast of the field}
\VS{19}Thou
sayest, Lo, thou hast
smitten the
Edomites; and thine
heart lifteth thee
up to
boast:
abide now at
home; why shouldest thou
meddle to
{\ADD{thine}}
hurt, that thou shouldest
fall,
{\ADD{even}} thou, and
Judah with thee?
\VS{20}But
Amaziah would not
hear; for it
{\ADD{came}} of
God, that he might
deliver them into the
hand
{\ADD{of their enemies}}, because they
sought after the
gods of
Edom.
\VS{21}So
Joash the
king of
Israel went
up; and they saw one
another in the
face,
{\ADD{both}} he and
Amaziah
king of
Judah, at
Beth-shemesh, which
{\ADD{belongeth}} to
Judah.
\VS{22}And
Judah was put to the
worse
before
Israel, and they
fled every
man to his
tent.\FTNT{***}{{\FR 25.22 }
put…: Heb. smitten}
\VS{23}And
Joash the
king of
Israel
took
Amaziah
king of
Judah, the
son of
Joash, the
son of
Jehoahaz, at
Beth-shemesh, and
brought him to
Jerusalem, and brake
down the
wall of
Jerusalem from the
gate of
Ephraim to the
corner
gate,
four
hundred
cubits.\FTNT{†††}{{\FR 25.23 }
the corner…: Heb. the gate of it that looketh}
\VS{24}And
{\ADD{he took}} all the
gold and the
silver, and all the
vessels that were
found in the
house of
God with
Obed-edom, and the
treasures of the
king’s
house, the
hostages also, and
returned to
Samaria.
\par }{\PP \VS{25}¶ And
Amaziah the
son of
Joash
king of
Judah
lived
after the
death of
Joash
son of
Jehoahaz
king of
Israel
fifteen
years.
\VS{26}Now the
rest of the
acts of
Amaziah,
first and
last, behold,
{\ADD{are}} they not
written in the
book of the
kings of
Judah and
Israel?
\par }{\PP \VS{27}¶ Now after the
time that
Amaziah did turn
away from
following the
{\ND{
LORD}} they
made a
conspiracy against him in
Jerusalem; and he
fled to
Lachish: but they
sent to
Lachish
after him, and
slew him there.\FTNT{‡‡‡}{{\FR 25.27 }
from…: Heb. from after}\FTNT{§§§}{{\FR 25.27 }
made…: Heb. conspired a conspiracy}
\VS{28}And they
brought him upon
horses, and
buried him with his
fathers in the
city of
Judah.\FTNT{*}{{\FR 25.28 }
Judah: that is, the city of David}

\par }\Chap{26}{\PP \VerseOne{1}Then all the
people of
Judah
took
Uzziah, who
{\ADD{was}}
sixteen
years
old, and made him
king in the room of his
father
Amaziah.\FTNT{*}{{\FR 26.1 }
Uzziah: or, Azariah}
\VS{2}He
built
Eloth, and
restored it to
Judah,
after that the
king
slept with his
fathers.
\VS{3}Sixteen
years
old
{\ADD{was}}
Uzziah when he began to
reign, and he
reigned
fifty and
two
years in
Jerusalem. His
mother’s
name also
{\ADD{was}}
Jecoliah of
Jerusalem.
\VS{4}And he
did
{\ADD{that which was}}
right in the
sight of the
{\ND{
LORD}}, according to all that his
father
Amaziah
did.
\VS{5}And he
sought
God in the
days of
Zechariah, who had
understanding in the
visions of
God: and as long
as he
sought the
{\ND{
LORD}},
God made him to
prosper.\FTNT{†}{{\FR 26.5 }
in the visions…: Heb. in the seeing of God}
\VS{6}And he went
forth and
warred against the
Philistines, and brake
down the
wall of
Gath, and the
wall of
Jabneh, and the
wall of
Ashdod, and
built
cities about
Ashdod, and among the
Philistines.\FTNT{‡}{{\FR 26.6 }
about…: or, in the country of Ashdod}
\VS{7}And
God
helped him against the
Philistines, and against the
Arabians that
dwelt in
Gur-baal, and the
Mehunims.
\VS{8}And the
Ammonites
gave
gifts to
Uzziah: and his
name spread
abroad
{\ADD{even}} to the
entering in of
Egypt; for he
strengthened
{\ADD{himself}}
exceedingly.\FTNT{§}{{\FR 26.8 }
spread…: Heb. went}
\VS{9}Moreover
Uzziah
built
towers in
Jerusalem at the
corner
gate, and at the
valley
gate, and at the
turning
{\ADD{of the wall}}, and
fortified them.\FTNT{**}{{\FR 26.9 }
fortified: or, repaired}
\VS{10}Also he
built
towers in the
desert, and
digged
many
wells: for he had
much
cattle, both in the low
country, and in the
plains:
husbandmen
{\ADD{also}}, and vine
dressers in the
mountains, and in
Carmel: for he
loved
husbandry.\FTNT{††}{{\FR 26.10 }
digged…: or, cut out many cisterns}\FTNT{‡‡}{{\FR 26.10 }
Carmel: or, fruitful fields}\FTNT{§§}{{\FR 26.10 }
husbandry: Heb. ground}
\VS{11}Moreover
Uzziah had an
host of
fighting
men, that went
out to
war by
bands, according to the
number of their
account by the
hand of
Jeiel the
scribe and
Maaseiah the
ruler, under the
hand of
Hananiah,
{\ADD{one}} of the
king’s
captains.
\VS{12}The whole
number of the
chief of the
fathers of the mighty
men of
valour
{\ADD{were}} two
thousand and
six
hundred.
\VS{13}And under their
hand
{\ADD{was}} an
army,
three
hundred
thousand and
seven
thousand and
five
hundred, that
made
war with
mighty
power, to
help the
king against the
enemy.\FTNT{***}{{\FR 26.13 }
an army: Heb. the power of an army}
\VS{14}And
Uzziah
prepared for them throughout all the
host
shields, and
spears, and
helmets, and
habergeons, and
bows, and
slings
{\ADD{to cast}}
stones.\FTNT{†††}{{\FR 26.14 }
slings…: Heb. stones of slings}
\VS{15}And he
made in
Jerusalem
engines,
invented by cunning
men, to be on the
towers and upon the
bulwarks, to
shoot
arrows and
great
stones withal. And his
name
spread far
abroad; for he was
marvellously
helped,
till he was
strong.\FTNT{‡‡‡}{{\FR 26.15 }
spread: Heb. went forth}
\par }{\PP \VS{16}¶ But when he was
strong, his
heart was lifted
up to
{\ADD{his}}
destruction: for he
transgressed against the
{\ND{
LORD}} his
God, and
went into the
temple of the
{\ND{
LORD}} to burn
incense upon the
altar of
incense.
\VS{17}And
Azariah the
priest went
in
after him, and with him
fourscore
priests of the
{\ND{
LORD}},
{\ADD{that were}}
valiant
men:
\VS{18}And they
withstood
Uzziah the
king, and
said unto him,
{\ADD{It appertaineth}} not unto thee,
Uzziah, to burn
incense unto the
{\ND{
LORD}}, but to the
priests the
sons of
Aaron, that are
consecrated to burn
incense: go
out of the
sanctuary; for thou hast
trespassed; neither
{\ADD{shall it be}} for thine
honour from the
{\ND{
LORD}}
God.
\VS{19}Then
Uzziah was
wroth, and
{\ADD{had}} a
censer in his
hand to burn
incense: and while he was
wroth with the
priests, the
leprosy even rose
up in his
forehead
before the
priests in the
house of the
{\ND{
LORD}}, from beside the
incense
altar.
\VS{20}And
Azariah the
chief
priest, and all the
priests,
looked upon him, and, behold, he
{\ADD{was}}
leprous in his
forehead, and they thrust him
out from thence; yea, himself
hasted also to go
out, because the
{\ND{
LORD}} had
smitten him.
\VS{21}And
Uzziah the
king was a
leper unto the
day of his
death, and dwelt
in a
several
house,
{\ADD{being}} a
leper; for he was cut
off from the
house of the
{\ND{
LORD}}: and
Jotham his
son
{\ADD{was}} over the
king’s
house,
judging the
people of the
land.\FTNT{§§§}{{\FR 26.21 }
several: Heb. free}
\par }{\PP \VS{22}¶ Now the
rest of the
acts of
Uzziah,
first and
last, did
Isaiah the
prophet, the
son of
Amoz,
write.
\VS{23}So
Uzziah
slept with his
fathers, and they
buried him with his
fathers in the
field of the
burial which
{\ADD{belonged}} to the
kings; for they
said, He
{\ADD{is}} a
leper: and
Jotham his
son
reigned in his stead.

\par }\Chap{27}{\PP \VerseOne{1}Jotham
{\ADD{was}}
twenty and
five
years
old when he began to
reign, and he
reigned
sixteen
years in
Jerusalem. His
mother’s
name also
{\ADD{was}}
Jerushah, the
daughter of
Zadok.
\VS{2}And he
did
{\ADD{that which was}}
right in the
sight of the
{\ND{
LORD}}, according to all that his
father
Uzziah
did: howbeit he
entered not into the
temple of the
{\ND{
LORD}}. And the
people did yet
corruptly.
\VS{3}He
built the
high
gate of the
house of the
{\ND{
LORD}}, and on the
wall of
Ophel he
built
much.\FTNT{*}{{\FR 27.3 }
Ophel: or, the tower}
\VS{4}Moreover he
built
cities in the
mountains of
Judah, and in the
forests he
built
castles and
towers.
\par }{\PP \VS{5}¶ He
fought also with the
king of the
Ammonites, and
prevailed against them. And the
children of
Ammon
gave him the same
year an
hundred
talents of
silver, and
ten
thousand
measures of
wheat, and
ten
thousand of
barley. So much did the
children of
Ammon
pay
unto him, both the
second
year, and the
third.\FTNT{†}{{\FR 27.5 }
So…: Heb. This}
\VS{6}So
Jotham became
mighty, because he
prepared his
ways
before the
{\ND{
LORD}} his
God.\FTNT{‡}{{\FR 27.6 }
prepared: or, established}
\par }{\PP \VS{7}¶ Now the
rest of the
acts of
Jotham, and all his
wars, and his
ways, lo, they
{\ADD{are}}
written in the
book of the
kings of
Israel and
Judah.
\VS{8}He was
five and
twenty
years
old when he began to
reign, and
reigned
sixteen
years in
Jerusalem.
\par }{\PP \VS{9}¶ And
Jotham
slept with his
fathers, and they
buried him in the
city of
David: and
Ahaz his
son
reigned in his stead.

\par }\Chap{28}{\PP \VerseOne{1}Ahaz
{\ADD{was}}
twenty
years
old when he began to
reign, and he
reigned
sixteen
years in
Jerusalem: but he
did not
{\ADD{that which was}}
right in the
sight of the
{\ND{
LORD}}, like
David his
father:
\VS{2}For he walked
in the
ways of the
kings of
Israel, and
made also molten
images for
Baalim.
\VS{3}Moreover he burnt
incense in the
valley of the
son of
Hinnom, and
burnt his
children in the
fire, after the
abominations of the
heathen whom the
{\ND{
LORD}} had cast
out
before the
children of
Israel.\FTNT{*}{{\FR 28.3 }
burnt…: or, offered sacrifice}
\VS{4}He
sacrificed also and burnt
incense in the high
places, and on the
hills, and under every
green
tree.
\VS{5}Wherefore the
{\ND{
LORD}} his
God
delivered him into the
hand of the
king of
Syria; and they
smote him, and carried
away a great
multitude of them
captives, and
brought
{\ADD{them}} to
Damascus. And he was also
delivered into the
hand of the
king of
Israel, who
smote him with a
great
slaughter.\FTNT{†}{{\FR 28.5 }
Damascus: Heb. Darmesek}
\par }{\PP \VS{6}¶ For
Pekah the
son of
Remaliah
slew in
Judah an
hundred and
twenty
thousand in
one
day,
{\ADD{which were}} all
valiant
men; because they had
forsaken the
{\ND{
LORD}}
God of their
fathers.\FTNT{‡}{{\FR 28.6 }
valiant…: Heb. sons of valour}
\VS{7}And
Zichri, a mighty
man of
Ephraim,
slew
Maaseiah the
king’s
son, and
Azrikam the
governor of the
house, and
Elkanah
{\ADD{that was}}
next to the
king.\FTNT{§}{{\FR 28.7 }
next…: Heb. the second to the king}
\VS{8}And the
children of
Israel carried away
captive of their
brethren two
hundred
thousand,
women,
sons, and
daughters, and took also
away
much
spoil from them, and
brought the
spoil to
Samaria.
\VS{9}But a
prophet of the
{\ND{
LORD}} was there, whose
name
{\ADD{was}}
Oded: and he went
out
before the
host that
came to
Samaria, and
said unto them, Behold, because the
{\ND{
LORD}}
God of your
fathers was
wroth with
Judah, he hath
delivered them into your
hand, and ye have
slain them in a
rage
{\ADD{that}} reacheth
up unto
heaven.
\VS{10}And now ye
purpose to keep
under the
children of
Judah and
Jerusalem for
bondmen and
bondwomen unto you:
{\ADD{but are there}} not with you,
even with you,
sins against the
{\ND{
LORD}} your
God?
\VS{11}Now
hear me therefore, and
deliver the
captives again, which ye have taken
captive of your
brethren: for the
fierce
wrath of the
{\ND{
LORD}}
{\ADD{is}} upon you.
\VS{12}Then
certain of the
heads of the
children of
Ephraim,
Azariah the
son of
Johanan,
Berechiah the
son of
Meshillemoth, and
Jehizkiah the
son of
Shallum, and
Amasa the
son of
Hadlai, stood
up against them that
came from the
war,
\VS{13}And
said unto them, Ye shall not bring
in the
captives hither: for whereas we have
offended against the
{\ND{
LORD}}
{\ADD{already}}, ye
intend to
add
{\ADD{more}} to our
sins and to our
trespass: for our
trespass is
great, and
{\ADD{there is}}
fierce
wrath against
Israel.
\VS{14}So the armed
men
left the
captives and the
spoil
before the
princes and all the
congregation.
\VS{15}And the
men which were
expressed by
name rose
up, and
took the
captives, and with the
spoil
clothed all that were
naked among them, and
arrayed them, and
shod them, and gave them to
eat and to
drink, and
anointed them, and
carried all the
feeble of them upon
asses, and
brought them to
Jericho, the
city of palm
trees,
to their
brethren: then they
returned to
Samaria.
\par }{\PP \VS{16}¶ At that
time did
king
Ahaz
send unto the
kings of
Assyria to
help him.
\VS{17}For again the
Edomites had
come and
smitten
Judah, and carried
away
captives.\FTNT{**}{{\FR 28.17 }
captives: Heb. a captivity}
\VS{18}The
Philistines also had
invaded the
cities of the low
country, and of the
south of
Judah, and had
taken
Beth-shemesh, and
Ajalon, and
Gederoth, and
Shocho with the
villages thereof, and
Timnah with the
villages thereof,
Gimzo also and the
villages thereof: and they
dwelt there.
\VS{19}For the
{\ND{
LORD}}
brought
Judah
low because of
Ahaz
king of
Israel; for he
made
Judah
naked, and
transgressed
sore against the
{\ND{
LORD}}.
\VS{20}And
Tilgath-pilneser
king of
Assyria
came unto him, and
distressed him, but
strengthened him not.
\VS{21}For
Ahaz took away a
portion
{\ADD{out}} of the
house of the
{\ND{
LORD}}, and
{\ADD{out}} of the
house of the
king, and of the
princes, and
gave
{\ADD{it}} unto the
king of
Assyria: but he
helped him not.
\par }{\PP \VS{22}¶ And in the
time of his
distress did he
trespass yet
more against the
{\ND{
LORD}}: this
{\ADD{is that}}
king
Ahaz.
\VS{23}For he
sacrificed unto the
gods of
Damascus, which
smote him: and he
said, Because the
gods of the
kings of
Syria
help them,
{\ADD{therefore}} will I
sacrifice to them, that they may
help me. But they were the
ruin of him, and of all
Israel.\FTNT{††}{{\FR 28.23 }
Damascus: Heb. Darmesek}
\VS{24}And
Ahaz gathered
together the
vessels of the
house of
God, and cut in
pieces the
vessels of the
house of
God, and shut
up the
doors of the
house of the
{\ND{
LORD}}, and he
made him
altars in every
corner of
Jerusalem.
\VS{25}And in every several
city of
Judah he
made high
places to burn
incense unto
other
gods, and provoked to
anger the
{\ND{
LORD}}
God of his
fathers.\FTNT{‡‡}{{\FR 28.25 }
to burn: or, to offer}
\par }{\PP \VS{26}¶ Now the
rest of his
acts and of all his
ways,
first and
last, behold, they
{\ADD{are}}
written in the
book of the
kings of
Judah and
Israel.
\VS{27}And
Ahaz
slept with his
fathers, and they
buried him in the
city,
{\ADD{even}} in
Jerusalem: but they
brought him not into the
sepulchres of the
kings of
Israel: and
Hezekiah his
son
reigned in his stead.

\par }\Chap{29}{\PP \VerseOne{1}Hezekiah began to
reign
{\ADD{when he was}}
five and
twenty
years
old, and he
reigned
nine and
twenty
years in
Jerusalem. And his
mother’s
name
{\ADD{was}}
Abijah, the
daughter of
Zechariah.
\VS{2}And he
did
{\ADD{that which was}}
right in the
sight of the
{\ND{
LORD}}, according to all that
David his
father had
done.
\par }{\PP \VS{3}¶ He in the
first
year of his
reign, in the
first
month,
opened the
doors of the
house of the
{\ND{
LORD}}, and
repaired them.
\VS{4}And he brought
in the
priests and the
Levites, and gathered them
together into the
east
street,
\VS{5}And
said unto them,
Hear me, ye
Levites,
sanctify now yourselves, and
sanctify the
house of the
{\ND{
LORD}}
God of your
fathers, and carry
forth the
filthiness out of the
holy
{\ADD{place}}.
\VS{6}For our
fathers have
trespassed, and
done
{\ADD{that which was}}
evil in the
eyes of the
{\ND{
LORD}} our
God, and have
forsaken him, and have turned
away their
faces from the
habitation of the
{\ND{
LORD}}, and
turned
{\ADD{their}}
backs.\FTNT{*}{{\FR 29.6 }
turned their…: Heb. given the neck}
\VS{7}Also they have shut
up the
doors of the
porch, and put
out the
lamps, and have not
burned
incense nor
offered burnt
offerings in the
holy
{\ADD{place}} unto the
God of
Israel.
\VS{8}Wherefore the
wrath of the
{\ND{
LORD}} was upon
Judah and
Jerusalem, and he hath
delivered them to
trouble, to
astonishment, and to
hissing, as ye
see with your
eyes.\FTNT{†}{{\FR 29.8 }
trouble: Heb. commotion}
\VS{9}For, lo, our
fathers have
fallen by the
sword, and our
sons and our
daughters and our
wives
{\ADD{are}} in
captivity for this.
\VS{10}Now
{\ADD{it is}} in mine
heart to
make a
covenant with the
{\ND{
LORD}}
God of
Israel, that his
fierce
wrath may turn
away from us.
\VS{11}My
sons, be not now
negligent: for the
{\ND{
LORD}} hath
chosen you to
stand
before him, to
serve him, and that ye should
minister unto him, and burn
incense.\FTNT{‡}{{\FR 29.11 }
be not…: or, be not now deceived}\FTNT{§}{{\FR 29.11 }
burn…: or, offer sacrifice}
\par }{\PP \VS{12}¶ Then the
Levites
arose,
Mahath the
son of
Amasai, and
Joel the
son of
Azariah, of the
sons of the
Kohathites: and of the
sons of
Merari,
Kish the
son of
Abdi, and
Azariah the
son of
Jehalelel: and of the
Gershonites;
Joah the
son of
Zimmah, and
Eden the
son of
Joah:
\VS{13}And of the
sons of
Elizaphan;
Shimri, and
Jeiel: and of the
sons of
Asaph;
Zechariah, and
Mattaniah:
\VS{14}And of the
sons of
Heman;
Jehiel, and
Shimei: and of the
sons of
Jeduthun;
Shemaiah, and
Uzziel.
\VS{15}And they
gathered their
brethren, and
sanctified themselves, and
came, according to the
commandment of the
king, by the
words of the
{\ND{
LORD}}, to
cleanse the
house of the
{\ND{
LORD}}.\FTNT{**}{{\FR 29.15 }
by the words…: or, in the business of the
{\ND{LORD}}}
\VS{16}And the
priests
went into the inner
part of the
house of the
{\ND{
LORD}}, to
cleanse
{\ADD{it}}, and brought
out all the
uncleanness that they
found in the
temple of the
{\ND{
LORD}} into the
court of the
house of the
{\ND{
LORD}}. And the
Levites
took
{\ADD{it}}, to carry
{\ADD{it}}
out
abroad into the
brook
Kidron.
\VS{17}Now they
began on the
first
{\ADD{day}} of the
first
month to
sanctify, and on the
eighth
day of the
month
came they to the
porch of the
{\ND{
LORD}}: so they
sanctified the
house of the
{\ND{
LORD}} in
eight
days; and in the
sixteenth day of the
first
month they made an
end.
\VS{18}Then they
went
in to
Hezekiah the
king, and
said, We have
cleansed all the
house of the
{\ND{
LORD}}, and the
altar of burnt
offering, with all the
vessels thereof, and the
shewbread
table, with all the
vessels thereof.
\VS{19}Moreover all the
vessels, which
king
Ahaz in his
reign did cast
away in his
transgression, have we
prepared and
sanctified, and, behold, they
{\ADD{are}}
before the
altar of the
{\ND{
LORD}}.
\par }{\PP \VS{20}¶ Then
Hezekiah the
king rose
early, and
gathered the
rulers of the
city, and went
up to the
house of the
{\ND{
LORD}}.
\VS{21}And they
brought
seven
bullocks, and
seven
rams, and
seven
lambs, and
seven
he
goats, for a sin
offering for the
kingdom, and for the
sanctuary, and for
Judah. And he
commanded the
priests the
sons of
Aaron to
offer
{\ADD{them}} on the
altar of the
{\ND{
LORD}}.
\VS{22}So they
killed the
bullocks, and the
priests
received the
blood, and
sprinkled
{\ADD{it}} on the
altar: likewise, when they had
killed the
rams, they
sprinkled the
blood upon the
altar: they
killed also the
lambs, and they
sprinkled the
blood upon the
altar.
\VS{23}And they brought
forth the he
goats
{\ADD{for}} the sin
offering
before the
king and the
congregation; and they
laid their
hands upon them:\FTNT{††}{{\FR 29.23 }
forth: Heb. near}
\VS{24}And the
priests
killed them, and they made
reconciliation with their
blood upon the
altar, to make an
atonement for all
Israel: for the
king
commanded
{\ADD{that}} the burnt
offering and the sin
offering
{\ADD{should be made}} for all
Israel.
\VS{25}And he
set the
Levites in the
house of the
{\ND{
LORD}} with
cymbals, with
psalteries, and with
harps, according to the
commandment of
David, and of
Gad the
king’s
seer, and
Nathan the
prophet: for
{\ADD{so was}} the
commandment
of the
{\ND{
LORD}}
by his
prophets.\FTNT{‡‡}{{\FR 29.25 }
commandment of the
{\ND{LORD}}: Heb. commandment by the hand of the
{\ND{LORD}}}\FTNT{§§}{{\FR 29.25 }
by: Heb. by the hand of}
\VS{26}And the
Levites
stood with the
instruments of
David, and the
priests with the
trumpets.
\VS{27}And
Hezekiah
commanded to
offer the burnt
offering upon the
altar. And
when the burnt
offering
began, the
song of the
{\ND{
LORD}}
began
{\ADD{also}} with the
trumpets, and with the
instruments
{\ADD{ordained}}
by
David
king of
Israel.\FTNT{***}{{\FR 29.27 }
when: Heb. in the time}\FTNT{†††}{{\FR 29.27 }
the instruments: Heb. hands of instruments}
\VS{28}And all the
congregation
worshipped, and the
singers
sang, and the
trumpeters
sounded:
{\ADD{and}} all
{\ADD{this continued}} until the burnt
offering was
finished.\FTNT{‡‡‡}{{\FR 29.28 }
singers: Heb. song}
\VS{29}And when they had made an
end of
offering, the
king and all that were
present with him
bowed themselves, and
worshipped.\FTNT{§§§}{{\FR 29.29 }
present: Heb. found}
\VS{30}Moreover
Hezekiah the
king and the
princes
commanded the
Levites to sing
praise unto the
{\ND{
LORD}} with the
words of
David, and of
Asaph the
seer. And they sang
praises with
gladness, and they bowed their
heads and
worshipped.
\VS{31}Then
Hezekiah
answered and
said, Now ye have
consecrated yourselves unto the
{\ND{
LORD}}, come
near and
bring
sacrifices and thank
offerings into the
house of the
{\ND{
LORD}}. And the
congregation brought
in
sacrifices and thank
offerings; and as many as were of a
free
heart burnt
offerings.\FTNT{*}{{\FR 29.31 }
consecrated…: or, filled your hand}
\VS{32}And the
number of the burnt
offerings, which the
congregation
brought, was threescore and
ten
bullocks, an
hundred
rams,
{\ADD{and}} two
hundred
lambs: all these
{\ADD{were}} for a burnt
offering to the
{\ND{
LORD}}.
\VS{33}And the consecrated
things
{\ADD{were}}
six
hundred
oxen and
three
thousand
sheep.
\VS{34}But the
priests were too
few, so that they
could not
flay all the burnt
offerings: wherefore their
brethren the
Levites did
help them, till the
work was
ended, and until the
{\ADD{other}}
priests had
sanctified themselves: for the
Levites
{\ADD{were}} more
upright in
heart to
sanctify themselves than the
priests.\FTNT{†}{{\FR 29.34 }
did help…: Heb. strengthened them}
\VS{35}And also the burnt
offerings
{\ADD{were}} in
abundance, with the
fat of the peace
offerings, and the drink
offerings for
{\ADD{every}} burnt
offering. So the
service of the
house of the
{\ND{
LORD}} was set in
order.
\VS{36}And
Hezekiah
rejoiced, and all the
people, that
God had
prepared the
people: for the
thing was
{\ADD{done}}
suddenly.

\par }\Chap{30}{\PP \VerseOne{1}And
Hezekiah
sent to all
Israel and
Judah, and
wrote
letters also to
Ephraim and
Manasseh, that they should
come to the
house of the
{\ND{
LORD}} at
Jerusalem, to
keep the
passover unto the
{\ND{
LORD}}
God of
Israel.
\VS{2}For the
king had taken
counsel, and his
princes, and all the
congregation in
Jerusalem, to
keep the
passover in the
second
month.
\VS{3}For they
could not
keep it at that
time, because the
priests had not
sanctified themselves
sufficiently, neither had the
people gathered themselves
together to
Jerusalem.
\VS{4}And the
thing
pleased the
king and all the
congregation.\FTNT{*}{{\FR 30.4 }
pleased…: Heb. was right in the eyes of the king}
\VS{5}So they
established a
decree to make
proclamation throughout all
Israel, from
Beer-sheba even to
Dan, that they should
come to
keep the
passover unto the
{\ND{
LORD}}
God of
Israel at
Jerusalem: for they had not
done
{\ADD{it}} of a
long
{\ADD{time in such sort}} as it was
written.
\VS{6}So the
posts
went with the
letters
from the
king and his
princes throughout all
Israel and
Judah, and according to the
commandment of the
king,
saying, Ye
children of
Israel, turn
again unto the
{\ND{
LORD}}
God of
Abraham,
Isaac, and
Israel, and he will
return to the
remnant of you, that are
escaped out of the
hand of the
kings of
Assyria.\FTNT{†}{{\FR 30.6 }
from: Heb. from the hand}
\VS{7}And be not ye like your
fathers, and like your
brethren, which
trespassed against the
{\ND{
LORD}}
God of their
fathers,
{\ADD{who}} therefore gave them
up to
desolation, as ye
see.
\VS{8}Now be ye not
stiffnecked, as your
fathers
{\ADD{were, but}}
yield
yourselves unto the
{\ND{
LORD}}, and
enter into his
sanctuary, which he hath
sanctified for
ever: and
serve the
{\ND{
LORD}} your
God, that the
fierceness of his
wrath may turn
away from you.\FTNT{‡}{{\FR 30.8 }
be ye…: Heb. harden not your necks}\FTNT{§}{{\FR 30.8 }
yield…: Heb. give the hand}
\VS{9}For if ye turn
again unto the
{\ND{
LORD}}, your
brethren and your
children
{\ADD{shall find}}
compassion
before them that lead them
captive, so that they shall come
again into this
land: for the
{\ND{
LORD}} your
God
{\ADD{is}}
gracious and
merciful, and will not turn
away
{\ADD{his}}
face from you, if ye
return unto him.
\VS{10}So the
posts
passed from
city to
city through the
country of
Ephraim and
Manasseh even unto
Zebulun: but they laughed them to
scorn, and
mocked them.
\VS{11}Nevertheless
divers of
Asher and
Manasseh and of
Zebulun
humbled themselves, and
came to
Jerusalem.
\VS{12}Also in
Judah the
hand of
God was to
give them
one
heart to
do the
commandment of the
king and of the
princes, by the
word of the
{\ND{
LORD}}.
\par }{\PP \VS{13}¶ And there
assembled at
Jerusalem
much
people to
keep the
feast of unleavened
bread in the
second
month, a
very
great
congregation.
\VS{14}And they
arose and took
away the
altars that
{\ADD{were}} in
Jerusalem, and all the altars for
incense took they
away, and
cast
{\ADD{them}} into the
brook
Kidron.
\VS{15}Then they
killed the
passover on the
fourteenth
{\ADD{day}} of the
second
month: and the
priests and the
Levites were
ashamed, and
sanctified themselves, and brought
in the burnt
offerings into the
house of the
{\ND{
LORD}}.
\VS{16}And they
stood in their
place after their
manner, according to the
law of
Moses the
man of
God: the
priests
sprinkled the
blood,
{\ADD{which they received}} of the
hand of the
Levites.\FTNT{**}{{\FR 30.16 }
their place: Heb. their standing}
\VS{17}For
{\ADD{there were}}
many in the
congregation that were not
sanctified: therefore the
Levites had the charge of the
killing of the
passovers for every one
{\ADD{that was}} not
clean, to
sanctify
{\ADD{them}} unto the
{\ND{
LORD}}.
\VS{18}For a
multitude of the
people,
{\ADD{even}}
many of
Ephraim, and
Manasseh,
Issachar, and
Zebulun, had
not
cleansed themselves, yet did they
eat the
passover otherwise than it was
written. But
Hezekiah
prayed for them,
saying, The
good
{\ND{
LORD}}
pardon every one
\VS{19}{\ADD{That}}
prepareth his
heart to
seek
God, the
{\ND{
LORD}}
God of his
fathers, though
{\ADD{he be}} not
{\ADD{cleansed}} according to the
purification of the
sanctuary.
\VS{20}And the
{\ND{
LORD}}
hearkened to
Hezekiah, and
healed the
people.
\par }{\PP \VS{21}And the
children of
Israel that were
present at
Jerusalem
kept the
feast of unleavened
bread
seven
days with
great
gladness: and the
Levites and the
priests
praised the
{\ND{
LORD}}
day by
day,
{\ADD{singing}} with
loud
instruments unto the
{\ND{
LORD}}.\FTNT{††}{{\FR 30.21 }
present: Heb. found}\FTNT{‡‡}{{\FR 30.21 }
loud…: Heb. instruments of strength}
\VS{22}And
Hezekiah
spake
comfortably unto all the
Levites that
taught the
good
knowledge of the
{\ND{
LORD}}: and they did
eat throughout the
feast
seven
days,
offering
peace
offerings, and making
confession to the
{\ND{
LORD}}
God of their
fathers.\FTNT{§§}{{\FR 30.22 }
comfortably…: Heb. to the heart of all}
\VS{23}And the whole
assembly took
counsel to
keep
other
seven
days: and they
kept
{\ADD{other}}
seven
days with
gladness.
\VS{24}For
Hezekiah
king of
Judah did
give to the
congregation a
thousand
bullocks and
seven
thousand
sheep; and the
princes
gave to the
congregation a
thousand
bullocks and
ten
thousand
sheep: and a great
number of
priests
sanctified themselves.\FTNT{***}{{\FR 30.24 }
did give: Heb. lifted up, or, offered}\FTNT{†††}{{\FR 30.24 }
gave: Heb. lifted up, or, offered}
\VS{25}And all the
congregation of
Judah, with the
priests and the
Levites, and all the
congregation that came
out of
Israel, and the
strangers that came
out of the
land of
Israel, and that
dwelt in
Judah,
rejoiced.
\VS{26}So there was
great
joy in
Jerusalem: for since the
time of
Solomon the
son of
David
king of
Israel
{\ADD{there was}} not the like in
Jerusalem.
\par }{\PP \VS{27}¶ Then the
priests the
Levites
arose and
blessed the
people: and their
voice was
heard, and their
prayer
came
{\ADD{up}} to his
holy dwelling
place,
{\ADD{even}} unto
heaven.\FTNT{‡‡‡}{{\FR 30.27 }
his holy…: Heb. the habitation of his holiness}

\par }\Chap{31}{\PP \VerseOne{1}Now when all this was
finished, all
Israel that were
present went
out to the
cities of
Judah, and
brake the
images in
pieces, and cut
down the
groves, and threw
down the high
places and the
altars out of all
Judah and
Benjamin, in
Ephraim also and
Manasseh, until they had utterly
destroyed them all. Then all the
children of
Israel
returned, every
man to his
possession, into their own
cities.\FTNT{*}{{\FR 31.1 }
present: Heb. found}\FTNT{†}{{\FR 31.1 }
images: Heb. statues}\FTNT{‡}{{\FR 31.1 }
until…: Heb. until to make an end}
\par }{\PP \VS{2}¶ And
Hezekiah
appointed the
courses of the
priests and the
Levites after their
courses, every
man
according to his
service, the
priests and
Levites for burnt
offerings and for peace
offerings, to
minister, and to give
thanks, and to
praise in the
gates of the
tents of the
{\ND{
LORD}}.
\VS{3}{\ADD{He appointed}} also the
king’s
portion of his
substance for the burnt
offerings,
{\ADD{to wit}}, for the
morning and
evening burnt
offerings, and the burnt
offerings for the
sabbaths, and for the new
moons, and for the set
feasts, as
{\ADD{it is}}
written in the
law of the
{\ND{
LORD}}.
\VS{4}Moreover he
commanded the
people that
dwelt in
Jerusalem to
give the
portion of the
priests and the
Levites, that they might be
encouraged in the
law of the
{\ND{
LORD}}.
\par }{\PP \VS{5}¶ And as soon as the
commandment came
abroad, the
children of
Israel
brought in
abundance the
firstfruits of
corn,
wine, and
oil, and
honey, and of all the
increase of the
field; and the
tithe of all
{\ADD{things}}
brought they in
abundantly.\FTNT{§}{{\FR 31.5 }
came…: Heb. brake forth}\FTNT{**}{{\FR 31.5 }
honey: or, dates}
\VS{6}And
{\ADD{concerning}} the
children of
Israel and
Judah, that
dwelt in the
cities of
Judah, they also brought
in the
tithe of
oxen and
sheep, and the
tithe of holy
things which were
consecrated unto the
{\ND{
LORD}} their
God, and
laid
{\ADD{them}} by
heaps.\FTNT{††}{{\FR 31.6 }
by heaps: Heb. heaps, heaps}
\VS{7}In the
third
month they
began to lay the
foundation of the
heaps, and
finished
{\ADD{them}} in the
seventh
month.
\VS{8}And when
Hezekiah and the
princes
came and
saw the
heaps, they
blessed the
{\ND{
LORD}}, and his
people
Israel.
\VS{9}Then
Hezekiah
questioned with the
priests and the
Levites concerning the
heaps.
\VS{10}And
Azariah the
chief
priest of the
house of
Zadok
answered him, and
said, Since
{\ADD{the people}}
began to
bring the
offerings into the
house of the
{\ND{
LORD}}, we have had
enough to
eat, and have
left
plenty: for the
{\ND{
LORD}} hath
blessed his
people; and that which is
left
{\ADD{is}} this great
store.
\par }{\PP \VS{11}¶ Then
Hezekiah
commanded to
prepare
chambers in the
house of the
{\ND{
LORD}}; and they
prepared
{\ADD{them}},\FTNT{‡‡}{{\FR 31.11 }
chambers: or, storehouses}
\VS{12}And brought
in the
offerings and the
tithes and the
dedicated
{\ADD{things}}
faithfully: over which
Cononiah the
Levite
{\ADD{was}}
ruler, and
Shimei his
brother
{\ADD{was}} the
next.
\VS{13}And
Jehiel, and
Azaziah, and
Nahath, and
Asahel, and
Jerimoth, and
Jozabad, and
Eliel, and
Ismachiah, and
Mahath, and
Benaiah,
{\ADD{were}}
overseers under the
hand of
Cononiah and
Shimei his
brother, at the
commandment of
Hezekiah the
king, and
Azariah the
ruler of the
house of
God.\FTNT{§§}{{\FR 31.13 }
under…: Heb. at the hand}
\VS{14}And
Kore the
son of
Imnah the
Levite, the
porter toward the
east,
{\ADD{was}} over the freewill
offerings of
God, to
distribute the
oblations of the
{\ND{
LORD}}, and the
most holy
things.
\VS{15}And next him
{\ADD{were}}
Eden, and
Miniamin, and
Jeshua, and
Shemaiah,
Amariah, and
Shecaniah, in the
cities of the
priests, in
{\ADD{their}} set
office, to
give to their
brethren
by
courses, as well to the
great as to the
small:\FTNT{***}{{\FR 31.15 }
next…: Heb. at his hand}\FTNT{†††}{{\FR 31.15 }
set…: or, trust}
\VS{16}Beside their
genealogy of
males, from
three
years
old and
upward,
{\ADD{even}} unto every one that
entereth into the
house of the
{\ND{
LORD}}, his
daily
portion for their
service in their
charges according to their
courses;
\VS{17}Both to the
genealogy of the
priests by the
house of their
fathers, and the
Levites from
twenty
years
old and
upward, in their
charges by their
courses;
\VS{18}And to the
genealogy of all their little
ones, their
wives, and their
sons, and their
daughters, through all the
congregation: for in their set
office they
sanctified themselves in
holiness:\FTNT{‡‡‡}{{\FR 31.18 }
set…: or, trust}
\VS{19}Also of the
sons of
Aaron the
priests,
{\ADD{which were}} in the
fields of the
suburbs of their
cities, in every several
city, the
men that were
expressed by
name, to
give
portions to all the
males among the
priests, and to all that were reckoned by
genealogies among the
Levites.
\par }{\PP \VS{20}¶ And thus
did
Hezekiah throughout all
Judah, and
wrought
{\ADD{that which was}}
good and
right and
truth
before the
{\ND{
LORD}} his
God.
\VS{21}And in every
work that he
began in the
service of the
house of
God, and in the
law, and in the
commandments, to
seek his
God, he
did
{\ADD{it}} with all his
heart, and
prospered.

\par }\Chap{32}{\PP \VerseOne{1}After these
things, and the
establishment thereof,
Sennacherib
king of
Assyria
came, and
entered into
Judah, and
encamped against the
fenced
cities, and
thought to
win them for himself.\FTNT{*}{{\FR 32.1 }
to win…: Heb. to break them up}
\VS{2}And when
Hezekiah
saw that
Sennacherib was
come, and that he was
purposed to
fight against
Jerusalem,\FTNT{†}{{\FR 32.2 }
he was…: Heb. his face was to war}
\VS{3}He took
counsel with his
princes and his mighty
men to
stop the
waters of the
fountains which
{\ADD{were}}
without the
city: and they did
help him.
\VS{4}So there was
gathered
much
people
together, who
stopped all the
fountains, and the
brook that
ran through the
midst of the
land,
saying, Why should the
kings of
Assyria
come, and
find
much
water?\FTNT{‡}{{\FR 32.4 }
ran: Heb. overflowed}
\VS{5}Also he
strengthened himself, and built
up all the
wall that was
broken, and raised
{\ADD{it}}
up to the
towers, and
another
wall
without, and
repaired
Millo
{\ADD{in}} the
city of
David, and
made
darts and
shields in
abundance.\FTNT{§}{{\FR 32.5 }
darts: or, swords, or, weapons}
\VS{6}And he
set
captains of
war over the
people, and gathered them
together to him in the
street of the
gate of the
city, and
spake
comfortably to them,
saying,\FTNT{**}{{\FR 32.6 }
spake…: Heb. he spoke to their heart}
\VS{7}Be
strong and
courageous, be not
afraid nor
dismayed
for the
king of
Assyria, nor for all the
multitude that
{\ADD{is}} with him: for
{\ADD{there be}}
more with us than with him:
\VS{8}With him
{\ADD{is}} an
arm of
flesh; but with us
{\ADD{is}} the
{\ND{
LORD}} our
God to
help us, and to
fight our
battles. And the
people
rested themselves upon the
words of
Hezekiah
king of
Judah.\FTNT{††}{{\FR 32.8 }
rested…: Heb. leaned}
\par }{\PP \VS{9}¶
After this did
Sennacherib
king of
Assyria
send his
servants to
Jerusalem, (but he
{\ADD{himself laid siege}} against
Lachish, and all his
power with him,) unto
Hezekiah
king of
Judah, and unto all
Judah that
{\ADD{were}} at
Jerusalem,
saying,\FTNT{‡‡}{{\FR 32.9 }
power: Heb. dominion}
\VS{10}Thus
saith
Sennacherib
king of
Assyria, Whereon do ye
trust, that ye
abide in the
siege in
Jerusalem?\FTNT{§§}{{\FR 32.10 }
siege: or, strong hold}
\VS{11}Doth not
Hezekiah
persuade you to
give over yourselves to
die by
famine and by
thirst,
saying, The
{\ND{
LORD}} our
God shall
deliver us out of the
hand of the
king of
Assyria?
\VS{12}Hath not the same
Hezekiah taken
away his high
places and his
altars, and
commanded
Judah and
Jerusalem,
saying, Ye shall
worship
before
one
altar, and burn
incense upon it?
\VS{13}Know ye not what I and my
fathers have
done unto all the
people of
{\ADD{other}}
lands? were the
gods of the
nations of those
lands any
ways
able to
deliver their
lands out of mine
hand?
\VS{14}Who
{\ADD{was there}} among all the
gods of those
nations that my
fathers utterly
destroyed, that
could
deliver his
people out of mine
hand, that your
God should be
able to
deliver you out of mine
hand?
\VS{15}Now therefore let
not
Hezekiah
deceive you, nor
persuade you on this manner, neither yet
believe him: for no
god of any
nation or
kingdom was
able to
deliver his
people out of mine
hand, and out of the
hand of my
fathers: how much less shall your
God
deliver you out of mine
hand?
\VS{16}And his
servants
spake yet
{\ADD{more}} against the
{\ND{
LORD}}
God, and against his
servant
Hezekiah.
\VS{17}He
wrote also
letters to
rail on the
{\ND{
LORD}}
God of
Israel, and to
speak against him,
saying, As the
gods of the
nations of
{\ADD{other}}
lands have not
delivered their
people out of mine
hand, so shall not the
God of
Hezekiah
deliver his
people out of mine
hand.
\VS{18}Then they
cried with a
loud
voice in the
Jews’ speech unto the
people of
Jerusalem that
{\ADD{were}} on the
wall, to
affright them, and to
trouble them; that they might
take the
city.
\VS{19}And they
spake
against the
God of
Jerusalem, as against the
gods of the
people of the
earth,
{\ADD{which were}} the
work of the
hands of
man.
\VS{20}And for this
{\ADD{cause}}
Hezekiah the
king, and the
prophet
Isaiah the
son of
Amoz,
prayed and
cried to
heaven.
\par }{\PP \VS{21}¶ And the
{\ND{
LORD}}
sent an
angel, which cut
off all the mighty
men of
valour, and the
leaders and
captains in the
camp of the
king of
Assyria. So he
returned with
shame of
face to his own
land. And when he was
come into the
house of his
god, they that came
forth of his own
bowels
slew him there with the
sword.\FTNT{***}{{\FR 32.21 }
slew…: Heb. made him fall}
\VS{22}Thus the
{\ND{
LORD}}
saved
Hezekiah and the
inhabitants of
Jerusalem from the
hand of
Sennacherib the
king of
Assyria, and from the
hand of all
{\ADD{other}}, and
guided them on every
side.
\VS{23}And
many
brought
gifts unto the
{\ND{
LORD}} to
Jerusalem, and
presents to
Hezekiah
king of
Judah: so that he was
magnified in the
sight of all
nations from
thenceforth.\FTNT{†††}{{\FR 32.23 }
presents: Heb. precious things}
\par }{\PP \VS{24}¶ In those
days
Hezekiah was
sick to the
death, and
prayed unto the
{\ND{
LORD}}: and he
spake unto him, and he
gave him a
sign.\FTNT{‡‡‡}{{\FR 32.24 }
gave…: or, wrought a miracle for him}
\VS{25}But
Hezekiah rendered not
again according to the
benefit
{\ADD{done}} unto him; for his
heart was lifted
up: therefore there was
wrath upon him, and upon
Judah and
Jerusalem.
\VS{26}Notwithstanding
Hezekiah
humbled himself for the
pride of his
heart,
{\ADD{both}} he and the
inhabitants of
Jerusalem, so that the
wrath of the
{\ND{
LORD}}
came not upon them in the
days of
Hezekiah.\FTNT{§§§}{{\FR 32.26 }
pride: Heb. lifting up}
\par }{\PP \VS{27}¶ And
Hezekiah had
exceeding
much
riches and
honour: and he
made himself
treasuries for
silver, and for
gold, and for
precious
stones, and for
spices, and for
shields, and for all manner of
pleasant
jewels;\FTNT{*}{{\FR 32.27 }
pleasant…: Heb. instruments of desire}
\VS{28}Storehouses also for the
increase of
corn, and
wine, and
oil; and
stalls for all manner of
beasts, and
cotes for
flocks.
\VS{29}Moreover he
provided him
cities, and
possessions of
flocks and
herds in
abundance: for
God had
given him
substance
very
much.
\VS{30}This same
Hezekiah also
stopped the
upper
watercourse of
Gihon, and brought it
straight
down to the west
side of the
city of
David. And
Hezekiah
prospered in all his
works.
\par }{\PP \VS{31}¶
Howbeit in
{\ADD{the business of}} the
ambassadors of the
princes of
Babylon, who
sent unto him to
enquire of the
wonder that was
{\ADD{done}} in the
land,
God
left him, to
try him, that he might
know all
{\ADD{that was}} in his
heart.\FTNT{†}{{\FR 32.31 }
ambassadors: Heb. interpreters}
\par }{\PP \VS{32}¶ Now the
rest of the
acts of
Hezekiah, and his
goodness, behold, they
{\ADD{are}}
written in the
vision of
Isaiah the
prophet, the
son of
Amoz,
{\ADD{and}} in the
book of the
kings of
Judah and
Israel.\FTNT{‡}{{\FR 32.32 }
goodness: Heb. kindnesses}
\VS{33}And
Hezekiah
slept with his
fathers, and they
buried him in the
chiefest of the
sepulchres of the
sons of
David: and all
Judah and the
inhabitants of
Jerusalem did
him
honour at his
death. And
Manasseh his
son
reigned in his stead.\FTNT{§}{{\FR 32.33 }
chiefest: or, highest}

\par }\Chap{33}{\PP \VerseOne{1}Manasseh
{\ADD{was}}
twelve
years
old when he began to
reign, and he
reigned
fifty and
five
years in
Jerusalem:
\VS{2}But
did
{\ADD{that which was}}
evil in the
sight of the
{\ND{
LORD}}, like unto the
abominations of the
heathen, whom the
{\ND{
LORD}} had cast
out
before the
children of
Israel.
\par }{\PP \VS{3}¶ For he
built
again the high
places which
Hezekiah his
father had broken
down, and he reared
up
altars for
Baalim, and
made
groves, and
worshipped all the
host of
heaven, and
served them.\FTNT{*}{{\FR 33.3 }
he built…: Heb. he returned and built}
\VS{4}Also he
built
altars in the
house of the
{\ND{
LORD}}, whereof the
{\ND{
LORD}} had
said, In
Jerusalem shall my
name be for
ever.
\VS{5}And he
built
altars for all the
host of
heaven in the
two
courts of the
house of the
{\ND{
LORD}}.
\VS{6}And he
caused his
children to pass
through the
fire in the
valley of the
son of
Hinnom: also he observed
times, and used
enchantments, and used
witchcraft, and
dealt with a familiar
spirit, and with
wizards: he
wrought
much
evil in the
sight of the
{\ND{
LORD}}, to provoke him to
anger.
\VS{7}And he
set a carved
image, the
idol which he had
made, in the
house of
God, of which
God had
said to
David and to
Solomon his
son, In this
house, and in
Jerusalem, which I have
chosen before all the
tribes of
Israel, will I
put my
name for
ever:
\VS{8}Neither will I any
more
remove the
foot of
Israel from
out of the
land which I have
appointed for your
fathers; so
that they will take
heed to
do all that I have
commanded them, according to the whole
law and the
statutes and the
ordinances by the
hand of
Moses.
\VS{9}So
Manasseh
made
Judah and the
inhabitants of
Jerusalem to
err,
{\ADD{and}} to
do
worse than the
heathen, whom the
{\ND{
LORD}} had
destroyed
before the
children of
Israel.
\VS{10}And the
{\ND{
LORD}}
spake to
Manasseh, and to his
people: but they would not
hearken.
\par }{\PP \VS{11}¶ Wherefore the
{\ND{
LORD}}
brought upon them the captains
of the host
of the
king of
Assyria, which
took
Manasseh among the
thorns, and bound
him with
fetters, and
carried him to
Babylon.\FTNT{†}{{\FR 33.11 }
of the king: Heb. which were the king’s}\FTNT{‡}{{\FR 33.11 }
fetters: or, chains}
\VS{12}And when he was in
affliction, he
besought the
{\ND{
LORD}} his
God, and
humbled himself
greatly
before the
God of his
fathers,
\VS{13}And
prayed unto him: and he was
intreated of him, and
heard his
supplication, and brought him
again to
Jerusalem into his
kingdom. Then
Manasseh
knew that the
{\ND{
LORD}} he
{\ADD{was}}
God.
\VS{14}Now after
this he
built a
wall
without the
city of
David, on the west
side of
Gihon, in the
valley, even to the entering
in at the
fish
gate, and
compassed about
Ophel, and raised it
up a very great
height, and
put
captains of
war in all the
fenced
cities of
Judah.\FTNT{§}{{\FR 33.14 }
Ophel: or, the tower}
\VS{15}And he took
away the
strange
gods, and the
idol out of the
house of the
{\ND{
LORD}}, and all the
altars that he had
built in the
mount of the
house of the
{\ND{
LORD}}, and in
Jerusalem, and
cast
{\ADD{them}}
out of the
city.
\VS{16}And he
repaired the
altar of the
{\ND{
LORD}}, and
sacrificed thereon
peace
offerings and thank
offerings, and
commanded
Judah to
serve the
{\ND{
LORD}}
God of
Israel.
\VS{17}Nevertheless the
people did
sacrifice still in the high
places,
{\ADD{yet}} unto the
{\ND{
LORD}} their
God only.
\par }{\PP \VS{18}¶ Now the
rest of the
acts of
Manasseh, and his
prayer unto his
God, and the
words of the
seers that
spake to him in the
name of the
{\ND{
LORD}}
God of
Israel, behold, they
{\ADD{are written}} in the
book of the
kings of
Israel.
\VS{19}His
prayer also, and
{\ADD{how God}} was
intreated of him, and all his
sin, and his
trespass, and the
places wherein he
built high
places, and set
up
groves and graven
images,
before he was
humbled: behold, they
{\ADD{are}}
written among the
sayings of the
seers.\FTNT{**}{{\FR 33.19 }
the seers: or, Hosai}
\par }{\PP \VS{20}¶ So
Manasseh
slept with his
fathers, and they
buried him in his own
house: and
Amon his
son
reigned in his stead.
\par }{\PP \VS{21}¶
Amon
{\ADD{was}}
two and
twenty
years
old when he began to
reign, and
reigned
two
years in
Jerusalem.
\VS{22}But he
did
{\ADD{that which was}}
evil in the
sight of the
{\ND{
LORD}}, as
did
Manasseh his
father: for
Amon
sacrificed unto all the carved
images which
Manasseh his
father had
made, and
served them;
\VS{23}And
humbled not himself
before the
{\ND{
LORD}}, as
Manasseh his
father had
humbled himself; but
Amon
trespassed more and
more.\FTNT{††}{{\FR 33.23 }
trespassed more and more: Heb. multiplied trespass}
\VS{24}And his
servants
conspired against him, and
slew him in his own
house.
\par }{\PP \VS{25}¶ But the
people of the
land
slew all them that had
conspired against
king
Amon; and the
people of the
land
made
Josiah his
son
king in his stead.

\par }\Chap{34}{\PP \VerseOne{1}Josiah
{\ADD{was}}
eight
years
old when he began to
reign, and he
reigned in
Jerusalem
one and
thirty
years.
\VS{2}And he
did
{\ADD{that which was}}
right in the
sight of the
{\ND{
LORD}}, and
walked in the
ways of
David his
father, and
declined
{\ADD{neither}} to the right
hand, nor to the
left.
\par }{\PP \VS{3}¶ For in the
eighth
year of his
reign, while he was yet
young, he
began to
seek after the
God of
David his
father: and in the
twelfth
year he
began to
purge
Judah and
Jerusalem from the high
places, and the
groves, and the carved
images, and the molten
images.
\VS{4}And they brake
down the
altars of
Baalim in his
presence; and the
images, that
{\ADD{were}} on high above
them, he cut
down; and the
groves, and the carved
images, and the molten
images, he brake in
pieces, and made
dust
{\ADD{of them}}, and
strowed
{\ADD{it}}
upon the
graves of them that had
sacrificed unto them.\FTNT{*}{{\FR 34.4 }
the images: or, the sun images}\FTNT{†}{{\FR 34.4 }
graves: Heb. face of the graves}
\VS{5}And he
burnt the
bones of the
priests upon their
altars, and
cleansed
Judah and
Jerusalem.
\VS{6}And
{\ADD{so did he}} in the
cities of
Manasseh, and
Ephraim, and
Simeon, even unto
Naphtali, with their
mattocks round
about.\FTNT{‡}{{\FR 34.6 }
mattocks: or, mauls}
\VS{7}And when he had broken
down the
altars and the
groves, and had
beaten the graven
images into
powder, and cut
down all the
idols throughout all the
land of
Israel, he
returned to
Jerusalem.\FTNT{§}{{\FR 34.7 }
into…: Heb. to make powder}
\par }{\PP \VS{8}¶ Now in the
eighteenth
year of his
reign, when he had
purged the
land, and the
house, he
sent
Shaphan the
son of
Azaliah, and
Maaseiah the
governor of the
city, and
Joah the
son of
Joahaz the
recorder, to
repair the
house of the
{\ND{
LORD}} his
God.
\VS{9}And when they
came to
Hilkiah the
high
priest, they
delivered the
money that was
brought into the
house of
God, which the
Levites that
kept the
doors had
gathered of the
hand of
Manasseh and
Ephraim, and of all the
remnant of
Israel, and of all
Judah and
Benjamin; and they
returned to
Jerusalem.
\VS{10}And they
put
{\ADD{it}} in the
hand of the
workmen that had the
oversight of the
house of the
{\ND{
LORD}}, and they
gave it to the
workmen that
wrought in the
house of the
{\ND{
LORD}}, to
repair and
amend the
house:
\VS{11}Even to the
artificers and
builders
gave they
{\ADD{it}}, to
buy
hewn
stone, and
timber for
couplings, and to
floor the
houses which the
kings of
Judah had
destroyed.\FTNT{**}{{\FR 34.11 }
to floor: or, to rafter}
\VS{12}And the
men
did the
work
faithfully: and the
overseers of them
{\ADD{were}}
Jahath and
Obadiah, the
Levites, of the
sons of
Merari; and
Zechariah and
Meshullam, of the sons of the
Kohathites, to set
{\ADD{it}}
forward; and
{\ADD{other of}} the
Levites, all that could
skill of
instruments of
musick.
\VS{13}Also
{\ADD{they were}} over the bearers of
burdens, and
{\ADD{were}}
overseers of all that
wrought the
work in any manner of
service: and of the
Levites
{\ADD{there were}}
scribes, and
officers, and
porters.
\par }{\PP \VS{14}¶ And when they brought
out the
money that was brought
into the
house of the
{\ND{
LORD}},
Hilkiah the
priest
found a
book of the
law of the
{\ND{
LORD}}
{\ADD{given}}
by
Moses.\FTNT{††}{{\FR 34.14 }
by: Heb. by the hand of}
\VS{15}And
Hilkiah
answered and
said to
Shaphan the
scribe, I have
found the
book of the
law in the
house of the
{\ND{
LORD}}. And
Hilkiah
delivered the
book to
Shaphan.
\VS{16}And
Shaphan
carried the
book to the
king, and
brought the
king
word
back again,
saying, All that was
committed
to thy
servants, they
do
{\ADD{it}}.\FTNT{‡‡}{{\FR 34.16 }
to thy…: Heb. to the hand of, etc}
\VS{17}And they have gathered
together the
money that was
found in the
house of the
{\ND{
LORD}}, and have
delivered it into the
hand of the
overseers, and to the
hand of the
workmen.\FTNT{§§}{{\FR 34.17 }
gathered together: Heb. poured out, or, melted}
\VS{18}Then
Shaphan the
scribe
told the
king,
saying,
Hilkiah the
priest hath
given me a
book. And
Shaphan
read it
before the
king.\FTNT{***}{{\FR 34.18 }
it: Heb. in it}
\VS{19}And it came to pass, when the
king had
heard the
words of the
law, that he
rent his
clothes.
\VS{20}And the
king
commanded
Hilkiah, and
Ahikam the
son of
Shaphan, and
Abdon the
son of
Micah, and
Shaphan the
scribe, and
Asaiah a
servant of the
king’s,
saying,\FTNT{†††}{{\FR 34.20 }
Abdon: or, Achbor}
\VS{21}Go,
enquire of the
{\ND{
LORD}} for me, and for them that are
left in
Israel and in
Judah, concerning the
words of the
book that is
found: for
great
{\ADD{is}} the
wrath of the
{\ND{
LORD}} that is poured
out upon us, because our
fathers have not
kept the
word of the
{\ND{
LORD}}, to
do after all that is
written in this
book.
\VS{22}And
Hilkiah, and
{\ADD{they}} that the
king
{\ADD{had appointed}},
went to
Huldah the
prophetess, the
wife of
Shallum the
son of
Tikvath, the
son of
Hasrah,
keeper of the
wardrobe; (now she
dwelt in
Jerusalem in the
college:) and they
spake to her to that
{\ADD{effect}}.\FTNT{‡‡‡}{{\FR 34.22 }
Hasrah: also called, Harhas}\FTNT{§§§}{{\FR 34.22 }
wardrobe: Heb. garments}\FTNT{*}{{\FR 34.22 }
in the college: or, in the school, or, in the second part}
\par }{\PP \VS{23}¶ And she
answered them, Thus
saith the
{\ND{
LORD}}
God of
Israel,
Tell ye the
man that
sent you to me,
\VS{24}Thus
saith the
{\ND{
LORD}}, Behold, I will
bring
evil upon this
place, and upon the
inhabitants thereof,
{\ADD{even}} all the
curses that are
written in the
book which they have
read
before the
king of
Judah:
\VS{25}Because they have
forsaken me, and have burned
incense unto
other
gods, that they might provoke me to
anger with all the
works of their
hands; therefore my
wrath shall be poured
out upon this
place, and shall not be
quenched.
\VS{26}And as
for the
king of
Judah, who
sent you to
enquire of the
{\ND{
LORD}}, so shall ye
say unto him, Thus
saith the
{\ND{
LORD}}
God of
Israel
{\ADD{concerning}} the
words which thou hast
heard;
\VS{27}Because thine
heart was
tender, and thou didst
humble thyself
before
God, when thou
heardest his
words against this
place, and against the
inhabitants thereof, and
humbledst thyself
before me, and didst
rend thy
clothes, and
weep
before me; I have even
heard
{\ADD{thee}} also,
saith the
{\ND{
LORD}}.
\VS{28}Behold, I will
gather thee to thy
fathers, and thou shalt be
gathered to thy
grave in
peace, neither shall thine
eyes
see all the
evil that I will
bring upon this
place, and upon the
inhabitants of the same. So they
brought the
king
word
again.
\par }{\PP \VS{29}¶ Then the
king
sent and gathered
together all the
elders of
Judah and
Jerusalem.
\VS{30}And the
king went
up into the
house of the
{\ND{
LORD}}, and all the
men of
Judah, and the
inhabitants of
Jerusalem, and the
priests, and the
Levites, and all the
people,
great and
small: and he
read in their
ears all the
words of the
book of the
covenant that was
found in the
house of the
{\ND{
LORD}}.\FTNT{†}{{\FR 34.30 }
great…: Heb. from great even to small}
\VS{31}And the
king
stood in his
place, and
made a
covenant
before the
{\ND{
LORD}}, to
walk
after the
{\ND{
LORD}}, and to
keep his
commandments, and his
testimonies, and his
statutes, with all his
heart, and with all his
soul, to
perform the
words of the
covenant which are
written in this
book.
\VS{32}And he caused all that were
present in
Jerusalem and
Benjamin to
stand
{\ADD{to it}}. And the
inhabitants of
Jerusalem
did according to the
covenant of
God, the
God of their
fathers.\FTNT{‡}{{\FR 34.32 }
present: Heb. found}
\VS{33}And
Josiah took
away all the
abominations out of all the
countries that
{\ADD{pertained}} to the
children of
Israel, and made all that were
present in
Israel to
serve,
{\ADD{even}} to
serve the
{\ND{
LORD}} their
God.
{\ADD{And}} all his
days they
departed not from
following the
{\ND{
LORD}}, the
God of their
fathers.\FTNT{§}{{\FR 34.33 }
present: Heb. found}\FTNT{**}{{\FR 34.33 }
from…: Heb. from after}

\par }\Chap{35}{\PP \VerseOne{1}Moreover
Josiah
kept a
passover unto the
{\ND{
LORD}} in
Jerusalem: and they
killed the
passover on the
fourteenth
{\ADD{day}} of the
first
month.
\VS{2}And he
set the
priests in their
charges, and
encouraged them to the
service of the
house of the
{\ND{
LORD}},
\VS{3}And
said unto the
Levites that
taught all
Israel, which were
holy unto the
{\ND{
LORD}},
Put the
holy
ark in the
house which
Solomon the
son of
David
king of
Israel did
build;
{\ADD{it shall}} not
{\ADD{be}} a
burden upon
{\ADD{your}}
shoulders:
serve now the
{\ND{
LORD}} your
God, and his
people
Israel,
\VS{4}And
prepare
{\ADD{yourselves}} by the
houses of your
fathers, after your
courses, according to the
writing of
David
king of
Israel, and according to the
writing of
Solomon his
son.
\VS{5}And stand
in the
holy
{\ADD{place}} according to the
divisions of the
families of the
fathers of your
brethren the
people, and
{\ADD{after}} the
division of the
families of the
Levites.\FTNT{*}{{\FR 35.5 }
the families of the fathers: Heb. the house of the fathers}\FTNT{†}{{\FR 35.5 }
the people: Heb. the sons of the people}
\VS{6}So
kill the
passover, and
sanctify yourselves, and
prepare your
brethren, that
{\ADD{they}} may
do according to the
word of the
{\ND{
LORD}} by the
hand of
Moses.
\VS{7}And
Josiah
gave to the
people, of the
flock,
lambs and
kids, all for the passover
offerings, for all that were
present, to the
number of
thirty
thousand, and
three
thousand
bullocks: these
{\ADD{were}} of the
king’s
substance.\FTNT{‡}{{\FR 35.7 }
gave: Heb. offered}
\VS{8}And his
princes
gave
willingly unto the
people, to the
priests, and to the
Levites:
Hilkiah and
Zechariah and
Jehiel,
rulers of the
house of
God,
gave unto the
priests for the passover
offerings two
thousand and
six
hundred
{\ADD{small cattle}}, and
three
hundred
oxen.\FTNT{§}{{\FR 35.8 }
gave willingly: Heb. offered, etc}
\VS{9}Conaniah also, and
Shemaiah and
Nethaneel, his
brethren, and
Hashabiah and
Jeiel and
Jozabad,
chief of the
Levites,
gave unto the
Levites for passover
offerings
five
thousand
{\ADD{small cattle}}, and
five
hundred
oxen.\FTNT{**}{{\FR 35.9 }
gave: Heb. offered}
\VS{10}So the
service was
prepared, and the
priests
stood in their
place, and the
Levites in their
courses, according to the
king’s
commandment.
\VS{11}And they
killed the
passover, and the
priests
sprinkled
{\ADD{the blood}} from their
hands, and the
Levites
flayed
{\ADD{them}}.
\VS{12}And they
removed the burnt
offerings, that they might
give according to the
divisions of the
families of the
people, to
offer unto the
{\ND{
LORD}}, as
{\ADD{it is}}
written in the
book of
Moses. And so
{\ADD{did they}} with the
oxen.
\VS{13}And they
roasted the
passover with
fire according to the
ordinance: but the
{\ADD{other}}
holy
{\ADD{offerings}}
sod they in
pots, and in
caldrons, and in
pans, and divided
{\ADD{them}}
speedily among all the
people.\FTNT{††}{{\FR 35.13 }
divided…: Heb. made them run}
\VS{14}And
afterward they made
ready for themselves, and for the
priests: because the
priests the
sons of
Aaron
{\ADD{were busied}} in
offering of burnt
offerings and the
fat until
night; therefore the
Levites
prepared for themselves, and for the
priests the
sons of
Aaron.
\VS{15}And the
singers the
sons of
Asaph
{\ADD{were}} in their
place, according to the
commandment of
David, and
Asaph, and
Heman, and
Jeduthun the
king’s
seer; and the
porters
{\ADD{waited}} at every
gate; they might not
depart from their
service; for their
brethren the
Levites
prepared for them.\FTNT{‡‡}{{\FR 35.15 }
place: Heb. station}
\VS{16}So all the
service of the
{\ND{
LORD}} was
prepared the same
day, to
keep the
passover, and to
offer burnt
offerings upon the
altar of the
{\ND{
LORD}}, according to the
commandment of
king
Josiah.
\VS{17}And the
children of
Israel that were
present
kept the
passover at that
time, and the
feast of unleavened
bread
seven
days.\FTNT{§§}{{\FR 35.17 }
present: Heb. found}
\VS{18}And there was no
passover
like to that
kept in
Israel from the
days of
Samuel the
prophet; neither
did all the
kings of
Israel
keep such a
passover as
Josiah
kept, and the
priests, and the
Levites, and all
Judah and
Israel that were
present, and the
inhabitants of
Jerusalem.\FTNT{***}{{\FR 35.18 }
present: Heb. found}
\VS{19}In the
eighteenth
year of the
reign of
Josiah was this
passover
kept.
\par }{\PP \VS{20}¶
After all this, when
Josiah had
prepared the
temple,
Necho
king of
Egypt came
up to
fight against
Carchemish by
Euphrates: and
Josiah went
out
against him.\FTNT{†††}{{\FR 35.20 }
temple: Heb. house}
\VS{21}But he
sent
ambassadors to him,
saying, What have I to do with thee, thou
king of
Judah?
{\ADD{I come}} not against thee this
day, but against the
house wherewith I have
war: for
God
commanded me to make
haste:
forbear thee from
{\ADD{meddling with}}
God, who
{\ADD{is}} with me, that he
destroy thee not.\FTNT{‡‡‡}{{\FR 35.21 }
the house…: Heb. the house of my war}
\VS{22}Nevertheless
Josiah would not
turn his
face from him, but
disguised himself, that he might
fight with him, and
hearkened not unto the
words of
Necho from the
mouth of
God, and
came to
fight in the
valley of
Megiddo.
\VS{23}And the
archers
shot at
king
Josiah; and the
king
said to his
servants, Have me
away; for I am
sore
wounded.\FTNT{§§§}{{\FR 35.23 }
wounded: Heb. made sick}
\VS{24}His
servants therefore
took him out of that
chariot, and
put him in the
second
chariot that he had; and they
brought him to
Jerusalem, and he
died, and was
buried in
{\ADD{one of}} the
sepulchres of his
fathers. And all
Judah and
Jerusalem
mourned for
Josiah.\FTNT{*}{{\FR 35.24 }
in one…: or, among the sepulchres}
\par }{\PP \VS{25}¶ And
Jeremiah
lamented for
Josiah: and all the singing
men and the singing
women
spake of
Josiah in their
lamentations to this
day, and
made them an
ordinance in
Israel: and, behold, they
{\ADD{are}}
written in the
lamentations.
\VS{26}Now the
rest of the
acts of
Josiah, and his
goodness, according to
{\ADD{that which was}}
written in the
law of the
{\ND{
LORD}},\FTNT{†}{{\FR 35.26 }
goodness: Heb. kindnesses}
\VS{27}And his
deeds,
first and
last, behold, they
{\ADD{are}}
written in the
book of the
kings of
Israel and
Judah.

\par }\Chap{36}{\PP \VerseOne{1}Then the
people of the
land
took
Jehoahaz the
son of
Josiah, and made him
king in his
father’s stead in
Jerusalem.
\VS{2}Jehoahaz
{\ADD{was}}
twenty and
three
years
old when he began to
reign, and he
reigned
three
months in
Jerusalem.
\VS{3}And the
king of
Egypt put him
down at
Jerusalem, and
condemned the
land in an
hundred
talents of
silver and a
talent of
gold.\FTNT{*}{{\FR 36.3 }
put…: Heb. removed him}
\VS{4}And the
king of
Egypt made
Eliakim his
brother
king over
Judah and
Jerusalem, and
turned his
name to
Jehoiakim. And
Necho
took
Jehoahaz his
brother, and
carried him to
Egypt.
\par }{\PP \VS{5}¶
Jehoiakim
{\ADD{was}}
twenty and
five
years
old when he began to
reign, and he
reigned
eleven
years in
Jerusalem: and he
did
{\ADD{that which was}}
evil in the
sight of the
{\ND{
LORD}} his
God.
\VS{6}Against him came
up
Nebuchadnezzar
king of
Babylon, and
bound him in
fetters, to
carry him to
Babylon.\FTNT{†}{{\FR 36.6 }
fetters: or, chains}
\VS{7}Nebuchadnezzar also
carried of the
vessels of the
house of the
{\ND{
LORD}} to
Babylon, and
put them in his
temple at
Babylon.
\VS{8}Now the
rest of the
acts of
Jehoiakim, and his
abominations which he
did, and that which was
found in him, behold, they
{\ADD{are}}
written in the
book of the
kings of
Israel and
Judah: and
Jehoiachin his
son
reigned in his stead.\FTNT{‡}{{\FR 36.8 }
Jehoiachin: also called, Jeconiah, or, Coniah}
\par }{\PP \VS{9}¶
Jehoiachin
{\ADD{was}}
eight
years
old when he began to
reign, and he
reigned
three
months and
ten
days in
Jerusalem: and he
did
{\ADD{that which was}}
evil in the
sight of the
{\ND{
LORD}}.
\VS{10}And when the
year was
expired,
king
Nebuchadnezzar
sent, and
brought him to
Babylon, with the
goodly
vessels of the
house of the
{\ND{
LORD}}, and made
Zedekiah his
brother
king over
Judah and
Jerusalem.\FTNT{§}{{\FR 36.10 }
when…: Heb. at the return of the year}\FTNT{**}{{\FR 36.10 }
goodly…: Heb. vessels of desire}\FTNT{††}{{\FR 36.10 }
Zedekiah…: or, Mattaniah, his father’s brother}
\par }{\PP \VS{11}¶
Zedekiah
{\ADD{was}}
one and
twenty
years
old when he began to
reign, and
reigned
eleven
years in
Jerusalem.
\VS{12}And he
did
{\ADD{that which was}}
evil in the
sight of the
{\ND{
LORD}} his
God,
{\ADD{and}}
humbled not himself
before
Jeremiah the
prophet
{\ADD{speaking}} from the
mouth of the
{\ND{
LORD}}.
\VS{13}And he also
rebelled against
king
Nebuchadnezzar, who had made him
swear by
God: but he
stiffened his
neck, and
hardened his
heart from
turning unto the
{\ND{
LORD}}
God of
Israel.
\par }{\PP \VS{14}¶ Moreover all the
chief of the
priests, and the
people,
transgressed
very
much after all the
abominations of the
heathen; and
polluted the
house of the
{\ND{
LORD}} which he had
hallowed in
Jerusalem.
\VS{15}And the
{\ND{
LORD}}
God of their
fathers
sent to them
by his
messengers, rising up
betimes, and
sending; because he had
compassion on his
people, and on his dwelling
place:\FTNT{‡‡}{{\FR 36.15 }
by: Heb. by the hand of}\FTNT{§§}{{\FR 36.15 }
betimes: that is, continually and carefully}
\VS{16}But they
mocked the
messengers of
God, and
despised his
words, and
misused his
prophets, until the
wrath of the
{\ND{
LORD}}
arose against his
people, till
{\ADD{there was}} no
remedy.\FTNT{***}{{\FR 36.16 }
remedy: Heb. healing}
\VS{17}Therefore he
brought upon them the
king of the
Chaldees, who
slew their young
men with the
sword in the
house of their
sanctuary, and had no
compassion upon young
man or
maiden, old
man, or him that stooped for
age: he
gave
{\ADD{them}} all into his
hand.
\VS{18}And all the
vessels of the
house of
God,
great and
small, and the
treasures of the
house of the
{\ND{
LORD}}, and the
treasures of the
king, and of his
princes; all
{\ADD{these}} he
brought to
Babylon.
\VS{19}And they
burnt the
house of
God, and brake
down the
wall of
Jerusalem, and
burnt all the
palaces thereof with
fire, and
destroyed all the
goodly
vessels thereof.
\VS{20}And them that had
escaped from the
sword carried he
away to
Babylon; where they were
servants to him and his
sons until the
reign of the
kingdom of
Persia:\FTNT{†††}{{\FR 36.20 }
them…: Heb. the remainder from the sword}
\VS{21}To
fulfil the
word of the
{\ND{
LORD}} by the
mouth of
Jeremiah, until the
land had
enjoyed her
sabbaths:
{\ADD{for}} as long
as she lay
desolate she kept
sabbath, to
fulfil threescore and
ten
years.
\par }{\PP \VS{22}¶ Now in the
first
year of
Cyrus
king of
Persia, that the
word of the
{\ND{
LORD}}
{\ADD{spoken}} by the
mouth of
Jeremiah might be
accomplished, the
{\ND{
LORD}} stirred
up the
spirit of
Cyrus
king of
Persia, that he made a
proclamation throughout all his
kingdom, and
{\ADD{put it}} also in
writing,
saying,
\VS{23}Thus
saith
Cyrus
king of
Persia, All the
kingdoms of the
earth hath the
{\ND{
LORD}}
God of
heaven
given me; and he hath
charged me to
build him an
house in
Jerusalem, which
{\ADD{is}} in
Judah. Who
{\ADD{is there}} among you of all his
people? The
{\ND{
LORD}} his
God
{\ADD{be}} with him, and let him go
up.
\par }