\NormalFont\ShortTitle{2 Corinthians}
{\MT THE SECOND EPISTLE OF PAUL THE APOSTLE TO THE CORINTHIANS

\par }\ChapOne{1}{\PP \VerseOne{1}Paul, an
apostle of
Jesus
Christ
by the
will of
God,
and
Timothy
{\ADD{our}}
brother, unto the
church of
God which
is
at
Corinth,
with
all the
saints which
are
in
all
Achaia:
\VS{2}Grace
{\ADD{be}} to
you
and
peace
from
God
our
Father,
and
{\ADD{from}} the
Lord
Jesus
Christ.
\VS{3}Blessed
{\ADD{be}}
God,
even the
Father of
our
Lord
Jesus
Christ, the
Father of
mercies,
and the
God of
all
comfort;
\VS{4}Who
comforteth
us
in
all
our
tribulation, that
we may be
able
to
comfort
them which
are
in
any
trouble,
by the comfort
wherewith we
ourselves are
comforted
of
God.
\VS{5}For
as the
sufferings of
Christ
abound
in
us,
so
our
consolation
also
aboundeth
by
Christ.
\VS{6}And
whether we be
afflicted,
{\ADD{it is}}
for
your
consolation
and
salvation,
which is
effectual
in the
enduring of the
same
sufferings
which
we
also
suffer: or
whether we be
comforted,
{\ADD{it is}}
for
your
consolation
and
salvation.
\VS{7}And our hope of you
{\ADD{is}} stedfast,
knowing,
that
as ye
are
partakers of the
sufferings,
so
{\ADD{shall ye be}}
also of the
consolation.
\VS{8}For we
would
not,
brethren,
have
you
ignorant
of
our
trouble which
came to
us
in
Asia,
that we were
pressed
out of
measure,
above
strength, insomuch
that
we
despaired
even of
life:
\VS{9}But
we
had the
sentence of
death
in
ourselves,
that we
should
not
trust
in
ourselves,
but
in
God
which
raiseth the
dead:
\VS{10}Who
delivered
us
from so
great a
death,
and doth
deliver:
in
whom we
trust
that he
will
yet
deliver
{\ADD{us}};
\VS{11}Ye
also helping
together by
prayer
for
us,
that for the
gift
{\ADD{bestowed}}
upon
us by the means
of
many
persons thanks may be
given
by
many
on our
behalf.
\par }{\PP \VS{12}For
our
rejoicing
is
this, the
testimony of
our
conscience,
that
in
simplicity
and
godly
sincerity,
not
with
fleshly
wisdom,
but
by the
grace of
God, we have had our
conversation
in the
world,
and more
abundantly
to
you-ward.
\VS{13}For we
write
none other
things unto
you,
than
what ye
read
or
acknowledge;
and I
trust ye shall
acknowledge
even
to the
end;
\VS{14}As
also ye have
acknowledged
us
in
part,
that we
are
your
rejoicing, even
as
ye
also
{\ADD{are}}
ours
in the
day of the
Lord
Jesus.
\par }{\PP \VS{15}And in
this
confidence I was
minded to
come
unto
you
before,
that ye might
have a
second
benefit;
\VS{16}And to
pass
by
you
into
Macedonia,
and to
come
again out
of
Macedonia
unto
you,
and
of
you to be brought on my
way
toward
Judaea.
\VS{17}When I
therefore
was
thus
minded,
did I
use
lightness?
or the
things that I
purpose, do I
purpose according
to the
flesh,
that
with
me there should
be
yea
yea,
and
nay
nay?
\VS{18}But
{\ADD{as}}
God
{\ADD{is}}
true,
our
word
toward
you
was
not
yea
and
nay.
\VS{19}For the
Son of
God,
Jesus
Christ,
who was
preached
among
you
by
us,
{\ADD{even}}
by
me
and
Silvanus
and
Timotheus,
was
not
yea
and
nay,
but
in
him
was
yea.
\VS{20}For
all the
promises of
God
in
him
{\ADD{are}}
yea,
and
in
him
Amen,
unto the
glory of
God
by
us.
\VS{21}Now he
which
stablisheth
us
with
you
in
Christ,
and hath
anointed
us,
{\ADD{is}}
God;
\VS{22}Who
hath
also
sealed
us,
and
given the
earnest of the
Spirit
in
our
hearts.
\VS{23}Moreover
I
call
God for a
record
upon
my
soul,
that to
spare
you I
came not as
yet
unto
Corinth.
\VS{24}Not for
that we have
dominion over
your
faith,
but
are
helpers of
your
joy:
for by
faith ye
stand.

\par }\Chap{2}{\PP \VerseOne{1}But I
determined
this with
myself, that I
would
not
come
again
to
you
in
heaviness.
\VS{2}For
if
I
make
you
sorry,
who
is
he
then that
maketh
me
glad,
but the same
which is made
sorry
by
me?
\VS{3}And I
wrote
this
same unto
you,
lest, when I
came, I should
have
sorrow
from them of
whom
I
ought to
rejoice; having
confidence
in
you
all,
that
my
joy
is
{\ADD{the joy}} of
you
all.
\VS{4}For out
of
much
affliction
and
anguish of
heart I
wrote unto
you
with
many
tears;
not
that ye should be
grieved,
but
that ye might
know the
love
which I
have more
abundantly
unto
you.
\VS{5}But
if
any have caused
grief, he
hath
not
grieved
me,
but
in
part:
that I
may
not
overcharge
you
all.
\VS{6}Sufficient to such a
man
{\ADD{is}}
this
punishment,
which
{\ADD{was inflicted}}
of
many.
\VS{7}So
that
contrariwise
ye
{\ADD{ought}}
rather to
forgive
{\ADD{him}},
and
comfort
{\ADD{him}}, lest
perhaps such a
one should be swallowed
up with
overmuch
sorrow.
\VS{8}Wherefore I
beseech
you that ye would
confirm
{\ADD{your}}
love
toward
him.
\VS{9}For
to this
end
also did I
write,
that I might
know the
proof of
you,
whether ye
be
obedient
in all
things.
\VS{10}To
whom ye
forgive any
thing,
I
{\ADD{forgive}}
also:
for
if
I
forgave any
thing, to
whom I
forgave
{\ADD{it}},
for your
sakes
{\ADD{forgave I it}}
in the
person of
Christ;
\VS{11}Lest
Satan should get an
advantage
of us:
for we
are
not
ignorant of
his
devices.
\VS{12}Furthermore, when I
came
to
Troas
to
{\ADD{preach}}
Christ’s
gospel,
and a
door was
opened unto
me
of the
Lord,
\VS{13}I
had
no
rest
in
my
spirit, because
I
found
not
Titus
my
brother:
but taking my
leave of
them, I went from
thence
into
Macedonia.
\VS{14}Now
thanks
{\ADD{be}} unto
God, which
always
causeth
us to triumph
in
Christ,
and maketh
manifest the
savour of
his
knowledge
by
us
in
every
place.
\VS{15}For we
are unto
God a sweet
savour of
Christ,
in them
that are
saved,
and
in them
that
perish:
\VS{16}To the
one
{\ADD{we are}} the
savour of
death
unto
death;
and to the
other the
savour of
life
unto
life.
And
who
{\ADD{is}}
sufficient
for these
things?
\VS{17}For we
are
not
as
many, which
corrupt the
word of
God:
but
as
of
sincerity,
but
as
of
God, in the
sight of
God speak
we
in
Christ.

\par }\Chap{3}{\PP \VerseOne{1}Do we
begin
again to
commend
ourselves?
or need
we,
as
some
{\ADD{others}},
epistles of
commendation
to
you,
or
{\ADD{letters}}
of
commendation from
you?
\VS{2}Ye
are
our
epistle
written
in
our
hearts,
known
and
read
of
all
men:
\VS{3}{\ADD{Forasmuch as ye are}}
manifestly
declared to
be the
epistle of
Christ
ministered
by
us,
written
not with
ink,
but with the
Spirit of the
living
God;
not
in
tables of
stone,
but
in
fleshy
tables of the
heart.
\par }{\PP \VS{4}And
such
trust have
we
through
Christ
to
God-ward:
\VS{5}Not
that we
are
sufficient
of
ourselves to
think any
thing
as
of
ourselves;
but
our
sufficiency
{\ADD{is}}
of
God;
\VS{6}Who
also hath
made
us able
ministers of the
new
testament;
not of the
letter,
but of the
spirit:
for the
letter
killeth,
but the
spirit giveth
life.
\VS{7}But
if the
ministration of
death,
written
{\ADD{and}}
engraven
in
stones,
was
glorious, so
that the
children of
Israel
could
not stedfastly
behold the
face of
Moses
for the
glory of
his
countenance;
which
{\ADD{glory}} was to be done
away:
\VS{8}How
shall
not the
ministration of the
spirit
be
rather
glorious?
\VS{9}For
if the
ministration of
condemnation
{\ADD{be}}
glory,
much
more
doth the
ministration of
righteousness
exceed
in
glory.
\VS{10}For
even that
which was made
glorious
had
no
glory
in
this
respect, by reason
of the
glory
that
excelleth.
\VS{11}For
if that
which is done
away
{\ADD{was}}
glorious,
much
more that
which
remaineth
{\ADD{is}}
glorious.
\VS{12}Seeing
then that we
have
such
hope, we
use
great plainness of
speech:
\VS{13}And
not
as
Moses,
{\ADD{which}}
put a
vail
over
his
face,
that the
children of
Israel
could
not stedfastly
look
to the
end of that
which is
abolished:
\VS{14}But
their
minds were
blinded:
for
until this
day
remaineth the
same
vail untaken
away
in the
reading of the
old
testament;
which
{\ADD{vail}} is done
away
in
Christ.
\VS{15}But even
unto this
day,
when
Moses is
read, the
vail
is
upon
their
heart.
\VS{16}Nevertheless
when it shall
turn
to the
Lord, the
vail shall be taken
away.
\VS{17}Now the
Lord
is that
Spirit:
and
where the
Spirit of the
Lord
{\ADD{is}},
there
{\ADD{is}}
liberty.
\VS{18}But
we
all,
with
open
face beholding as in a
glass the
glory of the
Lord, are
changed into the
same
image
from
glory
to
glory,
{\ADD{even}}
as by the
Spirit
of the
Lord.

\par }\Chap{4}{\PP \VerseOne{1}Therefore
seeing we
have
this
ministry,
as we have received
mercy, we
faint
not;
\VS{2}But have
renounced the hidden
things of
dishonesty,
not
walking
in
craftiness,
nor
handling the
word of
God
deceitfully;
but by
manifestation of the
truth
commending
ourselves
to
every
man’s
conscience in the sight
of
God.
\VS{3}But
if
our
gospel be
hid, it
is
hid
to
them that
are
lost:
\VS{4}In
whom the
god of
this
world hath
blinded the
minds of them which believe
not,
lest the
light of the
glorious
gospel of
Christ,
who
is the
image of
God, should
shine unto
them.
\VS{5}For we
preach
not
ourselves,
but
Christ
Jesus the
Lord;
and
ourselves
your
servants
for
Jesus’ sake.
\VS{6}For
God,
who
commanded the
light to
shine out
of
darkness, hath
shined
in
our
hearts,
to
{\ADD{give}} the
light of the
knowledge of the
glory of
God
in the
face of
Jesus
Christ.
\par }{\PP \VS{7}But we
have
this
treasure
in
earthen
vessels,
that the
excellency of the
power may
be of
God,
and
not
of
us.
\VS{8}{\ADD{We are}}
troubled
on every
side,
yet
not
distressed;
{\ADD{we are}}
perplexed,
but
not in
despair;
\VS{9}Persecuted,
but
not
forsaken; cast
down,
but
not
destroyed;
\VS{10}Always bearing
about
in the
body the
dying of the
Lord
Jesus,
that the
life
also of
Jesus might be made
manifest
in
our
body.
\VS{11}For
we which
live are
alway
delivered
unto
death
for
Jesus’ sake,
that the
life
also of
Jesus might be made
manifest
in
our
mortal
flesh.
\VS{12}So
then
death
worketh
in
us,
but
life
in
you.
\VS{13}We
having the
same
spirit of
faith, according
as it is
written, I
believed, and
therefore have I
spoken;
we
also
believe,
and
therefore
speak;
\VS{14}Knowing
that he
which raised
up the
Lord
Jesus shall raise
up
us
also
by
Jesus,
and shall
present
{\ADD{us}}
with
you.
\VS{15}For all
things
{\ADD{are}}
for your
sakes,
that the
abundant
grace
might
through the
thanksgiving of
many
redound
to the
glory of
God.
\VS{16}For which
cause we
faint
not;
but
though
our
outward
man
perish,
yet the
inward
{\ADD{man}} is
renewed
day
by
day.
\VS{17}For
our
light
affliction, which is but for a
moment,
worketh for
us a
far more
exceeding
{\ADD{and}}
eternal
weight of
glory;
\VS{18}While
we
look
not at the
things which are
seen,
but at the
things which
are
not
seen:
for the
things which are
seen
{\ADD{are}}
temporal;
but the
things which
are
not
seen
{\ADD{are}}
eternal.

\par }\Chap{5}{\PP \VerseOne{1}For we
know
that
if
our
earthly
house
of
{\ADD{this}}
tabernacle were
dissolved, we
have a
building of
God, an
house not made with
hands,
eternal
in the
heavens.
\VS{2}For
in
this we
groan, earnestly
desiring to be clothed
upon with
our
house which is
from
heaven:
\VS{3}If so be
that being
clothed we
shall
not be
found
naked.
\VS{4}For we that
are
in
{\ADD{this}}
tabernacle do
groan, being
burdened:
not
for
that we
would be
unclothed,
but
clothed upon,
that
mortality might be swallowed
up of
life.
\VS{5}Now he
that hath
wrought
us
for the
selfsame
thing
{\ADD{is}}
God,
who
also hath
given unto
us the
earnest of the
Spirit.
\VS{6}Therefore
{\ADD{we are}}
always
confident,
knowing
that,
whilst we are at
home
in the
body, we are
absent
from the
Lord:
\VS{7}(For we
walk
by
faith,
not
by
sight:)
\VS{8}We are
confident,
{\ADD{I say}},
and
willing
rather to be
absent
from the
body,
and to be
present
with the
Lord.
\VS{9}Wherefore we
labour,
that,
whether
present
or
absent, we may
be
accepted of
him.
\VS{10}For
we
must
all
appear
before the judgment
seat of
Christ;
that every
one may
receive the
things
{\ADD{done}}
in
{\ADD{his}}
body,
according to
that he hath
done,
whether
{\ADD{it be}}
good
or
bad.
\par }{\PP \VS{11}Knowing
therefore the
terror of the
Lord, we
persuade
men;
but we are made
manifest unto
God;
and I
trust
also are made
manifest
in
your
consciences.
\VS{12}For we
commend
not
ourselves
again unto
you,
but
give
you
occasion to
glory on
our
behalf,
that ye may
have somewhat
to
{\ADD{answer}}
them which
glory
in
appearance,
and
not in
heart.
\VS{13}For
whether we be beside
ourselves,
{\ADD{it is}} to
God: or
whether we be
sober,
{\ADD{it is}} for your
cause.
\VS{14}For the
love of
Christ
constraineth
us;
because we thus
judge,
that
if
one
died
for
all,
then
were
all
dead:
\VS{15}And
{\ADD{that}} he
died
for
all,
that
they which
live
should not
henceforth
live unto
themselves,
but unto
him which
died
for
them,
and rose
again.
\VS{16}Wherefore
henceforth
know
we no
man
after the
flesh:
yea,
though we have
known
Christ
after the
flesh,
yet now
henceforth know
we
{\ADD{him}} no
more.
\VS{17}Therefore
if any
man
{\ADD{be}}
in
Christ,
{\ADD{he is}} a
new
creature: old
things are passed
away;
behold, all
things are
become
new.
\VS{18}And all
things
{\ADD{are}}
of
God,
who hath
reconciled
us to
himself
by
Jesus
Christ,
and hath
given to
us the
ministry of
reconciliation;
\VS{19}To
wit,
that
God
was
in Christ,
reconciling the
world unto
himself,
not
imputing
their
trespasses unto
them;
and hath
committed
unto
us the
word of
reconciliation.
\VS{20}Now
then we are
ambassadors
for
Christ, as
though
God did
beseech
{\ADD{you}}
by
us: we
pray
{\ADD{you}}
in
Christ’s stead, be ye
reconciled to
God.
\VS{21}For he hath made
him
{\ADD{to be}}
sin
for
us, who
knew
no
sin;
that
we might be
made the
righteousness of
God
in
him.

\par }\Chap{6}{\PP \VerseOne{1}We
then,
{\ADD{as}} workers
together
{\ADD{with him}},
beseech
{\ADD{you}}
also that
ye
receive
not the
grace of
God
in
vain.
\VS{2}(For he
saith, I have
heard
thee in a
time
accepted,
and
in the
day of
salvation have I
succoured
thee:
behold,
now
{\ADD{is}} the
accepted
time;
behold,
now
{\ADD{is}} the
day of
salvation.)
\VS{3}Giving
no
offence
in any
thing,
that the
ministry
be
not
blamed:
\VS{4}But
in
all
{\ADD{things}}
approving
ourselves
as the
ministers of
God,
in
much
patience,
in
afflictions,
in
necessities,
in
distresses,
\VS{5}In
stripes,
in
imprisonments,
in
tumults,
in
labours,
in
watchings,
in
fastings;
\VS{6}By
pureness,
by
knowledge,
by
longsuffering,
by
kindness,
by the
Holy
Ghost,
by
love
unfeigned,
\VS{7}By the
word of
truth,
by the
power of
God,
by the
armour of
righteousness on the right
hand
and on the
left,
\VS{8}By
honour
and
dishonour,
by evil
report
and good
report:
as
deceivers,
and
{\ADD{yet}}
true;
\VS{9}As
unknown,
and
{\ADD{yet}} well
known;
as
dying,
and,
behold, we
live;
as
chastened,
and
not
killed;
\VS{10}As
sorrowful,
yet
alway
rejoicing;
as
poor,
yet
making
many
rich;
as
having
nothing,
and
{\ADD{yet}}
possessing all
things.
\par }{\PP \VS{11}O
{\ADD{ye}}
Corinthians,
our
mouth is
open
unto
you,
our
heart is
enlarged.
\VS{12}Ye
are
not
straitened
in
us,
but ye are
straitened
in your
own
bowels.
\VS{13}Now for a
recompence in the
same, (I
speak as
unto
{\ADD{my}}
children,)
be
ye
also
enlarged.
\VS{14}Be
ye
not unequally yoked
together with
unbelievers:
for
what
fellowship hath
righteousness
with
unrighteousness?
and
what
communion hath
light
with
darkness?
\VS{15}And
what
concord hath
Christ
with
Belial?
or
what
part hath he that
believeth
with an
infidel?
\VS{16}And
what
agreement hath the
temple of
God
with
idols?
for
ye
are the
temple of the
living
God;
as
God hath
said, I will
dwell
in
them,
and walk
in
{\ADD{them}};
and I will
be
their
God,
and
they shall
be
my
people.
\VS{17}Wherefore come
out
from
among
them,
and be ye
separate,
saith the
Lord,
and
touch
not the
unclean
{\ADD{thing}}; and
I will
receive
you,
\VS{18}And will
be a
Father
unto
you,
and
ye shall
be
my
sons
and
daughters,
saith the
Lord
Almighty.

\par }\Chap{7}{\PP \VerseOne{1}Having
therefore
these
promises, dearly
beloved, let us
cleanse
ourselves
from
all
filthiness of the
flesh
and
spirit,
perfecting
holiness
in the
fear of
God.
\par }{\PP \VS{2}Receive
us; we have
wronged no
man, we have
corrupted no
man, we have
defrauded no
man.
\VS{3}I
speak
not
{\ADD{this}}
to
condemn
{\ADD{you}}:
for I have said
before,
that ye
are
in
our
hearts
to
die
and live
with
{\ADD{you}}.
\VS{4}Great
{\ADD{is}}
my boldness of
speech
toward
you,
great
{\ADD{is}}
my
glorying
of
you: I am
filled with
comfort, I am
exceeding
joyful
in
all
our
tribulation.
\VS{5}For,
when we were
come
into
Macedonia,
our
flesh
had
no
rest,
but
we were
troubled
on every
side;
without
{\ADD{were}}
fightings,
within
{\ADD{were}}
fears.
\VS{6}Nevertheless
God, that
comforteth
those that are cast
down,
comforted
us
by the
coming of
Titus;
\VS{7}And
not
by
his
coming
only,
but
by the
consolation
wherewith he was
comforted
in
you, when he
told
us
your earnest
desire,
your
mourning,
your fervent
mind
toward
me; so
that
I
rejoiced the
more.
\VS{8}For though I
made
you sorry
with a
letter, I
do
not
repent,
though I did
repent:
for I
perceive
that the
same
epistle hath
made
you
sorry,
though
{\ADD{it were}}
but
for a
season.
\VS{9}Now I
rejoice,
not
that ye were made
sorry,
but
that ye
sorrowed
to
repentance:
for ye were made
sorry
after a godly
manner,
that ye might receive
damage
by
us
in
nothing.
\VS{10}For
godly
sorrow
worketh
repentance
to
salvation not to be repented
of:
but the
sorrow of the
world
worketh
death.
\VS{11}For
behold
this selfsame
thing,
that
ye
sorrowed
after a godly
sort,
what
carefulness it
wrought in
you,
yea,
{\ADD{what}} clearing of
yourselves,
yea,
{\ADD{what}}
indignation,
yea,
{\ADD{what}}
fear,
yea,
{\ADD{what}} vehement
desire,
yea,
{\ADD{what}}
zeal,
yea,
{\ADD{what}}
revenge!
In
all
{\ADD{things}} ye have
approved
yourselves to
be
clear
in this
matter.
\VS{12}Wherefore,
though I
wrote unto
you,
{\ADD{I did it}}
not for his
cause
that had done the
wrong,
nor for his
cause that suffered
wrong,
but that
our
care
for
you in the
sight of
God might
appear
unto
you.
\VS{13}Therefore we were
comforted
in
your
comfort: yea,
and
exceedingly the
more joyed
we
for the
joy of
Titus,
because
his
spirit was
refreshed
by
you
all.
\VS{14}For
if I have
boasted any
thing to
him
of
you, I
am
not
ashamed;
but
as we
spake all
things to
you
in
truth,
even
so
our
boasting,
which
{\ADD{I made}}
before
Titus, is
found a
truth.
\VS{15}And
his inward
affection
is more
abundant
toward
you, whilst he
remembereth the
obedience of
you
all,
how
with
fear
and
trembling ye
received
him.
\VS{16}I
rejoice
therefore
that I have
confidence
in
you
in
all
{\ADD{things.}}

\par }\Chap{8}{\PP \VerseOne{1}Moreover,
brethren, we
do
you to
wit of the
grace of
God
bestowed
on the
churches of
Macedonia;
\VS{2}How
that
in a
great
trial of
affliction the
abundance of
their
joy
and
their
deep
poverty
abounded
unto the
riches of
their
liberality.
\VS{3}For
to
{\ADD{their}}
power, I bear
record, yea,
and
beyond
{\ADD{their}}
power
{\ADD{they were}} willing of
themselves;
\VS{4}Praying
us
with
much
intreaty that
we would
receive the
gift,
and
{\ADD{take upon us}} the
fellowship of the
ministering
to the
saints.
\VS{5}And
{\ADD{this they did}},
not
as we
hoped,
but
first
gave their own
selves to the
Lord,
and unto
us
by the
will of
God.
\VS{6}Insomuch
that
we
desired
Titus,
that
as he had
begun,
so he would
also
finish
in
you the
same
grace
also.
\VS{7}Therefore,
as ye
abound
in
every
{\ADD{thing, in}}
faith,
and
utterance,
and
knowledge,
and
{\ADD{in}}
all
diligence,
and
{\ADD{in}}
your
love
to
us,
{\ADD{see}}
that ye
abound
in
this
grace
also.
\VS{8}I
speak
not
by
commandment,
but by
occasion of the
forwardness of
others,
and to
prove the
sincerity of
your
love.
\VS{9}For ye
know the
grace of
our
Lord
Jesus
Christ,
that, though he
was
rich, yet
for your
sakes he became
poor,
that
ye through
his
poverty might be
rich.
\VS{10}And
herein I
give
{\ADD{my}}
advice:
for
this is
expedient for
you,
who have begun
before,
not
only to
do,
but
also to be
forward a year
ago.
\VS{11}Now
therefore
perform the
doing
{\ADD{of it}};
that
as
{\ADD{there was}} a
readiness to
will,
so
{\ADD{there may be}} a
performance
also out
of that
which ye
have.
\VS{12}For
if there be
first a willing
mind,
{\ADD{it is}}
accepted according to
that a
man
hath,
{\ADD{and}}
not according
to that he
hath
not.
\VS{13}For
{\ADD{I mean}}
not
that other
men be
eased,
and
ye
burdened:
\VS{14}But by an equality,
{\ADD{that}} now at this time your abundance
{\ADD{may be a supply}} for their want,
that
their
abundance
also may
be
{\ADD{a supply}}
for
your
want:
that there may
be
equality:
\VS{15}As it is
written, He
that
{\ADD{had gathered}}
much
had
nothing
over;
and he
that
{\ADD{had gathered}}
little
had
no
lack.
\VS{16}But
thanks
{\ADD{be}} to
God,
which
put the
same earnest
care
into the
heart of
Titus
for
you.
\VS{17}For
indeed he
accepted the
exhortation;
but
being more
forward, of his own
accord he
went
unto
you.
\VS{18}And we have
sent
with
him the
brother,
whose
praise
{\ADD{is}}
in the
gospel
throughout
all the
churches;
\VS{19}And
not
{\ADD{that}}
only,
but who
was
also
chosen
of the
churches to travel
with
us
with
this
grace,
which is
administered
by
us
to the
glory of the
same
Lord,
and
{\ADD{declaration of}}
your ready
mind:
\VS{20}Avoiding
this, that
no
man should
blame
us
in
this
abundance which is
administered
by
us:
\VS{21}Providing
for honest
things,
not
only in the sight
of the
Lord,
but
also in the sight
of
men.
\VS{22}And we have sent
with
them
our
brother,
whom we
have
oftentimes
proved
diligent
in many
things,
but
now
much more
diligent, upon the
great
confidence
which
{\ADD{I have}}
in
you.
\VS{23}Whether
{\ADD{any do enquire}}
of
Titus,
{\ADD{he is}}
my
partner
and
fellowhelper
concerning
you:
or
our
brethren
{\ADD{be enquired of, they are}} the
messengers of the
churches,
{\ADD{and}} the
glory of
Christ.
\VS{24}Wherefore shew
ye
to
them,
and
before the
churches, the
proof of
your
love,
and of
our
boasting
on your
behalf.

\par }\Chap{9}{\PP \VerseOne{1}For as
touching the
ministering
to the
saints, it
is
superfluous for
me to
write to
you:
\VS{2}For I
know the
forwardness of your
mind, for
which I
boast
of
you to them of
Macedonia,
that
Achaia was
ready a year
ago;
and
your
zeal hath
provoked very
many.
\VS{3}Yet have I
sent the
brethren,
lest
our
boasting
of
you should be in
vain
in
this
behalf;
that,
as I
said, ye
may
be
ready:
\VS{4}Lest
haply
if they of
Macedonia
come
with
me,
and
find
you
unprepared,
we (that we
say
not,
ye) should be
ashamed
in this
same
confident
boasting.
\VS{5}Therefore I thought
it
necessary to
exhort the
brethren,
that they would go
before
unto
you,
and make up
beforehand
your
bounty, whereof ye had notice
before, that the
same might
be
ready,
as
{\ADD{a matter of}}
bounty,
and
not
as
{\ADD{of}}
covetousness.
\VS{6}But
this
{\ADD{I say}},
He which
soweth
sparingly shall
reap
also
sparingly;
and
he which
soweth
bountifully shall
reap
also
bountifully.
\VS{7}Every
man according
as he
purposeth in his
heart,
{\ADD{so let him give}};
not
grudgingly,
or
of
necessity:
for
God
loveth a
cheerful
giver.
\VS{8}And
God
{\ADD{is}}
able to make
all
grace
abound
toward
you;
that ye,
always
having
all
sufficiency
in
all
{\ADD{things}}, may
abound
to
every
good
work:
\VS{9}(As it is
written, He hath dispersed
abroad; he hath
given to the
poor:
his
righteousness
remaineth
for
ever.
\VS{10}Now he
that
ministereth
seed to the
sower
both
minister
bread
for
{\ADD{your}}
food,
and
multiply
your seed
sown,
and
increase the
fruits of
your
righteousness;)
\VS{11}Being
enriched
in every
thing
to
all
bountifulness,
which
causeth
through
us
thanksgiving to
God.
\VS{12}For the
administration of
this
service
not
only
supplieth the
want of the
saints,
but
is
abundant
also
by
many
thanksgivings unto
God;
\VS{13}Whiles by the
experiment of
this
ministration they
glorify
God
for
your
professed
subjection
unto the
gospel of
Christ,
and for
{\ADD{your}}
liberal
distribution
unto
them,
and
unto
all
{\ADD{men}};
\VS{14}And by
their
prayer
for
you, which long
after
you
for the
exceeding
grace of
God
in
you.
\VS{15}Thanks
{\ADD{be}}
unto
God
for
his
unspeakable
gift.

\par }\Chap{10}{\PP \VerseOne{1}Now
I
Paul
myself
beseech
you
by the
meekness
and
gentleness of
Christ,
who
in
presence
{\ADD{am}}
base
among
you,
but being
absent am
bold
toward
you:
\VS{2}But I
beseech
{\ADD{you}}, that I
may
not be
bold when I am
present with that
confidence, wherewith I
think to be
bold
against
some,
which
think of
us as
if we
walked according
to the
flesh.
\VS{3}For though we
walk
in the
flesh, we
do
not
war
after the
flesh:
\VS{4}(For the
weapons of
our
warfare
{\ADD{are}}
not
carnal,
but
mighty through
God
to the pulling
down of strong
holds;)
\VS{5}Casting
down
imaginations,
and
every high
thing that exalteth
itself
against the
knowledge of
God,
and bringing into
captivity
every
thought
to the
obedience of
Christ;
\VS{6}And
having
in a
readiness to
revenge
all
disobedience,
when
your
obedience is
fulfilled.
\VS{7}Do ye
look on
things
after the outward
appearance?
If any
man
trust to
himself that he
is
Christ’s, let
him
of
himself
think
this
again,
that,
as
he
{\ADD{is}}
Christ’s,
even
so
{\ADD{are}}
we
Christ’s.
\VS{8}For
though I should
boast
somewhat
more
of
our
authority,
which the
Lord hath
given
us
for
edification,
and
not
for
your
destruction, I
should
not be
ashamed:
\VS{9}That I
may
not
seem
as
if I would
terrify
you
by
letters.
\VS{10}For
{\ADD{his}}
letters, say
they,
{\ADD{are}}
weighty
and
powerful;
but
{\ADD{his}}
bodily
presence
{\ADD{is}}
weak,
and
{\ADD{his}}
speech
contemptible.
\VS{11}Let such an
one
think
this,
that, such
as we
are in
word
by
letters when we are
absent,
such
{\ADD{will we be}}
also in
deed when we are
present.
\VS{12}For we
dare
not make ourselves of the
number,
or
compare
ourselves with
some that
commend
themselves:
but
they
measuring
themselves
by
themselves,
and
comparing
themselves among
themselves,
are
not
wise.
\VS{13}But
we
will
not
boast
of
things
without
{\ADD{our}}
measure,
but according
to the
measure of the
rule
which
God hath
distributed to
us, a
measure to
reach
even
unto
you.
\VS{14}For we
stretch
not
ourselves
beyond
{\ADD{our measure}}, as
though we
reached
not
unto
you:
for we are
come as far
as to
you
also
in
{\ADD{preaching}} the
gospel of
Christ:
\VS{15}Not
boasting
of
things
without
{\ADD{our}}
measure,
{\ADD{that is}},
of other
men’s
labours;
but
having
hope, when
your
faith is
increased, that we shall be
enlarged
by
you
according
to
our
rule
abundantly,
\VS{16}To preach the
gospel
in
the
{\ADD{regions}}
beyond
you,
{\ADD{and}}
not to
boast
in another
man’s line of
things made
ready
to our
hand.
\VS{17}But
he that
glorieth, let him
glory
in the
Lord.
\VS{18}For
not he
that
commendeth
himself
is
approved,
but
whom the
Lord
commendeth.

\par }\Chap{11}{\PP \VerseOne{1}Would to
God ye could
bear with
me a
little
in
{\ADD{my}}
folly:
and
indeed
bear with
me.
\VS{2}For I am
jealous over
you with
godly
jealousy:
for I have
espoused
you to
one
husband, that I may
present
{\ADD{you as}} a
chaste
virgin to
Christ.
\VS{3}But I
fear, lest by any
means,
as the
serpent
beguiled
Eve
through
his
subtilty,
so
your
minds should be
corrupted
from the
simplicity that
is
in
Christ.
\VS{4}For
if
he that
cometh
preacheth
another
Jesus,
whom we
have
not
preached,
or
{\ADD{if}} ye
receive
another
spirit,
which ye have
not
received,
or
another
gospel,
which ye
have
not
accepted, ye
might
well bear
with
{\ADD{him}}.
\VS{5}For I
suppose I
was
not a whit
behind the
very
chiefest
apostles.
\VS{6}But
though
{\ADD{I be}}
rude in
speech,
yet
not in
knowledge;
but we have
been
throughly
made
manifest
among
you
in all
things.
\VS{7}Have I
committed an
offence in
abasing
myself
that
ye might be
exalted,
because I have
preached to
you the
gospel of
God
freely?
\VS{8}I
robbed
other
churches,
taking
wages
{\ADD{of them}}, to
do
you
service.
\VS{9}And when I was
present
with
you,
and
wanted, I was chargeable
to
no
man:
for that which was
lacking to
me the
brethren which
came
from
Macedonia
supplied:
and
in
all
{\ADD{things}} I have
kept
myself from being
burdensome unto
you,
and
{\ADD{so}} will I
keep
{\ADD{myself}}.
\VS{10}As the
truth of
Christ
is
in
me, no
man shall
stop
me of
this
boasting
in the
regions of
Achaia.
\VS{11}Wherefore?
because I
love
you
not?
God
knoweth.
\VS{12}But
what I
do, that I will
do,
that I may cut
off
occasion from
them which
desire
occasion;
that
wherein they
glory, they may be
found
even
as
we.
\VS{13}For
such
{\ADD{are}} false
apostles,
deceitful
workers, transforming
themselves
into the
apostles of
Christ.
\VS{14}And
no
marvel;
for
Satan
himself is
transformed
into an
angel of
light.
\VS{15}Therefore
{\ADD{it is}}
no great
thing
if
his
ministers
also be
transformed
as the
ministers of
righteousness;
whose
end shall
be according
to
their
works.
\par }{\PP \VS{16}I
say
again,
Let
no
man
think
me a
fool;
if
otherwise,
yet
as a
fool
receive
me,
that I may boast
myself a
little.
\VS{17}That
which I
speak, I
speak
{\ADD{it}}
not
after the
Lord,
but as it
were
foolishly,
in
this
confidence of
boasting.
\VS{18}Seeing
that
many
glory
after the
flesh,
I will
glory
also.
\VS{19}For ye
suffer
fools
gladly, seeing
ye
{\ADD{yourselves}} are
wise.
\VS{20}For ye
suffer,
if a
man
bring
you into
bondage,
if a
man
devour
{\ADD{you}},
if a
man
take
{\ADD{of you}},
if a
man exalt
himself,
if a
man
smite
you
on the
face.
\VS{21}I
speak as
concerning
reproach,
as
though
we had been
weak.
Howbeit
whereinsoever
any is
bold, (I
speak
foolishly,)
I am
bold
also.
\VS{22}Are
they
Hebrews?
so
{\ADD{am}}
I. Are
they
Israelites?
so
{\ADD{am}}
I. Are
they the
seed of
Abraham?
so
{\ADD{am}}
I.
\VS{23}Are
they
ministers of
Christ? (I
speak as a
fool)
I
{\ADD{am}}
more;
in
labours more
abundant,
in
stripes above
measure,
in
prisons more
frequent,
in
deaths
oft.
\VS{24}Of the
Jews five
times received
I
forty
{\ADD{stripes}}
save
one.
\VS{25}Thrice was I beaten with
rods,
once was I
stoned,
thrice I suffered
shipwreck, a night and a
day I have
been
in the
deep;
\VS{26}{\ADD{In}}
journeyings
often,
{\ADD{in}}
perils of
waters,
{\ADD{in}}
perils of
robbers,
{\ADD{in}}
perils
by
{\ADD{mine own}}
countrymen,
{\ADD{in}}
perils
by the
heathen,
{\ADD{in}}
perils
in the
city,
{\ADD{in}}
perils
in the
wilderness,
{\ADD{in}}
perils
in the
sea,
{\ADD{in}}
perils
among false
brethren;
\VS{27}In
weariness
and
painfulness,
in
watchings
often,
in
hunger
and
thirst,
in
fastings
often,
in
cold
and
nakedness.
\VS{28}Beside those things that are
without, that which
cometh upon
me
daily, the
care of
all the
churches.
\VS{29}Who is
weak,
and I
am
not
weak?
who is
offended,
and
I
burn
not?
\VS{30}If I must
needs
glory, I will
glory of the
things which concern
mine
infirmities.
\VS{31}The
God
and
Father of
our
Lord
Jesus
Christ, which
is
blessed
for
evermore,
knoweth
that I
lie
not.
\VS{32}In
Damascus the
governor under
Aretas the
king
kept the
city of the
Damascenes with a
garrison,
desirous to
apprehend
me:
\VS{33}And
through a
window
in a
basket was I let
down
by the
wall,
and
escaped
his
hands.

\par }\Chap{12}{\PP \VerseOne{1}It
is
not
expedient
for
me
doubtless to
glory. I will
come
to
visions
and
revelations of the
Lord.
\VS{2}I
knew a
man
in
Christ
above
fourteen years
ago, (whether
in the
body,
I
cannot
tell; or
whether out
of the
body,
I
cannot
tell:
God
knoweth;) such an
one caught
up
to the
third
heaven.
\VS{3}And I
knew
such a
man, (whether
in the
body,
or out
of the
body,
I
cannot
tell:
God
knoweth;)
\VS{4}How
that he was caught
up
into
paradise,
and
heard
unspeakable
words,
which it
is
not
lawful for a
man to
utter.
\VS{5}Of such an
one will I
glory:
yet
of
myself I
will
not
glory,
but
in
mine
infirmities.
\VS{6}For
though I would
desire to
glory, I
shall
not
be a
fool;
for I will
say the
truth:
but
{\ADD{now}} I
forbear,
lest any
man should
think
of
me
above that
which he
seeth
me
{\ADD{to be}},
or
{\ADD{that}} he
heareth
of
me.
\VS{7}And
lest I should be exalted above
measure through the
abundance of the
revelations, there was
given to
me a
thorn in the
flesh, the
messenger of
Satan
to
buffet
me,
lest I should be exalted above
measure.
\VS{8}For this
thing I
besought the
Lord
thrice,
that it might
depart
from
me.
\VS{9}And he
said unto
me,
{\WJ{My
grace
is sufficient
for thee:
for
my
strength
is made perfect
in
weakness.}} Most
gladly
therefore will
I
rather
glory
in
my
infirmities,
that the
power of
Christ may
rest
upon
me.
\VS{10}Therefore I take
pleasure
in
infirmities,
in
reproaches,
in
necessities,
in
persecutions,
in
distresses
for
Christ’s sake:
for
when I am
weak,
then am
I
strong.
\par }{\PP \VS{11}I am
become a
fool in
glorying;
ye have
compelled
me:
for
I
ought to have been
commended
of
you:
for in
nothing am I
behind the
very
chiefest
apostles,
though I
be
nothing.
\VS{12}Truly the
signs of an
apostle were
wrought
among
you
in
all
patience,
in
signs,
and
wonders,
and mighty
deeds.
\VS{13}For
what
is it
wherein ye were
inferior
to
other
churches,
except
{\ADD{it be}}
that
I
myself
was
not
burdensome to
you?
forgive
me
this
wrong.
\VS{14}Behold, the third
time I
am
ready to
come
to
you;
and I
will
not be
burdensome to
you:
for I
seek
not
yours,
but
you:
for the
children
ought
not to lay
up for the
parents,
but the
parents for the
children.
\VS{15}And
I will very
gladly
spend
and be
spent
for
you;
though the more
abundantly I
love
you, the
less I be
loved.
\VS{16}But be it
so,
I
did
not
burden
you:
nevertheless,
being
crafty, I
caught
you with
guile.
\VS{17}Did I make a
gain
of
you
by
any of
them
whom I
sent unto
you?
\VS{18}I
desired
Titus,
and
with
{\ADD{him}} I
sent a
brother.
Did
Titus make a
gain of
you? walked
we
not in the
same
spirit?
{\ADD{walked we}}
not in the
same
steps?
\VS{19}Again, think
ye
that we excuse
ourselves unto
you? we
speak
before
God
in
Christ:
but
{\ADD{we do}} all
things, dearly
beloved,
for
your
edifying.
\VS{20}For I
fear,
lest, when I
come, I
shall
not
find
you such
as I
would,
and
{\ADD{that}}
I shall be
found unto
you such
as ye
would
not:
lest
{\ADD{there be}}
debates,
envyings,
wraths,
strifes,
backbitings,
whisperings,
swellings,
tumults:
\VS{21}{\ADD{And}}
lest, when I
come
again,
my
God will
humble
me
among
you,
and
{\ADD{that}} I shall
bewail
many
which have sinned
already,
and
have
not
repented
of the
uncleanness
and
fornication
and
lasciviousness
which they have
committed.

\par }\Chap{13}{\PP \VerseOne{1}This
{\ADD{is}} the
third
{\ADD{time}} I am
coming
to
you.
In the
mouth of
two
or
three
witnesses
shall
every
word be
established.
\VS{2}I told you
before,
and foretell
you, as
if I were
present, the second
time;
and being
absent
now I
write to
them which heretofore have
sinned,
and to
all
other,
that,
if I
come
again, I
will
not
spare:
\VS{3}Since ye
seek a
proof of
Christ
speaking
in
me,
which
to
you-ward
is
not
weak,
but is
mighty
in
you.
\VS{4}For
though he was
crucified
through
weakness,
yet he
liveth
by the
power of
God.
For we
also are
weak
in
him,
but we shall
live
with
him
by the
power of
God
toward
you.
\VS{5}Examine
yourselves,
whether ye
be
in the
faith;
prove your own
selves. Know
ye
not your own
selves, how
that
Jesus
Christ
is
in
you,
except ye
be
reprobates?
\VS{6}But I
trust
that ye shall
know
that
we
are
not
reprobates.
\VS{7}Now I
pray
to
God
that
ye
do
no
evil;
not
that
we should
appear
approved,
but
that
ye should
do that
which is
honest,
though
we
be
as
reprobates.
\VS{8}For we can
do
nothing
against the
truth,
but
for the
truth.
\VS{9}For
we are
glad,
when we are
weak,
and
ye
are
strong:
and
this
also we
wish,
{\ADD{even}}
your
perfection.
\VS{10}Therefore I
write these
things
being
absent,
lest being
present I should
use
sharpness,
according to the
power
which the
Lord hath
given
me
to
edification,
and
not
to
destruction.
\par }{\PP \VS{11}Finally,
brethren,
farewell. Be
perfect, be of good
comfort,
be of
one
mind, live in
peace;
and the
God of
love
and
peace shall
be
with
you.
\VS{12}Greet one
another
with an
holy
kiss.
\par }{\PP \VS{13}All the
saints
salute
you.
\VS{14}The
grace of the
Lord
Jesus
Christ,
and the
love of
God,
and the
communion of the
Holy
Ghost,
{\ADD{be}}
with
you
all.
Amen.
\par }{\SH The second
{\ADD{epistle}} to the Corinthians was written from Philippi,
{\ADD{a city}} of Macedonia, by Titus and Lucas.
\par }