\NormalFont\ShortTitle{2 John}
{\MT THE SECOND EPISTLE OF JOHN

\par }\OneChap {\PP \VerseOne{1}The
elder unto the
elect
lady
and
her
children,
whom
I
love
in the
truth;
and
not
I
only,
but
also
all they that have
known the
truth;
\VS{2}For the
truth’s sake,
which
dwelleth
in
us,
and shall
be
with
us
for
ever.
\VS{3}Grace
be
with
you,
mercy,
{\ADD{and}}
peace,
from
God the
Father,
and
from the
Lord
Jesus
Christ, the
Son of the
Father,
in
truth
and
love.
\par }{\PP \VS{4}I
rejoiced
greatly
that I
found
of
thy
children
walking
in
truth,
as we have
received a
commandment
from the
Father.
\VS{5}And
now I
beseech
thee,
lady,
not
as though I
wrote a
new
commandment unto
thee,
but
that
which we
had
from the
beginning,
that we
love one
another.
\VS{6}And
this
is
love,
that we
walk
after
his
commandments.
This
is the
commandment,
That,
as ye have
heard
from the
beginning, ye should
walk
in
it.
\VS{7}For
many
deceivers are
entered
into the
world, who
confess
not that
Jesus
Christ is
come
in the
flesh.
This
is a
deceiver
and an
antichrist.
\VS{8}Look to
yourselves,
that we
lose
not those things
which we have
wrought,
but that we
receive
a
full
reward.
\VS{9}Whosoever
transgresseth,
and
abideth
not
in the
doctrine of
Christ,
hath
not
God. He that
abideth
in the
doctrine of
Christ,
he
hath
both the
Father
and the
Son.
\VS{10}If there
come
any
unto
you,
and
bring
not
this
doctrine,
receive
him
not
into
{\ADD{your}}
house,
neither
bid
him God
speed:
\VS{11}For he that
biddeth
him God
speed is
partaker
of
his
evil
deeds.
\par }{\PP \VS{12}Having
many things to
write unto
you, I
would
not
{\ADD{write}}
with
paper
and
ink:
but I
trust to
come
unto
you,
and
speak
face
to
face,
that
our
joy may
be
full.
\VS{13}The
children
of
thy
elect
sister
greet
thee.
Amen.
\par }