\NormalFont\ShortTitle{2 Samuel}
{\MT The Second Book of Samuel Otherwise Called the Second Book of the Kings

\par }\ChapOne{1}{\PP \VerseOne{1}Now it came to pass
after the
death of
Saul, when
David was
returned from the
slaughter of the
Amalekites, and
David had
abode
two
days in
Ziklag;
\VS{2}It came even to pass on the
third
day, that, behold, a
man
came out of the
camp from
Saul with his
clothes
rent, and
earth upon his
head: and
{\ADD{so}} it was, when he
came to
David, that he
fell to the
earth, and did
obeisance.
\VS{3}And
David
said unto him, From whence
comest thou? And he
said unto him, Out of the
camp of
Israel am I
escaped.
\VS{4}And
David
said unto him,
How went the
matter? I pray thee,
tell me. And he
answered, That the
people are
fled from the
battle, and
many of the
people also are
fallen and
dead; and
Saul and
Jonathan his
son are
dead also.\FTNT{*}{{\FR 1.4 }
How…: Heb. What was, etc}
\VS{5}And
David
said unto the young
man that
told him, How
knowest thou that
Saul and
Jonathan his
son be
dead?
\VS{6}And the young
man that
told him
said, As I
happened by
chance upon
mount
Gilboa, behold,
Saul
leaned upon his
spear; and, lo, the
chariots and
horsemen followed
hard after him.
\VS{7}And when he
looked
behind him, he
saw me, and
called unto me. And I
answered, Here
{\ADD{am}} I.\FTNT{†}{{\FR 1.7 }
Here…: Heb. Behold me}
\VS{8}And he
said unto me, Who
{\ADD{art}} thou? And I
answered him, I
{\ADD{am}} an
Amalekite.
\VS{9}He
said unto me again,
Stand, I pray thee, upon me, and
slay me: for
anguish is
come upon me, because my
life
{\ADD{is}} yet whole in me.\FTNT{‡}{{\FR 1.9 }
anguish…: or, my coat of mail (or, my embroidered coat) hindereth me, that my, etc}
\VS{10}So I
stood upon him, and
slew him, because I was
sure that he could not
live
after that he was
fallen: and I
took the
crown that
{\ADD{was}} upon his
head, and the
bracelet that
{\ADD{was}} on his
arm, and have
brought them hither unto my
lord.
\par }{\PP \VS{11}Then
David took
hold on his
clothes, and
rent them; and likewise all the
men that
{\ADD{were}} with him:
\VS{12}And they
mourned, and
wept, and
fasted until
even, for
Saul, and for
Jonathan his
son, and for the
people of the
{\ND{
LORD}}, and for the
house of
Israel; because they were
fallen by the
sword.
\par }{\PP \VS{13}¶ And
David
said unto the young
man that
told him, Whence
{\ADD{art}} thou? And he
answered, I
{\ADD{am}} the
son of a
stranger, an
Amalekite.
\VS{14}And
David
said unto him, How wast thou not
afraid to stretch
forth thine
hand to
destroy the
{\ND{
LORD’s}}
anointed?
\VS{15}And
David
called
one of the young
men, and
said, Go
near,
{\ADD{and}}
fall upon him. And he
smote him that he
died.
\VS{16}And
David
said unto him, Thy
blood
{\ADD{be}} upon thy
head; for thy
mouth hath
testified against thee,
saying, I have
slain the
{\ND{
LORD’s}}
anointed.
\par }{\PP \VS{17}¶ And
David
lamented with this
lamentation over
Saul and over
Jonathan his
son:
\VS{18}(Also he
bade them
teach the
children of
Judah
{\ADD{the use of}} the
bow: behold,
{\ADD{it is}}
written in the
book of
Jasher.)\FTNT{§}{{\FR 1.18 }
of Jasher: or, of the upright}
\VS{19}The
beauty of
Israel is
slain upon thy high
places: how are the
mighty
fallen!
\VS{20}Tell
{\ADD{it}} not in
Gath,
publish
{\ADD{it}} not in the
streets of
Askelon; lest the
daughters of the
Philistines
rejoice, lest the
daughters of the
uncircumcised
triumph.
\VS{21}Ye
mountains of
Gilboa,
{\ADD{let there be}} no
dew, neither
{\ADD{let there be}}
rain, upon you, nor
fields of
offerings: for there the
shield of the
mighty is vilely cast
away, the
shield of
Saul,
{\ADD{as though he had}} not
{\ADD{been}}
anointed with
oil.
\VS{22}From the
blood of the
slain, from the
fat of the
mighty, the
bow of
Jonathan
turned not
back, and the
sword of
Saul
returned not
empty.
\VS{23}Saul and
Jonathan
{\ADD{were}}
lovely and
pleasant in their
lives, and in their
death they were not
divided: they were
swifter than
eagles, they were
stronger than
lions.\FTNT{**}{{\FR 1.23 }
pleasant: or, sweet}
\VS{24}Ye
daughters of
Israel, weep
over
Saul, who
clothed you in
scarlet, with
{\ADD{other}}
delights, who
put on
ornaments of
gold upon your
apparel.
\VS{25}How are the
mighty
fallen in the
midst of the
battle! O
Jonathan,
{\ADD{thou wast}}
slain in thine high
places.
\VS{26}I am
distressed for thee, my
brother
Jonathan:
very
pleasant hast thou been unto me: thy
love to me was
wonderful, passing the
love of
women.
\VS{27}How are the
mighty
fallen, and the
weapons of
war
perished!

\par }\Chap{2}{\PP \VerseOne{1}And it came to pass after
this, that
David
enquired of the
{\ND{
LORD}},
saying, Shall I go
up into
any of the
cities of
Judah? And the
{\ND{
LORD}}
said unto him, Go
up. And
David
said, Whither shall I go
up? And he
said, Unto
Hebron.
\VS{2}So
David went
up thither, and his
two
wives also,
Ahinoam the
Jezreelitess, and
Abigail
Nabal’s
wife the
Carmelite.
\VS{3}And his
men that
{\ADD{were}} with him did
David bring
up, every
man with his
household: and they
dwelt in the
cities of
Hebron.
\VS{4}And the
men of
Judah
came, and there they
anointed
David
king over the
house of
Judah. And they
told
David,
saying,
{\ADD{That}} the
men of
Jabesh-gilead
{\ADD{were they}} that
buried
Saul.
\par }{\PP \VS{5}¶ And
David
sent
messengers unto the
men of
Jabesh-gilead, and
said unto them,
Blessed
{\ADD{be}} ye of the
{\ND{
LORD}}, that ye have
shewed this
kindness unto your
lord,
{\ADD{even}} unto
Saul, and have
buried him.
\VS{6}And now the
{\ND{
LORD}}
shew
kindness and
truth unto you: and I also will
requite you this
kindness, because ye have
done this
thing.
\VS{7}Therefore now let your
hands be
strengthened, and be ye
valiant: for your
master
Saul is
dead, and also the
house of
Judah have
anointed me
king over them.\FTNT{*}{{\FR 2.7 }
be ye…: Heb. be ye the sons of valour}
\par }{\PP \VS{8}¶ But
Abner the
son of
Ner,
captain of
Saul’s
host,
took
Ish-bosheth the
son of
Saul, and brought him
over to
Mahanaim;\FTNT{†}{{\FR 2.8 }
Saul’s host: Heb. the host which was Saul’s}\FTNT{‡}{{\FR 2.8 }
Ish-bosheth: or, Esh-baal}
\VS{9}And made him
king over
Gilead, and over the
Ashurites, and over
Jezreel, and over
Ephraim, and over
Benjamin, and over all
Israel.
\VS{10}Ish-bosheth
Saul’s
son
{\ADD{was}}
forty
years
old when he began to
reign over
Israel, and
reigned
two
years. But the
house of
Judah
followed
David.
\VS{11}And the
time that
David was
king in
Hebron over the
house of
Judah was
seven
years and
six
months.\FTNT{§}{{\FR 2.11 }
time: Heb. number of days}
\par }{\PP \VS{12}¶ And
Abner the
son of
Ner, and the
servants of
Ish-bosheth the
son of
Saul, went
out from
Mahanaim to
Gibeon.
\VS{13}And
Joab the
son of
Zeruiah, and the
servants of
David, went
out, and
met
together by the
pool of
Gibeon: and they sat
down, the one on the one side of the
pool, and the other on the other side of the
pool.\FTNT{**}{{\FR 2.13 }
together: Heb. them together}
\VS{14}And
Abner
said to
Joab, Let the young
men now
arise, and
play
before us. And
Joab
said, Let them
arise.
\VS{15}Then there
arose and went
over by
number
twelve of
Benjamin, which
{\ADD{pertained}} to
Ish-bosheth the
son of
Saul, and
twelve of the
servants of
David.
\VS{16}And they
caught every
one his
fellow by the
head, and
{\ADD{thrust}} his
sword in his
fellow’s
side; so they fell
down
together: wherefore that
place was
called
Helkath-hazzurim, which
{\ADD{is}} in
Gibeon.\FTNT{††}{{\FR 2.16 }
Helkath-hazzurim: that is, The field of strong men}
\VS{17}And there was a
very
sore
battle that
day; and
Abner was
beaten, and the
men of
Israel,
before the
servants of
David.
\par }{\PP \VS{18}¶ And there were
three
sons of
Zeruiah there,
Joab, and
Abishai, and
Asahel: and
Asahel
{\ADD{was as}}
light of
foot as
a
wild
roe.\FTNT{‡‡}{{\FR 2.18 }
of foot: Heb. of his feet}\FTNT{§§}{{\FR 2.18 }
as a wild…: Heb. as one of the roes that is in the field}
\VS{19}And
Asahel
pursued
after
Abner; and in
going he
turned not to the right
hand nor to the
left from
following
Abner.\FTNT{***}{{\FR 2.19 }
from…: Heb. from after Abner}
\VS{20}Then
Abner
looked
behind him, and
said,
{\ADD{Art}} thou
Asahel? And he
answered, I
{\ADD{am}}.
\VS{21}And
Abner
said to him, Turn thee
aside to thy right
hand or to thy
left, and lay thee
hold on
one of the young
men, and
take thee his
armour. But
Asahel
would not turn
aside from
following of him.\FTNT{†††}{{\FR 2.21 }
armour: or, spoil}
\VS{22}And
Abner
said
again to
Asahel, Turn thee
aside from
following me: wherefore should I
smite thee to the
ground? how then should I hold
up my
face to
Joab thy
brother?
\VS{23}Howbeit he
refused to turn
aside: wherefore
Abner with the hinder
end of the
spear
smote him
under the
fifth
{\ADD{rib}}, that the
spear came
out
behind him; and he fell
down there, and
died in the same
place: and it came to pass,
{\ADD{that}} as many as
came to the
place where
Asahel fell
down and
died stood
still.
\VS{24}Joab also and
Abishai
pursued
after
Abner: and the
sun went
down when they were
come to the
hill of
Ammah, that
{\ADD{lieth}}
before
Giah by the
way of the
wilderness of
Gibeon.
\par }{\PP \VS{25}¶ And the
children of
Benjamin gathered themselves
together
after
Abner, and became
one
troop, and
stood on the
top of an
hill.
\VS{26}Then
Abner
called to
Joab, and
said, Shall the
sword
devour for
ever?
knowest thou not that it will be
bitterness in the latter
end? how long shall it be then, ere thou
bid the
people
return from
following their
brethren?
\VS{27}And
Joab
said,
{\ADD{As}}
God
liveth,
unless thou hadst
spoken, surely then in the
morning the
people had gone
up every
one from
following his
brother.\FTNT{‡‡‡}{{\FR 2.27 }
in the…: Heb. from the morning}\FTNT{§§§}{{\FR 2.27 }
gone…: or, gone away}
\VS{28}So
Joab
blew a
trumpet, and all the
people stood
still, and
pursued
after
Israel no more, neither
fought they any
more.
\VS{29}And
Abner and his
men
walked all that
night through the
plain, and passed
over
Jordan, and went
through all
Bithron, and they
came to
Mahanaim.
\VS{30}And
Joab
returned from
following
Abner: and when he had
gathered all the
people
together, there
lacked of
David’s
servants
nineteen
men and
Asahel.
\VS{31}But the
servants of
David had
smitten of
Benjamin, and of
Abner’s
men,
{\ADD{so that}}
three
hundred and
threescore
men
died.
\par }{\PP \VS{32}¶ And they took
up
Asahel, and
buried him in the
sepulchre of his
father, which
{\ADD{was in}}
Beth-lehem. And
Joab and his
men
went all
night, and they came to
Hebron at break of
day.

\par }\Chap{3}{\PP \VerseOne{1}Now there was
long
war between the
house of
Saul and the
house of
David: but
David
waxed
stronger and
stronger, and the
house of
Saul
waxed
weaker and
weaker.
\par }{\PP \VS{2}¶ And unto
David were
sons
born in
Hebron: and his
firstborn was
Amnon, of
Ahinoam the
Jezreelitess;
\VS{3}And his
second,
Chileab, of
Abigail the
wife of
Nabal the
Carmelite; and the
third,
Absalom the
son of
Maacah the
daughter of
Talmai
king of
Geshur;\FTNT{*}{{\FR 3.3 }
Chileab: or, Daniel}
\VS{4}And the
fourth,
Adonijah the
son of
Haggith; and the
fifth,
Shephatiah the
son of
Abital;
\VS{5}And the
sixth,
Ithream, by
Eglah
David’s
wife. These were
born to
David in
Hebron.
\par }{\PP \VS{6}¶ And it came to pass, while there was
war between the
house of
Saul and the
house of
David, that
Abner made himself
strong for the
house of
Saul.
\par }{\PP \VS{7}And
Saul had a
concubine, whose
name
{\ADD{was}}
Rizpah, the
daughter of
Aiah: and
{\ADD{Ish-bosheth}}
said to
Abner, Wherefore hast thou gone
in unto my
father’s
concubine?
\VS{8}Then was
Abner
very
wroth for the
words of
Ish-bosheth, and
said,
{\ADD{Am}} I a
dog’s
head, which against
Judah do
shew
kindness this
day unto the
house of
Saul thy
father, to his
brethren, and to his
friends, and have not
delivered thee into the
hand of
David, that thou
chargest me to
day with a
fault concerning this
woman?
\VS{9}So
do
God to
Abner, and more
also, except, as the
{\ND{
LORD}} hath
sworn to
David, even so I
do to him;
\VS{10}To
translate the
kingdom from the
house of
Saul, and to set
up the
throne of
David over
Israel and over
Judah, from
Dan even to
Beer-sheba.
\VS{11}And he
could not
answer
Abner a
word again, because he
feared him.
\par }{\PP \VS{12}¶ And
Abner
sent
messengers to
David on his behalf,
saying, Whose
{\ADD{is}} the
land?
saying
{\ADD{also}},
Make thy
league with me, and, behold, my
hand
{\ADD{shall be}} with thee, to bring
about all
Israel unto thee.
\par }{\PP \VS{13}¶ And he
said,
Well; I will
make a
league with thee: but
one
thing I
require of thee, that
is, Thou shalt not
see my
face, except thou
first
bring
Michal
Saul’s
daughter, when thou
comest to
see my
face.\FTNT{†}{{\FR 3.13 }
that is: Heb. saying}
\VS{14}And
David
sent
messengers to
Ish-bosheth
Saul’s
son,
saying,
Deliver
{\ADD{me}} my
wife
Michal, which I
espoused to me for an
hundred
foreskins of the
Philistines.
\VS{15}And
Ish-bosheth
sent, and
took her from
{\ADD{her}}
husband,
{\ADD{even}} from
Phaltiel the
son of
Laish.\FTNT{‡}{{\FR 3.15 }
Phaltiel: also called, Phalti}
\VS{16}And her
husband
went with her along
weeping
behind her to
Bahurim. Then
said
Abner unto him,
Go,
return. And he
returned.\FTNT{§}{{\FR 3.16 }
along…: Heb. going and weeping}
\par }{\PP \VS{17}¶ And
Abner had
communication with the
elders of
Israel,
saying, Ye
sought for
David in
times
past
{\ADD{to be}}
king over you:\FTNT{**}{{\FR 3.17 }
in times…: Heb. both yesterday and the third day}
\VS{18}Now then
do
{\ADD{it:}} for the
{\ND{
LORD}} hath
spoken of
David,
saying, By the
hand of my
servant
David I will
save my
people
Israel out of the
hand of the
Philistines, and out of the
hand of all their
enemies.
\VS{19}And
Abner also
spake in the
ears of
Benjamin: and
Abner
went also to
speak in the
ears of
David in
Hebron all that
seemed
good to
Israel, and that seemed
good to the whole
house of
Benjamin.
\VS{20}So
Abner
came to
David to
Hebron, and
twenty
men with him. And
David
made
Abner and the
men that
{\ADD{were}} with him a
feast.
\VS{21}And
Abner
said unto
David, I will
arise and
go, and will
gather all
Israel unto my
lord the
king, that they may
make a
league with thee, and that thou mayest
reign over all that thine
heart
desireth. And
David
sent
Abner
away; and he
went in
peace.
\par }{\PP \VS{22}¶ And, behold, the
servants of
David and
Joab
came from
{\ADD{pursuing}} a
troop, and brought
in a
great
spoil with them: but
Abner
{\ADD{was}} not with
David in
Hebron; for he had sent him
away, and he was
gone in
peace.
\VS{23}When
Joab and all the
host that
{\ADD{was}} with him were
come, they
told
Joab,
saying,
Abner the
son of
Ner
came to the
king, and he hath sent him
away, and he is
gone in
peace.
\VS{24}Then
Joab
came to the
king, and
said, What hast thou
done? behold,
Abner
came unto thee; why
{\ADD{is}} it
{\ADD{that}} thou hast sent him
away, and he is
quite
gone?
\VS{25}Thou
knowest
Abner the
son of
Ner, that he
came to
deceive thee, and to
know thy going
out and thy coming
in, and to
know all that thou
doest.
\VS{26}And when
Joab was come
out from
David, he
sent
messengers
after
Abner, which brought him
again from the
well of
Sirah: but
David
knew
{\ADD{it}} not.
\VS{27}And when
Abner was
returned to
Hebron,
Joab took him
aside
in the
gate to
speak with him
quietly, and
smote him there under the
fifth
{\ADD{rib}}, that he
died, for the
blood of
Asahel his
brother.\FTNT{††}{{\FR 3.27 }
quietly: or, peaceably}
\par }{\PP \VS{28}¶ And
afterward when
David
heard
{\ADD{it}}, he
said, I and my
kingdom
{\ADD{are}}
guiltless before the
{\ND{
LORD}}
for
ever from the
blood of
Abner the
son of
Ner:\FTNT{‡‡}{{\FR 3.28 }
blood: Heb. bloods}
\VS{29}Let it
rest on the
head of
Joab, and on all his
father’s
house; and let there not
fail from the
house of
Joab one that hath an
issue, or that is a
leper, or that
leaneth on a
staff, or that
falleth on the
sword, or that
lacketh
bread.\FTNT{§§}{{\FR 3.29 }
fail: Heb. be cut off}
\VS{30}So
Joab and
Abishai his
brother
slew
Abner, because he had
slain their
brother
Asahel at
Gibeon in the
battle.
\par }{\PP \VS{31}¶ And
David
said to
Joab, and to all the
people that
{\ADD{were}} with him,
Rend your
clothes, and
gird you with
sackcloth, and
mourn
before
Abner. And
king
David
{\ADD{himself}}
followed the
bier.\FTNT{***}{{\FR 3.31 }
bier: Heb. bed}
\VS{32}And they
buried
Abner in
Hebron: and the
king lifted
up his
voice, and
wept at the
grave of
Abner; and all the
people
wept.
\VS{33}And the
king
lamented over
Abner, and
said,
Died
Abner as a
fool
dieth?
\VS{34}Thy
hands
{\ADD{were}} not
bound, nor thy
feet
put into
fetters: as a man
falleth
before
wicked
men,
{\ADD{so}}
fellest thou. And all the
people
wept
again over him.\FTNT{†††}{{\FR 3.34 }
wicked…: Heb. children of iniquity}
\VS{35}And when
all the
people
came to
cause
David to
eat
meat while it was
yet
day,
David
sware,
saying, So
do
God to me, and more
also, if I
taste
bread, or
ought else,
till the
sun be
down.
\VS{36}And all the
people took
notice
{\ADD{of it}}, and it
pleased them: as whatsoever the
king
did
pleased all the
people.\FTNT{‡‡‡}{{\FR 3.36 }
pleased them: Heb. was good in their eyes}
\VS{37}For all the
people and all
Israel
understood that
day that it was not of the
king to
slay
Abner the
son of
Ner.
\VS{38}And the
king
said unto his
servants,
Know ye not that there is a
prince and a great
man
fallen this
day in
Israel?
\VS{39}And I
{\ADD{am}} this
day
weak, though
anointed
king; and these
men the
sons of
Zeruiah
{\ADD{be}} too
hard for me: the
{\ND{
LORD}} shall
reward the doer of
evil according to his
wickedness.\FTNT{§§§}{{\FR 3.39 }
weak: Heb. tender}

\par }\Chap{4}{\PP \VerseOne{1}And when
Saul’s
son
heard that
Abner was
dead in
Hebron, his
hands were
feeble, and all the
Israelites were
troubled.
\VS{2}And
Saul’s
son
had
two
men
{\ADD{that were}}
captains of
bands: the
name of the
one
{\ADD{was}}
Baanah, and the
name of the
other
Rechab, the
sons of
Rimmon a
Beerothite, of the
children of
Benjamin: (for
Beeroth also was
reckoned to
Benjamin:\FTNT{*}{{\FR 4.2 }
other: Heb. second}
\VS{3}And the
Beerothites
fled to
Gittaim, and were
sojourners there until this
day.)
\VS{4}And
Jonathan,
Saul’s
son, had a
son
{\ADD{that was}}
lame of
{\ADD{his}}
feet. He was
five
years
old when the
tidings
came of
Saul and
Jonathan out of
Jezreel, and his
nurse took him
up, and
fled: and it came to pass, as she made
haste to
flee, that he
fell, and became
lame. And his
name
{\ADD{was}}
Mephibosheth.\FTNT{†}{{\FR 4.4 }
Mephibosheth: or, Merib-baal}
\VS{5}And the
sons of
Rimmon the
Beerothite,
Rechab and
Baanah,
went, and
came about the
heat of the
day to the
house of
Ish-bosheth, who
lay on a
bed at
noon.
\VS{6}And they came
thither into the
midst of the
house,
{\ADD{as though}} they would have
fetched
wheat; and they
smote him under the
fifth
{\ADD{rib:}} and
Rechab and
Baanah his
brother
escaped.
\VS{7}For when they
came into the
house, he
lay on his
bed in his
bedchamber, and they
smote him, and
slew him, and
beheaded him, and
took his
head, and gat them
away
through the
plain all
night.
\VS{8}And they
brought the
head of
Ish-bosheth unto
David to
Hebron, and
said to the
king, Behold the
head of
Ish-bosheth the
son of
Saul thine
enemy, which
sought thy
life; and the
{\ND{
LORD}}
hath
avenged my
lord the
king this
day of
Saul, and of his
seed.
\par }{\PP \VS{9}¶ And
David
answered
Rechab and
Baanah his
brother, the
sons of
Rimmon the
Beerothite, and
said unto them,
{\ADD{As}} the
{\ND{
LORD}}
liveth, who hath
redeemed my
soul out of all
adversity,
\VS{10}When one
told me,
saying, Behold,
Saul is
dead, thinking to have
brought good
tidings, I took
hold of him, and
slew him in
Ziklag, who
{\ADD{thought}} that I would have
given him a reward for his
tidings:\FTNT{‡}{{\FR 4.10 }
thinking…: Heb. he was in his own eyes as a bringer, etc}\FTNT{§}{{\FR 4.10 }
who…: or, which was the reward I gave him for his tidings}
\VS{11}How much more, when
wicked
men have
slain a
righteous
person in his own
house upon his
bed? shall I not therefore now
require his
blood of your
hand, and take you
away from the
earth?
\VS{12}And
David
commanded his young
men, and they
slew them, and cut
off their
hands and their
feet, and hanged
{\ADD{them}}
up over the
pool in
Hebron. But they
took the
head of
Ish-bosheth, and
buried
{\ADD{it}} in the
sepulchre of
Abner in
Hebron.

\par }\Chap{5}{\PP \VerseOne{1}Then
came all the
tribes of
Israel to
David unto
Hebron, and
spake,
saying, Behold, we
{\ADD{are}} thy
bone and thy
flesh.
\VS{2}Also in time
past, when
Saul was
king over us, thou wast he that leddest
out and broughtest
in
Israel: and the
{\ND{
LORD}}
said to thee, Thou shalt
feed my
people
Israel, and thou shalt be a
captain over
Israel.
\VS{3}So all the
elders of
Israel
came to the
king to
Hebron; and
king
David
made a
league with them in
Hebron
before the
{\ND{
LORD}}: and they
anointed
David
king over
Israel.
\par }{\PP \VS{4}¶
David
{\ADD{was}}
thirty
years
old when he began to
reign,
{\ADD{and}} he
reigned
forty
years.
\VS{5}In
Hebron he
reigned over
Judah
seven
years and
six
months: and in
Jerusalem he
reigned
thirty and
three
years over all
Israel and
Judah.
\par }{\PP \VS{6}¶ And the
king and his
men
went to
Jerusalem unto the
Jebusites, the
inhabitants of the
land: which
spake unto
David,
saying, Except thou take
away the
blind and the
lame, thou shalt not come
in hither:
thinking,
David cannot come
in hither.\FTNT{*}{{\FR 5.6 }
thinking…: or, saying David shall not, etc}
\VS{7}Nevertheless
David
took the strong
hold of
Zion: the same
{\ADD{is}} the
city of
David.
\VS{8}And
David
said on that
day, Whosoever getteth
up to the
gutter, and
smiteth the
Jebusites, and the
lame and the
blind,
{\ADD{that are}}
hated of
David’s
soul,
{\ADD{he shall be chief and captain}}. Wherefore they
said, The
blind and the
lame shall not
come into the
house.\FTNT{†}{{\FR 5.8 }
Wherefore…: or, Because they had said, even the blind and the lame, He shall not}
\VS{9}So
David
dwelt in the
fort, and
called it the
city of
David. And
David
built round
about from
Millo and
inward.
\VS{10}And
David
went
on, and grew
great, and the
{\ND{
LORD}}
God of
hosts
{\ADD{was}} with him.\FTNT{‡}{{\FR 5.10 }
went…: Heb. went, going and growing}
\par }{\PP \VS{11}¶ And
Hiram
king of
Tyre
sent
messengers to
David, and
cedar
trees, and
carpenters, and
masons: and they
built
David an
house.\FTNT{§}{{\FR 5.11 }
masons: Heb. hewers of the stone of the wall}
\VS{12}And
David
perceived that the
{\ND{
LORD}} had
established him
king over
Israel, and that he had
exalted his
kingdom for his
people
Israel’s sake.
\par }{\PP \VS{13}¶ And
David
took
{\ADD{him}} more
concubines and
wives out of
Jerusalem,
after he was
come from
Hebron: and there were yet
sons and
daughters
born to
David.
\VS{14}And these
{\ADD{be}} the
names of those that were
born unto him in
Jerusalem;
Shammua, and
Shobab, and
Nathan, and
Solomon,
\VS{15}Ibhar also, and
Elishua, and
Nepheg, and
Japhia,
\VS{16}And
Elishama, and
Eliada, and
Eliphalet.
\par }{\PP \VS{17}¶ But when the
Philistines
heard that they had
anointed
David
king over
Israel, all the
Philistines came
up to
seek
David; and
David
heard
{\ADD{of it}}, and went
down to the
hold.
\VS{18}The
Philistines also
came and
spread themselves in the
valley of
Rephaim.
\VS{19}And
David
enquired of the
{\ND{
LORD}},
saying, Shall I go
up to the
Philistines? wilt thou
deliver them into mine
hand? And the
{\ND{
LORD}}
said unto
David, Go
up: for I will
doubtless
deliver the
Philistines into thine
hand.
\VS{20}And
David
came to
Baal-perazim, and
David
smote them there, and
said, The
{\ND{
LORD}} hath broken
forth upon mine
enemies
before me, as the
breach of
waters. Therefore he
called the
name of that
place
Baal-perazim.\FTNT{**}{{\FR 5.20 }
Baal-perazim: that is, the plain of breaches}
\VS{21}And there they
left their
images, and
David and his
men
burned them.\FTNT{††}{{\FR 5.21 }
burned…: or, took them away}
\par }{\PP \VS{22}¶ And the
Philistines came
up yet
again, and
spread themselves in the
valley of
Rephaim.
\VS{23}And when
David
enquired of the
{\ND{
LORD}}, he
said, Thou shalt not go
up;
{\ADD{but}} fetch a
compass
behind them, and
come upon them over
against the mulberry
trees.
\VS{24}And let it be, when thou
hearest the
sound of a
going in the
tops of the mulberry
trees, that then thou shalt
bestir thyself: for then shall the
{\ND{
LORD}} go
out
before thee, to
smite the
host of the
Philistines.
\VS{25}And
David did
so, as the
{\ND{
LORD}} had
commanded him; and
smote the
Philistines from
Geba until thou
come to
Gazer.\FTNT{‡‡}{{\FR 5.25 }
Geba: also called, Gibeon}

\par }\Chap{6}{\PP \VerseOne{1}Again,
David gathered
together all
{\ADD{the}}
chosen
{\ADD{men}} of
Israel,
thirty
thousand.
\VS{2}And
David
arose, and
went with all the
people that
{\ADD{were}} with him from Baale of
Judah, to bring
up from thence the
ark of
God, whose
name is
called by the
name of the
{\ND{
LORD}} of
hosts that
dwelleth
{\ADD{between}} the
cherubims.\FTNT{*}{{\FR 6.2 }
Baale: or, Baalah, that is, Kirjath-jearim}\FTNT{†}{{\FR 6.2 }
whose name…: or, at which the name, even the name of the
{\ND{LORD}} of hosts, was called upon}
\VS{3}And they
set the
ark of
God upon a
new
cart, and
brought it out of the
house of
Abinadab that
{\ADD{was}} in
Gibeah: and
Uzzah and
Ahio, the
sons of
Abinadab,
drave the
new
cart.\FTNT{‡}{{\FR 6.3 }
set: Heb. made to ride}\FTNT{§}{{\FR 6.3 }
Gibeah: or, the hill}
\VS{4}And they
brought it out of the
house of
Abinadab which
{\ADD{was}} at
Gibeah,
accompanying the
ark of
God: and
Ahio
went
before the
ark.\FTNT{**}{{\FR 6.4 }
accompanying: Heb. with}
\VS{5}And
David and all the
house of
Israel
played
before the
{\ND{
LORD}} on all manner of
{\ADD{instruments made of}}
fir
wood, even on
harps, and on
psalteries, and on
timbrels, and on
cornets, and on
cymbals.
\par }{\PP \VS{6}¶ And when they
came to
Nachon’s
threshingfloor,
Uzzah put
forth
{\ADD{his hand}} to the
ark of
God, and took
hold of it; for the
oxen
shook
{\ADD{it}}.\FTNT{††}{{\FR 6.6 }
Nachon: also called Chidon}\FTNT{‡‡}{{\FR 6.6 }
shook it: or, stumbled}
\VS{7}And the
anger of the
{\ND{
LORD}} was
kindled against
Uzzah; and
God
smote him there for
{\ADD{his}}
error; and there he
died by the
ark of
God.\FTNT{§§}{{\FR 6.7 }
error: or, rashness}
\VS{8}And
David was
displeased, because the
{\ND{
LORD}} had
made a
breach upon
Uzzah: and he
called the name of the
place
Perez-uzzah to this
day.\FTNT{***}{{\FR 6.8 }
made: Heb. broken}\FTNT{†††}{{\FR 6.8 }
Perez-uzzah: that is, the breach of Uzzah}
\VS{9}And
David was
afraid of the
{\ND{
LORD}} that
day, and
said, How shall the
ark of the
{\ND{
LORD}}
come to me?
\VS{10}So
David
would not
remove the
ark of the
{\ND{
LORD}} unto him into the
city of
David: but
David carried it
aside into the
house of
Obed-edom the
Gittite.
\VS{11}And the
ark of the
{\ND{
LORD}}
continued in the
house of
Obed-edom the
Gittite
three
months: and the
{\ND{
LORD}}
blessed
Obed-edom, and all his
household.
\par }{\PP \VS{12}¶ And it was
told
king
David,
saying, The
{\ND{
LORD}} hath
blessed the
house of
Obed-edom, and all that
{\ADD{pertaineth}} unto him, because of the
ark of
God. So
David
went and brought
up the
ark of
God from the
house of
Obed-edom into the
city of
David with
gladness.
\VS{13}And it was
{\ADD{so}}, that when they that
bare the
ark of the
{\ND{
LORD}} had
gone
six
paces, he
sacrificed
oxen and
fatlings.
\VS{14}And
David
danced
before the
{\ND{
LORD}} with all
{\ADD{his}}
might; and
David
{\ADD{was}}
girded with a
linen
ephod.
\VS{15}So
David and all the
house of
Israel brought
up the
ark of the
{\ND{
LORD}} with
shouting, and with the
sound of the
trumpet.
\VS{16}And as the
ark of the
{\ND{
LORD}}
came into the
city of
David,
Michal
Saul’s
daughter
looked through a
window, and
saw
king
David
leaping and
dancing
before the
{\ND{
LORD}}; and she
despised him in her
heart.
\par }{\PP \VS{17}¶ And they
brought in the
ark of the
{\ND{
LORD}}, and
set it in his
place, in the
midst of the
tabernacle that
David had
pitched for it: and
David
offered burnt
offerings and peace
offerings
before the
{\ND{
LORD}}.\FTNT{‡‡‡}{{\FR 6.17 }
pitched: Heb. stretched}
\VS{18}And as soon as
David had made an
end of
offering burnt
offerings and peace
offerings, he
blessed the
people in the
name of the
{\ND{
LORD}} of
hosts.
\VS{19}And he
dealt among all the
people,
{\ADD{even}} among the whole
multitude of
Israel, as well to the
women as
men, to every
one
a
cake of
bread, and a good
piece
{\ADD{of flesh}}, and a
flagon
{\ADD{of wine}}. So all the
people
departed every
one to his
house.
\par }{\PP \VS{20}¶ Then
David
returned to
bless his
household. And
Michal the
daughter of
Saul came
out to
meet
David, and
said, How
glorious was the
king of
Israel to
day, who
uncovered himself to
day in the
eyes of the
handmaids of his
servants, as
one of the vain
fellows
shamelessly
uncovereth himself!\FTNT{§§§}{{\FR 6.20 }
shamelessly: or, openly}
\VS{21}And
David
said unto
Michal,
{\ADD{It was}}
before the
{\ND{
LORD}}, which
chose me before thy
father, and before all his
house, to
appoint me
ruler over the
people of the
{\ND{
LORD}}, over
Israel: therefore will I
play
before the
{\ND{
LORD}}.
\VS{22}And I will yet be more
vile than
thus, and will be
base in mine own
sight: and of the
maidservants which thou hast
spoken of, of them shall I be had in
honour.\FTNT{*}{{\FR 6.22 }
of the…: or, of the handmaids of my servants}
\VS{23}Therefore
Michal the
daughter of
Saul had no
child unto the
day of her
death.

\par }\Chap{7}{\PP \VerseOne{1}And it came to pass, when the
king
sat in his
house, and the
{\ND{
LORD}} had given him
rest round
about from all his
enemies;
\VS{2}That the
king
said unto
Nathan the
prophet,
See now, I
dwell in an
house of
cedar, but the
ark of
God
dwelleth
within
curtains.
\VS{3}And
Nathan
said to the
king,
Go,
do all that
{\ADD{is}} in thine
heart; for the
{\ND{
LORD}}
{\ADD{is}} with thee.
\par }{\PP \VS{4}¶ And it came to pass that
night, that the
word of the
{\ND{
LORD}} came unto
Nathan,
saying,
\VS{5}Go and
tell my
servant
David, Thus
saith the
{\ND{
LORD}}, Shalt thou
build me an
house for me to dwell
in?\FTNT{*}{{\FR 7.5 }
my servant…: Heb. to my servant, to David}
\VS{6}Whereas I have not
dwelt in
{\ADD{any}}
house since the
time that I brought
up the
children of
Israel out of
Egypt, even to this
day, but have
walked in a
tent and in a
tabernacle.
\VS{7}In all
{\ADD{the places}} wherein I have
walked with all the
children of
Israel
spake I a
word with
any of the
tribes of
Israel, whom I
commanded to
feed my
people
Israel,
saying, Why
build ye not me an
house of
cedar?\FTNT{†}{{\FR 7.7 }
any of…: any of the judges}
\VS{8}Now therefore so shalt thou
say unto my
servant
David, Thus
saith the
{\ND{
LORD}} of
hosts, I
took thee from the
sheepcote, from
following the
sheep, to be
ruler over my
people, over
Israel:\FTNT{‡}{{\FR 7.8 }
from following: Heb. from after}
\VS{9}And I was with thee whithersoever thou
wentest, and have cut
off all thine
enemies out of thy
sight, and have
made thee a
great
name, like unto the
name of the
great
{\ADD{men}} that
{\ADD{are}} in the
earth.\FTNT{§}{{\FR 7.9 }
out of…: Heb. from thy face}
\VS{10}Moreover I will
appoint a
place for my
people
Israel, and will
plant them, that they may
dwell in a place of their own, and
move no more; neither shall the
children of
wickedness
afflict them any
more, as
beforetime,
\VS{11}And as since the
time that I
commanded
judges
{\ADD{to be}} over my
people
Israel, and have caused thee to
rest from all thine
enemies. Also the
{\ND{
LORD}}
telleth thee that he will
make thee an
house.
\par }{\PP \VS{12}¶ And when thy
days be
fulfilled, and thou shalt
sleep with thy
fathers, I will set
up thy
seed
after thee, which shall
proceed out of thy
bowels, and I will
establish his
kingdom.
\VS{13}He shall
build an
house for my
name, and I will
stablish the
throne of his
kingdom
for
ever.
\VS{14}I will be his
father, and he shall be my
son. If he commit
iniquity, I will
chasten him with the
rod of
men, and with the
stripes of the
children of
men:
\VS{15}But my
mercy shall not depart
away from him, as I
took
{\ADD{it}} from
Saul, whom I put
away
before thee.
\VS{16}And thine
house and thy
kingdom shall be
established
for
ever
before thee: thy
throne shall be
established
for
ever.
\VS{17}According to all these
words, and according to all this
vision, so did
Nathan
speak unto
David.
\par }{\PP \VS{18}¶ Then
went
king
David in, and
sat
before the
{\ND{
LORD}}, and he
said, Who
{\ADD{am}} I, O
Lord
{\ND{
GOD}}? and what
{\ADD{is}} my
house, that thou hast
brought me
hitherto?
\VS{19}And this was yet a small
thing in thy
sight, O
Lord
{\ND{
GOD}}; but thou hast
spoken also of thy
servant’s
house for a great while to
come. And
{\ADD{is}} this the
manner of
man, O
Lord
{\ND{
GOD}}?\FTNT{**}{{\FR 7.19 }
manner: Heb. law}
\VS{20}And what
can
David
say
more unto thee? for thou,
Lord
{\ND{
GOD}},
knowest thy
servant.
\VS{21}For thy
word’s sake, and according to thine own
heart, hast thou
done all these great
things, to make thy
servant
know
{\ADD{them}}.
\VS{22}Wherefore thou art
great, O
{\ND{
LORD}}
God: for
{\ADD{there is}} none like thee, neither
{\ADD{is there any}}
God
beside thee, according to all that we have
heard with our
ears.
\VS{23}And what
one
nation in the
earth
{\ADD{is}} like thy
people,
{\ADD{even}} like
Israel, whom
God
went to
redeem for a
people to himself, and to
make him a
name, and to
do for you great
things and
terrible, for thy
land,
before thy
people, which thou
redeemedst to thee from
Egypt,
{\ADD{from}} the
nations and their
gods?
\VS{24}For thou hast
confirmed to thyself thy
people
Israel
{\ADD{to be}} a
people unto thee
for
ever: and thou,
{\ND{
LORD}}, art
become their
God.
\VS{25}And now, O
{\ND{
LORD}}
God, the
word that thou hast
spoken concerning thy
servant, and concerning his
house,
establish
{\ADD{it}}
for
ever, and
do as thou hast
said.
\VS{26}And let thy
name be
magnified
for
ever,
saying, The
{\ND{
LORD}} of
hosts
{\ADD{is}} the
God over
Israel: and let the
house of thy
servant
David be
established
before thee.
\VS{27}For thou, O
{\ND{
LORD}} of
hosts,
God of
Israel, hast
revealed to thy
servant,
saying, I will
build thee an
house: therefore hath thy
servant
found in his
heart to
pray this
prayer unto thee.\FTNT{††}{{\FR 7.27 }
revealed: Heb. opened the ear}
\VS{28}And now, O
Lord
{\ND{
GOD}}, thou
{\ADD{art}} that
God, and thy
words be
true, and thou hast
promised this
goodness unto thy
servant:
\VS{29}Therefore now let it
please thee to
bless the
house of thy
servant, that it may continue for
ever
before thee: for thou, O
Lord
{\ND{
GOD}}, hast
spoken
{\ADD{it:}} and with thy
blessing let the
house of thy
servant be
blessed for
ever.\FTNT{‡‡}{{\FR 7.29 }
let it…: Heb. be thou pleased and bless}

\par }\Chap{8}{\PP \VerseOne{1}And
after this it came to pass, that
David
smote the
Philistines, and
subdued them: and
David
took
Metheg-ammah out of the
hand of the
Philistines.\FTNT{*}{{\FR 8.1 }
Metheg-ammah: or, the bridle of Ammah}
\VS{2}And he
smote
Moab, and
measured them with a
line, casting them
down to the
ground; even with
two
lines
measured he to put to
death, and with one
full
line to keep
alive. And
{\ADD{so}} the
Moabites became
David’s
servants,
{\ADD{and}}
brought
gifts.
\par }{\PP \VS{3}¶
David
smote also
Hadadezer, the
son of
Rehob,
king of
Zobah, as he
went to
recover his
border at the
river
Euphrates.\FTNT{†}{{\FR 8.3 }
Hadadezer: or, Hadarezer}
\VS{4}And
David
took from him a
thousand
{\ADD{chariots}}, and
seven
hundred
horsemen, and
twenty
thousand
footmen: and
David
houghed all the
chariot
{\ADD{horses}}, but
reserved of them
{\ADD{for}} an
hundred
chariots.\FTNT{‡}{{\FR 8.4 }
from…: or, of his chariots}
\VS{5}And when the
Syrians of
Damascus
came to
succour
Hadadezer
king of
Zobah,
David
slew of the
Syrians
two and
twenty
thousand
men.
\VS{6}Then
David
put
garrisons in
Syria of
Damascus: and the
Syrians became
servants to
David,
{\ADD{and}}
brought
gifts. And the
{\ND{
LORD}}
preserved
David whithersoever he
went.
\VS{7}And
David
took the
shields of
gold that were on the
servants of
Hadadezer, and
brought them to
Jerusalem.
\VS{8}And from
Betah, and from
Berothai,
cities of
Hadadezer,
king
David
took
exceeding
much
brass.
\par }{\PP \VS{9}¶ When
Toi
king of
Hamath
heard that
David had
smitten all the
host of
Hadadezer,
\VS{10}Then
Toi
sent
Joram his
son unto
king
David, to
salute him, and to
bless him, because he had
fought against
Hadadezer, and
smitten him: for
Hadadezer had
wars with
Toi. And
{\ADD{Joram}} brought with
him
vessels of
silver, and
vessels of
gold, and
vessels of
brass:\FTNT{§}{{\FR 8.10 }
Joram: also called, Hadoram}\FTNT{**}{{\FR 8.10 }
salute…: Heb. ask him of peace}\FTNT{††}{{\FR 8.10 }
had wars…: Heb. was a man of wars with}\FTNT{‡‡}{{\FR 8.10 }
brought…: Heb. in his hand were}
\VS{11}Which also
king
David did
dedicate unto the
{\ND{
LORD}}, with the
silver and
gold that he had
dedicated of all
nations which he
subdued;
\VS{12}Of
Syria, and of
Moab, and of the
children of
Ammon, and of the
Philistines, and of
Amalek, and of the
spoil of
Hadadezer,
son of
Rehob,
king of
Zobah.
\VS{13}And
David
gat
{\ADD{him}} a
name when he
returned from
smiting of the
Syrians in the
valley of
salt,
{\ADD{being}}
eighteen
thousand
{\ADD{men}}.\FTNT{§§}{{\FR 8.13 }
smiting: Heb. his smiting}
\par }{\PP \VS{14}¶ And he
put
garrisons in
Edom; throughout all
Edom
put he
garrisons, and all they of
Edom became
David’s
servants. And the
{\ND{
LORD}}
preserved
David whithersoever he
went.
\par }{\PP \VS{15}And
David
reigned over all
Israel; and
David
executed
judgment and
justice unto all his
people.
\VS{16}And
Joab the
son of
Zeruiah
{\ADD{was}} over the
host; and
Jehoshaphat the
son of
Ahilud
{\ADD{was}}
recorder;\FTNT{***}{{\FR 8.16 }
recorder: or, remembrancer, or, writer of chronicles}
\VS{17}And
Zadok the
son of
Ahitub, and
Ahimelech the
son of
Abiathar,
{\ADD{were}} the
priests; and
Seraiah
{\ADD{was}} the
scribe;\FTNT{†††}{{\FR 8.17 }
scribe: or, secretary}
\VS{18}And
Benaiah the
son of
Jehoiada
{\ADD{was over}} both the
Cherethites and the
Pelethites; and
David’s
sons were chief
rulers.\FTNT{‡‡‡}{{\FR 8.18 }
chief rulers: or, princes}

\par }\Chap{9}{\PP \VerseOne{1}And
David
said, Is there
yet any that is
left of the
house of
Saul, that I may
shew him
kindness for
Jonathan’s sake?
\VS{2}And
{\ADD{there was}} of the
house of
Saul a
servant whose
name
{\ADD{was}}
Ziba. And when they had
called him unto
David, the
king
said unto him,
{\ADD{Art}} thou
Ziba? And he
said, Thy
servant
{\ADD{is he}}.
\VS{3}And the
king
said,
{\ADD{Is}} there
not yet
any of the
house of
Saul, that I may
shew the
kindness of
God unto him? And
Ziba
said unto the
king,
Jonathan hath yet a
son,
{\ADD{which is}}
lame on
{\ADD{his}}
feet.
\VS{4}And the
king
said unto him,
Where
{\ADD{is}} he? And
Ziba
said unto the
king, Behold, he
{\ADD{is}} in the
house of
Machir, the
son of
Ammiel, in
Lo-debar.
\par }{\PP \VS{5}¶ Then
king
David
sent, and
fetched him out of the
house of
Machir, the
son of
Ammiel, from
Lo-debar.
\VS{6}Now when
Mephibosheth, the
son of
Jonathan, the
son of
Saul, was
come unto
David, he
fell on his
face, and did
reverence. And
David
said,
Mephibosheth. And he
answered, Behold thy
servant!\FTNT{*}{{\FR 9.6 }
Mephibosheth: also called Merib-baal}
\par }{\PP \VS{7}¶ And
David
said unto him,
Fear not: for I will
surely
shew thee
kindness for
Jonathan thy
father’s sake, and will
restore thee all the
land of
Saul thy
father; and thou shalt
eat
bread at my
table
continually.
\VS{8}And he
bowed himself, and
said, What
{\ADD{is}} thy
servant, that thou shouldest
look upon such a
dead
dog as I
{\ADD{am}}?
\par }{\PP \VS{9}¶ Then the
king
called to
Ziba,
Saul’s
servant, and
said unto him, I have
given unto thy
master’s
son all that
pertained to
Saul and to all his
house.
\VS{10}Thou therefore, and thy
sons, and thy
servants, shall
till the
land for him, and thou shalt
bring in
{\ADD{the fruits}}, that thy
master’s
son may have
food to
eat: but
Mephibosheth thy
master’s
son shall
eat
bread
alway at my
table. Now
Ziba had
fifteen
sons and
twenty
servants.
\VS{11}Then
said
Ziba unto the
king, According to all that my
lord the
king hath
commanded his
servant, so shall thy
servant
do. As for
Mephibosheth,
{\ADD{said the king}}, he shall
eat at my
table, as
one of the
king’s
sons.
\VS{12}And
Mephibosheth had a
young
son, whose
name
{\ADD{was}}
Micha. And all that
dwelt in the
house of
Ziba
{\ADD{were}}
servants unto
Mephibosheth.
\VS{13}So
Mephibosheth
dwelt in
Jerusalem: for he did
eat
continually at the
king’s
table; and was
lame on
both his
feet.

\par }\Chap{10}{\PP \VerseOne{1}And it came to pass after
this, that the
king of the
children of
Ammon
died, and
Hanun his
son
reigned in his stead.
\VS{2}Then
said
David, I will
shew
kindness unto
Hanun the
son of
Nahash, as his
father
shewed
kindness unto me. And
David
sent to
comfort him by the
hand of his
servants for his
father. And
David’s
servants
came into the
land of the
children of
Ammon.
\VS{3}And the
princes of the
children of
Ammon
said unto
Hanun their
lord,
Thinkest thou that
David doth
honour thy
father, that he hath
sent
comforters unto thee? hath not
David
{\ADD{rather}}
sent his
servants unto thee,
to
search the
city, and to spy it
out, and to
overthrow it?\FTNT{*}{{\FR 10.3 }
Thinkest…: Heb. In thine eyes doth David}
\VS{4}Wherefore
Hanun
took
David’s
servants, and shaved
off the one
half of their
beards, and cut
off their
garments in the
middle,
{\ADD{even}} to their
buttocks, and sent them
away.
\VS{5}When they
told
{\ADD{it}} unto
David, he
sent to
meet them, because the
men were
greatly
ashamed: and the
king
said,
Tarry at
Jericho until your
beards be
grown, and
{\ADD{then}}
return.
\par }{\PP \VS{6}¶ And when the
children of
Ammon
saw that they
stank before
David, the
children of
Ammon
sent and
hired the
Syrians of
Beth-rehob, and the
Syrians of
Zoba,
twenty
thousand
footmen, and of
king
Maacah a
thousand
men, and of
Ish-tob
twelve
thousand
men.\FTNT{†}{{\FR 10.6 }
Ish-tob: or, the men of Tob}
\VS{7}And when
David
heard of
{\ADD{it}}, he
sent
Joab, and all the
host of the mighty
men.
\VS{8}And the
children of
Ammon came
out, and put the
battle in
array at the entering
in of the
gate: and the
Syrians of
Zoba, and of
Rehob, and
Ish-tob, and
Maacah,
{\ADD{were}} by
themselves in the
field.
\VS{9}When
Joab
saw that the
front of the
battle was against him
before and
behind, he
chose of all the
choice
{\ADD{men}} of
Israel, and put
{\ADD{them}} in
array
against the
Syrians:
\VS{10}And the
rest of the
people he
delivered into the
hand of
Abishai his
brother, that he might put
{\ADD{them}} in
array
against the
children of
Ammon.
\VS{11}And he
said, If the
Syrians be too
strong for me, then thou shalt
help me: but if the
children of
Ammon be too
strong for thee, then I will
come and
help thee.
\VS{12}Be of good
courage, and let us play the
men for our
people, and for the
cities of our
God: and the
{\ND{
LORD}}
do that which
seemeth him
good.
\VS{13}And
Joab drew
nigh, and the
people that
{\ADD{were}} with him, unto the
battle against the
Syrians: and they
fled
before him.
\VS{14}And when the
children of
Ammon
saw that the
Syrians were
fled, then
fled they also
before
Abishai, and
entered into the
city. So
Joab
returned from the
children of
Ammon, and
came to
Jerusalem.
\par }{\PP \VS{15}¶ And when the
Syrians
saw that they were
smitten
before
Israel, they
gathered themselves
together.
\VS{16}And
Hadarezer
sent, and brought
out the
Syrians that
{\ADD{were}}
beyond the
river: and they
came to
Helam; and
Shobach the
captain of the
host of
Hadarezer
{\ADD{went}}
before them.\FTNT{‡}{{\FR 10.16 }
the river: that is, Euphrates}\FTNT{§}{{\FR 10.16 }
Shobach: or, Shophach}
\VS{17}And when it was
told
David, he
gathered all
Israel
together, and passed
over
Jordan, and
came to
Helam. And the
Syrians set themselves in
array
against
David, and
fought with him.
\VS{18}And the
Syrians
fled
before
Israel; and
David
slew
{\ADD{the men of}}
seven
hundred
chariots of the
Syrians, and
forty
thousand
horsemen, and
smote
Shobach the
captain of their
host, who
died there.\FTNT{**}{{\FR 10.18 }
horsemen: also called, footmen}
\VS{19}And when all the
kings
{\ADD{that were}}
servants to
Hadarezer
saw that they were
smitten
before
Israel, they made
peace with
Israel, and
served them. So the
Syrians
feared to
help the
children of
Ammon any more.

\par }\Chap{11}{\PP \VerseOne{1}And it came to pass, after the
year was
expired, at the
time when
kings go
forth
{\ADD{to battle}}, that
David
sent
Joab, and his
servants with him, and all
Israel; and they
destroyed the
children of
Ammon, and
besieged
Rabbah. But
David tarried
still at
Jerusalem.\FTNT{*}{{\FR 11.1 }
after…: Heb. at the return of the year, or, in the spring}
\par }{\PP \VS{2}¶ And it came to pass in an
eveningtide, that
David
arose from off his
bed, and
walked upon the
roof of the
king’s
house: and from the
roof he
saw a
woman
washing herself; and the
woman
{\ADD{was}}
very
beautiful to look
upon.
\VS{3}And
David
sent and
enquired after the
woman. And
{\ADD{one}}
said,
{\ADD{Is}} not this
Bath-sheba, the
daughter of
Eliam, the
wife of
Uriah the
Hittite?\FTNT{†}{{\FR 11.3 }
Bath-sheba: or, Bath-shuah}\FTNT{‡}{{\FR 11.3 }
Eliam: or Ammiel}
\VS{4}And
David
sent
messengers, and
took her; and she came
in unto him, and he
lay with her; for she was
purified from her
uncleanness: and she
returned unto her
house.\FTNT{§}{{\FR 11.4 }
for she…: or, and when she had purified herself, etc, she returned}
\VS{5}And the
woman
conceived, and
sent and
told
David, and
said, I
{\ADD{am}} with
child.
\par }{\PP \VS{6}¶ And
David
sent to
Joab,
{\ADD{saying}},
Send me
Uriah the
Hittite. And
Joab
sent
Uriah to
David.
\VS{7}And when
Uriah was
come unto him,
David
demanded
{\ADD{of him}} how
Joab
did, and how the
people
did, and how the
war
prospered.\FTNT{**}{{\FR 11.7 }
how Joab…: Heb. of the peace of, etc}
\VS{8}And
David
said to
Uriah, Go
down to thy
house, and
wash thy
feet. And
Uriah departed
out of the
king’s
house, and there
followed him a
mess
{\ADD{of meat}} from the
king.\FTNT{††}{{\FR 11.8 }
followed…: Heb. went out after him}
\VS{9}But
Uriah
slept at the
door of the
king’s
house with all the
servants of his
lord, and went not
down to his
house.
\VS{10}And when they had
told
David,
saying,
Uriah went not
down unto his
house,
David
said unto
Uriah,
Camest thou not from
{\ADD{thy}}
journey? why
{\ADD{then}} didst thou not go
down unto thine
house?
\VS{11}And
Uriah
said unto
David, The
ark, and
Israel, and
Judah,
abide in
tents; and my
lord
Joab, and the
servants of my
lord, are
encamped in the
open
fields; shall
I then
go into mine
house, to
eat and to
drink, and to
lie with my
wife?
{\ADD{as}} thou
livest, and
{\ADD{as}} thy
soul
liveth, I will not
do this
thing.
\VS{12}And
David
said to
Uriah,
Tarry here to
day also, and to
morrow I will let thee
depart. So
Uriah
abode in
Jerusalem that
day, and the
morrow.
\VS{13}And when
David had
called him, he did
eat and
drink
before him; and he made him
drunk: and at
even he went
out to
lie on his
bed with the
servants of his
lord, but went not
down to his
house.
\par }{\PP \VS{14}¶ And it came to pass in the
morning, that
David
wrote a
letter to
Joab, and
sent
{\ADD{it}} by the
hand of
Uriah.
\VS{15}And he
wrote in the
letter,
saying,
Set ye
Uriah in the
forefront of the
hottest
battle, and
retire ye from
him, that he may be
smitten, and
die.\FTNT{‡‡}{{\FR 11.15 }
hottest: Heb. strong}\FTNT{§§}{{\FR 11.15 }
from him: Heb. from after him}
\VS{16}And it came to pass, when
Joab
observed the
city, that he
assigned
Uriah unto a
place where he
knew that
valiant
men
{\ADD{were}}.
\VS{17}And the
men of the
city went
out, and
fought with
Joab: and there
fell
{\ADD{some}} of the
people of the
servants of
David; and
Uriah the
Hittite
died also.
\par }{\PP \VS{18}¶ Then
Joab
sent and
told
David all the
things concerning the
war;
\VS{19}And
charged the
messenger,
saying, When thou hast made an
end of
telling the
matters of the
war unto the
king,
\VS{20}And if so be that the
king’s
wrath
arise, and he
say unto thee, Wherefore approached ye so
nigh unto the
city when ye did
fight?
knew ye not that they would
shoot from the
wall?
\VS{21}Who
smote
Abimelech the
son of
Jerubbesheth? did not a
woman
cast a
piece of a
millstone upon him from the
wall, that he
died in
Thebez? why went ye
nigh the
wall? then
say thou, Thy
servant
Uriah the
Hittite is
dead also.\FTNT{***}{{\FR 11.21 }
Jerubbesheth: also called, Jerubbaal}
\par }{\PP \VS{22}¶ So the
messenger
went, and
came and
shewed
David all that
Joab had
sent him for.
\VS{23}And the
messenger
said unto
David, Surely the
men
prevailed against us, and came
out unto us into the
field, and we were upon them even unto the
entering of the
gate.
\VS{24}And the
shooters
shot from off the
wall upon thy
servants; and
{\ADD{some}} of the
king’s
servants be
dead, and thy
servant
Uriah the
Hittite is
dead also.
\VS{25}Then
David
said unto the
messenger, Thus shalt thou
say unto
Joab, Let not this
thing
displease thee, for the
sword
devoureth one as well as
another: make thy
battle more
strong against the
city, and
overthrow it: and
encourage thou him.\FTNT{†††}{{\FR 11.25 }
displease…: Heb. be evil in thine eyes}\FTNT{‡‡‡}{{\FR 11.25 }
one…: Heb. so and such}
\par }{\PP \VS{26}¶ And when the
wife of
Uriah
heard that
Uriah her
husband was
dead, she
mourned for her
husband.
\VS{27}And when the
mourning was
past,
David
sent and
fetched her to his
house, and she became his
wife, and
bare him a
son. But the
thing that
David had
done
displeased the
{\ND{
LORD}}.\FTNT{§§§}{{\FR 11.27 }
displeased: Heb. was evil in the eyes of}

\par }\Chap{12}{\PP \VerseOne{1}And the
{\ND{
LORD}}
sent
Nathan unto
David. And he
came unto him, and
said unto him, There were
two
men in
one
city; the
one
rich, and the
other
poor.
\VS{2}The
rich
{\ADD{man}} had
exceeding
many
flocks and
herds:
\VS{3}But the
poor
{\ADD{man}} had nothing, save
one
little ewe
lamb, which he had
bought and nourished
up: and it grew
up
together with him, and with his
children; it did
eat of his own
meat, and
drank of his own
cup, and
lay in his
bosom, and was unto him as a
daughter.\FTNT{*}{{\FR 12.3 }
meat: Heb. morsel}
\VS{4}And there
came a
traveller unto the
rich
man, and he
spared to
take of his own
flock and of his own
herd, to
dress for the wayfaring
man that was
come unto him; but
took the
poor
man’s
lamb, and
dressed it for the
man that was
come to him.
\VS{5}And
David’s
anger was
greatly
kindled against the
man; and he
said to
Nathan,
{\ADD{As}} the
{\ND{
LORD}}
liveth, the
man that hath
done this
{\ADD{thing}} shall
surely
die:\FTNT{†}{{\FR 12.5 }
shall…: or, is worthy to die, or, is a son of death}
\VS{6}And he shall
restore the
lamb
fourfold,
because he
did this
thing, and
because he had no
pity.
\par }{\PP \VS{7}¶ And
Nathan
said to
David, Thou
{\ADD{art}} the
man. Thus
saith the
{\ND{
LORD}}
God of
Israel, I
anointed thee
king over
Israel, and I
delivered thee out of the
hand of
Saul;
\VS{8}And I
gave thee thy
master’s
house, and thy
master’s
wives into thy
bosom, and
gave thee the
house of
Israel and of
Judah; and if
{\ADD{that had been}} too
little, I would moreover have
given unto thee
such and such
things.
\VS{9}Wherefore hast thou
despised the
commandment of the
{\ND{
LORD}}, to
do
evil in his
sight? thou hast
killed
Uriah the
Hittite with the
sword, and hast
taken his
wife
{\ADD{to be}} thy
wife, and hast
slain him with the
sword of the
children of
Ammon.
\VS{10}Now therefore the
sword shall
never
depart from thine
house;
because thou hast
despised me, and hast
taken the
wife of
Uriah the
Hittite to be thy
wife.
\VS{11}Thus
saith the
{\ND{
LORD}}, Behold, I will raise
up
evil against thee out of thine own
house, and I will
take thy
wives before thine
eyes, and
give
{\ADD{them}} unto thy
neighbour, and he shall
lie with thy
wives in the
sight of this
sun.
\VS{12}For thou
didst
{\ADD{it}}
secretly: but I will
do this
thing before all
Israel, and before the
sun.
\VS{13}And
David
said unto
Nathan, I have
sinned against the
{\ND{
LORD}}. And
Nathan
said unto
David, The
{\ND{
LORD}} also hath put
away thy
sin; thou shalt not
die.
\VS{14}Howbeit, because by this
deed thou hast
given great
occasion to the
enemies of the
{\ND{
LORD}} to
blaspheme, the
child also
{\ADD{that is}}
born unto thee shall
surely
die.
\par }{\PP \VS{15}¶ And
Nathan
departed unto his
house. And the
{\ND{
LORD}}
struck the
child that
Uriah’s
wife
bare unto
David, and it was very
sick.
\VS{16}David therefore
besought
God for the
child; and
David
fasted, and went
in, and
lay all
night upon the
earth.\FTNT{‡}{{\FR 12.16 }
fasted: Heb. fasted a fast}
\VS{17}And the
elders of his
house
arose,
{\ADD{and went}} to him, to raise him
up from the
earth: but he
would not, neither did he
eat
bread with them.
\VS{18}And it came to pass on the
seventh
day, that the
child
died. And the
servants of
David
feared to
tell him that the
child was
dead: for they
said, Behold, while the
child was yet
alive, we
spake unto him, and he would not
hearken unto our
voice: how will he
then
vex
himself, if we
tell him that the
child is
dead?\FTNT{§}{{\FR 12.18 }
vex: Heb. do hurt}
\VS{19}But when
David
saw that his
servants
whispered,
David
perceived that the
child was
dead: therefore
David
said unto his
servants, Is the
child
dead? And they
said, He is
dead.
\VS{20}Then
David
arose from the
earth, and
washed, and
anointed
{\ADD{himself}}, and
changed his
apparel, and
came into the
house of the
{\ND{
LORD}}, and
worshipped: then he
came to his own
house; and when he
required, they
set
bread before him, and he did
eat.
\VS{21}Then
said his
servants unto him, What
thing
{\ADD{is}} this that thou hast
done? thou didst
fast and
weep for the
child,
{\ADD{while it was}}
alive; but when the
child was
dead, thou didst
rise and
eat
bread.
\VS{22}And he
said, While the
child was yet
alive, I
fasted and
wept: for I
said, Who can
tell
{\ADD{whether}}
{\ND{
GOD}} will be
gracious to me, that the
child may
live?
\VS{23}But now he is
dead, wherefore should I
fast?
can I bring him back
again? I shall
go to him, but he shall not
return to me.
\par }{\PP \VS{24}¶ And
David
comforted
Bath-sheba his
wife, and went
in unto her, and
lay with her: and she
bare a
son, and he
called his
name
Solomon: and the
{\ND{
LORD}}
loved him.
\VS{25}And he
sent by the
hand of
Nathan the
prophet; and he
called his
name
Jedidiah, because of the
{\ND{
LORD}}.\FTNT{**}{{\FR 12.25 }
Jedidiah: that is, Beloved of the
{\ND{LORD}}}
\par }{\PP \VS{26}¶ And
Joab
fought against
Rabbah of the
children of
Ammon, and
took the
royal
city.
\VS{27}And
Joab
sent
messengers to
David, and
said, I have
fought against
Rabbah, and have
taken the
city of
waters.
\VS{28}Now therefore
gather the
rest of the
people
together, and
encamp against the
city, and
take it: lest I
take the
city, and it be
called after my
name.\FTNT{††}{{\FR 12.28 }
it be called…: Heb. my name be called upon it}
\VS{29}And
David
gathered all the
people
together, and
went to
Rabbah, and
fought against it, and
took it.
\VS{30}And he
took their
king’s
crown from off his
head, the
weight whereof
{\ADD{was}} a
talent of
gold with the
precious
stones: and it was
{\ADD{set}} on
David’s
head. And he brought
forth the
spoil of the
city in
great
abundance.\FTNT{‡‡}{{\FR 12.30 }
in great…: Heb. very great}
\VS{31}And he brought
forth the
people that
{\ADD{were}} therein, and
put
{\ADD{them}} under
saws, and under
harrows of
iron, and under
axes of
iron, and made them
pass through the
brickkiln: and thus
did he unto all the
cities of the
children of
Ammon. So
David and all the
people
returned unto
Jerusalem.

\par }\Chap{13}{\PP \VerseOne{1}And it came to pass
after this, that
Absalom the
son of
David had a
fair
sister, whose
name
{\ADD{was}}
Tamar; and
Amnon the
son of
David
loved her.
\VS{2}And
Amnon was so
vexed, that he fell
sick for his
sister
Tamar; for she
{\ADD{was}} a
virgin; and
Amnon
thought it
hard for him to
do any
thing to her.\FTNT{*}{{\FR 13.2 }
Amnon thought…: Heb. it was marvellous, or, hidden in the eyes of Amnon}
\VS{3}But
Amnon had a
friend, whose
name
{\ADD{was}}
Jonadab, the
son of
Shimeah
David’s
brother: and
Jonadab
{\ADD{was}} a
very
subtil
man.
\VS{4}And he
said unto him, Why
{\ADD{art}} thou,
{\ADD{being}} the
king’s
son,
lean from
day to
day? wilt thou not
tell me? And
Amnon
said unto him, I
love
Tamar, my
brother
Absalom’s
sister.\FTNT{†}{{\FR 13.4 }
lean: Heb. thin}\FTNT{‡}{{\FR 13.4 }
from day…: Heb. morning by morning}
\VS{5}And
Jonadab
said unto him, Lay thee
down on thy
bed, and make thyself
sick: and when thy
father
cometh to
see thee,
say unto him, I pray thee, let my
sister
Tamar
come, and
give me
meat, and
dress the
meat in my
sight, that I may
see
{\ADD{it}}, and
eat
{\ADD{it}} at her
hand.
\par }{\PP \VS{6}¶ So
Amnon lay
down, and made himself
sick: and when the
king was
come to
see him,
Amnon
said unto the
king, I pray thee, let
Tamar my
sister
come, and
make me a
couple of
cakes in my
sight, that I may
eat at her
hand.
\VS{7}Then
David
sent
home to
Tamar,
saying,
Go now to thy
brother
Amnon’s
house, and
dress him
meat.
\VS{8}So
Tamar
went to her
brother
Amnon’s
house; and he was laid
down. And she
took
flour, and
kneaded
{\ADD{it}}, and made
cakes in his
sight, and did
bake the
cakes.\FTNT{§}{{\FR 13.8 }
flour: or, paste}
\VS{9}And she
took a
pan, and poured
{\ADD{them}}
out
before him; but he
refused to
eat. And
Amnon
said, Have
out all
men from me. And they went
out every
man from him.
\VS{10}And
Amnon
said unto
Tamar,
Bring the
meat into the
chamber, that I may
eat of thine
hand. And
Tamar
took the
cakes which she had
made, and
brought
{\ADD{them}} into the
chamber to
Amnon her
brother.
\VS{11}And when she had
brought
{\ADD{them}} unto him to
eat, he took
hold of her, and
said unto her,
Come
lie with me, my
sister.
\VS{12}And she
answered him, Nay, my
brother, do not
force me; for no such thing
ought to be
done in
Israel:
do not thou this
folly.\FTNT{**}{{\FR 13.12 }
force…: Heb. humble me}\FTNT{††}{{\FR 13.12 }
no such…: Heb. it ought not so to be done}
\VS{13}And I, whither shall I cause my
shame to
go? and as for thee, thou shalt be as
one of the
fools in
Israel. Now therefore, I pray thee,
speak unto the
king; for he will not
withhold me from thee.
\VS{14}Howbeit he
would not
hearken unto her
voice: but, being
stronger than she,
forced her, and
lay with her.
\par }{\PP \VS{15}¶ Then
Amnon
hated her
exceedingly; so that the
hatred wherewith he
hated her
{\ADD{was}}
greater than the
love wherewith he had
loved her. And
Amnon
said unto her,
Arise, be
gone.\FTNT{‡‡}{{\FR 13.15 }
exceedingly: Heb. with great hatred greatly}
\VS{16}And she
said unto him,
{\ADD{There is}} no
cause: this
evil in sending me
away
{\ADD{is}}
greater than the
other that thou
didst unto me. But he
would not
hearken unto her.
\VS{17}Then he
called his
servant that
ministered unto him, and
said,
Put now this
{\ADD{woman}}
out from me, and
bolt the
door
after her.
\VS{18}And
{\ADD{she had}} a
garment of divers
colours upon her: for with such
robes were the
king’s
daughters
{\ADD{that were}}
virgins
apparelled. Then his
servant
brought her
out, and
bolted the
door
after her.
\par }{\PP \VS{19}¶ And
Tamar
put
ashes on her
head, and
rent her
garment of divers
colours that
{\ADD{was}} on her, and
laid her
hand on her
head, and
went
on
crying.
\VS{20}And
Absalom her
brother
said unto her, Hath
Amnon thy
brother been with thee? but hold now thy
peace, my
sister: he
{\ADD{is}} thy
brother;
regard not this
thing. So
Tamar
remained
desolate in her
brother
Absalom’s
house.\FTNT{§§}{{\FR 13.20 }
Amnon: Heb. Aminon}\FTNT{***}{{\FR 13.20 }
regard…: Heb. set not thy heart}\FTNT{†††}{{\FR 13.20 }
desolate: Heb. and desolate}
\par }{\PP \VS{21}¶ But when
king
David
heard of all these
things, he was
very
wroth.
\VS{22}And
Absalom
spake unto his brother
Amnon neither
good nor
bad: for
Absalom
hated
Amnon,
because he had
forced his
sister
Tamar.
\par }{\PP \VS{23}¶ And it came to pass after two
full
years, that
Absalom had
sheepshearers in
Baal-hazor, which
{\ADD{is}} beside
Ephraim: and
Absalom
invited all the
king’s
sons.
\VS{24}And
Absalom
came to the
king, and
said, Behold now, thy
servant hath
sheepshearers;
let the
king, I beseech thee, and his
servants
go with thy
servant.
\VS{25}And the
king
said to
Absalom, Nay, my
son, let us not all now
go, lest we be
chargeable unto thee. And he
pressed him: howbeit he
would not
go, but
blessed him.
\VS{26}Then
said
Absalom, If not, I pray thee, let my
brother
Amnon
go with us. And the
king
said unto him, Why should he
go with thee?
\VS{27}But
Absalom
pressed him, that he let
Amnon and all the
king’s
sons
go with him.
\par }{\PP \VS{28}¶ Now
Absalom had
commanded his
servants,
saying,
Mark ye now when
Amnon’s
heart is
merry with
wine, and when I
say unto you,
Smite
Amnon; then
kill him,
fear not: have not I
commanded you? be
courageous, and be
valiant.\FTNT{‡‡‡}{{\FR 13.28 }
have…: or, will you not, since I have commanded you?}\FTNT{§§§}{{\FR 13.28 }
valiant: Heb. sons of valour}
\VS{29}And the
servants of
Absalom
did unto
Amnon as
Absalom had
commanded. Then all the
king’s
sons
arose, and every
man gat him
up upon his
mule, and
fled.\FTNT{*}{{\FR 13.29 }
gat…: Heb. rode}
\par }{\PP \VS{30}¶ And it came to pass, while they were in the
way, that
tidings
came to
David,
saying,
Absalom hath
slain all the
king’s
sons, and there is not
one of them
left.
\VS{31}Then the
king
arose, and
tare his
garments, and
lay on the
earth; and all his
servants
stood by with their
clothes
rent.
\VS{32}And
Jonadab, the
son of
Shimeah
David’s
brother,
answered and
said, Let not my
lord
suppose
{\ADD{that}} they have
slain all the young
men the
king’s
sons; for
Amnon only is
dead: for by the
appointment of
Absalom this hath been
determined from the
day that he
forced his
sister
Tamar.\FTNT{†}{{\FR 13.32 }
appointment: Heb. mouth}\FTNT{‡}{{\FR 13.32 }
determined: or, settled}
\VS{33}Now therefore let not my
lord the
king
take the
thing to his
heart, to
think that all the
king’s
sons are
dead: for
Amnon only is
dead.
\VS{34}But
Absalom
fled. And the young
man that kept the
watch lifted
up his
eyes, and
looked, and, behold, there
came
much
people by the
way of the
hill
side
behind him.
\VS{35}And
Jonadab
said unto the
king, Behold, the
king’s
sons
come: as thy
servant
said, so it is.\FTNT{§}{{\FR 13.35 }
as thy…: Heb. according to the word of thy servant}
\VS{36}And it came to pass, as soon as he had made an
end of
speaking, that, behold, the
king’s
sons
came, and lifted
up their
voice and
wept: and the
king also and all his
servants
wept
very
sore.\FTNT{**}{{\FR 13.36 }
very…: Heb. with a great weeping greatly}
\par }{\PP \VS{37}¶ But
Absalom
fled, and
went to
Talmai, the
son of
Ammihud,
king of
Geshur. And
{\ADD{David}}
mourned for his
son every
day.\FTNT{††}{{\FR 13.37 }
Ammihud: or, Ammihur}
\VS{38}So
Absalom
fled, and
went to
Geshur, and was there
three
years.
\VS{39}And
{\ADD{the soul of}}
king
David
longed to go
forth unto
Absalom: for he was
comforted concerning
Amnon, seeing he was
dead.\FTNT{‡‡}{{\FR 13.39 }
longed: or, was consumed}

\par }\Chap{14}{\PP \VerseOne{1}Now
Joab the
son of
Zeruiah
perceived that the
king’s
heart
{\ADD{was}} toward
Absalom.
\VS{2}And
Joab
sent to
Tekoah, and
fetched thence a
wise
woman, and
said unto her, I pray thee, feign thyself to be a
mourner, and put
on now
mourning
apparel, and
anoint not thyself with
oil, but be as a
woman that had a
long
time
mourned for the
dead:
\VS{3}And
come to the
king, and
speak on this
manner unto him. So
Joab
put the
words in her
mouth.
\par }{\PP \VS{4}¶ And when the
woman of
Tekoah
spake to the
king, she
fell on her
face to the
ground, and did
obeisance, and
said,
Help, O
king.\FTNT{*}{{\FR 14.4 }
Help: Heb. Save}
\VS{5}And the
king
said unto her, What aileth thee? And she
answered, I
{\ADD{am}}
indeed a
widow
woman, and mine
husband is
dead.
\VS{6}And thy
handmaid had
two
sons, and they
two strove
together in the
field, and
{\ADD{there was}} none to
part them, but the
one
smote the
other, and
slew him.\FTNT{†}{{\FR 14.6 }
none…: Heb. no deliverer between them}
\VS{7}And, behold, the whole
family is
risen against thine
handmaid, and they
said,
Deliver him that
smote his
brother, that we may
kill him, for the
life of his
brother whom he
slew; and we will
destroy the
heir also: and so they shall
quench my
coal which is
left, and shall not
leave to my
husband
{\ADD{neither}}
name nor
remainder
upon the
earth.\FTNT{‡}{{\FR 14.7 }
upon…: Heb. upon the face of the earth}
\VS{8}And the
king
said unto the
woman,
Go to thine
house, and I will give
charge concerning thee.
\VS{9}And the
woman of
Tekoah
said unto the
king, My
lord, O
king, the
iniquity
{\ADD{be}} on me, and on my
father’s
house: and the
king and his
throne
{\ADD{be}}
guiltless.
\VS{10}And the
king
said, Whosoever
saith
{\ADD{ought}} unto thee,
bring him to me, and he shall not
touch thee any
more.
\VS{11}Then
said she, I pray thee, let the
king
remember the
{\ND{
LORD}} thy
God, that thou wouldest not suffer the
revengers of
blood to
destroy any
more, lest they
destroy my
son. And he
said,
{\ADD{As}} the
{\ND{
LORD}}
liveth, there shall not one
hair of thy
son
fall to the
earth.\FTNT{§}{{\FR 14.11 }
that…: Heb. that the revenger of blood do not multiply to destroy}
\VS{12}Then the
woman
said, Let thine
handmaid, I pray thee,
speak
{\ADD{one}}
word unto my
lord the
king. And he
said, Say
on.
\VS{13}And the
woman
said, Wherefore then hast thou
thought such a thing against the
people of
God? for the
king doth
speak this
thing as one which is
faulty, in that the
king doth not fetch home
again his
banished.
\VS{14}For we must
needs
die, and
{\ADD{are}} as
water
spilt on the
ground, which cannot be gathered up
again; neither doth
God
respect
{\ADD{any}}
person: yet doth he
devise
means, that his
banished be not
expelled from him.\FTNT{**}{{\FR 14.14 }
neither…: or, because God hath not taken away his life, he hath also devised means, etc}
\VS{15}Now therefore that I am
come to
speak of this
thing unto my
lord the
king,
{\ADD{it is}} because the
people have made me
afraid: and thy
handmaid
said, I will now
speak unto the
king; it may be that the
king will
perform the
request of his
handmaid.
\VS{16}For the
king will
hear, to
deliver his
handmaid out of the
hand of the
man
{\ADD{that would}}
destroy me and my
son
together out of the
inheritance of
God.
\VS{17}Then thine
handmaid
said, The
word of my
lord the
king shall now be
comfortable: for as an
angel of
God, so
{\ADD{is}} my
lord the
king to
discern
good and
bad: therefore the
{\ND{
LORD}} thy
God will be with thee.\FTNT{††}{{\FR 14.17 }
comfortable: Heb. for rest}\FTNT{‡‡}{{\FR 14.17 }
to discern: Heb. to hear}
\VS{18}Then the
king
answered and
said unto the
woman,
Hide not from me, I pray thee, the
thing that I shall
ask thee. And the
woman
said, Let my
lord the
king now
speak.
\VS{19}And the
king
said,
{\ADD{Is not}} the
hand of
Joab with thee in all this? And the
woman
answered and
said,
{\ADD{As}} thy
soul
liveth, my
lord the
king,
none
can turn to the right
hand or to the
left from ought that my
lord the
king hath
spoken: for thy
servant
Joab, he
bade me, and he
put all these
words in the
mouth of thine
handmaid:
\VS{20}To fetch
about this
form of
speech hath thy
servant
Joab
done this
thing: and my
lord
{\ADD{is}}
wise, according to the
wisdom of an
angel of
God, to
know all
{\ADD{things}} that
{\ADD{are}} in the
earth.
\par }{\PP \VS{21}¶ And the
king
said unto
Joab, Behold now, I have
done this
thing:
go therefore,
bring the young
man
Absalom
again.
\VS{22}And
Joab
fell to the
ground on his
face, and
bowed himself, and
thanked the
king: and
Joab
said, To
day thy
servant
knoweth that I have
found
grace in thy
sight, my
lord, O
king, in that the
king hath
fulfilled the
request of his
servant.\FTNT{§§}{{\FR 14.22 }
thanked: Heb. blessed}\FTNT{***}{{\FR 14.22 }
his servant: or, thy servant}
\VS{23}So
Joab
arose and
went to
Geshur, and
brought
Absalom to
Jerusalem.
\VS{24}And the
king
said, Let him
turn to his own
house, and let him not
see my
face. So
Absalom
returned to his own
house, and
saw not the
king’s
face.
\par }{\PP \VS{25}¶ But in all
Israel there was
none to be so
much
praised as
Absalom for his
beauty: from the
sole of his
foot even to the crown of his
head there was no
blemish in him.\FTNT{†††}{{\FR 14.25 }
But…: Heb. And as Absalom there was not a beautiful man in all Israel to praise greatly}
\VS{26}And when he
polled his
head, (for it was at every
year’s
end that he
polled
{\ADD{it:}} because
{\ADD{the hair}} was
heavy on him, therefore he
polled it:) he
weighed the
hair of his
head at two
hundred
shekels after the
king’s
weight.
\VS{27}And unto
Absalom there were
born
three
sons, and
one
daughter, whose
name
{\ADD{was}}
Tamar: she was a
woman of a
fair
countenance.
\par }{\PP \VS{28}¶ So
Absalom
dwelt two
full
years in
Jerusalem, and
saw not the
king’s
face.
\VS{29}Therefore
Absalom
sent for
Joab, to have
sent him to the
king; but he
would not
come to him: and when he sent
again the second
time, he
would not
come.
\VS{30}Therefore he
said unto his
servants,
See,
Joab’s
field is
near
mine, and he hath
barley there;
go and
set it on
fire. And
Absalom’s
servants
set the
field on
fire.\FTNT{‡‡‡}{{\FR 14.30 }
near…: Heb. near my place}
\VS{31}Then
Joab
arose, and
came to
Absalom unto
{\ADD{his}}
house, and
said unto him, Wherefore have thy
servants
set my
field on
fire?
\VS{32}And
Absalom
answered
Joab, Behold, I
sent unto thee,
saying,
Come hither, that I may
send thee to the
king, to
say, Wherefore am I
come from
Geshur?
{\ADD{it had been}}
good for me
{\ADD{to have been}} there still: now therefore let me
see the
king’s
face; and if there
be
{\ADD{any}}
iniquity in me, let him
kill me.
\VS{33}So
Joab
came to the
king, and
told him: and when he had
called for
Absalom, he
came to the
king, and
bowed himself on his
face to the
ground
before the
king: and the
king
kissed
Absalom.

\par }\Chap{15}{\PP \VerseOne{1}And it came to pass after
this, that
Absalom
prepared him
chariots and
horses, and
fifty
men to
run
before him.
\VS{2}And
Absalom rose up
early, and
stood
beside the
way of the
gate: and it was
{\ADD{so}}, that when any
man that had a
controversy
came to the
king for
judgment, then
Absalom
called unto him, and
said, Of
what
city
{\ADD{art}} thou? And he
said, Thy
servant
{\ADD{is}} of
one of the
tribes of
Israel.\FTNT{*}{{\FR 15.2 }
came: Heb. to come}
\VS{3}And
Absalom
said unto him,
See, thy
matters
{\ADD{are}}
good and
right; but
{\ADD{there is}} no man
{\ADD{deputed}} of the
king to
hear thee.\FTNT{†}{{\FR 15.3 }
there…: or, none will hear thee from the king downward}
\VS{4}Absalom
said moreover, Oh that I were
made
judge in the
land, that every
man which hath any
suit or
cause might
come unto me, and I would do him
justice!
\VS{5}And it was
{\ADD{so}}, that when any
man came
nigh
{\ADD{to him}} to do him
obeisance, he put
forth his
hand, and
took him, and
kissed him.
\VS{6}And on this
manner
did
Absalom to all
Israel that
came to the
king for
judgment: so
Absalom
stole the
hearts of the
men of
Israel.
\par }{\PP \VS{7}¶ And it came to pass
after
forty
years, that
Absalom
said unto the
king, I pray thee, let me
go and
pay my
vow, which I have
vowed unto the
{\ND{
LORD}}, in
Hebron.
\VS{8}For thy
servant
vowed a
vow while I
abode at
Geshur in
Syria,
saying, If the
{\ND{
LORD}} shall bring me
again indeed to
Jerusalem, then I will
serve the
{\ND{
LORD}}.
\VS{9}And the
king
said unto him,
Go in
peace. So he
arose, and
went to
Hebron.
\par }{\PP \VS{10}¶ But
Absalom
sent
spies throughout all the
tribes of
Israel,
saying, As soon as ye
hear the
sound of the
trumpet, then ye shall
say,
Absalom
reigneth in
Hebron.
\VS{11}And with
Absalom
went two
hundred
men out of
Jerusalem,
{\ADD{that were}}
called; and they
went in their
simplicity, and they
knew not any
thing.
\VS{12}And
Absalom
sent for
Ahithophel the
Gilonite,
David’s
counsellor, from his
city,
{\ADD{even}} from
Giloh, while he
offered
sacrifices. And the
conspiracy was
strong; for the
people
increased
continually with
Absalom.
\par }{\PP \VS{13}¶ And there
came a
messenger to
David,
saying, The
hearts of the
men of
Israel are
after
Absalom.
\VS{14}And
David
said unto all his
servants that
{\ADD{were}} with him at
Jerusalem,
Arise, and let us
flee; for we shall not
{\ADD{else}}
escape
from
Absalom: make
speed to
depart, lest he
overtake us
suddenly, and
bring
evil upon us, and
smite the
city with the
edge of the
sword.\FTNT{‡}{{\FR 15.14 }
bring: Heb. thrust}
\VS{15}And the
king’s
servants
said unto the
king, Behold, thy
servants
{\ADD{are ready to do}} whatsoever my
lord the
king shall
appoint.\FTNT{§}{{\FR 15.15 }
appoint: Heb. choose}
\VS{16}And the
king went
forth, and all his
household
after him. And the
king
left
ten
women,
{\ADD{which were}}
concubines, to
keep the
house.\FTNT{**}{{\FR 15.16 }
after…: Heb. at his feet}
\VS{17}And the
king went
forth, and all the
people
after him, and
tarried in a
place that was far
off.
\VS{18}And all his
servants passed
on
beside him; and all the
Cherethites, and all the
Pelethites, and all the
Gittites,
six
hundred
men which
came
after him from
Gath, passed
on
before the
king.
\par }{\PP \VS{19}¶ Then
said the
king to
Ittai the
Gittite, Wherefore
goest thou also with us?
return to thy
place, and
abide with the
king: for thou
{\ADD{art}} a
stranger, and also an
exile.
\VS{20}Whereas thou
camest
{\ADD{but}}
yesterday, should I this
day make thee
go up and
down with us? seeing I
go whither I may,
return thou, and take
back thy
brethren:
mercy and
truth
{\ADD{be}} with thee.\FTNT{††}{{\FR 15.20 }
go up…: Heb. wander in going}
\VS{21}And
Ittai
answered the
king, and
said,
{\ADD{As}} the
{\ND{
LORD}}
liveth, and
{\ADD{as}} my
lord the
king
liveth, surely in what
place my
lord the
king shall be, whether in
death or
life, even there also will thy
servant be.
\VS{22}And
David
said to
Ittai,
Go and pass
over. And
Ittai the
Gittite passed
over, and all his
men, and all the little
ones that
{\ADD{were}} with him.
\VS{23}And all the
country
wept with a
loud
voice, and all the
people passed
over: the
king also himself passed
over the
brook
Kidron, and all the
people passed
over,
toward the
way of the
wilderness.\FTNT{‡‡}{{\FR 15.23 }
Kidron: Gr. Cedron}
\par }{\PP \VS{24}¶ And lo
Zadok also, and all the
Levites
{\ADD{were}} with him,
bearing the
ark of the
covenant of
God: and they set
down the
ark of
God; and
Abiathar went
up, until all the
people had
done
passing out of the
city.
\VS{25}And the
king
said unto
Zadok, Carry
back the
ark of
God into the
city: if I shall
find
favour in the
eyes of the
{\ND{
LORD}}, he will bring me
again, and
shew me
{\ADD{both}} it, and his
habitation:
\VS{26}But if he thus
say, I have no
delight in thee; behold,
{\ADD{here am}} I, let him
do to me as
seemeth
good unto him.
\VS{27}The
king
said also unto
Zadok the
priest,
{\ADD{Art not}} thou a
seer?
return into the
city in
peace, and your
two
sons with you,
Ahimaaz thy
son, and
Jonathan the
son of
Abiathar.
\VS{28}See, I will
tarry in the
plain of the
wilderness, until there
come
word from you to
certify me.
\VS{29}Zadok therefore and
Abiathar
carried the
ark of
God
again to
Jerusalem: and they
tarried there.
\par }{\PP \VS{30}¶ And
David went
up by the
ascent of
{\ADD{mount}}
Olivet, and
wept as he went
up, and had his
head
covered, and he
went
barefoot: and all the
people that
{\ADD{was}} with him
covered every
man his
head, and they went
up,
weeping as they went
up.\FTNT{§§}{{\FR 15.30 }
and wept…: Heb. going up and weeping}
\par }{\PP \VS{31}¶ And
{\ADD{one}}
told
David,
saying,
Ahithophel
{\ADD{is}} among the
conspirators with
Absalom. And
David
said, O
{\ND{
LORD}}, I pray thee, turn the
counsel of
Ahithophel into
foolishness.
\par }{\PP \VS{32}¶ And it came to pass, that
{\ADD{when}}
David was
come to the
top
{\ADD{of the mount}}, where he
worshipped
God, behold,
Hushai the
Archite came to
meet him with his
coat
rent, and
earth upon his
head:
\VS{33}Unto whom
David
said, If thou passest
on with me, then thou shalt be a
burden unto me:
\VS{34}But if thou
return to the
city, and
say unto
Absalom, I will be thy
servant, O
king;
{\ADD{as}} I
{\ADD{have been}} thy
father’s
servant
hitherto, so
{\ADD{will}} I now also
{\ADD{be}} thy
servant: then mayest thou for me
defeat the
counsel of
Ahithophel.
\VS{35}And
{\ADD{hast thou}} not there with thee
Zadok and
Abiathar the
priests? therefore it shall be,
{\ADD{that}} what
thing soever thou shalt
hear out of the
king’s
house, thou shalt
tell
{\ADD{it}} to
Zadok and
Abiathar the
priests.
\VS{36}Behold,
{\ADD{they have}} there with them their
two
sons,
Ahimaaz
Zadok’s
{\ADD{son}}, and
Jonathan
Abiathar’s
{\ADD{son;}} and by
them ye shall
send unto me every
thing that ye can
hear.
\VS{37}So
Hushai
David’s
friend
came into the
city, and
Absalom
came into
Jerusalem.

\par }\Chap{16}{\PP \VerseOne{1}And when
David was a
little
past the
top
{\ADD{of the hill}}, behold,
Ziba the
servant of
Mephibosheth
met him, with a
couple of
asses
saddled, and upon them two
hundred
{\ADD{loaves}} of
bread, and an
hundred bunches of
raisins, and an
hundred of summer
fruits, and a
bottle of
wine.
\VS{2}And the
king
said unto
Ziba, What meanest thou by these? And
Ziba
said, The
asses
{\ADD{be}} for the
king’s
household to ride
on; and the
bread and summer
fruit for the young
men to
eat; and the
wine, that such as be
faint in the
wilderness may
drink.
\VS{3}And the
king
said, And where
{\ADD{is}} thy
master’s
son? And
Ziba
said unto the
king, Behold, he
abideth at
Jerusalem: for he
said, To
day shall the
house of
Israel
restore me the
kingdom of my
father.
\VS{4}Then
said the
king to
Ziba, Behold, thine
{\ADD{are}} all that
{\ADD{pertained}} unto
Mephibosheth. And
Ziba
said, I
humbly beseech thee
{\ADD{that}} I may
find
grace in thy
sight, my
lord, O
king.\FTNT{*}{{\FR 16.4 }
I humbly…: Heb. I do obeisance}
\par }{\PP \VS{5}¶ And when
king
David
came to
Bahurim, behold, thence came
out a
man of the
family of the
house of
Saul, whose
name
{\ADD{was}}
Shimei, the
son of
Gera: he came
forth, and
cursed still as he
came.\FTNT{†}{{\FR 16.5 }
he came forth…: or, he still came forth and cursed}
\VS{6}And he
cast
stones at
David, and at all the
servants of
king
David: and all the
people and all the mighty
men
{\ADD{were}} on his right
hand and on his
left.
\VS{7}And thus
said
Shimei when he
cursed, Come
out, come
out, thou
bloody
man, and thou
man of
Belial:\FTNT{‡}{{\FR 16.7 }
bloody…: Heb. man of blood}
\VS{8}The
{\ND{
LORD}} hath
returned upon thee all the
blood of the
house of
Saul, in whose stead thou hast
reigned; and the
{\ND{
LORD}} hath
delivered the
kingdom into the
hand of
Absalom thy
son: and, behold, thou
{\ADD{art taken}} in thy
mischief, because thou
{\ADD{art}} a
bloody
man.\FTNT{§}{{\FR 16.8 }
behold…: Heb. behold thee in thy evil}
\par }{\PP \VS{9}¶ Then
said
Abishai the
son of
Zeruiah unto the
king, Why should this
dead
dog
curse my
lord the
king? let me go
over, I pray thee, and take
off his
head.
\VS{10}And the
king
said, What have I to do with you, ye
sons of
Zeruiah? so let him
curse, because the
{\ND{
LORD}} hath
said unto him,
Curse
David. Who shall then
say, Wherefore hast thou done
so?
\VS{11}And
David
said to
Abishai, and to all his
servants, Behold, my
son, which came
forth of my
bowels,
seeketh my
life: how much more now
{\ADD{may this}}
Benjamite
{\ADD{do it}}? let him
alone, and let him
curse; for the
{\ND{
LORD}} hath
bidden him.
\VS{12}It may be that the
{\ND{
LORD}} will
look on mine
affliction, and that the
{\ND{
LORD}} will
requite me
good for his
cursing this
day.\FTNT{**}{{\FR 16.12 }
affliction: or, tears: Heb. eye}
\VS{13}And as
David and his
men
went by the
way,
Shimei went
along on the
hill’s
side over
against him, and
cursed as he
went, and
threw
stones
at him, and
cast
dust.\FTNT{††}{{\FR 16.13 }
cast…: Heb. dusted him with dust}
\VS{14}And the
king, and all the
people that
{\ADD{were}} with him,
came
weary, and
refreshed themselves there.
\par }{\PP \VS{15}¶ And
Absalom, and all the
people the
men of
Israel,
came to
Jerusalem, and
Ahithophel with him.
\VS{16}And it came to pass, when
Hushai the
Archite,
David’s
friend, was
come unto
Absalom, that
Hushai
said unto
Absalom, God
save the
king, God
save the
king.\FTNT{‡‡}{{\FR 16.16 }
God…: Heb. Let the king live}
\VS{17}And
Absalom
said to
Hushai,
{\ADD{Is}} this thy
kindness to thy
friend? why
wentest thou not with thy
friend?
\VS{18}And
Hushai
said unto
Absalom, Nay; but whom the
{\ND{
LORD}}, and this
people, and all the
men of
Israel,
choose, his will I be, and with him will I
abide.
\VS{19}And
again, whom should I
serve?
{\ADD{should I}} not
{\ADD{serve}} in the
presence of his
son? as I have
served in thy
father’s
presence, so will I be in thy
presence.
\par }{\PP \VS{20}¶ Then
said
Absalom to
Ahithophel,
Give
counsel among you what we shall
do.
\VS{21}And
Ahithophel
said unto
Absalom, Go
in unto thy
father’s
concubines, which he hath
left to
keep the
house; and all
Israel shall
hear that thou art
abhorred of thy
father: then shall the
hands of all that
{\ADD{are}} with thee be
strong.
\VS{22}So they
spread
Absalom a
tent upon the top of the
house; and
Absalom went
in unto his
father’s
concubines in the
sight of all
Israel.
\VS{23}And the
counsel of
Ahithophel, which he
counselled in those
days,
{\ADD{was}} as if a
man had
enquired at the
oracle of
God: so
{\ADD{was}} all the
counsel of
Ahithophel both with
David and with
Absalom.\FTNT{§§}{{\FR 16.23 }
oracle: Heb. word}

\par }\Chap{17}{\PP \VerseOne{1}Moreover
Ahithophel
said unto
Absalom, Let me now choose
out
twelve
thousand
men, and I will
arise and
pursue
after
David this
night:
\VS{2}And I will
come upon him while he
{\ADD{is}}
weary and
weak
handed, and will make him
afraid: and all the
people that
{\ADD{are}} with him shall
flee; and I will
smite the
king only:
\VS{3}And I will bring
back all the
people unto thee: the
man whom thou
seekest
{\ADD{is}} as if all
returned:
{\ADD{so}} all the
people shall be in
peace.
\VS{4}And the
saying
pleased
Absalom
well, and all the
elders of
Israel.\FTNT{*}{{\FR 17.4 }
pleased…: Heb. was right in the eyes of, etc}
\VS{5}Then
said
Absalom,
Call now
Hushai the
Archite also, and let us
hear likewise what he
saith.\FTNT{†}{{\FR 17.5 }
what…: Heb. what is in his mouth}
\VS{6}And when
Hushai was
come to
Absalom,
Absalom
spake unto him,
saying,
Ahithophel hath
spoken after this
manner: shall we
do
{\ADD{after}} his
saying? if not;
speak thou.\FTNT{‡}{{\FR 17.6 }
his saying: Heb. his word}
\VS{7}And
Hushai
said unto
Absalom, The
counsel that
Ahithophel hath
given
{\ADD{is}} not
good at this
time.\FTNT{§}{{\FR 17.7 }
given: Heb. counselled}
\VS{8}For,
said
Hushai, thou
knowest thy
father and his
men, that they
{\ADD{be}} mighty
men, and they
{\ADD{be}}
chafed in their
minds, as a
bear
robbed of her whelps in the
field: and thy
father
{\ADD{is}} a
man of
war, and will not
lodge with the
people.\FTNT{**}{{\FR 17.8 }
chafed…: Heb. bitter of soul}
\VS{9}Behold, he is
hid now in
some
pit, or in
some
{\ADD{other}}
place: and it will come to pass, when some of them be
overthrown at the
first, that
whosoever
heareth it will
say, There is a
slaughter among the
people that
follow
Absalom.\FTNT{††}{{\FR 17.9 }
overthrown: Heb. fallen}
\VS{10}And he also
{\ADD{that is}}
valiant, whose
heart
{\ADD{is}} as the
heart of a
lion, shall
utterly
melt: for all
Israel
knoweth that thy
father
{\ADD{is}} a mighty
man, and
{\ADD{they}} which
{\ADD{be}} with him
{\ADD{are}}
valiant
men.
\VS{11}Therefore I
counsel that all
Israel be
generally
gathered unto thee, from
Dan even to
Beer-sheba, as the
sand that
{\ADD{is}} by the
sea for
multitude; and that thou
go to
battle in thine own
person.\FTNT{‡‡}{{\FR 17.11 }
that thou…: Heb. that thy face, or, presence go, etc}
\VS{12}So shall we
come upon him in
some
place where he shall be
found, and we will
light upon him as the
dew
falleth on the
ground: and of him and of all the
men that
{\ADD{are}} with him there shall not be
left so much
as
one.
\VS{13}Moreover,
if he be
gotten into a
city, then shall all
Israel
bring
ropes to that
city, and we will
draw it into the
river, until there be not
one small
stone
found there.
\VS{14}And
Absalom and all the
men of
Israel
said, The
counsel of
Hushai the
Archite
{\ADD{is}}
better than the
counsel of
Ahithophel. For the
{\ND{
LORD}} had
appointed to
defeat the
good
counsel of
Ahithophel, to the
intent that the
{\ND{
LORD}} might
bring
evil upon
Absalom.\FTNT{§§}{{\FR 17.14 }
appointed: Heb. commanded}
\par }{\PP \VS{15}¶ Then
said
Hushai unto
Zadok and to
Abiathar the
priests, Thus and thus did
Ahithophel
counsel
Absalom and the
elders of
Israel; and thus and thus have I
counselled.
\VS{16}Now therefore
send
quickly, and
tell
David,
saying,
Lodge not this
night in the
plains of the
wilderness, but
speedily pass
over; lest the
king be swallowed
up, and all the
people that
{\ADD{are}} with him.
\VS{17}Now
Jonathan and
Ahimaaz
stayed by
En-rogel; for they
might not be
seen to
come into the
city: and a
wench
went and
told them; and they
went and
told
king
David.
\VS{18}Nevertheless a
lad
saw them, and
told
Absalom: but they
went
both of them away
quickly, and
came to a
man’s
house in
Bahurim, which had a
well in his
court;
whither they went
down.
\VS{19}And the
woman
took and
spread a
covering over the
well’s
mouth, and
spread ground
corn thereon; and the
thing was not
known.
\VS{20}And when
Absalom’s
servants
came to the
woman to the
house, they
said, Where
{\ADD{is}}
Ahimaaz and
Jonathan? And the
woman
said unto them, They be gone
over the
brook of
water. And when they had
sought and could not
find
{\ADD{them}}, they
returned to
Jerusalem.
\VS{21}And it came to pass,
after they were
departed, that they came
up out of the
well, and
went and
told
king
David, and
said unto
David,
Arise, and pass
quickly
over the
water: for thus hath
Ahithophel
counselled against you.
\par }{\PP \VS{22}Then
David
arose, and all the
people that
{\ADD{were}} with him, and they passed
over
Jordan: by the
morning
light there
lacked not
one of them that was not gone
over
Jordan.
\par }{\PP \VS{23}¶ And when
Ahithophel
saw that his
counsel was not
followed, he
saddled
{\ADD{his}}
ass, and
arose, and gat him
home to his
house, to his
city, and put his
household in
order, and
hanged himself, and
died, and was
buried in the
sepulchre of his
father.\FTNT{***}{{\FR 17.23 }
followed: Heb. done}\FTNT{†††}{{\FR 17.23 }
put his…: Heb. gave charge concerning his house}
\VS{24}Then
David
came to
Mahanaim. And
Absalom passed
over
Jordan, he and all the
men of
Israel with him.
\par }{\PP \VS{25}¶ And
Absalom
made
Amasa captain of the
host instead of
Joab: which
Amasa
{\ADD{was}} a
man’s
son, whose
name
{\ADD{was}}
Ithra an
Israelite, that went
in to
Abigail the
daughter of
Nahash,
sister to
Zeruiah
Joab’s
mother.\FTNT{‡‡‡}{{\FR 17.25 }
Ithra an Israelite: or, Jether an Ishmaelite}\FTNT{§§§}{{\FR 17.25 }
Abigail: Heb. Abigal}\FTNT{*}{{\FR 17.25 }
Nahash: or, Jesse}
\VS{26}So
Israel and
Absalom
pitched in the
land of
Gilead.
\par }{\PP \VS{27}¶ And it came to pass, when
David was
come to
Mahanaim, that
Shobi the
son of
Nahash of
Rabbah of the
children of
Ammon, and
Machir the
son of
Ammiel of
Lo-debar, and
Barzillai the
Gileadite of
Rogelim,
\VS{28}Brought
beds, and
basons, and
earthen
vessels, and
wheat, and
barley, and
flour, and
parched
{\ADD{corn}}, and
beans, and
lentiles, and
parched
{\ADD{pulse}},\FTNT{†}{{\FR 17.28 }
basons: or, cups}
\VS{29}And
honey, and
butter, and
sheep, and
cheese of
kine, for
David, and for the
people that
{\ADD{were}} with him, to
eat: for they
said, The
people
{\ADD{is}}
hungry, and
weary, and
thirsty, in the
wilderness.

\par }\Chap{18}{\PP \VerseOne{1}And
David
numbered the
people that
{\ADD{were}} with him, and
set
captains of
thousands and
captains of
hundreds over them.
\VS{2}And
David sent
forth a third
part of the
people under the
hand of
Joab, and a third
part under the
hand of
Abishai the
son of
Zeruiah,
Joab’s
brother, and a third
part under the
hand of
Ittai the
Gittite. And the
king
said unto the
people, I will
surely go
forth with you
myself also.
\VS{3}But the
people
answered, Thou shalt not go
forth: for if we
flee
away, they will not
care for us; neither if
half of us
die, will they
care for us: but now
{\ADD{thou art}}
worth
ten
thousand of us: therefore now
{\ADD{it is}}
better that thou
succour us out of the
city.\FTNT{*}{{\FR 18.3 }
care…: Heb. set their heart on us}\FTNT{†}{{\FR 18.3 }
worth…: Heb. as ten thousand of us}\FTNT{‡}{{\FR 18.3 }
succour: Heb. be to succour or help}
\VS{4}And the
king
said unto them, What
seemeth you
best I will
do. And the
king
stood by the
gate
side, and all the
people came
out by
hundreds and by
thousands.
\VS{5}And the
king
commanded
Joab and
Abishai and
Ittai,
saying,
{\ADD{Deal}}
gently for my sake with the young
man,
{\ADD{even}} with
Absalom. And all the
people
heard when the
king gave all the
captains
charge
concerning
Absalom.
\par }{\PP \VS{6}¶ So the
people went
out into the
field
against
Israel: and the
battle was in the
wood of
Ephraim;
\VS{7}Where the
people of
Israel were
slain
before the
servants of
David, and there was there a
great
slaughter that
day of
twenty
thousand
{\ADD{men}}.
\VS{8}For the
battle was there
scattered over the
face of all the
country: and the
wood
devoured
more
people that
day than the
sword
devoured.\FTNT{§}{{\FR 18.8 }
devoured: Heb. multiplied to devour}
\par }{\PP \VS{9}¶ And
Absalom
met the
servants of
David. And
Absalom
rode upon a
mule, and the
mule
went under the thick
boughs of a
great
oak, and his
head caught
hold of the
oak, and he was taken
up between the
heaven and the
earth; and the
mule that
{\ADD{was}} under him went
away.
\VS{10}And a
certain
man
saw
{\ADD{it}}, and
told
Joab, and
said, Behold, I
saw
Absalom
hanged in an
oak.
\VS{11}And
Joab
said unto the
man that
told him, And, behold, thou
sawest
{\ADD{him}}, and why didst thou not
smite him there to the
ground? and I would have
given thee
ten
{\ADD{shekels}} of
silver, and
a
girdle.
\VS{12}And the
man
said unto
Joab,
Though I should
receive a
thousand
{\ADD{shekels}} of
silver in mine
hand,
{\ADD{yet}} would I not put
forth mine
hand against the
king’s
son: for in our
hearing the
king
charged thee and
Abishai and
Ittai,
saying,
Beware that none
{\ADD{touch}} the young
man
Absalom.\FTNT{**}{{\FR 18.12 }
receive: Heb. weigh upon mine hand}\FTNT{††}{{\FR 18.12 }
Beware…: Heb. Beware whosoever ye be of, etc}
\VS{13}Otherwise I should have
wrought
falsehood against mine own
life: for there is no
matter
hid from the
king, and thou thyself wouldest have
set thyself against
{\ADD{me}}.
\VS{14}Then
said
Joab, I may not
tarry thus with
thee. And he
took
three
darts in his
hand, and
thrust them through the
heart of
Absalom, while he
{\ADD{was}} yet
alive in the midst of the
oak.\FTNT{‡‡}{{\FR 18.14 }
with…: Heb. before thee}\FTNT{§§}{{\FR 18.14 }
midst: Heb. heart}
\VS{15}And
ten young
men that
bare
Joab’s
armour compassed
about and
smote
Absalom, and
slew him.
\VS{16}And
Joab
blew the
trumpet, and the
people
returned from
pursuing
after
Israel: for
Joab held
back the
people.
\VS{17}And they
took
Absalom, and
cast him into a
great
pit in the
wood, and
laid a
very
great
heap of
stones upon him: and all
Israel
fled every
one to his
tent.
\par }{\PP \VS{18}¶ Now
Absalom in his
lifetime had
taken and reared
up for himself a
pillar, which
{\ADD{is}} in the
king’s
dale: for he
said, I have no
son to keep my
name in
remembrance: and he
called the
pillar after his own
name: and it is
called unto this
day,
Absalom’s
place.
\par }{\PP \VS{19}¶ Then
said
Ahimaaz the
son of
Zadok, Let me now
run, and
bear the
king
tidings, how that the
{\ND{
LORD}} hath
avenged him of his
enemies.\FTNT{***}{{\FR 18.19 }
avenged…: Heb. judged him from the hand, etc}
\VS{20}And
Joab
said unto him,
Thou shalt not bear
tidings this
day, but thou shalt bear
tidings
another
day: but this
day thou shalt bear no
tidings, because the
king’s
son is
dead.\FTNT{†††}{{\FR 18.20 }
bear tidings this: Heb. be a man of tidings, etc}
\VS{21}Then
said
Joab to
Cushi,
Go
tell the
king what thou hast
seen. And
Cushi
bowed himself unto
Joab, and
ran.
\VS{22}Then
said
Ahimaaz the
son of
Zadok yet
again to
Joab, But
howsoever, let me, I pray thee, also
run
after
Cushi. And
Joab
said, Wherefore wilt thou
run, my
son, seeing that thou hast no
tidings
ready?\FTNT{‡‡‡}{{\FR 18.22 }
howsoever: Heb. be what may}\FTNT{§§§}{{\FR 18.22 }
ready: or, convenient}
\VS{23}But howsoever,
{\ADD{said he}}, let me
run. And he
said unto him,
Run. Then
Ahimaaz
ran by the
way of the
plain, and
overran
Cushi.
\VS{24}And
David
sat between the
two
gates: and the
watchman went
up to the
roof over the
gate unto the
wall, and lifted
up his
eyes, and
looked, and behold a
man
running alone.
\VS{25}And the
watchman
cried, and
told the
king. And the
king
said, If he
{\ADD{be}} alone,
{\ADD{there is}}
tidings in his
mouth. And he
came
apace, and drew
near.
\VS{26}And the
watchman
saw
another
man
running: and the
watchman
called unto the
porter, and
said, Behold
{\ADD{another}}
man
running alone. And the
king
said, He also bringeth
tidings.
\VS{27}And the
watchman
said, Me
thinketh the
running of the
foremost is like the
running of
Ahimaaz the
son of
Zadok. And the
king
said, He
{\ADD{is}} a
good
man, and
cometh with
good
tidings.\FTNT{*}{{\FR 18.27 }
Me…: Heb. I see the running}
\VS{28}And
Ahimaaz
called, and
said unto the
king, All is
well. And he fell
down to the
earth upon his
face before the
king, and
said,
Blessed
{\ADD{be}} the
{\ND{
LORD}} thy
God, which hath delivered
up the
men that lifted
up their
hand against my
lord the
king.\FTNT{†}{{\FR 18.28 }
All…: or, Peace be to thee: Heb. Peace}\FTNT{‡}{{\FR 18.28 }
delivered…: Heb. shut up}
\VS{29}And the
king
said, Is the young
man
Absalom
safe? And
Ahimaaz
answered, When
Joab
sent the
king’s
servant, and
{\ADD{me}} thy
servant, I
saw a
great
tumult, but I
knew not what
{\ADD{it was}}.\FTNT{§}{{\FR 18.29 }
Is…: Heb. Is there peace?}
\VS{30}And the
king
said
{\ADD{unto him}}, Turn
aside,
{\ADD{and}}
stand
here. And he turned
aside, and stood
still.
\VS{31}And, behold,
Cushi
came; and
Cushi
said,
Tidings, my
lord the
king: for the
{\ND{
LORD}} hath
avenged thee this
day of all
them that rose
up against thee.\FTNT{**}{{\FR 18.31 }
Tidings: Heb. Tidings is brought}
\VS{32}And the
king
said unto
Cushi, Is the young
man
Absalom
safe? And
Cushi
answered, The
enemies of my
lord the
king, and all that
rise against thee to do
{\ADD{thee}}
hurt, be as
{\ADD{that}} young
man
{\ADD{is}}.
\par }{\PP \VS{33}¶ And the
king was much
moved, and went
up to the
chamber over the
gate, and
wept: and as he
went, thus he
said, O my
son
Absalom, my
son, my
son
Absalom! would God I had
died for
thee, O
Absalom, my
son, my
son!

\par }\Chap{19}{\PP \VerseOne{1}And it was
told
Joab, Behold, the
king
weepeth and
mourneth for
Absalom.
\VS{2}And the
victory that
day was
{\ADD{turned}} into
mourning unto all the
people: for the
people
heard
say that
day how the
king was
grieved for his
son.\FTNT{*}{{\FR 19.2 }
victory: Heb. salvation, or, deliverance}
\VS{3}And the
people
gat them by
stealth that
day into the
city, as
people being
ashamed steal
away when they
flee in
battle.
\VS{4}But the
king
covered his
face, and the
king
cried with a
loud
voice, O my
son
Absalom, O
Absalom, my
son, my
son!
\VS{5}And
Joab
came into the
house to the
king, and
said, Thou hast
shamed this
day the
faces of all thy
servants, which this
day have
saved thy
life, and the
lives of thy
sons and of thy
daughters, and the
lives of thy
wives, and the
lives of thy
concubines;
\VS{6}In that thou
lovest thine
enemies, and
hatest thy
friends. For thou hast
declared this
day, that thou regardest neither
princes nor
servants: for this
day I
perceive, that
if
Absalom had
lived, and all we had
died this
day, then it had pleased thee
well.\FTNT{†}{{\FR 19.6 }
In…: Heb. By loving, etc}\FTNT{‡}{{\FR 19.6 }
that thou regardest…: Heb. that princes or servants are not to thee}
\VS{7}Now therefore
arise, go
forth, and
speak
comfortably unto thy
servants: for I
swear by the
{\ND{
LORD}}, if thou go not
forth, there will not
tarry
one with thee this
night: and that will be
worse unto thee than all the
evil that
befell thee from thy
youth until now.\FTNT{§}{{\FR 19.7 }
comfortably…: Heb. to the heart of thy servants}
\VS{8}Then the
king
arose, and
sat in the
gate. And they
told unto all the
people,
saying, Behold, the
king doth
sit in the
gate. And all the
people
came
before the
king: for
Israel had
fled every
man to his
tent.
\par }{\PP \VS{9}¶ And all the
people were at
strife throughout all the
tribes of
Israel,
saying, The
king
saved us out of the
hand of our
enemies, and he
delivered us out of the
hand of the
Philistines; and now he is
fled out of the
land for
Absalom.
\VS{10}And
Absalom, whom we
anointed over us, is
dead in
battle. Now therefore why
speak ye not a word of
bringing the
king
back?\FTNT{**}{{\FR 19.10 }
speak…: Heb. are ye silent?}
\par }{\PP \VS{11}¶ And
king
David
sent to
Zadok and to
Abiathar the
priests,
saying,
Speak unto the
elders of
Judah,
saying, Why are ye the
last to
bring the
king
back to his
house? seeing the
speech of all
Israel is
come to the
king,
{\ADD{even}} to his
house.
\VS{12}Ye
{\ADD{are}} my
brethren, ye
{\ADD{are}} my
bones and my
flesh: wherefore then are ye the
last to bring
back the
king?
\VS{13}And
say ye to
Amasa,
{\ADD{Art}} thou not of my
bone, and of my
flesh?
God
do so to me, and
more also, if thou be not
captain of the
host
before me
continually in the
room of
Joab.
\VS{14}And he
bowed the
heart of all the
men of
Judah, even as
{\ADD{the heart of}}
one
man; so that they
sent
{\ADD{this word}} unto the
king,
Return thou, and all thy
servants.
\VS{15}So the
king
returned, and
came to
Jordan. And
Judah
came to
Gilgal, to
go to
meet the
king, to
conduct the
king
over
Jordan.
\par }{\PP \VS{16}¶ And
Shimei the
son of
Gera, a
Benjamite, which
{\ADD{was}} of
Bahurim,
hasted and came
down with the
men of
Judah to
meet
king
David.
\VS{17}And
{\ADD{there were}} a
thousand
men of
Benjamin with him, and
Ziba the
servant of the
house of
Saul, and his
fifteen
sons and his
twenty
servants with him; and they went
over
Jordan
before the
king.
\VS{18}And there went
over a ferry
boat to carry
over the
king’s
household, and to
do what he
thought
good. And
Shimei the
son of
Gera fell
down
before the
king, as he was come
over
Jordan;\FTNT{††}{{\FR 19.18 }
what…: Heb. the good in his eyes}
\VS{19}And
said unto the
king, Let not my
lord
impute
iniquity unto me, neither do thou
remember that which thy
servant did
perversely the
day that my
lord the
king went
out of
Jerusalem, that the
king should
take it to his
heart.
\VS{20}For thy
servant doth
know that I have
sinned: therefore, behold, I am
come the
first this
day of all the
house of
Joseph to go
down to
meet my
lord the
king.
\VS{21}But
Abishai the
son of
Zeruiah
answered and
said, Shall not
Shimei be put to
death for this, because he
cursed the
{\ND{
LORD’s}}
anointed?
\VS{22}And
David
said, What have I to do with you, ye
sons of
Zeruiah, that ye should this
day be
adversaries unto me? shall there any
man be put to
death this
day in
Israel? for do not I
know that I
{\ADD{am}} this
day
king over
Israel?
\VS{23}Therefore the
king
said unto
Shimei, Thou shalt not
die. And the
king
sware unto him.
\par }{\PP \VS{24}¶ And
Mephibosheth the
son of
Saul came
down to
meet the
king, and had
neither
dressed his
feet, nor
trimmed his
beard, nor
washed his
clothes, from the
day the
king
departed until the
day he
came
{\ADD{again}} in
peace.
\VS{25}And it came to pass, when he was
come to
Jerusalem to
meet the
king, that the
king
said unto him, Wherefore
wentest not thou with me,
Mephibosheth?
\VS{26}And he
answered, My
lord, O
king, my
servant
deceived me: for thy
servant
said, I will
saddle me an
ass, that I may
ride thereon, and
go to the
king; because thy
servant
{\ADD{is}}
lame.
\VS{27}And he hath
slandered thy
servant unto my
lord the
king; but my
lord the
king
{\ADD{is}} as an
angel of
God:
do therefore
{\ADD{what is}}
good in thine
eyes.
\VS{28}For all
{\ADD{of}} my
father’s
house were
but
dead
men before my
lord the
king: yet didst thou
set thy
servant among them that did
eat at thine own
table. What
right therefore
have I yet to
cry any more unto the
king?\FTNT{‡‡}{{\FR 19.28 }
dead…: Heb. men of death}
\VS{29}And the
king
said unto him, Why
speakest thou any more of thy
matters? I have
said, Thou and
Ziba
divide the
land.
\VS{30}And
Mephibosheth
said unto the
king, Yea, let him
take all,
forasmuch
as my
lord the
king is come
again in
peace unto his own
house.
\par }{\PP \VS{31}¶ And
Barzillai the
Gileadite came
down from
Rogelim, and went
over
Jordan with the
king, to
conduct him over
Jordan.
\VS{32}Now
Barzillai was a
very
aged man,
{\ADD{even}}
fourscore
years
old: and he had provided the
king of
sustenance while he
lay at
Mahanaim; for he
{\ADD{was}} a
very
great
man.
\VS{33}And the
king
said unto
Barzillai, Come thou
over with me, and I will
feed thee with me in
Jerusalem.
\VS{34}And
Barzillai
said unto the
king, How
long have I to
live, that I should go
up with the
king unto
Jerusalem?\FTNT{§§}{{\FR 19.34 }
How…: Heb. How many days are the years of my life}
\VS{35}I
{\ADD{am}} this
day
fourscore
years
old:
{\ADD{and}} can I
discern between
good and
evil? can thy
servant
taste
what I
eat or
what I
drink? can I
hear any more the
voice of singing
men and singing
women? wherefore then should thy
servant be yet a
burden unto my
lord the
king?
\VS{36}Thy
servant will
go a little
way
over
Jordan with the
king: and why should the
king
recompense it me with such a
reward?
\VS{37}Let thy
servant, I pray thee, turn back
again, that I may
die in mine own
city,
{\ADD{and be buried}} by the
grave of my
father and of my
mother. But behold thy
servant
Chimham; let him go
over with my
lord the
king; and
do to him what shall seem
good unto thee.
\VS{38}And the
king
answered,
Chimham shall
go over with me, and I will
do to him that which shall seem
good unto thee: and whatsoever thou shalt
require of me,
{\ADD{that}} will I
do for thee.\FTNT{***}{{\FR 19.38 }
require: Heb. choose}
\VS{39}And all the
people went
over
Jordan. And when the
king was come
over, the
king
kissed
Barzillai, and
blessed him; and he
returned unto his own
place.
\par }{\PP \VS{40}Then the
king went
on to
Gilgal, and
Chimham went
on with him: and all the
people of
Judah
conducted the
king, and also
half the
people of
Israel.\FTNT{†††}{{\FR 19.40 }
Chimham: Heb. Chimhan}
\par }{\PP \VS{41}¶ And, behold, all the
men of
Israel
came to the
king, and
said unto the
king, Why have our
brethren the
men of
Judah stolen thee
away, and have
brought the
king, and his
household, and all
David’s
men with him, over
Jordan?
\VS{42}And all the
men of
Judah
answered the
men of
Israel, Because the
king
{\ADD{is}} near of
kin to us: wherefore then be ye
angry for this
matter? have we
eaten at
all of the
king’s
{\ADD{cost}}? or hath he
given us any
gift?
\VS{43}And the
men of
Israel
answered the
men of
Judah, and
said, We have
ten
parts in the
king, and
we have also more
{\ADD{right}} in
David than ye: why then did ye
despise us, that our
advice should not be
first had in bringing
back our
king? And the
words of the
men of
Judah were
fiercer than the
words of the
men of
Israel.\FTNT{‡‡‡}{{\FR 19.43 }
despise…: Heb. set us at light}

\par }\Chap{20}{\PP \VerseOne{1}And there
happened to be there a
man of
Belial, whose
name
{\ADD{was}}
Sheba, the
son of
Bichri, a
Benjamite: and he
blew a
trumpet, and
said, We have no
part in
David, neither have we
inheritance in the
son of
Jesse:
every
man to his
tents, O
Israel.
\VS{2}So every
man of
Israel went
up from
after
David,
{\ADD{and}}
followed
Sheba the
son of
Bichri: but the
men of
Judah
clave unto their
king, from
Jordan even to
Jerusalem.
\par }{\PP \VS{3}¶ And
David
came to his
house at
Jerusalem; and the
king
took the
ten
women
{\ADD{his}}
concubines, whom he had
left to
keep the
house, and
put them in
ward, and
fed them, but went not
in unto them. So they were shut
up unto the
day of their
death,
living in
widowhood.\FTNT{*}{{\FR 20.3 }
ward: Heb. an house of ward}\FTNT{†}{{\FR 20.3 }
shut…: Heb. bound}\FTNT{‡}{{\FR 20.3 }
living…: Heb. in widowhood of life}
\par }{\PP \VS{4}¶ Then
said the
king to
Amasa,
Assemble me the
men of
Judah within
three
days, and be thou here
present.\FTNT{§}{{\FR 20.4 }
Assemble: Heb. Call}
\VS{5}So
Amasa
went to
assemble
{\ADD{the men of}}
Judah: but he tarried
longer than the set
time which he had
appointed him.\FTNT{**}{{\FR 20.5 }
Assemble: Heb. Call}
\VS{6}And
David
said to
Abishai, Now shall
Sheba the
son of
Bichri do us more
harm than
{\ADD{did}}
Absalom:
take thou thy
lord’s
servants, and
pursue
after him, lest he
get him
fenced
cities, and
escape
us.\FTNT{††}{{\FR 20.6 }
escape…: Heb. deliver himself from our eyes}
\VS{7}And there went
out
after him
Joab’s
men, and the
Cherethites, and the
Pelethites, and all the mighty
men: and they went
out of
Jerusalem, to
pursue
after
Sheba the
son of
Bichri.
\VS{8}When they
{\ADD{were}} at the
great
stone which
{\ADD{is}} in
Gibeon,
Amasa
went
before them. And
Joab’s
garment that he had put
on was
girded unto him, and upon it a
girdle
{\ADD{with}} a
sword
fastened upon his
loins in the
sheath thereof; and as he went
forth it fell
out.
\VS{9}And
Joab
said to
Amasa,
{\ADD{Art}} thou in
health, my
brother? And
Joab
took
Amasa by the
beard with the
right
hand to
kiss him.
\VS{10}But
Amasa took no
heed to the
sword that
{\ADD{was}} in
Joab’s
hand: so he
smote him therewith in the
fifth
{\ADD{rib}}, and shed
out his
bowels to the
ground, and struck him not
again; and he
died. So
Joab and
Abishai his
brother
pursued
after
Sheba the
son of
Bichri.\FTNT{‡‡}{{\FR 20.10 }
struck…: Heb. doubled not his stroke}
\VS{11}And
one of
Joab’s
men
stood by him, and
said,
He that
favoureth
Joab, and he that
{\ADD{is}} for
David,
{\ADD{let him go}}
after
Joab.
\VS{12}And
Amasa
wallowed in
blood in the
midst of the
highway. And when the
man
saw that all the
people stood
still, he
removed
Amasa out of the
highway into the
field, and
cast a
cloth upon him, when he
saw that every one that
came by him stood
still.
\VS{13}When he was
removed out of the
highway, all the
people went
on
after
Joab, to
pursue
after
Sheba the
son of
Bichri.
\par }{\PP \VS{14}¶ And he
went through all the
tribes of
Israel unto
Abel, and to
Beth-maachah, and all the
Berites: and they were gathered
together, and
went also
after him.
\VS{15}And they
came and
besieged him in
Abel of
Beth-maachah, and they cast
up a
bank against the
city, and it
stood in the
trench: and all the
people that
{\ADD{were}} with
Joab
battered the
wall, to throw it
down.\FTNT{§§}{{\FR 20.15 }
in the trench: or, against the outmost wall}\FTNT{***}{{\FR 20.15 }
battered: Heb. marred to throw down}
\par }{\PP \VS{16}¶ Then
cried a
wise
woman out of the
city,
Hear,
hear;
say, I pray you, unto
Joab, Come
near hither, that I may
speak with thee.
\VS{17}And when he was come
near unto her, the
woman
said,
{\ADD{Art}} thou
Joab? And he
answered, I
{\ADD{am he}}. Then she
said unto him,
Hear the
words of thine
handmaid. And he
answered, I do
hear.
\VS{18}Then she
spake,
saying, They were
wont to
speak in old
time,
saying, They shall
surely
ask
{\ADD{counsel}} at
Abel: and so they
ended
{\ADD{the matter}}.\FTNT{†††}{{\FR 20.18 }
They were…: or, They plainly spake in the beginning, saying, Surely they will ask of Abel, and so make an end}
\VS{19}I
{\ADD{am one of them that are}}
peaceable
{\ADD{and}}
faithful in
Israel: thou
seekest to
destroy a
city and a
mother in
Israel: why wilt thou swallow
up the
inheritance of the
{\ND{
LORD}}?
\VS{20}And
Joab
answered and
said, Far be
it, far be
it from me, that I should swallow
up or
destroy.
\VS{21}The
matter
{\ADD{is}} not so: but a
man of
mount
Ephraim,
Sheba the
son of
Bichri by
name, hath lifted
up his
hand against the
king,
{\ADD{even}} against
David:
deliver him only, and I will
depart from the
city. And the
woman
said unto
Joab, Behold, his
head shall be
thrown to thee
over the
wall.\FTNT{‡‡‡}{{\FR 20.21 }
by name: Heb. by his name}
\VS{22}Then the
woman
went unto all the
people in her
wisdom. And they cut
off the
head of
Sheba the
son of
Bichri, and
cast
{\ADD{it}} out to
Joab. And he
blew a
trumpet, and they
retired from the
city, every
man to his
tent. And
Joab
returned to
Jerusalem unto the
king.\FTNT{§§§}{{\FR 20.22 }
retired: Heb. were scattered}
\par }{\PP \VS{23}¶ Now
Joab
{\ADD{was}} over all the
host of
Israel: and
Benaiah the
son of
Jehoiada
{\ADD{was}} over the
Cherethites and over the
Pelethites:
\VS{24}And
Adoram
{\ADD{was}} over the
tribute: and
Jehoshaphat the
son of
Ahilud
{\ADD{was}}
recorder:\FTNT{*}{{\FR 20.24 }
recorder: or, remembrancer}
\VS{25}And
Sheva
{\ADD{was}}
scribe: and
Zadok and
Abiathar
{\ADD{were}} the
priests:
\VS{26}And
Ira also the
Jairite was a chief
ruler about
David.\FTNT{†}{{\FR 20.26 }
a chief…: or, a prince}

\par }\Chap{21}{\PP \VerseOne{1}Then there was a
famine in the
days of
David
three
years,
year
after
year; and
David
enquired
of the
{\ND{
LORD}}. And the
{\ND{
LORD}}
answered,
{\ADD{It is}} for
Saul, and for
{\ADD{his}}
bloody
house, because he
slew the
Gibeonites.\FTNT{*}{{\FR 21.1 }
enquired…: Heb. sought the face, etc}
\VS{2}And the
king
called the
Gibeonites, and
said unto them; (now the
Gibeonites
{\ADD{were}} not of the
children of
Israel, but of the
remnant of the
Amorites; and the
children of
Israel had
sworn unto them: and
Saul
sought to
slay them in his
zeal to the
children of
Israel and
Judah.)
\VS{3}Wherefore
David
said unto the
Gibeonites, What shall I
do for you? and wherewith shall I make the
atonement, that ye may
bless the
inheritance of the
{\ND{
LORD}}?
\VS{4}And the
Gibeonites
said unto him, We will have no
silver nor
gold
of
Saul, nor of his
house; neither for us shalt thou
kill any
man in
Israel. And he
said, What ye shall
say,
{\ADD{that}} will I
do for you.\FTNT{†}{{\FR 21.4 }
We will…: or, It is not silver nor gold that we have to do with Saul or his house, neither pertains it to us to kill, etc}
\VS{5}And they
answered the
king, The
man that
consumed us, and that
devised against us
{\ADD{that}} we should be
destroyed from
remaining in any of the
coasts of
Israel,\FTNT{‡}{{\FR 21.5 }
devised…: or, cut us off}
\VS{6}Let
seven
men of his
sons be
delivered unto us, and we will hang them
up unto the
{\ND{
LORD}} in
Gibeah of
Saul,
{\ADD{whom}} the
{\ND{
LORD}} did
choose. And the
king
said, I will
give
{\ADD{them}}.\FTNT{§}{{\FR 21.6 }
whom…: or, chosen of the
{\ND{LORD}}}
\VS{7}But the
king
spared
Mephibosheth, the
son of
Jonathan the
son of
Saul, because of the
{\ND{
LORD’s}}
oath that
{\ADD{was}} between them, between
David and
Jonathan the
son of
Saul.
\VS{8}But the
king
took the
two
sons of
Rizpah the
daughter of
Aiah, whom she
bare unto
Saul,
Armoni and
Mephibosheth; and the
five
sons of
Michal the
daughter of
Saul, whom she brought
up for
Adriel the
son of
Barzillai the
Meholathite:\FTNT{**}{{\FR 21.8 }
Michal: or, Michal’s sister}\FTNT{††}{{\FR 21.8 }
brought…: Heb. bear to Adriel}
\VS{9}And he
delivered them into the
hands of the
Gibeonites, and they
hanged them in the
hill
before the
{\ND{
LORD}}: and they
fell
{\ADD{all}}
seven
together, and were put to
death in the
days of
harvest, in the
first
{\ADD{days}}, in the
beginning of
barley
harvest.
\par }{\PP \VS{10}¶ And
Rizpah the
daughter of
Aiah
took
sackcloth, and
spread it for her upon the
rock, from the
beginning of
harvest until
water
dropped upon them out of
heaven, and
suffered neither the
birds of the
air to
rest on them by
day, nor the
beasts of the
field by
night.
\VS{11}And it was
told
David what
Rizpah the
daughter of
Aiah, the
concubine of
Saul, had
done.
\par }{\PP \VS{12}¶ And
David
went and
took the
bones of
Saul and the
bones of
Jonathan his
son from the
men of
Jabesh-gilead, which had
stolen them from the
street of
Beth-shan, where the
Philistines had
hanged them,
when the
Philistines had
slain
Saul in
Gilboa:
\VS{13}And he brought
up from thence the
bones of
Saul and the
bones of
Jonathan his
son; and they
gathered the
bones of them that were
hanged.
\VS{14}And the
bones of
Saul and
Jonathan his
son
buried they in the
country of
Benjamin in
Zelah, in the
sepulchre of
Kish his
father: and they
performed all that the
king
commanded. And
after that
God was
intreated for the
land.
\par }{\PP \VS{15}¶ Moreover the
Philistines had yet
war again with
Israel; and
David went
down, and his
servants with him, and
fought against the
Philistines: and
David waxed
faint.
\VS{16}And
Ishbi-benob, which
{\ADD{was}} of the
sons of the
giant, the
weight of whose
spear
{\ADD{weighed}}
three
hundred
{\ADD{shekels}} of
brass in
weight, he being
girded with a
new
{\ADD{sword}},
thought to have
slain
David.\FTNT{‡‡}{{\FR 21.16 }
the giant: or, Rapha}\FTNT{§§}{{\FR 21.16 }
spear: Heb. the staff, or, the head}
\VS{17}But
Abishai the
son of
Zeruiah
succoured him, and
smote the
Philistine, and
killed him. Then the
men of
David
sware unto him,
saying, Thou shalt
go no more out with us to
battle, that thou
quench not the
light of
Israel.\FTNT{***}{{\FR 21.17 }
light: Heb. candle, or, lamp}
\VS{18}And it came to pass after
this, that there was again a
battle with the
Philistines at
Gob: then
Sibbechai the
Hushathite
slew
Saph, which
{\ADD{was}} of the
sons of the
giant.\FTNT{†††}{{\FR 21.18 }
the giant: or, Rapha}\FTNT{‡‡‡}{{\FR 21.18 }
Saph: or, Sippai}
\VS{19}And there was again a
battle in
Gob with the
Philistines, where
Elhanan the
son of
Jaare-oregim, a
Beth-lehemite,
slew
{\ADD{the brother of}}
Goliath the
Gittite, the
staff of whose
spear
{\ADD{was}} like a
weaver’s
beam.\FTNT{§§§}{{\FR 21.19 }
Jaare-oregim: or, Jair}
\VS{20}And there was yet a
battle in
Gath, where was a
man of
{\ADD{great}}
stature, that had on every
hand
six
fingers, and on every
foot
six
toes,
four and
twenty in
number; and he also was
born to the
giant.\FTNT{*}{{\FR 21.20 }
the giant: or, Rapha}
\VS{21}And when he
defied
Israel,
Jonathan the
son of
Shimea the
brother of
David
slew him.\FTNT{†}{{\FR 21.21 }
defied: or, reproached}\FTNT{‡}{{\FR 21.21 }
Shimea: also called, Shammah}
\VS{22}These
four were
born to the
giant in
Gath, and
fell by the
hand of
David, and by the
hand of his
servants.

\par }\Chap{22}{\PP \VerseOne{1}And
David
spake unto the
{\ND{
LORD}} the
words of this
song in the
day
{\ADD{that}} the
{\ND{
LORD}} had
delivered him out of the
hand of all his
enemies, and out of the
hand of
Saul:
\par }{\PP \VS{2}And he
said, The
{\ND{
LORD}}
{\ADD{is}} my
rock, and my
fortress, and my
deliverer;
\VS{3}The
God of my
rock; in him will I
trust:
{\ADD{he is}} my
shield, and the
horn of my
salvation, my high
tower, and my
refuge, my
saviour; thou
savest me from
violence.
\VS{4}I will
call on the
{\ND{
LORD}},
{\ADD{who is}} worthy to be
praised: so shall I be
saved from mine
enemies.
\VS{5}When the
waves of
death
compassed me, the
floods of ungodly
men made me
afraid;\FTNT{*}{{\FR 22.5 }
waves: or, pangs}\FTNT{†}{{\FR 22.5 }
ungodly…: Heb. Belial}
\VS{6}The
sorrows of
hell compassed me
about; the
snares of
death
prevented me;\FTNT{‡}{{\FR 22.6 }
sorrows: or, cords}
\VS{7}In my
distress I
called upon the
{\ND{
LORD}}, and
cried to my
God: and he did
hear my
voice out of his
temple, and my
cry
{\ADD{did enter}} into his
ears.
\VS{8}Then the
earth
shook and
trembled; the
foundations of
heaven
moved and
shook, because he was
wroth.
\VS{9}There went
up a
smoke out of his
nostrils, and
fire out of his
mouth
devoured:
coals were
kindled by it.\FTNT{§}{{\FR 22.9 }
out of his nostrils: Heb. by, etc}
\VS{10}He
bowed the
heavens also, and came
down; and
darkness
{\ADD{was}} under his
feet.
\VS{11}And he
rode upon a
cherub, and did
fly: and he was
seen upon the
wings of the
wind.
\VS{12}And he
made
darkness
pavilions round
about him,
dark
waters,
{\ADD{and}} thick
clouds of the
skies.\FTNT{**}{{\FR 22.12 }
dark…: Heb. binding of waters}
\VS{13}Through the
brightness before him were
coals of
fire
kindled.
\VS{14}The
{\ND{
LORD}}
thundered from
heaven, and the most
High
uttered his
voice.
\VS{15}And he sent
out
arrows, and
scattered them;
lightning, and
discomfited them.
\VS{16}And the
channels of the
sea
appeared, the
foundations of the
world were
discovered, at the
rebuking of the
{\ND{
LORD}}, at the
blast of the
breath of his
nostrils.
\VS{17}He
sent from
above, he
took me; he
drew me out of
many
waters;\FTNT{††}{{\FR 22.17 }
many: or, great}
\VS{18}He
delivered me from my
strong
enemy,
{\ADD{and}} from them that
hated me: for they were too
strong for me.
\VS{19}They
prevented me in the
day of my
calamity: but the
{\ND{
LORD}} was my
stay.
\VS{20}He brought me
forth also into a large
place: he
delivered me, because he
delighted in me.
\VS{21}The
{\ND{
LORD}}
rewarded me according to my
righteousness: according to the
cleanness of my
hands hath he
recompensed me.
\VS{22}For I have
kept the
ways of the
{\ND{
LORD}}, and have not wickedly
departed from my
God.
\VS{23}For all his
judgments
{\ADD{were}} before me: and
{\ADD{as for}} his
statutes, I did not
depart from them.
\VS{24}I was also
upright before him, and have
kept myself from mine
iniquity.\FTNT{‡‡}{{\FR 22.24 }
before: Heb. to}
\VS{25}Therefore the
{\ND{
LORD}} hath
recompensed me according to my
righteousness; according to my
cleanness in his eye
sight.\FTNT{§§}{{\FR 22.25 }
in…: Heb. before his eyes}
\VS{26}With the
merciful thou wilt shew thyself
merciful,
{\ADD{and}} with the
upright
man thou wilt shew thyself
upright.
\VS{27}With the
pure thou wilt shew thyself
pure; and with the
froward thou wilt shew thyself
unsavoury.\FTNT{***}{{\FR 22.27 }
shew thyself unsavoury: or, wrestle}
\VS{28}And the
afflicted
people thou wilt
save: but thine
eyes
{\ADD{are}} upon the
haughty,
{\ADD{that}} thou mayest bring
{\ADD{them}}
down.
\VS{29}For thou
{\ADD{art}} my
lamp, O
{\ND{
LORD}}: and the
{\ND{
LORD}} will
lighten my
darkness.\FTNT{†††}{{\FR 22.29 }
lamp: or, candle}
\VS{30}For by thee I have
run through a
troop: by my
God have I leaped
over a
wall.\FTNT{‡‡‡}{{\FR 22.30 }
run: or, broken}
\VS{31}{\ADD{As for}}
God, his
way
{\ADD{is}}
perfect; the
word of the
{\ND{
LORD}}
{\ADD{is}}
tried: he
{\ADD{is}} a
buckler to all them that
trust in him.\FTNT{§§§}{{\FR 22.31 }
tried: or, refined}
\VS{32}For who
{\ADD{is}}
God,
save the
{\ND{
LORD}}? and who
{\ADD{is}} a
rock,
save our
God?
\VS{33}God
{\ADD{is}} my
strength
{\ADD{and}}
power: and he
maketh my
way
perfect.\FTNT{*}{{\FR 22.33 }
maketh: Heb. riddeth, or, looseth}
\VS{34}He
maketh my
feet
like
hinds’
{\ADD{feet:}} and
setteth me upon my high
places.\FTNT{†}{{\FR 22.34 }
maketh: Heb. equalleth}
\VS{35}He
teacheth my
hands to
war; so that a
bow of
steel is
broken by mine
arms.\FTNT{‡}{{\FR 22.35 }
to war: Heb. for the war}
\VS{36}Thou hast also
given me the
shield of thy
salvation: and thy
gentleness hath made me
great.\FTNT{§}{{\FR 22.36 }
made…: Heb. multiplied me}
\VS{37}Thou hast
enlarged my
steps under me; so that my
feet did not
slip.\FTNT{**}{{\FR 22.37 }
feet: Heb. ankles}
\VS{38}I have
pursued mine
enemies, and
destroyed them; and turned not
again until I had
consumed them.
\VS{39}And I have
consumed them, and
wounded them, that they could not
arise: yea, they are
fallen under my
feet.
\VS{40}For thou hast
girded me with
strength to
battle: them that rose
up against me hast thou
subdued under me.\FTNT{††}{{\FR 22.40 }
subdued: Heb. caused to bow}
\VS{41}Thou hast also
given me the
necks of mine
enemies, that I might
destroy them that
hate me.
\VS{42}They
looked, but
{\ADD{there was}} none to
save;
{\ADD{even}} unto the
{\ND{
LORD}}, but he
answered them not.
\VS{43}Then did I
beat them as small as the
dust of the
earth, I did
stamp them as the
mire of the
street,
{\ADD{and}} did spread them
abroad.
\VS{44}Thou also hast
delivered me from the
strivings of my
people, thou hast
kept me
{\ADD{to be}}
head of the
heathen: a
people
{\ADD{which}} I
knew not shall
serve me.
\VS{45}Strangers shall
submit themselves unto me: as soon as they
hear, they shall be
obedient unto me.\FTNT{‡‡}{{\FR 22.45 }
Strangers: Heb. Sons of the stranger}\FTNT{§§}{{\FR 22.45 }
submit…: or, yield feigned obedience: Heb. lie}
\VS{46}Strangers shall fade
away, and they shall be
afraid out of their close
places.
\VS{47}The
{\ND{
LORD}}
liveth; and
blessed
{\ADD{be}} my
rock; and
exalted be the
God of the
rock of my
salvation.
\VS{48}It
{\ADD{is}}
God that
avengeth me, and that bringeth
down the
people under me,\FTNT{***}{{\FR 22.48 }
avengeth: Heb. giveth avengement for}
\VS{49}And that bringeth me
forth from mine
enemies: thou also hast lifted me up on
high above them that rose
up against me: thou hast
delivered me from the
violent
man.
\VS{50}Therefore I will give
thanks unto thee, O
{\ND{
LORD}}, among the
heathen, and I will sing
praises unto thy
name.
\VS{51}{\ADD{He is}} the
tower of
salvation for his
king: and
sheweth
mercy to his
anointed, unto
David, and to his
seed
for
evermore.

\par }\Chap{23}{\PP \VerseOne{1}Now these
{\ADD{be}} the
last
words of
David.
David the
son of
Jesse
said, and the
man
{\ADD{who was}} raised
up on
high, the
anointed of the
God of
Jacob, and the
sweet
psalmist of
Israel,
said,
\VS{2}The
Spirit of the
{\ND{
LORD}}
spake by me, and his
word
{\ADD{was}} in my
tongue.
\VS{3}The
God of
Israel
said, the
Rock of
Israel
spake to me, He that
ruleth over
men
{\ADD{must be}}
just,
ruling in the
fear of
God.\FTNT{*}{{\FR 23.3 }
He…: or, Be thou ruler, etc}
\VS{4}And
{\ADD{he shall be}} as the
light of the
morning,
{\ADD{when}} the
sun
riseth,
{\ADD{even}} a
morning
without
clouds;
{\ADD{as}} the tender
grass
{\ADD{springing}} out of the
earth by clear
shining after
rain.
\VS{5}Although my
house
{\ADD{be}} not so with
God; yet he hath
made with me an
everlasting
covenant,
ordered in all
{\ADD{things}}, and
sure: for
{\ADD{this is}} all my
salvation, and all
{\ADD{my}}
desire, although he make
{\ADD{it}} not to
grow.
\par }{\PP \VS{6}¶ But
{\ADD{the sons}} of
Belial
{\ADD{shall be}} all of them as
thorns thrust
away, because they cannot be
taken with
hands:
\VS{7}But the
man
{\ADD{that}} shall
touch them must be
fenced with
iron and the
staff of a
spear; and they shall be
utterly
burned with
fire in the
{\ADD{same}}
place.\FTNT{†}{{\FR 23.7 }
fenced: Heb. filled}
\par }{\PP \VS{8}¶ These
{\ADD{be}} the
names of the mighty
men whom
David had: The
Tachmonite that
sat in the
seat,
chief among the
captains; the same
{\ADD{was}}
Adino the
Eznite:
{\ADD{he lift up his spear}} against
eight
hundred, whom he
slew at
one
time.\FTNT{‡}{{\FR 23.8 }
The Tachmonite…: or, Joshebbassebet the Tachmonite, head of the three}\FTNT{§}{{\FR 23.8 }
whom he…: Heb. slain}\FTNT{**}{{\FR 23.8 }
1ch 11:11 he lift…: from whom he…: Heb. slain}
\VS{9}And
after him
{\ADD{was}}
Eleazar the
son of
Dodo the
Ahohite,
{\ADD{one}} of the
three mighty
men with
David, when they
defied the
Philistines
{\ADD{that}} were there gathered
together to
battle, and the
men of
Israel were gone
away:
\VS{10}He
arose, and
smote the
Philistines
until his
hand was
weary, and his
hand
clave unto the
sword: and the
{\ND{
LORD}}
wrought a
great
victory that
day; and the
people
returned
after him only to
spoil.
\VS{11}And
after him
{\ADD{was}}
Shammah the
son of
Agee the
Hararite. And the
Philistines were gathered
together into a
troop, where was a
piece of
ground
full of
lentiles: and the
people
fled
from the
Philistines.\FTNT{††}{{\FR 23.11 }
into…: or, for foraging}
\VS{12}But he
stood in the
midst of the
ground, and
defended it, and
slew the
Philistines: and the
{\ND{
LORD}}
wrought a
great
victory.
\VS{13}And
three of the
thirty
chief went
down, and
came to
David in the harvest
time unto the
cave of
Adullam: and the
troop of the
Philistines
pitched in the
valley of
Rephaim.\FTNT{‡‡}{{\FR 23.13 }
three…: or, the three captains over the thirty}
\VS{14}And
David
{\ADD{was}} then in an
hold, and the
garrison of the
Philistines
{\ADD{was}} then
{\ADD{in}}
Beth-lehem.
\VS{15}And
David
longed, and
said, Oh that one would give me
drink of the
water of the
well of
Beth-lehem, which
{\ADD{is}} by the
gate!
\VS{16}And the
three mighty
men brake
through the
host of the
Philistines, and
drew
water out of the
well of
Beth-lehem, that
{\ADD{was}} by the
gate, and
took
{\ADD{it}}, and
brought
{\ADD{it}} to
David: nevertheless he
would not
drink thereof, but poured it
out unto the
{\ND{
LORD}}.
\VS{17}And he
said, Be it
far from me, O
{\ND{
LORD}}, that I should
do this:
{\ADD{is not this}} the
blood of the
men that
went in jeopardy of their
lives? therefore he
would not
drink it. These things
did these
three mighty
men.
\VS{18}And
Abishai, the
brother of
Joab, the
son of
Zeruiah, was
chief among
three. And he lifted
up his
spear against
three
hundred,
{\ADD{and}}
slew
{\ADD{them}}, and had the
name among
three.\FTNT{§§}{{\FR 23.18 }
and slew them: Heb. slain}
\VS{19}Was he not most
honourable of
three? therefore he was their
captain: howbeit he
attained not unto the
{\ADD{first}}
three.
\VS{20}And
Benaiah the
son of
Jehoiada, the
son of a
valiant
man, of
Kabzeel, who had done
many
acts, he
slew
two lionlike
men of
Moab: he went
down also and
slew a
lion in the
midst of a
pit in
time of
snow:\FTNT{***}{{\FR 23.20 }
who…: Heb. great of acts}\FTNT{†††}{{\FR 23.20 }
lionlike…: Heb. lion of God}
\VS{21}And he
slew an
Egyptian, a
goodly
man: and the
Egyptian had a
spear in his
hand; but he went
down to him with a
staff, and
plucked the
spear out of the
Egyptian’s
hand, and
slew him with his own
spear.\FTNT{‡‡‡}{{\FR 23.21 }
a goodly…: Heb. a man of countenance, or, sight: also called, a man of great stature}
\VS{22}These
{\ADD{things}}
did
Benaiah the
son of
Jehoiada, and had the
name among
three mighty
men.
\VS{23}He was more
honourable than the
thirty, but he
attained not to the
{\ADD{first}}
three. And
David
set him over his
guard.\FTNT{§§§}{{\FR 23.23 }
more…: or, honourable among the thirty}\FTNT{*}{{\FR 23.23 }
guard: or, council: Heb. at his command}
\VS{24}Asahel the
brother of
Joab
{\ADD{was}} one of the
thirty;
Elhanan the
son of
Dodo of
Beth-lehem,
\VS{25}Shammah the
Harodite,
Elika the
Harodite,
\VS{26}Helez the
Paltite,
Ira the
son of
Ikkesh the
Tekoite,
\VS{27}Abiezer the
Anethothite,
Mebunnai the
Hushathite,
\VS{28}Zalmon the
Ahohite,
Maharai the
Netophathite,
\VS{29}Heleb the
son of
Baanah, a
Netophathite,
Ittai the
son of
Ribai out of
Gibeah of the
children of
Benjamin,
\VS{30}Benaiah the
Pirathonite,
Hiddai of the
brooks of
Gaash,\FTNT{†}{{\FR 23.30 }
brooks: or, valleys}
\VS{31}Abi-albon the
Arbathite,
Azmaveth the
Barhumite,
\VS{32}Eliahba the
Shaalbonite, of the
sons of
Jashen,
Jonathan,
\VS{33}Shammah the
Hararite,
Ahiam the
son of
Sharar the
Hararite,
\VS{34}Eliphelet the
son of
Ahasbai, the
son of the
Maachathite,
Eliam the
son of
Ahithophel the
Gilonite,
\VS{35}Hezrai the
Carmelite,
Paarai the
Arbite,
\VS{36}Igal the
son of
Nathan of
Zobah,
Bani the
Gadite,
\VS{37}Zelek the
Ammonite,
Naharai the
Beerothite,
armourbearer to
Joab the
son of
Zeruiah,
\VS{38}Ira an
Ithrite,
Gareb an
Ithrite,
\VS{39}Uriah the
Hittite:
thirty and
seven in all.

\par }\Chap{24}{\PP \VerseOne{1}And
again the
anger of the
{\ND{
LORD}} was
kindled against
Israel, and he
moved
David against them to
say,
Go,
number
Israel and
Judah.
\VS{2}For the
king
said to
Joab the
captain of the
host, which
{\ADD{was}} with him, Go now
through all the
tribes of
Israel, from
Dan even to
Beer-sheba, and
number ye the
people, that I may
know the
number of the
people.\FTNT{*}{{\FR 24.2 }
Go: or, Compass}
\VS{3}And
Joab
said unto the
king, Now the
{\ND{
LORD}} thy
God
add unto the
people, how many
soever they be, an
hundredfold, and that the
eyes of my
lord the
king may
see
{\ADD{it:}} but why doth my
lord the
king
delight in this
thing?
\VS{4}Notwithstanding the
king’s
word
prevailed against
Joab, and against the
captains of the
host. And
Joab and the
captains of the
host went
out from the
presence of the
king, to
number the
people of
Israel.
\par }{\PP \VS{5}¶ And they passed
over
Jordan, and
pitched in
Aroer, on the right
side of the
city that
{\ADD{lieth}} in the
midst of the
river of
Gad, and toward
Jazer:\FTNT{†}{{\FR 24.5 }
river: or, valley}
\VS{6}Then they
came to
Gilead, and to the
land of
Tahtim-hodshi; and they
came to
Dan-jaan, and
about to
Zidon,\FTNT{‡}{{\FR 24.6 }
land of Tahtim-hodshi: or, nether land newly inhabited}
\VS{7}And
came to the strong
hold of
Tyre, and to all the
cities of the
Hivites, and of the
Canaanites: and they went
out to the
south of
Judah,
{\ADD{even}} to
Beer-sheba.
\VS{8}So when they had
gone through all the
land, they
came to
Jerusalem at the
end of
nine
months and
twenty
days.
\VS{9}And
Joab gave
up the
sum of the
number of the
people unto the
king: and there were in
Israel
eight
hundred
thousand
valiant
men that
drew the
sword; and the
men of
Judah
{\ADD{were}}
five
hundred
thousand
men.
\par }{\PP \VS{10}¶ And
David’s
heart
smote him
after that he had
numbered the
people. And
David
said unto the
{\ND{
LORD}}, I have
sinned
greatly in that I have
done: and now, I beseech thee, O
{\ND{
LORD}}, take
away the
iniquity of thy
servant; for I have done
very
foolishly.
\VS{11}For when
David was
up in the
morning, the
word of the
{\ND{
LORD}} came unto the
prophet
Gad,
David’s
seer,
saying,
\VS{12}Go and
say unto
David, Thus
saith the
{\ND{
LORD}}, I
offer thee
three
{\ADD{things;}}
choose thee
one of them, that I may
{\ADD{do it}} unto
thee.
\VS{13}So
Gad
came to
David, and
told him, and
said unto him, Shall
seven
years of
famine
come unto thee in thy
land? or wilt thou
flee
three
months
before thine
enemies, while they
pursue thee? or that there be
three
days’
pestilence in thy
land? now
advise, and
see what
answer I shall
return to him that
sent me.
\VS{14}And
David
said unto
Gad, I am in a
great
strait: let us
fall now into the
hand of the
{\ND{
LORD}}; for his
mercies
{\ADD{are}}
great: and let me not
fall into the
hand of
man.\FTNT{§}{{\FR 24.14 }
are great: or, are many}
\par }{\PP \VS{15}¶ So the
{\ND{
LORD}}
sent a
pestilence upon
Israel from the
morning even to the
time
appointed: and there
died of the
people from
Dan even to
Beer-sheba
seventy
thousand
men.
\VS{16}And when the
angel stretched
out his
hand upon
Jerusalem to
destroy it, the
{\ND{
LORD}}
repented him of the
evil, and
said to the
angel that
destroyed the
people, It is
enough:
stay now thine
hand. And the
angel of the
{\ND{
LORD}} was by the
threshingplace of
Araunah the
Jebusite.\FTNT{**}{{\FR 24.16 }
Araunah: also called, Ornan}
\VS{17}And
David
spake unto the
{\ND{
LORD}} when he
saw the
angel that
smote the
people, and
said, Lo, I have
sinned, and I have done
wickedly: but these
sheep, what have they
done? let thine
hand, I pray thee, be against me, and against my
father’s
house.
\par }{\PP \VS{18}¶ And
Gad
came that
day to
David, and
said unto him, Go
up,
rear an
altar unto the
{\ND{
LORD}} in the
threshingfloor of
Araunah the
Jebusite.\FTNT{††}{{\FR 24.18 }
Araunah: Heb. Araniah}
\VS{19}And
David, according to the
saying of
Gad, went
up as the
{\ND{
LORD}}
commanded.
\VS{20}And
Araunah
looked, and
saw the
king and his
servants coming
on toward him: and
Araunah went
out, and
bowed himself before the
king on his
face upon the
ground.
\VS{21}And
Araunah
said, Wherefore is my
lord the
king
come to his
servant? And
David
said, To
buy the
threshingfloor of thee, to
build an
altar unto the
{\ND{
LORD}}, that the
plague may be
stayed from the
people.
\VS{22}And
Araunah
said unto
David, Let my
lord the
king
take and offer
up what
{\ADD{seemeth}}
good unto
him:
behold,
{\ADD{here be}}
oxen for burnt
sacrifice, and threshing
instruments and
{\ADD{other}}
instruments of the
oxen for
wood.
\VS{23}All these
{\ADD{things}} did
Araunah,
{\ADD{as}} a
king,
give unto the
king. And
Araunah
said unto the
king, The
{\ND{
LORD}} thy
God
accept thee.
\VS{24}And the
king
said unto
Araunah, Nay; but I will
surely
buy
{\ADD{it}} of thee at a
price: neither will I
offer burnt
offerings unto the
{\ND{
LORD}} my
God of that which doth cost me
nothing. So
David
bought the
threshingfloor and the
oxen for
fifty
shekels of
silver.
\VS{25}And
David
built there an
altar unto the
{\ND{
LORD}}, and
offered burnt
offerings and peace
offerings. So the
{\ND{
LORD}} was
intreated for the
land, and the
plague was
stayed from
Israel.
\par }