\NormalFont\ShortTitle{2 Timothy}
{\MT THE SECOND EPISTLE OF PAUL THE APOSTLE TO TIMOTHY

\par }\ChapOne{1}{\PP \VerseOne{1}Paul, an
apostle of
Jesus
Christ
by the
will of
God, according
to the
promise of
life which
is
in
Christ
Jesus,
\VS{2}To
Timothy,
{\ADD{my}} dearly
beloved
son:
Grace,
mercy,
{\ADD{and}}
peace,
from
God the
Father
and
Christ
Jesus
our
Lord.
\VS{3}I
thank
God,
whom I
serve
from
{\ADD{my}}
forefathers
with
pure
conscience,
that without
ceasing I
have
remembrance
of
thee
in
my
prayers
night
and
day;
\VS{4}Greatly
desiring to
see
thee, being
mindful of
thy
tears,
that I may be
filled with
joy;
\VS{5}When I
call to
remembrance the
unfeigned
faith that
is
in
thee,
which
dwelt
first
in
thy
grandmother
Lois,
and
thy
mother
Eunice;
and I am
persuaded
that
in
thee
also.
\VS{6}Wherefore I
put
thee in
remembrance that thou stir
up the
gift of
God,
which
is
in
thee
by the putting
on of
my
hands.
\VS{7}For
God
hath
not
given
us the
spirit of
fear;
but of
power,
and of
love,
and of a sound
mind.
\VS{8}Be
not
thou
therefore ashamed
of the
testimony of
our
Lord,
nor of
me
his
prisoner:
but be thou partaker of the
afflictions of the
gospel according
to the
power of
God;
\VS{9}Who hath
saved
us,
and
called
{\ADD{us}} with an
holy
calling,
not according
to
our
works,
but according
to his
own
purpose
and
grace, which was
given
us
in
Christ
Jesus
before the
world
began,
\VS{10}But
is
now made
manifest
by the
appearing of
our
Saviour
Jesus
Christ, who hath
abolished
death,
and hath
brought
life
and
immortality to
light
through the
gospel:
\VS{11}Where
unto
I am
appointed a
preacher,
and an
apostle,
and a
teacher of the
Gentiles.
\VS{12}For the
which cause
I
also suffer these
things:
nevertheless I
am
not
ashamed:
for I
know
whom I have
believed,
and am
persuaded
that he
is
able to
keep that
which
I have
committed unto him
against
that
day.
\VS{13}Hold
fast the
form of
sound
words,
which thou hast
heard
of
me,
in
faith
and
love which
is
in
Christ
Jesus.
\VS{14}That good
thing which was committed unto
thee
keep
by the
Holy
Ghost which
dwelleth
in
us.
\VS{15}This thou
knowest,
that
all they which
are
in
Asia be turned away
from
me; of
whom
are
Phygellus
and
Hermogenes.
\VS{16}The
Lord
give
mercy unto the
house of
Onesiphorus;
for
he
oft
refreshed
me,
and
was
not
ashamed of
my
chain:
\VS{17}But, when he
was
in
Rome, he
sought
me
out very
diligently,
and
found
{\ADD{me}}.
\VS{18}The
Lord
grant unto
him that he may
find
mercy
of the
Lord
in
that
day:
and in how many
things he
ministered unto me
at
Ephesus,
thou
knowest very
well.

\par }\Chap{2}{\PP \VerseOne{1}Thou
therefore,
my
son, be
strong
in the
grace that
is
in
Christ
Jesus.
\VS{2}And the things
that thou hast
heard
of
me
among
many
witnesses, the
same commit
thou to
faithful
men,
who shall
be able
to
teach
others
also.
\VS{3}Thou
therefore endure
hardness,
as a
good
soldier of
Jesus
Christ.
\VS{4}No
man that
warreth entangleth
himself with the affairs
of
{\ADD{this}}
life;
that he may
please him who hath chosen him to be a
soldier.
\VS{5}And
if a
man
also strive for
masteries,
{\ADD{yet}} is
he
not
crowned,
except he
strive
lawfully.
\VS{6}The
husbandman that
laboureth
must
be
first
partaker of the
fruits.
\VS{7}Consider
what I
say;
and the
Lord
give
thee
understanding
in all
things.
\VS{8}Remember
that
Jesus
Christ
of the
seed of
David was
raised
from the
dead according
to
my
gospel:
\VS{9}Where
in I suffer
trouble,
as an evil
doer,
{\ADD{even}}
unto
bonds;
but the
word of
God
is
not
bound.
\VS{10}Therefore I
endure all
things
for the
elect’s
sakes,
that
they
may
also
obtain the
salvation which
is
in
Christ
Jesus
with
eternal
glory.
\VS{11}{\ADD{It is}} a
faithful
saying:
For
if we be dead
with
{\ADD{him}}, we
shall
also live
with
{\ADD{him}}:
\VS{12}If we
suffer, we
shall
also reign
with
{\ADD{him}}:
if we
deny
{\ADD{him}}, he
also will
deny
us:
\VS{13}If we believe
not,
{\ADD{yet}}
he
abideth
faithful: he
can
not
deny
himself.
\VS{14}Of these
things
put
{\ADD{them}} in
remembrance,
charging
{\ADD{them}}
before the
Lord that they
strive
not about
words
to
no
profit,
{\ADD{but}}
to the
subverting of the
hearers.
\VS{15}Study to
shew
thyself
approved unto
God, a
workman that needeth not to be
ashamed, rightly
dividing the
word of
truth.
\VS{16}But
shun
profane
{\ADD{and}} vain
babblings:
for they will
increase
unto
more
ungodliness.
\VS{17}And
their
word
will
eat as
doth a
canker: of
whom
is
Hymenaeus
and
Philetus;
\VS{18}Who
concerning the
truth have
erred,
saying
that the
resurrection is
past
already;
and
overthrow the
faith of
some.
\VS{19}Nevertheless the
foundation of
God
standeth
sure,
having
this
seal, The
Lord
knoweth them that
are
his.
And,
Let every
one that
nameth the
name of
Christ
depart
from
iniquity.
\VS{20}But
in a
great
house there
are
not
only
vessels of
gold
and of
silver,
but
also of
wood
and of
earth;
and
some
to
honour,
and
some
to
dishonour.
\VS{21}If a
man
therefore
purge
himself
from
these, he shall
be a
vessel
unto
honour,
sanctified,
and meet
for the
master’s
use,
{\ADD{and}}
prepared
unto
every
good
work.
\VS{22}Flee
also
youthful
lusts:
but
follow
righteousness,
faith,
charity,
peace,
with them that call
on the
Lord out
of a
pure
heart.
\VS{23}But
foolish
and
unlearned
questions
avoid,
knowing
that they do
gender
strifes.
\VS{24}And the
servant of the
Lord
must
not
strive;
but
be
gentle
unto
all
{\ADD{men}}, apt to
teach,
patient,
\VS{25}In
meekness
instructing those that oppose
themselves;
if
God
peradventure will
give
them
repentance
to the
acknowledging of the
truth;
\VS{26}And
{\ADD{that}} they may recover
themselves out
of the
snare of the
devil, who are taken
captive
by
him
at
his
will.

\par }\Chap{3}{\PP \VerseOne{1}This
know
also,
that
in the
last
days
perilous
times shall
come.
\VS{2}For
men shall
be lovers of their own
selves,
covetous,
boasters,
proud,
blasphemers,
disobedient to
parents,
unthankful,
unholy,
\VS{3}Without natural
affection,
trucebreakers, false
accusers,
incontinent,
fierce, despisers of those that are
good,
\VS{4}Traitors,
heady,
highminded, lovers of
pleasures
more
than lovers of
God;
\VS{5}Having a
form of
godliness,
but
denying the
power
thereof:
from
such turn
away.
\VS{6}For
of this
sort
are they which
creep
into
houses,
and lead
captive silly
women
laden with
sins, led
away with
divers
lusts,
\VS{7}Ever
learning,
and
never
able to
come
to the
knowledge of the
truth.
\VS{8}Now
as
Jannes
and
Jambres
withstood
Moses,
so
do
these
also
resist the
truth:
men of
corrupt
minds,
reprobate
concerning the
faith.
\VS{9}But they shall
proceed
no
further:
for
their
folly shall
be
manifest unto
all
{\ADD{men}},
as
theirs
also
was.
\VS{10}But
thou hast fully
known
my
doctrine, manner of
life,
purpose,
faith,
longsuffering,
charity,
patience,
\VS{11}Persecutions,
afflictions,
which
came unto
me
at
Antioch,
at
Iconium,
at
Lystra;
what
persecutions I
endured:
but out
of
{\ADD{them}}
all the
Lord
delivered
me.
\VS{12}Yea,
and
all that
will
live
godly
in
Christ
Jesus shall suffer
persecution.
\VS{13}But
evil
men
and
seducers shall
wax worse and
worse,
deceiving,
and being
deceived.
\VS{14}But
continue
thou
in the things
which thou hast
learned
and hast been assured
of,
knowing
of
whom thou hast
learned
{\ADD{them}};
\VS{15}And
that
from a
child thou hast
known the
holy
scriptures, which are
able to
make
thee
wise
unto
salvation
through
faith which
is
in
Christ
Jesus.
\VS{16}All
scripture
{\ADD{is}} given by inspiration of
God,
and
{\ADD{is}}
profitable
for
doctrine,
for
reproof,
for
correction,
for
instruction
in
righteousness:
\VS{17}That the
man of
God may
be
perfect, throughly
furnished
unto
all
good
works.

\par }\Chap{4}{\PP \VerseOne{1}I
charge
{\ADD{thee}}
therefore
before
God,
and the
Lord
Jesus
Christ, who
shall
judge the
quick
and the
dead
at
his
appearing
and
his
kingdom;
\VS{2}Preach the
word; be
instant in
season, out of
season;
reprove,
rebuke,
exhort
with
all
longsuffering
and
doctrine.
\VS{3}For the
time will
come
when they
will
not
endure
sound
doctrine;
but
after their
own
lusts shall they
heap to
themselves
teachers, having
itching
ears;
\VS{4}And they shall turn
away
{\ADD{their}}
ears
from the
truth,
and shall be
turned
unto
fables.
\VS{5}But
watch
thou
in all
things, endure
afflictions,
do the
work of an
evangelist, make full
proof of
thy
ministry.
\VS{6}For
I
am
now ready to be
offered,
and the
time of
my
departure is at
hand.
\VS{7}I have
fought a
good
fight, I have
finished
{\ADD{my}}
course, I have
kept the
faith:
\VS{8}Henceforth there is laid
up for
me a
crown of
righteousness,
which the
Lord, the
righteous
judge, shall
give
me
at
that
day:
and
not to
me
only,
but unto
all
them
also that
love
his
appearing.
\par }{\PP \VS{9}Do thy
diligence to
come
shortly
unto
me:
\VS{10}For
Demas hath
forsaken
me, having
loved
this
present
world,
and is
departed
unto
Thessalonica;
Crescens
to
Galatia,
Titus
unto
Dalmatia.
\VS{11}Only
Luke
is
with
me.
Take
Mark,
and
bring him
with
thee:
for he
is
profitable to
me
for the
ministry.
\VS{12}And
Tychicus have I
sent
to
Ephesus.
\VS{13}The
cloke
that I
left
at
Troas
with
Carpus, when thou
comest,
bring
{\ADD{with thee}},
and the
books,
{\ADD{but}}
especially the
parchments.
\VS{14}Alexander the
coppersmith
did
me
much
evil: the
Lord
reward
him according
to
his
works:
\VS{15}Of
whom be
thou
ware
also;
for he
hath
greatly
withstood
our
words.
\VS{16}At
my
first
answer no
man stood
with
me,
but
all
{\ADD{men}}
forsook
me:
{\ADD{I pray God}} that it
may
not be
laid to their
charge.
\VS{17}Notwithstanding the
Lord
stood with
me,
and
strengthened
me;
that
by
me the
preaching might be fully
known,
and
{\ADD{that}}
all the
Gentiles might
hear:
and I was
delivered out
of the
mouth of the
lion.
\VS{18}And the
Lord shall
deliver
me
from
every
evil
work,
and will
preserve
{\ADD{me}}
unto
his
heavenly
kingdom: to
whom
{\ADD{be}}
glory
for
ever and
ever.
Amen.
\par }{\PP \VS{19}Salute
Prisca
and
Aquila,
and the
household of
Onesiphorus.
\VS{20}Erastus
abode
at
Corinth:
but
Trophimus have I
left
at
Miletum
sick.
\VS{21}Do thy
diligence to
come
before
winter.
Eubulus
greeteth
thee,
and
Pudens,
and
Linus,
and
Claudia,
and
all the
brethren.
\VS{22}The
Lord
Jesus
Christ
{\ADD{be}}
with
thy
spirit.
Grace
{\ADD{be}}
with
you.
Amen.
\par }{\SH The second
{\ADD{epistle}} unto Timotheus, ordained the first bishop of the church of the Ephesians, was written from Rome, when Paul was brought before Nero the second time.
\par }