\NormalFont\ShortTitle{Amos}
{\MT Amos

\par }\ChapOne{1}{\PP \VerseOne{1}The
words of
Amos, who was among the
herdmen of
Tekoa, which he
saw concerning
Israel in the
days of
Uzziah
king of
Judah, and in the
days of
Jeroboam the
son of
Joash
king of
Israel, two
years
before the
earthquake.
\VS{2}And he
said, The
{\ND{
LORD}} will
roar from
Zion, and
utter his
voice from
Jerusalem; and the
habitations of the
shepherds shall
mourn, and the
top of
Carmel shall
wither.
\par }{\PP \VS{3}Thus
saith the
{\ND{
LORD}}; For
three
transgressions of
Damascus, and for
four, I will not turn
away
{\ADD{the punishment}} thereof; because they have
threshed
Gilead with threshing
instruments of
iron:\FTNT{*}{{\FR 1.3 }
and…: or, yea, for four}\FTNT{†}{{\FR 1.3 }
turn…: or, convert it, or, let it be quiet}
\VS{4}But I will
send a
fire into the
house of
Hazael, which shall
devour the
palaces of
Ben-hadad.
\VS{5}I will
break also the
bar of
Damascus, and cut
off the
inhabitant from the
plain of
Aven, and him that
holdeth the
sceptre from the
house of
Eden: and the
people of
Syria shall go into
captivity unto
Kir,
saith the
{\ND{
LORD}}.\FTNT{‡}{{\FR 1.5 }
the plain…: or, Bikathaven}\FTNT{§}{{\FR 1.5 }
the house…: or, Beth-eden}
\par }{\PP \VS{6}¶ Thus
saith the
{\ND{
LORD}}; For
three
transgressions of
Gaza, and for
four, I will not turn
away
{\ADD{the punishment}} thereof; because they carried away
captive the
whole
captivity, to deliver
{\ADD{them}}
up to
Edom:\FTNT{**}{{\FR 1.6 }
away captive…: or, them away with an entire}
\VS{7}But I will
send a
fire on the
wall of
Gaza, which shall
devour the
palaces thereof:
\VS{8}And I will cut
off the
inhabitant from
Ashdod, and him that
holdeth the
sceptre from
Ashkelon, and I will
turn mine
hand against
Ekron: and the
remnant of the
Philistines shall
perish,
saith the
Lord
{\ND{
GOD}}.
\par }{\PP \VS{9}¶ Thus
saith the
{\ND{
LORD}}; For
three
transgressions of
Tyrus, and for
four, I will not turn
away
{\ADD{the punishment}} thereof; because they delivered
up the
whole
captivity to
Edom, and
remembered not the
brotherly
covenant:\FTNT{††}{{\FR 1.9 }
the brotherly…: Heb. the covenant of brethren}
\VS{10}But I will
send a
fire on the
wall of
Tyrus, which shall
devour the
palaces thereof.
\par }{\PP \VS{11}¶ Thus
saith the
{\ND{
LORD}}; For
three
transgressions of
Edom, and for
four, I will not turn
away
{\ADD{the punishment}} thereof; because he did
pursue his
brother with the
sword, and did cast
off all
pity, and his
anger did
tear
perpetually, and he
kept his
wrath for
ever:\FTNT{‡‡}{{\FR 1.11 }
did cast…: Heb. corrupted his compassions}
\VS{12}But I will
send a
fire upon
Teman, which shall
devour the
palaces of
Bozrah.
\par }{\PP \VS{13}¶ Thus
saith the
{\ND{
LORD}}; For
three
transgressions of the
children of
Ammon, and for
four, I will not turn
away
{\ADD{the punishment}} thereof; because they have ripped
up the women with
child of
Gilead, that they might
enlarge their
border:\FTNT{§§}{{\FR 1.13 }
ripped…: or, divided the mountains}
\VS{14}But I will
kindle a
fire in the
wall of
Rabbah, and it shall
devour the
palaces thereof, with
shouting in the
day of
battle, with a
tempest in the
day of the
whirlwind:
\VS{15}And their
king shall
go into
captivity, he and his
princes
together,
saith the
{\ND{
LORD}}.

\par }\Chap{2}{\PP \VerseOne{1}Thus
saith the
{\ND{
LORD}}; For
three
transgressions of
Moab, and for
four, I will not turn
away
{\ADD{the punishment}} thereof; because he
burned the
bones of the
king of
Edom into
lime:
\VS{2}But I will
send a
fire upon
Moab, and it shall
devour the
palaces of
Kerioth: and
Moab shall
die with
tumult, with
shouting,
{\ADD{and}} with the
sound of the
trumpet:
\VS{3}And I will cut
off the
judge from the
midst thereof, and will
slay all the
princes thereof with him,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{4}¶ Thus
saith the
{\ND{
LORD}}; For
three
transgressions of
Judah, and for
four, I will not turn
away
{\ADD{the punishment}} thereof; because they have
despised the
law of the
{\ND{
LORD}}, and have not
kept his
commandments, and their
lies caused them to
err,
after the which their
fathers have
walked:
\VS{5}But I will
send a
fire upon
Judah, and it shall
devour the
palaces of
Jerusalem.
\par }{\PP \VS{6}¶ Thus
saith the
{\ND{
LORD}}; For
three
transgressions of
Israel, and for
four, I will not turn
away
{\ADD{the punishment}} thereof; because they
sold the
righteous for
silver, and the
poor for a pair of
shoes;
\VS{7}That
pant after the
dust of the
earth on the
head of the
poor, and turn
aside the
way of the
meek: and a
man and his
father will
go in unto the
{\ADD{same}}
maid, to
profane my
holy
name:\FTNT{*}{{\FR 2.7 }
maid: or, young woman}
\VS{8}And they lay
{\ADD{themselves}}
down upon
clothes laid to
pledge
by every
altar, and they
drink the
wine of the
condemned
{\ADD{in}} the
house of their
god.\FTNT{†}{{\FR 2.8 }
the condemned: or, such as have fined, or, mulcted}
\par }{\PP \VS{9}¶ Yet
destroyed I the
Amorite
before them, whose
height
{\ADD{was}} like the
height of the
cedars, and he
{\ADD{was}}
strong as the
oaks; yet I
destroyed his
fruit from
above, and his
roots from beneath.
\VS{10}Also I brought you
up from the
land of
Egypt, and
led you
forty
years through the
wilderness, to
possess the
land of the
Amorite.
\VS{11}And I raised
up of your
sons for
prophets, and of your young
men for
Nazarites.
{\ADD{Is it}} not even thus, O ye
children of
Israel?
saith the
{\ND{
LORD}}.
\VS{12}But ye gave the
Nazarites
wine to
drink; and
commanded the
prophets,
saying,
Prophesy not.
\VS{13}Behold, I am
pressed under you, as a
cart is
pressed
{\ADD{that is}}
full of
sheaves.\FTNT{‡}{{\FR 2.13 }
I am…: or, I will press your place, as a cart full of sheaves presseth}
\VS{14}Therefore the
flight shall
perish from the
swift, and the
strong shall not
strengthen his
force, neither shall the
mighty
deliver
himself:\FTNT{§}{{\FR 2.14 }
himself: Heb. his soul, or, life}
\VS{15}Neither shall he
stand that
handleth the
bow; and
{\ADD{he that is}}
swift of
foot shall not
deliver
{\ADD{himself:}} neither shall he that
rideth the
horse
deliver
himself.
\VS{16}And
{\ADD{he that is}}
courageous among the
mighty shall flee
away
naked in that
day,
saith the
{\ND{
LORD}}.\FTNT{**}{{\FR 2.16 }
courageous: Heb. strong of his heart}

\par }\Chap{3}{\PP \VerseOne{1}Hear this
word that the
{\ND{
LORD}} hath
spoken against you, O
children of
Israel, against the whole
family which I brought
up from the
land of
Egypt,
saying,
\VS{2}You only have I
known of all the
families of the
earth: therefore I will
punish you for all your
iniquities.\FTNT{*}{{\FR 3.2 }
punish: Heb. visit upon}
\VS{3}Can
two
walk
together,
except they be
agreed?
\VS{4}Will a
lion
roar in the
forest, when he hath no
prey? will a young
lion cry
out of his
den, if he have
taken
nothing?\FTNT{†}{{\FR 3.4 }
cry: Heb. give forth his voice}
\VS{5}Can a
bird
fall in a
snare upon the
earth, where no
gin
{\ADD{is}} for him? shall
{\ADD{one}} take
up a
snare from the
earth, and have
taken nothing at
all?
\VS{6}Shall a
trumpet be
blown in the
city, and the
people not be
afraid? shall there be
evil in a
city, and the
{\ND{
LORD}} hath not
done
{\ADD{it}}?\FTNT{‡}{{\FR 3.6 }
be afraid: or, run together?}\FTNT{§}{{\FR 3.6 }
the
{\ND{LORD}}…: or, shall not the
{\ND{LORD}} do somewhat?}
\VS{7}Surely the
Lord
{\ND{
GOD}} will
do
nothing, but he
revealeth his
secret unto his
servants the
prophets.
\VS{8}The
lion hath
roared, who will
not
fear? the
Lord
{\ND{
GOD}} hath
spoken, who can but
prophesy?
\par }{\PP \VS{9}¶
Publish in the
palaces at
Ashdod, and in the
palaces in the
land of
Egypt, and
say,
Assemble yourselves upon the
mountains of
Samaria, and
behold the
great
tumults in the
midst thereof, and the
oppressed in the
midst thereof.\FTNT{**}{{\FR 3.9 }
oppressed: or, oppressions}
\VS{10}For they
know not to
do
right,
saith the
{\ND{
LORD}}, who
store up
violence and
robbery in their
palaces.\FTNT{††}{{\FR 3.10 }
robbery: or, spoil}
\VS{11}Therefore thus
saith the
Lord
{\ND{
GOD}}; An
adversary
{\ADD{there shall be}} even round
about the
land; and he shall bring
down thy
strength from thee, and thy
palaces shall be
spoiled.
\VS{12}Thus
saith the
{\ND{
LORD}}; As the
shepherd taketh
out of the
mouth of the
lion
two
legs, or a
piece of an
ear; so shall the
children of
Israel be taken
out that
dwell in
Samaria in the
corner of a
bed, and in
Damascus
{\ADD{in}} a
couch.\FTNT{‡‡}{{\FR 3.12 }
taketh: Heb. delivereth}\FTNT{§§}{{\FR 3.12 }
in Damascus…: or, on the bed’s feet}
\VS{13}Hear ye, and
testify in the
house of
Jacob,
saith the
Lord
{\ND{
GOD}}, the
God of
hosts,
\VS{14}That in the
day that I shall
visit the
transgressions of
Israel upon him I will also
visit the
altars of
Beth-el: and the
horns of the
altar shall be cut
off, and
fall to the
ground.\FTNT{***}{{\FR 3.14 }
visit: or, punish Israel for}
\VS{15}And I will
smite the
winter
house with the
summer
house; and the
houses of
ivory shall
perish, and the
great
houses shall have an
end,
saith the
{\ND{
LORD}}.

\par }\Chap{4}{\PP \VerseOne{1}Hear this
word, ye
kine of
Bashan, that
{\ADD{are}} in the
mountain of
Samaria, which
oppress the
poor, which
crush the
needy, which
say to their
masters,
Bring, and let us
drink.
\VS{2}The
Lord
{\ND{
GOD}} hath
sworn by his
holiness, that, lo, the
days shall
come upon you, that he will take you
away with
hooks, and your
posterity with
fishhooks.
\VS{3}And ye shall go
out at the
breaches,
every
{\ADD{cow at that which is}} before her; and ye shall
cast
{\ADD{them}} into the
palace,
saith the
{\ND{
LORD}}.\FTNT{*}{{\FR 4.3 }
cast…: or, cast away the things of the palace}
\par }{\PP \VS{4}¶
Come to
Beth-el, and
transgress; at
Gilgal
multiply
transgression; and
bring your
sacrifices every
morning,
{\ADD{and}} your
tithes after
three
years:\FTNT{†}{{\FR 4.4 }
three…: Heb. three years of days}
\VS{5}And
offer a sacrifice of
thanksgiving with
leaven, and
proclaim
{\ADD{and}}
publish the free
offerings: for
this
liketh you, O ye
children of
Israel,
saith the
Lord
{\ND{
GOD}}.\FTNT{‡}{{\FR 4.5 }
offer: Heb. offer by burning}\FTNT{§}{{\FR 4.5 }
this…: Heb. so ye love}
\par }{\PP \VS{6}¶ And I also have
given you
cleanness of
teeth in all your
cities, and
want of
bread in all your
places: yet have ye not
returned unto me,
saith the
{\ND{
LORD}}.
\VS{7}And also I have
withholden the
rain from you, when
{\ADD{there were}} yet
three
months to the
harvest: and I caused it to
rain upon
one
city, and caused it not to
rain upon
another
city:
one
piece was
rained upon, and the
piece whereupon it
rained not
withered.
\VS{8}So
two
{\ADD{or}}
three
cities
wandered unto
one
city, to
drink
water; but they were not
satisfied: yet have ye not
returned unto me,
saith the
{\ND{
LORD}}.
\VS{9}I have
smitten you with
blasting and
mildew: when your
gardens and your
vineyards and your fig
trees and your olive
trees
increased, the
palmerworm
devoured
{\ADD{them:}} yet have ye not
returned unto me,
saith the
{\ND{
LORD}}.\FTNT{**}{{\FR 4.9 }
when…: or, the multitude of your gardens, etc. did the palmerworm}
\VS{10}I have
sent among you the
pestilence after the
manner of
Egypt: your young
men have I
slain with the
sword, and have taken
away your
horses; and I have made the
stink of your
camps to come
up unto your
nostrils: yet have ye not
returned unto me,
saith the
{\ND{
LORD}}.\FTNT{††}{{\FR 4.10 }
after…: or, in the way}\FTNT{‡‡}{{\FR 4.10 }
and have…: Heb. with the captivity of your horses}
\VS{11}I have
overthrown
{\ADD{some}} of you, as
God
overthrew
Sodom and
Gomorrah, and ye were as a
firebrand plucked
out of the
burning: yet have ye not
returned unto me,
saith the
{\ND{
LORD}}.
\VS{12}Therefore thus will I
do unto thee, O
Israel:
{\ADD{and}}
because I will
do this unto thee,
prepare to
meet thy
God, O
Israel.
\VS{13}For, lo, he that
formeth the
mountains, and
createth the
wind, and
declareth unto
man what
{\ADD{is}} his
thought, that
maketh the
morning
darkness, and
treadeth upon the high
places of the
earth, The
{\ND{
LORD}}, The
God of
hosts,
{\ADD{is}} his
name.\FTNT{§§}{{\FR 4.13 }
wind: or, spirit}

\par }\Chap{5}{\PP \VerseOne{1}Hear ye this
word which I take
up against you,
{\ADD{even}} a
lamentation, O
house of
Israel.
\VS{2}The
virgin of
Israel is
fallen; she shall no
more
rise: she is
forsaken upon her
land;
{\ADD{there is}} none to raise her
up.
\VS{3}For thus
saith the
Lord
{\ND{
GOD}}; The
city that went
out
{\ADD{by}} a
thousand shall
leave an
hundred, and that which went
forth
{\ADD{by}} an
hundred shall
leave
ten, to the
house of
Israel.
\par }{\PP \VS{4}¶ For thus
saith the
{\ND{
LORD}} unto the
house of
Israel,
Seek ye me, and ye shall
live:
\VS{5}But
seek not
Beth-el, nor
enter into
Gilgal, and
pass not to
Beer-sheba: for
Gilgal shall
surely go into
captivity, and
Beth-el shall come to
nought.
\VS{6}Seek the
{\ND{
LORD}}, and ye shall
live; lest he break
out like
fire in the
house of
Joseph, and
devour
{\ADD{it}}, and
{\ADD{there be}} none to
quench
{\ADD{it}} in
Beth-el.
\VS{7}Ye who
turn
judgment to
wormwood, and leave
off
righteousness in the
earth,
\VS{8}{\ADD{Seek him}} that
maketh the seven
stars and
Orion, and
turneth the shadow of
death into the
morning, and maketh the
day
dark with
night: that
calleth for the
waters of the
sea, and poureth them
out upon the
face of the
earth: The
{\ND{
LORD}}
{\ADD{is}} his
name:
\VS{9}That
strengtheneth the
spoiled against the
strong, so that the
spoiled shall
come against the
fortress.\FTNT{*}{{\FR 5.9 }
spoiled: Heb. spoil}
\VS{10}They
hate him that
rebuketh in the
gate, and they
abhor him that
speaketh
uprightly.
\VS{11}Forasmuch therefore as your
treading
{\ADD{is}} upon the
poor, and ye
take from him
burdens of
wheat: ye have
built
houses of hewn
stone, but ye shall not
dwell in them; ye have
planted
pleasant
vineyards, but ye shall not
drink
wine of them.\FTNT{†}{{\FR 5.11 }
pleasant…: Heb. vineyards of desire}
\VS{12}For I
know your
manifold
transgressions and your
mighty
sins: they
afflict the
just, they
take a
bribe, and they turn
aside the
poor in the
gate
{\ADD{from their right}}.\FTNT{‡}{{\FR 5.12 }
a bribe: or, a ransom}
\VS{13}Therefore the
prudent shall keep
silence in that
time; for it
{\ADD{is}} an
evil
time.
\VS{14}Seek
good, and not
evil, that ye may
live: and so the
{\ND{
LORD}}, the
God of
hosts, shall be with you, as ye have
spoken.
\VS{15}Hate the
evil, and
love the
good, and
establish
judgment in the
gate: it may be that the
{\ND{
LORD}}
God of
hosts will be
gracious unto the
remnant of
Joseph.
\par }{\PP \VS{16}Therefore the
{\ND{
LORD}}, the
God of
hosts, the
Lord,
saith thus;
Wailing
{\ADD{shall be}} in all
streets; and they shall
say in all the
highways,
Alas!
alas! and they shall
call the
husbandman to
mourning, and such as are
skilful of
lamentation to
wailing.
\VS{17}And in all
vineyards
{\ADD{shall be}}
wailing: for I will
pass
through thee,
saith the
{\ND{
LORD}}.
\VS{18}Woe unto you that
desire the
day of the
{\ND{
LORD}}! to what
end
{\ADD{is}} it for you? the
day of the
{\ND{
LORD}}
{\ADD{is}}
darkness, and not
light.
\VS{19}As if a
man did
flee
from a
lion, and a
bear
met him; or
went into the
house, and
leaned his
hand on the
wall, and a
serpent
bit him.
\VS{20}{\ADD{Shall}} not the
day of the
{\ND{
LORD}}
{\ADD{be}}
darkness, and not
light? even very
dark, and no
brightness in it?
\par }{\PP \VS{21}¶ I
hate, I
despise your feast
days, and I will not
smell in your solemn
assemblies.\FTNT{§}{{\FR 5.21 }
in…: or, your holy days}
\VS{22}Though ye
offer me burnt
offerings and your meat
offerings, I will not
accept
{\ADD{them:}} neither will I
regard the peace
offerings of your fat
beasts.\FTNT{**}{{\FR 5.22 }
peace…: or, thank offerings}
\VS{23}Take thou
away from me the
noise of thy
songs; for I will not
hear the
melody of thy
viols.
\VS{24}But let
judgment run
down as
waters, and
righteousness as a
mighty
stream.\FTNT{††}{{\FR 5.24 }
run: Heb. roll}
\VS{25}Have ye
offered unto me
sacrifices and
offerings in the
wilderness
forty
years, O
house of
Israel?
\VS{26}But ye have
borne the
tabernacle of your
Moloch and
Chiun your
images, the
star of your
god, which ye
made to yourselves.\FTNT{‡‡}{{\FR 5.26 }
the tabernacle…: or, Siccuth your king}
\VS{27}Therefore will I cause you to go into
captivity
beyond
Damascus,
saith the
{\ND{
LORD}}, whose
name
{\ADD{is}} The
God of
hosts.

\par }\Chap{6}{\PP \VerseOne{1}Woe to them
{\ADD{that are}} at
ease in
Zion, and
trust in the
mountain of
Samaria,
{\ADD{which are}}
named
chief of the
nations, to whom the
house of
Israel
came!\FTNT{*}{{\FR 6.1 }
are at…: or, are secure}\FTNT{†}{{\FR 6.1 }
chief: or, firstfruits}
\VS{2}Pass ye unto
Calneh, and
see; and from thence
go ye to
Hamath the
great: then go
down to
Gath of the
Philistines:
{\ADD{be they}}
better than these
kingdoms? or their
border
greater than your
border?
\VS{3}Ye that put far
away the
evil
day, and cause the
seat of
violence to come
near;\FTNT{‡}{{\FR 6.3 }
seat: or, habitation}
\VS{4}That
lie upon
beds of
ivory, and
stretch themselves upon their
couches, and
eat the
lambs out of the
flock, and the
calves out of the
midst of the
stall;\FTNT{§}{{\FR 6.4 }
stretch…: or, abound with superfluities}
\VS{5}That
chant to the
sound of the
viol,
{\ADD{and}}
invent to themselves
instruments of
musick, like
David;\FTNT{**}{{\FR 6.5 }
chant: or, quaver}
\VS{6}That
drink
wine in
bowls, and
anoint themselves with the
chief
ointments: but they are not
grieved for the
affliction of
Joseph.\FTNT{††}{{\FR 6.6 }
wine…: or, in bowls of wine}\FTNT{‡‡}{{\FR 6.6 }
affliction: Heb. breach}
\par }{\PP \VS{7}¶ Therefore now shall they go
captive with the
first that go
captive, and the
banquet of them that
stretched themselves shall be
removed.
\par }{\PP \VS{8}The
Lord
{\ND{
GOD}} hath
sworn by
himself,
saith the
{\ND{
LORD}} the
God of
hosts, I
abhor the
excellency of
Jacob, and
hate his
palaces: therefore will I deliver
up the
city with all that is
therein.\FTNT{§§}{{\FR 6.8 }
that…: Heb. the fulness thereof}
\VS{9}And it shall come to pass, if there
remain
ten
men in
one
house, that they shall
die.
\VS{10}And a
man’s uncle shall take him
up, and he that
burneth him, to bring
out the
bones out of the
house, and shall
say unto him that
{\ADD{is}} by the
sides of the
house,
{\ADD{Is there}} yet
{\ADD{any}} with thee? and he shall
say,
No. Then shall he
say, Hold thy
tongue: for we may not make
mention of the
name of the
{\ND{
LORD}}.\FTNT{***}{{\FR 6.10 }
we…: or, they will not, or, have not}
\VS{11}For, behold, the
{\ND{
LORD}}
commandeth, and he will
smite the
great
house with
breaches, and the
little
house with
clefts.\FTNT{†††}{{\FR 6.11 }
breaches: or, droppings}
\par }{\PP \VS{12}¶ Shall
horses
run upon the
rock? will
{\ADD{one}}
plow
{\ADD{there}} with
oxen? for ye have
turned
judgment into
gall, and the
fruit of
righteousness into
hemlock:
\VS{13}Ye which
rejoice in a
thing of
nought, which
say, Have we not
taken to us
horns by our own
strength?
\VS{14}But, behold, I will raise
up against you a
nation, O
house of
Israel,
saith the
{\ND{
LORD}} the
God of
hosts; and they shall
afflict you from the entering
in of
Hemath unto the
river of the
wilderness.\FTNT{‡‡‡}{{\FR 6.14 }
river: or, valley}

\par }\Chap{7}{\PP \VerseOne{1}Thus hath the
Lord
{\ND{
GOD}}
shewed unto me; and, behold, he
formed
grasshoppers in the
beginning of the shooting
up of the latter
growth; and, lo,
{\ADD{it was}} the latter
growth
after the
king’s
mowings.\FTNT{*}{{\FR 7.1 }
grasshoppers: or, green worms}
\VS{2}And it came to pass,
{\ADD{that}} when they had made an
end of
eating the
grass of the
land, then I
said, O
Lord
{\ND{
GOD}},
forgive, I beseech thee: by whom shall
Jacob
arise? for he
{\ADD{is}}
small.\FTNT{†}{{\FR 7.2 }
by…: or, who of (or, for,) Jacob shall stand?}
\VS{3}The
{\ND{
LORD}}
repented for this: It shall not be,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{4}¶ Thus hath the
Lord
{\ND{
GOD}}
shewed unto me: and, behold, the
Lord
{\ND{
GOD}}
called to
contend by
fire, and it
devoured the
great
deep, and did eat
up a
part.
\VS{5}Then
said I, O
Lord
{\ND{
GOD}},
cease, I beseech thee: by whom shall
Jacob
arise? for he
{\ADD{is}}
small.\FTNT{‡}{{\FR 7.5 }
by…: or, who of (or, for,) Jacob shall stand?}
\VS{6}The
{\ND{
LORD}}
repented for this: This also shall not be,
saith the
Lord
{\ND{
GOD}}.
\par }{\PP \VS{7}¶ Thus he
shewed me: and, behold, the
Lord
stood upon a
wall
{\ADD{made}} by a
plumbline, with a
plumbline in his
hand.
\VS{8}And the
{\ND{
LORD}}
said unto me,
Amos, what
seest thou? And I
said, A
plumbline. Then
said the
Lord, Behold, I will
set a
plumbline in the
midst of my
people
Israel: I will not
again pass
by them any more:
\VS{9}And the high
places of
Isaac shall be
desolate, and the
sanctuaries of
Israel shall be laid
waste; and I will
rise against the
house of
Jeroboam with the
sword.
\par }{\PP \VS{10}¶ Then
Amaziah the
priest of
Beth-el
sent to
Jeroboam
king of
Israel,
saying,
Amos hath
conspired against thee in the
midst of the
house of
Israel: the
land is not
able to
bear all his
words.
\VS{11}For thus
Amos
saith,
Jeroboam shall
die by the
sword, and
Israel shall
surely be led away
captive out of their own
land.
\VS{12}Also
Amaziah
said unto
Amos, O thou
seer,
go, flee thee
away into the
land of
Judah, and there
eat
bread, and
prophesy there:
\VS{13}But
prophesy not
again any more at
Beth-el: for it
{\ADD{is}} the
king’s
chapel, and it
{\ADD{is}} the
king’s
court.\FTNT{§}{{\FR 7.13 }
chapel: or, sanctuary}\FTNT{**}{{\FR 7.13 }
king’s court: Heb. house of the kingdom}
\par }{\PP \VS{14}¶ Then
answered
Amos, and
said to
Amaziah, I
{\ADD{was}} no
prophet, neither
{\ADD{was}} I a
prophet’s
son; but I
{\ADD{was}} an
herdman, and a
gatherer of sycomore
fruit:\FTNT{††}{{\FR 7.14 }
sycomore…: or, wild figs}
\VS{15}And the
{\ND{
LORD}}
took me as I
followed the
flock, and the
{\ND{
LORD}}
said unto me,
Go,
prophesy unto my
people
Israel.\FTNT{‡‡}{{\FR 7.15 }
as…: Heb. from behind}
\par }{\PP \VS{16}¶ Now therefore
hear thou the
word of the
{\ND{
LORD}}: Thou
sayest,
Prophesy not against
Israel, and
drop not
{\ADD{thy word}} against the
house of
Isaac.
\VS{17}Therefore thus
saith the
{\ND{
LORD}}; Thy
wife shall be an
harlot in the
city, and thy
sons and thy
daughters shall
fall by the
sword, and thy
land shall be
divided by
line; and thou shalt
die
in a
polluted
land: and
Israel shall
surely go into
captivity forth of his
land.

\par }\Chap{8}{\PP \VerseOne{1}Thus hath the
Lord
{\ND{
GOD}}
shewed unto me: and behold a
basket of summer
fruit.
\VS{2}And he
said,
Amos, what
seest thou? And I
said, A
basket of summer
fruit. Then
said the
{\ND{
LORD}} unto me, The
end is
come upon my
people of
Israel; I will not
again
pass by them any more.
\VS{3}And the
songs of the
temple shall be
howlings in that
day,
saith the
Lord
{\ND{
GOD}}:
{\ADD{there shall be}}
many dead
bodies in every
place; they shall cast
{\ADD{them}}
forth with
silence.\FTNT{*}{{\FR 8.3 }
shall be howlings: Heb. shall howl}\FTNT{†}{{\FR 8.3 }
with…: Heb. be silent}
\par }{\PP \VS{4}¶
Hear this, O ye that swallow
up the
needy, even to
make the
poor of the
land to
fail,
\VS{5}Saying, When will the new
moon be
gone, that we may
sell
corn? and the
sabbath, that we may set
forth
wheat, making the
ephah
small, and the
shekel
great, and
falsifying the
balances by
deceit?\FTNT{‡}{{\FR 8.5 }
new…: or, month}\FTNT{§}{{\FR 8.5 }
set…: Heb. open}\FTNT{**}{{\FR 8.5 }
falsifying…: Heb. perverting the balances of deceit}
\VS{6}That we may
buy the
poor for
silver, and the
needy for a pair of
shoes;
{\ADD{yea}}, and
sell the
refuse of the
wheat?
\VS{7}The
{\ND{
LORD}} hath
sworn by the
excellency of
Jacob, Surely I will
never
forget any of their
works.
\VS{8}Shall not the
land
tremble for this, and every one
mourn that
dwelleth therein? and it shall rise
up wholly as a
flood; and it shall be cast
out and
drowned, as
{\ADD{by}} the
flood of
Egypt.
\VS{9}And it shall come to pass in that
day,
saith the
Lord
{\ND{
GOD}}, that I will cause the
sun to go
down at
noon, and I will
darken the
earth in the
clear
day:
\VS{10}And I will
turn your
feasts into
mourning, and all your
songs into
lamentation; and I will bring
up
sackcloth upon all
loins, and
baldness upon every
head; and I will
make it as the
mourning of an
only
{\ADD{son}}, and the
end thereof as a
bitter
day.
\par }{\PP \VS{11}¶ Behold, the
days
come,
saith the
Lord
{\ND{
GOD}}, that I will
send a
famine in the
land, not a
famine of
bread, nor a
thirst for
water, but of
hearing the
words of the
{\ND{
LORD}}:
\VS{12}And they shall
wander from
sea to
sea, and from the
north even to the
east, they shall run to and
fro to
seek the
word of the
{\ND{
LORD}}, and shall not
find
{\ADD{it}}.
\VS{13}In that
day shall the
fair
virgins and young
men
faint for
thirst.
\VS{14}They that
swear by the
sin of
Samaria, and
say, Thy
god, O
Dan,
liveth; and, The
manner of
Beer-sheba
liveth; even they shall
fall, and never rise
up again.\FTNT{††}{{\FR 8.14 }
manner: Heb. way}

\par }\Chap{9}{\PP \VerseOne{1}I
saw the
Lord
standing upon the
altar: and he
said,
Smite the lintel of the
door, that the
posts may
shake: and
cut them in the
head, all of them; and I will
slay the
last of them with the
sword: he that
fleeth of them shall not flee
away, and he that
escapeth of them shall not be
delivered.\FTNT{*}{{\FR 9.1 }
lintel: or, chapiter, or, knop}\FTNT{†}{{\FR 9.1 }
cut…: or, wound them}
\VS{2}Though they
dig into
hell, thence shall mine
hand
take them; though they climb
up to
heaven, thence will I bring them
down:
\VS{3}And though they
hide themselves in the
top of
Carmel, I will
search and take them
out thence; and though they be
hid from my
sight in the
bottom of the
sea, thence will I
command the
serpent, and he shall
bite them:
\VS{4}And though they
go into
captivity
before their
enemies, thence will I
command the
sword, and it shall
slay them: and I will
set mine
eyes upon them for
evil, and not for
good.
\VS{5}And the
Lord
{\ND{
GOD}} of
hosts
{\ADD{is}} he that
toucheth the
land, and it shall
melt, and all that
dwell therein shall
mourn: and it shall rise
up wholly like a
flood; and shall be
drowned, as
{\ADD{by}} the
flood of
Egypt.
\VS{6}{\ADD{It is}} he that
buildeth his
stories in the
heaven, and hath
founded his
troop in the
earth; he that
calleth for the
waters of the
sea, and poureth them
out upon the
face of the
earth: The
{\ND{
LORD}}
{\ADD{is}} his
name.\FTNT{‡}{{\FR 9.6 }
stories: or, spheres: Heb. ascensions}\FTNT{§}{{\FR 9.6 }
troop: or, bundle}
\VS{7}{\ADD{Are}} ye not as
children of the
Ethiopians unto me, O
children of
Israel?
saith the
{\ND{
LORD}}. Have not I brought
up
Israel out of the
land of
Egypt? and the
Philistines from
Caphtor, and the
Syrians from
Kir?
\VS{8}Behold, the
eyes of the
Lord
{\ND{
GOD}}
{\ADD{are}} upon the
sinful
kingdom, and I will
destroy it from off the
face of the
earth;
saving that I will not
utterly
destroy the
house of
Jacob,
saith the
{\ND{
LORD}}.
\VS{9}For, lo, I will
command, and I will
sift the
house of
Israel among all
nations, like as
{\ADD{corn}} is
sifted in a
sieve, yet shall not the least
grain
fall upon the
earth.\FTNT{**}{{\FR 9.9 }
sift: Heb. cause to move}\FTNT{††}{{\FR 9.9 }
grain: Heb. stone}
\VS{10}All the
sinners of my
people shall
die by the
sword, which
say, The
evil shall not
overtake nor
prevent us.
\par }{\PP \VS{11}¶ In that
day will I raise
up the
tabernacle of
David that is
fallen, and close
up the
breaches thereof; and I will raise
up his
ruins, and I will
build it as in the
days of
old:\FTNT{‡‡}{{\FR 9.11 }
close: Heb. hedge, or, wall}
\VS{12}That they may
possess the
remnant of
Edom, and of all the
heathen, which are
called by my
name,
saith the
{\ND{
LORD}} that
doeth this.\FTNT{§§}{{\FR 9.12 }
which…: Heb. upon whom my name is called}
\VS{13}Behold, the
days
come,
saith the
{\ND{
LORD}}, that the
plowman shall
overtake the
reaper, and the
treader of
grapes him that
soweth
seed; and the
mountains shall
drop sweet
wine, and all the
hills shall
melt.\FTNT{***}{{\FR 9.13 }
soweth: Heb. draweth forth}\FTNT{†††}{{\FR 9.13 }
sweet: or, new}
\VS{14}And I will bring
again the
captivity of my
people of
Israel, and they shall
build the
waste
cities, and
inhabit
{\ADD{them;}} and they shall
plant
vineyards, and
drink the
wine thereof; they shall also
make
gardens, and
eat the
fruit of them.
\VS{15}And I will
plant them upon their
land, and they shall no more be pulled
up out of their
land which I have
given them,
saith the
{\ND{
LORD}} thy
God.
\par }