\NormalFont\ShortTitle{Daniel}
{\MT The Book of Daniel

\par }\ChapOne{1}{\PP \VerseOne{1}In the
third
year of the
reign of
Jehoiakim
king of
Judah
came
Nebuchadnezzar
king of
Babylon unto
Jerusalem, and
besieged it.
\VS{2}And the
Lord
gave
Jehoiakim
king of
Judah into his
hand, with
part of the
vessels of the
house of
God: which he
carried into the
land of
Shinar to the
house of his
god; and he
brought the
vessels into the
treasure
house of his
god.
\par }{\PP \VS{3}¶ And the
king
spake unto
Ashpenaz the
master of his
eunuchs, that he should
bring
{\ADD{certain}} of the
children of
Israel, and of the
king’s
seed, and of the
princes;
\VS{4}Children in whom
{\ADD{was}} no
blemish, but
well
favoured, and
skilful in all
wisdom, and
cunning in
knowledge, and
understanding
science, and such as
{\ADD{had}}
ability in them to
stand in the
king’s
palace, and whom they might
teach the
learning and the
tongue of the
Chaldeans.
\VS{5}And the
king
appointed them a
daily
provision of the
king’s
meat, and of the
wine which he
drank: so
nourishing them
three
years, that at the
end thereof they might
stand
before the
king.\FTNT{*}{{\FR 1.5 }
the wine…: Heb. the wine of his drink}
\VS{6}Now among these were of the
children of
Judah,
Daniel,
Hananiah,
Mishael, and
Azariah:
\VS{7}Unto whom the
prince of the
eunuchs
gave
names: for he
gave unto
Daniel
{\ADD{the name}} of
Belteshazzar; and to
Hananiah, of
Shadrach; and to
Mishael, of
Meshach; and to
Azariah, of
Abed-nego.
\par }{\PP \VS{8}¶ But
Daniel
purposed in his
heart that he would not
defile himself with the
portion of the
king’s
meat, nor with the
wine which he
drank: therefore he
requested of the
prince of the
eunuchs that he might not
defile himself.
\VS{9}Now
God had
brought
Daniel into
favour and tender
love
with the
prince of the
eunuchs.
\VS{10}And the
prince of the
eunuchs
said unto
Daniel, I
fear my
lord the
king, who hath
appointed your
meat and your
drink: for why should he
see your
faces worse
liking than the
children which
{\ADD{are}} of your
sort? then shall ye make
{\ADD{me}}
endanger my
head to the
king.\FTNT{†}{{\FR 1.10 }
worse…: Heb. sadder}\FTNT{‡}{{\FR 1.10 }
sort: or, term, or, continuance?}
\VS{11}Then
said
Daniel to
Melzar, whom the
prince of the
eunuchs had
set over
Daniel,
Hananiah,
Mishael, and
Azariah,\FTNT{§}{{\FR 1.11 }
Melzar: or, the steward}
\VS{12}Prove thy
servants, I beseech thee,
ten
days; and let them
give us
pulse to
eat, and
water to
drink.\FTNT{**}{{\FR 1.12 }
pulse: Heb. of pulse}\FTNT{††}{{\FR 1.12 }
to eat…: Heb. that we may eat, etc}
\VS{13}Then let our
countenances be looked
upon
before thee, and the
countenance of the
children that
eat of the
portion of the
king’s
meat: and as thou
seest,
deal with thy
servants.
\VS{14}So he
consented to them in this
matter, and
proved them
ten
days.
\VS{15}And at the
end of
ten
days their
countenances
appeared
fairer and
fatter in
flesh than all the
children which did
eat the
portion of the
king’s
meat.
\VS{16}Thus
Melzar took
away the portion of their
meat, and the
wine that they should
drink; and
gave them
pulse.
\par }{\PP \VS{17}¶ As for these
four
children,
God
gave them
knowledge and
skill in all
learning and
wisdom: and
Daniel had
understanding in all
visions and
dreams.\FTNT{‡‡}{{\FR 1.17 }
Daniel…: or, he made Daniel understand}
\VS{18}Now at the
end of the
days that the
king had
said he should
bring them in, then the
prince of the
eunuchs
brought them in
before
Nebuchadnezzar.
\VS{19}And the
king
communed with them; and among them all was
found none like
Daniel,
Hananiah,
Mishael, and
Azariah: therefore
stood they
before the
king.
\VS{20}And in all
matters of
wisdom
{\ADD{and}}
understanding, that the
king
enquired of them, he
found them ten
times
better than all the
magicians
{\ADD{and}}
astrologers that
{\ADD{were}} in all his
realm.\FTNT{§§}{{\FR 1.20 }
wisdom…: Heb. wisdom of understanding}
\VS{21}And
Daniel continued
{\ADD{even}} unto the
first
year of
king
Cyrus.

\par }\Chap{2}{\PP \VerseOne{1}And in the
second
year of the
reign of
Nebuchadnezzar
Nebuchadnezzar
dreamed
dreams, wherewith his
spirit was
troubled, and his
sleep
brake from him.
\VS{2}Then the
king
commanded to
call the
magicians, and the
astrologers, and the
sorcerers, and the
Chaldeans, for to
shew the
king his
dreams. So they
came and
stood
before the
king.
\VS{3}And the
king
said unto them, I have
dreamed a
dream, and my
spirit was
troubled to
know the
dream.
\VS{4}Then
spake the
Chaldeans to the
king in
Syriack, O
king,
live for
ever:
tell thy
servants the
dream, and we will
shew the
interpretation.\FTNT{*}{{\FR 2.4 }
O king…: (Chaldee, to the end of chapter seven)}
\VS{5}The
king
answered and
said to the
Chaldeans, The
thing is
gone from
me:
if ye will
not make
known unto me the
dream, with the
interpretation thereof, ye shall be
cut in
pieces, and your
houses shall be
made a
dunghill.\FTNT{†}{{\FR 2.5 }
cut…: Chaldee, made pieces}
\VS{6}But
if ye
shew the
dream, and the
interpretation thereof, ye shall
receive
of
me
gifts and
rewards and
great
honour:
therefore
shew me the
dream, and the
interpretation thereof.\FTNT{‡}{{\FR 2.6 }
rewards: or, fee}
\VS{7}They
answered
again and
said, Let the
king
tell his
servants the
dream, and we will
shew the
interpretation of it.
\VS{8}The
king
answered and
said,
I
know
of
certainty that
ye would
gain the
time,
because ye
see the
thing is
gone
from me.\FTNT{§}{{\FR 2.8 }
gain: Chaldee, buy}
\VS{9}But
if ye will
not make
known unto me the
dream,
{\ADD{there is but}}
one
decree for you: for ye have
prepared
lying and
corrupt
words to
speak
before me,
till the
time be
changed:
therefore
tell me the
dream, and I shall
know
that ye can
shew me the
interpretation thereof.
\par }{\PP \VS{10}¶ The
Chaldeans
answered
before the
king, and
said, There
is
not a
man
upon the
earth that
can
shew the
king’s
matter:
therefore
{\ADD{there is}}
no
king,
lord, nor
ruler,
{\ADD{that}}
asked
such
things at
any
magician, or
astrologer, or
Chaldean.
\VS{11}And
{\ADD{it is}} a
rare
thing that the
king
requireth, and there
is
none
other that can
shew it
before the
king,
except the
gods, whose
dwelling
is
not
with
flesh.
\VS{12}For
this
cause the
king was
angry and
very
furious, and
commanded to
destroy
all the
wise
{\ADD{men}} of
Babylon.
\VS{13}And the
decree went
forth that the
wise
{\ADD{men}} should be
slain; and they
sought
Daniel and his
fellows to be
slain.
\par }{\PP \VS{14}¶
Then
Daniel
answered with
counsel and
wisdom to
Arioch the
captain
of the
king’s
guard, which was gone
forth to
slay the
wise
{\ADD{men}} of
Babylon:\FTNT{**}{{\FR 2.14 }
answered…: Chaldee, returned}\FTNT{††}{{\FR 2.14 }
captain…: or, chief marshal: Chaldee, chief of the executioners, or, slaughtermen}
\VS{15}He
answered and
said to
Arioch the
king’s
captain,
Why
{\ADD{is}} the
decree
{\ADD{so}}
hasty
from the
king?
Then
Arioch made the
thing
known to
Daniel.
\VS{16}Then
Daniel went
in, and
desired
of the
king that he would
give him
time, and that he would
shew the
king the
interpretation.
\VS{17}Then
Daniel
went to his
house, and made the
thing
known to
Hananiah,
Mishael, and
Azariah, his
companions:
\VS{18}That they would
desire
mercies
of the
God of
heaven
concerning
this
secret; that
Daniel and his
fellows should
not
perish
with the
rest of the
wise
{\ADD{men}} of
Babylon.\FTNT{‡‡}{{\FR 2.18 }
of the God: Chaldee, from before God}\FTNT{§§}{{\FR 2.18 }
that Daniel…: or, that they should not destroy Daniel, etc}
\par }{\PP \VS{19}¶
Then was the
secret
revealed unto
Daniel in a
night
vision.
Then
Daniel
blessed the
God of
heaven.
\VS{20}Daniel
answered and
said,
Blessed
be the
name of
God
for
ever and
ever: for
wisdom and
might
are his:
\VS{21}And he
changeth the
times and the
seasons: he
removeth
kings, and setteth
up
kings: he
giveth
wisdom unto the
wise, and
knowledge to them that
know
understanding:
\VS{22}He
revealeth the
deep and secret
things: he
knoweth
what
{\ADD{is}} in the
darkness, and the
light
dwelleth with
him.
\VS{23}I
thank thee, and
praise thee, O thou
God of my
fathers, who hast
given me
wisdom and
might, and hast made
known unto me
now what we
desired of
thee: for thou hast
{\ADD{now}} made
known unto us the
king’s
matter.
\par }{\PP \VS{24}¶
Therefore
Daniel went
in
unto
Arioch, whom the
king had
ordained to
destroy the
wise
{\ADD{men}} of
Babylon: he
went and
said
thus unto him;
Destroy
not the
wise
{\ADD{men}} of
Babylon: bring me
in
before the
king, and I will
shew unto the
king the
interpretation.
\VS{25}Then
Arioch brought
in
Daniel
before the
king in
haste, and
said
thus unto him, I have
found a
man of the
captives
of
Judah, that will make
known unto the
king the
interpretation.\FTNT{***}{{\FR 2.25 }
I have…: Chaldee, That I have found}\FTNT{†††}{{\FR 2.25 }
captives…: Chaldee, children of the captivity of Judah}
\VS{26}The
king
answered and
said to
Daniel, whose
name
{\ADD{was}}
Belteshazzar,
Art thou
able to make
known unto me the
dream which I have
seen, and the
interpretation thereof?
\VS{27}Daniel
answered in the
presence of the
king, and
said, The
secret which the
king hath
demanded
cannot the
wise
{\ADD{men}}, the
astrologers, the
magicians, the
soothsayers,
shew unto the
king;
\VS{28}But there
is a
God in
heaven that
revealeth
secrets, and maketh
known to the
king
Nebuchadnezzar
what shall
be in the
latter
days. Thy
dream, and the
visions of thy
head
upon thy
bed, are
these;\FTNT{‡‡‡}{{\FR 2.28 }
maketh…: Chaldee, hath made known}
\VS{29}As for
thee, O
king, thy
thoughts
came
{\ADD{into thy mind}}
upon thy
bed,
what should come to
pass
hereafter: and he that
revealeth
secrets maketh
known to thee
what shall come to
pass.\FTNT{§§§}{{\FR 2.29 }
came: Chaldee, came up}
\VS{30}But as for
me,
this
secret is
not
revealed to me for
{\ADD{any}}
wisdom that I
have more
than
any
living,
but for
{\ADD{their}}
sakes
that shall make
known the
interpretation to the
king, and that thou mightest
know the
thoughts of thy
heart.\FTNT{*}{{\FR 2.30 }
but for…: or, but for the intent that the interpretation may be made known to the king}
\par }{\PP \VS{31}¶
Thou, O
king,
sawest, and
behold a
great
image.
This
great
image, whose
brightness
{\ADD{was}}
excellent,
stood
before thee; and the
form thereof
{\ADD{was}}
terrible.\FTNT{†}{{\FR 2.31 }
sawest: Chaldee, wast seeing}
\VS{32}This
image’s
head
{\ADD{was}} of
fine
gold, his
breast and his
arms of
silver, his
belly and his
thighs of
brass,\FTNT{‡}{{\FR 2.32 }
thighs: or, sides}
\VS{33}His
legs of
iron, his
feet
part of
iron and
part of
clay.
\VS{34}Thou
sawest
till that a
stone was cut
out
without
hands, which
smote the
image
upon his
feet
{\ADD{that were}} of
iron and
clay, and
brake
them to
pieces.\FTNT{§}{{\FR 2.34 }
without…: or, which was not in hands}
\VS{35}Then was the
iron, the
clay, the
brass, the
silver, and the
gold, broken to
pieces
together, and
became like the
chaff
of the
summer
threshingfloors; and the
wind
carried
them
away,
that
no
place was
found for them: and the
stone that
smote the
image
became a
great
mountain, and
filled the
whole
earth.
\par }{\PP \VS{36}¶
This
{\ADD{is}} the
dream; and we will
tell the
interpretation thereof
before the
king.
\VS{37}Thou, O
king,
{\ADD{art}} a
king of
kings: for the
God of
heaven hath
given thee a
kingdom,
power, and
strength, and
glory.
\VS{38}And
wheresoever the
children of
men
dwell, the
beasts of the
field and the
fowls of the
heaven hath he
given into thine
hand, and hath made thee
ruler over them
all.
Thou
{\ADD{art}} this
head of
gold.
\VS{39}And
after thee shall
arise
another
kingdom
inferior to
thee, and
another
third
kingdom of
brass, which shall bear
rule over
all the
earth.
\VS{40}And the
fourth
kingdom shall
be
strong as
iron: forasmuch as
iron breaketh in
pieces and
subdueth
all
{\ADD{things:}} and
as
iron that
breaketh
all
these, shall it break in
pieces and
bruise.
\VS{41}And whereas thou
sawest the
feet and
toes,
part of
potters’
clay, and
part of
iron, the
kingdom shall
be
divided;
but there shall
be in it
of the
strength of the
iron, forasmuch
as thou
sawest the
iron
mixed with
miry
clay.
\VS{42}And
{\ADD{as}} the
toes of the
feet
{\ADD{were}}
part of
iron, and
part of
clay,
{\ADD{so}} the
kingdom shall
be
partly
strong, and
partly
broken.\FTNT{**}{{\FR 2.42 }
broken: or, brittle}
\VS{43}And
whereas thou
sawest
iron
mixed with
miry
clay, they shall mingle
themselves with the
seed of
men: but they
shall
not
cleave
one
to
another,
even as
iron is
not
mixed with
clay.\FTNT{††}{{\FR 2.43 }
one…: Chaldee, this with this}
\VS{44}And in the
days of
these
kings shall the
God of
heaven set
up a
kingdom, which shall
never be
destroyed: and the
kingdom shall
not be
left to
other
people,
{\ADD{but}} it shall break in
pieces and
consume
all
these
kingdoms, and
it shall
stand for
ever.\FTNT{‡‡}{{\FR 2.44 }
the days: Chaldee, their days}\FTNT{§§}{{\FR 2.44 }
the kingdom: Chaldee, the kingdom thereof}
\VS{45}Forasmuch
as thou
sawest that the
stone was cut
out of the
mountain
without
hands, and that it brake in
pieces the
iron, the
brass, the
clay, the
silver, and the
gold; the
great
God hath made
known to the
king
what shall come to
pass
hereafter: and the
dream
{\ADD{is}}
certain, and the
interpretation thereof
sure.\FTNT{***}{{\FR 2.45 }
without…: or, which was not in hands}\FTNT{†††}{{\FR 2.45 }
hereafter: Chaldee, after this}
\par }{\PP \VS{46}¶
Then the
king
Nebuchadnezzar
fell
upon his
face, and
worshipped
Daniel, and
commanded that they should
offer an
oblation and sweet
odours unto him.
\VS{47}The
king
answered unto
Daniel, and
said,
Of a
truth
{\ADD{it is}},
that your
God
{\ADD{is}} a
God of
gods, and a
Lord of
kings, and a
revealer of
secrets, seeing thou
couldest
reveal
this
secret.
\VS{48}Then the
king
made
Daniel a great
man, and
gave him
many
great
gifts, and made him
ruler
over the
whole
province of
Babylon, and
chief of the
governors
over
all the
wise
{\ADD{men}} of
Babylon.
\VS{49}Then
Daniel
requested
of the
king, and he
set
Shadrach,
Meshach, and
Abed-nego,
over the
affairs of the
province of
Babylon: but
Daniel
{\ADD{sat}} in the
gate of the
king.

\par }\Chap{3}{\PP \VerseOne{1}Nebuchadnezzar the
king
made an
image of
gold, whose
height
{\ADD{was}}
threescore
cubits,
{\ADD{and}} the
breadth thereof
six
cubits: he set it
up in the
plain of
Dura, in the
province of
Babylon.
\VS{2}Then
Nebuchadnezzar the
king
sent to gather
together the
princes, the
governors, and the
captains, the
judges, the
treasurers, the
counsellors, the
sheriffs, and
all the
rulers of the
provinces, to
come to the
dedication of the
image which
Nebuchadnezzar the
king had set
up.
\VS{3}Then the
princes, the
governors, and
captains, the
judges, the
treasurers, the
counsellors, the
sheriffs, and
all the
rulers of the
provinces, were gathered
together unto the
dedication of the
image that
Nebuchadnezzar the
king had set
up; and they
stood
before the
image that
Nebuchadnezzar had set
up.
\VS{4}Then an
herald
cried
aloud, To you it is
commanded, O
people,
nations, and
languages,\FTNT{*}{{\FR 3.4 }
aloud: Chaldee, with might}\FTNT{†}{{\FR 3.4 }
it…: Chaldee, they command}
\VS{5}{\ADD{That}} at what
time ye
hear the
sound of the
cornet,
flute,
harp,
sackbut,
psaltery,
dulcimer, and
all
kinds of
musick, ye fall
down and
worship the
golden
image that
Nebuchadnezzar the
king hath set
up:\FTNT{‡}{{\FR 3.5 }
dulcimer: or, singing: Chaldee, symphony}
\VS{6}And
whoso
falleth
not
down and
worshippeth shall the same
hour be
cast into the
midst of a
burning
fiery
furnace.
\VS{7}Therefore at that
time,
when
all the
people
heard the
sound of the
cornet,
flute,
harp,
sackbut,
psaltery, and
all
kinds of
musick,
all the
people, the
nations, and the
languages, fell
down
{\ADD{and}}
worshipped the
golden
image that
Nebuchadnezzar the
king had set
up.
\par }{\PP \VS{8}¶
Wherefore at
that
time
certain
Chaldeans came
near, and
accused the
Jews.
\VS{9}They
spake and
said to the
king
Nebuchadnezzar, O
king,
live for
ever.
\VS{10}Thou, O
king, hast
made a
decree, that
every
man that shall
hear the
sound of the
cornet,
flute,
harp,
sackbut,
psaltery, and
dulcimer, and
all
kinds of
musick, shall fall
down and
worship the
golden
image:\FTNT{§}{{\FR 3.10 }
dulcimer: or, singing: Chaldee, symphony}
\VS{11}And
whoso
falleth
not
down and
worshippeth,
{\ADD{that}} he should be
cast into the
midst of a
burning
fiery
furnace.
\VS{12}There
are
certain
Jews
whom thou hast
set
over the
affairs of the
province of
Babylon,
Shadrach,
Meshach, and
Abed-nego;
these
men, O
king,
have
not
regarded
thee: they
serve
not thy
gods,
nor
worship the
golden
image which thou hast set
up.\FTNT{**}{{\FR 3.12 }
have…: Chaldee, have set no regard upon thee}
\par }{\PP \VS{13}¶
Then
Nebuchadnezzar in
{\ADD{his}}
rage and
fury
commanded to
bring
Shadrach,
Meshach, and
Abed-nego.
Then they
brought
these
men
before the
king.
\VS{14}Nebuchadnezzar
spake and
said unto them,
{\ADD{Is it}}
true, O
Shadrach,
Meshach, and
Abed-nego, do
not
ye
serve my
gods,
nor
worship the
golden
image which I have
set up?\FTNT{††}{{\FR 3.14 }
true: or, of purpose}
\VS{15}Now
if ye
be
ready that at what
time ye
hear the
sound of the
cornet,
flute,
harp,
sackbut,
psaltery, and
dulcimer, and
all
kinds of
musick, ye fall
down and
worship the
image which I have
made;
{\ADD{well:}} but
if ye
worship
not, ye shall be
cast the same
hour into the
midst of a
burning
fiery
furnace; and
who
{\ADD{is}} that
God that shall
deliver you out
of my
hands?\FTNT{‡‡}{{\FR 3.15 }
dulcimer: or, singing: Chaldee, symphony}
\VS{16}Shadrach,
Meshach, and
Abed-nego,
answered and
said to the
king, O
Nebuchadnezzar,
we
{\ADD{are}}
not
careful to
answer thee
in
this
matter.
\VS{17}If it be
{\ADD{so}}, our
God whom
we
serve
is
able to
deliver us
from the
burning
fiery
furnace,
and he will
deliver
{\ADD{us}} out
of thine
hand, O
king.
\VS{18}But
if
not, be
it
known unto thee, O
king, that we
will
not
serve thy
gods,
nor
worship the
golden
image which thou hast set
up.
\par }{\PP \VS{19}¶
Then was
Nebuchadnezzar
full of
fury, and the
form of his
visage was
changed
against
Shadrach,
Meshach, and
Abed-nego:
{\ADD{therefore}} he
spake, and
commanded that they should
heat the
furnace
one seven
times
more
than it was
wont to be
heated.\FTNT{§§}{{\FR 3.19 }
full: Chaldee, filled}
\VS{20}And he
commanded the
most
mighty
men that
{\ADD{were}} in his
army to
bind
Shadrach,
Meshach, and
Abed-nego,
{\ADD{and}} to
cast
{\ADD{them}} into the
burning
fiery
furnace.\FTNT{***}{{\FR 3.20 }
most…: Chaldee, mighty of strength}
\VS{21}Then
these
men were
bound in their
coats, their
hosen, and their
hats, and their
{\ADD{other}}
garments, and were
cast into the
midst of the
burning
fiery
furnace.\FTNT{†††}{{\FR 3.21 }
coats: or, mantles}\FTNT{‡‡‡}{{\FR 3.21 }
hats: or, turbans}
\VS{22}Therefore
because the
king’s
commandment was
urgent, and the
furnace
exceeding
hot, the
flame of the
fire
slew
those
men that took
up
Shadrach,
Meshach, and
Abed-nego.\FTNT{§§§}{{\FR 3.22 }
commandment: Chaldee, word}\FTNT{*}{{\FR 3.22 }
flame: or, spark}
\VS{23}And
these
three
men,
Shadrach,
Meshach, and
Abed-nego, fell
down
bound into the
midst of the
burning
fiery
furnace.
\VS{24}Then
Nebuchadnezzar the
king was
astonied, and rose
up in
haste,
{\ADD{and}}
spake, and
said unto his
counsellors, Did
not we
cast
three
men
bound into the
midst of the
fire? They
answered and
said unto the
king,
True, O
king.\FTNT{†}{{\FR 3.24 }
counsellors: or, governors}
\VS{25}He
answered and
said,
Lo,
I
see
four
men
loose,
walking in the
midst of the
fire, and they
have
no
hurt; and the
form of the
fourth is
like the
Son of
God.\FTNT{‡}{{\FR 3.25 }
they…: Chaldee, there is no hurt in them}
\par }{\PP \VS{26}¶
Then
Nebuchadnezzar came
near to the
mouth of the
burning
fiery
furnace,
{\ADD{and}}
spake, and
said,
Shadrach,
Meshach, and
Abed-nego, ye
servants of the most
high
God, come
forth, and
come
{\ADD{hither}}.
Then
Shadrach,
Meshach, and
Abed-nego, came
forth
of the
midst of the
fire.\FTNT{§}{{\FR 3.26 }
mouth: Chaldee, door}
\VS{27}And the
princes,
governors, and
captains, and the
king’s
counsellors, being gathered
together,
saw
these
men, upon whose
bodies the
fire had
no
power,
nor was an
hair of their
head
singed,
neither were their
coats
changed,
nor the
smell of
fire had
passed on them.
\par }{\PP \VS{28}{\ADD{Then}}
Nebuchadnezzar
spake, and
said,
Blessed
{\ADD{be}} the
God of
Shadrach,
Meshach, and
Abed-nego, who hath
sent his
angel, and
delivered his
servants that
trusted in
him, and have
changed the
king’s
word, and
yielded their
bodies, that they might
not
serve
nor
worship
any
god,
except their own
God.
\VS{29}Therefore
I
make a
decree, That
every
people,
nation, and
language, which
speak any thing
amiss
against the
God of
Shadrach,
Meshach, and
Abed-nego, shall be
cut in
pieces, and their
houses shall be
made a
dunghill:
because there
is
no
other
God that
can
deliver after
this sort.\FTNT{**}{{\FR 3.29 }
I make…: Chaldee, a decree is made by me}\FTNT{††}{{\FR 3.29 }
any…: Chaldee, error}\FTNT{‡‡}{{\FR 3.29 }
cut…: Chaldee, made pieces}
\VS{30}Then the
king
promoted
Shadrach,
Meshach, and
Abed-nego, in the
province of
Babylon.\FTNT{§§}{{\FR 3.30 }
promoted: Chaldee, made to prosper}

\par }\Chap{4}{\PP \VerseOne{1}Nebuchadnezzar the
king, unto
all
people,
nations, and
languages, that
dwell in
all the
earth;
Peace be
multiplied unto you.
\VS{2}I
thought it
good to
shew the
signs and
wonders that the
high
God hath
wrought
toward me.\FTNT{*}{{\FR 4.2 }
I thought…: Chaldee, It was seemly before me}
\VS{3}How
great
{\ADD{are}} his
signs! and
how
mighty
{\ADD{are}} his
wonders! his
kingdom
{\ADD{is}} an
everlasting
kingdom, and his
dominion
{\ADD{is}}
from
generation to
generation.
\par }{\PP \VS{4}¶
I
Nebuchadnezzar
was at
rest in mine
house, and
flourishing in my
palace:
\VS{5}I
saw a
dream which made me
afraid, and the
thoughts
upon my
bed and the
visions of my
head
troubled me.
\VS{6}Therefore
made
I a
decree to
bring in
all the
wise
{\ADD{men}} of
Babylon
before me, that they might make
known unto me the
interpretation of the
dream.
\VS{7}Then
came in the
magicians, the
astrologers, the
Chaldeans, and the
soothsayers: and
I
told the
dream
before them; but they did
not make
known unto me the
interpretation thereof.
\par }{\PP \VS{8}¶ But
at the
last
Daniel came
in before
me, whose
name
{\ADD{was}}
Belteshazzar, according to the
name of my
god, and in whom
{\ADD{is}} the
spirit of the
holy
gods: and
before him I
told the
dream,
{\ADD{saying}},
\VS{9}O
Belteshazzar,
master of the
magicians, because
I
know that the
spirit of the
holy
gods
{\ADD{is}} in thee, and
no
secret
troubleth thee,
tell me the
visions of my
dream that I have
seen, and the
interpretation thereof.
\VS{10}Thus
{\ADD{were}} the
visions of mine
head
in my
bed; I
saw, and
behold a
tree in the
midst of the
earth, and the
height thereof
{\ADD{was}}
great.\FTNT{†}{{\FR 4.10 }
I saw: Chaldee, I was seeing}
\VS{11}The
tree
grew, and was
strong, and the
height thereof
reached unto
heaven, and the
sight thereof to the
end of
all the
earth:
\VS{12}The
leaves thereof
{\ADD{were}}
fair, and the
fruit thereof
much, and in it
{\ADD{was}}
meat for
all: the
beasts of the
field had
shadow
under it, and the
fowls of the
heaven
dwelt in the
boughs thereof, and
all
flesh was
fed of
it.
\VS{13}I
saw in the
visions of my
head
upon my
bed, and,
behold, a
watcher and an holy
one came
down
from
heaven;
\VS{14}He
cried
aloud, and
said
thus, Hew
down the
tree, and cut
off his
branches, shake
off his
leaves, and
scatter his
fruit: let the
beasts get
away
from under
it, and the
fowls
from his
branches:\FTNT{‡}{{\FR 4.14 }
aloud: Chaldee, with might}
\VS{15}Nevertheless
leave the
stump of his
roots in the
earth, even with a
band of
iron and
brass, in the tender
grass of the
field; and let it be
wet with the
dew of
heaven, and
{\ADD{let}} his
portion
{\ADD{be}}
with the
beasts in the
grass of the
earth:
\VS{16}Let his
heart be
changed
from
man’s, and let a
beast’s
heart be
given unto him; and let
seven
times
pass
over him.
\VS{17}This
matter
{\ADD{is}} by the
decree of the
watchers, and the
demand by the
word of the holy
ones:
to the
intent that the
living may
know that the most
High
ruleth in the
kingdom of
men, and
giveth it to
whomsoever he
will, and setteth
up
over it the
basest of
men.
\VS{18}This
dream
I
king
Nebuchadnezzar have
seen. Now
thou, O
Belteshazzar,
declare the
interpretation thereof, forasmuch
as
all the
wise
{\ADD{men}} of my
kingdom are
not
able to make
known unto me the
interpretation: but
thou
{\ADD{art}}
able; for the
spirit of the
holy
gods
{\ADD{is}} in thee.
\par }{\PP \VS{19}¶
Then
Daniel, whose
name
{\ADD{was}}
Belteshazzar, was
astonied for
one
hour, and his
thoughts
troubled him. The
king
spake, and
said,
Belteshazzar, let
not the
dream, or the
interpretation thereof,
trouble thee.
Belteshazzar
answered and
said, My
lord, the
dream
{\ADD{be}} to them that
hate thee, and the
interpretation thereof to thine
enemies.
\VS{20}The
tree that thou
sawest, which
grew, and was
strong, whose
height
reached unto the
heaven, and the
sight thereof to
all the
earth;
\VS{21}Whose
leaves
{\ADD{were}}
fair, and the
fruit thereof
much, and in it
{\ADD{was}}
meat for
all;
under which the
beasts of the
field
dwelt, and upon whose
branches the
fowls of the
heaven had their
habitation:
\VS{22}It
{\ADD{is}}
thou, O
king, that art
grown and become
strong: for thy
greatness is
grown, and
reacheth unto
heaven, and thy
dominion to the
end of the
earth.
\VS{23}And whereas the
king
saw a
watcher and an holy
one coming
down
from
heaven, and
saying,
Hew the
tree
down, and
destroy it;
yet
leave the
stump of the
roots thereof in the
earth, even with a
band of
iron and
brass, in the tender
grass of the
field; and let it be
wet with the
dew of
heaven, and
{\ADD{let}} his
portion
{\ADD{be}}
with the
beasts of the
field,
till
seven
times
pass
over him;
\VS{24}This
{\ADD{is}} the
interpretation, O
king, and
this
{\ADD{is}} the
decree of the most
High, which is
come
upon my
lord the
king:
\VS{25}That they shall
drive thee
from
men, and thy
dwelling shall
be
with the
beasts of the
field, and they shall make thee to
eat
grass as
oxen, and they shall
wet thee with the
dew of
heaven, and
seven
times shall
pass
over thee,
till thou
know that the most
High
ruleth in the
kingdom of
men, and
giveth it to
whomsoever he
will.
\VS{26}And whereas they
commanded to
leave the
stump of the
tree
roots; thy
kingdom shall be
sure unto thee,
after that thou shalt have
known that the
heavens do
rule.
\VS{27}Wherefore, O
king, let my
counsel be
acceptable unto
thee, and break
off thy
sins by
righteousness, and thine
iniquities by shewing
mercy to the
poor;
if it may
be a
lengthening of thy
tranquillity.\FTNT{§}{{\FR 4.27 }
a lengthening…: or, an healing of thine error}
\par }{\PP \VS{28}¶
All this
came
upon the
king
Nebuchadnezzar.
\VS{29}At the
end of
twelve
months he
walked
in the
palace of the
kingdom of
Babylon.\FTNT{**}{{\FR 4.29 }
in: or, upon}
\VS{30}The
king
spake, and
said, Is
not
this
great
Babylon, that
I have
built for the
house of the
kingdom by the
might of my
power, and for the
honour of my
majesty?
\VS{31}While the
word
{\ADD{was}} in the
king’s
mouth, there
fell a
voice
from
heaven,
{\ADD{saying}}, O
king
Nebuchadnezzar, to thee it is
spoken; The
kingdom is
departed
from thee.
\VS{32}And they shall
drive thee
from
men, and thy
dwelling
{\ADD{shall be}}
with the
beasts of the
field: they shall make thee to
eat
grass as
oxen, and
seven
times shall
pass
over thee,
until thou
know that the most
High
ruleth in the
kingdom of
men, and
giveth it to
whomsoever he
will.
\VS{33}The same
hour was the
thing
fulfilled
upon
Nebuchadnezzar: and he was
driven
from
men, and did
eat
grass as
oxen, and his
body was
wet with the
dew of
heaven,
till his
hairs were
grown like
eagles’
{\ADD{feathers}}, and his
nails like
birds’
{\ADD{claws}}.
\par }{\PP \VS{34}And at the
end of the
days
I
Nebuchadnezzar lifted
up mine
eyes unto
heaven, and mine
understanding
returned unto
me, and I
blessed the most
High, and I
praised and
honoured him that
liveth for
ever, whose
dominion
{\ADD{is}} an
everlasting
dominion, and his
kingdom
{\ADD{is}}
from
generation to
generation:
\VS{35}And
all the
inhabitants of the
earth
{\ADD{are}}
reputed as
nothing: and he
doeth according to his
will in the
army of
heaven, and
{\ADD{among}} the
inhabitants of the
earth: and
none
can
stay his
hand, or
say unto him,
What
doest thou?
\VS{36}At the same
time my
reason
returned unto
me; and for the
glory of my
kingdom, mine
honour and
brightness
returned unto
me; and my
counsellors and my
lords
sought unto
me; and I was
established in my
kingdom, and
excellent
majesty was
added unto me.
\VS{37}Now
I
Nebuchadnezzar
praise and
extol and
honour the
King of
heaven,
all whose
works
{\ADD{are}}
truth, and his
ways
judgment: and those that
walk in
pride he is
able to
abase.

\par }\Chap{5}{\PP \VerseOne{1}Belshazzar the
king
made a
great
feast to a
thousand of his
lords, and
drank
wine
before the
thousand.
\VS{2}Belshazzar, whiles he
tasted the
wine,
commanded to
bring the
golden and
silver
vessels which his
father
Nebuchadnezzar had
taken out
of the
temple which
{\ADD{was}} in
Jerusalem; that the
king, and his
princes, his
wives, and his
concubines, might
drink therein.\FTNT{*}{{\FR 5.2 }
father: or, grandfather}\FTNT{†}{{\FR 5.2 }
taken: Chaldee, brought forth}
\VS{3}Then they
brought the
golden
vessels that were
taken out
of the
temple of the
house of
God which
{\ADD{was}} at
Jerusalem; and the
king, and his
princes, his
wives, and his
concubines,
drank in them.
\VS{4}They
drank
wine, and
praised the
gods of
gold, and of
silver, of
brass, of
iron, of
wood, and of
stone.
\par }{\PP \VS{5}¶ In the same
hour came
forth
fingers of a
man’s
hand, and
wrote over
against the
candlestick
upon the
plaister of the
wall of the
king’s
palace: and the
king
saw the
part of the
hand that
wrote.
\VS{6}Then the
king’s
countenance was
changed, and his
thoughts
troubled him, so that the
joints of his
loins were
loosed, and his
knees
smote
one against
another.\FTNT{‡}{{\FR 5.6 }
countenance: Chaldee, brightnesses}\FTNT{§}{{\FR 5.6 }
was changed: Chaldee, changed it}\FTNT{**}{{\FR 5.6 }
joints: or, girdles: Chaldee, bindings, or, knots}
\VS{7}The
king
cried
aloud to
bring in the
astrologers, the
Chaldeans, and the
soothsayers.
{\ADD{And}} the
king
spake, and
said to the
wise
{\ADD{men}} of
Babylon,
Whosoever shall
read
this
writing, and
shew me the
interpretation thereof, shall be
clothed with
scarlet, and
{\ADD{have}} a
chain of
gold
about his
neck, and shall be the
third
ruler in the
kingdom.\FTNT{††}{{\FR 5.7 }
aloud: Chaldee, with might}\FTNT{‡‡}{{\FR 5.7 }
scarlet: or, purple}
\VS{8}Then
came in
all the
king’s
wise
{\ADD{men:}} but they
could
not
read the
writing, nor make
known to the
king the
interpretation thereof.
\VS{9}Then was
king
Belshazzar
greatly
troubled, and his
countenance was
changed in
him, and his
lords were
astonied.\FTNT{§§}{{\FR 5.9 }
countenance: Chaldee, brightnesses}
\par }{\PP \VS{10}¶
{\ADD{Now}} the
queen, by
reason of the
words of the
king and his
lords,
came into the
banquet
house:
{\ADD{and}} the
queen
spake and
said, O
king,
live for
ever: let
not thy
thoughts
trouble thee, nor let thy
countenance be
changed:
\VS{11}There
is a
man in thy
kingdom, in whom
{\ADD{is}} the
spirit of the
holy
gods; and in the
days of thy
father
light and
understanding and
wisdom, like the
wisdom of the
gods, was
found in him; whom the
king
Nebuchadnezzar thy
father, the
king,
{\ADD{I say}}, thy
father,
made
master of the
magicians,
astrologers,
Chaldeans,
{\ADD{and}}
soothsayers;\FTNT{***}{{\FR 5.11 }
father: or, grandfather}
\VS{12}Forasmuch
as an
excellent
spirit, and
knowledge, and
understanding,
interpreting of
dreams, and
shewing of hard
sentences, and
dissolving of
doubts, were
found in the same
Daniel, whom the
king
named
Belteshazzar:
now let
Daniel be
called, and he will
shew the
interpretation.\FTNT{†††}{{\FR 5.12 }
interpreting: or, of an interpreter}\FTNT{‡‡‡}{{\FR 5.12 }
dissolving: or, of a dissolver}\FTNT{§§§}{{\FR 5.12 }
doubts: Chaldee, knots}
\VS{13}Then was
Daniel brought
in
before the
king.
{\ADD{And}} the
king
spake and
said unto
Daniel,
{\ADD{Art}}
thou that
Daniel, which
{\ADD{art}}
of the
children of the
captivity of
Judah, whom the
king my
father
brought out
of
Jewry?\FTNT{*}{{\FR 5.13 }
father: or, grandfather}
\VS{14}I have even
heard of thee, that the
spirit of the
gods
{\ADD{is}} in
thee, and
{\ADD{that}}
light and
understanding and
excellent
wisdom is
found in thee.
\VS{15}And
now the
wise
{\ADD{men}}, the
astrologers, have been
brought in
before me, that they should
read
this
writing, and make
known unto me the
interpretation thereof: but they
could
not
shew the
interpretation of the
thing:
\VS{16}And
I have
heard of
thee, that thou
canst
make
interpretations, and
dissolve
doubts:
now
if thou
canst
read the
writing, and make
known to me the
interpretation thereof, thou shalt be
clothed with
scarlet, and
{\ADD{have}} a
chain of
gold
about thy
neck, and shalt be the
third
ruler in the
kingdom.\FTNT{†}{{\FR 5.16 }
make interpretations: Chaldee, interpret, etc}
\par }{\PP \VS{17}¶
Then
Daniel
answered and
said
before the
king, Let thy
gifts
be to thyself, and
give thy
rewards to
another;
yet I will
read the
writing unto the
king, and make
known to him the
interpretation.\FTNT{‡}{{\FR 5.17 }
rewards: or, fee}
\VS{18}O
thou
king, the most
high
God
gave
Nebuchadnezzar thy
father a
kingdom, and
majesty, and
glory, and
honour:
\VS{19}And
for the
majesty that he
gave him,
all
people,
nations, and
languages,
trembled and
feared
before
him: whom he
would he
slew; and whom he
would he kept
alive; and whom he
would he set
up; and whom he
would he put
down.
\VS{20}But when his
heart was lifted
up, and his
mind
hardened in
pride, he was
deposed
from his
kingly
throne, and they
took his
glory from
him:\FTNT{§}{{\FR 5.20 }
in pride: or, to deal proudly}\FTNT{**}{{\FR 5.20 }
deposed: Chaldee, made to come down}
\VS{21}And he was
driven
from the
sons of
men; and his
heart was
made
like the
beasts, and his
dwelling
{\ADD{was}} with the wild
asses: they
fed him with
grass like
oxen, and his
body was
wet with the
dew of
heaven;
till he
knew that the most
high
God
ruled in the
kingdom of
men, and
{\ADD{that}} he
appointeth
over it
whomsoever he
will.\FTNT{††}{{\FR 5.21 }
his heart…: or, he made his heart equal, etc}
\VS{22}And
thou his
son, O
Belshazzar, hast
not
humbled thine
heart,
though thou
knewest
all
this;
\VS{23}But hast lifted
up thyself
against the
Lord of
heaven; and they have
brought the
vessels of his
house
before thee, and
thou, and thy
lords, thy
wives, and thy
concubines, have
drunk
wine in them; and thou hast
praised the
gods of
silver, and
gold, of
brass,
iron,
wood, and
stone, which
see
not,
nor
hear,
nor
know: and the
God in whose
hand thy
breath
{\ADD{is}}, and whose
{\ADD{are}}
all thy
ways, hast thou
not
glorified:
\VS{24}Then was the
part of the
hand
sent
from
him; and
this
writing was
written.
\par }{\PP \VS{25}¶ And
this
{\ADD{is}} the
writing that was
written,
MENE,
MENE,
TEKEL,
UPHARSIN.
\VS{26}This
{\ADD{is}} the
interpretation of the
thing:
MENE;
God hath
numbered thy
kingdom, and
finished it.
\VS{27}TEKEL; Thou art
weighed in the
balances, and art
found
wanting.
\VS{28}PERES; Thy
kingdom is
divided, and
given to the
Medes and
Persians.
\VS{29}Then
commanded
Belshazzar, and they
clothed
Daniel with
scarlet, and
{\ADD{put}} a
chain of
gold
about his
neck, and made a
proclamation
concerning him, that he should
be the
third
ruler in the
kingdom.
\par }{\PP \VS{30}¶ In that
night was
Belshazzar the
king of the
Chaldeans
slain.
\VS{31}And
Darius the
Median
took the
kingdom,
{\ADD{being}} about
threescore and
two
years
old.\FTNT{‡‡}{{\FR 5.31 }
being…: Chaldee, he as the son of, etc}\FTNT{§§}{{\FR 5.31 }
about: or, now}

\par }\Chap{6}{\PP \VerseOne{1}It
pleased
Darius to
set
over the
kingdom an
hundred and
twenty
princes, which should
be over the
whole
kingdom;
\VS{2}And
over
these
three
presidents; of
whom
Daniel
{\ADD{was}}
first:
that the
princes
might
give
accounts unto them, and the
king should
have
no
damage.
\VS{3}Then
this
Daniel
was
preferred
above the
presidents and
princes,
because an
excellent
spirit
{\ADD{was}} in him; and the
king
thought to
set him
over the
whole
realm.
\par }{\PP \VS{4}¶
Then the
presidents and
princes
sought to
find
occasion against
Daniel
concerning the
kingdom;
but they
could
find
none
occasion
nor
fault;
forasmuch as he
{\ADD{was}}
faithful,
neither was there
any
error or
fault
found in
him.
\VS{5}Then
said
these
men, We shall
not
find
any
occasion
against
this
Daniel,
except we
find
{\ADD{it}} against him concerning the
law of his
God.
\par }{\PP \VS{6}Then
these
presidents and
princes assembled
together to the
king, and
said
thus unto
him,
King
Darius,
live for
ever.\FTNT{*}{{\FR 6.6 }
assembled…: or, came tumultuously}
\VS{7}All the
presidents of the
kingdom, the
governors, and the
princes, the
counsellors, and the
captains, have consulted
together to
establish a
royal
statute, and to make a
firm
decree, that whosoever shall
ask a
petition
of
any
God or
man
for
thirty
days,
save of
thee, O
king, he shall be
cast into the
den of
lions.\FTNT{†}{{\FR 6.7 }
decree: or, interdict}
\VS{8}Now, O
king,
establish the
decree, and
sign the
writing, that it be
not
changed, according to the
law of the
Medes and
Persians, which
altereth
not.\FTNT{‡}{{\FR 6.8 }
altereth not: Chaldee, passeth not}
\VS{9}Wherefore
king
Darius
signed the
writing and the
decree.
\par }{\PP \VS{10}¶ Now when
Daniel
knew that the
writing was
signed, he
went into his
house; and his
windows being
open in his
chamber
toward
Jerusalem, he
kneeled
upon his
knees
three
times a
day, and
prayed, and gave
thanks
before his
God,
as he
did
aforetime.
\par }{\PP \VS{11}Then
these
men
assembled, and
found
Daniel
praying and making
supplication
before his
God.
\VS{12}Then they came
near, and
spake
before the
king
concerning the
king’s
decree; Hast thou
not
signed a
decree, that
every
man that shall
ask
{\ADD{a petition}}
of
any
God or
man
within
thirty
days,
save of
thee, O
king, shall be
cast into the
den of
lions? The
king
answered and
said, The
thing
{\ADD{is}}
true, according to the
law of the
Medes and
Persians, which
altereth
not.
\VS{13}Then
answered they and
said
before the
king, That
Daniel, which
{\ADD{is}}
of the
children of the
captivity of
Judah,
regardeth
not
thee, O
king, nor the
decree that thou hast
signed, but
maketh his
petition
three
times a
day.
\VS{14}Then the
king, when he
heard
{\ADD{these}}
words, was
sore
displeased with himself, and
set
{\ADD{his}}
heart
on
Daniel to
deliver him: and he
laboured
till the going
down of the
sun to
deliver him.
\VS{15}Then
these
men
assembled
unto the
king, and
said unto the
king,
Know, O
king, that the
law of the
Medes and
Persians
{\ADD{is}}, That
no
decree nor
statute which the
king
establisheth may be
changed.
\VS{16}Then the
king
commanded, and they
brought
Daniel, and
cast
{\ADD{him}} into the
den of
lions.
{\ADD{Now}} the
king
spake and
said unto
Daniel, Thy
God whom
thou
servest
continually, he will
deliver thee.
\VS{17}And
a
stone was
brought, and
laid
upon the
mouth of the
den; and the
king
sealed it with his own
signet, and with the
signet of his
lords; that the
purpose might
not be
changed concerning
Daniel.
\par }{\PP \VS{18}¶
Then the
king
went to his
palace, and passed the
night
fasting:
neither were
instruments of musick
brought
before him: and his
sleep
went from
him.\FTNT{§}{{\FR 6.18 }
instruments…: or, table}
\VS{19}Then the
king
arose very
early in the
morning, and
went in
haste unto the
den of
lions.
\VS{20}And when he
came to the
den, he
cried with a
lamentable
voice unto
Daniel:
{\ADD{and}} the
king
spake and
said to
Daniel, O
Daniel,
servant of the
living
God, is thy
God, whom thou
servest
continually,
able to
deliver thee
from the
lions?
\VS{21}Then
said
Daniel
unto the
king, O
king,
live for
ever.
\VS{22}My
God hath
sent his
angel, and hath
shut the
lions’
mouths, that they have
not
hurt me: forasmuch
as
before him
innocency was
found in me; and
also
before thee, O
king, have I
done
no
hurt.
\VS{23}Then was the
king
exceeding
glad for
him, and
commanded that they should
take
Daniel
up out
of the
den. So
Daniel was taken
up out
of the
den, and
no
manner of
hurt was
found upon him, because he
believed in his
God.
\par }{\PP \VS{24}¶ And the
king
commanded, and they
brought
those
men which had
accused
Daniel, and they
cast
{\ADD{them}} into the
den of
lions,
them, their
children, and their
wives; and the
lions had the
mastery of them, and
brake
all their
bones in
pieces or
ever they
came at the
bottom of the
den.
\par }{\PP \VS{25}¶
Then
king
Darius
wrote unto
all
people,
nations, and
languages, that
dwell in
all the
earth;
Peace be
multiplied unto you.
\VS{26}I
make a
decree, That in
every
dominion of my
kingdom men
tremble and
fear
before the
God of
Daniel: for he
{\ADD{is}} the
living
God, and
stedfast for
ever, and his
kingdom
{\ADD{that}} which shall
not be
destroyed, and his
dominion
{\ADD{shall be even}}
unto the
end.
\VS{27}He
delivereth and
rescueth, and he
worketh
signs and
wonders in
heaven and in
earth, who hath
delivered
Daniel
from the
power of the
lions.\FTNT{**}{{\FR 6.27 }
power: Chaldee, hand}
\VS{28}So
this
Daniel
prospered in the
reign of
Darius, and in the
reign of
Cyrus the
Persian.

\par }\Chap{7}{\PP \VerseOne{1}In the
first
year of
Belshazzar
king of
Babylon
Daniel
had a
dream and
visions of his
head
upon his
bed:
then he
wrote the
dream,
{\ADD{and}}
told the
sum of the
matters.\FTNT{*}{{\FR 7.1 }
had: Chaldee, saw}\FTNT{†}{{\FR 7.1 }
matters: or, words}
\VS{2}Daniel
spake and
said, I
saw in my
vision
by
night, and,
behold, the
four
winds of the
heaven
strove upon the
great
sea.
\VS{3}And
four
great
beasts came
up
from the
sea,
diverse
one
from
another.
\VS{4}The
first
{\ADD{was}} like a
lion, and had
eagle’s
wings: I
beheld
till the
wings thereof were
plucked, and it was lifted
up
from the
earth, and made
stand
upon the
feet as a
man, and a
man’s
heart was
given to it.\FTNT{‡}{{\FR 7.4 }
and it: or, wherewith it}
\VS{5}And
behold
another
beast, a
second,
like to a
bear, and it raised
up itself on
one
side, and
{\ADD{it had}}
three
ribs in the
mouth of it
between the
teeth of it: and they
said
thus unto it,
Arise,
devour
much
flesh.\FTNT{§}{{\FR 7.5 }
it raised…: or, it raised up one dominion}
\VS{6}After
this I
beheld, and
lo
another, like a
leopard, which had
upon the
back of it
four
wings of a
fowl; the
beast had also
four
heads; and
dominion was
given to it.
\VS{7}After
this I
saw in the
night
visions, and
behold a
fourth
beast,
dreadful and
terrible, and
strong
exceedingly; and it had
great
iron
teeth: it
devoured and brake in
pieces, and
stamped the
residue with the
feet of it: and
it
{\ADD{was}}
diverse
from
all the
beasts that
{\ADD{were}}
before it; and it had
ten
horns.
\VS{8}I
considered the
horns, and,
behold, there came
up
among them
another
little
horn,
before whom there were
three
of the
first
horns plucked up by the
roots: and,
behold, in
this
horn
{\ADD{were}}
eyes like the
eyes of
man, and a
mouth
speaking great
things.
\par }{\PP \VS{9}¶ I
beheld
till the
thrones were cast
down, and the
Ancient of
days did
sit, whose
garment
{\ADD{was}}
white as
snow, and the
hair of his
head like the
pure
wool: his
throne
{\ADD{was like}} the
fiery
flame,
{\ADD{and}} his
wheels
{\ADD{as}}
burning
fire.
\VS{10}A
fiery
stream
issued and came
forth
from
before him:
thousand
thousands
ministered unto him, and ten
thousand times ten
thousand
stood
before him: the
judgment was
set, and the
books were
opened.
\VS{11}I
beheld
then
because of the
voice of the
great
words which the
horn
spake: I
beheld
{\ADD{even}}
till the
beast was
slain, and his
body
destroyed, and
given to the
burning
flame.
\VS{12}As concerning the
rest of the
beasts, they had their
dominion taken
away: yet their
lives were
prolonged
for a
season and
time.\FTNT{**}{{\FR 7.12 }
their lives…: Chaldee, a prolonging in life was given them}
\VS{13}I
saw in the
night
visions, and,
behold,
{\ADD{one}} like the
Son of
man
came
with the
clouds of
heaven, and
came
to the
Ancient of
days, and they brought him
near
before him.
\VS{14}And there was
given him
dominion, and
glory, and a
kingdom, that
all
people,
nations, and
languages, should
serve him: his
dominion
{\ADD{is}} an
everlasting
dominion, which shall
not pass
away, and his
kingdom
{\ADD{that}} which shall
not be
destroyed.
\par }{\PP \VS{15}¶
I
Daniel was
grieved in my
spirit in the
midst of
{\ADD{my}}
body, and the
visions of my
head
troubled me.\FTNT{††}{{\FR 7.15 }
body: Chaldee, sheath}
\VS{16}I came
near
unto
one
of them that stood
by, and
asked
him the
truth of
all
this. So he
told me, and made me
know the
interpretation of the
things.
\VS{17}These
great
beasts,
which are
four,
{\ADD{are}}
four
kings,
{\ADD{which}} shall
arise out
of the
earth.
\VS{18}But the
saints of the most
High shall
take the
kingdom, and
possess the
kingdom
for
ever, even
for
ever and
ever.\FTNT{‡‡}{{\FR 7.18 }
most…: Chaldee, high ones, that is, things, or, places}
\VS{19}Then I
would know the
truth
of the
fourth
beast, which
was
diverse
from
all the others,
exceeding
dreadful, whose
teeth
{\ADD{were of}}
iron, and his
nails
{\ADD{of}}
brass;
{\ADD{which}}
devoured, brake in
pieces, and
stamped the
residue with his
feet;\FTNT{§§}{{\FR 7.19 }
from…: Chaldee, from all those}
\VS{20}And
of the
ten
horns that
{\ADD{were}} in his
head, and
{\ADD{of}} the
other which came
up, and
before whom
three
fell; even
{\ADD{of}} that
horn
that had
eyes, and a
mouth that
spake very great
things, whose
look
{\ADD{was}}
more
stout
than his
fellows.
\VS{21}I
beheld, and the
same
horn
made
war
with the
saints, and
prevailed against them;
\VS{22}Until the
Ancient of
days
came, and
judgment was
given to the
saints of the most
High; and the
time
came that the
saints
possessed the
kingdom.
\VS{23}Thus he
said, The
fourth
beast shall
be the
fourth
kingdom upon
earth, which shall be
diverse
from
all
kingdoms, and shall
devour the
whole
earth, and shall tread it
down, and break it in
pieces.
\VS{24}And the
ten
horns out
of this
kingdom
{\ADD{are}}
ten
kings
{\ADD{that}} shall
arise: and
another shall
rise
after them; and he shall be
diverse
from the
first, and he shall
subdue
three
kings.
\VS{25}And he shall
speak
{\ADD{great}}
words
against the most
High, and shall wear
out the
saints of the most
High, and
think to
change
times and
laws: and they shall be
given into his
hand
until a
time and
times and the
dividing of
time.
\VS{26}But the
judgment shall
sit, and they shall take
away his
dominion, to
consume and to
destroy
{\ADD{it}}
unto the
end.
\VS{27}And the
kingdom and
dominion, and the
greatness of the
kingdom
under the
whole
heaven, shall be
given to the
people of the
saints of the most
High, whose
kingdom
{\ADD{is}} an
everlasting
kingdom, and
all
dominions shall
serve and
obey him.\FTNT{***}{{\FR 7.27 }
dominions: or, rulers}
\VS{28}Hitherto
{\ADD{is}} the
end of the
matter. As for
me
Daniel, my
cogitations
much
troubled me, and my
countenance
changed in
me: but I
kept the
matter in my
heart.

\par }\Chap{8}{\PP \VerseOne{1}In the
third
year of the
reign of
king
Belshazzar a
vision
appeared unto me,
{\ADD{even unto}} me
Daniel,
after that which
appeared unto me at the
first.
\VS{2}And I
saw in a
vision; and it came to pass, when I
saw, that I
{\ADD{was}} at
Shushan
{\ADD{in}} the
palace, which
{\ADD{is}} in the
province of
Elam; and I
saw in a
vision, and I was by the
river of
Ulai.
\VS{3}Then I lifted
up mine
eyes, and
saw, and, behold, there
stood
before the
river
a
ram which had
{\ADD{two}}
horns: and the
{\ADD{two}}
horns
{\ADD{were}}
high; but
one
{\ADD{was}}
higher than the
other, and the
higher came
up
last.\FTNT{*}{{\FR 8.3 }
the other: Heb. the second}
\VS{4}I
saw the
ram
pushing
westward, and
northward, and
southward; so that no
beasts might
stand
before him, neither
{\ADD{was there any}} that could
deliver out of his
hand; but he
did according to his
will, and became
great.
\VS{5}And as I was
considering, behold, an
he
goat
came from the
west on the
face of the whole
earth, and
touched not the
ground: and the
goat
{\ADD{had}} a
notable
horn between his
eyes.\FTNT{†}{{\FR 8.5 }
touched…: or, none touched him in the earth}\FTNT{‡}{{\FR 8.5 }
a notable…: Heb. a horn of sight}
\VS{6}And he
came to the
ram that
had
{\ADD{two}}
horns, which I had
seen
standing
before the
river, and
ran unto him in the
fury of his
power.
\VS{7}And I
saw him
come close
unto the
ram, and he was moved with
choler against him, and
smote the
ram, and
brake his
two
horns: and there was no
power in the
ram to
stand
before him, but he cast him
down to the
ground, and
stamped upon him: and there was none that could
deliver the
ram out of his
hand.
\VS{8}Therefore the
he
goat waxed
very
great: and when he was
strong, the
great
horn was
broken; and for it came
up
four notable
ones toward the
four
winds of
heaven.
\VS{9}And out of
one of them came
forth a
little
horn, which waxed
exceeding
great, toward the
south, and toward the
east, and toward the
pleasant
{\ADD{land}}.
\VS{10}And it waxed
great,
{\ADD{even}} to the
host of
heaven; and it cast
down
{\ADD{some}} of the
host and of the
stars to the
ground, and
stamped upon them.\FTNT{§}{{\FR 8.10 }
to the host: or, against the host}
\VS{11}Yea, he
magnified
{\ADD{himself}} even to the
prince of the
host, and by him the
daily
{\ADD{sacrifice}} was taken
away, and the
place of his
sanctuary was cast
down.\FTNT{**}{{\FR 8.11 }
to: or, against}\FTNT{††}{{\FR 8.11 }
by him: or, from him}
\VS{12}And an
host was
given
{\ADD{him}} against the
daily
{\ADD{sacrifice}} by reason of
transgression, and it cast
down the
truth to the
ground; and it
practised, and
prospered.\FTNT{‡‡}{{\FR 8.12 }
an host…: or, the host was given over for the transgression against the daily sacrifice}
\par }{\PP \VS{13}¶ Then I
heard
one
saint
speaking, and
another
saint
said unto that
certain
{\ADD{saint}} which
spake, How long
{\ADD{shall be}} the
vision
{\ADD{concerning}} the
daily
{\ADD{sacrifice}}, and the
transgression of
desolation, to
give both the
sanctuary and the
host to be trodden under
foot?\FTNT{§§}{{\FR 8.13 }
that…: or, the numberer of secrets, or, the wonderful numberer: Heb. Palmoni}\FTNT{***}{{\FR 8.13 }
of…: or, making desolate}
\VS{14}And he
said unto me, Unto two
thousand and
three
hundred
days; then shall the
sanctuary be
cleansed.\FTNT{†††}{{\FR 8.14 }
days: Heb. evening morning}\FTNT{‡‡‡}{{\FR 8.14 }
cleansed: Heb. justified}
\par }{\PP \VS{15}¶ And it came to pass, when I,
{\ADD{even}} I
Daniel, had
seen the
vision, and
sought for the
meaning, then, behold, there
stood before me as the
appearance of a
man.
\VS{16}And I
heard a
man’s
voice between
{\ADD{the banks of}}
Ulai, which
called, and
said,
Gabriel, make
this
{\ADD{man}} to
understand the
vision.
\VS{17}So he
came
near where I
stood: and when he
came, I was
afraid, and
fell upon my
face: but he
said unto me,
Understand, O
son of
man: for at the
time of the
end
{\ADD{shall be}} the
vision.
\VS{18}Now as he was
speaking with me, I was in a deep
sleep on my
face toward the
ground: but he
touched me, and
set me
upright.\FTNT{§§§}{{\FR 8.18 }
set…: Heb. made me stand upon my standing}
\VS{19}And he
said, Behold, I will make thee
know what shall be in the last
end of the
indignation: for at the time
appointed the
end
{\ADD{shall be}}.
\VS{20}The
ram which thou
sawest
having
{\ADD{two}}
horns
{\ADD{are}} the
kings of
Media and
Persia.
\VS{21}And the
rough
goat
{\ADD{is}} the
king of
Grecia: and the
great
horn that
{\ADD{is}} between his
eyes
{\ADD{is}} the
first
king.
\VS{22}Now that being
broken, whereas
four stood
up for it,
four
kingdoms shall stand
up out of the
nation, but not in his
power.
\VS{23}And in the latter
time of their
kingdom, when the
transgressors are come to the
full, a
king of
fierce
countenance, and
understanding dark
sentences, shall stand
up.\FTNT{*}{{\FR 8.23 }
are…: Heb. are accomplished}
\VS{24}And his
power shall be
mighty, but not by his own
power: and he shall
destroy
wonderfully, and shall
prosper, and
practise, and shall
destroy the
mighty and the
holy
people.\FTNT{†}{{\FR 8.24 }
holy…: Heb. people of the holy ones}
\VS{25}And through his
policy also he shall cause
craft to
prosper in his
hand; and he shall
magnify
{\ADD{himself}} in his
heart, and by
peace shall
destroy
many: he shall also stand
up against the
Prince of
princes; but he shall be
broken
without
hand.\FTNT{‡}{{\FR 8.25 }
peace: or, prosperity}
\VS{26}And the
vision of the
evening and the
morning which was
told
{\ADD{is}}
true: wherefore shut thou
up the
vision; for it
{\ADD{shall be}} for
many
days.
\VS{27}And I
Daniel
fainted, and was
sick
{\ADD{certain}}
days; afterward I rose
up, and
did the
king’s
business; and I was
astonished at the
vision, but none
understood
{\ADD{it}}.

\par }\Chap{9}{\PP \VerseOne{1}In the
first
year of
Darius the
son of
Ahasuerus, of the
seed of the
Medes, which was made
king over the
realm of the
Chaldeans;\FTNT{*}{{\FR 9.1 }
which: or, in which he}
\VS{2}In the
first
year of his
reign I
Daniel
understood by
books the
number of the
years, whereof the
word of the
{\ND{
LORD}} came to
Jeremiah the
prophet, that he would
accomplish
seventy
years in the
desolations of
Jerusalem.
\par }{\PP \VS{3}¶ And I
set my
face unto the
Lord
God, to
seek by
prayer and
supplications, with
fasting, and
sackcloth, and
ashes:
\par }{\PP \VS{4}And I
prayed unto the
{\ND{
LORD}} my
God, and made my
confession, and
said,
O
Lord, the
great and
dreadful
God,
keeping the
covenant and
mercy to them that
love him, and to them that
keep his
commandments;
\VS{5}We have
sinned, and have committed
iniquity, and have done
wickedly, and have
rebelled, even by
departing from thy
precepts and from thy
judgments:
\VS{6}Neither have we
hearkened unto thy
servants the
prophets, which
spake in thy
name to our
kings, our
princes, and our
fathers, and to all the
people of the
land.
\VS{7}O
Lord,
righteousness
{\ADD{belongeth}} unto thee, but unto us
confusion of
faces, as at this
day; to the
men of
Judah, and to the
inhabitants of
Jerusalem, and unto all
Israel,
{\ADD{that are}}
near, and
{\ADD{that are}} far
off, through all the
countries whither thou hast
driven them, because of their
trespass that they have
trespassed against thee.\FTNT{†}{{\FR 9.7 }
belongeth…: or, thou hast}
\VS{8}O
Lord, to us
{\ADD{belongeth}}
confusion of
face, to our
kings, to our
princes, and to our
fathers, because we have
sinned against thee.
\VS{9}To the
Lord our
God
{\ADD{belong}}
mercies and
forgivenesses, though we have
rebelled against him;
\VS{10}Neither have we
obeyed the
voice of the
{\ND{
LORD}} our
God, to
walk in his
laws, which he
set
before us
by his
servants the
prophets.
\VS{11}Yea, all
Israel have
transgressed thy
law, even by
departing, that they might not
obey thy
voice; therefore the
curse is
poured upon us, and the
oath that
{\ADD{is}}
written in the
law of
Moses the
servant of
God, because we have
sinned against him.
\VS{12}And he hath
confirmed his
words, which he
spake against us, and against our
judges that
judged us, by
bringing upon us a
great
evil: for under the whole
heaven hath not been
done as hath been
done upon
Jerusalem.
\VS{13}As
{\ADD{it is}}
written in the
law of
Moses, all this
evil is
come upon us: yet made we not our
prayer
before the
{\ND{
LORD}} our
God, that we might
turn from our
iniquities, and
understand thy
truth.\FTNT{‡}{{\FR 9.13 }
made…: Heb. intreated we not the face of the, etc}
\VS{14}Therefore hath the
{\ND{
LORD}}
watched upon the
evil, and
brought it upon us: for the
{\ND{
LORD}} our
God
{\ADD{is}}
righteous in all his
works which he
doeth: for we
obeyed not his
voice.
\VS{15}And now, O
Lord our
God, that hast
brought thy
people
forth out of the
land of
Egypt with a
mighty
hand, and hast
gotten thee
renown, as at this
day; we have
sinned, we have done
wickedly.\FTNT{§}{{\FR 9.15 }
gotten…: Heb. made thee a name}
\par }{\PP \VS{16}¶ O
Lord, according to all thy
righteousness, I beseech thee, let thine
anger and thy
fury be turned
away from thy
city
Jerusalem, thy
holy
mountain: because for our
sins, and for the
iniquities of our
fathers,
Jerusalem and thy
people
{\ADD{are become}} a
reproach to all
{\ADD{that are}}
about us.
\VS{17}Now therefore, O our
God,
hear the
prayer of thy
servant, and his
supplications, and cause thy
face to
shine upon thy
sanctuary that is
desolate, for the
Lord’s sake.
\VS{18}O my
God,
incline thine
ear, and
hear;
open thine
eyes, and
behold our
desolations, and the
city which is
called by thy
name: for we do not
present our
supplications
before thee for our
righteousnesses, but for thy
great
mercies.\FTNT{**}{{\FR 9.18 }
which…: Heb. whereupon thy name is called}\FTNT{††}{{\FR 9.18 }
present: Heb. cause to fall}
\VS{19}O
Lord,
hear; O
Lord,
forgive; O
Lord,
hearken and
do;
defer
not, for thine own sake, O my
God: for thy
city and thy
people are
called by thy
name.
\par }{\PP \VS{20}¶ And whiles I
{\ADD{was}}
speaking, and
praying, and
confessing my
sin and the
sin of my
people
Israel, and
presenting my
supplication
before the
{\ND{
LORD}} my
God for the
holy
mountain of my
God;
\VS{21}Yea, whiles I
{\ADD{was}}
speaking in
prayer, even the
man
Gabriel, whom I had
seen in the
vision at the
beginning, being caused to
fly
swiftly,
touched me about the
time of the
evening
oblation.\FTNT{‡‡}{{\FR 9.21 }
swiftly: Heb. with weariness, or, flight}
\VS{22}And he
informed
{\ADD{me}}, and
talked with me, and
said, O
Daniel, I am now come
forth to give thee
skill and
understanding.\FTNT{§§}{{\FR 9.22 }
to…: Heb. to make thee skilful of}
\VS{23}At the
beginning of thy
supplications the
commandment came
forth, and I am
come to
shew
{\ADD{thee;}} for thou
{\ADD{art}} greatly
beloved: therefore
understand the
matter, and
consider the
vision.\FTNT{***}{{\FR 9.23 }
commandment: Heb. word}\FTNT{†††}{{\FR 9.23 }
greatly…: Heb. a man of desires}
\VS{24}Seventy
weeks are
determined upon thy
people and upon thy
holy
city, to
finish the
transgression, and to make an
end of
sins, and to make
reconciliation for
iniquity, and to bring
in
everlasting
righteousness, and to seal
up the
vision and
prophecy, and to
anoint the
most
Holy.\FTNT{‡‡‡}{{\FR 9.24 }
finish: or, restrain}\FTNT{§§§}{{\FR 9.24 }
make an…: or, seal up}\FTNT{*}{{\FR 9.24 }
prophecy: Heb. prophet}
\VS{25}Know therefore and
understand,
{\ADD{that}} from the going
forth of the
commandment to
restore and to
build
Jerusalem unto the
Messiah the
Prince
{\ADD{shall be}}
seven
weeks, and
threescore and
two
weeks: the
street shall be
built
again, and the
wall, even in
troublous
times.\FTNT{†}{{\FR 9.25 }
be built…: Heb. return and be built}\FTNT{‡}{{\FR 9.25 }
wall: or, breach, or, ditch}\FTNT{§}{{\FR 9.25 }
troublous: Heb. strait of}
\VS{26}And
after
threescore and
two
weeks shall
Messiah be cut
off, but not for himself: and the
people of the
prince that shall
come shall
destroy the
city and the
sanctuary; and the
end thereof
{\ADD{shall be}} with a
flood, and unto the
end of the
war
desolations are
determined.\FTNT{**}{{\FR 9.26 }
but…: or, and shall have nothing}\FTNT{††}{{\FR 9.26 }
desolations…: or, it shall be cut off by desolations}
\VS{27}And he shall
confirm the
covenant with
many for
one
week: and in the
midst of the
week he shall cause the
sacrifice and the
oblation to
cease, and for the
overspreading of
abominations he shall make
{\ADD{it}}
desolate, even until the
consummation, and that
determined shall be
poured upon the
desolate.\FTNT{‡‡}{{\FR 9.27 }
the covenant: or, a covenant}\FTNT{§§}{{\FR 9.27 }
for the…: or, with the abominable armies}\FTNT{***}{{\FR 9.27 }
the desolate: or, the desolator}

\par }\Chap{10}{\PP \VerseOne{1}In the
third
year of
Cyrus
king of
Persia a
thing was
revealed unto
Daniel, whose
name was
called
Belteshazzar; and the
thing
{\ADD{was}}
true, but the time
appointed
{\ADD{was}}
long: and he
understood the
thing, and had
understanding of the
vision.\FTNT{*}{{\FR 10.1 }
long: Heb. great}
\VS{2}In those
days I
Daniel was
mourning
three
full
weeks.\FTNT{†}{{\FR 10.2 }
full…: Heb. weeks of days}
\VS{3}I
ate no
pleasant
bread, neither
came
flesh nor
wine in my
mouth, neither did I
anoint myself at
all, till
three
whole
weeks were
fulfilled.\FTNT{‡}{{\FR 10.3 }
pleasant…: Heb. bread of desires}
\VS{4}And in the
four and
twentieth
day of the
first
month, as I was by the
side of the
great
river, which
{\ADD{is}}
Hiddekel;
\VS{5}Then I lifted
up mine
eyes, and
looked, and behold a
certain
man
clothed in
linen, whose
loins
{\ADD{were}}
girded with fine
gold of
Uphaz:\FTNT{§}{{\FR 10.5 }
a…: Heb. one man}
\VS{6}His
body also
{\ADD{was}} like the
beryl, and his
face as the
appearance of
lightning, and his
eyes as
lamps of
fire, and his
arms and his
feet like in
colour to
polished
brass, and the
voice of his
words like the
voice of a
multitude.
\VS{7}And I
Daniel alone
saw the
vision: for the
men that were with me
saw not the
vision;
but a
great
quaking
fell upon them, so that they
fled to
hide themselves.
\VS{8}Therefore I was left
alone, and
saw this
great
vision, and there
remained no
strength in me: for my
comeliness was
turned in me into
corruption, and I
retained no
strength.\FTNT{**}{{\FR 10.8 }
comeliness: or, vigour}
\VS{9}Yet
heard I the
voice of his
words: and when I
heard the
voice of his
words, then was I in a deep
sleep on my
face, and my
face toward the
ground.
\par }{\PP \VS{10}¶ And, behold, an
hand
touched me, which
set me upon my
knees and
{\ADD{upon}} the
palms of my
hands.\FTNT{††}{{\FR 10.10 }
set: Heb. moved}
\VS{11}And he
said unto me, O
Daniel, a
man greatly
beloved,
understand the
words that I
speak unto thee, and
stand
upright: for unto thee am I now
sent. And when he had
spoken this
word unto me, I
stood
trembling.\FTNT{‡‡}{{\FR 10.11 }
greatly…: Heb. of desires}\FTNT{§§}{{\FR 10.11 }
upright: Heb. upon thy standing}
\VS{12}Then
said he unto me,
Fear not,
Daniel: for from the
first
day that thou didst
set thine
heart to
understand, and to
chasten thyself
before thy
God, thy
words were
heard, and I am
come for thy
words.
\VS{13}But the
prince of the
kingdom of
Persia
withstood me
one and
twenty
days: but, lo,
Michael,
one of the
chief
princes,
came to
help me; and I
remained there
with the
kings of
Persia.\FTNT{***}{{\FR 10.13 }
chief: or, first}
\VS{14}Now I am
come to make thee
understand what shall
befall thy
people in the
latter
days: for yet the
vision
{\ADD{is}} for
{\ADD{many}}
days.
\VS{15}And when he had
spoken such
words unto me, I
set my
face toward the
ground, and I became
dumb.
\VS{16}And, behold,
{\ADD{one}} like the
similitude of the
sons of
men
touched my
lips: then I
opened my
mouth, and
spake, and
said unto him that
stood before me, O my
lord, by the
vision my
sorrows are
turned upon me, and I have
retained no
strength.
\VS{17}For
how
can the
servant of this my
lord
talk with this my
lord? for as for me,
straightway there
remained no
strength in me, neither is there
breath
left in me.\FTNT{†††}{{\FR 10.17 }
the…: or, this servant of my lord}
\VS{18}Then there came
again and
touched me
{\ADD{one}} like the
appearance of a
man, and he
strengthened me,
\VS{19}And
said, O
man greatly
beloved,
fear not:
peace
{\ADD{be}} unto thee, be
strong, yea, be
strong. And when he had
spoken unto me, I was
strengthened, and
said, Let my
lord
speak; for thou hast
strengthened me.
\VS{20}Then
said he,
Knowest thou wherefore I
come unto thee? and now will I
return to
fight with the
prince of
Persia: and when I am gone
forth, lo, the
prince of
Grecia shall
come.
\VS{21}But I will
shew thee that which is
noted in the
scripture of
truth: and
{\ADD{there is}}
none that
holdeth with me in these things, but
Michael your
prince.\FTNT{‡‡‡}{{\FR 10.21 }
holdeth: Heb. strengtheneth himself}

\par }\Chap{11}{\PP \VerseOne{1}Also I in the
first
year of
Darius the
Mede,
{\ADD{even}} I,
stood to
confirm and to
strengthen him.
\VS{2}And now will I
shew thee the
truth. Behold, there shall stand
up yet
three
kings in
Persia; and the
fourth shall be
far
richer than
{\ADD{they}}
all: and by his
strength through his
riches he shall stir
up all against the
realm of
Grecia.
\VS{3}And a
mighty
king shall stand
up, that shall
rule with
great
dominion, and
do according to his
will.
\VS{4}And when he shall stand
up, his
kingdom shall be
broken, and shall be
divided toward the
four
winds of
heaven; and not to his
posterity, nor according to his
dominion which he
ruled: for his
kingdom shall be plucked
up, even for
others beside those.
\par }{\PP \VS{5}¶ And the
king of the
south shall be
strong, and
{\ADD{one}} of his
princes; and he shall be
strong above him, and have
dominion; his
dominion
{\ADD{shall be}} a
great
dominion.
\VS{6}And in the
end of
years they shall join themselves
together; for the
king’s
daughter of the
south shall
come to the
king of the
north to
make an
agreement: but she shall not
retain the
power of the
arm; neither shall he
stand, nor his
arm: but she shall be given
up, and they that
brought her, and he that
begat her, and he that
strengthened her in
{\ADD{these}}
times.\FTNT{*}{{\FR 11.6 }
join…: Heb. associate themselves}\FTNT{†}{{\FR 11.6 }
an…: Heb. rights}\FTNT{‡}{{\FR 11.6 }
he that begat…: or, whom she brought forth}
\VS{7}But out of a
branch of her
roots shall
{\ADD{one}} stand
up in his
estate, which shall
come with an
army, and shall
enter into the
fortress of the
king of the
north, and shall
deal against them, and shall
prevail:\FTNT{§}{{\FR 11.7 }
in…: or, in his place, or, office}
\VS{8}And shall also
carry
captives into
Egypt their
gods, with their
princes,
{\ADD{and}} with their
precious
vessels of
silver and of
gold; and he shall
continue
{\ADD{more}}
years than the
king of the
north.\FTNT{**}{{\FR 11.8 }
their precious…: Heb. vessels of their desire}
\VS{9}So the
king of the
south shall
come into
{\ADD{his}}
kingdom, and shall
return into his own
land.
\VS{10}But his
sons shall be stirred
up, and shall
assemble a
multitude of
great
forces: and
{\ADD{one}} shall
certainly
come, and
overflow, and pass
through: then shall he
return, and be stirred
up,
{\ADD{even}} to his
fortress.\FTNT{††}{{\FR 11.10 }
shall be…: or, shall war}\FTNT{‡‡}{{\FR 11.10 }
return…: or, be stirred up again}
\VS{11}And the
king of the
south shall be moved with
choler, and shall come
forth and
fight with him,
{\ADD{even}} with the
king of the
north: and he shall set
forth a
great
multitude; but the
multitude shall be
given into his
hand.
\VS{12}{\ADD{And}} when he hath taken
away the
multitude, his
heart shall be lifted
up; and he shall cast
down
{\ADD{many}} ten
thousands: but he shall not be
strengthened
{\ADD{by it}}.
\VS{13}For the
king of the
north shall
return, and shall set
forth a
multitude
greater than the
former, and shall
certainly
come
after
certain
years with a
great
army and with
much
riches.\FTNT{§§}{{\FR 11.13 }
after…: Heb. at the end of times, even years}
\VS{14}And in those
times there shall
many stand
up against the
king of the
south: also the
robbers of thy
people shall
exalt themselves to
establish the
vision; but they shall
fall.\FTNT{***}{{\FR 11.14 }
the robbers: Heb. the children of robbers}
\VS{15}So the
king of the
north shall
come, and cast
up a
mount, and
take the most
fenced
cities: and the
arms of the
south shall not
withstand, neither his
chosen
people, neither
{\ADD{shall there be any}}
strength to
withstand.\FTNT{†††}{{\FR 11.15 }
the most…: Heb. the city of munitions}\FTNT{‡‡‡}{{\FR 11.15 }
his…: Heb. the people of his choices}
\VS{16}But he that
cometh against him shall
do according to his own
will, and none shall
stand
before him: and he shall
stand in the
glorious
land, which by his
hand shall be
consumed.\FTNT{§§§}{{\FR 11.16 }
glorious…: or, goodly, etc.: Heb. land of ornament}
\VS{17}He shall also
set his
face to
enter with the
strength of his whole
kingdom, and upright
ones with him; thus shall he
do: and he shall
give him the
daughter of
women,
corrupting her: but she shall not
stand
{\ADD{on his side}}, neither be for him.\FTNT{*}{{\FR 11.17 }
upright…: or, much uprightness: or, equal conditions}\FTNT{†}{{\FR 11.17 }
corrupting: Heb. to corrupt}
\VS{18}After this shall he
turn his
face unto the
isles, and shall
take
many: but a
prince for his own behalf shall cause the
reproach offered by him to
cease;
without his own
reproach he shall cause
{\ADD{it}} to
turn upon him.\FTNT{‡}{{\FR 11.18 }
for…: Heb. for him}\FTNT{§}{{\FR 11.18 }
the reproach…: Heb. his reproach}
\VS{19}Then he shall
turn his
face toward the
fort of his own
land: but he shall
stumble and
fall, and not be
found.
\VS{20}Then shall stand
up in his
estate a
raiser of
taxes
{\ADD{in}} the
glory of the
kingdom: but within
few
days he shall be
destroyed, neither in
anger, nor in
battle.\FTNT{**}{{\FR 11.20 }
estate: or, place}\FTNT{††}{{\FR 11.20 }
a…: Heb. one that causeth an exacter to pass over}\FTNT{‡‡}{{\FR 11.20 }
anger: Heb. angers}
\par }{\PP \VS{21}And in his
estate shall stand
up a vile
person, to whom they shall not
give the
honour of the
kingdom: but he shall
come in
peaceably, and
obtain the
kingdom by
flatteries.\FTNT{§§}{{\FR 11.21 }
estate: or, place}
\VS{22}And with the
arms of a
flood shall they be
overflown from
before him, and shall be
broken; yea, also the
prince of the
covenant.
\VS{23}And after the
league
{\ADD{made}} with him he shall
work
deceitfully: for he shall come
up, and shall become
strong with a
small
people.
\VS{24}He shall
enter
peaceably even upon the fattest
places of the
province; and he shall
do
{\ADD{that}} which his
fathers have not
done, nor his
fathers’
fathers; he shall
scatter among them the
prey, and
spoil, and
riches:
{\ADD{yea}}, and he shall
forecast his
devices against the strong
holds, even for a
time.\FTNT{***}{{\FR 11.24 }
peaceably…: or, into the peaceable and fat, etc}\FTNT{†††}{{\FR 11.24 }
forecast…: Heb. think his thoughts}
\VS{25}And he shall stir
up his
power and his
courage against the
king of the
south with a
great
army; and the
king of the
south shall be stirred
up to
battle with a
very
great and
mighty
army; but he shall not
stand: for they shall
forecast
devices against him.
\VS{26}Yea, they that
feed of the portion of his
meat shall
destroy him, and his
army shall
overflow: and
many shall
fall down
slain.
\VS{27}And
both these
kings’
hearts
{\ADD{shall be}} to do
mischief, and they shall
speak
lies at
one
table; but it shall not
prosper: for yet the
end
{\ADD{shall be}} at the time
appointed.\FTNT{‡‡‡}{{\FR 11.27 }
hearts: Heb. their hearts}
\VS{28}Then shall he
return into his
land with
great
riches; and his
heart
{\ADD{shall be}} against the
holy
covenant; and he shall
do
{\ADD{exploits}}, and
return to his own
land.
\VS{29}At the time
appointed he shall
return, and
come toward the
south; but it shall not be as the
former, or as the
latter.
\par }{\PP \VS{30}¶ For the
ships of
Chittim shall
come against him: therefore he shall be
grieved, and
return, and have
indignation against the
holy
covenant: so shall he
do; he shall even
return, and have
intelligence with them that
forsake the
holy
covenant.
\VS{31}And
arms shall
stand on his part, and they shall
pollute the
sanctuary of
strength, and shall take
away the
daily
{\ADD{sacrifice}}, and they shall
place the
abomination that maketh
desolate.\FTNT{§§§}{{\FR 11.31 }
maketh…: or, astonisheth}
\VS{32}And such as do
wickedly against the
covenant shall he
corrupt by
flatteries: but the
people that do
know their
God shall be
strong, and
do
{\ADD{exploits}}.\FTNT{*}{{\FR 11.32 }
corrupt: or, cause to dissemble}
\VS{33}And they that
understand among the
people shall
instruct
many: yet they shall
fall by the
sword, and by
flame, by
captivity, and by
spoil,
{\ADD{many}}
days.
\VS{34}Now when they shall
fall, they shall be
holpen with a
little
help: but
many shall
cleave to them with
flatteries.
\VS{35}And
{\ADD{some}} of them of
understanding shall
fall, to
try them, and to
purge, and to make
{\ADD{them}}
white,
{\ADD{even}} to the
time of the
end: because
{\ADD{it is}} yet for a time
appointed.\FTNT{†}{{\FR 11.35 }
try them: or, try by them}
\VS{36}And the
king shall
do according to his
will; and he shall
exalt himself, and
magnify himself above every
god, and shall
speak marvellous
things against the
God of
gods, and shall
prosper till the
indignation be
accomplished: for that that is
determined shall be
done.
\VS{37}Neither shall he
regard the
God of his
fathers, nor the
desire of
women, nor
regard any
god: for he shall
magnify himself above all.
\VS{38}But in his
estate shall he
honour the
God of
forces: and a
god whom his
fathers
knew not shall he
honour with
gold, and
silver, and with
precious
stones, and pleasant
things.\FTNT{‡}{{\FR 11.38 }
in…: or, as for the Almighty God, in his seat he shall honour yea, he shall honour a god, whom, etc}\FTNT{§}{{\FR 11.38 }
estate: or, stead}\FTNT{**}{{\FR 11.38 }
forces: or, munitions: Heb. Mauzzim, or, God’s protectors}\FTNT{††}{{\FR 11.38 }
pleasant: Heb. things desired}
\VS{39}Thus shall he
do in the most
strong
holds with a
strange
god, whom he shall
acknowledge
{\ADD{and}}
increase with
glory: and he shall cause them to
rule over
many, and shall
divide the
land for
gain.\FTNT{‡‡}{{\FR 11.39 }
most…: Heb. fortresses of munitions}\FTNT{§§}{{\FR 11.39 }
gain: Heb. a price}
\VS{40}And at the
time of the
end shall the
king of the
south
push at him: and the
king of the
north shall come against him like a
whirlwind, with
chariots, and with
horsemen, and with
many
ships; and he shall
enter into the
countries, and shall
overflow and pass
over.
\VS{41}He shall
enter also into the
glorious
land, and
many
{\ADD{countries}} shall be
overthrown: but these shall
escape out of his
hand,
{\ADD{even}}
Edom, and
Moab, and the
chief of the
children of
Ammon.\FTNT{***}{{\FR 11.41 }
glorious…: or, goodly, etc.: Heb. land of delight, or, ornament}
\VS{42}He shall stretch
forth his
hand also upon the
countries: and the
land of
Egypt shall not
escape.\FTNT{†††}{{\FR 11.42 }
stretch…: Heb. send forth}
\VS{43}But he shall have
power over the
treasures of
gold and of
silver, and over all the
precious things of
Egypt: and the
Libyans and the
Ethiopians
{\ADD{shall be}} at his
steps.
\VS{44}But
tidings out of the
east and out of the
north shall
trouble him: therefore he shall go
forth with
great
fury to
destroy, and utterly to make
away
many.
\VS{45}And he shall
plant the
tabernacles of his
palace between the
seas in the
glorious
holy
mountain; yet he shall
come to his
end, and none shall
help him.\FTNT{‡‡‡}{{\FR 11.45 }
glorious…: or, goodly, etc.: Heb. mountain of delight of holiness}

\par }\Chap{12}{\PP \VerseOne{1}And at that
time shall
Michael stand
up, the
great
prince which
standeth for the
children of thy
people: and there shall
be a
time of
trouble, such as never was since there was a
nation
{\ADD{even}} to that same
time: and at that
time thy
people shall be
delivered, every one that shall be
found
written in the
book.
\VS{2}And
many of them that
sleep in the
dust of the
earth shall
awake, some to
everlasting
life, and some to
shame
{\ADD{and}}
everlasting
contempt.
\VS{3}And they that be
wise shall
shine as the
brightness of the
firmament; and they that turn
many to
righteousness as the
stars for
ever and
ever.\FTNT{*}{{\FR 12.3 }
wise: or, teachers}
\VS{4}But thou, O
Daniel, shut
up the
words, and
seal the
book,
{\ADD{even}} to the
time of the
end:
many shall run to and
fro, and
knowledge shall be
increased.
\par }{\PP \VS{5}¶ Then I
Daniel
looked, and, behold, there
stood
other
two, the
one on this
side of the
bank of the
river, and the
other on that side of the
bank of the
river.\FTNT{†}{{\FR 12.5 }
bank: Heb. lip}\FTNT{‡}{{\FR 12.5 }
bank: Heb. lip}
\VS{6}And
{\ADD{one}}
said to the
man
clothed in
linen, which
{\ADD{was}}
upon the
waters of the
river, How long
{\ADD{shall it be to}} the
end of these
wonders?\FTNT{§}{{\FR 12.6 }
upon…: or, from above}
\VS{7}And I
heard the
man
clothed in
linen, which
{\ADD{was}}
upon the
waters of the
river, when he held
up his right
hand and his left
hand unto
heaven, and
sware by him that
liveth for
ever that
{\ADD{it shall be}} for a
time,
times, and an
half; and when he shall have
accomplished to
scatter the
power of the
holy
people, all these
{\ADD{things}} shall be
finished.\FTNT{**}{{\FR 12.7 }
a time…: or, part}
\VS{8}And I
heard, but I
understood not: then
said I, O my
Lord, what
{\ADD{shall be}} the
end of these
{\ADD{things}}?
\VS{9}And he
said, Go thy
way,
Daniel: for the
words
{\ADD{are}} closed
up and
sealed till the
time of the
end.
\VS{10}Many shall be
purified, and made
white, and
tried; but the
wicked shall do
wickedly: and none of the
wicked shall
understand; but the
wise shall
understand.
\VS{11}And from the
time
{\ADD{that}} the
daily
{\ADD{sacrifice}} shall be taken
away, and the
abomination that maketh
desolate set
up,
{\ADD{there shall be}} a
thousand two
hundred and
ninety
days.\FTNT{††}{{\FR 12.11 }
the abomination: Heb. to set up the abomination}\FTNT{‡‡}{{\FR 12.11 }
maketh…: or, astonisheth}
\VS{12}Blessed
{\ADD{is}} he that
waiteth, and
cometh to the
thousand
three
hundred and
five and
thirty
days.
\VS{13}But go thou thy
way till the
end
{\ADD{be:}} for thou shalt
rest, and
stand in thy
lot at the
end of the
days.\FTNT{§§}{{\FR 12.13 }
for thou: or, and thou, etc}
\par }