\NormalFont\ShortTitle{Deuteronomy}
{\MT The Fifth Book of Moses, called Deuteronomy

\par }\ChapOne{1}{\PP \VerseOne{1}These
{\ADD{be}} the
words which
Moses
spake unto all
Israel on this
side
Jordan in the
wilderness, in the
plain over
against the
Red
{\ADD{sea}}, between
Paran, and
Tophel, and
Laban, and
Hazeroth, and
Dizahab.\FTNT{*}{{\FR 1.1 }
the Red…: or, Zuph}
\VS{2}({\ADD{There are}}
eleven
days’
{\ADD{journey}} from
Horeb by the
way of
mount
Seir unto
Kadesh-barnea.)
\VS{3}And it came to pass in the
fortieth
year, in the
eleventh
month, on the
first
{\ADD{day}} of the
month,
{\ADD{that}}
Moses
spake unto the
children of
Israel, according unto all that the
{\ND{
LORD}} had given him in
commandment unto them;
\VS{4}After he had
slain
Sihon the
king of the
Amorites, which
dwelt in
Heshbon, and
Og the
king of
Bashan, which
dwelt at
Astaroth in
Edrei:
\VS{5}On this
side
Jordan, in the
land of
Moab,
began
Moses to
declare this
law,
saying,
\VS{6}The
{\ND{
LORD}} our
God
spake unto us in
Horeb,
saying, Ye have
dwelt long
enough in this
mount:
\VS{7}Turn you, and take your
journey, and
go to the
mount of the
Amorites, and unto all
{\ADD{the places}}
nigh thereunto, in the
plain, in the
hills, and in the
vale, and in the
south, and by the
sea
side, to the
land of the
Canaanites, and unto
Lebanon, unto the
great
river, the
river
Euphrates.\FTNT{†}{{\FR 1.7 }
all…: Heb. all his neighbours}
\VS{8}Behold, I have
set the
land
before you:
go in and
possess the
land which the
{\ND{
LORD}}
sware unto your
fathers,
Abraham,
Isaac, and
Jacob, to
give unto them and to their
seed
after them.\FTNT{‡}{{\FR 1.8 }
set: Heb. given}
\par }{\PP \VS{9}¶ And I
spake unto you at that
time,
saying, I am not
able to
bear you myself
alone:
\VS{10}The
{\ND{
LORD}} your
God hath
multiplied you, and, behold, ye
{\ADD{are}} this
day as the
stars of
heaven for
multitude.
\VS{11}(The
{\ND{
LORD}}
God of your
fathers make you a
thousand
times so many more
as ye
{\ADD{are}}, and
bless you, as he hath
promised you!)
\VS{12}How can I myself alone
bear your
cumbrance, and your
burden, and your
strife?
\VS{13}Take you
wise
men, and
understanding, and
known among your
tribes, and I will
make them
rulers over you.\FTNT{§}{{\FR 1.13 }
Take: Heb. Give}
\VS{14}And ye
answered me, and
said, The
thing which thou hast
spoken
{\ADD{is}}
good
{\ADD{for us}} to
do.
\VS{15}So I
took the
chief of your
tribes,
wise
men, and
known, and
made them
heads over you,
captains over
thousands, and
captains over
hundreds, and
captains over
fifties, and
captains over
tens, and
officers among your
tribes.\FTNT{**}{{\FR 1.15 }
made: Heb. gave}
\VS{16}And I
charged your
judges at that
time,
saying,
Hear
{\ADD{the causes}} between your
brethren, and
judge
righteously between
{\ADD{every}}
man and his
brother, and the
stranger
{\ADD{that is}} with him.
\VS{17}Ye shall not
respect
persons in
judgment;
{\ADD{but}} ye shall
hear the
small as well as the
great; ye shall not be
afraid of the
face of
man; for the
judgment
{\ADD{is}}
God’s: and the
cause that is too
hard for you,
bring
{\ADD{it}} unto me, and I will
hear it.\FTNT{††}{{\FR 1.17 }
respect…: Heb. acknowledge faces}
\VS{18}And I
commanded you at that
time all the
things which ye should
do.
\par }{\PP \VS{19}¶ And when we
departed from
Horeb, we went
through all that
great and
terrible
wilderness, which ye
saw by the
way of the
mountain of the
Amorites, as the
{\ND{
LORD}} our
God
commanded us; and we
came to
Kadesh-barnea.
\VS{20}And I
said unto you, Ye are
come unto the
mountain of the
Amorites, which the
{\ND{
LORD}} our
God doth
give unto us.
\VS{21}Behold, the
{\ND{
LORD}} thy
God hath
set the
land
before thee: go
up
{\ADD{and}}
possess
{\ADD{it}}, as the
{\ND{
LORD}}
God of thy
fathers hath
said unto thee;
fear not, neither be
discouraged.
\par }{\PP \VS{22}¶ And ye came
near unto me every one of you, and
said, We will
send
men
before us, and they shall search us
out the
land, and
bring us
word
again by what
way we must go
up, and into what
cities we shall
come.
\VS{23}And the
saying pleased me
well: and I
took
twelve
men of you,
one of a
tribe:
\VS{24}And they
turned and went
up into the
mountain, and
came unto the
valley of
Eshcol, and searched it
out.
\VS{25}And they
took of the
fruit of the
land in their
hands, and brought
{\ADD{it}}
down unto us, and
brought us
word
again, and
said,
{\ADD{It is}} a
good
land which the
{\ND{
LORD}} our
God doth
give us.
\VS{26}Notwithstanding ye
would not go
up, but
rebelled against the
commandment of the
{\ND{
LORD}} your
God:
\VS{27}And ye
murmured in your
tents, and
said, Because the
{\ND{
LORD}}
hated us, he hath brought us
forth out of the
land of
Egypt, to
deliver us into the
hand of the
Amorites, to
destroy us.
\VS{28}Whither shall we go
up? our
brethren have
discouraged our
heart,
saying, The
people
{\ADD{is}}
greater and
taller than we; the
cities
{\ADD{are}}
great and walled
up to
heaven; and moreover we have
seen the
sons of the
Anakims there.\FTNT{‡‡}{{\FR 1.28 }
discouraged: Heb. melted}
\VS{29}Then I
said unto you,
Dread not, neither be
afraid of them.
\VS{30}The
{\ND{
LORD}} your
God which
goeth
before you, he shall
fight for you, according to all that he
did for you in
Egypt before your
eyes;
\VS{31}And in the
wilderness, where thou hast
seen how that the
{\ND{
LORD}} thy
God
bare thee, as a
man doth
bear his
son, in all the
way that ye
went, until ye
came into this
place.
\VS{32}Yet in this
thing ye did not
believe the
{\ND{
LORD}} your
God,
\VS{33}Who
went in the
way
before you, to search you
out a
place to pitch your
tents
{\ADD{in}}, in
fire by
night, to
shew you by what
way ye should
go, and in a
cloud by
day.
\VS{34}And the
{\ND{
LORD}}
heard the
voice of your
words, and was
wroth, and
sware,
saying,
\VS{35}Surely there shall not
one of these
men of this
evil
generation
see that
good
land, which I
sware to
give unto your
fathers,
\VS{36}Save
Caleb the
son of
Jephunneh; he shall
see it, and to him will I
give the
land
that he hath trodden
upon, and to his
children, because he hath
wholly
followed the
{\ND{
LORD}}.\FTNT{§§}{{\FR 1.36 }
wholly…: Heb. fulfilled to go after}
\VS{37}Also the
{\ND{
LORD}} was
angry with me for your
sakes,
saying, Thou also shalt not
go in
thither.
\VS{38}{\ADD{But}}
Joshua the
son of
Nun, which
standeth
before thee, he shall go
in thither:
encourage him: for he shall cause
Israel to
inherit it.
\VS{39}Moreover your little
ones, which ye
said should be a
prey, and your
children, which in that
day had no
knowledge between
good and
evil, they shall go
in thither, and unto them will I
give it, and they shall
possess it.
\VS{40}But
{\ADD{as for}} you,
turn you, and take your
journey into the
wilderness by the
way of the
Red
sea.
\VS{41}Then ye
answered and
said unto me, We have
sinned against the
{\ND{
LORD}}, we will go
up and
fight, according to all that the
{\ND{
LORD}} our
God
commanded us. And when ye had
girded on every
man his
weapons of
war, ye were
ready to go
up into the
hill.
\VS{42}And the
{\ND{
LORD}}
said unto me,
Say unto them, Go not
up, neither
fight; for I
{\ADD{am}} not
among you; lest ye be
smitten
before your
enemies.
\VS{43}So I
spake unto you; and ye would not
hear, but
rebelled against the
commandment of the
{\ND{
LORD}}, and
went
presumptuously
up into the
hill.\FTNT{***}{{\FR 1.43 }
went…: Heb. ye were presumptuous, and went up}
\VS{44}And the
Amorites, which
dwelt in that
mountain, came
out
against you, and
chased you, as
bees
do, and
destroyed you in
Seir,
{\ADD{even}} unto
Hormah.
\VS{45}And ye
returned and
wept
before the
{\ND{
LORD}}; but the
{\ND{
LORD}} would not
hearken to your
voice, nor give
ear unto you.
\VS{46}So ye
abode in
Kadesh
many
days, according unto the
days that ye
abode
{\ADD{there}}.

\par }\Chap{2}{\PP \VerseOne{1}Then we
turned, and took our
journey into the
wilderness by the
way of the
Red
sea, as the
{\ND{
LORD}}
spake unto me: and we
compassed
mount
Seir
many
days.
\VS{2}And the
{\ND{
LORD}}
spake unto me,
saying,
\VS{3}Ye have
compassed this
mountain long
enough:
turn you
northward.
\VS{4}And
command thou the
people,
saying, Ye
{\ADD{are}} to
pass through the
coast of your
brethren the
children of
Esau, which
dwell in
Seir; and they shall be
afraid of you: take ye
good
heed unto yourselves therefore:
\VS{5}Meddle not with them; for I will not
give you of their
land, no, not so much as a
foot
breadth; because I have
given
mount
Seir unto
Esau
{\ADD{for}} a
possession.\FTNT{*}{{\FR 2.5 }
no…: Heb. even to the treading of the sole of the foot}
\VS{6}Ye shall
buy
meat of them for
money, that ye may
eat; and ye shall also
buy
water of them for
money, that ye may
drink.
\VS{7}For the
{\ND{
LORD}} thy
God hath
blessed thee in all the
works of thy
hand: he
knoweth thy
walking through this
great
wilderness: these
forty
years the
{\ND{
LORD}} thy
God
{\ADD{hath been}} with thee; thou hast
lacked
nothing.
\par }{\PP \VS{8}And when we
passed by from our
brethren the
children of
Esau, which
dwelt in
Seir, through the
way of the
plain from
Elath, and from
Ezion-gaber, we
turned and
passed by the
way of the
wilderness of
Moab.
\VS{9}And the
{\ND{
LORD}}
said unto me,
Distress not the
Moabites, neither
contend with them in
battle: for I will not
give thee of their
land
{\ADD{for}} a
possession; because I have
given
Ar unto the
children of
Lot
{\ADD{for}} a
possession.\FTNT{†}{{\FR 2.9 }
Distress…: or, Use no hostility against Moab}
\VS{10}The
Emims
dwelt therein in times
past, a
people
great, and
many, and
tall, as the
Anakims;
\VS{11}Which also were
accounted
giants, as the
Anakims; but the
Moabites
call them
Emims.
\VS{12}The
Horims also
dwelt in
Seir
beforetime; but the
children of
Esau
succeeded them, when they had
destroyed them from
before them, and
dwelt in their stead; as
Israel
did unto the
land of his
possession, which the
{\ND{
LORD}}
gave unto them.\FTNT{‡}{{\FR 2.12 }
succeeded: Heb. inherited}\FTNT{§}{{\FR 2.12 }
stead: or, room}
\VS{13}Now rise
up,
{\ADD{said I}}, and get you
over the
brook
Zered. And we went
over the
brook
Zered.\FTNT{**}{{\FR 2.13 }
brook: or, valley}
\VS{14}And the
space in which we
came from
Kadesh-barnea, until we were come
over the
brook
Zered,
{\ADD{was}}
thirty and
eight
years; until all the
generation of the
men of
war were wasted
out from
among the
host, as the
{\ND{
LORD}}
sware unto them.\FTNT{††}{{\FR 2.14 }
brook: or, valley}
\VS{15}For indeed the
hand of the
{\ND{
LORD}} was against them, to
destroy them from
among the
host, until they were
consumed.
\par }{\PP \VS{16}¶ So it came to pass, when all the
men of
war were
consumed and
dead from
among the
people,
\VS{17}That the
{\ND{
LORD}}
spake unto me,
saying,
\VS{18}Thou art to pass
over through
Ar, the
coast of
Moab, this
day:
\VS{19}And
{\ADD{when}} thou comest
nigh over
against the
children of
Ammon,
distress them not, nor
meddle with them: for I will not
give thee of the
land of the
children of
Ammon
{\ADD{any}}
possession; because I have
given it unto the
children of
Lot
{\ADD{for}} a
possession.
\VS{20}(That also was
accounted a
land of
giants:
giants
dwelt therein in old
time; and the
Ammonites
call them
Zamzummims;\FTNT{‡‡}{{\FR 2.20 }
Zamzummims: also called, Zuzims}
\VS{21}A
people
great, and
many, and
tall, as the
Anakims; but the
{\ND{
LORD}}
destroyed them
before them; and they
succeeded them, and
dwelt in their stead:
\VS{22}As he
did to the
children of
Esau, which
dwelt in
Seir, when he
destroyed the
Horims from
before them; and they
succeeded them, and
dwelt in their stead even unto this
day:
\VS{23}And the
Avims which
dwelt in
Hazerim,
{\ADD{even}} unto
Azzah, the
Caphtorims, which came
forth out of
Caphtor,
destroyed them, and
dwelt in their stead.)
\par }{\PP \VS{24}¶ Rise ye
up, take your
journey, and pass
over the
river
Arnon:
behold, I have
given into thine
hand
Sihon the
Amorite,
king of
Heshbon, and his
land:
begin to
possess
{\ADD{it}}, and
contend with him in
battle.\FTNT{§§}{{\FR 2.24 }
begin…: Heb. begin, possess}
\VS{25}This
day will I
begin to
put the
dread of thee and the
fear of thee upon the
nations
{\ADD{that are}} under the whole
heaven, who shall
hear
report of thee, and shall
tremble, and be in
anguish because of
thee.
\par }{\PP \VS{26}¶ And I
sent
messengers out of the
wilderness of
Kedemoth unto
Sihon
king of
Heshbon with
words of
peace,
saying,
\VS{27}Let me
pass through thy
land: I will
go along by the high
way, I will neither
turn unto the right
hand nor to the
left.
\VS{28}Thou shalt
sell me
meat for
money, that I may
eat; and
give me
water for
money, that I may
drink: only I will pass
through on my
feet;
\VS{29}(As the
children of
Esau which
dwell in
Seir, and the
Moabites which
dwell in
Ar,
did unto me;) until I shall pass
over
Jordan into the
land which the
{\ND{
LORD}} our
God
giveth us.
\VS{30}But
Sihon
king of
Heshbon
would not let us
pass by him: for the
{\ND{
LORD}} thy
God
hardened his
spirit, and made his
heart
obstinate, that he might
deliver him into thy
hand, as
{\ADD{appeareth}} this
day.
\VS{31}And the
{\ND{
LORD}}
said unto me,
Behold, I have
begun to
give
Sihon and his
land
before thee:
begin to
possess, that thou mayest
inherit his
land.
\VS{32}Then
Sihon came
out
against us, he and all his
people, to
fight at
Jahaz.
\VS{33}And the
{\ND{
LORD}} our
God
delivered him
before us; and we
smote him, and his
sons, and all his
people.
\VS{34}And we
took all his
cities at that
time, and utterly
destroyed the
men, and the
women, and the little
ones, of every
city, we
left none to
remain:\FTNT{***}{{\FR 2.34 }
the men…: Heb. every city of men, and women, and little ones}
\VS{35}Only the
cattle we took for a
prey unto ourselves, and the
spoil of the
cities which we
took.
\VS{36}From
Aroer, which
{\ADD{is}} by the
brink of the
river of
Arnon, and
{\ADD{from}} the
city that
{\ADD{is}} by the
river, even unto
Gilead, there was not one
city too
strong for us: the
{\ND{
LORD}} our
God
delivered all unto
us:
\VS{37}Only unto the
land of the
children of
Ammon thou
camest not,
{\ADD{nor}} unto any
place of the
river
Jabbok, nor unto the
cities in the
mountains, nor unto whatsoever the
{\ND{
LORD}} our
God
forbad us.

\par }\Chap{3}{\PP \VerseOne{1}Then we
turned, and went
up the
way to
Bashan: and
Og the
king of
Bashan came
out
against us, he and all his
people, to
battle at
Edrei.
\VS{2}And the
{\ND{
LORD}}
said unto me,
Fear him not: for I will
deliver him, and all his
people, and his
land, into thy
hand; and thou shalt
do unto him as thou
didst unto
Sihon
king of the
Amorites, which
dwelt at
Heshbon.
\VS{3}So the
{\ND{
LORD}} our
God
delivered into our
hands
Og also, the
king of
Bashan, and all his
people: and we
smote him until none was
left to him
remaining.
\VS{4}And we
took all his
cities at that
time, there was not a
city which we
took not from them,
threescore
cities, all the
region of
Argob, the
kingdom of
Og in
Bashan.
\VS{5}All these
cities
{\ADD{were}}
fenced with
high
walls,
gates, and
bars; beside
unwalled
towns a
great
many.
\VS{6}And we utterly
destroyed them, as we
did unto
Sihon
king of
Heshbon, utterly
destroying the
men,
women, and
children, of every
city.
\VS{7}But all the
cattle, and the
spoil of the
cities, we took for a
prey to ourselves.
\VS{8}And we
took at that
time
out of the
hand of the
two
kings of the
Amorites the
land that
{\ADD{was}} on this
side
Jordan, from the
river of
Arnon unto
mount
Hermon;
\VS{9}({\ADD{Which}}
Hermon the
Sidonians
call
Sirion; and the
Amorites
call it
Shenir;)
\VS{10}All the
cities of the
plain, and all
Gilead, and all
Bashan, unto
Salchah and
Edrei,
cities of the
kingdom of
Og in
Bashan.
\VS{11}For only
Og
king of
Bashan
remained of the
remnant of
giants; behold, his
bedstead
{\ADD{was}} a
bedstead of
iron;
{\ADD{is}} it
not in
Rabbath of the
children of
Ammon?
nine
cubits
{\ADD{was}} the
length thereof, and
four
cubits the
breadth of it, after the
cubit of a
man.
\par }{\PP \VS{12}And this
land,
{\ADD{which}} we
possessed at that
time, from
Aroer, which
{\ADD{is}} by the
river
Arnon, and
half
mount
Gilead, and the
cities thereof,
gave I unto the
Reubenites and to the
Gadites.
\VS{13}And the
rest of
Gilead, and all
Bashan,
{\ADD{being}} the
kingdom of
Og,
gave I unto the
half
tribe of
Manasseh; all the
region of
Argob, with all
Bashan,
which was
called the
land of
giants.
\VS{14}Jair the
son of
Manasseh
took all the
country of
Argob unto the
coasts of
Geshuri and
Maachathi; and
called them after his own
name,
Bashan-havoth-jair, unto this
day.
\VS{15}And I
gave
Gilead unto
Machir.
\VS{16}And unto the
Reubenites and unto the
Gadites I
gave from
Gilead even unto the
river
Arnon
half the
valley, and the
border even unto the
river
Jabbok,
{\ADD{which is}} the
border of the
children of
Ammon;
\VS{17}The
plain also, and
Jordan, and the
coast
{\ADD{thereof}}, from
Chinnereth even unto the
sea of the
plain,
{\ADD{even}} the
salt
sea, under
Ashdoth-pisgah
eastward.\FTNT{*}{{\FR 3.17 }
Ashdoth-pisgah: or, the springs of Pisgah, or, the hill}
\par }{\PP \VS{18}¶ And I
commanded you at that
time,
saying, The
{\ND{
LORD}} your
God hath
given you this
land to
possess it: ye shall pass
over
armed
before your
brethren the
children of
Israel, all
{\ADD{that are}}
meet for the
war.\FTNT{†}{{\FR 3.18 }
meet…: Heb. sons of power}
\VS{19}But your
wives, and your little
ones, and your
cattle, ({\ADD{for}} I
know that ye have
much
cattle,) shall
abide in your
cities which I have
given you;
\VS{20}Until the
{\ND{
LORD}} have given
rest unto your
brethren, as well as unto you, and
{\ADD{until}} they also
possess the
land which the
{\ND{
LORD}} your
God hath
given them
beyond
Jordan: and
{\ADD{then}} shall ye
return every
man unto his
possession, which I have
given you.
\par }{\PP \VS{21}¶ And I
commanded
Joshua at that
time,
saying, Thine
eyes have
seen all that the
{\ND{
LORD}} your
God hath
done unto these
two
kings: so shall the
{\ND{
LORD}}
do unto all the
kingdoms whither thou
passest.
\VS{22}Ye shall not
fear them: for the
{\ND{
LORD}} your
God he shall
fight for you.
\VS{23}And I
besought the
{\ND{
LORD}} at that
time,
saying,
\VS{24}O
Lord
{\ND{
GOD}}, thou hast
begun to
shew thy
servant thy
greatness, and thy
mighty
hand: for what
God
{\ADD{is there}} in
heaven or in
earth, that can
do according to thy
works, and according to thy
might?
\VS{25}I pray thee, let me go
over, and
see the
good
land that
{\ADD{is}}
beyond
Jordan, that
goodly
mountain, and
Lebanon.
\VS{26}But the
{\ND{
LORD}} was
wroth with me for your sakes, and would not
hear me: and the
{\ND{
LORD}}
said unto me, Let it
suffice thee;
speak no
more unto me of this
matter.
\VS{27}Get thee
up into the
top of
Pisgah, and lift
up thine
eyes
westward, and
northward, and
southward, and
eastward, and
behold
{\ADD{it}} with thine
eyes: for thou shalt not go
over this
Jordan.\FTNT{‡}{{\FR 3.27 }
Pisgah: or, the hill}
\VS{28}But
charge
Joshua, and
encourage him, and
strengthen him: for he shall go
over
before this
people, and he shall cause them to
inherit the
land which thou shalt
see.
\VS{29}So we
abode in the
valley over
against
Beth-peor.

\par }\Chap{4}{\PP \VerseOne{1}Now therefore
hearken, O
Israel, unto the
statutes and unto the
judgments, which I
teach you, for to
do
{\ADD{them}}, that ye may
live, and go
in and
possess the
land which the
{\ND{
LORD}}
God of your
fathers
giveth you.
\VS{2}Ye shall not
add unto the
word which I
command you, neither shall ye
diminish
{\ADD{ought}} from it, that ye may
keep the
commandments of the
{\ND{
LORD}} your
God which I
command you.
\VS{3}Your
eyes have
seen what the
{\ND{
LORD}}
did because of
Baal-peor: for all the
men that
followed
Baal-peor, the
{\ND{
LORD}} thy
God hath
destroyed them from
among you.
\VS{4}But ye that did
cleave unto the
{\ND{
LORD}} your
God
{\ADD{are}}
alive every one of you this
day.
\VS{5}Behold, I have
taught you
statutes and
judgments, even
as the
{\ND{
LORD}} my
God
commanded me, that ye should
do so
in the
land whither ye
go to
possess it.
\VS{6}Keep therefore and
do
{\ADD{them;}} for this
{\ADD{is}} your
wisdom and your
understanding in the
sight of the
nations, which shall
hear all these
statutes, and
say, Surely this
great
nation
{\ADD{is}} a
wise and
understanding
people.
\VS{7}For what
nation
{\ADD{is there so}}
great, who
{\ADD{hath}}
God
{\ADD{so}}
nigh unto them, as the
{\ND{
LORD}} our
God
{\ADD{is}} in all
{\ADD{things that}} we
call upon him
{\ADD{for}}?
\VS{8}And what
nation
{\ADD{is there so}}
great, that hath
statutes and
judgments
{\ADD{so}}
righteous as all this
law, which I
set
before you this
day?
\VS{9}Only take
heed to thyself, and
keep thy
soul
diligently, lest thou
forget the
things which thine
eyes have
seen, and lest they
depart from thy
heart all the
days of thy
life: but
teach them thy
sons, and thy
sons’
sons;
\VS{10}{\ADD{Specially}} the
day that thou
stoodest
before the
{\ND{
LORD}} thy
God in
Horeb, when the
{\ND{
LORD}}
said unto me,
Gather me the
people
together, and I will make them
hear my
words, that they may
learn to
fear me all the
days that they shall
live upon the
earth, and
{\ADD{that}} they may
teach their
children.
\VS{11}And ye came
near and
stood under the
mountain; and the
mountain
burned with
fire unto the
midst of
heaven, with
darkness,
clouds, and thick
darkness.\FTNT{*}{{\FR 4.11 }
midst: Heb. heart}
\VS{12}And the
{\ND{
LORD}}
spake unto you out of the
midst of the
fire: ye
heard the
voice of the
words, but
saw no
similitude;
only
{\ADD{ye heard}} a
voice.\FTNT{†}{{\FR 4.12 }
only…: Heb. save a voice}
\VS{13}And he
declared unto you his
covenant, which he
commanded you to
perform,
{\ADD{even}}
ten
commandments; and he
wrote them upon
two
tables of
stone.
\par }{\PP \VS{14}¶ And the
{\ND{
LORD}}
commanded me at that
time to
teach you
statutes and
judgments, that ye might
do them in the
land whither ye go
over to
possess it.
\VS{15}Take ye therefore
good
heed unto
yourselves; for ye
saw no manner of
similitude on the
day
{\ADD{that}} the
{\ND{
LORD}}
spake unto you in
Horeb out of the
midst of the
fire:
\VS{16}Lest ye
corrupt
{\ADD{yourselves}}, and
make you a graven
image, the
similitude of any
figure, the
likeness of
male or
female,
\VS{17}The
likeness of any
beast that
{\ADD{is}} on the
earth, the
likeness of any
winged
fowl that
flieth in the
air,
\VS{18}The
likeness of any thing that
creepeth on the
ground, the
likeness of any
fish that
{\ADD{is}} in the
waters
beneath the
earth:
\VS{19}And lest thou lift
up thine
eyes unto
heaven, and when thou
seest the
sun, and the
moon, and the
stars,
{\ADD{even}} all the
host of
heaven, shouldest be
driven to
worship them, and
serve them, which the
{\ND{
LORD}} thy
God hath
divided unto all
nations under the whole
heaven.\FTNT{‡}{{\FR 4.19 }
divided: or, imparted}
\VS{20}But the
{\ND{
LORD}} hath
taken you, and brought you
forth out of the
iron
furnace,
{\ADD{even}} out of
Egypt, to be unto him a
people of
inheritance, as
{\ADD{ye are}} this
day.
\VS{21}Furthermore the
{\ND{
LORD}} was
angry with me for your
sakes, and
sware that I should not go
over
Jordan, and that I should not
go in unto that
good
land, which the
{\ND{
LORD}} thy
God
giveth thee
{\ADD{for}} an
inheritance:
\VS{22}But I must
die in this
land, I must not go
over
Jordan: but ye shall go
over, and
possess that
good
land.
\VS{23}Take
heed unto yourselves, lest ye
forget the
covenant of the
{\ND{
LORD}} your
God, which he
made with you, and
make you a graven
image,
{\ADD{or}} the
likeness of
any
{\ADD{thing}}, which the
{\ND{
LORD}} thy
God hath
forbidden thee.
\VS{24}For the
{\ND{
LORD}} thy
God
{\ADD{is}} a
consuming
fire,
{\ADD{even}} a
jealous
God.
\par }{\PP \VS{25}¶ When thou shalt
beget
children, and
children’s
children, and ye shall have remained
long in the
land, and shall
corrupt
{\ADD{yourselves}}, and
make a graven
image,
{\ADD{or}} the
likeness of any
{\ADD{thing}}, and shall
do
evil in the
sight of the
{\ND{
LORD}} thy
God, to provoke him to
anger:
\VS{26}I
call
heaven and
earth to
witness against you this
day, that ye shall
soon
utterly
perish from off the
land whereunto ye go
over
Jordan to
possess it; ye shall not
prolong
{\ADD{your}}
days upon it, but shall
utterly be
destroyed.
\VS{27}And the
{\ND{
LORD}} shall
scatter you among the
nations, and ye shall be
left
few in
number among the
heathen, whither the
{\ND{
LORD}} shall
lead you.
\VS{28}And there ye shall
serve
gods, the
work of
men’s
hands,
wood and
stone, which neither
see, nor
hear, nor
eat, nor
smell.
\VS{29}But if from thence thou shalt
seek the
{\ND{
LORD}} thy
God, thou shalt
find
{\ADD{him}}, if thou
seek him with all thy
heart and with all thy
soul.
\VS{30}When thou art in
tribulation, and all these
things are
come upon thee,
{\ADD{even}} in the
latter
days, if thou
turn to the
{\ND{
LORD}} thy
God, and shalt be
obedient unto his
voice;\FTNT{§}{{\FR 4.30 }
are…: Heb. have found thee}
\VS{31}(For the
{\ND{
LORD}} thy
God
{\ADD{is}} a
merciful
God;) he will not
forsake thee, neither
destroy thee, nor
forget the
covenant of thy
fathers which he
sware unto them.
\VS{32}For
ask now of the
days that are
past, which
were
before thee, since the
day that
God
created
man upon the
earth, and
{\ADD{ask}} from the one
side of
heaven unto the other, whether there hath been
{\ADD{any such thing}} as this
great
thing
{\ADD{is}}, or hath been
heard like it?
\VS{33}Did
{\ADD{ever}}
people
hear the
voice of
God speaking
out of the
midst of the
fire, as thou hast
heard, and
live?
\VS{34}Or hath
God
assayed to
go
{\ADD{and}}
take him a
nation from the
midst of
{\ADD{another}}
nation, by
temptations, by
signs, and by
wonders, and by
war, and by a
mighty
hand, and by a stretched
out
arm, and by
great
terrors, according to all that the
{\ND{
LORD}} your
God
did for you in
Egypt before your
eyes?
\VS{35}Unto thee it was
shewed, that thou mightest
know that the
{\ND{
LORD}} he
{\ADD{is}}
God;
{\ADD{there is}} none
else beside
him.
\VS{36}Out of
heaven he made thee to
hear his
voice, that he might
instruct thee: and upon
earth he
shewed thee his
great
fire; and thou
heardest his
words out of the
midst of the
fire.
\VS{37}And because he
loved thy
fathers, therefore he
chose their
seed
after them, and brought thee
out in his
sight with his
mighty
power out of
Egypt;
\VS{38}To drive
out
nations from
before thee
greater and
mightier than thou
{\ADD{art}}, to
bring thee in, to
give thee their
land
{\ADD{for}} an
inheritance, as
{\ADD{it is}} this
day.
\VS{39}Know therefore this
day, and
consider
{\ADD{it}} in thine
heart, that the
{\ND{
LORD}} he
{\ADD{is}}
God in
heaven
above, and upon the
earth beneath:
{\ADD{there is}} none else.
\VS{40}Thou shalt
keep therefore his
statutes, and his
commandments, which I
command thee this
day, that it may go
well with thee, and with thy
children
after thee, and that thou mayest
prolong
{\ADD{thy}}
days upon the
earth, which the
{\ND{
LORD}} thy
God
giveth thee, for ever.
\par }{\PP \VS{41}¶ Then
Moses
severed
three
cities on this
side
Jordan toward the
sunrising;
\VS{42}That the
slayer might
flee thither, which should
kill his
neighbour
unawares, and
hated him not in
times
past; and that
fleeing unto
one of
these
cities he might
live:
\VS{43}{\ADD{Namely}},
Bezer in the
wilderness, in the
plain
country, of the
Reubenites; and
Ramoth in
Gilead, of the
Gadites; and
Golan in
Bashan, of the
Manassites.
\par }{\PP \VS{44}¶ And this
{\ADD{is}} the
law which
Moses
set
before the
children of
Israel:
\VS{45}These
{\ADD{are}} the
testimonies, and the
statutes, and the
judgments, which
Moses
spake unto the
children of
Israel, after they came
forth out of
Egypt,
\VS{46}On this
side
Jordan, in the
valley over
against
Beth-peor, in the
land of
Sihon
king of the
Amorites, who
dwelt at
Heshbon, whom
Moses and the
children of
Israel
smote, after they were come
forth out of
Egypt:
\VS{47}And they
possessed his
land, and the
land of
Og
king of
Bashan,
two
kings of the
Amorites, which
{\ADD{were}} on this
side
Jordan toward the
sunrising;
\VS{48}From
Aroer, which
{\ADD{is}} by the
bank of the
river
Arnon, even unto
mount
Sion, which
{\ADD{is}}
Hermon,
\VS{49}And all the
plain on this
side
Jordan
eastward, even unto the
sea of the
plain, under the
springs of
Pisgah.

\par }\Chap{5}{\PP \VerseOne{1}And
Moses
called all
Israel, and
said unto them,
Hear, O
Israel, the
statutes and
judgments which I
speak in your
ears this
day, that ye may
learn them, and
keep, and
do them.\FTNT{*}{{\FR 5.1 }
keep…: Heb. keep to do them}
\VS{2}The
{\ND{
LORD}} our
God
made a
covenant with us in
Horeb.
\VS{3}The
{\ND{
LORD}}
made not this
covenant with our
fathers, but with
us,
{\ADD{even}} us, who
{\ADD{are}} all of
us here
alive this
day.
\VS{4}The
{\ND{
LORD}}
talked with you
face to
face in the
mount out of the
midst of the
fire,
\VS{5}(I
stood between the
{\ND{
LORD}} and you at that
time, to
shew you the
word of the
{\ND{
LORD}}: for ye were
afraid by
reason of the
fire, and went not
up into the
mount;)
saying,
\par }{\PP \VS{6}¶ I
{\ADD{am}} the
{\ND{
LORD}} thy
God, which brought thee
out of the
land of
Egypt, from the
house of
bondage.\FTNT{†}{{\FR 5.6 }
bondage: Heb. servants}
\VS{7}Thou shalt have none
other
gods
before me.
\VS{8}Thou shalt not
make thee
{\ADD{any}} graven
image,
{\ADD{or}} any
likeness
{\ADD{of any thing}} that
{\ADD{is}} in
heaven
above, or that
{\ADD{is}} in the
earth beneath, or that
{\ADD{is}} in the
waters beneath the
earth:
\VS{9}Thou shalt not bow
down thyself unto them, nor
serve them: for I the
{\ND{
LORD}} thy
God
{\ADD{am}} a
jealous
God,
visiting the
iniquity of the
fathers upon the
children unto the
third and
fourth
{\ADD{generation}} of them that
hate me,
\VS{10}And
shewing
mercy unto
thousands of them that
love me and
keep my
commandments.
\VS{11}Thou shalt not
take the
name of the
{\ND{
LORD}} thy
God in
vain: for the
{\ND{
LORD}} will not hold
{\ADD{him}}
guiltless that
taketh his
name in
vain.
\VS{12}Keep the
sabbath
day to
sanctify it, as the
{\ND{
LORD}} thy
God hath
commanded thee.
\VS{13}Six
days thou shalt
labour, and
do all thy
work:
\VS{14}But the
seventh
day
{\ADD{is}} the
sabbath of the
{\ND{
LORD}} thy
God:
{\ADD{in it}} thou shalt not
do any
work, thou, nor thy
son, nor thy
daughter, nor thy
manservant, nor thy
maidservant, nor thine
ox, nor thine
ass, nor any of thy
cattle, nor thy
stranger that
{\ADD{is}} within thy
gates; that thy
manservant and thy
maidservant may
rest as well as thou.
\VS{15}And
remember that thou wast a
servant in the
land of
Egypt, and
{\ADD{that}} the
{\ND{
LORD}} thy
God
brought thee out thence through a
mighty
hand and by a stretched
out
arm: therefore the
{\ND{
LORD}} thy
God
commanded thee to
keep the
sabbath
day.
\par }{\PP \VS{16}¶
Honour thy
father and thy
mother, as the
{\ND{
LORD}} thy
God hath
commanded thee; that thy
days may be
prolonged, and that it may go
well with thee, in the
land which the
{\ND{
LORD}} thy
God
giveth thee.
\VS{17}Thou shalt not
kill.
\VS{18}Neither shalt thou commit
adultery.
\VS{19}Neither shalt thou
steal.
\VS{20}Neither shalt thou
bear
false
witness against thy
neighbour.
\VS{21}Neither shalt thou
desire thy
neighbour’s
wife, neither shalt thou
covet thy
neighbour’s
house, his
field, or his
manservant, or his
maidservant, his
ox, or his
ass, or any
{\ADD{thing}} that
{\ADD{is}} thy
neighbour’s.
\par }{\PP \VS{22}¶ These
words the
{\ND{
LORD}}
spake unto all your
assembly in the
mount out of the
midst of the
fire, of the
cloud, and of the thick
darkness, with a
great
voice: and he added no
more. And he
wrote them in
two
tables of
stone, and
delivered them unto me.
\par }{\PP \VS{23}And it came to pass, when ye
heard the
voice out of the
midst of the
darkness, (for the
mountain did
burn with
fire,) that ye came
near unto me,
{\ADD{even}} all the
heads of your
tribes, and your
elders;
\VS{24}And ye
said, Behold, the
{\ND{
LORD}} our
God hath
shewed us his
glory and his
greatness, and we have
heard his
voice out of the
midst of the
fire: we have
seen this
day that
God doth
talk with
man, and he
liveth.
\VS{25}Now therefore why should we
die? for this
great
fire will
consume us: if we
hear the
voice of the
{\ND{
LORD}} our
God any
more, then we shall
die.\FTNT{‡}{{\FR 5.25 }
hear: Heb. add to hear}
\VS{26}For who
{\ADD{is there of}} all
flesh, that hath
heard the
voice of the
living
God
speaking out of the
midst of the
fire, as we
{\ADD{have}}, and
lived?
\VS{27}Go thou
near, and
hear all that the
{\ND{
LORD}} our
God shall
say: and
speak thou unto us all that the
{\ND{
LORD}} our
God shall
speak unto thee; and we will
hear
{\ADD{it}}, and
do
{\ADD{it}}.
\VS{28}And the
{\ND{
LORD}}
heard the
voice of your
words, when ye
spake unto me; and the
{\ND{
LORD}}
said unto me, I have
heard the
voice of the
words of this
people, which they have
spoken unto thee: they have
well said all that they have
spoken.
\VS{29}O that there were
such
an
heart in
them, that they would
fear me, and
keep all my
commandments
always, that it might be
well with them, and with their
children for
ever!
\VS{30}Go
say to them,
Get you into your
tents
again.
\VS{31}But as for thee,
stand thou here by
me, and I will
speak unto thee all the
commandments, and the
statutes, and the
judgments, which thou shalt
teach them, that they may
do
{\ADD{them}} in the
land which I
give them to
possess it.
\VS{32}Ye shall
observe to
do therefore as the
{\ND{
LORD}} your
God hath
commanded you: ye shall not turn
aside to the right
hand or to the
left.
\VS{33}Ye shall
walk in all the
ways which the
{\ND{
LORD}} your
God hath
commanded you, that ye may
live, and
{\ADD{that it may be}}
well with you, and
{\ADD{that}} ye may
prolong
{\ADD{your}}
days in the
land which ye shall
possess.

\par }\Chap{6}{\PP \VerseOne{1}Now these
{\ADD{are}} the
commandments, the
statutes, and the
judgments, which the
{\ND{
LORD}} your
God
commanded to
teach you, that ye might
do
{\ADD{them}} in the
land whither ye
go to
possess it:\FTNT{*}{{\FR 6.1 }
go: Heb. pass over}
\VS{2}That thou mightest
fear the
{\ND{
LORD}} thy
God, to
keep all his
statutes and his
commandments, which I
command thee, thou, and thy
son, and thy
son’s
son, all the
days of thy
life; and that thy
days may be
prolonged.
\par }{\PP \VS{3}¶
Hear therefore, O
Israel, and
observe to
do
{\ADD{it;}} that it may be
well with thee, and that ye may
increase
mightily, as the
{\ND{
LORD}}
God of thy
fathers hath
promised thee, in the
land that
floweth with
milk and
honey.
\par }{\PP \VS{4}Hear, O
Israel: The
{\ND{
LORD}} our
God
{\ADD{is}}
one
{\ND{
LORD}}:
\VS{5}And thou shalt
love the
{\ND{
LORD}} thy
God with all thine
heart, and with all thy
soul, and with all thy
might.
\VS{6}And these
words, which I
command thee this
day, shall be in thine
heart:
\VS{7}And thou shalt teach them
diligently unto thy
children, and shalt
talk of them when thou
sittest in thine
house, and when thou
walkest by the
way, and when thou liest
down, and when thou risest
up.\FTNT{†}{{\FR 6.7 }
teach: Heb. whet, or, sharpen}
\VS{8}And thou shalt
bind them for a
sign upon thine
hand, and they shall be as
frontlets between thine
eyes.
\VS{9}And thou shalt
write them upon the
posts of thy
house, and on thy
gates.
\VS{10}And it shall be, when the
{\ND{
LORD}} thy
God shall have
brought thee into the
land which he
sware unto thy
fathers, to
Abraham, to
Isaac, and to
Jacob, to
give thee
great and
goodly
cities, which thou
buildedst not,
\VS{11}And
houses
full of all
good
{\ADD{things}}, which thou
filledst not, and
wells
digged, which thou
diggedst not,
vineyards and olive
trees, which thou
plantedst not; when thou shalt have
eaten and be
full;
\VS{12}{\ADD{Then}}
beware lest thou
forget the
{\ND{
LORD}}, which brought thee
forth out of the
land of
Egypt, from the
house of
bondage.\FTNT{‡}{{\FR 6.12 }
bondage: Heb. bondmen or, servants}
\VS{13}Thou shalt
fear the
{\ND{
LORD}} thy
God, and
serve him, and shalt
swear by his
name.
\VS{14}Ye shall not
go
after
other
gods, of the
gods of the
people which
{\ADD{are}} round
about you;
\VS{15}(For the
{\ND{
LORD}} thy
God
{\ADD{is}} a
jealous
God
among you) lest the
anger of the
{\ND{
LORD}} thy
God be
kindled against thee, and
destroy thee from off the
face of the
earth.
\par }{\PP \VS{16}¶ Ye shall not
tempt the
{\ND{
LORD}} your
God, as ye
tempted
{\ADD{him}} in
Massah.
\par }{\PP \VS{17}Ye shall
diligently
keep the
commandments of the
{\ND{
LORD}} your
God, and his
testimonies, and his
statutes, which he hath
commanded thee.
\VS{18}And thou shalt
do
{\ADD{that which is}}
right and
good in the
sight of the
{\ND{
LORD}}: that it may be
well with thee, and that thou mayest go
in and
possess the
good
land which the
{\ND{
LORD}}
sware unto thy
fathers,
\VS{19}To cast
out all thine
enemies from
before thee, as the
{\ND{
LORD}} hath
spoken.
\VS{20}{\ADD{And}} when thy
son
asketh thee in time to
come,
saying, What
{\ADD{mean}} the
testimonies, and the
statutes, and the
judgments, which the
{\ND{
LORD}} our
God hath
commanded you?\FTNT{§}{{\FR 6.20 }
in…: Heb. to morrow}
\VS{21}Then thou shalt
say unto thy
son, We were
Pharaoh’s
bondmen in
Egypt; and the
{\ND{
LORD}} brought us
out of
Egypt with a
mighty
hand:
\VS{22}And the
{\ND{
LORD}}
shewed
signs and
wonders,
great and
sore, upon
Egypt, upon
Pharaoh, and upon all his
household, before our
eyes:\FTNT{**}{{\FR 6.22 }
sore: Heb. evil}
\VS{23}And he brought us
out from thence, that he might bring us
in, to
give us the
land which he
sware unto our
fathers.
\VS{24}And the
{\ND{
LORD}}
commanded us to
do all these
statutes, to
fear the
{\ND{
LORD}} our
God, for our
good
always, that he might preserve us
alive, as
{\ADD{it is}} at this
day.
\VS{25}And it shall be our
righteousness, if we
observe to
do all
these
commandments
before the
{\ND{
LORD}} our
God, as he hath
commanded us.

\par }\Chap{7}{\PP \VerseOne{1}When the
{\ND{
LORD}} thy
God shall
bring thee into the
land whither thou
goest to
possess it, and hath cast
out
many
nations
before thee, the
Hittites, and the
Girgashites, and the
Amorites, and the
Canaanites, and the
Perizzites, and the
Hivites, and the
Jebusites,
seven
nations
greater and
mightier than thou;
\VS{2}And when the
{\ND{
LORD}} thy
God shall
deliver them
before thee; thou shalt
smite them,
{\ADD{and}}
utterly
destroy them; thou shalt
make no
covenant with them, nor shew
mercy unto them:
\VS{3}Neither shalt thou make
marriages with them; thy
daughter thou shalt not
give unto his
son, nor his
daughter shalt thou
take unto thy
son.
\VS{4}For they will turn
away thy
son from
following me, that they may
serve
other
gods: so will the
anger of the
{\ND{
LORD}} be
kindled against you, and
destroy thee
suddenly.
\VS{5}But thus shall ye
deal with them; ye shall
destroy their
altars, and break
down their
images, and cut
down their
groves, and
burn their graven
images with
fire.\FTNT{*}{{\FR 7.5 }
their images: Heb. their statues, or, pillars}
\VS{6}For thou
{\ADD{art}} an
holy
people unto the
{\ND{
LORD}} thy
God: the
{\ND{
LORD}} thy
God hath
chosen thee to be a
special
people unto himself, above all
people that
{\ADD{are}} upon the
face of the
earth.
\VS{7}The
{\ND{
LORD}} did not set his
love upon you, nor
choose you, because ye were
more in number than any
people; for ye
{\ADD{were}} the
fewest of all
people:
\VS{8}But because the
{\ND{
LORD}}
loved you, and because he would
keep the
oath which he had
sworn unto your
fathers, hath the
{\ND{
LORD}} brought you
out with a
mighty
hand, and
redeemed you out of the
house of
bondmen, from the
hand of
Pharaoh
king of
Egypt.
\VS{9}Know therefore that the
{\ND{
LORD}} thy
God, he
{\ADD{is}}
God, the
faithful
God, which
keepeth
covenant and
mercy with them that
love him and
keep his
commandments to a
thousand
generations;
\VS{10}And
repayeth them that
hate him to their
face, to
destroy them: he will not be
slack to him that
hateth him, he will
repay him to his
face.
\VS{11}Thou shalt therefore
keep the
commandments, and the
statutes, and the
judgments, which I
command thee this
day, to
do them.
\par }{\PP \VS{12}¶ Wherefore it shall come to pass,
if ye
hearken to these
judgments, and
keep, and
do them, that the
{\ND{
LORD}} thy
God shall
keep unto thee the
covenant and the
mercy which he
sware unto thy
fathers:\FTNT{†}{{\FR 7.12 }
if: Heb. because}
\VS{13}And he will
love thee, and
bless thee, and
multiply thee: he will also
bless the
fruit of thy
womb, and the
fruit of thy
land, thy
corn, and thy
wine, and thine
oil, the
increase of thy
kine, and the
flocks of thy
sheep, in the
land which he
sware unto thy
fathers to
give thee.
\VS{14}Thou shalt be
blessed above all
people: there shall not be
male or female
barren among you, or among your
cattle.
\VS{15}And the
{\ND{
LORD}} will take
away from thee all
sickness, and will
put none of the
evil
diseases of
Egypt, which thou
knowest, upon thee; but will
lay them upon all
{\ADD{them}} that
hate thee.
\VS{16}And thou shalt
consume all the
people which the
{\ND{
LORD}} thy
God shall
deliver thee; thine
eye shall have no
pity upon them: neither shalt thou
serve their
gods; for that
{\ADD{will be}} a
snare unto thee.
\VS{17}If thou shalt
say in thine
heart, These
nations
{\ADD{are}}
more than I;
how
can I
dispossess them?
\VS{18}Thou shalt not be
afraid of them:
{\ADD{but}} shalt
well
remember what the
{\ND{
LORD}} thy
God
did unto
Pharaoh, and unto all
Egypt;
\VS{19}The
great
temptations which thine
eyes
saw, and the
signs, and the
wonders, and the
mighty
hand, and the stretched
out
arm, whereby the
{\ND{
LORD}} thy
God brought thee
out: so shall the
{\ND{
LORD}} thy
God
do unto all the
people of whom thou art
afraid.
\VS{20}Moreover the
{\ND{
LORD}} thy
God will
send the
hornet among them, until they that are
left, and
hide themselves
from thee, be
destroyed.
\VS{21}Thou shalt not be
affrighted
at them: for the
{\ND{
LORD}} thy
God
{\ADD{is}}
among you, a
mighty
God and
terrible.
\VS{22}And the
{\ND{
LORD}} thy
God will put
out
those
nations
before thee by
little and
little: thou
mayest not
consume them at
once, lest the
beasts of the
field
increase upon thee.\FTNT{‡}{{\FR 7.22 }
put…: Heb. pluck off}
\VS{23}But the
{\ND{
LORD}} thy
God shall
deliver them unto
thee, and shall
destroy them with a
mighty
destruction, until they be
destroyed.\FTNT{§}{{\FR 7.23 }
unto…: Heb. before thy face}
\VS{24}And he shall
deliver their
kings into thine
hand, and thou shalt
destroy their
name from under
heaven: there shall no
man be able to
stand
before thee, until thou have
destroyed them.
\VS{25}The graven
images of their
gods shall ye
burn with
fire: thou shalt not
desire the
silver or
gold
{\ADD{that is}} on them, nor
take
{\ADD{it}} unto thee, lest thou be
snared therein: for it
{\ADD{is}} an
abomination to the
{\ND{
LORD}} thy
God.
\VS{26}Neither shalt thou
bring an
abomination into thine
house, lest thou be a cursed
thing like it:
{\ADD{but}} thou shalt
utterly
detest it, and thou shalt
utterly
abhor it; for it
{\ADD{is}} a cursed
thing.

\par }\Chap{8}{\PP \VerseOne{1}All the
commandments which I
command thee this
day shall ye
observe to
do, that ye may
live, and
multiply, and go
in and
possess the
land which the
{\ND{
LORD}}
sware unto your
fathers.
\VS{2}And thou shalt
remember all the
way which the
{\ND{
LORD}} thy
God
led thee these
forty
years in the
wilderness, to
humble thee,
{\ADD{and}} to
prove thee, to
know what
{\ADD{was}} in thine
heart, whether thou wouldest
keep his
commandments, or no.
\VS{3}And he
humbled thee, and suffered thee to
hunger, and
fed thee with
manna, which thou
knewest not, neither did thy
fathers
know; that he might make thee
know that
man doth not
live by
bread
only, but by every
{\ADD{word}} that
proceedeth out of the
mouth of the
{\ND{
LORD}} doth
man
live.
\VS{4}Thy
raiment waxed not
old upon thee, neither did thy
foot
swell, these
forty
years.
\VS{5}Thou shalt also
consider
in thine
heart, that, as a
man
chasteneth his
son,
{\ADD{so}} the
{\ND{
LORD}} thy
God
chasteneth thee.
\VS{6}Therefore thou shalt
keep the
commandments of the
{\ND{
LORD}} thy
God, to
walk in his
ways, and to
fear him.
\VS{7}For the
{\ND{
LORD}} thy
God
bringeth thee into a
good
land, a
land of
brooks of
water, of
fountains and
depths that spring
out of
valleys and
hills;
\VS{8}A
land of
wheat, and
barley, and
vines, and fig
trees, and
pomegranates; a
land of
oil
olive, and
honey;\FTNT{*}{{\FR 8.8 }
of oil…: Heb. of olive tree of oil}
\VS{9}A
land
wherein thou shalt
eat
bread
without
scarceness, thou shalt not
lack any
{\ADD{thing}} in it; a
land whose
stones
{\ADD{are}}
iron, and out of whose
hills thou mayest
dig
brass.
\par }{\PP \VS{10}When thou hast
eaten and art
full, then thou shalt
bless the
{\ND{
LORD}} thy
God for the
good
land which he hath
given thee.
\VS{11}Beware that thou
forget not the
{\ND{
LORD}} thy
God, in not
keeping his
commandments, and his
judgments, and his
statutes, which I
command thee this
day:
\VS{12}Lest
{\ADD{when}} thou hast
eaten and art
full, and hast
built
goodly
houses, and
dwelt
{\ADD{therein;}}
\VS{13}And
{\ADD{when}} thy
herds and thy
flocks
multiply, and thy
silver and thy
gold is
multiplied, and all that thou hast is
multiplied;
\VS{14}Then thine
heart be lifted
up, and thou
forget the
{\ND{
LORD}} thy
God, which brought thee
forth out of the
land of
Egypt, from the
house of
bondage;
\VS{15}Who
led thee through that
great and
terrible
wilderness,
{\ADD{wherein were}}
fiery
serpents, and
scorpions, and
drought, where
{\ADD{there was}} no
water; who brought thee
forth
water out of the
rock of
flint;
\VS{16}Who
fed thee in the
wilderness with
manna, which thy
fathers
knew not, that he might
humble thee, and that he might
prove thee, to do thee
good at thy latter
end;
\VS{17}And thou
say in thine
heart, My
power and the
might of
{\ADD{mine}}
hand hath
gotten me this
wealth.
\VS{18}But thou shalt
remember the
{\ND{
LORD}} thy
God: for
{\ADD{it is}} he that
giveth thee
power to
get
wealth, that he may
establish his
covenant which he
sware unto thy
fathers, as
{\ADD{it is}} this
day.
\VS{19}And it shall be, if thou do at
all
forget the
{\ND{
LORD}} thy
God, and
walk
after
other
gods, and
serve them, and
worship them, I
testify against you this
day that ye shall
surely
perish.
\VS{20}As the
nations which the
{\ND{
LORD}}
destroyeth before your
face, so shall ye
perish;
because ye would not be
obedient unto the
voice of the
{\ND{
LORD}} your
God.

\par }\Chap{9}{\PP \VerseOne{1}Hear, O
Israel: Thou
{\ADD{art}} to pass
over
Jordan this
day, to go
in to
possess
nations
greater and
mightier than thyself,
cities
great and fenced
up to
heaven,
\VS{2}A
people
great and
tall, the
children of the
Anakims, whom thou
knowest, and
{\ADD{of whom}} thou hast
heard
{\ADD{say}}, Who can
stand
before the
children of
Anak!
\VS{3}Understand therefore this
day, that the
{\ND{
LORD}} thy
God
{\ADD{is}} he which goeth
over
before thee;
{\ADD{as}} a
consuming
fire he shall
destroy them, and he shall bring them
down before thy
face: so shalt thou drive them
out, and
destroy them
quickly, as the
{\ND{
LORD}} hath
said unto thee.
\VS{4}Speak not thou in thine
heart, after that the
{\ND{
LORD}} thy
God hath cast them
out from
before thee,
saying, For my
righteousness the
{\ND{
LORD}} hath brought me
in to
possess this
land: but for the
wickedness of these
nations the
{\ND{
LORD}} doth drive them
out from
before thee.
\VS{5}Not for thy
righteousness, or for the
uprightness of thine
heart, dost thou
go to
possess their
land: but for the
wickedness of these
nations the
{\ND{
LORD}} thy
God doth drive them
out from
before thee, and that he may
perform the
word which the
{\ND{
LORD}}
sware unto thy
fathers,
Abraham,
Isaac, and
Jacob.
\VS{6}Understand therefore, that the
{\ND{
LORD}} thy
God
giveth thee not this
good
land to
possess it for thy
righteousness; for thou
{\ADD{art}} a
stiffnecked
people.
\par }{\PP \VS{7}¶
Remember,
{\ADD{and}}
forget not, how thou provokedst the
{\ND{
LORD}} thy
God to
wrath in the
wilderness:
from the
day that thou didst depart
out of the
land of
Egypt, until ye
came unto this
place, ye have been
rebellious
against the
{\ND{
LORD}}.
\VS{8}Also in
Horeb ye
provoked the
{\ND{
LORD}} to
wrath, so that the
{\ND{
LORD}} was
angry with you to have
destroyed you.
\VS{9}When I was gone
up into the
mount to
receive the
tables of
stone,
{\ADD{even}} the
tables of the
covenant which the
{\ND{
LORD}}
made with you, then I
abode in the
mount
forty
days and
forty
nights, I neither did
eat
bread nor
drink
water:
\VS{10}And the
{\ND{
LORD}}
delivered unto me
two
tables of
stone
written with the
finger of
God; and on them
{\ADD{was written}} according to all the
words, which the
{\ND{
LORD}}
spake with you in the
mount out of the
midst of the
fire in the
day of the
assembly.
\VS{11}And it came to pass at the
end of
forty
days and
forty
nights,
{\ADD{that}} the
{\ND{
LORD}}
gave me the
two
tables of
stone,
{\ADD{even}} the
tables of the
covenant.
\VS{12}And the
{\ND{
LORD}}
said unto me,
Arise, get thee
down
quickly from hence; for thy
people which thou hast brought
forth out of
Egypt have
corrupted
{\ADD{themselves;}} they are
quickly turned
aside out of the
way which I
commanded them; they have
made them a molten
image.
\VS{13}Furthermore the
{\ND{
LORD}}
spake unto me,
saying, I have
seen this
people, and, behold, it
{\ADD{is}} a
stiffnecked
people:
\VS{14}Let me
alone, that I may
destroy them, and blot
out their
name from under
heaven: and I will
make
of thee a
nation
mightier and
greater than they.
\VS{15}So I
turned and came
down from the
mount, and the
mount
burned with
fire: and the
two
tables of the
covenant
{\ADD{were}} in my
two
hands.
\VS{16}And I
looked, and, behold, ye had
sinned against the
{\ND{
LORD}} your
God,
{\ADD{and}} had
made you a
molten
calf: ye had turned
aside
quickly out of the
way which the
{\ND{
LORD}} had
commanded you.
\VS{17}And I
took the
two
tables, and
cast them out
of my
two
hands, and
brake them before your
eyes.
\VS{18}And I fell
down
before the
{\ND{
LORD}}, as at the
first,
forty
days and
forty
nights: I did neither
eat
bread, nor
drink
water, because of all your
sins which ye
sinned, in
doing
wickedly in the
sight of the
{\ND{
LORD}}, to provoke him to
anger.
\VS{19}For I was
afraid
of the
anger and hot
displeasure, wherewith the
{\ND{
LORD}} was
wroth against you to
destroy you. But the
{\ND{
LORD}}
hearkened unto me at that
time also.
\VS{20}And the
{\ND{
LORD}} was
very
angry with
Aaron to have
destroyed him: and I
prayed for
Aaron also the same
time.
\VS{21}And I
took your
sin, the
calf which ye had
made, and
burnt it with
fire, and
stamped it,
{\ADD{and}}
ground
{\ADD{it}} very
small,
{\ADD{even}} until it was as
small as
dust: and I
cast the
dust thereof into the
brook that descended
out of the
mount.
\VS{22}And at
Taberah, and at
Massah, and at
Kibroth-hattaavah, ye provoked the
{\ND{
LORD}} to
wrath.
\VS{23}Likewise when the
{\ND{
LORD}}
sent you from
Kadesh-barnea,
saying, Go
up and
possess the
land which I have
given you; then ye
rebelled against the
commandment of the
{\ND{
LORD}} your
God, and ye
believed him not, nor
hearkened to his
voice.
\VS{24}Ye have been
rebellious against the
{\ND{
LORD}} from the
day that I
knew you.
\VS{25}Thus I fell
down
before the
{\ND{
LORD}}
forty
days and
forty
nights, as I fell
down
{\ADD{at the first;}} because the
{\ND{
LORD}} had
said he would
destroy you.
\VS{26}I
prayed therefore unto the
{\ND{
LORD}}, and
said, O
Lord
{\ND{
GOD}},
destroy not thy
people and thine
inheritance, which thou hast
redeemed through thy
greatness, which thou hast brought
forth out of
Egypt with a
mighty
hand.
\VS{27}Remember thy
servants,
Abraham,
Isaac, and
Jacob;
look not unto the
stubbornness of this
people, nor to their
wickedness, nor to their
sin:
\VS{28}Lest the
land whence thou broughtest us
out
say, Because the
{\ND{
LORD}} was not
able to
bring them into the
land which he
promised them, and because he
hated them, he hath brought them
out to
slay them in the
wilderness.
\VS{29}Yet they
{\ADD{are}} thy
people and thine
inheritance, which thou broughtest
out by thy
mighty
power and by thy stretched
out
arm.

\par }\Chap{10}{\PP \VerseOne{1}At that
time the
{\ND{
LORD}}
said unto me,
Hew thee
two
tables of
stone like unto the
first, and come
up unto me into the
mount, and
make thee an
ark of
wood.
\VS{2}And I will
write on the
tables the
words that were in the
first
tables which thou
brakest, and thou shalt
put them in the
ark.
\VS{3}And I
made an
ark
{\ADD{of}}
shittim
wood, and
hewed
two
tables of
stone like unto the
first, and went
up into the
mount, having the
two
tables in mine
hand.
\VS{4}And he
wrote on the
tables, according to the
first
writing, the
ten
commandments, which the
{\ND{
LORD}}
spake unto you in the
mount out of the
midst of the
fire in the
day of the
assembly: and the
{\ND{
LORD}}
gave them unto me.\FTNT{*}{{\FR 10.4 }
commandments: Heb. words}
\VS{5}And I
turned myself and came
down from the
mount, and
put the
tables in the
ark which I had
made; and there they be, as the
{\ND{
LORD}}
commanded me.
\par }{\PP \VS{6}¶ And the
children of
Israel took their
journey from Beeroth of the children of
Jaakan to
Mosera: there
Aaron
died, and there he was
buried; and
Eleazar his
son ministered in the
priest’s office in his stead.
\VS{7}From thence they
journeyed unto
Gudgodah; and from
Gudgodah to
Jotbath, a
land of
rivers of
waters.
\par }{\PP \VS{8}¶ At that
time the
{\ND{
LORD}}
separated the
tribe of
Levi, to
bear the
ark of the
covenant of the
{\ND{
LORD}}, to
stand
before the
{\ND{
LORD}} to
minister unto him, and to
bless in his
name, unto this
day.
\VS{9}Wherefore
Levi hath no
part nor
inheritance with his
brethren; the
{\ND{
LORD}}
{\ADD{is}} his
inheritance, according as the
{\ND{
LORD}} thy
God
promised him.
\VS{10}And I
stayed in the
mount, according to the
first
time,
forty
days and
forty
nights; and the
{\ND{
LORD}}
hearkened unto me at that
time also,
{\ADD{and}} the
{\ND{
LORD}}
would not
destroy thee.\FTNT{†}{{\FR 10.10 }
first…: or, former days}
\VS{11}And the
{\ND{
LORD}}
said unto me,
Arise,
take
{\ADD{thy}}
journey
before the
people, that they may go
in and
possess the
land, which I
sware unto their
fathers to
give unto them.\FTNT{‡}{{\FR 10.11 }
take…: Heb. go in journey}
\par }{\PP \VS{12}¶ And now,
Israel, what doth the
{\ND{
LORD}} thy
God
require of thee, but to
fear the
{\ND{
LORD}} thy
God, to
walk in all his
ways, and to
love him, and to
serve the
{\ND{
LORD}} thy
God with all thy
heart and with all thy
soul,
\VS{13}To
keep the
commandments of the
{\ND{
LORD}}, and his
statutes, which I
command thee this
day for thy
good?
\VS{14}Behold, the
heaven and the
heaven of
heavens
{\ADD{is}} the
{\ND{
LORD’s}} thy
God, the
earth
{\ADD{also}}, with all that therein
{\ADD{is}}.
\VS{15}Only the
{\ND{
LORD}} had a
delight in thy
fathers to
love them, and he
chose their
seed
after them,
{\ADD{even}} you above all
people, as
{\ADD{it is}} this
day.
\VS{16}Circumcise therefore the
foreskin of your
heart, and be no more
stiffnecked.
\VS{17}For the
{\ND{
LORD}} your
God
{\ADD{is}}
God of
gods, and
Lord of
lords, a
great
God, a
mighty, and a
terrible, which
regardeth not
persons, nor
taketh
reward:
\VS{18}He doth
execute the
judgment of the
fatherless and
widow, and
loveth the
stranger, in
giving him
food and
raiment.
\VS{19}Love ye therefore the
stranger: for ye were
strangers in the
land of
Egypt.
\VS{20}Thou shalt
fear the
{\ND{
LORD}} thy
God; him shalt thou
serve, and to him shalt thou
cleave, and
swear by his
name.
\VS{21}He
{\ADD{is}} thy
praise, and he
{\ADD{is}} thy
God, that hath
done for thee these
great and terrible
things, which thine
eyes have
seen.
\VS{22}Thy
fathers went
down into
Egypt with threescore and
ten
persons; and now the
{\ND{
LORD}} thy
God hath
made thee as the
stars of
heaven for
multitude.

\par }\Chap{11}{\PP \VerseOne{1}Therefore thou shalt
love the
{\ND{
LORD}} thy
God, and
keep his
charge, and his
statutes, and his
judgments, and his
commandments,
alway.
\VS{2}And
know ye this
day: for
{\ADD{I speak}} not with your
children which have not
known, and which have not
seen the
chastisement of the
{\ND{
LORD}} your
God, his
greatness, his
mighty
hand, and his stretched
out
arm,
\VS{3}And his
miracles, and his
acts, which he
did in the
midst of
Egypt unto
Pharaoh the
king of
Egypt, and unto all his
land;
\VS{4}And what he
did unto the
army of
Egypt, unto their
horses, and to their
chariots; how he made the
water of the
Red
sea to
overflow
them as they
pursued
after you, and
{\ADD{how}} the
{\ND{
LORD}} hath
destroyed them unto this
day;
\VS{5}And what he
did unto you in the
wilderness, until ye
came into this
place;
\VS{6}And what he
did unto
Dathan and
Abiram, the
sons of
Eliab, the
son of
Reuben: how the
earth
opened her
mouth, and swallowed them
up, and their
households, and their
tents, and all the
substance that
{\ADD{was}} in their
possession, in the
midst of all
Israel:\FTNT{*}{{\FR 11.6 }
substance…: or, living substance which followed them}\FTNT{†}{{\FR 11.6 }
was…: Heb. was at their feet}
\VS{7}But your
eyes have
seen all the
great
acts of the
{\ND{
LORD}} which he
did.
\par }{\PP \VS{8}Therefore shall ye
keep all the
commandments
which I
command you this
day, that ye may be
strong, and go
in and
possess the
land, whither ye
go to
possess it;
\VS{9}And that ye may
prolong
{\ADD{your}}
days in the
land, which the
{\ND{
LORD}}
sware unto your
fathers to
give unto them and to their
seed, a
land that
floweth with
milk and
honey.
\par }{\PP \VS{10}¶ For the
land, whither thou goest
in to
possess it,
{\ADD{is}} not as the
land of
Egypt, from whence ye came
out, where thou
sowedst thy
seed, and
wateredst
{\ADD{it}} with thy
foot, as a
garden of
herbs:
\VS{11}But the
land, whither ye
go to
possess it,
{\ADD{is}} a
land of
hills and
valleys,
{\ADD{and}}
drinketh
water of the
rain of
heaven:
\VS{12}A
land which the
{\ND{
LORD}} thy
God careth
for: the
eyes of the
{\ND{
LORD}} thy
God
{\ADD{are}}
always upon it, from the
beginning of the
year even unto the
end of the
year.\FTNT{‡}{{\FR 11.12 }
careth…: Heb. seeketh}
\par }{\PP \VS{13}¶ And it shall come to pass, if ye shall
hearken
diligently unto my
commandments which I
command you this
day, to
love the
{\ND{
LORD}} your
God, and to
serve him with all your
heart and with all your
soul,
\VS{14}That I will
give
{\ADD{you}} the
rain of your
land in his due
season, the first
rain and the latter
rain, that thou mayest
gather in thy
corn, and thy
wine, and thine
oil.
\VS{15}And I will
send
grass in thy
fields for thy
cattle, that thou mayest
eat and be
full.\FTNT{§}{{\FR 11.15 }
send: Heb. give}
\VS{16}Take
heed to yourselves, that your
heart be not
deceived, and ye turn
aside, and
serve
other
gods, and
worship them;
\VS{17}And
{\ADD{then}} the
{\ND{
LORD’s}}
wrath be
kindled against you, and he shut
up the
heaven, that there be no
rain, and that the
land
yield not her
fruit; and
{\ADD{lest}} ye
perish
quickly from off the
good
land which the
{\ND{
LORD}}
giveth you.
\par }{\PP \VS{18}¶ Therefore shall ye lay
up these my
words in your
heart and in your
soul, and
bind them for a
sign upon your
hand, that they may be as
frontlets between your
eyes.
\VS{19}And ye shall
teach them your
children,
speaking of them when thou
sittest in thine
house, and when thou
walkest by the
way, when thou
liest down, and when thou risest
up.
\VS{20}And thou shalt
write them upon the door
posts of thine
house, and upon thy
gates:
\VS{21}That your
days may be
multiplied, and the
days of your
children, in the
land which the
{\ND{
LORD}}
sware unto your
fathers to
give them, as the
days of
heaven upon the
earth.
\par }{\PP \VS{22}¶ For if ye shall
diligently
keep all these
commandments which I
command you, to
do them, to
love the
{\ND{
LORD}} your
God, to
walk in all his
ways, and to
cleave unto him;
\VS{23}Then will the
{\ND{
LORD}} drive
out all these
nations from
before you, and ye shall
possess
greater
nations and
mightier than yourselves.
\VS{24}Every
place whereon the
soles of your
feet shall
tread shall be yours: from the
wilderness and
Lebanon, from the
river, the
river
Euphrates, even unto the
uttermost
sea shall your
coast be.
\VS{25}There shall no
man be able to
stand
before you:
{\ADD{for}} the
{\ND{
LORD}} your
God shall
lay the
fear of you and the
dread of you
upon all the
land that ye shall
tread upon, as he hath
said unto you.
\par }{\PP \VS{26}¶
Behold, I
set
before you this
day a
blessing and a
curse;
\VS{27}A
blessing, if ye
obey the
commandments of the
{\ND{
LORD}} your
God, which I
command you this
day:
\VS{28}And a
curse, if ye will not
obey the
commandments of the
{\ND{
LORD}} your
God, but turn
aside out of the
way which I
command you this
day, to
go
after
other
gods, which ye have not
known.
\VS{29}And it shall come to pass, when the
{\ND{
LORD}} thy
God hath brought thee
in unto the
land whither thou
goest to
possess it, that thou shalt
put the
blessing upon
mount
Gerizim, and the
curse upon
mount
Ebal.
\VS{30}{\ADD{Are}} they not on the other
side
Jordan,
by the
way where the
sun goeth
down, in the
land of the
Canaanites, which
dwell in the
champaign over
against
Gilgal,
beside the
plains of
Moreh?
\VS{31}For ye shall pass
over
Jordan to go
in to
possess the
land which the
{\ND{
LORD}} your
God
giveth you, and ye shall
possess it, and
dwell therein.
\VS{32}And ye shall
observe to
do all the
statutes and
judgments which I
set
before you this
day.

\par }\Chap{12}{\PP \VerseOne{1}These
{\ADD{are}} the
statutes and
judgments, which ye shall
observe to
do in the
land, which the
{\ND{
LORD}}
God of thy
fathers
giveth thee to
possess it, all the
days that ye
live upon the
earth.
\VS{2}Ye shall
utterly
destroy all the
places, wherein the
nations
which ye shall
possess
served their
gods, upon the
high
mountains, and upon the
hills, and under every
green
tree:\FTNT{*}{{\FR 12.2 }
possess: or, inherit}
\VS{3}And ye shall
overthrow their
altars, and
break their
pillars, and
burn their
groves with
fire; and ye shall hew
down the graven
images of their
gods, and
destroy the
names of them out of that
place.\FTNT{†}{{\FR 12.3 }
overthrow: Heb. break down}
\VS{4}Ye shall not do
so unto the
{\ND{
LORD}} your
God.
\par }{\PP \VS{5}But unto the
place which the
{\ND{
LORD}} your
God shall
choose out of all your
tribes to
put his
name there,
{\ADD{even}} unto his
habitation shall ye
seek, and thither thou shalt
come:
\VS{6}And thither ye shall
bring your burnt
offerings, and your
sacrifices, and your
tithes, and heave
offerings of your
hand, and your
vows, and your freewill
offerings, and the
firstlings of your
herds and of your
flocks:
\VS{7}And there ye shall
eat
before the
{\ND{
LORD}} your
God, and ye shall
rejoice in all that ye
put your
hand unto, ye and your
households, wherein the
{\ND{
LORD}} thy
God hath
blessed thee.
\VS{8}Ye shall not
do after all
{\ADD{the things}} that we
do here this
day, every
man whatsoever
{\ADD{is}}
right in his own
eyes.
\VS{9}For ye are not as yet
come to the
rest and to the
inheritance, which the
{\ND{
LORD}} your
God
giveth you.
\VS{10}But
{\ADD{when}} ye go
over
Jordan, and
dwell in the
land which the
{\ND{
LORD}} your
God giveth you to
inherit, and
{\ADD{when}} he giveth you
rest from all your
enemies round
about, so that ye
dwell in
safety;
\VS{11}Then there shall be a
place which the
{\ND{
LORD}} your
God shall
choose to cause his
name to
dwell there; thither shall ye
bring all that I
command you; your burnt
offerings, and your
sacrifices, your
tithes, and the heave
offering of your
hand, and all your
choice
vows which ye
vow unto the
{\ND{
LORD}}:\FTNT{‡}{{\FR 12.11 }
your choice…: Heb. the choice of your vows}
\VS{12}And ye shall
rejoice
before the
{\ND{
LORD}} your
God, ye, and your
sons, and your
daughters, and your
menservants, and your
maidservants, and the
Levite that
{\ADD{is}} within your
gates;
forasmuch as he hath no
part nor
inheritance with you.
\VS{13}Take
heed to thyself that thou
offer not thy burnt
offerings in every
place that thou
seest:
\VS{14}But in the
place which the
{\ND{
LORD}} shall
choose in
one of thy
tribes, there thou shalt
offer thy burnt
offerings, and there thou shalt
do all that I
command thee.
\VS{15}Notwithstanding thou mayest
kill and
eat
flesh in all thy
gates, whatsoever thy
soul lusteth
after, according to the
blessing of the
{\ND{
LORD}} thy
God which he hath
given thee: the
unclean and the
clean may
eat thereof, as of the
roebuck, and as of the
hart.
\VS{16}Only ye shall not
eat the
blood; ye shall
pour it upon the
earth as
water.
\par }{\PP \VS{17}¶ Thou
mayest not
eat within thy
gates the
tithe of thy
corn, or of thy
wine, or of thy
oil, or the
firstlings of thy
herds or of thy
flock, nor any of thy
vows which thou
vowest, nor thy freewill
offerings, or heave
offering of thine
hand:
\VS{18}But thou must
eat them
before the
{\ND{
LORD}} thy
God in the
place which the
{\ND{
LORD}} thy
God shall
choose, thou, and thy
son, and thy
daughter, and thy
manservant, and thy
maidservant, and the
Levite that
{\ADD{is}} within thy
gates: and thou shalt
rejoice
before the
{\ND{
LORD}} thy
God in all that thou
puttest thine
hands unto.
\VS{19}Take
heed to thyself that thou
forsake not the
Levite as long as thou
livest upon the
earth.\FTNT{§}{{\FR 12.19 }
as long…: Heb. all thy days}
\par }{\PP \VS{20}¶ When the
{\ND{
LORD}} thy
God shall
enlarge thy
border, as he hath
promised thee, and thou shalt
say, I will
eat
flesh, because thy
soul
longeth to
eat
flesh; thou mayest
eat
flesh, whatsoever thy
soul lusteth
after.
\VS{21}If the
place which the
{\ND{
LORD}} thy
God hath
chosen to
put his
name there be too
far from thee, then thou shalt
kill of thy
herd and of thy
flock, which the
{\ND{
LORD}} hath
given thee, as I have
commanded thee, and thou shalt
eat in thy
gates whatsoever thy
soul lusteth
after.
\VS{22}Even as the
roebuck and the
hart is
eaten, so thou shalt
eat them: the
unclean and the
clean shall
eat
{\ADD{of}} them
alike.
\VS{23}Only be
sure that thou
eat not the
blood: for the
blood
{\ADD{is}} the
life; and thou mayest not
eat the
life with the
flesh.\FTNT{**}{{\FR 12.23 }
be…: Heb. be strong}
\VS{24}Thou shalt not
eat it; thou shalt
pour it upon the
earth as
water.
\VS{25}Thou shalt not
eat it; that it may go
well with thee, and with thy
children
after thee, when thou shalt
do
{\ADD{that which is}}
right in the
sight of the
{\ND{
LORD}}.
\VS{26}Only thy holy
things which thou hast, and thy
vows, thou shalt
take, and
go unto the
place which the
{\ND{
LORD}} shall
choose:
\VS{27}And thou shalt
offer thy burnt
offerings, the
flesh and the
blood, upon the
altar of the
{\ND{
LORD}} thy
God: and the
blood of thy
sacrifices shall be poured
out upon the
altar of the
{\ND{
LORD}} thy
God, and thou shalt
eat the
flesh.
\VS{28}Observe and
hear all these
words which I
command thee, that it may go
well with thee, and with thy
children
after thee
for
ever, when thou
doest
{\ADD{that which is}}
good and
right in the
sight of the
{\ND{
LORD}} thy
God.
\par }{\PP \VS{29}¶ When the
{\ND{
LORD}} thy
God shall cut
off the
nations from
before thee, whither thou
goest to
possess them, and thou
succeedest them, and
dwellest in their
land;\FTNT{††}{{\FR 12.29 }
succeedest…: Heb. inheritest, or, possessest them}
\VS{30}Take
heed to thyself that thou be not
snared by
following them,
after that they be
destroyed from
before thee; and that thou
enquire not after their
gods,
saying, How did these
nations
serve their
gods? even so will I
do
likewise.\FTNT{‡‡}{{\FR 12.30 }
by…: Heb. after them}
\VS{31}Thou shalt not
do so unto the
{\ND{
LORD}} thy
God: for every
abomination to the
{\ND{
LORD}}, which he
hateth, have they
done unto their
gods; for even their
sons and their
daughters they have
burnt in the
fire to their
gods.\FTNT{§§}{{\FR 12.31 }
to the: Heb. of the}
\VS{32}What
thing soever I
command you,
observe to
do it: thou shalt not
add thereto, nor
diminish from it.

\par }\Chap{13}{\PP \VerseOne{1}If there
arise
among you a
prophet, or a
dreamer of
dreams, and
giveth thee a
sign or a
wonder,
\VS{2}And the
sign or the
wonder come to
pass, whereof he
spake unto thee,
saying, Let us
go
after
other
gods, which thou hast not
known, and let us
serve them;
\VS{3}Thou shalt not
hearken unto the
words of that
prophet, or that
dreamer of
dreams: for the
{\ND{
LORD}} your
God
proveth you, to
know whether
ye
love the
{\ND{
LORD}} your
God with all your
heart and with all your
soul.
\VS{4}Ye shall
walk
after the
{\ND{
LORD}} your
God, and
fear him, and
keep his
commandments, and
obey his
voice, and ye shall
serve him, and
cleave unto him.
\VS{5}And that
prophet, or that
dreamer of
dreams, shall be put to
death; because he hath
spoken to
turn
{\ADD{you}} away from the
{\ND{
LORD}} your
God, which brought you
out of the
land of
Egypt, and
redeemed you out of the
house of
bondage, to
thrust thee out of the
way which the
{\ND{
LORD}} thy
God
commanded thee to walk
in. So shalt thou
put the
evil
away from the
midst of thee.\FTNT{*}{{\FR 13.5 }
to turn…: Heb. revolt against the
{\ND{LORD}}}
\par }{\PP \VS{6}¶ If thy
brother, the
son of thy
mother, or thy
son, or thy
daughter, or the
wife of thy
bosom, or thy
friend, which
{\ADD{is}} as thine own
soul,
entice thee
secretly,
saying, Let us
go and
serve
other
gods, which thou hast not
known, thou, nor thy
fathers;
\VS{7}{\ADD{Namely}}, of the
gods of the
people which
{\ADD{are}} round
about you,
nigh unto thee, or far
off from thee, from the
{\ADD{one}}
end of the
earth even unto the
{\ADD{other}}
end of the
earth;
\VS{8}Thou shalt not
consent unto him, nor
hearken unto him; neither shall thine
eye
pity him, neither shalt thou
spare, neither shalt thou
conceal him:
\VS{9}But thou shalt
surely
kill him; thine
hand shall be
first upon him to put him to
death, and
afterwards the
hand of all the
people.
\VS{10}And thou shalt
stone him with
stones, that he
die; because he hath
sought to thrust thee
away from the
{\ND{
LORD}} thy
God, which brought thee
out of the
land of
Egypt, from the
house of
bondage.\FTNT{†}{{\FR 13.10 }
bondage: Heb. bondmen}
\VS{11}And all
Israel shall
hear, and
fear, and shall
do no
more
any such
wickedness as this is
among you.
\par }{\PP \VS{12}¶ If thou shalt
hear
{\ADD{say}} in
one of thy
cities, which the
{\ND{
LORD}} thy
God hath
given thee to
dwell there,
saying,
\VS{13}{\ADD{Certain}}
men, the
children of
Belial, are gone
out from
among you, and have
withdrawn the
inhabitants of their
city,
saying, Let us
go and
serve
other
gods, which ye have not
known;\FTNT{‡}{{\FR 13.13 }
the children…: or, naughty men}
\VS{14}Then shalt thou
enquire, and make
search, and
ask
diligently; and, behold,
{\ADD{if it be}}
truth,
{\ADD{and}} the
thing
certain,
{\ADD{that}}
such
abomination is
wrought
among you;
\VS{15}Thou shalt
surely
smite the
inhabitants of that
city with the
edge of the
sword, destroying it
utterly, and all that
{\ADD{is}} therein, and the
cattle thereof, with the
edge of the
sword.
\VS{16}And thou shalt
gather all the
spoil of it into the
midst of the
street thereof, and shalt
burn with
fire the
city, and all the
spoil thereof every
whit, for the
{\ND{
LORD}} thy
God: and it shall be an
heap for
ever; it shall not be built
again.
\VS{17}And there shall
cleave
nought of the cursed
thing to thine
hand: that the
{\ND{
LORD}} may
turn from the
fierceness of his
anger, and
shew thee
mercy, and have
compassion upon thee, and
multiply thee, as he hath
sworn unto thy
fathers;\FTNT{§}{{\FR 13.17 }
cursed: or, devoted}
\VS{18}When thou shalt
hearken to the
voice of the
{\ND{
LORD}} thy
God, to
keep all his
commandments which I
command thee this
day, to
do
{\ADD{that which is}}
right in the
eyes of the
{\ND{
LORD}} thy
God.

\par }\Chap{14}{\PP \VerseOne{1}Ye
{\ADD{are}} the
children of the
{\ND{
LORD}} your
God: ye shall not
cut yourselves, nor
make any
baldness between your
eyes for the
dead.
\VS{2}For thou
{\ADD{art}} an
holy
people unto the
{\ND{
LORD}} thy
God, and the
{\ND{
LORD}} hath
chosen thee to be a
peculiar
people unto himself, above all the
nations that
{\ADD{are}}
upon the
earth.
\par }{\PP \VS{3}¶ Thou shalt not
eat any abominable
thing.
\VS{4}These
{\ADD{are}} the
beasts which ye shall
eat: the
ox, the
sheep, and the
goat,
\VS{5}The
hart, and the
roebuck, and the fallow
deer, and the wild
goat, and the
pygarg, and the wild
ox, and the
chamois.\FTNT{*}{{\FR 14.5 }
pygarg: or bison: Heb. dishon}
\VS{6}And every
beast that
parteth the
hoof, and
cleaveth the
cleft into
two
claws,
{\ADD{and}}
cheweth the
cud among the
beasts, that ye shall
eat.
\VS{7}Nevertheless these ye shall not
eat of them that
chew the
cud, or of them that
divide the
cloven
hoof;
{\ADD{as}} the
camel, and the
hare, and the
coney: for they
chew the
cud, but
divide not the
hoof;
{\ADD{therefore}} they
{\ADD{are}}
unclean unto you.
\VS{8}And the
swine, because it
divideth the
hoof, yet cheweth not the
cud, it
{\ADD{is}}
unclean unto you: ye shall not
eat of their
flesh, nor
touch their dead
carcase.
\par }{\PP \VS{9}¶ These ye shall
eat of all that
{\ADD{are}} in the
waters: all that have
fins and
scales shall ye
eat:
\VS{10}And whatsoever hath not
fins and
scales ye may not
eat; it
{\ADD{is}}
unclean unto you.
\par }{\PP \VS{11}¶
{\ADD{Of}} all
clean
birds ye shall
eat.
\VS{12}But these
{\ADD{are they}} of which ye shall not
eat: the
eagle, and the
ossifrage, and the
ospray,
\VS{13}And the
glede, and the
kite, and the
vulture after his
kind,
\VS{14}And every
raven after his
kind,
\VS{15}And the
owl, and the night
hawk, and the
cuckow, and the
hawk after his
kind,
\VS{16}The little
owl, and the great
owl, and the
swan,
\VS{17}And the
pelican, and the gier
eagle, and the
cormorant,
\VS{18}And the
stork, and the
heron after her
kind, and the
lapwing, and the
bat.
\VS{19}And every creeping
thing that
flieth
{\ADD{is}}
unclean unto you: they shall not be
eaten.
\VS{20}{\ADD{But of}} all
clean
fowls ye may
eat.
\par }{\PP \VS{21}¶ Ye shall not
eat
{\ADD{of}} any thing that dieth of
itself: thou shalt
give it unto the
stranger that
{\ADD{is}} in thy
gates, that he may
eat it; or thou mayest
sell it unto an
alien: for thou
{\ADD{art}} an
holy
people unto the
{\ND{
LORD}} thy
God. Thou shalt not
seethe a
kid in his
mother’s
milk.
\par }{\PP \VS{22}Thou shalt
truly
tithe all the
increase of thy
seed, that the
field bringeth
forth
year by
year.
\VS{23}And thou shalt
eat
before the
{\ND{
LORD}} thy
God, in the
place which he shall
choose to
place his
name there, the
tithe of thy
corn, of thy
wine, and of thine
oil, and the
firstlings of thy
herds and of thy
flocks; that thou mayest
learn to
fear the
{\ND{
LORD}} thy
God
always.
\VS{24}And if the
way be too
long for thee, so
that thou art not
able to
carry it;
{\ADD{or}} if the
place be too
far from thee, which the
{\ND{
LORD}} thy
God shall
choose to
set his
name there, when the
{\ND{
LORD}} thy
God hath
blessed thee:
\VS{25}Then shalt thou
turn
{\ADD{it}} into
money, and bind
up the
money in thine
hand, and shalt
go unto the
place which the
{\ND{
LORD}} thy
God shall
choose:
\VS{26}And thou shalt
bestow that
money for whatsoever thy
soul lusteth
after, for
oxen, or for
sheep, or for
wine, or for strong
drink, or for whatsoever thy
soul
desireth: and thou shalt
eat there
before the
{\ND{
LORD}} thy
God, and thou shalt
rejoice, thou, and thine
household,\FTNT{†}{{\FR 14.26 }
desireth: Heb. asketh of thee}
\VS{27}And the
Levite that
{\ADD{is}} within thy
gates; thou shalt not
forsake him; for he hath no
part nor
inheritance with thee.
\par }{\PP \VS{28}¶ At the
end of
three
years thou shalt bring
forth all the
tithe of thine
increase the same
year, and shalt lay
{\ADD{it}}
up within thy
gates:
\VS{29}And the
Levite, (because he hath no
part nor
inheritance with thee,) and the
stranger, and the
fatherless, and the
widow, which
{\ADD{are}} within thy
gates, shall
come, and shall
eat and be
satisfied; that the
{\ND{
LORD}} thy
God may
bless thee in all the
work of thine
hand which thou
doest.

\par }\Chap{15}{\PP \VerseOne{1}At the
end of
{\ADD{every}}
seven
years thou shalt
make a
release.
\VS{2}And this
{\ADD{is}} the
manner of the
release: Every
creditor that
lendeth
{\ADD{ought}} unto his
neighbour shall
release
{\ADD{it;}} he shall not
exact
{\ADD{it}} of his
neighbour, or of his
brother; because it is
called the
{\ND{
LORD’s}}
release.\FTNT{*}{{\FR 15.2 }
creditor: Heb. master of the lending of his hand}
\VS{3}Of a
foreigner thou mayest
exact
{\ADD{it again:}} but
{\ADD{that}} which is thine with thy
brother thine
hand shall
release;
\VS{4}Save when there shall be no
poor among you; for the
{\ND{
LORD}} shall
greatly
bless thee in the
land which the
{\ND{
LORD}} thy
God
giveth thee
{\ADD{for}} an
inheritance to
possess it:\FTNT{†}{{\FR 15.4 }
Save…: or, To the end that there be no poor among you}
\VS{5}Only if thou
carefully
hearken unto the
voice of the
{\ND{
LORD}} thy
God, to
observe to
do all these
commandments which I
command thee this
day.
\VS{6}For the
{\ND{
LORD}} thy
God
blesseth thee, as he
promised thee: and thou shalt
lend unto
many
nations, but thou shalt not
borrow; and thou shalt
reign over
many
nations, but they shall not
reign over thee.
\par }{\PP \VS{7}¶ If there be among you a poor
man of
one of thy
brethren within
any of thy
gates in thy
land which the
{\ND{
LORD}} thy
God
giveth thee, thou shalt not
harden thine
heart, nor
shut thine
hand from thy
poor
brother:
\VS{8}But thou shalt
open thine
hand
wide unto him, and shalt
surely
lend him
sufficient for his
need,
{\ADD{in that}} which he
wanteth.
\VS{9}Beware that there be not a
thought in thy
wicked
heart,
saying, The
seventh
year, the
year of
release, is at
hand; and thine
eye be
evil against thy
poor
brother, and thou
givest him nought; and he
cry unto the
{\ND{
LORD}} against thee, and it be
sin unto thee.\FTNT{‡}{{\FR 15.9 }
thought: Heb. word}\FTNT{§}{{\FR 15.9 }
wicked: Heb. Belial}
\VS{10}Thou shalt
surely
give him, and thine
heart shall not be
grieved when thou
givest unto him:
because that for this
thing the
{\ND{
LORD}} thy
God shall
bless thee in all thy
works, and in all that thou
puttest thine
hand unto.
\VS{11}For the
poor shall never
cease out
of the
land: therefore I
command thee,
saying, Thou shalt
open thine
hand
wide unto thy
brother, to thy
poor, and to thy
needy, in thy
land.
\par }{\PP \VS{12}¶
{\ADD{And}} if thy
brother, an Hebrew
man, or an Hebrew
woman, be
sold unto thee, and
serve thee
six
years; then in the
seventh
year thou shalt let him
go
free from thee.
\VS{13}And when thou
sendest him out
free from thee, thou shalt not let him go
away
empty:
\VS{14}Thou shalt
furnish him
liberally out of thy
flock, and out of thy
floor, and out of thy
winepress:
{\ADD{of that}} wherewith the
{\ND{
LORD}} thy
God hath
blessed thee thou shalt
give unto him.
\VS{15}And thou shalt
remember that thou wast a
bondman in the
land of
Egypt, and the
{\ND{
LORD}} thy
God
redeemed thee: therefore I
command thee this
thing to
day.
\VS{16}And it shall be, if he
say unto thee, I will not go
away from thee; because he
loveth thee and thine
house, because he is
well with thee;
\VS{17}Then thou shalt
take an
aul, and
thrust
{\ADD{it}} through his
ear unto the
door, and he shall be thy
servant for
ever. And also unto thy
maidservant thou shalt
do likewise.
\VS{18}It shall not seem
hard unto
thee, when thou sendest him
away
free from thee; for he hath been
worth a
double hired
servant
{\ADD{to thee}}, in
serving thee
six
years: and the
{\ND{
LORD}} thy
God shall
bless thee in all that thou
doest.
\par }{\PP \VS{19}¶ All the
firstling
males that
come of thy
herd and of thy
flock thou shalt
sanctify unto the
{\ND{
LORD}} thy
God: thou shalt do no
work with the
firstling of thy
bullock, nor
shear the
firstling of thy
sheep.
\VS{20}Thou shalt
eat
{\ADD{it}}
before the
{\ND{
LORD}} thy
God
year by
year in the
place which the
{\ND{
LORD}} shall
choose, thou and thy
household.
\VS{21}And if there be
{\ADD{any}}
blemish therein,
{\ADD{as if it be}}
lame, or
blind,
{\ADD{or have}} any
ill
blemish, thou shalt not
sacrifice it unto the
{\ND{
LORD}} thy
God.
\VS{22}Thou shalt
eat it within thy
gates: the
unclean and the
clean
{\ADD{person shall eat it}}
alike, as the
roebuck, and as the
hart.
\VS{23}Only thou shalt not
eat the
blood thereof; thou shalt
pour it upon the
ground as
water.

\par }\Chap{16}{\PP \VerseOne{1}Observe the
month of
Abib, and
keep the
passover unto the
{\ND{
LORD}} thy
God: for in the
month of
Abib the
{\ND{
LORD}} thy
God brought thee
forth out of
Egypt by
night.
\VS{2}Thou shalt therefore
sacrifice the
passover unto the
{\ND{
LORD}} thy
God, of the
flock and the
herd, in the
place which the
{\ND{
LORD}} shall
choose to
place his
name there.
\VS{3}Thou shalt
eat no leavened
bread with it;
seven
days shalt thou
eat unleavened
bread therewith,
{\ADD{even}} the
bread of
affliction; for thou camest
forth out of the
land of
Egypt in
haste: that thou mayest
remember the
day when thou camest
forth out of the
land of
Egypt all the
days of thy
life.
\VS{4}And there shall be no leavened
bread
seen with thee in all thy
coast
seven
days; neither shall there
{\ADD{any thing}} of the
flesh, which thou
sacrificedst the
first
day at
even,
remain all night until the
morning.
\VS{5}Thou
mayest not
sacrifice the
passover within
any of thy
gates, which the
{\ND{
LORD}} thy
God
giveth thee:\FTNT{*}{{\FR 16.5 }
sacrifice: or, kill}
\VS{6}But at the
place which the
{\ND{
LORD}} thy
God shall
choose to
place his
name in, there thou shalt
sacrifice the
passover at
even, at the going
down of the
sun, at the
season that thou camest
forth out of
Egypt.
\VS{7}And thou shalt
roast and
eat
{\ADD{it}} in the
place which the
{\ND{
LORD}} thy
God shall
choose: and thou shalt
turn in the
morning, and
go unto thy
tents.
\VS{8}Six
days thou shalt
eat unleavened
bread: and on the
seventh
day
{\ADD{shall be}} a solemn
assembly to the
{\ND{
LORD}} thy
God: thou shalt
do no
work
{\ADD{therein}}.\FTNT{†}{{\FR 16.8 }
solemn…: Heb. restraint}
\par }{\PP \VS{9}¶
Seven
weeks shalt thou
number unto thee:
begin to
number the
seven
weeks from
{\ADD{such time as}} thou
beginnest
{\ADD{to put}} the
sickle to the
corn.
\VS{10}And thou shalt
keep the
feast of
weeks unto the
{\ND{
LORD}} thy
God with a
tribute of a freewill
offering of thine
hand, which thou shalt
give
{\ADD{unto the
{\ND{Lord}} thy God}}, according as the
{\ND{
LORD}} thy
God hath
blessed thee:\FTNT{‡}{{\FR 16.10 }
a tribute: or, sufficiency}
\VS{11}And thou shalt
rejoice
before the
{\ND{
LORD}} thy
God, thou, and thy
son, and thy
daughter, and thy
manservant, and thy
maidservant, and the
Levite that
{\ADD{is}} within thy
gates, and the
stranger, and the
fatherless, and the
widow, that
{\ADD{are}}
among you, in the
place which the
{\ND{
LORD}} thy
God hath
chosen to
place his
name there.
\VS{12}And thou shalt
remember that thou wast a
bondman in
Egypt: and thou shalt
observe and
do these
statutes.
\par }{\PP \VS{13}¶ Thou shalt
observe the
feast of
tabernacles
seven
days, after that thou hast
gathered in thy
corn and thy
wine:\FTNT{§}{{\FR 16.13 }
corn…: Heb. floor, and thy winepress}
\VS{14}And thou shalt
rejoice in thy
feast, thou, and thy
son, and thy
daughter, and thy
manservant, and thy
maidservant, and the
Levite, the
stranger, and the
fatherless, and the
widow, that
{\ADD{are}} within thy
gates.
\VS{15}Seven
days shalt thou keep a solemn
feast unto the
{\ND{
LORD}} thy
God in the
place which the
{\ND{
LORD}} shall
choose: because the
{\ND{
LORD}} thy
God shall
bless thee in all thine
increase, and in all the
works of thine
hands, therefore thou shalt surely
rejoice.
\par }{\PP \VS{16}¶
Three
times in a
year shall all thy
males
appear
before the
{\ND{
LORD}} thy
God in the
place which he shall
choose; in the
feast of unleavened
bread, and in the
feast of
weeks, and in the
feast of
tabernacles: and they shall not
appear
before the
{\ND{
LORD}}
empty:
\VS{17}Every
man
{\ADD{shall give}} as he is
able, according to the
blessing of the
{\ND{
LORD}} thy
God which he hath
given thee.\FTNT{**}{{\FR 16.17 }
as…: Heb. according to the gift of his hand}
\par }{\PP \VS{18}¶
Judges and
officers shalt thou
make thee in all thy
gates, which the
{\ND{
LORD}} thy
God
giveth thee, throughout thy
tribes: and they shall
judge the
people with
just
judgment.
\VS{19}Thou shalt not
wrest
judgment; thou shalt not
respect
persons, neither
take a
gift: for a
gift doth
blind the
eyes of the
wise, and
pervert the
words of the
righteous.\FTNT{††}{{\FR 16.19 }
words: or, matters}
\VS{20}That which is
altogether
just shalt thou
follow, that thou mayest
live, and
inherit the
land which the
{\ND{
LORD}} thy
God
giveth thee.\FTNT{‡‡}{{\FR 16.20 }
That which…: Heb. Justice, justice}
\par }{\PP \VS{21}¶ Thou shalt not
plant thee a
grove of any
trees near
unto the
altar of the
{\ND{
LORD}} thy
God, which thou shalt
make thee.
\VS{22}Neither shalt thou set thee
up
{\ADD{any}}
image; which the
{\ND{
LORD}} thy
God
hateth.\FTNT{§§}{{\FR 16.22 }
image: or, statue, or, pillar}

\par }\Chap{17}{\PP \VerseOne{1}Thou shalt not
sacrifice unto the
{\ND{
LORD}} thy
God
{\ADD{any}}
bullock, or
sheep, wherein is
blemish,
{\ADD{or}} any
evilfavouredness: for that
{\ADD{is}} an
abomination unto the
{\ND{
LORD}} thy
God.\FTNT{*}{{\FR 17.1 }
sheep: or, goat}
\par }{\PP \VS{2}¶ If there be
found
among you, within
any of thy
gates which the
{\ND{
LORD}} thy
God
giveth thee,
man or
woman, that hath
wrought
wickedness in the
sight of the
{\ND{
LORD}} thy
God, in
transgressing his
covenant,
\VS{3}And hath
gone and
served
other
gods, and
worshipped them, either the
sun, or
moon, or any of the
host of
heaven, which I have not
commanded;
\VS{4}And it be
told thee, and thou hast
heard
{\ADD{of it}}, and
enquired
diligently, and, behold,
{\ADD{it be}}
true,
{\ADD{and}} the
thing
certain,
{\ADD{that}} such
abomination is
wrought in
Israel:
\VS{5}Then shalt thou bring
forth that
man or that
woman, which have
committed that
wicked
thing, unto thy
gates,
{\ADD{even}} that
man or that
woman, and shalt
stone them with
stones, till they
die.
\VS{6}At the
mouth of
two
witnesses, or
three
witnesses, shall he that is worthy of
death be put to
death;
{\ADD{but}} at the
mouth of
one
witness he shall not be put to
death.
\VS{7}The
hands of the
witnesses shall be
first upon him to put him to
death, and
afterward the
hands of all the
people. So thou shalt
put the
evil
away from
among you.
\par }{\PP \VS{8}¶ If there arise a
matter too
hard for thee in
judgment, between
blood and
blood, between
plea and
plea, and between
stroke and
stroke,
{\ADD{being}}
matters of
controversy within thy
gates: then shalt thou
arise, and get thee
up into the
place which the
{\ND{
LORD}} thy
God shall
choose;
\VS{9}And thou shalt
come unto the
priests the
Levites, and unto the
judge that shall be in those
days, and
enquire; and they shall
shew thee the
sentence of
judgment:
\VS{10}And thou shalt
do according to the
sentence, which they of that
place which the
{\ND{
LORD}} shall
choose shall
shew thee; and thou shalt
observe to
do according to all that they
inform thee:
\VS{11}According to the
sentence of the
law which they shall
teach thee, and according to the
judgment which they shall
tell thee, thou shalt
do: thou shalt not
decline from the
sentence which they shall
shew thee,
{\ADD{to}} the right
hand, nor
{\ADD{to}} the
left.
\VS{12}And the
man that will
do
presumptuously, and will not
hearken unto the
priest that
standeth to
minister there before the
{\ND{
LORD}} thy
God, or unto the
judge, even that
man shall
die: and thou shalt put
away the
evil from
Israel.\FTNT{†}{{\FR 17.12 }
and will…: Heb. not to hearken}
\VS{13}And all the
people shall
hear, and
fear, and do no more
presumptuously.
\par }{\PP \VS{14}¶ When thou art
come unto the
land which the
{\ND{
LORD}} thy
God
giveth thee, and shalt
possess it, and shalt
dwell therein, and shalt
say, I will
set a
king over me, like as all the
nations that
{\ADD{are}}
about me;
\VS{15}Thou shalt in any
wise
set
{\ADD{him}}
king over thee, whom the
{\ND{
LORD}} thy
God shall
choose:
{\ADD{one}} from
among thy
brethren shalt thou
set
king over thee: thou
mayest not
set a
stranger over thee, which
{\ADD{is}} not thy
brother.
\VS{16}But he shall not
multiply
horses to himself, nor cause the
people to
return to
Egypt, to the end that he should
multiply
horses: forasmuch as the
{\ND{
LORD}} hath
said unto you, Ye shall
henceforth
return no more that
way.
\VS{17}Neither shall he
multiply
wives to himself, that his
heart
turn not away: neither shall he
greatly
multiply to himself
silver and
gold.
\VS{18}And it shall be, when he
sitteth upon the
throne of his
kingdom, that he shall
write him a
copy of this
law in a
book out of
{\ADD{that which is}}
before the
priests the
Levites:
\VS{19}And it shall be with him, and he shall
read therein all the
days of his
life: that he may
learn to
fear the
{\ND{
LORD}} his
God, to
keep all the
words of this
law and these
statutes, to
do them:
\VS{20}That his
heart be not lifted
up above his
brethren, and that he turn not
aside from the
commandment,
{\ADD{to}} the right
hand, or
{\ADD{to}} the
left: to the end that he may
prolong
{\ADD{his}}
days in his
kingdom, he, and his
children, in the
midst of
Israel.

\par }\Chap{18}{\PP \VerseOne{1}The
priests the
Levites,
{\ADD{and}} all the
tribe of
Levi, shall have no
part nor
inheritance with
Israel: they shall
eat the
offerings of the
{\ND{
LORD}} made by
fire, and his
inheritance.
\VS{2}Therefore shall they have no
inheritance
among their
brethren: the
{\ND{
LORD}}
{\ADD{is}} their
inheritance, as he hath
said unto them.
\par }{\PP \VS{3}¶ And this shall be the
priest’s
due from the
people, from them that
offer a
sacrifice, whether
{\ADD{it be}}
ox or
sheep; and they shall
give unto the
priest the
shoulder, and the two
cheeks, and the
maw.
\VS{4}The
firstfruit
{\ADD{also}} of thy
corn, of thy
wine, and of thine
oil, and the
first of the
fleece of thy
sheep, shalt thou
give him.
\VS{5}For the
{\ND{
LORD}} thy
God hath
chosen him out of all thy
tribes, to
stand to
minister in the
name of the
{\ND{
LORD}}, him and his
sons for
ever.
\par }{\PP \VS{6}¶ And if a
Levite
come from
any of thy
gates out of all
Israel, where he
sojourned, and
come with all the
desire of his
mind unto the
place which the
{\ND{
LORD}} shall
choose;
\VS{7}Then he shall
minister in the
name of the
{\ND{
LORD}} his
God, as all his
brethren the
Levites
{\ADD{do}}, which
stand there
before the
{\ND{
LORD}}.
\VS{8}They shall have like
portions to
eat, beside that which cometh of the
sale of his
patrimony.\FTNT{*}{{\FR 18.8 }
that…: Heb. his sale by the fathers}
\par }{\PP \VS{9}¶ When thou art
come into the
land which the
{\ND{
LORD}} thy
God
giveth thee, thou shalt not
learn to
do after the
abominations of those
nations.
\VS{10}There shall not be
found among you
{\ADD{any one}} that maketh his
son or his
daughter to
pass through the
fire,
{\ADD{or}} that
useth
divination,
{\ADD{or}} an observer of
times, or an
enchanter, or a
witch,
\VS{11}Or a
charmer, or a
consulter with familiar
spirits, or a
wizard, or a
necromancer.
\VS{12}For all that
do these things
{\ADD{are}} an
abomination unto the
{\ND{
LORD}}: and
because of these
abominations the
{\ND{
LORD}} thy
God doth drive them
out from
before thee.
\VS{13}Thou shalt be
perfect with the
{\ND{
LORD}} thy
God.\FTNT{†}{{\FR 18.13 }
perfect: or, upright, or, sincere}
\VS{14}For these
nations, which thou shalt
possess,
hearkened unto observers of
times, and unto
diviners: but as for thee, the
{\ND{
LORD}} thy
God hath not
suffered thee so
{\ADD{to do}}.\FTNT{‡}{{\FR 18.14 }
possess: or, inherit}
\par }{\PP \VS{15}¶ The
{\ND{
LORD}} thy
God will raise
up unto thee a
Prophet from the
midst of thee, of thy
brethren, like unto me; unto him ye shall
hearken;
\VS{16}According to all that thou
desiredst of the
{\ND{
LORD}} thy
God in
Horeb in the
day of the
assembly,
saying, Let me not
hear
again the
voice of the
{\ND{
LORD}} my
God, neither let me
see this
great
fire any more, that I
die not.
\VS{17}And the
{\ND{
LORD}}
said unto me, They have
well
{\ADD{spoken that}} which they have
spoken.
\VS{18}I will raise them
up a
Prophet from
among their
brethren, like unto thee, and will
put my
words in his
mouth; and he shall
speak unto them all that I shall
command him.
\VS{19}And it shall come to pass,
{\ADD{that}}
whosoever will not
hearken unto my
words which he shall
speak in my
name, I will
require
{\ADD{it}} of him.
\VS{20}But the
prophet, which shall
presume to
speak a
word in my
name, which I have not
commanded him to
speak, or that shall
speak in the
name of
other
gods, even that
prophet shall
die.
\VS{21}And if thou
say in thine
heart, How shall we
know the
word which the
{\ND{
LORD}} hath not
spoken?
\VS{22}When a
prophet
speaketh in the
name of the
{\ND{
LORD}}, if the
thing follow not, nor come to
pass, that
{\ADD{is}} the
thing which the
{\ND{
LORD}} hath not
spoken,
{\ADD{but}} the
prophet hath
spoken it
presumptuously: thou shalt not be
afraid of him.

\par }\Chap{19}{\PP \VerseOne{1}When the
{\ND{
LORD}} thy
God hath cut
off the
nations, whose
land the
{\ND{
LORD}} thy
God
giveth thee, and thou
succeedest them, and
dwellest in their
cities, and in their
houses;\FTNT{*}{{\FR 19.1 }
succeedest: Heb. inheritest, or, possessest}
\VS{2}Thou shalt
separate
three
cities for thee in the
midst of thy
land, which the
{\ND{
LORD}} thy
God
giveth thee to
possess it.
\VS{3}Thou shalt
prepare thee a
way, and divide the
coasts of thy
land, which the
{\ND{
LORD}} thy
God giveth thee to
inherit, into three
parts, that every
slayer may
flee thither.
\par }{\PP \VS{4}¶ And this
{\ADD{is}} the
case of the
slayer, which shall
flee thither, that he may
live: Whoso
killeth his
neighbour
ignorantly, whom he
hated not in
time
past;\FTNT{†}{{\FR 19.4 }
in…: Heb. from yesterday the third day}
\VS{5}As when a man
goeth into the
wood with his
neighbour to
hew
wood, and his
hand fetcheth a
stroke with the
axe to cut
down the
tree, and the
head
slippeth from the
helve, and
lighteth upon his
neighbour, that he
die; he shall
flee unto
one of those
cities, and
live:\FTNT{‡}{{\FR 19.5 }
head: Heb. iron}\FTNT{§}{{\FR 19.5 }
helve: Heb. wood}\FTNT{**}{{\FR 19.5 }
lighteth…: Heb. findeth}
\VS{6}Lest the
avenger of the
blood
pursue the
slayer,
while his
heart is
hot, and
overtake him, because the
way is
long, and
slay
him; whereas he
{\ADD{was}} not
worthy of
death, inasmuch
as he
hated him not in
time
past.\FTNT{††}{{\FR 19.6 }
slay…: Heb. smite him in life}\FTNT{‡‡}{{\FR 19.6 }
in…: Heb. from yesterday the third day}
\VS{7}Wherefore I
command thee,
saying, Thou shalt
separate
three
cities for thee.
\VS{8}And if the
{\ND{
LORD}} thy
God
enlarge thy
coast, as he hath
sworn unto thy
fathers, and
give thee all the
land which he
promised to
give unto thy
fathers;
\VS{9}If thou shalt
keep all these
commandments to
do them, which I
command thee this
day, to
love the
{\ND{
LORD}} thy
God, and to
walk
ever in his
ways; then shalt thou
add
three
cities more for thee, beside these
three:
\VS{10}That
innocent
blood be not
shed
in thy
land, which the
{\ND{
LORD}} thy
God
giveth thee
{\ADD{for}} an
inheritance, and
{\ADD{so}}
blood be upon thee.
\par }{\PP \VS{11}¶ But if any
man
hate his
neighbour, and lie in
wait for him, and rise
up against him, and
smite him
mortally that he
die, and
fleeth into
one of
these
cities:\FTNT{§§}{{\FR 19.11 }
mortally: Heb. in life}
\VS{12}Then the
elders of his
city shall
send and
fetch him thence, and
deliver him into the
hand of the
avenger of
blood, that he may
die.
\VS{13}Thine
eye shall not
pity him, but thou shalt put
away
{\ADD{the guilt of}}
innocent
blood from
Israel, that it may go
well with thee.
\par }{\PP \VS{14}¶ Thou shalt not
remove thy
neighbour’s
landmark, which they of old
time have
set in thine
inheritance, which thou shalt
inherit in the
land that the
{\ND{
LORD}} thy
God
giveth thee to
possess it.
\par }{\PP \VS{15}¶
One
witness shall not rise
up against a
man for any
iniquity, or for any
sin, in any
sin that he
sinneth: at the
mouth of
two
witnesses, or at the
mouth of
three
witnesses, shall the
matter be
established.
\par }{\PP \VS{16}¶ If a
false
witness rise
up against any
man to
testify against him
{\ADD{that which is}}
wrong;\FTNT{***}{{\FR 19.16 }
that…: or, falling away}
\VS{17}Then
both the
men, between whom the
controversy
{\ADD{is}}, shall
stand
before the
{\ND{
LORD}},
before the
priests and the
judges, which shall be in those
days;
\VS{18}And the
judges shall make
diligent
inquisition: and, behold,
{\ADD{if}} the
witness
{\ADD{be}} a
false
witness,
{\ADD{and}} hath
testified
falsely against his
brother;
\VS{19}Then shall ye
do unto him, as he had
thought to have
done unto his
brother: so shalt thou
put the
evil
away from
among you.
\VS{20}And those which
remain shall
hear, and
fear, and shall
henceforth
commit no more
any such
evil
among you.
\VS{21}And thine
eye shall not
pity;
{\ADD{but}}
life
{\ADD{shall go}} for
life,
eye for
eye,
tooth for
tooth,
hand for
hand,
foot for
foot.

\par }\Chap{20}{\PP \VerseOne{1}When thou goest
out to
battle against thine
enemies, and
seest
horses, and
chariots,
{\ADD{and}} a
people
more than thou, be not
afraid of them: for the
{\ND{
LORD}} thy
God
{\ADD{is}} with thee, which brought thee
up out of the
land of
Egypt.
\VS{2}And it shall be, when ye are come
nigh unto the
battle, that the
priest shall
approach and
speak unto the
people,
\VS{3}And shall
say unto them,
Hear, O
Israel, ye
approach this
day unto
battle against your
enemies: let not your
hearts
faint,
fear not, and do not
tremble, neither be ye
terrified
because of them;\FTNT{*}{{\FR 20.3 }
faint: Heb. be tender}\FTNT{†}{{\FR 20.3 }
tremble: Heb. make haste}
\VS{4}For the
{\ND{
LORD}} your
God
{\ADD{is}} he that
goeth with you, to
fight for you against your
enemies, to
save you.
\par }{\PP \VS{5}¶ And the
officers shall
speak unto the
people,
saying, What
man
{\ADD{is there}} that hath
built a
new
house, and hath not
dedicated it? let him
go and
return to his
house, lest he
die in the
battle, and
another
man
dedicate it.
\VS{6}And what
man
{\ADD{is he}} that hath
planted a
vineyard, and hath not
{\ADD{yet}}
eaten of it? let him
{\ADD{also}}
go and
return unto his
house, lest he
die in the
battle, and
another
man
eat of it.\FTNT{‡}{{\FR 20.6 }
eaten…: Heb. made it common}
\VS{7}And what
man
{\ADD{is there}} that hath
betrothed a
wife, and hath not
taken her? let him
go and
return unto his
house, lest he
die in the
battle, and
another
man
take her.
\VS{8}And the
officers shall
speak
further unto the
people, and they shall
say, What
man
{\ADD{is there that is}}
fearful and
fainthearted? let him
go and
return unto his
house, lest his
brethren’s
heart
faint as well as his
heart.\FTNT{§}{{\FR 20.8 }
faint: Heb. melt}
\VS{9}And it shall be, when the
officers have made an
end of
speaking unto the
people, that they shall
make
captains of the
armies to
lead the
people.\FTNT{**}{{\FR 20.9 }
to lead…: Heb. to be in the head of the people}
\par }{\PP \VS{10}¶ When thou comest
nigh unto a
city to
fight against it, then
proclaim
peace unto it.
\VS{11}And it shall be, if it make thee
answer of
peace, and
open unto thee, then it shall be,
{\ADD{that}} all the
people
{\ADD{that is}}
found therein shall be
tributaries unto thee, and they shall
serve thee.
\VS{12}And if it will make no
peace with thee, but will
make
war against thee, then thou shalt
besiege it:
\VS{13}And when the
{\ND{
LORD}} thy
God hath
delivered it into thine
hands, thou shalt
smite every
male thereof with the
edge of the
sword:
\VS{14}But the
women, and the little
ones, and the
cattle, and all that is in the
city,
{\ADD{even}} all the
spoil thereof, shalt thou
take unto thyself; and thou shalt
eat the
spoil of thine
enemies, which the
{\ND{
LORD}} thy
God hath
given thee.\FTNT{††}{{\FR 20.14 }
take: Heb. spoil}
\VS{15}Thus shalt thou
do unto all the
cities
{\ADD{which are}}
very far
off from thee,
which
{\ADD{are}} not of the
cities of these
nations.
\VS{16}But of the
cities of these
people, which the
{\ND{
LORD}} thy
God doth
give thee
{\ADD{for}} an
inheritance, thou shalt save
alive nothing that
breatheth:
\VS{17}But thou shalt
utterly
destroy them;
{\ADD{namely}}, the
Hittites, and the
Amorites, the
Canaanites, and the
Perizzites, the
Hivites, and the
Jebusites; as the
{\ND{
LORD}} thy
God hath
commanded thee:
\VS{18}That they
teach you not to
do after all their
abominations, which they have
done unto their
gods; so should ye
sin against the
{\ND{
LORD}} your
God.
\par }{\PP \VS{19}¶ When thou shalt
besiege a
city a
long
time, in making
war against it to
take it, thou shalt not
destroy the
trees thereof by
forcing an
axe against them: for thou mayest
eat of them, and thou shalt not cut them
down (for the
tree of the
field
{\ADD{is}}
man’s
{\ADD{life}}) to
employ
{\ADD{them}} in the
siege:\FTNT{‡‡}{{\FR 20.19 }
for the…: or, for, O man, the tree of the field is to be employed in the siege}\FTNT{§§}{{\FR 20.19 }
to employ…: Heb. to go from before thee}
\VS{20}Only the
trees which thou
knowest that they
{\ADD{be}} not
trees for
meat, thou shalt
destroy and cut them
down; and thou shalt
build
bulwarks against the
city that
maketh
war with thee, until it be
subdued.\FTNT{***}{{\FR 20.20 }
it…: Heb. it come down}

\par }\Chap{21}{\PP \VerseOne{1}If
{\ADD{one}} be
found
slain in the
land which the
{\ND{
LORD}} thy
God
giveth thee to
possess it,
lying in the
field,
{\ADD{and}} it be not
known who hath
slain him:
\VS{2}Then thy
elders and thy
judges shall come
forth, and they shall
measure unto the
cities which
{\ADD{are}} round
about him that is
slain:
\VS{3}And it shall be,
{\ADD{that}} the
city
{\ADD{which is}}
next unto the slain
man, even the
elders of that
city shall
take an
heifer, which hath not been wrought
with,
{\ADD{and}} which hath not
drawn in the
yoke;
\VS{4}And the
elders of that
city shall bring
down the
heifer unto a
rough
valley, which is neither
eared nor
sown, and shall strike
off the
heifer’s
neck there in the
valley:
\VS{5}And the
priests the
sons of
Levi shall come
near; for them the
{\ND{
LORD}} thy
God hath
chosen to
minister unto him, and to
bless in the
name of the
{\ND{
LORD}}; and by their
word shall every
controversy and every
stroke be
{\ADD{tried:}}\FTNT{*}{{\FR 21.5 }
word: Heb. mouth}
\VS{6}And all the
elders of that
city,
{\ADD{that are}}
next unto the
slain
{\ADD{man}}, shall
wash their
hands over the
heifer that is
beheaded in the
valley:
\VS{7}And they shall
answer and
say, Our
hands have not
shed this
blood, neither have our
eyes
seen
{\ADD{it}}.
\VS{8}Be
merciful, O
{\ND{
LORD}}, unto thy
people
Israel, whom thou hast
redeemed, and
lay not
innocent
blood unto thy
people of
Israel’s
charge. And the
blood shall be
forgiven them.\FTNT{†}{{\FR 21.8 }
unto thy people of: Heb. in the midst, etc}
\VS{9}So shalt thou put
away the
{\ADD{guilt of}}
innocent
blood from
among you, when thou shalt
do
{\ADD{that which is}}
right in the
sight of the
{\ND{
LORD}}.
\par }{\PP \VS{10}¶ When thou goest
forth to
war against thine
enemies, and the
{\ND{
LORD}} thy
God hath
delivered them into thine
hands, and thou hast
taken them
captive,
\VS{11}And
seest among the
captives a
beautiful
woman, and hast a
desire unto her, that thou wouldest have
her to thy
wife;
\VS{12}Then thou shalt
bring her
home to thine
house; and she shall
shave her
head, and
pare her
nails;\FTNT{‡}{{\FR 21.12 }
pare: or, suffer to grow: Heb. make, or, dress}
\VS{13}And she shall
put the
raiment of her
captivity from off her, and shall
remain in thine
house, and
bewail her
father and her
mother a
full
month: and
after that thou shalt
go in unto her, and be her
husband, and she shall be thy
wife.
\VS{14}And it shall be, if thou have no
delight in her, then thou shalt let her
go whither she
will; but thou shalt not
sell her at
all for
money, thou shalt not make
merchandise of her,
because thou hast
humbled her.
\par }{\PP \VS{15}¶ If a
man have
two
wives,
one
beloved, and
another
hated, and they have
born him
children,
{\ADD{both}} the
beloved and the
hated; and
{\ADD{if}} the
firstborn
son be hers that was
hated:
\VS{16}Then it shall be,
when he maketh his
sons to
inherit
{\ADD{that}} which he hath,
{\ADD{that}} he
may not make the
son of the
beloved
firstborn
before the
son of the
hated,
{\ADD{which is indeed}} the
firstborn:
\VS{17}But he shall
acknowledge the
son of the
hated
{\ADD{for}} the
firstborn, by
giving him a
double
portion of all that he
hath: for he
{\ADD{is}} the
beginning of his
strength; the
right of the
firstborn
{\ADD{is}} his.\FTNT{§}{{\FR 21.17 }
that…: Heb. that is found with him}
\par }{\PP \VS{18}¶ If a
man have a
stubborn and
rebellious
son, which will not
obey the
voice of his
father, or the
voice of his
mother, and
{\ADD{that}}, when they have
chastened him, will not
hearken unto them:
\VS{19}Then shall his
father and his
mother lay
hold on him, and bring him
out unto the
elders of his
city, and unto the
gate of his
place;
\VS{20}And they shall
say unto the
elders of his
city, This our
son
{\ADD{is}}
stubborn and
rebellious, he will not
obey our
voice;
{\ADD{he is}} a
glutton, and a
drunkard.
\VS{21}And all the
men of his
city shall
stone him with
stones, that he
die: so shalt thou
put
evil
away from
among you; and all
Israel shall
hear, and
fear.
\par }{\PP \VS{22}¶ And if a
man have committed a
sin worthy
of
death, and he be to be put to
death, and thou
hang him on a
tree:
\VS{23}His
body shall not remain all
night upon the
tree, but thou shalt in any
wise
bury him that
day; (for he that is
hanged
{\ADD{is}}
accursed of
God;) that thy
land be not
defiled, which the
{\ND{
LORD}} thy
God
giveth thee
{\ADD{for}} an
inheritance.\FTNT{**}{{\FR 21.23 }
accursed…: Heb. the curse of God}

\par }\Chap{22}{\PP \VerseOne{1}Thou shalt not
see thy
brother’s
ox or his
sheep go
astray, and
hide thyself from them: thou shalt in any
case bring them
again unto thy
brother.
\VS{2}And if thy
brother
{\ADD{be}} not
nigh unto thee, or if thou
know him not, then thou shalt
bring it
unto thine own
house, and it shall be with thee until thy
brother
seek after it, and thou shalt restore it to him
again.
\VS{3}In like
manner shalt thou
do with his
ass; and so shalt thou
do with his
raiment; and with all lost
thing of thy
brother’s, which he hath
lost, and thou hast
found, shalt thou
do likewise: thou
mayest not
hide thyself.
\par }{\PP \VS{4}¶ Thou shalt not
see thy
brother’s
ass or his
ox fall
down by the
way, and
hide thyself from them: thou shalt
surely help him to lift
{\ADD{them}} up
again.
\par }{\PP \VS{5}¶ The
woman shall not wear that which
pertaineth unto a
man, neither shall a
man
put on a
woman’s
garment: for all that
do
so
{\ADD{are}}
abomination unto the
{\ND{
LORD}} thy
God.
\par }{\PP \VS{6}¶ If a
bird’s
nest
chance to be
before thee in the
way in any
tree, or on the
ground,
{\ADD{whether they be}} young
ones, or
eggs, and the
dam
sitting upon the
young, or upon the
eggs, thou shalt not
take the
dam with the
young:
\VS{7}{\ADD{But}} thou shalt in any
wise let the
dam
go, and
take the
young to thee; that it may be
well with thee, and
{\ADD{that}} thou mayest
prolong
{\ADD{thy}}
days.
\par }{\PP \VS{8}¶ When thou
buildest a
new
house, then thou shalt
make a
battlement for thy
roof, that thou
bring not
blood upon thine
house, if any
man
fall from thence.
\par }{\PP \VS{9}¶ Thou shalt not
sow thy
vineyard with divers
seeds: lest the
fruit of thy
seed which thou hast
sown, and the
fruit of thy
vineyard, be
defiled.\FTNT{*}{{\FR 22.9 }
fruit of thy seed: Heb. fulness of the seed}
\par }{\PP \VS{10}¶ Thou shalt not
plow with an
ox and an
ass
together.
\par }{\PP \VS{11}¶ Thou shalt not
wear a garment of divers
sorts,
{\ADD{as}} of
woollen and
linen
together.
\par }{\PP \VS{12}¶ Thou shalt
make thee
fringes upon the
four
quarters of thy
vesture, wherewith thou
coverest
{\ADD{thyself}}.\FTNT{†}{{\FR 22.12 }
quarters: Heb. wings}
\par }{\PP \VS{13}¶ If any
man
take a
wife, and go
in unto her, and
hate her,
\VS{14}And
give
occasions of
speech against her, and bring
up an
evil
name upon her, and
say, I
took this
woman, and when I
came to her, I
found her not a
maid:
\VS{15}Then shall the
father of the
damsel, and her
mother,
take and bring
forth
{\ADD{the tokens of}} the
damsel’s
virginity unto the
elders of the
city in the
gate:
\VS{16}And the
damsel’s
father shall
say unto the
elders, I
gave my
daughter unto this
man to
wife, and he
hateth her;
\VS{17}And, lo, he hath
given
occasions of
speech
{\ADD{against her}},
saying, I
found not thy
daughter a
maid; and yet these
{\ADD{are the tokens of}} my
daughter’s
virginity. And they shall
spread the
cloth
before the
elders of the
city.
\VS{18}And the
elders of that
city shall
take that
man and
chastise him;
\VS{19}And they shall
amerce him in an
hundred
{\ADD{shekels}} of
silver, and
give
{\ADD{them}} unto the
father of the
damsel, because he hath brought
up an
evil
name upon a
virgin of
Israel: and she shall be his
wife; he
may not put her
away all his
days.
\VS{20}But if this
thing be
true,
{\ADD{and the tokens of}}
virginity be not
found for the
damsel:
\VS{21}Then they shall bring
out the
damsel to the
door of her
father’s
house, and the
men of her
city shall
stone her with
stones that she
die: because she hath
wrought
folly in
Israel, to play the
whore in her
father’s
house: so shalt thou
put
evil
away from
among you.
\par }{\PP \VS{22}¶ If a
man be
found
lying with a
woman
married to an
husband, then they shall both of
them
die,
{\ADD{both}} the
man that
lay with the
woman, and the
woman: so shalt thou put
away
evil from
Israel.
\par }{\PP \VS{23}¶ If a
damsel
{\ADD{that is}} a
virgin be
betrothed unto an
husband, and a
man
find her in the
city, and
lie with her;
\VS{24}Then ye shall
bring them
both out unto the
gate of that
city, and ye shall
stone them with
stones that they
die; the
damsel,
because she
cried not,
{\ADD{being}} in the
city; and the
man,
because he hath
humbled his
neighbour’s
wife: so thou shalt put
away
evil from
among you.
\par }{\PP \VS{25}¶ But if a
man
find a
betrothed
damsel in the
field, and the
man
force her, and
lie with her: then the
man only that
lay with her shall
die:\FTNT{‡}{{\FR 22.25 }
force: or, take strong hold of}
\VS{26}But unto the
damsel thou shalt
do
nothing;
{\ADD{there is}} in the
damsel no
sin
{\ADD{worthy}} of
death: for as when a
man
riseth against his
neighbour, and
slayeth him, even so
{\ADD{is}} this
matter:
\VS{27}For he
found her in the
field,
{\ADD{and}} the
betrothed
damsel
cried, and
{\ADD{there was}} none to
save her.
\par }{\PP \VS{28}¶ If a
man
find a
damsel
{\ADD{that is}} a
virgin, which is not
betrothed, and lay
hold on her, and
lie with her, and they be
found;
\VS{29}Then the
man that
lay with her shall
give unto the
damsel’s
father
fifty
{\ADD{shekels}} of
silver, and she shall be his
wife; because he hath
humbled her, he
may not put her
away all his
days.
\par }{\PP \VS{30}¶ A
man shall not
take his
father’s
wife, nor
discover his
father’s
skirt.

\par }\Chap{23}{\PP \VerseOne{1}He that is wounded in the
stones, or hath his privy
member cut
off, shall not
enter into the
congregation of the
{\ND{
LORD}}.
\VS{2}A
bastard shall not
enter into the
congregation of the
{\ND{
LORD}}; even to his
tenth
generation shall he not
enter into the
congregation of the
{\ND{
LORD}}.
\VS{3}An
Ammonite or
Moabite shall not
enter into the
congregation of the
{\ND{
LORD}}; even to their
tenth
generation shall they not
enter into the
congregation of the
{\ND{
LORD}}
for
ever:
\VS{4}Because they
met you not with
bread and with
water in the
way, when ye came
forth out of
Egypt; and because they
hired against thee
Balaam the
son of
Beor of
Pethor of
Mesopotamia, to
curse thee.
\VS{5}Nevertheless the
{\ND{
LORD}} thy
God
would not
hearken unto
Balaam; but the
{\ND{
LORD}} thy
God
turned the
curse into a
blessing unto thee, because the
{\ND{
LORD}} thy
God
loved thee.
\VS{6}Thou shalt not
seek their
peace nor their
prosperity all thy
days for
ever.\FTNT{*}{{\FR 23.6 }
prosperity: Heb. good}
\par }{\PP \VS{7}¶ Thou shalt not
abhor an
Edomite; for he
{\ADD{is}} thy
brother: thou shalt not
abhor an
Egyptian; because thou wast a
stranger in his
land.
\VS{8}The
children that are
begotten of them shall
enter into the
congregation of the
{\ND{
LORD}} in their
third
generation.
\par }{\PP \VS{9}¶ When the
host goeth
forth against thine
enemies, then
keep thee from every
wicked
thing.
\par }{\PP \VS{10}¶ If there be among you any
man, that is not
clean by reason of uncleanness that
chanceth him by
night, then shall he
go
abroad out of the
camp, he shall not
come
within the
camp:
\VS{11}But it shall be, when
evening cometh
on, he shall
wash
{\ADD{himself}} with
water: and when the
sun is
down, he shall
come
into the
camp
{\ADD{again}}.\FTNT{†}{{\FR 23.11 }
cometh…: Heb. turneth toward}
\par }{\PP \VS{12}¶ Thou shalt have a place
also
without the
camp, whither thou shalt go
forth
abroad:
\VS{13}And thou shalt have a
paddle upon thy
weapon; and it shall be, when thou wilt
ease thyself
abroad, thou shalt
dig therewith, and shalt turn
back and
cover that which
cometh from thee:\FTNT{‡}{{\FR 23.13 }
wilt…: Heb. sittest down}
\VS{14}For the
{\ND{
LORD}} thy
God
walketh in the
midst of thy
camp, to
deliver thee, and to give
up thine
enemies
before thee; therefore shall thy
camp be
holy: that he
see no
unclean
thing in thee, and turn
away from
thee.\FTNT{§}{{\FR 23.14 }
unclean…: Heb. nakedness of any thing}
\par }{\PP \VS{15}¶ Thou shalt not
deliver unto his
master the
servant which is
escaped from his
master unto thee:
\VS{16}He shall
dwell with thee,
{\ADD{even}}
among you, in that
place which he shall
choose in
one of thy
gates, where it liketh him
best: thou shalt not
oppress him.\FTNT{**}{{\FR 23.16 }
liketh…: Heb. is good for him}
\par }{\PP \VS{17}¶ There shall be no
whore of the
daughters of
Israel, nor a
sodomite of the
sons of
Israel.\FTNT{††}{{\FR 23.17 }
whore: or, sodomitess}
\VS{18}Thou shalt not
bring the
hire of a
whore, or the
price of a
dog, into the
house of the
{\ND{
LORD}} thy
God for any
vow: for even
both these
{\ADD{are}}
abomination unto the
{\ND{
LORD}} thy
God.
\par }{\PP \VS{19}¶ Thou shalt not lend upon
usury to thy
brother;
usury of
money,
usury of
victuals,
usury of any
thing that is lent upon
usury:
\VS{20}Unto a
stranger thou mayest lend upon
usury; but unto thy
brother thou shalt not lend upon
usury: that the
{\ND{
LORD}} thy
God may
bless thee in all that thou
settest thine
hand to in the
land whither thou
goest to
possess it.
\par }{\PP \VS{21}¶ When thou shalt
vow a
vow unto the
{\ND{
LORD}} thy
God, thou shalt not
slack to
pay it: for the
{\ND{
LORD}} thy
God will
surely
require it of thee; and it would be
sin in thee.
\VS{22}But if thou shalt
forbear to
vow, it shall be no
sin in thee.
\VS{23}That which is gone
out of thy
lips thou shalt
keep and
perform;
{\ADD{even}} a freewill
offering, according as thou hast
vowed unto the
{\ND{
LORD}} thy
God, which thou hast
promised with thy
mouth.
\par }{\PP \VS{24}¶ When thou
comest into thy
neighbour’s
vineyard, then thou mayest
eat
grapes thy
fill at thine own
pleasure; but thou shalt not
put
{\ADD{any}} in thy
vessel.
\VS{25}When thou
comest into the standing
corn of thy
neighbour, then thou mayest
pluck the
ears with thine
hand; but thou shalt not
move a
sickle unto thy
neighbour’s standing
corn.

\par }\Chap{24}{\PP \VerseOne{1}When a
man hath
taken a
wife, and
married her, and it come to pass that she
find no
favour in his
eyes, because he hath
found
some
uncleanness in her: then let him
write her a
bill of
divorcement, and
give
{\ADD{it}} in her
hand, and
send her out of his
house.\FTNT{*}{{\FR 24.1 }
some…: Heb. matter of nakedness}\FTNT{†}{{\FR 24.1 }
divorcement: Heb. cutting off}
\VS{2}And when she is departed
out of his
house, she may
go and be
another
man’s
{\ADD{wife}}.
\VS{3}And
{\ADD{if}} the
latter
husband
hate her, and
write her a
bill of
divorcement, and
giveth
{\ADD{it}} in her
hand, and
sendeth her out of his
house; or if the
latter
husband
die, which
took her
{\ADD{to be}} his
wife;
\VS{4}Her
former
husband, which sent her
away,
may not
take her
again to be his
wife,
after that she is
defiled; for that
{\ADD{is}}
abomination
before the
{\ND{
LORD}}: and thou shalt not cause the
land to
sin, which the
{\ND{
LORD}} thy
God
giveth thee
{\ADD{for}} an
inheritance.
\par }{\PP \VS{5}¶ When a
man hath
taken a
new
wife, he shall not go
out to
war, neither shall he be
charged with any
business:
{\ADD{but}} he shall be
free at
home
one
year, and shall cheer
up his
wife which he hath
taken.\FTNT{‡}{{\FR 24.5 }
neither…: Heb. not any thing shall pass upon him}
\par }{\PP \VS{6}¶ No man shall
take the
nether or the upper
millstone to
pledge: for he
taketh
{\ADD{a man’s}}
life to
pledge.
\par }{\PP \VS{7}¶ If a
man be
found
stealing
any of his
brethren of the
children of
Israel, and maketh
merchandise of him, or
selleth him; then that
thief shall
die; and thou shalt
put
evil
away from
among you.
\par }{\PP \VS{8}¶ Take
heed in the
plague of
leprosy, that thou
observe
diligently, and
do according to all that the
priests the
Levites shall
teach you: as I
commanded them,
{\ADD{so}} ye shall
observe to
do.
\VS{9}Remember what the
{\ND{
LORD}} thy
God
did unto
Miriam by the
way, after that ye were come
forth out of
Egypt.
\par }{\PP \VS{10}¶ When thou dost
lend thy
brother
any
thing, thou shalt not
go into his
house to
fetch his
pledge.\FTNT{§}{{\FR 24.10 }
lend…: Heb. lend the loan of any thing to, etc}
\VS{11}Thou shalt
stand
abroad, and the
man to whom thou dost
lend shall bring
out the
pledge
abroad unto thee.
\VS{12}And if the
man
{\ADD{be}}
poor, thou shalt not
sleep with his
pledge:
\VS{13}In any
case thou shalt
deliver him the
pledge
again when the
sun goeth
down, that he may
sleep in his own
raiment, and
bless thee: and it shall be
righteousness unto thee
before the
{\ND{
LORD}} thy
God.
\par }{\PP \VS{14}¶ Thou shalt not
oppress an hired
servant
{\ADD{that is}}
poor and
needy,
{\ADD{whether he be}} of thy
brethren, or of thy
strangers that
{\ADD{are}} in thy
land within thy
gates:
\VS{15}At his
day thou shalt
give
{\ADD{him}} his
hire, neither shall the
sun go
down upon it; for he
{\ADD{is}}
poor, and
setteth his
heart upon it: lest he
cry against thee unto the
{\ND{
LORD}}, and it be
sin unto thee.\FTNT{**}{{\FR 24.15 }
setteth…: Heb. lifteth his soul unto it}
\VS{16}The
fathers shall not be put to
death for the
children, neither shall the
children be put to
death for the
fathers: every
man shall be put to
death for his own
sin.
\par }{\PP \VS{17}¶ Thou shalt not
pervert the
judgment of the
stranger,
{\ADD{nor}} of the
fatherless; nor
take a
widow’s
raiment to
pledge:
\VS{18}But thou shalt
remember that thou wast a
bondman in
Egypt, and the
{\ND{
LORD}} thy
God
redeemed thee thence: therefore I
command thee to
do this
thing.
\par }{\PP \VS{19}¶ When thou cuttest
down thine
harvest in thy
field, and hast
forgot a
sheaf in the
field, thou shalt not go
again to
fetch it: it shall be for the
stranger, for the
fatherless, and for the
widow: that the
{\ND{
LORD}} thy
God may
bless thee in all the
work of thine
hands.
\VS{20}When thou
beatest thine olive
tree, thou shalt not go over the
boughs
again: it shall be for the
stranger, for the
fatherless, and for the
widow.\FTNT{††}{{\FR 24.20 }
go…: Heb. bough it after thee}
\VS{21}When thou
gatherest the grapes of thy
vineyard, thou shalt not
glean
{\ADD{it}}
afterward: it shall be for the
stranger, for the
fatherless, and for the
widow.\FTNT{‡‡}{{\FR 24.21 }
afterward: Heb. after thee}
\VS{22}And thou shalt
remember that thou wast a
bondman in the
land of
Egypt: therefore I
command thee to
do this
thing.

\par }\Chap{25}{\PP \VerseOne{1}If there be a
controversy between
men, and they
come unto
judgment, that
{\ADD{the judges}} may
judge them; then they shall
justify the
righteous, and
condemn the
wicked.
\VS{2}And it shall be, if the wicked
man
{\ADD{be}}
worthy to be
beaten, that the
judge shall cause him to lie
down, and to be
beaten before his
face,
according to his
fault, by a certain
number.
\VS{3}Forty
stripes he may give him,
{\ADD{and}} not
exceed: lest,
{\ADD{if}} he should
exceed, and
beat him above these with
many
stripes, then thy
brother should seem
vile unto thee.
\par }{\PP \VS{4}¶ Thou shalt not
muzzle the
ox when he treadeth
out
{\ADD{the corn}}.\FTNT{*}{{\FR 25.4 }
treadeth…: Heb. thresheth}
\par }{\PP \VS{5}¶ If
brethren
dwell
together, and
one of them
die, and have no
child, the
wife of the
dead shall not marry
without unto a
stranger: her
husband’s brother shall go
in unto her, and
take her to him to
wife, and perform the duty of an
husband’s brother unto her.\FTNT{†}{{\FR 25.5 }
her husband’s…: or, her next kinsman}
\VS{6}And it shall be,
{\ADD{that}} the
firstborn which she
beareth shall
succeed in the
name of his
brother
{\ADD{which is}}
dead, that his
name be not put
out of
Israel.
\VS{7}And if the
man
like not to
take his
brother’s wife, then let his
brother’s wife go
up to the
gate unto the
elders, and
say, My
husband’s brother
refuseth to raise
up unto his
brother a
name in
Israel, he
will not perform the duty of my
husband’s brother.\FTNT{‡}{{\FR 25.7 }
brother’s: or, next kinsman’s}
\VS{8}Then the
elders of his
city shall
call him, and
speak unto him: and
{\ADD{if}} he
stand
{\ADD{to it}}, and
say, I
like not to
take her;
\VS{9}Then shall his
brother’s wife
come unto him in the
presence of the
elders, and
loose his
shoe from off his
foot, and
spit in his
face, and shall
answer and
say, So shall it be
done unto that
man that will not build
up his
brother’s
house.
\VS{10}And his
name shall be
called in
Israel, The
house of him that hath his
shoe
loosed.
\par }{\PP \VS{11}¶ When
men
strive
together one with
another, and the
wife of the
one draweth
near for to
deliver her
husband out of the
hand of him that
smiteth him, and putteth
forth her
hand, and
taketh him by the
secrets:
\VS{12}Then thou shalt cut
off her
hand, thine
eye shall not
pity
{\ADD{her}}.
\par }{\PP \VS{13}¶ Thou shalt not have in thy
bag divers
weights, a
great and a
small.\FTNT{§}{{\FR 25.13 }
divers…: Heb. a stone and a stone}
\VS{14}Thou shalt not have in thine
house divers
measures, a
great and a
small.\FTNT{**}{{\FR 25.14 }
divers…: Heb. an ephah and an ephah}
\VS{15}{\ADD{But}} thou shalt have a
perfect and
just
weight, a
perfect and
just
measure shalt thou have: that thy
days may be
lengthened in the
land which the
{\ND{
LORD}} thy
God
giveth thee.
\VS{16}For all that
do such
things,
{\ADD{and}} all that
do
unrighteously,
{\ADD{are}} an
abomination unto the
{\ND{
LORD}} thy
God.
\par }{\PP \VS{17}¶
Remember what
Amalek
did unto thee by the
way, when ye were come
forth out of
Egypt;
\VS{18}How he
met thee by the
way, and smote the
hindmost of thee,
{\ADD{even}} all
{\ADD{that were}}
feeble
behind thee, when thou
{\ADD{wast}}
faint and
weary; and he
feared not
God.
\VS{19}Therefore it shall be, when the
{\ND{
LORD}} thy
God hath given thee
rest from all thine
enemies round
about, in the
land which the
{\ND{
LORD}} thy
God
giveth thee
{\ADD{for}} an
inheritance to
possess it,
{\ADD{that}} thou shalt blot
out the
remembrance of
Amalek from under
heaven; thou shalt not
forget
{\ADD{it}}.

\par }\Chap{26}{\PP \VerseOne{1}And it shall be, when thou
{\ADD{art}} come
in unto the
land which the
{\ND{
LORD}} thy
God
giveth thee
{\ADD{for}} an
inheritance, and
possessest it, and
dwellest therein;
\VS{2}That thou shalt
take of the
first of all the
fruit of the
earth, which thou shalt
bring of thy
land that the
{\ND{
LORD}} thy
God
giveth thee, and shalt
put
{\ADD{it}} in a
basket, and shalt
go unto the
place which the
{\ND{
LORD}} thy
God shall
choose to
place his
name there.
\VS{3}And thou shalt
go unto the
priest that shall be in those
days, and
say unto him, I
profess this
day unto the
{\ND{
LORD}} thy
God, that I am
come unto the
country which the
{\ND{
LORD}}
sware unto our
fathers for to
give us.
\VS{4}And the
priest shall
take the
basket out of thine
hand, and set it
down
before the
altar of the
{\ND{
LORD}} thy
God.
\VS{5}And thou shalt
speak and
say
before the
{\ND{
LORD}} thy
God, A
Syrian ready to
perish
{\ADD{was}} my
father, and he went
down into
Egypt, and
sojourned there
with a
few, and became there a
nation,
great,
mighty, and
populous:
\VS{6}And the
Egyptians evil
entreated us, and
afflicted us, and
laid upon us
hard
bondage:
\VS{7}And when we
cried unto the
{\ND{
LORD}}
God of our
fathers, the
{\ND{
LORD}}
heard our
voice, and
looked on our
affliction, and our
labour, and our
oppression:
\VS{8}And the
{\ND{
LORD}} brought us
forth out of
Egypt with a
mighty
hand, and with an
outstretched
arm, and with
great
terribleness, and with
signs, and with
wonders:
\VS{9}And he hath
brought us into this
place, and hath
given us this
land,
{\ADD{even}} a
land that
floweth with
milk and
honey.
\VS{10}And now, behold, I have
brought the
firstfruits of the
land, which thou, O
{\ND{
LORD}}, hast
given me. And thou shalt
set it
before the
{\ND{
LORD}} thy
God, and
worship
before the
{\ND{
LORD}} thy
God:
\VS{11}And thou shalt
rejoice in every
good
{\ADD{thing}} which the
{\ND{
LORD}} thy
God hath
given unto thee, and unto thine
house, thou, and the
Levite, and the
stranger that
{\ADD{is}}
among you.
\par }{\PP \VS{12}¶ When thou hast made an
end of
tithing all the
tithes of thine
increase the
third
year,
{\ADD{which is}} the
year of
tithing, and hast
given
{\ADD{it}} unto the
Levite, the
stranger, the
fatherless, and the
widow, that they may
eat within thy
gates, and be
filled;
\VS{13}Then thou shalt
say
before the
{\ND{
LORD}} thy
God, I have brought
away the hallowed
things out of
{\ADD{mine}}
house, and also have
given them unto the
Levite, and unto the
stranger, to the
fatherless, and to the
widow, according to all thy
commandments which thou hast
commanded me: I have not
transgressed thy
commandments, neither have I
forgotten
{\ADD{them:}}
\VS{14}I have not
eaten thereof in my
mourning, neither have I taken
away
{\ADD{ought}} thereof for
{\ADD{any}}
unclean
{\ADD{use}}, nor
given
{\ADD{ought}} thereof for the
dead:
{\ADD{but}} I have
hearkened to the
voice of the
{\ND{
LORD}} my
God,
{\ADD{and}} have
done according to all that thou hast
commanded me.
\VS{15}Look
down from thy
holy
habitation, from
heaven, and
bless thy
people
Israel, and the
land which thou hast
given us, as thou
swarest unto our
fathers, a
land that
floweth with
milk and
honey.
\par }{\PP \VS{16}¶ This
day the
{\ND{
LORD}} thy
God hath
commanded thee to
do these
statutes and
judgments: thou shalt therefore
keep and
do them with all thine
heart, and with all thy
soul.
\VS{17}Thou hast
avouched the
{\ND{
LORD}} this
day to be thy
God, and to
walk in his
ways, and to
keep his
statutes, and his
commandments, and his
judgments, and to
hearken unto his
voice:
\VS{18}And the
{\ND{
LORD}} hath
avouched thee this
day to be his
peculiar
people, as he hath
promised thee, and that
{\ADD{thou}} shouldest
keep all his
commandments;
\VS{19}And to
make thee
high above all
nations which he hath
made, in
praise, and in
name, and in
honour; and that thou mayest be an
holy
people unto the
{\ND{
LORD}} thy
God, as he hath
spoken.

\par }\Chap{27}{\PP \VerseOne{1}And
Moses with the
elders of
Israel
commanded the
people,
saying,
Keep all the
commandments which I
command you this
day.
\VS{2}And it shall be on the
day when ye shall pass
over
Jordan unto the
land which the
{\ND{
LORD}} thy
God
giveth thee, that thou shalt set thee
up
great
stones, and
plaister them with
plaister:
\VS{3}And thou shalt
write upon them all the
words of this
law, when thou art passed
over, that thou mayest go
in unto the
land which the
{\ND{
LORD}} thy
God
giveth thee, a
land that
floweth with
milk and
honey; as the
{\ND{
LORD}}
God of thy
fathers hath
promised thee.
\VS{4}Therefore it shall be when ye be gone
over
Jordan,
{\ADD{that}} ye shall set
up these
stones, which I
command you this
day, in
mount
Ebal, and thou shalt
plaister them with
plaister.
\VS{5}And there shalt thou
build an
altar unto the
{\ND{
LORD}} thy
God, an
altar of
stones: thou shalt not lift
up
{\ADD{any}}
iron
{\ADD{tool}} upon them.
\VS{6}Thou shalt
build the
altar of the
{\ND{
LORD}} thy
God of
whole
stones: and thou shalt
offer burnt
offerings thereon unto the
{\ND{
LORD}} thy
God:
\VS{7}And thou shalt
offer peace
offerings, and shalt
eat there, and
rejoice
before the
{\ND{
LORD}} thy
God.
\VS{8}And thou shalt
write upon the
stones all the
words of this
law
very
plainly.
\par }{\PP \VS{9}¶ And
Moses and the
priests the
Levites
spake unto all
Israel,
saying, Take
heed, and
hearken, O
Israel; this
day thou art
become the
people of the
{\ND{
LORD}} thy
God.
\VS{10}Thou shalt therefore
obey the
voice of the
{\ND{
LORD}} thy
God, and
do his
commandments and his
statutes, which I
command thee this
day.
\par }{\PP \VS{11}¶ And
Moses
charged the
people the same
day,
saying,
\VS{12}These shall
stand upon
mount
Gerizim to
bless the
people, when ye are come
over
Jordan;
Simeon, and
Levi, and
Judah, and
Issachar, and
Joseph, and
Benjamin:
\VS{13}And these shall
stand upon
mount
Ebal to
curse;
Reuben,
Gad, and
Asher, and
Zebulun,
Dan, and
Naphtali.\FTNT{*}{{\FR 27.13 }
to curse: Heb. for a cursing}
\par }{\PP \VS{14}¶ And the
Levites shall
speak, and
say unto all the
men of
Israel with a
loud
voice,
\VS{15}Cursed
{\ADD{be}} the
man that
maketh
{\ADD{any}}
graven or molten
image, an
abomination unto the
{\ND{
LORD}}, the
work of the
hands of the
craftsman, and
putteth
{\ADD{it}} in
{\ADD{a}}
secret
{\ADD{place}}. And all the
people shall
answer and
say,
Amen.
\VS{16}Cursed
{\ADD{be}} he that setteth
light by his
father or his
mother. And all the
people shall
say,
Amen.
\VS{17}Cursed
{\ADD{be}} he that
removeth his
neighbour’s
landmark. And all the
people shall
say,
Amen.
\VS{18}Cursed
{\ADD{be}} he that maketh the
blind to
wander out of the
way. And all the
people shall
say,
Amen.
\VS{19}Cursed
{\ADD{be}} he that
perverteth the
judgment of the
stranger,
fatherless, and
widow. And all the
people shall
say,
Amen.
\VS{20}Cursed
{\ADD{be}} he that
lieth with his
father’s
wife; because he
uncovereth his
father’s
skirt. And all the
people shall
say,
Amen.
\VS{21}Cursed
{\ADD{be}} he that
lieth with any manner of
beast. And all the
people shall
say,
Amen.
\VS{22}Cursed
{\ADD{be}} he that
lieth with his
sister, the
daughter of his
father, or the
daughter of his
mother. And all the
people shall
say,
Amen.
\VS{23}Cursed
{\ADD{be}} he that
lieth with his mother in
law. And all the
people shall
say,
Amen.
\VS{24}Cursed
{\ADD{be}} he that
smiteth his
neighbour
secretly. And all the
people shall
say,
Amen.
\VS{25}Cursed
{\ADD{be}} he that
taketh
reward to
slay an
innocent
person. And all the
people shall
say,
Amen.
\VS{26}Cursed
{\ADD{be}} he that
confirmeth not
{\ADD{all}} the
words of this
law to
do them. And all the
people shall
say,
Amen.

\par }\Chap{28}{\PP \VerseOne{1}And it shall come to pass, if thou shalt
hearken
diligently unto the
voice of the
{\ND{
LORD}} thy
God, to
observe
{\ADD{and}} to
do all his
commandments which I
command thee this
day, that the
{\ND{
LORD}} thy
God will
set thee on
high above all
nations of the
earth:
\VS{2}And all these
blessings shall
come on thee, and
overtake thee, if thou shalt
hearken unto the
voice of the
{\ND{
LORD}} thy
God.
\VS{3}Blessed
{\ADD{shalt}} thou
{\ADD{be}} in the
city, and
blessed
{\ADD{shalt}} thou
{\ADD{be}} in the
field.
\VS{4}Blessed
{\ADD{shall be}} the
fruit of thy
body, and the
fruit of thy
ground, and the
fruit of thy
cattle, the
increase of thy
kine, and the
flocks of thy
sheep.
\VS{5}Blessed
{\ADD{shall be}} thy
basket and thy
store.\FTNT{*}{{\FR 28.5 }
store: or, dough, or, kneadingtroughs}
\VS{6}Blessed
{\ADD{shalt}} thou
{\ADD{be}} when thou comest
in, and
blessed
{\ADD{shalt}} thou
{\ADD{be}} when thou goest
out.
\VS{7}The
{\ND{
LORD}} shall
cause thine
enemies that rise
up against thee to be
smitten before thy
face: they shall come
out against thee
one
way, and
flee
before thee
seven
ways.
\VS{8}The
{\ND{
LORD}} shall
command the
blessing upon thee in thy
storehouses, and in all that thou
settest thine
hand unto; and he shall
bless thee in the
land which the
{\ND{
LORD}} thy
God
giveth thee.\FTNT{†}{{\FR 28.8 }
storehouses: or, barns}
\VS{9}The
{\ND{
LORD}} shall
establish thee an
holy
people unto himself, as he hath
sworn unto thee, if thou shalt
keep the
commandments of the
{\ND{
LORD}} thy
God, and
walk in his
ways.
\VS{10}And all
people of the
earth shall
see that thou art
called by the
name of the
{\ND{
LORD}}; and they shall be
afraid of thee.
\VS{11}And the
{\ND{
LORD}} shall make thee
plenteous in
goods, in the
fruit of thy
body, and in the
fruit of thy
cattle, and in the
fruit of thy
ground, in the
land which the
{\ND{
LORD}}
sware unto thy
fathers to
give thee.\FTNT{‡}{{\FR 28.11 }
in goods: or, for good}\FTNT{§}{{\FR 28.11 }
body: Heb. belly}
\VS{12}The
{\ND{
LORD}} shall
open unto thee his
good
treasure, the
heaven to
give the
rain unto thy
land in his
season, and to
bless all the
work of thine
hand: and thou shalt
lend unto
many
nations, and thou shalt not
borrow.
\VS{13}And the
{\ND{
LORD}} shall
make thee the
head, and not the
tail; and thou shalt be above
only, and thou shalt not be
beneath; if that thou
hearken unto the
commandments of the
{\ND{
LORD}} thy
God, which I
command thee this
day, to
observe and to
do
{\ADD{them:}}
\VS{14}And thou shalt not go
aside from any of the
words which I
command thee this
day,
{\ADD{to}} the right
hand, or
{\ADD{to}} the
left, to
go
after
other
gods to
serve them.
\par }{\PP \VS{15}¶ But it shall come to pass, if thou wilt not
hearken unto the
voice of the
{\ND{
LORD}} thy
God, to
observe to
do all his
commandments and his
statutes which I
command thee this
day; that all these
curses shall
come upon thee, and
overtake thee:
\VS{16}Cursed
{\ADD{shalt}} thou
{\ADD{be}} in the
city, and
cursed
{\ADD{shalt}} thou
{\ADD{be}} in the
field.
\VS{17}Cursed
{\ADD{shall be}} thy
basket and thy
store.
\VS{18}Cursed
{\ADD{shall be}} the
fruit of thy
body, and the
fruit of thy
land, the
increase of thy
kine, and the
flocks of thy
sheep.
\VS{19}Cursed
{\ADD{shalt}} thou
{\ADD{be}} when thou comest
in, and
cursed
{\ADD{shalt}} thou
{\ADD{be}} when thou goest
out.
\VS{20}The
{\ND{
LORD}} shall
send upon thee
cursing,
vexation, and
rebuke, in all that thou
settest thine
hand unto for to
do, until thou be
destroyed, and until thou
perish
quickly;
because of the
wickedness of thy
doings, whereby thou hast
forsaken me.\FTNT{**}{{\FR 28.20 }
for…: Heb. which thou wouldest do}
\VS{21}The
{\ND{
LORD}} shall make the
pestilence
cleave unto thee, until he have
consumed thee from off the
land, whither thou
goest to
possess it.
\VS{22}The
{\ND{
LORD}} shall
smite thee with a
consumption, and with a
fever, and with an
inflammation, and with an extreme
burning, and with the
sword, and with
blasting, and with
mildew; and they shall
pursue thee until thou
perish.\FTNT{††}{{\FR 28.22 }
sword: or, drought}
\VS{23}And thy
heaven that
{\ADD{is}} over thy
head shall be
brass, and the
earth that is under thee
{\ADD{shall be}}
iron.
\VS{24}The
{\ND{
LORD}} shall
make the
rain of thy
land
powder and
dust: from
heaven shall it come
down upon thee, until thou be
destroyed.
\VS{25}The
{\ND{
LORD}} shall
cause thee to be
smitten
before thine
enemies: thou shalt go
out
one
way against them, and
flee
seven
ways
before them: and shalt be
removed into all the
kingdoms of the
earth.\FTNT{‡‡}{{\FR 28.25 }
removed: Heb. for a removing}
\VS{26}And thy
carcase shall be
meat unto all
fowls of the
air, and unto the
beasts of the
earth, and no man shall fray
{\ADD{them}}
away.
\VS{27}The
{\ND{
LORD}} will
smite thee with the
botch of
Egypt, and with the
emerods, and with the
scab, and with the
itch, whereof thou
canst not be
healed.
\VS{28}The
{\ND{
LORD}} shall
smite thee with
madness, and
blindness, and
astonishment of
heart:
\VS{29}And thou shalt
grope at
noonday, as the
blind
gropeth in
darkness, and thou shalt not
prosper in thy
ways: and thou shalt be only
oppressed and
spoiled
evermore, and no man shall
save
{\ADD{thee}}.
\VS{30}Thou shalt
betroth a
wife, and
another
man shall
lie with her: thou shalt
build an
house, and thou shalt not
dwell therein: thou shalt
plant a
vineyard, and shalt not gather the
grapes thereof.\FTNT{§§}{{\FR 28.30 }
gather…: Heb. profane, or, use it as common meat}
\VS{31}Thine
ox
{\ADD{shall be}}
slain before thine
eyes, and thou shalt not
eat thereof: thine
ass
{\ADD{shall be}} violently taken
away from before thy
face, and shall not be
restored to thee: thy
sheep
{\ADD{shall be}}
given unto thine
enemies, and thou shalt have none to
rescue
{\ADD{them}}.\FTNT{***}{{\FR 28.31 }
shall not…: Heb. shall not return to thee}
\VS{32}Thy
sons and thy
daughters
{\ADD{shall be}}
given unto
another
people, and thine
eyes shall
look, and
fail
{\ADD{with longing}} for them all the
day long: and
{\ADD{there shall be}} no
might in thine
hand.
\VS{33}The
fruit of thy
land, and all thy
labours, shall a
nation which thou
knowest not eat
up; and thou shalt be only
oppressed and
crushed
alway:
\VS{34}So that thou shalt be
mad for the
sight of thine
eyes which thou shalt
see.
\VS{35}The
{\ND{
LORD}} shall
smite thee in the
knees, and in the
legs, with a
sore
botch that
cannot be
healed, from the
sole of thy
foot unto the top of thy
head.
\VS{36}The
{\ND{
LORD}} shall
bring thee, and thy
king which thou shalt
set over thee, unto a
nation which neither thou nor thy
fathers have
known; and there shalt thou
serve
other
gods,
wood and
stone.
\VS{37}And thou shalt become an
astonishment, a
proverb, and a
byword, among all
nations whither the
{\ND{
LORD}} shall
lead thee.
\VS{38}Thou shalt
carry
much
seed
out into the
field, and shalt
gather
{\ADD{but}}
little
in; for the
locust shall
consume it.
\VS{39}Thou shalt
plant
vineyards, and
dress
{\ADD{them}}, but shalt neither
drink
{\ADD{of}} the
wine, nor
gather
{\ADD{the grapes;}} for the
worms shall
eat them.
\VS{40}Thou shalt have olive
trees throughout all thy
coasts, but thou shalt not
anoint
{\ADD{thyself}} with the
oil; for thine
olive shall
cast
{\ADD{his fruit}}.
\VS{41}Thou shalt
beget
sons and
daughters, but thou shalt not enjoy them; for they shall
go into
captivity.\FTNT{†††}{{\FR 28.41 }
thou shalt not…: Heb. they shall not be thine}
\VS{42}All thy
trees and
fruit of thy
land shall the
locust
consume.\FTNT{‡‡‡}{{\FR 28.42 }
consume: or, possess}
\VS{43}The
stranger that
{\ADD{is}}
within thee shall get
up above thee
very
high; and thou shalt come
down
very
low.
\VS{44}He shall
lend to thee, and thou shalt not
lend to him: he shall be the
head, and thou shalt be the
tail.
\par }{\PP \VS{45}Moreover all these
curses shall
come upon thee, and shall
pursue thee, and
overtake thee, till thou be
destroyed; because thou
hearkenedst not unto the
voice of the
{\ND{
LORD}} thy
God, to
keep his
commandments and his
statutes which he
commanded thee:
\VS{46}And they shall be upon thee for a
sign and for a
wonder, and upon thy
seed
for
ever.
\VS{47}Because thou
servedst not the
{\ND{
LORD}} thy
God with
joyfulness, and with
gladness of
heart, for the
abundance of all
{\ADD{things;}}
\VS{48}Therefore shalt thou
serve thine
enemies which the
{\ND{
LORD}} shall
send against thee, in
hunger, and in
thirst, and in
nakedness, and in
want of all
{\ADD{things:}} and he shall
put a
yoke of
iron upon thy
neck, until he have
destroyed thee.
\VS{49}The
{\ND{
LORD}} shall
bring a
nation against thee from
far, from the
end of the
earth,
{\ADD{as swift}} as the
eagle
flieth; a
nation whose
tongue thou shalt not
understand;\FTNT{§§§}{{\FR 28.49 }
understand: Heb. hear}
\VS{50}A
nation of
fierce
countenance, which shall not
regard the
person of the
old, nor shew
favour to the
young:\FTNT{*}{{\FR 28.50 }
of fierce…: Heb. strong of face}
\VS{51}And he shall
eat the
fruit of thy
cattle, and the
fruit of thy
land, until thou be
destroyed: which
{\ADD{also}} shall not
leave thee
{\ADD{either}}
corn,
wine, or
oil,
{\ADD{or}} the
increase of thy
kine, or
flocks of thy
sheep, until he have
destroyed thee.
\VS{52}And he shall
besiege thee in all thy
gates, until thy
high and
fenced
walls come
down,
wherein thou
trustedst, throughout all thy
land: and he shall
besiege thee in all thy
gates throughout all thy
land, which the
{\ND{
LORD}} thy
God hath
given thee.
\VS{53}And thou shalt
eat the
fruit of thine own
body, the
flesh of thy
sons and of thy
daughters, which the
{\ND{
LORD}} thy
God hath
given thee, in the
siege, and in the
straitness, wherewith thine
enemies shall
distress thee:\FTNT{†}{{\FR 28.53 }
body: Heb. belly}
\VS{54}{\ADD{So that}} the
man
{\ADD{that is}}
tender among you, and
very
delicate, his
eye shall be
evil toward his
brother, and toward the
wife of his
bosom, and toward the
remnant of his
children which he shall
leave:
\VS{55}So that he will not
give to
any of them of the
flesh of his
children whom he shall
eat: because he hath nothing
left him in the
siege, and in the
straitness, wherewith thine
enemies shall
distress thee in all thy
gates.
\VS{56}The
tender and
delicate woman among you, which would not
adventure to
set the
sole of her
foot upon the
ground for
delicateness and
tenderness, her
eye shall be
evil toward the
husband of her
bosom, and toward her
son, and toward her
daughter,
\VS{57}And toward her young
one that cometh
out from between her
feet, and toward her
children which she shall
bear: for she shall
eat them for
want of all
{\ADD{things}}
secretly in the
siege and
straitness, wherewith thine
enemy shall
distress thee in thy
gates.\FTNT{‡}{{\FR 28.57 }
young one: Heb. afterbirth}
\VS{58}If thou wilt not
observe to
do all the
words of this
law that are
written in this
book, that thou mayest
fear this
glorious and
fearful
name, THE
{\ND{
LORD}} THY
GOD;
\VS{59}Then the
{\ND{
LORD}} will
make thy
plagues
wonderful, and the
plagues of thy
seed,
{\ADD{even}}
great
plagues, and of long
continuance, and
sore
sicknesses, and of long
continuance.
\VS{60}Moreover he will
bring upon thee all the
diseases of
Egypt, which thou wast
afraid
of; and they shall
cleave unto thee.
\VS{61}Also every
sickness, and every
plague, which
{\ADD{is}} not
written in the
book of this
law, them will the
{\ND{
LORD}}
bring upon thee, until thou be
destroyed.\FTNT{§}{{\FR 28.61 }
bring: Heb. cause to ascend}
\VS{62}And ye shall be
left
few in
number,
whereas ye were as the
stars of
heaven for
multitude; because thou wouldest not
obey the
voice of the
{\ND{
LORD}} thy
God.
\VS{63}And it shall come to pass,
{\ADD{that}} as the
{\ND{
LORD}}
rejoiced over you to do you
good, and to
multiply you; so the
{\ND{
LORD}} will
rejoice over you to
destroy you, and to bring you to
nought; and ye shall be
plucked from off the
land whither thou
goest to
possess it.
\VS{64}And the
{\ND{
LORD}} shall
scatter thee among all
people, from the one
end of the
earth even unto the
other; and there thou shalt
serve
other
gods, which neither thou nor thy
fathers have
known,
{\ADD{even}}
wood and
stone.
\VS{65}And among
these
nations shalt thou find no
ease, neither shall the
sole of thy
foot have
rest: but the
{\ND{
LORD}} shall
give thee there a
trembling
heart, and
failing of
eyes, and
sorrow of
mind:
\VS{66}And thy
life shall
hang in doubt
before thee; and thou shalt
fear
day and
night, and shalt have none
assurance of thy
life:
\VS{67}In the
morning thou shalt
say, Would God it
were
even! and at
even thou shalt
say, Would God it
were
morning! for the
fear of thine
heart wherewith thou shalt
fear, and for the
sight of thine
eyes which thou shalt
see.
\VS{68}And the
{\ND{
LORD}} shall
bring thee into
Egypt
again with
ships, by the
way whereof I
spake unto thee, Thou shalt
see it no more
again: and there ye shall be
sold unto your
enemies for
bondmen and
bondwomen, and no man shall
buy
{\ADD{you}}.

\par }\Chap{29}{\PP \VerseOne{1}These
{\ADD{are}} the
words of the
covenant, which the
{\ND{
LORD}}
commanded
Moses to
make with the
children of
Israel in the
land of
Moab, beside the
covenant which he
made with them in
Horeb.
\par }{\PP \VS{2}¶ And
Moses
called unto all
Israel, and
said unto them, Ye have
seen all that the
{\ND{
LORD}}
did before your
eyes in the
land of
Egypt unto
Pharaoh, and unto all his
servants, and unto all his
land;
\VS{3}The
great
temptations which thine
eyes have
seen, the
signs, and those
great
miracles:
\VS{4}Yet the
{\ND{
LORD}} hath not
given you an
heart to
perceive, and
eyes to
see, and
ears to
hear, unto this
day.
\VS{5}And I have
led you
forty
years in the
wilderness: your
clothes are not waxen
old upon you, and thy
shoe is not waxen
old upon thy
foot.
\VS{6}Ye have not
eaten
bread, neither have ye
drunk
wine or strong
drink: that ye might
know that I
{\ADD{am}} the
{\ND{
LORD}} your
God.
\VS{7}And when ye
came unto this
place,
Sihon the
king of
Heshbon, and
Og the
king of
Bashan, came
out
against us unto
battle, and we
smote them:
\VS{8}And we
took their
land, and
gave it for an
inheritance unto the
Reubenites, and to the
Gadites, and to the
half
tribe of
Manasseh.
\VS{9}Keep therefore the
words of this
covenant, and
do them, that ye may
prosper in all that ye
do.
\par }{\PP \VS{10}¶ Ye
stand this
day all of you
before the
{\ND{
LORD}} your
God; your
captains of your
tribes, your
elders, and your
officers,
{\ADD{with}} all the
men of
Israel,
\VS{11}Your little
ones, your
wives, and thy
stranger that
{\ADD{is}}
in thy
camp, from the
hewer of thy
wood unto the
drawer of thy
water:
\VS{12}That thou shouldest
enter into
covenant with the
{\ND{
LORD}} thy
God, and into his
oath, which the
{\ND{
LORD}} thy
God
maketh with thee this
day:\FTNT{*}{{\FR 29.12 }
enter: Heb. pass}
\VS{13}That he may
establish thee to
day for a
people unto himself, and
{\ADD{that}} he may be unto thee a
God, as he hath
said unto thee, and as he hath
sworn unto thy
fathers, to
Abraham, to
Isaac, and to
Jacob.
\VS{14}Neither with you only do I
make this
covenant and this
oath;
\VS{15}But with
{\ADD{him}} that
standeth here with us
this
day
before the
{\ND{
LORD}} our
God, and also with
{\ADD{him}} that
{\ADD{is}} not here with us this
day:
\VS{16}(For ye
know how we have
dwelt in the
land of
Egypt; and how we
came
through the
nations which ye passed
by;
\VS{17}And ye have
seen their
abominations, and their
idols,
wood and
stone,
silver and
gold, which
{\ADD{were}} among them:)\FTNT{†}{{\FR 29.17 }
idols: Heb. dungy gods}
\VS{18}Lest there should
be among you
man, or
woman, or
family, or
tribe, whose
heart turneth
away this
day from the
{\ND{
LORD}} our
God, to
go
{\ADD{and}}
serve the
gods of these
nations; lest there should
be among you a
root that
beareth
gall and
wormwood;\FTNT{‡}{{\FR 29.18 }
gall: or, a poisonous herb: Heb. rosh}
\VS{19}And it come to pass, when he
heareth the
words of this
curse, that he
bless himself in his
heart,
saying, I shall have
peace,
though I
walk in the
imagination of mine
heart, to
add
drunkenness to
thirst:\FTNT{§}{{\FR 29.19 }
imagination: or, stubbornness}\FTNT{**}{{\FR 29.19 }
drunkenness…: Heb. the drunken to the thirsty}
\VS{20}The
{\ND{
LORD}}
will not
spare him, but then the
anger of the
{\ND{
LORD}} and his
jealousy shall
smoke against that
man, and all the
curses that are
written in this
book shall
lie upon him, and the
{\ND{
LORD}} shall blot
out his
name from under
heaven.
\VS{21}And the
{\ND{
LORD}} shall
separate him unto
evil out of all the
tribes of
Israel, according to all the
curses of the
covenant that are
written in this
book of the
law:\FTNT{††}{{\FR 29.21 }
are: Heb. is}
\VS{22}So that the
generation to
come of your
children that shall rise
up
after you, and the
stranger that shall
come from a
far
land, shall
say, when they
see the
plagues of that
land, and the
sicknesses which the
{\ND{
LORD}} hath
laid upon it;\FTNT{‡‡}{{\FR 29.22 }
which…: Heb. wherewith the
{\ND{LORD}} hath made it sick}
\VS{23}{\ADD{And that}} the whole
land thereof
{\ADD{is}}
brimstone, and
salt,
{\ADD{and}}
burning,
{\ADD{that}} it is not
sown, nor
beareth, nor any
grass
groweth therein, like the
overthrow of
Sodom, and
Gomorrah,
Admah, and
Zeboim, which the
{\ND{
LORD}}
overthrew in his
anger, and in his
wrath:
\VS{24}Even all
nations shall
say, Wherefore hath the
{\ND{
LORD}}
done thus unto this
land? what
{\ADD{meaneth}} the
heat of this
great
anger?
\VS{25}Then men shall
say, Because they have
forsaken the
covenant of the
{\ND{
LORD}}
God of their
fathers, which he
made with them when he brought them
forth out of the
land of
Egypt:
\VS{26}For they
went and
served
other
gods, and
worshipped them,
gods whom they
knew not, and
{\ADD{whom}} he had not
given unto them:\FTNT{§§}{{\FR 29.26 }
whom he…: or, who had not given to them any portion}\FTNT{***}{{\FR 29.26 }
given: Heb. divided}
\VS{27}And the
anger of the
{\ND{
LORD}} was
kindled against this
land, to
bring upon it all the
curses that are
written in this
book:
\VS{28}And the
{\ND{
LORD}}
rooted them out of their
land in
anger, and in
wrath, and in
great
indignation, and
cast them into
another
land, as
{\ADD{it is}} this
day.
\VS{29}The
secret
{\ADD{things belong}} unto the
{\ND{
LORD}} our
God: but those
{\ADD{things which are}}
revealed
{\ADD{belong}} unto us and to our
children
for
ever, that
{\ADD{we}} may
do all the
words of this
law.

\par }\Chap{30}{\PP \VerseOne{1}And it shall come to pass, when all these
things are
come upon thee, the
blessing and the
curse, which I have
set
before thee, and thou shalt
call
{\ADD{them}} to
mind among all the
nations, whither the
{\ND{
LORD}} thy
God hath
driven thee,
\VS{2}And shalt
return unto the
{\ND{
LORD}} thy
God, and shalt
obey his
voice according to all that I
command thee this
day, thou and thy
children, with all thine
heart, and with all thy
soul;
\VS{3}That then the
{\ND{
LORD}} thy
God will
turn thy
captivity, and have
compassion upon thee, and will
return and
gather thee from all the
nations, whither the
{\ND{
LORD}} thy
God hath
scattered thee.
\VS{4}If
{\ADD{any}} of thine be driven
out unto the
outmost
{\ADD{parts}} of
heaven, from thence will the
{\ND{
LORD}} thy
God
gather thee, and from thence will he
fetch thee:
\VS{5}And the
{\ND{
LORD}} thy
God will
bring thee into the
land which thy
fathers
possessed, and thou shalt
possess it; and he will do thee
good, and
multiply thee above thy
fathers.
\VS{6}And the
{\ND{
LORD}} thy
God will
circumcise thine
heart, and the
heart of thy
seed, to
love the
{\ND{
LORD}} thy
God with all thine
heart, and with all thy
soul, that thou mayest
live.
\VS{7}And the
{\ND{
LORD}} thy
God will
put all these
curses upon thine
enemies, and on them that
hate thee, which
persecuted thee.
\VS{8}And thou shalt
return and
obey the
voice of the
{\ND{
LORD}}, and
do all his
commandments which I
command thee this
day.
\VS{9}And the
{\ND{
LORD}} thy
God will make thee
plenteous in every
work of thine
hand, in the
fruit of thy
body, and in the
fruit of thy
cattle, and in the
fruit of thy
land, for
good: for the
{\ND{
LORD}} will
again
rejoice over thee for
good, as he
rejoiced over thy
fathers:
\VS{10}If thou shalt
hearken unto the
voice of the
{\ND{
LORD}} thy
God, to
keep his
commandments and his
statutes which are
written in this
book of the
law,
{\ADD{and}} if thou
turn unto the
{\ND{
LORD}} thy
God with all thine
heart, and with all thy
soul.
\par }{\PP \VS{11}¶ For this
commandment which I
command thee this
day, it
{\ADD{is}} not
hidden from thee, neither
{\ADD{is}} it far
off.
\VS{12}It
{\ADD{is}} not in
heaven, that thou shouldest
say, Who shall go
up for us to
heaven, and
bring it unto us, that we may
hear it, and
do it?
\VS{13}Neither
{\ADD{is}} it
beyond the
sea, that thou shouldest
say, Who shall
go
over the
sea for us, and
bring it unto us, that we may
hear it, and
do it?
\VS{14}But the
word
{\ADD{is}}
very
nigh unto thee, in thy
mouth, and in thy
heart, that thou mayest
do it.
\par }{\PP \VS{15}¶
See, I have
set
before thee this
day
life and
good, and
death and
evil;
\VS{16}In that I
command thee this
day to
love the
{\ND{
LORD}} thy
God, to
walk in his
ways, and to
keep his
commandments and his
statutes and his
judgments, that thou mayest
live and
multiply: and the
{\ND{
LORD}} thy
God shall
bless thee in the
land whither thou
goest to
possess it.
\VS{17}But if thine
heart turn
away, so that thou wilt not
hear, but shalt be drawn
away, and
worship
other
gods, and
serve them;
\VS{18}I
denounce unto you this
day, that ye shall
surely
perish,
{\ADD{and that}} ye shall not
prolong
{\ADD{your}}
days upon the
land, whither thou passest
over
Jordan to
go to
possess it.
\VS{19}I
call
heaven and
earth to
record this
day against you,
{\ADD{that}} I have
set
before you
life and
death,
blessing and
cursing: therefore
choose
life, that both thou and thy
seed may
live:
\VS{20}That thou mayest
love the
{\ND{
LORD}} thy
God,
{\ADD{and}} that thou mayest
obey his
voice, and that thou mayest
cleave unto him: for he
{\ADD{is}} thy
life, and the
length of thy
days: that thou mayest
dwell in the
land which the
{\ND{
LORD}}
sware unto thy
fathers, to
Abraham, to
Isaac, and to
Jacob, to
give them.

\par }\Chap{31}{\PP \VerseOne{1}And
Moses
went and
spake these
words unto all
Israel.
\VS{2}And he
said unto them, I
{\ADD{am}} an
hundred and
twenty
years
old this
day; I
can no more go
out and come
in: also the
{\ND{
LORD}} hath
said unto me, Thou shalt not go
over this
Jordan.
\VS{3}The
{\ND{
LORD}} thy
God, he will go
over
before thee,
{\ADD{and}} he will
destroy these
nations from
before thee, and thou shalt
possess them:
{\ADD{and}}
Joshua, he shall go
over
before thee, as the
{\ND{
LORD}} hath
said.
\VS{4}And the
{\ND{
LORD}} shall
do unto them as he
did to
Sihon and to
Og,
kings of the
Amorites, and unto the
land of them, whom he
destroyed.
\VS{5}And the
{\ND{
LORD}} shall give them
up before your
face, that ye may
do unto them according unto all the
commandments which I have
commanded you.
\VS{6}Be
strong and of a good
courage,
fear not, nor be
afraid of
them: for the
{\ND{
LORD}} thy
God, he
{\ADD{it is}} that doth
go with thee; he will not
fail thee, nor
forsake thee.
\par }{\PP \VS{7}¶ And
Moses
called unto
Joshua, and
said unto him in the
sight of all
Israel, Be
strong and of a good
courage: for thou must
go with this
people unto the
land which the
{\ND{
LORD}} hath
sworn unto their
fathers to
give them; and thou shalt cause them to
inherit it.
\VS{8}And the
{\ND{
LORD}}, he
{\ADD{it is}} that doth
go
before thee; he will be with thee, he will not
fail thee, neither
forsake thee:
fear not, neither be
dismayed.
\par }{\PP \VS{9}¶ And
Moses
wrote this
law, and
delivered it unto the
priests the
sons of
Levi, which
bare the
ark of the
covenant of the
{\ND{
LORD}}, and unto all the
elders of
Israel.
\VS{10}And
Moses
commanded them,
saying, At the
end of
{\ADD{every}}
seven
years, in the
solemnity of the
year of
release, in the
feast of
tabernacles,
\VS{11}When all
Israel is
come to
appear
before the
{\ND{
LORD}} thy
God in the
place which he shall
choose, thou shalt
read this
law before all
Israel in their
hearing.
\VS{12}Gather the
people
together,
men, and
women, and
children, and thy
stranger that
{\ADD{is}} within thy
gates, that they may
hear, and that they may
learn, and
fear the
{\ND{
LORD}} your
God, and
observe to
do all the
words of this
law:
\VS{13}And
{\ADD{that}} their
children, which have not
known
{\ADD{any thing}}, may
hear, and
learn to
fear the
{\ND{
LORD}} your
God, as long
as ye
live in the
land whither ye go
over
Jordan to
possess it.
\par }{\PP \VS{14}¶ And the
{\ND{
LORD}}
said unto
Moses, Behold, thy
days
approach that thou must
die:
call
Joshua, and
present yourselves in the
tabernacle of the
congregation, that I may give him a
charge. And
Moses and
Joshua
went, and
presented themselves in the
tabernacle of the
congregation.
\VS{15}And the
{\ND{
LORD}}
appeared in the
tabernacle in a
pillar of a
cloud: and the
pillar of the
cloud
stood over the
door of the
tabernacle.
\par }{\PP \VS{16}¶ And the
{\ND{
LORD}}
said unto
Moses, Behold, thou shalt
sleep with thy
fathers; and this
people will rise
up, and go a
whoring
after the
gods of the
strangers of the
land, whither they
go
{\ADD{to be}}
among them, and will
forsake me, and
break my
covenant which I have
made with them.\FTNT{*}{{\FR 31.16 }
sleep: Heb. lie down}
\VS{17}Then my
anger shall be
kindled against them in that
day, and I will
forsake them, and I will
hide my
face from them, and they shall be
devoured, and
many
evils and
troubles shall
befall them; so that they will
say in that
day, Are not these
evils
come upon us, because our
God
{\ADD{is}} not
among us?\FTNT{†}{{\FR 31.17 }
befall: Heb. find}
\VS{18}And I will
surely
hide my
face in that
day for all the
evils which they shall have
wrought, in
that they are
turned unto
other
gods.
\VS{19}Now therefore
write ye this
song for you, and
teach it the
children of
Israel:
put it in their
mouths, that this
song may be a
witness for me against the
children of
Israel.
\VS{20}For when I shall have
brought them into the
land which I
sware unto their
fathers, that
floweth with
milk and
honey; and they shall have
eaten and
filled themselves, and waxen
fat; then will they
turn unto
other
gods, and
serve them, and
provoke me, and
break my
covenant.
\VS{21}And it shall come to pass, when
many
evils and
troubles are
befallen them, that this
song shall
testify
against them as a
witness; for it shall not be
forgotten out of the
mouths of their
seed: for I
know their
imagination which they go
about, even
now, before I have
brought them into the
land which I
sware.\FTNT{‡}{{\FR 31.21 }
against: Heb. before}\FTNT{§}{{\FR 31.21 }
go…: Heb. do}
\par }{\PP \VS{22}¶
Moses therefore
wrote this
song the same
day, and
taught it the
children of
Israel.
\VS{23}And he gave
Joshua the
son of
Nun a
charge, and
said, Be
strong and of a good
courage: for thou shalt
bring the
children of
Israel into the
land which I
sware unto them: and I will be with thee.
\par }{\PP \VS{24}¶ And it came to pass, when
Moses had made an
end of
writing the
words of this
law in a
book, until they were
finished,
\VS{25}That
Moses
commanded the
Levites, which
bare the
ark of the
covenant of the
{\ND{
LORD}},
saying,
\VS{26}Take this
book of the
law, and
put it in the
side of the
ark of the
covenant of the
{\ND{
LORD}} your
God, that it may be there for a
witness against thee.
\VS{27}For I
know thy
rebellion, and thy
stiff
neck: behold, while I am yet
alive with you this
day, ye have been
rebellious against the
{\ND{
LORD}}; and how much more
after my
death?
\par }{\PP \VS{28}¶
Gather unto me all the
elders of your
tribes, and your
officers, that I may
speak these
words in their
ears, and
call
heaven and
earth to
record against them.
\VS{29}For I
know that
after my
death ye will
utterly
corrupt
{\ADD{yourselves}}, and turn
aside from the
way which I have
commanded you; and
evil will
befall you in the
latter
days; because ye will
do
evil in the
sight of the
{\ND{
LORD}}, to provoke him to
anger through the
work of your
hands.
\VS{30}And
Moses
spake in the
ears of all the
congregation of
Israel the
words of this
song, until they were
ended.

\par }\Chap{32}{\PP \VerseOne{1}Give
ear, O ye
heavens, and I will
speak; and
hear, O
earth, the
words of my
mouth.
\VS{2}My
doctrine shall
drop as the
rain, my
speech shall
distil as the
dew, as the small
rain upon the tender
herb, and as the
showers upon the
grass:
\VS{3}Because I will
publish the
name of the
{\ND{
LORD}}:
ascribe ye
greatness unto our
God.
\VS{4}{\ADD{He is}} the
Rock, his
work
{\ADD{is}}
perfect: for all his
ways
{\ADD{are}}
judgment: a
God of
truth and without
iniquity,
just and
right
{\ADD{is}} he.
\VS{5}They have
corrupted themselves, their
spot
{\ADD{is}} not
{\ADD{the spot}} of his
children:
{\ADD{they are}} a
perverse and
crooked
generation.\FTNT{*}{{\FR 32.5 }
They have…: Heb. He hath corrupted to himself}\FTNT{†}{{\FR 32.5 }
their…: or, that they are not his children, that is their blot}
\VS{6}Do ye thus
requite the
{\ND{
LORD}}, O
foolish
people and
unwise?
{\ADD{is}} not he thy
father
{\ADD{that}} hath
bought thee? hath he not
made thee, and
established thee?
\par }{\PP \VS{7}¶
Remember the
days of
old,
consider the
years of
many
generations:
ask thy
father, and he will
shew thee; thy
elders, and they will
tell thee.\FTNT{‡}{{\FR 32.7 }
many…: Heb. generation and generation}
\VS{8}When the most
High
divided to the
nations their
inheritance, when he
separated the
sons of
Adam, he
set the
bounds of the
people according to the
number of the
children of
Israel.
\VS{9}For the
{\ND{
LORD’s}}
portion
{\ADD{is}} his
people;
Jacob
{\ADD{is}} the
lot of his
inheritance.\FTNT{§}{{\FR 32.9 }
lot: Heb. cord}
\VS{10}He
found him in a
desert
land, and in the
waste
howling
wilderness; he led him
about, he
instructed him, he
kept him as the
apple of his
eye.\FTNT{**}{{\FR 32.10 }
led: or, compassed}
\VS{11}As an
eagle stirreth
up her
nest,
fluttereth over her
young, spreadeth
abroad her
wings,
taketh them,
beareth them on her
wings:
\VS{12}{\ADD{So}} the
{\ND{
LORD}}
alone did
lead him, and
{\ADD{there was}} no
strange
god with him.
\VS{13}He made him
ride on the high
places of the
earth, that he might
eat the
increase of the
fields; and he made him to
suck
honey out of the
rock, and
oil out of the
flinty
rock;
\VS{14}Butter of
kine, and
milk of
sheep, with
fat of
lambs, and
rams of the
breed of
Bashan, and
goats, with the
fat of
kidneys of
wheat; and thou didst
drink the
pure
blood of the
grape.
\par }{\PP \VS{15}¶ But
Jeshurun waxed
fat, and
kicked: thou art waxen
fat, thou art grown
thick, thou art
covered
{\ADD{with fatness;}} then he
forsook
God
{\ADD{which}}
made him, and lightly
esteemed the
Rock of his
salvation.
\VS{16}They provoked him to
jealousy with
strange
{\ADD{gods}}, with
abominations provoked they him to
anger.
\VS{17}They
sacrificed unto
devils, not to
God; to
gods whom they
knew not, to
new
{\ADD{gods that}}
came
newly
up, whom your
fathers
feared not.\FTNT{††}{{\FR 32.17 }
not to…: or, which were not God}
\VS{18}Of the
Rock
{\ADD{that}}
begat thee thou art
unmindful, and hast
forgotten
God that
formed thee.
\par }{\PP \VS{19}And when the
{\ND{
LORD}}
saw
{\ADD{it}}, he
abhorred
{\ADD{them}}, because of the
provoking of his
sons, and of his
daughters.\FTNT{‡‡}{{\FR 32.19 }
abhorred: or, despised}
\VS{20}And he
said, I will
hide my
face from them, I will
see what their
end
{\ADD{shall be:}} for they
{\ADD{are}} a very
froward
generation,
children in whom
{\ADD{is}} no
faith.
\VS{21}They have moved me to
jealousy with
{\ADD{that which is}}
not
God; they have provoked me to
anger with their
vanities: and I will move them to
jealousy with
{\ADD{those which are}} not a
people; I will provoke them to
anger with a
foolish
nation.
\VS{22}For a
fire is
kindled in mine
anger, and shall
burn unto the
lowest
hell, and shall
consume the
earth with her
increase, and set on
fire the
foundations of the
mountains.\FTNT{§§}{{\FR 32.22 }
shall burn: or, hath burned}\FTNT{***}{{\FR 32.22 }
shall consume: or, hath consumed}
\VS{23}I will
heap
mischiefs upon them; I will
spend mine
arrows upon them.
\VS{24}{\ADD{They shall be}}
burnt with
hunger, and
devoured with burning
heat, and with
bitter
destruction: I will also
send the
teeth of
beasts upon them, with the
poison of
serpents of the
dust.\FTNT{†††}{{\FR 32.24 }
heat: Heb. coals}
\VS{25}The
sword
without, and
terror
within, shall
destroy both the young
man and the
virgin, the
suckling
{\ADD{also}} with the
man of gray
hairs.\FTNT{‡‡‡}{{\FR 32.25 }
within: Heb. from the chambers}\FTNT{§§§}{{\FR 32.25 }
destroy: Heb. bereave}
\par }{\PP \VS{26}I
said, I would scatter them into
corners, I would make the
remembrance of them to
cease from among
men:
\VS{27}Were it
not that I
feared the
wrath of the
enemy, lest their
adversaries should behave themselves
strangely,
{\ADD{and}} lest they should
say, Our
hand
{\ADD{is}}
high, and the
{\ND{
LORD}} hath not
done all this.\FTNT{*}{{\FR 32.27 }
Our…: or, Our high hand, and not the
{\ND{LORD}} hath done}
\VS{28}For they
{\ADD{are}} a
nation
void of
counsel, neither
{\ADD{is there any}}
understanding in them.
\VS{29}O
that they were
wise,
{\ADD{that}} they
understood this,
{\ADD{that}} they would
consider their latter
end!
\VS{30}How should
one
chase a
thousand, and
two put
ten
thousand to
flight,
except their
Rock had
sold
them, and the
{\ND{
LORD}} had shut them
up?
\VS{31}For their
rock
{\ADD{is}} not as our
Rock, even our
enemies themselves
{\ADD{being}}
judges.
\VS{32}For their
vine
{\ADD{is}} of the
vine of
Sodom, and of the
fields of
Gomorrah: their
grapes
{\ADD{are}}
grapes of
gall, their
clusters
{\ADD{are}}
bitter:\FTNT{†}{{\FR 32.32 }
of the vine: or, worse than the vine}
\VS{33}Their
wine
{\ADD{is}} the
poison of
dragons, and the
cruel
venom of
asps.
\VS{34}{\ADD{Is}} not this laid up in
store with me,
{\ADD{and}} sealed
up among my
treasures?
\VS{35}To me
{\ADD{belongeth}}
vengeance, and
recompence; their
foot shall
slide in
{\ADD{due}}
time: for the
day of their
calamity
{\ADD{is}} at
hand, and the things that shall
come upon them make
haste.
\VS{36}For the
{\ND{
LORD}} shall
judge his
people, and
repent himself for his
servants, when he
seeth that
{\ADD{their}}
power is
gone, and
{\ADD{there is}}
none shut
up, or
left.\FTNT{‡}{{\FR 32.36 }
power: Heb. hand}
\VS{37}And he shall
say, Where
{\ADD{are}} their
gods,
{\ADD{their}}
rock in whom they
trusted,
\VS{38}Which did
eat the
fat of their
sacrifices,
{\ADD{and}}
drank the
wine of their drink
offerings? let them rise
up and
help you,
{\ADD{and}} be your
protection.\FTNT{§}{{\FR 32.38 }
your…: Heb. an hiding for you}
\par }{\PP \VS{39}See now that I,
{\ADD{even}} I,
{\ADD{am}} he, and
{\ADD{there is}} no
god with me: I
kill, and I make
alive; I
wound, and I
heal: neither
{\ADD{is there any}} that can
deliver out of my
hand.
\VS{40}For I lift
up my
hand to
heaven, and
say, I
live for
ever.
\VS{41}If I
whet my
glittering
sword, and mine
hand take
hold on
judgment; I will
render
vengeance to mine
enemies, and will
reward them that
hate me.
\VS{42}I will make mine
arrows
drunk with
blood, and my
sword shall
devour
flesh;
{\ADD{and that}} with the
blood of the
slain and of the
captives, from the
beginning of
revenges upon the
enemy.
\VS{43}Rejoice, O ye
nations,
{\ADD{with}} his
people: for he will
avenge the
blood of his
servants, and will
render
vengeance to his
adversaries, and will be
merciful unto his
land,
{\ADD{and}} to his
people.\FTNT{**}{{\FR 32.43 }
Rejoice…: or, Praise his people, ye nations: or, Sing ye}
\par }{\PP \VS{44}¶ And
Moses
came and
spake all the
words of this
song in the
ears of the
people, he, and
Hoshea the
son of
Nun.\FTNT{††}{{\FR 32.44 }
Hoshea: or, Joshua}
\VS{45}And
Moses made an
end of
speaking all these
words to all
Israel:
\VS{46}And he
said unto them,
Set your
hearts unto all the
words which I
testify among you this
day, which ye shall
command your
children to
observe to
do, all the
words of this
law.
\VS{47}For it
{\ADD{is}} not a
vain
thing for you; because it
{\ADD{is}} your
life: and through this
thing ye shall
prolong
{\ADD{your}}
days in the
land, whither ye go
over
Jordan to
possess it.
\VS{48}And the
{\ND{
LORD}}
spake unto
Moses that
selfsame
day,
saying,
\VS{49}Get thee
up into this
mountain
Abarim,
{\ADD{unto}}
mount
Nebo, which
{\ADD{is}} in the
land of
Moab, that
{\ADD{is}} over
against
Jericho; and
behold the
land of
Canaan, which I
give unto the
children of
Israel for a
possession:
\VS{50}And
die in the
mount whither thou goest
up, and be
gathered unto thy
people; as
Aaron thy
brother
died in
mount
Hor, and was
gathered unto his
people:
\VS{51}Because ye
trespassed against me
among the
children of
Israel at the
waters of
Meribah-Kadesh, in the
wilderness of
Zin; because ye
sanctified me not in the
midst of the
children of
Israel.\FTNT{‡‡}{{\FR 32.51 }
Meribah-Kadesh: or, strife at Kadesh}
\VS{52}Yet thou shalt
see the
land before
{\ADD{thee;}} but thou shalt not
go thither unto the
land which I
give the
children of
Israel.

\par }\Chap{33}{\PP \VerseOne{1}And this
{\ADD{is}} the
blessing, wherewith
Moses the
man of
God
blessed the
children of
Israel
before his
death.
\VS{2}And he
said, The
{\ND{
LORD}}
came from
Sinai, and rose
up from
Seir unto them; he shined
forth from
mount
Paran, and he
came with ten
thousands of
saints: from his right
hand
{\ADD{went}} a
fiery
law for them.\FTNT{*}{{\FR 33.2 }
a fiery…: Heb. a fire of law}
\VS{3}Yea, he
loved the
people; all his
saints
{\ADD{are}} in thy
hand: and they sat
down at thy
feet;
{\ADD{every one}} shall
receive of thy
words.
\VS{4}Moses
commanded us a
law,
{\ADD{even}} the
inheritance of the
congregation of
Jacob.
\VS{5}And he was
king in
Jeshurun, when the
heads of the
people
{\ADD{and}} the
tribes of
Israel were
gathered
together.
\par }{\PP \VS{6}¶ Let
Reuben
live, and not
die; and let
{\ADD{not}} his
men be
few.
\par }{\PP \VS{7}¶ And this
{\ADD{is the blessing}} of
Judah: and he
said,
Hear,
{\ND{
LORD}}, the
voice of
Judah, and
bring him unto his
people: let his
hands be
sufficient for him; and be thou an
help
{\ADD{to him}} from his
enemies.
\par }{\PP \VS{8}¶ And of
Levi he
said,
{\ADD{Let}} thy
Thummim and thy
Urim
{\ADD{be}} with thy
holy
one, whom thou didst
prove at
Massah,
{\ADD{and with}} whom thou didst
strive at the
waters of
Meribah;
\VS{9}Who
said unto his
father and to his
mother, I have not
seen him; neither did he
acknowledge his
brethren, nor
knew his own
children: for they have
observed thy
word, and
kept thy
covenant.
\VS{10}They shall
teach
Jacob thy
judgments, and
Israel thy
law: they shall
put
incense
before thee, and whole burnt
sacrifice upon thine
altar.\FTNT{†}{{\FR 33.10 }
They shall teach: or, Let them teach}\FTNT{‡}{{\FR 33.10 }
they shall put: or, let them put}\FTNT{§}{{\FR 33.10 }
before…: Heb. at thy nose}
\VS{11}Bless,
{\ND{
LORD}}, his
substance, and
accept the
work of his
hands:
smite through the
loins of them that
rise against him, and of them that
hate him, that they
rise
not
again.
\par }{\PP \VS{12}¶
{\ADD{And}} of
Benjamin he
said, The
beloved of the
{\ND{
LORD}} shall
dwell in
safety by him;
{\ADD{and the
{\ND{Lord}}}} shall
cover him all the
day long, and he shall
dwell between his
shoulders.
\par }{\PP \VS{13}¶ And of
Joseph he
said,
Blessed of the
{\ND{
LORD}}
{\ADD{be}} his
land, for the precious
things of
heaven, for the
dew, and for the
deep that coucheth
beneath,
\VS{14}And for the
precious
fruits
{\ADD{brought forth}} by the
sun, and for the precious
things put
forth by the
moon,\FTNT{**}{{\FR 33.14 }
put: Heb. thrust}\FTNT{††}{{\FR 33.14 }
moon: Heb. moons}
\VS{15}And for the chief
things of the
ancient
mountains, and for the precious
things of the
lasting
hills,
\VS{16}And for the precious
things of the
earth and
fulness thereof, and
{\ADD{for}} the good
will of him that
dwelt in the
bush: let
{\ADD{the blessing}}
come upon the
head of
Joseph, and upon the top of the
head of him
{\ADD{that was}}
separated from his
brethren.
\VS{17}His
glory
{\ADD{is like}} the
firstling of his
bullock, and his
horns
{\ADD{are like}} the
horns of
unicorns: with them he shall
push the
people
together to the
ends of the
earth: and they
{\ADD{are}} the ten
thousands of
Ephraim, and they
{\ADD{are}} the
thousands of
Manasseh.\FTNT{‡‡}{{\FR 33.17 }
unicorns: Heb. an unicorn}
\par }{\PP \VS{18}¶ And of
Zebulun he
said,
Rejoice,
Zebulun, in thy going
out; and,
Issachar, in thy
tents.
\VS{19}They shall
call the
people unto the
mountain; there they shall
offer
sacrifices of
righteousness: for they shall
suck
{\ADD{of}} the
abundance of the
seas, and
{\ADD{of}}
treasures
hid in the
sand.
\par }{\PP \VS{20}¶ And of
Gad he
said,
Blessed
{\ADD{be}} he that
enlargeth
Gad: he
dwelleth as a
lion, and
teareth the
arm
with the crown of the
head.
\VS{21}And he
provided the first
part for himself, because there,
{\ADD{in}} a
portion of the
lawgiver,
{\ADD{was he}}
seated; and he
came with the
heads of the
people, he
executed the
justice of the
{\ND{
LORD}}, and his
judgments with
Israel.\FTNT{§§}{{\FR 33.21 }
seated: Heb. cieled}
\par }{\PP \VS{22}¶ And of
Dan he
said,
Dan
{\ADD{is}} a
lion’s
whelp: he shall
leap from
Bashan.
\par }{\PP \VS{23}¶ And of
Naphtali he
said, O
Naphtali,
satisfied with
favour, and
full with the
blessing of the
{\ND{
LORD}}:
possess thou the
west and the
south.
\par }{\PP \VS{24}¶ And of
Asher he
said,
{\ADD{Let}}
Asher
{\ADD{be}}
blessed with
children; let him be
acceptable to his
brethren, and let him
dip his
foot in
oil.
\VS{25}Thy
shoes
{\ADD{shall be}}
iron and
brass; and as thy
days,
{\ADD{so shall}} thy
strength
{\ADD{be}}.\FTNT{***}{{\FR 33.25 }
Thy shoes: or, Under thy shoes}
\par }{\PP \VS{26}¶
{\ADD{There is}} none like unto the
God of
Jeshurun,
{\ADD{who}}
rideth upon the
heaven in thy
help, and in his
excellency on the
sky.
\VS{27}The
eternal
God
{\ADD{is thy}}
refuge, and underneath
{\ADD{are}} the
everlasting
arms: and he shall thrust
out the
enemy from
before thee; and shall
say,
Destroy
{\ADD{them}}.
\VS{28}Israel then shall
dwell in
safety
alone: the
fountain of
Jacob
{\ADD{shall be}} upon a
land of
corn and
wine; also his
heavens shall drop
down
dew.
\VS{29}Happy
{\ADD{art}} thou, O
Israel: who
{\ADD{is}} like unto thee, O
people
saved by the
{\ND{
LORD}}, the
shield of thy
help, and who
{\ADD{is}} the
sword of thy
excellency! and thine
enemies shall be found
liars unto thee; and thou shalt
tread upon their high
places.\FTNT{†††}{{\FR 33.29 }
found…: or, subdued}

\par }\Chap{34}{\PP \VerseOne{1}And
Moses went
up from the
plains of
Moab unto the
mountain of
Nebo, to the
top of
Pisgah, that
{\ADD{is}} over
against
Jericho. And the
{\ND{
LORD}}
shewed him all the
land of
Gilead, unto
Dan,\FTNT{*}{{\FR 34.1 }
Pisgah: or, the hill}
\VS{2}And all
Naphtali, and the
land of
Ephraim, and
Manasseh, and all the
land of
Judah, unto the
utmost
sea,
\VS{3}And the
south, and the
plain of the
valley of
Jericho, the
city of palm
trees, unto
Zoar.
\VS{4}And the
{\ND{
LORD}}
said unto him, This
{\ADD{is}} the
land which I
sware unto
Abraham, unto
Isaac, and unto
Jacob,
saying, I will
give it unto thy
seed: I have caused thee to
see
{\ADD{it}} with thine
eyes, but thou shalt not go
over thither.
\par }{\PP \VS{5}¶ So
Moses the
servant of the
{\ND{
LORD}}
died there in the
land of
Moab, according to the
word of the
{\ND{
LORD}}.
\VS{6}And he
buried him in a
valley in the
land of
Moab, over
against
Beth-peor: but no
man
knoweth of his
sepulchre unto this
day.
\par }{\PP \VS{7}¶ And
Moses
{\ADD{was}} an
hundred and
twenty
years
old when he
died: his
eye was not
dim, nor his natural
force
abated.\FTNT{†}{{\FR 34.7 }
natural…: Heb. moisture}\FTNT{‡}{{\FR 34.7 }
abated: Heb. fled}
\par }{\PP \VS{8}¶ And the
children of
Israel
wept for
Moses in the
plains of
Moab
thirty
days: so the
days of
weeping
{\ADD{and}}
mourning for
Moses were
ended.
\par }{\PP \VS{9}¶ And
Joshua the
son of
Nun was
full of the
spirit of
wisdom; for
Moses had
laid his
hands upon him: and the
children of
Israel
hearkened unto him, and
did as the
{\ND{
LORD}}
commanded
Moses.
\par }{\PP \VS{10}¶ And there
arose not a
prophet
since in
Israel like unto
Moses, whom the
{\ND{
LORD}}
knew
face to
face,
\VS{11}In all the
signs and the
wonders, which the
{\ND{
LORD}}
sent him to
do in the
land of
Egypt to
Pharaoh, and to all his
servants, and to all his
land,
\VS{12}And in all that
mighty
hand, and in all the
great
terror which
Moses
shewed in the
sight of all
Israel.
\par }