\NormalFont\ShortTitle{Galatians}
{\MT THE EPISTLE OF PAUL THE APOSTLE TO THE GALATIANS

\par }\ChapOne{1}{\PP \VerseOne{1}Paul, an
apostle, (not
of
men,
neither
by
man,
but
by
Jesus
Christ,
and
God the
Father,
who
raised
him
from the
dead;)
\VS{2}And
all the
brethren which
are
with
me, unto the
churches of
Galatia:
\VS{3}Grace
{\ADD{be}} to
you
and
peace
from
God the
Father,
and
{\ADD{from}}
our
Lord
Jesus
Christ,
\VS{4}Who
gave
himself
for
our
sins,
that he might
deliver
us
from
this
present
evil
world, according
to the
will of
God
and
our
Father:
\VS{5}To
whom
{\ADD{be}}
glory
for
ever and
ever.
Amen.
\par }{\PP \VS{6}I
marvel
that ye
are
so
soon
removed
from him
that
called
you
into the
grace of
Christ
unto
another
gospel:
\VS{7}Which
is
not
another;
but there
be
some
that
trouble
you,
and
would
pervert the
gospel of
Christ.
\VS{8}But
though
we,
or an
angel
from
heaven, preach any other
gospel unto
you
than that
which we have
preached unto
you, let him
be
accursed.
\VS{9}As we said
before,
so say
I
now
again,
If
any
{\ADD{man}} preach any other
gospel unto
you
than
that ye have
received, let him
be
accursed.
\par }{\PP \VS{10}For do
I
now
persuade
men,
or
God?
or do I
seek to
please
men?
for
if
I
yet
pleased
men, I
should
not
be the
servant of
Christ.
\VS{11}But I
certify
you,
brethren,
that the
gospel
which was
preached
of
me
is
not
after
man.
\VS{12}For
I
neither
received
it
of
man,
neither was I
taught
{\ADD{it}},
but
by the
revelation of
Jesus
Christ.
\VS{13}For ye have heard
of
my
conversation in time
past
in the
Jews’ religion, how
that beyond
measure I
persecuted the
church of
God,
and
wasted
it:
\VS{14}And
profited
in the
Jews’ religion
above
many my
equals
in mine
own
nation,
being more
exceedingly
zealous of the
traditions of
my
fathers.
\VS{15}But
when it
pleased
God,
who
separated
me
from
my
mother’s
womb,
and
called
{\ADD{me}}
by
his
grace,
\VS{16}To
reveal
his
Son
in
me,
that I might
preach
him
among the
heathen;
immediately I
conferred
not with
flesh
and
blood:
\VS{17}Neither went I
up
to
Jerusalem
to them which
were
apostles
before
me;
but I
went
into
Arabia,
and
returned
again
unto
Damascus.
\VS{18}Then
after
three
years I went
up
to
Jerusalem to
see
Peter,
and
abode
with
him
fifteen
days.
\VS{19}But
other of the
apostles saw
I
none,
save
James
the
Lord’s
brother.
\VS{20}Now the things
which I
write unto
you,
behold,
before
God, I
lie
not.
\VS{21}Afterwards I
came
into the
regions of
Syria
and
Cilicia;
\VS{22}And
was
unknown by
face unto the
churches of
Judaea which
were
in
Christ:
\VS{23}But they
had
heard
only,
That he
which
persecuted
us in times
past
now
preacheth the
faith
which
once he
destroyed.
\VS{24}And they
glorified
God
in
me.

\par }\Chap{2}{\PP \VerseOne{1}Then
fourteen
years
after I went
up
again
to
Jerusalem
with
Barnabas, and
took
Titus
with
{\ADD{me}}
also.
\VS{2}And I went
up
by
revelation,
and
communicated unto
them that
gospel
which I
preach
among the
Gentiles,
but
privately to them
which were of
reputation, lest by any
means I should
run,
or had
run,
in
vain.
\VS{3}But
neither
Titus, who
was
with
me,
being a
Greek, was
compelled to be
circumcised:
\VS{4}And
that because
of false
brethren unawares brought
in,
who came in
privily to spy
out
our
liberty
which we
have
in
Christ
Jesus,
that they might
bring
us into
bondage:
\VS{5}To
whom we gave
place by
subjection, no,
not
for an
hour;
that the
truth of the
gospel might
continue
with
you.
\VS{6}But
of these
who
seemed to
be
somewhat, (whatsoever they
were, it
maketh
no
matter to
me:
God
accepteth
no
man’s
person:)
for they
who
seemed
{\ADD{to be somewhat}} in conference
added
nothing to
me:
\VS{7}But
contrariwise, when they
saw
that the
gospel of the
uncircumcision was committed unto
me,
as
{\ADD{the gospel}} of the
circumcision
{\ADD{was}} unto
Peter;
\VS{8}(For he
that wrought
effectually in
Peter
to the
apostleship of the
circumcision, the same was
mighty in
me
toward the
Gentiles:)
\VS{9}And when
James,
Cephas,
and
John,
who
seemed to
be
pillars,
perceived the
grace
that was
given unto
me, they
gave to
me
and
Barnabas the right
hands of
fellowship;
that
we
{\ADD{should go}}
unto the
heathen,
and
they
unto the
circumcision.
\VS{10}Only
{\ADD{they would}}
that we should
remember the
poor; the
same
which
I
also was
forward to
do.
\VS{11}But
when
Peter was
come
to
Antioch, I
withstood
him
to the
face,
because he
was to be
blamed.
\VS{12}For
before
that
certain
came
from
James, he did
eat
with the
Gentiles:
but
when they were
come, he
withdrew
and
separated
himself,
fearing them
which were
of the
circumcision.
\VS{13}And the
other
Jews
dissembled
likewise with
him; insomuch
that
Barnabas
also was carried away
with
their
dissimulation.
\VS{14}But
when I
saw
that they
walked
not
uprightly according
to the
truth of the
gospel, I
said unto
Peter
before
{\ADD{them}}
all,
If
thou,
being a
Jew,
livest after the manner of
Gentiles,
and
not as do the
Jews,
why compellest
thou the
Gentiles to live as do the
Jews?
\VS{15}We
{\ADD{who are}}
Jews by
nature,
and
not
sinners
of the
Gentiles,
\VS{16}Knowing
that a
man
is
not
justified
by the
works of the
law,
but
by the
faith of
Jesus
Christ,
even
we have
believed
in
Jesus
Christ,
that we might be
justified
by the
faith of
Christ,
and
not
by the
works of the
law: for
by the
works
of the
law
shall
no
flesh be
justified.
\VS{17}But
if, while we
seek to be
justified
by
Christ, we
ourselves
also are
found
sinners,
{\ADD{is}}
therefore
Christ the
minister of
sin? God
forbid.
\VS{18}For
if I
build
again the
things
which I
destroyed, I
make
myself a
transgressor.
\VS{19}For
I
through the
law am
dead to the
law,
that I might
live unto
God.
\VS{20}I am crucified
with
Christ:
nevertheless I
live; yet
not
I,
but
Christ
liveth
in
me:
and the life
which
I
now
live
in the
flesh I
live
by the
faith of the
Son of
God,
who
loved
me,
and
gave
himself
for
me.
\VS{21}I
do
not
frustrate the
grace of
God:
for
if
righteousness
{\ADD{come}}
by the
law,
then
Christ is
dead in
vain.

\par }\Chap{3}{\PP \VerseOne{1}O
foolish
Galatians,
who hath
bewitched
you, that ye
should
not
obey the
truth,
before
whose
eyes
Jesus
Christ hath been evidently set
forth,
crucified
among
you?
\VS{2}This
only
would I
learn
of
you, Received
ye the
Spirit
by the
works of the
law,
or
by the
hearing of
faith?
\VS{3}Are
ye
so
foolish? having
begun in the
Spirit, are
ye
now made
perfect by the
flesh?
\VS{4}Have ye
suffered so many
things in
vain?
if
{\ADD{it be}}
yet in
vain.
\VS{5}He
therefore
that
ministereth to
you the
Spirit,
and
worketh
miracles
among
you,
{\ADD{doeth he it}}
by the
works of the
law,
or
by the
hearing of
faith?
\VS{6}Even
as
Abraham
believed
God,
and it was
accounted to
him
for
righteousness.
\par }{\PP \VS{7}Know
ye
therefore
that they which
are
of
faith, the
same
are the children
of
Abraham.
\VS{8}And the
scripture,
foreseeing
that
God would
justify the
heathen
through
faith, preached before the
gospel unto
Abraham,
{\ADD{saying}},
In
thee
shall
all
nations be
blessed.
\VS{9}So
then
they which
be
of
faith are
blessed
with
faithful
Abraham.
\VS{10}For as many
as
are
of the
works of the
law
are
under the
curse:
for it is
written,
Cursed
{\ADD{is}} every
one
that
continueth
not
in all
things
which are
written
in the
book of the
law to
do
them.
\VS{11}But
that no
man is
justified
by the
law in the sight
of
God,
{\ADD{it is}}
evident:
for, The
just shall
live
by
faith.
\VS{12}And the
law
is
not
of
faith:
but, The
man that
doeth
them shall
live
in
them.
\VS{13}Christ hath
redeemed
us
from the
curse of the
law, being
made a
curse
for
us:
for it is
written,
Cursed
{\ADD{is}} every
one
that
hangeth
on a
tree:
\VS{14}That the
blessing of
Abraham might
come
on the
Gentiles
through
Jesus
Christ;
that we might
receive the
promise of the
Spirit
through
faith.
\VS{15}Brethren, I
speak after the manner
of
men;
Though
{\ADD{it be}}
but a
man’s
covenant, yet
{\ADD{if it be}}
confirmed, no
man
disannulleth,
or addeth
thereto.
\VS{16}Now to
Abraham
and
his
seed
were the
promises
made. He
saith
not,
And to
seeds,
as
of
many;
but
as
of
one,
And
to
thy
seed,
which
is
Christ.
\VS{17}And
this I
say,
{\ADD{that}} the
covenant, that was confirmed
before
of
God
in
Christ, the
law,
which
was four
hundred
and
thirty
years
after,
cannot
disannul,
that it should
make the
promise of none
effect.
\VS{18}For
if the
inheritance
{\ADD{be}}
of the
law,
{\ADD{it is}} no
more
of
promise:
but
God
gave
{\ADD{it}} to
Abraham
by
promise.
\VS{19}Wherefore
then
{\ADD{serveth}} the
law? It was
added
because of
transgressions,
till the
seed should
come to
whom the promise was
made;
{\ADD{and it was}}
ordained
by
angels
in the
hand of a
mediator.
\VS{20}Now a
mediator
is
not
{\ADD{a mediator}} of
one,
but
God
is
one.
\VS{21}{\ADD{Is}} the
law
then
against the
promises of
God? God
forbid:
for
if there had
been a
law
given
which could
have given
life,
verily
righteousness should have
been
by the
law.
\VS{22}But the
scripture hath
concluded
all
under
sin,
that the
promise
by
faith of
Jesus
Christ might be
given to them
that
believe.
\par }{\PP \VS{23}But
before
faith
came, we were
kept
under the
law, shut
up
unto the
faith
which should
afterwards be
revealed.
\VS{24}Wherefore the
law
was
our
schoolmaster
{\ADD{to bring us}}
unto
Christ,
that we might be
justified
by
faith.
\VS{25}But after
that
faith is
come, we
are no
longer
under a
schoolmaster.
\VS{26}For ye
are
all the
children of
God
by
faith
in
Christ
Jesus.
\VS{27}For as
many of
you
as have been
baptized
into
Christ have put
on
Christ.
\VS{28}There
is
neither
Jew
nor
Greek, there
is
neither
bond
nor
free, there
is
neither
male
nor
female:
for
ye
are
all
one
in
Christ
Jesus.
\VS{29}And
if
ye
{\ADD{be}}
Christ’s,
then are
ye
Abraham’s
seed,
and
heirs according
to the
promise.

\par }\Chap{4}{\PP \VerseOne{1}Now I
say,
{\ADD{That}} the
heir, as long
as he
is a
child,
differeth
nothing
from a
servant,
though he
be
lord of
all;
\VS{2}But
is
under
tutors
and
governors
until the time
appointed of the
father.
\VS{3}Even
so
we,
when we
were
children,
were in
bondage
under the
elements of the
world:
\VS{4}But
when the
fulness of the
time was
come,
God sent
forth
his
Son,
made
of a
woman,
made
under the
law,
\VS{5}To
redeem
them that
were
under the
law,
that we might
receive the adoption of
sons.
\VS{6}And
because ye
are
sons,
God hath sent
forth the
Spirit of
his
Son
into
your
hearts,
crying,
Abba,
Father.
\VS{7}Wherefore thou
art no
more a
servant,
but a
son;
and
if a
son,
then an
heir of
God
through
Christ.
\par }{\PP \VS{8}Howbeit
then, when ye knew
not
God, ye did
service unto them
which by
nature
are
no
gods.
\VS{9}But
now, after that ye have
known
God,
or
rather are
known
of
God,
how turn
ye
again
to the
weak
and
beggarly
elements,
whereunto ye
desire
again to be in
bondage?
\VS{10}Ye
observe
days,
and
months,
and
times,
and
years.
\VS{11}I am afraid
of
you,
lest I have
bestowed
upon
you
labour in
vain.
\par }{\PP \VS{12}Brethren, I
beseech
you,
be
as
I
{\ADD{am}}; for
I
{\ADD{am}}
as
ye
{\ADD{are}}: ye
have
not
injured
me at all.
\VS{13}Ye
know
how
through
infirmity of the
flesh I preached the
gospel unto
you at the
first.
\VS{14}And
my
temptation which
was
in
my
flesh ye
despised
not,
nor
rejected;
but
received
me
as an
angel of
God,
{\ADD{even}}
as
Christ
Jesus.
\VS{15}Where
is
then the
blessedness ye spake
of?
for I
bear
you
record,
that,
if
{\ADD{it had been}}
possible, ye would have plucked
out your
own
eyes, and have given
them to
me.
\VS{16}Am
I
therefore
become
your
enemy, because I
tell
you the
truth?
\VS{17}They zealously
affect
you,
{\ADD{but}}
not
well;
yea, they
would
exclude
you,
that ye might
affect
them.
\VS{18}But
{\ADD{it is}}
good to be zealously
affected
always
in
{\ADD{a}}
good
{\ADD{thing}},
and
not
only when
I am
present
with
you.
\VS{19}My little
children, of
whom I travail in
birth
again
until
Christ be
formed
in
you,
\VS{20}I
desire to be
present
with
you
now,
and to
change
my
voice;
for I stand in
doubt
of
you.
\par }{\PP \VS{21}Tell
me, ye
that
desire to
be
under the
law, do
ye
not
hear the
law?
\VS{22}For it is
written,
that
Abraham
had
two
sons, the
one
by a
bondmaid, the
other
by a
freewoman.
\VS{23}But
he
{\ADD{who was}}
of the
bondwoman was
born
after the
flesh;
but
he
of the
freewoman
{\ADD{was}}
by
promise.
\VS{24}Which
things
are an
allegory:
for
these
are
the
two
covenants; the
one
from the
mount
Sinai, which
gendereth
to
bondage,
which
is
Agar.
\VS{25}For
this
Agar
is
mount
Sinai
in
Arabia,
and answereth
to
Jerusalem
which
now
is,
and is in
bondage
with
her
children.
\VS{26}But
Jerusalem which
is
above
is
free,
which
is the
mother of
us
all.
\VS{27}For it is
written,
Rejoice,
{\ADD{thou}}
barren
that
bearest
not; break
forth
and
cry, thou
that
travailest
not:
for the desolate
hath
many
more
children
than she
which
hath an
husband.
\VS{28}Now
we,
brethren,
as Isaac
was,
are the
children of
promise.
\VS{29}But
as
then he
that was
born
after the
flesh
persecuted
him
{\ADD{that was born}}
after the
Spirit,
even
so
{\ADD{it is}}
now.
\VS{30}Nevertheless
what
saith the
scripture? Cast
out the
bondwoman
and
her
son:
for the
son of the
bondwoman
shall
not be
heir
with the
son of the
freewoman.
\VS{31}So
then,
brethren, we
are
not
children of the
bondwoman,
but of the
free.

\par }\Chap{5}{\PP \VerseOne{1}Stand
fast
therefore in the
liberty
wherewith
Christ hath
made
us
free,
and
be
not
entangled
again with the
yoke of
bondage.
\par }{\PP \VS{2}Behold,
I
Paul
say unto
you,
that
if ye be
circumcised,
Christ shall
profit
you
nothing.
\VS{3}For I
testify
again to
every
man that is
circumcised,
that he
is a
debtor to
do
the
whole
law.
\VS{4}Christ is
become
of no effect unto
you,
whosoever of you are
justified
by the
law; ye are fallen
from
grace.
\VS{5}For
we through the
Spirit wait
for the
hope of
righteousness
by
faith.
\VS{6}For
in
Jesus
Christ
neither
circumcision
availeth
any thing,
nor
uncircumcision;
but
faith which
worketh
by
love.
\VS{7}Ye did
run
well;
who did
hinder
you that ye
should
not
obey the
truth?
\VS{8}This
persuasion
{\ADD{cometh}}
not
of him
that
calleth
you.
\VS{9}A
little
leaven
leaveneth
the
whole
lump.
\VS{10}I have
confidence
in
you
through the
Lord,
that ye will
be
none
otherwise
minded:
but he
that
troubleth
you shall
bear
his
judgment,
whosoever he
be.
\VS{11}And
I,
brethren,
if
I
yet
preach
circumcision,
why do
I
yet suffer
persecution?
then
is the
offence of the
cross
ceased.
\VS{12}I
would they
were
even cut
off
which
trouble
you.
\VS{13}For,
brethren,
ye have been
called
unto
liberty;
only
{\ADD{use}}
not
liberty
for an
occasion to the
flesh,
but
by
love
serve one
another.
\VS{14}For
all the
law is
fulfilled
in
one
word,
{\ADD{even}}
in
this; Thou shalt
love
thy
neighbour
as
thyself.
\VS{15}But
if ye
bite
and
devour one
another, take
heed that ye
be
not
consumed
one
of
another.
\VS{16}{\ADD{This}} I
say
then,
Walk in the
Spirit,
and ye
shall
not
fulfil the
lust of the
flesh.
\VS{17}For the
flesh
lusteth
against the
Spirit,
and the
Spirit
against the
flesh:
and
these are
contrary the one to the
other: so
that
ye
cannot
do the
things
that ye
would.
\VS{18}But
if ye be
led of the
Spirit, ye
are
not
under the
law.
\VS{19}Now the
works of the
flesh
are
manifest,
which
are
{\ADD{these}};
Adultery,
fornication,
uncleanness,
lasciviousness,
\VS{20}Idolatry,
witchcraft,
hatred,
variance,
emulations,
wrath,
strife,
seditions,
heresies,
\VS{21}Envyings,
murders,
drunkenness,
revellings,
and
such
like: of the
which I
tell
you
before,
as I
have
also
told
{\ADD{you}} in time past,
that they
which
do such
things
shall
not
inherit the
kingdom of
God.
\VS{22}But the
fruit of the
Spirit
is
love,
joy,
peace,
longsuffering,
gentleness,
goodness,
faith,
\VS{23}Meekness,
temperance:
against
such there
is
no
law.
\VS{24}And they that
are
Christ’s have
crucified the
flesh
with the
affections
and
lusts.
\VS{25}If we
live in the
Spirit, let
us
also
walk in the
Spirit.
\VS{26}Let
us
not
be desirous of vain
glory,
provoking one
another,
envying one
another.

\par }\Chap{6}{\PP \VerseOne{1}Brethren,
if a
man be
overtaken
in a
fault,
ye which
are
spiritual,
restore such an
one
in the
spirit of
meekness;
considering
thyself,
lest
thou
also be
tempted.
\VS{2}Bear
ye one
another’s
burdens,
and
so
fulfil the
law of
Christ.
\VS{3}For
if a
man
think himself to
be
something, when he
is
nothing, he
deceiveth
himself.
\VS{4}But
let every
man
prove his
own
work,
and
then shall he
have
rejoicing
in
himself
alone,
and
not
in
another.
\VS{5}For every
man shall
bear his
own
burden.
\VS{6}Let him
that is
taught in the
word
communicate unto him
that
teacheth
in
all good
things.
\VS{7}Be
not
deceived;
God
is
not
mocked:
for
whatsoever a
man
soweth,
that shall
he
also
reap.
\VS{8}For he
that
soweth
to
his
flesh
shall
of the
flesh
reap
corruption;
but he
that
soweth
to the
Spirit
shall
of the
Spirit
reap
life
everlasting.
\VS{9}And let
us
not be
weary in
well
doing:
for in
due
season we shall
reap, if we
faint
not.
\VS{10}As we
have
therefore
opportunity, let us
do
good
unto
all
{\ADD{men}},
especially
unto them who
are of the
household of
faith.
\VS{11}Ye
see how
large a
letter I have
written unto
you with mine
own
hand.
\VS{12}As many
as
desire to make a fair
shew
in the
flesh,
they
constrain
you to be
circumcised;
only
lest they should suffer
persecution for the
cross of
Christ.
\VS{13}For
neither
they
themselves
who are
circumcised
keep the
law;
but
desire to
have
you
circumcised,
that they may
glory
in
your
flesh.
\VS{14}But God forbid
that
I should
glory,
save
in the
cross of
our
Lord
Jesus
Christ,
by
whom the
world is
crucified unto
me, and
I unto the
world.
\VS{15}For
in
Christ
Jesus
neither
circumcision
availeth any
thing,
nor
uncircumcision,
but a
new
creature.
\VS{16}And as many
as
walk according to
this
rule,
peace
{\ADD{be}}
on
them,
and
mercy,
and
upon the
Israel of
God.
\VS{17}From
henceforth
let no
man
trouble
me:
for
I
bear
in
my
body the
marks of the
Lord
Jesus.
\VS{18}Brethren, the
grace of
our
Lord
Jesus
Christ
{\ADD{be}}
with
your
spirit.
Amen.
\par }{\SH Unto the Galatians written from Rome.
\par }