\NormalFont\ShortTitle{Habakkuk}
{\MT Habakkuk

\par }\ChapOne{1}{\PP \VerseOne{1}The
burden which
Habakkuk the
prophet did
see.
\VS{2}O
{\ND{
LORD}}, how long shall I
cry, and thou wilt not
hear!
{\ADD{even}} cry
out unto thee
{\ADD{of}}
violence, and thou wilt not
save!
\VS{3}Why dost thou
shew me
iniquity, and cause
{\ADD{me}} to
behold
grievance? for
spoiling and
violence
{\ADD{are}} before me: and there are
{\ADD{that}} raise
up
strife and
contention.
\VS{4}Therefore the
law is
slacked, and
judgment doth
never go
forth: for the
wicked doth
compass about the
righteous; therefore
wrong
judgment
proceedeth.\FTNT{*}{{\FR 1.4 }
wrong: or, wrested}
\par }{\PP \VS{5}¶
Behold ye among the
heathen, and
regard, and
wonder
marvellously: for
{\ADD{I}} will
work a
work in your
days,
{\ADD{which}} ye will not
believe, though it be
told
{\ADD{you}}.
\VS{6}For, lo, I raise
up the
Chaldeans,
{\ADD{that}}
bitter and
hasty
nation, which shall
march through the
breadth of the
land, to
possess the
dwellingplaces
{\ADD{that are}} not theirs.\FTNT{†}{{\FR 1.6 }
breadth: Heb. breadths}
\VS{7}They
{\ADD{are}}
terrible and
dreadful: their
judgment and their
dignity shall
proceed of themselves.\FTNT{‡}{{\FR 1.7 }
their judgment…: or, from them shall proceed the judgment of these, and the captivity of these}
\VS{8}Their
horses also are
swifter than the
leopards, and are more
fierce than the
evening
wolves: and their
horsemen shall
spread themselves, and their
horsemen shall
come from
far; they shall
fly as the
eagle
{\ADD{that}}
hasteth to
eat.\FTNT{§}{{\FR 1.8 }
fierce: Heb. sharp}
\VS{9}They shall
come all for
violence: their
faces shall sup
up
{\ADD{as}} the east
wind, and they shall
gather the
captivity as the
sand.\FTNT{**}{{\FR 1.9 }
their…: or, the supping up of their faces, etc: or, their faces shall look (Heb. the opposition of their faces) toward the east}
\VS{10}And they shall
scoff at the
kings, and the
princes shall be a
scorn unto them: they shall
deride every strong
hold; for they shall
heap
dust, and
take it.
\VS{11}Then shall
{\ADD{his}}
mind
change, and he shall pass
over, and
offend,
{\ADD{imputing}}
this his
power unto his
god.
\par }{\PP \VS{12}¶
{\ADD{Art}} thou not from
everlasting, O
{\ND{
LORD}} my
God, mine Holy
One? we shall not
die. O
{\ND{
LORD}}, thou hast
ordained them for
judgment; and, O mighty
God, thou hast
established them for
correction.\FTNT{††}{{\FR 1.12 }
mighty…: Heb. rock}\FTNT{‡‡}{{\FR 1.12 }
established: Heb. founded}
\VS{13}{\ADD{Thou art}} of
purer
eyes than to
behold
evil, and
canst not
look on
iniquity: wherefore
lookest thou upon them that deal
treacherously,
{\ADD{and}} holdest thy
tongue when the
wicked
devoureth
{\ADD{the man that is}} more
righteous than he?\FTNT{§§}{{\FR 1.13 }
iniquity: or, grievance}
\VS{14}And
makest
men as the
fishes of the
sea, as the creeping
things,
{\ADD{that have}} no
ruler over them?\FTNT{***}{{\FR 1.14 }
creeping: or, moving}
\VS{15}They take
up all of them with the
angle, they
catch them in their
net, and
gather them in their
drag: therefore they
rejoice and are
glad.\FTNT{†††}{{\FR 1.15 }
drag: or, flue net}
\VS{16}Therefore they
sacrifice unto their
net, and burn
incense unto their
drag; because by
them their
portion
{\ADD{is}}
fat, and their
meat
plenteous.\FTNT{‡‡‡}{{\FR 1.16 }
plenteous: or, dainty: Heb. fat}
\VS{17}Shall they therefore
empty their
net, and not
spare
continually to
slay the
nations?

\par }\Chap{2}{\PP \VerseOne{1}I will
stand upon my
watch, and
set me upon the
tower, and will
watch to
see what he will
say unto me, and what I shall
answer when I am
reproved.\FTNT{*}{{\FR 2.1 }
tower: Heb. fenced place}\FTNT{†}{{\FR 2.1 }
unto me: or, in me}\FTNT{‡}{{\FR 2.1 }
when…: or, when I am argued with: Heb. upon my reproof, or, arguing}
\VS{2}And the
{\ND{
LORD}}
answered me, and
said,
Write the
vision, and make
{\ADD{it}}
plain upon
tables, that he may
run that
readeth it.
\VS{3}For the
vision
{\ADD{is}} yet for an appointed
time, but at the
end it shall
speak, and not
lie: though it
tarry,
wait for it; because it will
surely
come, it will not
tarry.
\VS{4}Behold, his
soul
{\ADD{which}} is lifted
up is not
upright in him: but the
just shall
live by his
faith.
\par }{\PP \VS{5}¶ Yea also, because he
transgresseth by
wine,
{\ADD{he is}} a
proud
man, neither keepeth at
home, who
enlargeth his
desire as
hell, and
{\ADD{is}} as
death, and cannot be
satisfied, but
gathereth unto him all
nations, and
heapeth unto him all
people:\FTNT{§}{{\FR 2.5 }
Yea…: or, How much more}
\VS{6}Shall not all these take
up a
parable against him, and a
taunting
proverb against him, and
say,
Woe to him that
increaseth
{\ADD{that which is}} not his! how long? and to him that
ladeth himself with thick
clay!\FTNT{**}{{\FR 2.6 }
Woe…: or, Ho, he}
\VS{7}Shall they not rise
up
suddenly that shall
bite thee, and
awake that shall
vex thee, and thou shalt be for
booties unto them?
\VS{8}Because thou hast
spoiled
many
nations, all the
remnant of the
people shall
spoil thee; because of
men’s
blood, and
{\ADD{for}} the
violence of the
land, of the
city, and of all that
dwell therein.\FTNT{††}{{\FR 2.8 }
blood: Heb. bloods}
\par }{\PP \VS{9}¶
Woe to him that
coveteth an
evil
covetousness to his
house, that he may
set his
nest on
high, that he may be
delivered from the
power of
evil!\FTNT{‡‡}{{\FR 2.9 }
coveteth…: or, gaineth and evil gain}\FTNT{§§}{{\FR 2.9 }
power…: Heb. palm of the hand}
\VS{10}Thou hast
consulted
shame to thy
house by cutting
off
many
people, and hast
sinned
{\ADD{against}} thy
soul.
\VS{11}For the
stone shall cry
out of the
wall, and the beam
out of the
timber shall
answer it.\FTNT{***}{{\FR 2.11 }
beam: or, piece, or, fastening}\FTNT{†††}{{\FR 2.11 }
answer it: or, witness against it}
\par }{\PP \VS{12}¶
Woe to him that
buildeth a
town with
blood, and
stablisheth a
city by
iniquity!\FTNT{‡‡‡}{{\FR 2.12 }
blood: Heb. bloods}
\VS{13}Behold,
{\ADD{is it}} not of the
{\ND{
LORD}} of
hosts that the
people shall
labour in the
very
fire, and the
people shall
weary themselves for
very
vanity?\FTNT{§§§}{{\FR 2.13 }
for…: or, in vain?}
\VS{14}For the
earth shall be
filled with the
knowledge of the
glory of the
{\ND{
LORD}}, as the
waters
cover the
sea.\FTNT{*}{{\FR 2.14 }
with…: or, by knowing the glory}
\par }{\PP \VS{15}¶
Woe unto him that giveth his
neighbour
drink, that
puttest thy
bottle to
{\ADD{him}}, and makest
{\ADD{him}}
drunken also, that thou mayest
look on their
nakedness!
\VS{16}Thou art
filled with
shame for
glory:
drink thou also, and let thy foreskin be
uncovered: the
cup of the
{\ND{
LORD’s}} right
hand shall be
turned unto thee, and shameful
spewing
{\ADD{shall be}} on thy
glory.\FTNT{†}{{\FR 2.16 }
with…: or, more with shame than with glory}
\VS{17}For the
violence of
Lebanon shall
cover thee, and the
spoil of
beasts,
{\ADD{which}} made them
afraid, because of
men’s
blood, and for the
violence of the
land, of the
city, and of all that
dwell therein.
\par }{\PP \VS{18}¶ What
profiteth the graven
image that the
maker thereof hath
graven it; the molten
image, and a
teacher of
lies, that the
maker of his
work
trusteth therein, to
make
dumb
idols?\FTNT{‡}{{\FR 2.18 }
maker of…: Heb. fashioner of his fashion}
\VS{19}Woe unto him that
saith to the
wood,
Awake; to the
dumb
stone,
Arise, it shall
teach! Behold, it
{\ADD{is}}
laid over with
gold and
silver, and
{\ADD{there is}} no
breath at all in the
midst of it.
\VS{20}But the
{\ND{
LORD}}
{\ADD{is}} in his
holy
temple: let all the
earth keep
silence
before him.\FTNT{§}{{\FR 2.20 }
let…: Heb. be silent all the earth before him}

\par }\Chap{3}{\PP \VerseOne{1}A
prayer of
Habakkuk the
prophet upon
Shigionoth.\FTNT{*}{{\FR 3.1 }
upon…: or, according to variable songs, or, tunes, called in Hebrew, Shigionoth}
\VS{2}O
{\ND{
LORD}}, I have
heard thy
speech,
{\ADD{and}} was
afraid: O
{\ND{
LORD}},
revive thy
work in the
midst of the
years, in the
midst of the
years make
known; in
wrath
remember
mercy.\FTNT{†}{{\FR 3.2 }
speech: Heb. report, or, hearing}\FTNT{‡}{{\FR 3.2 }
revive: or, preserve alive}
\par }{\PP \VS{3}God
came from
Teman, and the Holy
One from
mount
Paran.
Selah. His
glory
covered the
heavens, and the
earth was
full of his
praise.\FTNT{§}{{\FR 3.3 }
Teman: or, the south}
\VS{4}And
{\ADD{his}}
brightness was as the
light; he had
horns
{\ADD{coming}} out of his
hand: and there
{\ADD{was}} the
hiding of his
power.\FTNT{**}{{\FR 3.4 }
horns…: or, bright beams out of his side}
\VS{5}Before him
went the
pestilence, and burning
coals went
forth at his
feet.\FTNT{††}{{\FR 3.5 }
coals: or, diseases}
\VS{6}He
stood, and
measured the
earth: he
beheld, and drove
asunder the
nations; and the
everlasting
mountains were
scattered, the
perpetual
hills did
bow: his
ways
{\ADD{are}}
everlasting.
\VS{7}I
saw the
tents of
Cushan in
affliction:
{\ADD{and}} the
curtains of the
land of
Midian did
tremble.\FTNT{‡‡}{{\FR 3.7 }
Cushan: or, Ethiopia}\FTNT{§§}{{\FR 3.7 }
in…: or, under affliction, or, vanity}
\VS{8}Was the
{\ND{
LORD}}
displeased against the
rivers?
{\ADD{was}} thine
anger against the
rivers?
{\ADD{was}} thy
wrath against the
sea, that thou didst
ride upon thine
horses
{\ADD{and}} thy
chariots of
salvation?\FTNT{***}{{\FR 3.8 }
of…: or, were salvation?}
\VS{9}Thy
bow was made
quite
naked,
{\ADD{according}} to the
oaths of the
tribes,
{\ADD{even thy}}
word.
Selah. Thou didst
cleave the
earth with
rivers.\FTNT{†††}{{\FR 3.9 }
earth…: or, rivers of the earth}
\VS{10}The
mountains
saw thee,
{\ADD{and}} they
trembled: the
overflowing of the
water passed
by: the
deep
uttered his
voice,
{\ADD{and}} lifted
up his
hands on
high.
\VS{11}The
sun
{\ADD{and}}
moon stood
still in their
habitation: at the
light of thine
arrows they
went,
{\ADD{and}} at the
shining of thy
glittering
spear.\FTNT{‡‡‡}{{\FR 3.11 }
at the light…: or, thine arrows walked in the light}
\VS{12}Thou didst march
through the
land in
indignation, thou didst
thresh the
heathen in
anger.
\VS{13}Thou wentest
forth for the
salvation of thy
people,
{\ADD{even}} for
salvation with thine
anointed; thou
woundedst the
head out of the
house of the
wicked, by
discovering the
foundation unto the
neck.
Selah.\FTNT{§§§}{{\FR 3.13 }
by…: Heb. making naked}
\VS{14}Thou didst strike
through with his
staves the
head of his
villages: they came out as a
whirlwind to
scatter me: their
rejoicing
{\ADD{was}} as to
devour the
poor
secretly.\FTNT{*}{{\FR 3.14 }
came…: Heb. were tempestuous}
\VS{15}Thou didst
walk through the
sea with thine
horses,
{\ADD{through}} the
heap of
great
waters.\FTNT{†}{{\FR 3.15 }
heap: or, mud}
\par }{\PP \VS{16}When I
heard, my
belly
trembled; my
lips
quivered at the
voice:
rottenness
entered into my
bones, and I
trembled in myself, that I might
rest in the
day of
trouble: when he cometh
up unto the
people, he will invade them with his
troops.\FTNT{‡}{{\FR 3.16 }
invade…: or, cut them in pieces}
\par }{\PP \VS{17}¶ Although the fig
tree shall not
blossom, neither
{\ADD{shall}}
fruit
{\ADD{be}} in the
vines; the
labour of the
olive shall
fail, and the
fields shall
yield no
meat; the
flock shall be cut
off from the
fold, and
{\ADD{there shall be}} no
herd in the
stalls:\FTNT{§}{{\FR 3.17 }
fail: Heb. lie}
\VS{18}Yet I will
rejoice in the
{\ND{
LORD}}, I will
joy in the
God of my
salvation.
\VS{19}The
{\ND{
LORD}}
God
{\ADD{is}} my
strength, and he will
make my
feet like
hinds’
{\ADD{feet}}, and he will make me to
walk upon mine high
places. To the chief
singer on my stringed
instruments.\FTNT{**}{{\FR 3.19 }
chief…: or, overseer}\FTNT{††}{{\FR 3.19 }
stringed…: Heb. Neginoth}
\par }