\NormalFont\ShortTitle{Haggai}
{\MT Haggai

\par }\ChapOne{1}{\PP \VerseOne{1}In the
second
year of
Darius the
king, in the
sixth
month, in the
first
day of the
month, came the
word of the
{\ND{
LORD}}
by
Haggai the
prophet unto
Zerubbabel the
son of
Shealtiel,
governor of
Judah, and to
Joshua the
son of
Josedech, the
high
priest,
saying,
\FTNT{*}{{\FR 1.1 }
by: Heb. by the hand of}\FTNT{†}{{\FR 1.1 }
governor: or, captain}
\VS{2}Thus
speaketh the
{\ND{
LORD}} of
hosts,
saying, This
people
say, The
time is not
come, the
time that the
{\ND{
LORD’s}}
house should be
built.
\VS{3}Then came the
word of the
{\ND{
LORD}}
by
Haggai the
prophet,
saying,
\VS{4}{\ADD{Is it}}
time for you, O ye, to
dwell in your
cieled
houses, and this
house
{\ADD{lie}}
waste?
\VS{5}Now therefore thus
saith the
{\ND{
LORD}} of
hosts;
Consider your
ways.\FTNT{‡}{{\FR 1.5 }
Consider…: Heb. Set your heart on your ways}
\VS{6}Ye have
sown
much, and
bring in
little; ye
eat, but ye have not
enough; ye
drink, but ye are not filled with
drink; ye
clothe you, but there is none
warm; and he that earneth
wages earneth
wages
{\ADD{to put it}} into a
bag with
holes.\FTNT{§}{{\FR 1.6 }
with holes: Heb. pierced through}
\par }{\PP \VS{7}¶ Thus
saith the
{\ND{
LORD}} of
hosts;
Consider your
ways.\FTNT{**}{{\FR 1.7 }
Consider…: Heb. Set your heart on your ways}
\VS{8}Go
up to the
mountain, and
bring
wood, and
build the
house; and I will take
pleasure in it, and I will be
glorified,
saith the
{\ND{
LORD}}.
\VS{9}Ye
looked for
much, and, lo,
{\ADD{it came}} to
little; and when ye
brought
{\ADD{it}}
home, I did
blow upon it. Why?
saith the
{\ND{
LORD}} of
hosts.
Because of mine
house that
{\ADD{is}}
waste, and ye
run every
man unto his own
house.\FTNT{††}{{\FR 1.9 }
blow…: or, blow it away}
\VS{10}Therefore the
heaven over you is
stayed from
dew, and the
earth is
stayed
{\ADD{from}} her
fruit.
\VS{11}And I
called for a
drought upon the
land, and upon the
mountains, and upon the
corn, and upon the new
wine, and upon the
oil, and upon
{\ADD{that}} which the
ground bringeth
forth, and upon
men, and upon
cattle, and upon all the
labour of the
hands.
\par }{\PP \VS{12}¶ Then
Zerubbabel the
son of
Shealtiel, and
Joshua the
son of
Josedech, the
high
priest, with all the
remnant of the
people,
obeyed the
voice of the
{\ND{
LORD}} their
God, and the
words of
Haggai the
prophet, as the
{\ND{
LORD}} their
God had
sent him, and the
people did
fear
before the
{\ND{
LORD}}.
\VS{13}Then
spake
Haggai the
{\ND{
LORD’s}}
messenger in the
{\ND{
LORD’s}}
message unto the
people,
saying, I
{\ADD{am}} with you,
saith the
{\ND{
LORD}}.
\VS{14}And the
{\ND{
LORD}} stirred
up the
spirit of
Zerubbabel the
son of
Shealtiel,
governor of
Judah, and the
spirit of
Joshua the
son of
Josedech, the
high
priest, and the
spirit of all the
remnant of the
people; and they
came and
did
work in the
house of the
{\ND{
LORD}} of
hosts, their
God,
\VS{15}In the
four and
twentieth
day of the
sixth
month, in the
second
year of
Darius the
king.

\par }\Chap{2}{\PP \VerseOne{1}In the
seventh
{\ADD{month}}, in the
one and
twentieth
{\ADD{day}} of the
month, came the
word of the
{\ND{
LORD}}
by the
prophet
Haggai,
saying,
\FTNT{*}{{\FR 2.1 }
by: Heb. by the hand of}
\VS{2}Speak now to
Zerubbabel the
son of
Shealtiel,
governor of
Judah, and to
Joshua the
son of
Josedech, the
high
priest, and to the
residue of the
people,
saying,
\VS{3}Who
{\ADD{is}}
left among you that
saw this
house in her
first
glory? and how do ye
see it now?
{\ADD{is it}} not in your
eyes in
comparison of it as nothing?
\VS{4}Yet now be
strong, O
Zerubbabel,
saith the
{\ND{
LORD}}; and be
strong, O
Joshua,
son of
Josedech, the
high
priest; and be
strong, all ye
people of the
land,
saith the
{\ND{
LORD}}, and
work: for I
{\ADD{am}} with you,
saith the
{\ND{
LORD}} of
hosts:
\VS{5}{\ADD{According to}} the
word that I
covenanted with you when ye
came out of
Egypt, so my
spirit
remaineth
among you:
fear ye not.
\VS{6}For thus
saith the
{\ND{
LORD}} of
hosts; Yet
once, it
{\ADD{is}} a little
while, and I will
shake the
heavens, and the
earth, and the
sea, and the
dry
{\ADD{land;}}
\VS{7}And I will
shake all
nations, and the
desire of all
nations shall
come: and I will
fill this
house with
glory,
saith the
{\ND{
LORD}} of
hosts.
\VS{8}The
silver
{\ADD{is}} mine, and the
gold
{\ADD{is}} mine,
saith the
{\ND{
LORD}} of
hosts.
\VS{9}The
glory of this
latter
house shall be greater
than of the
former,
saith the
{\ND{
LORD}} of
hosts: and in this
place will I
give
peace,
saith the
{\ND{
LORD}} of
hosts.
\par }{\PP \VS{10}¶ In the
four and
twentieth
{\ADD{day}} of the
ninth
{\ADD{month}}, in the
second
year of
Darius, came the
word of the
{\ND{
LORD}}
by
Haggai the
prophet,
saying,
\VS{11}Thus
saith the
{\ND{
LORD}} of
hosts;
Ask now the
priests
{\ADD{concerning}} the
law,
saying,
\VS{12}If
one
bear
holy
flesh in the
skirt of his
garment, and with his
skirt do
touch
bread, or
pottage, or
wine, or
oil, or any
meat, shall it be
holy? And the
priests
answered and
said, No.
\VS{13}Then
said
Haggai, If
{\ADD{one that is}}
unclean by a dead
body
touch any of these, shall it be
unclean? And the
priests
answered and
said, It shall be
unclean.
\VS{14}Then
answered
Haggai, and
said, So
{\ADD{is}} this
people, and so
{\ADD{is}} this
nation
before me,
saith the
{\ND{
LORD}}; and so
{\ADD{is}} every
work of their
hands; and that which they
offer there
{\ADD{is}}
unclean.
\VS{15}And now, I pray you,
consider from this
day and
upward, from
before a
stone was
laid upon a
stone in the
temple of the
{\ND{
LORD}}:
\VS{16}Since those
{\ADD{days}} were, when
{\ADD{one}}
came to an
heap of
twenty
{\ADD{measures}}, there were
{\ADD{but}}
ten: when
{\ADD{one}}
came to the
pressfat for to draw
out
fifty
{\ADD{vessels}} out of the
press, there were
{\ADD{but}}
twenty.
\VS{17}I
smote you with
blasting and with
mildew and with
hail in all the
labours of your
hands; yet ye
{\ADD{turned}} not to me,
saith the
{\ND{
LORD}}.
\VS{18}Consider now from this
day and
upward, from the
four and
twentieth
day of the
ninth
{\ADD{month, even}} from the
day that the
foundation of the
{\ND{
LORD’s}}
temple was laid,
consider
{\ADD{it}}.
\VS{19}Is the
seed yet in the
barn? yea, as yet the
vine, and the fig
tree, and the
pomegranate, and the
olive
tree, hath not brought
forth: from this
day will I
bless
{\ADD{you}}.
\par }{\PP \VS{20}¶ And
again the
word of the
{\ND{
LORD}} came unto
Haggai in the
four and
twentieth
{\ADD{day}} of the
month,
saying,
\VS{21}Speak to
Zerubbabel,
governor of
Judah,
saying, I will
shake the
heavens and the
earth;
\VS{22}And I will
overthrow the
throne of
kingdoms, and I will
destroy the
strength of the
kingdoms of the
heathen; and I will
overthrow the
chariots, and those that
ride in them; and the
horses and their
riders shall come
down, every
one by the
sword of his
brother.
\VS{23}In that
day,
saith the
{\ND{
LORD}} of
hosts, will I
take thee, O
Zerubbabel, my
servant, the
son of
Shealtiel,
saith the
{\ND{
LORD}}, and will
make thee as a
signet: for I have
chosen thee,
saith the
{\ND{
LORD}} of
hosts.
\par }