\NormalFont\ShortTitle{Hebrews}
{\MT THE EPISTLE OF PAUL THE APOSTLE TO THE HEBREWS

\par }\ChapOne{1}{\PP \VerseOne{1}God,
who at sundry
times
and in divers
manners
spake in time
past unto the
fathers
by the
prophets,
\VS{2}Hath
in
these
last
days
spoken unto
us
by
{\ADD{his}}
Son,
whom he hath
appointed
heir of all
things,
by
whom
also he
made the
worlds;
\VS{3}Who
being the
brightness
of
{\ADD{his}}
glory,
and the express
image
of
his
person,
and
upholding all
things by the
word
of
his
power, when he
had
by
himself
purged
our
sins, sat
down
on the right
hand of the
Majesty
on
high;
\VS{4}Being
made so
much
better than the
angels,
as he hath by inheritance
obtained a more
excellent
name
than
they.
\VS{5}For unto
which of the
angels said
he at any
time,
Thou
art
my
Son, this
day
have
I
begotten
thee?
And
again,
I will
be
to
him a
Father,
and
he shall
be
to
me a
Son?
\VS{6}And
again,
when he bringeth
in the
firstbegotten
into the
world, he
saith,
And
let
all the
angels of
God
worship
him.
\VS{7}And
of the
angels he
saith,
Who
maketh
his
angels
spirits,
and
his
ministers a
flame of
fire.
\VS{8}But
unto the
Son
{\ADD{he saith}},
Thy
throne, O
God,
{\ADD{is}}
for
ever and
ever: a
sceptre of
righteousness
{\ADD{is}} the
sceptre
of
thy
kingdom.
\VS{9}Thou hast
loved
righteousness,
and
hated
iniquity;
therefore
God,
{\ADD{even}}
thy
God, hath
anointed
thee with the
oil of
gladness
above
thy
fellows.
\VS{10}And,
Thou,
Lord,
in the
beginning hast laid the
foundation of the
earth;
and the
heavens
are the
works
of
thine
hands:
\VS{11}They shall
perish;
but
thou
remainest;
and they
all shall wax
old as
doth a
garment;
\VS{12}And
as a
vesture shalt thou
fold
them
up,
and they shall be
changed:
but
thou
art the
same,
and
thy
years
shall
not
fail.
\VS{13}But
to
which of the
angels said
he at any
time,
Sit
on
my right
hand,
until I
make
thine
enemies
thy
footstool?
\VS{14}Are
they
not
all
ministering
spirits, sent
forth
to
minister
for them
who shall
be
heirs of
salvation?

\par }\Chap{2}{\PP \VerseOne{1}Therefore
we
ought to
give the more
earnest
heed to the things which we have
heard, lest at any
time we should
let
{\ADD{them}}
slip.
\VS{2}For
if the
word
spoken
by
angels
was
stedfast,
and
every
transgression
and
disobedience
received
a
just recompence of
reward;
\VS{3}How
shall
we
escape, if we
neglect so
great
salvation;
which
at the
first
began to be
spoken
by the
Lord, and was
confirmed
unto
us
by them that
heard
{\ADD{him}};
\VS{4}God
also
bearing
{\ADD{them}}
witness,
both with
signs
and
wonders,
and with
divers
miracles,
and
gifts of
the
Holy
Ghost, according
to his
own
will?
\par }{\PP \VS{5}For unto the
angels hath
he
not put in
subjection the
world to
come,
whereof we
speak.
\VS{6}But
one in a certain
place
testified,
saying,
What
is
man,
that thou art
mindful of
him?
or the
son of
man,
that thou
visitest
him?
\VS{7}Thou
madest
him a
little
lower
than the
angels; thou
crownedst
him with
glory
and
honour,
and didst
set
him
over the
works
of
thy
hands:
\VS{8}Thou hast
put all
things in
subjection
under
his
feet.
For
in
that he
put
all in subjection
under
him, he
left
nothing
{\ADD{that is}} not put
under
him.
But
now we
see not
yet all
things put
under
him.
\VS{9}But we
see Jesus,
who was
made a
little
lower
than the
angels
for the
suffering of
death,
crowned with
glory
and
honour;
that
he by the
grace of
God should
taste
death
for every
man.
\VS{10}For it
became
him,
for
whom
{\ADD{are}} all
things,
and
by
whom
{\ADD{are}} all
things, in
bringing
many
sons
unto
glory, to
make the
captain
of
their
salvation
perfect
through
sufferings.
\VS{11}For
both he that
sanctifieth
and they who are
sanctified
{\ADD{are}}
all
of
one:
for
which
cause he
is
not
ashamed to
call
them
brethren,
\VS{12}Saying, I will
declare
thy
name
unto
my
brethren,
in the
midst of the
church will I sing praise
unto
thee.
\VS{13}And
again,
I
will put my
trust
in
him.
And
again,
Behold
I
and the
children
which
God hath
given
me.
\VS{14}Forasmuch
then
as the
children are
partakers of
flesh
and
blood,
he
also
himself
likewise took
part of the
same;
that
through
death he might
destroy him that
had the
power of
death,
that
is, the
devil;
\VS{15}And
deliver
them
who through
fear of
death
were
all
their
lifetime
subject to
bondage.
\VS{16}For
verily he
took
not
on
{\ADD{him the nature of}}
angels;
but he took
on
{\ADD{him}} the seed
of
Abraham.
\VS{17}Wherefore
in all
things it behoved
him to be made
like
unto
{\ADD{his}}
brethren,
that he might
be
a
merciful
and
faithful high
priest in
things
{\ADD{pertaining}}
to
God,
to make reconciliation
for the
sins of the
people.
\VS{18}For
in
that
he
himself hath
suffered being
tempted, he is
able to
succour them that are
tempted.

\par }\Chap{3}{\PP \VerseOne{1}Wherefore,
holy
brethren,
partakers of the
heavenly
calling,
consider the
Apostle
and High
Priest
of
our
profession,
Christ
Jesus;
\VS{2}Who
was
faithful to him that
appointed
him,
as
also
Moses
{\ADD{was faithful}}
in
all
his
house.
\VS{3}For
this
{\ADD{man}} was counted
worthy
of
more
glory
than
Moses,
inasmuch
as he who hath
builded the
house
hath
more
honour than the
house.
\VS{4}For
every
house is
builded
by
some
{\ADD{man}};
but he that
built all
things
{\ADD{is}}
God.
\VS{5}And
Moses
verily
{\ADD{was}}
faithful
in
all
his
house,
as a
servant,
for a
testimony of those things which were to be spoken
after;
\VS{6}But
Christ
as a
son
over his
own
house;
whose
house are
we,
if
we hold
fast the
confidence
and the
rejoicing of the
hope
firm
unto the
end.
\VS{7}Wherefore (as
the
Holy
Ghost
saith, To
day
if ye will
hear
his
voice,
\VS{8}Harden
not
your
hearts,
as
in the
provocation,
in the
day of
temptation
in the
wilderness:
\VS{9}When
your
fathers
tempted
me,
proved
me,
and
saw
my
works
forty
years.
\VS{10}Wherefore I was grieved
with
that
generation,
and
said, They
do
alway
err
in
{\ADD{their}}
heart;
and
they
have
not
known
my
ways.
\VS{11}So I
sware
in
my
wrath, They
shall
not
enter
into
my
rest.)
\VS{12}Take
heed,
brethren,
lest there
be
in
any of
you
an
evil
heart of
unbelief,
in
departing
from the
living
God.
\VS{13}But
exhort one
another
daily,
while it is
called To
day;
lest
any
of
you be
hardened through the
deceitfulness of
sin.
\VS{14}For we are
made
partakers of
Christ,
if we
hold the
beginning of our
confidence
stedfast
unto the
end;
\VS{15}While it is
said, To
day
if ye will
hear
his
voice,
harden
not
your
hearts,
as
in the
provocation.
\VS{16}For
some, when they had
heard, did
provoke:
howbeit
not
all that
came out
of
Egypt
by
Moses.
\VS{17}But with
whom was he
grieved
forty
years?
{\ADD{was it}}
not with them that had
sinned,
whose
carcases
fell
in the
wilderness?
\VS{18}And to
whom sware
he
that they
should
not
enter
into
his
rest,
but to them that believed
not?
\VS{19}So we
see
that they
could
not enter
in because
of
unbelief.

\par }\Chap{4}{\PP \VerseOne{1}Let
us
therefore
fear,
lest, a
promise being
left
{\ADD{us}} of
entering
into
his
rest,
any
of
you should
seem to come short of
it.
\VS{2}For
unto us
was the gospel
preached, as well
as unto
them:
but the
word
preached
did
not
profit
them,
not being mixed
with
faith in them that
heard
{\ADD{it}}.
\VS{3}For
we which have
believed do
enter
into
rest,
as he
said,
As I have
sworn
in
my
wrath,
if they shall
enter
into
my
rest:
although the
works were
finished
from the
foundation of the
world.
\VS{4}For he
spake in a certain
place
of the
seventh
{\ADD{day}} on this
wise,
And
God did
rest
the
seventh
day
from
all
his
works.
\VS{5}And
in
this
{\ADD{place}}
again,
If they shall
enter
into
my
rest.
\VS{6}Seeing
therefore it
remaineth that
some must
enter
therein,
and they to whom it
was
first
preached
entered
not
in because
of
unbelief:
\VS{7}Again, he
limiteth a
certain
day,
saying
in
David, To
day,
after so
long a
time;
as it is
said, To
day
if ye will
hear
his
voice,
harden
not
your
hearts.
\VS{8}For
if
Jesus had
given
them
rest,
then would
he
not
afterward have
spoken
of
another
day.
\VS{9}There
remaineth
therefore a
rest to the
people of
God.
\VS{10}For he that is
entered
into
his
rest,
he
also hath
ceased
from his
own
works,
as
God
{\ADD{did}}
from
his.
\VS{11}Let us
labour
therefore to
enter
into
that
rest,
lest any
man
fall
after the
same
example of
unbelief.
\VS{12}For the
word of
God
{\ADD{is}}
quick,
and
powerful,
and
sharper
than
any
twoedged
sword,
piercing
even
to the dividing
asunder of
soul
and
spirit,
and of the
joints
and
marrow,
and
{\ADD{is}} a
discerner of the
thoughts
and
intents of the
heart.
\VS{13}Neither is
there any
creature that is not
manifest
in
his
sight:
but all
things
{\ADD{are}}
naked
and
opened unto the
eyes of
him
with
whom we
have to
do.
\VS{14}Seeing
then that we
have
a
great high
priest, that is passed
into the
heavens,
Jesus the
Son of
God, let us hold
fast
{\ADD{our}}
profession.
\VS{15}For we
have
not an high priest
which
cannot be touched with the feeling
of
our
infirmities;
but
was
in all
points
tempted like
as
{\ADD{we are, yet}}
without
sin.
\VS{16}Let
us
therefore
come
boldly unto the
throne of
grace,
that we may
obtain
mercy,
and
find
grace
to
help in time of
need.

\par }\Chap{5}{\PP \VerseOne{1}For
every high
priest
taken from
among
men is
ordained
for
men
in
things
{\ADD{pertaining}}
to
God,
that he may
offer
both
gifts
and
sacrifices
for
sins:
\VS{2}Who
can have compassion
on the
ignorant,
and
on them that are out of the
way; for
that
he
himself
also is compassed
with
infirmity.
\VS{3}And by
reason
hereof he
ought,
as
for the
people,
so
also
for
himself, to
offer
for
sins.
\VS{4}And
no
man
taketh this
honour unto
himself,
but he that is
called
of
God,
as
{\ADD{was}}
Aaron.
\VS{5}So
also
Christ
glorified
not
himself to be
made an high
priest;
but he that
said
unto
him,
Thou
art
my
Son, to
day
have
I
begotten
thee.
\VS{6}As he
saith
also
in
another
{\ADD{place}},
Thou
{\ADD{art}} a
priest
for
ever
after the
order of
Melchisedec.
\VS{7}Who
in the
days
of
his
flesh, when he had offered
up
prayers
and
supplications
with
strong
crying
and
tears
unto him that was
able to
save
him
from
death,
and was
heard
in that he
feared;
\VS{8}Though he
were a
Son,
yet learned
he
obedience
by the things
which he
suffered;
\VS{9}And being made
perfect, he
became the
author
of
eternal
salvation unto
all them that
obey
him;
\VS{10}Called
of
God an high
priest
after the
order of
Melchisedec.
\par }{\PP \VS{11}Of
whom we
have
many
things to
say,
and hard to be
uttered,
seeing ye
are
dull of
hearing.
\VS{12}For
when
for the
time ye
ought to
be
teachers, ye
have
need that
one
teach
you
again
which
{\ADD{be}}
the
first
principles of the
oracles of
God;
and are
become such as
have
need of
milk,
and
not
of
strong
meat.
\VS{13}For every
one that
useth
milk
{\ADD{is}}
unskilful in the
word of
righteousness:
for he
is a
babe.
\VS{14}But
strong
meat
belongeth to them that are of full
age,
{\ADD{even}} those
who by reason
of
use
have their
senses
exercised
to
discern
both
good
and
evil.

\par }\Chap{6}{\PP \VerseOne{1}Therefore
leaving the principles
of the
doctrine of
Christ, let us go
on
unto
perfection;
not
laying
again the
foundation of
repentance
from
dead
works,
and of
faith
toward
God,
\VS{2}Of the
doctrine of
baptisms,
and of laying
on of
hands,
and of
resurrection of the
dead,
and
of
eternal
judgment.
\VS{3}And
this will we
do,
if
God
permit.
\VS{4}For
{\ADD{it is}}
impossible for those who
were
once
enlightened,
and have
tasted of
the
heavenly
gift,
and were
made
partakers of
the
Holy
Ghost,
\VS{5}And have
tasted
the
good
word of
God,
and the
powers of the
world to
come,
\VS{6}If they shall fall
away, to renew
them
again
unto
repentance; seeing they
crucify to
themselves the
Son of
God
afresh,
and
put
{\ADD{him}} to an open
shame.
\VS{7}For the earth
which drinketh
in the
rain that
cometh
oft
upon
it,
and bringeth
forth
herbs
meet for
them
by
whom it is
dressed,
receiveth
blessing
from
God:
\VS{8}But that which
beareth
thorns
and
briers
{\ADD{is}}
rejected,
and
{\ADD{is}} nigh
unto
cursing;
whose
end
{\ADD{is}}
to be
burned.
\VS{9}But,
beloved, we are
persuaded better
things
of
you,
and things that
accompany
salvation,
though
we
thus
speak.
\VS{10}For
God
{\ADD{is}}
not
unrighteous to
forget
your
work
and
labour of
love,
which ye have
shewed
toward
his
name, in that ye have ministered
to the
saints,
and do
minister.
\VS{11}And we desire
that every
one of
you do
shew
the
same
diligence
to the full
assurance of
hope
unto the
end:
\VS{12}That ye
be
not
slothful,
but
followers of them
who
through
faith
and
patience
inherit the
promises.
\VS{13}For
when
God made
promise to
Abraham,
because he
could
swear
by
no
greater, he
sware
by
himself,
\VS{14}Saying,
Surely
blessing I will
bless
thee,
and
multiplying I will
multiply
thee.
\VS{15}And
so, after he had patiently
endured, he
obtained the
promise.
\VS{16}For
men
verily
swear
by the
greater:
and an
oath
for
confirmation
{\ADD{is}} to
them an
end
of
all
strife.
\VS{17}Where
in
God,
willing more
abundantly to
shew unto the
heirs of
promise the
immutability
of
his
counsel,
confirmed
{\ADD{it}} by an
oath:
\VS{18}That
by
two
immutable
things,
in
which
{\ADD{it was}}
impossible for
God to
lie, we might
have
a
strong
consolation, who have fled for
refuge to lay
hold upon the
hope set before
us:
\VS{19}Which
{\ADD{hope}} we
have
as an
anchor of the
soul,
both
sure
and
stedfast,
and which
entereth
into that
within the
veil;
\VS{20}Whither the
forerunner
is
for
us
entered,
{\ADD{even}}
Jesus,
made an high
priest
for
ever
after the
order of
Melchisedec.

\par }\Chap{7}{\PP \VerseOne{1}For
this
Melchisedec,
king of
Salem,
priest of
the most
high
God, who
met
Abraham
returning
from the
slaughter of the
kings,
and
blessed
him;
\VS{2}To
whom
also
Abraham
gave a tenth
part
of
all;
first being by
interpretation
King of
righteousness,
and after
that
also
King of
Salem,
which
is,
King of
peace;
\VS{3}Without
father, without
mother, without
descent,
having
neither
beginning of
days,
nor
end of
life;
but made
like unto the
Son of
God;
abideth a
priest
continually.
\VS{4}Now
consider how
great this
man
{\ADD{was}}, unto
whom
even the
patriarch
Abraham
gave the
tenth
of the
spoils.
\VS{5}And
verily they that
are
of the
sons of
Levi, who
receive the office of the
priesthood,
have a
commandment to take tithes
of the
people according
to the
law,
that
is,
of
their
brethren,
though they come
out
of the
loins of
Abraham:
\VS{6}But he whose descent
is
not
counted
from
them received
tithes of
Abraham,
and
blessed him that
had the
promises.
\VS{7}And
without
all
contradiction the
less is
blessed
of the
better.
\VS{8}And
here
men that
die
receive
tithes;
but
there
he
{\ADD{receiveth them}}, of whom it is
witnessed
that he
liveth.
\VS{9}And
as I may so
say,
Levi
also, who
receiveth
tithes, payed
tithes
in
Abraham.
\VS{10}For he
was
yet
in the
loins of his
father,
when
Melchisedec
met
him.
\VS{11}If
therefore
perfection
were
by the
Levitical
priesthood, (for
under
it the
people received the
law,)
what
further
need
{\ADD{was there}}
that
another
priest should
rise
after the
order of
Melchisedec,
and
not be
called
after the
order of
Aaron?
\VS{12}For the
priesthood being
changed, there is
made
of
necessity a
change
also of the
law.
\VS{13}For
he
of
whom these
things are
spoken
pertaineth
to
another
tribe,
of
which no
man gave
attendance at the
altar.
\VS{14}For
{\ADD{it is}}
evident
that
our
Lord
sprang out
of
Juda;
of
which
tribe
Moses
spake
nothing
concerning
priesthood.
\VS{15}And it
is
yet far
more
evident: for
that
after the
similitude of
Melchisedec there
ariseth
another
priest,
\VS{16}Who is
made,
not
after the
law of
a
carnal
commandment,
but
after the
power of
an
endless
life.
\VS{17}For he
testifieth,
Thou
{\ADD{art}} a
priest
for
ever
after the
order of
Melchisedec.
\VS{18}For there
is
verily a
disannulling of the
commandment going
before
for the
weakness
and
unprofitableness
thereof.
\VS{19}For the
law
made
nothing
perfect,
but the bringing
in of
a
better
hope
{\ADD{did}};
by the
which we draw
nigh unto
God.
\VS{20}And
inasmuch
as
not
without an
oath
{\ADD{he was made priest}}:
\VS{21}(For those
priests
were
made
without an
oath;
but
this
with an
oath
by him that
said
unto
him, The
Lord
sware
and
will
not
repent,
Thou
{\ADD{art}} a
priest
for
ever
after the
order of
Melchisedec:)
\VS{22}By so
much
was
Jesus
made a
surety of
a
better
testament.
\VS{23}And
they
truly
were
many
priests,
because they were not
suffered to
continue by reason of
death:
\VS{24}But
this
{\ADD{man}},
because
he
continueth
ever,
hath
an
unchangeable
priesthood.
\VS{25}Wherefore he is
able
also to
save
them
to the
uttermost that
come unto
God
by
him, seeing
he
ever
liveth
to make
intercession
for
them.
\VS{26}For
such an high
priest
became
us,
{\ADD{who is}}
holy,
harmless,
undefiled,
separate
from
sinners,
and
made
higher than the
heavens;
\VS{27}Who
needeth
not
daily,
as those high
priests, to offer
up
sacrifice,
first
for his
own
sins, and
then for the
people’s:
for
this he
did
once, when he offered
up
himself.
\VS{28}For the
law
maketh
men high
priests which
have
infirmity;
but the
word of the
oath, which
was
since the
law,
{\ADD{maketh}} the
Son, who is
consecrated
for
evermore.

\par }\Chap{8}{\PP \VerseOne{1}Now
of the things which we have
spoken
{\ADD{this is}} the
sum: We
have
such an high
priest,
who is
set
on the right
hand of the
throne of the
Majesty
in the
heavens;
\VS{2}A
minister of the
sanctuary,
and of
the
true
tabernacle,
which the
Lord
pitched,
and
not
man.
\VS{3}For
every high
priest is
ordained
to
offer
gifts
and
sacrifices:
wherefore
{\ADD{it is}} of
necessity
that this
man
have
somewhat
also to
offer.
\VS{4}For
if he
were
on
earth, he
should
not
be a
priest, seeing that there
are
priests that
offer
gifts
according to the
law:
\VS{5}Who
serve unto the
example
and
shadow of heavenly
things,
as
Moses was admonished of
God when he was
about to
make the
tabernacle:
for,
See, saith
he,
{\ADD{that}} thou
make all
things according
to the
pattern
shewed to
thee
in the
mount.
\VS{6}But
now hath he obtained
a more
excellent
ministry, by how
much
also he
is the
mediator of a
better
covenant,
which was
established
upon
better
promises.
\VS{7}For
if
that
first
{\ADD{covenant}} had
been
faultless, then should
no
place have been
sought for the
second.
\VS{8}For finding
fault with
them, he
saith,
Behold, the
days
come,
saith the
Lord,
when I will
make
a
new
covenant
with the
house of
Israel
and
with the
house of
Judah:
\VS{9}Not according
to the
covenant
that I
made
with
their
fathers
in the
day
when
I
took
them by the
hand to
lead
them
out
of the
land of
Egypt;
because
they
continued
not
in
my
covenant, and
I
regarded
them
not,
saith the
Lord.
\VS{10}For
this
{\ADD{is}} the
covenant
that I will
make with the
house of
Israel
after
those
days,
saith the
Lord; I
will
put
my
laws
into
their
mind,
and
write
them
in
their
hearts:
and I will
be to
them a
God,
and
they shall
be to
me a
people:
\VS{11}And they
shall
not
teach every
man
his
neighbour,
and every
man
his
brother,
saying,
Know the
Lord:
for
all shall
know
me,
from the
least
to the
greatest.
\VS{12}For I will
be
merciful
to
their
unrighteousness,
and
their
sins
and
their
iniquities will I
remember
no
more.
\VS{13}In that he
saith, A
new
{\ADD{covenant}}, he hath
made the
first
old.
Now that which
decayeth
and waxeth
old
{\ADD{is}}
ready to vanish
away.

\par }\Chap{9}{\PP \VerseOne{1}Then
verily the
first
{\ADD{covenant}}
had
also
ordinances of divine
service,
and
a
worldly
sanctuary.
\VS{2}For there
was a
tabernacle
made; the
first,
wherein
{\ADD{was}} the
candlestick,
and the
table,
and the
shewbread;
which is
called the
sanctuary.
\VS{3}And
after
the
second
veil, the
tabernacle which is
called the
Holiest of
all;
\VS{4}Which
had
the
golden
censer,
and the
ark of the
covenant
overlaid round
about with
gold,
wherein
{\ADD{was}}
the
golden
pot that
had
manna,
and
Aaron’s
rod that
budded,
and the
tables of the
covenant;
\VS{5}And
over
it the
cherubims of
glory
shadowing the
mercyseat;
of
which we
can
not
now
speak
particularly.
\VS{6}Now
when these
things
were
thus
ordained, the
priests
went
always
into
the
first
tabernacle,
accomplishing the
service
{\ADD{of God}}.
\VS{7}But
into the
second
{\ADD{went}} the high
priest
alone
once every
year,
not
without
blood,
which he
offered
for
himself,
and
{\ADD{for}} the
errors of the
people:
\VS{8}The
Holy
Ghost
this
signifying, that the
way into the holiest of
all
was not
yet made
manifest, while
as
the
first
tabernacle
was
yet
standing:
\VS{9}Which
{\ADD{was}} a
figure
for the
time then
present,
in
which were
offered
both
gifts
and
sacrifices,
that
could
not
make him that did the
service
perfect, as pertaining
to the
conscience;
\VS{10}{\ADD{Which stood}}
only
in
meats
and
drinks,
and
divers
washings,
and
carnal
ordinances,
imposed
{\ADD{on them}}
until the
time of
reformation.
\VS{11}But
Christ being
come an high
priest of good
things to
come,
by
a
greater
and more
perfect
tabernacle,
not made with
hands,
that is to
say,
not
of
this
building;
\VS{12}Neither
by the
blood of
goats
and
calves,
but
by his
own
blood he entered
in
once
into the holy
place, having
obtained
eternal
redemption
{\ADD{for us}}.
\VS{13}For
if the
blood of
bulls
and of
goats,
and the
ashes of an
heifer
sprinkling the
unclean,
sanctifieth
to the
purifying of the
flesh:
\VS{14}How
much
more
shall the
blood of
Christ,
who
through
the
eternal
Spirit
offered
himself without
spot to
God,
purge
your
conscience
from
dead
works
to
serve
the
living
God?
\VS{15}And
for this
cause he
is the
mediator of
the
new
testament,
that by
means of
death,
for the
redemption of the
transgressions
{\ADD{that were}}
under
the
first
testament,
they which are
called might
receive the
promise
of
eternal
inheritance.
\VS{16}For
where a
testament
{\ADD{is}}, there must also of necessity be the
death of the
testator.
\VS{17}For a
testament
{\ADD{is}} of
force
after men are
dead:
otherwise it is
of
no
strength at
all
while the
testator
liveth.
\VS{18}Whereupon
neither the
first
{\ADD{testament}} was
dedicated
without
blood.
\VS{19}For
when
Moses had
spoken
every
precept
to
all the
people according
to the
law, he
took the
blood of
calves
and of
goats,
with
water,
and
scarlet
wool,
and
hyssop, and
sprinkled
both the
book,
and
all the
people,
\VS{20}Saying,
This
{\ADD{is}} the
blood of the
testament
which
God hath
enjoined
unto
you.
\VS{21}Moreover he
sprinkled with
blood
both the
tabernacle,
and
all the
vessels of the
ministry.
\VS{22}And
almost all
things
are
by the
law
purged
with
blood;
and
without shedding of
blood
is
no
remission.
\VS{23}{\ADD{It was}}
therefore necessary
that the
patterns of
things
in the
heavens should be
purified with
these;
but the heavenly
things
themselves
with
better
sacrifices
than
these.
\VS{24}For
Christ
is
not
entered
into the holy
places made with
hands,
{\ADD{which are}} the
figures of the
true;
but
into
heaven
itself,
now to
appear in the
presence of
God
for
us:
\VS{25}Nor
yet
that he should
offer
himself
often,
as the high
priest
entereth
into the holy
place
every
year
with
blood of
others;
\VS{26}For
then
must
he
often have
suffered
since the
foundation of the
world:
but
now
once
in the
end of the
world hath he
appeared
to put
away
sin
by the
sacrifice of
himself.
\VS{27}And
as it is
appointed unto
men
once to
die,
but
after
this the
judgment:
\VS{28}So
Christ
was
once
offered
to
bear the
sins of
many; and unto them that look
for
him shall he
appear the second
time
without
sin
unto
salvation.

\par }\Chap{10}{\PP \VerseOne{1}For the
law
having a
shadow of good
things to
come,
{\ADD{and}}
not
the
very
image of the
things,
can
never
with
those
sacrifices
which they
offered year
by
year
continually
make the
comers thereunto
perfect.
\VS{2}For
then would
they
not have
ceased to be
offered?
because that the
worshippers
once
purged should have
had
no
more
conscience of
sins.
\VS{3}But
in
those
{\ADD{sacrifices there is}} a remembrance
again
{\ADD{made}} of
sins
every
year.
\VS{4}For
{\ADD{it is}} not
possible
that the
blood of
bulls
and of
goats should take
away
sins.
\VS{5}Wherefore when he
cometh
into the
world, he
saith,
Sacrifice
and
offering thou
wouldest
not,
but a
body hast thou
prepared
me:
\VS{6}In burnt
offerings
and
{\ADD{sacrifices}}
for
sin thou hast
had
no
pleasure.
\VS{7}Then said
I,
Lo, I
come (in the
volume of the
book it is
written
of
me,) to
do
thy
will, O
God.
\VS{8}Above
when he
said,
Sacrifice
and
offering
and burnt
offerings
and
{\ADD{offering}}
for
sin thou
wouldest
not,
neither hadst
pleasure
{\ADD{therein}};
which are
offered
by the
law;
\VS{9}Then said
he,
Lo, I
come to
do
thy
will, O
God. He taketh
away the
first,
that he may
establish the
second.
\VS{10}By
the
which
will we
are
sanctified
through the
offering of the
body of
Jesus
Christ
once
{\ADD{for all}}.
\VS{11}And
every
priest
standeth
daily
ministering
and
offering
oftentimes
the
same
sacrifices,
which
can
never take
away
sins:
\VS{12}But this
man, after he had
offered
one
sacrifice
for
sins
for
ever, sat
down
on the right
hand of
God;
\VS{13}From
henceforth
expecting
till
his
enemies be
made
his
footstool.
\VS{14}For by
one
offering he hath
perfected
for
ever them that are
sanctified.
\VS{15}{\ADD{Whereof}}
the
Holy
Ghost
also is a
witness to
us:
for
after that he had said
before,
\VS{16}This
{\ADD{is}} the
covenant
that I will
make
with
them
after
those
days,
saith the
Lord, I will
put
my
laws
into
their
hearts,
and
in
their
minds will I
write
them;
\VS{17}And
their
sins
and
iniquities will I
remember
no
more.
\VS{18}Now
where
remission of
these
{\ADD{is, there is}} no
more
offering
for
sin.
\par }{\PP \VS{19}Having
therefore,
brethren,
boldness
to enter
into the
holiest
by the
blood of
Jesus,
\VS{20}By
a
new
and
living
way,
which he hath
consecrated for
us,
through the
veil,
that is to
say,
his
flesh;
\VS{21}And
{\ADD{having}}
an
high
priest
over the
house of
God;
\VS{22}Let us draw
near
with
a
true
heart
in full
assurance of
faith,
having our
hearts
sprinkled
from
an
evil
conscience,
and our
bodies
washed
with
pure
water.
\VS{23}Let us hold
fast the
profession
of
{\ADD{our}}
faith without
wavering; (for
he
{\ADD{is}}
faithful that
promised;)
\VS{24}And let us
consider one
another
to
provoke unto
love
and
to
good
works:
\VS{25}Not
forsaking the
assembling of
ourselves
together,
as the
manner of
some
{\ADD{is}};
but
exhorting
{\ADD{one another}}:
and so
much the
more,
as ye
see the
day
approaching.
\VS{26}For
if
we
sin
wilfully
after that we have
received the
knowledge of the
truth, there
remaineth no
more
sacrifice
for
sins,
\VS{27}But a
certain
fearful looking
for of
judgment
and
fiery
indignation,
which
shall
devour the
adversaries.
\VS{28}He
that
despised
Moses’
law
died
without
mercy
under
two
or
three
witnesses:
\VS{29}Of how
much
sorer
punishment, suppose
ye, shall he be thought
worthy, who hath trodden under
foot the
Son of
God,
and hath
counted the
blood of the
covenant,
wherewith he was
sanctified, an unholy
thing,
and hath done despite
unto the
Spirit of
grace?
\VS{30}For we
know him that hath
said,
Vengeance
{\ADD{belongeth}} unto
me,
I will
recompense,
saith the
Lord.
And
again, The
Lord shall
judge
his
people.
\VS{31}{\ADD{It is}} a fearful
thing to
fall
into the
hands of
the
living
God.
\VS{32}But call to
remembrance
the
former
days,
in
which, after ye were
illuminated, ye
endured
a
great
fight of
afflictions;
\VS{33}Partly, whilst ye were made a
gazingstock
both by
reproaches
and
afflictions;
and
partly, whilst ye
became
companions of them that
were
so
used.
\VS{34}For ye had
compassion of
me in my
bonds,
and
took
joyfully the
spoiling
of
your
goods,
knowing
in
yourselves that ye
have
in
heaven
a
better
and an
enduring
substance.
\VS{35}Cast
not
away
therefore
your
confidence,
which
hath
great recompence of
reward.
\VS{36}For ye
have
need of
patience,
that, after ye have
done the
will of
God, ye might
receive the
promise.
\VS{37}For
yet a
little
while, and he that shall
come will
come,
and
will
not
tarry.
\VS{38}Now the
just shall
live
by
faith:
but
if
{\ADD{any man}} draw
back,
my
soul shall
have
no
pleasure
in
him.
\VS{39}But
we
are
not of them who draw
back
unto
perdition;
but of them that
believe
to the
saving of the
soul.

\par }\Chap{11}{\PP \VerseOne{1}Now
faith
is the
substance of things hoped
for, the
evidence of
things
not
seen.
\VS{2}For
by
it the
elders obtained a good
report.
\VS{3}Through
faith we
understand
that the
worlds were
framed by the
word of
God, so
that things which are
seen
were
not
made
of things which do
appear.
\VS{4}By
faith
Abel
offered unto
God
a more
excellent
sacrifice
than
Cain,
by
which he obtained
witness that he
was
righteous,
God
testifying
of
his
gifts:
and
by
it he being
dead
yet
speaketh.
\VS{5}By
faith
Enoch was
translated
that he
should
not
see
death;
and
was
not
found,
because
God had
translated
him:
for
before
his
translation he had this
testimony, that he
pleased
God.
\VS{6}But
without
faith
{\ADD{it is}}
impossible to
please
{\ADD{him}}:
for he that
cometh to
God
must
believe
that he
is,
and
{\ADD{that}} he
is a
rewarder of them that diligently
seek
him.
\VS{7}By
faith
Noah, being warned of
God
of
things
not
seen as
yet, moved with
fear,
prepared an
ark
to the
saving
of
his
house;
by the
which he
condemned the
world,
and
became
heir of the
righteousness which is
by
faith.
\VS{8}By
faith
Abraham, when he was
called to go
out
into a
place
which he should
after
receive
for an
inheritance,
obeyed;
and he went
out,
not
knowing
whither he
went.
\VS{9}By
faith he
sojourned
in the
land of
promise,
as
{\ADD{in}} a strange
country,
dwelling
in
tabernacles
with
Isaac
and
Jacob, the heirs with
him of
the
same
promise:
\VS{10}For he looked
for a
city which
hath
foundations,
whose
builder
and
maker
{\ADD{is}}
God.
\VS{11}Through
faith
also
Sara
herself
received
strength
to
conceive
seed,
and was delivered of a
child when she was
past
age,
because she
judged
him
faithful who had
promised.
\VS{12}Therefore sprang
there
even
of
one,
and
him as good as
dead,
{\ADD{so many}}
as the
stars of the
sky in
multitude,
and
as the
sand which
is
by
the
sea
shore
innumerable.
\VS{13}These
all
died
in
faith,
not having
received the
promises,
but having
seen
them afar
off,
and were persuaded
of
{\ADD{them}},
and
embraced
{\ADD{them}},
and
confessed
that they
were
strangers
and
pilgrims
on the
earth.
\VS{14}For they that
say such
things declare
plainly
that they
seek a
country.
\VS{15}And
truly,
if they had been
mindful of
that
{\ADD{country}}
from
whence they came
out, they might have
had
opportunity to have
returned.
\VS{16}But
now they
desire a
better
{\ADD{country}},
that
is, an
heavenly:
wherefore
God
is
not
ashamed to be
called
their
God:
for he hath
prepared for
them a
city.
\VS{17}By
faith
Abraham, when he was
tried, offered
up
Isaac:
and he that had
received the
promises offered
up his only
begotten
{\ADD{son}},
\VS{18}Of
whom it was
said,
That
in
Isaac
shall
thy
seed be
called:
\VS{19}Accounting
that
God
{\ADD{was}}
able to
raise
{\ADD{him}}
up,
even
from the
dead; from
whence
also he
received
him
in a
figure.
\VS{20}By
faith
Isaac
blessed
Jacob
and
Esau
concerning things to
come.
\VS{21}By
faith
Jacob, when he was a
dying,
blessed
both the
sons of
Joseph;
and
worshipped,
{\ADD{leaning}}
upon the
top
of
his
staff.
\VS{22}By
faith
Joseph, when he
died, made
mention
of the
departing of the
children of
Israel;
and gave
commandment
concerning
his
bones.
\VS{23}By
faith
Moses, when he was
born, was
hid three
months
of
his
parents,
because they
saw
{\ADD{he was}}
a
proper
child;
and they
were
not
afraid
of the
king’s
commandment.
\VS{24}By
faith
Moses, when he was
come to
years,
refused to be
called the
son
of
Pharaoh’s
daughter;
\VS{25}Choosing
rather to suffer affliction
with the
people of
God,
than to
enjoy the
pleasures of
sin for a
season;
\VS{26}Esteeming the
reproach of
Christ
greater
riches than the
treasures
in
Egypt:
for he had
respect
unto the recompence of the
reward.
\VS{27}By
faith he
forsook
Egypt,
not
fearing the
wrath of the
king:
for he
endured,
as
seeing him who is
invisible.
\VS{28}Through
faith he
kept the
passover,
and the
sprinkling of
blood,
lest he that
destroyed the
firstborn should
touch
them.
\VS{29}By
faith they passed
through
the
Red
sea
as
by
dry
{\ADD{land}}:
which the
Egyptians
assaying to
do were
drowned.
\VS{30}By
faith the
walls of
Jericho fell
down, after they were
compassed
about
seven
days.
\VS{31}By
faith the
harlot
Rahab
perished
not
with them that believed
not, when she had
received the
spies
with
peace.
\VS{32}And
what shall
I
more
say?
for the
time would
fail
me to
tell
of
Gedeon,
and
{\ADD{of}}
Barak,
and
{\ADD{of}}
Samson,
and
{\ADD{of}}
Jephthae;
{\ADD{of}}
David
also,
and
Samuel,
and
{\ADD{of}} the
prophets:
\VS{33}Who
through
faith
subdued
kingdoms,
wrought
righteousness,
obtained
promises,
stopped the
mouths of
lions,
\VS{34}Quenched the
violence of
fire,
escaped the
edge of the
sword, out
of
weakness were made
strong,
waxed
valiant
in
fight, turned to
flight the
armies of the
aliens.
\VS{35}Women
received
their
dead raised to life
again:
and
others were
tortured,
not
accepting
deliverance;
that they might
obtain
a
better
resurrection:
\VS{36}And
others
had
trial
of
{\ADD{cruel}}
mockings
and
scourgings,
yea,
moreover of
bonds
and
imprisonment:
\VS{37}They were
stoned, they were sawn
asunder, were
tempted, were
slain with the
sword: they wandered
about
in
sheepskins
and
goatskins; being
destitute,
afflicted,
tormented;
\VS{38}(Of
whom the
world
was
not
worthy:) they
wandered
in
deserts,
and
{\ADD{in}}
mountains,
and
{\ADD{in}}
dens
and
caves of the
earth.
\VS{39}And
these
all, having obtained a good
report
through
faith,
received
not the
promise:
\VS{40}God having
provided
some
better
thing
for
us,
that they
without
us
should
not be made
perfect.

\par }\Chap{12}{\PP \VerseOne{1}Wherefore
seeing
we
also are compassed about
with so
great a
cloud of
witnesses, let us lay
aside
every
weight,
and the
sin which doth so easily
beset
{\ADD{us}}, and let us
run
with
patience the
race that is set
before
us,
\VS{2}Looking
unto
Jesus the
author
and
finisher
of
{\ADD{our}}
faith;
who
for the
joy that was set
before
him
endured the
cross,
despising the
shame,
and is set
down
at the right
hand of the
throne of
God.
\VS{3}For
consider him that
endured
such
contradiction
of
sinners
against
himself,
lest ye be
wearied and
faint
in
your
minds.
\VS{4}Ye
have not
yet
resisted
unto
blood,
striving
against
sin.
\VS{5}And ye have
forgotten the
exhortation
which
speaketh unto
you
as unto
children,
My
son,
despise
not
thou the
chastening of the
Lord,
nor
faint when thou art
rebuked
of
him:
\VS{6}For
whom the
Lord
loveth he
chasteneth,
and
scourgeth
every
son
whom he
receiveth.
\VS{7}If ye
endure
chastening,
God
dealeth with
you
as with
sons;
for
what
son is
he
whom the
father
chasteneth
not?
\VS{8}But
if ye
be
without
chastisement,
whereof
all
are
partakers,
then are
ye
bastards,
and
not
sons.
\VS{9}Furthermore we have
had
fathers
of
our
flesh which
corrected
{\ADD{us}},
and we
gave
{\ADD{them}}
reverence: shall
we
not
much
rather be in subjection
unto the
Father of
spirits,
and
live?
\VS{10}For
they
verily
for a
few
days
chastened
{\ADD{us}}
after their
own
pleasure;
but
he
for
{\ADD{our}}
profit,
that
{\ADD{we}} might be
partakers
of
his
holiness.
\VS{11}Now
no
chastening
for the
present
seemeth to
be
joyous,
but
grievous:
nevertheless
afterward it
yieldeth
the
peaceable
fruit of
righteousness unto them which are
exercised
thereby.
\VS{12}Wherefore lift
up the
hands which hang
down,
and
the
feeble
knees;
\VS{13}And
make
straight
paths
for
your
feet,
lest that which is
lame be turned out of the
way;
but let
it
rather be
healed.
\VS{14}Follow
peace
with
all
{\ADD{men}},
and
holiness,
without
which no
man shall
see the
Lord:
\VS{15}Looking
diligently
lest any
man
fail
of the
grace of
God;
lest
any
root of
bitterness
springing
up
trouble
{\ADD{you}},
and
thereby
many be
defiled;
\VS{16}Lest
there
{\ADD{be}}
any
fornicator,
or profane
person,
as
Esau,
who
for
one morsel of
meat
sold
his
birthright.
\VS{17}For ye
know how
that
afterward, when he would
have
inherited the
blessing, he was
rejected:
for he
found
no
place of
repentance,
though he
sought
it
carefully
with
tears.
\VS{18}For ye
are
not come
unto the
mount that might be
touched,
and that
burned with
fire,
nor unto
blackness,
and
darkness,
and
tempest,
\VS{19}And the
sound of a
trumpet,
and the
voice of
words;
which
{\ADD{voice}} they that
heard
intreated
that the
word
should
not be
spoken to
them any more:
\VS{20}(For they
could
not
endure that which was
commanded, And if so much
as a
beast
touch the
mountain, it shall be
stoned,
or thrust
through with a
dart:
\VS{21}And
so
terrible
was the
sight,
{\ADD{that}}
Moses
said,
I exceedingly
fear
and
quake:)
\VS{22}But ye are
come unto
mount
Sion,
and unto the
city of
the
living
God,
the
heavenly
Jerusalem,
and to an innumerable
company of
angels,
\VS{23}To the general
assembly
and
church of the
firstborn, which are
written
in
heaven,
and to
God the
Judge of
all,
and to the
spirits of just
men made
perfect,
\VS{24}And to
Jesus the
mediator of
the
new
covenant,
and to the
blood of
sprinkling, that
speaketh better
things
than
{\ADD{that of}}
Abel.
\VS{25}See
that
ye
refuse
not him that
speaketh.
For
if
they
escaped
not who
refused him that
spake
on
earth,
much
more
{\ADD{shall not}}
we
{\ADD{escape}}, if we turn away
from him
that
{\ADD{speaketh}}
from
heaven:
\VS{26}Whose
voice
then
shook the
earth:
but
now he hath
promised,
saying,
Yet
once
more
I
shake
not the
earth
only,
but
also
heaven.
\VS{27}And
this
{\ADD{word}},
Yet
once
more,
signifieth the
removing of those things that are
shaken,
as of things that are
made,
that those things
which
cannot be
shaken may
remain.
\VS{28}Wherefore we
receiving a
kingdom which cannot be
moved, let us
have
grace,
whereby we may
serve
God
acceptably
with
reverence
and godly
fear:
\VS{29}For
our
God
{\ADD{is}}
a
consuming
fire.

\par }\Chap{13}{\PP \VerseOne{1}Let brotherly
love
continue.
\VS{2}Be
not
forgetful to entertain
strangers:
for
thereby
some have
entertained
angels
unawares.
\VS{3}Remember them that are in
bonds,
as bound with
them;
{\ADD{and}} them which suffer
adversity,
as
being
yourselves
also
in the
body.
\VS{4}Marriage
{\ADD{is}}
honourable
in
all,
and the
bed
undefiled:
but
whoremongers
and
adulterers
God will
judge.
\VS{5}{\ADD{Let your}}
conversation
{\ADD{be}} without
covetousness;
{\ADD{and be}}
content with such things as ye
have:
for
he hath
said, I
will
never
leave
thee,
nor
forsake
thee.
\VS{6}So
that
we
may
boldly
say, The
Lord
{\ADD{is}}
my
helper,
and I
will
not
fear
what
man shall
do unto
me.
\par }{\PP \VS{7}Remember them which have the rule
over
you,
who have
spoken unto
you the
word of
God:
whose
faith
follow,
considering the
end
of
{\ADD{their}}
conversation.
\VS{8}Jesus
Christ the
same
yesterday,
and to
day,
and
for
ever.
\VS{9}Be
not carried
about with
divers
and
strange
doctrines.
For
{\ADD{it is}} a good
thing
that the
heart be
established with
grace;
not with
meats,
which
have
not
profited them that have been
occupied
therein.
\VS{10}We
have an
altar,
whereof they
have
no
right to
eat which
serve the
tabernacle.
\VS{11}For the
bodies
of
those
beasts,
whose
blood is
brought
into the
sanctuary
by the high
priest
for
sin, are
burned
without the
camp.
\VS{12}Wherefore
Jesus
also,
that he might
sanctify the
people
with his
own
blood,
suffered
without the
gate.
\VS{13}Let us go
forth
therefore
unto
him
without the
camp,
bearing
his
reproach.
\VS{14}For
here have
we
no
continuing
city,
but we
seek one to
come.
\VS{15}By
him
therefore let us
offer the
sacrifice of
praise to
God
continually,
that
is, the
fruit
of
{\ADD{our}}
lips giving
thanks
to
his
name.
\VS{16}But to do
good
and to
communicate
forget
not:
for with
such
sacrifices
God is well
pleased.
\VS{17}Obey them that have the rule
over
you,
and submit
yourselves:
for
they
watch
for
your
souls,
as they that must
give
account,
that they may
do
it
with
joy,
and
not with
grief:
for
that
{\ADD{is}}
unprofitable for
you.
\par }{\PP \VS{18}Pray
for
us:
for we
trust we
have
a
good
conscience,
in all
things
willing to
live
honestly.
\VS{19}But I
beseech
{\ADD{you}} the
rather to
do
this,
that I may be
restored to
you the
sooner.
\VS{20}Now the
God of
peace, that brought
again
from the
dead
our
Lord
Jesus,
that
great
shepherd of the
sheep,
through the
blood of
the
everlasting
covenant,
\VS{21}Make
you
perfect
in
every
good
work
to
do
his
will,
working
in
you that which is
wellpleasing
in
his
sight,
through
Jesus
Christ; to
whom
{\ADD{be}}
glory
for
ever and
ever.
Amen.
\VS{22}And I
beseech
you,
brethren,
suffer the
word of
exhortation:
for I have written a
letter unto
you
in few
words.
\VS{23}Know
ye
that
{\ADD{our}}
brother
Timothy is set at
liberty;
with
whom,
if he
come
shortly, I will
see
you.
\VS{24}Salute
all them that have the rule
over
you,
and
all the
saints.
They
of
Italy
salute
you.
\VS{25}Grace
{\ADD{be}}
with
you
all.
Amen.
\par }{\SH Written to the Hebrews from Italy by Timothy.
\par }