\NormalFont\ShortTitle{Hosea}
{\MT Hosea

\par }\ChapOne{1}{\PP \VerseOne{1}The
word of the
{\ND{
LORD}} that came unto
Hosea, the
son of
Beeri, in the
days of
Uzziah,
Jotham,
Ahaz,
{\ADD{and}}
Hezekiah,
kings of
Judah, and in the
days of
Jeroboam the
son of
Joash,
king of
Israel.
\par }{\PP \VS{2}The
beginning of the
word of the
{\ND{
LORD}} by
Hosea. And the
{\ND{
LORD}}
said to
Hosea,
Go,
take unto thee a
wife of
whoredoms and
children of
whoredoms: for the
land hath committed
great
whoredom,
{\ADD{departing}}
from the
{\ND{
LORD}}.
\VS{3}So he
went and
took
Gomer the
daughter of
Diblaim; which
conceived, and
bare him a
son.
\VS{4}And the
{\ND{
LORD}}
said unto him,
Call his
name
Jezreel; for yet a
little
{\ADD{while}}, and I will
avenge the
blood of
Jezreel upon the
house of
Jehu, and will cause to
cease the
kingdom of the
house of
Israel.\FTNT{*}{{\FR 1.4 }
avenge: Heb. visit}
\VS{5}And it shall come to pass at that
day, that I will
break the
bow of
Israel in the
valley of
Jezreel.
\par }{\PP \VS{6}¶ And she conceived
again, and
bare a
daughter. And
{\ADD{God}}
said unto him,
Call her
name
Lo-ruhamah: for I will no
more have
mercy upon the
house of
Israel; but I will
utterly take them
away.\FTNT{†}{{\FR 1.6 }
Lo-ruhamah: that is, Not having obtained mercy}\FTNT{‡}{{\FR 1.6 }
no…: Heb. not add any more to}\FTNT{§}{{\FR 1.6 }
but…: or, that I should altogether pardon them}
\VS{7}But I will have
mercy upon the
house of
Judah, and will
save them by the
{\ND{
LORD}} their
God, and will not
save them by
bow, nor by
sword, nor by
battle, by
horses, nor by
horsemen.
\par }{\PP \VS{8}¶ Now when she had
weaned
Lo-ruhamah, she
conceived, and
bare a
son.
\VS{9}Then
said
{\ADD{God}},
Call his
name
Lo-ammi: for ye
{\ADD{are}} not my
people, and I will not be your
{\ADD{God}}.\FTNT{**}{{\FR 1.9 }
Lo-ammi: that is, Not my people}
\par }{\PP \VS{10}¶ Yet the
number of the
children of
Israel shall be as the
sand of the
sea, which cannot be
measured nor
numbered; and it shall come to pass,
{\ADD{that}} in the
place where it was
said unto them, Ye
{\ADD{are}} not my
people,
{\ADD{there}} it shall be
said unto them,
{\ADD{Ye are}} the
sons of the
living
God.\FTNT{††}{{\FR 1.10 }
in…: or, instead of that}
\VS{11}Then shall the
children of
Judah and the
children of
Israel be
gathered
together, and
appoint themselves
one
head, and they shall come
up out of the
land: for
great
{\ADD{shall be}} the
day of
Jezreel.

\par }\Chap{2}{\PP \VerseOne{1}Say ye unto your
brethren,
Ammi; and to your
sisters,
Ru-hamah.
\FTNT{*}{{\FR 2.1 }
Ammi: that is, My people}\FTNT{†}{{\FR 2.1 }
Ru-hamah: that is, Having obtained mercy}
\VS{2}Plead with your
mother,
plead: for she
{\ADD{is}} not my
wife, neither
{\ADD{am}} I her
husband: let her therefore put
away her
whoredoms out of her
sight, and her
adulteries from between her
breasts;
\VS{3}Lest I
strip her
naked, and
set her as in the
day that she was
born, and
make her as a
wilderness, and
set her like a
dry
land, and
slay her with
thirst.
\VS{4}And I will not have
mercy upon her
children; for they
{\ADD{be}} the
children of
whoredoms.
\VS{5}For their
mother hath played the
harlot: she that
conceived them hath done
shamefully: for she
said, I will
go
after my
lovers, that
give
{\ADD{me}} my
bread and my
water, my
wool and my
flax, mine
oil and my
drink.\FTNT{‡}{{\FR 2.5 }
drink: Heb. drinks}
\par }{\PP \VS{6}¶ Therefore, behold, I will
hedge up thy
way with
thorns, and
make a
wall, that she shall not
find her
paths.\FTNT{§}{{\FR 2.6 }
make…: Heb. wall a wall}
\VS{7}And she shall
follow after her
lovers, but she shall not
overtake them; and she shall
seek them, but shall not
find
{\ADD{them:}} then shall she
say, I will
go and
return to my
first
husband; for then
{\ADD{was it}}
better with me than
now.
\VS{8}For she did not
know that I
gave her
corn, and
wine, and
oil, and
multiplied her
silver and
gold,
{\ADD{which}} they
prepared for
Baal.\FTNT{**}{{\FR 2.8 }
wine: Heb. new wine}\FTNT{††}{{\FR 2.8 }
which…: or, wherewith they made Baal}
\VS{9}Therefore will I
return, and take
away my
corn in the
time thereof, and my
wine in the
season thereof, and will
recover my
wool and my
flax
{\ADD{given}} to
cover her
nakedness.\FTNT{‡‡}{{\FR 2.9 }
recover: or, take away}
\VS{10}And now will I
discover her
lewdness in the
sight of her
lovers, and
none shall
deliver her out of mine
hand.\FTNT{§§}{{\FR 2.10 }
lewdness: Heb. folly, or, villany}
\VS{11}I will also cause all her
mirth to
cease, her feast
days, her new
moons, and her
sabbaths, and all her solemn
feasts.
\VS{12}And I will
destroy her
vines and her fig
trees, whereof she hath
said, These
{\ADD{are}} my
rewards that my
lovers have
given me: and I will
make them a
forest, and the
beasts of the
field shall
eat them.\FTNT{***}{{\FR 2.12 }
destroy: Heb. make desolate}
\VS{13}And I will
visit upon her the
days of
Baalim, wherein she burned
incense to them, and she
decked herself with her
earrings and her
jewels, and she
went
after her
lovers, and
forgat me,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{14}¶ Therefore, behold, I will
allure her, and
bring her into the
wilderness, and
speak
comfortably unto her.\FTNT{†††}{{\FR 2.14 }
comfortably: or, friendly: Heb. to her heart}
\VS{15}And I will
give her her
vineyards from thence, and the
valley of
Achor for a
door of
hope: and she shall
sing there, as in the
days of her
youth, and as in the
day when she came
up out of the
land of
Egypt.
\VS{16}And it shall be at that
day,
saith the
{\ND{
LORD}},
{\ADD{that}} thou shalt
call me
Ishi; and shalt
call me no more
Baali.\FTNT{‡‡‡}{{\FR 2.16 }
Ishi: that is, My husband}\FTNT{§§§}{{\FR 2.16 }
Baali: that is, My lord}
\VS{17}For I will take
away the
names of
Baalim out of her
mouth, and they shall no more be
remembered by their
name.
\VS{18}And in that
day will I
make a
covenant for them with the
beasts of the
field, and with the
fowls of
heaven, and
{\ADD{with}} the creeping
things of the
ground: and I will
break the
bow and the
sword and the
battle out of the
earth, and will make them to lie
down
safely.
\VS{19}And I will
betroth thee unto me for
ever; yea, I will
betroth thee unto me in
righteousness, and in
judgment, and in
lovingkindness, and in
mercies.
\VS{20}I will even
betroth thee unto me in
faithfulness: and thou shalt
know the
{\ND{
LORD}}.
\VS{21}And it shall come to pass in that
day, I will
hear,
saith the
{\ND{
LORD}}, I will
hear the
heavens, and they shall
hear the
earth;
\VS{22}And the
earth shall
hear the
corn, and the
wine, and the
oil; and they shall
hear
Jezreel.
\VS{23}And I will
sow her unto me in the
earth; and I will have
mercy upon her that had not obtained
mercy; and I will
say to
{\ADD{them which were}} not my
people, Thou
{\ADD{art}} my
people; and they shall
say,
{\ADD{Thou art}} my
God.

\par }\Chap{3}{\PP \VerseOne{1}Then
said the
{\ND{
LORD}} unto me,
Go yet,
love a
woman
beloved of
{\ADD{her}}
friend, yet an
adulteress, according to the
love of the
{\ND{
LORD}} toward the
children of
Israel, who
look to
other
gods, and
love
flagons of
wine.\FTNT{*}{{\FR 3.1 }
of wine: Heb. of grapes}
\VS{2}So I
bought her to me for
fifteen
{\ADD{pieces}} of
silver, and
{\ADD{for}} an
homer of
barley, and an half
homer of
barley:\FTNT{†}{{\FR 3.2 }
half…: Heb. lethech}
\VS{3}And I
said unto her, Thou shalt
abide for me
many
days; thou shalt not play the
harlot, and thou shalt not
be for
{\ADD{another}}
man: so
{\ADD{will}} I also
{\ADD{be}} for thee.
\VS{4}For the
children of
Israel shall
abide
many
days without a
king, and without a
prince, and without a
sacrifice, and without an
image, and without an
ephod, and
{\ADD{without}}
teraphim:\FTNT{‡}{{\FR 3.4 }
image: Heb. a standing, or, statue, or, pillar}
\VS{5}Afterward shall the
children of
Israel
return, and
seek the
{\ND{
LORD}} their
God, and
David their
king; and shall
fear the
{\ND{
LORD}} and his
goodness in the
latter
days.

\par }\Chap{4}{\PP \VerseOne{1}Hear the
word of the
{\ND{
LORD}}, ye
children of
Israel: for the
{\ND{
LORD}} hath a
controversy with the
inhabitants of the
land, because
{\ADD{there is}} no
truth, nor
mercy, nor
knowledge of
God in the
land.
\VS{2}By
swearing, and
lying, and
killing, and
stealing, and committing
adultery, they break
out, and
blood
toucheth
blood.\FTNT{*}{{\FR 4.2 }
blood: Heb. bloods}
\VS{3}Therefore shall the
land
mourn, and every one that
dwelleth therein shall
languish, with the
beasts of the
field, and with the
fowls of
heaven; yea, the
fishes of the
sea also shall be taken
away.
\VS{4}Yet let no
man
strive, nor
reprove
another: for thy
people
{\ADD{are}} as they that
strive with the
priest.
\VS{5}Therefore shalt thou
fall in the
day, and the
prophet also shall
fall with thee in the
night, and I will
destroy thy
mother.\FTNT{†}{{\FR 4.5 }
destroy: Heb. cut off}
\par }{\PP \VS{6}¶ My
people are
destroyed for lack of
knowledge: because thou hast
rejected
knowledge, I will also
reject thee, that thou shalt be no
priest to me: seeing thou hast
forgotten the
law of thy
God, I will also
forget thy
children.\FTNT{‡}{{\FR 4.6 }
destroyed: Heb. cut off}
\VS{7}As they were
increased, so they
sinned against me:
{\ADD{therefore}} will I
change their
glory into
shame.
\VS{8}They eat
up the
sin of my
people, and they
set their
heart on their
iniquity.\FTNT{§}{{\FR 4.8 }
set…: Heb. lift up their soul to}
\VS{9}And there shall be, like
people, like
priest: and I will
punish them for their
ways, and
reward them their
doings.\FTNT{**}{{\FR 4.9 }
punish: Heb. visit upon}\FTNT{††}{{\FR 4.9 }
reward: Heb. cause to return}
\VS{10}For they shall
eat, and not have
enough: they shall commit
whoredom, and shall not
increase: because they have left
off to take
heed to the
{\ND{
LORD}}.
\VS{11}Whoredom and
wine and new
wine take
away the
heart.
\par }{\PP \VS{12}¶ My
people
ask counsel at their
stocks, and their
staff
declareth unto them: for the
spirit of
whoredoms hath caused
{\ADD{them}} to
err, and they have gone a
whoring from under their
God.
\VS{13}They
sacrifice upon the
tops of the
mountains, and burn
incense upon the
hills, under
oaks and
poplars and
elms, because the
shadow thereof
{\ADD{is}}
good: therefore your
daughters shall commit
whoredom, and your
spouses shall commit
adultery.
\VS{14}I will not
punish your
daughters when they commit
whoredom, nor your
spouses when they commit
adultery: for themselves are
separated with
whores, and they
sacrifice with
harlots: therefore the
people
{\ADD{that}} doth not
understand shall
fall.\FTNT{‡‡}{{\FR 4.14 }
I will not: or, Shall I not}\FTNT{§§}{{\FR 4.14 }
fall: or, be punished}
\par }{\PP \VS{15}¶ Though thou,
Israel, play the
harlot,
{\ADD{yet}} let not
Judah
offend; and
come not ye unto
Gilgal, neither go ye
up to
Beth-aven, nor
swear, The
{\ND{
LORD}}
liveth.
\VS{16}For
Israel slideth
back as a
backsliding
heifer: now the
{\ND{
LORD}} will
feed them as a
lamb in a large
place.
\VS{17}Ephraim
{\ADD{is}}
joined to
idols: let him
alone.
\VS{18}Their
drink is
sour: they have committed
whoredom
continually: her
rulers
{\ADD{with}}
shame do
love,
Give ye.\FTNT{***}{{\FR 4.18 }
sour: Heb. gone}\FTNT{†††}{{\FR 4.18 }
rulers: Heb. shields}
\VS{19}The
wind hath bound her
up in her
wings, and they shall be
ashamed because of their
sacrifices.

\par }\Chap{5}{\PP \VerseOne{1}Hear ye this, O
priests; and
hearken, ye
house of
Israel; and give ye
ear, O
house of the
king; for
judgment
{\ADD{is}} toward you, because ye have been a
snare on
Mizpah, and a
net
spread upon
Tabor.
\VS{2}And the
revolters are
profound to make
slaughter, though I
{\ADD{have been}} a
rebuker of them all.\FTNT{*}{{\FR 5.2 }
though: or, and}\FTNT{†}{{\FR 5.2 }
a rebuker: Heb. a correction}
\VS{3}I
know
Ephraim, and
Israel is not
hid from me: for now, O
Ephraim, thou committest
whoredom,
{\ADD{and}}
Israel is
defiled.
\VS{4}They will not
frame their
doings to
turn unto their
God: for the
spirit of
whoredoms
{\ADD{is}} in the
midst of them, and they have not
known the
{\ND{
LORD}}.\FTNT{‡}{{\FR 5.4 }
They will…: or, Their doings will not suffer them}\FTNT{§}{{\FR 5.4 }
frame: Heb. give}
\VS{5}And the
pride of
Israel doth
testify to his
face: therefore shall
Israel and
Ephraim
fall in their
iniquity;
Judah also shall
fall with them.
\VS{6}They shall
go with their
flocks and with their
herds to
seek the
{\ND{
LORD}}; but they shall not
find
{\ADD{him;}} he hath
withdrawn himself from them.
\VS{7}They have dealt
treacherously against the
{\ND{
LORD}}: for they have
begotten
strange
children: now shall a
month
devour them with their
portions.
\par }{\PP \VS{8}Blow ye the
cornet in
Gibeah,
{\ADD{and}} the
trumpet in
Ramah: cry
aloud
{\ADD{at}}
Beth-aven,
after thee, O
Benjamin.
\VS{9}Ephraim shall be
desolate in the
day of
rebuke: among the
tribes of
Israel have I made
known that which shall surely
be.
\VS{10}The
princes of
Judah were like them that
remove the
bound:
{\ADD{therefore}} I will pour
out my
wrath upon them like
water.
\VS{11}Ephraim
{\ADD{is}}
oppressed
{\ADD{and}}
broken in
judgment, because he
willingly
walked
after the
commandment.
\VS{12}Therefore
{\ADD{will}} I
{\ADD{be}} unto
Ephraim as a
moth, and to the
house of
Judah as
rottenness.\FTNT{**}{{\FR 5.12 }
rottenness: or, a worm}
\VS{13}When
Ephraim
saw his
sickness, and
Judah
{\ADD{saw}} his
wound, then
went
Ephraim to the
Assyrian, and
sent to
king
Jareb: yet
could he not
heal you, nor
cure you of your
wound.\FTNT{††}{{\FR 5.13 }
king Jareb: or, the king of Jareb: or, the king that should plead}
\VS{14}For I
{\ADD{will be}} unto
Ephraim as a
lion, and as a young
lion to the
house of
Judah: I,
{\ADD{even}} I, will
tear and go
away; I will take
away, and none shall
rescue
{\ADD{him}}.
\par }{\PP \VS{15}¶ I will
go
{\ADD{and}}
return to my
place, till they acknowledge their
offence, and
seek my
face: in their
affliction they will seek me
early.\FTNT{‡‡}{{\FR 5.15 }
acknowledge…: Heb. be guilty}

\par }\Chap{6}{\PP \VerseOne{1}Come, and let us
return unto the
{\ND{
LORD}}: for he hath
torn, and he will
heal us; he hath
smitten, and he will bind us
up.
\VS{2}After two
days will he
revive us: in the
third
day he will raise us
up, and we shall
live in his
sight.
\VS{3}Then shall we
know,
{\ADD{if}} we
follow on to
know the
{\ND{
LORD}}: his going
forth is
prepared as the
morning; and he shall
come unto us as the
rain, as the
latter
{\ADD{and}} former
rain unto the
earth.
\par }{\PP \VS{4}¶ O
Ephraim, what shall I
do unto thee? O
Judah, what shall I
do unto thee? for your
goodness
{\ADD{is}} as a
morning
cloud, and as the
early
dew it goeth
away.\FTNT{*}{{\FR 6.4 }
goodness: or, mercy, or, kindness}
\VS{5}Therefore have I
hewed
{\ADD{them}} by the
prophets; I have
slain them by the
words of my
mouth: and thy
judgments
{\ADD{are as}} the
light
{\ADD{that}} goeth
forth.\FTNT{†}{{\FR 6.5 }
and…: or, that thy judgments might be, etc}
\VS{6}For I
desired
mercy, and not
sacrifice; and the
knowledge of
God more than burnt
offerings.
\VS{7}But they like
men have
transgressed the
covenant: there have they dealt
treacherously against me.\FTNT{‡}{{\FR 6.7 }
men: or, Adam}
\VS{8}Gilead
{\ADD{is}} a
city of them that
work
iniquity,
{\ADD{and is}}
polluted with
blood.\FTNT{§}{{\FR 6.8 }
polluted: or, cunning for}
\VS{9}And as
troops of robbers
wait for a
man,
{\ADD{so}} the
company of
priests
murder in the
way by
consent: for they
commit
lewdness.\FTNT{**}{{\FR 6.9 }
by…: Heb. with one shoulder, or, to Shechem}\FTNT{††}{{\FR 6.9 }
lewdness: or, enormity}
\VS{10}I have
seen an horrible
thing in the
house of
Israel: there
{\ADD{is}} the
whoredom of
Ephraim,
Israel is
defiled.
\VS{11}Also, O
Judah, he hath
set an
harvest for thee, when I
returned the
captivity of my
people.

\par }\Chap{7}{\PP \VerseOne{1}When I would have
healed
Israel, then the
iniquity of
Ephraim was
discovered, and the
wickedness of
Samaria: for they
commit
falsehood; and the
thief cometh
in,
{\ADD{and}} the
troop of robbers
spoileth
without.\FTNT{*}{{\FR 7.1 }
wickedness: Heb. evils}\FTNT{†}{{\FR 7.1 }
spoileth: Heb. strippeth}
\VS{2}And they
consider not in their
hearts
{\ADD{that}} I
remember all their
wickedness: now their own
doings have beset them
about; they are before my
face.\FTNT{‡}{{\FR 7.2 }
consider…: Heb. say not to}
\VS{3}They make the
king
glad with their
wickedness, and the
princes with their
lies.
\VS{4}They
{\ADD{are}} all
adulterers, as an
oven
heated by the
baker,
{\ADD{who}}
ceaseth from
raising after he hath
kneaded the
dough, until it be
leavened.\FTNT{§}{{\FR 7.4 }
who…: or, the raiser will cease}\FTNT{**}{{\FR 7.4 }
raising: or, waking}
\VS{5}In the
day of our
king the
princes have made
{\ADD{him}}
sick with
bottles of
wine; he stretched
out his
hand with
scorners.\FTNT{††}{{\FR 7.5 }
bottles…: or, heat through wine}
\VS{6}For they have made
ready their
heart like an
oven, whiles they lie in
wait: their
baker
sleepeth all the
night; in the
morning it
burneth as a
flaming
fire.\FTNT{‡‡}{{\FR 7.6 }
made…: or, applied}
\VS{7}They are all
hot as an
oven, and have
devoured their
judges; all their
kings are
fallen:
{\ADD{there is}} none among them that
calleth unto me.
\par }{\PP \VS{8}Ephraim, he hath
mixed himself among the
people;
Ephraim is a
cake not
turned.
\VS{9}Strangers have
devoured his
strength, and he
knoweth
{\ADD{it}} not: yea, gray
hairs are here and
there upon him, yet he
knoweth not.\FTNT{§§}{{\FR 7.9 }
here…: Heb. sprinkled}
\VS{10}And the
pride of
Israel
testifieth to his
face: and they do not
return to the
{\ND{
LORD}} their
God, nor
seek him for all this.
\par }{\PP \VS{11}¶
Ephraim also is like a
silly
dove without
heart: they
call to
Egypt, they
go to
Assyria.
\VS{12}When they shall
go, I will
spread my
net upon them; I will bring them
down as the
fowls of the
heaven; I will
chastise them, as their
congregation hath
heard.
\VS{13}Woe unto them! for they have
fled from me:
destruction unto them! because they have
transgressed against me: though I have
redeemed them, yet they have
spoken
lies against me.\FTNT{***}{{\FR 7.13 }
destruction: Heb. spoil}
\VS{14}And they have not
cried unto me with their
heart, when they
howled upon their
beds: they
assemble themselves for
corn and
wine,
{\ADD{and}} they
rebel against me.
\VS{15}Though I have
bound
{\ADD{and}}
strengthened their
arms, yet do they
imagine
mischief against me.\FTNT{†††}{{\FR 7.15 }
have…: or, chastened}
\VS{16}They
return,
{\ADD{but}} not to the most
High: they are like a
deceitful
bow: their
princes shall
fall by the
sword for the
rage of their
tongue:
this
{\ADD{shall be}} their
derision in the
land of
Egypt.

\par }\Chap{8}{\PP \VerseOne{1}{\ADD{Set}} the
trumpet to thy
mouth.
{\ADD{He shall come}} as an
eagle against the
house of the
{\ND{
LORD}}, because they have
transgressed my
covenant, and
trespassed against my
law.\FTNT{*}{{\FR 8.1 }
thy: Heb. the roof of thy}
\VS{2}Israel shall
cry unto me, My
God, we
know thee.
\VS{3}Israel hath cast
off
{\ADD{the thing that is}}
good: the
enemy shall
pursue him.
\VS{4}They have set up
kings, but not by me: they have made
princes, and I
knew
{\ADD{it}} not: of their
silver and their
gold have they
made them
idols, that they may be cut
off.
\par }{\PP \VS{5}¶ Thy
calf, O
Samaria, hath
cast
{\ADD{thee}} off; mine
anger is
kindled against them: how long
{\ADD{will it be}}
ere they
attain to
innocency?
\VS{6}For from
Israel
{\ADD{was}} it also: the
workman
made it; therefore it
{\ADD{is}} not
God: but the
calf of
Samaria shall be broken in
pieces.
\VS{7}For they have
sown the
wind, and they shall
reap the
whirlwind: it hath no
stalk: the
bud shall
yield no
meal:
if so be it
yield, the
strangers shall swallow it
up.\FTNT{†}{{\FR 8.7 }
stalk: or, standing corn}
\par }{\PP \VS{8}Israel is swallowed
up: now shall they be among the
Gentiles as a
vessel wherein
{\ADD{is}} no
pleasure.
\VS{9}For they are gone
up to
Assyria, a wild
ass
alone by himself:
Ephraim hath
hired
lovers.\FTNT{‡}{{\FR 8.9 }
lovers: Heb. loves}
\VS{10}Yea, though they have
hired among the
nations, now will I
gather them, and they shall
sorrow a
little for the
burden of the
king of
princes.\FTNT{§}{{\FR 8.10 }
sorrow: or, begin}\FTNT{**}{{\FR 8.10 }
a…: or, in a little while}
\VS{11}Because
Ephraim hath made
many
altars to
sin,
altars shall be unto him to
sin.
\VS{12}I have
written to him the great
things of my
law,
{\ADD{but}} they were
counted as a strange
thing.
\VS{13}They
sacrifice
flesh
{\ADD{for}} the
sacrifices of mine
offerings, and
eat
{\ADD{it; but}} the
{\ND{
LORD}}
accepteth them not; now will he
remember their
iniquity, and
visit their
sins: they shall
return to
Egypt.\FTNT{††}{{\FR 8.13 }
They sacrifice…: or, In the sacrifices of mine offerings they, etc}
\VS{14}For
Israel hath
forgotten his
Maker, and
buildeth
temples; and
Judah hath
multiplied
fenced
cities: but I will
send a
fire upon his
cities, and it shall
devour the
palaces thereof.

\par }\Chap{9}{\PP \VerseOne{1}Rejoice not, O
Israel, for
joy, as
{\ADD{other}}
people: for thou hast gone a
whoring from thy
God, thou hast
loved a
reward upon every
cornfloor.\FTNT{*}{{\FR 9.1 }
upon: or, in, etc}
\VS{2}The
floor and the
winepress shall not
feed them, and the new
wine shall
fail in her.\FTNT{†}{{\FR 9.2 }
winepress: or, winefat}
\VS{3}They shall not
dwell in the
{\ND{
LORD’s}}
land; but
Ephraim shall
return to
Egypt, and they shall
eat
unclean
{\ADD{things}} in
Assyria.
\VS{4}They shall not
offer
wine
{\ADD{offerings}} to the
{\ND{
LORD}}, neither shall they be
pleasing unto him: their
sacrifices
{\ADD{shall be}} unto them as the
bread of
mourners; all that
eat thereof shall be
polluted: for their
bread for their
soul shall not
come into the
house of the
{\ND{
LORD}}.
\VS{5}What will ye
do in the
solemn
day, and in the
day of the
feast of the
{\ND{
LORD}}?
\VS{6}For, lo, they are
gone because of
destruction:
Egypt shall gather them
up,
Memphis shall
bury them: the
pleasant
{\ADD{places}} for their
silver,
nettles shall
possess them:
thorns
{\ADD{shall be}} in their
tabernacles.\FTNT{‡}{{\FR 9.6 }
destruction: Heb. spoil}\FTNT{§}{{\FR 9.6 }
the…: or, their silver shall be desired, the nettle, etc.: Heb. the desire}
\par }{\PP \VS{7}The
days of
visitation are
come, the
days of
recompence are
come;
Israel shall
know
{\ADD{it:}} the
prophet
{\ADD{is}} a
fool, the
spiritual
man
{\ADD{is}}
mad, for the
multitude of thine
iniquity, and the
great
hatred.\FTNT{**}{{\FR 9.7 }
spiritual…: Heb. man of the spirit}
\VS{8}The
watchman of
Ephraim
{\ADD{was}} with my
God:
{\ADD{but}} the
prophet
{\ADD{is}} a
snare of a
fowler in all his
ways,
{\ADD{and}}
hatred in the
house of his
God.\FTNT{††}{{\FR 9.8 }
in the: or, against the}
\VS{9}They have
deeply
corrupted
{\ADD{themselves}}, as in the
days of
Gibeah:
{\ADD{therefore}} he will
remember their
iniquity, he will
visit their
sins.
\VS{10}I
found
Israel like
grapes in the
wilderness; I
saw your
fathers as the
firstripe in the fig
tree at her first
time:
{\ADD{but}} they
went to
Baal-peor, and
separated themselves unto
{\ADD{that}}
shame; and
{\ADD{their}}
abominations were according as they
loved.
\par }{\PP \VS{11}{\ADD{As for}}
Ephraim, their
glory shall fly
away like a
bird, from the
birth, and from the
womb, and from the
conception.
\VS{12}Though they bring
up their
children, yet will I
bereave them,
{\ADD{that there shall}} not
{\ADD{be}} a
man
{\ADD{left:}} yea,
woe also to them when I
depart from them!
\VS{13}Ephraim, as I
saw
Tyrus,
{\ADD{is}}
planted in a pleasant
place: but
Ephraim shall bring
forth his
children to the
murderer.
\VS{14}Give them, O
{\ND{
LORD}}: what wilt thou
give?
give them a
miscarrying
womb and
dry
breasts.\FTNT{‡‡}{{\FR 9.14 }
miscarrying: Heb. that casteth the fruit}
\VS{15}All their
wickedness
{\ADD{is}} in
Gilgal: for there I
hated them: for the
wickedness of their
doings I will drive them
out of mine
house, I will
love them no
more: all their
princes
{\ADD{are}}
revolters.
\VS{16}Ephraim is
smitten, their
root is dried
up, they shall
bear no
fruit: yea, though they bring
forth, yet will I
slay
{\ADD{even}} the
beloved
{\ADD{fruit}} of their
womb.\FTNT{§§}{{\FR 9.16 }
the…: Heb. the desires}
\VS{17}My
God will cast them
away, because they did not
hearken unto him: and they shall be
wanderers among the
nations.

\par }\Chap{10}{\PP \VerseOne{1}Israel
{\ADD{is}} an
empty
vine, he bringeth
forth
fruit unto himself: according to the
multitude of his
fruit he hath
increased the
altars; according to the
goodness of his
land they have made
goodly
images.\FTNT{*}{{\FR 10.1 }
an…: or, a vine emptying the fruit which it giveth}\FTNT{†}{{\FR 10.1 }
images: Heb. statues, or, standing images}
\VS{2}Their
heart is
divided; now shall they be found
faulty: he shall break
down their
altars, he shall
spoil their
images.\FTNT{‡}{{\FR 10.2 }
Their heart…: or, He hath divided their heart}\FTNT{§}{{\FR 10.2 }
break…: Heb. behead}\FTNT{**}{{\FR 10.2 }
images: Heb. statues, or, standing images}
\VS{3}For now they shall
say, We have no
king, because we
feared not the
{\ND{
LORD}}; what then should a
king
do to us?
\VS{4}They have
spoken
words,
swearing
falsely in
making a
covenant: thus
judgment springeth
up as
hemlock in the
furrows of the
field.
\VS{5}The
inhabitants of
Samaria shall
fear because of the
calves of
Beth-aven: for the
people thereof shall
mourn over it, and the
priests thereof
{\ADD{that}}
rejoiced on it, for the
glory thereof, because it is
departed from it.\FTNT{††}{{\FR 10.5 }
the priests…: or, Chemarim}
\VS{6}It shall be also
carried unto
Assyria
{\ADD{for}} a
present to
king
Jareb:
Ephraim shall
receive
shame, and
Israel shall be
ashamed of his own
counsel.
\VS{7}{\ADD{As for}}
Samaria, her
king is cut
off as the
foam
upon the
water.\FTNT{‡‡}{{\FR 10.7 }
the water: Heb. the face of the water}
\VS{8}The high
places also of
Aven, the
sin of
Israel, shall be
destroyed: the
thorn and the
thistle shall come
up on their
altars; and they shall
say to the
mountains,
Cover us; and to the
hills,
Fall on us.
\par }{\PP \VS{9}O
Israel, thou hast
sinned from the
days of
Gibeah: there they
stood: the
battle in
Gibeah against the
children of
iniquity did not
overtake them.
\VS{10}{\ADD{It is}} in my
desire that I should
chastise them; and the
people shall be
gathered against them, when they shall
bind themselves in their
two
furrows.\FTNT{§§}{{\FR 10.10 }
when…: or, when I shall bind them for their two transgressions, or, in their two habitations}
\VS{11}And
Ephraim
{\ADD{is as}} an
heifer
{\ADD{that is}}
taught,
{\ADD{and}}
loveth to tread
out
{\ADD{the corn;}} but I passed
over upon her
fair
neck: I will make
Ephraim to
ride;
Judah shall
plow,
{\ADD{and}}
Jacob shall break his
clods.\FTNT{***}{{\FR 10.11 }
her…: Heb. the beauty of her neck}
\VS{12}Sow to yourselves in
righteousness,
reap
in
mercy; break
up your fallow
ground: for
{\ADD{it is}}
time to
seek the
{\ND{
LORD}}, till he
come and
rain
righteousness upon you.
\VS{13}Ye have
plowed
wickedness, ye have
reaped
iniquity; ye have
eaten the
fruit of
lies: because thou didst
trust in thy
way, in the
multitude of thy mighty
men.
\VS{14}Therefore shall a
tumult
arise among thy
people, and all thy
fortresses shall be
spoiled, as
Shalman
spoiled
Beth-arbel in the
day of
battle: the
mother was dashed in
pieces upon
{\ADD{her}}
children.
\VS{15}So shall
Beth-el
do unto you
because of your
great
wickedness: in a
morning shall the
king of
Israel
utterly be cut
off.\FTNT{†††}{{\FR 10.15 }
your…: Heb. the evil of your evil}

\par }\Chap{11}{\PP \VerseOne{1}When
Israel
{\ADD{was}} a
child, then I
loved him, and
called my
son out of
Egypt.
\VS{2}{\ADD{As}} they
called them, so they
went from
them: they
sacrificed unto
Baalim, and burned
incense to graven
images.
\VS{3}I
taught
Ephraim also to
go,
taking them by their
arms; but they
knew not that I
healed them.
\VS{4}I
drew them with
cords of a
man, with
bands of
love: and I was to them as they that take
off the
yoke on their
jaws, and I
laid
meat unto them.\FTNT{*}{{\FR 11.4 }
take off: Heb. lift up}
\par }{\PP \VS{5}¶ He shall not
return into the
land of
Egypt, but the
Assyrian shall be his
king, because they
refused to
return.
\VS{6}And the
sword shall
abide on his
cities, and shall
consume his
branches, and
devour
{\ADD{them}}, because of their own
counsels.
\VS{7}And my
people are
bent to
backsliding from me: though they
called them to the most
High, none at
all would
exalt
{\ADD{him}}.\FTNT{†}{{\FR 11.7 }
none…: Heb. together they exalted not}
\par }{\PP \VS{8}How shall I give thee
up,
Ephraim?
{\ADD{how}} shall I
deliver thee,
Israel? how shall I
make thee as
Admah?
{\ADD{how}} shall I
set thee as
Zeboim? mine
heart is
turned within me, my
repentings are
kindled
together.
\VS{9}I will not
execute the
fierceness of mine
anger, I will not
return to
destroy
Ephraim: for I
{\ADD{am}}
God, and not
man; the Holy
One in the
midst of thee: and I will not
enter into the
city.
\VS{10}They shall
walk
after the
{\ND{
LORD}}: he shall
roar like a
lion: when he shall
roar, then the
children shall
tremble from the
west.
\VS{11}They shall
tremble as a
bird out of
Egypt, and as a
dove out of the
land of
Assyria: and I will
place them in their
houses,
saith the
{\ND{
LORD}}.
\VS{12}Ephraim
compasseth me about with
lies, and the
house of
Israel with
deceit: but
Judah yet
ruleth with
God, and is
faithful with the
saints.\FTNT{‡}{{\FR 11.12 }
saints: or, most holy}

\par }\Chap{12}{\PP \VerseOne{1}Ephraim
feedeth on
wind, and
followeth after the east
wind: he
daily
increaseth
lies and
desolation; and they do
make a
covenant with the
Assyrians, and
oil is
carried into
Egypt.
\VS{2}The
{\ND{
LORD}} hath also a
controversy with
Judah, and will
punish
Jacob according to his
ways; according to his
doings will he
recompense him.\FTNT{*}{{\FR 12.2 }
punish: Heb. visit upon}
\par }{\PP \VS{3}¶ He took his
brother by the
heel in the
womb, and by his
strength he had
power with
God:\FTNT{†}{{\FR 12.3 }
had…: Heb. was a prince, or, behaved himself princely}
\VS{4}Yea, he had
power over the
angel, and
prevailed: he
wept, and made
supplication unto him: he
found him
{\ADD{in}}
Beth-el, and there he
spake with us;
\VS{5}Even the
{\ND{
LORD}}
God of
hosts; the
{\ND{
LORD}}
{\ADD{is}} his
memorial.
\VS{6}Therefore
turn thou to thy
God:
keep
mercy and
judgment, and
wait on thy
God
continually.
\par }{\PP \VS{7}¶
{\ADD{He is}} a
merchant, the
balances of
deceit
{\ADD{are}} in his
hand: he
loveth to
oppress.\FTNT{‡}{{\FR 12.7 }
a merchant: or, Canaan}\FTNT{§}{{\FR 12.7 }
oppress: or, deceive}
\VS{8}And
Ephraim
said, Yet I am become
rich, I have found me
out
substance:
{\ADD{in}} all my
labours they shall
find none
iniquity in me that
{\ADD{were}}
sin.\FTNT{**}{{\FR 12.8 }
in all…: or, all my labours suffice me not: he shall have punishment of iniquity in whom is sin}\FTNT{††}{{\FR 12.8 }
that: Heb. which}
\VS{9}And I
{\ADD{that am}} the
{\ND{
LORD}} thy
God from the
land of
Egypt will yet make thee to
dwell in
tabernacles, as in the
days of the solemn
feast.
\VS{10}I have also
spoken by the
prophets, and I have
multiplied
visions, and used
similitudes, by the
ministry of the
prophets.\FTNT{‡‡}{{\FR 12.10 }
ministry: Heb. hand}
\VS{11}{\ADD{Is there}}
iniquity
{\ADD{in}}
Gilead? surely they are
vanity: they
sacrifice
bullocks in
Gilgal; yea, their
altars
{\ADD{are}} as
heaps in the
furrows of the
fields.
\VS{12}And
Jacob
fled into the
country of
Syria, and
Israel
served for a
wife, and for a
wife he
kept
{\ADD{sheep}}.
\VS{13}And by a
prophet the
{\ND{
LORD}}
brought
Israel out of
Egypt, and by a
prophet was he
preserved.
\VS{14}Ephraim provoked
{\ADD{him}} to
anger most
bitterly: therefore shall he
leave his
blood upon him, and his
reproach shall his
Lord
return unto him.\FTNT{§§}{{\FR 12.14 }
most…: Heb. with bitternesses}\FTNT{***}{{\FR 12.14 }
blood: Heb. bloods}

\par }\Chap{13}{\PP \VerseOne{1}When
Ephraim
spake
trembling, he
exalted himself in
Israel; but when he
offended in
Baal, he
died.
\VS{2}And now they
sin more and
more, and have
made them molten
images of their
silver,
{\ADD{and}}
idols according to their own
understanding, all of it the
work of the
craftsmen: they
say of them, Let the
men that
sacrifice
kiss the
calves.\FTNT{*}{{\FR 13.2 }
they sin…: Heb. they add to sin}\FTNT{†}{{\FR 13.2 }
the men…: or, the sacrificers of men}
\VS{3}Therefore they shall be as the
morning
cloud, and as the
early
dew that passeth
away, as the
chaff
{\ADD{that}} is driven with the
whirlwind out of the
floor, and as the
smoke out of the
chimney.
\VS{4}Yet I
{\ADD{am}} the
{\ND{
LORD}} thy
God from the
land of
Egypt, and thou shalt
know no
god but
me: for
{\ADD{there is}} no
saviour beside
me.
\par }{\PP \VS{5}¶ I did
know thee in the
wilderness, in the
land of great
drought.\FTNT{‡}{{\FR 13.5 }
great…: Heb. droughts}
\VS{6}According to their
pasture, so were they
filled; they were
filled, and their
heart was
exalted; therefore have they
forgotten me.
\VS{7}Therefore I will be unto them as a
lion: as a
leopard by the
way will I
observe
{\ADD{them:}}
\VS{8}I will
meet them as a
bear
{\ADD{that is}}
bereaved
{\ADD{of her whelps}}, and will
rend the
caul of their
heart, and there will I
devour them like a
lion: the
wild
beast shall
tear them.\FTNT{§}{{\FR 13.8 }
wild…: Heb. beast of the field}
\par }{\PP \VS{9}¶ O
Israel, thou hast
destroyed thyself; but in me
{\ADD{is}} thine
help.\FTNT{**}{{\FR 13.9 }
is…: Heb. in thy help}
\VS{10}I
will be thy
king:
where
{\ADD{is any other}} that may
save thee in all thy
cities? and thy
judges of whom thou
saidst,
Give me a
king and
princes?\FTNT{††}{{\FR 13.10 }
I will…: rather, Where is thy king?}
\VS{11}I
gave thee a
king in mine
anger, and
took
{\ADD{him}} away in my
wrath.
\VS{12}The
iniquity of
Ephraim
{\ADD{is}} bound
up; his
sin
{\ADD{is}}
hid.
\VS{13}The
sorrows of a travailing
woman shall
come upon him: he
{\ADD{is}} an
unwise
son; for he should not
stay
long in
{\ADD{the place of}} the breaking
forth of
children.\FTNT{‡‡}{{\FR 13.13 }
long: Heb. a time}
\VS{14}I will
ransom them from the
power of the
grave; I will
redeem them from
death: O
death, I
will be thy
plagues; O
grave, I
will be thy
destruction:
repentance shall be
hid from mine
eyes.\FTNT{§§}{{\FR 13.14 }
power: Heb. hand}
\par }{\PP \VS{15}¶ Though he be
fruitful among
{\ADD{his}}
brethren, an east
wind shall
come, the
wind of the
{\ND{
LORD}} shall come
up from the
wilderness, and his
spring shall become
dry, and his
fountain shall be dried
up: he shall
spoil the
treasure of all
pleasant
vessels.\FTNT{***}{{\FR 13.15 }
pleasant…: Heb. vessels of desire}
\VS{16}Samaria shall become
desolate; for she hath
rebelled against her
God: they shall
fall by the
sword: their
infants shall be dashed in
pieces, and their women with
child shall be ripped
up.

\par }\Chap{14}{\PP \VerseOne{1}O
Israel,
return unto the
{\ND{
LORD}} thy
God; for thou hast
fallen by thine
iniquity.
\VS{2}Take with you
words, and
turn to the
{\ND{
LORD}}:
say unto him, Take
away all
iniquity, and
receive
{\ADD{us}}
graciously: so will we
render the
calves of our
lips.\FTNT{*}{{\FR 14.2 }
receive…: or, give good}
\VS{3}Asshur shall not
save us; we will not
ride upon
horses: neither will we
say any more to the
work of our
hands,
{\ADD{Ye are}} our
gods: for in thee the
fatherless findeth
mercy.
\par }{\PP \VS{4}¶ I will
heal their
backsliding, I will
love them
freely: for mine
anger is turned
away from him.
\VS{5}I will be as the
dew unto
Israel: he shall
grow as the
lily, and cast
forth his
roots as
Lebanon.\FTNT{†}{{\FR 14.5 }
grow: or, blossom}\FTNT{‡}{{\FR 14.5 }
cast…: Heb. strike}
\VS{6}His
branches shall
spread, and his
beauty shall be as the olive
tree, and his
smell as
Lebanon.\FTNT{§}{{\FR 14.6 }
spread: Heb. go}
\VS{7}They that
dwell under his
shadow shall
return; they shall
revive
{\ADD{as}} the
corn, and
grow as the
vine: the
scent thereof
{\ADD{shall be}} as the
wine of
Lebanon.\FTNT{**}{{\FR 14.7 }
grow: or, blossom}\FTNT{††}{{\FR 14.7 }
scent: or, memorial}
\par }{\PP \VS{8}Ephraim
{\ADD{shall say}}, What have I to do any more with
idols? I have
heard
{\ADD{him}}, and
observed him: I
{\ADD{am}} like a
green fir
tree. From me is thy
fruit
found.
\VS{9}Who
{\ADD{is}}
wise, and he shall
understand these
{\ADD{things}}?
prudent, and he shall
know them? for the
ways of the
{\ND{
LORD}}
{\ADD{are}}
right, and the
just shall
walk in them: but the
transgressors shall
fall therein.
\par }