\NormalFont\ShortTitle{Isaiah}
{\MT The Book of the Prophet Isaiah

\par }\ChapOne{1}{\PP \VerseOne{1}The
vision of
Isaiah the
son of
Amoz, which he
saw concerning
Judah and
Jerusalem in the
days of
Uzziah,
Jotham,
Ahaz,
{\ADD{and}}
Hezekiah,
kings of
Judah.
\par }{\PP \VS{2}Hear, O
heavens, and give
ear, O
earth: for the
{\ND{
LORD}} hath
spoken, I have
nourished and brought
up
children, and they have
rebelled against me.
\VS{3}The
ox
knoweth his
owner, and the
ass his
master’s
crib:
{\ADD{but}}
Israel doth not
know, my
people doth not
consider.
\VS{4}Ah
sinful
nation, a
people
laden with
iniquity, a
seed of
evildoers,
children that are
corrupters: they have
forsaken the
{\ND{
LORD}}, they have
provoked the Holy
One of
Israel unto
anger, they are gone
away
backward.\FTNT{*}{{\FR 1.4 }
laden: Heb. of heaviness}\FTNT{†}{{\FR 1.4 }
gone…: Heb. alienated, or, separated}
\par }{\PP \VS{5}¶ Why should ye be
stricken any more? ye will
revolt more and
more: the whole
head is
sick, and the whole
heart
faint.\FTNT{‡}{{\FR 1.5 }
revolt…: Heb. increase revolt}
\VS{6}From the
sole of the
foot even unto the
head
{\ADD{there is}} no
soundness in it;
{\ADD{but}}
wounds, and
bruises, and
putrifying
sores: they have not been
closed, neither bound
up, neither
mollified with
ointment.\FTNT{§}{{\FR 1.6 }
ointment: or, oil}
\VS{7}Your
country
{\ADD{is}}
desolate, your
cities
{\ADD{are}}
burned with
fire: your
land,
strangers
devour it in your presence, and
{\ADD{it is}}
desolate, as
overthrown by
strangers.\FTNT{**}{{\FR 1.7 }
overthrown…: Heb. the overthrow of}
\VS{8}And the
daughter of
Zion is
left as a
cottage in a
vineyard, as a
lodge in a garden of
cucumbers, as a
besieged
city.
\VS{9}Except the
{\ND{
LORD}} of
hosts had
left unto us a very
small
remnant, we should have
been as
Sodom,
{\ADD{and}} we should have been
like unto
Gomorrah.
\par }{\PP \VS{10}¶
Hear the
word of the
{\ND{
LORD}}, ye
rulers of
Sodom; give
ear unto the
law of our
God, ye
people of
Gomorrah.
\VS{11}To
what purpose
{\ADD{is}} the
multitude of your
sacrifices unto me?
saith the
{\ND{
LORD}}: I am
full of the burnt
offerings of
rams, and the
fat of fed
beasts; and I
delight not in the
blood of
bullocks, or of
lambs, or of he
goats.\FTNT{††}{{\FR 1.11 }
he goats: Heb. great he goats}
\VS{12}When ye
come to
appear
before me, who hath
required this at your
hand, to
tread my
courts?\FTNT{‡‡}{{\FR 1.12 }
to appear: Heb. to be seen}
\VS{13}Bring no
more
vain
oblations;
incense is an
abomination unto me; the new
moons and
sabbaths, the
calling of
assemblies, I cannot away
with;
{\ADD{it is}}
iniquity, even the solemn
meeting.\FTNT{§§}{{\FR 1.13 }
iniquity: or, grief}
\VS{14}Your new
moons and your appointed
feasts my
soul
hateth: they are a
trouble unto me; I am
weary to
bear
{\ADD{them}}.
\VS{15}And when ye spread
forth your
hands, I will
hide mine
eyes from you: yea, when ye make
many
prayers, I will not
hear: your
hands are
full of
blood.\FTNT{***}{{\FR 1.15 }
make…: Heb. multiply prayer}\FTNT{†††}{{\FR 1.15 }
blood: Heb. bloods}
\par }{\PP \VS{16}¶
Wash you, make you
clean; put
away the
evil of your
doings from
before mine
eyes;
cease to do
evil;
\VS{17}Learn to do
well;
seek
judgment,
relieve the
oppressed,
judge the
fatherless,
plead for the
widow.\FTNT{‡‡‡}{{\FR 1.17 }
relieve: or, righten}
\VS{18}Come
now, and let us reason
together,
saith the
{\ND{
LORD}}: though your
sins be as
scarlet, they shall be as
white as
snow; though they be
red like
crimson, they shall be as
wool.
\VS{19}If ye be
willing and
obedient, ye shall
eat the
good of the
land:
\VS{20}But if ye
refuse and
rebel, ye shall be
devoured with the
sword: for the
mouth of the
{\ND{
LORD}} hath
spoken
{\ADD{it}}.
\par }{\PP \VS{21}¶ How is the
faithful
city become an
harlot! it was
full of
judgment;
righteousness
lodged in it; but now
murderers.
\VS{22}Thy
silver is become
dross, thy
wine
mixed with
water:
\VS{23}Thy
princes
{\ADD{are}}
rebellious, and
companions of
thieves: every one
loveth
gifts, and
followeth after
rewards: they
judge not the
fatherless, neither doth the
cause of the
widow
come unto them.
\VS{24}Therefore
saith the
Lord, the
{\ND{
LORD}} of
hosts, the mighty
One of
Israel,
Ah, I will
ease me of mine
adversaries, and
avenge me of mine
enemies:
\par }{\PP \VS{25}¶ And I will
turn my
hand upon thee, and
purely purge
away thy
dross, and take
away all thy
tin:\FTNT{§§§}{{\FR 1.25 }
purely: Heb. according to pureness}
\VS{26}And I will
restore thy
judges as at the
first, and thy
counsellors as at the
beginning:
afterward thou shalt be
called, The
city of
righteousness, the
faithful
city.
\VS{27}Zion shall be
redeemed with
judgment, and her
converts with
righteousness.\FTNT{*}{{\FR 1.27 }
her…: or, they that return of her}
\par }{\PP \VS{28}¶ And the
destruction of the
transgressors and of the
sinners
{\ADD{shall be}}
together, and they that
forsake the
{\ND{
LORD}} shall be
consumed.\FTNT{†}{{\FR 1.28 }
destruction: Heb. breaking}
\VS{29}For they shall be
ashamed of the
oaks which ye have
desired, and ye shall be
confounded for the
gardens that ye have
chosen.
\VS{30}For ye shall be as an
oak whose
leaf
fadeth, and as a
garden that hath no
water.
\VS{31}And the
strong shall be as
tow, and the
maker of it as a
spark, and they shall
both
burn
together, and none shall
quench
{\ADD{them}}.\FTNT{‡}{{\FR 1.31 }
maker…: or, and his work}

\par }\Chap{2}{\PP \VerseOne{1}The
word that
Isaiah the
son of
Amoz
saw concerning
Judah and
Jerusalem.
\VS{2}And it shall come to pass in the
last
days,
{\ADD{that}} the
mountain of the
{\ND{
LORD’s}}
house shall be
established in the
top of the
mountains, and shall be
exalted above the
hills; and all
nations shall
flow unto it.\FTNT{*}{{\FR 2.2 }
established: or, prepared}
\VS{3}And
many
people shall
go and
say,
Come ye, and let us go
up to the
mountain of the
{\ND{
LORD}}, to the
house of the
God of
Jacob; and he will
teach us of his
ways, and we will
walk in his
paths: for out of
Zion shall go
forth the
law, and the
word of the
{\ND{
LORD}} from
Jerusalem.
\VS{4}And he shall
judge among the
nations, and shall
rebuke
many
people: and they shall
beat their
swords into
plowshares, and their
spears into
pruninghooks:
nation shall not lift
up
sword against
nation, neither shall they
learn
war any more.\FTNT{†}{{\FR 2.4 }
pruninghooks: or, scythes}
\VS{5}O
house of
Jacob,
come ye, and let us
walk in the
light of the
{\ND{
LORD}}.
\par }{\PP \VS{6}¶ Therefore thou hast
forsaken thy
people the
house of
Jacob, because they be
replenished from the
east, and
{\ADD{are}}
soothsayers like the
Philistines, and they
please themselves in the
children of
strangers.\FTNT{‡}{{\FR 2.6 }
from…: or, more than the}\FTNT{§}{{\FR 2.6 }
please…: or, abound with the}
\VS{7}Their
land also is
full of
silver and
gold, neither
{\ADD{is there any}}
end of their
treasures; their
land is also
full of
horses, neither
{\ADD{is there any}}
end of their
chariots:
\VS{8}Their
land also is
full of
idols; they
worship the
work of their own
hands, that which their own
fingers have
made:
\VS{9}And the mean
man boweth
down, and the great
man
humbleth himself: therefore
forgive them not.
\par }{\PP \VS{10}¶
Enter into the
rock, and
hide thee in the
dust,
for
fear of the
{\ND{
LORD}}, and for the
glory of his
majesty.
\VS{11}The
lofty
looks of
man shall be
humbled, and the
haughtiness of
men shall be bowed
down, and the
{\ND{
LORD}} alone shall be
exalted in that
day.
\VS{12}For the
day of the
{\ND{
LORD}} of
hosts
{\ADD{shall be}} upon every
{\ADD{one that is}}
proud and
lofty, and upon every
{\ADD{one that is}} lifted
up; and he shall be brought
low:
\VS{13}And upon all the
cedars of
Lebanon,
{\ADD{that are}}
high and lifted
up, and upon all the
oaks of
Bashan,
\VS{14}And upon all the
high
mountains, and upon all the
hills
{\ADD{that are}} lifted
up,
\VS{15}And upon every
high
tower, and upon every
fenced
wall,
\VS{16}And upon all the
ships of
Tarshish, and upon all
pleasant
pictures.\FTNT{**}{{\FR 2.16 }
pleasant…: Heb. pictures of desire}
\VS{17}And the
loftiness of
man shall be bowed
down, and the
haughtiness of
men shall be made
low: and the
{\ND{
LORD}} alone shall be
exalted in that
day.
\VS{18}And the
idols he shall
utterly
abolish.\FTNT{††}{{\FR 2.18 }
he…: or, shall utterly pass away}
\VS{19}And they shall
go into the
holes of the
rocks, and into the
caves of the
earth,
for
fear of the
{\ND{
LORD}}, and for the
glory of his
majesty, when he
ariseth to shake
terribly the
earth.\FTNT{‡‡}{{\FR 2.19 }
of the earth: Heb. of the dust}
\VS{20}In that
day a
man shall
cast his
idols of
silver, and his
idols of
gold, which they made
{\ADD{each one}} for
himself to
worship, to the
moles and to the
bats;\FTNT{§§}{{\FR 2.20 }
his idols of silver…: Heb. the idols of his silver, etc}\FTNT{***}{{\FR 2.20 }
each…: or, for him}
\VS{21}To
go into the
clefts of the
rocks, and into the
tops of the ragged
rocks,
for
fear of the
{\ND{
LORD}}, and for the
glory of his
majesty, when he
ariseth to shake
terribly the
earth.
\VS{22}Cease ye from
man, whose
breath
{\ADD{is}} in his
nostrils: for wherein is he to be accounted
of?

\par }\Chap{3}{\PP \VerseOne{1}For, behold, the
Lord, the
{\ND{
LORD}} of
hosts, doth take
away from
Jerusalem and from
Judah the
stay and the
staff, the whole
stay of
bread, and the
whole
stay of
water,
\VS{2}The mighty
man, and the
man of
war, the
judge, and the
prophet, and the
prudent, and the
ancient,
\VS{3}The
captain of
fifty, and the
honourable
man, and the
counsellor, and the
cunning
artificer, and the
eloquent
orator.\FTNT{*}{{\FR 3.3 }
the honourable…: Heb. a man eminent in countenance}\FTNT{†}{{\FR 3.3 }
eloquent…: or, skilful of speech}
\VS{4}And I will
give
children
{\ADD{to be}} their
princes, and
babes shall
rule over them.
\VS{5}And the
people shall be
oppressed, every
one by
another, and every
one by his
neighbour: the
child shall behave himself
proudly against the
ancient, and the
base against the
honourable.
\VS{6}When a
man shall take
hold of his
brother of the
house of his
father,
{\ADD{saying}}, Thou hast
clothing, be thou our
ruler, and
{\ADD{let}} this
ruin
{\ADD{be}} under thy
hand:
\VS{7}In that
day shall he
swear,
saying, I will not be an
healer; for in my
house
{\ADD{is}} neither
bread nor
clothing:
make me not a
ruler of the
people.\FTNT{‡}{{\FR 3.7 }
swear: Heb. lift up the hand}\FTNT{§}{{\FR 3.7 }
healer: Heb. binder up}
\VS{8}For
Jerusalem is
ruined, and
Judah is
fallen: because their
tongue and their
doings
{\ADD{are}} against the
{\ND{
LORD}}, to
provoke the
eyes of his
glory.
\par }{\PP \VS{9}¶ The
shew of their
countenance doth witness against
them; and they
declare their
sin as
Sodom, they
hide
{\ADD{it}} not.
Woe unto their
soul! for they have
rewarded
evil unto themselves.
\VS{10}Say ye to the
righteous, that
{\ADD{it shall be}}
well
{\ADD{with him:}} for they shall
eat the
fruit of their
doings.
\VS{11}Woe unto the
wicked!
{\ADD{it shall be}}
ill
{\ADD{with him:}} for the
reward of his
hands shall be
given him.\FTNT{**}{{\FR 3.11 }
given…: Heb. done to him}
\par }{\PP \VS{12}¶
{\ADD{As for}} my
people,
children
{\ADD{are}} their
oppressors, and
women
rule over them. O my
people, they which
lead thee cause
{\ADD{thee}} to
err, and
destroy the
way of thy
paths.\FTNT{††}{{\FR 3.12 }
lead…: or, call thee blessed}\FTNT{‡‡}{{\FR 3.12 }
destroy: Heb. swallow up}
\VS{13}The
{\ND{
LORD}} standeth
up to
plead, and
standeth to
judge the
people.
\VS{14}The
{\ND{
LORD}} will
enter into
judgment with the
ancients of his
people, and the
princes thereof: for ye have eaten
up the
vineyard; the
spoil of the
poor
{\ADD{is}} in your
houses.\FTNT{§§}{{\FR 3.14 }
eaten: or, burnt}
\VS{15}What mean ye
{\ADD{that}} ye
beat my
people to
pieces, and
grind the
faces of the
poor?
saith the
Lord
{\ND{
GOD}} of
hosts.
\par }{\PP \VS{16}¶ Moreover the
{\ND{
LORD}}
saith,
Because the
daughters of
Zion are
haughty, and
walk with stretched
forth
necks and
wanton
eyes,
walking and
mincing
{\ADD{as}} they
go, and making a
tinkling with their
feet:\FTNT{***}{{\FR 3.16 }
wanton…: Heb. deceiving with their eyes}\FTNT{†††}{{\FR 3.16 }
mincing: or, tripping nicely}
\VS{17}Therefore the
Lord will smite with a
scab the crown of the
head of the
daughters of
Zion, and the
{\ND{
LORD}} will
discover their secret
parts.\FTNT{‡‡‡}{{\FR 3.17 }
discover: Heb. make naked}
\VS{18}In that
day the
Lord will take
away the
bravery of
{\ADD{their}} tinkling
ornaments
{\ADD{about their feet}}, and
{\ADD{their}}
cauls, and
{\ADD{their}} round tires like the
moon,\FTNT{§§§}{{\FR 3.18 }
cauls: or, networks}
\VS{19}The
chains, and the
bracelets, and the
mufflers,\FTNT{*}{{\FR 3.19 }
chains: or, sweet balls}\FTNT{†}{{\FR 3.19 }
mufflers: or, spangled ornaments}
\VS{20}The
bonnets, and the ornaments of the
legs, and the
headbands, and the
tablets, and the
earrings,\FTNT{‡}{{\FR 3.20 }
tablets: Heb. houses of the soul}
\VS{21}The
rings, and
nose
jewels,
\VS{22}The changeable suits of
apparel, and the
mantles, and the
wimples, and the crisping
pins,
\VS{23}The
glasses, and the fine
linen, and the
hoods, and the
vails.
\VS{24}And it shall come to pass,
{\ADD{that}} instead of sweet
smell there shall be
stink; and instead of a
girdle a
rent; and instead of well
set
hair
baldness; and instead of a
stomacher a
girding of
sackcloth;
{\ADD{and}}
burning instead of
beauty.
\VS{25}Thy
men shall
fall by the
sword, and thy
mighty in the
war.\FTNT{§}{{\FR 3.25 }
mighty: Heb. might}
\VS{26}And her
gates shall
lament and
mourn; and she
{\ADD{being}}
desolate shall
sit upon the
ground.\FTNT{**}{{\FR 3.26 }
desolate: or, emptied: Heb. cleansed}

\par }\Chap{4}{\PP \VerseOne{1}And in that
day
seven
women shall take
hold of
one
man,
saying, We will
eat our own
bread, and
wear our own
apparel: only let us be
called by thy
name, to take
away our
reproach.\FTNT{*}{{\FR 4.1 }
let…: Heb. let thy name be called upon us}\FTNT{†}{{\FR 4.1 }
to take…: or, take thou away}
\par }{\PP \VS{2}In that
day shall the
branch of the
{\ND{
LORD}} be
beautiful and
glorious, and the
fruit of the
earth
{\ADD{shall be}}
excellent and comely
for them that are
escaped of
Israel.\FTNT{‡}{{\FR 4.2 }
beautiful…: Heb. beauty and glory}\FTNT{§}{{\FR 4.2 }
them…: Heb. the escaping}
\VS{3}And it shall come to pass,
{\ADD{that he that is}}
left in
Zion, and
{\ADD{he that}}
remaineth in
Jerusalem, shall be
called
holy,
{\ADD{even}} every one that is
written among the
living in
Jerusalem:\FTNT{**}{{\FR 4.3 }
among…: or, to life}
\VS{4}When the
Lord shall have washed
away the
filth of the
daughters of
Zion, and shall have
purged the
blood of
Jerusalem from the
midst thereof by the
spirit of
judgment, and by the
spirit of
burning.
\VS{5}And the
{\ND{
LORD}} will
create upon every dwelling
place of
mount
Zion, and upon her
assemblies, a
cloud and
smoke by
day, and the
shining of a
flaming
fire by
night: for upon all the
glory
{\ADD{shall be}} a
defence.\FTNT{††}{{\FR 4.5 }
upon all: or, above all}\FTNT{‡‡}{{\FR 4.5 }
defence: Heb. covering}
\VS{6}And there shall be a
tabernacle for a
shadow in the
daytime from the
heat, and for a place of
refuge, and for a
covert from
storm and from
rain.

\par }\Chap{5}{\PP \VerseOne{1}Now will I
sing to my
wellbeloved a
song of my
beloved touching his
vineyard. My
wellbeloved hath a
vineyard in a very
fruitful
hill:\FTNT{*}{{\FR 5.1 }
a very…: Heb. the horn of the son of oil}
\VS{2}And he
fenced it, and gathered out the
stones thereof, and
planted it with the choicest
vine, and
built a
tower in the
midst of it, and also
made a
winepress therein: and he
looked that it should bring
forth
grapes, and it brought
forth wild
grapes.\FTNT{†}{{\FR 5.2 }
fenced: or, made a wall about}\FTNT{‡}{{\FR 5.2 }
made: Heb. hewed}
\VS{3}And now, O
inhabitants of
Jerusalem, and
men of
Judah,
judge, I pray you, betwixt me and my
vineyard.
\VS{4}What could have been
done more to my
vineyard, that I have not
done in it?
wherefore, when I
looked that it should bring
forth
grapes, brought it
forth wild
grapes?
\VS{5}And now go to; I will
tell you what I will
do to my
vineyard: I will take
away the
hedge thereof, and it shall be eaten
up;
{\ADD{and}} break
down the
wall thereof, and it shall be trodden
down:\FTNT{§}{{\FR 5.5 }
trodden…: Heb. for a treading}
\VS{6}And I will
lay it
waste: it shall not be
pruned, nor
digged; but there shall come
up
briers and
thorns: I will also
command the
clouds that they
rain no
rain upon it.
\VS{7}For the
vineyard of the
{\ND{
LORD}} of
hosts
{\ADD{is}} the
house of
Israel, and the
men of
Judah his
pleasant
plant: and he
looked for
judgment, but behold
oppression; for
righteousness, but behold a
cry.\FTNT{**}{{\FR 5.7 }
his…: Heb. plant of his pleasures}\FTNT{††}{{\FR 5.7 }
oppression: Heb. a scab}
\par }{\PP \VS{8}¶
Woe unto them that
join
house to
house,
{\ADD{that}}
lay
field to
field, till
{\ADD{there be}}
no
place, that they may be
placed alone in the
midst of the
earth!\FTNT{‡‡}{{\FR 5.8 }
they: Heb. ye}
\VS{9}In mine
ears
{\ADD{said}} the
{\ND{
LORD}} of
hosts, Of a
truth
many
houses shall be
desolate,
{\ADD{even}}
great and
fair, without
inhabitant.\FTNT{§§}{{\FR 5.9 }
In…: or, This is in mine ears, saith the
{\ND{LORD}}, etc}\FTNT{***}{{\FR 5.9 }
Of a…: Heb. If not, etc}
\VS{10}Yea,
ten
acres of
vineyard shall
yield
one
bath, and the
seed of an
homer shall
yield an
ephah.
\par }{\PP \VS{11}¶
Woe unto them that rise up
early in the
morning,
{\ADD{that}} they may
follow strong
drink; that
continue until
night,
{\ADD{till}}
wine
inflame them!\FTNT{†††}{{\FR 5.11 }
inflame: of, pursue}
\VS{12}And the
harp, and the
viol, the
tabret, and
pipe, and
wine, are in their
feasts: but they
regard not the
work of the
{\ND{
LORD}}, neither
consider the
operation of his
hands.
\par }{\PP \VS{13}¶ Therefore my
people are gone into
captivity, because
{\ADD{they have}} no
knowledge: and their
honourable
men
{\ADD{are}}
famished, and their
multitude dried
up with
thirst.\FTNT{‡‡‡}{{\FR 5.13 }
honourable…: Heb. glory are men of famine}
\VS{14}Therefore
hell hath
enlarged
herself, and
opened her
mouth without
measure: and their
glory, and their
multitude, and their
pomp, and he that
rejoiceth, shall
descend into it.
\VS{15}And the mean
man shall be brought
down, and the mighty
man shall be
humbled, and the
eyes of the
lofty shall be
humbled:
\VS{16}But the
{\ND{
LORD}} of
hosts shall be
exalted in
judgment, and
God that is
holy shall be
sanctified in
righteousness.\FTNT{§§§}{{\FR 5.16 }
God…: or, the holy God: Heb. the God the holy}
\VS{17}Then shall the
lambs
feed after their
manner, and the waste
places of the fat
ones shall
strangers
eat.
\par }{\PP \VS{18}Woe unto them that
draw
iniquity with
cords of
vanity, and
sin as it were with a
cart
rope:
\VS{19}That
say, Let him make
speed,
{\ADD{and}}
hasten his
work, that we may
see
{\ADD{it:}} and let the
counsel of the Holy
One of
Israel draw
nigh and
come, that we may
know
{\ADD{it}}!
\par }{\PP \VS{20}¶
Woe unto them that
call
evil
good, and
good
evil; that
put
darkness for
light, and
light for
darkness; that
put
bitter for
sweet, and
sweet for
bitter!\FTNT{*}{{\FR 5.20 }
call…: Heb. say concerning evil, It is good, etc}
\VS{21}Woe unto
{\ADD{them that are}}
wise in their own
eyes, and
prudent in their own
sight!\FTNT{†}{{\FR 5.21 }
in their own sight: Heb. before their face}
\VS{22}Woe unto
{\ADD{them that are}}
mighty to
drink
wine, and
men of
strength to
mingle strong
drink:
\VS{23}Which
justify the
wicked
for
reward, and take
away the
righteousness of the
righteous from him!
\VS{24}Therefore as the
fire
devoureth the
stubble, and the
flame
consumeth the
chaff,
{\ADD{so}} their
root shall be as
rottenness, and their
blossom shall go
up as
dust: because they have cast
away the
law of the
{\ND{
LORD}} of
hosts, and
despised the
word of the Holy
One of
Israel.\FTNT{‡}{{\FR 5.24 }
fire: Heb. tongue of fire}
\VS{25}Therefore is the
anger of the
{\ND{
LORD}}
kindled against his
people, and he hath stretched
forth his
hand against them, and hath
smitten them: and the
hills did
tremble, and their
carcases
{\ADD{were}}
torn in the
midst of the
streets. For all this his
anger is not turned
away, but his
hand
{\ADD{is}} stretched
out still.\FTNT{§}{{\FR 5.25 }
torn: or, as dung}
\par }{\PP \VS{26}¶ And he will lift
up an
ensign to the
nations from
far, and will
hiss unto them from the
end of the
earth: and, behold, they shall
come with
speed
swiftly:
\VS{27}None shall be
weary nor
stumble among them; none shall
slumber nor
sleep; neither shall the
girdle of their
loins be
loosed, nor the
latchet of their
shoes be
broken:
\VS{28}Whose
arrows
{\ADD{are}}
sharp, and all their
bows
bent, their
horses’
hoofs shall be
counted like
flint, and their
wheels like a
whirlwind:
\VS{29}Their
roaring
{\ADD{shall be}} like a
lion, they shall
roar like young
lions: yea, they shall
roar, and lay
hold of the
prey, and shall carry
{\ADD{it}} away
safe, and none shall
deliver
{\ADD{it}}.
\VS{30}And in that
day they shall
roar against them like the
roaring of the
sea: and if
{\ADD{one}}
look unto the
land, behold
darkness
{\ADD{and}}
sorrow, and the
light is
darkened in the
heavens thereof.\FTNT{**}{{\FR 5.30 }
sorrow: or, distress}\FTNT{††}{{\FR 5.30 }
and the light…: or, when it is light, it shall be dark in the destructions thereof}

\par }\Chap{6}{\PP \VerseOne{1}In the
year that
king
Uzziah
died I
saw
also the
Lord
sitting upon a
throne,
high and lifted
up, and his
train
filled the
temple.\FTNT{*}{{\FR 6.1 }
his…: or, the skirts thereof}
\VS{2}Above it
stood the
seraphims: each
one had
six
wings; with
twain he
covered his
face, and with
twain he
covered his
feet, and with
twain he did
fly.
\VS{3}And one
cried unto another, and
said,
Holy,
holy,
holy,
{\ADD{is}} the
{\ND{
LORD}} of
hosts: the whole
earth
{\ADD{is}}
full of his
glory.\FTNT{†}{{\FR 6.3 }
one…: Heb. this cried to this}\FTNT{‡}{{\FR 6.3 }
the whole…: Heb. his glory is the fulness of the whole earth}
\VS{4}And the
posts of the
door
moved at the
voice of him that
cried, and the
house was
filled with
smoke.\FTNT{§}{{\FR 6.4 }
door: Heb. thresholds}
\par }{\PP \VS{5}¶ Then
said I,
Woe
{\ADD{is}} me! for I am
undone; because I
{\ADD{am}} a
man of
unclean
lips, and I
dwell in the
midst of a
people of
unclean
lips: for mine
eyes have
seen the
King, the
{\ND{
LORD}} of
hosts.\FTNT{**}{{\FR 6.5 }
undone: Heb. cut off}
\VS{6}Then
flew
one of the
seraphims unto me, having a live
coal in his
hand,
{\ADD{which}} he had
taken with the
tongs from off the
altar:\FTNT{††}{{\FR 6.6 }
having…: Heb. and in his hand a live coal}
\VS{7}And he
laid
{\ADD{it}} upon my
mouth, and
said, Lo, this hath
touched thy
lips; and thine
iniquity is taken
away, and thy
sin
purged.\FTNT{‡‡}{{\FR 6.7 }
laid…: Heb. caused it to touch}
\VS{8}Also I
heard the
voice of the
Lord,
saying, Whom shall I
send, and who will
go for us? Then
said I, Here
{\ADD{am}} I;
send me.\FTNT{§§}{{\FR 6.8 }
Here…: Heb. behold me}
\par }{\PP \VS{9}¶ And he
said,
Go, and
tell this
people,
Hear ye
indeed, but
understand not; and
see ye
indeed, but
perceive not.\FTNT{***}{{\FR 6.9 }
indeed, but understand: or, without ceasing, etc: Heb. in hearing, etc}
\VS{10}Make the
heart of this
people
fat, and make their
ears
heavy, and
shut their
eyes; lest they
see with their
eyes, and
hear with their
ears, and
understand with their
heart, and
convert, and be
healed.
\VS{11}Then
said I,
Lord, how long? And he
answered, Until the
cities be
wasted without
inhabitant, and the
houses without
man, and the
land be
utterly
desolate,\FTNT{†††}{{\FR 6.11 }
utterly…: Heb. desolate with desolation}
\VS{12}And the
{\ND{
LORD}} have
removed
men far
away, and
{\ADD{there be}} a
great
forsaking in the
midst of the
land.
\par }{\PP \VS{13}¶ But yet in it
{\ADD{shall be}} a
tenth, and
{\ADD{it}} shall
return, and shall be
eaten: as a teil
tree, and as an
oak, whose
substance
{\ADD{is}} in them, when they
cast
{\ADD{their leaves: so}} the
holy
seed
{\ADD{shall be}} the
substance thereof.\FTNT{‡‡‡}{{\FR 6.13 }
and it…: or, when it is returned, and hath been broused}\FTNT{§§§}{{\FR 6.13 }
substance: or, stock, or, stem}

\par }\Chap{7}{\PP \VerseOne{1}And it came to pass in the
days of
Ahaz the
son of
Jotham, the
son of
Uzziah,
king of
Judah,
{\ADD{that}}
Rezin the
king of
Syria, and
Pekah the
son of
Remaliah,
king of
Israel, went
up toward
Jerusalem to
war against it, but
could not
prevail against it.
\VS{2}And it was
told the
house of
David,
saying,
Syria is
confederate with
Ephraim. And his
heart was
moved, and the
heart of his
people, as the
trees of the
wood are
moved
with the
wind.\FTNT{*}{{\FR 7.2 }
is confederate…: Heb. resteth on}
\VS{3}Then
said the
{\ND{
LORD}} unto
Isaiah, Go
forth now to
meet
Ahaz, thou, and
Shear-jashub thy
son, at the
end of the
conduit of the
upper
pool in the
highway of the
fuller’s
field;\FTNT{†}{{\FR 7.3 }
Shear-jashub: that is, The remnant shall return}\FTNT{‡}{{\FR 7.3 }
highway: or, causeway}
\VS{4}And
say unto him, Take
heed, and be
quiet;
fear not, neither be
fainthearted for the
two
tails of these
smoking
firebrands, for the
fierce
anger of
Rezin with
Syria, and of the
son of
Remaliah.\FTNT{§}{{\FR 7.4 }
neither…: Heb. let not thy heart be tender}
\VS{5}Because
Syria,
Ephraim, and the
son of
Remaliah, have taken
evil
counsel against thee,
saying,
\VS{6}Let us go
up against
Judah, and
vex it, and let us make a
breach therein for us, and
set a
king in the
midst of it,
{\ADD{even}} the
son of
Tabeal:\FTNT{**}{{\FR 7.6 }
vex: or, waken}
\VS{7}Thus
saith the
Lord
{\ND{
GOD}}, It shall not
stand, neither shall it come to pass.
\VS{8}For the
head of
Syria
{\ADD{is}}
Damascus, and the
head of
Damascus
{\ADD{is}}
Rezin; and within
threescore and
five
years shall
Ephraim be
broken, that it be not a
people.\FTNT{††}{{\FR 7.8 }
that…: Heb. from a people}
\VS{9}And the
head of
Ephraim
{\ADD{is}}
Samaria, and the
head of
Samaria
{\ADD{is}}
Remaliah’s
son. If ye will not
believe, surely ye shall not be
established.\FTNT{‡‡}{{\FR 7.9 }
If…: or, Do ye not believe? it is because ye are not stable}
\par }{\PP \VS{10}¶ Moreover the
{\ND{
LORD}}
spake
again unto
Ahaz,
saying,\FTNT{§§}{{\FR 7.10 }
Moreover…: Heb. And the
{\ND{LORD}} added to speak}
\VS{11}Ask thee a
sign of the
{\ND{
LORD}} thy
God;
ask it either in the
depth, or in the
height
above.\FTNT{***}{{\FR 7.11 }
ask it…: or, make thy petition deep}
\VS{12}But
Ahaz
said, I will not
ask, neither will I
tempt the
{\ND{
LORD}}.
\VS{13}And he
said,
Hear ye now, O
house of
David;
{\ADD{Is it}} a small
thing for you to
weary
men, but will ye
weary my
God also?
\VS{14}Therefore the
Lord himself shall
give you a
sign; Behold, a
virgin shall
conceive, and
bear a
son, and shall
call his
name
Immanuel.\FTNT{†††}{{\FR 7.14 }
shall call: or, thou, O virgin, shalt call}
\VS{15}Butter and
honey shall he
eat, that he may
know to
refuse the
evil, and
choose the
good.
\VS{16}For before the
child shall
know to
refuse the
evil, and
choose the
good, the
land that thou
abhorrest shall
be
forsaken of
both her
kings.
\par }{\PP \VS{17}¶ The
{\ND{
LORD}} shall
bring upon thee, and upon thy
people, and upon thy
father’s
house,
days that have not
come, from the
day that
Ephraim
departed from
Judah;
{\ADD{even}} the
king of
Assyria.
\VS{18}And it shall come to pass in that
day,
{\ADD{that}} the
{\ND{
LORD}} shall
hiss for the
fly that
{\ADD{is}} in the uttermost
part of the
rivers of
Egypt, and for the
bee that
{\ADD{is}} in the
land of
Assyria.
\VS{19}And they shall
come, and shall
rest all of them in the
desolate
valleys, and in the
holes of the
rocks, and upon all
thorns, and upon all
bushes.\FTNT{‡‡‡}{{\FR 7.19 }
bushes: or, commendable trees}
\VS{20}In the same
day shall the
Lord
shave with a
razor that is
hired,
{\ADD{namely}}, by them
beyond the
river, by the
king of
Assyria, the
head, and the
hair of the
feet: and it shall also
consume the
beard.
\VS{21}And it shall come to pass in that
day,
{\ADD{that}} a
man shall
nourish a
young
cow, and
two
sheep;
\VS{22}And it shall come to pass, for the
abundance of
milk
{\ADD{that}} they shall
give he shall
eat
butter: for
butter and
honey shall every one
eat that is
left
in the
land.\FTNT{§§§}{{\FR 7.22 }
the land: Heb. the midst of the land}
\VS{23}And it shall come to pass in that
day,
{\ADD{that}} every
place shall be, where there were a
thousand
vines at a
thousand
silverlings, it shall
{\ADD{even}} be for
briers and
thorns.
\VS{24}With
arrows and with
bows shall
{\ADD{men}}
come thither; because all the
land shall become
briers and
thorns.
\VS{25}And
{\ADD{on}} all
hills that shall be
digged with the
mattock, there shall not come
thither the
fear of
briers and
thorns: but it shall be for the sending
forth of
oxen, and for the
treading of lesser
cattle.

\par }\Chap{8}{\PP \VerseOne{1}Moreover the
{\ND{
LORD}}
said unto me,
Take thee a
great
roll, and
write in it with a
man’s
pen concerning
Maher-shalal-hash-baz.\FTNT{*}{{\FR 8.1 }
Maher-shalal-hash-baz: Heb. In making speed to the spoil he hasteneth the prey, or, Make speed, etc}
\VS{2}And I
took unto me
faithful
witnesses to
record,
Uriah the
priest, and
Zechariah the
son of
Jeberechiah.
\VS{3}And I
went unto the
prophetess; and she
conceived, and
bare a
son. Then
said the
{\ND{
LORD}} to me,
Call his
name
Maher-shalal-hash-baz.\FTNT{†}{{\FR 8.3 }
went: Heb. approached}
\VS{4}For before the
child shall have
knowledge to
cry, My
father, and my
mother, the
riches of
Damascus and the
spoil of
Samaria shall be taken
away
before the
king of
Assyria.\FTNT{‡}{{\FR 8.4 }
the riches…: or, he that is before the king of Assyria shall take away the riches, etc}
\par }{\PP \VS{5}¶ The
{\ND{
LORD}}
spake also unto me
again,
saying,
\VS{6}Forasmuch as this
people
refuseth the
waters of
Shiloah that
go
softly, and
rejoice in
Rezin and
Remaliah’s
son;
\VS{7}Now therefore, behold, the
Lord bringeth
up upon them the
waters of the
river,
strong and
many,
{\ADD{even}} the
king of
Assyria, and all his
glory: and he shall come
up over all his
channels, and go
over all his
banks:
\VS{8}And he shall
pass through
Judah; he shall
overflow and go
over, he shall
reach
{\ADD{even}} to the
neck; and the stretching
out of his
wings shall
fill the
breadth of thy
land, O
Immanuel.\FTNT{§}{{\FR 8.8 }
stretching…: Heb. fulness of the breadth of thy land shall be the stretchings out of his wings}
\par }{\PP \VS{9}¶
Associate yourselves, O ye
people, and ye shall be broken in
pieces; and give
ear,
all ye of
far
countries:
gird yourselves, and ye shall be broken in
pieces;
gird yourselves, and ye shall be broken in
pieces.\FTNT{**}{{\FR 8.9 }
people, and: or, people, yet}
\VS{10}Take
counsel
together, and it shall come to
nought;
speak the
word, and it shall not
stand: for
God
{\ADD{is}} with us.
\par }{\PP \VS{11}¶ For the
{\ND{
LORD}}
spake thus to me with a
strong
hand, and
instructed me that I should not
walk in the
way of this
people,
saying,\FTNT{††}{{\FR 8.11 }
with…: Heb. in strength of hand}
\VS{12}Say ye not, A
confederacy, to all
{\ADD{them to}} whom this
people shall
say, A
confederacy; neither
fear ye their
fear, nor be
afraid.
\VS{13}Sanctify the
{\ND{
LORD}} of
hosts himself; and
{\ADD{let}} him
{\ADD{be}} your
fear, and
{\ADD{let}} him
{\ADD{be}} your
dread.
\VS{14}And he shall be for a
sanctuary; but for a
stone of
stumbling and for a
rock of
offence to
both the
houses of
Israel, for a
gin and for a
snare to the
inhabitants of
Jerusalem.
\VS{15}And
many among them shall
stumble, and
fall, and be
broken, and be
snared, and be
taken.
\par }{\PP \VS{16}Bind
up the
testimony,
seal the
law among my
disciples.
\VS{17}And I will
wait upon the
{\ND{
LORD}}, that
hideth his
face from the
house of
Jacob, and I will
look for him.
\VS{18}Behold, I and the
children whom the
{\ND{
LORD}} hath
given me
{\ADD{are}} for
signs and for
wonders in
Israel from the
{\ND{
LORD}} of
hosts, which
dwelleth in
mount
Zion.
\par }{\PP \VS{19}¶ And when they shall
say unto you,
Seek unto them that have familiar
spirits, and unto
wizards that
peep, and that
mutter: should not a
people
seek unto their
God? for the
living to the
dead?
\VS{20}To the
law and to the
testimony: if they
speak not according to this
word,
{\ADD{it is}} because
{\ADD{there is}} no
light in them.\FTNT{‡‡}{{\FR 8.20 }
no…: Heb. no morning}
\VS{21}And they shall
pass through it, hardly
bestead and
hungry: and it shall come to pass, that when they shall be
hungry, they shall
fret themselves, and
curse their
king and their
God, and
look
upward.
\VS{22}And they shall
look unto the
earth; and behold
trouble and
darkness,
dimness of
anguish; and
{\ADD{they shall be}}
driven to
darkness.

\par }\Chap{9}{\PP \VerseOne{1}Nevertheless the
dimness
{\ADD{shall}} not
{\ADD{be}} such as
{\ADD{was}} in her
vexation,
when at the
first he lightly
afflicted the
land of
Zebulun and the
land of
Naphtali, and
afterward did more grievously
afflict
{\ADD{her by}} the
way of the
sea,
beyond
Jordan, in
Galilee of the
nations.\FTNT{*}{{\FR 9.1 }
of the nations: or, populous}
\VS{2}The
people that
walked in
darkness have
seen a
great
light: they that
dwell in the
land of the shadow of
death, upon them hath the
light
shined.
\VS{3}Thou hast
multiplied the
nation,
{\ADD{and}} not
increased the
joy: they
joy
before thee according to the
joy in
harvest,
{\ADD{and}} as
{\ADD{men}}
rejoice when they
divide the
spoil.\FTNT{†}{{\FR 9.3 }
not: or, to him}
\VS{4}For thou hast
broken the
yoke of his
burden, and the
staff of his
shoulder, the
rod of his
oppressor, as in the
day of
Midian.\FTNT{‡}{{\FR 9.4 }
For…: or, When thou brakest}
\VS{5}For every
battle of the
warrior
{\ADD{is}} with confused
noise, and
garments
rolled in
blood; but
{\ADD{this}} shall be with
burning
{\ADD{and}}
fuel of
fire.\FTNT{§}{{\FR 9.5 }
For…: or, When the whole battle of the warrior was, etc}\FTNT{**}{{\FR 9.5 }
but…: or, and it was, etc}\FTNT{††}{{\FR 9.5 }
fuel: Heb. meat}
\VS{6}For unto us a
child is
born, unto us a
son is
given: and the
government shall be upon his
shoulder: and his
name shall be
called
Wonderful,
Counsellor, The
mighty
God, The
everlasting
Father, The
Prince of
Peace.
\VS{7}Of the
increase of
{\ADD{his}}
government and
peace
{\ADD{there shall be}} no
end, upon the
throne of
David, and upon his
kingdom, to
order it, and to
establish it with
judgment and with
justice from henceforth even
for
ever. The
zeal of the
{\ND{
LORD}} of
hosts will
perform this.
\par }{\PP \VS{8}¶ The
Lord
sent a
word into
Jacob, and it hath
lighted upon
Israel.
\VS{9}And all the
people shall
know,
{\ADD{even}}
Ephraim and the
inhabitant of
Samaria, that
say in the
pride and
stoutness of
heart,
\VS{10}The
bricks are fallen
down, but we will
build with hewn
stones: the
sycomores are cut
down, but we will
change
{\ADD{them into}}
cedars.
\VS{11}Therefore the
{\ND{
LORD}} shall set
up the
adversaries of
Rezin against him, and
join his
enemies
together;\FTNT{‡‡}{{\FR 9.11 }
join: Heb. mingle}
\VS{12}The
Syrians
before, and the
Philistines
behind; and they shall
devour
Israel with open
mouth. For all this his
anger is not turned
away, but his
hand
{\ADD{is}} stretched out
still.\FTNT{§§}{{\FR 9.12 }
open: Heb. whole}
\par }{\PP \VS{13}¶ For the
people
turneth not unto him that
smiteth them, neither do they
seek the
{\ND{
LORD}} of
hosts.
\VS{14}Therefore the
{\ND{
LORD}} will cut
off from
Israel
head and
tail,
branch and
rush, in
one
day.
\VS{15}The
ancient and
honourable,
he
{\ADD{is}} the
head; and the
prophet that
teacheth
lies, he
{\ADD{is}} the
tail.
\VS{16}For the
leaders of this
people cause
{\ADD{them}} to
err; and
{\ADD{they that are}}
led of them
{\ADD{are}}
destroyed.\FTNT{***}{{\FR 9.16 }
the leaders: or, they that call them blessed}\FTNT{†††}{{\FR 9.16 }
led of: or, called blessed of}\FTNT{‡‡‡}{{\FR 9.16 }
destroyed: Heb. swallowed up}
\VS{17}Therefore the
Lord shall have no
joy in their young
men, neither shall have
mercy on their
fatherless and
widows: for every one
{\ADD{is}} an
hypocrite and an
evildoer, and every
mouth
speaketh
folly. For all this his
anger is not turned
away, but his
hand
{\ADD{is}} stretched out
still.\FTNT{§§§}{{\FR 9.17 }
folly: or, villany}
\par }{\PP \VS{18}¶ For
wickedness
burneth as the
fire: it shall
devour the
briers and
thorns, and shall
kindle in the
thickets of the
forest, and they shall mount
up
{\ADD{like}} the lifting
up of
smoke.
\VS{19}Through the
wrath of the
{\ND{
LORD}} of
hosts is the
land
darkened, and the
people shall be as the
fuel of the
fire: no
man shall
spare his
brother.\FTNT{*}{{\FR 9.19 }
fuel: Heb. meat}
\VS{20}And he shall
snatch on the right
hand, and be
hungry; and he shall
eat on the left
hand, and they shall not be
satisfied: they shall
eat every
man the
flesh of his own
arm:\FTNT{†}{{\FR 9.20 }
snatch: Heb cut}
\VS{21}Manasseh,
Ephraim; and
Ephraim,
Manasseh:
{\ADD{and}} they
together
{\ADD{shall be}} against
Judah. For all this his
anger is not turned
away, but his
hand
{\ADD{is}} stretched out
still.

\par }\Chap{10}{\PP \VerseOne{1}Woe unto them that
decree
unrighteous
decrees, and that
write
grievousness
{\ADD{which}} they have
prescribed;\FTNT{*}{{\FR 10.1 }
that write…: or, to the writers that write grievousness}
\VS{2}To turn
aside the
needy from
judgment, and to take
away the
right from the
poor of my
people, that
widows may be their
prey, and
{\ADD{that}} they may
rob the
fatherless!
\VS{3}And what will ye
do in the
day of
visitation, and in the
desolation
{\ADD{which}} shall
come from
far? to whom will ye
flee for
help? and where will ye
leave your
glory?
\VS{4}Without me they
shall bow
down under the
prisoners, and they shall
fall under the
slain. For all this his
anger is not turned
away, but his
hand
{\ADD{is}} stretched out
still.
\par }{\PP \VS{5}¶
O
Assyrian, the
rod of mine
anger, and the
staff in their
hand is mine
indignation.\FTNT{†}{{\FR 10.5 }
O…: or, Woe to the Assyrian: Heb. Asshur}\FTNT{‡}{{\FR 10.5 }
and: or, though}
\VS{6}I will
send him against an
hypocritical
nation, and against the
people of my
wrath will I give him a
charge, to
take the
spoil, and to
take the
prey, and to tread them
down like the
mire of the
streets.\FTNT{§}{{\FR 10.6 }
tread…: Heb. lay them a treading}
\VS{7}Howbeit he
meaneth not so, neither doth his
heart
think so; but
{\ADD{it is}} in his
heart to
destroy and cut
off
nations not a
few.
\VS{8}For he
saith,
{\ADD{Are}} not my
princes
altogether
kings?
\VS{9}{\ADD{Is}} not
Calno as
Carchemish?
{\ADD{is}} not
Hamath as
Arpad?
{\ADD{is}} not
Samaria as
Damascus?
\VS{10}As my
hand hath
found the
kingdoms of the
idols, and whose graven
images did excel them of
Jerusalem and of
Samaria;
\VS{11}Shall I not, as I have
done unto
Samaria and her
idols, so
do to
Jerusalem and her
idols?
\VS{12}Wherefore it shall come to pass,
{\ADD{that}} when the
Lord hath
performed his whole
work upon
mount
Zion and on
Jerusalem, I will
punish the
fruit of the
stout
heart of the
king of
Assyria, and the
glory of his
high
looks.\FTNT{**}{{\FR 10.12 }
punish: Heb. visit upon}\FTNT{††}{{\FR 10.12 }
stout…: Heb. greatness of the heart}
\VS{13}For he
saith, By the
strength of my
hand I have
done
{\ADD{it}}, and by my
wisdom; for I am
prudent: and I have
removed the
bounds of the
people, and have
robbed their
treasures, and I have put
down the
inhabitants like a
valiant
{\ADD{man:}}\FTNT{‡‡}{{\FR 10.13 }
a valiant…: or, many people}
\VS{14}And my
hand hath
found as a
nest the
riches of the
people: and as one
gathereth
eggs
{\ADD{that are}}
left, have I
gathered all the
earth; and there was none that
moved the
wing, or
opened the
mouth, or
peeped.
\VS{15}Shall the
axe
boast itself against him that
heweth therewith?
{\ADD{or}} shall the
saw
magnify itself against him that
shaketh it? as if the
rod should
shake
{\ADD{itself}} against them that lift it
up,
{\ADD{or}} as if the
staff should lift
up
{\ADD{itself, as if it were}} no
wood.\FTNT{§§}{{\FR 10.15 }
the rod…: or, a rod should shake them that lift it up}\FTNT{***}{{\FR 10.15 }
itself, as if…: or, that which is not wood}
\VS{16}Therefore shall the
Lord, the
Lord of
hosts,
send among his fat
ones
leanness; and under his
glory he shall
kindle a
burning like the burning of a
fire.
\VS{17}And the
light of
Israel shall be for a
fire, and his Holy
One for a
flame: and it shall
burn and
devour his
thorns and his
briers in
one
day;
\VS{18}And shall
consume the
glory of his
forest, and of his fruitful
field, both
soul and
body: and they shall be as when a
standardbearer
fainteth.\FTNT{†††}{{\FR 10.18 }
both…: Heb. from the soul, and even to the flesh}
\VS{19}And the
rest of the
trees of his
forest shall be
few, that a
child may
write them.\FTNT{‡‡‡}{{\FR 10.19 }
few: Heb. number}
\par }{\PP \VS{20}¶ And it shall come to pass in that
day,
{\ADD{that}} the
remnant of
Israel, and such as are
escaped of the
house of
Jacob, shall no more
again
stay upon him that
smote them; but shall
stay upon the
{\ND{
LORD}}, the Holy
One of
Israel, in
truth.
\VS{21}The
remnant shall
return,
{\ADD{even}} the
remnant of
Jacob, unto the
mighty
God.
\VS{22}For though thy
people
Israel be as the
sand of the
sea,
{\ADD{yet}} a
remnant of them shall
return: the
consumption
decreed shall
overflow with
righteousness.\FTNT{§§§}{{\FR 10.22 }
of them: Heb. in, or, among, etc}\FTNT{*}{{\FR 10.22 }
with: or, in}
\VS{23}For the
Lord
{\ND{
GOD}} of
hosts shall
make a
consumption, even
determined, in the
midst of all the
land.
\par }{\PP \VS{24}¶ Therefore thus
saith the
Lord
{\ND{
GOD}} of
hosts, O my
people that
dwellest in
Zion, be not
afraid of the
Assyrian: he shall
smite thee with a
rod, and shall lift
up his
staff against thee, after the
manner of
Egypt.\FTNT{†}{{\FR 10.24 }
and shall…: or, but he shall lift up his staff for}
\VS{25}For yet a
very little
while, and the
indignation shall
cease, and mine
anger in their
destruction.
\VS{26}And the
{\ND{
LORD}} of
hosts shall stir
up a
scourge for him according to the
slaughter of
Midian at the
rock of
Oreb: and
{\ADD{as}} his
rod
{\ADD{was}} upon the
sea, so shall he lift it
up after the
manner of
Egypt.
\VS{27}And it shall come to pass in that
day,
{\ADD{that}} his
burden shall be taken
away from off thy
shoulder, and his
yoke from off thy
neck, and the
yoke shall be
destroyed
because of the
anointing.\FTNT{‡}{{\FR 10.27 }
be taken…: Heb. remove}
\VS{28}He is
come to
Aiath, he is
passed to
Migron; at
Michmash he hath laid
up his
carriages:
\VS{29}They are gone
over the
passage: they have taken up their
lodging at
Geba;
Ramah is
afraid;
Gibeah of
Saul is
fled.
\VS{30}Lift
up thy
voice, O
daughter
of
Gallim: cause it to be
heard unto
Laish, O
poor
Anathoth.\FTNT{§}{{\FR 10.30 }
Lift…: Heb. Cry shrill with}
\VS{31}Madmenah is
removed; the
inhabitants of
Gebim gather themselves to
flee.
\VS{32}As yet shall he
remain at
Nob that
day: he shall
shake his
hand
{\ADD{against}} the
mount of the
daughter of
Zion, the
hill of
Jerusalem.
\VS{33}Behold, the
Lord, the
{\ND{
LORD}} of
hosts, shall
lop the
bough with
terror: and the high
ones of
stature
{\ADD{shall be}} hewn
down, and the
haughty shall be
humbled.
\VS{34}And he shall cut
down the
thickets of the
forest with
iron, and
Lebanon shall
fall by a mighty
one.\FTNT{**}{{\FR 10.34 }
by…: or, mightily}

\par }\Chap{11}{\PP \VerseOne{1}And there shall come
forth a
rod out of the
stem of
Jesse, and a
Branch shall
grow out of his
roots:
\VS{2}And the
spirit of the
{\ND{
LORD}} shall
rest upon him, the
spirit of
wisdom and
understanding, the
spirit of
counsel and
might, the
spirit of
knowledge and of the
fear of the
{\ND{
LORD}};
\VS{3}And shall make him of quick
understanding in the
fear of the
{\ND{
LORD}}: and he shall not
judge after the
sight of his
eyes, neither
reprove after the
hearing of his
ears:\FTNT{*}{{\FR 11.3 }
quick…: Heb. scent or, smell}
\VS{4}But with
righteousness shall he
judge the
poor, and
reprove with
equity for the
meek of the
earth: and he shall
smite the
earth with the
rod of his
mouth, and with the
breath of his
lips shall he
slay the
wicked.\FTNT{†}{{\FR 11.4 }
reprove: or, argue}
\VS{5}And
righteousness shall be the
girdle of his
loins, and
faithfulness the
girdle of his
reins.
\VS{6}The
wolf also shall
dwell with the
lamb, and the
leopard shall lie
down with the
kid; and the
calf and the young
lion and the
fatling
together; and a
little
child shall
lead them.
\VS{7}And the
cow and the
bear shall
feed; their young
ones shall lie
down
together: and the
lion shall
eat
straw like the
ox.
\VS{8}And the sucking
child shall
play on the
hole of the
asp, and the weaned
child shall
put his
hand on the
cockatrice’
den.\FTNT{‡}{{\FR 11.8 }
cockatrice’: or, adder’s}
\VS{9}They shall not
hurt nor
destroy in all my
holy
mountain: for the
earth shall be
full of the
knowledge of the
{\ND{
LORD}}, as the
waters
cover the
sea.
\par }{\PP \VS{10}¶ And in that
day there shall be a
root of
Jesse, which shall
stand for an
ensign of the
people; to it shall the
Gentiles
seek: and his
rest shall be
glorious.\FTNT{§}{{\FR 11.10 }
glorious: Heb. glory}
\VS{11}And it shall come to pass in that
day,
{\ADD{that}} the
Lord shall
set his
hand
again the second
time to
recover the
remnant of his
people, which shall be
left, from
Assyria, and from
Egypt, and from
Pathros, and from
Cush, and from
Elam, and from
Shinar, and from
Hamath, and from the
islands of the
sea.
\VS{12}And he shall set
up an
ensign for the
nations, and shall
assemble the
outcasts of
Israel, and gather
together the
dispersed of
Judah from the
four
corners of the
earth.\FTNT{**}{{\FR 11.12 }
corners: Heb. wings}
\VS{13}The
envy also of
Ephraim shall
depart, and the
adversaries of
Judah shall be cut
off:
Ephraim shall not
envy
Judah, and
Judah shall not
vex
Ephraim.
\VS{14}But they shall
fly upon the
shoulders of the
Philistines toward the
west; they shall
spoil
them of the
east
together: they shall
lay their
hand upon
Edom and
Moab; and the
children of
Ammon shall
obey them.\FTNT{††}{{\FR 11.14 }
them of: Heb. the children of}\FTNT{‡‡}{{\FR 11.14 }
they shall lay…: Heb. Edom and Moab shall be the laying on of their hand}\FTNT{§§}{{\FR 11.14 }
shall obey…: Heb. their obedience}
\VS{15}And the
{\ND{
LORD}} shall utterly
destroy the
tongue of the
Egyptian
sea; and with his
mighty
wind shall he
shake his
hand over the
river, and shall
smite it in the
seven
streams, and make
{\ADD{men}} go
over
dryshod.\FTNT{***}{{\FR 11.15 }
dryshod: Heb. in shoes}
\VS{16}And there shall be an
highway for the
remnant of his
people,
which shall be
left, from
Assyria; like as it was to
Israel in the
day that he came
up out of the
land of
Egypt.

\par }\Chap{12}{\PP \VerseOne{1}And in that
day thou shalt
say, O
{\ND{
LORD}}, I will
praise thee: though thou wast
angry with me, thine
anger is turned
away, and thou
comfortedst me.
\VS{2}Behold,
God
{\ADD{is}} my
salvation; I will
trust, and not be
afraid: for the
{\ND{
LORD}}
{\ND{
JEHOVAH}}
{\ADD{is}} my
strength and
{\ADD{my}}
song; he also is become my
salvation.
\VS{3}Therefore with
joy shall ye
draw
water out of the
wells of
salvation.
\par }{\PP \VS{4}And in that
day shall ye
say,
Praise the
{\ND{
LORD}},
call upon his
name,
declare his
doings among the
people, make
mention that his
name is
exalted.\FTNT{*}{{\FR 12.4 }
call…: or, proclaim}
\VS{5}Sing unto the
{\ND{
LORD}}; for he hath
done excellent
things: this
{\ADD{is}}
known in all the
earth.
\VS{6}Cry
out and
shout, thou
inhabitant of
Zion: for
great
{\ADD{is}} the Holy
One of
Israel in the
midst of thee.\FTNT{†}{{\FR 12.6 }
inhabitant: Heb. inhabitress}

\par }\Chap{13}{\PP \VerseOne{1}The
burden of
Babylon, which
Isaiah the
son of
Amoz did
see.
\VS{2}Lift ye
up a
banner upon the
high
mountain,
exalt the
voice unto them,
shake the
hand, that they may
go into the
gates of the
nobles.
\VS{3}I have
commanded my sanctified
ones, I have also
called my mighty
ones for mine
anger,
{\ADD{even}} them that
rejoice in my
highness.
\VS{4}The
noise of a
multitude in the
mountains, like
as of a
great
people; a
tumultuous
noise of the
kingdoms of
nations gathered
together: the
{\ND{
LORD}} of
hosts
mustereth the
host of the
battle.\FTNT{*}{{\FR 13.4 }
like…: Heb. the likeness of}
\VS{5}They
come from a
far
country, from the
end of
heaven,
{\ADD{even}} the
{\ND{
LORD}}, and the
weapons of his
indignation, to
destroy the whole
land.
\par }{\PP \VS{6}¶
Howl ye; for the
day of the
{\ND{
LORD}}
{\ADD{is}} at
hand; it shall
come as a
destruction from the
Almighty.
\VS{7}Therefore shall all
hands be
faint, and every
man’s
heart shall
melt:\FTNT{†}{{\FR 13.7 }
be faint: or, fall down}
\VS{8}And they shall be
afraid:
pangs and
sorrows shall take
hold of them; they shall be in
pain as a woman that
travaileth: they shall be
amazed
one at
another; their
faces
{\ADD{shall be as}}
flames.\FTNT{‡}{{\FR 13.8 }
be amazed: Heb. wonder}\FTNT{§}{{\FR 13.8 }
one…: Heb. every man at his neighbour}\FTNT{**}{{\FR 13.8 }
flames: Heb. faces of the flames}
\VS{9}Behold, the
day of the
{\ND{
LORD}}
cometh,
cruel both with
wrath and
fierce
anger, to
lay the
land
desolate: and he shall
destroy the
sinners thereof out of it.
\VS{10}For the
stars of
heaven and the
constellations thereof shall not
give their
light: the
sun shall be
darkened in his going
forth, and the
moon shall not cause her
light to
shine.
\VS{11}And I will
punish the
world for
{\ADD{their}}
evil, and the
wicked for their
iniquity; and I will cause the
arrogancy of the
proud to
cease, and will lay
low the
haughtiness of the
terrible.
\VS{12}I will make a
man more
precious than fine
gold; even a
man than the golden
wedge of
Ophir.
\VS{13}Therefore I will
shake the
heavens, and the
earth shall
remove out of her
place, in the
wrath of the
{\ND{
LORD}} of
hosts, and in the
day of his
fierce
anger.
\VS{14}And it shall be as the
chased
roe, and as a
sheep that no man taketh
up: they shall every
man
turn to his own
people, and
flee every
one into his own
land.
\VS{15}Every one that is
found shall be thrust
through; and every one that is
joined
{\ADD{unto them}} shall
fall by the
sword.
\VS{16}Their
children also shall be dashed to
pieces before their
eyes; their
houses shall be
spoiled, and their
wives
ravished.
\VS{17}Behold, I will stir
up the
Medes against them, which shall not
regard
silver; and
{\ADD{as for}}
gold, they shall not
delight in it.
\VS{18}{\ADD{Their}}
bows also shall
dash the young
men to
pieces; and they shall have no
pity on the
fruit of the
womb; their
eye shall not
spare
children.
\par }{\PP \VS{19}¶ And
Babylon, the
glory of
kingdoms, the
beauty of the
Chaldees’
excellency, shall be as when
God
overthrew
Sodom and
Gomorrah.\FTNT{††}{{\FR 13.19 }
as…: Heb. as the overthrowing}
\VS{20}It shall
never be
inhabited, neither shall it be
dwelt in from
generation to
generation: neither shall the
Arabian pitch
tent there; neither shall the
shepherds make their
fold there.
\VS{21}But wild beasts of the
desert shall
lie there; and their
houses shall be
full of doleful
creatures; and
owls shall
dwell there, and
satyrs shall
dance there.\FTNT{‡‡}{{\FR 13.21 }
wild…: Heb. Ziim}\FTNT{§§}{{\FR 13.21 }
doleful…: Heb. Ochim}\FTNT{***}{{\FR 13.21 }
owls: or, ostriches: Heb. daughters of the owl}
\VS{22}And the wild beasts of the
islands shall
cry in their desolate
houses, and
dragons in
{\ADD{their}}
pleasant
palaces: and her
time
{\ADD{is}}
near to
come, and her
days shall not be
prolonged.\FTNT{†††}{{\FR 13.22 }
the wild…: Heb. Iim}\FTNT{‡‡‡}{{\FR 13.22 }
desolate…: or, palaces}

\par }\Chap{14}{\PP \VerseOne{1}For the
{\ND{
LORD}} will have
mercy on
Jacob, and will yet
choose
Israel, and
set them in their own
land: and the
strangers shall be
joined with them, and they shall
cleave to the
house of
Jacob.
\VS{2}And the
people shall
take them, and
bring them to their
place: and the
house of
Israel shall
possess them in the
land of the
{\ND{
LORD}} for
servants and
handmaids: and they shall take them
captives, whose
captives they were; and they shall
rule over their
oppressors.\FTNT{*}{{\FR 14.2 }
whose…: Heb. that had taken them captives}
\VS{3}And it shall come to pass in the
day that the
{\ND{
LORD}} shall give thee
rest from thy
sorrow, and from thy
fear, and from the
hard
bondage wherein thou wast made to
serve,
\par }{\PP \VS{4}¶ That thou shalt take
up this
proverb against the
king of
Babylon, and
say, How hath the
oppressor
ceased! the golden
city
ceased!\FTNT{†}{{\FR 14.4 }
proverb: or, taunting speech}\FTNT{‡}{{\FR 14.4 }
golden…: or, exactress of gold}
\VS{5}The
{\ND{
LORD}} hath
broken the
staff of the
wicked,
{\ADD{and}} the
sceptre of the
rulers.
\VS{6}He who
smote the
people in
wrath with a
continual
stroke, he that
ruled the
nations in
anger, is
persecuted,
{\ADD{and}}
none
hindereth.
\FTNT{§}{{\FR 14.6 }
a continual…: Heb. a stroke without removing}
\VS{7}The whole
earth is at
rest,
{\ADD{and}} is
quiet: they break
forth into
singing.
\VS{8}Yea, the fir
trees
rejoice at thee,
{\ADD{and}} the
cedars of
Lebanon,
{\ADD{saying}}, Since thou art laid
down, no
feller is come
up against us.
\VS{9}Hell from beneath is
moved for thee to
meet
{\ADD{thee}} at thy
coming: it stirreth
up the
dead for thee,
{\ADD{even}} all the chief
ones of the
earth; it hath raised
up from their
thrones all the
kings of the
nations.\FTNT{**}{{\FR 14.9 }
Hell: or, The grave}\FTNT{††}{{\FR 14.9 }
chief…: Heb. leaders, or, great goats}
\VS{10}All they shall
speak and
say unto thee, Art thou also become
weak as we? art thou become
like unto us?
\VS{11}Thy
pomp is brought
down to the
grave,
{\ADD{and}} the
noise of thy
viols: the
worm is
spread under thee, and the
worms
cover thee.
\VS{12}How art thou
fallen from
heaven, O
Lucifer,
son of the
morning!
{\ADD{how}} art thou cut
down to the
ground, which didst
weaken the
nations!\FTNT{‡‡}{{\FR 14.12 }
O Lucifer: or, O day star}
\VS{13}For thou hast
said in thine
heart, I will
ascend into
heaven, I will
exalt my
throne
above the
stars of
God: I will
sit also upon the
mount of the
congregation, in the
sides of the
north:
\VS{14}I will
ascend above the
heights of the
clouds; I will be
like the most
High.
\VS{15}Yet thou shalt be brought
down to
hell, to the
sides of the
pit.
\VS{16}They that
see thee shall narrowly
look upon thee,
{\ADD{and}}
consider thee,
{\ADD{saying, Is}} this the
man that made the
earth to
tremble, that did
shake
kingdoms;
\VS{17}{\ADD{That}}
made the
world as a
wilderness, and
destroyed the
cities thereof;
{\ADD{that}}
opened not the
house of his
prisoners?\FTNT{§§}{{\FR 14.17 }
opened…: or, did not let his prisoners loose homeward?}
\VS{18}All the
kings of the
nations,
{\ADD{even}} all of them,
lie in
glory, every
one in his own
house.
\VS{19}But thou art cast
out of thy
grave like an
abominable
branch,
{\ADD{and as}} the
raiment of those that are
slain, thrust
through with a
sword, that go
down to the
stones of the
pit; as a
carcase trodden under
feet.
\VS{20}Thou shalt not be
joined with them in
burial, because thou hast
destroyed thy
land,
{\ADD{and}}
slain thy
people: the
seed of
evildoers shall
never be
renowned.
\VS{21}Prepare
slaughter for his
children for the
iniquity of their
fathers; that they do
not
rise, nor
possess the
land, nor
fill the
face of the
world with
cities.
\VS{22}For I will rise
up against them,
saith the
{\ND{
LORD}} of
hosts, and cut
off from
Babylon the
name, and
remnant, and
son, and
nephew,
saith the
{\ND{
LORD}}.
\VS{23}I will also
make it a
possession for the
bittern, and
pools of
water: and I will
sweep it with the
besom of
destruction,
saith the
{\ND{
LORD}} of
hosts.
\par }{\PP \VS{24}¶ The
{\ND{
LORD}} of
hosts hath
sworn,
saying, Surely as I have
thought, so shall it come to pass; and as I have
purposed,
{\ADD{so}} shall it
stand:
\VS{25}That I will
break the
Assyrian in my
land, and upon my
mountains tread him under
foot: then shall his
yoke
depart from off them, and his
burden
depart from off their
shoulders.
\VS{26}This
{\ADD{is}} the
purpose that is
purposed upon the whole
earth: and this
{\ADD{is}} the
hand that is stretched
out upon all the
nations.
\VS{27}For the
{\ND{
LORD}} of
hosts hath
purposed, and who shall
disannul
{\ADD{it}}? and his
hand
{\ADD{is}} stretched
out, and who shall turn it
back?
\VS{28}In the
year that
king
Ahaz
died was this
burden.
\par }{\PP \VS{29}¶
Rejoice not thou, whole
Palestina, because the
rod of him that
smote thee is
broken: for out of the
serpent’s
root shall come
forth a
cockatrice, and his
fruit
{\ADD{shall be}} a
fiery
flying
serpent.\FTNT{***}{{\FR 14.29 }
cockatrice: or, adder}
\VS{30}And the
firstborn of the
poor shall
feed, and the
needy shall lie
down in
safety: and I will
kill thy
root with
famine, and he shall
slay thy
remnant.
\VS{31}Howl, O
gate;
cry, O
city; thou, whole
Palestina,
{\ADD{art}}
dissolved: for there shall
come from the
north a
smoke, and none
{\ADD{shall be}}
alone in his appointed
times.\FTNT{†††}{{\FR 14.31 }
none…: or, he shall not be alone}\FTNT{‡‡‡}{{\FR 14.31 }
appointed…: or, assemblies}
\VS{32}What shall
{\ADD{one}} then
answer the
messengers of the
nation? That the
{\ND{
LORD}} hath
founded
Zion, and the
poor of his
people shall
trust in it.\FTNT{§§§}{{\FR 14.32 }
trust…: or, betake themselves unto it}

\par }\Chap{15}{\PP \VerseOne{1}The
burden of
Moab. Because in the
night
Ar of
Moab is laid
waste,
{\ADD{and}} brought to
silence; because in the
night
Kir of
Moab is laid
waste,
{\ADD{and}} brought to
silence;\FTNT{*}{{\FR 15.1 }
brought…: or, cut off}
\VS{2}He is gone
up to
Bajith, and to
Dibon, the high
places, to
weep:
Moab shall
howl over
Nebo, and over
Medeba: on all their
heads
{\ADD{shall be}}
baldness,
{\ADD{and}} every
beard cut
off.
\VS{3}In their
streets they shall
gird themselves with
sackcloth: on the
tops of their houses, and in their
streets, every one shall
howl,
weeping
abundantly.\FTNT{†}{{\FR 15.3 }
weeping…: Heb. descending into weeping, or, coming down with weeping}
\VS{4}And
Heshbon shall
cry, and
Elealeh: their
voice shall be
heard
{\ADD{even}} unto
Jahaz: therefore the armed
soldiers of
Moab shall cry
out; his
life shall be
grievous unto him.
\VS{5}My
heart shall cry
out for
Moab; his
fugitives
{\ADD{shall flee}} unto
Zoar, an
heifer of three years
old: for by the mounting
up of
Luhith with
weeping shall they go it
up; for in the
way of
Horonaim they shall raise
up a
cry of
destruction.\FTNT{‡}{{\FR 15.5 }
his…: or, to the borders thereof, even to Zoar, as an heifer}\FTNT{§}{{\FR 15.5 }
destruction: Heb. breaking}
\par }{\PP \VS{6}For the
waters of
Nimrim shall be
desolate: for the
hay is withered
away, the
grass
faileth, there is no green
thing.\FTNT{**}{{\FR 15.6 }
desolate: Heb. desolations}
\VS{7}Therefore the
abundance they have
gotten, and that which they have laid
up, shall they carry
away to the
brook of the
willows.\FTNT{††}{{\FR 15.7 }
brook…: or, valley of the Arabians}
\VS{8}For the
cry is gone round
about the
borders of
Moab; the
howling thereof unto
Eglaim, and the
howling thereof unto
Beer-elim.
\VS{9}For the
waters of
Dimon shall be
full of
blood: for I will
bring
more upon
Dimon,
lions upon him that
escapeth of
Moab, and upon the
remnant of the
land.\FTNT{‡‡}{{\FR 15.9 }
more: Heb. additions}

\par }\Chap{16}{\PP \VerseOne{1}Send ye the
lamb to the
ruler of the
land from
Sela to the
wilderness, unto the
mount of the
daughter of
Zion.\FTNT{*}{{\FR 16.1 }
Sela: or, Petra: Heb. A rock}
\VS{2}For it shall be,
{\ADD{that}}, as a
wandering
bird cast
out of the
nest,
{\ADD{so}} the
daughters of
Moab shall be at the
fords of
Arnon.\FTNT{†}{{\FR 16.2 }
cast…: or, a nest forsaken}
\VS{3}Take
counsel,
execute
judgment;
make thy
shadow as the
night in the
midst of the
noonday;
hide the
outcasts;
bewray not him that
wandereth.\FTNT{‡}{{\FR 16.3 }
Take: Heb. Bring}
\VS{4}Let mine
outcasts
dwell with thee,
Moab;
be thou a
covert to them from the
face of the
spoiler: for the
extortioner is at an
end, the
spoiler
ceaseth, the
oppressors are
consumed out of the
land.\FTNT{§}{{\FR 16.4 }
extortioner: Heb. wringer}\FTNT{**}{{\FR 16.4 }
the oppressors: Heb. the treaders down}
\VS{5}And in
mercy shall the
throne be
established: and he shall
sit upon it in
truth in the
tabernacle of
David,
judging, and
seeking
judgment, and
hasting
righteousness.\FTNT{††}{{\FR 16.5 }
established: or, prepared}
\par }{\PP \VS{6}¶ We have
heard of the
pride of
Moab;
{\ADD{he is}}
very
proud:
{\ADD{even}} of his
haughtiness, and his
pride, and his
wrath:
{\ADD{but}} his
lies
{\ADD{shall}} not
{\ADD{be}} so.
\VS{7}Therefore shall
Moab
howl for
Moab, every one shall
howl: for the
foundations of
Kir-hareseth shall ye
mourn; surely
{\ADD{they are}}
stricken.\FTNT{‡‡}{{\FR 16.7 }
mourn: or, mutter}
\VS{8}For the
fields of
Heshbon
languish,
{\ADD{and}} the
vine of
Sibmah: the
lords of the
heathen have broken
down the principal
plants thereof, they are
come
{\ADD{even}} unto
Jazer, they
wandered
{\ADD{through}} the
wilderness: her
branches are stretched
out, they are gone
over the
sea.\FTNT{§§}{{\FR 16.8 }
stretched…: or, plucked up}
\par }{\PP \VS{9}¶ Therefore I will
bewail with the
weeping of
Jazer the
vine of
Sibmah: I will
water thee with my
tears, O
Heshbon, and
Elealeh: for the
shouting for thy summer
fruits and for thy
harvest is
fallen.\FTNT{***}{{\FR 16.9 }
the shouting…: or, the alarm is fallen upon, etc}
\VS{10}And
gladness is taken
away, and
joy out of the plentiful
field; and in the
vineyards there shall be no
singing, neither shall there be
shouting: the
treaders shall tread
out no
wine in
{\ADD{their}}
presses; I have made
{\ADD{their vintage}}
shouting to
cease.
\VS{11}Wherefore my
bowels shall
sound like an
harp for
Moab, and mine inward
parts for
Kir-haresh.
\par }{\PP \VS{12}¶ And it shall come to pass, when it is
seen that
Moab is
weary on the high
place, that he shall
come to his
sanctuary to
pray; but he shall not
prevail.
\VS{13}This
{\ADD{is}} the
word that the
{\ND{
LORD}} hath
spoken concerning
Moab since that
time.
\VS{14}But now the
{\ND{
LORD}} hath
spoken,
saying, Within
three
years, as the
years of an
hireling, and the
glory of
Moab shall be
contemned, with all that
great
multitude; and the
remnant
{\ADD{shall be}}
very
small
{\ADD{and}}
feeble.\FTNT{†††}{{\FR 16.14 }
feeble: or, not many}

\par }\Chap{17}{\PP \VerseOne{1}The
burden of
Damascus. Behold,
Damascus is taken
away from
{\ADD{being}} a
city, and it shall be a
ruinous
heap.
\VS{2}The
cities of
Aroer
{\ADD{are}}
forsaken: they shall be for
flocks, which shall lie
down, and none shall make
{\ADD{them}}
afraid.
\VS{3}The
fortress also shall
cease from
Ephraim, and the
kingdom from
Damascus, and the
remnant of
Syria: they shall be as the
glory of the
children of
Israel,
saith the
{\ND{
LORD}} of
hosts.
\VS{4}And in that
day it shall come to pass,
{\ADD{that}} the
glory of
Jacob shall be made
thin, and the
fatness of his
flesh shall wax
lean.
\VS{5}And it shall be as when the
harvestman
gathereth the
corn, and
reapeth the
ears with his
arm; and it shall be as he that
gathereth
ears in the
valley of
Rephaim.
\par }{\PP \VS{6}¶ Yet gleaning
grapes shall be
left in it, as the
shaking of an olive
tree,
two
{\ADD{or}}
three
berries in the
top of the uppermost
bough,
four
{\ADD{or}}
five in the outmost
fruitful
branches thereof,
saith the
{\ND{
LORD}}
God of
Israel.
\VS{7}At that
day shall a
man
look to his
Maker, and his
eyes shall have
respect to the
Holy One of
Israel.
\VS{8}And he shall not
look to the
altars, the
work of his
hands, neither shall
respect
{\ADD{that}} which his
fingers have
made, either the
groves, or the
images.\FTNT{*}{{\FR 17.8 }
images: or, sun images}
\par }{\PP \VS{9}¶ In that
day shall his
strong
cities be as a
forsaken
bough, and an uppermost
branch, which they
left
because of the
children of
Israel: and there shall be
desolation.
\VS{10}Because thou hast
forgotten the
God of thy
salvation, and hast not been
mindful of the
rock of thy
strength, therefore shalt thou
plant
pleasant
plants, and shalt
set it with
strange
slips:
\VS{11}In the
day shalt thou make thy
plant to
grow, and in the
morning shalt thou make thy
seed to
flourish:
{\ADD{but}} the
harvest
{\ADD{shall be}} a
heap in the
day of
grief and of
desperate
sorrow.\FTNT{†}{{\FR 17.11 }
a heap…: or, removed in the day of inheritance, and there shall be deadly sorrow}
\par }{\PP \VS{12}¶
Woe to the
multitude of
many
people,
{\ADD{which}} make a
noise like the
noise of the
seas; and to the
rushing of
nations,
{\ADD{that}} make a rushing like the
rushing of
mighty
waters!\FTNT{‡}{{\FR 17.12 }
multitude: or, noise}\FTNT{§}{{\FR 17.12 }
mighty: or, many}
\VS{13}The
nations shall
rush like the
rushing of
many
waters: but
{\ADD{God}} shall
rebuke them, and they shall
flee far
off, and shall be
chased as the
chaff of the
mountains
before the
wind, and like a rolling
thing
before the
whirlwind.\FTNT{**}{{\FR 17.13 }
a rolling…: or, thistledown}
\VS{14}And behold at
eveningtide
trouble;
{\ADD{and}} before the
morning he
{\ADD{is}} not. This
{\ADD{is}} the
portion of them that
spoil us, and the
lot of them that
rob us.

\par }\Chap{18}{\PP \VerseOne{1}Woe to the
land
shadowing with
wings, which
{\ADD{is}}
beyond the
rivers of
Ethiopia:
\VS{2}That
sendeth
ambassadors by the
sea, even in
vessels of
bulrushes
upon the
waters,
{\ADD{saying}},
Go, ye
swift
messengers, to a
nation
scattered and
peeled, to a
people
terrible from their beginning
hitherto; a
nation meted
out and trodden
down, whose
land the
rivers have
spoiled!\FTNT{*}{{\FR 18.2 }
scattered…: or, outspread and polished}\FTNT{†}{{\FR 18.2 }
meted…: or, that meteth out and treadeth down: Heb. of line, line, and treading under foot}\FTNT{‡}{{\FR 18.2 }
have…: or, despise}
\VS{3}All ye
inhabitants of the
world, and
dwellers on the
earth,
see ye, when he lifteth
up an
ensign on the
mountains; and when he
bloweth a
trumpet,
hear ye.
\VS{4}For
so the
{\ND{
LORD}}
said unto me, I will take my
rest, and I will
consider in my dwelling
place like a
clear
heat upon
herbs,
{\ADD{and}} like a
cloud of
dew in the
heat of
harvest.\FTNT{§}{{\FR 18.4 }
consider…: or, regard my set dwelling}\FTNT{**}{{\FR 18.4 }
upon…: or, after rain}
\VS{5}For
afore the
harvest, when the
bud is
perfect, and the sour
grape is
ripening in the
flower, he shall both cut
off the
sprigs with pruning
hooks, and take
away
{\ADD{and}} cut
down the
branches.
\VS{6}They shall be
left
together unto the
fowls of the
mountains, and to the
beasts of the
earth: and the
fowls shall
summer upon them, and all the
beasts of the
earth shall
winter upon them.
\par }{\PP \VS{7}¶ In that
time shall the
present be
brought unto the
{\ND{
LORD}} of
hosts of a
people
scattered and
peeled, and from a
people
terrible from their beginning
hitherto; a
nation meted
out and trodden under
foot, whose
land the
rivers have
spoiled, to the
place of the
name of the
{\ND{
LORD}} of
hosts, the
mount
Zion.\FTNT{††}{{\FR 18.7 }
scattered…: or, outspread and polished}

\par }\Chap{19}{\PP \VerseOne{1}The
burden of
Egypt. Behold, the
{\ND{
LORD}}
rideth upon a
swift
cloud, and shall
come into
Egypt: and the
idols of
Egypt shall be
moved at his
presence, and the
heart of
Egypt shall
melt in the
midst of it.
\VS{2}And I will
set the
Egyptians against the
Egyptians: and they shall
fight every
one against his
brother, and every
one against his
neighbour;
city against
city,
{\ADD{and}}
kingdom against
kingdom.\FTNT{*}{{\FR 19.2 }
set: Heb. mingle}
\VS{3}And the
spirit of
Egypt shall
fail in the
midst thereof; and I will
destroy the
counsel thereof: and they shall
seek to the
idols, and to the
charmers, and to them that have familiar
spirits, and to the
wizards.\FTNT{†}{{\FR 19.3 }
fail: Heb. be emptied}\FTNT{‡}{{\FR 19.3 }
destroy: Heb. swallow up}
\VS{4}And the
Egyptians will I give
over into the
hand of a
cruel
lord; and a
fierce
king shall
rule over them,
saith the
Lord, the
{\ND{
LORD}} of
hosts.\FTNT{§}{{\FR 19.4 }
give…: or, shut up}
\VS{5}And the
waters shall
fail from the
sea, and the
river shall be
wasted and dried
up.
\VS{6}And they shall
turn the
rivers far
away;
{\ADD{and}} the
brooks of
defence shall be
emptied and dried
up: the
reeds and
flags shall
wither.
\VS{7}The paper
reeds by the
brooks, by the
mouth of the
brooks, and every thing
sown by the
brooks, shall
wither, be driven
away, and be no
{\ADD{more}}.\FTNT{**}{{\FR 19.7 }
and be…: Heb. and shall not be}
\VS{8}The
fishers also shall
mourn, and all they that
cast
angle into the
brooks shall
lament, and they that
spread
nets
upon the
waters shall
languish.
\VS{9}Moreover they that
work in
fine
flax, and they that
weave
networks, shall be
confounded.\FTNT{††}{{\FR 19.9 }
networks: or, white works}
\VS{10}And they shall be
broken in the
purposes thereof, all that
make
sluices
{\ADD{and}}
ponds for
fish.\FTNT{‡‡}{{\FR 19.10 }
purposes: Heb. foundations}\FTNT{§§}{{\FR 19.10 }
for fish: Heb. of living things}
\par }{\PP \VS{11}¶ Surely the
princes of
Zoan
{\ADD{are}}
fools, the
counsel of the
wise
counsellors of
Pharaoh is become
brutish: how
say ye unto
Pharaoh, I
{\ADD{am}} the
son of the
wise, the
son of
ancient
kings?
\VS{12}Where
{\ADD{are}} they?
where
{\ADD{are}} thy
wise
{\ADD{men}}? and let them
tell thee now, and let them
know what the
{\ND{
LORD}} of
hosts hath
purposed upon
Egypt.
\VS{13}The
princes of
Zoan are become
fools, the
princes of
Noph are
deceived; they have also
seduced
Egypt,
{\ADD{even they that are}} the
stay of the
tribes thereof.\FTNT{***}{{\FR 19.13 }
they that…: or, governors: Heb. corners}
\VS{14}The
{\ND{
LORD}} hath
mingled a
perverse
spirit in the
midst thereof: and they have caused
Egypt to
err in every
work thereof, as a
drunken
{\ADD{man}}
staggereth in his
vomit.\FTNT{†††}{{\FR 19.14 }
a perverse…: Heb. a spirit of perversities}
\VS{15}Neither shall there be
{\ADD{any}}
work for
Egypt, which the
head or
tail,
branch or
rush, may
do.
\VS{16}In that
day shall
Egypt be like unto
women: and it shall be
afraid and
fear
because of the
shaking of the
hand of the
{\ND{
LORD}} of
hosts, which he
shaketh over it.
\VS{17}And the
land of
Judah shall be a
terror unto
Egypt, every one that maketh
mention thereof shall be
afraid in himself,
because of the
counsel of the
{\ND{
LORD}} of
hosts, which he hath
determined against it.
\par }{\PP \VS{18}¶ In that
day shall
five
cities in the
land of
Egypt
speak the
language of
Canaan, and
swear to the
{\ND{
LORD}} of
hosts;
one shall be
called, The
city of
destruction.\FTNT{‡‡‡}{{\FR 19.18 }
the language: Heb. the lip}\FTNT{§§§}{{\FR 19.18 }
of destruction: or, of Heres, or, of the sun}
\VS{19}In that
day shall there be an
altar to the
{\ND{
LORD}} in the
midst of the
land of
Egypt, and a
pillar
at the
border thereof to the
{\ND{
LORD}}.
\VS{20}And it shall be for a
sign and for a
witness unto the
{\ND{
LORD}} of
hosts in the
land of
Egypt: for they shall
cry unto the
{\ND{
LORD}}
because of the
oppressors, and he shall
send them a
saviour, and a great
one, and he shall
deliver them.
\VS{21}And the
{\ND{
LORD}} shall be
known to
Egypt, and the
Egyptians shall
know the
{\ND{
LORD}} in that
day, and shall
do
sacrifice and
oblation; yea, they shall
vow a
vow unto the
{\ND{
LORD}}, and
perform
{\ADD{it}}.
\VS{22}And the
{\ND{
LORD}} shall
smite
Egypt: he shall
smite and
heal
{\ADD{it:}} and they shall
return
{\ADD{even}} to the
{\ND{
LORD}}, and he shall be
intreated of them, and shall
heal them.
\par }{\PP \VS{23}¶ In that
day shall there be a
highway out of
Egypt to
Assyria, and the
Assyrian shall
come into
Egypt, and the
Egyptian into
Assyria, and the
Egyptians shall
serve with the
Assyrians.
\VS{24}In that
day shall
Israel be the
third with
Egypt and with
Assyria,
{\ADD{even}} a
blessing in the
midst of the
land:
\VS{25}Whom the
{\ND{
LORD}} of
hosts shall
bless,
saying,
Blessed
{\ADD{be}}
Egypt my
people, and
Assyria the
work of my
hands, and
Israel mine
inheritance.

\par }\Chap{20}{\PP \VerseOne{1}In the
year that
Tartan
came unto
Ashdod, (when
Sargon the
king of
Assyria
sent him,) and
fought against
Ashdod, and
took it;
\VS{2}At the same
time
spake the
{\ND{
LORD}}
by
Isaiah the
son of
Amoz,
saying,
Go and
loose the
sackcloth from off thy
loins, and put
off thy
shoe from thy
foot. And he did
so,
walking
naked and
barefoot.\FTNT{*}{{\FR 20.2 }
by: Heb. by the hand of}
\VS{3}And the
{\ND{
LORD}}
said, Like as my
servant
Isaiah hath
walked
naked and
barefoot
three
years
{\ADD{for}} a
sign and
wonder upon
Egypt and upon
Ethiopia;
\VS{4}So shall the
king of
Assyria lead
away the
Egyptians
prisoners, and the
Ethiopians
captives,
young and
old,
naked and
barefoot, even with
{\ADD{their}}
buttocks
uncovered, to the
shame of
Egypt.\FTNT{†}{{\FR 20.4 }
the Egyptians…: Heb. the captivity of Egypt}\FTNT{‡}{{\FR 20.4 }
shame: Heb. nakedness}
\VS{5}And they shall be
afraid and
ashamed of
Ethiopia their
expectation, and of
Egypt their
glory.
\VS{6}And the
inhabitant of this
isle shall
say in that
day, Behold,
such
{\ADD{is}} our
expectation, whither we
flee for
help to be
delivered
from the
king of
Assyria: and how shall we
escape?\FTNT{§}{{\FR 20.6 }
isle: or, country}

\par }\Chap{21}{\PP \VerseOne{1}The
burden of the
desert of the
sea. As
whirlwinds in the
south
pass through;
{\ADD{so}} it
cometh from the
desert, from a
terrible
land.
\VS{2}A
grievous
vision is
declared unto me; the treacherous
dealer dealeth
treacherously, and the
spoiler
spoileth. Go
up, O
Elam:
besiege, O
Media; all the
sighing thereof have I made to
cease.\FTNT{*}{{\FR 21.2 }
grievous: Heb. hard}
\VS{3}Therefore are my
loins
filled with
pain:
pangs have taken
hold upon me, as the
pangs of a woman that
travaileth: I was bowed
down at the
hearing
{\ADD{of it;}} I was
dismayed at the
seeing
{\ADD{of it}}.
\VS{4}My
heart
panted,
fearfulness
affrighted me: the
night of my
pleasure hath he
turned into
fear unto me.\FTNT{†}{{\FR 21.4 }
heart…: or, mind wandered}\FTNT{‡}{{\FR 21.4 }
turned: Heb. put}
\VS{5}Prepare the
table,
watch in the
watchtower,
eat,
drink:
arise, ye
princes,
{\ADD{and}}
anoint the
shield.
\VS{6}For thus hath the
Lord
said unto me,
Go,
set a
watchman, let him
declare what he
seeth.
\VS{7}And he
saw a
chariot
{\ADD{with}} a
couple of
horsemen, a
chariot of
asses,
{\ADD{and}} a
chariot of
camels; and he
hearkened
diligently with
much
heed:
\VS{8}And he
cried, A
lion: My
lord, I
stand
continually upon the
watchtower in the
daytime, and I am
set in my
ward whole
nights:\FTNT{§}{{\FR 21.8 }
he…: or, cried as a lion}\FTNT{**}{{\FR 21.8 }
whole…: or, every night}
\VS{9}And, behold, here
cometh a
chariot of
men,
{\ADD{with}} a
couple of
horsemen. And he
answered and
said,
Babylon is
fallen, is
fallen; and all the graven
images of her
gods he hath
broken unto the
ground.
\VS{10}O my
threshing, and the
corn of my
floor: that which I have
heard of the
{\ND{
LORD}} of
hosts, the
God of
Israel, have I
declared unto you.\FTNT{††}{{\FR 21.10 }
corn: Heb. son}
\par }{\PP \VS{11}¶ The
burden of
Dumah. He
calleth to me out of
Seir,
Watchman, what of the
night?
Watchman, what of the
night?
\VS{12}The
watchman
said, The
morning
cometh, and also the
night: if ye will
enquire,
enquire ye:
return,
come.
\par }{\PP \VS{13}¶ The
burden upon
Arabia. In the
forest in
Arabia shall ye
lodge, O ye travelling
companies of
Dedanim.
\VS{14}The
inhabitants of the
land of
Tema
brought
water to
him that was
thirsty, they
prevented with their
bread him that
fled.\FTNT{‡‡}{{\FR 21.14 }
brought: or, bring ye}
\VS{15}For they
fled
from the
swords,
from the
drawn
sword, and
from the
bent
bow, and
from the
grievousness of
war.\FTNT{§§}{{\FR 21.15 }
from the swords: or, for fear, etc: Heb. from the face, etc}
\VS{16}For thus hath the
Lord
said unto me, Within a
year, according to the
years of an
hireling, and all the
glory of
Kedar shall
fail:
\VS{17}And the
residue of the
number of
archers, the mighty
men of the
children of
Kedar, shall be
diminished: for the
{\ND{
LORD}}
God of
Israel hath
spoken
{\ADD{it}}.\FTNT{***}{{\FR 21.17 }
archers: Heb. bows}

\par }\Chap{22}{\PP \VerseOne{1}The
burden of the
valley of
vision. What aileth thee
now, that thou art wholly gone
up to the
housetops?
\VS{2}Thou that art
full of
stirs, a
tumultuous
city, a
joyous
city: thy
slain
{\ADD{men are}} not
slain with the
sword, nor
dead in
battle.
\VS{3}All thy
rulers are
fled
together, they are
bound by the
archers: all that are
found in thee are
bound
together,
{\ADD{which}} have
fled from
far.\FTNT{*}{{\FR 22.3 }
by…: Heb. of the bow}
\VS{4}Therefore
said I, Look
away from me; I will
weep
bitterly,
labour not to
comfort me, because of the
spoiling of the
daughter of my
people.\FTNT{†}{{\FR 22.4 }
weep…: Heb. be bitter in weeping}
\VS{5}For
{\ADD{it is}} a
day of
trouble, and of treading
down, and of
perplexity by the
Lord
{\ND{
GOD}} of
hosts in the
valley of
vision,
breaking down the
walls, and of
crying to the
mountains.
\VS{6}And
Elam
bare the
quiver with
chariots of
men
{\ADD{and}}
horsemen, and
Kir
uncovered the
shield.\FTNT{‡}{{\FR 22.6 }
uncovered: Heb. made naked}
\VS{7}And it shall come to pass,
{\ADD{that}} thy
choicest
valleys shall be
full of
chariots, and the
horsemen shall
set themselves in
array at the
gate.\FTNT{§}{{\FR 22.7 }
thy…: Heb. the choice of thy valleys}\FTNT{**}{{\FR 22.7 }
at: or, toward}
\par }{\PP \VS{8}¶ And he
discovered the
covering of
Judah, and thou didst
look in that
day to the
armour of the
house of the
forest.
\VS{9}Ye have
seen also the
breaches of the
city of
David, that they are
many: and ye gathered
together the
waters of the
lower
pool.
\VS{10}And ye have
numbered the
houses of
Jerusalem, and the
houses have ye broken
down to
fortify the
wall.
\VS{11}Ye
made also a
ditch between the two
walls for the
water of the
old
pool: but ye have not
looked unto the
maker thereof, neither had
respect unto him that
fashioned it long
ago.
\VS{12}And in that
day did the
Lord
{\ND{
GOD}} of
hosts
call to
weeping, and to
mourning, and to
baldness, and to
girding with
sackcloth:
\VS{13}And behold
joy and
gladness,
slaying
oxen, and
killing
sheep,
eating
flesh, and
drinking
wine: let us
eat and
drink; for to
morrow we shall
die.
\VS{14}And it was
revealed in mine
ears by the
{\ND{
LORD}} of
hosts, Surely this
iniquity shall not be
purged from you till ye
die,
saith the
Lord
{\ND{
GOD}} of
hosts.
\par }{\PP \VS{15}¶ Thus
saith the
Lord
{\ND{
GOD}} of
hosts,
Go,
get thee unto this
treasurer,
{\ADD{even}} unto
Shebna, which
{\ADD{is}} over the
house,
{\ADD{and say}},
\VS{16}What hast thou here? and whom hast thou here, that thou hast hewed thee
out a
sepulchre
here,
{\ADD{as}} he that heweth him
out a
sepulchre on
high,
{\ADD{and}} that
graveth an
habitation for himself in a
rock?\FTNT{††}{{\FR 22.16 }
as…: or, O he}
\VS{17}Behold, the
{\ND{
LORD}} will carry thee
away with a
mighty
captivity, and will
surely
cover thee.\FTNT{‡‡}{{\FR 22.17 }
will carry…: or, who covered thee with an excellent covering, and clothed thee gorgeously, (next verse) shall surely, etc}\FTNT{§§}{{\FR 22.17 }
a mighty…: Heb. the captivity of a man}
\VS{18}He will
surely violently
turn and
toss thee
{\ADD{like}} a
ball into a
large
country: there shalt thou
die, and there the
chariots of thy
glory
{\ADD{shall be}} the
shame of thy
lord’s
house.\FTNT{***}{{\FR 22.18 }
large: Heb. large of spaces}
\VS{19}And I will
drive thee from thy
station, and from thy
state shall he pull thee
down.
\par }{\PP \VS{20}¶ And it shall come to pass in that
day, that I will
call my
servant
Eliakim the
son of
Hilkiah:
\VS{21}And I will
clothe him with thy
robe, and
strengthen him with thy
girdle, and I will
commit thy
government into his
hand: and he shall be a
father to the
inhabitants of
Jerusalem, and to the
house of
Judah.
\VS{22}And the
key of the
house of
David will I
lay upon his
shoulder; so he shall
open, and none shall
shut; and he shall
shut, and none shall
open.
\VS{23}And I will
fasten him
{\ADD{as}} a
nail in a
sure
place; and he shall be for a
glorious
throne to his
father’s
house.
\VS{24}And they shall
hang upon him all the
glory of his
father’s
house, the
offspring and the
issue, all
vessels of
small quantity, from the
vessels of
cups, even to all the
vessels of
flagons.\FTNT{†††}{{\FR 22.24 }
vessels of flagons: or, instruments of viols}
\VS{25}In that
day,
saith the
{\ND{
LORD}} of
hosts, shall the
nail that is
fastened in the
sure
place be
removed, and be cut
down, and
fall; and the
burden that
{\ADD{was}} upon it shall be cut
off: for the
{\ND{
LORD}} hath
spoken
{\ADD{it}}.

\par }\Chap{23}{\PP \VerseOne{1}The
burden of
Tyre.
Howl, ye
ships of
Tarshish; for it is laid
waste, so that there is no
house, no entering
in: from the
land of
Chittim it is
revealed to them.
\VS{2}Be
still, ye
inhabitants of the
isle; thou whom the
merchants of
Zidon, that pass
over the
sea, have
replenished.\FTNT{*}{{\FR 23.2 }
still: Heb. silent}
\VS{3}And by
great
waters the
seed of
Sihor, the
harvest of the
river,
{\ADD{is}} her
revenue; and she is a
mart of
nations.
\VS{4}Be thou
ashamed, O
Zidon: for the
sea hath
spoken,
{\ADD{even}} the
strength of the
sea,
saying, I
travail not, nor bring forth
children, neither do I nourish
up young
men,
{\ADD{nor}} bring
up
virgins.
\VS{5}As at the
report concerning
Egypt,
{\ADD{so}} shall they be sorely
pained at the
report of
Tyre.
\VS{6}Pass ye
over to
Tarshish;
howl, ye
inhabitants of the
isle.
\VS{7}{\ADD{Is}} this your
joyous
{\ADD{city}}, whose
antiquity
{\ADD{is}} of
ancient
days? her own
feet shall
carry her afar
off to
sojourn.\FTNT{†}{{\FR 23.7 }
afar…: Heb. from afar off}
\VS{8}Who hath taken this
counsel against
Tyre, the
crowning
{\ADD{city}}, whose
merchants
{\ADD{are}}
princes, whose
traffickers
{\ADD{are}} the
honourable of the
earth?
\VS{9}The
{\ND{
LORD}} of
hosts hath
purposed it, to
stain the
pride of all
glory,
{\ADD{and}} to bring into
contempt all the
honourable of the
earth.\FTNT{‡}{{\FR 23.9 }
to stain: Heb. to pollute}
\VS{10}Pass through thy
land as a
river, O
daughter of
Tarshish:
{\ADD{there is}} no more
strength.\FTNT{§}{{\FR 23.10 }
strength: Heb. girdle}
\VS{11}He stretched
out his
hand over the
sea, he
shook the
kingdoms: the
{\ND{
LORD}} hath given a
commandment against the
merchant
{\ADD{city}}, to
destroy the
strong
holds thereof.\FTNT{**}{{\FR 23.11 }
against…: or, concerning a merchantman}\FTNT{††}{{\FR 23.11 }
the merchant city: Heb. Canaan}\FTNT{‡‡}{{\FR 23.11 }
strong…: or, strengths}
\VS{12}And he
said, Thou shalt no
more
rejoice, O thou
oppressed
virgin,
daughter of
Zidon:
arise, pass
over to
Chittim; there also shalt thou have no
rest.
\VS{13}Behold the
land of the
Chaldeans; this
people was not,
{\ADD{till}} the
Assyrian
founded it for them that dwell in the
wilderness: they set
up the
towers thereof, they raised
up the
palaces thereof;
{\ADD{and}} he
brought it to
ruin.
\VS{14}Howl, ye
ships of
Tarshish: for your
strength is laid
waste.
\par }{\PP \VS{15}And it shall come to pass in that
day, that
Tyre shall be
forgotten
seventy
years, according to the
days of
one
king: after the
end of
seventy
years shall
Tyre
sing as an
harlot.\FTNT{§§}{{\FR 23.15 }
shall Tyre…: Heb. it shall be unto Tyre as the song of an harlot}
\VS{16}Take an
harp, go
about the
city, thou
harlot that hast been
forgotten; make
sweet
melody, sing
many
songs, that thou mayest be
remembered.
\par }{\PP \VS{17}¶ And it shall come to pass after the
end of
seventy
years, that the
{\ND{
LORD}} will
visit
Tyre, and she shall
turn to her
hire, and shall commit
fornication with all the
kingdoms of the
world upon the
face of the
earth.
\VS{18}And her
merchandise and her
hire shall be
holiness to the
{\ND{
LORD}}: it shall not be
treasured nor laid
up; for her
merchandise shall be for them that
dwell
before the
{\ND{
LORD}}, to
eat
sufficiently, and for
durable
clothing.\FTNT{***}{{\FR 23.18 }
durable: Heb. old}

\par }\Chap{24}{\PP \VerseOne{1}Behold, the
{\ND{
LORD}} maketh the
earth
empty, and maketh it
waste, and
turneth it upside
down, and scattereth
abroad the
inhabitants thereof.\FTNT{*}{{\FR 24.1 }
turneth…: Heb. perverteth the face thereof}
\VS{2}And it shall be, as with the
people, so with the
priest; as with the
servant, so with his
master; as with the
maid, so with her
mistress; as with the
buyer, so with the
seller; as with the
lender, so with the
borrower; as with the taker of
usury,
so with the giver of
usury to him.\FTNT{†}{{\FR 24.2 }
priest: or, prince}
\VS{3}The
land shall be
utterly
emptied, and
utterly
spoiled: for the
{\ND{
LORD}} hath
spoken this
word.
\VS{4}The
earth
mourneth
{\ADD{and}} fadeth
away, the
world
languisheth
{\ADD{and}} fadeth
away, the
haughty
people of the
earth do
languish.\FTNT{‡}{{\FR 24.4 }
the haughty…: Heb. the height of the people}
\VS{5}The
earth also is
defiled under the
inhabitants thereof; because they have
transgressed the
laws,
changed the
ordinance,
broken the
everlasting
covenant.
\VS{6}Therefore hath the
curse
devoured the
earth, and they that
dwell therein are
desolate: therefore the
inhabitants of the
earth are
burned, and
few
men
left.
\VS{7}The new
wine
mourneth, the
vine
languisheth, all the
merryhearted do
sigh.
\VS{8}The
mirth of
tabrets
ceaseth, the
noise of them that
rejoice
endeth, the
joy of the
harp
ceaseth.
\VS{9}They shall not
drink
wine with a
song; strong
drink shall be
bitter to them that
drink it.
\VS{10}The
city of
confusion is broken
down: every
house is shut
up, that no man may come
in.
\VS{11}{\ADD{There is}} a
crying for
wine in the
streets; all
joy is
darkened, the
mirth of the
land is
gone.
\VS{12}In the
city is
left
desolation, and the
gate is
smitten with
destruction.
\par }{\PP \VS{13}¶ When thus it shall be in the
midst of the
land
among the
people,
{\ADD{there shall be}} as the
shaking of an olive
tree,
{\ADD{and}} as the gleaning
grapes when the
vintage is
done.
\VS{14}They shall lift
up their
voice, they shall
sing for the
majesty of the
{\ND{
LORD}}, they shall cry
aloud from the
sea.
\VS{15}Wherefore
glorify ye the
{\ND{
LORD}} in the
fires,
{\ADD{even}} the
name of the
{\ND{
LORD}}
God of
Israel in the
isles of the
sea.\FTNT{§}{{\FR 24.15 }
fires: or, valleys}
\par }{\PP \VS{16}¶ From the uttermost
part of the
earth have we
heard
songs,
{\ADD{even}}
glory to the
righteous. But I
said, My
leanness, my
leanness,
woe unto me! the treacherous
dealers have dealt
treacherously; yea, the treacherous
dealers have dealt
very
treacherously.\FTNT{**}{{\FR 24.16 }
uttermost…: Heb. wing}\FTNT{††}{{\FR 24.16 }
My leanness: Heb. Leanness to me, or, My secret to me}
\VS{17}Fear, and the
pit, and the
snare,
{\ADD{are}} upon thee, O
inhabitant of the
earth.
\VS{18}And it shall come to pass,
{\ADD{that}} he who
fleeth from the
noise of the
fear shall
fall into the
pit; and he that cometh
up out of the
midst of the
pit shall be
taken in the
snare: for the
windows from on
high are
open, and the
foundations of the
earth do
shake.
\VS{19}The
earth is
utterly broken
down, the
earth is
clean
dissolved, the
earth is
moved
exceedingly.
\VS{20}The
earth shall
reel to and
fro like a
drunkard, and shall be
removed like a
cottage; and the
transgression thereof shall be
heavy upon it; and it shall
fall, and not
rise
again.
\VS{21}And it shall come to pass in that
day,
{\ADD{that}} the
{\ND{
LORD}} shall
punish the
host of the high
ones
{\ADD{that are}} on
high, and the
kings of the
earth upon the
earth.\FTNT{‡‡}{{\FR 24.21 }
punish: Heb. visit upon}
\VS{22}And they shall be gathered
together,
{\ADD{as}}
prisoners are
gathered in the
pit, and shall be shut
up in the
prison, and after
many
days shall they be
visited.\FTNT{§§}{{\FR 24.22 }
as prisoners…: Heb. with the gathering of prisoners}\FTNT{***}{{\FR 24.22 }
pit: or, dungeon}\FTNT{†††}{{\FR 24.22 }
visited: or, found wanting}
\VS{23}Then the
moon shall be
confounded, and the
sun
ashamed, when the
{\ND{
LORD}} of
hosts shall
reign in
mount
Zion, and in
Jerusalem, and before his
ancients
gloriously.\FTNT{‡‡‡}{{\FR 24.23 }
before…: or, there shall be glory before his ancients}

\par }\Chap{25}{\PP \VerseOne{1}O
{\ND{
LORD}}, thou
{\ADD{art}} my
God; I will
exalt thee, I will
praise thy
name; for thou hast
done
wonderful
{\ADD{things; thy}}
counsels of
old
{\ADD{are}}
faithfulness
{\ADD{and}}
truth.
\VS{2}For thou hast
made of a
city an
heap;
{\ADD{of}} a
defenced
city a
ruin: a
palace of
strangers to be no
city; it shall
never be
built.
\VS{3}Therefore shall the
strong
people
glorify thee, the
city of the
terrible
nations shall
fear thee.
\VS{4}For thou hast been a
strength to the
poor, a
strength to the
needy in his
distress, a
refuge from the
storm, a
shadow from the
heat, when the
blast of the terrible
ones
{\ADD{is}} as a
storm
{\ADD{against}} the
wall.
\VS{5}Thou shalt bring
down the
noise of
strangers, as the
heat in a dry
place;
{\ADD{even}} the
heat with the
shadow of a
cloud: the
branch of the terrible
ones shall be brought
low.
\par }{\PP \VS{6}¶ And in this
mountain shall the
{\ND{
LORD}} of
hosts
make unto all
people a
feast of fat
things, a
feast of wines on the
lees, of fat
things full of
marrow, of wines on the
lees well
refined.
\VS{7}And he will
destroy in this
mountain the
face of the
covering cast
over all
people, and the
vail that is
spread over all
nations.\FTNT{*}{{\FR 25.7 }
destroy: Heb. swallow up}\FTNT{†}{{\FR 25.7 }
cast: Heb. covered}
\VS{8}He will swallow
up
death in
victory; and the
Lord
{\ND{
GOD}} will wipe
away
tears from off all
faces; and the
rebuke of his
people shall he take
away from off all the
earth: for the
{\ND{
LORD}} hath
spoken
{\ADD{it}}.
\par }{\PP \VS{9}¶ And it shall be
said in that
day, Lo, this
{\ADD{is}} our
God; we have
waited for him, and he will
save us: this
{\ADD{is}} the
{\ND{
LORD}}; we have
waited for him, we will be
glad and
rejoice in his
salvation.
\VS{10}For in this
mountain shall the
hand of the
{\ND{
LORD}}
rest, and
Moab shall be trodden
down under him, even as
straw is trodden
down for the
dunghill.\FTNT{‡}{{\FR 25.10 }
trodden down under: or, threshed, etc}\FTNT{§}{{\FR 25.10 }
trodden down for…: or, threshed in Madmenah}
\VS{11}And he shall spread
forth his
hands in the
midst of them, as he that
swimmeth spreadeth
forth
{\ADD{his hands}} to
swim: and he shall bring
down their
pride together with the
spoils of their
hands.
\VS{12}And the
fortress of the high
fort of thy
walls shall he bring
down, lay
low,
{\ADD{and}}
bring to the
ground,
{\ADD{even}} to the
dust.

\par }\Chap{26}{\PP \VerseOne{1}In that
day shall this
song be
sung in the
land of
Judah; We have a
strong
city;
salvation will
{\ADD{God}}
appoint
{\ADD{for}}
walls and
bulwarks.
\VS{2}Open ye the
gates, that the
righteous
nation which
keepeth the
truth may enter
in.\FTNT{*}{{\FR 26.2 }
truth: Heb. truths}
\VS{3}Thou wilt
keep
{\ADD{him}} in
perfect
peace,
{\ADD{whose}}
mind
{\ADD{is}}
stayed
{\ADD{on thee:}} because he
trusteth in thee.\FTNT{†}{{\FR 26.3 }
perfect…: Heb. peace, peace}\FTNT{‡}{{\FR 26.3 }
mind: or, thought, or, imagination}
\VS{4}Trust ye in the
{\ND{
LORD}} for
ever: for in the
{\ND{
LORD}}
{\ND{
JEHOVAH}}
{\ADD{is}}
everlasting
strength:\FTNT{§}{{\FR 26.4 }
everlasting…: Heb. the rock of ages}
\par }{\PP \VS{5}¶ For he bringeth
down them that
dwell on
high; the
lofty
city, he layeth it
low; he layeth it
low,
{\ADD{even}} to the
ground; he
bringeth it
{\ADD{even}} to the
dust.
\VS{6}The
foot shall tread it
down,
{\ADD{even}} the
feet of the
poor,
{\ADD{and}} the
steps of the
needy.
\VS{7}The
way of the
just
{\ADD{is}}
uprightness: thou, most
upright, dost
weigh the
path of the
just.
\VS{8}Yea, in the
way of thy
judgments, O
{\ND{
LORD}}, have we
waited for thee; the
desire of
{\ADD{our}}
soul
{\ADD{is}} to thy
name, and to the
remembrance of thee.
\VS{9}With my
soul have I
desired thee in the
night; yea, with my
spirit
within me will I seek thee
early: for when thy
judgments
{\ADD{are}} in the
earth, the
inhabitants of the
world will
learn
righteousness.
\VS{10}Let favour be
shewed to the
wicked,
{\ADD{yet}} will he
not
learn
righteousness: in the
land of
uprightness will he deal
unjustly, and will
not
behold the
majesty of the
{\ND{
LORD}}.
\VS{11}{\ND{
LORD}},
{\ADD{when}} thy
hand is lifted
up, they will not
see:
{\ADD{but}} they shall
see, and be
ashamed for
{\ADD{their}}
envy at the
people; yea, the
fire of thine
enemies shall
devour them.\FTNT{**}{{\FR 26.11 }
at…: or, toward thy people}
\par }{\PP \VS{12}¶
{\ND{
LORD}}, thou wilt
ordain
peace for us: for thou also hast
wrought all our
works in us.\FTNT{††}{{\FR 26.12 }
in us: or, for us}
\VS{13}O
{\ND{
LORD}} our
God,
{\ADD{other}}
lords
beside thee have had
dominion over us:
{\ADD{but}} by thee only will we make
mention of thy
name.
\VS{14}{\ADD{They are}}
dead, they shall not
live;
{\ADD{they are}}
deceased, they shall not
rise: therefore hast thou
visited and
destroyed them, and made all their
memory to
perish.
\VS{15}Thou hast
increased the
nation, O
{\ND{
LORD}}, thou hast
increased the
nation: thou art
glorified: thou hadst removed
{\ADD{it}}
far
{\ADD{unto}} all the
ends of the
earth.
\VS{16}{\ND{
LORD}}, in
trouble have they
visited thee, they poured
out a
prayer
{\ADD{when}} thy
chastening
{\ADD{was}} upon them.\FTNT{‡‡}{{\FR 26.16 }
prayer: Heb. secret speech}
\VS{17}Like as a woman with
child,
{\ADD{that}} draweth
near the time of her
delivery, is in
pain,
{\ADD{and}} crieth
out in her
pangs; so have we been in thy
sight, O
{\ND{
LORD}}.
\VS{18}We have been with
child, we have been in
pain, we have as it
were brought
forth
wind; we have not
wrought any
deliverance in the
earth;
neither have the
inhabitants of the
world
fallen.
\VS{19}Thy
dead
{\ADD{men}} shall
live,
{\ADD{together with}} my dead
body shall they
arise.
Awake and
sing, ye that
dwell in
dust: for thy
dew
{\ADD{is as}} the
dew of
herbs, and the
earth shall cast
out the
dead.
\par }{\PP \VS{20}¶
Come, my
people,
enter thou into thy
chambers, and
shut thy
doors about thee:
hide thyself as it were for a
little
moment, until the
indignation be
overpast.
\VS{21}For, behold, the
{\ND{
LORD}} cometh
out of his
place to
punish the
inhabitants of the
earth for their
iniquity: the
earth also shall
disclose her
blood, and shall no more
cover her
slain.\FTNT{§§}{{\FR 26.21 }
blood: Heb. bloods}

\par }\Chap{27}{\PP \VerseOne{1}In that
day the
{\ND{
LORD}} with his
sore and
great and
strong
sword shall
punish
leviathan the
piercing
serpent, even
leviathan that
crooked
serpent; and he shall
slay the
dragon that
{\ADD{is}} in the
sea.\FTNT{*}{{\FR 27.1 }
piercing: or, crossing like a bar}
\VS{2}In that
day
sing ye unto her, A
vineyard of red
wine.
\VS{3}I the
{\ND{
LORD}} do
keep it; I will
water it every
moment: lest
{\ADD{any}}
hurt it, I will
keep it
night and
day.
\VS{4}Fury
{\ADD{is}} not in me: who would
set the
briers
{\ADD{and}}
thorns against me in
battle? I would
go through them, I would
burn them
together.\FTNT{†}{{\FR 27.4 }
go…: or, march against}
\VS{5}Or let him take
hold of my
strength,
{\ADD{that}} he may
make
peace with me;
{\ADD{and}} he shall
make
peace with me.
\VS{6}He shall cause them that
come of
Jacob to take
root:
Israel shall
blossom and
bud, and
fill the
face of the
world with
fruit.
\par }{\PP \VS{7}¶ Hath he
smitten him, as he
smote those that
smote him?
{\ADD{or}} is he
slain according to the
slaughter of them that are
slain by him?\FTNT{‡}{{\FR 27.7 }
as…: Heb. according to the stroke of}
\VS{8}In
measure, when it shooteth
forth, thou wilt
debate with it: he
stayeth his
rough
wind in the
day of the east
wind.\FTNT{§}{{\FR 27.8 }
it shooteth…: or, thou sendest it forth}\FTNT{**}{{\FR 27.8 }
he…: or, when he removeth it with}
\VS{9}By
this therefore shall the
iniquity of
Jacob be
purged; and this
{\ADD{is}} all the
fruit to take
away his
sin; when he
maketh all the
stones of the
altar as
chalkstones that are beaten in
sunder, the
groves and
images shall not stand
up.\FTNT{††}{{\FR 27.9 }
images: or, sun images}
\VS{10}Yet the
defenced
city
{\ADD{shall be}}
desolate,
{\ADD{and}} the
habitation
forsaken, and
left like a
wilderness: there shall the
calf
feed, and there shall he lie
down, and
consume the
branches thereof.
\VS{11}When the
boughs thereof are
withered, they shall be broken
off: the
women
come,
{\ADD{and}} set them on
fire: for it
{\ADD{is}} a
people of no
understanding: therefore he that
made them will not have
mercy on them, and he that
formed them will shew them no
favour.
\par }{\PP \VS{12}¶ And it shall come to pass in that
day,
{\ADD{that}} the
{\ND{
LORD}} shall beat
off from the
channel of the
river unto the
stream of
Egypt, and ye shall be
gathered
one by
one, O ye
children of
Israel.
\VS{13}And it shall come to pass in that
day,
{\ADD{that}} the
great
trumpet shall be
blown, and they shall
come which were ready to
perish in the
land of
Assyria, and the
outcasts in the
land of
Egypt, and shall
worship the
{\ND{
LORD}} in the
holy
mount at
Jerusalem.

\par }\Chap{28}{\PP \VerseOne{1}Woe to the
crown of
pride, to the
drunkards of
Ephraim, whose
glorious
beauty
{\ADD{is}} a
fading
flower, which
{\ADD{are}} on the
head of the
fat
valleys of them that are
overcome with
wine!\FTNT{*}{{\FR 28.1 }
overcome: Heb. broken}
\VS{2}Behold, the
Lord hath a
mighty and strong
one,
{\ADD{which}} as a
tempest of
hail
{\ADD{and}} a
destroying
storm, as a
flood of
mighty
waters
overflowing, shall cast
down to the
earth with the
hand.
\VS{3}The
crown of
pride, the
drunkards of
Ephraim, shall be
trodden under
feet:\FTNT{†}{{\FR 28.3 }
under…: Heb. with feet}
\VS{4}And the
glorious
beauty, which
{\ADD{is}} on the
head of the
fat
valley, shall be a
fading
flower,
{\ADD{and}} as the hasty
fruit before the
summer; which
{\ADD{when}} he that
looketh upon it
seeth, while it is yet in his
hand he eateth it
up.\FTNT{‡}{{\FR 28.4 }
eateth: Heb. swalloweth}
\par }{\PP \VS{5}¶ In that
day shall the
{\ND{
LORD}} of
hosts be for a
crown of
glory, and for a
diadem of
beauty, unto the
residue of his
people,
\VS{6}And for a
spirit of
judgment to him that
sitteth in
judgment, and for
strength to them that
turn the
battle to the
gate.
\par }{\PP \VS{7}¶ But
they also have
erred through
wine, and through strong
drink are out of the
way; the
priest and the
prophet have
erred through strong
drink, they are swallowed
up
of
wine, they are out of the
way through strong
drink; they
err in
vision, they
stumble
{\ADD{in}}
judgment.
\VS{8}For all
tables are
full of
vomit
{\ADD{and}}
filthiness,
{\ADD{so that there is}} no
place
{\ADD{clean}}.
\par }{\PP \VS{9}¶ Whom shall he
teach
knowledge? and whom shall he make to
understand
doctrine?
{\ADD{them that are}}
weaned from the
milk,
{\ADD{and}}
drawn from the
breasts.\FTNT{§}{{\FR 28.9 }
doctrine: Heb. the hearing?}
\VS{10}For
precept
{\ADD{must be}} upon
precept,
precept upon
precept;
line upon
line,
line upon
line; here a
little,
{\ADD{and}} there a
little:\FTNT{**}{{\FR 28.10 }
must be: or, hath been}
\VS{11}For with
stammering
lips and
another
tongue will he
speak to this
people.\FTNT{††}{{\FR 28.11 }
stammering…: Heb. stammerings of lip}\FTNT{‡‡}{{\FR 28.11 }
will…: or, he hath spoken}
\VS{12}To whom he
said, This
{\ADD{is}} the
rest
{\ADD{wherewith}} ye may cause the
weary to
rest; and this
{\ADD{is}} the
refreshing: yet they
would not
hear.
\VS{13}But the
word of the
{\ND{
LORD}} was unto them
precept upon
precept,
precept upon
precept;
line upon
line,
line upon
line; here a
little,
{\ADD{and}} there a
little; that they might
go, and
fall
backward, and be
broken, and
snared, and
taken.
\par }{\PP \VS{14}¶ Wherefore
hear the
word of the
{\ND{
LORD}}, ye
scornful
men, that
rule this
people which
{\ADD{is}} in
Jerusalem.
\VS{15}Because ye have
said, We have
made a
covenant with
death, and with
hell are
we at
agreement; when the
overflowing
scourge shall pass
through, it shall not
come unto us: for we have
made
lies our
refuge, and under
falsehood have we
hid ourselves:
\par }{\PP \VS{16}¶ Therefore thus
saith the
Lord
{\ND{
GOD}}, Behold, I
lay in
Zion for a foundation a
stone, a
tried
stone, a
precious
corner
{\ADD{stone}}, a
sure
foundation: he that
believeth shall not make
haste.
\VS{17}Judgment also will I
lay to the
line, and
righteousness to the
plummet: and the
hail shall sweep
away the
refuge of
lies, and the
waters shall
overflow the hiding
place.
\par }{\PP \VS{18}¶ And your
covenant with
death shall be
disannulled, and your
agreement with
hell shall not
stand; when the
overflowing
scourge shall pass
through, then ye shall be trodden
down by it.\FTNT{§§}{{\FR 28.18 }
trodden…: Heb. a treading down to it}
\VS{19}From the
time that it goeth
forth it shall
take you: for
morning by
morning shall it pass
over, by
day and by
night: and it shall be a
vexation only
{\ADD{to}}
understand the
report.\FTNT{***}{{\FR 28.19 }
to…: or, when he shall make you to understand doctrine}
\VS{20}For the
bed is
shorter than that
{\ADD{a man}} can
stretch himself
{\ADD{on it:}} and the
covering
narrower than that he can
wrap himself
{\ADD{in it}}.
\VS{21}For the
{\ND{
LORD}} shall rise
up as
{\ADD{in}}
mount
Perazim, he shall be
wroth as
{\ADD{in}} the
valley of
Gibeon, that he may
do his
work, his
strange
work; and bring to
pass his
act, his
strange
act.
\VS{22}Now therefore be ye not
mockers, lest your
bands be made
strong: for I have
heard from the
Lord
{\ND{
GOD}} of
hosts a
consumption, even
determined upon the whole
earth.
\par }{\PP \VS{23}¶ Give ye
ear, and
hear my
voice;
hearken, and
hear my
speech.
\VS{24}Doth the
plowman
plow all
day to
sow? doth he
open and break the
clods of his
ground?
\VS{25}When he hath made
plain the
face thereof, doth he not cast
abroad the
fitches, and
scatter the
cummin, and
cast in the
principal
wheat and the
appointed
barley and the
rie in their
place?\FTNT{†††}{{\FR 28.25 }
the principal…: or, the wheat in the principal place, and barley in the appointed place}\FTNT{‡‡‡}{{\FR 28.25 }
rie: or, spelt}\FTNT{§§§}{{\FR 28.25 }
place: Heb. border?}
\VS{26}For his
God doth
instruct him to
discretion,
{\ADD{and}} doth
teach him.\FTNT{*}{{\FR 28.26 }
For…: or, And he bindeth it in such sort as his God doth teach him}
\VS{27}For the
fitches are not
threshed with a threshing
instrument, neither is a
cart
wheel turned
about upon the
cummin; but the
fitches are beaten
out with a
staff, and the
cummin with a
rod.
\VS{28}Bread
{\ADD{corn}} is
bruised; because he will not
ever be
threshing it, nor
break
{\ADD{it with}} the
wheel of his
cart, nor
bruise it
{\ADD{with}} his
horsemen.
\VS{29}This also cometh
forth from the
{\ND{
LORD}} of
hosts,
{\ADD{which}} is
wonderful in
counsel,
{\ADD{and}}
excellent in
working.

\par }\Chap{29}{\PP \VerseOne{1}Woe to
Ariel, to
Ariel, the
city
{\ADD{where}}
David
dwelt!
add ye
year to
year; let them
kill
sacrifices.\FTNT{*}{{\FR 29.1 }
Woe…: or, O Ariel, that is, the lion of God}\FTNT{†}{{\FR 29.1 }
the city: or, of the city}\FTNT{‡}{{\FR 29.1 }
kill: Heb. cut off the heads of}
\VS{2}Yet I will
distress
Ariel, and there shall be
heaviness and
sorrow: and it shall be unto me as
Ariel.
\VS{3}And I will
camp against thee round
about, and will lay
siege against thee with a
mount, and I will
raise
forts against thee.
\VS{4}And thou shalt be brought
down,
{\ADD{and}} shalt
speak out of the
ground, and thy
speech shall be
low out of the
dust, and thy
voice shall be, as of one that hath a familiar
spirit, out of the
ground, and thy
speech shall
whisper out of the
dust.\FTNT{§}{{\FR 29.4 }
whisper: Heb. peep, or, chirp}
\VS{5}Moreover the
multitude of thy
strangers shall be like
small
dust, and the
multitude of the terrible
ones
{\ADD{shall be}} as
chaff that passeth
away: yea, it shall be at an
instant
suddenly.
\VS{6}Thou shalt be
visited of the
{\ND{
LORD}} of
hosts with
thunder, and with
earthquake, and
great
noise, with
storm and
tempest, and the
flame of
devouring
fire.
\par }{\PP \VS{7}¶ And the
multitude of all the
nations that
fight against
Ariel, even all that
fight against her and her
munition, and that
distress her, shall be as a
dream of a
night
vision.
\VS{8}It shall even be as when an
hungry
{\ADD{man}}
dreameth, and, behold, he
eateth; but he
awaketh, and his
soul is
empty: or as
when a thirsty
man
dreameth, and, behold, he
drinketh; but he
awaketh, and, behold,
{\ADD{he is}}
faint, and his
soul hath
appetite: so shall the
multitude of all the
nations be, that
fight against
mount
Zion.
\par }{\PP \VS{9}¶
Stay yourselves, and
wonder; cry ye
out, and
cry: they are
drunken, but not with
wine; they
stagger, but not with strong
drink.\FTNT{**}{{\FR 29.9 }
cry ye…: or, take your pleasure, and riot}
\VS{10}For the
{\ND{
LORD}} hath poured
out upon you the
spirit of deep
sleep, and hath
closed your
eyes: the
prophets and your
rulers, the
seers hath he
covered.\FTNT{††}{{\FR 29.10 }
rulers: Heb. heads}
\VS{11}And the
vision of all is become unto you as the
words of a
book that is
sealed, which
{\ADD{men}}
deliver to one that is
learned,
saying,
Read this, I pray thee: and he
saith, I
cannot; for it
{\ADD{is}}
sealed:\FTNT{‡‡}{{\FR 29.11 }
book: or, letter}
\VS{12}And the
book is
delivered to him that is not
learned,
saying,
Read this, I pray thee: and he
saith, I am not
learned.
\par }{\PP \VS{13}¶ Wherefore the
Lord
said,
Forasmuch as this
people draw
near
{\ADD{me}} with their
mouth, and with their
lips do
honour me, but have
removed their
heart
far from me, and their
fear toward me is
taught by the
precept of
men:
\VS{14}Therefore, behold, I will
proceed to do a marvellous
work among this
people,
{\ADD{even}} a marvellous
work and a
wonder: for the
wisdom of their
wise
{\ADD{men}} shall
perish, and the
understanding of their
prudent
{\ADD{men}} shall be
hid.\FTNT{§§}{{\FR 29.14 }
proceed: Heb. add}
\VS{15}Woe unto them that seek
deep to
hide their
counsel from the
{\ND{
LORD}}, and their
works are in the
dark, and they
say, Who
seeth us? and who
knoweth us?
\VS{16}Surely your turning of things upside
down shall be
esteemed as the
potter’s
clay: for shall the
work
say of him that
made it, He
made me not? or shall the thing
framed
say of him that
framed it, He had no
understanding?
\par }{\PP \VS{17}{\ADD{Is}} it not yet a
very little
while, and
Lebanon shall be
turned into a fruitful
field, and the fruitful
field shall be
esteemed as a
forest?
\par }{\PP \VS{18}¶ And in that
day shall the
deaf
hear the
words of the
book, and the
eyes of the
blind shall
see out of
obscurity, and out of
darkness.
\VS{19}The
meek also shall
increase
{\ADD{their}}
joy in the
{\ND{
LORD}}, and the
poor among
men shall
rejoice in the Holy
One of
Israel.\FTNT{***}{{\FR 29.19 }
increase: Heb. add}
\VS{20}For the terrible
one is brought to
nought, and the
scorner is
consumed, and all that
watch for
iniquity are cut
off:
\VS{21}That make a
man an
offender for a
word, and lay a
snare for him that
reproveth in the
gate, and turn
aside the
just for a thing of
nought.
\VS{22}Therefore thus
saith the
{\ND{
LORD}}, who
redeemed
Abraham, concerning the
house of
Jacob,
Jacob shall not now be
ashamed, neither shall his
face now wax
pale.
\VS{23}But when he
seeth his
children, the
work of mine
hands, in the
midst of him, they shall
sanctify my
name, and
sanctify the Holy
One of
Jacob, and shall
fear the
God of
Israel.
\VS{24}They also that
erred in
spirit shall
come to
understanding, and they that
murmured shall
learn
doctrine.\FTNT{†††}{{\FR 29.24 }
come…: Heb. know understanding}

\par }\Chap{30}{\PP \VerseOne{1}Woe to the
rebellious
children,
saith the
{\ND{
LORD}}, that
take
counsel, but not of me; and that
cover with a
covering, but not of my
spirit, that they may
add
sin to
sin:
\VS{2}That
walk to go
down into
Egypt, and have not
asked at my
mouth; to
strengthen themselves in the
strength of
Pharaoh, and to
trust in the
shadow of
Egypt!
\VS{3}Therefore shall the
strength of
Pharaoh be your
shame, and the
trust in the
shadow of
Egypt
{\ADD{your}}
confusion.
\VS{4}For his
princes were at
Zoan, and his
ambassadors
came to
Hanes.
\VS{5}They were all
ashamed of a
people
{\ADD{that}} could not
profit them, nor be an
help nor
profit, but a
shame, and also a
reproach.
\VS{6}The
burden of the
beasts of the
south: into the
land of
trouble and
anguish, from whence
{\ADD{come}} the
young and old
lion, the
viper and
fiery
flying
serpent, they will
carry their
riches upon the
shoulders of young
asses, and their
treasures upon the
bunches of
camels, to a
people
{\ADD{that}} shall not
profit
{\ADD{them}}.
\VS{7}For the
Egyptians shall
help in
vain, and to no
purpose: therefore have I
cried concerning
this,
Their
strength
{\ADD{is}} to sit
still.\FTNT{*}{{\FR 30.7 }
concerning…: or, to her}
\par }{\PP \VS{8}¶ Now
go,
write it before them in a
table, and
note it in a
book, that it may be for the
time to
come
for
ever and
ever:\FTNT{†}{{\FR 30.8 }
the…: Heb. the latter day}
\VS{9}That this
{\ADD{is}} a
rebellious
people,
lying
children,
children
{\ADD{that}}
will not
hear the
law of the
{\ND{
LORD}}:
\VS{10}Which
say to the
seers,
See not; and to the
prophets,
Prophesy not unto us right
things,
speak unto us smooth
things,
prophesy
deceits:
\VS{11}Get you
out of the
way, turn
aside out of the
path, cause the Holy
One of
Israel to
cease from
before us.
\VS{12}Wherefore thus
saith the Holy
One of
Israel, Because ye
despise this
word, and
trust in
oppression and
perverseness, and
stay thereon:\FTNT{‡}{{\FR 30.12 }
oppression: or, fraud}
\VS{13}Therefore this
iniquity shall be to you as a
breach ready to
fall, swelling
out in a
high
wall, whose
breaking
cometh
suddenly at an
instant.
\VS{14}And he shall
break it as the
breaking of the
potters’
vessel that is broken in
pieces; he shall not
spare: so that there shall not be
found in the
bursting of it a
sherd to
take
fire from the
hearth, or to
take
water
{\ADD{withal}} out of the
pit.\FTNT{§}{{\FR 30.14 }
the potters’…: Heb. the bottle of potters}
\VS{15}For thus
saith the
Lord
{\ND{
GOD}}, the Holy
One of
Israel; In
returning and
rest shall ye be
saved; in
quietness and in
confidence shall be your
strength: and ye
would not.
\VS{16}But ye
said, No; for we will
flee upon
horses; therefore shall ye
flee: and, We will
ride upon the
swift; therefore shall they that
pursue you be
swift.
\VS{17}One
thousand
{\ADD{shall flee}}
at the
rebuke of
one;
at the
rebuke of
five shall ye
flee: till ye be
left as a
beacon upon the
top of a
mountain, and as an
ensign on an
hill.\FTNT{**}{{\FR 30.17 }
a beacon: or, a tree bereft of branches, or, boughs: or, a mast}
\par }{\PP \VS{18}¶ And therefore will the
{\ND{
LORD}}
wait, that he may be
gracious unto you, and therefore will he be
exalted, that he may have
mercy upon you: for the
{\ND{
LORD}}
{\ADD{is}} a
God of
judgment:
blessed
{\ADD{are}} all they that
wait for him.
\VS{19}For the
people shall
dwell in
Zion at
Jerusalem: thou shalt
weep no
more: he will be
very
gracious unto thee at the
voice of thy
cry; when he shall
hear it, he will
answer thee.
\VS{20}And
{\ADD{though}} the
Lord
give you the
bread of
adversity, and the
water of
affliction, yet shall not thy
teachers be removed into a
corner any more, but thine
eyes shall
see thy
teachers:\FTNT{††}{{\FR 30.20 }
affliction: or, oppression}
\VS{21}And thine
ears shall
hear a
word
behind thee,
saying, This
{\ADD{is}} the
way, walk ye in
it, when ye turn to the right
hand, and when ye turn to the
left.
\VS{22}Ye shall
defile also the
covering of thy graven
images of
silver, and the
ornament of thy molten
images of
gold: thou shalt cast them
away as a menstruous
cloth; thou shalt
say unto it, Get thee
hence.\FTNT{‡‡}{{\FR 30.22 }
thy graven…: Heb. the graven images of thy silver}\FTNT{§§}{{\FR 30.22 }
cast: Heb. scatter}
\VS{23}Then shall he
give the
rain of thy
seed, that thou shalt
sow the
ground withal; and
bread of the
increase of the
earth, and it shall be
fat and
plenteous: in that
day shall thy
cattle
feed in
large
pastures.
\VS{24}The
oxen likewise and the young
asses that
ear the
ground shall
eat
clean
provender, which hath been
winnowed with the
shovel and with the
fan.\FTNT{***}{{\FR 30.24 }
clean: or, savoury: Heb. leavened}
\VS{25}And there shall be upon every
high
mountain, and upon every
high
hill,
rivers
{\ADD{and}}
streams of
waters in the
day of the
great
slaughter, when the
towers
fall.\FTNT{†††}{{\FR 30.25 }
high hill: Heb. lifted up, etc}
\VS{26}Moreover the
light of the
moon shall be as the
light of the
sun, and the
light of the
sun shall be
sevenfold, as the
light of
seven
days, in the
day that the
{\ND{
LORD}} bindeth
up the
breach of his
people, and
healeth the
stroke of their
wound.
\par }{\PP \VS{27}¶ Behold, the
name of the
{\ND{
LORD}}
cometh from
far,
burning
{\ADD{with}} his
anger, and the
burden
{\ADD{thereof is}}
heavy: his
lips are
full of
indignation, and his
tongue as a
devouring
fire:\FTNT{‡‡‡}{{\FR 30.27 }
burden…: or, grievousness of flame}\FTNT{§§§}{{\FR 30.27 }
heavy: Heb. heaviness}
\VS{28}And his
breath, as an
overflowing
stream, shall reach to the
midst of the
neck, to
sift the
nations with the
sieve of
vanity: and
{\ADD{there shall be}} a
bridle in the
jaws of the
people, causing
{\ADD{them}} to
err.
\VS{29}Ye shall have a
song, as in the
night
{\ADD{when}} a
holy
solemnity is
kept; and
gladness of
heart, as when one
goeth with a
pipe to
come into the
mountain of the
{\ND{
LORD}}, to the mighty
One of
Israel.\FTNT{*}{{\FR 30.29 }
mighty…: Heb. Rock}
\VS{30}And the
{\ND{
LORD}} shall cause his
glorious
voice to be
heard, and shall
shew the lighting
down of his
arm, with the
indignation of
{\ADD{his}}
anger, and
{\ADD{with}} the
flame of a
devouring
fire,
{\ADD{with}}
scattering, and
tempest, and
hailstones.\FTNT{†}{{\FR 30.30 }
his glorious…: Heb. the glory of his voice}
\VS{31}For through the
voice of the
{\ND{
LORD}} shall the
Assyrian be beaten
down,
{\ADD{which}}
smote with a
rod.
\VS{32}And
{\ADD{in}} every
place where the
grounded
staff shall
pass, which the
{\ND{
LORD}} shall
lay upon him,
{\ADD{it}} shall be with
tabrets and
harps: and in
battles of
shaking will he
fight with it.\FTNT{‡}{{\FR 30.32 }
in every…: Heb. every passing of the rod founded}\FTNT{§}{{\FR 30.32 }
lay…: Heb. cause to rest upon}\FTNT{**}{{\FR 30.32 }
with it: or, against them}
\VS{33}For
Tophet
{\ADD{is}}
ordained of
old; yea, for the
king it is
prepared; he hath made
{\ADD{it}}
deep
{\ADD{and}}
large: the
pile thereof
{\ADD{is}}
fire and
much
wood; the
breath of the
{\ND{
LORD}}, like a
stream of
brimstone, doth
kindle it.\FTNT{††}{{\FR 30.33 }
of old: Heb. from yesterday}

\par }\Chap{31}{\PP \VerseOne{1}Woe to them that go
down to
Egypt for
help; and
stay on
horses, and
trust in
chariots, because
{\ADD{they are}}
many; and in
horsemen, because they are
very
strong; but they
look not unto the Holy
One of
Israel, neither
seek the
{\ND{
LORD}}!
\VS{2}Yet he also
{\ADD{is}}
wise, and will
bring
evil, and will not call
back his
words: but will
arise against the
house of the
evildoers, and against the
help of them that
work
iniquity.\FTNT{*}{{\FR 31.2 }
call…: Heb. remove}
\VS{3}Now the
Egyptians
{\ADD{are}}
men, and not
God; and their
horses
flesh, and not
spirit. When the
{\ND{
LORD}} shall stretch
out his
hand, both he that
helpeth shall
fall, and he that is
holpen shall fall
down, and they all shall
fail
together.
\VS{4}For thus hath the
{\ND{
LORD}}
spoken unto me, Like as the
lion and the young
lion
roaring on his
prey, when a
multitude of
shepherds is called
forth against him,
{\ADD{he}} will not be
afraid of their
voice, nor
abase himself for the
noise of them: so shall the
{\ND{
LORD}} of
hosts come
down to
fight for
mount
Zion, and for the
hill thereof.\FTNT{†}{{\FR 31.4 }
noise: or, multitude}
\VS{5}As
birds
flying, so will the
{\ND{
LORD}} of
hosts
defend
Jerusalem;
defending also he will
deliver
{\ADD{it; and}} passing
over he will
preserve it.
\par }{\PP \VS{6}¶
Turn ye unto
{\ADD{him from}} whom the
children of
Israel have
deeply
revolted.
\VS{7}For in that
day every
man shall cast
away his
idols of
silver, and his
idols of
gold, which your own
hands have
made unto you
{\ADD{for}} a
sin.\FTNT{‡}{{\FR 31.7 }
his idols of gold: Heb. the idols of his gold}
\par }{\PP \VS{8}¶ Then shall the
Assyrian
fall with the
sword, not of a mighty
man; and the
sword, not of a mean
man, shall
devour him: but he shall
flee
from the
sword, and his young
men shall be
discomfited.\FTNT{§}{{\FR 31.8 }
from: or, for fear of}\FTNT{**}{{\FR 31.8 }
discomfited: or, tributary: Heb. for melting, or, tribute}
\VS{9}And he shall pass
over to his strong
hold for
fear, and his
princes shall be
afraid of the
ensign,
saith the
{\ND{
LORD}}, whose
fire
{\ADD{is}} in
Zion, and his
furnace in
Jerusalem.\FTNT{††}{{\FR 31.9 }
he…: Heb. his rock shall pass away for fear}\FTNT{‡‡}{{\FR 31.9 }
his strong hold: or, his strength}

\par }\Chap{32}{\PP \VerseOne{1}Behold, a
king shall
reign in
righteousness, and
princes shall
rule in
judgment.
\VS{2}And a
man shall be as an hiding
place from the
wind, and a
covert from the
tempest; as
rivers of
water in a dry
place, as the
shadow of a
great
rock in a
weary
land.\FTNT{*}{{\FR 32.2 }
great: Heb. heavy}
\VS{3}And the
eyes of them that
see shall not be
dim, and the
ears of them that
hear shall
hearken.
\VS{4}The
heart also of the
rash shall
understand
knowledge, and the
tongue of the
stammerers shall be
ready to
speak
plainly.\FTNT{†}{{\FR 32.4 }
rash: Heb. hasty}\FTNT{‡}{{\FR 32.4 }
plainly: or, elegantly}
\VS{5}The vile
person shall be no more
called
liberal, nor the
churl
said
{\ADD{to be}}
bountiful.
\VS{6}For the vile
person will
speak
villany, and his
heart will
work
iniquity, to
practise
hypocrisy, and to
utter
error against the
{\ND{
LORD}}, to make
empty the
soul of the
hungry, and he will cause the
drink of the
thirsty to
fail.
\VS{7}The
instruments also of the
churl
{\ADD{are}}
evil: he
deviseth wicked
devices to
destroy the
poor with
lying
words, even when the
needy
speaketh
right.\FTNT{§}{{\FR 32.7 }
the needy…: or, he speaketh against the poor in judgment}
\VS{8}But the
liberal
deviseth liberal
things; and by liberal
things shall he
stand.\FTNT{**}{{\FR 32.8 }
stand: or, be established}
\par }{\PP \VS{9}¶ Rise
up, ye
women that are at
ease;
hear my
voice, ye
careless
daughters; give
ear unto my
speech.
\VS{10}Many
days and
years shall ye be
troubled, ye careless
women: for the
vintage shall
fail, the
gathering shall not
come.\FTNT{††}{{\FR 32.10 }
Many…: Heb. Days above a year}
\VS{11}Tremble, ye women that are at
ease; be
troubled, ye careless
ones:
strip you, and make you
bare, and
gird
{\ADD{sackcloth}} upon
{\ADD{your}}
loins.
\VS{12}They shall
lament for the
teats, for the
pleasant
fields, for the
fruitful
vine.\FTNT{‡‡}{{\FR 32.12 }
pleasant…: Heb. fields of desire}
\VS{13}Upon the
land of my
people shall come
up
thorns
{\ADD{and}}
briers; yea, upon all the
houses of
joy
{\ADD{in}} the
joyous
city:\FTNT{§§}{{\FR 32.13 }
yea…: or, burning upon}
\VS{14}Because the
palaces shall be
forsaken; the
multitude of the
city shall be
left; the
forts and
towers shall be for
dens
for
ever, a
joy of wild
asses, a
pasture of
flocks;\FTNT{***}{{\FR 32.14 }
forts…: or, clifts and watchtowers}
\VS{15}Until the
spirit be
poured upon us from on
high, and the
wilderness be a fruitful
field, and the fruitful
field be
counted for a
forest.
\VS{16}Then
judgment shall
dwell in the
wilderness, and
righteousness
remain in the fruitful
field.
\VS{17}And the
work of
righteousness shall be
peace; and the
effect of
righteousness
quietness and
assurance
for
ever.
\VS{18}And my
people shall
dwell in a
peaceable
habitation, and in
sure
dwellings, and in
quiet resting
places;
\VS{19}When it shall
hail, coming
down on the
forest; and the
city shall be
low in a low
place.\FTNT{†††}{{\FR 32.19 }
low in…: or, utterly abased}
\VS{20}Blessed
{\ADD{are}} ye that
sow beside all
waters, that send
forth
{\ADD{thither}} the
feet of the
ox and the
ass.

\par }\Chap{33}{\PP \VerseOne{1}Woe to thee that
spoilest, and thou
{\ADD{wast}} not
spoiled; and dealest
treacherously, and they dealt not
treacherously with thee! when thou shalt
cease to
spoil, thou shalt be
spoiled;
{\ADD{and}} when thou shalt make an
end to deal
treacherously, they shall deal
treacherously with thee.
\VS{2}O
{\ND{
LORD}}, be
gracious unto us; we have
waited for thee: be thou their
arm every
morning, our
salvation also in the
time of
trouble.
\VS{3}At the
noise of the
tumult the
people
fled; at the lifting
up of thyself the
nations were
scattered.
\VS{4}And your
spoil shall be
gathered
{\ADD{like}} the
gathering of the
caterpiller: as the running to and
fro of
locusts shall he
run upon them.
\VS{5}The
{\ND{
LORD}} is
exalted; for he
dwelleth on
high: he hath
filled
Zion with
judgment and
righteousness.
\VS{6}And
wisdom and
knowledge shall be the
stability of thy
times,
{\ADD{and}}
strength of
salvation: the
fear of the
{\ND{
LORD}}
{\ADD{is}} his
treasure.\FTNT{*}{{\FR 33.6 }
salvation: Heb. salvations}
\VS{7}Behold, their valiant
ones shall
cry
without: the
ambassadors of
peace shall
weep
bitterly.\FTNT{†}{{\FR 33.7 }
valiant…: or, messengers}
\VS{8}The
highways lie
waste, the wayfaring
man
ceaseth: he hath
broken the
covenant, he hath
despised the
cities, he
regardeth no
man.
\VS{9}The
earth
mourneth
{\ADD{and}}
languisheth:
Lebanon is
ashamed
{\ADD{and}} hewn
down:
Sharon is like a
wilderness; and
Bashan and
Carmel shake
off
{\ADD{their fruits}}.\FTNT{‡}{{\FR 33.9 }
hewn…: or, withered away}
\VS{10}Now will I
rise,
saith the
{\ND{
LORD}}; now will I be
exalted; now will I lift
up myself.
\VS{11}Ye shall
conceive
chaff, ye shall bring
forth
stubble: your
breath,
{\ADD{as}}
fire, shall
devour you.
\VS{12}And the
people shall be
{\ADD{as}} the
burnings of
lime:
{\ADD{as}}
thorns cut
up shall they be
burned in the
fire.
\par }{\PP \VS{13}¶
Hear, ye
{\ADD{that are}} far
off, what I have
done; and, ye
{\ADD{that are}}
near,
acknowledge my
might.
\VS{14}The
sinners in
Zion are
afraid;
fearfulness hath
surprised the
hypocrites. Who among us shall
dwell with the
devouring
fire? who among us shall
dwell with
everlasting
burnings?
\VS{15}He that
walketh
righteously, and
speaketh
uprightly; he that
despiseth the
gain of
oppressions, that
shaketh his
hands from
holding of
bribes, that
stoppeth his
ears from
hearing of
blood, and
shutteth his
eyes from
seeing
evil;\FTNT{§}{{\FR 33.15 }
righteously: Heb. in righteousnesses}\FTNT{**}{{\FR 33.15 }
uprightly: Heb. uprightnesses}\FTNT{††}{{\FR 33.15 }
oppressions: or, deceits}\FTNT{‡‡}{{\FR 33.15 }
blood: Heb. bloods}
\VS{16}He shall
dwell on
high: his place of
defence
{\ADD{shall be}} the
munitions of
rocks:
bread shall be
given him; his
waters
{\ADD{shall be}}
sure.\FTNT{§§}{{\FR 33.16 }
high: Heb. heights, or, high places}
\VS{17}Thine
eyes shall
see the
king in his
beauty: they shall
behold the
land that is very far
off.\FTNT{***}{{\FR 33.17 }
that…: Heb. of far distances}
\VS{18}Thine
heart shall
meditate
terror. Where
{\ADD{is}} the
scribe? where
{\ADD{is}} the
receiver? where
{\ADD{is}} he that
counted the
towers?\FTNT{†††}{{\FR 33.18 }
receiver: Heb. weigher?}
\VS{19}Thou shalt not
see a
fierce
people, a
people of a
deeper
speech than thou canst
perceive; of a
stammering
tongue,
{\ADD{that thou canst}} not
understand.\FTNT{‡‡‡}{{\FR 33.19 }
stammering: or, ridiculous}
\VS{20}Look upon
Zion, the
city of our
solemnities: thine
eyes shall
see
Jerusalem a
quiet
habitation, a
tabernacle
{\ADD{that}} shall not be taken
down; not one of the
stakes thereof shall
ever be
removed, neither shall any of the
cords thereof be
broken.
\VS{21}But there the
glorious
{\ND{
LORD}}
{\ADD{will be}} unto us a
place of
broad
rivers
{\ADD{and}}
streams; wherein shall
go
no
galley with
oars, neither shall
gallant
ship
pass thereby.\FTNT{§§§}{{\FR 33.21 }
of…: Heb. broad of spaces, or, hands}
\VS{22}For the
{\ND{
LORD}}
{\ADD{is}} our
judge, the
{\ND{
LORD}}
{\ADD{is}} our
lawgiver, the
{\ND{
LORD}}
{\ADD{is}} our
king; he will
save us.\FTNT{*}{{\FR 33.22 }
lawgiver: Heb. statutemaker}
\VS{23}Thy
tacklings are
loosed; they could not
well
strengthen their
mast, they could not
spread the
sail: then is the
prey of a
great
spoil
divided; the
lame
take the
prey.\FTNT{†}{{\FR 33.23 }
Thy…: or, They have forsaken thy tacklings}
\VS{24}And the
inhabitant shall not
say, I am
sick: the
people that
dwell therein
{\ADD{shall be}}
forgiven
{\ADD{their}}
iniquity.

\par }\Chap{34}{\PP \VerseOne{1}Come
near, ye
nations, to
hear; and
hearken, ye
people: let the
earth
hear, and all that is
therein; the
world, and all things that come
forth of it.\FTNT{*}{{\FR 34.1 }
all that…: Heb. the fulness thereof}
\VS{2}For the
indignation of the
{\ND{
LORD}}
{\ADD{is}} upon all
nations, and
{\ADD{his}}
fury upon all their
armies: he hath utterly
destroyed them, he hath
delivered them to the
slaughter.
\VS{3}Their
slain also shall be cast
out, and their
stink shall come
up out of their
carcases, and the
mountains shall be
melted with their
blood.
\VS{4}And all the
host of
heaven shall be
dissolved, and the
heavens shall be rolled
together as a
scroll: and all their
host shall fall
down, as the
leaf falleth
off from the
vine, and as a
falling
{\ADD{fig}} from the fig
tree.
\VS{5}For my
sword shall be
bathed in
heaven: behold, it shall come
down upon
Idumea, and upon the
people of my
curse, to
judgment.
\VS{6}The
sword of the
{\ND{
LORD}} is
filled with
blood, it is made
fat with
fatness,
{\ADD{and}} with the
blood of
lambs and
goats, with the
fat of the
kidneys of
rams: for the
{\ND{
LORD}} hath a
sacrifice in
Bozrah, and a
great
slaughter in the
land of
Idumea.
\VS{7}And the
unicorns shall come
down with them, and the
bullocks with the
bulls; and their
land shall be
soaked with
blood, and their
dust made
fat with
fatness.\FTNT{†}{{\FR 34.7 }
unicorns: or, rhinocerots}\FTNT{‡}{{\FR 34.7 }
soaked: or, drunken}
\VS{8}For
{\ADD{it is}} the
day of the
{\ND{
LORD’s}}
vengeance,
{\ADD{and}} the
year of
recompences for the
controversy of
Zion.
\par }{\PP \VS{9}And the
streams thereof shall be
turned into
pitch, and the
dust thereof into
brimstone, and the
land thereof shall become
burning
pitch.
\VS{10}It shall not be
quenched
night nor
day; the
smoke thereof shall go
up for
ever: from
generation to
generation it shall lie
waste; none shall pass
through it for
ever and
ever.
\par }{\PP \VS{11}¶ But the
cormorant and the
bittern shall
possess it; the
owl also and the
raven shall
dwell in it: and he shall stretch
out upon it the
line of
confusion, and the
stones of
emptiness.\FTNT{§}{{\FR 34.11 }
cormorant: or, pelican}
\VS{12}They shall
call the
nobles thereof to the
kingdom, but none
{\ADD{shall be}} there, and all her
princes shall be
nothing.
\VS{13}And
thorns shall come
up in her
palaces,
nettles and
brambles in the
fortresses thereof: and it shall be an
habitation of
dragons,
{\ADD{and}} a
court for
owls.\FTNT{**}{{\FR 34.13 }
owls: or, ostriches: Heb. daughters of the owl}
\VS{14}The wild beasts of the
desert shall also
meet with the wild beasts of the
island, and the
satyr shall
cry to his
fellow; the screech
owl also shall
rest there, and
find for herself a place of
rest.\FTNT{††}{{\FR 34.14 }
The wild beasts of the desert: Heb. Ziim}\FTNT{‡‡}{{\FR 34.14 }
the wild beasts of the island: Heb. Ijim}\FTNT{§§}{{\FR 34.14 }
screech…: or, night monster}
\VS{15}There shall the great
owl make her
nest, and
lay, and
hatch, and
gather under her
shadow: there shall the
vultures also be
gathered, every
one with her
mate.
\par }{\PP \VS{16}¶ Seek ye
out of the
book of the
{\ND{
LORD}}, and
read: no
one of
these shall
fail,
none shall
want her
mate: for my
mouth it hath
commanded, and his
spirit it hath
gathered them.
\VS{17}And he hath
cast the
lot for them, and his
hand hath
divided it unto them by
line: they shall
possess it
for
ever, from
generation to
generation shall they
dwell therein.

\par }\Chap{35}{\PP \VerseOne{1}The
wilderness and the solitary
place shall be
glad for them; and the
desert shall
rejoice, and
blossom as the
rose.
\VS{2}It shall
blossom
abundantly, and
rejoice even with
joy and
singing: the
glory of
Lebanon shall be
given unto it, the
excellency of
Carmel and
Sharon, they shall
see the
glory of the
{\ND{
LORD}},
{\ADD{and}} the
excellency of our
God.
\par }{\PP \VS{3}¶
Strengthen ye the
weak
hands, and
confirm the
feeble
knees.
\VS{4}Say to them
{\ADD{that are}} of a
fearful
heart, Be
strong,
fear not: behold, your
God will
come
{\ADD{with}}
vengeance,
{\ADD{even}}
God
{\ADD{with}} a
recompence; he will
come and
save you.\FTNT{*}{{\FR 35.4 }
fearful: Heb. hasty}
\par }{\PP \VS{5}Then the
eyes of the
blind shall be
opened, and the
ears of the
deaf shall be
unstopped.
\VS{6}Then shall the
lame
{\ADD{man}}
leap as an
hart, and the
tongue of the
dumb
sing: for in the
wilderness shall
waters break
out, and
streams in the
desert.
\VS{7}And the parched
ground shall become a
pool, and the thirsty
land
springs of
water: in the
habitation of
dragons, where each
lay,
{\ADD{shall be}}
grass with
reeds and
rushes.\FTNT{†}{{\FR 35.7 }
grass…: or, a court for reeds, etc}
\VS{8}And an
highway shall be there, and a
way, and it shall be
called The
way of
holiness; the
unclean shall not pass
over it; but it
{\ADD{shall be}} for those: the wayfaring
men, though
fools, shall not
err
{\ADD{therein}}.\FTNT{‡}{{\FR 35.8 }
but…: or, for he shall be with them}
\VS{9}No
lion shall be there, nor
{\ADD{any}}
ravenous
beast shall go
up thereon, it shall not be
found there; but the
redeemed shall
walk
{\ADD{there:}}
\VS{10}And the
ransomed of the
{\ND{
LORD}} shall
return, and
come to
Zion with
songs and
everlasting
joy upon their
heads: they shall
obtain
joy and
gladness, and
sorrow and
sighing shall flee
away.

\par }\Chap{36}{\PP \VerseOne{1}Now it came to pass in the
fourteenth
year of
king
Hezekiah,
{\ADD{that}}
Sennacherib
king of
Assyria came
up against all the
defenced
cities of
Judah, and
took them.
\VS{2}And the
king of
Assyria
sent
Rabshakeh from
Lachish to
Jerusalem unto
king
Hezekiah with a
great
army. And he
stood by the
conduit of the
upper
pool in the
highway of the
fuller’s
field.
\VS{3}Then came
forth unto him
Eliakim,
Hilkiah’s
son, which was over the
house, and
Shebna the
scribe, and
Joah,
Asaph’s
son, the
recorder.\FTNT{*}{{\FR 36.3 }
scribe: or, secretary}
\par }{\PP \VS{4}¶ And
Rabshakeh
said unto them,
Say ye now to
Hezekiah, Thus
saith the
great
king, the
king of
Assyria, What
confidence
{\ADD{is}} this wherein thou
trustest?
\VS{5}I
say,
{\ADD{sayest thou}}, (but
{\ADD{they are but}}
vain
words)
{\ADD{I have}}
counsel and
strength for
war: now on whom dost thou
trust, that thou
rebellest against me?\FTNT{†}{{\FR 36.5 }
vain…: Heb. a word of lips}\FTNT{‡}{{\FR 36.5 }
I have…: or, but counsel and strength are for the war}
\VS{6}Lo, thou
trustest in the
staff of this
broken
reed, on
Egypt; whereon if a
man
lean, it will
go into his
hand, and
pierce it: so
{\ADD{is}}
Pharaoh
king of
Egypt to all that
trust in him.
\VS{7}But if thou
say to me, We
trust in the
{\ND{
LORD}} our
God:
{\ADD{is it}} not he, whose high
places and whose
altars
Hezekiah hath taken
away, and
said to
Judah and to
Jerusalem, Ye shall
worship
before this
altar?
\VS{8}Now therefore give
pledges, I pray thee, to my
master the
king of
Assyria, and I will
give thee two
thousand
horses, if thou be
able on thy part to
set
riders upon them.\FTNT{§}{{\FR 36.8 }
pledges: or, hostages}
\VS{9}How then wilt thou turn
away the
face of
one
captain of the
least of my
master’s
servants, and put thy
trust on
Egypt for
chariots and for
horsemen?
\VS{10}And am I now come
up
without the
{\ND{
LORD}} against this
land to
destroy it? the
{\ND{
LORD}}
said unto me, Go
up against this
land, and
destroy it.
\par }{\PP \VS{11}¶ Then
said
Eliakim and
Shebna and
Joah unto
Rabshakeh,
Speak, I pray thee, unto thy
servants in the Syrian
language; for we
understand
{\ADD{it:}} and
speak not to us in the
Jews’ language, in the
ears of the
people that
{\ADD{are}} on the
wall.
\par }{\PP \VS{12}¶ But
Rabshakeh
said, Hath my
master
sent me to thy
master and to thee to
speak these
words?
{\ADD{hath he}} not
{\ADD{sent me}} to the
men that
sit upon the
wall, that they may
eat their own
dung, and
drink their own
piss with you?
\VS{13}Then
Rabshakeh
stood, and
cried with a
loud
voice in the
Jews’ language, and
said,
Hear ye the
words of the
great
king, the
king of
Assyria.
\VS{14}Thus
saith the
king, Let not
Hezekiah
deceive you: for he shall not be
able to
deliver you.
\VS{15}Neither let
Hezekiah make you
trust in the
{\ND{
LORD}},
saying, The
{\ND{
LORD}} will
surely
deliver us: this
city shall not be
delivered into the
hand of the
king of
Assyria.
\VS{16}Hearken not to
Hezekiah: for thus
saith the
king of
Assyria,
Make
{\ADD{an agreement}} with me
{\ADD{by}} a
present, and come
out to me: and
eat ye every
one of his
vine, and every
one of his fig
tree, and
drink ye every
one the
waters of his own
cistern;\FTNT{**}{{\FR 36.16 }
Make…: or, Seek my favour by a present: Heb. Make with me a blessing}
\VS{17}Until I
come and take you
away to a
land like your own
land, a
land of
corn and
wine, a
land of
bread and
vineyards.
\VS{18}{\ADD{Beware}} lest
Hezekiah
persuade you,
saying, The
{\ND{
LORD}} will
deliver us. Hath
any of the
gods of the
nations
delivered his
land out of the
hand of the
king of
Assyria?
\VS{19}Where
{\ADD{are}} the
gods of
Hamath and
Arphad? where
{\ADD{are}} the
gods of
Sepharvaim? and have they
delivered
Samaria out of my
hand?
\VS{20}Who
{\ADD{are they}} among all the
gods of these
lands, that have
delivered their
land out of my
hand, that the
{\ND{
LORD}} should
deliver
Jerusalem out of my
hand?
\VS{21}But they held their
peace, and
answered him not a
word: for the
king’s
commandment was,
saying,
Answer him not.
\par }{\PP \VS{22}¶ Then
came
Eliakim, the
son of
Hilkiah, that
{\ADD{was}} over the
household, and
Shebna the
scribe, and
Joah, the
son of
Asaph, the
recorder, to
Hezekiah with
{\ADD{their}}
clothes
rent, and
told him the
words of
Rabshakeh.

\par }\Chap{37}{\PP \VerseOne{1}And it came to pass, when
king
Hezekiah
heard
{\ADD{it}}, that he
rent his
clothes, and
covered himself with
sackcloth, and
went into the
house of the
{\ND{
LORD}}.
\VS{2}And he
sent
Eliakim, who
{\ADD{was}} over the
household, and
Shebna the
scribe, and the
elders of the
priests
covered with
sackcloth, unto
Isaiah the
prophet the
son of
Amoz.
\VS{3}And they
said unto him, Thus
saith
Hezekiah, This
day
{\ADD{is}} a
day of
trouble, and of
rebuke, and of
blasphemy: for the
children are
come to the
birth, and
{\ADD{there is}} not
strength to bring
forth.\FTNT{*}{{\FR 37.3 }
blasphemy: or, provocation}
\VS{4}It may be the
{\ND{
LORD}} thy
God will
hear the
words of
Rabshakeh, whom the
king of
Assyria his
master hath
sent to
reproach the
living
God, and will
reprove the
words which the
{\ND{
LORD}} thy
God hath
heard: wherefore lift
up
{\ADD{thy}}
prayer for the
remnant that is
left.\FTNT{†}{{\FR 37.4 }
left: Heb. found}
\VS{5}So the
servants of
king
Hezekiah
came to
Isaiah.
\par }{\PP \VS{6}¶ And
Isaiah
said unto them, Thus shall ye
say unto your
master, Thus
saith the
{\ND{
LORD}}, Be not
afraid
of the
words that thou hast
heard, wherewith the
servants of the
king of
Assyria have
blasphemed me.
\VS{7}Behold, I will
send a
blast upon him, and he shall
hear a
rumour, and
return to his own
land; and I will cause him to
fall by the
sword in his own
land.\FTNT{‡}{{\FR 37.7 }
send…: or, put a spirit into him}
\par }{\PP \VS{8}¶ So
Rabshakeh
returned, and
found the
king of
Assyria
warring against
Libnah: for he had
heard that he was
departed from
Lachish.
\VS{9}And he
heard
say concerning
Tirhakah
king of
Ethiopia, He is come
forth to make
war with thee. And when he
heard
{\ADD{it}}, he
sent
messengers to
Hezekiah,
saying,
\VS{10}Thus shall ye
speak to
Hezekiah
king of
Judah,
saying, Let not thy
God, in whom thou
trustest,
deceive thee,
saying,
Jerusalem shall not be
given into the
hand of the
king of
Assyria.
\VS{11}Behold, thou hast
heard what the
kings of
Assyria have
done to all
lands by destroying them
utterly; and shalt thou be
delivered?
\VS{12}Have the
gods of the
nations
delivered them which my
fathers have
destroyed,
{\ADD{as}}
Gozan, and
Haran, and
Rezeph, and the
children of
Eden which
{\ADD{were}} in
Telassar?
\VS{13}Where
{\ADD{is}} the
king of
Hamath, and the
king of
Arphad, and the
king of the
city of
Sepharvaim,
Hena, and
Ivah?
\par }{\PP \VS{14}¶ And
Hezekiah
received the
letter from the
hand of the
messengers, and
read it: and
Hezekiah went
up unto the
house of the
{\ND{
LORD}}, and
spread it
before the
{\ND{
LORD}}.
\VS{15}And
Hezekiah
prayed unto the
{\ND{
LORD}},
saying,
\VS{16}O
{\ND{
LORD}} of
hosts,
God of
Israel, that
dwellest
{\ADD{between}} the
cherubims, thou
{\ADD{art}} the
God,
{\ADD{even}} thou alone, of all the
kingdoms of the
earth: thou hast
made
heaven and
earth.
\VS{17}Incline thine
ear, O
{\ND{
LORD}}, and
hear;
open thine
eyes, O
{\ND{
LORD}}, and
see: and
hear all the
words of
Sennacherib, which hath
sent to
reproach the
living
God.
\VS{18}Of a
truth,
{\ND{
LORD}}, the
kings of
Assyria have laid
waste all the
nations, and their
countries,\FTNT{§}{{\FR 37.18 }
nations: Heb. lands}
\VS{19}And have
cast their
gods into the
fire: for they
{\ADD{were}} no
gods, but the
work of
men’s
hands,
wood and
stone: therefore they have
destroyed them.\FTNT{**}{{\FR 37.19 }
cast: Heb. given}
\VS{20}Now therefore, O
{\ND{
LORD}} our
God,
save us from his
hand, that all the
kingdoms of the
earth may
know that thou
{\ADD{art}} the
{\ND{
LORD}},
{\ADD{even}} thou only.
\par }{\PP \VS{21}¶ Then
Isaiah the
son of
Amoz
sent unto
Hezekiah,
saying, Thus
saith the
{\ND{
LORD}}
God of
Israel, Whereas thou hast
prayed to me against
Sennacherib
king of
Assyria:
\VS{22}This
{\ADD{is}} the
word which the
{\ND{
LORD}} hath
spoken concerning him; The
virgin, the
daughter of
Zion, hath
despised thee,
{\ADD{and}} laughed thee to
scorn; the
daughter of
Jerusalem hath
shaken her
head at
thee.
\VS{23}Whom hast thou
reproached and
blasphemed? and against whom hast thou
exalted
{\ADD{thy}}
voice, and lifted
up thine
eyes on
high?
{\ADD{even}} against the Holy
One of
Israel.
\VS{24}By thy
servants hast thou
reproached the
Lord, and hast
said, By the
multitude of my
chariots am I come
up to the
height of the
mountains, to the
sides of
Lebanon; and I will cut
down the
tall
cedars thereof,
{\ADD{and}} the
choice fir
trees thereof: and I will
enter into the
height of his
border,
{\ADD{and}} the
forest of his
Carmel.\FTNT{††}{{\FR 37.24 }
By thy…: Heb. By the hand of thy servants}\FTNT{‡‡}{{\FR 37.24 }
the tall…: Heb. the tallness of the cedars thereof and the choice of the fir trees thereof}\FTNT{§§}{{\FR 37.24 }
the forest…: or, the forest and his fruitful field}
\VS{25}I have
digged, and
drunk
water; and with the
sole of my
feet have I dried
up all the
rivers of the besieged
places.\FTNT{***}{{\FR 37.25 }
besieged: or, fenced and closed}
\VS{26}Hast thou not
heard long
ago,
{\ADD{how}} I have
done it;
{\ADD{and}} of
ancient
times, that I have
formed it? now have I
brought it to pass, that thou shouldest be to lay
waste
defenced
cities
{\ADD{into}}
ruinous
heaps.\FTNT{†††}{{\FR 37.26 }
Hast…: or, Hast thou not heard how I have made it long ago, and formed it of ancient times? should I now bring it to be laid waste, and defenced cities to be ruinous heaps?}
\VS{27}Therefore their
inhabitants
{\ADD{were}} of
small
power, they were
dismayed and
confounded: they were
{\ADD{as}} the
grass of the
field, and
{\ADD{as}} the
green
herb,
{\ADD{as}} the
grass on the
housetops, and
{\ADD{as corn}}
blasted
before it be grown
up.\FTNT{‡‡‡}{{\FR 37.27 }
of small…: Heb. short of hand}
\VS{28}But I
know thy
abode, and thy going
out, and thy coming
in, and thy
rage against me.\FTNT{§§§}{{\FR 37.28 }
abode: or, sitting}
\VS{29}Because thy
rage against me, and thy
tumult, is come
up into mine
ears, therefore will I
put my
hook in thy
nose, and my
bridle in thy
lips, and I will turn thee
back by the
way by which thou
camest.
\VS{30}And this
{\ADD{shall be}} a
sign unto thee, Ye shall
eat
{\ADD{this}}
year such as
groweth of itself; and the
second
year that which
springeth of the same: and in the
third
year
sow ye, and
reap, and
plant
vineyards, and
eat the
fruit thereof.
\VS{31}And the
remnant that is
escaped of the
house of
Judah shall
again take
root
downward, and
bear
fruit
upward:\FTNT{*}{{\FR 37.31 }
the remnant…: Heb. the escaping of the house of Judah that remaineth}
\VS{32}For out of
Jerusalem shall go
forth a
remnant, and they that
escape out of
mount
Zion: the
zeal of the
{\ND{
LORD}} of
hosts shall
do this.\FTNT{†}{{\FR 37.32 }
they…: Heb. the escaping}
\VS{33}Therefore thus
saith the
{\ND{
LORD}} concerning the
king of
Assyria, He shall not
come into this
city, nor
shoot an
arrow there, nor come
before it with
shields, nor
cast a
bank against it.
\VS{34}By the
way that he
came, by the same shall he
return, and shall not
come into this
city,
saith the
{\ND{
LORD}}.
\VS{35}For I will
defend this
city to
save it for mine own sake, and for my
servant
David’s sake.
\VS{36}Then the
angel of the
{\ND{
LORD}} went
forth, and
smote in the
camp of the
Assyrians a
hundred and
fourscore and
five
thousand: and when they arose
early in the
morning, behold, they
{\ADD{were}} all
dead
corpses.
\par }{\PP \VS{37}¶ So
Sennacherib
king of
Assyria
departed, and
went and
returned, and
dwelt at
Nineveh.
\VS{38}And it came to pass, as he was
worshipping in the
house of
Nisroch his
god, that
Adrammelech and
Sharezer his
sons
smote him with the
sword; and they
escaped into the
land of
Armenia: and
Esar-haddon his
son
reigned in his stead.\FTNT{‡}{{\FR 37.38 }
Armenia: Heb. Ararat}

\par }\Chap{38}{\PP \VerseOne{1}In those
days was
Hezekiah
sick unto
death. And
Isaiah the
prophet the
son of
Amoz
came unto him, and
said unto him, Thus
saith the
{\ND{
LORD}}, Set thine
house in
order: for thou shalt
die, and not
live.\FTNT{*}{{\FR 38.1 }
Set…: Heb. Give charge concerning thy house}
\VS{2}Then
Hezekiah
turned his
face toward the
wall, and
prayed unto the
{\ND{
LORD}},
\VS{3}And
said,
Remember now, O
{\ND{
LORD}}, I
beseech thee, how I have
walked
before thee in
truth and with a
perfect
heart, and have
done
{\ADD{that which is}}
good in thy
sight. And
Hezekiah
wept
sore.\FTNT{†}{{\FR 38.3 }
sore: Heb. with great weeping}
\par }{\PP \VS{4}¶ Then came the
word of the
{\ND{
LORD}} to
Isaiah,
saying,
\VS{5}Go, and
say to
Hezekiah, Thus
saith the
{\ND{
LORD}}, the
God of
David thy
father, I have
heard thy
prayer, I have
seen thy
tears: behold, I will
add unto thy
days
fifteen
years.
\VS{6}And I will
deliver thee and this
city out of the
hand of the
king of
Assyria: and I will
defend this
city.
\VS{7}And this
{\ADD{shall be}} a
sign unto thee from the
{\ND{
LORD}}, that the
{\ND{
LORD}} will
do this
thing that he hath
spoken;
\VS{8}Behold, I will bring
again the
shadow of the
degrees, which is gone
down in the
sun
dial of
Ahaz,
ten
degrees
backward. So the
sun
returned
ten
degrees, by which
degrees it was gone
down.\FTNT{‡}{{\FR 38.8 }
sun dial: Heb. degrees by, or, with the sun}
\par }{\PP \VS{9}¶ The
writing of
Hezekiah
king of
Judah, when he had been
sick, and was
recovered of his
sickness:
\VS{10}I
said in the cutting
off of my
days, I shall
go to the
gates of the
grave: I am
deprived of the
residue of my
years.
\VS{11}I
said, I shall not
see the
{\ND{
LORD}},
{\ADD{even}} the
{\ND{
LORD}}, in the
land of the
living: I shall
behold
man no more with the
inhabitants of the
world.
\VS{12}Mine
age is
departed, and is
removed from me as a
shepherd’s
tent: I have cut
off like a
weaver my
life: he will cut me
off with pining
sickness: from
day
{\ADD{even}} to
night wilt thou make an
end of me.\FTNT{§}{{\FR 38.12 }
with…: or, from the thrum}
\VS{13}I
reckoned till
morning,
{\ADD{that}}, as a
lion, so will he
break all my
bones: from
day
{\ADD{even}} to
night wilt thou make an
end of me.
\VS{14}Like a
crane
{\ADD{or}} a
swallow, so did I
chatter: I did
mourn as a
dove: mine
eyes
fail
{\ADD{with looking}}
upward: O
{\ND{
LORD}}, I am
oppressed;
undertake for me.\FTNT{**}{{\FR 38.14 }
undertake…: or, ease me}
\VS{15}What shall I
say? he hath both
spoken unto me, and himself hath
done
{\ADD{it:}} I shall go
softly all my
years in the
bitterness of my
soul.
\VS{16}O
Lord, by these
{\ADD{things men}}
live, and in all these
{\ADD{things is}} the
life of my
spirit: so wilt thou
recover me, and make me to
live.
\VS{17}Behold, for
peace I had great
bitterness: but thou hast in
love to my
soul
{\ADD{delivered it}} from the
pit of
corruption: for thou hast
cast all my
sins
behind thy
back.\FTNT{††}{{\FR 38.17 }
for peace…: or, on my peace came great bitterness}\FTNT{‡‡}{{\FR 38.17 }
thou hast in…: Heb. thou hast loved my soul from the pit}
\VS{18}For the
grave cannot
praise thee,
death can
{\ADD{not}}
celebrate thee: they that go
down into the
pit cannot
hope for thy
truth.
\VS{19}The
living, the
living, he shall
praise thee, as I
{\ADD{do}} this
day: the
father to the
children shall make
known thy
truth.
\VS{20}The
{\ND{
LORD}}
{\ADD{was ready}} to
save me: therefore we will sing my
songs to the stringed
instruments all the
days of our
life in the
house of the
{\ND{
LORD}}.
\VS{21}For
Isaiah had
said, Let them
take a
lump of
figs, and lay
{\ADD{it}} for a
plaister upon the
boil, and he shall
recover.
\VS{22}Hezekiah also had
said, What
{\ADD{is}} the
sign that I shall go
up to the
house of the
{\ND{
LORD}}?

\par }\Chap{39}{\PP \VerseOne{1}At that
time
Merodach-baladan, the
son of
Baladan,
king of
Babylon,
sent
letters and a
present to
Hezekiah: for he had
heard that he had been
sick, and was
recovered.
\VS{2}And
Hezekiah was
glad of them, and
shewed them the
house of his precious
things, the
silver, and the
gold, and the
spices, and the
precious
ointment, and all the
house of his
armour, and all that was
found in his
treasures: there was
nothing in his
house, nor in all his
dominion, that
Hezekiah
shewed them not.\FTNT{*}{{\FR 39.2 }
precious things: or, spicery}\FTNT{†}{{\FR 39.2 }
armour: or, jewels: Heb. vessels, or, instruments}
\par }{\PP \VS{3}¶ Then
came
Isaiah the
prophet unto
king
Hezekiah, and
said unto him, What
said these
men? and from
whence
came they unto thee? And
Hezekiah
said, They are
come from a
far
country unto me,
{\ADD{even}} from
Babylon.
\VS{4}Then
said he, What have they
seen in thine
house? And
Hezekiah
answered, All that
{\ADD{is}} in mine
house have they
seen: there is
nothing among my
treasures that I have not
shewed them.
\par }{\PP \VS{5}Then
said
Isaiah to
Hezekiah,
Hear the
word of the
{\ND{
LORD}} of
hosts:
\VS{6}Behold, the
days
come, that all that
{\ADD{is}} in thine
house, and
{\ADD{that}} which thy
fathers have laid up in
store until this
day, shall be
carried to
Babylon:
nothing shall be
left,
saith the
{\ND{
LORD}}.
\VS{7}And of thy
sons that shall
issue from thee, which thou shalt
beget, shall they take
away; and they shall be
eunuchs in the
palace of the
king of
Babylon.
\VS{8}Then
said
Hezekiah to
Isaiah,
Good
{\ADD{is}} the
word of the
{\ND{
LORD}} which thou hast
spoken. He
said moreover, For there shall be
peace and
truth in my
days.

\par }\Chap{40}{\PP \VerseOne{1}Comfort ye,
comfort ye my
people,
saith your
God.
\VS{2}Speak ye
comfortably to
Jerusalem, and
cry unto her, that her
warfare is
accomplished, that her
iniquity is
pardoned: for she hath
received of the
{\ND{
LORD’s}}
hand
double for all her
sins.\FTNT{*}{{\FR 40.2 }
comfortably: Heb. to the heart}\FTNT{†}{{\FR 40.2 }
warfare: or, appointed time}
\par }{\PP \VS{3}¶ The
voice of him that
crieth in the
wilderness,
Prepare ye the
way of the
{\ND{
LORD}}, make
straight in the
desert a
highway for our
God.
\VS{4}Every
valley shall be
exalted, and every
mountain and
hill shall be made
low: and the
crooked shall be made
straight, and the rough
places
plain:\FTNT{‡}{{\FR 40.4 }
straight: or, a straight place}\FTNT{§}{{\FR 40.4 }
plain: or, a plain place}
\VS{5}And the
glory of the
{\ND{
LORD}} shall be
revealed, and all
flesh shall
see
{\ADD{it}}
together: for the
mouth of the
{\ND{
LORD}} hath
spoken
{\ADD{it}}.
\VS{6}The
voice
said,
Cry. And he
said, What shall I
cry? All
flesh
{\ADD{is}}
grass, and all the
goodliness thereof
{\ADD{is}} as the
flower of the
field:
\VS{7}The
grass
withereth, the
flower
fadeth: because the
spirit of the
{\ND{
LORD}}
bloweth upon it:
surely the
people
{\ADD{is}}
grass.
\VS{8}The
grass
withereth, the
flower
fadeth: but the
word of our
God shall
stand for
ever.
\par }{\PP \VS{9}¶ O
Zion, that bringest good
tidings, get thee
up into the
high
mountain; O
Jerusalem, that bringest good
tidings, lift
up thy
voice with
strength; lift
{\ADD{it}}
up, be not
afraid;
say unto the
cities of
Judah, Behold your
God!\FTNT{**}{{\FR 40.9 }
O Zion…: or, O thou that tellest good tidings to Zion}\FTNT{††}{{\FR 40.9 }
O Jerusalem…: or, O thou that tellest good tidings to Jerusalem}
\VS{10}Behold, the
Lord
{\ND{
GOD}} will
come with
strong
{\ADD{hand}}, and his
arm shall
rule for him: behold, his
reward
{\ADD{is}} with him, and his
work
before him.\FTNT{‡‡}{{\FR 40.10 }
with strong…: or, against the strong}\FTNT{§§}{{\FR 40.10 }
his work: or, recompence for his work}
\VS{11}He shall
feed his
flock like a
shepherd: he shall
gather the
lambs with his
arm, and
carry
{\ADD{them}} in his
bosom,
{\ADD{and}} shall gently
lead those that are with
young.\FTNT{***}{{\FR 40.11 }
that…: or, that give suck}
\par }{\PP \VS{12}¶ Who hath
measured the
waters in the hollow of his
hand, and meted
out
heaven with the
span, and
comprehended the
dust of the
earth in a
measure, and
weighed the
mountains in
scales, and the
hills in a
balance?\FTNT{†††}{{\FR 40.12 }
a measure: Heb. a tierce}
\VS{13}Who hath
directed the
Spirit of the
{\ND{
LORD}}, or
{\ADD{being}} his
counsellor hath
taught him?\FTNT{‡‡‡}{{\FR 40.13 }
his…: Heb. man of his counsel}
\VS{14}With whom took he
counsel, and
{\ADD{who}}
instructed him, and
taught him in the
path of
judgment, and
taught him
knowledge, and
shewed to him the
way of
understanding?\FTNT{§§§}{{\FR 40.14 }
instructed…: Heb. made him understand}\FTNT{*}{{\FR 40.14 }
understanding: Heb. understandings?}
\VS{15}Behold, the
nations
{\ADD{are}} as a
drop of a
bucket, and are
counted as the small
dust of the
balance: behold, he taketh
up the
isles as a very little
thing.
\VS{16}And
Lebanon
{\ADD{is}}
not
sufficient to
burn, nor the
beasts thereof
sufficient for a burnt
offering.
\VS{17}All
nations before him
{\ADD{are}} as nothing; and they are
counted to him less than
nothing, and
vanity.
\par }{\PP \VS{18}¶ To whom then will ye
liken
God? or what
likeness will ye
compare unto him?
\VS{19}The
workman
melteth a graven
image, and the
goldsmith
spreadeth it over with
gold, and
casteth
silver
chains.
\VS{20}He that
{\ADD{is}} so
impoverished that he hath no
oblation
chooseth a
tree
{\ADD{that}} will not
rot; he
seeketh unto him a
cunning
workman to
prepare a graven
image,
{\ADD{that}} shall not be
moved.\FTNT{†}{{\FR 40.20 }
is so…: Heb. is poor of oblation}
\VS{21}Have ye not
known? have ye not
heard? hath it not been
told you from the
beginning? have ye not
understood from the
foundations of the
earth?
\VS{22}{\ADD{It is}} he that
sitteth upon the
circle of the
earth, and the
inhabitants thereof
{\ADD{are}} as
grasshoppers; that stretcheth
out the
heavens as a
curtain, and spreadeth them
out as a
tent to dwell
in:\FTNT{‡}{{\FR 40.22 }
It is…: or, Him that}
\VS{23}That
bringeth the
princes to nothing; he
maketh the
judges of the
earth as
vanity.
\VS{24}Yea, they shall not be
planted; yea, they shall not be
sown: yea, their
stock shall not take
root in the
earth: and he shall also
blow upon them, and they shall
wither, and the
whirlwind shall
take them away as
stubble.
\VS{25}To whom then will ye
liken me, or shall I be
equal?
saith the Holy
One.
\VS{26}Lift
up your
eyes on
high, and
behold who hath
created these
{\ADD{things}}, that bringeth
out their
host by
number: he
calleth them all by
names by the
greatness of his
might, for that
{\ADD{he is}}
strong in
power; not
one
faileth.
\par }{\PP \VS{27}Why
sayest thou, O
Jacob, and
speakest, O
Israel, My
way is
hid from the
{\ND{
LORD}}, and my
judgment is passed
over from my
God?
\par }{\PP \VS{28}¶ Hast thou not
known? hast thou not
heard,
{\ADD{that}} the
everlasting
God, the
{\ND{
LORD}}, the
Creator of the
ends of the
earth,
fainteth not, neither is
weary?
{\ADD{there is}} no
searching of his
understanding.
\VS{29}He
giveth
power to the
faint; and to
{\ADD{them that have}} no
might he
increaseth
strength.
\VS{30}Even the
youths shall
faint and be
weary, and the young
men shall
utterly
fall:
\VS{31}But they that
wait upon the
{\ND{
LORD}} shall
renew
{\ADD{their}}
strength; they shall mount
up with
wings as
eagles; they shall
run, and not be
weary;
{\ADD{and}} they shall
walk, and not
faint.\FTNT{§}{{\FR 40.31 }
renew: Heb. change}

\par }\Chap{41}{\PP \VerseOne{1}Keep
silence before me, O
islands; and let the
people
renew
{\ADD{their}}
strength: let them come
near; then let them
speak: let us come
near
together to
judgment.
\VS{2}Who raised
up the
righteous
{\ADD{man}} from the
east,
called him to his
foot,
gave the
nations
before him, and made
{\ADD{him}}
rule over
kings? he
gave
{\ADD{them}} as the
dust to his
sword,
{\ADD{and}} as
driven
stubble to his
bow.\FTNT{*}{{\FR 41.2 }
the righteous…: Heb. righteousness}
\VS{3}He
pursued them,
{\ADD{and}}
passed
safely;
{\ADD{even}} by the
way
{\ADD{that}} he had not
gone with his
feet.\FTNT{†}{{\FR 41.3 }
safely: Heb. in peace}
\VS{4}Who hath
wrought and
done
{\ADD{it}},
calling the
generations from the
beginning? I the
{\ND{
LORD}}, the
first, and with the
last; I
{\ADD{am}} he.
\VS{5}The
isles
saw
{\ADD{it}}, and
feared; the
ends of the
earth were
afraid, drew
near, and
came.
\VS{6}They
helped every
one his
neighbour; and
{\ADD{every one}}
said to his
brother, Be of good
courage.\FTNT{‡}{{\FR 41.6 }
Be…: Heb. Be strong}
\VS{7}So the
carpenter
encouraged the
goldsmith,
{\ADD{and}} he that
smootheth
{\ADD{with}} the
hammer him that
smote the
anvil,
saying, It
{\ADD{is}}
ready for the
sodering: and he
fastened it with
nails,
{\ADD{that}} it should not be
moved.\FTNT{§}{{\FR 41.7 }
goldsmith: or, founder}\FTNT{**}{{\FR 41.7 }
him…: or, the smiting}\FTNT{††}{{\FR 41.7 }
saying…: or, saying of the soder, It is good}
\VS{8}But thou,
Israel,
{\ADD{art}} my
servant,
Jacob whom I have
chosen, the
seed of
Abraham my
friend.
\VS{9}{\ADD{Thou}} whom I have
taken from the
ends of the
earth, and
called thee from the chief
men thereof, and
said unto thee, Thou
{\ADD{art}} my
servant; I have
chosen thee, and not cast thee
away.
\par }{\PP \VS{10}¶
Fear thou not; for I
{\ADD{am}} with thee: be not
dismayed; for I
{\ADD{am}} thy
God: I will
strengthen thee; yea, I will
help thee; yea, I will
uphold thee with the right
hand of my
righteousness.
\VS{11}Behold, all they that were
incensed against thee shall be
ashamed and
confounded: they shall be as nothing; and
they that
strive with thee shall
perish.\FTNT{‡‡}{{\FR 41.11 }
they that strive…: Heb. the men of thy strife}
\VS{12}Thou shalt
seek them, and shalt not
find them,
{\ADD{even}}
them that
contended with thee: they that
war against thee shall be as nothing, and as a thing of
nought.\FTNT{§§}{{\FR 41.12 }
them that…: Heb. the men of thy contention}\FTNT{***}{{\FR 41.12 }
they…: Heb. the men of thy war}
\VS{13}For I the
{\ND{
LORD}} thy
God will
hold thy right
hand,
saying unto thee,
Fear not; I will
help thee.
\VS{14}Fear not, thou
worm
Jacob,
{\ADD{and}} ye
men of
Israel; I will
help thee,
saith the
{\ND{
LORD}}, and thy
redeemer, the Holy
One of
Israel.\FTNT{†††}{{\FR 41.14 }
men: or, few men}
\VS{15}Behold, I will
make thee a
new
sharp
threshing instrument
having
teeth: thou shalt
thresh the
mountains, and beat
{\ADD{them}}
small, and shalt
make the
hills as
chaff.\FTNT{‡‡‡}{{\FR 41.15 }
teeth: Heb. mouths}
\VS{16}Thou shalt
fan them, and the
wind shall carry them
away, and the
whirlwind shall
scatter them: and thou shalt
rejoice in the
{\ND{
LORD}},
{\ADD{and}} shalt
glory in the Holy
One of
Israel.
\VS{17}{\ADD{When}} the
poor and
needy
seek
water, and
{\ADD{there is}} none,
{\ADD{and}} their
tongue
faileth for
thirst, I the
{\ND{
LORD}} will
hear them, I the
God of
Israel will not
forsake them.
\VS{18}I will
open
rivers in high
places, and
fountains in the
midst of the
valleys: I will
make the
wilderness a
pool of
water, and the
dry
land
springs of
water.
\VS{19}I will
plant in the
wilderness the
cedar, the shittah
tree, and the
myrtle, and the
oil
tree; I will
set in the
desert the fir
tree,
{\ADD{and}} the
pine, and the box
tree
together:
\VS{20}That they may
see, and
know, and
consider, and
understand
together, that the
hand of the
{\ND{
LORD}} hath
done this, and the Holy
One of
Israel hath
created it.
\par }{\PP \VS{21}Produce your
cause,
saith the
{\ND{
LORD}}; bring
forth your
strong
{\ADD{reasons}},
saith the
King of
Jacob.\FTNT{§§§}{{\FR 41.21 }
Produce: Heb. Cause to come near}
\VS{22}Let them bring
{\ADD{them}}
forth, and
shew us what shall
happen: let them
shew the former
things, what they
{\ADD{be}}, that we may
consider them, and
know the latter
end of them; or
declare us things for to
come.\FTNT{*}{{\FR 41.22 }
consider…: Heb. set our heart upon them}
\VS{23}Shew the things that are to
come
hereafter, that we may
know that ye
{\ADD{are}}
gods: yea, do
good, or do
evil, that we may be
dismayed, and
behold
{\ADD{it}}
together.
\VS{24}Behold, ye
{\ADD{are}} of
nothing, and your
work of
nought: an
abomination
{\ADD{is he that}}
chooseth you.\FTNT{†}{{\FR 41.24 }
of nothing: or, worse than nothing}\FTNT{‡}{{\FR 41.24 }
of nought: or, worse than of a viper}
\VS{25}I have raised
up
{\ADD{one}} from the
north, and he shall
come: from the
rising of the
sun shall he
call upon my
name: and he shall
come upon
princes as
{\ADD{upon}}
morter, and as the
potter
treadeth
clay.
\VS{26}Who hath
declared from the
beginning, that we may
know? and
beforetime, that we may
say,
{\ADD{He is}}
righteous? yea,
{\ADD{there is}} none that
sheweth, yea,
{\ADD{there is}} none that
declareth, yea,
{\ADD{there is}} none that
heareth your
words.
\VS{27}The
first
{\ADD{shall say}} to
Zion,
Behold,
behold them: and I will
give to
Jerusalem one that bringeth good
tidings.
\VS{28}For I
beheld, and
{\ADD{there was}} no
man; even among them, and
{\ADD{there was}} no
counsellor, that, when I
asked of them, could
answer a
word.\FTNT{§}{{\FR 41.28 }
answer: Heb. return}
\VS{29}Behold, they
{\ADD{are}} all
vanity; their
works
{\ADD{are}}
nothing: their molten
images
{\ADD{are}}
wind and
confusion.

\par }\Chap{42}{\PP \VerseOne{1}Behold my
servant, whom I
uphold; mine
elect,
{\ADD{in whom}} my
soul
delighteth; I have
put my
spirit upon him: he shall bring
forth
judgment to the
Gentiles.
\VS{2}He shall not
cry, nor lift
up, nor cause his
voice to be
heard in the
street.
\VS{3}A
bruised
reed shall he not
break, and the
smoking
flax shall he not
quench: he shall bring
forth
judgment unto
truth.\FTNT{*}{{\FR 42.3 }
smoking: or, dimly burning}\FTNT{†}{{\FR 42.3 }
quench: Heb. quench it}
\VS{4}He shall not
fail nor be
discouraged, till he have
set
judgment in the
earth: and the
isles shall
wait for his
law.\FTNT{‡}{{\FR 42.4 }
discouraged: Heb. broken}
\par }{\PP \VS{5}¶ Thus
saith
God the
{\ND{
LORD}}, he that
created the
heavens, and stretched them
out; he that spread
forth the
earth, and that which cometh
out of it; he that
giveth
breath unto the
people upon it, and
spirit to them that
walk therein:
\VS{6}I the
{\ND{
LORD}} have
called thee in
righteousness, and will
hold thine
hand, and will
keep thee, and
give thee for a
covenant of the
people, for a
light of the
Gentiles;
\VS{7}To
open the
blind
eyes, to bring
out the
prisoners from the
prison,
{\ADD{and}} them that
sit in
darkness out of the
prison
house.
\VS{8}I
{\ADD{am}} the
{\ND{
LORD}}: that
{\ADD{is}} my
name: and my
glory will I not
give to
another, neither my
praise to graven
images.
\VS{9}Behold, the former
things are come to
pass, and new
things do I
declare: before they spring
forth I
tell you of them.
\VS{10}Sing unto the
{\ND{
LORD}} a
new
song,
{\ADD{and}} his
praise from the
end of the
earth, ye that go
down to the
sea, and all that is
therein; the
isles, and the
inhabitants thereof.\FTNT{§}{{\FR 42.10 }
all…: Heb. the fulness thereof}
\VS{11}Let the
wilderness and the
cities thereof lift
up
{\ADD{their voice}}, the
villages
{\ADD{that}}
Kedar doth
inhabit: let the
inhabitants of the
rock
sing, let them
shout from the
top of the
mountains.
\VS{12}Let them
give
glory unto the
{\ND{
LORD}}, and
declare his
praise in the
islands.
\par }{\PP \VS{13}The
{\ND{
LORD}} shall go
forth as a mighty
man, he shall stir
up
jealousy like a
man of
war: he shall
cry, yea,
roar; he shall
prevail against his
enemies.\FTNT{**}{{\FR 42.13 }
prevail: or, behave himself mightily}
\VS{14}I have long
time holden my
peace; I have been
still,
{\ADD{and}}
refrained myself:
{\ADD{now}} will I
cry like a travailing
woman; I will
destroy and
devour at
once.\FTNT{††}{{\FR 42.14 }
devour: Heb. swallow, or, sup up}
\VS{15}I will make
waste
mountains and
hills, and dry
up all their
herbs; and I will
make the
rivers
islands, and I will dry
up the
pools.
\VS{16}And I will
bring the
blind by a
way
{\ADD{that}} they
knew not; I will
lead them in
paths
{\ADD{that}} they have not
known: I will
make
darkness
light
before them, and crooked
things
straight. These
things will I
do unto them, and not
forsake them.\FTNT{‡‡}{{\FR 42.16 }
straight: Heb. into straightness}
\par }{\PP \VS{17}¶ They shall be
turned
back, they shall be
greatly
ashamed, that
trust in graven
images, that
say to the molten
images, Ye
{\ADD{are}} our
gods.
\par }{\PP \VS{18}Hear, ye
deaf; and
look, ye
blind, that ye may
see.
\VS{19}Who
{\ADD{is}}
blind, but my
servant? or
deaf, as my
messenger
{\ADD{that}} I
sent? who
{\ADD{is}}
blind as
{\ADD{he that is}}
perfect, and
blind as the
{\ND{
LORD’s}}
servant?
\VS{20}Seeing many
things, but thou
observest not;
opening the
ears, but he
heareth not.
\VS{21}The
{\ND{
LORD}} is well
pleased for his
righteousness’ sake; he will
magnify the
law, and make
{\ADD{it}}
honourable.\FTNT{§§}{{\FR 42.21 }
it: or, him}
\VS{22}But this
{\ADD{is}} a
people
robbed and
spoiled;
{\ADD{they are}} all of them
snared in
holes, and they are
hid in
prison
houses: they are for a
prey, and none
delivereth; for a
spoil, and none
saith,
Restore.\FTNT{***}{{\FR 42.22 }
they are all…: or, in snaring all the young men of them}\FTNT{†††}{{\FR 42.22 }
for a spoil: Heb. a treading}
\VS{23}Who among you will give
ear to this?
{\ADD{who}} will
hearken and
hear for the time to
come?\FTNT{‡‡‡}{{\FR 42.23 }
for…: Heb. for the after time?}
\VS{24}Who
gave
Jacob for a
spoil, and
Israel to the
robbers? did not the
{\ND{
LORD}}, he against
whom we have
sinned? for they
would not
walk in his
ways, neither were they
obedient unto his
law.
\VS{25}Therefore he hath
poured upon him the
fury of his
anger, and the
strength of
battle: and it hath set him on
fire round
about, yet he
knew not; and it
burned him, yet he
laid
{\ADD{it}} not to
heart.

\par }\Chap{43}{\PP \VerseOne{1}But now thus
saith the
{\ND{
LORD}} that
created thee, O
Jacob, and he that
formed thee, O
Israel,
Fear not: for I have
redeemed thee, I have
called
{\ADD{thee}} by thy
name; thou
{\ADD{art}} mine.
\VS{2}When thou passest
through the
waters, I
{\ADD{will be}} with thee; and through the
rivers, they shall not
overflow thee: when thou
walkest
through the
fire, thou shalt not be
burned; neither shall the
flame
kindle upon thee.
\VS{3}For I
{\ADD{am}} the
{\ND{
LORD}} thy
God, the Holy
One of
Israel, thy
Saviour: I
gave
Egypt
{\ADD{for}} thy
ransom,
Ethiopia and
Seba for thee.
\VS{4}Since thou wast
precious in my
sight, thou hast been
honourable, and I have
loved thee: therefore will I
give
men for thee, and
people for thy
life.\FTNT{*}{{\FR 43.4 }
life: or, person}
\VS{5}Fear not: for I
{\ADD{am}} with thee: I will
bring thy
seed from the
east, and
gather thee from the
west;
\VS{6}I will
say to the
north, Give
up; and to the
south, Keep not
back:
bring my
sons from
far, and my
daughters from the
ends of the
earth;
\VS{7}{\ADD{Even}} every
one that is
called by my
name: for I have
created him for my
glory, I have
formed him; yea, I have
made him.
\par }{\PP \VS{8}¶ Bring
forth the
blind
people that
have
eyes, and the
deaf that have
ears.
\VS{9}Let all the
nations be
gathered
together, and let the
people be
assembled: who among them can
declare this, and
shew us former
things? let them bring
forth their
witnesses, that they may be
justified: or let them
hear, and
say,
{\ADD{It is}}
truth.
\VS{10}Ye
{\ADD{are}} my
witnesses,
saith the
{\ND{
LORD}}, and my
servant whom I have
chosen: that ye may
know and
believe me, and
understand that I
{\ADD{am}} he:
before me there was no
God
formed, neither shall there be
after me.\FTNT{†}{{\FR 43.10 }
no…: or, nothing formed of God}
\VS{11}I,
{\ADD{even}} I,
{\ADD{am}} the
{\ND{
LORD}}; and
beside me
{\ADD{there is}} no
saviour.
\VS{12}I have
declared, and have
saved, and I have
shewed, when
{\ADD{there was}} no
strange
{\ADD{god}} among you: therefore ye
{\ADD{are}} my
witnesses,
saith the
{\ND{
LORD}}, that I
{\ADD{am}}
God.
\VS{13}Yea, before the
day
{\ADD{was}} I
{\ADD{am}} he; and
{\ADD{there is}} none that can
deliver out of my
hand: I will
work, and who shall
let it?\FTNT{‡}{{\FR 43.13 }
let it: Heb. turn it back?}
\par }{\PP \VS{14}¶ Thus
saith the
{\ND{
LORD}}, your
redeemer, the Holy
One of
Israel; For your sake I have
sent to
Babylon, and have brought
down all their
nobles, and the
Chaldeans, whose
cry
{\ADD{is}} in the
ships.\FTNT{§}{{\FR 43.14 }
nobles: Heb. bars}
\VS{15}I
{\ADD{am}} the
{\ND{
LORD}}, your Holy
One, the
creator of
Israel, your
King.
\VS{16}Thus
saith the
{\ND{
LORD}}, which
maketh a
way in the
sea, and a
path in the
mighty
waters;
\VS{17}Which bringeth
forth the
chariot and
horse, the
army and the
power; they shall lie
down
together, they shall not
rise: they are
extinct, they are
quenched as
tow.
\par }{\PP \VS{18}¶
Remember ye not the former
things, neither
consider the things of
old.
\VS{19}Behold, I will
do a new
thing; now it shall spring
forth; shall ye not
know it? I will even
make a
way in the
wilderness,
{\ADD{and}}
rivers in the
desert.
\VS{20}The
beast of the
field shall
honour me, the
dragons and the
owls: because I
give
waters in the
wilderness,
{\ADD{and}}
rivers in the
desert, to give
drink to my
people, my
chosen.\FTNT{**}{{\FR 43.20 }
owls: or, ostriches: Heb. daughters of the owl}
\VS{21}This
people have I
formed for myself; they shall shew
forth my
praise.
\par }{\PP \VS{22}¶ But thou hast not
called upon me, O
Jacob; but thou hast been
weary of me, O
Israel.
\VS{23}Thou hast not
brought me the small
cattle of thy burnt
offerings; neither hast thou
honoured me with thy
sacrifices. I have not caused thee to
serve with an
offering, nor
wearied thee with
incense.\FTNT{††}{{\FR 43.23 }
small…: Heb. lambs, or, kids}
\VS{24}Thou hast
bought me no sweet
cane with
money, neither hast thou
filled me with the
fat of thy
sacrifices: but thou hast made me to
serve with thy
sins, thou hast
wearied me with thine
iniquities.\FTNT{‡‡}{{\FR 43.24 }
filled…: Heb. made me drunk, or, abundantly moistened}
\VS{25}I,
{\ADD{even}} I,
{\ADD{am}} he that blotteth
out thy
transgressions for mine own sake, and will not
remember thy
sins.
\VS{26}Put me in
remembrance: let us
plead
together:
declare thou, that thou mayest be
justified.
\VS{27}Thy
first
father hath
sinned, and thy
teachers have
transgressed against me.\FTNT{§§}{{\FR 43.27 }
teachers: Heb. interpreters}
\VS{28}Therefore I have
profaned the
princes of the
sanctuary, and have
given
Jacob to the
curse, and
Israel to
reproaches.\FTNT{***}{{\FR 43.28 }
princes…: or, holy princes}

\par }\Chap{44}{\PP \VerseOne{1}Yet now
hear, O
Jacob my
servant; and
Israel, whom I have
chosen:
\VS{2}Thus
saith the
{\ND{
LORD}} that
made thee, and
formed thee from the
womb,
{\ADD{which}} will
help thee;
Fear not, O
Jacob, my
servant; and thou,
Jesurun, whom I have
chosen.
\VS{3}For I will
pour
water upon him that is
thirsty, and
floods upon the dry
ground: I will
pour my
spirit upon thy
seed, and my
blessing upon thine
offspring:
\VS{4}And they shall spring
up
{\ADD{as}}
among the
grass, as
willows by the
water
courses.
\VS{5}One shall
say, I
{\ADD{am}} the
{\ND{
LORD’s}}; and another shall
call
{\ADD{himself}} by the
name of
Jacob; and another shall
subscribe
{\ADD{with}} his
hand unto the
{\ND{
LORD}}, and
surname
{\ADD{himself}} by the
name of
Israel.
\VS{6}Thus
saith the
{\ND{
LORD}} the
King of
Israel, and his
redeemer the
{\ND{
LORD}} of
hosts; I
{\ADD{am}} the
first, and I
{\ADD{am}} the
last; and
beside me
{\ADD{there is}} no
God.
\VS{7}And who, as I, shall
call, and shall
declare it, and set it in
order for me, since I
appointed the
ancient
people? and the things that are
coming, and shall
come, let them
shew unto them.
\VS{8}Fear ye not, neither be
afraid: have not I
told thee from that
time, and have
declared
{\ADD{it}}? ye
{\ADD{are}} even my
witnesses. Is
there a
God
beside me? yea,
{\ADD{there is}} no
God; I
know not
{\ADD{any}}.\FTNT{*}{{\FR 44.8 }
God; I: Heb. rock, etc}
\par }{\PP \VS{9}¶ They that
make a graven
image
{\ADD{are}} all of them
vanity; and their delectable
things shall not
profit; and they
{\ADD{are}} their own
witnesses; they
see not, nor
know; that they may be
ashamed.\FTNT{†}{{\FR 44.9 }
delectable: Heb. desirable}
\VS{10}Who hath
formed a
god, or
molten a graven
image
{\ADD{that}} is
profitable for
nothing?
\VS{11}Behold, all his
fellows shall be
ashamed: and the
workmen, they
{\ADD{are}} of
men: let them all be gathered
together, let them stand
up;
{\ADD{yet}} they shall
fear,
{\ADD{and}} they shall be
ashamed
together.
\VS{12}The
smith with the
tongs both
worketh in the
coals, and
fashioneth it with
hammers, and
worketh it with the
strength of his
arms: yea, he is
hungry, and his
strength
faileth: he
drinketh no
water, and is
faint.\FTNT{‡}{{\FR 44.12 }
with the tongs: or, with an axe}
\VS{13}The
carpenter stretcheth
out
{\ADD{his}}
rule; he marketh it
out with a
line; he
fitteth it with
planes, and he marketh it
out with the
compass, and
maketh it after the
figure of a
man, according to the
beauty of a
man; that it may
remain in the
house.
\VS{14}He heweth him
down
cedars, and
taketh the
cypress and the
oak, which he
strengtheneth for himself among the
trees of the
forest: he
planteth an
ash, and the
rain doth
nourish
{\ADD{it}}.\FTNT{§}{{\FR 44.14 }
strengtheneth: or, taketh courage}
\VS{15}Then shall it be for a
man to
burn: for he will
take thereof, and
warm himself; yea, he
kindleth
{\ADD{it}}, and
baketh
bread; yea, he
maketh a
god, and
worshippeth
{\ADD{it;}} he
maketh it a graven
image, and falleth
down thereto.
\VS{16}He
burneth
part thereof
in the
fire; with
part thereof he
eateth
flesh; he
roasteth
roast, and is
satisfied: yea, he
warmeth
{\ADD{himself}}, and
saith,
Aha, I am
warm, I have
seen the
fire:
\VS{17}And the
residue thereof he
maketh a
god,
{\ADD{even}} his graven
image: he falleth
down unto it, and
worshippeth
{\ADD{it}}, and
prayeth unto it, and
saith,
Deliver me; for thou
{\ADD{art}} my
god.
\VS{18}They have not
known nor
understood: for he hath
shut their
eyes, that they cannot
see;
{\ADD{and}} their
hearts, that they cannot
understand.\FTNT{**}{{\FR 44.18 }
shut: Heb. daubed}
\VS{19}And none
considereth in his
heart, neither
{\ADD{is there}}
knowledge nor
understanding to
say, I have
burned
part of it
in the
fire; yea, also I have
baked
bread upon the
coals thereof; I have
roasted
flesh, and
eaten
{\ADD{it:}} and shall I
make the
residue thereof an
abomination? shall I fall
down to the
stock of a
tree?\FTNT{††}{{\FR 44.19 }
considereth…: Heb. setteth to his heart}\FTNT{‡‡}{{\FR 44.19 }
the stock…: Heb. that which comes of a tree?}
\VS{20}He
feedeth on
ashes: a
deceived
heart hath turned him
aside, that he cannot
deliver his
soul, nor
say,
{\ADD{Is there}} not a
lie in my right
hand?
\par }{\PP \VS{21}¶
Remember these, O
Jacob and
Israel; for thou
{\ADD{art}} my
servant: I have
formed thee; thou
{\ADD{art}} my
servant: O
Israel, thou shalt not be
forgotten of me.
\VS{22}I have blotted
out, as a thick
cloud, thy
transgressions, and, as a
cloud, thy
sins:
return unto me; for I have
redeemed thee.
\VS{23}Sing, O ye
heavens; for the
{\ND{
LORD}} hath
done
{\ADD{it:}}
shout, ye lower
parts of the
earth: break
forth into
singing, ye
mountains, O
forest, and every
tree therein: for the
{\ND{
LORD}} hath
redeemed
Jacob, and
glorified himself in
Israel.
\VS{24}Thus
saith the
{\ND{
LORD}}, thy
redeemer, and he that
formed thee from the
womb, I
{\ADD{am}} the
{\ND{
LORD}} that
maketh all
{\ADD{things;}} that stretcheth
forth the
heavens alone; that spreadeth
abroad the
earth by myself;
\VS{25}That
frustrateth the
tokens of the
liars, and maketh
diviners
mad; that
turneth
wise
{\ADD{men}}
backward, and maketh their
knowledge
foolish;
\VS{26}That
confirmeth the
word of his
servant, and
performeth the
counsel of his
messengers; that
saith to
Jerusalem, Thou shalt be
inhabited; and to the
cities of
Judah, Ye shall be
built, and I will raise
up the decayed
places thereof:\FTNT{§§}{{\FR 44.26 }
decayed…: Heb. wastes}
\VS{27}That
saith to the
deep, Be
dry, and I will dry
up thy
rivers:
\VS{28}That
saith of
Cyrus,
{\ADD{He is}} my
shepherd, and shall
perform all my
pleasure: even
saying to
Jerusalem, Thou shalt be
built; and to the
temple, Thy foundation shall be
laid.

\par }\Chap{45}{\PP \VerseOne{1}Thus
saith the
{\ND{
LORD}} to his
anointed, to
Cyrus, whose right
hand I have
holden, to
subdue
nations
before him; and I will
loose the
loins of
kings, to
open
before him the two leaved
gates; and the
gates shall not be
shut;\FTNT{*}{{\FR 45.1 }
have…: or, strengthened}
\VS{2}I will
go
before thee, and
make the crooked
places
straight: I will break in
pieces the
gates of
brass, and cut in
sunder the
bars of
iron:
\VS{3}And I will
give thee the
treasures of
darkness, and hidden
riches of secret
places, that thou mayest
know that I, the
{\ND{
LORD}}, which
call
{\ADD{thee}} by thy
name,
{\ADD{am}} the
God of
Israel.
\VS{4}For
Jacob my
servant’s sake, and
Israel mine
elect, I have even
called thee by thy
name: I have
surnamed thee, though thou hast not
known me.
\par }{\PP \VS{5}¶ I
{\ADD{am}} the
{\ND{
LORD}}, and
{\ADD{there is}} none else,
{\ADD{there is}} no
God
beside me: I
girded thee, though thou hast not
known me:
\VS{6}That they may
know from the
rising of the
sun, and from the
west, that
{\ADD{there is}}
none
beside me. I
{\ADD{am}} the
{\ND{
LORD}}, and
{\ADD{there is}} none else.
\VS{7}I
form the
light, and
create
darkness: I
make
peace, and
create
evil: I the
{\ND{
LORD}}
do all these
{\ADD{things}}.
\VS{8}Drop
down, ye
heavens, from
above, and let the
skies pour
down
righteousness: let the
earth
open, and let them bring
forth
salvation, and let
righteousness spring
up
together; I the
{\ND{
LORD}} have
created it.
\VS{9}Woe unto him that
striveth with his
Maker!
{\ADD{Let}} the
potsherd
{\ADD{strive}} with the
potsherds of the
earth. Shall the
clay
say to him that
fashioneth it, What
makest thou? or thy
work, He hath no
hands?
\VS{10}Woe unto him that
saith unto
{\ADD{his}}
father, What
begettest thou? or to the
woman, What hast thou brought
forth?
\par }{\PP \VS{11}Thus
saith the
{\ND{
LORD}}, the Holy
One of
Israel, and his
Maker,
Ask me of things to
come concerning my
sons, and concerning the
work of my
hands
command ye me.
\VS{12}I have
made the
earth, and
created
man upon it: I,
{\ADD{even}} my
hands, have stretched
out the
heavens, and all their
host have I
commanded.
\VS{13}I have raised him
up in
righteousness, and I will
direct all his
ways: he shall
build my
city, and he shall let
go my
captives, not for
price nor
reward,
saith the
{\ND{
LORD}} of
hosts.\FTNT{†}{{\FR 45.13 }
direct: or, make straight}
\VS{14}Thus
saith the
{\ND{
LORD}}, The
labour of
Egypt, and
merchandise of
Ethiopia and of the
Sabeans,
men of
stature, shall come
over unto thee, and they shall be thine: they shall
come
after thee; in
chains they shall come
over, and they shall fall
down unto thee, they shall make
supplication unto thee,
{\ADD{saying}}, Surely
God
{\ADD{is}} in thee; and
{\ADD{there is}} none else,
{\ADD{there is}}
no
God.
\VS{15}Verily thou
{\ADD{art}} a
God that
hidest thyself, O
God of
Israel, the
Saviour.
\VS{16}They shall be
ashamed, and also
confounded, all of them: they shall
go to
confusion
together
{\ADD{that are}}
makers of
idols.
\VS{17}{\ADD{But}}
Israel shall be
saved in the
{\ND{
LORD}} with an
everlasting
salvation: ye shall not be
ashamed nor
confounded
world without
end.
\VS{18}For thus
saith the
{\ND{
LORD}} that
created the
heavens;
God himself that
formed the
earth and
made it; he hath
established it, he
created it not in
vain, he
formed it to be
inhabited: I
{\ADD{am}} the
{\ND{
LORD}}; and
{\ADD{there is}} none else.
\VS{19}I have not
spoken in
secret, in a
dark
place of the
earth: I
said not unto the
seed of
Jacob,
Seek ye me in
vain: I the
{\ND{
LORD}}
speak
righteousness, I
declare things that are
right.
\par }{\PP \VS{20}¶
Assemble yourselves and
come; draw
near
together, ye
{\ADD{that are}}
escaped of the
nations: they have no
knowledge that set
up the
wood of their graven
image, and
pray unto a
god
{\ADD{that}} cannot
save.
\VS{21}Tell ye, and bring
{\ADD{them}}
near; yea, let them take
counsel
together: who hath
declared this from ancient
time?
{\ADD{who}} hath
told it from that time?
{\ADD{have}} not I the
{\ND{
LORD}}? and
{\ADD{there is}} no
God else
beside me; a
just
God and a
Saviour;
{\ADD{there is}}
none
beside me.
\VS{22}Look unto me, and be ye
saved, all the
ends of the
earth: for I
{\ADD{am}}
God, and
{\ADD{there is}} none else.
\VS{23}I have
sworn by myself, the
word is gone
out of my
mouth
{\ADD{in}}
righteousness, and shall not
return, That unto me every
knee shall
bow, every
tongue shall
swear.
\VS{24}Surely, shall
{\ADD{one}}
say, in the
{\ND{
LORD}} have I
righteousness and
strength:
{\ADD{even}} to him shall
{\ADD{men}}
come; and all that are
incensed against him shall be
ashamed.\FTNT{‡}{{\FR 45.24 }
Surely…: or, Surely he shall say of me, In the
{\ND{LORD}} is all righteousness and strength}\FTNT{§}{{\FR 45.24 }
righteousness: Heb. righteousnesses}
\VS{25}In the
{\ND{
LORD}} shall all the
seed of
Israel be
justified, and shall
glory.

\par }\Chap{46}{\PP \VerseOne{1}Bel boweth
down,
Nebo
stoopeth, their
idols were upon the
beasts, and upon the
cattle: your
carriages
{\ADD{were}} heavy
loaden;
{\ADD{they are}} a
burden to the
weary
{\ADD{beast}}.
\VS{2}They
stoop, they bow
down
together; they
could not
deliver the
burden, but
themselves are
gone into
captivity.\FTNT{*}{{\FR 46.2 }
themselves: Heb. their soul}
\par }{\PP \VS{3}¶
Hearken unto me, O
house of
Jacob, and all the
remnant of the
house of
Israel, which are
borne
{\ADD{by me}} from the
belly, which are
carried from the
womb:
\VS{4}And
{\ADD{even}} to
{\ADD{your}} old
age I
{\ADD{am}} he; and
{\ADD{even}} to hoar
hairs will I
carry
{\ADD{you:}} I have
made, and I will
bear; even I will
carry, and will
deliver
{\ADD{you}}.
\par }{\PP \VS{5}¶ To whom will ye
liken me, and make
{\ADD{me}}
equal, and
compare me, that we may be
like?
\VS{6}They
lavish
gold out of the
bag, and
weigh
silver in the
balance,
{\ADD{and}}
hire a
goldsmith; and he
maketh it a
god: they fall
down, yea, they
worship.
\VS{7}They
bear him upon the
shoulder, they
carry him, and set him in his
place, and he
standeth; from his
place shall he not
remove: yea,
{\ADD{one}} shall
cry unto him, yet can he not
answer, nor
save him out of his
trouble.
\VS{8}Remember this, and shew yourselves
men: bring
{\ADD{it}}
again to
mind, O ye
transgressors.
\VS{9}Remember the former
things of
old: for I
{\ADD{am}}
God, and
{\ADD{there is}} none else;
{\ADD{I am}}
God, and
{\ADD{there is}}
none
like me,
\VS{10}Declaring the
end from the
beginning, and from ancient
times
{\ADD{the things}} that are not
{\ADD{yet}}
done,
saying, My
counsel shall
stand, and I will
do all my
pleasure:
\VS{11}Calling a ravenous
bird from the
east, the
man that executeth my
counsel from a
far
country: yea, I have
spoken
{\ADD{it}}, I will also
bring it to pass; I have
purposed
{\ADD{it}}, I will also
do it.\FTNT{†}{{\FR 46.11 }
that…: Heb. of my counsel}
\par }{\PP \VS{12}¶
Hearken unto me, ye
stouthearted, that
{\ADD{are}} far
from
righteousness:
\VS{13}I bring
near my
righteousness; it shall not be far
off, and my
salvation shall not
tarry: and I will
place
salvation in
Zion for
Israel my
glory.

\par }\Chap{47}{\PP \VerseOne{1}Come
down, and
sit in the
dust, O
virgin
daughter of
Babylon,
sit on the
ground:
{\ADD{there is}} no
throne, O
daughter of the
Chaldeans: for thou shalt no
more be
called
tender and
delicate.
\VS{2}Take the
millstones, and
grind
meal:
uncover thy
locks, make
bare the
leg,
uncover the
thigh, pass
over the
rivers.
\VS{3}Thy
nakedness shall be
uncovered, yea, thy
shame shall be
seen: I will
take
vengeance, and I will not
meet
{\ADD{thee as}} a
man.
\VS{4}{\ADD{As for}} our
redeemer, the
{\ND{
LORD}} of
hosts
{\ADD{is}} his
name, the Holy
One of
Israel.
\VS{5}Sit thou
silent, and
get thee into
darkness, O
daughter of the
Chaldeans: for thou shalt no
more be
called, The
lady of
kingdoms.
\par }{\PP \VS{6}¶ I was
wroth with my
people, I have
polluted mine
inheritance, and
given them into thine
hand: thou didst
shew them no
mercy; upon the
ancient hast thou
very
heavily laid thy
yoke.
\par }{\PP \VS{7}¶ And thou
saidst, I shall be a
lady for
ever:
{\ADD{so}} that thou didst not
lay these
{\ADD{things}} to thy
heart, neither didst
remember the latter
end of it.
\VS{8}Therefore
hear now this,
{\ADD{thou that art}} given to
pleasures, that
dwellest
carelessly, that
sayest in thine
heart, I
{\ADD{am}}, and none
else beside me; I shall not
sit
{\ADD{as}} a
widow, neither shall I
know the loss of
children:
\VS{9}But these
two
{\ADD{things}} shall
come to thee in a
moment in
one
day, the loss of
children, and
widowhood: they shall
come upon thee in their
perfection for the
multitude of thy
sorceries,
{\ADD{and}} for the
great
abundance of thine
enchantments.
\par }{\PP \VS{10}¶ For thou hast
trusted in thy
wickedness: thou hast
said, None
seeth me. Thy
wisdom and thy
knowledge, it hath
perverted thee; and thou hast
said in thine
heart, I
{\ADD{am}}, and none
else beside me.\FTNT{*}{{\FR 47.10 }
perverted…: or, caused thee to turn away}
\par }{\PP \VS{11}¶ Therefore shall
evil
come upon thee; thou shalt not
know from whence it
riseth: and
mischief shall
fall upon thee; thou shalt not be
able to put it
off: and
desolation shall
come upon thee
suddenly,
{\ADD{which}} thou shalt not
know.\FTNT{†}{{\FR 47.11 }
from…: Heb. the morning thereof}\FTNT{‡}{{\FR 47.11 }
put: Heb. expiate}
\VS{12}Stand now with thine
enchantments, and with the
multitude of thy
sorceries,
wherein thou hast
laboured from thy
youth; if so be thou shalt be
able to
profit, if so be thou mayest
prevail.
\VS{13}Thou art
wearied in the
multitude of thy
counsels. Let now the
astrologers, the
stargazers, the
monthly
prognosticators, stand
up, and
save thee from
{\ADD{these things}} that shall
come upon thee.\FTNT{§}{{\FR 47.13 }
astrologers: Heb. viewers of the heavens}\FTNT{**}{{\FR 47.13 }
the monthly…: Heb. that give knowledge concerning the months}
\VS{14}Behold, they shall be as
stubble; the
fire shall
burn them; they shall not
deliver
themselves from the
power of the
flame:
{\ADD{there shall}} not
{\ADD{be}} a
coal to
warm at,
{\ADD{nor}}
fire to
sit before it.\FTNT{††}{{\FR 47.14 }
themselves: Heb. their souls}
\VS{15}Thus shall they be unto thee with whom thou hast
laboured,
{\ADD{even}} thy
merchants, from thy
youth: they shall
wander every
one to his
quarter; none shall
save thee.

\par }\Chap{48}{\PP \VerseOne{1}Hear ye this, O
house of
Jacob, which are
called by the
name of
Israel, and are come
forth out of the
waters of
Judah, which
swear by the
name of the
{\ND{
LORD}}, and make
mention of the
God of
Israel,
{\ADD{but}} not in
truth, nor in
righteousness.
\VS{2}For they
call themselves of the
holy
city, and
stay themselves upon the
God of
Israel; The
{\ND{
LORD}} of
hosts
{\ADD{is}} his
name.
\VS{3}I have
declared the former
things from the
beginning; and they went
forth out of my
mouth, and I
shewed them; I
did
{\ADD{them}}
suddenly, and they came to
pass.
\VS{4}Because I
knew that thou
{\ADD{art}}
obstinate, and thy
neck
{\ADD{is}} an
iron
sinew, and thy
brow
brass;\FTNT{*}{{\FR 48.4 }
obstinate: Heb. hard}
\VS{5}I have even from the
beginning
declared
{\ADD{it}} to thee; before it came to
pass I
shewed
{\ADD{it}} thee: lest thou shouldest
say, Mine
idol hath
done them, and my graven
image, and my molten
image, hath
commanded them.
\VS{6}Thou hast
heard,
see all this; and will not ye
declare
{\ADD{it}}? I have
shewed thee new
things from this
time, even hidden
things, and thou didst not
know them.
\VS{7}They are
created now, and not from the beginning; even
before the
day when thou
heardest them not; lest thou shouldest
say, Behold, I
knew them.
\VS{8}Yea, thou
heardest not; yea, thou
knewest not; yea, from that time
{\ADD{that}} thine
ear was not
opened: for I
knew that thou wouldest deal
very
treacherously, and wast
called a
transgressor from the
womb.
\par }{\PP \VS{9}¶ For my
name’s sake will I
defer mine
anger, and for my
praise will I
refrain for thee, that I cut thee not
off.
\VS{10}Behold, I have
refined thee, but not with
silver; I have
chosen thee in the
furnace of
affliction.\FTNT{†}{{\FR 48.10 }
with…: or, for silver}
\VS{11}For mine own sake,
{\ADD{even}} for mine own sake, will I
do
{\ADD{it:}} for how should
{\ADD{my name}} be
polluted? and I will not
give my
glory unto
another.
\par }{\PP \VS{12}¶
Hearken unto me, O
Jacob and
Israel, my
called; I
{\ADD{am}} he; I
{\ADD{am}} the
first, I also
{\ADD{am}} the
last.
\VS{13}Mine
hand also hath laid the
foundation of the
earth, and my right
hand hath
spanned the
heavens:
{\ADD{when}} I
call unto them, they stand
up
together.\FTNT{‡}{{\FR 48.13 }
my…: or, the palm of my right hand hath spread out}
\VS{14}All ye,
assemble yourselves, and
hear; which among them hath
declared these
{\ADD{things}}? The
{\ND{
LORD}} hath
loved him: he will
do his
pleasure on
Babylon, and his
arm
{\ADD{shall be on}} the
Chaldeans.
\VS{15}I,
{\ADD{even}} I, have
spoken; yea, I have
called him: I have
brought him, and he shall make his
way
prosperous.
\par }{\PP \VS{16}¶
Come ye
near unto me,
hear ye this; I have not
spoken in
secret from the
beginning; from the
time that it was, there
{\ADD{am}} I: and now the
Lord
{\ND{
GOD}}, and his
Spirit, hath
sent me.
\VS{17}Thus
saith the
{\ND{
LORD}}, thy
Redeemer, the Holy
One of
Israel; I
{\ADD{am}} the
{\ND{
LORD}} thy
God which
teacheth thee to
profit, which
leadeth thee by the
way
{\ADD{that}} thou shouldest
go.
\VS{18}O
that thou hadst
hearkened to my
commandments! then had thy
peace been as a
river, and thy
righteousness as the
waves of the
sea:
\VS{19}Thy
seed also had been as the
sand, and the
offspring of thy
bowels like the
gravel thereof; his
name should not have been cut
off nor
destroyed from
before me.
\par }{\PP \VS{20}¶ Go ye
forth of
Babylon,
flee ye from the
Chaldeans, with a
voice of
singing
declare ye,
tell this,
utter it
{\ADD{even}} to the
end of the
earth;
say ye, The
{\ND{
LORD}} hath
redeemed his
servant
Jacob.
\VS{21}And they
thirsted not
{\ADD{when}} he
led them through the
deserts: he caused the
waters to
flow out of the
rock for them: he
clave the
rock also, and the
waters gushed
out.
\VS{22}{\ADD{There is}} no
peace,
saith the
{\ND{
LORD}}, unto the
wicked.

\par }\Chap{49}{\PP \VerseOne{1}Listen, O
isles, unto me; and
hearken, ye
people, from
far; The
{\ND{
LORD}} hath
called me from the
womb; from the
bowels of my
mother hath he made
mention of my
name.
\VS{2}And he hath
made my
mouth like a
sharp
sword; in the
shadow of his
hand hath he
hid me, and
made me a
polished
shaft; in his
quiver hath he
hid me;
\VS{3}And
said unto me, Thou
{\ADD{art}} my
servant, O
Israel, in whom I will be
glorified.
\VS{4}Then I
said, I have
laboured in
vain, I have
spent my
strength for
nought, and in
vain:
{\ADD{yet}}
surely my
judgment
{\ADD{is}} with the
{\ND{
LORD}}, and my
work with my
God.\FTNT{*}{{\FR 49.4 }
my work: or, my reward}
\par }{\PP \VS{5}¶ And now,
saith the
{\ND{
LORD}} that
formed me from the
womb
{\ADD{to be}} his
servant, to
bring
Jacob
again to him, Though
Israel be not
gathered, yet shall I be
glorious in the
eyes of the
{\ND{
LORD}}, and my
God shall be my
strength.\FTNT{†}{{\FR 49.5 }
Though…: or, That Israel may be gathered to him, and I may, etc}
\VS{6}And he
said, It is a light
thing that thou shouldest be my
servant to raise
up the
tribes of
Jacob, and to
restore the
preserved of
Israel: I will also
give thee for a
light to the
Gentiles, that thou mayest be my
salvation unto the
end of the
earth.\FTNT{‡}{{\FR 49.6 }
It is…: or, Art thou lighter than that thou shouldest, etc}\FTNT{§}{{\FR 49.6 }
preserved: or, desolations}
\par }{\PP \VS{7}Thus
saith the
{\ND{
LORD}}, the
Redeemer of
Israel,
{\ADD{and}} his Holy
One, to him whom
man
despiseth, to him whom the
nation
abhorreth, to a
servant of
rulers,
Kings shall
see and
arise,
princes also shall
worship, because of the
{\ND{
LORD}} that is
faithful,
{\ADD{and}} the Holy
One of
Israel, and he shall
choose thee.\FTNT{**}{{\FR 49.7 }
whom man…: or, that is despised in soul}
\VS{8}Thus
saith the
{\ND{
LORD}}, In an
acceptable
time have I
heard thee, and in a
day of
salvation have I
helped thee: and I will
preserve thee, and
give thee for a
covenant of the
people, to
establish the
earth, to cause to
inherit the
desolate
heritages;\FTNT{††}{{\FR 49.8 }
establish: or, raise up}
\VS{9}That thou mayest
say to the
prisoners, Go
forth; to them that
{\ADD{are}} in
darkness,
Shew yourselves. They shall
feed in the
ways, and their
pastures
{\ADD{shall be}} in all high
places.
\VS{10}They shall not
hunger nor
thirst; neither shall the
heat nor
sun
smite them: for he that hath
mercy on them shall
lead them, even by the
springs of
water shall he
guide them.
\VS{11}And I will
make all my
mountains a
way, and my
highways shall be
exalted.
\VS{12}Behold, these shall
come from
far: and, lo, these from the
north and from the
west; and these from the
land of
Sinim.
\par }{\PP \VS{13}¶
Sing, O
heavens; and be
joyful, O
earth; and break
forth into
singing, O
mountains: for the
{\ND{
LORD}} hath
comforted his
people, and will have
mercy upon his
afflicted.
\VS{14}But
Zion
said, The
{\ND{
LORD}} hath
forsaken me, and my
Lord hath
forgotten me.
\VS{15}Can a
woman
forget her sucking
child, that she should not have
compassion on the
son of her
womb? yea, they may
forget, yet will I not
forget thee.\FTNT{‡‡}{{\FR 49.15 }
that…: Heb. from having compassion}
\VS{16}Behold, I have
graven thee upon the palms of
{\ADD{my}}
hands; thy
walls
{\ADD{are}}
continually before me.
\VS{17}Thy
children shall make
haste; thy
destroyers and they that made thee
waste shall go
forth of thee.
\par }{\PP \VS{18}¶ Lift
up thine
eyes round
about, and
behold: all these gather themselves
together,
{\ADD{and}}
come to thee.
{\ADD{As}} I
live,
saith the
{\ND{
LORD}}, thou shalt surely
clothe thee with them all, as with an
ornament, and
bind them
{\ADD{on thee}}, as a
bride
{\ADD{doeth}}.
\VS{19}For thy
waste and thy desolate
places, and the
land of thy
destruction, shall even now be too
narrow by reason of the
inhabitants, and they that swallowed thee
up shall be far
away.
\VS{20}The
children which thou shalt have, after thou hast
lost the other, shall
say again in thine
ears, The
place
{\ADD{is}} too
strait for me:
give
place to me that I may
dwell.
\VS{21}Then shalt thou
say in thine
heart, Who hath
begotten me these, seeing I have lost my
children, and am
desolate, a
captive, and removing to and
fro? and who hath brought
up these? Behold, I was
left alone; these,
where
{\ADD{had}} they
{\ADD{been}}?
\VS{22}Thus
saith the
Lord
{\ND{
GOD}}, Behold, I will lift
up mine
hand to the
Gentiles, and set
up my
standard to the
people: and they shall
bring thy
sons in
{\ADD{their}}
arms, and thy
daughters shall be
carried upon
{\ADD{their}}
shoulders.\FTNT{§§}{{\FR 49.22 }
arms: Heb. bosom}
\VS{23}And
kings shall be thy nursing
fathers, and their
queens thy nursing
mothers: they shall bow
down to thee with
{\ADD{their}}
face toward the
earth, and lick
up the
dust of thy
feet; and thou shalt
know that I
{\ADD{am}} the
{\ND{
LORD}}: for they shall not be
ashamed that
wait for me.\FTNT{***}{{\FR 49.23 }
nursing fathers: Heb. nourishers}\FTNT{†††}{{\FR 49.23 }
queens: Heb. princesses}
\par }{\PP \VS{24}¶ Shall the
prey be
taken from the
mighty, or the
lawful
captive
delivered?\FTNT{‡‡‡}{{\FR 49.24 }
lawful…: Heb. captivity of the just}
\VS{25}But thus
saith the
{\ND{
LORD}}, Even the
captives of the
mighty shall be taken
away, and the
prey of the
terrible shall be
delivered: for I will
contend with him that
contendeth with thee, and I will
save thy
children.\FTNT{§§§}{{\FR 49.25 }
captives: Heb. captivity}
\VS{26}And I will
feed them that
oppress thee with their own
flesh; and they shall be
drunken with their own
blood, as with sweet
wine: and all
flesh shall
know that I the
{\ND{
LORD}}
{\ADD{am}} thy
Saviour and thy
Redeemer, the mighty
One of
Jacob.\FTNT{*}{{\FR 49.26 }
sweet: or, new}

\par }\Chap{50}{\PP \VerseOne{1}Thus
saith the
{\ND{
LORD}}, Where
{\ADD{is}} the
bill of your
mother’s
divorcement, whom I have put
away? or which of my
creditors
{\ADD{is it}} to whom I have
sold you? Behold, for your
iniquities have ye
sold yourselves, and for your
transgressions is your
mother put
away.
\VS{2}Wherefore, when I
came,
{\ADD{was there}} no
man? when I
called,
{\ADD{was there}} none to
answer? Is my
hand
shortened at
all, that it cannot
redeem? or have I no
power to
deliver? behold, at my
rebuke I dry
up the
sea, I
make the
rivers a
wilderness: their
fish
stinketh, because
{\ADD{there is}} no
water, and
dieth for
thirst.
\VS{3}I
clothe the
heavens with
blackness, and I
make
sackcloth their
covering.
\par }{\PP \VS{4}The
Lord
{\ND{
GOD}} hath
given me the
tongue of the
learned, that I should
know how to
speak a
word in season to
{\ADD{him that is}}
weary: he
wakeneth
morning by
morning, he
wakeneth mine
ear to
hear as the
learned.
\par }{\PP \VS{5}¶ The
Lord
{\ND{
GOD}} hath
opened mine
ear, and I was not
rebellious, neither turned
away
back.
\VS{6}I
gave my
back to the
smiters, and my
cheeks to them that plucked
off the hair: I
hid not my
face from
shame and
spitting.
\par }{\PP \VS{7}¶ For the
Lord
{\ND{
GOD}} will
help me; therefore shall I not be
confounded: therefore have I
set my
face like a
flint, and I
know that I shall not be
ashamed.
\VS{8}{\ADD{He is}}
near that
justifieth me; who will
contend with me? let us
stand
together: who
{\ADD{is}} mine
adversary? let him come
near to me.\FTNT{*}{{\FR 50.8 }
mine…: Heb. the master of my cause?}
\VS{9}Behold, the
Lord
{\ND{
GOD}} will
help me; who
{\ADD{is}} he
{\ADD{that}} shall
condemn me? lo, they all shall wax
old as a
garment; the
moth shall eat them
up.
\par }{\PP \VS{10}¶ Who
{\ADD{is}} among you that
feareth the
{\ND{
LORD}}, that
obeyeth the
voice of his
servant, that
walketh
{\ADD{in}}
darkness, and hath no
light? let him
trust in the
name of the
{\ND{
LORD}}, and
stay upon his
God.
\VS{11}Behold, all ye that
kindle a
fire, that compass
{\ADD{yourselves}}
about with
sparks:
walk in the
light of your
fire, and in the
sparks
{\ADD{that}} ye have
kindled. This shall ye have of mine
hand; ye shall lie
down in
sorrow.

\par }\Chap{51}{\PP \VerseOne{1}Hearken to me, ye that follow
after
righteousness, ye that
seek the
{\ND{
LORD}}:
look unto the
rock
{\ADD{whence}} ye are
hewn, and to the
hole of the
pit
{\ADD{whence}} ye are
digged.
\VS{2}Look unto
Abraham your
father, and unto
Sarah
{\ADD{that}}
bare you: for I
called him
alone, and
blessed him, and
increased him.
\VS{3}For the
{\ND{
LORD}} shall
comfort
Zion: he will
comfort all her waste
places; and he will
make her
wilderness like
Eden, and her
desert like the
garden of the
{\ND{
LORD}};
joy and
gladness shall be
found therein,
thanksgiving, and the
voice of
melody.
\par }{\PP \VS{4}¶
Hearken unto me, my
people; and give
ear unto me, O my
nation: for a
law shall
proceed from me, and I will make my
judgment to
rest for a
light of the
people.
\VS{5}My
righteousness
{\ADD{is}}
near; my
salvation is gone
forth, and mine
arms shall
judge the
people; the
isles shall
wait upon me, and on mine
arm shall they
trust.
\VS{6}Lift
up your
eyes to the
heavens, and
look upon the
earth beneath: for the
heavens shall vanish
away like
smoke, and the
earth shall wax
old like a
garment, and they that
dwell therein shall
die in
like
manner: but my
salvation shall be for
ever, and my
righteousness shall not be
abolished.
\par }{\PP \VS{7}¶
Hearken unto me, ye that
know
righteousness, the
people in whose
heart
{\ADD{is}} my
law;
fear ye not the
reproach of
men, neither be ye
afraid of their
revilings.
\VS{8}For the
moth shall eat them
up like a
garment, and the
worm shall
eat them like
wool: but my
righteousness shall be for
ever, and my
salvation from
generation to
generation.
\par }{\PP \VS{9}¶
Awake,
awake, put
on
strength, O
arm of the
{\ND{
LORD}};
awake, as in the
ancient
days, in the
generations of
old.
{\ADD{Art}} thou not it that hath
cut
Rahab,
{\ADD{and}}
wounded the
dragon?
\VS{10}{\ADD{Art}} thou not it which hath
dried the
sea, the
waters of the
great
deep; that hath
made the
depths of the
sea a
way for the
ransomed to pass
over?
\VS{11}Therefore the
redeemed of the
{\ND{
LORD}} shall
return, and
come with
singing unto
Zion; and
everlasting
joy
{\ADD{shall be}} upon their
head: they shall
obtain
gladness and
joy;
{\ADD{and}}
sorrow and
mourning shall flee
away.
\VS{12}I,
{\ADD{even}} I,
{\ADD{am}} he that
comforteth you: who
{\ADD{art}} thou, that thou shouldest be
afraid of a
man
{\ADD{that}} shall
die, and of the
son of
man
{\ADD{which}} shall be
made
{\ADD{as}}
grass;
\VS{13}And
forgettest the
{\ND{
LORD}} thy
maker, that hath stretched
forth the
heavens, and laid the
foundations of the
earth; and hast
feared
continually every
day
because of the
fury of the
oppressor, as
if he were
ready to
destroy? and where
{\ADD{is}} the
fury of the
oppressor?\FTNT{*}{{\FR 51.13 }
were…: or, made himself ready}
\VS{14}The captive
exile
hasteneth that he may be
loosed, and that he should not
die in the
pit, nor that his
bread should
fail.
\VS{15}But I
{\ADD{am}} the
{\ND{
LORD}} thy
God, that
divided the
sea, whose
waves
roared: The
{\ND{
LORD}} of
hosts
{\ADD{is}} his
name.
\VS{16}And I have
put my
words in thy
mouth, and I have
covered thee in the
shadow of mine
hand, that I may
plant the
heavens, and lay the
foundations of the
earth, and
say unto
Zion, Thou
{\ADD{art}} my
people.
\par }{\PP \VS{17}¶
Awake,
awake, stand
up, O
Jerusalem, which hast
drunk at the
hand of the
{\ND{
LORD}} the
cup of his
fury; thou hast
drunken the
dregs of the
cup of
trembling,
{\ADD{and}} wrung
{\ADD{them}}
out.
\VS{18}{\ADD{There is}} none to
guide her among all the
sons
{\ADD{whom}} she hath brought
forth; neither
{\ADD{is there any}} that
taketh her by the
hand of all the
sons
{\ADD{that}} she hath brought
up.
\VS{19}These
two
{\ADD{things}} are
come unto thee; who shall be
sorry for thee?
desolation, and
destruction, and the
famine, and the
sword: by whom shall I
comfort thee?\FTNT{†}{{\FR 51.19 }
are…: Heb. happened}\FTNT{‡}{{\FR 51.19 }
destruction: Heb. breaking}
\VS{20}Thy
sons have
fainted, they
lie at the
head of all the
streets, as a wild
bull in a
net: they are
full of the
fury of the
{\ND{
LORD}}, the
rebuke of thy
God.
\par }{\PP \VS{21}¶ Therefore
hear now this, thou
afflicted, and
drunken, but not with
wine:
\VS{22}Thus
saith thy
Lord the
{\ND{
LORD}}, and thy
God
{\ADD{that}}
pleadeth the cause of his
people, Behold, I have taken
out of thine
hand the
cup of
trembling,
{\ADD{even}} the
dregs of the
cup of my
fury; thou shalt no
more drink it
again:
\VS{23}But I will
put it into the
hand of them that
afflict thee; which have
said to thy
soul, Bow
down, that we may go
over: and thou hast
laid thy
body as the
ground, and as the
street, to them that went
over.

\par }\Chap{52}{\PP \VerseOne{1}Awake,
awake; put
on thy
strength, O
Zion; put
on thy
beautiful
garments, O
Jerusalem, the
holy
city: for henceforth there shall no
more
come into thee the
uncircumcised and the
unclean.
\VS{2}Shake thyself from the
dust;
arise,
{\ADD{and}} sit
down, O
Jerusalem:
loose thyself from the
bands of thy
neck, O
captive
daughter of
Zion.
\VS{3}For thus
saith the
{\ND{
LORD}}, Ye have
sold yourselves for
nought; and ye shall be
redeemed without
money.
\VS{4}For thus
saith the
Lord
{\ND{
GOD}}, My
people went
down
aforetime into
Egypt to
sojourn there; and the
Assyrian
oppressed them without
cause.
\VS{5}Now therefore, what have I here,
saith the
{\ND{
LORD}}, that my
people is taken
away for
nought? they that
rule over them make them to
howl,
saith the
{\ND{
LORD}}; and my
name
continually every
day
{\ADD{is}}
blasphemed.
\VS{6}Therefore my
people shall
know my
name: therefore
{\ADD{they shall know}} in that
day that I
{\ADD{am}} he that doth
speak: behold,
{\ADD{it is}} I.
\par }{\PP \VS{7}¶ How
beautiful upon the
mountains are the
feet of him that bringeth good
tidings, that
publisheth
peace; that bringeth good
tidings of
good, that
publisheth
salvation; that
saith unto
Zion, Thy
God
reigneth!
\VS{8}Thy
watchmen shall lift
up the
voice; with the
voice
together shall they
sing: for they shall
see
eye to
eye, when the
{\ND{
LORD}} shall bring
again
Zion.
\par }{\PP \VS{9}¶ Break forth into
joy,
sing
together, ye waste
places of
Jerusalem: for the
{\ND{
LORD}} hath
comforted his
people, he hath
redeemed
Jerusalem.
\VS{10}The
{\ND{
LORD}} hath made
bare his
holy
arm in the
eyes of all the
nations; and all the
ends of the
earth shall
see the
salvation of our
God.
\par }{\PP \VS{11}¶
Depart ye,
depart ye, go ye
out from thence,
touch no
unclean
{\ADD{thing;}} go ye
out of the
midst of her; be ye
clean, that
bear the
vessels of the
{\ND{
LORD}}.
\VS{12}For ye shall not go
out with
haste, nor
go by
flight: for the
{\ND{
LORD}} will
go
before you; and the
God of
Israel
{\ADD{will be}} your
rereward.\FTNT{*}{{\FR 52.12 }
be…: Heb. gather you up}
\par }{\PP \VS{13}¶ Behold, my
servant shall deal
prudently, he shall be
exalted and
extolled, and be
very
high.\FTNT{†}{{\FR 52.13 }
deal…: or, prosper}
\VS{14}As
many were
astonied at thee; his
visage was so
marred more than any
man, and his
form more than the
sons of
men:
\VS{15}So shall he
sprinkle
many
nations; the
kings shall
shut their
mouths at him: for
{\ADD{that}} which had not been
told them shall they
see; and
{\ADD{that}} which they had not
heard shall they
consider.

\par }\Chap{53}{\PP \VerseOne{1}Who hath
believed our
report? and to whom is the
arm of the
{\ND{
LORD}}
revealed?\FTNT{*}{{\FR 53.1 }
report: or, doctrine?: Heb. hearing?}
\VS{2}For he shall grow
up
before him as a tender
plant, and as a
root out of a
dry
ground: he hath no
form nor
comeliness; and when we shall
see him,
{\ADD{there is}} no
beauty that we should
desire him.
\VS{3}He is
despised and
rejected of
men; a
man of
sorrows, and
acquainted with
grief: and we hid as it
were
{\ADD{our}}
faces from him; he was
despised, and we
esteemed him not.\FTNT{†}{{\FR 53.3 }
we hid…: or, he hid as it were his face from us: Heb. as an hiding of faces from him, or, from us}
\par }{\PP \VS{4}¶
Surely he hath
borne our
griefs, and
carried our
sorrows: yet we did
esteem him
stricken,
smitten of
God, and
afflicted.
\VS{5}But he
{\ADD{was}}
wounded for our
transgressions,
{\ADD{he was}}
bruised for our
iniquities: the
chastisement of our
peace
{\ADD{was}} upon him; and with his
stripes we are
healed.\FTNT{‡}{{\FR 53.5 }
wounded: or, tormented}\FTNT{§}{{\FR 53.5 }
stripes: Heb. bruise}
\VS{6}All we like
sheep have gone
astray; we have
turned every
one to his own
way; and the
{\ND{
LORD}} hath
laid on him the
iniquity of us all.\FTNT{**}{{\FR 53.6 }
laid…: Heb. made the iniquity of us all to meet on him}
\VS{7}He was
oppressed, and he was
afflicted, yet he
opened not his
mouth: he is
brought as a
lamb to the
slaughter, and as a
sheep
before her
shearers is
dumb, so he
openeth not his
mouth.
\VS{8}He was
taken from
prison and from
judgment: and who shall
declare his
generation? for he was cut
off out of the
land of the
living: for the
transgression of my
people was he
stricken.\FTNT{††}{{\FR 53.8 }
from prison…: or, away by distress and judgment: but, etc}\FTNT{‡‡}{{\FR 53.8 }
was he…: Heb. was the stroke upon him}
\VS{9}And he
made his
grave with the
wicked, and with the
rich in his
death; because he had
done no
violence, neither
{\ADD{was any}}
deceit in his
mouth.\FTNT{§§}{{\FR 53.9 }
death: Heb. deaths}
\par }{\PP \VS{10}¶ Yet it
pleased the
{\ND{
LORD}} to
bruise him; he hath put
{\ADD{him}} to
grief: when thou shalt
make his
soul an offering for
sin, he shall
see
{\ADD{his}}
seed, he shall
prolong
{\ADD{his}}
days, and the
pleasure of the
{\ND{
LORD}} shall
prosper in his
hand.\FTNT{***}{{\FR 53.10 }
thou…: or, his soul shall make an offering}
\VS{11}He shall
see of the
travail of his
soul,
{\ADD{and}} shall be
satisfied: by his
knowledge shall my
righteous
servant
justify
many; for he shall
bear their
iniquities.
\VS{12}Therefore will I
divide him
{\ADD{a portion}} with the
great, and he shall
divide the
spoil with the
strong; because he hath poured
out his
soul unto
death: and he was
numbered with the
transgressors; and he
bare the
sin of
many, and made
intercession for the
transgressors.

\par }\Chap{54}{\PP \VerseOne{1}Sing, O
barren, thou
{\ADD{that}} didst not
bear; break
forth into
singing, and cry
aloud, thou
{\ADD{that}} didst not travail with
child: for
more
{\ADD{are}} the
children of the
desolate than the
children of the married
wife,
saith the
{\ND{
LORD}}.
\VS{2}Enlarge the
place of thy
tent, and let them stretch
forth the
curtains of thine
habitations:
spare not,
lengthen thy
cords, and
strengthen thy
stakes;
\VS{3}For thou shalt break
forth on the right
hand and on the
left; and thy
seed shall
inherit the
Gentiles, and make the
desolate
cities to be
inhabited.
\VS{4}Fear not; for thou shalt not be
ashamed: neither be thou
confounded; for thou shalt not be put to
shame: for thou shalt
forget the
shame of thy
youth, and shalt not
remember the
reproach of thy
widowhood any more.
\VS{5}For thy
Maker
{\ADD{is}} thine
husband; the
{\ND{
LORD}} of
hosts
{\ADD{is}} his
name; and thy
Redeemer the Holy
One of
Israel; The
God of the whole
earth shall he be
called.\FTNT{*}{{\FR 54.5 }
Maker: Heb. Makers}
\par }{\PP \VS{6}For the
{\ND{
LORD}} hath
called thee as a
woman
forsaken and
grieved in
spirit, and a
wife of
youth, when thou wast
refused,
saith thy
God.
\VS{7}For a
small
moment have I
forsaken thee; but with
great
mercies will I
gather thee.
\VS{8}In a
little
wrath I
hid my
face from thee for a
moment; but with
everlasting
kindness will I have
mercy on thee,
saith the
{\ND{
LORD}} thy
Redeemer.
\VS{9}For this
{\ADD{is as}} the
waters of
Noah unto me: for
{\ADD{as}} I have
sworn that the
waters of
Noah should no more go
over the
earth; so have I
sworn that I would not be
wroth with thee, nor
rebuke thee.
\VS{10}For the
mountains shall
depart, and the
hills be
removed; but my
kindness shall not
depart from thee, neither shall the
covenant of my
peace be
removed,
saith the
{\ND{
LORD}} that hath
mercy on thee.
\par }{\PP \VS{11}¶ O thou
afflicted, tossed with
tempest,
{\ADD{and}} not
comforted, behold, I will
lay thy
stones with fair
colours, and lay thy
foundations with
sapphires.
\VS{12}And I will
make thy
windows of
agates, and thy
gates of
carbuncles, and all thy
borders of
pleasant
stones.
\VS{13}And all thy
children
{\ADD{shall be}}
taught of the
{\ND{
LORD}}; and
great
{\ADD{shall be}} the
peace of thy
children.
\VS{14}In
righteousness shalt thou be
established: thou shalt be
far from
oppression; for thou shalt not
fear: and from
terror; for it shall not come
near thee.
\VS{15}Behold, they shall
surely gather
together,
{\ADD{but}}
not by me:
whosoever shall gather
together against thee shall
fall for thy sake.
\VS{16}Behold, I have
created the
smith that
bloweth the
coals in the
fire, and that bringeth
forth an
instrument for his
work; and I have
created the
waster to
destroy.
\par }{\PP \VS{17}¶ No
weapon that is
formed against thee shall
prosper; and every
tongue
{\ADD{that}} shall
rise against thee in
judgment thou shalt
condemn. This
{\ADD{is}} the
heritage of the
servants of the
{\ND{
LORD}}, and their
righteousness
{\ADD{is}} of me,
saith the
{\ND{
LORD}}.

\par }\Chap{55}{\PP \VerseOne{1}Ho, every one that
thirsteth,
come ye to the
waters, and he that hath no
money;
come ye,
buy, and
eat; yea,
come,
buy
wine and
milk without
money and without
price.
\VS{2}Wherefore do ye
spend
money for
{\ADD{that which is}} not
bread? and your
labour for
{\ADD{that which}}
satisfieth
not?
hearken
diligently unto me, and
eat ye
{\ADD{that which is}}
good, and let your
soul
delight itself in
fatness.\FTNT{*}{{\FR 55.2 }
spend: Heb. weigh}
\VS{3}Incline your
ear, and
come unto me:
hear, and your
soul shall
live; and I will
make an
everlasting
covenant with you,
{\ADD{even}} the
sure
mercies of
David.
\VS{4}Behold, I have
given him
{\ADD{for}} a
witness to the
people, a
leader and
commander to the
people.
\VS{5}Behold, thou shalt
call a
nation
{\ADD{that}} thou
knowest not, and
nations
{\ADD{that}}
knew not thee shall
run unto thee because of the
{\ND{
LORD}} thy
God, and for the Holy
One of
Israel; for he hath
glorified thee.
\par }{\PP \VS{6}¶
Seek ye the
{\ND{
LORD}} while he may be
found,
call ye upon him while he is
near:
\VS{7}Let the
wicked
forsake his
way, and the
unrighteous
man his
thoughts: and let him
return unto the
{\ND{
LORD}}, and he will have
mercy upon him; and to our
God, for he will
abundantly
pardon.\FTNT{†}{{\FR 55.7 }
the unrighteous…: Heb. the man of iniquity}\FTNT{‡}{{\FR 55.7 }
abundantly…: Heb. multiply to pardon}
\par }{\PP \VS{8}¶ For my
thoughts
{\ADD{are}} not your
thoughts, neither
{\ADD{are}} your
ways my
ways,
saith the
{\ND{
LORD}}.
\VS{9}For
{\ADD{as}} the
heavens are
higher than the
earth, so are my
ways
higher than your
ways, and my
thoughts than your
thoughts.
\VS{10}For as the
rain cometh
down, and the
snow from
heaven, and
returneth not thither, but
watereth the
earth, and maketh it bring
forth and
bud, that it may
give
seed to the
sower, and
bread to the
eater:
\VS{11}So shall my
word be that goeth
forth out of my
mouth: it shall not
return unto me
void, but it shall
accomplish that which I
please, and it shall
prosper
{\ADD{in the thing}} whereto I
sent it.
\VS{12}For ye shall go
out with
joy, and be led
forth with
peace: the
mountains and the
hills shall break
forth
before you into
singing, and all the
trees of the
field shall
clap
{\ADD{their}}
hands.
\VS{13}Instead of the
thorn shall come
up the fir
tree, and instead of the
brier shall come
up the myrtle
tree: and it shall be to the
{\ND{
LORD}} for a
name, for an
everlasting
sign
{\ADD{that}} shall not be cut
off.

\par }\Chap{56}{\PP \VerseOne{1}Thus
saith the
{\ND{
LORD}},
Keep ye
judgment, and
do
justice: for my
salvation
{\ADD{is}}
near to
come, and my
righteousness to be
revealed.\FTNT{*}{{\FR 56.1 }
judgment: or, equity}
\VS{2}Blessed
{\ADD{is}} the
man
{\ADD{that}}
doeth this, and the
son of
man
{\ADD{that}} layeth
hold on it; that
keepeth the
sabbath from
polluting it, and
keepeth his
hand from
doing any
evil.
\par }{\PP \VS{3}¶ Neither let the
son of the
stranger, that hath
joined himself to the
{\ND{
LORD}},
speak,
saying, The
{\ND{
LORD}} hath
utterly
separated me from his
people: neither let the
eunuch
say, Behold, I
{\ADD{am}} a
dry
tree.
\VS{4}For thus
saith the
{\ND{
LORD}} unto the
eunuchs that
keep my
sabbaths, and
choose
{\ADD{the things}} that
please me, and take
hold of my
covenant;
\VS{5}Even unto them will I
give in mine
house and within my
walls a
place and a
name
better than of
sons and of
daughters: I will
give them an
everlasting
name, that shall not be cut
off.
\VS{6}Also the
sons of the
stranger, that
join themselves to the
{\ND{
LORD}}, to
serve him, and to
love the
name of the
{\ND{
LORD}}, to be his
servants, every one that
keepeth the
sabbath from
polluting it, and taketh
hold of my
covenant;
\VS{7}Even them will I
bring to my
holy
mountain, and make them
joyful in my
house of
prayer: their burnt
offerings and their
sacrifices
{\ADD{shall be}}
accepted upon mine
altar; for mine
house shall be
called an
house of
prayer for all
people.
\VS{8}The
Lord
{\ND{
GOD}} which
gathereth the
outcasts of
Israel
saith, Yet will I
gather
{\ADD{others}} to him, beside those that are
gathered unto him.\FTNT{†}{{\FR 56.8 }
beside…: Heb. to his gathered}
\par }{\PP \VS{9}¶ All ye
beasts of the
field,
come to
devour,
{\ADD{yea}}, all ye
beasts in the
forest.
\VS{10}His
watchmen
{\ADD{are}}
blind: they are all
ignorant, they
{\ADD{are}} all
dumb
dogs, they
cannot
bark;
sleeping, lying
down,
loving to
slumber.\FTNT{‡}{{\FR 56.10 }
sleeping: or, dreaming, or, talking in their sleep}
\VS{11}Yea,
{\ADD{they are}}
greedy
dogs
{\ADD{which}}
can never
have
enough, and they
{\ADD{are}}
shepherds
{\ADD{that}} cannot
understand: they all
look to their own
way, every
one for his
gain, from his
quarter.\FTNT{§}{{\FR 56.11 }
greedy: Heb. strong of appetite}\FTNT{**}{{\FR 56.11 }
can…: Heb. know not to be satisfied}
\VS{12}Come ye,
{\ADD{say they}}, I will
fetch
wine, and we will
fill ourselves with strong
drink; and to
morrow shall be as this
day,
{\ADD{and}}
much
more
abundant.

\par }\Chap{57}{\PP \VerseOne{1}The
righteous
perisheth, and no
man
layeth
{\ADD{it}} to
heart: and
merciful
men
{\ADD{are}} taken
away, none
considering that the
righteous is taken
away
from the
evil
{\ADD{to come}}.\FTNT{*}{{\FR 57.1 }
merciful…: Heb. men of kindness, or, godliness}\FTNT{†}{{\FR 57.1 }
from…: or, from that which is evil}
\VS{2}He shall
enter into
peace: they shall
rest in their
beds,
{\ADD{each one}}
walking
{\ADD{in}} his
uprightness.\FTNT{‡}{{\FR 57.2 }
enter into: or, go in}\FTNT{§}{{\FR 57.2 }
in his…: or, before him}
\par }{\PP \VS{3}¶ But draw
near hither, ye
sons of the
sorceress, the
seed of the
adulterer and the
whore.
\VS{4}Against whom do ye
sport yourselves? against whom make ye a
wide
mouth,
{\ADD{and}} draw
out the
tongue?
{\ADD{are}} ye not
children of
transgression, a
seed of
falsehood,
\VS{5}Enflaming yourselves with
idols under every
green
tree,
slaying the
children in the
valleys under the
clifts of the
rocks?\FTNT{**}{{\FR 57.5 }
with…: or, among the oaks}
\VS{6}Among the
smooth
{\ADD{stones}} of the
stream
{\ADD{is}} thy
portion; they, they
{\ADD{are}} thy
lot: even to them hast thou
poured a drink
offering, thou hast
offered a meat
offering. Should I receive
comfort in these?
\VS{7}Upon a
lofty and
high
mountain hast thou
set thy
bed: even thither wentest thou
up to
offer
sacrifice.
\VS{8}Behind the
doors also and the
posts hast thou set
up thy
remembrance: for thou hast
discovered
{\ADD{thyself to another}} than me, and art gone
up; thou hast
enlarged thy
bed, and
made thee
{\ADD{a covenant}} with them; thou
lovedst their
bed
where thou
sawest
{\ADD{it}}.\FTNT{††}{{\FR 57.8 }
made…: or, hewed it for thyself larger than theirs}\FTNT{‡‡}{{\FR 57.8 }
where…: or, thou providedst room}
\VS{9}And thou
wentest to the
king with
ointment, and didst
increase thy
perfumes, and didst
send thy
messengers far
off, and didst
debase
{\ADD{thyself even}} unto
hell.\FTNT{§§}{{\FR 57.9 }
thou…: or, thou respectedst}
\VS{10}Thou art
wearied in the
greatness of thy
way;
{\ADD{yet}}
saidst thou not, There is no
hope: thou hast
found the
life of thine
hand; therefore thou wast not
grieved.\FTNT{***}{{\FR 57.10 }
life: or, living}
\VS{11}And of whom hast thou been
afraid or
feared, that thou hast
lied, and hast not
remembered me, nor
laid
{\ADD{it}} to thy
heart? have not I held my
peace even of
old, and thou
fearest me not?
\VS{12}I will
declare thy
righteousness, and thy
works; for they shall not
profit thee.
\par }{\PP \VS{13}¶ When thou
criest, let thy
companies
deliver thee; but the
wind shall carry them all
away;
vanity shall
take
{\ADD{them:}} but he that putteth his
trust in me shall
possess the
land, and shall
inherit my
holy
mountain;
\VS{14}And shall
say, Cast ye
up, cast ye
up,
prepare the
way, take
up the
stumblingblock out of the
way of my
people.
\VS{15}For thus
saith the
high and lofty
One that
inhabiteth
eternity, whose
name
{\ADD{is}}
Holy; I
dwell in the
high and
holy
{\ADD{place}}, with him also
{\ADD{that is}} of a
contrite and
humble
spirit, to
revive the
spirit of the
humble, and to
revive the
heart of the contrite
ones.
\VS{16}For I will not
contend for
ever, neither will I be
always
wroth: for the
spirit should
fail
before me, and the
souls
{\ADD{which}} I have
made.
\par }{\PP \VS{17}For the
iniquity of his
covetousness was I
wroth, and
smote him: I
hid me, and was
wroth, and he went
on
frowardly in the
way of his
heart.\FTNT{†††}{{\FR 57.17 }
frowardly: Heb. turning away}
\VS{18}I have
seen his
ways, and will
heal him: I will
lead him also, and
restore
comforts unto him and to his
mourners.
\VS{19}I
create the
fruit of the
lips;
Peace,
peace to
{\ADD{him that is}} far
off, and to
{\ADD{him that is}}
near,
saith the
{\ND{
LORD}}; and I will
heal him.
\VS{20}But the
wicked
{\ADD{are}} like the
troubled
sea, when it
cannot
rest, whose
waters cast
up
mire and
dirt.
\VS{21}{\ADD{There is}} no
peace,
saith my
God, to the
wicked.

\par }\Chap{58}{\PP \VerseOne{1}Cry
aloud,
spare not,
lift up thy
voice like a
trumpet, and
shew my
people their
transgression, and the
house of
Jacob their
sins.\FTNT{*}{{\FR 58.1 }
aloud: Heb. with the throat}
\VS{2}Yet they
seek me
daily, and
delight to
know my
ways, as a
nation that
did
righteousness, and
forsook not the
ordinance of their
God: they
ask of me the
ordinances of
justice; they take
delight in
approaching to
God.
\par }{\PP \VS{3}¶ Wherefore have we
fasted,
{\ADD{say they}}, and thou
seest not?
{\ADD{wherefore}} have we
afflicted our
soul, and thou takest no
knowledge? Behold, in the
day of your
fast ye
find
pleasure, and
exact all your
labours.\FTNT{†}{{\FR 58.3 }
labours: or, things wherewith ye grieve others: Heb. griefs}
\VS{4}Behold, ye
fast for
strife and
debate, and to
smite with the
fist of
wickedness: ye shall not
fast as
{\ADD{ye do this}}
day, to make your
voice to be
heard on
high.\FTNT{‡}{{\FR 58.4 }
ye shall…: or, ye fast not as this day}
\VS{5}Is it such a
fast that I have
chosen? a
day for a
man to
afflict his
soul?
{\ADD{is it}} to bow
down his
head as a
bulrush, and to
spread
sackcloth and
ashes
{\ADD{under him}}? wilt thou
call
this a
fast, and an
acceptable
day to the
{\ND{
LORD}}?\FTNT{§}{{\FR 58.5 }
a day…: or, to afflict his soul for a day?}
\VS{6}{\ADD{Is}} not this the
fast that I have
chosen? to
loose the
bands of
wickedness, to
undo the
heavy
burdens, and to let the
oppressed
go
free, and that ye
break every
yoke?\FTNT{**}{{\FR 58.6 }
the heavy…: Heb. the bundles of the yoke}\FTNT{††}{{\FR 58.6 }
oppressed: Heb. broken}
\VS{7}{\ADD{Is it}} not to
deal thy
bread to the
hungry, and that thou
bring the
poor that are cast
out to thy
house? when thou
seest the
naked, that thou
cover him; and that thou
hide not thyself from thine own
flesh?\FTNT{‡‡}{{\FR 58.7 }
cast…: or, afflicted}
\par }{\PP \VS{8}¶ Then shall thy
light break
forth as the
morning, and thine
health shall spring
forth
speedily: and thy
righteousness shall
go
before thee; the
glory of the
{\ND{
LORD}} shall be thy
rereward.\FTNT{§§}{{\FR 58.8 }
be…: Heb. gather thee up}
\VS{9}Then shalt thou
call, and the
{\ND{
LORD}} shall
answer; thou shalt
cry, and he shall
say, Here I
{\ADD{am}}. If thou take
away from the
midst of thee the
yoke, the putting
forth of the
finger, and
speaking
vanity;
\VS{10}And
{\ADD{if}} thou draw
out thy
soul to the
hungry, and
satisfy the
afflicted
soul; then shall thy
light
rise in
obscurity, and thy
darkness
{\ADD{be}} as the
noonday:
\VS{11}And the
{\ND{
LORD}} shall
guide thee
continually, and
satisfy thy
soul in
drought, and make
fat thy
bones: and thou shalt be like a
watered
garden, and like a
spring of
water, whose
waters
fail not.\FTNT{***}{{\FR 58.11 }
drought: Heb. droughts}\FTNT{†††}{{\FR 58.11 }
fail: Heb. lie, or, deceive}
\VS{12}And
{\ADD{they that shall be}} of thee shall
build the
old waste
places: thou shalt raise
up the
foundations of
many
generations; and thou shalt be
called, The
repairer of the
breach, The
restorer of
paths to dwell
in.
\par }{\PP \VS{13}¶ If thou turn
away thy
foot from the
sabbath,
{\ADD{from}}
doing thy
pleasure on my
holy
day; and
call the
sabbath a
delight, the
holy of the
{\ND{
LORD}},
honourable; and shalt
honour him, not
doing thine own
ways, nor
finding thine own
pleasure, nor
speaking
{\ADD{thine own}}
words:
\VS{14}Then shalt thou
delight thyself in the
{\ND{
LORD}}; and I will cause thee to
ride upon the high
places of the
earth, and
feed thee with the
heritage of
Jacob thy
father: for the
mouth of the
{\ND{
LORD}} hath
spoken
{\ADD{it}}.

\par }\Chap{59}{\PP \VerseOne{1}Behold, the
{\ND{
LORD’s}}
hand is not
shortened, that it cannot
save; neither his
ear
heavy, that it cannot
hear:
\VS{2}But your
iniquities have
separated
between you and your
God, and your
sins have
hid
{\ADD{his}}
face from you, that he will not
hear.\FTNT{*}{{\FR 59.2 }
have hid: or, have made him hide}
\VS{3}For your
hands are
defiled with
blood, and your
fingers with
iniquity; your
lips have
spoken
lies, your
tongue hath
muttered
perverseness.
\VS{4}None
calleth for
justice, nor
{\ADD{any}}
pleadeth for
truth: they
trust in
vanity, and
speak
lies; they
conceive
mischief, and bring
forth
iniquity.
\VS{5}They
hatch
cockatrice’
eggs, and
weave the
spider’s
web: he that
eateth of their
eggs
dieth, and that which is
crushed breaketh
out into a
viper.\FTNT{†}{{\FR 59.5 }
cockatrice’: or, adder’s}\FTNT{‡}{{\FR 59.5 }
crushed…: or, sprinkled is as if there brake out a viper}
\VS{6}Their
webs shall not become
garments, neither shall they
cover themselves with their
works: their
works
{\ADD{are}}
works of
iniquity, and the
act of
violence
{\ADD{is}} in their
hands.
\VS{7}Their
feet
run to
evil, and they make
haste to
shed
innocent
blood: their
thoughts
{\ADD{are}}
thoughts of
iniquity;
wasting and
destruction
{\ADD{are}} in their
paths.\FTNT{§}{{\FR 59.7 }
destruction: Heb. breaking}
\VS{8}The
way of
peace they
know not; and
{\ADD{there is}} no
judgment in their
goings: they have made them
crooked
paths: whosoever
goeth therein shall not
know
peace.\FTNT{**}{{\FR 59.8 }
judgment: or, right}
\par }{\PP \VS{9}¶ Therefore is
judgment
far from us, neither doth
justice
overtake us: we
wait for
light, but behold
obscurity; for
brightness,
{\ADD{but}} we
walk in
darkness.
\VS{10}We
grope for the
wall like the
blind, and we
grope as if
{\ADD{we had}} no
eyes: we
stumble at
noonday as in the
night;
{\ADD{we are}} in
desolate places as
dead
{\ADD{men}}.
\VS{11}We
roar all like
bears, and
mourn
sore like
doves: we
look for
judgment, but
{\ADD{there is}} none; for
salvation,
{\ADD{but}} it is far
off from us.
\VS{12}For our
transgressions are
multiplied before thee, and our
sins
testify against us: for our
transgressions
{\ADD{are}} with us; and
{\ADD{as for}} our
iniquities, we
know them;
\VS{13}In
transgressing and
lying against the
{\ND{
LORD}}, and departing
away
from our
God,
speaking
oppression and
revolt,
conceiving and
uttering from the
heart
words of
falsehood.
\VS{14}And
judgment is turned
away
backward, and
justice
standeth afar
off: for
truth is
fallen in the
street, and
equity
cannot
enter.
\VS{15}Yea,
truth
faileth; and he
{\ADD{that}}
departeth from
evil maketh himself a
prey: and the
{\ND{
LORD}}
saw
{\ADD{it}}, and it
displeased him that
{\ADD{there was}} no
judgment.\FTNT{††}{{\FR 59.15 }
maketh…: or, is accounted mad}\FTNT{‡‡}{{\FR 59.15 }
it displeased…: Heb. it was evil in his eyes}
\par }{\PP \VS{16}¶ And he
saw that
{\ADD{there was}} no
man, and
wondered that
{\ADD{there was}} no
intercessor: therefore his
arm brought
salvation unto him; and his
righteousness, it
sustained him.
\VS{17}For he put
on
righteousness as a
breastplate, and an
helmet of
salvation upon his
head; and he put
on the
garments of
vengeance
{\ADD{for}}
clothing, and was
clad with
zeal as a
cloke.
\VS{18}According
to
{\ADD{their}}
deeds, accordingly he will
repay,
fury to his
adversaries,
recompence to his
enemies; to the
islands he will
repay
recompence.\FTNT{§§}{{\FR 59.18 }
deeds: Heb. recompences}
\VS{19}So shall they
fear the
name of the
{\ND{
LORD}} from the
west, and his
glory from the
rising of the
sun. When the
enemy shall come
in like a
flood, the
Spirit of the
{\ND{
LORD}} shall lift up a
standard against him.\FTNT{***}{{\FR 59.19 }
lift…: or, put him to flight}
\par }{\PP \VS{20}¶ And the
Redeemer shall
come to
Zion, and unto them that
turn from
transgression in
Jacob,
saith the
{\ND{
LORD}}.
\VS{21}As for me, this
{\ADD{is}} my
covenant with them,
saith the
{\ND{
LORD}}; My
spirit that
{\ADD{is}} upon thee, and my
words which I have
put in thy
mouth, shall not
depart out of thy
mouth, nor out of the
mouth of thy
seed, nor out of the
mouth of thy
seed’s
seed,
saith the
{\ND{
LORD}}, from henceforth and
for
ever.

\par }\Chap{60}{\PP \VerseOne{1}Arise,
shine; for thy
light is
come, and the
glory of the
{\ND{
LORD}} is
risen upon thee.\FTNT{*}{{\FR 60.1 }
shine…: or, be enlightened; for thy light cometh}
\VS{2}For, behold, the
darkness shall
cover the
earth, and gross
darkness the
people: but the
{\ND{
LORD}} shall
arise upon thee, and his
glory shall be
seen upon thee.
\VS{3}And the
Gentiles shall
come to thy
light, and
kings to the
brightness of thy
rising.
\VS{4}Lift
up thine
eyes round
about, and
see: all they gather themselves
together, they
come to thee: thy
sons shall
come from
far, and thy
daughters shall be
nursed at
{\ADD{thy}}
side.
\VS{5}Then thou shalt
see, and flow
together, and thine
heart shall
fear, and be
enlarged; because the
abundance of the
sea shall be
converted unto thee, the
forces of the
Gentiles shall
come unto thee.\FTNT{†}{{\FR 60.5 }
abundance…: or, noise of the sea shall be turned toward thee}\FTNT{‡}{{\FR 60.5 }
forces: or, wealth}
\VS{6}The
multitude of
camels shall
cover thee, the
dromedaries of
Midian and
Ephah; all they from
Sheba shall
come: they shall
bring
gold and
incense; and they shall shew
forth the
praises of the
{\ND{
LORD}}.
\VS{7}All the
flocks of
Kedar shall be gathered
together unto thee, the
rams of
Nebaioth shall
minister unto thee: they shall come
up with
acceptance on mine
altar, and I will
glorify the
house of my
glory.
\VS{8}Who
{\ADD{are}} these
{\ADD{that}}
fly as a
cloud, and as the
doves to their
windows?
\par }{\PP \VS{9}Surely the
isles shall
wait for me, and the
ships of
Tarshish
first, to
bring thy
sons from
far, their
silver and their
gold with them, unto the
name of the
{\ND{
LORD}} thy
God, and to the Holy
One of
Israel, because he hath
glorified thee.
\VS{10}And the
sons of
strangers shall build
up thy
walls, and their
kings shall
minister unto thee: for in my
wrath I
smote thee, but in my
favour have I had
mercy on thee.
\VS{11}Therefore thy
gates shall be
open
continually; they shall not be
shut
day nor
night; that
{\ADD{men}} may
bring unto thee the
forces of the
Gentiles, and
{\ADD{that}} their
kings
{\ADD{may be}}
brought.\FTNT{§}{{\FR 60.11 }
forces: or, wealth}
\VS{12}For the
nation and
kingdom that will not
serve thee shall
perish; yea,
{\ADD{those}}
nations shall be
utterly
wasted.
\VS{13}The
glory of
Lebanon shall
come unto thee, the fir
tree, the pine
tree, and the
box
together, to
beautify the
place of my
sanctuary; and I will make the
place of my
feet
glorious.
\VS{14}The
sons also of them that
afflicted thee shall
come
bending unto thee; and all they that
despised thee shall bow themselves
down at the
soles of thy
feet; and they shall
call thee, The
city of the
{\ND{
LORD}}, The
Zion of the Holy
One of
Israel.
\par }{\PP \VS{15}Whereas thou hast been
forsaken and
hated, so that no man went
through
{\ADD{thee}}, I will
make thee an
eternal
excellency, a
joy of
many
generations.
\VS{16}Thou shalt also
suck the
milk of the
Gentiles, and shalt
suck the
breast of
kings: and thou shalt
know that I the
{\ND{
LORD}}
{\ADD{am}} thy
Saviour and thy
Redeemer, the mighty
One of
Jacob.
\VS{17}For
brass I will
bring
gold, and for
iron I will
bring
silver, and for
wood
brass, and for
stones
iron: I will also
make thy
officers
peace, and thine
exactors
righteousness.
\VS{18}Violence shall no more be
heard in thy
land,
wasting nor
destruction within thy
borders; but thou shalt
call thy
walls
Salvation, and thy
gates
Praise.
\VS{19}The
sun shall be no more thy
light by
day; neither for
brightness shall the
moon give
light unto thee: but the
{\ND{
LORD}} shall be unto thee an
everlasting
light, and thy
God thy
glory.
\VS{20}Thy
sun shall no more go
down; neither shall thy
moon
withdraw itself: for the
{\ND{
LORD}} shall be thine
everlasting
light, and the
days of thy
mourning shall be
ended.
\VS{21}Thy
people also
{\ADD{shall be}} all
righteous: they shall
inherit the
land for
ever, the
branch of my
planting, the
work of my
hands, that I may be
glorified.
\VS{22}A little
one shall become a
thousand, and a small
one a
strong
nation: I the
{\ND{
LORD}} will
hasten it in his
time.

\par }\Chap{61}{\PP \VerseOne{1}The
Spirit of the
Lord
{\ND{
GOD}}
{\ADD{is}} upon me; because the
{\ND{
LORD}} hath
anointed me to preach good
tidings unto the
meek; he hath
sent me to bind
up the
brokenhearted, to
proclaim
liberty to the
captives, and the opening of the
prison to
{\ADD{them that are}}
bound;
\VS{2}To
proclaim the
acceptable
year of the
{\ND{
LORD}}, and the
day of
vengeance of our
God; to
comfort all that
mourn;
\VS{3}To
appoint unto them that
mourn in
Zion, to
give unto them
beauty for
ashes, the
oil of
joy for
mourning, the
garment of
praise for the
spirit of
heaviness; that they might be
called
trees of
righteousness, the
planting of the
{\ND{
LORD}}, that he might be
glorified.
\par }{\PP \VS{4}¶ And they shall
build the
old
wastes, they shall raise
up the
former
desolations, and they shall
repair the
waste
cities, the
desolations of
many
generations.
\VS{5}And
strangers shall
stand and
feed your
flocks, and the
sons of the
alien
{\ADD{shall be}} your
plowmen and your
vinedressers.
\VS{6}But ye shall be
named the
Priests of the
{\ND{
LORD}}:
{\ADD{men}} shall
call you the
Ministers of our
God: ye shall
eat the
riches of the
Gentiles, and in their
glory shall ye
boast yourselves.
\par }{\PP \VS{7}¶ For your
shame
{\ADD{ye shall have}}
double; and
{\ADD{for}}
confusion they shall
rejoice in their
portion: therefore in their
land they shall
possess the
double:
everlasting
joy shall be unto them.
\VS{8}For I the
{\ND{
LORD}}
love
judgment, I
hate
robbery for burnt
offering; and I will
direct their
work in
truth, and I will
make an
everlasting
covenant with them.
\VS{9}And their
seed shall be
known among the
Gentiles, and their
offspring
among the
people: all that
see them shall
acknowledge them, that they
{\ADD{are}} the
seed
{\ADD{which}} the
{\ND{
LORD}} hath
blessed.
\par }{\PP \VS{10}I will
greatly
rejoice in the
{\ND{
LORD}}, my
soul shall be
joyful in my
God; for he hath
clothed me with the
garments of
salvation, he hath
covered me with the
robe of
righteousness, as a
bridegroom
decketh
{\ADD{himself}} with
ornaments, and as a
bride
adorneth
{\ADD{herself}} with her
jewels.\FTNT{*}{{\FR 61.10 }
decketh: Heb. decketh as a priest}
\VS{11}For as the
earth bringeth
forth her
bud, and as the
garden causeth the things that are
sown in it to spring
forth; so the
Lord
{\ND{
GOD}} will cause
righteousness and
praise to spring
forth before all the
nations.

\par }\Chap{62}{\PP \VerseOne{1}For
Zion’s sake will I not hold my
peace, and for
Jerusalem’s sake I will not
rest, until the
righteousness thereof go
forth as
brightness, and the
salvation thereof as a
lamp
{\ADD{that}}
burneth.
\VS{2}And the
Gentiles shall
see thy
righteousness, and all
kings thy
glory: and thou shalt be
called by a
new
name, which the
mouth of the
{\ND{
LORD}} shall
name.
\VS{3}Thou shalt also be a
crown of
glory in the
hand of the
{\ND{
LORD}}, and a
royal
diadem in the
hand of thy
God.
\VS{4}Thou shalt no more be
termed
Forsaken; neither shall thy
land any more be
termed
Desolate: but thou shalt be
called
Hephzi-bah, and thy
land
Beulah: for the
{\ND{
LORD}}
delighteth in thee, and thy
land shall be
married.\FTNT{*}{{\FR 62.4 }
Hephzi-bah: that is, My delight is in her}\FTNT{†}{{\FR 62.4 }
Beulah: that is, Married}
\par }{\PP \VS{5}¶ For
{\ADD{as}} a young
man
marrieth a
virgin,
{\ADD{so}} shall thy
sons
marry thee: and
{\ADD{as}} the
bridegroom
rejoiceth over the
bride,
{\ADD{so}} shall thy
God
rejoice over thee.\FTNT{‡}{{\FR 62.5 }
as the…: Heb. with the joy of the bridegroom}
\par }{\PP \VS{6}I have
set
watchmen upon thy
walls, O
Jerusalem,
{\ADD{which}} shall
never hold their
peace
day nor
night: ye that make
mention of the
{\ND{
LORD}}, keep not
silence,\FTNT{§}{{\FR 62.6 }
make…: or, are the
{\ND{LORD’s}} remembrancers}
\VS{7}And
give him no
rest, till he
establish, and till he
make
Jerusalem a
praise in the
earth.\FTNT{**}{{\FR 62.7 }
rest: Heb. silence}
\VS{8}The
{\ND{
LORD}} hath
sworn by his right
hand, and by the
arm of his
strength, Surely I
will no more
give thy
corn
{\ADD{to be}}
meat for thine
enemies; and the
sons of the
stranger shall not
drink thy
wine, for the which thou hast
laboured:\FTNT{††}{{\FR 62.8 }
Surely…: Heb. If I give, etc}
\VS{9}But they that have
gathered it shall
eat it, and
praise the
{\ND{
LORD}}; and they that have brought it
together shall
drink it in the
courts of my
holiness.
\par }{\PP \VS{10}¶ Go
through, go
through the
gates;
prepare ye the
way of the
people; cast
up, cast
up the
highway; gather
out the
stones; lift
up a
standard for the
people.
\VS{11}Behold, the
{\ND{
LORD}} hath
proclaimed unto the
end of the
world,
Say ye to the
daughter of
Zion, Behold, thy
salvation
cometh; behold, his
reward
{\ADD{is}} with him, and his
work
before him.\FTNT{‡‡}{{\FR 62.11 }
work: or, recompence}
\VS{12}And they shall
call them, The
holy
people, The
redeemed of the
{\ND{
LORD}}: and thou shalt be
called, Sought
out, A
city not
forsaken.

\par }\Chap{63}{\PP \VerseOne{1}Who
{\ADD{is}} this that
cometh from
Edom, with
dyed
garments from
Bozrah? this
{\ADD{that is}}
glorious in his
apparel,
travelling in the
greatness of his
strength? I that
speak in
righteousness,
mighty to
save.\FTNT{*}{{\FR 63.1 }
glorious: Heb. decked}
\VS{2}Wherefore
{\ADD{art thou}}
red in thine
apparel, and thy
garments like him that
treadeth in the
winefat?
\VS{3}I have
trodden the
winepress alone; and of the
people
{\ADD{there was}}
none with me: for I will
tread them in mine
anger, and
trample them in my
fury; and their
blood shall be
sprinkled upon my
garments, and I will
stain all my
raiment.
\VS{4}For the
day of
vengeance
{\ADD{is}} in mine
heart, and the
year of my
redeemed is
come.
\VS{5}And I
looked, and
{\ADD{there was}} none to
help; and I
wondered that
{\ADD{there was}} none to
uphold: therefore mine own
arm brought
salvation unto me; and my
fury, it
upheld me.
\VS{6}And I will tread
down the
people in mine
anger, and make them
drunk in my
fury, and I will bring
down their
strength to the
earth.
\par }{\PP \VS{7}¶ I will
mention the
lovingkindnesses of the
{\ND{
LORD}},
{\ADD{and}} the
praises of the
{\ND{
LORD}}, according to all that the
{\ND{
LORD}} hath
bestowed on us, and the
great
goodness toward the
house of
Israel, which he hath
bestowed on them according to his
mercies, and according to the
multitude of his
lovingkindnesses.
\VS{8}For he
said, Surely they
{\ADD{are}} my
people,
children
{\ADD{that}} will not
lie: so he was their
Saviour.
\VS{9}In all their
affliction he was
afflicted, and the
angel of his
presence
saved them: in his
love and in his
pity he
redeemed them; and he
bare them, and
carried them all the
days of
old.
\par }{\PP \VS{10}¶ But they
rebelled, and
vexed his
holy
Spirit: therefore he was
turned to be their
enemy,
{\ADD{and}} he
fought against them.
\VS{11}Then he
remembered the
days of
old,
Moses,
{\ADD{and}} his
people,
{\ADD{saying}}, Where
{\ADD{is}} he that brought them
up out of the
sea with the
shepherd of his
flock? where
{\ADD{is}} he that
put his
holy
Spirit
within him?\FTNT{†}{{\FR 63.11 }
shepherd: or, shepherds}
\VS{12}That
led
{\ADD{them}} by the right
hand of
Moses with his
glorious
arm,
dividing the
water
before them, to
make himself an
everlasting
name?
\VS{13}That
led them through the
deep, as an
horse in the
wilderness,
{\ADD{that}} they should not
stumble?
\VS{14}As a
beast goeth
down into the
valley, the
Spirit of the
{\ND{
LORD}} caused him to
rest: so didst thou
lead thy
people, to
make thyself a
glorious
name.
\par }{\PP \VS{15}¶ Look
down from
heaven, and
behold from the
habitation of thy
holiness and of thy
glory: where
{\ADD{is}} thy
zeal and thy
strength, the
sounding of thy
bowels and of thy
mercies toward me? are they
restrained?\FTNT{‡}{{\FR 63.15 }
the sounding: or, the multitude}
\VS{16}Doubtless thou
{\ADD{art}} our
father, though
Abraham be
ignorant of us, and
Israel
acknowledge us not: thou, O
{\ND{
LORD}},
{\ADD{art}} our
father, our
redeemer; thy
name
{\ADD{is}} from
everlasting.\FTNT{§}{{\FR 63.16 }
our redeemer…: or, our redeemer from everlasting is thy name}
\par }{\PP \VS{17}¶ O
{\ND{
LORD}}, why hast thou made us to
err from thy
ways,
{\ADD{and}}
hardened our
heart from thy
fear?
Return for thy
servants’ sake, the
tribes of thine
inheritance.
\VS{18}The
people of thy
holiness have
possessed
{\ADD{it}} but a little
while: our
adversaries have trodden
down thy
sanctuary.
\VS{19}We are
{\ADD{thine:}} thou
never barest
rule over them; they were not
called by thy
name.\FTNT{**}{{\FR 63.19 }
they…: Heb. thy name was not called upon them}

\par }\Chap{64}{\PP \VerseOne{1}Oh that thou wouldest
rend the
heavens, that thou wouldest come
down, that the
mountains might flow
down at thy
presence,
\VS{2}As
{\ADD{when}} the
melting
fire
burneth, the
fire causeth the
waters to
boil, to make thy
name
known to thine
adversaries,
{\ADD{that}} the
nations may
tremble at thy
presence!\FTNT{*}{{\FR 64.2 }
the melting…: Heb. the fire of meltings}
\VS{3}When thou
didst terrible
things
{\ADD{which}} we
looked not for, thou camest
down, the
mountains flowed
down at thy
presence.
\VS{4}For since the beginning of the
world
{\ADD{men}} have not
heard, nor perceived by the
ear, neither hath the
eye
seen, O
God,
beside thee,
{\ADD{what}} he hath
prepared for him that
waiteth for him.\FTNT{†}{{\FR 64.4 }
seen…: or, seen a God beside thee, which doeth so for him, etc}
\VS{5}Thou
meetest him that
rejoiceth and
worketh
righteousness,
{\ADD{those that}}
remember thee in thy
ways: behold, thou art
wroth; for we have
sinned: in those is
continuance, and we shall be
saved.
\par }{\PP \VS{6}But we are all as an
unclean
{\ADD{thing}}, and all our
righteousnesses
{\ADD{are}} as
filthy
rags; and we all do
fade as a
leaf; and our
iniquities, like the
wind, have taken us
away.
\VS{7}And
{\ADD{there is}} none that
calleth upon thy
name, that stirreth
up himself to take
hold of thee: for thou hast
hid thy
face from us, and hast
consumed us,
because of our
iniquities.\FTNT{‡}{{\FR 64.7 }
consumed: Heb. melted}\FTNT{§}{{\FR 64.7 }
because: Heb. by the hand}
\VS{8}But now, O
{\ND{
LORD}}, thou
{\ADD{art}} our
father; we
{\ADD{are}} the
clay, and thou our
potter; and we all
{\ADD{are}} the
work of thy
hand.
\par }{\PP \VS{9}¶ Be not
wroth very
sore, O
{\ND{
LORD}}, neither
remember
iniquity for
ever: behold,
see, we beseech thee, we
{\ADD{are}} all thy
people.
\VS{10}Thy
holy
cities are a
wilderness,
Zion is a
wilderness,
Jerusalem a
desolation.
\VS{11}Our
holy and our
beautiful
house, where our
fathers
praised thee, is burned
up with
fire: and all our pleasant
things are laid
waste.
\VS{12}Wilt thou
refrain thyself for these
{\ADD{things}}, O
{\ND{
LORD}}? wilt thou hold thy
peace, and
afflict us very
sore?

\par }\Chap{65}{\PP \VerseOne{1}I am
sought of
{\ADD{them that}}
asked
not
{\ADD{for me;}} I am
found of
{\ADD{them that}}
sought me
not: I
said, Behold me, behold me, unto a
nation
{\ADD{that}} was not
called by my
name.
\VS{2}I have spread
out my
hands all the
day unto a
rebellious
people, which
walketh in a
way
{\ADD{that was}} not
good,
after their own
thoughts;
\VS{3}A
people that provoketh me to
anger
continually to my
face; that
sacrificeth in
gardens, and burneth
incense upon altars of
brick;\FTNT{*}{{\FR 65.3 }
upon…: Heb. upon bricks}
\VS{4}Which
remain among the
graves, and
lodge in the
monuments, which
eat
swine’s
flesh, and
broth of
abominable
{\ADD{things is in}} their
vessels;\FTNT{†}{{\FR 65.4 }
broth: or, pieces}
\VS{5}Which
say,
Stand by thyself, come not
near to me; for I am
holier than thou. These
{\ADD{are}} a
smoke in my
nose, a
fire that
burneth all the
day.\FTNT{‡}{{\FR 65.5 }
nose: or, anger}
\VS{6}Behold,
{\ADD{it is}}
written
before me: I will not keep
silence, but will
recompense, even
recompense into their
bosom,
\VS{7}Your
iniquities, and the
iniquities of your
fathers
together,
saith the
{\ND{
LORD}}, which have burned
incense upon the
mountains, and
blasphemed me upon the
hills: therefore will I
measure their
former
work into their
bosom.
\par }{\PP \VS{8}¶ Thus
saith the
{\ND{
LORD}}, As the new
wine is
found in the
cluster, and
{\ADD{one}}
saith,
Destroy it not; for a
blessing
{\ADD{is}} in it: so will I
do for my
servants’ sakes, that I may not
destroy them all.
\VS{9}And I will bring
forth a
seed out of
Jacob, and out of
Judah an
inheritor of my
mountains: and mine
elect shall
inherit it, and my
servants shall
dwell there.
\VS{10}And
Sharon shall be a
fold of
flocks, and the
valley of
Achor a place for the
herds to lie down
in, for my
people that have
sought me.
\par }{\PP \VS{11}¶ But ye
{\ADD{are}} they that
forsake the
{\ND{
LORD}}, that
forget my
holy
mountain, that
prepare a
table for that
troop, and that
furnish the drink
offering unto that
number.\FTNT{§}{{\FR 65.11 }
troop: or, Gad}\FTNT{**}{{\FR 65.11 }
number: or, Meni, a pagan god}
\VS{12}Therefore will I
number you to the
sword, and ye shall all bow
down to the
slaughter: because when I
called, ye did not
answer; when I
spake, ye did not
hear; but
did
evil before mine
eyes, and did
choose
{\ADD{that}} wherein I
delighted not.
\VS{13}Therefore thus
saith the
Lord
{\ND{
GOD}}, Behold, my
servants shall
eat, but ye shall be
hungry: behold, my
servants shall
drink, but ye shall be
thirsty: behold, my
servants shall
rejoice, but ye shall be
ashamed:
\VS{14}Behold, my
servants shall
sing for
joy of
heart, but ye shall
cry for
sorrow of
heart, and shall
howl for
vexation of
spirit.\FTNT{††}{{\FR 65.14 }
vexation: Heb. breaking}
\VS{15}And ye shall
leave your
name for a
curse unto my
chosen: for the
Lord
{\ND{
GOD}} shall
slay thee, and
call his
servants by
another
name:
\VS{16}That he who
blesseth himself in the
earth shall
bless himself in the
God of
truth; and he that
sweareth in the
earth shall
swear by the
God of
truth; because the
former
troubles are
forgotten, and because they are
hid from mine
eyes.
\par }{\PP \VS{17}¶ For, behold, I
create
new
heavens and a
new
earth: and the
former shall not be
remembered, nor
come into
mind.\FTNT{‡‡}{{\FR 65.17 }
come…: Heb. come upon the heart}
\VS{18}But be ye
glad and
rejoice for
ever
{\ADD{in that}} which I
create: for, behold, I
create
Jerusalem a
rejoicing, and her
people a
joy.
\VS{19}And I will
rejoice in
Jerusalem, and
joy in my
people: and the
voice of
weeping shall be no more
heard in her, nor the
voice of
crying.
\VS{20}There shall be no more thence an
infant of
days, nor an old
man that hath not
filled his
days: for the
child shall
die an
hundred
years
old; but the
sinner
{\ADD{being}} an
hundred
years
old shall be
accursed.
\VS{21}And they shall
build
houses, and
inhabit
{\ADD{them;}} and they shall
plant
vineyards, and
eat the
fruit of them.
\VS{22}They shall not
build, and
another
inhabit; they shall not
plant, and
another
eat: for as the
days of a
tree
{\ADD{are}} the
days of my
people, and mine
elect shall long
enjoy the
work of their
hands.\FTNT{§§}{{\FR 65.22 }
shall long…: Heb. shall make them continue long, or, shall wear out}
\VS{23}They shall not
labour in
vain, nor bring
forth for
trouble; for they
{\ADD{are}} the
seed of the
blessed of the
{\ND{
LORD}}, and their
offspring with them.
\VS{24}And it shall come to pass, that before they
call, I will
answer; and while they are yet
speaking, I will
hear.
\VS{25}The
wolf and the
lamb shall
feed
together, and the
lion shall
eat
straw like the
bullock: and
dust
{\ADD{shall be}} the
serpent’s
meat. They shall not
hurt nor
destroy in all my
holy
mountain,
saith the
{\ND{
LORD}}.

\par }\Chap{66}{\PP \VerseOne{1}Thus
saith the
{\ND{
LORD}}, The
heaven
{\ADD{is}} my
throne, and the
earth
{\ADD{is}} my
footstool: where
{\ADD{is}} the
house that ye
build unto me? and where
{\ADD{is}} the
place of my
rest?
\VS{2}For all those
{\ADD{things}} hath mine
hand
made, and all those
{\ADD{things}} have been,
saith the
{\ND{
LORD}}: but to this
{\ADD{man}} will I
look,
{\ADD{even}} to
{\ADD{him that is}}
poor and of a
contrite
spirit, and
trembleth at my
word.
\VS{3}He that
killeth an
ox
{\ADD{is as if}} he
slew a
man; he that
sacrificeth a
lamb,
{\ADD{as if}} he cut
off a
dog’s
neck; he that
offereth an
oblation,
{\ADD{as if he offered}}
swine’s
blood; he that
burneth
incense,
{\ADD{as if}} he
blessed an
idol. Yea, they have
chosen their own
ways, and their
soul
delighteth in their
abominations.\FTNT{*}{{\FR 66.3 }
lamb: or, kid}\FTNT{†}{{\FR 66.3 }
burneth: Heb. maketh a memorial of}
\VS{4}I also will
choose their
delusions, and will
bring their
fears upon them; because when I
called, none did
answer; when I
spake, they did not
hear: but they
did
evil before mine
eyes, and
chose
{\ADD{that}} in which I
delighted not.\FTNT{‡}{{\FR 66.4 }
delusions: or, devices}
\par }{\PP \VS{5}¶
Hear the
word of the
{\ND{
LORD}}, ye that
tremble at his
word; Your
brethren that
hated you, that cast you
out for my
name’s sake,
said, Let the
{\ND{
LORD}} be
glorified: but he shall
appear to your
joy, and they shall be
ashamed.
\VS{6}A
voice of
noise from the
city, a
voice from the
temple, a
voice of the
{\ND{
LORD}} that
rendereth
recompence to his
enemies.
\VS{7}Before she
travailed, she brought
forth; before her
pain
came, she was
delivered of a man
child.
\VS{8}Who hath
heard such a thing? who hath
seen such things? Shall the
earth be made to bring
forth in
one
day?
{\ADD{or}} shall a
nation be
born at
once? for as soon
as
Zion
travailed, she brought
forth her
children.
\VS{9}Shall I bring to the
birth, and not cause to bring
forth?
saith the
{\ND{
LORD}}: shall
I cause to bring
forth, and
shut
{\ADD{the womb}}?
saith thy
God.\FTNT{§}{{\FR 66.9 }
not…: or, not beget?}
\VS{10}Rejoice ye with
Jerusalem, and be
glad with her, all ye that
love her:
rejoice for
joy with her, all ye that
mourn for her:
\VS{11}That ye may
suck, and be
satisfied with the
breasts of her
consolations; that ye may milk
out, and be
delighted with the
abundance of her
glory.\FTNT{**}{{\FR 66.11 }
abundance: or, brightness}
\VS{12}For thus
saith the
{\ND{
LORD}}, Behold, I will
extend
peace to her like a
river, and the
glory of the
Gentiles like a
flowing
stream: then shall ye
suck, ye shall be
borne upon
{\ADD{her}}
sides, and be
dandled upon
{\ADD{her}}
knees.
\VS{13}As
one whom his
mother
comforteth, so will I
comfort you; and ye shall be
comforted in
Jerusalem.
\VS{14}And when ye
see
{\ADD{this}}, your
heart shall
rejoice, and your
bones shall
flourish like an
herb: and the
hand of the
{\ND{
LORD}} shall be
known toward his
servants, and
{\ADD{his}}
indignation toward his
enemies.
\par }{\PP \VS{15}For, behold, the
{\ND{
LORD}} will
come with
fire, and with his
chariots like a
whirlwind, to
render his
anger with
fury, and his
rebuke with
flames of
fire.
\VS{16}For by
fire and by his
sword will the
{\ND{
LORD}}
plead with all
flesh: and the
slain of the
{\ND{
LORD}} shall be
many.
\VS{17}They that
sanctify themselves, and
purify themselves in the
gardens
behind
one
{\ADD{tree}} in the
midst,
eating
swine’s
flesh, and the
abomination, and the
mouse, shall be
consumed
together,
saith the
{\ND{
LORD}}.\FTNT{††}{{\FR 66.17 }
behind…: or, one after another}
\VS{18}For I
{\ADD{know}} their
works and their
thoughts: it shall
come, that I will
gather all
nations and
tongues; and they shall
come, and
see my
glory.
\VS{19}And I will
set a
sign among them, and I will
send those that
escape of them unto the
nations,
{\ADD{to}}
Tarshish,
Pul, and
Lud, that
draw the
bow,
{\ADD{to}}
Tubal, and
Javan,
{\ADD{to}} the
isles afar
off, that have not
heard my
fame, neither have
seen my
glory; and they shall
declare my
glory among the
Gentiles.
\VS{20}And they shall
bring all your
brethren
{\ADD{for}} an
offering unto the
{\ND{
LORD}} out of all
nations upon
horses, and in
chariots, and in
litters, and upon
mules, and upon swift
beasts, to my
holy
mountain
Jerusalem,
saith the
{\ND{
LORD}}, as the
children of
Israel
bring an
offering in a
clean
vessel into the
house of the
{\ND{
LORD}}.\FTNT{‡‡}{{\FR 66.20 }
litters: or, coaches}
\VS{21}And I will also
take of them for
priests
{\ADD{and}} for
Levites,
saith the
{\ND{
LORD}}.
\VS{22}For as the
new
heavens and the
new
earth, which I will
make, shall
remain
before me,
saith the
{\ND{
LORD}}, so shall your
seed and your
name
remain.
\VS{23}And it shall come to pass,
{\ADD{that}}
from one new
moon to
another, and
from one
sabbath to
another, shall all
flesh
come to
worship
before me,
saith the
{\ND{
LORD}}.\FTNT{§§}{{\FR 66.23 }
from one new…: Heb. from new moon to his new moon, and from sabbath to his sabbath}
\VS{24}And they shall go
forth, and
look upon the
carcases of the
men that have
transgressed against me: for their
worm shall not
die, neither shall their
fire be
quenched; and they shall be an
abhorring unto all
flesh.
\par }