\NormalFont\ShortTitle{Judges}
{\MT The Book of Judges

\par }\ChapOne{1}{\PP \VerseOne{1}Now
after the
death of
Joshua it came to pass, that the
children of
Israel
asked the
{\ND{
LORD}},
saying, Who shall go
up for us against the
Canaanites
first, to
fight against them?
\VS{2}And the
{\ND{
LORD}}
said,
Judah shall go
up: behold, I have
delivered the
land into his
hand.
\VS{3}And
Judah
said unto
Simeon his
brother, Come
up with me into my
lot, that we may
fight against the
Canaanites; and I likewise will
go with thee into thy
lot. So
Simeon
went with him.
\VS{4}And
Judah went
up; and the
{\ND{
LORD}}
delivered the
Canaanites and the
Perizzites into their
hand: and they
slew of them in
Bezek
ten
thousand
men.
\VS{5}And they
found
Adoni-bezek in
Bezek: and they
fought against him, and they
slew the
Canaanites and the
Perizzites.
\VS{6}But
Adoni-bezek
fled; and they
pursued
after him, and
caught him, and cut
off his
thumbs and his great
toes.
\VS{7}And
Adoni-bezek
said, Threescore and
ten
kings, having their
thumbs and their great
toes cut
off,
gathered
{\ADD{their meat}} under my
table: as I have
done, so
God hath
requited me. And they
brought him to
Jerusalem, and there he
died.\FTNT{*}{{\FR 1.7 }
their thumbs…: Heb. the thumbs of their hands and of their feet}\FTNT{†}{{\FR 1.7 }
gathered: or, gleaned}
\VS{8}Now the
children of
Judah had
fought against
Jerusalem, and had
taken it, and
smitten it with the
edge of the
sword, and
set the
city on
fire.
\par }{\PP \VS{9}¶ And
afterward the
children of
Judah went
down to
fight against the
Canaanites, that
dwelt in the
mountain, and in the
south, and in the
valley.\FTNT{‡}{{\FR 1.9 }
valley: or, low country}
\VS{10}And
Judah
went against the
Canaanites that
dwelt in
Hebron: (now the
name of
Hebron
before
{\ADD{was}}
Kirjath-arba:) and they
slew
Sheshai, and
Ahiman, and
Talmai.
\VS{11}And from thence he
went against the
inhabitants of
Debir: and the
name of
Debir
before
{\ADD{was}}
Kirjath-sepher:
\VS{12}And
Caleb
said, He that
smiteth
Kirjath-sepher, and
taketh it, to him will I
give
Achsah my
daughter to
wife.
\VS{13}And
Othniel the
son of
Kenaz,
Caleb’s
younger
brother,
took it: and he
gave him
Achsah his
daughter to
wife.
\VS{14}And it came to pass, when she
came
{\ADD{to him}}, that she
moved him to
ask of her
father a
field: and she
lighted from off
{\ADD{her}}
ass; and
Caleb
said unto her, What wilt thou?
\VS{15}And she
said unto him,
Give me a
blessing: for thou hast
given me a
south
land;
give me also
springs of
water. And
Caleb
gave her the
upper
springs and the
nether
springs.
\par }{\PP \VS{16}¶ And the
children of the
Kenite,
Moses’ father in
law, went
up out of the
city of palm
trees with the
children of
Judah into the
wilderness of
Judah, which
{\ADD{lieth}} in the
south of
Arad; and they
went and
dwelt
among the
people.
\VS{17}And
Judah
went with
Simeon his
brother, and they
slew the
Canaanites that
inhabited
Zephath, and utterly
destroyed it. And the
name of the
city was
called
Hormah.
\VS{18}Also
Judah
took
Gaza with the
coast thereof, and
Askelon with the
coast thereof, and
Ekron with the
coast thereof.
\VS{19}And the
{\ND{
LORD}} was with
Judah; and he drave
out
{\ADD{the inhabitants of}} the
mountain; but could
not drive
out the
inhabitants of the
valley, because they had
chariots of
iron.\FTNT{§}{{\FR 1.19 }
drave…: or, possessed the mountain}
\VS{20}And they
gave
Hebron unto
Caleb, as
Moses
said: and he
expelled thence the
three
sons of
Anak.
\par }{\PP \VS{21}And the
children of
Benjamin did not drive
out the
Jebusites that
inhabited
Jerusalem; but the
Jebusites
dwell with the
children of
Benjamin in
Jerusalem unto this
day.
\par }{\PP \VS{22}¶ And the
house of
Joseph, they also went
up against
Beth-el: and the
{\ND{
LORD}}
{\ADD{was}} with them.
\VS{23}And the
house of
Joseph sent to
descry
Beth-el. (Now the
name of the
city
before
{\ADD{was}}
Luz.)
\VS{24}And the
spies
saw a
man come
forth out of the
city, and they
said unto him,
Shew us, we pray thee, the
entrance into the
city, and we will
shew thee
mercy.
\VS{25}And when he
shewed them the
entrance into the
city, they
smote the
city with the
edge of the
sword; but they let
go the
man and all his
family.
\VS{26}And the
man
went into the
land of the
Hittites, and
built a
city, and
called the
name thereof
Luz: which
{\ADD{is}} the
name thereof unto this
day.
\par }{\PP \VS{27}¶ Neither did
Manasseh drive
out
{\ADD{the inhabitants of}}
Beth-shean and her
towns, nor
Taanach and her
towns, nor the
inhabitants of
Dor and her
towns, nor the
inhabitants of
Ibleam and her
towns, nor the
inhabitants of
Megiddo and her
towns: but the
Canaanites
would
dwell in that
land.
\VS{28}And it came to pass, when
Israel was
strong, that they
put the
Canaanites to
tribute, and did not
utterly drive them
out.
\par }{\PP \VS{29}¶ Neither did
Ephraim drive
out the
Canaanites that
dwelt in
Gezer; but the
Canaanites
dwelt in
Gezer
among them.
\par }{\PP \VS{30}¶ Neither did
Zebulun drive
out the
inhabitants of
Kitron, nor the
inhabitants of
Nahalol; but the
Canaanites
dwelt
among them, and became
tributaries.
\par }{\PP \VS{31}¶ Neither did
Asher drive
out the
inhabitants of
Accho, nor the
inhabitants of
Zidon, nor of
Ahlab, nor of
Achzib, nor of
Helbah, nor of
Aphik, nor of
Rehob:
\VS{32}But the
Asherites
dwelt
among the
Canaanites, the
inhabitants of the
land: for they did not drive them
out.
\par }{\PP \VS{33}¶ Neither did
Naphtali drive
out the
inhabitants of
Beth-shemesh, nor the
inhabitants of
Beth-anath; but he
dwelt
among the
Canaanites, the
inhabitants of the
land: nevertheless the
inhabitants of
Beth-shemesh and of
Beth-anath became
tributaries unto them.
\VS{34}And the
Amorites
forced the
children of
Dan into the
mountain: for they would not
suffer them to come
down to the
valley:
\VS{35}But the
Amorites
would
dwell in
mount
Heres in
Aijalon, and in
Shaalbim: yet the
hand of the
house of
Joseph
prevailed, so that they became
tributaries.\FTNT{**}{{\FR 1.35 }
prevailed: Heb. was heavy}
\VS{36}And the
coast of the
Amorites
{\ADD{was}} from the going
up to
Akrabbim, from the
rock, and
upward.\FTNT{††}{{\FR 1.36 }
the going…: or, Maaleh-acrabbim}

\par }\Chap{2}{\PP \VerseOne{1}And an
angel of the
{\ND{
LORD}} came
up from
Gilgal to
Bochim, and
said, I made you to go
up out of
Egypt, and have
brought you unto the
land which I
sware unto your
fathers; and I
said, I will
never
break my
covenant with you.\FTNT{*}{{\FR 2.1 }
angel: or, messenger}
\VS{2}And ye shall
make no
league with the
inhabitants of this
land; ye shall throw
down their
altars: but ye have not
obeyed my
voice: why have ye
done this?
\VS{3}Wherefore I also
said, I will not drive them
out from
before you; but they shall be
{\ADD{as thorns}} in your
sides, and their
gods shall be a
snare unto you.
\VS{4}And it came to pass, when the
angel of the
{\ND{
LORD}}
spake these
words unto all the
children of
Israel, that the
people lifted
up their
voice, and
wept.
\VS{5}And they
called the
name of that
place
Bochim: and they
sacrificed there unto the
{\ND{
LORD}}.\FTNT{†}{{\FR 2.5 }
Bochim: that is, Weepers}
\par }{\PP \VS{6}¶ And when
Joshua had let the
people
go, the
children of
Israel
went every
man unto his
inheritance to
possess the
land.
\VS{7}And the
people
served the
{\ND{
LORD}} all the
days of
Joshua, and all the
days of the
elders that
outlived
Joshua, who had
seen all the
great
works of the
{\ND{
LORD}}, that he
did for
Israel.\FTNT{‡}{{\FR 2.7 }
outlived: Heb. prolonged days after}
\VS{8}And
Joshua the
son of
Nun, the
servant of the
{\ND{
LORD}},
died,
{\ADD{being}} an
hundred and
ten
years
old.
\VS{9}And they
buried him in the
border of his
inheritance in
Timnath-heres, in the
mount of
Ephraim, on the north
side of the
hill
Gaash.\FTNT{§}{{\FR 2.9 }
Timnath-heres: also called, Timnath-serah}
\VS{10}And also all that
generation were
gathered unto their
fathers: and there
arose
another
generation
after them, which
knew not the
{\ND{
LORD}}, nor yet the
works which he had
done for
Israel.
\par }{\PP \VS{11}¶ And the
children of
Israel
did
evil in the
sight of the
{\ND{
LORD}}, and
served
Baalim:
\VS{12}And they
forsook the
{\ND{
LORD}}
God of their
fathers, which brought them
out of the
land of
Egypt, and
followed
other
gods, of the
gods of the
people that
{\ADD{were}} round
about them, and
bowed themselves unto them, and provoked the
{\ND{
LORD}} to
anger.
\VS{13}And they
forsook the
{\ND{
LORD}}, and
served
Baal and
Ashtaroth.
\par }{\PP \VS{14}¶ And the
anger of the
{\ND{
LORD}} was
hot against
Israel, and he
delivered them into the
hands of
spoilers that
spoiled them, and he
sold them into the
hands of their
enemies round
about, so that they
could not any
longer
stand
before their
enemies.
\VS{15}Whithersoever they went
out, the
hand of the
{\ND{
LORD}} was against them for
evil, as the
{\ND{
LORD}} had
said, and as the
{\ND{
LORD}} had
sworn unto them: and they were
greatly
distressed.
\par }{\PP \VS{16}¶ Nevertheless the
{\ND{
LORD}} raised
up
judges, which
delivered them out of the
hand of those that
spoiled them.\FTNT{**}{{\FR 2.16 }
delivered: Heb. saved}
\VS{17}And yet they would not
hearken unto their
judges, but they went a
whoring
after
other
gods, and
bowed themselves unto them: they
turned
quickly out of the
way which their
fathers walked
in,
obeying the
commandments of the
{\ND{
LORD}};
{\ADD{but}} they
did not so.
\VS{18}And when the
{\ND{
LORD}}
raised them up
judges, then the
{\ND{
LORD}} was with the
judge, and
delivered them out of the
hand of their
enemies all the
days of the
judge: for it
repented the
{\ND{
LORD}}
because of their
groanings by reason of them that
oppressed them and
vexed them.
\VS{19}And it came to pass, when the
judge was
dead,
{\ADD{that}} they
returned, and
corrupted
{\ADD{themselves}} more than their
fathers, in
following
other
gods to
serve them, and to bow
down unto them; they
ceased not from their own
doings, nor from their
stubborn
way.\FTNT{††}{{\FR 2.19 }
corrupted…: or, were corrupt}\FTNT{‡‡}{{\FR 2.19 }
ceased…: Heb. let nothing fall of their}
\par }{\PP \VS{20}¶ And the
anger of the
{\ND{
LORD}} was
hot against
Israel; and he
said, Because that this
people hath
transgressed my
covenant which I
commanded their
fathers, and have not
hearkened unto my
voice;
\VS{21}I also will not
henceforth drive
out
any from
before them of the
nations which
Joshua
left when he
died:
\VS{22}That through them I may
prove
Israel, whether they will
keep the
way of the
{\ND{
LORD}} to
walk therein, as their
fathers did
keep
{\ADD{it}}, or not.
\VS{23}Therefore the
{\ND{
LORD}}
left those
nations,
without driving them
out
hastily; neither
delivered he them into the
hand of
Joshua.\FTNT{§§}{{\FR 2.23 }
left: or, suffered}

\par }\Chap{3}{\PP \VerseOne{1}Now these
{\ADD{are}} the
nations which the
{\ND{
LORD}}
left, to
prove
Israel by them,
{\ADD{even}} as many
{\ADD{of Israel}} as had not
known all the
wars of
Canaan;
\VS{2}Only
that the
generations of the
children of
Israel might
know, to
teach them
war, at the least such as
before
knew nothing thereof;
\VS{3}{\ADD{Namely}},
five
lords of the
Philistines, and all the
Canaanites, and the
Sidonians, and the
Hivites that
dwelt in
mount
Lebanon, from
mount
Baal-hermon unto the entering
in of
Hamath.
\VS{4}And they were to
prove
Israel by them, to
know whether they would
hearken unto the
commandments of the
{\ND{
LORD}}, which he
commanded their
fathers by the
hand of
Moses.
\par }{\PP \VS{5}¶ And the
children of
Israel
dwelt
among the
Canaanites,
Hittites, and
Amorites, and
Perizzites, and
Hivites, and
Jebusites:
\VS{6}And they
took their
daughters to be their
wives, and
gave their
daughters to their
sons, and
served their
gods.
\VS{7}And the
children of
Israel
did
evil in the
sight of the
{\ND{
LORD}}, and
forgat the
{\ND{
LORD}} their
God, and
served
Baalim and the
groves.
\par }{\PP \VS{8}¶ Therefore the
anger of the
{\ND{
LORD}} was
hot against
Israel, and he
sold them into the
hand of
Chushan-rishathaim
king of
Mesopotamia: and the
children of
Israel
served
Chushan-rishathaim
eight
years.\FTNT{*}{{\FR 3.8 }
Mesopotamia: Heb. Aram-naharaim}
\VS{9}And when the
children of
Israel
cried unto the
{\ND{
LORD}}, the
{\ND{
LORD}} raised
up a
deliverer to the
children of
Israel, who
delivered them,
{\ADD{even}}
Othniel the
son of
Kenaz,
Caleb’s
younger
brother.\FTNT{†}{{\FR 3.9 }
deliverer: Heb. saviour}
\VS{10}And the
Spirit of the
{\ND{
LORD}} came upon him, and he
judged
Israel, and went
out to
war: and the
{\ND{
LORD}}
delivered
Chushan-rishathaim
king of
Mesopotamia into his
hand; and his
hand
prevailed against
Chushan-rishathaim.\FTNT{‡}{{\FR 3.10 }
came: Heb. was}\FTNT{§}{{\FR 3.10 }
Mesopotamia: Heb. Aram}
\VS{11}And the
land had
rest
forty
years. And
Othniel the
son of
Kenaz
died.
\par }{\PP \VS{12}¶ And the
children of
Israel
did
evil
again in the
sight of the
{\ND{
LORD}}: and the
{\ND{
LORD}}
strengthened
Eglon the
king of
Moab against
Israel, because they had
done
evil in the
sight of the
{\ND{
LORD}}.
\VS{13}And he
gathered unto him the
children of
Ammon and
Amalek, and
went and
smote
Israel, and
possessed the
city of palm
trees.
\VS{14}So the
children of
Israel
served
Eglon the
king of
Moab
eighteen
years.
\VS{15}But when the
children of
Israel
cried unto the
{\ND{
LORD}}, the
{\ND{
LORD}} raised them
up a
deliverer,
Ehud the
son of
Gera, a
Benjamite, a
man
lefthanded: and by
him the
children of
Israel
sent a
present unto
Eglon the
king of
Moab.\FTNT{**}{{\FR 3.15 }
a Benjamite: or, the son of Jemini}\FTNT{††}{{\FR 3.15 }
lefthanded: Heb. shut of his right hand}
\VS{16}But
Ehud
made him a
dagger which had
two
edges, of a
cubit
length; and he did
gird it under his
raiment upon his
right
thigh.
\VS{17}And he
brought the
present unto
Eglon
king of
Moab: and
Eglon
{\ADD{was}} a
very
fat
man.
\VS{18}And when he had made an
end to
offer the
present, he sent
away the
people that
bare the
present.
\VS{19}But he himself turned
again from the
quarries that
{\ADD{were}} by
Gilgal, and
said, I have a
secret
errand unto thee, O
king: who
said, Keep
silence. And all that
stood by him went
out from him.\FTNT{‡‡}{{\FR 3.19 }
quarries: or, graven images}
\VS{20}And
Ehud
came unto him; and he was
sitting in a
summer
parlour, which he had for himself alone. And
Ehud
said, I have a
message from
God unto thee. And he
arose out of
{\ADD{his}}
seat.\FTNT{§§}{{\FR 3.20 }
a summer…: Heb. a parlour of cooling}
\VS{21}And
Ehud put
forth his
left
hand, and
took the
dagger from his
right
thigh, and
thrust it into his
belly:
\VS{22}And the
haft also went
in
after the
blade; and the
fat
closed
upon the
blade, so that he could not
draw the
dagger out of his
belly; and the
dirt came
out.\FTNT{***}{{\FR 3.22 }
the dirt…: or, it came out at the buttocks}
\VS{23}Then
Ehud went
forth through the
porch, and
shut the
doors of the
parlour upon him, and
locked them.
\VS{24}When he was gone
out, his
servants
came; and when they
saw that, behold, the
doors of the
parlour
{\ADD{were}}
locked, they
said, Surely he
covereth his
feet in his
summer
chamber.\FTNT{†††}{{\FR 3.24 }
covereth…: or, doeth his easement}
\VS{25}And they
tarried till they were
ashamed: and, behold, he
opened not the
doors of the
parlour; therefore they
took a
key, and
opened
{\ADD{them:}} and, behold, their
lord
{\ADD{was}} fallen
down
dead on the
earth.
\VS{26}And
Ehud
escaped while they
tarried, and passed
beyond the
quarries, and
escaped unto
Seirath.
\VS{27}And it came to pass, when he was
come, that he
blew a
trumpet in the
mountain of
Ephraim, and the
children of
Israel went
down with him from the
mount, and he
before them.
\VS{28}And he
said unto them,
Follow
after me: for the
{\ND{
LORD}} hath
delivered your
enemies the
Moabites into your
hand. And they went
down
after him, and
took the
fords of
Jordan toward
Moab, and
suffered not a
man to pass
over.
\VS{29}And they
slew of
Moab at that
time about
ten
thousand
men, all
lusty, and all
men of
valour; and there
escaped not a
man.\FTNT{‡‡‡}{{\FR 3.29 }
lusty: Heb. fat}
\VS{30}So
Moab was
subdued that
day under the
hand of
Israel. And the
land had
rest
fourscore
years.
\par }{\PP \VS{31}¶ And
after him was
Shamgar the
son of
Anath, which
slew of the
Philistines
six
hundred
men with an
ox
goad: and he also
delivered
Israel.

\par }\Chap{4}{\PP \VerseOne{1}And the
children of
Israel
again
did
evil in the
sight of the
{\ND{
LORD}}, when
Ehud was
dead.
\VS{2}And the
{\ND{
LORD}}
sold them into the
hand of
Jabin
king of
Canaan, that
reigned in
Hazor; the
captain of whose
host
{\ADD{was}}
Sisera, which
dwelt in
Harosheth of the
Gentiles.
\VS{3}And the
children of
Israel
cried unto the
{\ND{
LORD}}: for he had
nine
hundred
chariots of
iron; and
twenty
years he
mightily
oppressed the
children of
Israel.
\par }{\PP \VS{4}¶ And
Deborah, a
prophetess, the
wife of
Lapidoth, she
judged
Israel at that
time.
\VS{5}And she
dwelt under the palm
tree of
Deborah between
Ramah and
Beth-el in
mount
Ephraim: and the
children of
Israel came
up to her for
judgment.
\VS{6}And she
sent and
called
Barak the
son of
Abinoam out of
Kedesh-naphtali, and
said unto him, Hath not the
{\ND{
LORD}}
God of
Israel
commanded,
{\ADD{saying}},
Go and
draw toward
mount
Tabor, and
take with thee
ten
thousand
men of the
children of
Naphtali and of the
children of
Zebulun?
\VS{7}And I will
draw unto thee to the
river
Kishon
Sisera, the
captain of
Jabin’s
army, with his
chariots and his
multitude; and I will
deliver him into thine
hand.
\VS{8}And
Barak
said unto her, If thou wilt
go with me, then I will
go: but if thou wilt not
go with me,
{\ADD{then}} I will not
go.
\VS{9}And she
said, I will
surely
go with thee:
notwithstanding the
journey that thou
takest shall not be for thine
honour; for the
{\ND{
LORD}} shall
sell
Sisera into the
hand of a
woman. And
Deborah
arose, and
went with
Barak to
Kedesh.
\par }{\PP \VS{10}¶ And
Barak
called
Zebulun and
Naphtali to
Kedesh; and he went
up with
ten
thousand
men at his
feet: and
Deborah went
up with him.
\VS{11}Now
Heber the
Kenite,
{\ADD{which was}} of the
children of
Hobab the father in
law of
Moses, had
severed himself from the
Kenites, and
pitched his
tent unto the
plain of
Zaanaim, which
{\ADD{is}} by
Kedesh.
\VS{12}And they
shewed
Sisera that
Barak the
son of
Abinoam was gone
up to
mount
Tabor.
\VS{13}And
Sisera gathered
together all his
chariots,
{\ADD{even}}
nine
hundred
chariots of
iron, and all the
people that
{\ADD{were}} with him, from
Harosheth of the
Gentiles unto the
river of
Kishon.\FTNT{*}{{\FR 4.13 }
together: Heb. by cry, or, proclamation}
\VS{14}And
Deborah
said unto
Barak,
Up; for this
{\ADD{is}} the
day in which the
{\ND{
LORD}} hath
delivered
Sisera into thine
hand: is not the
{\ND{
LORD}} gone
out
before thee? So
Barak went
down from
mount
Tabor, and
ten
thousand
men
after him.
\VS{15}And the
{\ND{
LORD}}
discomfited
Sisera, and all
{\ADD{his}}
chariots, and all
{\ADD{his}}
host, with the
edge of the
sword
before
Barak; so that
Sisera lighted
down off
{\ADD{his}}
chariot, and fled
away on his
feet.
\VS{16}But
Barak
pursued
after the
chariots, and after the
host, unto
Harosheth of the
Gentiles: and all the
host of
Sisera
fell upon the
edge of the
sword;
{\ADD{and}} there was not a
man
left.\FTNT{†}{{\FR 4.16 }
a man: Heb. unto one}
\par }{\PP \VS{17}Howbeit
Sisera fled
away on his
feet to the
tent of
Jael the
wife of
Heber the
Kenite: for
{\ADD{there was}}
peace between
Jabin the
king of
Hazor and the
house of
Heber the
Kenite.
\par }{\PP \VS{18}¶ And
Jael went
out to
meet
Sisera, and
said unto him, Turn
in, my
lord, turn
in to me;
fear not. And when he had turned
in unto her into the
tent, she
covered him with a
mantle.\FTNT{‡}{{\FR 4.18 }
mantle: or, rug, or, blanket}
\VS{19}And he
said unto her, Give me, I pray thee, a
little
water to
drink; for I am
thirsty. And she
opened a
bottle of
milk, and gave him
drink, and
covered him.
\VS{20}Again he
said unto her, Stand
in the
door of the
tent, and it shall be, when any
man doth
come and
enquire of thee, and
say, Is
there any
man
here? that thou shalt
say, No.
\VS{21}Then
Jael
Heber’s
wife
took a
nail of the
tent, and
took an
hammer in her
hand, and
went
softly unto him, and
smote the
nail into his
temples, and
fastened it into the
ground: for he was fast
asleep and
weary. So he
died.\FTNT{§}{{\FR 4.21 }
and took: Heb. and put}
\VS{22}And, behold, as
Barak
pursued
Sisera,
Jael came
out to
meet him, and
said unto him,
Come, and I will
shew thee the
man whom thou
seekest. And when he
came into her
{\ADD{tent}}, behold,
Sisera
lay
dead, and the
nail
{\ADD{was}} in his
temples.
\VS{23}So
God
subdued on that
day
Jabin the
king of
Canaan
before the
children of
Israel.
\VS{24}And the
hand of the
children of
Israel
prospered, and
prevailed against
Jabin the
king of
Canaan, until they had
destroyed
Jabin
king of
Canaan.\FTNT{**}{{\FR 4.24 }
prospered…: Heb. going went and was hard}

\par }\Chap{5}{\PP \VerseOne{1}Then
sang
Deborah and
Barak the
son of
Abinoam on that
day,
saying,
\VS{2}Praise ye the
{\ND{
LORD}} for the
avenging of
Israel, when the
people willingly
offered themselves.
\VS{3}Hear, O ye
kings; give
ear, O ye
princes; I,
{\ADD{even}} I, will
sing unto the
{\ND{
LORD}}; I will
sing
{\ADD{praise}} to the
{\ND{
LORD}}
God of
Israel.
\VS{4}{\ND{
LORD}}, when thou wentest
out of
Seir, when thou marchedst
out of the
field of
Edom, the
earth
trembled, and the
heavens
dropped, the
clouds also
dropped
water.
\VS{5}The
mountains
melted from
before the
{\ND{
LORD}},
{\ADD{even}} that
Sinai from
before the
{\ND{
LORD}}
God of
Israel.\FTNT{*}{{\FR 5.5 }
melted: Heb. flowed}
\par }{\PP \VS{6}In the
days of
Shamgar the
son of
Anath, in the
days of
Jael, the
highways were
unoccupied, and the
travellers
walked through
byways.\FTNT{†}{{\FR 5.6 }
travellers: Heb. walkers of paths}\FTNT{‡}{{\FR 5.6 }
byways: Heb. crooked ways}
\VS{7}{\ADD{The inhabitants of}} the
villages
ceased, they
ceased in
Israel, until that I
Deborah
arose, that I
arose a
mother in
Israel.
\VS{8}They
chose
new
gods; then
{\ADD{was}}
war in the
gates: was there a
shield or
spear
seen among
forty
thousand in
Israel?
\VS{9}My
heart
{\ADD{is}} toward the
governors of
Israel, that offered themselves
willingly among the
people.
Bless ye the
{\ND{
LORD}}.
\VS{10}Speak, ye that
ride on
white
asses, ye that
sit in
judgment, and
walk by the
way.\FTNT{§}{{\FR 5.10 }
Speak: or, Meditate}
\VS{11}{\ADD{They that are delivered}} from the
noise of
archers in the places of drawing
water, there shall they
rehearse the righteous
acts of the
{\ND{
LORD}},
{\ADD{even}} the righteous
acts
{\ADD{toward the inhabitants}} of his
villages in
Israel: then shall the
people of the
{\ND{
LORD}} go
down to the
gates.\FTNT{**}{{\FR 5.11 }
righteous…: Heb. righteousnesses}
\par }{\PP \VS{12}Awake,
awake,
Deborah:
awake,
awake,
utter a
song:
arise,
Barak, and lead thy
captivity
captive, thou
son of
Abinoam.
\VS{13}Then he made him that
remaineth have
dominion over the
nobles among the
people: the
{\ND{
LORD}} made me have
dominion over the
mighty.
\VS{14}Out of
Ephraim
{\ADD{was there}} a
root of them against
Amalek;
after thee,
Benjamin, among thy
people; out of
Machir came
down
governors, and out of
Zebulun they that
handle the
pen of the
writer.\FTNT{††}{{\FR 5.14 }
handle: Heb. draw with}
\VS{15}And the
princes of
Issachar
{\ADD{were}} with
Deborah; even
Issachar, and also
Barak: he was
sent on
foot into the
valley. For the
divisions of
Reuben
{\ADD{there were}}
great
thoughts of
heart.\FTNT{‡‡}{{\FR 5.15 }
foot: Heb. his feet}\FTNT{§§}{{\FR 5.15 }
thoughts: Heb. impressions}\FTNT{***}{{\FR 5.15 }
For: or, In}
\VS{16}Why
abodest thou
among the
sheepfolds, to
hear the
bleatings of the
flocks? For the
divisions of
Reuben
{\ADD{there were}}
great
searchings of
heart.\FTNT{†††}{{\FR 5.16 }
For: or, In}
\VS{17}Gilead
abode
beyond
Jordan: and why did
Dan
remain in
ships?
Asher
continued on the
sea
shore, and
abode in his
breaches.\FTNT{‡‡‡}{{\FR 5.17 }
shore: or, port}\FTNT{§§§}{{\FR 5.17 }
breaches: or, creeks}
\VS{18}Zebulun and
Naphtali
{\ADD{were}} a
people
{\ADD{that}}
jeoparded their
lives unto the
death in the high
places of the
field.\FTNT{*}{{\FR 5.18 }
jeoparded: Heb. exposed to reproach}
\VS{19}The
kings
came
{\ADD{and}}
fought, then
fought the
kings of
Canaan in
Taanach by the
waters of
Megiddo; they
took no
gain of
money.
\VS{20}They
fought from
heaven; the
stars in their
courses
fought against
Sisera.\FTNT{†}{{\FR 5.20 }
courses: Heb. paths}
\VS{21}The
river of
Kishon swept them
away, that
ancient
river, the
river
Kishon. O my
soul, thou hast trodden
down
strength.
\VS{22}Then were the
horsehoofs
broken by the means of the
pransings, the
pransings of their mighty
ones.\FTNT{‡}{{\FR 5.22 }
pransings: or, tramplings, or, plungings}
\VS{23}Curse ye
Meroz,
said the
angel of the
{\ND{
LORD}},
curse ye
bitterly the
inhabitants thereof; because they
came not to the
help of the
{\ND{
LORD}}, to the
help of the
{\ND{
LORD}} against the
mighty.
\par }{\PP \VS{24}Blessed above
women shall
Jael the
wife of
Heber the
Kenite be,
blessed shall she be above
women in the
tent.
\VS{25}He
asked
water,
{\ADD{and}} she
gave
{\ADD{him}}
milk; she brought
forth
butter in a
lordly
dish.
\VS{26}She
put her
hand to the
nail, and her right
hand to the
workmen’s
hammer; and with the hammer she
smote
Sisera, she smote
off his
head, when she had
pierced and stricken
through his
temples.\FTNT{§}{{\FR 5.26 }
with…: Heb. she hammered}
\VS{27}At her
feet he
bowed, he
fell, he lay
down: at her
feet he
bowed, he
fell:
where he
bowed, there he fell
down
dead.\FTNT{**}{{\FR 5.27 }
At: Heb. Between}\FTNT{††}{{\FR 5.27 }
dead: Heb. destroyed}
\VS{28}The
mother of
Sisera
looked out at a
window, and
cried through the
lattice, Why is his
chariot
{\ADD{so}}
long in
coming? why
tarry the
wheels of his
chariots?
\VS{29}Her
wise
ladies
answered her, yea, she
returned
answer to herself,\FTNT{‡‡}{{\FR 5.29 }
answer: Heb. her words}
\VS{30}Have they not
sped? have they
{\ADD{not}}
divided the
prey; to
every
man a
damsel
{\ADD{or}}
two; to
Sisera a
prey of divers
colours, a
prey of divers
colours of
needlework, of divers
colours of needlework on both
sides,
{\ADD{meet}} for the
necks of
{\ADD{them that take}} the
spoil?\FTNT{§§}{{\FR 5.30 }
every…: Heb. the head of a man}
\VS{31}So let all thine
enemies
perish, O
{\ND{
LORD}}: but
{\ADD{let}} them that
love him
{\ADD{be}} as the
sun when he goeth
forth in his
might. And the
land had
rest
forty
years.

\par }\Chap{6}{\PP \VerseOne{1}And the
children of
Israel
did
evil in the
sight of the
{\ND{
LORD}}: and the
{\ND{
LORD}}
delivered them into the
hand of
Midian
seven
years.
\VS{2}And the
hand of
Midian
prevailed against
Israel:
{\ADD{and}}
because of the
Midianites the
children of
Israel
made them the
dens which
{\ADD{are}} in the
mountains, and
caves, and strong
holds.\FTNT{*}{{\FR 6.2 }
prevailed: Heb. was strong}
\VS{3}And
{\ADD{so}} it was, when
Israel had
sown, that the
Midianites came
up, and the
Amalekites, and the
children of the
east, even they came
up against them;
\VS{4}And they
encamped against them, and
destroyed the
increase of the
earth, till thou
come unto
Gaza, and
left no
sustenance for
Israel, neither
sheep, nor
ox, nor
ass.\FTNT{†}{{\FR 6.4 }
sheep: or goat}
\VS{5}For they came
up with their
cattle and their
tents, and they
came
as
grasshoppers for
multitude;
{\ADD{for}} both they and their
camels were without
number: and they
entered into the
land to
destroy it.
\VS{6}And
Israel was
greatly
impoverished
because of the
Midianites; and the
children of
Israel
cried unto the
{\ND{
LORD}}.
\par }{\PP \VS{7}¶ And it came to pass, when the
children of
Israel
cried unto the
{\ND{
LORD}}
because of the
Midianites,
\VS{8}That the
{\ND{
LORD}}
sent a
prophet unto the
children of
Israel, which
said unto them, Thus
saith the
{\ND{
LORD}}
God of
Israel, I brought you
up from
Egypt, and brought you
forth out of the
house of
bondage;\FTNT{‡}{{\FR 6.8 }
a prophet: Heb. a man a prophet}
\VS{9}And I
delivered you out of the
hand of the
Egyptians, and out of the
hand of all that
oppressed you, and drave them
out from
before you, and
gave you their
land;
\VS{10}And I
said unto you, I
{\ADD{am}} the
{\ND{
LORD}} your
God;
fear not the
gods of the
Amorites, in whose
land ye
dwell: but ye have not
obeyed my
voice.
\par }{\PP \VS{11}¶ And there
came an
angel of the
{\ND{
LORD}}, and
sat under an
oak which
{\ADD{was}} in
Ophrah, that
{\ADD{pertained}} unto
Joash the
Abi-ezrite: and his
son
Gideon
threshed
wheat by the
winepress, to
hide
{\ADD{it}}
from the
Midianites.\FTNT{§}{{\FR 6.11 }
Gideon: Gr. Gedeon}\FTNT{**}{{\FR 6.11 }
to hide…: Heb. to cause it to flee}
\VS{12}And the
angel of the
{\ND{
LORD}}
appeared unto him, and
said unto him, The
{\ND{
LORD}}
{\ADD{is}} with thee, thou
mighty man of
valour.
\VS{13}And
Gideon
said unto him,
Oh my
Lord,
if the
{\ND{
LORD}} be with us, why then is all this
befallen us? and where
{\ADD{be}} all his
miracles which our
fathers
told us of,
saying, Did not the
{\ND{
LORD}} bring us
up from
Egypt? but now the
{\ND{
LORD}} hath
forsaken us, and
delivered us into the
hands of the
Midianites.
\VS{14}And the
{\ND{
LORD}}
looked upon him, and
said,
Go in this thy
might, and thou shalt
save
Israel from the
hand of the
Midianites: have not I
sent thee?
\VS{15}And he
said unto him,
Oh my
Lord,
wherewith shall I
save
Israel? behold, my
family
{\ADD{is}}
poor in
Manasseh, and I
{\ADD{am}} the
least in my
father’s
house.\FTNT{††}{{\FR 6.15 }
my family…: Heb. my thousand is the meanest}
\VS{16}And the
{\ND{
LORD}}
said unto him, Surely I will be with thee, and thou shalt
smite the
Midianites as
one
man.
\VS{17}And he
said unto him, If now I have
found
grace in thy
sight, then
shew me a
sign that thou
talkest with me.
\VS{18}Depart not hence, I pray thee, until I
come unto thee, and bring
forth my
present, and
set
{\ADD{it}}
before thee. And he
said, I will
tarry until thou come
again.\FTNT{‡‡}{{\FR 6.18 }
present: or, meat offering}
\par }{\PP \VS{19}¶ And
Gideon went
in, and made
ready a
kid, and unleavened
cakes of an
ephah of
flour: the
flesh he
put in a
basket, and he
put the
broth in a
pot, and brought
{\ADD{it}}
out unto him under the
oak, and
presented
{\ADD{it}}.\FTNT{§§}{{\FR 6.19 }
a kid: Heb. a kid of the goats}
\VS{20}And the
angel of
God
said unto him,
Take the
flesh and the unleavened
cakes, and
lay
{\ADD{them}} upon
this
rock, and pour
out the
broth. And he
did so.
\par }{\PP \VS{21}¶ Then the
angel of the
{\ND{
LORD}} put
forth the
end of the
staff that
{\ADD{was}} in his
hand, and
touched the
flesh and the unleavened
cakes; and there rose
up
fire out of the
rock, and
consumed the
flesh and the unleavened
cakes. Then the
angel of the
{\ND{
LORD}}
departed out of his
sight.
\VS{22}And when
Gideon
perceived that he
{\ADD{was}} an
angel of the
{\ND{
LORD}},
Gideon
said,
Alas, O
Lord
{\ND{
GOD}}! for
because I have
seen an
angel of the
{\ND{
LORD}}
face to
face.
\VS{23}And the
{\ND{
LORD}}
said unto him,
Peace
{\ADD{be}} unto thee;
fear not: thou shalt not
die.
\VS{24}Then
Gideon
built an
altar there unto the
{\ND{
LORD}}, and
called it
Jehovah-shalom: unto this
day it
{\ADD{is}} yet in
Ophrah of the
Abi-ezrites.\FTNT{***}{{\FR 6.24 }
Jehovah-shalom: that is, The
{\ND{LORD}} send peace}
\par }{\PP \VS{25}¶ And it came to pass the same
night, that the
{\ND{
LORD}}
said unto him,
Take thy
father’s
young
bullock, even the
second
bullock of
seven years
old, and throw
down the
altar of
Baal that thy
father hath, and cut
down the
grove that
{\ADD{is}} by it:\FTNT{†††}{{\FR 6.25 }
even: or, and}
\VS{26}And
build an
altar unto the
{\ND{
LORD}} thy
God upon the
top of this
rock, in the ordered
place, and
take the
second
bullock, and
offer a burnt
sacrifice with the
wood of the
grove which thou shalt cut
down.\FTNT{‡‡‡}{{\FR 6.26 }
rock: Heb. strong place}\FTNT{§§§}{{\FR 6.26 }
in the ordered…: or, in an orderly manner}
\VS{27}Then
Gideon
took
ten
men of his
servants, and
did as the
{\ND{
LORD}} had
said unto him: and
{\ADD{so}} it was, because he
feared his
father’s
household, and the
men of the
city, that he could not
do
{\ADD{it}} by
day, that he
did
{\ADD{it}} by
night.
\par }{\PP \VS{28}¶ And when the
men of the
city arose
early in the
morning, behold, the
altar of
Baal was cast
down, and the
grove was cut
down that
{\ADD{was}} by it, and the
second
bullock was
offered upon the
altar
{\ADD{that was}}
built.
\VS{29}And they
said
one to
another, Who hath
done this
thing? And when they
enquired and
asked, they
said,
Gideon the
son of
Joash hath
done this
thing.
\VS{30}Then the
men of the
city
said unto
Joash, Bring
out thy
son, that he may
die: because he hath cast
down the
altar of
Baal, and because he hath cut
down the
grove that
{\ADD{was}} by it.
\VS{31}And
Joash
said unto all that
stood against him, Will ye
plead for
Baal? will ye
save him? he that will
plead for him, let him be put to
death whilst
{\ADD{it is yet}}
morning: if he
{\ADD{be}} a
god, let him
plead for himself, because
{\ADD{one}} hath cast
down his
altar.
\VS{32}Therefore on that
day he
called him
Jerubbaal,
saying, Let
Baal
plead against him, because he hath thrown
down his
altar.\FTNT{*}{{\FR 6.32 }
Jerubbaal: that is, Let Baal plead}\FTNT{†}{{\FR 6.32 }
Jerubbesheth: that is, Let the shameful thing plead}
\par }{\PP \VS{33}¶ Then all the
Midianites and the
Amalekites and the
children of the
east were
gathered
together, and went
over, and
pitched in the
valley of
Jezreel.
\VS{34}But the
Spirit of the
{\ND{
LORD}}
came upon
Gideon, and he
blew a
trumpet; and
Abi-ezer was
gathered
after him.\FTNT{‡}{{\FR 6.34 }
came…: Heb. clothed}\FTNT{§}{{\FR 6.34 }
gathered: Heb. called}
\VS{35}And he
sent
messengers throughout all
Manasseh; who also was
gathered
after him: and he
sent
messengers unto
Asher, and unto
Zebulun, and unto
Naphtali; and they came
up to
meet them.\FTNT{**}{{\FR 6.35 }
gathered: Heb. called}
\par }{\PP \VS{36}¶ And
Gideon
said unto
God, If thou
wilt
save
Israel by mine
hand, as thou hast
said,
\VS{37}Behold, I will
put a
fleece of
wool in the
floor;
{\ADD{and}} if the
dew be on the
fleece only, and
{\ADD{it be}}
dry upon all the
earth
{\ADD{beside}}, then shall I
know that thou wilt
save
Israel by mine
hand, as thou hast
said.
\VS{38}And it was so: for he rose up
early on the
morrow, and
thrust the
fleece
together, and
wringed the
dew out of the
fleece, a
bowl
full of
water.
\VS{39}And
Gideon
said unto
God, Let not thine
anger be
hot against me, and I will
speak but this
once: let me
prove, I pray thee, but this
once with the
fleece; let it now be
dry only upon the
fleece, and upon all the
ground let there be
dew.
\VS{40}And
God
did so that
night: for it was
dry upon the
fleece only, and there was
dew on all the
ground.

\par }\Chap{7}{\PP \VerseOne{1}Then
Jerubbaal, who
{\ADD{is}}
Gideon, and all the
people that
{\ADD{were}} with him, rose up
early, and
pitched beside the well of
Harod: so that the
host of the
Midianites were on the north
side of them, by the
hill of
Moreh, in the
valley.
\VS{2}And the
{\ND{
LORD}}
said unto
Gideon, The
people that
{\ADD{are}} with thee
{\ADD{are}} too
many for me to
give the
Midianites into their
hands, lest
Israel
vaunt themselves against me,
saying, Mine own
hand hath
saved me.
\VS{3}Now therefore go
to,
proclaim in the
ears of the
people,
saying, Whosoever
{\ADD{is}}
fearful and
afraid, let him
return and depart
early from
mount
Gilead. And there
returned of the
people
twenty and
two
thousand; and there
remained
ten
thousand.
\VS{4}And the
{\ND{
LORD}}
said unto
Gideon, The
people
{\ADD{are}} yet
{\ADD{too}}
many; bring them
down unto the
water, and I will
try them for thee there: and it shall be,
{\ADD{that}} of whom I
say unto thee, This shall
go with thee, the same shall
go with thee; and of whomsoever I
say unto thee, This shall not
go with thee, the same shall not
go.
\VS{5}So he brought
down the
people unto the
water: and the
{\ND{
LORD}}
said unto
Gideon, Every one that
lappeth of the
water with his
tongue, as a
dog
lappeth, him shalt thou
set by himself; likewise every one that boweth
down upon his
knees to
drink.
\VS{6}And the
number of them that
lapped,
{\ADD{putting}} their
hand to their
mouth, were
three
hundred
men: but all the
rest of the
people bowed
down upon their
knees to
drink
water.
\VS{7}And the
{\ND{
LORD}}
said unto
Gideon, By the
three
hundred
men that
lapped will I
save you, and
deliver the
Midianites into thine
hand: and let all the
{\ADD{other}}
people
go every
man unto his
place.
\VS{8}So the
people
took
victuals in their
hand, and their
trumpets: and he
sent
all
{\ADD{the rest of}}
Israel every
man unto his
tent, and
retained those
three
hundred
men: and the
host of
Midian was beneath him in the
valley.
\par }{\PP \VS{9}¶ And it came to pass the same
night, that the
{\ND{
LORD}}
said unto him,
Arise, get thee
down unto the
host; for I have
delivered it into thine
hand.
\VS{10}But if thou
fear to go
down,
go thou with
Phurah thy
servant
down to the
host:
\VS{11}And thou shalt
hear what they
say; and
afterward shall thine
hands be
strengthened to go
down unto the
host. Then went he
down with
Phurah his
servant unto the
outside of the armed
men that
{\ADD{were}} in the
host.\FTNT{*}{{\FR 7.11 }
armed men: or, ranks by five}
\VS{12}And the
Midianites and the
Amalekites and all the
children of the
east
lay along in the
valley like
grasshoppers for
multitude; and their
camels
{\ADD{were}} without
number, as the
sand by the
sea
side for
multitude.
\VS{13}And when
Gideon was
come, behold,
{\ADD{there was}} a
man that
told a
dream unto his
fellow, and
said, Behold, I
dreamed a
dream, and, lo, a
cake of
barley
bread
tumbled into the
host of
Midian, and
came unto a
tent, and
smote it that it
fell, and
overturned it, that the
tent lay
along.
\VS{14}And his
fellow
answered and
said, This
{\ADD{is}} nothing else save the
sword of
Gideon the
son of
Joash, a
man of
Israel:
{\ADD{for}} into his
hand hath
God
delivered
Midian, and all the
host.
\par }{\PP \VS{15}¶ And it was
{\ADD{so}}, when
Gideon
heard the
telling of the
dream, and the
interpretation thereof, that he
worshipped, and
returned into the
host of
Israel, and
said,
Arise; for the
{\ND{
LORD}} hath
delivered into your
hand the
host of
Midian.\FTNT{†}{{\FR 7.15 }
interpretation: Heb. breaking}
\par }{\PP \VS{16}And he
divided the
three
hundred
men
{\ADD{into}}
three
companies, and he
put a
trumpet in every
man’s hand, with
empty
pitchers, and
lamps
within the
pitchers.\FTNT{‡}{{\FR 7.16 }
a trumpet…: Heb. trumpets in the hand of all of them}\FTNT{§}{{\FR 7.16 }
lamps: or, firebrands, or, torches}
\VS{17}And he
said unto them,
Look on me, and
do likewise: and, behold, when I
come to the
outside of the
camp, it shall be
{\ADD{that}}, as I
do, so shall ye
do.
\VS{18}When I
blow with a
trumpet, I and all that
{\ADD{are}} with me, then
blow ye the
trumpets also on every
side of all the
camp, and
say,
{\ADD{The sword}} of the
{\ND{
LORD}}, and of
Gideon.
\par }{\PP \VS{19}¶ So
Gideon, and the
hundred
men that
{\ADD{were}} with him,
came unto the
outside of the
camp in the
beginning of the
middle
watch; and they had
but
newly
set the
watch: and they
blew the
trumpets, and
brake the
pitchers that
{\ADD{were}} in their
hands.
\VS{20}And the
three
companies
blew the
trumpets, and
brake the
pitchers, and
held the
lamps in their
left
hands, and the
trumpets in their
right
hands to
blow
{\ADD{withal:}} and they
cried, The
sword of the
{\ND{
LORD}}, and of
Gideon.
\VS{21}And they
stood every
man in his place round
about the
camp: and all the
host
ran, and
cried, and
fled.
\VS{22}And the
three
hundred
blew the
trumpets, and the
{\ND{
LORD}}
set every
man’s
sword against his
fellow, even throughout all the
host: and the
host
fled to
Beth-shittah in
Zererath,
{\ADD{and}} to the
border of
Abel-meholah, unto
Tabbath.\FTNT{**}{{\FR 7.22 }
in: or, toward}\FTNT{††}{{\FR 7.22 }
border: Heb. lip}
\par }{\PP \VS{23}And the
men of
Israel gathered themselves
together out of
Naphtali, and out of
Asher, and out of all
Manasseh, and
pursued
after the
Midianites.
\par }{\PP \VS{24}¶ And
Gideon
sent
messengers throughout all
mount
Ephraim,
saying, Come
down
against the
Midianites, and
take before them the
waters unto
Beth-barah and
Jordan. Then all the
men of
Ephraim gathered themselves
together, and
took the
waters unto
Beth-barah and
Jordan.
\VS{25}And they
took
two
princes of the
Midianites,
Oreb and
Zeeb; and they
slew
Oreb upon the
rock
Oreb, and
Zeeb they
slew at the
winepress of
Zeeb, and
pursued
Midian, and
brought the
heads of
Oreb and
Zeeb to
Gideon on the other
side
Jordan.

\par }\Chap{8}{\PP \VerseOne{1}And the
men of
Ephraim
said unto him, Why hast
thou
served us
thus, that thou
calledst us not, when thou
wentest to
fight with the
Midianites? And they did
chide with him
sharply.\FTNT{*}{{\FR 8.1 }
Why…: Heb. What thing is this thou hast done unto us}\FTNT{†}{{\FR 8.1 }
sharply: Heb. strongly}
\VS{2}And he
said unto them, What have I
done now in comparison of you?
{\ADD{Is}} not the
gleaning of the grapes of
Ephraim
better than the
vintage of
Abi-ezer?
\VS{3}God hath
delivered into your
hands the
princes of
Midian,
Oreb and
Zeeb: and what was I
able to
do in comparison of you? Then their
anger was
abated toward him, when he had
said
that.\FTNT{‡}{{\FR 8.3 }
anger: Heb. spirit}
\par }{\PP \VS{4}¶ And
Gideon
came to
Jordan,
{\ADD{and}} passed
over, he, and the
three
hundred
men that
{\ADD{were}} with him,
faint, yet
pursuing
{\ADD{them}}.
\VS{5}And he
said unto the
men of
Succoth,
Give, I pray you,
loaves of
bread unto the
people that
follow me; for they
{\ADD{be}}
faint, and I am
pursuing
after
Zebah and
Zalmunna,
kings of
Midian.
\par }{\PP \VS{6}¶ And the
princes of
Succoth
said,
{\ADD{Are}} the
hands of
Zebah and
Zalmunna now in thine
hand, that we should
give
bread unto thine
army?
\VS{7}And
Gideon
said, Therefore when the
{\ND{
LORD}} hath
delivered
Zebah and
Zalmunna into mine
hand, then I will
tear your
flesh with the
thorns of the
wilderness and with
briers.\FTNT{§}{{\FR 8.7 }
tear: Heb. thresh}
\par }{\PP \VS{8}¶ And he went
up thence to
Penuel, and
spake unto them
likewise: and the
men of
Penuel
answered him as the
men of
Succoth had
answered
{\ADD{him}}.
\VS{9}And he
spake also unto the
men of
Penuel,
saying, When I come
again in
peace, I will break
down this
tower.
\par }{\PP \VS{10}¶ Now
Zebah and
Zalmunna
{\ADD{were}} in
Karkor, and their
hosts with them, about
fifteen
thousand
{\ADD{men}}, all that were
left of all the
hosts of the
children of the
east: for there
fell an
hundred and
twenty
thousand
men that
drew
sword.\FTNT{**}{{\FR 8.10 }
men that…: or, every one drawing a sword}
\par }{\PP \VS{11}¶ And
Gideon went
up by the
way of them that
dwelt in
tents on the
east of
Nobah and
Jogbehah, and
smote the
host: for the
host was
secure.
\VS{12}And when
Zebah and
Zalmunna
fled, he
pursued
after them, and
took the
two
kings of
Midian,
Zebah and
Zalmunna, and
discomfited all the
host.\FTNT{††}{{\FR 8.12 }
discomfited: Heb. terrified}
\par }{\PP \VS{13}¶ And
Gideon the
son of
Joash
returned from
battle
before the
sun
{\ADD{was up}},
\VS{14}And
caught a young
man of the
men of
Succoth, and
enquired of him: and he
described unto him the
princes of
Succoth, and the
elders thereof,
{\ADD{even}} threescore and
seventeen
men.\FTNT{‡‡}{{\FR 8.14 }
described: Heb. writ}
\VS{15}And he
came unto the
men of
Succoth, and
said, Behold
Zebah and
Zalmunna, with whom ye did
upbraid me,
saying,
{\ADD{Are}} the
hands of
Zebah and
Zalmunna now in thine
hand, that we should
give
bread unto thy
men
{\ADD{that are}}
weary?
\VS{16}And he
took the
elders of the
city, and
thorns of the
wilderness and
briers, and with them he
taught the
men of
Succoth.\FTNT{§§}{{\FR 8.16 }
taught: Heb. made to know}
\VS{17}And he beat
down the
tower of
Penuel, and
slew the
men of the
city.
\par }{\PP \VS{18}¶ Then
said he unto
Zebah and
Zalmunna,
What manner of
men
{\ADD{were they}} whom ye
slew at
Tabor? And they
answered, As
thou
{\ADD{art}}, so
{\ADD{were}} they; each
one
resembled the
children of a
king.\FTNT{***}{{\FR 8.18 }
resembled…: Heb. according to the form, etc}
\VS{19}And he
said, They
{\ADD{were}} my
brethren,
{\ADD{even}} the
sons of my
mother:
{\ADD{as}} the
{\ND{
LORD}}
liveth,
if ye had saved them
alive, I would not
slay you.
\VS{20}And he
said unto
Jether his
firstborn,
Up,
{\ADD{and}}
slay them. But the
youth
drew not his
sword: for he
feared, because he
{\ADD{was}} yet a
youth.
\VS{21}Then
Zebah and
Zalmunna
said,
Rise thou, and
fall upon us: for as the
man
{\ADD{is, so is}} his
strength. And
Gideon
arose, and
slew
Zebah and
Zalmunna, and took
away the
ornaments that
{\ADD{were}} on their
camels’
necks.\FTNT{†††}{{\FR 8.21 }
ornaments: or, ornaments like the moon}
\par }{\PP \VS{22}¶ Then the
men of
Israel
said unto
Gideon,
Rule thou over us, both thou, and thy
son, and thy
son’s
son also: for thou hast
delivered us from the
hand of
Midian.
\VS{23}And
Gideon
said unto them, I will not
rule over you, neither shall my
son
rule over you: the
{\ND{
LORD}} shall
rule over you.
\par }{\PP \VS{24}¶ And
Gideon
said unto them, I would
desire a
request of you, that ye would
give me every
man the
earrings of his
prey. (For they had
golden
earrings, because they
{\ADD{were}}
Ishmaelites.)
\VS{25}And they
answered, We will
willingly
give
{\ADD{them}}. And they
spread a
garment, and did
cast therein every
man the
earrings of his
prey.
\VS{26}And the
weight of the
golden
earrings that he
requested was a
thousand and
seven
hundred
{\ADD{shekels}} of
gold; beside
ornaments, and
collars, and
purple
raiment that
{\ADD{was}} on the
kings of
Midian, and beside the
chains that
{\ADD{were}} about their
camels’
necks.\FTNT{‡‡‡}{{\FR 8.26 }
collars: or, sweet jewels}
\VS{27}And
Gideon
made an
ephod thereof, and
put it in his
city,
{\ADD{even}} in
Ophrah: and all
Israel went thither a
whoring
after it: which thing became a
snare unto
Gideon, and to his
house.
\par }{\PP \VS{28}¶ Thus was
Midian
subdued
before the
children of
Israel, so that they lifted
up their
heads no
more. And the
country was in
quietness
forty
years in the
days of
Gideon.
\par }{\PP \VS{29}¶ And
Jerubbaal the
son of
Joash
went and
dwelt in his own
house.
\VS{30}And
Gideon had threescore and
ten
sons of his
body
begotten: for he had
many
wives.\FTNT{§§§}{{\FR 8.30 }
of his…: Heb. going out of his thigh}
\VS{31}And his
concubine that
{\ADD{was}} in
Shechem, she also
bare him a
son, whose
name he
called
Abimelech.\FTNT{*}{{\FR 8.31 }
called: Heb. set}
\par }{\PP \VS{32}¶ And
Gideon the
son of
Joash
died in a
good old
age, and was
buried in the
sepulchre of
Joash his
father, in
Ophrah of the
Abi-ezrites.
\VS{33}And it came to pass, as soon as
Gideon was
dead, that the
children of
Israel turned
again, and went a
whoring
after
Baalim, and
made
Baal-berith their
god.
\VS{34}And the
children of
Israel
remembered not the
{\ND{
LORD}} their
God, who had
delivered them out of the
hands of all their
enemies on every
side:
\VS{35}Neither
shewed they
kindness to the
house of
Jerubbaal,
{\ADD{namely}},
Gideon, according to all the
goodness which he had
shewed unto
Israel.

\par }\Chap{9}{\PP \VerseOne{1}And
Abimelech the
son of
Jerubbaal
went to
Shechem unto his
mother’s
brethren, and
communed with them, and with all the
family of the
house of his
mother’s
father,
saying,
\VS{2}Speak, I pray you, in the
ears of all the
men of
Shechem, Whether
{\ADD{is}}
better for you, either that all the
sons of
Jerubbaal,
{\ADD{which are}} threescore and
ten
persons,
reign over you, or that
one
reign over you?
remember also that I
{\ADD{am}} your
bone and your
flesh.\FTNT{*}{{\FR 9.2 }
Whether…: Heb. What is good? whether, etc}
\VS{3}And his
mother’s
brethren
spake of him in the
ears of all the
men of
Shechem all these
words: and their
hearts
inclined to
follow
Abimelech; for they
said, He
{\ADD{is}} our
brother.\FTNT{†}{{\FR 9.3 }
to follow: Heb. after}
\VS{4}And they
gave him threescore and
ten
{\ADD{pieces}} of
silver out of the
house of
Baal-berith, wherewith
Abimelech
hired
vain and
light
persons, which
followed
him.
\VS{5}And he
went unto his
father’s
house at
Ophrah, and
slew his
brethren the
sons of
Jerubbaal,
{\ADD{being}} threescore and
ten
persons, upon
one
stone: notwithstanding yet
Jotham the
youngest
son of
Jerubbaal was
left; for he
hid himself.
\VS{6}And all the
men of
Shechem gathered
together, and all the
house of
Millo, and
went, and
made
Abimelech
king, by the
plain of the
pillar that
{\ADD{was}} in
Shechem.\FTNT{‡}{{\FR 9.6 }
plain: or, oak}
\par }{\PP \VS{7}¶ And when they
told
{\ADD{it}} to
Jotham, he
went and
stood in the
top of
mount
Gerizim, and lifted
up his
voice, and
cried, and
said unto them,
Hearken unto me, ye
men of
Shechem, that
God may
hearken unto you.
\VS{8}The
trees went
forth
{\ADD{on a time}} to
anoint a
king over them; and they
said unto the olive
tree,
Reign thou over us.
\VS{9}But the olive
tree
said unto them, Should I
leave my
fatness, wherewith by me they
honour
God and
man, and
go to be
promoted over the
trees?\FTNT{§}{{\FR 9.9 }
go…: or, go up and down for other trees}
\VS{10}And the
trees
said to the fig
tree,
Come thou,
{\ADD{and}}
reign over us.
\VS{11}But the fig
tree
said unto them, Should I
forsake my
sweetness, and my
good
fruit, and
go to be
promoted over the
trees?
\VS{12}Then
said the
trees unto the
vine,
Come thou,
{\ADD{and}}
reign over us.
\VS{13}And the
vine
said unto them, Should I
leave my
wine, which
cheereth
God and
man, and
go to be
promoted over the
trees?
\VS{14}Then
said all the
trees unto the
bramble,
Come thou,
{\ADD{and}}
reign over us.\FTNT{**}{{\FR 9.14 }
bramble: or, thistle}
\VS{15}And the
bramble
said unto the
trees, If in
truth ye
anoint me
king over you,
{\ADD{then}}
come
{\ADD{and}} put your
trust in my
shadow: and if not, let
fire come
out of the
bramble, and
devour the
cedars of
Lebanon.
\VS{16}Now therefore, if ye have
done
truly and
sincerely, in that ye have made
Abimelech
king, and if ye have
dealt
well with
Jerubbaal and his
house, and have
done unto him according to the
deserving of his
hands;
\VS{17}(For my
father
fought for you, and
adventured his
life
far, and
delivered you out of the
hand of
Midian:\FTNT{††}{{\FR 9.17 }
adventured…: Heb. cast his life}
\VS{18}And ye are risen
up against my
father’s
house this
day, and have
slain his
sons, threescore and
ten
persons, upon
one
stone, and have made
Abimelech, the
son of his
maidservant,
king over the
men of
Shechem, because he
{\ADD{is}} your
brother;)
\VS{19}If ye then have
dealt
truly and
sincerely with
Jerubbaal and with his
house this
day,
{\ADD{then}}
rejoice ye in
Abimelech, and let him also
rejoice in you:
\VS{20}But if not, let
fire come
out from
Abimelech, and
devour the
men of
Shechem, and the
house of
Millo; and let
fire come
out from the
men of
Shechem, and from the
house of
Millo, and
devour
Abimelech.
\VS{21}And
Jotham ran
away, and
fled, and
went to
Beer, and
dwelt there, for
fear of
Abimelech his
brother.
\par }{\PP \VS{22}¶ When
Abimelech had
reigned
three
years over
Israel,
\VS{23}Then
God
sent an
evil
spirit between
Abimelech and the
men of
Shechem; and the
men of
Shechem dealt
treacherously with
Abimelech:
\VS{24}That the
cruelty
{\ADD{done}} to the threescore and
ten
sons of
Jerubbaal might
come, and their
blood be
laid upon
Abimelech their
brother, which
slew them; and upon the
men of
Shechem, which
aided him in the
killing of his
brethren.\FTNT{‡‡}{{\FR 9.24 }
aided…: Heb. strengthened his hands to kill}
\VS{25}And the
men of
Shechem
set liers in
wait for him in the
top of the
mountains, and they
robbed all that
came along that
way by them: and it was
told
Abimelech.
\VS{26}And
Gaal the
son of
Ebed
came with his
brethren, and went
over to
Shechem: and the
men of
Shechem put their
confidence in him.
\VS{27}And they went
out into the
fields, and
gathered their
vineyards, and
trode
{\ADD{the grapes}}, and
made
merry, and
went into the
house of their
god, and did
eat and
drink, and
cursed
Abimelech.\FTNT{§§}{{\FR 9.27 }
merry: or, songs}
\VS{28}And
Gaal the
son of
Ebed
said, Who
{\ADD{is}}
Abimelech, and who
{\ADD{is}}
Shechem, that we should
serve him?
{\ADD{is}} not
{\ADD{he}} the
son of
Jerubbaal? and
Zebul his
officer?
serve the
men of
Hamor the
father of
Shechem: for why should we
serve him?
\VS{29}And would to God this
people were
under my
hand! then would I
remove
Abimelech. And he
said to
Abimelech,
Increase thine
army, and come
out.
\par }{\PP \VS{30}¶ And when
Zebul the
ruler of the
city
heard the
words of
Gaal the
son of
Ebed, his
anger was
kindled.\FTNT{***}{{\FR 9.30 }
kindled: or, hot}
\VS{31}And he
sent
messengers unto
Abimelech
privily,
saying, Behold,
Gaal the
son of
Ebed and his
brethren be
come to
Shechem; and, behold, they
fortify the
city against thee.\FTNT{†††}{{\FR 9.31 }
privily: Heb. craftily or, to Tormah}
\VS{32}Now therefore
up by
night, thou and the
people that
{\ADD{is}} with thee, and lie in
wait in the
field:
\VS{33}And it shall be,
{\ADD{that}} in the
morning, as soon as the
sun is
up, thou shalt rise
early, and
set upon the
city: and, behold,
{\ADD{when}} he and the
people that
{\ADD{is}} with him come
out against thee, then mayest thou
do to them as
thou shalt find
occasion.\FTNT{‡‡‡}{{\FR 9.33 }
as thou shalt…: Heb. as thine hand shall find}
\par }{\PP \VS{34}¶ And
Abimelech rose
up, and all the
people that
{\ADD{were}} with him, by
night, and they laid
wait against
Shechem in
four
companies.
\VS{35}And
Gaal the
son of
Ebed went
out, and
stood in the
entering of the
gate of the
city: and
Abimelech rose
up, and the
people that
{\ADD{were}} with him, from lying in
wait.
\VS{36}And when
Gaal
saw the
people, he
said to
Zebul, Behold, there
come
people
down from the
top of the
mountains. And
Zebul
said unto him, Thou
seest the
shadow of the
mountains as
{\ADD{if they were}}
men.
\VS{37}And
Gaal
spake
again and
said, See there
come
people
down by the
middle of the
land, and
another
company
come along by the
plain
of
Meonenim.\FTNT{§§§}{{\FR 9.37 }
middle: Heb. navel}\FTNT{*}{{\FR 9.37 }
Meonenim: or, the regarders of times}
\VS{38}Then
said
Zebul unto him, Where
{\ADD{is}}
now thy
mouth, wherewith thou
saidst, Who
{\ADD{is}}
Abimelech, that we should
serve him?
{\ADD{is}} not this the
people that thou hast
despised? go
out, I pray
now, and
fight with them.
\VS{39}And
Gaal went
out
before the
men of
Shechem, and
fought with
Abimelech.
\VS{40}And
Abimelech
chased him, and he
fled
before him, and
many were
overthrown
{\ADD{and}}
wounded,
{\ADD{even}} unto the
entering of the
gate.
\VS{41}And
Abimelech
dwelt at
Arumah: and
Zebul thrust
out
Gaal and his
brethren, that they should not
dwell in
Shechem.
\VS{42}And it came to pass on the
morrow, that the
people went
out into the
field; and they
told
Abimelech.
\VS{43}And he
took the
people, and
divided them into
three
companies, and laid
wait in the
field, and
looked, and, behold, the
people
{\ADD{were}} come
forth out of the
city; and he rose
up against them, and
smote them.
\VS{44}And
Abimelech, and the
company that
{\ADD{was}} with him, rushed
forward, and
stood in the
entering of the
gate of the
city: and the
two
{\ADD{other}}
companies ran
upon all
{\ADD{the people}} that
{\ADD{were}} in the
fields, and
slew them.
\VS{45}And
Abimelech
fought against the
city all that
day; and he
took the
city, and
slew the
people that
{\ADD{was}} therein, and beat
down the
city, and
sowed it with
salt.
\par }{\PP \VS{46}¶ And when all the
men of the
tower of
Shechem
heard
{\ADD{that}}, they
entered into an
hold of the
house of the
god
Berith.
\VS{47}And it was
told
Abimelech, that all the
men of the
tower of
Shechem were gathered
together.
\VS{48}And
Abimelech gat him
up to
mount
Zalmon, he and all the
people that
{\ADD{were}} with him; and
Abimelech
took an
axe in his
hand, and cut
down a
bough from the
trees, and
took it, and
laid
{\ADD{it}} on his
shoulder, and
said unto the
people that
{\ADD{were}} with him, What ye have
seen me
do, make
haste,
{\ADD{and}}
do as
I
{\ADD{have done}}.\FTNT{†}{{\FR 9.48 }
me do: Heb. I have done}
\VS{49}And all the
people likewise cut
down every
man his
bough, and
followed
Abimelech, and
put
{\ADD{them}} to the
hold, and
set the
hold on
fire upon them; so that all the
men of the
tower of
Shechem
died also, about a
thousand
men and
women.
\par }{\PP \VS{50}¶ Then
went
Abimelech to
Thebez, and
encamped against
Thebez, and
took it.
\VS{51}But there was a
strong
tower
within the
city, and thither
fled all the
men and
women, and all
they of the
city, and
shut
{\ADD{it}} to them, and gat them
up to the
top of the
tower.
\VS{52}And
Abimelech
came unto the
tower, and
fought against it, and went
hard unto the
door of the
tower to
burn it with
fire.
\VS{53}And a
certain
woman
cast a
piece of a
millstone upon
Abimelech’s
head, and all to
brake his
skull.
\VS{54}Then he
called
hastily unto the young
man his
armourbearer, and
said unto him,
Draw thy
sword, and
slay me, that men
say not of me, A
woman
slew him. And his young
man thrust him
through, and he
died.
\VS{55}And when the
men of
Israel
saw that
Abimelech was
dead, they
departed every
man unto his
place.
\par }{\PP \VS{56}¶ Thus
God
rendered the
wickedness of
Abimelech, which he
did unto his
father, in
slaying his
seventy
brethren:
\VS{57}And all the
evil of the
men of
Shechem did
God
render upon their
heads: and upon them
came the
curse of
Jotham the
son of
Jerubbaal.

\par }\Chap{10}{\PP \VerseOne{1}And
after
Abimelech there
arose to
defend
Israel
Tola the
son of
Puah, the
son of
Dodo, a
man of
Issachar; and he
dwelt in
Shamir in
mount
Ephraim.\FTNT{*}{{\FR 10.1 }
defend: or, deliver: Heb. save}
\VS{2}And he
judged
Israel
twenty and
three
years, and
died, and was
buried in
Shamir.
\par }{\PP \VS{3}¶ And
after him
arose
Jair, a
Gileadite, and
judged
Israel
twenty and
two
years.
\VS{4}And he had
thirty
sons that
rode on
thirty ass
colts, and they had
thirty
cities, which are
called
Havoth-jair unto this
day, which
{\ADD{are}} in the
land of
Gilead.\FTNT{†}{{\FR 10.4 }
Havoth-jair: or, the villages of Jair}
\VS{5}And
Jair
died, and was
buried in
Camon.
\par }{\PP \VS{6}¶ And the
children of
Israel
did
evil
again in the
sight of the
{\ND{
LORD}}, and
served
Baalim, and
Ashtaroth, and the
gods of
Syria, and the
gods of
Zidon, and the
gods of
Moab, and the
gods of the
children of
Ammon, and the
gods of the
Philistines, and
forsook the
{\ND{
LORD}}, and
served not him.
\VS{7}And the
anger of the
{\ND{
LORD}} was
hot against
Israel, and he
sold them into the
hands of the
Philistines, and into the
hands of the
children of
Ammon.
\VS{8}And that
year they
vexed and
oppressed the
children of
Israel:
eighteen
years, all the
children of
Israel that
{\ADD{were}} on the other
side
Jordan in the
land of the
Amorites, which
{\ADD{is}} in
Gilead.\FTNT{‡}{{\FR 10.8 }
oppressed: Heb. crushed}
\VS{9}Moreover the
children of
Ammon passed
over
Jordan to
fight also against
Judah, and against
Benjamin, and against the
house of
Ephraim; so that
Israel was
sore
distressed.
\par }{\PP \VS{10}¶ And the
children of
Israel
cried unto the
{\ND{
LORD}},
saying, We have
sinned against thee, both because we have
forsaken our
God, and also
served
Baalim.
\VS{11}And the
{\ND{
LORD}}
said unto the
children of
Israel,
{\ADD{Did}} not
{\ADD{I deliver you}} from the
Egyptians, and from the
Amorites, from the
children of
Ammon, and from the
Philistines?
\VS{12}The
Zidonians also, and the
Amalekites, and the
Maonites, did
oppress you; and ye
cried to me, and I
delivered you out of their
hand.
\VS{13}Yet ye have
forsaken me, and
served
other
gods: wherefore I will
deliver you no
more.
\VS{14}Go and
cry unto the
gods which ye have
chosen; let
them
deliver you in the
time of your
tribulation.
\par }{\PP \VS{15}¶ And the
children of
Israel
said unto the
{\ND{
LORD}}, We have
sinned:
do thou unto us whatsoever
seemeth
good unto thee;
deliver us only, we pray thee, this
day.\FTNT{§}{{\FR 10.15 }
seemeth…: Heb. is good in thine eyes}
\VS{16}And they put
away the
strange
gods from
among them, and
served the
{\ND{
LORD}}: and his
soul was
grieved for the
misery of
Israel.\FTNT{**}{{\FR 10.16 }
strange…: Heb. gods of strangers}\FTNT{††}{{\FR 10.16 }
grieved: Heb. shortened}
\VS{17}Then the
children of
Ammon were gathered
together, and
encamped in
Gilead. And the
children of
Israel assembled themselves
together, and
encamped in
Mizpeh.\FTNT{‡‡}{{\FR 10.17 }
gathered: Heb. cried}
\VS{18}And the
people
{\ADD{and}}
princes of
Gilead
said
one to
another, What
man
{\ADD{is he}} that will
begin to
fight against the
children of
Ammon? he shall be
head over all the
inhabitants of
Gilead.

\par }\Chap{11}{\PP \VerseOne{1}Now
Jephthah the
Gileadite was a mighty
man of
valour, and he
{\ADD{was}} the
son of an
harlot: and
Gilead
begat
Jephthah.\FTNT{*}{{\FR 11.1 }
Jephthah: Gr. Jephthae}\FTNT{†}{{\FR 11.1 }
an harlot: Heb. a woman an harlot}
\VS{2}And
Gilead’s
wife
bare him
sons; and his
wife’s
sons grew
up, and they thrust
out
Jephthah, and
said unto him, Thou shalt not
inherit in our
father’s
house; for thou
{\ADD{art}} the
son of a
strange
woman.
\VS{3}Then
Jephthah
fled
from his
brethren, and
dwelt in the
land of
Tob: and there were
gathered
vain
men to
Jephthah, and went
out with him.\FTNT{‡}{{\FR 11.3 }
from: Heb. from the face of}
\par }{\PP \VS{4}¶ And it came to pass in process of
time, that the
children of
Ammon made
war against
Israel.\FTNT{§}{{\FR 11.4 }
in process…: Heb. after days}
\VS{5}And it was so, that when the
children of
Ammon made
war against
Israel, the
elders of
Gilead
went to
fetch
Jephthah out of the
land of
Tob:
\VS{6}And they
said unto
Jephthah,
Come, and be our
captain, that we may
fight with the
children of
Ammon.
\VS{7}And
Jephthah
said unto the
elders of
Gilead, Did not ye
hate me, and
expel me out of my
father’s
house? and why are ye
come unto me now when ye are in
distress?
\VS{8}And the
elders of
Gilead
said unto
Jephthah, Therefore we turn
again to thee now, that thou mayest
go with us, and
fight against the
children of
Ammon, and be our
head over all the
inhabitants of
Gilead.
\VS{9}And
Jephthah
said unto the
elders of
Gilead, If ye bring me home
again to
fight against the
children of
Ammon, and the
{\ND{
LORD}}
deliver them
before me, shall I be your
head?
\VS{10}And the
elders of
Gilead
said unto
Jephthah, The
{\ND{
LORD}} be
witness between us, if we
do not so according to thy
words.\FTNT{**}{{\FR 11.10 }
witness…: Heb. the hearer between us}
\VS{11}Then
Jephthah
went with the
elders of
Gilead, and the
people
made him
head and
captain over them: and
Jephthah
uttered all his
words
before the
{\ND{
LORD}} in
Mizpeh.
\par }{\PP \VS{12}¶ And
Jephthah
sent
messengers unto the
king of the
children of
Ammon,
saying, What hast thou to do with me, that thou art
come against me to
fight in my
land?
\VS{13}And the
king of the
children of
Ammon
answered unto the
messengers of
Jephthah, Because
Israel took
away my
land, when they came
up out of
Egypt, from
Arnon even unto
Jabbok, and unto
Jordan: now therefore restore those
{\ADD{lands}}
again
peaceably.
\VS{14}And
Jephthah
sent
messengers
again unto the
king of the
children of
Ammon:
\VS{15}And
said unto him, Thus
saith
Jephthah,
Israel took not
away the
land of
Moab, nor the
land of the
children of
Ammon:
\VS{16}But when
Israel came
up from
Egypt, and
walked through the
wilderness unto the
Red
sea, and
came to
Kadesh;
\VS{17}Then
Israel
sent
messengers unto the
king of
Edom,
saying, Let me, I pray thee, pass
through thy
land: but the
king of
Edom would not
hearken
{\ADD{thereto}}. And in like manner they
sent unto the
king of
Moab: but he
would not
{\ADD{consent:}} and
Israel
abode in
Kadesh.
\VS{18}Then they went
along through the
wilderness, and
compassed the
land of
Edom, and the
land of
Moab, and
came by the east
side of the
land of
Moab, and
pitched on the other
side of
Arnon, but
came not within the
border of
Moab: for
Arnon
{\ADD{was}} the
border of
Moab.
\VS{19}And
Israel
sent
messengers unto
Sihon
king of the
Amorites, the
king of
Heshbon; and
Israel
said unto him, Let us
pass, we pray thee, through thy
land
into my
place.
\VS{20}But
Sihon
trusted not
Israel to
pass through his
coast: but
Sihon
gathered all his
people
together, and
pitched in
Jahaz, and
fought against
Israel.
\VS{21}And the
{\ND{
LORD}}
God of
Israel
delivered
Sihon and all his
people into the
hand of
Israel, and they
smote them: so
Israel
possessed all the
land of the
Amorites, the
inhabitants of that
country.
\VS{22}And they
possessed all the
coasts of the
Amorites, from
Arnon even unto
Jabbok, and from the
wilderness even unto
Jordan.
\VS{23}So now the
{\ND{
LORD}}
God of
Israel hath
dispossessed the
Amorites from
before his
people
Israel, and shouldest thou
possess it?
\VS{24}Wilt not thou
possess that which
Chemosh thy
god giveth thee to
possess? So whomsoever the
{\ND{
LORD}} our
God shall drive
out from
before us, them will we
possess.
\VS{25}And now
{\ADD{art}} thou any thing
better than
Balak the
son of
Zippor,
king of
Moab? did he
ever
strive against
Israel, or did he
ever
fight against them,
\VS{26}While
Israel
dwelt in
Heshbon and her
towns, and in
Aroer and her
towns, and in all the
cities that
{\ADD{be}} along by the
coasts of
Arnon,
three
hundred
years? why therefore did ye not
recover
{\ADD{them}} within that
time?
\VS{27}Wherefore I have not
sinned against thee, but thou
doest me
wrong to
war against me: the
{\ND{
LORD}} the
Judge be
judge this
day between the
children of
Israel and the
children of
Ammon.
\VS{28}Howbeit the
king of the
children of
Ammon
hearkened not unto the
words of
Jephthah which he
sent him.
\par }{\PP \VS{29}¶ Then the
Spirit of the
{\ND{
LORD}} came upon
Jephthah, and he passed
over
Gilead, and
Manasseh, and passed
over
Mizpeh of
Gilead, and from
Mizpeh of
Gilead he passed
over
{\ADD{unto}} the
children of
Ammon.
\VS{30}And
Jephthah
vowed a
vow unto the
{\ND{
LORD}}, and
said, If thou shalt without
fail
deliver the
children of
Ammon into mine
hands,
\VS{31}Then it shall be, that
whatsoever cometh
forth of the
doors of my
house to
meet me, when I
return in
peace from the
children of
Ammon, shall surely be the
{\ND{
LORD’s}}, and I will offer it
up for a burnt
offering.\FTNT{††}{{\FR 11.31 }
whatsoever…: Heb. that which cometh forth, which shall come forth}\FTNT{‡‡}{{\FR 11.31 }
and…: or, or I will offer it, etc}
\par }{\PP \VS{32}¶ So
Jephthah passed
over unto the
children of
Ammon to
fight against them; and the
{\ND{
LORD}}
delivered them into his
hands.
\VS{33}And he
smote them from
Aroer, even till thou
come to
Minnith,
{\ADD{even}}
twenty
cities, and unto the
plain of the
vineyards, with a
very
great
slaughter. Thus the
children of
Ammon were
subdued
before the
children of
Israel.\FTNT{§§}{{\FR 11.33 }
the plain: or, Abel}
\par }{\PP \VS{34}¶ And
Jephthah
came to
Mizpeh unto his
house, and, behold, his
daughter came
out to
meet him with
timbrels and with
dances: and she
{\ADD{was his}} only
child; beside her he had neither
son
nor
daughter.\FTNT{***}{{\FR 11.34 }
beside…: or, he had not of his own either son or daughter: Heb. of himself}
\VS{35}And it came to pass, when he
saw her, that he
rent his
clothes, and
said,
Alas, my
daughter! thou hast
brought me
very
low, and thou art one of them that
trouble me: for I have
opened my
mouth unto the
{\ND{
LORD}}, and I
cannot go
back.
\VS{36}And she
said unto him, My
father,
{\ADD{if}} thou hast
opened thy
mouth unto the
{\ND{
LORD}},
do to me according to that
which hath
proceeded out of thy
mouth;
forasmuch as the
{\ND{
LORD}} hath
taken
vengeance for thee of thine
enemies,
{\ADD{even}} of the
children of
Ammon.
\VS{37}And she
said unto her
father, Let this
thing be
done for me: let me
alone
two
months, that I may go
up and
down upon the
mountains, and
bewail my
virginity, I and my
fellows.\FTNT{†††}{{\FR 11.37 }
go up…: Heb. go and go down}
\VS{38}And he
said,
Go. And he sent her
away
{\ADD{for}}
two
months: and she
went with her
companions, and
bewailed her
virginity upon the
mountains.
\VS{39}And it came to pass at the
end of
two
months, that she
returned unto her
father, who
did with her
{\ADD{according}} to his
vow which he had
vowed: and she
knew no
man. And it was a
custom in
Israel,\FTNT{‡‡‡}{{\FR 11.39 }
custom: or, ordinance}
\VS{40}{\ADD{That}} the
daughters of
Israel
went
yearly to
lament the
daughter of
Jephthah the
Gileadite
four
days in a
year.\FTNT{§§§}{{\FR 11.40 }
yearly: Heb. from year to year}\FTNT{*}{{\FR 11.40 }
to lament: or, to talk with}

\par }\Chap{12}{\PP \VerseOne{1}And the
men of
Ephraim gathered themselves
together, and
went
northward, and
said unto
Jephthah, Wherefore passedst thou
over to
fight against the
children of
Ammon, and didst not
call us to
go with thee? we will
burn thine
house upon thee with
fire.\FTNT{*}{{\FR 12.1 }
gathered…: Heb. were called}
\VS{2}And
Jephthah
said unto them, I and my
people
were at
great
strife with the
children of
Ammon; and when I
called you, ye
delivered me not out of their
hands.
\VS{3}And when I
saw that ye
delivered
{\ADD{me}} not, I
put my
life in my
hands, and passed
over against the
children of
Ammon, and the
{\ND{
LORD}}
delivered them into my
hand: wherefore then are ye come
up unto me this
day, to
fight against me?
\VS{4}Then
Jephthah gathered
together all the
men of
Gilead, and
fought with
Ephraim: and the
men of
Gilead
smote
Ephraim, because they
said, Ye
Gileadites
{\ADD{are}}
fugitives of
Ephraim
among the
Ephraimites,
{\ADD{and}}
among the
Manassites.
\VS{5}And the
Gileadites
took the
passages of
Jordan before the
Ephraimites: and it was
{\ADD{so}}, that when those
Ephraimites which were
escaped
said, Let me go
over; that the
men of
Gilead
said unto him,
{\ADD{Art}} thou an
Ephraimite? If he
said, Nay;
\VS{6}Then
said they unto him,
Say now
Shibboleth: and he
said
Sibboleth: for he could not
frame to
pronounce
{\ADD{it}}
right. Then they
took him, and
slew him at the
passages of
Jordan: and there
fell at that
time of the
Ephraimites
forty and
two
thousand.
\VS{7}And
Jephthah
judged
Israel
six
years. Then
died
Jephthah the
Gileadite, and was
buried in
{\ADD{one of}} the
cities of
Gilead.
\par }{\PP \VS{8}¶ And
after him
Ibzan of
Beth-lehem
judged
Israel.
\VS{9}And he had
thirty
sons, and
thirty
daughters,
{\ADD{whom}} he
sent
abroad, and
took in
thirty
daughters from
abroad for his
sons. And he
judged
Israel
seven
years.
\VS{10}Then
died
Ibzan, and was
buried at
Beth-lehem.
\par }{\PP \VS{11}¶ And
after him
Elon, a
Zebulonite,
judged
Israel; and he
judged
Israel
ten
years.
\VS{12}And
Elon the
Zebulonite
died, and was
buried in
Aijalon in the
country of
Zebulun.
\par }{\PP \VS{13}¶ And
after him
Abdon the
son of
Hillel, a
Pirathonite,
judged
Israel.
\VS{14}And he had
forty
sons and
thirty
nephews, that
rode on threescore and
ten ass
colts: and he
judged
Israel
eight
years.\FTNT{†}{{\FR 12.14 }
nephews: Heb. sons’ sons}
\VS{15}And
Abdon the
son of
Hillel the
Pirathonite
died, and was
buried in
Pirathon in the
land of
Ephraim, in the
mount of the
Amalekites.

\par }\Chap{13}{\PP \VerseOne{1}And the
children of
Israel
did
evil
again in the
sight of the
{\ND{
LORD}}; and the
{\ND{
LORD}}
delivered them into the
hand of the
Philistines
forty
years.\FTNT{*}{{\FR 13.1 }
did evil…: Heb. added to commit, etc}
\par }{\PP \VS{2}¶ And there was a
certain
man of
Zorah, of the
family of the
Danites, whose
name
{\ADD{was}}
Manoah; and his
wife
{\ADD{was}}
barren, and
bare not.
\VS{3}And the
angel of the
{\ND{
LORD}}
appeared unto the
woman, and
said unto her, Behold now, thou
{\ADD{art}}
barren, and
bearest not: but thou shalt
conceive, and
bear a
son.
\VS{4}Now therefore
beware, I pray thee, and
drink not
wine nor strong
drink, and
eat not any
unclean
{\ADD{thing:}}
\VS{5}For, lo, thou shalt
conceive, and
bear a
son; and no
razor shall
come on his
head: for the
child shall be a
Nazarite unto
God from the
womb: and he shall
begin to
deliver
Israel out of the
hand of the
Philistines.
\par }{\PP \VS{6}¶ Then the
woman
came and
told her
husband,
saying, A
man of
God
came unto me, and his
countenance
{\ADD{was}} like the
countenance of an
angel of
God,
very
terrible: but I
asked him not whence he
{\ADD{was}}, neither
told he me his
name:
\VS{7}But he
said unto me, Behold, thou shalt
conceive, and
bear a
son; and now
drink no
wine nor strong
drink, neither
eat any
unclean
{\ADD{thing:}} for the
child shall be a
Nazarite to
God from the
womb to the
day of his
death.
\par }{\PP \VS{8}¶ Then
Manoah
intreated the
{\ND{
LORD}}, and
said,
O my
Lord,
let the
man of
God which thou didst
send come
again unto us, and
teach us what we shall
do unto the
child that shall be
born.
\VS{9}And
God
hearkened to the
voice of
Manoah; and the
angel of
God came
again unto the
woman as she
sat in the
field: but
Manoah her
husband
{\ADD{was}} not with her.
\VS{10}And the
woman made
haste, and
ran, and
shewed her
husband, and
said unto him, Behold, the
man hath
appeared unto me, that
came unto me the
{\ADD{other}}
day.
\VS{11}And
Manoah
arose, and
went
after his
wife, and
came to the
man, and
said unto him,
{\ADD{Art}} thou the
man that
spakest unto the
woman? And he
said, I
{\ADD{am}}.
\VS{12}And
Manoah
said, Now let thy
words come to
pass. How shall we
order the
child, and
{\ADD{how}} shall we
do unto him?\FTNT{†}{{\FR 13.12 }
How shall we order…: Heb. What shall be the manner of the, etc}\FTNT{‡}{{\FR 13.12 }
how shall we do…: or, what shall he do?: Heb. what shall be his work?}
\VS{13}And the
angel of the
{\ND{
LORD}}
said unto
Manoah, Of all that I
said unto the
woman let her
beware.
\VS{14}She may not
eat of any
{\ADD{thing}} that
cometh of the
vine, neither let her
drink
wine or strong
drink, nor
eat any
unclean
{\ADD{thing:}} all that I
commanded her let her
observe.
\par }{\PP \VS{15}¶ And
Manoah
said unto the
angel of the
{\ND{
LORD}}, I pray thee, let us
detain thee, until we shall have made
ready a
kid
for thee.\FTNT{§}{{\FR 13.15 }
for thee: Heb. before thee}
\VS{16}And the
angel of the
{\ND{
LORD}}
said unto
Manoah, Though thou
detain me, I will not
eat of thy
bread: and
if thou wilt
offer a burnt
offering, thou must
offer it unto the
{\ND{
LORD}}. For
Manoah
knew not that he
{\ADD{was}} an
angel of the
{\ND{
LORD}}.
\VS{17}And
Manoah
said unto the
angel of the
{\ND{
LORD}}, What
{\ADD{is}} thy
name, that when thy
sayings come to
pass we may do thee
honour?
\VS{18}And the
angel of the
{\ND{
LORD}}
said unto him, Why
askest thou thus after my
name, seeing it
{\ADD{is}}
secret?\FTNT{**}{{\FR 13.18 }
secret: or, wonderful}
\VS{19}So
Manoah
took a
kid with a meat
offering, and
offered
{\ADD{it}} upon a
rock unto the
{\ND{
LORD}}: and
{\ADD{the angel}}
did
wondrously; and
Manoah and his
wife looked
on.
\VS{20}For it came to pass, when the
flame went
up toward
heaven from off the
altar, that the
angel of the
{\ND{
LORD}}
ascended in the
flame of the
altar. And
Manoah and his
wife looked
on
{\ADD{it}}, and fell
on their
faces to the
ground.
\VS{21}But the
angel of the
{\ND{
LORD}} did no
more
appear to
Manoah and to his
wife. Then
Manoah
knew that he
{\ADD{was}} an
angel of the
{\ND{
LORD}}.
\VS{22}And
Manoah
said unto his
wife, We shall
surely
die, because we have
seen
God.
\VS{23}But his
wife
said unto him,
If the
{\ND{
LORD}} were
pleased to
kill us, he would not have
received a burnt
offering and a meat
offering at our
hands, neither would he have
shewed us all these
{\ADD{things}}, nor would as at this
time have
told us
{\ADD{such things}} as
these.
\par }{\PP \VS{24}¶ And the
woman
bare a
son, and
called his
name
Samson: and the
child
grew, and the
{\ND{
LORD}}
blessed him.
\VS{25}And the
Spirit of the
{\ND{
LORD}}
began to move him at
times in the
camp of
Dan between
Zorah and
Eshtaol.\FTNT{††}{{\FR 13.25 }
the camp…: Heb. Mahaneh-dan}

\par }\Chap{14}{\PP \VerseOne{1}And
Samson went
down to
Timnath, and
saw a
woman in
Timnath of the
daughters of the
Philistines.
\VS{2}And he came
up, and
told his
father and his
mother, and
said, I have
seen a
woman in
Timnath of the
daughters of the
Philistines: now therefore
get her for me to
wife.
\VS{3}Then his
father and his
mother
said unto him,
{\ADD{Is there}}
never a
woman among the
daughters of thy
brethren, or among all my
people, that thou
goest to
take a
wife of the
uncircumcised
Philistines? And
Samson
said unto his
father,
Get her for me; for she pleaseth me
well.\FTNT{*}{{\FR 14.3 }
pleaseth…: Heb. is right in mine eyes}
\VS{4}But his
father and his
mother
knew not that it
{\ADD{was}} of the
{\ND{
LORD}}, that he
sought an
occasion against the
Philistines: for at that
time the
Philistines had
dominion over
Israel.
\par }{\PP \VS{5}¶ Then
went
Samson
down, and his
father and his
mother, to
Timnath, and
came to the
vineyards of
Timnath: and, behold, a
young
lion
roared
against him.\FTNT{†}{{\FR 14.5 }
against…: Heb. in meeting him}
\VS{6}And the
Spirit of the
{\ND{
LORD}} came
mightily upon him, and he
rent him as he would have
rent a
kid, and
{\ADD{he had}}
nothing in his
hand: but he
told not his
father or his
mother what he had
done.
\VS{7}And he went
down, and
talked with the
woman; and she
pleased
Samson
well.
\par }{\PP \VS{8}¶ And after a
time he
returned to
take her, and he turned
aside to
see the
carcase of the
lion: and, behold,
{\ADD{there was}} a
swarm of
bees and
honey in the
carcase of the
lion.
\VS{9}And he
took thereof in his
hands, and went
on
eating, and
came to his
father and
mother, and he
gave them, and they did
eat: but he
told not them that he had
taken the
honey out of the
carcase of the
lion.
\par }{\PP \VS{10}¶ So his
father went
down unto the
woman: and
Samson
made there a
feast; for so used the young
men to
do.
\VS{11}And it came to pass, when they
saw him, that they
brought
thirty
companions to be with him.
\par }{\PP \VS{12}¶ And
Samson
said unto them, I will now put
forth a
riddle unto you: if ye can
certainly
declare it me within the
seven
days of the
feast, and find
{\ADD{it}}
out, then I will
give you
thirty
sheets and
thirty
change of
garments:\FTNT{‡}{{\FR 14.12 }
sheets: or, shirts}
\VS{13}But if ye
cannot
declare
{\ADD{it}} me, then shall ye
give me
thirty
sheets and
thirty
change of
garments. And they
said unto him, Put
forth thy
riddle, that we may
hear it.\FTNT{§}{{\FR 14.13 }
sheets: or, shirts}
\VS{14}And he
said unto them, Out of the
eater came
forth
meat, and out of the
strong came
forth
sweetness. And they
could not in
three
days
expound the
riddle.
\VS{15}And it came to pass on the
seventh
day, that they
said unto
Samson’s
wife,
Entice thy
husband, that he may
declare unto us the
riddle, lest we
burn thee and thy
father’s
house with
fire: have ye
called us to take that we
have?
{\ADD{is it}} not
{\ADD{so}}?\FTNT{**}{{\FR 14.15 }
take…: Heb. possess us, or, impoverish us?}
\VS{16}And
Samson’s
wife
wept before him, and
said, Thou dost but
hate me, and
lovest me not: thou hast put
forth a
riddle unto the
children of my
people, and hast not
told
{\ADD{it}} me. And he
said unto her, Behold, I have not
told
{\ADD{it}} my
father nor my
mother, and shall I
tell
{\ADD{it}} thee?
\VS{17}And she
wept before him the
seven
days, while their
feast
lasted: and it came to pass on the
seventh
day, that he
told her, because she lay
sore upon him: and she
told the
riddle to the
children of her
people.\FTNT{††}{{\FR 14.17 }
the seven…: or, the rest of the seven days}
\VS{18}And the
men of the
city
said unto him on the
seventh
day before the
sun went
down, What
{\ADD{is}}
sweeter than
honey? and what
{\ADD{is}}
stronger than a
lion? And he
said unto them,
If ye had not
plowed with my
heifer, ye had not found
out my
riddle.
\par }{\PP \VS{19}¶ And the
Spirit of the
{\ND{
LORD}}
came upon him, and he went
down to
Ashkelon, and
slew
thirty
men of them, and
took their
spoil, and
gave
change of garments unto them which
expounded the
riddle. And his
anger was
kindled, and he went
up to his
father’s
house.\FTNT{‡‡}{{\FR 14.19 }
spoil: or, apparel}
\VS{20}But
Samson’s
wife was
{\ADD{given}} to his
companion, whom he had used as his
friend.

\par }\Chap{15}{\PP \VerseOne{1}But it came to pass within a while
after, in the
time of
wheat
harvest, that
Samson
visited his
wife with a
kid; and he
said, I will go
in to my
wife into the
chamber. But her
father would not
suffer him to go
in.
\VS{2}And her
father
said, I
verily
thought that thou hadst
utterly
hated her; therefore I
gave her to thy
companion:
{\ADD{is}} not her
younger
sister
fairer than she? take her, I pray thee, instead of her.\FTNT{*}{{\FR 15.2 }
take…: Heb. let her be thine}
\par }{\PP \VS{3}¶ And
Samson
said concerning them,
Now shall I be more
blameless than the
Philistines, though I
do them a
displeasure.\FTNT{†}{{\FR 15.3 }
more…: or, blameless from the Philistines though, etc}
\VS{4}And
Samson
went and
caught
three
hundred
foxes, and
took
firebrands, and
turned
tail to
tail, and
put
a
firebrand in the
midst between
two
tails.\FTNT{‡}{{\FR 15.4 }
firebrands: or, torches}
\VS{5}And when he had
set the
brands on
fire, he let
{\ADD{them}}
go into the standing
corn of the
Philistines, and burnt
up both the
shocks, and also the standing
corn, with the
vineyards
{\ADD{and}}
olives.
\par }{\PP \VS{6}¶ Then the
Philistines
said, Who hath
done this? And they
answered,
Samson, the son in
law of the
Timnite, because he had
taken his
wife, and
given her to his
companion. And the
Philistines came
up, and
burnt her and her
father with
fire.
\par }{\PP \VS{7}¶ And
Samson
said unto them,
Though ye have
done
this, yet will I be
avenged of you, and
after that I will
cease.
\VS{8}And he
smote them
hip
and
thigh with a
great
slaughter: and he went
down and
dwelt in the
top of the
rock
Etam.
\par }{\PP \VS{9}¶ Then the
Philistines went
up, and
pitched in
Judah, and
spread themselves in
Lehi.
\VS{10}And the
men of
Judah
said, Why are ye come
up against us? And they
answered, To
bind
Samson are we come
up, to
do to him as he hath
done to us.
\VS{11}Then
three
thousand
men of
Judah
went to the
top of the
rock
Etam, and
said to
Samson,
Knowest thou not that the
Philistines
{\ADD{are}}
rulers over us? what
{\ADD{is}} this
{\ADD{that}} thou hast
done unto us? And he
said unto them, As they
did unto me, so have I
done unto them.\FTNT{§}{{\FR 15.11 }
went: Heb. went down}
\VS{12}And they
said unto him, We are come
down to
bind thee, that we may
deliver thee into the
hand of the
Philistines. And
Samson
said unto them,
Swear unto me, that ye will not fall
upon me yourselves.
\VS{13}And they
spake unto him,
saying, No; but we will
bind thee
fast, and
deliver thee into their
hand: but
surely we will not
kill thee. And they
bound him with
two
new
cords, and brought him
up from the
rock.
\par }{\PP \VS{14}¶
{\ADD{And}} when he
came unto
Lehi, the
Philistines
shouted
against him: and the
Spirit of the
{\ND{
LORD}} came
mightily upon him, and the
cords that
{\ADD{were}} upon his
arms became as
flax that was
burnt with
fire, and his
bands
loosed from off his
hands.\FTNT{**}{{\FR 15.14 }
loosed: Heb. were melted}
\VS{15}And he
found a
new
jawbone of an
ass, and put
forth his
hand, and
took it, and
slew a
thousand
men therewith.\FTNT{††}{{\FR 15.15 }
new: Heb. moist}
\VS{16}And
Samson
said, With the
jawbone of an
ass,
heaps upon
heaps, with the
jaw of an
ass have I
slain a
thousand
men.\FTNT{‡‡}{{\FR 15.16 }
heaps upon…: Heb. an heap, two heaps}
\VS{17}And it came to pass, when he had made an
end of
speaking, that he cast
away the
jawbone out of his
hand, and
called that
place
Ramath-lehi.\FTNT{§§}{{\FR 15.17 }
Ramath-lehi: that is, the lifting up of the jawbone, or, casting away of the jawbone}
\par }{\PP \VS{18}¶ And he was
sore
athirst, and
called on the
{\ND{
LORD}}, and
said, Thou hast
given this
great
deliverance into the
hand of thy
servant: and now shall I
die for
thirst, and
fall into the
hand of the
uncircumcised?
\VS{19}But
God
clave an hollow
place that
{\ADD{was}} in the
jaw, and there
came
water thereout; and when he had
drunk, his
spirit came
again, and he
revived: wherefore he
called the
name thereof
En-hakkore, which
{\ADD{is}} in
Lehi unto this
day.\FTNT{***}{{\FR 15.19 }
En-hakkore: that is, the well of him that called or, cried}\FTNT{†††}{{\FR 15.19 }
the jaw: or, Lehi, as called in this chapter}
\VS{20}And he
judged
Israel in the
days of the
Philistines
twenty
years.

\par }\Chap{16}{\PP \VerseOne{1}Then
went
Samson to
Gaza, and
saw there an
harlot, and went
in unto her.\FTNT{*}{{\FR 16.1 }
harlot: Heb. a woman an harlot}
\VS{2}{\ADD{And it was told}} the
Gazites,
saying,
Samson is
come hither. And they compassed
{\ADD{him}}
in, and laid
wait for him all
night in the
gate of the
city, and were
quiet all the
night,
saying, In the
morning, when it is
day, we shall
kill him.\FTNT{†}{{\FR 16.2 }
quiet: Heb. silence}
\VS{3}And
Samson
lay till
midnight, and
arose at
midnight, and
took the
doors of the
gate of the
city, and the
two
posts, and went
away with them,
bar and all, and
put
{\ADD{them}} upon his
shoulders, and carried them
up to the
top of an
hill that
{\ADD{is}}
before
Hebron.\FTNT{‡}{{\FR 16.3 }
bar…: Heb. with the bar}
\par }{\PP \VS{4}¶ And it came to pass
afterward, that he
loved a
woman in the
valley of
Sorek, whose
name
{\ADD{was}}
Delilah.\FTNT{§}{{\FR 16.4 }
in…: or, by the brook}
\VS{5}And the
lords of the
Philistines came
up unto her, and
said unto her,
Entice him, and
see wherein his
great
strength
{\ADD{lieth}}, and by what
{\ADD{means}} we may
prevail against him, that we may
bind him to
afflict him: and we will
give thee every
one of us eleven
hundred
{\ADD{pieces}} of
silver.\FTNT{**}{{\FR 16.5 }
afflict: or, humble}
\par }{\PP \VS{6}¶ And
Delilah
said to
Samson,
Tell me, I pray thee, wherein thy
great
strength
{\ADD{lieth}}, and wherewith thou mightest be
bound to
afflict thee.
\VS{7}And
Samson
said unto her, If they
bind me with
seven
green
withs that were never
dried, then shall I be
weak, and be as
another
man.\FTNT{††}{{\FR 16.7 }
green…: or, new cords: Heb. moist}\FTNT{‡‡}{{\FR 16.7 }
another: Heb. one}
\VS{8}Then the
lords of the
Philistines brought
up to her
seven
green
withs which had not been
dried, and she
bound him with them.\FTNT{§§}{{\FR 16.8 }
green…: or, new cords: Heb. moist}
\VS{9}Now
{\ADD{there were}} men lying in
wait,
abiding with her in the
chamber. And she
said unto him, The
Philistines
{\ADD{be}} upon thee,
Samson. And he
brake the
withs, as a
thread of
tow is
broken when it
toucheth the
fire. So his
strength was not
known.\FTNT{***}{{\FR 16.9 }
toucheth: Heb. smelleth}
\VS{10}And
Delilah
said unto
Samson, Behold, thou hast
mocked me, and
told me
lies: now
tell me, I pray thee, wherewith thou mightest be
bound.
\VS{11}And he
said unto her, If they
bind me
fast with
new
ropes that never were
occupied, then shall I be
weak, and be as
another
man.\FTNT{†††}{{\FR 16.11 }
that never…: Heb. wherewith work hath not been done}
\VS{12}Delilah therefore
took
new
ropes, and
bound him therewith, and
said unto him, The
Philistines
{\ADD{be}} upon thee,
Samson. And
{\ADD{there were}} liers in
wait
abiding in the
chamber. And he
brake them from off his
arms like a
thread.
\VS{13}And
Delilah
said unto
Samson,
Hitherto thou hast
mocked me, and
told me
lies:
tell me wherewith thou mightest be
bound. And he
said unto her, If thou
weavest the
seven
locks of my
head with the
web.
\VS{14}And she
fastened
{\ADD{it}} with the
pin, and
said unto him, The
Philistines
{\ADD{be}} upon thee,
Samson. And he
awaked out of his
sleep, and went
away with the
pin of the
beam, and with the
web.
\par }{\PP \VS{15}¶ And she
said unto him,
How canst thou
say, I
love thee, when thine
heart
{\ADD{is}} not with me? thou hast
mocked me these
three
times, and hast not
told me wherein thy
great
strength
{\ADD{lieth}}.
\VS{16}And it came to pass, when she
pressed him
daily with her
words, and
urged him,
{\ADD{so}} that his
soul was
vexed unto
death;\FTNT{‡‡‡}{{\FR 16.16 }
vexed: Heb. shortened}
\VS{17}That he
told her all his
heart, and
said unto her, There hath not
come a
razor upon mine
head; for I
{\ADD{have been}} a
Nazarite unto
God from my
mother’s
womb: if I be
shaven, then my
strength will
go from me, and I shall become
weak, and be like any
{\ADD{other}}
man.
\par }{\PP \VS{18}And when
Delilah
saw that he had
told her all his
heart, she
sent and
called for the
lords of the
Philistines,
saying, Come
up this
once, for he hath
shewed me all his
heart. Then the
lords of the
Philistines came
up unto her, and
brought
money in their
hand.
\VS{19}And she made him
sleep upon her
knees; and she
called for a
man, and she caused him to shave
off the
seven
locks of his
head; and she
began to
afflict him, and his
strength
went from him.
\VS{20}And she
said, The
Philistines
{\ADD{be}} upon thee,
Samson. And he
awoke out of his
sleep, and
said, I will go
out as at other times
before, and
shake myself. And he
wist not that the
{\ND{
LORD}} was
departed from him.
\par }{\PP \VS{21}¶ But the
Philistines
took him, and put
out his
eyes, and brought him
down to
Gaza, and
bound him with fetters of
brass; and he did
grind in the
prison
house.\FTNT{§§§}{{\FR 16.21 }
put out: Heb. bored out}
\par }{\PP \VS{22}Howbeit the
hair of his
head
began to grow
again
after he was
shaven.\FTNT{*}{{\FR 16.22 }
after…: or, as when he was shaven}
\VS{23}Then the
lords of the
Philistines gathered them
together for to
offer a
great
sacrifice unto
Dagon their
god, and to
rejoice: for they
said, Our
god hath
delivered
Samson our
enemy into our
hand.
\VS{24}And when the
people
saw him, they
praised their
god: for they
said, Our
god hath
delivered into our
hands our
enemy, and the
destroyer of our
country, which
slew
many of us.\FTNT{†}{{\FR 16.24 }
which…: Heb. and who multiplied our slain}
\VS{25}And it came to pass, when their
hearts were
merry, that they
said,
Call for
Samson, that he may make us
sport. And they
called for
Samson out of the
prison
house; and he made
them
sport: and they
set him between the
pillars.\FTNT{‡}{{\FR 16.25 }
them: Heb. before them}
\VS{26}And
Samson
said unto the
lad that
held him by the
hand,
Suffer me that I may
feel the
pillars whereupon the
house
standeth, that I may
lean upon them.
\VS{27}Now the
house was
full of
men and
women; and all the
lords of the
Philistines
{\ADD{were}} there; and
{\ADD{there were}} upon the
roof about
three
thousand
men and
women, that
beheld while
Samson made
sport.
\VS{28}And
Samson
called unto the
{\ND{
LORD}}, and
said, O
Lord
{\ND{
GOD}},
remember me, I pray thee, and
strengthen me, I pray thee, only this
once, O
God, that I may be at
once
avenged of the
Philistines for my
two
eyes.
\VS{29}And
Samson took
hold of the
two
middle
pillars upon which the
house
stood, and on which it was borne
up, of the
one with his right
hand, and of the
other with his
left.\FTNT{§}{{\FR 16.29 }
on which…: or, he leaned on them}
\VS{30}And
Samson
said, Let me
die with the
Philistines. And he
bowed himself with
{\ADD{all his}}
might; and the
house
fell upon the
lords, and upon all the
people that
{\ADD{were}} therein. So the
dead which he
slew at his
death were
more than
{\ADD{they}} which he
slew in his
life.\FTNT{**}{{\FR 16.30 }
me: Heb. my soul}
\VS{31}Then his
brethren and all the
house of his
father came
down, and
took him, and brought
{\ADD{him}}
up, and
buried him between
Zorah and
Eshtaol in the
buryingplace of
Manoah his
father. And he
judged
Israel
twenty
years.

\par }\Chap{17}{\PP \VerseOne{1}And there was a
man of
mount
Ephraim, whose
name
{\ADD{was}}
Micah.
\VS{2}And he
said unto his
mother, The eleven
hundred
{\ADD{shekels}} of
silver that were
taken from thee, about which thou
cursedst, and
spakest of also in mine
ears, behold, the
silver
{\ADD{is}} with me; I
took it. And his
mother
said,
Blessed
{\ADD{be thou}} of the
{\ND{
LORD}}, my
son.
\VS{3}And when he had
restored the eleven
hundred
{\ADD{shekels}} of
silver to his
mother, his
mother
said, I had
wholly
dedicated the
silver unto the
{\ND{
LORD}} from my
hand for my
son, to
make a graven
image and a molten
image: now therefore I will
restore it unto thee.
\VS{4}Yet he
restored the
money unto his
mother; and his
mother
took two
hundred
{\ADD{shekels}} of
silver, and
gave them to the
founder, who
made thereof a graven
image and a molten
image: and they were in the
house of
Micah.
\VS{5}And the
man
Micah had an
house of
gods, and
made an
ephod, and
teraphim, and
consecrated
one of his
sons, who became his
priest.\FTNT{*}{{\FR 17.5 }
consecrated: Heb. filled the hand}
\VS{6}In those
days
{\ADD{there was}} no
king in
Israel,
{\ADD{but}} every
man
did
{\ADD{that which was}}
right in his own
eyes.
\par }{\PP \VS{7}¶ And there was a young
man out of
Beth-lehem-judah of the
family of
Judah, who
{\ADD{was}} a
Levite, and he
sojourned there.
\VS{8}And the
man
departed out of the
city from
Beth-lehem-judah to
sojourn where he could
find
{\ADD{a place:}} and he
came to
mount
Ephraim to the
house of
Micah, as he
journeyed.\FTNT{†}{{\FR 17.8 }
as he…: Heb. in making his way}
\VS{9}And
Micah
said unto him,
Whence
comest thou? And he
said unto him, I
{\ADD{am}} a
Levite of
Beth-lehem-judah, and I
go to
sojourn where I may
find
{\ADD{a place}}.
\VS{10}And
Micah
said unto him,
Dwell with me, and be unto me a
father and a
priest, and I will
give thee
ten
{\ADD{shekels}} of
silver by the
year, and a
suit of
apparel, and thy
victuals. So the
Levite went
in.\FTNT{‡}{{\FR 17.10 }
a suit…: or, a double suit, etc: Heb. an order of garments}
\VS{11}And the
Levite was
content to
dwell with the
man; and the young
man was unto him as
one of his
sons.
\VS{12}And
Micah
consecrated the
Levite; and the young
man became his
priest, and was in the
house of
Micah.
\VS{13}Then
said
Micah, Now
know I that the
{\ND{
LORD}} will do me
good, seeing I have a
Levite to
{\ADD{my}}
priest.

\par }\Chap{18}{\PP \VerseOne{1}In those
days
{\ADD{there was}} no
king in
Israel: and in those
days the
tribe of the
Danites
sought them an
inheritance to dwell
in; for unto that
day
{\ADD{all their}}
inheritance had not
fallen unto them
among the
tribes of
Israel.
\VS{2}And the
children of
Dan
sent of their
family
five
men from their
coasts,
men of
valour, from
Zorah, and from
Eshtaol, to
spy out the
land, and to
search it; and they
said unto them,
Go,
search the
land: who when they
came to
mount
Ephraim, to the
house of
Micah, they
lodged there.\FTNT{*}{{\FR 18.2 }
men of: Heb. sons of}
\VS{3}When they
{\ADD{were}} by the
house of
Micah, they
knew the
voice of the young
man the
Levite: and they turned
in thither, and
said unto him, Who
brought thee
hither? and what
makest thou in this
{\ADD{place}}? and what hast thou here?
\VS{4}And he
said unto them,
Thus and
thus
dealeth
Micah with me, and hath
hired me, and I am his
priest.
\VS{5}And they
said unto him, Ask
counsel, we pray thee, of
God, that we may
know whether our
way which we
go shall be
prosperous.
\VS{6}And the
priest
said unto them,
Go in
peace:
before the
{\ND{
LORD}}
{\ADD{is}} your
way wherein ye
go.
\par }{\PP \VS{7}¶ Then the
five
men
departed, and
came to
Laish, and
saw the
people that
{\ADD{were}}
therein, how they
dwelt
careless, after the
manner of the
Zidonians,
quiet and
secure; and
{\ADD{there was}} no
magistrate in the
land, that might put
{\ADD{them}} to
shame in
{\ADD{any}}
thing; and they
{\ADD{were}}
far from the
Zidonians, and had no
business with
{\ADD{any}}
man.\FTNT{†}{{\FR 18.7 }
Laish: also called, Leshem}\FTNT{‡}{{\FR 18.7 }
magistrate: Heb. possessor, or, heir of restraints}
\VS{8}And they
came unto their
brethren to
Zorah and
Eshtaol: and their
brethren
said unto them, What
{\ADD{say}} ye?
\VS{9}And they
said,
Arise, that we may go
up against them: for we have
seen the
land, and, behold, it
{\ADD{is}}
very
good: and
{\ADD{are}} ye
still? be not
slothful to
go,
{\ADD{and}} to
enter to
possess the
land.
\VS{10}When ye
go, ye shall
come unto a
people
secure, and to a
large
land: for
God hath
given it into your
hands; a
place where
{\ADD{there is}} no
want of any
thing that
{\ADD{is}} in the
earth.
\par }{\PP \VS{11}¶ And there
went from thence of the
family of the
Danites, out of
Zorah and out of
Eshtaol,
six
hundred
men
appointed with
weapons of
war.\FTNT{§}{{\FR 18.11 }
appointed: Heb. girded}
\VS{12}And they went
up, and
pitched in
Kirjath-jearim, in
Judah: wherefore they
called that
place
Mahaneh-dan unto this
day: behold,
{\ADD{it is}}
behind
Kirjath-jearim.
\VS{13}And they
passed thence unto
mount
Ephraim, and
came unto the
house of
Micah.
\par }{\PP \VS{14}¶ Then
answered the
five
men that
went to spy
out the
country of
Laish, and
said unto their
brethren, Do ye
know that there
is in these
houses an
ephod, and
teraphim, and a graven
image, and a molten
image? now therefore
consider what ye have to
do.
\VS{15}And they
turned thitherward, and
came to the
house of the young
man the
Levite,
{\ADD{even}} unto the
house of
Micah, and
saluted him.\FTNT{**}{{\FR 18.15 }
saluted…: Heb. asked him of peace}
\VS{16}And the
six
hundred
men
appointed with their
weapons of
war, which
{\ADD{were}} of the
children of
Dan,
stood by the
entering of the
gate.
\VS{17}And the
five
men that
went to spy
out the
land went
up,
{\ADD{and}} came
in thither,
{\ADD{and}}
took the graven
image, and the
ephod, and the
teraphim, and the molten
image: and the
priest
stood in the
entering of the
gate with the
six
hundred
men
{\ADD{that were}}
appointed with
weapons of
war.
\VS{18}And these
went into
Micah’s
house, and
fetched the carved
image, the
ephod, and the
teraphim, and the molten
image. Then
said the
priest unto them, What
do ye?
\VS{19}And they
said unto him, Hold thy
peace,
lay thine
hand upon thy
mouth, and
go with us, and be to
us a
father and a
priest:
{\ADD{is it}}
better for thee to be a
priest unto the
house of
one
man, or that thou be a
priest unto a
tribe and a
family in
Israel?
\VS{20}And the
priest’s
heart was
glad, and he
took the
ephod, and the
teraphim, and the graven
image, and went
in the
midst of the
people.
\VS{21}So they
turned and
departed, and
put the little
ones and the
cattle and the
carriage
before them.
\par }{\PP \VS{22}¶
{\ADD{And}} when they were a good
way from the
house of
Micah, the
men that
{\ADD{were}} in the
houses near to
Micah’s
house were gathered
together, and
overtook the
children of
Dan.
\VS{23}And they
cried unto the
children of
Dan. And they
turned their
faces, and
said unto
Micah, What aileth thee, that thou comest with such a
company?\FTNT{††}{{\FR 18.23 }
comest…: Heb. art gathered together?}
\VS{24}And he
said, Ye have taken
away my
gods which I
made, and the
priest, and ye are gone
away: and what have I more? and what
{\ADD{is}} this
{\ADD{that}} ye
say unto me, What aileth thee?
\VS{25}And the
children of
Dan
said unto him, Let not thy
voice be
heard among us, lest
angry
fellows
run upon thee, and thou
lose thy
life, with the
lives of thy
household.\FTNT{‡‡}{{\FR 18.25 }
angry: Heb. bitter of soul}
\VS{26}And the
children of
Dan
went their
way: and when
Micah
saw that they
{\ADD{were}} too
strong for him, he
turned and went
back unto his
house.
\par }{\PP \VS{27}And they
took
{\ADD{the things}} which
Micah had
made, and the
priest which he had, and
came unto
Laish, unto a
people
{\ADD{that were}} at
quiet and
secure: and they
smote them with the
edge of the
sword, and
burnt the
city with
fire.
\VS{28}And
{\ADD{there was}} no
deliverer, because it
{\ADD{was}}
far from
Zidon, and they had no
business with
{\ADD{any}}
man; and it was in the
valley that
{\ADD{lieth}} by
Beth-rehob. And they
built a
city, and
dwelt therein.
\VS{29}And they
called the
name of the
city
Dan, after the
name of
Dan their
father, who was
born unto
Israel:
howbeit the
name of the
city
{\ADD{was}}
Laish at the
first.
\par }{\PP \VS{30}¶ And the
children of
Dan set
up the graven
image: and
Jonathan, the
son of
Gershom, the
son of
Manasseh, he and his
sons were
priests to the
tribe of
Dan until the
day of the
captivity of the
land.
\VS{31}And they set them
up
Micah’s graven
image, which he
made, all the
time that the
house of
God was in
Shiloh.

\par }\Chap{19}{\PP \VerseOne{1}And it came to pass in those
days, when
{\ADD{there was}} no
king in
Israel, that there was a
certain
Levite
sojourning on the
side of
mount
Ephraim, who
took to him a
concubine out of
Beth-lehem-judah.\FTNT{*}{{\FR 19.1 }
a concubine: Heb. a woman a concubine, or, a wife a concubine}
\VS{2}And his
concubine played the
whore against him, and went
away from him unto her
father’s
house to
Beth-lehem-judah, and was there
four whole
months.\FTNT{†}{{\FR 19.2 }
four whole months: or, a year and four month: Heb. days, four months}
\VS{3}And her
husband
arose, and
went
after her, to
speak
friendly unto her,
{\ADD{and}} to bring her
again, having his
servant with him, and a
couple of
asses: and she
brought him into her
father’s
house: and when the
father of the
damsel
saw him, he
rejoiced to
meet him.\FTNT{‡}{{\FR 19.3 }
friendly…: Heb. to her heart}
\VS{4}And his father in
law, the
damsel’s
father,
retained him; and he
abode with him
three
days: so they did
eat and
drink, and
lodged there.
\par }{\PP \VS{5}¶ And it came to pass on the
fourth
day, when they arose
early in the
morning, that he rose
up to
depart: and the
damsel’s
father
said unto his son in
law,
Comfort thine
heart with a
morsel of
bread, and
afterward go your
way.\FTNT{§}{{\FR 19.5 }
Comfort: Heb. Strengthen}
\VS{6}And they sat
down, and did
eat and
drink
both of them
together: for the
damsel’s
father had
said unto the
man, Be
content, I pray thee, and tarry all
night, and let thine
heart be
merry.
\VS{7}And when the
man rose
up to
depart, his father in
law
urged him: therefore he
lodged there
again.
\VS{8}And he arose
early in the
morning on the
fifth
day to
depart: and the
damsel’s
father
said,
Comfort thine
heart, I pray thee. And they
tarried until
afternoon, and they did
eat
both of them.\FTNT{**}{{\FR 19.8 }
until afternoon: Heb. till the day declined}
\VS{9}And when the
man rose
up to
depart, he, and his
concubine, and his
servant, his father in
law, the
damsel’s
father,
said unto him, Behold, now the
day
draweth toward
evening, I pray you tarry all
night: behold, the
day groweth to an
end,
lodge here, that thine
heart may be
merry; and to
morrow get you
early on your
way, that thou mayest
go
home.\FTNT{††}{{\FR 19.9 }
draweth: Heb. is weak}\FTNT{‡‡}{{\FR 19.9 }
the day groweth to an end: Heb. it is the pitching time of the day}\FTNT{§§}{{\FR 19.9 }
home: Heb. to thy tent}
\VS{10}But the
man
would not tarry that
night, but he rose
up and
departed, and
came over
against
Jebus, which
{\ADD{is}}
Jerusalem; and
{\ADD{there were}} with him
two
asses
saddled, his
concubine also
{\ADD{was}} with him.\FTNT{***}{{\FR 19.10 }
over against: Heb. to over against}
\VS{11}{\ADD{And}} when they
{\ADD{were}} by
Jebus, the
day was
far
spent; and the
servant
said unto his
master,
Come, I pray thee, and let us turn
in into this
city of the
Jebusites, and
lodge in it.
\VS{12}And his
master
said unto him, We will not turn
aside hither into the
city of a
stranger, that
{\ADD{is}} not of the
children of
Israel; we will pass
over to
Gibeah.
\VS{13}And he
said unto his
servant,
Come, and let us draw
near to
one of these
places to lodge all
night, in
Gibeah, or in
Ramah.
\VS{14}And they passed
on and went their
way; and the
sun went
down upon them
{\ADD{when they were}}
by
Gibeah, which
{\ADD{belongeth}} to
Benjamin.
\VS{15}And they turned
aside thither, to go
in
{\ADD{and}} to
lodge in
Gibeah: and when he went
in, he sat him
down in a
street of the
city: for
{\ADD{there was}} no
man that
took them into his
house to
lodging.
\par }{\PP \VS{16}¶ And, behold, there
came an
old
man from his
work out of the
field at
even, which
{\ADD{was}}
also of
mount
Ephraim; and he
sojourned in
Gibeah: but the
men of the
place
{\ADD{were}}
Benjamites.
\VS{17}And when he had lifted
up his
eyes, he
saw a
wayfaring
man in the
street of the
city: and the
old
man
said, Whither
goest thou? and
whence
comest thou?
\VS{18}And he
said unto him, We
{\ADD{are}}
passing from
Beth-lehem-judah toward the
side of
mount
Ephraim; from thence
{\ADD{am}} I: and I
went to
Beth-lehem-judah, but I
{\ADD{am now}}
going to the
house of the
{\ND{
LORD}}; and there
{\ADD{is}} no
man that
receiveth me to
house.\FTNT{†††}{{\FR 19.18 }
receiveth: Heb. gathereth}
\VS{19}Yet there
is both
straw and
provender for our
asses; and there is
bread and
wine also for me, and for thy
handmaid, and for the young
man
{\ADD{which is}} with thy
servants:
{\ADD{there is}} no
want of any
thing.
\VS{20}And the
old
man
said,
Peace
{\ADD{be}} with thee;
howsoever
{\ADD{let}} all thy
wants
{\ADD{lie}} upon me; only
lodge not in the
street.
\VS{21}So he
brought him into his
house, and gave
provender unto the
asses: and they
washed their
feet, and did
eat and
drink.
\par }{\PP \VS{22}¶
{\ADD{Now}} as they were making their
hearts
merry, behold, the
men of the
city,
certain
sons of
Belial,
beset the
house round
about,
{\ADD{and}}
beat at the
door, and
spake to the
master of the
house, the
old
man,
saying, Bring
forth the
man that
came into thine
house, that we may
know him.
\VS{23}And the
man, the
master of the
house, went
out unto them, and
said unto them,
Nay, my
brethren,
{\ADD{nay}}, I pray you, do not
{\ADD{so}}
wickedly;
seeing that this
man is
come into mine
house,
do not this
folly.
\VS{24}Behold,
{\ADD{here is}} my
daughter a
maiden, and his
concubine; them I will bring
out now, and
humble ye them, and
do with them what
seemeth
good unto you: but unto this
man
do not
so
vile a
thing.\FTNT{‡‡‡}{{\FR 19.24 }
so vile…: Heb. the matter of this folly}
\VS{25}But the
men
would not
hearken to him: so the
man
took his
concubine, and
brought her
forth unto them; and they
knew her, and
abused her all the
night until the
morning: and when the
day began to
spring, they let her
go.
\VS{26}Then
came the
woman in the
dawning of the
day, and fell
down at the
door of the
man’s
house where her
lord
{\ADD{was}}, till it was
light.
\VS{27}And her
lord rose
up in the
morning, and
opened the
doors of the
house, and went
out to
go his
way: and, behold, the
woman his
concubine was fallen
down
{\ADD{at}} the
door of the
house, and her
hands
{\ADD{were}} upon the
threshold.
\VS{28}And he
said unto her,
Up, and let us be
going. But none
answered. Then the
man
took her
{\ADD{up}} upon an
ass, and the
man rose
up, and
gat him unto his
place.
\par }{\PP \VS{29}¶ And when he was
come into his
house, he
took a
knife, and laid
hold on his
concubine, and
divided her,
{\ADD{together}} with her
bones, into
twelve
pieces, and
sent her into all the
coasts of
Israel.
\VS{30}And it was
so, that all that
saw it
said, There was no such
deed done nor
seen from the
day that the
children of
Israel came
up out of the
land of
Egypt unto this
day:
consider of it, take
advice, and
speak
{\ADD{your minds}}.

\par }\Chap{20}{\PP \VerseOne{1}Then all the
children of
Israel went
out, and the
congregation was gathered
together as
one
man, from
Dan even to
Beer-sheba, with the
land of
Gilead, unto the
{\ND{
LORD}} in
Mizpeh.
\VS{2}And the
chief of all the
people,
{\ADD{even}} of all the
tribes of
Israel,
presented themselves in the
assembly of the
people of
God,
four
hundred
thousand
footmen that
drew
sword.
\VS{3}(Now the
children of
Benjamin
heard that the
children of
Israel were gone
up to
Mizpeh.) Then
said the
children of
Israel,
Tell
{\ADD{us}}, how
was this
wickedness?
\VS{4}And the
Levite, the
husband of the
woman that was
slain,
answered and
said, I
came into
Gibeah that
{\ADD{belongeth}} to
Benjamin, I and my
concubine, to
lodge.\FTNT{*}{{\FR 20.4 }
the Levite: Heb. the man the Levite}
\VS{5}And the
men of
Gibeah
rose against me, and
beset the
house round
about upon me by
night,
{\ADD{and}}
thought to have
slain me: and my
concubine have they
forced, that she is
dead.\FTNT{†}{{\FR 20.5 }
forced: Heb. humbled}
\VS{6}And I
took my
concubine, and cut her in
pieces, and
sent her throughout all the
country of the
inheritance of
Israel: for they have
committed
lewdness and
folly in
Israel.
\VS{7}Behold, ye
{\ADD{are}} all
children of
Israel;
give
here your
advice and
counsel.
\par }{\PP \VS{8}¶ And all the
people
arose as
one
man,
saying, We will not
any
{\ADD{of us}}
go to his
tent, neither will we
any
{\ADD{of us}}
turn into his
house.
\VS{9}But now this
{\ADD{shall be}} the
thing which we will
do to
Gibeah;
{\ADD{we will go up}} by
lot against it;
\VS{10}And we will
take
ten
men of an
hundred throughout all the
tribes of
Israel, and an
hundred of a
thousand, and a
thousand out of ten
thousand, to
fetch
victual for the
people, that they may
do, when they
come to
Gibeah of
Benjamin, according to all the
folly that they have
wrought in
Israel.
\VS{11}So all the
men of
Israel were
gathered against the
city, knit
together as
one
man.\FTNT{‡}{{\FR 20.11 }
knit…: Heb. fellows}
\par }{\PP \VS{12}¶ And the
tribes of
Israel
sent
men through all the
tribe of
Benjamin,
saying, What
wickedness
{\ADD{is}} this that is
done among you?
\VS{13}Now therefore
deliver
{\ADD{us}} the
men, the
children of
Belial, which
{\ADD{are}} in
Gibeah, that we may put them to
death, and put
away
evil from
Israel. But the
children of
Benjamin
would not
hearken to the
voice of their
brethren the
children of
Israel:
\VS{14}But the
children of
Benjamin gathered themselves
together out of the
cities unto
Gibeah, to go
out to
battle against the
children of
Israel.
\VS{15}And the
children of
Benjamin were
numbered at that
time out of the
cities
twenty and
six
thousand
men that
drew
sword, beside the
inhabitants of
Gibeah, which were
numbered
seven
hundred
chosen
men.
\VS{16}Among all this
people
{\ADD{there were}}
seven
hundred
chosen
men
lefthanded; every one could
sling
stones at an
hair
{\ADD{breadth}}, and not
miss.
\VS{17}And the
men of
Israel, beside
Benjamin, were
numbered
four
hundred
thousand
men that
drew
sword: all these
{\ADD{were}}
men of
war.
\par }{\PP \VS{18}¶ And the
children of
Israel
arose, and went
up to the
house of
God, and
asked counsel of
God, and
said,
Which of us shall go
up
first to the
battle against the
children of
Benjamin? And the
{\ND{
LORD}}
said,
Judah
{\ADD{shall go up}}
first.
\VS{19}And the
children of
Israel rose
up in the
morning, and
encamped against
Gibeah.
\VS{20}And the
men of
Israel went
out to
battle against
Benjamin; and the
men of
Israel put themselves in
array to
fight against them at
Gibeah.
\VS{21}And the
children of
Benjamin came
forth out of
Gibeah, and destroyed
down to the
ground of the
Israelites that
day
twenty and
two
thousand
men.
\VS{22}And the
people the
men of
Israel
encouraged themselves, and set their
battle
again in
array in the
place where they put themselves in
array the
first
day.
\VS{23}(And the
children of
Israel went
up and
wept
before the
{\ND{
LORD}} until
even, and
asked counsel of the
{\ND{
LORD}},
saying, Shall I go
up
again to
battle against the
children of
Benjamin my
brother? And the
{\ND{
LORD}}
said, Go
up against him.)
\VS{24}And the
children of
Israel came
near against the
children of
Benjamin the
second
day.
\VS{25}And
Benjamin went
forth
against them out of
Gibeah the
second
day, and destroyed
down to the
ground of the
children of
Israel again
eighteen
thousand
men; all these
drew the
sword.
\par }{\PP \VS{26}¶ Then all the
children of
Israel, and all the
people, went
up, and
came unto the
house of
God, and
wept, and
sat there
before the
{\ND{
LORD}}, and
fasted that
day until
even, and
offered burnt
offerings and peace
offerings
before the
{\ND{
LORD}}.
\VS{27}And the
children of
Israel
enquired of the
{\ND{
LORD}}, (for the
ark of the
covenant of
God
{\ADD{was}} there in those
days,
\VS{28}And
Phinehas, the
son of
Eleazar, the
son of
Aaron,
stood
before it in those
days,)
saying, Shall I yet
again go
out to
battle against the
children of
Benjamin my
brother, or shall I
cease? And the
{\ND{
LORD}}
said, Go
up; for to
morrow I will
deliver them into thine
hand.
\VS{29}And
Israel
set liers in
wait round
about
Gibeah.
\VS{30}And the
children of
Israel went
up against the
children of
Benjamin on the
third
day, and put themselves in
array against
Gibeah, as at other
times.
\VS{31}And the
children of
Benjamin went
out
against the
people,
{\ADD{and}} were drawn
away from the
city; and they
began to
smite of the
people,
{\ADD{and}}
kill, as at other
times, in the
highways, of which
one goeth
up to the
house of
God, and the
other to
Gibeah in the
field, about
thirty
men of
Israel.\FTNT{§}{{\FR 20.31 }
to smite…: Heb. to smite of the people wounded as at, etc}\FTNT{**}{{\FR 20.31 }
the house…: or, Beth-el}
\VS{32}And the
children of
Benjamin
said, They
{\ADD{are}} smitten
down
before us, as at the
first. But the
children of
Israel
said, Let us
flee, and
draw them from the
city unto the
highways.
\VS{33}And all the
men of
Israel rose
up out of their
place, and put themselves in
array at
Baal-tamar: and the liers in
wait of
Israel came
forth out of their
places,
{\ADD{even}} out of the
meadows of
Gibeah.
\VS{34}And there
came
against
Gibeah
ten
thousand
chosen
men out of all
Israel, and the
battle was
sore: but they
knew not that
evil
{\ADD{was}}
near them.
\VS{35}And the
{\ND{
LORD}}
smote
Benjamin
before
Israel: and the
children of
Israel
destroyed of the
Benjamites that
day
twenty and
five
thousand and an
hundred
men: all these
drew the
sword.
\VS{36}So the
children of
Benjamin
saw that they were
smitten: for the
men of
Israel
gave
place to the
Benjamites, because they
trusted unto the liers in
wait which they had
set beside
Gibeah.
\VS{37}And the liers in
wait
hasted, and
rushed upon
Gibeah; and the liers in
wait drew
{\ADD{themselves}}
along, and
smote all the
city with the
edge of the
sword.\FTNT{††}{{\FR 20.37 }
drew…: or, made a long sound with the trumpet}
\VS{38}Now there was an appointed
sign between the
men of
Israel
and the liers in
wait, that they should make a
great
flame with
smoke rise
up out of the
city.\FTNT{‡‡}{{\FR 20.38 }
sign: or, time}\FTNT{§§}{{\FR 20.38 }
and: Heb. with}\FTNT{***}{{\FR 20.38 }
flame: Heb. elevation}
\VS{39}And when the
men of
Israel
retired in the
battle,
Benjamin
began to
smite
{\ADD{and}}
kill of the
men of
Israel about
thirty
persons: for they
said,
Surely they are smitten
down
before us, as
{\ADD{in}} the
first
battle.\FTNT{†††}{{\FR 20.39 }
to smite…: Heb. to smite the wounded}
\VS{40}But when the
flame
began to arise
up out of the
city with a
pillar of
smoke, the
Benjamites
looked
behind them, and, behold, the
flame of the
city ascended
up to
heaven.\FTNT{‡‡‡}{{\FR 20.40 }
the flame of: Heb. the whole consumption of}
\VS{41}And when the
men of
Israel turned
again, the
men of
Benjamin were
amazed: for they
saw that
evil was
come upon them.\FTNT{§§§}{{\FR 20.41 }
was come…: Heb. touched them}
\VS{42}Therefore they
turned
{\ADD{their backs}}
before the
men of
Israel unto the
way of the
wilderness; but the
battle
overtook them; and them which
{\ADD{came}} out of the
cities they
destroyed in the
midst of them.
\VS{43}{\ADD{Thus}} they
inclosed the
Benjamites round
about,
{\ADD{and}}
chased them,
{\ADD{and}} trode them
down with
ease over
against
Gibeah toward the
sunrising.\FTNT{*}{{\FR 20.43 }
with ease…: or, from Menuchah, etc}\FTNT{†}{{\FR 20.43 }
over…: Heb. unto over against}
\VS{44}And there
fell of
Benjamin
eighteen
thousand
men; all these
{\ADD{were}}
men of
valour.
\VS{45}And they
turned and
fled toward the
wilderness unto the
rock of
Rimmon: and they
gleaned of them in the
highways
five
thousand
men; and
pursued hard
after them unto
Gidom, and
slew two
thousand
men of them.
\VS{46}So that all which
fell that
day of
Benjamin were
twenty and
five
thousand
men that
drew the
sword; all these
{\ADD{were}}
men of
valour.
\VS{47}But
six
hundred
men
turned and
fled to the
wilderness unto the
rock
Rimmon, and
abode in the
rock
Rimmon
four
months.
\VS{48}And the
men of
Israel turned
again upon the
children of
Benjamin, and
smote them with the
edge of the
sword, as well the
men of
{\ADD{every}}
city,
as the
beast, and all that came to
hand: also they
set on
fire all the
cities that they came
to.\FTNT{‡}{{\FR 20.48 }
came to hand: Heb. was found}\FTNT{§}{{\FR 20.48 }
they came…: Heb. were found}

\par }\Chap{21}{\PP \VerseOne{1}Now the
men of
Israel had
sworn in
Mizpeh,
saying, There shall not
any of us
give his
daughter unto
Benjamin to
wife.
\VS{2}And the
people
came to the
house of
God, and
abode there till
even
before
God, and lifted
up their
voices, and
wept
sore;
\VS{3}And
said, O
{\ND{
LORD}}
God of
Israel, why is this come to pass in
Israel, that there should be to
day
one
tribe
lacking in
Israel?
\VS{4}And it came to pass on the
morrow, that the
people rose
early, and
built there an
altar, and
offered burnt
offerings and peace
offerings.
\VS{5}And the
children of
Israel
said, Who
{\ADD{is there}} among all the
tribes of
Israel
that
came not up with the
congregation unto the
{\ND{
LORD}}? For they had made a
great
oath concerning him that came not
up to the
{\ND{
LORD}} to
Mizpeh,
saying, He shall
surely be put to
death.
\VS{6}And the
children of
Israel
repented them for
Benjamin their
brother, and
said, There is
one
tribe cut
off from
Israel this
day.
\VS{7}How shall we
do for
wives for them that
remain, seeing we have
sworn by the
{\ND{
LORD}} that we will not
give them of our
daughters to
wives?
\par }{\PP \VS{8}¶ And they
said, What
one
{\ADD{is there}} of the
tribes of
Israel that came not
up to
Mizpeh to the
{\ND{
LORD}}? And, behold, there
came
none to the
camp from
Jabesh-gilead to the
assembly.
\VS{9}For the
people were
numbered, and, behold,
{\ADD{there were}}
none of the
inhabitants of
Jabesh-gilead there.
\VS{10}And the
congregation
sent thither
twelve
thousand
men of the
valiantest, and
commanded them,
saying,
Go and
smite the
inhabitants of
Jabesh-gilead with the
edge of the
sword, with the
women and the
children.
\VS{11}And this
{\ADD{is}} the
thing that ye shall
do, Ye shall utterly
destroy every
male, and every
woman that hath
lain by
man.\FTNT{*}{{\FR 21.11 }
hath lain…: Heb. knoweth the lying with man}
\VS{12}And they
found among the
inhabitants of
Jabesh-gilead
four
hundred
young
virgins, that had
known no
man by
lying with any
male: and they
brought them unto the
camp to
Shiloh, which
{\ADD{is}} in the
land of
Canaan.\FTNT{†}{{\FR 21.12 }
young…: Heb. young women virgins}
\VS{13}And the whole
congregation
sent
{\ADD{some}} to
speak to the
children of
Benjamin that
{\ADD{were}} in the
rock
Rimmon, and to
call
peaceably unto them.\FTNT{‡}{{\FR 21.13 }
to speak…: Heb. and spake and called}\FTNT{§}{{\FR 21.13 }
call…: or, proclaim peace}
\VS{14}And
Benjamin came
again at that
time; and they
gave them
wives which they had saved
alive of the
women of
Jabesh-gilead: and yet so they
sufficed them not.
\VS{15}And the
people
repented them for
Benjamin, because that the
{\ND{
LORD}} had
made a
breach in the
tribes of
Israel.
\par }{\PP \VS{16}¶ Then the
elders of the
congregation
said, How shall we
do for
wives for them that
remain, seeing the
women are
destroyed out of
Benjamin?
\VS{17}And they
said,
{\ADD{There must be}} an
inheritance for them that be
escaped of
Benjamin, that a
tribe be not
destroyed out of
Israel.
\VS{18}Howbeit we
may not
give them
wives of our
daughters: for the
children of
Israel have
sworn,
saying,
Cursed
{\ADD{be}} he that
giveth a
wife to
Benjamin.
\VS{19}Then they
said, Behold,
{\ADD{there is}} a
feast of the
{\ND{
LORD}} in
Shiloh
yearly
{\ADD{in a place}} which
{\ADD{is}} on the north
side of
Beth-el, on the east
side of the
highway that goeth
up from
Beth-el to
Shechem, and on the
south of
Lebonah.\FTNT{**}{{\FR 21.19 }
yearly: Heb. from year to year}\FTNT{††}{{\FR 21.19 }
on the east…: or, toward the sunrising}\FTNT{‡‡}{{\FR 21.19 }
of the highway: or, on the highway}
\VS{20}Therefore they
commanded the
children of
Benjamin,
saying,
Go and lie in
wait in the
vineyards;
\VS{21}And
see, and, behold, if the
daughters of
Shiloh come
out to
dance in
dances, then come ye
out of the
vineyards, and
catch you every
man his
wife of the
daughters of
Shiloh, and
go to the
land of
Benjamin.
\VS{22}And it shall be, when their
fathers or their
brethren
come unto us to
complain, that we will
say unto them, Be
favourable unto them for our sakes: because we
reserved not to each
man his
wife in the
war: for ye did not
give unto them at this
time,
{\ADD{that}} ye should be
guilty.\FTNT{§§}{{\FR 21.22 }
Be favourable…: or, Gratify us in them}
\VS{23}And the
children of
Benjamin did
so, and
took
{\ADD{them}}
wives, according to their
number, of them that
danced, whom they
caught: and they
went and
returned unto their
inheritance, and
repaired the
cities, and
dwelt in them.
\VS{24}And the
children of
Israel
departed thence at that
time, every
man to his
tribe and to his
family, and they went
out from thence every
man to his
inheritance.
\VS{25}In those
days
{\ADD{there was}} no
king in
Israel: every
man
did
{\ADD{that which was}}
right in his own
eyes.
\par }