\NormalFont\ShortTitle{Jeremiah}
{\MT The Book of the Prophet Jeremiah

\par }\ChapOne{1}{\PP \VerseOne{1}The
words of
Jeremiah the
son of
Hilkiah, of the
priests that
{\ADD{were}} in
Anathoth in the
land of
Benjamin:
\VS{2}To whom the
word of the
{\ND{
LORD}} came in the
days of
Josiah the
son of
Amon
king of
Judah, in the
thirteenth
year of his
reign.
\VS{3}It came also in the
days of
Jehoiakim the
son of
Josiah
king of
Judah, unto the
end of the
eleventh
year of
Zedekiah the
son of
Josiah
king of
Judah, unto the carrying
away of
Jerusalem
captive in the
fifth
month.
\par }{\PP \VS{4}Then the
word of the
{\ND{
LORD}} came unto me,
saying,
\VS{5}Before I
formed thee in the
belly I
knew thee; and before thou camest
forth out of the
womb I
sanctified thee,
{\ADD{and}} I
ordained thee a
prophet unto the
nations.\FTNT{*}{{\FR 1.5 }
ordained: Heb. gave}
\VS{6}Then
said I,
Ah,
Lord
{\ND{
GOD}}! behold, I
cannot
speak: for I
{\ADD{am}} a
child.
\par }{\PP \VS{7}¶ But the
{\ND{
LORD}}
said unto me,
Say not, I
{\ADD{am}} a
child: for thou shalt
go to all that I shall
send thee, and whatsoever I
command thee thou shalt
speak.
\VS{8}Be not
afraid of their
faces: for I
{\ADD{am}} with thee to
deliver thee,
saith the
{\ND{
LORD}}.
\VS{9}Then the
{\ND{
LORD}} put
forth his
hand, and
touched my
mouth. And the
{\ND{
LORD}}
said unto me, Behold, I have
put my
words in thy
mouth.
\VS{10}See, I have this
day
set thee over the
nations and over the
kingdoms, to root
out, and to pull
down, and to
destroy, and to throw
down, to
build, and to
plant.
\par }{\PP \VS{11}¶ Moreover the
word of the
{\ND{
LORD}} came unto me,
saying,
Jeremiah, what
seest thou? And I
said, I
see a
rod of an almond
tree.
\VS{12}Then
said the
{\ND{
LORD}} unto me, Thou hast
well
seen: for I will
hasten my
word to
perform it.
\VS{13}And the
word of the
{\ND{
LORD}} came unto me the second
time,
saying, What
seest thou? And I
said, I
see a
seething
pot; and the
face thereof
{\ADD{is}}
toward the
north.\FTNT{†}{{\FR 1.13 }
toward…: Heb. from the face of the north}
\VS{14}Then the
{\ND{
LORD}}
said unto me, Out of the
north an
evil shall break
forth upon all the
inhabitants of the
land.\FTNT{‡}{{\FR 1.14 }
shall…: Heb. shall be opened}
\VS{15}For, lo, I will
call all the
families of the
kingdoms of the
north,
saith the
{\ND{
LORD}}; and they shall
come, and they shall
set every
one his
throne at the
entering of the
gates of
Jerusalem, and against all the
walls thereof round
about, and against all the
cities of
Judah.
\VS{16}And I will
utter my
judgments against them
touching all their
wickedness, who have
forsaken me, and have burned
incense unto
other
gods, and
worshipped the
works of their own
hands.
\par }{\PP \VS{17}¶ Thou therefore gird
up thy
loins, and
arise, and
speak unto them all that I
command thee: be not
dismayed at their
faces, lest I
confound thee
before them.\FTNT{§}{{\FR 1.17 }
confound: or, break to pieces}
\VS{18}For, behold,
I have
made thee this
day a
defenced
city, and an
iron
pillar, and
brasen
walls against the whole
land, against the
kings of
Judah, against the
princes thereof, against the
priests thereof, and against the
people of the
land.
\VS{19}And they shall
fight against thee; but they shall not
prevail against thee; for I
{\ADD{am}} with thee,
saith the
{\ND{
LORD}}, to
deliver thee.

\par }\Chap{2}{\PP \VerseOne{1}Moreover the
word of the
{\ND{
LORD}} came to me,
saying,
\VS{2}Go and
cry in the
ears of
Jerusalem,
saying, Thus
saith the
{\ND{
LORD}}; I
remember thee, the
kindness of thy
youth, the
love of thine
espousals, when thou
wentest
after me in the
wilderness, in a
land
{\ADD{that was}} not
sown.\FTNT{*}{{\FR 2.2 }
thee: or, for thy sake}
\VS{3}Israel
{\ADD{was}}
holiness unto the
{\ND{
LORD}},
{\ADD{and}} the
firstfruits of his
increase: all that
devour him shall
offend;
evil shall
come upon them,
saith the
{\ND{
LORD}}.
\VS{4}Hear ye the
word of the
{\ND{
LORD}}, O
house of
Jacob, and all the
families of the
house of
Israel:
\par }{\PP \VS{5}¶ Thus
saith the
{\ND{
LORD}}, What
iniquity have your
fathers
found in me, that they are gone
far from me, and have
walked
after
vanity, and are become
vain?
\VS{6}Neither
said they, Where
{\ADD{is}} the
{\ND{
LORD}} that brought us
up out of the
land of
Egypt, that
led us through the
wilderness, through a
land of
deserts and of
pits, through a
land of
drought, and of the shadow of
death, through a
land that no
man passed
through, and where no
man
dwelt?
\VS{7}And I
brought you into a
plentiful
country, to
eat the
fruit thereof and the
goodness thereof; but when ye
entered, ye
defiled my
land, and
made mine
heritage an
abomination.\FTNT{†}{{\FR 2.7 }
a plentiful…: or, the land of Carmel}
\VS{8}The
priests
said not, Where
{\ADD{is}} the
{\ND{
LORD}}? and they that
handle the
law
knew me not: the
pastors also
transgressed against me, and the
prophets
prophesied by
Baal, and
walked
after
{\ADD{things that}} do not
profit.
\par }{\PP \VS{9}¶ Wherefore I will yet
plead with you,
saith the
{\ND{
LORD}}, and with your
children’s
children will I
plead.
\VS{10}For pass
over the
isles of
Chittim, and
see; and
send unto
Kedar, and
consider
diligently, and
see if there be such a thing.\FTNT{‡}{{\FR 2.10 }
over: or, over to}
\VS{11}Hath a
nation
changed
{\ADD{their}}
gods, which
{\ADD{are}} yet no
gods? but my
people have
changed their
glory for
{\ADD{that which}} doth not
profit.
\VS{12}Be
astonished, O ye
heavens, at this, and be horribly
afraid, be ye
very
desolate,
saith the
{\ND{
LORD}}.
\VS{13}For my
people have
committed
two
evils; they have
forsaken me the
fountain of
living
waters,
{\ADD{and}} hewed them
out
cisterns,
broken
cisterns, that can
hold no
water.
\par }{\PP \VS{14}¶
{\ADD{Is}}
Israel a
servant?
{\ADD{is}} he a
homeborn
{\ADD{slave}}?
why is he
spoiled?\FTNT{§}{{\FR 2.14 }
spoiled: Heb. become a spoil?}
\VS{15}The young
lions
roared upon him,
{\ADD{and}}
yelled, and they
made his
land
waste: his
cities are
burned without
inhabitant.\FTNT{**}{{\FR 2.15 }
yelled: Heb. gave out their voice}
\VS{16}Also the
children of
Noph and
Tahapanes have
broken the crown of thy
head.\FTNT{††}{{\FR 2.16 }
have…: or, feed on thy crown}
\VS{17}Hast thou not
procured this unto thyself, in that thou hast
forsaken the
{\ND{
LORD}} thy
God,
when he
led thee by the
way?
\VS{18}And now what hast thou to do in the
way of
Egypt, to
drink the
waters of
Sihor? or what hast thou to do in the
way of
Assyria, to
drink the
waters of the
river?
\VS{19}Thine own
wickedness shall
correct thee, and thy
backslidings shall
reprove thee:
know therefore and
see that
{\ADD{it is}} an
evil
{\ADD{thing}} and
bitter, that thou hast
forsaken the
{\ND{
LORD}} thy
God, and that my
fear
{\ADD{is}} not in thee,
saith the
Lord
{\ND{
GOD}} of
hosts.
\par }{\PP \VS{20}¶ For of old
time I have
broken thy
yoke,
{\ADD{and}}
burst thy
bands; and thou
saidst, I will not
transgress; when upon every
high
hill and under every
green
tree thou
wanderest, playing the
harlot.\FTNT{‡‡}{{\FR 2.20 }
transgress: or, serve}
\VS{21}Yet I had
planted thee a noble
vine, wholly a
right
seed: how then art thou
turned into the degenerate
plant of a
strange
vine unto me?
\VS{22}For though thou
wash thee with
nitre, and take thee
much
soap,
{\ADD{yet}} thine
iniquity is
marked
before me,
saith the
Lord
{\ND{
GOD}}.
\VS{23}How canst thou
say, I am not
polluted, I have not
gone
after
Baalim?
see thy
way in the
valley,
know what thou hast
done:
{\ADD{thou art}} a
swift
dromedary
traversing her
ways;\FTNT{§§}{{\FR 2.23 }
thou art…: or, O swift dromedary}
\VS{24}A wild
ass
used to the
wilderness,
{\ADD{that}} snuffeth
up the
wind at her
pleasure; in her
occasion who can turn her
away? all they that
seek her will not
weary themselves; in her
month they shall
find her.\FTNT{***}{{\FR 2.24 }
A wild…: or, O wild ass, etc}\FTNT{†††}{{\FR 2.24 }
used: Heb. taught}\FTNT{‡‡‡}{{\FR 2.24 }
her pleasure: Heb. the desire of her heart}\FTNT{§§§}{{\FR 2.24 }
turn…: or, reverse it?}
\VS{25}Withhold thy
foot from being
unshod, and thy
throat from
thirst: but thou
saidst, There is no
hope: no; for I have
loved
strangers, and
after them will I
go.\FTNT{*}{{\FR 2.25 }
There…: or, Is the case desperate?}
\VS{26}As the
thief is
ashamed when he is
found, so is the
house of
Israel
ashamed; they, their
kings, their
princes, and their
priests, and their
prophets,
\VS{27}Saying to a
stock, Thou
{\ADD{art}} my
father; and to a
stone, Thou hast brought me
forth: for they have
turned
{\ADD{their}}
back unto me, and not
{\ADD{their}}
face: but in the
time of their
trouble they will
say,
Arise, and
save us.\FTNT{†}{{\FR 2.27 }
brought…: or, begotten me}\FTNT{‡}{{\FR 2.27 }
their back: Heb. the hinder part of the neck}
\VS{28}But where
{\ADD{are}} thy
gods that thou hast
made thee? let them
arise, if they can
save thee in the
time of thy
trouble: for
{\ADD{according to}} the
number of thy
cities are thy
gods, O
Judah.\FTNT{§}{{\FR 2.28 }
trouble: Heb. evil}
\par }{\PP \VS{29}Wherefore will ye
plead with me? ye all have
transgressed against me,
saith the
{\ND{
LORD}}.
\VS{30}In
vain have I
smitten your
children; they
received no
correction: your own
sword hath
devoured your
prophets, like a
destroying
lion.
\par }{\PP \VS{31}¶ O
generation,
see ye the
word of the
{\ND{
LORD}}. Have I been a
wilderness unto
Israel? a
land of
darkness? wherefore
say my
people, We are
lords; we will
come no more unto thee?\FTNT{**}{{\FR 2.31 }
We are…: Heb. We have dominion}
\VS{32}Can a
maid
forget her
ornaments,
{\ADD{or}} a
bride her
attire? yet my
people have
forgotten me
days without
number.
\VS{33}Why
trimmest thou thy
way to
seek
love? therefore hast thou also
taught the wicked
ones thy
ways.
\VS{34}Also in thy
skirts is
found the
blood of the
souls of the
poor
innocents: I have not
found it by secret
search, but upon all these.\FTNT{††}{{\FR 2.34 }
secret…: Heb. digging}
\VS{35}Yet thou
sayest, Because I am
innocent, surely his
anger shall
turn from me. Behold, I will
plead with thee, because thou
sayest, I have not
sinned.
\VS{36}Why gaddest thou
about so
much to
change thy
way? thou also shalt be
ashamed of
Egypt, as thou wast
ashamed of
Assyria.
\VS{37}Yea, thou shalt go
forth from him, and thine
hands upon thine
head: for the
{\ND{
LORD}} hath
rejected thy
confidences, and thou shalt not
prosper in them.

\par }\Chap{3}{\PP \VerseOne{1}They
say, If a
man put
away his
wife, and she
go from him, and become
another
man’s, shall he return unto her
again? shall not that
land be
greatly
polluted? but thou hast played the
harlot with
many
lovers; yet return
again to me,
saith the
{\ND{
LORD}}.\FTNT{*}{{\FR 3.1 }
They say: Heb. Saying}
\VS{2}Lift
up thine
eyes unto the high
places, and
see
where thou hast not been
lien with. In the
ways hast thou
sat for them, as the
Arabian in the
wilderness; and thou hast
polluted the
land with thy
whoredoms and with thy
wickedness.
\VS{3}Therefore the
showers have been
withholden, and there hath been no latter
rain; and thou hadst a
whore’s
forehead, thou
refusedst to be
ashamed.
\VS{4}Wilt thou not from this time
cry unto me, My
father, thou
{\ADD{art}} the
guide of my
youth?
\VS{5}Will he
reserve
{\ADD{his anger}} for
ever? will he
keep
{\ADD{it}} to the
end? Behold, thou hast
spoken and
done evil
things as thou
couldest.
\par }{\PP \VS{6}¶ The
{\ND{
LORD}}
said also unto me in the
days of
Josiah the
king, Hast thou
seen
{\ADD{that}} which
backsliding
Israel hath
done? she is gone
up upon every
high
mountain and
under every
green
tree, and there hath played the
harlot.
\VS{7}And I
said
after she had
done all these
{\ADD{things}},
Turn thou unto me. But she
returned not. And her
treacherous
sister
Judah
saw
{\ADD{it}}.
\VS{8}And I
saw, when for all the
causes whereby
backsliding
Israel committed
adultery I had put her
away, and
given her a
bill of
divorce; yet her
treacherous
sister
Judah
feared not, but
went and played the
harlot also.
\VS{9}And it came to pass through the
lightness of her
whoredom, that she
defiled the
land, and committed
adultery with
stones and with
stocks.\FTNT{†}{{\FR 3.9 }
lightness: or, fame}
\VS{10}And yet for all this her
treacherous
sister
Judah hath not
turned unto me with her whole
heart, but
feignedly,
saith the
{\ND{
LORD}}.\FTNT{‡}{{\FR 3.10 }
feignedly: Heb. in falsehood}
\VS{11}And the
{\ND{
LORD}}
said unto me, The
backsliding
Israel hath
justified
herself more than
treacherous
Judah.
\par }{\PP \VS{12}¶
Go and
proclaim these
words toward the
north, and
say,
Return, thou
backsliding
Israel,
saith the
{\ND{
LORD}};
{\ADD{and}} I will not cause mine
anger to
fall upon you: for I
{\ADD{am}}
merciful,
saith the
{\ND{
LORD}},
{\ADD{and}} I will not
keep
{\ADD{anger}} for
ever.
\VS{13}Only
acknowledge thine
iniquity, that thou hast
transgressed against the
{\ND{
LORD}} thy
God, and hast
scattered thy
ways to the
strangers under every
green
tree, and ye have not
obeyed my
voice,
saith the
{\ND{
LORD}}.
\VS{14}Turn, O
backsliding
children,
saith the
{\ND{
LORD}}; for I am
married unto you: and I will
take you
one of a
city, and
two of a
family, and I will
bring you to
Zion:
\VS{15}And I will
give you
pastors according to mine
heart, which shall
feed you with
knowledge and
understanding.
\VS{16}And it shall come to pass, when ye be
multiplied and
increased in the
land, in those
days,
saith the
{\ND{
LORD}}, they shall
say no more, The
ark of the
covenant of the
{\ND{
LORD}}: neither shall it
come to
mind: neither shall they
remember it; neither shall they
visit
{\ADD{it;}} neither shall
{\ADD{that}} be
done any more.\FTNT{§}{{\FR 3.16 }
come to mind: Heb. come upon the heart}\FTNT{**}{{\FR 3.16 }
that…: or, it be magnified}
\VS{17}At that
time they shall
call
Jerusalem the
throne of the
{\ND{
LORD}}; and all the
nations shall be
gathered unto it, to the
name of the
{\ND{
LORD}}, to
Jerusalem: neither shall they
walk any more
after the
imagination of their
evil
heart.\FTNT{††}{{\FR 3.17 }
imagination: or, stubbornness}
\VS{18}In those
days the
house of
Judah shall
walk with the
house of
Israel, and they shall
come
together out of the
land of the
north to the
land that I have given for an
inheritance unto your
fathers.\FTNT{‡‡}{{\FR 3.18 }
with: or, to}\FTNT{§§}{{\FR 3.18 }
given…: or, caused your fathers to possess}
\VS{19}But I
said, How shall I
put thee among the
children, and
give thee a
pleasant
land, a
goodly
heritage of the
hosts of
nations? and I
said, Thou shalt
call me, My
father; and shalt not turn
away from
me.\FTNT{***}{{\FR 3.19 }
from me: Heb. from after me}\FTNT{†††}{{\FR 3.19 }
a goodly…: Heb. an heritage of glory, or, beauty}\FTNT{‡‡‡}{{\FR 3.19 }
pleasant…: Heb. land of desire}
\par }{\PP \VS{20}¶
Surely
{\ADD{as}} a
wife
treacherously departeth from her
husband, so have ye dealt
treacherously with me, O
house of
Israel,
saith the
{\ND{
LORD}}.\FTNT{§§§}{{\FR 3.20 }
husband: Heb. friend}
\VS{21}A
voice was
heard upon the high
places,
weeping
{\ADD{and}}
supplications of the
children of
Israel: for they have
perverted their
way,
{\ADD{and}} they have
forgotten the
{\ND{
LORD}} their
God.
\VS{22}Return, ye
backsliding
children,
{\ADD{and}} I will
heal your
backslidings. Behold, we
come unto thee; for thou
{\ADD{art}} the
{\ND{
LORD}} our
God.
\VS{23}Truly in
vain
{\ADD{is salvation hoped for}} from the
hills,
{\ADD{and from}} the
multitude of
mountains:
truly in the
{\ND{
LORD}} our
God
{\ADD{is}} the
salvation of
Israel.
\VS{24}For
shame hath
devoured the
labour of our
fathers from our
youth; their
flocks and their
herds, their
sons and their
daughters.
\VS{25}We lie
down in our
shame, and our
confusion
covereth us: for we have
sinned against the
{\ND{
LORD}} our
God, we and our
fathers, from our
youth even unto this
day, and have not
obeyed the
voice of the
{\ND{
LORD}} our
God.

\par }\Chap{4}{\PP \VerseOne{1}If thou wilt
return, O
Israel,
saith the
{\ND{
LORD}},
return unto me: and if thou wilt put
away thine
abominations out of my
sight, then shalt thou not
remove.
\VS{2}And thou shalt
swear, The
{\ND{
LORD}}
liveth, in
truth, in
judgment, and in
righteousness; and the
nations shall
bless themselves in him, and in him shall they
glory.
\par }{\PP \VS{3}¶ For thus
saith the
{\ND{
LORD}} to the
men of
Judah and
Jerusalem, Break
up your fallow
ground, and
sow not among
thorns.
\VS{4}Circumcise yourselves to the
{\ND{
LORD}}, and take
away the
foreskins of your
heart, ye
men of
Judah and
inhabitants of
Jerusalem: lest my
fury come
forth like
fire, and
burn that none can
quench
{\ADD{it}},
because of the
evil of your
doings.
\par }{\PP \VS{5}Declare ye in
Judah, and
publish in
Jerusalem; and
say,
Blow ye the
trumpet in the
land:
cry, gather
together, and
say,
Assemble yourselves, and let us
go into the
defenced
cities.
\VS{6}Set
up the
standard toward
Zion:
retire,
stay not: for I will
bring
evil from the
north, and a
great
destruction.\FTNT{*}{{\FR 4.6 }
retire: or, strengthen}\FTNT{†}{{\FR 4.6 }
destruction: Heb. breaking}
\VS{7}The
lion is come
up from his
thicket, and the
destroyer of the
Gentiles is on his
way; he is gone
forth from his
place to
make thy
land
desolate;
{\ADD{and}} thy
cities shall be laid
waste, without an
inhabitant.
\VS{8}For this
gird you with
sackcloth,
lament and
howl: for the
fierce
anger of the
{\ND{
LORD}} is not turned
back from us.
\VS{9}And it shall come to pass at that
day,
saith the
{\ND{
LORD}},
{\ADD{that}} the
heart of the
king shall
perish, and the
heart of the
princes; and the
priests shall be
astonished, and the
prophets shall
wonder.
\VS{10}Then
said I,
Ah,
Lord
{\ND{
GOD}}!
surely thou hast
greatly
deceived this
people and
Jerusalem,
saying, Ye shall have
peace; whereas the
sword
reacheth unto the
soul.
\VS{11}At that
time shall it be
said to this
people and to
Jerusalem, A
dry
wind of the high
places in the
wilderness
toward the
daughter of my
people, not to
fan, nor to
cleanse,
\VS{12}{\ADD{Even}} a
full
wind from those
{\ADD{places}} shall
come unto me: now also will I
give
sentence against them.\FTNT{‡}{{\FR 4.12 }
a full…: or, a fuller wind than those}\FTNT{§}{{\FR 4.12 }
give…: Heb. utter judgments}
\VS{13}Behold, he shall come
up as
clouds, and his
chariots
{\ADD{shall be}} as a
whirlwind: his
horses are
swifter than
eagles.
Woe unto us! for we are
spoiled.
\VS{14}O
Jerusalem,
wash thine
heart from
wickedness, that thou mayest be
saved. How long shall thy
vain
thoughts
lodge
within thee?
\VS{15}For a
voice
declareth from
Dan, and
publisheth
affliction from
mount
Ephraim.
\VS{16}Make ye
mention to the
nations; behold,
publish against
Jerusalem,
{\ADD{that}}
watchers
come from a
far
country, and give
out their
voice against the
cities of
Judah.
\VS{17}As
keepers of a
field, are they against her round
about; because she hath been
rebellious against me,
saith the
{\ND{
LORD}}.
\VS{18}Thy
way and thy
doings have
procured these
{\ADD{things}} unto thee; this
{\ADD{is}} thy
wickedness, because it is
bitter, because it
reacheth unto thine
heart.
\par }{\PP \VS{19}¶ My
bowels, my
bowels! I am
pained at my
very
heart; my
heart maketh a
noise in me; I cannot hold my
peace, because thou hast
heard, O my
soul, the
sound of the
trumpet, the
alarm of
war.\FTNT{**}{{\FR 4.19 }
my very…: Heb. the walls of my heart}
\VS{20}Destruction upon
destruction is
cried; for the whole
land is
spoiled:
suddenly are my
tents
spoiled,
{\ADD{and}} my
curtains in a
moment.
\VS{21}How long shall I
see the
standard,
{\ADD{and}}
hear the
sound of the
trumpet?
\VS{22}For my
people
{\ADD{is}}
foolish, they have not
known me; they
{\ADD{are}}
sottish
children, and they have none
understanding: they
{\ADD{are}}
wise to do
evil, but to do
good they have no
knowledge.
\VS{23}I
beheld the
earth, and, lo,
{\ADD{it was}} without
form, and
void; and the
heavens, and they
{\ADD{had}} no
light.
\VS{24}I
beheld the
mountains, and, lo, they
trembled, and all the
hills moved
lightly.
\VS{25}I
beheld, and, lo,
{\ADD{there was}} no
man, and all the
birds of the
heavens were
fled.
\VS{26}I
beheld, and, lo, the fruitful
place
{\ADD{was}} a
wilderness, and all the
cities thereof were broken
down at the
presence of the
{\ND{
LORD}},
{\ADD{and}} by his
fierce
anger.
\VS{27}For thus hath the
{\ND{
LORD}}
said, The whole
land shall be
desolate; yet will I not
make a full
end.
\VS{28}For this shall the
earth
mourn, and the
heavens
above be
black: because I have
spoken
{\ADD{it}}, I have
purposed
{\ADD{it}}, and will not
repent, neither will I turn
back from it.
\VS{29}The whole
city shall
flee for the
noise of the
horsemen and
bowmen; they shall
go into
thickets, and climb
up upon the
rocks: every
city
{\ADD{shall be}}
forsaken, and not a
man
dwell
therein.
\VS{30}And
{\ADD{when}} thou
{\ADD{art}}
spoiled, what wilt thou
do? Though thou
clothest thyself with
crimson, though thou
deckest thee with
ornaments of
gold, though thou
rentest thy
face with
painting, in
vain shalt thou make thyself
fair;
{\ADD{thy}}
lovers will
despise thee, they will
seek thy
life.\FTNT{††}{{\FR 4.30 }
face: Heb. eyes}
\VS{31}For I have
heard a
voice as of a woman in
travail,
{\ADD{and}} the
anguish as of her that bringeth forth her first
child, the
voice of the
daughter of
Zion,
{\ADD{that}}
bewaileth herself,
{\ADD{that}}
spreadeth her
hands,
{\ADD{saying}},
Woe
{\ADD{is}} me now! for my
soul is
wearied because of
murderers.

\par }\Chap{5}{\PP \VerseOne{1}Run ye to and
fro through the
streets of
Jerusalem, and
see now, and
know, and
seek in the broad
places thereof, if ye can
find a
man, if there
be
{\ADD{any}} that
executeth
judgment, that
seeketh the
truth; and I will
pardon it.
\VS{2}And though they
say, The
{\ND{
LORD}}
liveth; surely they
swear
falsely.
\VS{3}O
{\ND{
LORD}},
{\ADD{are}} not thine
eyes upon the
truth? thou hast
stricken them, but they have not
grieved; thou hast
consumed them,
{\ADD{but}} they have
refused to
receive
correction: they have made their
faces
harder than a
rock; they have
refused to
return.
\VS{4}Therefore I
said, Surely these
{\ADD{are}}
poor; they are
foolish: for they
know not the
way of the
{\ND{
LORD}},
{\ADD{nor}} the
judgment of their
God.
\VS{5}I will
get me unto the great
men, and will
speak unto them; for they have
known the
way of the
{\ND{
LORD}},
{\ADD{and}} the
judgment of their
God: but these have
altogether
broken the
yoke,
{\ADD{and}}
burst the
bonds.
\VS{6}Wherefore a
lion out of the
forest shall
slay them,
{\ADD{and}} a
wolf of the
evenings shall
spoil them, a
leopard shall
watch over their
cities: every one that goeth
out thence
shall be torn in
pieces: because their
transgressions are
many,
{\ADD{and}} their
backslidings are
increased.\FTNT{*}{{\FR 5.6 }
evenings: or, deserts}\FTNT{†}{{\FR 5.6 }
are increased: Heb. are strong}
\par }{\PP \VS{7}¶
How shall I
pardon thee for
this? thy
children have
forsaken me, and
sworn by
{\ADD{them that are}}
no
gods: when I had fed them to the
full, they then committed
adultery, and assembled themselves by
troops in the
harlots’
houses.
\VS{8}They were
{\ADD{as}}
fed
horses in the
morning: every
one
neighed after his
neighbour’s
wife.
\VS{9}Shall I not
visit for these
{\ADD{things}}?
saith the
{\ND{
LORD}}: and shall not my
soul be
avenged on such a
nation as this?
\par }{\PP \VS{10}¶ Go ye
up upon her
walls, and
destroy; but
make not a full
end: take
away her
battlements; for they
{\ADD{are}} not the
{\ND{
LORD’s}}.
\VS{11}For the
house of
Israel and the
house of
Judah have dealt
very
treacherously against me,
saith the
{\ND{
LORD}}.
\VS{12}They have
belied the
{\ND{
LORD}}, and
said,
{\ADD{It is}} not he; neither shall
evil
come upon us; neither shall we
see
sword nor
famine:
\VS{13}And the
prophets shall become
wind, and the
word
{\ADD{is}} not in them: thus shall it be
done unto them.
\VS{14}Wherefore thus
saith the
{\ND{
LORD}}
God of
hosts, Because ye
speak this
word, behold, I will
make my
words in thy
mouth
fire, and this
people
wood, and it shall
devour them.
\VS{15}Lo, I will
bring a
nation upon you from
far, O
house of
Israel,
saith the
{\ND{
LORD}}: it
{\ADD{is}} a
mighty
nation, it
{\ADD{is}} an
ancient
nation, a
nation whose
language thou
knowest not, neither
understandest what they
say.
\VS{16}Their
quiver
{\ADD{is}} as an
open
sepulchre, they
{\ADD{are}} all mighty
men.
\VS{17}And they shall eat
up thine
harvest, and thy
bread,
{\ADD{which}} thy
sons and thy
daughters should
eat: they shall eat
up thy
flocks and thine
herds: they shall eat
up thy
vines and thy fig
trees: they shall
impoverish thy
fenced
cities, wherein
thou
trustedst, with the
sword.
\VS{18}Nevertheless in those
days,
saith the
{\ND{
LORD}}, I will not
make a full
end with you.
\par }{\PP \VS{19}¶ And it shall come to pass, when ye shall
say,
Wherefore
doeth the
{\ND{
LORD}} our
God all these
{\ADD{things}} unto us? then shalt thou
answer them, Like as ye have
forsaken me, and
served
strange
gods in your
land, so shall ye
serve
strangers in a
land
{\ADD{that is}} not yours.
\par }{\PP \VS{20}Declare this in the
house of
Jacob, and
publish it in
Judah,
saying,
\VS{21}Hear now this, O
foolish
people, and without
understanding; which have
eyes, and
see not; which have
ears, and
hear not:\FTNT{‡}{{\FR 5.21 }
understanding: Heb. heart}
\VS{22}Fear ye not me?
saith the
{\ND{
LORD}}: will ye not
tremble at my
presence, which have
placed the
sand
{\ADD{for}} the
bound of the
sea by a
perpetual
decree, that it cannot
pass it: and though the
waves thereof
toss themselves, yet can they not
prevail; though they
roar, yet can they not pass
over it?
\VS{23}But this
people hath a
revolting and a
rebellious
heart; they are
revolted and
gone.
\VS{24}Neither
say they in their
heart, Let us now
fear the
{\ND{
LORD}} our
God, that
giveth
rain, both the
former and the
latter, in his
season: he
reserveth unto us the
appointed
weeks of the
harvest.
\par }{\PP \VS{25}¶ Your
iniquities have turned
away these
{\ADD{things}}, and your
sins have
withholden
good
{\ADD{things}} from you.
\VS{26}For among my
people are
found
wicked
{\ADD{men:}} they lay
wait, as he that
setteth
snares; they
set a
trap, they
catch
men.\FTNT{§}{{\FR 5.26 }
they lay…: or, they pry as fowlers lie in wait}
\VS{27}As a
cage is
full of
birds, so
{\ADD{are}} their
houses
full of
deceit: therefore they are become
great, and waxen
rich.\FTNT{**}{{\FR 5.27 }
cage: or, coop}
\VS{28}They are waxen
fat, they
shine: yea, they
overpass the
deeds of the
wicked: they
judge not the
cause, the cause of the
fatherless, yet they
prosper; and the
right of the
needy do they not
judge.
\VS{29}Shall I not
visit for these
{\ADD{things}}?
saith the
{\ND{
LORD}}: shall not my
soul be
avenged on such a
nation as this?
\par }{\PP \VS{30}¶ A
wonderful and horrible
thing is
committed in the
land;\FTNT{††}{{\FR 5.30 }
A wonderful…: or, Astonishment and filthiness}
\VS{31}The
prophets
prophesy
falsely, and the
priests bear
rule by their
means; and my
people
love
{\ADD{to have it}} so: and what will ye
do in the
end thereof?\FTNT{‡‡}{{\FR 5.31 }
bear…: or, take into their hands}

\par }\Chap{6}{\PP \VerseOne{1}O ye
children of
Benjamin, gather yourselves to
flee out of the
midst of
Jerusalem, and
blow the
trumpet in
Tekoa, and set
up a
sign of fire in
Beth-haccerem: for
evil
appeareth out of the
north, and
great
destruction.
\VS{2}I have
likened the
daughter of
Zion to a
comely and
delicate
{\ADD{woman}}.\FTNT{*}{{\FR 6.2 }
comely: or, dwelling at home}
\VS{3}The
shepherds with their
flocks shall
come unto her; they shall
pitch
{\ADD{their}}
tents against her round
about; they shall
feed every
one in his
place.
\VS{4}Prepare ye
war against her;
arise, and let us go
up at
noon.
Woe unto us! for the
day goeth
away, for the
shadows of the
evening are stretched
out.
\VS{5}Arise, and let us
go by
night, and let us
destroy her
palaces.
\par }{\PP \VS{6}¶ For thus hath the
{\ND{
LORD}} of
hosts
said, Hew ye
down
trees, and
cast a
mount against
Jerusalem: this
{\ADD{is}} the
city to be
visited; she
{\ADD{is}} wholly
oppression in the
midst of her.\FTNT{†}{{\FR 6.6 }
cast…: or, pour out the engine of shot}
\VS{7}As a
fountain casteth
out her
waters, so she casteth
out her
wickedness:
violence and
spoil is
heard in her; before
me
continually
{\ADD{is}}
grief and
wounds.
\VS{8}Be thou
instructed, O
Jerusalem, lest my
soul
depart from thee; lest I
make thee
desolate, a
land not
inhabited.\FTNT{‡}{{\FR 6.8 }
depart: Heb. be loosed, or, disjointed}
\par }{\PP \VS{9}¶ Thus
saith the
{\ND{
LORD}} of
hosts, They shall
throughly
glean the
remnant of
Israel as a
vine: turn
back thine
hand as a
grapegatherer into the
baskets.
\VS{10}To whom shall I
speak, and give
warning, that they may
hear? behold, their
ear
{\ADD{is}}
uncircumcised, and they
cannot
hearken: behold, the
word of the
{\ND{
LORD}} is unto them a
reproach; they have no
delight in it.
\VS{11}Therefore I am
full of the
fury of the
{\ND{
LORD}}; I am
weary with holding
in: I will pour it
out upon the
children
abroad, and upon the
assembly of young
men
together: for even the
husband with the
wife shall be
taken, the
aged with
{\ADD{him that is}}
full of
days.
\VS{12}And their
houses shall be
turned unto
others,
{\ADD{with their}}
fields and
wives
together: for I will stretch
out my
hand upon the
inhabitants of the
land,
saith the
{\ND{
LORD}}.
\VS{13}For from the
least of them even unto the
greatest of them every one
{\ADD{is}}
given to
covetousness; and from the
prophet even unto the
priest every one
dealeth
falsely.
\VS{14}They have
healed also the
hurt
{\ADD{
of the daughter}} of my
people
slightly,
saying,
Peace,
peace; when
{\ADD{there is}} no
peace.\FTNT{§}{{\FR 6.14 }
hurt: Heb. bruise, or, breach}
\VS{15}Were they
ashamed when they had
committed
abomination?
nay, they were not at
all
ashamed,
neither
could they
blush: therefore they shall
fall among them that
fall: at the
time
{\ADD{that}} I
visit them they shall be cast
down,
saith the
{\ND{
LORD}}.
\VS{16}Thus
saith the
{\ND{
LORD}},
Stand ye in the
ways, and
see, and
ask for the
old
paths, where
{\ADD{is}} the
good
way, and
walk therein, and ye shall
find
rest for your
souls. But they
said, We will not
walk
{\ADD{therein}}.
\VS{17}Also I
set
watchmen over you,
{\ADD{saying}},
Hearken to the
sound of the
trumpet. But they
said, We will not
hearken.
\par }{\PP \VS{18}¶ Therefore
hear, ye
nations, and
know, O
congregation, what
{\ADD{is}} among them.
\VS{19}Hear, O
earth: behold, I will
bring
evil upon this
people,
{\ADD{even}} the
fruit of their
thoughts, because they have not
hearkened unto my
words, nor to my
law, but
rejected it.
\VS{20}To what purpose
cometh there to me
incense from
Sheba, and the
sweet
cane from a
far
country? your burnt
offerings
{\ADD{are}} not
acceptable, nor your
sacrifices
sweet unto me.
\VS{21}Therefore thus
saith the
{\ND{
LORD}}, Behold, I will
lay
stumblingblocks before this
people, and the
fathers and the
sons
together shall
fall upon them; the
neighbour and his
friend shall
perish.
\VS{22}Thus
saith the
{\ND{
LORD}}, Behold, a
people
cometh from the
north
country, and a
great
nation shall be
raised from the
sides of the
earth.
\VS{23}They shall lay
hold on
bow and
spear; they
{\ADD{are}}
cruel, and have no
mercy; their
voice
roareth like the
sea; and they
ride upon
horses, set in
array as
men for
war against thee, O
daughter of
Zion.
\VS{24}We have
heard the
fame thereof: our
hands wax
feeble:
anguish hath taken
hold of us,
{\ADD{and}}
pain, as of a woman in
travail.
\VS{25}Go not
forth into the
field, nor
walk by the
way; for the
sword of the
enemy
{\ADD{and}}
fear
{\ADD{is}} on every
side.
\par }{\PP \VS{26}¶ O
daughter of my
people,
gird
{\ADD{thee}} with
sackcloth, and
wallow thyself in
ashes:
make thee
mourning,
{\ADD{as for}} an only
son, most
bitter
lamentation: for the
spoiler shall
suddenly
come upon us.
\VS{27}I have
set thee
{\ADD{for}} a
tower
{\ADD{and}} a
fortress among my
people, that thou mayest
know and
try their
way.
\VS{28}They
{\ADD{are}} all
grievous
revolters,
walking with
slanders:
{\ADD{they are}}
brass and
iron; they
{\ADD{are}} all
corrupters.
\VS{29}The
bellows are
burned, the
lead is
consumed of the
fire; the
founder
melteth in
vain: for the
wicked are not plucked
away.
\VS{30}Reprobate
silver shall
{\ADD{men}}
call them, because the
{\ND{
LORD}} hath
rejected them.\FTNT{**}{{\FR 6.30 }
Reprobate…: or, Refuse silver}

\par }\Chap{7}{\PP \VerseOne{1}The
word that came to
Jeremiah from the
{\ND{
LORD}},
saying,
\VS{2}Stand in the
gate of the
{\ND{
LORD’s}}
house, and
proclaim there this
word, and
say,
Hear the
word of the
{\ND{
LORD}}, all
{\ADD{ye of}}
Judah, that enter
in at these
gates to
worship the
{\ND{
LORD}}.
\VS{3}Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel,
Amend your
ways and your
doings, and I will cause you to
dwell in this
place.
\VS{4}Trust ye not in
lying
words,
saying, The
temple of the
{\ND{
LORD}}, The
temple of the
{\ND{
LORD}}, The
temple of the
{\ND{
LORD}},
{\ADD{are}} these.
\VS{5}For if ye
throughly
amend your
ways and your
doings; if ye
throughly
execute
judgment between a
man and his
neighbour;
\VS{6}{\ADD{If}} ye
oppress not the
stranger, the
fatherless, and the
widow, and
shed not
innocent
blood in this
place, neither
walk
after
other
gods to your
hurt:
\VS{7}Then will I cause you to
dwell in this
place, in the
land that I
gave to your
fathers,
for
ever and
ever.
\par }{\PP \VS{8}¶ Behold, ye
trust in
lying
words, that cannot
profit.
\VS{9}Will ye
steal,
murder, and commit
adultery, and
swear
falsely, and burn
incense unto
Baal, and
walk
after
other
gods whom ye
know not;
\VS{10}And
come and
stand
before me in this
house, which is
called by my
name, and
say, We are
delivered to
do all these
abominations?\FTNT{*}{{\FR 7.10 }
which…: Heb. whereupon my name is called}
\VS{11}Is this
house, which is
called by my
name, become a
den of
robbers in your
eyes? Behold, even I have
seen
{\ADD{it}},
saith the
{\ND{
LORD}}.
\VS{12}But
go ye now unto my
place which
{\ADD{was}} in
Shiloh, where I
set my
name at the
first, and
see what I
did to it
for the
wickedness of my
people
Israel.
\VS{13}And now, because ye have
done all these
works,
saith the
{\ND{
LORD}}, and I
spake unto you, rising up
early and
speaking, but ye
heard not; and I
called you, but ye
answered not;
\VS{14}Therefore will I
do unto
{\ADD{this}}
house, which is
called by my
name, wherein ye
trust, and unto the
place which I
gave to you and to your
fathers, as I have
done to
Shiloh.
\VS{15}And I will cast you
out of my
sight, as I have cast
out all your
brethren,
{\ADD{even}} the whole
seed of
Ephraim.
\par }{\PP \VS{16}Therefore
pray not thou for this
people, neither lift
up
cry nor
prayer for
them, neither make
intercession to me: for I will not
hear thee.
\par }{\PP \VS{17}¶
Seest thou not what they
do in the
cities of
Judah and in the
streets of
Jerusalem?
\VS{18}The
children
gather
wood, and the
fathers
kindle the
fire, and the
women
knead
{\ADD{their}}
dough, to
make
cakes to the
queen of
heaven, and to pour
out drink
offerings unto
other
gods, that they may provoke me to
anger.\FTNT{†}{{\FR 7.18 }
queen…: or, frame, or, workmanship of heaven}
\VS{19}Do they provoke me to
anger?
saith the
{\ND{
LORD}}:
{\ADD{do they}} not
{\ADD{provoke}} themselves to the
confusion of their own
faces?
\VS{20}Therefore thus
saith the
Lord
{\ND{
GOD}}; Behold, mine
anger and my
fury shall be poured
out upon this
place, upon
man, and upon
beast, and upon the
trees of the
field, and upon the
fruit of the
ground; and it shall
burn, and shall not be
quenched.
\par }{\PP \VS{21}¶ Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel;
Put your burnt
offerings unto your
sacrifices, and
eat
flesh.
\VS{22}For I
spake not unto your
fathers, nor
commanded them in the
day that I brought them
out of the
land of
Egypt,
concerning burnt
offerings or
sacrifices:\FTNT{‡}{{\FR 7.22 }
concerning: Heb. concerning the matter of}
\VS{23}But this
thing
commanded I them,
saying,
Obey my
voice, and I will be your
God, and ye shall be my
people: and
walk ye in all the
ways that I have
commanded you, that it may be
well unto you.
\VS{24}But they
hearkened not, nor
inclined their
ear, but
walked in the
counsels
{\ADD{and}} in the
imagination of their
evil
heart, and went
backward, and not
forward.\FTNT{§}{{\FR 7.24 }
imagination: or, stubbornness}\FTNT{**}{{\FR 7.24 }
went: Heb. were}
\VS{25}Since the
day that your
fathers came forth
out of the
land of
Egypt unto this
day I have even
sent unto you all my
servants the
prophets,
daily rising up
early and
sending
{\ADD{them:}}
\VS{26}Yet they
hearkened not unto me, nor
inclined their
ear, but
hardened their
neck: they did
worse than their
fathers.
\VS{27}Therefore thou shalt
speak all these
words unto them; but they will not
hearken to thee: thou shalt also
call unto them; but they will not
answer thee.
\VS{28}But thou shalt
say unto them, This
{\ADD{is}} a
nation that
obeyeth not the
voice of the
{\ND{
LORD}} their
God, nor
receiveth
correction:
truth is
perished, and is cut
off from their
mouth.\FTNT{††}{{\FR 7.28 }
correction: or, instruction}
\par }{\PP \VS{29}¶ Cut
off thine
hair,
{\ADD{O Jerusalem}}, and cast
{\ADD{it}}
away, and take
up a
lamentation on high
places; for the
{\ND{
LORD}} hath
rejected and
forsaken the
generation of his
wrath.
\VS{30}For the
children of
Judah have
done
evil in my
sight,
saith the
{\ND{
LORD}}: they have
set their
abominations in the
house which is
called by my
name, to
pollute it.
\VS{31}And they have
built the high
places of
Tophet, which
{\ADD{is}} in the
valley of the
son of
Hinnom, to
burn their
sons and their
daughters in the
fire; which I
commanded
{\ADD{them}} not, neither
came it into my
heart.\FTNT{‡‡}{{\FR 7.31 }
came…: Heb. came it upon my heart}
\par }{\PP \VS{32}¶ Therefore, behold, the
days
come,
saith the
{\ND{
LORD}}, that it shall no more be
called
Tophet, nor the
valley of the
son of
Hinnom, but the
valley of
slaughter: for they shall
bury in
Tophet, till there be
no
place.
\VS{33}And the
carcases of this
people shall be
meat for the
fowls of the
heaven, and for the
beasts of the
earth; and none shall fray
{\ADD{them}}
away.
\VS{34}Then will I cause to
cease from the
cities of
Judah, and from the
streets of
Jerusalem, the
voice of
mirth, and the
voice of
gladness, the
voice of the
bridegroom, and the
voice of the
bride: for the
land shall be
desolate.

\par }\Chap{8}{\PP \VerseOne{1}At that
time,
saith the
{\ND{
LORD}}, they shall bring
out the
bones of the
kings of
Judah, and the
bones of his
princes, and the
bones of the
priests, and the
bones of the
prophets, and the
bones of the
inhabitants of
Jerusalem, out of their
graves:
\VS{2}And they shall
spread them before the
sun, and the
moon, and all the
host of
heaven, whom they have
loved, and whom they have
served, and
after whom they have
walked, and whom they have
sought, and whom they have
worshipped: they shall not be
gathered, nor be
buried; they shall be for
dung upon the
face of the
earth.
\VS{3}And
death shall be
chosen rather than
life by all the
residue of them that
remain of this
evil
family, which
remain in all the
places whither I have
driven them,
saith the
{\ND{
LORD}} of
hosts.
\par }{\PP \VS{4}¶ Moreover thou shalt
say unto them, Thus
saith the
{\ND{
LORD}}; Shall they
fall, and not
arise? shall he turn
away, and not
return?
\VS{5}Why
{\ADD{then}} is this
people of
Jerusalem slidden
back by a
perpetual
backsliding? they
hold fast
deceit, they
refuse to
return.
\VS{6}I
hearkened and
heard,
{\ADD{but}} they spake not
aright: no
man
repented him
of his
wickedness,
saying, What have I
done? every one
turned to his
course, as the
horse
rusheth into the
battle.
\VS{7}Yea, the
stork in the
heaven
knoweth her appointed
times; and the
turtle and the
crane and the
swallow
observe the
time of their
coming; but my
people
know not the
judgment of the
{\ND{
LORD}}.
\VS{8}How do ye
say, We
{\ADD{are}}
wise, and the
law of the
{\ND{
LORD}}
{\ADD{is}} with us? Lo,
certainly in
vain
made he
{\ADD{it;}} the
pen of the
scribes
{\ADD{is}} in
vain.\FTNT{*}{{\FR 8.8 }
in vain made…: or, the false pen of the scribes worketh for falsehood}
\VS{9}The
wise
{\ADD{men}} are
ashamed, they are
dismayed and
taken: lo, they have
rejected the
word of the
{\ND{
LORD}}; and what
wisdom
{\ADD{is}} in them?\FTNT{†}{{\FR 8.9 }
The wise…: or, Have they been ashamed, etc}\FTNT{‡}{{\FR 8.9 }
what…: Heb. the wisdom of what thing}
\VS{10}Therefore will I
give their
wives unto
others,
{\ADD{and}} their
fields to them that shall
inherit
{\ADD{them:}} for every one from the
least even unto the
greatest is
given to
covetousness, from the
prophet even unto the
priest every one
dealeth
falsely.
\VS{11}For they have
healed the
hurt of the
daughter of my
people
slightly,
saying,
Peace,
peace; when
{\ADD{there is}} no
peace.
\VS{12}Were they
ashamed when they had
committed
abomination? nay, they were not at
all
ashamed, neither
could they
blush: therefore shall they
fall among them that
fall: in the
time of their
visitation they shall be cast
down,
saith the
{\ND{
LORD}}.\FTNT{§}{{\FR 8.12 }
I will…: or, In gathering I will consume}
\par }{\PP \VS{13}¶ I will
surely
consume them,
saith the
{\ND{
LORD}}:
{\ADD{there shall be}} no
grapes on the
vine, nor
figs on the fig
tree, and the
leaf shall
fade; and
{\ADD{the things that}} I have
given them shall pass
away from them.
\VS{14}Why do we sit
still?
assemble yourselves, and let us
enter into the
defenced
cities, and let us be
silent there: for the
{\ND{
LORD}} our
God hath put us to
silence, and given us
water of
gall to
drink, because we have
sinned against the
{\ND{
LORD}}.\FTNT{**}{{\FR 8.14 }
gall: or, poison}
\VS{15}We
looked for
peace, but no
good
{\ADD{came; and}} for a
time of
health, and behold
trouble!
\VS{16}The
snorting of his
horses was
heard from
Dan: the whole
land
trembled at the
sound of the
neighing of his strong
ones; for they are
come, and have
devoured the
land, and
all that is in it; the
city, and those that
dwell therein.\FTNT{††}{{\FR 8.16 }
all…: Heb. the fulness thereof}
\VS{17}For, behold, I will
send
serpents,
cockatrices, among you, which
{\ADD{will}} not
{\ADD{be}}
charmed, and they shall
bite you,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{18}¶
{\ADD{When}} I would
comfort myself against
sorrow, my
heart
{\ADD{is}}
faint in me.\FTNT{‡‡}{{\FR 8.18 }
in: Heb. upon}
\VS{19}Behold the
voice of the
cry of the
daughter of my
people because of them that dwell in a
far
country:
{\ADD{Is}} not the
{\ND{
LORD}} in
Zion?
{\ADD{is}} not her
king in her? Why have they provoked me to
anger with their graven
images,
{\ADD{and}} with
strange
vanities?\FTNT{§§}{{\FR 8.19 }
them…: Heb. the country of them that are far off}
\VS{20}The
harvest is
past, the
summer is
ended, and we are not
saved.
\VS{21}For the
hurt of the
daughter of my
people am I
hurt; I am
black;
astonishment hath taken
hold on me.
\VS{22}{\ADD{Is there}} no
balm in
Gilead;
{\ADD{is there}} no
physician there? why
then is not the
health of the
daughter of my
people
recovered?\FTNT{***}{{\FR 8.22 }
recovered: Heb. gone up?}

\par }\Chap{9}{\PP \VerseOne{1}Oh
that my
head were
waters, and mine
eyes a
fountain of
tears, that I might
weep
day and
night for the
slain of the
daughter of my
people!\FTNT{*}{{\FR 9.1 }
Oh…: Heb. Who will give my head, etc}
\VS{2}Oh
that I had in the
wilderness a lodging
place of wayfaring
men; that I might
leave my
people, and
go from them! for they
{\ADD{be}} all
adulterers, an
assembly of treacherous
men.
\VS{3}And they
bend their
tongues
{\ADD{like}} their
bow
{\ADD{for}}
lies: but they are not
valiant for the
truth upon the
earth; for they
proceed from
evil to
evil, and they
know not me,
saith the
{\ND{
LORD}}.
\VS{4}Take ye
heed every
one of his
neighbour, and
trust ye not in any
brother: for every
brother will
utterly
supplant, and every
neighbour will
walk with
slanders.\FTNT{†}{{\FR 9.4 }
neighbour: or, friend}
\VS{5}And they will
deceive every
one his
neighbour, and will not
speak the
truth: they have
taught their
tongue to
speak
lies,
{\ADD{and}}
weary themselves to commit
iniquity.\FTNT{‡}{{\FR 9.5 }
deceive: or, mock}
\VS{6}Thine
habitation
{\ADD{is}} in the
midst of
deceit; through
deceit they
refuse to
know me,
saith the
{\ND{
LORD}}.
\VS{7}Therefore thus
saith the
{\ND{
LORD}} of
hosts, Behold, I will
melt them, and
try them; for how shall I
do
for the
daughter of my
people?
\VS{8}Their
tongue
{\ADD{is as}} an
arrow shot
out; it
speaketh
deceit:
{\ADD{one}}
speaketh
peaceably to his
neighbour with his
mouth, but in
heart he
layeth his
wait.\FTNT{§}{{\FR 9.8 }
in heart: Heb. in the midst of him}\FTNT{**}{{\FR 9.8 }
his wait: or, wait for him}
\par }{\PP \VS{9}¶ Shall I not
visit them for these
{\ADD{things}}?
saith the
{\ND{
LORD}}: shall not my
soul be
avenged on such a
nation as this?
\VS{10}For the
mountains will I take
up a
weeping and
wailing, and for the
habitations of the
wilderness a
lamentation, because they are burned
up, so that
none can
pass through
{\ADD{them;}} neither can
{\ADD{men}}
hear the
voice of the
cattle; both the
fowl of the
heavens and the
beast are
fled; they are
gone.\FTNT{††}{{\FR 9.10 }
habitations: or, pastures}\FTNT{‡‡}{{\FR 9.10 }
burned up: or, desolate}\FTNT{§§}{{\FR 9.10 }
both…: Heb. from the fowl even to, etc}
\VS{11}And I will
make
Jerusalem
heaps,
{\ADD{and}} a
den of
dragons; and I will
make the
cities of
Judah
desolate, without an
inhabitant.\FTNT{***}{{\FR 9.11 }
desolate: Heb. desolation}
\par }{\PP \VS{12}¶ Who
{\ADD{is}} the
wise
man, that may
understand this? and
{\ADD{who is he}} to whom the
mouth of the
{\ND{
LORD}} hath
spoken, that he may
declare it, for what the
land
perisheth
{\ADD{and}} is burned
up like a
wilderness, that none passeth
through?
\VS{13}And the
{\ND{
LORD}}
saith, Because they have
forsaken my
law which I
set
before them, and have not
obeyed my
voice, neither
walked therein;
\VS{14}But have
walked
after the
imagination of their own
heart, and
after
Baalim, which their
fathers
taught them:\FTNT{†††}{{\FR 9.14 }
imagination: or, stubbornness}
\VS{15}Therefore thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; Behold, I will
feed them,
{\ADD{even}} this
people, with
wormwood, and give them
water of
gall to
drink.
\VS{16}I will
scatter them also among the
heathen, whom neither they nor their
fathers have
known: and I will
send a
sword
after them, till I have
consumed them.
\par }{\PP \VS{17}¶ Thus
saith the
{\ND{
LORD}} of
hosts,
Consider ye, and
call for the mourning
women, that they may
come; and
send for
cunning
{\ADD{women}}, that they may
come:
\VS{18}And let them make
haste, and take
up a
wailing for us, that our
eyes may run
down with
tears, and our
eyelids gush
out with
waters.
\VS{19}For a
voice of
wailing is
heard out of
Zion, How are we
spoiled! we are
greatly
confounded, because we have
forsaken the
land, because our
dwellings have cast
{\ADD{us}}
out.
\VS{20}Yet
hear the
word of the
{\ND{
LORD}}, O ye
women, and let your
ear
receive the
word of his
mouth, and
teach your
daughters
wailing, and every
one her
neighbour
lamentation.
\VS{21}For
death is come
up into our
windows,
{\ADD{and}} is
entered into our
palaces, to cut
off the
children from
without,
{\ADD{and}} the young
men from the
streets.
\VS{22}Speak, Thus
saith the
{\ND{
LORD}}, Even the
carcases of
men shall
fall as
dung upon the
open
field, and as the
handful
after the
harvestman, and none shall
gather
{\ADD{them}}.
\par }{\PP \VS{23}¶ Thus
saith the
{\ND{
LORD}}, Let not the
wise
{\ADD{man}}
glory in his
wisdom, neither let the
mighty
{\ADD{man}}
glory in his
might, let not the
rich
{\ADD{man}}
glory in his
riches:
\VS{24}But let him that
glorieth
glory in this, that he
understandeth and
knoweth me, that I
{\ADD{am}} the
{\ND{
LORD}} which
exercise
lovingkindness,
judgment, and
righteousness, in the
earth: for in these
{\ADD{things}} I
delight,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{25}¶ Behold, the
days
come,
saith the
{\ND{
LORD}}, that I will
punish all
{\ADD{them which are}}
circumcised with the
uncircumcised;\FTNT{‡‡‡}{{\FR 9.25 }
punish: Heb. visit upon}
\VS{26}Egypt, and
Judah, and
Edom, and the
children of
Ammon, and
Moab, and all
{\ADD{that are}} in the
utmost
corners, that
dwell in the
wilderness: for all
{\ADD{these}}
nations
{\ADD{are}}
uncircumcised, and all the
house of
Israel
{\ADD{are}}
uncircumcised in the
heart.\FTNT{§§§}{{\FR 9.26 }
in the utmost…: Heb. cut off into corners, or, having the corners of their hair polled}

\par }\Chap{10}{\PP \VerseOne{1}Hear ye the
word which the
{\ND{
LORD}}
speaketh unto you, O
house of
Israel:
\VS{2}Thus
saith the
{\ND{
LORD}},
Learn not the
way of the
heathen, and be not
dismayed at the
signs of
heaven; for the
heathen are
dismayed at
them.
\VS{3}For the
customs of the
people
{\ADD{are}}
vain: for
{\ADD{one}}
cutteth a
tree out of the
forest, the
work of the
hands of the
workman, with the
axe.\FTNT{*}{{\FR 10.3 }
customs…: Heb. statutes, or, ordinances are vanity}
\VS{4}They
deck it with
silver and with
gold; they
fasten it with
nails and with
hammers, that it
move not.
\VS{5}They
{\ADD{are}}
upright as the palm
tree, but
speak not: they must
needs be
borne, because they cannot
go. Be not
afraid of them; for they cannot do
evil, neither also
{\ADD{is it}} in them to do
good.
\VS{6}Forasmuch as
{\ADD{there is}} none like unto thee, O
{\ND{
LORD}}; thou
{\ADD{art}}
great, and thy
name
{\ADD{is}}
great in
might.
\VS{7}Who would not
fear thee, O
King of
nations? for to thee doth it
appertain: forasmuch as among all the
wise
{\ADD{men}} of the
nations, and in all their
kingdoms,
{\ADD{there is}} none like unto thee.\FTNT{†}{{\FR 10.7 }
to…: or, it liketh thee}
\VS{8}But they are
altogether
brutish and
foolish: the
stock
{\ADD{is}} a
doctrine of
vanities.\FTNT{‡}{{\FR 10.8 }
altogether: Heb. in one, or, at once}
\VS{9}Silver spread into
plates is
brought from
Tarshish, and
gold from
Uphaz, the
work of the
workman, and of the
hands of the
founder:
blue and
purple
{\ADD{is}} their
clothing: they
{\ADD{are}} all the
work of
cunning
{\ADD{men}}.
\VS{10}But the
{\ND{
LORD}}
{\ADD{is}} the
true
God, he
{\ADD{is}} the
living
God, and an
everlasting
king: at his
wrath the
earth shall
tremble, and the
nations shall not be able to
abide his
indignation.\FTNT{§}{{\FR 10.10 }
true…: Heb. God of truth}\FTNT{**}{{\FR 10.10 }
living God: Heb. living Gods}\FTNT{††}{{\FR 10.10 }
everlasting…: Heb. king of eternity}
\VS{11}Thus shall ye
say unto them, The
gods that have
not
made the
heavens and the
earth,
{\ADD{even}} they shall
perish from the
earth, and from
under
these
heavens.
\VS{12}He hath
made the
earth by his
power, he hath
established the
world by his
wisdom, and hath stretched
out the
heavens by his
discretion.
\VS{13}When he
uttereth his
voice,
{\ADD{there is}} a
multitude of
waters in the
heavens, and he causeth the
vapours to
ascend from the
ends of the
earth; he
maketh
lightnings with
rain, and bringeth
forth the
wind out of his
treasures.\FTNT{‡‡}{{\FR 10.13 }
multitude: or, noise}\FTNT{§§}{{\FR 10.13 }
with: or, for}
\VS{14}Every
man is
brutish in
{\ADD{his}}
knowledge: every
founder is
confounded by the graven
image: for his molten
image
{\ADD{is}}
falsehood, and
{\ADD{there is}} no
breath in them.\FTNT{***}{{\FR 10.14 }
brutish in his knowledge: or, more brutish than to know}
\VS{15}They
{\ADD{are}}
vanity,
{\ADD{and}} the
work of
errors: in the
time of their
visitation they shall
perish.
\VS{16}The
portion of
Jacob
{\ADD{is}} not like them: for he
{\ADD{is}} the
former of all
{\ADD{things;}} and
Israel
{\ADD{is}} the
rod of his
inheritance: The
{\ND{
LORD}} of
hosts
{\ADD{is}} his
name.
\par }{\PP \VS{17}¶ Gather
up thy
wares out of the
land, O
inhabitant of the
fortress.\FTNT{†††}{{\FR 10.17 }
inhabitant: Heb. inhabitress}
\VS{18}For thus
saith the
{\ND{
LORD}}, Behold, I will
sling out the
inhabitants of the
land at this
once, and will
distress them, that they may
find
{\ADD{it so}}.
\par }{\PP \VS{19}¶
Woe is me for my
hurt! my
wound is
grievous: but I
said,
Truly this
{\ADD{is}} a
grief, and I must
bear it.
\VS{20}My
tabernacle is
spoiled, and all my
cords are
broken: my
children are gone
forth of me, and they
{\ADD{are}} not:
{\ADD{there is}} none to stretch
forth my
tent any more, and to set
up my
curtains.
\VS{21}For the
pastors are become
brutish, and have not
sought the
{\ND{
LORD}}: therefore they shall not
prosper, and all their
flocks shall be
scattered.
\VS{22}Behold, the
noise of the
bruit is
come, and a
great
commotion out of the
north
country, to
make the
cities of
Judah
desolate,
{\ADD{and}} a
den of
dragons.
\par }{\PP \VS{23}¶ O
{\ND{
LORD}}, I
know that the
way of
man
{\ADD{is}} not in himself:
{\ADD{it is}} not in
man that
walketh to
direct his
steps.
\VS{24}O
{\ND{
LORD}},
correct me, but with
judgment; not in thine
anger, lest thou bring me to
nothing.\FTNT{‡‡‡}{{\FR 10.24 }
bring…: Heb. diminish me}
\VS{25}Pour
out thy
fury upon the
heathen that
know thee not, and upon the
families that
call not on thy
name: for they have eaten
up
Jacob, and
devoured him, and
consumed him, and have made his
habitation
desolate.

\par }\Chap{11}{\PP \VerseOne{1}The
word that came to
Jeremiah from the
{\ND{
LORD}},
saying,
\VS{2}Hear ye the
words of this
covenant, and
speak unto the
men of
Judah, and to the
inhabitants of
Jerusalem;
\VS{3}And
say thou unto them, Thus
saith the
{\ND{
LORD}}
God of
Israel;
Cursed
{\ADD{be}} the
man that
obeyeth not the
words of this
covenant,
\VS{4}Which I
commanded your
fathers in the
day
{\ADD{that}} I brought them
forth out of the
land of
Egypt, from the
iron
furnace,
saying,
Obey my
voice, and
do them, according to all which I
command you: so shall ye be my
people, and I will be your
God:
\VS{5}That I may
perform the
oath which I have
sworn unto your
fathers, to
give them a
land
flowing with
milk and
honey, as
{\ADD{it is}} this
day. Then
answered I, and
said, So be
it, O
{\ND{
LORD}}.\FTNT{*}{{\FR 11.5 }
So…: Heb. Amen}
\VS{6}Then the
{\ND{
LORD}}
said unto me,
Proclaim all these
words in the
cities of
Judah, and in the
streets of
Jerusalem,
saying,
Hear ye the
words of this
covenant, and
do them.
\VS{7}For I
earnestly
protested unto your
fathers in the
day
{\ADD{that}} I brought them
up out of the
land of
Egypt,
{\ADD{even}} unto this
day, rising
early and
protesting,
saying,
Obey my
voice.
\VS{8}Yet they
obeyed not, nor
inclined their
ear, but
walked every
one in the
imagination of their
evil
heart: therefore I will
bring upon them all the
words of this
covenant, which I
commanded
{\ADD{them}} to
do; but they
did
{\ADD{them}} not.\FTNT{†}{{\FR 11.8 }
imagination: or, stubbornness}
\VS{9}And the
{\ND{
LORD}}
said unto me, A
conspiracy is
found among the
men of
Judah, and among the
inhabitants of
Jerusalem.
\VS{10}They are turned
back to the
iniquities of their
forefathers, which
refused to
hear my
words; and they
went
after
other
gods to
serve them: the
house of
Israel and the
house of
Judah have
broken my
covenant which I
made with their
fathers.
\par }{\PP \VS{11}¶ Therefore thus
saith the
{\ND{
LORD}}, Behold, I will
bring
evil upon them, which they shall not be
able to
escape; and though they shall
cry unto me, I will not
hearken unto them.\FTNT{‡}{{\FR 11.11 }
to escape: Heb. to go forth of}
\VS{12}Then shall the
cities of
Judah and
inhabitants of
Jerusalem
go, and
cry unto the
gods unto whom they offer
incense: but they shall not
save them at
all in the
time of their
trouble.\FTNT{§}{{\FR 11.12 }
trouble: Heb. evil}
\VS{13}For
{\ADD{according to}} the
number of thy
cities were thy
gods, O
Judah; and
{\ADD{according to}} the
number of the
streets of
Jerusalem have ye set
up
altars to
{\ADD{that}} shameful
thing,
{\ADD{even}}
altars to burn
incense unto
Baal.\FTNT{**}{{\FR 11.13 }
shameful…: Heb. shame}
\VS{14}Therefore
pray not thou for this
people, neither lift
up a
cry or
prayer for them: for I will not
hear
{\ADD{them}} in the
time that they
cry unto me for their
trouble.\FTNT{††}{{\FR 11.14 }
trouble: Heb. evil}
\VS{15}What hath my
beloved to do in mine
house,
{\ADD{seeing}} she hath
wrought
lewdness with
many, and the
holy
flesh is
passed from thee? when thou doest
evil, then thou
rejoicest.\FTNT{‡‡}{{\FR 11.15 }
What…: Heb. What is to my beloved in my house}\FTNT{§§}{{\FR 11.15 }
when…: or, when thy evil is}
\VS{16}The
{\ND{
LORD}}
called thy
name, A
green olive
tree,
fair,
{\ADD{and}} of
goodly
fruit: with the
noise of a
great
tumult he hath
kindled
fire upon it, and the
branches of it are
broken.
\VS{17}For the
{\ND{
LORD}} of
hosts, that
planted thee, hath
pronounced
evil against thee,
for the
evil of the
house of
Israel and of the
house of
Judah, which they have
done against themselves to provoke me to
anger in offering
incense unto
Baal.
\par }{\PP \VS{18}¶ And the
{\ND{
LORD}} hath given me
knowledge
{\ADD{of it}}, and I
know
{\ADD{it:}} then thou
shewedst me their
doings.
\VS{19}But I
{\ADD{was}} like a
lamb
{\ADD{or}} an
ox
{\ADD{that}} is
brought to the
slaughter; and I
knew not that they had
devised
devices against me,
{\ADD{saying}}, Let us
destroy the
tree with the
fruit thereof, and let us cut him
off from the
land of the
living, that his
name may be no more
remembered.\FTNT{***}{{\FR 11.19 }
the tree…: Heb. the stalk with his bread}
\VS{20}But, O
{\ND{
LORD}} of
hosts, that
judgest
righteously, that
triest the
reins and the
heart, let me
see thy
vengeance on them: for unto thee have I
revealed my
cause.
\VS{21}Therefore thus
saith the
{\ND{
LORD}} of the
men of
Anathoth, that
seek thy
life,
saying,
Prophesy not in the
name of the
{\ND{
LORD}}, that thou
die not by our
hand:
\VS{22}Therefore thus
saith the
{\ND{
LORD}} of
hosts, Behold, I will
punish them: the young
men shall
die by the
sword; their
sons and their
daughters shall
die by
famine:\FTNT{†††}{{\FR 11.22 }
punish: Heb. visit upon}
\VS{23}And there shall be no
remnant of them: for I will
bring
evil upon the
men of
Anathoth,
{\ADD{even}} the
year of their
visitation.

\par }\Chap{12}{\PP \VerseOne{1}Righteous
{\ADD{art}} thou, O
{\ND{
LORD}}, when I
plead with thee: yet let me
talk with thee of
{\ADD{thy}}
judgments: Wherefore doth the
way of the
wicked
prosper?
{\ADD{wherefore}} are all they
happy that deal
very
treacherously?\FTNT{*}{{\FR 12.1 }
talk…: or, reason the case with thee}
\VS{2}Thou hast
planted them, yea, they have taken
root: they
grow, yea, they bring
forth
fruit: thou
{\ADD{art}}
near in their
mouth, and far
from their
reins.\FTNT{†}{{\FR 12.2 }
they grow: Heb. they go on}
\VS{3}But thou, O
{\ND{
LORD}},
knowest me: thou hast
seen me, and
tried mine
heart toward thee: pull them
out like
sheep for the
slaughter, and
prepare them for the
day of
slaughter.\FTNT{‡}{{\FR 12.3 }
toward…: Heb. with thee}
\VS{4}How long shall the
land
mourn, and the
herbs of every
field
wither, for the
wickedness of them that
dwell therein? the
beasts are
consumed, and the
birds; because they
said, He shall not
see our last
end.
\par }{\PP \VS{5}¶ If thou hast
run with the
footmen, and they have
wearied thee, then how canst thou
contend with
horses? and
{\ADD{if}} in the
land of
peace,
{\ADD{wherein}} thou
trustedst,
{\ADD{they wearied thee}}, then how wilt thou
do in the
swelling of
Jordan?
\VS{6}For even thy
brethren, and the
house of thy
father, even they have dealt
treacherously with thee; yea, they have
called a
multitude
after thee:
believe them not, though they
speak fair
words unto thee.\FTNT{§}{{\FR 12.6 }
they have called…: or, they cried after thee fully}\FTNT{**}{{\FR 12.6 }
fair…: Heb. good things}
\par }{\PP \VS{7}¶ I have
forsaken mine
house, I have
left mine
heritage; I have
given the dearly
beloved of my
soul into the
hand of her
enemies.\FTNT{††}{{\FR 12.7 }
the dearly…: Heb. the love}
\VS{8}Mine
heritage is unto me as a
lion in the
forest; it crieth
out against me: therefore have I
hated it.\FTNT{‡‡}{{\FR 12.8 }
crieth…: or, yelleth: Heb. giveth out his voice}
\VS{9}Mine
heritage
{\ADD{is}} unto me
{\ADD{as}} a
speckled
bird, the
birds round
about
{\ADD{are}} against her;
come ye,
assemble all the
beasts of the
field,
come to
devour.\FTNT{§§}{{\FR 12.9 }
speckled: or, taloned}\FTNT{***}{{\FR 12.9 }
come to: or, cause them to come to}
\VS{10}Many
pastors have
destroyed my
vineyard, they have
trodden my
portion under
foot, they have
made my
pleasant
portion a
desolate
wilderness.\FTNT{†††}{{\FR 12.10 }
pleasant…: Heb. portion of desire}
\VS{11}They have
made it
desolate,
{\ADD{and being}}
desolate it
mourneth unto me; the whole
land is made
desolate, because no
man
layeth
{\ADD{it}} to
heart.
\VS{12}The
spoilers are
come upon all high
places through the
wilderness: for the
sword of the
{\ND{
LORD}} shall
devour from the
{\ADD{one}}
end of the
land even to the
{\ADD{other}}
end of the
land: no
flesh shall have
peace.
\VS{13}They have
sown
wheat, but shall
reap
thorns: they have put themselves to
pain,
{\ADD{but}} shall not
profit: and they shall be
ashamed of your
revenues because of the
fierce
anger of the
{\ND{
LORD}}.\FTNT{‡‡‡}{{\FR 12.13 }
they shall: or, ye shall}
\par }{\PP \VS{14}¶ Thus
saith the
{\ND{
LORD}} against all mine
evil
neighbours, that
touch the
inheritance which I have caused my
people
Israel to
inherit; Behold, I will pluck them
out of their
land, and pluck
out the
house of
Judah from
among them.
\VS{15}And it shall come to pass,
after that I have plucked them
out I will
return, and have
compassion on them, and will bring them
again, every
man to his
heritage, and every
man to his
land.
\VS{16}And it shall come to pass, if they will
diligently
learn the
ways of my
people, to
swear by my
name, The
{\ND{
LORD}}
liveth; as they
taught my
people to
swear by
Baal; then shall they be
built in the
midst of my
people.
\VS{17}But if they will not
obey, I will
utterly pluck
up and
destroy that
nation,
saith the
{\ND{
LORD}}.

\par }\Chap{13}{\PP \VerseOne{1}Thus
saith the
{\ND{
LORD}} unto me,
Go and
get thee a
linen
girdle, and
put it upon thy
loins, and
put it not in
water.
\VS{2}So I
got a
girdle according to the
word of the
{\ND{
LORD}}, and
put
{\ADD{it}} on my
loins.
\VS{3}And the
word of the
{\ND{
LORD}} came unto me the second
time,
saying,
\VS{4}Take the
girdle that thou hast
got, which
{\ADD{is}} upon thy
loins, and
arise,
go to
Euphrates, and
hide it there in a
hole of the
rock.
\VS{5}So I
went, and
hid it by
Euphrates, as the
{\ND{
LORD}}
commanded me.
\VS{6}And it came to pass
after
many
days, that the
{\ND{
LORD}}
said unto me,
Arise,
go to
Euphrates, and
take the
girdle from thence, which I
commanded thee to
hide there.
\VS{7}Then I
went to
Euphrates, and
digged, and
took the
girdle from the
place where I had
hid it: and, behold, the
girdle was
marred, it was
profitable for nothing.
\VS{8}Then the
word of the
{\ND{
LORD}} came unto me,
saying,
\VS{9}Thus
saith the
{\ND{
LORD}}, After this manner will I
mar the
pride of
Judah, and the
great
pride of
Jerusalem.
\VS{10}This
evil
people, which
refuse to
hear my
words, which
walk in the
imagination of their
heart, and
walk
after
other
gods, to
serve them, and to
worship them, shall even be as this
girdle, which is
good for nothing.\FTNT{*}{{\FR 13.10 }
imagination: or, stubbornness}
\VS{11}For as the
girdle
cleaveth to the
loins of a
man, so have I caused to
cleave unto me the whole
house of
Israel and the whole
house of
Judah,
saith the
{\ND{
LORD}}; that they might be unto me for a
people, and for a
name, and for a
praise, and for a
glory: but they would not
hear.
\par }{\PP \VS{12}¶ Therefore thou shalt
speak unto them this
word; Thus
saith the
{\ND{
LORD}}
God of
Israel, Every
bottle shall be
filled with
wine: and they shall
say unto thee, Do we not
certainly
know that every
bottle shall be
filled with
wine?
\VS{13}Then shalt thou
say unto them, Thus
saith the
{\ND{
LORD}}, Behold, I will
fill all the
inhabitants of this
land, even the
kings that
sit upon
David’s
throne, and the
priests, and the
prophets, and all the
inhabitants of
Jerusalem, with
drunkenness.
\VS{14}And I will
dash them
one against
another, even the
fathers and the
sons
together,
saith the
{\ND{
LORD}}: I will not
pity, nor
spare, nor have
mercy, but
destroy them.\FTNT{†}{{\FR 13.14 }
one…: Heb. a man against his brother}\FTNT{‡}{{\FR 13.14 }
but…: Heb. from destroying them}
\par }{\PP \VS{15}¶
Hear ye, and give
ear; be not
proud: for the
{\ND{
LORD}} hath
spoken.
\VS{16}Give
glory to the
{\ND{
LORD}} your
God, before he cause
darkness, and before your
feet
stumble upon the
dark
mountains, and, while ye
look for
light, he
turn it into the shadow of
death,
{\ADD{and}}
make
{\ADD{it}} gross
darkness.
\VS{17}But if ye will not
hear it, my
soul shall
weep in secret
places
for
{\ADD{your}}
pride; and mine
eye shall
weep
sore, and run
down with
tears, because the
{\ND{
LORD’s}}
flock is carried away
captive.
\VS{18}Say unto the
king and to the
queen,
Humble yourselves, sit
down: for your
principalities shall come
down,
{\ADD{even}} the
crown of your
glory.\FTNT{§}{{\FR 13.18 }
principalities: or, head tires}
\VS{19}The
cities of the
south shall be shut
up, and none shall
open
{\ADD{them:}}
Judah shall be carried away
captive all of it, it shall be
wholly carried away
captive.
\VS{20}Lift
up your
eyes, and
behold them that
come from the
north: where
{\ADD{is}} the
flock
{\ADD{that}} was
given thee, thy
beautiful
flock?
\VS{21}What wilt thou
say when he shall
punish thee? for thou hast
taught them
{\ADD{to be}}
captains,
{\ADD{and}} as
chief over thee: shall not
sorrows
take thee, as a
woman in
travail?\FTNT{**}{{\FR 13.21 }
punish: Heb visit upon}
\par }{\PP \VS{22}¶ And if thou
say in thine
heart, Wherefore
come these things upon me? For the
greatness of thine
iniquity are thy
skirts
discovered,
{\ADD{and}} thy
heels made
bare.\FTNT{††}{{\FR 13.22 }
made…: or, shall be violently taken away}
\VS{23}Can the
Ethiopian
change his
skin, or the
leopard his
spots?
{\ADD{then}}
may ye also do
good, that are
accustomed to do
evil.\FTNT{‡‡}{{\FR 13.23 }
accustomed: Heb. taught}
\VS{24}Therefore will I
scatter them as the
stubble that passeth
away by the
wind of the
wilderness.
\VS{25}This
{\ADD{is}} thy
lot, the
portion of thy
measures from me,
saith the
{\ND{
LORD}}; because thou hast
forgotten me, and
trusted in
falsehood.
\VS{26}Therefore will I
discover thy
skirts upon thy
face, that thy
shame may
appear.
\VS{27}I have
seen thine
adulteries, and thy
neighings, the
lewdness of thy
whoredom,
{\ADD{and}} thine
abominations on the
hills in the
fields.
Woe unto thee, O
Jerusalem! wilt thou not be made
clean? when
{\ADD{shall it}}
once
{\ADD{be}}?\FTNT{§§}{{\FR 13.27 }
when…: Heb. after when yet?}

\par }\Chap{14}{\PP \VerseOne{1}The
word of the
{\ND{
LORD}} that came to
Jeremiah
concerning the
dearth.\FTNT{*}{{\FR 14.1 }
the dearth: Heb. the words of the dearths, or, restraints}
\VS{2}Judah
mourneth, and the
gates thereof
languish; they are
black unto the
ground; and the
cry of
Jerusalem is gone
up.
\VS{3}And their
nobles have
sent their little
ones to the
waters: they
came to the
pits,
{\ADD{and}}
found no
water; they
returned with their
vessels
empty; they were
ashamed and
confounded, and
covered their
heads.
\VS{4}Because the
ground is
chapt, for there was no
rain in the
earth, the
plowmen were
ashamed, they
covered their
heads.
\VS{5}Yea, the
hind also
calved in the
field, and
forsook
{\ADD{it}}, because there was no
grass.
\VS{6}And the wild
asses did
stand in the high
places, they snuffed
up the
wind like
dragons; their
eyes did
fail, because
{\ADD{there was}} no
grass.
\par }{\PP \VS{7}¶ O
{\ND{
LORD}}, though our
iniquities
testify against us,
do thou
{\ADD{it}} for thy
name’s sake: for our
backslidings are
many; we have
sinned against thee.
\VS{8}O the
hope of
Israel, the
saviour thereof in
time of
trouble, why shouldest thou be as a
stranger in the
land, and as a wayfaring
man
{\ADD{that}} turneth
aside to tarry for a
night?
\VS{9}Why shouldest thou be as a
man
astonied, as a mighty
man
{\ADD{that}}
cannot
save? yet thou, O
{\ND{
LORD}},
{\ADD{art}} in the
midst of us, and we are
called by thy
name;
leave us not.\FTNT{†}{{\FR 14.9 }
we…: Heb. thy name is called upon us}
\par }{\PP \VS{10}¶ Thus
saith the
{\ND{
LORD}} unto this
people, Thus have they
loved to
wander, they have not
refrained their
feet, therefore the
{\ND{
LORD}} doth not
accept them; he will now
remember their
iniquity, and
visit their
sins.
\VS{11}Then
said the
{\ND{
LORD}} unto me,
Pray not for this
people for
{\ADD{their}}
good.
\VS{12}When they
fast, I will not
hear their
cry; and when they
offer burnt
offering and an
oblation, I will not
accept them: but I will
consume them by the
sword, and by the
famine, and by the
pestilence.
\par }{\PP \VS{13}¶ Then
said I,
Ah,
Lord
{\ND{
GOD}}! behold, the
prophets
say unto them, Ye shall not
see the
sword, neither shall ye have
famine; but I will
give you
assured
peace in this
place.\FTNT{‡}{{\FR 14.13 }
assured…: Heb. peace of truth}
\VS{14}Then the
{\ND{
LORD}}
said unto me, The
prophets
prophesy
lies in my
name: I
sent them not, neither have I
commanded them, neither
spake unto them: they
prophesy unto you a
false
vision and
divination, and a thing of
nought, and the
deceit of their
heart.
\VS{15}Therefore thus
saith the
{\ND{
LORD}} concerning the
prophets that
prophesy in my
name, and I
sent them not, yet they
say,
Sword and
famine shall not be in this
land; By
sword and
famine shall those
prophets be
consumed.
\VS{16}And the
people to whom they
prophesy shall be cast
out in the
streets of
Jerusalem
because of the
famine and the
sword; and they shall have none to
bury
them, them, their
wives, nor their
sons, nor their
daughters: for I will
pour their
wickedness upon them.
\par }{\PP \VS{17}¶ Therefore thou shalt
say this
word unto them; Let mine
eyes run
down with
tears
night and
day, and let them not
cease: for the
virgin
daughter of my
people is
broken with a
great
breach, with a
very
grievous
blow.
\VS{18}If I go
forth into the
field, then behold the
slain with the
sword! and if I
enter into the
city, then behold them that are
sick with
famine! yea, both the
prophet and the
priest go
about into a
land that they
know not.\FTNT{§}{{\FR 14.18 }
go about…: or, make merchandise against a land, and men acknowledge it not}
\VS{19}Hast thou
utterly
rejected
Judah? hath thy
soul
lothed
Zion? why hast thou
smitten us, and
{\ADD{there is}} no
healing for us? we
looked for
peace, and
{\ADD{there is}} no
good; and for the
time of
healing, and behold
trouble!
\VS{20}We
acknowledge, O
{\ND{
LORD}}, our
wickedness,
{\ADD{and}} the
iniquity of our
fathers: for we have
sinned against thee.
\VS{21}Do not
abhor
{\ADD{us}}, for thy
name’s sake, do not
disgrace the
throne of thy
glory:
remember,
break not thy
covenant with us.
\VS{22}Are
there
{\ADD{any}} among the
vanities of the
Gentiles that can cause
rain? or can the
heavens
give
showers?
{\ADD{art}} not thou he, O
{\ND{
LORD}} our
God? therefore we will
wait upon thee: for thou hast
made all these
{\ADD{things}}.

\par }\Chap{15}{\PP \VerseOne{1}Then
said the
{\ND{
LORD}} unto me, Though
Moses and
Samuel
stood
before me,
{\ADD{yet}} my
mind
{\ADD{could}} not
{\ADD{be}} toward this
people: cast
{\ADD{them}}
out of my
sight, and let them go
forth.
\VS{2}And it shall come to pass, if they
say unto thee, Whither shall we go
forth? then thou shalt
tell them, Thus
saith the
{\ND{
LORD}}; Such as
{\ADD{are}} for
death, to
death; and such as
{\ADD{are}} for the
sword, to the
sword; and such as
{\ADD{are}} for the
famine, to the
famine; and such as
{\ADD{are}} for the
captivity, to the
captivity.
\VS{3}And I will
appoint over them
four
kinds,
saith the
{\ND{
LORD}}: the
sword to
slay, and the
dogs to
tear, and the
fowls of the
heaven, and the
beasts of the
earth, to
devour and
destroy.\FTNT{*}{{\FR 15.3 }
kinds: Heb. families}
\VS{4}And I will
cause them to be
removed into all
kingdoms of the
earth,
because of
Manasseh the
son of
Hezekiah
king of
Judah, for
{\ADD{that}} which he
did in
Jerusalem.\FTNT{†}{{\FR 15.4 }
cause…: Heb. give them for a removing}
\VS{5}For who shall have
pity upon thee, O
Jerusalem? or who shall
bemoan thee? or who shall go
aside to
ask how thou
doest?\FTNT{‡}{{\FR 15.5 }
how…: Heb. of thy peace?}
\VS{6}Thou hast
forsaken me,
saith the
{\ND{
LORD}}, thou art
gone
backward: therefore will I stretch
out my
hand against thee, and
destroy thee; I am
weary with
repenting.
\VS{7}And I will
fan them with a
fan in the
gates of the
land; I will
bereave
{\ADD{them}} of children, I will
destroy my
people,
{\ADD{since}} they
return not from their
ways.\FTNT{§}{{\FR 15.7 }
children: or, whatsoever is dear}
\VS{8}Their
widows are
increased to me above the
sand of the
seas: I have
brought upon them against the
mother of the young
men a
spoiler at
noonday: I have caused
{\ADD{him}} to
fall upon it
suddenly, and
terrors upon the
city.\FTNT{**}{{\FR 15.8 }
the mother…: or, the mother city a young man spoiling, etc, or, the mother and the young men}
\VS{9}She that hath
borne
seven
languisheth: she hath given
up the
ghost; her
sun is
gone down while
{\ADD{it was}} yet
day: she hath been
ashamed and
confounded: and the
residue of them will I
deliver to the
sword
before their
enemies,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{10}¶
Woe is me, my
mother, that thou hast
borne me a
man of
strife and a
man of
contention to the whole
earth! I have neither lent on
usury, nor men have lent to me on
usury;
{\ADD{yet}} every one of them doth
curse me.
\VS{11}The
{\ND{
LORD}}
said,
Verily it shall be
well with thy
remnant;
verily I will cause the
enemy to
entreat thee
{\ADD{well}} in the
time of
evil and in the
time of
affliction.\FTNT{††}{{\FR 15.11 }
cause…: or, intreat the enemy for thee}
\VS{12}Shall
iron
break the
northern
iron and the
steel?
\VS{13}Thy
substance and thy
treasures will I
give to the
spoil without
price, and
{\ADD{that}} for all thy
sins, even in all thy
borders.
\VS{14}And I will make
{\ADD{thee}} to
pass with thine
enemies into a
land
{\ADD{which}} thou
knowest not: for a
fire is
kindled in mine
anger,
{\ADD{which}} shall
burn upon you.
\par }{\PP \VS{15}¶ O
{\ND{
LORD}}, thou
knowest:
remember me, and
visit me, and
revenge me of my
persecutors; take me not
away in thy
longsuffering:
know that for thy sake I have
suffered
rebuke.
\VS{16}Thy
words were
found, and I did
eat them; and thy
word was unto me the
joy and
rejoicing of mine
heart: for I am
called by thy
name, O
{\ND{
LORD}}
God of
hosts.\FTNT{‡‡}{{\FR 15.16 }
I am…: Heb. thy name is called upon me}
\VS{17}I
sat not in the
assembly of the
mockers, nor
rejoiced; I
sat
alone
because of thy
hand: for thou hast
filled me with
indignation.
\VS{18}Why is my
pain
perpetual, and my
wound
incurable,
{\ADD{which}}
refuseth to be
healed? wilt thou be altogether unto me as a
liar,
{\ADD{and as}}
waters
{\ADD{that}}
fail?\FTNT{§§}{{\FR 15.18 }
fail: Heb. be not sure?}
\par }{\PP \VS{19}¶ Therefore thus
saith the
{\ND{
LORD}}, If thou
return, then will I bring thee
again,
{\ADD{and}} thou shalt
stand
before me: and if thou take
forth the
precious from the
vile, thou shalt be as my
mouth: let them
return unto thee; but
return not thou unto them.
\VS{20}And I will
make thee unto this
people a
fenced
brasen
wall: and they shall
fight against thee, but they shall not
prevail against thee: for I
{\ADD{am}} with thee to
save thee and to
deliver thee,
saith the
{\ND{
LORD}}.
\VS{21}And I will
deliver thee out of the
hand of the
wicked, and I will
redeem thee out of the
hand of the
terrible.

\par }\Chap{16}{\PP \VerseOne{1}The
word of the
{\ND{
LORD}} came also unto me,
saying,
\VS{2}Thou shalt not
take thee a
wife, neither shalt thou have
sons or
daughters in this
place.
\VS{3}For thus
saith the
{\ND{
LORD}} concerning the
sons and concerning the
daughters that are
born in this
place, and concerning their
mothers that
bare them, and concerning their
fathers that
begat them in this
land;
\VS{4}They shall
die of
grievous
deaths; they shall not be
lamented; neither shall they be
buried;
{\ADD{but}} they shall be as
dung upon the
face of the
earth: and they shall be
consumed by the
sword, and by
famine; and their
carcases shall be
meat for the
fowls of
heaven, and for the
beasts of the
earth.
\VS{5}For thus
saith the
{\ND{
LORD}},
Enter not into the
house of
mourning, neither
go to
lament nor
bemoan them: for I have taken
away my
peace from this
people,
saith the
{\ND{
LORD}},
{\ADD{even}}
lovingkindness and
mercies.\FTNT{*}{{\FR 16.5 }
mourning: or, mourning feast}
\VS{6}Both the
great and the
small shall
die in this
land: they shall not be
buried, neither shall
{\ADD{men}}
lament for them, nor
cut themselves, nor make themselves
bald for them:
\VS{7}Neither shall
{\ADD{men}}
tear
{\ADD{themselves}} for them in
mourning, to
comfort them for the
dead; neither shall
{\ADD{men}} give them the
cup of
consolation to
drink for their
father or for their
mother.\FTNT{†}{{\FR 16.7 }
tear…: or, break bread for them}
\VS{8}Thou shalt not also
go into the
house of
feasting, to
sit with them to
eat and to
drink.
\VS{9}For thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; Behold, I will cause to
cease out of this
place in your
eyes, and in your
days, the
voice of
mirth, and the
voice of
gladness, the
voice of the
bridegroom, and the
voice of the
bride.
\par }{\PP \VS{10}¶ And it shall come to pass, when thou shalt
shew this
people all these
words, and they shall
say unto thee, Wherefore hath the
{\ND{
LORD}}
pronounced all this
great
evil against us? or what
{\ADD{is}} our
iniquity? or what
{\ADD{is}} our
sin that we have
committed against the
{\ND{
LORD}} our
God?
\VS{11}Then shalt thou
say unto them, Because your
fathers have
forsaken me,
saith the
{\ND{
LORD}}, and have
walked
after
other
gods, and have
served them, and have
worshipped them, and have
forsaken me, and have not
kept my
law;
\VS{12}And ye have
done
worse than your
fathers; for,
behold, ye
walk every
one
after the
imagination of his
evil
heart, that they may not
hearken unto me:\FTNT{‡}{{\FR 16.12 }
imagination: or, stubbornness}
\VS{13}Therefore will I
cast you out of this
land into a
land that ye
know not,
{\ADD{neither}} ye nor your
fathers; and there shall ye
serve
other
gods
day and
night; where I will not
shew you
favour.
\par }{\PP \VS{14}¶ Therefore, behold, the
days
come,
saith the
{\ND{
LORD}}, that it shall no more be
said, The
{\ND{
LORD}}
liveth, that brought
up the
children of
Israel out of the
land of
Egypt;
\VS{15}But, The
{\ND{
LORD}}
liveth, that brought
up the
children of
Israel from the
land of the
north, and from all the
lands whither he had
driven them: and I will bring them
again into their
land that I
gave unto their
fathers.
\par }{\PP \VS{16}¶ Behold, I will
send for
many
fishers,
saith the
{\ND{
LORD}}, and they shall
fish them; and
after will I
send for
many
hunters, and they shall
hunt them from every
mountain, and from every
hill, and out of the
holes of the
rocks.
\VS{17}For mine
eyes
{\ADD{are}} upon all their
ways: they are not
hid from my
face, neither is their
iniquity
hid
from mine
eyes.
\VS{18}And
first I will
recompense their
iniquity and their
sin
double; because they have
defiled my
land, they have
filled mine
inheritance with the
carcases of their
detestable and abominable
things.
\VS{19}O
{\ND{
LORD}}, my
strength, and my
fortress, and my
refuge in the
day of
affliction, the
Gentiles shall
come unto thee from the
ends of the
earth, and shall
say, Surely our
fathers have
inherited
lies,
vanity, and
{\ADD{things}} wherein
{\ADD{there is}} no
profit.
\VS{20}Shall a
man
make
gods unto himself, and they
{\ADD{are}} no
gods?
\VS{21}Therefore, behold, I will this
once cause them to
know, I will cause them to
know mine
hand and my
might; and they shall
know that my
name
{\ADD{is}} The
{\ND{
LORD}}.\FTNT{§}{{\FR 16.21 }
The
{\ND{LORD}}: or,
{\ND{JEHOVAH}}}

\par }\Chap{17}{\PP \VerseOne{1}The
sin of
Judah
{\ADD{is}}
written with a
pen of
iron,
{\ADD{and}} with the
point of a
diamond:
{\ADD{it is}}
graven upon the
table of their
heart, and upon the
horns of your
altars;\FTNT{*}{{\FR 17.1 }
point: Heb. nail}
\VS{2}Whilst their
children
remember their
altars and their
groves by the
green
trees upon the
high
hills.
\VS{3}O my
mountain in the
field, I will
give thy
substance
{\ADD{and}} all thy
treasures to the
spoil,
{\ADD{and}} thy high
places for
sin, throughout all thy
borders.
\VS{4}And thou, even thyself, shalt
discontinue from thine
heritage that I
gave thee; and I will cause thee to
serve thine
enemies in the
land which thou
knowest not: for ye have
kindled a
fire in mine
anger,
{\ADD{which}} shall
burn
for
ever.\FTNT{†}{{\FR 17.4 }
thyself: Heb. in thyself}
\par }{\PP \VS{5}¶ Thus
saith the
{\ND{
LORD}};
Cursed
{\ADD{be}} the
man that
trusteth in
man, and
maketh
flesh his
arm, and whose
heart
departeth from the
{\ND{
LORD}}.
\VS{6}For he shall be like the
heath in the
desert, and shall not
see when
good
cometh; but shall
inhabit the parched
places in the
wilderness,
{\ADD{in}} a
salt
land and not
inhabited.
\VS{7}Blessed
{\ADD{is}} the
man that
trusteth in the
{\ND{
LORD}}, and whose
hope the
{\ND{
LORD}} is.
\VS{8}For he shall be as a
tree
planted by the
waters, and
{\ADD{that}} spreadeth
out her
roots by the
river, and shall not
see when
heat
cometh, but her
leaf shall be
green; and shall not be
careful in the
year of
drought, neither shall
cease from
yielding
fruit.\FTNT{‡}{{\FR 17.8 }
drought: or, restraint}
\par }{\PP \VS{9}¶ The
heart
{\ADD{is}}
deceitful above all
{\ADD{things}}, and desperately
wicked: who can
know it?
\VS{10}I the
{\ND{
LORD}}
search the
heart,
{\ADD{I}}
try the
reins, even to
give every
man according to his
ways,
{\ADD{and}} according to the
fruit of his
doings.
\VS{11}{\ADD{As}} the
partridge
sitteth
{\ADD{on eggs}}, and
hatcheth
{\ADD{them}} not;
{\ADD{so}} he that
getteth
riches, and not by
right, shall
leave them in the
midst of his
days, and at his
end shall be a
fool.\FTNT{§}{{\FR 17.11 }
sitteth…: or, gathereth young which she hath not brought forth}
\par }{\PP \VS{12}¶ A
glorious
high
throne from the
beginning
{\ADD{is}} the
place of our
sanctuary.
\VS{13}O
{\ND{
LORD}}, the
hope of
Israel, all that
forsake thee shall be
ashamed,
{\ADD{and}} they that
depart from
me shall be
written in the
earth, because they have
forsaken the
{\ND{
LORD}}, the
fountain of
living
waters.
\VS{14}Heal me, O
{\ND{
LORD}}, and I shall be
healed;
save me, and I shall be
saved: for thou
{\ADD{art}} my
praise.
\par }{\PP \VS{15}¶ Behold, they
say unto me, Where
{\ADD{is}} the
word of the
{\ND{
LORD}}? let it
come now.
\VS{16}As for me, I have not
hastened from
{\ADD{being}} a
pastor to
follow thee: neither have I
desired the
woeful
day; thou
knowest: that which came
out of my
lips was
{\ADD{right}}
before
thee.\FTNT{**}{{\FR 17.16 }
to…: Heb. after thee}
\VS{17}Be not a
terror unto me: thou
{\ADD{art}} my
hope in the
day of
evil.
\VS{18}Let them be
confounded that
persecute me, but let not me be
confounded: let them be
dismayed, but let not me be
dismayed:
bring upon them the
day of
evil, and
destroy them with
double
destruction.\FTNT{††}{{\FR 17.18 }
destroy…: Heb. break them with a double breach}
\par }{\PP \VS{19}¶ Thus
said the
{\ND{
LORD}} unto me;
Go and
stand in the
gate of the
children of the
people, whereby the
kings of
Judah come
in, and by the which they go
out, and in all the
gates of
Jerusalem;
\VS{20}And
say unto them,
Hear ye the
word of the
{\ND{
LORD}}, ye
kings of
Judah, and all
Judah, and all the
inhabitants of
Jerusalem, that enter
in by these
gates:
\VS{21}Thus
saith the
{\ND{
LORD}}; Take
heed to
yourselves, and
bear no
burden on the
sabbath
day, nor
bring
{\ADD{it}} in by the
gates of
Jerusalem;
\VS{22}Neither carry
forth a
burden out of your
houses on the
sabbath
day, neither
do ye any
work, but
hallow ye the
sabbath
day, as I
commanded your
fathers.
\VS{23}But they
obeyed not, neither
inclined their
ear, but made their
neck
stiff, that they might not
hear, nor
receive
instruction.
\VS{24}And it shall come to pass, if ye
diligently
hearken unto me,
saith the
{\ND{
LORD}}, to
bring in no
burden through the
gates of this
city on the
sabbath
day, but
hallow the
sabbath
day, to
do
no
work therein;
\VS{25}Then shall there
enter into the
gates of this
city
kings and
princes
sitting upon the
throne of
David,
riding in
chariots and on
horses, they, and their
princes, the
men of
Judah, and the
inhabitants of
Jerusalem: and this
city shall
remain for
ever.
\VS{26}And they shall
come from the
cities of
Judah, and from the places
about
Jerusalem, and from the
land of
Benjamin, and from the
plain, and from the
mountains, and from the
south,
bringing burnt
offerings, and
sacrifices, and meat
offerings, and
incense, and
bringing sacrifices of
praise, unto the
house of the
{\ND{
LORD}}.
\VS{27}But if ye will not
hearken unto me to
hallow the
sabbath
day, and not to
bear a
burden, even entering
in at the
gates of
Jerusalem on the
sabbath
day; then will I
kindle a
fire in the
gates thereof, and it shall
devour the
palaces of
Jerusalem, and it shall not be
quenched.

\par }\Chap{18}{\PP \VerseOne{1}The
word which came to
Jeremiah from the
{\ND{
LORD}},
saying,
\VS{2}Arise, and go
down to the
potter’s
house, and there I will cause thee to
hear my
words.
\VS{3}Then I went
down to the
potter’s
house, and,
behold, he
wrought a
work on the
wheels.\FTNT{*}{{\FR 18.3 }
wheels: or, frames, or, seats}
\VS{4}And the
vessel that he
made of
clay was
marred in the
hand of the
potter: so he
made it
again
another
vessel, as
seemed
good to the
potter to
make
{\ADD{it}}.\FTNT{†}{{\FR 18.4 }
of clay…: or, was marred, as clay in the hand of the potter}\FTNT{‡}{{\FR 18.4 }
made it…: Heb. returned and made, etc}
\VS{5}Then the
word of the
{\ND{
LORD}} came to me,
saying,
\VS{6}O
house of
Israel,
cannot I
do with you as this
potter?
saith the
{\ND{
LORD}}. Behold, as the
clay
{\ADD{is}} in the
potter’s
hand, so
{\ADD{are}} ye in mine
hand, O
house of
Israel.
\VS{7}{\ADD{At what}}
instant I shall
speak concerning a
nation, and concerning a
kingdom, to pluck
up, and to pull
down, and to
destroy
{\ADD{it;}}
\VS{8}If that
nation, against whom I have
pronounced,
turn from their
evil, I will
repent of the
evil that I
thought to
do unto them.
\VS{9}And
{\ADD{at what}}
instant I shall
speak concerning a
nation, and concerning a
kingdom, to
build and to
plant
{\ADD{it;}}
\VS{10}If it
do
evil in my
sight, that it
obey not my
voice, then I will
repent of the
good, wherewith I
said I would
benefit them.
\par }{\PP \VS{11}¶ Now therefore go to,
speak to the
men of
Judah, and to the
inhabitants of
Jerusalem,
saying, Thus
saith the
{\ND{
LORD}}; Behold, I
frame
evil against you, and
devise a
device against you:
return ye now every
one from his
evil
way, and make your
ways and your
doings
good.
\VS{12}And they
said, There is no
hope: but we will
walk
after our own
devices, and we will every
one
do the
imagination of his
evil
heart.
\VS{13}Therefore thus
saith the
{\ND{
LORD}};
Ask ye now among the
heathen, who hath
heard such things: the
virgin of
Israel hath
done a
very horrible
thing.
\VS{14}Will
{\ADD{a man}}
leave the
snow of
Lebanon
{\ADD{which cometh}} from the
rock of the
field?
{\ADD{or}} shall the
cold
flowing
waters that come from another
place be
forsaken?\FTNT{§}{{\FR 18.14 }
the snow…: or, my fields for a rock, or for the snow of Lebanon? shall the running waters be forsaken for the strange cold waters?}
\VS{15}Because my
people hath
forgotten me, they have burned
incense to
vanity, and they have caused them to
stumble in their
ways
{\ADD{from}} the
ancient
paths, to
walk in
paths,
{\ADD{in}} a
way not cast
up;
\VS{16}To
make their
land
desolate,
{\ADD{and}} a
perpetual
hissing; every one that
passeth thereby shall be
astonished, and
wag his
head.
\VS{17}I will
scatter them as with an
east
wind
before the
enemy; I will
shew them the
back, and not the
face, in the
day of their
calamity.
\par }{\PP \VS{18}¶ Then
said they,
Come, and let us
devise
devices against
Jeremiah; for the
law shall not
perish from the
priest, nor
counsel from the
wise, nor the
word from the
prophet.
Come, and let us
smite him with the
tongue, and let us not give
heed to any of his
words.\FTNT{**}{{\FR 18.18 }
with…: or, for the tongue}
\VS{19}Give
heed to me, O
{\ND{
LORD}}, and
hearken to the
voice of them that
contend with me.
\VS{20}Shall
evil be
recompensed for
good? for they have
digged a
pit for my
soul.
Remember that I
stood
before thee to
speak
good for them,
{\ADD{and}} to turn
away thy
wrath from them.
\VS{21}Therefore deliver
up their
children to the
famine, and pour
out their
{\ADD{blood}} by the
force of the
sword; and let their
wives be
bereaved of their children, and
{\ADD{be}}
widows; and let their
men be
put to
death;
{\ADD{let}} their young
men
{\ADD{be}}
slain by the
sword in
battle.\FTNT{††}{{\FR 18.21 }
pour…: Heb. pour them out}
\VS{22}Let a
cry be
heard from their
houses, when thou shalt
bring a
troop
suddenly upon them: for they have
digged a
pit to
take me, and
hid
snares for my
feet.
\VS{23}Yet,
{\ND{
LORD}}, thou
knowest all their
counsel against me to
slay
{\ADD{me:}}
forgive not their
iniquity, neither blot
out their
sin from thy
sight, but let them be
overthrown
before thee;
deal
{\ADD{thus}} with them in the
time of thine
anger.\FTNT{‡‡}{{\FR 18.23 }
to slay…: Heb. for death}

\par }\Chap{19}{\PP \VerseOne{1}Thus
saith the
{\ND{
LORD}},
Go and
get a
potter’s
earthen
bottle, and
{\ADD{take}} of the
ancients of the
people, and of the
ancients of the
priests;
\VS{2}And go
forth unto the
valley of the
son of
Hinnom, which
{\ADD{is}} by the
entry of the
east
gate, and
proclaim there the
words that I shall
tell thee,\FTNT{*}{{\FR 19.2 }
the east…: Heb. the sun gate}
\VS{3}And
say,
Hear ye the
word of the
{\ND{
LORD}}, O
kings of
Judah, and
inhabitants of
Jerusalem; Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; Behold, I will
bring
evil upon this
place, the which whosoever
heareth, his
ears shall
tingle.
\VS{4}Because they have
forsaken me, and have
estranged this
place, and have burned
incense in it unto
other
gods, whom neither they nor their
fathers have
known, nor the
kings of
Judah, and have
filled this
place with the
blood of
innocents;
\VS{5}They have
built also the high
places of
Baal, to
burn their
sons with
fire
{\ADD{for}} burnt
offerings unto
Baal, which I
commanded not, nor
spake
{\ADD{it}}, neither
came
{\ADD{it}} into my
mind:
\VS{6}Therefore, behold, the
days
come,
saith the
{\ND{
LORD}}, that this
place shall no more be
called
Tophet, nor The
valley of the
son of
Hinnom, but The
valley of
slaughter.
\VS{7}And I will make
void the
counsel of
Judah and
Jerusalem in this
place; and I will cause them to
fall by the
sword
before their
enemies, and by the
hands of them that
seek their
lives: and their
carcases will I
give to be
meat for the
fowls of the
heaven, and for the
beasts of the
earth.
\VS{8}And I will
make this
city
desolate, and an
hissing; every one that
passeth thereby shall be
astonished and
hiss because of all the
plagues thereof.
\VS{9}And I will cause them to
eat the
flesh of their
sons and the
flesh of their
daughters, and they shall
eat every
one the
flesh of his
friend in the
siege and
straitness, wherewith their
enemies, and they that
seek their
lives, shall
straiten them.
\par }{\PP \VS{10}Then shalt thou
break the
bottle in the
sight of the
men that
go with thee,
\VS{11}And shalt
say unto them, Thus
saith the
{\ND{
LORD}} of
hosts; Even
so will I
break this
people and this
city, as
{\ADD{one}}
breaketh a
potter’s
vessel, that
cannot be made whole
again: and they shall
bury
{\ADD{them}} in
Tophet, till
{\ADD{there be}} no
place to
bury.\FTNT{†}{{\FR 19.11 }
be made…: Heb. be healed}
\VS{12}Thus will I
do unto this
place,
saith the
{\ND{
LORD}}, and to the
inhabitants thereof, and
{\ADD{even}}
make this
city as
Tophet:
\VS{13}And the
houses of
Jerusalem, and the
houses of the
kings of
Judah, shall be
defiled as the
place of
Tophet, because of all the
houses upon whose
roofs they have burned
incense unto all the
host of
heaven, and have poured
out drink
offerings unto
other
gods.
\VS{14}Then
came
Jeremiah from
Tophet, whither the
{\ND{
LORD}} had
sent him to
prophesy; and he
stood in the
court of the
{\ND{
LORD’s}}
house; and
said to all the
people,
\VS{15}Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; Behold, I will
bring upon this
city and upon all her
towns all the
evil that I have
pronounced against it, because they have
hardened their
necks, that they might not
hear my
words.

\par }\Chap{20}{\PP \VerseOne{1}Now
Pashur the
son of
Immer the
priest, who
{\ADD{was}} also
chief
governor in the
house of the
{\ND{
LORD}},
heard that
Jeremiah
prophesied these
things.
\VS{2}Then
Pashur
smote
Jeremiah the
prophet, and
put him in the
stocks that
{\ADD{were}} in the
high
gate of
Benjamin, which
{\ADD{was}} by the
house of the
{\ND{
LORD}}.
\VS{3}And it came to pass on the
morrow, that
Pashur brought
forth
Jeremiah out of the
stocks. Then
said
Jeremiah unto him, The
{\ND{
LORD}} hath not
called thy
name
Pashur, but
Magor-missabib.\FTNT{*}{{\FR 20.3 }
Magor-missabib: that is, Fear round about}
\VS{4}For thus
saith the
{\ND{
LORD}}, Behold, I will
make thee a
terror to thyself, and to all thy
friends: and they shall
fall by the
sword of their
enemies, and thine
eyes shall
behold
{\ADD{it:}} and I will
give all
Judah into the
hand of the
king of
Babylon, and he shall carry them
captive into
Babylon, and shall
slay them with the
sword.
\VS{5}Moreover I will
deliver all the
strength of this
city, and all the
labours thereof, and all the precious
things thereof, and all the
treasures of the
kings of
Judah will I
give into the
hand of their
enemies, which shall
spoil them, and
take them, and
carry them to
Babylon.
\VS{6}And thou,
Pashur, and all that
dwell in thine
house shall
go into
captivity: and thou shalt
come to
Babylon, and there thou shalt
die, and shalt be
buried there, thou, and all thy
friends, to whom thou hast
prophesied
lies.
\par }{\PP \VS{7}¶ O
{\ND{
LORD}}, thou hast
deceived me, and I was
deceived: thou art
stronger than I, and hast
prevailed: I am in
derision
daily, every one
mocketh me.\FTNT{†}{{\FR 20.7 }
was deceived: or, was enticed}
\VS{8}For
since I
spake, I cried
out, I
cried
violence and
spoil; because the
word of the
{\ND{
LORD}} was made a
reproach unto me, and a
derision,
daily.
\VS{9}Then I
said, I will not make
mention of him, nor
speak any more in his
name. But
{\ADD{his word}} was in mine
heart as a
burning
fire shut
up in my
bones, and I was
weary with
forbearing, and I
could not
{\ADD{stay}}.
\par }{\PP \VS{10}¶ For I
heard the
defaming of
many,
fear on every
side.
Report,
{\ADD{say they}}, and we will
report it. All my
familiars
watched for my
halting,
{\ADD{saying}}, Peradventure he will be
enticed, and we shall
prevail against him, and we shall
take our
revenge on him.\FTNT{‡}{{\FR 20.10 }
All…: Heb. Every man of my peace}
\VS{11}But the
{\ND{
LORD}}
{\ADD{is}} with me as a
mighty terrible
one: therefore my
persecutors shall
stumble, and they shall not
prevail: they shall be
greatly
ashamed; for they shall not
prosper:
{\ADD{their}}
everlasting
confusion shall never be
forgotten.
\VS{12}But, O
{\ND{
LORD}} of
hosts, that
triest the
righteous,
{\ADD{and}}
seest the
reins and the
heart, let me
see thy
vengeance on them: for unto thee have I
opened my
cause.
\VS{13}Sing unto the
{\ND{
LORD}},
praise ye the
{\ND{
LORD}}: for he hath
delivered the
soul of the
poor from the
hand of
evildoers.
\par }{\PP \VS{14}¶
Cursed
{\ADD{be}} the
day wherein I was
born: let not the
day wherein my
mother
bare me be
blessed.
\VS{15}Cursed
{\ADD{be}} the
man who brought
tidings to my
father,
saying, A
man
child is
born unto thee; making him
very
glad.
\VS{16}And let that
man be as the
cities which the
{\ND{
LORD}}
overthrew, and
repented not: and let him
hear the
cry in the
morning, and the
shouting at
noontide;
\VS{17}Because he
slew me not from the
womb; or that my
mother might have been my
grave, and her
womb
{\ADD{to be}}
always
great
{\ADD{with me}}.
\VS{18}Wherefore came I
forth out of the
womb to
see
labour and
sorrow, that my
days should be
consumed with
shame?

\par }\Chap{21}{\PP \VerseOne{1}The
word which came unto
Jeremiah from the
{\ND{
LORD}}, when
king
Zedekiah
sent unto him
Pashur the
son of
Melchiah, and
Zephaniah the
son of
Maaseiah the
priest,
saying,
\VS{2}Enquire, I pray thee, of the
{\ND{
LORD}} for
us; for
Nebuchadrezzar
king of
Babylon maketh
war against us; if so be that the
{\ND{
LORD}} will
deal with us according to all his wondrous
works, that he may go
up from us.
\par }{\PP \VS{3}¶ Then
said
Jeremiah unto them, Thus shall ye
say to
Zedekiah:
\VS{4}Thus
saith the
{\ND{
LORD}}
God of
Israel; Behold, I will turn
back the
weapons of
war that
{\ADD{are}} in your
hands, wherewith ye
fight against the
king of
Babylon, and
{\ADD{against}} the
Chaldeans, which
besiege you
without the
walls, and I will
assemble them into the
midst of this
city.
\VS{5}And I myself will
fight against you with an
outstretched
hand and with a
strong
arm, even in
anger, and in
fury, and in
great
wrath.
\VS{6}And I will
smite the
inhabitants of this
city, both
man and
beast: they shall
die of a
great
pestilence.
\VS{7}And
afterward,
saith the
{\ND{
LORD}}, I will
deliver
Zedekiah
king of
Judah, and his
servants, and the
people, and such as are
left in this
city from the
pestilence, from the
sword, and from the
famine, into the
hand of
Nebuchadrezzar
king of
Babylon, and into the
hand of their
enemies, and into the
hand of those that
seek their
life: and he shall
smite them with the
edge of the
sword; he shall not
spare them, neither have
pity, nor have
mercy.
\par }{\PP \VS{8}¶ And unto this
people thou shalt
say, Thus
saith the
{\ND{
LORD}}; Behold, I
set
before you the
way of
life, and the
way of
death.
\VS{9}He that
abideth in this
city shall
die by the
sword, and by the
famine, and by the
pestilence: but he that goeth
out, and
falleth to the
Chaldeans that
besiege you, he shall
live, and his
life shall be unto him for a
prey.
\VS{10}For I have
set my
face against this
city for
evil, and not for
good,
saith the
{\ND{
LORD}}: it shall be
given into the
hand of the
king of
Babylon, and he shall
burn it with
fire.
\par }{\PP \VS{11}¶ And touching the
house of the
king of
Judah,
{\ADD{say}},
Hear ye the
word of the
{\ND{
LORD}};
\VS{12}O
house of
David, thus
saith the
{\ND{
LORD}};
Execute
judgment in the
morning, and
deliver
{\ADD{him that is}}
spoiled out of the
hand of the
oppressor, lest my
fury go
out like
fire, and
burn that none can
quench
{\ADD{it}},
because of the
evil of your
doings.\FTNT{*}{{\FR 21.12 }
Execute: Heb. Judge}
\VS{13}Behold, I
{\ADD{am}} against thee, O
inhabitant of the
valley,
{\ADD{and}}
rock of the
plain,
saith the
{\ND{
LORD}}; which
say, Who shall come
down against us? or who shall
enter into our
habitations?\FTNT{†}{{\FR 21.13 }
inhabitant: Heb. inhabitress}
\VS{14}But I will
punish you according to the
fruit of your
doings,
saith the
{\ND{
LORD}}: and I will
kindle a
fire in the
forest thereof, and it shall
devour all things round
about it.\FTNT{‡}{{\FR 21.14 }
punish: Heb. visit upon}

\par }\Chap{22}{\PP \VerseOne{1}Thus
saith the
{\ND{
LORD}}; Go
down to the
house of the
king of
Judah, and
speak there this
word,
\VS{2}And
say,
Hear the
word of the
{\ND{
LORD}}, O
king of
Judah, that
sittest upon the
throne of
David, thou, and thy
servants, and thy
people that
enter in by these
gates:
\VS{3}Thus
saith the
{\ND{
LORD}};
Execute ye
judgment and
righteousness, and
deliver the
spoiled out of the
hand of the
oppressor: and do no
wrong, do no
violence to the
stranger, the
fatherless, nor the
widow, neither
shed
innocent
blood in this
place.
\VS{4}For if ye
do this
thing
indeed, then shall there enter
in by the
gates of this
house
kings
sitting upon the
throne of
David,
riding in
chariots and on
horses, he, and his
servants, and his
people.\FTNT{*}{{\FR 22.4 }
upon…: Heb. for David upon his throne}
\VS{5}But if ye will not
hear these
words, I
swear by myself,
saith the
{\ND{
LORD}}, that this
house shall become a
desolation.
\VS{6}For thus
saith the
{\ND{
LORD}} unto the
king’s
house of
Judah; Thou
{\ADD{art}}
Gilead unto me,
{\ADD{and}} the
head of
Lebanon:
{\ADD{yet}} surely I will
make thee a
wilderness,
{\ADD{and}}
cities
{\ADD{which}} are not
inhabited.
\VS{7}And I will
prepare
destroyers against thee, every
one with his
weapons: and they shall cut
down thy
choice
cedars, and
cast
{\ADD{them}} into the
fire.
\VS{8}And
many
nations shall
pass by this
city, and they shall
say every
man to his
neighbour, Wherefore hath the
{\ND{
LORD}}
done thus unto this
great
city?
\VS{9}Then they shall
answer, Because they have
forsaken the
covenant of the
{\ND{
LORD}} their
God, and
worshipped
other
gods, and
served them.
\par }{\PP \VS{10}¶
Weep ye not for the
dead, neither
bemoan him:
{\ADD{but}}
weep
sore for him that goeth
away: for he shall
return no more, nor
see his
native
country.
\VS{11}For thus
saith the
{\ND{
LORD}}
touching
Shallum the
son of
Josiah
king of
Judah, which
reigned instead of
Josiah his
father, which went
forth out of this
place; He shall not
return thither any more:
\VS{12}But he shall
die in the
place whither they have led him
captive, and shall
see this
land no more.
\par }{\PP \VS{13}¶
Woe unto him that
buildeth his
house
by
unrighteousness, and his
chambers by
wrong;
{\ADD{that}} useth his
neighbour’s
service without
wages, and
giveth him not for his
work;
\VS{14}That
saith, I will
build me a
wide
house and
large
chambers, and cutteth him
out
windows; and
{\ADD{it is}}
cieled with
cedar, and
painted with
vermilion.\FTNT{†}{{\FR 22.14 }
large: Heb. through-aired}\FTNT{‡}{{\FR 22.14 }
windows: or, my windows}
\VS{15}Shalt thou
reign, because thou
closest
{\ADD{thyself}} in
cedar? did not thy
father
eat and
drink, and
do
judgment and
justice,
{\ADD{and}} then
{\ADD{it was}}
well with him?
\VS{16}He
judged the
cause of the
poor and
needy; then
{\ADD{it was}}
well
{\ADD{with him: was}} not this to
know me?
saith the
{\ND{
LORD}}.
\VS{17}But thine
eyes and thine
heart
{\ADD{are}} not but for thy
covetousness, and for to
shed
innocent
blood, and for
oppression, and for
violence, to
do
{\ADD{it}}.\FTNT{§}{{\FR 22.17 }
violence: or, incursion}
\VS{18}Therefore thus
saith the
{\ND{
LORD}} concerning
Jehoiakim the
son of
Josiah
king of
Judah; They shall not
lament for him,
{\ADD{saying}},
Ah my
brother! or,
Ah
sister! they shall not
lament for him,
{\ADD{saying}},
Ah
lord! or,
Ah his
glory!
\VS{19}He shall be
buried with the
burial of an
ass,
drawn and cast
forth
beyond the
gates of
Jerusalem.
\par }{\PP \VS{20}¶ Go
up to
Lebanon, and
cry; and lift
up thy
voice in
Bashan, and
cry from the
passages: for all thy
lovers are
destroyed.
\VS{21}I
spake unto thee in thy
prosperity;
{\ADD{but}} thou
saidst, I will not
hear. This
{\ADD{hath been}} thy
manner from thy
youth, that thou
obeyedst not my
voice.\FTNT{**}{{\FR 22.21 }
prosperity: Heb. prosperities}
\VS{22}The
wind shall eat
up all thy
pastors, and thy
lovers shall
go into
captivity: surely then shalt thou be
ashamed and
confounded for all thy
wickedness.
\VS{23}O
inhabitant of
Lebanon, that makest thy
nest in the
cedars, how
gracious shalt thou be when
pangs
come upon thee, the
pain as of a woman in
travail!\FTNT{††}{{\FR 22.23 }
inhabitant: Heb. inhabitress}
\VS{24}{\ADD{As}} I
live,
saith the
{\ND{
LORD}}, though
Coniah the
son of
Jehoiakim
king of
Judah were the
signet upon my
right
hand, yet would I
pluck thee thence;
\VS{25}And I will
give thee into the
hand of them that
seek thy
life, and into the
hand
{\ADD{of them}} whose
face thou
fearest, even into the
hand of
Nebuchadrezzar
king of
Babylon, and into the
hand of the
Chaldeans.
\VS{26}And I will cast thee
out, and thy
mother that
bare thee, into
another
country, where ye were not
born; and there shall ye
die.
\VS{27}But to the
land whereunto they
desire to
return, thither shall they not
return.\FTNT{‡‡}{{\FR 22.27 }
desire: Heb. lift up their mind}
\VS{28}{\ADD{Is}} this
man
Coniah a
despised
broken
idol?
{\ADD{is he}} a
vessel wherein
{\ADD{is}} no
pleasure? wherefore are they cast
out, he and his
seed, and are
cast into a
land which they
know not?
\VS{29}O
earth,
earth,
earth,
hear the
word of the
{\ND{
LORD}}.
\VS{30}Thus
saith the
{\ND{
LORD}},
Write ye this
man
childless, a
man
{\ADD{that}} shall not
prosper in his
days: for no
man of his
seed shall
prosper,
sitting upon the
throne of
David, and
ruling any more in
Judah.

\par }\Chap{23}{\PP \VerseOne{1}Woe be unto the
pastors that
destroy and
scatter the
sheep of my
pasture!
saith the
{\ND{
LORD}}.
\VS{2}Therefore thus
saith the
{\ND{
LORD}}
God of
Israel against the
pastors that
feed my
people; Ye have
scattered my
flock, and driven them
away, and have not
visited them: behold, I will
visit upon you the
evil of your
doings,
saith the
{\ND{
LORD}}.
\VS{3}And I will
gather the
remnant of my
flock out of all
countries whither I have
driven them, and will bring them
again to their
folds; and they shall be
fruitful and
increase.
\VS{4}And I will set
up
shepherds over them which shall
feed them: and they shall
fear no more, nor be
dismayed, neither shall they be
lacking,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{5}¶ Behold, the
days
come,
saith the
{\ND{
LORD}}, that I will
raise unto
David a
righteous
Branch, and a
King shall
reign and
prosper, and shall
execute
judgment and
justice in the
earth.
\VS{6}In his
days
Judah shall be
saved, and
Israel shall
dwell
safely: and this
{\ADD{is}} his
name whereby he shall be
called, THE
{\ND{LORD}} OUR
RIGHTEOUSNESS.\FTNT{*}{{\FR 23.6 }
THE
{\ND{LORD}}…: Heb. Jehovah-tsidkenu}
\VS{7}Therefore, behold, the
days
come,
saith the
{\ND{
LORD}}, that they shall no more
say, The
{\ND{
LORD}}
liveth, which brought
up the
children of
Israel out of the
land of
Egypt;
\VS{8}But, The
{\ND{
LORD}}
liveth, which brought
up and which
led the
seed of the
house of
Israel out of the
north
country, and from all
countries whither I had
driven them; and they shall
dwell in their own
land.
\par }{\PP \VS{9}¶ Mine
heart
within me is
broken because of the
prophets; all my
bones
shake; I am like a
drunken
man, and like a
man whom
wine hath
overcome,
because of the
{\ND{
LORD}}, and because of the
words of his
holiness.
\VS{10}For the
land is
full of
adulterers; for
because of
swearing the
land
mourneth; the pleasant
places of the
wilderness are dried
up, and their
course is
evil, and their
force
{\ADD{is}} not right.\FTNT{†}{{\FR 23.10 }
swearing: or, cursing}\FTNT{‡}{{\FR 23.10 }
course: or, violence}
\VS{11}For both
prophet and
priest are
profane; yea, in my
house have I
found their
wickedness,
saith the
{\ND{
LORD}}.
\VS{12}Wherefore their
way shall be unto them as
slippery
{\ADD{ways}} in the
darkness: they shall be driven
on, and
fall therein: for I will
bring
evil upon them,
{\ADD{even}} the
year of their
visitation,
saith the
{\ND{
LORD}}.
\VS{13}And I have
seen
folly in the
prophets of
Samaria; they
prophesied in
Baal, and caused my
people
Israel to
err.\FTNT{§}{{\FR 23.13 }
folly: or, an absurd thing: Heb. unsavoury}
\VS{14}I have
seen also in the
prophets of
Jerusalem an horrible
thing: they commit
adultery, and
walk in
lies: they
strengthen also the
hands of
evildoers, that
none doth
return from his
wickedness: they are all of them unto me as
Sodom, and the
inhabitants thereof as
Gomorrah.\FTNT{**}{{\FR 23.14 }
an…: or, filthiness}
\VS{15}Therefore thus
saith the
{\ND{
LORD}} of
hosts concerning the
prophets; Behold, I will
feed them with
wormwood, and make them
drink the
water of
gall: for from the
prophets of
Jerusalem is
profaneness gone
forth into all the
land.\FTNT{††}{{\FR 23.15 }
profaneness: or, hypocrisy}
\VS{16}Thus
saith the
{\ND{
LORD}} of
hosts,
Hearken not unto the
words of the
prophets that
prophesy unto you: they make you
vain: they
speak a
vision of their own
heart,
{\ADD{and}} not out of the
mouth of the
{\ND{
LORD}}.
\VS{17}They
say
still unto them that
despise me, The
{\ND{
LORD}} hath
said, Ye shall have
peace; and they
say unto every one that
walketh after the
imagination of his own
heart, No
evil shall
come upon you.\FTNT{‡‡}{{\FR 23.17 }
imagination: or, stubbornness}
\VS{18}For who hath
stood in the
counsel of the
{\ND{
LORD}}, and hath
perceived and
heard his
word? who hath
marked his
word, and
heard
{\ADD{it}}?\FTNT{§§}{{\FR 23.18 }
counsel: or, secret}
\VS{19}Behold, a
whirlwind of the
{\ND{
LORD}} is gone
forth in
fury, even a
grievous
whirlwind: it shall fall
grievously upon the
head of the
wicked.
\VS{20}The
anger of the
{\ND{
LORD}} shall not
return, until he have
executed, and till he have
performed the
thoughts of his
heart: in the
latter
days ye shall
consider it
perfectly.
\VS{21}I have not
sent these
prophets, yet they
ran: I have not
spoken to them, yet they
prophesied.
\VS{22}But if they had
stood in my
counsel, and had caused my
people to
hear my
words, then they should have
turned them from their
evil
way, and from the
evil of their
doings.
\VS{23}{\ADD{Am}} I a
God at
hand,
saith the
{\ND{
LORD}}, and not a
God afar
off?
\VS{24}Can
any
hide himself in secret
places that I shall not
see him?
saith the
{\ND{
LORD}}. Do not I
fill
heaven and
earth?
saith the
{\ND{
LORD}}.
\VS{25}I have
heard what the
prophets
said, that
prophesy
lies in my
name,
saying, I have
dreamed, I have
dreamed.
\VS{26}How long shall
{\ADD{this}}
be in the
heart of the
prophets that
prophesy
lies? yea,
{\ADD{they are}}
prophets of the
deceit of their own
heart;
\VS{27}Which
think to cause my
people to
forget my
name by their
dreams which they
tell every
man to his
neighbour, as their
fathers have
forgotten my
name for
Baal.
\VS{28}The
prophet that hath a
dream, let him
tell a
dream; and he that hath my
word, let him
speak my
word
faithfully. What
{\ADD{is}} the
chaff to the
wheat?
saith the
{\ND{
LORD}}.\FTNT{***}{{\FR 23.28 }
that hath a dream: Heb. with whom is, etc}
\VS{29}{\ADD{Is}} not my
word
like as a
fire?
saith the
{\ND{
LORD}}; and like a
hammer
{\ADD{that}}
breaketh the
rock in
pieces?
\VS{30}Therefore, behold, I
{\ADD{am}} against the
prophets,
saith the
{\ND{
LORD}}, that
steal my
words every
one from his
neighbour.
\VS{31}Behold, I
{\ADD{am}} against the
prophets,
saith the
{\ND{
LORD}}, that
use their
tongues, and
say, He
saith.\FTNT{†††}{{\FR 23.31 }
that…: or, that smooth their tongues}
\VS{32}Behold, I
{\ADD{am}} against them that
prophesy
false
dreams,
saith the
{\ND{
LORD}}, and do
tell them, and cause my
people to
err by their
lies, and by their
lightness; yet I
sent them not, nor
commanded them: therefore they shall not
profit this
people at
all,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{33}¶ And when this
people, or the
prophet, or a
priest, shall
ask thee,
saying, What
{\ADD{is}} the
burden of the
{\ND{
LORD}}? thou shalt then
say unto them, What
burden? I will even
forsake you,
saith the
{\ND{
LORD}}.
\VS{34}And
{\ADD{as for}} the
prophet, and the
priest, and the
people, that shall
say, The
burden of the
{\ND{
LORD}}, I will even
punish that
man and his
house.\FTNT{‡‡‡}{{\FR 23.34 }
punish: Heb. visit upon}
\VS{35}Thus shall ye
say every
one to his
neighbour, and every
one to his
brother, What hath the
{\ND{
LORD}}
answered? and, What hath the
{\ND{
LORD}}
spoken?
\VS{36}And the
burden of the
{\ND{
LORD}} shall ye
mention no more: for every
man’s
word shall be his
burden; for ye have
perverted the
words of the
living
God, of the
{\ND{
LORD}} of
hosts our
God.
\VS{37}Thus shalt thou
say to the
prophet, What hath the
{\ND{
LORD}}
answered thee? and, What hath the
{\ND{
LORD}}
spoken?
\VS{38}But since ye
say, The
burden of the
{\ND{
LORD}}; therefore thus
saith the
{\ND{
LORD}}; Because ye
say this
word, The
burden of the
{\ND{
LORD}}, and I have
sent unto you,
saying, Ye shall not
say, The
burden of the
{\ND{
LORD}};
\VS{39}Therefore, behold, I, even I, will
utterly
forget you, and I will
forsake you, and the
city that I
gave you and your
fathers,
{\ADD{and cast you}} out of my
presence:
\VS{40}And I will
bring an
everlasting
reproach upon you, and a
perpetual
shame, which shall not be
forgotten.

\par }\Chap{24}{\PP \VerseOne{1}The
{\ND{
LORD}}
shewed me, and, behold,
two
baskets of
figs
{\ADD{were}}
set
before the
temple of the
{\ND{
LORD}},
after that
Nebuchadrezzar
king of
Babylon had carried away
captive
Jeconiah the
son of
Jehoiakim
king of
Judah, and the
princes of
Judah, with the
carpenters and
smiths, from
Jerusalem, and had
brought them to
Babylon.
\VS{2}One
basket
{\ADD{had}}
very
good
figs,
{\ADD{even}} like the
figs
{\ADD{that are}} first
ripe: and the
other
basket
{\ADD{had}}
very
naughty
figs, which could not be
eaten, they were so
bad.\FTNT{*}{{\FR 24.2 }
they…: Heb. for badness}
\VS{3}Then
said the
{\ND{
LORD}} unto me, What
seest thou,
Jeremiah? And I
said,
Figs; the
good
figs,
very
good; and the
evil,
very
evil, that cannot be
eaten, they are so
evil.
\par }{\PP \VS{4}¶ Again the
word of the
{\ND{
LORD}} came unto me,
saying,
\VS{5}Thus
saith the
{\ND{
LORD}}, the
God of
Israel; Like these
good
figs, so will I
acknowledge them that are carried away
captive of
Judah, whom I have
sent out of this
place into the
land of the
Chaldeans for
{\ADD{their}}
good.\FTNT{†}{{\FR 24.5 }
them…: Heb. the captivity}
\VS{6}For I will
set mine
eyes upon them for
good, and I will bring them
again to this
land: and I will
build them, and not pull
{\ADD{them}}
down; and I will
plant them, and not pluck
{\ADD{them}}
up.
\VS{7}And I will
give them an
heart to
know me, that I
{\ADD{am}} the
{\ND{
LORD}}: and they shall be my
people, and I will be their
God: for they shall
return unto me with their whole
heart.
\par }{\PP \VS{8}¶ And as the
evil
figs, which cannot be
eaten, they are so
evil; surely thus
saith the
{\ND{
LORD}}, So will I
give
Zedekiah the
king of
Judah, and his
princes, and the
residue of
Jerusalem, that
remain in this
land, and them that
dwell in the
land of
Egypt:
\VS{9}And I will
deliver them to be
removed into all the
kingdoms of the
earth for
{\ADD{their}}
hurt,
{\ADD{to be}} a
reproach and a
proverb, a
taunt and a
curse, in all
places whither I shall
drive them.\FTNT{‡}{{\FR 24.9 }
to be removed: Heb. for removing, or, vexation}
\VS{10}And I will
send the
sword, the
famine, and the
pestilence, among them, till they be
consumed from off the
land that I
gave unto them and to their
fathers.

\par }\Chap{25}{\PP \VerseOne{1}The
word that came to
Jeremiah concerning all the
people of
Judah in the
fourth
year of
Jehoiakim the
son of
Josiah
king of
Judah, that
{\ADD{was}} the
first
year of
Nebuchadrezzar
king of
Babylon;
\VS{2}The which
Jeremiah the
prophet
spake unto all the
people of
Judah, and to all the
inhabitants of
Jerusalem,
saying,
\VS{3}From the
thirteenth
year of
Josiah the
son of
Amon
king of
Judah, even unto this
day, that
{\ADD{is}} the
three and
twentieth
year, the
word of the
{\ND{
LORD}} hath come unto me, and I have
spoken unto you, rising
early and
speaking; but ye have not
hearkened.
\VS{4}And the
{\ND{
LORD}} hath
sent unto you all his
servants the
prophets, rising
early and
sending
{\ADD{them;}} but ye have not
hearkened, nor
inclined your
ear to
hear.
\VS{5}They
said, Turn ye
again now every
one from his
evil
way, and from the
evil of your
doings, and
dwell in the
land that the
{\ND{
LORD}} hath
given unto you and to your
fathers
for
ever and
ever:
\VS{6}And
go not
after
other
gods to
serve them, and to
worship them, and provoke me not to
anger with the
works of your
hands; and I will do you no
hurt.
\VS{7}Yet ye have not
hearkened unto me,
saith the
{\ND{
LORD}}; that ye might provoke me to
anger with the
works of your
hands to your own
hurt.
\par }{\PP \VS{8}¶ Therefore thus
saith the
{\ND{
LORD}} of
hosts; Because ye have not
heard my
words,
\VS{9}Behold, I will
send and
take all the
families of the
north,
saith the
{\ND{
LORD}}, and
Nebuchadrezzar the
king of
Babylon, my
servant, and will
bring them against this
land, and against the
inhabitants thereof, and against all these
nations round
about, and will utterly
destroy them, and
make them an
astonishment, and an
hissing, and
perpetual
desolations.
\VS{10}Moreover I will
take from them the
voice of
mirth, and the
voice of
gladness, the
voice of the
bridegroom, and the
voice of the
bride, the
sound of the
millstones, and the
light of the
candle.\FTNT{*}{{\FR 25.10 }
I will…: Heb. I will cause to perish from them}
\VS{11}And this whole
land shall be a
desolation,
{\ADD{and}} an
astonishment; and these
nations shall
serve the
king of
Babylon
seventy
years.
\par }{\PP \VS{12}¶ And it shall come to pass, when
seventy
years are
accomplished,
{\ADD{that}} I will
punish the
king of
Babylon, and that
nation,
saith the
{\ND{
LORD}}, for their
iniquity, and the
land of the
Chaldeans, and will
make it
perpetual
desolations.\FTNT{†}{{\FR 25.12 }
punish: Heb. visit upon}
\VS{13}And I will
bring upon that
land all my
words which I have
pronounced against it,
{\ADD{even}} all that is
written in this
book, which
Jeremiah hath
prophesied against all the
nations.
\VS{14}For
many
nations and
great
kings shall
serve themselves of them also: and I will
recompense them according to their
deeds, and according to the
works of their own
hands.
\par }{\PP \VS{15}¶ For thus
saith the
{\ND{
LORD}}
God of
Israel unto me;
Take the
wine
cup of this
fury at my
hand, and cause all the
nations, to whom I
send thee, to
drink it.
\VS{16}And they shall
drink, and be
moved, and be
mad,
because of the
sword that I will
send among them.
\VS{17}Then
took I the
cup at the
{\ND{
LORD’s}}
hand, and made all the
nations to
drink, unto whom the
{\ND{
LORD}} had
sent me:
\VS{18}{\ADD{To wit}},
Jerusalem, and the
cities of
Judah, and the
kings thereof, and the
princes thereof, to
make them a
desolation, an
astonishment, an
hissing, and a
curse; as
{\ADD{it is}} this
day;
\VS{19}Pharaoh
king of
Egypt, and his
servants, and his
princes, and all his
people;
\VS{20}And all the
mingled people, and all the
kings of the
land of
Uz, and all the
kings of the
land of the
Philistines, and
Ashkelon, and
Azzah, and
Ekron, and the
remnant of
Ashdod,
\VS{21}Edom, and
Moab, and the
children of
Ammon,
\VS{22}And all the
kings of
Tyrus, and all the
kings of
Zidon, and the
kings of the
isles which
{\ADD{are}}
beyond the
sea,\FTNT{‡}{{\FR 25.22 }
isles: or, region by the sea side}
\VS{23}Dedan, and
Tema, and
Buz, and all
{\ADD{that are}} in the
utmost
corners,\FTNT{§}{{\FR 25.23 }
that…: Heb. cut off into corners, or, having the corners of the hair polled}
\VS{24}And all the
kings of
Arabia, and all the
kings of the mingled
people that
dwell in the
desert,
\VS{25}And all the
kings of
Zimri, and all the
kings of
Elam, and all the
kings of the
Medes,
\VS{26}And all the
kings of the
north,
far and
near,
one with
another, and all the
kingdoms of the
world, which
{\ADD{are}} upon the
face of the
earth: and the
king of
Sheshach shall
drink
after them.
\VS{27}Therefore thou shalt
say unto them, Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel;
Drink ye, and be
drunken, and
spue, and
fall, and
rise no more,
because of the
sword which I will
send among you.
\VS{28}And it shall be, if they
refuse to
take the
cup at thine
hand to
drink, then shalt thou
say unto them, Thus
saith the
{\ND{
LORD}} of
hosts; Ye shall
certainly
drink.
\VS{29}For, lo, I
begin to bring
evil on the
city which is
called by my
name, and should ye be
utterly
unpunished? Ye shall not be
unpunished: for I will
call for a
sword upon all the
inhabitants of the
earth,
saith the
{\ND{
LORD}} of
hosts.\FTNT{**}{{\FR 25.29 }
which…: Heb. upon which my name is called}
\par }{\PP \VS{30}Therefore
prophesy thou against them all these
words, and
say unto them, The
{\ND{
LORD}} shall
roar from on
high, and
utter his
voice from his
holy
habitation; he shall
mightily
roar upon his
habitation; he shall
give a
shout, as they that
tread
{\ADD{the grapes}}, against all the
inhabitants of the
earth.
\VS{31}A
noise shall
come
{\ADD{even}} to the
ends of the
earth; for the
{\ND{
LORD}} hath a
controversy with the
nations, he will
plead with all
flesh; he will
give them
{\ADD{that are}}
wicked to the
sword,
saith the
{\ND{
LORD}}.
\VS{32}Thus
saith the
{\ND{
LORD}} of
hosts, Behold,
evil shall go
forth from
nation to
nation, and a
great
whirlwind shall be raised
up from the
coasts of the
earth.
\VS{33}And the
slain of the
{\ND{
LORD}} shall be at that
day from
{\ADD{one}}
end of the
earth even unto the
{\ADD{other}}
end of the
earth: they shall not be
lamented, neither
gathered, nor
buried; they shall be
dung
upon the
ground.
\par }{\PP \VS{34}¶
Howl, ye
shepherds, and
cry; and
wallow yourselves
{\ADD{in the ashes}}, ye
principal of the
flock: for the
days of your
slaughter and of your
dispersions are
accomplished; and ye shall
fall like a
pleasant
vessel.\FTNT{††}{{\FR 25.34 }
the days…: Heb. your days for slaughter}\FTNT{‡‡}{{\FR 25.34 }
a pleasant…: Heb. a vessel of desire}
\VS{35}And the
shepherds shall have no
way to
flee, nor the
principal of the
flock to
escape.\FTNT{§§}{{\FR 25.35 }
the shepherds…: Heb. flight shall perish from the shepherds, and escaping from, etc}
\VS{36}A
voice of the
cry of the
shepherds, and an
howling of the
principal of the
flock,
{\ADD{shall be heard:}} for the
{\ND{
LORD}} hath
spoiled their
pasture.
\VS{37}And the
peaceable
habitations are cut
down
because of the
fierce
anger of the
{\ND{
LORD}}.
\VS{38}He hath
forsaken his
covert, as the
lion: for their
land is
desolate
because of the
fierceness of the
oppressor, and
because of his
fierce
anger.\FTNT{***}{{\FR 25.38 }
desolate: Heb. a desolation}

\par }\Chap{26}{\PP \VerseOne{1}In the
beginning of the
reign of
Jehoiakim the
son of
Josiah
king of
Judah came this
word from the
{\ND{
LORD}},
saying,
\VS{2}Thus
saith the
{\ND{
LORD}};
Stand in the
court of the
{\ND{
LORD’s}}
house, and
speak unto all the
cities of
Judah, which
come to
worship in the
{\ND{
LORD’s}}
house, all the
words that I
command thee to
speak unto them;
diminish not a
word:
\VS{3}If so be they will
hearken, and
turn every
man from his
evil
way, that I may
repent me of the
evil, which I
purpose to
do unto them
because of the
evil of their
doings.
\VS{4}And thou shalt
say unto them, Thus
saith the
{\ND{
LORD}}; If ye will not
hearken to me, to
walk in my
law, which I have
set
before you,
\VS{5}To
hearken to the
words of my
servants the
prophets, whom I
sent unto you, both rising up
early, and
sending
{\ADD{them}}, but ye have not
hearkened;
\VS{6}Then will I
make this
house like
Shiloh, and will
make this
city a
curse to all the
nations of the
earth.
\par }{\PP \VS{7}So the
priests and the
prophets and all the
people
heard
Jeremiah
speaking these
words in the
house of the
{\ND{
LORD}}.
\par }{\PP \VS{8}¶ Now it came to pass, when
Jeremiah had made an
end of
speaking all that the
{\ND{
LORD}} had
commanded
{\ADD{him}} to
speak unto all the
people, that the
priests and the
prophets and all the
people
took him,
saying, Thou shalt
surely
die.
\VS{9}Why hast thou
prophesied in the
name of the
{\ND{
LORD}},
saying, This
house shall be like
Shiloh, and this
city shall be
desolate without an
inhabitant? And all the
people were
gathered against
Jeremiah in the
house of the
{\ND{
LORD}}.
\par }{\PP \VS{10}¶ When the
princes of
Judah
heard these
things, then they came
up from the
king’s
house unto the
house of the
{\ND{
LORD}}, and sat
down in the
entry of the
new
gate of the
{\ND{
LORD’s}}
{\ADD{house}}.\FTNT{*}{{\FR 26.10 }
in the…: or, at the door}
\VS{11}Then
spake the
priests and the
prophets unto the
princes and to all the
people,
saying, This
man
{\ADD{is}}
worthy to
die; for he hath
prophesied against this
city, as ye have
heard with your
ears.\FTNT{†}{{\FR 26.11 }
This man…: Heb. The judgment of death is for this man}
\par }{\PP \VS{12}¶ Then
spake
Jeremiah unto all the
princes and to all the
people,
saying, The
{\ND{
LORD}}
sent me to
prophesy against this
house and against this
city all the
words that ye have
heard.
\VS{13}Therefore now
amend your
ways and your
doings, and
obey the
voice of the
{\ND{
LORD}} your
God; and the
{\ND{
LORD}} will
repent him of the
evil that he hath
pronounced against you.
\VS{14}As for me, behold, I
{\ADD{am}} in your
hand:
do with me as
seemeth
good and
meet unto you.\FTNT{‡}{{\FR 26.14 }
as…: Heb. as it is good and right in your eyes}
\VS{15}But
know ye for
certain, that if ye put me to
death, ye shall surely
bring
innocent
blood upon yourselves, and upon this
city, and upon the
inhabitants thereof: for of a
truth the
{\ND{
LORD}} hath
sent me unto you to
speak all these
words in your
ears.
\par }{\PP \VS{16}¶ Then
said the
princes and all the
people unto the
priests and to the
prophets; This
man
{\ADD{is}} not
worthy to
die: for he hath
spoken to us in the
name of the
{\ND{
LORD}} our
God.
\VS{17}Then rose
up
certain of the
elders of the
land, and
spake to all the
assembly of the
people,
saying,
\VS{18}Micah the
Morasthite
prophesied in the
days of
Hezekiah
king of
Judah, and
spake to all the
people of
Judah,
saying, Thus
saith the
{\ND{
LORD}} of
hosts;
Zion shall be
plowed
{\ADD{like}} a
field, and
Jerusalem shall become
heaps, and the
mountain of the
house as the high
places of a
forest.
\VS{19}Did
Hezekiah
king of
Judah and all
Judah put him at
all to
death? did he not
fear the
{\ND{
LORD}}, and
besought the
{\ND{
LORD}}, and the
{\ND{
LORD}}
repented him of the
evil which he had
pronounced against them?
Thus might we
procure
great
evil against our
souls.\FTNT{§}{{\FR 26.19 }
the
{\ND{LORD}}, and the: Heb. the face of the
{\ND{LORD}}, etc}
\VS{20}And there was also a
man that
prophesied in the
name of the
{\ND{
LORD}},
Urijah the
son of
Shemaiah of
Kirjath-jearim, who
prophesied against this
city and against this
land according to all the
words of
Jeremiah:
\VS{21}And when
Jehoiakim the
king, with all his mighty
men, and all the
princes,
heard his
words, the
king
sought to put him to
death: but when
Urijah
heard it, he was
afraid, and
fled, and
went into
Egypt;
\VS{22}And
Jehoiakim the
king
sent
men into
Egypt,
{\ADD{namely}},
Elnathan the
son of
Achbor, and
{\ADD{certain}}
men with him into
Egypt.
\VS{23}And they fetched
forth
Urijah out of
Egypt, and
brought him unto
Jehoiakim the
king; who
slew him with the
sword, and
cast his dead
body into the
graves of the
common
people.\FTNT{**}{{\FR 26.23 }
common…: Heb. sons of the people}
\VS{24}Nevertheless the
hand of
Ahikam the
son of
Shaphan was with
Jeremiah, that they should not
give him into the
hand of the
people to put him to
death.

\par }\Chap{27}{\PP \VerseOne{1}In the
beginning of the
reign of
Jehoiakim the
son of
Josiah
king of
Judah came this
word unto
Jeremiah from the
{\ND{
LORD}},
saying,
\VS{2}Thus
saith the
{\ND{
LORD}} to me;
Make thee
bonds and
yokes, and
put them upon thy
neck,\FTNT{*}{{\FR 27.2 }
saith…: or, hath the
{\ND{LORD}} said}
\VS{3}And
send them to the
king of
Edom, and to the
king of
Moab, and to the
king of the
Ammonites, and to the
king of
Tyrus, and to the
king of
Zidon, by the
hand of the
messengers which
come to
Jerusalem unto
Zedekiah
king of
Judah;
\VS{4}And
command them to
say unto their
masters, Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; Thus shall ye
say unto your
masters;\FTNT{†}{{\FR 27.4 }
to say…: or, concerning their masters, saying}
\VS{5}I have
made the
earth, the
man and the
beast that
{\ADD{are}}
upon the
ground, by my
great
power and by my
outstretched
arm, and have
given it unto whom it
seemed
meet unto me.
\VS{6}And now have I
given all these
lands into the
hand of
Nebuchadnezzar the
king of
Babylon, my
servant; and the
beasts of the
field have I
given him also to
serve him.
\VS{7}And all
nations shall
serve him, and his
son, and his
son’s
son, until the very
time of his
land
come: and then
many
nations and
great
kings shall
serve themselves of him.
\VS{8}And it shall come to pass,
{\ADD{that}} the
nation and
kingdom which will not
serve the same
Nebuchadnezzar the
king of
Babylon, and that will not
put their
neck under the
yoke of the
king of
Babylon, that
nation will I
punish,
saith the
{\ND{
LORD}}, with the
sword, and with the
famine, and with the
pestilence, until I have
consumed them by his
hand.
\VS{9}Therefore
hearken not ye to your
prophets, nor to your
diviners, nor to your
dreamers, nor to your
enchanters, nor to your
sorcerers, which
speak unto you,
saying, Ye shall not
serve the
king of
Babylon:\FTNT{‡}{{\FR 27.9 }
dreamers: Heb. dreams}
\VS{10}For they
prophesy a
lie unto you, to remove you
far from your
land; and that I should drive you
out, and ye should
perish.
\VS{11}But the
nations that
bring their
neck under the
yoke of the
king of
Babylon, and
serve him, those will I let remain
still in their own
land,
saith the
{\ND{
LORD}}; and they shall
till it, and
dwell therein.
\par }{\PP \VS{12}¶ I
spake also to
Zedekiah
king of
Judah according to all these
words,
saying,
Bring your
necks under the
yoke of the
king of
Babylon, and
serve him and his
people, and
live.
\VS{13}Why will ye
die, thou and thy
people, by the
sword, by the
famine, and by the
pestilence, as the
{\ND{
LORD}} hath
spoken against the
nation that will not
serve the
king of
Babylon?
\VS{14}Therefore
hearken not unto the
words of the
prophets that
speak unto you,
saying, Ye shall not
serve the
king of
Babylon: for they
prophesy a
lie unto you.
\VS{15}For I have not
sent them,
saith the
{\ND{
LORD}}, yet they
prophesy a
lie in my
name; that I might drive you
out, and that ye might
perish, ye, and the
prophets that
prophesy unto you.\FTNT{§}{{\FR 27.15 }
a lie: Heb. in a lie, or, lyingly}
\VS{16}Also I
spake to the
priests and to all this
people,
saying, Thus
saith the
{\ND{
LORD}};
Hearken not to the
words of your
prophets that
prophesy unto you,
saying, Behold, the
vessels of the
{\ND{
LORD’s}}
house shall now
shortly be brought
again from
Babylon: for they
prophesy a
lie unto you.
\VS{17}Hearken not unto them;
serve the
king of
Babylon, and
live: wherefore should this
city be laid
waste?
\VS{18}But if they
{\ADD{be}}
prophets, and if the
word of the
{\ND{
LORD}}
be with them, let them now make
intercession to the
{\ND{
LORD}} of
hosts, that the
vessels which are
left in the
house of the
{\ND{
LORD}}, and
{\ADD{in}} the
house of the
king of
Judah, and at
Jerusalem,
go not to
Babylon.
\par }{\PP \VS{19}¶ For thus
saith the
{\ND{
LORD}} of
hosts concerning the
pillars, and concerning the
sea, and concerning the
bases, and concerning the
residue of the
vessels that
remain in this
city,
\VS{20}Which
Nebuchadnezzar
king of
Babylon
took not, when he carried away
captive
Jeconiah the
son of
Jehoiakim
king of
Judah from
Jerusalem to
Babylon, and all the
nobles of
Judah and
Jerusalem;
\VS{21}Yea, thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel, concerning the
vessels that
remain
{\ADD{in}} the
house of the
{\ND{
LORD}}, and
{\ADD{in}} the
house of the
king of
Judah and of
Jerusalem;
\VS{22}They shall be
carried to
Babylon, and there shall they be until the
day that I
visit them,
saith the
{\ND{
LORD}}; then will I
bring them up, and
restore them to this
place.

\par }\Chap{28}{\PP \VerseOne{1}And it came to pass the same
year, in the
beginning of the
reign of
Zedekiah
king of
Judah, in the
fourth
year,
{\ADD{and}} in the
fifth
month,
{\ADD{that}}
Hananiah the
son of
Azur the
prophet, which
{\ADD{was}} of
Gibeon,
spake unto me in the
house of the
{\ND{
LORD}}, in the
presence of the
priests and of all the
people,
saying,
\VS{2}Thus
speaketh the
{\ND{
LORD}} of
hosts, the
God of
Israel,
saying, I have
broken the
yoke of the
king of
Babylon.
\VS{3}Within two
full
years will I bring
again into this
place all the
vessels of the
{\ND{
LORD’s}}
house, that
Nebuchadnezzar
king of
Babylon took
away from this
place, and
carried them to
Babylon:\FTNT{*}{{\FR 28.3 }
two…: Heb. two years of days}
\VS{4}And I will bring
again to this
place
Jeconiah the
son of
Jehoiakim
king of
Judah, with all the
captives of
Judah, that
went into
Babylon,
saith the
{\ND{
LORD}}: for I will
break the
yoke of the
king of
Babylon.\FTNT{†}{{\FR 28.4 }
captives: Heb. captivity}
\par }{\PP \VS{5}¶ Then the
prophet
Jeremiah
said unto the
prophet
Hananiah in the
presence of the
priests, and in the
presence of all the
people that
stood in the
house of the
{\ND{
LORD}},
\VS{6}Even the
prophet
Jeremiah
said,
Amen: the
{\ND{
LORD}}
do so: the
{\ND{
LORD}}
perform thy
words which thou hast
prophesied, to bring
again the
vessels of the
{\ND{
LORD’s}}
house, and all that is carried away
captive, from
Babylon into this
place.
\VS{7}Nevertheless
hear thou now this
word that I
speak in thine
ears, and in the
ears of all the
people;
\VS{8}The
prophets that have been
before me and
before thee of
old
prophesied both against
many
countries, and against
great
kingdoms, of
war, and of
evil, and of
pestilence.
\VS{9}The
prophet which
prophesieth of
peace, when the
word of the
prophet shall come to
pass,
{\ADD{
then}} shall the
prophet be
known, that the
{\ND{
LORD}} hath
truly
sent him.
\par }{\PP \VS{10}¶ Then
Hananiah the
prophet
took the
yoke from off the
prophet
Jeremiah’s
neck, and
brake it.
\VS{11}And
Hananiah
spake in the
presence of all the
people,
saying, Thus
saith the
{\ND{
LORD}}; Even so will I
break the
yoke of
Nebuchadnezzar
king of
Babylon from the
neck of all
nations within the space of two
full
years. And the
prophet
Jeremiah
went his
way.
\par }{\PP \VS{12}¶ Then the
word of the
{\ND{
LORD}} came unto
Jeremiah
{\ADD{the prophet}},
after that
Hananiah the
prophet had
broken the
yoke from off the
neck of the
prophet
Jeremiah,
saying,
\VS{13}Go and
tell
Hananiah,
saying, Thus
saith the
{\ND{
LORD}}; Thou hast
broken the
yokes of
wood; but thou shalt
make for them
yokes of
iron.
\VS{14}For thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; I have
put a
yoke of
iron upon the
neck of all these
nations, that they may
serve
Nebuchadnezzar
king of
Babylon; and they shall
serve him: and I have
given him the
beasts of the
field also.
\par }{\PP \VS{15}¶ Then
said the
prophet
Jeremiah unto
Hananiah the
prophet,
Hear now,
Hananiah; The
{\ND{
LORD}} hath not
sent thee; but thou makest this
people to
trust in a
lie.
\VS{16}Therefore thus
saith the
{\ND{
LORD}}; Behold, I will
cast thee from off the
face of the
earth: this
year thou shalt
die, because thou hast
taught
rebellion against the
{\ND{
LORD}}.\FTNT{‡}{{\FR 28.16 }
rebellion: Heb. revolt}
\VS{17}So
Hananiah the
prophet
died the same
year in the
seventh
month.

\par }\Chap{29}{\PP \VerseOne{1}Now these
{\ADD{are}} the
words of the
letter that
Jeremiah the
prophet
sent from
Jerusalem unto the
residue of the
elders which were carried away
captives, and to the
priests, and to the
prophets, and to all the
people whom
Nebuchadnezzar had carried away
captive from
Jerusalem to
Babylon;
\VS{2}(After that
Jeconiah the
king, and the
queen, and the
eunuchs, the
princes of
Judah and
Jerusalem, and the
carpenters, and the
smiths, were
departed from
Jerusalem;)\FTNT{*}{{\FR 29.2 }
eunuchs: or, chamberlains}
\VS{3}By the
hand of
Elasah the
son of
Shaphan, and
Gemariah the
son of
Hilkiah, (whom
Zedekiah
king of
Judah
sent unto
Babylon to
Nebuchadnezzar
king of
Babylon)
saying,
\VS{4}Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel, unto all that are carried away
captives, whom I have caused to be carried
away from
Jerusalem unto
Babylon;
\VS{5}Build ye
houses, and
dwell
{\ADD{in them;}} and
plant
gardens, and
eat the
fruit of them;
\VS{6}Take ye
wives, and
beget
sons and
daughters; and
take
wives for your
sons, and
give your
daughters to
husbands, that they may
bear
sons and
daughters; that ye may be
increased there, and not
diminished.
\VS{7}And
seek the
peace of the
city whither I have caused you to be carried away
captives, and
pray unto the
{\ND{
LORD}} for it: for in the
peace thereof shall ye have
peace.
\par }{\PP \VS{8}¶ For thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; Let not your
prophets and your
diviners, that
{\ADD{be}} in the
midst of you,
deceive you, neither
hearken to your
dreams which ye cause to be
dreamed.
\VS{9}For they
prophesy
falsely unto you in my
name: I have not
sent them,
saith the
{\ND{
LORD}}.\FTNT{†}{{\FR 29.9 }
falsely: Heb. in a lie}
\par }{\PP \VS{10}¶ For thus
saith the
{\ND{
LORD}}, That
after
seventy
years be
accomplished at
Babylon I will
visit you, and
perform my
good
word toward you, in causing you to
return to this
place.
\VS{11}For I
know the
thoughts that I
think toward you,
saith the
{\ND{
LORD}},
thoughts of
peace, and not of
evil, to
give you an
expected
end.\FTNT{‡}{{\FR 29.11 }
expected…: Heb. end and expectation}
\VS{12}Then shall ye
call upon me, and ye shall
go and
pray unto me, and I will
hearken unto you.
\VS{13}And ye shall
seek me, and
find
{\ADD{me}}, when ye shall
search for me with all your
heart.
\VS{14}And I will be
found of you,
saith the
{\ND{
LORD}}: and I will turn
away your
captivity, and I will
gather you from all the
nations, and from all the
places whither I have
driven you,
saith the
{\ND{
LORD}}; and I will bring you
again into the
place whence I caused you to be carried away
captive.
\par }{\PP \VS{15}¶ Because ye have
said, The
{\ND{
LORD}} hath raised us
up
prophets in
Babylon;
\VS{16}{\ADD{Know}} that thus
saith the
{\ND{
LORD}} of the
king that
sitteth upon the
throne of
David, and of all the
people that
dwelleth in this
city,
{\ADD{and}} of your
brethren that are not gone
forth with you into
captivity;
\VS{17}Thus
saith the
{\ND{
LORD}} of
hosts; Behold, I will
send upon them the
sword, the
famine, and the
pestilence, and will
make them like
vile
figs, that cannot be
eaten, they are so
evil.
\VS{18}And I will
persecute
them with the
sword, with the
famine, and with the
pestilence, and will
deliver them to be
removed to all the
kingdoms of the
earth, to be a
curse, and an
astonishment, and an
hissing, and a
reproach, among all the
nations whither I have
driven them:\FTNT{§}{{\FR 29.18 }
to be a curse: Heb. for a curse}
\VS{19}Because they have not
hearkened to my
words,
saith the
{\ND{
LORD}}, which I
sent unto them by my
servants the
prophets, rising up
early and
sending
{\ADD{them;}} but ye would not
hear,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{20}¶
Hear ye therefore the
word of the
{\ND{
LORD}}, all ye of the
captivity, whom I have
sent from
Jerusalem to
Babylon:
\VS{21}Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel, of
Ahab the
son of
Kolaiah, and of
Zedekiah the
son of
Maaseiah, which
prophesy a
lie unto you in my
name; Behold, I will
deliver them into the
hand of
Nebuchadrezzar
king of
Babylon; and he shall
slay them before your
eyes;
\VS{22}And of them shall be taken
up a
curse by all the
captivity of
Judah which
{\ADD{are}} in
Babylon,
saying, The
{\ND{
LORD}}
make thee like
Zedekiah and like
Ahab, whom the
king of
Babylon
roasted in the
fire;
\VS{23}Because they have
committed
villany in
Israel, and have
committed
adultery with their
neighbours’
wives, and have
spoken
lying
words in my
name, which I have not
commanded them; even I
know, and
{\ADD{am}} a
witness,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{24}¶
{\ADD{Thus}} shalt thou also
speak to
Shemaiah the
Nehelamite,
saying,\FTNT{**}{{\FR 29.24 }
Nehelamite: or, dreamer}
\VS{25}Thus
speaketh the
{\ND{
LORD}} of
hosts, the
God of
Israel,
saying, Because thou hast
sent
letters in thy
name unto all the
people that
{\ADD{are}} at
Jerusalem, and to
Zephaniah the
son of
Maaseiah the
priest, and to all the
priests,
saying,
\VS{26}The
{\ND{
LORD}} hath
made thee
priest in the stead of
Jehoiada the
priest, that ye should be
officers in the
house of the
{\ND{
LORD}}, for every
man
{\ADD{that is}}
mad, and maketh himself a
prophet, that thou shouldest
put him in
prison, and in the
stocks.
\VS{27}Now therefore why hast thou not
reproved
Jeremiah of
Anathoth, which maketh himself a
prophet to you?
\VS{28}For therefore he
sent unto us
{\ADD{in}}
Babylon,
saying,
This
{\ADD{captivity is}}
long:
build ye
houses, and
dwell
{\ADD{in them;}} and
plant
gardens, and
eat the
fruit of them.
\VS{29}And
Zephaniah the
priest
read this
letter in the
ears of
Jeremiah the
prophet.
\par }{\PP \VS{30}¶ Then came the
word of the
{\ND{
LORD}} unto
Jeremiah,
saying,
\VS{31}Send to all them of the
captivity,
saying, Thus
saith the
{\ND{
LORD}} concerning
Shemaiah the
Nehelamite; Because that
Shemaiah hath
prophesied unto you, and I
sent him not, and he caused you to
trust in a
lie:
\VS{32}Therefore thus
saith the
{\ND{
LORD}}; Behold, I will
punish
Shemaiah the
Nehelamite, and his
seed: he shall not have a
man to
dwell
among this
people; neither shall he
behold the
good that I will
do for my
people,
saith the
{\ND{
LORD}}; because he hath
taught
rebellion against the
{\ND{
LORD}}.\FTNT{††}{{\FR 29.32 }
rebellion: Heb. revolt}

\par }\Chap{30}{\PP \VerseOne{1}The
word that came to
Jeremiah from the
{\ND{
LORD}},
saying,
\VS{2}Thus
speaketh the
{\ND{
LORD}}
God of
Israel,
saying,
Write thee all the
words that I have
spoken unto thee in a
book.
\VS{3}For, lo, the
days
come,
saith the
{\ND{
LORD}}, that I will bring
again the
captivity of my
people
Israel and
Judah,
saith the
{\ND{
LORD}}: and I will cause them to
return to the
land that I
gave to their
fathers, and they shall
possess it.
\par }{\PP \VS{4}¶ And these
{\ADD{are}} the
words that the
{\ND{
LORD}}
spake concerning
Israel and concerning
Judah.
\VS{5}For thus
saith the
{\ND{
LORD}}; We have
heard a
voice of
trembling, of
fear, and not of
peace.\FTNT{*}{{\FR 30.5 }
of fear…: or, there is fear, and not peace}
\VS{6}Ask ye now, and
see whether a
man doth travail with
child? wherefore do I
see every
man with his
hands on his
loins, as a woman in
travail, and all
faces are
turned into
paleness?\FTNT{†}{{\FR 30.6 }
a man: Heb. a male}
\VS{7}Alas! for that
day
{\ADD{is}}
great, so
that none
{\ADD{is}} like it: it
{\ADD{is}} even the
time of
Jacob’s
trouble; but he shall be
saved out of it.
\VS{8}For it shall come to pass in that
day,
saith the
{\ND{
LORD}} of
hosts,
{\ADD{that}} I will
break his
yoke from off thy
neck, and will
burst thy
bonds, and
strangers shall no more
serve themselves of him:
\VS{9}But they shall
serve the
{\ND{
LORD}} their
God, and
David their
king, whom I will raise
up unto them.
\par }{\PP \VS{10}¶ Therefore
fear thou not, O my
servant
Jacob,
saith the
{\ND{
LORD}}; neither be
dismayed, O
Israel: for, lo, I will
save thee from
afar, and thy
seed from the
land of their
captivity; and
Jacob shall
return, and shall be in
rest, and be
quiet, and none shall make
{\ADD{him}}
afraid.
\VS{11}For I
{\ADD{am}} with thee,
saith the
{\ND{
LORD}}, to
save thee: though I
make a full
end of all
nations whither I have
scattered thee, yet will I not
make a full
end of thee: but I will
correct thee in
measure, and will not leave thee
altogether
unpunished.
\VS{12}For thus
saith the
{\ND{
LORD}}, Thy
bruise
{\ADD{is}}
incurable,
{\ADD{and}} thy
wound
{\ADD{is}}
grievous.
\VS{13}{\ADD{There is}} none to
plead thy
cause, that thou mayest be bound
up: thou hast no
healing
medicines.\FTNT{‡}{{\FR 30.13 }
that…: Heb. for binding up, or, pressing}
\VS{14}All thy
lovers have
forgotten thee; they
seek thee not; for I have
wounded thee with the
wound of an
enemy, with the
chastisement of a cruel
one, for the
multitude of thine
iniquity;
{\ADD{because}} thy
sins were
increased.
\VS{15}Why
criest thou for thine
affliction? thy
sorrow
{\ADD{is}}
incurable for the
multitude of thine
iniquity:
{\ADD{because}} thy
sins were
increased, I have
done these things unto thee.
\VS{16}Therefore all they that
devour thee shall be
devoured; and all thine
adversaries, every one of them, shall
go into
captivity; and they that
spoil thee shall be a
spoil, and all that
prey upon thee will I
give for a
prey.
\VS{17}For I will
restore
health unto thee, and I will
heal thee of thy
wounds,
saith the
{\ND{
LORD}}; because they
called thee an
Outcast,
{\ADD{saying}}, This
{\ADD{is}}
Zion, whom no man seeketh
after.
\par }{\PP \VS{18}¶ Thus
saith the
{\ND{
LORD}}; Behold, I will bring
again the
captivity of
Jacob’s
tents, and have
mercy on his
dwellingplaces; and the
city shall be
builded upon her own
heap, and the
palace shall
remain after the
manner thereof.\FTNT{§}{{\FR 30.18 }
heap: or, little hill}
\VS{19}And out of them shall
proceed
thanksgiving and the
voice of them that make
merry: and I will
multiply them, and they shall not be
few; I will also
glorify them, and they shall not be
small.
\VS{20}Their
children also shall be as
aforetime, and their
congregation shall be
established
before me, and I will
punish all that
oppress them.
\VS{21}And their
nobles shall be of themselves, and their
governor shall
proceed from the
midst of them; and I will cause him to draw
near, and he shall
approach unto me: for who
{\ADD{is}} this that
engaged his
heart to
approach unto me?
saith the
{\ND{
LORD}}.
\VS{22}And ye shall be my
people, and I will be your
God.
\VS{23}Behold, the
whirlwind of the
{\ND{
LORD}} goeth
forth with
fury, a
continuing
whirlwind: it shall fall with
pain upon the
head of the
wicked.\FTNT{**}{{\FR 30.23 }
continuing: Heb. cutting}\FTNT{††}{{\FR 30.23 }
fall…: or, remain}
\VS{24}The
fierce
anger of the
{\ND{
LORD}} shall not
return, until he have
done
{\ADD{it}}, and until he have
performed the
intents of his
heart: in the
latter
days ye shall
consider it.

\par }\Chap{31}{\PP \VerseOne{1}At the same
time,
saith the
{\ND{
LORD}}, will I be the
God of all the
families of
Israel, and they shall be my
people.
\VS{2}Thus
saith the
{\ND{
LORD}}, The
people
{\ADD{which were}}
left of the
sword
found
grace in the
wilderness;
{\ADD{even}}
Israel, when I
went to cause him to
rest.
\VS{3}The
{\ND{
LORD}} hath
appeared of
old unto me,
{\ADD{saying}}, Yea, I have
loved thee with an
everlasting
love: therefore with
lovingkindness have I
drawn thee.\FTNT{*}{{\FR 31.3 }
of…: Heb. from afar}\FTNT{†}{{\FR 31.3 }
with lovingkindness…: have I extended lovingkindness unto thee}
\VS{4}Again I will
build thee, and thou shalt be
built, O
virgin of
Israel: thou shalt again be
adorned with thy
tabrets, and shalt go
forth in the
dances of them that make
merry.\FTNT{‡}{{\FR 31.4 }
tabrets: or, timbrels}
\VS{5}Thou shalt yet
plant
vines upon the
mountains of
Samaria: the
planters shall
plant, and shall eat
{\ADD{them}} as common
things.\FTNT{§}{{\FR 31.5 }
eat…: Heb. profane them}
\VS{6}For there shall
be a
day,
{\ADD{that}} the
watchmen upon the
mount
Ephraim shall
cry,
Arise ye, and let us go
up to
Zion unto the
{\ND{
LORD}} our
God.
\VS{7}For thus
saith the
{\ND{
LORD}};
Sing with
gladness for
Jacob, and
shout among the
chief of the
nations:
publish ye,
praise ye, and
say, O
{\ND{
LORD}},
save thy
people, the
remnant of
Israel.
\VS{8}Behold, I will
bring them from the
north
country, and
gather them from the
coasts of the
earth,
{\ADD{and}} with them the
blind and the
lame, the woman with
child and her that travaileth with
child
together: a
great
company shall
return thither.
\VS{9}They shall
come with
weeping, and with
supplications will I
lead them: I will cause them to
walk by the
rivers of
waters in a
straight
way, wherein they shall not
stumble: for I am a
father to
Israel, and
Ephraim
{\ADD{is}} my
firstborn.\FTNT{**}{{\FR 31.9 }
supplications: or, favours}
\par }{\PP \VS{10}¶
Hear the
word of the
{\ND{
LORD}}, O ye
nations, and
declare
{\ADD{it}} in the
isles afar
off, and
say, He that
scattered
Israel will
gather him, and
keep him, as a
shepherd
{\ADD{doth}} his
flock.
\VS{11}For the
{\ND{
LORD}} hath
redeemed
Jacob, and
ransomed him from the
hand of
{\ADD{him that was}}
stronger than he.
\VS{12}Therefore they shall
come and
sing in the
height of
Zion, and shall flow
together to the
goodness of the
{\ND{
LORD}}, for
wheat, and for
wine, and for
oil, and for the
young of the
flock and of the
herd: and their
soul shall be as a
watered
garden; and they shall not
sorrow any
more at all.
\VS{13}Then shall the
virgin
rejoice in the
dance, both young
men and
old
together: for I will
turn their
mourning into
joy, and will
comfort them, and make them
rejoice from their
sorrow.
\VS{14}And I will
satiate the
soul of the
priests with
fatness, and my
people shall be
satisfied with my
goodness,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{15}¶ Thus
saith the
{\ND{
LORD}}; A
voice was
heard in
Ramah,
lamentation,
{\ADD{and}}
bitter
weeping;
Rahel
weeping for her
children
refused to be
comforted for her
children, because they
{\ADD{were}} not.
\VS{16}Thus
saith the
{\ND{
LORD}};
Refrain thy
voice from
weeping, and thine
eyes from
tears: for thy
work shall
be
rewarded,
saith the
{\ND{
LORD}}; and they shall come
again from the
land of the
enemy.
\VS{17}And there
is
hope in thine
end,
saith the
{\ND{
LORD}}, that thy
children shall come
again to their own
border.
\par }{\PP \VS{18}¶ I have
surely
heard
Ephraim
bemoaning himself
{\ADD{thus;}} Thou hast
chastised me, and I was
chastised, as a
bullock
unaccustomed
{\ADD{to the yoke:}}
turn thou me, and I shall be
turned; for thou
{\ADD{art}} the
{\ND{
LORD}} my
God.
\VS{19}Surely
after that I was
turned, I
repented; and
after that I was
instructed, I
smote upon
{\ADD{my}}
thigh: I was
ashamed, yea, even
confounded, because I did
bear the
reproach of my
youth.
\VS{20}{\ADD{Is}}
Ephraim my
dear
son?
{\ADD{is he}} a
pleasant
child? for
since I
spake against him, I do
earnestly
remember him still: therefore my
bowels are
troubled for him; I will
surely have
mercy upon him,
saith the
{\ND{
LORD}}.\FTNT{††}{{\FR 31.20 }
are…: Heb. sound}
\VS{21}Set thee
up
waymarks,
make thee high
heaps:
set thine
heart toward the
highway,
{\ADD{even}} the
way
{\ADD{which}} thou
wentest: turn
again, O
virgin of
Israel, turn
again to these thy
cities.
\par }{\PP \VS{22}¶ How long wilt thou go
about, O thou
backsliding
daughter? for the
{\ND{
LORD}} hath
created a new
thing in the
earth, A
woman shall
compass a
man.
\VS{23}Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; As yet they shall
use this
speech in the
land of
Judah and in the
cities thereof, when I shall bring
again their
captivity; The
{\ND{
LORD}}
bless thee, O
habitation of
justice,
{\ADD{and}}
mountain of
holiness.
\VS{24}And there shall
dwell in
Judah itself, and in all the
cities thereof
together,
husbandmen, and they
{\ADD{that}} go
forth with
flocks.
\VS{25}For I have
satiated the
weary
soul, and I have
replenished every
sorrowful
soul.
\VS{26}Upon this I
awaked, and
beheld; and my
sleep was
sweet unto me.
\par }{\PP \VS{27}¶ Behold, the
days
come,
saith the
{\ND{
LORD}}, that I will
sow the
house of
Israel and the
house of
Judah with the
seed of
man, and with the
seed of
beast.
\VS{28}And it shall come to pass,
{\ADD{that}} like as I have watched
over them, to pluck
up, and to break
down, and to throw
down, and to
destroy, and to
afflict; so will I watch
over them, to
build, and to
plant,
saith the
{\ND{
LORD}}.
\VS{29}In those
days they shall
say no more, The
fathers have
eaten a sour
grape, and the
children’s
teeth are set on
edge.
\VS{30}But every
one shall
die for his own
iniquity: every
man that
eateth the sour
grape, his
teeth shall be set on
edge.
\par }{\PP \VS{31}¶ Behold, the
days
come,
saith the
{\ND{
LORD}}, that I will
make a
new
covenant with the
house of
Israel, and with the
house of
Judah:
\VS{32}Not according to the
covenant that I
made with their
fathers in the
day
{\ADD{that}} I
took them by the
hand to bring them
out of the
land of
Egypt; which my
covenant they
brake, although I was an
husband unto them,
saith the
{\ND{
LORD}}:\FTNT{‡‡}{{\FR 31.32 }
although…: or, should I have continued an husband unto them?}
\VS{33}But this
{\ADD{shall be}} the
covenant that I will
make with the
house of
Israel;
After those
days,
saith the
{\ND{
LORD}}, I will
put my
law in their inward
parts, and
write it in their
hearts; and will be their
God, and they shall be my
people.
\VS{34}And they shall
teach no more every
man his
neighbour, and every
man his
brother,
saying,
Know the
{\ND{
LORD}}: for they shall all
know me, from the
least of them unto the
greatest of them,
saith the
{\ND{
LORD}}: for I will
forgive their
iniquity, and I will
remember their
sin no more.
\par }{\PP \VS{35}¶ Thus
saith the
{\ND{
LORD}}, which
giveth the
sun for a
light by
day,
{\ADD{and}} the
ordinances of the
moon and of the
stars for a
light by
night, which
divideth the
sea when the
waves thereof
roar; The
{\ND{
LORD}} of
hosts
{\ADD{is}} his
name:
\VS{36}If those
ordinances
depart from
before me,
saith the
{\ND{
LORD}},
{\ADD{then}} the
seed of
Israel also shall
cease from being a
nation
before me for
ever.
\VS{37}Thus
saith the
{\ND{
LORD}}; If
heaven
above can be
measured, and the
foundations of the
earth searched
out
beneath, I will also cast
off all the
seed of
Israel for all that they have
done,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{38}¶ Behold, the
days
come,
saith the
{\ND{
LORD}}, that the
city shall be
built to the
{\ND{
LORD}} from the
tower of
Hananeel unto the
gate of the
corner.
\VS{39}And the
measuring
line shall yet go
forth over against it upon the
hill
Gareb, and shall compass
about to
Goath.
\VS{40}And the whole
valley of the dead
bodies, and of the
ashes, and all the
fields unto the
brook of
Kidron, unto the
corner of the
horse
gate toward the
east,
{\ADD{shall be}}
holy unto the
{\ND{
LORD}}; it shall not be plucked
up, nor thrown
down any more for
ever.

\par }\Chap{32}{\PP \VerseOne{1}The
word that came to
Jeremiah from the
{\ND{
LORD}} in the
tenth
year of
Zedekiah
king of
Judah, which
{\ADD{was}} the
eighteenth
year of
Nebuchadrezzar.
\VS{2}For then the
king of
Babylon’s
army
besieged
Jerusalem: and
Jeremiah the
prophet was shut
up in the
court of the
prison, which
{\ADD{was}} in the
king of
Judah’s
house.
\VS{3}For
Zedekiah
king of
Judah had shut him
up,
saying, Wherefore dost thou
prophesy, and
say, Thus
saith the
{\ND{
LORD}}, Behold, I will
give this
city into the
hand of the
king of
Babylon, and he shall
take it;
\VS{4}And
Zedekiah
king of
Judah shall not
escape out of the
hand of the
Chaldeans, but shall
surely be
delivered into the
hand of the
king of
Babylon, and shall
speak with him
mouth to
mouth, and his
eyes shall
behold his
eyes;
\VS{5}And he shall
lead
Zedekiah to
Babylon, and there shall he be until I
visit him,
saith the
{\ND{
LORD}}: though ye
fight with the
Chaldeans, ye shall not
prosper.
\par }{\PP \VS{6}¶ And
Jeremiah
said, The
word of the
{\ND{
LORD}} came unto me,
saying,
\VS{7}Behold,
Hanameel the
son of
Shallum thine
uncle shall
come unto thee,
saying,
Buy thee my
field that
{\ADD{is}} in
Anathoth: for the
right of
redemption
{\ADD{is}} thine to
buy
{\ADD{it}}.
\VS{8}So
Hanameel mine
uncle’s
son
came to me in the
court of the
prison according to the
word of the
{\ND{
LORD}}, and
said unto me,
Buy my
field, I pray thee, that
{\ADD{is}} in
Anathoth, which
{\ADD{is}} in the
country of
Benjamin: for the
right of
inheritance
{\ADD{is}} thine, and the
redemption
{\ADD{is}} thine;
buy
{\ADD{it}} for thyself. Then I
knew that this
{\ADD{was}} the
word of the
{\ND{
LORD}}.
\VS{9}And I
bought the
field of
Hanameel my
uncle’s
son, that
{\ADD{was}} in
Anathoth, and
weighed him the
money,
{\ADD{even}}
seventeen
shekels of
silver.\FTNT{*}{{\FR 32.9 }
seventeen…: or, seven shekels and ten pieces of silver}
\VS{10}And I
subscribed the
evidence, and
sealed
{\ADD{it}}, and
took
witnesses, and
weighed
{\ADD{him}} the
money in the
balances.\FTNT{†}{{\FR 32.10 }
subscribed…: Heb. wrote in the book}
\VS{11}So I
took the
evidence of the
purchase,
{\ADD{both}} that which was
sealed
{\ADD{according}} to the
law and
custom, and that which was
open:
\VS{12}And I
gave the
evidence of the
purchase unto
Baruch the
son of
Neriah, the
son of
Maaseiah, in the
sight of
Hanameel mine
uncle’s
{\ADD{son}}, and in the
presence of the
witnesses that
subscribed the
book of the
purchase,
before all the
Jews that
sat in the
court of the
prison.
\par }{\PP \VS{13}¶ And I
charged
Baruch
before them,
saying,
\VS{14}Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel;
Take these
evidences, this
evidence of the
purchase, both which is
sealed, and this
evidence which is
open; and
put them in an
earthen
vessel, that they may
continue
many
days.
\VS{15}For thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel;
Houses and
fields and
vineyards shall be possessed
again in this
land.
\par }{\PP \VS{16}¶ Now
when I had
delivered the
evidence of the
purchase unto
Baruch the
son of
Neriah, I
prayed unto the
{\ND{
LORD}},
saying,
\VS{17}Ah
Lord
{\ND{
GOD}}! behold, thou hast
made the
heaven and the
earth by thy
great
power and stretched
out
arm,
{\ADD{and}} there is
nothing too
hard for thee:\FTNT{‡}{{\FR 32.17 }
too…: or, hid from thee}
\VS{18}Thou
shewest
lovingkindness unto
thousands, and
recompensest the
iniquity of the
fathers into the
bosom of their
children
after them: the
Great, the
Mighty
God, the
{\ND{
LORD}} of
hosts,
{\ADD{is}} his
name,
\VS{19}Great in
counsel, and
mighty in
work: for thine
eyes
{\ADD{are}}
open upon all the
ways of the
sons of
men: to
give every
one according to his
ways, and according to the
fruit of his
doings:\FTNT{§}{{\FR 32.19 }
work: Heb. doing}
\VS{20}Which hast
set
signs and
wonders in the
land of
Egypt,
{\ADD{even}} unto this
day, and in
Israel, and among
{\ADD{other}}
men; and hast
made thee a
name, as at this
day;
\VS{21}And hast brought
forth thy
people
Israel out of the
land of
Egypt with
signs, and with
wonders, and with a
strong
hand, and with a stretched
out
arm, and with
great
terror;
\VS{22}And hast
given them this
land, which thou didst
swear to their
fathers to
give them, a
land
flowing with
milk and
honey;
\VS{23}And they came
in, and
possessed it; but they
obeyed not thy
voice, neither
walked in thy
law; they have
done nothing of all that thou
commandedst them to
do: therefore thou hast caused all this
evil to
come upon them:
\VS{24}Behold the
mounts, they are
come unto the
city to
take it; and the
city is
given into the
hand of the
Chaldeans, that
fight against it,
because of the
sword, and of the
famine, and of the
pestilence: and what thou hast
spoken is come to pass; and, behold, thou
seest
{\ADD{it}}.\FTNT{**}{{\FR 32.24 }
mounts: or, engines of shot}
\VS{25}And thou hast
said unto me, O
Lord
{\ND{
GOD}},
Buy thee the
field for
money, and
take
witnesses; for the
city is
given into the
hand of the
Chaldeans.\FTNT{††}{{\FR 32.25 }
for the: or, though the}
\par }{\PP \VS{26}¶ Then came the
word of the
{\ND{
LORD}} unto
Jeremiah,
saying,
\VS{27}Behold, I
{\ADD{am}} the
{\ND{
LORD}}, the
God of all
flesh: is there any
thing too
hard for me?
\VS{28}Therefore thus
saith the
{\ND{
LORD}}; Behold, I will
give this
city into the
hand of the
Chaldeans, and into the
hand of
Nebuchadrezzar
king of
Babylon, and he shall
take it:
\VS{29}And the
Chaldeans, that
fight against this
city, shall
come and
set
fire on this
city, and
burn it with the
houses, upon whose
roofs they have offered
incense unto
Baal, and poured
out drink
offerings unto
other
gods, to provoke me to
anger.
\VS{30}For the
children of
Israel and the
children of
Judah have only
done
evil
before me from their
youth: for the
children of
Israel have only provoked me to
anger with the
work of their
hands,
saith the
{\ND{
LORD}}.
\VS{31}For this
city hath been to me
{\ADD{as}} a provocation of mine
anger and of my
fury from the
day that they
built it even unto this
day; that I should
remove it from before my
face,\FTNT{‡‡}{{\FR 32.31 }
a provocation…: Heb. for my anger}
\VS{32}Because of all the
evil of the
children of
Israel and of the
children of
Judah, which they have
done to provoke me to
anger, they, their
kings, their
princes, their
priests, and their
prophets, and the
men of
Judah, and the
inhabitants of
Jerusalem.
\VS{33}And they have
turned unto me the
back, and not the
face: though I
taught them, rising up
early and
teaching
{\ADD{them}}, yet they have not
hearkened to
receive
instruction.\FTNT{§§}{{\FR 32.33 }
back: Heb. neck}
\VS{34}But they
set their
abominations in the
house, which is
called by my
name, to
defile it.
\VS{35}And they
built the high
places of
Baal, which
{\ADD{are}} in the
valley of the
son of
Hinnom, to cause their
sons and their
daughters to
pass through
{\ADD{the fire}} unto
Molech; which I
commanded them not, neither
came it into my
mind, that they should
do this
abomination, to cause
Judah to
sin.
\par }{\PP \VS{36}¶ And now therefore thus
saith the
{\ND{
LORD}}, the
God of
Israel, concerning this
city, whereof ye
say, It shall be
delivered into the
hand of the
king of
Babylon by the
sword, and by the
famine, and by the
pestilence;
\VS{37}Behold, I will gather them
out of all
countries, whither I have
driven them in mine
anger, and in my
fury, and in
great
wrath; and I will bring them
again unto this
place, and I will cause them to
dwell
safely:
\VS{38}And they shall be my
people, and I will be their
God:
\VS{39}And I will
give them
one
heart, and
one
way, that they may
fear me for
ever, for the
good of them, and of their
children
after them:\FTNT{***}{{\FR 32.39 }
for ever: Heb. all days}
\VS{40}And I will
make an
everlasting
covenant with them, that I will not turn
away from
them, to do them
good; but I will
put my
fear in their
hearts, that they shall not
depart from me.\FTNT{†††}{{\FR 32.40 }
from them: Heb. from after them}
\VS{41}Yea, I will
rejoice over them to do them
good, and I will
plant them in this
land
assuredly with my whole
heart and with my whole
soul.\FTNT{‡‡‡}{{\FR 32.41 }
assuredly: Heb. in truth, or, stability}
\VS{42}For thus
saith the
{\ND{
LORD}}; Like as I have
brought all this
great
evil upon this
people, so will I
bring upon them all the
good that I have
promised them.
\VS{43}And
fields shall be
bought in this
land, whereof ye
say,
{\ADD{It is}}
desolate without
man or
beast; it is
given into the
hand of the
Chaldeans.
\VS{44}Men shall
buy
fields for
money, and
subscribe
evidences, and
seal
{\ADD{them}}, and
take
witnesses in the
land of
Benjamin, and in the places
about
Jerusalem, and in the
cities of
Judah, and in the
cities of the
mountains, and in the
cities of the
valley, and in the
cities of the
south: for I will cause their
captivity to
return,
saith the
{\ND{
LORD}}.

\par }\Chap{33}{\PP \VerseOne{1}Moreover the
word of the
{\ND{
LORD}} came unto
Jeremiah the
second time, while he was yet shut
up in the
court of the
prison,
saying,
\VS{2}Thus
saith the
{\ND{
LORD}} the
maker thereof, the
{\ND{
LORD}} that
formed it, to
establish it; the
{\ND{
LORD}}
{\ADD{is}} his
name;\FTNT{*}{{\FR 33.2 }
the
{\ND{LORD}} is: or,
{\ND{JEHOVAH}}, etc}
\VS{3}Call unto me, and I will
answer thee, and
shew thee
great and mighty
things, which thou
knowest not.\FTNT{†}{{\FR 33.3 }
mighty: or, hidden}
\VS{4}For thus
saith the
{\ND{
LORD}}, the
God of
Israel, concerning the
houses of this
city, and concerning the
houses of the
kings of
Judah, which are thrown
down by the
mounts, and by the
sword;
\VS{5}They
come to
fight with the
Chaldeans, but
{\ADD{it is}} to
fill them with the dead
bodies of
men, whom I have
slain in mine
anger and in my
fury, and for all whose
wickedness I have
hid my
face from this
city.
\VS{6}Behold, I will
bring it
health and
cure, and I will
cure them, and will
reveal unto them the
abundance of
peace and
truth.
\VS{7}And I will cause the
captivity of
Judah and the
captivity of
Israel to
return, and will
build them, as at the
first.
\VS{8}And I will
cleanse them from all their
iniquity, whereby they have
sinned against me; and I will
pardon all their
iniquities, whereby they have
sinned, and whereby they have
transgressed against me.
\par }{\PP \VS{9}¶ And it shall be to me a
name of
joy, a
praise and an
honour before all the
nations of the
earth, which shall
hear all the
good that I
do unto them: and they shall
fear and
tremble for all the
goodness and for all the
prosperity that I
procure unto it.
\par }{\PP \VS{10}Thus
saith the
{\ND{
LORD}}; Again there shall be
heard in this
place, which ye
say
{\ADD{shall be}}
desolate without
man and without
beast,
{\ADD{even}} in the
cities of
Judah, and in the
streets of
Jerusalem, that are
desolate, without
man, and without
inhabitant, and without
beast,
\VS{11}The
voice of
joy, and the
voice of
gladness, the
voice of the
bridegroom, and the
voice of the
bride, the
voice of them that shall
say,
Praise the
{\ND{
LORD}} of
hosts: for the
{\ND{
LORD}}
{\ADD{is}}
good; for his
mercy
{\ADD{endureth}} for
ever:
{\ADD{and}} of them that shall
bring the sacrifice of
praise into the
house of the
{\ND{
LORD}}. For I will cause to
return the
captivity of the
land, as at the
first,
saith the
{\ND{
LORD}}.
\VS{12}Thus
saith the
{\ND{
LORD}} of
hosts; Again in this
place, which is
desolate without
man and without
beast, and in all the
cities thereof, shall be an
habitation of
shepherds causing
{\ADD{their}}
flocks to lie
down.
\VS{13}In the
cities of the
mountains, in the
cities of the
vale, and in the
cities of the
south, and in the
land of
Benjamin, and in the places
about
Jerusalem, and in the
cities of
Judah, shall the
flocks pass
again under the
hands of him that
telleth
{\ADD{them}},
saith the
{\ND{
LORD}}.
\VS{14}Behold, the
days
come,
saith the
{\ND{
LORD}}, that I will
perform that
good
thing which I have
promised unto the
house of
Israel and to the
house of
Judah.
\par }{\PP \VS{15}¶ In those
days, and at that
time, will I cause the
Branch of
righteousness to grow
up unto
David; and he shall
execute
judgment and
righteousness in the
land.
\VS{16}In those
days shall
Judah be
saved, and
Jerusalem shall
dwell
safely: and this
{\ADD{is the name}} wherewith she shall be
called, The
{\ND{LORD}} our
righteousness.\FTNT{‡}{{\FR 33.16 }
The
{\ND{LORD}}…: Heb. Jehovah-tsidkenu}
\par }{\PP \VS{17}¶ For thus
saith the
{\ND{
LORD}};
David shall
never
want a
man to
sit upon the
throne of the
house of
Israel;\FTNT{§}{{\FR 33.17 }
David…: Heb. There shall not be cut off from David}
\VS{18}Neither shall the
priests the
Levites
want a
man
before me to
offer burnt
offerings, and to
kindle meat
offerings, and to
do
sacrifice
continually.
\par }{\PP \VS{19}¶ And the
word of the
{\ND{
LORD}} came unto
Jeremiah,
saying,
\VS{20}Thus
saith the
{\ND{
LORD}}; If ye can
break my
covenant of the
day, and my
covenant of the
night, and that there should not be
day and
night in their
season;
\VS{21}{\ADD{Then}} may also my
covenant be
broken with
David my
servant, that he should not have a
son to
reign upon his
throne; and with the
Levites the
priests, my
ministers.
\VS{22}As the
host of
heaven cannot be
numbered, neither the
sand of the
sea
measured: so will I
multiply the
seed of
David my
servant, and the
Levites that
minister unto me.
\VS{23}Moreover the
word of the
{\ND{
LORD}} came to
Jeremiah,
saying,
\VS{24}Considerest thou not what this
people have
spoken,
saying, The
two
families which the
{\ND{
LORD}} hath
chosen, he hath even cast them
off? thus they have
despised my
people, that they should be no more a
nation
before them.
\VS{25}Thus
saith the
{\ND{
LORD}}; If my
covenant
{\ADD{be}} not with
day and
night,
{\ADD{and if}} I have not
appointed the
ordinances of
heaven and
earth;
\VS{26}Then will I cast
away the
seed of
Jacob, and
David my
servant,
{\ADD{so}} that I will not
take
{\ADD{any}} of his
seed
{\ADD{to be}}
rulers over the
seed of
Abraham,
Isaac, and
Jacob: for I will cause their
captivity to
return, and have
mercy on them.

\par }\Chap{34}{\PP \VerseOne{1}The
word which came unto
Jeremiah from the
{\ND{
LORD}}, when
Nebuchadnezzar
king of
Babylon, and all his
army, and all the
kingdoms of the
earth of his
dominion, and all the
people,
fought against
Jerusalem, and against all the
cities thereof,
saying,\FTNT{*}{{\FR 34.1 }
of his…: Heb. the dominion of his hand}
\VS{2}Thus
saith the
{\ND{
LORD}}, the
God of
Israel;
Go and
speak to
Zedekiah
king of
Judah, and
tell him, Thus
saith the
{\ND{
LORD}}; Behold, I will
give this
city into the
hand of the
king of
Babylon, and he shall
burn it with
fire:
\VS{3}And thou shalt not escape
out of his
hand, but shalt
surely be
taken, and
delivered into his
hand; and thine
eyes shall
behold the
eyes of the
king of
Babylon, and he shall
speak with thee
mouth to
mouth, and thou shalt
go to
Babylon.\FTNT{†}{{\FR 34.3 }
he shall…: Heb. his mouth shall speak to thy mouth}
\VS{4}Yet
hear the
word of the
{\ND{
LORD}}, O
Zedekiah
king of
Judah; Thus
saith the
{\ND{
LORD}} of thee, Thou shalt not
die by the
sword:
\VS{5}{\ADD{But}} thou shalt
die in
peace: and with the
burnings of thy
fathers, the
former
kings which were
before thee, so shall they
burn
{\ADD{odours}} for thee; and they will
lament thee,
{\ADD{saying}},
Ah
lord! for I have
pronounced the
word,
saith the
{\ND{
LORD}}.
\VS{6}Then
Jeremiah the
prophet
spake all these
words unto
Zedekiah
king of
Judah in
Jerusalem,
\VS{7}When the
king of
Babylon’s
army
fought against
Jerusalem, and against all the
cities of
Judah that were
left, against
Lachish, and against
Azekah: for these
defenced
cities
remained of the
cities of
Judah.
\par }{\PP \VS{8}¶
{\ADD{This is}} the
word that came unto
Jeremiah from the
{\ND{
LORD}},
after that the
king
Zedekiah had
made a
covenant with all the
people which
{\ADD{were}} at
Jerusalem, to
proclaim
liberty unto them;
\VS{9}That every
man should let his
manservant, and every
man his
maidservant,
{\ADD{being}} an
Hebrew or an
Hebrewess,
go
free; that none should
serve himself of them,
{\ADD{to wit}}, of a
Jew his
brother.
\VS{10}Now when all the
princes, and all the
people, which had
entered into the
covenant,
heard that every
one should let his
manservant, and every
one his
maidservant,
go
free, that none should
serve themselves of them any more, then they
obeyed, and let
{\ADD{them}}
go.
\VS{11}But
afterward they
turned, and caused the
servants and the
handmaids, whom they had let
go
free, to
return, and brought them into
subjection for
servants and for
handmaids.
\par }{\PP \VS{12}¶ Therefore the
word of the
{\ND{
LORD}} came to
Jeremiah from the
{\ND{
LORD}},
saying,
\VS{13}Thus
saith the
{\ND{
LORD}}, the
God of
Israel; I
made a
covenant with your
fathers in the
day that I brought them
forth out of the
land of
Egypt, out of the
house of
bondmen,
saying,
\VS{14}At the
end of
seven
years let ye
go every
man his
brother an
Hebrew, which hath been
sold unto thee; and when he hath
served thee
six
years, thou shalt let him
go
free from thee: but your
fathers
hearkened not unto me, neither
inclined their
ear.\FTNT{‡}{{\FR 34.14 }
hath been…: or, hath sold himself}
\VS{15}And ye were
now
turned, and had
done
right in my
sight, in
proclaiming
liberty every
man to his
neighbour; and ye had
made a
covenant
before me in the
house which is
called by my
name:\FTNT{§}{{\FR 34.15 }
now: Heb. to day}\FTNT{**}{{\FR 34.15 }
which…: Heb. whereupon my name is called}
\VS{16}But ye
turned and
polluted my
name, and caused every
man his
servant, and every
man his
handmaid, whom ye had
set at
liberty at their
pleasure, to
return, and brought them into
subjection, to be unto you for
servants and for
handmaids.
\VS{17}Therefore thus
saith the
{\ND{
LORD}}; Ye have not
hearkened unto me, in
proclaiming
liberty, every
one to his
brother, and every
man to his
neighbour: behold, I
proclaim a
liberty for you,
saith the
{\ND{
LORD}}, to the
sword, to the
pestilence, and to the
famine; and I will
make you to be
removed into all the
kingdoms of the
earth.\FTNT{††}{{\FR 34.17 }
to be…: Heb. for a removing}
\VS{18}And I will
give the
men that have
transgressed my
covenant, which have not
performed the
words of the
covenant which they had
made
before me, when they
cut the
calf in
twain, and
passed between the
parts thereof,
\VS{19}The
princes of
Judah, and the
princes of
Jerusalem, the
eunuchs, and the
priests, and all the
people of the
land, which
passed between the
parts of the
calf;
\VS{20}I will even
give them into the
hand of their
enemies, and into the
hand of them that
seek their
life: and their dead
bodies shall be for
meat unto the
fowls of the
heaven, and to the
beasts of the
earth.
\VS{21}And
Zedekiah
king of
Judah and his
princes will I
give into the
hand of their
enemies, and into the
hand of them that
seek their
life, and into the
hand of the
king of
Babylon’s
army, which are gone
up from you.
\VS{22}Behold, I will
command,
saith the
{\ND{
LORD}}, and cause them to
return to this
city; and they shall
fight against it, and
take it, and
burn it with
fire: and I will
make the
cities of
Judah a
desolation without an
inhabitant.

\par }\Chap{35}{\PP \VerseOne{1}The
word which came unto
Jeremiah from the
{\ND{
LORD}} in the
days of
Jehoiakim the
son of
Josiah
king of
Judah,
saying,
\VS{2}Go unto the
house of the
Rechabites, and
speak unto them, and
bring them into the
house of the
{\ND{
LORD}}, into
one of the
chambers, and give them
wine to
drink.
\VS{3}Then I
took
Jaazaniah the
son of
Jeremiah, the
son of
Habaziniah, and his
brethren, and all his
sons, and the whole
house of the
Rechabites;
\VS{4}And I
brought them into the
house of the
{\ND{
LORD}}, into the
chamber of the
sons of
Hanan, the
son of
Igdaliah, a
man of
God, which
{\ADD{was}}
by the
chamber of the
princes, which
{\ADD{was}}
above the
chamber of
Maaseiah the
son of
Shallum, the
keeper of the
door:\FTNT{*}{{\FR 35.4 }
door: Heb. threshold, or, vessel}
\VS{5}And I
set
before the
sons of the
house of the
Rechabites
pots
full of
wine, and
cups, and I
said unto them,
Drink ye
wine.
\VS{6}But they
said, We will
drink no
wine: for
Jonadab the
son of
Rechab our
father
commanded us,
saying, Ye shall
drink no
wine,
{\ADD{neither ye}}, nor your
sons
for
ever:
\VS{7}Neither shall ye
build
house, nor
sow
seed, nor
plant
vineyard, nor have
{\ADD{any:}} but all your
days ye shall
dwell in
tents; that ye may
live
many
days
in the
land where ye
{\ADD{be}}
strangers.
\VS{8}Thus have we
obeyed the
voice of
Jonadab the
son of
Rechab our
father in all that he hath
charged us, to
drink no
wine all our
days, we, our
wives, our
sons, nor our
daughters;
\VS{9}Nor to
build
houses for us to dwell
in: neither have we
vineyard, nor
field, nor
seed:
\VS{10}But we have dwelt
in
tents, and have
obeyed, and
done according to all that
Jonadab our
father
commanded us.
\VS{11}But it came to pass, when
Nebuchadrezzar
king of
Babylon came
up into the
land, that we
said,
Come, and let us
go to
Jerusalem for
fear of the
army of the
Chaldeans, and for
fear of the
army of the
Syrians: so we
dwell at
Jerusalem.
\par }{\PP \VS{12}¶ Then came the
word of the
{\ND{
LORD}} unto
Jeremiah,
saying,
\VS{13}Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel;
Go and
tell the
men of
Judah and the
inhabitants of
Jerusalem, Will ye not
receive
instruction to
hearken to my
words?
saith the
{\ND{
LORD}}.
\VS{14}The
words of
Jonadab the
son of
Rechab, that he
commanded his
sons not to
drink
wine, are
performed; for unto this
day they
drink none, but
obey their
father’s
commandment: notwithstanding I have
spoken unto you, rising
early and
speaking; but ye
hearkened not unto me.
\VS{15}I have
sent also unto you all my
servants the
prophets, rising up
early and
sending
{\ADD{them}},
saying,
Return ye now every
man from his
evil
way, and
amend your
doings, and
go not
after
other
gods to
serve them, and ye shall dwell
in the
land which I have
given to you and to your
fathers: but ye have not
inclined your
ear, nor
hearkened unto me.
\VS{16}Because the
sons of
Jonadab the
son of
Rechab have
performed the
commandment of their
father, which he
commanded them; but this
people hath not
hearkened unto me:
\VS{17}Therefore thus
saith the
{\ND{
LORD}}
God of
hosts, the
God of
Israel; Behold, I will
bring upon
Judah and upon all the
inhabitants of
Jerusalem all the
evil that I have
pronounced against them: because I have
spoken unto them, but they have not
heard; and I have
called unto them, but they have not
answered.
\par }{\PP \VS{18}¶ And
Jeremiah
said unto the
house of the
Rechabites, Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; Because ye have
obeyed the
commandment of
Jonadab your
father, and
kept all his
precepts, and
done according unto all that he hath
commanded you:
\VS{19}Therefore thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel;
Jonadab the
son of
Rechab shall not
want a
man to
stand
before me for
ever.\FTNT{†}{{\FR 35.19 }
Jonadab…: Heb. There shall not a man be cut off from Jonadab the son of Rechab to stand, etc}

\par }\Chap{36}{\PP \VerseOne{1}And it came to pass in the
fourth
year of
Jehoiakim the
son of
Josiah
king of
Judah,
{\ADD{that}} this
word came unto
Jeremiah from the
{\ND{
LORD}},
saying,
\VS{2}Take thee a
roll of a
book, and
write therein all the
words that I have
spoken unto thee against
Israel, and against
Judah, and against all the
nations, from the
day I
spake unto thee, from the
days of
Josiah, even unto this
day.
\VS{3}It may be that the
house of
Judah will
hear all the
evil which I
purpose to
do unto them; that they may
return every
man from his
evil
way; that I may
forgive their
iniquity and their
sin.
\VS{4}Then
Jeremiah
called
Baruch the
son of
Neriah: and
Baruch
wrote from the
mouth of
Jeremiah all the
words of the
{\ND{
LORD}}, which he had
spoken unto him, upon a
roll of a
book.
\VS{5}And
Jeremiah
commanded
Baruch,
saying, I
{\ADD{am}} shut
up; I
cannot
go into the
house of the
{\ND{
LORD}}:
\VS{6}Therefore
go thou, and
read in the
roll, which thou hast
written from my
mouth, the
words of the
{\ND{
LORD}} in the
ears of the
people in the
{\ND{
LORD’s}}
house upon the
fasting
day: and also thou shalt
read them in the
ears of all
Judah that come
out of their
cities.
\VS{7}It may be they will
present their
supplication
before the
{\ND{
LORD}}, and will
return every
one from his
evil
way: for
great
{\ADD{is}} the
anger and the
fury that the
{\ND{
LORD}} hath
pronounced against this
people.\FTNT{*}{{\FR 36.7 }
they…: Heb. their supplication shall fail}
\VS{8}And
Baruch the
son of
Neriah
did according to all that
Jeremiah the
prophet
commanded him,
reading in the
book the
words of the
{\ND{
LORD}} in the
{\ND{
LORD’s}}
house.
\par }{\PP \VS{9}And it came to pass in the
fifth
year of
Jehoiakim the
son of
Josiah
king of
Judah, in the
ninth
month,
{\ADD{that}} they
proclaimed a
fast
before the
{\ND{
LORD}} to all the
people in
Jerusalem, and to all the
people that
came from the
cities of
Judah unto
Jerusalem.
\VS{10}Then
read
Baruch in the
book the
words of
Jeremiah in the
house of the
{\ND{
LORD}}, in the
chamber of
Gemariah the
son of
Shaphan the
scribe, in the
higher
court, at the
entry of the
new
gate of the
{\ND{
LORD’s}}
house, in the
ears of all the
people.\FTNT{†}{{\FR 36.10 }
entry: or, door}
\par }{\PP \VS{11}¶ When
Michaiah the
son of
Gemariah, the
son of
Shaphan, had
heard out of the
book all the
words of the
{\ND{
LORD}},
\VS{12}Then he went
down into the
king’s
house, into the
scribe’s
chamber: and, lo, all the
princes
sat there,
{\ADD{even}}
Elishama the
scribe, and
Delaiah the
son of
Shemaiah, and
Elnathan the
son of
Achbor, and
Gemariah the
son of
Shaphan, and
Zedekiah the
son of
Hananiah, and all the
princes.
\VS{13}Then
Michaiah
declared unto them all the
words that he had
heard, when
Baruch
read the
book in the
ears of the
people.
\VS{14}Therefore all the
princes
sent
Jehudi the
son of
Nethaniah, the
son of
Shelemiah, the
son of
Cushi, unto
Baruch,
saying,
Take in thine
hand the
roll wherein thou hast
read in the
ears of the
people, and
come. So
Baruch the
son of
Neriah
took the
roll in his
hand, and
came unto them.
\VS{15}And they
said unto him, Sit
down now, and
read it in our
ears. So
Baruch
read
{\ADD{it}} in their
ears.
\VS{16}Now it came to pass, when they had
heard all the
words, they were
afraid
both
one and
other, and
said unto
Baruch, We will
surely
tell the
king of all these
words.
\VS{17}And they
asked
Baruch,
saying,
Tell us now, How didst thou
write all these
words at his
mouth?
\VS{18}Then
Baruch
answered them, He
pronounced all these
words unto me with his
mouth, and I
wrote
{\ADD{them}} with
ink in the
book.
\VS{19}Then
said the
princes unto
Baruch,
Go,
hide thee, thou and
Jeremiah; and let no
man
know
where ye be.
\par }{\PP \VS{20}¶ And they went
in to the
king into the
court, but they laid
up the
roll in the
chamber of
Elishama the
scribe, and
told all the
words in the
ears of the
king.
\VS{21}So the
king
sent
Jehudi to
fetch the
roll: and he
took it out of
Elishama the
scribe’s
chamber. And
Jehudi
read it in the
ears of the
king, and in the
ears of all the
princes which
stood
beside the
king.
\VS{22}Now the
king
sat in the
winterhouse in the
ninth
month: and
{\ADD{there was a fire}} on the
hearth
burning
before him.
\VS{23}And it came to pass,
{\ADD{that}} when
Jehudi had
read
three or
four
leaves, he
cut it with the
penknife, and
cast
{\ADD{it}} into the
fire that
{\ADD{was}} on the
hearth, until all the
roll was
consumed in the
fire that
{\ADD{was}} on the
hearth.
\VS{24}Yet they were not
afraid, nor
rent their
garments,
{\ADD{neither}} the
king, nor any of his
servants that
heard all these
words.
\VS{25}Nevertheless
Elnathan and
Delaiah and
Gemariah had made
intercession to the
king that he would not
burn the
roll: but he would not
hear them.
\VS{26}But the
king
commanded
Jerahmeel the
son of
Hammelech, and
Seraiah the
son of
Azriel, and
Shelemiah the
son of
Abdeel, to
take
Baruch the
scribe and
Jeremiah the
prophet: but the
{\ND{
LORD}}
hid them.\FTNT{‡}{{\FR 36.26 }
of Hammelech: or, of the king}
\par }{\PP \VS{27}¶ Then the
word of the
{\ND{
LORD}} came to
Jeremiah,
after that the
king had
burned the
roll, and the
words which
Baruch
wrote at the
mouth of
Jeremiah,
saying,
\VS{28}Take thee
again
another
roll, and
write in it all the
former
words that were in the
first
roll, which
Jehoiakim the
king of
Judah hath
burned.
\VS{29}And thou shalt
say to
Jehoiakim
king of
Judah, Thus
saith the
{\ND{
LORD}}; Thou hast
burned this
roll,
saying, Why hast thou
written therein,
saying, The
king of
Babylon shall
certainly
come and
destroy this
land, and shall cause to
cease from thence
man and
beast?
\VS{30}Therefore thus
saith the
{\ND{
LORD}} of
Jehoiakim
king of
Judah; He shall have none to
sit upon the
throne of
David: and his dead
body shall be cast
out in the
day to the
heat, and in the
night to the
frost.
\VS{31}And I will
punish him and his
seed and his
servants for their
iniquity; and I will
bring upon them, and upon the
inhabitants of
Jerusalem, and upon the
men of
Judah, all the
evil that I have
pronounced against them; but they
hearkened not.\FTNT{§}{{\FR 36.31 }
punish: Heb. visit upon}
\par }{\PP \VS{32}¶ Then
took
Jeremiah
another
roll, and
gave it to
Baruch the
scribe, the
son of
Neriah; who
wrote therein from the
mouth of
Jeremiah all the
words of the
book which
Jehoiakim
king of
Judah had
burned in the
fire: and there were
added besides unto them
many
like
words.\FTNT{**}{{\FR 36.32 }
like: Heb. as they}

\par }\Chap{37}{\PP \VerseOne{1}And
king
Zedekiah the
son of
Josiah
reigned instead of
Coniah the
son of
Jehoiakim, whom
Nebuchadrezzar
king of
Babylon made
king in the
land of
Judah.
\VS{2}But neither he, nor his
servants, nor the
people of the
land, did
hearken unto the
words of the
{\ND{
LORD}}, which he
spake
by the
prophet
Jeremiah.\FTNT{*}{{\FR 37.2 }
by…: Heb. by the hand of the prophet}
\VS{3}And
Zedekiah the
king
sent
Jehucal the
son of
Shelemiah and
Zephaniah the
son of
Maaseiah the
priest to the
prophet
Jeremiah,
saying,
Pray now unto the
{\ND{
LORD}} our
God for us.
\VS{4}Now
Jeremiah came
in and went
out
among the
people: for they had not
put him into
prison.
\VS{5}Then
Pharaoh’s
army was come
forth out of
Egypt: and when the
Chaldeans that
besieged
Jerusalem
heard
tidings of them, they
departed from
Jerusalem.
\par }{\PP \VS{6}¶ Then came the
word of the
{\ND{
LORD}} unto the
prophet
Jeremiah,
saying,
\VS{7}Thus
saith the
{\ND{
LORD}}, the
God of
Israel; Thus shall ye
say to the
king of
Judah, that
sent you unto me to
enquire of me; Behold,
Pharaoh’s
army, which is come
forth to
help you, shall
return to
Egypt into their own
land.
\VS{8}And the
Chaldeans shall come
again, and
fight against this
city, and
take it, and
burn it with
fire.
\VS{9}Thus
saith the
{\ND{
LORD}};
Deceive not
yourselves,
saying, The
Chaldeans shall
surely
depart from us: for they shall not
depart.\FTNT{†}{{\FR 37.9 }
yourselves: Heb. your souls}
\VS{10}For though ye had
smitten the whole
army of the
Chaldeans that
fight against you, and there
remained
{\ADD{but}}
wounded
men among them,
{\ADD{yet}} should they rise
up every
man in his
tent, and
burn this
city with
fire.\FTNT{‡}{{\FR 37.10 }
wounded: Heb. thrust through}
\par }{\PP \VS{11}¶ And it came to pass, that when the
army of the
Chaldeans was broken
up from
Jerusalem for
fear of
Pharaoh’s
army,\FTNT{§}{{\FR 37.11 }
broken…: Heb. made to ascend}
\VS{12}Then
Jeremiah went
forth out of
Jerusalem to
go into the
land of
Benjamin, to
separate himself thence in the
midst of the
people.\FTNT{**}{{\FR 37.12 }
separate…: or, to slip away from thence in the midst of the people}
\VS{13}And when he was in the
gate of
Benjamin, a
captain of the
ward
{\ADD{was}} there, whose
name
{\ADD{was}}
Irijah, the
son of
Shelemiah, the
son of
Hananiah; and he
took
Jeremiah the
prophet,
saying, Thou fallest
away to the
Chaldeans.
\VS{14}Then
said
Jeremiah,
{\ADD{It is}}
false; I fall not
away to the
Chaldeans. But he
hearkened not to him: so
Irijah
took
Jeremiah, and
brought him to the
princes.\FTNT{††}{{\FR 37.14 }
false: Heb. falsehood, or, a lie}
\VS{15}Wherefore the
princes were
wroth with
Jeremiah, and
smote him, and
put him in
prison in the
house of
Jonathan the
scribe: for they had
made that the
prison.
\par }{\PP \VS{16}¶ When
Jeremiah was
entered into the
dungeon, and into the
cabins, and
Jeremiah had
remained there
many
days;\FTNT{‡‡}{{\FR 37.16 }
cabins: or, cells}
\VS{17}Then
Zedekiah the
king
sent, and took him
out: and the
king
asked him
secretly in his
house, and
said, Is there
{\ADD{any}}
word from the
{\ND{
LORD}}? And
Jeremiah
said, There
is: for,
said he, thou shalt be
delivered into the
hand of the
king of
Babylon.
\VS{18}Moreover
Jeremiah
said unto
king
Zedekiah, What have I
offended against thee, or against thy
servants, or against this
people, that ye have
put me in
prison?
\VS{19}Where
{\ADD{are}}
now your
prophets which
prophesied unto you,
saying, The
king of
Babylon shall not
come against you, nor against this
land?
\VS{20}Therefore
hear now, I pray thee, O my
lord the
king: let my
supplication, I pray thee, be
accepted
before thee; that thou cause me not to
return to the
house of
Jonathan the
scribe, lest I
die there.\FTNT{§§}{{\FR 37.20 }
let…: Heb. let my supplication fall}
\VS{21}Then
Zedekiah the
king
commanded that they should
commit
Jeremiah into the
court of the
prison, and that they should
give him
daily a
piece of
bread out of the
bakers’
street, until all the
bread in the
city were
spent. Thus
Jeremiah
remained in the
court of the
prison.

\par }\Chap{38}{\PP \VerseOne{1}Then
Shephatiah the
son of
Mattan, and
Gedaliah the
son of
Pashur, and
Jucal the
son of
Shelemiah, and
Pashur the
son of
Malchiah,
heard the
words that
Jeremiah had
spoken unto all the
people,
saying,
\VS{2}Thus
saith the
{\ND{
LORD}}, He that
remaineth in this
city shall
die by the
sword, by the
famine, and by the
pestilence: but he that goeth
forth to the
Chaldeans shall
live; for he shall have his
life for a
prey, and shall
live.
\VS{3}Thus
saith the
{\ND{
LORD}}, This
city shall
surely be
given into the
hand of the
king of
Babylon’s
army, which shall
take it.
\VS{4}Therefore the
princes
said unto the
king, We beseech thee, let this
man be put to
death: for
thus he
weakeneth the
hands of the
men of
war that
remain in this
city, and the
hands of all the
people, in
speaking such
words unto them: for this
man
seeketh not the
welfare of this
people, but the
hurt.\FTNT{*}{{\FR 38.4 }
welfare: Heb. peace}
\VS{5}Then
Zedekiah the
king
said, Behold, he
{\ADD{is}} in your
hand: for the
king
{\ADD{is}} not
{\ADD{he that}}
can do
{\ADD{any}}
thing against you.
\VS{6}Then
took they
Jeremiah, and
cast him into the
dungeon of
Malchiah the
son of
Hammelech, that
{\ADD{was}} in the
court of the
prison: and they let
down
Jeremiah with
cords. And in the
dungeon
{\ADD{there was}} no
water, but
mire: so
Jeremiah
sunk in the
mire.\FTNT{†}{{\FR 38.6 }
of Hammelech: or, of the king}
\par }{\PP \VS{7}¶ Now when
Ebed-melech the
Ethiopian,
one of the
eunuchs which was in the
king’s
house,
heard that they had
put
Jeremiah in the
dungeon; the
king then
sitting in the
gate of
Benjamin;
\VS{8}Ebed-melech went
forth out of the
king’s
house, and
spake to the
king,
saying,
\VS{9}My
lord the
king, these
men have done
evil in all that they have
done to
Jeremiah the
prophet, whom they have
cast into the
dungeon; and he is like to
die
for
hunger in the
place where he is: for
{\ADD{there is}} no more
bread in the
city.\FTNT{‡}{{\FR 38.9 }
he is like…: Heb. he will die}
\VS{10}Then the
king
commanded
Ebed-melech the
Ethiopian,
saying,
Take from hence
thirty
men with
thee, and take
up
Jeremiah the
prophet out of the
dungeon, before he
die.\FTNT{§}{{\FR 38.10 }
with thee: Heb. in thine hand}
\VS{11}So
Ebed-melech
took the
men with
him, and
went into the
house of the
king under the
treasury, and
took thence
old cast
clouts and
old rotten
rags, and let them
down by
cords into the
dungeon to
Jeremiah.
\VS{12}And
Ebed-melech the
Ethiopian
said unto
Jeremiah,
Put now
{\ADD{these}}
old cast
clouts and rotten
rags under thine
armholes under the
cords. And
Jeremiah
did so.
\VS{13}So they drew
up
Jeremiah with
cords, and took him
up out of the
dungeon: and
Jeremiah
remained in the
court of the
prison.
\par }{\PP \VS{14}¶ Then
Zedekiah the
king
sent, and
took
Jeremiah the
prophet unto him into the
third
entry that
{\ADD{is}} in the
house of the
{\ND{
LORD}}: and the
king
said unto
Jeremiah, I will
ask thee a
thing;
hide nothing from me.\FTNT{**}{{\FR 38.14 }
third: or, principal}
\VS{15}Then
Jeremiah
said unto
Zedekiah, If I
declare
{\ADD{it}} unto thee, wilt thou not
surely put me to
death? and if I give thee
counsel, wilt thou not
hearken unto me?
\VS{16}So
Zedekiah the
king
sware
secretly unto
Jeremiah,
saying,
{\ADD{As}} the
{\ND{
LORD}}
liveth, that
made us this
soul, I will not put thee to
death, neither will I
give thee into the
hand of these
men that
seek thy
life.
\VS{17}Then
said
Jeremiah unto
Zedekiah, Thus
saith the
{\ND{
LORD}}, the
God of
hosts, the
God of
Israel; If thou wilt
assuredly go
forth unto the
king of
Babylon’s
princes, then thy
soul shall
live, and this
city shall not be
burned with
fire; and thou shalt
live, and thine
house:
\VS{18}But if thou wilt not go
forth to the
king of
Babylon’s
princes, then shall this
city be
given into the
hand of the
Chaldeans, and they shall
burn it with
fire, and thou shalt not escape
out of their
hand.
\VS{19}And
Zedekiah the
king
said unto
Jeremiah, I am
afraid of the
Jews that are
fallen to the
Chaldeans, lest they
deliver me into their
hand, and they
mock me.
\VS{20}But
Jeremiah
said, They shall not
deliver
{\ADD{thee}}.
Obey, I beseech thee, the
voice of the
{\ND{
LORD}}, which I
speak unto thee: so it shall be
well unto thee, and thy
soul shall
live.
\VS{21}But if thou
refuse to go
forth, this
{\ADD{is}} the
word that the
{\ND{
LORD}} hath
shewed me:
\VS{22}And, behold, all the
women that are
left in the
king of
Judah’s
house
{\ADD{shall be}} brought
forth to the
king of
Babylon’s
princes, and those
{\ADD{women}} shall
say, Thy
friends have set thee
on, and have
prevailed against thee: thy
feet are
sunk in the
mire,
{\ADD{and}} they are turned
away
back.\FTNT{††}{{\FR 38.22 }
Thy friends: Heb. Men of thy peace}
\VS{23}So they shall bring
out all thy
wives and thy
children to the
Chaldeans: and thou shalt not escape
out of their
hand, but shalt be
taken by the
hand of the
king of
Babylon: and thou shalt cause this
city to be
burned with
fire.\FTNT{‡‡}{{\FR 38.23 }
thou shalt cause…: Heb. thou shalt burn, etc}
\par }{\PP \VS{24}¶ Then
said
Zedekiah unto
Jeremiah, Let no
man
know of these
words, and thou shalt not
die.
\VS{25}But if the
princes
hear that I have
talked with thee, and they
come unto thee, and
say unto thee,
Declare unto us now what thou hast
said unto the
king,
hide it not from us, and we will not put thee to
death; also what the
king
said unto thee:
\VS{26}Then thou shalt
say unto them, I
presented my
supplication
before the
king, that he would not cause me to
return to
Jonathan’s
house, to
die there.
\VS{27}Then
came all the
princes unto
Jeremiah, and
asked him: and he
told them according to all these
words that the
king had
commanded. So they left off
speaking with him; for the
matter was not
perceived.\FTNT{§§}{{\FR 38.27 }
they…: Heb. they were silent from him}
\VS{28}So
Jeremiah
abode in the
court of the
prison until the
day that
Jerusalem was
taken: and he was
{\ADD{there}} when
Jerusalem was
taken.

\par }\Chap{39}{\PP \VerseOne{1}In the
ninth
year of
Zedekiah
king of
Judah, in the
tenth
month,
came
Nebuchadrezzar
king of
Babylon and all his
army against
Jerusalem, and they
besieged it.
\VS{2}{\ADD{And}} in the
eleventh
year of
Zedekiah, in the
fourth
month, the
ninth
{\ADD{day}} of the
month, the
city was broken
up.
\VS{3}And all the
princes of the
king of
Babylon came
in, and
sat in the
middle
gate,
{\ADD{even}}
Nergal-sharezer,
Samgar-nebo,
Sarsechim,
Rab-saris,
Nergal-sharezer,
Rab-mag, with all the
residue of the
princes of the
king of
Babylon.
\par }{\PP \VS{4}¶ And it came to pass,
{\ADD{that}} when
Zedekiah the
king of
Judah
saw them, and all the
men of
war, then they
fled, and went
forth out of the
city by
night, by the
way of the
king’s
garden, by the
gate betwixt the two
walls: and he went
out the
way of the
plain.
\VS{5}But the
Chaldeans’
army
pursued
after them, and
overtook
Zedekiah in the
plains of
Jericho: and when they had
taken him, they brought him
up to
Nebuchadnezzar
king of
Babylon to
Riblah in the
land of
Hamath, where he
gave
judgment upon him.\FTNT{*}{{\FR 39.5 }
gave…: Heb. spake with him judgments}
\VS{6}Then the
king of
Babylon
slew the
sons of
Zedekiah in
Riblah before his
eyes: also the
king of
Babylon
slew all the
nobles of
Judah.
\VS{7}Moreover he put
out
Zedekiah’s
eyes, and
bound him with
chains, to
carry him to
Babylon.\FTNT{†}{{\FR 39.7 }
with…: Heb. with two brasen chains, or, fetters}
\par }{\PP \VS{8}¶ And the
Chaldeans
burned the
king’s
house, and the
houses of the
people, with
fire, and brake
down the
walls of
Jerusalem.
\VS{9}Then
Nebuzar-adan the
captain of the
guard carried away
captive into
Babylon the
remnant of the
people that
remained in the
city, and those that fell
away, that
fell to him, with the
rest of the
people that
remained.\FTNT{‡}{{\FR 39.9 }
captain…: or, chief marshal: Heb. chief of the executioners, or, slaughtermen}
\VS{10}But
Nebuzar-adan the
captain of the
guard
left of the
poor of the
people, which had
nothing, in the
land of
Judah, and
gave them
vineyards and
fields at the same
time.\FTNT{§}{{\FR 39.10 }
captain…: or, chief marshal: Heb. chief of the executioners, or, slaughtermen}\FTNT{**}{{\FR 39.10 }
at the…: Heb. in that day}
\par }{\PP \VS{11}¶ Now
Nebuchadrezzar
king of
Babylon gave
charge concerning
Jeremiah
to
Nebuzar-adan the
captain of the
guard,
saying,\FTNT{††}{{\FR 39.11 }
to: Heb. by the hand of}\FTNT{‡‡}{{\FR 39.11 }
captain…: or, chief marshal: Heb. chief of the executioners, or, slaughtermen}
\VS{12}Take him, and look
well to him, and
do him no
harm; but
do unto him even as he shall
say unto thee.\FTNT{§§}{{\FR 39.12 }
look…: Heb. set thine eyes upon him}
\VS{13}So
Nebuzar-adan the
captain of the
guard
sent, and
Nebushasban,
Rab-saris, and
Nergal-sharezer,
Rab-mag, and all the
king of
Babylon’s
princes;\FTNT{***}{{\FR 39.13 }
captain…: or, chief marshal: Heb. chief of the executioners, or, slaughtermen}
\VS{14}Even they
sent, and
took
Jeremiah out of the
court of the
prison, and
committed him unto
Gedaliah the
son of
Ahikam the
son of
Shaphan, that he should
carry him
home: so he
dwelt
among the
people.
\par }{\PP \VS{15}¶ Now the
word of the
{\ND{
LORD}} came unto
Jeremiah, while he was shut
up in the
court of the
prison,
saying,
\VS{16}Go and
speak to
Ebed-melech the
Ethiopian,
saying, Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; Behold, I will
bring my
words upon this
city for
evil, and not for
good; and they shall be
{\ADD{accomplished}} in that
day
before thee.
\VS{17}But I will
deliver thee in that
day,
saith the
{\ND{
LORD}}: and thou shalt not be
given into the
hand of the
men of
whom thou
{\ADD{art}}
afraid.
\VS{18}For I will
surely
deliver thee, and thou shalt not
fall by the
sword, but thy
life shall be for a
prey unto thee: because thou hast put thy
trust in me,
saith the
{\ND{
LORD}}.

\par }\Chap{40}{\PP \VerseOne{1}The
word that came to
Jeremiah from the
{\ND{
LORD}},
after that
Nebuzar-adan the
captain of the
guard had let him
go from
Ramah, when he had
taken him being
bound in
chains
among all that were carried away
captive of
Jerusalem and
Judah, which were carried away
captive unto
Babylon.\FTNT{*}{{\FR 40.1 }
chains: or, manicles}
\VS{2}And the
captain of the
guard
took
Jeremiah, and
said unto him, The
{\ND{
LORD}} thy
God hath
pronounced this
evil upon this
place.
\VS{3}Now the
{\ND{
LORD}} hath
brought
{\ADD{it}}, and
done according as he hath
said: because ye have
sinned against the
{\ND{
LORD}}, and have not
obeyed his
voice, therefore this
thing is come upon you.
\VS{4}And now, behold, I
loose thee this
day from the
chains which
{\ADD{were}} upon thine
hand. If it seem
good unto thee to
come with me into
Babylon,
come; and I will look
well unto thee: but if it seem
ill unto thee to
come with me into
Babylon,
forbear:
behold, all the
land
{\ADD{is}}
before thee: whither it
seemeth
good and
convenient for thee to
go, thither
go.\FTNT{†}{{\FR 40.4 }
were…: or, are upon thine hand}\FTNT{‡}{{\FR 40.4 }
I will…: Heb. I will set mine eye upon thee}
\VS{5}Now while he was not yet gone
back,
{\ADD{he said}}, Go
back also to
Gedaliah the
son of
Ahikam the
son of
Shaphan, whom the
king of
Babylon hath made
governor over the
cities of
Judah, and
dwell with him
among the
people: or
go wheresoever it seemeth
convenient unto thee to
go. So the
captain of the
guard
gave him
victuals and a
reward, and let him
go.
\VS{6}Then
went
Jeremiah unto
Gedaliah the
son of
Ahikam to
Mizpah; and
dwelt with him
among the
people that were
left in the
land.
\par }{\PP \VS{7}¶ Now when all the
captains of the
forces which
{\ADD{were}} in the
fields,
{\ADD{even}} they and their
men,
heard that the
king of
Babylon had made
Gedaliah the
son of
Ahikam
governor in the
land, and had
committed unto him
men, and
women, and
children, and of the
poor of the
land, of them that were not carried away
captive to
Babylon;
\VS{8}Then they
came to
Gedaliah to
Mizpah, even
Ishmael the
son of
Nethaniah, and
Johanan and
Jonathan the
sons of
Kareah, and
Seraiah the
son of
Tanhumeth, and the
sons of
Ephai the
Netophathite, and
Jezaniah the
son of a
Maachathite, they and their
men.
\VS{9}And
Gedaliah the
son of
Ahikam the
son of
Shaphan
sware unto them and to their
men,
saying,
Fear not to
serve the
Chaldeans: dwell
in the
land, and
serve the
king of
Babylon, and it shall be
well with you.
\VS{10}As for me, behold, I will
dwell at
Mizpah to
serve the
Chaldeans, which will
come unto us: but ye,
gather ye
wine, and summer
fruits, and
oil, and
put
{\ADD{them}} in your
vessels, and
dwell in your
cities that ye have
taken.\FTNT{§}{{\FR 40.10 }
to serve: Heb. to stand before}
\VS{11}Likewise when all the
Jews that
{\ADD{were}} in
Moab, and among the
Ammonites, and in
Edom, and that
{\ADD{were}} in all the
countries,
heard that the
king of
Babylon had
left a
remnant of
Judah, and that he had
set over them
Gedaliah the
son of
Ahikam the
son of
Shaphan;
\VS{12}Even all the
Jews returned
out of all
places whither they were
driven, and
came to the
land of
Judah, to
Gedaliah, unto
Mizpah, and
gathered
wine and summer
fruits
very
much.
\par }{\PP \VS{13}¶ Moreover
Johanan the
son of
Kareah, and all the
captains of the
forces that
{\ADD{were}} in the
fields,
came to
Gedaliah to
Mizpah,
\VS{14}And
said unto him, Dost thou
certainly
know that
Baalis the
king of the
Ammonites hath
sent
Ishmael the
son of
Nethaniah to
slay thee? But
Gedaliah the
son of
Ahikam
believed them not.\FTNT{**}{{\FR 40.14 }
to slay…: Heb. to strike thee in soul?}
\VS{15}Then
Johanan the
son of
Kareah
spake to
Gedaliah in
Mizpah
secretly,
saying, Let me
go, I pray thee, and I will
slay
Ishmael the
son of
Nethaniah, and no
man shall
know
{\ADD{it:}} wherefore should he
slay thee, that all the
Jews which are
gathered unto thee should be
scattered, and the
remnant in
Judah
perish?
\VS{16}But
Gedaliah the
son of
Ahikam
said unto
Johanan the
son of
Kareah, Thou shalt not
do this
thing: for thou
speakest
falsely of
Ishmael.

\par }\Chap{41}{\PP \VerseOne{1}Now it came to pass in the
seventh
month,
{\ADD{that}}
Ishmael the
son of
Nethaniah the
son of
Elishama, of the
seed
royal, and the
princes of the
king, even
ten
men with him,
came unto
Gedaliah the
son of
Ahikam to
Mizpah; and there they did
eat
bread
together in
Mizpah.
\VS{2}Then
arose
Ishmael the
son of
Nethaniah, and the
ten
men that were with him, and
smote
Gedaliah the
son of
Ahikam the
son of
Shaphan with the
sword, and
slew him, whom the
king of
Babylon had made
governor over the
land.
\VS{3}Ishmael also
slew all the
Jews that were with him,
{\ADD{even}} with
Gedaliah, at
Mizpah, and the
Chaldeans that were
found there,
{\ADD{and}} the
men of
war.
\VS{4}And it came to pass the
second
day after he had
slain
Gedaliah, and no
man
knew
{\ADD{it}},
\VS{5}That there
came
certain from
Shechem, from
Shiloh, and from
Samaria,
{\ADD{even}}
fourscore
men, having their
beards
shaven, and their
clothes
rent, and having
cut themselves, with
offerings and
incense in their
hand, to
bring
{\ADD{them}} to the
house of the
{\ND{
LORD}}.
\VS{6}And
Ishmael the
son of
Nethaniah went
forth from
Mizpah to
meet them,
weeping all
along as he
went: and it came to pass, as he
met them, he
said unto them,
Come to
Gedaliah the
son of
Ahikam.\FTNT{*}{{\FR 41.6 }
weeping…: Heb. in going and weeping}
\VS{7}And it was
{\ADD{so}}, when they
came into the
midst of the
city, that
Ishmael the
son of
Nethaniah
slew them,
{\ADD{and cast them}} into the
midst of the
pit, he, and the
men that
{\ADD{were}} with him.
\VS{8}But
ten
men were
found among them that
said unto
Ishmael,
Slay us not: for we
have
treasures in the
field, of
wheat, and of
barley, and of
oil, and of
honey. So he
forbare, and
slew them not
among their
brethren.
\VS{9}Now the
pit wherein
Ishmael had
cast all the dead
bodies of the
men, whom he had
slain
because of
Gedaliah,
{\ADD{was}} it which
Asa the
king had
made for
fear of
Baasha
king of
Israel:
{\ADD{and}}
Ishmael the
son of
Nethaniah
filled it with
{\ADD{them that were}}
slain.\FTNT{†}{{\FR 41.9 }
because…: or, near Gedaliah: Heb. by the hand, or, by the side of Gedaliah}
\VS{10}Then
Ishmael carried away
captive all the
residue of the
people that
{\ADD{were}} in
Mizpah,
{\ADD{even}} the
king’s
daughters, and all the
people that
remained in
Mizpah, whom
Nebuzar-adan the
captain of the
guard had
committed to
Gedaliah the
son of
Ahikam: and
Ishmael the
son of
Nethaniah carried them away
captive, and
departed to go
over to the
Ammonites.
\par }{\PP \VS{11}¶ But when
Johanan the
son of
Kareah, and all the
captains of the
forces that
{\ADD{were}} with him,
heard of all the
evil that
Ishmael the
son of
Nethaniah had
done,
\VS{12}Then they
took all the
men, and
went to
fight with
Ishmael the
son of
Nethaniah, and
found him by the
great
waters that
{\ADD{are}} in
Gibeon.
\VS{13}Now it came to pass,
{\ADD{that}} when all the
people which
{\ADD{were}} with
Ishmael
saw
Johanan the
son of
Kareah, and all the
captains of the
forces that
{\ADD{were}} with him, then they were
glad.
\VS{14}So all the
people that
Ishmael had carried away
captive from
Mizpah cast
about and
returned, and
went unto
Johanan the
son of
Kareah.
\VS{15}But
Ishmael the
son of
Nethaniah
escaped
from
Johanan with
eight
men, and
went to the
Ammonites.
\VS{16}Then
took
Johanan the
son of
Kareah, and all the
captains of the
forces that
{\ADD{were}} with him, all the
remnant of the
people whom he had
recovered from
Ishmael the
son of
Nethaniah, from
Mizpah,
after
{\ADD{that}} he had
slain
Gedaliah the
son of
Ahikam,
{\ADD{even}}
mighty
men of
war, and the
women, and the
children, and the
eunuchs, whom he had brought
again from
Gibeon:
\VS{17}And they
departed, and dwelt
in the
habitation of
Chimham, which is
by
Beth-lehem, to
go to
enter into
Egypt,
\VS{18}Because of the
Chaldeans: for they were
afraid of them, because
Ishmael the
son of
Nethaniah had
slain
Gedaliah the
son of
Ahikam, whom the
king of
Babylon made
governor in the
land.

\par }\Chap{42}{\PP \VerseOne{1}Then all the
captains of the
forces, and
Johanan the
son of
Kareah, and
Jezaniah the
son of
Hoshaiah, and all the
people from the
least even unto the
greatest, came
near,
\VS{2}And
said unto
Jeremiah the
prophet, Let, we beseech thee, our
supplication be
accepted
before thee, and
pray for us unto the
{\ND{
LORD}} thy
God,
{\ADD{even}} for all this
remnant; (for we are
left
{\ADD{but}} a
few of
many, as thine
eyes do
behold us:)\FTNT{*}{{\FR 42.2 }
Let…: or, Let our supplication fall before thee}
\VS{3}That the
{\ND{
LORD}} thy
God may
shew us the
way wherein we may
walk, and the
thing that we may
do.
\VS{4}Then
Jeremiah the
prophet
said unto them, I have
heard
{\ADD{you;}} behold, I will
pray unto the
{\ND{
LORD}} your
God according to your
words; and it shall come to pass,
{\ADD{that}} whatsoever
thing the
{\ND{
LORD}} shall
answer you, I will
declare
{\ADD{it}} unto you; I will
keep
nothing
back from you.
\VS{5}Then they
said to
Jeremiah, The
{\ND{
LORD}} be a
true and
faithful
witness between us, if we
do not even according to all
things for the which the
{\ND{
LORD}} thy
God shall
send thee to us.
\VS{6}Whether
{\ADD{it be}}
good, or whether
{\ADD{it be}}
evil, we will
obey the
voice of the
{\ND{
LORD}} our
God, to whom
we
send thee; that it may be
well with us, when we
obey the
voice of the
{\ND{
LORD}} our
God.
\par }{\PP \VS{7}¶ And it came to pass
after
ten
days, that the
word of the
{\ND{
LORD}} came unto
Jeremiah.
\VS{8}Then
called he
Johanan the
son of
Kareah, and all the
captains of the
forces which
{\ADD{were}} with him, and all the
people from the
least even to the
greatest,
\VS{9}And
said unto them, Thus
saith the
{\ND{
LORD}}, the
God of
Israel, unto whom ye
sent me to
present your
supplication
before him;
\VS{10}If ye will
still
abide in this
land, then will I
build you, and not pull
{\ADD{you}}
down, and I will
plant you, and not pluck
{\ADD{you}}
up: for I
repent me of the
evil that I have
done unto you.
\VS{11}Be not
afraid
of the
king of
Babylon, of
whom ye are
afraid; be not
afraid of him,
saith the
{\ND{
LORD}}: for I
{\ADD{am}} with you to
save you, and to
deliver you from his
hand.
\VS{12}And I will
shew
mercies unto you, that he may have
mercy upon you, and cause you to
return to your own
land.
\par }{\PP \VS{13}¶ But if ye
say, We will not
dwell in this
land, neither
obey the
voice of the
{\ND{
LORD}} your
God,
\VS{14}Saying, No; but we will
go into the
land of
Egypt, where we shall
see no
war, nor
hear the
sound of the
trumpet, nor have
hunger of
bread; and there will we
dwell:
\VS{15}And now therefore
hear the
word of the
{\ND{
LORD}}, ye
remnant of
Judah; Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; If ye
wholly
set your
faces to
enter into
Egypt, and
go to
sojourn there;
\VS{16}Then it shall come to pass,
{\ADD{that}} the
sword, which ye
feared, shall
overtake you there in the
land of
Egypt, and the
famine, whereof ye were
afraid, shall follow
close
after you there in
Egypt; and there ye shall
die.\FTNT{†}{{\FR 42.16 }
shall follow…: Heb. shall cleave after you}
\VS{17}So shall it be with all the
men that
set their
faces to
go into
Egypt to
sojourn there; they shall
die by the
sword, by the
famine, and by the
pestilence: and none of them shall
remain or
escape
from the
evil that I will
bring upon them.\FTNT{‡}{{\FR 42.17 }
So…: Heb. So shall all the men be}
\VS{18}For thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; As mine
anger and my
fury hath been poured
forth upon the
inhabitants of
Jerusalem; so shall my
fury be poured
forth upon you, when ye shall
enter into
Egypt: and ye shall be an
execration, and an
astonishment, and a
curse, and a
reproach; and ye shall
see this
place no more.
\par }{\PP \VS{19}¶ The
{\ND{
LORD}} hath
said concerning you, O ye
remnant of
Judah;
Go ye not into
Egypt:
know
certainly that I have
admonished you this
day.\FTNT{§}{{\FR 42.19 }
admonished…: Heb. testified against you}
\VS{20}For ye
dissembled in your
hearts, when ye
sent me unto the
{\ND{
LORD}} your
God,
saying,
Pray for us unto the
{\ND{
LORD}} our
God; and according unto all that the
{\ND{
LORD}} our
God shall
say, so
declare unto us, and we will
do
{\ADD{it}}.\FTNT{**}{{\FR 42.20 }
ye dissembled…: or, ye have used deceit against your souls}
\VS{21}And
{\ADD{now}} I have this
day
declared
{\ADD{it}} to you; but ye have not
obeyed the
voice of the
{\ND{
LORD}} your
God, nor any
{\ADD{thing}} for the which he hath
sent me unto you.
\VS{22}Now therefore
know
certainly that ye shall
die by the
sword, by the
famine, and by the
pestilence, in the
place whither ye
desire to
go
{\ADD{and}} to
sojourn.\FTNT{††}{{\FR 42.22 }
to go…: or, to go to sojourn}

\par }\Chap{43}{\PP \VerseOne{1}And it came to pass,
{\ADD{that}} when
Jeremiah had made an
end of
speaking unto all the
people all the
words of the
{\ND{
LORD}} their
God, for which the
{\ND{
LORD}} their
God had
sent him to them,
{\ADD{even}} all these
words,
\VS{2}Then
spake
Azariah the
son of
Hoshaiah, and
Johanan the
son of
Kareah, and all the
proud
men,
saying unto
Jeremiah, Thou
speakest
falsely: the
{\ND{
LORD}} our
God hath not
sent thee to
say,
Go not into
Egypt to
sojourn there:
\VS{3}But
Baruch the
son of
Neriah setteth thee
on against us, for
to
deliver us into the
hand of the
Chaldeans, that they might put us to
death, and carry us away
captives into
Babylon.
\VS{4}So
Johanan the
son of
Kareah, and all the
captains of the
forces, and all the
people,
obeyed not the
voice of the
{\ND{
LORD}}, to
dwell in the
land of
Judah.
\VS{5}But
Johanan the
son of
Kareah, and all the
captains of the
forces,
took all the
remnant of
Judah, that were
returned from all
nations, whither they had been
driven, to
dwell in the
land of
Judah;
\VS{6}{\ADD{Even}}
men, and
women, and
children, and the
king’s
daughters, and every
person that
Nebuzar-adan the
captain of the
guard had
left with
Gedaliah the
son of
Ahikam the
son of
Shaphan, and
Jeremiah the
prophet, and
Baruch the
son of
Neriah.
\VS{7}So they
came into the
land of
Egypt: for they
obeyed not the
voice of the
{\ND{
LORD}}: thus
came they
{\ADD{even}} to
Tahpanhes.
\par }{\PP \VS{8}¶ Then came the
word of the
{\ND{
LORD}} unto
Jeremiah in
Tahpanhes,
saying,
\VS{9}Take
great
stones in thine
hand, and
hide them in the
clay in the
brickkiln, which
{\ADD{is}} at the
entry of
Pharaoh’s
house in
Tahpanhes, in the
sight of the
men of
Judah;
\VS{10}And
say unto them, Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; Behold, I will
send and
take
Nebuchadrezzar the
king of
Babylon, my
servant, and will
set his
throne
upon these
stones that I have
hid; and he shall
spread his royal
pavilion over them.
\VS{11}And when he
cometh, he shall
smite the
land of
Egypt,
{\ADD{and deliver}} such
{\ADD{as are}} for
death to
death; and such
{\ADD{as are}} for
captivity to
captivity; and such
{\ADD{as are}} for the
sword to the
sword.
\VS{12}And I will
kindle a
fire in the
houses of the
gods of
Egypt; and he shall
burn them, and carry them away
captives: and he shall
array himself with the
land of
Egypt, as a
shepherd putteth
on his
garment; and he shall go
forth from thence in
peace.
\VS{13}He shall
break also the
images of
Beth-shemesh, that
{\ADD{is}} in the
land of
Egypt; and the
houses of the
gods of the
Egyptians shall he
burn with
fire.\FTNT{*}{{\FR 43.13 }
images: Heb. statues, or, standing images}\FTNT{†}{{\FR 43.13 }
Beth-shemesh: or, The house of the sun}

\par }\Chap{44}{\PP \VerseOne{1}The
word that came to
Jeremiah concerning all the
Jews which
dwell in the
land of
Egypt, which
dwell at
Migdol, and at
Tahpanhes, and at
Noph, and in the
country of
Pathros,
saying,
\VS{2}Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; Ye have
seen all the
evil that I have
brought upon
Jerusalem, and upon all the
cities of
Judah; and, behold, this
day they
{\ADD{are}} a
desolation, and no man
dwelleth therein,
\VS{3}Because of their
wickedness which they have
committed to provoke me to
anger, in that they
went to burn
incense,
{\ADD{and}} to
serve
other
gods, whom they
knew not,
{\ADD{neither}} they, ye, nor your
fathers.
\VS{4}Howbeit I
sent unto you all my
servants the
prophets, rising
early and
sending
{\ADD{them}},
saying, Oh,
do not this
abominable
thing that I
hate.
\VS{5}But they
hearkened not, nor
inclined their
ear to
turn from their
wickedness, to burn no
incense unto
other
gods.
\VS{6}Wherefore my
fury and mine
anger was poured
forth, and was
kindled in the
cities of
Judah and in the
streets of
Jerusalem; and they are
wasted
{\ADD{and}}
desolate, as at this
day.
\VS{7}Therefore now thus
saith the
{\ND{
LORD}}, the
God of
hosts, the
God of
Israel; Wherefore
commit ye
{\ADD{this}}
great
evil against your
souls, to cut
off from you
man and
woman,
child and
suckling,
out of
Judah, to
leave you none to
remain;\FTNT{*}{{\FR 44.7 }
out…: Heb. out of the midst of Judah}
\VS{8}In that ye provoke me unto
wrath with the
works of your
hands, burning
incense unto
other
gods in the
land of
Egypt, whither ye be
gone to
dwell, that ye might cut yourselves
off, and that ye might be a
curse and a
reproach among all the
nations of the
earth?
\VS{9}Have ye
forgotten the
wickedness of your
fathers, and the
wickedness of the
kings of
Judah, and the
wickedness of their
wives, and your own
wickedness, and the
wickedness of your
wives, which they have
committed in the
land of
Judah, and in the
streets of
Jerusalem?\FTNT{†}{{\FR 44.9 }
wickedness…: Heb. wickednesses, or, punishments, etc}
\VS{10}They are not
humbled
{\ADD{even}} unto this
day, neither have they
feared, nor
walked in my
law, nor in my
statutes, that I
set
before you and
before your
fathers.\FTNT{‡}{{\FR 44.10 }
humbled: Heb. contrite}
\par }{\PP \VS{11}¶ Therefore thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; Behold, I will
set my
face against you for
evil, and to cut
off all
Judah.
\VS{12}And I will
take the
remnant of
Judah, that have
set their
faces to
go into the
land of
Egypt to
sojourn there, and they shall all be
consumed,
{\ADD{and}}
fall in the
land of
Egypt; they shall
{\ADD{even}} be
consumed by the
sword
{\ADD{and}} by the
famine: they shall
die, from the
least even unto the
greatest, by the
sword and by the
famine: and they shall be an
execration,
{\ADD{and}} an
astonishment, and a
curse, and a
reproach.
\VS{13}For I will
punish them that
dwell in the
land of
Egypt, as I have
punished
Jerusalem, by the
sword, by the
famine, and by the
pestilence:
\VS{14}So that none of the
remnant of
Judah, which are
gone into the
land of
Egypt to
sojourn there, shall
escape or
remain, that they should
return into the
land of
Judah, to the which they
have a
desire to
return to
dwell there: for none shall
return but such as shall
escape.\FTNT{§}{{\FR 44.14 }
have…: Heb. lift up their soul}
\par }{\PP \VS{15}¶ Then all the
men which
knew that their
wives had burned
incense unto
other
gods, and all the
women that stood
by, a
great
multitude, even all the
people that
dwelt in the
land of
Egypt, in
Pathros,
answered
Jeremiah,
saying,
\VS{16}{\ADD{As for}} the
word that thou hast
spoken unto us in the
name of the
{\ND{
LORD}}, we will not
hearken unto thee.
\VS{17}But we will
certainly
do whatsoever
thing goeth
forth out of our own
mouth, to burn
incense unto the
queen of
heaven, and to pour
out drink
offerings unto her, as we have
done, we, and our
fathers, our
kings, and our
princes, in the
cities of
Judah, and in the
streets of
Jerusalem: for
{\ADD{then}} had we
plenty of
victuals, and were
well, and
saw no
evil.\FTNT{**}{{\FR 44.17 }
queen…: or, frame of heaven}\FTNT{††}{{\FR 44.17 }
victuals: Heb. bread}
\VS{18}But since we left
off to burn
incense to the
queen of
heaven, and to pour
out drink
offerings unto her, we have
wanted all
{\ADD{things}}, and have been
consumed by the
sword and by the
famine.
\VS{19}And when we burned
incense to the
queen of
heaven, and poured
out drink
offerings unto her, did we
make her
cakes to
worship her, and pour
out drink
offerings unto her,
without our
men?\FTNT{‡‡}{{\FR 44.19 }
men: or, husbands?}
\par }{\PP \VS{20}¶ Then
Jeremiah
said unto all the
people, to the
men, and to the
women, and to all the
people which had given him
{\ADD{that}}
answer,
saying,
\VS{21}The
incense that ye
burned in the
cities of
Judah, and in the
streets of
Jerusalem, ye, and your
fathers, your
kings, and your
princes, and the
people of the
land, did not the
{\ND{
LORD}}
remember them, and
came it
{\ADD{not}} into his
mind?
\VS{22}So that the
{\ND{
LORD}}
could no longer
bear,
because of the
evil of your
doings,
{\ADD{and}}
because of the
abominations which ye have
committed; therefore is your
land a
desolation, and an
astonishment, and a
curse, without an
inhabitant, as at this
day.
\VS{23}Because ye have burned
incense, and because ye have
sinned against the
{\ND{
LORD}}, and have not
obeyed the
voice of the
{\ND{
LORD}}, nor
walked in his
law, nor in his
statutes, nor in his
testimonies; therefore this
evil is
happened unto you, as at this
day.
\VS{24}Moreover
Jeremiah
said unto all the
people, and to all the
women,
Hear the
word of the
{\ND{
LORD}}, all
Judah that
{\ADD{are}} in the
land of
Egypt:
\VS{25}Thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel,
saying; Ye and your
wives have both
spoken with your
mouths, and
fulfilled with your
hand,
saying, We will
surely
perform our
vows that we have
vowed, to burn
incense to the
queen of
heaven, and to pour
out drink
offerings unto her: ye will
surely
accomplish your
vows, and
surely
perform your
vows.
\VS{26}Therefore
hear ye the
word of the
{\ND{
LORD}}, all
Judah that
dwell in the
land of
Egypt; Behold, I have
sworn by my
great
name,
saith the
{\ND{
LORD}}, that my
name shall no more be
named in the
mouth of any
man of
Judah in all the
land of
Egypt,
saying, The
Lord
{\ND{
GOD}}
liveth.
\VS{27}Behold, I will
watch over them for
evil, and not for
good: and all the
men of
Judah that
{\ADD{are}} in the
land of
Egypt shall be
consumed by the
sword and by the
famine, until there be an
end of them.
\VS{28}Yet a
small
number that
escape the
sword shall return
out of the
land of
Egypt into the
land of
Judah, and all the
remnant of
Judah, that are
gone into the
land of
Egypt to
sojourn there, shall
know whose
words shall
stand, mine, or theirs.\FTNT{§§}{{\FR 44.28 }
mine…: Heb. from me, or, from them}
\par }{\PP \VS{29}¶ And this
{\ADD{shall be}} a
sign unto you,
saith the
{\ND{
LORD}}, that I will
punish you in this
place, that ye may
know that my
words shall
surely
stand against you for
evil:
\VS{30}Thus
saith the
{\ND{
LORD}}; Behold, I will
give
Pharaoh-hophra
king of
Egypt into the
hand of his
enemies, and into the
hand of them that
seek his
life; as I
gave
Zedekiah
king of
Judah into the
hand of
Nebuchadrezzar
king of
Babylon, his
enemy, and that
sought his
life.

\par }\Chap{45}{\PP \VerseOne{1}The
word that
Jeremiah the
prophet
spake unto
Baruch the
son of
Neriah, when he had
written these
words in a
book at the
mouth of
Jeremiah, in the
fourth
year of
Jehoiakim the
son of
Josiah
king of
Judah,
saying,
\VS{2}Thus
saith the
{\ND{
LORD}}, the
God of
Israel, unto thee, O
Baruch;
\VS{3}Thou didst
say,
Woe is me now! for the
{\ND{
LORD}} hath
added
grief to my
sorrow; I
fainted in my
sighing, and I
find no
rest.
\par }{\PP \VS{4}¶ Thus shalt thou
say unto him, The
{\ND{
LORD}}
saith thus; Behold,
{\ADD{that}} which I have
built will I break
down, and that which I have
planted I will pluck
up, even this whole
land.
\VS{5}And
seekest thou great
things for thyself?
seek
{\ADD{them}} not: for, behold, I will
bring
evil upon all
flesh,
saith the
{\ND{
LORD}}: but thy
life will I
give unto thee for a
prey in all
places whither thou
goest.

\par }\Chap{46}{\PP \VerseOne{1}The
word of the
{\ND{
LORD}} which came to
Jeremiah the
prophet against the
Gentiles;
\VS{2}Against
Egypt, against the
army of
Pharaoh-necho
king of
Egypt, which was by the
river
Euphrates in
Carchemish, which
Nebuchadrezzar
king of
Babylon
smote in the
fourth
year of
Jehoiakim the
son of
Josiah
king of
Judah.
\VS{3}Order ye the
buckler and
shield, and draw
near to
battle.
\VS{4}Harness the
horses; and get
up, ye
horsemen, and stand
forth with
{\ADD{your}}
helmets;
furbish the
spears,
{\ADD{and}} put
on the
brigandines.
\VS{5}Wherefore have I
seen them
dismayed
{\ADD{and}}
turned away
back? and their mighty
ones are beaten
down, and are
fled
apace, and look not
back:
{\ADD{for}}
fear
{\ADD{was}} round
about,
saith the
{\ND{
LORD}}.\FTNT{*}{{\FR 46.5 }
beaten…: Heb. broken in pieces}\FTNT{†}{{\FR 46.5 }
fled…: Heb. fled a flight}
\VS{6}Let not the
swift flee
away, nor the mighty
man
escape; they shall
stumble, and
fall toward the
north
by the
river
Euphrates.
\VS{7}Who
{\ADD{is}} this
{\ADD{that}} cometh
up as a
flood, whose
waters are
moved as the
rivers?
\VS{8}Egypt riseth
up like a
flood, and
{\ADD{his}}
waters are
moved like the
rivers; and he
saith, I will go
up,
{\ADD{and}} will
cover the
earth; I will
destroy the
city and the
inhabitants thereof.
\VS{9}Come
up, ye
horses; and
rage, ye
chariots; and let the mighty
men come
forth; the
Ethiopians and the
Libyans, that
handle the
shield; and the
Lydians, that
handle
{\ADD{and}}
bend the
bow.\FTNT{‡}{{\FR 46.9 }
the Ethiopians: Heb. Cush}\FTNT{§}{{\FR 46.9 }
the Libyans: Heb. Put}
\VS{10}For this
{\ADD{is}} the
day of the
Lord
{\ND{
GOD}} of
hosts, a
day of
vengeance, that he may
avenge him of his
adversaries: and the
sword shall
devour, and it shall be
satiate and made
drunk with their
blood: for the
Lord
{\ND{
GOD}} of
hosts hath a
sacrifice in the
north
country by the
river
Euphrates.
\VS{11}Go
up into
Gilead, and
take
balm, O
virgin, the
daughter of
Egypt: in
vain shalt thou use
many
medicines;
{\ADD{for}} thou shalt not be
cured.\FTNT{**}{{\FR 46.11 }
thou shalt…: Heb. no cure shall be unto thee}
\par }{\PP \VS{12}The
nations have
heard of thy
shame, and thy
cry hath
filled the
land: for the mighty
man hath
stumbled against the
mighty,
{\ADD{and}} they are
fallen
both
together.
\par }{\PP \VS{13}¶ The
word that the
{\ND{
LORD}}
spake to
Jeremiah the
prophet, how
Nebuchadrezzar
king of
Babylon should
come
{\ADD{and}}
smite the
land of
Egypt.
\VS{14}Declare ye in
Egypt, and
publish in
Migdol, and
publish in
Noph and in
Tahpanhes:
say ye, Stand
fast, and
prepare thee; for the
sword shall
devour round
about thee.
\VS{15}Why are thy
valiant
{\ADD{men}} swept
away? they
stood not, because the
{\ND{
LORD}} did
drive them.
\VS{16}He made
many to
fall, yea,
one
fell upon
another: and they
said,
Arise, and let us go
again to our own
people, and to the
land of our
nativity,
from the
oppressing
sword.\FTNT{††}{{\FR 46.16 }
made…: Heb. multiplied the faller}
\VS{17}They did
cry there,
Pharaoh
king of
Egypt
{\ADD{is but}} a
noise; he hath
passed the time
appointed.
\VS{18}{\ADD{As}} I
live,
saith the
King, whose
name
{\ADD{is}} the
{\ND{
LORD}} of
hosts, Surely as
Tabor
{\ADD{is}} among the
mountains, and as
Carmel by the
sea,
{\ADD{so}} shall he
come.
\VS{19}O thou
daughter
dwelling in
Egypt,
furnish thyself to go into
captivity: for
Noph shall be
waste and
desolate without an
inhabitant.\FTNT{‡‡}{{\FR 46.19 }
furnish…: Heb. make thee instruments of captivity}
\VS{20}Egypt
{\ADD{is like}} a very
fair
heifer,
{\ADD{but}}
destruction
cometh; it cometh
out of the
north.
\VS{21}Also her hired
men
{\ADD{are}} in the
midst of her like
fatted
bullocks; for they also are turned
back,
{\ADD{and}} are fled
away
together: they did not
stand, because the
day of their
calamity was
come upon them,
{\ADD{and}} the
time of their
visitation.\FTNT{§§}{{\FR 46.21 }
fatted…: Heb. bullocks of the stall}
\VS{22}The
voice thereof shall
go like a
serpent; for they shall
march with an
army, and
come against her with
axes, as
hewers of
wood.
\VS{23}They shall cut
down her
forest,
saith the
{\ND{
LORD}}, though it cannot be
searched; because they are
more than the
grasshoppers, and
{\ADD{are}}
innumerable.
\VS{24}The
daughter of
Egypt shall be
confounded; she shall be
delivered into the
hand of the
people of the
north.
\VS{25}The
{\ND{
LORD}} of
hosts, the
God of
Israel,
saith; Behold, I will
punish the
multitude of
No, and
Pharaoh, and
Egypt, with their
gods, and their
kings; even
Pharaoh, and
{\ADD{all}} them that
trust in him:\FTNT{***}{{\FR 46.25 }
multitude: or, nourisher: Heb. Amon}
\VS{26}And I will
deliver them into the
hand of those that
seek their
lives, and into the
hand of
Nebuchadrezzar
king of
Babylon, and into the
hand of his
servants: and
afterward it shall be
inhabited, as in the
days of
old,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{27}¶ But
fear not thou, O my
servant
Jacob, and be not
dismayed, O
Israel: for, behold, I will
save thee from afar
off, and thy
seed from the
land of their
captivity; and
Jacob shall
return, and be in
rest and at
ease, and none shall make
{\ADD{him}}
afraid.
\VS{28}Fear thou not, O
Jacob my
servant,
saith the
{\ND{
LORD}}: for I
{\ADD{am}} with thee; for I will
make a full
end of all the
nations whither I have
driven thee: but I will not
make a full
end of thee, but
correct thee in
measure; yet will I not leave thee
wholly
unpunished.\FTNT{†††}{{\FR 46.28 }
not leave…: or, not utterly cut thee off}

\par }\Chap{47}{\PP \VerseOne{1}The
word of the
{\ND{
LORD}} that came to
Jeremiah the
prophet against the
Philistines, before that
Pharaoh
smote
Gaza.\FTNT{*}{{\FR 47.1 }
Gaza: Heb. Azzah}
\VS{2}Thus
saith the
{\ND{
LORD}}; Behold,
waters rise
up out of the
north, and shall be an
overflowing
flood, and shall
overflow the
land, and all that is
therein; the
city, and them that
dwell therein: then the
men shall
cry, and all the
inhabitants of the
land shall
howl.\FTNT{†}{{\FR 47.2 }
all that…: Heb. the fulness thereof}
\VS{3}At the
noise of the
stamping of the
hoofs of his
strong
{\ADD{horses}}, at the
rushing of his
chariots,
{\ADD{and at}} the
rumbling of his
wheels, the
fathers shall not look
back to
{\ADD{their}}
children for
feebleness of
hands;
\VS{4}Because of the
day that
cometh to
spoil all the
Philistines,
{\ADD{and}} to cut
off from
Tyrus and
Zidon every
helper that
remaineth: for the
{\ND{
LORD}} will
spoil the
Philistines, the
remnant of the
country of
Caphtor.\FTNT{‡}{{\FR 47.4 }
the country: Heb. the isle}
\VS{5}Baldness is
come upon
Gaza;
Ashkelon is cut
off
{\ADD{with}} the
remnant of their
valley: how long wilt thou
cut thyself?
\VS{6}O thou
sword of the
{\ND{
LORD}}, how long
{\ADD{will it be}}
ere thou be
quiet? put
up thyself into thy
scabbard,
rest, and be
still.\FTNT{§}{{\FR 47.6 }
put…: Heb. gather thyself}
\VS{7}How can it be
quiet, seeing the
{\ND{
LORD}} hath given it a
charge against
Ashkelon, and against the
sea
shore? there hath he
appointed it.\FTNT{**}{{\FR 47.7 }
How…: Heb. How canst thou}

\par }\Chap{48}{\PP \VerseOne{1}Against
Moab thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel;
Woe unto
Nebo! for it is
spoiled:
Kiriathaim is
confounded
{\ADD{and}}
taken:
Misgab is
confounded and
dismayed.\FTNT{*}{{\FR 48.1 }
Misgab: or, the high place}
\VS{2}{\ADD{There shall be}} no more
praise of
Moab: in
Heshbon they have
devised
evil against it;
come, and let us cut it
off from
{\ADD{being}} a
nation. Also thou shalt be cut
down, O
Madmen; the
sword shall
pursue
thee.\FTNT{†}{{\FR 48.2 }
be cut…: or, be brought to silence}\FTNT{‡}{{\FR 48.2 }
pursue…: Heb. go after thee}
\VS{3}A
voice of
crying
{\ADD{shall be}} from
Horonaim,
spoiling and
great
destruction.
\VS{4}Moab is
destroyed; her little
ones have caused a
cry to be
heard.
\VS{5}For in the going
up of
Luhith
continual
weeping shall go
up; for in the going
down of
Horonaim the
enemies have
heard a
cry of
destruction.\FTNT{§}{{\FR 48.5 }
continual…: Heb. weeping with weeping}
\VS{6}Flee,
save your
lives, and
be like the
heath in the
wilderness.\FTNT{**}{{\FR 48.6 }
the heath: or, a naked tree}
\par }{\PP \VS{7}¶ For because thou hast
trusted in thy
works and in thy
treasures, thou shalt also be
taken: and
Chemosh shall go
forth into
captivity
{\ADD{with}} his
priests and his
princes
together.
\VS{8}And the
spoiler shall
come upon every
city, and no
city shall
escape: the
valley also shall
perish, and the
plain shall be
destroyed, as the
{\ND{
LORD}} hath
spoken.
\VS{9}Give
wings unto
Moab, that it may
flee and get
away: for the
cities thereof shall be
desolate, without any to
dwell
therein.
\VS{10}Cursed
{\ADD{be}} he that
doeth the
work of the
{\ND{
LORD}}
deceitfully, and
cursed
{\ADD{be}} he that keepeth
back his
sword from
blood.\FTNT{††}{{\FR 48.10 }
deceitfully: or, negligently}
\par }{\PP \VS{11}¶
Moab hath been at
ease from his
youth, and he hath
settled on his
lees, and hath not been
emptied from
vessel to
vessel, neither hath he
gone into
captivity: therefore his
taste
remained in him, and his
scent is not
changed.\FTNT{‡‡}{{\FR 48.11 }
remained: Heb. stood}
\VS{12}Therefore, behold, the
days
come,
saith the
{\ND{
LORD}}, that I will
send unto him
wanderers, that shall cause him to
wander, and shall
empty his
vessels, and
break their
bottles.
\VS{13}And
Moab shall be
ashamed of
Chemosh, as the
house of
Israel was
ashamed of
Beth-el their
confidence.
\par }{\PP \VS{14}¶ How
say ye, We
{\ADD{are}}
mighty and
strong
men for the
war?
\VS{15}Moab is
spoiled, and gone
up
{\ADD{out of}} her
cities, and his
chosen young
men are gone
down to the
slaughter,
saith the
King, whose
name
{\ADD{is}} the
{\ND{
LORD}} of
hosts.\FTNT{§§}{{\FR 48.15 }
his…: Heb. the choice of}
\VS{16}The
calamity of
Moab
{\ADD{is}}
near to
come, and his
affliction
hasteth
fast.
\VS{17}All ye that are
about him,
bemoan him; and all ye that
know his
name,
say, How is the
strong
staff
broken,
{\ADD{and}} the
beautiful
rod!
\VS{18}Thou
daughter that dost
inhabit
Dibon, come
down from
{\ADD{thy}}
glory, and
sit in
thirst; for the
spoiler of
Moab shall
come upon thee,
{\ADD{and}} he shall
destroy thy strong
holds.
\VS{19}O
inhabitant of
Aroer,
stand by the
way, and
espy;
ask him that
fleeth, and her that
escapeth,
{\ADD{and}}
say, What is
done?\FTNT{***}{{\FR 48.19 }
inhabitant: Heb. inhabitress}
\VS{20}Moab is
confounded; for it is broken
down:
howl and
cry;
tell ye it in
Arnon, that
Moab is
spoiled,
\VS{21}And
judgment is
come upon the
plain
country; upon
Holon, and upon
Jahazah, and upon
Mephaath,
\VS{22}And upon
Dibon, and upon
Nebo, and upon
Beth-diblathaim,
\VS{23}And upon
Kiriathaim, and upon
Beth-gamul, and upon
Beth-meon,
\VS{24}And upon
Kerioth, and upon
Bozrah, and upon all the
cities of the
land of
Moab,
far or
near.
\VS{25}The
horn of
Moab is cut
off, and his
arm is
broken,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{26}¶ Make ye him
drunken: for he
magnified
{\ADD{himself}} against the
{\ND{
LORD}}:
Moab also shall
wallow in his
vomit, and he also shall be in
derision.
\VS{27}For was not
Israel a
derision unto thee? was he
found among
thieves? for
since thou
spakest of him, thou
skippedst for joy.\FTNT{†††}{{\FR 48.27 }
skippedst…: or, movedst thyself}
\VS{28}O ye that
dwell in
Moab,
leave the
cities, and
dwell in the
rock, and be like the
dove
{\ADD{that}} maketh her
nest in the
sides of the
hole’s
mouth.
\VS{29}We have
heard the
pride of
Moab, (he is
exceeding
proud) his
loftiness, and his
arrogancy, and his
pride, and the
haughtiness of his
heart.
\VS{30}I
know his
wrath,
saith the
{\ND{
LORD}}; but
{\ADD{it shall}} not
{\ADD{be}} so; his
lies shall not so
effect
{\ADD{it}}.\FTNT{‡‡‡}{{\FR 48.30 }
his lies…: or, those on whom he stayeth (Heb. his bars) do not right}
\VS{31}Therefore will I
howl for
Moab, and I will cry
out for all
Moab;
{\ADD{mine heart}} shall
mourn for the
men of
Kir-heres.
\VS{32}O
vine of
Sibmah, I will
weep for thee with the
weeping of
Jazer: thy
plants are gone
over the
sea, they
reach
{\ADD{even}} to the
sea of
Jazer: the
spoiler is
fallen upon thy summer
fruits and upon thy
vintage.
\VS{33}And
joy and
gladness is
taken from the plentiful
field, and from the
land of
Moab; and I have caused
wine to
fail from the
winepresses: none shall
tread with
shouting;
{\ADD{their}}
shouting
{\ADD{shall be}} no
shouting.
\VS{34}From the
cry of
Heshbon
{\ADD{even}} unto
Elealeh,
{\ADD{and even}} unto
Jahaz, have they
uttered their
voice, from
Zoar
{\ADD{even}} unto
Horonaim,
{\ADD{as}} an
heifer of three years
old: for the
waters also of
Nimrim shall be
desolate.\FTNT{§§§}{{\FR 48.34 }
desolate: Heb. desolations}
\VS{35}Moreover I will cause to
cease in
Moab,
saith the
{\ND{
LORD}}, him that
offereth in the high
places, and him that burneth
incense to his
gods.
\VS{36}Therefore mine
heart shall
sound for
Moab like
pipes, and mine
heart shall
sound like
pipes for the
men of
Kir-heres: because the
riches
{\ADD{that}} he hath
gotten are
perished.
\VS{37}For every
head
{\ADD{shall be}}
bald, and every
beard
clipped: upon all the
hands
{\ADD{shall be}}
cuttings, and upon the
loins
sackcloth.\FTNT{*}{{\FR 48.37 }
clipped: Heb. diminished}
\VS{38}{\ADD{There shall be}}
lamentation generally upon all the
housetops of
Moab, and in the
streets thereof: for I have
broken
Moab like a
vessel wherein
{\ADD{is}} no
pleasure,
saith the
{\ND{
LORD}}.
\VS{39}They shall
howl,
{\ADD{saying}}, How is it broken
down! how hath
Moab
turned the
back with
shame! so shall
Moab be a
derision and a
dismaying to all them
about him.\FTNT{†}{{\FR 48.39 }
back: Heb. neck}
\VS{40}For thus
saith the
{\ND{
LORD}}; Behold, he shall
fly as an
eagle, and shall
spread his
wings over
Moab.
\VS{41}Kerioth is
taken, and the strong
holds are
surprised, and the mighty
men’s
hearts in
Moab at that
day shall be as the
heart of a
woman in her
pangs.\FTNT{‡}{{\FR 48.41 }
Kerioth: or, The cities}
\VS{42}And
Moab shall be
destroyed from
{\ADD{being}} a
people, because he hath
magnified
{\ADD{himself}} against the
{\ND{
LORD}}.
\VS{43}Fear, and the
pit, and the
snare,
{\ADD{shall be}} upon thee, O
inhabitant of
Moab,
saith the
{\ND{
LORD}}.
\VS{44}He that
fleeth
from the
fear shall
fall into the
pit; and he that getteth
up out of the
pit shall be
taken in the
snare: for I will
bring upon it,
{\ADD{even}} upon
Moab, the
year of their
visitation,
saith the
{\ND{
LORD}}.
\VS{45}They that
fled
stood under the
shadow of
Heshbon because of the
force: but a
fire shall come
forth out of
Heshbon, and a
flame from the
midst of
Sihon, and shall
devour the
corner of
Moab, and the crown of the
head of the
tumultuous
ones.\FTNT{§}{{\FR 48.45 }
tumultuous…: Heb. children of noise}
\VS{46}Woe be unto thee, O
Moab! the
people of
Chemosh
perisheth: for thy
sons are
taken
captives, and thy
daughters
captives.\FTNT{**}{{\FR 48.46 }
captives, and: Heb. in captivity, etc}
\par }{\PP \VS{47}¶ Yet will I bring
again the
captivity of
Moab in the
latter
days,
saith the
{\ND{
LORD}}. Thus
far
{\ADD{is}} the
judgment of
Moab.

\par }\Chap{49}{\PP \VerseOne{1}Concerning the
Ammonites, thus
saith the
{\ND{
LORD}}; Hath
Israel no
sons? hath he no
heir? why
{\ADD{then}} doth their
king
inherit
Gad, and his
people
dwell in his
cities?\FTNT{*}{{\FR 49.1 }
Concerning: or, Against}\FTNT{†}{{\FR 49.1 }
their king: or, Melcom}
\VS{2}Therefore, behold, the
days
come,
saith the
{\ND{
LORD}}, that I will cause an
alarm of
war to be
heard in
Rabbah of the
Ammonites; and it shall be a
desolate
heap, and her
daughters shall be
burned with
fire: then shall
Israel be
heir unto them that were his
heirs,
saith the
{\ND{
LORD}}.
\VS{3}Howl, O
Heshbon, for
Ai is
spoiled:
cry, ye
daughters of
Rabbah,
gird you with
sackcloth;
lament, and run to and
fro by the
hedges; for their
king shall
go into
captivity,
{\ADD{and}} his
priests and his
princes
together.\FTNT{‡}{{\FR 49.3 }
their king: or, Melcom}
\VS{4}Wherefore
gloriest thou in the
valleys, thy
flowing
valley, O
backsliding
daughter? that
trusted in her
treasures,
{\ADD{saying}}, Who shall
come unto me?\FTNT{§}{{\FR 49.4 }
thy…: or, thy valley floweth away}
\VS{5}Behold, I will
bring a
fear upon thee,
saith the
Lord
{\ND{
GOD}} of
hosts, from all those that be
about thee; and ye shall be driven
out every
man right
forth; and none shall gather
up him that
wandereth.
\VS{6}And
afterward I will bring
again the
captivity of the
children of
Ammon,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{7}¶ Concerning
Edom, thus
saith the
{\ND{
LORD}} of
hosts;
{\ADD{Is}}
wisdom no more in
Teman? is
counsel
perished from the
prudent? is their
wisdom
vanished?
\VS{8}Flee ye, turn
back,
dwell
deep, O
inhabitants of
Dedan; for I will
bring the
calamity of
Esau upon him, the
time
{\ADD{that}} I will
visit him.\FTNT{**}{{\FR 49.8 }
turn…: or, they are turned back}
\VS{9}If
grapegatherers
come to thee, would they not
leave
{\ADD{some}} gleaning
grapes? if
thieves by
night, they will
destroy till they have
enough.\FTNT{††}{{\FR 49.9 }
till…: Heb. their sufficiency}
\VS{10}But I have made
Esau
bare, I have
uncovered his secret
places, and he shall not be
able to
hide himself: his
seed is
spoiled, and his
brethren, and his
neighbours, and he
{\ADD{is}} not.
\VS{11}Leave thy fatherless
children, I will preserve
{\ADD{them}}
alive; and let thy
widows
trust in me.
\VS{12}For thus
saith the
{\ND{
LORD}}; Behold, they whose
judgment
{\ADD{was}} not to
drink of the
cup have
assuredly
drunken; and
{\ADD{art}} thou he
{\ADD{that}} shall
altogether go
unpunished? thou shalt not go
unpunished, but thou shalt
surely
drink
{\ADD{of it}}.
\VS{13}For I have
sworn by myself,
saith the
{\ND{
LORD}}, that
Bozrah shall become a
desolation, a
reproach, a
waste, and a
curse; and all the
cities thereof shall be
perpetual
wastes.
\VS{14}I have
heard a
rumour from the
{\ND{
LORD}}, and an
ambassador is
sent unto the
heathen,
{\ADD{saying}}, Gather ye
together, and
come against her, and rise
up to the
battle.
\VS{15}For, lo, I will
make thee
small among the
heathen,
{\ADD{and}}
despised among
men.
\VS{16}Thy
terribleness hath
deceived thee,
{\ADD{and}} the
pride of thine
heart, O thou that
dwellest in the
clefts of the
rock, that
holdest the
height of the
hill: though thou shouldest make thy
nest as
high as the
eagle, I will bring thee
down from thence,
saith the
{\ND{
LORD}}.
\VS{17}Also
Edom shall be a
desolation: every one that
goeth by it shall be
astonished, and shall
hiss at all the
plagues thereof.
\VS{18}As in the
overthrow of
Sodom and
Gomorrah and the
neighbour
{\ADD{cities}} thereof,
saith the
{\ND{
LORD}}, no
man shall
abide there, neither shall a
son of
man
dwell in it.
\VS{19}Behold, he shall come
up like a
lion from the
swelling of
Jordan against the
habitation of the
strong: but I will
suddenly make him run
away from her: and who
{\ADD{is}} a
chosen
{\ADD{man, that}} I may
appoint over her? for who
{\ADD{is}} like me? and who will appoint me the
time? and who
{\ADD{is}} that
shepherd that will
stand
before me?\FTNT{‡‡}{{\FR 49.19 }
appoint me…: or, convent me in judgment?}
\VS{20}Therefore
hear the
counsel of the
{\ND{
LORD}}, that he hath
taken against
Edom; and his
purposes, that he hath
purposed against the
inhabitants of
Teman: Surely the
least of the
flock shall draw them
out: surely he shall make their
habitations
desolate with them.
\VS{21}The
earth is
moved at the
noise of their
fall, at the
cry the
noise thereof was
heard in the
Red
sea.\FTNT{§§}{{\FR 49.21 }
Red sea: Heb. Weedy sea}
\VS{22}Behold, he shall come
up and
fly as the
eagle, and
spread his
wings over
Bozrah: and at that
day shall the
heart of the mighty
men of
Edom be as the
heart of a
woman in her
pangs.
\par }{\PP \VS{23}¶ Concerning
Damascus.
Hamath is
confounded, and
Arpad: for they have
heard
evil
tidings: they are
fainthearted;
{\ADD{there is}}
sorrow on the
sea; it
cannot be
quiet.\FTNT{***}{{\FR 49.23 }
fainthearted: Heb. melted}\FTNT{†††}{{\FR 49.23 }
on…: or, as on the sea}
\VS{24}Damascus is waxed
feeble,
{\ADD{and}}
turneth herself to
flee, and
fear hath
seized on
{\ADD{her:}}
anguish and
sorrows have
taken her, as a woman in
travail.
\VS{25}How is the
city of
praise not
left, the
city of my
joy!
\VS{26}Therefore her young
men shall
fall in her
streets, and all the
men of
war shall be cut
off in that
day,
saith the
{\ND{
LORD}} of
hosts.
\VS{27}And I will
kindle a
fire in the
wall of
Damascus, and it shall
consume the
palaces of
Ben-hadad.
\par }{\PP \VS{28}¶ Concerning
Kedar, and concerning the
kingdoms of
Hazor, which
Nebuchadrezzar
king of
Babylon shall
smite, thus
saith the
{\ND{
LORD}};
Arise ye, go
up to
Kedar, and
spoil the
men of the
east.
\VS{29}Their
tents and their
flocks shall they take
away: they shall
take to themselves their
curtains, and all their
vessels, and their
camels; and they shall
cry unto them,
Fear
{\ADD{is}} on every
side.
\par }{\PP \VS{30}¶
Flee,
get you far
off,
dwell
deep, O ye
inhabitants of
Hazor,
saith the
{\ND{
LORD}}; for
Nebuchadrezzar
king of
Babylon hath taken
counsel against you, and hath
conceived a
purpose against you.\FTNT{‡‡‡}{{\FR 49.30 }
get…: Heb. flit greatly}
\VS{31}Arise, get you
up unto the
wealthy
nation, that
dwelleth without
care,
saith the
{\ND{
LORD}}, which have neither
gates nor
bars,
{\ADD{which}}
dwell
alone.\FTNT{§§§}{{\FR 49.31 }
wealthy: or, that is at ease}
\VS{32}And their
camels shall be a
booty, and the
multitude of their
cattle a
spoil: and I will
scatter into all
winds them
{\ADD{that are}} in the
utmost
corners; and I will
bring their
calamity from all
sides thereof,
saith the
{\ND{
LORD}}.\FTNT{*}{{\FR 49.32 }
in…: Heb. cut off into corners, or, that have the corners of their hair polled}
\VS{33}And
Hazor shall be a
dwelling for
dragons,
{\ADD{and}} a
desolation
for
ever: there shall no
man
abide there, nor
{\ADD{any}}
son of
man
dwell in it.
\par }{\PP \VS{34}¶ The
word of the
{\ND{
LORD}} that came to
Jeremiah the
prophet against
Elam in the
beginning of the
reign of
Zedekiah
king of
Judah,
saying,
\VS{35}Thus
saith the
{\ND{
LORD}} of
hosts; Behold, I will
break the
bow of
Elam, the
chief of their
might.
\VS{36}And upon
Elam will I
bring the
four
winds from the
four
quarters of
heaven, and will
scatter them toward all those
winds; and there shall be no
nation whither the
outcasts of
Elam shall not
come.
\VS{37}For I will cause
Elam to be
dismayed
before their
enemies, and
before them that
seek their
life: and I will
bring
evil upon them,
{\ADD{even}} my
fierce
anger,
saith the
{\ND{
LORD}}; and I will
send the
sword
after them, till I have
consumed them:
\VS{38}And I will
set my
throne in
Elam, and will
destroy from thence the
king and the
princes,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{39}¶ But it shall come to pass in the
latter
days,
{\ADD{that}} I will bring
again the
captivity of
Elam,
saith the
{\ND{
LORD}}.

\par }\Chap{50}{\PP \VerseOne{1}The
word that the
{\ND{
LORD}}
spake against
Babylon
{\ADD{and}} against the
land of the
Chaldeans
by
Jeremiah the
prophet.\FTNT{*}{{\FR 50.1 }
by…: Heb. by the hand of Jeremiah}
\VS{2}Declare ye among the
nations, and
publish, and set
up a
standard;
publish,
{\ADD{and}}
conceal not:
say,
Babylon is
taken,
Bel is
confounded,
Merodach is broken in
pieces; her
idols are
confounded, her
images are broken in
pieces.\FTNT{†}{{\FR 50.2 }
set up: Heb. lift up}
\VS{3}For out of the
north there cometh
up a
nation against her, which shall
make her
land
desolate, and none shall
dwell therein: they shall
remove, they shall
depart, both
man and
beast.
\par }{\PP \VS{4}¶ In those
days, and in that
time,
saith the
{\ND{
LORD}}, the
children of
Israel shall
come, they and the
children of
Judah
together,
going and
weeping: they shall
go, and
seek the
{\ND{
LORD}} their
God.
\VS{5}They shall
ask the
way to
Zion with their
faces
thitherward,
{\ADD{saying}},
Come, and let us
join ourselves to the
{\ND{
LORD}} in a
perpetual
covenant
{\ADD{that}} shall not be
forgotten.
\VS{6}My
people hath been
lost
sheep: their
shepherds have caused them to go
astray, they have turned them
away
{\ADD{on}} the
mountains: they have
gone from
mountain to
hill, they have
forgotten their
restingplace.\FTNT{‡}{{\FR 50.6 }
restingplace: Heb. place to lie down in}
\VS{7}All that
found them have
devoured them: and their
adversaries
said, We
offend not, because they have
sinned against the
{\ND{
LORD}}, the
habitation of
justice, even the
{\ND{
LORD}}, the
hope of their
fathers.
\VS{8}Remove out of the
midst of
Babylon, and go
forth out of the
land of the
Chaldeans, and be as the he
goats
before the
flocks.
\par }{\PP \VS{9}¶ For, lo, I will
raise and cause to come
up against
Babylon an
assembly of
great
nations from the
north
country: and they shall set themselves in
array against her; from thence she shall be
taken: their
arrows
{\ADD{shall be}} as of a
mighty expert
man; none shall
return in
vain.\FTNT{§}{{\FR 50.9 }
expert…: or, destroyer}
\VS{10}And
Chaldea shall be a
spoil: all that
spoil her shall be
satisfied,
saith the
{\ND{
LORD}}.
\VS{11}Because ye were
glad, because ye
rejoiced, O ye
destroyers of mine
heritage, because ye are grown
fat as the
heifer at
grass, and
bellow as
bulls;\FTNT{**}{{\FR 50.11 }
fat: Heb. big, or, corpulent}\FTNT{††}{{\FR 50.11 }
bellow…: or, neigh as steeds}
\VS{12}Your
mother shall be
sore
confounded; she that
bare you shall be
ashamed: behold, the
hindermost of the
nations
{\ADD{shall be}} a
wilderness, a dry
land, and a
desert.
\VS{13}Because of the
wrath of the
{\ND{
LORD}} it shall not be
inhabited, but it shall be wholly
desolate: every one that
goeth by
Babylon shall be
astonished, and
hiss at all her
plagues.
\VS{14}Put yourselves in
array against
Babylon round
about: all ye that
bend the
bow,
shoot at her,
spare no
arrows: for she hath
sinned against the
{\ND{
LORD}}.
\VS{15}Shout against her round
about: she hath
given her
hand: her
foundations are
fallen, her
walls are thrown
down: for it
{\ADD{is}} the
vengeance of the
{\ND{
LORD}}: take
vengeance upon her; as she hath
done,
do unto her.
\VS{16}Cut
off the
sower from
Babylon, and him that
handleth the
sickle in the
time of
harvest: for
fear of the
oppressing
sword they shall
turn every
one to his
people, and they shall
flee every
one to his own
land.\FTNT{‡‡}{{\FR 50.16 }
sickle: or, scythe}
\par }{\PP \VS{17}¶
Israel
{\ADD{is}} a
scattered
sheep; the
lions have driven
{\ADD{him}}
away:
first the
king of
Assyria hath
devoured him; and
last this
Nebuchadrezzar
king of
Babylon hath broken his
bones.
\VS{18}Therefore thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; Behold, I will
punish the
king of
Babylon and his
land, as I have
punished the
king of
Assyria.
\VS{19}And I will
bring
Israel
again to his
habitation, and he shall
feed on
Carmel and
Bashan, and his
soul shall be
satisfied upon
mount
Ephraim and
Gilead.
\VS{20}In those
days, and in that
time,
saith the
{\ND{
LORD}}, the
iniquity of
Israel shall be sought
for, and
{\ADD{there shall be}} none; and the
sins of
Judah, and they shall not be
found: for I will
pardon them whom I
reserve.
\par }{\PP \VS{21}¶ Go
up against the
land of
Merathaim,
{\ADD{even}} against it, and against the
inhabitants of
Pekod:
waste and utterly
destroy
after them,
saith the
{\ND{
LORD}}, and
do according to all that I have
commanded thee.\FTNT{§§}{{\FR 50.21 }
of Merathaim: or, of the rebels}\FTNT{***}{{\FR 50.21 }
Pekod: or, Visitation}
\VS{22}A
sound of
battle
{\ADD{is}} in the
land, and of
great
destruction.
\VS{23}How is the
hammer of the whole
earth cut
asunder and
broken! how is
Babylon become a
desolation among the
nations!
\VS{24}I have laid a
snare for thee, and thou art also
taken, O
Babylon, and thou wast not
aware: thou art
found, and also
caught, because thou hast
striven against the
{\ND{
LORD}}.
\VS{25}The
{\ND{
LORD}} hath
opened his
armoury, and hath brought
forth the
weapons of his
indignation: for this
{\ADD{is}} the
work of the
Lord
{\ND{
GOD}} of
hosts in the
land of the
Chaldeans.
\VS{26}Come against her from the utmost
border,
open her
storehouses: cast her
up as
heaps, and destroy her
utterly: let nothing of her be
left.\FTNT{†††}{{\FR 50.26 }
from…: Heb. from the end}\FTNT{‡‡‡}{{\FR 50.26 }
cast…: or, tread her}
\VS{27}Slay all her
bullocks; let them go
down to the
slaughter:
woe unto them! for their
day is
come, the
time of their
visitation.
\VS{28}The
voice of them that
flee and escape
out of the
land of
Babylon, to
declare in
Zion the
vengeance of the
{\ND{
LORD}} our
God, the
vengeance of his
temple.
\VS{29}Call
together the
archers against
Babylon: all ye that
bend the
bow,
camp against it round
about; let none thereof
escape:
recompense her according to her
work; according to all that she hath
done,
do unto her: for she hath been
proud against the
{\ND{
LORD}}, against the Holy
One of
Israel.
\VS{30}Therefore shall her young
men
fall in the
streets, and all her
men of
war shall be cut
off in that
day,
saith the
{\ND{
LORD}}.
\VS{31}Behold, I
{\ADD{am}} against thee,
{\ADD{O thou}} most
proud,
saith the
Lord
{\ND{
GOD}} of
hosts: for thy
day is
come, the
time
{\ADD{that}} I will
visit thee.\FTNT{§§§}{{\FR 50.31 }
most…: Heb. pride}
\VS{32}And the most
proud shall
stumble and
fall, and none shall raise him
up: and I will
kindle a
fire in his
cities, and it shall
devour all round
about him.\FTNT{*}{{\FR 50.32 }
the…: Heb. pride}
\par }{\PP \VS{33}¶ Thus
saith the
{\ND{
LORD}} of
hosts; The
children of
Israel and the
children of
Judah
{\ADD{were}}
oppressed
together: and all that took them
captives held them
fast; they
refused to let them
go.
\VS{34}Their
Redeemer
{\ADD{is}}
strong; the
{\ND{
LORD}} of
hosts
{\ADD{is}} his
name: he shall
throughly
plead their
cause, that he may give
rest to the
land, and
disquiet the
inhabitants of
Babylon.
\par }{\PP \VS{35}¶ A
sword
{\ADD{is}} upon the
Chaldeans,
saith the
{\ND{
LORD}}, and upon the
inhabitants of
Babylon, and upon her
princes, and upon her
wise
{\ADD{men}}.
\VS{36}A
sword
{\ADD{is}} upon the
liars; and they shall
dote: a
sword
{\ADD{is}} upon her mighty
men; and they shall be
dismayed.\FTNT{†}{{\FR 50.36 }
liars: or, chief stays: Heb. bars}
\VS{37}A
sword
{\ADD{is}} upon their
horses, and upon their
chariots, and upon all the mingled
people that
{\ADD{are}} in the
midst of her; and they shall become as
women: a
sword
{\ADD{is}} upon her
treasures; and they shall be
robbed.
\VS{38}A
drought
{\ADD{is}} upon her
waters; and they shall be dried
up: for it
{\ADD{is}} the
land of graven
images, and they are
mad upon
{\ADD{their}}
idols.
\VS{39}Therefore the wild beasts of the
desert with the wild beasts of the
islands shall
dwell
{\ADD{there}}, and the
owls shall
dwell therein: and it shall be no more
inhabited for
ever; neither shall it be
dwelt in from
generation to
generation.
\VS{40}As
God
overthrew
Sodom and
Gomorrah and the
neighbour
{\ADD{cities}} thereof,
saith the
{\ND{
LORD}};
{\ADD{so}} shall no
man
abide there, neither shall any
son of
man
dwell therein.
\VS{41}Behold, a
people shall
come from the
north, and a
great
nation, and
many
kings shall be raised
up from the
coasts of the
earth.
\VS{42}They shall
hold the
bow and the
lance: they
{\ADD{are}}
cruel, and will not shew
mercy: their
voice shall
roar like the
sea, and they shall
ride upon
horses,
{\ADD{every one}} put in
array, like a
man to the
battle, against thee, O
daughter of
Babylon.
\VS{43}The
king of
Babylon hath
heard the
report of them, and his
hands waxed
feeble:
anguish took
hold of him,
{\ADD{and}}
pangs as of a woman in
travail.
\VS{44}Behold, he shall come
up like a
lion from the
swelling of
Jordan unto the
habitation of the
strong: but I will
make them
suddenly run
away from her: and who
{\ADD{is}} a
chosen
{\ADD{man, that}} I may
appoint over her? for who
{\ADD{is}} like me? and who will appoint me the
time? and who
{\ADD{is}} that
shepherd that will
stand
before me?\FTNT{‡}{{\FR 50.44 }
appoint me…: or, convent me to plead?}
\VS{45}Therefore
hear ye the
counsel of the
{\ND{
LORD}}, that he hath
taken against
Babylon; and his
purposes, that he hath
purposed against the
land of the
Chaldeans: Surely the
least of the
flock shall draw them
out: surely he shall make
{\ADD{their}}
habitation
desolate with them.
\VS{46}At the
noise of the
taking of
Babylon the
earth is
moved, and the
cry is
heard among the
nations.

\par }\Chap{51}{\PP \VerseOne{1}Thus
saith the
{\ND{
LORD}}; Behold, I will raise
up against
Babylon, and against them that
dwell in the
midst of them that rise
up against me, a
destroying
wind;\FTNT{*}{{\FR 51.1 }
midst: Heb. heart}
\VS{2}And will
send unto
Babylon
fanners, that shall
fan her, and shall
empty her
land: for in the
day of
trouble they shall be against her round
about.
\VS{3}Against
{\ADD{him that}}
bendeth let the
archer
bend his
bow, and against
{\ADD{him that}} lifteth himself
up in his
brigandine: and
spare ye not her young
men; destroy ye
utterly all her
host.
\VS{4}Thus the
slain shall
fall in the
land of the
Chaldeans, and
{\ADD{they that are}} thrust
through in her
streets.
\VS{5}For
Israel
{\ADD{hath}} not
{\ADD{been}}
forsaken, nor
Judah of his
God, of the
{\ND{
LORD}} of
hosts; though their
land was
filled with
sin against the Holy
One of
Israel.
\VS{6}Flee
out of the
midst of
Babylon, and
deliver every
man his
soul: be not cut
off in her
iniquity; for this
{\ADD{is}} the
time of the
{\ND{
LORD’s}}
vengeance; he will
render unto her a
recompence.
\VS{7}Babylon
{\ADD{hath been}} a
golden
cup in the
{\ND{
LORD’s}}
hand, that made all the
earth
drunken: the
nations have
drunken of her
wine; therefore the
nations are
mad.
\VS{8}Babylon is
suddenly
fallen and
destroyed:
howl for her;
take
balm for her
pain, if so be she may be
healed.
\VS{9}We would have
healed
Babylon, but she is not
healed:
forsake her, and let us
go every
one into his own
country: for her
judgment
reacheth unto
heaven, and is lifted
up
{\ADD{even}} to the
skies.
\VS{10}The
{\ND{
LORD}} hath brought
forth our
righteousness:
come, and let us
declare in
Zion the
work of the
{\ND{
LORD}} our
God.
\VS{11}Make
bright the
arrows;
gather the
shields: the
{\ND{
LORD}} hath raised
up the
spirit of the
kings of the
Medes: for his
device
{\ADD{is}} against
Babylon, to
destroy it; because it
{\ADD{is}} the
vengeance of the
{\ND{
LORD}}, the
vengeance of his
temple.\FTNT{†}{{\FR 51.11 }
bright: Heb. pure}
\VS{12}Set
up the
standard upon the
walls of
Babylon, make the
watch
strong, set
up the
watchmen,
prepare the
ambushes: for the
{\ND{
LORD}} hath both
devised and
done that which he
spake against the
inhabitants of
Babylon.\FTNT{‡}{{\FR 51.12 }
ambushes: Heb. liers in wait}
\VS{13}O thou that
dwellest upon
many
waters,
abundant in
treasures, thine
end is
come,
{\ADD{and}} the
measure of thy
covetousness.
\VS{14}The
{\ND{
LORD}} of
hosts hath
sworn by
himself,
{\ADD{saying}}, Surely I will
fill thee with
men, as with
caterpillers; and they shall lift
up a
shout against thee.\FTNT{§}{{\FR 51.14 }
by himself: Heb. by his soul}\FTNT{**}{{\FR 51.14 }
lift up: Heb. utter}
\VS{15}He hath
made the
earth by his
power, he hath
established the
world by his
wisdom, and hath stretched
out the
heaven by his
understanding.
\VS{16}When he
uttereth
{\ADD{his}}
voice,
{\ADD{there is}} a
multitude of
waters in the
heavens; and he causeth the
vapours to
ascend from the
ends of the
earth: he
maketh
lightnings with
rain, and bringeth
forth the
wind out of his
treasures.\FTNT{††}{{\FR 51.16 }
multitude: or, noise}
\VS{17}Every
man is
brutish by
{\ADD{his}}
knowledge; every
founder is
confounded by the graven
image: for his molten
image
{\ADD{is}}
falsehood, and
{\ADD{there is}} no
breath in them.\FTNT{‡‡}{{\FR 51.17 }
is brutish…: or, is more brutish than to know}
\VS{18}They
{\ADD{are}}
vanity, the
work of
errors: in the
time of their
visitation they shall
perish.
\VS{19}The
portion of
Jacob
{\ADD{is}} not like them; for he
{\ADD{is}} the
former of all things: and
{\ADD{Israel is}} the
rod of his
inheritance: the
{\ND{
LORD}} of
hosts
{\ADD{is}} his
name.
\VS{20}Thou
{\ADD{art}} my battle
axe
{\ADD{and}}
weapons of
war: for with thee will I break in
pieces the
nations, and with thee will I
destroy
kingdoms;\FTNT{§§}{{\FR 51.20 }
with thee: or, in thee, or, by thee}
\VS{21}And with thee will I break in
pieces the
horse and his
rider; and with thee will I break in
pieces the
chariot and his
rider;
\VS{22}With thee also will I break in
pieces
man and
woman; and with thee will I break in
pieces
old and
young; and with thee will I break in
pieces the young
man and the
maid;
\VS{23}I will also break in
pieces with thee the
shepherd and his
flock; and with thee will I break in
pieces the
husbandman and his yoke of
oxen; and with thee will I break in
pieces
captains and
rulers.
\VS{24}And I will
render unto
Babylon and to all the
inhabitants of
Chaldea all their
evil that they have
done in
Zion in your
sight,
saith the
{\ND{
LORD}}.
\VS{25}Behold, I
{\ADD{am}} against thee, O
destroying
mountain,
saith the
{\ND{
LORD}}, which
destroyest all the
earth: and I will stretch
out mine
hand upon thee, and roll thee
down from the
rocks, and will
make thee a
burnt
mountain.
\VS{26}And they shall not
take of thee a
stone for a
corner, nor a
stone for
foundations; but thou shalt be
desolate for
ever,
saith the
{\ND{
LORD}}.\FTNT{***}{{\FR 51.26 }
desolate…: Heb. everlasting desolations}
\VS{27}Set ye
up a
standard in the
land,
blow the
trumpet among the
nations,
prepare the
nations against her, call
together against her the
kingdoms of
Ararat,
Minni, and
Ashchenaz;
appoint a
captain against her; cause the
horses to come
up as the
rough
caterpillers.
\VS{28}Prepare against her the
nations with the
kings of the
Medes, the
captains thereof, and all the
rulers thereof, and all the
land of his
dominion.
\VS{29}And the
land shall
tremble and
sorrow: for every
purpose of the
{\ND{
LORD}} shall be
performed against
Babylon, to
make the
land of
Babylon a
desolation without an
inhabitant.
\VS{30}The mighty
men of
Babylon have
forborn to
fight, they have
remained in
{\ADD{their}}
holds: their
might hath
failed; they became as
women: they have
burned her
dwellingplaces; her
bars are
broken.
\VS{31}One
post shall
run to
meet
another, and one
messenger to
meet
another, to
shew the
king of
Babylon that his
city is
taken at
{\ADD{one}}
end,
\VS{32}And that the
passages are
stopped, and the
reeds they have
burned with
fire, and the
men of
war are
affrighted.
\VS{33}For thus
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel; The
daughter of
Babylon
{\ADD{is}} like a
threshingfloor,
{\ADD{it is}}
time to
thresh her: yet a little
while, and the
time of her
harvest shall
come.\FTNT{†††}{{\FR 51.33 }
it is…: or, in the time that he thresheth her}
\VS{34}Nebuchadrezzar the
king of
Babylon hath
devoured me, he hath
crushed me, he hath
made me an
empty
vessel, he hath swallowed me
up like a
dragon, he hath
filled his
belly with my
delicates, he hath cast me
out.
\VS{35}The
violence done to me and to my
flesh
{\ADD{be}} upon
Babylon, shall the
inhabitant of
Zion
say; and my
blood upon the
inhabitants of
Chaldea, shall
Jerusalem
say.\FTNT{‡‡‡}{{\FR 51.35 }
The violence…: Heb. My violence}\FTNT{§§§}{{\FR 51.35 }
flesh: or, remainder}\FTNT{*}{{\FR 51.35 }
inhabitant: Heb. inhabitress}
\VS{36}Therefore thus
saith the
{\ND{
LORD}}; Behold, I will
plead thy
cause, and take
vengeance for
thee; and I will dry
up her
sea, and make her
springs
dry.
\VS{37}And
Babylon shall become
heaps, a
dwellingplace for
dragons, an
astonishment, and an
hissing, without an
inhabitant.
\VS{38}They shall
roar
together like
lions: they shall
yell as
lions’
whelps.\FTNT{†}{{\FR 51.38 }
yell: or, shake themselves}
\VS{39}In their
heat I will
make their
feasts, and I will make them
drunken, that they may
rejoice, and
sleep a
perpetual
sleep, and not
wake,
saith the
{\ND{
LORD}}.
\VS{40}I will bring them
down like
lambs to the
slaughter, like
rams with he
goats.
\VS{41}How is
Sheshach
taken! and how is the
praise of the whole
earth
surprised! how is
Babylon become an
astonishment among the
nations!
\VS{42}The
sea is come
up upon
Babylon: she is
covered with the
multitude of the
waves thereof.
\VS{43}Her
cities are a
desolation, a
dry
land, and a
wilderness, a
land wherein no
man
dwelleth, neither doth
{\ADD{any}}
son of
man
pass
thereby.
\VS{44}And I will
punish
Bel in
Babylon, and I will bring
forth out of his
mouth that which he hath swallowed
up: and the
nations shall not flow
together any more unto him: yea, the
wall of
Babylon shall
fall.
\VS{45}My
people, go ye
out of the
midst of her, and
deliver ye every
man his
soul from the
fierce
anger of the
{\ND{
LORD}}.
\VS{46}And lest your
heart
faint, and ye
fear for the
rumour that shall be
heard in the
land; a
rumour shall both
come
{\ADD{one}}
year, and
after that in
{\ADD{another}}
year
{\ADD{shall come}} a
rumour, and
violence in the
land,
ruler against
ruler.\FTNT{‡}{{\FR 51.46 }
lest: or, let not}
\VS{47}Therefore, behold, the
days
come, that I will do
judgment upon the graven
images of
Babylon: and her whole
land shall be
confounded, and all her
slain shall
fall in the
midst of her.\FTNT{§}{{\FR 51.47 }
do…: Heb. visit upon}
\VS{48}Then the
heaven and the
earth, and all that
{\ADD{is}} therein, shall
sing for
Babylon: for the
spoilers shall
come unto her from the
north,
saith the
{\ND{
LORD}}.
\VS{49}As
Babylon
{\ADD{hath caused}} the
slain of
Israel to
fall, so at
Babylon shall
fall the
slain of all the
earth.\FTNT{**}{{\FR 51.49 }
As…: or, Both Babylon is to fall, O ye slain of Israel, and with Babylon, etc}\FTNT{††}{{\FR 51.49 }
the earth: or, the country}
\VS{50}Ye that have
escaped the
sword, go
away, stand not
still:
remember the
{\ND{
LORD}} afar
off, and let
Jerusalem
come into your
mind.
\VS{51}We are
confounded, because we have
heard
reproach:
shame hath
covered our
faces: for
strangers are
come into the
sanctuaries of the
{\ND{
LORD’s}}
house.
\VS{52}Wherefore, behold, the
days
come,
saith the
{\ND{
LORD}}, that I will do
judgment upon her graven
images: and through all her
land the
wounded shall
groan.
\VS{53}Though
Babylon should mount
up to
heaven, and though she should
fortify the
height of her
strength,
{\ADD{
yet}} from me shall
spoilers
come unto her,
saith the
{\ND{
LORD}}.
\VS{54}A
sound of a
cry
{\ADD{cometh}} from
Babylon, and
great
destruction from the
land of the
Chaldeans:
\VS{55}Because the
{\ND{
LORD}} hath
spoiled
Babylon, and
destroyed out of her the
great
voice; when her
waves do
roar like
great
waters, a
noise of their
voice is
uttered:
\VS{56}Because the
spoiler is
come upon her,
{\ADD{even}} upon
Babylon, and her mighty
men are
taken, every one of their
bows is
broken: for the
{\ND{
LORD}}
God of
recompences shall
surely
requite.
\VS{57}And I will make
drunk her
princes, and her
wise
{\ADD{men}}, her
captains, and her
rulers, and her mighty
men: and they shall
sleep a
perpetual
sleep, and not
wake,
saith the
King, whose
name
{\ADD{is}} the
{\ND{
LORD}} of
hosts.
\VS{58}Thus
saith the
{\ND{
LORD}} of
hosts; The
broad
walls of
Babylon shall be
utterly
broken, and her
high
gates shall be
burned with
fire; and the
people shall
labour in
vain, and the
folk
in the
fire, and they shall be
weary.\FTNT{‡‡}{{\FR 51.58 }
The broad…: or, The walls of broad Babylon}\FTNT{§§}{{\FR 51.58 }
broken: or, made naked}
\par }{\PP \VS{59}¶ The
word which
Jeremiah the
prophet
commanded
Seraiah the
son of
Neriah, the
son of
Maaseiah, when he
went with
Zedekiah the
king of
Judah into
Babylon in the
fourth
year of his
reign. And
{\ADD{this}}
Seraiah
{\ADD{was}} a
quiet
prince.\FTNT{***}{{\FR 51.59 }
with: or, on the behalf of}\FTNT{†††}{{\FR 51.59 }
quiet…: or, prince of Menucha, or, chief chamberlain}
\VS{60}So
Jeremiah
wrote in
a
book all the
evil that should
come upon
Babylon,
{\ADD{even}} all these
words that are
written against
Babylon.
\VS{61}And
Jeremiah
said to
Seraiah, When thou
comest to
Babylon, and shalt
see, and shalt
read all these
words;
\VS{62}Then shalt thou
say, O
{\ND{
LORD}}, thou hast
spoken against this
place, to cut it
off, that none shall
remain in it, neither
man nor
beast, but that it shall be
desolate for
ever.\FTNT{‡‡‡}{{\FR 51.62 }
desolate: Heb. desolations}
\VS{63}And it shall be, when thou hast made an
end of
reading this
book,
{\ADD{that}} thou shalt
bind a
stone to it, and
cast it into the
midst of
Euphrates:
\VS{64}And thou shalt
say, Thus shall
Babylon
sink, and shall not
rise
from the
evil that I will
bring upon her: and they shall be
weary. Thus far
{\ADD{are}} the
words of
Jeremiah.

\par }\Chap{52}{\PP \VerseOne{1}Zedekiah
{\ADD{was}}
one and
twenty
years
old when he began to
reign, and he
reigned
eleven
years in
Jerusalem. And his
mother’s
name
{\ADD{was}}
Hamutal the
daughter of
Jeremiah of
Libnah.\FTNT{*}{{\FR 52.1 }
began…: Heb. reigned}
\VS{2}And he
did
{\ADD{that which was}}
evil in the
eyes of the
{\ND{
LORD}}, according to all that
Jehoiakim had
done.
\VS{3}For
through the
anger of the
{\ND{
LORD}} it came to pass in
Jerusalem and
Judah, till he had cast them
out from his
presence, that
Zedekiah
rebelled against the
king of
Babylon.
\par }{\PP \VS{4}¶ And it came to pass in the
ninth
year of his
reign, in the
tenth
month, in the
tenth
{\ADD{day}} of the
month,
{\ADD{that}}
Nebuchadrezzar
king of
Babylon
came, he and all his
army, against
Jerusalem, and
pitched against it, and
built
forts against it round
about.
\VS{5}So the
city was
besieged unto the
eleventh
year of
king
Zedekiah.
\VS{6}And in the
fourth
month, in the
ninth
{\ADD{day}} of the
month, the
famine was
sore in the
city, so that there was no
bread for the
people of the
land.
\VS{7}Then the
city was broken
up, and all the
men of
war
fled, and went
forth out of the
city by
night by the
way of the
gate between the two
walls, which
{\ADD{was}} by the
king’s
garden; (now the
Chaldeans
{\ADD{were}} by the
city round
about:) and they
went by the
way of the
plain.
\par }{\PP \VS{8}¶ But the
army of the
Chaldeans
pursued
after the
king, and
overtook
Zedekiah in the
plains of
Jericho; and all his
army was
scattered from him.
\VS{9}Then they
took the
king, and carried him
up unto the
king of
Babylon to
Riblah in the
land of
Hamath; where he
gave
judgment upon him.
\VS{10}And the
king of
Babylon
slew the
sons of
Zedekiah before his
eyes: he
slew also all the
princes of
Judah in
Riblah.
\VS{11}Then he put
out the
eyes of
Zedekiah; and the
king of
Babylon
bound him in
chains, and
carried him to
Babylon, and
put him in
prison till the
day of his
death.\FTNT{†}{{\FR 52.11 }
put out: Heb. blinded}\FTNT{‡}{{\FR 52.11 }
chains: or, fetters}\FTNT{§}{{\FR 52.11 }
prison: Heb. house of the wards}
\par }{\PP \VS{12}¶ Now in the
fifth
month, in the
tenth
{\ADD{day}} of the
month, which
{\ADD{was}} the
nineteenth
year of
Nebuchadrezzar
king of
Babylon,
came
Nebuzar-adan,
captain of the
guard,
{\ADD{which}}
served the
king of
Babylon, into
Jerusalem,\FTNT{**}{{\FR 52.12 }
captain…: or, chief marshal: Heb. chief of the executioners, or, slaughtermen}\FTNT{††}{{\FR 52.12 }
served: Heb. stood before}
\VS{13}And
burned the
house of the
{\ND{
LORD}}, and the
king’s
house; and all the
houses of
Jerusalem, and all the
houses of the
great
{\ADD{men}},
burned he with
fire:
\VS{14}And all the
army of the
Chaldeans, that
{\ADD{were}} with the
captain of the
guard, brake
down all the
walls of
Jerusalem round
about.
\VS{15}Then
Nebuzar-adan the
captain of the
guard carried away
captive
{\ADD{certain}} of the
poor of the
people, and the
residue of the
people that
remained in the
city, and those that fell
away, that
fell to the
king of
Babylon, and the
rest of the
multitude.
\VS{16}But
Nebuzar-adan the
captain of the
guard
left
{\ADD{certain}} of the
poor of the
land for
vinedressers and for
husbandmen.
\VS{17}Also the
pillars of
brass that
{\ADD{were}} in the
house of the
{\ND{
LORD}}, and the
bases, and the
brasen
sea that
{\ADD{was}} in the
house of the
{\ND{
LORD}}, the
Chaldeans
brake, and
carried all the
brass of them to
Babylon.
\VS{18}The
caldrons also, and the
shovels, and the
snuffers, and the
bowls, and the
spoons, and all the
vessels of
brass wherewith they
ministered, took they
away.\FTNT{‡‡}{{\FR 52.18 }
shovels: or, instruments to remove the ashes}\FTNT{§§}{{\FR 52.18 }
bowls: or, basons}
\VS{19}And the
basons, and the
firepans, and the
bowls, and the
caldrons, and the
candlesticks, and the
spoons, and the
cups;
{\ADD{that}} which
{\ADD{was}} of
gold
{\ADD{in}}
gold, and
{\ADD{that}} which
{\ADD{was}} of
silver
{\ADD{in}}
silver,
took the
captain of the
guard
away.\FTNT{***}{{\FR 52.19 }
firepans: or, censers}
\VS{20}The
two
pillars,
one
sea, and
twelve
brasen
bulls that
{\ADD{were}} under the
bases, which
king
Solomon had
made in the
house of the
{\ND{
LORD}}: the
brass of all these
vessels was without
weight.\FTNT{†††}{{\FR 52.20 }
the brass: Heb. their brass}
\VS{21}And
{\ADD{concerning}} the
pillars, the
height of
one
pillar
{\ADD{was}}
eighteen
cubits; and a
fillet of
twelve
cubits did
compass it; and the
thickness thereof
{\ADD{was}}
four
fingers:
{\ADD{it was}}
hollow.\FTNT{‡‡‡}{{\FR 52.21 }
fillet: Heb. thread}
\VS{22}And a
chapiter of
brass
{\ADD{was}} upon it; and the
height of
one
chapiter
{\ADD{was}}
five
cubits, with
network and
pomegranates upon the
chapiters round
about, all
{\ADD{of}}
brass. The
second
pillar also and the
pomegranates
{\ADD{were}} like unto these.
\VS{23}And there were
ninety and
six
pomegranates on a
side;
{\ADD{and}} all the
pomegranates upon the
network
{\ADD{were}} an
hundred round
about.
\par }{\PP \VS{24}¶ And the
captain of the
guard
took
Seraiah the
chief
priest, and
Zephaniah the
second
priest, and the
three
keepers of the
door:\FTNT{§§§}{{\FR 52.24 }
door: Heb. threshold}
\VS{25}He
took also out of the
city
an
eunuch, which had the
charge of the
men of
war; and
seven
men of them that were
near the
king’s
person, which were
found in the
city; and the
principal
scribe of the
host, who
mustered the
people of the
land; and
threescore
men of the
people of the
land, that were
found in the
midst of the
city.\FTNT{*}{{\FR 52.25 }
were near…: Heb. saw the face of the king}\FTNT{†}{{\FR 52.25 }
principal…: or, scribe of the captain of the host}
\VS{26}So
Nebuzar-adan the
captain of the
guard
took them, and
brought them to the
king of
Babylon to
Riblah.
\VS{27}And the
king of
Babylon
smote them, and put them to
death in
Riblah in the
land of
Hamath. Thus
Judah was carried away
captive out of his own
land.
\VS{28}This
{\ADD{is}} the
people whom
Nebuchadrezzar carried away
captive: in the
seventh
year
three
thousand
Jews and
three and
twenty:
\VS{29}In the
eighteenth
year of
Nebuchadrezzar he carried away
captive from
Jerusalem
eight
hundred
thirty and
two
persons:\FTNT{‡}{{\FR 52.29 }
persons: Heb. souls}
\VS{30}In the
three and
twentieth
year of
Nebuchadrezzar
Nebuzar-adan the
captain of the
guard carried away
captive of the
Jews
seven
hundred
forty and
five
persons: all the
persons
{\ADD{were}}
four
thousand and
six
hundred.
\par }{\PP \VS{31}¶ And it came to pass in the
seven and
thirtieth
year of the
captivity of
Jehoiachin
king of
Judah, in the
twelfth
month, in the
five and
twentieth
{\ADD{day}} of the
month,
{\ADD{that}}
Evil-merodach
king of
Babylon in the
{\ADD{first}}
year of his
reign lifted
up the
head of
Jehoiachin
king of
Judah, and brought him
forth out of
prison,
\VS{32}And
spake
kindly unto him, and
set his
throne
above the
throne of the
kings that
{\ADD{were}} with him in
Babylon,\FTNT{§}{{\FR 52.32 }
kindly…: Heb. good things with him}
\VS{33}And
changed his
prison
garments: and he did
continually
eat
bread
before him all the
days of his
life.
\VS{34}And
{\ADD{for}} his
diet, there was a
continual
diet
given him of the
king of
Babylon,
every
day a
portion until the
day of his
death, all the
days of his
life.\FTNT{**}{{\FR 52.34 }
every…: Heb. the matter of the day in his day}
\par }