\NormalFont\ShortTitle{Job}
{\MT The Book of Job

\par }\ChapOne{1}{\PP \VerseOne{1}There was a
man in the
land of
Uz, whose
name
{\ADD{was}}
Job; and that
man was
perfect and
upright, and one that
feared
God, and
eschewed
evil.
\VS{2}And there were
born unto him
seven
sons and
three
daughters.
\VS{3}His
substance also was
seven
thousand
sheep, and
three
thousand
camels, and
five
hundred
yoke of
oxen, and
five
hundred she
asses, and a
very
great
household; so that this
man was the
greatest of all the
men of the
east.\FTNT{*}{{\FR 1.3 }
substance: or, cattle}\FTNT{†}{{\FR 1.3 }
household: or, husbandry}\FTNT{‡}{{\FR 1.3 }
men: Heb. sons}
\par }{\PP \VS{4}And his
sons
went and
feasted
{\ADD{in their}}
houses, every
one his
day; and
sent and
called for their
three
sisters to
eat and to
drink with them.
\VS{5}And it was so, when the
days of
{\ADD{their}}
feasting were gone
about, that
Job
sent and
sanctified them, and rose up
early in the
morning, and
offered burnt
offerings
{\ADD{according}} to the
number of them all: for
Job
said, It may
be that my
sons have
sinned, and
cursed
God in their
hearts. Thus
did
Job
continually.\FTNT{§}{{\FR 1.5 }
continually: Heb. all the days}
\par }{\PP \VS{6}¶ Now there was a
day when the
sons of
God
came to
present themselves before the
{\ND{
LORD}}, and
Satan
came also
among them.\FTNT{**}{{\FR 1.6 }
Satan: Heb. the adversary}\FTNT{††}{{\FR 1.6 }
among: Heb. in the midst of}
\VS{7}And the
{\ND{
LORD}}
said unto
Satan,
Whence
comest thou? Then
Satan
answered the
{\ND{
LORD}}, and
said, From going to and
fro in the
earth, and from walking up and
down in it.
\VS{8}And the
{\ND{
LORD}}
said unto
Satan, Hast thou
considered my
servant
Job, that
{\ADD{there is}} none like him in the
earth, a
perfect and an
upright
man, one that
feareth
God, and
escheweth
evil?\FTNT{‡‡}{{\FR 1.8 }
considered: Heb. set thy heart on}
\VS{9}Then
Satan
answered the
{\ND{
LORD}}, and
said, Doth
Job
fear
God for
nought?
\VS{10}Hast not thou made an
hedge about him, and
about his
house, and about all that he hath on every
side? thou hast
blessed the
work of his
hands, and his
substance is
increased in the
land.\FTNT{§§}{{\FR 1.10 }
substance: or, cattle}
\VS{11}But put
forth thine
hand now, and
touch all that he hath, and he will
curse thee to thy
face.\FTNT{***}{{\FR 1.11 }
and he…: Heb. if he curse thee not to thy face}
\VS{12}And the
{\ND{
LORD}}
said unto
Satan, Behold, all that he hath
{\ADD{is}} in thy
power; only upon himself put not
forth thine
hand. So
Satan went
forth from the
presence of the
{\ND{
LORD}}.\FTNT{†††}{{\FR 1.12 }
power: Heb. hand}
\par }{\PP \VS{13}¶ And there was a
day when his
sons and his
daughters
{\ADD{were}}
eating and
drinking
wine in their
eldest
brother’s
house:
\VS{14}And there
came a
messenger unto
Job, and
said, The
oxen were
plowing, and the
asses
feeding
beside them:
\VS{15}And the
Sabeans
fell
{\ADD{upon them}}, and took them
away; yea, they have
slain the
servants with the
edge of the
sword; and I only am
escaped alone to
tell thee.
\VS{16}While he
{\ADD{was}} yet
speaking, there
came also another, and
said, The
fire of
God is
fallen from
heaven, and hath burned
up the
sheep, and the
servants, and
consumed them; and I only am
escaped alone to
tell thee.\FTNT{‡‡‡}{{\FR 1.16 }
The fire…: or, A great fire}
\VS{17}While he
{\ADD{was}} yet
speaking, there
came also another, and
said, The
Chaldeans made
out
three
bands, and
fell upon the
camels, and have carried them
away, yea, and
slain the
servants with the
edge of the
sword; and I only am
escaped alone to
tell thee.\FTNT{§§§}{{\FR 1.17 }
fell: Heb. rushed}
\VS{18}While he
{\ADD{was}} yet
speaking, there
came also another, and
said, Thy
sons and thy
daughters
{\ADD{were}}
eating and
drinking
wine in their
eldest
brother’s
house:
\VS{19}And, behold, there
came a
great
wind
from the
wilderness, and
smote the
four
corners of the
house, and it
fell upon the young
men, and they are
dead; and I only am
escaped alone to
tell thee.\FTNT{*}{{\FR 1.19 }
from…: Heb. from aside, etc}
\par }{\PP \VS{20}Then
Job
arose, and
rent his
mantle, and
shaved his
head, and fell
down upon the
ground, and
worshipped,\FTNT{†}{{\FR 1.20 }
mantle: or, robe}
\VS{21}And
said,
Naked came I
out of my
mother’s
womb, and
naked shall I
return thither: the
{\ND{
LORD}}
gave, and the
{\ND{
LORD}} hath taken
away;
blessed be the
name of the
{\ND{
LORD}}.
\VS{22}In all this
Job
sinned not, nor
charged
God
foolishly.\FTNT{‡}{{\FR 1.22 }
charged…: or, attributed folly to God}

\par }\Chap{2}{\PP \VerseOne{1}Again there was a
day when the
sons of
God
came to
present themselves before the
{\ND{
LORD}}, and
Satan
came also
among them to
present himself before the
{\ND{
LORD}}.
\VS{2}And the
{\ND{
LORD}}
said unto
Satan, From
whence
comest thou? And
Satan
answered the
{\ND{
LORD}}, and
said, From going to and
fro in the
earth, and from walking up and
down in it.
\VS{3}And the
{\ND{
LORD}}
said unto
Satan, Hast
thou
considered my
servant
Job, that
{\ADD{there is}} none like him in the
earth, a
perfect and an
upright
man, one that
feareth
God, and
escheweth
evil? and still he holdeth
fast his
integrity, although thou
movedst me against him, to
destroy him without
cause.\FTNT{*}{{\FR 2.3 }
to destroy…: Heb. to swallow him up}
\VS{4}And
Satan
answered the
{\ND{
LORD}}, and
said,
Skin for
skin, yea, all that a
man hath will he
give for his
life.
\VS{5}But put
forth thine
hand now, and
touch his
bone and his
flesh, and he will
curse thee to thy
face.
\VS{6}And the
{\ND{
LORD}}
said unto
Satan, Behold, he
{\ADD{is}} in thine
hand; but
save his
life.\FTNT{†}{{\FR 2.6 }
but: or, only}
\par }{\PP \VS{7}¶ So
went
Satan
forth from the
presence of the
{\ND{
LORD}}, and
smote
Job with
sore
boils from the
sole of his
foot unto his
crown.
\VS{8}And he
took him a
potsherd to
scrape himself withal; and he sat
down
among the
ashes.
\par }{\PP \VS{9}¶ Then
said his
wife unto him, Dost thou still
retain thine
integrity?
curse
God, and
die.
\VS{10}But he
said unto her, Thou
speakest as
one of the foolish
women
speaketh.
What? shall we
receive
good at the hand of
God, and shall we not
receive
evil? In all this did not
Job
sin with his
lips.
\par }{\PP \VS{11}¶ Now when
Job’s
three
friends
heard of all this
evil that was
come upon him, they
came every
one from his own
place;
Eliphaz the
Temanite, and
Bildad the
Shuhite, and
Zophar the
Naamathite: for they had made an
appointment
together to
come to
mourn with him and to
comfort him.
\VS{12}And when they lifted
up their
eyes afar
off, and
knew him not, they lifted
up their
voice, and
wept; and they
rent every
one his
mantle, and
sprinkled
dust upon their
heads toward
heaven.
\VS{13}So they sat
down with him upon the
ground
seven
days and
seven
nights, and none
spake a
word unto him: for they
saw that
{\ADD{his}}
grief was
very
great.

\par }\Chap{3}{\PP \VerseOne{1}After this
opened
Job his
mouth, and
cursed his
day.
\VS{2}And
Job
spake, and
said,\FTNT{*}{{\FR 3.2 }
spake: Heb. answered}
\VS{3}Let the
day
perish wherein I was
born, and the
night
{\ADD{in which}} it was
said, There is a man
child
conceived.
\VS{4}Let that
day be
darkness; let not
God
regard it from
above, neither let the
light
shine upon it.
\VS{5}Let
darkness and the shadow of
death
stain it; let a
cloud
dwell upon it; let the
blackness of the
day
terrify it.\FTNT{†}{{\FR 3.5 }
stain: or, challenge}\FTNT{‡}{{\FR 3.5 }
let the…: or, let them terrify it, as those who have a bitter day}
\VS{6}As
{\ADD{for}} that
night, let
darkness
seize upon it; let it not be
joined unto the
days of the
year, let it not
come into the
number of the
months.\FTNT{§}{{\FR 3.6 }
let it not be…: or, let it not rejoice among the days}
\VS{7}Lo, let that
night be
solitary, let no joyful
voice
come therein.
\VS{8}Let them
curse it that
curse the
day, who are
ready to raise
up their
mourning.\FTNT{**}{{\FR 3.8 }
their…: or, leviathan}
\VS{9}Let the
stars of the
twilight thereof be
dark; let it
look for
light, but
{\ADD{have}} none; neither let it
see the
dawning of the
day:\FTNT{††}{{\FR 3.9 }
the dawning…: Heb. the eyelids of the morning}
\VS{10}Because it shut not
up the
doors of my
{\ADD{mother’s}}
womb, nor
hid
sorrow from mine
eyes.
\par }{\PP \VS{11}Why
died I not from the
womb?
{\ADD{why}} did I
{\ADD{not}} give up the
ghost when I came
out of the
belly?
\VS{12}Why did the
knees
prevent me? or why the
breasts that I should
suck?
\VS{13}For now should I have lain
still and been
quiet, I should have
slept: then had I been at
rest,
\VS{14}With
kings and
counsellors of the
earth, which
built desolate
places for themselves;
\VS{15}Or with
princes that had
gold, who
filled their
houses with
silver:
\VS{16}Or as an
hidden untimely
birth I had not been; as
infants
{\ADD{which}} never
saw
light.
\VS{17}There the
wicked
cease
{\ADD{from}}
troubling; and there the
weary be at
rest.\FTNT{‡‡}{{\FR 3.17 }
weary: Heb. wearied in strength}
\VS{18}{\ADD{There}} the
prisoners
rest
together; they
hear not the
voice of the
oppressor.
\VS{19}The
small and
great are there; and the
servant
{\ADD{is}}
free from his
master.
\par }{\PP \VS{20}Wherefore is
light
given to him that is in
misery, and
life unto the
bitter
{\ADD{in}}
soul;
\VS{21}Which
long for
death, but it
{\ADD{cometh}} not; and
dig for it more than for hid
treasures;\FTNT{§§}{{\FR 3.21 }
long: Heb. wait}
\VS{22}Which
rejoice
exceedingly,
{\ADD{and}} are
glad, when they can
find the
grave?
\VS{23}{\ADD{Why is light given}} to a
man whose
way is
hid, and whom
God hath hedged
in?
\VS{24}For my
sighing
cometh
before I
eat, and my
roarings are poured
out like the
waters.\FTNT{***}{{\FR 3.24 }
I eat: Heb. my meat}
\VS{25}For the thing which I
greatly
feared is
come upon me, and that which I was
afraid of is
come unto me.\FTNT{†††}{{\FR 3.25 }
the thing…: Heb. I feared a fear, and it came upon me}
\VS{26}I was not in
safety, neither had I
rest, neither was I
quiet; yet
trouble
came.

\par }\Chap{4}{\PP \VerseOne{1}Then
Eliphaz the
Temanite
answered and
said,
\VS{2}{\ADD{If}} we
assay to
commune with thee, wilt thou be
grieved? but who
can
withhold himself from
speaking?\FTNT{*}{{\FR 4.2 }
to…: Heb. a word}\FTNT{†}{{\FR 4.2 }
withhold…: Heb. refrain from words?}
\VS{3}Behold, thou hast
instructed
many, and thou hast
strengthened the
weak
hands.
\VS{4}Thy
words have
upholden him that was
falling, and thou hast
strengthened the
feeble
knees.
\VS{5}But now it is
come upon thee, and thou
faintest; it
toucheth thee, and thou art
troubled.
\VS{6}{\ADD{Is}} not
{\ADD{this}} thy
fear, thy
confidence, thy
hope, and the
uprightness of thy
ways?
\par }{\PP \VS{7}Remember, I pray thee, who
{\ADD{ever}}
perished, being
innocent? or
where were the
righteous cut
off?
\VS{8}Even as I have
seen, they that
plow
iniquity, and
sow
wickedness,
reap the same.
\VS{9}By the
blast of
God they
perish, and by the
breath of his
nostrils are they
consumed.\FTNT{‡}{{\FR 4.9 }
by the breath…: that is, by his anger}
\VS{10}The
roaring of the
lion, and the
voice of the fierce
lion, and the
teeth of the young
lions, are
broken.
\VS{11}The old
lion
perisheth for
lack of
prey, and the stout
lion’s
whelps are scattered
abroad.
\par }{\PP \VS{12}Now a
thing was secretly
brought to me, and mine
ear
received a
little thereof.\FTNT{§}{{\FR 4.12 }
secretly: Heb. by stealth}
\VS{13}In
thoughts from the
visions of the
night, when deep
sleep
falleth on
men,
\VS{14}Fear
came upon me, and
trembling, which made
all my
bones to
shake.\FTNT{**}{{\FR 4.14 }
came…: Heb. met}\FTNT{††}{{\FR 4.14 }
all: Heb. the multitude of}
\VS{15}Then a
spirit
passed before my
face; the
hair of my
flesh stood
up:
\VS{16}It stood
still, but I could not
discern the
form thereof: an
image
{\ADD{was}} before mine
eyes,
{\ADD{there was}}
silence, and I
heard a
voice,
{\ADD{saying}},\FTNT{‡‡}{{\FR 4.16 }
there…: or, I heard a still voice}
\VS{17}Shall mortal
man be more
just than
God? shall a
man be more
pure than his
maker?
\VS{18}Behold, he put no
trust in his
servants; and his
angels he
charged with
folly:\FTNT{§§}{{\FR 4.18 }
and his…: or, nor in his angels, in whom he put light}
\VS{19}How much
less
{\ADD{in}} them that
dwell in
houses of
clay, whose
foundation
{\ADD{is}} in the
dust,
{\ADD{which}} are
crushed
before the
moth?
\VS{20}They are
destroyed from
morning to
evening: they
perish for
ever without any
regarding
{\ADD{it}}.\FTNT{***}{{\FR 4.20 }
destroyed: Heb. beaten in pieces}
\VS{21}Doth not their
excellency
{\ADD{which is}} in them go
away? they
die, even without
wisdom.

\par }\Chap{5}{\PP \VerseOne{1}Call now, if there
be any that will
answer thee; and to which of the
saints wilt thou
turn?\FTNT{*}{{\FR 5.1 }
turn: or, look?}
\VS{2}For
wrath
killeth the foolish
man, and
envy
slayeth the silly
one.\FTNT{†}{{\FR 5.2 }
envy: or, indignation}
\VS{3}I have
seen the
foolish taking
root: but
suddenly I
cursed his
habitation.
\VS{4}His
children are
far from
safety, and they are
crushed in the
gate, neither
{\ADD{is there}} any to
deliver
{\ADD{them}}.
\VS{5}Whose
harvest the
hungry eateth
up, and
taketh it even
out of the
thorns, and the
robber swalloweth
up their
substance.
\par }{\PP \VS{6}Although
affliction cometh not
forth of the
dust, neither doth
trouble spring
out of the
ground;\FTNT{‡}{{\FR 5.6 }
affliction: or, iniquity}
\VS{7}Yet
man is
born unto
trouble, as the
sparks
fly
upward.\FTNT{§}{{\FR 5.7 }
trouble: or, labour}\FTNT{**}{{\FR 5.7 }
sparks…: Heb. the sons of the burning coal lift up to fly}
\VS{8}I would
seek unto
God, and unto
God would I
commit my
cause:
\VS{9}Which
doeth great
things and
unsearchable; marvellous
things without
number:\FTNT{††}{{\FR 5.9 }
unsearchable: Heb. there is no search}\FTNT{‡‡}{{\FR 5.9 }
without…: Heb. till there be no number}
\VS{10}Who
giveth
rain
upon the
earth, and
sendeth
waters
upon the
fields:\FTNT{§§}{{\FR 5.10 }
fields: Heb. outplaces}
\VS{11}To set
up on
high those that be
low; that those which
mourn may be
exalted to
safety.
\VS{12}He
disappointeth the
devices of the
crafty, so that their
hands cannot
perform
{\ADD{their}}
enterprise.\FTNT{***}{{\FR 5.12 }
their enterprise: or, any thing}
\VS{13}He
taketh the
wise in their own
craftiness: and the
counsel of the
froward is carried
headlong.
\VS{14}They
meet with
darkness in the
daytime, and
grope in the
noonday as in the
night.\FTNT{†††}{{\FR 5.14 }
meet…: or, run into}
\VS{15}But he
saveth the
poor from the
sword, from their
mouth, and from the
hand of the
mighty.
\VS{16}So the
poor hath
hope, and
iniquity
stoppeth her
mouth.
\par }{\PP \VS{17}Behold,
happy
{\ADD{is}} the
man whom
God
correcteth: therefore
despise not thou the
chastening of the
Almighty:
\VS{18}For he maketh
sore, and bindeth
up: he
woundeth, and his
hands make
whole.
\VS{19}He shall
deliver thee in
six
troubles: yea, in
seven there shall no
evil
touch thee.
\VS{20}In
famine he shall
redeem thee from
death: and in
war from the
power of the
sword.\FTNT{‡‡‡}{{\FR 5.20 }
power: Heb. hands}
\VS{21}Thou shalt be
hid from the
scourge of the
tongue: neither shalt thou be
afraid of
destruction when it
cometh.\FTNT{§§§}{{\FR 5.21 }
from…: or, when the tongue scourgeth}
\VS{22}At
destruction and
famine thou shalt
laugh: neither shalt thou be
afraid of the
beasts of the
earth.
\VS{23}For thou shalt be in
league with the
stones of the
field: and the
beasts of the
field shall be at
peace with thee.
\VS{24}And thou shalt
know that thy
tabernacle
{\ADD{shall be}} in
peace; and thou shalt
visit thy
habitation, and shalt not
sin.\FTNT{*}{{\FR 5.24 }
thy tabernacle…: or, peace is thy tabernacle}\FTNT{†}{{\FR 5.24 }
sin: or, err}
\VS{25}Thou shalt
know also that thy
seed
{\ADD{shall be}}
great, and thine
offspring as the
grass of the
earth.\FTNT{‡}{{\FR 5.25 }
great: or, much}
\VS{26}Thou shalt
come to
{\ADD{thy}}
grave in a full
age, like as a shock of
corn cometh
in in his
season.\FTNT{§}{{\FR 5.26 }
cometh in: Heb. ascendeth}
\VS{27}Lo this, we have
searched it, so it
{\ADD{is;}}
hear it, and
know thou
{\ADD{it}} for thy good.\FTNT{**}{{\FR 5.27 }
for…: Heb. for thyself}

\par }\Chap{6}{\PP \VerseOne{1}But
Job
answered and
said,
\VS{2}Oh
that my
grief were
throughly
weighed, and my
calamity
laid in the
balances
together!\FTNT{*}{{\FR 6.2 }
laid: Heb. lifted up}
\VS{3}For now it would be
heavier than the
sand of the
sea: therefore my
words are swallowed
up.\FTNT{†}{{\FR 6.3 }
my words…: that is, I want words to express my grief}
\VS{4}For the
arrows of the
Almighty
{\ADD{are}}
within me, the
poison whereof drinketh
up my
spirit: the
terrors of
God do set themselves in
array against me.
\VS{5}Doth the wild
ass
bray when he hath
grass? or
loweth the
ox over his
fodder?\FTNT{‡}{{\FR 6.5 }
when he…: Heb. at grass?}
\VS{6}Can that which is
unsavoury be
eaten without
salt? or is
there
{\ADD{any}}
taste in the
white of an
egg?
\VS{7}The things
{\ADD{that}} my
soul
refused to
touch
{\ADD{are}} as my
sorrowful
meat.
\par }{\PP \VS{8}Oh that I might
have my
request; and that
God would
grant
{\ADD{me}} the thing that I long
for!\FTNT{§}{{\FR 6.8 }
the thing…: Heb. my expectation}
\VS{9}Even that it would
please
God to
destroy me; that he would let
loose his
hand, and cut me
off!
\VS{10}Then should I yet have
comfort; yea, I would
harden myself in
sorrow: let him not
spare; for I have not
concealed the
words of the Holy
One.
\VS{11}What
{\ADD{is}} my
strength, that I should
hope? and what
{\ADD{is}} mine
end, that I should
prolong my
life?
\VS{12}{\ADD{Is}} my
strength the
strength of
stones? or
{\ADD{is}} my
flesh of
brass?\FTNT{**}{{\FR 6.12 }
of brass: Heb. brasen?}
\VS{13}{\ADD{Is}} not my
help in me? and is
wisdom driven
quite from me?
\par }{\PP \VS{14}To him that is
afflicted
pity
{\ADD{should be shewed}} from his
friend; but he
forsaketh the
fear of the
Almighty.\FTNT{††}{{\FR 6.14 }
is afflicted: Heb. melteth}
\VS{15}My
brethren have dealt
deceitfully as a
brook,
{\ADD{and}} as the
stream of
brooks they pass
away;
\VS{16}Which are
blackish by reason of the
ice,
{\ADD{and}} wherein the
snow is
hid:
\VS{17}What
time they wax
warm, they
vanish: when it is
hot, they are consumed
out of their
place.\FTNT{‡‡}{{\FR 6.17 }
vanish: Heb. are cut off}\FTNT{§§}{{\FR 6.17 }
when…: Heb. in the heat thereof}\FTNT{***}{{\FR 6.17 }
consumed: Heb. extinguished}
\VS{18}The
paths of their
way are turned
aside; they
go to
nothing, and
perish.
\VS{19}The
troops of
Tema
looked, the
companies of
Sheba
waited for them.
\VS{20}They were
confounded because they had
hoped; they
came thither, and were
ashamed.
\VS{21}For now ye are nothing; ye
see
{\ADD{my}} casting
down, and are
afraid.\FTNT{†††}{{\FR 6.21 }
ye are…: or, ye are like to them: Heb. to it}\FTNT{‡‡‡}{{\FR 6.21 }
nothing: Heb. not}
\par }{\PP \VS{22}Did I
say,
Bring unto me? or, Give a
reward for me of your
substance?
\VS{23}Or,
Deliver me from the
enemy’s
hand? or,
Redeem me from the
hand of the
mighty?
\VS{24}Teach me, and I will hold my
tongue: and cause me to
understand wherein I have
erred.
\VS{25}How
forcible are
right
words! but what doth your
arguing
reprove?
\VS{26}Do ye
imagine to
reprove
words, and the
speeches of one that is
desperate,
{\ADD{which are}} as
wind?
\VS{27}Yea, ye
overwhelm the
fatherless, and ye
dig
{\ADD{a pit}} for your
friend.\FTNT{§§§}{{\FR 6.27 }
ye overwhelm: Heb. ye cause to fall upon}
\VS{28}Now therefore be
content,
look upon me; for
{\ADD{it is}}
evident unto you if I
lie.\FTNT{*}{{\FR 6.28 }
evident…: Heb. before your face}
\VS{29}Return, I pray you, let it not be
iniquity; yea, return
again, my
righteousness
{\ADD{is}} in it.\FTNT{†}{{\FR 6.29 }
in it: that is, in this matter}
\VS{30}Is
there
iniquity in my
tongue? cannot my
taste
discern perverse
things?\FTNT{‡}{{\FR 6.30 }
my taste: Heb. my palate}

\par }\Chap{7}{\PP \VerseOne{1}{\ADD{Is there}} not an appointed
time to
man upon
earth?
{\ADD{are not}} his
days also like the
days of an
hireling?\FTNT{*}{{\FR 7.1 }
an appointed…: or, a warfare}
\VS{2}As a
servant earnestly
desireth the
shadow, and as an
hireling
looketh for
{\ADD{the reward of}} his
work:\FTNT{†}{{\FR 7.2 }
earnestly…: Heb. gapeth after}
\VS{3}So am I made to
possess
months of
vanity, and
wearisome
nights are
appointed to me.
\VS{4}When I lie
down, I
say, When shall I
arise, and the
night be
gone? and I am
full of tossings to and
fro unto the dawning of the
day.\FTNT{‡}{{\FR 7.4 }
the night…: Heb. the evening be measured?}
\VS{5}My
flesh is
clothed with
worms and
clods of
dust; my
skin is
broken, and become
loathsome.
\VS{6}My
days are
swifter than a
weaver’s shuttle, and are
spent
without
hope.
\par }{\PP \VS{7}O
remember that my
life
{\ADD{is}}
wind: mine
eye shall no
more
see
good.\FTNT{§}{{\FR 7.7 }
shall…: Heb. shall not return}\FTNT{**}{{\FR 7.7 }
see: to see, that is, to enjoy}
\VS{8}The
eye of him that hath
seen me shall
see me no
{\ADD{more:}} thine
eyes
{\ADD{are}} upon me, and I
{\ADD{am}} not.\FTNT{††}{{\FR 7.8 }
I am…: that is, I can live no longer}
\VS{9}{\ADD{As}} the
cloud is
consumed and vanisheth
away: so he that goeth
down to the
grave shall come
up no
{\ADD{more}}.
\VS{10}He shall
return no more to his
house, neither shall his
place
know him any more.
\VS{11}Therefore I will not
refrain my
mouth; I will
speak in the
anguish of my
spirit; I will
complain in the
bitterness of my
soul.
\VS{12}{\ADD{Am}} I a
sea, or a
whale, that thou
settest a
watch over me?
\VS{13}When I
say, My
bed shall
comfort me, my
couch shall
ease my
complaint;
\VS{14}Then thou
scarest me with
dreams, and
terrifiest me through
visions:
\VS{15}So that my
soul
chooseth
strangling,
{\ADD{and}}
death rather than my
life.\FTNT{‡‡}{{\FR 7.15 }
life: Heb. bones}
\VS{16}I
loathe
{\ADD{it;}} I would not
live
alway: let me
alone; for my
days
{\ADD{are}}
vanity.
\par }{\PP \VS{17}What
{\ADD{is}}
man, that thou shouldest
magnify him? and that thou shouldest
set thine
heart upon him?
\VS{18}And
{\ADD{that}} thou shouldest
visit him every
morning,
{\ADD{and}}
try him every
moment?
\VS{19}How
long wilt thou not
depart from me, nor let me
alone till I swallow
down my
spittle?
\VS{20}I have
sinned; what shall I
do unto thee, O thou
preserver of
men? why hast thou
set me as a
mark against thee, so that I am a
burden to myself?
\VS{21}And why dost thou not
pardon my
transgression, and take
away mine
iniquity? for now shall I
sleep in the
dust; and thou shalt seek me in the
morning, but I
{\ADD{shall}} not
{\ADD{be}}.

\par }\Chap{8}{\PP \VerseOne{1}Then
answered
Bildad the
Shuhite, and
said,
\VS{2}How
long wilt thou
speak these
{\ADD{things}}? and
{\ADD{how long shall}} the
words of thy
mouth
{\ADD{be like}} a
strong
wind?
\VS{3}Doth
God
pervert
judgment? or doth the
Almighty
pervert
justice?
\VS{4}If thy
children have
sinned against him, and he have cast them
away
for their
transgression;\FTNT{*}{{\FR 8.4 }
for…: Heb. in the hand of their transgression}
\VS{5}If thou wouldest
seek unto
God
betimes, and make thy
supplication to the
Almighty;
\VS{6}If thou
{\ADD{wert}}
pure and
upright; surely now he would
awake for thee, and make the
habitation of thy
righteousness
prosperous.
\VS{7}Though thy
beginning was
small, yet thy latter
end should
greatly
increase.
\par }{\PP \VS{8}For
enquire, I pray thee, of the
former
age, and
prepare thyself to the
search of their
fathers:
\VS{9}(For we
{\ADD{are but of}}
yesterday, and
know nothing, because our
days upon
earth
{\ADD{are}} a
shadow:)\FTNT{†}{{\FR 8.9 }
nothing: Heb. not}
\VS{10}Shall not they
teach thee,
{\ADD{and}}
tell thee, and
utter
words out of their
heart?
\VS{11}Can the
rush grow
up without
mire?
can the
flag
grow
without
water?
\VS{12}Whilst it
{\ADD{is}} yet in his
greenness,
{\ADD{and}} not cut
down, it
withereth
before any
{\ADD{other}}
herb.
\VS{13}So
{\ADD{are}} the
paths of all that
forget
God; and the
hypocrite’s
hope shall
perish:
\VS{14}Whose
hope shall be cut
off, and whose
trust
{\ADD{shall be}} a
spider’s
web.\FTNT{‡}{{\FR 8.14 }
web: Heb. house}
\VS{15}He shall
lean upon his
house, but it shall not
stand: he shall
hold it fast, but it shall not
endure.
\VS{16}He
{\ADD{is}}
green
before the
sun, and his
branch shooteth
forth in his
garden.
\VS{17}His
roots are wrapped
about the
heap,
{\ADD{and}}
seeth the
place of
stones.
\VS{18}If he
destroy him from his
place, then
{\ADD{it}} shall
deny him,
{\ADD{saying}}, I have not
seen thee.
\VS{19}Behold, this
{\ADD{is}} the
joy of his
way, and out of the
earth shall
others
grow.
\par }{\PP \VS{20}Behold,
God will not cast
away a
perfect
{\ADD{man}}, neither will he
help the evil
doers:\FTNT{§}{{\FR 8.20 }
help…: Heb. take the ungodly by the hand}\FTNT{**}{{\FR 8.20 }
come: Heb. shall not be}
\VS{21}Till he
fill thy
mouth with
laughing, and thy
lips with
rejoicing.\FTNT{††}{{\FR 8.21 }
rejoicing: Heb. shouting for joy}
\VS{22}They that
hate thee shall be
clothed with
shame; and the
dwelling place of the
wicked shall come to
nought.\FTNT{‡‡}{{\FR 8.22 }
shall come…: Heb. shall not be}

\par }\Chap{9}{\PP \VerseOne{1}Then
Job
answered and
said,
\VS{2}I
know
{\ADD{it is}} so of a
truth: but how should
man be
just with
God?\FTNT{*}{{\FR 9.2 }
with God: or, before God?}
\VS{3}If he
will
contend with him, he cannot
answer him
one of a
thousand.
\VS{4}{\ADD{He is}}
wise in
heart, and
mighty in
strength: who hath
hardened
{\ADD{himself}} against him, and hath
prospered?
\VS{5}Which
removeth the
mountains, and they
know not: which
overturneth them in his
anger.
\VS{6}Which
shaketh the
earth out of her
place, and the
pillars thereof
tremble.
\VS{7}Which
commandeth the
sun, and it
riseth not; and sealeth
up the
stars.
\VS{8}Which alone spreadeth
out the
heavens, and
treadeth upon the
waves of the
sea.\FTNT{†}{{\FR 9.8 }
waves: Heb. heights}
\VS{9}Which
maketh
Arcturus,
Orion, and
Pleiades, and the
chambers of the
south.\FTNT{‡}{{\FR 9.9 }
Arcturus…: Heb. Ash, Cesil, and Cimah}
\VS{10}Which
doeth great
things
past finding
out; yea, and
wonders without
number.
\VS{11}Lo, he
goeth by me, and I
see
{\ADD{him}} not: he passeth
on also, but I
perceive him not.
\VS{12}Behold, he taketh
away, who can
hinder him? who will
say unto him, What
doest thou?\FTNT{§}{{\FR 9.12 }
hinder…: Heb. turn him away?}
\VS{13}{\ADD{If}}
God will not
withdraw his
anger, the
proud
helpers do
stoop under him.\FTNT{**}{{\FR 9.13 }
proud…: Heb. helpers of pride, or, strength}
\par }{\PP \VS{14}How much less shall I
answer him,
{\ADD{and}} choose
out my
words
{\ADD{to reason}} with him?
\VS{15}Whom, though I were
righteous,
{\ADD{yet}} would I not
answer,
{\ADD{but}} I would make
supplication to my
judge.
\VS{16}If I had
called, and he had
answered me;
{\ADD{yet}} would I not
believe that he had
hearkened unto my
voice.
\VS{17}For he
breaketh me with a
tempest, and
multiplieth my
wounds without
cause.
\VS{18}He will not
suffer me to
take my
breath, but
filleth me with
bitterness.
\VS{19}If
{\ADD{I speak}} of
strength, lo,
{\ADD{he is}}
strong: and if of
judgment, who shall
set me a time
{\ADD{to plead}}?
\VS{20}If I
justify myself, mine own
mouth shall
condemn me:
{\ADD{if I say}}, I
{\ADD{am}}
perfect, it shall also prove me
perverse.
\VS{21}{\ADD{Though}} I
{\ADD{were}}
perfect,
{\ADD{yet}} would I not
know my
soul: I would
despise my
life.
\par }{\PP \VS{22}This
{\ADD{is}}
one
{\ADD{thing}}, therefore I
said
{\ADD{it}}, He
destroyeth the
perfect and the
wicked.
\VS{23}If the
scourge
slay
suddenly, he will
laugh at the
trial of the
innocent.
\VS{24}The
earth is
given into the
hand of the
wicked: he
covereth the
faces of the
judges thereof; if not,
where,
{\ADD{and}} who
{\ADD{is}} he?
\par }{\PP \VS{25}Now my
days are
swifter than a
post: they flee
away, they
see no
good.
\VS{26}They are passed
away
as the
swift
ships: as the
eagle
{\ADD{that}}
hasteth to the
prey.\FTNT{††}{{\FR 9.26 }
swift…: or, ships of Ebeh: Heb. ships of desire}
\VS{27}If I
say, I will
forget my
complaint, I will leave
off my
heaviness, and
comfort
{\ADD{myself:}}
\VS{28}I am
afraid of all my
sorrows, I
know that thou wilt not hold me
innocent.
\VS{29}{\ADD{If}} I be
wicked, why then
labour I in
vain?
\VS{30}If I
wash myself
with snow
water, and make my
hands
never so
clean;
\VS{31}Yet shalt thou
plunge me in the
ditch, and mine own
clothes shall
abhor me.\FTNT{‡‡}{{\FR 9.31 }
abhor…: or, make me to be abhorred}
\VS{32}For
{\ADD{he is}} not a
man, as I
{\ADD{am, that}} I should
answer him,
{\ADD{and}} we should
come
together in
judgment.
\VS{33}Neither
is there any
daysman betwixt us,
{\ADD{that}} might
lay his
hand upon us
both.\FTNT{§§}{{\FR 9.33 }
any…: Heb. one that should argue}\FTNT{***}{{\FR 9.33 }
daysman: or, umpire}
\VS{34}Let him
take his
rod
away from me, and let not his
fear
terrify me:
\VS{35}{\ADD{Then}} would I
speak, and not
fear him; but
{\ADD{it is}} not so with me.\FTNT{†††}{{\FR 9.35 }
it is…: Heb. I am not so with myself}

\par }\Chap{10}{\PP \VerseOne{1}My
soul is
weary of my
life; I will
leave my
complaint upon myself; I will
speak in the
bitterness of my
soul.\FTNT{*}{{\FR 10.1 }
weary…: or, cut off while I live}
\VS{2}I will
say unto
God, Do not
condemn me;
shew me wherefore thou
contendest with me.
\VS{3}{\ADD{Is it}}
good unto thee that thou shouldest
oppress, that thou shouldest
despise the
work of thine
hands, and
shine upon the
counsel of the
wicked?\FTNT{†}{{\FR 10.3 }
work: Heb. labour}
\VS{4}Hast thou
eyes of
flesh? or
seest thou as
man
seeth?
\VS{5}{\ADD{Are}} thy
days as the
days of
man?
{\ADD{are}} thy
years as
man’s
days,
\VS{6}That thou
enquirest after mine
iniquity, and
searchest after my
sin?
\VS{7}Thou
knowest that I am not
wicked; and
{\ADD{there is}} none that can
deliver out of thine
hand.\FTNT{‡}{{\FR 10.7 }
Thou…: Heb. It is upon thy knowledge}
\par }{\PP \VS{8}Thine
hands have
made me and
fashioned me
together round
about; yet thou dost
destroy me.\FTNT{§}{{\FR 10.8 }
have…: Heb. took pains about me}
\VS{9}Remember, I beseech thee, that thou hast
made me as the
clay; and wilt thou
bring me into
dust
again?
\VS{10}Hast thou not poured me
out as
milk, and
curdled me like
cheese?
\VS{11}Thou hast
clothed me with
skin and
flesh, and hast
fenced me with
bones and
sinews.\FTNT{**}{{\FR 10.11 }
fenced: Heb. hedged}
\VS{12}Thou hast
granted me
life and
favour, and thy
visitation hath
preserved my
spirit.
\VS{13}And these
{\ADD{things}} hast thou
hid in thine
heart: I
know that this
{\ADD{is}} with thee.
\par }{\PP \VS{14}If I
sin, then thou
markest me, and thou wilt not
acquit me from mine
iniquity.
\VS{15}If I be
wicked,
woe unto me; and
{\ADD{if}} I be
righteous,
{\ADD{yet}} will I not lift
up my
head.
{\ADD{I am}}
full of
confusion; therefore
see thou mine
affliction;
\VS{16}For it
increaseth. Thou
huntest me as a fierce
lion: and
again thou shewest thyself
marvellous upon me.
\VS{17}Thou
renewest thy
witnesses against me, and
increasest thine
indignation upon
me;
changes and
war
{\ADD{are}} against me.\FTNT{††}{{\FR 10.17 }
witnesses: that is, plagues}
\VS{18}Wherefore then hast thou brought me
forth out of the
womb? Oh that I had given up the
ghost, and no
eye had
seen me!
\VS{19}I should have been as though I had not been; I should have been
carried from the
womb to the
grave.
\VS{20}{\ADD{Are}} not my
days
few?
cease
{\ADD{then, and}} let me
alone, that I may take
comfort a
little,
\VS{21}Before I
go
{\ADD{whence}} I shall not
return,
{\ADD{even}} to the
land of
darkness and the shadow of
death;
\VS{22}A
land of
darkness, as
darkness
{\ADD{itself; and}} of the shadow of
death, without any
order, and
{\ADD{where}} the
light
{\ADD{is}} as
darkness.

\par }\Chap{11}{\PP \VerseOne{1}Then
answered
Zophar the
Naamathite, and
said,
\VS{2}Should not the
multitude of
words be
answered? and should a
man full of
talk be
justified?\FTNT{*}{{\FR 11.2 }
full…: Heb. of lips}
\VS{3}Should thy
lies
make
men hold their
peace? and when thou
mockest, shall no man make thee
ashamed?\FTNT{†}{{\FR 11.3 }
lies: or, devices}
\VS{4}For thou hast
said, My
doctrine
{\ADD{is}}
pure, and I am
clean in thine
eyes.
\VS{5}But oh
that
God would
speak, and
open his
lips against thee;
\VS{6}And that he would
shew thee the
secrets of
wisdom, that
{\ADD{they are}}
double to that which
is!
Know therefore that
God
exacteth of thee
{\ADD{less}} than thine
iniquity
{\ADD{deserveth}}.
\par }{\PP \VS{7}Canst thou by
searching find
out
God? canst thou
find out the
Almighty unto
perfection?
\VS{8}{\ADD{It is}} as
high as
heaven; what canst thou
do?
deeper than
hell; what canst thou
know?\FTNT{‡}{{\FR 11.8 }
as high…: Heb. the heights of heaven}
\VS{9}The
measure thereof
{\ADD{is}}
longer than the
earth, and
broader than the
sea.
\VS{10}If he cut
off, and shut
up, or gather
together, then who can
hinder him?\FTNT{§}{{\FR 11.10 }
cut…: or, make a change}\FTNT{**}{{\FR 11.10 }
hinder…: Heb. turn him away?}
\VS{11}For he
knoweth
vain
men: he
seeth
wickedness also; will he not then
consider
{\ADD{it}}?
\VS{12}For
vain
man would be
wise, though
man be
born
{\ADD{like}} a wild
ass’s
colt.\FTNT{††}{{\FR 11.12 }
vain: Heb. empty}
\par }{\PP \VS{13}If thou
prepare thine
heart, and
stretch out thine
hands toward him;
\VS{14}If
iniquity
{\ADD{be}} in thine
hand, put it far
away, and let not
wickedness
dwell in thy
tabernacles.
\VS{15}For then shalt thou lift
up thy
face without
spot; yea, thou shalt be
stedfast, and shalt not
fear:
\VS{16}Because thou shalt
forget
{\ADD{thy}}
misery,
{\ADD{and}}
remember
{\ADD{it}} as
waters
{\ADD{that}} pass
away:
\VS{17}And
{\ADD{thine}}
age shall be
clearer than the
noonday; thou shalt shine
forth, thou shalt be as the
morning.\FTNT{‡‡}{{\FR 11.17 }
be clearer…: Heb. arise above}
\VS{18}And thou shalt be
secure, because there
is
hope; yea, thou shalt
dig
{\ADD{about thee, and}} thou shalt take thy
rest in
safety.
\VS{19}Also thou shalt lie
down, and none shall make
{\ADD{thee}}
afraid; yea,
many shall make
suit unto
thee.\FTNT{§§}{{\FR 11.19 }
make suit…: Heb. intreat thy face}
\VS{20}But the
eyes of the
wicked shall
fail, and they shall not
escape, and their
hope
{\ADD{shall be as}} the giving
up of the
ghost.\FTNT{***}{{\FR 11.20 }
they shall…: Heb. flight shall perish from them}\FTNT{†††}{{\FR 11.20 }
the giving…: or, a puff of breath}

\par }\Chap{12}{\PP \VerseOne{1}And
Job
answered and
said,
\VS{2}No
doubt but ye
{\ADD{are}} the
people, and
wisdom shall
die with you.
\VS{3}But I have
understanding as well as you; I
{\ADD{am}} not
inferior to you: yea, who knoweth not such
things as these?\FTNT{*}{{\FR 12.3 }
understanding: Heb. an heart}\FTNT{†}{{\FR 12.3 }
I am…: Heb. I fall not lower than you}\FTNT{‡}{{\FR 12.3 }
who…: Heb. with whom are not such as these?}
\VS{4}I am
{\ADD{as}} one
mocked of his
neighbour, who
calleth upon
God, and he
answereth him: the
just
upright
{\ADD{man is}} laughed to
scorn.
\VS{5}He that is
ready to
slip with
{\ADD{his}}
feet
{\ADD{is as}} a
lamp
despised in the
thought of him that is at
ease.
\par }{\PP \VS{6}The
tabernacles of
robbers
prosper, and they that
provoke
God are
secure; into whose
hand
God
bringeth
{\ADD{abundantly}}.
\VS{7}But
ask now the
beasts, and they shall
teach thee; and the
fowls of the
air, and they shall
tell thee:
\VS{8}Or
speak to the
earth, and it shall
teach thee: and the
fishes of the
sea shall
declare unto thee.
\VS{9}Who
knoweth not in all these that the
hand of the
{\ND{
LORD}} hath
wrought this?
\VS{10}In whose
hand
{\ADD{is}} the
soul of every living
thing, and the
breath of all
mankind.\FTNT{§}{{\FR 12.10 }
soul: or, life}\FTNT{**}{{\FR 12.10 }
all…: Heb. all flesh of man}
\VS{11}Doth not the
ear
try
words? and the
mouth
taste his
meat?\FTNT{††}{{\FR 12.11 }
mouth: Heb. palate}
\par }{\PP \VS{12}With the
ancient
{\ADD{is}}
wisdom; and in
length of
days
understanding.
\VS{13}With him
{\ADD{is}}
wisdom and
strength, he hath
counsel and
understanding.\FTNT{‡‡}{{\FR 12.13 }
With…: that is, With God}
\VS{14}Behold, he breaketh
down, and it cannot be built
again: he shutteth
up a
man, and there can be no
opening.\FTNT{§§}{{\FR 12.14 }
up: Heb. upon}
\VS{15}Behold, he
withholdeth the
waters, and they dry
up: also he sendeth them
out, and they
overturn the
earth.
\VS{16}With him
{\ADD{is}}
strength and
wisdom: the
deceived and the
deceiver
{\ADD{are}} his.
\VS{17}He
leadeth
counsellors away
spoiled, and maketh the
judges
fools.
\VS{18}He
looseth the
bond of
kings, and
girdeth their
loins with a
girdle.
\VS{19}He
leadeth
princes away
spoiled, and
overthroweth the
mighty.
\VS{20}He removeth
away the
speech of the
trusty, and taketh
away the
understanding of the
aged.\FTNT{***}{{\FR 12.20 }
speech…: Heb. lip of the faithful}
\VS{21}He
poureth
contempt upon
princes, and
weakeneth the
strength of the
mighty.\FTNT{†††}{{\FR 12.21 }
weakeneth…: or, looseth the girdle of the strong}
\VS{22}He
discovereth deep
things out of
darkness, and bringeth
out to
light the shadow of
death.
\VS{23}He
increaseth the
nations, and
destroyeth them: he
enlargeth the
nations, and
straiteneth them
{\ADD{again}}.\FTNT{‡‡‡}{{\FR 12.23 }
straiteneth: Heb. leadeth in}
\VS{24}He taketh
away the
heart of the
chief of the
people of the
earth, and causeth them to
wander in a
wilderness
{\ADD{where there is}} no
way.
\VS{25}They
grope in the
dark without
light, and he maketh them to
stagger like
{\ADD{a}}
drunken
{\ADD{man}}.\FTNT{§§§}{{\FR 12.25 }
stagger: Heb. wander}

\par }\Chap{13}{\PP \VerseOne{1}Lo, mine
eye hath
seen all
{\ADD{this}}, mine
ear hath
heard and
understood it.
\VS{2}What ye
know,
{\ADD{the same}} do I
know also: I
{\ADD{am}} not
inferior unto you.
\VS{3}Surely I would
speak to the
Almighty, and I
desire to
reason with
God.
\VS{4}But ye
{\ADD{are}}
forgers of
lies, ye
{\ADD{are}} all
physicians of no
value.
\VS{5}O
that ye would
altogether hold your
peace! and it should be your
wisdom.
\VS{6}Hear now my
reasoning, and
hearken to the
pleadings of my
lips.
\VS{7}Will ye
speak
wickedly for
God? and
talk
deceitfully for him?
\VS{8}Will ye
accept his
person? will ye
contend for
God?
\VS{9}Is it
good that he should search you
out? or as one
man
mocketh another, do ye
{\ADD{so}}
mock him?
\VS{10}He will
surely
reprove you, if ye do
secretly
accept
persons.
\VS{11}Shall not his
excellency make you
afraid? and his
dread
fall upon you?
\VS{12}Your
remembrances
{\ADD{are}}
like unto
ashes, your
bodies to
bodies of
clay.
\par }{\PP \VS{13}Hold your
peace, let me alone, that I may
speak, and let
come on me what
{\ADD{will}}.\FTNT{*}{{\FR 13.13 }
Hold…: Heb. Be silent from me}
\VS{14}Wherefore do I
take my
flesh in my
teeth, and
put my
life in mine
hand?
\VS{15}Though he
slay me, yet will I
trust in him: but I will
maintain mine own
ways
before him.\FTNT{†}{{\FR 13.15 }
maintain: Heb. prove, or, argue}
\VS{16}He also
{\ADD{shall be}} my
salvation: for an
hypocrite shall not
come
before him.
\VS{17}Hear
diligently my
speech, and my
declaration with your
ears.
\VS{18}Behold now, I have
ordered
{\ADD{my}}
cause; I
know that I shall be
justified.
\VS{19}Who
{\ADD{is}} he
{\ADD{that}} will
plead with me? for now, if I hold my
tongue, I shall give up the
ghost.
\VS{20}Only
do not
two
{\ADD{things}} unto me: then will I not
hide myself from
thee.
\VS{21}Withdraw thine
hand
far from me: and let not thy
dread make me
afraid.
\VS{22}Then
call thou, and I will
answer: or let me
speak, and
answer thou me.
\par }{\PP \VS{23}How many
{\ADD{are}} mine
iniquities and
sins? make me to
know my
transgression and my
sin.
\VS{24}Wherefore
hidest thou thy
face, and
holdest me for thine
enemy?
\VS{25}Wilt thou
break a
leaf driven to and
fro? and wilt thou
pursue the
dry
stubble?
\VS{26}For thou
writest bitter
things against me, and makest me to
possess the
iniquities of my
youth.
\VS{27}Thou
puttest my
feet also in the
stocks, and lookest
narrowly unto all my
paths; thou settest a
print upon the
heels of my
feet.\FTNT{‡}{{\FR 13.27 }
lookest…: Heb. observest}\FTNT{§}{{\FR 13.27 }
heels: Heb. roots}
\VS{28}And he, as a rotten
thing,
consumeth, as a
garment that is
moth
eaten.

\par }\Chap{14}{\PP \VerseOne{1}Man
{\ADD{that is}}
born of a
woman
{\ADD{is}} of
few
days, and
full of
trouble.\FTNT{*}{{\FR 14.1 }
few…: Heb. short of days}
\VS{2}He cometh
forth like a
flower, and is cut
down: he
fleeth also as a
shadow, and
continueth not.
\VS{3}And dost thou
open thine
eyes upon such an
one, and
bringest me into
judgment with thee?
\VS{4}Who can
bring a
clean
{\ADD{thing}} out of an
unclean? not
one.\FTNT{†}{{\FR 14.4 }
can…: Heb. will give}
\VS{5}Seeing his
days
{\ADD{are}}
determined, the
number of his
months
{\ADD{are}} with thee, thou hast
appointed his
bounds that he cannot
pass;
\VS{6}Turn from him, that he may
rest, till he shall
accomplish, as an
hireling, his
day.\FTNT{‡}{{\FR 14.6 }
rest: Heb. cease}
\par }{\PP \VS{7}For there
is
hope of a
tree, if it be cut
down, that it will sprout
again, and that the tender
branch thereof will not
cease.
\VS{8}Though the
root thereof wax
old in the
earth, and the
stock thereof
die in the
ground;
\VS{9}{\ADD{Yet}} through the
scent of
water it will
bud, and bring
forth
boughs like a
plant.
\VS{10}But
man
dieth, and wasteth
away: yea,
man giveth up the
ghost, and where
{\ADD{is}} he?\FTNT{§}{{\FR 14.10 }
wasteth…: Heb. is weakened, or, cut off}
\VS{11}{\ADD{As}} the
waters
fail from the
sea, and the
flood
decayeth and drieth
up:
\VS{12}So
man lieth
down, and
riseth not: till the
heavens
{\ADD{be}} no
more, they shall not
awake, nor be raised
out of their
sleep.
\VS{13}O
that thou wouldest
hide me in the
grave, that thou wouldest keep me
secret, until thy
wrath be
past, that thou wouldest
appoint me a set
time, and
remember me!
\VS{14}If a
man
die, shall he
live
{\ADD{again}}? all the
days of my appointed
time will I
wait, till my
change
come.
\VS{15}Thou shalt
call, and I will
answer thee: thou wilt have a
desire to the
work of thine
hands.
\par }{\PP \VS{16}For now thou
numberest my
steps: dost thou not
watch over my
sin?
\VS{17}My
transgression
{\ADD{is}} sealed
up in a
bag, and thou sewest
up mine
iniquity.
\VS{18}And
surely the
mountain
falling cometh to
nought, and the
rock is
removed out of his
place.\FTNT{**}{{\FR 14.18 }
cometh…: Heb. fadeth}
\VS{19}The
waters
wear the
stones: thou washest
away the things which
grow
{\ADD{out}} of the
dust of the
earth; and thou
destroyest the
hope of
man.\FTNT{††}{{\FR 14.19 }
washest…: Heb. overflowest}
\VS{20}Thou
prevailest for
ever against him, and he
passeth: thou
changest his
countenance, and sendest him
away.
\VS{21}His
sons come to
honour, and he
knoweth
{\ADD{it}} not; and they are brought
low, but he
perceiveth
{\ADD{it}} not of them.
\VS{22}But his
flesh upon him shall have
pain, and his
soul within him shall
mourn.

\par }\Chap{15}{\PP \VerseOne{1}Then
answered
Eliphaz the
Temanite, and
said,
\VS{2}Should a wise
man
utter
vain
knowledge, and
fill his
belly with the east
wind?\FTNT{*}{{\FR 15.2 }
vain…: Heb. knowledge of wind}
\VS{3}Should he
reason with
unprofitable
talk? or with
speeches wherewith he can do no
good?
\VS{4}Yea, thou castest
off
fear, and
restrainest
prayer
before
God.\FTNT{†}{{\FR 15.4 }
castest…: Heb. makest void}\FTNT{‡}{{\FR 15.4 }
prayer: or, speech}
\VS{5}For thy
mouth
uttereth thine
iniquity, and thou
choosest the
tongue of the
crafty.\FTNT{§}{{\FR 15.5 }
uttereth: Heb. teacheth}
\VS{6}Thine own
mouth
condemneth thee, and not I: yea, thine own
lips
testify against thee.
\VS{7}{\ADD{Art}} thou the
first
man
{\ADD{that}} was
born? or wast thou
made
before the
hills?
\VS{8}Hast thou
heard the
secret of
God? and dost thou
restrain
wisdom to thyself?
\VS{9}What
knowest thou, that we
know not?
{\ADD{what}}
understandest thou, which
{\ADD{is}} not in us?
\VS{10}With us
{\ADD{are}} both the
grayheaded and very
aged men,
much
elder than thy
father.
\VS{11}{\ADD{Are}} the
consolations of
God
small with thee? is there any
secret
thing with thee?
\VS{12}Why doth thine
heart carry thee
away? and what do thy
eyes wink
at,
\VS{13}That thou
turnest thy
spirit against
God, and lettest
{\ADD{such}}
words go
out of thy
mouth?
\VS{14}What
{\ADD{is}}
man, that he should be
clean? and
{\ADD{he which is}}
born of a
woman, that he should be
righteous?
\VS{15}Behold, he putteth no
trust in his
saints; yea, the
heavens are not
clean in his
sight.
\VS{16}How much more
abominable and
filthy
{\ADD{is}}
man, which
drinketh
iniquity like
water?
\par }{\PP \VS{17}I will
shew thee,
hear me; and that
{\ADD{which}} I have
seen I will
declare;
\VS{18}Which
wise men have
told from their
fathers, and have not
hid
{\ADD{it:}}
\VS{19}Unto whom alone the
earth was
given, and no
stranger
passed
among them.
\VS{20}The
wicked man
travaileth with pain all
{\ADD{his}}
days, and the
number of
years is
hidden to the
oppressor.
\VS{21}A
dreadful
sound
{\ADD{is}} in his
ears: in
prosperity the
destroyer shall
come upon him.\FTNT{**}{{\FR 15.21 }
A dreadful…: Heb. A sound of fears}
\VS{22}He
believeth not that he shall
return out of
darkness, and he is
waited for of the
sword.
\VS{23}He wandereth
abroad for
bread,
{\ADD{saying}}, Where
{\ADD{is it}}? he
knoweth that the
day of
darkness is
ready at his
hand.
\VS{24}Trouble and
anguish shall make him
afraid; they shall
prevail against him, as a
king
ready to the
battle.
\VS{25}For he stretcheth
out his
hand against
God, and
strengtheneth himself against the
Almighty.
\VS{26}He
runneth upon him,
{\ADD{even}} on
{\ADD{his}}
neck, upon the
thick
bosses of his
bucklers:
\VS{27}Because he
covereth his
face with his
fatness, and
maketh collops of
fat on
{\ADD{his}}
flanks.
\VS{28}And he
dwelleth in
desolate
cities,
{\ADD{and}} in
houses which no man
inhabiteth, which are
ready to become
heaps.
\VS{29}He shall not be
rich, neither shall his
substance
continue, neither shall he
prolong the
perfection thereof upon the
earth.
\VS{30}He shall not
depart out of
darkness; the
flame shall dry
up his
branches, and by the
breath of his
mouth shall he go
away.
\VS{31}Let not him that is
deceived
trust in
vanity: for
vanity shall be his
recompence.
\VS{32}It shall be
accomplished
before his
time, and his
branch shall not be
green.\FTNT{††}{{\FR 15.32 }
accomplished: or, cut off}
\VS{33}He shall shake
off his unripe
grape as the
vine, and shall cast
off his
flower as the
olive.
\VS{34}For the
congregation of
hypocrites
{\ADD{shall be}}
desolate, and
fire shall
consume the
tabernacles of
bribery.
\VS{35}They
conceive
mischief, and bring
forth
vanity, and their
belly
prepareth
deceit.\FTNT{‡‡}{{\FR 15.35 }
vanity: or, iniquity}

\par }\Chap{16}{\PP \VerseOne{1}Then
Job
answered and
said,
\VS{2}I have
heard many such
things:
miserable
comforters
{\ADD{are}} ye all.\FTNT{*}{{\FR 16.2 }
miserable: or, troublesome}
\VS{3}Shall
vain
words have an
end? or what
emboldeneth thee that thou
answerest?\FTNT{†}{{\FR 16.3 }
vain…: Heb. words of wind}
\VS{4}I also could
speak as ye
{\ADD{do:}}
if your
soul
were in my
soul’s stead, I could heap
up
words against you, and
shake mine
head
at you.
\VS{5}{\ADD{But}} I would
strengthen you
with my
mouth, and the
moving of my
lips should
asswage
{\ADD{your grief}}.
\par }{\PP \VS{6}Though I
speak, my
grief is not
asswaged: and
{\ADD{though}} I
forbear, what am I
eased?\FTNT{‡}{{\FR 16.6 }
what…: Heb. what goeth from me?}
\VS{7}But now he hath made me
weary: thou hast made
desolate all my
company.
\VS{8}And thou hast filled me with
wrinkles,
{\ADD{which}} is a
witness
{\ADD{against me:}} and my
leanness rising
up in me beareth
witness to my
face.
\VS{9}He
teareth
{\ADD{me}} in his
wrath, who
hateth me: he
gnasheth upon me with his
teeth; mine
enemy
sharpeneth his
eyes upon me.
\VS{10}They have
gaped upon me with their
mouth; they have
smitten me upon the
cheek
reproachfully; they have
gathered themselves
together against me.
\VS{11}God hath
delivered me to the
ungodly, and turned me
over into the
hands of the
wicked.\FTNT{§}{{\FR 16.11 }
hath…: Heb. hath shut me up}
\VS{12}I was at
ease, but he hath broken me
asunder: he hath also
taken
{\ADD{me}} by my
neck, and shaken me to
pieces, and set me
up for his
mark.
\VS{13}His
archers compass me round
about, he
cleaveth my
reins
asunder, and doth not
spare; he poureth
out my
gall upon the
ground.
\VS{14}He
breaketh me with
breach
upon
breach, he
runneth upon me like a
giant.
\VS{15}I have
sewed
sackcloth upon my
skin, and
defiled my
horn in the
dust.
\VS{16}My
face is
foul with
weeping, and on my
eyelids
{\ADD{is}} the shadow of
death;
\par }{\PP \VS{17}Not for
{\ADD{any}}
injustice in mine
hands: also my
prayer
{\ADD{is}}
pure.
\VS{18}O
earth,
cover not thou my
blood, and let my
cry have no
place.
\VS{19}Also now, behold, my
witness
{\ADD{is}} in
heaven, and my
record
{\ADD{is}} on
high.\FTNT{**}{{\FR 16.19 }
on high: Heb. in the high places}
\VS{20}My
friends
scorn me:
{\ADD{but}} mine
eye poureth
out
{\ADD{tears}} unto
God.\FTNT{††}{{\FR 16.20 }
scorn me: Heb. are my scorners}
\VS{21}O that one might
plead for a
man with
God, as a
man
{\ADD{pleadeth}} for his
neighbour!\FTNT{‡‡}{{\FR 16.21 }
neighbour: or, friend}
\VS{22}When a
few
years are
come, then I shall
go the
way
{\ADD{whence}} I shall not
return.\FTNT{§§}{{\FR 16.22 }
a few…: Heb. years of number}

\par }\Chap{17}{\PP \VerseOne{1}My
breath is
corrupt, my
days are
extinct, the
graves
{\ADD{are ready}} for me.\FTNT{*}{{\FR 17.1 }
breath…: or, spirit is spent}
\VS{2}{\ADD{Are there}} not
mockers with me? and doth not mine
eye
continue in their
provocation?\FTNT{†}{{\FR 17.2 }
continue: Heb. lodge}
\VS{3}Lay
down now, put me in a
surety with thee; who
{\ADD{is}} he
{\ADD{that}} will
strike
hands with me?
\VS{4}For thou hast
hid their
heart from
understanding: therefore shalt thou not
exalt
{\ADD{them}}.
\VS{5}He that
speaketh
flattery to
{\ADD{his}}
friends, even the
eyes of his
children shall
fail.
\VS{6}He hath
made me also a
byword of the
people; and
aforetime I was as a
tabret.\FTNT{‡}{{\FR 17.6 }
aforetime: or, before them}
\VS{7}Mine
eye also is
dim by reason of
sorrow, and all my
members
{\ADD{are}} as a
shadow.\FTNT{§}{{\FR 17.7 }
my members: or, my thoughts}
\VS{8}Upright
{\ADD{men}} shall be
astonied at this, and the
innocent shall stir
up himself against the
hypocrite.
\VS{9}The
righteous also shall
hold on his
way, and he that hath
clean
hands shall
be stronger and
stronger.\FTNT{**}{{\FR 17.9 }
be…: Heb. add strength}
\par }{\PP \VS{10}But as for you all, do ye
return, and come
now: for I cannot
find
{\ADD{one}}
wise
{\ADD{man}} among you.
\VS{11}My
days are
past, my
purposes are broken
off,
{\ADD{even}} the
thoughts of my
heart.\FTNT{††}{{\FR 17.11 }
the thoughts: Heb. the possessions}
\VS{12}They
change the
night into
day: the
light
{\ADD{is}}
short
because of
darkness.\FTNT{‡‡}{{\FR 17.12 }
short: Heb. near}
\VS{13}If I
wait, the
grave
{\ADD{is}} mine
house: I have
made my
bed in the
darkness.
\VS{14}I have
said to
corruption, Thou
{\ADD{art}} my
father: to the
worm,
{\ADD{Thou art}} my
mother, and my
sister.\FTNT{§§}{{\FR 17.14 }
said: Heb. cried, or, called}
\VS{15}And where
{\ADD{is}}
now my
hope? as for my
hope, who shall
see it?
\VS{16}They shall go
down to the
bars of the
pit, when
{\ADD{our}}
rest
together
{\ADD{is}} in the
dust.

\par }\Chap{18}{\PP \VerseOne{1}Then
answered
Bildad the
Shuhite, and
said,
\VS{2}How
long
{\ADD{will it be ere}} ye
make an
end of
words?
mark, and
afterwards we will
speak.
\VS{3}Wherefore are we
counted as
beasts,
{\ADD{and}} reputed
vile in your
sight?
\VS{4}He
teareth
himself in his
anger: shall the
earth be
forsaken for thee? and shall the
rock be
removed out of his
place?\FTNT{*}{{\FR 18.4 }
himself: Heb. his soul}
\par }{\PP \VS{5}Yea, the
light of the
wicked shall be put
out, and the
spark of his
fire shall not
shine.
\VS{6}The
light shall be
dark in his
tabernacle, and his
candle shall be put
out with him.\FTNT{†}{{\FR 18.6 }
candle: or, lamp}
\VS{7}The
steps of his
strength shall be
straitened, and his own
counsel shall cast him
down.
\VS{8}For he is
cast into a
net by his own
feet, and he
walketh upon a
snare.
\VS{9}The
gin shall
take
{\ADD{him}} by the
heel,
{\ADD{and}} the
robber shall
prevail against him.
\VS{10}The
snare
{\ADD{is}}
laid for him in the
ground, and a
trap for him in the
way.\FTNT{‡}{{\FR 18.10 }
laid: Heb. hidden}
\par }{\PP \VS{11}Terrors shall make him
afraid on every
side, and shall
drive him to his
feet.\FTNT{§}{{\FR 18.11 }
drive: Heb. scatter}
\VS{12}His
strength shall be
hungerbitten, and
destruction
{\ADD{shall be}}
ready at his
side.
\VS{13}It shall
devour the
strength of his
skin:
{\ADD{even}} the
firstborn of
death shall
devour his
strength.\FTNT{**}{{\FR 18.13 }
strength: Heb. bars}
\VS{14}His
confidence shall be rooted
out of his
tabernacle, and it shall
bring him to the
king of
terrors.
\VS{15}It shall
dwell in his
tabernacle, because
{\ADD{it is}}
none of his:
brimstone shall be
scattered upon his
habitation.
\VS{16}His
roots shall be dried
up beneath, and
above shall his
branch be cut
off.
\VS{17}His
remembrance shall
perish from the
earth, and he shall have no
name in the
street.
\VS{18}He shall be
driven from
light into
darkness, and
chased out of the
world.\FTNT{††}{{\FR 18.18 }
He…: Heb. They shall drive him}
\VS{19}He shall neither have
son nor
nephew among his
people, nor any
remaining in his
dwellings.
\VS{20}They that come
after
{\ADD{him}} shall be
astonied at his
day, as they that went
before were
affrighted.\FTNT{‡‡}{{\FR 18.20 }
went…: or, lived with him}\FTNT{§§}{{\FR 18.20 }
were…: Heb. laid hold on horror}
\VS{21}Surely such
{\ADD{are}} the
dwellings of the
wicked, and this
{\ADD{is}} the
place
{\ADD{of him that}}
knoweth not
God.

\par }\Chap{19}{\PP \VerseOne{1}Then
Job
answered and
said,
\VS{2}How long will ye
vex my
soul, and
break me in pieces with
words?
\VS{3}These
ten
times have ye
reproached me: ye are not
ashamed
{\ADD{that}} ye make yourselves
strange to me.\FTNT{*}{{\FR 19.3 }
make…: or, harden yourselves against me}
\VS{4}And be it
indeed
{\ADD{that}} I have
erred, mine
error
remaineth with myself.
\VS{5}If
indeed ye will
magnify
{\ADD{yourselves}} against me, and
plead against me my
reproach:
\VS{6}Know
now that
God hath
overthrown me, and hath
compassed me with his
net.
\VS{7}Behold, I cry
out of
wrong, but I am not
heard: I cry
aloud, but
{\ADD{there is}} no
judgment.\FTNT{†}{{\FR 19.7 }
wrong: or, violence}
\par }{\PP \VS{8}He hath fenced
up my
way that I cannot
pass, and he hath
set
darkness in my
paths.
\VS{9}He hath
stripped me of my
glory, and
taken the
crown
{\ADD{from}} my
head.
\VS{10}He hath
destroyed me on every
side, and I am
gone: and mine
hope hath he
removed like a
tree.
\VS{11}He hath also
kindled his
wrath against me, and he
counteth me unto him as
{\ADD{one of}} his
enemies.
\VS{12}His
troops
come
together, and raise
up their
way against me, and
encamp round
about my
tabernacle.
\VS{13}He hath
put my
brethren
far from me, and mine
acquaintance are verily
estranged from me.
\VS{14}My
kinsfolk have
failed, and my familiar
friends have
forgotten me.
\VS{15}They that
dwell in mine
house, and my
maids,
count me for a
stranger: I am an
alien in their
sight.
\VS{16}I
called my
servant, and he gave
{\ADD{me}} no
answer; I
intreated him
with my
mouth.
\VS{17}My
breath is
strange to my
wife, though I
intreated for the
children’s
{\ADD{sake}} of mine own
body.\FTNT{‡}{{\FR 19.17 }
mine…: Heb. my belly}
\VS{18}Yea, young
children
despised me; I
arose, and they
spake against me.\FTNT{§}{{\FR 19.18 }
young…: or, the wicked}
\VS{19}All my
inward
friends
abhorred me: and they whom I
loved are
turned against me.\FTNT{**}{{\FR 19.19 }
my…: Heb. the men of my secret}
\VS{20}My
bone
cleaveth to my
skin and to my
flesh, and I am
escaped with the
skin of my
teeth.\FTNT{††}{{\FR 19.20 }
and to: or, as to}
\VS{21}Have
pity upon me, have
pity upon me, O ye my
friends; for the
hand of
God hath
touched me.
\VS{22}Why do ye
persecute me as
God, and are not
satisfied with my
flesh?
\par }{\PP \VS{23}Oh
that my
words were
now
written! oh
that they were
printed in a
book!\FTNT{‡‡}{{\FR 19.23 }
Oh…: Heb. Who will give, etc}
\VS{24}That they were
graven with an
iron
pen and
lead in the
rock for
ever!
\VS{25}For I
know
{\ADD{that}} my
redeemer
liveth, and
{\ADD{that}} he shall
stand at the
latter
{\ADD{day}} upon the
earth:
\VS{26}And
{\ADD{though}}
after my
skin
{\ADD{worms}}
destroy this
{\ADD{body}}, yet in my
flesh shall I
see
God:\FTNT{§§}{{\FR 19.26 }
And…: or, After I shall awake, though this body be destroyed, yet out of my flesh}
\VS{27}Whom I shall
see for myself, and mine
eyes shall
behold, and not
another;
{\ADD{though}} my
reins be
consumed
within me.\FTNT{***}{{\FR 19.27 }
another: Heb. a stranger}\FTNT{†††}{{\FR 19.27 }
though…: or, my reins within me are consumed with earnest desire (for that day)}\FTNT{‡‡‡}{{\FR 19.27 }
within…: Heb. in my bosom}
\VS{28}But ye should
say, Why
persecute we him, seeing the
root of the
matter is
found in me?\FTNT{§§§}{{\FR 19.28 }
seeing…: or, and what root of matter is found in me?}
\VS{29}Be ye
afraid of the
sword: for
wrath
{\ADD{bringeth}} the
punishments
of the
sword, that ye may
know
{\ADD{there is}} a
judgment.

\par }\Chap{20}{\PP \VerseOne{1}Then
answered
Zophar the
Naamathite, and
said,
\VS{2}Therefore do my
thoughts cause me to
answer, and for
{\ADD{this}} I make
haste.\FTNT{*}{{\FR 20.2 }
I make…: Heb. my haste is in me}
\VS{3}I have
heard the
check of my
reproach, and the
spirit of my
understanding causeth me to
answer.
\VS{4}Knowest thou
{\ADD{not}} this of
old, since
man was
placed upon
earth,
\VS{5}That the
triumphing of the
wicked
{\ADD{is}}
short, and the
joy of the
hypocrite
{\ADD{but}} for a
moment?\FTNT{†}{{\FR 20.5 }
short: Heb. from near}
\VS{6}Though his
excellency mount
up to the
heavens, and his
head
reach unto the
clouds;\FTNT{‡}{{\FR 20.6 }
clouds: Heb. cloud}
\VS{7}{\ADD{Yet}} he shall
perish for
ever like his own
dung: they which have
seen him shall
say,
Where
{\ADD{is}} he?
\VS{8}He shall fly
away as a
dream, and shall not be
found: yea, he shall be chased
away as a
vision of the
night.
\VS{9}The
eye also
{\ADD{which}}
saw him shall
{\ADD{see him}} no
more; neither shall his
place any more
behold him.
\par }{\PP \VS{10}His
children shall seek to
please the
poor, and his
hands shall
restore their
goods.\FTNT{§}{{\FR 20.10 }
His children…: or, The poor shall oppress his children}
\VS{11}His
bones are
full
{\ADD{of the sin}} of his
youth, which shall lie
down with him in the
dust.
\VS{12}Though
wickedness be
sweet in his
mouth,
{\ADD{though}} he
hide it under his
tongue;
\VS{13}{\ADD{Though}} he
spare it, and
forsake it not; but keep it
still
within his
mouth:\FTNT{**}{{\FR 20.13 }
within…: Heb. in the midst of his palate}
\VS{14}{\ADD{Yet}} his
meat in his
bowels is
turned,
{\ADD{it is}} the
gall of
asps
within him.
\VS{15}He hath swallowed
down
riches, and he shall vomit them up
again:
God shall cast them
out of his
belly.
\VS{16}He shall
suck the
poison of
asps: the
viper’s
tongue shall
slay him.
\VS{17}He shall not
see the
rivers, the
floods, the
brooks of
honey and
butter.\FTNT{††}{{\FR 20.17 }
the floods…: or, streaming brooks}
\VS{18}That which he
laboured for shall he
restore, and shall not swallow
{\ADD{it}}
down: according to
{\ADD{his}}
substance
{\ADD{shall}} the
restitution
{\ADD{be}}, and he shall not
rejoice
{\ADD{therein}}.\FTNT{‡‡}{{\FR 20.18 }
his…: Heb. the substance of his exchange}
\VS{19}Because he hath
oppressed
{\ADD{and}} hath
forsaken the
poor;
{\ADD{because}} he hath violently taken
away an
house which he
builded not;\FTNT{§§}{{\FR 20.19 }
oppressed: Heb. crushed}
\VS{20}Surely he shall not
feel
quietness in his
belly, he shall not
save of that which he
desired.\FTNT{***}{{\FR 20.20 }
feel: Heb. know}
\VS{21}There shall none of his
meat be
left; therefore shall no man
look for his
goods.\FTNT{†††}{{\FR 20.21 }
none…: or, be none left for his meat}
\VS{22}In the
fulness of his
sufficiency he shall be in
straits: every
hand of the
wicked shall
come upon him.\FTNT{‡‡‡}{{\FR 20.22 }
wicked: or, troublesome}
\par }{\PP \VS{23}{\ADD{When}} he is about to
fill his
belly,
{\ADD{God}} shall
cast the
fury of his
wrath upon him, and shall
rain
{\ADD{it}} upon him while he is
eating.
\VS{24}He shall
flee from the
iron
weapon,
{\ADD{and}} the
bow of
steel shall strike him
through.
\VS{25}It is
drawn, and cometh
out of the
body; yea, the glittering
sword cometh
out of his
gall:
terrors
{\ADD{are}} upon him.
\VS{26}All
darkness
{\ADD{shall be}}
hid in his secret
places: a
fire not
blown shall
consume him; it shall go
ill with him that is
left in his
tabernacle.
\VS{27}The
heaven shall
reveal his
iniquity; and the
earth shall rise
up against him.
\VS{28}The
increase of his
house shall
depart,
{\ADD{and his goods}} shall flow
away in the
day of his
wrath.
\VS{29}This
{\ADD{is}} the
portion of a
wicked
man from
God, and the
heritage
appointed unto him by
God.\FTNT{§§§}{{\FR 20.29 }
appointed…: Heb. of his decree from God}

\par }\Chap{21}{\PP \VerseOne{1}But
Job
answered and
said,
\VS{2}Hear
diligently my
speech, and let this be your
consolations.
\VS{3}Suffer me that I may
speak; and
after that I have
spoken, mock
on.
\VS{4}As for me,
{\ADD{is}} my
complaint to
man? and if
{\ADD{it were so}}, why should not my
spirit be
troubled?\FTNT{*}{{\FR 21.4 }
troubled: Heb. shortened?}
\VS{5}Mark me, and be
astonished, and
lay
{\ADD{your}}
hand upon
{\ADD{your}}
mouth.\FTNT{†}{{\FR 21.5 }
Mark…: Heb. Look unto me}
\VS{6}Even when I
remember I am
afraid, and
trembling taketh hold
on my
flesh.
\par }{\PP \VS{7}Wherefore do the
wicked
live, become
old, yea, are
mighty in
power?
\VS{8}Their
seed is
established in their
sight with them, and their
offspring before their
eyes.
\VS{9}Their
houses
{\ADD{are}}
safe from
fear, neither
{\ADD{is}} the
rod of
God upon them.\FTNT{‡}{{\FR 21.9 }
safe…: Heb. peace from}
\VS{10}Their
bull
gendereth, and
faileth not; their
cow
calveth, and casteth not her
calf.
\VS{11}They send
forth their little
ones like a
flock, and their
children
dance.
\VS{12}They
take the
timbrel and
harp, and
rejoice at the
sound of the
organ.
\VS{13}They
spend their
days in
wealth, and in a
moment go
down to the
grave.\FTNT{§}{{\FR 21.13 }
in wealth: or, in mirth}
\VS{14}Therefore they
say unto
God,
Depart from us; for we
desire not the
knowledge of thy
ways.
\VS{15}What
{\ADD{is}} the
Almighty, that we should
serve him? and what
profit should we have, if we
pray unto him?
\VS{16}Lo, their
good
{\ADD{is}} not in their
hand: the
counsel of the
wicked is
far from me.
\par }{\PP \VS{17}How oft is the
candle of the
wicked put
out! and
{\ADD{how oft}}
cometh their
destruction upon them!
{\ADD{God}}
distributeth
sorrows in his
anger.\FTNT{**}{{\FR 21.17 }
candle: or, lamp}
\VS{18}They are as
stubble
before the
wind, and as
chaff that the
storm carrieth
away.\FTNT{††}{{\FR 21.18 }
carrieth: Heb. stealeth}
\VS{19}God layeth
up his
iniquity for his
children: he
rewardeth him, and he shall
know
{\ADD{it}}.\FTNT{‡‡}{{\FR 21.19 }
his iniquity: that is, the punishment of his iniquity}
\VS{20}His
eyes shall
see his
destruction, and he shall
drink of the
wrath of the
Almighty.
\VS{21}For what
pleasure
{\ADD{hath}} he in his
house
after him, when the
number of his
months is cut off in the
midst?
\VS{22}Shall
{\ADD{any}}
teach
God
knowledge? seeing he
judgeth those that are
high.
\VS{23}One
dieth in his
full
strength, being wholly at
ease and
quiet.\FTNT{§§}{{\FR 21.23 }
his…: Heb. his very, or, the strength of, his perfection}
\VS{24}His
breasts are
full of
milk, and his
bones are
moistened with
marrow.\FTNT{***}{{\FR 21.24 }
breasts: or, milk pails}
\VS{25}And another
dieth in the
bitterness of his
soul, and never
eateth with
pleasure.
\VS{26}They shall lie
down
alike in the
dust, and the
worms shall
cover them.
\par }{\PP \VS{27}Behold, I
know your
thoughts, and the
devices
{\ADD{which}} ye wrongfully
imagine against me.
\VS{28}For ye
say, Where
{\ADD{is}} the
house of the
prince? and where
{\ADD{are}} the
dwelling
places of the
wicked?\FTNT{†††}{{\FR 21.28 }
the dwelling…: Heb. the tent of the tabernacles}
\VS{29}Have ye not
asked them that
go by the
way? and do ye not
know their
tokens,
\VS{30}That the
wicked is
reserved to the
day of
destruction? they shall be brought
forth to the
day of
wrath.\FTNT{‡‡‡}{{\FR 21.30 }
wrath: Heb. wraths}
\VS{31}Who shall
declare his
way to his
face? and who shall
repay him
{\ADD{what}} he hath
done?
\VS{32}Yet shall he be
brought to the
grave, and shall
remain in the
tomb.\FTNT{§§§}{{\FR 21.32 }
grave: Heb. graves}\FTNT{*}{{\FR 21.32 }
remain…: Heb. watch in the heap}
\VS{33}The
clods of the
valley shall be
sweet unto him, and every
man shall
draw
after him, as
{\ADD{there are}}
innumerable
before him.
\VS{34}How then
comfort ye me in
vain, seeing in your
answers there
remaineth
falsehood?\FTNT{†}{{\FR 21.34 }
falsehood: Heb. transgression?}

\par }\Chap{22}{\PP \VerseOne{1}Then
Eliphaz the
Temanite
answered and
said,
\VS{2}Can a
man be
profitable unto
God,
as he that is
wise may be
profitable unto himself?\FTNT{*}{{\FR 22.2 }
as he…: or, if he may be profitable, doth his good success depend thereon?}
\VS{3}{\ADD{Is it}} any
pleasure to the
Almighty, that thou art
righteous? or
{\ADD{is it}}
gain
{\ADD{to him}}, that thou makest thy
ways
perfect?
\VS{4}Will he
reprove thee for
fear of thee? will he
enter with thee into
judgment?
\par }{\PP \VS{5}{\ADD{Is}} not thy
wickedness
great? and thine
iniquities
infinite?
\VS{6}For thou hast taken a
pledge from thy
brother for
nought, and
stripped the
naked of their
clothing.\FTNT{†}{{\FR 22.6 }
the naked…: Heb. the clothes of the naked}
\VS{7}Thou hast not given
water to the
weary to
drink, and thou hast
withholden
bread from the
hungry.
\VS{8}But
{\ADD{as for}} the
mighty
man, he had the
earth; and the honourable
man
dwelt in it.\FTNT{‡}{{\FR 22.8 }
mighty…: Heb. man of arm}\FTNT{§}{{\FR 22.8 }
honourable…: Heb. eminent, or, accepted for countenance}
\VS{9}Thou hast
sent
widows away
empty, and the
arms of the
fatherless have been
broken.
\VS{10}Therefore
snares
{\ADD{are}} round
about thee, and
sudden
fear
troubleth thee;
\VS{11}Or
darkness,
{\ADD{that}} thou canst not
see; and
abundance of
waters
cover thee.
\VS{12}{\ADD{Is}} not
God in the
height of
heaven? and
behold the
height of the
stars, how
high they are!\FTNT{**}{{\FR 22.12 }
height of the stars: Heb. head of the stars}
\VS{13}And thou
sayest, How doth
God
know? can he
judge through the dark
cloud?\FTNT{††}{{\FR 22.13 }
How: or, What}
\VS{14}Thick
clouds
{\ADD{are}} a
covering to him, that he
seeth not; and he
walketh in the
circuit of
heaven.
\par }{\PP \VS{15}Hast thou
marked the
old
way which
wicked
men have
trodden?
\VS{16}Which were cut
down out of
time, whose
foundation was
overflown with a
flood:\FTNT{‡‡}{{\FR 22.16 }
whose…: Heb. a flood was poured upon their foundation}
\VS{17}Which
said unto
God,
Depart from us: and what can the
Almighty
do for them?\FTNT{§§}{{\FR 22.17 }
for: or, to}
\VS{18}Yet he
filled their
houses with
good
{\ADD{things:}} but the
counsel of the
wicked is
far from me.
\VS{19}The
righteous
see
{\ADD{it}}, and are
glad: and the
innocent laugh them to
scorn.
\VS{20}Whereas our
substance is not cut
down, but the
remnant of them the
fire
consumeth.\FTNT{***}{{\FR 22.20 }
substance: or, estate}\FTNT{†††}{{\FR 22.20 }
the remnant…: or, their excellency}
\par }{\PP \VS{21}Acquaint now thyself with him, and be at
peace: thereby
good shall
come unto thee.\FTNT{‡‡‡}{{\FR 22.21 }
him: that is, God}
\VS{22}Receive, I pray thee, the
law from his
mouth, and lay
up his
words in thine
heart.
\VS{23}If thou
return to the
Almighty, thou shalt be built
up, thou shalt put
away
iniquity
far from thy
tabernacles.
\VS{24}Then shalt thou lay
up
gold
as
dust, and the
{\ADD{gold}} of
Ophir as the
stones of the
brooks.\FTNT{§§§}{{\FR 22.24 }
as dust: or, on the dust}
\VS{25}Yea, the
Almighty shall be thy
defence, and thou shalt have
plenty of
silver.\FTNT{*}{{\FR 22.25 }
defence: or, gold}\FTNT{†}{{\FR 22.25 }
plenty…: Heb. silver of strength}
\VS{26}For then shalt thou have thy
delight in the
Almighty, and shalt lift
up thy
face unto
God.
\VS{27}Thou shalt make thy
prayer unto him, and he shall
hear thee, and thou shalt
pay thy
vows.
\VS{28}Thou shalt also
decree a
thing, and it shall be
established unto thee: and the
light shall
shine upon thy
ways.
\VS{29}When
{\ADD{men}} are cast
down, then thou shalt
say,
{\ADD{There is}} lifting
up; and he shall
save the
humble
person.\FTNT{‡}{{\FR 22.29 }
the humble…: Heb. him that hath low eyes}
\VS{30}He shall
deliver the
island of the
innocent: and it is
delivered by the
pureness of thine
hands.\FTNT{§}{{\FR 22.30 }
He shall…: or, The innocent shall deliver the island}

\par }\Chap{23}{\PP \VerseOne{1}Then
Job
answered and
said,
\VS{2}Even to
day
{\ADD{is}} my
complaint
bitter: my
stroke is
heavier than my
groaning.\FTNT{*}{{\FR 23.2 }
stroke: Heb. hand}
\VS{3}Oh
that I
knew where I might
find him!
{\ADD{that}} I might
come
{\ADD{even}} to his
seat!
\VS{4}I would
order
{\ADD{my}}
cause
before him, and
fill my
mouth with
arguments.
\VS{5}I would
know the
words
{\ADD{which}} he would
answer me, and
understand what he would
say unto me.
\VS{6}Will he
plead
against me with
{\ADD{his}}
great
power? No; but he would
put
{\ADD{strength}} in me.
\VS{7}There the
righteous might
dispute with him; so should I be
delivered for
ever from my
judge.
\par }{\PP \VS{8}Behold, I
go
forward, but he
{\ADD{is}} not
{\ADD{there;}} and
backward, but I cannot
perceive him:
\VS{9}On the left
hand, where he doth
work, but I cannot
behold
{\ADD{him:}} he
hideth himself on the right
hand, that I cannot
see
{\ADD{him:}}
\VS{10}But he
knoweth the
way that I
take:
{\ADD{when}} he hath
tried me, I shall come
forth as
gold.\FTNT{†}{{\FR 23.10 }
that…: Heb. that is with me}
\VS{11}My
foot hath
held his
steps, his
way have I
kept, and not
declined.
\VS{12}Neither have I gone
back from the
commandment of his
lips; I have
esteemed the
words of his
mouth more than my
necessary
{\ADD{food}}.\FTNT{‡}{{\FR 23.12 }
esteemed: Heb. hid, or, laid up}\FTNT{§}{{\FR 23.12 }
my…: or, my appointed portion}
\par }{\PP \VS{13}But he
{\ADD{is}} in
one
{\ADD{mind}}, and who can
turn him? and
{\ADD{what}} his
soul
desireth, even
{\ADD{that}} he
doeth.
\VS{14}For he
performeth
{\ADD{the thing that is}}
appointed for me: and
many
such
{\ADD{things are}} with him.
\VS{15}Therefore am I
troubled at his
presence: when I
consider, I am
afraid of him.
\VS{16}For
God maketh my
heart
soft, and the
Almighty
troubleth me:
\VS{17}Because I was not cut
off
before the
darkness,
{\ADD{neither}} hath he
covered the
darkness from my
face.

\par }\Chap{24}{\PP \VerseOne{1}Why, seeing
times are not
hidden from the
Almighty, do they that
know him not
see his
days?
\VS{2}{\ADD{Some}}
remove the
landmarks; they violently take
away
flocks, and
feed
{\ADD{thereof}}.\FTNT{*}{{\FR 24.2 }
feed…: or, feed them}
\VS{3}They drive
away the
ass of the
fatherless, they
take the
widow’s
ox for a
pledge.
\VS{4}They
turn the
needy out of the
way: the
poor of the
earth
hide themselves
together.
\VS{5}Behold,
{\ADD{as}} wild
asses in the
desert, go they
forth to their
work; rising
betimes for a
prey: the
wilderness
{\ADD{yieldeth}}
food for them
{\ADD{and}} for
{\ADD{their}}
children.
\VS{6}They
reap
{\ADD{every one}} his
corn in the
field: and they
gather the
vintage of the
wicked.\FTNT{†}{{\FR 24.6 }
corn: Heb. mingled corn, or, dredge}\FTNT{‡}{{\FR 24.6 }
they gather…: Heb. the wicked gather the vintage}
\VS{7}They cause the
naked to
lodge without
clothing, that
{\ADD{they have}} no
covering in the
cold.
\VS{8}They are
wet with the
showers of the
mountains, and
embrace the
rock for want of a
shelter.
\VS{9}They
pluck the
fatherless from the
breast, and take a
pledge of the
poor.
\VS{10}They cause
{\ADD{him}} to
go
naked without
clothing, and they take
away the
sheaf
{\ADD{from}} the
hungry;
\VS{11}{\ADD{Which}} make
oil
within their
walls,
{\ADD{and}}
tread
{\ADD{their}}
winepresses, and suffer
thirst.
\VS{12}Men
groan from out of the
city, and the
soul of the
wounded crieth
out: yet
God
layeth not
folly
{\ADD{to them}}.
\par }{\PP \VS{13}They are of those that
rebel against the
light; they
know not the
ways thereof, nor
abide in the
paths thereof.
\VS{14}The
murderer
rising with the
light
killeth the
poor and
needy, and in the
night is as a
thief.
\VS{15}The
eye also of the
adulterer
waiteth for the
twilight,
saying, No
eye shall
see me: and
disguiseth
{\ADD{his}}
face.\FTNT{§}{{\FR 24.15 }
disguiseth…: Heb. setteth his face in secret}
\VS{16}In the
dark they dig
through
houses,
{\ADD{which}} they had
marked for themselves in the
daytime: they
know not the
light.
\VS{17}For the
morning
{\ADD{is}} to them even
as the shadow of
death: if
{\ADD{one}}
know
{\ADD{them, they are in}} the
terrors of the shadow of
death.
\par }{\PP \VS{18}He
{\ADD{is}}
swift as the
waters; their
portion is
cursed in the
earth: he
beholdeth not the
way of the
vineyards.
\VS{19}Drought and
heat
consume the
snow
waters:
{\ADD{so doth}} the
grave
{\ADD{those which}} have
sinned.\FTNT{**}{{\FR 24.19 }
consume: Heb. violently take}
\VS{20}The
womb shall
forget him; the
worm shall feed
sweetly on him; he shall be no more
remembered; and
wickedness shall be
broken as a
tree.
\VS{21}He evil
entreateth the
barren
{\ADD{that}}
beareth not: and doeth not
good to the
widow.
\VS{22}He
draweth also the
mighty with his
power: he riseth
up, and no
{\ADD{man}} is
sure of
life.\FTNT{††}{{\FR 24.22 }
no…: or, he trusteth not his own life}
\VS{23}{\ADD{Though}} it be
given him
{\ADD{to be}} in
safety, whereon he
resteth; yet his
eyes
{\ADD{are}} upon their
ways.
\VS{24}They are
exalted for a little
while, but are gone and brought
low; they are taken
out of the
way as all
{\ADD{other}}, and cut
off as the
tops of the ears of
corn.\FTNT{‡‡}{{\FR 24.24 }
are gone: Heb. are not}\FTNT{§§}{{\FR 24.24 }
taken…: Heb. closed up}
\VS{25}And if
{\ADD{it be}} not
{\ADD{so}}
now, who will
make me a
liar, and
make my
speech nothing
worth?

\par }\Chap{25}{\PP \VerseOne{1}Then
answered
Bildad the
Shuhite, and
said,
\VS{2}Dominion and
fear
{\ADD{are}} with him, he
maketh
peace in his high
places.
\VS{3}Is
there any
number of his
armies? and upon whom doth not his
light
arise?
\VS{4}How then can
man be
justified with
God? or how can he be
clean
{\ADD{that is}}
born of a
woman?
\VS{5}Behold even to the
moon, and it
shineth not; yea, the
stars are not
pure in his
sight.
\VS{6}How much less
man,
{\ADD{that is}} a
worm? and the
son of
man,
{\ADD{which is}} a
worm?

\par }\Chap{26}{\PP \VerseOne{1}But
Job
answered and
said,
\VS{2}How hast thou
helped
{\ADD{him that is}} without
power?
{\ADD{how}}
savest thou the
arm
{\ADD{that hath}}
no
strength?
\VS{3}How hast thou
counselled
{\ADD{him that hath}}
no
wisdom? and
{\ADD{how}} hast thou
plentifully
declared the thing as it
is?
\VS{4}To whom hast thou
uttered
words? and whose
spirit
came from thee?
\par }{\PP \VS{5}Dead
{\ADD{things}} are
formed from under the
waters, and the
inhabitants thereof.\FTNT{*}{{\FR 26.5 }
and the…: or, with the inhabitants}
\VS{6}Hell
{\ADD{is}}
naked before him, and
destruction hath no
covering.
\VS{7}He stretcheth
out the
north over the empty
place,
{\ADD{and}}
hangeth the
earth upon
nothing.
\VS{8}He bindeth
up the
waters in his thick
clouds; and the
cloud is not
rent under them.
\VS{9}He holdeth
back the
face of his
throne,
{\ADD{and}}
spreadeth his
cloud upon it.
\VS{10}He hath
compassed the
waters with
bounds, until the
day and
night come to an
end.\FTNT{†}{{\FR 26.10 }
until…: Heb. until the end of light with darkness}
\VS{11}The
pillars of
heaven
tremble and are
astonished at his
reproof.
\VS{12}He
divideth the
sea with his
power, and by his
understanding he
smiteth through the
proud.\FTNT{‡}{{\FR 26.12 }
the proud: Heb. pride}
\VS{13}By his
spirit he hath
garnished the
heavens; his
hand hath
formed the
crooked
serpent.
\VS{14}Lo, these
{\ADD{are}}
parts of his
ways: but how
little a
portion is
heard of him? but the
thunder of his
power who can
understand?

\par }\Chap{27}{\PP \VerseOne{1}Moreover
Job
continued his
parable, and
said,\FTNT{*}{{\FR 27.1 }
continued: Heb. added to take up}
\VS{2}{\ADD{As}}
God
liveth,
{\ADD{who}} hath taken
away my
judgment; and the
Almighty,
{\ADD{who}} hath
vexed my
soul;\FTNT{†}{{\FR 27.2 }
vexed…: Heb. made my soul bitter}
\VS{3}All the
while my
breath
{\ADD{is}} in me, and the
spirit of
God
{\ADD{is}} in my
nostrils;\FTNT{‡}{{\FR 27.3 }
the spirit…: that is, the breath which God gave him}
\VS{4}My
lips shall not
speak
wickedness, nor my
tongue
utter
deceit.
\VS{5}God
forbid that I should
justify you: till I
die I will not
remove mine
integrity from me.
\VS{6}My
righteousness I
hold fast, and will not let it
go: my
heart shall not
reproach
{\ADD{me}} so long as I
live.\FTNT{§}{{\FR 27.6 }
so long…: Heb. from my days}
\par }{\PP \VS{7}Let mine
enemy be as the
wicked, and he that riseth
up against me as the
unrighteous.
\VS{8}For what
{\ADD{is}} the
hope of the
hypocrite, though he hath
gained, when
God taketh
away his
soul?
\VS{9}Will
God
hear his
cry when
trouble
cometh upon him?
\VS{10}Will he
delight himself in the
Almighty? will he
always
call upon
God?
\par }{\PP \VS{11}I will
teach you by the
hand of
God:
{\ADD{that}} which
{\ADD{is}} with the
Almighty will I not
conceal.\FTNT{**}{{\FR 27.11 }
by…: or, being in the hand, etc}
\VS{12}Behold, all ye yourselves have
seen
{\ADD{it;}} why then are ye thus
altogether
vain?
\VS{13}This
{\ADD{is}} the
portion of a
wicked
man with
God, and the
heritage of
oppressors,
{\ADD{which}} they shall
receive of the
Almighty.
\VS{14}If his
children be
multiplied,
{\ADD{it is}}
for the
sword: and his
offspring shall not be
satisfied with
bread.
\VS{15}Those that
remain of him shall be
buried in
death: and his
widows shall not
weep.
\VS{16}Though he heap
up
silver as the
dust, and
prepare
raiment as the
clay;
\VS{17}He may
prepare
{\ADD{it}}, but the
just shall put
{\ADD{it}}
on, and the
innocent shall
divide the
silver.
\VS{18}He
buildeth his
house as a
moth, and as a
booth
{\ADD{that}} the
keeper
maketh.
\VS{19}The
rich man shall lie
down, but he shall not be
gathered: he
openeth his
eyes, and he
{\ADD{is}} not.
\VS{20}Terrors take
hold on him as
waters, a
tempest stealeth him
away in the
night.
\VS{21}The
east wind carrieth him
away, and he
departeth: and as a storm
hurleth him out of his
place.
\VS{22}For
{\ADD{God}} shall
cast upon him, and not
spare: he would
fain
flee out of his
hand.\FTNT{††}{{\FR 27.22 }
he…: Heb. in fleeing he would flee}
\VS{23}{\ADD{Men}} shall
clap their
hands at him, and shall
hiss him out of his
place.

\par }\Chap{28}{\PP \VerseOne{1}Surely there is a
vein for the
silver, and a
place for
gold
{\ADD{where}} they
fine
{\ADD{it}}.\FTNT{*}{{\FR 28.1 }
vein: or, mine}
\VS{2}Iron is
taken out of the
earth, and
brass
{\ADD{is}}
molten
{\ADD{out of}} the
stone.\FTNT{†}{{\FR 28.2 }
earth: or, dust}
\VS{3}He
setteth an
end to
darkness, and searcheth
out all
perfection: the
stones of
darkness, and the shadow of
death.
\VS{4}The
flood breaketh
out from the
inhabitant;
{\ADD{even the waters}}
forgotten of the
foot: they are dried
up, they are gone
away from
men.
\VS{5}{\ADD{As for}} the
earth, out of it
cometh
bread: and under it is turned
up as it were
fire.
\VS{6}The
stones of it
{\ADD{are}} the
place of
sapphires: and it hath
dust of
gold.\FTNT{‡}{{\FR 28.6 }
dust…: or, gold ore}
\VS{7}{\ADD{There is}} a
path which no
fowl
knoweth, and which the
vulture’s
eye hath not
seen:
\VS{8}The
lion’s
whelps have not
trodden it, nor the fierce
lion
passed by it.
\VS{9}He putteth
forth his
hand upon the
rock; he
overturneth the
mountains by the
roots.\FTNT{§}{{\FR 28.9 }
rock: or, flint}
\VS{10}He cutteth
out
rivers among the
rocks; and his
eye
seeth every precious
thing.
\VS{11}He
bindeth the
floods from
overflowing; and
{\ADD{the thing that is}}
hid bringeth he
forth to
light.\FTNT{**}{{\FR 28.11 }
overflowing: Heb. weeping}
\VS{12}But
where shall
wisdom be
found? and where
{\ADD{is}} the
place of
understanding?
\VS{13}Man
knoweth not the
price thereof; neither is it
found in the
land of the
living.
\par }{\PP \VS{14}The
depth
saith, It
{\ADD{is}} not in me: and the
sea
saith,
{\ADD{It is}} not with me.
\VS{15}It cannot be
gotten for
gold, neither shall
silver be
weighed
{\ADD{for}} the
price thereof.\FTNT{††}{{\FR 28.15 }
It…: Heb. Fine gold shall not be given for it}
\VS{16}It cannot be
valued with the
gold of
Ophir, with the
precious
onyx, or the
sapphire.
\VS{17}The
gold and the
crystal cannot
equal it: and the
exchange of it
{\ADD{shall not be for}}
jewels of fine
gold.\FTNT{‡‡}{{\FR 28.17 }
jewels…: or, vessels of}
\VS{18}No
mention shall be made of
coral, or of
pearls: for the
price of
wisdom
{\ADD{is}} above
rubies.\FTNT{§§}{{\FR 28.18 }
coral: or, Ramoth}
\VS{19}The
topaz of
Ethiopia shall not
equal it, neither shall it be
valued with
pure
gold.
\par }{\PP \VS{20}Whence then
cometh
wisdom? and where
{\ADD{is}} the
place of
understanding?
\VS{21}Seeing it is
hid from the
eyes of all
living, and kept
close from the
fowls of the
air.\FTNT{***}{{\FR 28.21 }
air: or, heaven}
\VS{22}Destruction and
death
say, We have
heard the
fame thereof with our
ears.
\VS{23}God
understandeth the
way thereof, and he
knoweth the
place thereof.
\VS{24}For he
looketh to the
ends of the
earth,
{\ADD{and}}
seeth under the whole
heaven;
\VS{25}To
make the
weight for the
winds; and he
weigheth the
waters by
measure.
\VS{26}When he
made a
decree for the
rain, and a
way for the
lightning of the
thunder:
\VS{27}Then did he
see it, and
declare it; he
prepared it, yea, and searched it
out.\FTNT{†††}{{\FR 28.27 }
declare…: or, number it}
\VS{28}And unto
man he
said, Behold, the
fear of the
Lord, that
{\ADD{is}}
wisdom; and to
depart from
evil
{\ADD{is}}
understanding.

\par }\Chap{29}{\PP \VerseOne{1}Moreover
Job
continued his
parable, and
said,\FTNT{*}{{\FR 29.1 }
continued: Heb. added to take up}
\VS{2}Oh
that I were as
{\ADD{in}}
months
past, as
{\ADD{in}} the
days
{\ADD{when}}
God
preserved me;
\VS{3}When his
candle
shined upon my
head,
{\ADD{and when}} by his
light I
walked
{\ADD{through}}
darkness;\FTNT{†}{{\FR 29.3 }
candle: or, lamp}
\VS{4}As I was in the
days of my
youth, when the
secret of
God
{\ADD{was}} upon my
tabernacle;
\VS{5}When the
Almighty
{\ADD{was}} yet with me,
{\ADD{when}} my
children
{\ADD{were}}
about me;
\VS{6}When I
washed my
steps with
butter, and the
rock poured me
out
rivers of
oil;\FTNT{‡}{{\FR 29.6 }
me: Heb. with me}
\par }{\PP \VS{7}When I went
out to the
gate through the
city,
{\ADD{when}} I
prepared my
seat in the
street!
\VS{8}The young
men
saw me, and
hid themselves: and the
aged
arose,
{\ADD{and}} stood
up.
\VS{9}The
princes
refrained
talking, and
laid
{\ADD{their}}
hand on their
mouth.
\VS{10}The
nobles
held their
peace, and their
tongue
cleaved to the roof of their
mouth.\FTNT{§}{{\FR 29.10 }
The nobles…: Heb. The voice of the nobles was hid}
\VS{11}When the
ear
heard
{\ADD{me}}, then it
blessed me; and when the
eye
saw
{\ADD{me}}, it gave
witness to me:
\VS{12}Because I
delivered the
poor that
cried, and the
fatherless, and
{\ADD{him that had}} none to
help him.
\VS{13}The
blessing of him that was ready to
perish
came upon me: and I caused the
widow’s
heart to sing for
joy.
\VS{14}I
put on
righteousness, and it
clothed me: my
judgment
{\ADD{was}} as a
robe and a
diadem.
\VS{15}I was
eyes to the
blind, and
feet
{\ADD{was}} I to the
lame.
\VS{16}I
{\ADD{was}} a
father to the
poor: and the
cause
{\ADD{which}} I
knew not I searched
out.
\VS{17}And I
brake the
jaws of the
wicked, and
plucked the
spoil out of his
teeth.\FTNT{**}{{\FR 29.17 }
the jaws: Heb. the jawteeth, or, the grinders}\FTNT{††}{{\FR 29.17 }
plucked: Heb. cast}
\par }{\PP \VS{18}Then I
said, I shall
die in my
nest, and I shall
multiply
{\ADD{my}}
days as the
sand.
\VS{19}My
root
{\ADD{was}} spread
out by the
waters, and the
dew lay all
night upon my
branch.\FTNT{‡‡}{{\FR 29.19 }
spread…: Heb. opened}
\VS{20}My
glory
{\ADD{was}}
fresh in
me, and my
bow was
renewed in my
hand.\FTNT{§§}{{\FR 29.20 }
fresh: Heb. new}\FTNT{***}{{\FR 29.20 }
renewed: Heb. changed}
\VS{21}Unto me
{\ADD{men}} gave
ear, and
waited, and kept
silence
at my
counsel.
\VS{22}After my
words they spake not
again; and my
speech
dropped upon them.
\VS{23}And they
waited for me as for the
rain; and they
opened their
mouth wide
{\ADD{as}} for the latter
rain.
\VS{24}{\ADD{If}} I
laughed on them, they
believed
{\ADD{it}} not; and the
light of my
countenance they cast not
down.
\VS{25}I
chose out their
way, and
sat
chief, and
dwelt as a
king in the
army, as one
{\ADD{that}}
comforteth the
mourners.

\par }\Chap{30}{\PP \VerseOne{1}But now
{\ADD{they that are}}
younger than I have me in
derision, whose
fathers I would have
disdained to have
set with the
dogs of my
flock.\FTNT{*}{{\FR 30.1 }
younger…: Heb. of fewer days than I}
\VS{2}Yea,
whereto
{\ADD{might}} the
strength of their
hands
{\ADD{profit}} me, in whom old
age was
perished?
\VS{3}For
want and
famine
{\ADD{they were}}
solitary;
fleeing into the
wilderness in former
time
desolate and
waste.\FTNT{†}{{\FR 30.3 }
solitary: or, dark as the night}\FTNT{‡}{{\FR 30.3 }
in…: Heb. yesternight}
\VS{4}Who cut
up
mallows by the
bushes, and
juniper
roots
{\ADD{for}} their
meat.
\VS{5}They were driven
forth from
among
{\ADD{men}}, (they
cried after them as
{\ADD{after}} a
thief;)
\VS{6}To
dwell in the
clifts of the
valleys,
{\ADD{in}}
caves of the
earth, and
{\ADD{in}} the
rocks.\FTNT{§}{{\FR 30.6 }
caves: Heb. holes}
\VS{7}Among the
bushes they
brayed; under the
nettles they were gathered
together.
\VS{8}{\ADD{They were}}
children of
fools, yea,
children of base
men: they were
viler than the
earth.\FTNT{**}{{\FR 30.8 }
base…: Heb. men of no name}
\VS{9}And now am I their
song, yea, I am their
byword.
\VS{10}They
abhor me, they flee
far from me, and
spare not to
spit in my
face.\FTNT{††}{{\FR 30.10 }
and…: Heb. and withhold not spittle from}
\VS{11}Because he hath
loosed my
cord, and
afflicted me, they have also let
loose the
bridle
before me.
\VS{12}Upon
{\ADD{my}}
right
{\ADD{hand}}
rise the
youth; they push
away my
feet, and they raise
up against me the
ways of their
destruction.
\VS{13}They
mar my
path, they set
forward my
calamity, they have no
helper.
\VS{14}They
came
{\ADD{upon me}} as a
wide
breaking in
{\ADD{of waters:}}
in the
desolation they
rolled themselves
{\ADD{upon me}}.
\par }{\PP \VS{15}Terrors are
turned upon me: they
pursue my
soul as the
wind: and my
welfare passeth
away as a
cloud.\FTNT{‡‡}{{\FR 30.15 }
my soul: Heb. my principal one}
\VS{16}And now my
soul is poured
out upon me; the
days of
affliction have taken
hold upon me.
\VS{17}My
bones are
pierced in me in the night
season: and my
sinews take no
rest.
\VS{18}By the
great
force
{\ADD{of my disease}} is my
garment
changed: it bindeth me
about as the
collar of my
coat.
\VS{19}He hath
cast me into the
mire, and I am become
like
dust and
ashes.
\VS{20}I
cry unto thee, and thou dost not
hear me: I stand
up, and thou
regardest me
{\ADD{not}}.
\VS{21}Thou art
become
cruel to me: with thy
strong
hand thou
opposest thyself against me.\FTNT{§§}{{\FR 30.21 }
become…: Heb. turned to be cruel}\FTNT{***}{{\FR 30.21 }
thy…: Heb. the strength of thy hand}
\VS{22}Thou liftest me
up to the
wind; thou causest me to
ride
{\ADD{upon it}}, and
dissolvest my
substance.\FTNT{†††}{{\FR 30.22 }
substance: or, wisdom}
\VS{23}For I
know
{\ADD{that}} thou wilt
bring me
{\ADD{to}}
death, and
{\ADD{to}} the
house
appointed for all
living.
\VS{24}Howbeit he will not stretch
out
{\ADD{his}}
hand to the
grave, though they
cry in his
destruction.\FTNT{‡‡‡}{{\FR 30.24 }
grave: Heb. heap}
\VS{25}Did not I
weep for him that was in
trouble? was
{\ADD{not}} my
soul
grieved for the
poor?\FTNT{§§§}{{\FR 30.25 }
in trouble: Heb. hard of day?}
\VS{26}When I
looked for
good, then
evil
came
{\ADD{unto me:}} and when I
waited for
light, there
came
darkness.
\VS{27}My
bowels
boiled, and
rested not: the
days of
affliction
prevented me.
\VS{28}I
went
mourning without the
sun: I stood
up,
{\ADD{and}} I
cried in the
congregation.
\VS{29}I am a
brother to
dragons, and a
companion to
owls.\FTNT{*}{{\FR 30.29 }
owls: or, ostriches}
\VS{30}My
skin is
black upon me, and my
bones are
burned with
heat.
\VS{31}My
harp also is
{\ADD{turned}} to
mourning, and my
organ into the
voice of them that
weep.

\par }\Chap{31}{\PP \VerseOne{1}I
made a
covenant with mine
eyes; why then should I
think upon a
maid?
\VS{2}For what
portion of
God
{\ADD{is there}} from
above? and
{\ADD{what}}
inheritance of the
Almighty from on
high?
\VS{3}{\ADD{Is}} not
destruction to the
wicked? and a
strange
{\ADD{punishment}} to the
workers of
iniquity?
\VS{4}Doth not he
see my
ways, and
count all my
steps?
\VS{5}If I have
walked with
vanity, or if my
foot hath
hasted to
deceit;
\VS{6}Let me be
weighed in an
even
balance, that
God may
know mine
integrity.\FTNT{*}{{\FR 31.6 }
Let…: Heb. Let him weigh me in balances of justice}
\VS{7}If my
step hath turned
out of the
way, and mine
heart
walked
after mine
eyes, and if any
blot hath
cleaved to mine
hands;
\VS{8}{\ADD{Then}} let me
sow, and let
another
eat; yea, let my
offspring be rooted
out.
\par }{\PP \VS{9}If mine
heart have been
deceived by a
woman, or
{\ADD{if}} I have laid
wait at my
neighbour’s
door;
\VS{10}{\ADD{Then}} let my
wife
grind unto
another, and let
others bow
down upon her.
\VS{11}For this
{\ADD{is}} an heinous
crime; yea, it
{\ADD{is}} an
iniquity
{\ADD{to be punished by}} the
judges.
\VS{12}For it
{\ADD{is}} a
fire
{\ADD{that}}
consumeth to
destruction, and would root
out all mine
increase.
\VS{13}If I did
despise the
cause of my
manservant or of my
maidservant, when they
contended with me;
\VS{14}What then shall I
do when
God riseth
up? and when he
visiteth, what shall I
answer him?
\VS{15}Did not he that
made me in the
womb
make him? and did not
one
fashion us in the
womb?\FTNT{†}{{\FR 31.15 }
did not one…: or, did he not fashion us in one womb?}
\par }{\PP \VS{16}If I have
withheld the
poor from
{\ADD{their}}
desire, or have caused the
eyes of the
widow to
fail;
\VS{17}Or have
eaten my
morsel myself alone, and the
fatherless hath not
eaten thereof;
\VS{18}(For from my
youth he was brought
up with me, as
{\ADD{with}} a
father, and I have
guided her from my
mother’s
womb;)\FTNT{‡}{{\FR 31.18 }
her: that is, the widow}
\VS{19}If I have
seen any
perish for want of
clothing, or any
poor without
covering;
\VS{20}If his
loins have not
blessed me, and
{\ADD{if}} he were
{\ADD{not}}
warmed with the
fleece of my
sheep;
\VS{21}If I have lifted
up my
hand against the
fatherless, when I
saw my
help in the
gate:
\VS{22}{\ADD{Then}} let mine
arm
fall from my shoulder
blade, and mine
arm be
broken from the
bone.\FTNT{§}{{\FR 31.22 }
bone: or, chanelbone}
\VS{23}For
destruction
{\ADD{from}}
God
{\ADD{was}} a
terror to me, and by reason of his
highness I
could not endure.
\par }{\PP \VS{24}If I have
made
gold my
hope, or have
said to the fine
gold,
{\ADD{Thou art}} my
confidence;
\VS{25}If I
rejoiced because my
wealth
{\ADD{was}}
great, and because mine
hand had
gotten
much;\FTNT{**}{{\FR 31.25 }
gotten…: Heb. found much}
\VS{26}If I
beheld the
sun when it
shined, or the
moon
walking
{\ADD{in}}
brightness;\FTNT{††}{{\FR 31.26 }
sun: Heb. light}\FTNT{‡‡}{{\FR 31.26 }
in…: Heb. bright}
\VS{27}And my
heart hath been
secretly
enticed, or my
mouth hath
kissed my
hand:\FTNT{§§}{{\FR 31.27 }
my mouth…: Heb. my hand hath kissed my mouth}
\VS{28}This also
{\ADD{were}} an
iniquity
{\ADD{to be punished by}} the
judge: for I should have
denied the
God
{\ADD{that is}}
above.
\VS{29}If I
rejoiced at the
destruction of him that
hated me, or lifted
up myself when
evil
found him:
\VS{30}Neither have I
suffered my
mouth to
sin by
wishing a
curse to his
soul.\FTNT{***}{{\FR 31.30 }
mouth: Heb. palate}
\VS{31}If the
men of my
tabernacle
said not, Oh that we
had of his
flesh! we cannot be
satisfied.
\VS{32}The
stranger did not
lodge in the
street:
{\ADD{but}} I
opened my
doors to the
traveller.\FTNT{†††}{{\FR 31.32 }
traveller: or, way}
\par }{\PP \VS{33}If I
covered my
transgressions as
Adam, by
hiding mine
iniquity in my
bosom:\FTNT{‡‡‡}{{\FR 31.33 }
as Adam: or, after the manner of men}
\VS{34}Did I
fear a
great
multitude, or did the
contempt of
families
terrify me, that I kept
silence,
{\ADD{and}} went not
out of the
door?
\VS{35}Oh
that one would
hear me! behold, my
desire
{\ADD{is, that}} the
Almighty would
answer me, and
{\ADD{that}} mine
adversary had
written a
book.\FTNT{§§§}{{\FR 31.35 }
my…: or, my sign is that the Almighty will}
\VS{36}Surely I would
take it upon my
shoulder,
{\ADD{and}}
bind it
{\ADD{as}} a
crown to me.
\VS{37}I would
declare unto him the
number of my
steps; as a
prince would I go
near unto him.
\VS{38}If my
land
cry against me, or that the
furrows
likewise thereof
complain;\FTNT{*}{{\FR 31.38 }
complain: Heb. weep}
\VS{39}If I have
eaten the
fruits thereof without
money, or have caused the
owners thereof to
lose their
life:\FTNT{†}{{\FR 31.39 }
fruits: Heb. strength}\FTNT{‡}{{\FR 31.39 }
the owners…: Heb. the soul of the owners thereof to expire, or, breathe out}
\VS{40}Let
thistles
grow instead of
wheat, and
cockle instead of
barley. The
words of
Job are
ended.\FTNT{§}{{\FR 31.40 }
cockle: or, noisome weeds}

\par }\Chap{32}{\PP \VerseOne{1}So these
three
men
ceased to
answer
Job, because he
{\ADD{was}}
righteous in his own
eyes.\FTNT{*}{{\FR 32.1 }
to…: Heb. from answering}
\VS{2}Then was
kindled the
wrath of
Elihu the
son of
Barachel the
Buzite, of the
kindred of
Ram: against
Job was his
wrath
kindled, because he
justified
himself rather than
God.\FTNT{†}{{\FR 32.2 }
himself: Heb. his soul}
\VS{3}Also against his
three
friends was his
wrath
kindled, because they had
found no
answer, and
{\ADD{yet}} had
condemned
Job.
\VS{4}Now
Elihu had
waited till
Job had
spoken, because they
{\ADD{were}}
elder than he.\FTNT{‡}{{\FR 32.4 }
waited…: Heb. expected Job in words}\FTNT{§}{{\FR 32.4 }
elder: Heb. elder for days}
\VS{5}When
Elihu
saw that
{\ADD{there was}} no
answer in the
mouth of
{\ADD{these}}
three
men, then his
wrath was
kindled.
\par }{\PP \VS{6}And
Elihu the
son of
Barachel the
Buzite
answered and
said, I
{\ADD{am}}
young, and ye
{\ADD{are}} very
old; wherefore I was
afraid, and
durst not
shew you mine
opinion.\FTNT{**}{{\FR 32.6 }
young: Heb. few of days}\FTNT{††}{{\FR 32.6 }
durst…: Heb. feared}
\VS{7}I
said,
Days should
speak, and
multitude of
years should
teach
wisdom.
\VS{8}But
{\ADD{there is}} a
spirit in
man: and the
inspiration of the
Almighty giveth them
understanding.
\VS{9}Great
men are not
{\ADD{always}}
wise: neither do the
aged
understand
judgment.
\VS{10}Therefore I
said,
Hearken to me; I also will
shew mine
opinion.
\VS{11}Behold, I
waited for your
words; I gave
ear to your
reasons, whilst ye searched
out what to
say.\FTNT{‡‡}{{\FR 32.11 }
reasons: Heb. understandings}\FTNT{§§}{{\FR 32.11 }
what…: Heb. words}
\VS{12}Yea, I
attended unto you, and, behold,
{\ADD{there was}} none of you that
convinced
Job,
{\ADD{or}} that
answered his
words:
\VS{13}Lest ye should
say, We have found
out
wisdom:
God thrusteth him
down, not
man.
\VS{14}Now he hath not
directed
{\ADD{his}}
words against me: neither will I
answer him with your
speeches.\FTNT{***}{{\FR 32.14 }
directed: or, ordered}
\par }{\PP \VS{15}They were
amazed, they
answered no more: they left
off
speaking.\FTNT{†††}{{\FR 32.15 }
left…: Heb. removed speeches from themselves}
\VS{16}When I had
waited, (for they
spake not, but stood
still,
{\ADD{and}}
answered no more;)
\VS{17}{\ADD{I said}}, I will
answer also my
part, I also will
shew mine
opinion.
\VS{18}For I am
full of
matter, the
spirit
within me
constraineth me.\FTNT{‡‡‡}{{\FR 32.18 }
matter: Heb. words}\FTNT{§§§}{{\FR 32.18 }
spirit…: Heb. spirit of my belly}
\VS{19}Behold, my
belly
{\ADD{is}} as
wine
{\ADD{which}} hath no
vent; it is ready to
burst like
new
bottles.\FTNT{*}{{\FR 32.19 }
hath…: Heb. is not opened}
\VS{20}I will
speak, that I may be
refreshed: I will
open my
lips and
answer.\FTNT{†}{{\FR 32.20 }
be…: Heb. breathe}
\VS{21}Let me not, I pray you,
accept any
man’s
person, neither let me give flattering
titles unto
man.
\VS{22}For I
know not to give flattering
titles;
{\ADD{in so doing}} my
maker would
soon take me
away.

\par }\Chap{33}{\PP \VerseOne{1}Wherefore,
Job, I pray thee,
hear my
speeches, and
hearken to all my
words.
\VS{2}Behold, now I have
opened my
mouth, my
tongue hath
spoken in my
mouth.\FTNT{*}{{\FR 33.2 }
in my mouth: Heb. in my palate}
\VS{3}My
words
{\ADD{shall be of}} the
uprightness of my
heart: and my
lips shall
utter
knowledge
clearly.
\VS{4}The
Spirit of
God hath
made me, and the
breath of the
Almighty hath given me
life.
\VS{5}If thou
canst
answer me, set
{\ADD{thy words}} in
order
before me, stand
up.
\VS{6}Behold, I
{\ADD{am}} according to thy
wish in
God’s stead: I also am
formed out of the
clay.\FTNT{†}{{\FR 33.6 }
wish: Heb. mouth}\FTNT{‡}{{\FR 33.6 }
formed: Heb. cut}
\VS{7}Behold, my
terror shall not make thee
afraid, neither shall my
hand be
heavy upon thee.
\par }{\PP \VS{8}Surely thou hast
spoken in mine
hearing, and I have
heard the
voice of
{\ADD{thy}}
words,
{\ADD{saying}},\FTNT{§}{{\FR 33.8 }
hearing: Heb. ears}
\VS{9}I am
clean without
transgression, I
{\ADD{am}}
innocent; neither
{\ADD{is there}}
iniquity in me.
\VS{10}Behold, he
findeth
occasions against me, he
counteth me for his
enemy,
\VS{11}He
putteth my
feet in the
stocks, he
marketh all my
paths.
\VS{12}Behold,
{\ADD{in}} this thou art not
just: I will
answer thee, that
God is
greater than
man.
\VS{13}Why dost thou
strive against him? for he giveth not
account of any of his
matters.\FTNT{**}{{\FR 33.13 }
he giveth…: Heb. he answereth not}
\par }{\PP \VS{14}For
God
speaketh
once, yea
twice,
{\ADD{yet man}}
perceiveth it not.
\VS{15}In a
dream, in a
vision of the
night, when
deep sleep
falleth upon
men, in
slumberings upon the
bed;
\VS{16}Then he
openeth the
ears of
men, and
sealeth their
instruction,\FTNT{††}{{\FR 33.16 }
he…: Heb. he revealeth, or, uncovereth}
\VS{17}That he may
withdraw
man
{\ADD{from his}}
purpose, and
hide
pride from
man.\FTNT{‡‡}{{\FR 33.17 }
purpose: Heb. work}
\VS{18}He keepeth
back his
soul from the
pit, and his
life from
perishing by the
sword.\FTNT{§§}{{\FR 33.18 }
from perishing: Heb. from passing}
\par }{\PP \VS{19}He is
chastened also with
pain upon his
bed, and the
multitude of his
bones with
strong
{\ADD{pain:}}
\VS{20}So that his
life
abhorreth
bread, and his
soul
dainty
meat.\FTNT{***}{{\FR 33.20 }
dainty…: Heb. meat of desire}
\VS{21}His
flesh is consumed
away, that it cannot be
seen; and his
bones
{\ADD{that}} were not
seen stick
out.
\VS{22}Yea, his
soul draweth
near unto the
grave, and his
life to the
destroyers.
\VS{23}If
there be a
messenger with him, an
interpreter,
one among a
thousand, to
shew unto
man his
uprightness:
\VS{24}Then he is
gracious unto him, and
saith,
Deliver him from going
down to the
pit: I have
found a
ransom.\FTNT{†††}{{\FR 33.24 }
a ransom: or, an atonement}
\VS{25}His
flesh shall be
fresher than a
child’s: he shall
return to the
days of his
youth:\FTNT{‡‡‡}{{\FR 33.25 }
a child’s: Heb. childhood}
\VS{26}He shall
pray unto
God, and he will be
favourable unto him: and he shall
see his
face with
joy: for he will
render unto
man his
righteousness.
\VS{27}He
looketh upon
men, and
{\ADD{if any}}
say, I have
sinned, and
perverted
{\ADD{that which was}}
right, and it
profited me not;\FTNT{§§§}{{\FR 33.27 }
He…: or, He shall look upon men, and say}
\VS{28}He will
deliver his
soul from
going into the
pit, and his
life shall
see the
light.\FTNT{*}{{\FR 33.28 }
He…: or, He hath delivered my soul, etc, and my life}
\par }{\PP \VS{29}Lo, all these
{\ADD{things}}
worketh
God
oftentimes with
man,\FTNT{†}{{\FR 33.29 }
oftentimes: Heb. twice and thrice}
\VS{30}To bring
back his
soul from the
pit, to be
enlightened with the
light of the
living.
\VS{31}Mark
well, O
Job,
hearken unto me: hold thy
peace, and I will
speak.
\VS{32}If thou
hast any thing to
say,
answer me:
speak, for I
desire to
justify thee.
\VS{33}If not,
hearken unto me: hold thy
peace, and I shall
teach thee
wisdom.

\par }\Chap{34}{\PP \VerseOne{1}Furthermore
Elihu
answered and
said,
\VS{2}Hear my
words, O ye
wise
{\ADD{men;}} and give
ear unto me, ye that have
knowledge.
\VS{3}For the
ear
trieth
words, as the
mouth
tasteth
meat.\FTNT{*}{{\FR 34.3 }
mouth: Heb. palate}
\VS{4}Let us
choose to us
judgment: let us
know among ourselves what
{\ADD{is}}
good.
\VS{5}For
Job hath
said, I am
righteous: and
God hath taken
away my
judgment.
\VS{6}Should I
lie against my
right? my
wound
{\ADD{is}}
incurable without
transgression.\FTNT{†}{{\FR 34.6 }
my wound: Heb. mine arrow}
\VS{7}What
man
{\ADD{is}} like
Job,
{\ADD{who}} drinketh
up
scorning like
water?
\VS{8}Which
goeth in
company with the
workers of
iniquity, and
walketh with
wicked
men.
\VS{9}For he hath
said, It
profiteth a
man nothing that he should
delight himself with
God.
\par }{\PP \VS{10}Therefore
hearken unto me, ye
men of
understanding: far be
it from
God,
{\ADD{that he should do}}
wickedness; and
{\ADD{from}} the
Almighty,
{\ADD{that he should commit}}
iniquity.\FTNT{‡}{{\FR 34.10 }
men…: Heb. men of heart}
\VS{11}For the
work of a
man shall he
render unto him, and cause every
man to
find according to
{\ADD{his}}
ways.
\VS{12}Yea,
surely
God will not do
wickedly, neither will the
Almighty
pervert
judgment.
\VS{13}Who hath given him a
charge over the
earth? or who hath
disposed the whole
world?\FTNT{§}{{\FR 34.13 }
the whole: Heb. all of it?}
\VS{14}If he
set his
heart upon man,
{\ADD{if}} he
gather unto himself his
spirit and his
breath;\FTNT{**}{{\FR 34.14 }
man: Heb. him}
\VS{15}All
flesh shall
perish
together, and
man shall turn
again unto
dust.
\par }{\PP \VS{16}If now
{\ADD{thou hast}}
understanding,
hear this:
hearken to the
voice of my
words.
\VS{17}Shall even he that
hateth
right
govern? and wilt thou
condemn him that is
most
just?\FTNT{††}{{\FR 34.17 }
govern: Heb. bind?}
\VS{18}{\ADD{Is it fit}} to
say to a
king,
{\ADD{Thou art}}
wicked?
{\ADD{and}} to
princes,
{\ADD{Ye are}}
ungodly?
\VS{19}{\ADD{How much less to him}} that
accepteth not the
persons of
princes, nor
regardeth the
rich more
than the
poor? for they all
{\ADD{are}} the
work of his
hands.
\VS{20}In a
moment shall they
die, and the
people shall be
troubled at
midnight, and pass
away: and the
mighty shall be taken
away without
hand.\FTNT{‡‡}{{\FR 34.20 }
the mighty…: Heb. they shall take away the mighty}
\VS{21}For his
eyes
{\ADD{are}} upon the
ways of
man, and he
seeth all his
goings.
\VS{22}{\ADD{There is}} no
darkness, nor shadow of
death, where the
workers of
iniquity may
hide themselves.
\VS{23}For he will not
lay upon
man more
{\ADD{than right;}} that he should
enter into
judgment with
God.\FTNT{§§}{{\FR 34.23 }
enter: Heb. go}
\VS{24}He shall break in
pieces mighty
men without
number, and
set
others in their stead.\FTNT{***}{{\FR 34.24 }
number: Heb. searching out}
\VS{25}Therefore he
knoweth their
works, and he
overturneth
{\ADD{them}} in the
night, so that they are
destroyed.\FTNT{†††}{{\FR 34.25 }
destroyed: Heb. crushed}
\VS{26}He
striketh them
as wicked
men in the
open
sight of others;\FTNT{‡‡‡}{{\FR 34.26 }
open…: Heb. place of beholders}
\VS{27}Because they turned
back from
him, and would not
consider any of his
ways:\FTNT{§§§}{{\FR 34.27 }
him: Heb. after him}
\VS{28}So that they cause the
cry of the
poor to
come unto him, and he
heareth the
cry of the
afflicted.
\VS{29}When he giveth
quietness, who then can make
trouble? and when he
hideth
{\ADD{his}}
face, who then can
behold him? whether
{\ADD{it be done}} against a
nation, or against a
man
only:
\VS{30}That the
hypocrite
reign not, lest the
people be
ensnared.
\par }{\PP \VS{31}Surely it is meet to be
said unto
God, I have
borne
{\ADD{chastisement}}, I will not
offend
{\ADD{any more:}}
\VS{32}{\ADD{That which}} I
see
not
teach thou me: if I have
done
iniquity, I will
do no more.
\VS{33}{\ADD{Should it be}} according to thy mind? he will
recompense it, whether thou
refuse, or whether thou
choose; and not I: therefore
speak what thou
knowest.\FTNT{*}{{\FR 34.33 }
according…: Heb. from with thee?}
\VS{34}Let
men of
understanding
tell me, and let a
wise
man
hearken unto me.\FTNT{†}{{\FR 34.34 }
of…: Heb. of heart}
\VS{35}Job hath
spoken without
knowledge, and his
words
{\ADD{were}} without
wisdom.
\VS{36}My
desire
{\ADD{is that}}
Job may be
tried unto the
end because of
{\ADD{his}}
answers for
wicked
men.\FTNT{‡}{{\FR 34.36 }
My…: or, My father, let Job be tried}
\VS{37}For he
addeth
rebellion unto his
sin, he
clappeth
{\ADD{his hands}} among us, and
multiplieth his
words against
God.

\par }\Chap{35}{\PP \VerseOne{1}Elihu
spake moreover, and
said,
\VS{2}Thinkest thou this to be
right,
{\ADD{that}} thou
saidst, My
righteousness
{\ADD{is}} more than
God’s?
\VS{3}For thou
saidst, What
advantage will it be unto thee?
{\ADD{and}}, What
profit shall I have,
{\ADD{if I be cleansed}} from my
sin?\FTNT{*}{{\FR 35.3 }
if…: or, by it more than by my sin}
\VS{4}I will
answer thee, and thy
companions with thee.\FTNT{†}{{\FR 35.4 }
answer…: Heb. return to thee words}
\VS{5}Look unto the
heavens, and
see; and
behold the
clouds
{\ADD{which}} are
higher than thou.
\VS{6}If thou
sinnest, what
doest thou against him? or
{\ADD{if}} thy
transgressions be
multiplied, what
doest thou unto him?
\VS{7}If thou be
righteous, what
givest thou him? or what
receiveth he of thine
hand?
\VS{8}Thy
wickedness
{\ADD{may hurt}} a
man as thou
{\ADD{art;}} and thy
righteousness
{\ADD{may profit}} the
son of
man.
\par }{\PP \VS{9}By reason of the
multitude of
oppressions they make
{\ADD{the oppressed}} to
cry: they cry
out by reason of the
arm of the
mighty.
\VS{10}But none
saith, Where
{\ADD{is}}
God my
maker, who
giveth
songs in the
night;
\VS{11}Who
teacheth us more than the
beasts of the
earth, and maketh us
wiser than the
fowls of
heaven?
\VS{12}There they
cry, but none giveth
answer,
because of the
pride of evil
men.
\VS{13}Surely
God will not
hear
vanity, neither will the
Almighty
regard it.
\par }{\PP \VS{14}Although thou
sayest thou shalt not
see him,
{\ADD{yet}}
judgment
{\ADD{is}}
before him; therefore
trust thou in him.
\VS{15}But now, because
{\ADD{it is}} not
{\ADD{so}}, he hath
visited in his
anger; yet he
knoweth
{\ADD{it}} not in
great
extremity:\FTNT{‡}{{\FR 35.15 }
he hath: that is, God hath}\FTNT{§}{{\FR 35.15 }
he knoweth: that is, Job knoweth}
\VS{16}Therefore doth
Job
open his
mouth in
vain; he
multiplieth
words
without
knowledge.

\par }\Chap{36}{\PP \VerseOne{1}Elihu also
proceeded, and
said,
\VS{2}Suffer me a
little, and I will
shew thee that
{\ADD{I have}} yet to
speak on
God’s behalf.\FTNT{*}{{\FR 36.2 }
I have…: Heb. there are yet words for God}
\VS{3}I will
fetch my
knowledge from
afar, and will
ascribe
righteousness to my
Maker.
\VS{4}For
truly my
words
{\ADD{shall}} not
{\ADD{be}}
false: he that is
perfect in
knowledge
{\ADD{is}} with thee.
\par }{\PP \VS{5}Behold,
God
{\ADD{is}}
mighty, and
despiseth not
{\ADD{any: he is}}
mighty in
strength
{\ADD{and}}
wisdom.\FTNT{†}{{\FR 36.5 }
wisdom: Heb. heart}
\VS{6}He preserveth not the
life of the
wicked: but
giveth
right to the
poor.\FTNT{‡}{{\FR 36.6 }
poor: or, afflicted}
\VS{7}He
withdraweth not his
eyes from the
righteous: but with
kings
{\ADD{are they}} on the
throne; yea, he doth
establish them for
ever, and they are
exalted.
\VS{8}And if
{\ADD{they be}}
bound in
fetters,
{\ADD{and}} be
holden in
cords of
affliction;
\VS{9}Then he
sheweth them their
work, and their
transgressions that they have
exceeded.
\VS{10}He
openeth also their
ear to
discipline, and
commandeth that they
return from
iniquity.
\VS{11}If they
obey and
serve
{\ADD{him}}, they shall
spend their
days in
prosperity, and their
years in
pleasures.
\VS{12}But if they
obey not, they shall
perish by the
sword, and they shall
die without
knowledge.\FTNT{§}{{\FR 36.12 }
perish: Heb. pass away}
\VS{13}But the
hypocrites in
heart heap
up
wrath: they
cry not when he
bindeth them.
\VS{14}They
die in
youth, and their
life
{\ADD{is}} among the
unclean.\FTNT{**}{{\FR 36.14 }
They…: Heb. Their soul dieth}\FTNT{††}{{\FR 36.14 }
unclean: or, sodomites}
\par }{\PP \VS{15}He
delivereth the
poor in his
affliction, and
openeth their
ears in
oppression.\FTNT{‡‡}{{\FR 36.15 }
poor: or, afflicted}
\VS{16}Even so would he have
removed thee out of the
strait
{\ADD{into}} a broad
place,
where
{\ADD{there is}} no
straitness; and that which should be
set on thy
table
{\ADD{should be}}
full of
fatness.\FTNT{§§}{{\FR 36.16 }
that…: Heb. the rest of thy table}
\VS{17}But thou hast
fulfilled the
judgment of the
wicked:
judgment and
justice take
hold
{\ADD{on thee}}.\FTNT{***}{{\FR 36.17 }
take…: or, should uphold thee}
\VS{18}Because
{\ADD{there is}}
wrath,
{\ADD{beware}} lest he take thee
away with
{\ADD{his}}
stroke: then a
great
ransom cannot
deliver thee.\FTNT{†††}{{\FR 36.18 }
deliver…: Heb. turn thee aside}
\VS{19}Will he
esteem thy
riches?
{\ADD{no}}, not
gold, nor all the
forces of
strength.
\VS{20}Desire not the
night, when
people are cut
off in their place.
\VS{21}Take
heed,
regard not
iniquity: for this hast thou
chosen rather than
affliction.
\VS{22}Behold,
God
exalteth by his
power: who
teacheth like him?
\VS{23}Who hath
enjoined him his
way? or who can
say, Thou hast
wrought
iniquity?
\par }{\PP \VS{24}Remember that thou
magnify his
work, which
men
behold.
\VS{25}Every
man may
see it;
man may
behold
{\ADD{it}} afar
off.
\VS{26}Behold,
God
{\ADD{is}}
great, and we
know
{\ADD{him}} not, neither can the
number of his
years be searched
out.
\VS{27}For he maketh
small the
drops of
water: they pour
down
rain according to the
vapour thereof:
\VS{28}Which the
clouds do
drop
{\ADD{and}}
distil upon
man
abundantly.
\VS{29}Also can
{\ADD{any}}
understand the
spreadings of the
clouds,
{\ADD{or}} the
noise of his
tabernacle?
\VS{30}Behold, he
spreadeth his
light upon it, and
covereth the
bottom of the
sea.\FTNT{‡‡‡}{{\FR 36.30 }
bottom: Heb. roots}
\VS{31}For by them
judgeth he the
people; he
giveth
meat in
abundance.
\VS{32}With
clouds he
covereth the
light; and
commandeth it
{\ADD{not to shine}} by
{\ADD{the cloud}} that cometh
betwixt.
\VS{33}The
noise thereof
sheweth concerning it, the
cattle also concerning the
vapour.\FTNT{§§§}{{\FR 36.33 }
the vapour: Heb. that which goeth up}

\par }\Chap{37}{\PP \VerseOne{1}At
this also my
heart
trembleth, and is moved
out of his
place.
\VS{2}Hear
attentively the
noise of his
voice, and the
sound
{\ADD{that}} goeth
out of his
mouth.\FTNT{*}{{\FR 37.2 }
Hear…: Heb. Hear in hearing}
\VS{3}He
directeth
it under the whole
heaven, and his
lightning unto the
ends of the
earth.\FTNT{†}{{\FR 37.3 }
lightning: Heb. light}\FTNT{‡}{{\FR 37.3 }
ends: Heb. wings}
\VS{4}After it a
voice
roareth: he
thundereth with the
voice of his
excellency; and he will not
stay them when his
voice is
heard.
\VS{5}God
thundereth
marvellously with his
voice; great
things
doeth he, which we cannot
comprehend.
\par }{\PP \VS{6}For he
saith to the
snow, Be
thou
{\ADD{on}} the
earth; likewise to the
small
rain, and to the
great
rain of his
strength.\FTNT{§}{{\FR 37.6 }
likewise…: Heb. and to the shower of rain, and to the showers of rain of his strength}
\VS{7}He sealeth
up the
hand of every
man; that all
men may
know his
work.
\VS{8}Then the
beasts
go
into
dens, and
remain in their
places.
\VS{9}Out of the
south
cometh the
whirlwind: and
cold out of the
north.\FTNT{**}{{\FR 37.9 }
south: Heb. chamber}\FTNT{††}{{\FR 37.9 }
north: Heb. scattering winds}
\VS{10}By the
breath of
God
frost is
given: and the
breadth of the
waters is
straitened.
\VS{11}Also by
watering he
wearieth the thick
cloud: he
scattereth his
bright
cloud:\FTNT{‡‡}{{\FR 37.11 }
his…: Heb. the cloud of his light}
\VS{12}And it is
turned round
about by his
counsels: that they may
do whatsoever he
commandeth them upon the
face of the
world in the
earth.
\VS{13}He causeth it to
come, whether for
correction, or for his
land, or for
mercy.\FTNT{§§}{{\FR 37.13 }
correction: Heb. a rod}
\par }{\PP \VS{14}Hearken unto this, O
Job: stand
still, and
consider the
wondrous works of
God.
\VS{15}Dost thou
know when
God
disposed them, and caused the
light of his
cloud to
shine?
\VS{16}Dost thou
know the
balancings of the
clouds, the wondrous
works of him which is
perfect in
knowledge?
\VS{17}How thy
garments
{\ADD{are}}
warm, when he
quieteth the
earth by the
south
{\ADD{wind}}?
\VS{18}Hast thou with him spread
out the
sky,
{\ADD{which is}}
strong,
{\ADD{and}} as a
molten looking
glass?
\VS{19}Teach us what we shall
say unto him;
{\ADD{for}} we cannot
order
{\ADD{our speech}} by
reason of
darkness.
\VS{20}Shall it be
told him that I
speak? if a
man
speak, surely he shall be swallowed
up.
\par }{\PP \VS{21}And now
{\ADD{men}}
see not the
bright
light which
{\ADD{is}} in the
clouds: but the
wind
passeth, and
cleanseth them.
\VS{22}Fair
weather
cometh out of the
north: with
God
{\ADD{is}}
terrible
majesty.\FTNT{***}{{\FR 37.22 }
Fair…: Heb. Gold}
\VS{23}{\ADD{Touching}} the
Almighty, we cannot find him
out:
{\ADD{he is}}
excellent in
power, and in
judgment, and in
plenty of
justice: he will not
afflict.
\VS{24}Men do therefore
fear him: he
respecteth not any
{\ADD{that are}}
wise of
heart.

\par }\Chap{38}{\PP \VerseOne{1}Then the
{\ND{
LORD}}
answered
Job out of the
whirlwind, and
said,
\VS{2}Who
{\ADD{is}} this that
darkeneth
counsel by
words without
knowledge?
\VS{3}Gird
up now thy
loins like a
man; for I will
demand of thee, and
answer thou me.\FTNT{*}{{\FR 38.3 }
answer…: Heb. make me know}
\par }{\PP \VS{4}Where wast thou when I laid the
foundations of the
earth?
declare, if thou
hast
understanding.\FTNT{†}{{\FR 38.4 }
hast…: Heb. knowest understanding}
\VS{5}Who hath
laid the
measures thereof, if thou
knowest? or who hath
stretched the
line upon it?
\VS{6}Whereupon are the
foundations thereof
fastened? or who
laid the
corner
stone thereof;\FTNT{‡}{{\FR 38.6 }
foundations: Heb. sockets}\FTNT{§}{{\FR 38.6 }
fastened: Heb. made to sink?}
\VS{7}When the
morning
stars
sang
together, and all the
sons of
God
shouted for joy?
\VS{8}Or
{\ADD{who}} shut
up the
sea with
doors, when it brake
forth,
{\ADD{as if}} it had issued
out of the
womb?
\VS{9}When I
made the
cloud the
garment thereof, and thick
darkness a
swaddlingband for it,
\VS{10}And brake
up for it my
decreed
{\ADD{place}}, and
set
bars and
doors,\FTNT{**}{{\FR 38.10 }
brake…: or, established my decree upon it}
\VS{11}And
said,
Hitherto shalt thou
come, but no
further: and here shall thy
proud
waves be
stayed?\FTNT{††}{{\FR 38.11 }
thy…: Heb. the pride of thy waves}
\par }{\PP \VS{12}Hast thou
commanded the
morning since thy
days;
{\ADD{and}} caused the
dayspring to
know his
place;
\VS{13}That it might take
hold of the
ends of the
earth, that the
wicked might be shaken
out of it?\FTNT{‡‡}{{\FR 38.13 }
ends: Heb. wings}
\VS{14}It is
turned as
clay
{\ADD{to}} the
seal; and they
stand as a
garment.
\VS{15}And from the
wicked their
light is
withholden, and the
high
arm shall be
broken.
\VS{16}Hast thou
entered into the
springs of the
sea? or hast thou
walked in the
search of the
depth?
\VS{17}Have the
gates of
death been
opened unto thee? or hast thou
seen the
doors of the shadow of
death?
\VS{18}Hast thou
perceived the
breadth of the
earth?
declare if thou
knowest it all.
\VS{19}Where
{\ADD{is}} the
way
{\ADD{where}}
light
dwelleth? and
{\ADD{as for}}
darkness, where
{\ADD{is}} the
place thereof,
\VS{20}That thou shouldest
take it to the
bound thereof, and that thou shouldest
know the
paths
{\ADD{to}} the
house thereof?\FTNT{§§}{{\FR 38.20 }
to the bound: or, at, etc}
\VS{21}Knowest thou
{\ADD{it}}, because thou wast then
born? or
{\ADD{because}} the
number of thy
days
{\ADD{is}}
great?
\VS{22}Hast thou
entered into the
treasures of the
snow? or hast thou
seen the
treasures of the
hail,
\VS{23}Which I have
reserved against the
time of
trouble, against the
day of
battle and
war?
\VS{24}By
what
way is the
light
parted,
{\ADD{which}}
scattereth the east
wind upon the
earth?
\par }{\PP \VS{25}Who hath
divided a
watercourse for the overflowing of
waters, or a
way for the
lightning of
thunder;
\VS{26}To cause it to
rain on the
earth,
{\ADD{where}} no
man
{\ADD{is; on}} the
wilderness, wherein
{\ADD{there is}} no
man;
\VS{27}To
satisfy the
desolate and
waste
{\ADD{ground;}} and to cause the
bud of the tender
herb to spring
forth?
\VS{28}Hath the
rain a
father? or who hath
begotten the
drops of
dew?
\VS{29}Out of whose
womb
came the
ice? and the hoary
frost of
heaven, who hath
gendered it?
\VS{30}The
waters are
hid as
{\ADD{with}} a
stone, and the
face of the
deep is
frozen.\FTNT{***}{{\FR 38.30 }
is…: Heb. is taken}
\VS{31}Canst thou
bind the sweet
influences of
Pleiades, or
loose the
bands of
Orion?\FTNT{†††}{{\FR 38.31 }
Pleiades: or, the seven stars: Heb. Cimah}\FTNT{‡‡‡}{{\FR 38.31 }
Orion: Heb. Cesil?}
\VS{32}Canst thou bring
forth
Mazzaroth in his
season? or canst thou
guide
Arcturus with his
sons?\FTNT{§§§}{{\FR 38.32 }
Mazzaroth: or, the twelve signs}\FTNT{*}{{\FR 38.32 }
guide: Heb. guide them}
\VS{33}Knowest thou the
ordinances of
heaven? canst thou
set the
dominion thereof in the
earth?
\VS{34}Canst thou lift
up thy
voice to the
clouds, that
abundance of
waters may
cover thee?
\VS{35}Canst thou
send
lightnings, that they may
go, and
say unto thee, Here we
{\ADD{are}}?\FTNT{†}{{\FR 38.35 }
Here…: Heb. Behold us?}
\VS{36}Who hath
put
wisdom in the inward
parts? or who hath
given
understanding to the
heart?
\VS{37}Who can
number the
clouds in
wisdom? or who can
stay the
bottles of
heaven,\FTNT{‡}{{\FR 38.37 }
stay: Heb. cause to lie down}
\VS{38}When the
dust
groweth into
hardness, and the
clods cleave fast
together?\FTNT{§}{{\FR 38.38 }
groweth…: or, is turned into mire: Heb. is poured}
\VS{39}Wilt thou
hunt the
prey for the
lion? or
fill the
appetite of the young
lions,\FTNT{**}{{\FR 38.39 }
the appetite: Heb. the life}
\VS{40}When they
couch in
{\ADD{their}}
dens,
{\ADD{and}}
abide in the
covert
to lie in
wait?
\VS{41}Who
provideth for the
raven his
food? when his young
ones
cry unto
God, they
wander for
lack of
meat.

\par }\Chap{39}{\PP \VerseOne{1}Knowest thou the
time when the wild
goats of the
rock bring
forth?
{\ADD{or}} canst thou
mark when the
hinds do
calve?
\VS{2}Canst thou
number the
months
{\ADD{that}} they
fulfil? or
knowest thou the
time when they bring
forth?
\VS{3}They
bow themselves, they bring
forth their young
ones, they cast
out their
sorrows.
\VS{4}Their young
ones are in good
liking, they grow
up with
corn; they go
forth, and
return not unto them.
\VS{5}Who hath sent
out the wild
ass
free? or who hath
loosed the
bands of the wild
ass?
\VS{6}Whose
house I have
made the
wilderness, and the
barren land his
dwellings.\FTNT{*}{{\FR 39.6 }
barren…: Heb. salt places}
\VS{7}He
scorneth the
multitude of the
city, neither
regardeth he the
crying of the
driver.\FTNT{†}{{\FR 39.7 }
of the driver: Heb. of the exactor}
\VS{8}The
range of the
mountains
{\ADD{is}} his
pasture, and he
searcheth
after every green
thing.
\VS{9}Will the
unicorn be
willing to
serve thee, or
abide by thy
crib?
\VS{10}Canst thou
bind the
unicorn with his
band in the
furrow? or will he
harrow the
valleys
after thee?
\VS{11}Wilt thou
trust him, because his
strength
{\ADD{is}}
great? or wilt thou
leave thy
labour to him?
\VS{12}Wilt thou
believe him, that he will bring
home thy
seed, and
gather
{\ADD{it into}} thy
barn?
\par }{\PP \VS{13}{\ADD{Gavest thou}} the
goodly
wings unto the
peacocks? or
wings and
feathers unto the
ostrich?\FTNT{‡}{{\FR 39.13 }
wings and…: or, the feathers of the stork and ostrich}
\VS{14}Which
leaveth her
eggs in the
earth, and
warmeth them in
dust,
\VS{15}And
forgetteth that the
foot may
crush them, or that the
wild
beast may
break them.
\VS{16}She is
hardened against her young
ones, as though
{\ADD{they were}}
not hers: her
labour is in
vain without
fear;
\VS{17}Because
God hath
deprived her of
wisdom, neither hath he
imparted to her
understanding.
\VS{18}What
time she lifteth
up herself on
high, she
scorneth the
horse and his
rider.
\par }{\PP \VS{19}Hast thou
given the
horse
strength? hast thou
clothed his
neck with
thunder?
\VS{20}Canst thou make him
afraid as a
grasshopper? the
glory of his
nostrils
{\ADD{is}}
terrible.\FTNT{§}{{\FR 39.20 }
terrible: Heb. terror}
\VS{21}He
paweth in the
valley, and
rejoiceth in
{\ADD{his}}
strength: he goeth
on to
meet the armed
men.\FTNT{**}{{\FR 39.21 }
He paweth: or, His feet dig}\FTNT{††}{{\FR 39.21 }
the armed…: Heb. the armour}
\VS{22}He
mocketh at
fear, and is not
affrighted; neither turneth he
back
from the
sword.
\VS{23}The
quiver
rattleth against him, the
glittering
spear and the
shield.
\VS{24}He
swalloweth the
ground with
fierceness and
rage: neither
believeth he that
{\ADD{it is}} the
sound of the
trumpet.
\VS{25}He
saith
among the
trumpets,
Ha,
ha; and he
smelleth the
battle afar
off, the
thunder of the
captains, and the
shouting.
\par }{\PP \VS{26}Doth the
hawk
fly by thy
wisdom,
{\ADD{and}}
stretch her
wings toward the
south?
\VS{27}Doth the
eagle mount
up at thy
command, and
make her
nest on
high?\FTNT{‡‡}{{\FR 39.27 }
at…: Heb. by thy mouth}
\VS{28}She
dwelleth and
abideth on the
rock, upon the
crag of the
rock, and the strong
place.
\VS{29}From thence she
seeketh the
prey,
{\ADD{and}} her
eyes
behold afar
off.
\VS{30}Her young
ones also suck
up
blood: and where the
slain
{\ADD{are}}, there
{\ADD{is}}
she.

\par }\Chap{40}{\PP \VerseOne{1}Moreover the
{\ND{
LORD}}
answered
Job, and
said,
\VS{2}Shall he that
contendeth with the
Almighty
instruct
{\ADD{him}}? he that
reproveth
God, let him
answer it.
\par }{\PP \VS{3}¶ Then
Job
answered the
{\ND{
LORD}}, and
said,
\VS{4}Behold, I am
vile; what shall I
answer thee? I will
lay mine
hand
upon my
mouth.
\VS{5}Once have I
spoken; but I will not
answer: yea,
twice; but I will proceed no
further.
\par }{\PP \VS{6}¶ Then
answered the
{\ND{
LORD}} unto
Job out of the
whirlwind, and
said,
\VS{7}Gird
up thy
loins now like a
man: I will
demand of thee, and
declare thou unto me.
\VS{8}Wilt thou also
disannul my
judgment? wilt thou
condemn me, that thou mayest be
righteous?
\VS{9}Hast thou an
arm like
God? or canst thou
thunder with a
voice like him?
\VS{10}Deck thyself now
{\ADD{with}}
majesty and
excellency; and
array thyself with
glory and
beauty.
\VS{11}Cast
abroad the
rage of thy
wrath: and
behold every one
{\ADD{that is}}
proud, and
abase him.
\VS{12}Look on every one
{\ADD{that is}}
proud,
{\ADD{and}} bring him
low; and tread
down the
wicked in their place.
\VS{13}Hide them in the
dust
together;
{\ADD{and}}
bind their
faces in
secret.
\VS{14}Then will I also
confess unto thee that thine own right
hand can
save thee.
\par }{\PP \VS{15}¶ Behold now
behemoth, which I
made with thee; he
eateth
grass as an
ox.\FTNT{*}{{\FR 40.15 }
behemoth: probably an extinct animal of some kind}
\VS{16}Lo now, his
strength
{\ADD{is}} in his
loins, and his
force
{\ADD{is}} in the
navel of his
belly.
\VS{17}He
moveth his
tail like a
cedar: the
sinews of his
stones are wrapped
together.\FTNT{†}{{\FR 40.17 }
He…: or, He setteth up}
\VS{18}His
bones
{\ADD{are as}}
strong pieces of
brass; his
bones
{\ADD{are}} like
bars of
iron.
\VS{19}He
{\ADD{is}} the
chief of the
ways of
God: he that
made him can
make his
sword to
approach
{\ADD{unto him}}.
\VS{20}Surely the
mountains bring him
forth
food, where all the
beasts of the
field
play.
\VS{21}He
lieth under the shady
trees, in the
covert of the
reed, and
fens.
\VS{22}The shady
trees
cover him
{\ADD{with}} their
shadow; the
willows of the
brook compass him
about.
\VS{23}Behold, he drinketh
up a
river,
{\ADD{and}}
hasteth not: he
trusteth that he can draw
up
Jordan into his
mouth.\FTNT{‡}{{\FR 40.23 }
he drinketh up: Heb. he oppresseth}
\VS{24}He
taketh it with his
eyes:
{\ADD{his}}
nose pierceth
through
snares.\FTNT{§}{{\FR 40.24 }
He…: or, Will any take him in his sight, or, bore his nose with a gin?}

\par }\Chap{41}{\PP \VerseOne{1}Canst thou draw
out
leviathan with an
hook? or his
tongue with a
cord
{\ADD{which}} thou lettest
down?\FTNT{*}{{\FR 41.1 }
leviathan: probably an extinct animal of some kind}\FTNT{†}{{\FR 41.1 }
which…: Heb. which thou drownest?}
\VS{2}Canst thou
put an
hook into his
nose? or
bore his
jaw through with a
thorn?
\VS{3}Will he make
many
supplications unto thee? will he
speak
soft
{\ADD{words}} unto thee?
\VS{4}Will he
make a
covenant with thee? wilt thou
take him for a
servant for
ever?
\VS{5}Wilt thou
play with him as
{\ADD{with}} a
bird? or wilt thou
bind him for thy
maidens?
\VS{6}Shall the
companions make a
banquet of him? shall they
part him among the
merchants?
\VS{7}Canst thou
fill his
skin with barbed
irons? or his
head with
fish
spears?
\VS{8}Lay thine
hand upon him,
remember the
battle, do no
more.
\VS{9}Behold, the
hope of him is in
vain: shall not
{\ADD{one}} be cast
down even at the
sight of him?
\VS{10}None
{\ADD{is so}}
fierce that dare stir him
up: who then is able to
stand
before me?
\par }{\PP \VS{11}Who hath
prevented me, that I should
repay
{\ADD{him? whatsoever is}} under the whole
heaven is mine.
\VS{12}I will not
conceal his
parts, nor his
power, nor his
comely
proportion.
\VS{13}Who can
discover the
face of his
garment?
{\ADD{or}} who can
come
{\ADD{to him}} with his
double
bridle?\FTNT{‡}{{\FR 41.13 }
with: or, within}
\VS{14}Who can
open the
doors of his
face? his
teeth
{\ADD{are}}
terrible round
about.
\VS{15}{\ADD{His}}
scales
{\ADD{are his}}
pride, shut up
together
{\ADD{as with}} a
close
seal.\FTNT{§}{{\FR 41.15 }
scales: Heb. strong pieces of shields}
\VS{16}One is so
near to
another, that no
air can
come between them.
\VS{17}They are
joined
one to
another, they stick
together, that they cannot be
sundered.
\VS{18}By his
neesings a
light doth
shine, and his
eyes
{\ADD{are}} like the
eyelids of the
morning.
\VS{19}Out of his
mouth
go burning
lamps,
{\ADD{and}}
sparks of
fire leap
out.
\VS{20}Out of his
nostrils
goeth
smoke, as
{\ADD{out}} of a
seething
pot or
caldron.
\VS{21}His
breath
kindleth
coals, and a
flame goeth
out of his
mouth.
\VS{22}In his
neck
remaineth
strength, and
sorrow is turned into
joy
before him.\FTNT{**}{{\FR 41.22 }
is turned into joy: Heb. rejoiceth}
\VS{23}The
flakes of his
flesh are joined
together: they are
firm in themselves; they cannot be
moved.\FTNT{††}{{\FR 41.23 }
flakes: Heb. fallings}
\VS{24}His
heart is as
firm as a
stone; yea, as
hard as a
piece of the
nether
{\ADD{millstone}}.
\VS{25}When he raiseth
up himself, the
mighty are
afraid: by reason of
breakings they
purify themselves.
\VS{26}The
sword of him that
layeth at him cannot
hold: the
spear, the
dart, nor the
habergeon.\FTNT{‡‡}{{\FR 41.26 }
habergeon: or, breastplate}
\VS{27}He
esteemeth
iron as
straw,
{\ADD{and}}
brass as
rotten
wood.
\VS{28}The
arrow cannot make him
flee:
slingstones are
turned with him into
stubble.
\VS{29}Darts are
counted as
stubble: he
laugheth at the
shaking of a
spear.
\VS{30}Sharp
stones
{\ADD{are}} under him: he
spreadeth sharp pointed
things upon the
mire.\FTNT{§§}{{\FR 41.30 }
Sharp stones: Heb. Sharp pieces of potsherd}
\VS{31}He maketh the
deep to
boil like a
pot: he
maketh the
sea like a pot of
ointment.
\VS{32}He maketh a
path to
shine
after him;
{\ADD{one}} would
think the
deep
{\ADD{to be}}
hoary.
\VS{33}Upon
earth there is not his
like, who is
made
without
fear.\FTNT{***}{{\FR 41.33 }
is made without fear: or, behave themselves without fear}
\VS{34}He
beholdeth all
high
{\ADD{things:}} he
{\ADD{is}} a
king over all the
children of
pride.

\par }\Chap{42}{\PP \VerseOne{1}Then
Job
answered the
{\ND{
LORD}}, and
said,
\VS{2}I
know that thou canst
do every
{\ADD{thing}}, and
{\ADD{that}} no
thought can be
withholden from thee.\FTNT{*}{{\FR 42.2 }
no thought can be…: or, no thought of thine can be hindered}
\VS{3}Who
{\ADD{is}} he that
hideth
counsel without
knowledge? therefore have I
uttered that I
understood not; things too
wonderful for me, which I
knew not.
\VS{4}Hear, I beseech thee, and I will
speak: I will
demand of thee, and
declare thou unto me.
\VS{5}I have
heard of thee by the
hearing of the
ear: but now mine
eye
seeth thee.
\VS{6}Wherefore I
abhor
{\ADD{myself}}, and
repent in
dust and
ashes.
\par }{\PP \VS{7}¶ And it was
{\ADD{so}}, that
after the
{\ND{
LORD}} had
spoken these
words unto
Job, the
{\ND{
LORD}}
said to
Eliphaz the
Temanite, My
wrath is
kindled against thee, and against thy
two
friends: for ye have not
spoken of me
{\ADD{the thing that is}}
right, as my
servant
Job
{\ADD{hath}}.
\VS{8}Therefore
take unto you now
seven
bullocks and
seven
rams, and
go to my
servant
Job, and offer
up for yourselves a burnt
offering; and my
servant
Job shall
pray for you: for
him will I
accept: lest I
deal with you
{\ADD{after your}}
folly, in that ye have not
spoken of me
{\ADD{the thing which is}}
right, like my
servant
Job.\FTNT{†}{{\FR 42.8 }
him: Heb. his face, or, person}
\VS{9}So
Eliphaz the
Temanite and
Bildad the
Shuhite
{\ADD{and}}
Zophar the
Naamathite
went, and
did according as the
{\ND{
LORD}}
commanded them: the
{\ND{
LORD}} also
accepted
Job.\FTNT{‡}{{\FR 42.9 }
Job: Heb. the face of Job}
\par }{\PP \VS{10}And the
{\ND{
LORD}}
turned the
captivity of
Job, when he
prayed for his
friends: also the
{\ND{
LORD}}
gave
Job twice as
much as he had before.\FTNT{§}{{\FR 42.10 }
gave…: Heb. added all that had been to Job unto the double}
\VS{11}Then
came there unto him all his
brethren, and all his
sisters, and all they that had been of his
acquaintance
before, and did
eat
bread with him in his
house: and they
bemoaned him, and
comforted him over all the
evil that the
{\ND{
LORD}} had
brought upon him: every
man also
gave him
a piece of
money, and every
one an
earring of
gold.
\VS{12}So the
{\ND{
LORD}}
blessed the latter
end of
Job more than his
beginning: for he had
fourteen
thousand
sheep, and
six
thousand
camels, and a
thousand
yoke of
oxen, and a
thousand she
asses.
\VS{13}He had also
seven
sons and
three
daughters.
\VS{14}And he
called the
name of the
first,
Jemima; and the
name of the
second,
Kezia; and the
name of the
third,
Keren-happuch.
\VS{15}And in all the
land were no
women
found
{\ADD{so}}
fair as the
daughters of
Job: and their
father
gave them
inheritance
among their
brethren.
\VS{16}After this
lived
Job an
hundred and
forty
years, and
saw his
sons, and his
sons’
sons,
{\ADD{even}}
four
generations.
\VS{17}So
Job
died,
{\ADD{being}}
old and
full of
days.
\par }