\NormalFont\ShortTitle{Jonah}
{\MT Jonah

\par }\ChapOne{1}{\PP \VerseOne{1}Now the
word of the
{\ND{
LORD}} came unto
Jonah the
son of
Amittai,
saying,\FTNT{*}{{\FR 1.1 }
Jonah: Gr. Jonas}
\VS{2}Arise,
go to
Nineveh, that
great
city, and
cry against it; for their
wickedness is come
up
before me.
\VS{3}But
Jonah rose
up to
flee unto
Tarshish from the
presence of the
{\ND{
LORD}}, and went
down to
Joppa; and he
found a
ship
going to
Tarshish: so he
paid the
fare thereof, and went
down into it, to
go with them unto
Tarshish from the
presence of the
{\ND{
LORD}}.
\par }{\PP \VS{4}¶ But the
{\ND{
LORD}} sent
out a
great
wind into the
sea, and there was a
mighty
tempest in the
sea, so that the
ship was
like to be
broken.\FTNT{†}{{\FR 1.4 }
sent out: Heb. cast forth}\FTNT{‡}{{\FR 1.4 }
was like…: Heb. thought to be broken}
\VS{5}Then the
mariners were
afraid, and
cried every
man unto his
god, and cast
forth the
wares that
{\ADD{were}} in the
ship into the
sea, to
lighten
{\ADD{it}} of them. But
Jonah was gone
down into the
sides of the
ship; and he
lay, and was fast
asleep.
\VS{6}So the
shipmaster
came to him, and
said unto him, What meanest thou, O
sleeper?
arise,
call upon thy
God, if so be that
God will
think upon us, that we
perish not.
\VS{7}And they
said every
one to his
fellow,
Come, and let us
cast
lots, that we may
know for whose
cause this
evil
{\ADD{is}} upon us. So they
cast
lots, and the
lot
fell upon
Jonah.
\VS{8}Then
said they unto him,
Tell us, we pray thee, for whose cause this
evil
{\ADD{is}} upon us; What
{\ADD{is}} thine
occupation? and
whence
comest thou? what
{\ADD{is}} thy
country? and of what
people
{\ADD{art}} thou?
\VS{9}And he
said unto them, I
{\ADD{am}} an
Hebrew; and I
fear the
{\ND{
LORD}}, the
God of
heaven, which hath
made the
sea and the
dry
{\ADD{land}}.\FTNT{§}{{\FR 1.9 }
the
{\ND{LORD}}: or,
{\ND{JEHOVAH}}}
\VS{10}Then were the
men
exceedingly
afraid, and
said unto him, Why hast thou
done this? For the
men
knew that he
fled from the
presence of the
{\ND{
LORD}}, because he had
told them.\FTNT{**}{{\FR 1.10 }
exceedingly…: Heb. with great fear}
\par }{\PP \VS{11}¶ Then
said they unto him, What shall we
do unto thee, that the
sea may be
calm unto us? for the
sea
wrought, and was
tempestuous.\FTNT{††}{{\FR 1.11 }
may…: Heb. may be silent from us}\FTNT{‡‡}{{\FR 1.11 }
wrought…: or, grew more and more tempestuous: Heb. went}
\VS{12}And he
said unto them, Take me
up, and cast me
forth into the
sea; so shall the
sea be
calm unto you: for I
know that for my
sake this
great
tempest
{\ADD{is}} upon you.
\VS{13}Nevertheless the
men rowed
hard to
bring
{\ADD{it}} to the
land; but they
could not: for the
sea
wrought, and was
tempestuous against them.\FTNT{§§}{{\FR 1.13 }
rowed: Heb. digged}\FTNT{***}{{\FR 1.13 }
wrought…: or, grew more and more tempestuous: Heb. went}
\VS{14}Wherefore they
cried unto the
{\ND{
LORD}}, and
said, We beseech
thee, O
{\ND{
LORD}}, we beseech thee, let us not
perish for this
man’s
life, and
lay not upon us
innocent
blood: for thou, O
{\ND{
LORD}}, hast
done as it
pleased thee.
\VS{15}So they took
up
Jonah, and cast him
forth into the
sea: and the
sea
ceased from her
raging.\FTNT{†††}{{\FR 1.15 }
ceased: Heb. stood}
\VS{16}Then the
men
feared the
{\ND{
LORD}}
exceedingly, and
offered a
sacrifice unto the
{\ND{
LORD}}, and
made
vows.\FTNT{‡‡‡}{{\FR 1.16 }
offered…: Heb. sacrifice unto the
{\ND{LORD}}, and vowed vows}
\par }{\PP \VS{17}¶ Now the
{\ND{
LORD}} had
prepared a
great
fish to swallow
up
Jonah. And
Jonah was in the
belly of the
fish
three
days and
three
nights.\FTNT{§§§}{{\FR 1.17 }
belly: Heb. bowels}

\par }\Chap{2}{\PP \VerseOne{1}Then
Jonah
prayed unto the
{\ND{
LORD}} his
God out of the
fish’s
belly,
\VS{2}And
said, I
cried by reason of mine
affliction unto the
{\ND{
LORD}}, and he
heard me; out of the
belly of
hell
cried I,
{\ADD{and}} thou
heardest my
voice.\FTNT{*}{{\FR 2.2 }
by…: or, out of mine affliction}\FTNT{†}{{\FR 2.2 }
hell: or, the grave}
\VS{3}For thou hadst
cast me into the
deep, in the
midst of the
seas; and the
floods
compassed me about: all thy
billows and thy
waves passed
over me.\FTNT{‡}{{\FR 2.3 }
midst: Heb. heart}
\VS{4}Then I
said, I am cast
out of thy
sight; yet I will
look
again toward thy
holy
temple.
\VS{5}The
waters compassed me
about,
{\ADD{even}} to the
soul: the
depth closed me round
about, the
weeds were wrapped
about my
head.
\VS{6}I went
down to the
bottoms of the
mountains; the
earth with her
bars
{\ADD{was}} about me for
ever: yet hast thou brought
up my
life from
corruption, O
{\ND{
LORD}} my
God.\FTNT{§}{{\FR 2.6 }
bottoms: Heb. cuttings off}\FTNT{**}{{\FR 2.6 }
corruption: or, the pit}
\VS{7}When my
soul
fainted within me I
remembered the
{\ND{
LORD}}: and my
prayer came
in unto thee, into thine
holy
temple.
\VS{8}They that
observe
lying
vanities
forsake their own
mercy.
\VS{9}But I will
sacrifice unto thee with the
voice of
thanksgiving; I will
pay
{\ADD{that}} that I have
vowed.
Salvation
{\ADD{is}} of the
{\ND{
LORD}}.
\par }{\PP \VS{10}¶ And the
{\ND{
LORD}}
spake unto the
fish, and it vomited
out
Jonah upon the
dry
{\ADD{land}}.

\par }\Chap{3}{\PP \VerseOne{1}And the
word of the
{\ND{
LORD}} came unto
Jonah the
second time,
saying,
\VS{2}Arise,
go unto
Nineveh, that
great
city, and
preach unto it the
preaching that I
bid thee.
\VS{3}So
Jonah
arose, and
went unto
Nineveh, according to the
word of the
{\ND{
LORD}}. Now
Nineveh was an
exceeding
great
city of
three
days’
journey.\FTNT{*}{{\FR 3.3 }
exceeding: Heb. of God}
\VS{4}And
Jonah
began to
enter into the
city
a
day’s
journey, and he
cried, and
said, Yet
forty
days, and
Nineveh shall be
overthrown.
\par }{\PP \VS{5}¶ So the
people of
Nineveh
believed
God, and
proclaimed a
fast, and
put on
sackcloth, from the
greatest of them even to the
least of them.
\VS{6}For
word
came unto the
king of
Nineveh, and he
arose from his
throne, and he
laid his
robe from him, and
covered
{\ADD{him}} with
sackcloth, and
sat in
ashes.
\VS{7}And he caused
{\ADD{it}} to be
proclaimed and
published through
Nineveh by the
decree of the
king and his
nobles,
saying, Let neither
man nor
beast,
herd nor
flock,
taste any
thing: let them not
feed, nor
drink
water:\FTNT{†}{{\FR 3.7 }
published: Heb. said}\FTNT{‡}{{\FR 3.7 }
nobles: Heb. great men}
\VS{8}But let
man and
beast be
covered with
sackcloth, and
cry
mightily unto
God: yea, let them
turn every
one from his
evil
way, and from the
violence that
{\ADD{is}} in their
hands.
\VS{9}Who can
tell
{\ADD{if}}
God will
turn and
repent, and turn
away from his
fierce
anger, that we
perish not?
\par }{\PP \VS{10}¶ And
God
saw their
works, that they
turned from their
evil
way; and
God
repented of the
evil, that he had
said that he would
do unto them; and he
did
{\ADD{it}} not.

\par }\Chap{4}{\PP \VerseOne{1}But it
displeased
Jonah
exceedingly, and he was very
angry.
\VS{2}And he
prayed unto the
{\ND{
LORD}}, and
said, I
pray thee, O
{\ND{
LORD}},
{\ADD{was}} not this my
saying, when I was yet in my
country? Therefore I
fled
before unto
Tarshish: for I
knew that thou
{\ADD{art}} a
gracious
God, and
merciful,
slow to
anger, and of
great
kindness, and
repentest thee of the
evil.
\VS{3}Therefore now, O
{\ND{
LORD}},
take, I beseech thee, my
life from me; for
{\ADD{it is}}
better for me to
die than to
live.
\par }{\PP \VS{4}¶ Then
said the
{\ND{
LORD}}, Doest thou
well to be
angry?\FTNT{*}{{\FR 4.4 }
Doest…: or, Art thou greatly angry?}
\par }{\PP \VS{5}So
Jonah went
out of the
city, and
sat on the east
side of the
city, and there
made him a
booth, and
sat under it in the
shadow, till he might
see what would become of the
city.
\VS{6}And the
{\ND{
LORD}}
God
prepared a
gourd, and made
{\ADD{it}} to come
up over
Jonah, that it might be a
shadow over his
head, to
deliver him from his
grief. So
Jonah was
exceeding
glad of the
gourd.\FTNT{†}{{\FR 4.6 }
gourd: or, palmcrist: Heb. Kikajon}\FTNT{‡}{{\FR 4.6 }
was…: Heb. rejoiced with great joy}
\VS{7}But
God
prepared a
worm when the
morning
rose the next
day, and it
smote the
gourd that it
withered.
\VS{8}And it came to pass, when the
sun did
arise, that
God
prepared a
vehement
east
wind; and the
sun
beat upon the
head of
Jonah, that he
fainted, and
wished in
himself to
die, and
said,
{\ADD{It is}}
better for me to
die than to
live.\FTNT{§}{{\FR 4.8 }
vehement: or, silent}
\VS{9}And
God
said to
Jonah, Doest thou
well to be
angry for the
gourd? And he
said, I do
well to be
angry,
{\ADD{even}} unto
death.\FTNT{**}{{\FR 4.9 }
Doest…: or, Art thou greatly angry?}\FTNT{††}{{\FR 4.9 }
I do well…: or, I am greatly angry}
\VS{10}Then
said the
{\ND{
LORD}}, Thou hast had
pity on the
gourd, for the which thou hast not
laboured, neither madest it
grow; which came up in a
night, and
perished in a
night:\FTNT{‡‡}{{\FR 4.10 }
had pity: or, spared}\FTNT{§§}{{\FR 4.10 }
came…: Heb. was the son of the night}
\VS{11}And should not I
spare
Nineveh, that
great
city, wherein
are
more than
sixscore
thousand
persons that cannot
discern between their right
hand and their left
hand; and
{\ADD{also}}
much
cattle?
\par }