\NormalFont\ShortTitle{Joshua}
{\MT The Book of Joshua

\par }\ChapOne{1}{\PP \VerseOne{1}Now
after the
death of
Moses the
servant of the
{\ND{
LORD}} it came to pass, that the
{\ND{
LORD}}
spake unto
Joshua the
son of
Nun,
Moses’
minister,
saying,
\VS{2}Moses my
servant is
dead; now therefore
arise, go
over this
Jordan, thou, and all this
people, unto the
land which I do
give to them,
{\ADD{even}} to the
children of
Israel.
\VS{3}Every
place that the
sole of your
foot shall tread
upon, that have I
given unto you, as I
said unto
Moses.
\VS{4}From the
wilderness and this
Lebanon even unto the
great
river, the
river
Euphrates, all the
land of the
Hittites, and unto the
great
sea toward the going
down of the
sun, shall be your
coast.
\VS{5}There shall not any
man be able to
stand
before thee all the
days of thy
life: as I was with
Moses,
{\ADD{so}} I will be with thee: I will not
fail thee, nor
forsake thee.
\VS{6}Be
strong and of a good
courage: for unto this
people shalt thou divide for an
inheritance the
land, which I
sware unto their
fathers to
give them.\FTNT{*}{{\FR 1.6 }
unto this…: or, thou shalt cause this people to inherit the land, etc}
\VS{7}Only be thou
strong and
very
courageous, that thou mayest
observe to
do according to all the
law, which
Moses my
servant
commanded thee:
turn not from it
{\ADD{to}} the right
hand or
{\ADD{to}} the
left, that thou mayest
prosper whithersoever thou
goest.\FTNT{†}{{\FR 1.7 }
prosper: or, do wisely}
\VS{8}This
book of the
law shall not
depart out of thy
mouth; but thou shalt
meditate therein
day and
night, that thou mayest
observe to
do according to all that is
written therein: for then thou shalt
make thy
way
prosperous, and then thou shalt have good
success.\FTNT{‡}{{\FR 1.8 }
have…: or, do wisely}
\VS{9}Have not I
commanded thee? Be
strong and of a good
courage; be not
afraid, neither be thou
dismayed: for the
{\ND{
LORD}} thy
God
{\ADD{is}} with thee whithersoever thou
goest.
\par }{\PP \VS{10}¶ Then
Joshua
commanded the
officers of the
people,
saying,
\VS{11}Pass
through the
host, and
command the
people,
saying,
Prepare you
victuals; for within
three
days ye shall pass
over this
Jordan, to go
in to
possess the
land, which the
{\ND{
LORD}} your
God
giveth you to
possess it.
\par }{\PP \VS{12}¶ And to the
Reubenites, and to the
Gadites, and to
half the
tribe of
Manasseh,
spake
Joshua,
saying,
\VS{13}Remember the
word which
Moses the
servant of the
{\ND{
LORD}}
commanded you,
saying, The
{\ND{
LORD}} your
God hath given you
rest, and hath
given you this
land.
\VS{14}Your
wives, your little
ones, and your
cattle, shall
remain in the
land which
Moses
gave you on this
side
Jordan; but ye shall
pass
before your
brethren
armed, all the
mighty men of
valour, and
help them;\FTNT{§}{{\FR 1.14 }
armed: Heb. marshalled by five}
\VS{15}Until the
{\ND{
LORD}} have
given your
brethren
rest, as
{\ADD{he hath given}} you, and they also have
possessed the
land which the
{\ND{
LORD}} your
God
giveth them: then ye shall
return unto the
land of your
possession, and
enjoy it, which
Moses the
{\ND{
LORD’s}}
servant
gave you on this
side
Jordan toward the
sunrising.
\par }{\PP \VS{16}¶ And they
answered
Joshua,
saying, All that thou
commandest us we will
do, and
whithersoever thou
sendest us, we will
go.
\VS{17}According as we
hearkened unto
Moses in all things, so will we
hearken unto thee: only the
{\ND{
LORD}} thy
God be with thee, as he was with
Moses.
\VS{18}Whosoever
{\ADD{he be}} that doth
rebel against thy
commandment, and will not
hearken unto thy
words in all that thou
commandest him, he shall be put to
death: only be
strong and of a good
courage.

\par }\Chap{2}{\PP \VerseOne{1}And
Joshua the
son of
Nun
sent out of
Shittim
two
men to
spy
secretly,
saying,
Go
view the
land, even
Jericho. And they
went, and
came into an
harlot’s
house,
named
Rahab, and
lodged there.\FTNT{*}{{\FR 2.1 }
sent: or, had sent}\FTNT{†}{{\FR 2.1 }
lodged: Heb. lay}
\VS{2}And it was
told the
king of
Jericho,
saying, Behold, there
came
men in hither to
night of the
children of
Israel to search
out the
country.
\VS{3}And the
king of
Jericho
sent unto
Rahab,
saying, Bring
forth the
men that are
come to thee, which are
entered into thine
house: for they be
come to search
out all the
country.
\VS{4}And the
woman
took the
two
men, and
hid them, and
said thus, There
came
men unto me, but I
wist not
whence they
{\ADD{were:}}
\VS{5}And it came to pass
{\ADD{about the time}} of
shutting of the
gate, when it was
dark, that the
men went
out: whither the
men
went I
wot not:
pursue
after them
quickly; for ye shall
overtake them.
\VS{6}But she had brought them
up to the
roof of the house, and
hid them with the
stalks of
flax, which she had laid in
order upon the
roof.
\VS{7}And the
men
pursued
after them the
way to
Jordan unto the
fords: and as soon as they which
pursued
after them were gone
out, they
shut the
gate.
\par }{\PP \VS{8}¶ And before they were laid
down, she came
up unto them upon the
roof;
\VS{9}And she
said unto the
men, I
know that the
{\ND{
LORD}} hath
given you the
land, and that your
terror is
fallen upon us, and that all the
inhabitants of the
land
faint
because of you.\FTNT{‡}{{\FR 2.9 }
faint: Heb. melt}
\VS{10}For we have
heard how the
{\ND{
LORD}} dried
up the
water of the
Red
sea for
you, when ye came
out of
Egypt; and what ye
did unto the
two
kings of the
Amorites, that
{\ADD{were}} on the other
side
Jordan,
Sihon and
Og, whom ye utterly
destroyed.
\VS{11}And as soon as we had
heard
{\ADD{these things}}, our
hearts did
melt, neither did there
remain any more
courage in any
man,
because of you: for the
{\ND{
LORD}} your
God, he
{\ADD{is}}
God in
heaven
above, and in
earth beneath.\FTNT{§}{{\FR 2.11 }
did there…: Heb. rose up}
\VS{12}Now therefore, I pray you,
swear unto me by the
{\ND{
LORD}},
since I have
shewed you
kindness, that ye will also
shew
kindness unto my
father’s
house, and
give me a
true
token:
\VS{13}And
{\ADD{that}} ye will save
alive my
father, and my
mother, and my
brethren, and my
sisters, and all that they have, and
deliver our
lives from
death.
\VS{14}And the
men
answered her, Our
life for
yours, if ye
utter not this our
business. And it shall be, when the
{\ND{
LORD}} hath
given us the
land, that we will
deal
kindly and
truly with thee.\FTNT{**}{{\FR 2.14 }
for…: Heb. instead of you to die}
\VS{15}Then she let them
down by a
cord
through the
window: for her
house
{\ADD{was}} upon the
town
wall, and she
dwelt upon the
wall.
\VS{16}And she
said unto them,
Get you to the
mountain, lest the
pursuers
meet you; and
hide yourselves there
three
days, until the
pursuers be
returned: and
afterward may ye
go your
way.
\VS{17}And the
men
said unto her, We
{\ADD{will be}}
blameless of this thine
oath which thou hast made us
swear.
\VS{18}Behold,
{\ADD{when}} we
come into the
land, thou shalt
bind this
line of
scarlet
thread in the
window which thou didst let us down
by: and thou shalt
bring thy
father, and thy
mother, and thy
brethren, and all thy
father’s
household,
home unto thee.\FTNT{††}{{\FR 2.18 }
bring: Heb. gather}
\VS{19}And it shall be,
{\ADD{that}}
whosoever shall go
out of the
doors of thy
house into the
street, his
blood
{\ADD{shall be}} upon his
head, and we
{\ADD{will be}}
guiltless: and whosoever shall be with thee in the
house, his
blood
{\ADD{shall be}} on our
head, if
{\ADD{any}}
hand be upon him.
\VS{20}And if thou
utter this our
business, then we will be
quit of thine
oath which thou hast made us to
swear.
\VS{21}And she
said, According unto your
words, so
{\ADD{be}} it. And she sent them
away, and they
departed: and she
bound the
scarlet
line in the
window.
\par }{\PP \VS{22}And they
went, and
came unto the
mountain, and
abode there
three
days, until the
pursuers were
returned: and the
pursuers
sought
{\ADD{them}} throughout all the
way, but
found
{\ADD{them}} not.
\par }{\PP \VS{23}¶ So the
two
men
returned, and
descended from the
mountain, and passed
over, and
came to
Joshua the
son of
Nun, and
told him all
{\ADD{things}} that
befell them:
\VS{24}And they
said unto
Joshua,
Truly the
{\ND{
LORD}} hath
delivered into our
hands all the
land; for even all the
inhabitants of the
country do
faint
because of us.\FTNT{‡‡}{{\FR 2.24 }
faint: Heb. melt}

\par }\Chap{3}{\PP \VerseOne{1}And
Joshua rose
early in the
morning; and they
removed from
Shittim, and
came to
Jordan, he and all the
children of
Israel, and
lodged there before they passed
over.
\VS{2}And it came to pass
after
three
days, that the
officers
went
through the
host;
\VS{3}And they
commanded the
people,
saying, When ye
see the
ark of the
covenant of the
{\ND{
LORD}} your
God, and the
priests the
Levites
bearing it, then ye shall
remove from your
place, and
go
after it.
\VS{4}Yet there shall be a
space between you and it, about two
thousand
cubits by
measure: come not
near unto it, that ye may
know the
way by which ye must
go: for ye have not
passed
{\ADD{this}}
way
heretofore.\FTNT{*}{{\FR 3.4 }
heretofore: Heb. since yesterday, and the third day}
\VS{5}And
Joshua
said unto the
people,
Sanctify yourselves: for to
morrow the
{\ND{
LORD}} will
do
wonders
among you.
\VS{6}And
Joshua
spake unto the
priests,
saying, Take
up the
ark of the
covenant, and pass
over
before the
people. And they took
up the
ark of the
covenant, and
went
before the
people.
\par }{\PP \VS{7}¶ And the
{\ND{
LORD}}
said unto
Joshua, This
day will I
begin to
magnify thee in the
sight of all
Israel, that they may
know that, as I was with
Moses,
{\ADD{so}} I will be with thee.
\VS{8}And thou shalt
command the
priests that
bear the
ark of the
covenant,
saying, When ye are
come to the
brink of the
water of
Jordan, ye shall stand
still in
Jordan.
\par }{\PP \VS{9}¶ And
Joshua
said unto the
children of
Israel, Come
hither, and
hear the
words of the
{\ND{
LORD}} your
God.
\VS{10}And
Joshua
said, Hereby ye shall
know that the
living
God
{\ADD{is}}
among you, and
{\ADD{that}} he will without
fail drive
out from
before you the
Canaanites, and the
Hittites, and the
Hivites, and the
Perizzites, and the
Girgashites, and the
Amorites, and the
Jebusites.
\VS{11}Behold, the
ark of the
covenant of the
Lord of all the
earth passeth
over
before you into
Jordan.
\VS{12}Now therefore
take you
twelve
men out of the
tribes of
Israel, out of
every
tribe a
man.
\VS{13}And it shall come to pass, as soon as the
soles of the
feet of the
priests that
bear the
ark of the
{\ND{
LORD}}, the
Lord of all the
earth, shall
rest in the
waters of
Jordan,
{\ADD{that}} the
waters of
Jordan shall be cut
off
{\ADD{from}} the
waters that come
down from
above; and they shall
stand upon
an
heap.
\par }{\PP \VS{14}¶ And it came to pass, when the
people
removed from their
tents, to pass
over
Jordan, and the
priests
bearing the
ark of the
covenant
before the
people;
\VS{15}And as they that
bare the
ark were
come unto
Jordan, and the
feet of the
priests that
bare the
ark were
dipped in the
brim of the
water, (for
Jordan
overfloweth all his
banks all the
time of
harvest,)
\VS{16}That the
waters which came
down from
above
stood
{\ADD{and}} rose
up upon
an
heap
very
far from the
city
Adam, that
{\ADD{is}}
beside
Zaretan: and those that came
down toward the
sea of the
plain,
{\ADD{even}} the
salt
sea,
failed,
{\ADD{and}} were cut
off: and the
people passed
over right against
Jericho.
\VS{17}And the
priests that
bare the
ark of the
covenant of the
{\ND{
LORD}}
stood
firm on dry
ground in the
midst of
Jordan, and all the
Israelites passed
over on dry
ground, until all the
people were
passed
clean
over
Jordan.

\par }\Chap{4}{\PP \VerseOne{1}And it came to pass, when all the
people were
clean passed
over
Jordan, that the
{\ND{
LORD}}
spake unto
Joshua,
saying,
\VS{2}Take you
twelve
men out of the
people, out of
every
tribe a
man,
\VS{3}And
command ye them,
saying,
Take you hence out of the
midst of
Jordan, out of the place where the
priests’
feet
stood
firm,
twelve
stones, and ye shall carry them
over with you, and
leave them in the lodging
place, where ye shall
lodge this
night.
\VS{4}Then
Joshua
called the
twelve
men, whom he had
prepared of the
children of
Israel, out of
every
tribe a
man:
\VS{5}And
Joshua
said unto them, Pass
over
before the
ark of the
{\ND{
LORD}} your
God into the
midst of
Jordan, and take ye
up
every
man of you a
stone upon his
shoulder, according unto the
number of the
tribes of the
children of
Israel:
\VS{6}That this may be a
sign
among you,
{\ADD{that}} when your
children
ask
{\ADD{their fathers}} in time to
come,
saying, What
{\ADD{mean}} ye by these
stones?\FTNT{*}{{\FR 4.6 }
in time…: Heb. to morrow}
\VS{7}Then ye shall
answer them, That the
waters of
Jordan were cut
off
before the
ark of the
covenant of the
{\ND{
LORD}}; when it passed
over
Jordan, the
waters of
Jordan were cut
off: and these
stones shall be for a
memorial unto the
children of
Israel
for
ever.
\VS{8}And the
children of
Israel
did so as
Joshua
commanded, and took
up
twelve
stones out of the
midst of
Jordan, as the
{\ND{
LORD}}
spake unto
Joshua, according to the
number of the
tribes of the
children of
Israel, and carried them
over with them unto the place where they
lodged, and laid them
down there.
\VS{9}And
Joshua set
up
twelve
stones in the
midst of
Jordan, in the place where the
feet of the
priests which
bare the
ark of the
covenant
stood: and they are there unto this
day.
\par }{\PP \VS{10}¶ For the
priests which
bare the
ark
stood in the
midst of
Jordan, until every
thing was
finished that the
{\ND{
LORD}}
commanded
Joshua to
speak unto the
people, according to all that
Moses
commanded
Joshua: and the
people
hasted and passed
over.
\VS{11}And it came to pass, when all the
people were
clean passed
over, that the
ark of the
{\ND{
LORD}} passed
over, and the
priests, in the
presence of the
people.
\VS{12}And the
children of
Reuben, and the
children of
Gad, and
half the
tribe of
Manasseh, passed
over
armed
before the
children of
Israel, as
Moses
spake unto them:
\VS{13}About
forty
thousand
prepared for
war passed
over
before the
{\ND{
LORD}} unto
battle, to the
plains of
Jericho.\FTNT{†}{{\FR 4.13 }
prepared: or, ready armed}
\par }{\PP \VS{14}¶ On that
day the
{\ND{
LORD}}
magnified
Joshua in the
sight of all
Israel; and they
feared him, as they
feared
Moses, all the
days of his
life.
\VS{15}And the
{\ND{
LORD}}
spake unto
Joshua,
saying,
\VS{16}Command the
priests that
bear the
ark of the
testimony, that they come
up out of
Jordan.
\VS{17}Joshua therefore
commanded the
priests,
saying, Come ye
up out of
Jordan.
\VS{18}And it came to pass, when the
priests that
bare the
ark of the
covenant of the
{\ND{
LORD}} were come
up out of the
midst of
Jordan,
{\ADD{and}} the
soles of the
priests’
feet were lifted
up unto the dry
land, that the
waters of
Jordan
returned unto their
place, and
flowed over all his
banks,
as
{\ADD{they did}}
before.\FTNT{‡}{{\FR 4.18 }
lifted…: Heb. plucked up}\FTNT{§}{{\FR 4.18 }
flowed: Heb. went}
\par }{\PP \VS{19}¶ And the
people came
up out of
Jordan on the
tenth
{\ADD{day}} of the
first
month, and
encamped in
Gilgal, in the
east
border of
Jericho.
\par }{\PP \VS{20}And those
twelve
stones, which they took
out of
Jordan, did
Joshua
pitch in
Gilgal.
\VS{21}And he
spake unto the
children of
Israel,
saying, When your
children shall
ask their
fathers in time to
come,
saying, What
{\ADD{mean}} these
stones?\FTNT{**}{{\FR 4.21 }
in time…: Heb. to morrow}
\VS{22}Then ye shall let your
children
know,
saying,
Israel came
over this
Jordan on dry
land.
\VS{23}For the
{\ND{
LORD}} your
God dried
up the
waters of
Jordan from
before you, until ye were passed
over, as the
{\ND{
LORD}} your
God
did to the
Red
sea, which he dried
up from
before us, until we were gone
over:
\VS{24}That all the
people of the
earth might
know the
hand of the
{\ND{
LORD}}, that it
{\ADD{is}}
mighty: that ye might
fear the
{\ND{
LORD}} your
God for
ever.\FTNT{††}{{\FR 4.24 }
for ever: Heb. all days}

\par }\Chap{5}{\PP \VerseOne{1}And it came to pass, when all the
kings of the
Amorites, which
{\ADD{were}} on the
side of
Jordan
westward, and all the
kings of the
Canaanites, which
{\ADD{were}} by the
sea,
heard that the
{\ND{
LORD}} had dried
up the
waters of
Jordan from
before the
children of
Israel, until we were passed
over, that their
heart
melted, neither was there
spirit in them any more,
because of the
children of
Israel.
\par }{\PP \VS{2}¶ At that
time the
{\ND{
LORD}}
said unto
Joshua,
Make thee
sharp
knives, and
circumcise
again the
children of
Israel the second
time.\FTNT{*}{{\FR 5.2 }
sharp…: or, knives of flints}
\VS{3}And
Joshua
made him
sharp
knives, and
circumcised the
children of
Israel at the
hill of the
foreskins.\FTNT{†}{{\FR 5.3 }
sharp…: or, knives of flints}\FTNT{‡}{{\FR 5.3 }
the hill…: or, Gibeah-haaraloth}
\VS{4}And this
{\ADD{is}} the
cause why
Joshua did
circumcise: All the
people that came
out of
Egypt,
{\ADD{that were}}
males,
{\ADD{even}} all the
men of
war,
died in the
wilderness by the
way, after they came
out of
Egypt.
\VS{5}Now all the
people that came
out were
circumcised: but all the
people
{\ADD{that were}}
born in the
wilderness by the
way as they came
forth out of
Egypt,
{\ADD{them}} they had not
circumcised.
\VS{6}For the
children of
Israel
walked
forty
years in the
wilderness, till all the
people
{\ADD{that were}}
men of
war, which came
out of
Egypt, were
consumed, because they
obeyed not the
voice of the
{\ND{
LORD}}: unto whom the
{\ND{
LORD}}
sware that he would not
shew them the
land, which the
{\ND{
LORD}}
sware unto their
fathers that he would
give us, a
land that
floweth with
milk and
honey.
\VS{7}And their
children,
{\ADD{whom}} he raised
up in their stead, them
Joshua
circumcised: for they were
uncircumcised, because they had not
circumcised them by the
way.
\VS{8}And it came to pass, when they had
done
circumcising all the
people, that they
abode in their places in the
camp, till they were
whole.\FTNT{§}{{\FR 5.8 }
they had…: Heb. the people had made an end to be circumcised}
\VS{9}And the
{\ND{
LORD}}
said unto
Joshua, This
day have I rolled
away the
reproach of
Egypt from off you. Wherefore the
name of the
place is
called
Gilgal unto this
day.\FTNT{**}{{\FR 5.9 }
Gilgal: that is Rolling}
\par }{\PP \VS{10}¶ And the
children of
Israel
encamped in
Gilgal, and
kept the
passover on the
fourteenth
day of the
month at
even in the
plains of
Jericho.
\VS{11}And they did
eat of the old
corn of the
land on the
morrow after the
passover, unleavened
cakes, and
parched
{\ADD{corn}} in the
selfsame
day.
\par }{\PP \VS{12}¶ And the
manna
ceased on the
morrow after they had
eaten of the old
corn of the
land; neither had the
children of
Israel
manna any more; but they did
eat of the
fruit of the
land of
Canaan that
year.
\par }{\PP \VS{13}¶ And it came to pass, when
Joshua was by
Jericho, that he lifted
up his
eyes and
looked, and, behold, there
stood a
man over against him with his
sword
drawn in his
hand: and
Joshua
went unto him, and
said unto him,
{\ADD{Art}} thou for us, or for our
adversaries?
\VS{14}And he
said, Nay; but
{\ADD{as}}
captain of the
host of the
{\ND{
LORD}} am I now
come. And
Joshua
fell on his
face to the
earth, and did
worship, and
said unto him, What
saith my
lord unto his
servant?\FTNT{††}{{\FR 5.14 }
captain: or, prince}
\VS{15}And the
captain of the
{\ND{
LORD’s}}
host
said unto
Joshua,
Loose thy
shoe from off thy
foot; for the
place whereon thou
standest
{\ADD{is}}
holy. And
Joshua
did so.

\par }\Chap{6}{\PP \VerseOne{1}Now
Jericho was
straitly shut
up
because of the
children of
Israel: none went
out, and none came
in.\FTNT{*}{{\FR 6.1 }
was…: Heb. did shut up, and was shut up}
\VS{2}And the
{\ND{
LORD}}
said unto
Joshua,
See, I have
given into thine
hand
Jericho, and the
king thereof,
{\ADD{and}} the mighty
men of
valour.
\VS{3}And ye shall
compass the
city, all
{\ADD{ye}}
men of
war,
{\ADD{and}} go round
about the
city
once. Thus shalt thou
do
six
days.
\VS{4}And
seven
priests shall
bear
before the
ark
seven
trumpets of
rams’ horns: and the
seventh
day ye shall
compass the
city
seven
times, and the
priests shall
blow with the
trumpets.
\VS{5}And it shall come to pass, that when they make a
long
{\ADD{blast}} with the
ram’s
horn,
{\ADD{and}} when ye
hear the
sound of the
trumpet, all the
people shall
shout with a
great
shout; and the
wall of the
city shall fall
down
flat, and the
people shall ascend
up every
man straight before him.\FTNT{†}{{\FR 6.5 }
flat: Heb. under it}
\par }{\PP \VS{6}¶ And
Joshua the
son of
Nun
called the
priests, and
said unto them, Take
up the
ark of the
covenant, and let
seven
priests
bear
seven
trumpets of
rams’ horns
before the
ark of the
{\ND{
LORD}}.
\VS{7}And he
said unto the
people, Pass
on, and
compass the
city, and let him that is
armed pass
on
before the
ark of the
{\ND{
LORD}}.
\par }{\PP \VS{8}¶ And it came to pass, when
Joshua had
spoken unto the
people, that the
seven
priests
bearing the
seven
trumpets of
rams’ horns passed
on
before the
{\ND{
LORD}}, and
blew with the
trumpets: and the
ark of the
covenant of the
{\ND{
LORD}}
followed
them.
\par }{\PP \VS{9}¶ And the
armed men
went
before the
priests that
blew with the
trumpets, and the
rereward
came
after the
ark,
{\ADD{the priests}} going
on, and
blowing with the
trumpets.\FTNT{‡}{{\FR 6.9 }
rereward: Heb. gathering host}
\VS{10}And
Joshua had
commanded the
people,
saying, Ye shall not
shout, nor make any
noise with your
voice, neither shall
{\ADD{any}}
word
proceed out of your
mouth, until the
day I
bid you
shout; then shall ye
shout.\FTNT{§}{{\FR 6.10 }
any noise…: Heb. your voice to be heard}
\VS{11}So the
ark of the
{\ND{
LORD}}
compassed the
city, going
about
{\ADD{it}}
once: and they
came into the
camp, and
lodged in the
camp.
\par }{\PP \VS{12}¶ And
Joshua rose
early in the
morning, and the
priests took
up the
ark of the
{\ND{
LORD}}.
\VS{13}And
seven
priests
bearing
seven
trumpets of
rams’ horns
before the
ark of the
{\ND{
LORD}} went
on
continually, and
blew with the
trumpets: and the armed
men
went
before them; but the
rereward
came
after the
ark of the
{\ND{
LORD}},
{\ADD{the priests}} going
on, and
blowing with the
trumpets.
\VS{14}And the
second
day they
compassed the
city
once, and
returned into the
camp: so they
did
six
days.
\VS{15}And it came to pass on the
seventh
day, that they rose
early about the
dawning of the
day, and
compassed the
city after the same
manner
seven
times: only on that
day they
compassed the
city
seven
times.
\VS{16}And it came to pass at the
seventh
time, when the
priests
blew with the
trumpets,
Joshua
said unto the
people,
Shout; for the
{\ND{
LORD}} hath
given you the
city.
\par }{\PP \VS{17}¶ And the
city shall be
accursed,
{\ADD{even}} it, and all that
{\ADD{are}} therein, to the
{\ND{
LORD}}: only
Rahab the
harlot shall
live, she and all that
{\ADD{are}} with her in the
house, because she
hid the
messengers that we
sent.\FTNT{**}{{\FR 6.17 }
accursed: or, devoted}
\VS{18}And ye, in any
wise
keep
{\ADD{yourselves}} from the accursed
thing, lest ye make
{\ADD{yourselves}}
accursed, when ye
take of the accursed
thing, and
make the
camp of
Israel a
curse, and
trouble it.\FTNT{††}{{\FR 6.18 }
accursed: or, devoted}\FTNT{‡‡}{{\FR 6.18 }
a curse: or, devoted}
\VS{19}But all the
silver, and
gold, and
vessels of
brass and
iron,
{\ADD{are}}
consecrated unto the
{\ND{
LORD}}: they shall
come into the
treasury of the
{\ND{
LORD}}.\FTNT{§§}{{\FR 6.19 }
consecrated: Heb. holiness}
\VS{20}So the
people
shouted when
{\ADD{the priests}}
blew with the
trumpets: and it came to pass, when the
people
heard the
sound of the
trumpet, and the
people
shouted with a
great
shout, that the
wall fell down
flat, so that the
people went
up into the
city, every
man straight before him, and they
took the
city.\FTNT{***}{{\FR 6.20 }
flat: Heb. under it}
\VS{21}And they utterly
destroyed all that
{\ADD{was}} in the
city, both
man and
woman,
young and
old, and
ox, and
sheep, and
ass, with the
edge of the
sword.
\VS{22}But
Joshua had
said unto the
two
men that had spied
out the
country,
Go into the
harlot’s
house, and bring
out thence the
woman, and all that she hath, as ye
sware unto her.
\VS{23}And the young
men that were
spies went
in, and brought
out
Rahab, and her
father, and her
mother, and her
brethren, and all that she had; and they brought
out all her
kindred, and
left them
without the
camp of
Israel.\FTNT{†††}{{\FR 6.23 }
kindred: Heb. families}
\VS{24}And they
burnt the
city with
fire, and all that
{\ADD{was}} therein: only the
silver, and the
gold, and the
vessels of
brass and of
iron, they
put into the
treasury of the
house of the
{\ND{
LORD}}.
\VS{25}And
Joshua
saved
Rahab the
harlot
alive, and her
father’s
household, and all that she had; and she
dwelleth
in
Israel
{\ADD{even}} unto this
day; because she
hid the
messengers, which
Joshua
sent to spy
out
Jericho.
\par }{\PP \VS{26}¶ And
Joshua
adjured
{\ADD{them}} at that
time,
saying,
Cursed
{\ADD{be}} the
man
before the
{\ND{
LORD}}, that riseth
up and
buildeth this
city
Jericho: he shall lay the
foundation thereof in his
firstborn, and in his
youngest
{\ADD{son}} shall he set
up the
gates of it.
\VS{27}So the
{\ND{
LORD}} was with
Joshua; and his
fame was
{\ADD{noised}} throughout all the
country.

\par }\Chap{7}{\PP \VerseOne{1}But the
children of
Israel
committed a
trespass in the accursed
thing: for
Achan, the
son of
Carmi, the
son of
Zabdi, the
son of
Zerah, of the
tribe of
Judah,
took of the accursed
thing: and the
anger of the
{\ND{
LORD}} was
kindled against the
children of
Israel.\FTNT{*}{{\FR 7.1 }
Achan: also called, Achar}\FTNT{†}{{\FR 7.1 }
Zabdi: also called, Zimri}
\VS{2}And
Joshua
sent
men from
Jericho to
Ai, which
{\ADD{is}}
beside
Beth-aven, on the east
side of
Beth-el, and
spake unto them,
saying, Go
up and
view the
country. And the
men went
up and
viewed
Ai.
\VS{3}And they
returned to
Joshua, and
said unto him, Let not all the
people go
up; but let about two or
three
thousand
men go
up and
smite
Ai;
{\ADD{and}} make not all the
people to
labour thither; for they
{\ADD{are but}}
few.\FTNT{‡}{{\FR 7.3 }
about…: Heb. about two thousand men, or, about three thousand men}
\VS{4}So there went
up thither of the
people about
three
thousand
men: and they
fled
before the
men of
Ai.
\VS{5}And the
men of
Ai
smote of them about
thirty and
six
men: for they
chased them
{\ADD{from}}
before the
gate
{\ADD{even}} unto
Shebarim, and
smote them in the going
down: wherefore the
hearts of the
people
melted, and became as
water.\FTNT{§}{{\FR 7.5 }
in…: or, in Morad}
\par }{\PP \VS{6}¶ And
Joshua
rent his
clothes, and
fell to the
earth upon his
face
before the
ark of the
{\ND{
LORD}} until the
eventide, he and the
elders of
Israel, and
put
dust upon their
heads.
\VS{7}And
Joshua
said, Alas,
O
Lord
{\ND{
GOD}}, wherefore hast thou at
all
brought this
people
over
Jordan, to
deliver us into the
hand of the
Amorites, to
destroy us? would to
God we had been
content, and
dwelt on the other
side
Jordan!
\VS{8}O
Lord, what shall I
say,
when
Israel
turneth their
backs
before their
enemies!\FTNT{**}{{\FR 7.8 }
backs: Heb. necks}
\VS{9}For the
Canaanites and all the
inhabitants of the
land shall
hear
{\ADD{of it}}, and shall environ us
round, and cut
off our
name from the
earth: and what wilt thou
do unto thy
great
name?
\par }{\PP \VS{10}¶ And the
{\ND{
LORD}}
said unto
Joshua, Get thee
up; wherefore
liest thou thus upon thy
face?\FTNT{††}{{\FR 7.10 }
liest: Heb. fallest}
\VS{11}Israel hath
sinned, and they have also
transgressed my
covenant which I
commanded them: for they have even
taken of the accursed
thing, and have also
stolen, and
dissembled also, and they have
put
{\ADD{it}} even among their own
stuff.
\VS{12}Therefore the
children of
Israel
could not
stand
before their
enemies,
{\ADD{but}}
turned
{\ADD{their}}
backs
before their
enemies, because they were
accursed: neither will I be with you any
more,
except ye
destroy the
accursed from
among you.
\VS{13}Up,
sanctify the
people, and
say,
Sanctify yourselves against to
morrow: for thus
saith the
{\ND{
LORD}}
God of
Israel,
{\ADD{There is}} an accursed
thing in the
midst of thee, O
Israel: thou
canst not
stand
before thine
enemies, until ye take
away the accursed
thing from
among you.
\VS{14}In the
morning therefore ye shall be
brought according to your
tribes: and it shall be,
{\ADD{that}} the
tribe which the
{\ND{
LORD}}
taketh shall
come according to the
families
{\ADD{thereof;}} and the
family which the
{\ND{
LORD}} shall
take shall
come by
households; and the
household which the
{\ND{
LORD}} shall
take shall
come
man by
man.
\VS{15}And it shall be,
{\ADD{that}} he that is
taken with the accursed
thing shall be
burnt with
fire, he and all that he hath: because he hath
transgressed the
covenant of the
{\ND{
LORD}}, and because he hath
wrought
folly in
Israel.\FTNT{‡‡}{{\FR 7.15 }
folly: or, wickedness}
\par }{\PP \VS{16}¶ So
Joshua rose up
early in the
morning, and
brought
Israel by their
tribes; and the
tribe of
Judah was
taken:
\VS{17}And he
brought the
family of
Judah; and he
took the
family of the
Zarhites: and he
brought the
family of the
Zarhites
man by
man; and
Zabdi was
taken:
\VS{18}And he
brought his
household
man by
man; and
Achan, the
son of
Carmi, the
son of
Zabdi, the
son of
Zerah, of the
tribe of
Judah, was
taken.
\VS{19}And
Joshua
said unto
Achan, My
son,
give, I pray thee,
glory to the
{\ND{
LORD}}
God of
Israel, and
make
confession unto him; and
tell me now what thou hast
done;
hide
{\ADD{it}} not from me.
\VS{20}And
Achan
answered
Joshua, and
said,
Indeed I have
sinned against the
{\ND{
LORD}}
God of
Israel, and thus and thus have I
done:
\VS{21}When I
saw among the
spoils
a
goodly
Babylonish
garment, and two
hundred
shekels of
silver, and
a
wedge of
gold of
fifty
shekels
weight, then I
coveted them, and
took them; and, behold, they
{\ADD{are}}
hid in the
earth in the
midst of my
tent, and the
silver under it.\FTNT{§§}{{\FR 7.21 }
wedge: Heb. tongue}
\par }{\PP \VS{22}¶ So
Joshua
sent
messengers, and they
ran unto the
tent; and, behold,
{\ADD{it was}}
hid in his
tent, and the
silver under it.
\VS{23}And they
took them out of the
midst of the
tent, and
brought them unto
Joshua, and unto all the
children of
Israel, and laid them
out
before the
{\ND{
LORD}}.\FTNT{***}{{\FR 7.23 }
laid: Heb. poured}
\VS{24}And
Joshua, and all
Israel with him,
took
Achan the
son of
Zerah, and the
silver, and the
garment, and the
wedge of
gold, and his
sons, and his
daughters, and his
oxen, and his
asses, and his
sheep, and his
tent, and all that he had: and they
brought them unto the
valley of
Achor.
\VS{25}And
Joshua
said,
Why hast thou
troubled us? the
{\ND{
LORD}} shall
trouble thee this
day. And all
Israel
stoned him with
stones, and
burned them with
fire, after they had
stoned them with
stones.
\VS{26}And they
raised over him a
great
heap of
stones unto this
day. So the
{\ND{
LORD}}
turned from the
fierceness of his
anger. Wherefore the
name of that
place was
called, The
valley of
Achor, unto this
day.\FTNT{†††}{{\FR 7.26 }
Achor: that is, Trouble}

\par }\Chap{8}{\PP \VerseOne{1}And the
{\ND{
LORD}}
said unto
Joshua,
Fear not, neither be thou
dismayed:
take all the
people of
war with thee, and
arise, go
up to
Ai:
see, I have
given into thy
hand the
king of
Ai, and his
people, and his
city, and his
land:
\VS{2}And thou shalt
do to
Ai and her
king as thou
didst unto
Jericho and her
king: only the
spoil thereof, and the
cattle thereof, shall ye take for a
prey unto yourselves:
lay thee an
ambush for the
city
behind it.
\par }{\PP \VS{3}¶ So
Joshua
arose, and all the
people of
war, to go
up against
Ai: and
Joshua chose
out
thirty
thousand
mighty
men of
valour, and sent them
away by
night.
\VS{4}And he
commanded them,
saying,
Behold, ye shall lie in
wait against the
city,
{\ADD{even}}
behind the
city:
go not
very
far from the
city, but be ye all
ready:
\VS{5}And I, and all the
people that
{\ADD{are}} with me, will
approach unto the
city: and it shall come to pass, when they come
out
against us, as at the
first, that we will
flee
before them,
\VS{6}(For they will come
out
after us) till we have
drawn them from the
city; for they will
say, They
flee
before us, as at the
first: therefore we will
flee
before them.\FTNT{*}{{\FR 8.6 }
drawn: Heb. pulled}
\VS{7}Then ye shall rise
up from the
ambush, and
seize upon the
city: for the
{\ND{
LORD}} your
God will
deliver it into your
hand.
\VS{8}And it shall be, when ye have
taken the
city,
{\ADD{that}} ye shall
set the
city on
fire: according to the
commandment of the
{\ND{
LORD}} shall ye
do.
See, I have
commanded you.
\par }{\PP \VS{9}¶
Joshua therefore sent them
forth: and they
went to lie in
ambush, and
abode between
Beth-el and
Ai, on the west
side of
Ai: but
Joshua
lodged that
night
among the
people.
\VS{10}And
Joshua rose up
early in the
morning, and
numbered the
people, and went
up, he and the
elders of
Israel,
before the
people to
Ai.
\VS{11}And all the
people,
{\ADD{even the people}} of
war that
{\ADD{were}} with him, went
up, and drew
nigh, and
came before the
city, and
pitched on the north
side of
Ai: now
{\ADD{there was}} a
valley between them and
Ai.
\VS{12}And he
took about
five
thousand
men, and
set them to lie in
ambush between
Beth-el and
Ai, on the west
side of the
city.\FTNT{†}{{\FR 8.12 }
of…: or, of Ai}
\VS{13}And when they had
set the
people,
{\ADD{even}} all the
host that
{\ADD{was}} on the
north of the
city, and their liers in
wait on the
west of the
city,
Joshua
went that
night into the
midst of the
valley.\FTNT{‡}{{\FR 8.13 }
liers…: Heb. lying in wait}
\par }{\PP \VS{14}¶ And it came to pass, when the
king of
Ai
saw
{\ADD{it}}, that they
hasted and rose up
early, and the
men of the
city went
out
against
Israel to
battle, he and all his
people, at a time
appointed,
before the
plain; but he
wist not that
{\ADD{there were}} liers in
ambush against him
behind the
city.
\VS{15}And
Joshua and all
Israel made as if they were
beaten
before them, and
fled by the
way of the
wilderness.
\VS{16}And all the
people that
{\ADD{were}} in
Ai were
called together to
pursue
after them: and they
pursued
after
Joshua, and were drawn
away from the
city.
\VS{17}And there was not a
man
left in
Ai or
Beth-el, that went not
out
after
Israel: and they
left the
city
open, and
pursued after
Israel.
\VS{18}And the
{\ND{
LORD}}
said unto
Joshua, Stretch
out the
spear that
{\ADD{is}} in thy
hand toward
Ai; for I will
give it into thine
hand. And
Joshua stretched
out the
spear that
{\ADD{he had}} in his
hand toward the
city.
\VS{19}And the
ambush
arose
quickly out of their
place, and they
ran as soon as he had stretched
out his
hand: and they
entered into the
city, and
took it, and
hasted and
set the
city on
fire.
\VS{20}And when the
men of
Ai
looked
behind them, they
saw, and, behold, the
smoke of the
city ascended
up to
heaven, and they had no
power to
flee this way or that
way: and the
people that
fled to the
wilderness turned
back upon the
pursuers.\FTNT{§}{{\FR 8.20 }
power: Heb. hand}
\VS{21}And when
Joshua and all
Israel
saw that the
ambush had
taken the
city, and that the
smoke of the
city
ascended, then they turned
again, and
slew the
men of
Ai.
\VS{22}And the
other issued
out of the
city
against them; so they were in the
midst of
Israel, some on this side, and some on that side: and they
smote them, so
that they
let none of them
remain or
escape.
\par }{\PP \VS{23}And the
king of
Ai they
took
alive, and
brought him to
Joshua.
\VS{24}And it came to pass, when
Israel had made an
end of
slaying all the
inhabitants of
Ai in the
field, in the
wilderness wherein they
chased them, and when they were all
fallen on the
edge of the
sword, until they were
consumed, that all the
Israelites
returned unto
Ai, and
smote it with the
edge of the
sword.
\VS{25}And
{\ADD{so}} it was,
{\ADD{that}} all that
fell that
day, both of
men and
women,
{\ADD{were}}
twelve
thousand,
{\ADD{even}} all the
men of
Ai.
\VS{26}For
Joshua
drew not his
hand
back, wherewith he stretched
out the
spear, until he had utterly
destroyed all the
inhabitants of
Ai.
\VS{27}Only the
cattle and the
spoil of that
city
Israel took for a
prey unto themselves, according unto the
word of the
{\ND{
LORD}} which he
commanded
Joshua.
\VS{28}And
Joshua
burnt
Ai, and
made it an
heap for
ever,
{\ADD{even}} a
desolation unto this
day.
\VS{29}And the
king of
Ai he
hanged on a
tree until
eventide: and as soon as the
sun was
down,
Joshua
commanded that they should
take his
carcase
down from the
tree, and
cast it at the
entering of the
gate of the
city, and
raise thereon a
great
heap of
stones,
{\ADD{that remaineth}} unto this
day.
\par }{\PP \VS{30}¶ Then
Joshua
built an
altar unto the
{\ND{
LORD}}
God of
Israel in
mount
Ebal,
\VS{31}As
Moses the
servant of the
{\ND{
LORD}}
commanded the
children of
Israel, as it is
written in the
book of the
law of
Moses, an
altar of
whole
stones, over which no man hath lift
up
{\ADD{any}}
iron: and they
offered thereon burnt
offerings unto the
{\ND{
LORD}}, and
sacrificed peace
offerings.
\par }{\PP \VS{32}¶ And he
wrote there upon the
stones a
copy of the
law of
Moses, which he
wrote in the
presence of the
children of
Israel.
\VS{33}And all
Israel, and their
elders, and
officers, and their
judges,
stood on this side the
ark and on that side before the
priests the
Levites, which
bare the
ark of the
covenant of the
{\ND{
LORD}}, as well the
stranger, as he that was
born among them;
half of them
over
against
mount
Gerizim, and
half of them over
against
mount
Ebal; as
Moses the
servant of the
{\ND{
LORD}} had
commanded
before, that they should
bless the
people of
Israel.
\VS{34}And
afterward he
read all the
words of the
law, the
blessings and
cursings, according to all that is
written in the
book of the
law.
\VS{35}There was not a
word of all that
Moses
commanded, which
Joshua
read not before all the
congregation of
Israel, with the
women, and the little
ones, and the
strangers that were
conversant
among them.\FTNT{**}{{\FR 8.35 }
were…: Heb. walked}

\par }\Chap{9}{\PP \VerseOne{1}And it came to pass, when all the
kings which
{\ADD{were}} on this
side
Jordan, in the
hills, and in the
valleys, and in all the
coasts of the
great
sea over
against
Lebanon, the
Hittite, and the
Amorite, the
Canaanite, the
Perizzite, the
Hivite, and the
Jebusite,
heard
{\ADD{thereof;}}
\VS{2}That they
gathered themselves
together, to
fight with
Joshua and with
Israel, with
one
accord.\FTNT{*}{{\FR 9.2 }
accord: Heb. mouth}
\par }{\PP \VS{3}¶ And when the
inhabitants of
Gibeon
heard what
Joshua had
done unto
Jericho and to
Ai,
\VS{4}They did
work
wilily, and
went and made as if they had been
ambassadors, and
took
old
sacks upon their
asses, and
wine
bottles,
old, and
rent, and bound
up;
\VS{5}And
old
shoes and
clouted upon their
feet, and
old
garments upon them; and all the
bread of their
provision was
dry
{\ADD{and}}
mouldy.
\VS{6}And they
went to
Joshua unto the
camp at
Gilgal, and
said unto him, and to the
men of
Israel, We be
come from a
far
country: now therefore
make ye a
league with us.
\VS{7}And the
men of
Israel
said unto the
Hivites,
Peradventure ye
dwell
among us; and how shall we
make a
league with you?
\VS{8}And they
said unto
Joshua, We
{\ADD{are}} thy
servants. And
Joshua
said unto them, Who
{\ADD{are}} ye? and from
whence
come ye?
\VS{9}And they
said unto him, From a
very
far
country thy
servants are
come because of the
name of the
{\ND{
LORD}} thy
God: for we have
heard the
fame of him, and all that he
did in
Egypt,
\VS{10}And all that he
did to the
two
kings of the
Amorites, that
{\ADD{were}}
beyond
Jordan, to
Sihon
king of
Heshbon, and to
Og
king of
Bashan, which
{\ADD{was}} at
Ashtaroth.
\VS{11}Wherefore our
elders and all the
inhabitants of our
country
spake to us,
saying,
Take
victuals with
you for the
journey, and
go to
meet them, and
say unto them, We
{\ADD{are}} your
servants: therefore now
make ye a
league with us.\FTNT{†}{{\FR 9.11 }
with you: Heb. in your hand}
\VS{12}This our
bread we took
hot
{\ADD{for}} our
provision out of our
houses on the
day we came
forth to
go unto you; but now, behold, it is
dry, and it is
mouldy:
\VS{13}And these
bottles of
wine, which we
filled,
{\ADD{were}}
new; and, behold, they be
rent: and these our
garments and our
shoes are become
old by reason of the
very
long
journey.
\VS{14}And the
men
took of their
victuals, and
asked not
{\ADD{counsel}} at the
mouth of the
{\ND{
LORD}}.\FTNT{‡}{{\FR 9.14 }
the men…: or, they received the men by reason of their victuals}
\par }{\PP \VS{15}And
Joshua
made
peace with them, and
made a
league with them, to let them
live: and the
princes of the
congregation
sware unto them.
\par }{\PP \VS{16}¶ And it came to pass at the
end of
three
days
after they had
made a
league with them, that they
heard that they
{\ADD{were}} their
neighbours, and
{\ADD{that}} they
dwelt
among them.
\VS{17}And the
children of
Israel
journeyed, and
came unto their
cities on the
third
day. Now their
cities
{\ADD{were}}
Gibeon, and
Chephirah, and
Beeroth, and
Kirjath-jearim.
\VS{18}And the
children of
Israel
smote them not, because the
princes of the
congregation had
sworn unto them by the
{\ND{
LORD}}
God of
Israel. And all the
congregation
murmured against the
princes.
\VS{19}But all the
princes
said unto all the
congregation, We have
sworn unto them by the
{\ND{
LORD}}
God of
Israel: now therefore we
may not
touch them.
\VS{20}This we will
do to them; we will even let them
live, lest
wrath be upon us, because of the
oath which we
sware unto them.
\VS{21}And the
princes
said unto them, Let them
live; but let them be
hewers of
wood and
drawers of
water unto all the
congregation; as the
princes had
promised them.
\par }{\PP \VS{22}¶ And
Joshua
called for them, and he
spake unto them,
saying, Wherefore have ye
beguiled us,
saying, We
{\ADD{are}}
very
far from you; when ye
dwell
among us?
\VS{23}Now therefore ye
{\ADD{are}}
cursed, and there shall none of you be
freed from being
bondmen, and
hewers of
wood and
drawers of
water for the
house of my
God.\FTNT{§}{{\FR 9.23 }
none…: Heb. not be cut off from you}
\VS{24}And they
answered
Joshua, and
said, Because it was
certainly
told thy
servants, how that the
{\ND{
LORD}} thy
God
commanded his
servant
Moses to
give you all the
land, and to
destroy all the
inhabitants of the
land from
before you, therefore we were
sore
afraid of our
lives
because of you, and have
done this
thing.
\VS{25}And now, behold,
we
{\ADD{are}} in thine
hand: as it
seemeth
good and
right unto thee to
do unto us,
do.
\VS{26}And so
did he unto them, and
delivered them out of the
hand of the
children of
Israel, that they
slew them not.
\VS{27}And
Joshua
made them that
day
hewers of
wood and
drawers of
water for the
congregation, and for the
altar of the
{\ND{
LORD}}, even unto this
day, in the
place which he should
choose.\FTNT{**}{{\FR 9.27 }
made: Heb. gave, or, delivered to be}

\par }\Chap{10}{\PP \VerseOne{1}Now it came to pass, when
Adoni-zedek
king of
Jerusalem had
heard
how
Joshua had
taken
Ai, and had utterly
destroyed it; as he had
done to
Jericho and her
king, so he had
done to
Ai and her
king; and how the
inhabitants of
Gibeon had made
peace with
Israel, and were
among them;
\VS{2}That they
feared
greatly, because
Gibeon
{\ADD{was}} a
great
city, as
one of the
royal
cities, and because it
{\ADD{was}}
greater than
Ai, and all the
men thereof
{\ADD{were}}
mighty.\FTNT{*}{{\FR 10.2 }
royal…: Heb. cities of the kingdom}
\VS{3}Wherefore
Adoni-zedek
king of
Jerusalem
sent unto
Hoham
king of
Hebron, and unto
Piram
king of
Jarmuth, and unto
Japhia
king of
Lachish, and unto
Debir
king of
Eglon,
saying,
\VS{4}Come
up unto me, and
help me, that we may
smite
Gibeon: for it hath made
peace with
Joshua and with the
children of
Israel.
\VS{5}Therefore the
five
kings of the
Amorites, the
king of
Jerusalem, the
king of
Hebron, the
king of
Jarmuth, the
king of
Lachish, the
king of
Eglon, gathered themselves
together, and went
up, they and all their
hosts, and
encamped before
Gibeon, and made
war against it.
\par }{\PP \VS{6}¶ And the
men of
Gibeon
sent unto
Joshua to the
camp to
Gilgal,
saying,
Slack not thy
hand from thy
servants; come
up to us
quickly, and
save us, and
help us: for all the
kings of the
Amorites that
dwell in the
mountains are gathered
together against us.
\par }{\PP \VS{7}So
Joshua
ascended from
Gilgal, he, and all the
people of
war with him, and all the
mighty men of
valour.
\par }{\PP \VS{8}¶ And the
{\ND{
LORD}}
said unto
Joshua,
Fear them not: for I have
delivered them into thine
hand; there shall not a
man of them
stand
before thee.
\VS{9}Joshua therefore
came unto them
suddenly,
{\ADD{and}} went
up from
Gilgal all
night.
\VS{10}And the
{\ND{
LORD}}
discomfited them
before
Israel, and
slew them with a
great
slaughter at
Gibeon, and
chased them along the
way that goeth
up to
Beth-horon, and
smote them to
Azekah, and unto
Makkedah.
\VS{11}And it came to pass, as they
fled from
before
Israel,
{\ADD{and}} were in the going
down to
Beth-horon, that the
{\ND{
LORD}} cast
down
great
stones from
heaven upon them unto
Azekah, and they
died:
{\ADD{they were}}
more which
died with
hailstones than
{\ADD{they}}
whom the
children of
Israel
slew with the
sword.
\par }{\PP \VS{12}¶ Then
spake
Joshua to the
{\ND{
LORD}} in the
day when the
{\ND{
LORD}} delivered
up the
Amorites
before the
children of
Israel, and he
said in the
sight of
Israel,
Sun, stand thou
still upon
Gibeon; and thou,
Moon, in the
valley of
Ajalon.\FTNT{†}{{\FR 10.12 }
stand…: Heb. be silent}
\VS{13}And the
sun stood
still, and the
moon
stayed, until the
people had
avenged themselves upon their
enemies.
{\ADD{Is}} not
this
written in the
book of
Jasher? So the
sun stood
still in the
midst of
heaven, and
hasted not to go
down about a
whole
day.\FTNT{‡}{{\FR 10.13 }
Jasher: or, the upright?}
\VS{14}And there was no
day like that
before it or
after it, that the
{\ND{
LORD}}
hearkened unto the
voice of a
man: for the
{\ND{
LORD}}
fought for
Israel.
\par }{\PP \VS{15}¶ And
Joshua
returned, and all
Israel with him, unto the
camp to
Gilgal.
\VS{16}But these
five
kings
fled, and
hid themselves in a
cave at
Makkedah.
\VS{17}And it was
told
Joshua,
saying, The
five
kings are
found
hid in a
cave at
Makkedah.
\VS{18}And
Joshua
said,
Roll
great
stones upon the
mouth of the
cave, and
set
men by it for to
keep them:
\VS{19}And
stay ye not,
{\ADD{but}}
pursue
after your
enemies, and smite the
hindmost of them;
suffer them not to
enter into their
cities: for the
{\ND{
LORD}} your
God hath
delivered them into your
hand.\FTNT{§}{{\FR 10.19 }
smite…: Heb. cut off the tail}
\VS{20}And it came to pass, when
Joshua and the
children of
Israel had made an
end of
slaying them with a
very
great
slaughter, till they were
consumed, that the
rest
{\ADD{which}}
remained of them
entered into
fenced
cities.
\VS{21}And all the
people
returned to the
camp to
Joshua at
Makkedah in
peace: none
moved his
tongue against
any of the
children of
Israel.
\VS{22}Then
said
Joshua,
Open the
mouth of the
cave, and bring
out those
five
kings unto me out of the
cave.
\VS{23}And they
did so, and brought
forth those
five
kings unto him out of the
cave, the
king of
Jerusalem, the
king of
Hebron, the
king of
Jarmuth, the
king of
Lachish,
{\ADD{and}} the
king of
Eglon.
\VS{24}And it came to pass, when they brought
out those
kings unto
Joshua, that
Joshua
called for all the
men of
Israel, and
said unto the
captains of the
men of
war which
went with him, Come
near,
put your
feet upon the
necks of these
kings. And they came
near, and
put their
feet upon the
necks of them.
\VS{25}And
Joshua
said unto them,
Fear not, nor be
dismayed, be
strong and of good
courage: for thus shall the
{\ND{
LORD}}
do to all your
enemies against whom ye
fight.
\VS{26}And
afterward
Joshua
smote them, and
slew them, and
hanged them on
five
trees: and they were
hanging upon the
trees until the
evening.
\VS{27}And it came to pass at the
time of the going
down of the
sun,
{\ADD{that}}
Joshua
commanded, and they took them
down off the
trees, and
cast them into the
cave wherein they had been
hid, and
laid
great
stones in the
cave’s
mouth,
{\ADD{which remain}} until this
very
day.
\par }{\PP \VS{28}¶ And that
day
Joshua
took
Makkedah, and
smote it with the
edge of the
sword, and the
king thereof he utterly
destroyed, them, and all the
souls that
{\ADD{were}} therein; he
let none
remain: and he
did to the
king of
Makkedah as he
did unto the
king of
Jericho.
\VS{29}Then
Joshua
passed from
Makkedah, and all
Israel with him, unto
Libnah, and
fought against
Libnah:
\VS{30}And the
{\ND{
LORD}}
delivered it also, and the
king thereof, into the
hand of
Israel; and he
smote it with the
edge of the
sword, and all the
souls that
{\ADD{were}} therein; he
let none
remain in it; but
did unto the
king thereof as he
did unto the
king of
Jericho.
\par }{\PP \VS{31}¶ And
Joshua
passed from
Libnah, and all
Israel with him, unto
Lachish, and
encamped against it, and
fought against it:
\VS{32}And the
{\ND{
LORD}}
delivered
Lachish into the
hand of
Israel, which
took it on the
second
day, and
smote it with the
edge of the
sword, and all the
souls that
{\ADD{were}} therein, according to all that he had
done to
Libnah.
\par }{\PP \VS{33}¶ Then
Horam
king of
Gezer came
up to
help
Lachish; and
Joshua
smote him and his
people, until he had
left him none
remaining.
\par }{\PP \VS{34}¶ And from
Lachish
Joshua
passed unto
Eglon, and all
Israel with him; and they
encamped against it, and
fought against it:
\VS{35}And they
took it on that
day, and
smote it with the
edge of the
sword, and all the
souls that
{\ADD{were}} therein he utterly
destroyed that
day, according to all that he had
done to
Lachish.
\VS{36}And
Joshua went
up from
Eglon, and all
Israel with him, unto
Hebron; and they
fought against it:
\VS{37}And they
took it, and
smote it with the
edge of the
sword, and the
king thereof, and all the
cities thereof, and all the
souls that
{\ADD{were}} therein; he
left none
remaining, according to all that he had
done to
Eglon; but
destroyed it utterly, and all the
souls that
{\ADD{were}} therein.
\par }{\PP \VS{38}¶ And
Joshua
returned, and all
Israel with him, to
Debir; and
fought against it:
\VS{39}And he
took it, and the
king thereof, and all the
cities thereof; and they
smote them with the
edge of the
sword, and utterly
destroyed all the
souls that
{\ADD{were}} therein; he
left none
remaining: as he had
done to
Hebron, so he
did to
Debir, and to the
king thereof; as he had
done also to
Libnah, and to her
king.
\par }{\PP \VS{40}¶ So
Joshua
smote all the
country of the
hills, and of the
south, and of the
vale, and of the
springs, and all their
kings: he
left none
remaining, but utterly
destroyed all that
breathed, as the
{\ND{
LORD}}
God of
Israel
commanded.
\VS{41}And
Joshua
smote them from
Kadesh-barnea even unto
Gaza, and all the
country of
Goshen, even unto
Gibeon.
\VS{42}And all these
kings and their
land did
Joshua
take at
one
time, because the
{\ND{
LORD}}
God of
Israel
fought for
Israel.
\VS{43}And
Joshua
returned, and all
Israel with him, unto the
camp to
Gilgal.

\par }\Chap{11}{\PP \VerseOne{1}And it came to pass, when
Jabin
king of
Hazor had
heard
{\ADD{those things}}, that he
sent to
Jobab
king of
Madon, and to the
king of
Shimron, and to the
king of
Achshaph,
\VS{2}And to the
kings that
{\ADD{were}} on the
north of the
mountains, and of the
plains
south of
Chinneroth, and in the
valley, and in the
borders of
Dor on the
west,
\VS{3}{\ADD{And to}} the
Canaanite on the
east and on the
west, and
{\ADD{to}} the
Amorite, and the
Hittite, and the
Perizzite, and the
Jebusite in the
mountains, and
{\ADD{to}} the
Hivite under
Hermon in the
land of
Mizpeh.
\VS{4}And they went
out, they and all their
hosts with them,
much
people, even as the
sand that
{\ADD{is}} upon the
sea
shore in
multitude, with
horses and
chariots
very
many.
\VS{5}And when all these
kings were met
together, they
came and
pitched
together at the
waters of
Merom, to
fight against
Israel.\FTNT{*}{{\FR 11.5 }
met…: Heb. assembled by appointment}
\par }{\PP \VS{6}¶ And the
{\ND{
LORD}}
said unto
Joshua, Be not
afraid
because of them: for to
morrow about this
time will I deliver them
up all
slain
before
Israel: thou shalt
hough their
horses, and
burn their
chariots with
fire.
\VS{7}So
Joshua
came, and all the
people of
war with him, against them by the
waters of
Merom
suddenly; and they fell
upon them.
\VS{8}And the
{\ND{
LORD}}
delivered them into the
hand of
Israel, who
smote them, and
chased them unto
great
Zidon, and unto
Misrephoth-maim, and unto the
valley of
Mizpeh
eastward; and they
smote them, until they
left them none
remaining.\FTNT{†}{{\FR 11.8 }
great Zidon: or, Zidonrabbah}\FTNT{‡}{{\FR 11.8 }
Misrephoth-maim: or, Salt pits: Heb. Burnings of waters}
\VS{9}And
Joshua
did unto them as the
{\ND{
LORD}}
bade him: he
houghed their
horses, and
burnt their
chariots with
fire.
\par }{\PP \VS{10}¶ And
Joshua at that
time turned
back, and
took
Hazor, and
smote the
king thereof with the
sword: for
Hazor
beforetime was the
head of all those
kingdoms.
\VS{11}And they
smote all the
souls that
{\ADD{were}} therein with the
edge of the
sword, utterly
destroying
{\ADD{them:}} there was not any
left to
breathe: and he
burnt
Hazor with
fire.\FTNT{§}{{\FR 11.11 }
any…: Heb. any breath}
\VS{12}And all the
cities of those
kings, and all the
kings of them, did
Joshua
take, and
smote them with the
edge of the
sword,
{\ADD{and}} he utterly
destroyed them, as
Moses the
servant of the
{\ND{
LORD}}
commanded.
\VS{13}But
{\ADD{as for}} the
cities that
stood still in their
strength,
Israel
burned none of them,
save
Hazor only;
{\ADD{that}} did
Joshua
burn.\FTNT{**}{{\FR 11.13 }
in their…: Heb. on their heap}
\VS{14}And all the
spoil of these
cities, and the
cattle, the
children of
Israel took for a
prey unto themselves; but every
man they
smote with the
edge of the
sword, until they had
destroyed them, neither
left they any to
breathe.
\par }{\PP \VS{15}¶ As the
{\ND{
LORD}}
commanded
Moses his
servant, so did
Moses
command
Joshua, and so
did
Joshua; he
left
nothing
undone of all that the
{\ND{
LORD}}
commanded
Moses.\FTNT{††}{{\FR 11.15 }
left…: Heb. removed nothing}
\VS{16}So
Joshua
took all that
land, the
hills, and all the south
country, and all the
land of
Goshen, and the
valley, and the
plain, and the
mountain of
Israel, and the
valley of the same;
\VS{17}{\ADD{Even}} from the
mount
Halak, that goeth
up to
Seir, even unto
Baal-gad in the
valley of
Lebanon under
mount
Hermon: and all their
kings he
took, and
smote them, and
slew them.\FTNT{‡‡}{{\FR 11.17 }
the mount…: or, the smooth mountain}
\VS{18}Joshua
made
war a
long
time with all those
kings.
\VS{19}There was not a
city that made
peace with the
children of
Israel, save the
Hivites the
inhabitants of
Gibeon: all
{\ADD{other}} they
took in
battle.
\VS{20}For it was of the
{\ND{
LORD}} to
harden their
hearts, that they should come
against
Israel in
battle, that he might destroy them
utterly,
{\ADD{and}} that they might have no
favour, but that he might
destroy them, as the
{\ND{
LORD}}
commanded
Moses.
\par }{\PP \VS{21}¶ And at that
time
came
Joshua, and cut
off the
Anakims from the
mountains, from
Hebron, from
Debir, from
Anab, and from all the
mountains of
Judah, and from all the
mountains of
Israel:
Joshua destroyed them
utterly with their
cities.
\VS{22}There was none of the
Anakims
left in the
land of the
children of
Israel: only in
Gaza, in
Gath, and in
Ashdod, there
remained.
\VS{23}So
Joshua
took the whole
land, according to all that the
{\ND{
LORD}}
said unto
Moses; and
Joshua
gave it for an
inheritance unto
Israel according to their
divisions by their
tribes. And the
land
rested from
war.

\par }\Chap{12}{\PP \VerseOne{1}Now these
{\ADD{are}} the
kings of the
land, which the
children of
Israel
smote, and
possessed their
land on the other
side
Jordan toward the
rising of the
sun, from the
river
Arnon unto
mount
Hermon, and all the
plain on the
east:
\VS{2}Sihon
king of the
Amorites, who
dwelt in
Heshbon,
{\ADD{and}}
ruled from
Aroer, which
{\ADD{is}} upon the
bank of the
river
Arnon, and from the
middle of the
river, and from
half
Gilead, even unto the
river
Jabbok,
{\ADD{which is}} the
border of the
children of
Ammon;
\VS{3}And from the
plain to the
sea of
Chinneroth on the
east, and unto the
sea of the
plain,
{\ADD{even}} the
salt
sea on the
east, the
way to
Beth-jeshimoth; and from the
south, under
Ashdoth-pisgah:\FTNT{*}{{\FR 12.3 }
the south: or, Teman}\FTNT{†}{{\FR 12.3 }
Ashdoth-pisgah: or, the springs of Pisgah, or, the hill}
\par }{\PP \VS{4}¶ And the
coast of
Og
king of
Bashan,
{\ADD{which was}} of the
remnant of the
giants, that
dwelt at
Ashtaroth and at
Edrei,
\VS{5}And
reigned in
mount
Hermon, and in
Salcah, and in all
Bashan, unto the
border of the
Geshurites and the
Maachathites, and
half
Gilead, the
border of
Sihon
king of
Heshbon.
\VS{6}Them did
Moses the
servant of the
{\ND{
LORD}} and the
children of
Israel
smite: and
Moses the
servant of the
{\ND{
LORD}}
gave it
{\ADD{for}} a
possession unto the
Reubenites, and the
Gadites, and the
half
tribe of
Manasseh.
\par }{\PP \VS{7}¶ And these
{\ADD{are}} the
kings of the
country which
Joshua and the
children of
Israel
smote on this
side
Jordan on the
west, from
Baal-gad in the
valley of
Lebanon even unto the
mount
Halak, that goeth
up to
Seir; which
Joshua
gave unto the
tribes of
Israel
{\ADD{for}} a
possession according to their
divisions;
\VS{8}In the
mountains, and in the
valleys, and in the
plains, and in the
springs, and in the
wilderness, and in the south
country; the
Hittites, the
Amorites, and the
Canaanites, the
Perizzites, the
Hivites, and the
Jebusites:
\par }{\PP \VS{9}¶ The
king of
Jericho,
one; the
king of
Ai, which
{\ADD{is}}
beside
Beth-el,
one;
\VS{10}The
king of
Jerusalem,
one; the
king of
Hebron,
one;
\VS{11}The
king of
Jarmuth,
one; the
king of
Lachish,
one;
\VS{12}The
king of
Eglon,
one; the
king of
Gezer,
one;
\VS{13}The
king of
Debir,
one; the
king of
Geder,
one;
\VS{14}The
king of
Hormah,
one; the
king of
Arad,
one;
\VS{15}The
king of
Libnah,
one; the
king of
Adullam,
one;
\VS{16}The
king of
Makkedah,
one; the
king of
Beth-el,
one;
\VS{17}The
king of
Tappuah,
one; the
king of
Hepher,
one;
\VS{18}The
king of
Aphek,
one; the
king of
Lasharon,
one;\FTNT{‡}{{\FR 12.18 }
Lasharon: or, Sharon}
\VS{19}The
king of
Madon,
one; the
king of
Hazor,
one;
\VS{20}The
king of
Shimron-meron,
one; the
king of
Achshaph,
one;
\VS{21}The
king of
Taanach,
one; the
king of
Megiddo,
one;
\VS{22}The
king of
Kedesh,
one; the
king of
Jokneam of
Carmel,
one;
\VS{23}The
king of
Dor in the
coast of
Dor,
one; the
king of the
nations of
Gilgal,
one;
\VS{24}The
king of
Tirzah,
one: all the
kings
thirty and
one.

\par }\Chap{13}{\PP \VerseOne{1}Now
Joshua was
old
{\ADD{and}}
stricken in
years; and the
{\ND{
LORD}}
said unto him, Thou art
old
{\ADD{and}}
stricken in
years, and there
remaineth yet
very
much
land to be
possessed.\FTNT{*}{{\FR 13.1 }
to…: Heb. to possess it}
\VS{2}This
{\ADD{is}} the
land that yet
remaineth: all the
borders of the
Philistines, and all
Geshuri,
\VS{3}From
Sihor, which
{\ADD{is}}
before
Egypt, even unto the
borders of
Ekron
northward,
{\ADD{which}} is
counted to the
Canaanite:
five
lords of the
Philistines; the
Gazathites, and the
Ashdothites, the
Eshkalonites, the
Gittites, and the
Ekronites; also the
Avites:
\VS{4}From the
south, all the
land of the
Canaanites, and
Mearah that
{\ADD{is}} beside the
Sidonians, unto
Aphek, to the
borders of the
Amorites:\FTNT{†}{{\FR 13.4 }
Mearah: or, the cave}
\VS{5}And the
land of the
Giblites, and all
Lebanon, toward the
sunrising, from
Baal-gad under
mount
Hermon unto the
entering into
Hamath.
\VS{6}All the
inhabitants of the hill
country from
Lebanon unto
Misrephoth-maim,
{\ADD{and}} all the
Sidonians, them will I drive
out from
before the
children of
Israel: only divide thou it by
lot unto the
Israelites for an
inheritance, as I have
commanded thee.
\par }{\PP \VS{7}Now therefore
divide this
land for an
inheritance unto the
nine
tribes, and the
half
tribe of
Manasseh,
\VS{8}With whom the
Reubenites and the
Gadites have
received their
inheritance, which
Moses
gave them,
beyond
Jordan
eastward,
{\ADD{even}} as
Moses the
servant of the
{\ND{
LORD}}
gave them;
\VS{9}From
Aroer, that
{\ADD{is}} upon the
bank of the
river
Arnon, and the
city that
{\ADD{is}} in the
midst of the
river, and all the
plain of
Medeba unto
Dibon;
\VS{10}And all the
cities of
Sihon
king of the
Amorites, which
reigned in
Heshbon, unto the
border of the
children of
Ammon;
\VS{11}And
Gilead, and the
border of the
Geshurites and
Maachathites, and all
mount
Hermon, and all
Bashan unto
Salcah;
\VS{12}All the
kingdom of
Og in
Bashan, which
reigned in
Ashtaroth and in
Edrei, who
remained of the
remnant of the
giants: for these did
Moses
smite, and cast them
out.
\VS{13}Nevertheless the
children of
Israel
expelled not the
Geshurites, nor the
Maachathites: but the
Geshurites and the
Maachathites
dwell
among the
Israelites until this
day.
\VS{14}Only unto the
tribe of
Levi he
gave none
inheritance; the
sacrifices of the
{\ND{
LORD}}
God of
Israel made by
fire
{\ADD{are}} their
inheritance, as he
said unto them.
\par }{\PP \VS{15}¶ And
Moses
gave unto the
tribe of the
children of
Reuben
{\ADD{inheritance}} according to their
families.
\VS{16}And their
coast was from
Aroer, that
{\ADD{is}} on the
bank of the
river
Arnon, and the
city that
{\ADD{is}} in the
midst of the
river, and all the
plain by
Medeba;
\VS{17}Heshbon, and all her
cities that
{\ADD{are}} in the
plain;
Dibon, and
Bamoth-baal, and
Beth-baal-meon,\FTNT{‡}{{\FR 13.17 }
Bamoth-baal…: or, the high places of Baal, and house of Baal-meon}
\VS{18}And
Jahazah, and
Kedemoth, and
Mephaath,
\VS{19}And
Kirjathaim, and
Sibmah, and
Zareth-shahar in the
mount of the
valley,
\VS{20}And
Beth-peor, and
Ashdoth-pisgah, and
Beth-jeshimoth,\FTNT{§}{{\FR 13.20 }
Ashdoth-pisgah: or, springs of Pisgah, or, the hill}
\VS{21}And all the
cities of the
plain, and all the
kingdom of
Sihon
king of the
Amorites, which
reigned in
Heshbon, whom
Moses
smote with the
princes of
Midian,
Evi, and
Rekem, and
Zur, and
Hur, and
Reba,
{\ADD{which were}}
dukes of
Sihon,
dwelling in the
country.
\par }{\PP \VS{22}¶
Balaam also the
son of
Beor, the
soothsayer, did the
children of
Israel
slay with the
sword among
them that were
slain by them.\FTNT{**}{{\FR 13.22 }
soothsayer: or, diviner}
\VS{23}And the
border of the
children of
Reuben was
Jordan, and the
border
{\ADD{thereof}}. This
{\ADD{was}} the
inheritance of the
children of
Reuben after their
families, the
cities and the
villages thereof.
\VS{24}And
Moses
gave
{\ADD{inheritance}} unto the
tribe of
Gad,
{\ADD{even}} unto the
children of
Gad according to their
families.
\VS{25}And their
coast was
Jazer, and all the
cities of
Gilead, and
half the
land of the
children of
Ammon, unto
Aroer that
{\ADD{is}}
before
Rabbah;
\VS{26}And from
Heshbon unto
Ramath-mizpeh, and
Betonim; and from
Mahanaim unto the
border of
Debir;
\VS{27}And in the
valley,
Beth-aram, and
Beth-nimrah, and
Succoth, and
Zaphon, the
rest of the
kingdom of
Sihon
king of
Heshbon,
Jordan and
{\ADD{his}}
border,
{\ADD{even}} unto the
edge of the
sea of
Chinnereth on the other
side
Jordan
eastward.
\VS{28}This
{\ADD{is}} the
inheritance of the
children of
Gad after their
families, the
cities, and their
villages.
\par }{\PP \VS{29}¶ And
Moses
gave
{\ADD{inheritance}} unto the
half
tribe of
Manasseh: and
{\ADD{this}} was
{\ADD{the possession}} of the
half
tribe of the
children of
Manasseh by their
families.
\VS{30}And their
coast was from
Mahanaim, all
Bashan, all the
kingdom of
Og
king of
Bashan, and all the
towns of
Jair, which
{\ADD{are}} in
Bashan,
threescore
cities:
\VS{31}And
half
Gilead, and
Ashtaroth, and
Edrei,
cities of the
kingdom of
Og in
Bashan,
{\ADD{were pertaining}} unto the
children of
Machir the
son of
Manasseh,
{\ADD{even}} to the one
half of the
children of
Machir by their
families.
\VS{32}These
{\ADD{are the countries}} which
Moses did distribute for
inheritance in the
plains of
Moab, on the other
side
Jordan, by
Jericho,
eastward.
\VS{33}But unto the
tribe of
Levi
Moses
gave not
{\ADD{any}}
inheritance: the
{\ND{
LORD}}
God of
Israel
{\ADD{was}} their
inheritance, as he
said unto them.

\par }\Chap{14}{\PP \VerseOne{1}And these
{\ADD{are the countries}} which the
children of
Israel
inherited in the
land of
Canaan, which
Eleazar the
priest, and
Joshua the
son of
Nun, and the
heads of the
fathers of the
tribes of the
children of
Israel, distributed for
inheritance to them.
\VS{2}By
lot
{\ADD{was}} their
inheritance, as the
{\ND{
LORD}}
commanded by the
hand of
Moses, for the
nine
tribes, and
{\ADD{for}} the
half
tribe.
\VS{3}For
Moses had
given the
inheritance of
two
tribes and an
half
tribe on the other
side
Jordan: but unto the
Levites he
gave none
inheritance
among them.
\VS{4}For the
children of
Joseph were
two
tribes,
Manasseh and
Ephraim: therefore they
gave no
part unto the
Levites in the
land, save
cities to
dwell
{\ADD{in}}, with their
suburbs for their
cattle and for their
substance.
\VS{5}As the
{\ND{
LORD}}
commanded
Moses, so the
children of
Israel
did, and they
divided the
land.
\par }{\PP \VS{6}¶ Then the
children of
Judah
came unto
Joshua in
Gilgal: and
Caleb the
son of
Jephunneh the
Kenezite
said unto him, Thou
knowest the
thing that the
{\ND{
LORD}}
said unto
Moses the
man of
God concerning
me and
thee in
Kadesh-barnea.
\VS{7}Forty
years
old
{\ADD{was}} I when
Moses the
servant of the
{\ND{
LORD}}
sent me from
Kadesh-barnea to espy
out the
land; and I
brought him
word
again as
{\ADD{it was}} in mine
heart.
\VS{8}Nevertheless my
brethren that went
up with me made the
heart of the
people
melt: but I
wholly
followed the
{\ND{
LORD}} my
God.
\VS{9}And
Moses
sware on that
day,
saying, Surely the
land whereon thy
feet have
trodden shall be thine
inheritance, and thy
children’s
for
ever, because thou hast
wholly
followed the
{\ND{
LORD}} my
God.
\VS{10}And now, behold, the
{\ND{
LORD}} hath kept me
alive, as he
said, these
forty and
five
years, even
since the
{\ND{
LORD}}
spake this
word unto
Moses, while
{\ADD{the children of}}
Israel
wandered in the
wilderness: and now, lo, I
{\ADD{am}} this
day
fourscore and
five
years
old.\FTNT{*}{{\FR 14.10 }
wandered: Heb. walked}
\VS{11}As yet I
{\ADD{am as}}
strong this
day as
{\ADD{I was}} in the
day that
Moses
sent me: as my
strength
{\ADD{was}} then, even so
{\ADD{is}} my
strength now, for
war, both to go
out, and to come
in.
\VS{12}Now therefore
give me this
mountain, whereof the
{\ND{
LORD}}
spake in that
day; for thou
heardest in that
day how the
Anakims
{\ADD{were}} there, and
{\ADD{that}} the
cities
{\ADD{were}}
great
{\ADD{and}}
fenced: if so
be the
{\ND{
LORD}}
{\ADD{will be}}
with me, then I shall be able to drive them
out, as the
{\ND{
LORD}}
said.
\VS{13}And
Joshua
blessed him, and
gave unto
Caleb the
son of
Jephunneh
Hebron for an
inheritance.
\VS{14}Hebron therefore became the
inheritance of
Caleb the
son of
Jephunneh the
Kenezite unto this
day, because that he
wholly
followed the
{\ND{
LORD}}
God of
Israel.
\VS{15}And the
name of
Hebron
before
{\ADD{was}}
Kirjath-arba;
{\ADD{which Arba was}} a
great
man among the
Anakims. And the
land had
rest from
war.

\par }\Chap{15}{\PP \VerseOne{1}{\ADD{This}} then was the
lot of the
tribe of the
children of
Judah by their
families;
{\ADD{even}} to the
border of
Edom the
wilderness of
Zin
southward
{\ADD{was}} the uttermost
part of the south
coast.
\VS{2}And their
south
border was from the
shore of the
salt
sea, from the
bay that
looketh
southward:\FTNT{*}{{\FR 15.2 }
bay: Heb. tongue}
\VS{3}And it went
out to the south
side to
Maaleh-acrabbim, and
passed along to
Zin, and ascended
up on the south
side unto
Kadesh-barnea, and
passed along to
Hezron, and went
up to
Adar, and fetched a
compass to
Karkaa:\FTNT{†}{{\FR 15.3 }
Maaleh-acrabbim: or, the going up to Acrabbim}
\VS{4}{\ADD{From thence}} it
passed toward
Azmon, and went
out unto the
river of
Egypt; and the goings
out of that
coast were at the
sea: this shall be your
south
coast.
\VS{5}And the
east
border
{\ADD{was}} the
salt
sea,
{\ADD{even}} unto the
end of
Jordan. And
{\ADD{their}}
border in the
north
quarter
{\ADD{was}} from the
bay of the
sea at the uttermost
part of
Jordan:
\VS{6}And the
border went
up to
Beth-hogla, and
passed along by the
north of
Beth-arabah; and the
border went
up to the
stone of
Bohan the
son of
Reuben:
\VS{7}And the
border went
up toward
Debir from the
valley of
Achor, and so
northward,
looking toward
Gilgal, that
{\ADD{is}}
before the going
up to
Adummim, which
{\ADD{is}} on the south
side of the
river: and the
border
passed toward the
waters of
En-shemesh, and the goings
out thereof were at
En-rogel:
\VS{8}And the
border went
up by the
valley of the
son of
Hinnom unto the
south
side of the
Jebusite; the same
{\ADD{is}}
Jerusalem: and the
border went
up to the
top of the
mountain that
{\ADD{lieth}}
before the
valley of
Hinnom
westward, which
{\ADD{is}} at the
end of the
valley of the
giants
northward:
\VS{9}And the
border was
drawn from the
top of the
hill unto the
fountain of the
water of
Nephtoah, and went
out to the
cities of
mount
Ephron; and the
border was
drawn to
Baalah, which
{\ADD{is}}
Kirjath-jearim:
\VS{10}And the
border
compassed from
Baalah
westward unto
mount
Seir, and
passed along unto the
side of
mount
Jearim, which
{\ADD{is}}
Chesalon, on the north
side, and went
down to
Beth-shemesh, and passed
on to
Timnah:
\VS{11}And the
border went
out unto the
side of
Ekron
northward: and the
border was
drawn to
Shicron, and passed
along to
mount
Baalah, and went
out unto
Jabneel; and the goings
out of the
border were at the
sea.
\VS{12}And the
west
border
{\ADD{was}} to the
great
sea, and the
coast
{\ADD{thereof}}. This
{\ADD{is}} the
coast of the
children of
Judah round
about according to their
families.
\par }{\PP \VS{13}¶ And unto
Caleb the
son of
Jephunneh he
gave a
part
among the
children of
Judah,
according to the
commandment of the
{\ND{
LORD}} to
Joshua,
{\ADD{even}} the
city of
Arba the
father of
Anak, which
{\ADD{city is}}
Hebron.\FTNT{‡}{{\FR 15.13 }
the city…: or, Kirjath-arba}
\VS{14}And
Caleb
drove thence the
three
sons of
Anak,
Sheshai, and
Ahiman, and
Talmai, the
children of
Anak.
\VS{15}And he went
up thence to the
inhabitants of
Debir: and the
name of
Debir
before
{\ADD{was}}
Kirjath-sepher.
\par }{\PP \VS{16}¶ And
Caleb
said, He that
smiteth
Kirjath-sepher, and
taketh it, to him will I
give
Achsah my
daughter to
wife.
\VS{17}And
Othniel the
son of
Kenaz, the
brother of
Caleb,
took it: and he
gave him
Achsah his
daughter to
wife.
\VS{18}And it came to pass, as she
came
{\ADD{unto him}}, that she
moved him to
ask of her
father a
field: and she lighted
off
{\ADD{her}}
ass; and
Caleb
said unto her, What wouldest thou?
\VS{19}Who
answered,
Give me a
blessing; for thou hast
given me a
south
land;
give me also
springs of
water. And he
gave her the
upper
springs, and the
nether
springs.
\par }{\PP \VS{20}This
{\ADD{is}} the
inheritance of the
tribe of the
children of
Judah according to their
families.
\VS{21}And the
uttermost
cities of the
tribe of the
children of
Judah toward the
coast of
Edom
southward were
Kabzeel, and
Eder, and
Jagur,
\VS{22}And
Kinah, and
Dimonah, and
Adadah,
\VS{23}And
Kedesh, and
Hazor, and
Ithnan,
\VS{24}Ziph, and
Telem, and
Bealoth,
\VS{25}And
Hazor,
Hadattah, and
Kerioth,
{\ADD{and}}
Hezron, which
{\ADD{is}}
Hazor,
\VS{26}Amam, and
Shema, and
Moladah,
\VS{27}And
Hazar-gaddah, and
Heshmon, and
Beth-palet,
\VS{28}And
Hazar-shual, and
Beer-sheba, and
Bizjothjah,
\VS{29}Baalah, and
Iim, and
Azem,
\VS{30}And
Eltolad, and
Chesil, and
Hormah,
\VS{31}And
Ziklag, and
Madmannah, and
Sansannah,
\VS{32}And
Lebaoth, and
Shilhim, and
Ain, and
Rimmon: all the
cities
{\ADD{are}}
twenty and
nine, with their
villages:
\VS{33}{\ADD{And}} in the
valley,
Eshtaol, and
Zoreah, and
Ashnah,
\VS{34}And
Zanoah, and
En-gannim,
Tappuah, and
Enam,
\VS{35}Jarmuth, and
Adullam,
Socoh, and
Azekah,
\VS{36}And
Sharaim, and
Adithaim, and
Gederah, and
Gederothaim;
fourteen
cities with their
villages:\FTNT{§}{{\FR 15.36 }
and Gederothaim: or, or Gederothaim}
\VS{37}Zenan, and
Hadashah, and
Migdal-gad,
\VS{38}And
Dilean, and
Mizpeh, and
Joktheel,
\VS{39}Lachish, and
Bozkath, and
Eglon,
\VS{40}And
Cabbon, and
Lahmam, and
Kithlish,
\VS{41}And
Gederoth,
Beth-dagon, and
Naamah, and
Makkedah;
sixteen
cities with their
villages:
\VS{42}Libnah, and
Ether, and
Ashan,
\VS{43}And
Jiphtah, and
Ashnah, and
Nezib,
\VS{44}And
Keilah, and
Achzib, and
Mareshah;
nine
cities with their
villages:
\VS{45}Ekron, with her
towns and her
villages:
\VS{46}From
Ekron even unto the
sea, all that
{\ADD{lay}}
near
Ashdod, with their
villages:\FTNT{**}{{\FR 15.46 }
near: Heb. by the place of}
\VS{47}Ashdod with her
towns and her
villages,
Gaza with her
towns and her
villages, unto the
river of
Egypt, and the
great
sea, and the
border
{\ADD{thereof:}}
\par }{\PP \VS{48}¶ And in the
mountains,
Shamir, and
Jattir, and
Socoh,
\VS{49}And
Dannah, and
Kirjath-sannah, which
{\ADD{is}}
Debir,
\VS{50}And
Anab, and
Eshtemoh, and
Anim,
\VS{51}And
Goshen, and
Holon, and
Giloh;
eleven
cities with their
villages:
\VS{52}Arab, and
Dumah, and
Eshean,
\VS{53}And
Janum, and
Beth-tappuah, and
Aphekah,\FTNT{††}{{\FR 15.53 }
Janum: or, Janus}
\VS{54}And
Humtah, and
Kirjath-arba, which
{\ADD{is}}
Hebron, and
Zior;
nine
cities with their
villages:
\VS{55}Maon,
Carmel, and
Ziph, and
Juttah,
\VS{56}And
Jezreel, and
Jokdeam, and
Zanoah,
\VS{57}Cain,
Gibeah, and
Timnah;
ten
cities with their
villages:
\VS{58}Halhul,
Beth-zur, and
Gedor,
\VS{59}And
Maarath, and
Beth-anoth, and
Eltekon;
six
cities with their
villages:
\VS{60}Kirjath-baal, which
{\ADD{is}}
Kirjath-jearim, and
Rabbah;
two
cities with their
villages:
\VS{61}In the
wilderness,
Beth-arabah,
Middin, and
Secacah,
\VS{62}And
Nibshan, and the city of
Salt, and
En-gedi;
six
cities with their
villages.
\par }{\PP \VS{63}¶ As for the
Jebusites the
inhabitants of
Jerusalem, the
children of
Judah
could not drive them
out: but the
Jebusites
dwell with the
children of
Judah at
Jerusalem unto this
day.

\par }\Chap{16}{\PP \VerseOne{1}And the
lot of the
children of
Joseph
fell from
Jordan by
Jericho, unto the
water of
Jericho on the
east, to the
wilderness that goeth
up from
Jericho throughout
mount
Beth-el,\FTNT{*}{{\FR 16.1 }
fell: Heb. went forth}
\VS{2}And goeth
out from
Beth-el to
Luz, and passeth
along unto the
borders of
Archi to
Ataroth,
\VS{3}And goeth
down
westward to the
coast of
Japhleti, unto the
coast of
Beth-horon the
nether, and to
Gezer: and the goings
out thereof are at the
sea.
\VS{4}So the
children of
Joseph,
Manasseh and
Ephraim, took their
inheritance.
\par }{\PP \VS{5}¶ And the
border of the
children of
Ephraim according to their
families was
{\ADD{thus:}} even the
border of their
inheritance on the east
side was
Ataroth-addar, unto
Beth-horon the
upper;
\VS{6}And the
border went
out toward the
sea to
Michmethah on the north
side; and the
border went
about
eastward unto
Taanath-shiloh, and
passed by it on the
east to
Janohah;
\VS{7}And it went
down from
Janohah to
Ataroth, and to
Naarath, and
came to
Jericho, and went
out at
Jordan.
\VS{8}The
border went
out from
Tappuah
westward unto the
river
Kanah; and the goings
out thereof were at the
sea. This
{\ADD{is}} the
inheritance of the
tribe of the
children of
Ephraim by their
families.
\VS{9}And the
separate
cities for the
children of
Ephraim
{\ADD{were}}
among the
inheritance of the
children of
Manasseh, all the
cities with their
villages.
\VS{10}And they drave not
out the
Canaanites that
dwelt in
Gezer: but the
Canaanites
dwell
among the
Ephraimites unto this
day, and
serve under
tribute.

\par }\Chap{17}{\PP \VerseOne{1}There was also a
lot for the
tribe of
Manasseh; for he
{\ADD{was}} the
firstborn of
Joseph;
{\ADD{to wit}}, for
Machir the
firstborn of
Manasseh, the
father of
Gilead: because he was a
man of
war, therefore he had
Gilead and
Bashan.
\VS{2}There was also
{\ADD{a lot}} for the
rest of the
children of
Manasseh by their
families; for the
children of
Abiezer, and for the
children of
Helek, and for the
children of
Asriel, and for the
children of
Shechem, and for the
children of
Hepher, and for the
children of
Shemida: these
{\ADD{were}} the
male
children of
Manasseh the
son of
Joseph by their
families.\FTNT{*}{{\FR 17.2 }
Abiezer: also called, Jeezer}
\par }{\PP \VS{3}¶ But
Zelophehad, the
son of
Hepher, the
son of
Gilead, the
son of
Machir, the
son of
Manasseh, had no
sons, but
daughters: and these
{\ADD{are}} the
names of his
daughters,
Mahlah, and
Noah,
Hoglah,
Milcah, and
Tirzah.
\VS{4}And they came
near
before
Eleazar the
priest, and
before
Joshua the
son of
Nun, and
before the
princes,
saying, The
{\ND{
LORD}}
commanded
Moses to
give us an
inheritance
among our
brethren. Therefore according to the
commandment of the
{\ND{
LORD}} he
gave them an
inheritance
among the
brethren of their
father.
\VS{5}And there
fell
ten
portions to
Manasseh, beside the
land of
Gilead and
Bashan, which
{\ADD{were}} on the other
side
Jordan;
\VS{6}Because the
daughters of
Manasseh
had an
inheritance
among his
sons: and the
rest of
Manasseh’s
sons had the
land of
Gilead.
\par }{\PP \VS{7}¶ And the
coast of
Manasseh was from
Asher to
Michmethah, that
{\ADD{lieth}}
before
Shechem; and the
border went
along on the right
hand unto the
inhabitants of
En-tappuah.
\VS{8}{\ADD{Now}}
Manasseh had the
land of
Tappuah: but
Tappuah on the
border of
Manasseh
{\ADD{belonged}} to the
children of
Ephraim;
\VS{9}And the
coast
descended unto the
river
Kanah,
southward of the
river: these
cities of
Ephraim
{\ADD{are}}
among the
cities of
Manasseh: the
coast of
Manasseh also
{\ADD{was}} on the north
side of the
river, and the
outgoings of it were at the
sea:\FTNT{†}{{\FR 17.9 }
river Kanah: or, brook of reeds}
\VS{10}Southward
{\ADD{it was}}
Ephraim’s, and
northward
{\ADD{it was}}
Manasseh’s, and the
sea is his
border; and they met
together in
Asher on the
north, and in
Issachar on the
east.
\VS{11}And
Manasseh had in
Issachar and in
Asher
Beth-shean and her
towns, and
Ibleam and her
towns, and the
inhabitants of
Dor and her
towns, and the
inhabitants of
Endor and her
towns, and the
inhabitants of
Taanach and her
towns, and the
inhabitants of
Megiddo and her
towns,
{\ADD{even}}
three
countries.
\VS{12}Yet the
children of
Manasseh
could not drive
out
{\ADD{the inhabitants of}} those
cities; but the
Canaanites
would
dwell in that
land.
\VS{13}Yet it came to pass, when the
children of
Israel were waxen
strong, that they
put the
Canaanites to
tribute; but did not
utterly drive them
out.
\par }{\PP \VS{14}And the
children of
Joseph
spake unto
Joshua,
saying, Why hast thou
given me
{\ADD{but}}
one
lot and
one
portion to
inherit, seeing I
{\ADD{am}} a
great
people, forasmuch
as the
{\ND{
LORD}} hath
blessed me
hitherto?
\VS{15}And
Joshua
answered them, If thou
{\ADD{be}} a
great
people,
{\ADD{then}} get thee
up to the
wood
{\ADD{country}}, and cut
down for thyself there in the
land of the
Perizzites and of the
giants, if
mount
Ephraim be too
narrow for thee.\FTNT{‡}{{\FR 17.15 }
giants: or, Rephaims}
\VS{16}And the
children of
Joseph
said, The
hill is not
enough for us: and all the
Canaanites that
dwell in the
land of the
valley have
chariots of
iron,
{\ADD{both they}} who
{\ADD{are}} of
Beth-shean and her
towns, and
{\ADD{they}} who
{\ADD{are}} of the
valley of
Jezreel.
\VS{17}And
Joshua
spake unto the
house of
Joseph,
{\ADD{even}} to
Ephraim and to
Manasseh,
saying, Thou
{\ADD{art}} a
great
people, and hast
great
power: thou shalt not have
one
lot
{\ADD{only:}}
\VS{18}But the
mountain shall be thine; for it
{\ADD{is}} a
wood, and thou shalt cut it
down: and the
outgoings of it shall be thine: for thou shalt drive
out the
Canaanites, though they have
iron
chariots,
{\ADD{and}} though they
{\ADD{be}}
strong.

\par }\Chap{18}{\PP \VerseOne{1}And the whole
congregation of the
children of
Israel assembled
together at
Shiloh, and set
up the
tabernacle of the
congregation there. And the
land was
subdued
before them.
\par }{\PP \VS{2}And there
remained among the
children of
Israel
seven
tribes, which had not yet
received their
inheritance.
\VS{3}And
Joshua
said unto the
children of
Israel, How long
{\ADD{are}} ye
slack to
go to
possess the
land, which the
{\ND{
LORD}}
God of your
fathers hath
given you?
\VS{4}Give
out from among you
three
men for
{\ADD{each}}
tribe: and I will
send them, and they shall
rise, and
go through the
land, and
describe it
according to the
inheritance of them; and they shall
come
{\ADD{again}} to me.
\VS{5}And they shall
divide it into
seven
parts:
Judah shall
abide in their
coast on the
south, and the
house of
Joseph shall
abide in their
coasts on the
north.
\VS{6}Ye shall therefore
describe the
land
{\ADD{into}}
seven
parts, and
bring
{\ADD{the description}} hither to me, that I may
cast
lots for you here
before the
{\ND{
LORD}} our
God.
\VS{7}But the
Levites have no
part
among you; for the
priesthood of the
{\ND{
LORD}}
{\ADD{is}} their
inheritance: and
Gad, and
Reuben, and
half the
tribe of
Manasseh, have
received their
inheritance
beyond
Jordan on the
east, which
Moses the
servant of the
{\ND{
LORD}}
gave them.
\par }{\PP \VS{8}¶ And the
men
arose, and went
away: and
Joshua
charged them that
went to
describe the
land,
saying,
Go and
walk through the
land, and
describe it, and come
again to me, that I may here
cast
lots for you
before the
{\ND{
LORD}} in
Shiloh.
\VS{9}And the
men
went and passed
through the
land, and
described it by
cities into
seven
parts in a
book, and
came
{\ADD{again}} to
Joshua to the
host at
Shiloh.
\par }{\PP \VS{10}¶ And
Joshua
cast
lots for them in
Shiloh
before the
{\ND{
LORD}}: and there
Joshua
divided the
land unto the
children of
Israel according to their
divisions.
\par }{\PP \VS{11}¶ And the
lot of the
tribe of the
children of
Benjamin came
up according to their
families: and the
coast of their
lot came
forth between the
children of
Judah and the
children of
Joseph.
\VS{12}And their
border on the
north
side was from
Jordan; and the
border went
up to the
side of
Jericho on the
north side, and went
up through the
mountains
westward; and the goings
out thereof were at the
wilderness of
Beth-aven.
\VS{13}And the
border went
over from thence toward
Luz, to the
side of
Luz, which
{\ADD{is}}
Beth-el,
southward; and the
border
descended to
Ataroth-adar, near the
hill that
{\ADD{lieth}} on the south
side of the
nether
Beth-horon.
\VS{14}And the
border was
drawn
{\ADD{thence}}, and
compassed the
corner of the
sea
southward, from the
hill that
{\ADD{lieth}}
before
Beth-horon
southward; and the goings
out thereof were at
Kirjath-baal, which
{\ADD{is}}
Kirjath-jearim, a
city of the
children of
Judah: this
{\ADD{was}} the
west
quarter.
\VS{15}And the
south
quarter
{\ADD{was}} from the
end of
Kirjath-jearim, and the
border went
out on the
west, and went
out to the
well of
waters of
Nephtoah:
\VS{16}And the
border came
down to the
end of the
mountain that
{\ADD{lieth}}
before the
valley of the
son of
Hinnom,
{\ADD{and}} which
{\ADD{is}} in the
valley of the
giants on the
north, and
descended to the
valley of
Hinnom, to the
side of
Jebusi on the
south, and
descended to
En-rogel,
\VS{17}And was
drawn from the
north, and went
forth to
En-shemesh, and went
forth toward
Geliloth, which
{\ADD{is}} over
against the going
up of
Adummim, and
descended to the
stone of
Bohan the
son of
Reuben,
\VS{18}And passed
along toward the
side over
against
Arabah
northward, and went
down unto
Arabah:\FTNT{*}{{\FR 18.18 }
Arabah: or, the plain}
\VS{19}And the
border passed
along to the
side of
Beth-hoglah
northward: and the
outgoings of the
border were at the
north
bay of the
salt
sea at the
south
end of
Jordan: this
{\ADD{was}} the
south
coast.\FTNT{†}{{\FR 18.19 }
bay: Heb. tongue}
\VS{20}And
Jordan was the
border of it on the
east
side. This
{\ADD{was}} the
inheritance of the
children of
Benjamin, by the
coasts thereof round
about, according to their
families.
\VS{21}Now the
cities of the
tribe of the
children of
Benjamin according to their
families were
Jericho, and
Beth-hoglah, and the
valley of
Keziz,
\VS{22}And
Beth-arabah, and
Zemaraim, and
Beth-el,
\VS{23}And
Avim, and
Parah, and
Ophrah,
\VS{24}And
Chephar-haammonai, and
Ophni, and
Gaba;
twelve
cities with their
villages:
\VS{25}Gibeon, and
Ramah, and
Beeroth,
\VS{26}And
Mizpeh, and
Chephirah, and
Mozah,
\VS{27}And
Rekem, and
Irpeel, and
Taralah,
\VS{28}And
Zelah,
Eleph, and
Jebusi, which
{\ADD{is}}
Jerusalem,
Gibeath,
{\ADD{and}}
Kirjath;
fourteen
cities with their
villages. This
{\ADD{is}} the
inheritance of the
children of
Benjamin according to their
families.

\par }\Chap{19}{\PP \VerseOne{1}And the
second
lot came
forth to
Simeon,
{\ADD{even}} for the
tribe of the
children of
Simeon according to their
families: and their
inheritance was
within the
inheritance of the
children of
Judah.
\VS{2}And they had in their
inheritance
Beer-sheba, or
Sheba, and
Moladah,
\VS{3}And
Hazar-shual, and
Balah, and
Azem,
\VS{4}And
Eltolad, and
Bethul, and
Hormah,
\VS{5}And
Ziklag, and
Beth-marcaboth, and
Hazar-susah,
\VS{6}And
Beth-lebaoth, and
Sharuhen;
thirteen
cities and their
villages:
\VS{7}Ain,
Remmon, and
Ether, and
Ashan;
four
cities and their
villages:
\VS{8}And all the
villages that
{\ADD{were}} round
about these
cities to
Baalath-beer,
Ramath of the
south. This
{\ADD{is}} the
inheritance of the
tribe of the
children of
Simeon according to their
families.
\VS{9}Out of the
portion of the
children of
Judah
{\ADD{was}} the
inheritance of the
children of
Simeon: for the
part of the
children of
Judah was too
much for them: therefore the
children of
Simeon had their
inheritance
within the
inheritance of them.
\par }{\PP \VS{10}¶ And the
third
lot came
up for the
children of
Zebulun according to their
families: and the
border of their
inheritance was unto
Sarid:
\VS{11}And their
border went
up toward the
sea, and
Maralah, and
reached to
Dabbasheth, and
reached to the
river that
{\ADD{is}}
before
Jokneam;
\VS{12}And
turned from
Sarid
eastward toward the
sunrising unto the
border of
Chisloth-tabor, and then goeth
out to
Daberath, and goeth
up to
Japhia,
\VS{13}And from thence passeth on
along on the
east to
Gittah-hepher, to
Ittah-kazin, and goeth
out to
Remmon-methoar to
Neah;\FTNT{*}{{\FR 19.13 }
methoar: or, which is drawn}
\VS{14}And the
border
compasseth it on the north
side to
Hannathon: and the
outgoings thereof are in the
valley of
Jiphthah-el:
\VS{15}And
Kattath, and
Nahallal, and
Shimron, and
Idalah, and
Beth-lehem:
twelve
cities with their
villages.
\VS{16}This
{\ADD{is}} the
inheritance of the
children of
Zebulun according to their
families, these
cities with their
villages.
\par }{\PP \VS{17}¶
{\ADD{And}} the
fourth
lot came
out to
Issachar, for the
children of
Issachar according to their
families.
\VS{18}And their
border was toward
Jezreel, and
Chesulloth, and
Shunem,
\VS{19}And
Hapharaim, and
Shion, and
Anaharath,
\VS{20}And
Rabbith, and
Kishion, and
Abez,
\VS{21}And
Remeth, and
En-gannim, and
En-haddah, and
Beth-pazzez;
\VS{22}And the
coast
reacheth to
Tabor, and
Shahazimah, and
Beth-shemesh; and the
outgoings of their
border were at
Jordan:
sixteen
cities with their
villages.
\VS{23}This
{\ADD{is}} the
inheritance of the
tribe of the
children of
Issachar according to their
families, the
cities and their
villages.
\par }{\PP \VS{24}¶ And the
fifth
lot came
out for the
tribe of the
children of
Asher according to their
families.
\VS{25}And their
border was
Helkath, and
Hali, and
Beten, and
Achshaph,
\VS{26}And
Alammelech, and
Amad, and
Misheal; and
reacheth to
Carmel
westward, and to
Shihor-libnath;
\VS{27}And
turneth toward the
sunrising to
Beth-dagon, and
reacheth to
Zebulun, and to the
valley of
Jiphthah-el toward the north
side of
Beth-emek, and
Neiel, and goeth
out to
Cabul on the left
hand,
\VS{28}And
Hebron, and
Rehob, and
Hammon, and
Kanah,
{\ADD{even}} unto
great
Zidon;
\VS{29}And
{\ADD{then}} the
coast
turneth to
Ramah, and to the
strong
city
Tyre; and the
coast
turneth to
Hosah; and the
outgoings thereof are at the
sea from the
coast to
Achzib:\FTNT{†}{{\FR 19.29 }
Tyre: Heb. Tzor}
\VS{30}Ummah also, and
Aphek, and
Rehob:
twenty and
two
cities with their
villages.
\VS{31}This
{\ADD{is}} the
inheritance of the
tribe of the
children of
Asher according to their
families, these
cities with their
villages.
\par }{\PP \VS{32}¶ The
sixth
lot came
out to the
children of
Naphtali,
{\ADD{even}} for the
children of
Naphtali according to their
families.
\VS{33}And their
coast was from
Heleph, from
Allon to
Zaanannim, and
Adami,
Nekeb, and
Jabneel, unto
Lakum; and the
outgoings thereof were at
Jordan:
\VS{34}And
{\ADD{then}} the
coast
turneth
westward to
Aznoth-tabor, and goeth
out from thence to
Hukkok, and
reacheth to
Zebulun on the south
side, and
reacheth to
Asher on the west
side, and to
Judah upon
Jordan toward the
sunrising.
\VS{35}And the
fenced
cities
{\ADD{are}}
Ziddim,
Zer, and
Hammath,
Rakkath, and
Chinnereth,
\VS{36}And
Adamah, and
Ramah, and
Hazor,
\VS{37}And
Kedesh, and
Edrei, and
En-hazor,
\VS{38}And
Iron, and
Migdal-el,
Horem, and
Beth-anath, and
Beth-shemesh;
nineteen
cities with their
villages.
\VS{39}This
{\ADD{is}} the
inheritance of the
tribe of the
children of
Naphtali according to their
families, the
cities and their
villages.
\par }{\PP \VS{40}¶
{\ADD{And}} the
seventh
lot came
out for the
tribe of the
children of
Dan according to their
families.
\VS{41}And the
coast of their
inheritance was
Zorah, and
Eshtaol, and
Ir-shemesh,
\VS{42}And
Shaalabbin, and
Ajalon, and
Jethlah,
\VS{43}And
Elon, and
Thimnathah, and
Ekron,
\VS{44}And
Eltekeh, and
Gibbethon, and
Baalath,
\VS{45}And
Jehud, and
Bene-berak, and
Gath-rimmon,
\VS{46}And
Me-jarkon, and
Rakkon, with the
border
before
Japho.\FTNT{‡}{{\FR 19.46 }
before: or, over against}\FTNT{§}{{\FR 19.46 }
Japho: or, Joppa}
\VS{47}And the
coast of the
children of
Dan went
out
{\ADD{too little}} for them: therefore the
children of
Dan went
up to
fight against
Leshem, and
took it, and
smote it with the
edge of the
sword, and
possessed it, and
dwelt therein, and
called
Leshem,
Dan, after the
name of
Dan their
father.
\VS{48}This
{\ADD{is}} the
inheritance of the
tribe of the
children of
Dan according to their
families, these
cities with their
villages.
\par }{\PP \VS{49}¶ When they had made an
end of dividing the
land for
inheritance by their
coasts, the
children of
Israel
gave an
inheritance to
Joshua the
son of
Nun
among them:
\VS{50}According to the
word of the
{\ND{
LORD}} they
gave him the
city which he
asked,
{\ADD{even}}
Timnath-serah in
mount
Ephraim: and he
built the
city, and
dwelt therein.
\VS{51}These
{\ADD{are}} the
inheritances, which
Eleazar the
priest, and
Joshua the
son of
Nun, and the
heads of the
fathers of the
tribes of the
children of
Israel, divided for an
inheritance by
lot in
Shiloh
before the
{\ND{
LORD}}, at the
door of the
tabernacle of the
congregation. So they made an
end of
dividing the
country.

\par }\Chap{20}{\PP \VerseOne{1}The
{\ND{
LORD}} also
spake unto
Joshua,
saying,
\VS{2}Speak to the
children of
Israel,
saying, Appoint
out for you
cities of
refuge, whereof I
spake unto you by the
hand of
Moses:
\VS{3}That the
slayer that
killeth
{\ADD{any}}
person
unawares
{\ADD{and}}
unwittingly may
flee thither: and they shall be your
refuge from the
avenger of
blood.
\VS{4}And when he that doth
flee unto
one of those
cities shall
stand at the
entering of the
gate of the
city, and shall
declare his
cause in the
ears of the
elders of that
city, they shall
take him into the
city unto them, and
give him a
place, that he may
dwell among them.
\VS{5}And if the
avenger of
blood
pursue
after him, then they shall not
deliver the
slayer up into his
hand; because he
smote his
neighbour
unwittingly, and
hated him not
beforetime.
\VS{6}And he shall
dwell in that
city, until he
stand
before the
congregation for
judgment,
{\ADD{and}} until the
death of the
high
priest
that shall be in those
days: then shall the
slayer
return, and
come unto his own
city, and unto his own
house, unto the
city from whence he
fled.
\par }{\PP \VS{7}¶ And they
appointed
Kedesh in
Galilee in
mount
Naphtali, and
Shechem in
mount
Ephraim, and
Kirjath-arba, which
{\ADD{is}}
Hebron, in the
mountain of
Judah.\FTNT{*}{{\FR 20.7 }
appointed: Heb. sanctified}
\VS{8}And on the other
side
Jordan by
Jericho
eastward, they
assigned
Bezer in the
wilderness upon the
plain out of the
tribe of
Reuben, and
Ramoth in
Gilead out of the
tribe of
Gad, and
Golan in
Bashan out of the
tribe of
Manasseh.
\VS{9}These were the
cities
appointed for all the
children of
Israel, and for the
stranger that
sojourneth
among them, that whosoever
killeth
{\ADD{any}}
person at
unawares might
flee thither, and not
die by the
hand of the
avenger of
blood, until he
stood
before the
congregation.

\par }\Chap{21}{\PP \VerseOne{1}Then came
near the
heads of the
fathers of the
Levites unto
Eleazar the
priest, and unto
Joshua the
son of
Nun, and unto the
heads of the
fathers of the
tribes of the
children of
Israel;
\VS{2}And they
spake unto them at
Shiloh in the
land of
Canaan,
saying, The
{\ND{
LORD}}
commanded by the
hand of
Moses to
give us
cities to dwell
in, with the
suburbs thereof for our
cattle.
\VS{3}And the
children of
Israel
gave unto the
Levites out of their
inheritance, at the
commandment of the
{\ND{
LORD}}, these
cities and their
suburbs.
\VS{4}And the
lot came
out for the
families of the
Kohathites: and the
children of
Aaron the
priest,
{\ADD{which were}} of the
Levites, had by
lot out of the
tribe of
Judah, and out of the
tribe of
Simeon, and out of the
tribe of
Benjamin,
thirteen
cities.
\VS{5}And the
rest of the
children of
Kohath
{\ADD{had}} by
lot out of the
families of the
tribe of
Ephraim, and out of the
tribe of
Dan, and out of the
half
tribe of
Manasseh,
ten
cities.
\VS{6}And the
children of
Gershon
{\ADD{had}} by
lot out of the
families of the
tribe of
Issachar, and out of the
tribe of
Asher, and out of the
tribe of
Naphtali, and out of the
half
tribe of
Manasseh in
Bashan,
thirteen
cities.
\VS{7}The
children of
Merari by their
families
{\ADD{had}} out of the
tribe of
Reuben, and out of the
tribe of
Gad, and out of the
tribe of
Zebulun,
twelve
cities.
\VS{8}And the
children of
Israel
gave by
lot unto the
Levites these
cities with their
suburbs, as the
{\ND{
LORD}}
commanded by the
hand of
Moses.
\par }{\PP \VS{9}¶ And they
gave out of the
tribe of the
children of
Judah, and out of the
tribe of the
children of
Simeon, these
cities which are
{\ADD{here}}
mentioned by
name,\FTNT{*}{{\FR 21.9 }
mentioned: Heb. called}
\VS{10}Which the
children of
Aaron,
{\ADD{being}} of the
families of the
Kohathites,
{\ADD{who were}} of the
children of
Levi, had: for theirs was the
first
lot.
\VS{11}And they
gave them the
city of
Arba the
father of
Anak, which
{\ADD{city is}}
Hebron, in the
hill
{\ADD{country}} of
Judah, with the
suburbs thereof round
about it.\FTNT{†}{{\FR 21.11 }
the city…: or, Kirjath-arba}
\VS{12}But the
fields of the
city, and the
villages thereof,
gave they to
Caleb the
son of
Jephunneh for his
possession.
\par }{\PP \VS{13}¶ Thus they
gave to the
children of
Aaron the
priest
Hebron with her
suburbs,
{\ADD{to be}} a
city of
refuge for the
slayer; and
Libnah with her
suburbs,
\VS{14}And
Jattir with her
suburbs, and
Eshtemoa with her
suburbs,
\VS{15}And
Holon with her
suburbs, and
Debir with her
suburbs,\FTNT{‡}{{\FR 21.15 }
Holon: also called, Hilen}
\VS{16}And
Ain with her
suburbs, and
Juttah with her
suburbs,
{\ADD{and}}
Beth-shemesh with her
suburbs;
nine
cities out of those
two
tribes.\FTNT{§}{{\FR 21.16 }
Ain: also called, Ashan}
\VS{17}And out of the
tribe of
Benjamin,
Gibeon with her
suburbs,
Geba with her
suburbs,\FTNT{**}{{\FR 21.17 }
Geba: also called, Gaba}
\VS{18}Anathoth with her
suburbs, and
Almon with her
suburbs;
four
cities.\FTNT{††}{{\FR 21.18 }
Almon: also called, Alemeth}
\VS{19}All the
cities of the
children of
Aaron, the
priests,
{\ADD{were}}
thirteen
cities with their
suburbs.
\par }{\PP \VS{20}¶ And the
families of the
children of
Kohath, the
Levites which
remained of the
children of
Kohath, even they had the
cities of their
lot out of the
tribe of
Ephraim.
\VS{21}For they
gave them
Shechem with her
suburbs in
mount
Ephraim,
{\ADD{to be}} a
city of
refuge for the
slayer; and
Gezer with her
suburbs,
\VS{22}And
Kibzaim with her
suburbs, and
Beth-horon with her
suburbs;
four
cities.
\VS{23}And out of the
tribe of
Dan,
Eltekeh with her
suburbs,
Gibbethon with her
suburbs,
\VS{24}Aijalon with her
suburbs,
Gath-rimmon with her
suburbs;
four
cities.
\VS{25}And out of the
half
tribe of
Manasseh,
Tanach with her
suburbs, and
Gath-rimmon with her
suburbs;
two
cities.
\VS{26}All the
cities
{\ADD{were}}
ten with their
suburbs for the
families of the
children of
Kohath that
remained.
\par }{\PP \VS{27}¶ And unto the
children of
Gershon, of the
families of the
Levites, out of the
{\ADD{other}}
half
tribe of
Manasseh
{\ADD{they gave}}
Golan in
Bashan with her
suburbs,
{\ADD{to be}} a
city of
refuge for the
slayer; and
Beesh-terah with her
suburbs;
two
cities.
\VS{28}And out of the
tribe of
Issachar,
Kishon with her
suburbs,
Dabareh with her
suburbs,
\VS{29}Jarmuth with her
suburbs,
En-gannim with her
suburbs;
four
cities.
\VS{30}And out of the
tribe of
Asher,
Mishal with her
suburbs,
Abdon with her
suburbs,
\VS{31}Helkath with her
suburbs, and
Rehob with her
suburbs;
four
cities.
\VS{32}And out of the
tribe of
Naphtali,
Kedesh in
Galilee with her
suburbs,
{\ADD{to be}} a
city of
refuge for the
slayer; and
Hammoth-dor with her
suburbs, and
Kartan with her
suburbs;
three
cities.
\VS{33}All the
cities of the
Gershonites according to their
families
{\ADD{were}}
thirteen
cities with their
suburbs.
\par }{\PP \VS{34}¶ And unto the
families of the
children of
Merari, the
rest of the
Levites, out of the
tribe of
Zebulun,
Jokneam with her
suburbs, and
Kartah with her
suburbs,
\VS{35}Dimnah with her
suburbs,
Nahalal with her
suburbs;
four
cities.
\VS{36}And out of the tribe of
Reuben,
Bezer with her suburbs, and
Jahazah with her suburbs,
\VS{37}Kedemoth with her suburbs, and
Mephaath with her suburbs;
four
cities.
\VS{38}And out of the
tribe of
Gad,
Ramoth in
Gilead with her
suburbs,
{\ADD{to be}} a
city of
refuge for the
slayer; and
Mahanaim with her
suburbs,
\VS{39}Heshbon with her
suburbs,
Jazer with her
suburbs;
four
cities in all.
\VS{40}So all the
cities for the
children of
Merari by their
families, which were
remaining of the
families of the
Levites, were
{\ADD{by}} their
lot
twelve
cities.
\VS{41}All the
cities of the
Levites
within the
possession of the
children of
Israel
{\ADD{were}}
forty and
eight
cities with their
suburbs.
\VS{42}These
cities were every
one with their
suburbs round
about them: thus
{\ADD{were}} all these
cities.
\par }{\PP \VS{43}¶ And the
{\ND{
LORD}}
gave unto
Israel all the
land which he
sware to
give unto their
fathers; and they
possessed it, and
dwelt therein.
\VS{44}And the
{\ND{
LORD}} gave them
rest round
about, according to all that he
sware unto their
fathers: and there
stood not a
man of all their
enemies
before them; the
{\ND{
LORD}}
delivered all their
enemies into their
hand.
\VS{45}There
failed not
ought of any
good
thing which the
{\ND{
LORD}} had
spoken unto the
house of
Israel; all came to
pass.

\par }\Chap{22}{\PP \VerseOne{1}Then
Joshua
called the
Reubenites, and the
Gadites, and the
half
tribe of
Manasseh,
\VS{2}And
said unto them, Ye have
kept all that
Moses the
servant of the
{\ND{
LORD}}
commanded you, and have
obeyed my
voice in all that I
commanded you:
\VS{3}Ye have not
left your
brethren these
many
days unto this
day, but have
kept the
charge of the
commandment of the
{\ND{
LORD}} your
God.
\VS{4}And now the
{\ND{
LORD}} your
God hath given
rest unto your
brethren, as he
promised them: therefore now
return ye, and
get you unto your
tents,
{\ADD{and}} unto the
land of your
possession, which
Moses the
servant of the
{\ND{
LORD}}
gave you on the other
side
Jordan.
\VS{5}But take
diligent
heed to
do the
commandment and the
law, which
Moses the
servant of the
{\ND{
LORD}}
charged you, to
love the
{\ND{
LORD}} your
God, and to
walk in all his
ways, and to
keep his
commandments, and to
cleave unto him, and to
serve him with all your
heart and with all your
soul.
\VS{6}So
Joshua
blessed them, and sent them
away: and they
went unto their
tents.
\par }{\PP \VS{7}¶ Now to the
{\ADD{one}}
half of the
tribe of
Manasseh
Moses had
given
{\ADD{possession}} in
Bashan: but unto the
{\ADD{other}}
half thereof
gave
Joshua
among their
brethren on this
side
Jordan
westward. And when
Joshua sent them
away also unto their
tents, then he
blessed them,
\VS{8}And he
spake unto them,
saying,
Return with
much
riches unto your
tents, and with
very
much
cattle, with
silver, and with
gold, and with
brass, and with
iron, and with
very
much
raiment:
divide the
spoil of your
enemies with your
brethren.
\par }{\PP \VS{9}¶ And the
children of
Reuben and the
children of
Gad and the
half
tribe of
Manasseh
returned, and
departed from the
children of
Israel out of
Shiloh, which
{\ADD{is}} in the
land of
Canaan, to
go unto the
country of
Gilead, to the
land of their
possession, whereof they were
possessed, according to the
word of the
{\ND{
LORD}} by the
hand of
Moses.
\par }{\PP \VS{10}¶ And when they
came unto the
borders of
Jordan, that
{\ADD{are}} in the
land of
Canaan, the
children of
Reuben and the
children of
Gad and the
half
tribe of
Manasseh
built there an
altar by
Jordan, a
great
altar to see
to.
\par }{\PP \VS{11}¶ And the
children of
Israel
heard
say, Behold, the
children of
Reuben and the
children of
Gad and the
half
tribe of
Manasseh have
built an
altar over
against the
land of
Canaan, in the
borders of
Jordan, at the
passage of the
children of
Israel.
\VS{12}And when the
children of
Israel
heard
{\ADD{of it}}, the whole
congregation of the
children of
Israel gathered themselves
together at
Shiloh, to go
up to
war against them.
\VS{13}And the
children of
Israel
sent unto the
children of
Reuben, and to the
children of
Gad, and to the
half
tribe of
Manasseh, into the
land of
Gilead,
Phinehas the
son of
Eleazar the
priest,
\VS{14}And with him
ten
princes, of
each
chief
house a
prince throughout all the
tribes of
Israel; and each
one
{\ADD{was}} an
head of the
house of their
fathers among the
thousands of
Israel.\FTNT{*}{{\FR 22.14 }
chief…: Heb. house of the father}
\par }{\PP \VS{15}¶ And they
came unto the
children of
Reuben, and to the
children of
Gad, and to the
half
tribe of
Manasseh, unto the
land of
Gilead, and they
spake with them,
saying,
\VS{16}Thus
saith the whole
congregation of the
{\ND{
LORD}}, What
trespass
{\ADD{is}} this that ye have
committed against the
God of
Israel, to turn
away this
day from
following the
{\ND{
LORD}}, in that ye have
builded you an
altar, that ye might
rebel this
day against the
{\ND{
LORD}}?
\VS{17}{\ADD{Is}} the
iniquity of
Peor too
little for us, from which we are not
cleansed until this
day, although there was a
plague in the
congregation of the
{\ND{
LORD}},
\VS{18}But that ye must turn
away this
day from
following the
{\ND{
LORD}}? and it will be,
{\ADD{seeing}} ye
rebel to
day against the
{\ND{
LORD}}, that to
morrow he will be
wroth with the whole
congregation of
Israel.
\VS{19}Notwithstanding, if the
land of your
possession
{\ADD{be}}
unclean,
{\ADD{then}} pass ye
over unto the
land of the
possession of the
{\ND{
LORD}}, wherein the
{\ND{
LORD’s}}
tabernacle
dwelleth, and take
possession
among us: but
rebel not against the
{\ND{
LORD}},
nor
rebel against us, in
building you an
altar
beside the
altar of the
{\ND{
LORD}} our
God.
\VS{20}Did not
Achan the
son of
Zerah
commit a
trespass in the accursed
thing, and
wrath
fell on all the
congregation of
Israel? and that
man
perished not
alone in his
iniquity.
\par }{\PP \VS{21}¶ Then the
children of
Reuben and the
children of
Gad and the
half
tribe of
Manasseh
answered, and
said unto the
heads of the
thousands of
Israel,
\VS{22}The
{\ND{
LORD}}
God of
gods, the
{\ND{
LORD}}
God of
gods, he
knoweth, and
Israel he shall
know; if
{\ADD{it be}} in
rebellion, or if in
transgression against the
{\ND{
LORD}}, (save us not this
day,)
\VS{23}That we have
built us an
altar to
turn from
following the
{\ND{
LORD}}, or if to
offer thereon burnt
offering or meat
offering, or if to
offer
peace
offerings thereon, let the
{\ND{
LORD}} himself
require
{\ADD{it;}}
\VS{24}And if we have not
{\ADD{rather}}
done it for
fear of
{\ADD{this}}
thing,
saying, In time to
come your
children might
speak unto our
children,
saying, What have ye to do with the
{\ND{
LORD}}
God of
Israel?\FTNT{†}{{\FR 22.24 }
In time…: Heb. To morrow}
\VS{25}For the
{\ND{
LORD}} hath
made
Jordan a
border between us and you, ye
children of
Reuben and
children of
Gad; ye have no
part in the
{\ND{
LORD}}: so shall your
children
make our
children
cease
from
fearing the
{\ND{
LORD}}.
\VS{26}Therefore we
said, Let us now
prepare to
build us an
altar, not for burnt
offering, nor for
sacrifice:
\VS{27}But
{\ADD{that}} it
{\ADD{may be}} a
witness between us, and you, and our
generations
after us, that we might
do the
service of the
{\ND{
LORD}}
before him with our burnt
offerings, and with our
sacrifices, and with our peace
offerings; that your
children may not
say to our
children in time to
come, Ye have no
part in the
{\ND{
LORD}}.
\VS{28}Therefore
said we, that it shall be, when they should
{\ADD{so}}
say to us or to our
generations in time to
come, that we may
say
{\ADD{again}},
Behold the
pattern of the
altar of the
{\ND{
LORD}}, which our
fathers
made, not for burnt
offerings, nor for
sacrifices; but it
{\ADD{is}} a
witness between us and you.
\VS{29}God
forbid that we should
rebel against the
{\ND{
LORD}}, and
turn this
day from
following the
{\ND{
LORD}}, to
build an
altar for burnt
offerings, for meat
offerings, or for
sacrifices, beside the
altar of the
{\ND{
LORD}} our
God that
{\ADD{is}}
before his
tabernacle.
\par }{\PP \VS{30}¶ And when
Phinehas the
priest, and the
princes of the
congregation and
heads of the
thousands of
Israel which
{\ADD{were}} with him,
heard the
words that the
children of
Reuben and the
children of
Gad and the
children of
Manasseh
spake, it
pleased them.\FTNT{‡}{{\FR 22.30 }
pleased…: Heb. was good in their eyes}
\VS{31}And
Phinehas the
son of
Eleazar the
priest
said unto the
children of
Reuben, and to the
children of
Gad, and to the
children of
Manasseh, This
day we
perceive that the
{\ND{
LORD}}
{\ADD{is}}
among us, because ye have not
committed this
trespass against the
{\ND{
LORD}}:
now ye have
delivered the
children of
Israel out of the
hand of the
{\ND{
LORD}}.\FTNT{§}{{\FR 22.31 }
now: Heb. then}
\par }{\PP \VS{32}¶ And
Phinehas the
son of
Eleazar the
priest, and the
princes,
returned from the
children of
Reuben, and from the
children of
Gad, out of the
land of
Gilead, unto the
land of
Canaan, to the
children of
Israel, and
brought them
word
again.
\VS{33}And the
thing
pleased the
children of
Israel; and the
children of
Israel
blessed
God, and did not
intend to go
up against them in
battle, to
destroy the
land wherein the
children of
Reuben and
Gad
dwelt.
\VS{34}And the
children of
Reuben and the
children of
Gad
called the
altar
{\ADD{Ed:}} for it
{\ADD{shall be}} a
witness between us that the
{\ND{
LORD}}
{\ADD{is}}
God.\FTNT{**}{{\FR 22.34 }
Ed: that is, A witness}

\par }\Chap{23}{\PP \VerseOne{1}And it came to pass a
long
time
after that the
{\ND{
LORD}} had given
rest unto
Israel from all their
enemies round
about, that
Joshua waxed
old
{\ADD{and}}
stricken in
age.\FTNT{*}{{\FR 23.1 }
stricken…: Heb. come into days}
\VS{2}And
Joshua
called for all
Israel,
{\ADD{and}} for their
elders, and for their
heads, and for their
judges, and for their
officers, and
said unto them, I am
old
{\ADD{and}}
stricken in
age:
\VS{3}And ye have
seen all that the
{\ND{
LORD}} your
God hath
done unto all these
nations
because of you; for the
{\ND{
LORD}} your
God
{\ADD{is}} he that hath
fought for you.
\VS{4}Behold, I have
divided unto you by lot these
nations that
remain, to be an
inheritance for your
tribes, from
Jordan, with all the
nations that I have cut
off, even unto the
great
sea
westward.\FTNT{†}{{\FR 23.4 }
westward: Heb. at the sunset}
\VS{5}And the
{\ND{
LORD}} your
God, he shall
expel them from
before you, and
drive them from out of your
sight; and ye shall
possess their
land, as the
{\ND{
LORD}} your
God hath
promised unto you.
\VS{6}Be ye therefore
very
courageous to
keep and to
do all that is
written in the
book of the
law of
Moses, that ye
turn not aside therefrom
{\ADD{to}} the right
hand or
{\ADD{to}} the
left;
\VS{7}That ye
come not among these
nations, these that
remain among you; neither make
mention of the
name of their
gods, nor cause to
swear
{\ADD{by them}}, neither
serve them, nor
bow yourselves unto them:
\VS{8}But
cleave unto the
{\ND{
LORD}} your
God, as ye have
done unto this
day.\FTNT{‡}{{\FR 23.8 }
But cleave…: or, For if ye will cleave, etc}
\VS{9}For the
{\ND{
LORD}} hath driven
out from
before you
great
nations and
strong: but
{\ADD{as for}} you, no
man hath been able to
stand
before you unto this
day.\FTNT{§}{{\FR 23.9 }
For the
{\ND{LORD}}…: or, Then the
{\ND{LORD}} will drive}
\VS{10}One
man of you shall
chase a
thousand: for the
{\ND{
LORD}} your
God, he
{\ADD{it is}} that
fighteth for you, as he hath
promised you.
\par }{\PP \VS{11}Take
good
heed therefore unto
yourselves, that ye
love the
{\ND{
LORD}} your
God.\FTNT{**}{{\FR 23.11 }
yourselves: Heb. your souls}
\VS{12}Else if ye do in any
wise go
back, and
cleave unto the
remnant of these
nations,
{\ADD{even}} these that
remain among you, and shall make
marriages with them, and go
in unto them, and they to you:
\VS{13}Know for a
certainty that the
{\ND{
LORD}} your
God will no
more drive
out
{\ADD{any of}} these
nations from
before you; but they shall be
snares and
traps unto you, and
scourges in your
sides, and
thorns in your
eyes, until ye
perish from off this
good
land which the
{\ND{
LORD}} your
God hath
given you.
\VS{14}And, behold, this
day I
{\ADD{am}}
going the
way of all the
earth: and ye
know in all your
hearts and in all your
souls, that not
one
thing hath
failed of all the
good
things which the
{\ND{
LORD}} your
God
spake concerning you; all are come to
pass unto you,
{\ADD{and}} not
one
thing hath
failed thereof.
\VS{15}Therefore it shall come to pass,
{\ADD{that}} as all
good
things are
come upon you, which the
{\ND{
LORD}} your
God
promised you; so shall the
{\ND{
LORD}}
bring upon you all
evil
things, until he have
destroyed you from off this
good
land which the
{\ND{
LORD}} your
God hath
given you.
\VS{16}When ye have
transgressed the
covenant of the
{\ND{
LORD}} your
God, which he
commanded you, and have
gone and
served
other
gods, and
bowed yourselves to them; then shall the
anger of the
{\ND{
LORD}} be
kindled against you, and ye shall
perish
quickly from off the
good
land which he hath
given unto you.

\par }\Chap{24}{\PP \VerseOne{1}And
Joshua
gathered all the
tribes of
Israel to
Shechem, and
called for the
elders of
Israel, and for their
heads, and for their
judges, and for their
officers; and they
presented themselves
before
God.
\VS{2}And
Joshua
said unto all the
people, Thus
saith the
{\ND{
LORD}}
God of
Israel, Your
fathers
dwelt on the other
side of the
flood in old
time,
{\ADD{even}}
Terah, the
father of
Abraham, and the
father of
Nachor: and they
served
other
gods.
\VS{3}And I
took your
father
Abraham from the other
side of the
flood, and
led him throughout all the
land of
Canaan, and
multiplied his
seed, and
gave him
Isaac.
\VS{4}And I
gave unto
Isaac
Jacob and
Esau: and I
gave unto
Esau
mount
Seir, to
possess it; but
Jacob and his
children went
down into
Egypt.
\VS{5}I
sent
Moses also and
Aaron, and I
plagued
Egypt, according to that
which I
did
among them: and
afterward I brought you
out.
\VS{6}And I
brought your
fathers out of
Egypt: and ye
came unto the
sea; and the
Egyptians
pursued
after your
fathers with
chariots and
horsemen unto the
Red
sea.
\VS{7}And when they
cried unto the
{\ND{
LORD}}, he
put
darkness between you and the
Egyptians, and
brought the
sea upon them, and
covered them; and your
eyes have
seen what I have
done in
Egypt: and ye
dwelt in the
wilderness a
long
season.
\VS{8}And I
brought you into the
land of the
Amorites, which
dwelt on the other
side
Jordan; and they
fought with you: and I
gave them into your
hand, that ye might
possess their
land; and I
destroyed them from
before you.
\VS{9}Then
Balak the
son of
Zippor,
king of
Moab,
arose and
warred against
Israel, and
sent and
called
Balaam the
son of
Beor to
curse you:
\VS{10}But I
would not
hearken unto
Balaam; therefore he
blessed you
still: so I
delivered you out of his
hand.
\VS{11}And ye went
over
Jordan, and
came unto
Jericho: and the
men of
Jericho
fought against you, the
Amorites, and the
Perizzites, and the
Canaanites, and the
Hittites, and the
Girgashites, the
Hivites, and the
Jebusites; and I
delivered them into your
hand.
\VS{12}And I
sent the
hornet
before you, which drave them
out from
before you,
{\ADD{even}} the
two
kings of the
Amorites;
{\ADD{but}} not with thy
sword, nor with thy
bow.
\VS{13}And I have
given you a
land for which ye did not
labour, and
cities which ye
built not, and ye
dwell in them; of the
vineyards and
oliveyards which ye
planted not do ye
eat.
\par }{\PP \VS{14}¶ Now therefore
fear the
{\ND{
LORD}}, and
serve him in
sincerity and in
truth: and put
away the
gods which your
fathers
served on the other
side of the
flood, and in
Egypt; and
serve ye the
{\ND{
LORD}}.
\par }{\PP \VS{15}And if it seem
evil unto
you to
serve the
{\ND{
LORD}},
choose you this
day
whom ye will
serve; whether the
gods which your
fathers
served that
{\ADD{were}} on the other
side of the
flood, or the
gods of the
Amorites, in whose
land ye
dwell: but as for me and my
house, we will
serve the
{\ND{
LORD}}.
\VS{16}And the
people
answered and
said, God
forbid that we should
forsake the
{\ND{
LORD}}, to
serve
other
gods;
\VS{17}For the
{\ND{
LORD}} our
God, he
{\ADD{it is}} that brought us
up and our
fathers out of the
land of
Egypt, from the
house of
bondage, and which
did those
great
signs in our
sight, and
preserved us in all the
way wherein we
went, and among all the
people
through whom we
passed:
\VS{18}And the
{\ND{
LORD}} drave
out from
before us all the
people, even the
Amorites which
dwelt in the
land:
{\ADD{therefore}} will we also
serve the
{\ND{
LORD}}; for he
{\ADD{is}} our
God.
\VS{19}And
Joshua
said unto the
people, Ye
cannot
serve the
{\ND{
LORD}}: for he
{\ADD{is}} an
holy
God; he
{\ADD{is}} a
jealous
God; he will not
forgive your
transgressions nor your
sins.\FTNT{*}{{\FR 24.19 }
holy God: Heb. holy Gods}
\VS{20}If ye
forsake the
{\ND{
LORD}}, and
serve
strange
gods, then he will
turn and do you
hurt, and
consume you,
after that he hath done you
good.
\VS{21}And the
people
said unto
Joshua, Nay; but we will
serve the
{\ND{
LORD}}.
\VS{22}And
Joshua
said unto the
people, Ye
{\ADD{are}}
witnesses against yourselves that ye have
chosen you the
{\ND{
LORD}}, to
serve him. And they
said,
{\ADD{We are}}
witnesses.
\VS{23}Now therefore put
away,
{\ADD{said he}}, the
strange
gods which
{\ADD{are}}
among you, and
incline your
heart unto the
{\ND{
LORD}}
God of
Israel.
\VS{24}And the
people
said unto
Joshua, The
{\ND{
LORD}} our
God will we
serve, and his
voice will we
obey.
\VS{25}So
Joshua
made a
covenant with the
people that
day, and
set them a
statute and an
ordinance in
Shechem.
\par }{\PP \VS{26}¶ And
Joshua
wrote these
words in the
book of the
law of
God, and
took a
great
stone, and set it
up there under an
oak, that
{\ADD{was}} by the
sanctuary of the
{\ND{
LORD}}.
\VS{27}And
Joshua
said unto all the
people, Behold, this
stone shall be a
witness unto us; for it hath
heard all the
words of the
{\ND{
LORD}} which he
spake unto us: it shall be therefore a
witness unto you, lest ye
deny your
God.
\VS{28}So
Joshua let the
people
depart, every
man unto his
inheritance.
\par }{\PP \VS{29}¶ And it came to pass
after these
things, that
Joshua the
son of
Nun, the
servant of the
{\ND{
LORD}},
died,
{\ADD{being}} an
hundred and
ten
years
old.
\VS{30}And they
buried him in the
border of his
inheritance in
Timnath-serah, which
{\ADD{is}} in
mount
Ephraim, on the north
side of the
hill of
Gaash.
\VS{31}And
Israel
served the
{\ND{
LORD}} all the
days of
Joshua, and all the
days of the
elders that
overlived
Joshua, and which had
known all the
works of the
{\ND{
LORD}}, that he had
done for
Israel.\FTNT{†}{{\FR 24.31 }
overlived…: Heb. prolonged their days after Joshua}
\par }{\PP \VS{32}¶ And the
bones of
Joseph, which the
children of
Israel brought
up out of
Egypt,
buried they in
Shechem, in a
parcel of
ground which
Jacob
bought of the
sons of
Hamor the
father of
Shechem for an
hundred pieces of
silver: and it became the
inheritance of the
children of
Joseph.\FTNT{‡}{{\FR 24.32 }
pieces…: or, lambs}
\VS{33}And
Eleazar the
son of
Aaron
died; and they
buried him in a
hill
{\ADD{that pertained to}}
Phinehas his
son, which was
given him in
mount
Ephraim.
\par }