\NormalFont\ShortTitle{Lamentations}
{\MT The Lamentations of Jeremiah

\par }\ChapOne{1}{\PP \VerseOne{1}How doth the
city
sit
solitary,
{\ADD{that was}}
full of
people!
{\ADD{how}} is she become as a
widow! she
{\ADD{that was}}
great among the
nations,
{\ADD{and}}
princess among the
provinces,
{\ADD{how}} is she become
tributary!
\VS{2}She
weepeth
sore in the
night, and her
tears
{\ADD{are}} on her
cheeks: among all her
lovers she hath none to
comfort
{\ADD{her:}} all her
friends have dealt
treacherously with her, they are become her
enemies.
\VS{3}Judah is gone into
captivity because of
affliction, and because of
great
servitude: she
dwelleth among the
heathen, she
findeth no
rest: all her
persecutors
overtook her between the
straits.\FTNT{*}{{\FR 1.3 }
because of great…: Heb. for the greatness of servitude}
\VS{4}The
ways of
Zion do
mourn, because none
come to the solemn
feasts: all her
gates are
desolate: her
priests
sigh, her
virgins are
afflicted, and she
{\ADD{is}} in
bitterness.
\VS{5}Her
adversaries are the
chief, her
enemies
prosper; for the
{\ND{
LORD}} hath
afflicted her for the
multitude of her
transgressions: her
children are
gone into
captivity
before the
enemy.
\VS{6}And from the
daughter of
Zion all her
beauty is
departed: her
princes are become like
harts
{\ADD{that}}
find no
pasture, and they are
gone without
strength
before the
pursuer.
\VS{7}Jerusalem
remembered in the
days of her
affliction and of her
miseries all her pleasant
things that she had in the
days of
old, when her
people
fell into the
hand of the
enemy, and none did
help her: the
adversaries
saw her,
{\ADD{and}} did
mock at her
sabbaths.\FTNT{†}{{\FR 1.7 }
pleasant: or, desirable}
\VS{8}Jerusalem hath
grievously
sinned; therefore she is
removed: all that
honoured her
despise her, because they have
seen her
nakedness: yea, she
sigheth, and
turneth
backward.\FTNT{‡}{{\FR 1.8 }
is…: Heb. is become a removing, or, wandering}
\VS{9}Her
filthiness
{\ADD{is}} in her
skirts; she
remembereth not her last
end; therefore she came
down
wonderfully: she had no
comforter. O
{\ND{
LORD}},
behold my
affliction: for the
enemy hath
magnified
{\ADD{himself}}.
\VS{10}The
adversary hath spread
out his
hand upon all her pleasant
things: for she hath
seen
{\ADD{that}} the
heathen
entered into her
sanctuary, whom thou didst
command
{\ADD{that}} they should not
enter into thy
congregation.\FTNT{§}{{\FR 1.10 }
pleasant: or, desirable}
\VS{11}All her
people
sigh, they
seek
bread; they have
given their pleasant
things for
meat to
relieve the
soul:
see, O
{\ND{
LORD}}, and
consider; for I am become
vile.\FTNT{**}{{\FR 1.11 }
to…: or, to make the soul to come again}
\par }{\PP \VS{12}¶
{\ADD{Is it}} nothing to you, all ye that
pass
by?
behold, and
see if there
be any
sorrow like unto my
sorrow, which is
done unto me, wherewith the
{\ND{
LORD}} hath
afflicted
{\ADD{me}} in the
day of his
fierce
anger.\FTNT{††}{{\FR 1.12 }
Is it…: or, It is nothing}\FTNT{‡‡}{{\FR 1.12 }
pass by: Heb. pass by the way?}
\VS{13}From
above hath he
sent
fire into my
bones, and it
prevaileth against them: he hath
spread a
net for my
feet, he hath
turned me
back: he hath
made me
desolate
{\ADD{and}}
faint all the
day.
\VS{14}The
yoke of my
transgressions is
bound by his
hand: they are
wreathed,
{\ADD{and}} come
up upon my
neck: he hath made my
strength to
fall, the
Lord hath
delivered me into
{\ADD{their}}
hands,
{\ADD{from whom}} I am not
able to rise
up.
\VS{15}The
Lord hath trodden under
foot all my
mighty
{\ADD{men}} in the
midst of me: he hath
called an
assembly against me to
crush my young
men: the
Lord hath
trodden the
virgin, the
daughter of
Judah,
{\ADD{as}} in a
winepress.\FTNT{§§}{{\FR 1.15 }
the virgin…: or, the winepress of the virgin, etc}
\VS{16}For these
{\ADD{things}} I
weep; mine
eye, mine
eye runneth
down with
water, because the
comforter that should
relieve my
soul is
far from me: my
children are
desolate, because the
enemy
prevailed.\FTNT{***}{{\FR 1.16 }
relieve: Heb. bring back}
\VS{17}Zion spreadeth
forth her
hands,
{\ADD{and there is}} none to
comfort her: the
{\ND{
LORD}} hath
commanded concerning
Jacob,
{\ADD{that}} his
adversaries
{\ADD{should be}} round
about him:
Jerusalem is as a menstruous
woman among them.
\par }{\PP \VS{18}¶ The
{\ND{
LORD}} is
righteous; for I have
rebelled against his
commandment:
hear, I pray you, all
people, and
behold my
sorrow: my
virgins and my young
men are
gone into
captivity.\FTNT{†††}{{\FR 1.18 }
commandment: Heb. mouth}
\VS{19}I
called for my
lovers,
{\ADD{but}} they
deceived me: my
priests and mine
elders gave up the
ghost in the
city, while they
sought their
meat to
relieve their
souls.
\VS{20}Behold, O
{\ND{
LORD}}; for I
{\ADD{am}} in
distress: my
bowels are
troubled; mine
heart is
turned
within me; for I have
grievously
rebelled:
abroad the
sword
bereaveth, at
home
{\ADD{there is}} as
death.
\VS{21}They have
heard that I
sigh:
{\ADD{there is}} none to
comfort me: all mine
enemies have
heard of my
trouble; they are
glad that thou hast
done
{\ADD{it:}} thou wilt
bring the
day
{\ADD{that}} thou hast
called, and they shall be
like unto me.\FTNT{‡‡‡}{{\FR 1.21 }
called: or, proclaimed}
\VS{22}Let all their
wickedness
come
before thee; and
do unto them, as thou hast
done unto me for all my
transgressions: for my
sighs
{\ADD{are}}
many, and my
heart
{\ADD{is}}
faint.

\par }\Chap{2}{\PP \VerseOne{1}How hath the
Lord
covered the
daughter of
Zion with a
cloud in his
anger,
{\ADD{and}} cast
down from
heaven unto the
earth the
beauty of
Israel, and
remembered not his
footstool in the
day of his
anger!
\VS{2}The
Lord hath swallowed
up all the
habitations of
Jacob, and hath not
pitied: he hath thrown
down in his
wrath the strong
holds of the
daughter of
Judah; he hath brought
{\ADD{them}}
down to the
ground: he hath
polluted the
kingdom and the
princes thereof.\FTNT{*}{{\FR 2.2 }
brought…: Heb. made to touch}
\VS{3}He hath cut
off in
{\ADD{his}}
fierce
anger all the
horn of
Israel: he hath
drawn
back his right
hand from
before the
enemy, and he
burned against
Jacob like a
flaming
fire,
{\ADD{which}}
devoureth round
about.
\VS{4}He hath
bent his
bow like an
enemy: he
stood with his right
hand as an
adversary, and
slew all
{\ADD{that were}}
pleasant to the
eye in the
tabernacle of the
daughter of
Zion: he poured
out his
fury like
fire.\FTNT{†}{{\FR 2.4 }
all…: Heb. all the desirable of the eye}
\VS{5}The
Lord was as an
enemy: he hath swallowed
up
Israel, he hath swallowed
up all her
palaces: he hath
destroyed his strong
holds, and hath
increased in the
daughter of
Judah
mourning and
lamentation.
\VS{6}And he hath violently taken
away his
tabernacle, as
{\ADD{if it were of}} a
garden: he hath
destroyed his places of the
assembly: the
{\ND{
LORD}} hath caused the solemn
feasts and
sabbaths to be
forgotten in
Zion, and hath
despised in the
indignation of his
anger the
king and the
priest.\FTNT{‡}{{\FR 2.6 }
tabernacle: or, hedge}
\VS{7}The
Lord hath cast
off his
altar, he hath
abhorred his
sanctuary, he hath given
up into the
hand of the
enemy the
walls of her
palaces; they have
made a
noise in the
house of the
{\ND{
LORD}}, as in the
day of a solemn
feast.\FTNT{§}{{\FR 2.7 }
given up: Heb. shut up}
\VS{8}The
{\ND{
LORD}} hath
purposed to
destroy the
wall of the
daughter of
Zion: he hath stretched
out a
line, he hath not
withdrawn his
hand from
destroying: therefore he made the
rampart and the
wall to
lament; they
languished
together.\FTNT{**}{{\FR 2.8 }
destroying: Heb. swallowing up}
\VS{9}Her
gates are
sunk into the
ground; he hath
destroyed and
broken her
bars: her
king and her
princes
{\ADD{are}} among the
Gentiles: the
law
{\ADD{is}} no
{\ADD{more;}} her
prophets also
find no
vision from the
{\ND{
LORD}}.
\par }{\PP \VS{10}The
elders of the
daughter of
Zion
sit upon the
ground,
{\ADD{and}} keep
silence: they have cast
up
dust upon their
heads; they have
girded themselves with
sackcloth: the
virgins of
Jerusalem hang
down their
heads to the
ground.
\VS{11}Mine
eyes do
fail with
tears, my
bowels are
troubled, my
liver is
poured upon the
earth, for the
destruction of the
daughter of my
people; because the
children and the
sucklings
swoon in the
streets of the
city.\FTNT{††}{{\FR 2.11 }
swoon: or, faint}
\VS{12}They
say to their
mothers, Where
{\ADD{is}}
corn and
wine? when they
swooned as the
wounded in the
streets of the
city, when their
soul was poured
out into their
mothers’
bosom.
\VS{13}What thing shall I take to
witness for thee? what
thing shall I
liken to thee, O
daughter of
Jerusalem? what shall I
equal to thee, that I may
comfort thee, O
virgin
daughter of
Zion? for thy
breach
{\ADD{is}}
great like the
sea: who can
heal thee?
\VS{14}Thy
prophets have
seen
vain and foolish
things for thee: and they have not
discovered thine
iniquity, to turn
away thy
captivity; but have
seen for thee
false
burdens and causes of
banishment.
\VS{15}All that
pass by
clap
{\ADD{their}}
hands at
thee; they
hiss and
wag their
head at the
daughter of
Jerusalem,
{\ADD{saying, Is}} this the
city that
{\ADD{men}}
call The
perfection of
beauty, The
joy of the whole
earth?\FTNT{‡‡}{{\FR 2.15 }
by: Heb. by the way}
\VS{16}All thine
enemies have
opened their
mouth against thee: they
hiss and
gnash the
teeth: they
say, We have swallowed
{\ADD{her}}
up:
certainly this
{\ADD{is}} the
day that we looked
for; we have
found, we have
seen
{\ADD{it}}.
\VS{17}The
{\ND{
LORD}} hath
done
{\ADD{that}} which he had
devised; he hath
fulfilled his
word that he had
commanded in the
days of
old: he hath thrown
down, and hath not
pitied: and he hath caused
{\ADD{thine}}
enemy to
rejoice over thee, he hath set
up the
horn of thine
adversaries.
\VS{18}Their
heart
cried unto the
Lord, O
wall of the
daughter of
Zion, let
tears run
down like a
river
day and
night:
give thyself no
rest; let not the
apple of thine
eye
cease.
\VS{19}Arise, cry
out in the
night: in the
beginning of the
watches pour
out thine
heart like
water
before the
face of the
Lord: lift
up thy
hands toward him for the
life of thy young
children, that
faint for
hunger in the
top of every
street.
\par }{\PP \VS{20}¶
Behold, O
{\ND{
LORD}}, and
consider to whom thou hast
done
this. Shall the
women
eat their
fruit,
{\ADD{and}}
children of a span
long? shall the
priest and the
prophet be
slain in the
sanctuary of the
Lord?\FTNT{§§}{{\FR 2.20 }
of a span…: or, swaddled with their hands?}
\VS{21}The
young and the
old
lie on the
ground in the
streets: my
virgins and my young
men are
fallen by the
sword; thou hast
slain
{\ADD{them}} in the
day of thine
anger; thou hast
killed,
{\ADD{and}} not
pitied.
\VS{22}Thou hast
called as in a
solemn
day my
terrors round
about, so that in the
day of the
{\ND{
LORD’s}}
anger none
escaped nor
remained: those that I have
swaddled and brought
up hath mine
enemy
consumed.

\par }\Chap{3}{\PP \VerseOne{1}I
{\ADD{am}} the
man
{\ADD{that}} hath
seen
affliction by the
rod of his
wrath.
\VS{2}He hath
led me, and
brought
{\ADD{me into}}
darkness, but not
{\ADD{into}}
light.
\VS{3}Surely against me is he
turned; he
turneth his
hand
{\ADD{against me}} all the
day.
\VS{4}My
flesh and my
skin hath he made
old; he hath
broken my
bones.
\VS{5}He hath builded
against me, and
compassed
{\ADD{me}} with
gall and
travail.
\VS{6}He hath
set me in dark
places, as
{\ADD{they that be}}
dead of
old.
\VS{7}He hath
hedged me about, that I cannot get
out: he hath made my
chain
heavy.
\VS{8}Also when I
cry and
shout, he shutteth
out my
prayer.
\VS{9}He hath
inclosed my
ways with hewn
stone, he hath made my
paths
crooked.
\VS{10}He
{\ADD{was}} unto me
{\ADD{as}} a
bear lying in
wait,
{\ADD{and as}} a
lion in secret
places.
\VS{11}He hath turned
aside my
ways, and pulled me in
pieces: he hath
made me
desolate.
\VS{12}He hath
bent his
bow, and
set me as a
mark for the
arrow.
\VS{13}He hath caused the
arrows of his
quiver to
enter into my
reins.\FTNT{*}{{\FR 3.13 }
arrows: Heb. sons}
\VS{14}I was a
derision to all my
people;
{\ADD{and}} their
song all the
day.
\VS{15}He hath
filled me with
bitterness, he hath made me
drunken with
wormwood.\FTNT{†}{{\FR 3.15 }
bitterness: Heb. bitternesses}
\VS{16}He hath also
broken my
teeth with gravel
stones, he hath
covered me with
ashes.\FTNT{‡}{{\FR 3.16 }
covered…: or, rolled me in the ashes}
\VS{17}And thou hast
removed my
soul far
off from
peace: I
forgat
prosperity.\FTNT{§}{{\FR 3.17 }
prosperity: Heb. good}
\VS{18}And I
said, My
strength and my
hope is
perished from the
{\ND{
LORD}}:
\VS{19}Remembering mine
affliction and my
misery, the
wormwood and the
gall.\FTNT{**}{{\FR 3.19 }
Remembering: or, Remember}
\VS{20}My
soul hath
{\ADD{them}}
still in
remembrance, and is
humbled in me.\FTNT{††}{{\FR 3.20 }
humbled: Heb. bowed}
\par }{\PP \VS{21}This I
recall to my
mind, therefore have I
hope.\FTNT{‡‡}{{\FR 3.21 }
recall…: Heb. make to return to my heart}
\par }{\PP \VS{22}¶
{\ADD{It is of}} the
{\ND{
LORD’s}}
mercies that we are not
consumed, because his
compassions
fail not.
\VS{23}{\ADD{They are}}
new every
morning:
great
{\ADD{is}} thy
faithfulness.
\VS{24}The
{\ND{
LORD}}
{\ADD{is}} my
portion,
saith my
soul; therefore will I
hope in him.
\VS{25}The
{\ND{
LORD}}
{\ADD{is}}
good unto them that
wait for him, to the
soul
{\ADD{that}}
seeketh him.
\VS{26}{\ADD{It is}}
good that
{\ADD{a man}} should both
hope and quietly
wait for the
salvation of the
{\ND{
LORD}}.
\VS{27}{\ADD{It is}}
good for a
man that he
bear the
yoke in his
youth.
\VS{28}He
sitteth
alone and keepeth
silence, because he hath
borne
{\ADD{it}} upon him.
\VS{29}He
putteth his
mouth in the
dust; if so be there may
be
hope.
\VS{30}He
giveth
{\ADD{his}}
cheek to him that
smiteth him: he is filled
full with
reproach.
\VS{31}For the
Lord will not cast
off for
ever:
\VS{32}But though he cause
grief, yet will he have
compassion according to the
multitude of his
mercies.
\VS{33}For he doth not
afflict
willingly nor
grieve the
children of
men.\FTNT{§§}{{\FR 3.33 }
willingly: Heb. from his heart}
\VS{34}To
crush under his
feet all the
prisoners of the
earth,
\VS{35}To turn
aside the
right of a
man before the
face of the most
High,\FTNT{***}{{\FR 3.35 }
the most High: or, a superior}
\VS{36}To
subvert a
man in his
cause, the
Lord
approveth not.\FTNT{†††}{{\FR 3.36 }
approveth not: or, seeth not}
\par }{\PP \VS{37}¶ Who
{\ADD{is}} he
{\ADD{that}}
saith, and it cometh to pass,
{\ADD{when}} the
Lord
commandeth
{\ADD{it}} not?
\VS{38}Out of the
mouth of the most
High
proceedeth not
evil and
good?
\VS{39}Wherefore doth a
living
man
complain, a
man for the punishment of his
sins?\FTNT{‡‡‡}{{\FR 3.39 }
complain: or, murmur}
\VS{40}Let us
search and
try our
ways, and turn
again to the
{\ND{
LORD}}.
\VS{41}Let us lift
up our
heart with
{\ADD{our}}
hands unto
God in the
heavens.
\par }{\PP \VS{42}We have
transgressed and have
rebelled: thou hast not
pardoned.
\VS{43}Thou hast
covered with
anger, and
persecuted us: thou hast
slain, thou hast not
pitied.
\VS{44}Thou hast
covered thyself with a
cloud, that
{\ADD{our}}
prayer should not pass
through.
\VS{45}Thou hast
made us
{\ADD{as}} the
offscouring and
refuse in the
midst of the
people.
\VS{46}All our
enemies have
opened their
mouths against us.
\VS{47}Fear and a
snare is come upon us,
desolation and
destruction.
\VS{48}Mine
eye runneth
down with
rivers of
water for the
destruction of the
daughter of my
people.
\VS{49}Mine
eye trickleth
down, and
ceaseth not, without any
intermission,
\VS{50}Till the
{\ND{
LORD}} look
down, and
behold from
heaven.
\VS{51}Mine
eye
affecteth mine
heart because of all the
daughters of my
city.\FTNT{§§§}{{\FR 3.51 }
mine heart: Heb. my soul}\FTNT{*}{{\FR 3.51 }
because…: or, more than all}
\VS{52}Mine
enemies
chased me
sore, like a
bird, without
cause.
\VS{53}They have cut
off my
life in the
dungeon, and
cast a
stone upon me.
\VS{54}Waters flowed
over mine
head;
{\ADD{then}} I
said, I am cut
off.
\par }{\PP \VS{55}¶ I
called upon thy
name, O
{\ND{
LORD}}, out of the
low
dungeon.
\VS{56}Thou hast
heard my
voice:
hide not thine
ear at my
breathing, at my
cry.
\VS{57}Thou drewest
near in the
day
{\ADD{that}} I
called upon thee: thou
saidst,
Fear not.
\VS{58}O
Lord, thou hast
pleaded the
causes of my
soul; thou hast
redeemed my
life.
\VS{59}O
{\ND{
LORD}}, thou hast
seen my
wrong:
judge thou my
cause.
\VS{60}Thou hast
seen all their
vengeance
{\ADD{and}} all their
imaginations against me.
\VS{61}Thou hast
heard their
reproach, O
{\ND{
LORD}},
{\ADD{and}} all their
imaginations against me;
\VS{62}The
lips of those that rose
up against me, and their
device against me all the
day.
\VS{63}Behold their sitting
down, and their rising
up; I
{\ADD{am}} their
musick.
\par }{\PP \VS{64}¶
Render unto them a
recompence, O
{\ND{
LORD}}, according to the
work of their
hands.
\VS{65}Give them
sorrow of
heart, thy
curse unto them.\FTNT{†}{{\FR 3.65 }
sorrow…: or, obstinacy of heart}
\VS{66}Persecute and
destroy them in
anger from under the
heavens of the
{\ND{
LORD}}.

\par }\Chap{4}{\PP \VerseOne{1}How is the
gold become
dim!
{\ADD{how}} is the
most fine
gold
changed! the
stones of the
sanctuary are poured
out in the
top of every
street.
\VS{2}The
precious
sons of
Zion,
comparable to fine
gold, how are they
esteemed as
earthen
pitchers, the
work of the
hands of the
potter!
\VS{3}Even the sea
monsters draw
out the
breast, they give
suck to their young
ones: the
daughter of my
people
{\ADD{is become}}
cruel, like the
ostriches in the
wilderness.\FTNT{*}{{\FR 4.3 }
sea…: or, sea calves}
\VS{4}The
tongue of the sucking
child
cleaveth to the roof of his
mouth for
thirst: the young
children
ask
bread,
{\ADD{and}} no man
breaketh
{\ADD{it}} unto them.
\VS{5}They that did
feed
delicately are
desolate in the
streets: they that were brought
up in
scarlet
embrace
dunghills.
\VS{6}For the punishment of the
iniquity of the
daughter of my
people is
greater than the
punishment of the sin of
Sodom, that was
overthrown as in a
moment, and no
hands
stayed on her.\FTNT{†}{{\FR 4.6 }
punishment of the iniquity: or, iniquity}
\VS{7}Her
Nazarites were
purer than
snow, they were
whiter than
milk, they were more
ruddy in
body than
rubies, their
polishing
{\ADD{was}} of
sapphire:
\VS{8}Their
visage is
blacker than a
coal; they are not
known in the
streets: their
skin
cleaveth to their
bones; it is
withered, it is become like a
stick.\FTNT{‡}{{\FR 4.8 }
blacker…: Heb. darker than blackness}
\VS{9}{\ADD{They that be}}
slain with the
sword are
better than
{\ADD{they that be}}
slain with
hunger: for
these pine
away,
stricken through for
{\ADD{want of}} the
fruits of the
field.\FTNT{§}{{\FR 4.9 }
pine…: Heb. flow out}
\VS{10}The
hands of the
pitiful
women have
sodden their own
children: they were their
meat in the
destruction of the
daughter of my
people.
\VS{11}The
{\ND{
LORD}} hath
accomplished his
fury; he hath poured
out his
fierce
anger, and hath
kindled a
fire in
Zion, and it hath
devoured the
foundations thereof.
\VS{12}The
kings of the
earth, and all the
inhabitants of the
world, would not have
believed that the
adversary and the
enemy should have
entered into the
gates of
Jerusalem.
\par }{\PP \VS{13}¶ For the
sins of her
prophets,
{\ADD{and}} the
iniquities of her
priests, that have
shed the
blood of the
just in the
midst of her,
\VS{14}They have
wandered
{\ADD{as}}
blind
{\ADD{men}} in the
streets, they have
polluted themselves with
blood, so that men
could
not
touch their
garments.\FTNT{**}{{\FR 4.14 }
so…: or, in that they could not but touch}
\VS{15}They
cried unto them,
Depart ye;
{\ADD{it is}}
unclean;
depart,
depart,
touch not: when they fled
away and
wandered, they
said among the
heathen, They shall no
more
sojourn
{\ADD{there}}.\FTNT{††}{{\FR 4.15 }
it…: or, ye polluted}
\VS{16}The
anger of the
{\ND{
LORD}} hath
divided them; he will no
more
regard them: they
respected not the
persons of the
priests, they
favoured not the
elders.\FTNT{‡‡}{{\FR 4.16 }
anger: or, face}
\VS{17}As for us, our
eyes as yet
failed for our
vain
help: in our
watching we have
watched for a
nation
{\ADD{that}} could not
save
{\ADD{us}}.
\VS{18}They
hunt our
steps, that we cannot
go in our
streets: our
end is
near, our
days are
fulfilled; for our
end is
come.
\VS{19}Our
persecutors are
swifter than the
eagles of the
heaven: they
pursued us upon the
mountains, they laid
wait for us in the
wilderness.
\VS{20}The
breath of our
nostrils, the
anointed of the
{\ND{
LORD}}, was
taken in their
pits, of whom we
said, Under his
shadow we shall
live among the
heathen.
\par }{\PP \VS{21}¶
Rejoice and be
glad, O
daughter of
Edom, that
dwellest in the
land of
Uz; the
cup also shall
pass through unto thee: thou shalt be
drunken, and shalt make thyself
naked.
\par }{\PP \VS{22}¶ The punishment of thine
iniquity is
accomplished, O
daughter of
Zion; he will no
more carry thee away into
captivity: he will
visit thine
iniquity, O
daughter of
Edom; he will
discover thy
sins.\FTNT{§§}{{\FR 4.22 }
The…: or, Thine iniquity}\FTNT{***}{{\FR 4.22 }
discover…: or, carry thee captive for thy sins}

\par }\Chap{5}{\PP \VerseOne{1}Remember, O
{\ND{
LORD}}, what is come upon us:
consider, and
behold our
reproach.
\VS{2}Our
inheritance is
turned to
strangers, our
houses to
aliens.
\VS{3}We are
orphans and
fatherless, our
mothers
{\ADD{are}} as
widows.
\VS{4}We have
drunken our
water for
money; our
wood is
sold unto us.\FTNT{*}{{\FR 5.4 }
is…: Heb. cometh for price}
\VS{5}Our
necks
{\ADD{are}} under
persecution: we
labour,
{\ADD{and}} have no
rest.\FTNT{†}{{\FR 5.5 }
Our…: Heb. On our necks are we persecuted}
\VS{6}We have
given the
hand
{\ADD{to}} the
Egyptians,
{\ADD{and to}} the
Assyrians, to be
satisfied with
bread.
\VS{7}Our
fathers have
sinned,
{\ADD{and are}}
not; and we have
borne their
iniquities.
\VS{8}Servants have
ruled over us:
{\ADD{there is}} none that doth
deliver
{\ADD{us}} out of their
hand.
\VS{9}We
gat our
bread with
{\ADD{the peril of}} our
lives
because of the
sword of the
wilderness.
\VS{10}Our
skin was
black like an
oven
because of the
terrible
famine.\FTNT{‡}{{\FR 5.10 }
terrible: or, terrors, or, storms}
\VS{11}They
ravished the
women in
Zion,
{\ADD{and}} the
maids in the
cities of
Judah.
\VS{12}Princes are hanged
up by their
hand: the
faces of
elders were not
honoured.
\VS{13}They
took the young
men to
grind, and the
children
fell under the
wood.
\VS{14}The
elders have
ceased from the
gate, the young
men from their
musick.
\VS{15}The
joy of our
heart is
ceased; our
dance is
turned into
mourning.
\VS{16}The
crown is
fallen
{\ADD{from}} our
head:
woe unto us, that we have
sinned!\FTNT{§}{{\FR 5.16 }
The…: Heb. The crown of our head is fallen}
\par }{\PP \VS{17}For this our
heart is
faint; for these
{\ADD{things}} our
eyes are
dim.
\VS{18}Because of the
mountain of
Zion, which is
desolate, the
foxes
walk upon it.
\VS{19}Thou, O
{\ND{
LORD}},
remainest for
ever; thy
throne from
generation to
generation.
\VS{20}Wherefore dost thou
forget us for
ever,
{\ADD{and}}
forsake us so
long
time?\FTNT{**}{{\FR 5.20 }
so…: Heb. for length of days?}
\VS{21}Turn thou us unto thee, O
{\ND{
LORD}}, and we shall be
turned;
renew our
days as of
old.
\VS{22}But thou hast
utterly
rejected us; thou art
very
wroth against us.\FTNT{††}{{\FR 5.22 }
But…: or, For wilt thou utterly reject us?}
\par }