\NormalFont\ShortTitle{Malachi}
{\MT Malachi

\par }\ChapOne{1}{\PP \VerseOne{1}The
burden of the
word of the
{\ND{
LORD}} to
Israel
by
Malachi.\FTNT{*}{{\FR 1.1 }
by…: Heb. by the hand of}
\VS{2}I have
loved you,
saith the
{\ND{
LORD}}. Yet ye
say, Wherein hast thou
loved us?
{\ADD{Was}} not
Esau
Jacob’s
brother?
saith the
{\ND{
LORD}}: yet I
loved
Jacob,
\VS{3}And I
hated
Esau, and
laid his
mountains and his
heritage
waste for the
dragons of the
wilderness.
\VS{4}Whereas
Edom
saith, We are
impoverished, but we will
return and
build the desolate
places; thus
saith the
{\ND{
LORD}} of
hosts, They shall
build, but I will throw
down; and they shall
call them, The
border of
wickedness, and, The
people against whom the
{\ND{
LORD}} hath
indignation
for
ever.
\VS{5}And your
eyes shall
see, and ye shall
say, The
{\ND{
LORD}} will be
magnified from the
border of
Israel.\FTNT{†}{{\FR 1.5 }
from: or, upon: Heb. from upon}
\par }{\PP \VS{6}¶ A
son
honoureth
{\ADD{his}}
father, and a
servant his
master: if then I
{\ADD{be}} a
father, where
{\ADD{is}} mine
honour? and if I
{\ADD{be}} a
master, where
{\ADD{is}} my
fear?
saith the
{\ND{
LORD}} of
hosts unto you, O
priests, that
despise my
name. And ye
say, Wherein have we
despised thy
name?
\VS{7}Ye
offer
polluted
bread upon mine
altar; and ye
say, Wherein have we
polluted thee? In that ye
say, The
table of the
{\ND{
LORD}}
{\ADD{is}}
contemptible.\FTNT{‡}{{\FR 1.7 }
offer…: or, bring unto, etc}
\VS{8}And if ye
offer the
blind for
sacrifice,
{\ADD{is it}} not
evil? and if ye
offer the
lame and
sick,
{\ADD{is it}} not
evil?
offer it now unto thy
governor; will he be
pleased with thee, or
accept thy
person?
saith the
{\ND{
LORD}} of
hosts.\FTNT{§}{{\FR 1.8 }
for: Heb. to}
\VS{9}And now, I pray you,
beseech
God that he will be
gracious unto us: this hath been by your
means: will he
regard your
persons?
saith the
{\ND{
LORD}} of
hosts.\FTNT{**}{{\FR 1.9 }
God: Heb. the face of God}\FTNT{††}{{\FR 1.9 }
by…: Heb. from your hand}
\VS{10}Who
{\ADD{is there}} even among you that would
shut the
doors
{\ADD{for nought?}} neither do ye
kindle
{\ADD{fire}} on mine
altar for
nought. I have no
pleasure in you,
saith the
{\ND{
LORD}} of
hosts, neither will I
accept an
offering at your
hand.
\VS{11}For from the
rising of the
sun even unto the going
down of the same my
name
{\ADD{shall be}}
great among the
Gentiles; and in every
place
incense
{\ADD{shall be}}
offered unto my
name, and a
pure
offering: for my
name
{\ADD{shall be}}
great among the
heathen,
saith the
{\ND{
LORD}} of
hosts.
\par }{\PP \VS{12}¶ But ye have
profaned it, in that ye
say, The
table of the
{\ND{
LORD}}
{\ADD{is}}
polluted; and the
fruit thereof,
{\ADD{even}} his
meat,
{\ADD{is}}
contemptible.
\VS{13}Ye
said also, Behold, what a
weariness
{\ADD{is it!}} and ye have
snuffed at it,
saith the
{\ND{
LORD}} of
hosts; and ye
brought
{\ADD{that which was}}
torn, and the
lame, and the
sick; thus ye
brought an
offering: should I
accept this of your
hand?
saith the
{\ND{
LORD}}.\FTNT{‡‡}{{\FR 1.13 }
and ye have…: or, whereas ye might have blown it away}
\VS{14}But
cursed
{\ADD{be}} the
deceiver,
which hath in his
flock a
male, and
voweth, and
sacrificeth unto the
Lord a corrupt
thing: for I
{\ADD{am}} a
great
King,
saith the
{\ND{
LORD}} of
hosts, and my
name
{\ADD{is}}
dreadful among the
heathen.\FTNT{§§}{{\FR 1.14 }
which…: Heb. in whose flock is}

\par }\Chap{2}{\PP \VerseOne{1}And now, O ye
priests, this
commandment
{\ADD{is}} for you.
\VS{2}If ye will not
hear, and if ye will not
lay
{\ADD{it}} to
heart, to
give
glory unto my
name,
saith the
{\ND{
LORD}} of
hosts, I will even
send a
curse upon you, and I will
curse your
blessings: yea, I have
cursed them already, because ye do not
lay
{\ADD{it}} to
heart.
\VS{3}Behold, I will
corrupt your
seed, and
spread
dung upon your
faces,
{\ADD{even}} the
dung of your solemn
feasts; and
{\ADD{one}} shall take you
away with it.\FTNT{*}{{\FR 2.3 }
corrupt: or, reprove}\FTNT{†}{{\FR 2.3 }
spread: Heb. scatter}\FTNT{‡}{{\FR 2.3 }
one…: or, it shall take you away to it}
\VS{4}And ye shall
know that I have
sent this
commandment unto you, that my
covenant might be with
Levi,
saith the
{\ND{
LORD}} of
hosts.
\VS{5}My
covenant was with him of
life and
peace; and I
gave them to him
{\ADD{for}} the
fear wherewith he
feared me, and was
afraid
before my
name.
\VS{6}The
law of
truth was in his
mouth, and
iniquity was not
found in his
lips: he
walked with me in
peace and
equity, and did
turn
many
away from
iniquity.
\VS{7}For the
priest’s
lips should
keep
knowledge, and they should
seek the
law at his
mouth: for he
{\ADD{is}} the
messenger of the
{\ND{
LORD}} of
hosts.
\VS{8}But ye are
departed out of the
way; ye have caused
many to
stumble at the
law; ye have
corrupted the
covenant of
Levi,
saith the
{\ND{
LORD}} of
hosts.\FTNT{§}{{\FR 2.8 }
stumble at: or, fall in}
\VS{9}Therefore have I also
made you
contemptible and
base before all the
people,
according as ye have not
kept my
ways, but have been
partial in the
law.\FTNT{**}{{\FR 2.9 }
have been…: or, lifted up the face against: Heb. accepted faces}
\par }{\PP \VS{10}Have we not all
one
father? hath not
one
God
created us? why do we deal
treacherously every
man against his
brother, by
profaning the
covenant of our
fathers?
\par }{\PP \VS{11}¶
Judah hath dealt
treacherously, and an
abomination is
committed in
Israel and in
Jerusalem; for
Judah hath
profaned the
holiness of the
{\ND{
LORD}} which he
loved, and hath
married the
daughter of a
strange
god.\FTNT{††}{{\FR 2.11 }
loved: or, ought to love}
\VS{12}The
{\ND{
LORD}} will cut
off the
man that
doeth this, the
master and the
scholar, out of the
tabernacles of
Jacob, and him that
offereth an
offering unto the
{\ND{
LORD}} of
hosts.\FTNT{‡‡}{{\FR 2.12 }
the master…: or, him that waketh, and him that answereth}
\VS{13}And this have ye
done
again,
covering the
altar of the
{\ND{
LORD}} with
tears, with
weeping, and with crying
out, insomuch that he
regardeth not the
offering any more, or
receiveth
{\ADD{it}} with good
will at your
hand.
\par }{\PP \VS{14}¶ Yet ye
say,
Wherefore? Because the
{\ND{
LORD}} hath been
witness between thee and the
wife of thy
youth, against whom thou hast dealt
treacherously: yet
{\ADD{is}} she thy
companion, and the
wife of thy
covenant.
\VS{15}And did not he
make
one? Yet had he the
residue of the
spirit. And wherefore
one? That he might
seek a
godly
seed. Therefore take
heed to your
spirit, and let none deal
treacherously against the
wife of his
youth.\FTNT{§§}{{\FR 2.15 }
residue: or, excellency}\FTNT{***}{{\FR 2.15 }
godly…: Heb. seed of God}\FTNT{†††}{{\FR 2.15 }
treacherously: or, unfaithfully}
\VS{16}For the
{\ND{
LORD}}, the
God of
Israel,
saith that he
hateth putting
away: for
{\ADD{one}}
covereth
violence with his
garment,
saith the
{\ND{
LORD}} of
hosts: therefore take
heed to your
spirit, that ye deal not
treacherously.\FTNT{‡‡‡}{{\FR 2.16 }
that he…: or, if he hate her, put her away}\FTNT{§§§}{{\FR 2.16 }
putting…: Heb. to put away}
\par }{\PP \VS{17}¶ Ye have
wearied the
{\ND{
LORD}} with your
words. Yet ye
say, Wherein have we
wearied
{\ADD{him?}} When ye
say, Every one that
doeth
evil
{\ADD{is}}
good in the
sight of the
{\ND{
LORD}}, and he
delighteth in them; or, Where
{\ADD{is}} the
God of
judgment?

\par }\Chap{3}{\PP \VerseOne{1}Behold, I will
send my
messenger, and he shall
prepare the
way
before me: and the
Lord, whom ye
seek, shall
suddenly
come to his
temple, even the
messenger of the
covenant, whom ye
delight in: behold, he shall
come,
saith the
{\ND{
LORD}} of
hosts.
\VS{2}But who may
abide the
day of his
coming? and who shall
stand when he
appeareth? for he
{\ADD{is}} like a
refiner’s
fire, and like
fullers’
soap:
\VS{3}And he shall
sit
{\ADD{as}} a
refiner and
purifier of
silver: and he shall
purify the
sons of
Levi, and
purge them as
gold and
silver, that they may
offer unto the
{\ND{
LORD}} an
offering in
righteousness.
\VS{4}Then shall the
offering of
Judah and
Jerusalem be
pleasant unto the
{\ND{
LORD}}, as in the
days of
old, and as in
former
years.\FTNT{*}{{\FR 3.4 }
former: or, ancient}
\VS{5}And I will come
near to you to
judgment; and I will be a
swift
witness against the
sorcerers, and against the
adulterers, and against
false
swearers, and against those that
oppress the
hireling in
{\ADD{his}}
wages, the
widow, and the
fatherless, and that turn
aside the
stranger
{\ADD{from his right}}, and
fear not me,
saith the
{\ND{
LORD}} of
hosts.\FTNT{†}{{\FR 3.5 }
oppress: or, defraud}
\VS{6}For I
{\ADD{am}} the
{\ND{
LORD}}, I
change not; therefore ye
sons of
Jacob are not
consumed.
\par }{\PP \VS{7}¶ Even from the
days of your
fathers ye are gone
away from mine
ordinances, and have not
kept
{\ADD{them}}.
Return unto me, and I will
return unto you,
saith the
{\ND{
LORD}} of
hosts. But ye
said, Wherein shall we
return?
\par }{\PP \VS{8}¶ Will a
man
rob
God? Yet ye have
robbed me. But ye
say, Wherein have we
robbed thee? In
tithes and
offerings.
\VS{9}Ye
{\ADD{are}}
cursed with a
curse: for ye have
robbed me,
{\ADD{even}} this whole
nation.
\VS{10}Bring ye all the
tithes into the
storehouse, that there may be
meat in mine
house, and
prove me now
herewith,
saith the
{\ND{
LORD}} of
hosts, if I will not
open you the
windows of
heaven, and pour you
out a
blessing, that
{\ADD{there shall}} not
{\ADD{be room}}
enough
{\ADD{to receive it}}.\FTNT{‡}{{\FR 3.10 }
pour…: Heb. empty out}
\VS{11}And I will
rebuke the
devourer for your sakes, and he shall not
destroy the
fruits of your
ground; neither shall your
vine cast her
fruit before the time in the
field,
saith the
{\ND{
LORD}} of
hosts.\FTNT{§}{{\FR 3.11 }
destroy: Heb. corrupt}
\VS{12}And all
nations shall call you
blessed: for ye shall be a
delightsome
land,
saith the
{\ND{
LORD}} of
hosts.
\par }{\PP \VS{13}¶ Your
words have been
stout against me,
saith the
{\ND{
LORD}}. Yet ye
say, What have we
spoken
{\ADD{so much}} against thee?
\VS{14}Ye have
said, It
{\ADD{is}}
vain to
serve
God: and what
profit
{\ADD{is it}} that we have
kept his
ordinance, and that we have
walked
mournfully
before the
{\ND{
LORD}} of
hosts?\FTNT{**}{{\FR 3.14 }
ordinance: Heb. observation}\FTNT{††}{{\FR 3.14 }
mournfully: Heb. in black}
\VS{15}And now we
call the
proud
happy; yea, they that
work
wickedness are set
up; yea,
{\ADD{they that}}
tempt
God are even
delivered.\FTNT{‡‡}{{\FR 3.15 }
are set up: Heb. are built}
\par }{\PP \VS{16}¶ Then they that
feared the
{\ND{
LORD}} spake
often
one to
another: and the
{\ND{
LORD}}
hearkened, and
heard
{\ADD{it}}, and a
book of
remembrance was
written
before him for them that
feared the
{\ND{
LORD}}, and that
thought upon his
name.
\VS{17}And they shall be mine,
saith the
{\ND{
LORD}} of
hosts, in that
day when I make
up my
jewels; and I will
spare them, as a
man
spareth his own
son that
serveth him.\FTNT{§§}{{\FR 3.17 }
jewels: or, special treasure}
\VS{18}Then shall ye
return, and
discern between the
righteous and the
wicked, between him that
serveth
God and him that
serveth him not.

\par }\Chap{4}{\PP \VerseOne{1}For, behold, the
day
cometh, that shall
burn
as an oven;
and all the proud, yea, and all that
do
wickedly,
shall be stubble:
and the day that
cometh
shall burn them up,
saith the
{\ND{
LORD}}
of hosts,
that it shall leave
them neither root
nor branch.
\par }{\PP \VS{2}¶
But unto you that fear
my name
shall the Sun
of righteousness
arise
with healing
in his wings;
and ye shall go forth,
and grow up
as calves
of the stall.
\VS{3}And ye shall tread down
the wicked;
for they shall be ashes
under the soles
of your feet
in the day
that I shall do
{\ADD{this}},
saith the
{\ND{
LORD}}
of hosts.
\par }{\PP \VS{4}¶
Remember
ye the law
of Moses
my servant,
which I commanded
unto him in Horeb
for all Israel,
{\ADD{with}}
the statutes
and judgments.
\par }{\PP \VS{5}¶ Behold, I will
send
you Elijah
the prophet
before
the coming
of the great
and dreadful
day of the
{\ND{
LORD}}:
\VS{6}And he shall turn
the heart
of the fathers
to the children,
and the heart
of the children
to their fathers,
lest I come
and smite
the earth
with a curse.
\par }