\NormalFont\ShortTitle{Matthew}
{\MT THE GOSPEL ACCORDING TO ST. MATTHEW

\par }\ChapOne{1}{\PP \VerseOne{1}The
book of the
generation of
Jesus
Christ, the
son of
David, the
son of
Abraham.
\VS{2}Abraham
begat
Isaac;
and
Isaac
begat
Jacob;
and
Jacob
begat
Judas
and
his
brethren;
\VS{3}And
Judas
begat
Phares
and
Zara
of
Thamar;
and
Phares
begat
Esrom;
and
Esrom
begat
Aram;
\VS{4}And
Aram
begat
Aminadab;
and
Aminadab
begat
Naasson;
and
Naasson
begat
Salmon;
\VS{5}And
Salmon
begat
Booz
of
Rachab;
and
Booz
begat
Obed
of
Ruth;
and
Obed
begat
Jesse;
\VS{6}And
Jesse
begat
David the
king;
and
David the
king
begat
Solomon
of
her
{\ADD{that had been the wife}} of
Urias;
\VS{7}And
Solomon
begat
Roboam;
and
Roboam
begat
Abia;
and
Abia
begat
Asa;
\VS{8}And
Asa
begat
Josaphat;
and
Josaphat
begat
Joram;
and
Joram
begat
Ozias;
\VS{9}And
Ozias
begat
Joatham;
and
Joatham
begat
Achaz;
and
Achaz
begat
Ezekias;
\VS{10}And
Ezekias
begat
Manasses;
and
Manasses
begat
Amon;
and
Amon
begat
Josias;
\VS{11}And
Josias
begat
Jechonias
and his
brethren, about the
time
they were carried away
to
Babylon:
\VS{12}And
after they were brought
to
Babylon,
Jechonias
begat
Salathiel;
and
Salathiel
begat
Zorobabel;
\VS{13}And
Zorobabel
begat
Abiud;
and
Abiud
begat
Eliakim;
and
Eliakim
begat
Azor;
\VS{14}And
Azor
begat
Sadoc;
and
Sadoc
begat
Achim;
and
Achim
begat
Eliud;
\VS{15}And
Eliud
begat
Eleazar;
and
Eleazar
begat
Matthan;
and
Matthan
begat
Jacob;
\VS{16}And
Jacob
begat
Joseph the
husband of
Mary,
of
whom was
born
Jesus, who is
called
Christ.
\VS{17}So
all the
generations
from
Abraham
to
David
{\ADD{are}}
fourteen
generations;
and
from
David
until the carrying away
into
Babylon
{\ADD{are}}
fourteen
generations;
and
from the carrying away
into
Babylon
unto
Christ
{\ADD{are}}
fourteen
generations.
\par }{\PP \VS{18}¶
Now the
birth of
Jesus
Christ
was on this
wise: When
as
his
mother
Mary was
espoused to
Joseph,
before
they came
together, she was
found
with
child
of
the
Holy
Ghost.
\VS{19}Then
Joseph
her
husband,
being a
just
{\ADD{man}},
and
not
willing to
make
her a publick
example, was
minded to
put
her
away
privily.
\VS{20}But
while
he thought
on these
things,
behold, the
angel of the
Lord
appeared unto
him
in a
dream,
saying,
Joseph, thou
son of
David,
fear
not to take unto
thee
Mary
thy
wife: for
that
which is
conceived
in
her
is
of
the
Holy
Ghost.
\VS{21}And she shall bring
forth a
son,
and thou shalt
call
his
name
JESUS:
for
he shall
save
his
people
from
their
sins.
\VS{22}Now
all
this was
done, that
it might be
fulfilled
which was
spoken
of the
Lord
by the
prophet,
saying,
\VS{23}Behold, a
virgin shall
be
with
child,
and shall bring
forth a
son,
and they shall
call
his
name
Emmanuel,
which being
interpreted
is,
God
with
us.
\VS{24}Then
Joseph being
raised
from
sleep
did
as the
angel of the
Lord had
bidden
him,
and took unto
him
his
wife:
\VS{25}And
knew
her
not
till she had brought
forth
her
firstborn
son:
and he
called
his
name
JESUS.

\par }\Chap{2}{\PP \VerseOne{1}Now
when
Jesus was
born
in
Bethlehem of
Judaea
in the
days of
Herod the
king,
behold, there
came wise
men
from the
east
to
Jerusalem,
\VS{2}Saying,
Where
is
he that is
born
King of the
Jews?
for we have
seen
his
star
in the
east,
and are
come to
worship
him.
\VS{3}When
Herod the
king had
heard
{\ADD{these things}}, he was
troubled,
and
all
Jerusalem
with
him.
\VS{4}And
when he had
gathered
all the chief
priests
and
scribes of the
people
together, he
demanded
of
them
where
Christ should be
born.
\VS{5}And they
said unto
him,
In
Bethlehem of
Judaea:
for
thus it is
written
by the
prophet,
\VS{6}And
thou
Bethlehem,
{\ADD{in}} the
land of
Juda,
art
not the
least
among the
princes of
Juda:
for out
of
thee shall
come a
Governor,
that shall
rule
my
people
Israel.
\VS{7}Then
Herod, when he
had
privily
called the wise
men,
enquired
of
them
diligently what
time the
star
appeared.
\VS{8}And he
sent
them
to
Bethlehem, and
said,
Go and
search
diligently
for the young
child;
and
when ye have
found
{\ADD{him}},
bring
me word
again,
that
I may
come and
worship
him
also.
\VS{9}When they had
heard the
king, they
departed;
and,
lo, the
star,
which they
saw
in the
east, went
before
them,
till it
came and
stood
over
where the young
child
was.
\VS{10}When they
saw the
star, they
rejoiced with
exceeding
great
joy.
\par }{\PP \VS{11}¶
And when they were
come
into the
house, they
saw the young
child
with
Mary
his
mother,
and fell
down, and
worshipped
him: and
when they had
opened
their
treasures, they
presented unto
him
gifts;
gold,
and
frankincense,
and
myrrh.
\VS{12}And being warned of
God
in a
dream that they
should
not
return
to
Herod, they
departed
into their
own
country
another
way.
\VS{13}And
when
they were
departed,
behold, the
angel of the
Lord
appeareth to
Joseph
in a
dream,
saying,
Arise, and
take the young
child
and
his
mother,
and
flee
into
Egypt,
and be
thou
there
until I
bring
thee
word:
for
Herod
will
seek the young
child to
destroy
him.
\VS{14}When he
arose, he
took the young
child
and
his
mother by
night,
and
departed
into
Egypt:
\VS{15}And
was
there
until the
death of
Herod:
that it might be
fulfilled which was
spoken
of the
Lord
by the
prophet,
saying, Out
of
Egypt have I
called
my
son.
\par }{\PP \VS{16}¶
Then
Herod, when he
saw
that he was
mocked
of the wise
men,
was
exceeding
wroth,
and sent
forth, and
slew
all the
children that
were
in
Bethlehem,
and
in
all the
coasts
thereof,
from two years
old
and
under, according
to the
time
which he had diligently
enquired
of the wise
men.
\VS{17}Then was
fulfilled that which was
spoken
by
Jeremy the
prophet,
saying,
\VS{18}In
Rama was
there a
voice
heard,
lamentation,
and
weeping,
and
great
mourning,
Rachel
weeping
{\ADD{for}}
her
children,
and
would
not be
comforted,
because they
are
not.
\par }{\PP \VS{19}¶ But
when
Herod was
dead,
behold, an
angel of the
Lord
appeareth
in a
dream to
Joseph
in
Egypt,
\VS{20}Saying,
Arise, and
take the young
child
and
his
mother,
and
go
into the
land of
Israel:
for they are
dead
which
sought the young
child’s
life.
\VS{21}And he
arose, and
took the young
child
and
his
mother,
and
came
into the
land of
Israel.
\VS{22}But when he
heard
that
Archelaus did
reign
in
Judaea in the room
of
his
father
Herod, he was
afraid to
go
thither:
notwithstanding, being warned of
God
in a
dream, he turned
aside
into the
parts of
Galilee:
\VS{23}And he
came and
dwelt
in a
city
called
Nazareth:
that it might be
fulfilled
which was
spoken
by the
prophets, He shall be
called a
Nazarene.

\par }\Chap{3}{\PP \VerseOne{1}In
those
days
came
John the
Baptist,
preaching
in the
wilderness of
Judaea,
\VS{2}And
saying, Repent
ye:
for the
kingdom of
heaven is at
hand.
\VS{3}For
this
is he
that was
spoken
of by the
prophet
Esaias,
saying, The
voice of one
crying
in the
wilderness, Prepare
ye the
way of the
Lord,
make
his
paths
straight.
\VS{4}And the
same
John
had
his
raiment
of
camel’s
hair,
and a
leathern
girdle
about
his
loins;
and
his
meat
was
locusts
and
wild
honey.
\VS{5}Then went
out
to
him
Jerusalem,
and
all
Judaea,
and
all the region round
about
Jordan,
\VS{6}And were
baptized
of
him
in
Jordan,
confessing
their
sins.
\par }{\PP \VS{7}¶
But when he
saw
many of the
Pharisees
and
Sadducees
come
to
his
baptism, he
said unto
them, O
generation of
vipers,
who hath
warned
you to
flee
from the
wrath to
come?
\VS{8}Bring
forth
therefore
fruits
meet for
repentance:
\VS{9}And
think
not to
say
within
yourselves, We
have
Abraham
to
{\ADD{our}}
father:
for I
say unto
you,
that
God is
able
of
these
stones to raise
up
children unto
Abraham.
\VS{10}And
now
also the
axe is
laid
unto the
root of the
trees:
therefore
every
tree which
bringeth
not
forth
good
fruit is hewn
down,
and
cast
into the
fire.
\VS{11}I
indeed
baptize
you
with
water
unto
repentance:
but he
that
cometh
after
me
is mightier
than
I,
whose
shoes I
am
not
worthy to
bear:
he shall
baptize
you
with the
Holy
Ghost,
and
{\ADD{with}}
fire:
\VS{12}Whose
fan
{\ADD{is}}
in
his
hand,
and he will throughly
purge
his
floor,
and
gather
his
wheat
into the
garner;
but he will burn
up the
chaff with
unquenchable
fire.
\par }{\PP \VS{13}¶
Then
cometh
Jesus
from
Galilee
to
Jordan
unto
John, to be
baptized
of
him.
\VS{14}But
John
forbad
him,
saying,
I
have
need to be
baptized
of
thee,
and
comest
thou
to
me?
\VS{15}And
Jesus
answering
said
unto
him,
{\WJ{Suffer
{\ADD{
it to be so}}
now:
for
thus
it becometh
us
to fulfil
all
righteousness.}}
Then he
suffered
him.
\VS{16}And
Jesus, when he was
baptized, went
up
straightway
out of the
water:
and,
lo, the
heavens were
opened unto
him,
and he
saw the
Spirit of
God
descending
like a
dove,
and
lighting
upon
him:
\VS{17}And
lo a
voice
from
heaven,
saying,
This
is
my
beloved
Son,
in
whom I am well
pleased.

\par }\Chap{4}{\PP \VerseOne{1}Then
was
Jesus led
up
of the
Spirit
into the
wilderness to be
tempted
of the
devil.
\VS{2}And when he had
fasted
forty
days
and
forty
nights, he
was
afterward an
hungred.
\VS{3}And when the
tempter
came to
him, he
said,
If thou
be the
Son of
God,
command
that
these
stones be
made
bread.
\VS{4}But he
answered and
said,
{\WJ{It is written,
Man
shall
not
live
by
bread
alone,
but
by
every
word
that proceedeth
out
of the mouth
of God.}}
\VS{5}Then the
devil
taketh
him
up
into the
holy
city,
and
setteth
him
on a
pinnacle of the
temple,
\VS{6}And
saith unto
him,
If thou
be the
Son of
God,
cast
thyself
down:
for it is
written, He shall
give
his
angels
charge
concerning
thee:
and
in
{\ADD{their}}
hands they shall
bear
thee
up, lest at any
time thou
dash
thy
foot
against a
stone.
\VS{7}Jesus
said unto
him,
{\WJ{It is written
again,
Thou shalt
not
tempt
the Lord
thy
God.}}
\VS{8}Again, the
devil
taketh
him
up
into an
exceeding
high
mountain,
and
sheweth
him
all the
kingdoms of the
world,
and the
glory of
them;
\VS{9}And
saith unto
him,
All these
things will I
give
thee,
if thou wilt fall
down and
worship
me.
\VS{10}Then
saith
Jesus unto
him,
{\WJ{Get thee hence,
Satan:
for
it is written,
Thou shalt worship
the Lord
thy
God,
and
him
only
shalt thou serve.}}
\VS{11}Then the
devil
leaveth
him,
and,
behold,
angels
came
and
ministered unto
him.
\par }{\PP \VS{12}¶
Now when
Jesus had
heard
that
John was cast into
prison, he
departed
into
Galilee;
\VS{13}And
leaving
Nazareth, he
came and
dwelt
in
Capernaum,
which is upon the sea
coast,
in the
borders of
Zabulon
and
Nephthalim:
\VS{14}That it might be
fulfilled
which was
spoken
by
Esaias the
prophet,
saying,
\VS{15}The
land of
Zabulon,
and the
land of
Nephthalim,
{\ADD{by}} the
way of the
sea,
beyond
Jordan,
Galilee of the
Gentiles;
\VS{16}The
people
which
sat
in
darkness
saw
great
light;
and to
them
which
sat
in the
region
and
shadow of
death
light is sprung
up.
\par }{\PP \VS{17}¶
From that
time
Jesus
began to
preach,
and to
say,
{\WJ{Repent:
for
the kingdom
of heaven
is at hand.}}
\par }{\PP \VS{18}¶
And
Jesus,
walking
by the
sea of
Galilee,
saw
two
brethren,
Simon
called
Peter,
and
Andrew
his
brother,
casting a
net
into the
sea:
for they
were
fishers.
\VS{19}And he
saith unto
them,
{\WJ{Follow
me,
and
I will make
you
fishers
of men.}}
\VS{20}And
they
straightway
left
{\ADD{their}}
nets, and
followed
him.
\VS{21}And going
on from
thence, he
saw
other
two
brethren,
James
{\ADD{the son}} of
Zebedee,
and
John
his
brother,
in a
ship
with
Zebedee
their
father,
mending
their
nets;
and he
called
them.
\VS{22}And they
immediately
left the
ship
and
their
father, and
followed
him.
\par }{\PP \VS{23}¶
And
Jesus went
about
all
Galilee,
teaching
in
their
synagogues,
and
preaching the
gospel of the
kingdom,
and
healing all manner
of
sickness
and all manner
of
disease
among the
people.
\VS{24}And
his
fame
went
throughout
all
Syria:
and they
brought unto
him
all sick
people that were
taken with
divers
diseases
and
torments,
and those which were possessed with
devils,
and those which were
lunatick,
and those that had the
palsy;
and he
healed
them.
\VS{25}And there
followed
him great
multitudes of
people
from
Galilee,
and
{\ADD{from}}
Decapolis,
and
{\ADD{from}}
Jerusalem,
and
{\ADD{from}}
Judaea,
and
{\ADD{from}}
beyond
Jordan.

\par }\Chap{5}{\PP \VerseOne{1}And
seeing the
multitudes, he went
up
into a
mountain:
and
when
he was
set,
his
disciples came
unto
him:
\VS{2}And he
opened
his
mouth, and
taught
them,
saying,
\VS{3}{\WJ{Blessed
{\ADD{are}}
the poor
in spirit:
for
theirs
is
the kingdom
of heaven.}}
\VS{4}{\WJ{Blessed
{\ADD{are}}
they that mourn:
for
they
shall be comforted.}}
\VS{5}{\WJ{Blessed
{\ADD{are}}
the meek:
for
they
shall inherit
the earth.}}
\VS{6}{\WJ{Blessed
{\ADD{are}}
they which do hunger
and
thirst
after righteousness:
for
they
shall be filled.}}
\VS{7}{\WJ{Blessed
{\ADD{are}}
the merciful:
for
they
shall obtain mercy.}}
\VS{8}{\WJ{Blessed
{\ADD{are}}
the pure
in heart:
for
they
shall see
God.}}
\VS{9}{\WJ{Blessed
{\ADD{are}}
the peacemakers:
for
they
shall be called
the children
of God.}}
\VS{10}{\WJ{Blessed
{\ADD{are}}
they which are persecuted
for
righteousness’
sake:
for
theirs
is
the kingdom
of heaven.}}
\VS{11}{\WJ{Blessed
are ye,
when
{\ADD{men}}
shall revile
you,
and
persecute
{\ADD{you}},
and
shall say
all manner of
evil
against
you
falsely,
for
my
sake.}}
\VS{12}{\WJ{Rejoice,
and
be exceeding glad:
for
great
{\ADD{is}}
your
reward
in
heaven:
for
so
persecuted they
the prophets
which were
before
you.}}
\par }{\PP \VS{13}{\WJ{¶
Ye
are
the salt
of the earth:
but
if
the salt
have lost his savour,
wherewith
shall it be salted?
it is
thenceforth
good
for
nothing,
but
to be cast
out,
and
to be trodden under foot
of
men.}}
\VS{14}{\WJ{Ye
are
the light
of the world.
A city
that is set
on
an hill
cannot
be hid.}}
\VS{15}{\WJ{Neither do men
light
a candle,
and
put
it
under
a bushel,
but
on
a candlestick;
and
it giveth light
unto all
that are
in
the house.}}
\VS{16}{\WJ{Let
your
light
so
shine
before
men,
that
they may see
your
good
works,
and
glorify
your
Father
which is
in
heaven.}}
\par }{\PP \VS{17}{\WJ{¶
Think
not
that
I am come
to destroy
the law,
or
the prophets:
I am
not
come
to destroy,
but
to fulfil.}}
\VS{18}{\WJ{For
verily
I say
unto you,
Till
heaven
and
earth
pass,
one
jot
or
one
tittle
shall
in no
wise
pass
from
the law,
till
all
be fulfilled.}}
\VS{19}{\WJ{Whosoever
therefore
shall break
one
of these
least
commandments,
and
shall teach
men
so, he
shall be called
the least
in
the kingdom
of heaven:
but
whosoever
shall do
and
teach
{\ADD{them}},
the same
shall be called
great
in
the kingdom
of heaven.}}
\VS{20}{\WJ{For
I say
unto you,
That
except
your
righteousness
shall exceed
{\ADD{the righteousness}}
of the scribes
and
Pharisees,
ye shall
in no
case
enter
into
the kingdom
of heaven.}}
\par }{\PP \VS{21}{\WJ{¶
Ye have heard
that
it was said
by them of old time,
Thou shalt
not
kill;
and
whosoever
shall kill
shall be
in danger
of the judgment:}}
\VS{22}{\WJ{But
I
say
unto you,
That
whosoever
is angry
with
his
brother
without a cause
shall be
in danger
of the judgment:
and
whosoever
shall say
to
his
brother,
Raca,
shall be
in danger
of the council:
but
whosoever
shall say,
Thou fool,
shall be
in danger
of
hell
fire.}}
\VS{23}{\WJ{Therefore
if
thou bring
thy
gift
to
the altar,
and there
rememberest
that
thy
brother
hath
ought
against
thee;}}
\VS{24}{\WJ{Leave
there
thy
gift
before
the altar,
and
go thy way;
first
be reconciled
to
thy
brother,
and
then
come
and offer
thy
gift.}}
\VS{25}{\WJ{Agree with
thine
adversary
quickly,
whiles
thou art
in
the way
with
him;
lest at any time
the adversary
deliver
thee
to the judge,
and
the judge
deliver
thee
to the officer,
and
thou be cast
into
prison.}}
\VS{26}{\WJ{Verily
I say
unto thee, Thou shalt
by no
means
come out
thence,
till
thou hast paid
the uttermost
farthing.}}
\par }{\PP \VS{27}{\WJ{¶
Ye have heard
that
it was said
by them of old time,
Thou shalt
not
commit adultery:}}
\VS{28}{\WJ{But
I
say
unto you,
That
whosoever
looketh
on a woman
to
lust after
her
hath committed adultery
with her
already
in
his
heart.}}
\VS{29}{\WJ{And
if
thy
right
eye
offend
thee,
pluck
it
out,
and
cast
{\ADD{it}}
from
thee:
for
it is profitable
for thee
that
one
of
thy
members
should perish,
and
not
{\ADD{that}}
thy
whole
body
should be cast
into
hell.}}
\VS{30}{\WJ{And
if
thy
right
hand
offend
thee,
cut
it
off,
and
cast
{\ADD{it}}
from
thee:
for
it is profitable
for thee
that
one
of
thy
members
should perish,
and
not
{\ADD{that}}
thy
whole
body
should be cast
into
hell.}}
\VS{31}{\WJ{It hath been said,
Whosoever
shall put away
his
wife,
let him give
her
a writing of divorcement:}}
\VS{32}{\WJ{But
I
say
unto you,
That
whosoever
shall put away
his
wife,
saving
for the cause
of fornication,
causeth
her
to commit adultery:
and
whosoever
shall marry
her that is divorced
committeth adultery.}}
\par }{\PP \VS{33}{\WJ{¶
Again,
ye have heard
that
it hath been said
by them of old time,
Thou shalt
not
forswear thyself,
but
shalt perform
unto the Lord
thine
oaths:}}
\VS{34}{\WJ{But
I
say
unto you,
Swear
not
at all;
neither
by
heaven;
for
it is
God’s
throne:}}
\VS{35}{\WJ{Nor
by
the earth;
for
it is
his
footstool:
neither
by
Jerusalem;
for
it is
the city
of the
great
King.}}
\VS{36}{\WJ{Neither
shalt thou swear
by
thy
head,
because
thou canst
not
make
one
hair
white
or
black.}}
\VS{37}{\WJ{But
let
your
communication
be,
Yea,
yea;
Nay,
nay:
for
whatsoever is more than
these
cometh
of
evil.}}
\par }{\PP \VS{38}{\WJ{¶
Ye have heard
that
it hath been said,
An eye
for
an eye,
and
a tooth
for
a tooth:}}
\VS{39}{\WJ{But
I
say
unto you,
That ye resist
not
evil:
but
whosoever
shall smite
thee
on
thy
right
cheek,
turn
to him
the other
also.}}
\VS{40}{\WJ{And if
any man will
sue
thee
at the law,
and
take away
thy
coat,
let
him
have
{\ADD{thy}}
cloke
also.}}
\VS{41}{\WJ{And
whosoever
shall compel
thee
to go
a
mile,
go
with
him
twain.}}
\VS{42}{\WJ{Give
to him that asketh
thee,
and
from
him that would
borrow
of thee
turn
not
thou away.}}
\par }{\PP \VS{43}{\WJ{¶
Ye have heard
that
it hath been said,
Thou shalt love
thy
neighbour,
and
hate
thine
enemy.}}
\VS{44}{\WJ{But
I
say
unto you,
Love
your
enemies,
bless
them that curse
you,
do
good
to them that hate
you,
and
pray
for
them which despitefully use
you,
and
persecute
you;}}
\VS{45}{\WJ{That
ye may be
the children
of your
Father
which is
in
heaven:
for
he maketh
his
sun
to rise
on
the evil
and
on the good,
and
sendeth rain
on
the just
and
on the unjust.}}
\VS{46}{\WJ{For
if
ye love
them which love
you,
what
reward
have ye?
do
not
even
the publicans
the same?}}
\VS{47}{\WJ{And
if
ye salute
your
brethren
only,
what
do ye
more
{\ADD{than others}}?
do
not
even
the publicans
so?}}
\VS{48}{\WJ{Be
ye
therefore
perfect,
even as
your
Father
which is
in
heaven
is
perfect.}}

\par }\Chap{6}{\PP \VerseOne{1}{\WJ{Take heed that
ye do
not
your
alms
before
men,
to
be seen
of them:
otherwise
ye have
no
reward
of
your
Father
which is
in
heaven.}}
\VS{2}{\WJ{Therefore
when
thou doest
{\ADD{thine}}
alms,
do not
sound a trumpet
before
thee,
as
the hypocrites
do
in
the synagogues
and
in
the streets,
that
they may have glory
of
men.
Verily
I say
unto you,
They have
their
reward.}}
\VS{3}{\WJ{But
when
thou
doest
alms,
let
not
thy
left hand
know
what
thy
right hand
doeth:}}
\VS{4}{\WJ{That
thine
alms
may be
in
secret:
and
thy
Father
which seeth
in
secret
himself
shall reward
thee
openly.}}
\par }{\PP \VS{5}{\WJ{¶
And
when
thou prayest,
thou shalt
not
be
as
the hypocrites
{\ADD{are}}:
for
they love
to pray
standing
in
the synagogues
and
in
the corners
of the streets,
that
they
may be seen
of men.
Verily
I say
unto you,
They have
their
reward.}}
\VS{6}{\WJ{But
thou,
when
thou prayest,
enter
into
thy
closet,
and
when thou hast shut
thy
door,
pray
to
thy
Father
which is
in
secret;
and
thy
Father
which seeth
in
secret
shall reward
thee
openly.}}
\VS{7}{\WJ{But
when ye pray,
use
not
vain repetitions,
as
the heathen
{\ADD{do}}:
for
they think
that
they shall be heard
for
their
much speaking.}}
\VS{8}{\WJ{Be
not
ye therefore
like
unto them:
for
your
Father
knoweth
what things
ye have
need of,
before
ye
ask
him.}}
\VS{9}{\WJ{After this manner
therefore
pray
ye:
Our
Father
which art
in
heaven,
Hallowed be
thy
name.}}
\VS{10}{\WJ{Thy
kingdom
come.
Thy
will
be done
in
earth,
as
{\ADD{it is}}
in
heaven.}}
\VS{11}{\WJ{Give
us
this day
our
daily
bread.}}
\VS{12}{\WJ{And
forgive
us
our
debts,
as
we
forgive
our
debtors.}}
\VS{13}{\WJ{And
lead
us
not
into
temptation,
but
deliver
us
from
evil:
For
thine
is
the kingdom,
and
the power,
and
the glory,
for
ever.
Amen.}}
\VS{14}{\WJ{For
if
ye forgive
men
their
trespasses,
your
heavenly
Father
will
also
forgive
you:}}
\VS{15}{\WJ{But
if
ye forgive
not
men
their
trespasses,
neither
will
your
Father
forgive
your
trespasses.}}
\par }{\PP \VS{16}{\WJ{¶
Moreover
when
ye fast,
be
not,
as
the hypocrites,
of a sad countenance:
for
they disfigure
their
faces,
that
they may appear
unto men
to fast.
Verily
I say
unto you,
They have
their
reward.}}
\VS{17}{\WJ{But
thou,
when thou fastest,
anoint
thine
head,
and
wash
thy
face;}}
\VS{18}{\WJ{That
thou appear
not
unto men
to fast,
but
unto
thy
Father
which is
in
secret:
and
thy
Father,
which seeth
in
secret,
shall reward
thee
openly.}}
\par }{\PP \VS{19}{\WJ{¶
Lay
not
up
for yourselves
treasures
upon
earth,
where
moth
and
rust
doth corrupt,
and
where
thieves
break through
and
steal:}}
\VS{20}{\WJ{But
lay up
for yourselves
treasures
in
heaven,
where
neither
moth
nor
rust
doth corrupt,
and
where
thieves
do
not
break through
nor
steal:}}
\VS{21}{\WJ{For
where
your
treasure
is,
there
will
your
heart
be
also.}}
\VS{22}{\WJ{The light
of the body
is
the eye:
if
therefore
thine
eye
be
single,
thy
whole
body
shall be
full of light.}}
\VS{23}{\WJ{But
if
thine
eye
be
evil,
thy
whole
body
shall be
full of darkness.
If
therefore
the light
that is
in
thee
be
darkness,
how great
{\ADD{is}}
that darkness!}}
\par }{\PP \VS{24}{\WJ{¶
No man
can
serve
two
masters:
for
either
he will hate
the one,
and
love
the other;
or else
he will hold to
the one,
and
despise
the other.
Ye cannot
serve
God
and
mammon.}}
\VS{25}{\WJ{Therefore
I say
unto you,
Take
no
thought
for
your
life,
what
ye shall eat,
or
what
ye shall drink;
nor yet
for
your
body,
what
ye shall put on.
Is
not
the life
more
than meat,
and
the body
than raiment?}}
\VS{26}{\WJ{Behold
the fowls
of the air:
for
they sow
not,
neither
do they reap,
nor
gather
into
barns;
yet
your
heavenly
Father
feedeth
them.
Are
ye
not
much
better
than they?}}
\VS{27}{\WJ{Which
of
you
by taking thought
can
add
one
cubit
unto
his
stature?}}
\VS{28}{\WJ{And
why
take ye thought
for
raiment?
Consider
the lilies
of the field,
how
they grow;
they toil
not,
neither
do they spin:}}
\VS{29}{\WJ{And yet
I say
unto you,
That
even
Solomon
in
all
his
glory
was
not
arrayed
like
one
of these.}}
\VS{30}{\WJ{Wherefore,
if
God
so
clothe
the grass
of the field,
which
to day
is,
and
to morrow
is cast
into
the oven,
{\ADD{shall he}}
not
much
more
{\ADD{clothe}}
you,
O ye of little faith?}}
\VS{31}{\WJ{Therefore
take
no
thought,
saying,
What
shall we eat?
or,
What
shall we drink?
or,
Wherewithal
shall we be clothed?}}
\VS{32}{\WJ{(For
after all
these things
do
the Gentiles
seek:)
for
your
heavenly
Father
knoweth
that
ye have need
of
all
these things.}}
\VS{33}{\WJ{But
seek ye
first
the kingdom
of God,
and
his
righteousness;
and
all
these things
shall be added
unto you.}}
\VS{34}{\WJ{Take
therefore
no
thought
for
the morrow:
for
the morrow
shall take thought
for the things
of itself.
Sufficient
unto the day
{\ADD{is}}
the evil
thereof.}}

\par }\Chap{7}{\PP \VerseOne{1}{\WJ{Judge
not,
that
ye be
not
judged.}}
\VS{2}{\WJ{For
with
what
judgment
ye judge,
ye shall be judged:
and
with
what
measure
ye mete,
it shall be measured
to you
again.}}
\VS{3}{\WJ{And
why
beholdest thou
the mote
that is
in
thy
brother’s
eye,
but
considerest
not
the beam that is
in
thine own
eye?}}
\VS{4}{\WJ{Or
how
wilt thou say
to
thy
brother,
Let
me pull out
the mote
out
of
thine
eye;
and,
behold,
a beam
{\ADD{is}}
in
thine own
eye?}}
\VS{5}{\WJ{Thou hypocrite,
first
cast out
the beam
out
of
thine own
eye;
and
then
shalt thou see clearly
to cast out
the mote
out
of
thy
brother’s
eye.}}
\par }{\PP \VS{6}{\WJ{¶
Give
not
that which is holy
unto the dogs,
neither
cast ye
your
pearls
before
swine,
lest
they trample
them
under
their
feet,
and
turn again and
rend
you.}}
\par }{\PP \VS{7}{\WJ{¶
Ask,
and
it shall be given
you;
seek,
and
ye shall find;
knock,
and
it shall be opened
unto you:}}
\VS{8}{\WJ{For
every
one that asketh
receiveth;
and
he that seeketh
findeth;
and
to him that knocketh
it shall be opened.}}
\VS{9}{\WJ{Or
what
man
is there
of
you,
whom
if
his
son
ask
bread,
will he give
him
a stone?}}
\VS{10}{\WJ{Or
if
he ask
a fish,
will he give
him
a serpent?}}
\VS{11}{\WJ{If
ye
then,
being
evil,
know how
to give
good
gifts
unto
your
children,
how much
more
shall
your
Father
which is
in
heaven
give
good things
to them that
ask
him?}}
\VS{12}{\WJ{Therefore
all things
whatsoever
ye would
that
men
should do
to you,
do
ye
even
so
to them:
for
this
is
the law
and
the prophets.}}
\par }{\PP \VS{13}{\WJ{¶
Enter ye
in at
the strait
gate:
for
wide
{\ADD{is}}
the gate,
and
broad
{\ADD{is}}
the way,
that leadeth
to
destruction,
and
many
there be
which go in
thereat:}}
\VS{14}{\WJ{Because
strait
{\ADD{is}}
the gate,
and
narrow
{\ADD{is}}
the way,
which leadeth
unto
life,
and
few
there be
that find
it.}}
\par }{\PP \VS{15}{\WJ{¶
Beware
of
false prophets,
which
come
to
you
in
sheep’s
clothing,
but
inwardly
they are
ravening
wolves.}}
\VS{16}{\WJ{Ye shall know
them
by
their
fruits.
Do
men gather
grapes
of
thorns,
or
figs
of
thistles?}}
\VS{17}{\WJ{Even so
every
good
tree
bringeth forth
good
fruit;
but
a corrupt
tree
bringeth forth
evil
fruit.}}
\VS{18}{\WJ{A good
tree
cannot
bring forth
evil
fruit,
neither
{\ADD{can}}
a corrupt
tree
bring forth
good
fruit.}}
\VS{19}{\WJ{Every
tree
that bringeth
not
forth
good
fruit
is hewn down,
and
cast
into
the fire.}}
\VS{20}{\WJ{Wherefore
by
their
fruits
ye shall know
them.}}
\par }{\PP \VS{21}{\WJ{¶
Not
every
one that saith
unto me,
Lord,
Lord,
shall enter
into
the kingdom
of heaven;
but
he that doeth
the will
of my
Father
which is in
heaven.}}
\VS{22}{\WJ{Many
will say
to me
in
that
day,
Lord,
Lord,
have we
not
prophesied
in
thy
name?
and
in
thy
name
have cast out
devils?
and
in
thy
name
done
many
wonderful works?}}
\VS{23}{\WJ{And
then
will I profess
unto them,
I
never
knew
you:
depart
from
me,
ye that work
iniquity.}}
\par }{\PP \VS{24}{\WJ{¶
Therefore
whosoever
heareth
these
sayings
of mine,
and
doeth
them,
I will liken
him
unto a
wise
man,
which
built
his
house
upon
a rock:}}
\VS{25}{\WJ{And
the rain
descended,
and
the floods
came,
and
the winds
blew,
and
beat upon
that
house;
and
it fell
not:
for
it was founded
upon
a rock.}}
\VS{26}{\WJ{And
every
one that heareth
these
sayings
of mine,
and
doeth
them
not,
shall be likened
unto a
foolish
man,
which
built
his
house
upon
the sand:}}
\VS{27}{\WJ{And
the rain
descended,
and
the floods
came,
and
the winds
blew,
and
beat upon
that
house;
and
it fell:
and
great
was
the fall
of it.}}
\VS{28}And it came to
pass,
when
Jesus had
ended
these
sayings, the
people were
astonished
at
his
doctrine:
\VS{29}For
he
taught
them
as
{\ADD{one}}
having
authority,
and
not
as the
scribes.

\par }\Chap{8}{\PP \VerseOne{1}When
he was come
down
from the
mountain,
great
multitudes
followed
him.
\VS{2}And,
behold, there
came a
leper and
worshipped
him,
saying,
Lord,
if thou
wilt, thou
canst
make
me
clean.
\VS{3}And
Jesus put
forth
{\ADD{his}}
hand, and
touched
him,
saying,
{\WJ{I will;
be thou clean.}}
And
immediately
his
leprosy was
cleansed.
\VS{4}And
Jesus
saith unto
him,
{\WJ{See
thou tell
no man;
but
go thy way,
shew
thyself
to the priest,
and
offer
the gift
that
Moses
commanded,
for
a testimony
unto them.}}
\par }{\PP \VS{5}¶
And
when
Jesus was
entered
into
Capernaum, there
came unto
him a
centurion,
beseeching
him,
\VS{6}And
saying,
Lord,
my
servant
lieth
at
home sick of the
palsy,
grievously
tormented.
\VS{7}And
Jesus
saith unto
him,
{\WJ{I
will come and
heal
him.}}
\VS{8}The
centurion
answered
and
said,
Lord, I
am
not
worthy
that thou shouldest
come
under
my
roof:
but
speak the
word
only,
and
my
servant shall be
healed.
\VS{9}For
I
am a
man
under
authority,
having
soldiers
under
me:
and I
say to
this
{\ADD{man}},
Go,
and he
goeth;
and to
another,
Come,
and he
cometh;
and
to
my
servant,
Do
this,
and he
doeth
{\ADD{it}}.
\VS{10}When
Jesus
heard
{\ADD{it}}, he
marvelled,
and
said to them that
followed,
{\WJ{Verily
I say
unto you,
I have
not
found
so great
faith, no,
not
in
Israel.}}
\VS{11}{\WJ{And
I say
unto you,
That
many
shall come
from
the east
and
west,
and
shall sit down
with
Abraham,
and
Isaac,
and
Jacob,
in
the kingdom
of heaven.}}
\VS{12}{\WJ{But
the children
of the kingdom
shall be cast out
into
outer
darkness:
there
shall be
weeping
and
gnashing
of teeth.}}
\VS{13}And
Jesus
said unto the
centurion,
{\WJ{Go thy way;
and
as
thou hast believed,
{\ADD{so}}
be it done
unto thee.
And
his
servant
was healed
in
the
selfsame
hour.}}
\par }{\PP \VS{14}¶
And
when
Jesus was
come
into
Peter’s
house, he
saw
his
wife’s mother
laid,
and sick of a
fever.
\VS{15}And he
touched
her
hand,
and the
fever
left
her:
and she
arose,
and
ministered unto
them.
\par }{\PP \VS{16}¶
When the
even was
come, they
brought unto
him
many that were possessed with
devils:
and he cast
out the
spirits
with
{\ADD{his}}
word,
and
healed
all that
were
sick:
\VS{17}That it might be
fulfilled which was
spoken
by
Esaias the
prophet,
saying,
Himself
took
our
infirmities,
and
bare
{\ADD{our}}
sicknesses.
\par }{\PP \VS{18}¶
Now
when
Jesus
saw
great
multitudes
about
him, he gave
commandment to
depart
unto the other
side.
\VS{19}And a
certain
scribe
came, and
said unto
him,
Master, I will
follow
thee
whithersoever thou
goest.
\VS{20}And
Jesus
saith unto
him,
{\WJ{The foxes
have
holes,
and
the birds
of the air
{\ADD{have}}
nests;
but
the Son
of man
hath
not
where
to lay
{\ADD{his}}
head.}}
\VS{21}And
another
of
his
disciples
said unto
him,
Lord,
suffer
me
first to
go
and
bury
my
father.
\VS{22}But
Jesus
said unto
him,
{\WJ{Follow
me;
and
let
the dead
bury
their
dead.}}
\par }{\PP \VS{23}¶
And
when
he was
entered
into a
ship,
his
disciples
followed
him.
\VS{24}And,
behold, there
arose a
great
tempest
in the
sea, insomuch
that the
ship was
covered
with the
waves:
but
he was
asleep.
\VS{25}And
his
disciples came
to
{\ADD{him}}, and
awoke
him,
saying,
Lord,
save
us: we
perish.
\VS{26}And he
saith unto
them,
{\WJ{Why
are ye
fearful,
O ye of little faith?}}
Then he
arose, and
rebuked the
winds
and the
sea;
and there
was a
great
calm.
\VS{27}But the
men
marvelled,
saying, What manner of
man
is
this,
that
even the
winds
and the
sea
obey
him!
\par }{\PP \VS{28}¶
And
when
he was
come
to the other
side
into the
country of the
Gergesenes, there
met
him
two possessed with
devils, coming
out
of the
tombs,
exceeding
fierce, so
that
no
man
might
pass
by
that
way.
\VS{29}And,
behold, they cried
out,
saying,
What have
we to
do with
thee,
Jesus, thou
Son of
God? art thou
come
hither to
torment
us
before the
time?
\VS{30}And there
was a good way
off
from
them an
herd of
many
swine
feeding.
\VS{31}So the
devils
besought
him,
saying,
If thou
cast
us
out,
suffer
us to go
away
into the
herd of
swine.
\VS{32}And he
said unto
them,
{\WJ{Go.}}
And when they were come
out, they
went
into the
herd of
swine:
and,
behold, the
whole
herd of
swine ran
violently
down a steep
place
into the
sea,
and
perished
in the
waters.
\VS{33}And they that kept
them
fled,
and went their
ways
into the
city, and
told every
thing,
and what was
befallen to the possessed of the
devils.
\VS{34}And,
behold, the
whole
city came
out
to
meet
Jesus:
and when they
saw
him, they
besought
{\ADD{him}}
that he would
depart
out
of
their
coasts.

\par }\Chap{9}{\PP \VerseOne{1}And he
entered
into a
ship, and passed
over,
and
came
into his
own
city.
\VS{2}And,
behold, they
brought to
him a man sick of the
palsy,
lying
on a
bed:
and
Jesus
seeing
their
faith
said unto the sick of the
palsy;
{\WJ{Son,
be of good cheer;
thy
sins
be forgiven
thee.}}
\VS{3}And,
behold,
certain of the
scribes
said
within
themselves,
This
{\ADD{man}}
blasphemeth.
\VS{4}And
Jesus
knowing
their
thoughts
said,
{\WJ{Wherefore
think
ye
evil
in
your
hearts?}}
\VS{5}{\WJ{For
whether
is
easier,
to say,
{\ADD{Thy}}
sins
be forgiven
thee;
or
to say,
Arise,
and
walk?}}
\VS{6}{\WJ{But
that
ye may know
that
the Son
of man
hath
power
on
earth
to forgive
sins,}} (then saith
he to the sick of the
palsy,)
{\WJ{Arise,
take up
thy
bed,
and
go
unto
thine
house.}}
\VS{7}And he
arose, and
departed
to
his
house.
\VS{8}But
when the
multitudes
saw
{\ADD{it}}, they
marvelled,
and
glorified
God, which had
given
such
power unto
men.
\par }{\PP \VS{9}¶
And
as
Jesus passed
forth from
thence, he
saw a
man,
named
Matthew,
sitting
at the receipt of
custom:
and he
saith unto
him,
{\WJ{Follow
me.}}
And he arose,
and
followed
him.
\par }{\PP \VS{10}¶
And it came to
pass,
as
Jesus sat at
meat
in the
house,
behold,
many
publicans
and
sinners came
and sat down
with
him
and
his
disciples.
\VS{11}And
when the
Pharisees
saw
{\ADD{it}}, they
said
unto
his
disciples,
Why
eateth
your
Master
with
publicans
and
sinners?
\VS{12}But
when
Jesus
heard
{\ADD{that}}, he
said unto
them,
{\WJ{They that be whole
need
not
a physician,
but
they that are sick.}}
\VS{13}{\WJ{But
go ye and
learn
what
{\ADD{that}}
meaneth,
I will have
mercy,
and
not
sacrifice:
for
I am
not
come
to call
the righteous,
but
sinners
to
repentance.}}
\par }{\PP \VS{14}¶
Then came
to
him the
disciples of
John,
saying,
Why
do
we
and the
Pharisees
fast
oft,
but
thy
disciples
fast
not?
\VS{15}And
Jesus
said unto
them,
{\WJ{Can
the children
of the bridechamber
mourn,
as
long as
the bridegroom
is
with
them?
but
the days
will come,
when
the bridegroom
shall be taken
from
them,
and
then
shall they fast.}}
\VS{16}{\WJ{No man
putteth
a piece
of new
cloth
unto
an
old
garment,
for
that which is put in to fill
it
up
taketh
from
the garment,
and
the rent
is made
worse.}}
\VS{17}{\WJ{Neither
do men put
new
wine
into
old
bottles:
else
the bottles
break,
and
the wine
runneth out,
and
the bottles
perish:
but
they put
new
wine
into
new
bottles,
and
both
are preserved.}}
\par }{\PP \VS{18}¶
While
he
spake these
things unto
them,
behold, there
came a certain
ruler,
and
worshipped
him,
saying,
My
daughter
is even
now
dead:
but come
and
lay
thy
hand
upon
her,
and she shall
live.
\VS{19}And
Jesus arose,
and
followed
him,
and
{\ADD{so did}}
his
disciples.
\par }{\PP \VS{20}¶
And,
behold, a
woman, which was diseased with an issue of
blood
twelve
years,
came
behind
{\ADD{him}}, and
touched the
hem
of
his
garment:
\VS{21}For she
said
within
herself,
If I
may
but
touch
his
garment, I shall be
whole.
\VS{22}But
Jesus turned him
about,
and when he
saw
her, he
said,
{\WJ{Daughter,
be of good comfort;
thy
faith
hath made
thee
whole.}}
And the
woman was made
whole
from
that
hour.
\VS{23}And
when
Jesus
came
into the
ruler’s
house,
and
saw the
minstrels
and the
people making a
noise,
\VS{24}He
said unto
them,
{\WJ{Give place:
for
the maid
is
not
dead,
but
sleepeth.}}
And they
laughed
him to
scorn.
\VS{25}But
when the
people were put
forth, he went in,
and
took
her by the
hand,
and the
maid
arose.
\VS{26}And the
fame
hereof went
abroad
into
all
that
land.
\par }{\PP \VS{27}¶
And
when
Jesus
departed
thence,
two blind
men
followed
him,
crying,
and
saying,
{\ADD{Thou}}
Son of
David, have mercy
on
us.
\VS{28}And when he was
come
into the
house, the blind
men
came to
him:
and
Jesus
saith unto
them,
{\WJ{Believe ye
that
I am able
to do
this?}} They
said unto
him,
Yea,
Lord.
\VS{29}Then touched
he
their
eyes,
saying,
{\WJ{According to
your
faith
be it
unto you.}}
\VS{30}And
their
eyes were
opened;
and
Jesus straitly
charged
them,
saying,
{\WJ{See
{\ADD{that}}
no man
know
{\ADD{it}}.}}
\VS{31}But
they, when they were
departed, spread
abroad
his
fame
in
all
that
country.
\par }{\PP \VS{32}¶
As
they went
out,
behold, they
brought to
him a
dumb
man possessed with a
devil.
\VS{33}And
when the
devil was cast
out, the
dumb
spake:
and the
multitudes
marvelled,
saying, It
was
never
so
seen
in
Israel.
\VS{34}But the
Pharisees
said, He casteth
out
devils
through the
prince of the
devils.
\VS{35}And
Jesus went
about
all the
cities
and
villages,
teaching
in
their
synagogues,
and
preaching the
gospel of the
kingdom,
and
healing
every
sickness
and
every
disease
among the
people.
\par }{\PP \VS{36}¶
But when he
saw the
multitudes, he was moved with
compassion
on
them,
because
they
fainted,
and
were scattered
abroad,
as
sheep
having
no
shepherd.
\VS{37}Then saith
he
unto
his
disciples,
{\WJ{The harvest
truly
{\ADD{is}}
plenteous,
but
the labourers
{\ADD{are}}
few;}}
\VS{38}{\WJ{Pray ye
therefore
the Lord
of the harvest,
that
he will send forth
labourers
into
his
harvest.}}

\par }\Chap{10}{\PP \VerseOne{1}And when he had called
unto
{\ADD{him}}
his
twelve
disciples, he
gave
them
power
{\ADD{against}}
unclean
spirits,
to
cast
them
out,
and to
heal all manner
of
sickness
and all manner
of
disease.
\VS{2}Now the
names of
the
twelve
apostles
are
these; The
first,
Simon, who is
called
Peter,
and
Andrew
his
brother;
James
{\ADD{the son}} of
Zebedee,
and
John
his
brother;
\VS{3}Philip,
and
Bartholomew;
Thomas,
and
Matthew the
publican;
James
{\ADD{the son}} of
Alphaeus,
and
Lebbaeus, whose surname
was
Thaddaeus;
\VS{4}Simon the
Canaanite,
and
Judas
Iscariot,
who
also
betrayed
him.
\VS{5}These
twelve
Jesus sent
forth, and
commanded
them,
saying,
{\WJ{Go
not
into
the way
of the Gentiles,
and
into
{\ADD{any}}
city
of the Samaritans
enter ye
not:}}
\VS{6}{\WJ{But
go
rather
to
the lost
sheep
of the house
of Israel.}}
\VS{7}{\WJ{And
as ye go,
preach,
saying,
The kingdom
of heaven
is at hand.}}
\VS{8}{\WJ{Heal
the sick,
cleanse
the lepers,
raise
the dead,
cast out
devils:
freely
ye have received,
freely
give.}}
\VS{9}{\WJ{Provide
neither
gold,
nor
silver,
nor
brass
in
your
purses,}}
\VS{10}{\WJ{Nor
scrip
for
{\ADD{your}}
journey,
neither
two
coats,
neither
shoes,
nor yet
staves:
for
the workman
is
worthy
of his
meat.}}
\VS{11}{\WJ{And
into
whatsoever
city
or
town
ye shall enter,
enquire
who
in
it
is
worthy;
and there
abide
till
ye go thence.}}
\VS{12}{\WJ{And
when ye come
into
an house,
salute
it.}}
\VS{13}{\WJ{And
if
the house
be
worthy,
let
your
peace
come
upon
it:
but
if
it be
not
worthy,
let
your
peace
return
to
you.}}
\VS{14}{\WJ{And
whosoever
shall
not
receive
you,
nor
hear
your
words,
when ye depart out
of
that
house
or
city,
shake off
the dust
of
your
feet.}}
\VS{15}{\WJ{Verily
I say
unto you,
It shall be
more tolerable
for the land
of Sodom
and
Gomorrha
in
the day
of judgment,
than
for
that
city.}}
\par }{\PP \VS{16}{\WJ{¶
Behold,
I
send
you
forth
as
sheep
in
the midst
of wolves:
be ye
therefore
wise
as
serpents,
and
harmless
as
doves.}}
\VS{17}{\WJ{But
beware
of
men:
for
they will deliver
you
up to
the councils,
and
they will scourge
you
in
their
synagogues;}}
\VS{18}{\WJ{And
ye shall be brought
before
governors
and
kings
for
my
sake,
for
a testimony
against them
and
the Gentiles.}}
\VS{19}{\WJ{But
when
they deliver
you
up,
take
no
thought
how
or
what
ye shall speak:
for
it shall be given
you
in
that same
hour
what
ye shall speak.}}
\VS{20}{\WJ{For
it is
not
ye
that speak,
but
the Spirit
of
your
Father
which speaketh
in
you.}}
\VS{21}{\WJ{And
the brother
shall deliver up
the brother
to
death,
and
the father
the child:
and
the children
shall rise up
against
{\ADD{their}}
parents,
and
cause
them
to be put to death.}}
\VS{22}{\WJ{And
ye shall be
hated
of
all
{\ADD{men}}
for
my
name’s
sake:
but
he
that endureth
to
the end
shall be saved.}}
\VS{23}{\WJ{But
when
they persecute
you
in
this
city,
flee ye
into
another:
for
verily
I say
unto you,
Ye shall
not
have gone over
the cities
of Israel,
till
the Son
of man
be come.}}
\VS{24}{\WJ{The disciple
is
not
above
{\ADD{his}}
master,
nor
the servant
above
his
lord.}}
\VS{25}{\WJ{It is enough
for the disciple
that
he be
as
his
master,
and
the servant
as
his
lord.
If
they have called
the master of the house
Beelzebub,
how much
more
{\ADD{shall they call}}
them of
his
household?}}
\VS{26}{\WJ{Fear
them
not
therefore:
for
there is
nothing
covered,
that
shall
not
be revealed;
and
hid,
that
shall
not
be known.}}
\VS{27}{\WJ{What
I tell
you
in
darkness,
{\ADD{that}}
speak ye
in
light:
and
what
ye hear
in
the ear,
{\ADD{that}}
preach ye
upon
the housetops.}}
\VS{28}{\WJ{And
fear
not
them which kill
the body,
but
are
not
able
to kill
the soul:
but
rather
fear
him which is able
to destroy
both
soul
and
body
in
hell.}}
\VS{29}{\WJ{Are
not
two
sparrows
sold
for a farthing?
and
one
of
them
shall
not
fall
on
the ground
without
your
Father.}}
\VS{30}{\WJ{But
the
very
hairs
of
your
head
are
all
numbered.}}
\VS{31}{\WJ{Fear ye
not
therefore,
ye
are of more value
than
many
sparrows.}}
\VS{32}{\WJ{Whosoever
therefore
shall confess
me
before
men,
him
will
I
confess
also
before
my
Father
which is
in
heaven.}}
\VS{33}{\WJ{But
whosoever
shall deny
me
before
men,
him
will
I also
deny
before
my
Father
which is
in
heaven.}}
\VS{34}{\WJ{Think
not
that
I am come
to send
peace
on
earth:
I came
not
to send
peace,
but
a sword.}}
\VS{35}{\WJ{For
I am come
to set
a man
at variance
against
his
father,
and
the daughter
against
her
mother,
and
the daughter in law
against
her
mother in law.}}
\VS{36}{\WJ{And
a man’s
foes
{\ADD{shall be}}
they of
his own
household.}}
\VS{37}{\WJ{He that loveth
father
or
mother
more than
me
is
not
worthy
of me:
and
he that loveth
son
or
daughter
more than
me
is
not
worthy
of me.}}
\VS{38}{\WJ{And
he
that
taketh
not
his
cross,
and
followeth
after
me,
is
not
worthy
of me.}}
\VS{39}{\WJ{He that findeth
his
life
shall lose
it:
and
he that loseth
his
life
for
my
sake
shall find
it.}}
\par }{\PP \VS{40}{\WJ{¶
He that receiveth
you
receiveth
me,
and
he that receiveth
me
receiveth
him that sent
me.}}
\VS{41}{\WJ{He that receiveth
a prophet
in
the name
of a prophet
shall receive
a prophet’s
reward;
and
he that receiveth
a righteous man
in
the name
of a righteous man
shall receive
a righteous man’s
reward.}}
\VS{42}{\WJ{And
whosoever
shall give to drink
unto
one of
these
little ones
a cup
of cold
{\ADD{water}}
only
in
the name
of a disciple,
verily
I say
unto you,
he shall
in no
wise
lose
his
reward.}}

\par }\Chap{11}{\PP \VerseOne{1}And it came to
pass,
when
Jesus had made an end
of
commanding
his
twelve
disciples, he
departed
thence to
teach
and to
preach
in
their
cities.
\VS{2}Now
when
John had
heard
in the
prison the
works of
Christ, he
sent
two
of
his
disciples,
\VS{3}And
said unto
him,
Art
thou he that should
come,
or do we look
for
another?
\VS{4}Jesus
answered and
said unto
them,
{\WJ{Go
and shew
John
again
those things which
ye do hear
and
see:}}
\VS{5}{\WJ{The blind
receive their sight,
and
the lame
walk,
the lepers
are cleansed,
and
the deaf
hear,
the dead
are raised up,
and
the poor
have the gospel preached to them.}}
\VS{6}{\WJ{And
blessed
is
{\ADD{he}},
whosoever
shall
not
be offended
in
me.}}
\par }{\PP \VS{7}¶ And
as
they
departed,
Jesus
began to
say unto the
multitudes
concerning
John,
{\WJ{What
went ye out
into
the wilderness
to see?
A reed
shaken
with
the wind?}}
\VS{8}{\WJ{But
what
went ye out
for to see?
A man
clothed
in
soft
raiment?
behold,
they that
wear
soft
{\ADD{clothing}}
are
in
kings’
houses.}}
\VS{9}{\WJ{But
what
went ye out
for to see?
A prophet?
yea,
I say
unto you,
and
more than
a prophet.}}
\VS{10}{\WJ{For
this
is
{\ADD{he}},
of
whom
it is written,
Behold,
I
send
my
messenger
before
thy
face,
which
shall prepare
thy
way
before
thee.}}
\VS{11}{\WJ{Verily
I say
unto you,
Among
them that are born
of women
there hath
not
risen
a greater than
John
the Baptist:
notwithstanding
he that is least
in
the kingdom
of heaven
is
greater than
he.}}
\VS{12}{\WJ{And
from
the days
of John
the Baptist
until
now
the kingdom
of heaven
suffereth violence,
and
the violent
take
it
by force.}}
\VS{13}{\WJ{For
all
the prophets
and
the law
prophesied
until
John.}}
\VS{14}{\WJ{And
if
ye will
receive
{\ADD{it}},
this
is
Elias,
which
was for
to come.}}
\VS{15}{\WJ{He that hath
ears
to hear,
let him hear.}}
\par }{\PP \VS{16}{\WJ{¶
But
whereunto
shall I liken
this
generation?
It is
like unto
children
sitting
in
the markets,
and
calling unto
their
fellows,}}
\VS{17}{\WJ{And
saying,
We have piped
unto you,
and
ye have
not
danced;
we have mourned
unto you,
and
ye have
not
lamented.}}
\VS{18}{\WJ{For
John
came
neither
eating
nor
drinking,
and
they say,
He hath
a devil.}}
\VS{19}{\WJ{The Son
of man
came
eating
and
drinking,
and
they say,
Behold
a man
gluttonous,
and
a winebibber,
a friend
of publicans
and
sinners.
But
wisdom
is justified
of
her
children.}}
\par }{\PP \VS{20}¶
Then began
he to
upbraid the
cities
wherein
most
of
his mighty
works were
done,
because they
repented
not:
\VS{21}{\WJ{Woe
unto thee,
Chorazin!
woe
unto thee,
Bethsaida!
for
if
the mighty works,
which were done
in
you,
had been done
in
Tyre
and
Sidon,
they
would have
repented
long ago
in
sackcloth
and
ashes.}}
\VS{22}{\WJ{But
I say
unto you,
It shall be
more tolerable
for Tyre
and
Sidon
at
the day
of judgment,
than
for you.}}
\VS{23}{\WJ{And
thou,
Capernaum,
which art exalted
unto
heaven,
shalt be brought down
to
hell:
for
if
the mighty works,
which have been done
in
thee,
had been done
in
Sodom,
it
would have
remained
until
this day.}}
\VS{24}{\WJ{But
I say
unto you,
That
it shall be
more tolerable
for the land
of Sodom
in
the day
of judgment,
than
for thee.}}
\par }{\PP \VS{25}¶
At
that
time
Jesus
answered and
said,
{\WJ{I thank
thee,
O Father,
Lord
of heaven
and
earth,
because
thou hast hid
these things
from
the wise
and
prudent,
and
hast revealed
them
unto babes.}}
\VS{26}{\WJ{Even so,
Father:
for
so
it seemed
good
in
thy sight.}}
\VS{27}{\WJ{All things
are delivered
unto me
of
my
Father:
and
no man
knoweth
the Son,
but
the Father;
neither
knoweth
any man
the Father,
save
the Son,
and
{\ADD{he}}
to whomsoever
the Son
will
reveal
{\ADD{him.}}}}
\par }{\PP \VS{28}{\WJ{¶
Come
unto
me,
all
{\ADD{ye}}
that labour
and
are heavy laden,
and I
will give
you
rest.}}
\VS{29}{\WJ{Take
my
yoke
upon
you,
and
learn
of
me;
for
I am
meek
and
lowly
in heart:
and
ye shall find
rest
unto
your
souls.}}
\VS{30}{\WJ{For
my
yoke
{\ADD{is}}
easy,
and
my
burden
is
light.}}

\par }\Chap{12}{\PP \VerseOne{1}At
that
time
Jesus went
on the sabbath
day
through the
corn;
and
his
disciples were an
hungred,
and
began to
pluck the ears of
corn,
and to
eat.
\VS{2}But
when the
Pharisees
saw
{\ADD{it}}, they
said unto
him,
Behold,
thy
disciples
do that
which
is
not
lawful to
do
upon the sabbath
day.
\VS{3}But
he
said unto
them,
{\WJ{Have ye
not
read
what
David
did,
when
he
was an hungred,
and
they that were
with
him;}}
\VS{4}{\WJ{How
he entered
into
the house
of God,
and
did eat
the
shewbread,
which
was
not
lawful
for him
to eat,
neither
for them which were
with
him,
but
only
for the priests?}}
\VS{5}{\WJ{Or
have ye
not
read
in
the law,
how that
on the sabbath days
the priests
in
the temple
profane
the sabbath,
and
are
blameless?}}
\VS{6}{\WJ{But
I say
unto you,
That
in this place
is
{\ADD{one}}
greater than
the temple.}}
\VS{7}{\WJ{But
if
ye had known
what
{\ADD{this}}
meaneth,
I will have
mercy,
and
not
sacrifice,
ye
would
not
have condemned
the guiltless.}}
\VS{8}{\WJ{For
the Son
of man
is
Lord
even
of the sabbath day.}}
\VS{9}And when he was
departed
thence, he
went
into
their
synagogue:
\par }{\PP \VS{10}¶
And,
behold, there
was a
man which
had
{\ADD{his}}
hand
withered.
And they
asked
him,
saying, Is it
lawful to
heal on the sabbath
days?
that they might
accuse
him.
\VS{11}And
he
said unto
them,
{\WJ{What
man
shall there be
among
you,
that
shall have
one
sheep,
and
if
it
fall
into
a pit
on the sabbath day,
will he
not
lay hold on
it,
and
lift
{\ADD{it}}
out?}}
\VS{12}{\WJ{How much
then
is
a man
better
than a sheep?
Wherefore
it is lawful
to do
well
on the sabbath days.}}
\VS{13}Then saith
he to the
man,
{\WJ{Stretch forth
thine
hand.}}
And he
stretched
{\ADD{it}}
forth;
and it was
restored
whole, like
as the
other.
\par }{\PP \VS{14}¶
Then the
Pharisees went
out, and
held a
council
against
him,
how they might
destroy
him.
\VS{15}But
when
Jesus
knew
{\ADD{it}}, he withdrew
himself from
thence:
and
great
multitudes
followed
him,
and he
healed
them
all;
\VS{16}And
charged
them
that they
should
not
make
him
known:
\VS{17}That it might be
fulfilled which was
spoken
by
Esaias the
prophet,
saying,
\VS{18}Behold
my
servant,
whom I have
chosen;
my
beloved,
in
whom
my
soul is well
pleased: I will
put
my
spirit
upon
him,
and he shall
shew
judgment to the
Gentiles.
\VS{19}He
shall
not
strive,
nor
cry;
neither
shall any
man
hear
his
voice
in the
streets.
\VS{20}A
bruised
reed shall
he
not
break,
and
smoking
flax shall
he
not
quench,
till he send
forth
judgment
unto
victory.
\VS{21}And
in
his
name
shall the
Gentiles
trust.
\par }{\PP \VS{22}¶
Then was
brought unto
him one possessed with a
devil,
blind,
and
dumb:
and he
healed
him, insomuch
that the
blind
and
dumb
both
spake
and
saw.
\VS{23}And
all the
people were
amazed,
and
said,
Is
not
this the
son of
David?
\VS{24}But
when the
Pharisees
heard
{\ADD{it}}, they
said,
This
{\ADD{fellow}}
doth
not cast
out
devils,
but
by
Beelzebub the
prince of the
devils.
\VS{25}And
Jesus
knew
their
thoughts, and
said unto
them,
{\WJ{Every
kingdom
divided
against
itself
is brought to desolation;
and
every
city
or
house
divided
against
itself
shall
not
stand:}}
\VS{26}{\WJ{And
if
Satan
cast out
Satan,
he is divided
against
himself;
how
shall
then
his
kingdom
stand?}}
\VS{27}{\WJ{And
if
I
by
Beelzebub
cast out
devils,
by
whom do
your
children
cast
{\ADD{them}}
out?
therefore
they
shall be
your
judges.}}
\VS{28}{\WJ{But
if
I
cast out
devils
by
the Spirit
of God,
then
the kingdom
of God
is come
unto
you.}}
\VS{29}{\WJ{Or else
how
can
one
enter
into
a strong man’s
house,
and
spoil
his
goods,
except
he
first
bind
the strong man?
and
then
he will spoil
his
house.}}
\VS{30}{\WJ{He that is
not
with
me
is
against
me;
and
he that gathereth
not
with
me
scattereth abroad.}}
\par }{\PP \VS{31}{\WJ{¶
Wherefore
I say
unto you,
All manner of
sin
and
blasphemy
shall be forgiven
unto men:
but the
blasphemy
{\ADD{against}}
the
{\ADD{Holy}}
Ghost
shall
not
be forgiven
unto men.}}
\VS{32}{\WJ{And
whosoever
speaketh
a word
against
the Son
of man,
it shall be forgiven
him:
but
whosoever
speaketh
against
the
Holy
Ghost,
it shall
not
be forgiven
him,
neither
in
this
world,
neither
in
the
{\ADD{world}}
to come.}}
\VS{33}{\WJ{Either
make
the tree
good,
and
his
fruit
good;
or else
make
the tree
corrupt,
and
his
fruit
corrupt:
for
the tree
is known
by
{\ADD{his}}
fruit.}}
\VS{34}{\WJ{O generation
of vipers,
how
can ye,
being
evil,
speak
good things?
for
out
of the abundance
of the heart
the mouth
speaketh.}}
\VS{35}{\WJ{A
good
man
out
of the good
treasure
of the heart
bringeth forth
good things:
and
an
evil
man
out
of the
evil
treasure
bringeth forth
evil things.}}
\VS{36}{\WJ{But
I say
unto you,
That
every
idle
word
that
men
shall speak,
they shall give
account
thereof
in
the day
of judgment.}}
\VS{37}{\WJ{For
by
thy
words
thou shalt be justified,
and
by
thy
words
thou shalt be condemned.}}
\par }{\PP \VS{38}¶
Then
certain of the
scribes
and of the
Pharisees
answered,
saying,
Master, we
would
see a
sign
from
thee.
\VS{39}But
he
answered and
said unto
them,
{\WJ{An
evil
and
adulterous
generation
seeketh after
a sign;
and
there shall
no
sign
be given
to it,
but
the sign
of the prophet
Jonas:}}
\VS{40}{\WJ{For
as
Jonas
was
three
days
and
three
nights
in
the whale’s
belly;
so
shall
the Son
of man
be
three
days
and
three
nights
in
the heart
of the earth.}}
\VS{41}{\WJ{The men
of Nineveh
shall rise
in
judgment
with
this
generation,
and
shall condemn
it:
because
they repented
at
the preaching
of Jonas;
and,
behold,
a greater
than Jonas
{\ADD{is}}
here.}}
\VS{42}{\WJ{The queen
of the south
shall rise up
in
the judgment
with
this
generation,
and
shall condemn
it:
for
she came
from
the uttermost parts
of the earth
to hear
the wisdom
of Solomon;
and,
behold,
a greater
than Solomon
{\ADD{is}}
here.}}
\VS{43}{\WJ{When
the
unclean
spirit
is gone
out of
a man,
he walketh
through
dry
places,
seeking
rest,
and
findeth
none.}}
\VS{44}{\WJ{Then
he saith,
I will return
into
my
house
from whence
I came out;
and when
he is come,
he findeth
{\ADD{it}}
empty,
swept,
and
garnished.}}
\VS{45}{\WJ{Then
goeth he,
and
taketh
with
himself
seven
other
spirits
more wicked
than himself,
and
they enter in
and dwell
there:
and
the last
{\ADD{state}}
of
that
man
is
worse
than the first.
Even so
shall it be
also
unto
this
wicked
generation.}}
\par }{\PP \VS{46}¶
While
he
yet
talked to the
people,
behold,
{\ADD{his}}
mother
and
his
brethren
stood
without,
desiring to
speak with
him.
\VS{47}Then
one
said unto
him,
Behold,
thy
mother
and
thy
brethren
stand
without,
desiring to
speak with
thee.
\VS{48}But
he
answered and
said unto him that
told
him,
{\WJ{Who
is
my
mother?
and
who
are
my
brethren?}}
\VS{49}And he stretched
forth
his
hand
toward
his
disciples, and
said,
{\WJ{Behold
my
mother
and
my
brethren!}}
\VS{50}{\WJ{For
whosoever
shall do
the will
of
my
Father
which is
in
heaven,
the same
is
my
brother,
and
sister,
and
mother.}}

\par }\Chap{13}{\PP \VerseOne{1}The
same
day
went
Jesus out
of the
house, and
sat
by the sea
side.
\VS{2}And
great
multitudes were gathered
together
unto
him, so
that
he
went
into a
ship, and
sat;
and
the
whole
multitude
stood
on the
shore.
\VS{3}And he
spake many
things unto
them
in
parables,
saying,
{\WJ{Behold,
a sower
went forth
to sow;}}
\VS{4}{\WJ{And
when
he
sowed,
some
{\ADD{seeds}}
fell
by
the way side,
and
the fowls
came
and
devoured
them
up:}}
\VS{5}{\WJ{Some
fell
upon
stony places,
where
they had
not
much
earth:
and
forthwith
they sprung up,
because
they had
no
deepness
of earth:}}
\VS{6}{\WJ{And
when the sun
was up,
they were scorched;
and
because
they had
no
root,
they withered away.}}
\VS{7}{\WJ{And
some
fell
among
thorns;
and
the thorns
sprung up,
and
choked
them:}}
\VS{8}{\WJ{But
other
fell
into
good
ground,
and
brought forth
fruit,
some
an hundredfold,
some
sixtyfold,
some
thirtyfold.}}
\VS{9}{\WJ{Who hath
ears
to hear,
let him hear.}}
\VS{10}And the
disciples
came, and
said unto
him,
Why speakest
thou unto
them
in
parables?
\VS{11}He
answered and
said unto
them,
{\WJ{Because
it is given
unto you
to know
the mysteries
of the kingdom
of heaven,
but
to them
it is
not
given.}}
\VS{12}{\WJ{For
whosoever
hath,
to him
shall be given,
and
he shall have more abundance:
but
whosoever
hath
not,
from
him
shall be taken away
even
that
he hath.}}
\VS{13}{\WJ{Therefore
speak I
to them
in
parables:
because
they
seeing
see
not;
and
hearing
they hear
not,
neither
do they understand.}}
\VS{14}{\WJ{And
in
them
is fulfilled
the prophecy
of Esaias,
which saith,
By hearing
ye shall hear,
and
shall
not
understand;
and
seeing
ye shall see,
and
shall
not
perceive:}}
\VS{15}{\WJ{For
this
people’s
heart
is waxed gross,
and
{\ADD{their}}
ears
are dull of
hearing,
and
their
eyes
they have closed;
lest at any time
they should see
with
{\ADD{their}}
eyes,
and
hear
with
{\ADD{their}}
ears,
and
should understand
with
{\ADD{their}}
heart,
and
should be converted,
and
I should heal
them.}}
\VS{16}{\WJ{But
blessed
{\ADD{are}}
your
eyes,
for
they see:
and
your
ears,
for
they hear.}}
\VS{17}{\WJ{For
verily
I say
unto you,
That
many
prophets
and
righteous
{\ADD{men}}
have desired
to see
{\ADD{those things}}
which
ye see,
and
have
not
seen
{\ADD{them}};
and
to hear
{\ADD{those things}}
which
ye hear,
and
have
not
heard
{\ADD{them.}}}}
\par }{\PP \VS{18}{\WJ{¶
Hear
ye
therefore
the parable
of the sower.}}
\VS{19}{\WJ{When any one
heareth
the word
of the kingdom,
and
understandeth
{\ADD{it}}
not,
then cometh
the wicked
{\ADD{one}},
and
catcheth away
that which was sown
in
his
heart.
This
is
he which received seed
by
the way side.}}
\VS{20}{\WJ{But
he that received the seed
into
stony places,
the same
is
he that heareth
the word,
and
anon
with
joy
receiveth
it;}}
\VS{21}{\WJ{Yet
hath he
not
root
in
himself,
but
dureth
for a while:
for when
tribulation
or
persecution
ariseth
because of
the word,
by and by
he is offended.}}
\VS{22}{\WJ{He
also
that received seed
among
the thorns
is
he
that heareth
the word;
and
the care
of this
world,
and
the deceitfulness
of riches,
choke
the word,
and
he becometh
unfruitful.}}
\VS{23}{\WJ{But
he that received seed
into
the
good
ground
is
he
that heareth
the word,
and
understandeth
{\ADD{it}};
which
also
beareth fruit,
and
bringeth forth,
some
an hundredfold,
some
sixty,
some
thirty.}}
\par }{\PP \VS{24}¶
Another
parable put he
forth unto
them,
saying,
{\WJ{The kingdom
of heaven
is likened unto
a man
which sowed
good
seed
in
his
field:}}
\VS{25}{\WJ{But
while
men
slept,
his
enemy
came
and
sowed
tares
among
the wheat,
and
went his way.}}
\VS{26}{\WJ{But
when
the blade
was sprung up,
and
brought forth
fruit,
then
appeared
the tares
also.}}
\VS{27}{\WJ{So
the servants
of the householder
came
and said
unto him,
Sir,
didst
not
thou sow
good
seed
in
thy
field?
from whence
then
hath it
tares?}}
\VS{28}{\WJ{He said
unto them,
An enemy
hath done
this.
The servants
said
unto him,
Wilt thou
then
that we go
and gather
them
up?}}
\VS{29}{\WJ{But
he said,
Nay;
lest
while ye gather up
the tares,
ye root up also
the wheat
with
them.}}
\VS{30}{\WJ{Let
both
grow together
until
the harvest:
and
in
the time
of harvest
I will say
to the reapers,
Gather ye together
first
the tares,
and
bind
them
in
bundles
to
burn
them:
but
gather
the wheat
into
my
barn.}}
\par }{\PP \VS{31}¶
Another
parable put he
forth unto
them,
saying,
{\WJ{The kingdom
of heaven
is
like to
a grain
of mustard seed,
which
a man
took,
and sowed
in
his
field:}}
\VS{32}{\WJ{Which
indeed
is
the least
of all
seeds:
but
when
it is grown,
it is
the greatest
among herbs,
and
becometh
a tree,
so that
the birds
of the air
come
and
lodge
in
the branches
thereof.}}
\par }{\PP \VS{33}¶
Another
parable spake
he unto
them;
{\WJ{The kingdom
of heaven
is
like unto
leaven,
which
a woman
took,
and hid
in
three
measures
of meal,
till
the whole
was leavened.}}
\VS{34}All these
things
spake
Jesus unto the
multitude
in
parables;
and
without a
parable spake
he
not unto
them:
\VS{35}That it might be
fulfilled which was
spoken
by the
prophet,
saying, I will
open
my
mouth
in
parables; I will
utter things which have been kept
secret
from the
foundation of the
world.
\VS{36}Then
Jesus
sent the
multitude
away, and
went
into the
house:
and
his
disciples
came unto
him,
saying,
Declare unto
us the
parable of the
tares of the
field.
\VS{37}He
answered and
said unto
them,
{\WJ{He that soweth
the good
seed
is
the Son
of man;}}
\VS{38}{\WJ{The field
is
the world;
the good
seed
are
the children
of the kingdom;
but
the tares
are
the children
of the wicked
{\ADD{one}};}}
\VS{39}{\WJ{The enemy
that sowed
them
is
the devil;
the harvest
is
the end
of the world;
and
the reapers
are
the angels.}}
\VS{40}{\WJ{As
therefore
the tares
are gathered
and
burned
in the fire;
so
shall it be
in
the end
of this
world.}}
\VS{41}{\WJ{The Son
of man
shall send forth
his
angels,
and
they shall gather
out of
his
kingdom
all
things that offend,
and
them which do
iniquity;}}
\VS{42}{\WJ{And
shall cast
them
into
a furnace
of fire:
there
shall be
wailing
and
gnashing
of teeth.}}
\VS{43}{\WJ{Then
shall
the righteous
shine forth
as
the sun
in
the kingdom
of
their
Father.
Who hath
ears
to hear,
let him hear.}}
\par }{\PP \VS{44}{\WJ{¶
Again,
the kingdom
of heaven
is
like unto
treasure
hid
in
a field;
the which
when a man
hath found,
he hideth,
and
for
joy
thereof
goeth
and
selleth
all
that
he hath,
and
buyeth
that
field.}}
\par }{\PP \VS{45}{\WJ{¶
Again,
the kingdom
of heaven
is
like unto
a merchant
man,
seeking
goodly
pearls:}}
\VS{46}{\WJ{Who,
when he had found
one
pearl
of great price,
went
and sold
all
that
he had,
and
bought
it.}}
\par }{\PP \VS{47}{\WJ{¶
Again,
the kingdom
of heaven
is
like unto
a net,
that was cast
into
the sea,
and
gathered
of
every
kind:}}
\VS{48}{\WJ{Which,
when
it was full,
they drew
to
shore,
and
sat down,
and gathered
the good
into
vessels,
but
cast
the bad
away.}}
\VS{49}{\WJ{So
shall it be
at
the end
of the world:
the angels
shall come forth,
and
sever
the wicked
from
among
the just,}}
\VS{50}{\WJ{And
shall cast
them
into
the furnace
of fire:
there
shall be
wailing
and
gnashing
of teeth.}}
\VS{51}Jesus
saith unto
them,
{\WJ{Have ye understood
all
these things?}} They
say unto
him,
Yea,
Lord.
\VS{52}Then said
he unto
them,
{\WJ{Therefore
every
scribe
{\ADD{which is}}
instructed
unto
the kingdom
of heaven
is
like unto
a man
{\ADD{that is}}
an householder,
which
bringeth forth
out of
his
treasure
{\ADD{things}}
new
and
old.}}
\par }{\PP \VS{53}¶
And it came to
pass,
{\ADD{that}}
when
Jesus had
finished
these
parables, he
departed
thence.
\VS{54}And when he was
come
into his
own
country, he
taught
them
in
their
synagogue, insomuch
that
they were
astonished,
and
said,
Whence hath
this
{\ADD{man}}
this
wisdom,
and
{\ADD{these}} mighty
works?
\VS{55}Is
not
this
the
carpenter’s
son?
is
not
his
mother
called
Mary?
and
his
brethren,
James,
and
Joses,
and
Simon,
and
Judas?
\VS{56}And
his
sisters, are
they
not
all
with
us?
Whence
then hath
this
{\ADD{man}}
all these
things?
\VS{57}And they were
offended
in
him.
But
Jesus
said unto
them,
{\WJ{A prophet
is
not
without honour,
save
in
his own
country,
and
in
his own
house.}}
\VS{58}And he
did
not
many mighty
works
there because
of
their
unbelief.

\par }\Chap{14}{\PP \VerseOne{1}At
that
time
Herod the
tetrarch
heard of the
fame of
Jesus,
\VS{2}And
said
unto
his
servants,
This
is
John the
Baptist;
he is
risen
from the
dead;
and
therefore mighty
works do shew forth
themselves
in
him.
\par }{\PP \VS{3}¶
For
Herod had laid hold
on
John, and
bound
him,
and
put
{\ADD{him}}
in
prison
for
Herodias’ sake,
his
brother
Philip’s
wife.
\VS{4}For
John
said unto
him, It
is
not
lawful for
thee to
have
her.
\VS{5}And
when
he would
have put him to
death, he
feared the
multitude,
because they
counted
him
as a
prophet.
\VS{6}But
when
Herod’s
birthday was
kept, the
daughter of
Herodias
danced
before
them,
and
pleased
Herod.
\VS{7}Whereupon he
promised
with an
oath to
give
her
whatsoever
she would
ask.
\VS{8}And she, being before
instructed
of
her
mother,
said,
Give
me
here
John
Baptist’s
head
in a
charger.
\VS{9}And the
king was
sorry:
nevertheless
for the
oath’s sake,
and them which sat with him at
meat, he
commanded
{\ADD{it}} to be
given
{\ADD{her}}.
\VS{10}And he
sent, and
beheaded
John
in the
prison.
\VS{11}And
his
head was
brought
in a
charger,
and
given to the
damsel:
and she
brought
{\ADD{it}}
to
her
mother.
\VS{12}And
his
disciples
came, and took
up the
body,
and
buried
it,
and
went and
told
Jesus.
\par }{\PP \VS{13}¶
When
Jesus
heard
{\ADD{of it}}, he
departed
thence
by
ship
into a
desert
place
apart: and
when the
people had
heard
{\ADD{thereof}}, they
followed
him on
foot out
of the
cities.
\VS{14}And
Jesus went
forth, and
saw a
great
multitude,
and was moved with
compassion
toward
them,
and he
healed
their
sick.
\par }{\PP \VS{15}¶
And when it
was
evening,
his
disciples
came to
him,
saying, This
is a
desert
place,
and the
time
is
now
past;
send the
multitude
away,
that they
may
go
into the
villages, and
buy
themselves
victuals.
\VS{16}But
Jesus
said unto
them,
{\WJ{They
need
not
depart;
give
ye
them
to eat.}}
\VS{17}And they
say unto
him, We
have
here
but
five
loaves,
and
two
fishes.
\VS{18}He
said,
{\WJ{Bring
them
hither
to me.}}
\VS{19}And he
commanded the
multitude to sit
down
on the
grass,
and
took
the
five
loaves,
and
the
two
fishes, and looking
up
to
heaven, he
blessed,
and
brake, and
gave the
loaves
to
{\ADD{his}}
disciples,
and the
disciples to the
multitude.
\VS{20}And they
did
all
eat,
and were
filled:
and they took
up of the
fragments that
remained
twelve
baskets
full.
\VS{21}And they that had
eaten
were
about five
thousand
men,
beside
women
and
children.
\par }{\PP \VS{22}¶
And
straightway
Jesus
constrained
his
disciples to
get
into a
ship,
and to go
before
him
unto the other
side,
while he
sent the
multitudes
away.
\VS{23}And when he had
sent the
multitudes
away, he went
up
into a
mountain
apart to
pray: and
when the
evening was
come, he
was
there
alone.
\VS{24}But the
ship
was
now in the
midst of the
sea,
tossed
with
waves:
for the
wind
was
contrary.
\VS{25}And in the
fourth
watch of the
night
Jesus
went
unto
them,
walking
on the
sea.
\VS{26}And
when the
disciples
saw
him
walking
on the
sea, they were
troubled,
saying, It
is a
spirit;
and they cried
out
for
fear.
\VS{27}But
straightway
Jesus
spake unto
them,
saying,
{\WJ{Be of good cheer;
it is
I;
be
not
afraid.}}
\VS{28}And
Peter
answered
him and
said,
Lord,
if it
be
thou,
bid
me
come
unto
thee
on the
water.
\VS{29}And he
said,
{\WJ{Come.}} And
when
Peter was come
down
out of the
ship, he
walked
on the
water, to
go
to
Jesus.
\VS{30}But when he
saw the
wind
boisterous, he was
afraid;
and
beginning to
sink, he
cried,
saying,
Lord,
save
me.
\VS{31}And
immediately
Jesus stretched
forth
{\ADD{his}}
hand, and
caught
him,
and
said unto
him,
{\WJ{O thou of little faith,
wherefore
didst thou doubt?}}
\VS{32}And when
they were
come
into the
ship, the
wind
ceased.
\VS{33}Then they that
were
in the
ship
came and
worshipped
him,
saying, Of a
truth thou
art the
Son of
God.
\par }{\PP \VS{34}¶
And when they were gone
over, they
came
into the
land of
Gennesaret.
\VS{35}And
when the
men of
that
place had knowledge
of
him, they sent
out
into
all
that country round
about,
and
brought unto
him
all that
were
diseased;
\VS{36}And
besought
him
that they
might
only
touch the
hem of
his
garment:
and as many
as
touched were made perfectly
whole.

\par }\Chap{15}{\PP \VerseOne{1}Then
came to
Jesus
scribes
and
Pharisees, which were
of
Jerusalem,
saying,
\VS{2}Why
do
thy
disciples
transgress the
tradition of the
elders?
for they
wash
not
their
hands
when they
eat
bread.
\VS{3}But he
answered and
said unto
them,
{\WJ{Why
do
ye
also
transgress
the commandment
of God
by
your
tradition?}}
\VS{4}{\WJ{For
God
commanded,
saying,
Honour
thy
father
and
mother:
and,
He that curseth
father
or
mother,
let him die
the death.}}
\VS{5}{\WJ{But
ye
say,
Whosoever
shall say
to
{\ADD{his}}
father
or
{\ADD{his}}
mother,
{\ADD{It is}}
a gift,
by whatsoever
thou mightest be profited
by
me;}}
\VS{6}{\WJ{And
honour
not
his
father
or
his
mother,
{\ADD{he shall be free}}.
Thus
have ye made
the commandment
of God
of none effect
by
your
tradition.}}
\VS{7}{\WJ{{\ADD{Ye}}
hypocrites,
well
did
Esaias
prophesy
of
you,
saying,}}
\VS{8}{\WJ{This
people
draweth nigh
unto me
with
their
mouth,
and
honoureth
me
with
{\ADD{their}}
lips;
but
their
heart
is
far
from
me.}}
\VS{9}{\WJ{But
in vain
they do worship
me,
teaching
{\ADD{for}}
doctrines
the commandments
of men.}}
\par }{\PP \VS{10}¶
And he
called the
multitude, and
said unto
them,
{\WJ{Hear,
and
understand:}}
\VS{11}{\WJ{Not
that which goeth
into
the mouth
defileth
a man;
but
that which cometh out
of
the mouth,
this
defileth
a man.}}
\VS{12}Then
came
his
disciples, and
said unto
him, Knowest
thou
that the
Pharisees were
offended, after they
heard this
saying?
\VS{13}But he
answered and
said,
{\WJ{Every
plant,
which
my
heavenly
Father
hath
not
planted,
shall be rooted up.}}
\VS{14}{\WJ{Let
them
alone:
they be
blind
leaders
of the blind.
And
if
the blind
lead
the blind,
both
shall fall
into
the ditch.}}
\VS{15}Then
answered
Peter and
said unto
him,
Declare unto
us
this
parable.
\VS{16}And
Jesus
said,
{\WJ{Are
ye
also
yet
without understanding?}}
\VS{17}{\WJ{Do not
ye
yet
understand,
that
whatsoever
entereth
in at
the mouth
goeth
into
the belly,
and
is cast out
into
the draught?}}
\VS{18}{\WJ{But
those things which proceed
out
of the mouth
come forth
from
the heart;
and they
defile
the man.}}
\VS{19}{\WJ{For
out
of the heart
proceed
evil
thoughts,
murders,
adulteries,
fornications,
thefts,
false witness,
blasphemies:}}
\VS{20}{\WJ{These
are
{\ADD{the things}}
which defile
a man:
but
to eat
with
unwashen
hands
defileth
not
a man.}}
\par }{\PP \VS{21}¶
Then
Jesus
went
thence, and
departed
into the
coasts of
Tyre
and
Sidon.
\VS{22}And,
behold, a
woman of
Canaan
came
out of
the
same
coasts, and
cried unto
him,
saying, Have
mercy on
me, O
Lord,
{\ADD{thou}}
Son of
David;
my
daughter
is
grievously vexed with a
devil.
\VS{23}But he
answered
her
not a
word.
And
his
disciples
came and
besought
him,
saying,
Send
her
away;
for she
crieth
after
us.
\VS{24}But he
answered and
said,
{\WJ{I am
not
sent
but
unto
the lost
sheep
of the house
of Israel.}}
\VS{25}Then came
she and
worshipped
him,
saying,
Lord,
help
me.
\VS{26}But he
answered and
said,
{\WJ{It is
not
meet
to take
the children’s
bread,
and
to cast
{\ADD{it}}
to dogs.}}
\VS{27}And she
said,
Truth,
Lord:
yet the
dogs
eat
of the
crumbs which
fall
from
their
masters’
table.
\VS{28}Then
Jesus
answered and
said unto
her,
{\WJ{O
woman,
great
{\ADD{is}}
thy
faith:
be it
unto thee
even as
thou wilt.}}
And
her
daughter was made
whole
from that
very
hour.
\VS{29}And
Jesus
departed from
thence, and
came
nigh unto the
sea of
Galilee;
and went
up
into a
mountain, and sat
down
there.
\VS{30}And
great
multitudes came
unto
him,
having
with
them
{\ADD{those that were}}
lame,
blind,
dumb,
maimed,
and
many
others,
and
cast
them
down
at
Jesus’
feet;
and he
healed
them:
\VS{31}Insomuch
that the
multitude
wondered, when they
saw the
dumb to
speak, the
maimed to be
whole, the
lame to
walk,
and the
blind to
see:
and they
glorified the
God of
Israel.
\par }{\PP \VS{32}¶
Then
Jesus
called
his
disciples
{\ADD{unto him}}, and
said,
{\WJ{I have compassion
on
the multitude,
because
they continue with
me
now
three
days,
and
have
nothing
to
eat:
and
I will
not
send
them
away
fasting,
lest
they faint
in
the way.}}
\VS{33}And
his
disciples
say unto
him,
Whence should we
have so
much
bread
in the
wilderness,
as to
fill so
great a
multitude?
\VS{34}And
Jesus
saith unto
them,
{\WJ{How many
loaves
have ye?}}
And they
said,
Seven,
and a
few little
fishes.
\VS{35}And he
commanded the
multitude to sit
down
on the
ground.
\VS{36}And he
took
the
seven
loaves
and the
fishes, and gave
thanks, and
brake
{\ADD{them}},
and
gave
to
his
disciples,
and the
disciples to the
multitude.
\VS{37}And they
did
all
eat,
and were
filled:
and they took
up of the
broken
{\ADD{meat}} that was
left
seven
baskets
full.
\VS{38}And they that did
eat
were four
thousand
men,
beside
women
and
children.
\VS{39}And he sent
away the
multitude, and
took
ship,
and
came
into the
coasts of
Magdala.

\par }\Chap{16}{\PP \VerseOne{1}The
Pharisees
also
with the
Sadducees
came, and
tempting
desired
him that he would
shew
them a
sign
from
heaven.
\VS{2}He
answered
and
said unto
them,
{\WJ{When it is
evening,
ye say,
{\ADD{It will be}}
fair weather:
for
the sky
is red.}}
\VS{3}{\WJ{And
in the morning,
{\ADD{It will be}}
foul weather
to day:
for
the sky
is red
and lowring. O
{\ADD{ye}}
hypocrites,
ye can
discern
the face
of the sky;
but
can ye
not
{\ADD{discern}}
the signs
of the times?}}
\VS{4}{\WJ{A wicked
and
adulterous
generation
seeketh after
a sign;
and
there shall
no
sign
be given
unto it,
but
the sign
of the prophet
Jonas.}}
And he
left
them, and
departed.
\VS{5}And
when
his
disciples were
come
to the other
side, they had
forgotten to
take
bread.
\par }{\PP \VS{6}¶
Then
Jesus
said unto
them,
{\WJ{Take heed
and
beware
of
the leaven
of the Pharisees
and
of the Sadducees.}}
\VS{7}And they
reasoned
among
themselves,
saying,
{\ADD{It is}}
because we have
taken
no
bread.
\VS{8}{\ADD{Which}}
when
Jesus
perceived, he
said unto
them,
{\WJ{O ye of little faith,
why
reason ye
among
yourselves,
because
ye have brought
no
bread?}}
\VS{9}{\WJ{Do ye
not yet
understand,
neither
remember
the
five
loaves
of the five thousand,
and
how many
baskets
ye took up?}}
\VS{10}{\WJ{Neither
the
seven
loaves
of the four thousand,
and
how many
baskets
ye took up?}}
\VS{11}{\WJ{How is it that
ye do
not
understand
that
I spake
{\ADD{it}}
not
to you
concerning
bread,
that ye should beware
of
the leaven
of the Pharisees
and
of the Sadducees?}}
\VS{12}Then understood
they how
that he
bade
{\ADD{them}}
not
beware
of the
leaven of
bread,
but
of the
doctrine of the
Pharisees
and of the
Sadducees.
\par }{\PP \VS{13}¶
When
Jesus
came
into the
coasts of
Caesarea
Philippi, he
asked
his
disciples,
saying,
{\WJ{Whom
do
men
say that
I
the Son
of man
am?}}
\VS{14}And they
said,
Some
{\ADD{say that thou art}}
John the
Baptist:
some,
Elias;
and
others,
Jeremias,
or
one of the
prophets.
\VS{15}He
saith unto
them,
{\WJ{But
whom
say
ye
that
I
am?}}
\VS{16}And
Simon
Peter
answered and
said,
Thou
art the
Christ, the
Son of
the
living
God.
\VS{17}And
Jesus
answered and
said unto
him,
{\WJ{Blessed
art thou,
Simon
Bar-jona:
for
flesh
and
blood
hath
not
revealed
{\ADD{it}}
unto thee,
but
my
Father
which is
in
heaven.}}
\VS{18}{\WJ{And
I
say
also
unto thee,
That
thou
art
Peter,
and
upon
this
rock
I will build
my
church;
and
the gates
of hell
shall
not
prevail against
it.}}
\VS{19}{\WJ{And
I will give
unto thee
the keys
of the kingdom
of heaven:
and
whatsoever
thou shalt bind
on
earth
shall be
bound
in
heaven:
and
whatsoever
thou shalt loose
on
earth
shall be
loosed
in
heaven.}}
\VS{20}Then charged
he
his
disciples
that they should
tell no
man
that
he
was
Jesus the
Christ.
\par }{\PP \VS{21}¶
From that
time
forth
began
Jesus to
shew unto
his
disciples, how
that
he
must
go
unto
Jerusalem,
and
suffer many
things
of the
elders
and chief
priests
and
scribes,
and be
killed,
and be raised
again
the
third
day.
\VS{22}Then
Peter
took
him, and began
to
rebuke
him,
saying, Be it
far from
thee,
Lord:
this
shall
not
be unto
thee.
\VS{23}But he
turned, and
said unto
Peter,
{\WJ{Get thee
behind
me,
Satan:
thou art
an offence
unto me:
for
thou savourest
not
the things that be
of God,
but
those that be
of men.}}
\par }{\PP \VS{24}¶
Then
said
Jesus
unto
his
disciples,
{\WJ{If
any
{\ADD{man}}
will
come
after
me,
let him deny
himself,
and
take up
his
cross,
and
follow
me.}}
\VS{25}{\WJ{For
whosoever
will
save
his
life
shall lose
it:
and
whosoever
will lose
his
life
for
my
sake
shall find
it.}}
\VS{26}{\WJ{For
what
is
a man
profited,
if
he shall gain
the
whole
world,
and
lose
his own
soul?
or
what
shall
a man
give
in exchange
for
his
soul?}}
\VS{27}{\WJ{For
the Son
of man
shall
come
in
the glory
of
his
Father
with
his
angels;
and
then
he shall reward
every man
according to
his
works.}}
\VS{28}{\WJ{Verily
I say
unto you,
There be
some
standing
here,
which
shall
not
taste
of death,
till
they see
the Son
of man
coming
in
his
kingdom.}}

\par }\Chap{17}{\PP \VerseOne{1}And
after
six
days
Jesus
taketh
Peter,
James,
and
John
his
brother,
and
bringeth
them
up
into
an
high
mountain
apart,
\VS{2}And was
transfigured
before
them:
and
his
face did
shine
as the
sun,
and
his
raiment
was
white
as the
light.
\VS{3}And,
behold, there
appeared unto
them
Moses
and
Elias
talking
with
him.
\VS{4}Then
answered
Peter, and
said unto
Jesus,
Lord, it
is
good for
us to
be
here:
if thou
wilt, let us
make
here
three
tabernacles;
one for
thee,
and
one for
Moses,
and
one for
Elias.
\VS{5}While
he
yet
spake,
behold,
a
bright
cloud
overshadowed
them:
and
behold a
voice out
of the
cloud, which
said,
This
is
my
beloved
Son,
in
whom I am well
pleased; hear
ye
him.
\VS{6}And
when the
disciples
heard
{\ADD{it}}, they
fell
on
their
face,
and
were
sore
afraid.
\VS{7}And
Jesus
came and
touched
them,
and
said,
{\WJ{Arise,
and
be
not
afraid.}}
\VS{8}And when they had lifted
up
their
eyes, they
saw no
man,
save
Jesus
only.
\VS{9}And
as
they came
down
from the
mountain,
Jesus
charged
them,
saying,
{\WJ{Tell
the vision
to no man,
until
the Son
of man
be risen again
from
the dead.}}
\VS{10}And
his
disciples
asked
him,
saying,
Why
then
say the
scribes
that
Elias
must
first
come?
\VS{11}And
Jesus
answered and
said unto
them,
{\WJ{Elias
truly
shall
first
come,
and
restore
all things.}}
\VS{12}{\WJ{But
I say
unto you,
That
Elias
is come
already,
and
they knew
him
not,
but
have done
unto
him
whatsoever
they listed.
Likewise
shall
also
the Son
of man
suffer
of
them.}}
\VS{13}Then the
disciples
understood
that he
spake unto
them
of
John the
Baptist.
\par }{\PP \VS{14}¶
And
when
they were
come
to the
multitude, there
came to
him
a
{\ADD{certain}}
man, kneeling
down to
him,
and
saying,
\VS{15}Lord, have
mercy
on
my
son:
for he is
lunatick,
and
sore
vexed:
for
ofttimes he
falleth
into the
fire,
and
oft
into the
water.
\VS{16}And I
brought
him
to
thy
disciples,
and they
could
not
cure
him.
\VS{17}Then
Jesus
answered and
said,
{\WJ{O
faithless
and
perverse
generation,
how
long
shall I be
with
you?
how
long
shall I suffer
you?
bring
him
hither
to me.}}
\VS{18}And
Jesus
rebuked the
devil;
and
he
departed
out of
him:
and the
child was
cured
from that
very
hour.
\VS{19}Then
came the
disciples to
Jesus
apart, and
said,
Why
could
not
we
cast
him
out?
\VS{20}And
Jesus
said unto
them,
{\WJ{Because
of
your
unbelief:
for
verily
I say
unto you,
If
ye have
faith
as
a grain
of mustard seed,
ye shall say
unto
this
mountain,
Remove
hence
to yonder place;
and
it shall remove;
and
nothing
shall be impossible
unto you.}}
\VS{21}{\WJ{Howbeit
this
kind
goeth
not
out
but
by
prayer
and
fasting.}}
\par }{\PP \VS{22}¶
And
while
they
abode
in
Galilee,
Jesus
said unto
them,
{\WJ{The Son
of man
shall
be betrayed
into
the hands
of men:}}
\VS{23}{\WJ{And
they shall kill
him,
and
the
third
day
he shall be raised again.}}
And they
were
exceeding
sorry.
\par }{\PP \VS{24}¶
And
when
they were
come
to
Capernaum, they that
received
tribute
{\ADD{money}}
came to
Peter,
and
said,
Doth
not
your
master
pay
tribute?
\VS{25}He
saith,
Yes.
And
when he was
come
into the
house,
Jesus
prevented
him,
saying,
{\WJ{What
thinkest
thou,
Simon?
of
whom
do
the kings
of the earth
take
custom
or
tribute?
of
their own
children,
or
of
strangers?}}
\VS{26}Peter
saith unto
him,
Of
strangers.
Jesus
saith unto
him,
{\WJ{Then
are
the children
free.}}
\VS{27}{\WJ{Notwithstanding,
lest
we should offend
them,
go thou
to
the sea,
and cast
an hook,
and
take up
the fish
that
first
cometh up;
and
when thou hast opened
his
mouth,
thou shalt find
a piece of money:
that
take,
and give
unto them
for
me
and
thee.}}

\par }\Chap{18}{\PP \VerseOne{1}At the
same
time
came the
disciples unto
Jesus,
saying,
Who
is the
greatest
in the
kingdom of
heaven?
\VS{2}And
Jesus
called a little
child unto
him, and
set
him
in the
midst of
them,
\VS{3}And
said,
{\WJ{Verily
I say
unto you,
Except
ye be converted,
and
become
as
little children,
ye shall
not
enter
into
the kingdom
of heaven.}}
\VS{4}{\WJ{Whosoever
therefore
shall humble
himself
as
this
little child,
the same
is
greatest
in
the kingdom
of heaven.}}
\VS{5}{\WJ{And
whoso
shall receive
one
such
little child
in
my
name
receiveth
me.}}
\VS{6}{\WJ{But
whoso
shall offend
one
of
these
little ones
which believe
in
me,
it were better
for him
that
a millstone
were hanged
about
his
neck,
and
{\ADD{that}}
he were drowned
in
the depth
of the sea.}}
\par }{\PP \VS{7}{\WJ{¶
Woe
unto the world
because
of offences!
for
it
must needs
be
that
offences
come;
but
woe
to
that
man
by
whom
the offence
cometh!}}
\VS{8}{\WJ{Wherefore
if
thy
hand
or
thy
foot
offend
thee,
cut
them
off,
and
cast
{\ADD{them}}
from
thee:
it is
better
for thee
to enter
into
life
halt
or
maimed,
rather than
having
two
hands
or
two
feet
to be cast
into
everlasting
fire.}}
\VS{9}{\WJ{And
if
thine
eye
offend
thee,
pluck
it
out,
and
cast
{\ADD{it}}
from
thee:
it is
better
for thee
to enter
into
life
with one eye,
rather than
having
two
eyes
to be cast
into
hell
fire.}}
\VS{10}{\WJ{Take heed that
ye despise
not
one
of
these
little ones;
for
I say
unto you,
That
in
heaven
their
angels
do
always
behold
the face
of
my
Father
which is
in
heaven.}}
\VS{11}{\WJ{For
the Son
of man
is come
to save
that which was lost.}}
\VS{12}{\WJ{How
think
ye?
if
a
man
have
an hundred
sheep,
and
one
of
them
be gone astray,
doth he
not
leave
the ninety and nine,
and goeth
into
the mountains,
and seeketh
that which is gone astray?}}
\VS{13}{\WJ{And
if
so be that
he find
it,
verily
I say
unto you,
he rejoiceth
more
of
that
{\ADD{sheep}},
than
of
the ninety and nine
which went
not
astray.}}
\VS{14}{\WJ{Even so
it is
not
the will
of
your
Father
which is
in
heaven,
that
one
of
these
little ones
should perish.}}
\par }{\PP \VS{15}{\WJ{¶
Moreover
if
thy
brother
shall trespass
against
thee,
go
and
tell
him
his fault
between
thee
and
him
alone:
if
he shall hear
thee,
thou hast gained
thy
brother.}}
\VS{16}{\WJ{But
if
he will
not
hear
{\ADD{thee, then}}
take
with
thee
one
or
two
more,
that
in
the mouth
of
two
or
three
witnesses
every
word
may be established.}}
\VS{17}{\WJ{And
if
he shall neglect to hear
them,
tell
{\ADD{it}}
unto the church:
but
if
he neglect to hear
the church,
let him be
unto thee
as
an heathen man
and
a publican.}}
\VS{18}{\WJ{Verily
I say
unto you,
Whatsoever
ye shall bind
on
earth
shall be
bound
in
heaven:
and
whatsoever
ye shall loose
on
earth
shall be
loosed
in
heaven.}}
\VS{19}{\WJ{Again
I say
unto you,
That
if
two
of you
shall agree
on
earth
as touching
any
thing
that
they shall ask,
it shall be done
for them
of
my
Father
which is
in
heaven.}}
\VS{20}{\WJ{For
where
two
or
three
are
gathered together
in
my
name,
there
am I
in
the midst
of them.}}
\par }{\PP \VS{21}¶
Then
came
Peter to
him, and
said,
Lord, how
oft
shall
my
brother
sin
against
me,
and I
forgive
him?
till seven
times?
\VS{22}Jesus
saith unto
him,
{\WJ{I say
not
unto thee,
Until
seven times:
but,
Until
seventy times
seven.}}
\par }{\PP \VS{23}{\WJ{¶
Therefore
is
the kingdom
of heaven
likened
unto a certain
king,
which
would
take
account
of
his
servants.}}
\VS{24}{\WJ{And
when
he
had begun
to reckon,
one
was brought
unto him,
which owed him
ten thousand
talents.}}
\VS{25}{\WJ{But forasmuch as
he
had
not
to pay,
his
lord
commanded
him
to be sold,
and
his
wife,
and
children,
and
all
that
he had,
and
payment to be made.}}
\VS{26}{\WJ{The servant
therefore
fell down,
and worshipped
him,
saying,
Lord,
have patience
with
me,
and
I will pay
thee
all.}}
\VS{27}{\WJ{Then
the lord
of
that
servant
was moved with compassion,
and loosed
him,
and
forgave
him
the debt.}}
\VS{28}{\WJ{But
the
same
servant
went out,
and found
one
of
his
fellowservants,
which
owed
him
an hundred
pence:
and
he laid hands
on him,
and took him by the throat,
saying,
Pay
me
that
thou owest.}}
\VS{29}{\WJ{And
his
fellowservant
fell down
at
his
feet,
and besought
him,
saying,
Have patience
with
me,
and
I will pay
thee
all.}}
\VS{30}{\WJ{And
he would
not:
but
went and
cast
him
into
prison,
till
he should pay
the debt.}}
\VS{31}{\WJ{So
when
his
fellowservants
saw
what was done,
they were
very
sorry,
and
came
and told
unto
their
lord
all
that was done.}}
\VS{32}{\WJ{Then
his
lord,
after that he had called
him,
said
unto him,
O thou
wicked
servant,
I forgave
thee
all
that
debt,
because
thou desiredst
me:}}
\VS{33}{\WJ{Shouldest
not
thou
also
have had compassion
on
thy
fellowservant,
even
as
I
had pity
on thee?}}
\VS{34}{\WJ{And
his
lord
was wroth,
and delivered
him
to the tormentors,
till
he should pay
all
that was due
unto him.}}
\VS{35}{\WJ{So likewise
shall
my
heavenly
Father
do
also
unto you,
if
ye
from
your
hearts
forgive
not
every one
his
brother
their
trespasses.}}

\par }\Chap{19}{\PP \VerseOne{1}And it came to
pass,
{\ADD{that}}
when
Jesus had
finished
these
sayings, he
departed
from
Galilee,
and
came
into the
coasts of
Judaea
beyond
Jordan;
\VS{2}And
great
multitudes
followed
him;
and he
healed
them
there.
\par }{\PP \VS{3}¶ The
Pharisees
also
came unto
him,
tempting
him,
and
saying unto
him, Is
it
lawful for a
man to put
away
his
wife
for
every
cause?
\VS{4}And he
answered and
said unto
them,
{\WJ{Have ye
not
read,
that
he which made
{\ADD{them}}
at
the beginning
made
them
male
and
female,}}
\VS{5}{\WJ{And
said,
For
this
cause
shall
a man
leave
father
and
mother,
and
shall cleave
to
his
wife:
and
they
twain
shall be
one
flesh?}}
\VS{6}{\WJ{Wherefore
they are
no more
twain,
but
one
flesh.
What
therefore
God
hath joined together,
let
not
man
put asunder.}}
\VS{7}They
say unto
him,
Why
did
Moses
then
command to
give a
writing of
divorcement,
and to
put
her
away?
\VS{8}He
saith unto
them,
{\WJ{Moses
because
of the hardness of
your
hearts
suffered
you
to put away
your
wives:
but
from
the beginning
it was
not
so.}}
\VS{9}{\WJ{And
I say
unto you,
Whosoever
shall put away
his
wife,
except
{\ADD{it be}}
for
fornication,
and
shall marry
another,
committeth adultery:
and
whoso marrieth
her which is put away
doth commit adultery.}}
\par }{\PP \VS{10}¶
His
disciples
say unto
him,
If the
case of the
man
be
so
with
{\ADD{his}}
wife, it
is
not
good to
marry.
\VS{11}But he
said unto
them,
{\WJ{All
{\ADD{men}}
cannot
receive
this
saying,
save
{\ADD{they}}
to whom
it is given.}}
\VS{12}{\WJ{For
there are
some
eunuchs,
which
were
so
born
from
{\ADD{their}}
mother’s
womb:
and
there are
some
eunuchs,
which
were made eunuchs
of
men:
and
there be
eunuchs,
which
have made
themselves
eunuchs
for
the kingdom
of heaven’s
sake.
He that is able
to receive
{\ADD{it}},
let him receive
{\ADD{it.}}}}
\par }{\PP \VS{13}¶
Then were there
brought unto
him little
children,
that he should
put
{\ADD{his}}
hands on
them,
and
pray:
and the
disciples
rebuked
them.
\VS{14}But
Jesus
said,
{\WJ{Suffer
little children,
and
forbid
them
not,
to come
unto
me:
for
of such
is
the kingdom
of heaven.}}
\VS{15}And he
laid
{\ADD{his}}
hands on
them, and
departed
thence.
\par }{\PP \VS{16}¶
And,
behold,
one
came and
said unto
him,
Good
Master,
what good
thing shall I
do,
that I may
have
eternal
life?
\VS{17}And he
said unto
him,
{\WJ{Why
callest thou
me
good?
{\ADD{there is}}
none
good
but
one,
{\ADD{that is}},
God:
but
if
thou wilt
enter
into
life,
keep
the commandments.}}
\VS{18}He
saith unto
him,
Which?
Jesus
said,
{\WJ{Thou shalt
do
no
murder, Thou
shalt
not commit
adultery, Thou
shalt
not
steal, Thou
shalt
not bear false
witness,}}
\VS{19}{\WJ{Honour
thy
father
and
{\ADD{thy}}
mother:
and,
Thou shalt love
thy
neighbour
as
thyself.}}
\VS{20}The young
man
saith unto
him,
All these
things have I
kept
from
my
youth
up:
what lack
I
yet?
\VS{21}Jesus
said unto
him,
{\WJ{If
thou wilt
be
perfect,
go
{\ADD{and}}
sell
that
thou
hast,
and
give
to the poor,
and
thou shalt have
treasure
in
heaven:
and
come
{\ADD{and}}
follow
me.}}
\VS{22}But
when the young
man
heard that
saying, he went
away
sorrowful:
for he
had
great
possessions.
\par }{\PP \VS{23}¶
Then
said
Jesus
unto
his
disciples,
{\WJ{Verily
I say
unto you,
That
a rich man
shall
hardly
enter
into
the kingdom
of heaven.}}
\VS{24}{\WJ{And
again
I say
unto you,
It is
easier
for
a camel
to go
through
the eye
of a needle,
than
for
a rich man
to enter
into
the kingdom
of God.}}
\VS{25}When
his
disciples
heard
{\ADD{it}}, they
were
exceedingly
amazed,
saying,
Who
then
can be
saved?
\VS{26}But
Jesus
beheld
{\ADD{them}}, and
said unto
them,
{\WJ{With
men
this
is
impossible;
but
with
God
all things
are
possible.}}
\par }{\PP \VS{27}¶
Then
answered
Peter and
said unto
him,
Behold,
we have
forsaken
all,
and
followed
thee;
what
shall
we
have
therefore?
\VS{28}And
Jesus
said unto
them,
{\WJ{Verily
I say
unto you,
That
ye
which have followed
me,
in
the regeneration
when
the Son
of man
shall sit
in
the throne
of
his
glory,
ye
also
shall sit
upon
twelve
thrones,
judging
the twelve
tribes
of Israel.}}
\VS{29}{\WJ{And
every
one
that hath forsaken
houses,
or
brethren,
or
sisters,
or
father,
or
mother,
or
wife,
or
children,
or
lands,
for
my
name’s
sake,
shall receive
an hundredfold,
and
shall inherit
everlasting
life.}}
\VS{30}{\WJ{But
many
{\ADD{that are}}
first
shall be
last;
and
the last
{\ADD{shall be}}
first.}}

\par }\Chap{20}{\PP \VerseOne{1}{\WJ{For
the kingdom
of heaven
is
like
unto a man
{\ADD{that is}}
an householder,
which
went out
early
in the morning
to hire
labourers
into
his
vineyard.}}
\VS{2}{\WJ{And
when he had agreed
with
the labourers
for
a penny
a day,
he sent
them
into
his
vineyard.}}
\VS{3}{\WJ{And
he went out
about
the
third
hour,
and saw
others
standing
idle
in
the marketplace,}}
\VS{4}{\WJ{And
said unto them;
Go
ye
also
into
the vineyard,
and
whatsoever
is
right
I will give
you.
And
they went their way.}}
\VS{5}{\WJ{Again
he went out
about
the
sixth
and
ninth
hour,
and did
likewise.}}
\VS{6}{\WJ{And
about
the
eleventh
hour
he went out,
and found
others
standing
idle,
and
saith
unto them,
Why
stand ye
here
all
the day
idle?}}
\VS{7}{\WJ{They say
unto him,
Because
no man
hath hired
us.
He saith
unto them,
Go
ye
also
into
the vineyard;
and
whatsoever
is
right,
{\ADD{that}}
shall ye receive.}}
\VS{8}{\WJ{So
when
even
was come,
the lord
of the vineyard
saith
unto
his
steward,
Call
the labourers,
and
give
them
{\ADD{their}}
hire,
beginning
from
the last
unto
the first.}}
\VS{9}{\WJ{And
when they came
that
{\ADD{were hired}}
about
the
eleventh
hour,
they received
every man
a penny.}}
\VS{10}{\WJ{But
when
the first
came,
they supposed
that
they should have received
more;
and
they
likewise
received
every man
a penny.}}
\VS{11}{\WJ{And
when they had received
{\ADD{it}},
they murmured
against
the goodman of the house,}}
\VS{12}{\WJ{Saying,
These
last
have wrought
{\ADD{but}}
one
hour,
and
thou hast made
them
equal
unto us,
which have borne
the burden
and
heat
of the day.}}
\VS{13}{\WJ{But
he answered
one
of them,
and said,
Friend,
I do
thee
no
wrong:
didst
not
thou agree
with me
for a penny?}}
\VS{14}{\WJ{Take
{\ADD{that}}
thine
{\ADD{is}},
and
go thy way:
I will
give
unto
this
last,
even
as
unto thee.}}
\VS{15}{\WJ{Is it
not
lawful
for me
to do
what
I will
with
mine own?
Is
thine
eye
evil,
because
I
am
good?}}
\VS{16}{\WJ{So
the last
shall be
first,
and
the first
last:
for
many
be
called,
but
few
chosen.}}
\par }{\PP \VS{17}¶
And
Jesus going
up
to
Jerusalem
took the
twelve
disciples
apart
in the
way,
and
said unto
them,
\VS{18}{\WJ{Behold,
we go up
to
Jerusalem;
and
the Son
of man
shall be betrayed
unto the chief priests
and
unto the scribes,
and
they shall condemn
him
to death,}}
\VS{19}{\WJ{And
shall deliver
him
to the Gentiles
to
mock,
and
to scourge,
and
to crucify
{\ADD{him}}:
and
the
third
day
he shall rise again.}}
\par }{\PP \VS{20}¶
Then
came to
him the
mother
of
Zebedee’s
children
with
her
sons,
worshipping
{\ADD{him}},
and
desiring a certain
thing
of
him.
\VS{21}And he
said unto
her,
{\WJ{What
wilt thou?}} She
saith unto
him,
Grant
that
these
my
two
sons may
sit, the
one
on
thy right
hand,
and the
other
on the
left,
in
thy
kingdom.
\VS{22}But
Jesus
answered and
said,
{\WJ{Ye know
not
what
ye ask.
Are ye able
to drink
of the cup
that
I
shall
drink of,
and
to be baptized
with the baptism
that
I
am baptized with?}} They
say unto
him, We are
able.
\VS{23}And he
saith unto
them,
{\WJ{Ye shall drink
indeed
of
my
cup,
and
be baptized
with the baptism
that
I
am baptized with:
but
to sit
on
my
right hand,
and
on
my
left,
is
not
mine
to give,
but
{\ADD{it shall be given to them}}
for whom
it is prepared
of
my
Father.}}
\VS{24}And
when the
ten
heard
{\ADD{it}}, they were moved with
indignation
against the
two
brethren.
\VS{25}But
Jesus
called
them
{\ADD{unto him}}, and
said,
{\WJ{Ye know
that
the princes
of the Gentiles
exercise dominion over
them,
and
they that are great
exercise authority upon
them.}}
\VS{26}{\WJ{But
it shall
not
be
so
among
you:
but
whosoever
will be
great
among
you,
let him
be
your
minister;}}
\VS{27}{\WJ{And
whosoever
will
be
chief
among
you,
let him be
your
servant:}}
\VS{28}{\WJ{Even as
the Son
of man
came
not
to be ministered unto,
but
to minister,
and
to give
his
life
a ransom
for
many.}}
\VS{29}And
as
they
departed
from
Jericho,
a
great
multitude
followed
him.
\par }{\PP \VS{30}¶
And,
behold,
two blind
men
sitting
by the way
side, when they
heard
that
Jesus passed
by, cried
out,
saying, Have mercy
on
us, O
Lord,
{\ADD{thou}}
Son of
David.
\VS{31}And the
multitude
rebuked
them,
because they should hold their
peace:
but they
cried the
more,
saying, Have mercy
on
us, O
Lord,
{\ADD{thou}}
Son of
David.
\VS{32}And
Jesus stood
still, and
called
them,
and
said,
{\WJ{What
will ye
that I shall do
unto you?}}
\VS{33}They
say unto
him,
Lord,
that
our
eyes may be
opened.
\VS{34}So
Jesus had
compassion
{\ADD{on them}}, and
touched
their
eyes:
and
immediately
their
eyes received
sight,
and they
followed
him.

\par }\Chap{21}{\PP \VerseOne{1}And
when they drew
nigh
unto
Jerusalem,
and were
come
to
Bethphage,
unto the
mount of
Olives,
then
sent
Jesus
two
disciples,
\VS{2}Saying unto
them,
{\WJ{Go
into
the village
over against
you,
and
straightway
ye shall find
an ass
tied,
and
a colt
with
her:
loose
{\ADD{them}},
and bring
{\ADD{them}}
unto me.}}
\VS{3}{\WJ{And
if
any
{\ADD{man}}
say
ought
unto you,
ye shall say,
The Lord
hath
need
of them;
and
straightway
he will send
them.}}
\VS{4}All
this was
done,
that it might be
fulfilled which was
spoken
by the
prophet,
saying,
\VS{5}Tell
ye the
daughter of
Sion,
Behold,
thy
King
cometh unto
thee,
meek,
and
sitting
upon an
ass,
and a
colt the
foal of an
ass.
\VS{6}And the
disciples
went,
and
did
as
Jesus
commanded
them,
\VS{7}And
brought the
ass,
and the
colt,
and
put
on
them
their
clothes,
and they
set
{\ADD{him}}
thereon.
\VS{8}And
a very
great
multitude
spread
their
garments
in the
way;
others cut
down
branches
from the
trees,
and
strawed
{\ADD{them}}
in the
way.
\VS{9}And the
multitudes that went
before,
and that
followed,
cried,
saying,
Hosanna to the
Son of
David:
Blessed
{\ADD{is}} he that
cometh
in the
name of the
Lord;
Hosanna
in the
highest.
\VS{10}And
when
he was
come
into
Jerusalem,
all the
city was
moved,
saying,
Who
is
this?
\VS{11}And the
multitude
said,
This
is
Jesus the
prophet
of
Nazareth of
Galilee.
\par }{\PP \VS{12}¶
And
Jesus
went
into the
temple of
God,
and cast
out
all them that
sold
and
bought
in the
temple,
and
overthrew the
tables of the
moneychangers,
and the
seats of them that
sold
doves,
\VS{13}And
said unto
them,
{\WJ{It is written,
My
house
shall be called
the house
of prayer;
but
ye
have made
it
a den
of thieves.}}
\VS{14}And the
blind
and the
lame
came to
him
in the
temple;
and he
healed
them.
\VS{15}And
when the chief
priests
and
scribes
saw the wonderful
things
that he
did,
and the
children
crying
in the
temple,
and
saying,
Hosanna to the
Son of
David; they were sore
displeased,
\VS{16}And
said unto
him, Hearest
thou
what
these
say?
And
Jesus
saith unto
them,
{\WJ{Yea;
have ye
never
read,
Out
of the mouth
of babes
and
sucklings
thou hast perfected
praise?}}
\par }{\PP \VS{17}¶
And he
left
them, and
went
out of the
city
into
Bethany;
and he
lodged
there.
\VS{18}Now in the
morning as he
returned
into the
city, he
hungered.
\VS{19}And when he
saw
a fig
tree
in the
way, he
came
to
it,
and
found
nothing
thereon,
but
leaves
only,
and
said unto
it,
{\WJ{Let
no
fruit
grow
on
thee
henceforward for
ever.}}
And
presently the fig
tree withered
away.
\VS{20}And
when the
disciples
saw
{\ADD{it}}, they
marvelled,
saying,
How
soon
is the fig
tree withered
away!
\VS{21}Jesus
answered and
said unto
them,
{\WJ{Verily
I say
unto you,
If
ye have
faith,
and
doubt
not,
ye shall
not
only
do
this
{\ADD{which is done}}
to the fig tree,
but
also if
ye shall say
unto
this
mountain,
Be thou removed,
and
be thou cast
into
the sea;
it shall be done.}}
\VS{22}{\WJ{And
all things,
whatsoever
ye shall ask
in
prayer,
believing,
ye shall receive.}}
\par }{\PP \VS{23}¶
And
when
he was
come
into the
temple, the chief
priests
and the
elders of the
people
came unto
him as he was
teaching, and
said,
By
what
authority doest
thou these
things?
and
who
gave
thee
this
authority?
\VS{24}And
Jesus
answered and
said unto
them,
{\WJ{I also
will ask
you
one
thing,
which
if
ye tell
me,
I in like wise
will tell
you
by
what
authority
I do
these things.}}
\VS{25}{\WJ{The baptism
of John,
whence
was it?
from
heaven,
or
of
men?}}
And they
reasoned
with
themselves,
saying,
If we shall
say,
From
heaven; he will
say unto
us,
Why did
ye
not
then
believe
him?
\VS{26}But
if we shall
say,
Of
men; we
fear the
people;
for
all
hold
John
as a
prophet.
\VS{27}And they
answered
Jesus, and
said,
We
cannot
tell.
And
he
said unto
them,
{\WJ{Neither
tell
I
you
by
what
authority
I do
these things.}}
\par }{\PP \VS{28}{\WJ{¶
But
what
think
ye?
A certain man
had
two
sons;
and
he came
to the first,
and said,
Son,
go
work
to day
in
my
vineyard.}}
\VS{29}{\WJ{He answered
and said,
I will
not:
but
afterward
he repented,
and went.}}
\VS{30}{\WJ{And
he came
to the second,
and said
likewise.
And
he answered
and said,
I
{\ADD{go}},
sir:
and
went
not.}}
\VS{31}{\WJ{Whether
of
them twain
did
the will
of his father?}} They
say unto
him, The
first.
Jesus
saith unto
them,
{\WJ{Verily
I say
unto you,
That
the publicans
and
the harlots
go
into
the kingdom
of God
before
you.}}
\VS{32}{\WJ{For
John
came
unto
you
in
the way
of righteousness,
and
ye believed
him
not:
but
the publicans
and
the harlots
believed
him:
and
ye,
when ye had seen
{\ADD{it}},
repented
not
afterward,
that
ye might believe
him.}}
\par }{\PP \VS{33}{\WJ{¶
Hear
another
parable:
There was
a certain
householder,
which
planted
a vineyard,
and
hedged
it
round about,
and
digged
a winepress
in
it,
and
built
a tower,
and
let
it
out
to husbandmen,
and
went into a far country:}}
\VS{34}{\WJ{And
when
the time
of the fruit
drew near,
he sent
his
servants
to
the husbandmen,
that they might receive
the fruits
of it.}}
\VS{35}{\WJ{And
the husbandmen
took
his
servants,
and beat
one,
and
killed
another,
and
stoned
another.}}
\VS{36}{\WJ{Again,
he sent
other
servants
more
than the first:
and
they did
unto them
likewise.}}
\VS{37}{\WJ{But
last of all
he sent
unto
them
his
son,
saying,
They will reverence
my
son.}}
\VS{38}{\WJ{But
when
the husbandmen
saw
the son,
they said
among
themselves,
This
is
the heir;
come,
let us kill
him,
and
let us seize
on
his
inheritance.}}
\VS{39}{\WJ{And
they caught
him,
and cast
{\ADD{him}}
out
of the vineyard,
and
slew
{\ADD{him}}.}}
\VS{40}{\WJ{When
the lord
therefore
of the vineyard
cometh,
what
will he do
unto
those
husbandmen?}}
\VS{41}They
say unto
him, He
will
miserably
destroy
those wicked
men,
and will let
out
{\ADD{his}}
vineyard
unto
other
husbandmen,
which shall
render
him the
fruits
in
their
seasons.
\VS{42}Jesus
saith unto
them,
{\WJ{Did ye
never
read
in
the scriptures,
The stone
which
the builders
rejected,
the same
is become
the head
of the corner:
this
is
the Lord’s
doing,
and
it is
marvellous
in
our
eyes?}}
\VS{43}{\WJ{Therefore
say I
unto you,
The kingdom
of God
shall be taken
from
you,
and
given
to a nation
bringing forth
the fruits
thereof.}}
\VS{44}{\WJ{And
whosoever shall fall
on
this
stone
shall be broken:
but
on
whomsoever
it shall fall,
it will grind
him
to powder.}}
\VS{45}And
when the chief
priests
and
Pharisees had
heard
his
parables, they
perceived
that he
spake
of
them.
\VS{46}But when they
sought to lay hands
on
him, they
feared the
multitude,
because they
took
him
for a
prophet.

\par }\Chap{22}{\PP \VerseOne{1}And
Jesus
answered and
spake unto
them
again
by
parables, and
said,
\VS{2}{\WJ{The kingdom
of heaven
is like unto
a certain
king,
which
made
a marriage
for
his
son,}}
\VS{3}{\WJ{And
sent forth
his
servants
to call
them that were bidden
to
the wedding:
and
they would
not
come.}}
\VS{4}{\WJ{Again,
he sent forth
other
servants,
saying,
Tell
them which are bidden,
Behold,
I have prepared
my
dinner:
my
oxen
and
{\ADD{my}}
fatlings
{\ADD{are}}
killed,
and
all things
{\ADD{are}}
ready:
come
unto
the marriage.}}
\VS{5}{\WJ{But
they made light of
{\ADD{it}},
and went their ways,
one
to
his
farm,
another
to
his
merchandise:}}
\VS{6}{\WJ{And
the remnant
took
his
servants,
and entreated them spitefully,
and
slew
{\ADD{them}}.}}
\VS{7}{\WJ{But
when
the king
heard
{\ADD{thereof}},
he was wroth:
and
he sent forth
his
armies,
and destroyed
those
murderers,
and
burned up
their
city.}}
\VS{8}{\WJ{Then
saith he
to
his
servants,
The wedding
is
ready,
but
they which were bidden
were
not
worthy.}}
\VS{9}{\WJ{Go ye
therefore
into
the highways,
and
as many
as
ye shall find,
bid
to
the marriage.}}
\VS{10}{\WJ{So
those
servants
went out
into
the highways,
and gathered together
all
as many as
they found,
both
bad
and
good:
and
the wedding
was furnished
with guests.}}
\par }{\PP \VS{11}{\WJ{¶
And
when
the king
came in
to see
the guests,
he saw
there
a man
which had
not
on
a wedding
garment:}}
\VS{12}{\WJ{And
he saith
unto him,
Friend,
how
camest thou in
hither
not
having
a
wedding
garment?
And
he was speechless.}}
\VS{13}{\WJ{Then
said
the king
to the servants,
Bind
him
hand
and
foot,
and take
him
away,
and
cast
{\ADD{him}}
into
outer
darkness;
there
shall be
weeping
and
gnashing
of teeth.}}
\VS{14}{\WJ{For
many
are
called,
but
few
{\ADD{are}}
chosen.}}
\par }{\PP \VS{15}¶
Then
went the
Pharisees, and
took
counsel
how they might
entangle
him
in
{\ADD{his}}
talk.
\VS{16}And they sent
out unto
him
their
disciples
with the
Herodians,
saying,
Master, we
know
that thou
art
true,
and
teachest the
way of
God
in
truth,
neither
carest
thou
for
any
{\ADD{man}}:
for thou
regardest
not the
person of
men.
\VS{17}Tell
us
therefore,
What
thinkest
thou? Is it
lawful to
give
tribute unto
Caesar,
or
not?
\VS{18}But
Jesus
perceived
their
wickedness, and
said,
{\WJ{Why
tempt ye
me,
{\ADD{ye}}
hypocrites?}}
\VS{19}{\WJ{Shew
me
the tribute
money.}}
And they
brought unto
him a
penny.
\VS{20}And he
saith unto
them,
{\WJ{Whose
{\ADD{is}}
this
image
and
superscription?}}
\VS{21}They
say unto
him,
Cæsar’s.
Then saith
he unto
them,
{\WJ{Render
therefore
unto Cæsar
the things which are Cæsar’s;
and
unto God
the things that are God’s.}}
\VS{22}When they had
heard
{\ADD{these words}}, they
marvelled,
and
left
him, and went their
way.
\par }{\PP \VS{23}¶
The
same
day
came to
him the
Sadducees, which say
that there
is
no
resurrection,
and
asked
him,
\VS{24}Saying,
Master,
Moses
said,
If a
man
die,
having
no
children,
his
brother shall
marry
his
wife,
and raise
up
seed
unto
his
brother.
\VS{25}Now there
were
with
us
seven
brethren:
and the
first, when he had married a
wife,
deceased,
and,
having
no
issue,
left
his
wife
unto
his
brother:
\VS{26}Likewise the
second
also,
and the
third,
unto the
seventh.
\VS{27}And
last of
all the
woman
died
also.
\VS{28}Therefore
in the
resurrection
whose
wife shall she
be of the
seven?
for
they
all
had
her.
\VS{29}Jesus
answered and
said unto
them,
{\WJ{Ye do err,
not
knowing
the scriptures,
nor
the power
of God.}}
\VS{30}{\WJ{For
in
the resurrection
they
neither
marry,
nor
are given in marriage,
but
are
as
the angels
of God
in
heaven.}}
\VS{31}{\WJ{But
as touching
the resurrection
of the dead,
have ye
not
read
that which was spoken
unto you
by
God,
saying,}}
\VS{32}{\WJ{I
am
the God
of Abraham,
and
the God
of Isaac,
and
the God
of Jacob?
God
is
not
the God
of the dead,
but
of the living.}}
\VS{33}And
when the
multitude
heard
{\ADD{this}}, they were
astonished
at
his
doctrine.
\par }{\PP \VS{34}¶
But
when the
Pharisees had
heard
that he had
put the
Sadducees to
silence, they were
gathered
together.
\VS{35}Then
one
of
them,
{\ADD{which was}} a
lawyer,
asked
{\ADD{him a question}},
tempting
him,
and
saying,
\VS{36}Master,
which
{\ADD{is}}
the
great
commandment
in the
law?
\VS{37}Jesus
said unto
him,
{\WJ{Thou shalt love
the Lord
thy
God
with
all
thy
heart,
and
with
all
thy
soul,
and
with
all
thy
mind.}}
\VS{38}{\WJ{This
is
the first
and
great
commandment.}}
\VS{39}{\WJ{And
the second
{\ADD{is}}
like
unto it,
Thou shalt love
thy
neighbour
as
thyself.}}
\VS{40}{\WJ{On
these
two
commandments
hang
all
the law
and
the prophets.}}
\par }{\PP \VS{41}¶
While the
Pharisees were gathered
together,
Jesus
asked
them,
\VS{42}Saying,
{\WJ{What
think
ye
of
Christ?
whose
son
is he?}} They
say unto
him,
{\ADD{The Son}} of
David.
\VS{43}He
saith unto
them,
{\WJ{How
then
doth
David
in
spirit
call
him
Lord,
saying,}}
\VS{44}{\WJ{The
{\ND{
LORD}}
said
unto
my
Lord,
Sit thou
on
my
right hand,
till
I make
thine
enemies
thy
footstool?}}
\VS{45}{\WJ{If
David
then
call
him
Lord,
how
is he
his
son?}}
\VS{46}And no
man was
able to
answer
him a
word,
neither
durst
any
{\ADD{man}}
from
that
day
forth
ask
him any
more
{\ADD{questions.}}

\par }\Chap{23}{\PP \VerseOne{1}Then
spake
Jesus to the
multitude,
and
to
his
disciples,
\VS{2}Saying,
{\WJ{The scribes
and
the Pharisees
sit
in
Moses’
seat:}}
\VS{3}{\WJ{All
therefore
whatsoever
they bid
you
observe,
{\ADD{that}}
observe
and
do;
but
do
not
ye
after
their
works:
for
they say,
and
do
not.}}
\VS{4}{\WJ{For
they bind
heavy
burdens
and
grievous to be borne,
and
lay
{\ADD{them}}
on
men’s
shoulders;
but
they themselves will
not
move
them
with one of
their
fingers.}}
\VS{5}{\WJ{But
all
their
works
they do
for
to be seen
of men:
they make broad
their
phylacteries,
and
enlarge
the borders
of
their
garments,}}
\VS{6}{\WJ{And
love
the uppermost rooms
at
feasts,
and
the chief seats
in
the synagogues,}}
\VS{7}{\WJ{And
greetings
in
the markets,
and
to be called
of
men,
Rabbi,
Rabbi.}}
\VS{8}{\WJ{But
be
not
ye
called
Rabbi:
for
one
is
your
Master,
{\ADD{even}}
Christ;
and
all
ye
are
brethren.}}
\VS{9}{\WJ{And
call
no
{\ADD{man}}
your
father
upon
the earth:
for
one
is
your
Father,
which is
in
heaven.}}
\VS{10}{\WJ{Neither
be ye called
masters:
for
one
is
your
Master,
{\ADD{even}}
Christ.}}
\VS{11}{\WJ{But
he that is greatest
among you
shall be
your
servant.}}
\VS{12}{\WJ{And
whosoever
shall exalt
himself
shall be abased;
and
he that
shall humble
himself
shall be exalted.}}
\par }{\PP \VS{13}{\WJ{¶
But woe
unto you,
scribes
and
Pharisees,
hypocrites!
for
ye shut up
the kingdom
of heaven
against
men:
for
ye
neither
go in
{\ADD{yourselves}},
neither
suffer ye
them that are entering
to go in.}}
\VS{14}{\WJ{Woe
unto you,
scribes
and
Pharisees,
hypocrites!
for
ye devour
widows’
houses,
and
for a pretence
make
long
prayer:
therefore
ye shall receive
the greater
damnation.}}
\VS{15}{\WJ{Woe
unto you,
scribes
and
Pharisees,
hypocrites!
for
ye compass
sea
and
land
to make
one
proselyte,
and
when
he is made,
ye make
him
twofold more
the child
of hell
than yourselves.}}
\VS{16}{\WJ{Woe
unto you,
{\ADD{ye}}
blind
guides,
which say,
Whosoever
shall swear
by
the temple,
it is
nothing;
but
whosoever
shall swear
by
the gold
of the temple,
he is a debtor!}}
\VS{17}{\WJ{{\ADD{Ye}}
fools
and
blind:
for
whether
is
greater,
the gold,
or
the temple
that sanctifieth
the gold?}}
\VS{18}{\WJ{And,
Whosoever
shall swear
by
the altar,
it is
nothing;
but
whosoever
sweareth
by
the gift
that is upon
it,
he is guilty.}}
\VS{19}{\WJ{{\ADD{Ye}}
fools
and
blind:
for
whether
{\ADD{is}}
greater,
the gift,
or
the altar
that sanctifieth
the gift?}}
\VS{20}{\WJ{Whoso
therefore
shall swear
by
the altar,
sweareth
by
it,
and
by
all things
thereon.}}
\VS{21}{\WJ{And
whoso shall swear
by
the temple,
sweareth
by
it,
and
by
him that dwelleth
therein.}}
\VS{22}{\WJ{And
he that shall swear
by
heaven,
sweareth
by
the throne
of God,
and
by
him that sitteth
thereon.}}
\VS{23}{\WJ{Woe
unto you,
scribes
and
Pharisees,
hypocrites!
for
ye pay tithe
of mint
and
anise
and
cummin,
and
have omitted
the weightier
{\ADD{matters}}
of the law,
judgment,
mercy,
and
faith:
these
ought ye
to have done,
and
not
to leave
the other
undone.}}
\VS{24}{\WJ{{\ADD{Ye}}
blind
guides,
which strain at
a gnat,
and
swallow
a camel.}}
\VS{25}{\WJ{Woe
unto you,
scribes
and
Pharisees,
hypocrites!
for
ye make clean
the outside
of the cup
and
of the platter,
but
within
they are full
of
extortion
and
excess.}}
\VS{26}{\WJ{{\ADD{Thou}}
blind
Pharisee,
cleanse
first
that
{\ADD{which is}}
within
the cup
and
platter,
that
the outside
of them
may be
clean
also.}}
\VS{27}{\WJ{Woe
unto you,
scribes
and
Pharisees,
hypocrites!
for
ye are like unto
whited
sepulchres,
which
indeed
appear
beautiful
outward,
but
are
within
full
of
dead
{\ADD{men’s}}
bones,
and
of
all
uncleanness.}}
\VS{28}{\WJ{Even so
ye
also
outwardly
appear
righteous
unto men,
but
within
ye are
full
of hypocrisy
and
iniquity.}}
\VS{29}{\WJ{Woe
unto you,
scribes
and
Pharisees,
hypocrites!
because
ye build
the tombs
of the prophets,
and
garnish
the sepulchres
of the righteous,}}
\VS{30}{\WJ{And
say,
If
we had been
in
the days
of
our
fathers,
we
would
not
have been
partakers
with them
in
the blood
of the prophets.}}
\VS{31}{\WJ{Wherefore
ye be witnesses
unto yourselves,
that
ye are
the children
of them which killed
the prophets.}}
\VS{32}{\WJ{Fill
ye
up
then
the measure
of
your
fathers.}}
\VS{33}{\WJ{{\ADD{Ye}}
serpents,
{\ADD{ye}}
generation
of vipers,
how can
ye escape
the damnation
of hell?}}
\par }{\PP \VS{34}{\WJ{¶
Wherefore,
behold,
I
send
unto
you
prophets,
and
wise men,
and
scribes:
and
{\ADD{some}}
of
them
ye shall kill
and
crucify;
and
{\ADD{some}}
of
them
shall ye scourge
in
your
synagogues,
and
persecute
{\ADD{them}}
from
city
to
city:}}
\VS{35}{\WJ{That
upon
you
may come
all
the righteous
blood
shed
upon
the earth,
from
the blood
of
righteous
Abel
unto
the blood
of Zacharias
son
of Barachias,
whom
ye slew
between
the temple
and
the altar.}}
\VS{36}{\WJ{Verily
I say
unto you,
All
these things
shall come
upon
this
generation.}}
\VS{37}{\WJ{O Jerusalem,
Jerusalem,
{\ADD{thou}}
that killest
the prophets,
and
stonest
them which are sent
unto
thee,
how often
would I have
gathered
thy
children
together,
even
as
a hen
gathereth
her
chickens
under
{\ADD{her}}
wings,
and
ye would
not!}}
\VS{38}{\WJ{Behold,
your
house
is left
unto you
desolate.}}
\VS{39}{\WJ{For
I say
unto you,
Ye shall
not
see
me
henceforth,
till
ye shall say,
Blessed
{\ADD{is}}
he that cometh
in
the name
of the Lord.}}

\par }\Chap{24}{\PP \VerseOne{1}And
Jesus went
out, and
departed
from the
temple:
and
his
disciples came
to
{\ADD{him}} for to
shew
him the
buildings of the
temple.
\VS{2}And
Jesus
said unto
them,
{\WJ{See ye
not
all
these things?
verily
I say
unto you,
There shall
not
be left
here
one stone
upon
another,
that
shall
not
be thrown down.}}
\par }{\PP \VS{3}¶
And
as
he
sat
upon the
mount of
Olives, the
disciples
came unto
him
privately,
saying,
Tell
us,
when
shall these
things
be?
and
what
{\ADD{shall be}} the
sign
of
thy
coming,
and of the
end of the
world?
\VS{4}And
Jesus
answered and
said unto
them,
{\WJ{Take heed that
no
man
deceive
you.}}
\VS{5}{\WJ{For
many
shall come
in
my
name,
saying,
I
am
Christ;
and
shall deceive
many.}}
\VS{6}{\WJ{And
ye shall
hear
of wars
and
rumours
of wars:
see that
ye be
not
troubled:
for
all
{\ADD{these things}}
must
come to pass,
but
the end
is
not yet.}}
\VS{7}{\WJ{For
nation
shall rise
against
nation,
and
kingdom
against
kingdom:
and
there shall be
famines,
and
pestilences,
and
earthquakes,
in divers
places.}}
\VS{8}{\WJ{All
these
{\ADD{are}}
the beginning
of sorrows.}}
\VS{9}{\WJ{Then
shall they deliver
you
up to be
afflicted,
and
shall kill
you:
and
ye shall be
hated
of
all
nations
for
my
name’s
sake.}}
\VS{10}{\WJ{And
then
shall
many
be offended,
and
shall betray
one another,
and
shall hate
one another.}}
\VS{11}{\WJ{And
many
false prophets
shall rise,
and
shall deceive
many.}}
\VS{12}{\WJ{And
because
iniquity
shall abound,
the love
of many
shall wax cold.}}
\VS{13}{\WJ{But
he that shall endure
unto
the end,
the same
shall be saved.}}
\VS{14}{\WJ{And
this
gospel
of the kingdom
shall be preached
in
all
the world
for
a witness
unto
all
nations;
and
then
shall
the end
come.}}
\VS{15}{\WJ{When
ye
therefore
shall see
the abomination
of desolation,
spoken of
by
Daniel
the prophet,
stand
in
the
holy
place, (whoso readeth,
let him understand:)}}
\VS{16}{\WJ{Then
let them
which be
in
Judaea
flee
into
the mountains:}}
\VS{17}{\WJ{Let him
which is
on
the housetop
not
come down
to take
any thing
out of
his
house:}}
\VS{18}{\WJ{Neither
let him
which is
in
the field
return
back
to take
his
clothes.}}
\VS{19}{\WJ{And
woe
unto them that are
with child,
and
to them that give suck
in
those
days!}}
\VS{20}{\WJ{But
pray ye
that
your
flight
be
not
in the winter,
neither
on
the sabbath day:}}
\VS{21}{\WJ{For
then
shall be
great
tribulation,
such as
was
not
since
the beginning
of the world
to
this time,
no,
nor
ever
shall be.}}
\VS{22}{\WJ{And
except
those
days
should be shortened,
there should
no
flesh
be saved:
but
for
the elect’s
sake
those
days
shall be shortened.}}
\VS{23}{\WJ{Then
if
any man
shall say
unto you,
Lo,
here
{\ADD{is}}
Christ,
or
there;
believe
{\ADD{it}}
not.}}
\VS{24}{\WJ{For
there shall arise
false Christs,
and
false prophets,
and
shall shew
great
signs
and
wonders;
insomuch that,
if
{\ADD{it were}}
possible,
they shall deceive
the
very
elect.}}
\VS{25}{\WJ{Behold,
I have told
you
before.}}
\VS{26}{\WJ{Wherefore
if
they shall say
unto you,
Behold,
he is
in
the desert;
go
not
forth:
behold,
{\ADD{he is}}
in
the secret chambers;
believe
{\ADD{it}}
not.}}
\VS{27}{\WJ{For
as
the lightning
cometh
out
of the east,
and
shineth
even unto
the west;
so
shall
also
the coming
of the Son
of man
be.}}
\VS{28}{\WJ{For
wheresoever
the carcase
is,
there
will
the eagles
be gathered together.}}
\par }{\PP \VS{29}{\WJ{¶
Immediately
after
the tribulation
of
those
days
shall
the sun
be darkened,
and
the moon
shall
not
give
her
light,
and
the stars
shall fall
from
heaven,
and
the powers
of the heavens
shall be shaken:}}
\VS{30}{\WJ{And
then
shall appear
the sign
of the Son
of man
in
heaven:
and
then
shall
all
the tribes
of the earth
mourn,
and
they shall see
the Son
of man
coming
in
the clouds
of heaven
with
power
and
great
glory.}}
\VS{31}{\WJ{And
he shall send
his
angels
with
a
great
sound
of a trumpet,
and
they shall gather together
his
elect
from
the
four
winds,
from
one end
of heaven
to
the other.}}
\VS{32}{\WJ{Now
learn
a parable
of
the fig tree;
When
his
branch
is
yet
tender,
and
putteth forth
leaves,
ye know
that
summer
{\ADD{is}}
nigh:}}
\VS{33}{\WJ{So
likewise
ye,
when
ye shall see
all
these things,
know
that
it is
near,
{\ADD{even}}
at
the doors.}}
\VS{34}{\WJ{Verily
I say
unto you,
This
generation
shall
not
pass,
till
all
these things
be fulfilled.}}
\VS{35}{\WJ{Heaven
and
earth
shall pass away,
but
my
words
shall
not
pass away.}}
\par }{\PP \VS{36}{\WJ{¶
But
of
that
day
and
hour
knoweth
no
{\ADD{man}}, no,
not
the angels
of heaven,
but
my
Father
only.}}
\VS{37}{\WJ{But
as
the days
of Noe
{\ADD{were}},
so
shall
also
the coming
of the Son
of man
be.}}
\VS{38}{\WJ{For
as
in
the days
that
were
before
the flood
they were eating
and
drinking,
marrying
and
giving in marriage,
until
the day
that
Noe
entered
into
the ark,}}
\VS{39}{\WJ{And
knew
not
until
the flood
came,
and
took
them all
away;
so
shall
also
the coming
of the Son
of man
be.}}
\VS{40}{\WJ{Then
shall
two
be
in
the field;
the one
shall be taken,
and
the other
left.}}
\VS{41}{\WJ{Two
{\ADD{women shall be}}
grinding
at
the mill;
the one
shall be taken,
and
the other
left.}}
\par }{\PP \VS{42}{\WJ{¶
Watch
therefore:
for ye know
not
what
hour
your
Lord
doth come.}}
\VS{43}{\WJ{But
know
this,
that
if
the goodman of the house
had known
in
what
watch
the thief
would come,
he would have watched,
and
would
not
have suffered
his
house
to be broken up.}}
\VS{44}{\WJ{Therefore
be
ye
also
ready:
for
in
such
an hour
as ye think
not
the Son
of man
cometh.}}
\VS{45}{\WJ{Who
then
is
a faithful
and
wise
servant,
whom
his
lord
hath made ruler
over
his
household,
to give
them
meat
in
due season?}}
\VS{46}{\WJ{Blessed
{\ADD{is}}
that
servant,
whom
his
lord
when he cometh
shall find
so
doing.}}
\VS{47}{\WJ{Verily
I say
unto you,
That
he shall make
him
ruler
over
all
his
goods.}}
\VS{48}{\WJ{But
and if
that
evil
servant
shall say
in
his
heart,
My
lord
delayeth
his coming;}}
\VS{49}{\WJ{And
shall begin
to smite
{\ADD{his}}
fellowservants,
and
to eat
and
drink
with
the drunken;}}
\VS{50}{\WJ{The lord
of
that
servant
shall come
in
a day
when
he looketh
not
for
{\ADD{him}},
and
in
an hour
that
he is
not
aware of,}}
\VS{51}{\WJ{And
shall cut
him
asunder,
and
appoint
{\ADD{him}}
his
portion
with
the hypocrites:
there
shall be
weeping
and
gnashing
of teeth.}}

\par }\Chap{25}{\PP \VerseOne{1}{\WJ{Then shall
the kingdom
of heaven
be likened
unto
ten
virgins,
which
took
their
lamps,
and went
forth
to meet
the bridegroom.}}
\VS{2}{\WJ{And
five
of
them
were
wise,
and
five
{\ADD{were}}
foolish.}}
\VS{3}{\WJ{They that
{\ADD{were}}
foolish
took
their
lamps,
and took
no
oil
with
them:}}
\VS{4}{\WJ{But
the wise
took
oil
in
their
vessels
with
their
lamps.}}
\VS{5}{\WJ{While
the bridegroom
tarried,
they
all
slumbered
and
slept.}}
\VS{6}{\WJ{And at
midnight
there was
a cry
made,
Behold,
the bridegroom
cometh;
go ye
out
to meet
him.}}
\VS{7}{\WJ{Then
all
those
virgins
arose,
and
trimmed
their
lamps.}}
\VS{8}{\WJ{And
the foolish
said
unto the wise,
Give
us
of
your
oil;
for
our
lamps
are gone out.}}
\VS{9}{\WJ{But
the wise
answered,
saying,
{\ADD{Not so}};
lest
there be
not
enough
for us
and
you:
but
go ye
rather
to
them that sell,
and
buy
for yourselves.}}
\VS{10}{\WJ{And
while
they
went
to buy,
the bridegroom
came;
and
they that were ready
went in
with
him
to
the marriage:
and
the door
was shut.}}
\VS{11}{\WJ{Afterward
came
also
the
other
virgins,
saying,
Lord,
Lord,
open
to us.}}
\VS{12}{\WJ{But
he answered
and said,
Verily
I say
unto you,
I know
you
not.}}
\VS{13}{\WJ{Watch
therefore,
for
ye know
neither
the day
nor
the hour
wherein
the Son
of man
cometh.}}
\par }{\PP \VS{14}{\WJ{¶
For
{\ADD{the kingdom of heaven is}}
as
a man
travelling into a far country,
{\ADD{who}}
called
his own
servants,
and
delivered
unto them
his
goods.}}
\VS{15}{\WJ{And
unto one
he gave
five
talents,
to another
two,
and
to another
one;
to every man
according to
his several
ability;
and
straightway
took his journey.}}
\VS{16}{\WJ{Then
he that had received
the five
talents
went
and traded
with
the same,
and
made
{\ADD{them}}
other
five
talents.}}
\VS{17}{\WJ{And
likewise
he that
{\ADD{had received}}
two,
he
also
gained
other
two.}}
\VS{18}{\WJ{But
he that had received
one
went
and digged
in
the earth,
and
hid
his
lord’s
money.}}
\VS{19}{\WJ{After
a
long
time
the lord
of
those
servants
cometh,
and
reckoneth
with
them.}}
\VS{20}{\WJ{And so
he that had received
five
talents
came
and brought
other
five
talents,
saying,
Lord,
thou deliveredst
unto me
five
talents:
behold,
I have gained
beside
them
five
talents
more.}}
\VS{21}{\WJ{His
lord
said
unto him,
Well done,
{\ADD{thou}}
good
and
faithful
servant:
thou hast been
faithful
over
a few things,
I will make
thee
ruler
over
many things:
enter thou
into
the joy
of
thy
lord.}}
\VS{22}{\WJ{He
also
that had received
two
talents
came
and said,
Lord,
thou deliveredst
unto me
two
talents:
behold,
I have gained
two
other
talents
beside
them.}}
\VS{23}{\WJ{His
lord
said
unto him,
Well done,
good
and
faithful
servant;
thou hast been
faithful
over
a few things,
I will make
thee
ruler
over
many things:
enter thou
into
the joy
of
thy
lord.}}
\VS{24}{\WJ{Then
he which had received
the
one
talent
came
and said,
Lord,
I knew
thee
that
thou art
an
hard
man,
reaping
where
thou hast
not
sown,
and
gathering
where
thou hast
not
strawed:}}
\VS{25}{\WJ{And
I was afraid,
and went
and hid
thy
talent
in
the earth:
lo,
{\ADD{there}}
thou hast
{\ADD{that is}}
thine.}}
\VS{26}{\WJ{His
lord
answered
and said
unto him,
{\ADD{Thou}}
wicked
and
slothful
servant,
thou knewest
that
I reap
where
I sowed
not,
and
gather
where
I have
not
strawed:}}
\VS{27}{\WJ{Thou
oughtest
therefore
to have put
my
money
to the exchangers,
and
{\ADD{then}}
at my coming
I
should have received
mine own
with
usury.}}
\VS{28}{\WJ{Take
therefore
the talent
from
him,
and
give
{\ADD{it}}
unto him which hath
ten
talents.}}
\VS{29}{\WJ{For
unto
every
one that hath
shall be given,
and
he shall have abundance:
but
from
him
that hath
not
shall be taken
away
even
that which
he hath.}}
\VS{30}{\WJ{And
cast ye
the unprofitable
servant
into
outer
darkness:
there
shall be
weeping
and
gnashing
of teeth.}}
\par }{\PP \VS{31}{\WJ{¶
When
the Son
of man
shall come
in
his
glory,
and
all
the
holy
angels
with
him,
then
shall he sit
upon
the throne
of
his
glory:}}
\VS{32}{\WJ{And
before
him
shall be gathered
all
nations:
and
he shall separate
them
one
from
another,
as
a shepherd
divideth
{\ADD{his}}
sheep
from
the goats:}}
\VS{33}{\WJ{And
he shall set
the sheep
on
his
right hand,
but
the goats
on
the left.}}
\VS{34}{\WJ{Then
shall
the King
say
unto them
on
his
right hand, Come,
ye
blessed
of
my
Father,
inherit
the kingdom
prepared
for you
from
the foundation
of the world:}}
\VS{35}{\WJ{For
I was an hungred,
and
ye gave
me
meat:
I was thirsty,
and
ye gave
me
drink:
I was
a stranger,
and
ye took
me
in:}}
\VS{36}{\WJ{Naked,
and
ye clothed
me:
I was sick,
and
ye visited
me:
I was
in
prison,
and
ye came
unto
me.}}
\VS{37}{\WJ{Then
shall
the righteous
answer
him,
saying,
Lord,
when
saw we
thee
an hungred,
and
fed
{\ADD{thee}}?
or
thirsty,
and
gave thee drink?}}
\VS{38}{\WJ{When
saw we
thee
a stranger,
and
took thee in?
or
naked,
and
clothed
{\ADD{thee}}?}}
\VS{39}{\WJ{Or
when
saw we
thee
sick,
or
in
prison,
and
came
unto
thee?}}
\VS{40}{\WJ{And
the King
shall answer
and say
unto them,
Verily
I say
unto you,
Inasmuch
as
ye have done
{\ADD{it}}
unto
one
of the least
of these
my
brethren,
ye have done
{\ADD{it}}
unto me.}}
\VS{41}{\WJ{Then
shall he say
also
unto them
on
the left hand,
Depart
from
me,
ye cursed,
into
everlasting
fire,
prepared
for the devil
and
his
angels:}}
\VS{42}{\WJ{For
I was an hungred,
and
ye gave
me
no
meat:
I was thirsty,
and
ye gave
me
no
drink:}}
\VS{43}{\WJ{I was
a stranger,
and
ye took
me
not
in:
naked,
and
ye clothed
me
not:
sick,
and
in
prison,
and
ye visited
me
not.}}
\VS{44}{\WJ{Then
shall
they
also
answer
him,
saying,
Lord,
when
saw we
thee
an hungred,
or
athirst,
or
a stranger,
or
naked,
or
sick,
or
in
prison,
and
did
not
minister
unto thee?}}
\VS{45}{\WJ{Then
shall he answer
them,
saying,
Verily
I say
unto you,
Inasmuch
as
ye did
{\ADD{it}}
not
to one
of the least
of these,
ye did
{\ADD{it}}
not
to me.}}
\VS{46}{\WJ{And
these
shall go away
into
everlasting
punishment:
but
the righteous
into
life
eternal.}}

\par }\Chap{26}{\PP \VerseOne{1}And it came to
pass,
when
Jesus had
finished
all
these
sayings, he
said
unto
his
disciples,
\VS{2}{\WJ{Ye know
that
after
two
days
is
{\ADD{the feast of}}
the passover,
and
the Son
of man
is betrayed
to
be crucified.}}
\VS{3}Then assembled
together the chief
priests,
and the
scribes,
and the
elders of the
people,
unto the
palace of the high
priest, who was
called
Caiaphas,
\VS{4}And
consulted
that they might
take
Jesus by
subtilty,
and
kill
{\ADD{him}}.
\VS{5}But they
said,
Not
on the
feast
{\ADD{day}},
lest there
be an
uproar
among the
people.
\par }{\PP \VS{6}¶
Now
when
Jesus
was
in
Bethany,
in the
house of
Simon the
leper,
\VS{7}There came
unto
him a
woman
having an alabaster
box
of very
precious
ointment,
and poured
it
on
his
head, as he
sat
{\ADD{at meat}}.
\VS{8}But
when
his
disciples
saw
{\ADD{it}}, they had
indignation,
saying,
To what
purpose
{\ADD{is}}
this
waste?
\VS{9}For
this
ointment
might have been
sold for
much,
and
given to the
poor.
\VS{10}When
Jesus
understood
{\ADD{it}}, he
said unto
them,
{\WJ{Why
trouble
ye
the woman?
for
she hath wrought
a
good
work
upon
me.}}
\VS{11}{\WJ{For
ye have
the poor
always
with
you;
but
me
ye have
not
always.}}
\VS{12}{\WJ{For in that
she
hath poured
this
ointment
on
my
body,
she did
{\ADD{it}}
for
my
burial.}}
\VS{13}{\WJ{Verily
I say
unto you,
Wheresoever
this
gospel
shall be preached
in
the
whole
world,
{\ADD{there}}
shall
also this,
that
this woman
hath done,
be told
for
a memorial
of her.}}
\par }{\PP \VS{14}¶
Then
one of the
twelve,
called
Judas
Iscariot,
went
unto the chief
priests,
\VS{15}And
said
{\ADD{unto them}},
What will
ye
give
me, and
I will
deliver
him unto
you?
And they
covenanted with
him
for
thirty pieces of
silver.
\VS{16}And
from that
time he
sought
opportunity
to
betray
him.
\par }{\PP \VS{17}¶
Now the
first
{\ADD{day}} of the feast of unleavened
bread the
disciples
came to
Jesus,
saying unto
him,
Where wilt thou
that we
prepare for
thee to
eat the
passover?
\VS{18}And he
said,
{\WJ{Go
into
the city
to
such a man,
and
say
unto him,
The Master
saith,
My
time
is
at hand;
I will keep
the passover
at
thy house
with
my
disciples.}}
\VS{19}And the
disciples
did
as
Jesus had
appointed
them;
and they made
ready the
passover.
\VS{20}Now
when the
even was
come, he sat
down
with the
twelve.
\VS{21}And
as
they did
eat, he
said,
{\WJ{Verily
I say
unto you,
that
one
of
you
shall betray
me.}}
\VS{22}And they
were
exceeding
sorrowful, and
began every
one of
them to
say unto
him,
Lord, is
it
I?
\VS{23}And he
answered and
said,
{\WJ{He that dippeth
{\ADD{his}}
hand
with
me
in
the dish,
the same
shall betray
me.}}
\VS{24}{\WJ{The Son
of man
goeth
as
it is written
of
him:
but
woe
unto
that
man
by
whom
the Son
of man
is betrayed!
it had been
good
for
that
man
if
he
had
not
been born.}}
\VS{25}Then
Judas, which
betrayed
him,
answered and
said,
Master, is
it
I? He
said unto
him,
{\WJ{Thou
hast said.}}
\par }{\PP \VS{26}¶
And
as
they were
eating,
Jesus
took
bread,
and
blessed
{\ADD{it}}, and
brake
{\ADD{it}},
and
gave
{\ADD{it}} to the
disciples,
and
said,
{\WJ{Take,
eat;
this
is
my
body.}}
\VS{27}And he
took the
cup,
and gave
thanks, and
gave
{\ADD{it}} to
them,
saying,
{\WJ{Drink ye
all
of
it;}}
\VS{28}{\WJ{For
this
is
my
blood
of the
new
testament,
which is shed
for
many
for
the remission
of sins.}}
\VS{29}{\WJ{But
I say
unto you,
I will
not
drink
henceforth
of
this
fruit
of the vine,
until
that
day
when
I drink
it
new
with
you
in
my
Father’s
kingdom.}}
\VS{30}And when they had sung an
hymn, they went
out
into the
mount of
Olives.
\VS{31}Then
saith
Jesus unto
them,
{\WJ{All
ye
shall be offended
because of
me
this
night:
for
it is written,
I will smite
the shepherd,
and
the sheep
of the flock
shall be scattered abroad.}}
\VS{32}{\WJ{But
after
I
am risen again,
I will go before
you
into
Galilee.}}
\VS{33}Peter
answered
and
said unto
him,
Though
all
{\ADD{men}} shall be
offended because
of
thee,
{\ADD{yet}}
will
I
never be
offended.
\VS{34}Jesus
said unto
him,
{\WJ{Verily
I say
unto thee,
That
this
night,
before
the cock
crow,
thou shalt deny
me
thrice.}}
\VS{35}Peter
said unto
him,
Though
I
should
die
with
thee,
yet will
I
not
deny
thee.
Likewise
also
said
all the
disciples.
\par }{\PP \VS{36}¶
Then
cometh
Jesus
with
them
unto a
place
called
Gethsemane,
and
saith unto the
disciples,
{\WJ{Sit ye
here,
while
I go
and pray
yonder.}}
\VS{37}And he took with
him
Peter
and
the
two
sons of
Zebedee, and
began to be
sorrowful
and very
heavy.
\VS{38}Then saith
he unto
them,
{\WJ{My
soul
is
exceeding sorrowful,
even unto
death:
tarry ye
here,
and
watch
with
me.}}
\VS{39}And he
went a little
further, and
fell
on
his
face, and
prayed,
saying,
{\WJ{O
my
Father,
if
it be
possible,
let
this
cup
pass
from
me:
nevertheless
not
as
I
will,
but
as
thou
{\ADD{wilt}}.}}
\VS{40}And he
cometh
unto the
disciples,
and
findeth
them
asleep,
and
saith unto
Peter,
{\WJ{What,
could ye
not
watch
with
me
one
hour?}}
\VS{41}{\WJ{Watch
and
pray,
that
ye enter
not
into
temptation:
the spirit
indeed
{\ADD{is}}
willing,
but
the flesh
{\ADD{is}}
weak.}}
\VS{42}He went
away
again the second
time, and
prayed,
saying,
{\WJ{O
my
Father,
if
this
cup
may
not
pass away
from
me,
except
I drink
it,
thy
will
be done.}}
\VS{43}And he
came and
found
them
asleep
again:
for
their
eyes
were
heavy.
\VS{44}And he
left
them, and went
away
again, and
prayed the third
time,
saying
the
same
words.
\VS{45}Then cometh
he
to
his
disciples,
and
saith unto
them,
{\WJ{Sleep on
now,
and
take your rest:
behold,
the hour
is at hand,
and
the Son
of man
is betrayed
into
the hands
of sinners.}}
\VS{46}{\WJ{Rise,
let us be going:
behold,
he is at hand
that doth betray
me.}}
\par }{\PP \VS{47}¶
And
while
he
yet
spake,
lo,
Judas,
one of the
twelve,
came,
and
with
him
a
great
multitude
with
swords
and
staves,
from the chief
priests
and
elders of the
people.
\VS{48}Now he that
betrayed
him
gave
them a
sign,
saying,
Whomsoever I shall
kiss, that
same is
he:
hold
him
fast.
\VS{49}And
forthwith he came
to
Jesus, and
said,
Hail,
master;
and
kissed
him.
\VS{50}And
Jesus
said unto
him,
{\WJ{Friend,
wherefore
art thou come?}}
Then came they, and
laid
hands
on
Jesus,
and
took
him.
\VS{51}And,
behold,
one of them which
were
with
Jesus stretched
out
{\ADD{his}}
hand, and
drew
his
sword,
and
struck a
servant of the high
priest’s, and smote
off
his
ear.
\VS{52}Then
said
Jesus unto
him,
{\WJ{Put up again
thy
sword
into
his
place:
for
all
they that take
the sword
shall perish
with
the sword.}}
\VS{53}{\WJ{Thinkest thou
that
I cannot
now
pray
to
my
Father,
and
he shall presently give
me
more
than
twelve
legions
of angels?}}
\VS{54}{\WJ{But how
then
shall
the scriptures
be fulfilled,
that
thus
it must
be?}}
\VS{55}In that
same
hour
said
Jesus to the
multitudes,
{\WJ{Are ye come out
as
against
a thief
with
swords
and
staves
for to take
me?
I sat
daily
with
you
teaching
in
the temple,
and
ye laid
no
hold on
me.}}
\VS{56}{\WJ{But
all
this
was done,
that
the scriptures
of the prophets
might be fulfilled.}}
Then
all the
disciples
forsook
him, and
fled.
\par }{\PP \VS{57}¶
And they that had laid hold
on
Jesus
led
{\ADD{him}}
away
to
Caiaphas the high
priest,
where the
scribes
and the
elders were
assembled.
\VS{58}But
Peter
followed
him
afar
off
unto the high
priest’s
palace,
and
went
in, and
sat
with the
servants, to
see the
end.
\VS{59}Now the chief
priests,
and
elders,
and
all the
council,
sought false
witness
against
Jesus,
to
put
him to
death;
\VS{60}But
found
none: yea,
though
many false
witnesses
came,
{\ADD{yet}} found
they
none. At the
last
came
two false
witnesses,
\VS{61}And
said,
This
{\ADD{fellow}}
said, I am
able to
destroy the
temple of
God,
and to
build
it
in
three
days.
\VS{62}And the high
priest
arose, and
said unto
him, Answerest
thou
nothing?
what
{\ADD{is it which}}
these witness
against
thee?
\VS{63}But
Jesus held his
peace.
And the high
priest
answered and
said unto
him, I
adjure
thee
by
the
living
God,
that thou
tell
us
whether
thou
be the
Christ, the
Son of
God.
\VS{64}Jesus
saith unto
him,
{\WJ{Thou
hast said:
nevertheless
I say
unto you,
Hereafter
shall ye see
the Son
of man
sitting
on
the right hand
of power,
and
coming
in
the clouds
of heaven.}}
\VS{65}Then the high
priest
rent
his
clothes,
saying, He hath spoken
blasphemy;
what
further
need have
we of
witnesses?
behold,
now ye have
heard
his
blasphemy.
\VS{66}What
think
ye? They
answered
and
said, He
is guilty
of
death.
\VS{67}Then did they
spit
in
his
face,
and
buffeted
him;
and
others smote him with the palms of their
hands,
\VS{68}Saying,
Prophesy unto
us, thou
Christ,
Who is
he that
smote
thee?
\par }{\PP \VS{69}¶ Now
Peter
sat
without
in the
palace:
and
a
damsel came
unto
him,
saying,
Thou
also
wast
with
Jesus of
Galilee.
\VS{70}But he
denied
before
{\ADD{them}}
all,
saying, I
know
not
what thou
sayest.
\VS{71}And
when
he was gone
out
into the
porch,
another
{\ADD{maid}}
saw
him,
and
said unto them that were
there,
This
{\ADD{fellow}}
was
also
with
Jesus of
Nazareth.
\VS{72}And
again he
denied
with an
oath, I
do
not
know the
man.
\VS{73}And
after a
while came
unto
{\ADD{him}} they that stood
by, and
said to
Peter,
Surely
thou
also
art
{\ADD{one}}
of
them;
for
thy
speech
bewrayeth
thee.
\VS{74}Then began
he to
curse
and to
swear,
{\ADD{saying}}, I
know
not the
man.
And
immediately the
cock
crew.
\VS{75}And
Peter
remembered the
word of
Jesus, which
said unto
him,
Before the
cock
crow, thou shalt
deny
me
thrice.
And he
went
out, and
wept
bitterly.

\par }\Chap{27}{\PP \VerseOne{1}When the
morning was
come,
all the chief
priests
and
elders of the
people
took
counsel
against
Jesus
to
put
him to
death:
\VS{2}And when they had
bound
him, they
led
{\ADD{him}}
away,
and
delivered
him to
Pontius
Pilate the
governor.
\par }{\PP \VS{3}¶
Then
Judas, which had
betrayed
him, when he
saw
that he was
condemned, repented
himself, and brought
again the
thirty pieces of
silver to the chief
priests
and
elders,
\VS{4}Saying, I have
sinned in that I have
betrayed the
innocent
blood.
And
they
said,
What
{\ADD{is that}}
to
us?
see
thou
{\ADD{to that}}.
\VS{5}And he cast
down the pieces of
silver
in the
temple, and
departed,
and
went and hanged
himself.
\VS{6}And the chief
priests
took the silver
pieces, and
said, It
is
not
lawful for to
put
them
into the
treasury,
because it
is the
price of
blood.
\VS{7}And they
took
counsel, and
bought
with
them the
potter’s
field, to
bury
strangers
in.
\VS{8}Wherefore
that
field was
called, The
field of
blood,
unto this
day.
\VS{9}Then was
fulfilled that which was
spoken
by
Jeremy the
prophet,
saying,
And they
took
the
thirty pieces of
silver, the
price of him that was
valued,
whom they of the
children
of
Israel did
value;
\VS{10}And
gave
them
for
the
potter’s
field,
as the
Lord
appointed
me.
\VS{11}And
Jesus
stood
before the
governor:
and the
governor
asked
him,
saying,
Art
thou the
King of the
Jews?
And
Jesus
said unto
him,
{\WJ{Thou
sayest.}}
\VS{12}And
when
he was
accused
of the chief
priests
and
elders, he
answered
nothing.
\VS{13}Then
said
Pilate unto
him, Hearest
thou
not how many
things they witness
against
thee?
\VS{14}And he
answered
him
to
never
a
word; insomuch
that the
governor
marvelled
greatly.
\VS{15}Now
at
{\ADD{that}}
feast the
governor was
wont to
release
unto the
people a
prisoner,
whom they
would.
\VS{16}And they
had
then a
notable
prisoner,
called
Barabbas.
\VS{17}Therefore
when
they were gathered
together,
Pilate
said unto
them,
Whom will
ye that I
release unto
you?
Barabbas,
or
Jesus which is
called
Christ?
\VS{18}For he
knew
that
for
envy they had
delivered
him.
\par }{\PP \VS{19}¶
When
he was set
down
on the judgment
seat,
his
wife
sent
unto
him,
saying,
Have
thou nothing to
do
with
that just
man:
for I have
suffered many
things this
day
in a
dream because
of
him.
\VS{20}But the chief
priests
and
elders
persuaded the
multitude
that they should
ask
Barabbas,
and
destroy
Jesus.
\VS{21}The
governor
answered
and
said unto
them,
Whether
of the
twain will
ye that I
release unto
you?
They
said,
Barabbas.
\VS{22}Pilate
saith unto
them,
What shall I
do
then with
Jesus which is
called
Christ?
{\ADD{They}}
all say
unto
him, Let him be
crucified.
\VS{23}And the
governor
said,
Why,
what
evil hath he
done?
But
they cried
out the
more,
saying, Let him be
crucified.
\par }{\PP \VS{24}¶
When
Pilate
saw
that he could
prevail
nothing,
but
{\ADD{that}}
rather a
tumult was
made, he
took
water, and
washed
{\ADD{his}}
hands
before the
multitude,
saying, I
am
innocent
of the
blood of
this just
person:
see
ye
{\ADD{to it}}.
\VS{25}Then
answered
all the
people, and
said,
His
blood
{\ADD{be}}
on
us,
and
on
our
children.
\par }{\PP \VS{26}¶
Then released
he
Barabbas unto
them:
and when he had
scourged
Jesus, he
delivered
{\ADD{him}}
to be
crucified.
\VS{27}Then the
soldiers of the
governor
took
Jesus
into the common
hall, and
gathered
unto
him the
whole
band
{\ADD{of soldiers}}.
\VS{28}And they
stripped
him, and put
on
him a
scarlet
robe.
\par }{\PP \VS{29}¶
And when they had
platted a
crown
of
thorns, they
put
{\ADD{it}}
upon
his
head,
and a
reed
in
his right
hand:
and they bowed the
knee
before
him, and
mocked
him,
saying,
Hail,
King of the
Jews!
\VS{30}And they
spit
upon
him, and
took the
reed,
and
smote
him
on the
head.
\VS{31}And
after that they had
mocked
him, they
took the
robe
off from
him,
and
put his
own
raiment
on
him,
and
led
him
away
to
crucify
{\ADD{him}}.
\VS{32}And as they came
out, they
found a
man of
Cyrene,
Simon by
name:
him they
compelled
to
bear
his
cross.
\VS{33}And when they were
come
unto a
place
called
Golgotha,
that
is to
say, a
place of a
skull,
\par }{\PP \VS{34}¶ They
gave
him
vinegar to
drink
mingled
with
gall:
and when he had
tasted
{\ADD{thereof}}, he
would
not
drink.
\VS{35}And they
crucified
him, and
parted
his
garments,
casting
lots:
that it might be
fulfilled which was
spoken
by the
prophet, They
parted
my
garments among
them,
and
upon
my
vesture did they
cast
lots.
\VS{36}And sitting
down they
watched
him
there;
\VS{37}And set
up
over
his
head
his
accusation
written,
THIS
IS
JESUS THE
KING OF THE
JEWS.
\VS{38}Then were
there
two
thieves crucified
with
him,
one
on the right
hand,
and
another
on the
left.
\par }{\PP \VS{39}¶
And they that passed
by
reviled
him,
wagging
their
heads,
\VS{40}And
saying, Thou that
destroyest the
temple,
and
buildest
{\ADD{it}}
in
three
days,
save
thyself.
If thou
be the
Son of
God, come
down
from the
cross.
\VS{41}Likewise
also the chief
priests
mocking
{\ADD{him}},
with the
scribes
and
elders,
said,
\VS{42}He
saved
others;
himself
he
cannot
save.
If he
be the
King of
Israel, let
him
now come
down
from the
cross,
and we will
believe
him.
\VS{43}He
trusted
in
God; let him
deliver
him
now,
if he will
have
him:
for he
said, I
am the
Son of
God.
\VS{44}The
thieves
also, which were crucified
with
him,
cast the
same in
his
teeth.
\VS{45}Now
from the
sixth
hour there
was
darkness
over
all the
land
unto the
ninth
hour.
\VS{46}And
about
the
ninth
hour
Jesus cried
with
a
loud
voice,
saying,
{\WJ{Eli,
Eli,
lama
sabachthani?}}
that is to
say,
My
God,
my
God,
why hast thou
forsaken
me?
\VS{47}Some of them that
stood
there, when they
heard
{\ADD{that}},
said,
This
{\ADD{man}} calleth
for
Elias.
\VS{48}And
straightway
one
of
them
ran,
and
took a
spunge,
and
filled
{\ADD{it}} with
vinegar,
and
put
{\ADD{it}} on a
reed, and
gave
him to
drink.
\VS{49}The
rest
said, Let
be, let us
see
whether
Elias will
come to
save
him.
\par }{\PP \VS{50}¶
Jesus, when he had
cried
again
with
a
loud
voice, yielded
up the
ghost.
\VS{51}And,
behold, the
veil of the
temple was
rent
in
twain
from the
top
to the
bottom;
and the
earth did
quake,
and the
rocks
rent;
\VS{52}And the
graves were
opened;
and
many
bodies of the
saints which
slept
arose,
\VS{53}And came
out
of the
graves
after
his
resurrection, and
went
into
the
holy
city,
and
appeared unto
many.
\VS{54}Now
when the
centurion,
and they that
were
with
him,
watching
Jesus,
saw the
earthquake,
and those things that were
done, they
feared
greatly,
saying,
Truly
this
was the
Son of
God.
\VS{55}And
many
women
were
there
beholding
afar
off,
which
followed
Jesus
from
Galilee,
ministering unto
him:
\VS{56}Among
which
was
Mary
Magdalene,
and
Mary the
mother of
James
and
Joses,
and the
mother of
Zebedee’s
children.
\VS{57}When the
even was
come, there
came a
rich
man
of
Arimathæa,
named
Joseph,
who
also
himself
was
Jesus’
disciple:
\VS{58}He went
to
Pilate, and
begged the
body of
Jesus.
Then
Pilate
commanded the
body to be
delivered.
\VS{59}And
when
Joseph had
taken the
body, he
wrapped
it
in a
clean linen
cloth,
\VS{60}And
laid
it
in his
own
new
tomb,
which he had hewn
out
in the
rock:
and he
rolled
a
great
stone to the
door of the
sepulchre, and
departed.
\VS{61}And
there
was
Mary
Magdalene,
and the
other
Mary,
sitting over
against the
sepulchre.
\par }{\PP \VS{62}¶
Now the next
day,
that
followed the day of the
preparation, the chief
priests
and
Pharisees came
together
unto
Pilate,
\VS{63}Saying,
Sir, we
remember
that
that
deceiver
said, while he
was
yet
alive,
After
three
days I will rise
again.
\VS{64}Command
therefore
that the
sepulchre be made
sure
until
the
third
day,
lest
his
disciples
come by
night, and
steal him
away,
and
say unto the
people, He is
risen
from the
dead:
so the
last
error shall
be worse
than the
first.
\VS{65}Pilate
said unto
them, Ye
have a
watch: go your
way,
make
{\ADD{it}} as
sure
as ye
can.
\VS{66}So
they
went, and
made the
sepulchre
sure,
sealing the
stone, and
setting a
watch.

\par }\Chap{28}{\PP \VerseOne{1}In the
end of the
sabbath, as it began to
dawn
toward the
first
{\ADD{day}} of the
week,
came
Mary
Magdalene
and the
other
Mary to
see the
sepulchre.
\VS{2}And,
behold, there
was
a
great
earthquake:
for the
angel of the
Lord
descended
from
heaven, and
came and rolled
back the
stone
from the
door,
and
sat
upon
it.
\VS{3}His
countenance
was
like
lightning,
and
his
raiment
white
as
snow:
\VS{4}And
for
fear of
him the
keepers did
shake,
and
became
as
dead
{\ADD{men}}.
\VS{5}And the
angel
answered and
said unto the
women,
Fear
not
ye:
for I
know
that ye
seek
Jesus, which was
crucified.
\VS{6}He
is
not
here:
for he is
risen,
as he
said.
Come,
see the
place
where the
Lord
lay.
\VS{7}And
go
quickly, and
tell
his
disciples
that he is
risen
from the
dead;
and,
behold, he goeth
before
you
into
Galilee;
there shall ye
see
him:
lo, I have
told
you.
\VS{8}And they
departed
quickly
from the
sepulchre
with
fear
and
great
joy; and did
run to
bring
his
disciples
word.
\par }{\PP \VS{9}¶
And
as they
went to
tell
his
disciples,
behold,
Jesus
met
them,
saying,
{\WJ{All hail.}}
And they
came and
held
him by the
feet,
and
worshipped
him.
\VS{10}Then
said
Jesus unto
them,
{\WJ{Be
not
afraid:
go
tell
my
brethren
that
they go
into
Galilee,
and there
shall they see
me.}}
\par }{\PP \VS{11}¶
Now
when
they were
going,
behold,
some of the
watch
came
into the
city, and
shewed unto the chief
priests
all the things that were
done.
\VS{12}And when they were
assembled
with the
elders,
and had
taken
counsel, they
gave
large
money unto the
soldiers,
\VS{13}Saying, Say
ye,
His
disciples
came by
night, and
stole
him
{\ADD{away}}
while
we
slept.
\VS{14}And
if
this
come
to the
governor’s
ears,
we will
persuade
him,
and
secure
you.
\VS{15}So
they
took the
money, and
did
as they were
taught:
and
this
saying is commonly
reported
among the
Jews
until this
day.
\par }{\PP \VS{16}¶
Then
the
eleven
disciples went
away
into
Galilee,
into a
mountain
where
Jesus had
appointed
them.
\VS{17}And when they
saw
him, they
worshipped
him:
but
some
doubted.
\VS{18}And
Jesus
came and
spake unto
them,
saying,
{\WJ{All
power
is given
unto me
in
heaven
and
in
earth.}}
\par }{\PP \VS{19}{\WJ{¶
Go ye
therefore,
and teach
all
nations,
baptizing
them
in
the name
of the Father,
and
of the Son,
and
of the
Holy
Ghost:}}
\VS{20}{\WJ{Teaching
them
to observe
all things
whatsoever
I have commanded
you:
and,
lo,
I
am
with
you
alway,
{\ADD{even}}
unto
the end
of the world.}}
Amen.
\par }