\NormalFont\ShortTitle{Micah}
{\MT Micah

\par }\ChapOne{1}{\PP \VerseOne{1}The
word of the
{\ND{
LORD}} that came to
Micah the
Morasthite in the
days of
Jotham,
Ahaz,
{\ADD{and}}
Hezekiah,
kings of
Judah, which he
saw concerning
Samaria and
Jerusalem.
\VS{2}Hear, all ye
people;
hearken, O
earth, and all that
therein is: and let the
Lord
{\ND{
GOD}} be
witness against you, the
Lord from his
holy
temple.\FTNT{*}{{\FR 1.2 }
all ye…: Heb. ye people, all of them}\FTNT{†}{{\FR 1.2 }
all that…: Heb. the fulness thereof}
\VS{3}For, behold, the
{\ND{
LORD}} cometh
forth out of his
place, and will come
down, and
tread upon the high
places of the
earth.
\VS{4}And the
mountains shall be
molten under him, and the
valleys shall be
cleft, as
wax
before the
fire,
{\ADD{and}} as the
waters
{\ADD{that are}}
poured down a steep
place.\FTNT{‡}{{\FR 1.4 }
a steep…: Heb. a descent}
\VS{5}For the
transgression of
Jacob
{\ADD{is}} all this, and for the
sins of the
house of
Israel. What
{\ADD{is}} the
transgression of
Jacob?
{\ADD{is it}} not
Samaria? and what
{\ADD{are}} the high
places of
Judah?
{\ADD{are they}} not
Jerusalem?
\VS{6}Therefore I will
make
Samaria as an
heap of the
field,
{\ADD{and}} as
plantings of a
vineyard: and I will pour
down the
stones thereof into the
valley, and I will
discover the
foundations thereof.
\VS{7}And all the graven
images thereof shall be beaten to
pieces, and all the
hires thereof shall be
burned with the
fire, and all the
idols thereof will I
lay
desolate: for she
gathered
{\ADD{it}} of the
hire of an
harlot, and they shall
return to the
hire of an
harlot.
\par }{\PP \VS{8}Therefore I will
wail and
howl, I will
go
stripped and
naked: I will
make a
wailing like the
dragons, and
mourning as the
owls.\FTNT{§}{{\FR 1.8 }
owls: Heb. daughters of the owl}
\VS{9}For her
wound
{\ADD{is}}
incurable; for it is
come unto
Judah; he is
come unto the
gate of my
people,
{\ADD{even}} to
Jerusalem.\FTNT{**}{{\FR 1.9 }
her…: or, she is grievously sick of her wounds}
\par }{\PP \VS{10}¶
Declare ye
{\ADD{it}} not at
Gath,
weep ye not at
all: in the house of
Aphrah
roll thyself in the
dust.\FTNT{††}{{\FR 1.10 }
Aphrah: that is, Dust}
\VS{11}Pass ye
away, thou
inhabitant of
Saphir, having thy
shame
naked: the
inhabitant of
Zaanan came not
forth in the
mourning of
Beth-ezel; he shall
receive of you his
standing.\FTNT{‡‡}{{\FR 1.11 }
thou…: or, thou that dwellest fairly}\FTNT{§§}{{\FR 1.11 }
inhabitant: Heb. inhabitress}\FTNT{***}{{\FR 1.11 }
Zaanan: or, The country of flocks}\FTNT{†††}{{\FR 1.11 }
Beth-ezel: or, A place near}
\VS{12}For the
inhabitant of
Maroth waited
carefully for
good: but
evil came
down from the
{\ND{
LORD}} unto the
gate of
Jerusalem.\FTNT{‡‡‡}{{\FR 1.12 }
inhabitant: Heb. inhabitress}\FTNT{§§§}{{\FR 1.12 }
waited…: or, was grieved}
\VS{13}O thou
inhabitant of
Lachish,
bind the
chariot to the swift
beast: she
{\ADD{is}} the
beginning of the
sin to the
daughter of
Zion: for the
transgressions of
Israel were
found in thee.\FTNT{*}{{\FR 1.13 }
inhabitant: Heb. inhabitress}
\VS{14}Therefore shalt thou
give
presents to
Moresheth-gath: the
houses of
Achzib
{\ADD{shall be}} a
lie to the
kings of
Israel.\FTNT{†}{{\FR 1.14 }
to…: or, for Moresheth-gath}\FTNT{‡}{{\FR 1.14 }
Achzib: that is, A lie}
\VS{15}Yet will I
bring an
heir unto thee, O
inhabitant of
Mareshah: he shall
come unto
Adullam the
glory of
Israel.\FTNT{§}{{\FR 1.15 }
inhabitant: Heb. inhabitress}\FTNT{**}{{\FR 1.15 }
he…: or, the glory of Israel shall, etc}
\VS{16}Make thee
bald, and
poll thee for thy
delicate
children;
enlarge thy
baldness as the
eagle; for they are gone into
captivity from thee.

\par }\Chap{2}{\PP \VerseOne{1}Woe to them that
devise
iniquity, and
work
evil upon their
beds! when the
morning is
light, they
practise it, because it
is in the
power of their
hand.
\VS{2}And they
covet
fields, and take
{\ADD{them}} by
violence; and
houses, and take
{\ADD{them}}
away: so they
oppress a
man and his
house, even a
man and his
heritage.\FTNT{*}{{\FR 2.2 }
oppress: or, defraud}
\VS{3}Therefore thus
saith the
{\ND{
LORD}}; Behold, against this
family do I
devise an
evil, from which ye shall not
remove your
necks; neither shall ye
go
haughtily: for this
time
{\ADD{is}}
evil.
\par }{\PP \VS{4}¶ In that
day shall
{\ADD{one}} take
up a
parable against you, and
lament with a
doleful
lamentation,
{\ADD{and}}
say, We be
utterly
spoiled: he hath
changed the
portion of my
people: how hath he
removed
{\ADD{it}} from me! turning
away he hath
divided our
fields.\FTNT{†}{{\FR 2.4 }
a doleful…: Heb. a lamentation of lamentations}\FTNT{‡}{{\FR 2.4 }
turning…: or, instead of restoring}
\VS{5}Therefore thou shalt have none that shall
cast a
cord by
lot in the
congregation of the
{\ND{
LORD}}.
\par }{\PP \VS{6}Prophesy ye not,
{\ADD{say they to them that}}
prophesy: they shall not
prophesy to them,
{\ADD{that}} they shall not
take
shame.\FTNT{§}{{\FR 2.6 }
Prophesy ye…: or, Prophesy not as they prophesy: Heb. Drop, etc}
\par }{\PP \VS{7}¶ O
{\ADD{thou that art}}
named the
house of
Jacob, is the
spirit of the
{\ND{
LORD}}
straitened?
{\ADD{are}} these his
doings? do not my
words do
good to him that
walketh
uprightly?\FTNT{**}{{\FR 2.7 }
straitened: or, shortened?}\FTNT{††}{{\FR 2.7 }
uprightly: Heb. upright?}
\VS{8}Even of
late my
people is risen
up as an
enemy: ye pull
off the
robe
with the
garment from them that
pass by
securely as men
averse from
war.\FTNT{‡‡}{{\FR 2.8 }
of late: Heb. yesterday}\FTNT{§§}{{\FR 2.8 }
with the: Heb. over against a}
\VS{9}The
women of my
people have ye cast
out from their
pleasant
houses; from their
children have ye taken
away my
glory for
ever.\FTNT{***}{{\FR 2.9 }
women: or, wives}
\VS{10}Arise ye, and
depart; for this
{\ADD{is}} not
{\ADD{your}}
rest: because it is
polluted, it shall
destroy
{\ADD{you}}, even with a
sore
destruction.
\VS{11}If a
man
walking in the
spirit and
falsehood do
lie,
{\ADD{saying}}, I will
prophesy unto thee of
wine and of strong
drink; he shall even be the
prophet of this
people.\FTNT{†††}{{\FR 2.11 }
walking…: or, walk with the wind, and lie falsely}
\par }{\PP \VS{12}¶ I will
surely
assemble, O
Jacob, all of thee; I will
surely
gather the
remnant of
Israel; I will
put them
together as the
sheep of
Bozrah, as the
flock in the
midst of their
fold: they shall make great
noise by reason of
{\ADD{the multitude of}}
men.
\VS{13}The
breaker is come
up
before them: they have broken
up, and have passed
through the
gate, and are gone
out by it: and their
king shall
pass
before them, and the
{\ND{
LORD}} on the
head of them.

\par }\Chap{3}{\PP \VerseOne{1}And I
said,
Hear, I pray you, O
heads of
Jacob, and ye
princes of the
house of
Israel;
{\ADD{Is it}} not for you to
know
judgment?
\VS{2}Who
hate the
good, and
love the
evil; who pluck
off their
skin from off them, and their
flesh from off their
bones;
\VS{3}Who also
eat the
flesh of my
people, and
flay their
skin from off them; and they
break their
bones, and chop them in
pieces, as for the
pot, and as
flesh
within the
caldron.
\VS{4}Then shall they
cry unto the
{\ND{
LORD}}, but he will not
hear them: he will even
hide his
face from them at that
time, as they have behaved themselves
ill in their
doings.
\par }{\PP \VS{5}¶ Thus
saith the
{\ND{
LORD}} concerning the
prophets that make my
people
err, that
bite with their
teeth, and
cry,
Peace; and he that
putteth not into their
mouths, they even
prepare
war against him.
\VS{6}Therefore
night
{\ADD{shall be}} unto you, that ye shall not have a
vision; and it shall be
dark unto you, that ye shall not
divine; and the
sun shall go
down over the
prophets, and the
day shall be
dark over them.\FTNT{*}{{\FR 3.6 }
that…: vision: Heb. from a vision}\FTNT{†}{{\FR 3.6 }
that…: divine: Heb. from divining}
\VS{7}Then shall the
seers be
ashamed, and the
diviners
confounded: yea, they shall all
cover their
lips; for
{\ADD{there is}} no
answer of
God.\FTNT{‡}{{\FR 3.7 }
lips: Heb. upper lip}
\par }{\PP \VS{8}¶ But
truly I am
full of
power by the
spirit of the
{\ND{
LORD}}, and of
judgment, and of
might, to
declare unto
Jacob his
transgression, and to
Israel his
sin.
\VS{9}Hear this, I pray you, ye
heads of the
house of
Jacob, and
princes of the
house of
Israel, that
abhor
judgment, and
pervert all
equity.
\VS{10}They build
up
Zion with
blood, and
Jerusalem with
iniquity.\FTNT{§}{{\FR 3.10 }
blood: Heb. bloods}
\VS{11}The
heads thereof
judge for
reward, and the
priests thereof
teach for
hire, and the
prophets thereof
divine for
money: yet will they
lean upon the
{\ND{
LORD}}, and
say,
{\ADD{Is}} not the
{\ND{
LORD}}
among us? none
evil can
come upon us.\FTNT{**}{{\FR 3.11 }
and say: Heb. saying}
\VS{12}Therefore shall
Zion for your
sake be
plowed
{\ADD{as}} a
field, and
Jerusalem shall become
heaps, and the
mountain of the
house as the high
places of the
forest.

\par }\Chap{4}{\PP \VerseOne{1}But in the
last
days it shall come to pass,
{\ADD{that}} the
mountain of the
house of the
{\ND{
LORD}} shall be
established in the
top of the
mountains, and it shall be
exalted above the
hills; and
people shall
flow unto it.
\VS{2}And
many
nations shall
come, and
say,
Come, and let us go
up to the
mountain of the
{\ND{
LORD}}, and to the
house of the
God of
Jacob; and he will
teach us of his
ways, and we will
walk in his
paths: for the
law shall go
forth of
Zion, and the
word of the
{\ND{
LORD}} from
Jerusalem.
\par }{\PP \VS{3}¶ And he shall
judge among
many
people, and
rebuke
strong
nations afar
off; and they shall
beat their
swords into
plowshares, and their
spears into
pruninghooks:
nation shall not lift
up a
sword against
nation, neither shall they
learn
war any more.\FTNT{*}{{\FR 4.3 }
pruninghooks: or, scythes}
\VS{4}But they shall
sit every
man under his
vine and under his fig
tree; and none shall make
{\ADD{them}}
afraid: for the
mouth of the
{\ND{
LORD}} of
hosts hath
spoken
{\ADD{it}}.
\VS{5}For all
people will
walk every
one in the
name of his
god, and we will
walk in the
name of the
{\ND{
LORD}} our
God for
ever and
ever.
\VS{6}In that
day,
saith the
{\ND{
LORD}}, will I
assemble her that
halteth, and I will
gather her that is driven
out, and her that I have
afflicted;
\VS{7}And I will
make her that
halted a
remnant, and her that was cast far
off a
strong
nation: and the
{\ND{
LORD}} shall
reign over them in
mount
Zion from henceforth, even
for
ever.
\par }{\PP \VS{8}¶ And thou, O
tower of the
flock, the strong
hold of the
daughter of
Zion, unto thee shall it
come, even the
first
dominion; the
kingdom shall
come to the
daughter of
Jerusalem.
\VS{9}Now why dost thou
cry out
aloud?
{\ADD{is there}} no
king in thee? is thy
counsellor
perished? for
pangs have
taken thee as a woman in
travail.
\VS{10}Be in
pain, and labour to bring
forth, O
daughter of
Zion, like a woman in
travail: for now shalt thou go
forth out of the
city, and thou shalt
dwell in the
field, and thou shalt
go
{\ADD{even}} to
Babylon; there shalt thou be
delivered; there the
{\ND{
LORD}} shall
redeem thee from the
hand of thine
enemies.
\par }{\PP \VS{11}¶ Now also
many
nations are
gathered against thee, that
say, Let her be
defiled, and let our
eye
look upon
Zion.
\VS{12}But they
know not the
thoughts of the
{\ND{
LORD}}, neither
understand they his
counsel: for he shall
gather them as the
sheaves into the
floor.
\VS{13}Arise and
thresh, O
daughter of
Zion: for I will
make thine
horn
iron, and I will
make thy
hoofs
brass: and thou shalt beat in
pieces
many
people: and I will
consecrate their
gain unto the
{\ND{
LORD}}, and their
substance unto the
Lord of the whole
earth.

\par }\Chap{5}{\PP \VerseOne{1}Now
gather thyself in
troops, O
daughter of
troops: he hath
laid
siege against us: they shall
smite the
judge of
Israel with a
rod upon the
cheek.
\VS{2}But thou,
Beth-lehem
Ephratah,
{\ADD{though}} thou be
little among the
thousands of
Judah,
{\ADD{yet}} out of thee shall he come
forth unto me
{\ADD{that is}} to be
ruler in
Israel; whose goings
forth
{\ADD{have been}} from of
old, from
everlasting.\FTNT{*}{{\FR 5.2 }
everlasting: Heb. the days of eternity}
\VS{3}Therefore will he give them
up, until the
time
{\ADD{that}} she which
travaileth hath brought
forth: then the
remnant of his
brethren shall
return unto the
children of
Israel.
\par }{\PP \VS{4}¶ And he shall
stand and
feed in the
strength of the
{\ND{
LORD}}, in the
majesty of the
name of the
{\ND{
LORD}} his
God; and they shall
abide: for now shall he be
great unto the
ends of the
earth.\FTNT{†}{{\FR 5.4 }
feed: or, rule}
\VS{5}And this
{\ADD{man}} shall be the
peace, when the
Assyrian shall
come into our
land: and when he shall
tread in our
palaces, then shall we
raise against him
seven
shepherds, and
eight
principal
men.\FTNT{‡}{{\FR 5.5 }
principal: Heb. princes of}
\VS{6}And they shall
waste the
land of
Assyria with the
sword, and the
land of
Nimrod in the
entrances thereof: thus shall he
deliver
{\ADD{us}} from the
Assyrian, when he
cometh into our
land, and when he
treadeth within our
borders.\FTNT{§}{{\FR 5.6 }
waste: Heb. eat up}\FTNT{**}{{\FR 5.6 }
in the…: or, with her own naked swords}
\par }{\PP \VS{7}And the
remnant of
Jacob shall be in the
midst of
many
people as a
dew from the
{\ND{
LORD}}, as the
showers upon the
grass, that
tarrieth not for
man, nor
waiteth for the
sons of
men.
\par }{\PP \VS{8}¶ And the
remnant of
Jacob shall be among the
Gentiles in the
midst of
many
people as a
lion among the
beasts of the
forest, as a young
lion among the
flocks of
sheep: who, if he go
through, both treadeth
down, and teareth in
pieces, and none can
deliver.\FTNT{††}{{\FR 5.8 }
sheep: or, goats}
\VS{9}Thine
hand shall be lifted
up upon thine
adversaries, and all thine
enemies shall be cut
off.
\VS{10}And it shall come to pass in that
day,
saith the
{\ND{
LORD}}, that I will cut
off thy
horses out of the
midst of thee, and I will
destroy thy
chariots:
\VS{11}And I will cut
off the
cities of thy
land, and throw
down all thy strong
holds:
\VS{12}And I will cut
off
witchcrafts out of thine
hand; and thou shalt have no
{\ADD{more}}
soothsayers:
\VS{13}Thy graven
images also will I cut
off, and thy standing
images out of the
midst of thee; and thou shalt no more
worship the
work of thine
hands.\FTNT{‡‡}{{\FR 5.13 }
standing…: or, statues}
\VS{14}And I will pluck
up thy
groves out of the
midst of thee: so will I
destroy thy
cities.\FTNT{§§}{{\FR 5.14 }
cities: or, enemies}
\VS{15}And I will
execute
vengeance in
anger and
fury upon the
heathen, such as they have not
heard.

\par }\Chap{6}{\PP \VerseOne{1}Hear ye now what the
{\ND{
LORD}}
saith;
Arise,
contend thou before the
mountains, and let the
hills
hear thy
voice.\FTNT{*}{{\FR 6.1 }
before: or, with}
\VS{2}Hear ye, O
mountains, the
{\ND{
LORD’s}}
controversy, and ye
strong
foundations of the
earth: for the
{\ND{
LORD}} hath a
controversy with his
people, and he will
plead with
Israel.
\VS{3}O my
people, what have I
done unto thee? and wherein have I
wearied thee?
testify against me.
\VS{4}For I brought thee
up out of the
land of
Egypt, and
redeemed thee out of the
house of
servants; and I
sent
before thee
Moses,
Aaron, and
Miriam.
\VS{5}O my
people,
remember now what
Balak
king of
Moab
consulted, and what
Balaam the
son of
Beor
answered him from
Shittim unto
Gilgal; that ye may
know the
righteousness of the
{\ND{
LORD}}.
\par }{\PP \VS{6}¶ Wherewith shall I come
before the
{\ND{
LORD}},
{\ADD{and}}
bow myself before the
high
God? shall I come
before him with burnt
offerings, with
calves of a
year
old?\FTNT{†}{{\FR 6.6 }
of a…: Heb. sons of a year?}
\VS{7}Will the
{\ND{
LORD}} be
pleased with
thousands of
rams,
{\ADD{or}} with ten
thousands of
rivers of
oil? shall I
give my
firstborn
{\ADD{for}} my
transgression, the
fruit of my
body
{\ADD{for}} the
sin of my
soul?\FTNT{‡}{{\FR 6.7 }
body: Heb. belly}
\VS{8}He hath
shewed thee, O
man, what
{\ADD{is}}
good; and what doth the
{\ND{
LORD}}
require of thee, but to
do
justly, and to
love
mercy, and to
walk
humbly with thy
God?\FTNT{§}{{\FR 6.8 }
walk…: Heb. humble thyself to walk}
\par }{\PP \VS{9}The
{\ND{
LORD’s}}
voice
crieth unto the
city, and
{\ADD{the man of}}
wisdom shall
see thy
name:
hear ye the
rod, and who hath
appointed it.\FTNT{**}{{\FR 6.9 }
the man…: or, thy name shall see that which is}
\par }{\PP \VS{10}¶ Are
there yet the
treasures of
wickedness in the
house of the
wicked, and the
scant
measure
{\ADD{that is}}
abominable?\FTNT{††}{{\FR 6.10 }
Are…: or, Is there yet unto every man an house of the, etc}\FTNT{‡‡}{{\FR 6.10 }
scant…: Heb. measure of leanness}
\VS{11}Shall I count
{\ADD{them}}
pure with the
wicked
balances, and with the
bag of
deceitful
weights?\FTNT{§§}{{\FR 6.11 }
count…: or, be pure with, etc}
\VS{12}For the rich
men thereof are
full of
violence, and the
inhabitants thereof have
spoken
lies, and their
tongue
{\ADD{is}}
deceitful in their
mouth.
\VS{13}Therefore also will I make
{\ADD{thee}}
sick in
smiting thee, in making
{\ADD{thee}}
desolate because of thy
sins.
\VS{14}Thou shalt
eat, but not be
satisfied; and thy casting
down
{\ADD{shall be}} in the
midst of thee; and thou shalt take
hold, but shalt not
deliver; and
{\ADD{that}} which thou
deliverest will I give
up to the
sword.
\VS{15}Thou shalt
sow, but thou shalt not
reap; thou shalt
tread the
olives, but thou shalt not
anoint thee with
oil; and sweet
wine, but shalt not
drink
wine.
\par }{\PP \VS{16}¶ For the
statutes of
Omri are
kept, and all the
works of the
house of
Ahab, and ye
walk in their
counsels; that I should
make thee a
desolation, and the
inhabitants thereof an
hissing: therefore ye shall
bear the
reproach of my
people.\FTNT{***}{{\FR 6.16 }
For the…: or, For he doth much keep the, etc}\FTNT{†††}{{\FR 6.16 }
desolation: or, astonishment}

\par }\Chap{7}{\PP \VerseOne{1}Woe is me! for I am as when they have
gathered the summer
fruits, as the
grapegleanings of the
vintage:
{\ADD{there is}} no
cluster to
eat: my
soul
desired the firstripe
fruit.\FTNT{*}{{\FR 7.1 }
when…: Heb. the gatherings of summer}
\VS{2}The
good
{\ADD{man}} is
perished out of the
earth: and
{\ADD{there is}} none
upright among
men: they all lie in
wait for
blood; they
hunt every
man his
brother with a
net.\FTNT{†}{{\FR 7.2 }
good: or, godly, or, merciful}
\par }{\PP \VS{3}¶ That they may do
evil with both
hands
earnestly, the
prince
asketh, and the
judge
{\ADD{asketh}} for a
reward; and the
great
{\ADD{man}}, he
uttereth his
mischievous
desire: so they wrap it
up.\FTNT{‡}{{\FR 7.3 }
his…: Heb. the mischief of his soul}
\VS{4}The
best of them
{\ADD{is}} as a
brier: the most
upright
{\ADD{is sharper}} than a thorn
hedge: the
day of thy
watchmen
{\ADD{and}} thy
visitation
cometh; now shall be their
perplexity.
\par }{\PP \VS{5}¶
Trust ye not in a
friend, put ye not
confidence in a
guide:
keep the
doors of thy
mouth from her that
lieth in thy
bosom.
\VS{6}For the
son
dishonoureth the
father, the
daughter riseth
up against her
mother, the daughter in
law against her mother in
law; a
man’s
enemies
{\ADD{are}} the
men of his own
house.
\par }{\PP \VS{7}Therefore I will
look unto the
{\ND{
LORD}}; I will
wait for the
God of my
salvation: my
God will
hear me.
\par }{\PP \VS{8}¶
Rejoice not against me, O mine
enemy: when I
fall, I shall
arise; when I
sit in
darkness, the
{\ND{
LORD}}
{\ADD{shall be}} a
light unto me.
\VS{9}I will
bear the
indignation of the
{\ND{
LORD}}, because I have
sinned against him, until he
plead my
cause, and
execute
judgment for me: he will bring me
forth to the
light,
{\ADD{and}} I shall
behold his
righteousness.
\VS{10}Then
{\ADD{she that is}} mine
enemy shall
see
{\ADD{it}}, and
shame shall
cover her which
said unto me, Where is the
{\ND{
LORD}} thy
God? mine
eyes shall
behold her: now shall she be trodden
down as the
mire of the
streets.\FTNT{§}{{\FR 7.10 }
Then…: or, And thou wilt see her that is mine enemy, and cover her with shame}\FTNT{**}{{\FR 7.10 }
shall she…: Heb. she shall be for a treading down}
\VS{11}{\ADD{In}} the
day that thy
walls are to be
built,
{\ADD{in}} that
day shall the
decree be far
removed.
\VS{12}{\ADD{In}} that
day
{\ADD{also}} he shall
come even to thee from
Assyria, and
{\ADD{from}} the
fortified
cities, and from the
fortress even to the
river, and from
sea to
sea, and
{\ADD{from}}
mountain to
mountain.\FTNT{††}{{\FR 7.12 }
and from the fortified cities: or, even to the fortified cities}
\VS{13}Notwithstanding the
land shall be
desolate because of them that
dwell therein, for the
fruit of their
doings.\FTNT{‡‡}{{\FR 7.13 }
Notwithstanding: or, After that it hath been}
\par }{\PP \VS{14}¶
Feed thy
people with thy
rod, the
flock of thine
heritage, which
dwell
solitarily
{\ADD{in}} the
wood, in the
midst of
Carmel: let them
feed
{\ADD{in}}
Bashan and
Gilead, as in the
days of
old.\FTNT{§§}{{\FR 7.14 }
Feed: or, Rule}
\VS{15}According to the
days of thy
coming out of the
land of
Egypt will I
shew unto him
marvellous
{\ADD{things}}.
\par }{\PP \VS{16}¶ The
nations shall
see and be
confounded at all their
might: they shall
lay
{\ADD{their}}
hand upon
{\ADD{their}}
mouth, their
ears shall be
deaf.
\VS{17}They shall
lick the
dust like a
serpent, they shall
move out of their
holes like
worms of the
earth: they shall be
afraid of the
{\ND{
LORD}} our
God, and shall
fear because of thee.\FTNT{***}{{\FR 7.17 }
worms: or, creeping things}
\VS{18}Who
{\ADD{is}} a
God like unto thee, that
pardoneth
iniquity, and passeth
by the
transgression of the
remnant of his
heritage? he
retaineth not his
anger for
ever, because he
delighteth
{\ADD{in}}
mercy.
\VS{19}He will turn
again, he will have
compassion upon us; he will
subdue our
iniquities; and thou wilt
cast all their
sins into the
depths of the
sea.
\VS{20}Thou wilt
perform the
truth to
Jacob,
{\ADD{and}} the
mercy to
Abraham, which thou hast
sworn unto our
fathers from the
days of
old.
\par }