\NormalFont\ShortTitle{Mark}
{\MT THE GOSPEL ACCORDING TO ST. MARK

\par }\ChapOne{1}{\PP \VerseOne{1}The
beginning of the
gospel of
Jesus
Christ, the
Son of
God;
\VS{2}As it is
written
in the
prophets,
Behold,
I
send
my
messenger
before
thy
face,
which shall
prepare
thy
way
before
thee.
\VS{3}The
voice of one
crying
in the
wilderness, Prepare
ye the
way of the
Lord,
make
his
paths
straight.
\VS{4}John
did
baptize
in the
wilderness,
and
preach the
baptism of
repentance
for the
remission of
sins.
\VS{5}And there went
out
unto
him
all the
land of
Judæa,
and they of
Jerusalem,
and
were
all
baptized
of
him
in the
river of
Jordan,
confessing
their
sins.
\VS{6}And
John
was clothed
with
camel’s
hair,
and with a
girdle of a
skin
about
his
loins;
and he did
eat
locusts
and
wild
honey;
\VS{7}And
preached,
saying, There
cometh one mightier
than
I
after
me, the
latchet of
whose
shoes I
am
not
worthy
to stoop
down and
unloose.
\VS{8}I
indeed have
baptized
you
with
water:
but
he shall
baptize
you
with
the
Holy
Ghost.
\VS{9}And it came to
pass
in
those
days, that
Jesus
came
from
Nazareth of
Galilee,
and was
baptized
of
John
in
Jordan.
\VS{10}And
straightway coming
up out
of the
water, he
saw the
heavens
opened,
and the
Spirit
like a
dove
descending
upon
him:
\VS{11}And there
came a
voice
from
heaven,
{\ADD{saying}},
Thou
art
my
beloved
Son,
in
whom I am well
pleased.
\VS{12}And
immediately the
Spirit
driveth
him
into the
wilderness.
\VS{13}And he
was
there
in the
wilderness
forty
days,
tempted
of
Satan;
and
was
with the wild
beasts;
and the
angels
ministered unto
him.
\VS{14}Now
after that
John was put in
prison,
Jesus
came
into
Galilee,
preaching the
gospel of the
kingdom of
God,
\VS{15}And
saying,
{\WJ{The time
is fulfilled,
and
the kingdom
of God
is at hand:
repent ye,
and
believe
the gospel.}}
\VS{16}Now as he
walked
by the
sea of
Galilee, he
saw
Simon
and
Andrew
his
brother
casting a
net
into the
sea:
for they
were
fishers.
\VS{17}And
Jesus
said unto
them,
{\WJ{Come ye
after
me,
and
I will make
you
to become
fishers
of men.}}
\VS{18}And
straightway they
forsook
their
nets, and
followed
him.
\VS{19}And when he had
gone a little
further
thence, he
saw
James the
{\ADD{son}} of
Zebedee,
and
John
his
brother,
who
also
were
in the
ship
mending their
nets.
\VS{20}And
straightway he
called
them:
and they
left
their
father
Zebedee
in the
ship
with the hired
servants, and
went
after
him.
\VS{21}And they
went
into
Capernaum;
and
straightway on the sabbath
day he
entered
into the
synagogue, and
taught.
\VS{22}And they were
astonished
at
his
doctrine:
for
he
taught
them
as one that
had
authority,
and
not
as the
scribes.
\VS{23}And there
was
in
their
synagogue a
man
with an
unclean
spirit;
and he cried
out,
\VS{24}Saying,
Let
{\ADD{us}}
alone; what
have we to
do
with
thee, thou
Jesus of
Nazareth? art thou
come to
destroy
us? I
know
thee
who thou
art, the Holy
One of
God.
\VS{25}And
Jesus
rebuked
him,
saying,
{\WJ{Hold thy peace,
and
come out
of
him.}}
\VS{26}And
when the
unclean
spirit had
torn
him,
and
cried with a
loud
voice, he came
out
of
him.
\VS{27}And they
were
all
amazed, insomuch
that they
questioned
among
themselves,
saying,
What thing
is
this?
what
new
doctrine
{\ADD{is}}
this?
for
with
authority commandeth
he
even the
unclean
spirits,
and they do
obey
him.
\VS{28}And
immediately
his
fame spread
abroad
throughout
all the region round
about
Galilee.
\VS{29}And
forthwith, when they were come
out
of the
synagogue, they
entered
into the
house of
Simon
and
Andrew,
with
James
and
John.
\VS{30}But
Simon’s
wife’s mother
lay sick of a
fever,
and
anon they
tell
him
of
her.
\VS{31}And he
came and
took
her by the
hand, and
lifted
her
up;
and
immediately the
fever
left
her,
and she
ministered unto
them.
\VS{32}And at
even,
when the
sun
did
set, they
brought
unto
him
all that
were
diseased,
and
them that were possessed with
devils.
\VS{33}And
all the
city
was gathered
together
at the
door.
\VS{34}And he
healed
many that
were
sick of
divers
diseases,
and cast
out
many
devils;
and
suffered
not the
devils to
speak,
because they
knew
him.
\VS{35}And in the
morning, rising
up a great
while before
day, he went
out,
and
departed
into a
solitary
place, and
there
prayed.
\VS{36}And
Simon
and they that
were
with
him followed
after
him.
\VS{37}And when they had
found
him, they
said unto
him,
All
{\ADD{men}} seek
for
thee.
\VS{38}And he
said unto
them,
{\WJ{Let us go
into
the next
towns,
that
I may preach
there also:
for
therefore
came I forth.}}
\VS{39}And
he
preached
in
their
synagogues
throughout
all
Galilee,
and cast
out
devils.
\VS{40}And there
came a
leper
to
him,
beseeching
him,
and kneeling
down to
him,
and
saying unto
him,
If thou
wilt, thou canst
make
me
clean.
\VS{41}And
Jesus, moved with
compassion, put
forth
{\ADD{his}}
hand, and
touched
him,
and
saith unto
him,
{\WJ{I will;
be thou clean.}}
\VS{42}And as soon
as
he had
spoken,
immediately the
leprosy
departed
from
him,
and he was
cleansed.
\VS{43}And he straitly
charged
him,
and
forthwith
sent
him
away;
\VS{44}And
saith unto
him,
{\WJ{See
thou say
nothing
to any man:
but
go thy way,
shew
thyself
to the priest,
and
offer
for
thy
cleansing
those things which
Moses
commanded,
for
a testimony
unto them.}}
\VS{45}But he went
out, and
began to
publish
{\ADD{it}}
much,
and to blaze
abroad the
matter, insomuch
that
Jesus
could no
more
openly
enter
into the
city,
but
was
without
in
desert
places:
and they
came
to
him from every
quarter.

\par }\Chap{2}{\PP \VerseOne{1}And
again he
entered
into
Capernaum
after
{\ADD{some}}
days;
and it was
noised
that he
was
in the
house.
\VS{2}And
straightway
many were gathered
together, insomuch
that there was no
room to
receive
{\ADD{them}}, no, not so much
as
about the
door:
and he
preached the
word unto
them.
\VS{3}And they
come
unto
him,
bringing one sick of the
palsy, which was
borne
of
four.
\VS{4}And when they
could
not come nigh
unto
him
for the
press, they
uncovered the
roof
where he
was:
and when they had
broken
{\ADD{it}}
up, they let
down the
bed
wherein the sick of the
palsy
lay.
\VS{5}When
Jesus
saw
their
faith, he
said unto the sick of the
palsy,
{\WJ{Son,
thy
sins
be forgiven
thee.}}
\VS{6}But there
were
certain of the
scribes
sitting
there,
and
reasoning
in
their
hearts,
\VS{7}Why
doth
this
{\ADD{man}}
thus
speak
blasphemies?
who
can
forgive
sins
but
God
only?
\VS{8}And
immediately
when
Jesus
perceived
in
his
spirit
that
they
so
reasoned
within
themselves, he
said unto
them,
{\WJ{Why
reason ye
these things
in
your
hearts?}}
\VS{9}{\WJ{Whether
is it
easier
to say
to the sick of the palsy,
{\ADD{Thy}}
sins
be forgiven
thee;
or
to say,
Arise,
and
take up
thy
bed,
and
walk?}}
\VS{10}{\WJ{But
that
ye may know
that
the Son
of man
hath
power
on
earth
to forgive
sins,}} (he
saith to the sick of the
palsy,)
\VS{11}{\WJ{I say
unto thee,
Arise,
and
take up
thy
bed,
and
go thy way
into
thine
house.}}
\VS{12}And
immediately he
arose, took
up the
bed,
and went
forth
before them
all; insomuch
that they were
all
amazed,
and
glorified
God,
saying,
We
never
saw it on this
fashion.
\VS{13}And he went
forth
again
by the sea
side;
and
all the
multitude
resorted
unto
him,
and he
taught
them.
\VS{14}And as he passed
by, he
saw
Levi
the
{\ADD{son}} of
Alphæus
sitting
at the receipt of
custom,
and
said unto
him,
{\WJ{Follow
me.}}
And
he arose
and
followed
him.
\VS{15}And it came to
pass, that,
as
Jesus sat at
meat
in
his
house,
many
publicans
and
sinners
sat
also
together with
Jesus
and
his
disciples:
for there
were
many,
and they
followed
him.
\VS{16}And when the
scribes
and
Pharisees
saw
him
eat
with
publicans
and
sinners, they
said
unto
his
disciples, How is
it
that he
eateth
and
drinketh
with
publicans
and
sinners?
\VS{17}When
Jesus
heard
{\ADD{it}}, he
saith unto
them,
{\WJ{They that are whole
have
no
need
of the physician,
but
they that are
sick:
I came
not
to call
the righteous,
but
sinners
to
repentance.}}
\VS{18}And the
disciples of
John
and of the
Pharisees
used to
fast:
and they
come
and
say unto
him,
Why do the
disciples of
John
and of the
Pharisees
fast,
but
thy
disciples
fast
not?
\VS{19}And
Jesus
said unto
them,
{\WJ{Can
the children
of the bridechamber
fast,
while
the bridegroom
is
with
them?
as long as
they have
the bridegroom
with
them,
they cannot
fast.}}
\VS{20}{\WJ{But
the days
will come,
when
the bridegroom
shall be taken away
from
them,
and
then
shall they fast
in
those
days.}}
\VS{21}{\WJ{No man
also
seweth
a piece
of new
cloth
on
an
old
garment:
else
the new piece
that filled
it
up
taketh away from
the old,
and
the rent
is made
worse.}}
\VS{22}{\WJ{And
no man
putteth
new
wine
into
old
bottles:
else
the
new
wine
doth burst
the bottles,
and
the wine
is spilled,
and
the bottles
will be marred:
but
new
wine
must be put
into
new
bottles.}}
\VS{23}And it came to
pass,
that
he
went
through the corn
fields
on the sabbath
day;
and
his
disciples
began, as they
went, to
pluck the ears of
corn.
\VS{24}And the
Pharisees
said unto
him,
Behold,
why do
they
on the sabbath
day that
which
is
not
lawful?
\VS{25}And
he
said unto
them,
{\WJ{Have ye
never
read
what
David
did,
when
he had
need,
and
was an hungred,
he,
and
they that were
with
him?}}
\VS{26}{\WJ{How
he went
into
the house
of God
in the days of
Abiathar
the high priest,
and
did eat
the shewbread,
which
is
not
lawful
to eat
but
for the priests,
and
gave
also
to them which
were
with
him?}}
\VS{27}And he
said unto
them,
{\WJ{The sabbath
was made
for
man,
and not
man
for
the sabbath:}}
\VS{28}{\WJ{Therefore
the Son
of man
is
Lord
also
of the sabbath.}}

\par }\Chap{3}{\PP \VerseOne{1}And he
entered
again
into the
synagogue;
and there
was a
man
there which
had a
withered
hand.
\VS{2}And they
watched
him,
whether he would
heal
him on the sabbath
day;
that they might
accuse
him.
\VS{3}And he
saith unto the
man which
had the
withered
hand,
{\WJ{Stand
forth.}}
\VS{4}And he
saith unto
them,
{\WJ{Is it lawful
to do good
on the sabbath days,
or
to do evil?
to save
life,
or
to kill?}}
But they held their
peace.
\VS{5}And when he had looked round
about on
them
with
anger, being
grieved
for the
hardness
of
their
hearts, he
saith unto the
man,
{\WJ{Stretch forth
thine
hand.}}
And he
stretched
{\ADD{it}}
out:
and
his
hand was
restored
whole
as the
other.
\VS{6}And the
Pharisees went
forth, and
straightway
took
counsel
with the
Herodians
against
him,
how they might
destroy
him.
\VS{7}But
Jesus withdrew
himself
with
his
disciples
to the
sea:
and a
great
multitude
from
Galilee
followed
him,
and
from
Judæa,
\VS{8}And
from
Jerusalem,
and
from
Idumaea,
and
{\ADD{from}}
beyond
Jordan;
and
they
about
Tyre
and
Sidon, a
great
multitude, when they had
heard what great
things he
did,
came
unto
him.
\VS{9}And he
spake
to
his
disciples,
that a small
ship should
wait on
him because
of the
multitude,
lest they should
throng
him.
\VS{10}For he had
healed
many; insomuch
that they pressed
upon
him for
to
touch
him, as many
as
had
plagues.
\VS{11}And
unclean
spirits,
when they
saw
him, fell down
before
him,
and
cried,
saying,
Thou
art the
Son of
God.
\VS{12}And
he
straitly
charged
them
that they
should
not
make
him
known.
\VS{13}And he goeth
up
into a
mountain,
and
calleth
{\ADD{unto him}}
whom
he
would:
and they
came
unto
him.
\VS{14}And he
ordained
twelve,
that they should
be
with
him,
and
that he might
send
them
forth to
preach,
\VS{15}And to
have
power to
heal
sicknesses,
and to cast
out
devils:
\VS{16}And
Simon
he
surnamed Peter;
\VS{17}And
James
the
{\ADD{son}} of
Zebedee,
and
John the
brother of
James;
and
he
surnamed
them
Boanerges,
which
is, The
sons of
thunder:
\VS{18}And
Andrew,
and
Philip,
and
Bartholomew,
and
Matthew,
and
Thomas,
and
James
the
{\ADD{son}} of
Alphaeus,
and
Thaddaeus,
and
Simon the
Canaanite,
\VS{19}And
Judas
Iscariot,
which
also
betrayed
him:
and they
went
into an
house.
\VS{20}And the
multitude cometh
together
again, so
that
they
could
not so much
as
eat
bread.
\VS{21}And
when
his
friends
heard
{\ADD{of it}}, they went
out to lay
hold on
him:
for they
said, He is beside
himself.
\par }{\PP \VS{22}¶
And the
scribes
which came
down
from
Jerusalem
said, He
hath
Beelzebub,
and
by the
prince of the
devils casteth he
out
devils.
\VS{23}And he
called
them
{\ADD{unto him}}, and
said unto
them
in
parables,
{\WJ{How
can
Satan
cast out
Satan?}}
\VS{24}{\WJ{And
if
a kingdom
be divided
against
itself,
that
kingdom
cannot
stand.}}
\VS{25}{\WJ{And
if
a house
be divided
against
itself,
that
house
cannot
stand.}}
\VS{26}{\WJ{And
if
Satan
rise up
against
himself,
and
be divided,
he
cannot
stand,
but
hath
an end.}}
\VS{27}{\WJ{No
man
can
enter
into
a strong man’s
house,
and spoil
his
goods,
except
he will
first
bind
the strong man;
and
then
he will spoil
his
house.}}
\VS{28}{\WJ{Verily
I say
unto you,
All
sins
shall be forgiven
unto the sons
of men,
and
blasphemies
wherewith
soever
they shall blaspheme:}}
\VS{29}{\WJ{But
he
that
shall blaspheme
against
the Holy
Ghost
hath
never
forgiveness,
but
is
in danger
of eternal
damnation:}}
\VS{30}Because they
said, He
hath an
unclean
spirit.
\par }{\PP \VS{31}¶ There
came
then
his
brethren
and
his
mother,
and,
standing
without,
sent
unto
him,
calling
him.
\VS{32}And the
multitude
sat
about
him,
and they
said unto
him,
Behold,
thy
mother
and
thy
brethren
without
seek for
thee.
\VS{33}And he
answered
them,
saying,
{\WJ{Who
is
my
mother,
or
my
brethren?}}
\VS{34}And he
looked round
about on them which
sat
about
him, and
said,
{\WJ{Behold
my
mother
and
my
brethren!}}
\VS{35}{\WJ{For
whosoever
shall do
the will
of God,
the same
is
my
brother,
and
my
sister,
and
mother.}}

\par }\Chap{4}{\PP \VerseOne{1}And he
began
again to
teach
by the sea
side:
and there was
gathered
unto
him
a
great
multitude, so
that
he
entered
into a
ship, and
sat
in the
sea;
and
the
whole
multitude
was
by the
sea
on the
land.
\VS{2}And he
taught
them many
things
by
parables,
and
said unto
them
in
his
doctrine,
\VS{3}{\WJ{Hearken;
Behold,
there went out
a sower
to sow:}}
\VS{4}{\WJ{And
it came to pass,
as
he sowed,
some
fell
by
the way side,
and
the fowls
of the air
came
and devoured
it
up.}}
\VS{5}{\WJ{And
some
fell
on
stony ground,
where
it had
not
much
earth;
and
immediately
it sprang up,
because
it had
no
depth
of earth:}}
\VS{6}{\WJ{But
when
the sun
was up,
it was scorched;
and
because
it had
no
root,
it withered away.}}
\VS{7}{\WJ{And
some
fell
among
thorns,
and
the thorns
grew up,
and
choked
it,
and
it yielded
no
fruit.}}
\VS{8}{\WJ{And
other
fell
on
good
ground,
and
did yield
fruit
that sprang up
and
increased;
and
brought forth,
some
thirty,
and
some
sixty,
and
some
an hundred.}}
\VS{9}And he
said unto
them,
{\WJ{He that hath
ears
to hear,
let him hear.}}
\VS{10}And
when he
was
alone, they that were
about
him
with the
twelve
asked of
him the
parable.
\VS{11}And he
said unto
them,
{\WJ{Unto you
it is given
to know
the mystery
of the kingdom
of God:
but
unto them
that are without,
all
{\ADD{these}}
things
are done
in
parables:}}
\VS{12}{\WJ{That
seeing
they may see,
and
not
perceive;
and
hearing
they may hear,
and
not
understand;
lest at any time
they should be converted,
and
{\ADD{their}}
sins
should be forgiven
them.}}
\VS{13}And he
said unto
them,
{\WJ{Know ye
not
this
parable?
and
how then
will ye know
all
parables?}}
\par }{\PP \VS{14}{\WJ{¶
The sower
soweth
the word.}}
\VS{15}{\WJ{And
these
are they
by
the way side,
where
the word
is sown;
but
when
they have heard,
Satan
cometh
immediately,
and
taketh away
the word
that was sown
in
their
hearts.}}
\VS{16}{\WJ{And
these
are they
likewise
which are sown
on
stony ground;
who,
when
they have heard
the word,
immediately
receive
it
with
gladness;}}
\VS{17}{\WJ{And
have
no
root
in
themselves,
and so
endure
but
for a time:
afterward,
when
affliction
or
persecution
ariseth
for
the word’s sake,
immediately
they are offended.}}
\VS{18}{\WJ{And
these
are they
which are sown
among
thorns;
such as hear
the word,}}
\VS{19}{\WJ{And
the cares
of this
world,
and
the deceitfulness
of riches,
and
the lusts
of
other things
entering in,
choke
the word,
and
it becometh
unfruitful.}}
\VS{20}{\WJ{And
these
are
they which are sown
on
good
ground;
such as
hear
the word,
and
receive
{\ADD{it}},
and
bring forth fruit,
some
thirtyfold,
some
sixty,
and
some
an hundred.}}
\par }{\PP \VS{21}¶
And he
said unto
them,
{\WJ{Is
a candle
brought
to
be put
under a
bushel,
or
under
a bed?
and not
to
be set
on
a candlestick?}}
\VS{22}{\WJ{For
there is
nothing
hid,
which
shall
not
be manifested;
neither
was any thing kept
secret,
but
that
it should come
abroad.}}
\VS{23}{\WJ{If
any man
have
ears
to hear,
let him hear.}}
\VS{24}And he
said unto
them,
{\WJ{Take heed
what
ye hear:
with
what
measure
ye mete,
it shall be measured
to you:
and
unto you
that hear
shall more be given.}}
\VS{25}{\WJ{For
he that
hath,
to him
shall be given:
and
he that
hath
not,
from
him
shall be taken
even
that which
he hath.}}
\par }{\PP \VS{26}¶
And he
said,
{\WJ{So
is
the kingdom
of God,
as
if
a man
should cast
seed
into
the ground;}}
\VS{27}{\WJ{And
should sleep,
and
rise
night
and
day,
and
the seed
should spring
and
grow up,
he
knoweth
not
how.}}
\VS{28}{\WJ{For
the earth
bringeth forth fruit
of herself;
first
the blade,
then
the ear,
after that
the
full
corn
in
the ear.}}
\VS{29}{\WJ{But
when
the fruit
is brought forth,
immediately
he putteth in
the sickle,
because
the harvest
is come.}}
\par }{\PP \VS{30}¶
And he
said,
{\WJ{Whereunto
shall we liken
the kingdom
of God?
or
with
what
comparison
shall we compare
it?}}
\VS{31}{\WJ{{\ADD{It is}}
like
a grain
of mustard seed,
which,
when
it is sown
in
the earth,
is
less than
all
the seeds
that be
in
the earth:}}
\VS{32}{\WJ{But
when
it is sown,
it groweth up,
and
becometh
greater than
all
herbs,
and
shooteth out
great
branches;
so that
the fowls
of the air
may
lodge
under
the shadow
of it.}}
\VS{33}And
with
many
such
parables spake
he the
word unto
them,
as they were
able to
hear
{\ADD{it}}.
\VS{34}But
without a
parable spake
he
not unto
them:
and
when they were
alone, he
expounded all
things
to
his
disciples.
\VS{35}And
the
same
day,
when the
even was
come, he
saith unto
them,
{\WJ{Let us pass over
unto
the other side.}}
\VS{36}And when they had sent
away the
multitude, they
took
him even
as he
was
in the
ship.
And there
were
also
with
him
other little
ships.
\VS{37}And there
arose
a
great
storm of
wind,
and the
waves
beat
into the
ship, so
that
it
was
now
full.
\VS{38}And
he
was
in the hinder part of the
ship,
asleep
on a
pillow:
and they
awake
him,
and
say unto
him,
Master,
carest
thou
not
that we
perish?
\VS{39}And he
arose,
and
rebuked the
wind,
and
said unto the
sea,
{\WJ{Peace,
be still.}}
And the
wind
ceased,
and there
was
a
great
calm.
\VS{40}And he
said unto
them,
{\WJ{Why
are ye
so
fearful?
how is it that
ye have
no
faith?}}
\VS{41}And they
feared
exceedingly,
and
said
one
to
another,
What
manner of man
is
this,
that
even the
wind
and the
sea
obey
him?

\par }\Chap{5}{\PP \VerseOne{1}And they
came over
unto the other
side of the
sea,
into the
country of the
Gadarenes.
\VS{2}And
when
he was come
out
of the
ship,
immediately there
met
him
out of the
tombs a
man
with
an
unclean
spirit,
\VS{3}Who
had
{\ADD{his}}
dwelling
among the
tombs;
and no
man
could
bind
him, no,
not with
chains:
\VS{4}Because
that
he had
been
often
bound with
fetters
and
chains,
and the
chains had been plucked
asunder
by
him,
and the
fetters broken in
pieces:
neither
could
any
{\ADD{man}}
tame
him.
\VS{5}And
always,
night
and
day, he
was
in the
mountains,
and
in the
tombs,
crying,
and
cutting
himself with
stones.
\VS{6}But when he
saw
Jesus
afar
off, he
ran
and
worshipped
him,
\VS{7}And
cried with
a
loud
voice, and
said,
What
have
I to do
with
thee,
Jesus,
{\ADD{thou}}
Son of
the most
high
God? I
adjure
thee by
God, that thou
torment
me
not.
\VS{8}For he
said unto
him,
{\WJ{Come out
of
the man,
{\ADD{thou}}
unclean
spirit.}}
\VS{9}And he
asked
him,
{\WJ{What
{\ADD{is}}
thy
name?}}
And he
answered,
saying,
My
name
{\ADD{is}}
Legion:
for we
are
many.
\VS{10}And he
besought
him
much
that he
would
not
send
them away out
of the
country.
\VS{11}Now there
was
there
nigh unto the
mountains
a
great
herd of
swine
feeding.
\VS{12}And
all the
devils
besought
him,
saying,
Send
us
into the
swine,
that we may
enter
into
them.
\VS{13}And
forthwith
Jesus
gave
them
leave.
And
the
unclean
spirits went
out, and
entered
into the
swine:
and the
herd ran
violently
down a steep
place
into the
sea, (they
were
about two
thousand;)
and were
choked
in the
sea.
\VS{14}And they that
fed the
swine
fled,
and
told
{\ADD{it}}
in the
city,
and
in the
country.
And they went
out to
see
what it
was that was
done.
\VS{15}And they
come
to
Jesus,
and
see him that was possessed with the
devil, and
had the
legion,
sitting,
and
clothed,
and in his right
mind:
and they were
afraid.
\VS{16}And they that
saw
{\ADD{it}}
told
them
how it befell to him that was possessed with the
devil,
and
{\ADD{also}}
concerning the
swine.
\VS{17}And they
began to
pray
him to
depart
out
of
their
coasts.
\VS{18}And
when
he was
come
into the
ship, he that had been possessed with the
devil
prayed
him
that he might
be
with
him.
\VS{19}Howbeit
Jesus
suffered
him
not,
but
saith unto
him,
{\WJ{Go
home
to
thy friends,
and
tell
them
how great things
the Lord
hath done
for thee,
and
hath had compassion on
thee.}}
\VS{20}And he
departed,
and
began to
publish
in
Decapolis how great
things
Jesus had
done for
him:
and
all
{\ADD{men}} did
marvel.
\VS{21}And
when
Jesus was passed
over
again
by
ship
unto the other
side,
much
people
gathered
unto
him:
and he
was nigh
unto the
sea.
\VS{22}And,
behold, there
cometh
one of the rulers of the
synagogue,
Jairus by
name;
and when he
saw
him, he
fell
at
his
feet,
\VS{23}And
besought
him
greatly,
saying,
My little
daughter
lieth at the point of
death:
{\ADD{I pray thee}},
come and
lay thy
hands on
her,
that she may be
healed;
and she shall
live.
\VS{24}And
{\ADD{Jesus}}
went
with
him;
and
much
people
followed
him,
and
thronged
him.
\VS{25}And
a
certain
woman, which
had an
issue of
blood
twelve
years,
\VS{26}And had
suffered many
things
of
many
physicians,
and had
spent
all
that
she
had,
and
was
nothing
bettered,
but
rather
grew
worse,
\VS{27}When she had
heard
of
Jesus,
came
in the
press
behind, and
touched
his
garment.
\VS{28}For she
said,
If I may
touch
but
his
clothes, I shall be
whole.
\VS{29}And
straightway the
fountain
of
her
blood was dried
up;
and she
felt in her
body
that she was
healed
of that
plague.
\VS{30}And
Jesus,
immediately
knowing
in
himself
that
virtue had gone
out
of
him, turned him
about
in the
press, and
said,
{\WJ{Who
touched
my
clothes?}}
\VS{31}And
his
disciples
said unto
him, Thou
seest the
multitude
thronging
thee,
and sayest
thou,
Who
touched
me?
\VS{32}And he looked round
about to
see her that had
done this
thing.
\VS{33}But the
woman
fearing
and
trembling,
knowing
what was
done
in
her,
came
and fell down
before
him,
and
told
him
all the
truth.
\VS{34}And he
said unto
her,
{\WJ{Daughter,
thy
faith
hath made
thee
whole;
go
in
peace,
and
be
whole
of
thy
plague.}}
\VS{35}While
he
yet
spake, there
came
from the ruler of the
synagogue’s
{\ADD{house certain}} which
said,
Thy
daughter is
dead:
why troublest
thou the
Master any
further?
\VS{36}As soon
as
Jesus
heard the
word that was
spoken, he
saith unto the ruler of the
synagogue,
{\WJ{Be
not
afraid,
only
believe.}}
\VS{37}And he
suffered no
man to
follow
him,
save
Peter,
and
James,
and
John the
brother of
James.
\VS{38}And he
cometh
to the
house of the ruler of the
synagogue,
and
seeth the
tumult, and them that
wept
and
wailed
greatly.
\VS{39}And when he was come
in, he
saith unto
them,
{\WJ{Why
make ye this ado,
and
weep?
the damsel
is
not
dead,
but
sleepeth.}}
\VS{40}And they
laughed
him to
scorn.
But when he had put
them
all
out, he
taketh the
father
and the
mother of the
damsel,
and them that were
with
him,
and entereth
in
where the
damsel
was
lying.
\VS{41}And he
took the
damsel by the
hand, and
said unto
her,
{\WJ{Talitha
cumi;}}
which
is, being
interpreted,
Damsel, I
say unto
thee,
arise.
\VS{42}And
straightway the
damsel
arose,
and
walked;
for she
was
{\ADD{of the age}}
of
twelve
years.
And they were
astonished
with a
great
astonishment.
\VS{43}And he
charged
them
straitly
that no
man should
know
it;
and
commanded that something should be
given
her to
eat.

\par }\Chap{6}{\PP \VerseOne{1}And he went
out from
thence,
and
came
into his
own
country;
and
his
disciples
follow
him.
\VS{2}And when the sabbath
day was
come, he
began to
teach
in the
synagogue:
and
many
hearing
{\ADD{him}} were
astonished,
saying, From
whence hath
this
{\ADD{man}} these
things?
and
what
wisdom
{\ADD{is}} this
which is
given unto
him,
that
even
such mighty
works are
wrought
by
his
hands?
\VS{3}Is
not
this the
carpenter, the
son of
Mary, the
brother of
James,
and
Joses,
and of
Juda,
and
Simon?
and
are
not
his
sisters
here
with
us?
And they were
offended
at
him.
\VS{4}But
Jesus
said unto
them,
{\WJ{A prophet
is
not
without honour,
but
in
his own
country,
and
among
his own kin,
and
in
his own
house.}}
\VS{5}And he
could
there
do
no mighty
work, save
that he
laid his
hands upon a
few sick
folk, and
healed
{\ADD{them}}.
\VS{6}And he
marvelled
because of
their
unbelief.
And he
went round
about the
villages,
teaching.
\par }{\PP \VS{7}¶
And he
called
{\ADD{unto him}} the
twelve,
and
began to
send
them
forth by
two and
two;
and
gave
them
power over
unclean
spirits;
\VS{8}And
commanded
them
that they should
take
nothing
for
{\ADD{their}}
journey,
save a
staff
only;
no
scrip,
no
bread,
no
money
in
{\ADD{their}}
purse:
\VS{9}But
{\ADD{be}}
shod with
sandals;
and
not put
on
two
coats.
\VS{10}And he
said unto
them,
{\WJ{In what place soever
ye enter
into
an house,
there
abide
till
ye depart
from that place.}}
\VS{11}{\WJ{And
whosoever
shall
not
receive
you,
nor
hear
you,
when ye depart
thence,
shake off
the dust
under
your
feet
for
a testimony
against them.
Verily
I say
unto you,
It shall be
more tolerable
for Sodom
and
Gomorrha
in
the day
of judgment,
than
for that
city.}}
\VS{12}And they went
out, and
preached
that men should
repent.
\VS{13}And they cast
out
many
devils,
and
anointed with
oil
many that were
sick,
and
healed
{\ADD{them}}.
\VS{14}And
king
Herod
heard
{\ADD{of him}}; (for
his
name
was spread
abroad:)
and he
said,
That
John the
Baptist was
risen
from the
dead,
and
therefore
mighty works do shew forth
themselves
in
him.
\VS{15}Others
said,
That it
is
Elias.
And
others
said,
That it
is a
prophet,
or
as
one of the
prophets.
\VS{16}But when
Herod
heard
{\ADD{thereof}}, he
said,
It
is
John,
whom
I
beheaded:
he is
risen
from the
dead.
\VS{17}For
Herod
himself had sent
forth and laid
hold upon
John,
and
bound
him
in
prison
for
Herodias’ sake,
his
brother
Philip’s
wife:
for he had
married
her.
\VS{18}For
John had
said unto
Herod, It
is
not
lawful for
thee to
have
thy
brother’s
wife.
\VS{19}Therefore
Herodias had a
quarrel against
him,
and
would have
killed
him;
but she
could
not:
\VS{20}For
Herod
feared
John,
knowing that
he was a
just
man
and an
holy,
and
observed
him;
and when he
heard
him, he
did many
things,
and
heard
him
gladly.
\VS{21}And
when a
convenient
day was
come, that
Herod on
his
birthday
made a
supper to
his
lords, high
captains,
and
chief
{\ADD{estates}} of
Galilee;
\VS{22}And when the
daughter of the
said
Herodias came
in,
and
danced,
and
pleased
Herod
and
them that sat with
him, the
king
said unto the
damsel,
Ask of
me
whatsoever thou
wilt,
and I will
give
{\ADD{it}}
thee.
\VS{23}And he
sware unto
her,
Whatsoever thou shalt
ask of
me, I will
give
{\ADD{it}}
thee,
unto the
half of
my
kingdom.
\VS{24}And she went
forth, and
said unto
her
mother,
What shall I
ask?
And she
said, The
head of
John the
Baptist.
\VS{25}And she came
in
straightway
with
haste
unto the
king, and
asked,
saying, I
will
that thou
give
me by and
by
in a
charger the
head of
John the
Baptist.
\VS{26}And the
king
was exceeding
sorry;
{\ADD{yet}}
for his
oath’s sake,
and for their sakes
which sat with
him, he
would
not
reject
her.
\VS{27}And
immediately the
king
sent an
executioner, and
commanded
his
head to be
brought:
and he
went and
beheaded
him
in the
prison,
\VS{28}And
brought
his
head
in a
charger,
and
gave
it to the
damsel:
and the
damsel
gave
it to
her
mother.
\VS{29}And when
his
disciples
heard
{\ADD{of it}}, they
came
and took
up
his
corpse,
and
laid
it
in a
tomb.
\VS{30}And the
apostles gathered themselves
together
unto
Jesus,
and
told
him all
things,
both
what they had
done,
and
what they had
taught.
\VS{31}And he
said unto
them,
{\WJ{Come
ye
yourselves
apart
into
a desert
place,
and
rest
a while:}}
for there
were
many
coming
and
going,
and they
had
no
leisure so much as to
eat.
\VS{32}And they
departed
into a
desert
place by
ship
privately.
\VS{33}And the
people
saw
them
departing,
and
many
knew
him,
and
ran
afoot
thither
out of
all
cities,
and
outwent
them,
and came
together
unto
him.
\VS{34}And
Jesus, when he came
out,
saw
much
people,
and was moved with
compassion
toward
them,
because they
were
as
sheep
not
having a
shepherd:
and he
began to
teach
them many
things.
\VS{35}And when the
day
was
now far
spent,
his
disciples
came unto
him, and
said, This
is a
desert
place,
and
now the
time
{\ADD{is}} far
passed:
\VS{36}Send
them
away,
that they may
go
into the
country round
about,
and into the
villages, and
buy
themselves
bread:
for they
have
nothing to
eat.
\VS{37}He
answered and
said unto
them,
{\WJ{Give
ye
them
to eat.}}
And they
say unto
him, Shall we
go and
buy two
hundred
pennyworth of
bread,
and
give
them to
eat?
\VS{38}He
saith unto
them,
{\WJ{How many
loaves
have ye?
go
and
see.}}
And when they
knew, they
say,
Five,
and
two
fishes.
\VS{39}And he
commanded
them to
make
all sit
down by
companies
upon the
green
grass.
\VS{40}And they sat
down in
ranks,
by
hundreds,
and
by
fifties.
\VS{41}And when he had
taken the
five
loaves
and the
two
fishes, he
looked up
to
heaven,
and
blessed, and
brake the
loaves,
and
gave
{\ADD{them}} to
his
disciples
to set
before
them;
and the
two
fishes divided he
among them
all.
\VS{42}And they
did
all
eat,
and were
filled.
\VS{43}And they took
up
twelve
baskets
full of the
fragments,
and
of the
fishes.
\VS{44}And they
that did
eat of the
loaves
were
about five
thousand
men.
\VS{45}And
straightway he
constrained
his
disciples to
get
into the
ship,
and to
go
to the other
side before
unto
Bethsaida,
while
he sent
away the
people.
\VS{46}And when he had
sent
them
away, he
departed
into a
mountain to
pray.
\VS{47}And when
even was
come, the
ship
was
in the
midst of the
sea,
and
he
alone
on the
land.
\VS{48}And he
saw
them
toiling
in
rowing;
for the
wind
was
contrary unto
them:
and
about the
fourth
watch of the
night he
cometh
unto
them,
walking
upon the
sea,
and would
have passed
by
them.
\VS{49}But when they
saw
him
walking
upon the
sea, they
supposed it had
been a
spirit,
and cried
out:
\VS{50}For
they
all
saw
him,
and were
troubled.
And
immediately he
talked
with
them,
and
saith unto
them,
{\WJ{Be of good cheer:
it is
I;
be
not
afraid.}}
\VS{51}And he went
up
unto
them
into the
ship;
and the
wind
ceased:
and they were
sore
amazed
in
themselves
beyond
measure,
and
wondered.
\VS{52}For they
considered
not
{\ADD{the miracle}}
of the
loaves:
for
their
heart
was
hardened.
\VS{53}And when they had passed
over, they
came
into the
land of
Gennesaret,
and drew to the
shore.
\VS{54}And when
they were come
out
of the
ship,
straightway they
knew
him,
\VS{55}And ran
through
that
whole region round
about, and
began to carry
about
in
beds those
that
were
sick,
where they
heard he
was.
\VS{56}And
whithersoever he
entered,
into
villages,
or
cities,
or
country, they
laid the
sick
in the
streets,
and
besought
him
that they might
touch if it were
but the
border of
his
garment:
and as many
as
touched
him were made
whole.

\par }\Chap{7}{\PP \VerseOne{1}Then came
together
unto
him the
Pharisees,
and
certain of the
scribes, which
came
from
Jerusalem.
\VS{2}And when they
saw
some
of
his
disciples
eat
bread
with
defiled,
that is to
say, with
unwashen,
hands, they found
fault.
\VS{3}For the
Pharisees,
and
all the
Jews,
except they
wash
{\ADD{their}}
hands
oft,
eat
not,
holding the
tradition of the
elders.
\VS{4}And
{\ADD{when they come}}
from the
market,
except they
wash, they
eat
not.
And
many other
things there
be,
which they have
received to
hold,
{\ADD{as}} the
washing of
cups,
and
pots, brasen
vessels,
and of
tables.
\VS{5}Then the
Pharisees
and
scribes
asked
him,
Why
walk
not
thy
disciples according
to the
tradition of the
elders,
but
eat
bread
with
unwashen
hands?
\VS{6}He
answered
and
said unto
them,
{\WJ{Well
hath
Esaias
prophesied
of
you
hypocrites,
as
it is written,
This
people
honoureth
me
with
{\ADD{their}}
lips,
but
their
heart
is
far
from
me.}}
\VS{7}{\WJ{Howbeit
in vain
do they worship
me,
teaching
{\ADD{for}}
doctrines
the commandments
of men.}}
\VS{8}{\WJ{For
laying aside
the commandment
of God,
ye hold
the tradition
of men,
{\ADD{as}}
the washing
of pots
and
cups:
and
many
other
such
like things
ye do.}}
\VS{9}And he
said unto
them,
{\WJ{Full well
ye reject
the commandment
of God,
that
ye may keep
your own
tradition.}}
\VS{10}{\WJ{For
Moses
said,
Honour
thy
father
and
thy
mother;
and,
Whoso curseth
father
or
mother,
let him die
the death:}}
\VS{11}{\WJ{But
ye
say,
If
a man
shall say
to his father
or
mother,
{\ADD{It is}}
Corban,
that
is to say,
a gift,
by
whatsoever
thou mightest be profited
by
me;
{\ADD{he shall be free}}.}}
\VS{12}{\WJ{And
ye suffer
him
no more
to do
ought
for
his
father
or
his
mother;}}
\VS{13}{\WJ{Making
the word
of God
of none effect
through your
tradition,
which
ye have delivered:
and
many
such
like things
do ye.}}
\par }{\PP \VS{14}¶
And when he had
called
all the
people
{\ADD{unto him}}, he
said unto
them,
{\WJ{Hearken
unto me
every one
{\ADD{of you}},
and
understand:}}
\VS{15}{\WJ{There is
nothing
from without
a man,
that
entering
into
him
can
defile
him:
but
the things which come out
of
him,
those
are they
that defile
the man.}}
\VS{16}{\WJ{If
any man
have
ears
to hear,
let him hear.}}
\VS{17}And
when he was
entered
into the
house
from the
people,
his
disciples
asked
him
concerning the
parable.
\VS{18}And he
saith unto
them,
{\WJ{Are
ye
so
without understanding
also?
Do ye
not
perceive,
that
whatsoever
thing from without
entereth
into
the man,
{\ADD{it}}
cannot
defile
him;}}
\VS{19}{\WJ{Because
it entereth
not
into
his
heart,
but
into
the belly,
and
goeth out
into
the draught,
purging
all
meats?}}
\VS{20}And he
said,
{\WJ{That which cometh out
of
the man,
that
defileth
the man.}}
\VS{21}{\WJ{For
from within,
out
of the heart
of men,
proceed
evil
thoughts,
adulteries,
fornications,
murders,}}
\VS{22}{\WJ{Thefts,
covetousness,
wickedness,
deceit,
lasciviousness,
an
evil
eye,
blasphemy,
pride,
foolishness:}}
\VS{23}{\WJ{All
these
evil things
come
from within,
and
defile
the man.}}
\par }{\PP \VS{24}¶
And from
thence he
arose, and
went
into the
borders of
Tyre
and
Sidon,
and
entered
into an
house, and would
have no
man
know
{\ADD{it}}:
but he
could
not be
hid.
\VS{25}For
a
{\ADD{certain}}
woman,
whose young
daughter
had
an
unclean
spirit,
heard
of
him, and
came and
fell
at
his
feet:
\VS{26}The
woman
was a
Greek, a
Syrophenician by
nation;
and she
besought
him
that he would cast
forth the
devil
out
of
her
daughter.
\VS{27}But
Jesus
said unto
her,
{\WJ{Let
the children
first
be filled:
for
it is
not
meet
to take
the children’s
bread,
and
to cast
{\ADD{it}}
unto the dogs.}}
\VS{28}And she
answered
and
said unto
him,
Yes,
Lord:
yet the
dogs
under the
table
eat
of the
children’s
crumbs.
\VS{29}And he
said unto
her,
{\WJ{For
this
saying
go thy way;
the devil
is gone out
of
thy
daughter.}}
\VS{30}And when she was
come
to
her
house, she
found the
devil gone
out,
and her
daughter
laid
upon the
bed.
\par }{\PP \VS{31}¶
And
again,
departing
from the
coasts of
Tyre
and
Sidon, he
came
unto the
sea of
Galilee,
through the
midst of the
coasts of
Decapolis.
\VS{32}And they
bring unto
him one that was
deaf, and had an impediment in his
speech;
and they
beseech
him
to
put his
hand
upon
him.
\VS{33}And he
took
him
aside
from the
multitude, and
put
his
fingers
into
his
ears,
and he
spit, and
touched
his
tongue;
\VS{34}And looking
up
to
heaven, he
sighed,
and
saith unto
him,
{\WJ{Ephphatha,}}
that
is, Be
opened.
\VS{35}And
straightway
his
ears were
opened,
and the
string
of
his
tongue was
loosed,
and he
spake
plain.
\VS{36}And he
charged
them
that they should
tell no
man:
but the
more
he
charged
them, so much the
more a great
deal they
published
{\ADD{it}};
\VS{37}And
were beyond
measure
astonished,
saying, He hath
done all
things
well: he
maketh
both the
deaf to
hear,
and the
dumb to
speak.

\par }\Chap{8}{\PP \VerseOne{1}In
those
days the
multitude
being very
great,
and
having
nothing to
eat,
Jesus
called
his
disciples
{\ADD{unto him}}, and
saith unto
them,
\VS{2}{\WJ{I have compassion
on
the multitude,
because
they have
now
been with
me
three
days,
and
have
nothing
to eat:}}
\VS{3}{\WJ{And
if
I send
them
away
fasting
to
their own
houses,
they will faint
by
the way:
for
divers
of them
came
from far.}}
\VS{4}And
his
disciples
answered
him, From
whence
can a
man
satisfy
these
{\ADD{men}} with
bread
here
in the
wilderness?
\VS{5}And he
asked
them,
{\WJ{How many
loaves
have ye?}}
And they
said,
Seven.
\VS{6}And he
commanded the
people to sit
down
on the
ground:
and he
took the
seven
loaves, and gave
thanks, and
brake,
and
gave
to
his
disciples
to set
before
{\ADD{them}};
and they did
set
{\ADD{them}}
before the
people.
\VS{7}And they
had a
few small
fishes:
and he
blessed, and
commanded to
set
them
also
before
{\ADD{them}}.
\VS{8}So they did
eat,
and were
filled:
and they took
up of the
broken
{\ADD{meat}} that was
left
seven
baskets.
\VS{9}And they that had
eaten
were
about four
thousand:
and he
sent
them
away.
\par }{\PP \VS{10}¶
And
straightway he
entered
into a
ship
with
his
disciples, and
came
into the
parts of
Dalmanutha.
\VS{11}And the
Pharisees came
forth,
and
began to
question with
him,
seeking
of
him a
sign
from
heaven,
tempting
him.
\VS{12}And he sighed
deeply
in
his
spirit, and
saith,
{\WJ{Why
doth
this
generation
seek after
a sign?
verily
I say
unto you,
There shall
no
sign
be given
unto
this
generation.}}
\VS{13}And he
left
them, and
entering
into the
ship
again
departed
to the other
side.
\par }{\PP \VS{14}¶
Now
{\ADD{the disciples}} had
forgotten to
take
bread,
neither had
they
in the
ship
with
them more than
one
loaf.
\VS{15}And he
charged
them,
saying,
{\WJ{Take heed,
beware
of
the leaven
of the Pharisees,
and
{\ADD{of}}
the leaven
of Herod.}}
\VS{16}And they
reasoned
among
themselves,
saying,
{\ADD{It is}}
because we
have
no
bread.
\VS{17}And
when
Jesus
knew
{\ADD{it}}, he
saith unto
them,
{\WJ{Why
reason ye,
because
ye have
no
bread?
perceive ye
not yet,
neither
understand?
have ye
your
heart
yet
hardened?}}
\VS{18}{\WJ{Having
eyes,
see ye
not?
and
having
ears,
hear ye
not?
and
do ye
not
remember?}}
\VS{19}{\WJ{When
I brake
the five
loaves
among
five thousand,
how many
baskets
full
of fragments
took ye up?}} They
say unto
him,
Twelve.
\VS{20}{\WJ{And
when
the seven
among
four thousand,
how many
baskets
full
of fragments
took ye up?}}
And they
said,
Seven.
\VS{21}And he
said unto
them,
{\WJ{How is it that
ye do
not
understand?}}
\par }{\PP \VS{22}¶
And he
cometh
to
Bethsaida;
and they
bring a blind
man unto
him,
and
besought
him
to
touch
him.
\VS{23}And he
took the blind
man by the
hand, and
led
him out
of the
town;
and when he had
spit
on
his
eyes, and
put his
hands upon
him, he
asked
him
if he
saw
ought.
\VS{24}And he looked
up, and
said, I
see
men
as
trees,
walking.
\VS{25}After
that he
put
{\ADD{his}}
hands
again
upon
his
eyes,
and
made
him look
up:
and he was
restored,
and
saw every
man
clearly.
\VS{26}And he
sent
him
away
to
his
house,
saying,
{\WJ{Neither
go
into
the town,
nor
tell
{\ADD{it}}
to any
in
the town.}}
\par }{\PP \VS{27}¶
And
Jesus went
out,
and
his
disciples,
into the
towns of
Cæsarea
Philippi:
and
by the
way he
asked
his
disciples,
saying unto
them,
{\WJ{Whom
do
men
say
that I
am?}}
\VS{28}And they
answered,
John the
Baptist:
but
some
{\ADD{say}},
Elias;
and
others,
One of the
prophets.
\VS{29}And
he
saith unto
them,
{\WJ{But
whom
say
ye
that I
am?}}
And
Peter
answereth and
saith unto
him,
Thou
art the
Christ.
\VS{30}And he
charged
them
that they should
tell no
man
of
him.
\VS{31}And he
began to
teach
them,
that the
Son of
man
must
suffer many
things,
and be
rejected
of the
elders,
and
{\ADD{of}} the chief
priests,
and
scribes,
and be
killed,
and
after
three
days rise
again.
\VS{32}And he
spake that
saying
openly.
And
Peter
took
him, and
began to
rebuke
him.
\VS{33}But when he had turned
about
and looked
on
his
disciples, he
rebuked
Peter,
saying,
{\WJ{Get thee
behind
me,
Satan:
for thou
savourest
not
the things
that be of God,
but
the things
that be of men.}}
\par }{\PP \VS{34}¶
And when he had
called the
people
{\ADD{unto him}}
with
his
disciples
also, he
said unto
them,
{\WJ{Whosoever
will
come
after
me,
let him deny
himself,
and
take up
his
cross,
and
follow
me.}}
\VS{35}{\WJ{For
whosoever
will
save
his
life
shall lose
it;
but
whosoever
shall lose
his
life
for
my sake
and
the gospel’s,
the same
shall save
it.}}
\VS{36}{\WJ{For
what
shall it profit
a man,
if
he shall gain
the
whole
world,
and
lose
his own
soul?}}
\VS{37}{\WJ{Or
what
shall
a man
give
in exchange for
his
soul?}}
\VS{38}{\WJ{Whosoever
therefore
shall be ashamed
of me
and of
my
words
in
this
adulterous
and
sinful
generation;
of him
also
shall
the Son
of man
be ashamed,
when
he cometh
in
the glory
of
his
Father
with
the
holy
angels.}}

\par }\Chap{9}{\PP \VerseOne{1}And he
said unto
them,
{\WJ{Verily
I say
unto you,
That
there be
some
of them that stand
here,
which
shall
not
taste
of death,
till
they have seen
the kingdom
of God
come
with
power.}}
\par }{\PP \VS{2}¶
And
after
six
days
Jesus
taketh
{\ADD{with him}}
Peter,
and
James,
and
John,
and
leadeth
them
up
into
an
high
mountain
apart
by
themselves:
and he was
transfigured
before
them.
\VS{3}And
his
raiment
became
shining,
exceeding
white
as
snow; so
as
no
fuller
on
earth
can white
them.
\VS{4}And there
appeared unto
them
Elias
with
Moses:
and they
were
talking with
Jesus.
\VS{5}And
Peter
answered and
said to
Jesus,
Master, it
is
good for
us to
be
here:
and let us
make
three
tabernacles;
one for
thee,
and
one for
Moses,
and
one for
Elias.
\VS{6}For he
wist
not
what to
say;
for they
were sore
afraid.
\VS{7}And there
was a
cloud that
overshadowed
them:
and a
voice
came out
of the
cloud,
saying,
This
is
my
beloved
Son:
hear
him.
\VS{8}And
suddenly, when they had looked round
about, they
saw no
man any
more,
save
Jesus
only
with
themselves.
\VS{9}And
as they came
down
from the
mountain, he
charged
them
that they should
tell no
man what
things they had
seen,
till the
Son of
man were
risen
from the
dead.
\VS{10}And they
kept that
saying
with
themselves, questioning one with
another
what the
rising
from the
dead should
mean.
\par }{\PP \VS{11}¶
And they
asked
him,
saying,
Why
say the
scribes
that
Elias
must
first
come?
\VS{12}And he
answered and
told
them,
{\WJ{Elias
verily
cometh
first,
and restoreth
all things;
and
how
it is written
of
the Son
of man,
that
he must suffer
many things,
and
be set at nought.}}
\VS{13}{\WJ{But
I say
unto you,
That
Elias
is
indeed
come,
and
they have done
unto him
whatsoever
they listed,
as
it is written
of
him.}}
\par }{\PP \VS{14}¶
And when he
came
to
{\ADD{his}}
disciples, he
saw
a
great
multitude
about
them,
and the
scribes
questioning with
them.
\VS{15}And
straightway
all the
people, when they
beheld
him, were greatly
amazed,
and running
to
{\ADD{him}}
saluted
him.
\VS{16}And he
asked the
scribes,
{\WJ{What
question ye
with
them?}}
\VS{17}And
one
of the
multitude
answered and
said,
Master, I have
brought
unto
thee
my
son, which
hath
a
dumb
spirit;
\VS{18}And
wheresoever he
taketh
him, he
teareth
him:
and he
foameth,
and
gnasheth
with
his
teeth,
and pineth
away:
and I
spake
to
thy
disciples
that they should
cast
him
out;
and they
could
not.
\VS{19}He
answereth
him,
and
saith,
{\WJ{O
faithless
generation,
how long
shall I be
with
you?
how long
shall I suffer
you?
bring
him
unto
me.}}
\VS{20}And they
brought
him
unto
him:
and when he
saw
him,
straightway the
spirit
tare
him;
and he
fell
on the
ground, and
wallowed
foaming.
\VS{21}And he
asked
his
father,
{\WJ{How long
is it
ago
since
this
came
unto him?}}
And he
said, Of a
child.
\VS{22}And
ofttimes it hath
cast
him
into the
fire,
and
into the
waters,
to
destroy
him:
but
if thou canst
do any
thing, have
compassion
on
us, and
help
us.
\VS{23}Jesus
said unto
him,
{\WJ{If
thou canst
believe,
all things
{\ADD{are}}
possible
to him that believeth.}}
\VS{24}And
straightway the
father of the
child cried
out, and
said
with
tears,
Lord, I
believe; help
thou
mine
unbelief.
\VS{25}When
Jesus
saw
that the
people came running
together, he
rebuked
the
foul
spirit,
saying unto
him,
{\WJ{{\ADD{Thou}}
dumb
and
deaf
spirit,
I
charge
thee,
come out
of
him,
and
enter
no more
into
him.}}
\VS{26}And
{\ADD{the spirit}}
cried,
and
rent
him
sore, and came
out of him:
and he
was
as one
dead; insomuch
that
many
said, He is
dead.
\VS{27}But
Jesus
took
him by the
hand, and
lifted
him
up;
and he
arose.
\VS{28}And
when he was
come
into the
house,
his
disciples
asked
him
privately,
Why
could
not
we
cast
him
out?
\VS{29}And he
said unto
them,
{\WJ{This
kind
can
come forth
by
nothing,
but
by
prayer
and
fasting.}}
\par }{\PP \VS{30}¶
And they
departed
thence, and
passed
through
Galilee;
and he
would
not
that any
man should
know
{\ADD{it}}.
\VS{31}For he
taught
his
disciples,
and
said unto
them,
{\WJ{The Son
of man
is delivered
into
the hands
of men,
and
they shall kill
him;
and
after that he is killed,
he shall rise
the
third
day.}}
\VS{32}But
they understood
not that
saying,
and were
afraid to
ask
him.
\par }{\PP \VS{33}¶
And he
came
to
Capernaum:
and
being
in the
house he
asked
them,
{\WJ{What was it that
ye disputed
among
yourselves
by
the way?}}
\VS{34}But
they held their
peace:
for
by the
way they had
disputed
among
themselves,
who
{\ADD{should be}} the
greatest.
\VS{35}And he sat
down, and
called the
twelve,
and
saith unto
them,
{\WJ{If
any man
desire
to be
first,
{\ADD{the same}}
shall be
last
of all,
and
servant
of all.}}
\VS{36}And he
took a
child, and
set
him
in the
midst of
them:
and when he had
taken
him in his
arms, he
said unto
them,
\VS{37}{\WJ{Whosoever
shall receive
one
of such
children
in
my
name,
receiveth
me:
and
whosoever
shall receive
me,
receiveth
not
me,
but
him that sent
me.}}
\par }{\PP \VS{38}¶
And
John
answered
him,
saying,
Master, we
saw
one casting
out
devils
in
thy
name, and
he
followeth
not
us:
and we
forbad
him,
because he
followeth
not
us.
\VS{39}But
Jesus
said,
{\WJ{Forbid
him
not:
for
there is
no man
which
shall do
a miracle
in
my
name,
that
can
lightly
speak evil
of me.}}
\VS{40}{\WJ{For
he that
is
not
against
us
is
on
our
part.}}
\VS{41}{\WJ{For
whosoever
shall give
you
a cup
of water
to drink
in
my
name,
because
ye belong
to Christ,
verily
I say
unto you,
he shall
not
lose
his
reward.}}
\VS{42}{\WJ{And
whosoever
shall offend
one
of
{\ADD{these}}
little
ones that believe
in
me,
it is
better
for him
that
a millstone
were hanged
about
his
neck,
and
he were cast
into
the sea.}}
\VS{43}{\WJ{And if thy hand offend thee, cut it off: it is better for thee to enter into life maimed, than having two hands to go into hell, into the fire that never shall be quenched:}}
\VS{44}{\WJ{Where
their
worm
dieth
not,
and
the fire
is
not
quenched.}}
\VS{45}{\WJ{And
if
thy
foot
offend
thee,
cut
it
off:
it is
better
for thee
to enter
halt
into
life,
than
having
two
feet
to be cast
into
hell,
into
the fire
that never shall be quenched:}}
\VS{46}{\WJ{Where
their
worm
dieth
not,
and
the fire
is
not
quenched.}}
\VS{47}{\WJ{And
if
thine
eye
offend
thee,
pluck
it
out:
it is
better
for thee
to enter
into
the kingdom
of God
with one eye,
than
having
two
eyes
to be cast
into
hell
fire:}}
\VS{48}{\WJ{Where
their
worm
dieth
not,
and
the fire
is
not
quenched.}}
\VS{49}{\WJ{For
every one
shall be salted
with fire,
and
every
sacrifice
shall be salted
with salt.}}
\VS{50}{\WJ{Salt
{\ADD{is}}
good:
but
if
the salt
have lost his saltness,
wherewith
will ye season
it?
Have
salt
in
yourselves,
and
have peace
one
with
another.}}

\par }\Chap{10}{\PP \VerseOne{1}And he
arose from
thence, and
cometh
into the
coasts of
Judæa
by the farther
side of
Jordan:
and the
people
resort
unto
him
again;
and,
as he was
wont, he
taught
them
again.
\par }{\PP \VS{2}¶
And the
Pharisees came
to
him, and
asked
him, Is it
lawful for a
man to put
away
{\ADD{his}}
wife?
tempting
him.
\VS{3}And he
answered and
said unto
them,
{\WJ{What
did
Moses
command
you?}}
\VS{4}And they
said,
Moses
suffered to
write a
bill of
divorcement,
and to put her
away.
\VS{5}And
Jesus
answered and
said unto
them,
{\WJ{For
the hardness of
your
heart
he wrote
you
this
precept.}}
\VS{6}{\WJ{But
from
the beginning
of the creation
God
made
them
male
and
female.}}
\VS{7}{\WJ{For
this
cause
shall
a man
leave
his
father
and
mother,
and
cleave
to
his
wife;}}
\VS{8}{\WJ{And
they
twain
shall be
one
flesh:
so then
they are
no more
twain,
but
one
flesh.}}
\VS{9}{\WJ{What
therefore
God
hath joined together,
let
not
man
put asunder.}}
\VS{10}And
in the
house
his
disciples
asked
him
again
of the
same
{\ADD{matter}}.
\VS{11}And he
saith unto
them,
{\WJ{Whosoever
shall put away
his
wife,
and
marry
another,
committeth adultery
against
her.}}
\VS{12}{\WJ{And
if
a woman
shall put away
her
husband,
and
be married to
another,
she committeth adultery.}}
\par }{\PP \VS{13}¶
And they
brought young
children to
him,
that he should
touch
them:
and
{\ADD{his}}
disciples
rebuked those that
brought
{\ADD{them}}.
\VS{14}But
when
Jesus
saw
{\ADD{it}}, he was much
displeased,
and
said unto
them,
{\WJ{Suffer
the little children
to come
unto
me,
and
forbid
them
not:
for
of such
is
the kingdom
of God.}}
\VS{15}{\WJ{Verily
I say
unto you,
Whosoever
shall
not
receive
the kingdom
of God
as
a little child,
he shall
not
enter
therein.}}
\VS{16}And he
took
them up in his
arms,
put
{\ADD{his}}
hands
upon
them, and
blessed
them.
\par }{\PP \VS{17}¶
And
when
he was gone
forth
into the
way, there
came
one
running,
and
kneeled to
him, and
asked
him,
Good
Master,
what shall I
do
that I may
inherit
eternal
life?
\VS{18}And
Jesus
said unto
him,
{\WJ{Why
callest thou
me
good?
{\ADD{there is}}
none
good
but
one,
{\ADD{that is}},
God.}}
\VS{19}{\WJ{Thou knowest
the commandments,
Do
not
commit adultery,
Do
not
kill,
Do
not
steal,
Do
not
bear false witness,
Defraud
not,
Honour
thy
father
and
mother.}}
\VS{20}And he
answered and
said unto
him,
Master,
all
these have I
observed
from
my
youth.
\VS{21}Then
Jesus
beholding
him
loved
him,
and
said unto
him,
{\WJ{One thing
thou
lackest:
go thy way,
sell
whatsoever
thou hast,
and
give
to the poor,
and
thou shalt have
treasure
in
heaven:
and
come,
take up
the cross,
and follow
me.}}
\VS{22}And he was
sad
at that
saying, and went
away
grieved:
for
he
had
great
possessions.
\par }{\PP \VS{23}¶
And
Jesus looked round
about, and
saith
unto
his
disciples,
{\WJ{How
hardly
shall
they that have
riches
enter
into
the kingdom
of God!}}
\VS{24}And the
disciples were
astonished
at
his
words.
But
Jesus
answereth
again, and
saith unto
them,
{\WJ{Children,
how
hard
is it
for them that trust
in
riches
to enter
into
the kingdom
of God!}}
\VS{25}{\WJ{It is
easier
for a camel
to go
through
the eye
of a needle,
than
for a rich man
to enter
into
the kingdom
of God.}}
\VS{26}And they were
astonished out of
measure,
saying
among
themselves,
Who
then
can be
saved?
\VS{27}And
Jesus looking
upon
them
saith,
{\WJ{With
men
{\ADD{it is}}
impossible,
but
not
with
God:
for
with
God
all things
are
possible.}}
\par }{\PP \VS{28}¶
Then
Peter
began to
say unto
him,
Lo,
we have
left
all,
and have
followed
thee.
\VS{29}And
Jesus
answered and
said,
{\WJ{Verily
I say
unto you,
There is
no man
that
hath left
house,
or
brethren,
or
sisters,
or
father,
or
mother,
or
wife,
or
children,
or
lands,
for
my
sake,
and
the gospel’s,}}
\VS{30}{\WJ{But
he shall receive
an hundredfold
now
in
this
time,
houses,
and
brethren,
and
sisters,
and
mothers,
and
children,
and
lands,
with
persecutions;
and
in
the world
to come
eternal
life.}}
\VS{31}{\WJ{But
many
{\ADD{that are}}
first
shall be
last;
and
the last
first.}}
\par }{\PP \VS{32}¶
And they
were
in the
way going
up
to
Jerusalem;
and
Jesus went
before
them:
and they were
amazed;
and as they
followed, they were
afraid.
And he
took
again the
twelve, and
began to
tell
them what things
should
happen unto
him,
\VS{33}{\ADD{Saying}},
{\WJ{Behold,
we go up
to
Jerusalem;
and
the Son
of man
shall be delivered
unto the chief priests,
and
unto the scribes;
and
they shall condemn
him
to death,
and
shall deliver
him
to the Gentiles:}}
\VS{34}{\WJ{And
they shall mock
him,
and
shall scourge
him,
and
shall spit upon
him,
and
shall kill
him:
and
the
third
day
he shall rise again.}}
\par }{\PP \VS{35}¶
And
James
and
John, the
sons of
Zebedee, come
unto
him,
saying,
Master, we
would
that thou shouldest
do for
us
whatsoever we shall
desire.
\VS{36}And he
said unto
them,
{\WJ{What
would ye that
I
should do
for you?}}
\VS{37}They
said unto
him,
Grant unto
us
that we may
sit,
one
on
thy right
hand,
and the
other
on
thy left
hand,
in
thy
glory.
\VS{38}But
Jesus
said unto
them,
{\WJ{Ye know
not
what
ye ask:
can ye
drink of
the cup
that
I
drink of?
and
be baptized with
the baptism
that
I
am baptized with?}}
\VS{39}And they
said unto
him, We
can.
And
Jesus
said unto
them,
{\WJ{Ye shall
indeed
drink of
the cup
that
I
drink of;
and
with the baptism
that
I
am baptized withal
shall ye be baptized:}}
\VS{40}{\WJ{But
to sit
on
my
right hand
and
on
my
left hand
is
not
mine
to give;
but
{\ADD{it shall be given to them}}
for whom
it is prepared.}}
\VS{41}And
when the
ten
heard
{\ADD{it}}, they
began to be much
displeased
with
James
and
John.
\VS{42}But
Jesus
called
them
{\ADD{to him}}, and
saith unto
them,
{\WJ{Ye know
that
they which are accounted
to rule over
the Gentiles
exercise lordship over
them;
and
their
great ones
exercise authority upon
them.}}
\VS{43}{\WJ{But
so
shall it
not
be
among
you:
but
whosoever
will
be
great
among
you,
shall be
your
minister:}}
\VS{44}{\WJ{And
whosoever
of you
will
be
the chiefest,
shall be
servant
of all.}}
\VS{45}{\WJ{For
even
the Son
of man
came
not
to be ministered unto,
but
to minister,
and
to give
his
life
a ransom
for
many.}}
\par }{\PP \VS{46}¶
And they
came
to
Jericho:
and
as
he went
out
of
Jericho
with
his
disciples
and
a
great number of
people,
blind
Bartimæus, the
son of
Timæus,
sat
by the highway
side
begging.
\VS{47}And when he
heard
that it
was
Jesus of
Nazareth, he
began to cry
out,
and
say,
Jesus,
{\ADD{thou}}
Son of
David, have mercy
on
me.
\VS{48}And
many
charged
him
that he should hold his
peace:
but he
cried the
more a great
deal,
{\ADD{Thou}}
Son of
David, have mercy
on
me.
\VS{49}And
Jesus stood
still, and
commanded
him to be
called.
And they
call the blind
man,
saying unto
him, Be of good
comfort,
rise; he
calleth
thee.
\VS{50}And he, casting
away
his
garment,
rose, and
came
to
Jesus.
\VS{51}And
Jesus
answered and
said unto
him,
{\WJ{What
wilt thou
that
I should do
unto thee?}} The blind
man
said unto
him,
Lord,
that I might receive my
sight.
\VS{52}And
Jesus
said unto
him,
{\WJ{Go thy way;
thy
faith
hath made
thee
whole.}}
And
immediately he received his
sight,
and
followed
Jesus
in the
way.

\par }\Chap{11}{\PP \VerseOne{1}And
when they came
nigh
to
Jerusalem,
unto
Bethphage
and
Bethany,
at the
mount of
Olives, he sendeth
forth
two of
his
disciples,
\VS{2}And
saith unto
them,
{\WJ{Go your way
into
the village
over against
you:
and
as soon as
ye be entered
into
it,
ye shall find
a colt
tied,
whereon
never
man
sat;
loose
him,
and bring
{\ADD{him}}.}}
\VS{3}{\WJ{And
if
any man
say
unto you,
Why
do ye
this?
say ye
that
the Lord
hath
need
of him;
and
straightway
he will send
him
hither.}}
\VS{4}And they went their
way,
and
found the
colt
tied
by the
door
without
in a place where two ways
met;
and they
loose
him.
\VS{5}And
certain of them
that
stood
there
said unto
them,
What do
ye,
loosing the
colt?
\VS{6}And they
said unto
them even
as
Jesus had
commanded:
and they
let
them
go.
\VS{7}And they
brought the
colt
to
Jesus,
and
cast
their
garments on
him;
and he
sat
upon
him.
\VS{8}And
many
spread
their
garments
in the
way:
and
others cut
down
branches
off the
trees,
and
strawed
{\ADD{them}}
in the
way.
\VS{9}And they
that went
before,
and they
that
followed,
cried,
saying,
Hosanna;
Blessed
{\ADD{is}} he
that
cometh
in the
name of the
Lord:
\VS{10}Blessed
{\ADD{be}} the
kingdom of
our
father
David,
that
cometh
in the
name of the
Lord:
Hosanna
in the
highest.
\VS{11}And
Jesus
entered
into
Jerusalem,
and
into the
temple: and
when he had looked round about
upon all
things, and
now the
eventide was
come, he went
out
unto
Bethany
with the
twelve.
\par }{\PP \VS{12}¶
And on the
morrow, when
they were
come
from
Bethany, he was
hungry:
\VS{13}And
seeing a fig
tree afar
off
having
leaves, he
came,
if
haply he might
find any
thing
thereon: and
when he
came
to
it, he
found
nothing
but
leaves;
for the
time of
figs
was
not
{\ADD{yet}}.
\VS{14}And
Jesus
answered and
said unto
it,
{\WJ{No man
eat
fruit
of
thee
hereafter
for
ever.}}
And
his
disciples
heard
{\ADD{it.}}
\par }{\PP \VS{15}¶
And they
come
to
Jerusalem:
and
Jesus
went
into the
temple, and
began to cast
out them
that
sold
and
bought
in the
temple,
and
overthrew the
tables of the
moneychangers,
and the
seats of them
that
sold
doves;
\VS{16}And
would
not
suffer
that any
man should
carry
{\ADD{any}}
vessel
through the
temple.
\VS{17}And he
taught,
saying unto
them,
{\WJ{Is it
not
written,
My
house
shall be called
of all
nations
the house
of prayer?
but
ye
have made
it
a den
of thieves.}}
\VS{18}And the
scribes
and chief
priests
heard
{\ADD{it}},
and
sought
how they might
destroy
him:
for they
feared
him,
because
all the
people was
astonished
at
his
doctrine.
\VS{19}And
when
even was
come, he
went out
of the
city.
\par }{\PP \VS{20}¶
And in the
morning, as they passed
by, they
saw the fig
tree dried
up
from the
roots.
\VS{21}And
Peter calling to
remembrance
saith unto
him,
Master,
behold, the fig
tree
which thou
cursedst is withered
away.
\VS{22}And
Jesus
answering
saith unto
them,
{\WJ{Have
faith
in God.}}
\VS{23}{\WJ{For
verily
I say
unto you,
That
whosoever
shall say
unto this
mountain,
Be thou removed,
and
be thou cast
into
the sea;
and
shall
not
doubt
in
his
heart,
but
shall believe
that
those things which
he
saith
shall come to pass;
he shall have
whatsoever
he saith.}}
\VS{24}{\WJ{Therefore
I say
unto you,
What things soever
ye desire,
when ye pray,
believe
that
ye receive
{\ADD{them}},
and
ye
shall have
{\ADD{them}}.}}
\VS{25}{\WJ{And
when
ye stand
praying,
forgive,
if
ye have
ought
against
any:
that
your
Father
also
which is
in
heaven
may forgive
you
your
trespasses.}}
\VS{26}{\WJ{But
if
ye
do
not
forgive,
neither
will
your
Father
which is
in
heaven
forgive
your
trespasses.}}
\par }{\PP \VS{27}¶
And they
come
again
to
Jerusalem:
and as
he was
walking
in the
temple, there
come
to
him the chief
priests,
and the
scribes,
and the
elders,
\VS{28}And
say unto
him,
By
what
authority doest
thou these
things?
and
who
gave
thee
this
authority
to
do these
things?
\VS{29}And
Jesus
answered and
said unto
them,
{\WJ{I will
also
ask
of you
one
question,
and
answer
me,
and
I will tell
you
by
what
authority
I do
these things.}}
\VS{30}{\WJ{The baptism
of John,
was
{\ADD{it}}
from
heaven,
or
of
men?
answer
me.}}
\VS{31}And they
reasoned
with
themselves,
saying,
If we shall
say,
From
heaven; he will
say,
Why
then did
ye
not
believe
him?
\VS{32}But
if we shall
say,
Of
men; they
feared the
people:
for
all
{\ADD{men}}
counted
John,
that he
was a
prophet
indeed.
\VS{33}And they
answered and
said unto
Jesus,
We
cannot
tell.
And
Jesus
answering
saith unto
them,
{\WJ{Neither
do I
tell
you
by
what
authority
I do
these things.}}

\par }\Chap{12}{\PP \VerseOne{1}And he
began to
speak unto
them
by
parables.
{\WJ{A
{\ADD{certain}}
man
planted
a vineyard,
and
set
an hedge about
{\ADD{it}},
and
digged
{\ADD{a place for}}
the winefat,
and
built
a tower,
and
let
it
out
to husbandmen,
and
went into a far country.}}
\VS{2}{\WJ{And
at the season
he sent
to
the husbandmen
a servant,
that
he might receive
from
the husbandmen
of
the fruit
of the vineyard.}}
\VS{3}{\WJ{And
they caught
{\ADD{him}},
and
beat
him,
and sent
{\ADD{him}}
away
empty.}}
\VS{4}{\WJ{And
again
he sent
unto
them
another
servant;
and at him
they cast stones,
and wounded
{\ADD{him}}
in the head,
and
sent
{\ADD{him}}
away
shamefully handled.}}
\VS{5}{\WJ{And
again
he sent
another;
and him
they killed,
and
many
others;
beating
some,
and
killing
some.}}
\VS{6}{\WJ{Having
yet
therefore
one
son,
his
wellbeloved,
he sent
him
also
last
unto
them,
saying,
They will reverence
my
son.}}
\VS{7}{\WJ{But
those
husbandmen
said
among
themselves,
This
is
the heir;
come,
let us kill
him,
and
the inheritance
shall be
ours.}}
\VS{8}{\WJ{And
they took
him,
and killed
{\ADD{him}},
and
cast
{\ADD{him}}
out of
the vineyard.}}
\VS{9}{\WJ{What
shall
therefore
the lord
of the vineyard
do?
he will come
and
destroy
the husbandmen,
and
will give
the vineyard
unto others.}}
\VS{10}{\WJ{And have ye
not
read
this
scripture;
The stone
which
the builders
rejected
is
become
the head
of the corner:}}
\VS{11}{\WJ{This
was
the Lord’s
doing,
and
it is
marvellous
in
our
eyes?}}
\VS{12}And they
sought to lay
hold on
him,
but
feared the
people:
for they
knew
that he had
spoken the
parable
against
them: and they
left
him,
and went their
way.
\par }{\PP \VS{13}¶
And they
send
unto
him
certain of the
Pharisees
and of the
Herodians,
to
catch
him
in
{\ADD{his}}
words.
\VS{14}And
when they were
come, they
say unto
him,
Master, we
know
that thou
art
true,
and
carest
for no
man:
for thou
regardest
not the
person of
men,
but
teachest the
way of
God
in
truth: Is it
lawful to
give
tribute to
Cæsar,
or
not?
\VS{15}Shall we
give,
or shall
we
not
give?
But
he,
knowing
their
hypocrisy,
said unto
them,
{\WJ{Why
tempt ye
me?
bring
me
a penny,
that
I may see
{\ADD{it}}.}}
\VS{16}And they
brought
{\ADD{it}}.
And he
saith unto
them,
{\WJ{Whose
{\ADD{is}}
this
image
and
superscription?}}
And they
said unto
him,
Cæsar’s.
\VS{17}And
Jesus
answering
said unto
them,
{\WJ{Render
to Cæsar
the things that are
Cæsar’s,
and
to God
the things that are
God’s.}}
And they
marvelled
at
him.
\par }{\PP \VS{18}¶
Then
come
unto
him the
Sadducees,
which
say there
is
no
resurrection;
and they
asked
him,
saying,
\VS{19}Master,
Moses
wrote unto
us,
If a
man’s
brother
die,
and
leave
{\ADD{his}}
wife
{\ADD{behind him}},
and
leave
no
children,
that
his
brother should
take
his
wife,
and raise
up
seed unto
his
brother.
\VS{20}Now there
were
seven
brethren:
and the
first
took a
wife,
and
dying
left
no
seed.
\VS{21}And the
second
took
her,
and
died,
neither
left
he any
seed:
and the
third
likewise.
\VS{22}And the
seven
had
her,
and
left
no
seed:
last of
all the
woman
died
also.
\VS{23}In the
resurrection
therefore,
when they shall
rise,
whose
wife shall she
be of
them?
for the
seven
had
her to
wife.
\VS{24}And
Jesus
answering
said unto
them,
{\WJ{Do ye
not
therefore
err,
because
ye know
not
the scriptures,
neither
the power
of God?}}
\VS{25}{\WJ{For
when
they shall rise
from
the dead,
they
neither
marry,
nor
are given in marriage;
but
are
as
the angels
which are
in
heaven.}}
\VS{26}{\WJ{And
as touching
the dead,
that
they rise:
have ye
not
read
in
the book
of Moses,
how
in
the bush
God
spake
unto him,
saying,
I
{\ADD{am}}
the God
of Abraham,
and
the God
of Isaac,
and
the God
of Jacob?}}
\VS{27}{\WJ{He is
not
the God
of the dead,
but
the God
of the living:
ye
therefore
do
greatly
err.}}
\par }{\PP \VS{28}¶
And
one of the
scribes
came, and having
heard
them reasoning
together, and
perceiving
that he had
answered
them
well,
asked
him,
Which
is the
first
commandment of
all?
\VS{29}And
Jesus
answered
him,
{\WJ{The first
of all
the commandments
{\ADD{is}},
Hear,
O Israel;
The Lord
our
God
is
one
Lord:}}
\VS{30}{\WJ{And
thou shalt love
the Lord
thy
God
with
all
thy
heart,
and
with
all
thy
soul,
and
with
all
thy
mind,
and
with
all
thy
strength:
this
{\ADD{is}}
the first
commandment.}}
\VS{31}{\WJ{And
the second
{\ADD{is}}
like,
{\ADD{namely}}
this,
Thou shalt love
thy
neighbour
as
thyself.
There is
none
other
commandment
greater
than these.}}
\VS{32}And the
scribe
said unto
him,
Well,
Master, thou hast
said the
truth:
for there
is
one
God;
and there
is
none
other
but
he:
\VS{33}And to
love
him
with
all the
heart,
and
with
all the
understanding,
and
with
all the
soul,
and
with
all the
strength,
and to
love
{\ADD{his}}
neighbour
as
himself,
is
more than
all whole burnt
offerings
and
sacrifices.
\VS{34}And when
Jesus
saw
that he
answered
discreetly, he
said unto
him,
{\WJ{Thou art
not
far
from
the kingdom
of God.}}
And no
man after
that
durst
ask
him
{\ADD{any question.}}
\par }{\PP \VS{35}¶
And
Jesus
answered and
said, while he
taught
in the
temple,
{\WJ{How
say
the scribes
that
Christ
is
the Son
of David?}}
\VS{36}{\WJ{For
David
himself
said
by
the Holy
Ghost, The
{\ND{
LORD}}
said
to my
Lord,
Sit thou
on
my
right hand,
till
I make
thine
enemies
thy
footstool.}}
\VS{37}{\WJ{David
therefore
himself
calleth
him
Lord;
and
whence
is he
{\ADD{then}}
his
son?
And
the common
people
heard
him
gladly.}}
\par }{\PP \VS{38}¶
And he
said unto
them
in
his
doctrine,
{\WJ{Beware
of
the scribes,
which
love
to go
in
long clothing,
and
{\ADD{love}}
salutations
in
the marketplaces,}}
\VS{39}{\WJ{And
the chief seats
in
the synagogues,
and
the uppermost rooms
at
feasts:}}
\VS{40}{\WJ{Which
devour
widows’
houses,
and
for a pretence
make
long
prayers:
these
shall receive
greater
damnation.}}
\par }{\PP \VS{41}¶
And
Jesus
sat over
against the
treasury, and
beheld
how the
people
cast
money
into the
treasury:
and
many that were
rich
cast in
much.
\VS{42}And there
came a
certain
poor
widow, and she threw
in
two
mites,
which
make a
farthing.
\VS{43}And he
called
{\ADD{unto him}}
his
disciples, and
saith unto
them,
{\WJ{Verily
I say
unto you,
That
this
poor
widow
hath cast
more
in,
than all
they which
have cast
into
the treasury:}}
\VS{44}{\WJ{For
all
{\ADD{they}}
did cast in
of
their
abundance;
but
she
of
her
want
did cast in
all
that
she had,
{\ADD{even}}
all
her
living.}}

\par }\Chap{13}{\PP \VerseOne{1}And as
he went
out
of the
temple,
one of
his
disciples
saith unto
him,
Master,
see what manner
of
stones
and
what
buildings
{\ADD{are here}}!
\VS{2}And
Jesus
answering
said unto
him,
{\WJ{Seest thou
these
great
buildings?
there shall
not
be left
one stone
upon
another,
that
shall
not
be thrown down.}}
\VS{3}And as
he
sat
upon the
mount of
Olives over
against the
temple,
Peter
and
James
and
John
and
Andrew
asked
him
privately,
\VS{4}Tell
us,
when
shall these
things
be?
and
what
{\ADD{
shall be}} the
sign
when
all these
things shall be
fulfilled?
\VS{5}And
Jesus
answering
them
began to
say,
{\WJ{Take heed
lest
any
{\ADD{man}}
deceive
you:}}
\VS{6}{\WJ{For
many
shall come
in
my
name,
saying,
I
am
{\ADD{Christ}};
and
shall deceive
many.}}
\VS{7}{\WJ{And
when
ye shall hear
of wars
and
rumours
of wars,
be ye
not
troubled:
for
{\ADD{such things}}
must needs be;
but
the end
{\ADD{shall}} not
{\ADD{be}}
yet.}}
\VS{8}{\WJ{For
nation
shall rise
against
nation,
and
kingdom
against
kingdom:
and
there shall be
earthquakes
in
divers places,
and
there shall be
famines
and
troubles:
these
{\ADD{are}}
the beginnings
of sorrows.}}
\par }{\PP \VS{9}{\WJ{¶
But
take heed
to yourselves:
for
they shall deliver
you
up
to
councils;
and
in
the synagogues
ye shall be beaten:
and
ye shall be brought
before
rulers
and
kings
for
my sake,
for
a testimony
against them.}}
\VS{10}{\WJ{And
the gospel
must
first
be published
among
all
nations.}}
\VS{11}{\WJ{But
when
they shall lead
{\ADD{
you}},
and deliver you up,
take
no
thought beforehand
what
ye shall speak,
neither
do ye premeditate:
but
whatsoever
shall be given
you
in
that
hour,
that
speak ye:
for
it is
not
ye that
speak,
but
the Holy
Ghost.}}
\VS{12}{\WJ{Now
the brother
shall betray
the brother
to
death,
and
the father
the son;
and
children
shall rise up
against
{\ADD{their}}
parents,
and
shall cause
them
to be put to death.}}
\VS{13}{\WJ{And
ye shall be
hated
of
all
{\ADD{men}}
for
my
name’s sake:
but
he that
shall endure
unto
the end,
the same
shall be saved.}}
\par }{\PP \VS{14}{\WJ{¶
But
when
ye shall see
the abomination
of desolation,
spoken of
by
Daniel
the prophet,
standing
where
it ought
not, (let him that
readeth
understand,)
then
let them that be
in
Judæa
flee
to
the mountains:}}
\VS{15}{\WJ{And
let him that is
on
the housetop
not
go down
into
the house,
neither
enter
{\ADD{therein}},
to take
any thing
out of
his
house:}}
\VS{16}{\WJ{And
let him
that is
in
the field
not
turn
back
again
for to take up
his
garment.}}
\VS{17}{\WJ{But
woe
to them
that are
with
child,
and
to them
that give suck
in
those
days!}}
\VS{18}{\WJ{And
pray ye
that
your
flight
be
not
in the winter.}}
\VS{19}{\WJ{For
{\ADD{in}}
those
days
shall be
affliction,
such as
was
not
from
the beginning
of the creation
which
God
created
unto
this time,
neither
shall be.}}
\VS{20}{\WJ{And
except
that the Lord
had shortened
those days,
no
flesh
should be saved:
but
for
the elect’s sake,
whom
he hath chosen,
he hath shortened
the days.}}
\VS{21}{\WJ{And
then
if
any man
shall say
to you,
Lo,
here
{\ADD{is}}
Christ;
or,
lo,
{\ADD{he is}}
there;
believe
{\ADD{him}}
not:}}
\VS{22}{\WJ{For
false Christs
and
false prophets
shall rise,
and
shall shew
signs
and
wonders,
to
seduce,
if
{\ADD{it were}}
possible,
even
the elect.}}
\VS{23}{\WJ{But
take
ye
heed:
behold,
I have foretold
you
all things.}}
\par }{\PP \VS{24}{\WJ{¶
But
in
those
days,
after
that
tribulation,
the sun
shall be darkened,
and
the moon
shall
not
give
her
light,}}
\VS{25}{\WJ{And
the stars
of heaven
shall
fall,
and
the powers
that are
in
heaven
shall be shaken.}}
\VS{26}{\WJ{And
then
shall they see
the Son
of man
coming
in
the clouds
with
great
power
and
glory.}}
\VS{27}{\WJ{And
then
shall he send
his
angels,
and
shall gather together
his
elect
from
the four
winds,
from
the uttermost part
of the earth
to
the uttermost part
of heaven.}}
\VS{28}{\WJ{Now
learn
a parable
of
the fig tree;
When
her
branch
is
yet
tender,
and
putteth forth
leaves,
ye know
that
summer
is
near:}}
\VS{29}{\WJ{So
ye
in like manner,
when
ye shall see
these things
come to pass,
know
that
it is
nigh,
{\ADD{even}}
at
the doors.}}
\VS{30}{\WJ{Verily
I say
unto you,
that
this
generation
shall
not
pass,
till
all
these things
be done.}}
\VS{31}{\WJ{Heaven
and
earth
shall pass away:
but
my
words
shall
not
pass away.}}
\par }{\PP \VS{32}{\WJ{¶
But
of
that
day
and
{\ADD{that}}
hour
knoweth
no man, no,
not
the angels
which are
in
heaven,
neither
the Son,
but
the Father.}}
\VS{33}{\WJ{Take ye heed,
watch
and
pray:
for
ye know
not
when
the time
is.}}
\VS{34}{\WJ{{\ADD{For the Son of man is}}
as
a man
taking a far journey,
who left
his
house,
and
gave
authority
to his
servants,
and
to every man
his
work,
and
commanded
the porter
to
watch.}}
\VS{35}{\WJ{Watch ye
therefore:
for
ye know
not
when
the master
of the house
cometh,
at even,
or
at midnight,
or
at the cockcrowing,
or
in the morning:}}
\VS{36}{\WJ{Lest
coming
suddenly
he find
you
sleeping.}}
\VS{37}{\WJ{And
what
I say
unto you
I say
unto all,
Watch.}}

\par }\Chap{14}{\PP \VerseOne{1}After
two
days
was
{\ADD{the feast of}} the
passover,
and of unleavened
bread:
and the chief
priests
and the
scribes
sought
how they might
take
him
by
craft, and
put
{\ADD{him}} to
death.
\VS{2}But they
said,
Not
on the
feast
{\ADD{day}},
lest there
be an
uproar of the
people.
\par }{\PP \VS{3}¶
And
being
in
Bethany
in the
house of
Simon the
leper, as
he sat at
meat, there
came a
woman
having an alabaster
box of
ointment of
spikenard very
precious;
and she
brake the
box, and
poured
{\ADD{it}}
on
his
head.
\VS{4}And there
were
some that had
indignation
within
themselves,
and
said,
Why
was
this
waste of the
ointment
made?
\VS{5}For
it
might have been
sold for more
than three
hundred
pence,
and have been
given to the
poor.
And they murmured
against
her.
\VS{6}And
Jesus
said,
{\WJ{Let
her
alone;
why
trouble ye
her?
she hath wrought
a good
work
on
me.}}
\VS{7}{\WJ{For
ye have
the poor
with
you
always,
and
whensoever
ye will
ye may
do
them
good:
but
me
ye have
not
always.}}
\VS{8}{\WJ{She
hath done
what
she could:
she is come aforehand
to
anoint
my
body
to the burying.}}
\VS{9}{\WJ{Verily
I say
unto you,
Wheresoever
this
gospel
shall be preached
throughout
the whole
world,
{\ADD{this}}
also
that
she
hath done
shall be spoken of
for
a memorial
of her.}}
\par }{\PP \VS{10}¶
And
Judas
Iscariot,
one of the
twelve,
went
unto the chief
priests,
to
betray
him unto
them.
\VS{11}And when
they
heard
{\ADD{it}}, they were
glad,
and
promised to
give
him
money.
And he
sought
how he
might
conveniently
betray
him.
\par }{\PP \VS{12}¶
And the
first
day of unleavened
bread,
when they
killed the
passover,
his
disciples
said unto
him,
Where wilt
thou that we
go and
prepare
that thou mayest
eat the
passover?
\VS{13}And he sendeth
forth
two of
his
disciples,
and
saith unto
them,
{\WJ{Go ye
into
the city,
and
there shall meet
you
a man
bearing
a pitcher
of water:
follow
him.}}
\VS{14}{\WJ{And
wheresoever
he shall go in,
say ye
to the goodman of the house,
The Master
saith,
Where
is
the guestchamber,
where
I shall eat
the passover
with
my
disciples?}}
\VS{15}{\WJ{And
he
will shew
you
a large
upper room
furnished
{\ADD{and}}
prepared:
there
make ready
for us.}}
\VS{16}And
his
disciples went
forth,
and
came
into the
city,
and
found
as he had
said unto
them:
and they made
ready the
passover.
\VS{17}And
in the
evening he
cometh
with the
twelve.
\VS{18}And as
they
sat
and did
eat,
Jesus
said,
{\WJ{Verily
I say
unto you,
One
of
you
which
eateth
with
me
shall betray
me.}}
\VS{19}And they
began to be
sorrowful,
and to
say unto
him
one
by
one,
{\ADD{Is}}
it
I?
and
another
{\ADD{said, Is}}
it
I?
\VS{20}And he
answered and
said unto
them,
{\WJ{{\ADD{It is}}
one
of
the twelve,
that
dippeth
with
me
in
the dish.}}
\VS{21}{\WJ{The Son
of man
indeed
goeth,
as
it is written
of
him:
but
woe
to that
man
by
whom
the Son
of man
is betrayed!
good
were it
for that
man
if
he
had
never
been born.}}
\par }{\PP \VS{22}¶
And as
they did
eat,
Jesus
took
bread, and
blessed, and
brake
{\ADD{it}},
and
gave to
them,
and
said,
{\WJ{Take,
eat:
this
is
my
body.}}
\VS{23}And he
took the
cup, and when he had given
thanks, he
gave
{\ADD{it}} to
them:
and
they
all
drank
of
it.
\VS{24}And he
said unto
them,
{\WJ{This
is
my
blood
of the new
testament,
which
is shed
for
many.}}
\VS{25}{\WJ{Verily
I say
unto you,
I will drink
no more
of
the fruit
of the vine,
until
that
day
that
I drink
it
new
in
the kingdom
of God.}}
\par }{\PP \VS{26}¶
And when they had sung an
hymn, they went
out
into the
mount of
Olives.
\VS{27}And
Jesus
saith unto
them,
{\WJ{All
ye shall be offended
because
of me
this
night:
for
it is written,
I will smite
the shepherd,
and
the sheep
shall be scattered.}}
\VS{28}{\WJ{But
after
that I
am risen,
I will go before
you
into
Galilee.}}
\VS{29}But
Peter
said unto
him,
Although
all shall be
offended,
yet
{\ADD{will}}
not
I.
\VS{30}And
Jesus
saith unto
him,
{\WJ{Verily
I say
unto thee,
That
this day,
{\ADD{even}}
in
this
night,
before
the cock
crow
twice,
thou shalt deny
me
thrice.}}
\VS{31}But he
spake the
more
vehemently,
If
I
should
die
with
thee, I
will
not
deny
thee in any
wise.
Likewise
also said
they
all.
\VS{32}And they
came
to a
place which
was
named
Gethsemane:
and he
saith to
his
disciples,
{\WJ{Sit ye
here,
while
I shall pray.}}
\VS{33}And he
taketh
with
him
Peter
and
James
and
John,
and
began to be sore
amazed,
and to be very
heavy;
\VS{34}And
saith unto
them,
{\WJ{My
soul
is
exceeding sorrowful
unto
death:
tarry ye
here,
and
watch.}}
\VS{35}And he went
forward a
little, and
fell
on the
ground,
and
prayed
that,
if it
were
possible, the
hour might
pass
from
him.
\VS{36}And he
said,
{\WJ{Abba,
Father,
all things
{\ADD{are}}
possible
unto thee;
take away
this
cup
from
me:
nevertheless
not
what
I
will,
but
what
thou wilt.}}
\VS{37}And he
cometh,
and
findeth
them
sleeping,
and
saith unto
Peter,
{\WJ{Simon,
sleepest thou?
couldest
not
thou watch
one
hour?}}
\VS{38}{\WJ{Watch ye
and
pray,
lest
ye enter
into
temptation.
The spirit
truly
{\ADD{is}}
ready,
but
the flesh
{\ADD{is}}
weak.}}
\VS{39}And
again he went
away, and
prayed, and
spake the
same
words.
\VS{40}And when he
returned, he
found
them
asleep
again, (for
their
eyes
were
heavy,)
neither wist
they
what to
answer
him.
\VS{41}And he
cometh the third
time,
and
saith unto
them,
{\WJ{Sleep on
now,
and
take
{\ADD{your}}
rest:
it is enough,
the hour
is come;
behold,
the Son
of man
is betrayed
into
the hands
of sinners.}}
\VS{42}{\WJ{Rise up,
let us go;
lo,
he that
betrayeth
me
is at hand.}}
\par }{\PP \VS{43}¶
And
immediately, while
he
yet
spake,
cometh
Judas,
one of the
twelve,
and
with
him a
great
multitude
with
swords
and
staves,
from the chief
priests
and the
scribes
and the
elders.
\VS{44}And he
that
betrayed
him had
given
them a
token,
saying,
Whomsoever I shall
kiss, that
same is
he;
take
him,
and
lead
{\ADD{him}}
away
safely.
\VS{45}And as soon as he was
come, he
goeth
straightway to
him, and
saith,
Master,
master;
and
kissed
him.
\par }{\PP \VS{46}¶
And they
laid
their
hands
on
him,
and
took
him.
\VS{47}And
one of them
that stood
by
drew a
sword, and
smote a
servant of the high
priest,
and cut
off
his
ear.
\VS{48}And
Jesus
answered and
said unto
them,
{\WJ{Are ye come out,
as
against
a thief,
with
swords
and
{\ADD{with}}
staves
to take
me?}}
\VS{49}{\WJ{I was
daily
with
you
in
the temple
teaching,
and
ye took
me
not:
but
the scriptures
must be fulfilled.}}
\VS{50}And
they
all
forsook
him, and
fled.
\VS{51}And there
followed
him a
certain young
man,
having a linen
cloth
cast
about
{\ADD{his}}
naked
{\ADD{body}};
and the young
men laid
hold on
him:
\VS{52}And he
left the linen
cloth, and
fled
from
them
naked.
\par }{\PP \VS{53}¶
And they
led
Jesus
away
to the high
priest:
and with
him were
assembled
all the chief
priests
and the
elders
and the
scribes.
\VS{54}And
Peter
followed
him afar
off,
even
into the
palace of the high
priest:
and he
sat
with the
servants,
and warmed
himself
at the
fire.
\VS{55}And the chief
priests
and
all the
council sought
for
witness
against
Jesus to
put
him
to
death;
and
found
none.
\VS{56}For
many bare false
witness
against
him,
but their
witness
agreed
not
together.
\VS{57}And there
arose
certain, and bare false
witness
against
him,
saying,
\VS{58}We
heard
him
say,
I will
destroy
this
temple
that is made with
hands,
and
within
three
days I will
build
another made without
hands.
\VS{59}But
neither
so
did
their
witness
agree
together.
\VS{60}And the high
priest stood
up
in the
midst, and
asked
Jesus,
saying, Answerest
thou
nothing?
what
{\ADD{is it which}}
these witness
against
thee?
\VS{61}But he held his
peace,
and
answered
nothing.
Again the high
priest
asked
him,
and
said unto
him,
Art
thou the
Christ, the
Son of the
Blessed?
\VS{62}And
Jesus
said,
{\WJ{I
am:
and
ye shall see
the Son
of man
sitting
on
the right hand
of power,
and
coming
in
the clouds
of heaven.}}
\VS{63}Then the high
priest
rent
his
clothes, and
saith,
What
need we
any
further
witnesses?
\VS{64}Ye have
heard the
blasphemy:
what
think
ye?
And
they
all
condemned
him to
be guilty
of
death.
\VS{65}And
some
began to
spit on
him,
and to
cover
his
face,
and to
buffet
him,
and to
say unto
him,
Prophesy:
and the
servants did
strike
him with the palms of their
hands.
\par }{\PP \VS{66}¶
And as
Peter
was
beneath
in the
palace, there
cometh
one of the
maids of the high
priest:
\VS{67}And when she
saw
Peter warming
himself, she
looked upon
him, and
said,
And thou
also
wast
with
Jesus of
Nazareth.
\VS{68}But he
denied,
saying, I
know
not,
neither understand
I
what
thou
sayest.
And he
went
out
into the
porch;
and the
cock
crew.
\VS{69}And a
maid
saw
him
again, and
began to
say to them
that stood
by,
This
is
{\ADD{one}}
of
them.
\VS{70}And he denied
it
again.
And a
little
after, they
that stood
by
said
again to
Peter,
Surely thou
art
{\ADD{one}}
of
them:
for thou
art a
Galilaean,
and
thy
speech
agreeth
{\ADD{thereto}}.
\VS{71}But he
began to
curse
and to
swear,
{\ADD{saying}}, I
know
not
this
man of
whom ye
speak.
\VS{72}And the second
time the
cock
crew.
And
Peter called to
mind the
word
that
Jesus
said unto
him,
{\WJ{Before
the cock
crow
twice,
thou shalt deny
me
thrice.}}
And when he thought
thereon, he
wept.

\par }\Chap{15}{\PP \VerseOne{1}And
straightway
in the
morning the chief
priests
held a
consultation
with the
elders
and
scribes
and the
whole
council, and
bound
Jesus, and
carried
{\ADD{him}}
away,
and
delivered
{\ADD{him}} to
Pilate.
\VS{2}And
Pilate
asked
him,
Art
thou the
King of the
Jews?
And he
answering
said unto
him,
{\WJ{Thou
sayest
{\ADD{it}}.}}
\VS{3}And the chief
priests
accused
him of many
things: but he answered nothing.
\VS{4}And
Pilate
asked
him
again,
saying, Answerest
thou
nothing?
behold how many
things they witness
against
thee.
\VS{5}But
Jesus
yet
answered
nothing; so
that
Pilate
marvelled.
\VS{6}Now
at
{\ADD{that}}
feast he
released unto
them
one
prisoner,
whomsoever they
desired.
\VS{7}And there
was
{\ADD{one}}
named
Barabbas,
{\ADD{which lay}}
bound
with them
that had made insurrection with
him,
who had
committed
murder
in the
insurrection.
\VS{8}And the
multitude crying
aloud
began to
desire
{\ADD{him to do}}
as he
had
ever
done unto
them.
\VS{9}But
Pilate
answered
them,
saying, Will
ye that I
release unto
you the
King of the
Jews?
\VS{10}For he
knew
that the chief
priests had
delivered
him
for
envy.
\VS{11}But the chief
priests
moved the
people,
that he
should
rather
release
Barabbas unto
them.
\VS{12}And
Pilate
answered and
said
again unto
them,
What will
ye
then that I shall
do
{\ADD{unto him}}
whom ye
call the
King of the
Jews?
\VS{13}And they cried
out
again,
Crucify
him.
\VS{14}Then
Pilate
said unto
them,
Why,
what
evil hath he
done?
And they cried
out the more
exceedingly,
Crucify
him.
\par }{\PP \VS{15}¶
And
{\ADD{so}}
Pilate,
willing to
content the
people,
released
Barabbas unto
them,
and
delivered
Jesus, when he had
scourged
{\ADD{him}},
to be
crucified.
\VS{16}And the
soldiers
led
him
away
into the
hall,
called
Prætorium;
and they call
together the
whole
band.
\VS{17}And they
clothed
him with
purple, and
platted a
crown of
thorns,
and put it
about
his
{\ADD{head}},
\VS{18}And
began to
salute
him,
Hail,
King of the
Jews!
\VS{19}And they
smote
him on the
head with a
reed,
and did
spit upon
him,
and
bowing
{\ADD{their}}
knees
worshipped
him.
\VS{20}And
when they had
mocked
him, they took
off the
purple
from
him,
and
put his
own
clothes on
him,
and
led
him
out
to
crucify
him.
\VS{21}And they
compel
one
Simon a
Cyrenian, who passed
by, coming
out
of the
country, the
father of
Alexander
and
Rufus,
to
bear
his
cross.
\VS{22}And they
bring
him
unto the
place
Golgotha,
which
is, being
interpreted, The
place of a
skull.
\VS{23}And they
gave
him to
drink
wine mingled with
myrrh:
but he
received
{\ADD{it}}
not.
\VS{24}And when they had
crucified
him, they
parted
his
garments,
casting
lots
upon
them,
what every
man should
take.
\VS{25}And it
was the
third
hour,
and they
crucified
him.
\VS{26}And the
superscription of
his
accusation
was written
over, THE
KING OF THE
JEWS.
\VS{27}And
with
him they
crucify
two
thieves; the
one
on his right
hand,
and the
other
on
his
left.
\VS{28}And the
scripture was
fulfilled,
which
saith,
And he was
numbered
with the
transgressors.
\VS{29}And they
that passed
by railed
on
him,
wagging
their
heads,
and
saying,
Ah, thou
that
destroyest the
temple,
and
buildest
{\ADD{it}}
in
three
days,
\VS{30}Save
thyself,
and come
down
from the
cross.
\VS{31}Likewise
also the chief
priests
mocking
said
among
themselves
with the
scribes, He
saved
others;
himself he
cannot
save.
\VS{32}Let
Christ the
King of
Israel
descend
now
from the
cross,
that we may
see
and
believe.
And they
that were
crucified with
him
reviled
him.
\VS{33}And when the
sixth
hour was
come, there
was
darkness
over the
whole
land
until the
ninth
hour.
\VS{34}And at the
ninth
hour
Jesus
cried with a
loud
voice,
saying,
{\WJ{Eloi,
Eloi,
lama
sabachthani?}}
which
is, being
interpreted,
My
God,
my
God,
why hast thou
forsaken
me?
\VS{35}And
some of them
that stood
by, when they
heard
{\ADD{it}},
said,
Behold, he
calleth
Elias.
\VS{36}And
one
ran
and
filled a
spunge
full of
vinegar,
and
put
{\ADD{it}} on a
reed, and
gave
him to
drink,
saying, Let
alone; let us
see
whether
Elias will
come to
take
him
down.
\VS{37}And
Jesus
cried with a
loud
voice, and gave up the
ghost.
\VS{38}And the
veil of the
temple was
rent
in
twain
from the
top
to the
bottom.
\par }{\PP \VS{39}¶
And when the
centurion,
which
stood
over
against
him,
saw
that
he
so cried
out, and gave up the
ghost, he
said,
Truly
this
man
was the
Son of
God.
\VS{40}There
were
also
women looking
on
afar
off:
among
whom
was
Mary
Magdalene,
and
Mary the
mother of
James the
less
and of
Joses,
and
Salome;
\VS{41}(Who
also,
when he
was
in
Galilee,
followed
him,
and
ministered unto
him;)
and
many other
women
which came
up with
him
unto
Jerusalem.
\par }{\PP \VS{42}¶
And
now when the
even was
come,
because it
was the
preparation,
that
is, the day before the
sabbath,
\VS{43}Joseph
of
Arimathaea, an
honourable
counsellor,
which
also waited
for the
kingdom of
God,
came, and went
in
boldly
unto
Pilate,
and
craved the
body of
Jesus.
\VS{44}And
Pilate
marvelled
if he
were
already
dead:
and
calling
{\ADD{unto him}} the
centurion, he
asked
him
whether he had
been any
while
dead.
\VS{45}And when he
knew
{\ADD{it}}
of the
centurion, he
gave the
body to
Joseph.
\VS{46}And he
bought fine
linen,
and
took
him
down, and wrapped him
in the
linen,
and
laid
him
in a
sepulchre
which
was
hewn out
of a
rock,
and
rolled a
stone
unto the
door of the
sepulchre.
\VS{47}And
Mary
Magdalene
and
Mary
{\ADD{the mother}} of
Joses
beheld
where he was
laid.

\par }\Chap{16}{\PP \VerseOne{1}And when the
sabbath was
past,
Mary
Magdalene,
and
Mary
the
{\ADD{mother}} of
James,
and
Salome, had
bought sweet
spices,
that they might
come and
anoint
him.
\VS{2}And
very early in the
morning the
first
{\ADD{day}} of the
week, they
came
unto the
sepulchre at the
rising of the
sun.
\VS{3}And they
said
among
themselves,
Who shall
roll
us
away the
stone
from the
door of the
sepulchre?
\VS{4}And when they
looked, they
saw
that the
stone was rolled
away:
for it
was
very
great.
\VS{5}And
entering
into the
sepulchre, they
saw a young
man
sitting
on the right
side,
clothed in a
long
white
garment;
and they were
affrighted.
\VS{6}And he
saith unto
them,
Be
not
affrighted: Ye
seek
Jesus of
Nazareth,
which was
crucified: he is
risen; he
is
not
here:
behold the
place
where they
laid
him.
\VS{7}But go your
way,
tell
his
disciples
and
Peter
that he goeth
before
you
into
Galilee:
there shall ye
see
him,
as he
said unto
you.
\VS{8}And they went
out
quickly, and
fled
from the
sepulchre;
for
they
trembled
and were
amazed:
neither said
they any
thing to
any
{\ADD{man}};
for they were
afraid.
\par }{\PP \VS{9}¶ Now
when
{\ADD{Jesus}} was
risen
early the
first
{\ADD{day}} of the
week, he
appeared
first to
Mary
Magdalene, out
of
whom he had
cast
seven
devils.
\VS{10}{\ADD{And}} she
went and
told them
that had
been
with
him, as they
mourned
and
wept.
\VS{11}And
they, when they had
heard
that he was
alive,
and had been
seen
of
her, believed
not.
\par }{\PP \VS{12}¶
After
that he
appeared
in
another
form unto
two
of
them, as they
walked, and
went
into the
country.
\VS{13}And
they
went and
told
{\ADD{it}} unto the
residue:
neither believed
they
them.
\par }{\PP \VS{14}¶
Afterward he
appeared unto the
eleven as
they sat at
meat,
and
upbraided them with
their
unbelief
and hardness of
heart,
because they
believed
not them
which had
seen
him after he was
risen.
\VS{15}And he
said unto
them,
{\WJ{Go ye
into
all
the world,
and preach
the gospel
to every
creature.}}
\VS{16}{\WJ{He
that believeth
and
is baptized
shall be saved;
but
he
that believeth not
shall be damned.}}
\VS{17}{\WJ{And
these
signs
shall follow
them
that believe;
In
my
name
shall they cast out
devils;
they shall speak
with new
tongues;}}
\VS{18}{\WJ{They shall take up
serpents;
and if
they drink
any
deadly thing,
it shall
not
hurt
them;
they shall lay
hands
on
the sick,
and
they shall
recover.}}
\par }{\PP \VS{19}¶
So
then
after the
Lord had
spoken unto
them, he was received
up
into
heaven,
and
sat
on the right
hand of
God.
\VS{20}And
they went
forth, and
preached every
where, the
Lord working
with
{\ADD{them}},
and
confirming the
word
with
signs
following.
Amen.
\par }