\NormalFont\ShortTitle{Nahum}
{\MT Nahum

\par }\ChapOne{1}{\PP \VerseOne{1}The
burden of
Nineveh. The
book of the
vision of
Nahum the
Elkoshite.
\par }{\PP \VS{2}God
{\ADD{is}}
jealous, and the
{\ND{
LORD}}
revengeth; the
{\ND{
LORD}}
revengeth, and
{\ADD{is}}
furious; the
{\ND{
LORD}} will take
vengeance on his
adversaries, and he
reserveth
{\ADD{wrath}} for his
enemies.\FTNT{*}{{\FR 1.2 }
God…: or, The
{\ND{LORD}} is a jealous God, and a revenger, etc}\FTNT{†}{{\FR 1.2 }
is furious: Heb. that hath fury}
\VS{3}The
{\ND{
LORD}}
{\ADD{is}}
slow to
anger, and
great in
power, and will not at
all
acquit
{\ADD{the wicked:}} the
{\ND{
LORD}} hath his
way in the
whirlwind and in the
storm, and the
clouds
{\ADD{are}} the
dust of his
feet.
\VS{4}He
rebuketh the
sea, and maketh it
dry, and drieth
up all the
rivers:
Bashan
languisheth, and
Carmel, and the
flower of
Lebanon
languisheth.
\VS{5}The
mountains
quake at him, and the
hills
melt, and the
earth is
burned at his
presence, yea, the
world, and all that
dwell therein.
\VS{6}Who can
stand
before his
indignation? and who can
abide in the
fierceness of his
anger? his
fury is poured
out like
fire, and the
rocks are thrown
down by him.\FTNT{‡}{{\FR 1.6 }
abide: Heb. stand up}
\VS{7}The
{\ND{
LORD}}
{\ADD{is}}
good, a strong
hold in the
day of
trouble; and he
knoweth them that
trust in him.\FTNT{§}{{\FR 1.7 }
strong hold: or, strength}
\VS{8}But with an
overrunning
flood he will
make an utter
end of the
place thereof, and
darkness shall
pursue his
enemies.
\par }{\PP \VS{9}What do ye
imagine against the
{\ND{
LORD}}? he will
make an utter
end:
affliction shall not rise
up the second
time.
\VS{10}For while
{\ADD{they be}} folden
together
{\ADD{as}}
thorns, and while they are
drunken
{\ADD{as}}
drunkards, they shall be
devoured as
stubble
fully
dry.
\VS{11}There is
{\ADD{one}} come
out of thee, that
imagineth
evil against the
{\ND{
LORD}}, a
wicked
counsellor.\FTNT{**}{{\FR 1.11 }
a wicked…: Heb. a counsellor of Belial}
\VS{12}Thus
saith the
{\ND{
LORD}}; Though
{\ADD{they be}}
quiet, and likewise
many, yet thus shall they be cut
down, when he shall pass
through. Though I have
afflicted thee, I will
afflict thee no more.\FTNT{††}{{\FR 1.12 }
Though…: or, If they would have been at peace, so should they have been many, and so should they have been shorn, and he should have passed away}\FTNT{‡‡}{{\FR 1.12 }
cut down: Heb. shorn}
\VS{13}For now will I
break his
yoke from off thee, and will
burst thy
bonds in
sunder.
\VS{14}And the
{\ND{
LORD}} hath given a
commandment concerning thee,
{\ADD{that}} no more of thy
name be
sown: out of the
house of thy
gods will I cut
off the graven
image and the molten
image: I will
make thy
grave; for thou art
vile.
\VS{15}Behold upon the
mountains the
feet of him that bringeth good
tidings, that
publisheth
peace! O
Judah,
keep thy solemn
feasts,
perform thy
vows: for the
wicked shall no
more pass
through thee; he is utterly cut
off.\FTNT{§§}{{\FR 1.15 }
keep…: Heb. feast}\FTNT{***}{{\FR 1.15 }
the wicked: Heb. Belial}

\par }\Chap{2}{\PP \VerseOne{1}He that dasheth in
pieces is come
up before thy
face:
keep the
munition,
watch the
way, make
{\ADD{thy}}
loins
strong,
fortify
{\ADD{thy}}
power
mightily.\FTNT{*}{{\FR 2.1 }
He…: or, The disperser, or, hammer}
\VS{2}For the
{\ND{
LORD}} hath turned
away the
excellency of
Jacob, as the
excellency of
Israel: for the
emptiers have emptied them
out, and
marred their vine
branches.\FTNT{†}{{\FR 2.2 }
the excellency of Jacob…: or, the pride of Jacob as the pride, etc}
\VS{3}The
shield of his mighty
men is made
red, the
valiant
men
{\ADD{are}} in
scarlet: the
chariots
{\ADD{shall be}} with
flaming
torches in the
day of his
preparation, and the fir
trees shall be terribly
shaken.\FTNT{‡}{{\FR 2.3 }
in scarlet: or, dyed scarlet}\FTNT{§}{{\FR 2.3 }
flaming: or, fiery}
\VS{4}The
chariots shall
rage in the
streets, they shall justle one against
another in the broad
ways: they shall
seem like
torches, they shall
run like the
lightnings.\FTNT{**}{{\FR 2.4 }
they: Heb. their show}
\VS{5}He shall
recount his
worthies: they shall
stumble in their
walk; they shall make
haste to the
wall thereof, and the
defence shall be
prepared.\FTNT{††}{{\FR 2.5 }
worthies: or, gallants}\FTNT{‡‡}{{\FR 2.5 }
defence: Heb. covering, or, coverer}
\VS{6}The
gates of the
rivers shall be
opened, and the
palace shall be
dissolved.\FTNT{§§}{{\FR 2.6 }
dissolved: or, molten}
\VS{7}And
Huzzab shall be led away
captive, she shall be brought
up, and her
maids shall
lead
{\ADD{her}} as with the
voice of
doves,
tabering upon their
breasts.\FTNT{***}{{\FR 2.7 }
Huzzab: or, that which was established, or, there was a stand made}\FTNT{†††}{{\FR 2.7 }
led…: or, discovered}
\VS{8}But
Nineveh
{\ADD{is}} of
old like a
pool of
water: yet they shall flee
away.
Stand,
stand,
{\ADD{shall they cry;}} but none shall look
back.\FTNT{‡‡‡}{{\FR 2.8 }
of old: or, from the days that she hath been}\FTNT{§§§}{{\FR 2.8 }
look back: or, cause them to turn}
\VS{9}Take ye the
spoil of
silver, take the
spoil of
gold: for
{\ADD{there is}} none
end of the
store
{\ADD{and}}
glory out of all the
pleasant
furniture.\FTNT{*}{{\FR 2.9 }
for…: or, and their infinite store, etc}\FTNT{†}{{\FR 2.9 }
pleasant…: Heb. vessels of desire}
\VS{10}She is
empty, and
void, and
waste: and the
heart
melteth, and the
knees smite
together, and much
pain
{\ADD{is}} in all
loins, and the
faces of them all
gather
blackness.
\par }{\PP \VS{11}Where
{\ADD{is}} the
dwelling of the
lions, and the
feedingplace of the young
lions, where the
lion,
{\ADD{even}} the old
lion,
walked,
{\ADD{and}} the
lion’s
whelp, and none made
{\ADD{them}}
afraid?
\VS{12}The
lion did tear in
pieces
enough for his
whelps, and
strangled for his
lionesses, and
filled his
holes with
prey, and his
dens with
ravin.
\VS{13}Behold, I
{\ADD{am}} against thee,
saith the
{\ND{
LORD}} of
hosts, and I will
burn her
chariots in the
smoke, and the
sword shall
devour thy young
lions: and I will cut
off thy
prey from the
earth, and the
voice of thy
messengers shall no more be
heard.

\par }\Chap{3}{\PP \VerseOne{1}Woe to the
bloody
city! it
{\ADD{is}} all
full of
lies
{\ADD{and}}
robbery; the
prey
departeth not;\FTNT{*}{{\FR 3.1 }
bloody…: Heb. city of bloods}
\VS{2}The
noise of a
whip, and the
noise of the
rattling of the
wheels, and of the
pransing
horses, and of the
jumping
chariots.
\VS{3}The
horseman lifteth
up both the
bright
sword and the
glittering
spear: and
{\ADD{there is}} a
multitude of
slain, and a
great number of
carcases; and
{\ADD{there is}} none
end of
{\ADD{their}}
corpses; they
stumble upon their
corpses:\FTNT{†}{{\FR 3.3 }
the bright…: Heb. the flame of the sword, and the lightning of the spear}
\VS{4}Because of the
multitude of the
whoredoms of the
wellfavoured
harlot, the
mistress of
witchcrafts, that
selleth
nations through her
whoredoms, and
families through her
witchcrafts.
\VS{5}Behold, I
{\ADD{am}} against thee,
saith the
{\ND{
LORD}} of
hosts; and I will
discover thy
skirts upon thy
face, and I will
shew the
nations thy
nakedness, and the
kingdoms thy
shame.
\VS{6}And I will
cast abominable
filth upon thee, and make thee
vile, and will
set thee as a
gazingstock.
\VS{7}And it shall come to pass,
{\ADD{that}} all they that
look upon thee shall
flee from thee, and
say,
Nineveh is laid
waste: who will
bemoan her?
whence shall I
seek
comforters for thee?
\par }{\PP \VS{8}Art thou
better than
populous
No, that was
situate among the
rivers,
{\ADD{that had}} the
waters round
about it, whose
rampart
{\ADD{was}} the
sea,
{\ADD{and}} her
wall
{\ADD{was}} from the
sea?\FTNT{‡}{{\FR 3.8 }
populous…: or, nourishing, etc: Heb. No Amon}
\VS{9}Ethiopia and
Egypt
{\ADD{were}} her
strength, and
{\ADD{it was}}
infinite;
Put and
Lubim were thy
helpers.\FTNT{§}{{\FR 3.9 }
thy helpers: Heb. in thy help}
\VS{10}Yet
{\ADD{was}} she carried
away, she
went into
captivity: her young
children also were dashed in
pieces at the
top of all the
streets: and they
cast
lots for her honourable
men, and all her great
men were
bound in
chains.
\VS{11}Thou also shalt be
drunken: thou shalt be
hid, thou also shalt
seek
strength because of the
enemy.
\VS{12}All thy strong
holds
{\ADD{shall be like}} fig
trees with the firstripe
figs: if they be
shaken, they shall even
fall into the
mouth of the
eater.
\VS{13}Behold, thy
people in the
midst of thee
{\ADD{are}}
women: the
gates of thy
land shall be set
wide
open unto thine
enemies: the
fire shall
devour thy
bars.
\VS{14}Draw thee
waters for the
siege,
fortify thy strong
holds:
go into
clay, and
tread the
morter, make
strong the
brickkiln.
\VS{15}There shall the
fire
devour thee; the
sword shall cut thee
off, it shall eat thee
up like the
cankerworm: make thyself
many as the
cankerworm, make thyself
many as the
locusts.
\VS{16}Thou hast
multiplied thy
merchants above the
stars of
heaven: the
cankerworm
spoileth, and flieth
away.\FTNT{**}{{\FR 3.16 }
spoileth: or, spreadeth himself}
\VS{17}Thy
crowned
{\ADD{are}} as the
locusts, and thy
captains as the great
grasshoppers, which
camp in the
hedges in the
cold
day,
{\ADD{but}} when the
sun
ariseth they flee
away, and their
place is not
known
where they
{\ADD{are}}.
\VS{18}Thy
shepherds
slumber, O
king of
Assyria: thy
nobles shall
dwell
{\ADD{in the dust:}} thy
people is
scattered upon the
mountains, and no man
gathereth
{\ADD{them}}.\FTNT{††}{{\FR 3.18 }
nobles: or, valiant ones}
\VS{19}{\ADD{There is}} no
healing of thy
bruise; thy
wound is
grievous: all that
hear the
bruit of thee shall
clap the
hands over thee: for upon whom hath not thy
wickedness
passed
continually?\FTNT{‡‡}{{\FR 3.19 }
healing: Heb. wrinkling}
\par }