\NormalFont\ShortTitle{Nehemiah}
{\MT The Book of Nehemiah

\par }\ChapOne{1}{\PP \VerseOne{1}The
words of
Nehemiah the
son of
Hachaliah. And it came to pass in the
month
Chisleu, in the
twentieth
year, as I was in
Shushan the
palace,
\VS{2}That
Hanani,
one of my
brethren,
came, he and
{\ADD{certain}}
men of
Judah; and I
asked them concerning the
Jews that had
escaped, which were
left of the
captivity, and concerning
Jerusalem.
\VS{3}And they
said unto me, The
remnant that are
left of the
captivity there in the
province
{\ADD{are}} in
great
affliction and
reproach: the
wall of
Jerusalem also
{\ADD{is}} broken
down, and the
gates thereof are
burned with
fire.
\par }{\PP \VS{4}¶ And it came to pass, when I
heard these
words, that I sat
down and
wept, and
mourned
{\ADD{certain}}
days, and
fasted, and
prayed
before the
God of
heaven,
\par }{\PP \VS{5}And
said, I
beseech thee, O
{\ND{
LORD}}
God of
heaven, the
great and
terrible
God, that
keepeth
covenant and
mercy for them that
love him and
observe his
commandments:
\VS{6}Let thine
ear now be
attentive, and thine
eyes
open, that thou mayest
hear the
prayer of thy
servant, which I
pray
before thee
now,
day and
night, for the
children of
Israel thy
servants, and
confess the
sins of the
children of
Israel, which we have
sinned against thee: both I and my
father’s
house have
sinned.
\VS{7}We have dealt
very
corruptly against thee, and have not
kept the
commandments, nor the
statutes, nor the
judgments, which thou
commandedst thy
servant
Moses.
\VS{8}Remember, I beseech thee, the
word that thou
commandedst thy
servant
Moses,
saying,
{\ADD{If}} ye
transgress, I will scatter you
abroad among the
nations:
\VS{9}But
{\ADD{if}} ye
turn unto me, and
keep my
commandments, and
do them; though there were of you cast
out unto the uttermost
part of the
heaven,
{\ADD{yet}} will I
gather them from thence, and will
bring them unto the
place that I have
chosen to
set my
name there.
\VS{10}Now these
{\ADD{are}} thy
servants and thy
people, whom thou hast
redeemed by thy
great
power, and by thy
strong
hand.
\VS{11}O
Lord, I
beseech thee, let now thine
ear be
attentive to the
prayer of thy
servant, and to the
prayer of thy
servants, who
desire to
fear thy
name: and
prosper, I pray thee, thy
servant this
day, and
grant him
mercy in the
sight of this
man. For I was the
king’s
cupbearer.

\par }\Chap{2}{\PP \VerseOne{1}And it came to pass in the
month
Nisan, in the
twentieth
year of
Artaxerxes the
king,
{\ADD{that}}
wine
{\ADD{was}}
before him: and I took
up the
wine, and
gave
{\ADD{it}} unto the
king. Now I had not been
{\ADD{beforetime}}
sad in his
presence.
\VS{2}Wherefore the
king
said unto me, Why
{\ADD{is}} thy
countenance
sad, seeing thou
{\ADD{art}} not
sick? this
{\ADD{is}} nothing
{\ADD{else}} but
sorrow of
heart. Then I was
very
sore
afraid,
\VS{3}And
said unto the
king, Let the
king
live for
ever: why should not my
countenance be
sad, when the
city, the
place of my
fathers’
sepulchres,
{\ADD{lieth}}
waste, and the
gates thereof are
consumed with
fire?
\VS{4}Then the
king
said unto me, For what dost thou make
request? So I
prayed to the
God of
heaven.
\VS{5}And I
said unto the
king, If it
please the
king, and if thy
servant have found
favour in thy
sight, that thou wouldest
send me unto
Judah, unto the
city of my
fathers’
sepulchres, that I may
build it.
\VS{6}And the
king
said unto me, (the
queen also
sitting by
him,) For how long shall thy
journey be? and when wilt thou
return? So it
pleased the
king to
send me; and I
set him a
time.\FTNT{*}{{\FR 2.6 }
queen: Heb. wife}
\VS{7}Moreover I
said unto the
king, If it
please the
king, let
letters be
given me to the
governors
beyond the
river, that they may convey me
over till I
come into
Judah;
\VS{8}And a
letter unto
Asaph the
keeper of the
king’s
forest, that he may
give me
timber to make
beams for the
gates of the
palace which
{\ADD{appertained}} to the
house, and for the
wall of the
city, and for the
house that I shall enter
into. And the
king
granted me, according to the
good
hand of my
God upon me.
\par }{\PP \VS{9}¶ Then I
came to the
governors
beyond the
river, and
gave them the
king’s
letters. Now the
king had
sent
captains of the
army and
horsemen with me.
\VS{10}When
Sanballat the
Horonite, and
Tobiah the
servant, the
Ammonite,
heard
{\ADD{of it}}, it
grieved them
exceedingly that there was
come a
man to
seek the
welfare of the
children of
Israel.
\VS{11}So I
came to
Jerusalem, and was there
three
days.
\par }{\PP \VS{12}¶ And I
arose in the
night, I and
some
few
men with me; neither
told I
{\ADD{any}}
man what my
God had
put in my
heart to
do at
Jerusalem: neither
{\ADD{was there any}}
beast with me, save the
beast that I rode
upon.
\VS{13}And I went
out by
night by the
gate of the
valley, even
before the
dragon
well, and to the
dung
port, and
viewed the
walls of
Jerusalem, which were broken
down, and the
gates thereof were
consumed with
fire.
\VS{14}Then I went
on to the
gate of the
fountain, and to the
king’s
pool: but
{\ADD{there was}} no
place for the
beast
{\ADD{that was}} under me to
pass.
\VS{15}Then went I
up in the
night by the
brook, and
viewed the
wall, and turned
back, and
entered by the
gate of the
valley, and
{\ADD{so}}
returned.
\VS{16}And the
rulers
knew not whither I
went, or what I
did; neither had I as
yet
told
{\ADD{it}} to the
Jews, nor to the
priests, nor to the
nobles, nor to the
rulers, nor to the
rest that
did the
work.
\par }{\PP \VS{17}¶ Then
said I unto them, Ye
see the
distress that we
{\ADD{are}} in, how
Jerusalem
{\ADD{lieth}}
waste, and the
gates thereof are
burned with
fire:
come, and let us build
up the
wall of
Jerusalem, that we be no more a
reproach.
\VS{18}Then I
told them of the
hand of my
God which was
good upon me; as also the
king’s
words that he had
spoken unto me. And they
said, Let us rise
up and
build. So they
strengthened their
hands for
{\ADD{this}}
good
{\ADD{work}}.
\VS{19}But when
Sanballat the
Horonite, and
Tobiah the
servant, the
Ammonite, and
Geshem the
Arabian,
heard
{\ADD{it}}, they laughed us to
scorn, and
despised us, and
said, What
{\ADD{is}} this
thing that ye
do? will ye
rebel against the
king?
\VS{20}Then
answered I them, and
said unto them, The
God of
heaven, he will
prosper us; therefore we his
servants will
arise and
build: but ye have no
portion, nor
right, nor
memorial, in
Jerusalem.

\par }\Chap{3}{\PP \VerseOne{1}Then
Eliashib the
high
priest rose
up with his
brethren the
priests, and they
builded the
sheep
gate; they
sanctified it, and set
up the
doors of it; even unto the
tower of
Meah they
sanctified it, unto the
tower of
Hananeel.
\VS{2}And next unto
him
builded the
men of
Jericho. And next to them
builded
Zaccur the
son of
Imri.\FTNT{*}{{\FR 3.2 }
next unto…: Heb. at his hand}
\VS{3}But the
fish
gate did the
sons of
Hassenaah
build, who
{\ADD{also}} laid the
beams thereof, and set
up the
doors thereof, the
locks thereof, and the
bars thereof.
\VS{4}And next unto
them
repaired
Meremoth the
son of
Urijah, the
son of
Koz. And next unto
them
repaired
Meshullam the
son of
Berechiah, the
son of
Meshezabeel. And next unto
them
repaired
Zadok the
son of
Baana.
\VS{5}And next unto
them the
Tekoites
repaired; but their
nobles
put not their
necks to the
work of their
Lord.
\VS{6}Moreover the
old
gate
repaired
Jehoiada the
son of
Paseah, and
Meshullam the
son of
Besodeiah; they laid the
beams thereof, and set
up the
doors thereof, and the
locks thereof, and the
bars thereof.
\VS{7}And next unto
them
repaired
Melatiah the
Gibeonite, and
Jadon the
Meronothite, the
men of
Gibeon, and of
Mizpah, unto the
throne of the
governor on this
side the
river.
\VS{8}Next unto
him
repaired
Uzziel the
son of
Harhaiah, of the
goldsmiths. Next unto
him also
repaired
Hananiah the
son of
{\ADD{one of}} the
apothecaries, and they
fortified
Jerusalem unto the
broad
wall.\FTNT{†}{{\FR 3.8 }
fortified…: or, left Jerusalem unto the broad wall}
\VS{9}And next unto
them
repaired
Rephaiah the
son of
Hur, the
ruler of the
half
part of
Jerusalem.
\VS{10}And next unto
them
repaired
Jedaiah the
son of
Harumaph, even over against his
house. And next unto
him
repaired
Hattush the
son of
Hashabniah.
\VS{11}Malchijah the
son of
Harim, and
Hashub the
son of
Pahath-moab,
repaired the
other
piece, and the
tower of the
furnaces.\FTNT{‡}{{\FR 3.11 }
other…: Heb. second measure}
\VS{12}And next unto
him
repaired
Shallum the
son of
Halohesh, the
ruler of the
half
part of
Jerusalem, he and his
daughters.
\VS{13}The
valley
gate
repaired
Hanun, and the
inhabitants of
Zanoah; they
built it, and set
up the
doors thereof, the
locks thereof, and the
bars thereof, and a
thousand
cubits on the
wall unto the
dung
gate.
\VS{14}But the
dung
gate
repaired
Malchiah the
son of
Rechab, the
ruler of
part of
Beth-haccerem; he
built it, and set
up the
doors thereof, the
locks thereof, and the
bars thereof.
\VS{15}But the
gate of the
fountain
repaired
Shallun the
son of
Col-hozeh, the
ruler of
part of
Mizpah; he
built it, and
covered it, and set
up the
doors thereof, the
locks thereof, and the
bars thereof, and the
wall of the
pool of
Siloah by the
king’s
garden, and unto the
stairs that go
down from the
city of
David.
\VS{16}After him
repaired
Nehemiah the
son of
Azbuk, the
ruler of the
half
part of
Beth-zur, unto
{\ADD{the place}} over against the
sepulchres of
David, and to the
pool that was
made, and unto the
house of the
mighty.
\VS{17}After him
repaired the
Levites,
Rehum the
son of
Bani. Next unto
him
repaired
Hashabiah, the
ruler of the
half
part of
Keilah, in his
part.
\VS{18}After him
repaired their
brethren,
Bavai the
son of
Henadad, the
ruler of the
half
part of
Keilah.
\VS{19}And next to
him
repaired
Ezer the
son of
Jeshua, the
ruler of
Mizpah,
another
piece over against the going
up to the
armoury at the
turning
{\ADD{of the wall}}.
\VS{20}After him
Baruch the
son of
Zabbai
earnestly
repaired the
other
piece, from the
turning
{\ADD{of the wall}} unto the
door of the
house of
Eliashib the
high
priest.\FTNT{§}{{\FR 3.20 }
Zabbai: also called, Zaccai}
\VS{21}After him
repaired
Meremoth the
son of
Urijah the
son of
Koz
another
piece, from the
door of the
house of
Eliashib even to the
end of the
house of
Eliashib.
\VS{22}And
after him
repaired the
priests, the
men of the
plain.
\VS{23}After him
repaired
Benjamin and
Hashub over against their
house.
After him
repaired
Azariah the
son of
Maaseiah the
son of
Ananiah
by his
house.
\VS{24}After him
repaired
Binnui the
son of
Henadad
another
piece, from the
house of
Azariah unto the
turning
{\ADD{of the wall}}, even unto the
corner.
\VS{25}Palal the
son of
Uzai, over against the
turning
{\ADD{of the wall}}, and the
tower which lieth
out from the
king’s
high
house, that
{\ADD{was}} by the
court of the
prison.
After him
Pedaiah the
son of
Parosh.
\VS{26}Moreover the
Nethinims
dwelt in
Ophel, unto
{\ADD{the place}} over against the
water
gate toward the
east, and the
tower that lieth
out.\FTNT{**}{{\FR 3.26 }
dwelt…: or, which dwelt in Ophel, repaired unto}\FTNT{††}{{\FR 3.26 }
Ophel: or, tower}
\VS{27}After them the
Tekoites
repaired
another
piece, over against the
great
tower that lieth
out, even unto the
wall of
Ophel.
\VS{28}From above the
horse
gate
repaired the
priests, every
one over
against his
house.
\VS{29}After them
repaired
Zadok the
son of
Immer over against his
house.
After him
repaired also
Shemaiah the
son of
Shechaniah, the
keeper of the
east
gate.
\VS{30}After him
repaired
Hananiah the
son of
Shelemiah, and
Hanun the
sixth
son of
Zalaph,
another
piece.
After him
repaired
Meshullam the
son of
Berechiah over against his
chamber.
\VS{31}After him
repaired
Malchiah the
goldsmith’s
son unto the
place of the
Nethinims, and of the
merchants, over against the
gate
Miphkad, and to the going
up of the
corner.\FTNT{‡‡}{{\FR 3.31 }
going…: or, corner chamber}
\VS{32}And between the going
up of the
corner unto the
sheep
gate
repaired the
goldsmiths and the
merchants.\FTNT{§§}{{\FR 3.32 }
going…: or, corner chamber}

\par }\Chap{4}{\PP \VerseOne{1}But it came to pass, that when
Sanballat
heard that we
builded the
wall, he was
wroth, and took
great
indignation, and
mocked the
Jews.
\VS{2}And he
spake
before his
brethren and the
army of
Samaria, and
said, What
do these
feeble
Jews? will they
fortify themselves? will they
sacrifice? will they make an
end in a
day? will they
revive the
stones out of the
heaps of the
rubbish which are
burned?\FTNT{*}{{\FR 4.2 }
fortify…: Heb. leave to themselves}
\VS{3}Now
Tobiah the
Ammonite
{\ADD{was}} by
him, and he
said, Even that which they
build, if a
fox go
up, he shall even break
down their
stone
wall.
\VS{4}Hear, O our
God; for we are
despised: and
turn their
reproach upon their own
head, and
give them for a
prey in the
land of
captivity:\FTNT{†}{{\FR 4.4 }
despised: Heb. despite}
\VS{5}And
cover not their
iniquity, and let not their
sin be blotted
out from
before thee: for they have provoked
{\ADD{thee}} to
anger before the
builders.
\VS{6}So
built we the
wall; and all the
wall was joined
together unto the
half thereof: for the
people had a
mind to
work.
\par }{\PP \VS{7}¶ But it came to pass,
{\ADD{that}} when
Sanballat, and
Tobiah, and the
Arabians, and the
Ammonites, and the
Ashdodites,
heard that the
walls of
Jerusalem were made
up,
{\ADD{and}} that the
breaches
began to be
stopped, then they were
very
wroth,
\FTNT{‡}{{\FR 4.7 }
were made up: Heb. ascended}
\VS{8}And
conspired all of them
together to
come
{\ADD{and}} to
fight against
Jerusalem, and to
hinder it.\FTNT{§}{{\FR 4.8 }
to hinder…: Heb. to make an error to it}
\VS{9}Nevertheless we made our
prayer unto our
God, and
set a
watch against them
day and
night,
because of them.
\VS{10}And
Judah
said, The
strength of the bearers of
burdens is
decayed, and
{\ADD{there is}}
much
rubbish; so that we are not
able to
build the
wall.
\VS{11}And our
adversaries
said, They shall not
know, neither
see, till we
come in the
midst
among them, and
slay them, and cause the
work to
cease.
\VS{12}And it came to pass, that when the
Jews which
dwelt by
them
came, they
said unto us
ten
times, From all
places whence ye shall
return unto us
{\ADD{they will be upon you}}.\FTNT{**}{{\FR 4.12 }
From all…: or, That from all places ye must return to us}
\par }{\PP \VS{13}¶ Therefore
set I in the
lower
places
behind the
wall,
{\ADD{and}} on the higher
places, I even
set the
people after their
families with their
swords, their
spears, and their
bows.\FTNT{††}{{\FR 4.13 }
in the…: Heb. from the lower parts of the place, etc}
\VS{14}And I
looked, and rose
up, and
said unto the
nobles, and to the
rulers, and to the
rest of the
people, Be not ye
afraid of
them:
remember the
Lord,
{\ADD{which is}}
great and
terrible, and
fight for your
brethren, your
sons, and your
daughters, your
wives, and your
houses.
\VS{15}And it came to pass, when our
enemies
heard that it was
known unto us, and
God had
brought their
counsel to
nought, that we
returned all of us to the
wall, every
one unto his
work.
\par }{\PP \VS{16}And it came to pass from that
time forth,
{\ADD{that}} the
half of my
servants
wrought in the
work, and the other
half of them
held both the
spears, the
shields, and the
bows, and the
habergeons; and the
rulers
{\ADD{were}}
behind all the
house of
Judah.
\VS{17}They which
builded on the
wall, and they that
bare
burdens, with those that
laded,
{\ADD{every one}} with
one of his
hands
wrought in the
work, and with the
other
{\ADD{hand}}
held a
weapon.
\VS{18}For the
builders, every
one had his
sword
girded by his
side, and
{\ADD{so}}
builded. And he that
sounded the
trumpet
{\ADD{was}} by
me.\FTNT{‡‡}{{\FR 4.18 }
by his side: Heb. on his loins}
\par }{\PP \VS{19}¶ And I
said unto the
nobles, and to the
rulers, and to the
rest of the
people, The
work
{\ADD{is}}
great and
large, and we are
separated upon the
wall,
one far
from
another.
\VS{20}In what
place
{\ADD{therefore}} ye
hear the
sound of the
trumpet,
resort ye thither unto us: our
God shall
fight for us.
\VS{21}So we
laboured in the
work: and
half of them
held the
spears from the
rising of the
morning till the
stars
appeared.
\VS{22}Likewise at the same
time
said I unto the
people, Let every
one with his
servant
lodge
within
Jerusalem, that in the
night they may be a
guard to us, and
labour on the
day.
\VS{23}So neither I, nor my
brethren, nor my
servants, nor the
men of the
guard which
followed me, none of
us put
off our
clothes,
{\ADD{saving that}} every
one put them
off for
washing.\FTNT{§§}{{\FR 4.23 }
saving…: or, every one went with his weapon for water}

\par }\Chap{5}{\PP \VerseOne{1}And there was a
great
cry of the
people and of their
wives against their
brethren the
Jews.
\VS{2}For there
were that
said, We, our
sons, and our
daughters,
{\ADD{are}}
many: therefore we take
up
corn
{\ADD{for them}}, that we may
eat, and
live.
\VS{3}{\ADD{Some}} also there
were that
said, We have
mortgaged our
lands,
vineyards, and
houses, that we might
buy
corn, because of the
dearth.
\VS{4}There
were also that
said, We have
borrowed
money for the
king’s
tribute,
{\ADD{and that upon}} our
lands and
vineyards.
\VS{5}Yet now our
flesh
{\ADD{is}} as the
flesh of our
brethren, our
children as their
children: and, lo, we bring into
bondage our
sons and our
daughters to be
servants, and
{\ADD{some}} of our
daughters
are brought unto
bondage
{\ADD{already:}} neither
{\ADD{is it}} in our
power
{\ADD{to redeem them;}} for other
men have our
lands and
vineyards.
\par }{\PP \VS{6}¶ And I was
very
angry when I
heard their
cry and these
words.
\VS{7}Then I
consulted with
myself, and I
rebuked the
nobles, and the
rulers, and
said unto them, Ye
exact
usury, every
one of his
brother. And I
set a
great
assembly against them.\FTNT{*}{{\FR 5.7 }
I consulted…: Heb. my heart consulted in me}
\VS{8}And I
said unto them, We after our
ability have
redeemed our
brethren the
Jews, which were
sold unto the
heathen; and will ye even
sell your
brethren? or shall they be
sold unto us? Then held they their
peace, and
found
nothing
{\ADD{to answer}}.
\VS{9}Also I
said, It
{\ADD{is}} not
good
that ye
do: ought ye not to
walk in the
fear of our
God because of the
reproach of the
heathen our
enemies?
\VS{10}I likewise,
{\ADD{and}} my
brethren, and my
servants, might
exact of them
money and
corn: I pray you, let us leave
off this
usury.
\VS{11}Restore, I pray you, to them, even this
day, their
lands, their
vineyards, their
oliveyards, and their
houses, also the
hundredth
{\ADD{part}} of the
money, and of the
corn, the
wine, and the
oil, that ye
exact of them.
\VS{12}Then
said they, We will
restore
{\ADD{them}}, and will
require nothing of them; so will we
do as thou
sayest. Then I
called the
priests, and took an
oath of them, that they should
do according to this
promise.
\VS{13}Also I
shook my
lap, and
said, So
God shake
out every
man from his
house, and from his
labour, that
performeth not this
promise, even thus be he shaken
out, and
emptied. And all the
congregation
said,
Amen, and
praised the
{\ND{
LORD}}. And the
people
did according to this
promise.\FTNT{†}{{\FR 5.13 }
emptied: Heb. empty, or, void}
\par }{\PP \VS{14}¶ Moreover from the
time that I was
appointed to be their
governor in the
land of
Judah, from the
twentieth
year even unto the
two and
thirtieth
year of
Artaxerxes the
king,
{\ADD{that is}},
twelve
years, I and my
brethren have not
eaten the
bread of the
governor.
\VS{15}But the
former
governors that
{\ADD{had been}}
before me were
chargeable unto the
people, and had
taken of them
bread and
wine,
beside
forty
shekels of
silver; yea,
even their
servants bare
rule over the
people: but so
did not I,
because of the
fear of
God.
\VS{16}Yea, also I
continued in the
work of this
wall, neither
bought we any
land: and all my
servants
{\ADD{were}}
gathered thither unto the
work.
\VS{17}Moreover
{\ADD{there were}} at my
table an
hundred and
fifty of the
Jews and
rulers, beside those that
came unto us from among the
heathen that
{\ADD{are}} about
us.
\VS{18}Now
{\ADD{that}} which was
prepared
{\ADD{for me}}
daily
{\ADD{was}}
one
ox
{\ADD{and}}
six
choice
sheep; also
fowls were
prepared for me, and once in
ten
days
store of all sorts of
wine: yet for
all this
required not I the
bread of the
governor, because the
bondage was
heavy upon this
people.
\VS{19}Think upon me, my
God, for
good,
{\ADD{according}} to all that I have
done for this
people.

\par }\Chap{6}{\PP \VerseOne{1}Now it came to pass, when
Sanballat, and
Tobiah, and
Geshem the
Arabian, and the
rest of our
enemies,
heard that I had
builded the
wall, and
{\ADD{that}} there was no
breach
left therein; (though at that
time I had not set
up the
doors upon the
gates;)
\VS{2}That
Sanballat and
Geshem
sent unto me,
saying,
Come, let us
meet
together in
{\ADD{some one of}} the
villages in the
plain of
Ono. But they
thought to
do me
mischief.
\VS{3}And I
sent
messengers unto them,
saying, I
{\ADD{am}}
doing a
great
work, so that I
cannot come
down: why should the
work
cease,
whilst I
leave it, and come
down to you?
\VS{4}Yet they
sent unto me
four
times after this
sort; and I
answered them after the same
manner.
\VS{5}Then
sent
Sanballat his
servant unto me in like
manner the
fifth
time with an
open
letter in his
hand;
\VS{6}Wherein
{\ADD{was}}
written, It is
reported among the
heathen, and
Gashmu
saith
{\ADD{it, that}} thou and the
Jews
think to
rebel: for which
cause thou
buildest the
wall, that thou mayest
be their
king, according to these
words.\FTNT{*}{{\FR 6.6 }
Gashmu: also called, Geshem}
\VS{7}And thou hast also
appointed
prophets to
preach of thee at
Jerusalem,
saying,
{\ADD{There is}} a
king in
Judah: and now shall it be
reported to the
king according to these
words.
Come now therefore, and let us take
counsel
together.
\VS{8}Then I
sent unto him,
saying, There
are no such
things
done as thou
sayest, but thou
feignest them out of thine own
heart.
\VS{9}For they all made us
afraid,
saying, Their
hands shall be
weakened from the
work, that it be not
done. Now therefore,
{\ADD{O God}},
strengthen my
hands.
\par }{\PP \VS{10}Afterward I
came unto the
house of
Shemaiah the
son of
Delaiah the
son of
Mehetabeel, who
{\ADD{was}} shut
up; and he
said, Let us meet
together in the
house of
God,
within the
temple, and let us
shut the
doors of the
temple: for they will
come to
slay thee; yea, in the
night will they
come to
slay thee.
\VS{11}And I
said, Should
such a
man as I
flee? and who
{\ADD{is there}}, that,
{\ADD{being}} as I
{\ADD{am}}, would
go into the
temple to save his
life? I will not go
in.
\VS{12}And, lo, I
perceived that
God had not
sent him; but that he
pronounced this
prophecy against me: for
Tobiah and
Sanballat had
hired him.
\VS{13}Therefore
{\ADD{was}} he
hired, that I should be
afraid, and do
so, and
sin, and
{\ADD{that}} they might have
{\ADD{matter}} for an
evil
report, that they might
reproach me.
\VS{14}My
God,
think thou upon
Tobiah and
Sanballat according to these their
works, and on the
prophetess
Noadiah, and the
rest of the
prophets, that would have put me in
fear.
\par }{\PP \VS{15}¶ So the
wall was
finished in the
twenty and
fifth
{\ADD{day}} of
{\ADD{the month}}
Elul, in
fifty and
two
days.
\VS{16}And it came to pass, that when all our
enemies
heard
{\ADD{thereof}}, and all the
heathen that
{\ADD{were}} about
us
saw
{\ADD{these things}}, they were
much cast
down in their own
eyes: for they
perceived that this
work was
wrought of our
God.
\par }{\PP \VS{17}¶ Moreover in those
days the
nobles of
Judah
sent
many
letters unto
Tobiah, and
{\ADD{the letters}} of
Tobiah
came unto them.\FTNT{†}{{\FR 6.17 }
sent…: Heb. multiplied their letters passing to Tobiah}
\VS{18}For
{\ADD{there were}}
many in
Judah
sworn unto him, because he
{\ADD{was}} the son in
law of
Shechaniah the
son of
Arah; and his
son
Johanan had
taken the
daughter of
Meshullam the
son of
Berechiah.
\VS{19}Also they
reported his good
deeds
before me, and
uttered my
words to him.
{\ADD{And}}
Tobiah
sent
letters to put me in
fear.\FTNT{‡}{{\FR 6.19 }
words: or, matters}

\par }\Chap{7}{\PP \VerseOne{1}Now it came to pass, when the
wall was
built, and I had set
up the
doors, and the
porters and the
singers and the
Levites were
appointed,
\VS{2}That I gave my
brother
Hanani, and
Hananiah the
ruler of the
palace,
charge over
Jerusalem: for he
{\ADD{was}} a
faithful
man, and
feared
God above
many.
\VS{3}And I
said unto them, Let not the
gates of
Jerusalem be
opened until the
sun be
hot; and while they stand
by, let them
shut the
doors, and
bar
{\ADD{them:}} and
appoint
watches of the
inhabitants of
Jerusalem, every
one in his
watch, and every
one
{\ADD{to be}} over against his
house.
\VS{4}Now the
city
{\ADD{was}}
large and
great: but the
people
{\ADD{were}}
few
therein, and the
houses
{\ADD{were}} not
builded.\FTNT{*}{{\FR 7.4 }
large: Heb. broad in spaces}
\par }{\PP \VS{5}¶ And my
God
put into mine
heart to gather
together the
nobles, and the
rulers, and the
people, that they might be reckoned by
genealogy. And I
found a
register of the
genealogy of them which came
up at the
first, and
found
written therein,
\VS{6}These
{\ADD{are}} the
children of the
province, that went
up out of the
captivity, of those that had been carried
away, whom
Nebuchadnezzar the
king of
Babylon had carried
away, and came
again to
Jerusalem and to
Judah, every
one unto his
city;
\VS{7}Who
came with
Zerubbabel,
Jeshua,
Nehemiah,
Azariah,
Raamiah,
Nahamani,
Mordecai,
Bilshan,
Mispereth,
Bigvai,
Nehum,
Baanah. The
number,
{\ADD{I say}}, of the
men of the
people of
Israel
{\ADD{was this;}}\FTNT{†}{{\FR 7.7 }
Azariah: also called, Seraiah}
\VS{8}The
children of
Parosh,
two
thousand an
hundred
seventy and
two.
\VS{9}The
children of
Shephatiah,
three
hundred
seventy and
two.
\VS{10}The
children of
Arah,
six
hundred
fifty and
two.
\VS{11}The
children of
Pahath-moab, of the
children of
Jeshua and
Joab, two
thousand and
eight
hundred
{\ADD{and}}
eighteen.
\VS{12}The
children of
Elam, a
thousand two
hundred
fifty and
four.
\VS{13}The
children of
Zattu,
eight
hundred
forty and
five.
\VS{14}The
children of
Zaccai,
seven
hundred and
threescore.
\VS{15}The
children of
Binnui,
six
hundred
forty and
eight.\FTNT{‡}{{\FR 7.15 }
Binnui: also called, Bani}
\VS{16}The
children of
Bebai,
six
hundred
twenty and
eight.
\VS{17}The
children of
Azgad,
two
thousand
three
hundred
twenty and
two.
\VS{18}The
children of
Adonikam,
six
hundred
threescore and
seven.
\VS{19}The
children of
Bigvai, two
thousand
threescore and
seven.
\VS{20}The
children of
Adin,
six
hundred
fifty and
five.
\VS{21}The
children of
Ater of
Hezekiah,
ninety and
eight.
\VS{22}The
children of
Hashum,
three
hundred
twenty and
eight.
\VS{23}The
children of
Bezai,
three
hundred
twenty and
four.
\VS{24}The
children of
Hariph, an
hundred and
twelve.\FTNT{§}{{\FR 7.24 }
Hariph: also called, Jora}
\VS{25}The
children of
Gibeon,
ninety and
five.\FTNT{**}{{\FR 7.25 }
Gibeon: also called, Gibbar}
\VS{26}The
men of
Beth-lehem and
Netophah, an
hundred
fourscore and
eight.
\VS{27}The
men of
Anathoth, an
hundred
twenty and
eight.
\VS{28}The
men of
Beth-azmaveth,
forty and
two.\FTNT{††}{{\FR 7.28 }
Beth-azmaveth: also called, Azmaveth}
\VS{29}The
men of
Kirjath-jearim,
Chephirah, and
Beeroth,
seven
hundred
forty and
three.\FTNT{‡‡}{{\FR 7.29 }
Kirjath-jearim: also called, Kirjath-arim}
\VS{30}The
men of
Ramah and
Geba,
six
hundred
twenty and
one.
\VS{31}The
men of
Michmas, an
hundred and
twenty and
two.
\VS{32}The
men of
Beth-el and
Ai, an
hundred
twenty and
three.
\VS{33}The
men of the
other
Nebo,
fifty and
two.
\VS{34}The
children of the
other
Elam, a
thousand two
hundred
fifty and
four.
\VS{35}The
children of
Harim,
three
hundred and
twenty.
\VS{36}The
children of
Jericho,
three
hundred
forty and
five.
\VS{37}The
children of
Lod,
Hadid, and
Ono,
seven
hundred
twenty and
one.
\VS{38}The
children of
Senaah,
three
thousand
nine
hundred and
thirty.
\par }{\PP \VS{39}¶ The
priests: the
children of
Jedaiah, of the
house of
Jeshua,
nine
hundred
seventy and
three.
\VS{40}The
children of
Immer, a
thousand
fifty and
two.
\VS{41}The
children of
Pashur, a
thousand two
hundred
forty and
seven.
\VS{42}The
children of
Harim, a
thousand and
seventeen.
\par }{\PP \VS{43}¶ The
Levites: the
children of
Jeshua, of
Kadmiel,
{\ADD{and}} of the
children of
Hodevah,
seventy and
four.\FTNT{§§}{{\FR 7.43 }
Hodevah: also called, Hodaviah or Judah}
\par }{\PP \VS{44}¶ The
singers: the
children of
Asaph, an
hundred
forty and
eight.
\par }{\PP \VS{45}¶ The
porters: the
children of
Shallum, the
children of
Ater, the
children of
Talmon, the
children of
Akkub, the
children of
Hatita, the
children of
Shobai, an
hundred
thirty and
eight.
\par }{\PP \VS{46}¶ The
Nethinims: the
children of
Ziha, the
children of
Hashupha, the
children of
Tabbaoth,
\VS{47}The
children of
Keros, the
children of
Sia, the
children of
Padon,
\VS{48}The
children of
Lebana, the
children of
Hagaba, the
children of
Shalmai,
\VS{49}The
children of
Hanan, the
children of
Giddel, the
children of
Gahar,
\VS{50}The
children of
Reaiah, the
children of
Rezin, the
children of
Nekoda,
\VS{51}The
children of
Gazzam, the
children of
Uzza, the
children of
Phaseah,
\VS{52}The
children of
Besai, the
children of
Meunim, the
children of
Nephishesim,
\VS{53}The
children of
Bakbuk, the
children of
Hakupha, the
children of
Harhur,
\VS{54}The
children of
Bazlith, the
children of
Mehida, the
children of
Harsha,
\VS{55}The
children of
Barkos, the
children of
Sisera, the
children of
Tamah,
\VS{56}The
children of
Neziah, the
children of
Hatipha.
\par }{\PP \VS{57}¶ The
children of
Solomon’s
servants: the
children of
Sotai, the
children of
Sophereth, the
children of
Perida,
\VS{58}The
children of
Jaala, the
children of
Darkon, the
children of
Giddel,
\VS{59}The
children of
Shephatiah, the
children of
Hattil, the
children of Pochereth of
Zebaim, the
children of
Amon.\FTNT{***}{{\FR 7.59 }
Amon: also called, Ami}
\VS{60}All the
Nethinims, and the
children of
Solomon’s
servants,
{\ADD{were}}
three
hundred
ninety and
two.
\VS{61}And these
{\ADD{were}} they which went
up
{\ADD{also}} from
Tel-melah,
Tel-haresha,
Cherub,
Addon, and
Immer: but they
could not
shew their
father’s
house, nor their
seed, whether they
{\ADD{were}} of
Israel.\FTNT{†††}{{\FR 7.61 }
seed: or, pedigree}
\VS{62}The
children of
Delaiah, the
children of
Tobiah, the
children of
Nekoda,
six
hundred
forty and
two.
\par }{\PP \VS{63}¶ And of the
priests: the
children of
Habaiah, the
children of
Koz, the
children of
Barzillai, which
took
{\ADD{one}} of the
daughters of
Barzillai the
Gileadite to
wife, and was
called after their
name.
\VS{64}These
sought their
register
{\ADD{among}} those that were reckoned by
genealogy, but it was not
found: therefore were they, as
polluted, put from the
priesthood.
\VS{65}And the
Tirshatha
said unto them, that they should not
eat of the
most holy
things, till there
stood
{\ADD{up}} a
priest with
Urim and
Thummim.\FTNT{‡‡‡}{{\FR 7.65 }
the Tirshatha: or, the governor}
\par }{\PP \VS{66}¶ The whole
congregation
together
{\ADD{was}}
forty and two
thousand
three
hundred and
threescore,
\VS{67}Beside their
manservants and their
maidservants, of whom
{\ADD{there were}}
seven
thousand
three
hundred
thirty and
seven: and they had two
hundred
forty and
five singing
men and singing
women.
\VS{68}Their
horses,
seven
hundred
thirty and
six: their
mules, two
hundred
forty and
five:
\VS{69}{\ADD{Their}}
camels,
four
hundred
thirty and
five:
six
thousand
seven
hundred and
twenty
asses.
\par }{\PP \VS{70}¶ And some
of the
chief of the
fathers
gave unto the
work. The
Tirshatha
gave to the
treasure a
thousand
drams of
gold,
fifty
basons,
five
hundred and
thirty
priests’
garments.\FTNT{§§§}{{\FR 7.70 }
some: Heb. part.}\FTNT{*}{{\FR 7.70 }
The Tirshatha: or, the governor}
\VS{71}And
{\ADD{some}} of the
chief of the
fathers
gave to the
treasure of the
work
twenty
thousand
drams of
gold, and two
thousand and two
hundred
pound of
silver.
\VS{72}And
{\ADD{that}} which the
rest of the
people
gave
{\ADD{was}}
twenty
thousand
drams of
gold, and two
thousand
pound of
silver, and
threescore and
seven
priests’
garments.
\VS{73}So the
priests, and the
Levites, and the
porters, and the
singers, and
{\ADD{some}} of the
people, and the
Nethinims, and all
Israel,
dwelt in their
cities; and when the
seventh
month
came, the
children of
Israel
{\ADD{were}} in their
cities.

\par }\Chap{8}{\PP \VerseOne{1}And all the
people gathered themselves
together as
one
man into the
street that
{\ADD{was}}
before the
water
gate; and they
spake unto
Ezra the
scribe to
bring the
book of the
law of
Moses, which the
{\ND{
LORD}} had
commanded to
Israel.
\VS{2}And
Ezra the
priest
brought the
law
before the
congregation both of
men and
women, and all that could
hear with
understanding, upon the
first
day of the
seventh
month.\FTNT{*}{{\FR 8.2 }
that…: Heb. that understood in hearing}
\VS{3}And he
read therein
before the
street that
{\ADD{was}}
before the
water
gate from the
morning until
midday, before the
men and the
women, and those that could
understand; and the
ears of all the
people
{\ADD{were attentive}} unto the
book of the
law.\FTNT{†}{{\FR 8.3 }
from…: Heb. from the light}
\VS{4}And
Ezra the
scribe
stood upon a
pulpit of
wood, which they had
made for the
purpose; and
beside him
stood
Mattithiah, and
Shema, and
Anaiah, and
Urijah, and
Hilkiah, and
Maaseiah, on his right
hand; and on his left
hand,
Pedaiah, and
Mishael, and
Malchiah, and
Hashum, and
Hashbadana,
Zechariah,
{\ADD{and}}
Meshullam.\FTNT{‡}{{\FR 8.4 }
pulpit…: Heb. tower of wood}
\VS{5}And
Ezra
opened the
book in the
sight of all the
people; (for he was above all the
people;) and when he
opened it, all the
people stood
up:\FTNT{§}{{\FR 8.5 }
sight: Heb. eyes}
\VS{6}And
Ezra
blessed the
{\ND{
LORD}}, the
great
God. And all the
people
answered,
Amen,
Amen, with lifting
up their
hands: and they
bowed their heads, and
worshipped the
{\ND{
LORD}} with
{\ADD{their}}
faces to the
ground.
\VS{7}Also
Jeshua, and
Bani, and
Sherebiah,
Jamin,
Akkub,
Shabbethai,
Hodijah,
Maaseiah,
Kelita,
Azariah,
Jozabad,
Hanan,
Pelaiah, and the
Levites, caused the
people to
understand the
law: and the
people
{\ADD{stood}} in their
place.
\VS{8}So they
read in the
book in the
law of
God
distinctly, and
gave the
sense, and caused
{\ADD{them}} to
understand the
reading.
\par }{\PP \VS{9}¶ And
Nehemiah, which
{\ADD{is}} the
Tirshatha, and
Ezra the
priest the
scribe, and the
Levites that
taught the
people,
said unto all the
people, This
day
{\ADD{is}}
holy unto the
{\ND{
LORD}} your
God;
mourn not, nor
weep. For all the
people
wept, when they
heard the
words of the
law.\FTNT{**}{{\FR 8.9 }
the Tirshatha: or, the governor}
\VS{10}Then he
said unto them, Go your
way,
eat the
fat, and
drink the
sweet, and
send
portions unto them for whom nothing is
prepared: for
{\ADD{this}}
day
{\ADD{is}}
holy unto our
Lord: neither be ye
sorry; for the
joy of the
{\ND{
LORD}} is your
strength.
\VS{11}So the
Levites
stilled all the
people,
saying, Hold your
peace, for the
day
{\ADD{is}}
holy; neither be ye
grieved.
\VS{12}And all the
people went their
way to
eat, and to
drink, and to
send
portions, and to
make
great
mirth, because they had
understood the
words that were
declared unto them.
\par }{\PP \VS{13}¶ And on the
second
day were gathered
together the
chief of the
fathers of all the
people, the
priests, and the
Levites, unto
Ezra the
scribe, even to
understand the
words of the
law.\FTNT{††}{{\FR 8.13 }
to understand…: or, that they might instruct in the words of the law}
\VS{14}And they
found
written in the
law which the
{\ND{
LORD}} had
commanded
by
Moses, that the
children of
Israel should
dwell in
booths in the
feast of the
seventh
month:\FTNT{‡‡}{{\FR 8.14 }
by: Heb. by the hand of}
\VS{15}And that they should
publish and
proclaim in all their
cities, and in
Jerusalem,
saying, Go
forth unto the
mount, and
fetch
olive
branches, and
pine
branches, and
myrtle
branches, and
palm
branches, and
branches of
thick
trees, to
make
booths, as
{\ADD{it is}}
written.
\par }{\PP \VS{16}¶ So the
people went
forth, and
brought
{\ADD{them}}, and
made themselves
booths, every
one upon the roof of his
house, and in their
courts, and in the
courts of the
house of
God, and in the
street of the
water
gate, and in the
street of the
gate of
Ephraim.
\VS{17}And all the
congregation of them that were come
again out of the
captivity
made
booths, and
sat under the
booths: for since the
days of
Jeshua the
son of
Nun unto that
day had not the
children of
Israel done
so. And there was
very
great
gladness.
\VS{18}Also
day by
day, from the
first
day unto the
last
day, he
read in the
book of the
law of
God. And they
kept the
feast
seven
days; and on the
eighth
day
{\ADD{was}} a solemn
assembly, according unto the
manner.\FTNT{§§}{{\FR 8.18 }
a solemn…: Heb. a restraint}

\par }\Chap{9}{\PP \VerseOne{1}Now in the
twenty and
fourth
day of this
month the
children of
Israel were
assembled with
fasting, and with
sackclothes, and
earth upon them.
\VS{2}And the
seed of
Israel
separated themselves from all
strangers, and
stood and
confessed their
sins, and the
iniquities of their
fathers.\FTNT{*}{{\FR 9.2 }
strangers: Heb. strange children}
\VS{3}And they stood
up in their
place, and
read in the
book of the
law of the
{\ND{
LORD}} their
God
{\ADD{one}} fourth
part of the
day; and
{\ADD{another}} fourth
part they
confessed, and
worshipped the
{\ND{
LORD}} their
God.
\par }{\PP \VS{4}¶ Then stood
up upon the
stairs, of the
Levites,
Jeshua, and
Bani,
Kadmiel,
Shebaniah,
Bunni,
Sherebiah,
Bani,
{\ADD{and}}
Chenani, and
cried with a
loud
voice unto the
{\ND{
LORD}} their
God.\FTNT{†}{{\FR 9.4 }
stairs: or, scaffold}
\VS{5}Then the
Levites,
Jeshua, and
Kadmiel,
Bani,
Hashabniah,
Sherebiah,
Hodijah,
Shebaniah,
{\ADD{and}}
Pethahiah,
said, Stand
up
{\ADD{and}}
bless the
{\ND{
LORD}} your
God for
ever and
ever: and
blessed be thy
glorious
name, which is
exalted above all
blessing and
praise.
\VS{6}Thou,
{\ADD{even}} thou,
{\ADD{art}}
{\ND{
LORD}} alone; thou hast
made
heaven, the
heaven of
heavens, with all their
host, the
earth, and all
{\ADD{things}} that
{\ADD{are}} therein, the
seas, and all that
{\ADD{is}} therein, and thou
preservest them all; and the
host of
heaven
worshippeth thee.
\VS{7}Thou
{\ADD{art}} the
{\ND{
LORD}} the
God, who didst
choose
Abram, and broughtest him
forth out of
Ur of the
Chaldees, and
gavest him the
name of
Abraham;
\VS{8}And
foundest his
heart
faithful
before thee, and
madest a
covenant with him to
give the
land of the
Canaanites, the
Hittites, the
Amorites, and the
Perizzites, and the
Jebusites, and the
Girgashites, to
give
{\ADD{it, I say}}, to his
seed, and hast
performed thy
words; for thou
{\ADD{art}}
righteous:
\VS{9}And didst
see the
affliction of our
fathers in
Egypt, and
heardest their
cry by the
Red
sea;
\VS{10}And
shewedst
signs and
wonders upon
Pharaoh, and on all his
servants, and on all the
people of his
land: for thou
knewest that they dealt
proudly against them. So didst thou
get thee a
name, as
{\ADD{it is}} this
day.
\VS{11}And thou didst
divide the
sea
before them, so that they went
through the
midst of the
sea on the dry
land; and their
persecutors thou
threwest into the
deeps, as a
stone into the
mighty
waters.
\VS{12}Moreover thou
leddest them in the
day by a
cloudy
pillar; and in the
night by a
pillar of
fire, to give them
light in the
way wherein they should
go.
\VS{13}Thou camest
down also upon
mount
Sinai, and
spakest with them from
heaven, and
gavest them
right
judgments, and
true
laws,
good
statutes and
commandments:\FTNT{‡}{{\FR 9.13 }
true…: Heb. laws of truth}
\VS{14}And madest
known unto them thy
holy
sabbath, and
commandedst them
precepts,
statutes, and
laws, by the
hand of
Moses thy
servant:
\VS{15}And
gavest them
bread from
heaven for their
hunger, and broughtest
forth
water for them out of the
rock for their
thirst, and
promisedst them that they should go
in to
possess the
land which thou hadst
sworn to
give them.\FTNT{§}{{\FR 9.15 }
which…: Heb. which thou hadst lift up thine hand to give them}
\VS{16}But they and our
fathers dealt
proudly, and
hardened their
necks, and
hearkened not to thy
commandments,
\VS{17}And
refused to
obey, neither were
mindful of thy
wonders that thou
didst among them; but
hardened their
necks, and in their
rebellion
appointed a
captain to
return to their
bondage: but thou
{\ADD{art}} a
God ready to
pardon,
gracious and
merciful,
slow to
anger, and of
great
kindness, and
forsookest them not.\FTNT{**}{{\FR 9.17 }
a God…: Heb. a God of pardons}
\VS{18}Yea, when they had
made them a
molten
calf, and
said, This
{\ADD{is}} thy
God that brought thee
up out of
Egypt, and had
wrought
great
provocations;
\VS{19}Yet thou in thy
manifold
mercies
forsookest them not in the
wilderness: the
pillar of the
cloud
departed not from them by
day, to
lead them in the
way; neither the
pillar of
fire by
night, to shew them
light, and the
way wherein they should
go.
\VS{20}Thou
gavest also thy
good
spirit to
instruct them, and
withheldest not thy
manna from their
mouth, and
gavest them
water for their
thirst.
\VS{21}Yea,
forty
years didst thou
sustain them in the
wilderness,
{\ADD{so that}} they
lacked nothing; their
clothes waxed not
old, and their
feet
swelled not.
\VS{22}Moreover thou
gavest them
kingdoms and
nations, and didst
divide them into
corners: so they
possessed the
land of
Sihon, and the
land of the
king of
Heshbon, and the
land of
Og
king of
Bashan.
\VS{23}Their
children also
multipliedst thou as the
stars of
heaven, and
broughtest them into the
land, concerning which thou hadst
promised to their
fathers, that they should go
in to
possess
{\ADD{it}}.
\VS{24}So the
children went
in and
possessed the
land, and thou
subduedst
before them the
inhabitants of the
land, the
Canaanites, and
gavest them into their
hands, with their
kings, and the
people of the
land, that they might
do with them as they
would.\FTNT{††}{{\FR 9.24 }
as they…: Heb. according to their will}
\VS{25}And they
took
strong
cities, and a
fat
land, and
possessed
houses
full of all
goods,
wells
digged,
vineyards, and
oliveyards, and
fruit
trees in
abundance: so they did
eat, and were
filled, and became
fat, and
delighted themselves in thy
great
goodness.\FTNT{‡‡}{{\FR 9.25 }
wells: or, cisterns}\FTNT{§§}{{\FR 9.25 }
fruit…: Heb. tree of food}
\VS{26}Nevertheless they were
disobedient, and
rebelled against thee, and
cast thy
law
behind their
backs, and
slew thy
prophets which
testified against them to
turn them to thee, and they
wrought
great
provocations.
\VS{27}Therefore thou
deliveredst them into the
hand of their
enemies, who
vexed them: and in the
time of their
trouble, when they
cried unto thee, thou
heardest
{\ADD{them}} from
heaven; and according to thy
manifold
mercies thou
gavest them
saviours, who
saved them out of the
hand of their
enemies.
\VS{28}But after they had
rest, they
did
evil
again
before thee: therefore
leftest thou them in the
hand of their
enemies, so that they had the
dominion over them: yet when they
returned, and
cried unto thee, thou
heardest
{\ADD{them}} from
heaven; and
many
times didst thou
deliver them according to thy
mercies;\FTNT{***}{{\FR 9.28 }
they did…: Heb. they returned to do evil}
\VS{29}And
testifiedst against them, that thou mightest bring them
again unto thy
law: yet they dealt
proudly, and
hearkened not unto thy
commandments, but
sinned against thy
judgments, (which if a
man
do, he shall
live in them;) and
withdrew the
shoulder, and
hardened their
neck, and would not
hear.\FTNT{†††}{{\FR 9.29 }
withdrew…: Heb. they gave a withdrawing shoulder}
\VS{30}Yet
many
years didst thou
forbear them, and
testifiedst against them by thy
spirit
in thy
prophets: yet would they not give
ear: therefore
gavest thou them into the
hand of the
people of the
lands.\FTNT{‡‡‡}{{\FR 9.30 }
forbear…: Heb. protract over them}\FTNT{§§§}{{\FR 9.30 }
in thy…: Heb. in the hand of thy prophets}
\VS{31}Nevertheless for thy
great
mercies’ sake thou
didst not utterly
consume them, nor
forsake them; for thou
{\ADD{art}} a
gracious and
merciful
God.
\VS{32}Now therefore, our
God, the
great, the
mighty, and the
terrible
God, who
keepest
covenant and
mercy, let not all the
trouble seem
little
before thee, that hath
come upon us, on our
kings, on our
princes, and on our
priests, and on our
prophets, and on our
fathers, and on all thy
people, since the
time of the
kings of
Assyria unto this
day.\FTNT{*}{{\FR 9.32 }
trouble: Heb. weariness}\FTNT{†}{{\FR 9.32 }
that…: Heb. that hath found us}
\VS{33}Howbeit thou
{\ADD{art}}
just in all that is
brought upon us; for thou hast
done
right, but we have done
wickedly:
\VS{34}Neither have our
kings, our
princes, our
priests, nor our
fathers,
kept thy
law, nor
hearkened unto thy
commandments and thy
testimonies, wherewith thou didst
testify against them.
\VS{35}For they have not
served thee in their
kingdom, and in thy
great
goodness that thou
gavest them, and in the
large and
fat
land which thou
gavest
before them, neither
turned they from their
wicked
works.
\VS{36}Behold, we
{\ADD{are}}
servants this
day, and
{\ADD{for}} the
land that thou
gavest unto our
fathers to
eat the
fruit thereof and the
good thereof, behold, we
{\ADD{are}}
servants in it:
\VS{37}And it yieldeth
much
increase unto the
kings whom thou hast
set over us because of our
sins: also they have
dominion over our
bodies, and over our
cattle, at their
pleasure, and we
{\ADD{are}} in
great
distress.
\VS{38}And because of all this we
make a
sure
{\ADD{covenant}}, and
write
{\ADD{it;}} and our
princes,
Levites,
{\ADD{and}}
priests,
seal
{\ADD{unto it}}.\FTNT{‡}{{\FR 9.38 }
seal: Heb. are at the sealing, or, sealed}

\par }\Chap{10}{\PP \VerseOne{1}Now those that
sealed
{\ADD{were}},
Nehemiah, the
Tirshatha, the
son of
Hachaliah, and
Zidkijah,\FTNT{*}{{\FR 10.1 }
those…: Heb. at the sealings}\FTNT{†}{{\FR 10.1 }
the Tirshatha: or, the governor}
\VS{2}Seraiah,
Azariah,
Jeremiah,
\VS{3}Pashur,
Amariah,
Malchijah,
\VS{4}Hattush,
Shebaniah,
Malluch,
\VS{5}Harim,
Meremoth,
Obadiah,
\VS{6}Daniel,
Ginnethon,
Baruch,
\VS{7}Meshullam,
Abijah,
Mijamin,
\VS{8}Maaziah,
Bilgai,
Shemaiah: these
{\ADD{were}} the
priests.
\VS{9}And the
Levites: both
Jeshua the
son of
Azaniah,
Binnui of the
sons of
Henadad,
Kadmiel;
\VS{10}And their
brethren,
Shebaniah,
Hodijah,
Kelita,
Pelaiah,
Hanan,
\VS{11}Micha,
Rehob,
Hashabiah,
\VS{12}Zaccur,
Sherebiah,
Shebaniah,
\VS{13}Hodijah,
Bani,
Beninu.
\VS{14}The
chief of the
people;
Parosh,
Pahath-moab,
Elam,
Zatthu,
Bani,
\VS{15}Bunni,
Azgad,
Bebai,
\VS{16}Adonijah,
Bigvai,
Adin,
\VS{17}Ater,
Hizkijah,
Azzur,
\VS{18}Hodijah,
Hashum,
Bezai,
\VS{19}Hariph,
Anathoth,
Nebai,
\VS{20}Magpiash,
Meshullam,
Hezir,
\VS{21}Meshezabeel,
Zadok,
Jaddua,
\VS{22}Pelatiah,
Hanan,
Anaiah,
\VS{23}Hoshea,
Hananiah,
Hashub,
\VS{24}Hallohesh,
Pileha,
Shobek,
\VS{25}Rehum,
Hashabnah,
Maaseiah,
\VS{26}And
Ahijah,
Hanan,
Anan,
\VS{27}Malluch,
Harim,
Baanah.
\par }{\PP \VS{28}¶ And the
rest of the
people, the
priests, the
Levites, the
porters, the
singers, the
Nethinims, and all they that had
separated themselves from the
people of the
lands unto the
law of
God, their
wives, their
sons, and their
daughters, every one having
knowledge, and having
understanding;
\VS{29}They
clave to their
brethren, their
nobles, and
entered into a
curse, and into an
oath, to
walk in
God’s
law, which was
given
by
Moses the
servant of
God, and to
observe and
do all the
commandments of the
{\ND{
LORD}} our
Lord, and his
judgments and his
statutes;\FTNT{‡}{{\FR 10.29 }
by: Heb. by the hand of}
\VS{30}And that we would not
give our
daughters unto the
people of the
land, nor
take their
daughters for our
sons:
\VS{31}And
{\ADD{if}} the
people of the
land
bring
ware or any
victuals on the
sabbath
day to
sell,
{\ADD{that}} we would not
buy it of them on the
sabbath, or on the
holy
day: and
{\ADD{that}} we would
leave the
seventh
year, and the
exaction of every
debt.\FTNT{§}{{\FR 10.31 }
every…: Heb. every hand}
\par }{\PP \VS{32}Also we
made
ordinances for us, to
charge ourselves
yearly with the third
part of a
shekel for the
service of the
house of our
God;
\VS{33}For the
shewbread, and for the
continual meat
offering, and for the
continual burnt
offering, of the
sabbaths, of the new
moons, for the set
feasts, and for the
holy
{\ADD{things}}, and for the sin
offerings to make an
atonement for
Israel, and
{\ADD{for}} all the
work of the
house of our
God.
\VS{34}And we
cast the
lots among the
priests, the
Levites, and the
people, for the
wood
offering, to
bring
{\ADD{it}} into the
house of our
God, after the
houses of our
fathers, at
times
appointed
year by
year, to
burn upon the
altar of the
{\ND{
LORD}} our
God, as
{\ADD{it is}}
written in the
law:
\VS{35}And to
bring the
firstfruits of our
ground, and the
firstfruits of all
fruit of all
trees,
year by
year, unto the
house of the
{\ND{
LORD}}:
\VS{36}Also the
firstborn of our
sons, and of our
cattle, as
{\ADD{it is}}
written in the
law, and the
firstlings of our
herds and of our
flocks, to
bring to the
house of our
God, unto the
priests that
minister in the
house of our
God:
\VS{37}And
{\ADD{that}} we should
bring the
firstfruits of our
dough, and our
offerings, and the
fruit of all manner of
trees, of
wine and of
oil, unto the
priests, to the
chambers of the
house of our
God; and the
tithes of our
ground unto the
Levites, that the same
Levites might have the
tithes in all the
cities of our
tillage.
\VS{38}And the
priest the
son of
Aaron shall be with the
Levites, when the
Levites take
tithes: and the
Levites shall bring
up the
tithe of the
tithes unto the
house of our
God, to the
chambers, into the
treasure
house.
\VS{39}For the
children of
Israel and the
children of
Levi shall
bring the
offering of the
corn, of the new
wine, and the
oil, unto the
chambers, where
{\ADD{are}} the
vessels of the
sanctuary, and the
priests that
minister, and the
porters, and the
singers: and we will not
forsake the
house of our
God.

\par }\Chap{11}{\PP \VerseOne{1}And the
rulers of the
people
dwelt at
Jerusalem: the
rest of the
people also
cast
lots, to
bring
one of
ten to
dwell in
Jerusalem the
holy
city, and
nine
parts
{\ADD{to dwell}} in
{\ADD{other}}
cities.
\VS{2}And the
people
blessed all the
men, that willingly
offered themselves to
dwell at
Jerusalem.
\par }{\PP \VS{3}¶ Now these
{\ADD{are}} the
chief of the
province that
dwelt in
Jerusalem: but in the
cities of
Judah
dwelt every
one in his
possession in their
cities,
{\ADD{to wit}},
Israel, the
priests, and the
Levites, and the
Nethinims, and the
children of
Solomon’s
servants.
\VS{4}And at
Jerusalem
dwelt
{\ADD{certain}} of the
children of
Judah, and of the
children of
Benjamin. Of the
children of
Judah;
Athaiah the
son of
Uzziah, the
son of
Zechariah, the
son of
Amariah, the
son of
Shephatiah, the
son of
Mahalaleel, of the
children of
Perez;\FTNT{*}{{\FR 11.4 }
Perez: also called, Pharez}
\VS{5}And
Maaseiah the
son of
Baruch, the
son of
Col-hozeh, the
son of
Hazaiah, the
son of
Adaiah, the
son of
Joiarib, the
son of
Zechariah, the
son of
Shiloni.
\VS{6}All the
sons of
Perez that
dwelt at
Jerusalem
{\ADD{were}}
four
hundred
threescore and
eight
valiant
men.
\VS{7}And these
{\ADD{are}} the
sons of
Benjamin;
Sallu the
son of
Meshullam, the
son of
Joed, the
son of
Pedaiah, the
son of
Kolaiah, the
son of
Maaseiah, the
son of
Ithiel, the
son of
Jesaiah.
\VS{8}And
after him
Gabbai,
Sallai,
nine
hundred
twenty and
eight.
\VS{9}And
Joel the
son of
Zichri
{\ADD{was}} their
overseer: and
Judah the
son of
Senuah
{\ADD{was}}
second over the
city.
\VS{10}Of the
priests:
Jedaiah the
son of
Joiarib,
Jachin.
\VS{11}Seraiah the
son of
Hilkiah, the
son of
Meshullam, the
son of
Zadok, the
son of
Meraioth, the
son of
Ahitub,
{\ADD{was}} the
ruler of the
house of
God.
\VS{12}And their
brethren that
did the
work of the
house
{\ADD{were}}
eight
hundred
twenty and
two: and
Adaiah the
son of
Jeroham, the
son of
Pelaliah, the
son of
Amzi, the
son of
Zechariah, the
son of
Pashur, the
son of
Malchiah,
\VS{13}And his
brethren,
chief of the
fathers,
two
hundred
forty and
two: and
Amashai the
son of
Azareel, the
son of
Ahasai, the
son of
Meshillemoth, the
son of
Immer,
\VS{14}And their
brethren, mighty
men of
valour, an
hundred
twenty and
eight: and their
overseer
{\ADD{was}}
Zabdiel, the
son of
{\ADD{one of}} the great
men.\FTNT{†}{{\FR 11.14 }
the son…: or, the son of Haggedolim}
\VS{15}Also of the
Levites:
Shemaiah the
son of
Hashub, the
son of
Azrikam, the
son of
Hashabiah, the
son of
Bunni;
\VS{16}And
Shabbethai and
Jozabad, of the
chief of the
Levites,
{\ADD{had}} the oversight of the
outward
business of the
house of
God.\FTNT{‡}{{\FR 11.16 }
had…: Heb. were over}
\VS{17}And
Mattaniah the
son of
Micha, the
son of
Zabdi, the
son of
Asaph,
{\ADD{was}} the
principal to
begin the
thanksgiving in
prayer: and
Bakbukiah the
second among his
brethren, and
Abda the
son of
Shammua, the
son of
Galal, the
son of
Jeduthun.
\VS{18}All the
Levites in the
holy
city
{\ADD{were}} two
hundred
fourscore and
four.
\VS{19}Moreover the
porters,
Akkub,
Talmon, and their
brethren that
kept the
gates,
{\ADD{were}} an
hundred
seventy and
two.\FTNT{§}{{\FR 11.19 }
the gates: Heb. at the gates}
\par }{\PP \VS{20}¶ And the
residue of
Israel, of the
priests,
{\ADD{and}} the
Levites,
{\ADD{were}} in all the
cities of
Judah, every
one in his
inheritance.
\VS{21}But the
Nethinims
dwelt in
Ophel: and
Ziha and
Gispa
{\ADD{were}} over the
Nethinims.\FTNT{**}{{\FR 11.21 }
Ophel: or, the tower}
\VS{22}The
overseer also of the
Levites at
Jerusalem
{\ADD{was}}
Uzzi the
son of
Bani, the
son of
Hashabiah, the
son of
Mattaniah, the
son of
Micha. Of the
sons of
Asaph, the
singers
{\ADD{were}}
over the
business of the
house of
God.
\VS{23}For
{\ADD{it was}} the
king’s
commandment concerning them, that a certain
portion should be for the
singers,
due for every
day.\FTNT{††}{{\FR 11.23 }
a certain…: or, a sure ordinance}
\VS{24}And
Pethahiah the
son of
Meshezabeel, of the
children of
Zerah the
son of
Judah,
{\ADD{was}} at the
king’s
hand in all
matters concerning the
people.\FTNT{‡‡}{{\FR 11.24 }
Zerah: also called, Zarah}
\VS{25}And for the
villages, with their
fields,
{\ADD{some}} of the
children of
Judah
dwelt at
Kirjath-arba, and
{\ADD{in}} the
villages thereof, and at
Dibon, and
{\ADD{in}} the
villages thereof, and at
Jekabzeel, and
{\ADD{in}} the
villages thereof,
\VS{26}And at
Jeshua, and at
Moladah, and at
Beth-phelet,
\VS{27}And at
Hazar-shual, and at
Beer-sheba, and
{\ADD{in}} the
villages thereof,
\VS{28}And at
Ziklag, and at
Mekonah, and in the
villages thereof,
\VS{29}And at
En-rimmon, and at
Zareah, and at
Jarmuth,
\VS{30}Zanoah,
Adullam, and
{\ADD{in}} their
villages, at
Lachish, and the
fields thereof, at
Azekah, and
{\ADD{in}} the
villages thereof. And they
dwelt from
Beer-sheba unto the
valley of
Hinnom.
\VS{31}The
children also of
Benjamin from
Geba
{\ADD{dwelt}} at
Michmash, and
Aija, and
Beth-el, and
{\ADD{in}} their
villages,\FTNT{§§}{{\FR 11.31 }
from: or, of}\FTNT{***}{{\FR 11.31 }
at: or, to}
\VS{32}{\ADD{And}} at
Anathoth,
Nob,
Ananiah,
\VS{33}Hazor,
Ramah,
Gittaim,
\VS{34}Hadid,
Zeboim,
Neballat,
\VS{35}Lod, and
Ono, the
valley of
craftsmen.
\VS{36}And of the
Levites
{\ADD{were}}
divisions
{\ADD{in}}
Judah,
{\ADD{and}} in
Benjamin.

\par }\Chap{12}{\PP \VerseOne{1}Now these
{\ADD{are}} the
priests and the
Levites that went
up with
Zerubbabel the
son of
Shealtiel, and
Jeshua:
Seraiah,
Jeremiah,
Ezra,
\VS{2}Amariah,
Malluch,
Hattush,\FTNT{*}{{\FR 12.2 }
Malluch: also called, Melicu}
\VS{3}Shechaniah,
Rehum,
Meremoth,\FTNT{†}{{\FR 12.3 }
Shechaniah: also called, Shebaniah}\FTNT{‡}{{\FR 12.3 }
Rehum: also called, Harim}\FTNT{§}{{\FR 12.3 }
Meremoth: also called, Meraioth}
\VS{4}Iddo,
Ginnetho,
Abijah,\FTNT{**}{{\FR 12.4 }
Ginnetho: also called, Ginnethon}
\VS{5}Miamin,
Maadiah,
Bilgah,\FTNT{††}{{\FR 12.5 }
Miamin: also called, Miniamin}\FTNT{‡‡}{{\FR 12.5 }
Maadiah: also called, Moadiah}
\VS{6}Shemaiah, and
Joiarib,
Jedaiah,
\VS{7}Sallu,
Amok,
Hilkiah,
Jedaiah. These
{\ADD{were}} the
chief of the
priests and of their
brethren in the
days of
Jeshua.\FTNT{§§}{{\FR 12.7 }
Sallu: also called, Sallai}
\VS{8}Moreover the
Levites:
Jeshua,
Binnui,
Kadmiel,
Sherebiah,
Judah,
{\ADD{and}}
Mattaniah,
{\ADD{which was}} over the
thanksgiving, he and his
brethren.\FTNT{***}{{\FR 12.8 }
the thanksgiving: that is, the psalms of thanksgiving}
\VS{9}Also
Bakbukiah and
Unni, their
brethren,
{\ADD{were}} over against them in the
watches.
\par }{\PP \VS{10}¶ And
Jeshua
begat
Joiakim,
Joiakim also
begat
Eliashib, and
Eliashib
begat
Joiada,
\VS{11}And
Joiada
begat
Jonathan, and
Jonathan
begat
Jaddua.
\VS{12}And in the
days of
Joiakim were
priests, the
chief of the
fathers: of
Seraiah,
Meraiah; of
Jeremiah,
Hananiah;
\VS{13}Of
Ezra,
Meshullam; of
Amariah,
Jehohanan;
\VS{14}Of
Melicu,
Jonathan; of
Shebaniah,
Joseph;
\VS{15}Of
Harim,
Adna; of
Meraioth,
Helkai;
\VS{16}Of
Iddo,
Zechariah; of
Ginnethon,
Meshullam;
\VS{17}Of
Abijah,
Zichri; of
Miniamin, of
Moadiah,
Piltai;
\VS{18}Of
Bilgah,
Shammua; of
Shemaiah,
Jehonathan;
\VS{19}And of
Joiarib,
Mattenai; of
Jedaiah,
Uzzi;
\VS{20}Of
Sallai,
Kallai; of
Amok,
Eber;
\VS{21}Of
Hilkiah,
Hashabiah; of
Jedaiah,
Nethaneel.
\par }{\PP \VS{22}¶ The
Levites in the
days of
Eliashib,
Joiada, and
Johanan, and
Jaddua,
{\ADD{were}}
recorded
chief of the
fathers: also the
priests, to the
reign of
Darius the
Persian.
\VS{23}The
sons of
Levi, the
chief of the
fathers,
{\ADD{were}}
written in the
book of the
chronicles, even until the
days of
Johanan the
son of
Eliashib.
\VS{24}And the
chief of the
Levites:
Hashabiah,
Sherebiah, and
Jeshua the
son of
Kadmiel, with their
brethren over against them, to
praise
{\ADD{and}} to give
thanks, according to the
commandment of
David the
man of
God,
ward over
against
ward.
\VS{25}Mattaniah, and
Bakbukiah,
Obadiah,
Meshullam,
Talmon,
Akkub,
{\ADD{were}}
porters
keeping the
ward at the
thresholds of the
gates.\FTNT{†††}{{\FR 12.25 }
thresholds: or, treasuries, or, assemblies}
\VS{26}These
{\ADD{were}} in the
days of
Joiakim the
son of
Jeshua, the
son of
Jozadak, and in the
days of
Nehemiah the
governor, and of
Ezra the
priest, the
scribe.
\par }{\PP \VS{27}¶ And at the
dedication of the
wall of
Jerusalem they
sought the
Levites out of all their
places, to
bring them to
Jerusalem, to
keep the
dedication with
gladness, both with
thanksgivings, and with
singing,
{\ADD{with}}
cymbals,
psalteries, and with
harps.
\VS{28}And the
sons of the
singers gathered themselves
together, both out of the plain
country round
about
Jerusalem, and from the
villages of
Netophathi;
\VS{29}Also from the
house of
Gilgal, and out of the
fields of
Geba and
Azmaveth: for the
singers had
builded them
villages round
about
Jerusalem.
\VS{30}And the
priests and the
Levites
purified themselves, and
purified the
people, and the
gates, and the
wall.
\VS{31}Then I brought
up the
princes of
Judah upon the
wall, and
appointed
two
great
{\ADD{companies of them that gave}}
thanks,
{\ADD{whereof one}}
went on the right
hand upon the
wall toward the
dung
gate:
\VS{32}And
after them
went
Hoshaiah, and
half of the
princes of
Judah,
\VS{33}And
Azariah,
Ezra, and
Meshullam,
\VS{34}Judah, and
Benjamin, and
Shemaiah, and
Jeremiah,
\VS{35}And
{\ADD{certain}} of the
priests’
sons with
trumpets;
{\ADD{namely}},
Zechariah the
son of
Jonathan, the
son of
Shemaiah, the
son of
Mattaniah, the
son of
Michaiah, the
son of
Zaccur, the
son of
Asaph:
\VS{36}And his
brethren,
Shemaiah, and
Azarael,
Milalai,
Gilalai,
Maai,
Nethaneel, and
Judah,
Hanani, with the
musical
instruments of
David the
man of
God, and
Ezra the
scribe
before them.
\VS{37}And at the
fountain
gate, which was over against them, they went
up by the
stairs of the
city of
David, at the going
up of the
wall, above the
house of
David, even unto the
water
gate
eastward.
\VS{38}And the
other
{\ADD{company of them that gave}}
thanks
went over
against
{\ADD{them}}, and I
after them, and the
half of the
people upon the
wall, from beyond the
tower of the
furnaces even unto the
broad
wall;
\VS{39}And from above the
gate of
Ephraim, and above the
old
gate, and above the
fish
gate, and the
tower of
Hananeel, and the
tower of
Meah, even unto the
sheep
gate: and they stood
still in the
prison
gate.
\VS{40}So
stood the
two
{\ADD{companies of them that gave}}
thanks in the
house of
God, and I, and the
half of the
rulers with me:
\VS{41}And the
priests;
Eliakim,
Maaseiah,
Miniamin,
Michaiah,
Elioenai,
Zechariah,
{\ADD{and}}
Hananiah, with
trumpets;
\VS{42}And
Maaseiah, and
Shemaiah, and
Eleazar, and
Uzzi, and
Jehohanan, and
Malchijah, and
Elam, and
Ezer. And the
singers sang
loud, with
Jezrahiah
{\ADD{their}}
overseer.\FTNT{‡‡‡}{{\FR 12.42 }
sang…: Heb. made their voice to be heard}
\VS{43}Also that
day they
offered
great
sacrifices, and
rejoiced: for
God had made them
rejoice with
great
joy: the
wives also and the
children
rejoiced: so that the
joy of
Jerusalem was
heard even afar
off.
\par }{\PP \VS{44}¶ And at that
time were
some
appointed over the
chambers for the
treasures, for the
offerings, for the
firstfruits, and for the
tithes, to
gather into them out of the
fields of the
cities the
portions of the
law for the
priests and
Levites: for
Judah
rejoiced for the
priests and for the
Levites that
waited.\FTNT{§§§}{{\FR 12.44 }
of the law: that is, appointed by the law}\FTNT{*}{{\FR 12.44 }
for Judah…: Heb. for the joy of Judah}\FTNT{†}{{\FR 12.44 }
waited: Heb. stood}
\VS{45}And both the
singers and the
porters
kept the
ward of their
God, and the
ward of the
purification, according to the
commandment of
David,
{\ADD{and}} of
Solomon his
son.
\VS{46}For in the
days of
David and
Asaph of
old
{\ADD{there were}}
chief of the
singers, and
songs of
praise and
thanksgiving unto
God.
\VS{47}And all
Israel in the
days of
Zerubbabel, and in the
days of
Nehemiah,
gave the
portions of the
singers and the
porters, every
day his
portion: and they
sanctified
{\ADD{holy things}} unto the
Levites; and the
Levites
sanctified
{\ADD{them}} unto the
children of
Aaron.\FTNT{‡}{{\FR 12.47 }
sanctified: that is, set apart}

\par }\Chap{13}{\PP \VerseOne{1}On that
day they
read in the
book of
Moses in the
audience of the
people; and therein was
found
written, that the
Ammonite and the
Moabite should not
come into the
congregation of
God
for
ever;\FTNT{*}{{\FR 13.1 }
they…: Heb. there was read}\FTNT{†}{{\FR 13.1 }
audience: Heb. ears}
\VS{2}Because they
met not the
children of
Israel with
bread and with
water, but
hired
Balaam against them, that he should
curse them: howbeit our
God
turned the
curse into a
blessing.
\VS{3}Now it came to pass, when they had
heard the
law, that they
separated from
Israel all the mixed
multitude.
\par }{\PP \VS{4}¶ And
before this,
Eliashib the
priest, having the
oversight of the
chamber of the
house of our
God,
{\ADD{was}}
allied unto
Tobiah:\FTNT{‡}{{\FR 13.4 }
having…: Heb. being set over}
\VS{5}And he had
prepared for him a
great
chamber, where
aforetime they
laid the meat
offerings, the
frankincense, and the
vessels, and the
tithes of the
corn, the new
wine, and the
oil, which was
commanded
{\ADD{to be given}} to the
Levites, and the
singers, and the
porters; and the
offerings of the
priests.\FTNT{§}{{\FR 13.5 }
which…: Heb. the commandment of the Levites}
\VS{6}But in all this
{\ADD{time}} was not I at
Jerusalem: for in the
two and
thirtieth
year of
Artaxerxes
king of
Babylon
came I unto the
king, and
after certain
days obtained I
leave of the
king:\FTNT{**}{{\FR 13.6 }
after…: Heb. at the end of days}\FTNT{††}{{\FR 13.6 }
obtained…: or, I earnestly requested}
\VS{7}And I
came to
Jerusalem, and
understood of the
evil that
Eliashib
did for
Tobiah, in
preparing him a
chamber in the
courts of the
house of
God.
\VS{8}And it
grieved me
sore: therefore I cast
forth all the
household
stuff of
Tobiah out
of the
chamber.
\VS{9}Then I
commanded, and they
cleansed the
chambers: and thither brought I
again the
vessels of the
house of
God, with the meat
offering and the
frankincense.
\par }{\PP \VS{10}¶ And I
perceived that the
portions of the
Levites had not been
given
{\ADD{them:}} for the
Levites and the
singers, that
did the
work, were
fled every
one to his
field.
\VS{11}Then
contended I with the
rulers, and
said, Why is the
house of
God
forsaken? And I gathered them
together, and
set them in their
place.\FTNT{‡‡}{{\FR 13.11 }
place: Heb. standing}
\VS{12}Then
brought all
Judah the
tithe of the
corn and the new
wine and the
oil unto the
treasuries.\FTNT{§§}{{\FR 13.12 }
treasuries: or, storehouses}
\VS{13}And I made
treasurers over the
treasuries,
Shelemiah the
priest, and
Zadok the
scribe, and of the
Levites,
Pedaiah: and next to
them
{\ADD{was}}
Hanan the
son of
Zaccur, the
son of
Mattaniah: for they were
counted
faithful, and their office
{\ADD{was}} to
distribute unto their
brethren.\FTNT{***}{{\FR 13.13 }
treasuries: or, storehouses}\FTNT{†††}{{\FR 13.13 }
next…: Heb. at their hand}\FTNT{‡‡‡}{{\FR 13.13 }
their office…: Heb. it was upon them}
\VS{14}Remember me, O my
God, concerning this, and wipe not
out my
good deeds that I have
done for the
house of my
God, and for the
offices thereof.\FTNT{§§§}{{\FR 13.14 }
good…: Heb. kindnesses}\FTNT{*}{{\FR 13.14 }
offices: or, observations}
\par }{\PP \VS{15}¶ In those
days
saw I in
Judah
{\ADD{some}}
treading wine
presses on the
sabbath, and bringing
in
sheaves, and
lading
asses; as also
wine,
grapes, and
figs, and all
{\ADD{manner of}}
burdens, which they
brought into
Jerusalem on the
sabbath
day: and I
testified
{\ADD{against them}} in the
day wherein they
sold
victuals.
\VS{16}There
dwelt men of
Tyre also therein, which
brought
fish, and all manner of
ware, and
sold on the
sabbath unto the
children of
Judah, and in
Jerusalem.
\VS{17}Then I
contended with the
nobles of
Judah, and
said unto them, What
evil
thing
{\ADD{is}} this that ye
do, and
profane the
sabbath
day?
\VS{18}Did not your
fathers thus, and
did not our
God
bring all this
evil upon us, and upon this
city? yet ye
bring
more
wrath upon
Israel by
profaning the
sabbath.
\VS{19}And it came to pass, that when the
gates of
Jerusalem began to be
dark
before the
sabbath, I
commanded that the
gates should be
shut, and
charged that they should not be
opened till
after the
sabbath: and
{\ADD{some}} of my
servants
set I at the
gates,
{\ADD{that}} there should no
burden be brought
in on the
sabbath
day.
\VS{20}So the
merchants and
sellers of all kind of
ware
lodged
without
Jerusalem
once or
twice.
\VS{21}Then I
testified against them, and
said unto them, Why
lodge ye
about the
wall? if ye do
{\ADD{so}}
again, I will
lay
hands on you. From that
time forth
came they no
{\ADD{more}} on the
sabbath.\FTNT{†}{{\FR 13.21 }
about: Heb. before}
\VS{22}And I
commanded the
Levites that they should
cleanse themselves, and
{\ADD{that}} they should
come
{\ADD{and}}
keep the
gates, to
sanctify the
sabbath
day.
Remember me, O my
God,
{\ADD{concerning}} this also, and
spare me according to the
greatness of thy
mercy.\FTNT{‡}{{\FR 13.22 }
greatness: or, multitude}
\par }{\PP \VS{23}¶ In those
days also
saw I
Jews
{\ADD{that}} had
married
wives of
Ashdod, of
Ammon,
{\ADD{and}} of
Moab:\FTNT{§}{{\FR 13.23 }
had…: Heb. had made to dwell with them}
\VS{24}And their
children
spake
half in the speech of
Ashdod, and
could not
speak in the
Jews’ language, but according to the
language of
each
people.\FTNT{**}{{\FR 13.24 }
could…: Heb. they discerned not to speak}\FTNT{††}{{\FR 13.24 }
of each…: Heb. of people and people}
\VS{25}And I
contended with them, and
cursed them, and
smote
certain of them, and plucked off their
hair, and made them
swear by
God,
{\ADD{saying}}, Ye shall not
give your
daughters unto their
sons, nor
take their
daughters unto your
sons, or for yourselves.\FTNT{‡‡}{{\FR 13.25 }
cursed: or, reviled}
\VS{26}Did not
Solomon
king of
Israel
sin by these things? yet among
many
nations was there no
king like him, who was
beloved of his
God, and
God
made him
king over all
Israel: nevertheless even
him did
outlandish
women cause to
sin.
\VS{27}Shall we then
hearken unto you to
do all this
great
evil, to
transgress against our
God in
marrying
strange
wives?
\VS{28}And
{\ADD{one}} of the
sons of
Joiada, the
son of
Eliashib the
high
priest,
{\ADD{was}} son in
law to
Sanballat the
Horonite: therefore I
chased him from me.
\VS{29}Remember them, O my
God, because they have
defiled the
priesthood, and the
covenant of the
priesthood, and of the
Levites.\FTNT{§§}{{\FR 13.29 }
because…: Heb. for the defilings}
\VS{30}Thus
cleansed I them from all
strangers, and
appointed the
wards of the
priests and the
Levites, every
one in his
business;
\VS{31}And for the
wood
offering, at
times
appointed, and for the
firstfruits.
Remember me, O my
God, for
good.
\par }