\NormalFont\ShortTitle{Philemon}
{\MT THE EPISTLE OF PAUL THE APOSTLE TO PHILEMON

\par }\OneChap {\PP \VerseOne{1}Paul, a
prisoner of
Jesus
Christ,
and
Timothy
{\ADD{our}}
brother, unto
Philemon
our dearly
beloved,
and
fellowlabourer,
\VS{2}And
to
{\ADD{our}}
beloved
Apphia,
and
Archippus
our
fellowsoldier,
and to the
church
in
thy
house:
\VS{3}Grace to
you,
and
peace,
from
God
our
Father
and the
Lord
Jesus
Christ.
\par }{\PP \VS{4}I
thank
my
God,
making
mention of
thee
always
in
my
prayers,
\VS{5}Hearing of
thy
love
and
faith,
which thou
hast
toward the
Lord
Jesus,
and
toward
all
saints;
\VS{6}That the
communication of
thy
faith may
become
effectual
by the
acknowledging of
every good
thing which
is
in
you
in
Christ
Jesus.
\VS{7}For we
have
great
joy
and
consolation
in
thy
love,
because the
bowels of the
saints are
refreshed
by
thee,
brother.
\VS{8}Wherefore, though I might
be
much
bold
in
Christ to
enjoin
thee that which is
convenient,
\VS{9}Yet
for
love’s
sake
I
rather
beseech
{\ADD{thee}},
being such an
one
as
Paul the
aged,
and
now
also a
prisoner of
Jesus
Christ.
\VS{10}I
beseech
thee
for
my
son
Onesimus,
whom I have
begotten
in
my
bonds:
\VS{11}Which in time
past
was to
thee
unprofitable,
but
now
profitable to
thee
and to
me:
\VS{12}Whom I have sent
again:
thou
therefore
receive
him,
that
is, mine
own
bowels:
\VS{13}Whom
I
would have
retained
with
me,
that
in
thy
stead he might have
ministered unto
me
in the
bonds of the
gospel:
\VS{14}But
without
thy
mind would
I
do
nothing;
that
thy
benefit should
not
be as it
were
of
necessity,
but
willingly.
\VS{15}For
perhaps
he
therefore
departed
for a
season,
that thou shouldest
receive
him for
ever;
\VS{16}Not
now
as a
servant,
but
above a
servant, a
brother
beloved,
specially to
me,
but how
much
more unto
thee,
both
in the
flesh,
and
in the
Lord?
\VS{17}If thou
count
me
therefore a
partner,
receive
him
as
myself.
\VS{18}If he hath
wronged
thee,
or
oweth
{\ADD{thee}}
ought,
put
that on
mine
account;
\VS{19}I
Paul have
written
{\ADD{it}} with mine
own
hand,
I will
repay
{\ADD{it}}:
albeit I
do
not
say to
thee
how thou
owest unto
me
even thine own
self
besides.
\VS{20}Yea,
brother,
let
me have
joy of
thee
in the
Lord:
refresh
my
bowels
in the
Lord.
\par }{\PP \VS{21}Having
confidence
in
thy
obedience I
wrote unto
thee,
knowing
that thou wilt
also
do more
than I
say.
\VS{22}But
withal
prepare
me
also a
lodging:
for I
trust
that
through
your
prayers I shall be
given unto
you.
\par }{\PP \VS{23}There
salute
thee
Epaphras,
my
fellowprisoner
in
Christ
Jesus;
\VS{24}Marcus,
Aristarchus,
Demas,
Lucas,
my
fellowlabourers.
\VS{25}The
grace of
our
Lord
Jesus
Christ
{\ADD{be}}
with
your
spirit.
Amen.
\par }{\SH Written from Rome to Philemon, by Onesimus, a servant.
\par }