\NormalFont\ShortTitle{Philippians}
{\MT THE EPISTLE OF PAUL THE APOSTLE TO THE PHILIPPIANS

\par }\ChapOne{1}{\PP \VerseOne{1}Paul
and
Timotheus, the
servants of
Jesus
Christ, to
all the
saints
in
Christ
Jesus which
are
at
Philippi,
with the
bishops
and
deacons:
\VS{2}Grace
{\ADD{be}} unto
you,
and
peace,
from
God
our
Father,
and
{\ADD{from}} the
Lord
Jesus
Christ.
\VS{3}I
thank
my
God
upon
every
remembrance of
you,
\VS{4}Always
in
every
prayer of
mine
for
you
all
making
request
with
joy,
\VS{5}For
your
fellowship
in the
gospel
from
the
first
day
until
now;
\VS{6}Being confident
of
this very
thing,
that he which hath
begun
a
good
work
in
you will
perform
{\ADD{it}}
until the
day of
Jesus
Christ:
\VS{7}Even
as it
is
meet for
me to
think
this
of
you
all,
because
I
have
you
in my
heart; inasmuch
as
both
in
my
bonds,
and in the
defence
and
confirmation of the
gospel,
ye
all
are
partakers
of
my
grace.
\VS{8}For
God
is
my
record, how
greatly I
long after
you
all
in the
bowels of
Jesus
Christ.
\VS{9}And
this I
pray,
that
your
love may
abound
yet
more
and
more
in
knowledge
and
{\ADD{in}}
all
judgment;
\VS{10}That
ye may approve
things that are
excellent;
that ye may
be
sincere
and without
offence
till the
day of
Christ;
\VS{11}Being
filled with the
fruits of
righteousness, which
are
by
Jesus
Christ,
unto the
glory
and
praise of
God.
\par }{\PP \VS{12}But I
would
ye should
understand,
brethren,
that the
things
{\ADD{which happened}}
unto
me have fallen
out
rather
unto the
furtherance of the
gospel;
\VS{13}So
that
my
bonds
in
Christ
are
manifest
in
all the
palace,
and
in
all
other
{\ADD{places}};
\VS{14}And
many of the
brethren
in the
Lord, waxing
confident
by
my
bonds,
are much
more
bold to
speak the
word without
fear.
\VS{15}Some
indeed
preach
Christ
even
of
envy
and
strife;
and
some
also
of good
will:
\VS{16}The
one
preach
Christ
of
contention,
not
sincerely,
supposing to
add
affliction
to
my
bonds:
\VS{17}But the
other
of
love,
knowing
that I am
set
for the
defence of the
gospel.
\VS{18}What
then?
notwithstanding,
every
way,
whether in
pretence,
or in
truth,
Christ is
preached;
and
I
therein do
rejoice,
yea,
and will
rejoice.
\VS{19}For I
know
that
this shall
turn
to
my
salvation
through
your
prayer,
and the
supply of the
Spirit of
Jesus
Christ,
\VS{20}According
to
my earnest
expectation
and
{\ADD{my}}
hope,
that
in
nothing I shall be
ashamed,
but
{\ADD{that}}
with
all
boldness,
as
always,
{\ADD{so}}
now
also
Christ shall be
magnified
in
my
body,
whether
{\ADD{it be}}
by
life,
or
by
death.
\VS{21}For to
me to
live
{\ADD{is}}
Christ,
and to
die
{\ADD{is}}
gain.
\VS{22}But
if I
live
in the
flesh,
this
{\ADD{is}} the
fruit
of
my
labour:
yet
what I shall
choose I
wot
not.
\VS{23}For I am in a
strait
betwixt
two,
having a
desire
to
depart,
and to
be
with
Christ; which
is
far
better:
\VS{24}Nevertheless to
abide
in the
flesh
{\ADD{is}} more
needful
for
you.
\VS{25}And
having
this
confidence, I
know
that I shall
abide
and continue
with
you
all
for
your
furtherance
and
joy of
faith;
\VS{26}That
your
rejoicing may be more
abundant
in
Jesus
Christ
for
me
by
my
coming
to
you
again.
\VS{27}Only let your conversation
be as it
becometh the
gospel of
Christ:
that
whether I
come
and
see
you, or
else be
absent, I may
hear of
your
affairs,
that ye stand
fast
in
one
spirit,
with
one
mind striving
together for the
faith of the
gospel;
\VS{28}And
in
nothing
terrified
by your
adversaries:
which
is to
them an evident
token of
perdition,
but to
you of
salvation,
and
that
of
God.
\VS{29}For unto
you it is
given in the
behalf of
Christ,
not
only to
believe
on
him,
but
also to
suffer
for
his
sake;
\VS{30}Having
the
same
conflict
which ye
saw
in
me,
and
now
hear
{\ADD{to be}}
in
me.

\par }\Chap{2}{\PP \VerseOne{1}If
{\ADD{there be}}
therefore
any
consolation
in
Christ,
if
any
comfort of
love,
if
any
fellowship of the
Spirit,
if
any
bowels
and
mercies,
\VS{2}Fulfil
ye
my
joy,
that ye
be
likeminded,
having
the
same
love,
{\ADD{being}} of one
accord,
of
one
mind.
\VS{3}{\ADD{Let}}
nothing
{\ADD{be done}}
through
strife
or
vainglory;
but in lowliness of
mind
let
each
esteem
other
better than
themselves.
\VS{4}Look
not every
man on his own
things,
but every
man
also on the things of
others.
\VS{5}Let
this mind
be
in
you, which
was
also
in
Christ
Jesus:
\VS{6}Who,
being
in the
form of
God, thought
it
not
robbery to
be
equal with
God:
\VS{7}But
made
himself of no
reputation, and took upon
him the
form of a
servant, and was
made
in the
likeness of
men:
\VS{8}And being
found in
fashion
as a
man, he
humbled
himself, and
became
obedient
unto
death,
even the
death of the
cross.
\VS{9}Wherefore
God
also hath highly
exalted
him,
and
given
him a
name which
is
above
every
name:
\VS{10}That
at the
name of
Jesus
every
knee should
bow,
of
{\ADD{things}} in
heaven,
and
{\ADD{things}} in
earth,
and
{\ADD{things}} under the
earth;
\VS{11}And
{\ADD{that}}
every
tongue should
confess
that
Jesus
Christ
{\ADD{is}}
Lord,
to the
glory of
God the
Father.
\par }{\PP \VS{12}Wherefore,
my
beloved,
as ye
have
always
obeyed,
not
as
in
my
presence
only,
but
now
much
more
in
my
absence, work
out your
own
salvation
with
fear
and
trembling.
\VS{13}For it
is
God
which
worketh
in
you
both to
will
and to
do
of
{\ADD{his}} good
pleasure.
\VS{14}Do all
things
without
murmurings
and
disputings:
\VS{15}That ye may
be
blameless
and
harmless, the
sons of
God, without
rebuke,
in the
midst of
a
crooked
and
perverse
nation,
among
whom ye
shine
as
lights
in the
world;
\VS{16}Holding
forth the
word of
life;
that
I may
rejoice
in the
day of
Christ,
that I
have
not
run
in
vain,
neither
laboured
in
vain.
\VS{17}Yea,
and
if I be
offered
upon the
sacrifice
and
service
of
your
faith, I
joy,
and
rejoice with
you
all.
\VS{18}For the same
cause
also
do
ye
joy,
and rejoice
with
me.
\par }{\PP \VS{19}But I
trust
in the
Lord
Jesus to
send
Timotheus
shortly unto
you,
that I
also may be of good
comfort, when I
know
your
state.
\VS{20}For I
have no
man
likeminded,
who
will
naturally care
for
your
state.
\VS{21}For
all
seek their
own,
not the things which
are
Jesus
Christ’s.
\VS{22}But ye
know the
proof of
him,
that,
as a
son with the
father, he hath
served
with
me
in the
gospel.
\VS{23}Him
therefore I
hope to
send
presently, so soon
as I shall
see how it will go with me.
\VS{24}But I
trust
in the
Lord
that
I
also
myself shall
come
shortly.
\VS{25}Yet I supposed
it
necessary to
send
to
you
Epaphroditus,
my
brother,
and companion in
labour,
and
fellowsoldier,
but
your
messenger,
and he that
ministered
to
my
wants.
\VS{26}For he longed
after
you
all,
and was full of
heaviness, because
that ye had
heard
that he had been
sick.
\VS{27}For
indeed he was
sick nigh
unto
death:
but
God had mercy
on
him;
and
not on
him
only,
but on
me
also,
lest I should
have
sorrow
upon
sorrow.
\VS{28}I
sent
him
therefore the more
carefully,
that, when ye
see
him
again, ye may
rejoice, and that
I may
be the less
sorrowful.
\VS{29}Receive
him
therefore
in the
Lord
with
all
gladness;
and
hold
such in
reputation:
\VS{30}Because
for the
work of
Christ he was
nigh
unto
death, not
regarding his
life, to
supply
your
lack of
service
toward
me.

\par }\Chap{3}{\PP \VerseOne{1}Finally,
my
brethren,
rejoice
in the
Lord. To
write the same
things to
you, to
me
indeed
{\ADD{is}}
not
grievous,
but for
you
{\ADD{it is}}
safe.
\par }{\PP \VS{2}Beware of
dogs,
beware
of
evil
workers,
beware of the
concision.
\VS{3}For
we
are the
circumcision, which
worship
God in the
spirit,
and
rejoice
in
Christ
Jesus,
and
have
no
confidence
in the
flesh.
\VS{4}Though
I
might
also
have
confidence
in the
flesh.
If
any
other
man
thinketh that he hath whereof he might
trust
in the
flesh,
I
more:
\VS{5}Circumcised the eighth
day,
of the
stock of
Israel,
{\ADD{of}} the
tribe of
Benjamin, an
Hebrew
of the
Hebrews; as
touching the
law, a
Pharisee;
\VS{6}Concerning
zeal,
persecuting the
church;
touching the
righteousness
which
is
in the
law,
blameless.
\VS{7}But what
things
were
gain to
me,
those I
counted
loss
for
Christ.
\VS{8}Yea
doubtless,
and I
count all
things
{\ADD{but}}
loss
for the
excellency of the
knowledge of
Christ
Jesus
my
Lord:
for
whom I have suffered the loss
of all
things,
and do
count them
{\ADD{but}}
dung,
that I may
win
Christ,
\VS{9}And be
found
in
him,
not
having mine
own
righteousness, which
is
of the
law,
but that which
is
through the
faith of
Christ, the
righteousness which
is
of
God
by
faith:
\VS{10}That I may
know
him,
and the
power
of
his
resurrection,
and the
fellowship
of
his
sufferings, being made conformable
unto
his
death;
\VS{11}If by any
means I might
attain
unto the
resurrection of the
dead.
\VS{12}Not as
though I
had
already
attained,
either
were
already
perfect:
but I follow
after,
if
that I may
apprehend that
for
which
also I am
apprehended
of
Christ
Jesus.
\VS{13}Brethren,
I
count
not
myself to have
apprehended:
but
{\ADD{this}} one
thing
{\ADD{I do}},
forgetting those things which
are
behind,
and reaching
forth unto those things which
are
before,
\VS{14}I
press
toward the
mark
for the
prize of
the
high
calling of
God
in
Christ
Jesus.
\VS{15}Let
us
therefore, as many
as be
perfect, be
thus
minded:
and
if in any
thing ye
be
otherwise
minded,
God shall
reveal
even
this unto
you.
\VS{16}Nevertheless,
whereto we have already
attained, let us
walk by
the
same
rule, let us
mind the same
thing.
\VS{17}Brethren,
be followers
together of
me,
and
mark them which
walk so
as ye
have
us
for an
ensample.
\VS{18}(For
many
walk, of
whom I have
told
you
often,
and
now tell
you
even
weeping,
{\ADD{that they are}} the
enemies of the
cross of
Christ:
\VS{19}Whose
end
{\ADD{is}}
destruction,
whose
God
{\ADD{is their}}
belly,
and
{\ADD{whose}}
glory
{\ADD{is}}
in
their
shame,
who
mind earthly
things.)
\VS{20}For
our
conversation
is
in
heaven;
from
whence
also we
look for the
Saviour, the
Lord
Jesus
Christ:
\VS{21}Who shall
change
our
vile
body,
that
it may
be fashioned like
unto
his
glorious
body, according
to the
working
whereby
he is
able
even to
subdue all
things unto
himself.

\par }\Chap{4}{\PP \VerseOne{1}Therefore,
my
brethren dearly
beloved
and longed
for,
my
joy
and
crown,
so stand
fast
in the
Lord,
{\ADD{my}} dearly
beloved.
\VS{2}I
beseech
Euodias,
and
beseech
Syntyche, that they be of the same mind
in the
Lord.
\VS{3}And I
intreat
thee
also,
true
yokefellow,
help those
women
which
laboured with
me
in the
gospel,
with
Clement
also,
and
{\ADD{with}}
other
my
fellowlabourers,
whose
names
{\ADD{are}}
in the
book of
life.
\VS{4}Rejoice
in the
Lord
alway:
{\ADD{and}}
again I
say,
Rejoice.
\VS{5}Let
your
moderation be
known
unto
all
men. The
Lord
{\ADD{is}} at
hand.
\VS{6}Be
careful for
nothing;
but
in every
thing by
prayer
and
supplication
with
thanksgiving
let
your
requests be made
known
unto
God.
\VS{7}And the
peace of
God, which
passeth
all
understanding, shall
keep
your
hearts
and
minds
through
Christ
Jesus.
\VS{8}Finally,
brethren, whatsoever
things
are
true, whatsoever
things
{\ADD{are}}
honest, whatsoever
things
{\ADD{are}}
just, whatsoever
things
{\ADD{are}}
pure, whatsoever
things
{\ADD{are}}
lovely, whatsoever
things
{\ADD{are}} of good
report;
if
{\ADD{there be}}
any
virtue,
and
if
{\ADD{there be}}
any
praise, think
on these
things.
\VS{9}Those
things,
which ye
have
both
learned,
and
received,
and
heard,
and
seen
in
me,
do:
and the
God of
peace shall
be
with
you.
\par }{\PP \VS{10}But I
rejoiced
in the
Lord
greatly,
that
now at the
last your
care
of
me hath
flourished again;
wherein ye
were
also
careful,
but ye lacked
opportunity.
\VS{11}Not
that I
speak in respect
of
want:
for
I have
learned,
in
whatsoever state I
am,
{\ADD{therewith}} to
be
content.
\VS{12}I
know
both
how to be
abased,
and I know
how to
abound: every
where
and
in all
things I am
instructed
both to be
full
and to be
hungry,
both to
abound
and to suffer
need.
\VS{13}I can
do all
things
through
Christ
which
strengtheneth
me.
\VS{14}Notwithstanding ye
have
well
done, that ye did
communicate
with
my
affliction.
\VS{15}Now
ye
Philippians
know
also,
that
in the
beginning of the
gospel,
when I
departed
from
Macedonia,
no
church communicated
with
me
as
concerning
giving
and
receiving,
but
ye
only.
\VS{16}For
even
in
Thessalonica ye
sent
once
and
again
unto
my
necessity.
\VS{17}Not
because I
desire a
gift:
but I
desire
fruit that may
abound
to
your
account.
\VS{18}But I
have
all,
and
abound: I am
full, having
received
of
Epaphroditus the
things
{\ADD{which were sent}}
from
you, an
odour of a sweet
smell, a
sacrifice
acceptable,
wellpleasing to
God.
\VS{19}But
my
God shall
supply
all
your
need according
to
his
riches
in
glory
by
Christ
Jesus.
\VS{20}Now unto
God
and
our
Father
{\ADD{be}}
glory
for
ever and
ever.
Amen.
\par }{\PP \VS{21}Salute
every
saint
in
Christ
Jesus. The brethren which
are
with
me
greet
you.
\VS{22}All the
saints
salute
you,
chiefly they that
are
of
Caesar’s
household.
\VS{23}The
grace
of
our
Lord
Jesus
Christ
{\ADD{be}}
with
you
all.
Amen.
\par }{\SH It was written to the Philippians from Rome by Epaphroditus.
\par }