\NormalFont\ShortTitle{Proverbs}
{\MT The Proverbs

\par }\ChapOne{1}{\PP \VerseOne{1}The
proverbs of
Solomon the
son of
David,
king of
Israel;
\VS{2}To
know
wisdom and
instruction; to
perceive the
words of
understanding;
\VS{3}To
receive the
instruction of
wisdom,
justice, and
judgment, and
equity;\FTNT{*}{{\FR 1.3 }
equity: Heb. equities}
\VS{4}To
give
subtilty to the
simple, to the young
man
knowledge and
discretion.\FTNT{†}{{\FR 1.4 }
discretion: or, advisement}
\VS{5}A
wise
{\ADD{man}} will
hear, and will
increase
learning; and a man of
understanding shall
attain unto wise
counsels:
\VS{6}To
understand a
proverb, and the
interpretation; the
words of the
wise, and their dark
sayings.\FTNT{‡}{{\FR 1.6 }
the interpretation: or, an eloquent speech}
\par }{\PP \VS{7}¶ The
fear of the
{\ND{
LORD}}
{\ADD{is}} the
beginning of
knowledge:
{\ADD{but}}
fools
despise
wisdom and
instruction.\FTNT{§}{{\FR 1.7 }
the beginning: or, the principal part}
\VS{8}My
son,
hear the
instruction of thy
father, and
forsake not the
law of thy
mother:
\VS{9}For they
{\ADD{shall be}} an
ornament of
grace unto thy
head, and
chains about thy
neck.\FTNT{**}{{\FR 1.9 }
an…: Heb. an adding}
\par }{\PP \VS{10}¶ My
son, if
sinners
entice thee,
consent thou not.
\VS{11}If they
say,
Come with us, let us lay
wait for
blood, let us
lurk privily for the
innocent without
cause:
\VS{12}Let us swallow them
up
alive as the
grave; and
whole, as those that go
down into the
pit:
\VS{13}We shall
find all
precious
substance, we shall
fill our
houses with
spoil:
\VS{14}Cast in thy
lot
among us; let us all have
one
purse:
\VS{15}My
son,
walk not thou in the
way with them;
refrain thy
foot from their
path:
\VS{16}For their
feet
run to
evil, and make
haste to
shed
blood.
\VS{17}Surely in
vain the
net is
spread in the
sight of any
bird.\FTNT{††}{{\FR 1.17 }
in the…: Heb. in the eyes of every thing that hath a wing}
\VS{18}And they lay
wait for their
{\ADD{own}}
blood; they lurk
privily for their
{\ADD{own}}
lives.
\VS{19}So
{\ADD{are}} the
ways of every one that is
greedy of
gain;
{\ADD{which}} taketh
away the
life of the
owners thereof.
\par }{\PP \VS{20}¶
Wisdom
crieth
without; she
uttereth her
voice in the
streets:\FTNT{‡‡}{{\FR 1.20 }
Wisdom: Heb. Wisdoms, that is, Excellent wisdom}
\VS{21}She
crieth in the chief
place of
concourse, in the
openings of the
gates: in the
city she
uttereth her
words,
{\ADD{saying}},
\VS{22}How long, ye simple
ones, will ye
love
simplicity? and the
scorners
delight in their
scorning, and
fools
hate
knowledge?
\VS{23}Turn you at my
reproof: behold, I will pour
out my
spirit unto you, I will make
known my
words unto you.
\par }{\PP \VS{24}¶ Because I have
called, and ye
refused; I have stretched
out my
hand, and no man
regarded;
\VS{25}But ye have set at
nought all my
counsel, and
would none of my
reproof:
\VS{26}I also will
laugh at your
calamity; I will
mock when your
fear
cometh;
\VS{27}When your
fear
cometh as
desolation, and your
destruction
cometh as a
whirlwind; when
distress and
anguish
cometh upon you.
\VS{28}Then shall they
call upon me, but I will not
answer; they shall seek me
early, but they shall not
find me:
\VS{29}For that they
hated
knowledge, and did not
choose the
fear of the
{\ND{
LORD}}:
\VS{30}They
would none of my
counsel: they
despised all my
reproof.
\VS{31}Therefore shall they
eat of the
fruit of their own
way, and be
filled with their own
devices.
\VS{32}For the turning
away of the
simple shall
slay them, and the
prosperity of
fools shall
destroy them.\FTNT{§§}{{\FR 1.32 }
turning…: or, ease of the simple}
\VS{33}But whoso
hearkeneth unto me shall
dwell
safely, and shall be
quiet from
fear of
evil.

\par }\Chap{2}{\PP \VerseOne{1}My
son, if thou wilt
receive my
words, and
hide my
commandments with thee;
\VS{2}So that thou
incline thine
ear unto
wisdom,
{\ADD{and}}
apply thine
heart to
understanding;
\VS{3}Yea, if thou
criest after
knowledge,
{\ADD{and}} liftest
up thy
voice for
understanding;\FTNT{*}{{\FR 2.3 }
liftest…: Heb. givest thy voice}
\VS{4}If thou
seekest her as
silver, and
searchest for her as
{\ADD{for}} hid
treasures;
\VS{5}Then shalt thou
understand the
fear of the
{\ND{
LORD}}, and
find the
knowledge of
God.
\VS{6}For the
{\ND{
LORD}}
giveth
wisdom: out of his
mouth
{\ADD{cometh}}
knowledge and
understanding.
\VS{7}He layeth
up sound
wisdom for the
righteous:
{\ADD{he is}} a
buckler to them that
walk
uprightly.
\VS{8}He
keepeth the
paths of
judgment, and
preserveth the
way of his
saints.
\VS{9}Then shalt thou
understand
righteousness, and
judgment, and
equity;
{\ADD{yea}}, every
good
path.
\par }{\PP \VS{10}¶ When
wisdom
entereth into thine
heart, and
knowledge is
pleasant unto thy
soul;
\VS{11}Discretion shall
preserve thee,
understanding shall
keep thee:
\VS{12}To
deliver thee from the
way of the
evil
{\ADD{man}}, from the
man that
speaketh froward
things;
\VS{13}Who
leave the
paths of
uprightness, to
walk in the
ways of
darkness;
\VS{14}Who
rejoice to
do
evil,
{\ADD{and}}
delight in the
frowardness of the
wicked;
\VS{15}Whose
ways
{\ADD{are}}
crooked, and
{\ADD{they}}
froward in their
paths:
\VS{16}To
deliver thee from the
strange
woman,
{\ADD{even}} from the
stranger
{\ADD{which}}
flattereth with her
words;
\VS{17}Which
forsaketh the
guide of her
youth, and
forgetteth the
covenant of her
God.
\VS{18}For her
house
inclineth unto
death, and her
paths unto the
dead.
\VS{19}None that
go unto her return
again, neither take they
hold of the
paths of
life.
\VS{20}That thou mayest
walk in the
way of
good
{\ADD{men}}, and
keep the
paths of the
righteous.
\VS{21}For the
upright shall
dwell in the
land, and the
perfect shall
remain in it.
\VS{22}But the
wicked shall be cut
off from the
earth, and the
transgressors shall be rooted
out of it.\FTNT{†}{{\FR 2.22 }
rooted: or, plucked up}

\par }\Chap{3}{\PP \VerseOne{1}My
son,
forget not my
law; but let thine
heart
keep my
commandments:
\VS{2}For
length of
days, and
long
life, and
peace, shall they
add to thee.\FTNT{*}{{\FR 3.2 }
long…: Heb. years of life}
\VS{3}Let not
mercy and
truth
forsake thee:
bind them about thy
neck;
write them upon the
table of thine
heart:
\VS{4}So shalt thou
find
favour and
good
understanding in the
sight of
God and
man.\FTNT{†}{{\FR 3.4 }
good…: or, good success}
\par }{\PP \VS{5}¶
Trust in the
{\ND{
LORD}} with all thine
heart; and
lean not unto thine own
understanding.
\VS{6}In all thy
ways
acknowledge him, and he shall
direct thy
paths.
\par }{\PP \VS{7}¶ Be not
wise in thine own
eyes:
fear the
{\ND{
LORD}}, and
depart from
evil.
\VS{8}It shall be
health to thy
navel, and
marrow to thy
bones.\FTNT{‡}{{\FR 3.8 }
health: Heb. medicine}\FTNT{§}{{\FR 3.8 }
marrow: Heb. watering, or, moistening}
\VS{9}Honour the
{\ND{
LORD}} with thy
substance, and with the
firstfruits of all thine
increase:
\VS{10}So shall thy
barns be
filled with
plenty, and thy
presses shall burst
out with new
wine.
\par }{\PP \VS{11}¶ My
son,
despise not the
chastening of the
{\ND{
LORD}}; neither be
weary of his
correction:
\VS{12}For whom the
{\ND{
LORD}}
loveth he
correcteth; even as a
father the
son
{\ADD{in whom}} he
delighteth.
\par }{\PP \VS{13}¶
Happy
{\ADD{is}} the
man
{\ADD{that}}
findeth
wisdom, and the
man
{\ADD{that}}
getteth
understanding.\FTNT{**}{{\FR 3.13 }
that getteth…: Heb. that draweth out understanding}
\VS{14}For the
merchandise of it
{\ADD{is}}
better than the
merchandise of
silver, and the
gain thereof than fine
gold.
\VS{15}She
{\ADD{is}} more
precious than
rubies: and all the things thou canst
desire are not to be
compared unto her.
\VS{16}Length of
days
{\ADD{is}} in her right
hand;
{\ADD{and}} in her left
hand
riches and
honour.
\VS{17}Her
ways
{\ADD{are}}
ways of
pleasantness, and all her
paths
{\ADD{are}}
peace.
\VS{18}She
{\ADD{is}} a
tree of
life to them that lay
hold upon her: and
happy
{\ADD{is every one}} that
retaineth her.
\VS{19}The
{\ND{
LORD}} by
wisdom hath
founded the
earth; by
understanding hath he
established the
heavens.\FTNT{††}{{\FR 3.19 }
established: or, prepared}
\VS{20}By his
knowledge the
depths are broken
up, and the
clouds drop
down the
dew.
\par }{\PP \VS{21}¶ My
son, let not them
depart from thine
eyes:
keep sound
wisdom and
discretion:
\VS{22}So shall they be
life unto thy
soul, and
grace to thy
neck.
\VS{23}Then shalt thou
walk in thy
way
safely, and thy
foot shall not
stumble.
\VS{24}When thou liest
down, thou shalt not be
afraid: yea, thou shalt lie
down, and thy
sleep shall be
sweet.
\VS{25}Be not
afraid of
sudden
fear, neither of the
desolation of the
wicked, when it
cometh.
\VS{26}For the
{\ND{
LORD}} shall be thy
confidence, and shall
keep thy
foot from being
taken.
\par }{\PP \VS{27}¶
Withhold not
good from them to whom it is
due, when it is in the
power of thine
hand to
do
{\ADD{it}}.\FTNT{‡‡}{{\FR 3.27 }
them…: Heb. the owners thereof}
\VS{28}Say not unto thy
neighbour,
Go, and come
again, and to
morrow I will
give; when thou
hast it by thee.
\VS{29}Devise not
evil against thy
neighbour, seeing he
dwelleth
securely by thee.\FTNT{§§}{{\FR 3.29 }
Devise…: or, Practise no evil}
\par }{\PP \VS{30}¶
Strive not with a
man without
cause, if he have
done thee no
harm.
\par }{\PP \VS{31}¶
Envy thou not the
oppressor, and
choose none of his
ways.\FTNT{***}{{\FR 3.31 }
the oppressor: Heb. a man of violence}
\VS{32}For the
froward
{\ADD{is}}
abomination to the
{\ND{
LORD}}: but his
secret
{\ADD{is}} with the
righteous.
\par }{\PP \VS{33}¶ The
curse of the
{\ND{
LORD}}
{\ADD{is}} in the
house of the
wicked: but he
blesseth the
habitation of the
just.
\VS{34}Surely he
scorneth the
scorners: but he
giveth
grace unto the
lowly.
\VS{35}The
wise shall
inherit
glory: but
shame shall be the
promotion of
fools.\FTNT{†††}{{\FR 3.35 }
shall be…: Heb. exalteth the fools}

\par }\Chap{4}{\PP \VerseOne{1}Hear, ye
children, the
instruction of a
father, and
attend to
know
understanding.
\VS{2}For I
give you
good
doctrine,
forsake ye not my
law.
\VS{3}For I was my
father’s
son,
tender and
only
{\ADD{beloved}} in the
sight of my
mother.
\VS{4}He
taught me also, and
said unto me, Let thine
heart
retain my
words:
keep my
commandments, and
live.
\VS{5}Get
wisdom,
get
understanding:
forget
{\ADD{it}} not; neither
decline from the
words of my
mouth.
\VS{6}Forsake her not, and she shall
preserve thee:
love her, and she shall
keep thee.
\VS{7}Wisdom
{\ADD{is}} the principal
thing;
{\ADD{therefore}}
get
wisdom: and with all thy
getting get
understanding.
\VS{8}Exalt her, and she shall
promote thee: she shall bring thee to
honour, when thou dost
embrace her.
\VS{9}She shall
give to thine
head an
ornament of
grace: a
crown of
glory shall she
deliver to thee.\FTNT{*}{{\FR 4.9 }
a crown…: or, she shall compass thee with a crown of glory}
\VS{10}Hear, O my
son, and
receive my
sayings; and the
years of thy
life shall be
many.
\VS{11}I have
taught thee in the
way of
wisdom; I have
led thee in
right
paths.
\VS{12}When thou
goest, thy
steps shall not be
straitened; and when thou
runnest, thou shalt not
stumble.
\VS{13}Take fast
hold of
instruction; let
{\ADD{her}} not
go:
keep her; for she
{\ADD{is}} thy
life.
\par }{\PP \VS{14}¶
Enter not into the
path of the
wicked, and
go not in the
way of
evil
{\ADD{men}}.
\VS{15}Avoid it,
pass not by it,
turn from it, and pass
away.
\VS{16}For they
sleep not, except they have done
mischief; and their
sleep is taken
away, unless they cause
{\ADD{some}} to
fall.
\VS{17}For they
eat the
bread of
wickedness, and
drink the
wine of
violence.
\VS{18}But the
path of the
just
{\ADD{is}} as the
shining
light, that
shineth
more and more unto the
perfect
day.
\VS{19}The
way of the
wicked
{\ADD{is}} as
darkness: they
know not at what they
stumble.
\par }{\PP \VS{20}¶ My
son,
attend to my
words;
incline thine
ear unto my
sayings.
\VS{21}Let them not
depart from thine
eyes;
keep them in the
midst of thine
heart.
\VS{22}For they
{\ADD{are}}
life unto those that
find them, and
health to all their
flesh.\FTNT{†}{{\FR 4.22 }
health: Heb. medicine}
\par }{\PP \VS{23}¶
Keep thy
heart with all
diligence; for out of it
{\ADD{are}} the
issues of
life.\FTNT{‡}{{\FR 4.23 }
with…: Heb. above all keeping}
\VS{24}Put
away from thee a
froward
mouth, and
perverse
lips put
far from thee.\FTNT{§}{{\FR 4.24 }
a froward…: Heb. frowardness of mouth and perverseness of lips}
\VS{25}Let thine
eyes
look right
on, and let thine
eyelids look
straight before thee.
\VS{26}Ponder the
path of thy
feet, and let all thy
ways be
established.\FTNT{**}{{\FR 4.26 }
let…: or, all thy ways shall be ordered aright}
\VS{27}Turn not to the right
hand nor to the
left:
remove thy
foot from
evil.

\par }\Chap{5}{\PP \VerseOne{1}My
son,
attend unto my
wisdom,
{\ADD{and}}
bow thine
ear to my
understanding:
\VS{2}That thou mayest
regard
discretion, and
{\ADD{that}} thy
lips may
keep
knowledge.
\par }{\PP \VS{3}¶ For the
lips of a strange
woman
drop
{\ADD{as}} an
honeycomb, and her
mouth
{\ADD{is}}
smoother than
oil:\FTNT{*}{{\FR 5.3 }
mouth: Heb. palate}
\VS{4}But her
end is
bitter as
wormwood,
sharp as a
twoedged
sword.
\VS{5}Her
feet go
down to
death; her
steps take
hold on
hell.
\VS{6}Lest thou shouldest
ponder the
path of
life, her
ways are
moveable,
{\ADD{that}} thou canst not
know
{\ADD{them}}.
\VS{7}Hear me now therefore, O ye
children, and
depart not from the
words of my
mouth.
\VS{8}Remove thy
way
far from her, and come not
nigh the
door of her
house:
\VS{9}Lest thou
give thine
honour unto
others, and thy
years unto the
cruel:
\VS{10}Lest
strangers be
filled with thy
wealth; and thy
labours
{\ADD{be}} in the
house of a
stranger;\FTNT{†}{{\FR 5.10 }
thy wealth: Heb. thy strength}
\VS{11}And thou
mourn at the
last, when thy
flesh and thy
body are
consumed,
\VS{12}And
say, How have I
hated
instruction, and my
heart
despised
reproof;
\VS{13}And have not
obeyed the
voice of my
teachers, nor
inclined mine
ear to them that
instructed me!
\VS{14}I was
almost in all
evil in the
midst of the
congregation and
assembly.
\par }{\PP \VS{15}¶
Drink
waters out of thine own
cistern, and running
waters out
of thine own
well.
\VS{16}Let thy
fountains be
dispersed
abroad,
{\ADD{and}}
rivers of
waters in the
streets.
\VS{17}Let them be only thine own, and not
strangers’ with thee.
\VS{18}Let thy
fountain be
blessed: and
rejoice with the
wife of thy
youth.
\VS{19}{\ADD{Let her be as}} the
loving
hind and
pleasant
roe; let her
breasts
satisfy thee at all
times; and be thou
ravished
always with her
love.\FTNT{‡}{{\FR 5.19 }
satisfy…: Heb. water thee}\FTNT{§}{{\FR 5.19 }
be thou…: Heb. err thou always in her love}
\VS{20}And why wilt thou, my
son, be
ravished with a strange
woman, and
embrace the
bosom of a
stranger?
\VS{21}For the
ways of
man
{\ADD{are}}
before the
eyes of the
{\ND{
LORD}}, and he
pondereth all his
goings.
\par }{\PP \VS{22}¶ His own
iniquities shall
take the
wicked himself, and he shall be
holden with the
cords of his
sins.\FTNT{**}{{\FR 5.22 }
sins: Heb. sin}
\VS{23}He shall
die without
instruction; and in the
greatness of his
folly he shall go
astray.

\par }\Chap{6}{\PP \VerseOne{1}My
son, if thou be
surety for thy
friend,
{\ADD{if}} thou hast
stricken thy
hand with a
stranger,
\VS{2}Thou art
snared with the
words of thy
mouth, thou art
taken with the
words of thy
mouth.
\VS{3}Do this
now, my
son, and
deliver thyself, when thou art
come into the
hand of thy
friend;
go,
humble thyself, and make
sure thy
friend.\FTNT{*}{{\FR 6.3 }
and make…: or, so shalt thou prevail with thy friend}
\VS{4}Give not
sleep to thine
eyes, nor
slumber to thine
eyelids.
\VS{5}Deliver thyself as a
roe from the
hand
{\ADD{of the hunter}}, and as a
bird from the
hand of the
fowler.
\par }{\PP \VS{6}¶
Go to the
ant, thou
sluggard;
consider her
ways, and be
wise:
\VS{7}Which having no
guide,
overseer, or
ruler,
\VS{8}Provideth her
meat in the
summer,
{\ADD{and}}
gathereth her
food in the
harvest.
\VS{9}How long wilt thou
sleep, O
sluggard? when wilt thou
arise out of thy
sleep?
\VS{10}{\ADD{Yet}} a
little
sleep, a
little
slumber, a
little
folding of the
hands to
sleep:
\VS{11}So shall thy
poverty
come as one that
travelleth, and thy
want as an
armed
man.
\par }{\PP \VS{12}¶ A
naughty
person, a
wicked
man,
walketh with a
froward
mouth.
\VS{13}He
winketh with his
eyes, he
speaketh with his
feet, he
teacheth with his
fingers;
\VS{14}Frowardness
{\ADD{is}} in his
heart, he
deviseth
mischief
continually; he
soweth
discord.\FTNT{†}{{\FR 6.14 }
soweth: Heb. casteth forth}
\VS{15}Therefore shall his
calamity
come
suddenly;
suddenly shall he be
broken without
remedy.
\par }{\PP \VS{16}¶ These
six
{\ADD{things}} doth the
{\ND{
LORD}}
hate: yea,
seven
{\ADD{are}} an
abomination unto
him:\FTNT{‡}{{\FR 6.16 }
unto…: Heb. of his soul}
\VS{17}A
proud
look, a
lying
tongue, and
hands that
shed
innocent
blood,\FTNT{§}{{\FR 6.17 }
A proud…: Heb. Haughty eyes}
\VS{18}An
heart that
deviseth
wicked
imaginations,
feet that be
swift in
running to
mischief,
\VS{19}A
false
witness
{\ADD{that}}
speaketh
lies, and he that
soweth
discord among
brethren.
\par }{\PP \VS{20}¶ My
son,
keep thy
father’s
commandment, and
forsake not the
law of thy
mother:
\VS{21}Bind them
continually upon thine
heart,
{\ADD{and}}
tie them about thy
neck.
\VS{22}When thou
goest, it shall
lead thee; when thou
sleepest, it shall
keep thee; and
{\ADD{when}} thou
awakest, it shall
talk with thee.
\VS{23}For the
commandment
{\ADD{is}} a
lamp; and the
law
{\ADD{is}}
light; and
reproofs of
instruction
{\ADD{are}} the
way of
life:\FTNT{**}{{\FR 6.23 }
lamp: or, candle}
\VS{24}To
keep thee from the
evil
woman, from the
flattery of the
tongue of a strange
woman.\FTNT{††}{{\FR 6.24 }
of the…: or, of the strange tongue}
\VS{25}Lust not after her
beauty in thine
heart; neither let her
take thee with her
eyelids.
\VS{26}For by means
of a
whorish
woman
{\ADD{a man is brought}} to a
piece of
bread: and the
adulteress will
hunt for the
precious
life.\FTNT{‡‡}{{\FR 6.26 }
the adulteress: Heb. the woman of a man, or, a man’s wife}
\VS{27}Can a
man
take
fire in his
bosom, and his
clothes not be
burned?
\VS{28}Can
one
go upon hot
coals, and his
feet not be
burned?
\VS{29}So he that goeth
in to his
neighbour’s
wife; whosoever
toucheth her shall not be
innocent.
\VS{30}{\ADD{Men}} do not
despise a
thief, if he
steal to
satisfy his
soul when he is
hungry;
\VS{31}But
{\ADD{if}} he be
found, he shall
restore
sevenfold; he shall
give all the
substance of his
house.
\VS{32}{\ADD{But}} whoso committeth
adultery with a
woman
lacketh
understanding: he
{\ADD{that}}
doeth it
destroyeth his own
soul.\FTNT{§§}{{\FR 6.32 }
understanding: Heb. heart}
\VS{33}A
wound and
dishonour shall he
get; and his
reproach shall not be wiped
away.
\VS{34}For
jealousy
{\ADD{is}} the
rage of a
man: therefore he will not
spare in the
day of
vengeance.
\VS{35}He will not
regard any
ransom; neither will he rest
content, though thou givest
many
gifts.\FTNT{***}{{\FR 6.35 }
He will…: Heb. He will not accept the face of any ransom}

\par }\Chap{7}{\PP \VerseOne{1}My
son,
keep my
words, and lay
up my
commandments with thee.
\VS{2}Keep my
commandments, and
live; and my
law as the
apple of thine
eye.
\VS{3}Bind them upon thy
fingers,
write them upon the
table of thine
heart.
\VS{4}Say unto
wisdom, Thou
{\ADD{art}} my
sister; and
call
understanding
{\ADD{thy}}
kinswoman:
\VS{5}That they may
keep thee from the
strange
woman, from the
stranger
{\ADD{which}}
flattereth with her
words.
\par }{\PP \VS{6}¶ For at the
window of my
house I
looked through my
casement,
\VS{7}And
beheld among the simple
ones, I
discerned among the
youths, a young
man
void of
understanding,\FTNT{*}{{\FR 7.7 }
the youths: Heb. the sons}
\VS{8}Passing
through the
street
near her
corner; and he
went the
way to her
house,
\VS{9}In the
twilight, in the
evening, in the
black and
dark
night:\FTNT{†}{{\FR 7.9 }
in the evening: Heb. in the evening of days}
\VS{10}And, behold, there
met him a
woman
{\ADD{with}} the
attire of an
harlot, and
subtil of
heart.
\VS{11}(She
{\ADD{is}}
loud and
stubborn; her
feet
abide not in her
house:
\VS{12}Now
{\ADD{is she}}
without,
now in the
streets, and lieth in
wait
at every
corner.)
\VS{13}So she
caught him, and
kissed him,
{\ADD{and}} with an
impudent
face
said unto him,\FTNT{‡}{{\FR 7.13 }
with…: Heb. she strengthened her face and said}
\VS{14}{\ADD{I have}}
peace
offerings with me; this
day have I
payed my
vows.\FTNT{§}{{\FR 7.14 }
I have…: Heb. Peace offerings are upon me}
\VS{15}Therefore came I
forth to
meet thee, diligently to
seek thy
face, and I have
found thee.
\VS{16}I have
decked my
bed with
coverings of tapestry, with
carved
{\ADD{works}}, with fine
linen of
Egypt.
\VS{17}I have
perfumed my
bed with
myrrh,
aloes, and
cinnamon.
\VS{18}Come, let us take our
fill of
love until the
morning: let us
solace ourselves with
loves.
\VS{19}For the
goodman
{\ADD{is}} not at
home, he is
gone a
long
journey:
\VS{20}He hath
taken a
bag of
money with
him,
{\ADD{and}} will
come
home at the
day
appointed.\FTNT{**}{{\FR 7.20 }
with…: Heb. in his hand}\FTNT{††}{{\FR 7.20 }
the day…: or, the new moon}
\VS{21}With her
much fair
speech she caused him to
yield, with the
flattering of her
lips she
forced him.
\VS{22}He
goeth
after her
straightway, as an
ox
goeth to the
slaughter, or as a
fool to the
correction of the
stocks;\FTNT{‡‡}{{\FR 7.22 }
straightway: Heb. suddenly}
\VS{23}Till a
dart
strike through his
liver; as a
bird
hasteth to the
snare, and
knoweth not that it
{\ADD{is}} for his
life.
\par }{\PP \VS{24}¶
Hearken unto me now therefore, O ye
children, and
attend to the
words of my
mouth.
\VS{25}Let not thine
heart
decline to her
ways, go not
astray in her
paths.
\VS{26}For she hath cast
down
many
wounded: yea, many
strong
{\ADD{men}} have been
slain by her.
\VS{27}Her
house
{\ADD{is}} the
way to
hell, going
down to the
chambers of
death.

\par }\Chap{8}{\PP \VerseOne{1}Doth not
wisdom
cry? and
understanding put
forth her
voice?
\VS{2}She
standeth in the
top of high
places, by the
way in the
places of the
paths.
\VS{3}She
crieth
at the
gates, at the
entry of the
city, at the
coming in at the
doors.
\VS{4}Unto you, O
men, I
call; and my
voice
{\ADD{is}} to the
sons of
man.
\VS{5}O ye
simple,
understand
wisdom: and, ye
fools, be ye of an
understanding
heart.
\VS{6}Hear; for I will
speak of excellent
things; and the
opening of my
lips
{\ADD{shall be}} right
things.
\VS{7}For my
mouth shall
speak
truth; and
wickedness
{\ADD{is}} an
abomination to my
lips.\FTNT{*}{{\FR 8.7 }
an…: Heb. the abomination of my lips}
\VS{8}All the
words of my
mouth
{\ADD{are}} in
righteousness;
{\ADD{there is}} nothing
froward or
perverse in them.\FTNT{†}{{\FR 8.8 }
froward: Heb. wreathed}
\VS{9}They
{\ADD{are}} all
plain to him that
understandeth, and
right to them that
find
knowledge.
\VS{10}Receive my
instruction, and not
silver; and
knowledge rather than
choice
gold.
\VS{11}For
wisdom
{\ADD{is}}
better than
rubies; and all the things that may be
desired are not to be
compared to it.
\par }{\PP \VS{12}I
wisdom
dwell with
prudence, and find
out
knowledge of witty
inventions.\FTNT{‡}{{\FR 8.12 }
prudence: or, subtilty}
\VS{13}The
fear of the
{\ND{
LORD}}
{\ADD{is}} to
hate
evil:
pride, and
arrogancy, and the
evil
way, and the
froward
mouth, do I
hate.
\VS{14}Counsel
{\ADD{is}} mine, and sound
wisdom: I
{\ADD{am}}
understanding; I have
strength.
\VS{15}By me
kings
reign, and
princes
decree
justice.
\VS{16}By me
princes
rule, and
nobles,
{\ADD{even}} all the
judges of the
earth.
\VS{17}I
love them that
love me; and those that seek me
early shall
find me.
\VS{18}Riches and
honour
{\ADD{are}} with me;
{\ADD{yea}},
durable
riches and
righteousness.
\VS{19}My
fruit
{\ADD{is}}
better than
gold, yea, than fine
gold; and my
revenue than
choice
silver.
\VS{20}I
lead in the
way of
righteousness, in the
midst of the
paths of
judgment:\FTNT{§}{{\FR 8.20 }
lead: or, walk}
\VS{21}That I may cause those that
love me to
inherit
substance; and I will
fill their
treasures.
\par }{\PP \VS{22}The
{\ND{
LORD}}
possessed me in the
beginning of his
way,
before his
works of old.
\VS{23}I was set
up from
everlasting, from the
beginning, or
ever the
earth was.
\VS{24}When
{\ADD{there were}} no
depths, I was brought
forth; when
{\ADD{there were}} no
fountains
abounding with
water.
\VS{25}Before the
mountains were
settled,
before the
hills was I brought
forth:
\VS{26}While as yet he had not
made the
earth, nor the
fields, nor the highest
part of the
dust of the
world.\FTNT{**}{{\FR 8.26 }
fields: or, open places}\FTNT{††}{{\FR 8.26 }
the highest…: or, the chief part}
\VS{27}When he
prepared the
heavens, I
{\ADD{was}} there: when he
set a
compass upon the
face of the
depth:\FTNT{‡‡}{{\FR 8.27 }
a compass: or, a circle}
\VS{28}When he
established the
clouds
above: when he
strengthened the
fountains of the
deep:
\VS{29}When he
gave to the
sea his
decree, that the
waters should not
pass his
commandment: when he
appointed the
foundations of the
earth:
\VS{30}Then I was by
him,
{\ADD{as}} one brought
up
{\ADD{with him:}} and I was
daily
{\ADD{his}}
delight,
rejoicing
always
before him;
\VS{31}Rejoicing in the habitable
part of his
earth; and my
delights
{\ADD{were}} with the
sons of
men.
\par }{\PP \VS{32}Now therefore
hearken unto me, O ye
children: for
blessed
{\ADD{are they that}}
keep my
ways.
\VS{33}Hear
instruction, and be
wise, and
refuse it not.
\VS{34}Blessed
{\ADD{is}} the
man that
heareth me,
watching
daily at my
gates,
waiting at the
posts of my
doors.
\VS{35}For whoso
findeth me
findeth
life, and shall
obtain
favour of the
{\ND{
LORD}}.\FTNT{§§}{{\FR 8.35 }
obtain: Heb. bring forth}
\VS{36}But he that
sinneth against me
wrongeth his own
soul: all they that
hate me
love
death.

\par }\Chap{9}{\PP \VerseOne{1}Wisdom hath
builded her
house, she hath hewn
out her
seven
pillars:
\VS{2}She hath
killed her
beasts; she hath
mingled her
wine; she hath also
furnished her
table.\FTNT{*}{{\FR 9.2 }
her beasts: Heb. her killing}
\VS{3}She hath sent
forth her
maidens: she
crieth
upon the highest
places of the
city,
\VS{4}Whoso
{\ADD{is}}
simple, let him turn
in hither:
{\ADD{as for}} him that
wanteth
understanding, she
saith to him,
\VS{5}Come,
eat of my
bread, and
drink of the
wine
{\ADD{which}} I have
mingled.
\VS{6}Forsake the
foolish, and
live; and
go in the
way of
understanding.
\VS{7}He that
reproveth a
scorner
getteth to himself
shame: and he that
rebuketh a
wicked
{\ADD{man getteth}} himself a
blot.
\VS{8}Reprove not a
scorner, lest he
hate thee:
rebuke a wise
man, and he will
love thee.
\VS{9}Give
{\ADD{instruction}} to a
wise
{\ADD{man}}, and he will be yet
wiser:
teach a
just
{\ADD{man}}, and he will
increase in
learning.
\VS{10}The
fear of the
{\ND{
LORD}}
{\ADD{is}} the
beginning of
wisdom: and the
knowledge of the
holy
{\ADD{is}}
understanding.
\VS{11}For by me thy
days shall be
multiplied, and the
years of thy
life shall be
increased.
\VS{12}If thou be
wise, thou shalt be
wise for thyself: but
{\ADD{if}} thou
scornest, thou alone shalt
bear
{\ADD{it}}.
\par }{\PP \VS{13}¶ A
foolish
woman
{\ADD{is}}
clamorous:
{\ADD{she is}}
simple,
and
knoweth
nothing.
\VS{14}For she
sitteth at the
door of her
house, on a
seat in the high
places of the
city,
\VS{15}To
call
passengers who go
right on their
ways:
\VS{16}Whoso
{\ADD{is}}
simple, let him turn
in hither: and
{\ADD{as for}} him that
wanteth
understanding, she
saith to him,
\VS{17}Stolen
waters are
sweet, and
bread
{\ADD{eaten}} in
secret is
pleasant.\FTNT{†}{{\FR 9.17 }
eaten…: Heb. of secrecies}
\VS{18}But he
knoweth not that the
dead
{\ADD{are}} there;
{\ADD{and that}} her
guests
{\ADD{are}} in the
depths of
hell.

\par }\Chap{10}{\PP \VerseOne{1}The
proverbs of
Solomon. A
wise
son maketh a
glad
father: but a
foolish
son
{\ADD{is}} the
heaviness of his
mother.
\par }{\PP \VS{2}Treasures of
wickedness
profit nothing: but
righteousness
delivereth from
death.
\VS{3}The
{\ND{
LORD}} will not
suffer the
soul of the
righteous to
famish: but he casteth
away the
substance of the
wicked.\FTNT{*}{{\FR 10.3 }
the substance…: or, the wicked for their wickedness}
\VS{4}He becometh
poor that
dealeth
{\ADD{with}} a
slack
hand: but the
hand of the
diligent maketh
rich.
\VS{5}He that
gathereth in
summer
{\ADD{is}} a
wise
son:
{\ADD{but}} he that
sleepeth in
harvest
{\ADD{is}} a
son that causeth
shame.
\VS{6}Blessings
{\ADD{are}} upon the
head of the
just: but
violence
covereth the
mouth of the
wicked.
\VS{7}The
memory of the
just
{\ADD{is}}
blessed: but the
name of the
wicked shall
rot.
\VS{8}The
wise in
heart will
receive
commandments: but a
prating
fool shall
fall.\FTNT{†}{{\FR 10.8 }
a prating…: Heb. a fool of lips}\FTNT{‡}{{\FR 10.8 }
fall: Heb. be beaten}
\VS{9}He that
walketh
uprightly
walketh
surely: but he that
perverteth his
ways shall be
known.
\VS{10}He that
winketh with the
eye
causeth
sorrow: but a
prating
fool shall
fall.\FTNT{§}{{\FR 10.10 }
fall: Heb. be beaten}
\VS{11}The
mouth of a
righteous
{\ADD{man is}} a
well of
life: but
violence
covereth the
mouth of the
wicked.
\VS{12}Hatred stirreth
up
strifes: but
love
covereth all
sins.
\VS{13}In the
lips of him that hath
understanding
wisdom is
found: but a
rod
{\ADD{is}} for the
back of him that is
void of
understanding.\FTNT{**}{{\FR 10.13 }
of understanding: Heb. of heart}
\VS{14}Wise
{\ADD{men}} lay
up
knowledge: but the
mouth of the
foolish
{\ADD{is}}
near
destruction.
\VS{15}The rich
man’s
wealth
{\ADD{is}} his
strong
city: the
destruction of the
poor
{\ADD{is}} their
poverty.
\VS{16}The
labour of the
righteous
{\ADD{tendeth}} to
life: the
fruit of the
wicked to
sin.
\VS{17}He
{\ADD{is in}} the
way of
life that
keepeth
instruction: but he that
refuseth
reproof
erreth.\FTNT{††}{{\FR 10.17 }
erreth: or, causeth to err}
\VS{18}He that
hideth
hatred
{\ADD{with}}
lying
lips, and he that
uttereth a
slander,
{\ADD{is}} a
fool.
\VS{19}In the
multitude of
words there
wanteth not
sin: but he that
refraineth his
lips
{\ADD{is}}
wise.
\par }{\PP \VS{20}The
tongue of the
just
{\ADD{is as}}
choice
silver: the
heart of the
wicked
{\ADD{is}} little
worth.
\VS{21}The
lips of the
righteous
feed
many: but
fools
die for
want of
wisdom.\FTNT{‡‡}{{\FR 10.21 }
of wisdom: Heb. of heart}
\VS{22}The
blessing of the
{\ND{
LORD}}, it maketh
rich, and he
addeth no
sorrow with it.
\VS{23}{\ADD{It is}} as
sport to a
fool to
do
mischief: but a
man of
understanding hath
wisdom.
\par }{\PP \VS{24}The
fear of the
wicked, it shall
come upon him: but the
desire of the
righteous shall be
granted.
\VS{25}As the
whirlwind
passeth, so
{\ADD{is}} the
wicked no
{\ADD{more:}} but the
righteous
{\ADD{is}} an
everlasting
foundation.
\VS{26}As
vinegar to the
teeth, and as
smoke to the
eyes, so
{\ADD{is}} the
sluggard to them that
send him.
\par }{\PP \VS{27}The
fear of the
{\ND{
LORD}}
prolongeth
days: but the
years of the
wicked shall be
shortened.\FTNT{§§}{{\FR 10.27 }
prolongeth: Heb. addeth}
\VS{28}The
hope of the
righteous
{\ADD{shall be}}
gladness: but the
expectation of the
wicked shall
perish.
\par }{\PP \VS{29}The
way of the
{\ND{
LORD}}
{\ADD{is}}
strength to the
upright: but
destruction
{\ADD{shall be}} to the
workers of
iniquity.
\VS{30}The
righteous shall
never be
removed: but the
wicked shall not
inhabit the
earth.
\par }{\PP \VS{31}The
mouth of the
just bringeth
forth
wisdom: but the
froward
tongue shall be cut
out.
\VS{32}The
lips of the
righteous
know what is
acceptable: but the
mouth of the
wicked
{\ADD{speaketh}}
frowardness.\FTNT{***}{{\FR 10.32 }
frowardness: Heb. frowardnesses}

\par }\Chap{11}{\PP \VerseOne{1}A
false
balance
{\ADD{is}}
abomination to the
{\ND{
LORD}}: but a
just
weight
{\ADD{is}} his
delight.\FTNT{*}{{\FR 11.1 }
A false…: Heb. Balances of deceit}\FTNT{†}{{\FR 11.1 }
just…: Heb. perfect stone}
\VS{2}{\ADD{When}}
pride
cometh, then
cometh
shame: but with the
lowly
{\ADD{is}}
wisdom.
\VS{3}The
integrity of the
upright shall
guide them: but the
perverseness of
transgressors shall
destroy them.
\VS{4}Riches
profit not in the
day of
wrath: but
righteousness
delivereth from
death.
\par }{\PP \VS{5}The
righteousness of the
perfect shall
direct his
way: but the
wicked shall
fall by his own
wickedness.\FTNT{‡}{{\FR 11.5 }
direct: Heb. rectify}
\VS{6}The
righteousness of the
upright shall
deliver them: but
transgressors shall be
taken in
{\ADD{their own}}
naughtiness.
\VS{7}When a
wicked
man
dieth,
{\ADD{his}}
expectation shall
perish: and the
hope of
unjust
{\ADD{men}}
perisheth.
\VS{8}The
righteous is delivered
out of
trouble, and the
wicked
cometh in his stead.
\VS{9}An
hypocrite with
{\ADD{his}}
mouth
destroyeth his
neighbour: but through
knowledge shall the
just be
delivered.
\par }{\PP \VS{10}When it goeth
well with the
righteous, the
city
rejoiceth: and when the
wicked
perish,
{\ADD{there is}}
shouting.
\VS{11}By the
blessing of the
upright the
city is
exalted: but it is
overthrown by the
mouth of the
wicked.
\par }{\PP \VS{12}He that is
void of
wisdom
despiseth his
neighbour: but a
man of
understanding holdeth his
peace.\FTNT{§}{{\FR 11.12 }
void…: Heb. destitute of heart}
\VS{13}A
talebearer
revealeth
secrets: but he that is of a
faithful
spirit
concealeth the
matter.\FTNT{**}{{\FR 11.13 }
A talebearer: Heb. He that walketh, being a talebearer}
\VS{14}Where no
counsel
{\ADD{is}}, the
people
fall: but in the
multitude of
counsellors
{\ADD{there is}}
safety.
\VS{15}He that is
surety for a
stranger shall
smart
{\ADD{for it:}} and he that
hateth
suretiship is
sure.\FTNT{††}{{\FR 11.15 }
smart: Heb. be sore broken}\FTNT{‡‡}{{\FR 11.15 }
suretiship: Heb. those that strike hands}
\VS{16}A
gracious
woman
retaineth
honour: and
strong
{\ADD{men}}
retain
riches.
\VS{17}The
merciful
man doeth
good to his own
soul: but
{\ADD{he that is}}
cruel
troubleth his own
flesh.
\VS{18}The
wicked
worketh a
deceitful
work: but to him that
soweth
righteousness
{\ADD{shall be}} a
sure
reward.
\VS{19}As
righteousness
{\ADD{tendeth}} to
life: so he that
pursueth
evil
{\ADD{pursueth it}} to his own
death.
\VS{20}They that are of a
froward
heart
{\ADD{are}}
abomination to the
{\ND{
LORD}}: but
{\ADD{such as are}}
upright in
{\ADD{their}}
way
{\ADD{are}} his
delight.
\VS{21}{\ADD{Though}}
hand
{\ADD{join}} in
hand, the
wicked shall not be
unpunished: but the
seed of the
righteous shall be
delivered.
\VS{22}{\ADD{As}} a
jewel of
gold in a
swine’s
snout,
{\ADD{so is}} a
fair
woman which is
without
discretion.\FTNT{§§}{{\FR 11.22 }
is without: Heb. departeth from}
\VS{23}The
desire of the
righteous
{\ADD{is}} only
good:
{\ADD{but}} the
expectation of the
wicked
{\ADD{is}}
wrath.
\VS{24}There
is that
scattereth, and yet
increaseth; and
{\ADD{there is}} that
withholdeth more than is
meet, but
{\ADD{it tendeth}} to
poverty.
\VS{25}The
liberal
soul shall be made
fat: and he that
watereth shall be
watered also himself.\FTNT{***}{{\FR 11.25 }
liberal…: Heb. soul of blessing}
\VS{26}He that
withholdeth
corn, the
people shall
curse him: but
blessing
{\ADD{shall be}} upon the
head of him that
selleth
{\ADD{it}}.
\VS{27}He that diligently
seeketh
good
procureth
favour: but he that
seeketh
mischief, it shall
come unto him.
\VS{28}He that
trusteth in his
riches shall
fall: but the
righteous shall
flourish as a
branch.
\VS{29}He that
troubleth his own
house shall
inherit the
wind: and the
fool
{\ADD{shall be}}
servant to the
wise of
heart.
\VS{30}The
fruit of the
righteous
{\ADD{is}} a
tree of
life; and he that
winneth
souls
{\ADD{is}}
wise.\FTNT{†††}{{\FR 11.30 }
winneth: Heb. taketh}
\VS{31}Behold, the
righteous shall be
recompensed in the
earth: much more the
wicked and the
sinner.

\par }\Chap{12}{\PP \VerseOne{1}Whoso
loveth
instruction
loveth
knowledge: but he that
hateth
reproof
{\ADD{is}}
brutish.
\VS{2}A
good
{\ADD{man}}
obtaineth
favour of the
{\ND{
LORD}}: but a
man of wicked
devices will he
condemn.
\VS{3}A
man shall not be
established by
wickedness: but the
root of the
righteous shall not be
moved.
\VS{4}A
virtuous
woman
{\ADD{is}} a
crown to her
husband: but she that maketh
ashamed
{\ADD{is}} as
rottenness in his
bones.
\VS{5}The
thoughts of the
righteous
{\ADD{are}}
right:
{\ADD{but}} the
counsels of the
wicked
{\ADD{are}}
deceit.
\VS{6}The
words of the
wicked
{\ADD{are}} to lie in
wait for
blood: but the
mouth of the
upright shall
deliver them.
\VS{7}The
wicked are
overthrown, and
{\ADD{are}} not: but the
house of the
righteous shall
stand.
\VS{8}A
man shall be
commended
according to his
wisdom: but he that is of a
perverse
heart shall be
despised.\FTNT{*}{{\FR 12.8 }
of a…: Heb. perverse of heart}
\VS{9}{\ADD{He that is}}
despised, and hath a
servant,
{\ADD{is}}
better than he that
honoureth himself, and
lacketh
bread.
\VS{10}A
righteous
{\ADD{man}}
regardeth the
life of his
beast: but the tender
mercies of the
wicked
{\ADD{are}}
cruel.\FTNT{†}{{\FR 12.10 }
tender…: or, bowels}
\VS{11}He that
tilleth his
land shall be
satisfied with
bread: but he that
followeth
vain
{\ADD{persons is}}
void of
understanding.
\VS{12}The
wicked
desireth the
net of
evil
{\ADD{men:}} but the
root of the
righteous
yieldeth
{\ADD{fruit}}.\FTNT{‡}{{\FR 12.12 }
the net: or, the fortress}
\VS{13}The
wicked is
snared by the
transgression of
{\ADD{his}}
lips: but the
just shall come
out of
trouble.\FTNT{§}{{\FR 12.13 }
The wicked…: Heb. The snare of the wicked is in the transgression of lips}
\VS{14}A
man shall be
satisfied with
good by the
fruit of
{\ADD{his}}
mouth: and the
recompence of a
man’s
hands shall be
rendered unto him.
\VS{15}The
way of a
fool
{\ADD{is}}
right in his own
eyes: but he that
hearkeneth unto
counsel
{\ADD{is}}
wise.
\VS{16}A
fool’s
wrath is
presently
known: but a
prudent
{\ADD{man}}
covereth
shame.\FTNT{**}{{\FR 12.16 }
presently: Heb. in that day}
\VS{17}{\ADD{He that}}
speaketh
truth sheweth
forth
righteousness: but a
false
witness
deceit.
\VS{18}There
is that
speaketh like the
piercings of a
sword: but the
tongue of the
wise
{\ADD{is}}
health.
\VS{19}The
lip of
truth shall be
established for
ever: but a
lying
tongue
{\ADD{is}} but for a
moment.
\VS{20}Deceit
{\ADD{is}} in the
heart of them that
imagine
evil: but to the
counsellors of
peace
{\ADD{is}}
joy.
\VS{21}There shall no
evil
happen to the
just: but the
wicked shall be
filled with
mischief.
\VS{22}Lying
lips
{\ADD{are}}
abomination to the
{\ND{
LORD}}: but they that
deal
truly
{\ADD{are}} his
delight.
\VS{23}A
prudent
man
concealeth
knowledge: but the
heart of
fools
proclaimeth
foolishness.
\VS{24}The
hand of the
diligent shall bear
rule: but the
slothful shall be under
tribute.\FTNT{††}{{\FR 12.24 }
slothful: or, deceitful}
\VS{25}Heaviness in the
heart of
man maketh it
stoop: but a
good
word maketh it
glad.
\VS{26}The
righteous
{\ADD{is}} more
excellent than his
neighbour: but the
way of the
wicked
seduceth them.\FTNT{‡‡}{{\FR 12.26 }
excellent: or, abundant}
\VS{27}The
slothful
{\ADD{man}}
roasteth not that which he took in
hunting: but the
substance of a
diligent
man
{\ADD{is}}
precious.
\VS{28}In the
way of
righteousness
{\ADD{is}}
life; and
{\ADD{in}} the
pathway
{\ADD{thereof there is}} no
death.

\par }\Chap{13}{\PP \VerseOne{1}A
wise
son
{\ADD{heareth}} his
father’s
instruction: but a
scorner
heareth not
rebuke.
\VS{2}A
man shall
eat
good by the
fruit of
{\ADD{his}}
mouth: but the
soul of the
transgressors
{\ADD{shall eat}}
violence.
\VS{3}He that
keepeth his
mouth
keepeth his
life:
{\ADD{but}} he that openeth
wide his
lips shall have
destruction.
\VS{4}The
soul of the
sluggard
desireth, and
{\ADD{hath}} nothing: but the
soul of the
diligent shall be made
fat.
\VS{5}A
righteous
{\ADD{man}}
hateth
lying: but a
wicked
{\ADD{man}} is
loathsome, and cometh to
shame.
\VS{6}Righteousness
keepeth
{\ADD{him that is}}
upright in the
way: but
wickedness
overthroweth the
sinner.\FTNT{*}{{\FR 13.6 }
the sinner: Heb. sin}
\VS{7}There
is that maketh himself
rich, yet
{\ADD{hath}} nothing:
{\ADD{there is}} that maketh himself
poor,
yet
{\ADD{hath}}
great
riches.
\VS{8}The
ransom of a
man’s
life
{\ADD{are}} his
riches: but the
poor
heareth not
rebuke.
\VS{9}The
light of the
righteous
rejoiceth: but the
lamp of the
wicked shall be put
out.\FTNT{†}{{\FR 13.9 }
lamp: or, candle}
\VS{10}Only by
pride
cometh
contention: but with the well
advised
{\ADD{is}}
wisdom.
\VS{11}Wealth
{\ADD{gotten}} by
vanity shall be
diminished: but he that
gathereth by
labour shall
increase.\FTNT{‡}{{\FR 13.11 }
by labour: Heb. with the hand}
\VS{12}Hope
deferred maketh the
heart
sick: but
{\ADD{when}} the
desire
cometh,
{\ADD{it is}} a
tree of
life.
\VS{13}Whoso
despiseth the
word shall be
destroyed: but he that
feareth the
commandment shall be
rewarded.\FTNT{§}{{\FR 13.13 }
shall be rewarded: or, shall be in peace}
\VS{14}The
law of the
wise
{\ADD{is}} a
fountain of
life, to
depart from the
snares of
death.
\VS{15}Good
understanding
giveth
favour: but the
way of
transgressors
{\ADD{is}}
hard.
\VS{16}Every
prudent
{\ADD{man}}
dealeth with
knowledge: but a
fool layeth
open
{\ADD{his}}
folly.\FTNT{**}{{\FR 13.16 }
layeth…: Heb. spreadeth}
\VS{17}A
wicked
messenger
falleth into
mischief: but a
faithful
ambassador
{\ADD{is}}
health.\FTNT{††}{{\FR 13.17 }
a faithful…: Heb. an ambassador of faithfulness}
\VS{18}Poverty and
shame
{\ADD{shall be to}} him that
refuseth
instruction: but he that
regardeth
reproof shall be
honoured.
\VS{19}The
desire
accomplished is
sweet to the
soul: but
{\ADD{it is}}
abomination to
fools to
depart from
evil.
\VS{20}He that
walketh with
wise
{\ADD{men}} shall be
wise: but a
companion of
fools shall be
destroyed.\FTNT{‡‡}{{\FR 13.20 }
destroyed: Heb. broken}
\VS{21}Evil
pursueth
sinners: but to the
righteous
good shall be
repayed.
\VS{22}A
good
{\ADD{man}} leaveth an
inheritance to his
children’s
children: and the
wealth of the
sinner
{\ADD{is}} laid
up for the
just.
\VS{23}Much
food
{\ADD{is in}} the
tillage of the
poor: but there
is
{\ADD{that is}}
destroyed for
want of
judgment.
\VS{24}He that
spareth his
rod
hateth his
son: but he that
loveth him
chasteneth him
betimes.
\VS{25}The
righteous
eateth to the
satisfying of his
soul: but the
belly of the
wicked shall
want.

\par }\Chap{14}{\PP \VerseOne{1}Every
wise
woman
buildeth her
house: but the
foolish plucketh it
down with her
hands.
\VS{2}He that
walketh in his
uprightness
feareth the
{\ND{
LORD}}: but
{\ADD{he that is}}
perverse in his
ways
despiseth him.
\VS{3}In the
mouth of the
foolish
{\ADD{is}} a
rod of
pride: but the
lips of the
wise shall
preserve them.
\VS{4}Where no
oxen
{\ADD{are}}, the
crib
{\ADD{is}}
clean: but
much
increase
{\ADD{is}} by the
strength of the
ox.
\VS{5}A
faithful
witness will not
lie: but a
false
witness will
utter
lies.
\VS{6}A
scorner
seeketh
wisdom, and
{\ADD{findeth it}} not: but
knowledge
{\ADD{is}}
easy unto him that
understandeth.
\VS{7}Go from the
presence of a
foolish
man, when thou
perceivest not
{\ADD{in him}} the
lips of
knowledge.
\VS{8}The
wisdom of the
prudent
{\ADD{is}} to
understand his
way: but the
folly of
fools
{\ADD{is}}
deceit.
\VS{9}Fools make a
mock at
sin: but among the
righteous
{\ADD{there is}}
favour.
\VS{10}The
heart
knoweth his
own
bitterness; and a
stranger doth not
intermeddle with his
joy.\FTNT{*}{{\FR 14.10 }
his own…: Heb. the bitterness of his soul}
\VS{11}The
house of the
wicked shall be
overthrown: but the
tabernacle of the
upright shall
flourish.
\VS{12}There
is a
way which seemeth
right
unto a
man, but the
end thereof
{\ADD{are}} the
ways of
death.
\VS{13}Even in
laughter the
heart is
sorrowful; and the
end of that
mirth
{\ADD{is}}
heaviness.
\VS{14}The
backslider in
heart shall be
filled with his own
ways: and a
good
man
{\ADD{shall be satisfied}} from himself.
\VS{15}The
simple
believeth every
word: but the
prudent
{\ADD{man}} looketh
well to his
going.
\VS{16}A
wise
{\ADD{man}}
feareth, and
departeth from
evil: but the
fool
rageth, and is
confident.
\VS{17}{\ADD{He that is}}
soon
angry
dealeth
foolishly: and a
man of wicked
devices is
hated.
\VS{18}The
simple
inherit
folly: but the
prudent are
crowned with
knowledge.
\VS{19}The
evil
bow
before the
good; and the
wicked at the
gates of the
righteous.
\VS{20}The
poor is
hated even of his own
neighbour: but the
rich
{\ADD{hath}}
many
friends.\FTNT{†}{{\FR 14.20 }
the rich…: Heb. many are the lovers of the rich}
\VS{21}He that
despiseth his
neighbour
sinneth: but he that hath
mercy on the
poor,
happy
{\ADD{is}} he.
\VS{22}Do they not
err that
devise
evil? but
mercy and
truth
{\ADD{shall be}} to them that
devise
good.
\VS{23}In all
labour there is
profit: but the
talk of the
lips
{\ADD{tendeth}} only to
penury.
\VS{24}The
crown of the
wise
{\ADD{is}} their
riches:
{\ADD{but}} the
foolishness of
fools
{\ADD{is}}
folly.
\VS{25}A
true
witness
delivereth
souls: but a
deceitful
{\ADD{witness}}
speaketh
lies.
\par }{\PP \VS{26}In the
fear of the
{\ND{
LORD}}
{\ADD{is}}
strong
confidence: and his
children shall have a place of
refuge.
\VS{27}The
fear of the
{\ND{
LORD}}
{\ADD{is}} a
fountain of
life, to
depart from the
snares of
death.
\VS{28}In the
multitude of
people
{\ADD{is}} the
king’s
honour: but in the
want of
people
{\ADD{is}} the
destruction of the
prince.
\VS{29}{\ADD{He that is}}
slow to
wrath
{\ADD{is}} of
great
understanding: but
{\ADD{he that is}}
hasty of
spirit
exalteth
folly.\FTNT{‡}{{\FR 14.29 }
hasty…: Heb. short of spirit}
\VS{30}A
sound
heart
{\ADD{is}} the
life of the
flesh: but
envy the
rottenness of the
bones.
\VS{31}He that
oppresseth the
poor
reproacheth his
Maker: but he that
honoureth him hath
mercy on the
poor.
\VS{32}The
wicked is driven
away in his
wickedness: but the
righteous hath
hope in his
death.
\VS{33}Wisdom
resteth in the
heart of him that hath
understanding: but
{\ADD{that which is}} in the
midst of
fools is made
known.
\VS{34}Righteousness
exalteth a
nation: but
sin
{\ADD{is}} a
reproach to any
people.\FTNT{§}{{\FR 14.34 }
to any…: Heb. to nations}
\VS{35}The
king’s
favour
{\ADD{is}} toward a
wise
servant: but his
wrath is
{\ADD{against}} him that causeth
shame.

\par }\Chap{15}{\PP \VerseOne{1}A
soft
answer
turneth away
wrath: but
grievous
words stir
up
anger.
\VS{2}The
tongue of the
wise useth
knowledge
aright: but the
mouth of
fools poureth
out
foolishness.\FTNT{*}{{\FR 15.2 }
poureth: Heb. belcheth, or, bubbleth}
\VS{3}The
eyes of the
{\ND{
LORD}}
{\ADD{are}} in every
place,
beholding the
evil and the
good.
\VS{4}A
wholesome
tongue
{\ADD{is}} a
tree of
life: but
perverseness therein
{\ADD{is}} a
breach in the
spirit.\FTNT{†}{{\FR 15.4 }
A wholesome…: Heb. The healing of the tongue}
\VS{5}A
fool
despiseth his
father’s
instruction: but he that
regardeth
reproof is
prudent.
\VS{6}In the
house of the
righteous
{\ADD{is}}
much
treasure: but in the
revenues of the
wicked is
trouble.
\VS{7}The
lips of the
wise
disperse
knowledge: but the
heart of the
foolish
{\ADD{doeth}} not so.
\VS{8}The
sacrifice of the
wicked
{\ADD{is}} an
abomination to the
{\ND{
LORD}}: but the
prayer of the
upright
{\ADD{is}} his
delight.
\VS{9}The
way of the
wicked
{\ADD{is}} an
abomination unto the
{\ND{
LORD}}: but he
loveth him that
followeth after
righteousness.
\VS{10}Correction
{\ADD{is}}
grievous unto him that
forsaketh the
way:
{\ADD{and}} he that
hateth
reproof shall
die.\FTNT{‡}{{\FR 15.10 }
Correction: or, Instruction}
\VS{11}Hell and
destruction
{\ADD{are}} before the
{\ND{
LORD}}: how much more then the
hearts of the
children of
men?
\VS{12}A
scorner
loveth not one that
reproveth him: neither will he
go unto the
wise.
\VS{13}A
merry
heart maketh a
cheerful
countenance: but by
sorrow of the
heart the
spirit is
broken.
\VS{14}The
heart of him that hath
understanding
seeketh
knowledge: but the
mouth of
fools
feedeth on
foolishness.
\VS{15}All the
days of the
afflicted
{\ADD{are}}
evil: but he that is of a
merry
heart
{\ADD{hath}} a
continual
feast.
\par }{\PP \VS{16}Better
{\ADD{is}}
little with the
fear of the
{\ND{
LORD}} than
great
treasure and
trouble therewith.
\VS{17}Better
{\ADD{is}} a
dinner of
herbs where
love is, than a
stalled
ox and
hatred therewith.
\VS{18}A
wrathful
man stirreth
up
strife: but
{\ADD{he that is}}
slow to
anger
appeaseth
strife.
\VS{19}The
way of the
slothful
{\ADD{man is}} as an
hedge of
thorns: but the
way of the
righteous
{\ADD{is}} made
plain.\FTNT{§}{{\FR 15.19 }
is made…: Heb. is raised up as a causey}
\VS{20}A
wise
son maketh a
glad
father: but a
foolish
man
despiseth his
mother.
\VS{21}Folly
{\ADD{is}}
joy to
{\ADD{him that is}}
destitute of
wisdom: but a
man of
understanding
walketh
uprightly.\FTNT{**}{{\FR 15.21 }
destitute…: Heb. void of heart}
\VS{22}Without
counsel
purposes are
disappointed: but in the
multitude of
counsellors they are
established.
\VS{23}A
man hath
joy by the
answer of his
mouth: and a
word
{\ADD{spoken}} in due
season, how
good
{\ADD{is it}}!\FTNT{††}{{\FR 15.23 }
due…: Heb. his season}
\VS{24}The
way of
life
{\ADD{is}}
above to the
wise, that he may
depart from
hell
beneath.
\VS{25}The
{\ND{
LORD}} will
destroy the
house of the
proud: but he will
establish the
border of the
widow.
\VS{26}The
thoughts of the
wicked
{\ADD{are}} an
abomination to the
{\ND{
LORD}}: but
{\ADD{the words}} of the
pure
{\ADD{are}}
pleasant
words.\FTNT{‡‡}{{\FR 15.26 }
pleasant…: Heb. words of pleasantness}
\VS{27}He that is
greedy of
gain
troubleth his own
house; but he that
hateth
gifts shall
live.
\VS{28}The
heart of the
righteous
studieth to
answer: but the
mouth of the
wicked
poureth out evil
things.
\VS{29}The
{\ND{
LORD}}
{\ADD{is}}
far from the
wicked: but he
heareth the
prayer of the
righteous.
\VS{30}The
light of the
eyes
rejoiceth the
heart:
{\ADD{and}} a
good
report
maketh the
bones
fat.
\VS{31}The
ear that
heareth the
reproof of
life
abideth
among the
wise.
\VS{32}He that
refuseth
instruction
despiseth his own
soul: but he that
heareth
reproof
getteth
understanding.\FTNT{§§}{{\FR 15.32 }
instruction: or, correction}\FTNT{***}{{\FR 15.32 }
heareth: or, obeyeth}\FTNT{†††}{{\FR 15.32 }
getteth…: Heb. possesseth an heart}
\VS{33}The
fear of the
{\ND{
LORD}}
{\ADD{is}} the
instruction of
wisdom; and
before
honour
{\ADD{is}}
humility.

\par }\Chap{16}{\PP \VerseOne{1}The
preparations of the
heart in
man, and the
answer of the
tongue,
{\ADD{is}} from the
{\ND{
LORD}}.\FTNT{*}{{\FR 16.1 }
preparations: or, disposings}
\VS{2}All the
ways of a
man
{\ADD{are}}
clean in his own
eyes; but the
{\ND{
LORD}}
weigheth the
spirits.
\VS{3}Commit thy
works unto the
{\ND{
LORD}}, and thy
thoughts shall be
established.\FTNT{†}{{\FR 16.3 }
Commit: Heb. Roll}
\VS{4}The
{\ND{
LORD}} hath
made all
{\ADD{things}} for
himself: yea, even the
wicked for the
day of
evil.
\VS{5}Every one
{\ADD{that is}}
proud in
heart
{\ADD{is}} an
abomination to the
{\ND{
LORD}}:
{\ADD{though}}
hand
{\ADD{join}} in
hand, he shall not be
unpunished.\FTNT{‡}{{\FR 16.5 }
unpunished: Heb. held innocent}
\VS{6}By
mercy and
truth
iniquity is
purged: and by the
fear of the
{\ND{
LORD}}
{\ADD{men}}
depart from
evil.
\VS{7}When a
man’s
ways
please the
{\ND{
LORD}}, he maketh even his
enemies to be at
peace with him.
\VS{8}Better
{\ADD{is}} a
little with
righteousness than
great
revenues without
right.
\VS{9}A
man’s
heart
deviseth his
way: but the
{\ND{
LORD}}
directeth his
steps.
\VS{10}A divine
sentence
{\ADD{is}} in the
lips of the
king: his
mouth
transgresseth not in
judgment.\FTNT{§}{{\FR 16.10 }
A divine…: Heb. Divination}
\VS{11}A
just
weight and
balance
{\ADD{are}} the
{\ND{
LORD’s}}: all the
weights of the
bag
{\ADD{are}} his
work.\FTNT{**}{{\FR 16.11 }
the weights: Heb. the stones}
\VS{12}{\ADD{It is}} an
abomination to
kings to
commit
wickedness: for the
throne is
established by
righteousness.
\VS{13}Righteous
lips
{\ADD{are}} the
delight of
kings; and they
love him that
speaketh
right.
\par }{\PP \VS{14}The
wrath of a
king
{\ADD{is as}}
messengers of
death: but a
wise
man will
pacify it.
\VS{15}In the
light of the
king’s
countenance
{\ADD{is}}
life; and his
favour
{\ADD{is}} as a
cloud of the latter
rain.
\VS{16}How much
better
{\ADD{is it}} to
get
wisdom than
gold! and to
get
understanding rather to be
chosen than
silver!
\VS{17}The
highway of the
upright
{\ADD{is}} to
depart from
evil: he that
keepeth his
way
preserveth his
soul.
\VS{18}Pride
{\ADD{goeth}}
before
destruction, and an
haughty
spirit
before a
fall.
\VS{19}Better
{\ADD{it is to be}} of an
humble
spirit with the
lowly, than to
divide the
spoil with the
proud.
\VS{20}He that handleth a
matter
wisely shall
find
good: and whoso
trusteth in the
{\ND{
LORD}},
happy
{\ADD{is}} he.\FTNT{††}{{\FR 16.20 }
handleth…: or, understandeth a matter}
\VS{21}The
wise in
heart shall be
called
prudent: and the
sweetness of the
lips
increaseth
learning.
\VS{22}Understanding
{\ADD{is}} a
wellspring of
life unto him that
hath it: but the
instruction of
fools
{\ADD{is}}
folly.
\VS{23}The
heart of the
wise
teacheth his
mouth, and
addeth
learning to his
lips.\FTNT{‡‡}{{\FR 16.23 }
teacheth: Heb. maketh wise}
\VS{24}Pleasant
words
{\ADD{are as}} an
honeycomb,
sweet to the
soul, and
health to the
bones.
\VS{25}There
is a
way that seemeth
right
unto a
man, but the
end thereof
{\ADD{are}} the
ways of
death.
\VS{26}He that
laboureth
laboureth for himself; for his
mouth
craveth it of him.\FTNT{§§}{{\FR 16.26 }
He…: Heb. The soul of him that}\FTNT{***}{{\FR 16.26 }
craveth…: Heb. boweth unto him}
\par }{\PP \VS{27}An
ungodly
man diggeth
up
evil: and in his
lips
{\ADD{there is}} as a
burning
fire.\FTNT{†††}{{\FR 16.27 }
An…: Heb. A man of Belial}
\VS{28}A
froward
man
soweth
strife: and a
whisperer
separateth chief
friends.\FTNT{‡‡‡}{{\FR 16.28 }
soweth: Heb. sendeth forth}
\par }{\PP \VS{29}A
violent
man
enticeth his
neighbour, and
leadeth him into the
way
{\ADD{that is}} not
good.
\VS{30}He
shutteth his
eyes to
devise froward
things:
moving his
lips he
bringeth
evil to
pass.
\VS{31}The hoary
head
{\ADD{is}} a
crown of
glory,
{\ADD{if}} it be
found in the
way of
righteousness.
\VS{32}{\ADD{He that is}}
slow to
anger
{\ADD{is}}
better than the
mighty; and he that
ruleth his
spirit than he that
taketh a
city.
\VS{33}The
lot is
cast into the
lap; but the whole
disposing thereof
{\ADD{is}} of the
{\ND{
LORD}}.

\par }\Chap{17}{\PP \VerseOne{1}Better
{\ADD{is}} a
dry
morsel, and
quietness therewith, than an
house
full of
sacrifices
{\ADD{with}}
strife.\FTNT{*}{{\FR 17.1 }
sacrifices: or, good cheer}
\VS{2}A
wise
servant shall have
rule over a
son that causeth
shame, and shall have
part of the
inheritance
among the
brethren.
\VS{3}The fining
pot
{\ADD{is}} for
silver, and the
furnace for
gold: but the
{\ND{
LORD}}
trieth the
hearts.
\VS{4}A wicked
doer giveth
heed to
false
lips;
{\ADD{and}} a
liar giveth
ear to a
naughty
tongue.
\VS{5}Whoso
mocketh the
poor
reproacheth his
Maker:
{\ADD{and}} he that is
glad at
calamities shall not be
unpunished.\FTNT{†}{{\FR 17.5 }
unpunished: Heb. held innocent}
\VS{6}Children’s
children
{\ADD{are}} the
crown of old
men; and the
glory of
children
{\ADD{are}} their
fathers.
\VS{7}Excellent
speech
becometh not a
fool: much less do
lying
lips a
prince.\FTNT{‡}{{\FR 17.7 }
Excellent…: Heb. A lip of excellency}\FTNT{§}{{\FR 17.7 }
lying…: Heb. a lip of lying}
\VS{8}A
gift
{\ADD{is as}} a
precious
stone in the
eyes of him that
hath it: whithersoever it
turneth, it
prospereth.\FTNT{**}{{\FR 17.8 }
a precious…: Heb. a stone of grace}
\VS{9}He that
covereth a
transgression
seeketh
love; but he that
repeateth a
matter
separateth
{\ADD{very}}
friends.\FTNT{††}{{\FR 17.9 }
seeketh: or, procureth}
\VS{10}A
reproof
entereth more into a wise
man than an
hundred
stripes into a
fool.\FTNT{‡‡}{{\FR 17.10 }
entereth…: or, aweth more a wise man, than to strike a fool an hundred times}
\VS{11}An
evil
{\ADD{man}}
seeketh only
rebellion: therefore a
cruel
messenger shall be
sent against him.
\VS{12}Let a
bear
robbed of her whelps
meet a
man, rather
than a
fool in his
folly.
\VS{13}Whoso
rewardeth
evil for
good,
evil shall not
depart from his
house.
\VS{14}The
beginning of
strife
{\ADD{is as}} when one letteth
out
water: therefore leave
off
contention,
before it be
meddled with.
\VS{15}He that
justifieth the
wicked, and he that
condemneth the
just, even they
both
{\ADD{are}}
abomination to the
{\ND{
LORD}}.
\VS{16}Wherefore
{\ADD{is there}} a
price in the
hand of a
fool to
get
wisdom, seeing
{\ADD{he hath}} no
heart
{\ADD{to it}}?
\VS{17}A
friend
loveth at all
times, and a
brother is
born for
adversity.
\VS{18}A
man
void of
understanding
striketh
hands,
{\ADD{and}}
becometh
surety in the
presence of his
friend.\FTNT{§§}{{\FR 17.18 }
understanding: Heb. heart}
\VS{19}He
loveth
transgression that
loveth
strife:
{\ADD{and}} he that
exalteth his
gate
seeketh
destruction.
\VS{20}He that hath a
froward
heart
findeth no
good: and he that hath a
perverse
tongue
falleth into
mischief.\FTNT{***}{{\FR 17.20 }
He that hath a froward…: Heb. The froward of heart}
\VS{21}He that
begetteth a
fool
{\ADD{doeth it}} to his
sorrow: and the
father of a
fool hath no
joy.
\VS{22}A
merry
heart doeth
good
{\ADD{like}} a
medicine: but a
broken
spirit
drieth the
bones.\FTNT{†††}{{\FR 17.22 }
like: or, to}
\VS{23}A
wicked
{\ADD{man}}
taketh a
gift out of the
bosom to
pervert the
ways of
judgment.
\VS{24}Wisdom
{\ADD{is}} before him that hath
understanding; but the
eyes of a
fool
{\ADD{are}} in the
ends of the
earth.
\VS{25}A
foolish
son
{\ADD{is}} a
grief to his
father, and
bitterness to her that
bare him.
\VS{26}Also to
punish the
just
{\ADD{is}} not
good,
{\ADD{nor}} to
strike
princes for
equity.
\par }{\PP \VS{27}He that
hath
knowledge
spareth his
words:
{\ADD{and}} a
man of
understanding is of an
excellent
spirit.\FTNT{‡‡‡}{{\FR 17.27 }
an…: or, a cool}
\VS{28}Even a
fool, when he holdeth his
peace, is
counted
wise:
{\ADD{and}} he that
shutteth his
lips
{\ADD{is esteemed}} a man of
understanding.

\par }\Chap{18}{\PP \VerseOne{1}Through
desire a man, having
separated himself,
seeketh
{\ADD{and}}
intermeddleth with all
wisdom.\FTNT{*}{{\FR 18.1 }
Through…: or, He that separateth himself seeketh according to his desire, and intermeddleth in every business}
\VS{2}A
fool hath no
delight in
understanding, but that his
heart may
discover itself.
\VS{3}When the
wicked
cometh,
{\ADD{then}}
cometh also
contempt, and with
ignominy
reproach.
\VS{4}The
words of a
man’s
mouth
{\ADD{are as}}
deep
waters,
{\ADD{and}} the
wellspring of
wisdom
{\ADD{as}} a
flowing
brook.
\VS{5}{\ADD{It is}} not
good to
accept the
person of the
wicked, to
overthrow the
righteous in
judgment.
\par }{\PP \VS{6}A
fool’s
lips
enter into
contention, and his
mouth
calleth for
strokes.
\VS{7}A
fool’s
mouth
{\ADD{is}} his
destruction, and his
lips
{\ADD{are}} the
snare of his
soul.
\VS{8}The
words of a
talebearer
{\ADD{are}} as
wounds, and they go
down into the innermost
parts of the
belly.\FTNT{†}{{\FR 18.8 }
talebearer: or, whisperer}\FTNT{‡}{{\FR 18.8 }
as wounds: or, like as when men are wounded}\FTNT{§}{{\FR 18.8 }
innermost…: Heb. chambers}
\VS{9}He also that is
slothful in his
work is
brother to him that is a
great
waster.
\VS{10}The
name of the
{\ND{
LORD}}
{\ADD{is}} a
strong
tower: the
righteous
runneth into it, and is
safe.\FTNT{**}{{\FR 18.10 }
safe: Heb. set aloft}
\VS{11}The rich
man’s
wealth
{\ADD{is}} his
strong
city, and as an
high
wall in his own
conceit.
\VS{12}Before
destruction the
heart of
man is
haughty, and
before
honour
{\ADD{is}}
humility.
\VS{13}He that
answereth a
matter before he
heareth
{\ADD{it}}, it
{\ADD{is}}
folly and
shame unto him.\FTNT{††}{{\FR 18.13 }
answereth…: Heb. returneth a word}
\VS{14}The
spirit of a
man will
sustain his
infirmity; but a
wounded
spirit who can
bear?
\VS{15}The
heart of the
prudent
getteth
knowledge; and the
ear of the
wise
seeketh
knowledge.
\VS{16}A
man’s
gift maketh
room for him, and
bringeth him
before great
men.
\VS{17}{\ADD{He that is}}
first in his own
cause
{\ADD{seemeth}}
just; but his
neighbour
cometh and
searcheth him.
\VS{18}The
lot causeth
contentions to
cease, and
parteth between the
mighty.
\VS{19}A
brother
offended
{\ADD{is harder to be won}} than a
strong
city: and
{\ADD{their}}
contentions
{\ADD{are}} like the
bars of a
castle.
\VS{20}A
man’s
belly shall be
satisfied with the
fruit of his
mouth;
{\ADD{and}} with the
increase of his
lips shall he be
filled.
\VS{21}Death and
life
{\ADD{are}} in the
power of the
tongue: and they that
love it shall
eat the
fruit thereof.
\VS{22}{\ADD{Whoso}}
findeth a
wife
findeth a
good
{\ADD{thing}}, and
obtaineth
favour of the
{\ND{
LORD}}.
\VS{23}The
poor
useth
intreaties; but the
rich
answereth
roughly.
\VS{24}A
man
{\ADD{that hath}}
friends must shew himself
friendly: and there
is a
friend
{\ADD{that}} sticketh
closer than a
brother.

\par }\Chap{19}{\PP \VerseOne{1}Better
{\ADD{is}} the
poor that
walketh in his
integrity, than
{\ADD{he that is}}
perverse in his
lips, and is a
fool.
\VS{2}Also,
{\ADD{that}} the
soul
{\ADD{be}} without
knowledge,
{\ADD{it is}} not
good; and he that
hasteth with
{\ADD{his}}
feet
sinneth.
\VS{3}The
foolishness of
man
perverteth his
way: and his
heart
fretteth against the
{\ND{
LORD}}.
\VS{4}Wealth
maketh
many
friends; but the
poor is
separated from his
neighbour.
\VS{5}A
false
witness shall not be
unpunished, and
{\ADD{he that}}
speaketh
lies shall not
escape.\FTNT{*}{{\FR 19.5 }
unpunished: Heb. held innocent}
\par }{\PP \VS{6}Many will
intreat the
favour of the
prince: and every man
{\ADD{is}} a
friend to
him that giveth
gifts.\FTNT{†}{{\FR 19.6 }
him…: Heb. a man of gifts}
\VS{7}All the
brethren of the
poor do
hate him: how much more do his
friends go
far from him? he
pursueth
{\ADD{them with}}
words,
{\ADD{yet}} they
{\ADD{are}} wanting
{\ADD{to him}}.
\VS{8}He that
getteth
wisdom
loveth his own
soul: he that
keepeth
understanding shall
find
good.\FTNT{‡}{{\FR 19.8 }
wisdom: Heb. an heart}
\VS{9}A
false
witness shall not be
unpunished, and
{\ADD{he that}}
speaketh
lies shall
perish.
\VS{10}Delight is not
seemly for a
fool; much less for a
servant to have
rule over
princes.
\VS{11}The
discretion of a
man
deferreth his
anger; and
{\ADD{it is}} his
glory to pass
over a
transgression.\FTNT{§}{{\FR 19.11 }
discretion: or, prudence}
\VS{12}The
king’s
wrath
{\ADD{is}} as the
roaring of a
lion; but his
favour
{\ADD{is}} as
dew upon the
grass.
\VS{13}A
foolish
son
{\ADD{is}} the
calamity of his
father: and the
contentions of a
wife
{\ADD{are}} a
continual
dropping.
\VS{14}House and
riches
{\ADD{are}} the
inheritance of
fathers: and a
prudent
wife
{\ADD{is}} from the
{\ND{
LORD}}.
\VS{15}Slothfulness
casteth into a deep
sleep; and an
idle
soul shall suffer
hunger.
\VS{16}He that
keepeth the
commandment
keepeth his own
soul;
{\ADD{but}} he that
despiseth his
ways shall
die.
\VS{17}He that hath
pity upon the
poor
lendeth unto the
{\ND{
LORD}}; and that which he hath
given will he pay him
again.\FTNT{**}{{\FR 19.17 }
that which…: or, his deed}
\VS{18}Chasten thy
son while there
is
hope, and let not thy
soul
spare for his
crying.\FTNT{††}{{\FR 19.18 }
for…: or, to his destruction: or, to cause him to die}
\VS{19}A
man of
great
wrath shall
suffer
punishment: for if thou
deliver
{\ADD{him}}, yet thou must do it
again.\FTNT{‡‡}{{\FR 19.19 }
do: Heb. add}
\VS{20}Hear
counsel, and
receive
instruction, that thou mayest be
wise in thy latter
end.
\VS{21}{\ADD{There are}}
many
devices in a
man’s
heart; nevertheless the
counsel of the
{\ND{
LORD}}, that shall
stand.
\VS{22}The
desire of a
man
{\ADD{is}} his
kindness: and a poor
man
{\ADD{is}}
better
than a
liar.
\VS{23}The
fear of the
{\ND{
LORD}}
{\ADD{tendeth}} to
life: and
{\ADD{he that hath it}} shall
abide
satisfied; he shall not be
visited with
evil.
\VS{24}A
slothful
{\ADD{man}}
hideth his
hand in
{\ADD{his}}
bosom, and will not so much as
bring it to his
mouth
again.
\VS{25}Smite a
scorner, and the
simple will
beware: and
reprove one that hath
understanding,
{\ADD{and}} he will
understand
knowledge.\FTNT{§§}{{\FR 19.25 }
will beware: Heb. will be cunning}
\VS{26}He that
wasteth
{\ADD{his}}
father,
{\ADD{and}} chaseth
away
{\ADD{his}}
mother,
{\ADD{is}} a
son that causeth
shame, and bringeth
reproach.
\VS{27}Cease, my
son, to
hear the
instruction
{\ADD{that causeth}} to
err from the
words of
knowledge.
\VS{28}An
ungodly
witness
scorneth
judgment: and the
mouth of the
wicked
devoureth
iniquity.\FTNT{***}{{\FR 19.28 }
An…: Heb. A witness of Belial}
\VS{29}Judgments are
prepared for
scorners, and
stripes for the
back of
fools.

\par }\Chap{20}{\PP \VerseOne{1}Wine
{\ADD{is}} a
mocker, strong
drink
{\ADD{is}}
raging: and whosoever is
deceived thereby is not
wise.
\VS{2}The
fear of a
king
{\ADD{is}} as the
roaring of a
lion:
{\ADD{whoso}} provoketh him to
anger
sinneth
{\ADD{against}} his own
soul.
\VS{3}{\ADD{It is}} an
honour for a
man to
cease from
strife: but every
fool will be
meddling.
\VS{4}The
sluggard will not
plow by reason of the
cold;
{\ADD{therefore}} shall he
beg in
harvest, and
{\ADD{have}} nothing.\FTNT{*}{{\FR 20.4 }
cold: or, winter}
\VS{5}Counsel in the
heart of
man
{\ADD{is like}}
deep
water; but a
man of
understanding will draw it
out.
\VS{6}Most
men will
proclaim every
one his own
goodness: but a
faithful
man who can
find?\FTNT{†}{{\FR 20.6 }
goodness: or, bounty}
\VS{7}The
just
{\ADD{man}}
walketh in his
integrity: his
children
{\ADD{are}}
blessed
after him.
\VS{8}A
king that sitteth
in the
throne of
judgment scattereth
away all
evil with his
eyes.
\VS{9}Who can
say, I have made my
heart
clean, I am
pure from my
sin?
\VS{10}Divers
weights,
{\ADD{and}} divers
measures,
both of them
{\ADD{are}}
alike
abomination to the
{\ND{
LORD}}.\FTNT{‡}{{\FR 20.10 }
Divers weights: Heb. A stone and a stone}\FTNT{§}{{\FR 20.10 }
divers measures: Heb. an ephah and an ephah}
\VS{11}Even a
child is
known by his
doings, whether his
work
{\ADD{be}}
pure, and whether
{\ADD{it be}}
right.
\VS{12}The
hearing
ear, and the
seeing
eye, the
{\ND{
LORD}} hath
made even
both of them.
\VS{13}Love not
sleep, lest thou come to
poverty;
open thine
eyes,
{\ADD{and}} thou shalt be
satisfied with
bread.
\VS{14}{\ADD{It is}}
naught,
{\ADD{it is}}
naught,
saith the
buyer: but when he is
gone his way, then he
boasteth.
\VS{15}There
is
gold, and a
multitude of
rubies: but the
lips of
knowledge
{\ADD{are}} a
precious
jewel.
\VS{16}Take his
garment that is
surety
{\ADD{for}} a
stranger: and take a
pledge of him for a strange
woman.
\VS{17}Bread of
deceit
{\ADD{is}}
sweet to a
man; but
afterwards his
mouth shall be
filled with
gravel.\FTNT{**}{{\FR 20.17 }
deceit: Heb. lying, or, falsehood}
\VS{18}{\ADD{Every}}
purpose is
established by
counsel: and with good
advice
make
war.
\VS{19}He that goeth
about
{\ADD{as}} a
talebearer
revealeth
secrets: therefore
meddle not with him that
flattereth with his
lips.\FTNT{††}{{\FR 20.19 }
flattereth: or, enticeth}
\VS{20}Whoso
curseth his
father or his
mother, his
lamp shall be put
out in
obscure
darkness.\FTNT{‡‡}{{\FR 20.20 }
lamp: or, candle}
\VS{21}An
inheritance
{\ADD{may be}} gotten
hastily at the
beginning; but the
end thereof shall not be
blessed.
\VS{22}Say not thou, I will
recompense
evil;
{\ADD{but}} wait
on the
{\ND{
LORD}}, and he shall
save thee.
\VS{23}Divers
weights
{\ADD{are}} an
abomination unto the
{\ND{
LORD}}; and a
false
balance
{\ADD{is}} not
good.\FTNT{§§}{{\FR 20.23 }
a false…: Heb. balance of deceit}
\VS{24}Man’s
goings
{\ADD{are}} of the
{\ND{
LORD}}; how can a
man then
understand his own
way?
\VS{25}{\ADD{It is}} a
snare to the
man
{\ADD{who}}
devoureth
{\ADD{that which is}}
holy, and
after
vows to make
enquiry.
\VS{26}A
wise
king
scattereth the
wicked, and
bringeth the
wheel over them.
\VS{27}The
spirit of
man
{\ADD{is}} the
candle of the
{\ND{
LORD}},
searching all the inward
parts of the
belly.\FTNT{***}{{\FR 20.27 }
candle: or, lamp}
\VS{28}Mercy and
truth
preserve the
king: and his
throne is
upholden by
mercy.
\VS{29}The
glory of young
men
{\ADD{is}} their
strength: and the
beauty of old
men
{\ADD{is}} the gray
head.
\VS{30}The
blueness of a
wound
cleanseth away
evil: so
{\ADD{do}}
stripes the inward
parts of the
belly.\FTNT{†††}{{\FR 20.30 }
cleanseth…: Heb. is a purging medicine against}

\par }\Chap{21}{\PP \VerseOne{1}The
king’s
heart
{\ADD{is}} in the
hand of the
{\ND{
LORD}},
{\ADD{as}} the
rivers of
water: he
turneth it whithersoever he
will.
\VS{2}Every
way of a
man
{\ADD{is}}
right in his own
eyes: but the
{\ND{
LORD}}
pondereth the
hearts.
\VS{3}To
do
justice and
judgment
{\ADD{is}} more
acceptable to the
{\ND{
LORD}} than
sacrifice.
\VS{4}An
high
look, and a
proud
heart,
{\ADD{and}} the
plowing of the
wicked,
{\ADD{is}}
sin.\FTNT{*}{{\FR 21.4 }
An…: Heb. Haughtiness of eyes}\FTNT{†}{{\FR 21.4 }
the plowing: or, the light}
\VS{5}The
thoughts of the
diligent
{\ADD{tend}} only to
plenteousness; but of every one
{\ADD{that is}}
hasty only to
want.
\VS{6}The
getting of
treasures by a
lying
tongue
{\ADD{is}} a
vanity tossed to and fro
of them that
seek
death.
\VS{7}The
robbery of the
wicked shall
destroy them; because they
refuse to
do
judgment.\FTNT{‡}{{\FR 21.7 }
destroy…: Heb. saw them, or, dwell with them}
\VS{8}The
way of
man
{\ADD{is}}
froward and
strange: but
{\ADD{as for}} the
pure, his
work
{\ADD{is}}
right.
\VS{9}{\ADD{It is}}
better to
dwell in a
corner of the
housetop, than with a
brawling
woman in a
wide
house.\FTNT{§}{{\FR 21.9 }
a brawling…: Heb. a woman of contentions}\FTNT{**}{{\FR 21.9 }
a wide…: Heb. an house of society}
\VS{10}The
soul of the
wicked
desireth
evil: his
neighbour findeth no
favour in his
eyes.\FTNT{††}{{\FR 21.10 }
findeth…: Heb. is not favoured}
\VS{11}When the
scorner is
punished, the
simple is made
wise: and when the
wise is
instructed, he
receiveth
knowledge.
\VS{12}The
righteous
{\ADD{man}} wisely
considereth the
house of the
wicked:
{\ADD{but God}}
overthroweth the
wicked for
{\ADD{their}}
wickedness.
\VS{13}Whoso
stoppeth his
ears at the
cry of the
poor, he also shall
cry himself, but shall not be
heard.
\VS{14}A
gift in
secret
pacifieth
anger: and a
reward in the
bosom
strong
wrath.
\VS{15}{\ADD{It is}}
joy to the
just to
do
judgment: but
destruction
{\ADD{shall be}} to the
workers of
iniquity.
\VS{16}The
man that
wandereth out of the
way of
understanding shall
remain in the
congregation of the
dead.
\VS{17}He that
loveth
pleasure
{\ADD{shall be}} a
poor
man: he that
loveth
wine and
oil shall not be
rich.\FTNT{‡‡}{{\FR 21.17 }
pleasure: or, sport}
\VS{18}The
wicked
{\ADD{shall be}} a
ransom for the
righteous, and the
transgressor for the
upright.
\VS{19}{\ADD{It is}}
better to
dwell in the
wilderness, than with a
contentious and an
angry
woman.\FTNT{§§}{{\FR 21.19 }
in…: Heb. in the land of the desert}
\VS{20}{\ADD{There is}}
treasure to be
desired and
oil in the
dwelling of the
wise; but a
foolish
man spendeth it
up.
\VS{21}He that
followeth after
righteousness and
mercy
findeth
life,
righteousness, and
honour.
\VS{22}A
wise
{\ADD{man}}
scaleth the
city of the
mighty, and casteth
down the
strength of the
confidence thereof.
\VS{23}Whoso
keepeth his
mouth and his
tongue
keepeth his
soul from
troubles.
\VS{24}Proud
{\ADD{and}}
haughty
scorner
{\ADD{is}} his
name, who
dealeth in
proud
wrath.\FTNT{***}{{\FR 21.24 }
in proud…: Heb. in the wrath of pride}
\par }{\PP \VS{25}The
desire of the
slothful
killeth him; for his
hands
refuse to
labour.
\VS{26}He
coveteth
greedily all the
day long: but the
righteous
giveth and
spareth not.
\VS{27}The
sacrifice of the
wicked
{\ADD{is}}
abomination: how much more,
{\ADD{when}} he
bringeth it with a wicked
mind?\FTNT{†††}{{\FR 21.27 }
with…: Heb. in wickedness?}
\VS{28}A
false
witness shall
perish: but the
man that
heareth
speaketh
constantly.\FTNT{‡‡‡}{{\FR 21.28 }
A…: Heb. A witness of lies}
\VS{29}A
wicked
man
hardeneth his
face: but
{\ADD{as for}} the
upright, he
directeth his
way.\FTNT{§§§}{{\FR 21.29 }
directeth: or, considereth}
\par }{\PP \VS{30}{\ADD{There is}} no
wisdom nor
understanding nor
counsel against the
{\ND{
LORD}}.
\VS{31}The
horse
{\ADD{is}}
prepared against the
day of
battle: but
safety
{\ADD{is}} of the
{\ND{
LORD}}.\FTNT{*}{{\FR 21.31 }
safety: or, victory}

\par }\Chap{22}{\PP \VerseOne{1}A
{\ADD{good}}
name
{\ADD{is}} rather to be
chosen than
great
riches,
{\ADD{and}}
loving
favour rather than
silver and
gold.\FTNT{*}{{\FR 22.1 }
loving…: or, favour is better than, etc}
\VS{2}The
rich and
poor meet
together: the
{\ND{
LORD}}
{\ADD{is}} the
maker of them all.
\VS{3}A
prudent
{\ADD{man}}
foreseeth the
evil, and
hideth himself: but the
simple pass
on, and are
punished.
\VS{4}By
humility
{\ADD{and}} the
fear of the
{\ND{
LORD}}
{\ADD{are}}
riches, and
honour, and
life.\FTNT{†}{{\FR 22.4 }
By…: or, The reward of humility, etc}
\VS{5}Thorns
{\ADD{and}}
snares
{\ADD{are}} in the
way of the
froward: he that doth
keep his
soul shall be
far from them.
\VS{6}Train
up a
child in the
way he should
go: and when he is
old, he will not
depart from it.\FTNT{‡}{{\FR 22.6 }
Train…: or, Catechise}\FTNT{§}{{\FR 22.6 }
in…: Heb. in his way}
\VS{7}The
rich
ruleth over the
poor, and the
borrower
{\ADD{is}}
servant to the
lender.\FTNT{**}{{\FR 22.7 }
the lender: Heb. the man that lendeth}
\VS{8}He that
soweth
iniquity shall
reap
vanity: and the
rod of his
anger shall
fail.\FTNT{††}{{\FR 22.8 }
the rod…: or, with the rod of his anger he shall be consumed}
\VS{9}He that hath a
bountiful
eye shall be
blessed; for he
giveth of his
bread to the
poor.\FTNT{‡‡}{{\FR 22.9 }
He that…: Heb. Good of eye}
\VS{10}Cast
out the
scorner, and
contention shall go
out; yea,
strife and
reproach shall
cease.
\VS{11}He that
loveth
pureness of
heart,
{\ADD{for}} the
grace of his
lips the
king
{\ADD{shall be}} his
friend.\FTNT{§§}{{\FR 22.11 }
for…: or, and hath grace in his lips}
\VS{12}The
eyes of the
{\ND{
LORD}}
preserve
knowledge, and he
overthroweth the
words of the
transgressor.\FTNT{***}{{\FR 22.12 }
the words: or, the matters}
\VS{13}The
slothful
{\ADD{man}}
saith,
{\ADD{There is}} a
lion
without, I shall be
slain
in the
streets.
\VS{14}The
mouth of strange
women
{\ADD{is}} a
deep
pit: he that is
abhorred of the
{\ND{
LORD}} shall
fall therein.
\VS{15}Foolishness
{\ADD{is}}
bound in the
heart of a
child;
{\ADD{but}} the
rod of
correction shall drive it
far from him.
\VS{16}He that
oppresseth the
poor to
increase his
{\ADD{riches, and}} he that
giveth to the
rich,
{\ADD{shall}} surely
{\ADD{come}} to
want.
\par }{\PP \VS{17}Bow
down thine
ear, and
hear the
words of the
wise, and
apply thine
heart unto my
knowledge.
\VS{18}For
{\ADD{it is}} a pleasant
thing if thou
keep them
within thee; they shall
withal be
fitted in thy
lips.\FTNT{†††}{{\FR 22.18 }
within…: Heb. in thy belly}
\VS{19}That thy
trust may be in the
{\ND{
LORD}}, I have made
known to thee this
day, even to thee.\FTNT{‡‡‡}{{\FR 22.19 }
even…: or, trust thou also}
\VS{20}Have not I
written to
thee excellent
things in
counsels and
knowledge,
\VS{21}That I might make thee
know the
certainty of the
words of
truth; that thou mightest
answer the
words of
truth to them that
send unto thee?\FTNT{§§§}{{\FR 22.21 }
to them…: or, to those that send thee?}
\par }{\PP \VS{22}Rob not the
poor, because he
{\ADD{is}}
poor: neither
oppress the
afflicted in the
gate:
\VS{23}For the
{\ND{
LORD}} will
plead their
cause, and
spoil the
soul of those that
spoiled them.
\par }{\PP \VS{24}Make no
friendship with an
angry
man; and with a
furious
man thou shalt not
go:
\VS{25}Lest thou
learn his
ways, and
get a
snare to thy
soul.
\par }{\PP \VS{26}Be not thou
{\ADD{one}} of them that
strike
hands,
{\ADD{or}} of them that are
sureties for
debts.
\VS{27}If thou hast nothing to
pay, why should he take
away thy
bed from under thee?
\VS{28}Remove not the
ancient
landmark, which thy
fathers have
set.\FTNT{*}{{\FR 22.28 }
landmark: or, bound}
\VS{29}Seest thou a
man
diligent in his
business? he shall
stand
before
kings; he shall not
stand
before
mean
{\ADD{men}}.\FTNT{†}{{\FR 22.29 }
mean…: Heb. obscure men}

\par }\Chap{23}{\PP \VerseOne{1}When thou
sittest to
eat with a
ruler,
consider
diligently what
{\ADD{is}}
before thee:
\VS{2}And
put a
knife to thy
throat, if thou
{\ADD{be}} a man
given to
appetite.
\VS{3}Be not
desirous of his
dainties: for they
{\ADD{are}}
deceitful
meat.
\par }{\PP \VS{4}Labour not to be
rich:
cease from thine own
wisdom.
\VS{5}Wilt thou
set thine
eyes upon that which is not? for
{\ADD{riches}}
certainly
make themselves
wings; they fly
away as an
eagle toward
heaven.\FTNT{*}{{\FR 23.5 }
set…: Heb. cause thine eyes to fly upon}
\par }{\PP \VS{6}Eat thou not the
bread of
{\ADD{him that hath}} an
evil
eye, neither
desire thou his dainty
meats:
\VS{7}For as he
thinketh in his
heart, so
{\ADD{is}} he:
Eat and
drink,
saith he to thee; but his
heart
{\ADD{is}} not with thee.
\VS{8}The
morsel
{\ADD{which}} thou hast
eaten shalt thou vomit
up, and
lose thy
sweet
words.
\VS{9}Speak not in the
ears of a
fool: for he will
despise the
wisdom of thy
words.
\par }{\PP \VS{10}Remove not the
old
landmark; and
enter not into the
fields of the
fatherless:\FTNT{†}{{\FR 23.10 }
landmark: or, bound}
\VS{11}For their
redeemer
{\ADD{is}}
mighty; he shall
plead their
cause with thee.
\par }{\PP \VS{12}Apply thine
heart unto
instruction, and thine
ears to the
words of
knowledge.
\VS{13}Withhold not
correction from the
child: for
{\ADD{if}} thou
beatest him with the
rod, he shall not
die.
\VS{14}Thou shalt
beat him with the
rod, and shalt
deliver his
soul from
hell.
\VS{15}My
son, if thine
heart be
wise, my
heart shall
rejoice, even
mine.\FTNT{‡}{{\FR 23.15 }
even…: or, even I will rejoice}
\VS{16}Yea, my
reins shall
rejoice, when thy
lips
speak right
things.
\par }{\PP \VS{17}Let not thine
heart
envy
sinners: but
{\ADD{be thou}} in the
fear of the
{\ND{
LORD}} all the
day long.
\VS{18}For surely there
is an
end; and thine
expectation shall not be cut
off.\FTNT{§}{{\FR 23.18 }
end: or, reward}
\par }{\PP \VS{19}Hear thou, my
son, and be
wise, and
guide thine
heart in the
way.
\VS{20}Be not among
winebibbers; among riotous
eaters of
flesh:\FTNT{**}{{\FR 23.20 }
of flesh: Heb. of their flesh}
\VS{21}For the
drunkard and the
glutton shall come to
poverty: and
drowsiness shall
clothe
{\ADD{a man}} with
rags.
\VS{22}Hearken unto thy
father that
begat thee, and
despise not thy
mother when she is
old.
\VS{23}Buy the
truth, and
sell
{\ADD{it}} not;
{\ADD{also}}
wisdom, and
instruction, and
understanding.
\VS{24}The
father of the
righteous shall
greatly
rejoice: and he that
begetteth a
wise
{\ADD{child}} shall have
joy of him.
\VS{25}Thy
father and thy
mother shall be
glad, and she that
bare thee shall
rejoice.
\VS{26}My
son,
give me thine
heart, and let thine
eyes
observe my
ways.
\VS{27}For a
whore
{\ADD{is}} a
deep
ditch; and a strange
woman
{\ADD{is}} a
narrow
pit.
\VS{28}She also lieth in
wait as
{\ADD{for}} a
prey, and
increaseth the
transgressors among
men.\FTNT{††}{{\FR 23.28 }
as for…: or, as a robber}
\par }{\PP \VS{29}Who hath
woe? who hath
sorrow? who hath
contentions? who hath
babbling? who hath
wounds without
cause? who hath
redness of
eyes?
\VS{30}They that tarry
long at the
wine; they that
go to
seek mixed
wine.
\VS{31}Look not thou upon the
wine when it is
red, when it
giveth his
colour in the
cup,
{\ADD{when}} it
moveth itself
aright.
\VS{32}At the
last it
biteth like a
serpent, and
stingeth like an
adder.\FTNT{‡‡}{{\FR 23.32 }
an…: or, a cockatrice}
\VS{33}Thine
eyes shall
behold strange
women, and thine
heart shall
utter perverse
things.
\VS{34}Yea, thou shalt be as he that lieth
down in the
midst of the
sea, or as he that
lieth upon the
top of a
mast.\FTNT{§§}{{\FR 23.34 }
the midst…: Heb. the heart of the sea}
\VS{35}They have
stricken me,
{\ADD{shalt thou say, and}} I was not
sick; they have
beaten me,
{\ADD{and}} I
felt
{\ADD{it}} not: when shall I
awake? I will
seek it
yet again.\FTNT{***}{{\FR 23.35 }
I felt…: Heb. I knew it not}

\par }\Chap{24}{\PP \VerseOne{1}Be not thou
envious against
evil
men, neither
desire to be with them.
\VS{2}For their
heart
studieth
destruction, and their
lips
talk of
mischief.
\par }{\PP \VS{3}Through
wisdom is an
house
builded; and by
understanding it is
established:
\VS{4}And by
knowledge shall the
chambers be
filled with all
precious and
pleasant
riches.
\VS{5}A
wise
man
{\ADD{is}}
strong; yea, a
man of
knowledge
increaseth
strength.\FTNT{*}{{\FR 24.5 }
is strong: Heb. is in strength}\FTNT{†}{{\FR 24.5 }
increaseth…: Heb. strengtheneth might}
\VS{6}For by wise
counsel thou shalt
make thy
war: and in
multitude of
counsellors
{\ADD{there is}}
safety.
\par }{\PP \VS{7}Wisdom
{\ADD{is}} too
high for a
fool: he
openeth not his
mouth in the
gate.
\VS{8}He that
deviseth to do
evil shall be
called a
mischievous
person.
\VS{9}The
thought of
foolishness
{\ADD{is}}
sin: and the
scorner
{\ADD{is}} an
abomination to
men.
\VS{10}{\ADD{If}} thou
faint in the
day of
adversity, thy
strength
{\ADD{is}}
small.\FTNT{‡}{{\FR 24.10 }
small: Heb. narrow}
\par }{\PP \VS{11}If thou
forbear to
deliver
{\ADD{them that are}}
drawn unto
death, and
{\ADD{those that are}}
ready to be
slain;
\VS{12}If thou
sayest, Behold, we
knew it not; doth not he that
pondereth the
heart
consider
{\ADD{it}}? and he that
keepeth thy
soul, doth
{\ADD{not}} he
know
{\ADD{it}}? and shall
{\ADD{not}} he
render to
{\ADD{every}}
man according to his
works?
\par }{\PP \VS{13}My
son,
eat thou
honey, because
{\ADD{it is}}
good; and the
honeycomb,
{\ADD{which is}}
sweet to thy
taste:\FTNT{§}{{\FR 24.13 }
to…: Heb. upon thy palate}
\VS{14}So
{\ADD{shall}} the
knowledge of
wisdom
{\ADD{be}} unto thy
soul: when thou hast
found
{\ADD{it}}, then there shall
be a
reward, and thy
expectation shall not be cut
off.
\par }{\PP \VS{15}Lay not
wait, O
wicked
{\ADD{man}}, against the
dwelling of the
righteous;
spoil not his resting
place:
\VS{16}For a
just
{\ADD{man}}
falleth seven
times, and riseth up
again: but the
wicked shall
fall into
mischief.
\par }{\PP \VS{17}Rejoice not when thine
enemy
falleth, and let not thine
heart be
glad when he
stumbleth:
\VS{18}Lest the
{\ND{
LORD}}
see
{\ADD{it}}, and it
displease him, and he turn
away his
wrath from him.\FTNT{**}{{\FR 24.18 }
it displease…: Heb. it be evil in his eyes}
\par }{\PP \VS{19}Fret not thyself because of
evil
{\ADD{men}}, neither be thou
envious at the
wicked;\FTNT{††}{{\FR 24.19 }
Fret…: or, Keep not company with the wicked}
\VS{20}For there shall be no
reward to the
evil
{\ADD{man;}} the
candle of the
wicked shall be put
out.\FTNT{‡‡}{{\FR 24.20 }
candle: or, lamp}
\par }{\PP \VS{21}My
son,
fear thou the
{\ND{
LORD}} and the
king:
{\ADD{and}}
meddle not with them that are given to
change:\FTNT{§§}{{\FR 24.21 }
them…: Heb. changers}
\VS{22}For their
calamity shall
rise
suddenly; and who
knoweth the
ruin of them
both?
\par }{\PP \VS{23}These
{\ADD{things}} also
{\ADD{belong}} to the
wise.
{\ADD{It is}} not
good to have
respect of
persons in
judgment.
\VS{24}He that
saith unto the
wicked, Thou
{\ADD{art}}
righteous; him shall the
people
curse,
nations shall
abhor him:
\VS{25}But to them that
rebuke
{\ADD{him}} shall be
delight, and a
good
blessing shall
come upon them.\FTNT{***}{{\FR 24.25 }
a good…: Heb. a blessing of good}
\VS{26}{\ADD{Every man}} shall
kiss
{\ADD{his}}
lips that
giveth a
right
answer.\FTNT{†††}{{\FR 24.26 }
that…: Heb. that answereth right words}
\VS{27}Prepare thy
work
without, and make it
fit for thyself in the
field; and
afterwards
build thine
house.
\par }{\PP \VS{28}Be not a
witness against thy
neighbour without
cause; and
deceive
{\ADD{not}} with thy
lips.
\VS{29}Say not, I will
do so to him as he hath
done to me: I will
render to the
man according to his
work.
\par }{\PP \VS{30}I
went by the
field of the
slothful, and by the
vineyard of the
man
void of
understanding;
\VS{31}And, lo, it was all grown
over with
thorns,
{\ADD{and}}
nettles had
covered the
face thereof, and the
stone
wall thereof was broken
down.
\VS{32}Then I
saw,
{\ADD{and}} considered
{\ADD{it}}
well: I
looked upon
{\ADD{it, and}}
received
instruction.\FTNT{‡‡‡}{{\FR 24.32 }
considered…: Heb. set my heart}
\VS{33}{\ADD{Yet}} a
little
sleep, a
little
slumber, a
little
folding of the
hands to
sleep:
\VS{34}So shall thy
poverty
come
{\ADD{as}} one that
travelleth; and thy
want as an
armed
man.\FTNT{§§§}{{\FR 24.34 }
an…: Heb. a man of shield}

\par }\Chap{25}{\PP \VerseOne{1}These
{\ADD{are}} also
proverbs of
Solomon, which the
men of
Hezekiah
king of
Judah copied
out.
\par }{\PP \VS{2}{\ADD{It is}} the
glory of
God to
conceal a
thing: but the
honour of
kings
{\ADD{is}} to search
out a
matter.
\VS{3}The
heaven for
height, and the
earth for
depth, and the
heart of
kings
{\ADD{is}}
unsearchable.\FTNT{*}{{\FR 25.3 }
is…: Heb. there is no searching}
\par }{\PP \VS{4}Take
away the
dross from the
silver, and there shall come
forth a
vessel for the
finer.
\VS{5}Take
away the
wicked
{\ADD{from}}
before the
king, and his
throne shall be
established in
righteousness.
\par }{\PP \VS{6}Put not
forth thyself in the
presence of the
king, and
stand not in the
place of
great
{\ADD{men:}}\FTNT{†}{{\FR 25.6 }
Put…: Heb. Set not out thy glory}
\VS{7}For
better
{\ADD{it is}} that it be
said unto thee, Come
up hither; than that thou shouldest be put
lower in the
presence of the
prince whom thine
eyes have
seen.
\par }{\PP \VS{8}Go not
forth
hastily to
strive, lest
{\ADD{thou know not}} what to
do in the
end thereof, when thy
neighbour hath put thee to
shame.
\VS{9}Debate thy
cause with thy
neighbour
{\ADD{himself;}} and
discover not a
secret to
another:\FTNT{‡}{{\FR 25.9 }
a secret…: or, the secret of}
\VS{10}Lest he that
heareth
{\ADD{it}} put thee to
shame, and thine
infamy turn not
away.
\par }{\PP \VS{11}A
word
fitly
spoken
{\ADD{is like}}
apples of
gold in
pictures of
silver.\FTNT{§}{{\FR 25.11 }
fitly…: Heb. spoken upon his wheels}
\VS{12}{\ADD{As}} an
earring of
gold, and an
ornament of fine
gold,
{\ADD{so is}} a
wise
reprover upon an
obedient
ear.
\VS{13}As the
cold of
snow in the
time of
harvest,
{\ADD{so is}} a
faithful
messenger to them that
send him: for he
refresheth the
soul of his
masters.
\VS{14}Whoso
boasteth himself of a
false
gift
{\ADD{is like}}
clouds and
wind without
rain.\FTNT{**}{{\FR 25.14 }
of a…: Heb. in a gift of falsehood}
\VS{15}By
long
forbearing is a
prince
persuaded, and a
soft
tongue
breaketh the
bone.
\VS{16}Hast thou
found
honey?
eat so much as is
sufficient for thee, lest thou be
filled therewith, and
vomit it.
\VS{17}Withdraw thy
foot from thy
neighbour’s
house; lest he be
weary of thee, and
{\ADD{so}}
hate thee.\FTNT{††}{{\FR 25.17 }
Withdraw…: or, Let thy foot be seldom in}\FTNT{‡‡}{{\FR 25.17 }
weary…: Heb. full of thee}
\VS{18}A
man that
beareth
false
witness against his
neighbour
{\ADD{is}} a
maul, and a
sword, and a
sharp
arrow.
\VS{19}Confidence in an unfaithful
man in
time of
trouble
{\ADD{is like}} a
broken
tooth, and a
foot out of
joint.
\VS{20}{\ADD{As}} he that taketh
away a
garment in
cold
weather,
{\ADD{and as}}
vinegar upon
nitre, so
{\ADD{is}} he that
singeth
songs to an
heavy
heart.
\par }{\PP \VS{21}If thine
enemy be
hungry, give him
bread to
eat; and if he be
thirsty, give him
water to
drink:
\VS{22}For thou shalt
heap coals of
fire upon his
head, and the
{\ND{
LORD}} shall
reward thee.
\VS{23}The
north
wind driveth
away
rain: so
{\ADD{doth}} an
angry
countenance a
backbiting
tongue.\FTNT{§§}{{\FR 25.23 }
driveth…: or, bringeth forth rain: so doth a backbiting tongue an angry countenance}
\VS{24}{\ADD{It is}}
better to
dwell in the
corner of the
housetop, than with a
brawling
woman and in a
wide
house.
\VS{25}{\ADD{As}}
cold
waters to a
thirsty
soul, so
{\ADD{is}}
good
news from a
far
country.
\VS{26}A
righteous man falling
down
before the
wicked
{\ADD{is as}} a
troubled
fountain, and a
corrupt
spring.
\VS{27}{\ADD{It is}} not
good to
eat
much
honey: so
{\ADD{for men}} to
search their own
glory
{\ADD{is not}}
glory.
\VS{28}He that
{\ADD{hath}} no
rule over his own
spirit
{\ADD{is like}} a
city
{\ADD{that is}} broken
down,
{\ADD{and}} without
walls.

\par }\Chap{26}{\PP \VerseOne{1}As
snow in
summer, and as
rain in
harvest, so
honour is not
seemly for a
fool.
\VS{2}As the
bird by
wandering, as the
swallow by
flying, so the
curse
causeless shall not
come.
\VS{3}A
whip for the
horse, a
bridle for the
ass, and a
rod for the
fool’s
back.
\par }{\PP \VS{4}Answer not a
fool according to his
folly, lest thou also be
like unto him.
\VS{5}Answer a
fool according to his
folly, lest he be
wise in his own
conceit.\FTNT{*}{{\FR 26.5 }
conceit: Heb. eyes}
\par }{\PP \VS{6}He that
sendeth a
message by the
hand of a
fool cutteth
off the
feet,
{\ADD{and}}
drinketh
damage.\FTNT{†}{{\FR 26.6 }
damage: or, violence}
\VS{7}The
legs of the
lame are not
equal: so
{\ADD{is}} a
parable in the
mouth of
fools.\FTNT{‡}{{\FR 26.7 }
are…: Heb. are lifted up}
\VS{8}As he that
bindeth a
stone in a
sling, so
{\ADD{is}} he that
giveth
honour to a
fool.\FTNT{§}{{\FR 26.8 }
bindeth…: or, putteth a precious stone in an heap of stones}
\VS{9}{\ADD{As}} a
thorn goeth
up into the
hand of a
drunkard, so
{\ADD{is}} a
parable in the
mouth of
fools.
\VS{10}The
great
{\ADD{God}} that
formed all
{\ADD{things}} both
rewardeth the
fool, and
rewardeth
transgressors.\FTNT{**}{{\FR 26.10 }
The great…: or, A great man grieveth all, and he hireth the fool, he hireth also transgressors}
\VS{11}As a
dog
returneth to his
vomit,
{\ADD{so}} a
fool
returneth to his
folly.\FTNT{††}{{\FR 26.11 }
returneth to his folly: Heb. iterateth his folly}
\VS{12}Seest thou a
man
wise in his own
conceit?
{\ADD{there is}} more
hope of a
fool than of him.
\VS{13}The
slothful
{\ADD{man}}
saith,
{\ADD{There is}} a
lion in the
way; a
lion
{\ADD{is}} in the
streets.
\VS{14}{\ADD{As}} the
door
turneth upon his
hinges, so
{\ADD{doth}} the
slothful upon his
bed.
\VS{15}The
slothful
hideth his
hand in
{\ADD{his}}
bosom; it
grieveth him to bring it
again to his
mouth.\FTNT{‡‡}{{\FR 26.15 }
it grieveth…: or, he is weary}
\VS{16}The
sluggard
{\ADD{is}}
wiser in his own
conceit than seven
men that can
render a
reason.
\VS{17}He that passeth
by,
{\ADD{and}}
meddleth with
strife
{\ADD{belonging}} not to him,
{\ADD{is like}} one that
taketh a
dog by the
ears.\FTNT{§§}{{\FR 26.17 }
meddleth: or, is enraged}
\par }{\PP \VS{18}As a
mad
{\ADD{man}} who
casteth
firebrands,
arrows, and
death,\FTNT{***}{{\FR 26.18 }
firebrands: Heb. flames, or, sparks}
\VS{19}So
{\ADD{is}} the
man
{\ADD{that}}
deceiveth his
neighbour, and
saith, Am not I in
sport?
\par }{\PP \VS{20}Where
no
wood is,
{\ADD{there}} the
fire goeth
out: so where
{\ADD{there is}} no
talebearer, the
strife
ceaseth.\FTNT{†††}{{\FR 26.20 }
Where no…: Heb. Without wood}\FTNT{‡‡‡}{{\FR 26.20 }
talebearer: or, whisperer}\FTNT{§§§}{{\FR 26.20 }
ceaseth: Heb. is silent}
\VS{21}{\ADD{As}}
coals
{\ADD{are}} to burning
coals, and
wood to
fire; so
{\ADD{is}} a
contentious
man to
kindle
strife.
\VS{22}The
words of a
talebearer
{\ADD{are}} as
wounds, and they go
down into the innermost
parts of the
belly.\FTNT{*}{{\FR 26.22 }
innermost…: Heb. chambers}
\VS{23}Burning
lips and a
wicked
heart
{\ADD{are like}} a
potsherd
covered with
silver
dross.
\par }{\PP \VS{24}He that
hateth
dissembleth with his
lips, and layeth
up
deceit
within him;\FTNT{†}{{\FR 26.24 }
dissembleth: or, is known}
\VS{25}When he
speaketh
fair,
believe him not: for
{\ADD{there are}}
seven
abominations in his
heart.\FTNT{‡}{{\FR 26.25 }
speaketh…: Heb. maketh his voice gracious}
\VS{26}{\ADD{Whose}}
hatred is
covered by
deceit, his
wickedness shall be
shewed before the
{\ADD{whole}}
congregation.\FTNT{§}{{\FR 26.26 }
by…: or, in secret}
\VS{27}Whoso
diggeth a
pit shall
fall therein: and he that
rolleth a
stone, it will
return upon him.
\VS{28}A
lying
tongue
hateth
{\ADD{those that are}}
afflicted by it; and a
flattering
mouth
worketh
ruin.

\par }\Chap{27}{\PP \VerseOne{1}Boast not thyself of to
morrow; for thou
knowest not what a
day may bring
forth.\FTNT{*}{{\FR 27.1 }
to…: Heb. to morrow day}
\VS{2}Let another
man
praise thee, and not thine own
mouth; a
stranger, and not thine own
lips.
\par }{\PP \VS{3}A
stone
{\ADD{is}}
heavy, and the
sand
weighty; but a
fool’s
wrath
{\ADD{is}}
heavier than them
both.\FTNT{†}{{\FR 27.3 }
heavy: Heb. heaviness}
\VS{4}Wrath
{\ADD{is}}
cruel, and
anger
{\ADD{is}}
outrageous; but who
{\ADD{is}} able to
stand
before
envy?\FTNT{‡}{{\FR 27.4 }
Wrath…: Heb. Wrath is cruelty, and anger an overflowing}\FTNT{§}{{\FR 27.4 }
envy: or, jealousy?}
\par }{\PP \VS{5}Open
rebuke
{\ADD{is}}
better than
secret
love.
\VS{6}Faithful
{\ADD{are}} the
wounds of a
friend; but the
kisses of an
enemy
{\ADD{are}}
deceitful.\FTNT{**}{{\FR 27.6 }
deceitful: or, earnest, or, frequent}
\VS{7}The
full
soul
loatheth an
honeycomb; but to the
hungry
soul every bitter
thing is
sweet.\FTNT{††}{{\FR 27.7 }
loatheth: Heb. treadeth under foot}
\VS{8}As a
bird that
wandereth from her
nest, so
{\ADD{is}} a
man that
wandereth from his
place.
\par }{\PP \VS{9}Ointment and
perfume
rejoice the
heart: so
{\ADD{doth}} the
sweetness of a
man’s friend by
hearty
counsel.\FTNT{‡‡}{{\FR 27.9 }
by…: Heb. from the counsel of the soul}
\VS{10}Thine own
friend, and thy
father’s
friend,
forsake not; neither
go into thy
brother’s
house in the
day of thy
calamity:
{\ADD{for}}
better
{\ADD{is}} a
neighbour
{\ADD{that is}}
near than a
brother far
off.
\VS{11}My
son, be
wise, and make my
heart
glad, that I may
answer him that
reproacheth me.
\VS{12}A
prudent
{\ADD{man}}
foreseeth the
evil,
{\ADD{and}}
hideth himself;
{\ADD{but}} the
simple pass
on,
{\ADD{and}} are
punished.
\VS{13}Take his
garment that is
surety for a
stranger, and take a
pledge of him for a strange
woman.
\VS{14}He that
blesseth his
friend with a
loud
voice, rising
early in the
morning, it shall be
counted a
curse to him.
\par }{\PP \VS{15}A
continual
dropping in a very
rainy
day and a
contentious
woman are
alike.
\VS{16}Whosoever
hideth her
hideth the
wind, and the
ointment of his right
hand,
{\ADD{which}}
bewrayeth
{\ADD{itself}}.
\VS{17}Iron
sharpeneth
iron; so a
man
sharpeneth the
countenance of his
friend.
\VS{18}Whoso
keepeth the fig
tree shall
eat the
fruit thereof: so he that
waiteth on his
master shall be
honoured.
\VS{19}As in
water
face
{\ADD{answereth}} to
face, so the
heart of
man to
man.
\VS{20}Hell and
destruction are
never
full; so the
eyes of
man are never
satisfied.\FTNT{§§}{{\FR 27.20 }
never: Heb. not}
\VS{21}{\ADD{As}} the fining
pot for
silver, and the
furnace for
gold; so
{\ADD{is}} a
man
to his
praise.
\VS{22}Though thou shouldest
bray a
fool in a
mortar
among
wheat with a
pestle,
{\ADD{yet}} will not his
foolishness
depart from him.
\par }{\PP \VS{23}Be thou
diligent to
know the
state of thy
flocks,
{\ADD{and}}
look
well to thy
herds.\FTNT{***}{{\FR 27.23 }
look…: Heb. set thy heart}
\VS{24}For
riches
{\ADD{are}} not for
ever: and doth the
crown
{\ADD{endure}} to
every
generation?\FTNT{†††}{{\FR 27.24 }
riches: Heb. strength}\FTNT{‡‡‡}{{\FR 27.24 }
to…: Heb. to generation and generation?}
\VS{25}The
hay
appeareth, and the tender
grass
sheweth itself, and
herbs of the
mountains are
gathered.
\VS{26}The
lambs
{\ADD{are}} for thy
clothing, and the
goats
{\ADD{are}} the
price of the
field.
\VS{27}And
{\ADD{thou shalt have}}
goats’
milk
enough for thy
food, for the
food of thy
household, and
{\ADD{for}} the
maintenance for thy
maidens.\FTNT{§§§}{{\FR 27.27 }
maintenance: Heb. life}

\par }\Chap{28}{\PP \VerseOne{1}The
wicked
flee when no man
pursueth: but the
righteous are
bold as a
lion.
\VS{2}For the
transgression of a
land
many
{\ADD{are}} the
princes thereof: but by a
man of
understanding
{\ADD{and}}
knowledge the
state
{\ADD{thereof}} shall be
prolonged.\FTNT{*}{{\FR 28.2 }
by…: or, by men of understanding and wisdom shall they likewise be prolonged}
\VS{3}A
poor
man that
oppresseth the
poor
{\ADD{is like}} a
sweeping
rain which leaveth no
food.\FTNT{†}{{\FR 28.3 }
which…: Heb. without food}
\VS{4}They that
forsake the
law
praise the
wicked: but such as
keep the
law
contend with them.
\VS{5}Evil
men
understand not
judgment: but they that
seek the
{\ND{
LORD}}
understand all
{\ADD{things}}.
\VS{6}Better
{\ADD{is}} the
poor that
walketh in his
uprightness, than
{\ADD{he that is}}
perverse
{\ADD{in his}}
ways, though he
{\ADD{be}}
rich.
\VS{7}Whoso
keepeth the
law
{\ADD{is}} a
wise
son: but he that is a
companion of
riotous
{\ADD{men}}
shameth his
father.\FTNT{‡}{{\FR 28.7 }
is a companion…: or, feedeth gluttons}
\VS{8}He that by
usury and unjust
gain
increaseth his
substance, he shall
gather it for him that will
pity the
poor.\FTNT{§}{{\FR 28.8 }
unjust…: Heb. by increase}
\VS{9}He that turneth
away his
ear from
hearing the
law, even his
prayer
{\ADD{shall be}}
abomination.
\VS{10}Whoso causeth the
righteous to go
astray in an
evil
way, he shall
fall himself into his own
pit: but the
upright shall have
good
{\ADD{things}} in
possession.
\VS{11}The
rich
man
{\ADD{is}}
wise in his own
conceit; but the
poor that hath
understanding searcheth him
out.\FTNT{**}{{\FR 28.11 }
own…: Heb. eyes}
\VS{12}When
righteous
{\ADD{men}} do
rejoice,
{\ADD{there is}}
great
glory: but when the
wicked
rise, a
man is
hidden.\FTNT{††}{{\FR 28.12 }
hidden: or, sought for}
\VS{13}He that
covereth his
sins shall not
prosper: but whoso
confesseth and
forsaketh
{\ADD{them}} shall have
mercy.
\VS{14}Happy
{\ADD{is}} the
man that
feareth
alway: but he that
hardeneth his
heart shall
fall into
mischief.
\VS{15}{\ADD{As}} a
roaring
lion, and a
ranging
bear;
{\ADD{so is}} a
wicked
ruler over the
poor
people.
\VS{16}The
prince that
wanteth
understanding
{\ADD{is}} also a
great
oppressor:
{\ADD{but}} he that
hateth
covetousness shall
prolong
{\ADD{his}}
days.
\VS{17}A
man that doeth
violence to the
blood of
{\ADD{any}}
person shall
flee to the
pit; let no man
stay him.
\VS{18}Whoso
walketh
uprightly shall be
saved: but
{\ADD{he that is}}
perverse
{\ADD{in his}}
ways shall
fall at
once.
\VS{19}He that
tilleth his
land shall have
plenty of
bread: but he that
followeth after
vain
{\ADD{persons}} shall have poverty
enough.
\VS{20}A
faithful
man shall
abound with
blessings: but he that maketh
haste to be
rich shall not be
innocent.\FTNT{‡‡}{{\FR 28.20 }
innocent: or, unpunished}
\VS{21}To have
respect of
persons
{\ADD{is}} not
good: for for a
piece of
bread
{\ADD{that}}
man will
transgress.
\VS{22}He that
hasteth to be
rich
{\ADD{hath}} an
evil
eye, and
considereth not that
poverty shall
come upon him.\FTNT{§§}{{\FR 28.22 }
hasteth…: or, hath and evil eye hasteth to be rich}
\VS{23}He that
rebuketh a
man
afterwards shall
find more
favour than he that
flattereth with the
tongue.
\VS{24}Whoso
robbeth his
father or his
mother, and
saith,
{\ADD{It is}} no
transgression; the same
{\ADD{is}} the
companion of a
destroyer.\FTNT{***}{{\FR 28.24 }
a destroyer: Heb. a man destroying}
\VS{25}He that is of a
proud
heart stirreth
up
strife: but he that putteth his
trust in the
{\ND{
LORD}} shall be made
fat.
\VS{26}He that
trusteth in his own
heart is a
fool: but whoso
walketh
wisely, he shall be
delivered.
\VS{27}He that
giveth unto the
poor shall not
lack: but he that
hideth his
eyes shall have
many a
curse.
\VS{28}When the
wicked
rise,
men
hide themselves: but when they
perish, the
righteous
increase.

\par }\Chap{29}{\PP \VerseOne{1}He, that being often
reproved
hardeneth
{\ADD{his}}
neck, shall
suddenly be
destroyed, and that without
remedy.\FTNT{*}{{\FR 29.1 }
He…: Heb. A man of reproofs}
\VS{2}When the
righteous are in
authority, the
people
rejoice: but when the
wicked beareth
rule, the
people
mourn.
\FTNT{†}{{\FR 29.2 }
in…: or, increased}
\VS{3}Whoso
loveth
wisdom
rejoiceth his
father: but he that keepeth
company with
harlots
spendeth
{\ADD{his}}
substance.
\VS{4}The
king by
judgment
establisheth the
land: but
he that receiveth
gifts
overthroweth it.\FTNT{‡}{{\FR 29.4 }
he…: Heb. a man of oblations}
\VS{5}A
man that
flattereth his
neighbour
spreadeth a
net for his
feet.
\VS{6}In the
transgression of an
evil
man
{\ADD{there is}} a
snare: but the
righteous doth
sing and
rejoice.
\VS{7}The
righteous
considereth the
cause of the
poor:
{\ADD{but}} the
wicked
regardeth not to
know
{\ADD{it}}.
\VS{8}Scornful
men
bring a
city into a
snare: but
wise
{\ADD{men}} turn
away
wrath.\FTNT{§}{{\FR 29.8 }
bring…: or, set a city on fire}
\VS{9}{\ADD{If}} a
wise
man
contendeth with a
foolish
man, whether he
rage or
laugh,
{\ADD{there is}} no
rest.
\VS{10}The
bloodthirsty
hate the
upright: but the
just
seek his
soul.\FTNT{**}{{\FR 29.10 }
The bloodthirsty: Heb. Men of blood}
\VS{11}A
fool
uttereth all his
mind: but a
wise
{\ADD{man}}
keepeth it in till
afterwards.
\VS{12}If a
ruler
hearken to
lies, all his
servants
{\ADD{are}}
wicked.
\VS{13}The
poor and the
deceitful
man meet
together: the
{\ND{
LORD}}
lighteneth
both their
eyes.\FTNT{††}{{\FR 29.13 }
the deceitful…: or, the usurer}
\VS{14}The
king that
faithfully
judgeth the
poor, his
throne shall be
established for
ever.
\VS{15}The
rod and
reproof
give
wisdom: but a
child
left
{\ADD{to himself}} bringeth his
mother to
shame.
\VS{16}When the
wicked are
multiplied,
transgression
increaseth: but the
righteous shall
see their
fall.
\VS{17}Correct thy
son, and he shall give thee
rest; yea, he shall
give
delight unto thy
soul.
\VS{18}Where
{\ADD{there is}} no
vision, the
people
perish: but he that
keepeth the
law,
happy
{\ADD{is}} he.\FTNT{‡‡}{{\FR 29.18 }
perish: or, is made naked}
\VS{19}A
servant will not be
corrected by
words: for though he
understand he will not
answer.
\VS{20}Seest thou a
man
{\ADD{that is}}
hasty in his
words?
{\ADD{there is}} more
hope of a
fool than of him.\FTNT{§§}{{\FR 29.20 }
words: or, matters?}
\VS{21}He that delicately bringeth
up his
servant from a
child shall have him become
{\ADD{his}}
son at the
length.
\VS{22}An
angry
man stirreth
up
strife, and a
furious
man aboundeth
in
transgression.
\VS{23}A
man’s
pride shall bring him
low: but
honour shall
uphold the
humble in
spirit.
\VS{24}Whoso is
partner with a
thief
hateth his own
soul: he
heareth
cursing, and
bewrayeth
{\ADD{it}} not.
\VS{25}The
fear of
man
bringeth a
snare: but whoso putteth his
trust in the
{\ND{
LORD}} shall be
safe.\FTNT{***}{{\FR 29.25 }
safe: Heb. set on high}
\VS{26}Many
seek the
ruler’s
favour; but
{\ADD{every}}
man’s
judgment
{\ADD{cometh}} from the
{\ND{
LORD}}.\FTNT{†††}{{\FR 29.26 }
the ruler’s…: Heb. the face of a ruler}
\VS{27}An
unjust
man
{\ADD{is}} an
abomination to the
just: and
{\ADD{he that is}}
upright in the
way
{\ADD{is}}
abomination to the
wicked.

\par }\Chap{30}{\PP \VerseOne{1}The
words of
Agur the
son of
Jakeh,
{\ADD{even}} the
prophecy: the
man
spake unto
Ithiel, even unto
Ithiel and
Ucal,
\VS{2}Surely I
{\ADD{am}} more
brutish than
{\ADD{any}}
man, and have not the
understanding of a
man.
\VS{3}I neither
learned
wisdom, nor
have the
knowledge of the
holy.\FTNT{*}{{\FR 30.3 }
have: Heb. know}
\VS{4}Who hath ascended
up into
heaven, or
descended? who hath
gathered the
wind in his
fists? who hath
bound the
waters in a
garment? who hath
established all the
ends of the
earth? what
{\ADD{is}} his
name, and what
{\ADD{is}} his
son’s
name, if thou canst
tell?
\VS{5}Every
word of
God
{\ADD{is}}
pure: he
{\ADD{is}} a
shield unto them that put their
trust in him.\FTNT{†}{{\FR 30.5 }
pure: Heb. purified}
\VS{6}Add thou not unto his
words, lest he
reprove thee, and thou be found a
liar.
\par }{\PP \VS{7}Two
{\ADD{things}} have I
required of thee;
deny me
{\ADD{them}} not before I
die:\FTNT{‡}{{\FR 30.7 }
deny…: Heb. withhold not from me}
\VS{8}Remove
far from me
vanity and
lies:
give me neither
poverty nor
riches;
feed me with
food
convenient for me:\FTNT{§}{{\FR 30.8 }
convenient…: Heb. of my allowance}
\VS{9}Lest I be
full, and
deny
{\ADD{thee}}, and
say, Who
{\ADD{is}} the
{\ND{
LORD}}? or lest I be
poor, and
steal, and
take the
name of my
God
{\ADD{in vain}}.\FTNT{**}{{\FR 30.9 }
deny…: Heb. belie thee}
\par }{\PP \VS{10}Accuse not a
servant unto his
master, lest he
curse thee, and thou be found
guilty.\FTNT{††}{{\FR 30.10 }
Accuse…: Heb. Hurt not with thy tongue}
\VS{11}{\ADD{There is}} a
generation
{\ADD{that}}
curseth their
father, and doth not
bless their
mother.
\VS{12}{\ADD{There is}} a
generation
{\ADD{that are}}
pure in their own
eyes, and
{\ADD{yet}} is not
washed from their
filthiness.
\VS{13}{\ADD{There is}} a
generation, O how
lofty are their
eyes! and their
eyelids are lifted
up.
\VS{14}{\ADD{There is}} a
generation, whose
teeth
{\ADD{are as}}
swords, and their jaw
teeth
{\ADD{as}}
knives, to
devour the
poor from off the
earth, and the
needy from
{\ADD{among}}
men.
\par }{\PP \VS{15}The
horseleach hath
two
daughters,
{\ADD{crying}},
Give,
give. There are
three
{\ADD{things that}} are never
satisfied,
{\ADD{yea}},
four
{\ADD{things}}
say not,
{\ADD{It is}}
enough:\FTNT{‡‡}{{\FR 30.15 }
It is…: Heb. Wealth}
\VS{16}The
grave; and the
barren
womb; the
earth
{\ADD{that}} is not
filled with
water; and the
fire
{\ADD{that}}
saith not,
{\ADD{It is}}
enough.
\VS{17}The
eye
{\ADD{that}}
mocketh at
{\ADD{his}}
father, and
despiseth to
obey
{\ADD{his}}
mother, the
ravens of the
valley shall pick it
out, and the
young
eagles shall
eat it.\FTNT{§§}{{\FR 30.17 }
the valley: or, the brook}
\par }{\PP \VS{18}There be
three
{\ADD{things which}} are too
wonderful for me, yea,
four which I
know not:
\VS{19}The
way of an
eagle in the
air; the
way of a
serpent upon a
rock; the
way of a
ship in the
midst of the
sea; and the
way of a
man with a
maid.\FTNT{***}{{\FR 30.19 }
midst: Heb. heart}
\VS{20}Such
{\ADD{is}} the
way of an
adulterous
woman; she
eateth, and
wipeth her
mouth, and
saith, I have
done no
wickedness.
\VS{21}For
three
{\ADD{things}} the
earth is
disquieted, and for
four
{\ADD{which}} it
cannot
bear:
\VS{22}For a
servant when he
reigneth; and a
fool when he is
filled with
meat;
\VS{23}For an
odious
{\ADD{woman}} when she is
married; and an
handmaid that is
heir to her
mistress.
\par }{\PP \VS{24}There be
four
{\ADD{things which are}}
little upon the
earth, but they
{\ADD{are}}
exceeding
wise:\FTNT{†††}{{\FR 30.24 }
exceeding…: Heb. wise, made wise}
\VS{25}The
ants
{\ADD{are}} a
people not
strong, yet they
prepare their
meat in the
summer;
\VS{26}The
conies
{\ADD{are but}} a
feeble
folk, yet
make they their
houses in the
rocks;
\VS{27}The
locusts have no
king, yet go they
forth all of them by
bands;\FTNT{‡‡‡}{{\FR 30.27 }
by…: Heb. gathered together}
\VS{28}The
spider taketh
hold with her
hands, and is in
kings’
palaces.
\par }{\PP \VS{29}There be
three
{\ADD{things}} which
go
well, yea,
four are
comely in
going:
\VS{30}A
lion
{\ADD{which is}}
strongest among
beasts, and turneth not
away
for any;
\VS{31}A
greyhound; an he
goat
also; and a
king, against whom
{\ADD{there is}} no rising
up.\FTNT{§§§}{{\FR 30.31 }
greyhound: or, horse: Heb. girt in the loins}
\VS{32}If thou hast done
foolishly in lifting
up thyself, or if thou hast thought
evil,
{\ADD{lay}} thine
hand upon thy
mouth.
\VS{33}Surely the
churning of
milk bringeth
forth
butter, and the
wringing of the
nose bringeth
forth
blood: so the
forcing of
wrath bringeth
forth
strife.

\par }\Chap{31}{\PP \VerseOne{1}The
words of
king
Lemuel, the
prophecy that his
mother
taught him.
\VS{2}What, my
son? and what, the
son of my
womb? and what, the
son of my
vows?
\VS{3}Give not thy
strength unto
women, nor thy
ways to that which
destroyeth
kings.
\VS{4}{\ADD{It is}} not for
kings, O
Lemuel,
{\ADD{it is}} not for
kings to
drink
wine; nor
for
princes strong
drink:
\VS{5}Lest they
drink, and
forget the
law, and
pervert the
judgment of any of the
afflicted.\FTNT{*}{{\FR 31.5 }
pervert: Heb. alter}\FTNT{†}{{\FR 31.5 }
of any…: Heb. of all the sons of affliction}
\VS{6}Give strong
drink unto him that is ready to
perish, and
wine unto those that be of
heavy
hearts.\FTNT{‡}{{\FR 31.6 }
of…: Heb. bitter of soul}
\VS{7}Let him
drink, and
forget his
poverty, and
remember his
misery no more.
\VS{8}Open thy
mouth for the
dumb in the
cause of all such as are
appointed to
destruction.\FTNT{§}{{\FR 31.8 }
such…: Heb. the sons of destruction}
\VS{9}Open thy
mouth,
judge
righteously, and
plead the cause of the
poor and
needy.
\par }{\PP \VS{10}¶ Who can
find a
virtuous
woman? for her
price
{\ADD{is}}
far above
rubies.
\VS{11}The
heart of her
husband doth safely
trust in her, so that he shall have no
need of
spoil.
\VS{12}She will
do him
good and not
evil all the
days of her
life.
\VS{13}She
seeketh
wool, and
flax, and
worketh
willingly with her
hands.
\VS{14}She is like the
merchants’
ships; she
bringeth her
food from
afar.
\VS{15}She
riseth also while it is yet
night, and
giveth
meat to her
household, and a
portion to her
maidens.
\VS{16}She
considereth a
field, and
buyeth it: with the
fruit of her
hands she
planteth a
vineyard.\FTNT{**}{{\FR 31.16 }
buyeth: Heb. taketh}
\VS{17}She
girdeth her
loins with
strength, and
strengtheneth her
arms.
\VS{18}She
perceiveth that her
merchandise
{\ADD{is}}
good: her
candle goeth not
out by
night.\FTNT{††}{{\FR 31.18 }
She…: Heb. She tasteth}
\VS{19}She
layeth her
hands to the
spindle, and her
hands
hold the
distaff.
\VS{20}She stretcheth
out her
hand to the
poor; yea, she reacheth
forth her
hands to the
needy.\FTNT{‡‡}{{\FR 31.20 }
She…: Heb. She spreadeth}
\VS{21}She is not
afraid of the
snow for her
household: for all her
household
{\ADD{are}}
clothed with
scarlet.\FTNT{§§}{{\FR 31.21 }
scarlet: or, double garments}
\VS{22}She
maketh herself coverings of
tapestry; her
clothing
{\ADD{is}}
silk and
purple.
\VS{23}Her
husband is
known in the
gates, when he
sitteth among the
elders of the
land.
\VS{24}She
maketh fine
linen, and
selleth
{\ADD{it;}} and
delivereth
girdles unto the
merchant.
\VS{25}Strength and
honour
{\ADD{are}} her
clothing; and she shall
rejoice in
time to
come.
\VS{26}She
openeth her
mouth with
wisdom; and in her
tongue
{\ADD{is}} the
law of
kindness.
\VS{27}She looketh
well to the
ways of her
household, and
eateth not the
bread of
idleness.
\VS{28}Her
children arise
up, and call her
blessed; her
husband
{\ADD{also}}, and he
praiseth her.
\VS{29}Many
daughters have
done
virtuously, but thou
excellest them all.\FTNT{***}{{\FR 31.29 }
have…: or, have gotten riches}
\VS{30}Favour
{\ADD{is}}
deceitful, and
beauty
{\ADD{is}}
vain:
{\ADD{but}} a
woman
{\ADD{that}}
feareth the
{\ND{
LORD}}, she shall be
praised.
\VS{31}Give her of the
fruit of her
hands; and let her own
works
praise her in the
gates.
\par }