\NormalFont\ShortTitle{Psalms}
{\MT The Book of Psalms

\par }\PsalmChap{1}{\Q \VerseOne{1}Blessed
{\ADD{is}} the
man that
walketh not in the
counsel of the
ungodly, nor
standeth in the
way of
sinners, nor
sitteth in the
seat of the
scornful.\FTNT{*}{{\FR 1.1 }
ungodly: or, wicked}
\par }{\Q \VS{2}But his
delight
{\ADD{is}} in the
law of the
{\ND{
LORD}}; and in his
law doth he
meditate
day and
night.
\par }{\Q \VS{3}And he shall be like a
tree
planted by the
rivers of
water, that bringeth
forth his
fruit in his
season; his
leaf also shall not
wither; and whatsoever he
doeth shall
prosper.\FTNT{†}{{\FR 1.3 }
wither: Heb. fade}
\par }{\BB \par }{\Q \VS{4}The
ungodly
{\ADD{are}} not so: but
{\ADD{are}} like the
chaff which the
wind driveth
away.
\par }{\Q \VS{5}Therefore the
ungodly shall not
stand in the
judgment, nor
sinners in the
congregation of the
righteous.
\par }{\Q \VS{6}For the
{\ND{
LORD}}
knoweth the
way of the
righteous: but the
way of the
ungodly shall
perish.

\par }\PsalmChap{2}{\Q \VerseOne{1}Why do the
heathen
rage, and the
people
imagine a vain
thing?\FTNT{*}{{\FR 2.1 }
rage: or, tumultuously assemble}\FTNT{†}{{\FR 2.1 }
imagine: Heb. meditate}
\par }{\Q \VS{2}The
kings of the
earth
set themselves, and the
rulers take
counsel
together, against the
{\ND{
LORD}}, and against his
anointed,
{\ADD{saying}},
\par }{\Q \VS{3}Let us
break their
bands
asunder, and cast
away their
cords from us.
\par }{\Q \VS{4}He that
sitteth in the
heavens shall
laugh: the
Lord shall have them in
derision.
\par }{\Q \VS{5}Then shall he
speak unto them in his
wrath, and
vex them in his sore
displeasure.\FTNT{‡}{{\FR 2.5 }
vex: or, trouble}
\par }{\Q \VS{6}Yet have I
set my
king upon my
holy
hill of
Zion.\FTNT{§}{{\FR 2.6 }
set: Heb. anointed}\FTNT{**}{{\FR 2.6 }
upon…: Heb. upon Zion, the hill of my holiness}
\par }{\BB \par }{\Q \VS{7}I will
declare the
decree: the
{\ND{
LORD}} hath
said unto me, Thou
{\ADD{art}} my
Son; this
day have I
begotten thee.\FTNT{††}{{\FR 2.7 }
the decree: or, for a decree}
\par }{\Q \VS{8}Ask of me, and I shall
give
{\ADD{thee}} the
heathen
{\ADD{for}} thine
inheritance, and the uttermost
parts of the
earth
{\ADD{for}} thy
possession.
\par }{\Q \VS{9}Thou shalt
break them with a
rod of
iron; thou shalt dash them in
pieces like a
potter’s
vessel.
\par }{\BB \par }{\Q \VS{10}Be
wise now therefore, O ye
kings: be
instructed, ye
judges of the
earth.
\par }{\Q \VS{11}Serve the
{\ND{
LORD}} with
fear, and
rejoice with
trembling.
\par }{\Q \VS{12}Kiss the
Son, lest he be
angry, and ye
perish
{\ADD{from}} the
way, when his
wrath is
kindled but a
little.
Blessed
{\ADD{are}} all they that put their
trust in him.

\par }\PsalmChap{3}{\D A
Psalm of
David, when he
fled
from
Absalom his
son.
\par }{\Q \VerseOne{1}{\ND{
LORD}}, how are they
increased that
trouble me!
many
{\ADD{are}} they that rise
up against me.
\par }{\Q \VS{2}Many
{\ADD{there be}} which
say of my
soul,
{\ADD{There is}} no
help for him in
God.
Selah.
\par }{\Q \VS{3}But thou, O
{\ND{
LORD}},
{\ADD{art}} a
shield for me; my
glory, and the lifter
up of mine
head.\FTNT{*}{{\FR 3.3 }
for: or, about}
\par }{\BB \par }{\Q \VS{4}I
cried unto the
{\ND{
LORD}} with my
voice, and he
heard me out of his
holy
hill.
Selah.
\par }{\Q \VS{5}I laid me
down and
slept; I
awaked; for the
{\ND{
LORD}}
sustained me.
\par }{\Q \VS{6}I will not be
afraid of ten
thousands of
people, that have
set
{\ADD{themselves}} against me round
about.
\par }{\Q \VS{7}Arise, O
{\ND{
LORD}};
save me, O my
God: for thou hast
smitten all mine
enemies
{\ADD{upon}} the cheek
bone; thou hast
broken the
teeth of the
ungodly.
\par }{\Q \VS{8}Salvation
{\ADD{belongeth}} unto the
{\ND{
LORD}}: thy
blessing
{\ADD{is}} upon thy
people.
Selah.

\par }\PsalmChap{4}{\D To the chief
Musician on
Neginoth, A
Psalm of
David.
\par }{\Q \VerseOne{1}Hear me when I
call, O
God of my
righteousness: thou hast
enlarged me
{\ADD{when I was}} in
distress; have
mercy upon me, and
hear my
prayer.\FTNT{*}{{\FR 4.1 }
chief…: or, overseer}\FTNT{†}{{\FR 4.1 }
have…: or, be gracious unto me}
\par }{\Q \VS{2}O ye
sons of
men, how
long
{\ADD{will ye turn}} my
glory into
shame?
{\ADD{how long}} will ye
love
vanity,
{\ADD{and}}
seek after
leasing?
Selah.
\par }{\Q \VS{3}But
know that the
{\ND{
LORD}} hath set
apart him that is
godly for himself: the
{\ND{
LORD}} will
hear when I
call unto him.
\par }{\Q \VS{4}Stand in
awe, and
sin not:
commune with your own
heart upon your
bed, and be
still.
Selah.
\par }{\Q \VS{5}Offer the
sacrifices of
righteousness, and put your
trust in the
{\ND{
LORD}}.
\par }{\BB \par }{\Q \VS{6}{\ADD{There be}}
many that
say, Who will
shew us
{\ADD{any}}
good?
{\ND{
LORD}}, lift thou
up the
light of thy
countenance upon us.
\par }{\Q \VS{7}Thou hast
put
gladness in my
heart, more than in the
time
{\ADD{that}} their
corn and their
wine
increased.
\par }{\Q \VS{8}I will
both lay me
down in
peace, and
sleep: for thou,
{\ND{
LORD}},
only makest me
dwell in
safety.

\par }\PsalmChap{5}{\D To the chief
Musician upon
Nehiloth, A
Psalm of
David.
\par }{\Q \VerseOne{1}Give
ear to my
words, O
{\ND{
LORD}},
consider my
meditation.
\par }{\Q \VS{2}Hearken unto the
voice of my
cry, my
King, and my
God: for unto thee will I
pray.
\par }{\Q \VS{3}My
voice shalt thou
hear in the
morning, O
{\ND{
LORD}}; in the
morning will I
direct
{\ADD{my prayer}} unto thee, and will look
up.
\par }{\Q \VS{4}For thou
{\ADD{art}} not a
God that hath
pleasure in
wickedness: neither shall
evil
dwell with thee.
\par }{\Q \VS{5}The
foolish shall not
stand in thy
sight: thou
hatest all
workers of
iniquity.\FTNT{*}{{\FR 5.5 }
in…: Heb. before thine eyes}
\par }{\Q \VS{6}Thou shalt
destroy them that
speak
leasing: the
{\ND{
LORD}} will
abhor the
bloody and
deceitful
man.\FTNT{†}{{\FR 5.6 }
bloody…: Heb. man of bloods and deceit}
\par }{\BB \par }{\Q \VS{7}But as for me, I will
come
{\ADD{into}} thy
house in the
multitude of thy
mercy:
{\ADD{and}} in thy
fear will I
worship toward thy
holy
temple.\FTNT{‡}{{\FR 5.7 }
thy holy…: Heb. the temple of thy holiness}
\par }{\Q \VS{8}Lead me, O
{\ND{
LORD}}, in thy
righteousness because of mine
enemies; make thy
way
straight before my
face.\FTNT{§}{{\FR 5.8 }
mine…: Heb. those which observe me}
\par }{\Q \VS{9}For
{\ADD{there is}} no
faithfulness in their
mouth; their
inward part
{\ADD{is}} very
wickedness; their
throat
{\ADD{is}} an
open
sepulchre; they
flatter with their
tongue.\FTNT{**}{{\FR 5.9 }
faithfulness: or, stedfastness}\FTNT{††}{{\FR 5.9 }
their mouth: Heb. his mouth, that is, the mouth of any of them}\FTNT{‡‡}{{\FR 5.9 }
very…: Heb. wickednesses}
\par }{\Q \VS{10}Destroy thou them, O
God; let them
fall by their own
counsels; cast them
out in the
multitude of their
transgressions; for they have
rebelled against thee.\FTNT{§§}{{\FR 5.10 }
Destroy…: or, Make them guilty}\FTNT{***}{{\FR 5.10 }
by…: or, from their counsels}
\par }{\Q \VS{11}But let all those that put their
trust in thee
rejoice: let them
ever shout for
joy, because thou
defendest them: let them also that
love thy
name be
joyful in thee.\FTNT{†††}{{\FR 5.11 }
defendest…: Heb. coverest over, or, protectest them}
\par }{\Q \VS{12}For thou,
{\ND{
LORD}}, wilt
bless the
righteous; with
favour wilt thou
compass him as
{\ADD{with}} a
shield.\FTNT{‡‡‡}{{\FR 5.12 }
compass: Heb. crown}

\par }\PsalmChap{6}{\D To the chief
Musician on
Neginoth upon
Sheminith, A
Psalm of
David.
\par }{\Q \VerseOne{1}O
{\ND{
LORD}},
rebuke me not in thine
anger, neither
chasten me in thy hot
displeasure.\FTNT{*}{{\FR 6.1 }
chief…: or, overseer}\FTNT{†}{{\FR 6.1 }
Sheminith: or, the eighth}
\par }{\Q \VS{2}Have
mercy upon me, O
{\ND{
LORD}}; for I
{\ADD{am}}
weak: O
{\ND{
LORD}},
heal me; for my
bones are
vexed.
\par }{\Q \VS{3}My
soul is also
sore
vexed: but thou, O
{\ND{
LORD}}, how long?
\par }{\Q \VS{4}Return, O
{\ND{
LORD}},
deliver my
soul: oh
save me for thy
mercies’ sake.
\par }{\Q \VS{5}For in
death
{\ADD{there is}} no
remembrance of thee: in the
grave who shall give thee
thanks?
\par }{\Q \VS{6}I am
weary with my
groaning; all the
night make I my
bed to
swim; I
water my
couch with my
tears.\FTNT{‡}{{\FR 6.6 }
all…: or, every night}
\par }{\Q \VS{7}Mine
eye is
consumed because of
grief; it waxeth
old because of all mine
enemies.
\par }{\BB \par }{\Q \VS{8}Depart from me, all ye
workers of
iniquity; for the
{\ND{
LORD}} hath
heard the
voice of my
weeping.
\par }{\Q \VS{9}The
{\ND{
LORD}} hath
heard my
supplication; the
{\ND{
LORD}} will
receive my
prayer.
\par }{\Q \VS{10}Let all mine
enemies be
ashamed and
sore
vexed: let them
return
{\ADD{and}} be
ashamed
suddenly.

\par }\PsalmChap{7}{\D Shiggaion of
David, which he
sang unto the
{\ND{
LORD}}, concerning the
words of
Cush the
Benjamite.
\par }{\Q \VerseOne{1}O
{\ND{
LORD}} my
God, in thee do I put my
trust:
save me from all them that
persecute me, and
deliver me:\FTNT{*}{{\FR 7.1 }
words: or, business}
\par }{\Q \VS{2}Lest he
tear my
soul like a
lion, rending
{\ADD{it}} in
pieces, while
{\ADD{there is}} none to
deliver.\FTNT{†}{{\FR 7.2 }
none…: Heb. not a deliverer}
\par }{\Q \VS{3}O
{\ND{
LORD}} my
God, if I have
done this; if there
be
iniquity in my
hands;
\par }{\Q \VS{4}If I have
rewarded
evil unto him that was at
peace with me; (yea, I have
delivered him that without
cause is mine
enemy:)
\par }{\Q \VS{5}Let the
enemy
persecute my
soul, and
take
{\ADD{it;}} yea, let him tread
down my
life upon the
earth, and
lay mine
honour in the
dust.
Selah.
\par }{\Q \VS{6}Arise, O
{\ND{
LORD}}, in thine
anger, lift
up thyself because of the
rage of mine
enemies: and
awake for me
{\ADD{to}} the
judgment
{\ADD{that}} thou hast
commanded.
\par }{\Q \VS{7}So shall the
congregation of the
people compass thee
about: for their sakes therefore
return thou on
high.
\par }{\Q \VS{8}The
{\ND{
LORD}} shall
judge the
people:
judge me, O
{\ND{
LORD}}, according to my
righteousness, and according to mine
integrity
{\ADD{that is}} in me.
\par }{\Q \VS{9}Oh let the
wickedness of the
wicked come to an
end; but
establish the
just: for the
righteous
God
trieth the
hearts and
reins.
\par }{\BB \par }{\Q \VS{10}My
defence
{\ADD{is}} of
God, which
saveth the
upright in
heart.\FTNT{‡}{{\FR 7.10 }
My…: Heb. My buckler is upon God}
\par }{\Q \VS{11}God
judgeth the
righteous, and
God is
angry
{\ADD{with the wicked}} every
day.\FTNT{§}{{\FR 7.11 }
judgeth…: or, is a righteous judge}
\par }{\Q \VS{12}If he
turn not, he will
whet his
sword; he hath
bent his
bow, and made it
ready.
\par }{\Q \VS{13}He hath also
prepared for him the
instruments of
death; he
ordaineth his
arrows against the
persecutors.
\par }{\Q \VS{14}Behold, he
travaileth with
iniquity, and hath
conceived
mischief, and brought
forth
falsehood.
\par }{\Q \VS{15}He
made a
pit, and
digged it, and is
fallen into the
ditch
{\ADD{which}} he
made.\FTNT{**}{{\FR 7.15 }
He made a pit: Heb. He hath digged a pit}
\par }{\Q \VS{16}His
mischief shall
return upon his own
head, and his
violent dealing shall come
down upon his own
pate.
\par }{\Q \VS{17}I will
praise the
{\ND{
LORD}} according to his
righteousness: and will sing
praise to the
name of the
{\ND{
LORD}} most
high.

\par }\PsalmChap{8}{\D To the chief
Musician upon
Gittith, A
Psalm of
David.
\par }{\Q \VerseOne{1}O
{\ND{
LORD}} our
Lord, how
excellent
{\ADD{is}} thy
name in all the
earth! who hast
set thy
glory above the
heavens.
\par }{\Q \VS{2}Out of the
mouth of
babes and
sucklings hast thou
ordained
strength because of thine
enemies, that thou mightest
still the
enemy and the
avenger.\FTNT{*}{{\FR 8.2 }
ordained: Heb. founded}
\par }{\BB \par }{\Q \VS{3}When I
consider thy
heavens, the
work of thy
fingers, the
moon and the
stars, which thou hast
ordained;
\par }{\Q \VS{4}What is
man, that thou art
mindful of him? and the
son of
man, that thou
visitest him?
\par }{\Q \VS{5}For thou hast made him a
little
lower than the
angels, and hast
crowned him with
glory and
honour.
\par }{\Q \VS{6}Thou madest him to have
dominion over the
works of thy
hands; thou hast
put all
{\ADD{things}} under his
feet:
\par }{\Q \VS{7}All
sheep and
oxen, yea, and the
beasts of the
field;\FTNT{†}{{\FR 8.7 }
All…: Heb. Flocks and oxen all of them}
\par }{\Q \VS{8}The
fowl of the
air, and the
fish of the
sea,
{\ADD{and whatsoever}} passeth
through the
paths of the
seas.
\par }{\Q \VS{9}O
{\ND{
LORD}} our
Lord, how
excellent
{\ADD{is}} thy
name in all the
earth!

\par }\PsalmChap{9}{\D To the chief
Musician upon
Muthlabben, A
Psalm of
David.
\par }{\Q \VerseOne{1}I will
praise
{\ADD{thee}}, O
{\ND{
LORD}}, with my whole
heart; I will shew
forth all thy marvellous
works.
\par }{\Q \VS{2}I will be
glad and
rejoice in thee: I will sing
praise to thy
name, O thou most
High.
\par }{\Q \VS{3}When mine
enemies are
turned
back, they shall
fall and
perish at thy
presence.
\par }{\Q \VS{4}For thou hast
maintained my
right and my
cause; thou
satest in the
throne
judging
right.\FTNT{*}{{\FR 9.4 }
maintained…: Heb. made my judgment}\FTNT{†}{{\FR 9.4 }
judging right: Heb. judging in righteousness}
\par }{\Q \VS{5}Thou hast
rebuked the
heathen, thou hast
destroyed the
wicked, thou hast put
out their
name for
ever and
ever.
\par }{\Q \VS{6}O thou
enemy,
destructions are come to a
perpetual
end: and thou hast
destroyed
cities; their
memorial is
perished with
them.\FTNT{‡}{{\FR 9.6 }
O thou…: or, The destructions of the enemy are come to a perpetual end: and their cities hast thou destroyed, etc}
\par }{\Q \VS{7}But the
{\ND{
LORD}} shall
endure for
ever: he hath
prepared his
throne for
judgment.
\par }{\Q \VS{8}And he shall
judge the
world in
righteousness, he shall minister
judgment to the
people in
uprightness.
\par }{\Q \VS{9}The
{\ND{
LORD}} also will be a
refuge for the
oppressed, a
refuge in
times of
trouble.\FTNT{§}{{\FR 9.9 }
a refuge: Heb. an high place}
\par }{\Q \VS{10}And they that
know thy
name will put their
trust in thee: for thou,
{\ND{
LORD}}, hast not
forsaken them that
seek thee.
\par }{\BB \par }{\Q \VS{11}Sing
praises to the
{\ND{
LORD}}, which
dwelleth in
Zion:
declare among the
people his
doings.
\par }{\Q \VS{12}When he maketh
inquisition for
blood, he
remembereth them: he
forgetteth not the
cry of the
humble.\FTNT{**}{{\FR 9.12 }
humble: or, afflicted}
\par }{\Q \VS{13}Have
mercy upon me, O
{\ND{
LORD}};
consider my
trouble
{\ADD{which I suffer}} of them that
hate me, thou that liftest me
up from the
gates of
death:
\par }{\Q \VS{14}That I may shew
forth all thy
praise in the
gates of the
daughter of
Zion: I will
rejoice in thy
salvation.
\par }{\Q \VS{15}The
heathen are sunk
down in the
pit
{\ADD{that}} they
made: in the
net
which they
hid is their own
foot
taken.
\par }{\Q \VS{16}The
{\ND{
LORD}} is
known
{\ADD{by}} the
judgment
{\ADD{which}} he
executeth: the
wicked is
snared in the
work of his own
hands.
Higgaion.
Selah.\FTNT{††}{{\FR 9.16 }
Higgaion: that is, Meditation}
\par }{\Q \VS{17}The
wicked shall be
turned into
hell,
{\ADD{and}} all the
nations that
forget
God.
\par }{\Q \VS{18}For the
needy shall not
alway be
forgotten: the
expectation of the
poor shall
{\ADD{not}}
perish for
ever.
\par }{\Q \VS{19}Arise, O
{\ND{
LORD}}; let not
man
prevail: let the
heathen be
judged in thy
sight.
\par }{\Q \VS{20}Put them in
fear, O
{\ND{
LORD}}:
{\ADD{that}} the
nations may
know themselves
{\ADD{to be but}}
men.
Selah.

\par }\PsalmChap{10}{\Q \VerseOne{1}Why
standest thou afar
off, O
{\ND{
LORD}}?
{\ADD{why}}
hidest thou
{\ADD{thyself}} in
times of
trouble?
\par }{\Q \VS{2}The
wicked in
{\ADD{his}}
pride doth
persecute the
poor: let them be
taken in the
devices
that they have
imagined.\FTNT{*}{{\FR 10.2 }
The wicked…: Heb. In the pride of the wicked he doth persecute}
\par }{\Q \VS{3}For the
wicked
boasteth of his
heart’s
desire, and
blesseth the
covetous,
{\ADD{whom}} the
{\ND{
LORD}}
abhorreth.\FTNT{†}{{\FR 10.3 }
heart’s: Heb. soul’s}\FTNT{‡}{{\FR 10.3 }
blesseth…: or, the covetous blesseth himself, he abhorreth the
{\ND{LORD}}}
\par }{\Q \VS{4}The
wicked, through the
pride of his
countenance, will not
seek
{\ADD{after God:}}
God
{\ADD{is}} not in all his
thoughts.\FTNT{§}{{\FR 10.4 }
God is…: or, all his thoughts are, There is no God}
\par }{\Q \VS{5}His
ways are
always
grievous; thy
judgments
{\ADD{are}} far
above out of his sight:
{\ADD{as for}} all his
enemies, he
puffeth at them.
\par }{\Q \VS{6}He hath
said in his
heart, I shall not be
moved: for
{\ADD{I shall}}
never
{\ADD{be}} in
adversity.\FTNT{**}{{\FR 10.6 }
never: Heb. unto generation and generation}
\par }{\Q \VS{7}His
mouth is
full of
cursing and
deceit and
fraud: under his
tongue
{\ADD{is}}
mischief and
vanity.\FTNT{††}{{\FR 10.7 }
deceit: Heb. deceits}\FTNT{‡‡}{{\FR 10.7 }
vanity: or, iniquity}
\par }{\Q \VS{8}He
sitteth in the lurking
places of the
villages: in the secret
places doth he
murder the
innocent: his
eyes are privily
set against the
poor.\FTNT{§§}{{\FR 10.8 }
are…: Heb. hide themselves}
\par }{\Q \VS{9}He lieth in
wait
secretly as a
lion in his
den: he lieth in
wait to
catch the
poor: he doth
catch the
poor, when he
draweth him into his
net.\FTNT{***}{{\FR 10.9 }
secretly: Heb. in the secret places}
\par }{\Q \VS{10}He
croucheth,
{\ADD{and}}
humbleth himself, that the
poor may
fall by his strong
ones.\FTNT{†††}{{\FR 10.10 }
He…: Heb. He breaketh himself}\FTNT{‡‡‡}{{\FR 10.10 }
by…: or, into his strong parts}
\par }{\Q \VS{11}He hath
said in his
heart,
God hath
forgotten: he
hideth his
face; he will
never
see
{\ADD{it}}.
\par }{\BB \par }{\Q \VS{12}Arise, O
{\ND{
LORD}}; O
God, lift
up thine
hand:
forget not the
humble.\FTNT{§§§}{{\FR 10.12 }
humble: or, afflicted}
\par }{\Q \VS{13}Wherefore doth the
wicked
contemn
God? he hath
said in his
heart, Thou wilt not
require
{\ADD{it}}.
\par }{\Q \VS{14}Thou hast
seen
{\ADD{it;}} for thou
beholdest
mischief and
spite, to
requite
{\ADD{it}} with thy
hand: the
poor
committeth himself unto thee; thou art the
helper of the
fatherless.\FTNT{*}{{\FR 10.14 }
committeth: Heb. leaveth}
\par }{\Q \VS{15}Break thou the
arm of the
wicked and the
evil
{\ADD{man:}} seek
out his
wickedness
{\ADD{till}} thou
find
none.
\par }{\Q \VS{16}The
{\ND{
LORD}}
{\ADD{is}}
King for
ever and
ever: the
heathen are
perished out of his
land.
\par }{\Q \VS{17}{\ND{
LORD}}, thou hast
heard the
desire of the
humble: thou wilt
prepare their
heart, thou wilt cause thine
ear to
hear:\FTNT{†}{{\FR 10.17 }
prepare: or, establish}
\par }{\Q \VS{18}To
judge the
fatherless and the
oppressed, that the
man of the
earth may no
more
oppress.\FTNT{‡}{{\FR 10.18 }
oppress: or, terrify}

\par }\PsalmChap{11}{\D To the chief
Musician,
{\ADD{A Psalm}} of
David.
\par }{\Q \VerseOne{1}In the
{\ND{
LORD}} put I my
trust: how
say ye to my
soul,
Flee
{\ADD{as}} a
bird to your
mountain?
\par }{\Q \VS{2}For, lo, the
wicked
bend
{\ADD{their}}
bow, they make
ready their
arrow upon the
string, that they may
privily
shoot at the
upright in
heart.\FTNT{*}{{\FR 11.2 }
privily: Heb. in darkness}
\par }{\Q \VS{3}If the
foundations be
destroyed, what can the
righteous
do?
\par }{\BB \par }{\Q \VS{4}The
{\ND{
LORD}}
{\ADD{is}} in his
holy
temple, the
{\ND{
LORD’s}}
throne
{\ADD{is}} in
heaven: his
eyes
behold, his
eyelids
try, the
children of
men.
\par }{\Q \VS{5}The
{\ND{
LORD}}
trieth the
righteous: but the
wicked and him that
loveth
violence his
soul
hateth.
\par }{\Q \VS{6}Upon the
wicked he shall
rain
snares,
fire and
brimstone, and an
horrible
tempest:
{\ADD{this shall be}} the
portion of their
cup.\FTNT{†}{{\FR 11.6 }
an horrible…: or, a burning tempest}
\par }{\Q \VS{7}For the
righteous
{\ND{
LORD}}
loveth
righteousness; his
countenance doth
behold the
upright.

\par }\PsalmChap{12}{\D To the chief
Musician upon
Sheminith, A
Psalm of
David.
\par }{\Q \VerseOne{1}Help,
{\ND{
LORD}}; for the godly
man
ceaseth; for the
faithful
fail from among the
children of
men.\FTNT{*}{{\FR 12.1 }
upon…: or, upon the eighth}\FTNT{†}{{\FR 12.1 }
Help: or, Save}
\par }{\Q \VS{2}They
speak
vanity every
one with his
neighbour:
{\ADD{with}}
flattering
lips
{\ADD{and}} with a
double
heart do they
speak.\FTNT{‡}{{\FR 12.2 }
a double…: Heb. an heart and an heart}
\par }{\Q \VS{3}The
{\ND{
LORD}} shall cut
off all
flattering
lips,
{\ADD{and}} the
tongue that
speaketh
proud things:\FTNT{§}{{\FR 12.3 }
proud: Heb. great}
\par }{\Q \VS{4}Who have
said, With our
tongue will we
prevail; our
lips
{\ADD{are}} our own: who
{\ADD{is}}
lord over us?\FTNT{**}{{\FR 12.4 }
are…: Heb. are with us}
\par }{\Q \VS{5}For the
oppression of the
poor, for the
sighing of the
needy, now will I
arise,
saith the
{\ND{
LORD}}; I will
set
{\ADD{him}} in
safety
{\ADD{from him that}}
puffeth at him.\FTNT{††}{{\FR 12.5 }
puffeth…: or, would ensnare him}
\par }{\Q \VS{6}The
words of the
{\ND{
LORD}}
{\ADD{are}}
pure
words:
{\ADD{as}}
silver
tried in a
furnace of
earth,
purified seven
times.
\par }{\Q \VS{7}Thou shalt
keep them, O
{\ND{
LORD}}, thou shalt
preserve them from
this
generation for
ever.\FTNT{‡‡}{{\FR 12.7 }
them from: Heb. him, etc: that is, every one of them, etc}
\par }{\Q \VS{8}The
wicked
walk on every
side, when the
vilest
men are
exalted.\FTNT{§§}{{\FR 12.8 }
men: Heb. of the sons of the men}

\par }\PsalmChap{13}{\D To the chief
Musician, A
Psalm of
David.
\par }{\Q \VerseOne{1}How long wilt thou
forget me, O
{\ND{
LORD}}? for
ever? how long wilt thou
hide thy
face from me?\FTNT{*}{{\FR 13.1 }
chief…: or, overseer}
\par }{\Q \VS{2}How long shall I
take
counsel in my
soul,
{\ADD{having}}
sorrow in my
heart
daily? how long shall mine
enemy be
exalted over me?
\par }{\Q \VS{3}Consider
{\ADD{and}}
hear me, O
{\ND{
LORD}} my
God:
lighten mine
eyes, lest I
sleep the
{\ADD{sleep of}}
death;
\par }{\Q \VS{4}Lest mine
enemy
say, I have
prevailed against him;
{\ADD{and}} those that
trouble me
rejoice when I am
moved.
\par }{\Q \VS{5}But I have
trusted in thy
mercy; my
heart shall
rejoice in thy
salvation.
\par }{\Q \VS{6}I will
sing unto the
{\ND{
LORD}}, because he hath dealt
bountifully with me.

\par }\PsalmChap{14}{\D To the chief
Musician,
{\ADD{A Psalm}} of
David.
\par }{\Q \VerseOne{1}The
fool hath
said in his
heart,
{\ADD{There is}} no
God. They are
corrupt, they have done
abominable
works,
{\ADD{there is}} none that
doeth
good.
\par }{\Q \VS{2}The
{\ND{
LORD}} looked
down from
heaven upon the
children of
men, to
see if
there were any that did
understand,
{\ADD{and}}
seek
God.
\par }{\Q \VS{3}They are all gone
aside, they are
{\ADD{all}}
together become
filthy:
{\ADD{there is}} none that
doeth
good, no, not
one.\FTNT{*}{{\FR 14.3 }
filthy: Heb. stinking}
\par }{\BB \par }{\Q \VS{4}Have all the
workers of
iniquity no
knowledge? who eat
up my
people
{\ADD{as}} they
eat
bread, and
call not upon the
{\ND{
LORD}}.
\par }{\Q \VS{5}There were they in
great
fear: for
God
{\ADD{is}} in the
generation of the
righteous.\FTNT{†}{{\FR 14.5 }
were…: Heb. they feared a fear}
\par }{\Q \VS{6}Ye have
shamed the
counsel of the
poor, because the
{\ND{
LORD}}
{\ADD{is}} his
refuge.
\par }{\Q \VS{7}Oh
that the
salvation of
Israel
{\ADD{were come}} out of
Zion! when the
{\ND{
LORD}} bringeth
back the
captivity of his
people,
Jacob shall
rejoice,
{\ADD{and}}
Israel shall be
glad.\FTNT{‡}{{\FR 14.7 }
Oh…: Heb. Who will give}

\par }\PsalmChap{15}{\D A
Psalm of
David.
\par }{\Q \VerseOne{1}{\ND{
LORD}}, who shall
abide in thy
tabernacle? who shall
dwell in thy
holy
hill?\FTNT{*}{{\FR 15.1 }
abide: Heb. sojourn}
\par }{\Q \VS{2}He that
walketh
uprightly, and
worketh
righteousness, and
speaketh the
truth in his
heart.
\par }{\Q \VS{3}{\ADD{He that}}
backbiteth not with his
tongue, nor
doeth
evil to his
neighbour, nor taketh
up a
reproach against his
neighbour.\FTNT{†}{{\FR 15.3 }
taketh…: or, receiveth, or, endureth}
\par }{\Q \VS{4}In whose
eyes a vile
person is
contemned; but he
honoureth them that
fear the
{\ND{
LORD}}.
{\ADD{He that}}
sweareth to
{\ADD{his own}}
hurt, and
changeth not.
\par }{\Q \VS{5}{\ADD{He that}} putteth not
out his
money to
usury, nor
taketh
reward against the
innocent. He that
doeth these
{\ADD{things}} shall
never be
moved.

\par }\PsalmChap{16}{\D Michtam of
David.
\par }{\Q \VerseOne{1}Preserve me, O
God: for in thee do I put my
trust.\FTNT{*}{{\FR 16.1 }
Michtam: or, A golden Psalm}
\par }{\Q \VS{2}{\ADD{O my soul}}, thou hast
said unto the
{\ND{
LORD}}, Thou
{\ADD{art}} my
Lord: my
goodness
{\ADD{extendeth}} not to thee;
\par }{\Q \VS{3}{\ADD{But}} to the
saints
that
{\ADD{are}} in the
earth, and
{\ADD{to}} the
excellent, in whom
{\ADD{is}} all my
delight.
\par }{\Q \VS{4}Their
sorrows shall be
multiplied
{\ADD{that}}
hasten
{\ADD{after}}
another
{\ADD{god:}} their drink
offerings of
blood will I not
offer,
nor take
up their
names into my
lips.\FTNT{†}{{\FR 16.4 }
hasten…: or, give gifts to another}
\par }{\Q \VS{5}The
{\ND{
LORD}}
{\ADD{is}} the
portion of mine
inheritance and of my
cup: thou
maintainest my
lot.\FTNT{‡}{{\FR 16.5 }
of mine…: Heb. of my part}
\par }{\Q \VS{6}The
lines are
fallen unto me in
pleasant
{\ADD{places;}} yea, I have a
goodly
heritage.
\par }{\Q \VS{7}I will
bless the
{\ND{
LORD}}, who hath given me
counsel: my
reins also
instruct me in the night
seasons.
\par }{\BB \par }{\Q \VS{8}I have
set the
{\ND{
LORD}}
always before me: because
{\ADD{he is}} at my right
hand, I shall not be
moved.
\par }{\Q \VS{9}Therefore my
heart is
glad, and my
glory
rejoiceth: my
flesh also shall
rest in
hope.\FTNT{§}{{\FR 16.9 }
rest…: Heb. dwell confidently}
\par }{\Q \VS{10}For thou wilt not
leave my
soul in
hell; neither wilt thou
suffer thine Holy
One to
see
corruption.
\par }{\Q \VS{11}Thou wilt
shew me the
path of
life: in thy
presence
{\ADD{is}}
fulness of
joy; at thy right
hand
{\ADD{there are}}
pleasures for
evermore.

\par }\PsalmChap{17}{\D A
Prayer of
David.
\par }{\Q \VerseOne{1}Hear the
right, O
{\ND{
LORD}},
attend unto my
cry, give
ear unto my
prayer,
{\ADD{that goeth}}
not out of
feigned
lips.\FTNT{*}{{\FR 17.1 }
the right: Heb. justice}\FTNT{†}{{\FR 17.1 }
not…: Heb. without lips of deceit}
\par }{\Q \VS{2}Let my
sentence come
forth from thy
presence; let thine
eyes
behold the things that are
equal.
\par }{\Q \VS{3}Thou hast
proved mine
heart; thou hast
visited
{\ADD{me}} in the
night; thou hast
tried me,
{\ADD{and}} shalt
find nothing; I am
purposed
{\ADD{that}} my
mouth shall not
transgress.
\par }{\Q \VS{4}Concerning the
works of
men, by the
word of thy
lips I have
kept
{\ADD{me from}} the
paths of the
destroyer.
\par }{\Q \VS{5}Hold
up my
goings in thy
paths,
{\ADD{that}} my
footsteps
slip not.\FTNT{‡}{{\FR 17.5 }
slip…: Heb. be not moved}
\par }{\Q \VS{6}I have
called upon thee, for thou wilt
hear me, O
God:
incline thine
ear unto me,
{\ADD{
and hear}} my
speech.
\par }{\Q \VS{7}Shew thy
marvellous
lovingkindness, O thou that
savest by thy right
hand them which put their
trust
{\ADD{in thee}} from those that rise
up
{\ADD{against them}}.\FTNT{§}{{\FR 17.7 }
by thy…: or, them which trust in thee from those that rise up against thy right hand}
\par }{\BB \par }{\Q \VS{8}Keep me as the
apple of the
eye,
hide me under the
shadow of thy
wings,
\par }{\Q \VS{9}From the
wicked
that
oppress me,
{\ADD{from}} my
deadly
enemies,
{\ADD{who}} compass me
about.\FTNT{**}{{\FR 17.9 }
oppress: Heb. waste}\FTNT{††}{{\FR 17.9 }
my…: Heb. my enemies against the soul}
\par }{\Q \VS{10}They are
inclosed in their own
fat: with their
mouth they
speak
proudly.
\par }{\Q \VS{11}They have now
compassed us in our
steps: they have
set their
eyes bowing
down to the
earth;
\par }{\Q \VS{12}Like as a
lion
{\ADD{that}} is
greedy of his
prey, and as it were a young
lion
lurking in secret
places.\FTNT{‡‡}{{\FR 17.12 }
Like…: Heb. The likeness of him (that is, of every one of them) is as a lion that desireth to ravin}\FTNT{§§}{{\FR 17.12 }
lurking: Heb. sitting}
\par }{\Q \VS{13}Arise, O
{\ND{
LORD}},
disappoint
him, cast him
down:
deliver my
soul from the
wicked,
{\ADD{which is}} thy
sword:\FTNT{***}{{\FR 17.13 }
disappoint…: Heb. prevent his face}\FTNT{†††}{{\FR 17.13 }
which is: or, by}
\par }{\Q \VS{14}From
men
{\ADD{which are}} thy
hand, O
{\ND{
LORD}}, from
men of the
world,
{\ADD{which have}} their
portion in
{\ADD{this}}
life, and whose
belly thou
fillest with thy
hid
{\ADD{treasure:}} they are
full of
children, and
leave the
rest of their
{\ADD{substance}} to their
babes.\FTNT{‡‡‡}{{\FR 17.14 }
which are: or, by}\FTNT{§§§}{{\FR 17.14 }
they…: or, their children are full}
\par }{\Q \VS{15}As for me, I will
behold thy
face in
righteousness: I shall be
satisfied, when I
awake, with thy
likeness.

\par }\PsalmChap{18}{\D To the chief
Musician,
{\ADD{A Psalm}} of
David, the
servant of the
{\ND{
LORD}}, who
spake unto the
{\ND{
LORD}} the
words of this
song in the
day
{\ADD{that}} the
{\ND{
LORD}}
delivered him from the
hand of all his
enemies, and from the
hand of
Saul: And he
said,
\par }{\Q \VerseOne{1}I will
love thee, O
{\ND{
LORD}}, my
strength.
\par }{\Q \VS{2}The
{\ND{
LORD}}
{\ADD{is}} my
rock, and my
fortress, and my
deliverer; my
God, my
strength, in whom I will
trust; my
buckler, and the
horn of my
salvation,
{\ADD{and}} my high
tower.\FTNT{*}{{\FR 18.2 }
my strength: Heb. my rock}
\par }{\Q \VS{3}I will
call upon the
{\ND{
LORD}},
{\ADD{who is worthy}} to be
praised: so shall I be
saved from mine
enemies.
\par }{\Q \VS{4}The
sorrows of
death
compassed me, and the
floods of ungodly
men made me
afraid.\FTNT{†}{{\FR 18.4 }
ungodly men: Heb. Belial}
\par }{\Q \VS{5}The
sorrows of
hell compassed me
about: the
snares of
death
prevented me.\FTNT{‡}{{\FR 18.5 }
sorrows: or, cords}
\par }{\Q \VS{6}In my
distress I
called upon the
{\ND{
LORD}}, and
cried unto my
God: he
heard my
voice out of his
temple, and my
cry
came
before him,
{\ADD{even}} into his
ears.
\par }{\Q \VS{7}Then the
earth
shook and
trembled; the
foundations also of the
hills
moved and were
shaken, because he was
wroth.
\par }{\Q \VS{8}There went
up a
smoke out of his
nostrils, and
fire out of his
mouth
devoured:
coals were
kindled by it.\FTNT{§}{{\FR 18.8 }
out of his nostrils: Heb. by his, etc}
\par }{\Q \VS{9}He
bowed the
heavens also, and came
down: and
darkness
{\ADD{was}} under his
feet.
\par }{\Q \VS{10}And he
rode upon a
cherub, and did
fly: yea, he did
fly upon the
wings of the
wind.
\par }{\Q \VS{11}He
made
darkness his secret
place; his
pavilion round
about him
{\ADD{were}}
dark
waters
{\ADD{and}} thick
clouds of the
skies.
\par }{\Q \VS{12}At the
brightness
{\ADD{that was}} before him his thick
clouds
passed,
hail
{\ADD{stones}} and
coals of
fire.
\par }{\Q \VS{13}The
{\ND{
LORD}} also
thundered in the
heavens, and the
Highest
gave his
voice;
hail
{\ADD{stones}} and
coals of
fire.
\par }{\Q \VS{14}Yea, he sent
out his
arrows, and
scattered them; and he shot
out
lightnings, and
discomfited them.
\par }{\Q \VS{15}Then the
channels of
waters were
seen, and the
foundations of the
world were
discovered at thy
rebuke, O
{\ND{
LORD}}, at the
blast of the
breath of thy
nostrils.
\par }{\Q \VS{16}He
sent from
above, he
took me, he
drew me out of
many
waters.\FTNT{**}{{\FR 18.16 }
many: or, great}
\par }{\Q \VS{17}He
delivered me from my
strong
enemy, and from them which
hated me: for they were too
strong for me.
\par }{\Q \VS{18}They
prevented me in the
day of my
calamity: but the
{\ND{
LORD}} was my
stay.
\par }{\Q \VS{19}He brought me
forth also into a large
place; he
delivered me, because he
delighted in me.
\par }{\BB \par }{\Q \VS{20}The
{\ND{
LORD}}
rewarded me according to my
righteousness; according to the
cleanness of my
hands hath he
recompensed me.
\par }{\Q \VS{21}For I have
kept the
ways of the
{\ND{
LORD}}, and have not wickedly
departed from my
God.
\par }{\Q \VS{22}For all his
judgments
{\ADD{were}} before me, and I did not put
away his
statutes from me.
\par }{\Q \VS{23}I was also
upright before him, and I
kept myself from mine
iniquity.\FTNT{††}{{\FR 18.23 }
before: Heb. with}
\par }{\Q \VS{24}Therefore hath the
{\ND{
LORD}}
recompensed me according to my
righteousness, according to the
cleanness of my
hands in his
eyesight.\FTNT{‡‡}{{\FR 18.24 }
in…: Heb. before his eyes}
\par }{\Q \VS{25}With the
merciful thou wilt shew thyself
merciful; with an
upright
man thou wilt shew thyself
upright;
\par }{\Q \VS{26}With the
pure thou wilt shew thyself
pure; and with the
froward thou wilt shew thyself
froward.\FTNT{§§}{{\FR 18.26 }
shew thyself froward: or, wrestle}
\par }{\Q \VS{27}For thou wilt
save the
afflicted
people; but wilt bring
down
high
looks.
\par }{\Q \VS{28}For thou wilt
light my
candle: the
{\ND{
LORD}} my
God will
enlighten my
darkness.\FTNT{***}{{\FR 18.28 }
candle: or, lamp}
\par }{\BB \par }{\Q \VS{29}For by thee I have run
through a
troop; and by my
God have I leaped
over a
wall.\FTNT{†††}{{\FR 18.29 }
run…: or, broken}
\par }{\Q \VS{30}{\ADD{As for}}
God, his
way
{\ADD{is}}
perfect: the
word of the
{\ND{
LORD}} is
tried: he
{\ADD{is}} a
buckler to all those that
trust in him.\FTNT{‡‡‡}{{\FR 18.30 }
tried: or, refined}
\par }{\Q \VS{31}For who
{\ADD{is}}
God
save the
{\ND{
LORD}}? or who
{\ADD{is}} a
rock
save our
God?
\par }{\Q \VS{32}{\ADD{It is}}
God that
girdeth me with
strength, and
maketh my
way
perfect.
\par }{\Q \VS{33}He
maketh my
feet like
hinds’
{\ADD{feet}}, and
setteth me upon my high
places.
\par }{\Q \VS{34}He
teacheth my
hands to
war, so that a
bow of
steel is
broken by mine
arms.
\par }{\Q \VS{35}Thou hast also
given me the
shield of thy
salvation: and thy right
hand hath holden me
up, and thy
gentleness hath made me
great.\FTNT{§§§}{{\FR 18.35 }
thy gentleness…: or, with thy meekness thou hast multiplied me}
\par }{\Q \VS{36}Thou hast
enlarged my
steps under me, that my
feet did not
slip.\FTNT{*}{{\FR 18.36 }
my feet: Heb. mine ankles}
\par }{\Q \VS{37}I have
pursued mine
enemies, and
overtaken them: neither did I turn
again till they were
consumed.
\par }{\Q \VS{38}I have
wounded them that they were not
able to
rise: they are
fallen under my
feet.
\par }{\Q \VS{39}For thou hast
girded me with
strength unto the
battle: thou hast
subdued under me those that rose
up against me.\FTNT{†}{{\FR 18.39 }
subdued: Heb. caused to bow}
\par }{\Q \VS{40}Thou hast also
given me the
necks of mine
enemies; that I might
destroy them that
hate me.
\par }{\Q \VS{41}They
cried, but
{\ADD{there was}} none to
save
{\ADD{them: even}} unto the
{\ND{
LORD}}, but he
answered them not.
\par }{\Q \VS{42}Then did I
beat them small as the
dust
before the
wind: I did cast them
out as the
dirt in the
streets.
\par }{\Q \VS{43}Thou hast
delivered me from the
strivings of the
people;
{\ADD{and}} thou hast
made me the
head of the
heathen: a
people
{\ADD{whom}} I have not
known shall
serve me.
\par }{\Q \VS{44}As soon as they
hear of me, they shall
obey me: the
strangers shall
submit themselves unto me.\FTNT{‡}{{\FR 18.44 }
As soon…: Heb. At the hearing of the ear}\FTNT{§}{{\FR 18.44 }
submit…: or, yield feigned obedience}\FTNT{**}{{\FR 18.44 }
strangers: Heb. sons of the stranger}
\par }{\Q \VS{45}The
strangers shall fade
away, and be
afraid out of their close
places.\FTNT{††}{{\FR 18.45 }
strangers: Heb. sons of the stranger}
\par }{\Q \VS{46}The
{\ND{
LORD}}
liveth; and
blessed
{\ADD{be}} my
rock; and let the
God of my
salvation be
exalted.
\par }{\Q \VS{47}{\ADD{It is}}
God that
avengeth me, and
subdueth the
people under me.\FTNT{‡‡}{{\FR 18.47 }
avengeth: Heb. giveth avengements for}\FTNT{§§}{{\FR 18.47 }
subdueth: or, destroyeth}
\par }{\Q \VS{48}He
delivereth me from mine
enemies: yea, thou liftest me
up above those that rise
up against me: thou hast
delivered me from the
violent
man.\FTNT{***}{{\FR 18.48 }
violent…: Heb. man of violence}
\par }{\Q \VS{49}Therefore will I give
thanks unto thee, O
{\ND{
LORD}}, among the
heathen, and sing
praises unto thy
name.\FTNT{†††}{{\FR 18.49 }
give…: or, confess}
\par }{\Q \VS{50}Great
deliverance giveth he to his
king; and
sheweth
mercy to his
anointed, to
David, and to his
seed
for
evermore.

\par }\PsalmChap{19}{\D To the chief
Musician, A
Psalm of
David.
\par }{\Q \VerseOne{1}The
heavens
declare the
glory of
God; and the
firmament sheweth
his
handywork.
\par }{\Q \VS{2}Day unto
day
uttereth
speech, and
night unto
night
sheweth
knowledge.
\par }{\Q \VS{3}{\ADD{There is}} no
speech nor
language,
{\ADD{where}} their
voice is not
heard.\FTNT{*}{{\FR 19.3 }
where…: or, without these their voice is heard: Heb. without their voice heard}
\par }{\Q \VS{4}Their
line is gone
out through all the
earth, and their
words to the
end of the
world. In them hath he
set a
tabernacle for the
sun,\FTNT{†}{{\FR 19.4 }
line: or, rule, or, direction}
\par }{\Q \VS{5}Which
{\ADD{is}} as a
bridegroom coming
out of his
chamber,
{\ADD{and}}
rejoiceth as a strong
man to
run a
race.
\par }{\Q \VS{6}His going
forth
{\ADD{is}} from the
end of the
heaven, and his
circuit unto the
ends of it: and there is nothing
hid from the
heat thereof.
\par }{\BB \par }{\Q \VS{7}The
law of the
{\ND{
LORD}}
{\ADD{is}}
perfect,
converting the
soul: the
testimony of the
{\ND{
LORD}}
{\ADD{is}}
sure, making
wise the
simple.\FTNT{‡}{{\FR 19.7 }
law: or, doctrine}\FTNT{§}{{\FR 19.7 }
converting: or, restoring}
\par }{\Q \VS{8}The
statutes of the
{\ND{
LORD}}
{\ADD{are}}
right,
rejoicing the
heart: the
commandment of the
{\ND{
LORD}}
{\ADD{is}}
pure,
enlightening the
eyes.
\par }{\Q \VS{9}The
fear of the
{\ND{
LORD}}
{\ADD{is}}
clean,
enduring for
ever: the
judgments of the
{\ND{
LORD}}
{\ADD{are}}
true
{\ADD{and}}
righteous
altogether.\FTNT{**}{{\FR 19.9 }
true: Heb. truth}
\par }{\Q \VS{10}More to be
desired
{\ADD{are they}} than
gold, yea, than
much fine
gold:
sweeter also than
honey and the
honeycomb.\FTNT{††}{{\FR 19.10 }
the honeycomb: Heb. the dropping of honeycombs}
\par }{\Q \VS{11}Moreover by them is thy
servant
warned:
{\ADD{and}} in
keeping of them
{\ADD{there is}}
great
reward.
\par }{\Q \VS{12}Who can
understand
{\ADD{his}}
errors?
cleanse thou me from
secret
{\ADD{faults}}.
\par }{\Q \VS{13}Keep
back thy
servant also from
presumptuous
{\ADD{sins;}} let them not have
dominion over me: then shall I be
upright, and I shall be
innocent from the
great
transgression.\FTNT{‡‡}{{\FR 19.13 }
the great: or, much}
\par }{\Q \VS{14}Let the
words of my
mouth, and the
meditation of my
heart, be
acceptable in thy
sight, O
{\ND{
LORD}}, my
strength, and my
redeemer.\FTNT{§§}{{\FR 19.14 }
strength: Heb. rock}

\par }\PsalmChap{20}{\D To the chief
Musician, A
Psalm of
David.
\par }{\Q \VerseOne{1}The
{\ND{
LORD}}
hear thee in the
day of
trouble; the
name of the
God of
Jacob
defend thee;\FTNT{*}{{\FR 20.1 }
defend…: Heb. set thee on an high place}
\par }{\Q \VS{2}Send thee
help from the
sanctuary, and
strengthen thee out of
Zion;\FTNT{†}{{\FR 20.2 }
thee help: Heb. thy help}\FTNT{‡}{{\FR 20.2 }
strengthen: Heb. support}
\par }{\Q \VS{3}Remember all thy
offerings, and
accept thy burnt
sacrifice;
Selah.\FTNT{§}{{\FR 20.3 }
accept: Heb. turn to ashes: or, make fat}
\par }{\Q \VS{4}Grant thee according to thine own
heart, and
fulfil all thy
counsel.
\par }{\Q \VS{5}We will
rejoice in thy
salvation, and in the
name of our
God we will set up
{\ADD{our}}
banners: the
{\ND{
LORD}}
fulfil all thy
petitions.
\par }{\BB \par }{\Q \VS{6}Now
know I that the
{\ND{
LORD}}
saveth his
anointed; he will
hear him from his
holy
heaven with the
saving
strength of his right
hand.\FTNT{**}{{\FR 20.6 }
his holy…: Heb. the heaven of his holiness}\FTNT{††}{{\FR 20.6 }
with…: Heb. by the strength of the salvation of}
\par }{\Q \VS{7}Some
{\ADD{trust}} in
chariots, and some in
horses: but we will
remember the
name of the
{\ND{
LORD}} our
God.
\par }{\Q \VS{8}They are brought
down and
fallen: but we are
risen, and stand
upright.
\par }{\Q \VS{9}Save,
{\ND{
LORD}}: let the
king
hear us
when we
call.

\par }\PsalmChap{21}{\D To the chief
Musician, A
Psalm of
David.
\par }{\Q \VerseOne{1}The
king shall
joy in thy
strength, O
{\ND{
LORD}}; and in thy
salvation how
greatly shall he
rejoice!
\par }{\Q \VS{2}Thou hast
given him his
heart’s
desire, and hast not
withholden the
request of his
lips.
Selah.
\par }{\Q \VS{3}For thou
preventest him with the
blessings of
goodness: thou
settest a
crown of pure
gold on his
head.
\par }{\Q \VS{4}He
asked
life of thee,
{\ADD{and}} thou
gavest
{\ADD{it}} him,
{\ADD{even}}
length of
days for
ever and
ever.
\par }{\Q \VS{5}His
glory
{\ADD{is}}
great in thy
salvation:
honour and
majesty hast thou
laid upon him.
\par }{\Q \VS{6}For thou hast
made him most
blessed for
ever: thou hast
made him exceeding
glad with thy
countenance.\FTNT{*}{{\FR 21.6 }
made him most…: Heb. set him to be blessings}\FTNT{†}{{\FR 21.6 }
made him exceeding…: Heb. make him glad with joy}
\par }{\BB \par }{\Q \VS{7}For the
king
trusteth in the
{\ND{
LORD}}, and through the
mercy of the most
High he shall not be
moved.
\par }{\Q \VS{8}Thine
hand shall find
out all thine
enemies: thy right
hand shall find
out those that
hate thee.
\par }{\Q \VS{9}Thou shalt
make them as a
fiery
oven in the
time of thine
anger: the
{\ND{
LORD}} shall swallow them
up in his
wrath, and the
fire shall
devour them.
\par }{\Q \VS{10}Their
fruit shalt thou
destroy from the
earth, and their
seed from among the
children of
men.
\par }{\Q \VS{11}For they
intended
evil against thee: they
imagined a mischievous
device,
{\ADD{which}} they are not
able
{\ADD{to perform}}.
\par }{\Q \VS{12}Therefore shalt thou
make them turn their
back,
{\ADD{when}} thou shalt make
ready
{\ADD{thine arrows}} upon thy
strings against the
face of them.\FTNT{‡}{{\FR 21.12 }
shalt thou…: or, thou shalt set them as a butt}\FTNT{§}{{\FR 21.12 }
back: Heb. shoulder}
\par }{\Q \VS{13}Be thou
exalted,
{\ND{
LORD}}, in thine own
strength:
{\ADD{so}} will we
sing and
praise thy
power.

\par }\PsalmChap{22}{\D To the chief
Musician upon
Aijeleth
Shahar, A
Psalm of
David.
\par }{\Q \VerseOne{1}My
God, my
God, why hast thou
forsaken me?
{\ADD{why art thou so}}
far from
helping me,
{\ADD{and from}} the
words of my
roaring?\FTNT{*}{{\FR 22.1 }
Aijeleth…: or, the hind of the morning}\FTNT{†}{{\FR 22.1 }
helping…: Heb. my salvation}
\par }{\Q \VS{2}O my
God, I
cry in the
daytime, but thou
hearest not; and in the night
season, and am not
silent.\FTNT{‡}{{\FR 22.2 }
am…: Heb. there is no silence to me}
\par }{\Q \VS{3}But thou
{\ADD{art}}
holy,
{\ADD{O thou}} that
inhabitest the
praises of
Israel.
\par }{\Q \VS{4}Our
fathers
trusted in thee: they
trusted, and thou didst
deliver them.
\par }{\Q \VS{5}They
cried unto thee, and were
delivered: they
trusted in thee, and were not
confounded.
\par }{\Q \VS{6}But I
{\ADD{am}} a
worm, and no
man; a
reproach of
men, and
despised of the
people.
\par }{\Q \VS{7}All they that
see me laugh me to
scorn: they shoot
out the
lip, they
shake the
head,
{\ADD{saying}},\FTNT{§}{{\FR 22.7 }
shoot…: Heb. open}
\par }{\Q \VS{8}He
trusted on the
{\ND{
LORD}}
{\ADD{that}} he would
deliver him: let him
deliver him, seeing he
delighted in him.\FTNT{**}{{\FR 22.8 }
He trusted…: Heb. He rolled himself on}\FTNT{††}{{\FR 22.8 }
seeing…: or, if he delight in}
\par }{\Q \VS{9}But thou
{\ADD{art}} he that
took me out of the
womb: thou didst make me
hope
{\ADD{when I was}} upon my
mother’s
breasts.\FTNT{‡‡}{{\FR 22.9 }
didst…: or, kept me in safety}
\par }{\Q \VS{10}I was
cast upon thee from the
womb: thou
{\ADD{art}} my
God from my
mother’s
belly.
\par }{\BB \par }{\Q \VS{11}Be not
far from me; for
trouble
{\ADD{is}}
near; for
{\ADD{there is}} none to
help.\FTNT{§§}{{\FR 22.11 }
none…: Heb. not a helper}
\par }{\Q \VS{12}Many
bulls have
compassed me:
strong
{\ADD{bulls}} of
Bashan have beset me
round.
\par }{\Q \VS{13}They
gaped upon me
{\ADD{with}} their
mouths,
{\ADD{as}} a
ravening and a
roaring
lion.\FTNT{***}{{\FR 22.13 }
gaped…: Heb. opened their mouths against me}
\par }{\Q \VS{14}I am poured
out like
water, and all my
bones are out of
joint: my
heart is like
wax; it is
melted in the
midst of my
bowels.\FTNT{†††}{{\FR 22.14 }
out of…: or, sundered}
\par }{\Q \VS{15}My
strength is dried
up like a
potsherd; and my
tongue
cleaveth to my
jaws; and thou hast
brought me into the
dust of
death.
\par }{\Q \VS{16}For
dogs have
compassed me: the
assembly of the
wicked have
inclosed me: they
pierced my
hands and my
feet.
\par }{\Q \VS{17}I may
tell all my
bones: they
look
{\ADD{and}}
stare upon me.
\par }{\Q \VS{18}They
part my
garments among them, and
cast
lots upon my
vesture.
\par }{\Q \VS{19}But be not thou
far from me, O
{\ND{
LORD}}: O my
strength,
haste thee to
help me.
\par }{\Q \VS{20}Deliver my
soul from the
sword; my
darling from the
power of the
dog.\FTNT{‡‡‡}{{\FR 22.20 }
my darling: Heb. my only one}\FTNT{§§§}{{\FR 22.20 }
power: Heb. hand}
\par }{\Q \VS{21}Save me from the
lion’s
mouth: for thou hast
heard me from the
horns of the
unicorns.
\par }{\BB \par }{\Q \VS{22}I will
declare thy
name unto my
brethren: in the
midst of the
congregation will I
praise thee.
\par }{\Q \VS{23}Ye that
fear the
{\ND{
LORD}},
praise him; all ye the
seed of
Jacob,
glorify him; and
fear him, all ye the
seed of
Israel.
\par }{\Q \VS{24}For he hath not
despised nor
abhorred the
affliction of the
afflicted; neither hath he
hid his
face from him; but when he
cried unto him, he
heard.
\par }{\Q \VS{25}My
praise
{\ADD{shall be}} of thee in the
great
congregation: I will
pay my
vows before them that
fear him.
\par }{\Q \VS{26}The
meek shall
eat and be
satisfied: they shall
praise the
{\ND{
LORD}} that
seek him: your
heart shall
live for
ever.
\par }{\Q \VS{27}All the
ends of the
world shall
remember and
turn unto the
{\ND{
LORD}}: and all the
kindreds of the
nations shall
worship
before thee.
\par }{\Q \VS{28}For the
kingdom
{\ADD{is}} the
{\ND{
LORD’s}}: and he
{\ADD{is}} the
governor among the
nations.
\par }{\Q \VS{29}All
{\ADD{they that be}}
fat upon
earth shall
eat and
worship: all they that go
down to the
dust shall
bow
before him: and none can keep
alive his own
soul.
\par }{\Q \VS{30}A
seed shall
serve him; it shall be
accounted to the
Lord for a
generation.
\par }{\Q \VS{31}They shall
come, and shall
declare his
righteousness unto a
people that shall be
born, that he hath
done
{\ADD{this}}.

\par }\PsalmChap{23}{\D A
Psalm of
David.
\par }{\Q \VerseOne{1}The
{\ND{
LORD}}
{\ADD{is}} my
shepherd; I shall not
want.
\par }{\Q \VS{2}He maketh me to lie
down in
green
pastures: he
leadeth me beside the
still
waters.\FTNT{*}{{\FR 23.2 }
green…: Heb. pastures of tender grass}\FTNT{†}{{\FR 23.2 }
still…: Heb. waters of quietness}
\par }{\Q \VS{3}He
restoreth my
soul: he
leadeth me in the
paths of
righteousness for his
name’s sake.
\par }{\Q \VS{4}Yea, though I
walk through the
valley of the shadow of
death, I will
fear no
evil: for thou
{\ADD{art}} with me; thy
rod and thy
staff they
comfort me.
\par }{\Q \VS{5}Thou
preparest a
table
before me in the presence of mine
enemies: thou
anointest my
head with
oil; my
cup runneth
over.\FTNT{‡}{{\FR 23.5 }
anointest: Heb. makest fat}
\par }{\Q \VS{6}Surely
goodness and
mercy shall
follow me all the
days of my
life: and I will
dwell in the
house of the
{\ND{
LORD}} for
ever.\FTNT{§}{{\FR 23.6 }
for ever: Heb. to length of days}

\par }\PsalmChap{24}{\D A
Psalm of
David.
\par }{\Q \VerseOne{1}The
earth
{\ADD{is}} the
{\ND{
LORD’s}}, and the
fulness thereof; the
world, and they that
dwell therein.
\par }{\Q \VS{2}For he hath
founded it upon the
seas, and
established it upon the
floods.
\par }{\BB \par }{\Q \VS{3}Who shall
ascend into the
hill of the
{\ND{
LORD}}? or who shall
stand in his
holy
place?
\par }{\Q \VS{4}He that hath
clean
hands, and a
pure
heart; who hath not lifted
up his
soul unto
vanity, nor
sworn
deceitfully.\FTNT{*}{{\FR 24.4 }
He…: Heb. The clean of hands}
\par }{\Q \VS{5}He shall
receive the
blessing from the
{\ND{
LORD}}, and
righteousness from the
God of his
salvation.
\par }{\Q \VS{6}This
{\ADD{is}} the
generation of them that
seek him, that
seek thy
face, O
Jacob.
Selah.\FTNT{†}{{\FR 24.6 }
O Jacob: or, O God of Jacob}
\par }{\BB \par }{\Q \VS{7}Lift
up your
heads, O ye
gates; and be ye lift
up, ye
everlasting
doors; and the
King of
glory shall come
in.
\par }{\Q \VS{8}Who
{\ADD{is}} this
King of
glory? The
{\ND{
LORD}}
strong and
mighty, the
{\ND{
LORD}}
mighty in
battle.
\par }{\Q \VS{9}Lift
up your
heads, O ye
gates; even lift
{\ADD{them}}
up, ye
everlasting
doors; and the
King of
glory shall come
in.
\par }{\Q \VS{10}Who is this
King of
glory? The
{\ND{
LORD}} of
hosts, he
{\ADD{is}} the
King of
glory.
Selah.

\par }\PsalmChap{25}{\D {\ADD{A Psalm}} of
David.
\par }{\Q \VerseOne{1}Unto thee, O
{\ND{
LORD}}, do I lift
up my
soul.
\par }{\Q \VS{2}O my
God, I
trust in thee: let me not be
ashamed, let not mine
enemies
triumph over me.
\par }{\Q \VS{3}Yea, let none that
wait on thee be
ashamed: let them be
ashamed which
transgress without
cause.
\par }{\Q \VS{4}Shew me thy
ways, O
{\ND{
LORD}};
teach me thy
paths.
\par }{\Q \VS{5}Lead me in thy
truth, and
teach me: for thou
{\ADD{art}} the
God of my
salvation; on thee do I
wait all the
day.
\par }{\Q \VS{6}Remember, O
{\ND{
LORD}}, thy tender
mercies and thy
lovingkindnesses; for they
{\ADD{have been}} ever of
old.\FTNT{*}{{\FR 25.6 }
tender…: Heb. bowels}
\par }{\Q \VS{7}Remember not the
sins of my
youth, nor my
transgressions: according to thy
mercy
remember thou me for thy
goodness’ sake, O
{\ND{
LORD}}.
\par }{\BB \par }{\Q \VS{8}Good and
upright
{\ADD{is}} the
{\ND{
LORD}}: therefore will he
teach
sinners in the
way.
\par }{\Q \VS{9}The
meek will he
guide in
judgment: and the
meek will he
teach his
way.
\par }{\Q \VS{10}All the
paths of the
{\ND{
LORD}}
{\ADD{are}}
mercy and
truth unto such as
keep his
covenant and his
testimonies.
\par }{\Q \VS{11}For thy
name’s sake, O
{\ND{
LORD}},
pardon mine
iniquity; for it
{\ADD{is}}
great.
\par }{\Q \VS{12}What
man
{\ADD{is}} he that
feareth the
{\ND{
LORD}}? him shall he
teach in the
way
{\ADD{that}} he shall
choose.
\par }{\Q \VS{13}His
soul shall
dwell at
ease; and his
seed shall
inherit the
earth.\FTNT{†}{{\FR 25.13 }
dwell…: Heb. lodge in goodness}
\par }{\Q \VS{14}The
secret of the
{\ND{
LORD}}
{\ADD{is}} with them that
fear him; and he will
shew them his
covenant.\FTNT{‡}{{\FR 25.14 }
and…: or, and his covenant to make them know it}
\par }{\BB \par }{\Q \VS{15}Mine
eyes
{\ADD{are}}
ever toward the
{\ND{
LORD}}; for he shall
pluck my
feet out of the
net.\FTNT{§}{{\FR 25.15 }
pluck: Heb. bring forth}
\par }{\Q \VS{16}Turn thee unto me, and have
mercy upon me; for I
{\ADD{am}}
desolate and
afflicted.
\par }{\Q \VS{17}The
troubles of my
heart are
enlarged:
{\ADD{O}} bring thou me
out of my
distresses.
\par }{\Q \VS{18}Look upon mine
affliction and my
pain; and
forgive all my
sins.
\par }{\Q \VS{19}Consider mine
enemies; for they are
many; and they
hate me with
cruel
hatred.\FTNT{**}{{\FR 25.19 }
cruel…: Heb. hatred of violence}
\par }{\Q \VS{20}O
keep my
soul, and
deliver me: let me not be
ashamed; for I put my
trust in thee.
\par }{\Q \VS{21}Let
integrity and
uprightness
preserve me; for I
wait on thee.
\par }{\Q \VS{22}Redeem
Israel, O
God, out of all his
troubles.

\par }\PsalmChap{26}{\D {\ADD{A Psalm}} of
David.
\par }{\Q \VerseOne{1}Judge me, O
{\ND{
LORD}}; for I have
walked in mine
integrity: I have
trusted also in the
{\ND{
LORD}};
{\ADD{therefore}} I shall not
slide.
\par }{\Q \VS{2}Examine me, O
{\ND{
LORD}}, and
prove me;
try my
reins and my
heart.
\par }{\Q \VS{3}For thy
lovingkindness
{\ADD{is}} before mine
eyes: and I have
walked in thy
truth.
\par }{\Q \VS{4}I have not
sat with
vain
persons, neither will I go
in with
dissemblers.
\par }{\Q \VS{5}I have
hated the
congregation of evil
doers; and will not
sit with the
wicked.
\par }{\BB \par }{\Q \VS{6}I will
wash mine
hands in
innocency: so will I
compass thine
altar, O
{\ND{
LORD}}:
\par }{\Q \VS{7}That I may
publish with the
voice of
thanksgiving, and
tell of all thy wondrous
works.
\par }{\Q \VS{8}{\ND{
LORD}}, I have
loved the
habitation of thy
house, and the
place where thine
honour
dwelleth.\FTNT{*}{{\FR 26.8 }
where…: Heb. of the tabernacle of thy honour}
\par }{\Q \VS{9}Gather not my
soul with
sinners, nor my
life with
bloody
men:\FTNT{†}{{\FR 26.9 }
Gather…: or, Take not away}\FTNT{‡}{{\FR 26.9 }
bloody…: Heb. men of blood}
\par }{\Q \VS{10}In whose
hands
{\ADD{is}}
mischief, and their right
hand is
full of
bribes.\FTNT{§}{{\FR 26.10 }
full…: Heb. filled with}
\par }{\Q \VS{11}But as for me, I will
walk in mine
integrity:
redeem me, and be
merciful unto me.
\par }{\Q \VS{12}My
foot
standeth in an even
place: in the
congregations will I
bless the
{\ND{
LORD}}.

\par }\PsalmChap{27}{\D {\ADD{A Psalm}} of
David.
\par }{\Q \VerseOne{1}The
{\ND{
LORD}}
{\ADD{is}} my
light and my
salvation; whom shall I
fear? the
{\ND{
LORD}}
{\ADD{is}} the
strength of my
life; of whom shall I be
afraid?
\par }{\Q \VS{2}When the
wicked,
{\ADD{even}} mine
enemies and my
foes,
came upon me to eat
up my
flesh, they
stumbled and
fell.\FTNT{*}{{\FR 27.2 }
came…: Heb. approached against me}
\par }{\Q \VS{3}Though an
host should
encamp against me, my
heart shall not
fear: though
war should
rise against me, in this
{\ADD{will}} I
{\ADD{be}}
confident.
\par }{\Q \VS{4}One
{\ADD{thing}} have I
desired of the
{\ND{
LORD}}, that will I
seek after; that I may
dwell in the
house of the
{\ND{
LORD}} all the
days of my
life, to
behold the
beauty of the
{\ND{
LORD}}, and to
enquire in his
temple.\FTNT{†}{{\FR 27.4 }
the beauty: or, the delight}
\par }{\Q \VS{5}For in the
time of
trouble he shall
hide me in his
pavilion: in the
secret of his
tabernacle shall he
hide me; he shall set me
up upon a
rock.
\par }{\Q \VS{6}And now shall mine
head be lifted
up above mine
enemies round
about me: therefore will I
offer in his
tabernacle
sacrifices of
joy; I will
sing, yea, I will sing
praises unto the
{\ND{
LORD}}.\FTNT{‡}{{\FR 27.6 }
joy: Heb. shouting}
\par }{\BB \par }{\Q \VS{7}Hear, O
{\ND{
LORD}},
{\ADD{when}} I
cry with my
voice: have
mercy also upon me, and
answer me.
\par }{\Q \VS{8}{\ADD{When thou saidst}},
Seek ye my
face; my
heart
said unto thee, Thy
face,
{\ND{
LORD}}, will I
seek.\FTNT{§}{{\FR 27.8 }
When…: or, My heart said unto thee, Let my face seek thy face, etc}
\par }{\Q \VS{9}Hide not thy
face
{\ADD{far}} from me;
put not thy
servant
away in
anger: thou hast been my
help;
leave me not, neither
forsake me, O
God of my
salvation.
\par }{\Q \VS{10}When my
father and my
mother
forsake me, then the
{\ND{
LORD}} will take me
up.\FTNT{**}{{\FR 27.10 }
take…: Heb. gather me}
\par }{\Q \VS{11}Teach me thy
way, O
{\ND{
LORD}}, and
lead me in a
plain
path, because of mine
enemies.\FTNT{††}{{\FR 27.11 }
a plain…: Heb. a way of plainness}\FTNT{‡‡}{{\FR 27.11 }
mine…: Heb. those which observe me}
\par }{\Q \VS{12}Deliver me not over unto the
will of mine
enemies: for
false
witnesses are risen
up against me, and such as breathe
out
cruelty.
\par }{\Q \VS{13}{\ADD{I had fainted}},
unless I had
believed to
see the
goodness of the
{\ND{
LORD}} in the
land of the
living.
\par }{\Q \VS{14}Wait on the
{\ND{
LORD}}: be of good
courage, and he shall
strengthen thine
heart:
wait, I say, on the
{\ND{
LORD}}.

\par }\PsalmChap{28}{\D {\ADD{A Psalm}} of
David.
\par }{\Q \VerseOne{1}Unto thee will I
cry, O
{\ND{
LORD}} my
rock; be not
silent to me: lest,
{\ADD{if}} thou be
silent to me, I
become like them that go
down into the
pit.\FTNT{*}{{\FR 28.1 }
to me: Heb. from me}
\par }{\Q \VS{2}Hear the
voice of my
supplications, when I
cry unto thee, when I lift
up my
hands toward thy
holy
oracle.\FTNT{†}{{\FR 28.2 }
thy…: or, the oracle of thy sanctuary}
\par }{\Q \VS{3}Draw me not
away with the
wicked, and with the
workers of
iniquity, which
speak
peace to their
neighbours, but
mischief
{\ADD{is}} in their
hearts.
\par }{\Q \VS{4}Give them according to their
deeds, and according to the
wickedness of their
endeavours:
give them after the
work of their
hands;
render to them their
desert.
\par }{\Q \VS{5}Because they
regard not the
works of the
{\ND{
LORD}}, nor the
operation of his
hands, he shall
destroy them, and not build them
up.
\par }{\BB \par }{\Q \VS{6}Blessed
{\ADD{be}} the
{\ND{
LORD}}, because he hath
heard the
voice of my
supplications.
\par }{\Q \VS{7}The
{\ND{
LORD}}
{\ADD{is}} my
strength and my
shield; my
heart
trusted in him, and I am
helped: therefore my
heart greatly
rejoiceth; and with my
song will I
praise him.
\par }{\Q \VS{8}The
{\ND{
LORD}}
{\ADD{is}} their
strength, and he
{\ADD{is}} the
saving
strength of his
anointed.\FTNT{‡}{{\FR 28.8 }
their: or, his}\FTNT{§}{{\FR 28.8 }
saving…: Heb. strength of salvations}
\par }{\Q \VS{9}Save thy
people, and
bless thine
inheritance:
feed them also, and lift them
up for
ever.\FTNT{**}{{\FR 28.9 }
feed: or, rule}

\par }\PsalmChap{29}{\D A
Psalm of
David.
\par }{\Q \VerseOne{1}Give unto the
{\ND{
LORD}}, O ye
mighty,
give unto the
{\ND{
LORD}}
glory and
strength.\FTNT{*}{{\FR 29.1 }
ye mighty: Heb. ye sons of the mighty}
\par }{\Q \VS{2}Give unto the
{\ND{
LORD}} the
glory due unto his
name;
worship the
{\ND{
LORD}} in the
beauty of
holiness.\FTNT{†}{{\FR 29.2 }
the glory…: Heb. the honour of his name}\FTNT{‡}{{\FR 29.2 }
in…: or, in his glorious sanctuary}
\par }{\Q \VS{3}The
voice of the
{\ND{
LORD}}
{\ADD{is}} upon the
waters: the
God of
glory
thundereth: the
{\ND{
LORD}}
{\ADD{is}} upon
many
waters.\FTNT{§}{{\FR 29.3 }
many: or, great}
\par }{\Q \VS{4}The
voice of the
{\ND{
LORD}}
{\ADD{is}}
powerful; the
voice of the
{\ND{
LORD}}
{\ADD{is}} full of
majesty.\FTNT{**}{{\FR 29.4 }
powerful: Heb. in power}\FTNT{††}{{\FR 29.4 }
full…: Heb. in majesty}
\par }{\Q \VS{5}The
voice of the
{\ND{
LORD}}
breaketh the
cedars; yea, the
{\ND{
LORD}}
breaketh the
cedars of
Lebanon.
\par }{\Q \VS{6}He maketh them also to
skip like a
calf;
Lebanon and
Sirion like a
young
unicorn.
\par }{\Q \VS{7}The
voice of the
{\ND{
LORD}}
divideth the
flames of
fire.\FTNT{‡‡}{{\FR 29.7 }
divideth: Heb. cutteth out}
\par }{\Q \VS{8}The
voice of the
{\ND{
LORD}}
shaketh the
wilderness; the
{\ND{
LORD}}
shaketh the
wilderness of
Kadesh.
\par }{\Q \VS{9}The
voice of the
{\ND{
LORD}} maketh the
hinds to
calve, and
discovereth the
forests: and in his
temple doth every one
speak of
{\ADD{his}}
glory.\FTNT{§§}{{\FR 29.9 }
to calve: or, to be in pain}\FTNT{***}{{\FR 29.9 }
doth…: or, every whit of it uttereth, etc}
\par }{\Q \VS{10}The
{\ND{
LORD}}
sitteth upon the
flood; yea, the
{\ND{
LORD}}
sitteth
King for
ever.
\par }{\Q \VS{11}The
{\ND{
LORD}} will
give
strength unto his
people; the
{\ND{
LORD}} will
bless his
people with
peace.

\par }\PsalmChap{30}{\D A
Psalm
{\ADD{and}}
Song
{\ADD{at}} the
dedication of the
house of
David.
\par }{\Q \VerseOne{1}I will
extol thee, O
{\ND{
LORD}}; for thou hast lifted me
up, and hast not made my
foes to
rejoice over me.
\par }{\Q \VS{2}O
{\ND{
LORD}} my
God, I
cried unto thee, and thou hast
healed me.
\par }{\Q \VS{3}O
{\ND{
LORD}}, thou hast brought
up my
soul from the
grave: thou hast kept me
alive, that I should not go
down to the
pit.
\par }{\Q \VS{4}Sing unto the
{\ND{
LORD}}, O ye
saints of his, and give
thanks at the
remembrance of his
holiness.\FTNT{*}{{\FR 30.4 }
at…: or, to the memorial}
\par }{\Q \VS{5}For his
anger
{\ADD{endureth but}} a
moment; in his
favour
{\ADD{is}}
life:
weeping may
endure for a
night, but
joy
{\ADD{cometh}} in the
morning.\FTNT{†}{{\FR 30.5 }
his anger…: Heb. there is but a moment in his anger}\FTNT{‡}{{\FR 30.5 }
for a night: Heb. in the evening}\FTNT{§}{{\FR 30.5 }
joy: Heb. singing}
\par }{\BB \par }{\Q \VS{6}And in my
prosperity I
said, I shall
never be
moved.
\par }{\Q \VS{7}{\ND{
LORD}}, by thy
favour thou hast made my
mountain to
stand
strong: thou didst
hide thy
face,
{\ADD{and}} I was
troubled.\FTNT{**}{{\FR 30.7 }
made…: Heb. settled strength for my mountain}
\par }{\Q \VS{8}I
cried to thee, O
{\ND{
LORD}}; and unto the
{\ND{
LORD}} I made
supplication.
\par }{\Q \VS{9}What
profit
{\ADD{is there}} in my
blood, when I go
down to the
pit? Shall the
dust
praise thee? shall it
declare thy
truth?
\par }{\Q \VS{10}Hear, O
{\ND{
LORD}}, and have
mercy upon me:
{\ND{
LORD}}, be thou my
helper.
\par }{\Q \VS{11}Thou hast
turned for me my
mourning into
dancing: thou hast put
off my
sackcloth, and
girded me with
gladness;
\par }{\Q \VS{12}To the end that
{\ADD{my}}
glory may
sing praise to thee, and not be
silent. O
{\ND{
LORD}} my
God, I will give
thanks unto thee for
ever.\FTNT{††}{{\FR 30.12 }
my glory: that is, my tongue, or, my soul}

\par }\PsalmChap{31}{\D To the chief
Musician, A
Psalm of
David.
\par }{\Q \VerseOne{1}In thee, O
{\ND{
LORD}}, do I put my
trust; let me
never be
ashamed:
deliver me in thy
righteousness.
\par }{\Q \VS{2}Bow
down thine
ear to me;
deliver me
speedily: be thou my
strong
rock, for an
house of
defence to
save me.\FTNT{*}{{\FR 31.2 }
my…: Heb. to me for a rock of strength}
\par }{\Q \VS{3}For thou
{\ADD{art}} my
rock and my
fortress; therefore for thy
name’s sake
lead me, and
guide me.
\par }{\Q \VS{4}Pull me
out of the
net
that they have laid
privily for me: for thou
{\ADD{art}} my
strength.
\par }{\Q \VS{5}Into thine
hand I
commit my
spirit: thou hast
redeemed me, O
{\ND{
LORD}}
God of
truth.
\par }{\Q \VS{6}I have
hated them that
regard
lying
vanities: but I
trust in the
{\ND{
LORD}}.
\par }{\Q \VS{7}I will be
glad and
rejoice in thy
mercy: for thou hast
considered my
trouble; thou hast
known my
soul in
adversities;
\par }{\Q \VS{8}And hast not shut me
up into the
hand of the
enemy: thou hast
set my
feet in a large
room.
\par }{\BB \par }{\Q \VS{9}Have
mercy upon me, O
{\ND{
LORD}}, for I am in
trouble: mine
eye is
consumed with
grief,
{\ADD{yea}}, my
soul and my
belly.
\par }{\Q \VS{10}For my
life is
spent with
grief, and my
years with
sighing: my
strength
faileth because of mine
iniquity, and my
bones are
consumed.
\par }{\Q \VS{11}I was a
reproach among all mine
enemies, but
especially among my
neighbours, and a
fear to mine
acquaintance: they that did
see me
without
fled from me.
\par }{\Q \VS{12}I am
forgotten as a dead
man out of
mind: I am like a
broken
vessel.\FTNT{†}{{\FR 31.12 }
a broken…: Heb. a vessel that perisheth}
\par }{\Q \VS{13}For I have
heard the
slander of
many:
fear
{\ADD{was}} on every
side: while they took
counsel
together against me, they
devised to take
away my
life.
\par }{\Q \VS{14}But I
trusted in thee, O
{\ND{
LORD}}: I
said, Thou
{\ADD{art}} my
God.
\par }{\Q \VS{15}My
times
{\ADD{are}} in thy
hand:
deliver me from the
hand of mine
enemies, and from them that
persecute me.
\par }{\Q \VS{16}Make thy
face to
shine upon thy
servant:
save me for thy
mercies’ sake.
\par }{\Q \VS{17}Let me not be
ashamed, O
{\ND{
LORD}}; for I have
called upon thee: let the
wicked be
ashamed,
{\ADD{and}} let them be
silent in the
grave.\FTNT{‡}{{\FR 31.17 }
silent…: or, cut off for}
\par }{\Q \VS{18}Let the
lying
lips be put to
silence; which
speak
grievous things
proudly and
contemptuously against the
righteous.\FTNT{§}{{\FR 31.18 }
grievous…: Heb. a hard thing}
\par }{\BB \par }{\Q \VS{19}{\ADD{Oh}} how
great
{\ADD{is}} thy
goodness, which thou hast laid
up for them that
fear thee;
{\ADD{which}} thou hast
wrought for them that
trust in thee before the
sons of
men!
\par }{\Q \VS{20}Thou shalt
hide them in the
secret of thy
presence from the
pride of
man: thou shalt keep them
secretly in a
pavilion from the
strife of
tongues.
\par }{\Q \VS{21}Blessed
{\ADD{be}} the
{\ND{
LORD}}: for he hath shewed me his
marvellous
kindness in a
strong
city.\FTNT{**}{{\FR 31.21 }
strong: or, fenced}
\par }{\Q \VS{22}For I
said in my
haste, I am cut
off from
before thine
eyes:
nevertheless thou
heardest the
voice of my
supplications when I
cried unto thee.
\par }{\Q \VS{23}O
love the
{\ND{
LORD}}, all ye his
saints:
{\ADD{for}} the
{\ND{
LORD}}
preserveth the
faithful, and
plentifully
rewardeth the
proud
doer.
\par }{\Q \VS{24}Be of good
courage, and he shall
strengthen your
heart, all ye that
hope in the
{\ND{
LORD}}.

\par }\PsalmChap{32}{\D {\ADD{A Psalm}} of
David,
Maschil.
\par }{\Q \VerseOne{1}Blessed
{\ADD{is he whose}}
transgression
{\ADD{is}}
forgiven,
{\ADD{whose}}
sin
{\ADD{is}}
covered.\FTNT{*}{{\FR 32.1 }
A Psalm…: or, A Psalm of David giving instruction}
\par }{\Q \VS{2}Blessed
{\ADD{is}} the
man unto whom the
{\ND{
LORD}}
imputeth not
iniquity, and in whose
spirit
{\ADD{there is}} no
guile.
\par }{\Q \VS{3}When I kept
silence, my
bones waxed
old through my
roaring all the
day long.
\par }{\Q \VS{4}For
day and
night thy
hand was
heavy upon me: my
moisture is
turned into the
drought of
summer.
Selah.
\par }{\Q \VS{5}I
acknowledged my
sin unto thee, and mine
iniquity have I not
hid. I
said, I will
confess my
transgressions unto the
{\ND{
LORD}}; and thou
forgavest the
iniquity of my
sin.
Selah.
\par }{\Q \VS{6}For this shall every one that is
godly
pray unto thee in a
time when thou mayest be
found: surely in the
floods of
great
waters they shall not come
nigh unto him.\FTNT{†}{{\FR 32.6 }
when…: Heb. of finding}
\par }{\BB \par }{\Q \VS{7}Thou
{\ADD{art}} my hiding
place; thou shalt
preserve me from
trouble; thou shalt
compass me about with
songs of
deliverance.
Selah.
\par }{\Q \VS{8}I will
instruct thee and
teach thee in the
way
which thou shalt
go: I will
guide thee with mine
eye.\FTNT{‡}{{\FR 32.8 }
guide…: Heb. counsel thee, mine eye shall be upon thee}
\par }{\Q \VS{9}Be ye not as the
horse,
{\ADD{or}} as the
mule,
{\ADD{which}} have no
understanding: whose
mouth must be
held in with
bit and
bridle,
lest they come
near unto thee.
\par }{\Q \VS{10}Many
sorrows
{\ADD{shall be}} to the
wicked: but he that
trusteth in the
{\ND{
LORD}},
mercy shall
compass him about.
\par }{\Q \VS{11}Be
glad in the
{\ND{
LORD}}, and
rejoice, ye
righteous: and
shout for joy, all
{\ADD{ye that are}}
upright in
heart.

\par }\PsalmChap{33}{\Q \VerseOne{1}Rejoice in the
{\ND{
LORD}}, O ye
righteous:
{\ADD{for}}
praise is
comely for the
upright.
\par }{\Q \VS{2}Praise the
{\ND{
LORD}} with
harp:
sing unto him with the
psaltery
{\ADD{and}} an instrument of ten
strings.
\par }{\Q \VS{3}Sing unto him a
new
song;
play
skilfully with a loud
noise.
\par }{\Q \VS{4}For the
word of the
{\ND{
LORD}}
{\ADD{is}}
right; and all his
works
{\ADD{are done}} in
truth.
\par }{\Q \VS{5}He
loveth
righteousness and
judgment: the
earth is
full of the
goodness of the
{\ND{
LORD}}.\FTNT{*}{{\FR 33.5 }
goodness: or, mercy}
\par }{\Q \VS{6}By the
word of the
{\ND{
LORD}} were the
heavens
made; and all the
host of them by the
breath of his
mouth.
\par }{\Q \VS{7}He
gathereth the
waters of the
sea together as an
heap: he layeth
up the
depth in
storehouses.
\par }{\Q \VS{8}Let all the
earth
fear the
{\ND{
LORD}}: let all the
inhabitants of the
world stand in
awe of him.
\par }{\Q \VS{9}For he
spake, and it was
{\ADD{done;}} he
commanded, and it stood
fast.
\par }{\Q \VS{10}The
{\ND{
LORD}}
bringeth the
counsel of the
heathen to
nought: he
maketh the
devices of the
people of none
effect.\FTNT{†}{{\FR 33.10 }
bringeth: Heb. maketh frustrate}
\par }{\Q \VS{11}The
counsel of the
{\ND{
LORD}}
standeth for
ever, the
thoughts of his
heart to
all
generations.\FTNT{‡}{{\FR 33.11 }
to all…: Heb. to generation and generation}
\par }{\BB \par }{\Q \VS{12}Blessed
{\ADD{is}} the
nation whose
God
{\ADD{is}} the
{\ND{
LORD}};
{\ADD{and}} the
people
{\ADD{whom}} he hath
chosen for his own
inheritance.
\par }{\Q \VS{13}The
{\ND{
LORD}}
looketh from
heaven; he
beholdeth all the
sons of
men.
\par }{\Q \VS{14}From the
place of his
habitation he
looketh upon all the
inhabitants of the
earth.
\par }{\Q \VS{15}He
fashioneth their
hearts
alike; he
considereth all their
works.
\par }{\Q \VS{16}There is no
king
saved by the
multitude of an
host: a mighty
man is not
delivered by
much
strength.
\par }{\Q \VS{17}An
horse
{\ADD{is}} a vain
thing for
safety: neither shall he
deliver
{\ADD{any}} by his
great
strength.
\par }{\Q \VS{18}Behold, the
eye of the
{\ND{
LORD}}
{\ADD{is}} upon them that
fear him, upon them that
hope in his
mercy;
\par }{\Q \VS{19}To
deliver their
soul from
death, and to keep them
alive in
famine.
\par }{\Q \VS{20}Our
soul
waiteth for the
{\ND{
LORD}}: he
{\ADD{is}} our
help and our
shield.
\par }{\Q \VS{21}For our
heart shall
rejoice in him, because we have
trusted in his
holy
name.
\par }{\Q \VS{22}Let thy
mercy, O
{\ND{
LORD}}, be upon us, according as we
hope in thee.

\par }\PsalmChap{34}{\D {\ADD{A Psalm}} of
David, when he
changed his
behaviour
before
Abimelech; who drove him
away, and he
departed.
\par }{\Q \VerseOne{1}I will
bless the
{\ND{
LORD}} at all
times: his
praise
{\ADD{shall}}
continually
{\ADD{be}} in my
mouth.\FTNT{*}{{\FR 34.1 }
Abimelech: or, Achish}
\par }{\Q \VS{2}My
soul shall make her
boast in the
{\ND{
LORD}}: the
humble shall
hear
{\ADD{thereof}}, and be
glad.
\par }{\Q \VS{3}O
magnify the
{\ND{
LORD}} with me, and let us
exalt his
name
together.
\par }{\Q \VS{4}I
sought the
{\ND{
LORD}}, and he
heard me, and
delivered me from all my
fears.
\par }{\Q \VS{5}They
looked unto him, and were
lightened: and their
faces were not
ashamed.\FTNT{†}{{\FR 34.5 }
were lightened: or, they flowed unto him}
\par }{\Q \VS{6}This poor
man
cried, and the
{\ND{
LORD}}
heard
{\ADD{him}}, and
saved him out of all his
troubles.
\par }{\Q \VS{7}The
angel of the
{\ND{
LORD}}
encampeth round
about them that
fear him, and
delivereth them.
\par }{\Q \VS{8}O
taste and
see that the
{\ND{
LORD}}
{\ADD{is}}
good:
blessed
{\ADD{is}} the
man
{\ADD{that}}
trusteth in him.
\par }{\Q \VS{9}O
fear the
{\ND{
LORD}}, ye his
saints: for
{\ADD{there is}} no
want to them that
fear him.
\par }{\Q \VS{10}The young
lions do
lack, and suffer
hunger: but they that
seek the
{\ND{
LORD}} shall not
want any
good
{\ADD{thing}}.
\par }{\BB \par }{\Q \VS{11}Come, ye
children,
hearken unto me: I will
teach you the
fear of the
{\ND{
LORD}}.
\par }{\Q \VS{12}What
man
{\ADD{is he that}}
desireth
life,
{\ADD{and}}
loveth
{\ADD{many}}
days, that he may
see
good?
\par }{\Q \VS{13}Keep thy
tongue from
evil, and thy
lips from
speaking
guile.
\par }{\Q \VS{14}Depart from
evil, and
do
good;
seek
peace, and
pursue it.
\par }{\Q \VS{15}The
eyes of the
{\ND{
LORD}}
{\ADD{are}} upon the
righteous, and his
ears
{\ADD{are open}} unto their
cry.
\par }{\Q \VS{16}The
face of the
{\ND{
LORD}}
{\ADD{is}} against them that
do
evil, to cut
off the
remembrance of them from the
earth.
\par }{\Q \VS{17}{\ADD{The righteous}}
cry, and the
{\ND{
LORD}}
heareth, and
delivereth them out of all their
troubles.
\par }{\Q \VS{18}The
{\ND{
LORD}}
{\ADD{is}}
nigh unto them that are of a
broken
heart; and
saveth such as be of a
contrite
spirit.\FTNT{‡}{{\FR 34.18 }
unto…: Heb. to the broken of heart}\FTNT{§}{{\FR 34.18 }
of a contrite…: Heb. contrite of spirit}
\par }{\Q \VS{19}Many
{\ADD{are}} the
afflictions of the
righteous: but the
{\ND{
LORD}}
delivereth him out of them all.
\par }{\Q \VS{20}He
keepeth all his
bones: not
one of
them is
broken.
\par }{\Q \VS{21}Evil shall
slay the
wicked: and they that
hate the
righteous shall be
desolate.\FTNT{**}{{\FR 34.21 }
desolate: or, guilty}
\par }{\Q \VS{22}The
{\ND{
LORD}}
redeemeth the
soul of his
servants: and none of them that
trust in him shall be
desolate.\FTNT{††}{{\FR 34.22 }
desolate: or, guilty}

\par }\PsalmChap{35}{\D {\ADD{A Psalm}} of
David.
\par }{\Q \VerseOne{1}Plead
{\ADD{my cause}}, O
{\ND{
LORD}}, with them that
strive with me:
fight against them that
fight against me.
\par }{\Q \VS{2}Take
hold of
shield and
buckler, and stand
up for mine
help.
\par }{\Q \VS{3}Draw
out also the
spear, and
stop
{\ADD{the way}}
against them that
persecute me:
say unto my
soul, I
{\ADD{am}} thy
salvation.
\par }{\Q \VS{4}Let them be
confounded and put to
shame that
seek after my
soul: let them be
turned
back and brought to
confusion that
devise my
hurt.
\par }{\Q \VS{5}Let them be as
chaff
before the
wind: and let the
angel of the
{\ND{
LORD}}
chase
{\ADD{them}}.
\par }{\Q \VS{6}Let their
way be
dark and
slippery: and let the
angel of the
{\ND{
LORD}}
persecute them.\FTNT{*}{{\FR 35.6 }
dark…: Heb. darkness and slipperiness}
\par }{\Q \VS{7}For without
cause have they
hid for me their
net
{\ADD{in}} a
pit,
{\ADD{which}} without
cause they have
digged for my
soul.
\par }{\Q \VS{8}Let
destruction
come upon him at
unawares; and let his
net that he hath
hid
catch himself: into that very
destruction let him
fall.\FTNT{†}{{\FR 35.8 }
at…: Heb. which he knoweth not of}
\par }{\Q \VS{9}And my
soul shall be
joyful in the
{\ND{
LORD}}: it shall
rejoice in his
salvation.
\par }{\Q \VS{10}All my
bones shall
say,
{\ND{
LORD}}, who
{\ADD{is}} like unto thee, which
deliverest the
poor from him that is too
strong for him, yea, the
poor and the
needy from him that
spoileth him?
\par }{\BB \par }{\Q \VS{11}False
witnesses did rise
up; they laid to my
charge
{\ADD{things}} that I
knew not.\FTNT{‡}{{\FR 35.11 }
False…: Heb. Witnesses of wrong}\FTNT{§}{{\FR 35.11 }
they…: Heb. they asked me}
\par }{\Q \VS{12}They
rewarded me
evil for
good
{\ADD{to}} the
spoiling of my
soul.\FTNT{**}{{\FR 35.12 }
spoiling: Heb. depriving}
\par }{\Q \VS{13}But as for me, when they were
sick, my
clothing
{\ADD{was}}
sackcloth: I
humbled my
soul with
fasting; and my
prayer
returned into mine own
bosom.\FTNT{††}{{\FR 35.13 }
humbled: or, afflicted}
\par }{\Q \VS{14}I
behaved myself as though
{\ADD{he had been}} my
friend
{\ADD{or}}
brother: I bowed
down
heavily, as one that
mourneth
{\ADD{for his}}
mother.\FTNT{‡‡}{{\FR 35.14 }
behaved myself: Heb. walked}\FTNT{§§}{{\FR 35.14 }
as though…: Heb. as a friend, as a brother to me}
\par }{\Q \VS{15}But in mine
adversity they
rejoiced, and gathered themselves
together:
{\ADD{yea}}, the
abjects gathered themselves
together against me, and I
knew
{\ADD{it}} not; they did
tear
{\ADD{me}}, and
ceased not:\FTNT{***}{{\FR 35.15 }
adversity: Heb. halting}
\par }{\Q \VS{16}With
hypocritical
mockers in
feasts, they
gnashed upon me with their
teeth.
\par }{\BB \par }{\Q \VS{17}Lord, how long wilt thou look
on?
rescue my
soul from their
destructions, my
darling from the
lions.\FTNT{†††}{{\FR 35.17 }
darling: Heb. only one}
\par }{\Q \VS{18}I will give thee
thanks in the
great
congregation: I will
praise thee among
much
people.\FTNT{‡‡‡}{{\FR 35.18 }
much: Heb. strong}
\par }{\Q \VS{19}Let not them that are mine
enemies
wrongfully
rejoice over me:
{\ADD{neither}} let them
wink with the
eye that
hate me without a
cause.\FTNT{§§§}{{\FR 35.19 }
wrongfully: Heb. falsely}
\par }{\Q \VS{20}For they
speak not
peace: but they
devise
deceitful
matters against
{\ADD{them that are}}
quiet in the
land.
\par }{\Q \VS{21}Yea, they opened their
mouth
wide against me,
{\ADD{and}}
said,
Aha,
aha, our
eye hath
seen
{\ADD{it}}.
\par }{\Q \VS{22}{\ADD{This}} thou hast
seen, O
{\ND{
LORD}}: keep not
silence: O
Lord, be not
far from me.
\par }{\Q \VS{23}Stir
up thyself, and
awake to my
judgment,
{\ADD{even}} unto my
cause, my
God and my
Lord.
\par }{\Q \VS{24}Judge me, O
{\ND{
LORD}} my
God, according to thy
righteousness; and let them not
rejoice over me.
\par }{\Q \VS{25}Let them not
say in their
hearts,
Ah, so would we have
it: let them not
say, We have swallowed him
up.\FTNT{*}{{\FR 35.25 }
Ah…: Heb. Ah, ah, our soul}
\par }{\Q \VS{26}Let them be
ashamed and brought to
confusion
together that
rejoice at mine
hurt: let them be
clothed with
shame and
dishonour that
magnify
{\ADD{themselves}} against me.
\par }{\Q \VS{27}Let them shout for
joy, and be
glad, that
favour my righteous
cause: yea, let them
say
continually, Let the
{\ND{
LORD}} be
magnified, which hath
pleasure in the
prosperity of his
servant.\FTNT{†}{{\FR 35.27 }
righteous…: Heb. righteousness}
\par }{\Q \VS{28}And my
tongue shall
speak of thy
righteousness
{\ADD{and}} of thy
praise all the
day long.

\par }\PsalmChap{36}{\D To the chief
Musician,
{\ADD{A Psalm}} of
David the
servant of the
{\ND{
LORD}}.
\par }{\Q \VerseOne{1}The
transgression of the
wicked
saith
within my
heart,
{\ADD{that there is}} no
fear of
God before his
eyes.
\par }{\Q \VS{2}For he
flattereth himself in his own
eyes, until his
iniquity be
found to be
hateful.\FTNT{*}{{\FR 36.2 }
until…: Heb. to find his iniquity to hate}
\par }{\Q \VS{3}The
words of his
mouth
{\ADD{are}}
iniquity and
deceit: he hath left
off to be
wise,
{\ADD{and}} to do
good.
\par }{\Q \VS{4}He
deviseth
mischief upon his
bed; he
setteth himself in a
way
{\ADD{that is}} not
good; he
abhorreth not
evil.\FTNT{†}{{\FR 36.4 }
mischief: or, vanity}
\par }{\BB \par }{\Q \VS{5}Thy
mercy, O
{\ND{
LORD}},
{\ADD{is}} in the
heavens;
{\ADD{and}} thy
faithfulness
{\ADD{reacheth}} unto the
clouds.
\par }{\Q \VS{6}Thy
righteousness
{\ADD{is}} like the
great
mountains; thy
judgments
{\ADD{are}} a
great
deep: O
{\ND{
LORD}}, thou
preservest
man and
beast.\FTNT{‡}{{\FR 36.6 }
the…: Heb. the mountains of God}
\par }{\Q \VS{7}How
excellent
{\ADD{is}} thy
lovingkindness, O
God! therefore the
children of
men put their
trust under the
shadow of thy
wings.\FTNT{§}{{\FR 36.7 }
excellent: Heb. precious}
\par }{\Q \VS{8}They shall be abundantly
satisfied with the
fatness of thy
house; and thou shalt make them
drink of the
river of thy
pleasures.\FTNT{**}{{\FR 36.8 }
abundantly…: Heb. watered}
\par }{\Q \VS{9}For with thee
{\ADD{is}} the
fountain of
life: in thy
light shall we
see
light.
\par }{\Q \VS{10}O
continue thy
lovingkindness unto them that
know thee; and thy
righteousness to the
upright in
heart.\FTNT{††}{{\FR 36.10 }
continue: Heb. draw out at length}
\par }{\Q \VS{11}Let not the
foot of
pride
come against me, and let not the
hand of the
wicked
remove me.
\par }{\Q \VS{12}There are the
workers of
iniquity
fallen: they are cast
down, and shall not be
able to
rise.

\par }\PsalmChap{37}{\D {\ADD{A Psalm}} of
David.
\par }{\Q \VerseOne{1}Fret not thyself because of
evildoers, neither be thou
envious against the
workers of
iniquity.
\par }{\Q \VS{2}For they shall
soon be cut
down like the
grass, and
wither as the
green
herb.
\par }{\Q \VS{3}Trust in the
{\ND{
LORD}}, and
do
good;
{\ADD{so}} shalt thou
dwell in the
land, and
verily thou shalt be
fed.\FTNT{*}{{\FR 37.3 }
verily: Heb. in truth, or, stableness}
\par }{\Q \VS{4}Delight thyself also in the
{\ND{
LORD}}; and he shall
give thee the
desires of thine
heart.
\par }{\Q \VS{5}Commit thy
way unto the
{\ND{
LORD}};
trust also in him; and he shall bring
{\ADD{it}} to
pass.\FTNT{†}{{\FR 37.5 }
Commit…: Heb. Roll thy way upon}
\par }{\Q \VS{6}And he shall bring
forth thy
righteousness as the
light, and thy
judgment as the
noonday.
\par }{\BB \par }{\Q \VS{7}Rest in the
{\ND{
LORD}}, and wait
patiently for him:
fret not thyself because of him who
prospereth in his
way, because of the
man who
bringeth wicked
devices to
pass.\FTNT{‡}{{\FR 37.7 }
Rest in: Heb. Be silent to}
\par }{\Q \VS{8}Cease from
anger, and
forsake
wrath:
fret not thyself in any
wise to do
evil.
\par }{\Q \VS{9}For
evildoers shall be cut
off: but those that
wait upon the
{\ND{
LORD}}, they shall
inherit the
earth.
\par }{\Q \VS{10}For yet a little
while, and the
wicked
{\ADD{shall}} not
{\ADD{be:}} yea, thou shalt diligently
consider his
place, and it
{\ADD{shall}} not
{\ADD{be}}.
\par }{\Q \VS{11}But the
meek shall
inherit the
earth; and shall
delight themselves in the
abundance of
peace.
\par }{\Q \VS{12}The
wicked
plotteth against the
just, and
gnasheth upon him with his
teeth.\FTNT{§}{{\FR 37.12 }
plotteth: or, practiseth}
\par }{\Q \VS{13}The
Lord shall
laugh at him: for he
seeth that his
day is
coming.
\par }{\Q \VS{14}The
wicked have drawn
out the
sword, and have
bent their
bow, to cast
down the
poor and
needy,
{\ADD{and}} to
slay such as be of
upright
conversation.\FTNT{**}{{\FR 37.14 }
such…: Heb. the upright of way}
\par }{\Q \VS{15}Their
sword shall
enter into their own
heart, and their
bows shall be
broken.
\par }{\Q \VS{16}A
little that a righteous
man hath
{\ADD{is}}
better than the
riches of
many
wicked.
\par }{\Q \VS{17}For the
arms of the
wicked shall be
broken: but the
{\ND{
LORD}}
upholdeth the
righteous.
\par }{\Q \VS{18}The
{\ND{
LORD}}
knoweth the
days of the
upright: and their
inheritance shall be for
ever.
\par }{\Q \VS{19}They shall not be
ashamed in the
evil
time: and in the
days of
famine they shall be
satisfied.
\par }{\Q \VS{20}But the
wicked shall
perish, and the
enemies of the
{\ND{
LORD}}
{\ADD{shall be}} as the
fat of
lambs: they shall
consume; into
smoke shall they consume
away.\FTNT{††}{{\FR 37.20 }
the fat: Heb. the preciousness}
\par }{\BB \par }{\Q \VS{21}The
wicked
borroweth, and payeth not
again: but the
righteous sheweth
mercy, and
giveth.
\par }{\Q \VS{22}For
{\ADD{such as be}}
blessed of him shall
inherit the
earth; and
{\ADD{they that be}}
cursed of him shall be cut
off.
\par }{\Q \VS{23}The
steps of a
{\ADD{good}}
man are
ordered by the
{\ND{
LORD}}: and he
delighteth in his
way.\FTNT{‡‡}{{\FR 37.23 }
ordered: or, established}
\par }{\Q \VS{24}Though he
fall, he shall not be utterly cast
down: for the
{\ND{
LORD}}
upholdeth
{\ADD{him with}} his
hand.
\par }{\Q \VS{25}I have been
young, and
{\ADD{now}} am
old; yet have I not
seen the
righteous
forsaken, nor his
seed
begging
bread.
\par }{\Q \VS{26}{\ADD{He is}}
ever
merciful, and
lendeth; and his
seed
{\ADD{is}}
blessed.\FTNT{§§}{{\FR 37.26 }
ever: Heb. all the day}
\par }{\Q \VS{27}Depart from
evil, and
do
good; and
dwell for
evermore.
\par }{\Q \VS{28}For the
{\ND{
LORD}}
loveth
judgment, and
forsaketh not his
saints; they are
preserved for
ever: but the
seed of the
wicked shall be cut
off.
\par }{\Q \VS{29}The
righteous shall
inherit the
land, and
dwell therein for
ever.
\par }{\Q \VS{30}The
mouth of the
righteous
speaketh
wisdom, and his
tongue
talketh of
judgment.
\par }{\Q \VS{31}The
law of his
God
{\ADD{is}} in his
heart; none of his
steps shall
slide.\FTNT{***}{{\FR 37.31 }
steps: or, goings}
\par }{\Q \VS{32}The
wicked
watcheth the
righteous, and
seeketh to
slay him.
\par }{\Q \VS{33}The
{\ND{
LORD}} will not
leave him in his
hand, nor
condemn him when he is
judged.
\par }{\BB \par }{\Q \VS{34}Wait on the
{\ND{
LORD}}, and
keep his
way, and he shall
exalt thee to
inherit the
land: when the
wicked are cut
off, thou shalt
see
{\ADD{it}}.
\par }{\Q \VS{35}I have
seen the
wicked in great
power, and
spreading himself like a
green bay
tree.\FTNT{†††}{{\FR 37.35 }
a green…: or, a green tree that groweth in his own soil}
\par }{\Q \VS{36}Yet he passed
away, and, lo, he
{\ADD{was}} not: yea, I
sought him, but he could not be
found.
\par }{\Q \VS{37}Mark the
perfect
{\ADD{man}}, and
behold the
upright: for the
end of
{\ADD{that}}
man
{\ADD{is}}
peace.
\par }{\Q \VS{38}But the
transgressors shall be
destroyed
together: the
end of the
wicked shall be cut
off.
\par }{\Q \VS{39}But the
salvation of the
righteous
{\ADD{is}} of the
{\ND{
LORD}}:
{\ADD{he is}} their
strength in the
time of
trouble.
\par }{\Q \VS{40}And the
{\ND{
LORD}} shall
help them, and
deliver them: he shall
deliver them from the
wicked, and
save them, because they
trust in him.

\par }\PsalmChap{38}{\D A
Psalm of
David, to bring to
remembrance.
\par }{\Q \VerseOne{1}O
{\ND{
LORD}},
rebuke me not in thy
wrath: neither
chasten me in thy hot
displeasure.
\par }{\Q \VS{2}For thine
arrows stick
fast in me, and thy
hand presseth me
sore.
\par }{\Q \VS{3}{\ADD{There is}} no
soundness in my
flesh
because of thine
anger; neither
{\ADD{is there any}}
rest in my
bones
because of my
sin.\FTNT{*}{{\FR 38.3 }
rest: Heb. peace, or, health}
\par }{\Q \VS{4}For mine
iniquities are gone
over mine
head: as an
heavy
burden they are too
heavy for me.
\par }{\Q \VS{5}My
wounds
stink
{\ADD{and}} are
corrupt
because of my
foolishness.
\par }{\Q \VS{6}I am
troubled; I am bowed
down
greatly; I
go
mourning all the
day long.\FTNT{†}{{\FR 38.6 }
troubled: Heb. wried}
\par }{\Q \VS{7}For my
loins are
filled with a
loathsome
{\ADD{disease:}} and
{\ADD{there is}} no
soundness in my
flesh.
\par }{\Q \VS{8}I am
feeble and
sore
broken: I have
roared by reason of the
disquietness of my
heart.
\par }{\Q \VS{9}Lord, all my
desire
{\ADD{is}} before thee; and my
groaning is not
hid from thee.
\par }{\Q \VS{10}My
heart
panteth, my
strength
faileth me: as for the
light of mine
eyes,
it also is
gone from me.\FTNT{‡}{{\FR 38.10 }
is gone…: Heb. is not with me}
\par }{\Q \VS{11}My
lovers and my
friends
stand aloof
from my
sore; and my
kinsmen
stand afar
off.\FTNT{§}{{\FR 38.11 }
sore: Heb. stroke}\FTNT{**}{{\FR 38.11 }
my kinsmen: or, my neighbours}
\par }{\BB \par }{\Q \VS{12}They also that
seek after my
life lay
snares
{\ADD{for me:}} and they that
seek my
hurt
speak mischievous
things, and
imagine
deceits all the
day long.
\par }{\Q \VS{13}But I, as a
deaf
{\ADD{man}},
heard not; and
{\ADD{I was}} as a dumb
man
{\ADD{that}}
openeth not his
mouth.
\par }{\Q \VS{14}Thus I was as a
man that
heareth not, and in whose
mouth
{\ADD{are}} no
reproofs.
\par }{\Q \VS{15}For in thee, O
{\ND{
LORD}}, do I
hope: thou wilt
hear, O
Lord my
God.\FTNT{††}{{\FR 38.15 }
in…: or, thee do I wait for}\FTNT{‡‡}{{\FR 38.15 }
hear: or, answer}
\par }{\Q \VS{16}For I
said,
{\ADD{Hear me}}, lest
{\ADD{otherwise}} they should
rejoice over me: when my
foot
slippeth, they
magnify
{\ADD{themselves}} against me.
\par }{\Q \VS{17}For I
{\ADD{am}}
ready to
halt, and my
sorrow
{\ADD{is}}
continually before me.\FTNT{§§}{{\FR 38.17 }
to halt: Heb. for halting}
\par }{\Q \VS{18}For I will
declare mine
iniquity; I will be
sorry for my
sin.
\par }{\Q \VS{19}But mine
enemies
{\ADD{are}}
lively,
{\ADD{and}} they are
strong: and they that
hate me
wrongfully are
multiplied.\FTNT{***}{{\FR 38.19 }
are lively…: Heb. being living, are strong}
\par }{\Q \VS{20}They also that
render
evil for
good are mine
adversaries; because I
follow
{\ADD{the thing that}}
good
{\ADD{is}}.
\par }{\Q \VS{21}Forsake me not, O
{\ND{
LORD}}: O my
God, be not
far from me.
\par }{\Q \VS{22}Make
haste to
help me, O
Lord my
salvation.\FTNT{†††}{{\FR 38.22 }
to…: Heb. for my help}

\par }\PsalmChap{39}{\D To the chief
Musician,
{\ADD{even}} to
Jeduthun, A
Psalm of
David.
\par }{\Q \VerseOne{1}I
said, I will take
heed to my
ways, that I
sin not with my
tongue: I will
keep my
mouth with a
bridle, while the
wicked is before me.\FTNT{*}{{\FR 39.1 }
my mouth…: Heb. a bridle, or, muzzle for my mouth}
\par }{\Q \VS{2}I was
dumb with
silence, I held my
peace,
{\ADD{even}} from
good; and my
sorrow was
stirred.\FTNT{†}{{\FR 39.2 }
stirred: Heb. troubled}
\par }{\Q \VS{3}My
heart was
hot
within me, while I was
musing the
fire
burned:
{\ADD{
then}}
spake I with my
tongue,
\par }{\Q \VS{4}{\ND{
LORD}}, make me to
know mine
end, and the
measure of my
days, what it
{\ADD{is; that}} I may
know how
frail I
{\ADD{am}}.\FTNT{‡}{{\FR 39.4 }
how…: or, what time I have here}
\par }{\Q \VS{5}Behold, thou hast
made my
days
{\ADD{as}} an
handbreadth; and mine
age
{\ADD{is}} as nothing before thee: verily
every
man at his best
state
{\ADD{is}}
altogether
vanity.
Selah.\FTNT{§}{{\FR 39.5 }
at…: Heb. settled}
\par }{\Q \VS{6}Surely every
man
walketh in a vain
shew: surely they are
disquieted in
vain: he heapeth
up
{\ADD{riches}}, and
knoweth not who shall
gather them.\FTNT{**}{{\FR 39.6 }
a vain…: Heb. an image}
\par }{\BB \par }{\Q \VS{7}And now,
Lord, what
wait I for? my
hope
{\ADD{is}} in thee.
\par }{\Q \VS{8}Deliver me from all my
transgressions:
make me not the
reproach of the
foolish.
\par }{\Q \VS{9}I was
dumb, I
opened not my
mouth; because thou
didst
{\ADD{it}}.
\par }{\Q \VS{10}Remove thy
stroke
away from me: I am
consumed by the
blow of thine
hand.\FTNT{††}{{\FR 39.10 }
blow: Heb. conflict}
\par }{\Q \VS{11}When thou with
rebukes dost
correct
man for
iniquity, thou makest his
beauty to consume
away like a
moth: surely every
man
{\ADD{is}}
vanity.
Selah.\FTNT{‡‡}{{\FR 39.11 }
his…: Heb. that which is to be desired in him to melt away}
\par }{\Q \VS{12}Hear my
prayer, O
{\ND{
LORD}}, and give
ear unto my
cry; hold not thy
peace at my
tears: for I
{\ADD{am}} a
stranger with thee,
{\ADD{and}} a
sojourner, as all my
fathers
{\ADD{were}}.
\par }{\Q \VS{13}O
spare me, that I may recover
strength, before I go
hence, and be no more.

\par }\PsalmChap{40}{\D To the chief
Musician, A
Psalm of
David.
\par }{\Q \VerseOne{1}I
waited
patiently for the
{\ND{
LORD}}; and he
inclined unto me, and
heard my
cry.\FTNT{*}{{\FR 40.1 }
I waited…: Heb. In waiting I waited}
\par }{\Q \VS{2}He brought me
up also out of an
horrible
pit, out of the
miry
clay, and
set my
feet upon a
rock,
{\ADD{and}}
established my
goings.\FTNT{†}{{\FR 40.2 }
an…: Heb. a pit of noise}
\par }{\Q \VS{3}And he hath
put a
new
song in my
mouth,
{\ADD{even}}
praise unto our
God:
many shall
see
{\ADD{it}}, and
fear, and shall
trust in the
{\ND{
LORD}}.
\par }{\Q \VS{4}Blessed
{\ADD{is}} that
man that
maketh the
{\ND{
LORD}} his
trust, and
respecteth not the
proud, nor such as turn
aside to
lies.
\par }{\Q \VS{5}Many, O
{\ND{
LORD}} my
God,
{\ADD{are}} thy
wonderful works
{\ADD{which}} thou hast
done, and thy
thoughts
{\ADD{which are}} to us-ward: they cannot be reckoned up in
order unto thee:
{\ADD{if}} I would
declare and
speak
{\ADD{of them}}, they are
more than can be
numbered.\FTNT{‡}{{\FR 40.5 }
they cannot…: or, none can order them unto thee}
\par }{\BB \par }{\Q \VS{6}Sacrifice and
offering thou didst not
desire; mine
ears hast thou
opened: burnt
offering and sin
offering hast thou not
required.\FTNT{§}{{\FR 40.6 }
opened: Heb. digged}
\par }{\Q \VS{7}Then
said I, Lo, I
come: in the
volume of the
book
{\ADD{it is}}
written of me,
\par }{\Q \VS{8}I
delight to
do thy
will, O my
God: yea, thy
law
{\ADD{is}}
within my
heart.\FTNT{**}{{\FR 40.8 }
within…: Heb. in the midst of my bowels}
\par }{\Q \VS{9}I have
preached
righteousness in the
great
congregation: lo, I have not
refrained my
lips, O
{\ND{
LORD}}, thou
knowest.
\par }{\Q \VS{10}I have not
hid thy
righteousness
within my
heart; I have
declared thy
faithfulness and thy
salvation: I have not
concealed thy
lovingkindness and thy
truth from the
great
congregation.
\par }{\BB \par }{\Q \VS{11}Withhold not thou thy tender
mercies from me, O
{\ND{
LORD}}: let thy
lovingkindness and thy
truth
continually
preserve me.
\par }{\Q \VS{12}For
innumerable
evils have
compassed me about: mine
iniquities have taken
hold upon me, so that I am not
able to look
up; they are
more than the
hairs of mine
head: therefore my
heart
faileth me.\FTNT{††}{{\FR 40.12 }
faileth: Heb. forsaketh}
\par }{\Q \VS{13}Be
pleased, O
{\ND{
LORD}}, to
deliver me: O
{\ND{
LORD}}, make
haste to
help me.
\par }{\Q \VS{14}Let them be
ashamed and
confounded
together that
seek after my
soul to
destroy it; let them be
driven
backward and put to
shame that
wish me
evil.
\par }{\Q \VS{15}Let them be
desolate for a
reward of their
shame that
say unto me,
Aha,
aha.
\par }{\Q \VS{16}Let all those that
seek thee
rejoice and be
glad in thee: let such as
love thy
salvation
say
continually, The
{\ND{
LORD}} be
magnified.
\par }{\Q \VS{17}But I
{\ADD{am}}
poor and
needy;
{\ADD{yet}} the
Lord
thinketh upon me: thou
{\ADD{art}} my
help and my
deliverer; make no
tarrying, O my
God.

\par }\PsalmChap{41}{\D To the chief
Musician, A
Psalm of
David.
\par }{\Q \VerseOne{1}Blessed
{\ADD{is}} he that
considereth the
poor: the
{\ND{
LORD}} will
deliver him in
time of
trouble.\FTNT{*}{{\FR 41.1 }
the poor: or, the weak, or, sick}\FTNT{†}{{\FR 41.1 }
in time…: Heb. in the day of evil}
\par }{\Q \VS{2}The
{\ND{
LORD}} will
preserve him, and keep him
alive;
{\ADD{and}} he shall be
blessed upon the
earth: and thou wilt not
deliver him unto the
will of his
enemies.\FTNT{‡}{{\FR 41.2 }
thou…: or, do not thou deliver}
\par }{\Q \VS{3}The
{\ND{
LORD}} will
strengthen him upon the
bed of
languishing: thou wilt
make all his
bed in his
sickness.\FTNT{§}{{\FR 41.3 }
make: Heb. turn}
\par }{\Q \VS{4}I
said,
{\ND{
LORD}}, be
merciful unto me:
heal my
soul; for I have
sinned against thee.
\par }{\BB \par }{\Q \VS{5}Mine
enemies
speak
evil of me, When shall he
die, and his
name
perish?
\par }{\Q \VS{6}And if he
come to
see
{\ADD{me}}, he
speaketh
vanity: his
heart
gathereth
iniquity to itself;
{\ADD{when}} he
goeth
abroad, he
telleth
{\ADD{it}}.
\par }{\Q \VS{7}All that
hate me
whisper
together against me: against me do they
devise my
hurt.\FTNT{**}{{\FR 41.7 }
my…: Heb. evil to me}
\par }{\Q \VS{8}An
evil
disease,
{\ADD{say they}}, cleaveth
fast unto him: and
{\ADD{now}} that he
lieth he shall rise
up no
more.\FTNT{††}{{\FR 41.8 }
An…: Heb. A thing of Belial}
\par }{\Q \VS{9}Yea, mine own
familiar
friend, in whom I
trusted, which did
eat of my
bread, hath lifted
up
{\ADD{his}}
heel against me.\FTNT{‡‡}{{\FR 41.9 }
mine…: Heb. the man of my peace}\FTNT{§§}{{\FR 41.9 }
lifted…: Heb. magnified}
\par }{\Q \VS{10}But thou, O
{\ND{
LORD}}, be
merciful unto me, and raise me
up, that I may
requite them.
\par }{\Q \VS{11}By this I
know that thou
favourest me, because mine
enemy doth not
triumph over me.
\par }{\Q \VS{12}And as for me, thou
upholdest me in mine
integrity, and
settest me before thy
face for
ever.
\par }{\Q \VS{13}Blessed
{\ADD{be}} the
{\ND{
LORD}}
God of
Israel from
everlasting, and to
everlasting.
Amen, and
Amen.

\par }\PsalmChap{42}{\D To the chief
Musician,
Maschil, for the
sons of
Korah.
\par }{\Q \VerseOne{1}As the
hart
panteth after the
water
brooks, so
panteth my
soul after thee, O
God.\FTNT{*}{{\FR 42.1 }
Maschil…: or, A Psalm giving instruction of the sons, etc}\FTNT{†}{{\FR 42.1 }
panteth: Heb. brayeth}
\par }{\Q \VS{2}My
soul
thirsteth for
God, for the
living
God: when shall I
come and
appear
before
God?
\par }{\Q \VS{3}My
tears have been my
meat
day and
night, while they
continually
say unto me, Where
{\ADD{is}} thy
God?
\par }{\Q \VS{4}When I
remember these
{\ADD{things}}, I pour
out my
soul in me: for I had
gone with the
multitude, I
went with them to the
house of
God, with the
voice of
joy and
praise, with a
multitude that kept
holyday.
\par }{\Q \VS{5}Why art thou cast
down, O my
soul? and
{\ADD{why}} art thou
disquieted in me?
hope thou in
God: for I shall yet
praise him
{\ADD{for}} the
help of his
countenance.\FTNT{‡}{{\FR 42.5 }
cast: Heb. bowed}\FTNT{§}{{\FR 42.5 }
praise: or, give thanks}\FTNT{**}{{\FR 42.5 }
for the…: or, his presence is salvation}
\par }{\BB \par }{\Q \VS{6}O my
God, my
soul is cast
down within me: therefore will I
remember thee from the
land of
Jordan, and of the
Hermonites, from the
hill
Mizar.\FTNT{††}{{\FR 42.6 }
the hill…: or, the little hill}
\par }{\Q \VS{7}Deep
calleth unto
deep at the
noise of thy
waterspouts: all thy
waves and thy
billows are
gone over me.
\par }{\Q \VS{8}{\ADD{Yet}} the
{\ND{
LORD}} will
command his
lovingkindness in the
daytime, and in the
night his
song
{\ADD{shall be}} with me,
{\ADD{and}} my
prayer unto the
God of my
life.
\par }{\Q \VS{9}I will
say unto
God my
rock, Why hast thou
forgotten me? why
go I
mourning because of the
oppression of the
enemy?
\par }{\Q \VS{10}{\ADD{As}} with a
sword in my
bones, mine
enemies
reproach me; while they
say
daily unto me, Where
{\ADD{is}} thy
God?\FTNT{‡‡}{{\FR 42.10 }
sword: or, killing}
\par }{\Q \VS{11}Why art thou cast
down, O my
soul? and why art thou
disquieted within me?
hope thou in
God: for I shall yet
praise him,
{\ADD{who is}} the
health of my
countenance, and my
God.

\par }\PsalmChap{43}{\Q \VerseOne{1}Judge me, O
God, and
plead my
cause against an
ungodly
nation: O
deliver me from the
deceitful and
unjust
man.\FTNT{*}{{\FR 43.1 }
ungodly: or, unmerciful}\FTNT{†}{{\FR 43.1 }
the deceitful…: Heb. a man of deceit and iniquity}
\par }{\Q \VS{2}For thou
{\ADD{art}} the
God of my
strength: why dost thou cast me
off? why
go I
mourning because of the
oppression of the
enemy?
\par }{\Q \VS{3}O send
out thy
light and thy
truth: let them
lead me; let them
bring me unto thy
holy
hill, and to thy
tabernacles.
\par }{\Q \VS{4}Then will I
go unto the
altar of
God, unto
God my
exceeding
joy: yea, upon the
harp will I
praise thee, O
God my
God.\FTNT{‡}{{\FR 43.4 }
my exceeding…: Heb. the gladness of my joy}
\par }{\Q \VS{5}Why art thou cast
down, O my
soul? and why art thou
disquieted within me?
hope in
God: for I shall yet
praise him,
{\ADD{who is}} the
health of my
countenance, and my
God.

\par }\PsalmChap{44}{\D To the chief
Musician for the
sons of
Korah,
Maschil.
\par }{\Q \VerseOne{1}We have
heard with our
ears, O
God, our
fathers have
told us,
{\ADD{what}}
work thou
didst in their
days, in the
times of
old.\FTNT{*}{{\FR 44.1 }
Maschil: or, of instruction}
\par }{\Q \VS{2}{\ADD{How}} thou didst drive
out the
heathen with thy
hand, and
plantedst them;
{\ADD{how}} thou didst
afflict the
people, and cast them
out.
\par }{\Q \VS{3}For they
got not the
land in
possession by their own
sword, neither did their own
arm
save them: but thy right
hand, and thine
arm, and the
light of thy
countenance, because thou hadst a
favour unto them.
\par }{\Q \VS{4}Thou art my
King, O
God:
command
deliverances for
Jacob.
\par }{\Q \VS{5}Through thee will we push
down our
enemies: through thy
name will we tread them
under that rise
up against us.
\par }{\Q \VS{6}For I will not
trust in my
bow, neither shall my
sword
save me.
\par }{\Q \VS{7}But thou hast
saved us from our
enemies, and hast put them to
shame that
hated us.
\par }{\Q \VS{8}In
God we
boast all the
day long, and
praise thy
name for
ever.
Selah.
\par }{\BB \par }{\Q \VS{9}But thou hast cast
off, and put us to
shame; and goest not
forth with our
armies.
\par }{\Q \VS{10}Thou makest us to
turn
back from the
enemy: and they which
hate us
spoil for themselves.
\par }{\Q \VS{11}Thou hast
given us like
sheep
{\ADD{appointed}} for
meat; and hast
scattered us among the
heathen.\FTNT{†}{{\FR 44.11 }
like…: Heb. as sheep of meat}
\par }{\Q \VS{12}Thou
sellest thy
people for
nought, and dost not
increase
{\ADD{thy wealth}} by their
price.\FTNT{‡}{{\FR 44.12 }
for…: Heb. without riches}
\par }{\Q \VS{13}Thou
makest us a
reproach to our
neighbours, a
scorn and a
derision to them that are round
about us.
\par }{\Q \VS{14}Thou
makest us a
byword among the
heathen, a
shaking of the
head among the
people.
\par }{\Q \VS{15}My
confusion
{\ADD{is}}
continually before me, and the
shame of my
face hath
covered me,
\par }{\Q \VS{16}For the
voice of him that
reproacheth and
blasphemeth; by
reason of the
enemy and
avenger.
\par }{\BB \par }{\Q \VS{17}All this is
come upon us; yet have we not
forgotten thee, neither have we dealt
falsely in thy
covenant.
\par }{\Q \VS{18}Our
heart is not
turned
back, neither have our
steps
declined from thy
way;\FTNT{§}{{\FR 44.18 }
steps: or, goings}
\par }{\Q \VS{19}Though thou hast sore
broken us in the
place of
dragons, and
covered us with the shadow of
death.
\par }{\Q \VS{20}If we have
forgotten the
name of our
God, or stretched
out our
hands to a
strange
god;
\par }{\Q \VS{21}Shall not
God search this
out? for he
knoweth the
secrets of the
heart.
\par }{\Q \VS{22}Yea, for thy sake are we
killed all the
day long; we are
counted as
sheep for the
slaughter.
\par }{\Q \VS{23}Awake, why
sleepest thou, O
Lord?
arise, cast
{\ADD{us}} not
off for
ever.
\par }{\Q \VS{24}Wherefore
hidest thou thy
face,
{\ADD{and}}
forgettest our
affliction and our
oppression?
\par }{\Q \VS{25}For our
soul is bowed
down to the
dust: our
belly
cleaveth unto the
earth.
\par }{\Q \VS{26}Arise for our
help, and
redeem us for thy
mercies’ sake.\FTNT{**}{{\FR 44.26 }
for our…: Heb. a help for us}

\par }\PsalmChap{45}{\D To the chief
Musician upon
Shoshannim, for the
sons of
Korah,
Maschil, A
Song of
loves.
\par }{\Q \VerseOne{1}My
heart is
inditing a
good
matter: I
speak of the things which I have
made touching the
king: my
tongue
{\ADD{is}} the
pen of a
ready
writer.\FTNT{*}{{\FR 45.1 }
Maschil: or, of instruction}\FTNT{†}{{\FR 45.1 }
is inditing: Heb. boileth, or, bubbleth up}
\par }{\Q \VS{2}Thou art
fairer than the
children of
men:
grace is
poured into thy
lips: therefore
God hath
blessed thee for
ever.
\par }{\Q \VS{3}Gird thy
sword upon
{\ADD{thy}}
thigh, O
{\ADD{most}}
mighty, with thy
glory and thy
majesty.
\par }{\Q \VS{4}And in thy
majesty
ride
prosperously
because of
truth and
meekness
{\ADD{and}}
righteousness; and thy right
hand shall
teach thee terrible
things.\FTNT{‡}{{\FR 45.4 }
ride…: Heb. prosper thou, ride thou}
\par }{\Q \VS{5}Thine
arrows
{\ADD{are}}
sharp in the
heart of the
king’s
enemies;
{\ADD{whereby}} the
people
fall under thee.
\par }{\BB \par }{\Q \VS{6}Thy
throne, O
God,
{\ADD{is}} for
ever and
ever: the
sceptre of thy
kingdom
{\ADD{is}} a
right
sceptre.
\par }{\Q \VS{7}Thou
lovest
righteousness, and
hatest
wickedness: therefore
God, thy
God, hath
anointed thee with the
oil of
gladness above thy
fellows.
\par }{\Q \VS{8}All thy
garments
{\ADD{smell}} of
myrrh, and
aloes,
{\ADD{and}}
cassia, out of the
ivory
palaces,
whereby they have made thee
glad.
\par }{\Q \VS{9}Kings’
daughters
{\ADD{were}} among thy honourable
women: upon thy right
hand did
stand the
queen in
gold of
Ophir.
\par }{\BB \par }{\Q \VS{10}Hearken, O
daughter, and
consider, and
incline thine
ear;
forget also thine own
people, and thy
father’s
house;
\par }{\Q \VS{11}So shall the
king greatly
desire thy
beauty: for he
{\ADD{is}} thy
Lord; and
worship thou him.
\par }{\Q \VS{12}And the
daughter of
Tyre
{\ADD{shall be there}} with a
gift;
{\ADD{even}} the
rich among the
people shall
intreat thy
favour.\FTNT{§}{{\FR 45.12 }
favour: Heb. face}
\par }{\Q \VS{13}The
king’s
daughter
{\ADD{is}} all
glorious
within: her
clothing
{\ADD{is}} of
wrought
gold.
\par }{\Q \VS{14}She shall be
brought unto the
king in raiment of
needlework: the
virgins her
companions that
follow her shall be
brought unto thee.
\par }{\Q \VS{15}With
gladness and
rejoicing shall they be
brought: they shall
enter into the
king’s
palace.
\par }{\Q \VS{16}Instead of thy
fathers shall be thy
children, whom thou mayest
make
princes in all the
earth.
\par }{\Q \VS{17}I will make thy
name to be
remembered in
all
generations: therefore shall the
people
praise thee for
ever and
ever.

\par }\PsalmChap{46}{\D To the chief
Musician for the
sons of
Korah, A
Song upon
Alamoth.
\par }{\Q \VerseOne{1}God
{\ADD{is}} our
refuge and
strength, a
very
present
help in
trouble.\FTNT{*}{{\FR 46.1 }
for: or, of}
\par }{\Q \VS{2}Therefore will not we
fear, though the
earth be
removed, and though the
mountains be
carried into the
midst of the
sea;\FTNT{†}{{\FR 46.2 }
midst…: Heb. heart of the seas}
\par }{\Q \VS{3}{\ADD{Though}} the
waters thereof
roar
{\ADD{and}} be
troubled,
{\ADD{though}} the
mountains
shake with the
swelling thereof.
Selah.
\par }{\Q \VS{4}{\ADD{There is}} a
river, the
streams whereof shall make
glad the
city of
God, the
holy
{\ADD{place}} of the
tabernacles of the most
High.
\par }{\Q \VS{5}God
{\ADD{is}} in the
midst of her; she shall not be
moved:
God shall
help her,
{\ADD{and that}}
right
early.\FTNT{‡}{{\FR 46.5 }
and…: Heb. when the morning appeareth}
\par }{\BB \par }{\Q \VS{6}The
heathen
raged, the
kingdoms were
moved: he
uttered his
voice, the
earth
melted.
\par }{\Q \VS{7}The
{\ND{
LORD}} of
hosts
{\ADD{is}} with us; the
God of
Jacob
{\ADD{is}} our
refuge.
Selah.\FTNT{§}{{\FR 46.7 }
our…: Heb. an high place for us}
\par }{\Q \VS{8}Come,
behold the
works of the
{\ND{
LORD}}, what
desolations he hath
made in the
earth.
\par }{\Q \VS{9}He maketh
wars to
cease unto the
end of the
earth; he
breaketh the
bow, and
cutteth the
spear in
sunder; he
burneth the
chariot in the
fire.
\par }{\Q \VS{10}Be
still, and
know that I
{\ADD{am}}
God: I will be
exalted among the
heathen, I will be
exalted in the
earth.
\par }{\Q \VS{11}The
{\ND{
LORD}} of
hosts
{\ADD{is}} with us; the
God of
Jacob
{\ADD{is}} our
refuge.
Selah.

\par }\PsalmChap{47}{\D To the chief
Musician, A
Psalm for the
sons of
Korah.
\par }{\Q \VerseOne{1}O
clap your
hands, all ye
people;
shout unto
God with the
voice of
triumph.\FTNT{*}{{\FR 47.1 }
for: or, of}
\par }{\Q \VS{2}For the
{\ND{
LORD}} most
high
{\ADD{is}}
terrible;
{\ADD{he is}} a
great
King over all the
earth.
\par }{\Q \VS{3}He shall
subdue the
people under us, and the
nations under our
feet.
\par }{\Q \VS{4}He shall
choose our
inheritance for us, the
excellency of
Jacob whom he
loved.
Selah.
\par }{\BB \par }{\Q \VS{5}God is gone
up with a
shout, the
{\ND{
LORD}} with the
sound of a
trumpet.
\par }{\Q \VS{6}Sing
praises to
God, sing
praises: sing
praises unto our
King, sing
praises.
\par }{\Q \VS{7}For
God
{\ADD{is}} the
King of all the
earth: sing ye
praises with
understanding.\FTNT{†}{{\FR 47.7 }
with: or, every one that hath}
\par }{\Q \VS{8}God
reigneth over the
heathen:
God
sitteth upon the
throne of his
holiness.
\par }{\Q \VS{9}The
princes of the
people are gathered
together,
{\ADD{even}} the
people of the
God of
Abraham: for the
shields of the
earth
{\ADD{belong}} unto
God: he is
greatly
exalted.\FTNT{‡}{{\FR 47.9 }
princes…: or, voluntary of the people are gathered unto the people of the God of Abraham}

\par }\PsalmChap{48}{\D A
Song
{\ADD{and}}
Psalm for the
sons of
Korah.
\par }{\Q \VerseOne{1}Great
{\ADD{is}} the
{\ND{
LORD}}, and
greatly to be
praised in the
city of our
God,
{\ADD{in}} the
mountain of his
holiness.\FTNT{*}{{\FR 48.1 }
for: or, of}
\par }{\Q \VS{2}Beautiful for
situation, the
joy of the whole
earth,
{\ADD{is}}
mount
Zion,
{\ADD{on}} the
sides of the
north, the
city of the
great
King.
\par }{\Q \VS{3}God is
known in her
palaces for a
refuge.
\par }{\Q \VS{4}For, lo, the
kings were
assembled, they passed
by
together.
\par }{\Q \VS{5}They
saw
{\ADD{it, and}} so they
marvelled; they were
troubled,
{\ADD{and}} hasted
away.
\par }{\Q \VS{6}Fear took
hold upon them there,
{\ADD{and}}
pain, as of a woman in
travail.
\par }{\Q \VS{7}Thou
breakest the
ships of
Tarshish with an
east
wind.
\par }{\BB \par }{\Q \VS{8}As we have
heard, so have we
seen in the
city of the
{\ND{
LORD}} of
hosts, in the
city of our
God:
God will
establish it
for
ever.
Selah.
\par }{\Q \VS{9}We have
thought of thy
lovingkindness, O
God, in the
midst of thy
temple.
\par }{\Q \VS{10}According to thy
name, O
God, so
{\ADD{is}} thy
praise unto the
ends of the
earth: thy right
hand is
full of
righteousness.
\par }{\Q \VS{11}Let
mount
Zion
rejoice, let the
daughters of
Judah be
glad, because of thy
judgments.
\par }{\Q \VS{12}Walk
about
Zion, and go round
about her:
tell the
towers thereof.
\par }{\Q \VS{13}Mark ye
well her
bulwarks,
consider her
palaces; that ye may
tell
{\ADD{it}} to the
generation
following.\FTNT{†}{{\FR 48.13 }
Mark…: Heb. Set your heart to}\FTNT{‡}{{\FR 48.13 }
consider: or, raise up}
\par }{\Q \VS{14}For this
God
{\ADD{is}} our
God for
ever and
ever: he will be our
guide
{\ADD{even}} unto
death.

\par }\PsalmChap{49}{\D To the chief
Musician, A
Psalm for the
sons of
Korah.
\par }{\Q \VerseOne{1}Hear this, all
{\ADD{ye}}
people; give
ear, all
{\ADD{ye}}
inhabitants of the
world:\FTNT{*}{{\FR 49.1 }
for: or, of}
\par }{\Q \VS{2}Both
low and
high,
rich and
poor,
together.
\par }{\Q \VS{3}My
mouth shall
speak of
wisdom; and the
meditation of my
heart
{\ADD{shall be}} of
understanding.
\par }{\Q \VS{4}I will
incline mine
ear to a
parable: I will
open my dark
saying upon the
harp.
\par }{\Q \VS{5}Wherefore should I
fear in the
days of
evil,
{\ADD{when}} the
iniquity of my
heels shall
compass me about?
\par }{\BB \par }{\Q \VS{6}They that
trust in their
wealth, and
boast themselves in the
multitude of their
riches;
\par }{\Q \VS{7}None
{\ADD{of them}}
can by any
means
redeem his
brother, nor
give to
God a
ransom for him:
\par }{\Q \VS{8}(For the
redemption of their
soul
{\ADD{is}}
precious, and it
ceaseth for
ever:)
\par }{\Q \VS{9}That he should still
live for
ever,
{\ADD{and}} not
see
corruption.
\par }{\Q \VS{10}For he
seeth
{\ADD{that}} wise
men
die,
likewise the
fool and the brutish
person
perish, and
leave their
wealth to
others.
\par }{\Q \VS{11}Their
inward thought
{\ADD{is, that}} their
houses
{\ADD{shall continue}} for
ever,
{\ADD{and}} their
dwelling places to
all
generations; they
call
{\ADD{their}}
lands after their own
names.\FTNT{†}{{\FR 49.11 }
all…: Heb. generation and generation}
\par }{\Q \VS{12}Nevertheless
man
{\ADD{being}} in
honour
abideth not: he is
like the
beasts
{\ADD{that}}
perish.
\par }{\Q \VS{13}This their
way
{\ADD{is}} their
folly: yet their
posterity
approve their
sayings.
Selah.\FTNT{‡}{{\FR 49.13 }
approve…: Heb. delight in their mouth}
\par }{\Q \VS{14}Like
sheep they are
laid in the
grave;
death shall
feed on them; and the
upright shall have
dominion over them in the
morning; and their
beauty shall
consume in the
grave from their
dwelling.\FTNT{§}{{\FR 49.14 }
beauty: or, strength}\FTNT{**}{{\FR 49.14 }
in the grave from…: or, the grave being an habitation to every one of them}
\par }{\BB \par }{\Q \VS{15}But
God will
redeem my
soul from the
power of the
grave: for he shall
receive me.
Selah.\FTNT{††}{{\FR 49.15 }
power: Heb. hand}\FTNT{‡‡}{{\FR 49.15 }
the grave: or, hell}
\par }{\Q \VS{16}Be not thou
afraid when
one is made
rich, when the
glory of his
house is
increased;
\par }{\Q \VS{17}For when he
dieth he shall carry nothing
away: his
glory shall not
descend
after him.
\par }{\Q \VS{18}Though while he
lived he
blessed his
soul: and
{\ADD{men}} will
praise thee, when thou doest
well to thyself.\FTNT{§§}{{\FR 49.18 }
while…: Heb. in his life}
\par }{\Q \VS{19}He shall
go to the
generation of his
fathers; they shall
never
see
light.\FTNT{***}{{\FR 49.19 }
He…: Heb. The soul shall}
\par }{\Q \VS{20}Man
{\ADD{that is}} in
honour, and
understandeth not, is
like the
beasts
{\ADD{that}}
perish.

\par }\PsalmChap{50}{\D A
Psalm of
Asaph.
\par }{\Q \VerseOne{1}The
mighty
God,
{\ADD{even}} the
{\ND{
LORD}}, hath
spoken, and
called the
earth from the
rising of the
sun unto the going
down thereof.\FTNT{*}{{\FR 50.1 }
of…: or, for Asaph}
\par }{\Q \VS{2}Out of
Zion, the
perfection of
beauty,
God hath
shined.
\par }{\Q \VS{3}Our
God shall
come, and shall not keep
silence: a
fire shall
devour
before him, and it shall be
very
tempestuous round
about him.
\par }{\Q \VS{4}He shall
call to the
heavens from
above, and to the
earth, that he may
judge his
people.
\par }{\Q \VS{5}Gather my
saints together unto me; those that have
made a
covenant with me by
sacrifice.
\par }{\Q \VS{6}And the
heavens shall
declare his
righteousness: for
God
{\ADD{is}}
judge himself.
Selah.
\par }{\BB \par }{\Q \VS{7}Hear, O my
people, and I will
speak; O
Israel, and I will
testify against thee: I
{\ADD{am}}
God,
{\ADD{even}} thy
God.
\par }{\Q \VS{8}I will not
reprove thee for thy
sacrifices or thy burnt
offerings,
{\ADD{to have been}}
continually before me.
\par }{\Q \VS{9}I will
take no
bullock out of thy
house,
{\ADD{nor}} he
goats out of thy
folds.
\par }{\Q \VS{10}For every
beast of the
forest
{\ADD{is}} mine,
{\ADD{and}} the
cattle upon a
thousand
hills.
\par }{\Q \VS{11}I
know all the
fowls of the
mountains: and the wild
beasts of the
field
{\ADD{are}}
mine.\FTNT{†}{{\FR 50.11 }
mine: Heb. with me}
\par }{\Q \VS{12}If I were
hungry, I would not
tell thee: for the
world
{\ADD{is}} mine, and the
fulness thereof.
\par }{\Q \VS{13}Will I
eat the
flesh of
bulls, or
drink the
blood of
goats?
\par }{\Q \VS{14}Offer unto
God
thanksgiving; and
pay thy
vows unto the most
High:
\par }{\Q \VS{15}And
call upon me in the
day of
trouble: I will
deliver thee, and thou shalt
glorify me.
\par }{\BB \par }{\Q \VS{16}But unto the
wicked
God
saith, What hast thou to do to
declare my
statutes, or
{\ADD{that}} thou shouldest
take my
covenant in thy
mouth?
\par }{\Q \VS{17}Seeing thou
hatest
instruction, and
castest my
words
behind thee.
\par }{\Q \VS{18}When thou
sawest a
thief, then thou
consentedst with him, and hast been
partaker with
adulterers.\FTNT{‡}{{\FR 50.18 }
hast…: Heb. thy portion was with}
\par }{\Q \VS{19}Thou
givest thy
mouth to
evil, and thy
tongue
frameth
deceit.\FTNT{§}{{\FR 50.19 }
givest: Heb. sendest}
\par }{\Q \VS{20}Thou
sittest
{\ADD{and}}
speakest against thy
brother;
thou
slanderest thine own
mother’s
son.
\par }{\Q \VS{21}These
{\ADD{things}} hast thou
done, and I kept
silence; thou
thoughtest that I was
altogether
{\ADD{such an one}} as thyself:
{\ADD{but}} I will
reprove thee, and set
{\ADD{them}} in
order before thine
eyes.
\par }{\Q \VS{22}Now
consider this, ye that
forget
God, lest I tear
{\ADD{you}} in
pieces, and
{\ADD{there be}} none to
deliver.
\par }{\Q \VS{23}Whoso
offereth
praise
glorifieth me: and to him that
ordereth
{\ADD{his}}
conversation
{\ADD{aright}} will I
shew the
salvation of
God.\FTNT{**}{{\FR 50.23 }
that…: Heb. that disposeth his way}

\par }\PsalmChap{51}{\D To the chief
Musician, A
Psalm of
David, when
Nathan the
prophet
came unto him, after he had gone
in to
Bath-sheba.
\par }{\Q \VerseOne{1}Have
mercy upon me, O
God, according to thy
lovingkindness: according unto the
multitude of thy tender
mercies blot
out my
transgressions.
\par }{\Q \VS{2}Wash me
throughly from mine
iniquity, and
cleanse me from my
sin.
\par }{\Q \VS{3}For I
acknowledge my
transgressions: and my
sin
{\ADD{is}}
ever before me.
\par }{\Q \VS{4}Against thee, thee only, have I
sinned, and
done
{\ADD{this}}
evil in thy
sight: that thou mightest be
justified when thou
speakest,
{\ADD{and}} be
clear when thou
judgest.
\par }{\Q \VS{5}Behold, I was
shapen in
iniquity; and in
sin did my
mother
conceive me.\FTNT{*}{{\FR 51.5 }
conceive…: Heb. warm me}
\par }{\Q \VS{6}Behold, thou
desirest
truth in the inward
parts: and in the
hidden
{\ADD{part}} thou shalt make me to
know
wisdom.
\par }{\BB \par }{\Q \VS{7}Purge me with
hyssop, and I shall be
clean:
wash me, and I shall be
whiter than
snow.
\par }{\Q \VS{8}Make me to
hear
joy and
gladness;
{\ADD{that}} the
bones
{\ADD{which}} thou hast
broken may
rejoice.
\par }{\Q \VS{9}Hide thy
face from my
sins, and blot
out all mine
iniquities.
\par }{\Q \VS{10}Create in me a
clean
heart, O
God; and
renew a
right
spirit
within me.\FTNT{†}{{\FR 51.10 }
right: or, constant}
\par }{\Q \VS{11}Cast me not
away from thy
presence; and
take not thy
holy
spirit from me.
\par }{\Q \VS{12}Restore unto me the
joy of thy
salvation; and
uphold me
{\ADD{with thy}}
free
spirit.
\par }{\Q \VS{13}{\ADD{Then}} will I
teach
transgressors thy
ways; and
sinners shall be
converted unto thee.
\par }{\BB \par }{\Q \VS{14}Deliver me from
bloodguiltiness, O
God, thou
God of my
salvation:
{\ADD{and}} my
tongue shall sing
aloud of thy
righteousness.\FTNT{‡}{{\FR 51.14 }
bloodguiltiness: Heb. bloods}
\par }{\Q \VS{15}O
Lord,
open thou my
lips; and my
mouth shall shew
forth thy
praise.
\par }{\Q \VS{16}For thou
desirest not
sacrifice; else would I
give
{\ADD{it:}} thou
delightest not in burnt
offering.\FTNT{§}{{\FR 51.16 }
else…: or, that I should}
\par }{\Q \VS{17}The
sacrifices of
God
{\ADD{are}} a
broken
spirit: a
broken and a
contrite
heart, O
God, thou wilt not
despise.
\par }{\Q \VS{18}Do
good in thy good
pleasure unto
Zion:
build thou the
walls of
Jerusalem.
\par }{\Q \VS{19}Then shalt thou be
pleased with the
sacrifices of
righteousness, with burnt
offering and
whole burnt
offering: then shall they
offer
bullocks upon thine
altar.

\par }\PsalmChap{52}{\D To the chief
Musician,
Maschil,
{\ADD{A Psalm}} of
David, when
Doeg the
Edomite
came and
told
Saul, and
said unto him,
David is
come to the
house of
Ahimelech.
\par }{\Q \VerseOne{1}Why
boastest thou thyself in
mischief, O
mighty man? the
goodness of
God
{\ADD{endureth}}
continually.\FTNT{*}{{\FR 52.1 }
Maschil: or, of instruction}
\par }{\Q \VS{2}Thy
tongue
deviseth
mischiefs; like a
sharp
razor,
working
deceitfully.
\par }{\Q \VS{3}Thou
lovest
evil more than
good;
{\ADD{and}}
lying rather than to
speak
righteousness.
Selah.
\par }{\Q \VS{4}Thou
lovest all
devouring
words, O
{\ADD{thou}}
deceitful
tongue.\FTNT{†}{{\FR 52.4 }
O thou…: or, and the deceitful tongue}
\par }{\Q \VS{5}God shall likewise
destroy thee for
ever, he shall take thee
away, and pluck thee
out of
{\ADD{thy}} dwelling
place, and root thee
out of the
land of the
living.
Selah.\FTNT{‡}{{\FR 52.5 }
destroy…: Heb. beat thee down}
\par }{\BB \par }{\Q \VS{6}The
righteous also shall
see, and
fear, and shall
laugh at him:
\par }{\Q \VS{7}Lo,
{\ADD{this is}} the
man
{\ADD{that}}
made not
God his
strength; but
trusted in the
abundance of his
riches,
{\ADD{and}}
strengthened himself in his
wickedness.\FTNT{§}{{\FR 52.7 }
wickedness: or, substance}
\par }{\Q \VS{8}But I
{\ADD{am}} like a
green olive
tree in the
house of
God: I
trust in the
mercy of
God for
ever and
ever.
\par }{\Q \VS{9}I will
praise thee for
ever, because thou hast
done
{\ADD{it:}} and I will
wait on thy
name; for
{\ADD{it is}}
good before thy
saints.

\par }\PsalmChap{53}{\D To the chief
Musician upon
Mahalath,
Maschil,
{\ADD{A Psalm}} of
David.
\par }{\Q \VerseOne{1}The
fool hath
said in his
heart,
{\ADD{There is}} no
God.
Corrupt are they, and have done
abominable
iniquity:
{\ADD{there is}} none that
doeth
good.\FTNT{*}{{\FR 53.1 }
Maschil: or, of instruction}
\par }{\Q \VS{2}God looked
down from
heaven upon the
children of
men, to
see if there
were
{\ADD{any}} that did
understand, that did
seek
God.
\par }{\Q \VS{3}Every one of them is gone
back: they are
altogether become
filthy;
{\ADD{there is}} none that
doeth
good, no, not
one.
\par }{\Q \VS{4}Have the
workers of
iniquity no
knowledge? who eat
up my
people
{\ADD{as}} they
eat
bread: they have not
called upon
God.
\par }{\Q \VS{5}There were they in
great
fear,
{\ADD{where}} no
fear was: for
God hath
scattered the
bones of him that
encampeth
{\ADD{against}} thee: thou hast put
{\ADD{them}} to
shame, because
God hath
despised them.\FTNT{†}{{\FR 53.5 }
were…: Heb. they feared a fear}
\par }{\Q \VS{6}Oh
that the
salvation of
Israel
{\ADD{were come}} out of
Zion! When
God bringeth
back the
captivity of his
people,
Jacob shall
rejoice,
{\ADD{and}}
Israel shall be
glad.\FTNT{‡}{{\FR 53.6 }
Oh that…: Heb. Who will give salvation, etc}

\par }\PsalmChap{54}{\D To the chief
Musician on
Neginoth,
Maschil,
{\ADD{A Psalm}} of
David, when the
Ziphims
came and
said to
Saul, Doth not
David
hide himself with us?
\par }{\Q \VerseOne{1}Save me, O
God, by thy
name, and
judge me by thy
strength.\FTNT{*}{{\FR 54.1 }
chief…: or, overseer}\FTNT{†}{{\FR 54.1 }
Maschil: or, of instruction}
\par }{\Q \VS{2}Hear my
prayer, O
God; give
ear to the
words of my
mouth.
\par }{\Q \VS{3}For
strangers are risen
up against me, and
oppressors
seek after my
soul: they have not
set
God before them.
Selah.
\par }{\BB \par }{\Q \VS{4}Behold,
God
{\ADD{is}} mine
helper: the
Lord
{\ADD{is}} with them that
uphold my
soul.
\par }{\Q \VS{5}He shall
reward
evil unto mine
enemies: cut them
off in thy
truth.\FTNT{‡}{{\FR 54.5 }
mine…: Heb. those that observe me}
\par }{\Q \VS{6}I will
freely
sacrifice unto thee: I will
praise thy
name, O
{\ND{
LORD}}; for
{\ADD{it is}}
good.
\par }{\Q \VS{7}For he hath
delivered me out of all
trouble: and mine
eye hath
seen
{\ADD{his desire}} upon mine
enemies.

\par }\PsalmChap{55}{\D To the chief
Musician on
Neginoth,
Maschil,
{\ADD{A Psalm}} of
David.
\par }{\Q \VerseOne{1}Give
ear to my
prayer, O
God; and
hide not thyself from my
supplication.\FTNT{*}{{\FR 55.1 }
chief…: or, overseer}\FTNT{†}{{\FR 55.1 }
Maschil: or, of instruction}
\par }{\Q \VS{2}Attend unto me, and
hear me: I
mourn in my
complaint, and make a
noise;
\par }{\Q \VS{3}Because of the
voice of the
enemy,
because of the
oppression of the
wicked: for they
cast
iniquity upon me, and in
wrath they
hate me.
\par }{\Q \VS{4}My
heart is sore
pained
within me: and the
terrors of
death are
fallen upon me.
\par }{\Q \VS{5}Fearfulness and
trembling are
come upon me, and
horror hath
overwhelmed me.\FTNT{‡}{{\FR 55.5 }
overwhelmed: Heb. covered}
\par }{\Q \VS{6}And I
said, Oh
that I had
wings like a
dove!
{\ADD{for then}} would I fly
away, and be at
rest.
\par }{\Q \VS{7}Lo,
{\ADD{then}} would I
wander far
off,
{\ADD{and}}
remain in the
wilderness.
Selah.
\par }{\Q \VS{8}I would
hasten my
escape from the
windy
storm
{\ADD{and}}
tempest.
\par }{\BB \par }{\Q \VS{9}Destroy, O
Lord,
{\ADD{and}}
divide their
tongues: for I have
seen
violence and
strife in the
city.
\par }{\Q \VS{10}Day and
night they go
about it upon the
walls thereof:
mischief also and
sorrow
{\ADD{are}} in the
midst of it.
\par }{\Q \VS{11}Wickedness
{\ADD{is}} in the
midst thereof:
deceit and
guile
depart not from her
streets.
\par }{\Q \VS{12}For
{\ADD{it was}} not an
enemy
{\ADD{that}}
reproached me; then I could have
borne
{\ADD{it:}} neither
{\ADD{was it}} he that
hated me
{\ADD{that}} did
magnify
{\ADD{himself}} against me; then I would have
hid myself from him:
\par }{\Q \VS{13}But
{\ADD{it was}} thou, a
man mine
equal, my
guide, and mine
acquaintance.\FTNT{§}{{\FR 55.13 }
mine equal: Heb. according to my rank}
\par }{\Q \VS{14}We took
sweet
counsel
together,
{\ADD{and}}
walked unto the
house of
God in
company.\FTNT{**}{{\FR 55.14 }
We…: Heb. Who sweetened counsel}
\par }{\Q \VS{15}Let
death
seize upon them,
{\ADD{and}} let them go
down
quick into
hell: for
wickedness
{\ADD{is}} in their
dwellings,
{\ADD{and}}
among them.\FTNT{††}{{\FR 55.15 }
hell: or, the grave}
\par }{\BB \par }{\Q \VS{16}As for me, I will
call upon
God; and the
{\ND{
LORD}} shall
save me.
\par }{\Q \VS{17}Evening, and
morning, and at
noon, will I
pray, and cry
aloud: and he shall
hear my
voice.
\par }{\Q \VS{18}He hath
delivered my
soul in
peace from the
battle
{\ADD{that was}} against me: for there were
many with me.
\par }{\Q \VS{19}God shall
hear, and
afflict them, even he that
abideth of
old.
Selah. Because they have no
changes, therefore they
fear not
God.\FTNT{‡‡}{{\FR 55.19 }
Because…: or, With whom also there be no changes, yet they}
\par }{\Q \VS{20}He hath put
forth his
hands against such as be at
peace with him: he hath
broken his
covenant.\FTNT{§§}{{\FR 55.20 }
broken: Heb. profaned}
\par }{\Q \VS{21}{\ADD{The words}} of his
mouth were
smoother than
butter, but
war
{\ADD{was}} in his
heart: his
words were
softer than
oil, yet
{\ADD{were}} they drawn
swords.
\par }{\Q \VS{22}Cast thy
burden upon the
{\ND{
LORD}}, and he shall
sustain thee: he shall
never
suffer the
righteous to be
moved.\FTNT{***}{{\FR 55.22 }
burden: or, gift}
\par }{\Q \VS{23}But thou, O
God, shalt bring them
down into the
pit of
destruction:
bloody and
deceitful
men shall not live out
half their
days; but I will
trust in thee.\FTNT{†††}{{\FR 55.23 }
bloody…: Heb. men of bloods and deceit}\FTNT{‡‡‡}{{\FR 55.23 }
shall…: Heb. shall not half their days}

\par }\PsalmChap{56}{\D To the chief
Musician upon
Jonath-elem-rechokim,
Michtam of
David, when the
Philistines
took him in
Gath.
\par }{\Q \VerseOne{1}Be
merciful unto me, O
God: for
man would swallow me
up; he
fighting
daily
oppresseth me.\FTNT{*}{{\FR 56.1 }
Michtam…: or, A golden Psalm of David}
\par }{\Q \VS{2}Mine
enemies would
daily
swallow
{\ADD{me}} up: for
{\ADD{they be}}
many that
fight against me, O thou most
High.\FTNT{†}{{\FR 56.2 }
enemies: Heb. observers}
\par }{\Q \VS{3}What
time I am
afraid, I will
trust in thee.
\par }{\Q \VS{4}In
God I will
praise his
word, in
God I have put my
trust; I will not
fear what
flesh can
do unto me.
\par }{\Q \VS{5}Every
day they
wrest my
words: all their
thoughts
{\ADD{are}} against me for
evil.
\par }{\Q \VS{6}They gather themselves
together, they
hide themselves, they
mark my
steps, when they
wait for my
soul.
\par }{\Q \VS{7}Shall they
escape by
iniquity? in
{\ADD{thine}}
anger cast
down the
people, O
God.
\par }{\BB \par }{\Q \VS{8}Thou
tellest my
wanderings:
put thou my
tears into thy
bottle:
{\ADD{are they}} not in thy
book?
\par }{\Q \VS{9}When I
cry
{\ADD{unto thee}}, then shall mine
enemies
turn
back: this I
know; for
God
{\ADD{is}} for me.
\par }{\Q \VS{10}In
God will I
praise
{\ADD{his}}
word: in the
{\ND{
LORD}} will I
praise
{\ADD{his}}
word.
\par }{\Q \VS{11}In
God have I put my
trust: I will not be
afraid what
man can
do unto me.
\par }{\Q \VS{12}Thy
vows
{\ADD{are}} upon me, O
God: I will
render
praises unto thee.
\par }{\Q \VS{13}For thou hast
delivered my
soul from
death:
{\ADD{wilt}} not
{\ADD{thou deliver}} my
feet from
falling, that I may
walk
before
God in the
light of the
living?

\par }\PsalmChap{57}{\D To the chief
Musician,
Al-taschith,
Michtam of
David, when he
fled
from
Saul in the
cave.
\par }{\Q \VerseOne{1}Be
merciful unto me, O
God, be
merciful unto me: for my
soul
trusteth in thee: yea, in the
shadow of thy
wings will I make my
refuge, until
{\ADD{these}}
calamities be
overpast.\FTNT{*}{{\FR 57.1 }
Al-taschith: or, Destroy not}\FTNT{†}{{\FR 57.1 }
Michtam: or, A golden Psalm}
\par }{\Q \VS{2}I will
cry unto
God most
high; unto
God that
performeth
{\ADD{all things}} for me.
\par }{\Q \VS{3}He shall
send from
heaven, and
save me
{\ADD{from}} the
reproach of him that would swallow me
up.
Selah.
God shall send
forth his
mercy and his
truth.\FTNT{‡}{{\FR 57.3 }
from the…: or, he reproacheth him that}
\par }{\Q \VS{4}My
soul
{\ADD{is}}
among
lions:
{\ADD{and}} I
lie
{\ADD{even among}} them that are set on
fire,
{\ADD{even}} the
sons of
men, whose
teeth
{\ADD{are}}
spears and
arrows, and their
tongue a
sharp
sword.
\par }{\Q \VS{5}Be thou
exalted, O
God, above the
heavens;
{\ADD{let}} thy
glory
{\ADD{be}} above all the
earth.
\par }{\Q \VS{6}They have
prepared a
net for my
steps; my
soul is bowed
down: they have
digged a
pit
before me, into the
midst whereof they are
fallen
{\ADD{themselves}}.
Selah.
\par }{\BB \par }{\Q \VS{7}My
heart is
fixed, O
God, my
heart is
fixed: I will
sing and give
praise.\FTNT{§}{{\FR 57.7 }
fixed: or, prepared}
\par }{\Q \VS{8}Awake
up, my
glory;
awake,
psaltery and
harp: I
{\ADD{myself}} will
awake
early.
\par }{\Q \VS{9}I will
praise thee, O
Lord, among the
people: I will
sing unto thee among the
nations.
\par }{\Q \VS{10}For thy
mercy
{\ADD{is}}
great unto the
heavens, and thy
truth unto the
clouds.
\par }{\Q \VS{11}Be thou
exalted, O
God, above the
heavens:
{\ADD{let}} thy
glory
{\ADD{be}} above all the
earth.

\par }\PsalmChap{58}{\D To the chief
Musician,
Al-taschith,
Michtam of
David.
\par }{\Q \VerseOne{1}Do ye
indeed
speak
righteousness, O
congregation? do ye
judge
uprightly, O ye
sons of
men?\FTNT{*}{{\FR 58.1 }
Al-taschith…: or, Destroy not}\FTNT{†}{{\FR 58.1 }
Michtam: or, A golden Psalm}
\par }{\Q \VS{2}Yea, in
heart ye
work
wickedness; ye
weigh the
violence of your
hands in the
earth.
\par }{\Q \VS{3}The
wicked are
estranged from the
womb: they go
astray as soon as they be
born,
speaking
lies.\FTNT{‡}{{\FR 58.3 }
as soon…: Heb. from the belly}
\par }{\Q \VS{4}Their
poison
{\ADD{is}}
like the
poison of a
serpent:
{\ADD{they are}} like the
deaf
adder
{\ADD{that}}
stoppeth her
ear;\FTNT{§}{{\FR 58.4 }
like the poison: Heb. according to the likeness, etc}\FTNT{**}{{\FR 58.4 }
adder: or, asp}
\par }{\Q \VS{5}Which will not
hearken to the
voice of
charmers,
charming never so
wisely.\FTNT{††}{{\FR 58.5 }
charming…: or, be the charmer never so cunning}
\par }{\BB \par }{\Q \VS{6}Break their
teeth, O
God, in their
mouth: break
out the great
teeth of the young
lions, O
{\ND{
LORD}}.
\par }{\Q \VS{7}Let them melt
away as
waters
{\ADD{which}} run
continually:
{\ADD{when}} he
bendeth
{\ADD{his bow to shoot}} his
arrows, let them be as cut in
pieces.
\par }{\Q \VS{8}As a
snail
{\ADD{which}}
melteth, let
{\ADD{every one of them}} pass
away:
{\ADD{like}} the untimely
birth of a
woman,
{\ADD{that}} they may not
see the
sun.
\par }{\Q \VS{9}Before your
pots can
feel the
thorns, he shall take them away as with a
whirlwind, both
living, and in
{\ADD{his}}
wrath.\FTNT{‡‡}{{\FR 58.9 }
both…: Heb. as living as wrath}
\par }{\Q \VS{10}The
righteous shall
rejoice when he
seeth the
vengeance: he shall
wash his
feet in the
blood of the
wicked.
\par }{\Q \VS{11}So that a
man shall
say, Verily
{\ADD{there is}} a
reward for the
righteous: verily he
is a
God that
judgeth in the
earth.\FTNT{§§}{{\FR 58.11 }
a reward…: Heb. fruit of the, etc}

\par }\PsalmChap{59}{\D To the chief
Musician,
Al-taschith,
Michtam of
David; when
Saul
sent, and they
watched the
house to kill
him.
\par }{\Q \VerseOne{1}Deliver me from mine
enemies, O my
God:
defend me from them that rise
up against me.\FTNT{*}{{\FR 59.1 }
Al-taschith…: or, Destroy not}\FTNT{†}{{\FR 59.1 }
Michtam: or, A golden Psalm}\FTNT{‡}{{\FR 59.1 }
defend…: Heb. set me on high}
\par }{\Q \VS{2}Deliver me from the
workers of
iniquity, and
save me from
bloody
men.
\par }{\Q \VS{3}For, lo, they lie in
wait for my
soul: the
mighty are
gathered against me; not
{\ADD{for}} my
transgression, nor
{\ADD{for}} my
sin, O
{\ND{
LORD}}.
\par }{\Q \VS{4}They
run and
prepare themselves without
{\ADD{my}}
fault:
awake to
help me, and
behold.\FTNT{§}{{\FR 59.4 }
help: Heb. meet}
\par }{\Q \VS{5}Thou therefore, O
{\ND{
LORD}}
God of
hosts, the
God of
Israel,
awake to
visit all the
heathen: be not
merciful to any
wicked
transgressors.
Selah.
\par }{\Q \VS{6}They
return at
evening: they make a
noise like a
dog, and go round
about the
city.
\par }{\Q \VS{7}Behold, they belch
out with their
mouth:
swords
{\ADD{are}} in their
lips: for who,
{\ADD{say they}}, doth
hear?
\par }{\BB \par }{\Q \VS{8}But thou, O
{\ND{
LORD}}, shalt
laugh at them; thou shalt have all the
heathen in
derision.
\par }{\Q \VS{9}{\ADD{Because of}} his
strength will I
wait upon thee: for
God
{\ADD{is}} my
defence.\FTNT{**}{{\FR 59.9 }
defence: Heb. high place}
\par }{\Q \VS{10}The
God of my
mercy shall
prevent me:
God shall let me
see
{\ADD{my desire}} upon mine
enemies.\FTNT{††}{{\FR 59.10 }
enemies: Heb. observers}
\par }{\Q \VS{11}Slay them not, lest my
people
forget:
scatter them by thy
power; and bring them
down, O
Lord our
shield.
\par }{\Q \VS{12}{\ADD{For}} the
sin of their
mouth
{\ADD{and}} the
words of their
lips let them even be
taken in their
pride: and for
cursing and
lying
{\ADD{which}} they
speak.
\par }{\Q \VS{13}Consume
{\ADD{them}} in
wrath,
consume
{\ADD{them}}, that they
{\ADD{may}} not
{\ADD{be:}} and let them
know that
God
ruleth in
Jacob unto the
ends of the
earth.
Selah.
\par }{\Q \VS{14}And at
evening let them
return;
{\ADD{and}} let them make a
noise like a
dog, and go round
about the
city.
\par }{\Q \VS{15}Let them wander up and
down for
meat, and
grudge if they be not
satisfied.\FTNT{‡‡}{{\FR 59.15 }
for meat: Heb. to eat}\FTNT{§§}{{\FR 59.15 }
and grudge…: or, if they be not satisfied, then they will stay all night}
\par }{\Q \VS{16}But I will
sing of thy
power; yea, I will sing
aloud of thy
mercy in the
morning: for thou hast been my
defence and
refuge in the
day of my
trouble.
\par }{\Q \VS{17}Unto thee, O my
strength, will I
sing: for
God
{\ADD{is}} my
defence,
{\ADD{and}} the
God of my
mercy.

\par }\PsalmChap{60}{\D To the chief
Musician upon
Shushan-eduth,
Michtam of
David, to
teach; when he
strove with
Aram-naharaim and with
Aram-zobah, when
Joab
returned, and
smote of
Edom in the
valley of
salt
twelve
thousand.
\par }{\Q \VerseOne{1}O
God, thou hast cast us
off, thou hast
scattered us, thou hast been
displeased; O turn thyself to us
again.\FTNT{*}{{\FR 60.1 }
Michtam: or, A golden Psalm}\FTNT{†}{{\FR 60.1 }
scattered: Heb. broken}
\par }{\Q \VS{2}Thou hast made the
earth to
tremble; thou hast
broken it:
heal the
breaches thereof; for it
shaketh.
\par }{\Q \VS{3}Thou hast
shewed thy
people hard
things: thou hast made us to
drink the
wine of
astonishment.
\par }{\Q \VS{4}Thou hast
given a
banner to them that
fear thee, that it may be
displayed
because of the
truth.
Selah.
\par }{\Q \VS{5}That thy
beloved may be
delivered;
save
{\ADD{with}} thy right
hand, and
hear me.
\par }{\BB \par }{\Q \VS{6}God hath
spoken in his
holiness; I will
rejoice, I will
divide
Shechem, and mete
out the
valley of
Succoth.
\par }{\Q \VS{7}Gilead
{\ADD{is}} mine, and
Manasseh
{\ADD{is}} mine;
Ephraim also
{\ADD{is}} the
strength of mine
head;
Judah
{\ADD{is}} my
lawgiver;
\par }{\Q \VS{8}Moab
{\ADD{is}} my
washpot; over
Edom will I cast
out my
shoe:
Philistia,
triumph thou because of me.\FTNT{‡}{{\FR 60.8 }
triumph…: or, triumph thou over me: (by an irony)}
\par }{\Q \VS{9}Who will
bring me
{\ADD{into}} the
strong
city? who will
lead me into
Edom?\FTNT{§}{{\FR 60.9 }
strong…: Heb. city of strength?}
\par }{\Q \VS{10}{\ADD{Wilt}} not thou, O
God,
{\ADD{which}} hadst cast us
off? and
{\ADD{thou}}, O
God,
{\ADD{which}} didst not go
out with our
armies?
\par }{\Q \VS{11}Give us
help from
trouble: for
vain
{\ADD{is}} the
help of
man.\FTNT{**}{{\FR 60.11 }
help of man: Heb. salvation, etc}
\par }{\Q \VS{12}Through
God we shall
do
valiantly: for he
{\ADD{it is that}} shall tread
down our
enemies.

\par }\PsalmChap{61}{\D To the chief
Musician upon
Neginah,
{\ADD{A Psalm}} of
David.
\par }{\Q \VerseOne{1}Hear my
cry, O
God;
attend unto my
prayer.
\par }{\Q \VS{2}From the
end of the
earth will I
cry unto thee, when my
heart is
overwhelmed:
lead me to the
rock
{\ADD{that}} is
higher than I.
\par }{\Q \VS{3}For thou hast been a
shelter for me,
{\ADD{and}} a
strong
tower
from the
enemy.
\par }{\Q \VS{4}I will
abide in thy
tabernacle for
ever: I will
trust in the
covert of thy
wings.
Selah.\FTNT{*}{{\FR 61.4 }
trust: or, make my refuge}
\par }{\BB \par }{\Q \VS{5}For thou, O
God, hast
heard my
vows: thou hast
given
{\ADD{me}} the
heritage of those that
fear thy
name.
\par }{\Q \VS{6}Thou wilt
prolong the
king’s
life:
{\ADD{and}} his
years as
many
generations.\FTNT{†}{{\FR 61.6 }
Thou…: Heb. Thou shalt add days to the days of the king}\FTNT{‡}{{\FR 61.6 }
as…: Heb. as generation and generation}
\par }{\Q \VS{7}He shall
abide
before
God for
ever: O
prepare
mercy and
truth,
{\ADD{which}} may
preserve him.
\par }{\Q \VS{8}So will I
sing praise unto thy
name for
ever, that I may
daily
perform my
vows.

\par }\PsalmChap{62}{\D To the chief
Musician, to
Jeduthun, A
Psalm of
David.
\par }{\Q \VerseOne{1}Truly my
soul
waiteth upon
God: from him
{\ADD{cometh}} my
salvation.\FTNT{*}{{\FR 62.1 }
Truly: or, Only}\FTNT{†}{{\FR 62.1 }
waiteth: Heb. is silent}
\par }{\Q \VS{2}He only
{\ADD{is}} my
rock and my
salvation;
{\ADD{he is}} my
defence; I shall not be
greatly
moved.\FTNT{‡}{{\FR 62.2 }
defence: Heb. high place}
\par }{\Q \VS{3}How long will ye imagine
mischief against a
man? ye shall be
slain all of you: as a
bowing
wall
{\ADD{shall ye be, and as}} a
tottering
fence.
\par }{\Q \VS{4}They only
consult to cast
{\ADD{him}}
down from his
excellency: they
delight in
lies: they
bless with their
mouth, but they
curse
inwardly.
Selah.\FTNT{§}{{\FR 62.4 }
inwardly: Heb. in their inward parts}
\par }{\Q \VS{5}My
soul,
wait thou only upon
God; for my
expectation
{\ADD{is}} from him.
\par }{\Q \VS{6}He only
{\ADD{is}} my
rock and my
salvation:
{\ADD{he is}} my
defence; I shall not be
moved.
\par }{\Q \VS{7}In
God
{\ADD{is}} my
salvation and my
glory: the
rock of my
strength,
{\ADD{and}} my
refuge,
{\ADD{is}} in
God.
\par }{\BB \par }{\Q \VS{8}Trust in him at all
times;
{\ADD{ye}}
people, pour
out your
heart
before him:
God
{\ADD{is}} a
refuge for us.
Selah.
\par }{\Q \VS{9}Surely
men of low degree
{\ADD{are}}
vanity,
{\ADD{and}}
men of high degree
{\ADD{are}} a
lie: to be
laid in the
balance, they
{\ADD{are}}
altogether
{\ADD{lighter}} than
vanity.\FTNT{**}{{\FR 62.9 }
altogether: or, alike}
\par }{\Q \VS{10}Trust not in
oppression, and become not
vain in
robbery: if
riches
increase,
set not your
heart
{\ADD{upon them}}.
\par }{\Q \VS{11}God hath
spoken
once;
twice have I
heard
this; that
power
{\ADD{belongeth}} unto
God.\FTNT{††}{{\FR 62.11 }
power: or, strength}
\par }{\Q \VS{12}Also unto thee, O
Lord,
{\ADD{belongeth}}
mercy: for thou
renderest to every
man according to his
work.

\par }\PsalmChap{63}{\D A
Psalm of
David, when he was in the
wilderness of
Judah.
\par }{\Q \VerseOne{1}O
God, thou
{\ADD{art}} my
God; early will I
seek thee: my
soul
thirsteth for thee, my
flesh
longeth for thee in a
dry and
thirsty
land, where
no
water is;\FTNT{*}{{\FR 63.1 }
thirsty: Heb. weary}\FTNT{†}{{\FR 63.1 }
where…: without water}
\par }{\Q \VS{2}To
see thy
power and thy
glory, so
{\ADD{as}} I have
seen thee in the
sanctuary.
\par }{\BB \par }{\Q \VS{3}Because thy
lovingkindness
{\ADD{is}}
better than
life, my
lips shall
praise thee.
\par }{\Q \VS{4}Thus will I
bless thee while I
live: I will lift
up my
hands in thy
name.
\par }{\Q \VS{5}My
soul shall be
satisfied as
{\ADD{with}}
marrow and
fatness; and my
mouth shall
praise
{\ADD{thee}} with
joyful
lips:\FTNT{‡}{{\FR 63.5 }
marrow: Heb. fatness}
\par }{\Q \VS{6}When I
remember thee upon my
bed,
{\ADD{and}}
meditate on thee in the
{\ADD{night}}
watches.
\par }{\BB \par }{\Q \VS{7}Because thou hast been my
help, therefore in the
shadow of thy
wings will I
rejoice.
\par }{\Q \VS{8}My
soul followeth
hard
after thee: thy right
hand
upholdeth me.
\par }{\Q \VS{9}But those
{\ADD{that}}
seek my
soul, to
destroy
{\ADD{it}}, shall
go into the lower
parts of the
earth.
\par }{\Q \VS{10}They shall
fall by the
sword: they shall be a
portion for
foxes.\FTNT{§}{{\FR 63.10 }
They shall fall…: Heb. They shall make him run out like water by the hands of}
\par }{\Q \VS{11}But the
king shall
rejoice in
God; every one that
sweareth by him shall
glory: but the
mouth of them that
speak
lies shall be
stopped.

\par }\PsalmChap{64}{\D To the chief
Musician, A
Psalm of
David.
\par }{\Q \VerseOne{1}Hear my
voice, O
God, in my
prayer:
preserve my
life from
fear of the
enemy.
\par }{\Q \VS{2}Hide me from the secret
counsel of the
wicked; from the
insurrection of the
workers of
iniquity:
\par }{\Q \VS{3}Who
whet their
tongue like a
sword,
{\ADD{and}}
bend
{\ADD{their bows to shoot}} their
arrows,
{\ADD{even}}
bitter
words:
\par }{\Q \VS{4}That they may
shoot in
secret at the
perfect:
suddenly do they
shoot at him, and
fear not.
\par }{\Q \VS{5}They
encourage themselves
{\ADD{in}} an
evil
matter: they
commune of
laying
snares
privily; they
say, Who shall
see them?\FTNT{*}{{\FR 64.5 }
matter: or, speech}\FTNT{†}{{\FR 64.5 }
of…: Heb. to hide his snares}
\par }{\Q \VS{6}They search
out
iniquities; they
accomplish a
diligent
search: both the
inward
{\ADD{thought}} of every
one
{\ADD{of them}}, and the
heart,
{\ADD{is}}
deep.\FTNT{‡}{{\FR 64.6 }
they…: or, we are consumed by that which they have throughly searched}\FTNT{§}{{\FR 64.6 }
a diligent…: Heb. a search searched}
\par }{\BB \par }{\Q \VS{7}But
God shall
shoot at them
{\ADD{with}} an
arrow;
suddenly shall they be
wounded.\FTNT{**}{{\FR 64.7 }
shall they…: Heb. their wound shall be}
\par }{\Q \VS{8}So they shall make their own
tongue to
fall upon themselves: all that
see them shall flee
away.
\par }{\Q \VS{9}And all
men shall
fear, and shall
declare the
work of
God; for they shall wisely
consider of his
doing.
\par }{\Q \VS{10}The
righteous shall be
glad in the
{\ND{
LORD}}, and shall
trust in him; and all the
upright in
heart shall
glory.

\par }\PsalmChap{65}{\D To the chief
Musician, A
Psalm
{\ADD{and}}
Song of
David.
\par }{\Q \VerseOne{1}Praise
waiteth for thee, O
God, in
Sion: and unto thee shall the
vow be
performed.\FTNT{*}{{\FR 65.1 }
waiteth: Heb. is silent}
\par }{\Q \VS{2}O thou that
hearest
prayer, unto thee shall all
flesh
come.
\par }{\Q \VS{3}Iniquities
prevail against me:
{\ADD{as for}} our
transgressions, thou shalt purge them
away.\FTNT{†}{{\FR 65.3 }
Iniquities: Heb. Words, or, Matters of iniquities}
\par }{\Q \VS{4}Blessed
{\ADD{is the man whom}} thou
choosest, and causest to
approach
{\ADD{unto thee, that}} he may
dwell in thy
courts: we shall be
satisfied with the
goodness of thy
house,
{\ADD{even}} of thy
holy
temple.
\par }{\Q \VS{5}{\ADD{By}} terrible
things in
righteousness wilt thou
answer us, O
God of our
salvation;
{\ADD{who art}} the
confidence of all the
ends of the
earth, and of them that are afar
off
{\ADD{upon}} the
sea:
\par }{\BB \par }{\Q \VS{6}Which by his
strength setteth
fast the
mountains;
{\ADD{being}}
girded with
power:
\par }{\Q \VS{7}Which
stilleth the
noise of the
seas, the
noise of their
waves, and the
tumult of the
people.
\par }{\Q \VS{8}They also that
dwell in the uttermost
parts are
afraid at thy
tokens: thou makest the
outgoings of the
morning and
evening to
rejoice.\FTNT{‡}{{\FR 65.8 }
rejoice: or, sing}
\par }{\Q \VS{9}Thou
visitest the
earth, and
waterest it: thou
greatly
enrichest it with the
river of
God,
{\ADD{which}} is
full of
water: thou
preparest them
corn, when thou hast so
provided for it.\FTNT{§}{{\FR 65.9 }
and…: or, after thou hadst made it to desire rain}
\par }{\Q \VS{10}Thou
waterest the
ridges thereof
abundantly: thou
settlest the
furrows thereof: thou makest it
soft with
showers: thou
blessest the
springing thereof.\FTNT{**}{{\FR 65.10 }
settlest: or, causest rain to descend into}\FTNT{††}{{\FR 65.10 }
makest…: Heb. dissolvest it}
\par }{\Q \VS{11}Thou
crownest the
year with thy
goodness; and thy
paths
drop
fatness.\FTNT{‡‡}{{\FR 65.11 }
with: Heb. of}
\par }{\Q \VS{12}They
drop
{\ADD{upon}} the
pastures of the
wilderness: and the little
hills
rejoice on every
side.\FTNT{§§}{{\FR 65.12 }
rejoice…: Heb. are girded with joy}
\par }{\Q \VS{13}The
pastures are
clothed with
flocks; the
valleys also are covered
over with
corn; they shout for
joy, they also
sing.

\par }\PsalmChap{66}{\D To the chief
Musician, A
Song
{\ADD{or}}
Psalm.
\par }{\Q \VerseOne{1}Make a joyful
noise unto
God, all ye
lands:\FTNT{*}{{\FR 66.1 }
all…: Heb. all the earth}
\par }{\Q \VS{2}Sing
forth the
honour of his
name:
make his
praise
glorious.
\par }{\Q \VS{3}Say unto
God, How
terrible
{\ADD{art thou in}} thy
works! through the
greatness of thy
power shall thine
enemies
submit themselves unto thee.\FTNT{†}{{\FR 66.3 }
submit…: or, yield feigned obedience: Heb. lie}
\par }{\Q \VS{4}All the
earth shall
worship thee, and shall
sing unto thee; they shall
sing
{\ADD{to}} thy
name.
Selah.
\par }{\Q \VS{5}Come and
see the
works of
God:
{\ADD{he is}}
terrible
{\ADD{in his}}
doing toward the
children of
men.
\par }{\Q \VS{6}He
turned the
sea into
dry
{\ADD{land:}} they
went through the
flood on
foot: there did we
rejoice in him.
\par }{\Q \VS{7}He
ruleth by his
power for
ever; his
eyes
behold the
nations: let not the
rebellious
exalt themselves.
Selah.
\par }{\BB \par }{\Q \VS{8}O
bless our
God, ye
people, and make the
voice of his
praise to be
heard:
\par }{\Q \VS{9}Which
holdeth our
soul in
life, and
suffereth not our
feet to be
moved.\FTNT{‡}{{\FR 66.9 }
holdeth: Heb. putteth}
\par }{\Q \VS{10}For thou, O
God, hast
proved us: thou hast
tried us, as
silver is
tried.
\par }{\Q \VS{11}Thou
broughtest us into the
net; thou
laidst
affliction upon our
loins.
\par }{\Q \VS{12}Thou hast caused
men to
ride over our
heads; we
went through
fire and through
water: but thou broughtest us
out into a
wealthy
{\ADD{place}}.\FTNT{§}{{\FR 66.12 }
wealthy: Heb. moist}
\par }{\BB \par }{\Q \VS{13}I will
go into thy
house with burnt
offerings: I will
pay thee my
vows,
\par }{\Q \VS{14}Which my
lips have
uttered, and my
mouth hath
spoken, when I was in
trouble.\FTNT{**}{{\FR 66.14 }
uttered: Heb. opened}
\par }{\Q \VS{15}I will
offer unto thee burnt
sacrifices of
fatlings, with the
incense of
rams; I will
offer
bullocks with
goats.
Selah.\FTNT{††}{{\FR 66.15 }
fatlings: Heb. marrow}
\par }{\Q \VS{16}Come
{\ADD{and}}
hear, all ye that
fear
God, and I will
declare what he hath
done for my
soul.
\par }{\Q \VS{17}I
cried unto him with my
mouth, and he was
extolled
with my
tongue.
\par }{\Q \VS{18}If I
regard
iniquity in my
heart, the
Lord will not
hear
{\ADD{me:}}
\par }{\Q \VS{19}{\ADD{But}}
verily
God hath
heard
{\ADD{me;}} he hath
attended to the
voice of my
prayer.
\par }{\Q \VS{20}Blessed
{\ADD{be}}
God, which hath not turned
away my
prayer, nor his
mercy from me.

\par }\PsalmChap{67}{\D To the chief
Musician on
Neginoth, A
Psalm
{\ADD{or}}
Song.
\par }{\Q \VerseOne{1}God be
merciful unto us, and
bless us;
{\ADD{and}} cause his
face to
shine upon us;
Selah.\FTNT{*}{{\FR 67.1 }
chief…: or, overseer}\FTNT{†}{{\FR 67.1 }
upon: Heb. with}
\par }{\Q \VS{2}That thy
way may be
known upon
earth, thy saving
health among all
nations.
\par }{\Q \VS{3}Let the
people
praise thee, O
God; let all the
people
praise thee.
\par }{\Q \VS{4}O let the
nations be
glad and sing for
joy: for thou shalt
judge the
people
righteously, and
govern the
nations upon
earth.
Selah.\FTNT{‡}{{\FR 67.4 }
govern: Heb. lead}
\par }{\Q \VS{5}Let the
people
praise thee, O
God; let all the
people
praise thee.
\par }{\Q \VS{6}{\ADD{Then}} shall the
earth
yield her
increase;
{\ADD{and}}
God,
{\ADD{even}} our own
God, shall
bless us.
\par }{\Q \VS{7}God shall
bless us; and all the
ends of the
earth shall
fear him.

\par }\PsalmChap{68}{\D To the chief
Musician, A
Psalm
{\ADD{or}}
Song of
David.
\par }{\Q \VerseOne{1}Let
God
arise, let his
enemies be
scattered: let them also that
hate him
flee
before him.\FTNT{*}{{\FR 68.1 }
before…: Heb. from his face}
\par }{\Q \VS{2}As
smoke is driven
away,
{\ADD{so}} drive
{\ADD{them}}
away: as
wax
melteth
before the
fire,
{\ADD{so}} let the
wicked
perish at the
presence of
God.
\par }{\Q \VS{3}But let the
righteous be
glad; let them
rejoice
before
God: yea, let them
exceedingly
rejoice.\FTNT{†}{{\FR 68.3 }
exceedingly…: Heb. rejoice with gladness}
\par }{\Q \VS{4}Sing unto
God, sing
praises to his
name:
extol him that
rideth upon the
heavens by his
name
{\ND{
Jah}}, and
rejoice
before him.
\par }{\Q \VS{5}A
father of the
fatherless, and a
judge of the
widows,
{\ADD{is}}
God in his
holy
habitation.
\par }{\Q \VS{6}God
setteth the
solitary in
families: he bringeth
out those which are
bound with
chains: but the
rebellious
dwell in a
dry
{\ADD{land}}.\FTNT{‡}{{\FR 68.6 }
in families: Heb. in a house}
\par }{\BB \par }{\Q \VS{7}O
God, when thou wentest
forth
before thy
people, when thou didst
march through the
wilderness;
Selah:
\par }{\Q \VS{8}The
earth
shook, the
heavens also
dropped at the
presence of
God:
{\ADD{even}}
Sinai
itself
{\ADD{was moved}} at the
presence of
God, the
God of
Israel.
\par }{\Q \VS{9}Thou, O
God, didst
send a
plentiful
rain, whereby thou didst
confirm thine
inheritance, when it was
weary.\FTNT{§}{{\FR 68.9 }
send: Heb. shake out}\FTNT{**}{{\FR 68.9 }
confirm: Heb. confirm it}
\par }{\Q \VS{10}Thy
congregation hath
dwelt therein: thou, O
God, hast
prepared of thy
goodness for the
poor.
\par }{\Q \VS{11}The
Lord
gave the
word:
great
{\ADD{was}} the
company of those that
published
{\ADD{it}}.\FTNT{††}{{\FR 68.11 }
company: Heb. army}
\par }{\Q \VS{12}Kings of
armies did
flee
apace: and she that
tarried at
home
divided the
spoil.\FTNT{‡‡}{{\FR 68.12 }
did…: Heb. did flee, did flee}
\par }{\Q \VS{13}Though ye have
lien among the
pots,
{\ADD{yet shall ye be as}} the
wings of a
dove
covered with
silver, and her
feathers with
yellow
gold.
\par }{\Q \VS{14}When the
Almighty
scattered
kings in it, it was
{\ADD{white}} as
snow in
Salmon.\FTNT{§§}{{\FR 68.14 }
in it…: or, for her, she}
\par }{\BB \par }{\Q \VS{15}The
hill of
God
{\ADD{is as}} the
hill of
Bashan; an
high
hill
{\ADD{as}} the
hill of
Bashan.
\par }{\Q \VS{16}Why
leap ye, ye
high
hills?
{\ADD{this is}} the
hill
{\ADD{which}}
God
desireth to dwell
in; yea, the
{\ND{
LORD}} will
dwell
{\ADD{in it}} for
ever.
\par }{\Q \VS{17}The
chariots of
God
{\ADD{are}} twenty
thousand,
{\ADD{even}}
thousands of
angels: the
Lord
{\ADD{is}} among them,
{\ADD{as in}}
Sinai, in the
holy
{\ADD{place}}.\FTNT{***}{{\FR 68.17 }
even…: or, even many thousands}
\par }{\Q \VS{18}Thou hast
ascended on
high, thou hast led
captivity
captive: thou hast
received
gifts for
men; yea,
{\ADD{for}} the
rebellious also, that the
{\ND{
LORD}}
God might
dwell
{\ADD{among them}}.\FTNT{†††}{{\FR 68.18 }
for men: Heb. in the man}
\par }{\Q \VS{19}Blessed
{\ADD{be}} the
Lord,
{\ADD{who}}
daily
loadeth us
{\ADD{with benefits, even}} the
God of our
salvation.
Selah.
\par }{\Q \VS{20}{\ADD{He that is}} our
God
{\ADD{is}} the
God of
salvation; and unto
{\ND{
GOD}} the
Lord
{\ADD{belong}} the
issues from
death.
\par }{\Q \VS{21}But
God shall
wound the
head of his
enemies,
{\ADD{and}} the
hairy
scalp of such an one as goeth on
still in his
trespasses.
\par }{\BB \par }{\Q \VS{22}The
Lord
said, I will bring
again from
Bashan, I will bring
{\ADD{my people}}
again from the
depths of the
sea:
\par }{\Q \VS{23}That thy
foot may be
dipped in the
blood of
{\ADD{thine}}
enemies,
{\ADD{and}} the
tongue of thy
dogs in the same.\FTNT{‡‡‡}{{\FR 68.23 }
dipped: or, red}
\par }{\Q \VS{24}They have
seen thy
goings, O
God;
{\ADD{even}} the
goings of my
God, my
King, in the
sanctuary.
\par }{\Q \VS{25}The
singers went
before, the players on
instruments
{\ADD{followed}}
after;
among
{\ADD{them were}} the
damsels playing with
timbrels.
\par }{\Q \VS{26}Bless ye
God in the
congregations,
{\ADD{even}} the
Lord, from the
fountain of
Israel.\FTNT{§§§}{{\FR 68.26 }
from…: or, ye that are of the fountain of Israel}
\par }{\Q \VS{27}There
{\ADD{is}}
little
Benjamin
{\ADD{with}} their
ruler, the
princes of
Judah
{\ADD{and}} their
council, the
princes of
Zebulun,
{\ADD{and}} the
princes of
Naphtali.\FTNT{*}{{\FR 68.27 }
and their…: or, with their company}
\par }{\Q \VS{28}Thy
God hath
commanded thy
strength:
strengthen, O
God, that
which thou hast
wrought for us.
\par }{\Q \VS{29}Because of thy
temple at
Jerusalem shall
kings
bring
presents unto thee.
\par }{\Q \VS{30}Rebuke the
company of
spearmen, the
multitude of the
bulls, with the
calves of the
people,
{\ADD{till every one}}
submit himself with
pieces of
silver:
scatter thou the
people
{\ADD{that}}
delight in
war.\FTNT{†}{{\FR 68.30 }
the company…: or, the beasts of the reeds}\FTNT{‡}{{\FR 68.30 }
scatter…: or, he scattereth}
\par }{\Q \VS{31}Princes shall come
out of
Egypt;
Ethiopia shall soon stretch
out her
hands unto
God.
\par }{\BB \par }{\Q \VS{32}Sing unto
God, ye
kingdoms of the
earth; O sing
praises unto the
Lord;
Selah:
\par }{\Q \VS{33}To him that
rideth upon the
heavens of
heavens,
{\ADD{which were}} of
old; lo, he doth send
out his
voice,
{\ADD{and that}} a
mighty
voice.\FTNT{§}{{\FR 68.33 }
send…: Heb. give}
\par }{\Q \VS{34}Ascribe ye
strength unto
God: his
excellency
{\ADD{is}} over
Israel, and his
strength
{\ADD{is}} in the
clouds.\FTNT{**}{{\FR 68.34 }
clouds: or, heavens}
\par }{\Q \VS{35}O
God,
{\ADD{thou art}}
terrible out of thy holy
places: the
God of
Israel
{\ADD{is}} he that
giveth
strength and
power unto
{\ADD{his}}
people.
Blessed
{\ADD{be}}
God.

\par }\PsalmChap{69}{\D To the chief
Musician upon
Shoshannim,
{\ADD{A Psalm}} of
David.
\par }{\Q \VerseOne{1}Save me, O
God; for the
waters are
come in unto
{\ADD{my}}
soul.
\par }{\Q \VS{2}I
sink in
deep
mire, where
{\ADD{there is}} no
standing: I am
come into
deep
waters, where the
floods
overflow me.\FTNT{*}{{\FR 69.2 }
deep mire: Heb. the mire of depth}\FTNT{†}{{\FR 69.2 }
deep waters: Heb. depth of waters}
\par }{\Q \VS{3}I am
weary of my
crying: my
throat is
dried: mine
eyes
fail while I
wait for my
God.
\par }{\Q \VS{4}They that
hate me without a
cause are
more than the
hairs of mine
head: they that would
destroy me,
{\ADD{being}} mine
enemies
wrongfully, are
mighty: then I
restored
{\ADD{that}} which I took not
away.
\par }{\Q \VS{5}O
God, thou
knowest my
foolishness; and my
sins are not
hid from thee.\FTNT{‡}{{\FR 69.5 }
sins: Heb. guiltiness}
\par }{\Q \VS{6}Let not them that
wait on thee, O
Lord
{\ND{
GOD}} of
hosts, be
ashamed for my sake: let not those that
seek thee be
confounded for my sake, O
God of
Israel.
\par }{\Q \VS{7}Because for thy sake I have
borne
reproach;
shame hath
covered my
face.
\par }{\Q \VS{8}I am become a
stranger unto my
brethren, and an
alien unto my
mother’s
children.
\par }{\Q \VS{9}For the
zeal of thine
house hath eaten me
up; and the
reproaches of them that
reproached thee are
fallen upon me.
\par }{\Q \VS{10}When I
wept,
{\ADD{and chastened}} my
soul with
fasting, that was to my
reproach.
\par }{\Q \VS{11}I
made
sackcloth also my
garment; and I became a
proverb to them.
\par }{\Q \VS{12}They that sit
in the
gate
speak against me; and I
{\ADD{was}} the
song of the
drunkards.\FTNT{§}{{\FR 69.12 }
drunkards: Heb. drinkers of strong drink}
\par }{\BB \par }{\Q \VS{13}But as for me, my
prayer
{\ADD{is}} unto thee, O
{\ND{
LORD}},
{\ADD{in}} an
acceptable
time: O
God, in the
multitude of thy
mercy
hear me, in the
truth of thy
salvation.
\par }{\Q \VS{14}Deliver me out of the
mire, and let me not
sink: let me be
delivered from them that
hate me, and out of the
deep
waters.
\par }{\Q \VS{15}Let not the
waterflood
overflow me, neither let the
deep swallow me
up, and let not the
pit
shut her
mouth upon me.
\par }{\Q \VS{16}Hear me, O
{\ND{
LORD}}; for thy
lovingkindness
{\ADD{is}}
good:
turn unto me according to the
multitude of thy tender
mercies.
\par }{\Q \VS{17}And
hide not thy
face from thy
servant; for I am in
trouble:
hear me
speedily.\FTNT{**}{{\FR 69.17 }
hear…: Heb. make haste to hear me}
\par }{\Q \VS{18}Draw
nigh unto my
soul,
{\ADD{and}}
redeem it:
deliver me because of mine
enemies.
\par }{\Q \VS{19}Thou hast
known my
reproach, and my
shame, and my
dishonour: mine
adversaries
{\ADD{are}} all before thee.
\par }{\Q \VS{20}Reproach hath
broken my
heart; and I am full of
heaviness: and I
looked
{\ADD{for some}} to take
pity, but
{\ADD{there was}} none; and for
comforters, but I
found none.\FTNT{††}{{\FR 69.20 }
to take…: Heb. to lament with me}
\par }{\Q \VS{21}They
gave me also
gall for my
meat; and in my
thirst they gave me
vinegar to
drink.
\par }{\BB \par }{\Q \VS{22}Let their
table become a
snare
before them: and
{\ADD{that which should have been}} for
{\ADD{their}}
welfare,
{\ADD{let it become}} a
trap.
\par }{\Q \VS{23}Let their
eyes be
darkened, that they
see not; and make their
loins
continually to
shake.
\par }{\Q \VS{24}Pour
out thine
indignation upon them, and let thy
wrathful
anger take
hold of them.
\par }{\Q \VS{25}Let their
habitation be
desolate;
{\ADD{and}} let none
dwell in their
tents.\FTNT{‡‡}{{\FR 69.25 }
their habitation: Heb. their palace}\FTNT{§§}{{\FR 69.25 }
let none…: Heb. let there not be a dweller}
\par }{\Q \VS{26}For they
persecute
{\ADD{him}} whom thou hast
smitten; and they
talk to the
grief of those whom thou hast
wounded.\FTNT{***}{{\FR 69.26 }
those…: Heb. thy wounded}
\par }{\Q \VS{27}Add
iniquity unto their
iniquity: and let them not
come into thy
righteousness.\FTNT{†††}{{\FR 69.27 }
iniquity unto…: or, punishment of iniquity, etc}
\par }{\Q \VS{28}Let them be
blotted out of the
book of the
living, and not be
written with the
righteous.
\par }{\Q \VS{29}But I
{\ADD{am}}
poor and
sorrowful: let thy
salvation, O
God, set me up on
high.
\par }{\BB \par }{\Q \VS{30}I will
praise the
name of
God with a
song, and will
magnify him with
thanksgiving.
\par }{\Q \VS{31}{\ADD{This}} also shall
please the
{\ND{
LORD}}
better than an
ox
{\ADD{or}}
bullock that hath
horns and
hoofs.
\par }{\Q \VS{32}The
humble shall
see
{\ADD{this, and}} be
glad: and your
heart shall
live that
seek
God.\FTNT{‡‡‡}{{\FR 69.32 }
humble: or, meek}
\par }{\Q \VS{33}For the
{\ND{
LORD}}
heareth the
poor, and
despiseth not his
prisoners.
\par }{\Q \VS{34}Let the
heaven and
earth
praise him, the
seas, and every thing that
moveth therein.\FTNT{§§§}{{\FR 69.34 }
moveth: Heb. creepeth}
\par }{\Q \VS{35}For
God will
save
Zion, and will
build the
cities of
Judah: that they may
dwell there, and have it in
possession.
\par }{\Q \VS{36}The
seed also of his
servants shall
inherit it: and they that
love his
name shall
dwell therein.

\par }\PsalmChap{70}{\D To the chief
Musician,
{\ADD{A Psalm}} of
David, to bring to
remembrance.
\par }{\Q \VerseOne{1}{\ADD{Make haste}}, O
God, to
deliver me; make
haste to
help me, O
{\ND{
LORD}}.\FTNT{*}{{\FR 70.1 }
to help…: Heb. to my help}
\par }{\Q \VS{2}Let them be
ashamed and
confounded that
seek after my
soul: let them be
turned
backward, and put to
confusion, that
desire my
hurt.
\par }{\Q \VS{3}Let them be turned
back for a
reward of their
shame that
say,
Aha,
aha.
\par }{\Q \VS{4}Let all those that
seek thee
rejoice and be
glad in thee: and let such as
love thy
salvation
say
continually, Let
God be
magnified.
\par }{\Q \VS{5}But I
{\ADD{am}}
poor and
needy: make
haste unto me, O
God: thou
{\ADD{art}} my
help and my
deliverer; O
{\ND{
LORD}}, make no
tarrying.

\par }\PsalmChap{71}{\Q \VerseOne{1}In thee, O
{\ND{
LORD}}, do I put my
trust: let me
never be put to
confusion.
\par }{\Q \VS{2}Deliver me in thy
righteousness, and cause me to
escape:
incline thine
ear unto me, and
save me.
\par }{\Q \VS{3}Be thou my
strong
habitation, whereunto I may
continually
resort: thou hast given
commandment to
save me; for thou
{\ADD{art}} my
rock and my
fortress.\FTNT{*}{{\FR 71.3 }
Be…: Heb. Be thou to me for a rock of habitation}
\par }{\Q \VS{4}Deliver me, O my
God, out of the
hand of the
wicked, out of the
hand of the
unrighteous and cruel
man.
\par }{\Q \VS{5}For thou
{\ADD{art}} my
hope, O
Lord
{\ND{
GOD}}:
{\ADD{thou art}} my
trust from my
youth.
\par }{\Q \VS{6}By thee have I been holden
up from the
womb: thou art he that
took me out of my
mother’s
bowels: my
praise
{\ADD{shall be}}
continually of thee.
\par }{\Q \VS{7}I am as a
wonder unto
many; but thou
{\ADD{art}} my
strong
refuge.
\par }{\Q \VS{8}Let my
mouth be
filled
{\ADD{with}} thy
praise
{\ADD{and with}} thy
honour all the
day.
\par }{\Q \VS{9}Cast me not
off in the
time of old
age;
forsake me not when my
strength
faileth.
\par }{\Q \VS{10}For mine
enemies
speak against me; and they that lay
wait for my
soul take
counsel
together,\FTNT{†}{{\FR 71.10 }
lay…: Heb. watch, or, observe}
\par }{\Q \VS{11}Saying,
God hath
forsaken him:
persecute and
take him; for
{\ADD{there is}} none to
deliver
{\ADD{him}}.
\par }{\Q \VS{12}O
God, be not
far from me: O my
God, make
haste for my
help.
\par }{\Q \VS{13}Let them be
confounded
{\ADD{and}}
consumed that are
adversaries to my
soul; let them be
covered
{\ADD{with}}
reproach and
dishonour that
seek my
hurt.
\par }{\BB \par }{\Q \VS{14}But I will
hope
continually, and will yet
praise thee
more and
more.
\par }{\Q \VS{15}My
mouth shall shew
forth thy
righteousness
{\ADD{and}} thy
salvation all the
day; for I
know not the
numbers
{\ADD{thereof}}.
\par }{\Q \VS{16}I will go
in the
strength of the
Lord
{\ND{
GOD}}: I will make
mention of thy
righteousness,
{\ADD{even}} of thine only.
\par }{\Q \VS{17}O
God, thou hast
taught me from my
youth: and hitherto have I
declared thy wondrous
works.
\par }{\Q \VS{18}Now also
when I am
old and
grayheaded, O
God,
forsake me not; until I have
shewed thy
strength unto
{\ADD{this}}
generation,
{\ADD{and}} thy
power to every one
{\ADD{that}} is to
come.\FTNT{‡}{{\FR 71.18 }
when…: Heb. unto old age and gray hairs}\FTNT{§}{{\FR 71.18 }
thy strength: Heb. thine arm}
\par }{\Q \VS{19}Thy
righteousness also, O
God,
{\ADD{is}} very
high, who hast
done great
things: O
God, who
{\ADD{is}} like unto thee!
\par }{\Q \VS{20}{\ADD{Thou}}, which hast
shewed me
great and
sore
troubles, shalt
quicken me
again, and shalt bring me
up
again from the
depths of the
earth.
\par }{\Q \VS{21}Thou shalt
increase my
greatness, and
comfort me on every
side.
\par }{\Q \VS{22}I will also
praise thee with the
psaltery,
{\ADD{even}} thy
truth, O my
God: unto thee will I
sing with the
harp, O thou Holy
One of
Israel.\FTNT{**}{{\FR 71.22 }
the psaltery: Heb. the instrument of psaltery}
\par }{\Q \VS{23}My
lips shall greatly
rejoice when I
sing unto thee; and my
soul, which thou hast
redeemed.
\par }{\Q \VS{24}My
tongue also shall
talk of thy
righteousness all the
day long: for they are
confounded, for they are brought unto
shame, that
seek my
hurt.

\par }\PsalmChap{72}{\D {\ADD{A Psalm}} for
Solomon.
\par }{\Q \VerseOne{1}Give the
king thy
judgments, O
God, and thy
righteousness unto the
king’s
son.\FTNT{*}{{\FR 72.1 }
for: or, of}
\par }{\BB \par }{\Q \VS{2}He shall
judge thy
people with
righteousness, and thy
poor with
judgment.
\par }{\Q \VS{3}The
mountains shall
bring
peace to the
people, and the little
hills, by
righteousness.
\par }{\Q \VS{4}He shall
judge the
poor of the
people, he shall
save the
children of the
needy, and shall break in
pieces the
oppressor.
\par }{\Q \VS{5}They shall
fear thee as long
as the
sun and
moon
endure, throughout
all
generations.
\par }{\Q \VS{6}He shall come
down like
rain upon the
mown grass: as
showers
{\ADD{that}}
water the
earth.
\par }{\Q \VS{7}In his
days shall the
righteous
flourish; and
abundance of
peace so long as the
moon endureth.\FTNT{†}{{\FR 72.7 }
so long…: Heb. till there be no moon}
\par }{\Q \VS{8}He shall have
dominion also from
sea to
sea, and from the
river unto the
ends of the
earth.
\par }{\Q \VS{9}They that dwell in the
wilderness shall
bow
before him; and his
enemies shall
lick the
dust.
\par }{\Q \VS{10}The
kings of
Tarshish and of the
isles shall
bring
presents: the
kings of
Sheba and
Seba shall
offer
gifts.
\par }{\Q \VS{11}Yea, all
kings shall fall
down before him: all
nations shall
serve him.
\par }{\Q \VS{12}For he shall
deliver the
needy when he
crieth; the
poor also, and
{\ADD{him}} that hath no
helper.
\par }{\Q \VS{13}He shall
spare the
poor and
needy, and shall
save the
souls of the
needy.
\par }{\Q \VS{14}He shall
redeem their
soul from
deceit and
violence: and
precious shall their
blood be in his
sight.
\par }{\Q \VS{15}And he shall
live, and to him shall be
given of the
gold of
Sheba: prayer also shall be
made for him
continually;
{\ADD{and}}
daily shall he be
praised.\FTNT{‡}{{\FR 72.15 }
shall be given: Heb. one shall give}
\par }{\Q \VS{16}There shall be an
handful of
corn in the
earth upon the
top of the
mountains; the
fruit thereof shall
shake like
Lebanon: and
{\ADD{they}} of the
city shall
flourish like
grass of the
earth.
\par }{\Q \VS{17}His
name shall endure for
ever: his
name shall be
continued as long
as the
sun: and
{\ADD{men}} shall be
blessed in him: all
nations shall call him
blessed.\FTNT{§}{{\FR 72.17 }
endure: Heb. be}\FTNT{**}{{\FR 72.17 }
his name shall be…: Heb. shall be as a son to continue his father’s name for ever}
\par }{\BB \par }{\Q \VS{18}Blessed
{\ADD{be}} the
{\ND{
LORD}}
God, the
God of
Israel, who only
doeth wondrous
things.
\par }{\Q \VS{19}And
blessed
{\ADD{be}} his
glorious
name for
ever: and let the whole
earth be
filled
{\ADD{with}} his
glory;
Amen, and
Amen.
\par }{\Q \VS{20}The
prayers of
David the
son of
Jesse are
ended.

\par }\PsalmChap{73}{\D A
Psalm of
Asaph.
\par }{\Q \VerseOne{1}Truly
God
{\ADD{is}}
good to
Israel,
{\ADD{even}} to such as are of a
clean
heart.\FTNT{*}{{\FR 73.1 }
of: or, for}\FTNT{†}{{\FR 73.1 }
Truly: or, Yet}\FTNT{‡}{{\FR 73.1 }
of…: Heb. clean of heart}
\par }{\Q \VS{2}But as for me, my
feet were
almost
gone; my
steps had well
nigh
slipped.
\par }{\Q \VS{3}For I was
envious at the
foolish,
{\ADD{when}} I
saw the
prosperity of the
wicked.
\par }{\Q \VS{4}For
{\ADD{there are}} no
bands in their
death: but their
strength
{\ADD{is}}
firm.\FTNT{§}{{\FR 73.4 }
firm: Heb. fat}
\par }{\Q \VS{5}They
{\ADD{are}} not in
trouble
{\ADD{as other}}
men; neither are they
plagued
like
{\ADD{other}}
men.\FTNT{**}{{\FR 73.5 }
in…: Heb. in the trouble of other men}\FTNT{††}{{\FR 73.5 }
like: Heb. with}
\par }{\Q \VS{6}Therefore
pride compasseth them about as a
chain;
violence
covereth them
{\ADD{as}} a
garment.
\par }{\Q \VS{7}Their
eyes stand
out with
fatness: they have
more than
heart could
wish.\FTNT{‡‡}{{\FR 73.7 }
have…: Heb. pass the thoughts of the heart}
\par }{\Q \VS{8}They are
corrupt, and
speak
wickedly
{\ADD{concerning}}
oppression: they
speak
loftily.
\par }{\Q \VS{9}They
set their
mouth against the
heavens, and their
tongue
walketh through the
earth.
\par }{\Q \VS{10}Therefore his
people
return
hither: and
waters of a
full
{\ADD{cup}} are wrung
out to them.
\par }{\Q \VS{11}And they
say, How doth
God
know? and is
there
knowledge in the most
High?
\par }{\Q \VS{12}Behold, these
{\ADD{are}} the
ungodly, who
prosper in the
world; they
increase
{\ADD{in}}
riches.
\par }{\Q \VS{13}Verily I have
cleansed my
heart
{\ADD{in}}
vain, and
washed my
hands in
innocency.
\par }{\Q \VS{14}For all the
day long have I been
plagued, and
chastened every
morning.\FTNT{§§}{{\FR 73.14 }
chastened: Heb. my chastisement was}
\par }{\BB \par }{\Q \VS{15}If I
say, I will
speak
thus; behold, I should
offend
{\ADD{against}} the
generation of thy
children.
\par }{\Q \VS{16}When I
thought to
know this, it
{\ADD{was}} too
painful for
me;\FTNT{***}{{\FR 73.16 }
too…: Heb. labour in mine eyes}
\par }{\Q \VS{17}Until I
went into the
sanctuary of
God;
{\ADD{then}}
understood I their
end.
\par }{\Q \VS{18}Surely thou didst
set them in slippery
places: thou castedst them
down into
destruction.
\par }{\Q \VS{19}How are they
{\ADD{brought}} into
desolation, as in a
moment! they are
utterly
consumed with
terrors.
\par }{\Q \VS{20}As a
dream when
{\ADD{one}}
awaketh;
{\ADD{so}}, O
Lord, when thou
awakest, thou shalt
despise their
image.
\par }{\BB \par }{\Q \VS{21}Thus my
heart was
grieved, and I was
pricked in my
reins.
\par }{\Q \VS{22}So
foolish
{\ADD{was}} I, and
ignorant: I was
{\ADD{as}} a
beast before thee.\FTNT{†††}{{\FR 73.22 }
ignorant: Heb. I knew not}\FTNT{‡‡‡}{{\FR 73.22 }
before: Heb. with}
\par }{\Q \VS{23}Nevertheless I
{\ADD{am}}
continually with thee: thou hast
holden
{\ADD{me}} by my
right
hand.
\par }{\Q \VS{24}Thou shalt
guide me with thy
counsel, and
afterward
receive me
{\ADD{to}}
glory.
\par }{\Q \VS{25}Whom have I in
heaven
{\ADD{but thee}}? and
{\ADD{there is}} none upon
earth
{\ADD{that}} I
desire beside thee.
\par }{\Q \VS{26}My
flesh and my
heart
faileth:
{\ADD{but}}
God
{\ADD{is}} the
strength of my
heart, and my
portion for
ever.\FTNT{§§§}{{\FR 73.26 }
strength: Heb. rock}
\par }{\Q \VS{27}For, lo, they that are
far from thee shall
perish: thou hast
destroyed all them that go a
whoring from thee.
\par }{\Q \VS{28}But
{\ADD{it is}}
good for me to draw
near to
God: I have
put my
trust in the
Lord
{\ND{
GOD}}, that I may
declare all thy
works.

\par }\PsalmChap{74}{\D Maschil of
Asaph.
\par }{\Q \VerseOne{1}O
God, why hast thou cast
{\ADD{us}}
off for
ever?
{\ADD{why}} doth thine
anger
smoke against the
sheep of thy
pasture?\FTNT{*}{{\FR 74.1 }
Maschil…: or, A Psalm for Asaph to give instruction}
\par }{\Q \VS{2}Remember thy
congregation,
{\ADD{which}} thou hast
purchased of
old; the
rod of thine
inheritance,
{\ADD{which}} thou hast
redeemed; this
mount
Zion, wherein thou hast
dwelt.\FTNT{†}{{\FR 74.2 }
rod: or, tribe}
\par }{\Q \VS{3}Lift
up thy
feet unto the
perpetual
desolations;
{\ADD{even}} all
{\ADD{that}} the
enemy hath done
wickedly in the
sanctuary.
\par }{\Q \VS{4}Thine
enemies
roar in the
midst of thy
congregations; they set
up their
ensigns
{\ADD{for}}
signs.
\par }{\Q \VS{5}{\ADD{A man}} was
famous according as he had lifted
up
axes upon the
thick
trees.
\par }{\Q \VS{6}But now they break
down the carved
work thereof at
once with
axes and
hammers.
\par }{\Q \VS{7}They have
cast
fire into thy
sanctuary, they have
defiled
{\ADD{by casting down}} the dwelling
place of thy
name to the
ground.\FTNT{‡}{{\FR 74.7 }
They have cast…: Heb. They have sent thy sanctuary unto the fire}
\par }{\Q \VS{8}They
said in their
hearts, Let us
destroy them
together: they have burned
up all the
synagogues of
God in the
land.\FTNT{§}{{\FR 74.8 }
destroy: Heb. break}
\par }{\Q \VS{9}We
see not our
signs:
{\ADD{there is}} no more any
prophet: neither
{\ADD{is there}} among us any that
knoweth how
long.
\par }{\Q \VS{10}O
God, how long shall the
adversary
reproach? shall the
enemy
blaspheme thy
name for
ever?
\par }{\Q \VS{11}Why
withdrawest thou thy
hand, even thy right
hand?
pluck
{\ADD{it}} out
of thy
bosom.
\par }{\BB \par }{\Q \VS{12}For
God
{\ADD{is}} my
King of
old,
working
salvation in the
midst of the
earth.
\par }{\Q \VS{13}Thou didst
divide the
sea by thy
strength: thou
brakest the
heads of the
dragons in the
waters.\FTNT{**}{{\FR 74.13 }
divide: Heb. break}\FTNT{††}{{\FR 74.13 }
dragons: or, whales}
\par }{\Q \VS{14}Thou
brakest the
heads of
leviathan in pieces,
{\ADD{and}}
gavest him
{\ADD{to be}}
meat to the
people inhabiting the
wilderness.
\par }{\Q \VS{15}Thou didst
cleave the
fountain and the
flood: thou driedst
up
mighty
rivers.\FTNT{‡‡}{{\FR 74.15 }
mighty…: Heb. rivers of strength}
\par }{\Q \VS{16}The
day
{\ADD{is}} thine, the
night also
{\ADD{is}} thine: thou hast
prepared the
light and the
sun.
\par }{\Q \VS{17}Thou hast
set all the
borders of the
earth: thou hast
made
summer and
winter.\FTNT{§§}{{\FR 74.17 }
made: Heb. made them}
\par }{\BB \par }{\Q \VS{18}Remember this,
{\ADD{that}} the
enemy hath
reproached, O
{\ND{
LORD}}, and
{\ADD{that}} the
foolish
people have
blasphemed thy
name.
\par }{\Q \VS{19}O
deliver not the
soul of thy
turtledove unto the
multitude
{\ADD{of the wicked:}}
forget not the
congregation of thy
poor for
ever.
\par }{\Q \VS{20}Have
respect unto the
covenant: for the
dark places of the
earth are
full of the
habitations of
cruelty.
\par }{\Q \VS{21}O let not the
oppressed
return
ashamed: let the
poor and
needy
praise thy
name.
\par }{\Q \VS{22}Arise, O
God,
plead thine own
cause:
remember how the foolish
man
reproacheth thee
daily.
\par }{\Q \VS{23}Forget not the
voice of thine
enemies: the
tumult of those that rise
up against thee
increaseth
continually.\FTNT{***}{{\FR 74.23 }
increaseth: Heb. ascendeth}

\par }\PsalmChap{75}{\D To the chief
Musician,
Al-taschith, A
Psalm
{\ADD{or}}
Song of
Asaph.
\par }{\Q \VerseOne{1}Unto thee, O
God, do we give
thanks,
{\ADD{unto thee}} do we give
thanks: for
{\ADD{that}} thy
name is
near thy wondrous
works
declare.\FTNT{*}{{\FR 75.1 }
Al-taschith: or, Destroy not}\FTNT{†}{{\FR 75.1 }
of: or, for}
\par }{\Q \VS{2}When I shall
receive the
congregation I will
judge
uprightly.\FTNT{‡}{{\FR 75.2 }
receive…: or, take a set time}
\par }{\Q \VS{3}The
earth and all the
inhabitants thereof are
dissolved: I bear
up the
pillars of it.
Selah.
\par }{\Q \VS{4}I
said unto the
fools, Deal not
foolishly: and to the
wicked, Lift not
up the
horn:
\par }{\Q \VS{5}Lift not
up your
horn on
high:
speak
{\ADD{not with}} a
stiff
neck.
\par }{\BB \par }{\Q \VS{6}For
promotion
{\ADD{cometh}} neither from the
east, nor from the
west, nor from the
south.\FTNT{§}{{\FR 75.6 }
south: Heb. desert}
\par }{\Q \VS{7}But
God
{\ADD{is}} the
judge: he putteth
down one, and setteth
up another.
\par }{\Q \VS{8}For in the
hand of the
{\ND{
LORD}}
{\ADD{there is}} a
cup, and the
wine is
red; it is
full of
mixture; and he poureth
out of the
same: but the
dregs thereof, all the
wicked of the
earth shall wring
{\ADD{them}}
out,
{\ADD{and}}
drink
{\ADD{them}}.
\par }{\Q \VS{9}But I will
declare for
ever; I will sing
praises to the
God of
Jacob.
\par }{\Q \VS{10}All the
horns of the
wicked also will I cut
off;
{\ADD{but}} the
horns of the
righteous shall be
exalted.

\par }\PsalmChap{76}{\D To the chief
Musician on
Neginoth, A
Psalm
{\ADD{or}}
Song of
Asaph.
\par }{\Q \VerseOne{1}In
Judah
{\ADD{is}}
God
known: his
name
{\ADD{is}}
great in
Israel.\FTNT{*}{{\FR 76.1 }
chief…: or, overseer}\FTNT{†}{{\FR 76.1 }
of: or, for}
\par }{\Q \VS{2}In
Salem also is his
tabernacle, and his dwelling
place in
Zion.
\par }{\Q \VS{3}There
brake he the
arrows of the
bow, the
shield, and the
sword, and the
battle.
Selah.
\par }{\Q \VS{4}Thou
{\ADD{art}} more
glorious
{\ADD{and}}
excellent than the
mountains of
prey.
\par }{\Q \VS{5}The
stouthearted are
spoiled, they have
slept their
sleep: and none of the
men of
might have
found their
hands.
\par }{\Q \VS{6}At thy
rebuke, O
God of
Jacob, both the
chariot and
horse are cast into a dead
sleep.
\par }{\BB \par }{\Q \VS{7}Thou,
{\ADD{even}} thou,
{\ADD{art}} to be
feared: and who may
stand in thy
sight
when once thou art
angry?
\par }{\Q \VS{8}Thou didst cause
judgment to be
heard from
heaven; the
earth
feared, and was
still,
\par }{\Q \VS{9}When
God
arose to
judgment, to
save all the
meek of the
earth.
Selah.
\par }{\Q \VS{10}Surely the
wrath of
man shall
praise thee: the
remainder of
wrath shalt thou
restrain.
\par }{\Q \VS{11}Vow, and
pay unto the
{\ND{
LORD}} your
God: let all that be round
about him
bring
presents unto him that ought to be
feared.\FTNT{‡}{{\FR 76.11 }
unto him…: Heb. to fear}
\par }{\Q \VS{12}He shall cut
off the
spirit of
princes:
{\ADD{he is}}
terrible to the
kings of the
earth.

\par }\PsalmChap{77}{\D To the chief
Musician, to
Jeduthun, A
Psalm of
Asaph.
\par }{\Q \VerseOne{1}I
cried unto
God with my
voice,
{\ADD{even}} unto
God with my
voice; and he gave
ear unto me.\FTNT{*}{{\FR 77.1 }
of: or, for}
\par }{\Q \VS{2}In the
day of my
trouble I
sought the
Lord: my
sore
ran in the
night, and
ceased not: my
soul
refused to be
comforted.\FTNT{†}{{\FR 77.2 }
sore: Heb. hand}
\par }{\Q \VS{3}I
remembered
God, and was
troubled: I
complained, and my
spirit was
overwhelmed.
Selah.
\par }{\Q \VS{4}Thou
holdest mine
eyes
waking: I am so
troubled that I cannot
speak.
\par }{\Q \VS{5}I have
considered the
days of
old, the
years of ancient
times.
\par }{\Q \VS{6}I call to
remembrance my
song in the
night: I
commune with mine own
heart: and my
spirit made diligent
search.
\par }{\Q \VS{7}Will the
Lord cast
off for
ever? and
will he be
favourable no more?
\par }{\Q \VS{8}Is his
mercy clean
gone for
ever? doth
{\ADD{his}}
promise
fail for
evermore?\FTNT{‡}{{\FR 77.8 }
for evermore: Heb. to generation and generation?}
\par }{\Q \VS{9}Hath
God
forgotten to be
gracious? hath he in
anger shut
up his tender
mercies?
Selah.
\par }{\Q \VS{10}And I
said, This
{\ADD{is}} my
infirmity:
{\ADD{but I will remember}} the
years of the right
hand of the most
High.
\par }{\BB \par }{\Q \VS{11}I will
remember the
works of the
{\ND{
LORD}}: surely I will
remember thy
wonders of
old.
\par }{\Q \VS{12}I will
meditate also of all thy
work, and
talk of thy
doings.
\par }{\Q \VS{13}Thy
way, O
God,
{\ADD{is}} in the
sanctuary: who
{\ADD{is so}}
great a
God as
{\ADD{our}}
God?
\par }{\Q \VS{14}Thou
{\ADD{art}} the
God that
doest
wonders: thou hast
declared thy
strength among the
people.
\par }{\Q \VS{15}Thou hast with
{\ADD{thine}}
arm
redeemed thy
people, the
sons of
Jacob and
Joseph.
Selah.
\par }{\Q \VS{16}The
waters
saw thee, O
God, the
waters
saw thee; they were
afraid: the
depths also were
troubled.
\par }{\Q \VS{17}The
clouds poured
out
water: the
skies sent
out a
sound: thine
arrows also went
abroad.\FTNT{§}{{\FR 77.17 }
poured…: Heb. were poured forth with water}
\par }{\Q \VS{18}The
voice of thy
thunder
{\ADD{was}} in the
heaven: the
lightnings
lightened the
world: the
earth
trembled and
shook.
\par }{\Q \VS{19}Thy
way
{\ADD{is}} in the
sea, and thy
path in the
great
waters, and thy
footsteps are not
known.
\par }{\Q \VS{20}Thou
leddest thy
people like a
flock by the
hand of
Moses and
Aaron.

\par }\PsalmChap{78}{\D Maschil of
Asaph.
\par }{\Q \VerseOne{1}Give
ear, O my
people,
{\ADD{to}} my
law:
incline your
ears to the
words of my
mouth.\FTNT{*}{{\FR 78.1 }
Maschil…: or, A Psalm for Asaph to give instruction}
\par }{\Q \VS{2}I will
open my
mouth in a
parable: I will
utter dark
sayings of
old:
\par }{\Q \VS{3}Which we have
heard and
known, and our
fathers have
told us.
\par }{\Q \VS{4}We will not
hide
{\ADD{them}} from their
children,
shewing to the
generation to
come the
praises of the
{\ND{
LORD}}, and his
strength, and his wonderful
works that he hath
done.
\par }{\Q \VS{5}For he
established a
testimony in
Jacob, and
appointed a
law in
Israel, which he
commanded our
fathers, that they should make them
known to their
children:
\par }{\Q \VS{6}That the
generation to
come might
know
{\ADD{them, even}} the
children
{\ADD{which}} should be
born;
{\ADD{who}} should
arise and
declare
{\ADD{them}} to their
children:
\par }{\Q \VS{7}That they might
set their
hope in
God, and not
forget the
works of
God, but
keep his
commandments:
\par }{\Q \VS{8}And might not be as their
fathers, a
stubborn and
rebellious
generation; a
generation
{\ADD{that}}
set not their
heart
aright, and whose
spirit was not
stedfast with
God.\FTNT{†}{{\FR 78.8 }
that…: Heb. that prepared not their heart}
\par }{\BB \par }{\Q \VS{9}The
children of
Ephraim,
{\ADD{being}}
armed,
{\ADD{and}}
carrying
bows, turned
back in the
day of
battle.\FTNT{‡}{{\FR 78.9 }
carrying: Heb. throwing forth}
\par }{\Q \VS{10}They
kept not the
covenant of
God, and
refused to
walk in his
law;
\par }{\Q \VS{11}And
forgat his
works, and his
wonders that he had
shewed them.
\par }{\Q \VS{12}Marvellous
things
did he in the
sight of their
fathers, in the
land of
Egypt,
{\ADD{in}} the
field of
Zoan.
\par }{\Q \VS{13}He
divided the
sea, and caused them to pass
through; and he made the
waters to
stand as an
heap.
\par }{\Q \VS{14}In the
daytime also he
led them with a
cloud, and all the
night with a
light of
fire.
\par }{\Q \VS{15}He
clave the
rocks in the
wilderness, and gave
{\ADD{them}}
drink as
{\ADD{out of}} the
great
depths.
\par }{\Q \VS{16}He
brought
streams also out of the
rock, and caused
waters to run
down like
rivers.
\par }{\Q \VS{17}And they
sinned yet
more against him by
provoking the most
High in the
wilderness.
\par }{\Q \VS{18}And they
tempted
God in their
heart by
asking
meat for their
lust.
\par }{\Q \VS{19}Yea, they
spake against
God; they
said,
Can
God
furnish a
table in the
wilderness?\FTNT{§}{{\FR 78.19 }
furnish: Heb. order}
\par }{\Q \VS{20}Behold, he
smote the
rock, that the
waters gushed
out, and the
streams
overflowed;
can he
give
bread also?
can he
provide
flesh for his
people?
\par }{\Q \VS{21}Therefore the
{\ND{
LORD}}
heard
{\ADD{this}}, and was
wroth: so a
fire was
kindled against
Jacob, and
anger also came
up against
Israel;
\par }{\Q \VS{22}Because they
believed not in
God, and
trusted not in his
salvation:
\par }{\Q \VS{23}Though he had
commanded the
clouds from
above, and
opened the
doors of
heaven,
\par }{\Q \VS{24}And had rained
down
manna upon them to
eat, and had
given them of the
corn of
heaven.
\par }{\Q \VS{25}Man did
eat
angels’
food: he
sent them
meat to the
full.\FTNT{**}{{\FR 78.25 }
Man…: or, Every one did eat the bread of the mighty}
\par }{\Q \VS{26}He caused an east
wind to
blow in the
heaven: and by his
power he brought
in the south
wind.\FTNT{††}{{\FR 78.26 }
to blow: Heb. to go}
\par }{\Q \VS{27}He
rained
flesh also upon them as
dust, and
feathered
fowls like as the
sand of the
sea:\FTNT{‡‡}{{\FR 78.27 }
feathered…: Heb. fowl of wing}
\par }{\Q \VS{28}And he let
{\ADD{it}}
fall in the
midst of their
camp, round
about their
habitations.
\par }{\Q \VS{29}So they did
eat, and were
well
filled: for he
gave them their own
desire;
\par }{\Q \VS{30}They were not
estranged from their
lust. But while their
meat
{\ADD{was}} yet in their
mouths,
\par }{\Q \VS{31}The
wrath of
God
came upon them, and
slew the
fattest of them, and smote
down the
chosen
{\ADD{men}} of
Israel.\FTNT{§§}{{\FR 78.31 }
smote…: Heb. made to bow}\FTNT{***}{{\FR 78.31 }
chosen…: or, young men}
\par }{\Q \VS{32}For all this they
sinned still, and
believed not for his wondrous
works.
\par }{\Q \VS{33}Therefore their
days did he
consume in
vanity, and their
years in
trouble.
\par }{\Q \VS{34}When he
slew them, then they
sought him: and they
returned and enquired
early after
God.
\par }{\Q \VS{35}And they
remembered that
God
{\ADD{was}} their
rock, and the
high
God their
redeemer.
\par }{\Q \VS{36}Nevertheless they did
flatter him with their
mouth, and they
lied unto him with their
tongues.
\par }{\Q \VS{37}For their
heart was not
right with him, neither were they
stedfast in his
covenant.
\par }{\Q \VS{38}But he,
{\ADD{being}} full of
compassion,
forgave
{\ADD{their}}
iniquity, and
destroyed
{\ADD{them}} not: yea, many a
time
turned he his
anger
away, and did not stir
up all his
wrath.
\par }{\Q \VS{39}For he
remembered that they
{\ADD{were but}}
flesh; a
wind that passeth
away, and cometh not
again.
\par }{\BB \par }{\Q \VS{40}How oft did they
provoke him in the
wilderness,
{\ADD{and}}
grieve him in the
desert!\FTNT{†††}{{\FR 78.40 }
provoke: or, rebel against}
\par }{\Q \VS{41}Yea, they turned
back and
tempted
God, and
limited the Holy
One of
Israel.
\par }{\Q \VS{42}They
remembered not his
hand,
{\ADD{nor}} the
day when he
delivered them from the
enemy.\FTNT{‡‡‡}{{\FR 78.42 }
from…: or, from affliction}
\par }{\Q \VS{43}How he had
wrought his
signs in
Egypt, and his
wonders in the
field of
Zoan:\FTNT{§§§}{{\FR 78.43 }
wrought: Heb. set}
\par }{\Q \VS{44}And had
turned their
rivers into
blood; and their
floods, that they could not
drink.
\par }{\Q \VS{45}He
sent divers sorts of
flies among them, which
devoured them; and
frogs, which
destroyed them.
\par }{\Q \VS{46}He
gave also their
increase unto the
caterpiller, and their
labour unto the
locust.
\par }{\Q \VS{47}He
destroyed their
vines with
hail, and their sycomore
trees with
frost.\FTNT{*}{{\FR 78.47 }
destroyed: Heb. killed}\FTNT{†}{{\FR 78.47 }
frost: or, great hailstones}
\par }{\Q \VS{48}He gave
up their
cattle also to the
hail, and their
flocks to hot
thunderbolts.\FTNT{‡}{{\FR 78.48 }
gave…: Heb. shut up}\FTNT{§}{{\FR 78.48 }
hot…: or, lightnings}
\par }{\Q \VS{49}He
cast upon them the
fierceness of his
anger,
wrath, and
indignation, and
trouble, by
sending
evil
angels
{\ADD{among them}}.
\par }{\Q \VS{50}He
made a
way to his
anger; he
spared not their
soul from
death, but
gave their
life
over to the
pestilence;\FTNT{**}{{\FR 78.50 }
He made…: Heb. He weighed a path}\FTNT{††}{{\FR 78.50 }
life…: or, beasts to the murrain}
\par }{\Q \VS{51}And
smote all the
firstborn in
Egypt; the
chief of
{\ADD{their}}
strength in the
tabernacles of
Ham:
\par }{\Q \VS{52}But made his own
people to go
forth like
sheep, and
guided them in the
wilderness like a
flock.
\par }{\Q \VS{53}And he
led them on
safely, so that they
feared not: but the
sea
overwhelmed their
enemies.\FTNT{‡‡}{{\FR 78.53 }
overwhelmed: Heb. covered}
\par }{\Q \VS{54}And he
brought them to the
border of his
sanctuary,
{\ADD{even to}} this
mountain,
{\ADD{which}} his right
hand had
purchased.
\par }{\Q \VS{55}He cast
out the
heathen also
before them, and
divided them an
inheritance by
line, and made the
tribes of
Israel to
dwell in their
tents.
\par }{\Q \VS{56}Yet they
tempted and
provoked the most
high
God, and
kept not his
testimonies:
\par }{\Q \VS{57}But turned
back, and dealt
unfaithfully like their
fathers: they were turned
aside like a
deceitful
bow.
\par }{\Q \VS{58}For they provoked him to
anger with their high
places, and moved him to
jealousy with their graven
images.
\par }{\Q \VS{59}When
God
heard
{\ADD{this}}, he was
wroth, and
greatly
abhorred
Israel:
\par }{\Q \VS{60}So that he
forsook the
tabernacle of
Shiloh, the
tent
{\ADD{which}} he
placed among
men;
\par }{\Q \VS{61}And
delivered his
strength into
captivity, and his
glory into the
enemy’s
hand.
\par }{\Q \VS{62}He
gave his
people
over also unto the
sword; and was
wroth with his
inheritance.
\par }{\Q \VS{63}The
fire
consumed their young
men; and their
maidens were not given to
marriage.\FTNT{§§}{{\FR 78.63 }
given…: Heb. praised}
\par }{\Q \VS{64}Their
priests
fell by the
sword; and their
widows made no
lamentation.
\par }{\Q \VS{65}Then the
Lord
awaked as one out of
sleep,
{\ADD{and}} like a mighty
man that
shouteth by reason of
wine.
\par }{\Q \VS{66}And he
smote his
enemies in the hinder
parts: he
put them to a
perpetual
reproach.
\par }{\Q \VS{67}Moreover he
refused the
tabernacle of
Joseph, and
chose not the
tribe of
Ephraim:
\par }{\Q \VS{68}But
chose the
tribe of
Judah, the
mount
Zion which he
loved.
\par }{\Q \VS{69}And he
built his
sanctuary like
high
{\ADD{palaces}}, like the
earth which he hath
established for
ever.\FTNT{***}{{\FR 78.69 }
established: Heb. founded}
\par }{\Q \VS{70}He
chose
David also his
servant, and
took him from the
sheepfolds:
\par }{\Q \VS{71}From
following the ewes great with
young he
brought him to
feed
Jacob his
people, and
Israel his
inheritance.\FTNT{†††}{{\FR 78.71 }
following: Heb. after}
\par }{\Q \VS{72}So he
fed them according to the
integrity of his
heart; and
guided them by the
skilfulness of his
hands.

\par }\PsalmChap{79}{\D A
Psalm of
Asaph.
\par }{\Q \VerseOne{1}O
God, the
heathen are
come into thine
inheritance; thy
holy
temple have they
defiled; they have
laid
Jerusalem on
heaps.\FTNT{*}{{\FR 79.1 }
of: or, for}
\par }{\Q \VS{2}The dead
bodies of thy
servants have they
given
{\ADD{to be}}
meat unto the
fowls of the
heaven, the
flesh of thy
saints unto the
beasts of the
earth.
\par }{\Q \VS{3}Their
blood have they
shed like
water round
about
Jerusalem; and
{\ADD{there was}} none to
bury
{\ADD{them}}.
\par }{\Q \VS{4}We are become a
reproach to our
neighbours, a
scorn and
derision to them that are round
about us.
\par }{\Q \VS{5}How long,
{\ND{
LORD}}? wilt thou be
angry for
ever? shall thy
jealousy
burn like
fire?
\par }{\BB \par }{\Q \VS{6}Pour
out thy
wrath upon the
heathen that have not
known thee, and upon the
kingdoms that have not
called upon thy
name.
\par }{\Q \VS{7}For they have
devoured
Jacob, and laid
waste his dwelling
place.
\par }{\Q \VS{8}O
remember not against us
former
iniquities: let thy tender
mercies
speedily
prevent us: for we are
brought
very
low.\FTNT{†}{{\FR 79.8 }
former…: or, the iniquities of them that were before us}
\par }{\Q \VS{9}Help us, O
God of our
salvation,
for the
glory of thy
name: and
deliver us, and purge
away our
sins, for thy
name’s sake.
\par }{\Q \VS{10}Wherefore should the
heathen
say, Where
{\ADD{is}} their
God? let him be
known among the
heathen in our
sight
{\ADD{by}} the
revenging of the
blood of thy
servants
{\ADD{which is}}
shed.\FTNT{‡}{{\FR 79.10 }
revenging: Heb. vengeance}
\par }{\Q \VS{11}Let the
sighing of the
prisoner
come
before thee; according to the
greatness of thy
power
preserve thou those that are
appointed to
die;\FTNT{§}{{\FR 79.11 }
thy…: Heb. thine arm}\FTNT{**}{{\FR 79.11 }
preserve…: Heb. reserve the children of death}
\par }{\Q \VS{12}And
render unto our
neighbours
sevenfold into their
bosom their
reproach, wherewith they have
reproached thee, O
Lord.
\par }{\Q \VS{13}So we thy
people and
sheep of thy
pasture will give thee
thanks for
ever: we will shew
forth thy
praise to
all
generations.\FTNT{††}{{\FR 79.13 }
to all…: Heb. to generation and generation}

\par }\PsalmChap{80}{\D To the chief
Musician upon
Shoshannim-eduth, A
Psalm of
Asaph.
\par }{\Q \VerseOne{1}Give
ear, O
Shepherd of
Israel, thou that
leadest
Joseph like a
flock; thou that
dwellest
{\ADD{between}} the
cherubims, shine
forth.\FTNT{*}{{\FR 80.1 }
of: or, for}
\par }{\Q \VS{2}Before
Ephraim and
Benjamin and
Manasseh stir
up thy
strength, and
come
{\ADD{and}}
save us.\FTNT{†}{{\FR 80.2 }
come…: Heb. come for salvation to us}
\par }{\Q \VS{3}Turn us
again, O
God, and cause thy
face to
shine; and we shall be
saved.
\par }{\Q \VS{4}O
{\ND{
LORD}}
God of
hosts, how long wilt thou be
angry against the
prayer of thy
people?\FTNT{‡}{{\FR 80.4 }
be…: Heb. smoke}
\par }{\Q \VS{5}Thou
feedest them with the
bread of
tears; and givest them
tears to
drink in great
measure.
\par }{\Q \VS{6}Thou
makest us a
strife unto our
neighbours: and our
enemies
laugh among themselves.
\par }{\Q \VS{7}Turn us
again, O
God of
hosts, and cause thy
face to
shine; and we shall be
saved.
\par }{\BB \par }{\Q \VS{8}Thou hast
brought a
vine out of
Egypt: thou hast cast
out the
heathen, and
planted it.
\par }{\Q \VS{9}Thou
preparedst
{\ADD{room}}
before it, and didst cause it to take
deep
root, and it
filled the
land.
\par }{\Q \VS{10}The
hills were
covered with the
shadow of it, and the
boughs thereof
{\ADD{were like}} the
goodly
cedars.\FTNT{§}{{\FR 80.10 }
goodly…: Heb. cedars of God}
\par }{\Q \VS{11}She sent
out her
boughs unto the
sea, and her
branches unto the
river.
\par }{\Q \VS{12}Why hast thou
{\ADD{then}} broken
down her
hedges, so that all they which
pass by the
way do
pluck her?
\par }{\Q \VS{13}The
boar out of the
wood doth
waste it, and the wild
beast of the
field doth
devour it.
\par }{\Q \VS{14}Return, we beseech thee, O
God of
hosts: look
down from
heaven, and
behold, and
visit this
vine;
\par }{\Q \VS{15}And the
vineyard which thy right
hand hath
planted, and the
branch
{\ADD{that}} thou madest
strong for thyself.
\par }{\Q \VS{16}{\ADD{It is}}
burned with
fire,
{\ADD{it is}} cut
down: they
perish at the
rebuke of thy
countenance.
\par }{\Q \VS{17}Let thy
hand be upon the
man of thy right
hand, upon the
son of
man
{\ADD{whom}} thou madest
strong for thyself.
\par }{\Q \VS{18}So will not we go
back from thee:
quicken us, and we will
call upon thy
name.
\par }{\Q \VS{19}Turn us
again, O
{\ND{
LORD}}
God of
hosts, cause thy
face to
shine; and we shall be
saved.

\par }\PsalmChap{81}{\D To the chief
Musician upon
Gittith,
{\ADD{A Psalm}} of
Asaph.
\par }{\Q \VerseOne{1}Sing
aloud unto
God our
strength: make a joyful
noise unto the
God of
Jacob.\FTNT{*}{{\FR 81.1 }
of Asaph: or, for Asaph}
\par }{\Q \VS{2}Take a
psalm, and
bring hither the
timbrel, the
pleasant
harp with the
psaltery.
\par }{\Q \VS{3}Blow
up the
trumpet in the new
moon, in the time
appointed, on our solemn
feast
day.
\par }{\Q \VS{4}For this
{\ADD{was}} a
statute for
Israel,
{\ADD{and}} a
law of the
God of
Jacob.
\par }{\Q \VS{5}This he
ordained in
Joseph
{\ADD{for}} a
testimony, when he went
out through the
land of
Egypt:
{\ADD{where}} I
heard a
language
{\ADD{that}} I
understood not.\FTNT{†}{{\FR 81.5 }
through: or, against}
\par }{\Q \VS{6}I
removed his
shoulder from the
burden: his
hands were
delivered from the
pots.\FTNT{‡}{{\FR 81.6 }
were…: Heb. passed away}
\par }{\Q \VS{7}Thou
calledst in
trouble, and I
delivered thee; I
answered thee in the secret
place of
thunder: I
proved thee at the
waters of
Meribah.
Selah.\FTNT{§}{{\FR 81.7 }
Meribah: or, Strife}
\par }{\BB \par }{\Q \VS{8}Hear, O my
people, and I will
testify unto thee: O
Israel, if thou wilt
hearken unto me;
\par }{\Q \VS{9}There shall no
strange
god be in thee; neither shalt thou
worship any
strange
god.
\par }{\Q \VS{10}I
{\ADD{am}} the
{\ND{
LORD}} thy
God, which
brought thee out of the
land of
Egypt: open thy
mouth
wide, and I will
fill it.
\par }{\Q \VS{11}But my
people would not
hearken to my
voice; and
Israel
would none of me.
\par }{\Q \VS{12}So I gave them
up unto their own
hearts’
lust:
{\ADD{and}} they
walked in their own
counsels.\FTNT{**}{{\FR 81.12 }
unto…: or, to the hardness of their hearts, or, imagination}
\par }{\Q \VS{13}Oh that my
people had
hearkened unto me,
{\ADD{and}}
Israel had
walked in my
ways!
\par }{\Q \VS{14}I should
soon have
subdued their
enemies, and
turned my
hand against their
adversaries.
\par }{\Q \VS{15}The
haters of the
{\ND{
LORD}} should have
submitted themselves unto him: but their
time should have endured for
ever.\FTNT{††}{{\FR 81.15 }
submitted…: or, yielded feigned obedience: Heb. lied}
\par }{\Q \VS{16}He should have
fed them also with the
finest of the
wheat: and with
honey out of the
rock should I have
satisfied thee.\FTNT{‡‡}{{\FR 81.16 }
finest…: Heb. fat of wheat}

\par }\PsalmChap{82}{\D A
Psalm of
Asaph.
\par }{\Q \VerseOne{1}God
standeth in the
congregation of the
mighty; he
judgeth
among the
gods.\FTNT{*}{{\FR 82.1 }
of Asaph: or, for Asaph}
\par }{\Q \VS{2}How long will ye
judge
unjustly, and
accept the
persons of the
wicked?
Selah.
\par }{\Q \VS{3}Defend the
poor and
fatherless: do
justice to the
afflicted and
needy.\FTNT{†}{{\FR 82.3 }
Defend: Heb. Judge}
\par }{\Q \VS{4}Deliver the
poor and
needy:
rid
{\ADD{them}} out of the
hand of the
wicked.
\par }{\Q \VS{5}They
know not, neither will they
understand; they
walk on in
darkness: all the
foundations of the
earth are out of
course.\FTNT{‡}{{\FR 82.5 }
out…: Heb. moved}
\par }{\BB \par }{\Q \VS{6}I have
said, Ye
{\ADD{are}}
gods; and all of you
{\ADD{are}}
children of the most
High.
\par }{\Q \VS{7}But ye shall
die like
men, and
fall like
one of the
princes.
\par }{\Q \VS{8}Arise, O
God,
judge the
earth: for thou shalt
inherit all
nations.

\par }\PsalmChap{83}{\D A
Song
{\ADD{or}}
Psalm of
Asaph.
\par }{\Q \VerseOne{1}Keep not thou
silence, O
God: hold not thy
peace, and be not
still, O
God.\FTNT{*}{{\FR 83.1 }
of Asaph: or, for Asaph}
\par }{\Q \VS{2}For, lo, thine
enemies make a
tumult: and they that
hate thee have lifted
up the
head.
\par }{\Q \VS{3}They have taken
crafty
counsel against thy
people, and
consulted against thy hidden
ones.
\par }{\Q \VS{4}They have
said,
Come, and let us cut them
off from
{\ADD{being}} a
nation; that the
name of
Israel may be no more in
remembrance.
\par }{\Q \VS{5}For they have
consulted
together with one
consent: they
are
confederate against thee:\FTNT{†}{{\FR 83.5 }
consent: Heb. heart}
\par }{\Q \VS{6}The
tabernacles of
Edom, and the
Ishmaelites; of
Moab, and the
Hagarenes;
\par }{\Q \VS{7}Gebal, and
Ammon, and
Amalek; the
Philistines with the
inhabitants of
Tyre;
\par }{\Q \VS{8}Assur also is
joined with them: they have
holpen the
children of
Lot.
Selah.\FTNT{‡}{{\FR 83.8 }
holpen: Heb. been an arm to}
\par }{\BB \par }{\Q \VS{9}Do unto them as
{\ADD{unto}} the
Midianites; as
{\ADD{to}}
Sisera, as
{\ADD{to}}
Jabin, at the
brook of
Kison:
\par }{\Q \VS{10}{\ADD{Which}}
perished at
Endor: they became
{\ADD{as}}
dung for the
earth.
\par }{\Q \VS{11}Make their
nobles like
Oreb, and like
Zeeb: yea, all their
princes as
Zebah, and as
Zalmunna:
\par }{\Q \VS{12}Who
said, Let us take to ourselves the
houses of
God in
possession.
\par }{\Q \VS{13}O my
God,
make them like a
wheel; as the
stubble
before the
wind.
\par }{\Q \VS{14}As the
fire
burneth a
wood, and as the
flame
setteth the
mountains on
fire;
\par }{\Q \VS{15}So
persecute them with thy
tempest, and make them
afraid with thy
storm.
\par }{\Q \VS{16}Fill their
faces with
shame; that they may
seek thy
name, O
{\ND{
LORD}}.
\par }{\Q \VS{17}Let them be
confounded and
troubled for
ever; yea, let them be put to
shame, and
perish:
\par }{\Q \VS{18}That
{\ADD{men}} may
know that thou, whose
name alone
{\ADD{is}}
{\ND{
JEHOVAH}},
{\ADD{art}} the most
high over all the
earth.

\par }\PsalmChap{84}{\D To the chief
Musician upon
Gittith, A
Psalm for the
sons of
Korah.
\par }{\Q \VerseOne{1}How
amiable
{\ADD{are}} thy
tabernacles, O
{\ND{
LORD}} of
hosts!\FTNT{*}{{\FR 84.1 }
for the sons: or, of the sons}
\par }{\Q \VS{2}My
soul
longeth, yea, even
fainteth for the
courts of the
{\ND{
LORD}}: my
heart and my
flesh crieth
out for the
living
God.
\par }{\Q \VS{3}Yea, the
sparrow hath
found an
house, and the
swallow a
nest for herself, where she may
lay her
young,
{\ADD{even}} thine
altars, O
{\ND{
LORD}} of
hosts, my
King, and my
God.
\par }{\Q \VS{4}Blessed
{\ADD{are}} they that
dwell in thy
house: they will be still
praising thee.
Selah.
\par }{\Q \VS{5}Blessed
{\ADD{is}} the
man whose
strength
{\ADD{is}} in thee; in whose
heart
{\ADD{are}} the
ways
{\ADD{of them}}.
\par }{\Q \VS{6}{\ADD{Who}}
passing through the
valley of
Baca
make it a
well; the
rain also
filleth the
pools.\FTNT{†}{{\FR 84.6 }
Baca…: or, mulberry trees make him a well, etc}\FTNT{‡}{{\FR 84.6 }
filleth: Heb. covereth}
\par }{\Q \VS{7}They
go from
strength to
strength,
{\ADD{every one of them}} in
Zion
appeareth before
God.\FTNT{§}{{\FR 84.7 }
strength to…: or, company to company}
\par }{\BB \par }{\Q \VS{8}O
{\ND{
LORD}}
God of
hosts,
hear my
prayer: give
ear, O
God of
Jacob.
Selah.
\par }{\Q \VS{9}Behold, O
God our
shield, and
look upon the
face of thine
anointed.
\par }{\Q \VS{10}For a
day in thy
courts
{\ADD{is}}
better than a
thousand. I had
rather be a
doorkeeper in the
house of my
God, than to
dwell in the
tents of
wickedness.\FTNT{**}{{\FR 84.10 }
I had…: Heb. I would choose rather to sit at the threshold}
\par }{\Q \VS{11}For the
{\ND{
LORD}}
God
{\ADD{is}} a
sun and
shield: the
{\ND{
LORD}} will
give
grace and
glory: no
good
{\ADD{thing}} will he
withhold from them that
walk
uprightly.
\par }{\Q \VS{12}O
{\ND{
LORD}} of
hosts,
blessed
{\ADD{is}} the
man that
trusteth in thee.

\par }\PsalmChap{85}{\D To the chief
Musician, A
Psalm for the
sons of
Korah.
\par }{\Q \VerseOne{1}{\ND{
LORD}}, thou hast been
favourable unto thy
land: thou hast brought
back the
captivity of
Jacob.\FTNT{*}{{\FR 85.1 }
for the sons: or, of the sons}\FTNT{†}{{\FR 85.1 }
favourable: or, well pleased}
\par }{\Q \VS{2}Thou hast
forgiven the
iniquity of thy
people, thou hast
covered all their
sin.
Selah.
\par }{\Q \VS{3}Thou hast taken
away all thy
wrath: thou hast
turned
{\ADD{thyself}} from the
fierceness of thine
anger.\FTNT{‡}{{\FR 85.3 }
thyself…: or, thine anger from waxing hot}
\par }{\Q \VS{4}Turn us, O
God of our
salvation, and cause thine
anger toward us to
cease.
\par }{\Q \VS{5}Wilt thou be
angry with us for
ever? wilt thou draw
out thine
anger to
all
generations?
\par }{\Q \VS{6}Wilt thou not
revive us
again: that thy
people may
rejoice in thee?
\par }{\Q \VS{7}Shew us thy
mercy, O
{\ND{
LORD}}, and
grant us thy
salvation.
\par }{\BB \par }{\Q \VS{8}I will
hear what
God the
{\ND{
LORD}} will
speak: for he will
speak
peace unto his
people, and to his
saints: but let them not turn
again to
folly.
\par }{\Q \VS{9}Surely his
salvation
{\ADD{is}}
nigh them that
fear him; that
glory may
dwell in our
land.
\par }{\Q \VS{10}Mercy and
truth are met
together;
righteousness and
peace have
kissed
{\ADD{each other}}.
\par }{\Q \VS{11}Truth shall spring
out of the
earth; and
righteousness shall look
down from
heaven.
\par }{\Q \VS{12}Yea, the
{\ND{
LORD}} shall
give
{\ADD{that which is}}
good; and our
land shall
yield her
increase.
\par }{\Q \VS{13}Righteousness shall
go
before him; and shall
set
{\ADD{us}} in the
way of his
steps.

\par }\PsalmChap{86}{\D A
Prayer of
David.
\par }{\Q \VerseOne{1}Bow
down thine
ear, O
{\ND{
LORD}},
hear me: for I
{\ADD{am}}
poor and
needy.\FTNT{*}{{\FR 86.1 }
A Prayer…: or, A Prayer, being a Psalm of David}
\par }{\Q \VS{2}Preserve my
soul; for I
{\ADD{am}}
holy: O thou my
God,
save thy
servant that
trusteth in thee.\FTNT{†}{{\FR 86.2 }
holy: or, one whom thou favourest}
\par }{\Q \VS{3}Be
merciful unto me, O
Lord: for I
cry unto thee
daily.\FTNT{‡}{{\FR 86.3 }
daily: or, all the day}
\par }{\Q \VS{4}Rejoice the
soul of thy
servant: for unto thee, O
Lord, do I lift
up my
soul.
\par }{\Q \VS{5}For thou,
Lord,
{\ADD{art}}
good, and ready to
forgive; and
plenteous in
mercy unto all them that
call upon thee.
\par }{\Q \VS{6}Give
ear, O
{\ND{
LORD}}, unto my
prayer; and
attend to the
voice of my
supplications.
\par }{\Q \VS{7}In the
day of my
trouble I will
call upon thee: for thou wilt
answer me.
\par }{\BB \par }{\Q \VS{8}Among the
gods
{\ADD{there is}} none like unto thee, O
Lord; neither
{\ADD{are there any works}} like unto thy
works.
\par }{\Q \VS{9}All
nations whom thou hast
made shall
come and
worship
before thee, O
Lord; and shall
glorify thy
name.
\par }{\Q \VS{10}For thou
{\ADD{art}}
great, and
doest wondrous
things: thou
{\ADD{art}}
God alone.
\par }{\Q \VS{11}Teach me thy
way, O
{\ND{
LORD}}; I will
walk in thy
truth:
unite my
heart to
fear thy
name.
\par }{\Q \VS{12}I will
praise thee, O
Lord my
God, with all my
heart: and I will
glorify thy
name for
evermore.
\par }{\Q \VS{13}For
great
{\ADD{is}} thy
mercy toward me: and thou hast
delivered my
soul from the
lowest
hell.\FTNT{§}{{\FR 86.13 }
hell: or, grave}
\par }{\Q \VS{14}O
God, the
proud are
risen against me, and the
assemblies of
violent
{\ADD{men}} have
sought after my
soul; and have not
set thee before them.\FTNT{**}{{\FR 86.14 }
violent: Heb. terrible}
\par }{\Q \VS{15}But thou, O
Lord,
{\ADD{art}} a
God full of
compassion, and
gracious,
longsuffering, and
plenteous in
mercy and
truth.
\par }{\Q \VS{16}O
turn unto me, and have
mercy upon me;
give thy
strength unto thy
servant, and
save the
son of thine
handmaid.
\par }{\Q \VS{17}Shew me a
token for
good; that they which
hate me may
see
{\ADD{it}}, and be
ashamed: because thou,
{\ND{
LORD}}, hast
holpen me, and
comforted me.

\par }\PsalmChap{87}{\D A
Psalm
{\ADD{or}}
Song for the
sons of
Korah.
\par }{\Q \VerseOne{1}His
foundation
{\ADD{is}} in the
holy
mountains.\FTNT{*}{{\FR 87.1 }
for the sons: or, of the sons}
\par }{\Q \VS{2}The
{\ND{
LORD}}
loveth the
gates of
Zion more than all the
dwellings of
Jacob.
\par }{\Q \VS{3}Glorious
things are
spoken of thee, O
city of
God.
Selah.
\par }{\BB \par }{\Q \VS{4}I will make
mention of
Rahab and
Babylon to them that
know me: behold
Philistia, and
Tyre, with
Ethiopia; this
{\ADD{man}} was
born there.
\par }{\Q \VS{5}And of
Zion it shall be
said, This and that
man was
born in her: and the
highest himself shall
establish her.
\par }{\Q \VS{6}The
{\ND{
LORD}} shall
count, when he writeth
up the
people,
{\ADD{that}} this
{\ADD{man}} was
born there.
Selah.
\par }{\Q \VS{7}As well the
singers as the
players on instruments
{\ADD{shall be there:}} all my
springs
{\ADD{are}} in thee.

\par }\PsalmChap{88}{\D A
Song
{\ADD{or}}
Psalm for the
sons of
Korah, to the chief
Musician upon
Mahalath
Leannoth,
Maschil of
Heman the
Ezrahite.
\par }{\Q \VerseOne{1}O
{\ND{
LORD}}
God of my
salvation, I have
cried
day
{\ADD{and}}
night before thee:\FTNT{*}{{\FR 88.1 }
for the sons: or, of the sons}\FTNT{†}{{\FR 88.1 }
Maschil…: or, A Psalm of Heman the Ezrahite, giving instruction}
\par }{\Q \VS{2}Let my
prayer
come
before thee:
incline thine
ear unto my
cry;
\par }{\Q \VS{3}For my
soul is
full of
troubles: and my
life draweth
nigh unto the
grave.
\par }{\Q \VS{4}I am
counted with them that go
down into the
pit: I am as a
man
{\ADD{that hath}} no
strength:
\par }{\Q \VS{5}Free among the
dead, like the
slain that
lie in the
grave, whom thou
rememberest no more: and they are cut
off from thy
hand.\FTNT{‡}{{\FR 88.5 }
from: or, by}
\par }{\Q \VS{6}Thou hast
laid me in the
lowest
pit, in
darkness, in the
deeps.
\par }{\Q \VS{7}Thy
wrath lieth
hard upon me, and thou hast
afflicted
{\ADD{me}} with all thy
waves.
Selah.
\par }{\Q \VS{8}Thou hast put
away mine
acquaintance
far from me; thou hast
made me an
abomination unto them:
{\ADD{I am}} shut
up, and I cannot come
forth.
\par }{\Q \VS{9}Mine
eye
mourneth by reason of
affliction:
{\ND{
LORD}}, I have
called
daily upon thee, I have stretched
out my
hands unto thee.
\par }{\BB \par }{\Q \VS{10}Wilt thou
shew
wonders to the
dead? shall the
dead
arise
{\ADD{and}}
praise thee?
Selah.
\par }{\Q \VS{11}Shall thy
lovingkindness be
declared in the
grave?
{\ADD{or}} thy
faithfulness in
destruction?
\par }{\Q \VS{12}Shall thy
wonders be
known in the
dark? and thy
righteousness in the
land of
forgetfulness?
\par }{\Q \VS{13}But unto thee have I
cried, O
{\ND{
LORD}}; and in the
morning shall my
prayer
prevent thee.
\par }{\Q \VS{14}{\ND{
LORD}}, why castest thou
off my
soul?
{\ADD{why}}
hidest thou thy
face from me?
\par }{\Q \VS{15}I
{\ADD{am}}
afflicted and ready to
die from
{\ADD{my}}
youth up:
{\ADD{while}} I
suffer thy
terrors I am
distracted.
\par }{\Q \VS{16}Thy fierce
wrath goeth
over me; thy
terrors have cut me
off.
\par }{\Q \VS{17}They came round
about me
daily like
water; they
compassed me about
together.\FTNT{§}{{\FR 88.17 }
daily: or, all the day}
\par }{\Q \VS{18}Lover and
friend hast thou put
far from me,
{\ADD{and}} mine
acquaintance into
darkness.

\par }\PsalmChap{89}{\D Maschil of
Ethan the
Ezrahite.
\par }{\Q \VerseOne{1}I will
sing of the
mercies of the
{\ND{
LORD}} for
ever: with my
mouth will I make
known thy
faithfulness to
all
generations.\FTNT{*}{{\FR 89.1 }
Maschil…: or, A Psalm for Ethan the Ezrahite, to give instruction}\FTNT{†}{{\FR 89.1 }
to all…: Heb. to generation and generation}
\par }{\Q \VS{2}For I have
said,
Mercy shall be built
up for
ever: thy
faithfulness shalt thou
establish in the very
heavens.
\par }{\Q \VS{3}I have
made a
covenant with my
chosen, I have
sworn unto
David my
servant,
\par }{\Q \VS{4}Thy
seed will I
establish
for
ever, and build
up thy
throne to
all
generations.
Selah.
\par }{\BB \par }{\Q \VS{5}And the
heavens shall
praise thy
wonders, O
{\ND{
LORD}}: thy
faithfulness also in the
congregation of the
saints.
\par }{\Q \VS{6}For who in the
heaven can be
compared unto the
{\ND{
LORD}}?
{\ADD{who}} among the
sons of the
mighty can be
likened unto the
{\ND{
LORD}}?
\par }{\Q \VS{7}God is
greatly to be
feared in the
assembly of the
saints, and to be had in
reverence of all
{\ADD{them that are}}
about him.
\par }{\Q \VS{8}O
{\ND{
LORD}}
God of
hosts, who
{\ADD{is}} a
strong
{\ND{
LORD}} like unto thee? or to thy
faithfulness round
about thee?
\par }{\Q \VS{9}Thou
rulest the
raging of the
sea: when the
waves thereof
arise, thou
stillest them.
\par }{\Q \VS{10}Thou hast
broken
Rahab in
pieces, as one that is
slain; thou hast
scattered thine
enemies with thy
strong
arm.\FTNT{‡}{{\FR 89.10 }
Rahab: or, Egypt}\FTNT{§}{{\FR 89.10 }
thy…: Heb. the arm of thy strength}
\par }{\Q \VS{11}The
heavens
{\ADD{are}} thine, the
earth also
{\ADD{is}} thine:
{\ADD{as for}} the
world and the
fulness thereof, thou hast
founded them.\FTNT{**}{{\FR 89.11 }
the fulness…: or, all it containeth}
\par }{\Q \VS{12}The
north and the
south thou hast
created them:
Tabor and
Hermon shall
rejoice in thy
name.
\par }{\Q \VS{13}Thou hast a
mighty
arm:
strong is thy
hand,
{\ADD{and}}
high is thy right
hand.\FTNT{††}{{\FR 89.13 }
a…: Heb. an arm with might}
\par }{\Q \VS{14}Justice and
judgment
{\ADD{are}} the
habitation of thy
throne:
mercy and
truth shall
go before thy
face.\FTNT{‡‡}{{\FR 89.14 }
habitation: or, establishment}
\par }{\BB \par }{\Q \VS{15}Blessed
{\ADD{is}} the
people that
know the joyful
sound: they shall
walk, O
{\ND{
LORD}}, in the
light of thy
countenance.
\par }{\Q \VS{16}In thy
name shall they
rejoice all the
day: and in thy
righteousness shall they be
exalted.
\par }{\Q \VS{17}For thou
{\ADD{art}} the
glory of their
strength: and in thy
favour our
horn shall be
exalted.
\par }{\Q \VS{18}For the
{\ND{
LORD}}
{\ADD{is}} our
defence; and the Holy
One of
Israel
{\ADD{is}} our
king.\FTNT{§§}{{\FR 89.18 }
the
{\ND{LORD}}…: or, our shield is of the
{\ND{LORD}}, and our king is of the Holy One of Israel}
\par }{\BB \par }{\Q \VS{19}Then thou
spakest in
vision to thy holy
one, and
saidst, I have
laid
help upon
{\ADD{one that is}}
mighty; I have
exalted
{\ADD{one}}
chosen out of the
people.
\par }{\Q \VS{20}I have
found
David my
servant; with my
holy
oil have I
anointed him:
\par }{\Q \VS{21}With whom my
hand shall be
established: mine
arm also shall
strengthen him.
\par }{\Q \VS{22}The
enemy shall not
exact upon him; nor the
son of
wickedness
afflict him.
\par }{\Q \VS{23}And I will beat
down his
foes before his
face, and
plague them that
hate him.
\par }{\Q \VS{24}But my
faithfulness and my
mercy
{\ADD{shall be}} with him: and in my
name shall his
horn be
exalted.
\par }{\Q \VS{25}I will
set his
hand also in the
sea, and his right
hand in the
rivers.
\par }{\Q \VS{26}He shall
cry unto me, Thou
{\ADD{art}} my
father, my
God, and the
rock of my
salvation.
\par }{\Q \VS{27}Also I will
make him
{\ADD{my}}
firstborn,
higher than the
kings of the
earth.
\par }{\Q \VS{28}My
mercy will I
keep for him for
evermore, and my
covenant shall stand
fast with him.
\par }{\Q \VS{29}His
seed also will I
make
{\ADD{to endure}} for
ever, and his
throne as the
days of
heaven.
\par }{\Q \VS{30}If his
children
forsake my
law, and
walk not in my
judgments;
\par }{\Q \VS{31}If they
break my
statutes, and
keep not my
commandments;\FTNT{***}{{\FR 89.31 }
break: Heb. profane}
\par }{\Q \VS{32}Then will I
visit their
transgression with the
rod, and their
iniquity with
stripes.
\par }{\Q \VS{33}Nevertheless my
lovingkindness will I not utterly
take from him, nor suffer my
faithfulness to
fail.\FTNT{†††}{{\FR 89.33 }
will…: Heb. I will not make void from him}\FTNT{‡‡‡}{{\FR 89.33 }
to fail: Heb. to lie}
\par }{\Q \VS{34}My
covenant will I not
break, nor
alter the thing that is gone
out of my
lips.
\par }{\Q \VS{35}Once have I
sworn by my
holiness that I will not
lie unto
David.\FTNT{§§§}{{\FR 89.35 }
that…: Heb. if I lie}
\par }{\Q \VS{36}His
seed shall endure for
ever, and his
throne as the
sun before me.
\par }{\Q \VS{37}It shall be
established for
ever as the
moon, and
{\ADD{as}} a
faithful
witness in
heaven.
Selah.
\par }{\BB \par }{\Q \VS{38}But thou hast cast
off and
abhorred, thou hast been
wroth with thine
anointed.
\par }{\Q \VS{39}Thou hast made
void the
covenant of thy
servant: thou hast
profaned his
crown
{\ADD{by casting it}} to the
ground.
\par }{\Q \VS{40}Thou hast broken
down all his
hedges; thou hast
brought his strong
holds to
ruin.
\par }{\Q \VS{41}All that pass
by the
way
spoil him: he is a
reproach to his
neighbours.
\par }{\Q \VS{42}Thou hast set
up the right
hand of his
adversaries; thou hast made all his
enemies to
rejoice.
\par }{\Q \VS{43}Thou hast also
turned the
edge of his
sword, and hast not made him to
stand in the
battle.
\par }{\Q \VS{44}Thou hast made his
glory to
cease, and
cast his
throne
down to the
ground.\FTNT{*}{{\FR 89.44 }
glory: Heb. brightness}
\par }{\Q \VS{45}The
days of his
youth hast thou
shortened: thou hast
covered him with
shame.
Selah.
\par }{\Q \VS{46}How long,
{\ND{
LORD}}? wilt thou
hide thyself for
ever? shall thy
wrath
burn like
fire?
\par }{\Q \VS{47}Remember how
short my time is: wherefore hast thou
made all
men in
vain?
\par }{\Q \VS{48}What
man
{\ADD{is he that}}
liveth, and shall not
see
death? shall he
deliver his
soul from the
hand of the
grave?
Selah.
\par }{\Q \VS{49}Lord, where
{\ADD{are}} thy
former
lovingkindnesses,
{\ADD{which}} thou
swarest unto
David in thy
truth?
\par }{\Q \VS{50}Remember,
Lord, the
reproach of thy
servants;
{\ADD{how}} I do
bear in my
bosom
{\ADD{the reproach of}} all the
mighty
people;
\par }{\Q \VS{51}Wherewith thine
enemies have
reproached, O
{\ND{
LORD}}; wherewith they have
reproached the
footsteps of thine
anointed.
\par }{\Q \VS{52}Blessed
{\ADD{be}} the
{\ND{
LORD}} for
evermore.
Amen, and
Amen.

\par }\PsalmChap{90}{\D A
Prayer of
Moses the
man of
God.
\par }{\Q \VerseOne{1}Lord, thou hast been our dwelling
place in
all
generations.\FTNT{*}{{\FR 90.1 }
A Prayer…: or, A Prayer, being a Psalm of Moses}\FTNT{†}{{\FR 90.1 }
in…: Heb. in generation and generation}
\par }{\Q \VS{2}Before the
mountains were brought
forth, or ever thou hadst
formed the
earth and the
world, even from
everlasting
to
everlasting, thou
{\ADD{art}}
God.
\par }{\Q \VS{3}Thou
turnest
man to
destruction; and
sayest,
Return, ye
children of
men.
\par }{\Q \VS{4}For a
thousand
years in thy
sight
{\ADD{are but}}
as
yesterday when it is
past, and
{\ADD{as}} a
watch in the
night.\FTNT{‡}{{\FR 90.4 }
when…: or, when he hath passed them}
\par }{\Q \VS{5}Thou carriest them away as with a
flood; they are
{\ADD{as}} a
sleep: in the
morning
{\ADD{they are}} like
grass
{\ADD{which}} groweth
up.\FTNT{§}{{\FR 90.5 }
groweth…: or, is changed}
\par }{\Q \VS{6}In the
morning it
flourisheth, and groweth
up; in the
evening it is cut
down, and
withereth.
\par }{\BB \par }{\Q \VS{7}For we are
consumed by thine
anger, and by thy
wrath are we
troubled.
\par }{\Q \VS{8}Thou hast
set our
iniquities before thee, our
secret
{\ADD{sins}} in the
light of thy
countenance.
\par }{\Q \VS{9}For all our
days are passed
away in thy
wrath: we
spend our
years as a
tale
{\ADD{that is told}}.\FTNT{**}{{\FR 90.9 }
passed…: Heb. turned away}\FTNT{††}{{\FR 90.9 }
as a…: or, as a meditation}
\par }{\Q \VS{10}The
days of our
years
{\ADD{are}}
threescore
years and
ten; and if by reason of
strength
{\ADD{they be}}
fourscore
years, yet
{\ADD{is}} their
strength
labour and
sorrow; for it is
soon cut
off, and we fly
away.\FTNT{‡‡}{{\FR 90.10 }
The days…: Heb. As for the days of our years, in them are seventy years}
\par }{\Q \VS{11}Who
knoweth the
power of thine
anger? even according to thy
fear,
{\ADD{so is}} thy
wrath.
\par }{\BB \par }{\Q \VS{12}So
teach
{\ADD{us}} to
number our
days, that we may
apply
{\ADD{our}}
hearts unto
wisdom.\FTNT{§§}{{\FR 90.12 }
apply: Heb. cause to come}
\par }{\Q \VS{13}Return, O
{\ND{
LORD}}, how long? and let it
repent thee concerning thy
servants.
\par }{\Q \VS{14}O
satisfy us
early with thy
mercy; that we may
rejoice and be
glad all our
days.
\par }{\Q \VS{15}Make us
glad according to the
days
{\ADD{wherein}} thou hast
afflicted us,
{\ADD{and}} the
years
{\ADD{wherein}} we have
seen
evil.
\par }{\Q \VS{16}Let thy
work
appear unto thy
servants, and thy
glory unto their
children.
\par }{\Q \VS{17}And let the
beauty of the
{\ND{
LORD}} our
God be upon us: and
establish thou the
work of our
hands upon us; yea, the
work of our
hands
establish thou it.

\par }\PsalmChap{91}{\Q \VerseOne{1}He that
dwelleth in the
secret place of the most
High shall
abide under the
shadow of the
Almighty.\FTNT{*}{{\FR 91.1 }
abide: Heb. lodge}
\par }{\Q \VS{2}I will
say of the
{\ND{
LORD}},
{\ADD{He is}} my
refuge and my
fortress: my
God; in him will I
trust.
\par }{\Q \VS{3}Surely he shall
deliver thee from the
snare of the
fowler,
{\ADD{and}} from the
noisome
pestilence.
\par }{\Q \VS{4}He shall
cover thee with his
feathers, and under his
wings shalt thou
trust: his
truth
{\ADD{shall be thy}}
shield and
buckler.
\par }{\Q \VS{5}Thou shalt not be
afraid for the
terror by
night;
{\ADD{nor}} for the
arrow
{\ADD{that}}
flieth by
day;
\par }{\Q \VS{6}{\ADD{Nor}} for the
pestilence
{\ADD{that}}
walketh in
darkness;
{\ADD{nor}} for the
destruction
{\ADD{that}}
wasteth at
noonday.
\par }{\Q \VS{7}A
thousand shall
fall at thy
side, and ten
thousand at thy right
hand;
{\ADD{but}} it shall not come
nigh thee.
\par }{\Q \VS{8}Only with thine
eyes shalt thou
behold and
see the
reward of the
wicked.
\par }{\BB \par }{\Q \VS{9}Because thou hast
made the
{\ND{
LORD}},
{\ADD{which is}} my
refuge,
{\ADD{even}} the most
High, thy
habitation;
\par }{\Q \VS{10}There shall no
evil
befall thee, neither shall any
plague come
nigh thy
dwelling.
\par }{\Q \VS{11}For he shall give his
angels
charge over thee, to
keep thee in all thy
ways.
\par }{\Q \VS{12}They shall bear thee
up in
{\ADD{their}}
hands, lest thou
dash thy
foot against a
stone.
\par }{\Q \VS{13}Thou shalt
tread upon the
lion and
adder: the young
lion and the
dragon shalt thou trample under
feet.\FTNT{†}{{\FR 91.13 }
adder: or, asp}
\par }{\Q \VS{14}Because he hath set his
love upon me, therefore will I
deliver him: I will set him on
high, because he hath
known my
name.
\par }{\Q \VS{15}He shall
call upon me, and I will
answer him: I
{\ADD{will be}} with him in
trouble; I will
deliver him, and
honour him.
\par }{\Q \VS{16}With
long
life will I
satisfy him, and
shew him my
salvation.\FTNT{‡}{{\FR 91.16 }
long…: Heb. length of days}

\par }\PsalmChap{92}{\D A
Psalm
{\ADD{or}}
Song for the
sabbath
day.
\par }{\Q \VerseOne{1}{\ADD{It is a}}
good
{\ADD{thing}} to give
thanks unto the
{\ND{
LORD}}, and to sing
praises unto thy
name, O most
High:
\par }{\Q \VS{2}To shew
forth thy
lovingkindness in the
morning, and thy
faithfulness every
night,\FTNT{*}{{\FR 92.2 }
every…: Heb. in the nights}
\par }{\Q \VS{3}Upon an instrument of ten
strings, and upon the
psaltery; upon the
harp with a solemn
sound.\FTNT{†}{{\FR 92.3 }
the harp…: or, the solemn sound with the harp}\FTNT{‡}{{\FR 92.3 }
a solemn…: Heb. Higgaion}
\par }{\Q \VS{4}For thou,
{\ND{
LORD}}, hast made me
glad through thy
work: I will
triumph in the
works of thy
hands.
\par }{\Q \VS{5}O
{\ND{
LORD}}, how
great are thy
works!
{\ADD{and}} thy
thoughts are
very
deep.
\par }{\Q \VS{6}A
brutish
man
knoweth not; neither doth a
fool
understand this.
\par }{\BB \par }{\Q \VS{7}When the
wicked
spring as the
grass, and when all the
workers of
iniquity do
flourish;
{\ADD{it is}} that they shall be
destroyed for
ever:
\par }{\Q \VS{8}But thou,
{\ND{
LORD}},
{\ADD{art most}}
high for
evermore.
\par }{\Q \VS{9}For, lo, thine
enemies, O
{\ND{
LORD}}, for, lo, thine
enemies shall
perish; all the
workers of
iniquity shall be
scattered.
\par }{\Q \VS{10}But my
horn shalt thou
exalt like
{\ADD{the horn of}} an
unicorn: I shall be
anointed with
fresh
oil.
\par }{\Q \VS{11}Mine
eye also shall
see
{\ADD{my desire}} on mine
enemies,
{\ADD{and}} mine
ears shall
hear
{\ADD{my desire}} of the
wicked that rise
up against me.
\par }{\Q \VS{12}The
righteous shall
flourish like the palm
tree: he shall
grow like a
cedar in
Lebanon.
\par }{\Q \VS{13}Those that be
planted in the
house of the
{\ND{
LORD}} shall
flourish in the
courts of our
God.
\par }{\Q \VS{14}They shall still bring forth
fruit in old
age; they shall be
fat and
flourishing;\FTNT{§}{{\FR 92.14 }
flourishing: Heb. green}
\par }{\Q \VS{15}To
shew that the
{\ND{
LORD}}
{\ADD{is}}
upright:
{\ADD{he is}} my
rock, and
{\ADD{there is}} no
unrighteousness in him.

\par }\PsalmChap{93}{\Q \VerseOne{1}The
{\ND{
LORD}}
reigneth, he is
clothed with
majesty; the
{\ND{
LORD}} is
clothed with
strength,
{\ADD{wherewith}} he hath
girded himself: the
world also is
stablished, that it cannot be
moved.
\par }{\Q \VS{2}Thy
throne
{\ADD{is}}
established of
old: thou
{\ADD{art}} from
everlasting.\FTNT{*}{{\FR 93.2 }
of old: Heb. from then}
\par }{\Q \VS{3}The
floods have lifted
up, O
{\ND{
LORD}}, the
floods have lifted
up their
voice; the
floods lift
up their
waves.
\par }{\Q \VS{4}The
{\ND{
LORD}} on
high
{\ADD{is}}
mightier than the
noise of
many
waters,
{\ADD{yea, than}} the
mighty
waves of the
sea.
\par }{\Q \VS{5}Thy
testimonies are
very
sure:
holiness
becometh thine
house, O
{\ND{
LORD}}, for
ever.\FTNT{†}{{\FR 93.5 }
for ever: Heb. to length of days}

\par }\PsalmChap{94}{\Q \VerseOne{1}O
{\ND{
LORD}}
God, to whom
vengeance belongeth; O
God, to whom
vengeance belongeth,
shew thyself.\FTNT{*}{{\FR 94.1 }
God…: Heb. God of revenges}\FTNT{†}{{\FR 94.1 }
shew…: Heb. shine forth}
\par }{\Q \VS{2}Lift
up thyself, thou
judge of the
earth:
render a
reward to the
proud.
\par }{\Q \VS{3}{\ND{
LORD}}, how long shall the
wicked, how long shall the
wicked
triumph?
\par }{\Q \VS{4}{\ADD{How long}} shall they
utter
{\ADD{and}}
speak hard
things?
{\ADD{and}} all the
workers of
iniquity
boast themselves?
\par }{\Q \VS{5}They break in
pieces thy
people, O
{\ND{
LORD}}, and
afflict thine
heritage.
\par }{\Q \VS{6}They
slay the
widow and the
stranger, and
murder the
fatherless.
\par }{\Q \VS{7}Yet they
say, The
{\ND{
LORD}} shall not
see, neither shall the
God of
Jacob
regard
{\ADD{it}}.
\par }{\Q \VS{8}Understand, ye
brutish among the
people: and
{\ADD{ye}}
fools, when will ye be
wise?
\par }{\Q \VS{9}He that
planted the
ear, shall he not
hear? he that
formed the
eye, shall he not
see?
\par }{\Q \VS{10}He that
chastiseth the
heathen, shall not he
correct? he that
teacheth
man
knowledge,
{\ADD{shall not he know}}?
\par }{\Q \VS{11}The
{\ND{
LORD}}
knoweth the
thoughts of
man, that they
{\ADD{are}}
vanity.
\par }{\BB \par }{\Q \VS{12}Blessed
{\ADD{is}} the
man whom thou
chastenest, O
{\ND{
LORD}}, and
teachest him out of thy
law;
\par }{\Q \VS{13}That thou mayest give him
rest from the
days of
adversity, until the
pit be
digged for the
wicked.
\par }{\Q \VS{14}For the
{\ND{
LORD}} will not cast
off his
people, neither will he
forsake his
inheritance.
\par }{\Q \VS{15}But
judgment shall
return unto
righteousness: and all the
upright in
heart shall
follow it.\FTNT{‡}{{\FR 94.15 }
shall follow…: Heb. shall be after it}
\par }{\Q \VS{16}Who will rise
up for me against the
evildoers?
{\ADD{or}} who will stand
up for me against the
workers of
iniquity?
\par }{\Q \VS{17}Unless the
{\ND{
LORD}}
{\ADD{had been}} my
help, my
soul had
almost
dwelt in
silence.\FTNT{§}{{\FR 94.17 }
almost: or, quickly}
\par }{\Q \VS{18}When I
said, My
foot
slippeth; thy
mercy, O
{\ND{
LORD}}, held me
up.
\par }{\Q \VS{19}In the
multitude of my
thoughts
within me thy
comforts
delight my
soul.
\par }{\Q \VS{20}Shall the
throne of
iniquity have
fellowship with thee, which
frameth
mischief by a
law?
\par }{\Q \VS{21}They gather themselves
together against the
soul of the
righteous, and
condemn the
innocent
blood.
\par }{\Q \VS{22}But the
{\ND{
LORD}} is my
defence; and my
God
{\ADD{is}} the
rock of my
refuge.
\par }{\Q \VS{23}And he shall
bring upon them their own
iniquity, and shall cut them
off in their own
wickedness;
{\ADD{yea}}, the
{\ND{
LORD}} our
God shall cut them
off.

\par }\PsalmChap{95}{\Q \VerseOne{1}O
come, let us
sing unto the
{\ND{
LORD}}: let us make a joyful
noise to the
rock of our
salvation.
\par }{\Q \VS{2}Let us
come before his
presence with
thanksgiving, and make a joyful
noise unto him with
psalms.\FTNT{*}{{\FR 95.2 }
come…: Heb. prevent his face}
\par }{\Q \VS{3}For the
{\ND{
LORD}}
{\ADD{is}} a
great
God, and a
great
King above all
gods.
\par }{\Q \VS{4}In his
hand
{\ADD{are}} the deep
places of the
earth: the
strength of the
hills
{\ADD{is}} his also.\FTNT{†}{{\FR 95.4 }
In…: Heb. In whose}\FTNT{‡}{{\FR 95.4 }
the strength…: or, the heights of the hills are his}
\par }{\Q \VS{5}The
sea
{\ADD{is}} his, and he
made it: and his
hands
formed the
dry
{\ADD{land}}.\FTNT{§}{{\FR 95.5 }
The sea…: Heb. Whose the sea is}
\par }{\Q \VS{6}O
come, let us
worship and bow
down: let us
kneel
before the
{\ND{
LORD}} our
maker.
\par }{\BB \par }{\Q \VS{7}For he
{\ADD{is}} our
God; and we
{\ADD{are}} the
people of his
pasture, and the
sheep of his
hand. To
day if ye will
hear his
voice,
\par }{\Q \VS{8}Harden not your
heart, as in the
provocation,
{\ADD{and}} as
{\ADD{in}} the
day of
temptation in the
wilderness:\FTNT{**}{{\FR 95.8 }
provocation: Heb. contention}
\par }{\Q \VS{9}When your
fathers
tempted me,
proved me, and
saw my
work.
\par }{\Q \VS{10}Forty
years long was I
grieved with
{\ADD{this}}
generation, and
said, It
{\ADD{is}} a
people that do
err in their
heart, and they have not
known my
ways:
\par }{\Q \VS{11}Unto whom I
sware in my
wrath that they should not
enter into my
rest.\FTNT{††}{{\FR 95.11 }
that…: Heb. if they enter into my rest}

\par }\PsalmChap{96}{\Q \VerseOne{1}O
sing unto the
{\ND{
LORD}} a
new
song:
sing unto the
{\ND{
LORD}}, all the
earth.
\par }{\Q \VS{2}Sing unto the
{\ND{
LORD}},
bless his
name; shew
forth his
salvation from
day to
day.
\par }{\Q \VS{3}Declare his
glory among the
heathen, his
wonders among all
people.
\par }{\Q \VS{4}For the
{\ND{
LORD}}
{\ADD{is}}
great, and
greatly to be
praised: he
{\ADD{is}} to be
feared above all
gods.
\par }{\Q \VS{5}For all the
gods of the
nations
{\ADD{are}}
idols: but the
{\ND{
LORD}}
made the
heavens.
\par }{\Q \VS{6}Honour and
majesty
{\ADD{are}}
before him:
strength and
beauty
{\ADD{are}} in his
sanctuary.
\par }{\Q \VS{7}Give unto the
{\ND{
LORD}}, O ye
kindreds of the
people,
give unto the
{\ND{
LORD}}
glory and
strength.
\par }{\Q \VS{8}Give unto the
{\ND{
LORD}} the
glory
{\ADD{due unto}} his
name:
bring an
offering, and
come into his
courts.\FTNT{*}{{\FR 96.8 }
due…: Heb. of his name}
\par }{\Q \VS{9}O
worship the
{\ND{
LORD}} in the
beauty of
holiness:
fear
before him, all the
earth.\FTNT{†}{{\FR 96.9 }
in the…: or, in the glorious sanctuary}
\par }{\BB \par }{\Q \VS{10}Say among the
heathen
{\ADD{that}} the
{\ND{
LORD}}
reigneth: the
world also shall be
established that it shall not be
moved: he shall
judge the
people
righteously.
\par }{\Q \VS{11}Let the
heavens
rejoice, and let the
earth be
glad; let the
sea
roar, and the
fulness thereof.\FTNT{‡}{{\FR 96.11 }
the fulness…: or, all it containeth}
\par }{\Q \VS{12}Let the
field be
joyful, and all that
{\ADD{is}} therein: then shall all the
trees of the
wood
rejoice
\par }{\Q \VS{13}Before the
{\ND{
LORD}}: for he
cometh, for he
cometh to
judge the
earth: he shall
judge the
world with
righteousness, and the
people with his
truth.

\par }\PsalmChap{97}{\Q \VerseOne{1}The
{\ND{
LORD}}
reigneth; let the
earth
rejoice; let the
multitude of
isles be
glad
{\ADD{thereof}}.\FTNT{*}{{\FR 97.1 }
multitude…: Heb. many, or, great isles}
\par }{\Q \VS{2}Clouds and
darkness
{\ADD{are}} round
about him:
righteousness and
judgment
{\ADD{are}} the
habitation of his
throne.\FTNT{†}{{\FR 97.2 }
habitation: or, establishment}
\par }{\Q \VS{3}A
fire
goeth
before him, and burneth
up his
enemies round
about.
\par }{\Q \VS{4}His
lightnings
enlightened the
world: the
earth
saw, and
trembled.
\par }{\Q \VS{5}The
hills
melted like
wax at the
presence of the
{\ND{
LORD}}, at the
presence of the
Lord of the whole
earth.
\par }{\Q \VS{6}The
heavens
declare his
righteousness, and all the
people
see his
glory.
\par }{\Q \VS{7}Confounded be all they that
serve graven
images, that
boast themselves of
idols:
worship him, all
{\ADD{ye}}
gods.
\par }{\BB \par }{\Q \VS{8}Zion
heard, and was
glad; and the
daughters of
Judah
rejoiced because of thy
judgments, O
{\ND{
LORD}}.
\par }{\Q \VS{9}For thou,
{\ND{
LORD}},
{\ADD{art}} high
above all the
earth: thou art
exalted far
above all
gods.
\par }{\Q \VS{10}Ye that
love the
{\ND{
LORD}},
hate
evil: he
preserveth the
souls of his
saints; he
delivereth them out of the
hand of the
wicked.
\par }{\Q \VS{11}Light is
sown for the
righteous, and
gladness for the
upright in
heart.
\par }{\Q \VS{12}Rejoice in the
{\ND{
LORD}}, ye
righteous; and give
thanks at the
remembrance of his
holiness.\FTNT{‡}{{\FR 97.12 }
at…: or, to the memorial}

\par }\PsalmChap{98}{\D A
Psalm.
\par }{\Q \VerseOne{1}O
sing unto the
{\ND{
LORD}} a
new
song; for he hath
done marvellous
things: his right
hand, and his
holy
arm, hath gotten him the
victory.
\par }{\Q \VS{2}The
{\ND{
LORD}} hath made
known his
salvation: his
righteousness hath he openly
shewed in the
sight of the
heathen.\FTNT{*}{{\FR 98.2 }
openly…: or, revealed}
\par }{\Q \VS{3}He hath
remembered his
mercy and his
truth toward the
house of
Israel: all the
ends of the
earth have
seen the
salvation of our
God.
\par }{\BB \par }{\Q \VS{4}Make a joyful
noise unto the
{\ND{
LORD}}, all the
earth: make a loud
noise, and
rejoice, and
sing praise.
\par }{\Q \VS{5}Sing unto the
{\ND{
LORD}} with the
harp; with the
harp, and the
voice of a
psalm.
\par }{\Q \VS{6}With
trumpets and
sound of
cornet make a joyful
noise
before the
{\ND{
LORD}}, the
King.
\par }{\Q \VS{7}Let the
sea
roar, and the
fulness thereof; the
world, and they that
dwell therein.\FTNT{†}{{\FR 98.7 }
the fulness…: or, all it containeth}
\par }{\Q \VS{8}Let the
floods
clap
{\ADD{their}}
hands: let the
hills be
joyful
together
\par }{\Q \VS{9}Before the
{\ND{
LORD}}; for he
cometh to
judge the
earth: with
righteousness shall he
judge the
world, and the
people with
equity.

\par }\PsalmChap{99}{\Q \VerseOne{1}The
{\ND{
LORD}}
reigneth; let the
people
tremble: he
sitteth
{\ADD{between}} the
cherubims; let the
earth be
moved.\FTNT{*}{{\FR 99.1 }
be moved: Heb. stagger}
\par }{\Q \VS{2}The
{\ND{
LORD}}
{\ADD{is}}
great in
Zion; and he
{\ADD{is}}
high above all the
people.
\par }{\Q \VS{3}Let them
praise thy
great and
terrible
name;
{\ADD{for}} it
{\ADD{is}}
holy.
\par }{\Q \VS{4}The
king’s
strength also
loveth
judgment; thou dost
establish
equity, thou
executest
judgment and
righteousness in
Jacob.
\par }{\Q \VS{5}Exalt ye the
{\ND{
LORD}} our
God, and
worship at his
footstool;
{\ADD{for}} he
{\ADD{is}}
holy.\FTNT{†}{{\FR 99.5 }
he is…: or, it is holy}
\par }{\BB \par }{\Q \VS{6}Moses and
Aaron among his
priests, and
Samuel among them that
call upon his
name; they
called upon the
{\ND{
LORD}}, and he
answered them.
\par }{\Q \VS{7}He
spake unto them in the
cloudy
pillar: they
kept his
testimonies, and the
ordinance
{\ADD{that}} he
gave them.
\par }{\Q \VS{8}Thou
answeredst them, O
{\ND{
LORD}} our
God: thou wast a
God that
forgavest them, though thou tookest
vengeance of their
inventions.
\par }{\Q \VS{9}Exalt the
{\ND{
LORD}} our
God, and
worship at his
holy
hill; for the
{\ND{
LORD}} our
God
{\ADD{is}}
holy.

\par }\PsalmChap{100}{\D A
Psalm of
praise.
\par }{\Q \VerseOne{1}Make a joyful
noise unto the
{\ND{
LORD}}, all ye
lands.\FTNT{*}{{\FR 100.1 }
praise: or, thanksgiving}\FTNT{†}{{\FR 100.1 }
all…: Heb. all the earth}
\par }{\Q \VS{2}Serve the
{\ND{
LORD}} with
gladness:
come before his
presence with
singing.
\par }{\Q \VS{3}Know ye that the
{\ND{
LORD}} he
{\ADD{is}}
God:
{\ADD{it is}} he
{\ADD{that}} hath
made us, and not we ourselves;
{\ADD{we are}} his
people, and the
sheep of his
pasture.\FTNT{‡}{{\FR 100.3 }
and not…: or, and his we are}
\par }{\Q \VS{4}Enter into his
gates with
thanksgiving,
{\ADD{and}} into his
courts with
praise: be
thankful unto him,
{\ADD{and}}
bless his
name.
\par }{\Q \VS{5}For the
{\ND{
LORD}}
{\ADD{is}}
good; his
mercy
{\ADD{is}}
everlasting; and his
truth
{\ADD{endureth}} to
all
generations.\FTNT{§}{{\FR 100.5 }
to all…: Heb. to generation and generation}

\par }\PsalmChap{101}{\D A
Psalm of
David.
\par }{\Q \VerseOne{1}I will
sing of
mercy and
judgment: unto thee, O
{\ND{
LORD}}, will I
sing.
\par }{\Q \VS{2}I will behave myself
wisely in a
perfect
way. O when wilt thou
come unto me? I will
walk
within my
house with a
perfect
heart.
\par }{\Q \VS{3}I will
set no
wicked
thing before mine
eyes: I
hate the
work of them that turn
aside;
{\ADD{it}} shall not
cleave to me.\FTNT{*}{{\FR 101.3 }
wicked…: Heb. thing of Belial}
\par }{\Q \VS{4}A
froward
heart shall
depart from me: I will not
know a
wicked
{\ADD{person}}.
\par }{\Q \VS{5}Whoso
privily
slandereth his
neighbour, him will I cut
off: him that hath an
high
look and a
proud
heart will not I
suffer.
\par }{\Q \VS{6}Mine
eyes
{\ADD{shall be}} upon the
faithful of the
land, that they may
dwell with me: he that
walketh in a
perfect
way, he shall
serve me.\FTNT{†}{{\FR 101.6 }
in a…: or, perfect in the way}
\par }{\Q \VS{7}He that
worketh
deceit shall not
dwell
within my
house: he that
telleth
lies shall not
tarry in my
sight.\FTNT{‡}{{\FR 101.7 }
shall not tarry: Heb. shall not be established}
\par }{\Q \VS{8}I will
early
destroy all the
wicked of the
land; that I may cut
off all
wicked
doers from the
city of the
{\ND{
LORD}}.

\par }\PsalmChap{102}{\D A
Prayer of the
afflicted, when he is
overwhelmed, and poureth
out his
complaint
before the
{\ND{
LORD}}.
\par }{\Q \VerseOne{1}Hear my
prayer, O
{\ND{
LORD}}, and let my
cry
come unto thee.\FTNT{*}{{\FR 102.1 }
of: or, for}
\par }{\Q \VS{2}Hide not thy
face from me in the
day
{\ADD{when}} I am in
trouble;
incline thine
ear unto me: in the
day
{\ADD{when}} I
call
answer me
speedily.
\par }{\Q \VS{3}For my
days are
consumed like
smoke, and my
bones are
burned as an
hearth.\FTNT{†}{{\FR 102.3 }
like…: or, (as some read) into smoke}
\par }{\Q \VS{4}My
heart is
smitten, and
withered like
grass; so that I
forget to
eat my
bread.
\par }{\Q \VS{5}By reason of the
voice of my
groaning my
bones
cleave to my
skin.\FTNT{‡}{{\FR 102.5 }
skin: or, flesh}
\par }{\Q \VS{6}I am
like a
pelican of the
wilderness: I am like an
owl of the
desert.
\par }{\Q \VS{7}I
watch, and am as a
sparrow
alone upon the house
top.
\par }{\Q \VS{8}Mine
enemies
reproach me all the
day;
{\ADD{and}} they that are
mad against me are
sworn against me.
\par }{\Q \VS{9}For I have
eaten
ashes like
bread, and
mingled my
drink with
weeping,
\par }{\Q \VS{10}Because of thine
indignation and thy
wrath: for thou hast lifted me
up, and cast me
down.
\par }{\Q \VS{11}My
days
{\ADD{are}} like a
shadow that
declineth; and I am
withered like
grass.
\par }{\BB \par }{\Q \VS{12}But thou, O
{\ND{
LORD}}, shalt
endure for
ever; and thy
remembrance unto
all
generations.
\par }{\Q \VS{13}Thou shalt
arise,
{\ADD{and}} have
mercy upon
Zion: for the
time to
favour her, yea, the set
time, is
come.
\par }{\Q \VS{14}For thy
servants take
pleasure in her
stones, and
favour the
dust thereof.
\par }{\Q \VS{15}So the
heathen shall
fear the
name of the
{\ND{
LORD}}, and all the
kings of the
earth thy
glory.
\par }{\Q \VS{16}When the
{\ND{
LORD}} shall build
up
Zion, he shall
appear in his
glory.
\par }{\Q \VS{17}He will
regard the
prayer of the
destitute, and not
despise their
prayer.
\par }{\Q \VS{18}This shall be
written for the
generation to
come: and the
people which shall be
created shall
praise the
{\ND{
LORD}}.
\par }{\Q \VS{19}For he hath looked
down from the
height of his
sanctuary; from
heaven did the
{\ND{
LORD}}
behold the
earth;
\par }{\Q \VS{20}To
hear the
groaning of the
prisoner; to
loose those that are
appointed to
death;\FTNT{§}{{\FR 102.20 }
those…: Heb. the children of death}
\par }{\Q \VS{21}To
declare the
name of the
{\ND{
LORD}} in
Zion, and his
praise in
Jerusalem;
\par }{\Q \VS{22}When the
people are
gathered
together, and the
kingdoms, to
serve the
{\ND{
LORD}}.
\par }{\BB \par }{\Q \VS{23}He
weakened my
strength in the
way; he
shortened my
days.\FTNT{**}{{\FR 102.23 }
weakened: Heb. afflicted}
\par }{\Q \VS{24}I
said, O my
God, take me not
away in the
midst of my
days: thy
years
{\ADD{are}} throughout
all
generations.
\par }{\Q \VS{25}Of
old hast thou laid the
foundation of the
earth: and the
heavens
{\ADD{are}} the
work of thy
hands.
\par }{\Q \VS{26}They shall
perish, but thou shalt
endure: yea, all of them shall wax
old like a
garment; as a
vesture shalt thou
change them, and they shall be
changed:\FTNT{††}{{\FR 102.26 }
endure: Heb. stand}
\par }{\Q \VS{27}But thou
{\ADD{art}} the same, and thy
years shall have no
end.
\par }{\Q \VS{28}The
children of thy
servants shall
continue, and their
seed shall be
established
before thee.

\par }\PsalmChap{103}{\D {\ADD{A Psalm}} of
David.
\par }{\Q \VerseOne{1}Bless the
{\ND{
LORD}}, O my
soul: and all that is
within me,
{\ADD{bless}} his
holy
name.
\par }{\Q \VS{2}Bless the
{\ND{
LORD}}, O my
soul, and
forget not all his
benefits:
\par }{\Q \VS{3}Who
forgiveth all thine
iniquities; who
healeth all thy
diseases;
\par }{\Q \VS{4}Who
redeemeth thy
life from
destruction; who
crowneth thee with
lovingkindness and tender
mercies;
\par }{\Q \VS{5}Who
satisfieth thy
mouth with
good
{\ADD{things; so that}} thy
youth is
renewed like the
eagle’s.
\par }{\BB \par }{\Q \VS{6}The
{\ND{
LORD}}
executeth
righteousness and
judgment for all that are
oppressed.
\par }{\Q \VS{7}He made
known his
ways unto
Moses, his
acts unto the
children of
Israel.
\par }{\Q \VS{8}The
{\ND{
LORD}}
{\ADD{is}}
merciful and
gracious,
slow to
anger, and
plenteous in
mercy.\FTNT{*}{{\FR 103.8 }
plenteous…: Heb. great in mercy}
\par }{\Q \VS{9}He will not
always
chide: neither will he
keep
{\ADD{his anger}} for
ever.
\par }{\Q \VS{10}He hath not
dealt with us after our
sins; nor
rewarded us according to our
iniquities.
\par }{\Q \VS{11}For as the
heaven is high
above the
earth,
{\ADD{so}}
great is his
mercy toward them that
fear him.\FTNT{†}{{\FR 103.11 }
as…: Heb. according to the height of the heaven}
\par }{\Q \VS{12}As far
as the
east is from the
west,
{\ADD{so}} far hath he
removed our
transgressions from us.
\par }{\Q \VS{13}Like as a
father
pitieth
{\ADD{his}}
children,
{\ADD{so}} the
{\ND{
LORD}}
pitieth them that
fear him.
\par }{\Q \VS{14}For he
knoweth our
frame; he
remembereth that we
{\ADD{are}}
dust.
\par }{\Q \VS{15}{\ADD{As for}}
man, his
days
{\ADD{are}} as
grass: as a
flower of the
field, so he
flourisheth.
\par }{\Q \VS{16}For the
wind passeth
over it, and it is gone; and the
place thereof shall
know it no more.\FTNT{‡}{{\FR 103.16 }
it is…: Heb. it is not}
\par }{\Q \VS{17}But the
mercy of the
{\ND{
LORD}}
{\ADD{is}} from
everlasting
to
everlasting upon them that
fear him, and his
righteousness unto
children’s
children;
\par }{\Q \VS{18}To such as
keep his
covenant, and to those that
remember his
commandments to
do them.
\par }{\BB \par }{\Q \VS{19}The
{\ND{
LORD}} hath
prepared his
throne in the
heavens; and his
kingdom
ruleth over all.
\par }{\Q \VS{20}Bless the
{\ND{
LORD}}, ye his
angels, that
excel in
strength, that
do his
commandments,
hearkening unto the
voice of his
word.\FTNT{§}{{\FR 103.20 }
that excel…: Heb. mighty in strength}
\par }{\Q \VS{21}Bless ye the
{\ND{
LORD}}, all
{\ADD{ye}} his
hosts;
{\ADD{ye}}
ministers of his, that
do his
pleasure.
\par }{\Q \VS{22}Bless the
{\ND{
LORD}}, all his
works in all
places of his
dominion:
bless the
{\ND{
LORD}}, O my
soul.

\par }\PsalmChap{104}{\Q \VerseOne{1}Bless the
{\ND{
LORD}}, O my
soul. O
{\ND{
LORD}} my
God, thou art
very
great; thou art
clothed with
honour and
majesty.
\par }{\Q \VS{2}Who
coverest
{\ADD{thyself}} with
light as
{\ADD{with}} a
garment: who stretchest
out the
heavens like a
curtain:
\par }{\Q \VS{3}Who layeth the
beams of his
chambers in the
waters: who
maketh the
clouds his
chariot: who
walketh upon the
wings of the
wind:
\par }{\Q \VS{4}Who
maketh his
angels
spirits; his
ministers a
flaming
fire:
\par }{\Q \VS{5}{\ADD{Who}}
laid the
foundations of the
earth,
{\ADD{that}} it should not be
removed for
ever.\FTNT{*}{{\FR 104.5 }
Who…: Heb. He hath founded the earth upon her bases}
\par }{\Q \VS{6}Thou
coveredst it with the
deep as
{\ADD{with}} a
garment: the
waters
stood above the
mountains.
\par }{\Q \VS{7}At thy
rebuke they
fled; at the
voice of thy
thunder they
hasted away.
\par }{\Q \VS{8}They go
up by the
mountains; they go
down by the
valleys unto the
place
which thou hast
founded for them.\FTNT{†}{{\FR 104.8 }
They go up…: or, The mountains ascend, the valleys descend}
\par }{\Q \VS{9}Thou hast
set a
bound that they may not pass
over; that they turn not
again to
cover the
earth.
\par }{\BB \par }{\Q \VS{10}He
sendeth the
springs into the
valleys,
{\ADD{which}}
run among the
hills.\FTNT{‡}{{\FR 104.10 }
He: Heb. Who}\FTNT{§}{{\FR 104.10 }
run: Heb. walk}
\par }{\Q \VS{11}They give
drink to every
beast of the
field: the wild
asses
quench their
thirst.\FTNT{**}{{\FR 104.11 }
quench: Heb. break}
\par }{\Q \VS{12}By them shall the
fowls of the
heaven have their
habitation,
{\ADD{which}}
sing
among the
branches.\FTNT{††}{{\FR 104.12 }
sing: Heb. give a voice}
\par }{\Q \VS{13}He
watereth the
hills from his
chambers: the
earth is
satisfied with the
fruit of thy
works.
\par }{\Q \VS{14}He causeth the
grass to
grow for the
cattle, and
herb for the
service of
man: that he may bring
forth
food out of the
earth;
\par }{\Q \VS{15}And
wine
{\ADD{that}} maketh
glad the
heart of
man,
{\ADD{and}}
oil to make
{\ADD{his}}
face to
shine, and
bread
{\ADD{which}}
strengtheneth
man’s
heart.\FTNT{‡‡}{{\FR 104.15 }
oil…: Heb. to make his face shine with oil, or, more than oil}
\par }{\Q \VS{16}The
trees of the
{\ND{
LORD}} are
full
{\ADD{of sap;}} the
cedars of
Lebanon, which he hath
planted;
\par }{\Q \VS{17}Where the
birds make their
nests:
{\ADD{as for}} the
stork, the fir
trees
{\ADD{are}} her
house.
\par }{\Q \VS{18}The
high
hills
{\ADD{are}} a
refuge for the wild
goats;
{\ADD{and}} the
rocks for the
conies.
\par }{\BB \par }{\Q \VS{19}He
appointed the
moon for
seasons: the
sun
knoweth his going
down.
\par }{\Q \VS{20}Thou
makest
darkness, and it is
night: wherein all the
beasts of the
forest do
creep
{\ADD{forth}}.\FTNT{§§}{{\FR 104.20 }
beasts…: Heb. beasts thereof do trample on the forest}
\par }{\Q \VS{21}The young
lions
roar after their
prey, and
seek their
meat from
God.
\par }{\Q \VS{22}The
sun
ariseth, they gather themselves
together, and lay them
down in their
dens.
\par }{\Q \VS{23}Man goeth
forth unto his
work and to his
labour until the
evening.
\par }{\Q \VS{24}O
{\ND{
LORD}}, how
manifold are thy
works! in
wisdom hast thou
made them all: the
earth is
full of thy
riches.
\par }{\Q \VS{25}{\ADD{So is}} this
great and
wide
sea, wherein
{\ADD{are}} things
creeping
innumerable, both
small and
great
beasts.
\par }{\Q \VS{26}There
go the
ships:
{\ADD{there is}} that
leviathan,
{\ADD{whom}} thou hast
made to
play therein.\FTNT{***}{{\FR 104.26 }
made: Heb. formed}
\par }{\Q \VS{27}These
wait all upon thee; that thou mayest
give
{\ADD{them}} their
meat in due
season.
\par }{\Q \VS{28}{\ADD{That}} thou
givest them they
gather: thou
openest thine
hand, they are
filled with
good.
\par }{\Q \VS{29}Thou
hidest thy
face, they are
troubled: thou takest
away their
breath, they
die, and
return to their
dust.
\par }{\Q \VS{30}Thou
sendest forth thy
spirit, they are
created: and thou
renewest the
face of the
earth.
\par }{\BB \par }{\Q \VS{31}The
glory of the
{\ND{
LORD}} shall endure for
ever: the
{\ND{
LORD}} shall
rejoice in his
works.\FTNT{†††}{{\FR 104.31 }
endure: Heb. be}
\par }{\Q \VS{32}He
looketh on the
earth, and it
trembleth: he
toucheth the
hills, and they
smoke.
\par }{\Q \VS{33}I will
sing unto the
{\ND{
LORD}} as long as I
live: I will
sing praise to my
God while I have my
being.
\par }{\Q \VS{34}My
meditation of him shall be
sweet: I will be
glad in the
{\ND{
LORD}}.
\par }{\Q \VS{35}Let the
sinners be
consumed out of the
earth, and let the
wicked be no more.
Bless thou the
{\ND{
LORD}}, O my
soul.
Praise ye the
{\ND{
LORD}}.

\par }\PsalmChap{105}{\Q \VerseOne{1}O give
thanks unto the
{\ND{
LORD}};
call upon his
name: make
known his
deeds among the
people.
\par }{\Q \VS{2}Sing unto him, sing
psalms unto him:
talk ye of all his wondrous
works.
\par }{\Q \VS{3}Glory ye in his
holy
name: let the
heart of them
rejoice that
seek the
{\ND{
LORD}}.
\par }{\Q \VS{4}Seek the
{\ND{
LORD}}, and his
strength:
seek his
face
evermore.
\par }{\Q \VS{5}Remember his marvellous
works that he hath
done; his
wonders, and the
judgments of his
mouth;
\par }{\Q \VS{6}O ye
seed of
Abraham his
servant, ye
children of
Jacob his
chosen.
\par }{\Q \VS{7}He
{\ADD{is}} the
{\ND{
LORD}} our
God: his
judgments
{\ADD{are}} in all the
earth.
\par }{\BB \par }{\Q \VS{8}He hath
remembered his
covenant for
ever, the
word
{\ADD{which}} he
commanded to a
thousand
generations.
\par }{\Q \VS{9}Which
{\ADD{covenant}} he
made with
Abraham, and his
oath unto
Isaac;
\par }{\Q \VS{10}And
confirmed the same unto
Jacob for a
law,
{\ADD{and}} to
Israel
{\ADD{for}} an
everlasting
covenant:
\par }{\Q \VS{11}Saying, Unto thee will I
give the
land of
Canaan, the
lot of your
inheritance:\FTNT{*}{{\FR 105.11 }
lot: Heb. cord}
\par }{\Q \VS{12}When they were
{\ADD{but}} a few
men in
number; yea, very
few, and
strangers in it.
\par }{\Q \VS{13}When they
went from one
nation to another, from
{\ADD{one}}
kingdom to
another
people;
\par }{\Q \VS{14}He
suffered no
man to do them
wrong: yea, he
reproved
kings for their sakes;
\par }{\Q \VS{15}{\ADD{Saying}},
Touch not mine
anointed, and do my
prophets no
harm.
\par }{\Q \VS{16}Moreover he
called for a
famine upon the
land: he
brake the whole
staff of
bread.
\par }{\Q \VS{17}He
sent a
man
before them,
{\ADD{even}}
Joseph,
{\ADD{who}} was
sold for a
servant:
\par }{\Q \VS{18}Whose
feet they
hurt with
fetters:
he was
laid in
iron:\FTNT{†}{{\FR 105.18 }
he…: Heb. his soul came into iron}
\par }{\Q \VS{19}Until the
time that his
word
came: the
word of the
{\ND{
LORD}}
tried him.
\par }{\Q \VS{20}The
king
sent and
loosed him;
{\ADD{even}} the
ruler of the
people, and let him go
free.
\par }{\Q \VS{21}He
made him
lord of his
house, and
ruler of all his
substance:\FTNT{‡}{{\FR 105.21 }
substance: Heb. possession}
\par }{\Q \VS{22}To
bind his
princes at his
pleasure; and teach his
senators
wisdom.
\par }{\Q \VS{23}Israel also
came into
Egypt; and
Jacob
sojourned in the
land of
Ham.
\par }{\Q \VS{24}And he
increased his
people
greatly; and made them
stronger than their
enemies.
\par }{\BB \par }{\Q \VS{25}He
turned their
heart to
hate his
people, to deal
subtilly with his
servants.
\par }{\Q \VS{26}He
sent
Moses his
servant;
{\ADD{and}}
Aaron whom he had
chosen.
\par }{\Q \VS{27}They
shewed his
signs among them, and
wonders in the
land of
Ham.\FTNT{§}{{\FR 105.27 }
his…: Heb. words of his signs}
\par }{\Q \VS{28}He
sent
darkness, and made it
dark; and they
rebelled not against his
word.
\par }{\Q \VS{29}He
turned their
waters into
blood, and
slew their
fish.
\par }{\Q \VS{30}Their
land brought
forth
frogs in
abundance, in the
chambers of their
kings.
\par }{\Q \VS{31}He
spake, and there
came divers sorts of
flies,
{\ADD{and}}
lice in all their
coasts.
\par }{\Q \VS{32}He
gave them
hail for
rain,
{\ADD{and}}
flaming
fire in their
land.\FTNT{**}{{\FR 105.32 }
them…: Heb. their rain hail}
\par }{\Q \VS{33}He
smote their
vines also and their fig
trees; and
brake the
trees of their
coasts.
\par }{\Q \VS{34}He
spake, and the
locusts
came, and
caterpillers, and that without
number,
\par }{\Q \VS{35}And did eat
up all the
herbs in their
land, and
devoured the
fruit of their
ground.
\par }{\Q \VS{36}He
smote also all the
firstborn in their
land, the
chief of all their
strength.
\par }{\Q \VS{37}He brought them
forth also with
silver and
gold: and
{\ADD{there was}} not one
feeble
{\ADD{person}} among their
tribes.
\par }{\Q \VS{38}Egypt was
glad when they
departed: for the
fear of them
fell upon them.
\par }{\Q \VS{39}He
spread a
cloud for a
covering; and
fire to give
light in the
night.
\par }{\Q \VS{40}{\ADD{The people}}
asked, and he
brought
quails, and
satisfied them with the
bread of
heaven.
\par }{\Q \VS{41}He
opened the
rock, and the
waters gushed
out; they
ran in the dry
places
{\ADD{like}} a
river.
\par }{\Q \VS{42}For he
remembered his
holy
promise,
{\ADD{and}}
Abraham his
servant.
\par }{\Q \VS{43}And he brought
forth his
people with
joy,
{\ADD{and}} his
chosen with
gladness:\FTNT{††}{{\FR 105.43 }
gladness: Heb. singing}
\par }{\Q \VS{44}And
gave them the
lands of the
heathen: and they
inherited the
labour of the
people;
\par }{\Q \VS{45}That they might
observe his
statutes, and
keep his
laws.
Praise ye the
{\ND{
LORD}}.\FTNT{‡‡}{{\FR 105.45 }
Praise…: Heb. Hallelujah}

\par }\PsalmChap{106}{\Q \VerseOne{1}Praise ye the
{\ND{
LORD}}. O give
thanks unto the
{\ND{
LORD}}; for
{\ADD{he is}}
good: for his
mercy
{\ADD{endureth}} for
ever.\FTNT{*}{{\FR 106.1 }
Praise…: Heb. Hallelujah}
\par }{\Q \VS{2}Who can
utter the mighty
acts of the
{\ND{
LORD}}?
{\ADD{who}} can shew
forth all his
praise?
\par }{\Q \VS{3}Blessed
{\ADD{are}} they that
keep
judgment,
{\ADD{and}} he that
doeth
righteousness at all
times.
\par }{\Q \VS{4}Remember me, O
{\ND{
LORD}}, with the
favour
{\ADD{that thou bearest unto}} thy
people: O
visit me with thy
salvation;
\par }{\Q \VS{5}That I may
see the
good of thy
chosen, that I may
rejoice in the
gladness of thy
nation, that I may
glory with thine
inheritance.
\par }{\BB \par }{\Q \VS{6}We have
sinned with our
fathers, we have committed
iniquity, we have done
wickedly.
\par }{\Q \VS{7}Our
fathers
understood not thy
wonders in
Egypt; they
remembered not the
multitude of thy
mercies; but
provoked
{\ADD{him}} at the
sea,
{\ADD{even}} at the
Red
sea.
\par }{\Q \VS{8}Nevertheless he
saved them for his
name’s sake, that he might make his mighty
power to be
known.
\par }{\Q \VS{9}He
rebuked the
Red
sea also, and it was dried
up: so he
led them through the
depths, as through the
wilderness.
\par }{\Q \VS{10}And he
saved them from the
hand of him that
hated
{\ADD{them}}, and
redeemed them from the
hand of the
enemy.
\par }{\Q \VS{11}And the
waters
covered their
enemies: there was not
one of them
left.
\par }{\Q \VS{12}Then
believed they his
words; they
sang his
praise.
\par }{\BB \par }{\Q \VS{13}They
soon
forgat his
works; they
waited not for his
counsel:\FTNT{†}{{\FR 106.13 }
They soon…: Heb. They made haste, they forgat}
\par }{\Q \VS{14}But
lusted
exceedingly in the
wilderness, and
tempted
God in the
desert.\FTNT{‡}{{\FR 106.14 }
lusted…: Heb. lusted a lust}
\par }{\Q \VS{15}And he
gave them their
request; but
sent
leanness into their
soul.
\par }{\Q \VS{16}They
envied
Moses also in the
camp,
{\ADD{and}}
Aaron the
saint of the
{\ND{
LORD}}.
\par }{\Q \VS{17}The
earth
opened and swallowed
up
Dathan, and
covered the
company of
Abiram.
\par }{\Q \VS{18}And a
fire was
kindled in their
company; the
flame burned
up the
wicked.
\par }{\Q \VS{19}They
made a
calf in
Horeb, and
worshipped the molten
image.
\par }{\Q \VS{20}Thus they
changed their
glory into the
similitude of an
ox that
eateth
grass.
\par }{\Q \VS{21}They
forgat
God their
saviour, which had
done great
things in
Egypt;
\par }{\Q \VS{22}Wondrous
works in the
land of
Ham,
{\ADD{and}} terrible
things by the
Red
sea.
\par }{\Q \VS{23}Therefore he
said that he would
destroy them, had
not
Moses his
chosen
stood
before him in the
breach, to turn
away his
wrath, lest he should
destroy
{\ADD{them}}.
\par }{\Q \VS{24}Yea, they
despised the
pleasant
land, they
believed not his
word:\FTNT{§}{{\FR 106.24 }
the pleasant…: Heb. a land of desire}
\par }{\Q \VS{25}But
murmured in their
tents,
{\ADD{and}}
hearkened not unto the
voice of the
{\ND{
LORD}}.
\par }{\Q \VS{26}Therefore he lifted
up his
hand against them, to
overthrow them in the
wilderness:\FTNT{**}{{\FR 106.26 }
to overthrow: Heb. to make them fall}
\par }{\Q \VS{27}To
overthrow their
seed also among the
nations, and to
scatter them in the
lands.\FTNT{††}{{\FR 106.27 }
To overthrow: Heb. To make them fall}
\par }{\Q \VS{28}They
joined themselves also unto
Baal-peor, and
ate the
sacrifices of the
dead.
\par }{\Q \VS{29}Thus they provoked
{\ADD{him}} to
anger with their
inventions: and the
plague brake
in upon them.
\par }{\Q \VS{30}Then stood
up
Phinehas, and executed
judgment: and
{\ADD{so}} the
plague was
stayed.
\par }{\Q \VS{31}And that was
counted unto him for
righteousness unto
all
generations
for
evermore.
\par }{\Q \VS{32}They
angered
{\ADD{him}} also at the
waters of
strife, so that it went
ill with
Moses for their sakes:
\par }{\Q \VS{33}Because they
provoked his
spirit, so that he spake
unadvisedly with his
lips.
\par }{\BB \par }{\Q \VS{34}They did not
destroy the
nations, concerning whom the
{\ND{
LORD}}
commanded them:
\par }{\Q \VS{35}But were
mingled among the
heathen, and
learned their
works.
\par }{\Q \VS{36}And they
served their
idols: which were a
snare unto them.
\par }{\Q \VS{37}Yea, they
sacrificed their
sons and their
daughters unto
devils,
\par }{\Q \VS{38}And
shed
innocent
blood,
{\ADD{even}} the
blood of their
sons and of their
daughters, whom they
sacrificed unto the
idols of
Canaan: and the
land was
polluted with
blood.
\par }{\Q \VS{39}Thus were they
defiled with their own
works, and went a
whoring with their own
inventions.
\par }{\Q \VS{40}Therefore was the
wrath of the
{\ND{
LORD}}
kindled against his
people, insomuch that he
abhorred his own
inheritance.
\par }{\Q \VS{41}And he
gave them into the
hand of the
heathen; and they that
hated them
ruled over them.
\par }{\Q \VS{42}Their
enemies also
oppressed them, and they were brought into
subjection under their
hand.
\par }{\Q \VS{43}Many
times did he
deliver them; but they
provoked
{\ADD{him}} with their
counsel, and were brought
low for their
iniquity.\FTNT{‡‡}{{\FR 106.43 }
brought…: or, impoverished, or, weakened}
\par }{\Q \VS{44}Nevertheless he
regarded their
affliction, when he
heard their
cry:
\par }{\Q \VS{45}And he
remembered for them his
covenant, and
repented according to the
multitude of his
mercies.
\par }{\Q \VS{46}He
made them also to be
pitied
of all those that carried them
captives.
\par }{\Q \VS{47}Save us, O
{\ND{
LORD}} our
God, and
gather us from among the
heathen, to give
thanks unto thy
holy
name,
{\ADD{and}} to
triumph in thy
praise.
\par }{\Q \VS{48}Blessed
{\ADD{be}} the
{\ND{
LORD}}
God of
Israel from
everlasting to
everlasting: and let all the
people
say,
Amen.
Praise ye the
{\ND{
LORD}}.\FTNT{§§}{{\FR 106.48 }
Praise…: Heb. Hallelujah}

\par }\PsalmChap{107}{\Q \VerseOne{1}O give
thanks unto the
{\ND{
LORD}}, for
{\ADD{he is}}
good: for his
mercy
{\ADD{endureth}} for
ever.
\par }{\Q \VS{2}Let the
redeemed of the
{\ND{
LORD}}
say
{\ADD{so}}, whom he hath
redeemed from the
hand of the
enemy;
\par }{\Q \VS{3}And
gathered them out of the
lands, from the
east, and from the
west, from the
north, and from the
south.\FTNT{*}{{\FR 107.3 }
south: Heb. sea}
\par }{\Q \VS{4}They
wandered in the
wilderness in a
solitary
way; they
found no
city to
dwell in.
\par }{\Q \VS{5}Hungry and
thirsty, their
soul
fainted in them.
\par }{\Q \VS{6}Then they
cried unto the
{\ND{
LORD}} in their
trouble,
{\ADD{and}} he
delivered them out of their
distresses.
\par }{\Q \VS{7}And he led them
forth by the
right
way, that they might
go to a
city of
habitation.
\par }{\Q \VS{8}Oh that
{\ADD{men}} would
praise the
{\ND{
LORD}}
{\ADD{for}} his
goodness, and
{\ADD{for}} his wonderful
works to the
children of
men!
\par }{\Q \VS{9}For he
satisfieth the
longing
soul, and
filleth the
hungry
soul with
goodness.
\par }{\BB \par }{\Q \VS{10}Such as
sit in
darkness and in the shadow of
death,
{\ADD{being}}
bound in
affliction and
iron;
\par }{\Q \VS{11}Because they
rebelled against the
words of
God, and
contemned the
counsel of the most
High:
\par }{\Q \VS{12}Therefore he brought
down their
heart with
labour; they fell
down, and
{\ADD{there was}} none to
help.
\par }{\Q \VS{13}Then they
cried unto the
{\ND{
LORD}} in their
trouble,
{\ADD{and}} he
saved them out of their
distresses.
\par }{\Q \VS{14}He brought them
out of
darkness and the shadow of
death, and
brake their
bands in
sunder.
\par }{\Q \VS{15}Oh that
{\ADD{men}} would
praise the
{\ND{
LORD}}
{\ADD{for}} his
goodness, and
{\ADD{for}} his wonderful
works to the
children of
men!
\par }{\Q \VS{16}For he hath
broken the
gates of
brass, and
cut the
bars of
iron in
sunder.
\par }{\BB \par }{\Q \VS{17}Fools
because of their
transgression, and because of their
iniquities, are
afflicted.
\par }{\Q \VS{18}Their
soul
abhorreth all manner of
meat; and they draw
near unto the
gates of
death.
\par }{\Q \VS{19}Then they
cry unto the
{\ND{
LORD}} in their
trouble,
{\ADD{and}} he
saveth them out of their
distresses.
\par }{\Q \VS{20}He
sent his
word, and
healed them, and
delivered
{\ADD{them}} from their
destructions.
\par }{\Q \VS{21}Oh that
{\ADD{men}} would
praise the
{\ND{
LORD}}
{\ADD{for}} his
goodness, and
{\ADD{for}} his wonderful
works to the
children of
men!
\par }{\Q \VS{22}And let them
sacrifice the
sacrifices of
thanksgiving, and
declare his
works with
rejoicing.\FTNT{†}{{\FR 107.22 }
rejoicing: Heb. singing}
\par }{\BB \par }{\Q \VS{23}They that go
down to the
sea in
ships, that
do
business in
great
waters;
\par }{\Q \VS{24}These
see the
works of the
{\ND{
LORD}}, and his
wonders in the
deep.
\par }{\Q \VS{25}For he
commandeth, and
raiseth the
stormy
wind, which lifteth
up the
waves thereof.\FTNT{‡}{{\FR 107.25 }
raiseth: Heb. maketh to stand}
\par }{\Q \VS{26}They mount
up to the
heaven, they go
down again to the
depths: their
soul is
melted because of
trouble.
\par }{\Q \VS{27}They reel to and
fro, and
stagger like a drunken
man, and are at their
wits’
end.
\FTNT{§}{{\FR 107.27 }
are…: Heb. all their wisdom is swallowed up}
\par }{\Q \VS{28}Then they
cry unto the
{\ND{
LORD}} in their
trouble, and he bringeth them
out of their
distresses.
\par }{\Q \VS{29}He
maketh the
storm a
calm, so that the
waves thereof are
still.
\par }{\Q \VS{30}Then are they
glad because they be
quiet; so he
bringeth them unto their
desired
haven.
\par }{\Q \VS{31}Oh that
{\ADD{men}} would
praise the
{\ND{
LORD}}
{\ADD{for}} his
goodness, and
{\ADD{for}} his wonderful
works to the
children of
men!
\par }{\Q \VS{32}Let them
exalt him also in the
congregation of the
people, and
praise him in the
assembly of the
elders.
\par }{\BB \par }{\Q \VS{33}He
turneth
rivers into a
wilderness, and the
watersprings into dry
ground;
\par }{\Q \VS{34}A
fruitful
land into
barrenness, for the
wickedness of them that
dwell therein.\FTNT{**}{{\FR 107.34 }
barrenness: Heb. saltiness}
\par }{\Q \VS{35}He
turneth the
wilderness into a
standing
water, and
dry
ground into
watersprings.
\par }{\Q \VS{36}And there he maketh the
hungry to
dwell, that they may
prepare a
city for
habitation;
\par }{\Q \VS{37}And
sow the
fields, and
plant
vineyards, which may
yield
fruits of
increase.
\par }{\Q \VS{38}He
blesseth them also, so that they are
multiplied
greatly; and
suffereth not their
cattle to
decrease.
\par }{\Q \VS{39}Again, they are
minished and brought
low through
oppression,
affliction, and
sorrow.
\par }{\Q \VS{40}He
poureth
contempt upon
princes, and causeth them to
wander in the
wilderness,
{\ADD{where there is}} no
way.\FTNT{††}{{\FR 107.40 }
wilderness: or, void place}
\par }{\Q \VS{41}Yet setteth he the
poor on
high from
affliction, and
maketh
{\ADD{him}}
families like a
flock.\FTNT{‡‡}{{\FR 107.41 }
from: or, after}
\par }{\Q \VS{42}The
righteous shall
see
{\ADD{it}}, and
rejoice: and all
iniquity shall
stop her
mouth.
\par }{\Q \VS{43}Whoso
{\ADD{is}}
wise, and will
observe these
{\ADD{things}}, even they shall
understand the
lovingkindness of the
{\ND{
LORD}}.

\par }\PsalmChap{108}{\D A
Song
{\ADD{or}}
Psalm of
David.
\par }{\Q \VerseOne{1}O
God, my
heart is
fixed; I will
sing and give
praise, even with my
glory.
\par }{\Q \VS{2}Awake,
psaltery and
harp: I
{\ADD{myself}} will
awake
early.
\par }{\Q \VS{3}I will
praise thee, O
{\ND{
LORD}}, among the
people: and I will sing
praises unto thee among the
nations.
\par }{\Q \VS{4}For thy
mercy
{\ADD{is}}
great above the
heavens: and thy
truth
{\ADD{reacheth}} unto the
clouds.\FTNT{*}{{\FR 108.4 }
clouds: or, skies}
\par }{\Q \VS{5}Be thou
exalted, O
God, above the
heavens: and thy
glory above all the
earth;
\par }{\BB \par }{\Q \VS{6}That thy
beloved may be
delivered:
save
{\ADD{with}} thy right
hand, and
answer me.
\par }{\Q \VS{7}God hath
spoken in his
holiness; I will
rejoice, I will
divide
Shechem, and mete
out the
valley of
Succoth.
\par }{\Q \VS{8}Gilead
{\ADD{is}} mine;
Manasseh
{\ADD{is}} mine;
Ephraim also
{\ADD{is}} the
strength of mine
head;
Judah
{\ADD{is}} my
lawgiver;
\par }{\Q \VS{9}Moab
{\ADD{is}} my
washpot; over
Edom will I cast
out my
shoe; over
Philistia will I
triumph.
\par }{\Q \VS{10}Who will
bring me into the
strong
city? who will
lead me into
Edom?
\par }{\Q \VS{11}{\ADD{Wilt}} not
{\ADD{thou}}, O
God,
{\ADD{who}} hast cast us
off? and wilt not thou, O
God, go
forth with our
hosts?
\par }{\Q \VS{12}Give us
help from
trouble: for
vain
{\ADD{is}} the
help of
man.
\par }{\Q \VS{13}Through
God we shall
do
valiantly: for he
{\ADD{it is that}} shall tread
down our
enemies.

\par }\PsalmChap{109}{\D To the chief
Musician, A
Psalm of
David.
\par }{\Q \VerseOne{1}Hold not thy
peace, O
God of my
praise;
\par }{\Q \VS{2}For the
mouth of the
wicked and the
mouth of the
deceitful are
opened against me: they have
spoken against me with a
lying
tongue.\FTNT{*}{{\FR 109.2 }
of the deceitful: Heb. of deceit}\FTNT{†}{{\FR 109.2 }
are opened: Heb. have opened themselves}
\par }{\Q \VS{3}They
compassed me about also with
words of
hatred; and
fought against me without a
cause.
\par }{\Q \VS{4}For my
love they are my
adversaries: but I
{\ADD{give myself unto}}
prayer.
\par }{\Q \VS{5}And they have
rewarded me
evil for
good, and
hatred for my
love.
\par }{\BB \par }{\Q \VS{6}Set thou a wicked
man over him: and let
Satan
stand at his right
hand.\FTNT{‡}{{\FR 109.6 }
Satan: or, an adversary}
\par }{\Q \VS{7}When he shall be
judged, let him
be
condemned: and let his
prayer become
sin.\FTNT{§}{{\FR 109.7 }
be condemned: Heb. go out guilty, or, wicked}
\par }{\Q \VS{8}Let his
days be
few;
{\ADD{and}} let
another
take his
office.\FTNT{**}{{\FR 109.8 }
office: or, charge}
\par }{\Q \VS{9}Let his
children be
fatherless, and his
wife a
widow.
\par }{\Q \VS{10}Let his
children be
continually
vagabonds, and
beg: let them
seek
{\ADD{their bread}} also out of their desolate
places.
\par }{\Q \VS{11}Let the
extortioner
catch all that he hath; and let the
strangers
spoil his
labour.
\par }{\Q \VS{12}Let there be none to
extend
mercy unto him: neither let there be any to
favour his fatherless
children.
\par }{\Q \VS{13}Let his
posterity be cut
off;
{\ADD{and}} in the
generation
following let their
name be blotted
out.
\par }{\Q \VS{14}Let the
iniquity of his
fathers be
remembered with the
{\ND{
LORD}}; and let not the
sin of his
mother be blotted
out.
\par }{\Q \VS{15}Let them be before the
{\ND{
LORD}}
continually, that he may cut
off the
memory of them from the
earth.
\par }{\Q \VS{16}Because that he
remembered not to
shew
mercy, but
persecuted the
poor and
needy
man, that he might even
slay the
broken in
heart.
\par }{\Q \VS{17}As he
loved
cursing, so let it
come unto him: as he
delighted not in
blessing, so let it be
far from him.
\par }{\Q \VS{18}As he
clothed himself with
cursing like as with his
garment, so let it
come into his
bowels like
water, and like
oil into his
bones.\FTNT{††}{{\FR 109.18 }
into his bowels: Heb. within him}
\par }{\Q \VS{19}Let it be unto him as the
garment
{\ADD{which}}
covereth him, and for a
girdle wherewith he is
girded
continually.
\par }{\Q \VS{20}{\ADD{Let}} this
{\ADD{be}} the
reward of mine
adversaries from the
{\ND{
LORD}}, and of them that
speak
evil against my
soul.
\par }{\BB \par }{\Q \VS{21}But
do thou for me, O
{\ND{
GOD}} the
Lord, for thy
name’s sake: because thy
mercy
{\ADD{is}}
good,
deliver thou me.
\par }{\Q \VS{22}For I
{\ADD{am}}
poor and
needy, and my
heart is
wounded
within me.
\par }{\Q \VS{23}I am
gone like the
shadow when it
declineth: I am tossed up and
down as the
locust.
\par }{\Q \VS{24}My
knees are
weak through
fasting; and my
flesh
faileth of
fatness.
\par }{\Q \VS{25}I became also a
reproach unto them:
{\ADD{when}} they
looked upon me they
shaked their
heads.
\par }{\Q \VS{26}Help me, O
{\ND{
LORD}} my
God: O
save me according to thy
mercy:
\par }{\Q \VS{27}That they may
know that this
{\ADD{is}} thy
hand;
{\ADD{that}} thou,
{\ND{
LORD}}, hast
done it.
\par }{\Q \VS{28}Let them
curse, but
bless thou: when they
arise, let them be
ashamed; but let thy
servant
rejoice.
\par }{\Q \VS{29}Let mine
adversaries be
clothed with
shame, and let them
cover themselves with their own
confusion, as with a
mantle.
\par }{\Q \VS{30}I will
greatly
praise the
{\ND{
LORD}} with my
mouth; yea, I will
praise him
among the
multitude.
\par }{\Q \VS{31}For he shall
stand at the right
hand of the
poor, to
save
{\ADD{him}} from those that
condemn his
soul.\FTNT{‡‡}{{\FR 109.31 }
those…: Heb. the judges of}

\par }\PsalmChap{110}{\D A
Psalm of
David.
\par }{\Q \VerseOne{1}The
{\ND{
LORD}}
said unto my
Lord,
Sit thou at my right
hand, until I
make thine
enemies thy
footstool.
\par }{\Q \VS{2}The
{\ND{
LORD}} shall
send the
rod of thy
strength out of
Zion:
rule thou in the
midst of thine
enemies.
\par }{\Q \VS{3}Thy
people
{\ADD{shall be}}
willing in the
day of thy
power, in the
beauties of
holiness from the
womb of the
morning: thou hast the
dew of thy
youth.\FTNT{*}{{\FR 110.3 }
from…: or, more than the womb of the morning: thou shalt have, etc}
\par }{\Q \VS{4}The
{\ND{
LORD}} hath
sworn, and will not
repent, Thou
{\ADD{art}} a
priest for
ever after the
order of
Melchizedek.
\par }{\BB \par }{\Q \VS{5}The
Lord at thy right
hand shall
strike through
kings in the
day of his
wrath.
\par }{\Q \VS{6}He shall
judge among the
heathen, he shall
fill
{\ADD{the places}} with the dead
bodies; he shall
wound the
heads over
many
countries.\FTNT{†}{{\FR 110.6 }
many: or, great}
\par }{\Q \VS{7}He shall
drink of the
brook in the
way: therefore shall he lift
up the
head.

\par }\PsalmChap{111}{\Q \VerseOne{1}Praise ye the
{\ND{
LORD}}. I will
praise the
{\ND{
LORD}} with
{\ADD{my}} whole
heart, in the
assembly of the
upright, and
{\ADD{in}} the
congregation.\FTNT{*}{{\FR 111.1 }
Praise ye…: Heb. Hallelujah}
\par }{\Q \VS{2}The
works of the
{\ND{
LORD}}
{\ADD{are}}
great, sought
out of all them that have
pleasure therein.
\par }{\Q \VS{3}His
work
{\ADD{is}}
honourable and
glorious: and his
righteousness
endureth for
ever.
\par }{\Q \VS{4}He hath
made his wonderful
works to be
remembered: the
{\ND{
LORD}}
{\ADD{is}}
gracious and full of
compassion.
\par }{\Q \VS{5}He hath
given
meat unto them that
fear him: he will
ever be
mindful of his
covenant.\FTNT{†}{{\FR 111.5 }
meat: Heb. prey}
\par }{\BB \par }{\Q \VS{6}He hath
shewed his
people the
power of his
works, that he may
give them the
heritage of the
heathen.
\par }{\Q \VS{7}The
works of his
hands
{\ADD{are}}
verity and
judgment; all his
commandments
{\ADD{are}}
sure.
\par }{\Q \VS{8}They stand
fast for
ever and
ever,
{\ADD{and are}}
done in
truth and
uprightness.\FTNT{‡}{{\FR 111.8 }
stand…: Heb. are established}
\par }{\Q \VS{9}He
sent
redemption unto his
people: he hath
commanded his
covenant for
ever:
holy and
reverend
{\ADD{is}} his
name.
\par }{\Q \VS{10}The
fear of the
{\ND{
LORD}}
{\ADD{is}} the
beginning of
wisdom: a
good
understanding have all they that
do
{\ADD{his commandments:}} his
praise
endureth for
ever.\FTNT{§}{{\FR 111.10 }
a good…: or, good success}\FTNT{**}{{\FR 111.10 }
his commandments: Heb. them}

\par }\PsalmChap{112}{\Q \VerseOne{1}Praise ye the
{\ND{
LORD}}.
Blessed
{\ADD{is}} the
man
{\ADD{that}}
feareth the
{\ND{
LORD}},
{\ADD{that}}
delighteth
greatly in his
commandments.\FTNT{*}{{\FR 112.1 }
Praise…: Heb. Hallelujah}
\par }{\Q \VS{2}His
seed shall be
mighty upon
earth: the
generation of the
upright shall be
blessed.
\par }{\Q \VS{3}Wealth and
riches
{\ADD{shall be}} in his
house: and his
righteousness
endureth for
ever.
\par }{\Q \VS{4}Unto the
upright there
ariseth
light in the
darkness:
{\ADD{he is}}
gracious, and full of
compassion, and
righteous.
\par }{\Q \VS{5}A
good
man sheweth
favour, and
lendeth: he will
guide his
affairs with
discretion.\FTNT{†}{{\FR 112.5 }
discretion: Heb. judgment}
\par }{\BB \par }{\Q \VS{6}Surely he shall not be
moved for
ever: the
righteous shall be in
everlasting
remembrance.
\par }{\Q \VS{7}He shall not be
afraid of
evil
tidings: his
heart is
fixed,
trusting in the
{\ND{
LORD}}.
\par }{\Q \VS{8}His
heart
{\ADD{is}}
established, he shall not be
afraid, until he
see
{\ADD{his desire}} upon his
enemies.
\par }{\Q \VS{9}He hath
dispersed, he hath
given to the
poor; his
righteousness
endureth for
ever; his
horn shall be
exalted with
honour.
\par }{\Q \VS{10}The
wicked shall
see
{\ADD{it}}, and be
grieved; he shall
gnash with his
teeth, and melt
away: the
desire of the
wicked shall
perish.

\par }\PsalmChap{113}{\Q \VerseOne{1}Praise ye the
{\ND{
LORD}}.
Praise, O ye
servants of the
{\ND{
LORD}},
praise the
name of the
{\ND{
LORD}}.\FTNT{*}{{\FR 113.1 }
Praise ye…: Heb. Hallelujah}
\par }{\Q \VS{2}Blessed be the
name of the
{\ND{
LORD}} from this time
forth and
for
evermore.
\par }{\Q \VS{3}From the
rising of the
sun unto the going
down of the same the
{\ND{
LORD’s}}
name
{\ADD{is}} to be
praised.
\par }{\Q \VS{4}The
{\ND{
LORD}}
{\ADD{is}}
high above all
nations,
{\ADD{and}} his
glory above the
heavens.
\par }{\Q \VS{5}Who
{\ADD{is}} like unto the
{\ND{
LORD}} our
God, who
dwelleth on
high,\FTNT{†}{{\FR 113.5 }
dwelleth…: Heb. exalteth himself to dwell}
\par }{\Q \VS{6}Who
humbleth
{\ADD{himself}} to
behold
{\ADD{the things that are}} in
heaven, and in the
earth!
\par }{\Q \VS{7}He raiseth
up the
poor out of the
dust,
{\ADD{and}}
lifteth the
needy out of the
dunghill;
\par }{\Q \VS{8}That he may
set
{\ADD{him}} with
princes,
{\ADD{even}} with the
princes of his
people.
\par }{\Q \VS{9}He maketh the barren
woman to
keep
house,
{\ADD{and to be}} a
joyful
mother of
children.
Praise ye the
{\ND{
LORD}}.\FTNT{‡}{{\FR 113.9 }
to keep…: Heb. to dwell in an house}

\par }\PsalmChap{114}{\Q \VerseOne{1}When
Israel went
out of
Egypt, the
house of
Jacob from a
people of strange
language;
\par }{\Q \VS{2}Judah was his
sanctuary,
{\ADD{and}}
Israel his
dominion.
\par }{\Q \VS{3}The
sea
saw
{\ADD{it}}, and
fled:
Jordan was
driven
back.
\par }{\Q \VS{4}The
mountains
skipped like
rams,
{\ADD{and}} the little
hills like
lambs.
\par }{\Q \VS{5}What
{\ADD{ailed}} thee, O thou
sea, that thou
fleddest? thou
Jordan,
{\ADD{that}} thou wast
driven
back?
\par }{\Q \VS{6}Ye
mountains,
{\ADD{that}} ye
skipped like
rams;
{\ADD{and}} ye little
hills, like
lambs?
\par }{\Q \VS{7}Tremble, thou
earth, at the
presence of the
Lord, at the
presence of the
God of
Jacob;
\par }{\Q \VS{8}Which
turned the
rock
{\ADD{into}} a
standing
water, the
flint into a
fountain of
waters.

\par }\PsalmChap{115}{\Q \VerseOne{1}Not unto us, O
{\ND{
LORD}}, not unto us, but unto thy
name
give
glory, for thy
mercy,
{\ADD{and}} for thy
truth’s sake.
\par }{\Q \VS{2}Wherefore should the
heathen
say, Where
{\ADD{is}} now their
God?
\par }{\Q \VS{3}But our
God
{\ADD{is}} in the
heavens: he hath
done whatsoever he hath
pleased.
\par }{\Q \VS{4}Their
idols
{\ADD{are}}
silver and
gold, the
work of
men’s
hands.
\par }{\Q \VS{5}They have
mouths, but they
speak not:
eyes have they, but they
see not:
\par }{\Q \VS{6}They have
ears, but they
hear not:
noses have they, but they
smell not:
\par }{\Q \VS{7}They have
hands, but they
handle not:
feet have they, but they
walk not: neither
speak they through their
throat.
\par }{\Q \VS{8}They that
make them are like unto them;
{\ADD{so is}} every one that
trusteth in them.
\par }{\BB \par }{\Q \VS{9}O
Israel,
trust thou in the
{\ND{
LORD}}: he
{\ADD{is}} their
help and their
shield.
\par }{\Q \VS{10}O
house of
Aaron,
trust in the
{\ND{
LORD}}: he
{\ADD{is}} their
help and their
shield.
\par }{\Q \VS{11}Ye that
fear the
{\ND{
LORD}},
trust in the
{\ND{
LORD}}: he
{\ADD{is}} their
help and their
shield.
\par }{\Q \VS{12}The
{\ND{
LORD}} hath been
mindful of us: he will
bless
{\ADD{us;}} he will
bless the
house of
Israel; he will
bless the
house of
Aaron.
\par }{\Q \VS{13}He will
bless them that
fear the
{\ND{
LORD}},
{\ADD{both}}
small and
great.\FTNT{*}{{\FR 115.13 }
and: Heb. with}
\par }{\Q \VS{14}The
{\ND{
LORD}} shall
increase you more and more, you and your
children.
\par }{\Q \VS{15}Ye
{\ADD{are}}
blessed of the
{\ND{
LORD}} which
made
heaven and
earth.
\par }{\Q \VS{16}The
heaven,
{\ADD{even}} the
heavens,
{\ADD{are}} the
{\ND{
LORD’s}}: but the
earth hath he
given to the
children of
men.
\par }{\Q \VS{17}The
dead
praise not the
{\ND{
LORD}}, neither any that go
down into
silence.
\par }{\Q \VS{18}But we will
bless the
{\ND{
LORD}} from this time forth and
for
evermore.
Praise the
{\ND{
LORD}}.

\par }\PsalmChap{116}{\Q \VerseOne{1}I
love the
{\ND{
LORD}}, because he hath
heard my
voice
{\ADD{and}} my
supplications.
\par }{\Q \VS{2}Because he hath
inclined his
ear unto me, therefore will I
call upon
{\ADD{him}} as long as I
live.\FTNT{*}{{\FR 116.2 }
as long…: Heb. in my days}
\par }{\Q \VS{3}The
sorrows of
death
compassed me, and the
pains of
hell gat
hold upon me: I
found
trouble and
sorrow.\FTNT{†}{{\FR 116.3 }
gat…: Heb. found me}
\par }{\Q \VS{4}Then
called I upon the
name of the
{\ND{
LORD}}; O
{\ND{
LORD}}, I
beseech thee,
deliver my
soul.
\par }{\Q \VS{5}Gracious
{\ADD{is}} the
{\ND{
LORD}}, and
righteous; yea, our
God
{\ADD{is}}
merciful.
\par }{\Q \VS{6}The
{\ND{
LORD}}
preserveth the
simple: I was brought
low, and he
helped me.
\par }{\Q \VS{7}Return unto thy
rest, O my
soul; for the
{\ND{
LORD}} hath dealt
bountifully with thee.
\par }{\Q \VS{8}For thou hast
delivered my
soul from
death, mine
eyes from
tears,
{\ADD{and}} my
feet from
falling.
\par }{\Q \VS{9}I will
walk
before the
{\ND{
LORD}} in the
land of the
living.
\par }{\BB \par }{\Q \VS{10}I
believed, therefore have I
spoken: I was
greatly
afflicted:
\par }{\Q \VS{11}I
said in my
haste, All
men
{\ADD{are}}
liars.
\par }{\Q \VS{12}What shall I
render unto the
{\ND{
LORD}}
{\ADD{for}} all his
benefits toward me?
\par }{\Q \VS{13}I will
take the
cup of
salvation, and
call upon the
name of the
{\ND{
LORD}}.
\par }{\Q \VS{14}I will
pay my
vows unto the
{\ND{
LORD}} now in the presence of all his
people.
\par }{\Q \VS{15}Precious in the
sight of the
{\ND{
LORD}}
{\ADD{is}} the
death of his
saints.
\par }{\Q \VS{16}O
{\ND{
LORD}},
truly I
{\ADD{am}} thy
servant; I
{\ADD{am}} thy
servant,
{\ADD{and}} the
son of thine
handmaid: thou hast
loosed my
bonds.
\par }{\Q \VS{17}I will
offer to thee the
sacrifice of
thanksgiving, and will
call upon the
name of the
{\ND{
LORD}}.
\par }{\Q \VS{18}I will
pay my
vows unto the
{\ND{
LORD}} now in the presence of all his
people,
\par }{\Q \VS{19}In the
courts of the
{\ND{
LORD’s}}
house, in the
midst of thee, O
Jerusalem.
Praise ye the
{\ND{
LORD}}.

\par }\PsalmChap{117}{\Q \VerseOne{1}O
praise the
{\ND{
LORD}}, all ye
nations:
praise him, all ye
people.
\par }{\Q \VS{2}For his
merciful
kindness is
great toward us: and the
truth of the
{\ND{
LORD}}
{\ADD{endureth}} for
ever.
Praise ye the
{\ND{
LORD}}.

\par }\PsalmChap{118}{\Q \VerseOne{1}O give
thanks unto the
{\ND{
LORD}}; for
{\ADD{he is}}
good: because his
mercy
{\ADD{endureth}} for
ever.
\par }{\Q \VS{2}Let
Israel now
say, that his
mercy
{\ADD{endureth}} for
ever.
\par }{\Q \VS{3}Let the
house of
Aaron now
say, that his
mercy
{\ADD{endureth}} for
ever.
\par }{\Q \VS{4}Let them now that
fear the
{\ND{
LORD}}
say, that his
mercy
{\ADD{endureth}} for
ever.
\par }{\Q \VS{5}I
called upon the
{\ND{
LORD}} in
distress: the
{\ND{
LORD}}
answered me,
{\ADD{and set me}} in a large
place.\FTNT{*}{{\FR 118.5 }
in distress: Heb. out of distress}
\par }{\Q \VS{6}The
{\ND{
LORD}}
{\ADD{is}} on my side; I will not
fear: what
can
man
do unto me?\FTNT{†}{{\FR 118.6 }
on…: Heb. for me}
\par }{\Q \VS{7}The
{\ND{
LORD}} taketh my part with them that
help me: therefore shall I
see
{\ADD{my desire}} upon them that
hate me.
\par }{\Q \VS{8}{\ADD{It is}}
better to
trust in the
{\ND{
LORD}} than to put
confidence in
man.
\par }{\Q \VS{9}{\ADD{It is}}
better to
trust in the
{\ND{
LORD}} than to put
confidence in
princes.
\par }{\Q \VS{10}All
nations compassed me
about: but in the
name of the
{\ND{
LORD}} will I
destroy them.\FTNT{‡}{{\FR 118.10 }
destroy…: Heb. cut them off}
\par }{\Q \VS{11}They compassed me
about; yea, they compassed me
about: but in the
name of the
{\ND{
LORD}} I will
destroy them.
\par }{\Q \VS{12}They compassed me
about like
bees; they are
quenched as the
fire of
thorns: for in the
name of the
{\ND{
LORD}} I will
destroy them.\FTNT{§}{{\FR 118.12 }
destroy: Heb. cut down}
\par }{\Q \VS{13}Thou hast
thrust
sore at me that I might
fall: but the
{\ND{
LORD}}
helped me.
\par }{\Q \VS{14}The
{\ND{
LORD}}
{\ADD{is}} my
strength and
song, and is become my
salvation.
\par }{\Q \VS{15}The
voice of
rejoicing and
salvation
{\ADD{is}} in the
tabernacles of the
righteous: the right
hand of the
{\ND{
LORD}}
doeth
valiantly.
\par }{\Q \VS{16}The right
hand of the
{\ND{
LORD}} is
exalted: the right
hand of the
{\ND{
LORD}}
doeth
valiantly.
\par }{\Q \VS{17}I shall not
die, but
live, and
declare the
works of the
{\ND{
LORD}}.
\par }{\Q \VS{18}The
{\ND{
LORD}} hath
chastened me
sore: but he hath not given me
over unto
death.
\par }{\BB \par }{\Q \VS{19}Open to me the
gates of
righteousness: I will
go into them,
{\ADD{and}} I will
praise the
{\ND{
LORD}}:
\par }{\Q \VS{20}This
gate of the
{\ND{
LORD}}, into which the
righteous shall
enter.
\par }{\Q \VS{21}I will
praise thee: for thou hast
heard me, and art become my
salvation.
\par }{\Q \VS{22}The
stone
{\ADD{which}} the
builders
refused is become the
head
{\ADD{stone}} of the
corner.
\par }{\Q \VS{23}This is the
{\ND{
LORD’s}} doing; it
{\ADD{is}}
marvellous in our
eyes.\FTNT{**}{{\FR 118.23 }
the
{\ND{LORD’s}}…: Heb. from the
{\ND{LORD}}}
\par }{\Q \VS{24}This
{\ADD{is}} the
day
{\ADD{which}} the
{\ND{
LORD}} hath
made; we will
rejoice and be
glad in it.
\par }{\Q \VS{25}Save now, I
beseech thee, O
{\ND{
LORD}}: O
{\ND{
LORD}}, I
beseech thee, send now
prosperity.
\par }{\Q \VS{26}Blessed
{\ADD{be}} he that
cometh in the
name of the
{\ND{
LORD}}: we have
blessed you out of the
house of the
{\ND{
LORD}}.
\par }{\Q \VS{27}God
{\ADD{is}} the
{\ND{
LORD}}, which hath shewed us
light:
bind the
sacrifice with
cords,
{\ADD{even}} unto the
horns of the
altar.
\par }{\Q \VS{28}Thou
{\ADD{art}} my
God, and I will
praise thee:
{\ADD{thou art}} my
God, I will
exalt thee.
\par }{\Q \VS{29}O give
thanks unto the
{\ND{
LORD}}; for
{\ADD{he is}}
good: for his
mercy
{\ADD{endureth}} for
ever.

\par }\PsalmChap{119}{\SH {\TL{ א ALEPH.}}
\par }{\Q \VerseOne{1}Blessed
{\ADD{are}} the
undefiled in the
way, who
walk in the
law of the
{\ND{
LORD}}.\FTNT{*}{{\FR 119.1 }
undefiled: or, perfect, or, sincere}
\par }{\Q \VS{2}Blessed
{\ADD{are}} they that
keep his
testimonies,
{\ADD{and that}}
seek him with the whole
heart.
\par }{\Q \VS{3}They also
do no
iniquity: they
walk in his
ways.
\par }{\BB \par }{\Q \VS{4}Thou hast
commanded
{\ADD{us}} to
keep thy
precepts
diligently.
\par }{\Q \VS{5}O
that my
ways were
directed to
keep thy
statutes!
\par }{\Q \VS{6}Then shall I not be
ashamed, when I have
respect unto all thy
commandments.
\par }{\BB \par }{\Q \VS{7}I will
praise thee with
uprightness of
heart, when I shall have
learned thy
righteous
judgments.\FTNT{†}{{\FR 119.7 }
thy…: Heb. judgments of thy righteousness}
\par }{\Q \VS{8}I will
keep thy
statutes: O
forsake me not
utterly.
\par }{\SH {\TL{ ב BETH.}}
\par }{\Q \VS{9}Wherewithal shall a young
man
cleanse his
way? by taking
heed
{\ADD{thereto}} according to thy
word.
\par }{\Q \VS{10}With my whole
heart have I
sought thee: O let me not
wander from thy
commandments.
\par }{\Q \VS{11}Thy
word have I
hid in mine
heart, that I might not
sin against thee.
\par }{\Q \VS{12}Blessed
{\ADD{art}} thou, O
{\ND{
LORD}}:
teach me thy
statutes.
\par }{\BB \par }{\Q \VS{13}With my
lips have I
declared all the
judgments of thy
mouth.
\par }{\Q \VS{14}I have
rejoiced in the
way of thy
testimonies, as
{\ADD{much as}}
in all
riches.
\par }{\Q \VS{15}I will
meditate in thy
precepts, and have
respect unto thy
ways.
\par }{\Q \VS{16}I will
delight myself in thy
statutes: I will not
forget thy
word.
\par }{\SH {\TL{ ג GIMEL.}}
\par }{\Q \VS{17}Deal
bountifully with thy
servant,
{\ADD{that}} I may
live, and
keep thy
word.
\par }{\Q \VS{18}Open thou mine
eyes, that I may
behold wondrous
things out of thy
law.\FTNT{‡}{{\FR 119.18 }
Open: Heb. Reveal}
\par }{\Q \VS{19}I
{\ADD{am}} a
stranger in the
earth:
hide not thy
commandments from me.
\par }{\Q \VS{20}My
soul
breaketh for the
longing
{\ADD{that it hath}} unto thy
judgments at all
times.
\par }{\Q \VS{21}Thou hast
rebuked the
proud
{\ADD{that are}}
cursed, which do
err from thy
commandments.
\par }{\Q \VS{22}Remove from me
reproach and
contempt; for I have
kept thy
testimonies.
\par }{\Q \VS{23}Princes also did
sit
{\ADD{and}}
speak against me:
{\ADD{but}} thy
servant did
meditate in thy
statutes.
\par }{\Q \VS{24}Thy
testimonies also
{\ADD{are}} my
delight
{\ADD{and}} my
counsellors.\FTNT{§}{{\FR 119.24 }
my counsellors: Heb. men of my counsel}
\par }{\SH {\TL{ ד DALETH.}}
\par }{\Q \VS{25}My
soul
cleaveth unto the
dust:
quicken thou me according to thy
word.
\par }{\BB \par }{\Q \VS{26}I have
declared my
ways, and thou
heardest me:
teach me thy
statutes.
\par }{\Q \VS{27}Make me to
understand the
way of thy
precepts: so shall I
talk of thy wondrous
works.
\par }{\BB \par }{\Q \VS{28}My
soul
melteth for
heaviness:
strengthen thou me according unto thy
word.\FTNT{**}{{\FR 119.28 }
melteth: Heb. droppeth}
\par }{\Q \VS{29}Remove from me the
way of
lying: and grant me thy
law
graciously.
\par }{\BB \par }{\Q \VS{30}I have
chosen the
way of
truth: thy
judgments have I
laid
{\ADD{before me}}.
\par }{\Q \VS{31}I have
stuck unto thy
testimonies: O
{\ND{
LORD}}, put me not to
shame.
\par }{\Q \VS{32}I will
run the
way of thy
commandments, when thou shalt
enlarge my
heart.
\par }{\BB \par }{\SH {\TL{ ה HE.}}
\par }{\Q \VS{33}Teach me, O
{\ND{
LORD}}, the
way of thy
statutes; and I shall
keep it
{\ADD{unto}} the
end.
\par }{\Q \VS{34}Give me
understanding, and I shall
keep thy
law; yea, I shall
observe it with
{\ADD{my}} whole
heart.
\par }{\BB \par }{\Q \VS{35}Make me to
go in the
path of thy
commandments; for therein do I
delight.
\par }{\Q \VS{36}Incline my
heart unto thy
testimonies, and not to
covetousness.
\par }{\Q \VS{37}Turn
away mine
eyes from
beholding
vanity;
{\ADD{and}}
quicken thou me in thy
way.\FTNT{††}{{\FR 119.37 }
Turn…: Heb. Make to pass}
\par }{\Q \VS{38}Stablish thy
word unto thy
servant, who
{\ADD{is devoted}} to thy
fear.
\par }{\Q \VS{39}Turn
away my
reproach which I
fear: for thy
judgments
{\ADD{are}}
good.
\par }{\Q \VS{40}Behold, I have
longed after thy
precepts:
quicken me in thy
righteousness.
\par }{\BB \par }{\SH {\TL{ ו VAU.}}
\par }{\Q \VS{41}Let thy
mercies
come also unto me, O
{\ND{
LORD}},
{\ADD{even}} thy
salvation, according to thy
word.
\par }{\Q \VS{42}So shall I have
wherewith to
answer him that
reproacheth me: for I
trust in thy
word.\FTNT{‡‡}{{\FR 119.42 }
So…: or, So shall I answer him that reproveth me in a thing}
\par }{\BB \par }{\Q \VS{43}And
take not the
word of
truth
utterly out of my
mouth; for I have
hoped in thy
judgments.
\par }{\Q \VS{44}So shall I
keep thy
law
continually for
ever and
ever.
\par }{\BB \par }{\Q \VS{45}And I will
walk at
liberty: for I
seek thy
precepts.\FTNT{§§}{{\FR 119.45 }
at liberty: Heb. at large}
\par }{\Q \VS{46}I will
speak of thy
testimonies also before
kings, and will not be
ashamed.
\par }{\Q \VS{47}And I will
delight myself in thy
commandments, which I have
loved.
\par }{\Q \VS{48}My
hands also will I lift
up unto thy
commandments, which I have
loved; and I will
meditate in thy
statutes.
\par }{\SH {\TL{ ז ZAIN.}}
\par }{\Q \VS{49}Remember the
word unto thy
servant, upon which thou hast caused me to
hope.
\par }{\Q \VS{50}This
{\ADD{is}} my
comfort in my
affliction: for thy
word hath
quickened me.
\par }{\Q \VS{51}The
proud have had me
greatly in
derision:
{\ADD{yet}} have I not
declined from thy
law.
\par }{\Q \VS{52}I
remembered thy
judgments of
old, O
{\ND{
LORD}}; and have
comforted myself.
\par }{\Q \VS{53}Horror hath taken
hold upon me because of the
wicked that
forsake thy
law.
\par }{\Q \VS{54}Thy
statutes have been my
songs in the
house of my
pilgrimage.
\par }{\BB \par }{\Q \VS{55}I have
remembered thy
name, O
{\ND{
LORD}}, in the
night, and have
kept thy
law.
\par }{\Q \VS{56}This I had, because I
kept thy
precepts.
\par }{\SH {\TL{ ח CHETH.}}
\par }{\Q \VS{57}{\ADD{Thou art}} my
portion, O
{\ND{
LORD}}: I have
said that I would
keep thy
words.
\par }{\Q \VS{58}I
intreated thy
favour with
{\ADD{my}} whole
heart: be
merciful unto me according to thy
word.\FTNT{***}{{\FR 119.58 }
favour: Heb. face}
\par }{\BB \par }{\Q \VS{59}I
thought on my
ways, and
turned my
feet unto thy
testimonies.
\par }{\Q \VS{60}I made
haste, and
delayed not to
keep thy
commandments.
\par }{\Q \VS{61}The
bands of the
wicked have
robbed me:
{\ADD{but}} I have not
forgotten thy
law.\FTNT{†††}{{\FR 119.61 }
bands: or, companies}
\par }{\Q \VS{62}At
midnight I will
rise to give
thanks unto thee because of thy
righteous
judgments.
\par }{\Q \VS{63}I
{\ADD{am}} a
companion of all
{\ADD{them}} that
fear thee, and of them that
keep thy
precepts.
\par }{\Q \VS{64}The
earth, O
{\ND{
LORD}}, is
full of thy
mercy:
teach me thy
statutes.
\par }{\BB \par }{\SH {\TL{ ט TETH.}}
\par }{\Q \VS{65}Thou hast
dealt
well with thy
servant, O
{\ND{
LORD}}, according unto thy
word.
\par }{\Q \VS{66}Teach me
good
judgment and
knowledge: for I have
believed thy
commandments.
\par }{\Q \VS{67}Before I was
afflicted I went
astray: but now have I
kept thy
word.
\par }{\Q \VS{68}Thou
{\ADD{art}}
good, and doest
good;
teach me thy
statutes.
\par }{\BB \par }{\Q \VS{69}The
proud have
forged a
lie against me:
{\ADD{but}} I will
keep thy
precepts with
{\ADD{my}} whole
heart.
\par }{\Q \VS{70}Their
heart is as
fat as
grease;
{\ADD{but}} I
delight in thy
law.
\par }{\Q \VS{71}{\ADD{It is}}
good for me that I have been
afflicted; that I might
learn thy
statutes.
\par }{\Q \VS{72}The
law of thy
mouth
{\ADD{is}}
better unto me than
thousands of
gold and
silver.
\par }{\SH {\TL{ י JOD.}}
\par }{\Q \VS{73}Thy
hands have
made me and
fashioned me: give me
understanding, that I may
learn thy
commandments.
\par }{\Q \VS{74}They that
fear thee will be
glad when they
see me; because I have
hoped in thy
word.
\par }{\Q \VS{75}I
know, O
{\ND{
LORD}}, that thy
judgments
{\ADD{are}}
right, and
{\ADD{that}} thou in
faithfulness hast
afflicted me.\FTNT{‡‡‡}{{\FR 119.75 }
right: Heb. righteousness}
\par }{\BB \par }{\Q \VS{76}Let, I pray thee, thy merciful
kindness be for my
comfort, according to thy
word unto thy
servant.\FTNT{§§§}{{\FR 119.76 }
for…: Heb. to comfort me}
\par }{\Q \VS{77}Let thy tender
mercies
come unto me, that I may
live: for thy
law
{\ADD{is}} my
delight.
\par }{\BB \par }{\Q \VS{78}Let the
proud be
ashamed; for they dealt
perversely with me without a
cause:
{\ADD{but}} I will
meditate in thy
precepts.
\par }{\Q \VS{79}Let those that
fear thee
turn unto me, and those that have
known thy
testimonies.
\par }{\Q \VS{80}Let my
heart be
sound in thy
statutes; that I be not
ashamed.
\par }{\BB \par }{\SH {\TL{ כ CAPH.}}
\par }{\Q \VS{81}My
soul
fainteth for thy
salvation:
{\ADD{but}} I
hope in thy
word.
\par }{\Q \VS{82}Mine
eyes
fail for thy
word,
saying, When wilt thou
comfort me?
\par }{\Q \VS{83}For I am become like a
bottle in the
smoke;
{\ADD{yet}} do I not
forget thy
statutes.
\par }{\Q \VS{84}How many
{\ADD{are}} the
days of thy
servant? when wilt thou
execute
judgment on them that
persecute me?
\par }{\BB \par }{\Q \VS{85}The
proud have
digged
pits for me, which
{\ADD{are}} not after thy
law.
\par }{\Q \VS{86}All thy
commandments
{\ADD{are}}
faithful: they
persecute me
wrongfully;
help thou me.\FTNT{*}{{\FR 119.86 }
faithful: Heb. faithfulness}
\par }{\Q \VS{87}They had
almost
consumed me upon
earth; but I
forsook not thy
precepts.
\par }{\Q \VS{88}Quicken me after thy
lovingkindness; so shall I
keep the
testimony of thy
mouth.
\par }{\BB \par }{\SH {\TL{ ל LAMED.}}
\par }{\Q \VS{89}For
ever, O
{\ND{
LORD}}, thy
word is
settled in
heaven.
\par }{\Q \VS{90}Thy
faithfulness
{\ADD{is}} unto
all
generations: thou hast
established the
earth, and it
abideth.\FTNT{†}{{\FR 119.90 }
unto…: Heb. to generation and generation}\FTNT{‡}{{\FR 119.90 }
abideth: Heb. standeth}
\par }{\Q \VS{91}They
continue this
day according to thine
ordinances: for all
{\ADD{are}} thy
servants.
\par }{\Q \VS{92}Unless thy
law
{\ADD{had been}} my
delights, I should then have
perished in mine
affliction.
\par }{\Q \VS{93}I will
never
forget thy
precepts: for with them thou hast
quickened me.
\par }{\Q \VS{94}I
{\ADD{am}} thine,
save me; for I have
sought thy
precepts.
\par }{\Q \VS{95}The
wicked have
waited for me to
destroy me:
{\ADD{but}} I will
consider thy
testimonies.
\par }{\Q \VS{96}I have
seen an
end of all
perfection:
{\ADD{but}} thy
commandment
{\ADD{is}}
exceeding
broad.
\par }{\SH {\TL{ מ MEM.}}
\par }{\Q \VS{97}O how love
I thy
law! it
{\ADD{is}} my
meditation all the
day.
\par }{\BB \par }{\Q \VS{98}Thou through thy
commandments hast made me
wiser than mine
enemies: for they
{\ADD{are}}
ever with me.\FTNT{§}{{\FR 119.98 }
they…: Heb. it is ever with me}
\par }{\Q \VS{99}I have more
understanding than all my
teachers: for thy
testimonies
{\ADD{are}} my
meditation.
\par }{\Q \VS{100}I
understand more than the
ancients, because I
keep thy
precepts.
\par }{\Q \VS{101}I have
refrained my
feet from every
evil
way, that I might
keep thy
word.
\par }{\Q \VS{102}I have not
departed from thy
judgments: for thou hast
taught me.
\par }{\BB \par }{\Q \VS{103}How
sweet are thy
words unto my
taste!
{\ADD{yea, sweeter}} than
honey to my
mouth!\FTNT{**}{{\FR 119.103 }
taste: Heb. palate}
\par }{\Q \VS{104}Through thy
precepts I get
understanding: therefore I
hate every
false
way.
\par }{\SH {\TL{ נ NUN.}}
\par }{\Q \VS{105}Thy
word
{\ADD{is}} a
lamp unto my
feet, and a
light unto my
path.\FTNT{††}{{\FR 119.105 }
lamp: or, candle}
\par }{\Q \VS{106}I have
sworn, and I will
perform
{\ADD{it}}, that I will
keep thy
righteous
judgments.
\par }{\Q \VS{107}I am
afflicted very
much:
quicken me, O
{\ND{
LORD}}, according unto thy
word.
\par }{\Q \VS{108}Accept, I beseech thee, the freewill
offerings of my
mouth, O
{\ND{
LORD}}, and
teach me thy
judgments.
\par }{\BB \par }{\Q \VS{109}My
soul
{\ADD{is}}
continually in my
hand: yet do I not
forget thy
law.
\par }{\Q \VS{110}The
wicked have
laid a
snare for me: yet I
erred not from thy
precepts.
\par }{\BB \par }{\Q \VS{111}Thy
testimonies have I taken as an
heritage for
ever: for they
{\ADD{are}} the
rejoicing of my
heart.
\par }{\Q \VS{112}I have
inclined mine
heart to
perform thy
statutes
alway,
{\ADD{even unto}} the
end.\FTNT{‡‡}{{\FR 119.112 }
to perform: Heb. to do}
\par }{\SH {\TL{ ס SAMECH.}}
\par }{\Q \VS{113}I
hate
{\ADD{vain}}
thoughts: but thy
law do I
love.
\par }{\Q \VS{114}Thou
{\ADD{art}} my hiding
place and my
shield: I
hope in thy
word.
\par }{\Q \VS{115}Depart from me, ye
evildoers: for I will
keep the
commandments of my
God.
\par }{\BB \par }{\Q \VS{116}Uphold me according unto thy
word, that I may
live: and let me not be
ashamed of my
hope.
\par }{\Q \VS{117}Hold thou me
up, and I shall be
safe: and I will have
respect unto thy
statutes
continually.
\par }{\BB \par }{\Q \VS{118}Thou hast trodden
down all them that
err from thy
statutes: for their
deceit
{\ADD{is}}
falsehood.
\par }{\Q \VS{119}Thou puttest
away all the
wicked of the
earth
{\ADD{like}}
dross: therefore I
love thy
testimonies.\FTNT{§§}{{\FR 119.119 }
puttest…: Heb. causest to cease}
\par }{\Q \VS{120}My
flesh
trembleth for
fear of thee; and I am
afraid of thy
judgments.
\par }{\BB \par }{\SH {\TL{ ע AIN.}}
\par }{\Q \VS{121}I have
done
judgment and
justice:
leave me not to mine
oppressors.
\par }{\Q \VS{122}Be
surety for thy
servant for
good: let not the
proud
oppress me.
\par }{\Q \VS{123}Mine
eyes
fail for thy
salvation, and for the
word of thy
righteousness.
\par }{\BB \par }{\Q \VS{124}Deal with thy
servant according unto thy
mercy, and
teach me thy
statutes.
\par }{\Q \VS{125}I
{\ADD{am}} thy
servant; give me
understanding, that I may
know thy
testimonies.
\par }{\Q \VS{126}{\ADD{It is}}
time for
{\ADD{thee}},
{\ND{
LORD}}, to
work:
{\ADD{for}} they have made
void thy
law.
\par }{\BB \par }{\Q \VS{127}Therefore I
love thy
commandments above
gold; yea, above fine
gold.
\par }{\Q \VS{128}Therefore I esteem all
{\ADD{thy}}
precepts
{\ADD{concerning}} all
{\ADD{things to be}}
right;
{\ADD{and}} I
hate every
false
way.
\par }{\SH {\TL{ פ PE.}}
\par }{\Q \VS{129}Thy
testimonies
{\ADD{are}}
wonderful: therefore doth my
soul
keep them.
\par }{\Q \VS{130}The
entrance of thy
words giveth
light; it giveth
understanding unto the
simple.
\par }{\Q \VS{131}I
opened my
mouth, and
panted: for I
longed for thy
commandments.
\par }{\Q \VS{132}Look thou upon me, and be
merciful unto me, as thou
usest to do unto those that
love thy
name.\FTNT{***}{{\FR 119.132 }
as thou…: Heb. according to the custom toward those, etc}
\par }{\Q \VS{133}Order my
steps in thy
word: and let not any
iniquity have
dominion over me.
\par }{\Q \VS{134}Deliver me from the
oppression of
man: so will I
keep thy
precepts.
\par }{\Q \VS{135}Make thy
face to
shine upon thy
servant; and
teach me thy
statutes.
\par }{\Q \VS{136}Rivers of
waters run
down mine
eyes, because they
keep not thy
law.
\par }{\BB \par }{\SH {\TL{ צ TZADDI.}}
\par }{\Q \VS{137}Righteous
{\ADD{art}} thou, O
{\ND{
LORD}}, and
upright
{\ADD{are}} thy
judgments.
\par }{\Q \VS{138}Thy
testimonies
{\ADD{that}} thou hast
commanded
{\ADD{are}}
righteous and
very
faithful.\FTNT{†††}{{\FR 119.138 }
righteous: Heb. righteousness}\FTNT{‡‡‡}{{\FR 119.138 }
faithful: Heb. faithfulness}
\par }{\Q \VS{139}My
zeal hath
consumed me, because mine
enemies have
forgotten thy
words.\FTNT{§§§}{{\FR 119.139 }
consumed…: Heb. cut me off}
\par }{\Q \VS{140}Thy
word
{\ADD{is}}
very
pure: therefore thy
servant
loveth it.\FTNT{*}{{\FR 119.140 }
pure: Heb. tried, or, refined}
\par }{\Q \VS{141}I
{\ADD{am}}
small and
despised:
{\ADD{yet}} do not I
forget thy
precepts.
\par }{\Q \VS{142}Thy
righteousness
{\ADD{is}} an
everlasting
righteousness, and thy
law
{\ADD{is}} the
truth.
\par }{\BB \par }{\Q \VS{143}Trouble and
anguish have taken
hold on me:
{\ADD{yet}} thy
commandments
{\ADD{are}} my
delights.\FTNT{†}{{\FR 119.143 }
taken…: Heb. found me}
\par }{\Q \VS{144}The
righteousness of thy
testimonies
{\ADD{is}}
everlasting: give me
understanding, and I shall
live.
\par }{\BB \par }{\SH {\TL{ ק KOPH.}}
\par }{\Q \VS{145}I
cried with
{\ADD{my}} whole
heart;
hear me, O
{\ND{
LORD}}: I will
keep thy
statutes.
\par }{\Q \VS{146}I
cried unto thee;
save me, and I shall
keep thy
testimonies.\FTNT{‡}{{\FR 119.146 }
and I…: or, that I may keep}
\par }{\BB \par }{\Q \VS{147}I
prevented the dawning of the
morning, and
cried: I
hoped in thy
word.
\par }{\Q \VS{148}Mine
eyes
prevent the
{\ADD{night}}
watches, that I might
meditate in thy
word.
\par }{\Q \VS{149}Hear my
voice according unto thy
lovingkindness: O
{\ND{
LORD}},
quicken me according to thy
judgment.
\par }{\BB \par }{\Q \VS{150}They draw
nigh that
follow after
mischief: they are
far from thy
law.
\par }{\Q \VS{151}Thou
{\ADD{art}}
near, O
{\ND{
LORD}}; and all thy
commandments
{\ADD{are}}
truth.
\par }{\Q \VS{152}Concerning thy
testimonies, I have
known of
old that thou hast
founded them for
ever.
\par }{\BB \par }{\SH {\TL{ ר RESH.}}
\par }{\Q \VS{153}Consider mine
affliction, and
deliver me: for I do not
forget thy
law.
\par }{\Q \VS{154}Plead my
cause, and
deliver me:
quicken me according to thy
word.
\par }{\Q \VS{155}Salvation
{\ADD{is}}
far from the
wicked: for they
seek not thy
statutes.
\par }{\Q \VS{156}Great
{\ADD{are}} thy tender
mercies, O
{\ND{
LORD}}:
quicken me according to thy
judgments.\FTNT{§}{{\FR 119.156 }
Great: or, Many}
\par }{\Q \VS{157}Many
{\ADD{are}} my
persecutors and mine
enemies;
{\ADD{yet}} do I not
decline from thy
testimonies.
\par }{\Q \VS{158}I
beheld the
transgressors, and was
grieved; because they
kept not thy
word.
\par }{\Q \VS{159}Consider how I
love thy
precepts:
quicken me, O
{\ND{
LORD}}, according to thy
lovingkindness.
\par }{\Q \VS{160}Thy
word
{\ADD{is}}
true
{\ADD{from}} the
beginning: and every one of thy
righteous
judgments
{\ADD{endureth}} for
ever.\FTNT{**}{{\FR 119.160 }
Thy word…: Heb. The beginning of thy word is true}
\par }{\SH {\TL{ ש SCHIN.}}
\par }{\Q \VS{161}Princes have
persecuted me without a
cause: but my
heart standeth in
awe of thy
word.
\par }{\Q \VS{162}I
rejoice at thy
word, as one that
findeth
great
spoil.
\par }{\Q \VS{163}I
hate and
abhor
lying:
{\ADD{but}} thy
law do I
love.
\par }{\Q \VS{164}Seven
times a
day do I
praise thee because of thy
righteous
judgments.
\par }{\Q \VS{165}Great
peace have they which
love thy
law: and nothing shall
offend them.\FTNT{††}{{\FR 119.165 }
nothing…: Heb. they shall have no stumblingblock}
\par }{\Q \VS{166}{\ND{
LORD}}, I have
hoped for thy
salvation, and
done thy
commandments.
\par }{\BB \par }{\Q \VS{167}My
soul hath
kept thy
testimonies; and I
love them
exceedingly.
\par }{\Q \VS{168}I have
kept thy
precepts and thy
testimonies: for all my
ways
{\ADD{are}} before thee.
\par }{\BB \par }{\SH {\TL{ ת TAU.}}
\par }{\Q \VS{169}Let my
cry come
near
before thee, O
{\ND{
LORD}}: give me
understanding according to thy
word.
\par }{\Q \VS{170}Let my
supplication
come
before thee:
deliver me according to thy
word.
\par }{\Q \VS{171}My
lips shall
utter
praise, when thou hast
taught me thy
statutes.
\par }{\Q \VS{172}My
tongue shall
speak of thy
word: for all thy
commandments
{\ADD{are}}
righteousness.
\par }{\BB \par }{\Q \VS{173}Let thine
hand
help me; for I have
chosen thy
precepts.
\par }{\Q \VS{174}I have
longed for thy
salvation, O
{\ND{
LORD}}; and thy
law
{\ADD{is}} my
delight.
\par }{\Q \VS{175}Let my
soul
live, and it shall
praise thee; and let thy
judgments
help me.
\par }{\Q \VS{176}I have gone
astray like a
lost
sheep;
seek thy
servant; for I do not
forget thy
commandments.

\par }\PsalmChap{120}{\D A
Song of
degrees.
\par }{\Q \VerseOne{1}In my
distress I
cried unto the
{\ND{
LORD}}, and he
heard me.
\par }{\Q \VS{2}Deliver my
soul, O
{\ND{
LORD}}, from
lying
lips,
{\ADD{and}} from a
deceitful
tongue.
\par }{\Q \VS{3}What shall be
given unto thee? or what shall be
done unto thee, thou
false
tongue?\FTNT{*}{{\FR 120.3 }
What shall be given…: or, What shall the deceitful tongue give unto thee? or, what shall is profit thee?}\FTNT{†}{{\FR 120.3 }
done: Heb. added}
\par }{\Q \VS{4}Sharp
arrows of the
mighty, with
coals of
juniper.\FTNT{‡}{{\FR 120.4 }
Sharp…: or, It is as the sharp arrows of the mighty man, with coals of juniper}
\par }{\BB \par }{\Q \VS{5}Woe is me, that I
sojourn in
Mesech,
{\ADD{that}} I
dwell in the
tents of
Kedar!
\par }{\Q \VS{6}My
soul hath
long
dwelt with him that
hateth
peace.
\par }{\Q \VS{7}I
{\ADD{am for}}
peace: but when I
speak, they
{\ADD{are}} for
war.\FTNT{§}{{\FR 120.7 }
for peace: or, a man of peace}

\par }\PsalmChap{121}{\D A
Song of
degrees.
\par }{\Q \VerseOne{1}I will lift
up mine
eyes unto the
hills, from
whence
cometh my
help.\FTNT{*}{{\FR 121.1 }
I will…: or, Shall I lift up mine eyes to the hills? whence should my help come?}
\par }{\Q \VS{2}My
help
{\ADD{cometh}} from the
{\ND{
LORD}}, which
made
heaven and
earth.
\par }{\Q \VS{3}He will not
suffer thy
foot to be
moved: he that
keepeth thee will not
slumber.
\par }{\Q \VS{4}Behold, he that
keepeth
Israel shall neither
slumber nor
sleep.
\par }{\Q \VS{5}The
{\ND{
LORD}}
{\ADD{is}} thy
keeper: the
{\ND{
LORD}}
{\ADD{is}} thy
shade upon thy
right
hand.
\par }{\Q \VS{6}The
sun shall not
smite thee by
day, nor the
moon by
night.
\par }{\Q \VS{7}The
{\ND{
LORD}} shall
preserve thee from all
evil: he shall
preserve thy
soul.
\par }{\Q \VS{8}The
{\ND{
LORD}} shall
preserve thy going
out and thy coming
in from this time forth, and even
for
evermore.

\par }\PsalmChap{122}{\D A
Song of
degrees of
David.
\par }{\Q \VerseOne{1}I was
glad when they
said unto me, Let us
go into the
house of the
{\ND{
LORD}}.
\par }{\Q \VS{2}Our
feet shall
stand within thy
gates, O
Jerusalem.
\par }{\Q \VS{3}Jerusalem is
builded as a
city that is
compact
together:
\par }{\Q \VS{4}Whither the
tribes go
up, the
tribes of the
{\ND{
LORD}}, unto the
testimony of
Israel, to give
thanks unto the
name of the
{\ND{
LORD}}.
\par }{\Q \VS{5}For there are
set
thrones of
judgment, the
thrones of the
house of
David.\FTNT{*}{{\FR 122.5 }
are…: Heb. do sit}
\par }{\BB \par }{\Q \VS{6}Pray for the
peace of
Jerusalem: they shall
prosper that
love thee.
\par }{\Q \VS{7}Peace be within thy
walls,
{\ADD{and}}
prosperity within thy
palaces.
\par }{\Q \VS{8}For my
brethren and
companions’ sakes, I will now
say,
Peace
{\ADD{be}} within thee.
\par }{\Q \VS{9}Because of the
house of the
{\ND{
LORD}} our
God I will
seek thy
good.

\par }\PsalmChap{123}{\D A
Song of
degrees.
\par }{\Q \VerseOne{1}Unto thee lift I
up mine
eyes, O thou that
dwellest in the
heavens.
\par }{\Q \VS{2}Behold, as the
eyes of
servants
{\ADD{look}} unto the
hand of their
masters,
{\ADD{and}} as the
eyes of a
maiden unto the
hand of her
mistress; so our
eyes
{\ADD{wait}} upon the
{\ND{
LORD}} our
God, until that he have
mercy upon us.
\par }{\Q \VS{3}Have
mercy upon us, O
{\ND{
LORD}}, have
mercy upon us: for we are
exceedingly
filled with
contempt.
\par }{\Q \VS{4}Our
soul is
exceedingly
filled with the
scorning of those that are at
ease,
{\ADD{and}} with the
contempt of the
proud.

\par }\PsalmChap{124}{\D A
Song of
degrees of
David.
\par }{\Q \VerseOne{1}If
{\ADD{it had not been}} the
{\ND{
LORD}} who was on our side, now may
Israel
say;
\par }{\Q \VS{2}If
{\ADD{it had not been}} the
{\ND{
LORD}} who was on our side, when
men rose
up against us:
\par }{\Q \VS{3}Then they had swallowed us
up
quick, when their
wrath was
kindled against us:
\par }{\Q \VS{4}Then the
waters had
overwhelmed us, the
stream had gone
over our
soul:
\par }{\Q \VS{5}Then the
proud
waters had gone
over our
soul.
\par }{\BB \par }{\Q \VS{6}Blessed
{\ADD{be}} the
{\ND{
LORD}}, who hath not
given us
{\ADD{as}} a
prey to their
teeth.
\par }{\Q \VS{7}Our
soul is
escaped as a
bird out of the
snare of the
fowlers: the
snare is
broken, and we are
escaped.
\par }{\Q \VS{8}Our
help
{\ADD{is}} in the
name of the
{\ND{
LORD}}, who
made
heaven and
earth.

\par }\PsalmChap{125}{\D A
Song of
degrees.
\par }{\Q \VerseOne{1}They that
trust in the
{\ND{
LORD}}
{\ADD{shall be}} as
mount
Zion,
{\ADD{which}} cannot be
removed,
{\ADD{but}}
abideth for
ever.
\par }{\Q \VS{2}As the
mountains
{\ADD{are}} round
about
Jerusalem, so the
{\ND{
LORD}}
{\ADD{is}} round
about his
people from henceforth even
for
ever.
\par }{\Q \VS{3}For the
rod of the
wicked shall not
rest upon the
lot of the
righteous;
lest the
righteous put
forth their
hands unto
iniquity.\FTNT{*}{{\FR 125.3 }
the wicked: Heb. wickedness}
\par }{\BB \par }{\Q \VS{4}Do
good, O
{\ND{
LORD}}, unto
{\ADD{those that be}}
good, and
{\ADD{to them that are}}
upright in their
hearts.
\par }{\Q \VS{5}As for such as turn
aside unto their crooked
ways, the
{\ND{
LORD}} shall lead them
forth with the
workers of
iniquity:
{\ADD{but}}
peace
{\ADD{shall be}} upon
Israel.

\par }\PsalmChap{126}{\D A
Song of
degrees.
\par }{\Q \VerseOne{1}When the
{\ND{
LORD}} turned
again the
captivity of
Zion, we were like them that
dream.\FTNT{*}{{\FR 126.1 }
turned…: Heb. returned the returning}
\par }{\Q \VS{2}Then was our
mouth
filled with
laughter, and our
tongue with
singing: then
said they among the
heathen, The
{\ND{
LORD}} hath
done great
things for them.\FTNT{†}{{\FR 126.2 }
hath…: Heb. hath magnified to do with them}
\par }{\Q \VS{3}The
{\ND{
LORD}} hath
done great
things for us;
{\ADD{whereof}} we are
glad.
\par }{\BB \par }{\Q \VS{4}Turn
again our
captivity, O
{\ND{
LORD}}, as the
streams in the
south.
\par }{\Q \VS{5}They that
sow in
tears shall
reap in
joy.\FTNT{‡}{{\FR 126.5 }
joy: or, singing}
\par }{\Q \VS{6}He that
goeth
forth and
weepeth,
bearing
precious
seed, shall
doubtless
come again with
rejoicing,
bringing his
sheaves
{\ADD{with him}}.\FTNT{§}{{\FR 126.6 }
precious…: or, seed basket}

\par }\PsalmChap{127}{\D A
Song of
degrees for
Solomon.
\par }{\Q \VerseOne{1}Except the
{\ND{
LORD}}
build the
house, they
labour in
vain that
build it: except the
{\ND{
LORD}}
keep the
city, the
watchman
waketh
{\ADD{but}} in
vain.\FTNT{*}{{\FR 127.1 }
for…: or, of Solomon}\FTNT{†}{{\FR 127.1 }
that…: Heb. that are builders of it in it}
\par }{\Q \VS{2}{\ADD{It is}}
vain for you to rise
up
early, to sit
up
late, to
eat the
bread of
sorrows:
{\ADD{for}} so he
giveth his
beloved
sleep.
\par }{\Q \VS{3}Lo,
children
{\ADD{are}} an
heritage of the
{\ND{
LORD}}:
{\ADD{and}} the
fruit of the
womb
{\ADD{is his}}
reward.
\par }{\Q \VS{4}As
arrows
{\ADD{are}} in the
hand of a mighty
man; so
{\ADD{are}}
children of the
youth.
\par }{\Q \VS{5}Happy
{\ADD{is}} the
man that hath his
quiver
full of them: they shall not be
ashamed, but they shall
speak with the
enemies in the
gate.\FTNT{‡}{{\FR 127.5 }
his…: Heb. filled his quiver with them}\FTNT{§}{{\FR 127.5 }
speak…: or, subdue, or, destroy}

\par }\PsalmChap{128}{\D A
Song of
degrees.
\par }{\Q \VerseOne{1}Blessed
{\ADD{is}} every one that
feareth the
{\ND{
LORD}}; that
walketh in his
ways.
\par }{\Q \VS{2}For thou shalt
eat the
labour of thine
hands:
happy
{\ADD{shalt}} thou
{\ADD{be}}, and
{\ADD{it shall be}}
well with thee.
\par }{\Q \VS{3}Thy
wife
{\ADD{shall be}} as a
fruitful
vine by the
sides of thine
house: thy
children like
olive
plants round
about thy
table.
\par }{\Q \VS{4}Behold, that thus shall the
man be
blessed that
feareth the
{\ND{
LORD}}.
\par }{\Q \VS{5}The
{\ND{
LORD}} shall
bless thee out of
Zion: and thou shalt
see the
good of
Jerusalem all the
days of thy
life.
\par }{\Q \VS{6}Yea, thou shalt
see thy
children’s
children,
{\ADD{and}}
peace upon
Israel.

\par }\PsalmChap{129}{\D A
Song of
degrees.
\par }{\Q \VerseOne{1}Many a
time have they
afflicted me from my
youth, may
Israel now
say:\FTNT{*}{{\FR 129.1 }
Many…: or, Much}
\par }{\Q \VS{2}Many a
time have they
afflicted me from my
youth: yet they have not
prevailed against me.\FTNT{†}{{\FR 129.2 }
Many…: or, Much}
\par }{\Q \VS{3}The
plowers
plowed upon my
back: they made
long their
furrows.
\par }{\Q \VS{4}The
{\ND{
LORD}}
{\ADD{is}}
righteous: he hath cut
asunder the
cords of the
wicked.
\par }{\BB \par }{\Q \VS{5}Let them all be
confounded and
turned
back that
hate
Zion.
\par }{\Q \VS{6}Let them be as the
grass
{\ADD{upon}} the
housetops, which
withereth
afore it groweth
up:
\par }{\Q \VS{7}Wherewith the
mower
filleth not his
hand; nor he that bindeth
sheaves his
bosom.
\par }{\Q \VS{8}Neither do they which go
by
say, The
blessing of the
{\ND{
LORD}}
{\ADD{be}} upon you: we
bless you in the
name of the
{\ND{
LORD}}.

\par }\PsalmChap{130}{\D A
Song of
degrees.
\par }{\Q \VerseOne{1}Out of the
depths have I
cried unto thee, O
{\ND{
LORD}}.
\par }{\Q \VS{2}Lord,
hear my
voice: let thine
ears be
attentive to the
voice of my
supplications.
\par }{\Q \VS{3}If thou,
{\ND{
LORD}}, shouldest
mark
iniquities, O
Lord, who shall
stand?
\par }{\Q \VS{4}But
{\ADD{there is}}
forgiveness with thee, that thou mayest be
feared.
\par }{\BB \par }{\Q \VS{5}I
wait for the
{\ND{
LORD}}, my
soul doth
wait, and in his
word do I
hope.
\par }{\Q \VS{6}My
soul
{\ADD{waiteth}} for the
Lord more than they that
watch for the
morning:
{\ADD{I say, more than}} they that
watch for the
morning.\FTNT{*}{{\FR 130.6 }
I say…: or, which watch unto}
\par }{\Q \VS{7}Let
Israel
hope in the
{\ND{
LORD}}: for with the
{\ND{
LORD}}
{\ADD{there is}}
mercy, and with him
{\ADD{is}}
plenteous
redemption.
\par }{\Q \VS{8}And he shall
redeem
Israel from all his
iniquities.

\par }\PsalmChap{131}{\D A
Song of
degrees of
David.
\par }{\Q \VerseOne{1}{\ND{
LORD}}, my
heart is not
haughty, nor mine
eyes
lofty: neither do I
exercise myself in great
matters, or in things too
high for me.\FTNT{*}{{\FR 131.1 }
exercise…: Heb. walk}\FTNT{†}{{\FR 131.1 }
high: Heb. wonderful}
\par }{\Q \VS{2}Surely I have
behaved and
quieted
myself, as a child that is
weaned of his
mother: my
soul
{\ADD{is}} even as a weaned
child.\FTNT{‡}{{\FR 131.2 }
myself: Heb. my soul}
\par }{\Q \VS{3}Let
Israel
hope in the
{\ND{
LORD}} from henceforth and
for
ever.\FTNT{§}{{\FR 131.3 }
henceforth: Heb. now}

\par }\PsalmChap{132}{\D A
Song of
degrees.
\par }{\Q \VerseOne{1}{\ND{
LORD}},
remember
David,
{\ADD{and}} all his
afflictions:
\par }{\Q \VS{2}How he
sware unto the
{\ND{
LORD}},
{\ADD{and}}
vowed unto the
mighty
{\ADD{God}} of
Jacob;
\par }{\Q \VS{3}Surely I will not
come into the
tabernacle of my
house, nor go
up into my
bed;
\par }{\Q \VS{4}I will not
give
sleep to mine
eyes,
{\ADD{or}}
slumber to mine
eyelids,
\par }{\Q \VS{5}Until I find
out a
place for the
{\ND{
LORD}}, an
habitation for the
mighty
{\ADD{God}} of
Jacob.\FTNT{*}{{\FR 132.5 }
an habitation: Heb. habitations}
\par }{\Q \VS{6}Lo, we
heard of it at
Ephratah: we
found it in the
fields of the
wood.
\par }{\Q \VS{7}We will
go into his
tabernacles: we will
worship at his
footstool.
\par }{\Q \VS{8}Arise, O
{\ND{
LORD}}, into thy
rest; thou, and the
ark of thy
strength.
\par }{\Q \VS{9}Let thy
priests be
clothed with
righteousness; and let thy
saints shout for
joy.
\par }{\Q \VS{10}For thy
servant
David’s sake turn not
away the
face of thine
anointed.
\par }{\BB \par }{\Q \VS{11}The
{\ND{
LORD}} hath
sworn
{\ADD{in}}
truth unto
David; he will not
turn from it; Of the
fruit of thy
body will I
set upon thy
throne.\FTNT{†}{{\FR 132.11 }
body: Heb. belly}
\par }{\Q \VS{12}If thy
children will
keep my
covenant and my
testimony
that I shall
teach them, their
children shall also
sit upon thy
throne for
evermore.
\par }{\Q \VS{13}For the
{\ND{
LORD}} hath
chosen
Zion; he hath
desired
{\ADD{it}} for his
habitation.
\par }{\Q \VS{14}This
{\ADD{is}} my
rest for
ever: here will I
dwell; for I have
desired it.
\par }{\Q \VS{15}I will
abundantly
bless her
provision: I will
satisfy her
poor with
bread.\FTNT{‡}{{\FR 132.15 }
abundantly: or, surely}
\par }{\Q \VS{16}I will also
clothe her
priests with
salvation: and her
saints shall shout
aloud for
joy.
\par }{\Q \VS{17}There will I
make the
horn of
David to
bud: I have
ordained a
lamp for mine
anointed.\FTNT{§}{{\FR 132.17 }
lamp: or, candle}
\par }{\Q \VS{18}His
enemies will I
clothe with
shame: but upon himself shall his
crown
flourish.

\par }\PsalmChap{133}{\D A
Song of
degrees of
David.
\par }{\Q \VerseOne{1}Behold, how
good and how
pleasant
{\ADD{it is}} for
brethren to
dwell
together in
unity!\FTNT{*}{{\FR 133.1 }
together…: Heb. even together}
\par }{\Q \VS{2}{\ADD{It is}} like the
precious
ointment upon the
head, that ran
down upon the
beard,
{\ADD{even}}
Aaron’s
beard: that went
down to the
skirts of his
garments;
\par }{\Q \VS{3}As the
dew of
Hermon,
{\ADD{and as the dew}} that
descended upon the
mountains of
Zion: for there the
{\ND{
LORD}}
commanded the
blessing,
{\ADD{even}}
life for
evermore.

\par }\PsalmChap{134}{\D A
Song of
degrees.
\par }{\Q \VerseOne{1}Behold,
bless ye the
{\ND{
LORD}}, all
{\ADD{ye}}
servants of the
{\ND{
LORD}}, which by
night
stand in the
house of the
{\ND{
LORD}}.
\par }{\Q \VS{2}Lift
up your
hands
{\ADD{in}} the
sanctuary, and
bless the
{\ND{
LORD}}.\FTNT{*}{{\FR 134.2 }
the sanctuary: or, holiness}
\par }{\Q \VS{3}The
{\ND{
LORD}} that
made
heaven and
earth
bless thee out of
Zion.

\par }\PsalmChap{135}{\Q \VerseOne{1}Praise ye the
{\ND{
LORD}}.
Praise ye the
name of the
{\ND{
LORD}};
praise
{\ADD{him}}, O ye
servants of the
{\ND{
LORD}}.
\par }{\Q \VS{2}Ye that
stand in the
house of the
{\ND{
LORD}}, in the
courts of the
house of our
God,
\par }{\Q \VS{3}Praise the
{\ND{
LORD}}; for the
{\ND{
LORD}}
{\ADD{is}}
good: sing
praises unto his
name; for
{\ADD{it is}}
pleasant.
\par }{\Q \VS{4}For the
{\ND{
LORD}} hath
chosen
Jacob unto himself,
{\ADD{and}}
Israel for his peculiar
treasure.
\par }{\BB \par }{\Q \VS{5}For I
know that the
{\ND{
LORD}}
{\ADD{is}}
great, and
{\ADD{that}} our
Lord
{\ADD{is}} above all
gods.
\par }{\Q \VS{6}Whatsoever the
{\ND{
LORD}}
pleased,
{\ADD{that}}
did he in
heaven, and in
earth, in the
seas, and all deep
places.
\par }{\Q \VS{7}He causeth the
vapours to
ascend from the
ends of the
earth; he
maketh
lightnings for the
rain; he
bringeth the
wind out of his
treasuries.
\par }{\Q \VS{8}Who
smote the
firstborn of
Egypt, both of
man and
beast.\FTNT{*}{{\FR 135.8 }
both…: Heb. from man unto beast}
\par }{\Q \VS{9}{\ADD{Who}}
sent
tokens and
wonders into the
midst of thee, O
Egypt, upon
Pharaoh, and upon all his
servants.
\par }{\Q \VS{10}Who
smote
great
nations, and
slew
mighty
kings;
\par }{\Q \VS{11}Sihon
king of the
Amorites, and
Og
king of
Bashan, and all the
kingdoms of
Canaan:
\par }{\Q \VS{12}And
gave their
land
{\ADD{for}} an
heritage, an
heritage unto
Israel his
people.
\par }{\Q \VS{13}Thy
name, O
{\ND{
LORD}},
{\ADD{endureth}} for
ever;
{\ADD{and}} thy
memorial, O
{\ND{
LORD}}, throughout
all
generations.\FTNT{†}{{\FR 135.13 }
throughout…: Heb. to generation and generation}
\par }{\Q \VS{14}For the
{\ND{
LORD}} will
judge his
people, and he will
repent himself concerning his
servants.
\par }{\BB \par }{\Q \VS{15}The
idols of the
heathen
{\ADD{are}}
silver and
gold, the
work of
men’s
hands.
\par }{\Q \VS{16}They have
mouths, but they
speak not;
eyes have they, but they
see not;
\par }{\Q \VS{17}They have
ears, but they
hear not; neither is
there
{\ADD{any}}
breath in their
mouths.
\par }{\Q \VS{18}They that
make them are like unto them:
{\ADD{so is}} every one that
trusteth in them.
\par }{\Q \VS{19}Bless the
{\ND{
LORD}}, O
house of
Israel:
bless the
{\ND{
LORD}}, O
house of
Aaron:
\par }{\Q \VS{20}Bless the
{\ND{
LORD}}, O
house of
Levi: ye that
fear the
{\ND{
LORD}},
bless the
{\ND{
LORD}}.
\par }{\Q \VS{21}Blessed be the
{\ND{
LORD}} out of
Zion, which
dwelleth at
Jerusalem.
Praise ye the
{\ND{
LORD}}.

\par }\PsalmChap{136}{\Q \VerseOne{1}O give
thanks unto the
{\ND{
LORD}}; for
{\ADD{he is}}
good: for his
mercy
{\ADD{endureth}} for
ever.
\par }{\Q \VS{2}O give
thanks unto the
God of
gods: for his
mercy
{\ADD{endureth}} for
ever.
\par }{\Q \VS{3}O give
thanks to the
Lord of
lords: for his
mercy
{\ADD{endureth}} for
ever.
\par }{\Q \VS{4}To him who alone
doeth
great
wonders: for his
mercy
{\ADD{endureth}} for
ever.
\par }{\Q \VS{5}To him that by
wisdom
made the
heavens: for his
mercy
{\ADD{endureth}} for
ever.
\par }{\Q \VS{6}To him that stretched
out the
earth above the
waters: for his
mercy
{\ADD{endureth}} for
ever.
\par }{\Q \VS{7}To him that
made
great
lights: for his
mercy
{\ADD{endureth}} for
ever:
\par }{\Q \VS{8}The
sun to
rule by
day: for his
mercy
{\ADD{endureth}} for
ever:\FTNT{*}{{\FR 136.8 }
to rule…: Heb. for the rulings by day}
\par }{\Q \VS{9}The
moon and
stars to
rule by
night: for his
mercy
{\ADD{endureth}} for
ever.
\par }{\BB \par }{\Q \VS{10}To him that
smote
Egypt in their
firstborn: for his
mercy
{\ADD{endureth}} for
ever:
\par }{\Q \VS{11}And brought
out
Israel from
among them: for his
mercy
{\ADD{endureth}} for
ever:
\par }{\Q \VS{12}With a
strong
hand, and with a stretched
out
arm: for his
mercy
{\ADD{endureth}} for
ever.
\par }{\Q \VS{13}To him which
divided the
Red
sea into
parts: for his
mercy
{\ADD{endureth}} for
ever:
\par }{\Q \VS{14}And made
Israel to pass
through the
midst of it: for his
mercy
{\ADD{endureth}} for
ever:
\par }{\Q \VS{15}But
overthrew
Pharaoh and his
host in the
Red
sea: for his
mercy
{\ADD{endureth}} for
ever.\FTNT{†}{{\FR 136.15 }
overthrew: Heb. shaked off}
\par }{\Q \VS{16}To him which
led his
people through the
wilderness: for his
mercy
{\ADD{endureth}} for
ever.
\par }{\Q \VS{17}To him which
smote
great
kings: for his
mercy
{\ADD{endureth}} for
ever:
\par }{\Q \VS{18}And
slew
famous
kings: for his
mercy
{\ADD{endureth}} for
ever:
\par }{\Q \VS{19}Sihon
king of the
Amorites: for his
mercy
{\ADD{endureth}} for
ever:
\par }{\Q \VS{20}And
Og the
king of
Bashan: for his
mercy
{\ADD{endureth}} for
ever:
\par }{\Q \VS{21}And
gave their
land for an
heritage: for his
mercy
{\ADD{endureth}} for
ever:
\par }{\Q \VS{22}{\ADD{Even}} an
heritage unto
Israel his
servant: for his
mercy
{\ADD{endureth}} for
ever.
\par }{\BB \par }{\Q \VS{23}Who
remembered us in our low
estate: for his
mercy
{\ADD{endureth}} for
ever:
\par }{\Q \VS{24}And hath
redeemed us from our
enemies: for his
mercy
{\ADD{endureth}} for
ever.
\par }{\Q \VS{25}Who
giveth
food to all
flesh: for his
mercy
{\ADD{endureth}} for
ever.
\par }{\Q \VS{26}O give
thanks unto the
God of
heaven: for his
mercy
{\ADD{endureth}} for
ever.

\par }\PsalmChap{137}{\Q \VerseOne{1}By the
rivers of
Babylon, there we sat
down, yea, we
wept, when we
remembered
Zion.
\par }{\Q \VS{2}We
hanged our
harps upon the
willows in the
midst thereof.
\par }{\Q \VS{3}For there they that carried us away
captive
required of us a
song; and they that
wasted us
{\ADD{required of us}}
mirth,
{\ADD{saying}},
Sing us
{\ADD{one}} of the
songs of
Zion.\FTNT{*}{{\FR 137.3 }
a song: Heb. the words of a song}\FTNT{†}{{\FR 137.3 }
wasted…: Heb. laid us on heaps}
\par }{\Q \VS{4}How shall we
sing the
{\ND{
LORD’s}}
song in a
strange
land?\FTNT{‡}{{\FR 137.4 }
strange…: Heb. land of a stranger?}
\par }{\Q \VS{5}If I
forget thee, O
Jerusalem, let my right
hand
forget
{\ADD{her cunning}}.
\par }{\Q \VS{6}If I do not
remember thee, let my
tongue
cleave to the roof of my
mouth; if I
prefer not
Jerusalem above my
chief
joy.\FTNT{§}{{\FR 137.6 }
my chief…: Heb. the head of my joy}
\par }{\BB \par }{\Q \VS{7}Remember, O
{\ND{
LORD}}, the
children of
Edom in the
day of
Jerusalem; who
said,
Rase
{\ADD{it}},
rase
{\ADD{it, even}} to the
foundation thereof.\FTNT{**}{{\FR 137.7 }
Rase it: Heb. Make bare}
\par }{\Q \VS{8}O
daughter of
Babylon, who art to be
destroyed;
happy
{\ADD{shall he be}}, that
rewardeth thee as thou hast
served us.\FTNT{††}{{\FR 137.8 }
destroyed: Heb. wasted}\FTNT{‡‡}{{\FR 137.8 }
rewardeth…: Heb. recompenseth unto thee thy deed which thou didst to us}
\par }{\Q \VS{9}Happy
{\ADD{shall he be}}, that
taketh and
dasheth thy little
ones against the
stones.\FTNT{§§}{{\FR 137.9 }
the stones: Heb. the rock}

\par }\PsalmChap{138}{\D {\ADD{A Psalm}} of
David.
\par }{\Q \VerseOne{1}I will
praise thee with my whole
heart: before the
gods will I sing
praise unto thee.
\par }{\Q \VS{2}I will
worship toward thy
holy
temple, and
praise thy
name for thy
lovingkindness and for thy
truth: for thou hast
magnified thy
word above all thy
name.
\par }{\Q \VS{3}In the
day when I
cried thou
answeredst me,
{\ADD{and}}
strengthenedst me
{\ADD{with}}
strength in my
soul.
\par }{\Q \VS{4}All the
kings of the
earth shall
praise thee, O
{\ND{
LORD}}, when they
hear the
words of thy
mouth.
\par }{\Q \VS{5}Yea, they shall
sing in the
ways of the
{\ND{
LORD}}: for
great
{\ADD{is}} the
glory of the
{\ND{
LORD}}.
\par }{\BB \par }{\Q \VS{6}Though the
{\ND{
LORD}}
{\ADD{be}}
high, yet hath he
respect unto the
lowly: but the
proud he
knoweth afar
off.
\par }{\Q \VS{7}Though I
walk in the
midst of
trouble, thou wilt
revive me: thou shalt stretch
forth thine
hand against the
wrath of mine
enemies, and thy right
hand shall
save me.
\par }{\Q \VS{8}The
{\ND{
LORD}} will
perfect
{\ADD{that which}} concerneth me: thy
mercy, O
{\ND{
LORD}},
{\ADD{endureth}} for
ever:
forsake not the
works of thine own
hands.

\par }\PsalmChap{139}{\D To the chief
Musician, A
Psalm of
David.
\par }{\Q \VerseOne{1}O
{\ND{
LORD}}, thou hast
searched me, and
known
{\ADD{me}}.
\par }{\Q \VS{2}Thou
knowest my
downsitting and mine
uprising, thou
understandest my
thought afar
off.
\par }{\Q \VS{3}Thou
compassest my
path and my lying
down, and art
acquainted
{\ADD{with}} all my
ways.\FTNT{*}{{\FR 139.3 }
compassest: or, winnowest}
\par }{\Q \VS{4}For
{\ADD{there is}} not a
word in my
tongue,
{\ADD{but}}, lo, O
{\ND{
LORD}}, thou
knowest it altogether.
\par }{\Q \VS{5}Thou hast
beset me
behind and
before, and
laid thine
hand upon me.
\par }{\Q \VS{6}{\ADD{Such}}
knowledge
{\ADD{is}} too
wonderful for me; it is
high, I
cannot
{\ADD{attain}} unto it.
\par }{\BB \par }{\Q \VS{7}Whither shall I
go from thy
spirit? or whither shall I
flee from thy
presence?
\par }{\Q \VS{8}If I ascend
up into
heaven, thou
{\ADD{art}} there: if I make my
bed in
hell, behold, thou
{\ADD{art there}}.
\par }{\Q \VS{9}{\ADD{If}} I
take the
wings of the
morning,
{\ADD{and}}
dwell in the uttermost
parts of the
sea;
\par }{\Q \VS{10}Even there shall thy
hand
lead me, and thy right
hand shall
hold me.
\par }{\Q \VS{11}If I
say, Surely the
darkness shall
cover me; even the
night shall be
light about
me.
\par }{\Q \VS{12}Yea, the
darkness
hideth not from thee; but the
night
shineth as the
day: the
darkness and the
light
{\ADD{are}} both alike
{\ADD{to thee}}.\FTNT{†}{{\FR 139.12 }
hideth…: Heb. darkeneth not}\FTNT{‡}{{\FR 139.12 }
the darkness and…: Heb. as is the darkness, so is the light}
\par }{\Q \VS{13}For thou hast
possessed my
reins: thou hast
covered me in my
mother’s
womb.
\par }{\Q \VS{14}I will
praise thee; for I am
fearfully
{\ADD{and}} wonderfully
made:
marvellous
{\ADD{are}} thy
works; and
{\ADD{that}} my
soul
knoweth right
well.\FTNT{§}{{\FR 139.14 }
right…: Heb. greatly}
\par }{\Q \VS{15}My
substance was not
hid from thee, when I was
made in
secret,
{\ADD{and}} curiously
wrought in the lowest
parts of the
earth.\FTNT{**}{{\FR 139.15 }
substance: or, strength, or, body}
\par }{\Q \VS{16}Thine
eyes did
see my substance, yet being
unperfect; and in thy
book all
{\ADD{my members}} were
written,
{\ADD{which}} in
continuance were
fashioned, when
{\ADD{as yet there was}}
none of them.\FTNT{††}{{\FR 139.16 }
all…: Heb. all of them}\FTNT{‡‡}{{\FR 139.16 }
which…: or, what days they should be fashioned}
\par }{\BB \par }{\Q \VS{17}How
precious also are thy
thoughts unto me, O
God! how
great is the
sum of them!
\par }{\Q \VS{18}{\ADD{If}} I should
count them, they are more in
number than the
sand: when I
awake, I am
still with thee.
\par }{\Q \VS{19}Surely thou wilt
slay the
wicked, O
God:
depart from me therefore, ye
bloody
men.
\par }{\Q \VS{20}For they
speak against thee
wickedly,
{\ADD{and}} thine
enemies
take
{\ADD{thy name}} in
vain.
\par }{\Q \VS{21}Do not I
hate them, O
{\ND{
LORD}}, that
hate thee? and am not I
grieved with those that rise
up against thee?
\par }{\Q \VS{22}I
hate them with
perfect
hatred: I count them mine
enemies.
\par }{\Q \VS{23}Search me, O
God, and
know my
heart:
try me, and
know my
thoughts:
\par }{\Q \VS{24}And
see if
{\ADD{there be any}}
wicked
way in me, and
lead me in the
way
everlasting.\FTNT{§§}{{\FR 139.24 }
wicked…: Heb. way of pain, or, grief}

\par }\PsalmChap{140}{\D To the chief
Musician, A
Psalm of
David.
\par }{\Q \VerseOne{1}Deliver me, O
{\ND{
LORD}}, from the
evil
man:
preserve me from the
violent
man;\FTNT{*}{{\FR 140.1 }
violent…: Heb. man of violences}
\par }{\Q \VS{2}Which
imagine
mischiefs in
{\ADD{their}}
heart;
continually are they gathered
together
{\ADD{for}}
war.
\par }{\Q \VS{3}They have
sharpened their
tongues like a
serpent;
adders’
poison
{\ADD{is}} under their
lips.
Selah.
\par }{\Q \VS{4}Keep me, O
{\ND{
LORD}}, from the
hands of the
wicked;
preserve me from the
violent
man; who have
purposed to
overthrow my
goings.
\par }{\Q \VS{5}The
proud have
hid a
snare for me, and
cords; they have
spread a
net by the
wayside; they have
set
gins for me.
Selah.
\par }{\Q \VS{6}I
said unto the
{\ND{
LORD}}, Thou
{\ADD{art}} my
God:
hear the
voice of my
supplications, O
{\ND{
LORD}}.
\par }{\Q \VS{7}O
{\ND{
GOD}} the
Lord, the
strength of my
salvation, thou hast
covered my
head in the
day of
battle.
\par }{\BB \par }{\Q \VS{8}Grant not, O
{\ND{
LORD}}, the
desires of the
wicked:
further not his wicked
device;
{\ADD{lest}} they
exalt themselves.
Selah.\FTNT{†}{{\FR 140.8 }
lest…: or, let them not be exalted}
\par }{\Q \VS{9}{\ADD{As for}} the
head of those that compass me
about, let the
mischief of their own
lips
cover them.
\par }{\Q \VS{10}Let burning
coals
fall upon them: let them be
cast into the
fire; into deep
pits, that they rise not up
again.
\par }{\Q \VS{11}Let not an evil
speaker be
established in the
earth:
evil shall
hunt the
violent
man to
overthrow
{\ADD{him}}.\FTNT{‡}{{\FR 140.11 }
an…: or, an evil speaker (Heb. a man of tongue), a wicked man of violence, be established in the earth: let him be hunted to his overthrow}
\par }{\Q \VS{12}I
know that the
{\ND{
LORD}} will
maintain the
cause of the
afflicted,
{\ADD{and}} the
right of the
poor.
\par }{\Q \VS{13}Surely the
righteous shall give
thanks unto thy
name: the
upright shall
dwell in thy
presence.

\par }\PsalmChap{141}{\D A
Psalm of
David.
\par }{\Q \VerseOne{1}{\ND{
LORD}}, I
cry unto thee: make
haste unto me; give
ear unto my
voice, when I
cry unto thee.
\par }{\Q \VS{2}Let my
prayer be set
forth
before thee
{\ADD{as}}
incense;
{\ADD{and}} the lifting
up of my
hands
{\ADD{as}} the
evening
sacrifice.\FTNT{*}{{\FR 141.2 }
set…: Heb. directed}
\par }{\Q \VS{3}Set a
watch, O
{\ND{
LORD}}, before my
mouth;
keep the
door of my
lips.
\par }{\Q \VS{4}Incline not my
heart to
{\ADD{any}}
evil
thing, to
practise
wicked
works with
men that
work
iniquity: and let me not
eat of their
dainties.
\par }{\BB \par }{\Q \VS{5}Let the
righteous
smite me;
{\ADD{it shall be}} a
kindness: and let him
reprove me;
{\ADD{it shall be}} an
excellent
oil,
{\ADD{which}} shall not
break my
head: for yet my
prayer also
{\ADD{shall be}} in their
calamities.\FTNT{†}{{\FR 141.5 }
me; it shall be a…: or, me kindly, and reprove me; let not their precious oil break, etc}
\par }{\Q \VS{6}When their
judges are
overthrown in
stony
places, they shall
hear my
words; for they are
sweet.
\par }{\Q \VS{7}Our
bones are
scattered at the
grave’s
mouth, as when one
cutteth and
cleaveth
{\ADD{wood}} upon the
earth.
\par }{\Q \VS{8}But mine
eyes
{\ADD{are}} unto thee, O
{\ND{
GOD}} the
Lord: in thee is my
trust;
leave not my
soul
destitute.\FTNT{‡}{{\FR 141.8 }
leave…: Heb. make not my soul bare}
\par }{\Q \VS{9}Keep me from the
snares
{\ADD{which}} they have
laid for me, and the
gins of the
workers of
iniquity.
\par }{\Q \VS{10}Let the
wicked
fall into their own
nets, whilst that I
withal
escape.\FTNT{§}{{\FR 141.10 }
escape: Heb. pass over}

\par }\PsalmChap{142}{\D Maschil of
David; A
Prayer when he was in the
cave.
\par }{\Q \VerseOne{1}I
cried unto the
{\ND{
LORD}} with my
voice; with my
voice unto the
{\ND{
LORD}} did I make my
supplication.\FTNT{*}{{\FR 142.1 }
Maschil…: or, A Psalm of David, giving instruction}
\par }{\Q \VS{2}I poured
out my
complaint
before him; I
shewed
before him my
trouble.
\par }{\Q \VS{3}When my
spirit was
overwhelmed within me, then thou
knewest my
path. In the
way
wherein I
walked have they privily
laid a
snare for me.
\par }{\BB \par }{\Q \VS{4}I
looked on
{\ADD{my}} right
hand, and
beheld, but
{\ADD{there was}} no man that would
know me:
refuge
failed me; no man
cared for my
soul.\FTNT{†}{{\FR 142.4 }
I looked…: or, Look on the right hand, and see}\FTNT{‡}{{\FR 142.4 }
failed…: Heb. perished from me}\FTNT{§}{{\FR 142.4 }
cared…: Heb. sought after}
\par }{\Q \VS{5}I
cried unto thee, O
{\ND{
LORD}}: I
said, Thou
{\ADD{art}} my
refuge
{\ADD{and}} my
portion in the
land of the
living.
\par }{\Q \VS{6}Attend unto my
cry; for I am
brought
very
low:
deliver me from my
persecutors; for they are
stronger than I.
\par }{\Q \VS{7}Bring my
soul out of
prison, that I may
praise thy
name: the
righteous shall compass me
about; for thou shalt deal
bountifully with me.

\par }\PsalmChap{143}{\D A
Psalm of
David.
\par }{\Q \VerseOne{1}Hear my
prayer, O
{\ND{
LORD}}, give
ear to my
supplications: in thy
faithfulness
answer me,
{\ADD{and}} in thy
righteousness.
\par }{\Q \VS{2}And
enter not into
judgment with thy
servant: for in thy
sight shall no man
living be
justified.
\par }{\Q \VS{3}For the
enemy hath
persecuted my
soul; he hath
smitten my
life
down to the
ground; he hath made me to
dwell in
darkness, as those that have been
long
dead.
\par }{\Q \VS{4}Therefore is my
spirit
overwhelmed within me; my
heart
within me is
desolate.
\par }{\Q \VS{5}I
remember the
days of
old; I
meditate on all thy
works; I
muse on the
work of thy
hands.
\par }{\Q \VS{6}I stretch
forth my
hands unto thee: my
soul
{\ADD{thirsteth}} after thee, as a
thirsty
land.
Selah.
\par }{\BB \par }{\Q \VS{7}Hear me
speedily, O
{\ND{
LORD}}: my
spirit
faileth:
hide not thy
face from me, lest I be
like unto them that go
down into the
pit.\FTNT{*}{{\FR 143.7 }
lest…: or, for I am become like, etc}
\par }{\Q \VS{8}Cause me to
hear thy
lovingkindness in the
morning; for in thee do I
trust: cause me to
know the
way
wherein I should
walk; for I lift
up my
soul unto thee.
\par }{\Q \VS{9}Deliver me, O
{\ND{
LORD}}, from mine
enemies: I flee unto thee to
hide me.\FTNT{†}{{\FR 143.9 }
flee…: Heb. hide me with thee}
\par }{\Q \VS{10}Teach me to
do thy
will; for thou
{\ADD{art}} my
God: thy
spirit
{\ADD{is}}
good;
lead me into the
land of
uprightness.
\par }{\Q \VS{11}Quicken me, O
{\ND{
LORD}}, for thy
name’s sake: for thy
righteousness’ sake
bring my
soul out of
trouble.
\par }{\Q \VS{12}And of thy
mercy cut
off mine
enemies, and
destroy all them that
afflict my
soul: for I
{\ADD{am}} thy
servant.

\par }\PsalmChap{144}{\D {\ADD{A Psalm}} of
David.
\par }{\Q \VerseOne{1}Blessed
{\ADD{be}} the
{\ND{
LORD}} my
strength, which
teacheth my
hands to
war,
{\ADD{and}} my
fingers to
fight:\FTNT{*}{{\FR 144.1 }
strength: Heb. rock}\FTNT{†}{{\FR 144.1 }
to war…: Heb. to the war, etc}
\par }{\Q \VS{2}My
goodness, and my
fortress; my high
tower, and my
deliverer; my
shield, and
{\ADD{he}} in whom I
trust; who
subdueth my
people under me.\FTNT{‡}{{\FR 144.2 }
My goodness: or, My mercy}
\par }{\Q \VS{3}{\ND{
LORD}}, what
{\ADD{is}}
man, that thou takest
knowledge of him!
{\ADD{or}} the
son of
man, that thou makest
account of him!
\par }{\Q \VS{4}Man is
like to
vanity: his
days
{\ADD{are}} as a
shadow that passeth
away.
\par }{\Q \VS{5}Bow thy
heavens, O
{\ND{
LORD}}, and come
down:
touch the
mountains, and they shall
smoke.
\par }{\Q \VS{6}Cast
forth
lightning, and
scatter them: shoot
out thine
arrows, and
destroy them.
\par }{\Q \VS{7}Send thine
hand from
above;
rid me, and
deliver me out of
great
waters, from the
hand of
strange
children;\FTNT{§}{{\FR 144.7 }
hand from: Heb. hands from}
\par }{\Q \VS{8}Whose
mouth
speaketh
vanity, and their right
hand
{\ADD{is}} a right
hand of
falsehood.
\par }{\BB \par }{\Q \VS{9}I will
sing a
new
song unto thee, O
God: upon a
psaltery
{\ADD{and}} an instrument of ten
strings will I sing
praises unto thee.
\par }{\Q \VS{10}{\ADD{It is he}} that
giveth
salvation unto
kings: who
delivereth
David his
servant from the
hurtful
sword.\FTNT{**}{{\FR 144.10 }
salvation: or, victory}
\par }{\Q \VS{11}Rid me, and
deliver me from the
hand of
strange
children, whose
mouth
speaketh
vanity, and their right
hand
{\ADD{is}} a right
hand of
falsehood:
\par }{\Q \VS{12}That our
sons
{\ADD{may be}} as
plants grown
up in their
youth;
{\ADD{that}} our
daughters
{\ADD{may be}} as corner
stones,
polished
{\ADD{after}} the
similitude of a
palace:\FTNT{††}{{\FR 144.12 }
polished: Heb. cut}
\par }{\Q \VS{13}{\ADD{That}} our
garners
{\ADD{may be}}
full,
affording all manner of
store:
{\ADD{that}} our
sheep may bring forth
thousands and ten
thousands in our
streets:\FTNT{‡‡}{{\FR 144.13 }
all…: Heb. from kind to kind}
\par }{\Q \VS{14}{\ADD{That}} our
oxen
{\ADD{may be}} strong to
labour;
{\ADD{that there be}} no breaking
in, nor going
out; that
{\ADD{there be}} no
complaining in our
streets.\FTNT{§§}{{\FR 144.14 }
strong…: Heb. able to bear burdens, or, loaden with flesh}
\par }{\Q \VS{15}Happy
{\ADD{is that}}
people, that is in such a
case:
{\ADD{yea}},
happy
{\ADD{is that}}
people, whose
God
{\ADD{is}} the
{\ND{
LORD}}.

\par }\PsalmChap{145}{\D David’s
{\ADD{Psalm}} of
praise.
\par }{\Q \VerseOne{1}I will
extol thee, my
God, O
king; and I will
bless thy
name for
ever and
ever.
\par }{\Q \VS{2}Every
day will I
bless thee; and I will
praise thy
name for
ever and
ever.
\par }{\Q \VS{3}Great
{\ADD{is}} the
{\ND{
LORD}}, and
greatly to be
praised; and his
greatness
{\ADD{is}}
unsearchable.\FTNT{*}{{\FR 145.3 }
and his…: Heb. and of his greatness there is no search}
\par }{\Q \VS{4}One
generation shall
praise thy
works to another, and shall
declare thy mighty
acts.
\par }{\Q \VS{5}I will
speak of the
glorious
honour of thy
majesty, and of thy
wondrous
works.\FTNT{†}{{\FR 145.5 }
works: Heb. things, or, words}
\par }{\Q \VS{6}And
{\ADD{men}} shall
speak of the
might of thy terrible
acts: and I will
declare thy
greatness.\FTNT{‡}{{\FR 145.6 }
declare: Heb. declare it}
\par }{\Q \VS{7}They shall abundantly
utter the
memory of thy
great
goodness, and shall
sing of thy
righteousness.
\par }{\Q \VS{8}The
{\ND{
LORD}}
{\ADD{is}}
gracious, and full of
compassion;
slow to
anger, and of
great
mercy.\FTNT{§}{{\FR 145.8 }
of great…: Heb. great in mercy}
\par }{\Q \VS{9}The
{\ND{
LORD}}
{\ADD{is}}
good to all: and his tender
mercies
{\ADD{are}} over all his
works.
\par }{\BB \par }{\Q \VS{10}All thy
works shall
praise thee, O
{\ND{
LORD}}; and thy
saints shall
bless thee.
\par }{\Q \VS{11}They shall
speak of the
glory of thy
kingdom, and
talk of thy
power;
\par }{\Q \VS{12}To make
known to the
sons of
men his mighty
acts, and the
glorious
majesty of his
kingdom.
\par }{\Q \VS{13}Thy
kingdom
{\ADD{is}} an
everlasting
kingdom, and thy
dominion
{\ADD{endureth}} throughout
all
generations.\FTNT{**}{{\FR 145.13 }
an…: Heb. a kingdom of all ages}
\par }{\Q \VS{14}The
{\ND{
LORD}}
upholdeth all that
fall, and raiseth
up all
{\ADD{those that be}} bowed
down.
\par }{\Q \VS{15}The
eyes of all
wait upon thee; and thou
givest them their
meat in due
season.\FTNT{††}{{\FR 145.15 }
wait…: or, look unto}
\par }{\Q \VS{16}Thou
openest thine
hand, and
satisfiest the
desire of every living
thing.
\par }{\Q \VS{17}The
{\ND{
LORD}}
{\ADD{is}}
righteous in all his
ways, and
holy in all his
works.\FTNT{‡‡}{{\FR 145.17 }
holy: or, merciful, or, bountiful}
\par }{\Q \VS{18}The
{\ND{
LORD}}
{\ADD{is}}
nigh unto all them that
call upon him, to all that
call upon him in
truth.
\par }{\Q \VS{19}He will
fulfil the
desire of them that
fear him: he also will
hear their
cry, and will
save them.
\par }{\Q \VS{20}The
{\ND{
LORD}}
preserveth all them that
love him: but all the
wicked will he
destroy.
\par }{\Q \VS{21}My
mouth shall
speak the
praise of the
{\ND{
LORD}}: and let all
flesh
bless his
holy
name for
ever and
ever.

\par }\PsalmChap{146}{\Q \VerseOne{1}Praise ye the
{\ND{
LORD}}.
Praise the
{\ND{
LORD}}, O my
soul.\FTNT{*}{{\FR 146.1 }
Praise ye…: Heb. Hallelujah}
\par }{\Q \VS{2}While I
live will I
praise the
{\ND{
LORD}}: I will sing
praises unto my
God while I have any being.
\par }{\Q \VS{3}Put not your
trust in
princes,
{\ADD{nor}} in the
son of
man, in whom
{\ADD{there is}} no
help.\FTNT{†}{{\FR 146.3 }
help: or, salvation}
\par }{\Q \VS{4}His
breath goeth
forth, he
returneth to his
earth; in that very
day his
thoughts
perish.
\par }{\BB \par }{\Q \VS{5}Happy
{\ADD{is he}} that
{\ADD{hath}} the
God of
Jacob for his
help, whose
hope
{\ADD{is}} in the
{\ND{
LORD}} his
God:
\par }{\Q \VS{6}Which
made
heaven, and
earth, the
sea, and all that therein
{\ADD{is:}} which
keepeth
truth for
ever:
\par }{\Q \VS{7}Which
executeth
judgment for the
oppressed: which
giveth
food to the
hungry. The
{\ND{
LORD}}
looseth the
prisoners:
\par }{\Q \VS{8}The
{\ND{
LORD}}
openeth
{\ADD{the eyes of}} the
blind: the
{\ND{
LORD}}
raiseth them that are bowed
down: the
{\ND{
LORD}}
loveth the
righteous:
\par }{\Q \VS{9}The
{\ND{
LORD}}
preserveth the
strangers; he
relieveth the
fatherless and
widow: but the
way of the
wicked he turneth upside
down.
\par }{\Q \VS{10}The
{\ND{
LORD}} shall
reign for
ever,
{\ADD{even}} thy
God, O
Zion, unto
all
generations.
Praise ye the
{\ND{
LORD}}.

\par }\PsalmChap{147}{\Q \VerseOne{1}Praise ye the
{\ND{
LORD}}: for
{\ADD{it is}}
good to sing
praises unto our
God; for
{\ADD{it is}}
pleasant;
{\ADD{and}}
praise is
comely.
\par }{\Q \VS{2}The
{\ND{
LORD}} doth build
up
Jerusalem: he gathereth
together the
outcasts of
Israel.
\par }{\Q \VS{3}He
healeth the
broken in
heart, and bindeth
up their
wounds.\FTNT{*}{{\FR 147.3 }
wounds: Heb. griefs}
\par }{\Q \VS{4}He
telleth the
number of the
stars; he
calleth them all by
{\ADD{their}}
names.
\par }{\Q \VS{5}Great
{\ADD{is}} our
Lord, and of
great
power: his
understanding
{\ADD{is}}
infinite.\FTNT{†}{{\FR 147.5 }
his…: Heb. of his understanding there is no number}
\par }{\Q \VS{6}The
{\ND{
LORD}} lifteth
up the
meek: he
casteth the
wicked
down to the
ground.
\par }{\Q \VS{7}Sing unto the
{\ND{
LORD}} with
thanksgiving; sing
praise upon the
harp unto our
God:
\par }{\Q \VS{8}Who
covereth the
heaven with
clouds, who
prepareth
rain for the
earth, who maketh
grass to
grow upon the
mountains.
\par }{\Q \VS{9}He
giveth to the
beast his
food,
{\ADD{and}} to the
young
ravens which
cry.
\par }{\Q \VS{10}He
delighteth not in the
strength of the
horse: he taketh not
pleasure in the
legs of a
man.
\par }{\Q \VS{11}The
{\ND{
LORD}} taketh
pleasure in them that
fear him, in those that
hope in his
mercy.
\par }{\BB \par }{\Q \VS{12}Praise the
{\ND{
LORD}}, O
Jerusalem;
praise thy
God, O
Zion.
\par }{\Q \VS{13}For he hath
strengthened the
bars of thy
gates; he hath
blessed thy
children
within thee.
\par }{\Q \VS{14}He
maketh
peace
{\ADD{in}} thy
borders,
{\ADD{and}}
filleth thee with the
finest of the
wheat.\FTNT{‡}{{\FR 147.14 }
He…: Heb. Who maketh thy border peace}\FTNT{§}{{\FR 147.14 }
finest…: Heb. fat of wheat}
\par }{\Q \VS{15}He sendeth
forth his
commandment
{\ADD{upon}}
earth: his
word
runneth very
swiftly.
\par }{\Q \VS{16}He
giveth
snow like
wool: he
scattereth the
hoarfrost like
ashes.
\par }{\Q \VS{17}He casteth
forth his
ice like
morsels: who can
stand
before his
cold?
\par }{\Q \VS{18}He sendeth
out his
word, and
melteth them: he causeth his
wind to
blow,
{\ADD{and}} the
waters
flow.
\par }{\Q \VS{19}He
sheweth his
word unto
Jacob, his
statutes and his
judgments unto
Israel.\FTNT{**}{{\FR 147.19 }
his word: Heb. his words}
\par }{\Q \VS{20}He hath not
dealt so with any
nation: and
{\ADD{as for his}}
judgments, they have not
known them.
Praise ye the
{\ND{
LORD}}.

\par }\PsalmChap{148}{\Q \VerseOne{1}Praise ye the
{\ND{
LORD}}.
Praise ye the
{\ND{
LORD}} from the
heavens:
praise him in the
heights.\FTNT{*}{{\FR 148.1 }
Praise ye the
{\ND{LORD}}. Praise: Heb. Hallelujah, etc}
\par }{\Q \VS{2}Praise ye him, all his
angels:
praise ye him, all his
hosts.
\par }{\Q \VS{3}Praise ye him,
sun and
moon:
praise him, all ye
stars of
light.
\par }{\Q \VS{4}Praise him, ye
heavens of
heavens, and ye
waters that
{\ADD{be}} above the
heavens.
\par }{\Q \VS{5}Let them
praise the
name of the
{\ND{
LORD}}: for he
commanded, and they were
created.
\par }{\Q \VS{6}He hath also
stablished them for
ever and
ever: he hath
made a
decree which shall not
pass.
\par }{\BB \par }{\Q \VS{7}Praise the
{\ND{
LORD}} from the
earth, ye
dragons, and all
deeps:
\par }{\Q \VS{8}Fire, and
hail;
snow, and
vapour;
stormy
wind
fulfilling his
word:
\par }{\Q \VS{9}Mountains, and all
hills;
fruitful
trees, and all
cedars:
\par }{\Q \VS{10}Beasts, and all
cattle; creeping
things, and
flying
fowl:\FTNT{†}{{\FR 148.10 }
flying…: Heb. birds of wing}
\par }{\Q \VS{11}Kings of the
earth, and all
people;
princes, and all
judges of the
earth:
\par }{\Q \VS{12}Both young
men, and
maidens; old
men, and
children:
\par }{\Q \VS{13}Let them
praise the
name of the
{\ND{
LORD}}: for his
name alone is
excellent; his
glory
{\ADD{is}} above the
earth and
heaven.\FTNT{‡}{{\FR 148.13 }
excellent: Heb. exalted}
\par }{\Q \VS{14}He also
exalteth the
horn of his
people, the
praise of all his
saints;
{\ADD{even}} of the
children of
Israel, a
people
near unto him.
Praise ye the
{\ND{
LORD}}.

\par }\PsalmChap{149}{\Q \VerseOne{1}Praise ye the
{\ND{
LORD}}.
Sing unto the
{\ND{
LORD}} a
new
song,
{\ADD{and}} his
praise in the
congregation of
saints.\FTNT{*}{{\FR 149.1 }
Praise…: Heb. Hallelujah}
\par }{\Q \VS{2}Let
Israel
rejoice in him that
made him: let the
children of
Zion be
joyful in their
King.\FTNT{†}{{\FR 149.2 }
in him…: Heb. in his Makers}
\par }{\Q \VS{3}Let them
praise his
name in the
dance: let them sing
praises unto him with the
timbrel and
harp.\FTNT{‡}{{\FR 149.3 }
in…: or, with the pipe}
\par }{\Q \VS{4}For the
{\ND{
LORD}} taketh
pleasure in his
people: he will
beautify the
meek with
salvation.
\par }{\Q \VS{5}Let the
saints be
joyful in
glory: let them sing
aloud upon their
beds.
\par }{\BB \par }{\Q \VS{6}{\ADD{Let}} the
high
{\ADD{praises}} of
God
{\ADD{be}} in their
mouth, and a
twoedged
sword in their
hand;\FTNT{§}{{\FR 149.6 }
mouth: Heb. throat}
\par }{\Q \VS{7}To
execute
vengeance upon the
heathen,
{\ADD{and}}
punishments upon the
people;
\par }{\Q \VS{8}To
bind their
kings with
chains, and their
nobles with
fetters of
iron;
\par }{\Q \VS{9}To
execute upon them the
judgment
written: this
honour have all his
saints.
Praise ye the
{\ND{
LORD}}.

\par }\PsalmChap{150}{\Q \VerseOne{1}Praise ye the
{\ND{
LORD}}.
Praise
God in his
sanctuary:
praise him in the
firmament of his
power.\FTNT{*}{{\FR 150.1 }
Praise ye…: Heb. Hallelujah}
\par }{\Q \VS{2}Praise him for his mighty
acts:
praise him according to his
excellent
greatness.
\par }{\Q \VS{3}Praise him with the
sound of the
trumpet:
praise him with the
psaltery and
harp.\FTNT{†}{{\FR 150.3 }
trumpet: or, cornet}
\par }{\Q \VS{4}Praise him with the
timbrel and
dance:
praise him with stringed
instruments and
organs.\FTNT{‡}{{\FR 150.4 }
dance: or, pipe}
\par }{\Q \VS{5}Praise him upon the
loud
cymbals:
praise him upon the high
sounding
cymbals.
\par }{\Q \VS{6}Let every thing that hath
breath
praise the
{\ND{
LORD}}.
Praise ye the
{\ND{
LORD}}.
\par }