\NormalFont\ShortTitle{Revelation}
{\MT THE REVELATION OF ST. JOHN THE DIVINE

\par }\ChapOne{1}{\PP \VerseOne{1}The
Revelation of
Jesus
Christ,
which
God
gave unto
him, to
shew
unto
his
servants things
which
must
shortly come to
pass;
and
he sent
and
signified
{\ADD{it}}
by
his
angel
unto
his
servant
John:
\VS{2}Who bare
record of the
word of
God,
and of the
testimony of
Jesus
Christ,
and of all things
that he
saw.
\VS{3}Blessed
{\ADD{is}} he that
readeth,
and they that
hear the
words of this
prophecy,
and
keep those things which are
written
therein:
for the
time
{\ADD{is}} at
hand.
\VS{4}John to
the
seven
churches which are
in
Asia:
Grace
{\ADD{be}} unto
you,
and
peace,
from him which
is,
and which
was,
and which is to
come;
and
from
the
seven
Spirits
which
are
before
his
throne;
\VS{5}And
from
Jesus
Christ,
{\ADD{who is}}
the
faithful
witness,
{\ADD{and}} the first
begotten
of the
dead,
and the
prince of the
kings of the
earth. Unto him that
loved
us,
and
washed
us
from
our
sins
in his
own
blood,
\VS{6}And hath
made
us
kings
and
priests unto
God
and
his
Father; to
him
{\ADD{be}}
glory
and
dominion
for
ever and
ever.
Amen.
\VS{7}Behold, he
cometh
with
clouds;
and
every
eye shall
see
him,
and
they
{\ADD{also}}
which
pierced
him:
and
all
kindreds of the
earth shall
wail because
of
him. Even
so,
Amen.
\VS{8}{\WJ{I
am
Alpha
and
Omega,
the beginning
and
the ending,}}
saith the
Lord, which
is,
and which
was,
and which is to
come, the
Almighty.
\par }{\PP \VS{9}I
John, who
also am
your
brother,
and
companion
in
tribulation,
and
in the
kingdom
and
patience of
Jesus
Christ,
was
in the
isle that is
called
Patmos,
for the
word of
God,
and
for the
testimony of
Jesus
Christ.
\VS{10}I
was
in the
Spirit
on
the
Lord’s
day,
and
heard
behind
me
a
great
voice,
as of a
trumpet,
\VS{11}Saying,
{\WJ{I
am
Alpha
and
Omega,
the first
and
the last:}}
and,
{\WJ{What
thou seest,
write
in
a book,
and
send
{\ADD{it}}
unto the
seven
churches
which are
in
Asia;
unto
Ephesus,
and
unto
Smyrna,
and
unto
Pergamos,
and
unto
Thyatira,
and
unto
Sardis,
and
unto
Philadelphia,
and
unto
Laodicea.}}
\VS{12}And I
turned to
see the
voice
that
spake
with
me.
And being
turned, I
saw
seven golden
candlesticks;
\VS{13}And
in the
midst of
the
seven
candlesticks
{\ADD{one}}
like unto the
Son of
man,
clothed with a garment down to the
foot,
and
girt
about the
paps with
a
golden
girdle.
\VS{14}His
head
and
{\ADD{his}}
hairs
{\ADD{were}}
white
like
wool,
as
white as
snow;
and
his
eyes
{\ADD{were}}
as a
flame of
fire;
\VS{15}And
his
feet
like unto fine
brass, as
if they
burned
in a
furnace;
and
his
voice
as the
sound
of
many
waters.
\VS{16}And he
had
in
his right
hand
seven
stars:
and
out
of
his
mouth
went
a
sharp
twoedged
sword:
and
his
countenance
{\ADD{was}}
as the
sun
shineth
in
his
strength.
\VS{17}And
when I
saw
him, I
fell
at
his
feet
as
dead.
And he
laid
his
right
hand
upon
me,
saying unto
me,
{\WJ{Fear
not;
I
am
the first
and
the last:}}
\VS{18}{\WJ{I
{\ADD{am}}
he that liveth,
and
was
dead;
and,
behold,
I am
alive
for
evermore,
Amen;
and
have
the keys
of hell
and
of death.}}
\par }{\PP \VS{19}{\WJ{Write
the things which
thou hast seen,
and
the things which
are,
and
the things which
shall
be
hereafter;}}
\VS{20}{\WJ{The mystery
of the
seven
stars
which
thou sawest
in
my
right hand,
and
the
seven
golden
candlesticks.
The seven
stars
are
the angels
of the
seven
churches:
and
the
seven
candlesticks
which
thou sawest
are
the
seven
churches.}}

\par }\Chap{2}{\PP \VerseOne{1}{\WJ{Unto the angel
of the church
of Ephesus
write;
These things
saith
he that holdeth
the
seven
stars
in
his
right hand,
who walketh
in
the midst
of the
seven
golden
candlesticks;}}
\VS{2}{\WJ{I know
thy
works,
and
thy
labour,
and
thy
patience,
and
how
thou canst
not
bear
them which are evil:
and
thou hast tried
them which say
they are
apostles,
and
are
not,
and
hast found
them
liars:}}
\VS{3}{\WJ{And
hast borne,
and
hast
patience,
and
for
my
name’s
sake
hast laboured,
and
hast
not
fainted.}}
\VS{4}{\WJ{Nevertheless
I have
{\ADD{somewhat}}
against
thee,
because
thou hast left
thy
first
love.}}
\VS{5}{\WJ{Remember
therefore
from whence
thou art fallen,
and
repent,
and
do
the
first
works;
or else
I will come
unto thee
quickly,
and
will remove
thy
candlestick
out
of
his
place,
except
thou repent.}}
\VS{6}{\WJ{But
this
thou hast,
that
thou hatest
the deeds
of the Nicolaitans,
which
I also
hate.}}
\VS{7}{\WJ{He that hath
an ear,
let him hear
what
the Spirit
saith
unto the churches;
To
him
that overcometh
will I give
to eat
of
the tree
of life,
which
is
in
the midst
of the paradise
of God.}}
\par }{\PP \VS{8}{\WJ{And
unto the angel
of the church
in Smyrna
write;
These things
saith
the first
and
the last,
which
was
dead,
and
is alive;}}
\VS{9}{\WJ{I know
thy
works,
and
tribulation,
and
poverty, (but
thou art
rich)
and
{\ADD{I know}}
the blasphemy
of them which say
they
are
Jews,
and
are
not,
but
{\ADD{are}}
the synagogue
of Satan.}}
\VS{10}{\WJ{Fear
none
of those things which
thou shalt
suffer:
behold,
the devil
shall
cast
{\ADD{some}}
of
you
into
prison,
that
ye may be tried;
and
ye shall have
tribulation
ten
days:
be thou
faithful
unto
death,
and
I will give
thee
a crown
of life.}}
\VS{11}{\WJ{He that hath
an ear,
let him hear
what
the Spirit
saith
unto the churches;
He that overcometh
shall
not
be hurt
of
the second
death.}}
\par }{\PP \VS{12}{\WJ{And
to the angel
of the church
in
Pergamos
write;
These things
saith
he which hath
the
sharp
sword
with two edges;}}
\VS{13}{\WJ{I know
thy
works,
and
where
thou dwellest,
{\ADD{even}}
where
Satan’s
seat
{\ADD{is}}:
and
thou holdest fast
my
name,
and
hast
not
denied
my
faith,
even
in
those days
wherein
Antipas
{\ADD{was}}
my
faithful
martyr,
who
was slain
among
you,
where
Satan
dwelleth.}}
\VS{14}{\WJ{But
I have
a few things
against
thee,
because
thou hast
there
them that hold
the doctrine
of Balaam,
who
taught
Balac
to cast
a stumblingblock
before
the children
of Israel,
to eat
things sacrificed unto idols,
and
to commit fornication.}}
\VS{15}{\WJ{So
hast
thou
also
them that hold
the doctrine
of the Nicolaitans,
which thing
I hate.}}
\VS{16}{\WJ{Repent;
or
else
I will come
unto thee
quickly,
and
will fight
against
them
with
the sword
of
my
mouth.}}
\VS{17}{\WJ{He that hath
an ear,
let him hear
what
the Spirit
saith
unto the churches;
To him
that overcometh
will I give
to eat
of
the
hidden
manna,
and
will give
him a
white
stone,
and
in
the stone
a
new
name
written,
which
no man
knoweth
saving
he that receiveth
{\ADD{it.}}}}
\par }{\PP \VS{18}{\WJ{And
unto the angel
of the church
in
Thyatira
write;
These things
saith
the Son
of God,
who hath
his
eyes
like unto
a flame
of fire,
and
his
feet
{\ADD{are}}
like
fine brass;}}
\VS{19}{\WJ{I know
thy
works,
and
charity,
and
service,
and
faith,
and
thy
patience,
and
thy
works;
and
the last
{\ADD{to be}}
more
than the first.}}
\VS{20}{\WJ{Notwithstanding
I have
a few things
against
thee,
because
thou sufferest
that woman
Jezebel,
which calleth
herself
a prophetess,
to teach
and
to seduce
my
servants
to commit fornication,
and
to eat
things sacrificed unto idols.}}
\VS{21}{\WJ{And
I gave
her
space
to
repent
of
her
fornication;
and
she repented
not.}}
\VS{22}{\WJ{Behold,
I
will cast
her
into
a bed,
and
them that commit adultery
with
her
into
great
tribulation,
except
they repent
of
their
deeds.}}
\VS{23}{\WJ{And
I will kill
her
children
with
death;
and
all
the churches
shall know
that
I
am
he which searcheth
the reins
and
hearts:
and
I will give
unto
every one of
you
according to
your
works.}}
\VS{24}{\WJ{But
unto you
I say,
and
unto the rest
in
Thyatira,
as many as
have
not
this
doctrine,
and
which
have
not
known
the depths
of Satan,
as
they speak;
I will put
upon
you
none
other
burden.}}
\VS{25}{\WJ{But
that which
ye have
{\ADD{already}}
hold fast
till
I come.}}
\VS{26}{\WJ{And
he that overcometh,
and
keepeth
my
works
unto
the end,
to him
will I give
power
over
the nations:}}
\VS{27}{\WJ{And
he shall rule
them
with
a rod
of iron;
as
the vessels
of a potter
shall they be broken to shivers:
even as
I
received
of
my
Father.}}
\VS{28}{\WJ{And
I will give
him
the
morning
star.}}
\VS{29}{\WJ{He that hath
an ear,
let him hear
what
the Spirit
saith
unto the churches.}}

\par }\Chap{3}{\PP \VerseOne{1}{\WJ{And
unto the angel
of the church
in
Sardis
write;
These things
saith
he that hath
the
seven
Spirits
of God,
and
the
seven
stars;
I know
thy
works,
that
thou hast
a name
that
thou livest,
and
art
dead.}}
\VS{2}{\WJ{Be
watchful,
and
strengthen
the things which remain,
that
are ready to
die:
for
I have
not
found
thy
works
perfect
before
God.}}
\VS{3}{\WJ{Remember
therefore
how
thou hast received
and
heard,
and
hold fast,
and
repent.
If
therefore
thou shalt
not
watch,
I will come
on
thee
as
a thief,
and thou shalt
not
know
what
hour
I will come
upon
thee.}}
\VS{4}{\WJ{Thou hast
a few
names
even
in
Sardis
which
have
not
defiled
their
garments;
and
they shall walk
with
me
in
white:
for
they are
worthy.}}
\VS{5}{\WJ{He that overcometh,
the same
shall be clothed
in
white
raiment;
and
I will
not
blot out
his
name
out of
the book
of life,
but
I will confess
his
name
before
my
Father,
and
before
his
angels.}}
\VS{6}{\WJ{He that hath
an ear,
let him hear
what
the Spirit
saith
unto the churches.}}
\par }{\PP \VS{7}{\WJ{And
to the angel
of the church
in
Philadelphia
write;
These things
saith
he that is holy,
he that is true,
he that hath
the key
of David,
he that openeth,
and
no man
shutteth;
and
shutteth,
and
no man
openeth;}}
\VS{8}{\WJ{I know
thy
works:
behold,
I have set
before
thee
an
open
door,
and
no man
can
shut
it:
for
thou hast
a little
strength,
and
hast kept
my
word,
and
hast
not
denied
my
name.}}
\VS{9}{\WJ{Behold,
I will make
them
of
the synagogue
of Satan,
which say
they
are
Jews,
and
are
not,
but
do lie;
behold,
I will make
them
to
come
and
worship
before
thy
feet,
and to
know
that
I
have loved
thee.}}
\VS{10}{\WJ{Because
thou hast kept
the word
of
my
patience,
I also
will keep
thee
from
the hour
of temptation,
which shall
come
upon
all
the world,
to try
them that dwell
upon
the earth.}}
\VS{11}{\WJ{Behold,
I come
quickly:
hold
that
fast
which
thou hast,
that
no man
take
thy
crown.}}
\VS{12}{\WJ{Him
that overcometh
will I make
a pillar
in
the temple
of
my
God,
and
he shall go
no
more
out:
and
I will write
upon
him
the name
of
my
God,
and
the name
of the city
of
my
God,
{\ADD{which is}}
new
Jerusalem,
which
cometh down
out of
heaven
from
my
God:
and
{\ADD{I will write upon him}}
my
new
name.}}
\VS{13}{\WJ{He that hath
an ear,
let him hear
what
the Spirit
saith
unto the churches.}}
\par }{\PP \VS{14}{\WJ{And
unto the angel
of the church
of the Laodiceans
write;
These things
saith
the Amen,
the
faithful
and
true
witness,
the beginning
of the creation
of God;}}
\VS{15}{\WJ{I know
thy
works,
that
thou art
neither
cold
nor
hot:
I would
thou wert
cold
or
hot.}}
\VS{16}{\WJ{So then
because
thou art
lukewarm,
and
neither
cold
nor
hot,
I will
spue
thee
out of
my
mouth.}}
\VS{17}{\WJ{Because
thou sayest,
I am
rich,
and
increased with goods,
and
have
need
of nothing;
and
knowest
not
that
thou
art
wretched,
and
miserable,
and
poor,
and
blind,
and
naked:}}
\VS{18}{\WJ{I counsel
thee
to buy
of
me
gold
tried
in
the fire,
that
thou mayest be rich;
and
white
raiment,
that
thou mayest be clothed,
and
{\ADD{that}}
the shame
of
thy
nakedness
do not
appear;
and
anoint
thine
eyes
with eyesalve,
that
thou mayest see.}}
\VS{19}{\WJ{As many as
I
love,
I
rebuke
and
chasten:
be zealous
therefore,
and
repent.}}
\VS{20}{\WJ{Behold,
I stand
at
the door,
and
knock:
if
any man
hear
my
voice,
and
open
the door,
I will come in
to
him,
and
will sup
with
him,
and
he
with
me.}}
\VS{21}{\WJ{To him
that overcometh
will I grant
to sit
with
me
in
my
throne,
even as
I also
overcame,
and
am set down
with
my
Father
in
his
throne.}}
\VS{22}{\WJ{He that hath
an ear,
let him hear
what
the Spirit
saith
unto the churches.}}

\par }\Chap{4}{\PP \VerseOne{1}After
this I
looked,
and,
behold, a
door
{\ADD{was}}
opened
in
heaven:
and the
first
voice
which I
heard
{\ADD{was}} as it
were of a
trumpet
talking
with
me; which
said, Come
up
hither,
and I will
shew
thee
things
which
must
be
hereafter.
\VS{2}And
immediately I
was
in the
spirit:
and,
behold, a
throne was
set
in
heaven,
and
{\ADD{one}}
sat
on the
throne.
\VS{3}And
he that
sat
was to look
upon
like a
jasper
and a
sardine
stone:
and
{\ADD{there was}} a
rainbow round
about the
throne, in
sight like
unto an
emerald.
\VS{4}And round
about the
throne
{\ADD{were}}
four
and
twenty
seats:
and
upon the
seats I
saw
four
and
twenty
elders
sitting,
clothed
in
white
raiment;
and they
had
on
their
heads
crowns of
gold.
\VS{5}And out
of the
throne
proceeded
lightnings
and
thunderings
and
voices:
and
{\ADD{there were}}
seven
lamps of
fire
burning
before the
throne,
which
are the
seven
Spirits of
God.
\VS{6}And
before the
throne
{\ADD{there was}} a
sea of
glass like
unto
crystal:
and
in the
midst of the
throne,
and round
about the
throne,
{\ADD{were}}
four
beasts full
of
eyes
before
and
behind.
\VS{7}And the
first
beast
{\ADD{was}}
like a
lion,
and the
second
beast
like a
calf,
and the
third
beast
had a
face
as a
man,
and the
fourth
beast
{\ADD{was}}
like a
flying
eagle.
\VS{8}And the
four
beasts
had
each of
them
six
wings
about
{\ADD{him}};
and
{\ADD{they were}}
full of
eyes
within:
and they
rest
not
day
and
night,
saying,
Holy,
holy,
holy,
Lord
God
Almighty,
which
was,
and
is,
and is to
come.
\VS{9}And
when
those
beasts
give
glory
and
honour
and
thanks to
him that
sat
on the
throne,
who
liveth
for
ever and
ever,
\VS{10}The
four
and
twenty
elders fall
down
before
him that
sat
on the
throne,
and
worship
him that
liveth
for
ever and
ever,
and
cast
their
crowns
before the
throne,
saying,
\VS{11}Thou
art
worthy, O
Lord, to
receive
glory
and
honour
and
power:
for
thou hast
created all
things,
and
for
thy
pleasure they
are
and were
created.

\par }\Chap{5}{\PP \VerseOne{1}And I
saw
in the right
hand of
him that
sat
on the
throne a
book
written
within
and on the
backside,
sealed with
seven
seals.
\VS{2}And I
saw a
strong
angel
proclaiming with a
loud
voice,
Who
is
worthy to
open the
book,
and to
loose the
seals
thereof?
\VS{3}And no
man
in
heaven,
nor
in
earth,
neither
under the
earth, was
able to
open the
book,
neither to
look
thereon.
\VS{4}And
I
wept
much,
because no
man was
found
worthy to
open
and to
read the
book,
neither to
look
thereon.
\VS{5}And
one
of the
elders
saith unto
me,
Weep
not:
behold, the
Lion
of the
tribe of
Juda, the
Root of
David, hath
prevailed to
open the
book,
and to
loose the
seven
seals
thereof.
\VS{6}And I
beheld,
and,
lo,
in the
midst of the
throne
and of the
four
beasts,
and
in the
midst of the
elders,
stood a
Lamb
as it had been
slain,
having
seven
horns
and
seven
eyes,
which
are the
seven
Spirits of
God sent
forth
into
all the
earth.
\VS{7}And he
came
and
took the
book out
of the right
hand of
him that
sat
upon the
throne.
\VS{8}And
when he had
taken the
book, the
four
beasts
and
four
{\ADD{and}}
twenty
elders fell
down
before the
Lamb,
having
every one of them
harps,
and
golden
vials
full of
odours,
which
are the
prayers of
saints.
\VS{9}And they
sung a
new
song,
saying, Thou
art
worthy to
take the
book,
and to
open the
seals
thereof:
for thou wast
slain,
and hast
redeemed
us to
God
by
thy
blood
out of
every
kindred,
and
tongue,
and
people,
and
nation;
\VS{10}And hast
made
us unto
our
God
kings
and
priests:
and we shall
reign
on the
earth.
\VS{11}And I
beheld,
and I
heard the
voice of
many
angels round
about the
throne
and the
beasts
and the
elders:
and the
number of
them was ten thousand
times ten
thousand,
and
thousands of
thousands;
\VS{12}Saying with a
loud
voice,
Worthy
is the
Lamb that was
slain to
receive
power,
and
riches,
and
wisdom,
and
strength,
and
honour,
and
glory,
and
blessing.
\VS{13}And
every
creature
which
is
in
heaven,
and
on the
earth,
and
under the
earth,
and such
as
are
in the
sea,
and
all that are
in
them, heard
I
saying,
Blessing,
and
honour,
and
glory,
and
power,
{\ADD{be}} unto
him that
sitteth
upon the
throne,
and unto the
Lamb
for
ever and
ever.
\VS{14}And the
four
beasts
said,
Amen.
And the
four
{\ADD{and}}
twenty
elders fell
down
and
worshipped him that
liveth
for
ever and
ever.

\par }\Chap{6}{\PP \VerseOne{1}And I
saw
when the
Lamb
opened
one
of the
seals,
and I
heard, as it
were the
noise of
thunder,
one
of the
four
beasts
saying,
Come
and
see.
\VS{2}And I
saw,
and
behold a
white
horse:
and
he that
sat
on
him
had a
bow;
and a
crown was
given unto
him:
and he went
forth
conquering,
and
to
conquer.
\VS{3}And
when he had
opened the
second
seal, I
heard the
second
beast
say,
Come
and
see.
\VS{4}And there went
out
another
horse
{\ADD{that was}}
red:
and
{\ADD{power}} was
given to
him that
sat
thereon to
take
peace
from the
earth,
and
that they should
kill one
another:
and there was
given unto
him a
great
sword.
\VS{5}And
when he had
opened the
third
seal, I
heard the
third
beast
say,
Come
and
see.
And I
beheld,
and
lo a
black
horse;
and
he that
sat
on
him
had a pair of
balances
in
his
hand.
\VS{6}And I
heard a
voice
in the
midst of the
four
beasts
say, A
measure of
wheat for a
penny,
and
three
measures of
barley for a
penny;
and
{\ADD{see}} thou
hurt
not the
oil
and the
wine.
\VS{7}And
when he had
opened the
fourth
seal, I
heard the
voice of the
fourth
beast
say,
Come
and
see.
\VS{8}And I
looked,
and
behold a
pale
horse:
and
his
name
that
sat
on
him was
Death,
and
Hell
followed
with
him.
And
power was
given unto
them
over the fourth
part of the
earth, to
kill
with
sword,
and
with
hunger,
and
with
death,
and
with the
beasts of the
earth.
\VS{9}And
when he had
opened the
fifth
seal, I
saw
under the
altar the
souls of
them that were
slain
for the
word of
God,
and
for the
testimony
which they
held:
\VS{10}And they
cried with a
loud
voice,
saying,
How
long, O
Lord,
holy
and
true, dost
thou
not
judge
and
avenge
our
blood
on them that
dwell
on the
earth?
\VS{11}And
white
robes were
given unto every one of
them;
and it was
said unto
them,
that they should
rest
yet for a
little
season,
until
their
fellowservants
also
and
their
brethren, that should be
killed
as
they
{\ADD{were}}, should
be
fulfilled.
\VS{12}And I
beheld
when he had
opened the
sixth
seal,
and,
lo, there
was a
great
earthquake;
and the
sun
became
black
as
sackcloth of
hair,
and the
moon
became
as
blood;
\VS{13}And the
stars of
heaven
fell
unto the
earth, even
as a fig
tree
casteth
her untimely
figs, when she is
shaken
of a
mighty
wind.
\VS{14}And the
heaven
departed
as a
scroll when it is rolled
together;
and
every
mountain
and
island were
moved
out of
their
places.
\VS{15}And the
kings of the
earth,
and the
great
men,
and the rich
men,
and the chief
captains,
and the mighty
men,
and
every
bondman,
and
every free
man,
hid
themselves
in the
dens
and
in the
rocks of the
mountains;
\VS{16}And
said to the
mountains
and
rocks,
Fall
on
us,
and
hide
us
from the
face of him that
sitteth
on the
throne,
and
from the
wrath of the
Lamb:
\VS{17}For the
great
day of
his
wrath is
come;
and
who shall be
able to
stand?

\par }\Chap{7}{\PP \VerseOne{1}And
after these
things I
saw
four
angels
standing
on the
four
corners of the
earth,
holding the
four
winds of the
earth,
that the
wind should
not
blow
on the
earth,
nor
on the
sea,
nor
on
any
tree.
\VS{2}And I
saw
another
angel
ascending
from the
east,
having the
seal of the
living
God:
and he
cried with a
loud
voice to the
four
angels, to
whom it was
given to
hurt the
earth
and the
sea,
\VS{3}Saying,
Hurt
not the
earth,
neither the
sea,
nor the
trees,
till we have
sealed the
servants of
our
God
in
their
foreheads.
\VS{4}And I
heard the
number of
them which were
sealed:
{\ADD{and there were}}
sealed an
hundred
{\ADD{and}}
forty
{\ADD{and}}
four
thousand
of
all the
tribes of the
children of
Israel.
\VS{5}Of the
tribe of
Juda
{\ADD{were}}
sealed
twelve
thousand.
Of the
tribe of
Reuben
{\ADD{were}}
sealed
twelve
thousand.
Of the
tribe of
Gad
{\ADD{were}}
sealed
twelve
thousand.
\VS{6}Of the
tribe of
Aser
{\ADD{were}}
sealed
twelve
thousand.
Of the
tribe of
Nepthalim
{\ADD{were}}
sealed
twelve
thousand.
Of the
tribe of
Manasses
{\ADD{were}}
sealed
twelve
thousand.
\VS{7}Of the
tribe of
Simeon
{\ADD{were}}
sealed
twelve
thousand.
Of the
tribe of
Levi
{\ADD{were}}
sealed
twelve
thousand.
Of the
tribe of
Issachar
{\ADD{were}}
sealed
twelve
thousand.
\VS{8}Of the
tribe of
Zabulon
{\ADD{were}}
sealed
twelve
thousand.
Of the
tribe of
Joseph
{\ADD{were}}
sealed
twelve
thousand.
Of the
tribe of
Benjamin
{\ADD{were}}
sealed
twelve
thousand.
\VS{9}After
this I
beheld,
and,
lo, a
great
multitude,
which
no
man
could
number,
of
all
nations,
and
kindreds,
and
people,
and
tongues,
stood
before the
throne,
and
before the
Lamb,
clothed with
white
robes,
and
palms
in
their
hands;
\VS{10}And
cried with a
loud
voice,
saying,
Salvation to
our
God
which
sitteth
upon the
throne,
and unto the
Lamb.
\VS{11}And
all the
angels
stood round
about the
throne,
and
{\ADD{about}} the
elders
and the
four
beasts,
and
fell
before the
throne
on
their
faces,
and
worshipped
God,
\VS{12}Saying,
Amen:
Blessing,
and
glory,
and
wisdom,
and
thanksgiving,
and
honour,
and
power,
and
might,
{\ADD{be}} unto
our
God
for
ever and
ever.
Amen.
\VS{13}And
one
of the
elders
answered,
saying unto
me, What are
these
which
are
arrayed in
white
robes?
and whence
came
they?
\VS{14}And I
said unto
him,
Sir,
thou
knowest.
And he
said to
me,
These
are they which
came
out of
great
tribulation,
and have
washed
their
robes,
and
made
them
white
in the
blood of the
Lamb.
\VS{15}Therefore are
they
before the
throne of
God,
and
serve
him
day
and
night
in
his
temple:
and
he that
sitteth
on the
throne shall
dwell
among
them.
\VS{16}They shall
hunger
no
more,
neither
thirst
any
more;
neither shall the
sun
light
on
them,
nor
any
heat.
\VS{17}For the
Lamb
which is
in the
midst of the
throne shall
feed
them,
and shall
lead
them
unto
living
fountains of
waters:
and
God shall wipe
away
all
tears
from
their
eyes.

\par }\Chap{8}{\PP \VerseOne{1}And
when he had
opened the
seventh
seal, there
was
silence
in
heaven
about the space of half an
hour.
\VS{2}And I
saw the
seven
angels
which
stood
before
God;
and to
them were
given
seven
trumpets.
\VS{3}And
another
angel
came
and
stood
at the
altar,
having a
golden
censer;
and there was
given unto
him
much
incense,
that he should
offer
{\ADD{it}} with the
prayers of
all
saints
upon the
golden
altar which
was
before the
throne.
\VS{4}And the
smoke of the
incense,
{\ADD{which came}} with the
prayers of the
saints, ascended
up
before
God
out of the
angel’s
hand.
\VS{5}And the
angel
took the
censer,
and
filled
it with
fire
of the
altar,
and
cast
{\ADD{it}}
into the
earth:
and there
were
voices,
and
thunderings,
and
lightnings,
and an
earthquake.
\VS{6}And the
seven
angels which
had the
seven
trumpets
prepared
themselves
to
sound.
\par }{\PP \VS{7}The
first
angel
sounded,
and there
followed
hail
and
fire
mingled with
blood,
and they were
cast
upon the
earth:
and the third
part of
trees was burnt
up,
and
all
green
grass was burnt
up.
\VS{8}And the
second
angel
sounded,
and as it
were a
great
mountain
burning with
fire was
cast
into the
sea:
and the third
part of the
sea
became
blood;
\VS{9}And the third
part of the
creatures which
were
in the
sea, and
had
life,
died;
and the third
part of the
ships were
destroyed.
\VS{10}And the
third
angel
sounded,
and there
fell a
great
star
from
heaven,
burning as it
were a
lamp,
and it
fell
upon the third
part of the
rivers,
and
upon the
fountains of
waters;
\VS{11}And the
name of the
star is
called
Wormwood:
and the
third part of the
waters
became
wormwood;
and
many
men
died
of the waters,
because they were made
bitter.
\VS{12}And the
fourth
angel
sounded,
and the third
part of the
sun was
smitten,
and the third
part of the
moon,
and the third
part of the
stars; so
as the third
part of
them was
darkened,
and the
day
shone
not for a third
part of
it,
and the
night
likewise.
\VS{13}And I
beheld,
and
heard
an
angel
flying through the midst of
heaven,
saying
with a
loud
voice,
Woe,
woe,
woe, to
the
inhabiters
of the
earth by
reason of the
other
voices of the
trumpet of the
three
angels, which are
yet to
sound!

\par }\Chap{9}{\PP \VerseOne{1}And the
fifth
angel
sounded,
and I
saw a
star
fall
from
heaven
unto the
earth:
and to
him was
given the
key of the
bottomless
pit.
\VS{2}And he
opened the
bottomless
pit;
and there
arose a
smoke
out of the
pit,
as the
smoke of a
great
furnace;
and the
sun
and the
air were
darkened by
reason of the
smoke of the
pit.
\VS{3}And there
came
out of the
smoke
locusts
upon the
earth:
and unto
them was
given
power,
as the
scorpions of the
earth
have
power.
\VS{4}And it was
commanded
them
that they
should
not
hurt the
grass of the
earth,
neither
any green
thing,
neither
any
tree;
but
only
those
men which
have
not the
seal of
God
in
their
foreheads.
\VS{5}And to
them it was
given
that they
should
not
kill
them,
but
that they should be
tormented
five
months:
and
their
torment
{\ADD{was}}
as the
torment of a
scorpion,
when he
striketh a
man.
\VS{6}And
in
those
days
shall
men
seek
death,
and
shall
not
find
it;
and shall
desire to
die,
and
death shall
flee
from
them.
\VS{7}And the
shapes of the
locusts
{\ADD{were}} like
unto
horses
prepared
unto
battle;
and
on
their
heads
{\ADD{were}} as it
were
crowns
like
gold,
and
their
faces
{\ADD{were}}
as the
faces of
men.
\VS{8}And they
had
hair
as the
hair of
women,
and
their
teeth
were
as
{\ADD{the teeth}} of
lions.
\VS{9}And they
had
breastplates, as it
were
breastplates of
iron;
and the
sound of
their
wings
{\ADD{was}}
as the
sound of
chariots of
many
horses
running
to
battle.
\VS{10}And they
had
tails like
unto
scorpions,
and there
were
stings
in
their
tails:
and
their
power
{\ADD{
was}} to
hurt
men
five
months.
\VS{11}And they
had a
king
over
them,
{\ADD{which is}} the
angel of the bottomless
pit, whose
name in the Hebrew
tongue
{\ADD{is}}
Abaddon,
but
in the Greek
tongue
hath
{\ADD{
his}}
name
Apollyon.
\VS{12}One
woe is
past;
{\ADD{and}},
behold, there
come
two
woes
more
hereafter.
\par }{\PP \VS{13}And the
sixth
angel
sounded,
and I
heard
a
voice
from the
four
horns of the
golden
altar which is
before
God,
\VS{14}Saying to the
sixth
angel
which
had the
trumpet,
Loose the
four
angels
which are
bound
in the
great
river
Euphrates.
\VS{15}And the
four
angels were
loosed,
which were
prepared
for an
hour,
and a
day,
and a
month,
and a
year, for
to
slay the third
part of
men.
\VS{16}And the
number of the
army of the
horsemen
{\ADD{were}}
two hundred
thousand
thousand:
and I
heard the
number of
them.
\VS{17}And
thus I
saw the
horses
in the
vision,
and
them that
sat
on
them,
having
breastplates of
fire,
and of
jacinth,
and
brimstone:
and the
heads of the
horses
{\ADD{were}}
as the
heads of
lions;
and
out of
their
mouths
issued
fire
and
smoke
and
brimstone.
\VS{18}By
these
three was the third
part of
men
killed,
by the
fire,
and
by the
smoke,
and
by the
brimstone,
which
issued
out of
their
mouths.
\VS{19}For
their
power
is
in
their
mouth,
and
in
their
tails:
for
their
tails
{\ADD{were}} like
unto
serpents, and
had
heads,
and
with
them they do
hurt.
\VS{20}And the
rest of the
men
which
were
not
killed
by
these
plagues yet
repented
not
of the
works of
their
hands,
that they
should
not
worship
devils,
and
idols of
gold,
and
silver,
and
brass,
and
stone,
and of
wood:
which
neither
can
see,
nor
hear,
nor
walk:
\VS{21}Neither repented
they
of
their
murders,
nor
of
their
sorceries,
nor
of
their
fornication,
nor
of
their
thefts.

\par }\Chap{10}{\PP \VerseOne{1}And I
saw
another
mighty
angel come
down
from
heaven, clothed
with a
cloud:
and a
rainbow
{\ADD{was}}
upon his
head,
and
his
face
{\ADD{was}} as it
were the
sun,
and
his
feet
as
pillars of
fire:
\VS{2}And he
had
in
his
hand a little
book
open:
and he
set
his
right
foot
upon the
sea,
and
{\ADD{his}}
left
{\ADD{foot}}
on the
earth,
\VS{3}And
cried with a
loud
voice,
as
{\ADD{when}} a
lion
roareth:
and
when he had
cried,
seven
thunders
uttered
their
voices.
\VS{4}And
when the
seven
thunders had
uttered
their
voices, I was
about to
write:
and I
heard a
voice
from
heaven
saying unto
me, Seal
up those things
which the
seven
thunders
uttered,
and
write
them
not.
\VS{5}And the
angel
which I
saw
stand
upon the
sea
and
upon the
earth lifted
up
his
hand
to
heaven,
\VS{6}And
sware
by him that
liveth
for
ever and
ever,
who
created
heaven,
and the things
that therein
are,
and the
earth,
and the things
that therein
are,
and the
sea,
and the things
which are
therein,
that there should
be
time
no
longer:
\VS{7}But
in the
days of the
voice of the
seventh
angel,
when he shall
begin to
sound, the
mystery of
God should be
finished,
as he hath
declared to
his
servants the
prophets.
\VS{8}And the
voice
which I
heard
from
heaven
spake
unto
me
again,
and
said,
Go
{\ADD{and}}
take the little
book
which is
open
in the
hand of the
angel
which
standeth
upon the
sea
and
upon the
earth.
\VS{9}And I
went
unto the
angel, and
said unto
him,
Give
me the little
book.
And he
said unto
me,
Take
{\ADD{it}},
and
eat
it
up;
and it shall
make
thy belly
bitter,
but it shall
be
in
thy
mouth
sweet
as
honey.
\VS{10}And I
took the little
book
out of the
angel’s
hand,
and
ate
it
up;
and it
was
in
my
mouth
sweet
as
honey:
and as soon
as I had
eaten
it,
my
belly was
bitter.
\VS{11}And he
said unto
me,
Thou
must
prophesy
again
before
many
peoples,
and
nations,
and
tongues,
and
kings.

\par }\Chap{11}{\PP \VerseOne{1}And there was
given
me a
reed like
unto a
rod: and the angel stood,
saying,
Rise,
and
measure the
temple of
God,
and the
altar,
and them
that
worship
therein.
\VS{2}But the
court which is
without the
temple
leave
out,
and
measure
it
not;
for it is
given unto the
Gentiles:
and the
holy
city shall they tread under
foot
forty
{\ADD{and}}
two
months.
\VS{3}And I will
give
{\ADD{power}}
unto
my
two
witnesses,
and they shall
prophesy a
thousand two
hundred
{\ADD{and}}
threescore
days,
clothed in
sackcloth.
\VS{4}These
are the
two olive
trees,
and the
two
candlesticks
standing
before the
God of the
earth.
\VS{5}And
if any
man
will
hurt
them,
fire
proceedeth out
of
their
mouth,
and
devoureth
their
enemies:
and
if any
man
will
hurt
them,
he
must in this
manner be
killed.
\VS{6}These
have
power to
shut
heaven,
that it
rain
not
in the
days of
their
prophecy:
and
have
power
over
waters to
turn
them
to
blood,
and to
smite the
earth with
all
plagues, as
often
as they
will.
\VS{7}And
when they shall have
finished
their
testimony, the
beast that
ascendeth out
of the bottomless
pit shall
make
war
against
them,
and shall
overcome
them,
and
kill
them.
\VS{8}And
their dead
bodies
{\ADD{shall lie}}
in the
street of the
great
city,
which
spiritually is
called
Sodom
and
Egypt,
where
also
our
Lord was
crucified.
\VS{9}And they
of the
people
and
kindreds
and
tongues
and
nations shall
see
their dead
bodies
three
days
and an
half,
and
shall
not
suffer
their dead
bodies to be
put
in
graves.
\VS{10}And
they that
dwell
upon the
earth shall
rejoice
over
them,
and make
merry,
and shall
send
gifts one to
another;
because
these
two
prophets
tormented
them that
dwelt
on the
earth.
\VS{11}And
after
three
days
and an
half the
Spirit
of
life from
God
entered
into
them,
and they
stood
upon
their
feet;
and
great
fear
fell
upon
them which
saw
them.
\VS{12}And they
heard a
great
voice
from
heaven
saying unto
them, Come
up
hither.
And they ascended
up
to
heaven
in a
cloud;
and
their
enemies
beheld
them.
\VS{13}And the
same
hour was
there a
great
earthquake,
and the tenth
part of the
city
fell,
and
in the
earthquake were
slain of
men
seven
thousand:
and the
remnant
were
affrighted,
and
gave
glory to the
God of
heaven.
\VS{14}The
second
woe is
past;
{\ADD{and}},
behold, the
third
woe
cometh
quickly.
\VS{15}And the
seventh
angel
sounded;
and there
were
great
voices
in
heaven,
saying, The
kingdoms of this
world are
become
{\ADD{the kingdoms}} of
our
Lord,
and of
his
Christ;
and he shall
reign
for
ever and
ever.
\VS{16}And the
four
and
twenty
elders, which
sat
before
God
on
their
seats,
fell
upon
their
faces,
and
worshipped
God,
\VS{17}Saying, We
give
thee
thanks, O
Lord
God
Almighty, which
art,
and
wast,
and art to
come;
because thou hast taken to
thee
thy
great
power,
and hast
reigned.
\VS{18}And the
nations were
angry,
and
thy
wrath is
come,
and the
time of the
dead, that they should be
judged,
and that thou shouldest
give
reward unto
thy
servants the
prophets,
and to the
saints,
and
them that
fear
thy
name,
small
and
great;
and shouldest
destroy
them which
destroy the
earth.
\VS{19}And the
temple of
God was
opened
in
heaven,
and there was
seen
in
his
temple the
ark of
his
testament:
and there
were
lightnings,
and
voices,
and
thunderings,
and an
earthquake,
and
great
hail.

\par }\Chap{12}{\PP \VerseOne{1}And there
appeared a
great
wonder
in
heaven; a
woman clothed
with the
sun,
and the
moon
under
her
feet,
and
upon
her
head a
crown of
twelve
stars:
\VS{2}And she
being
with
child
cried, travailing in
birth,
and
pained to be
delivered.
\VS{3}And there
appeared
another
wonder
in
heaven;
and
behold a
great
red
dragon,
having
seven
heads
and
ten
horns,
and
seven
crowns
upon
his
heads.
\VS{4}And
his
tail
drew the third
part of the
stars of
heaven,
and did
cast
them
to the
earth:
and the
dragon
stood
before the
woman which was
ready to be
delivered,
for to
devour
her
child as soon
as it was
born.
\VS{5}And she brought
forth a
man
child,
who was
to
rule
all
nations
with a
rod of
iron:
and
her
child was caught
up
unto
God,
and
{\ADD{to}}
his
throne.
\VS{6}And the
woman
fled
into the
wilderness,
where she
hath a
place
prepared
of
God,
that they should
feed
her
there a
thousand two
hundred
{\ADD{and}}
threescore
days.
\VS{7}And there
was
war
in
heaven:
Michael
and
his
angels
fought
against the
dragon;
and the
dragon
fought
and
his
angels,
\VS{8}And
prevailed
not;
neither
was
their
place
found any
more
in
heaven.
\VS{9}And the
great
dragon was cast
out, that
old
serpent,
called the
Devil,
and
Satan, which
deceiveth the
whole
world: he was cast
out
into the
earth,
and
his
angels were cast
out
with
him.
\VS{10}And I
heard a
loud
voice
saying
in
heaven,
Now is
come
salvation,
and
strength,
and the
kingdom of
our
God,
and the
power of
his
Christ:
for the
accuser of
our
brethren is cast
down,
which
accused
them
before
our
God
day
and
night.
\VS{11}And
they
overcame
him
by the
blood of the
Lamb,
and
by the
word of
their
testimony;
and they
loved
not
their
lives
unto the
death.
\VS{12}Therefore
rejoice,
{\ADD{ye}}
heavens,
and
ye that
dwell
in
them.
Woe to the
inhabiters of the
earth
and of the
sea!
for the
devil is come
down
unto
you,
having
great
wrath,
because he knoweth
that he
hath but a
short
time.
\VS{13}And
when the
dragon
saw
that he was
cast
unto the
earth, he
persecuted the
woman
which brought
forth the
man
{\ADD{child}}.
\VS{14}And to the
woman were
given
two
wings of a
great
eagle,
that she might
fly
into the
wilderness,
into
her
place,
where she is
nourished
for a
time,
and
times,
and
half a
time,
from the
face of the
serpent.
\VS{15}And the
serpent cast
out
of
his
mouth
water
as a
flood
after the
woman,
that he might
cause
her to be carried away of the
flood.
\VS{16}And the
earth
helped the
woman,
and the
earth
opened
her
mouth,
and swallowed
up the
flood
which the
dragon cast
out
of
his
mouth.
\VS{17}And the
dragon was
wroth
with the
woman,
and
went to
make
war
with the
remnant of
her
seed,
which
keep the
commandments of
God,
and
have the
testimony of
Jesus
Christ.

\par }\Chap{13}{\PP \VerseOne{1}And I
stood
upon the
sand of the
sea,
and
saw a
beast rise
up out
of the
sea,
having
seven
heads
and
ten
horns,
and
upon
his
horns
ten
crowns,
and
upon
his
heads the
name of
blasphemy.
\VS{2}And the
beast
which I
saw
was like
unto a
leopard,
and
his
feet were
as
{\ADD{the feet}} of a
bear,
and
his
mouth
as the
mouth of a
lion:
and the
dragon
gave
him
his
power,
and
his
seat,
and
great
authority.
\VS{3}And I
saw
one of
his
heads as it
were
wounded
to
death;
and
his
deadly
wound was
healed:
and
all the
world
wondered
after the
beast.
\VS{4}And they
worshipped the
dragon
which
gave
power unto the
beast:
and they
worshipped the
beast,
saying,
Who
{\ADD{is}} like
unto the
beast?
who is
able to make
war
with
him?
\VS{5}And there was
given unto
him a
mouth
speaking great
things
and
blasphemies;
and
power was
given unto
him to
continue
forty
{\ADD{and}}
two
months.
\VS{6}And he
opened
his
mouth
in
blasphemy
against
God, to
blaspheme
his
name,
and
his
tabernacle,
and
them that
dwell
in
heaven.
\VS{7}And it was
given unto
him to
make
war
with the
saints,
and to
overcome
them:
and
power was
given
him
over
all
kindreds,
and
tongues,
and
nations.
\VS{8}And
all
that
dwell
upon the
earth shall
worship
him,
whose
names
are
not
written
in the
book of
life of the
Lamb
slain
from the
foundation of the
world.
\VS{9}If any
man
have an
ear, let him
hear.
\VS{10}He
that leadeth
into
captivity shall
go
into
captivity:
he
that
killeth
with the
sword
must
be
killed
with the
sword.
Here
is the
patience
and the
faith of the
saints.
\VS{11}And I
beheld
another
beast coming
up out
of the
earth;
and he
had
two
horns
like a
lamb,
and he
spake
as a
dragon.
\VS{12}And he
exerciseth
all the
power of the
first
beast
before
him,
and
causeth the
earth
and
them which
dwell
therein
to
worship the
first
beast,
whose
deadly
wound was
healed.
\VS{13}And he
doeth
great
wonders,
so
that he
maketh
fire come
down
from
heaven
on the
earth in the
sight of
men,
\VS{14}And
deceiveth
them that
dwell
on the
earth
by
{\ADD{the means of}}
those
miracles
which
he had
power to
do in the sight
of the
beast;
saying to
them that
dwell
on the
earth, that they should
make an
image to the
beast,
which
had the
wound by a
sword,
and did
live.
\VS{15}And
he had
power to
give
life unto the
image of the
beast,
that the
image of the
beast
should
both
speak,
and
cause
that as
many
as
would
not
worship the
image of the
beast should be
killed.
\VS{16}And he
causeth
all,
both
small
and
great,
rich
and
poor,
free
and
bond,
to
receive a
mark
in
their
right
hand,
or
in
their
foreheads:
\VS{17}And
that
no
man
might
buy
or
sell, save
he
that
had the
mark,
or the
name of the
beast,
or the
number of
his
name.
\VS{18}Here
is
wisdom. Let
him that
hath
understanding
count the
number of the
beast:
for it
is the
number of a
man;
and
his
number
{\ADD{is}} Six hundred
threescore
{\ADD{and}}
six.

\par }\Chap{14}{\PP \VerseOne{1}And I
looked,
and,
lo, a
Lamb
stood
on the
mount
Sion,
and
with
him an
hundred
forty
{\ADD{and}}
four
thousand,
having
his
Father’s
name
written
in
their
foreheads.
\VS{2}And I
heard a
voice
from
heaven,
as the
voice of
many
waters,
and
as the
voice of a
great
thunder:
and I
heard the
voice of
harpers
harping
with
their
harps:
\VS{3}And they
sung as it
were a
new
song
before the
throne,
and
before the
four
beasts,
and the
elders:
and no
man
could
learn
that
song
but the
hundred
{\ADD{and}}
forty
{\ADD{and}}
four
thousand,
which were
redeemed
from the
earth.
\VS{4}These
are they
which
were
not
defiled
with
women;
for they
are
virgins.
These
are they
which
follow the
Lamb
whithersoever he
goeth.
These were
redeemed
from among
men,
{\ADD{being}} the
firstfruits unto
God
and to the
Lamb.
\VS{5}And
in
their
mouth was
found
no
guile:
for they
are without
fault
before the
throne of
God.
\par }{\PP \VS{6}And I
saw
another
angel
fly
in the midst of
heaven,
having the
everlasting
gospel to
preach unto
them that
dwell
on the
earth,
and to
every
nation,
and
kindred,
and
tongue,
and
people,
\VS{7}Saying
with a
loud
voice,
Fear
God,
and
give
glory to
him;
for the
hour of
his
judgment is
come:
and
worship
him that
made
heaven,
and
earth,
and the
sea,
and the
fountains of
waters.
\VS{8}And there
followed
another
angel,
saying,
Babylon is
fallen, is
fallen, that
great
city,
because she
made
all
nations
drink
of the
wine of the
wrath of
her
fornication.
\VS{9}And the
third
angel
followed
them,
saying
with a
loud
voice,
If any
man
worship the
beast
and
his
image,
and
receive
{\ADD{his}}
mark
in
his
forehead,
or
in
his
hand,
\VS{10}The
same shall
drink
of the
wine of the
wrath of
God,
which is poured
out without
mixture
into the
cup of
his
indignation;
and he shall be
tormented
with
fire
and
brimstone in the
presence of the
holy
angels,
and in the presence
of the
Lamb:
\VS{11}And the
smoke of
their
torment ascendeth
up
for
ever
and
ever:
and they
have
no
rest
day nor
night,
who
worship the
beast
and
his
image,
and
whosoever
receiveth the
mark of
his
name.
\VS{12}Here
is the
patience of the
saints:
here
{\ADD{are}}
they that
keep the
commandments of
God,
and the
faith of
Jesus.
\VS{13}And I
heard a
voice
from
heaven
saying unto
me,
Write,
Blessed
{\ADD{are}} the
dead
which
die
in the
Lord from
henceforth:
Yea,
saith the
Spirit,
that they may
rest
from
their
labours;
and
their
works do
follow
them.
\par }{\PP \VS{14}And I
looked,
and
behold a
white
cloud,
and
upon the
cloud
{\ADD{one}}
sat like
unto the
Son of
man,
having
on
his
head a
golden
crown,
and
in
his
hand a
sharp
sickle.
\VS{15}And
another
angel
came out
of the
temple,
crying
with a
loud
voice to
him that
sat
on the
cloud, Thrust
in
thy
sickle,
and
reap:
for the
time is
come for
thee to
reap;
for the
harvest of the
earth is
ripe.
\VS{16}And
he that
sat
on the
cloud
thrust in
his
sickle
on the
earth;
and the
earth was
reaped.
\VS{17}And
another
angel
came out
of the
temple which
is
in
heaven,
he
also
having a
sharp
sickle.
\VS{18}And
another
angel
came out
from the
altar, which
had
power
over
fire;
and
cried with a
loud
cry to
him that
had the
sharp
sickle,
saying, Thrust
in
thy
sharp
sickle,
and
gather the
clusters of the
vine of the earth;
for
her
grapes are fully
ripe.
\VS{19}And the
angel thrust
in
his
sickle
into the
earth,
and
gathered the
vine of the
earth,
and
cast
{\ADD{it}}
into the
great
winepress of the
wrath of
God.
\VS{20}And the
winepress was
trodden
without the
city,
and
blood
came out
of the
winepress, even
unto the
horse
bridles, by the space
of a
thousand
{\ADD{and}} six
hundred
furlongs.

\par }\Chap{15}{\PP \VerseOne{1}And I
saw
another
sign
in
heaven,
great
and
marvellous,
seven
angels
having the
seven
last
plagues;
for
in
them is filled
up the
wrath of
God.
\VS{2}And I
saw as it
were a
sea of
glass
mingled with
fire:
and them
that had gotten the
victory
over the
beast,
and
over
his
image,
and
over
his
mark,
{\ADD{and}}
over the
number of
his
name,
stand
on the
sea of
glass,
having the
harps of
God.
\VS{3}And they
sing the
song of
Moses the
servant of
God,
and the
song of the
Lamb,
saying,
Great
and
marvellous
{\ADD{are}}
thy
works,
Lord
God
Almighty;
just
and
true
{\ADD{are}}
thy
ways, thou
King of
saints.
\VS{4}Who
shall
not
fear
thee, O
Lord,
and
glorify
thy
name?
for
{\ADD{thou}}
only
{\ADD{art}}
holy:
for
all
nations shall
come
and
worship
before
thee;
for
thy
judgments are made
manifest.
\VS{5}And
after
that I
looked,
and,
behold, the
temple of the
tabernacle of the
testimony
in
heaven was
opened:
\VS{6}And the
seven
angels
came out
of the
temple,
having the
seven
plagues, clothed
in
pure
and
white
linen,
and having
their
breasts
girded with
golden
girdles.
\VS{7}And
one
of the
four
beasts
gave unto the
seven
angels
seven
golden
vials
full of the
wrath of
God, who
liveth
for
ever and
ever.
\VS{8}And the
temple was
filled with
smoke
from the
glory of
God,
and
from
his
power;
and no
man was
able to
enter
into the
temple,
till the
seven
plagues of the
seven
angels were
fulfilled.

\par }\Chap{16}{\PP \VerseOne{1}And I
heard a
great
voice out
of the
temple
saying to the
seven
angels, Go your
ways,
and pour
out the
vials of the
wrath of
God
upon the
earth.
\VS{2}And the
first
went,
and poured
out
his
vial
upon the
earth;
and there
fell a
noisome
and
grievous
sore
upon the
men
which
had the
mark of the
beast,
and
{\ADD{upon}}
them which
worshipped
his
image.
\VS{3}And the
second
angel poured
out
his
vial
upon the
sea;
and it
became
as the
blood of a
dead
{\ADD{man}}:
and
every
living
soul
died
in the
sea.
\VS{4}And the
third
angel poured
out
his
vial
upon the
rivers
and
fountains of
waters;
and they
became
blood.
\VS{5}And I
heard the
angel of the
waters
say, Thou
art
righteous, O
Lord, which
art,
and
wast,
and shalt
be,
because thou hast
judged
thus.
\VS{6}For they have
shed the
blood of
saints
and
prophets,
and thou hast
given
them
blood to
drink;
for they
are
worthy.
\VS{7}And I
heard
another out
of the
altar
say, Even
so,
Lord
God
Almighty,
true
and
righteous
{\ADD{are}}
thy
judgments.
\VS{8}And the
fourth
angel poured
out
his
vial
upon the
sun;
and
power was
given unto
him to
scorch
men
with
fire.
\VS{9}And
men were
scorched with
great
heat,
and
blasphemed the
name of
God,
which
hath
power
over
these
plagues:
and they
repented
not to
give
him
glory.
\VS{10}And the
fifth
angel poured
out
his
vial
upon the
seat of the
beast;
and
his
kingdom
was full of
darkness;
and they
gnawed
their
tongues
for
pain,
\VS{11}And
blasphemed the
God of
heaven
because of
their
pains
and
their
sores,
and
repented
not
of
their
deeds.
\VS{12}And the
sixth
angel poured
out
his
vial
upon the
great
river
Euphrates;
and the
water
thereof was dried
up,
that the
way of the
kings
of the
east might be
prepared.
\VS{13}And I
saw
three
unclean
spirits
like
frogs
{\ADD{come}} out
of the
mouth of the
dragon,
and out
of the
mouth of the
beast,
and out
of the
mouth of the false
prophet.
\VS{14}For they
are the
spirits of
devils,
working
miracles,
{\ADD{which}} go
forth
unto the
kings of the
earth
and of the
whole
world, to
gather
them
to the
battle of
that
great
day of
God
Almighty.
\VS{15}Behold, I
come
as a
thief.
Blessed
{\ADD{is}}
he that
watcheth,
and
keepeth
his
garments,
lest he
walk
naked,
and they
see
his
shame.
\VS{16}And he
gathered
them
together
into a
place
called in the Hebrew
tongue
Armageddon.
\VS{17}And the
seventh
angel poured
out
his
vial
into the
air;
and there
came a
great
voice
out of the
temple of
heaven,
from the
throne,
saying, It is
done.
\VS{18}And there
were
voices,
and
thunders,
and
lightnings;
and there
was a
great
earthquake, such
as
was
not
since
men
were
upon the
earth, so
mighty an
earthquake,
{\ADD{and}}
so
great.
\VS{19}And the
great
city was
divided
into
three
parts,
and the
cities of the
nations
fell:
and
great
Babylon came in
remembrance
before
God, to
give unto
her the
cup of the
wine of the
fierceness of
his
wrath.
\VS{20}And
every
island fled
away,
and the
mountains
were
not
found.
\VS{21}And there
fell
upon
men a
great
hail out
of
heaven,
{\ADD{every stone}}
about the weight of a
talent:
and
men
blasphemed
God because
of the
plague of the
hail;
for the
plague
thereof
was
exceeding
great.

\par }\Chap{17}{\PP \VerseOne{1}And there
came
one
of the
seven
angels which
had the
seven
vials,
and
talked
with
me,
saying unto
me, Come
hither; I will
shew unto
thee the
judgment of the
great
whore that
sitteth
upon
many
waters:
\VS{2}With
whom the
kings of the
earth have committed
fornication,
and the
inhabitants of the
earth have been made
drunk
with the
wine of
her
fornication.
\VS{3}So he
carried
me away
in the
spirit
into the
wilderness:
and I
saw a
woman
sit
upon a scarlet
coloured
beast,
full of
names of
blasphemy,
having
seven
heads
and
ten
horns.
\VS{4}And the
woman was arrayed
in
purple
and scarlet
colour,
and
decked with
gold
and
precious
stones
and
pearls,
having a
golden
cup
in
her
hand
full of
abominations
and
filthiness of
her
fornication:
\VS{5}And
upon
her
forehead
{\ADD{was}} a
name
written,
MYSTERY,
BABYLON THE
GREAT, THE
MOTHER OF
HARLOTS
AND
ABOMINATIONS OF THE
EARTH.
\VS{6}And I
saw the
woman
drunken
with the
blood of the
saints,
and
with the
blood of the
martyrs of
Jesus:
and when I
saw
her, I
wondered with
great
admiration.
\VS{7}And the
angel said
unto
me,
Wherefore didst thou
marvel?
I will
tell
thee the
mystery of the
woman,
and of the
beast that
carrieth
her,
which
hath the
seven
heads
and
ten
horns.
\VS{8}The
beast
that thou
sawest
was,
and
is
not;
and
shall
ascend out
of the bottomless
pit,
and
go
into
perdition:
and they
that
dwell
on the
earth shall
wonder,
whose
names
were
not
written
in the
book of
life
from the
foundation of the
world, when they
behold the
beast
that
was,
and
is
not, and
yet
is.
\VS{9}And
here
{\ADD{is}} the
mind which
hath
wisdom. The
seven
heads
are
seven
mountains,
on
which the
woman
sitteth.
\VS{10}And there
are
seven
kings:
five are
fallen,
and
one
is,
{\ADD{and}} the
other is not
yet
come;
and
when
he
cometh, he
must
continue a short
space.
\VS{11}And the
beast
that
was,
and
is
not,
even
he
is the
eighth,
and
is
of the
seven,
and
goeth
into
perdition.
\VS{12}And the
ten
horns
which thou
sawest
are
ten
kings,
which have
received
no
kingdom as
yet;
but
receive
power
as
kings
one
hour
with the
beast.
\VS{13}These
have
one
mind,
and shall
give
their
power
and
strength unto the
beast.
\VS{14}These shall make
war
with the
Lamb,
and the
Lamb shall
overcome
them:
for he
is
Lord of
lords,
and
King of
kings:
and they that
are
with
him
{\ADD{are}}
called,
and
chosen,
and
faithful.
\VS{15}And he
saith unto
me, The
waters
which thou
sawest,
where the
whore
sitteth,
are
peoples,
and
multitudes,
and
nations,
and
tongues.
\VS{16}And the
ten
horns
which thou
sawest
upon the
beast,
these shall
hate the
whore,
and shall
make
her
desolate
and
naked,
and shall
eat
her
flesh,
and
burn
her
with
fire.
\VS{17}For
God hath
put
in
their
hearts to
fulfil
his
will,
and to
agree,
and
give
their
kingdom unto the
beast,
until the
words of
God shall be
fulfilled.
\VS{18}And the
woman
which thou
sawest
is that
great
city,
which
reigneth
over the
kings of the
earth.

\par }\Chap{18}{\PP \VerseOne{1}And
after these
things I
saw
another
angel come
down
from
heaven,
having
great
power;
and the
earth was
lightened
with
his
glory.
\VS{2}And he
cried
mightily
with a
strong
voice,
saying,
Babylon the
great is
fallen, is
fallen,
and is
become the
habitation of
devils,
and the
hold of
every
foul
spirit,
and a
cage of
every
unclean
and
hateful
bird.
\VS{3}For
all
nations have
drunk
of the
wine of the
wrath of
her
fornication,
and the
kings of the
earth have committed
fornication
with
her,
and the
merchants of the
earth are waxed
rich
through the
abundance of
her
delicacies.
\VS{4}And I
heard
another
voice
from
heaven,
saying,
Come out
of
her,
my
people,
that ye
be
not
partakers of
her
sins,
and
that ye
receive
not
of
her
plagues.
\VS{5}For
her
sins have
reached
unto
heaven,
and
God hath
remembered
her
iniquities.
\VS{6}Reward
her
even
as
she
rewarded
you,
and
double unto
her
double according
to
her
works:
in the
cup
which she hath
filled
fill to
her
double.
\VS{7}How
much she hath
glorified
herself,
and lived
deliciously, so
much
torment
and
sorrow
give
her:
for she
saith
in
her
heart, I
sit a
queen,
and
am
no
widow,
and shall
see
no
sorrow.
\VS{8}Therefore
shall
her
plagues
come
in
one
day,
death,
and
mourning,
and
famine;
and she shall be utterly
burned
with
fire:
for
strong
{\ADD{is}} the
Lord
God who
judgeth
her.
\VS{9}And the
kings of the
earth,
who have committed
fornication
and lived
deliciously
with
her, shall
bewail
her,
and
lament
for
her,
when they shall
see the
smoke of
her
burning,
\VS{10}Standing
afar
off
for the
fear of
her
torment,
saying,
Alas,
alas,
that
great
city
Babylon, that
mighty
city!
for
in
one
hour is
thy
judgment
come.
\VS{11}And the
merchants of the
earth shall
weep
and
mourn
over
her;
for no
man
buyeth
their
merchandise any
more:
\VS{12}The
merchandise of
gold,
and
silver,
and
precious
stones,
and of
pearls,
and fine
linen,
and
purple,
and
silk,
and
scarlet,
and
all
thyine
wood,
and all
manner
vessels of
ivory,
and all
manner
vessels
of most
precious
wood,
and of
brass,
and
iron,
and
marble,
\VS{13}And
cinnamon,
and
odours,
and
ointments,
and
frankincense,
and
wine,
and
oil,
and fine
flour,
and
wheat,
and
beasts,
and
sheep,
and
horses,
and
chariots,
and
slaves,
and
souls of
men.
\VS{14}And the
fruits
that
thy
soul lusted
after are
departed
from
thee,
and all
things which were
dainty
and
goodly are
departed
from
thee,
and thou shalt
find
them no
more at
all.
\VS{15}The
merchants of these
things,
which were made
rich
by
her, shall
stand
afar
off
for the
fear of
her
torment,
weeping
and
wailing,
\VS{16}And
saying,
Alas,
alas, that
great
city, that was
clothed in fine
linen,
and
purple,
and
scarlet,
and
decked
with
gold,
and
precious
stones,
and
pearls!
\VS{17}For in
one
hour so
great
riches is come to
nought.
And
every
shipmaster,
and
all the
company
in
ships,
and
sailors,
and as many
as trade
by
sea,
stood
afar
off,
\VS{18}And
cried when they
saw the
smoke of
her
burning,
saying,
What
{\ADD{city is}} like
unto this
great
city!
\VS{19}And they
cast
dust
on
their
heads,
and
cried,
weeping
and
wailing,
saying,
Alas,
alas,
that
great
city,
wherein were made
rich
all that
had
ships
in the
sea by reason
of
her
costliness!
for in
one
hour is she made
desolate.
\VS{20}Rejoice
over
her,
{\ADD{thou}}
heaven,
and
{\ADD{ye}}
holy
apostles
and
prophets;
for
God
hath
avenged
you
on
her.
\VS{21}And
a
mighty
angel took
up a
stone
like a
great
millstone,
and
cast
{\ADD{it}}
into the
sea,
saying,
Thus with
violence
shall that
great
city
Babylon be thrown
down,
and shall be
found
no
more at
all.
\VS{22}And the
voice of
harpers,
and
musicians,
and of
pipers,
and
trumpeters, shall be
heard
no
more at
all
in
thee;
and
no
craftsman, of
whatsoever
craft
{\ADD{he be}}, shall be
found
any
more
in
thee;
and the
sound of a
millstone shall be
heard
no
more at
all
in
thee;
\VS{23}And the
light of a
candle shall
shine
no
more at
all
in
thee;
and the
voice of the
bridegroom
and of the
bride shall be
heard
no
more at
all
in
thee:
for
thy
merchants
were the great
men of the
earth;
for
by
thy
sorceries
were
all
nations
deceived.
\VS{24}And
in
her was
found the
blood of
prophets,
and of
saints,
and of
all that were
slain
upon the
earth.

\par }\Chap{19}{\PP \VerseOne{1}And
after these
things I
heard a
great
voice of
much
people
in
heaven,
saying,
Alleluia;
Salvation,
and
glory,
and
honour,
and
power, unto the
Lord
our
God:
\VS{2}For
true
and
righteous
{\ADD{are}}
his
judgments:
for he hath
judged the
great
whore,
which did
corrupt the
earth
with
her
fornication,
and hath
avenged the
blood of
his
servants
at
her
hand.
\VS{3}And
again they
said,
Alleluia.
And
her
smoke rose
up
for
ever and
ever.
\VS{4}And the
four
and
twenty
elders
and the
four
beasts fell
down
and
worshipped
God that
sat
on the
throne,
saying,
Amen;
Alleluia.
\VS{5}And a
voice came
out
of the
throne,
saying,
Praise
our
God,
all ye
his
servants,
and ye
that
fear
him,
both
small
and
great.
\VS{6}And I
heard as it
were the
voice of a
great
multitude,
and
as the
voice of
many
waters,
and
as the
voice of
mighty
thunderings,
saying,
Alleluia:
for the
Lord
God
omnipotent
reigneth.
\VS{7}Let us be
glad
and
rejoice,
and
give
honour to
him:
for the
marriage of the
Lamb is
come,
and
his
wife hath
made
herself
ready.
\VS{8}And to
her was
granted
that she should be
arrayed in fine
linen,
clean
and
white:
for the fine
linen
is the
righteousness of
saints.
\VS{9}And he
saith unto
me,
Write,
Blessed
{\ADD{are}} they
which are
called
unto the
marriage
supper of the
Lamb.
And he
saith unto
me,
These
are the
true
sayings of
God.
\VS{10}And I
fell
at
his
feet to
worship
him.
And he
said unto
me,
See
{\ADD{thou do it}}
not: I
am
thy
fellowservant,
and of
thy
brethren that
have the
testimony of
Jesus:
worship
God:
for the
testimony of
Jesus
is the
spirit of
prophecy.
\VS{11}And I
saw
heaven
opened,
and
behold a
white
horse;
and he that
sat
upon
him
{\ADD{was}}
called
Faithful
and
True,
and
in
righteousness he doth
judge
and make
war.
\VS{12}His
eyes
{\ADD{were}}
as a
flame of
fire,
and
on
his
head
{\ADD{were}}
many
crowns;
and he
had a
name
written,
that no
man
knew,
but he
himself.
\VS{13}And
he
{\ADD{was}} clothed
with a
vesture
dipped in
blood:
and
his
name is
called The
Word of
God.
\VS{14}And the
armies
{\ADD{which were}}
in
heaven
followed
him
upon
white
horses, clothed
in fine
linen,
white
and
clean.
\VS{15}And out
of
his
mouth
goeth a
sharp
sword,
that
with
it he should
smite the
nations:
and
he shall
rule
them
with a
rod of
iron:
and
he
treadeth the
winepress of the
fierceness
and
wrath of
Almighty
God.
\VS{16}And he
hath
on
{\ADD{his}}
vesture
and
on
his
thigh a name
written,
KING OF
KINGS,
AND
LORD OF
LORDS.
\par }{\PP \VS{17}And I
saw
an
angel
standing
in the
sun;
and he
cried with a
loud
voice,
saying to
all the
fowls that
fly
in the midst of
heaven,
Come
and gather yourselves
together
unto the
supper of the
great
God;
\VS{18}That ye may
eat the
flesh of
kings,
and the
flesh of
captains,
and the
flesh of mighty
men,
and the
flesh of
horses,
and of
them that
sit
on
them,
and the
flesh of
all
{\ADD{men, both}}
free
and
bond,
both
small
and
great.
\VS{19}And I
saw the
beast,
and the
kings of the
earth,
and
their
armies, gathered
together to
make
war
against
him that
sat
on the
horse,
and
against
his
army.
\VS{20}And the
beast was
taken,
and
with
him the false
prophet that
wrought
miracles
before
him,
with
which he
deceived
them that had
received the
mark of the
beast,
and
them that
worshipped
his
image. These
both were
cast
alive
into a
lake of
fire
burning
with
brimstone.
\VS{21}And the
remnant were
slain
with the
sword of
him that
sat
upon the
horse,
which
{\ADD{sword}}
proceeded out
of
his
mouth:
and
all the
fowls were
filled
with
their
flesh.

\par }\Chap{20}{\PP \VerseOne{1}And I
saw an
angel come
down
from
heaven,
having the
key of the bottomless
pit
and a
great
chain
in
his
hand.
\VS{2}And he laid hold
on the
dragon, that
old
serpent,
which
is the
Devil,
and
Satan,
and
bound
him a
thousand
years,
\VS{3}And
cast
him
into the bottomless
pit,
and
shut
him
up,
and set a
seal
upon
him,
that he should
deceive the
nations
no
more,
till the
thousand
years should be
fulfilled:
and
after
that
he must
be
loosed a
little
season.
\VS{4}And I
saw
thrones,
and they
sat
upon
them,
and
judgment was
given unto
them:
and
{\ADD{I saw}} the
souls of
them that were
beheaded
for the
witness of
Jesus,
and
for the
word of
God,
and
which
had
not
worshipped the
beast,
neither
his
image,
neither had
received
{\ADD{his}}
mark
upon
their
foreheads,
or
in
their
hands;
and they
lived
and
reigned
with
Christ a
thousand
years.
\VS{5}But the
rest of the
dead
lived
not
again
until the
thousand
years were
finished.
This
{\ADD{is}} the
first
resurrection.
\VS{6}Blessed
and
holy
{\ADD{is}} he that
hath
part
in the
first
resurrection:
on
such the
second
death
hath
no
power,
but they shall
be
priests of
God
and of
Christ,
and shall
reign
with
him a
thousand
years.
\VS{7}And
when the
thousand
years are
expired,
Satan shall be
loosed out
of
his
prison,
\VS{8}And shall go
out to
deceive the
nations which
are
in the
four
quarters of the
earth,
Gog
and
Magog, to
gather
them
together
to
battle: the
number of
whom
{\ADD{is}}
as the
sand of the
sea.
\VS{9}And they went
up
on the
breadth of the
earth,
and
compassed the
camp of the
saints
about,
and the
beloved
city:
and
fire came
down
from
God out
of
heaven,
and
devoured
them.
\VS{10}And the
devil that
deceived
them was
cast
into the
lake of
fire
and
brimstone,
where the
beast
and the false
prophet
{\ADD{are}},
and shall be
tormented
day
and
night
for
ever and
ever.
\par }{\PP \VS{11}And I
saw a
great
white
throne,
and
him that
sat
on
it,
from
whose
face the
earth
and the
heaven fled
away;
and there was
found
no
place for
them.
\VS{12}And I
saw the
dead,
small
and
great,
stand
before
God;
and the
books were
opened:
and
another
book was
opened,
which
is
{\ADD{the book}} of
life:
and the
dead were
judged out
of those
things which were
written
in the
books, according
to
their
works.
\VS{13}And the
sea gave
up the
dead which were
in
it;
and
death
and
hell delivered
up the
dead which were
in
them:
and they were
judged every
man according
to
their
works.
\VS{14}And
death
and
hell were
cast
into the
lake of
fire.
This
is the
second
death.
\VS{15}And
whosoever
was
not
found
written
in the
book of
life was
cast
into the
lake of
fire.

\par }\Chap{21}{\PP \VerseOne{1}And I
saw a
new
heaven
and a
new
earth:
for the
first
heaven
and the
first
earth were passed
away;
and there
was
no
more
sea.
\VS{2}And
I
John
saw the
holy
city,
new
Jerusalem, coming
down
from
God out
of
heaven,
prepared
as a
bride
adorned for
her
husband.
\VS{3}And I
heard a
great
voice out
of
heaven
saying,
Behold, the
tabernacle of
God
{\ADD{is}}
with
men,
and he will
dwell
with
them,
and
they shall
be
his
people,
and
God
himself shall
be
with
them,
{\ADD{and be}}
their
God.
\VS{4}And
God shall wipe
away
all
tears
from
their
eyes;
and there shall
be
no
more
death,
neither
sorrow,
nor
crying,
neither shall there
be any
more
pain:
for the former
things are passed
away.
\VS{5}And
he that
sat
upon the
throne
said,
Behold, I
make all
things
new.
And he
said unto
me,
Write:
for
these
words
are
true
and
faithful.
\VS{6}And he
said unto
me, It is
done.
I
am
Alpha
and
Omega, the
beginning
and the
end.
I will
give unto him that is
athirst
of the
fountain of the
water of
life
freely.
\VS{7}He that
overcometh shall
inherit all
things;
and I will
be
his
God,
and
he shall
be
my
son.
\VS{8}But the
fearful,
and
unbelieving,
and the
abominable,
and
murderers,
and
whoremongers,
and
sorcerers,
and
idolaters,
and
all
liars, shall have
their
part
in the
lake
which
burneth with
fire
and
brimstone:
which
is the
second
death.
\par }{\PP \VS{9}And there
came
unto
me
one of the
seven
angels which
had the
seven
vials
full of the
seven
last
plagues,
and
talked
with
me,
saying, Come
hither, I will
shew
thee the
bride, the
Lamb’s
wife.
\VS{10}And he
carried
me
away
in the
spirit
to a
great
and
high
mountain,
and
shewed
me that
great
city, the
holy
Jerusalem,
descending out
of
heaven
from
God,
\VS{11}Having the
glory of
God:
and
her
light
{\ADD{was}} like
unto a
stone most
precious, even
like a
jasper
stone, clear as
crystal;
\VS{12}And
had a
wall
great
and
high,
{\ADD{and}}
had
twelve
gates,
and
at the
gates
twelve
angels,
and
names written
thereon,
which
are
{\ADD{the names}} of the
twelve
tribes of the
children of
Israel:
\VS{13}On the
east
three
gates;
on the
north
three
gates;
on the
south
three
gates; and
on the
west
three
gates.
\VS{14}And the
wall of the
city
had
twelve
foundations,
and
in
them the
names of the
twelve
apostles of the
Lamb.
\VS{15}And
he that
talked
with
me
had a
golden
reed
to
measure the
city,
and the
gates
thereof,
and the
wall
thereof.
\VS{16}And the
city
lieth
foursquare,
and the
length
is as
large
as the
breadth:
and he
measured the
city with the
reed,
twelve
thousand
furlongs. The
length
and the
breadth
and the
height of
it
are
equal.
\VS{17}And he
measured the
wall
thereof, an
hundred
{\ADD{and}}
forty
{\ADD{and}}
four
cubits,
{\ADD{according to}} the
measure of a
man,
that
is, of the
angel.
\VS{18}And the
building of the
wall of
it
was
{\ADD{of}}
jasper:
and the
city
{\ADD{was}}
pure
gold, like
unto
clear
glass.
\VS{19}And the
foundations of the
wall of the
city
{\ADD{were}}
garnished with all
manner of
precious
stones. The
first
foundation
{\ADD{was}}
jasper; the
second,
sapphire; the
third, a
chalcedony; the
fourth, an
emerald;
\VS{20}The
fifth,
sardonyx; the
sixth,
sardius; the
seventh,
chrysolite; the
eighth,
beryl; the
ninth, a
topaz; the
tenth, a
chrysoprasus; the
eleventh, a
jacinth; the
twelfth, an
amethyst.
\VS{21}And the
twelve
gates
{\ADD{were}}
twelve
pearls;
every
several
gate
was
of
one
pearl:
and the
street of the
city
{\ADD{was}}
pure
gold, as it
were
transparent
glass.
\VS{22}And I
saw
no
temple
therein:
for the
Lord
God
Almighty
and the
Lamb
are the
temple of
it.
\VS{23}And the
city
had
no
need of the
sun,
neither of the
moon,
to
shine
in
it:
for the
glory of
God did
lighten
it,
and the
Lamb
{\ADD{is}} the
light
thereof.
\VS{24}And the
nations of
them which are
saved shall
walk
in the
light of
it:
and the
kings of the
earth do
bring
their
glory
and
honour
into
it.
\VS{25}And the
gates of
it
shall
not be
shut at
all by
day:
for there shall
be
no
night
there.
\VS{26}And they shall
bring the
glory
and
honour of the
nations
into
it.
\VS{27}And there
shall in no
wise
enter
into
it any
thing that
defileth,
neither
{\ADD{whatsoever}}
worketh
abomination,
or
{\ADD{maketh}} a
lie:
but
they which are
written
in the
Lamb’s
book of
life.

\par }\Chap{22}{\PP \VerseOne{1}And he
shewed
me a
pure
river of
water of
life,
clear
as
crystal,
proceeding out
of the
throne of
God
and of the
Lamb.
\VS{2}In the
midst of the
street of
it,
and on either
side of the
river,
{\ADD{was there}} the
tree of
life, which
bare
twelve
{\ADD{manner of}}
fruits,
{\ADD{and}}
yielded
her
fruit
every
month:
and the
leaves of the
tree
{\ADD{were}}
for the
healing of the
nations.
\VS{3}And there shall
be
no
more
curse:
but the
throne of
God
and of the
Lamb shall
be
in
it;
and
his
servants shall
serve
him:
\VS{4}And they shall
see
his
face;
and
his
name
{\ADD{shall be}}
in
their
foreheads.
\VS{5}And there shall
be
no
night
there;
and
they
need
no
candle,
neither
light of the
sun;
for the
Lord
God
giveth
them
light:
and they shall
reign
for
ever and
ever.
\par }{\PP \VS{6}And he
said unto
me,
These
sayings
{\ADD{are}}
faithful
and
true:
and the
Lord
God of the
holy
prophets
sent
his
angel to
shew unto
his
servants the things
which
must
shortly be
done.
\VS{7}{\WJ{Behold,
I come
quickly:
blessed
{\ADD{is}}
he
that keepeth
the sayings
of the prophecy
of this
book.}}
\VS{8}And
I
John
saw these
things,
and
heard
{\ADD{them}}.
And
when I had
heard
and
seen, I fell
down to
worship
before the
feet of the
angel which
shewed
me these
things.
\VS{9}Then saith
he unto
me,
See
{\ADD{thou do it}}
not:
for I
am
thy
fellowservant,
and of
thy
brethren the
prophets,
and of
them which
keep the
sayings of
this
book:
worship
God.
\VS{10}And he
saith unto
me,
Seal
not the
sayings of the
prophecy of
this
book:
for the
time
is at
hand.
\VS{11}He that is
unjust, let him be
unjust
still:
and he which is
filthy, let him be
filthy
still:
and he that is
righteous, let him be
righteous
still:
and he that is
holy, let him be
holy
still.
\VS{12}{\WJ{And,
behold,
I come
quickly;
and
my
reward
{\ADD{is}}
with
me,
to give
every man
according as
his
work
shall be.}}
\VS{13}{\WJ{I
am
Alpha
and
Omega,
the beginning
and
the end,
the first
and
the last.}}
\VS{14}Blessed
{\ADD{are}}
they that
do
his
commandments,
that
they may
have
right
to the
tree of
life,
and may enter
in through the
gates
into the
city.
\VS{15}For
without
{\ADD{are}}
dogs,
and
sorcerers,
and
whoremongers,
and
murderers,
and
idolaters,
and
whosoever
loveth
and
maketh a
lie.
\VS{16}{\WJ{I
Jesus
have sent
mine
angel
to testify
unto you
these things
in
the churches.
I
am
the root
and
the offspring
of David,
{\ADD{and}}
the bright
and
morning
star.}}
\VS{17}And the
Spirit
and the
bride
say,
Come.
And let him that
heareth
say,
Come.
And let him that is
athirst
come.
And whosoever
will, let him
take the
water of
life
freely.
\par }{\PP \VS{18}For I
testify unto every
man that
heareth the
words of the
prophecy of
this
book,
If any
man shall
add
unto these
things,
God shall
add
unto
him the
plagues that are
written
in
this
book:
\VS{19}And
if any
man shall take
away
from the
words of the
book of
this
prophecy,
God shall take
away
his
part
out of the
book of
life,
and out
of the
holy
city,
and
{\ADD{from}} the
things which are
written
in
this
book.
\VS{20}He
which
testifieth these
things
saith,
{\WJ{Surely
I come
quickly.}}
Amen. Even
so,
come,
Lord
Jesus.
\VS{21}The
grace of
our
Lord
Jesus
Christ
{\ADD{be}}
with
you
all.
Amen.
\par }