\NormalFont\ShortTitle{Song of Solomon}
{\MT The Song of Solomon

\par }\ChapOne{1}{\PP \VerseOne{1}The
song of
songs, which
{\ADD{is}}
Solomon’s.
\par }{\PP \VS{2}Let him
kiss me with the
kisses of his
mouth: for thy
love
{\ADD{is}}
better than
wine.\FTNT{*}{{\FR 1.2 }
thy…: Heb. thy loves}
\VS{3}Because of the
savour of thy
good
ointments thy
name
{\ADD{is as}}
ointment poured
forth, therefore do the
virgins
love thee.
\VS{4}Draw me, we will
run
after thee: the
king hath
brought me into his
chambers: we will be
glad and
rejoice in thee, we will
remember thy
love more than
wine: the
upright
love thee.\FTNT{†}{{\FR 1.4 }
the upright…: or, they love thee uprightly}
\VS{5}I
{\ADD{am}}
black, but
comely, O ye
daughters of
Jerusalem, as the
tents of
Kedar, as the
curtains of
Solomon.
\VS{6}Look not upon me, because I
{\ADD{am}}
black, because the
sun hath
looked upon me: my
mother’s
children were
angry with me; they
made me the
keeper of the
vineyards;
{\ADD{but}} mine own
vineyard have I not
kept.
\par }{\PP \VS{7}Tell me, O thou whom my
soul
loveth, where thou
feedest,
where thou makest
{\ADD{thy flock}} to
rest at
noon:
for why should I be as one that turneth
aside by the
flocks of thy
companions?\FTNT{‡}{{\FR 1.7 }
as one…: or, as one that is veiled}
\par }{\PP \VS{8}¶ If thou
know not, O thou
fairest among
women, go thy way
forth by the
footsteps of the
flock, and
feed thy
kids beside the
shepherds’
tents.
\VS{9}I have
compared thee, O my
love, to a company of
horses in
Pharaoh’s
chariots.
\VS{10}Thy
cheeks are
comely with
rows
{\ADD{of jewels}}, thy
neck with
chains
{\ADD{of gold}}.
\VS{11}We will
make thee
borders of
gold with
studs of
silver.
\par }{\PP \VS{12}¶ While the
king
{\ADD{sitteth}} at his
table, my
spikenard sendeth
forth the
smell thereof.
\VS{13}A
bundle of
myrrh
{\ADD{is}} my
wellbeloved unto me; he shall
lie all night betwixt my
breasts.
\VS{14}My
beloved
{\ADD{is}} unto me
{\ADD{as}} a
cluster of
camphire in the
vineyards of
En-gedi.\FTNT{§}{{\FR 1.14 }
camphire: or, cypress}
\VS{15}Behold, thou
{\ADD{art}}
fair, my
love; behold, thou
{\ADD{art}}
fair; thou
{\ADD{hast}}
doves’
eyes.\FTNT{**}{{\FR 1.15 }
my love: or, my companion}
\VS{16}Behold, thou
{\ADD{art}}
fair, my
beloved, yea,
pleasant: also our
bed
{\ADD{is}}
green.
\VS{17}The
beams of our
house
{\ADD{are}}
cedar,
{\ADD{and}} our
rafters of
fir.\FTNT{††}{{\FR 1.17 }
rafters: or, galleries}

\par }\Chap{2}{\PP \VerseOne{1}I
{\ADD{am}} the
rose of
Sharon,
{\ADD{and}} the
lily of the
valleys.
\VS{2}As the
lily among
thorns, so
{\ADD{is}} my
love among the
daughters.
\par }{\PP \VS{3}As the apple
tree among the
trees of the
wood, so
{\ADD{is}} my
beloved among the
sons. I
sat down under his
shadow with great
delight, and his
fruit
{\ADD{was}}
sweet to my
taste.\FTNT{*}{{\FR 2.3 }
I sat…: Heb. I delighted and sat down, etc}\FTNT{†}{{\FR 2.3 }
taste: Heb. palate}
\VS{4}He
brought me to the
banqueting
house, and his
banner over me
{\ADD{was}}
love.\FTNT{‡}{{\FR 2.4 }
banqueting…: Heb. house of wine}
\VS{5}Stay me with
flagons,
comfort me with
apples: for I
{\ADD{am}}
sick of
love.\FTNT{§}{{\FR 2.5 }
comfort…: Heb. straw me with apples}
\VS{6}His left
hand
{\ADD{is}} under my
head, and his right
hand doth
embrace me.
\VS{7}I
charge you, O ye
daughters of
Jerusalem, by the
roes,
and by the
hinds of the
field, that ye stir not
up, nor
awake
{\ADD{my}}
love, till he
please.\FTNT{**}{{\FR 2.7 }
I charge…: Heb. I adjure you}
\par }{\PP \VS{8}¶ The
voice of my
beloved! behold, he
cometh
leaping upon the
mountains,
skipping upon the
hills.
\VS{9}My
beloved is
like a
roe or a
young
hart: behold, he
standeth
behind our
wall, he looketh
forth at the
windows,
shewing himself through the
lattice.\FTNT{††}{{\FR 2.9 }
shewing…: Heb. flourishing}
\VS{10}My
beloved
spake, and
said unto me, Rise
up, my
love, my fair
one, and come
away.
\VS{11}For, lo, the
winter is
past, the
rain is
over
{\ADD{and}}
gone;
\VS{12}The
flowers
appear on the
earth; the
time of the
singing
{\ADD{of birds}} is
come, and the
voice of the
turtle is
heard in our
land;
\VS{13}The fig
tree putteth
forth her green
figs, and the
vines
{\ADD{with}} the tender
grape
give a
{\ADD{good}}
smell.
Arise, my
love, my fair
one, and come
away.
\par }{\PP \VS{14}¶ O my
dove,
{\ADD{that art}} in the
clefts of the
rock, in the
secret
{\ADD{places}} of the
stairs, let me
see thy
countenance, let me
hear thy
voice; for
sweet
{\ADD{is}} thy
voice, and thy
countenance
{\ADD{is}}
comely.
\VS{15}Take us the
foxes, the
little
foxes, that
spoil the
vines: for our
vines
{\ADD{have}} tender
grapes.
\par }{\PP \VS{16}¶ My
beloved
{\ADD{is}} mine, and I
{\ADD{am}} his: he
feedeth among the
lilies.
\VS{17}Until the
day
break, and the
shadows flee
away,
turn, my
beloved, and be thou
like a
roe or a
young
hart upon the
mountains of
Bether.\FTNT{‡‡}{{\FR 2.17 }
of Bether: or, of division}

\par }\Chap{3}{\PP \VerseOne{1}By
night on my
bed I
sought him whom my
soul
loveth: I
sought him, but I
found him not.
\VS{2}I will
rise now, and go
about the
city in the
streets, and in the broad
ways I will
seek him whom my
soul
loveth: I
sought him, but I
found him not.
\VS{3}The
watchmen that go
about the
city
found me:
{\ADD{to whom I said}},
Saw ye him whom my
soul
loveth?
\VS{4}{\ADD{It was}} but a
little that I
passed from them, but I
found him whom my
soul
loveth: I
held him, and would not let him
go, until I had
brought him into my
mother’s
house, and into the
chamber of her that
conceived me.
\VS{5}I
charge you, O ye
daughters of
Jerusalem, by the
roes,
and by the
hinds of the
field, that ye stir not
up, nor
awake
{\ADD{my}}
love, till he
please.
\par }{\PP \VS{6}¶ Who
{\ADD{is}} this that
cometh out of the
wilderness like
pillars of
smoke,
perfumed with
myrrh and
frankincense, with all
powders of the
merchant?
\par }{\PP \VS{7}Behold his
bed, which
{\ADD{is}}
Solomon’s;
threescore
valiant men
{\ADD{are}}
about it, of the
valiant of
Israel.
\VS{8}They all
hold
swords,
{\ADD{being}}
expert in
war: every
man
{\ADD{hath}} his
sword upon his
thigh because of
fear in the
night.
\VS{9}King
Solomon
made himself a
chariot of the
wood of
Lebanon.\FTNT{*}{{\FR 3.9 }
a chariot: or, a bed}
\VS{10}He
made the
pillars thereof
{\ADD{of}}
silver, the
bottom thereof
{\ADD{of}}
gold, the
covering of it
{\ADD{of}}
purple, the
midst thereof being
paved
{\ADD{with}}
love, for the
daughters of
Jerusalem.
\VS{11}Go
forth, O ye
daughters of
Zion, and
behold
king
Solomon with the
crown wherewith his
mother
crowned him in the
day of his
espousals, and in the
day of the
gladness of his
heart.

\par }\Chap{4}{\PP \VerseOne{1}Behold, thou
{\ADD{art}}
fair, my
love; behold, thou
{\ADD{art}}
fair; thou
{\ADD{hast}}
doves’
eyes
within thy
locks: thy
hair
{\ADD{is}} as a
flock of
goats, that
appear from
mount
Gilead.\FTNT{*}{{\FR 4.1 }
that…: or, that eat of, etc}
\VS{2}Thy
teeth
{\ADD{are}} like a
flock
{\ADD{of sheep that are even}}
shorn, which came
up from the
washing; whereof every one bear
twins, and none
{\ADD{is}}
barren among them.
\VS{3}Thy
lips
{\ADD{are}} like a
thread of
scarlet, and thy
speech
{\ADD{is}}
comely: thy
temples
{\ADD{are}} like a
piece of a
pomegranate
within thy
locks.
\VS{4}Thy
neck
{\ADD{is}} like the
tower of
David
builded for an
armoury, whereon there
hang a
thousand
bucklers, all
shields of mighty
men.
\VS{5}Thy
two
breasts
{\ADD{are}} like
two
young
roes that are
twins, which
feed among the
lilies.
\VS{6}Until the
day
break, and the
shadows flee
away, I will
get me to the
mountain of
myrrh, and to the
hill of
frankincense.\FTNT{†}{{\FR 4.6 }
break: Heb. breathe}
\VS{7}Thou
{\ADD{art}} all
fair, my
love;
{\ADD{there is}} no
spot in thee.
\par }{\PP \VS{8}¶
Come with me from
Lebanon,
{\ADD{my}}
spouse, with me from
Lebanon:
look from the
top of
Amana, from the
top of
Shenir and
Hermon, from the
lions’
dens, from the
mountains of the
leopards.
\VS{9}Thou hast ravished my
heart, my
sister,
{\ADD{my}}
spouse; thou hast ravished my
heart with
one of thine
eyes, with
one
chain of thy
neck.\FTNT{‡}{{\FR 4.9 }
ravished: or, taken away}
\VS{10}How
fair is thy
love, my
sister,
{\ADD{my}}
spouse! how much
better is thy
love than
wine! and the
smell of thine
ointments than all
spices!
\VS{11}Thy
lips, O
{\ADD{my}}
spouse,
drop
{\ADD{as}} the
honeycomb:
honey and
milk
{\ADD{are}} under thy
tongue; and the
smell of thy
garments
{\ADD{is}} like the
smell of
Lebanon.
\VS{12}A
garden
inclosed
{\ADD{is}} my
sister,
{\ADD{my}}
spouse; a
spring shut
up, a
fountain
sealed.
\FTNT{§}{{\FR 4.12 }
inclosed: Heb. barred}\FTNT{**}{{\FR 4.12 }
shut up: Heb. barred}
\VS{13}Thy
plants
{\ADD{are}} an
orchard of
pomegranates, with
pleasant
fruits;
camphire, with
spikenard,\FTNT{††}{{\FR 4.13 }
camphire: or, cypress}
\VS{14}Spikenard and
saffron;
calamus and
cinnamon, with all
trees of
frankincense;
myrrh and
aloes, with all the
chief
spices:
\par }{\PP \VS{15}A
fountain of
gardens, a
well of
living
waters, and
streams from
Lebanon.
\par }{\PP \VS{16}¶
Awake, O north
wind; and
come, thou
south;
blow upon my
garden,
{\ADD{that}} the
spices thereof may flow
out. Let my
beloved
come into his
garden, and
eat his
pleasant
fruits.

\par }\Chap{5}{\PP \VerseOne{1}I am
come into my
garden, my
sister,
{\ADD{my}}
spouse: I have
gathered my
myrrh with my
spice; I have
eaten my
honeycomb with my
honey; I have
drunk my
wine with my
milk:
eat, O
friends;
drink, yea, drink
abundantly, O
beloved.\FTNT{*}{{\FR 5.1 }
yea…: or, and be drunken with loves}
\par }{\PP \VS{2}¶ I
sleep, but my
heart
waketh:
{\ADD{it is}} the
voice of my
beloved that
knocketh,
{\ADD{saying}},
Open to me, my
sister, my
love, my
dove, my
undefiled: for my
head is
filled with
dew,
{\ADD{and}} my
locks with the
drops of the
night.
\VS{3}I have put
off my
coat;
how shall I put it
on? I have
washed my
feet;
how shall I
defile them?
\VS{4}My
beloved put
in his
hand
by the
hole
{\ADD{of the door}}, and my
bowels were
moved for him.\FTNT{†}{{\FR 5.4 }
for him: or, (as some read) in me}
\VS{5}I rose
up to
open to my
beloved; and my
hands
dropped
{\ADD{with}}
myrrh, and my
fingers
{\ADD{with}} sweet
smelling
myrrh, upon the
handles of the
lock.\FTNT{‡}{{\FR 5.5 }
sweet…: Heb. passing, or, running about}
\VS{6}I
opened to my
beloved; but my
beloved had
withdrawn himself,
{\ADD{and}} was
gone: my
soul
failed when he
spake: I
sought him, but I could not
find him; I
called him, but he gave me no
answer.
\VS{7}The
watchmen that went
about the
city
found me, they
smote me, they
wounded me; the
keepers of the
walls took
away my
veil from me.
\VS{8}I
charge you, O
daughters of
Jerusalem, if ye
find my
beloved, that ye
tell him, that I
{\ADD{am}}
sick of
love.\FTNT{§}{{\FR 5.8 }
that ye: Heb. what, etc}
\par }{\PP \VS{9}¶ What
{\ADD{is}} thy
beloved more than
{\ADD{another}}
beloved, O thou
fairest among
women? what
{\ADD{is}} thy
beloved more than
{\ADD{another}}
beloved, that
thou dost so
charge us?
\VS{10}My
beloved
{\ADD{is}}
white and
ruddy, the
chiefest among ten
thousand.\FTNT{**}{{\FR 5.10 }
the chiefest: Heb. a standard-bearer}
\VS{11}His
head
{\ADD{is as}} the
most fine
gold, his
locks
{\ADD{are}}
bushy,
{\ADD{and}}
black as a
raven.\FTNT{††}{{\FR 5.11 }
bushy: or, curled}
\VS{12}His
eyes
{\ADD{are}} as
{\ADD{the eyes}} of
doves by the
rivers of
waters,
washed with
milk,
{\ADD{and}} fitly
set.\FTNT{‡‡}{{\FR 5.12 }
fitly…: Heb. sitting in fulness, that is, fitly placed, and set as a precious stone in the foil of a ring}
\VS{13}His
cheeks
{\ADD{are}} as a
bed of
spices,
{\ADD{as}}
sweet
flowers: his
lips
{\ADD{like}}
lilies,
dropping sweet
smelling
myrrh.\FTNT{§§}{{\FR 5.13 }
sweet flowers: or, towers of perfumes}
\VS{14}His
hands
{\ADD{are as}}
gold
rings
set with the
beryl: his
belly
{\ADD{is as}}
bright
ivory
overlaid
{\ADD{with}}
sapphires.
\VS{15}His
legs
{\ADD{are as}}
pillars of
marble,
set upon
sockets of fine
gold: his
countenance
{\ADD{is}} as
Lebanon,
excellent as the
cedars.
\VS{16}His
mouth
{\ADD{is}} most
sweet: yea, he
{\ADD{is}} altogether
lovely. This
{\ADD{is}} my
beloved, and this
{\ADD{is}} my
friend, O
daughters of
Jerusalem.\FTNT{***}{{\FR 5.16 }
mouth: Heb. palate}

\par }\Chap{6}{\PP \VerseOne{1}Whither is thy
beloved
gone, O thou
fairest among
women? whither is thy
beloved turned
aside? that we may
seek him with thee.
\VS{2}My
beloved is gone
down into his
garden, to the
beds of
spices, to
feed in the
gardens, and to
gather
lilies.
\VS{3}I
{\ADD{am}} my
beloved’s, and my
beloved
{\ADD{is}} mine: he
feedeth among the
lilies.
\par }{\PP \VS{4}¶ Thou
{\ADD{art}}
beautiful, O my
love, as
Tirzah,
comely as
Jerusalem,
terrible as
{\ADD{an army}} with
banners.
\VS{5}Turn
away thine
eyes from me, for
they have
overcome me: thy
hair
{\ADD{is}} as a
flock of
goats that
appear from
Gilead.\FTNT{*}{{\FR 6.5 }
overcome…: or, puffed me up}
\VS{6}Thy
teeth
{\ADD{are}} as a
flock of
sheep which go
up from the
washing, whereof every one beareth
twins, and
{\ADD{there is}} not one
barren among them.
\VS{7}As a
piece of a
pomegranate
{\ADD{are}} thy
temples
within thy
locks.
\VS{8}There are
threescore
queens, and
fourscore
concubines, and
virgins without
number.
\VS{9}My
dove, my
undefiled is
{\ADD{but}}
one; she
{\ADD{is}} the
{\ADD{only}}
one of her
mother, she
{\ADD{is}} the
choice
{\ADD{one}} of her that
bare her. The
daughters
saw her, and
blessed her;
{\ADD{yea}}, the
queens and the
concubines, and they
praised her.
\par }{\PP \VS{10}¶ Who
{\ADD{is}} she
{\ADD{that}} looketh
forth as the
morning,
fair as the
moon,
clear as the
sun,
{\ADD{and}}
terrible as
{\ADD{an army}} with
banners?
\par }{\PP \VS{11}I went
down into the
garden of
nuts to
see the
fruits of the
valley,
{\ADD{and}} to
see whether the
vine
flourished,
{\ADD{and}} the
pomegranates
budded.
\VS{12}Or ever I was
aware, my
soul
made me
{\ADD{like}} the
chariots of
Amminadib.\FTNT{†}{{\FR 6.12 }
Or ever…: Heb. I knew not}\FTNT{‡}{{\FR 6.12 }
made…: or, set me on the chariots of my willing people}
\VS{13}Return,
return, O
Shulamite;
return,
return, that we may
look upon thee. What will ye
see in the
Shulamite? As it were the
company of two
armies.\FTNT{§}{{\FR 6.13 }
of…: or, of Mahanaim}

\par }\Chap{7}{\PP \VerseOne{1}How
beautiful are thy
feet with
shoes, O
prince’s
daughter! the
joints of thy
thighs
{\ADD{are}} like
jewels, the
work of the
hands of a cunning
workman.
\VS{2}Thy
navel
{\ADD{is like}} a
round
goblet,
{\ADD{which}}
wanteth not
liquor: thy
belly
{\ADD{is like}} an
heap of
wheat set
about with
lilies.\FTNT{*}{{\FR 7.2 }
liquor: Heb. mixture}
\VS{3}Thy
two
breasts
{\ADD{are}} like
two
young
roes
{\ADD{that are}}
twins.
\VS{4}Thy
neck
{\ADD{is}} as a
tower of
ivory; thine
eyes
{\ADD{like}} the
fishpools in
Heshbon, by the
gate of
Bath-rabbim: thy
nose
{\ADD{is}} as the
tower of
Lebanon which
looketh
toward
Damascus.
\VS{5}Thine
head upon thee
{\ADD{is}} like
Carmel, and the
hair of thine
head like
purple; the
king
{\ADD{is}}
held in the
galleries.\FTNT{†}{{\FR 7.5 }
Carmel: or, crimson}\FTNT{‡}{{\FR 7.5 }
held: Heb. bound}
\VS{6}How
fair and how
pleasant art thou, O
love, for
delights!
\VS{7}This thy
stature is
like to a palm
tree, and thy
breasts to
clusters
{\ADD{of grapes}}.
\VS{8}I
said, I will go
up to the palm
tree, I will take
hold of the
boughs thereof: now also thy
breasts shall be as
clusters of the
vine, and the
smell of thy
nose like
apples;
\VS{9}And the roof of thy
mouth like the
best
wine for my
beloved, that
goeth
{\ADD{down}}
sweetly, causing the
lips of those that are
asleep to
speak.\FTNT{§}{{\FR 7.9 }
sweetly: Heb. straightly}\FTNT{**}{{\FR 7.9 }
of those…: or, of the ancient}
\par }{\PP \VS{10}¶ I
{\ADD{am}} my
beloved’s, and his
desire
{\ADD{is}} toward me.
\VS{11}Come, my
beloved, let us go
forth into the
field; let us
lodge in the
villages.
\VS{12}Let us get up
early to the
vineyards; let us
see if the
vine
flourish,
{\ADD{whether}} the tender
grape
appear,
{\ADD{and}} the
pomegranates bud
forth: there will I
give thee my
loves.\FTNT{††}{{\FR 7.12 }
appear: Heb. open}
\VS{13}The
mandrakes
give a
smell, and at our
gates
{\ADD{are}} all manner of
pleasant
{\ADD{fruits}},
new and
old,
{\ADD{which}} I have laid
up for thee, O my
beloved.

\par }\Chap{8}{\PP \VerseOne{1}O
that thou
{\ADD{wert}} as my
brother, that
sucked the
breasts of my
mother!
{\ADD{when}} I should
find thee
without, I would
kiss thee; yea, I should not be
despised.\FTNT{*}{{\FR 8.1 }
I should not…: Heb. they should not despise me}
\VS{2}I would
lead thee,
{\ADD{and}}
bring thee into my
mother’s
house,
{\ADD{who}} would
instruct me: I would cause thee to
drink of
spiced
wine of the
juice of my
pomegranate.
\VS{3}His left
hand
{\ADD{should be}} under my
head, and his right
hand should
embrace me.
\VS{4}I
charge you, O
daughters of
Jerusalem, that ye stir not
up, nor
awake
{\ADD{my}}
love, until he
please.\FTNT{†}{{\FR 8.4 }
that…: Heb. why should ye stir up, or, why, etc}
\par }{\PP \VS{5}Who
{\ADD{is}} this that cometh
up from the
wilderness,
leaning upon her
beloved? I
raised thee up under the apple
tree: there thy
mother brought thee
forth: there she brought thee
forth
{\ADD{that}}
bare thee.
\par }{\PP \VS{6}¶
Set me as a
seal upon thine
heart, as a
seal upon thine
arm: for
love
{\ADD{is}}
strong as
death;
jealousy
{\ADD{is}}
cruel as the
grave: the
coals thereof
{\ADD{are}}
coals of
fire,
{\ADD{which hath a}} most vehement
flame.\FTNT{‡}{{\FR 8.6 }
cruel: Heb. hard}
\VS{7}Many
waters
cannot
quench
love, neither can the
floods
drown it: if
{\ADD{a}}
man would
give all the
substance of his
house for
love, it would
utterly be
contemned.
\par }{\PP \VS{8}¶ We have a
little
sister, and she hath no
breasts: what shall we
do for our
sister in the
day when she shall be spoken
for?
\VS{9}If she
{\ADD{be}} a
wall, we will
build upon her a
palace of
silver: and if she
{\ADD{be}} a
door, we will
inclose her with
boards of
cedar.
\VS{10}I
{\ADD{am}} a
wall, and my
breasts like
towers: then was I in his
eyes as one that
found
favour.\FTNT{§}{{\FR 8.10 }
favour: Heb. peace}
\VS{11}Solomon had a
vineyard at
Baal-hamon; he let
out the
vineyard unto
keepers; every
one for the
fruit thereof was to
bring a
thousand
{\ADD{pieces}} of
silver.
\VS{12}My
vineyard, which
{\ADD{is}} mine,
{\ADD{is}}
before me: thou, O
Solomon,
{\ADD{must have}} a
thousand, and those that
keep the
fruit thereof two
hundred.
\par }{\PP \VS{13}Thou that
dwellest in the
gardens, the
companions
hearken to thy
voice: cause me to
hear
{\ADD{it}}.
\par }{\PP \VS{14}¶ Make
haste, my
beloved, and be thou
like to a
roe or to a
young
hart upon the
mountains of
spices.\FTNT{**}{{\FR 8.14 }
Make…: Heb. Flee away}
\par }