\NormalFont\ShortTitle{Titus}
{\MT THE EPISTLE OF PAUL THE APOSTLE TO TITUS

\par }\ChapOne{1}{\PP \VerseOne{1}Paul, a
servant of
God,
and an
apostle of
Jesus
Christ, according
to the
faith
of
God’s
elect,
and the
acknowledging of the
truth
which is
after
godliness;
\VS{2}In
hope of
eternal
life,
which
God, that cannot
lie,
promised
before the
world
began;
\VS{3}But
hath in
due
times
manifested
his
word
through
preaching,
which is committed
unto
me according
to the
commandment of
God
our
Saviour;
\VS{4}To
Titus,
{\ADD{mine}}
own
son
after the
common
faith:
Grace,
mercy,
{\ADD{and}}
peace,
from
God the
Father
and the
Lord
Jesus
Christ
our
Saviour.
\par }{\PP \VS{5}For
this
cause left
I
thee
in
Crete,
that thou shouldest set in
order the
things that are
wanting,
and
ordain
elders in
every
city,
as
I had
appointed
thee:
\VS{6}If
any
be
blameless, the
husband of
one
wife,
having
faithful
children
not
accused of
riot
or
unruly.
\VS{7}For a
bishop
must
be
blameless,
as the
steward of
God;
not
selfwilled,
not soon
angry,
not given to
wine,
no
striker,
not given to filthy
lucre;
\VS{8}But a lover of
hospitality, a lover of good
men,
sober,
just,
holy,
temperate;
\VS{9}Holding
fast the
faithful
word
as he hath been
taught,
that he may
be
able
by
sound
doctrine
both to
exhort
and to
convince the
gainsayers.
\VS{10}For there
are
many
unruly
and vain
talkers
and
deceivers,
specially
they
of the
circumcision:
\VS{11}Whose
mouths
must be
stopped,
who
subvert
whole
houses,
teaching things
which they
ought
not,
for
filthy
lucre’s
sake.
\VS{12}One
of
themselves,
{\ADD{even}} a
prophet of
their
own,
said, The
Cretians
{\ADD{are}}
alway
liars,
evil
beasts,
slow
bellies.
\VS{13}This
witness
is
true.
Wherefore
rebuke
them
sharply,
that they may be
sound
in the
faith;
\VS{14}Not giving heed
to
Jewish
fables,
and
commandments of
men, that turn
from the
truth.
\VS{15}Unto the
pure all
things
{\ADD{are}}
pure:
but unto them that are
defiled
and
unbelieving
{\ADD{is}}
nothing
pure;
but
even
their
mind
and
conscience is
defiled.
\VS{16}They
profess that they
know
God;
but in
works they
deny
{\ADD{him}},
being
abominable,
and
disobedient,
and
unto
every
good
work
reprobate.

\par }\Chap{2}{\PP \VerseOne{1}But
speak
thou the things
which
become
sound
doctrine:
\VS{2}That the aged
men
be
sober,
grave,
temperate,
sound in
faith, in
charity, in
patience.
\VS{3}The aged
women
likewise,
that
{\ADD{they be}}
in
behaviour as becometh
holiness,
not false
accusers,
not
given to
much
wine, teachers of good
things;
\VS{4}That they may
teach the young
women to be
sober, to love their
husbands, to love their
children,
\VS{5}{\ADD{To be}}
discreet,
chaste, keepers at
home,
good,
obedient to their
own
husbands,
that the
word of
God
be
not
blasphemed.
\VS{6}Young
men
likewise
exhort to be sober
minded.
\VS{7}In all
things
shewing
thyself a
pattern of
good
works:
in
doctrine
{\ADD{shewing}}
uncorruptness,
gravity,
sincerity,
\VS{8}Sound
speech, that cannot be
condemned;
that he that
is
of the contrary
part may be
ashamed,
having
no evil
thing to
say
of
you.
\VS{9}{\ADD{Exhort}}
servants to
be
obedient unto their
own
masters,
{\ADD{and}} to
please
{\ADD{them}}
well
in
all
{\ADD{things}};
not answering
again;
\VS{10}Not
purloining,
but
shewing
all
good
fidelity;
that they may
adorn the
doctrine of
God
our
Saviour
in all
things.
\VS{11}For the
grace of
God that
bringeth
salvation hath
appeared to
all
men,
\VS{12}Teaching
us
that,
denying
ungodliness
and
worldly
lusts, we should
live
soberly,
righteously,
and
godly,
in
this
present
world;
\VS{13}Looking
for
that
blessed
hope,
and the
glorious
appearing of the
great
God
and
our
Saviour
Jesus
Christ;
\VS{14}Who
gave
himself
for
us,
that he might
redeem
us
from
all
iniquity,
and
purify unto
himself a
peculiar
people,
zealous of
good
works.
\VS{15}These
things
speak,
and
exhort,
and
rebuke
with
all
authority.
Let no
man
despise
thee.

\par }\Chap{3}{\PP \VerseOne{1}Put
them in
mind to be subject
to
principalities
and
powers, to obey
magistrates, to
be
ready
to
every
good
work,
\VS{2}To speak evil
of no
man, to
be no
brawlers,
{\ADD{but}}
gentle,
shewing
all
meekness
unto
all
men.
\VS{3}For we
ourselves
also
were
sometimes
foolish,
disobedient,
deceived,
serving
divers
lusts
and
pleasures,
living
in
malice
and
envy,
hateful,
{\ADD{and}}
hating one
another.
\VS{4}But after
that the
kindness
and
love of
God
our
Saviour toward
man
appeared,
\VS{5}Not
by
works
of
righteousness
which
we have
done,
but according
to
his
mercy he
saved
us,
by the
washing of
regeneration,
and
renewing of the
Holy
Ghost;
\VS{6}Which he
shed
on
us
abundantly
through
Jesus
Christ
our
Saviour;
\VS{7}That being
justified by
his
grace, we should be
made
heirs according
to the
hope of
eternal
life.
\VS{8}{\ADD{This is}} a
faithful
saying,
and these
things I
will
that
thou affirm
constantly,
that they which have
believed in
God might be
careful to
maintain
good
works. These
things
are
good
and
profitable unto
men.
\VS{9}But
avoid
foolish
questions,
and
genealogies,
and
contentions,
and
strivings about the
law;
for they
are
unprofitable
and
vain.
\VS{10}A
man that is an
heretick
after the
first
and
second
admonition
reject;
\VS{11}Knowing
that he that is
such is
subverted,
and
sinneth,
being condemned of
himself.
\par }{\PP \VS{12}When I shall
send
Artemas
unto
thee,
or
Tychicus, be
diligent to
come
unto
me
to
Nicopolis:
for I have
determined
there to
winter.
\VS{13}Bring
Zenas the
lawyer
and
Apollos on their
journey
diligently,
that
nothing be
wanting unto
them.
\VS{14}And
let
ours
also
learn to
maintain
good
works
for
necessary
uses,
that they
be
not
unfruitful.
\VS{15}All that
are
with
me
salute
thee.
Greet them that
love
us
in the
faith.
Grace
{\ADD{be}}
with
you
all.
Amen.
\par }{\SH It was written to Titus, ordained the first bishop of the church of the Cretians, from Nicopolis of Macedonia.
\par }