\NormalFont\ShortTitle{Zechariah}
{\MT Zechariah

\par }\ChapOne{1}{\PP \VerseOne{1}In the
eighth
month, in the
second
year of
Darius, came the
word of the
{\ND{
LORD}} unto
Zechariah, the
son of
Berechiah, the
son of
Iddo the
prophet,
saying,
\VS{2}The
{\ND{
LORD}} hath been
sore
displeased with your
fathers.\FTNT{*}{{\FR 1.2 }
sore…: Heb. with displeasure}
\VS{3}Therefore
say thou unto them, Thus
saith the
{\ND{
LORD}} of
hosts;
Turn ye unto me,
saith the
{\ND{
LORD}} of
hosts, and I will
turn unto you,
saith the
{\ND{
LORD}} of
hosts.
\VS{4}Be ye not as your
fathers, unto whom the
former
prophets have
cried,
saying, Thus
saith the
{\ND{
LORD}} of
hosts;
Turn ye now from your
evil
ways, and
{\ADD{from}} your
evil
doings: but they did not
hear, nor
hearken unto me,
saith the
{\ND{
LORD}}.
\VS{5}Your
fathers, where
{\ADD{are}} they? and the
prophets, do they
live for
ever?
\VS{6}But my
words and my
statutes, which I
commanded my
servants the
prophets, did they not take
hold of your
fathers? and they
returned and
said, Like as the
{\ND{
LORD}} of
hosts
thought to
do unto us, according to our
ways, and according to our
doings, so hath he
dealt with us.\FTNT{†}{{\FR 1.6 }
take…: or, overtake}
\par }{\PP \VS{7}¶ Upon the
four and
twentieth
day of the
eleventh
month, which
{\ADD{is}} the
month
Sebat, in the
second
year of
Darius, came the
word of the
{\ND{
LORD}} unto
Zechariah, the
son of
Berechiah, the
son of
Iddo the
prophet,
saying,
\VS{8}I
saw by
night, and behold a
man
riding upon a
red
horse, and he
stood among the myrtle
trees that
{\ADD{were}} in the
bottom; and
behind him
{\ADD{were there}}
red
horses,
speckled, and
white.\FTNT{‡}{{\FR 1.8 }
speckled: or, bay}
\VS{9}Then
said I, O my
lord, what
{\ADD{are}} these? And the
angel that
talked with me
said unto me, I will
shew thee what these
{\ADD{be}}.
\VS{10}And the
man that
stood among the myrtle
trees
answered and
said, These
{\ADD{are they}} whom the
{\ND{
LORD}} hath
sent to
walk to and fro through the
earth.
\VS{11}And they
answered the
angel of the
{\ND{
LORD}} that
stood among the myrtle
trees, and
said, We have walked to and
fro through the
earth, and, behold, all the
earth sitteth
still, and is at
rest.
\par }{\PP \VS{12}¶ Then the
angel of the
{\ND{
LORD}}
answered and
said, O
{\ND{
LORD}} of
hosts, how long wilt thou not have
mercy on
Jerusalem and on the
cities of
Judah, against which thou hast had
indignation these threescore and
ten
years?
\VS{13}And the
{\ND{
LORD}}
answered the
angel that
talked with me
{\ADD{with}}
good
words
{\ADD{and}}
comfortable
words.
\VS{14}So the
angel that
communed with me
said unto me,
Cry thou,
saying, Thus
saith the
{\ND{
LORD}} of
hosts; I am
jealous for
Jerusalem and for
Zion with a
great
jealousy.
\VS{15}And I am
very
sore
displeased with the
heathen
{\ADD{that are}} at
ease: for I was but a
little
displeased, and they
helped forward the
affliction.
\VS{16}Therefore thus
saith the
{\ND{
LORD}}; I am
returned to
Jerusalem with
mercies: my
house shall be
built in it,
saith the
{\ND{
LORD}} of
hosts, and a
line shall be stretched
forth upon
Jerusalem.
\VS{17}Cry yet,
saying, Thus
saith the
{\ND{
LORD}} of
hosts; My
cities through
prosperity shall yet be spread
abroad; and the
{\ND{
LORD}} shall yet
comfort
Zion, and shall yet
choose
Jerusalem.\FTNT{§}{{\FR 1.17 }
prosperity: Heb. good}
\par }{\PP \VS{18}¶ Then lifted I
up mine
eyes, and
saw, and behold
four
horns.
\VS{19}And I
said unto the
angel that
talked with me, What
{\ADD{be}} these? And he
answered me, These
{\ADD{are}} the
horns which have
scattered
Judah,
Israel, and
Jerusalem.
\VS{20}And the
{\ND{
LORD}}
shewed me
four
carpenters.
\VS{21}Then
said I, What
come these to
do? And he
spake,
saying, These
{\ADD{are}} the
horns which have
scattered
Judah, so
that no
man did lift
up his
head: but these are
come to
fray them, to cast
out the
horns of the
Gentiles, which lifted
up
{\ADD{their}}
horn over the
land of
Judah to
scatter it.

\par }\Chap{2}{\PP \VerseOne{1}I lifted
up mine
eyes again, and
looked, and behold a
man with a
measuring
line in his
hand.
\VS{2}Then
said I, Whither
goest thou? And he
said unto me, To
measure
Jerusalem, to
see
what
{\ADD{is}} the
breadth thereof, and what
{\ADD{is}} the
length thereof.
\VS{3}And, behold, the
angel that
talked with me went
forth, and
another
angel went
out to
meet him,
\VS{4}And
said unto him,
Run,
speak to
this young
man,
saying,
Jerusalem shall be
inhabited
{\ADD{as}} towns without
walls for the
multitude of
men and
cattle
therein:
\VS{5}For I,
saith the
{\ND{
LORD}}, will be unto her a
wall of
fire round
about, and will be the
glory in the
midst of her.
\par }{\PP \VS{6}¶
Ho,
ho,
{\ADD{come forth}}, and
flee from the
land of the
north,
saith the
{\ND{
LORD}}: for I have spread you
abroad as the
four
winds of the
heaven,
saith the
{\ND{
LORD}}.
\VS{7}Deliver thyself,
O
Zion, that
dwellest
{\ADD{with}} the
daughter of
Babylon.
\VS{8}For thus
saith the
{\ND{
LORD}} of
hosts;
After the
glory hath he
sent me unto the
nations which
spoiled you: for he that
toucheth you
toucheth the
apple of his
eye.
\VS{9}For, behold, I will
shake mine
hand upon them, and they shall be a
spoil to their
servants: and ye shall
know that the
{\ND{
LORD}} of
hosts hath
sent me.
\par }{\PP \VS{10}¶
Sing and
rejoice, O
daughter of
Zion: for, lo, I
come, and I will
dwell in the
midst of thee,
saith the
{\ND{
LORD}}.
\VS{11}And
many
nations shall be
joined to the
{\ND{
LORD}} in that
day, and shall be my
people: and I will
dwell in the
midst of thee, and thou shalt
know that the
{\ND{
LORD}} of
hosts hath
sent me unto thee.
\VS{12}And the
{\ND{
LORD}} shall
inherit
Judah his
portion in the
holy
land, and shall
choose
Jerusalem again.
\VS{13}Be
silent, O all
flesh,
before the
{\ND{
LORD}}: for he is raised
up out of his
holy
habitation.\FTNT{*}{{\FR 2.13 }
his…: Heb. the habitation of his holiness}

\par }\Chap{3}{\PP \VerseOne{1}And he
shewed me
Joshua the
high
priest
standing
before the
angel of the
{\ND{
LORD}}, and
Satan
standing at his right
hand to
resist him.\FTNT{*}{{\FR 3.1 }
Satan: that is, an adversary}\FTNT{†}{{\FR 3.1 }
to…: Heb. to be his adversary}
\VS{2}And the
{\ND{
LORD}}
said unto
Satan, The
{\ND{
LORD}}
rebuke thee, O
Satan; even the
{\ND{
LORD}} that hath
chosen
Jerusalem
rebuke thee:
{\ADD{is}} not this a
brand
plucked out of the
fire?
\VS{3}Now
Joshua was
clothed with
filthy
garments, and
stood
before the
angel.
\VS{4}And he
answered and
spake unto those that
stood
before him,
saying, Take
away the
filthy
garments from him. And unto him he
said,
Behold, I have caused thine
iniquity to
pass from thee, and I will
clothe thee with change of
raiment.
\VS{5}And I
said, Let them
set a
fair
mitre upon his
head. So they
set a
fair
mitre upon his
head, and
clothed him with
garments. And the
angel of the
{\ND{
LORD}} stood
by.
\VS{6}And the
angel of the
{\ND{
LORD}}
protested unto
Joshua,
saying,
\VS{7}Thus
saith the
{\ND{
LORD}} of
hosts; If thou wilt
walk in my
ways, and if thou wilt
keep my
charge, then thou shalt also
judge my
house, and shalt also
keep my
courts, and I will
give thee places to
walk among these that stand
by.\FTNT{‡}{{\FR 3.7 }
charge: or, ordinance}\FTNT{§}{{\FR 3.7 }
places…: Heb. walks}
\par }{\PP \VS{8}Hear now, O
Joshua the
high
priest, thou, and thy
fellows that
sit
before thee: for they
{\ADD{are}}
men wondered
at: for, behold, I will bring
forth my
servant the
BRANCH.\FTNT{**}{{\FR 3.8 }
wondered…: Heb. of wonder, or, sign}
\VS{9}For behold the
stone that I have
laid
before
Joshua; upon
one
stone
{\ADD{shall be}}
seven
eyes: behold, I will
engrave the
graving thereof,
saith the
{\ND{
LORD}} of
hosts, and I will
remove the
iniquity of that
land in
one
day.
\VS{10}In that
day,
saith the
{\ND{
LORD}} of
hosts, shall ye
call every
man his
neighbour under the
vine and under the fig
tree.

\par }\Chap{4}{\PP \VerseOne{1}And the
angel that
talked with me came
again, and
waked me, as a
man that is
wakened out of his
sleep,
\VS{2}And
said unto me, What
seest thou? And I
said, I have
looked, and behold a
candlestick all
{\ADD{of}}
gold, with a
bowl upon the
top of it, and his
seven
lamps thereon, and
seven
pipes to the
seven
lamps, which
{\ADD{are}} upon the
top thereof:\FTNT{*}{{\FR 4.2 }
a bowl: Heb. her bowl}\FTNT{†}{{\FR 4.2 }
seven pipes…: or, seven several pipes to the lamps}
\VS{3}And
two olive
trees by it,
one upon the
right
{\ADD{side}} of the
bowl, and the
other upon the
left
{\ADD{side}} thereof.
\VS{4}So I
answered and
spake to the
angel that
talked with me,
saying, What
{\ADD{are}} these, my
lord?
\VS{5}Then the
angel that
talked with me
answered and
said unto me,
Knowest thou not what these be? And I
said, No, my
lord.
\VS{6}Then he
answered and
spake unto me,
saying, This
{\ADD{is}} the
word of the
{\ND{
LORD}} unto
Zerubbabel,
saying, Not by
might, nor by
power, but by my
spirit,
saith the
{\ND{
LORD}} of
hosts.\FTNT{‡}{{\FR 4.6 }
might: or, army}
\VS{7}Who
{\ADD{art}} thou, O
great
mountain?
before
Zerubbabel
{\ADD{thou shalt become}} a
plain: and he shall bring
forth the
headstone
{\ADD{thereof with}}
shoutings,
{\ADD{crying}},
Grace,
grace unto it.
\VS{8}Moreover the
word of the
{\ND{
LORD}} came unto me,
saying,
\VS{9}The
hands of
Zerubbabel have laid the
foundation of this
house; his
hands shall also
finish it; and thou shalt
know that the
{\ND{
LORD}} of
hosts hath
sent me unto you.
\VS{10}For who hath
despised the
day of small
things? for they shall
rejoice, and shall
see the
plummet in the
hand of
Zerubbabel
{\ADD{with}} those
seven; they
{\ADD{are}} the
eyes of the
{\ND{
LORD}}, which run to and
fro through the whole
earth.\FTNT{§}{{\FR 4.10 }
for they…: or, since the seven eyes of the
{\ND{LORD}} shall rejoice}\FTNT{**}{{\FR 4.10 }
plummet: Heb. stone of tin}
\par }{\PP \VS{11}¶ Then
answered I, and
said unto him, What
{\ADD{are}} these
two olive
trees upon the
right
{\ADD{side}} of the
candlestick and upon the
left
{\ADD{side}} thereof?
\VS{12}And I
answered
again, and
said unto him, What
{\ADD{be these}}
two
olive
branches which
through the
two
golden
pipes
empty the
golden
{\ADD{oil}} out of themselves?\FTNT{††}{{\FR 4.12 }
through: Heb. by the hand of}\FTNT{‡‡}{{\FR 4.12 }
empty…: or, empty out of themselves oil into the gold}\FTNT{§§}{{\FR 4.12 }
the golden oil: Heb. the gold}
\VS{13}And he
answered me and
said,
Knowest thou not what these
{\ADD{be}}? And I
said, No, my
lord.
\VS{14}Then
said he, These
{\ADD{are}} the
two
anointed
ones, that
stand by the
Lord of the whole
earth.\FTNT{***}{{\FR 4.14 }
anointed…: Heb. sons of oil}

\par }\Chap{5}{\PP \VerseOne{1}Then I
turned, and lifted
up mine
eyes, and
looked, and behold a
flying
roll.
\VS{2}And he
said unto me, What
seest thou? And I
answered, I
see a
flying
roll; the
length thereof
{\ADD{is}}
twenty
cubits, and the
breadth thereof
ten
cubits.
\VS{3}Then
said he unto me, This
{\ADD{is}} the
curse that goeth
forth over the
face of the whole
earth: for every one that
stealeth shall be cut
off
{\ADD{as}} on this side
according to it; and every one that
sweareth shall be cut
off
{\ADD{as}} on that side
according to it.\FTNT{*}{{\FR 5.3 }
every one that stealeth…: or, every one of this people that stealeth holdeth himself guiltless, as it doth}
\VS{4}I will bring it
forth,
saith the
{\ND{
LORD}} of
hosts, and it shall
enter into the
house of the
thief, and into the
house of him that
sweareth
falsely by my
name: and it shall
remain in the
midst of his
house, and shall
consume it with the
timber thereof and the
stones thereof.
\par }{\PP \VS{5}¶ Then the
angel that
talked with me went
forth, and
said unto me, Lift
up now thine
eyes, and
see what
{\ADD{is}} this that goeth
forth.
\VS{6}And I
said, What
{\ADD{is}} it? And he
said, This
{\ADD{is}} an
ephah that goeth
forth. He
said moreover, This
{\ADD{is}} their
resemblance through all the
earth.
\VS{7}And, behold, there was lifted
up a
talent of
lead: and this
{\ADD{is}}
a
woman that
sitteth in the
midst of the
ephah.\FTNT{†}{{\FR 5.7 }
talent: or, weighty piece}
\VS{8}And he
said, This
{\ADD{is}}
wickedness. And he
cast it into the
midst of the
ephah; and he
cast the
weight of
lead upon the
mouth thereof.
\VS{9}Then lifted I
up mine
eyes, and
looked, and, behold, there came
out
two
women, and the
wind
{\ADD{was}} in their
wings; for they
had
wings like the
wings of a
stork: and they lifted
up the
ephah between the
earth and the
heaven.
\VS{10}Then
said I to the
angel that
talked with me, Whither do these
bear the
ephah?
\VS{11}And he
said unto me, To
build it an
house in the
land of
Shinar: and it shall be
established, and
set there upon her own
base.

\par }\Chap{6}{\PP \VerseOne{1}And I
turned, and lifted
up mine
eyes, and
looked, and, behold, there
came
four
chariots
out from between
two
mountains; and the
mountains
{\ADD{were}}
mountains of
brass.
\VS{2}In the
first
chariot
{\ADD{were}}
red
horses; and in the
second
chariot
black
horses;
\VS{3}And in the
third
chariot
white
horses; and in the
fourth
chariot
grisled and
bay
horses.\FTNT{*}{{\FR 6.3 }
bay: or, strong}
\VS{4}Then I
answered and
said unto the
angel that
talked with me, What
{\ADD{are}} these, my
lord?
\VS{5}And the
angel
answered and
said unto me, These
{\ADD{are}} the
four
spirits of the
heavens, which go
forth from
standing before the
Lord of all the
earth.\FTNT{†}{{\FR 6.5 }
spirits: or, winds}
\VS{6}The
black
horses which
{\ADD{are}} therein go
forth into the
north
country; and the
white go
forth
after them; and the
grisled go
forth toward the
south
country.
\VS{7}And the
bay went
forth, and
sought to
go that they might walk to and
fro through the
earth: and he
said, Get you
hence, walk to and
fro through the
earth. So they walked to and
fro through the
earth.
\VS{8}Then
cried he
upon me, and
spake unto me,
saying,
Behold, these that
go toward the
north
country have
quieted my
spirit in the
north
country.
\par }{\PP \VS{9}¶ And the
word of the
{\ND{
LORD}} came unto me,
saying,
\VS{10}Take of
{\ADD{them of}} the
captivity,
{\ADD{even}} of
Heldai, of
Tobijah, and of
Jedaiah, which are
come from
Babylon, and
come thou the same
day, and
go into the
house of
Josiah the
son of
Zephaniah;
\VS{11}Then
take
silver and
gold, and
make
crowns, and
set
{\ADD{them}} upon the
head of
Joshua the
son of
Josedech, the
high
priest;
\VS{12}And
speak unto him,
saying, Thus
speaketh the
{\ND{
LORD}} of
hosts,
saying, Behold the
man whose
name
{\ADD{is}} The
BRANCH; and he shall grow
up out of his place, and he shall
build the
temple of the
{\ND{
LORD}}:\FTNT{‡}{{\FR 6.12 }
grow…: or, branch up from under him}
\VS{13}Even he shall
build the
temple of the
{\ND{
LORD}}; and he shall
bear the
glory, and shall
sit and
rule upon his
throne; and he shall be a
priest upon his
throne: and the
counsel of
peace shall be between them
both.
\VS{14}And the
crowns shall be to
Helem, and to
Tobijah, and to
Jedaiah, and to
Hen the
son of
Zephaniah, for a
memorial in the
temple of the
{\ND{
LORD}}.
\VS{15}And they
{\ADD{that are}} far
off shall
come and
build in the
temple of the
{\ND{
LORD}}, and ye shall
know that the
{\ND{
LORD}} of
hosts hath
sent me unto you. And
{\ADD{this}} shall come to pass, if ye will
diligently
obey the
voice of the
{\ND{
LORD}} your
God.

\par }\Chap{7}{\PP \VerseOne{1}And it came to pass in the
fourth
year of
king
Darius,
{\ADD{that}} the
word of the
{\ND{
LORD}} came unto
Zechariah in the
fourth
{\ADD{day}} of the
ninth
month,
{\ADD{even}} in
Chisleu;
\VS{2}When they had
sent unto the
house of
God
Sherezer and
Regem-melech, and their
men, to
pray
before the
{\ND{
LORD}},\FTNT{*}{{\FR 7.2 }
pray…: Heb. intreat the face of}
\VS{3}{\ADD{And}} to
speak unto the
priests which
{\ADD{were}} in the
house of the
{\ND{
LORD}} of
hosts, and to the
prophets,
saying, Should I
weep in the
fifth
month,
separating myself, as I have
done these so many
years?
\par }{\PP \VS{4}¶ Then came the
word of the
{\ND{
LORD}} of
hosts unto me,
saying,
\VS{5}Speak unto all the
people of the
land, and to the
priests,
saying, When ye
fasted and
mourned in the
fifth and
seventh
{\ADD{month}}, even those
seventy
years, did ye at
all
fast unto me,
{\ADD{even}} to me?
\VS{6}And when ye did
eat, and when ye did
drink, did not ye
eat
{\ADD{for yourselves}}, and
drink
{\ADD{for yourselves}}?\FTNT{†}{{\FR 7.6 }
did not ye: or, be not ye they that}
\VS{7}{\ADD{Should ye}} not
{\ADD{hear}} the
words which the
{\ND{
LORD}} hath
cried
by the
former
prophets, when
Jerusalem was
inhabited and in
prosperity, and the
cities thereof round about
her, when
{\ADD{men}}
inhabited the
south and the
plain?\FTNT{‡}{{\FR 7.7 }
Should…: or, Are not these the words}\FTNT{§}{{\FR 7.7 }
by: Heb. by the hand of}
\par }{\PP \VS{8}¶ And the
word of the
{\ND{
LORD}} came unto
Zechariah,
saying,
\VS{9}Thus
speaketh the
{\ND{
LORD}} of
hosts,
saying,
Execute
true
judgment, and
shew
mercy and
compassions every
man to his
brother:\FTNT{**}{{\FR 7.9 }
Execute…: Heb. Judge judgment of truth}
\VS{10}And
oppress not the
widow, nor the
fatherless, the
stranger, nor the
poor; and let none of you
imagine
evil
against his
brother in your
heart.
\VS{11}But they
refused to
hearken, and pulled
away the
shoulder, and
stopped their
ears, that they should not
hear.\FTNT{††}{{\FR 7.11 }
pulled…: Heb. they gave a backsliding shoulder}\FTNT{‡‡}{{\FR 7.11 }
stopped: Heb. made heavy}
\VS{12}Yea, they
made their
hearts
{\ADD{as}} an adamant
stone, lest they should
hear the
law, and the
words which the
{\ND{
LORD}} of
hosts hath
sent in his
spirit
by the
former
prophets: therefore came a
great
wrath from the
{\ND{
LORD}} of
hosts.\FTNT{§§}{{\FR 7.12 }
by: Heb. by the hand of}
\VS{13}Therefore it is come to pass,
{\ADD{that}} as he
cried, and they would not
hear; so they
cried, and I would not
hear,
saith the
{\ND{
LORD}} of
hosts:
\VS{14}But I scattered them with a
whirlwind among all the
nations whom they
knew not. Thus the
land was
desolate
after them, that no man passed
through nor
returned: for they
laid the
pleasant
land
desolate.\FTNT{***}{{\FR 7.14 }
pleasant…: Heb. land of desire}

\par }\Chap{8}{\PP \VerseOne{1}Again the
word of the
{\ND{
LORD}} of
hosts came
{\ADD{to me}},
saying,
\VS{2}Thus
saith the
{\ND{
LORD}} of
hosts; I was
jealous for
Zion with
great
jealousy, and I was
jealous for her with
great
fury.
\VS{3}Thus
saith the
{\ND{
LORD}}; I am
returned unto
Zion, and will
dwell in the
midst of
Jerusalem: and
Jerusalem shall be
called a
city of
truth; and the
mountain of the
{\ND{
LORD}} of
hosts the
holy
mountain.
\VS{4}Thus
saith the
{\ND{
LORD}} of
hosts; There shall yet old
men and old
women
dwell in the
streets of
Jerusalem, and every
man with his
staff in his
hand for
very
age.\FTNT{*}{{\FR 8.4 }
very…: Heb. multitude of days}
\VS{5}And the
streets of the
city shall be
full of
boys and
girls
playing in the
streets thereof.
\VS{6}Thus
saith the
{\ND{
LORD}} of
hosts; If it be
marvellous in the
eyes of the
remnant of this
people in these
days, should it also be
marvellous in mine
eyes?
saith the
{\ND{
LORD}} of
hosts.\FTNT{†}{{\FR 8.6 }
marvellous: or, hard, or, difficult}
\VS{7}Thus
saith the
{\ND{
LORD}} of
hosts; Behold, I will
save my
people from the
east
country, and from the
west
country;\FTNT{‡}{{\FR 8.7 }
the west…: Heb. the country of the going down of the sun}
\VS{8}And I will
bring them, and they shall
dwell in the
midst of
Jerusalem: and they shall be my
people, and I will be their
God, in
truth and in
righteousness.
\par }{\PP \VS{9}¶ Thus
saith the
{\ND{
LORD}} of
hosts; Let your
hands be
strong, ye that
hear in these
days these
words by the
mouth of the
prophets, which
{\ADD{were}} in the
day
{\ADD{that}} the
foundation of the
house of the
{\ND{
LORD}} of
hosts was
laid, that the
temple might be
built.
\VS{10}For
before these
days there
was no
hire for
man, nor any
hire for
beast; neither
{\ADD{was there any}}
peace to him that went
out or came
in because of the
affliction: for I
set all
men every
one against his
neighbour.\FTNT{§}{{\FR 8.10 }
there was…: or, the hire of man became nothing}
\VS{11}But now I
{\ADD{will}} not
{\ADD{be}} unto the
residue of this
people as in the
former
days,
saith the
{\ND{
LORD}} of
hosts.
\VS{12}For the
seed
{\ADD{shall be}}
prosperous; the
vine shall
give her
fruit, and the
ground shall
give her
increase, and the
heavens shall
give their
dew; and I will cause the
remnant of this
people to
possess all these
{\ADD{things}}.\FTNT{**}{{\FR 8.12 }
prosperous: Heb. of peace}
\VS{13}And it shall come to pass,
{\ADD{that}} as ye were a
curse among the
heathen, O
house of
Judah, and
house of
Israel; so will I
save you, and ye shall be a
blessing:
fear not,
{\ADD{but}} let your
hands be
strong.
\VS{14}For thus
saith the
{\ND{
LORD}} of
hosts; As I
thought to
punish you, when your
fathers provoked me to
wrath,
saith the
{\ND{
LORD}} of
hosts, and I
repented not:
\VS{15}So
again have I
thought in these
days to do
well unto
Jerusalem and to the
house of
Judah:
fear ye not.
\par }{\PP \VS{16}¶ These
{\ADD{are}} the
things that ye shall
do;
Speak ye every
man the
truth to his
neighbour;
execute the
judgment of
truth and
peace in your
gates:\FTNT{††}{{\FR 8.16 }
execute…: Heb. judge truth, and the judgment of peace}
\VS{17}And let
none of you
imagine
evil in your
hearts against his
neighbour; and
love no
false
oath: for all these
{\ADD{are things}} that I
hate,
saith the
{\ND{
LORD}}.
\par }{\PP \VS{18}¶ And the
word of the
{\ND{
LORD}} of
hosts came unto me,
saying,
\VS{19}Thus
saith the
{\ND{
LORD}} of
hosts; The
fast of the
fourth
{\ADD{month}}, and the
fast of the
fifth, and the
fast of the
seventh, and the
fast of the
tenth, shall be to the
house of
Judah
joy and
gladness, and
cheerful
feasts; therefore
love the
truth and
peace.\FTNT{‡‡}{{\FR 8.19 }
feasts: or, solemn, or, set times}
\VS{20}Thus
saith the
{\ND{
LORD}} of
hosts;
{\ADD{It shall}} yet
{\ADD{come to pass}}, that there shall
come
people, and the
inhabitants of
many
cities:
\VS{21}And the
inhabitants of
one
{\ADD{city}} shall
go to
another,
saying, Let us
go
speedily to
pray
before the
{\ND{
LORD}}, and to
seek the
{\ND{
LORD}} of
hosts: I will
go also.\FTNT{§§}{{\FR 8.21 }
speedily: or, continually: Heb. going}\FTNT{***}{{\FR 8.21 }
to pray…: Heb. to intreat the face of}
\VS{22}Yea,
many
people and
strong
nations shall
come to
seek the
{\ND{
LORD}} of
hosts in
Jerusalem, and to
pray
before the
{\ND{
LORD}}.\FTNT{†††}{{\FR 8.22 }
to pray…: Heb. to intreat the face of}
\VS{23}Thus
saith the
{\ND{
LORD}} of
hosts; In those
days
{\ADD{it shall come to pass}}, that
ten
men shall take
hold out of all
languages of the
nations, even shall take
hold of the
skirt of him that is a
Jew,
saying, We will
go with you: for we have
heard
{\ADD{that}}
God
{\ADD{is}} with you.

\par }\Chap{9}{\PP \VerseOne{1}The
burden of the
word of the
{\ND{
LORD}} in the
land of
Hadrach, and
Damascus
{\ADD{shall be}} the
rest thereof: when the
eyes of
man, as of all the
tribes of
Israel,
{\ADD{shall be}} toward the
{\ND{
LORD}}.
\VS{2}And
Hamath also shall
border thereby;
Tyrus, and
Zidon, though it be
very
wise.
\VS{3}And
Tyrus did
build herself a strong
hold, and heaped
up
silver as the
dust, and fine
gold as the
mire of the
streets.
\VS{4}Behold, the
Lord will cast her
out, and he will
smite her
power in the
sea; and she shall be
devoured with
fire.
\VS{5}Ashkelon shall
see
{\ADD{it}}, and
fear;
Gaza also
{\ADD{shall see it}}, and be
very
sorrowful, and
Ekron; for her
expectation shall be
ashamed; and the
king shall
perish from
Gaza, and
Ashkelon shall not be
inhabited.
\VS{6}And a
bastard shall
dwell in
Ashdod, and I will cut
off the
pride of the
Philistines.
\VS{7}And I will take
away his
blood out of his
mouth, and his
abominations from between his
teeth: but he that
remaineth, even he,
{\ADD{shall be}} for our
God, and he shall be as a
governor in
Judah, and
Ekron as a
Jebusite.\FTNT{*}{{\FR 9.7 }
blood: Heb. bloods}
\VS{8}And I will
encamp about mine
house because of the
army, because of him that passeth
by, and because of him that
returneth: and no
oppressor shall pass
through them any more: for now have I
seen with mine
eyes.
\par }{\PP \VS{9}¶
Rejoice
greatly, O
daughter of
Zion;
shout, O
daughter of
Jerusalem: behold, thy
King
cometh unto thee: he
{\ADD{is}}
just, and having
salvation;
lowly, and
riding upon an
ass, and upon a
colt the
foal of an
ass.\FTNT{†}{{\FR 9.9 }
having…: or, saving himself}
\VS{10}And I will cut
off the
chariot from
Ephraim, and the
horse from
Jerusalem, and the
battle
bow shall be cut
off: and he shall
speak
peace unto the
heathen: and his
dominion
{\ADD{shall be}} from
sea
{\ADD{even}} to
sea, and from the
river
{\ADD{even}} to the
ends of the
earth.
\VS{11}As for thee also, by the
blood of thy
covenant I have sent
forth thy
prisoners out of the
pit wherein
{\ADD{is}} no
water.\FTNT{‡}{{\FR 9.11 }
by…: or, whose covenant is by blood}
\par }{\PP \VS{12}¶
Turn you to the strong
hold, ye
prisoners of
hope: even to
day do I
declare
{\ADD{that}} I will
render
double unto thee;
\VS{13}When I have
bent
Judah for me,
filled the
bow with
Ephraim, and raised
up thy
sons, O
Zion, against thy
sons, O
Greece, and
made thee as the
sword of a mighty
man.
\VS{14}And the
{\ND{
LORD}} shall be
seen over them, and his
arrow shall go
forth as the
lightning: and the
Lord
{\ND{
GOD}} shall
blow the
trumpet, and shall
go with
whirlwinds of the
south.
\VS{15}The
{\ND{
LORD}} of
hosts shall
defend them; and they shall
devour, and
subdue with
sling
stones; and they shall
drink,
{\ADD{and}} make a
noise as through
wine; and they shall be
filled like
bowls,
{\ADD{and}} as the
corners of the
altar.\FTNT{§}{{\FR 9.15 }
with…: or, the stones of the sling}\FTNT{**}{{\FR 9.15 }
be…: or, fill both the bowls, etc}
\VS{16}And the
{\ND{
LORD}} their
God shall
save them in that
day as the
flock of his
people: for they
{\ADD{shall be as}} the
stones of a
crown, lifted up as an
ensign upon his
land.
\VS{17}For how great
{\ADD{is}} his
goodness, and how great
{\ADD{is}} his
beauty!
corn shall make the young
men
cheerful, and new
wine the
maids.\FTNT{††}{{\FR 9.17 }
cheerful: or, grow, or, speak}

\par }\Chap{10}{\PP \VerseOne{1}Ask ye of the
{\ND{
LORD}}
rain in the
time of the latter
rain;
{\ADD{so}} the
{\ND{
LORD}} shall
make bright
clouds, and
give them
showers of
rain, to every
one
grass in the
field.\FTNT{*}{{\FR 10.1 }
bright…: or, lightnings}
\VS{2}For the
idols have
spoken
vanity, and the
diviners have
seen a
lie, and have
told
false
dreams; they
comfort in
vain: therefore they
went their way as a
flock, they were
troubled, because
{\ADD{there was}} no
shepherd.\FTNT{†}{{\FR 10.2 }
idols: Heb. teraphims}\FTNT{‡}{{\FR 10.2 }
were…: or, answered that, etc}
\VS{3}Mine
anger was
kindled against the
shepherds, and I
punished the
goats: for the
{\ND{
LORD}} of
hosts hath
visited his
flock the
house of
Judah, and hath
made them as his
goodly
horse in the
battle.\FTNT{§}{{\FR 10.3 }
punished: Heb. visited upon}
\VS{4}Out of him came
forth the
corner, out of him the
nail, out of him the
battle
bow, out of him every
oppressor
together.
\par }{\PP \VS{5}¶ And they shall be as
mighty
{\ADD{men}}, which tread
down
{\ADD{their enemies}} in the
mire of the
streets in the
battle: and they shall
fight, because the
{\ND{
LORD}}
{\ADD{is}} with them, and the
riders on
horses shall be
confounded.\FTNT{**}{{\FR 10.5 }
the riders…: or, they shall make the riders on horses ashamed}
\VS{6}And I will
strengthen the
house of
Judah, and I will
save the
house of
Joseph, and I will bring them again to
place them; for I have
mercy upon them: and they shall be as
though I had not cast them
off: for I
{\ADD{am}} the
{\ND{
LORD}} their
God, and will
hear them.
\VS{7}And
{\ADD{they of}}
Ephraim shall be like a
mighty
{\ADD{man}}, and their
heart shall
rejoice as through
wine: yea, their
children shall
see
{\ADD{it}}, and be
glad; their
heart shall
rejoice in the
{\ND{
LORD}}.
\VS{8}I will
hiss for them, and
gather them; for I have
redeemed them: and they shall
increase as they have
increased.
\VS{9}And I will
sow them among the
people: and they shall
remember me in far
countries; and they shall
live with their
children, and turn
again.
\VS{10}I will bring them
again also out of the
land of
Egypt, and
gather them out of
Assyria; and I will
bring them into the
land of
Gilead and
Lebanon; and
{\ADD{place}} shall not be
found for them.
\VS{11}And he shall pass
through the
sea with
affliction, and shall
smite the
waves in the
sea, and all the
deeps of the
river shall dry
up: and the
pride of
Assyria shall be brought
down, and the
sceptre of
Egypt shall depart
away.
\VS{12}And I will
strengthen them in the
{\ND{
LORD}}; and they shall walk up and
down in his
name,
saith the
{\ND{
LORD}}.

\par }\Chap{11}{\PP \VerseOne{1}Open thy
doors, O
Lebanon, that the
fire may
devour thy
cedars.
\VS{2}Howl, fir
tree; for the
cedar is
fallen; because the
mighty are
spoiled:
howl, O ye
oaks of
Bashan; for the
forest of the
vintage is come
down.\FTNT{*}{{\FR 11.2 }
mighty: or, gallants}\FTNT{†}{{\FR 11.2 }
the forest…: or, the defenced forest}
\par }{\PP \VS{3}¶
{\ADD{There is}} a
voice of the
howling of the
shepherds; for their
glory is
spoiled: a
voice of the
roaring of young
lions; for the
pride of
Jordan is
spoiled.
\par }{\PP \VS{4}Thus
saith the
{\ND{
LORD}} my
God;
Feed the
flock of the
slaughter;
\VS{5}Whose
possessors
slay them, and hold themselves not
guilty: and they that
sell them
say,
Blessed
{\ADD{be}} the
{\ND{
LORD}}; for I am
rich: and their own
shepherds
pity them not.
\VS{6}For I will no more
pity the
inhabitants of the
land,
saith the
{\ND{
LORD}}: but, lo, I will
deliver the
men every
one into his
neighbour’s
hand, and into the
hand of his
king: and they shall
smite the
land, and out of their
hand I will not
deliver
{\ADD{them}}.\FTNT{‡}{{\FR 11.6 }
deliver: Heb. make to be found}
\VS{7}And I will
feed the
flock of
slaughter,
{\ADD{even}}
you, O
poor of the
flock. And I
took unto me
two
staves; the
one I
called
Beauty, and the
other I
called
Bands; and I
fed the
flock.\FTNT{§}{{\FR 11.7 }
even…: or, verily the poor}\FTNT{**}{{\FR 11.7 }
Bands: or, Binders}
\VS{8}Three
shepherds also I cut
off in
one
month; and my
soul
lothed them, and their
soul also
abhorred me.\FTNT{††}{{\FR 11.8 }
lothed…: Heb. was straightened for them}
\VS{9}Then
said I, I will not
feed you: that that
dieth, let it
die; and that that is to be cut
off, let it be cut
off; and let the
rest
eat every
one the
flesh of
another.\FTNT{‡‡}{{\FR 11.9 }
another: Heb. his fellow, or, neighbour}
\par }{\PP \VS{10}¶ And I
took my
staff,
{\ADD{even}}
Beauty, and cut it
asunder, that I might
break my
covenant which I had
made with all the
people.
\VS{11}And it was
broken in that
day: and so the
poor of the
flock that
waited upon me
knew that it
{\ADD{was}} the
word of the
{\ND{
LORD}}.\FTNT{§§}{{\FR 11.11 }
so…: or, the poor of the flock, etc. certainly knew}
\VS{12}And I
said unto them, If ye
think
good,
give
{\ADD{me}} my
price; and if not,
forbear. So they
weighed for my
price
thirty
{\ADD{pieces}} of
silver.\FTNT{***}{{\FR 11.12 }
If ye…: Heb. If it be good in your eyes}
\VS{13}And the
{\ND{
LORD}}
said unto me,
Cast it unto the
potter: a
goodly
price that I was prised
at of them. And I
took the
thirty
{\ADD{pieces}} of
silver, and
cast them to the
potter in the
house of the
{\ND{
LORD}}.
\VS{14}Then I cut
asunder mine
other
staff,
{\ADD{even}}
Bands, that I might
break the
brotherhood between
Judah and
Israel.\FTNT{†††}{{\FR 11.14 }
Bands: or, Binders}
\par }{\PP \VS{15}¶ And the
{\ND{
LORD}}
said unto me,
Take unto thee yet the
instruments of a
foolish
shepherd.
\VS{16}For, lo, I will raise
up a
shepherd in the
land,
{\ADD{which}} shall not
visit those that be cut
off, neither shall
seek the young
one, nor
heal that that is
broken, nor
feed that that
standeth still: but he shall
eat the
flesh of the
fat, and
tear their
claws in
pieces.\FTNT{‡‡‡}{{\FR 11.16 }
cut off: or, hidden}\FTNT{§§§}{{\FR 11.16 }
feed: or, bear}
\VS{17}Woe to the
idol
shepherd that
leaveth the
flock! the
sword
{\ADD{shall be}} upon his
arm, and upon his
right
eye: his
arm shall be
clean dried
up, and his
right
eye shall be
utterly
darkened.

\par }\Chap{12}{\PP \VerseOne{1}The
burden of the
word of the
{\ND{
LORD}} for
Israel,
saith the
{\ND{
LORD}}, which
stretcheth forth the
heavens, and layeth the
foundation of the
earth, and
formeth the
spirit of
man
within him.
\VS{2}Behold, I will
make
Jerusalem a
cup of
trembling unto all the
people round
about, when they shall be in the
siege both against
Judah
{\ADD{and}} against
Jerusalem.\FTNT{*}{{\FR 12.2 }
trembling: or, slumber, or, poison}\FTNT{†}{{\FR 12.2 }
when…: or, and also against Judah shall he be which shall be in siege against Jerusalem}
\par }{\PP \VS{3}¶ And in that
day will I
make
Jerusalem a
burdensome
stone for all
people: all that
burden themselves with it shall be cut in
pieces, though all the
people of the
earth be gathered
together against it.
\VS{4}In that
day,
saith the
{\ND{
LORD}}, I will
smite every
horse with
astonishment, and his
rider with
madness: and I will
open mine
eyes upon the
house of
Judah, and will
smite every
horse of the
people with
blindness.
\VS{5}And the
governors of
Judah shall
say in their
heart, The
inhabitants of
Jerusalem
{\ADD{shall be}} my
strength in the
{\ND{
LORD}} of
hosts their
God.\FTNT{‡}{{\FR 12.5 }
The inhabitants…: or, There is strength to me and to the inhabitants, etc}
\par }{\PP \VS{6}¶ In that
day will I
make the
governors of
Judah like an
hearth of
fire among the
wood, and like a
torch of
fire in a
sheaf; and they shall
devour all the
people round
about, on the right
hand and on the
left: and
Jerusalem shall be
inhabited again in her own place,
{\ADD{even}} in
Jerusalem.
\VS{7}The
{\ND{
LORD}} also shall
save the
tents of
Judah
first, that the
glory of the
house of
David and the
glory of the
inhabitants of
Jerusalem do not
magnify
{\ADD{themselves}} against
Judah.
\VS{8}In that
day shall the
{\ND{
LORD}}
defend the
inhabitants of
Jerusalem; and he that is
feeble among them at that
day shall be as
David; and the
house of
David
{\ADD{shall be}} as
God, as the
angel of the
{\ND{
LORD}}
before them.\FTNT{§}{{\FR 12.8 }
feeble: or, abject: Heb. fallen}
\par }{\PP \VS{9}¶ And it shall come to pass in that
day,
{\ADD{that}} I will
seek to
destroy all the
nations that
come against
Jerusalem.
\VS{10}And I will
pour upon the
house of
David, and upon the
inhabitants of
Jerusalem, the
spirit of
grace and of
supplications: and they shall
look upon me whom they have
pierced, and they shall
mourn for him, as one
mourneth for
{\ADD{his}}
only
{\ADD{son}}, and shall be in
bitterness for him, as one that is in
bitterness for
{\ADD{his}}
firstborn.
\VS{11}In that
day shall there be a
great
mourning in
Jerusalem, as the
mourning of
Hadadrimmon in the
valley of
Megiddon.
\VS{12}And the
land shall
mourn,
every
family
apart; the
family of the
house of
David apart, and their
wives apart; the
family of the
house of
Nathan apart, and their
wives apart;\FTNT{**}{{\FR 12.12 }
every…: Heb. families, families}
\VS{13}The
family of the
house of
Levi apart, and their
wives apart; the
family of
Shimei apart, and their
wives apart;\FTNT{††}{{\FR 12.13 }
of Shimei: or, of Simeon, as LXX}
\VS{14}All the
families that
remain, every
family apart, and their
wives apart.

\par }\Chap{13}{\PP \VerseOne{1}In that
day there shall be a
fountain
opened to the
house of
David and to the
inhabitants of
Jerusalem for
sin and for
uncleanness.\FTNT{*}{{\FR 13.1 }
uncleanness: Heb. separation for uncleanness}
\par }{\PP \VS{2}¶ And it shall come to pass in that
day,
saith the
{\ND{
LORD}} of
hosts,
{\ADD{that}} I will cut
off the
names of the
idols out of the
land, and they shall no more be
remembered: and also I will cause the
prophets and the
unclean
spirit to
pass out of the
land.
\VS{3}And it shall come to pass,
{\ADD{that}} when
any shall yet
prophesy, then his
father and his
mother that
begat him shall
say unto him, Thou shalt not
live; for thou
speakest
lies in the
name of the
{\ND{
LORD}}: and his
father and his
mother that
begat him shall
thrust him through when he
prophesieth.
\VS{4}And it shall come to pass in that
day,
{\ADD{that}} the
prophets shall be
ashamed every
one of his
vision, when he hath
prophesied; neither shall they
wear a
rough
garment to
deceive:\FTNT{†}{{\FR 13.4 }
a rough…: Heb. a garment of hair}\FTNT{‡}{{\FR 13.4 }
deceive: Heb. lie}
\VS{5}But he shall
say, I
{\ADD{am}} no
prophet, I
{\ADD{am}} an
husbandman; for
man taught me to keep
cattle from my
youth.
\VS{6}And
{\ADD{one}} shall
say unto him, What
{\ADD{are}} these
wounds in thine
hands? Then he shall
answer,
{\ADD{Those}} with which I was
wounded
{\ADD{in}} the
house of my
friends.
\par }{\PP \VS{7}¶
Awake, O
sword, against my
shepherd, and against the
man
{\ADD{that is}} my
fellow,
saith the
{\ND{
LORD}} of
hosts:
smite the
shepherd, and the
sheep shall be
scattered: and I will
turn mine
hand upon the little
ones.
\VS{8}And it shall come to pass,
{\ADD{that}} in all the
land,
saith the
{\ND{
LORD}},
two
parts therein shall be cut
off
{\ADD{and}}
die; but the
third shall be
left therein.
\VS{9}And I will
bring the third
part through the
fire, and will
refine them as
silver is
refined, and will
try them as
gold is
tried: they shall
call on my
name, and I will
hear them: I will
say, It
{\ADD{is}} my
people: and they shall
say, The
{\ND{
LORD}}
{\ADD{is}} my
God.

\par }\Chap{14}{\PP \VerseOne{1}Behold, the
day of the
{\ND{
LORD}}
cometh, and thy
spoil shall be
divided in the
midst of thee.
\VS{2}For I will
gather all
nations against
Jerusalem to
battle; and the
city shall be
taken, and the
houses
rifled, and the
women
ravished; and
half of the
city shall go
forth into
captivity, and the
residue of the
people shall not be cut
off from the
city.
\VS{3}Then shall the
{\ND{
LORD}} go
forth, and
fight against those
nations, as
when he
fought in the
day of
battle.
\par }{\PP \VS{4}¶ And his
feet shall
stand in that
day upon the
mount of
Olives, which
{\ADD{is}}
before
Jerusalem on the
east, and the
mount of
Olives shall
cleave in the
midst thereof toward the
east and toward the
west,
{\ADD{and there shall be}} a
very
great
valley; and
half of the
mountain shall
remove toward the
north, and
half of it toward the
south.
\VS{5}And ye shall
flee
{\ADD{to}} the
valley of the
mountains; for the
valley of the
mountains shall
reach unto
Azal: yea, ye shall
flee, like as ye
fled from
before the
earthquake in the
days of
Uzziah
king of
Judah: and the
{\ND{
LORD}} my
God shall
come,
{\ADD{and}} all the
saints with thee.\FTNT{*}{{\FR 14.5 }
the mountains: or, my mountains}\FTNT{†}{{\FR 14.5 }
for the…: or, when he shall touch the valley of the mountains to the place he separated}
\VS{6}And it shall come to pass in that
day,
{\ADD{that}} the
light shall not be
clear,
{\ADD{nor}}
dark:\FTNT{‡}{{\FR 14.6 }
that the…: that is, it shall not be clear in some places, and dark in other places of the world}\FTNT{§}{{\FR 14.6 }
clear: Heb. precious}\FTNT{**}{{\FR 14.6 }
dark: Heb. thickness}
\VS{7}But it shall be
one
day which shall be
known to the
{\ND{
LORD}}, not
day, nor
night: but it shall come to pass,
{\ADD{that}} at
evening
time it shall be
light.\FTNT{††}{{\FR 14.7 }
it shall be…: or, the day shall be one}
\par }{\PP \VS{8}And it shall be in that
day,
{\ADD{that}}
living
waters shall go
out from
Jerusalem;
half of them toward the
former
sea, and
half of them toward the
hinder
sea: in
summer and in
winter shall it be.\FTNT{‡‡}{{\FR 14.8 }
former: or, eastern}
\VS{9}And the
{\ND{
LORD}} shall be
king over all the
earth: in that
day shall there be
one
{\ND{
LORD}}, and his
name
one.
\VS{10}All the
land shall be
turned as a
plain from
Geba to
Rimmon
south of
Jerusalem: and it shall be lifted
up, and
inhabited in her place, from
Benjamin’s
gate unto the
place of the
first
gate, unto the
corner
gate, and
{\ADD{from}} the
tower of
Hananeel unto the
king’s
winepresses.\FTNT{§§}{{\FR 14.10 }
turned: or, compassed}\FTNT{***}{{\FR 14.10 }
inhabited: or, shall abide}
\VS{11}And
{\ADD{men}} shall
dwell in it, and there shall be no more utter
destruction; but
Jerusalem shall be
safely
inhabited.\FTNT{†††}{{\FR 14.11 }
shall be: or, shall abide}
\par }{\PP \VS{12}¶ And this shall be the
plague wherewith the
{\ND{
LORD}} will
smite all the
people that have
fought against
Jerusalem; Their
flesh shall consume
away while they
stand upon their
feet, and their
eyes shall consume
away in their
holes, and their
tongue shall consume
away in their
mouth.
\VS{13}And it shall come to pass in that
day,
{\ADD{that}} a
great
tumult from the
{\ND{
LORD}} shall be among them; and they shall lay
hold every
one on the
hand of his
neighbour, and his
hand shall rise
up against the
hand of his
neighbour.
\VS{14}And
Judah also shall
fight at
Jerusalem; and the
wealth of all the
heathen round
about shall be gathered
together,
gold, and
silver, and
apparel, in
great
abundance.\FTNT{‡‡‡}{{\FR 14.14 }
Judah…: or, thou also, O Judah shalt}\FTNT{§§§}{{\FR 14.14 }
at: or, against}
\VS{15}And so shall be the
plague of the
horse, of the
mule, of the
camel, and of the
ass, and of all the
beasts that shall be in
these
tents, as this
plague.
\par }{\PP \VS{16}¶ And it shall come to pass,
{\ADD{that}} every one that is
left of all the
nations which
came against
Jerusalem shall even go
up
from
year to
year to
worship the
King, the
{\ND{
LORD}} of
hosts, and to
keep the
feast of
tabernacles.
\VS{17}And it shall be,
{\ADD{that}} whoso will not come
up of
{\ADD{all}} the
families of the
earth unto
Jerusalem to
worship the
King, the
{\ND{
LORD}} of
hosts, even upon them shall be no
rain.
\VS{18}And if the
family of
Egypt go not
up, and
come not, that
{\ADD{have}} no
{\ADD{rain;}} there shall be the
plague, wherewith the
{\ND{
LORD}} will
smite the
heathen that come not
up to
keep the
feast of
tabernacles.\FTNT{*}{{\FR 14.18 }
that have no: Heb. upon whom there is not}
\VS{19}This shall be the
punishment of
Egypt, and the
punishment of all
nations that come not
up to
keep the
feast of
tabernacles.\FTNT{†}{{\FR 14.19 }
punishment: or, sin}
\par }{\PP \VS{20}¶ In that
day shall there be upon the
bells of the
horses,
HOLINESS UNTO THE
{\ND{
LORD}}; and the
pots in the
{\ND{
LORD’s}}
house shall be like the
bowls
before the
altar.\FTNT{‡}{{\FR 14.20 }
bells: or, bridles}
\VS{21}Yea, every
pot in
Jerusalem and in
Judah shall be
holiness unto the
{\ND{
LORD}} of
hosts: and all they that
sacrifice shall
come and
take of them, and
seethe therein: and in that
day there shall be no more the
Canaanite in the
house of the
{\ND{
LORD}} of
hosts.
\par }