\NormalFont\ShortTitle{Zephaniah}
{\MT Zephaniah

\par }\ChapOne{1}{\PP \VerseOne{1}The
word of the
{\ND{
LORD}} which came unto
Zephaniah the
son of
Cushi, the
son of
Gedaliah, the
son of
Amariah, the
son of
Hizkiah, in the
days of
Josiah the
son of
Amon,
king of
Judah.
\VS{2}I will
utterly
consume all
{\ADD{things}} from off the
land,
saith the
{\ND{
LORD}}.\FTNT{*}{{\FR 1.2 }
I will…: Heb. By taking away I will make an end}\FTNT{†}{{\FR 1.2 }
the land: Heb. the face of the land}
\VS{3}I will
consume
man and
beast; I will
consume the
fowls of the
heaven, and the
fishes of the
sea, and the
stumblingblocks with the
wicked; and I will cut
off
man from
off the
land,
saith the
{\ND{
LORD}}.\FTNT{‡}{{\FR 1.3 }
stumblingblocks: or, idols}
\VS{4}I will also stretch
out mine
hand upon
Judah, and upon all the
inhabitants of
Jerusalem; and I will cut
off the
remnant of
Baal from this
place,
{\ADD{and}} the
name of the
Chemarims with the
priests;
\VS{5}And them that
worship the
host of
heaven upon the
housetops; and them that
worship
{\ADD{and}} that
swear by the
{\ND{
LORD}}, and that
swear by
Malcham;\FTNT{§}{{\FR 1.5 }
by the…: or, to the
{\ND{LORD}}}
\VS{6}And them that are turned
back
from the
{\ND{
LORD}}; and
{\ADD{those}} that have not
sought the
{\ND{
LORD}}, nor
enquired for him.
\par }{\PP \VS{7}Hold thy
peace at the
presence of the
Lord
{\ND{
GOD}}: for the
day of the
{\ND{
LORD}}
{\ADD{is}} at
hand: for the
{\ND{
LORD}} hath
prepared a
sacrifice, he hath
bid his
guests.\FTNT{**}{{\FR 1.7 }
bid: Heb. sanctified, or, prepared}
\VS{8}And it shall come to pass in the
day of the
{\ND{
LORD’s}}
sacrifice, that I will
punish the
princes, and the
king’s
children, and all such as are
clothed with
strange
apparel.\FTNT{††}{{\FR 1.8 }
punish: Heb. visit upon}
\VS{9}In the same
day also will I
punish all those that
leap on the
threshold, which
fill their
masters’
houses with
violence and
deceit.
\VS{10}And it shall come to pass in that
day,
saith the
{\ND{
LORD}},
{\ADD{that there shall be}} the
noise of a
cry from the
fish
gate, and an
howling from the
second, and a
great
crashing from the
hills.
\VS{11}Howl, ye
inhabitants of
Maktesh, for all the
merchant
people are cut
down; all they that
bear
silver are cut
off.
\VS{12}And it shall come to pass at that
time,
{\ADD{that}} I will
search
Jerusalem with
candles, and
punish the
men that are
settled on their
lees: that
say in their
heart, The
{\ND{
LORD}} will not do
good, neither will he do
evil.\FTNT{‡‡}{{\FR 1.12 }
settled: Heb. curded, or, thickened}
\VS{13}Therefore their
goods shall become a
booty, and their
houses a
desolation: they shall also
build
houses, but not
inhabit
{\ADD{them;}} and they shall
plant
vineyards, but not
drink the
wine thereof.
\par }{\PP \VS{14}The
great
day of the
{\ND{
LORD}}
{\ADD{is}}
near,
{\ADD{it is}}
near, and
hasteth
greatly,
{\ADD{even}} the
voice of the
day of the
{\ND{
LORD}}: the mighty
man shall
cry there
bitterly.
\VS{15}That
day
{\ADD{is}} a
day of
wrath, a
day of
trouble and
distress, a
day of
wasteness and
desolation, a
day of
darkness and
gloominess, a
day of
clouds and thick
darkness,
\VS{16}A
day of the
trumpet and
alarm against the
fenced
cities, and against the
high
towers.
\VS{17}And I will bring
distress upon
men, that they shall
walk like blind
men, because they have
sinned against the
{\ND{
LORD}}: and their
blood shall be poured
out as
dust, and their
flesh as the
dung.
\VS{18}Neither their
silver nor their
gold shall be
able to
deliver them in the
day of the
{\ND{
LORD’s}}
wrath; but the whole
land shall be
devoured by the
fire of his
jealousy: for he shall
make even a
speedy
riddance of all them that
dwell in the
land.

\par }\Chap{2}{\PP \VerseOne{1}Gather yourselves
together, yea, gather
together, O
nation not
desired;\FTNT{*}{{\FR 2.1 }
not…: or, not desirous}
\VS{2}Before the
decree bring
forth,
{\ADD{before}} the
day
pass as the
chaff,
before the
fierce
anger of the
{\ND{
LORD}}
come upon you, before the
day of the
{\ND{
LORD’s}}
anger
come upon you.
\VS{3}Seek ye the
{\ND{
LORD}}, all ye
meek of the
earth, which have
wrought his
judgment;
seek
righteousness,
seek
meekness: it may
be ye shall be
hid in the
day of the
{\ND{
LORD’s}}
anger.
\par }{\PP \VS{4}¶ For
Gaza shall be
forsaken, and
Ashkelon a
desolation: they shall drive
out
Ashdod at the noon
day, and
Ekron shall be rooted
up.
\VS{5}Woe unto the
inhabitants of the
sea
coast, the
nation of the
Cherethites! the
word of the
{\ND{
LORD}}
{\ADD{is}} against you; O
Canaan, the
land of the
Philistines, I will even
destroy thee, that there shall be no
inhabitant.
\VS{6}And the
sea
coast shall be
dwellings
{\ADD{and}}
cottages for
shepherds, and
folds for
flocks.
\VS{7}And the
coast shall be for the
remnant of the
house of
Judah; they shall
feed thereupon: in the
houses of
Ashkelon shall they lie
down in the
evening: for the
{\ND{
LORD}} their
God shall
visit them, and turn
away their
captivity.\FTNT{†}{{\FR 2.7 }
for the
{\ND{LORD}}: or, when, etc}
\par }{\PP \VS{8}¶ I have
heard the
reproach of
Moab, and the
revilings of the
children of
Ammon, whereby they have
reproached my
people, and
magnified
{\ADD{themselves}} against their
border.
\VS{9}Therefore
{\ADD{as}} I
live,
saith the
{\ND{
LORD}} of
hosts, the
God of
Israel, Surely
Moab shall be as
Sodom, and the
children of
Ammon as
Gomorrah,
{\ADD{even}} the
breeding of
nettles, and
saltpits, and a
perpetual
desolation: the
residue of my
people shall
spoil them, and the
remnant of my
people shall
possess them.
\VS{10}This shall they have for their
pride, because they have
reproached and
magnified
{\ADD{themselves}} against the
people of the
{\ND{
LORD}} of
hosts.
\VS{11}The
{\ND{
LORD}}
{\ADD{will be}}
terrible unto them: for he will
famish all the
gods of the
earth; and
{\ADD{men}} shall
worship him, every
one from his
place,
{\ADD{even}} all the
isles of the
heathen.\FTNT{‡}{{\FR 2.11 }
famish: Heb. make lean}
\par }{\PP \VS{12}¶ Ye
Ethiopians also,
ye
{\ADD{shall be}}
slain by my
sword.
\VS{13}And he will stretch
out his
hand against the
north, and
destroy
Assyria; and will
make
Nineveh a
desolation,
{\ADD{and}}
dry like a
wilderness.
\VS{14}And
flocks shall lie
down in the
midst of her, all the
beasts of the
nations: both the
cormorant and the
bittern shall
lodge in the upper
lintels of it;
{\ADD{their}}
voice shall
sing in the
windows;
desolation
{\ADD{shall be}} in the
thresholds: for he shall
uncover the cedar
work.\FTNT{§}{{\FR 2.14 }
cormorant: or, pelican}\FTNT{**}{{\FR 2.14 }
upper…: or, knops, or, chapiters}\FTNT{††}{{\FR 2.14 }
for…: or, when he hath uncovered}
\VS{15}This
{\ADD{is}} the
rejoicing
city that
dwelt
carelessly, that
said in her
heart, I
{\ADD{am}}, and
{\ADD{there is}} none
beside me: how is she become a
desolation, a place for
beasts to lie down
in! every one that passeth
by her shall
hiss,
{\ADD{and}}
wag his
hand.

\par }\Chap{3}{\PP \VerseOne{1}Woe to her that is
filthy and
polluted, to the
oppressing
city!\FTNT{*}{{\FR 3.1 }
her…: or, gluttonous: Heb. craw}
\VS{2}She
obeyed not the
voice; she
received not
correction; she
trusted not in the
{\ND{
LORD}}; she drew not
near to her
God.\FTNT{†}{{\FR 3.2 }
correction: or, instruction}
\VS{3}Her
princes
within her
{\ADD{are}}
roaring
lions; her
judges
{\ADD{are}}
evening
wolves; they gnaw not the
bones till the
morrow.
\VS{4}Her
prophets
{\ADD{are}}
light
{\ADD{and}}
treacherous
persons: her
priests have
polluted the
sanctuary, they have done
violence to the
law.
\VS{5}The
just
{\ND{
LORD}}
{\ADD{is}} in the
midst thereof; he will not
do
iniquity:
every
morning doth he
bring his
judgment to
light, he
faileth not; but the
unjust
knoweth no
shame.\FTNT{‡}{{\FR 3.5 }
every…: Heb. morning by morning}
\VS{6}I have cut
off the
nations: their
towers are
desolate; I made their
streets
waste, that none passeth
by: their
cities are
destroyed, so that there is no
man, that there is none
inhabitant.\FTNT{§}{{\FR 3.6 }
towers: or, corners}
\VS{7}I
said, Surely thou wilt
fear me, thou wilt
receive
instruction; so their
dwelling should not be cut
off,
howsoever I
punished them:
but they rose
early,
{\ADD{and}}
corrupted all their
doings.
\par }{\PP \VS{8}¶ Therefore
wait ye upon me,
saith the
{\ND{
LORD}}, until the
day that I rise
up to the
prey: for my
determination
{\ADD{is}} to
gather the
nations, that I may
assemble the
kingdoms, to
pour upon them mine
indignation,
{\ADD{even}} all my
fierce
anger: for all the
earth shall be
devoured with the
fire of my
jealousy.
\VS{9}For then will I
turn to the
people a
pure
language, that they may all
call upon the
name of the
{\ND{
LORD}}, to
serve him with
one
consent.\FTNT{**}{{\FR 3.9 }
language: Heb. lip}\FTNT{††}{{\FR 3.9 }
consent: Heb. shoulder}
\VS{10}From
beyond the
rivers of
Ethiopia my
suppliants,
{\ADD{even}} the
daughter of my
dispersed, shall
bring mine
offering.
\VS{11}In that
day shalt thou not be
ashamed for all thy
doings, wherein thou hast
transgressed against me: for then I will take
away out of the
midst of thee them that
rejoice in thy
pride, and thou shalt no
more be
haughty because of my
holy
mountain.\FTNT{‡‡}{{\FR 3.11 }
because…: Heb. in my holy}
\VS{12}I will also
leave in the
midst of thee an
afflicted and
poor
people, and they shall
trust in the
name of the
{\ND{
LORD}}.
\VS{13}The
remnant of
Israel shall not
do
iniquity, nor
speak
lies; neither shall a
deceitful
tongue be
found in their
mouth: for they shall
feed and lie
down, and none shall make
{\ADD{them}}
afraid.
\par }{\PP \VS{14}¶
Sing, O
daughter of
Zion;
shout, O
Israel; be
glad and
rejoice with all the
heart, O
daughter of
Jerusalem.
\VS{15}The
{\ND{
LORD}} hath taken
away thy
judgments, he hath cast
out thine
enemy: the
king of
Israel,
{\ADD{even}} the
{\ND{
LORD}},
{\ADD{is}} in the
midst of thee: thou shalt not
see
evil any more.
\VS{16}In that
day it shall be
said to
Jerusalem,
Fear thou not:
{\ADD{and to}}
Zion, Let not thine
hands be
slack.\FTNT{§§}{{\FR 3.16 }
slack: or, faint}
\VS{17}The
{\ND{
LORD}} thy
God in the
midst of thee
{\ADD{is}}
mighty; he will
save, he will
rejoice over thee with
joy; he will
rest in his
love, he will
joy over thee with
singing.\FTNT{***}{{\FR 3.17 }
rest: Heb. be silent}
\VS{18}I will
gather
{\ADD{them that are}}
sorrowful for the solemn
assembly,
{\ADD{who}} are of thee,
{\ADD{to whom}} the
reproach of it
{\ADD{was}} a
burden.\FTNT{†††}{{\FR 3.18 }
reproach…: Heb. the burden upon it was reproach}
\VS{19}Behold, at that
time I will
undo all that
afflict thee: and I will
save her that
halteth, and
gather her that was driven
out; and I will
get them
praise and
fame in every
land where they have been put to
shame.\FTNT{‡‡‡}{{\FR 3.19 }
get…: Heb. set them for a praise}\FTNT{§§§}{{\FR 3.19 }
where…: Heb. of their shame}
\VS{20}At that
time will I
bring you
{\ADD{again}}, even in the
time that I
gather you: for I will
make you a
name and a
praise among all
people of the
earth, when I turn
back your
captivity before your
eyes,
saith the
{\ND{
LORD}}.
\par }